### 2014-11-10 sharp-sigma-plus.tex
document added

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm} 
\usepackage{fullpage}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bm}

\def\A{{\frak A}}

\newtheorem{theo}{Theorem}
\newtheorem{lemma}[theo]{Lemma}
\newtheorem{claim}[theo]{Claim}
\newtheorem{coro}[theo]{Corollary}

\begin{document}

\begin{flushleft}
{\footnotesize Pontificia Universidad Cat\'olica de Chile - Facultad de Ingenier\'ia\\
Departamento de Ciencias de la Computaci\'on\\
Proyecto de Investigaci\'on\\
Profesor Gu\'ia: Marcelo Arenas\\
}
\end{flushleft}
\begin{center}
{ \LARGE \bf
  Some properties of $\#\Sigma_1^{+}$
}
\end{center}

The $\#\Sigma_1^{+}$ class is defined as an extension of $\#\Sigma_1$ which adds the possibility of adding arbitrarily quantified sub-formulas that don't include counted predicates.

Formally, let $\vec{P}$ be a sequence of predicates and $\vec{x},\vec{z}$ sequences of variables. Let $\varphi(\vec{S},\vec{x},\vec{z})$ be a formula where the variables in $\vec{P}$ are not in a quantified sub-formula. Note that sub-formulas that include any other predicate are not restricted at all. A counting problem $F \in \#\Sigma_1^{+}$ if given a structure $\A$
$$
F(\A) = \vert \{ <\!\!\vec{P},\vec{z}\!\!> :\A \models \exists \vec{x} \varphi(\vec{S},\vec{x},\vec{z}) \} \vert
$$ 
Some of its properties are:

\begin{theo}
The decision version of a counting problem in $\#\Sigma_1^{+}$ is in \text{P}.
\end{theo}
\begin{proof}
Let $F$ be a function in $\#\Sigma_1^{+}$. Then there is a formula $\varphi$ such that:
$$
F(\A) = \vert \{ <\!\!\vec{P},\vec{z}\!\!> :\A \models \exists \vec{x} \varphi(\vec{S},\vec{x},\vec{z}) \} \vert
$$ 
where $\A$ is a finite ordered structure and $\vec{z}$ is an $m$-tuple. For each $\vec{z} \in A^m$, every arbitrarily quantified sub-formula which does not include $\vec{P}$ can be evaluated in polynomial time. Let $\varphi'$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula by a contradiction. Note that $\varphi'$ is quantifier-free.

Then, we compute a similar counting function:
$$
F'(\A) = \vert \{ <\!\!\vec{P},(\vec{z},\vec{x})\!\!> :\A \models \varphi(\vec{S},\vec{x},\vec{z}) \} \vert
$$ 
which is in $\#\Sigma_0$. Note that for every $\A$, $F(\A) > 0$ iff $F'(\A) > 0$, so computing $F'$ is enough to solve the decision version of $F$. However, Saluja and Subrahmanyam showed that any counting problem in $\#\Sigma_0$ is computable in polynomial time \footnote[1]{(paper)}, therefore, the decision version of $F$ is in P.
\end{proof}

\begin{theo}
$\#\Sigma_1$ is closed under substraction $\Rightarrow$ P = NP.
\end{theo}
\begin{proof}
\#3DNF $\in \#\Sigma_1$. Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment in a \#3DNF instance. Suppose that $F_{2^n}-F_{\#3DNF} \in \#\Sigma_1$. Then we can compute in polynomial time whether an arbitrary #3DNF formula is a tautology or not, a problem that is co-NP-complete.
\end{proof}

\begin{theo}
$\#\Sigma_1 \subsetneq \#\Sigma_1^{+}$
\end{theo}
\begin{proof}
We will show that the $\#\Sigma_1^{+}$ function defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall x_2 S(x_2)$ is not in $\#\Sigma_1$.
\end{proof}

\end{document}

### 2014-11-10 sharp-sigma-plus.tex
document added

@@
 Departamento de Ciencias de la Computaci\'on\\
-Proyecto de Investigaci\'on\\
-Profesor Gu\'ia: Marcelo Arenas\\
 }

### 2014-11-10 sharp-sigma-plus.tex
document

@@
 Suppose that $F_{2^n}-F_{\#3DNF} \in \#\Sigma_1$.
-Then we can compute in polynomial time whether an arbitrary #3DNF formula is a tautology or not, a problem that is co-NP-complete.
+This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete.
+However, as we showed previously (Theorem 1), it's also in P.
+Then, co-NP $\subseteq$ P.
 \end{proof}

### 2014-11-17 sharp-sigma-plus.tex
f-1 iniciado

@@
 
-The $\#\Sigma_1^{+}$ class is defined as an extension of $\#\Sigma_1$ which adds the possibility of adding arbitrarily quantified sub-formulas that don't include counted predicates.
+The $\#\Sigma_1^{+}$ class is defined as an extension of $\#\Sigma_1$ which includes the possibility of adding arbitrarily quantified sub-formulas that don't include counted predicates.
 
@@
 
+\begin{theo}
+$\#\Sigma_1$ is closed under substraction of 1 element.
+\end{theo}
+\begin{proof}
+The counting set has three possible ways of counting. Counting only variables, only predicates, and predicates with variables. This separates the proof in three cases:
+\begin{enumerate}
+\item The counting set is:
+$$
+F(\A) = \vert \{ <\!\!\vec{z}\!\!> :\A \models \exists \vec{x} \varphi(\vec{x},\vec{z}) \} \vert
+$$
+our goal here is to eliminate the lexicographically smallest sequence of variables, which we can do easily. Let $F'$ be:
+$$
+F'(\A) = \vert \{ <\!\!\vec{z}\!\!> :\A \models \exists \vec{x}\left( \varphi(\vec{x},\vec{z}) \wedge \exists \vec{z'} (\varphi(\vec{x},\vec{z'}) \wedge \varphi_{\lneq}(\vec{z'},\vec{z} ) ) \right) \} \vert
+$$
+where $\varphi_{\lneq}(\vec{z'},\vec{z})$ is a formula that indicates that $\vec{z'}$ is lexicographically smaller than $\vec{z}$.
+Thus, $F'(\A)$ will count exactly one element less than $F(\A)$ if $F(\A)>0$.
+\item The counting set is:
+$$
+F(\A) = \vert \{ <\!\!\vec{P}\!\!> :\A \models \exists \vec{x} \varphi(\vec{x},\vec{P}) \} \vert
+$$
+similarly, we would like to eliminate the lexicographically smallest predicate $P$ that holds the formula true with the exact number of tuples.
+First we treat quantified sub-formulas as literals, and take a DNF formula that's equivalent to $\varphi$.
+Then
+$$
+F(\A) = \vert \{ <\!\!\vec{P}\!\!> :\A \models 
+\exists \vec{x} \varphi_1(\vec{x},\vec{P}) \vee
+\exists \vec{x} \varphi_2(\vec{x},\vec{P}) \vee
+... \vee
+\exists \vec{x} \varphi_k(\vec{x},\vec{P})
+ \} \vert
+$$
+Note that
+$$
+\varphi_i(\vec{x},\vec{P}) = \vec{P_1}(\vec{x}) \wedge \neg \vec{P_0}(\vec{x}) \wedge \varphi'_i(\vec{x})
+$$
+where $\vec{P_0},\vec{P_1}$ are subsequences of $\vec{P}$ and $\varphi'_i$ is arbitrarily quantified.
+We can assume every predicate in $\vec{P},$ uses all variables in $\vec{x}$ and they are not permutated(?).
+Let $\psi_i(x)$ be:
+$$
+\psi_i(\vec{x}) = \exists x' (x' \neq x \wedge P_1(x')) \vee \exists x' (\varphi'_i(x') \wedge \varphi_{\lneq}(x',x))
+$$
+then 
+$$
+\exists \vec{x} \left( \varphi_i(\vec{x},\vec{P}) \wedge \psi_i(\vec{x}) \right)
+$$
+will have exactly one predicate less than $\exists \vec{x} \varphi_1(\vec{x},\vec{P})$.
+\end{enumerate}
+\end{proof}
+
 \end{document}

### 2014-11-17 sharp-sigma-plus.tex
f-1 avance

@@
 similarly, we would like to eliminate the lexicographically smallest predicate $P$ that holds the formula true with the exact number of tuples.
-First we treat quantified sub-formulas as literals, and take a DNF formula that's equivalent to $\varphi$.
+First we treat sub-formulas without $\vec{P}$ as literals, and take a DNF formula that's equivalent to $\varphi$.
 Then
@@
 $$
-where $\vec{P_0},\vec{P_1}$ are subsequences of $\vec{P}$ and $\varphi'_i$ is arbitrarily quantified.
-We can assume every predicate in $\vec{P},$ uses all variables in $\vec{x}$ and they are not permutated(?).
-Let $\psi_i(x)$ be:
+where $\vec{P_0},\vec{P_1}$ are disjunct subsequences of $\vec{P}$ and $\varphi'_i$ is arbitrarily quantified.
+We can assume every predicate in $\vec{P}$ uses all variables in $\vec{x}$ and they are not permutated \footnote[1]{revisar esto}.
+Hence, if there is at least one assignation to $\vec{x}$ that satisfies $\varphi'_i(\vec{x})$, there is a predicate that has exactly one tuple that is lexicographically minimal.
+Thus, if
 $$
@@
 $$
-will have exactly one predicate less than $\exists \vec{x} \varphi_1(\vec{x},\vec{P})$.
+will be satisfied by exactly one predicate less than $\exists \vec{x} \varphi_1(\vec{x},\vec{P})$.
+If $\varphi'_i(\vec{x})$ is unsatisfiable, we eliminate one predicate on the next $\varphi_i$.
+If this sub-formula is also unsatisfiable, we move to the next, and so on.
+Therefore:
+\begin{multline*}
+F(\A) = \vert \{ <\!\!\vec{P}\!\!> :\A \models 
+\exists \vec{x} \left( \varphi_1(\vec{x},\vec{P}) \wedge \psi_1(\vec{x}) \right) \vee
+\exists \vec{x} \left( \varphi_2(\vec{x},\vec{P}) \wedge \left( (\forall \vec{x'} \neg \varphi'_1(\vec{x'})) \rightarrow \psi_2(\vec{x}) \right)\right) \vee \\
+... \vee
+\exists \vec{x} \left( \varphi_k(\vec{x},\vec{P}) \wedge \left( (\forall \vec{x'} (\neg \varphi'_1(\vec{x'}) \wedge \neg \varphi'_2(\vec{x'}) \wedge ...
+\wedge \neg \varphi'_{k-1}(\vec{x'})) ) \rightarrow \psi_k(\vec{x}) \right)\right)
+ \} \vert
+\end{multline}
 \end{enumerate}

### 2014-11-17 sharp-sigma-plus.tex
f-1 bosquejado

@@
 \begin{multline*}
-F(\A) = \vert \{ <\!\!\vec{P}\!\!> :\A \models 
+F_{-1}(\A) = \vert \{ <\!\!\vec{P}\!\!> :\A \models 
 \exists \vec{x} \left( \varphi_1(\vec{x},\vec{P}) \wedge \psi_1(\vec{x}) \right) \vee
@@
 \end{multline}
+eliminates exactly one predicate of $F(\A)$.
 \end{enumerate}
+\item The counting set is:
+$$
+F(\A) = \vert \{ <\!\!\vec{P},\vec{z}\!\!> :\A \models \exists \vec{x} \varphi(\vec{x}, \vec{P},\vec{z}) \} \vert
+$$
+Then, we going to isolate the minimal predicate $\vec{P}$ that holds the formula true and eliminate the lexicographically smallest $\vec{z}$ that satisfies it.
+We mix both previous strategies.
+For each DNF disjunct $\varphi_i(\vec{x}, \vec{z})$ we add:
+$$
+\psi_i(\vec{x}, \vec{z}) = \exists x' (x' \neq x \wedge P_1(x')) \vee \exists x' (\varphi'_i(x') \wedge \varphi_{\lneq}(x',x)) \vee \exists z'(\varphi_i(\vec{x}, \vec{z}) \wedge \varphi_{\lneq}(z',z))
+$$
+and proceed analogously.
 \end{proof}

### 2014-11-24 sharp-sigma-plus.tex
not perm

--- 2014-11-17 sharp-sigma-plus.tex
@@
 $$
-F(\A) = \vert \{ <\!\!\vec{P},\vec{z}\!\!> :\A \models \exists \vec{x} \varphi(\vec{S},\vec{x},\vec{z}) \} \vert
+F(\A) = \vert \{ \langle\vec{P},\vec{z}\rangle :\A \models \exists \vec{x} \varphi(\vec{S},\vec{x},\vec{z}) \} \vert
 $$ 
@@
 $$
-F'(\A) = \vert \{ <\!\!\vec{P},(\vec{z},\vec{x})\!\!> :\A \models \varphi(\vec{S},\vec{x},\vec{z}) \} \vert
+F'(\A) = \vert \{ \langle\vec{P},(\vec{z},\vec{x})\rangle :\A \models \varphi(\vec{S},\vec{x},\vec{z}) \} \vert
 $$ 
@@
 $$
-F(\A) = \vert \{ <\!\!\vec{z}\!\!> :\A \models \exists \vec{x} \varphi(\vec{x},\vec{z}) \} \vert
+F(\A) = \vert \{ \langle\vec{z}\rangle :\A \models \exists \vec{x} \ \varphi(\vec{x},\vec{z}) \} \vert
 $$
@@
 $$
-F'(\A) = \vert \{ <\!\!\vec{z}\!\!> :\A \models \exists \vec{x}\left( \varphi(\vec{x},\vec{z}) \wedge \exists \vec{z'} (\varphi(\vec{x},\vec{z'}) \wedge \varphi_{\lneq}(\vec{z'},\vec{z} ) ) \right) \} \vert
+F'(\A) = \vert \{ \langle\vec{z}\rangle :\A \models \exists \vec{x}\left( \varphi(\vec{x},\vec{z}) \wedge \exists \vec{z'} (\varphi(\vec{x},\vec{z'}) \wedge \varphi_{\lneq}(\vec{z'},\vec{z} ) ) \right) \} \vert
 $$
@@
 $$
-F(\A) = \vert \{ <\!\!\vec{P}\!\!> :\A \models \exists \vec{x} \varphi(\vec{x},\vec{P}) \} \vert
+F(\A) = \vert \{ \langle\vec{P}\rangle :\A \models \exists \vec{x} \ \varphi(\vec{x},\vec{P}) \} \vert
 $$
@@
 $$
-F(\A) = \vert \{ <\!\!\vec{P}\!\!> :\A \models 
+F(\A) = \vert \{ \langle\vec{P}\rangle :\A \models 
 \exists \vec{x} \varphi_1(\vec{x},\vec{P}) \vee
@@
 $$
-\varphi_i(\vec{x},\vec{P}) = \vec{P_1}(\vec{x}) \wedge \neg \vec{P_0}(\vec{x}) \wedge \varphi'_i(\vec{x})
+\varphi_i(\vec{x},\vec{P}) = \vec{Q}(\vec{x}) \wedge \neg \vec{R}(\vec{x}) \wedge \varphi'_i(\vec{x})
 $$
-where $\vec{P_0},\vec{P_1}$ are disjunct subsequences of $\vec{P}$ and $\varphi'_i$ is arbitrarily quantified.
-We can assume every predicate in $\vec{P}$ uses all variables in $\vec{x}$ and they are not permutated \footnote[1]{revisar esto}.
-Hence, if there is at least one assignation to $\vec{x}$ that satisfies $\varphi'_i(\vec{x})$, there is a predicate that has exactly one tuple that is lexicographically minimal.
+where $\vec{Q},\vec{R}$ are subsequences of $\vec{P}$ and $\varphi'_i$ is arbitrarily quantified.
+We can assume that variables in $\vec{x}$ appear in at most one predicate since we can replace common variables with new ones and add appropriate equals in $\varphi'_i$.
+Hence, if there is at least one assignment to $\vec{x}$ that satisfies $\varphi'_i(\vec{x})$, there is a predicate that has exactly one tuple that is lexicographically minimal.
 Thus, if
 $$
-\psi_i(\vec{x}) = \exists x' (x' \neq x \wedge P_1(x')) \vee \exists x' (\varphi'_i(x') \wedge \varphi_{\lneq}(x',x))
+\psi_i(\vec{x}) = \exists x' (x' \neq x \wedge Q(x')) \vee \exists x' (\varphi'_i(x') \wedge \varphi_{\lneq}(x',x))
 $$
@@
 If $\varphi'_i(\vec{x})$ is unsatisfiable, we eliminate one predicate on the next $\varphi_i$.
-If this sub-formula is also unsatisfiable, we move to the next, and so on.
+If this sub-formula is also unsatisfiable, we move to the next one, and so on.
 Therefore:
 \begin{multline*}
-F(\A) = \vert \{ <\!\!\vec{P}\!\!> :\A \models 
+F_{-1}(\A) = \vert \{ \langle\vec{P}\rangle :\A \models 
 \exists \vec{x} \left( \varphi_1(\vec{x},\vec{P}) \wedge \psi_1(\vec{x}) \right) \vee
@@
 \end{multline}
+eliminates exactly one predicate of $F(\A)$.
 \end{enumerate}
+\item The counting set is:
+$$
+F(\A) = \vert \{ \langle\vec{P},\vec{z}\rangle :\A \models \exists \vec{x} \ \varphi(\vec{x}, \vec{P},\vec{z}) \} \vert
+$$
+Then, we going to isolate the minimal predicate $\vec{P}$ that holds the formula true and eliminate the lexicographically smallest $\vec{z}$ that satisfies it.
+We mix both previous strategies.
+For each DNF disjunct $\varphi_i(\vec{x}, \vec{z})$ we add:
+$$
+\psi_i(\vec{x}, \vec{z}) = \exists x' (x' \neq x \wedge Q(x')) \vee \exists x' (\varphi'_i(x') \wedge \varphi_{\lneq}(x',x)) \vee \exists z'(\varphi_i(\vec{x}, \vec{z}) \wedge \varphi_{\lneq}(z',z))
+$$
+and proceed analogously.
 \end{proof}

### 2014-12-01 sharp-sigma-plus.tex
f-1 v.1

@@
 $$
-Note that
+We can also suppose that all of this formulas are satisfied by a single total order for $\vec{x}$. Therefore
 $$
-\varphi_i(\vec{x},\vec{P}) = \vec{Q}(\vec{x}) \wedge \neg \vec{R}(\vec{x}) \wedge \varphi'_i(\vec{x})
+\varphi_i(\vec{x},\vec{P}) = \bigwedge_{i=1}^n P_i(\vec{x_i}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x})
 $$
-where $\vec{Q},\vec{R}$ are subsequences of $\vec{P}$ and $\varphi'_i$ is arbitrarily quantified.
-We can assume that variables in $\vec{x}$ appear in at most one predicate since we can replace common variables with new ones and add appropriate equals in $\varphi'_i$.
-Hence, if there is at least one assignment to $\vec{x}$ that satisfies $\varphi'_i(\vec{x})$, there is a predicate that has exactly one tuple that is lexicographically minimal.
-Thus, if
-$$
-\psi_i(\vec{x}) = \exists x' (x' \neq x \wedge Q(x')) \vee \exists x' (\varphi'_i(x') \wedge \varphi_{\lneq}(x',x))
-$$
-then 
-$$
-\exists \vec{x} \left( \varphi_i(\vec{x},\vec{P}) \wedge \psi_i(\vec{x}) \right)
-$$
-will be satisfied by exactly one predicate less than $\exists \vec{x} \varphi_1(\vec{x},\vec{P})$.
-If $\varphi'_i(\vec{x})$ is unsatisfiable, we eliminate one predicate on the next $\varphi_i$.
-If this sub-formula is also unsatisfiable, we move to the next one, and so on.
-Therefore:
+where $P_i$ and $\overline{P}$ are subsequences of $\vec{P}$, $\vec{x_i}$ and $\vec{y}$ are subsequences of $\vec{x}$, $\theta$ defines a total order for $\vec{x}$, and $\varphi_{\vec{S}}$ involves only closed predicates in $\vec{S}$.
+We now define
 \begin{multline*}
-F_{-1}(\A) = \vert \{ \langle\vec{P}\rangle :\A \models 
-\exists \vec{x} \left( \varphi_1(\vec{x},\vec{P}) \wedge \psi_1(\vec{x}) \right) \vee
-\exists \vec{x} \left( \varphi_2(\vec{x},\vec{P}) \wedge \left( (\forall \vec{x'} \neg \varphi'_1(\vec{x'})) \rightarrow \psi_2(\vec{x}) \right)\right) \vee \\
-... \vee
-\exists \vec{x} \left( \varphi_k(\vec{x},\vec{P}) \wedge \left( (\forall \vec{x'} (\neg \varphi'_1(\vec{x'}) \wedge \neg \varphi'_2(\vec{x'}) \wedge ...
-\wedge \neg \varphi'_{k-1}(\vec{x'})) ) \rightarrow \psi_k(\vec{x}) \right)\right)
- \} \vert
+\alpha_{\min}^i(\vec{x_1}',...,\vec{x_n}') = \\ \exists\vec{y}\left( \alpha(\vec{x_1}',...,\vec{x_n}',\vec{y})\wedge \forall\vec{x_1}''...\vec{x_n}''\vec{y}'(\alpha(\vec{x_1}'',...,\vec{x_n}'',\vec{y}')\rightarrow (\vec{x_1}',...,\vec{x_n}') \leq (\vec{x_1}'',...,\vec{x_n}'')\right)
 \end{multline}
-eliminates exactly one predicate of $F(\A)$.
+Where $\alpha(\vec{x}) = \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x})$. Our new formula would be
+\begin{multline*}
+\varphi_i(\vec{x},\vec{P}) = \bigwedge_{i=1}^n P_i(\vec{x_i}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x})\wedge\alpha_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge \\
+\left(\bigvee_{P\in\overline{P}}\bigvee_{P_i = P}\neg P(\vec{x_i}') \vee \bigvee_{P\in\overline{P}} \exists \vec{x}''\left(P(\vec{x}'') \bigwedge_{P_i = P} \vec{x}'' \neq \vec{x_i}'\right)\right)
+\end{multline}
+and $\vec{x_1}',...,\vec{x_n}'$ are added to the existential quantifier. This will be satisfied by exactly one predicate less than $\exists \vec{x} \varphi_1(\vec{x},\vec{P})$.
+
+If $\alpha^i(\vec{x})$ is unsatisfiable, we eliminate one predicate on the next $\varphi_i$. If this sub-formula is also unsatisfiable, we move to the next one, and so on. Therefore:
+
 \end{enumerate}

### 2014-12-07 sharp-sigma-plus.tex
f-1 solo preds

@@
 $$
-We can also suppose that all of this formulas are satisfied by a single total order for $\vec{x}$. Therefore
+We can also suppose that all of this formulas are satisfied by a single total order for $\vec{x}$. We can manually eliminate unsatisfiable sub-formulas such as
 $$
-\varphi_i(\vec{x},\vec{P}) = \bigwedge_{i=1}^n P_i(\vec{x_i}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x})
+\exists x,y(P(x,y) \wedge \neg P(x,x) \wedge x = y)
 $$
-where $P_i$ and $\overline{P}$ are subsequences of $\vec{P}$, $\vec{x_i}$ and $\vec{y}$ are subsequences of $\vec{x}$, $\theta$ defines a total order for $\vec{x}$, and $\varphi_{\vec{S}}$ involves only closed predicates in $\vec{S}$.
+Therefore,
+$$
+\varphi_1(\vec{x},\vec{P}) = \bigwedge_{i=1}^n P_j(\vec{x_j}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x})
+$$
+where $P_i$ and $\overline{P}$ are subsequences of $\vec{P}$, $P_i$ is of arity $a_i$ and $\vec{x_i}$ is an $a_i$-tuple of variables in $\vec{x}$, $\vec{y}$ is also a subsequence of $\vec{x}$, $\theta(\vec{x})$ defines a total order for $\vec{x}$, and $\varphi_{\vec{S}}(\vec{x})$ involves only closed predicates.
+Without unsatisfiable formulas, we can assure that if there is an assignation $\vec{x}$ that satisfies $\alpha(\vec{x}) = \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x})$, there is at least one possible assignation to $\vec{P}$.
 We now define
@@
 \end{multline}
-Where $\alpha(\vec{x}) = \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x})$. Our new formula would be
+Note that $\alpha_{\min}^i$ is satisfied only by the lexicographically smallest assignation $\vec{x}^*$ to $\vec{x}$ that satisfies both $\theta(\vec{x})$ and $\varphi_{\vec{S}}(\vec{x})$.
+Our new formula is
 \begin{multline*}
-\varphi_i(\vec{x},\vec{P}) = \bigwedge_{i=1}^n P_i(\vec{x_i}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x})\wedge\alpha_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge \\
-\left(\bigvee_{P\in\overline{P}}\bigvee_{P_i = P}\neg P(\vec{x_i}') \vee \bigvee_{P\in\overline{P}} \exists \vec{x}''\left(P(\vec{x}'') \bigwedge_{P_i = P} \vec{x}'' \neq \vec{x_i}'\right)\right)
+\varphi_1'(\vec{x},\vec{P}) = \bigwedge_{i=1}^n P_i(\vec{x_i}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x})\wedge\exists\vec{x_1}',...,\vec{x_n}'º\Bigg(\alpha_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge \\
+\bigg(\bigvee_{P\in\vec{P}}\bigvee_{\substack{i\in\{1,...,n\}\\P_i = P}}\neg P(\vec{x_i}') \vee \bigvee_{P\in\vec{P}} \exists \vec{x}''\Big( P(\vec{x}'') \bigwedge_{P_i = P} \vec{x}'' \neq \vec{x_i}'\Big) \bigg) \Bigg)
 \end{multline}
-and $\vec{x_1}',...,\vec{x_n}'$ are added to the existential quantifier. This will be satisfied by exactly one predicate less than $\exists \vec{x} \varphi_1(\vec{x},\vec{P})$.
+where $\vec{x_i}'$ is a an $a_i$-tuple of variables in $\vec{x}$.
+If there is an assignation to satisfy $\alpha^j(\vec{x})$, this formula will be satisfied by exactly one predicate less than $\exists \vec{x} \varphi_1(\vec{x},\vec{P})$.
+Let $\vec{P}^*$ be the excluded assignation to $\vec{P}$.
+Suppose there is another assignation $\vec{P}^*'$ to $\vec{P}$ that is also excluded by the new sub-formula.
+Then, $\vec{x}^*$ is such that every tuple $\vec{x}_i^*$ is in $\vec{P}_i^*'$ and $\vec{x}^*$ includes every tuple in $\vec{P}^*'$.
+If that's the case, $\vec{P}^*'$ contains exactly the same tuples as $\vec{P}^*$, and are thus the same.
+Now we need to exclude that assignation to the next disjuncts, which we do by adding the same sub-formula.
+We also need to include the case where there are no possible assignations to $\vec{x}$ for the first disjunct, so we add this sub-formula:
+\begin{multline*}
+\forall\vec{x}\left(\neg\alpha^1_{\min}(\vec{x})\right) \to \exists\vec{x_1}',...,\vec{x_n}'º\Bigg(\alpha^2_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge \\
+\bigg(\bigvee_{P\in\vec{P}}\bigvee_{\substack{i\in\{1,...,n\}\\P_i = P}}\neg P(\vec{x_i}') \vee \bigvee_{P\in\vec{P}} \exists \vec{x}''\Big( P(\vec{x}'') \bigwedge_{P_i = P} \vec{x}'' \neq \vec{x_i}'\Big) \bigg) \Bigg)
+\end{multline}
+The full disjunct is as follows:
+\begin{multline*}
+\varphi_j'(\vec{x},\vec{P}) = \bigwedge_{i=1}^n P_i(\vec{x_i}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\vec{x}) \wedge \varphi_{\vec{S}}(\vec{x}) \wedge \Bigg(\Big(\forall\vec{x}\left(\neg\alpha^1_{\min}(\vec{x})\right)\wedge ...
+\wedge \forall\vec{x}\left(\neg\alpha^{j-1}_{\min}(\vec{x})\right)\Big) \to \\
+\exists\vec{x_1}',...,\vec{x_n}'º\Bigg(\alpha^j_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge
+\bigg(\bigvee_{P\in\vec{P}}\bigvee_{\substack{i\in\{1,...,n\}\\P_i = P}}\neg P(\vec{x_i}') \vee \bigvee_{P\in\vec{P}} \exists \vec{x}''\Big( P(\vec{x}'') \bigwedge_{P_i = P} \vec{x}'' \neq \vec{x_i}'\Big) \bigg) \Bigg) \Bigg)
+\end{multline}
+which, if all of the previous disjuncts have no possible assignations, eliminates the least assignation to $\vec{P}$ in that disjunct. Finally, the counting set
+$$
+F'(\A) = \vert \{ \langle\vec{P}\rangle :\A \models 
+\exists \vec{x} \varphi'_1(\vec{x},\vec{P}) \vee
+\exists \vec{x} \varphi'_2(\vec{x},\vec{P}) \vee
+... \vee
+\exists \vec{x} \varphi'_k(\vec{x},\vec{P})
+ \} \vert
+$$
+counts exactly one assignation to $\vec{P}$ less than $F(\A)$.
 
-If $\alpha^i(\vec{x})$ is unsatisfiable, we eliminate one predicate on the next $\varphi_i$. If this sub-formula is also unsatisfiable, we move to the next one, and so on. Therefore:
-
-\end{enumerate}
 \item The counting set is:
@@
 $$
-Then, we going to isolate the minimal predicate $\vec{P}$ that holds the formula true and eliminate the lexicographically smallest $\vec{z}$ that satisfies it.
-We mix both previous strategies.
-For each DNF disjunct $\varphi_i(\vec{x}, \vec{z})$ we add:
-$$
-\psi_i(\vec{x}, \vec{z}) = \exists x' (x' \neq x \wedge Q(x')) \vee \exists x' (\varphi'_i(x') \wedge \varphi_{\lneq}(x',x)) \vee \exists z'(\varphi_i(\vec{x}, \vec{z}) \wedge \varphi_{\lneq}(z',z))
-$$
-and proceed analogously.
+Then, we going to isolate the minimal predicate $\vec{P}$ that holds the formula true and eliminate the lexicographically smallest $\vec{z}$ that satisfies it. We mix both previous strategies.
+\end{enumerate}
 \end{proof}

### 2014-12-15 sharp-sigma-one-plus.tex
f-1 v2

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm} 
\usepackage{fullpage}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bm}

\def\A{{\frak A}}
\def\C{{\cal C}}
\def\L{{\cal L}}
\def\Q{{\cal Q}}
\def\F{{\cal F}}
\def\P{\vec{P}}
\def\S{\vec{S}}
\def\x{\vec{x}}
\def\z{\vec{z}}

\newtheorem{theo}{Theorem}
\newtheorem{lemma}[theo]{Lemma}
\newtheorem{claim}[theo]{Claim}
\newtheorem{coro}[theo]{Corollary}

\begin{document}

\begin{center}
{ \LARGE \bf
  Some properties of $\#\Sigma_1^{+}$
}
\end{center}

We define the vocabulary $\L = \langle \P, \S, \leq \rangle$, where $\P$ is a $r$-tuple of predicates of arity $a_1,a_2,...,a_r$, and $\S$ is a $t$-tuple of predicates of arity $b_1,b_2,...,b_t$. A counting problem $\Q$ in $\#\Sigma_1$ is defined by a quantifier free $\L$-formula $\varphi(\x,\P,\z)$ where its function $\F_\Q(\A)$ is
\begin{eqnarray*}
\F_\Q(\A) &=& \mid \{ \langle \P, \z \rangle \mid \A \models \exists \x \: \varphi(\x,\P,\z) \} \mid
\end{eqnarray*}
for every $\L$-structure $\A = \langle A, \S^\A, \leq^\A \rangle$, where $\x = (x_1,x_2,...,x_c)$, $\z = (z_1,z_2,...,z_d)$, and $\leq^\A$ is a total order for $A$. We refer to predicates in $\P$ and variables in $\z$ as "open", and predicates in $\S$ and variables in $\x$ as "closed". The $\#\Sigma_1^{+}$ class is defined as an extension of $\#\Sigma_1$ which includes the possibility of adding arbitrarily quantified $\L$-sub-formulas $\psi(\x,\z)$ which may contain closed predicates but no open ones.

Some properties of the $\#\Sigma_1^{+}$ class are:

\begin{theo}
The decision version of a counting problem in $\#\Sigma_1^{+}$ is in \text{P}.
\end{theo}
\begin{proof}
Let $F$ be a function in $\#\Sigma_1^{+}$. Then there is a formula $\varphi$ such that
\begin{eqnarray*}
F(\A) &=& \mid \{ \langle\P,\z\rangle \mid \A \models \exists \x \varphi(\S,\x,\z) \} \mid,
\end{eqnarray*}
where $\A$ is a finite ordered structure and $\z$ is an $m$-tuple. For each $\z \in A^m$, every arbitrarily quantified sub-formula which does not include $\P$ can be evaluated in polynomial time. Let $\varphi'$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula by a contradiction. Note that $\varphi'$ is quantifier-free.

Then, we compute a similar counting function,
\begin{eqnarray*}
F'(\A) &=& \mid \{ \langle\P,(\z,\x)\rangle \mid \A \models \varphi(\S,\x,\z) \} \mid
\end{eqnarray*}
which is in $\#\Sigma_0$. Note that for every $\A$, $F(\A) > 0$ iff $F'(\A) > 0$, so computing $F'$ is enough to solve the decision version of $F$. However, Saluja and Subrahmanyam showed that any counting problem in $\#\Sigma_0$ is computable in polynomial time \footnote[1]{(paper)}, therefore, the decision version of $F$ is in P.
\end{proof}

\begin{theo}
$\#\Sigma_1$ is closed under substraction $\Rightarrow$ P = NP.
\end{theo}
\begin{proof}
\#3DNF $\in \#\Sigma_1$. Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment in a \#3DNF instance. Suppose that $F_{2^n}-F_{\#3DNF} \in \#\Sigma_1$. This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete. However, as we showed previously (Theorem 1), it's also in P. Then, co-NP $\subseteq$ P.
\end{proof}

\begin{theo}
$\#\Sigma_1 \subsetneq \#\Sigma_1^{+}$
\end{theo}
\begin{proof}
We will show that the $\#\Sigma_1^{+}$ function defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall x_2 S(x_2)$ is not in $\#\Sigma_1$.
\end{proof}

%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 
%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1

\begin{theo}
$\#\Sigma_1^+$ is closed under substraction\footnote[1]{A counting class $\C$ is closed under substraction if given a counting problem $\Q \in \C$ there exists $\Q' \in \C$ where $$
\F_{\Q'}(\A) =
\begin{cases}
\F_\Q(\A)-1, & \text{if }\F_{\Q}(\A) > 0 \\
0, & \text{if }\F_{\Q}(\A) = 0 
\end{cases}
$$ }.
\end{theo}
\begin{proof}
The counting set has three possible ways of counting. Counting only variables, only predicates, and predicates with variables. This separates the proof in three cases:
\begin{enumerate}

% CASE 1

\item Let $\Q \in \#\Sigma_1^+$, which is defined by an $\L$-formula $\varphi(\x,\z)$. Its function would be
\begin{eqnarray*}
\F_\Q(\A) &=& \mid \{ \langle\z\rangle \mid \A \models \exists \x \ \varphi(\x,\z) \} \mid.
\end{eqnarray*}
Our goal here is to eliminate the lexicographically smallest sequence of variables, which we can do easily. Let $\F_{\Q'}$ be
\begin{eqnarray*}
\F_{\Q'}(\A) &=& \mid \{ \langle\z\rangle \mid \A \models \exists \x\left( \varphi(\x,\z) \wedge \exists \vec{z'} (\varphi(\x,\vec{z'}) \wedge \varphi_<(\vec{z'},\z ) ) \right) \} \mid,
\end{eqnarray*}
where
\begin{eqnarray*}
\varphi_<(\vec{z'},\z) &=& \bigvee_{i = 1}^d \left( \bigwedge_{j=1}^{i-1} z_j = z_j' \wedge z_i < z_i' \right).
\end{eqnarray*}
This formula is true if $\vec{z'}$ is lexicographically smaller than $\z$. Thus, $\F_{\Q'}(\A)$ will count exactly one element less than $\F_\Q(\A)$ if $\F_\Q(\A)>0$.

%CASE 2

\item Let $\Q \in \#\Sigma_1^+$, which is defined by an $\L$-formula $\varphi(\x,\P)$. Then
\begin{eqnarray*}
\F_\Q(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid.
\end{eqnarray*}
Suppose that
\begin{eqnarray*}
\varphi(\x,\P) &=& \bigwedge_{i=1}^n P_j(\vec{x_j}) \wedge \varphi_{\P}^{-}(\vec{y}) \wedge \theta(\x) \wedge \varphi_{\S}(\x)
\end{eqnarray*}
where $P_1,P_2,...,P_n$ is a subsequence of $\P$, $P_i$ is of arity $a_i'$ and $\vec{x_i}$ is an $a_i'$-tuple of variables in $\x$ for all $i \in \{1,...,n\}$, $\theta(\x)$ defines a total order for $\x$, and $\varphi_{\S}(\x)$ involves only closed predicates. We can assume $(\vec{x_1},\vec{x_2},...,\vec{x_n},\vec{y}) = \x$ by adding equivalences to $\theta(\x)$. For example, if the function is defined by
\begin{eqnarray*}
\exists x,y \left( P(x,y) \wedge \neg P(x,x) \wedge x < y \wedge \forall z\big( S(x,z) \big) \right),
\end{eqnarray*}
we can use
\begin{eqnarray*}
\exists x_1,x_2,x_3,x_4 \left( P(x_1,x_2) \wedge \neg P(x_3,x_4) \wedge (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 ) \wedge \forall z\big( S(x_1,z) \big) \right).
\end{eqnarray*}
Similarly, we would like to eliminate the lexicographically smallest predicate $P$ that satisfies the formula. We now define
\begin{multline*}
\alpha_{\min}(\vec{x_1}',...,\vec{x_n}') = \\ \exists\vec{y}\left( \alpha(\vec{x_1}',...,\vec{x_n}',\vec{y})\wedge \forall\vec{x_1}''...\vec{x_n}''\vec{y}'(\alpha(\vec{x_1}'',...,\vec{x_n}'',\vec{y}')\rightarrow \varphi_<((\vec{x_1}',...,\vec{x_n}'),(\vec{x_1}'',...,\vec{x_n}''))\right),
\end{multline*}
where $\alpha(\x) = \theta(\x) \wedge \varphi_{\S}(\x)$. Note that $\alpha_{\min}^i$ is satisfied only by the lexicographically smallest assignation $\x^*$ to $\x$ that satisfies both $\theta(\x)$ and $\varphi_{\S}(\x)$. Our new formula is

\begin{multline*}
\varphi'(\x,\P) = \bigwedge_{i=1}^n P_i(\vec{x_i}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\x) \wedge \varphi_{\S}(\x)\wedge\exists\vec{x_1}',...,\vec{x_n}'º\Bigg(\alpha_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge \\ \bigg(\bigvee_{i = 1}^{n}\neg P_i(\vec{x_i}') \vee \bigvee_{P\in\P} \exists \x''\Big( P(\x'') \wedge \bigwedge_{P_i = P} \x'' \neq \vec{x_i}'\Big) \bigg) \Bigg)
\end{multline*}

where $\vec{x_i}'$ is a an $a_i$-tuple of variables in $\x$. If there is an assignation to satisfy $\alpha^j(\x)$, this formula will be satisfied by exactly one predicate less than $\exists \x \varphi_1(\x,\P)$. Let $\P^*$ be the excluded assignation to $\P$. Suppose there is another assignation $\P^*'$ to $\P$ that is also excluded by the new sub-formula. Then, $\x^*$ is such that every tuple $\x_i^*$ is in $\P_i^*'$ and $\x^*$ includes every tuple in $\P^*'$. If that's the case, $\P^*'$ contains exactly the same tuples as $\P^*$, and are thus the same.

To use this result for the general case, let
\begin{eqnarray*}
\F_\Q(\A) &=& \mid \{ \langle\P\rangle :\A \models 
\exists \x \varphi_1(\x,\P) \vee
\exists \x \varphi_2(\x,\P) \vee
... \vee
\exists \x \varphi_k(\x,\P)
 \} \mid
\end{eqnarray*}
where $\varphi(\x,\P) \equiv \varphi_1(\x,\P) \vee \varphi_2(\x,\P) \vee ...  \vee \varphi_k(\x,\P) $.

Now we need to exclude that assignation to the next disjuncts, which we do by adding the same sub-formula. We also need to include the case where there are no possible assignations to $\x$ for the first disjunct, so we add the sub-formula
\begin{multline*}
\forall\x\left(\neg\alpha^1_{\min}(\x)\right) \to \exists\vec{x_1}',...,\vec{x_n}'\Bigg(\alpha^2_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge \\
\bigg(\bigvee_{P\in\P}\bigvee_{\substack{i\in\{1,...,n\}\\P_i = P}}\neg P(\vec{x_i}') \vee \bigvee_{P\in\P} \exists \x''\Big( P(\x'') \bigwedge_{P_i = P} \x'' \neq \vec{x_i}'\Big) \bigg) \Bigg)
\end{multline*}
The full disjunct is as follows,
\begin{multline*}
\varphi_j'(\x,\P) = \bigwedge_{i=1}^n P_i(\vec{x_i}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\x) \wedge \varphi_{\S}(\x) \wedge \Bigg(\Big(\forall\x\left(\neg\alpha^1_{\min}(\x)\right)\wedge ... \wedge \forall\x\left(\neg\alpha^{j-1}_{\min}(\x)\right)\Big) \to \\ 
\exists\vec{x_1}',...,\vec{x_n}'\Bigg(\alpha^j_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge
\bigg(\bigvee_{P\in\P}\bigvee_{\substack{i\in\{1,...,n\}\\P_i = P}}\neg P(\vec{x_i}') \vee \bigvee_{P\in\P} \exists \x''\Big( P(\x'') \bigwedge_{P_i = P} \x'' \neq \vec{x_i}'\Big) \bigg) \Bigg) \Bigg)
\end{multline*}
which, if all of the previous disjuncts have no possible assignations, eliminates the least assignation to $\P$ in that disjunct. Finally, the counting set
\begin{eqnarray*}
\F_{\Q'}(\A) &=& \mid \{ \langle\P\rangle \mid \A \models 
\exists \x \varphi'_1(\x,\P) \vee
\exists \x \varphi'_2(\x,\P) \vee
... \vee
\exists \x \varphi'_k(\x,\P)
 \} \mid
\end{eqnarray*}
counts exactly one assignation to $\P$ less than $F(\A)$.

Therefore,

% CASE 3

\item The counting set is:
\begin{eqnarray*}
F(\A) &=& \mid \{ \langle\P,\z\rangle \mid \A \models \exists \x \ \varphi(\x, \P,\z) \} \mid
\end{eqnarray*}
Then, we going to isolate the minimal predicate $\P$ that holds the formula true and eliminate the lexicographically smallest $\z$ that satisfies it. We mix both previous strategies.
\end{enumerate}
\end{proof}

\end{document}

### 2014-12-15 sharp-sigma-one-plus.tex
Solving problem with '

@@
 for every $\L$-structure $\A = \langle A, \S^\A, \leq^\A \rangle$, where $\x = (x_1,x_2,...,x_c)$, $\z = (z_1,z_2,...,z_d)$, and $\leq^\A$ is a total order for $A$.
-We refer to predicates in $\P$ and variables in $\z$ as "open", and predicates in $\S$ and variables in $\x$ as "closed".
+We refer to predicates in $\P$ and variables in $\z$ aa ``open", and predicates in $\S$ and variables in $\x$ as "closed".
 The $\#\Sigma_1^{+}$ class is defined as an extension of $\#\Sigma_1$ which includes the possibility of adding arbitrarily quantified $\L$-sub-formulas $\psi(\x,\z)$ which may contain closed predicates but no open ones.
@@
 \begin{multline*}
-\varphi'(\x,\P) = \bigwedge_{i=1}^n P_i(\vec{x_i}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\x) \wedge \varphi_{\S}(\x)\wedge\exists\vec{x_1}',...,\vec{x_n}'º\Bigg(\alpha_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge \\ \bigg(\bigvee_{i = 1}^{n}\neg P_i(\vec{x_i}') \vee \bigvee_{P\in\P} \exists \x''\Big( P(\x'') \wedge \bigwedge_{P_i = P} \x'' \neq \vec{x_i}'\Big) \bigg) \Bigg)
+\varphi'(\x,\P) = \bigwedge_{i=1}^n P_i(\vec{x_i}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\x) \wedge \varphi_{\S}(\x)\wedge\exists\vec{x_1}',...,\vec{x_n}' \Bigg(\alpha_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge \\ \bigg(\bigvee_{i = 1}^{n}\neg P_i(\vec{x_i}') \vee \bigvee_{P\in\P} \exists \x''\Big( P(\x'') \wedge \bigwedge_{P_i = P} \x'' \neq \vec{x_i}'\Big) \bigg) \Bigg)
 \end{multline*}
@@
 Let $\P^*$ be the excluded assignation to $\P$.
-Suppose there is another assignation $\P^*'$ to $\P$ that is also excluded by the new sub-formula.
-Then, $\x^*$ is such that every tuple $\x_i^*$ is in $\P_i^*'$ and $\x^*$ includes every tuple in $\P^*'$.
-If that's the case, $\P^*'$ contains exactly the same tuples as $\P^*$, and are thus the same.
+Suppose there is another assignation $\P^{*'}$ to $\P$ that is also excluded by the new sub-formula.
+Then, $\x^*$ is such that every tuple $\x_i^*$ is in $\P_i^{*'}$ and $\x^*$ includes every tuple in $\P^{*'}$.
+If that's the case, $\P^{*'}$ contains exactly the same tuples as $\P^*$, and are thus the same.
 

### 2014-12-18 sharp-sigma-one-plus.tex
18 dic

--- 2014-12-15 sharp-sigma-one-plus.tex
@@
 \def\L{{\cal L}}
-\def\Q{{\cal Q}}
 \def\F{{\cal F}}
 \def\P{\vec{P}}
+\def\Q{\vec{Q}}
 \def\S{\vec{S}}
+\def\s{\vec{s}}
+\def\u{\vec{u}}
+\def\v{\vec{v}}
+\def\w{\vec{w}}
 \def\x{\vec{x}}
+\def\y{\vec{y}}
 \def\z{\vec{z}}
@@
 
-We define the vocabulary $\L = \langle \P, \S, \leq \rangle$, where $\P$ is a $r$-tuple of predicates of arity $a_1,a_2,...,a_r$, and $\S$ is a $t$-tuple of predicates of arity $b_1,b_2,...,b_t$.
-A counting problem $\Q$ in $\#\Sigma_1$ is defined by a quantifier free $\L$-formula $\varphi(\x,\P,\z)$ where its function $\F_\Q(\A)$ is
+We define the vocabulary $\L = \{ P_1, ..., P_r, S_1,...,S_t, \leq \}$, where $P_1, ..., P_r$ have arity $a_1,...,a_r$, and $S_1,...,S_t$ have arity $b_1,...,b_t$.
+Let $\L\prime = \{ S_1,...,S_t, \leq \}$, and let
 \begin{eqnarray*}
-\F_\Q(\A) &=& \mid \{ \langle \P, \z \rangle \mid \A \models \exists \x \: \varphi(\x,\P,\z) \} \mid
+\textsc{Struct[\L]} &=& \{\A \mid \A \text{ is an } \L\prime \text{-structure with a finite domain } A \text{ such that} \\
+&& \leq \text{ is interpreted as a total order for } A \}.
 \end{eqnarray*}
-for every $\L$-structure $\A = \langle A, \S^\A, \leq^\A \rangle$, where $\x = (x_1,x_2,...,x_c)$, $\z = (z_1,z_2,...,z_d)$, and $\leq^\A$ is a total order for $A$.
-We refer to predicates in $\P$ and variables in $\z$ as "open", and predicates in $\S$ and variables in $\x$ as "closed".
-The $\#\Sigma_1^{+}$ class is defined as an extension of $\#\Sigma_1$ which includes the possibility of adding arbitrarily quantified $\L$-sub-formulas $\psi(\x,\z)$ which may contain closed predicates but no open ones.
-
-Some properties of the $\#\Sigma_1^{+}$ class are:
+A quantifier-free $\L$-formula is defined by the following grammar:
+\begin{eqnarray*}
+\varphi &::=& x = y \mid S_i(x_1,...,x_{b_i}), i \in \{1,...,t\} \mid x \leq y \mid \\
+&& P_i(x_1,...,x_{a_i}), i \in \{1,...,r\} \mid \\ 
+&& (\neg \varphi) \mid (\varphi \wedge \varphi) \mid (\varphi \vee \varphi),
+\end{eqnarray*}
+where $x, y$ are variables. We now define a quantifier-free$^{+}$ $\L$-formula as such:
+\begin{eqnarray*}
+\varphi &::=& \alpha, \alpha \text{ is a FO-formula over } \L\prime  \mid \\
+&& P_i(x_1,...,x_{a_i}), i \in \{1,...,r\} \mid \\
+&& (\neg \varphi) \mid (\varphi \wedge \varphi) \mid (\varphi \vee \varphi).
+\end{eqnarray*}
+Let $\P = (P_1, ..., P_r)$ and $\S = (S_1,...,S_t)$.
+A function $\F:\textsc{Struct[\L]} \to \mathbb{N}$ is in $\#\Sigma_1^+$ if there exists a quantifier-free$^{+}$ $\L$-formula $\varphi(\x,\P,\z)$ such that
+\begin{eqnarray*}
+\F &=& \mid \{ \langle \P, \z \rangle \mid \A \models \exists \x \: \varphi(\x,\P,\z) \} \mid
+\end{eqnarray*}
+for every $\A \in \textsc{Struct[\L]}$. We refer to predicates in $\P$ and variables in $\z$ as ``open", and predicates in $\S$ and variables in $\x$ as ``closed".
 
 \begin{theo}
-The decision version of a counting problem in $\#\Sigma_1^{+}$ is in \text{P}.
+The decision version of a function in $\#\Sigma_1^{+}$ is in \text{P}.
 \end{theo}
@@
 
-Then, we compute a similar counting function,
+Then, we compute a similar function,
 \begin{eqnarray*}
@@
 %% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 
-
+A counting class $\C$ is {\em closed under substraction by one} if for every function $\F \in \C$ there exists $\F' \in \C$ where 
+\begin{eqnarray*}
+\F'(\A) =
+\begin{cases}
+\F(\A)-1, & \text{if }\F(\A) > 0 \\
+0, & \text{if }\F(\A) = 0.
+\end{cases}
+\end{eqnarray*}
 \begin{theo}
-$\#\Sigma_1^+$ is closed under substraction\footnote[1]{A counting class $\C$ is closed under substraction if given a counting problem $\Q \in \C$ there exists $\Q' \in \C$ where $$
-\F_{\Q'}(\A) =
-\begin{cases}
-\F_\Q(\A)-1, & \text{if }\F_{\Q}(\A) > 0 \\
-0, & \text{if }\F_{\Q}(\A) = 0 
-\end{cases}
-$$ }.
+$\#\Sigma_1^+$ is closed under substraction by one
 \end{theo}
 \begin{proof}
-The counting set has three possible ways of counting. Counting only variables, only predicates, and predicates with variables. This separates the proof in three cases:
+A function in $\#\Sigma_1^+$ has three possible ways of counting. Counting only variables, only predicates, and predicates with variables. This separates the proof in three cases:
 \begin{enumerate}
@@
 
-\item Let $\Q \in \#\Sigma_1^+$, which is defined by an $\L$-formula $\varphi(\x,\z)$. Its function would be
+\item Let $\F \in \#\Sigma_1^+$, which is defined by an $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,...,z_d)$.:
 \begin{eqnarray*}
-\F_\Q(\A) &=& \mid \{ \langle\z\rangle \mid \A \models \exists \x \ \varphi(\x,\z) \} \mid.
+\F(\A) &=& \mid \{ \langle\z\rangle \mid \A \models \exists \x \ \varphi(\x,\z) \} \mid.
 \end{eqnarray*}
-Our goal here is to eliminate the lexicographically smallest sequence of variables, which we can do easily. Let $\F_{\Q'}$ be
+Our goal here is to eliminate the lexicographically smallest sequence of variables, which we can do easily. First, let $\y = (y_1,...,y_k)$, $\y\prime = (y_1\prime,...,y_k\prime)$ and
 \begin{eqnarray*}
-\F_{\Q'}(\A) &=& \mid \{ \langle\z\rangle \mid \A \models \exists \x\left( \varphi(\x,\z) \wedge \exists \vec{z'} (\varphi(\x,\vec{z'}) \wedge \varphi_<(\vec{z'},\z ) ) \right) \} \mid,
+\varphi_{k,<}(\y\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j = y_j\prime \wedge y_i < y_i\prime \right).
 \end{eqnarray*}
-where
+This formula is true if $\y\prime$ is lexicographically smaller than $\y$. Now, let $\F'$ be
 \begin{eqnarray*}
-\varphi_<(\vec{z'},\z) &=& \bigvee_{i = 1}^d \left( \bigwedge_{j=1}^{i-1} z_j = z_j' \wedge z_i < z_i' \right).
+\F'(\A) &=& \mid \{ \langle\z\rangle \mid \A \models \exists \x\left( \varphi(\x,\z) \wedge \exists \z\prime (\varphi(\x,\z\prime) \wedge \varphi_{d,<}(\z\prime,\z ) ) \right) \} \mid.
 \end{eqnarray*}
-This formula is true if $\vec{z'}$ is lexicographically smaller than $\z$. Thus, $\F_{\Q'}(\A)$ will count exactly one element less than $\F_\Q(\A)$ if $\F_\Q(\A)>0$.
+Thus, if $\F>0$, $\F'(\A)$ will count exactly one element less than $\F(\A)$, and if $\F(\A)=0$, $\A \not\models\varphi(\x,\z)$ so $\F'(\A)=0$.
 
@@
 
-\item Let $\Q \in \#\Sigma_1^+$, which is defined by an $\L$-formula $\varphi(\x,\P)$. Then
+\item Let $\F \in \#\Sigma_1^+$, which is defined by an $\L$-formula $\varphi(\x,\P)$ where $\x = (x_1,...,x_d)$. Then
 \begin{eqnarray*}
-\F_\Q(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid.
+\F &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid.
 \end{eqnarray*}
-Suppose that
+For the time being, suppose that
 \begin{eqnarray*}
-\varphi(\x,\P) &=& \bigwedge_{i=1}^n P_j(\vec{x_j}) \wedge \varphi_{\P}^{-}(\vec{y}) \wedge \theta(\x) \wedge \varphi_{\S}(\x)
+\varphi(\x,\P) &=& \bigwedge_{i=1}^n Q_j(\x_j) \wedge \varphi_{\P}^{-}(\y) \wedge \theta(\x) \wedge \varphi_{\S}(\x)
 \end{eqnarray*}
-where $P_1,P_2,...,P_n$ is a subsequence of $\P$, $P_i$ is of arity $a_i'$ and $\vec{x_i}$ is an $a_i'$-tuple of variables in $\x$ for all $i \in \{1,...,n\}$, $\theta(\x)$ defines a total order for $\x$, and $\varphi_{\S}(\x)$ involves only closed predicates.
-We can assume $(\vec{x_1},\vec{x_2},...,\vec{x_n},\vec{y}) = \x$ by adding equivalences to $\theta(\x)$.
-For example, if the function is defined by
+where $Q_i$ is in $\P$, $Q_i$ is of arity $c_i$ and $\x_i$ is an $c_i$-tuple of variables in $\x$ for all $i \in \{1,...,n\}$, $\y$ is a $k$-tuple of variables in $\x$, $\varphi_{\P}^{-}(\y)$ is a conjunction of negated predicates in $\P$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\varphi_{\S}(\x)$ involves only closed predicates.
+We also assume that $(\x_1,...,\x_n,\y) = \x$.
+For example,
 \begin{eqnarray*}
-\exists x,y \left( P(x,y) \wedge \neg P(x,x) \wedge x < y \wedge \forall z\big( S(x,z) \big) \right),
-\end{eqnarray*}
-we can use
-\begin{eqnarray*}
-\exists x_1,x_2,x_3,x_4 \left( P(x_1,x_2) \wedge \neg P(x_3,x_4) \wedge (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 ) \wedge \forall z\big( S(x_1,z) \big) \right).
+\exists x_1\exists x_2\exists x_3\exists x_4 \left( P(x_1,x_2) \wedge \neg P(x_3,x_4) \wedge (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 ) \wedge \forall z\big( S(x_1,z) \big) \right).
 \end{eqnarray*}
@@
 \begin{multline*}
-\alpha_{\min}(\vec{x_1}',...,\vec{x_n}') = \\ \exists\vec{y}\left( \alpha(\vec{x_1}',...,\vec{x_n}',\vec{y})\wedge \forall\vec{x_1}''...\vec{x_n}''\vec{y}'(\alpha(\vec{x_1}'',...,\vec{x_n}'',\vec{y}')\rightarrow \varphi_<((\vec{x_1}',...,\vec{x_n}'),(\vec{x_1}'',...,\vec{x_n}''))\right),
+\alpha_{\min}(\u_1,...,\u_n) = \\ \exists\y\left( \alpha(\u_1,...,\u_n,\y)\wedge \forall\v_1...\v_n\y\prime(\alpha(\v_1,...,\v_n,\y\prime)\rightarrow \varphi_{n,<}((\u_1,...,\u_n),(\v_1,...,\v_n))\right),
 \end{multline*}
 where $\alpha(\x) = \theta(\x) \wedge \varphi_{\S}(\x)$.
-Note that $\alpha_{\min}^i$ is satisfied only by the lexicographically smallest assignation $\x^*$ to $\x$ that satisfies both $\theta(\x)$ and $\varphi_{\S}(\x)$.
+Note that $\alpha_{\min}$ is satisfied only by the lexicographically smallest assignment $\sigma$ to $(\x_1,...,\x_n)$ such that $\theta(\sigma(\x_1),...,\sigma(\x_n),\y)$ and $\varphi_{\S}(\sigma(\x_1),...,\sigma(\x_n),\y)$ are true for some $\y$.
 Our new formula is
@@
 \begin{multline*}
-\varphi'(\x,\P) = \bigwedge_{i=1}^n P_i(\vec{x_i}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\x) \wedge \varphi_{\S}(\x)\wedge\exists\vec{x_1}',...,\vec{x_n}'º\Bigg(\alpha_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge \\ \bigg(\bigvee_{i = 1}^{n}\neg P_i(\vec{x_i}') \vee \bigvee_{P\in\P} \exists \x''\Big( P(\x'') \wedge \bigwedge_{P_i = P} \x'' \neq \vec{x_i}'\Big) \bigg) \Bigg)
+\varphi\prime(\x,\P) = \bigwedge_{i=1}^n Q_i(\x_i) \wedge \varphi^{-}_{\overline{P}}(\y) \wedge \theta(\x) \wedge \varphi_{\S}(\x)\wedge\exists\u_1,...,\u_n\Bigg(\alpha_{\min}(\u_1,...,\u_n) \wedge \\ \bigg(\bigvee_{i = 1}^{n}\neg P_i(\u_i) \vee \bigvee_{P\in\P} \exists \v\Big( P(\v) \wedge \bigwedge_{\substack{i\in\{1,...,n\} \\ P_i = P}} \v \neq \u_i\Big) \bigg) \Bigg)
 \end{multline*}
 
-where $\vec{x_i}'$ is a an $a_i$-tuple of variables in $\x$.
-If there is an assignation to satisfy $\alpha^j(\x)$, this formula will be satisfied by exactly one predicate less than $\exists \x \varphi_1(\x,\P)$.
-Let $\P^*$ be the excluded assignation to $\P$.
-Suppose there is another assignation $\P^*'$ to $\P$ that is also excluded by the new sub-formula.
-Then, $\x^*$ is such that every tuple $\x_i^*$ is in $\P_i^*'$ and $\x^*$ includes every tuple in $\P^*'$.
-If that's the case, $\P^*'$ contains exactly the same tuples as $\P^*$, and are thus the same.
+where $\u_i$ is a an $a_i$-tuple of variables in $\x$.
+We will show that if there is an assignation to satisfy $\alpha(\x)$, this formula will be satisfied by exactly one predicate less than $\exists \x \: \varphi(\x,\P)$.
+Suppose $\sigma$ is such an assignation.
+Let $\pi$ be the excluded assignation to $\P$.
+Suppose there is another assignation $\pi\prime$ to $\P$ that is also excluded by the new sub-formula.
+Then, $\sigma(\x)$ is such that every tuple $\s_i = \sigma(x_i)$ is in $\pi\prime(\P)$ and $\sigma(x)$ includes every tuple in $\pi\prime(\P)$.
+If that is the case, $\pi\prime(\P)$ contains exactly the same tuples as $\pi(\P)$, and are thus the same.
+Naturally, if $\varphi(\x,\P)$ is unsatisfiable for every assignation to $\x$ and $\P$, $\varphi\prime(\x,\P)$ also is.
 
@@
 \begin{eqnarray*}
-\F_\Q(\A) &=& \mid \{ \langle\P\rangle :\A \models 
+\F(\A) &=& \mid \{ \langle\P\rangle \mid \A \models 
 \exists \x \varphi_1(\x,\P) \vee
@@
 \end{eqnarray*}
-where $\varphi(\x,\P) \equiv \varphi_1(\x,\P) \vee \varphi_2(\x,\P) \vee ...  \vee \varphi_k(\x,\P) $.
-
+where $\varphi(\x,\P) \equiv \varphi_1(\x,\P) \vee \varphi_2(\x,\P) \vee ...
+\vee \varphi_k(\x,\P) $.
 Now we need to exclude that assignation to the next disjuncts, which we do by adding the same sub-formula.
@@
 \forall\x\left(\neg\alpha^1_{\min}(\x)\right) \to \exists\vec{x_1}',...,\vec{x_n}'\Bigg(\alpha^2_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge \\
-\bigg(\bigvee_{P\in\P}\bigvee_{\substack{i\in\{1,...,n\}\\P_i = P}}\neg P(\vec{x_i}') \vee \bigvee_{P\in\P} \exists \x''\Big( P(\x'') \bigwedge_{P_i = P} \x'' \neq \vec{x_i}'\Big) \bigg) \Bigg)
+\bigg(\bigvee_{i = 1}^{n}\neg P(\vec{x_i}') \vee \bigvee_{P\in\P} \exists \x''\Big( P(\x'') \bigwedge_{\substack{i\in\{1,...,n\} \\ P_i = P}} \x'' \neq \vec{x_i}'\Big) \bigg) \Bigg)
 \end{multline*}
@@
 \begin{multline*}
-\varphi_j'(\x,\P) = \bigwedge_{i=1}^n P_i(\vec{x_i}) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\x) \wedge \varphi_{\S}(\x) \wedge \Bigg(\Big(\forall\x\left(\neg\alpha^1_{\min}(\x)\right)\wedge ...
+\varphi_j\prime(\x,\P) = \bigwedge_{i=1}^n P_i(\x_i) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\x) \wedge \varphi_{\S}(\x) \wedge \Bigg(\Big(\forall\x\left(\neg\alpha^1_{\min}(\x)\right)\wedge ...
 \wedge \forall\x\left(\neg\alpha^{j-1}_{\min}(\x)\right)\Big) \to \\
-\exists\vec{x_1}',...,\vec{x_n}'\Bigg(\alpha^j_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge
-\bigg(\bigvee_{P\in\P}\bigvee_{\substack{i\in\{1,...,n\}\\P_i = P}}\neg P(\vec{x_i}') \vee \bigvee_{P\in\P} \exists \x''\Big( P(\x'') \bigwedge_{P_i = P} \x'' \neq \vec{x_i}'\Big) \bigg) \Bigg) \Bigg)
+\exists\u_1,...,\u_n\Bigg(\alpha^j_{\min}(\u_1,...,\u_n) \wedge
+\bigg(\bigvee_{i = 1}^n\neg P(\u_i) \vee \bigvee_{P\in\P} \exists \v\Big( P(\v) \bigwedge_{\substack{i\in\{1,...,n\} \\ P_i = P}} \v \neq \u_i\Big) \bigg) \Bigg) \Bigg)
 \end{multline*}
-which, if all of the previous disjuncts have no possible assignations, eliminates the least assignation to $\P$ in that disjunct. Finally, the counting set
+which, if all of the previous disjuncts have no possible assignations, eliminates the least assignation to $\P$ in that disjunct. Finally,
 \begin{eqnarray*}
-\F_{\Q'}(\A) &=& \mid \{ \langle\P\rangle \mid \A \models 
-\exists \x \varphi'_1(\x,\P) \vee
-\exists \x \varphi'_2(\x,\P) \vee
+\F\prime(\A) &=& \mid \{ \langle\P\rangle \mid \A \models 
+\exists \x \varphi\prime_1(\x,\P) \vee
+\exists \x \varphi\prime_2(\x,\P) \vee
 ... \vee
-\exists \x \varphi'_k(\x,\P)
+\exists \x \varphi\prime_k(\x,\P)
  \} \mid

### 2014-12-18 sharp-sigma-one-plus.tex

@@
 \begin{eqnarray*}
-\textsc{Struct[\L]} &=& \{\A \mid \A \text{ is an } \L\prime \text{-structure with a finite domain } A \text{ such that} \\
+\textsc{Struct}[\L] &=& \{\A \mid \A \text{ is an } \L\prime \text{-structure with a finite domain } A \text{ such that} \\
 && \leq \text{ is interpreted as a total order for } A \}.

### 2014-12-22 sharp-sigma-one-plus.tex
22 dic

@@
 \def\C{{\cal C}}
+\def\F{{\cal F}}
 \def\L{{\cal L}}
-\def\F{{\cal F}}
+\def\N{\mathbb{N}}
 \def\P{\vec{P}}
@@
 \def\S{\vec{S}}
+\def\X{\vec{X}}
+\def\Y{\vec{Y}}
+\def\Z{\vec{Z}}
+\def\d{\vec{d}}
+\def\e{\vec{e}}
 \def\s{\vec{s}}
@@
 
-We define the vocabulary $\L = \{ P_1, ..., P_r, S_1,...,S_t, \leq \}$, where $P_1, ..., P_r$ have arity $a_1,...,a_r$, and $S_1,...,S_t$ have arity $b_1,...,b_t$.
-Let $\L\prime = \{ S_1,...,S_t, \leq \}$, and let
+We define the vocabulary $\L = \{ S_1,...,S_t, \leq \}$, where $S_1,...,S_t$ have arity $b_1,...,b_t$. Let
 \begin{eqnarray*}
-\textsc{Struct}[\L] &=& \{\A \mid \A \text{ is an } \L\prime \text{-structure with a finite domain } A \text{ such that} \\
+\textsc{Struct}[\L] &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
 && \leq \text{ is interpreted as a total order for } A \}.
 \end{eqnarray*}
-A quantifier-free $\L$-formula is defined by the following grammar:
+We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$. A quantifier-free $\L$-formula is defined by the following grammar:
 \begin{eqnarray*}
-\varphi &::=& x = y \mid S_i(x_1,...,x_{b_i}), i \in \{1,...,t\} \mid x \leq y \mid \\
-&& P_i(x_1,...,x_{a_i}), i \in \{1,...,r\} \mid \\ 
-&& (\neg \varphi) \mid (\varphi \wedge \varphi) \mid (\varphi \vee \varphi),
+\varphi &::=& x = y \ \mid \ S_i(x_1,...,x_{b_i}), i \in \{1,...,t\} \ \mid \ x \leq y \ \mid \\
+&& X_i(x_1,...,x_{a_i}), i\in\N \ \mid \\ 
+&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
 \end{eqnarray*}
-where $x, y$ are variables. We now define a quantifier-free$^{+}$ $\L$-formula as such:
+where $y$ and $x_i$ are first order variables for every $i$. We now define an extended quantifier-free $\L$-formula as such:
 \begin{eqnarray*}
-\varphi &::=& \alpha, \alpha \text{ is a FO-formula over } \L\prime  \mid \\
-&& P_i(x_1,...,x_{a_i}), i \in \{1,...,r\} \mid \\
-&& (\neg \varphi) \mid (\varphi \wedge \varphi) \mid (\varphi \vee \varphi).
+\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L^\prime  \ \mid \\
+&& X_i(x_1,...,x_{a_i}), i\in\N \ \mid \ \\
+&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
 \end{eqnarray*}
-Let $\P = (P_1, ..., P_r)$ and $\S = (S_1,...,S_t)$.
-A function $\F:\textsc{Struct[\L]} \to \mathbb{N}$ is in $\#\Sigma_1^+$ if there exists a quantifier-free$^{+}$ $\L$-formula $\varphi(\x,\P,\z)$ such that
+Let $\X = (X_1,...,X_r)$ and $\P = (P_1,...,P_r)$, where $P_i$ is a predicate of arity $a_i$, for every $i$.
+Let $\z$ be a tuple of variables and $\e$ be a tuple of constants with the same dimension as $\z$.
+A function $f:\textsc{Struct}[\L] \to \mathbb{N}$ is in $\#\Sigma_1^+$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\X,\z)$ such that
 \begin{eqnarray*}
-\F &=& \mid \{ \langle \P, \z \rangle \mid \A \models \exists \x \: \varphi(\x,\P,\z) \} \mid
+f(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \exists \x \: \varphi(\x,\P,\e) \} \mid,
 \end{eqnarray*}
-for every $\A \in \textsc{Struct[\L]}$. We refer to predicates in $\P$ and variables in $\z$ as ``open", and predicates in $\S$ and variables in $\x$ as ``closed".
+for every $\A \in \textsc{Struct}[\L]$.
 
@@
 \begin{proof}
-Let $F$ be a function in $\#\Sigma_1^{+}$. Then there is a formula $\varphi$ such that
+Let $f$ be a function in $\#\Sigma_1^{+}$. Then there is a formula $\varphi(\x,\X,\z)$ such that
 \begin{eqnarray*}
-F(\A) &=& \mid \{ \langle\P,\z\rangle \mid \A \models \exists \x \varphi(\S,\x,\z) \} \mid,
+f(\A) &=& \mid \{ \langle\P,\z\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\z) \} \mid,
 \end{eqnarray*}
-where $\A$ is a finite ordered structure and $\z$ is an $m$-tuple.
-For each $\z \in A^m$, every arbitrarily quantified sub-formula which does not include $\P$ can be evaluated in polynomial time.
-Let $\varphi'$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula by a contradiction.
-Note that $\varphi'$ is quantifier-free.
-
+where $\A \in \textsc{Struct}[\L]$ and $\z$ is an $m$-tuple.
+For each $\z \in A^m$, every arbitrarily quantified first order sub-formula can be evaluated in polynomial time.
+Let $\varphi^\prime(\x,\P,\z)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula by a contradiction.
+Note that $\varphi^\prime(\x,\P,\z)$ is quantifier-free.
 Then, we compute a similar function,
 \begin{eqnarray*}
-F'(\A) &=& \mid \{ \langle\P,(\z,\x)\rangle \mid \A \models \varphi(\S,\x,\z) \} \mid
+f'(\A) &=& \mid \{ \langle\P,(\z,\x)\rangle \mid \A \models \varphi(\x,\P,\z) \} \mid
 \end{eqnarray*}
-which is in $\#\Sigma_0$.
-Note that for every $\A$, $F(\A) > 0$ iff $F'(\A) > 0$, so computing $F'$ is enough to solve the decision version of $F$.
-However, Saluja and Subrahmanyam showed that any counting problem in $\#\Sigma_0$ is computable in polynomial time \footnote[1]{(paper)}, therefore, the decision version of $F$ is in P.
+which is in $\#\Sigma_0$\footnote[1]{(paper)}.
+Note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
+However, it is showed in $^{[1]}$ that any counting problem in $\#\Sigma_0$ is computable in polynomial time, therefore, the decision version of $f$ is in P.
 \end{proof}
@@
 This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete.
-However, as we showed previously (Theorem 1), it's also in P.
+However, as we showed previously (Theorem 1), it is also in P.
 Then, co-NP $\subseteq$ P.
@@
 %% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1
-A counting class $\C$ is {\em closed under substraction by one} if for every function $\F \in \C$ there exists $\F' \in \C$ where 
+
+A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$ there exists $f' \in \F$ such that 
 \begin{eqnarray*}
-\F'(\A) =
+f'(\A) =
 \begin{cases}
-\F(\A)-1, & \text{if }\F(\A) > 0 \\
-0, & \text{if }\F(\A) = 0.
+f(\A)-1, & \text{if }f(\A) > 0 \\
+0, & \text{if }f(\A) = 0.
 \end{cases}
@@
 \begin{theo}
-$\#\Sigma_1^+$ is closed under substraction by one
+$\#\Sigma_1^+$ is closed under substraction by one.
 \end{theo}
@@
 
-\item Let $\F \in \#\Sigma_1^+$, which is defined by an $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,...,z_d)$.:
+\item Let $f \in \#\Sigma_1^+$, which is defined by an $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,...,z_d)$. That is,
 \begin{eqnarray*}
-\F(\A) &=& \mid \{ \langle\z\rangle \mid \A \models \exists \x \ \varphi(\x,\z) \} \mid.
+f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
 \end{eqnarray*}
-Our goal here is to eliminate the lexicographically smallest sequence of variables, which we can do easily. First, let $\y = (y_1,...,y_k)$, $\y\prime = (y_1\prime,...,y_k\prime)$ and
+for every $A \in \textsc{Struct}[\L]$.
+Our goal here is to eliminate the lexicographically smallest sequence of variables, which we can do easily.
+First, let $\y = (y_1,...,y_k)$, $\y\,^\prime = (y_1^\prime,...,y_k^\prime)$ and
 \begin{eqnarray*}
-\varphi_{k,<}(\y\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j = y_j\prime \wedge y_i < y_i\prime \right).
+\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j = y_j^\prime \wedge y_i < y_i^\prime \right).
 \end{eqnarray*}
-This formula is true if $\y\prime$ is lexicographically smaller than $\y$. Now, let $\F'$ be
+This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
 \begin{eqnarray*}
-\F'(\A) &=& \mid \{ \langle\z\rangle \mid \A \models \exists \x\left( \varphi(\x,\z) \wedge \exists \z\prime (\varphi(\x,\z\prime) \wedge \varphi_{d,<}(\z\prime,\z ) ) \right) \} \mid.
+\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
 \end{eqnarray*}
-Thus, if $\F>0$, $\F'(\A)$ will count exactly one element less than $\F(\A)$, and if $\F(\A)=0$, $\A \not\models\varphi(\x,\z)$ so $\F'(\A)=0$.
+Therefore, if $f(\A)>0$, $f'(\A)$ will count exactly one element less than $f(\A)$; and if $f(\A)=0$, $\A \not\models\exists x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists x\,\varphi^\prime(\x,\e)$ for every $\e$.
+Thus, $f'(\A)=0$.
 
@@
 
-\item Let $\F \in \#\Sigma_1^+$, which is defined by an $\L$-formula $\varphi(\x,\P)$ where $\x = (x_1,...,x_d)$. Then
+\item Let $f \in \#\Sigma_1^+$, which is defined by an $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,...,x_d)$ and $\X = (X_1,...,X_r)$. Then
 \begin{eqnarray*}
-\F &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid.
+f &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid,
 \end{eqnarray*}
-For the time being, suppose that
+for every $\A \in \textsc{Struct}[\L]$. For the time being, suppose that
 \begin{eqnarray*}
-\varphi(\x,\P) &=& \bigwedge_{i=1}^n Q_j(\x_j) \wedge \varphi_{\P}^{-}(\y) \wedge \theta(\x) \wedge \varphi_{\S}(\x)
+\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n Y_i(\x_i) \right) \wedge \varphi_{\X}^{-}(\y) \wedge \theta(\x) \wedge \beta(\x)
 \end{eqnarray*}
-where $Q_i$ is in $\P$, $Q_i$ is of arity $c_i$ and $\x_i$ is an $c_i$-tuple of variables in $\x$ for all $i \in \{1,...,n\}$, $\y$ is a $k$-tuple of variables in $\x$, $\varphi_{\P}^{-}(\y)$ is a conjunction of negated predicates in $\P$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\varphi_{\S}(\x)$ involves only closed predicates.
+where $Y_i$ is in $\X$, $Y_i$ is of arity $c_i$ and $\x_i$ is an $c_i$-tuple of variables in $\x$ for all $i \in \{1,...,n\}$, $\y$ is a $d$-tuple of variables in $\x$, $\varphi_{\X}^{-}(\y)$ is a conjunction of negated predicates in $\P$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$.
 We also assume that $(\x_1,...,\x_n,\y) = \x$.
-For example,
+As an example, we will show that the following formula is of this form:
 \begin{eqnarray*}
-\exists x_1\exists x_2\exists x_3\exists x_4 \left( P(x_1,x_2) \wedge \neg P(x_3,x_4) \wedge (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 ) \wedge \forall z\big( S(x_1,z) \big) \right).
+\varphi^{*}(\x,\X) &=& \left( X_1(x_1,x_2) \wedge \neg X_1(x_3,x_4) \wedge (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 ) \wedge \forall z\big( S_1(x_1,z) \big) \right),
 \end{eqnarray*}
-Similarly, we would like to eliminate the lexicographically smallest predicate $P$ that satisfies the formula. We now define
+where $\x = (x_1,x_2,x_3,x_4)$ and $\X = (X_1)$.
+Here, $n = 1$, $Y_1 = X_1$, $\x_1 = (x_1,x_2)$ and $\y = (x_3,x_4)$.
+$\varphi_{\X}^{-}(\y) = \neg X_1(x_3,x_4)$, $\theta^*(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4)$ which defines a total order on the partition $\{\{x_1,x_3,x_4\},\{x_2\}\}$, and $\beta^*(\x) = \forall z\big( S_1(x_1,z) \big)$.
+
+Similarly to the previous proof, we would like to eliminate the lexicographically smallest tuple of predicates that satisfies the formula.
+Let $\u_i$ be a $c_i$-tuple of variables for every $i \in \{1,...,n\}$, and let $m = \sum_{i = i}^n c_i$ be the number of variables of $(\x_1,...,\x_n)$.
+We now define
 \begin{multline*}
-\alpha_{\min}(\u_1,...,\u_n) = \\ \exists\y\left( \alpha(\u_1,...,\u_n,\y)\wedge \forall\v_1...\v_n\y\prime(\alpha(\v_1,...,\v_n,\y\prime)\rightarrow \varphi_{n,<}((\u_1,...,\u_n),(\v_1,...,\v_n))\right),
+\alpha_{\min}(\u_1,...,\u_n) = \\ \exists\y\left( \alpha(\u_1,...,\u_n,\y)\wedge \forall\v_1...\v_n\w(\alpha(\v_1,...,\v_n,\w)\rightarrow \varphi_{m,<}((\u_1,...,\u_n),(\v_1,...,\v_n))\right),
 \end{multline*}
-where $\alpha(\x) = \theta(\x) \wedge \varphi_{\S}(\x)$.
-Note that $\alpha_{\min}$ is satisfied only by the lexicographically smallest assignment $\sigma$ to $(\x_1,...,\x_n)$ such that $\theta(\sigma(\x_1),...,\sigma(\x_n),\y)$ and $\varphi_{\S}(\sigma(\x_1),...,\sigma(\x_n),\y)$ are true for some $\y$.
+where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
+Note that $\alpha_{\min}$ is satisfied only by the lexicographically \linebreak smallest assignment $\d$ to $(x_1,...,x_n)$ such that $\exists\y\,(\theta(d_1,...,d_n,\y)\wedge\beta(d_1,...,d_n,\y))$ is true.
 Our new formula is
+\begin{multline*}
+\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n Y_i(\x_i) \right) \wedge \varphi_{\X}^{-}(\y) \wedge \theta(\x) \wedge \beta(\x)\wedge\exists\u_1,...,\u_n\bigg[\alpha_{\min}(\u_1,...,\u_n) \wedge \\ \bigg(\bigvee_{i = 1}^{n}\neg Y_i(\u_i) \vee \bigvee_{Y\in\X} \exists \v\Big( Y(\v) \wedge \bigwedge_{i\in[1,n]: Y_i = Y} \v \neq \u_i\Big) \bigg) \bigg].
+\end{multline*}
+We will show that if there is an assignment to satisfy $\exists\x\,\varphi(\x,\X)$, this formula will be satisfied by exactly one predicate less than $\exists\x\,\varphi(\x,\X)$.
+Suppose $\d$ is the lexicographically smallest assignment which satisfies $\alpha(\x)$.
+Let $\d_i$ be the respective assignment to $x_i$, for every $i\in\{1,...,n\}$.
+Consider now the tuple $\P = (P_i,...,P_r)$ where $P_i = \{d_j \mid j\in\{1,...,n\} \text{ and } X_i = Y_j \}$.
+We will now show that this tuple of predicates is such that $\exists\x\,\varphi(\x,\P)$ is true.
+Suppose that it's not, that is, there is no assignation to $\x$ such that $\varphi(\x,\P)$ is true.
+Since $\d$ satisfies $\left( \bigwedge_{i=1}^n Y_i(\x_i) \right)$ and $\theta(\x) \wedge \beta(\x)$, then $\varphi_{\X}^{-}(\d)$ must be false.
+Therefore, there is a negated predicate $Y$ in $\varphi_{\X}^{-}(\x)$ such that its respective variables in $\y$ are equal to an $x_i$ where $Y_i = Y$.
+Then, since $\theta(\x)$ defines a partition of $\x$, these variables are equal, and thus, $\exists\x\,\varphi(\x,\X)$ is unsatisfiable ($\rightarrow\leftarrow$).
+$\P$ is excluded by the new formula since every $d_i$ is in its respective predicate, which does not satisfy the first part of the new formula, and does not have any tuples that are not in $\d$.
+$\P$ is also the only tuple excluded by the new formula since any other tuple that satisfies it would have at least the same tuples as $\P$, and would not have any more, and thus would be equal to $\P$.
+Therefore, if $\exists\x\,\varphi(\x,\X)$ is satisfiable, $\exists\x\,\varphi^\prime(\x,\X)$ is satisfied by exactly one predicate less.
+Naturally, if $\varphi(\x,\X)$ is unsatisfiable for every assignment to $\x$ and $\X$, $\varphi^\prime(\x,\X)$ also is.
 
-\begin{multline*}
-\varphi\prime(\x,\P) = \bigwedge_{i=1}^n Q_i(\x_i) \wedge \varphi^{-}_{\overline{P}}(\y) \wedge \theta(\x) \wedge \varphi_{\S}(\x)\wedge\exists\u_1,...,\u_n\Bigg(\alpha_{\min}(\u_1,...,\u_n) \wedge \\ \bigg(\bigvee_{i = 1}^{n}\neg P_i(\u_i) \vee \bigvee_{P\in\P} \exists \v\Big( P(\v) \wedge \bigwedge_{\substack{i\in\{1,...,n\} \\ P_i = P}} \v \neq \u_i\Big) \bigg) \Bigg)
-\end{multline*}
-
-where $\u_i$ is a an $a_i$-tuple of variables in $\x$.
-We will show that if there is an assignation to satisfy $\alpha(\x)$, this formula will be satisfied by exactly one predicate less than $\exists \x \: \varphi(\x,\P)$.
-Suppose $\sigma$ is such an assignation.
-Let $\pi$ be the excluded assignation to $\P$.
-Suppose there is another assignation $\pi\prime$ to $\P$ that is also excluded by the new sub-formula.
-Then, $\sigma(\x)$ is such that every tuple $\s_i = \sigma(x_i)$ is in $\pi\prime(\P)$ and $\sigma(x)$ includes every tuple in $\pi\prime(\P)$.
-If that is the case, $\pi\prime(\P)$ contains exactly the same tuples as $\pi(\P)$, and are thus the same.
-Naturally, if $\varphi(\x,\P)$ is unsatisfiable for every assignation to $\x$ and $\P$, $\varphi\prime(\x,\P)$ also is.
+%% Let $\P$ and $\P^\prime$ be tuples of predicates such that $\exists\x\,\varphi(\x,\P)$ and $\exists\x\,\varphi(\x,\P^\prime)$ are true but $\exists\x\,\varphi^\prime(\x,\P)$ and $\exists\x\,\varphi^\prime(\x,\P^\prime)$ are not.
+Let $Q_i$ and $Q_i^\prime$ be the assignment of $\P$ and $\P^\prime$ to $Y_i$, respectively, for every $i\in\{i,...n\}$.
+Then, every tuple $\d_i$ is in its respective relation $Q_i$ in both $\P$ and $\P^\prime$ and $(d_1,...,d_n)$ includes every tuple in $P^\prime$.
+If that is the case, $\P^\prime$ contains exactly the same tuples as $\P$, and are thus the same.
 
@@
 \begin{eqnarray*}
-\F(\A) &=& \mid \{ \langle\P\rangle \mid \A \models 
+f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models 
 \exists \x \varphi_1(\x,\P) \vee
@@
 \vee \varphi_k(\x,\P) $.
-Now we need to exclude that assignation to the next disjuncts, which we do by adding the same sub-formula.
-We also need to include the case where there are no possible assignations to $\x$ for the first disjunct, so we add the sub-formula
+This is possible if we consider all first order formulas as literals, and find an equivalent DNF formula.
+Now we need to exclude that assignment to the next disjuncts, which we do by adding the same sub-formula.
+We also need to include the case where there are no possible assignments to $\x$ for the first disjunct, so we add this sub-formula to $\varphi^2(\x,\X)$:
 \begin{multline*}
-\forall\x\left(\neg\alpha^1_{\min}(\x)\right) \to \exists\vec{x_1}',...,\vec{x_n}'\Bigg(\alpha^2_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge \\
-\bigg(\bigvee_{i = 1}^{n}\neg P(\vec{x_i}') \vee \bigvee_{P\in\P} \exists \x''\Big( P(\x'') \bigwedge_{\substack{i\in\{1,...,n\} \\ P_i = P}} \x'' \neq \vec{x_i}'\Big) \bigg) \Bigg)
+\exists\x\left(\varphi^1(\x,\X)\right) \vee \exists\vec{x_1}',...,\vec{x_n}'\Bigg(\alpha^2_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge \\
+\bigg(\bigvee_{i = 1}^{n}\neg Y^2_i(\vec{x_i}') \vee \bigvee_{P\in\P} \exists \x''\Big( P(\x'') \bigwedge_{\substack{i\in\{1,...,n\} \\ P_i = P}} \x'' \neq \vec{x_i}'\Big) \bigg) \Bigg).
 \end{multline*}
@@
 \begin{multline*}
-\varphi_j\prime(\x,\P) = \bigwedge_{i=1}^n P_i(\x_i) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\x) \wedge \varphi_{\S}(\x) \wedge \Bigg(\Big(\forall\x\left(\neg\alpha^1_{\min}(\x)\right)\wedge ...
+\varphi_j^\prime(\x,\X) = \bigwedge_{i=1}^n Y_i(\x_i) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\x) \wedge \varphi_{\S}(\x) \wedge \Bigg(\Big(\forall\x\left(\neg\alpha^1_{\min}(\x)\right)\wedge ...
 \wedge \forall\x\left(\neg\alpha^{j-1}_{\min}(\x)\right)\Big) \to \\
@@
 \end{multline*}
-which, if all of the previous disjuncts have no possible assignations, eliminates the least assignation to $\P$ in that disjunct. Finally,
+which, if all of the previous disjuncts have no possible assignments, eliminates the least assignment to $\P$ in that disjunct. Finally,
 \begin{eqnarray*}
-\F\prime(\A) &=& \mid \{ \langle\P\rangle \mid \A \models 
-\exists \x \varphi\prime_1(\x,\P) \vee
-\exists \x \varphi\prime_2(\x,\P) \vee
+f^\prime(\A) &=& \mid \{ \langle\P\rangle \mid \A \models 
+\exists \x \varphi^\prime_1(\x,\P) \vee
+\exists \x \varphi^\prime_2(\x,\P) \vee
 ... \vee
-\exists \x \varphi\prime_k(\x,\P)
+\exists \x \varphi^\prime_k(\x,\P)
  \} \mid
 \end{eqnarray*}
-counts exactly one assignation to $\P$ less than $F(\A)$.
+counts exactly one assignment to $\P$ less than $F(\A)$.
 

### 2014-12-24 sharp-sigma-one-plus.tex
24 dic

@@
 \def\Q{\vec{Q}}
+\def\R{\vec{R}}
 \def\S{\vec{S}}
@@
 \def\Z{\vec{Z}}
-\def\d{\vec{d}}
-\def\e{\vec{e}}
+%% a - arity of \X / arity of assignations \P to \X
+%% b - arity of predicates in \S
+%% c - arity of auxiliar predicates/variables
+\def\d{\vec{d}} %% counted elements
+\def\e{\vec{e}} %% counted elements
+%% f - counting function
+%% g - other functions
+%% h - other functions
+%% i - index
+%% j - index
+%% k - emergency index / size of tuple
+%% l - emergency index / size of tuple
+%% m - size of variable tuple
+%% n - size of predicate tuple
+%% o - not used
+%% p - 
+%% q - 
+%% r - size of \X / \P
 \def\s{\vec{s}}
-\def\u{\vec{u}}
-\def\v{\vec{v}}
-\def\w{\vec{w}}
-\def\x{\vec{x}}
-\def\y{\vec{y}}
-\def\z{\vec{z}}
+%% t - size of \S
+\def\u{\vec{u}} %% auxiliary variables
+\def\v{\vec{v}} %% auxiliary variables
+\def\w{\vec{w}} %% auxiliary variables
+\def\x{\vec{x}} %% quantified variables
+\def\y{\vec{y}} %% auxiliary variables
+\def\z{\vec{z}} %% open variables
+\def\ep{\vec{\varepsilon}}
+\def\ga{\vec{\gamma}}
 
@@
 
-We define the vocabulary $\L = \{ S_1,...,S_t, \leq \}$, where $S_1,...,S_t$ have arity $b_1,...,b_t$. Let
+We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$. A quantifier-free $\L$-formula is defined by the following grammar:
+We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in N$ there are infinite variables in ${\cal X}$ of arity $n$.
+A quantifier-free $\L$-formula is defined by the following grammar:
 \begin{eqnarray*}
-\varphi &::=& x = y \ \mid \ S_i(x_1,...,x_{b_i}), i \in \{1,...,t\} \ \mid \ x \leq y \ \mid \\
-&& X_i(x_1,...,x_{a_i}), i\in\N \ \mid \\ 
+\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
+&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
 && (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
 \end{eqnarray*}
-where $y$ and $x_i$ are first order variables for every $i$. We now define an extended quantifier-free $\L$-formula as such:
+where $x,y$ and $x_i$ are first order variables for every $i$. We now define an extended quantifier-free $\L$-formula as follows:
 \begin{eqnarray*}
-\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L^\prime  \ \mid \\
-&& X_i(x_1,...,x_{a_i}), i\in\N \ \mid \ \\
+\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
+&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
 && (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
 \end{eqnarray*}
-Let $\X = (X_1,...,X_r)$ and $\P = (P_1,...,P_r)$, where $P_i$ is a predicate of arity $a_i$, for every $i$.
-Let $\z$ be a tuple of variables and $\e$ be a tuple of constants with the same dimension as $\z$.
+Let $\X = (X_1,\dots,X_r)$ and $\P = (P_1,\dots,P_r)$, where $P_i$ is a predicate of arity $a_i$, for every $i$.
+Let $\z$ be a tuple of variables and $\e$ be a tuple of constants with the same number of elements as $\z$.
 A function $f:\textsc{Struct}[\L] \to \mathbb{N}$ is in $\#\Sigma_1^+$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\X,\z)$ such that
@@
 \end{eqnarray*}
-for every $\A \in \textsc{Struct}[\L]$.
+for every $\A \in \textsc{Struct}[\L]$.\\
 
+The {\em decision version} of a function $f$ is the language $L_f = \{\A \mid f(\A) > 0\}$.
 \begin{theo}
-The decision version of a function in $\#\Sigma_1^{+}$ is in \text{P}.
+The decision version of a function in $\#\Sigma_1^{+}$ is in \textsc{P}.
 \end{theo}
@@
 \begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\z\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\z) \} \mid,
+f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
 \end{eqnarray*}
-where $\A \in \textsc{Struct}[\L]$ and $\z$ is an $m$-tuple.
-For each $\z \in A^m$, every arbitrarily quantified first order sub-formula can be evaluated in polynomial time.
-Let $\varphi^\prime(\x,\P,\z)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula by a contradiction.
-Note that $\varphi^\prime(\x,\P,\z)$ is quantifier-free.
-Then, we compute a similar function,
+where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \textsc{Struct}[\L]$, $\z$ is an $m$-tuple of variables, $\e$ is an $m$-tuple of elements and $\x$ is a $k$-tuple of variables.
+Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible evaluations for $\z$.
+Let $\ell = \vert A \vert^k$ and $\d_1,\dots,d_\ell\in A^k$ be all possible evaluations for $\x$.
+Let $\varphi_\A(\X)$ be defined as follows:
 \begin{eqnarray*}
-f'(\A) &=& \mid \{ \langle\P,(\z,\x)\rangle \mid \A \models \varphi(\x,\P,\z) \} \mid
+\varphi_\A(\X) = \varphi(\d_1,\X,\e_1) \vee \cdots \vee \varphi(\d_1,\X,\e_n) \vee \varphi(\d_2,\X,\e_1) \vee \cdots \vee \varphi(\d_n,\X,\e_n).
 \end{eqnarray*}
-which is in $\#\Sigma_0$\footnote[1]{(paper)}.
-Note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
+Note that $\varphi_\A(\X)$ will have at least one assignation for $\X$ iff $f(\A)>0$.
+Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction.
+Note that $\psi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$.
+Then, let $f'$ be defined as follows:
+\begin{eqnarray*}
+f'(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \psi_\A(\P) \} \mid.
+\end{eqnarray*}
+Note that $f' \in\#\Sigma_0$ \footnote[1]{(paper)}.
+Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
 However, it is showed in $^{[1]}$ that any counting problem in $\#\Sigma_0$ is computable in polynomial time, therefore, the decision version of $f$ is in P.
@@
 \begin{proof}
-We will show that the $\#\Sigma_1^{+}$ function defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall x_2 S(x_2)$ is not in $\#\Sigma_1$.
+We will show that the $\#\Sigma_1^{+}$ function defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
 \end{proof}
@@
 
-A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$ there exists $f' \in \F$ such that 
+A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, there exists $f' \in \F$ such that 
 \begin{eqnarray*}
@@
 \begin{proof}
-A function in $\#\Sigma_1^+$ has three possible ways of counting. Counting only variables, only predicates, and predicates with variables. This separates the proof in three cases:
+A function in $\#\Sigma_1^+$ has three possible ways of counting: counting only variables, only predicates, and predicates with variables.
+Let us separate each of this cases on the classes $\#\Sigma_1^{+(a)}$, $\#\Sigma_1^{+(b)}$ and $\#\Sigma_1^{+(c)}$ respectively:
 \begin{enumerate}
@@
 
-\item Let $f \in \#\Sigma_1^+$, which is defined by an $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,...,z_d)$. That is,
+\item Let $f \in \#\Sigma_1^{+(a)}$, which is defined by an $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-for every $A \in \textsc{Struct}[\L]$.
+for every $\A \in \textsc{Struct}[\L]$.
 Our goal here is to eliminate the lexicographically smallest sequence of variables, which we can do easily.
-First, let $\y = (y_1,...,y_k)$, $\y\,^\prime = (y_1^\prime,...,y_k^\prime)$ and
+First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-Therefore, if $f(\A)>0$, $f'(\A)$ will count exactly one element less than $f(\A)$; and if $f(\A)=0$, $\A \not\models\exists x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists x\,\varphi^\prime(\x,\e)$ for every $\e$.
-Thus, $f'(\A)=0$.
+If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
+Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$, therefore $f'(\A)=0$.
+We conclude that $\#\Sigma_1^{+(a)}$ is closed under substraction by one.
 
@@
 
-\item Let $f \in \#\Sigma_1^+$, which is defined by an $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,...,x_d)$ and $\X = (X_1,...,X_r)$. Then
+\item Let $f \in \#\Sigma_1^{+(b)}$, which is defined by an $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. Then
 \begin{eqnarray*}
-f &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid,
+f &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
 \end{eqnarray*}
 for every $\A \in \textsc{Struct}[\L]$. For the time being, suppose that
-\begin{eqnarray*}
-\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n Y_i(\x_i) \right) \wedge \varphi_{\X}^{-}(\y) \wedge \theta(\x) \wedge \beta(\x)
-\end{eqnarray*}
-where $Y_i$ is in $\X$, $Y_i$ is of arity $c_i$ and $\x_i$ is an $c_i$-tuple of variables in $\x$ for all $i \in \{1,...,n\}$, $\y$ is a $d$-tuple of variables in $\x$, $\varphi_{\X}^{-}(\y)$ is a conjunction of negated predicates in $\P$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$.
-We also assume that $(\x_1,...,\x_n,\y) = \x$.
-As an example, we will show that the following formula is of this form:
-\begin{eqnarray*}
-\varphi^{*}(\x,\X) &=& \left( X_1(x_1,x_2) \wedge \neg X_1(x_3,x_4) \wedge (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 ) \wedge \forall z\big( S_1(x_1,z) \big) \right),
-\end{eqnarray*}
+\begin{eqnarray}
+\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n Y_i(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
+\end{eqnarray}
+where $Y_i$ is in $\X$, $Y_i$ is of arity $c_i$ and $\x_i$ is a $c_i$-tuple of variables in $\x$ for all $i \in \{1,\dots,n\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\P$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$.
+We also assume that $(\x_1,\dots,\x_n,\y) = \x$.
+As an example, the following formula is of this form:
+\begin{align*}
+\varphi(\x,\X) =  X_1(x_1,x_2) \wedge \neg X_1(x_3,x_4) \wedge (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 ) \wedge \forall z\big( S_1(x_1,z) \big),
+\end{align*}
 where $\x = (x_1,x_2,x_3,x_4)$ and $\X = (X_1)$.
 Here, $n = 1$, $Y_1 = X_1$, $\x_1 = (x_1,x_2)$ and $\y = (x_3,x_4)$.
-$\varphi_{\X}^{-}(\y) = \neg X_1(x_3,x_4)$, $\theta^*(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4)$ which defines a total order on the partition $\{\{x_1,x_3,x_4\},\{x_2\}\}$, and $\beta^*(\x) = \forall z\big( S_1(x_1,z) \big)$.
+Moreover $\varphi^{-}(\X,\y) = \neg X_1(x_3,x_4)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4)$, which defines a total order on the partition $\{\{x_1,x_3,x_4\},\{x_2\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z) \big)$.
 
-Similarly to the previous proof, we would like to eliminate the lexicographically smallest tuple of predicates that satisfies the formula.
-Let $\u_i$ be a $c_i$-tuple of variables for every $i \in \{1,...,n\}$, and let $m = \sum_{i = i}^n c_i$ be the number of variables of $(\x_1,...,\x_n)$.
+Similarly to the previous proof, we would like to eliminate the lexicographically smallest tuple of predicates that satisfies the formula \eqref{f1}.
+Let $\u_i$ be a $c_i$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = i}^n c_i$ be the number of variables of $(\x_1,\dots,\x_n)$.
 We now define
 \begin{multline*}
-\alpha_{\min}(\u_1,...,\u_n) = \\ \exists\y\left( \alpha(\u_1,...,\u_n,\y)\wedge \forall\v_1...\v_n\w(\alpha(\v_1,...,\v_n,\w)\rightarrow \varphi_{m,<}((\u_1,...,\u_n),(\v_1,...,\v_n))\right),
+\alpha_{\min}(\u_1,\dots,\u_n) = \\ \exists\y\left( \alpha(\u_1,\dots,\u_n,\y)\wedge \forall\v_1\cdots\forall\v_n\w(\alpha(\v_1,\dots,\v_n,\w)\rightarrow \varphi_{m,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\right),
 \end{multline*}
 where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
-Note that $\alpha_{\min}$ is satisfied only by the lexicographically \linebreak smallest assignment $\d$ to $(x_1,...,x_n)$ such that $\exists\y\,(\theta(d_1,...,d_n,\y)\wedge\beta(d_1,...,d_n,\y))$ is true.
+Note that $\alpha_{\min}$ is satisfied only by the lexicographically \linebreak smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\ell)$ and $\A\models\beta(\d_1,\dots,\d_n,\ell)$ for some $\ell \in A^p$.
+Let $\d = (\d_1,\dots,\d_n)$.
 Our new formula is
-\begin{multline*}
-\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n Y_i(\x_i) \right) \wedge \varphi_{\X}^{-}(\y) \wedge \theta(\x) \wedge \beta(\x)\wedge\exists\u_1,...,\u_n\bigg[\alpha_{\min}(\u_1,...,\u_n) \wedge \\ \bigg(\bigvee_{i = 1}^{n}\neg Y_i(\u_i) \vee \bigvee_{Y\in\X} \exists \v\Big( Y(\v) \wedge \bigwedge_{i\in[1,n]: Y_i = Y} \v \neq \u_i\Big) \bigg) \bigg].
-\end{multline*}
-We will show that if there is an assignment to satisfy $\exists\x\,\varphi(\x,\X)$, this formula will be satisfied by exactly one predicate less than $\exists\x\,\varphi(\x,\X)$.
-Suppose $\d$ is the lexicographically smallest assignment which satisfies $\alpha(\x)$.
-Let $\d_i$ be the respective assignment to $x_i$, for every $i\in\{1,...,n\}$.
-Consider now the tuple $\P = (P_i,...,P_r)$ where $P_i = \{d_j \mid j\in\{1,...,n\} \text{ and } X_i = Y_j \}$.
-We will now show that this tuple of predicates is such that $\exists\x\,\varphi(\x,\P)$ is true.
-Suppose that it's not, that is, there is no assignation to $\x$ such that $\varphi(\x,\P)$ is true.
-Since $\d$ satisfies $\left( \bigwedge_{i=1}^n Y_i(\x_i) \right)$ and $\theta(\x) \wedge \beta(\x)$, then $\varphi_{\X}^{-}(\d)$ must be false.
-Therefore, there is a negated predicate $Y$ in $\varphi_{\X}^{-}(\x)$ such that its respective variables in $\y$ are equal to an $x_i$ where $Y_i = Y$.
-Then, since $\theta(\x)$ defines a partition of $\x$, these variables are equal, and thus, $\exists\x\,\varphi(\x,\X)$ is unsatisfiable ($\rightarrow\leftarrow$).
-$\P$ is excluded by the new formula since every $d_i$ is in its respective predicate, which does not satisfy the first part of the new formula, and does not have any tuples that are not in $\d$.
-$\P$ is also the only tuple excluded by the new formula since any other tuple that satisfies it would have at least the same tuples as $\P$, and would not have any more, and thus would be equal to $\P$.
-Therefore, if $\exists\x\,\varphi(\x,\X)$ is satisfiable, $\exists\x\,\varphi^\prime(\x,\X)$ is satisfied by exactly one predicate less.
-Naturally, if $\varphi(\x,\X)$ is unsatisfiable for every assignment to $\x$ and $\X$, $\varphi^\prime(\x,\X)$ also is.
+\begin{multline}
+\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n Y_i(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge\exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \\ \bigg(\bigg(\bigvee_{i = 1}^{n}\neg Y_i(\u_i) \bigg) \vee \bigvee_{Y\in\X} \exists \v\Big( Y(\v) \wedge \bigwedge_{i\in[1,n]: Y_i = Y} \v \neq \u_i\Big) \bigg) \bigg] \label{f2}.
+\end{multline}
+We will now show a result by which the main proof will follow.
+Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignation $\R$ to $\X$.
+Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment which satisfies $\alpha(\x)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
+Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \{\d_j \mid j\in\{1,\dots,n\} \text{ and } X_i = Y_j \}$.
+We will show that this assignation to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P)$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that does this.
+\begin{enumerate}
+\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$.
+That is, there is no assignment $\s$ to $\x$ such that $\varphi(\s,\P)$ is true.
+Let $Q_i$ be the respective assignment of $\P$ to $Y_i$, for $i\in\{1,\dots,n\}$.
+Since $\d$ is such that $\A\models \bigwedge_{i=1}^n Q_i(\d_i)$, $\A\models\theta(\d)$ and  $\A\models\beta(\d)$, then $\A\not\models\varphi^{-}(\P,\ep)$.
+Therefore, there is a predicate $Q_i$ which contains a $c_i$-tuple $\ga$ in $\ep$ such that $\neg \Q_i(\ga)$ appears in $\varphi^{-}(\P,\ep)$.
+Then, there exists an $i\in\{1,\cdots,n\}$ such that $\neg Y_i(\z)$ appears in $\varphi^{-}(\X,\x)$, where $\z$ is in $\y$.
+We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \z < \x_i$ or (3) $\theta(\x)\models \z > \x_i$.
+Considering that (2) and (3) are not possible since $\A\models\theta(\d)$, then $\theta(\x)\models \z = \x_i$.
+If this is the case, then $\A\not\models\exists\x\,\varphi(\x,\R)$ for every possible assignation $\R$ to $\X$, which leads to a contradiction.
+\item[(b)] By the construction of $\P$, we see that 
+$$\A\not\models\bigvee_{i = 1}^{n}\neg Q_i(\d_i) \text{ and that } \A\not\models\bigvee_{Y\in\P} \exists \v\Big( Y(\v) \wedge \bigwedge_{i\in[1,n]: Y_i = Y} \v \neq \d_i\Big).$$ However, the only possible assignment to $\alpha_{\min}(\x_1,\dots,\x_n)$ is $(\d_1,\dots,\d_n)$.
+Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
+\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
+If $\P^\prime$ is missing any tuple of $\P$ in some predicate $\P_i$, then $\A\models \neg Q_i(\d_i)$, so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
+Otherwise, if some predicate $\P^\prime_i$ in $\P^\prime$ has any tuple that $\P^_i$ does not have, then $$\A\models\bigvee_{Y\in\P} \exists \v\Big( Y(\v) \wedge \bigwedge_{i\in[1,n]: Y_i = Y} \v \neq \d_i\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
+On both cases, we see follow to a contradiction.
+\end{enumerate}
+Second, assume that there is no assignation $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
+Let $\P$ be an arbitrary assignation to $\X$.
+Since $\A\not\models\exists\x\,\varphi(\x,\P)$, we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
+It follows that there is no assignation $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
 
-%% Let $\P$ and $\P^\prime$ be tuples of predicates such that $\exists\x\,\varphi(\x,\P)$ and $\exists\x\,\varphi(\x,\P^\prime)$ are true but $\exists\x\,\varphi^\prime(\x,\P)$ and $\exists\x\,\varphi^\prime(\x,\P^\prime)$ are not.
-Let $Q_i$ and $Q_i^\prime$ be the assignment of $\P$ and $\P^\prime$ to $Y_i$, respectively, for every $i\in\{i,...n\}$.
-Then, every tuple $\d_i$ is in its respective relation $Q_i$ in both $\P$ and $\P^\prime$ and $(d_1,...,d_n)$ includes every tuple in $P^\prime$.
-If that is the case, $\P^\prime$ contains exactly the same tuples as $\P$, and are thus the same.
-
-To use this result for the general case, let
+We now follow to the general case, in which $\varphi(\x,\X,\z)$ is an arbirtary extended quantifier-free $\L$-formula. We introduce the formulas $\varphi_i(\x,\X),$ for $i\in\{1,\dots,k\}$ such that:
 \begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models 
-\exists \x \varphi_1(\x,\P) \vee
-\exists \x \varphi_2(\x,\P) \vee
-... \vee
-\exists \x \varphi_k(\x,\P)
- \} \mid
+\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X).
 \end{eqnarray*}
-where $\varphi(\x,\P) \equiv \varphi_1(\x,\P) \vee \varphi_2(\x,\P) \vee ...
-\vee \varphi_k(\x,\P) $.
-This is possible if we consider all first order formulas as literals, and find an equivalent DNF formula.
+We do this by finding an equivalent DNF formula to $\varphi(\x,\X)$ which considers FO-formulas over $\L$ as literals.
 Now we need to exclude that assignment to the next disjuncts, which we do by adding the same sub-formula.
@@
 \begin{multline*}
-\exists\x\left(\varphi^1(\x,\X)\right) \vee \exists\vec{x_1}',...,\vec{x_n}'\Bigg(\alpha^2_{\min}(\vec{x_1}',...,\vec{x_n}') \wedge \\
-\bigg(\bigvee_{i = 1}^{n}\neg Y^2_i(\vec{x_i}') \vee \bigvee_{P\in\P} \exists \x''\Big( P(\x'') \bigwedge_{\substack{i\in\{1,...,n\} \\ P_i = P}} \x'' \neq \vec{x_i}'\Big) \bigg) \Bigg).
+\exists\x\left(\varphi^1(\x,\X)\right) \vee \exists\vec{x_1}',\dots,\vec{x_n}'\Bigg(\alpha^2_{\min}(\vec{x_1}',\dots,\vec{x_n}') \wedge \\
+\bigg(\bigvee_{i = 1}^{n}\neg Y^2_i(\vec{x_i}') \vee \bigvee_{P\in\P} \exists \x''\Big( P(\x'') \bigwedge_{\substack{i\in\{1,\dots,n\} \\ P_i = P}} \x'' \neq \vec{x_i}'\Big) \bigg) \Bigg).
 \end{multline*}
@@
 \begin{multline*}
-\varphi_j^\prime(\x,\X) = \bigwedge_{i=1}^n Y_i(\x_i) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\x) \wedge \varphi_{\S}(\x) \wedge \Bigg(\Big(\forall\x\left(\neg\alpha^1_{\min}(\x)\right)\wedge ...
-\wedge \forall\x\left(\neg\alpha^{j-1}_{\min}(\x)\right)\Big) \to \\
-\exists\u_1,...,\u_n\Bigg(\alpha^j_{\min}(\u_1,...,\u_n) \wedge
-\bigg(\bigvee_{i = 1}^n\neg P(\u_i) \vee \bigvee_{P\in\P} \exists \v\Big( P(\v) \bigwedge_{\substack{i\in\{1,...,n\} \\ P_i = P}} \v \neq \u_i\Big) \bigg) \Bigg) \Bigg)
+\varphi_j^\prime(\x,\X) = \bigwedge_{i=1}^n Y_i(\x_i) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\x) \wedge \varphi_{\S}(\x) \wedge \Bigg(\Big(\forall\x\left(\neg\alpha^1_{\min}(\x)\right)\wedge \dots \wedge \forall\x\left(\neg\alpha^{j-1}_{\min}(\x)\right)\Big) \to \\
+\exists\u_1,\dots,\u_n\Bigg(\alpha^j_{\min}(\u_1,\dots,\u_n) \wedge
+\bigg(\bigvee_{i = 1}^n\neg P(\u_i) \vee \bigvee_{P\in\P} \exists \v\Big( P(\v) \bigwedge_{\substack{i\in\{1,\dots,n\} \\ P_i = P}} \v \neq \u_i\Big) \bigg) \Bigg) \Bigg)
 \end{multline*}
@@
 \exists \x \varphi^\prime_2(\x,\P) \vee
-... \vee
+\dots \vee
 \exists \x \varphi^\prime_k(\x,\P)

### 2014-12-31 sharp-sigma-one-plus.tex
31 dic

@@
 
+\def\dotminus{\mathbin{\ooalign{\hss\raise1ex\hbox{.}\hss\cr
+  \mathsurround=0pt$-$}}}
+
+\def\E1{\#\Sigma_1^{+}}
+\def\Ea{\#\Sigma_1^{+,1}}
+\def\Eb{\#\Sigma_1^{+,2}}
+\def\Ec{\#\Sigma_1^{+,1,2}}
+
+\def\Truc{\textsc{Struct}[\L]}
+
 \def\A{{\frak A}}
@@
 \def\P{\vec{P}}
-\def\Q{\vec{Q}}
 \def\R{\vec{R}}
@@
 \def\Z{\vec{Z}}
-%% a - arity of \X / arity of assignations \P to \X
+%% a - arity of \X / arity of assignments \P to \X
 %% b - arity of predicates in \S
@@
 %% l - emergency index / size of tuple
+\def\l{\vec{\ell}}
 %% m - size of variable tuple
@@
 \def\z{\vec{z}} %% open variables
-\def\ep{\vec{\varepsilon}}
-\def\ga{\vec{\gamma}}
+\def\ep{\vec{o}}
+\def\ga{\vec{p}}
 
@@
 { \LARGE \bf
-  Some properties of $\#\Sigma_1^{+}$
+  Some properties of $\E1$
 }
@@
 \begin{eqnarray*}
-\textsc{Struct}[\L] &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
+\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
 && \leq \text{ is interpreted as a total order for } A \}.
 \end{eqnarray*}
-We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in N$ there are infinite variables in ${\cal X}$ of arity $n$.
+We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$.
 A quantifier-free $\L$-formula is defined by the following grammar:
@@
 \end{eqnarray*}
-Let $\X = (X_1,\dots,X_r)$ and $\P = (P_1,\dots,P_r)$, where $P_i$ is a predicate of arity $a_i$, for every $i$.
-Let $\z$ be a tuple of variables and $\e$ be a tuple of constants with the same number of elements as $\z$.
-A function $f:\textsc{Struct}[\L] \to \mathbb{N}$ is in $\#\Sigma_1^+$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\X,\z)$ such that
+Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables.
+For every $\L$-formula $\psi(\Y,\y)$, we define $f_{\psi(\Y,\y)}$ as follows:
 \begin{eqnarray*}
-f(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \exists \x \: \varphi(\x,\P,\e) \} \mid,
+f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
 \end{eqnarray*}
-for every $\A \in \textsc{Struct}[\L]$.\\
+for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, where $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.
+
+Let $\X = (X_1,\dots,X_r)$.
+Let $\z$ be a tuple of variables.
+A function $f:\Truc \to \mathbb{N}$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\X,\z)$ such that $f = f_{\exists \x \: \varphi(\x,\X,\z)}.$\\
 
@@
 \begin{theo}
-The decision version of a function in $\#\Sigma_1^{+}$ is in \textsc{P}.
+The decision version of a function in $\E1$ is in \textsc{P}.
 \end{theo}
 \begin{proof}
-Let $f$ be a function in $\#\Sigma_1^{+}$. Then there is a formula $\varphi(\x,\X,\z)$ such that
+Let $f$ be a function in $\E1$. Then there is a formula $\varphi(\x,\X,\z)$ such that
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \textsc{Struct}[\L]$, $\z$ is an $m$-tuple of variables, $\e$ is an $m$-tuple of elements and $\x$ is a $k$-tuple of variables.
+where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $\z$ is an $m$-tuple of variables, $\e$ is an $m$-tuple of elements and $\x$ is a $k$-tuple of variables.
 Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible evaluations for $\z$.
@@
 \end{eqnarray*}
-Note that $\varphi_\A(\X)$ will have at least one assignation for $\X$ iff $f(\A)>0$.
+Note that $\varphi_\A(\X)$ will have at least one assignment for $\X$ iff $f(\A)>0$.
 Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction.
@@
 \begin{theo}
-$\#\Sigma_1 \subsetneq \#\Sigma_1^{+}$
+$\#\Sigma_1 \subsetneq \E1$
 \end{theo}
 \begin{proof}
-We will show that the $\#\Sigma_1^{+}$ function defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
+We will show that the $\E1$ function defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
 \end{proof}
 
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1
-
-A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, there exists $f' \in \F$ such that 
+For a given function $f$, we define $f \dotminus 1$ as follows:
 \begin{eqnarray*}
-f'(\A) =
+f \dotminus 1(\A) =
 \begin{cases}
@@
 \end{eqnarray*}
+for every structure $\A$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, $f \dotminus 1 \in \F$.
+
+%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 
+%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1
+
 \begin{theo}
-$\#\Sigma_1^+$ is closed under substraction by one.
+$\E1$ is closed under substraction by one.
 \end{theo}
 \begin{proof}
-A function in $\#\Sigma_1^+$ has three possible ways of counting: counting only variables, only predicates, and predicates with variables.
-Let us separate each of this cases on the classes $\#\Sigma_1^{+(a)}$, $\#\Sigma_1^{+(b)}$ and $\#\Sigma_1^{+(c)}$ respectively:
+The class $\E1$ can be partitioned in three different classes:  $\Ea, \Eb$ and $\Ec.$ We define these three classes as follows:
+\begin{enumerate}
+\item A function $f \in \Ea$ is defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$ such that $f = f_{\exists \x \: \varphi(\x,\z)}.$
+\item A function $f \in \Eb$ is defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ such that $f = f_{\exists \x \: \varphi(\x,\X)}.$
+\item A function $f \in \Ec$ is defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$ such that $f = f_{\exists \x \: \varphi(\x,\X,\z)}.$
+\end{enumerate}
+We separate the proof in three cases, one for each one of these classes:
 \begin{enumerate}
@@
 
-\item Let $f \in \#\Sigma_1^{+(a)}$, which is defined by an $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
+\item Let $f \in \Ea$, which is defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$, such that $f = f_{\varphi(\x,\z)}$. Equivalently,
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-for every $\A \in \textsc{Struct}[\L]$.
+for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\e \in A^d$.
 Our goal here is to eliminate the lexicographically smallest sequence of variables, which we can do easily.
@@
 \begin{eqnarray*}
-\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j = y_j^\prime \wedge y_i < y_i^\prime \right).
+\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
 \end{eqnarray*}
@@
 If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
-Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$, therefore $f'(\A)=0$.
-We conclude that $\#\Sigma_1^{+(a)}$ is closed under substraction by one.
+Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore $f'(\A)=0$.
+As we see that $f' = f\dotminus 1$, we conclude that $\Ea$ is closed under substraction by one.
 
@@
 
-\item Let $f \in \#\Sigma_1^{+(b)}$, which is defined by an $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. Then
+\item Let $f \in \Eb$, which is defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$, such that $f = f_{\exists \x \: \varphi(\x,\X)}$.
+Equivalently,
 \begin{eqnarray*}
-f &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
+f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
 \end{eqnarray*}
-for every $\A \in \textsc{Struct}[\L]$. For the time being, suppose that
+for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $Y_i$ is in $\X$, $Y_i$ is of arity $c_i$ and $\x_i$ is a $c_i$-tuple of variables in $\x$ for all $i \in \{1,\dots,n\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\P$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$.
+where $Y_i$ is in $\X$, $Y_i$ is of arity $c_i$ and $\x_i$ is a $c_i$-tuple of variables in $\x$ for all $i \in \{1,\dots,n\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$.
 We also assume that $(\x_1,\dots,\x_n,\y) = \x$.
@@
 Here, $n = 1$, $Y_1 = X_1$, $\x_1 = (x_1,x_2)$ and $\y = (x_3,x_4)$.
-Moreover $\varphi^{-}(\X,\y) = \neg X_1(x_3,x_4)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4)$, which defines a total order on the partition $\{\{x_1,x_3,x_4\},\{x_2\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z) \big)$.
+Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_3,x_4)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4)$, which defines a total order on the partition $\{\{x_1,x_3,x_4\},\{x_2\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z) \big)$.
 
@@
 \begin{multline*}
-\alpha_{\min}(\u_1,\dots,\u_n) = \\ \exists\y\left( \alpha(\u_1,\dots,\u_n,\y)\wedge \forall\v_1\cdots\forall\v_n\w(\alpha(\v_1,\dots,\v_n,\w)\rightarrow \varphi_{m,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\right),
+\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to \varphi_{m,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
 \end{multline*}
 where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
-Note that $\alpha_{\min}$ is satisfied only by the lexicographically \linebreak smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\ell)$ and $\A\models\beta(\d_1,\dots,\d_n,\ell)$ for some $\ell \in A^p$.
+Note that $\alpha_{\min}$ is satisfied only by the lexicographically \linebreak smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
 Let $\d = (\d_1,\dots,\d_n)$.
@@
 We will now show a result by which the main proof will follow.
-Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignation $\R$ to $\X$.
-Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment which satisfies $\alpha(\x)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
+\begin{lemma}  \label{lemmaone}
+$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
+\end{lemma}
+\begin{proof}
+Let $\A \in \Truc$.
+Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
+Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
 Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \{\d_j \mid j\in\{1,\dots,n\} \text{ and } X_i = Y_j \}$.
-We will show that this assignation to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P)$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that does this.
+We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
+For the following procedure, let $Q_i = P_j$ if $Y_i = X_j$, for every $i \in \{1,\ldots,n\}$ and $j \in \{1,\ldots,r\}.$
 \begin{enumerate}
@@
 That is, there is no assignment $\s$ to $\x$ such that $\varphi(\s,\P)$ is true.
-Let $Q_i$ be the respective assignment of $\P$ to $Y_i$, for $i\in\{1,\dots,n\}$.
-Since $\d$ is such that $\A\models \bigwedge_{i=1}^n Q_i(\d_i)$, $\A\models\theta(\d)$ and  $\A\models\beta(\d)$, then $\A\not\models\varphi^{-}(\P,\ep)$.
-Therefore, there is a predicate $Q_i$ which contains a $c_i$-tuple $\ga$ in $\ep$ such that $\neg \Q_i(\ga)$ appears in $\varphi^{-}(\P,\ep)$.
-Then, there exists an $i\in\{1,\cdots,n\}$ such that $\neg Y_i(\z)$ appears in $\varphi^{-}(\X,\x)$, where $\z$ is in $\y$.
+Since $\d$ is such that $\A\models \bigwedge_{i=1}^n Q_i(\d_i)$ and $\A\models \alpha(\d)$, then $\A\models\theta(\d)$ and $\A\models\beta(\d)$, and so, $\A\not\models\varphi^{-}(\P,\ep)$.
+Therefore, there is a predicate $Q_i$ which contains a $c_i$-tuple $\ga$ in $\ep$ such that $\neg Q_i(\ga)$ appears in $\varphi^{-}(\P,\ep)$.
+Then, there exists an $i\in\{1,\ldots,n\}$ such that $\neg Y_i(\z)$ appears in $\varphi^{-}(\X,\x)$, where $\z$ is in $\y$.
 We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \z < \x_i$ or (3) $\theta(\x)\models \z > \x_i$.
-Considering that (2) and (3) are not possible since $\A\models\theta(\d)$, then $\theta(\x)\models \z = \x_i$.
-If this is the case, then $\A\not\models\exists\x\,\varphi(\x,\R)$ for every possible assignation $\R$ to $\X$, which leads to a contradiction.
+Considering that (2) and (3) are not possible since $\A\models\theta(\d)$, and $\z$ and $\x_i$ are assigned the value $\d_i$, then $\theta(\x)\models \z = \x_i$.
+But if this is the case, then $Y_i(\x_i), \neg Y_i(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X).$ As consequence, $\A\not\models\exists\x\,\varphi(\x,\R)$ for every possible assignment $\R$ to $\X$, which leads to a contradiction.
 \item[(b)] By the construction of $\P$, we see that 
@@
 If $\P^\prime$ is missing any tuple of $\P$ in some predicate $\P_i$, then $\A\models \neg Q_i(\d_i)$, so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-Otherwise, if some predicate $\P^\prime_i$ in $\P^\prime$ has any tuple that $\P^_i$ does not have, then $$\A\models\bigvee_{Y\in\P} \exists \v\Big( Y(\v) \wedge \bigwedge_{i\in[1,n]: Y_i = Y} \v \neq \d_i\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-On both cases, we see follow to a contradiction.
+Otherwise, if some predicate $\P^\prime_i$ in $\P^\prime$ has any tuple that $\P_i$ does not have, then $$\A\models\bigvee_{Y\in\P} \exists \v\Big( Y(\v) \wedge \bigwedge_{i\in[1,n]: Y_i = Y} \v \neq \d_i\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
+On both cases, we have a contradiction.
 \end{enumerate}
-Second, assume that there is no assignation $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
-Let $\P$ be an arbitrary assignation to $\X$.
+Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
+Let $\P$ be an arbitrary assignment to $\X$.
 Since $\A\not\models\exists\x\,\varphi(\x,\P)$, we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignation $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
+It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
+\end{proof}
 
-We now follow to the general case, in which $\varphi(\x,\X,\z)$ is an arbirtary extended quantifier-free $\L$-formula. We introduce the formulas $\varphi_i(\x,\X),$ for $i\in\{1,\dots,k\}$ such that:
+We now follow to the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula. We introduce the formulas $\varphi_i(\x,\X),$ for $i\in\{1,\dots,k\}$ such that:
 \begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X).
+\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X),
+\end{eqnarray*}
+where, for every $i\in\{1,\ldots,k\}$,
+\begin{eqnarray*}
+\varphi_j(\x,\X) &=& \left( \bigwedge_{i=1}^n Y^j_i(\x^j_i) \right) \wedge \varphi^{-}_j(\X,\y^j) \wedge \theta_j(\x) \wedge \beta_j(\x).
 \end{eqnarray*}
 We do this by finding an equivalent DNF formula to $\varphi(\x,\X)$ which considers FO-formulas over $\L$ as literals.
-Now we need to exclude that assignment to the next disjuncts, which we do by adding the same sub-formula.
-We also need to include the case where there are no possible assignments to $\x$ for the first disjunct, so we add this sub-formula to $\varphi^2(\x,\X)$:
+$\x^j_i$, $Y^j_i$, $\varphi^{-}_j(\X,\y^j)$, $\theta_j(\x)$, $\beta_j(\x)$, and $\alpha_j(\x)$ are defined analogously to the particular proof.
+
+Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\P$ such that $\A\models\varphi_2(\x,\P)$, and so on.
+As we already know how to exclude that assignment in the first disjunct, we now move to the next disjuncts.
+We now define
 \begin{multline*}
-\exists\x\left(\varphi^1(\x,\X)\right) \vee \exists\vec{x_1}',\dots,\vec{x_n}'\Bigg(\alpha^2_{\min}(\vec{x_1}',\dots,\vec{x_n}') \wedge \\
-\bigg(\bigvee_{i = 1}^{n}\neg Y^2_i(\vec{x_i}') \vee \bigvee_{P\in\P} \exists \x''\Big( P(\x'') \bigwedge_{\substack{i\in\{1,\dots,n\} \\ P_i = P}} \x'' \neq \vec{x_i}'\Big) \bigg) \Bigg).
+\psi_j(\x,\X) = \exists\u_1\cdots\exists\u_n\bigg[\alpha^j_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{i = 1}^{n}\neg Y^j_i(\u_i) \bigg) \vee \bigvee_{Y\in\X} \exists \v\Big( Y(\v) \wedge \bigwedge_{i\in[1,n]: Y^j_i = Y} \v \neq \u_i\Big) \bigg) \bigg],
 \end{multline*}
-The full disjunct is as follows,
-\begin{multline*}
-\varphi_j^\prime(\x,\X) = \bigwedge_{i=1}^n Y_i(\x_i) \wedge \varphi_{\overline{P}}(\vec{y}) \wedge \theta(\x) \wedge \varphi_{\S}(\x) \wedge \Bigg(\Big(\forall\x\left(\neg\alpha^1_{\min}(\x)\right)\wedge \dots \wedge \forall\x\left(\neg\alpha^{j-1}_{\min}(\x)\right)\Big) \to \\
-\exists\u_1,\dots,\u_n\Bigg(\alpha^j_{\min}(\u_1,\dots,\u_n) \wedge
-\bigg(\bigvee_{i = 1}^n\neg P(\u_i) \vee \bigvee_{P\in\P} \exists \v\Big( P(\v) \bigwedge_{\substack{i\in\{1,\dots,n\} \\ P_i = P}} \v \neq \u_i\Big) \bigg) \Bigg) \Bigg)
-\end{multline*}
-which, if all of the previous disjuncts have no possible assignments, eliminates the least assignment to $\P$ in that disjunct. Finally,
-\begin{eqnarray*}
-f^\prime(\A) &=& \mid \{ \langle\P\rangle \mid \A \models 
-\exists \x \varphi^\prime_1(\x,\P) \vee
-\exists \x \varphi^\prime_2(\x,\P) \vee
-\dots \vee
-\exists \x \varphi^\prime_k(\x,\P)
- \} \mid
-\end{eqnarray*}
-counts exactly one assignment to $\P$ less than $F(\A)$.
-
-Therefore,
+where $\alpha^j_{\min}$ is defined analogously. Our new formula $\varphi_j^\prime(\x,\X)$ is defined as follows:
+\begin{multline}
+\varphi_j^\prime(\x,\X) = \left( \bigwedge_{i=1}^n Y_i(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x) \wedge \\ \Bigg[\big(\exists\w(\varphi_1(\w,\X)\wedge\psi_1(\w,\X))\big)\vee \cdots \vee \big(\exists\w(\varphi_{j-1}(\w,\X)\wedge\psi_{j-1}(\w,\X))\big) \vee \psi_j(\x,\X) \Bigg] \label{f3}.
+\end{multline}
+Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$.
+Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least in $\{1,\ldots,k\}$ such that there exists at least an assignment $\R$ to $\X$ for which $\A\models\exists\x\:\varphi_q(\x,\R).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone}.
+We will show that for every $j\in\{1,\ldots,k\}$, $\A\not\models\varphi_j^\prime(\x,\P).$ Since for any $i\in\{1,\ldots,q-1\}$ there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_i(\x,\R)$, then $\A\not\models\exists\x\:\varphi_1(\x,\P),\ldots,\A\not\models\exists\x\:\varphi_{q-1}(\x,\P),$ so it follows that $\A\not\models\exists\x\:\varphi^\prime_1(\x,\P),\ldots,\A\not\models\exists\x\:\varphi^\prime_{q-1}(\x,\P).$  Analogously to Lemma \ref{lemmaone}, we note that $\P$ is such that $\A\not\models\exists\x\:\psi_q(\x,\P),$ and then $\A\not\models\exists\x\:\varphi^\prime_q(\x,\P)$
 

### 2015-01-03 sharp-sigma-one-plus.tex
6 ene v1

@@
 \def\E1{\#\Sigma_1^{+}}
-\def\Ea{\#\Sigma_1^{+,1}}
-\def\Eb{\#\Sigma_1^{+,2}}
-\def\Ec{\#\Sigma_1^{+,1,2}}
 
@@
 Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables.
-For every $\L$-formula $\psi(\Y,\y)$, we define $f_{\psi(\Y,\y)}$ as follows:
+For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, where $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.
+for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $a_1,\ldots,a_q$, and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.
 
-Let $\X = (X_1,\dots,X_r)$.
-Let $\z$ be a tuple of variables.
-A function $f:\Truc \to \mathbb{N}$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\X,\z)$ such that $f = f_{\exists \x \: \varphi(\x,\X,\z)}.$\\
+Let $\X = (X_1,\dots,X_r)$, and let $\z$ be a tuple of variables.
+A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\X,\z)$ such that $f = f_{\exists \x \: \varphi(\x,\X,\z)}.$\\
 
-The {\em decision version} of a function $f$ is the language $L_f = \{\A \mid f(\A) > 0\}$.
+The {\em decision version} of a function $f$ is defined by the language $L_f = \{\A \mid f(\A) > 0\}$.
 \begin{theo}
@@
 \end{eqnarray*}
-for every structure $\A$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, $f \dotminus 1 \in \F$.
+for every structure $\A$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
 
@@
 \begin{proof}
-The class $\E1$ can be partitioned in three different classes:  $\Ea, \Eb$ and $\Ec.$ We define these three classes as follows:
-\begin{enumerate}
-\item A function $f \in \Ea$ is defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$ such that $f = f_{\exists \x \: \varphi(\x,\z)}.$
-\item A function $f \in \Eb$ is defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ such that $f = f_{\exists \x \: \varphi(\x,\X)}.$
-\item A function $f \in \Ec$ is defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$ such that $f = f_{\exists \x \: \varphi(\x,\X,\z)}.$
-\end{enumerate}
-We separate the proof in three cases, one for each one of these classes:
+For every $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$, let $\z = (z_1,\dots,z_d)$ and let $\X = (X_1,\dots,X_r)$.
+There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$.
+This separates the proof in three cases:
 \begin{enumerate}
@@
 
-\item Let $f \in \Ea$, which is defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$, such that $f = f_{\varphi(\x,\z)}$. Equivalently,
+\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
 \begin{eqnarray*}
@@
 for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\e \in A^d$.
-Our goal here is to eliminate the lexicographically smallest sequence of variables, which we can do easily.
+Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily.
 First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
@@
 If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
-Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore $f'(\A)=0$.
-As we see that $f' = f\dotminus 1$, we conclude that $\Ea$ is closed under substraction by one.
+Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$.
+Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E1.$
 
@@
 
-\item Let $f \in \Eb$, which is defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$, such that $f = f_{\exists \x \: \varphi(\x,\X)}$.
-Equivalently,
+\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
 \begin{eqnarray*}
@@
 \begin{eqnarray}
-\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n Y_i(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
+\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
 \end{eqnarray}
-where $Y_i$ is in $\X$, $Y_i$ is of arity $c_i$ and $\x_i$ is a $c_i$-tuple of variables in $\x$ for all $i \in \{1,\dots,n\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$.
+where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$.
 We also assume that $(\x_1,\dots,\x_n,\y) = \x$.
@@
 where $\x = (x_1,x_2,x_3,x_4)$ and $\X = (X_1)$.
-Here, $n = 1$, $Y_1 = X_1$, $\x_1 = (x_1,x_2)$ and $\y = (x_3,x_4)$.
-Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_3,x_4)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4)$, which defines a total order on the partition $\{\{x_1,x_3,x_4\},\{x_2\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z) \big)$.
+Here, $n = 1$, $\lambda(1) = \lambda(2) = 1$, $\x_1 = (x_1,x_2)$ and $\y = (x_3,x_4)$.
+Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_3,x_4)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4)$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z) \big)$.
 
 Similarly to the previous proof, we would like to eliminate the lexicographically smallest tuple of predicates that satisfies the formula \eqref{f1}.
-Let $\u_i$ be a $c_i$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = i}^n c_i$ be the number of variables of $(\x_1,\dots,\x_n)$.
+Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = i}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
 We now define
@@
 Note that $\alpha_{\min}$ is satisfied only by the lexicographically \linebreak smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
-Let $\d = (\d_1,\dots,\d_n)$.
 Our new formula is
 \begin{multline}
-\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n Y_i(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge\exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \\ \bigg(\bigg(\bigvee_{i = 1}^{n}\neg Y_i(\u_i) \bigg) \vee \bigvee_{Y\in\X} \exists \v\Big( Y(\v) \wedge \bigwedge_{i\in[1,n]: Y_i = Y} \v \neq \u_i\Big) \bigg) \bigg] \label{f2}.
+\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge\exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \\ \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg] \label{f2}.
 \end{multline}
-We will now show a result by which the main proof will follow.
+We now show a result by which the main proof will follow.
 \begin{lemma}  \label{lemmaone}
@@
 Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \{\d_j \mid j\in\{1,\dots,n\} \text{ and } X_i = Y_j \}$.
+Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{\lambda(j)=i}\{\d_j\}$.
 We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
-For the following procedure, let $Q_i = P_j$ if $Y_i = X_j$, for every $i \in \{1,\ldots,n\}$ and $j \in \{1,\ldots,r\}.$
 \begin{enumerate}
 \item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$.
-That is, there is no assignment $\s$ to $\x$ such that $\varphi(\s,\P)$ is true.
-Since $\d$ is such that $\A\models \bigwedge_{i=1}^n Q_i(\d_i)$ and $\A\models \alpha(\d)$, then $\A\models\theta(\d)$ and $\A\models\beta(\d)$, and so, $\A\not\models\varphi^{-}(\P,\ep)$.
-Therefore, there is a predicate $Q_i$ which contains a $c_i$-tuple $\ga$ in $\ep$ such that $\neg Q_i(\ga)$ appears in $\varphi^{-}(\P,\ep)$.
-Then, there exists an $i\in\{1,\ldots,n\}$ such that $\neg Y_i(\z)$ appears in $\varphi^{-}(\X,\x)$, where $\z$ is in $\y$.
+That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$).
+Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$.
+Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$.
 We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \z < \x_i$ or (3) $\theta(\x)\models \z > \x_i$.
-Considering that (2) and (3) are not possible since $\A\models\theta(\d)$, and $\z$ and $\x_i$ are assigned the value $\d_i$, then $\theta(\x)\models \z = \x_i$.
-But if this is the case, then $Y_i(\x_i), \neg Y_i(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X).$ As consequence, $\A\not\models\exists\x\,\varphi(\x,\R)$ for every possible assignment $\R$ to $\X$, which leads to a contradiction.
+Considering that (2) and (3) are not possible since $\A\models\theta(\d)$, and both $\z$ and $\x_i$ are assigned the value $\d_i$, then $\theta(\x)\models \z = \x_i$.
+But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$ from which we conclude that $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which leads to a contradiction.
 \item[(b)] By the construction of $\P$, we see that 
-$$\A\not\models\bigvee_{i = 1}^{n}\neg Q_i(\d_i) \text{ and that } \A\not\models\bigvee_{Y\in\P} \exists \v\Big( Y(\v) \wedge \bigwedge_{i\in[1,n]: Y_i = Y} \v \neq \d_i\Big).$$ However, the only possible assignment to $\alpha_{\min}(\x_1,\dots,\x_n)$ is $(\d_1,\dots,\d_n)$.
+$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big).$$ However, the only possible assignment to $\alpha_{\min}(\x_1,\dots,\x_n)$ is $(\d_1,\dots,\d_n)$.
 Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
 \item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-If $\P^\prime$ is missing any tuple of $\P$ in some predicate $\P_i$, then $\A\models \neg Q_i(\d_i)$, so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-Otherwise, if some predicate $\P^\prime_i$ in $\P^\prime$ has any tuple that $\P_i$ does not have, then $$\A\models\bigvee_{Y\in\P} \exists \v\Big( Y(\v) \wedge \bigwedge_{i\in[1,n]: Y_i = Y} \v \neq \d_i\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
+We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$.
+Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $\P^\prime$, then $\A\models \neg P^\prime_i(\d_i)$.
+And so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
+Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $\P_i$ does not have.
+If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
 On both cases, we have a contradiction.
 \end{enumerate}
+With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\R)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\R)}(\A) = f_{\exists\x\,\varphi(\x,\R)}(\A)-1.$
+
 Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
@@
 It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
+And so, every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\R)}(\A))=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\R)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\R)} = f_{\exists\x\,\varphi(\x,\R)}\dotminus 1.$
 \end{proof}
@@
 Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least in $\{1,\ldots,k\}$ such that there exists at least an assignment $\R$ to $\X$ for which $\A\models\exists\x\:\varphi_q(\x,\R).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone}.
-We will show that for every $j\in\{1,\ldots,k\}$, $\A\not\models\varphi_j^\prime(\x,\P).$ Since for any $i\in\{1,\ldots,q-1\}$ there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_i(\x,\R)$, then $\A\not\models\exists\x\:\varphi_1(\x,\P),\ldots,\A\not\models\exists\x\:\varphi_{q-1}(\x,\P),$ so it follows that $\A\not\models\exists\x\:\varphi^\prime_1(\x,\P),\ldots,\A\not\models\exists\x\:\varphi^\prime_{q-1}(\x,\P).$  Analogously to Lemma \ref{lemmaone}, we note that $\P$ is such that $\A\not\models\exists\x\:\psi_q(\x,\P),$ and then $\A\not\models\exists\x\:\varphi^\prime_q(\x,\P)$
+We will show that for every $j\in\{1,\ldots,k\}$, $\A\not\models\varphi_j^\prime(\x,\P).$ Since for any $i\in\{1,\ldots,q-1\}$ there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_i(\x,\R)$, then $\A\not\models\exists\x\:\varphi_1(\x,\P),\ldots,\A\not\models\exists\x\:\varphi_{q-1}(\x,\P),$ so it follows that $\A\not\models\exists\x\:\varphi^\prime_1(\x,\P),\ldots,\A\not\models\exists\x\:\varphi^\prime_{q-1}(\x,\P).$  Analogously to Lemma \ref{lemmaone}, we note that $\P$ is such that $\A\not\models\exists\x\:\psi_q(\x,\P),$ and then $\A\not\models\exists\x\:\varphi^\prime_q(\x,\P)$.
 

### 2015-01-05 sharp-sigma-one-plus.tex
6 ene v2

@@
 Similarly to the previous proof, we would like to eliminate the lexicographically smallest tuple of predicates that satisfies the formula \eqref{f1}.
-Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = i}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
+Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
 We now define
@@
 \end{enumerate}
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\R)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\R)}(\A) = f_{\exists\x\,\varphi(\x,\R)}(\A)-1.$
+With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
 
@@
 It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
-And so, every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\R)}(\A))=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\R)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\R)} = f_{\exists\x\,\varphi(\x,\R)}\dotminus 1.$
+And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
 \end{proof}
 
-We now follow to the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula. We introduce the formulas $\varphi_i(\x,\X),$ for $i\in\{1,\dots,k\}$ such that:
+We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
+By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\hat{\varphi}_i(\x,\X)$ with $i\in\{1,\ldots,\hat{k}\}$ such that
 \begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X),
+\varphi(\x,\X) &\equiv& \hat{\varphi}_1(\x,\X) \vee \hat{\varphi}_2(\x,\X) \vee \dots  \vee \hat{\varphi}_{\hat{k}}(\x,\X),
 \end{eqnarray*}
-where, for every $i\in\{1,\ldots,k\}$,
+where, for every $i\in\{1,\ldots,\hat{k}\}$, 
 \begin{eqnarray*}
-\varphi_j(\x,\X) &=& \left( \bigwedge_{i=1}^n Y^j_i(\x^j_i) \right) \wedge \varphi^{-}_j(\X,\y^j) \wedge \theta_j(\x) \wedge \beta_j(\x).
+\hat{\varphi}_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \hat{\varphi}^{-}_i(\X,\y_i)  \wedge \beta_i(\x).
 \end{eqnarray*}
-We do this by finding an equivalent DNF formula to $\varphi(\x,\X)$ which considers FO-formulas over $\L$ as literals.
-$\x^j_i$, $Y^j_i$, $\varphi^{-}_j(\X,\y^j)$, $\theta_j(\x)$, $\beta_j(\x)$, and $\alpha_j(\x)$ are defined analogously to the particular proof.
+The function $\lambda_i$ is defined analogously to the first part of the proof, the tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x.$ The formulas $\hat{\varphi}^{-}_i(\X,\y_i)$, $\beta_i(\x)$, are also defined analogously.
+Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables.
+Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$.
+Note that for every formula $\psi(\x)$,
+\begin{eqnarray*}
+\exists\x\:\psi(\x) &\equiv& \exists\x(\psi(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\psi(\x)\wedge\theta^{g(d)}(\x)).
+\end{eqnarray*}
+The formulas we will be using are $\varphi_i(\x,\X)$, for $i\in\{1,\ldots,k\},$ which are defined as follows: 
+\begin{eqnarray*}
+\varphi_1(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^1(\x), \\
+& \vdots & \\
+\varphi_{g(d)}(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
+\varphi_{g(d)+1}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^1(\x), \\
+& \vdots & \\
+\varphi_{2\cdot g(d)}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
+& \vdots & \\
+\varphi_{(\hat{k}-1) g(d)+1}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^1(\x), \\
+& \vdots & \\
+\varphi_{\hat{k} \cdot g(d)}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^{g(d)}(\x).
+\end{eqnarray*}
+Note that $k = \hat{k} \cdot g(d).$ Now we can define the formula like we did on the first part of the proof. For every $i\in\{1,\ldots,k\}$,
+\begin{eqnarray*}
+\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \hat{\varphi}^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
+\end{eqnarray*}
+where $\theta_i(\x) = \theta^{\,i\:(\!\!\!\!\mod g(d))}(\x).$ Also, note that
+\begin{eqnarray*}
+\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X).
+\end{eqnarray*}
 
@@
 As we already know how to exclude that assignment in the first disjunct, we now move to the next disjuncts.
-We now define
+Similarly to the first part of the proof, let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
 \begin{multline*}
-\psi_j(\x,\X) = \exists\u_1\cdots\exists\u_n\bigg[\alpha^j_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{i = 1}^{n}\neg Y^j_i(\u_i) \bigg) \vee \bigvee_{Y\in\X} \exists \v\Big( Y(\v) \wedge \bigwedge_{i\in[1,n]: Y^j_i = Y} \v \neq \u_i\Big) \bigg) \bigg],
+\alpha^i_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_n,\w)\wedge\bigvee_{j=1}^n(\u_j\neq\v_j)\big)\to \varphi_{m_i,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
 \end{multline*}
-where $\alpha^j_{\min}$ is defined analogously. Our new formula $\varphi_j^\prime(\x,\X)$ is defined as follows:
+where $\u_j,\v_j,$ for $\{1,\ldots,n_i\}$ and $\w$ have the adequate number of variables. Also, let
+\begin{multline*}
+\psi_i(\X) = \exists\u_1\cdots\exists\u_n\bigg[\alpha^j_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
+\end{multline*}
+Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
 \begin{multline}
-\varphi_j^\prime(\x,\X) = \left( \bigwedge_{i=1}^n Y_i(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x) \wedge \\ \Bigg[\big(\exists\w(\varphi_1(\w,\X)\wedge\psi_1(\w,\X))\big)\vee \cdots \vee \big(\exists\w(\varphi_{j-1}(\w,\X)\wedge\psi_{j-1}(\w,\X))\big) \vee \psi_j(\x,\X) \Bigg] \label{f3}.
+\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\varphi_1(\v,\X)\vee\psi_2(\X)) \wedge \cdots \wedge \\ (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{i-1}(\v,\X)\vee\psi_i(\X)).
 \end{multline}
 Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$.
-Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least in $\{1,\ldots,k\}$ such that there exists at least an assignment $\R$ to $\X$ for which $\A\models\exists\x\:\varphi_q(\x,\R).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone}.
+Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least in $\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_q(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone}.
+As we did in Lemma \ref{lemmaone} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
+\begin{enumerate}
+\item[a)] We know that $\A\models\exists\x\:\varphi_q(\x,\P),$ so it immediately follows that $\A\models\exists\x\:v.$
+\item[b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
+(1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assigment to $\X$ for any of these formulas.
+(2) We can use the proof in Lemma \ref{lemmaone} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$.
+However, $\A\not\models\exists\x\:\hat{\psi}_q(\P)$ so $\A\not\models\exists\x\:\varphi^\prime_i(\P)$.
+And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
+\item[c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (1) and (2).
+\end{enumerate}
 We will show that for every $j\in\{1,\ldots,k\}$, $\A\not\models\varphi_j^\prime(\x,\P).$ Since for any $i\in\{1,\ldots,q-1\}$ there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_i(\x,\R)$, then $\A\not\models\exists\x\:\varphi_1(\x,\P),\ldots,\A\not\models\exists\x\:\varphi_{q-1}(\x,\P),$ so it follows that $\A\not\models\exists\x\:\varphi^\prime_1(\x,\P),\ldots,\A\not\models\exists\x\:\varphi^\prime_{q-1}(\x,\P).$  Analogously to Lemma \ref{lemmaone}, we note that $\P$ is such that $\A\not\models\exists\x\:\psi_q(\x,\P),$ and then $\A\not\models\exists\x\:\varphi^\prime_q(\x,\P)$.

### 2015-01-06 sharp-sigma-one-plus.tex
6 ene

@@
 \begin{multline*}
-\psi_i(\X) = \exists\u_1\cdots\exists\u_n\bigg[\alpha^j_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
+\psi_i(\X) = \forall\x\:\neg\alpha(\x) \vee \exists\u_1\cdots\exists\u_n\bigg[\alpha^i_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
 \end{multline*}
@@
 Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$.
-Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least in $\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_q(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone}.
+Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone} for this sub-formula.
 As we did in Lemma \ref{lemmaone} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
 \begin{enumerate}
-\item[a)] We know that $\A\models\exists\x\:\varphi_q(\x,\P),$ so it immediately follows that $\A\models\exists\x\:v.$
-\item[b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
-(1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assigment to $\X$ for any of these formulas.
+\item[(a)] We know that $\A\models\exists\x\:\varphi_q(\x,\P),$ so it immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
+\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
+(1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
 (2) We can use the proof in Lemma \ref{lemmaone} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$.
@@
 And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
-\item[c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (1) and (2).
+\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
+As we deduce from the part (c) of Lemma \ref{lemmaone}, $\P$ is the only assignment to $\X$ such that $\A\not\models\exists\x\:\psi_q(\P)$, so necessarily $\A\models\exists\x\:\psi_q(\P^\prime)$.
+Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ Let $q^\prime$ be such that $\A\models\exists\x\:\varphi_{q^\prime}(\x,\P^\prime).$ By the choice of $q,$ then $q^\prime \geq q.$
 \end{enumerate}
-We will show that for every $j\in\{1,\ldots,k\}$, $\A\not\models\varphi_j^\prime(\x,\P).$ Since for any $i\in\{1,\ldots,q-1\}$ there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_i(\x,\R)$, then $\A\not\models\exists\x\:\varphi_1(\x,\P),\ldots,\A\not\models\exists\x\:\varphi_{q-1}(\x,\P),$ so it follows that $\A\not\models\exists\x\:\varphi^\prime_1(\x,\P),\ldots,\A\not\models\exists\x\:\varphi^\prime_{q-1}(\x,\P).$  Analogously to Lemma \ref{lemmaone}, we note that $\P$ is such that $\A\not\models\exists\x\:\psi_q(\x,\P),$ and then $\A\not\models\exists\x\:\varphi^\prime_q(\x,\P)$.
 

### 2015-01-21 sharp-sigma-one-plus.tex
21 ene v0.1

@@
 \end{eqnarray*}
-The formulas we will be using are $\varphi_i(\x,\X)$, for $i\in\{1,\ldots,k\},$ which are defined as follows: 
+We define the following formulas $\check{\varphi}_i(\x,\X)$, for $i\in\{1,\ldots,\check{k}\},$ as follows: 
 \begin{eqnarray*}
-\varphi_1(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^1(\x), \\
+\check{\varphi}_1(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^1(\x), \\
 & \vdots & \\
-\varphi_{g(d)}(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
-\varphi_{g(d)+1}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^1(\x), \\
+\check{\varphi}_{g(d)}(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
+\check{\varphi}_{g(d)+1}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^1(\x), \\
 & \vdots & \\
-\varphi_{2\cdot g(d)}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
+\check{\varphi}_{2\cdot g(d)}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
 & \vdots & \\
-\varphi_{(\hat{k}-1) g(d)+1}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^1(\x), \\
+\check{\varphi}_{(\hat{k}-1) g(d)+1}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^1(\x), \\
 & \vdots & \\
-\varphi_{\hat{k} \cdot g(d)}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^{g(d)}(\x).
+\check{\varphi}_{\hat{k} \cdot g(d)}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^{g(d)}(\x).
 \end{eqnarray*}
-Note that $k = \hat{k} \cdot g(d).$ Now we can define the formula like we did on the first part of the proof. For every $i\in\{1,\ldots,k\}$,
+Note that $\check{k} = \hat{k} \cdot g(d).$ Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L-$structure $\A$, that is, every $i\in\{1,\ldots,\check{k}\}$ such that for every assignment $\s$ to $\x,$ $f_{\check{\varphi}_i(\s,\X)} = 0.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
+We order the remaining disjuncts and define the disjuncts $\varphi_i(\x,\X)$ such that for every $j\in\{1,\ldots,\check{k}\}$ where $\check{\varphi}_j(\x,\X)$ is satisfiable, there is an $i\in\{1,\ldots,k\}$ such that $\varphi_i(\x,\X) = \check{\varphi}_j(\x,\X).$
+
+As seen in lemma \ref{lemmaone}, for every $\L-$structure $\A$ and $i\in\{1,\ldots,k\}$, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$ but $f_{\varphi_i(\s,\X)}(\A) = 0,$ then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) = 0$, so it follows that for every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) > 0.$
+
+Now we can define the formula like we did on the first part of the proof. For every $i\in\{1,\ldots,k\}$,
 \begin{eqnarray*}
-\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \hat{\varphi}^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
+\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
 \end{eqnarray*}
-where $\theta_i(\x) = \theta^{\,i\:(\!\!\!\!\mod g(d))}(\x).$ Also, note that
+ Also, note that
 \begin{eqnarray*}
@@
 \begin{multline*}
-\psi_i(\X) = \forall\x\:\neg\alpha(\x) \vee \exists\u_1\cdots\exists\u_n\bigg[\alpha^i_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
+\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_n\bigg[\alpha^i_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
 \end{multline*}
@@
 \begin{multline}
-\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\varphi_1(\v,\X)\vee\psi_2(\X)) \wedge \cdots \wedge \\ (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{i-1}(\v,\X)\vee\psi_i(\X)).
+\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge \\ (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
 \end{multline}
@@
 \item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
-As we deduce from the part (c) of Lemma \ref{lemmaone}, $\P$ is the only assignment to $\X$ such that $\A\not\models\exists\x\:\psi_q(\P)$, so necessarily $\A\models\exists\x\:\psi_q(\P^\prime)$.
-Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ Let $q^\prime$ be such that $\A\models\exists\x\:\varphi_{q^\prime}(\x,\P^\prime).$ By the choice of $q,$ then $q^\prime \geq q.$
+As we deduce from the part (c) of Lemma \ref{lemmaone}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
+Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ Let $q^\prime$ be such that $\A\models\exists\x\:\varphi_{q^\prime}(\x,\P^\prime).$ By the choice of $q,$ then $q^\prime \geq q,$ but since $\P \neq \P^\prime,$ then $q^\prime > q.$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x),$ so $\A\models\psi_i(\P^\prime)$ for each $i$.
+But also, $\A\models\exists\x\:\alpha_q(\x),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which is a contradiction.
 \end{enumerate}
+
+With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
+
+Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$.
+Let $\P$ be an arbitrary assignment to $\X$.
+Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
+It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
+And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
 

### 2015-01-21 sharp-sigma-one-plus.tex
21 ene

@@
 
-\item The counting set is:
+\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_d)$. That is,
 \begin{eqnarray*}
-F(\A) &=& \mid \{ \langle\P,\z\rangle \mid \A \models \exists \x \ \varphi(\x, \P,\z) \} \mid
+f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
 \end{eqnarray*}
-Then, we going to isolate the minimal predicate $\P$ that holds the formula true and eliminate the lexicographically smallest $\z$ that satisfies it. We mix both previous strategies.
+for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^d$.
+We define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2:
+\begin{eqnarray*}
+\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\v_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
+\end{eqnarray*}
+where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_i + c_i = a_i$ for $i\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, $v_i$ has $q_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\v) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
+The formulas $\varphi^{-}_i(\X,\y_i,\v_i)$, $\beta_i(\x,\z)$, are also defined analogously.
+
 \end{enumerate}

### 2015-01-26 sharp-sigma-one-plus.tex
26 ene

@@
 %% c - arity of auxiliar predicates/variables
+\def\c{\vec{c}} %% super auxiliar elements
 \def\d{\vec{d}} %% counted elements
@@
 %% t - size of \S
+\def\t{\vec{t}}
 \def\u{\vec{u}} %% auxiliary variables
@@
 \end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $a_1,\ldots,a_q$, and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.
+for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, and $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.
 
-Let $\X = (X_1,\dots,X_r)$, and let $\z$ be a tuple of variables.
-A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\X,\z)$ such that $f = f_{\exists \x \: \varphi(\x,\X,\z)}.$\\
+Let $\Y = (Y_1,\dots,Y_r)$, and let $\y$ be a tuple of variables.
+A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\exists \x \: \varphi(\x,\Y,\y)}.$\\
 
@@
 \end{eqnarray*}
-for every structure $\A$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
+for every $\L$-structure $\A \in \Truc$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
 
@@
 \begin{proof}
-For every $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$, let $\z = (z_1,\dots,z_d)$ and let $\X = (X_1,\dots,X_r)$.
+Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$.
+Without loss of generality, let $\z = (z_1,\dots,z_d)$ and let $\X = (X_1,\dots,X_r)$.
 There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$.
@@
 \end{eqnarray}
-where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$.
+where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
+Note that $\theta(\x)$ also mentions all variables in $\x$.
 We also assume that $(\x_1,\dots,\x_n,\y) = \x$.
 As an example, the following formula is of this form:
-\begin{align*}
-\varphi(\x,\X) =  X_1(x_1,x_2) \wedge \neg X_1(x_3,x_4) \wedge (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 ) \wedge \forall z\big( S_1(x_1,z) \big),
-\end{align*}
-where $\x = (x_1,x_2,x_3,x_4)$ and $\X = (X_1)$.
-Here, $n = 1$, $\lambda(1) = \lambda(2) = 1$, $\x_1 = (x_1,x_2)$ and $\y = (x_3,x_4)$.
-Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_3,x_4)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4)$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z) \big)$.
+\begin{multline*}
+\varphi(\x,\X) =  X_1(x_1,x_2) \wedge X_3(x_3) \wedge X_2(x_4,x_5) \wedge X_3(x_6) \wedge \neg X_1(x_7,x_8) \wedge \\ (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 ) \wedge \\ \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big),
+\end{multline*}
+where $\x = (x_1,x_2,x_3,x_4,x_5,x_6,x_7,x_8)$ and $\X = (X_1,X_2,X_3)$.
+Here, $n = 4$, $\lambda(1) = 1$, $\lambda(2) = \lambda(4) = 3$ and $\lambda(3) = 2$, $\x_1 = (x_1,x_2)$, $\x_2 = (x_3)$, $\x_3 = (x_4,x_5)$, $\x_4 = (x_6)$ and $\y = (x_7,x_8)$.
+Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_7,x_8)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 )$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2,x_5,x_8\},\{x_6,x_7\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big)$.
 
@@
 \begin{multline}
-\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge\exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \\ \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg] \label{f2}.
+\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge \\ \exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg] \label{f2}.
 \end{multline}
 We now show a result by which the main proof will follow.
-\begin{lemma}  \label{lemmaone}
+\begin{lemma}  \label{first}
 $f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
@@
 Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$.
-We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \z < \x_i$ or (3) $\theta(\x)\models \z > \x_i$.
-Considering that (2) and (3) are not possible since $\A\models\theta(\d)$, and both $\z$ and $\x_i$ are assigned the value $\d_i$, then $\theta(\x)\models \z = \x_i$.
-But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$ from which we conclude that $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which leads to a contradiction.
-\item[(b)] By the construction of $\P$, we see that 
-$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big).$$ However, the only possible assignment to $\alpha_{\min}(\x_1,\dots,\x_n)$ is $(\d_1,\dots,\d_n)$.
-Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
+We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \varphi_{<,a_i}(\z,\x_i)$ or (3) $\theta(\x)\models \varphi_{<,a_i}(\x_i,\z)$.
+Considering that (2) and (3) are not possible given that both $\z$ and $\x_i$ are assigned the value $\d_i$ and $\A\models\theta(\d)$, we have that $\theta(\x)\models \z = \x_i$.
+But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$, from which we conclude that $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which contradicts the initial assumption that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
+\item[(b)] Note that if $\A \models \alpha_{\min}(\c_1,\ldots,\c_n)$, then necessarily $\c_i = \d_i$ for $i\in\{1,\ldots,n\}.$ However, by the construction of $\P$, we see that 
+$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \d_j\Big).$$ Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
 \item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
 We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$.
-Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $\P^\prime$, then $\A\models \neg P^\prime_i(\d_i)$.
-And so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $\P_i$ does not have.
+Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $P^\prime_i$.
+Then $\A\models \neg P^\prime_i(\d_i)$, and so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
+Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $P_i$ does not have.
 If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
@@
 \end{enumerate}
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
+With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
 
@@
 It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
+And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ it holds that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1,$ which was to be shown.
 \end{proof}
@@
 We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
-By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\hat{\varphi}_i(\x,\X)$ with $i\in\{1,\ldots,\hat{k}\}$ such that
+By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\gamma_i(\x,\X)$ with $i\in\{1,\ldots,\ell\}$ such that
 \begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \hat{\varphi}_1(\x,\X) \vee \hat{\varphi}_2(\x,\X) \vee \dots  \vee \hat{\varphi}_{\hat{k}}(\x,\X),
+\varphi(\x,\X) &\equiv& \gamma_1(\x,\X) \vee \gamma_2(\x,\X) \vee \dots  \vee \gamma_{\ell}(\x,\X),
 \end{eqnarray*}
-where, for every $i\in\{1,\ldots,\hat{k}\}$, 
+where, for every $i\in\{1,\ldots,\ell\}$, 
 \begin{eqnarray*}
-\hat{\varphi}_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \hat{\varphi}^{-}_i(\X,\y_i)  \wedge \beta_i(\x).
+\gamma_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \gamma^{-}_i(\X,\y_i)  \wedge \delta_i(\x).
 \end{eqnarray*}
-The function $\lambda_i$ is defined analogously to the first part of the proof, the tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x.$ The formulas $\hat{\varphi}^{-}_i(\X,\y_i)$, $\beta_i(\x)$, are also defined analogously.
+The function $\lambda_i$ is defined analogously to the first part of the proof, the tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x.$ The formulas $\gamma^{-}_i(\X,\y_i)$, $\delta_i(\x)$, are also defined analogously to the first part of the proof (refer to \eqref{f1}).
 Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables.
 Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$.
-Note that for every formula $\psi(\x)$,
+Note that for every formula $\eta(\x)$,
 \begin{eqnarray*}
-\exists\x\:\psi(\x) &\equiv& \exists\x(\psi(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\psi(\x)\wedge\theta^{g(d)}(\x)).
+\exists\x\:\eta(\x) &\equiv& \exists\x(\eta(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\eta(\x)\wedge\theta^{g(d)}(\x)).
 \end{eqnarray*}
-We define the following formulas $\check{\varphi}_i(\x,\X)$, for $i\in\{1,\ldots,\check{k}\},$ as follows: 
+We define the following formulas $\xi_i(\x,\X)$, for $i\in\{1,\ldots,m\},$ as follows: 
 \begin{eqnarray*}
-\check{\varphi}_1(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^1(\x), \\
+\xi_1(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^1(\x), \\
 & \vdots & \\
-\check{\varphi}_{g(d)}(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
-\check{\varphi}_{g(d)+1}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^1(\x), \\
+\xi_{g(d)}(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
+\xi_{g(d)+1}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^1(\x), \\
 & \vdots & \\
-\check{\varphi}_{2\cdot g(d)}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
+\xi_{2\cdot g(d)}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
 & \vdots & \\
-\check{\varphi}_{(\hat{k}-1) g(d)+1}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^1(\x), \\
+\xi_{(\ell-1) g(d)+1}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^1(\x), \\
 & \vdots & \\
-\check{\varphi}_{\hat{k} \cdot g(d)}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^{g(d)}(\x).
+\xi_{\ell \cdot g(d)}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^{g(d)}(\x).
 \end{eqnarray*}
-Note that $\check{k} = \hat{k} \cdot g(d).$ Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L-$structure $\A$, that is, every $i\in\{1,\ldots,\check{k}\}$ such that for every assignment $\s$ to $\x,$ $f_{\check{\varphi}_i(\s,\X)} = 0.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
-We order the remaining disjuncts and define the disjuncts $\varphi_i(\x,\X)$ such that for every $j\in\{1,\ldots,\check{k}\}$ where $\check{\varphi}_j(\x,\X)$ is satisfiable, there is an $i\in\{1,\ldots,k\}$ such that $\varphi_i(\x,\X) = \check{\varphi}_j(\x,\X).$
+Note that $m = \ell \cdot g(d).$ Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\Truc.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
+We use a function $\rho:\{1,\ldots,k\}\to\{1,\ldots,m\}$ such that $\varphi_i(\x,\X) = \xi_{\rho(i)}(\x,\X)$ is satisfiable, for every $i\in\{1,\ldots,k\}.$
 
-As seen in lemma \ref{lemmaone}, for every $\L-$structure $\A$ and $i\in\{1,\ldots,k\}$, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$ but $f_{\varphi_i(\s,\X)}(\A) = 0,$ then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) = 0$, so it follows that for every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) > 0.$
+\begin{claim}
+For every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, then $f_{\varphi_i(\s,\X)}(\A) > 0.$
+\end{claim}
+\begin{proof}
+By contradiction, assume that there is an assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, and $f_{\varphi_i(\s,\X)}(\A) = 0.$ That is, there is no assignment $\R$ to $\X$ such that $\A\models\varphi_i(\s,\R).$ Since $\A\models\alpha_i(\s)$, it holds that $\A\models\theta_i(\s)$.
+Therefore, as we showed in part a) of Lemma \ref{first}, there is no assignment $\s^\prime$ to $\x$ such that $f_{\varphi_i(\s,\X)}(\A) > 0.$ This means $\varphi_i(\x,\X)$ is unsatisfiable, but, as we mentioned previously, all of such disjuncts have been eliminated, which is a contradiction.
+\end{proof}
 
@@
 \end{eqnarray*}
+where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\y_i$ is a $p_i$-tuple of variables in $\x$, $\varphi_i^{-}(\X,\y_i)$ is a conjunction of negated predicates in $\X$, $\theta_i(\x)$ defines a total order on a partition of $\x$, and $\beta_i(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
 Also, note that
@@
 
-Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\P$ such that $\A\models\varphi_2(\x,\P)$, and so on.
+Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\exists\x\:\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\P$ such that $\A\models\exists\x\:\varphi_2(\x,\P)$, and so on.
 As we already know how to exclude that assignment in the first disjunct, we now move to the next disjuncts.
@@
 \begin{multline*}
-\alpha^i_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_n,\w)\wedge\bigvee_{j=1}^n(\u_j\neq\v_j)\big)\to \varphi_{m_i,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
+\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_{n_i},\y)\wedge \\ \forall\v_1\cdots\forall\v_{n_i}\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_{n_i},\w)\wedge\bigvee_{j=1}^{n_i}(\u_j\neq\v_j)\big)\to \varphi_{m_i,<}((\u_1,\dots,\u_{n_i}),(\v_1,\dots,\v_{n_i}))\Big)\Big],
 \end{multline*}
@@
 \begin{multline*}
-\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_n\bigg[\alpha^i_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
+\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
 \end{multline*}
@@
 Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$.
-Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone} for this sub-formula.
-As we did in Lemma \ref{lemmaone} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
+Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{first} for this sub-formula.
+As we did in Lemma \ref{first} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
 \begin{enumerate}
@@
 (1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
-(2) We can use the proof in Lemma \ref{lemmaone} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$.
+(2) We can use the proof in Lemma \ref{first} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$.
 However, $\A\not\models\exists\x\:\hat{\psi}_q(\P)$ so $\A\not\models\exists\x\:\varphi^\prime_i(\P)$.
@@
 \item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
-As we deduce from the part (c) of Lemma \ref{lemmaone}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
+As we deduce from the part (c) of Lemma \ref{first}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
 Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ Let $q^\prime$ be such that $\A\models\exists\x\:\varphi_{q^\prime}(\x,\P^\prime).$ By the choice of $q,$ then $q^\prime \geq q,$ but since $\P \neq \P^\prime,$ then $q^\prime > q.$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x),$ so $\A\models\psi_i(\P^\prime)$ for each $i$.
@@
 
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_d)$. That is,
+\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^d$.
+for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$.
 We define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2:
@@
 
+In this case, we mix both previous strategies, that is, we are going to {\em isolate} the lexicographically smallest predicate that satisfies the first satisfiable disjunct, and exclude the lexicographically smallest tuple that satisfies the evaluated disjunct.
+
+Let
+\begin{multline*}
+\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\u\exists\w(\alpha_i(\u,\w) \wedge \varphi_{<,p}(\w,\z)) \\.
+\end{multline*}
+
 \end{enumerate}

### 2015-01-30 sharp-sigma-one-plus.tex
30 ene

@@
 \end{eqnarray*}
-The function $\lambda_i$ is defined analogously to the first part of the proof, the tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x.$ The formulas $\gamma^{-}_i(\X,\y_i)$, $\delta_i(\x)$, are also defined analogously to the first part of the proof (refer to \eqref{f1}).
+The function $\lambda_i$ is defined analogously to $\lambda$ of the first part of the proof.
+The tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y_i) = \x.$ The formulas $\gamma^{-}_i(\X,\y_i)$, $\delta_i(\x)$ are defined analogously to $\varphi^{-}(\X,\y)$ and $\beta(\x)$\footnote[2]{Note that $\delta_i(\x)$ can include subformulas of the form $(x_j = x_j)$.} respectively (see \eqref{f1}).
 Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables.
@@
 \end{eqnarray*}
-We define the following formulas $\xi_i(\x,\X)$, for $i\in\{1,\ldots,m\},$ as follows: 
+We define the following formulas $\xi_i(\x,\X)$, for $i\in\{1,\ldots,m\}$ where $m = \ell \cdot g(d)$, as follows: 
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-Note that $m = \ell \cdot g(d).$ Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\Truc.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
-We use a function $\rho:\{1,\ldots,k\}\to\{1,\ldots,m\}$ such that $\varphi_i(\x,\X) = \xi_{\rho(i)}(\x,\X)$ is satisfiable, for every $i\in\{1,\ldots,k\}.$
+Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\Truc.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
+We use an injective\footnote[3]{we need this function to be injective because that way we can assure each one of the $k$ satisfiable disjuncts to be represented} function $\rho:\{1,\ldots,k\}\to\{1,\ldots,m\}$ such that $\varphi_i(\x,\X) = \xi_{\rho(i)}(\x,\X)$ is satisfiable, for every $i\in\{1,\ldots,k\}.$
+
+We can conclude that
+\begin{eqnarray*}
+\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X),
+\end{eqnarray*}
+and for every $i\in\{1,\ldots,k\}$,
+\begin{eqnarray*}
+\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
+\end{eqnarray*}
+where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\y_i$ is a $p_i$-tuple of variables in $\x$, $\varphi_i^{-}(\X,\y_i)$ is a conjunction of negated predicates in $\X$, $\theta_i(\x)$ defines a total order on a partition of $\x$, and $\beta_i(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
+Let $\alpha_i(\x) = \theta_i(\x)\wedge\beta_i(\x).$
 
 \begin{claim}
-For every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, then $f_{\varphi_i(\s,\X)}(\A) > 0.$
+Let $\A\in\Truc.$ For every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, it holds that $f_{\varphi_i(\s,\X)}(\A) > 0.$
 \end{claim}
@@
 \end{proof}
-
-Now we can define the formula like we did on the first part of the proof. For every $i\in\{1,\ldots,k\}$,
-\begin{eqnarray*}
-\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
-\end{eqnarray*}
-where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\y_i$ is a $p_i$-tuple of variables in $\x$, $\varphi_i^{-}(\X,\y_i)$ is a conjunction of negated predicates in $\X$, $\theta_i(\x)$ defines a total order on a partition of $\x$, and $\beta_i(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
-Also, note that
-\begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X).
-\end{eqnarray*}
 
@@
 \end{multline*}
-where $\u_j,\v_j,$ for $\{1,\ldots,n_i\}$ and $\w$ have the adequate number of variables. Also, let
+where $\u_j,\v_j,$ for $\{1,\ldots,n_i\}$ and $\w$ have the corresponding number of variables. Also, let
 \begin{multline*}
@@
 \end{multline*}
-Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
+Note that $\psi_i(\X)$ excludes only the lexicographically smallest predicate $\P$ such that $\A\models\exists\x\:\varphi_i(\X,\x),$ if there is at least one.
+In other words, every assignment $\P^\prime \neq \P$ is such that $\A\models\psi_i(\P^\prime).$ Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
 \begin{multline}
@@
 Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$.
-Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{first} for this sub-formula.
+Suppose first that there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least one assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\d = (\d_1,\dots,\d_{n_q},\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha_q(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n_q\}$, and $\ep$ is an assignment for $\y$.
+Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{\lambda_q(j)=i}\{\d_j\}$.
 As we did in Lemma \ref{first} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
 \begin{enumerate}
-\item[(a)] We know that $\A\models\exists\x\:\varphi_q(\x,\P),$ so it immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
+\item[(a)] As we showed in part (a) of Lemma \ref{first}, this assignment $\P$ to $\X$ is such that $\A\models\exists\x\:\varphi_q(\x,\P),$ if and only if there is at least one such an assignment.
+However, as we showed in Claim Six, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_q(\s)$, then there {\em is} such an assignment to $\X$.
+The assignment $\d$ to $\x$ satisfies that so we see that $\A\models\exists\x\:\varphi_q(\x,\P)$.
+It immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
+
 \item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
 (1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
-(2) We can use the proof in Lemma \ref{first} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$.
-However, $\A\not\models\exists\x\:\hat{\psi}_q(\P)$ so $\A\not\models\exists\x\:\varphi^\prime_i(\P)$.
+(2) We can use the proof in Lemma \ref{first} to see that $\A\not\models\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears as a conjunct in $\varphi^\prime_i(\x,\X)$.
+However, since $\A\not\models\exists\v\:\varphi_i(\v,\P)$ for $i\in\{1,\ldots,q-1\}$ and $\A\not\models\psi_q(\P)$, it holds that $\A\not\models\hat{\psi}_q(\P)$, so $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$.
 And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
+
 \item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
 As we deduce from the part (c) of Lemma \ref{first}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
-Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ Let $q^\prime$ be such that $\A\models\exists\x\:\varphi_{q^\prime}(\x,\P^\prime).$ By the choice of $q,$ then $q^\prime \geq q,$ but since $\P \neq \P^\prime,$ then $q^\prime > q.$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x),$ so $\A\models\psi_i(\P^\prime)$ for each $i$.
-But also, $\A\models\exists\x\:\alpha_q(\x),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which is a contradiction.
+Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x),$ so $\A\models\psi_i(\P^\prime)$ for each $i$.
+But also, $\A\models\exists\x\:\psi_q(\P^\prime),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which is a contradiction.
 \end{enumerate}
@@
 Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
+Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\chi(\x,\P))$ for any formula $\chi(\x,\P)$.
 It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
+And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Therefore, $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1 = f\dotminus 1.$
+
+Finally, let $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}.$ We conclude that $f^\prime = f\dotminus 1\in\E1$
 
@@
 for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$.
-We define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2:
+We define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2, where
 \begin{eqnarray*}
-\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\v_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
+\varphi(\x,\X,\z) \equiv \varphi_1(\x,\X,\z) \vee \cdots \vee \varphi_k(\x,\X,\z),
 \end{eqnarray*}
-where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_i + c_i = a_i$ for $i\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, $v_i$ has $q_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\v) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
-The formulas $\varphi^{-}_i(\X,\y_i,\v_i)$, $\beta_i(\x,\z)$, are also defined analogously.
+and
+\begin{eqnarray*}
+\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\w_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
+\end{eqnarray*}
+where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\ell} + c_{\ell} = a_{\ell}$ for $\ell\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, $w_i$ has $q_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\w_i) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
+The formulas $\varphi^{-}_i(\X,\y_i,\w_i)$, $\beta_i(\x,\z)$, are also defined analogously.
 
@@
 
-Let
+Let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$. Let
 \begin{multline*}
-\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\u\exists\w(\alpha_i(\u,\w) \wedge \varphi_{<,p}(\w,\z)) \\.
+\alpha^i_{\min}(\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}) = \exists\y\exists\w\Big[ \alpha_i(\x_1,\dots,\x_{n_i},\y,\z_1,\dots,\z_{n_i},\w)\wedge \\ \forall\u_1\cdots\forall\u_{n_i}\forall\s\forall\v_1\cdots\forall\v_{n_i}\forall\t\Big(\big(\alpha_i(\u_1,\dots,\u_{n_i},\s,\v_1,\dots,\v_{n_i},\t)\wedge\bigvee_{j=1}^{n_i}(\x_j\neq\u_j \vee \z_j\neq\v_j)\big)\to \\ \varphi_{m_i,<}((\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}),(\u_1,\dots,\u_{n_i},\v_1,\dots,\v_{n_i}))\Big)\Big],
 \end{multline*}
+and let
+\begin{multline*}
+\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\w(\exists\u\:\varphi(\u,\X,\w) \wedge \varphi_{<,p}(\w,\z)).
+\end{multline*}
+Let $\P$ and $\e$ be assignments to $\X$ and $\z$.
+Note that $\psi_i(\X,\z)$ excludes $\P$ and $\e$ only if $\P$ is the lexicographically smallest predicate such that $\A\models\exists\x\:\varphi_i(\x,\P,\d)$, for some assignment $\d$ to $\z$, and $\e$ is the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_i(\x,\P,\e).$ We define $\varphi_i^\prime(\x,\X,\z)$ in the same way as case 2:
+\begin{multline*}
+\varphi_i^\prime(\x,\X,\z) = \varphi_i(\x,\X,\z) \wedge \psi_1(\X,\z) \wedge (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\psi_2(\X,\z)) \wedge \cdots \wedge \\ (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\cdots\vee\exists\u\exists\v\:\alpha_{i-1}(\u,\v)\vee\psi_i(\X,\z)).
+\end{multline*}
+Finally, let $\varphi^\prime(\x,\X,\z) = \bigvee_{i = 1}^k \varphi_i^\prime(\x,\X,\z)$.
+
+Let $q$ be the least $i\in\{1,\ldots,k\}$ such that $\A\models\exists\x\:\varphi_i(\x,\R,\d)$ for some assignment $R$ to $\X$ and some assignment $\d$ to $\z$.
+Let $\P$ be the lexicographically smallest predicate such that $\A\models\exists\x\:\varphi_q(\x,\P,\d^\prime)$ for some assignment $\d^\prime$ to $\z$.
+Let $\e$ be the lexicographically smallest assignment to $z$ such that $\A\models\exists\x\:\varphi_q(\x,\P,\e).$ This formula is such that (a) $\A\models\exists\x\:\varphi(\x,\P,\e)$, (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P,\e)$ and $\P$ and $\e$ are the only assignments that satisfy (a) and (b).
+The proof of this is analogous to Case 2.
+We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X,\z)} = f_{\exists\x\,\varphi(\x,\X,\z)}\dotminus 1.$
 

### 2015-04-08 tex/case1.tex
8 abril

--- 2015-01-21 sharp-sigma-one-plus.tex
@@
-\documentclass[12pt]{article}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsthm} 
-\usepackage{fullpage}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{bm}
-
-\def\dotminus{\mathbin{\ooalign{\hss\raise1ex\hbox{.}\hss\cr
-  \mathsurround=0pt$-$}}}
-
-\def\E1{\#\Sigma_1^{+}}
-
-\def\Truc{\textsc{Struct}[\L]}
-
-\def\A{{\frak A}}
-\def\C{{\cal C}}
-\def\F{{\cal F}}
-\def\L{{\cal L}}
-\def\N{\mathbb{N}}
-\def\P{\vec{P}}
-\def\R{\vec{R}}
-\def\S{\vec{S}}
-\def\X{\vec{X}}
-\def\Y{\vec{Y}}
-\def\Z{\vec{Z}}
-%% a - arity of \X / arity of assignments \P to \X
-%% b - arity of predicates in \S
-%% c - arity of auxiliar predicates/variables
-\def\d{\vec{d}} %% counted elements
-\def\e{\vec{e}} %% counted elements
-%% f - counting function
-%% g - other functions
-%% h - other functions
-%% i - index
-%% j - index
-%% k - emergency index / size of tuple
-%% l - emergency index / size of tuple
-\def\l{\vec{\ell}}
-%% m - size of variable tuple
-%% n - size of predicate tuple
-%% o - not used
-%% p - 
-%% q - 
-%% r - size of \X / \P
-\def\s{\vec{s}}
-%% t - size of \S
-\def\u{\vec{u}} %% auxiliary variables
-\def\v{\vec{v}} %% auxiliary variables
-\def\w{\vec{w}} %% auxiliary variables
-\def\x{\vec{x}} %% quantified variables
-\def\y{\vec{y}} %% auxiliary variables
-\def\z{\vec{z}} %% open variables
-\def\ep{\vec{o}}
-\def\ga{\vec{p}}
-
-\newtheorem{theo}{Theorem}
-\newtheorem{lemma}[theo]{Lemma}
-\newtheorem{claim}[theo]{Claim}
-\newtheorem{coro}[theo]{Corollary}
-
-\begin{document}
-
-\begin{center}
-{ \LARGE \bf
-  Some properties of $\E1$
-}
-\end{center}
-
-We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
-\begin{eqnarray*}
-\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
-&& \leq \text{ is interpreted as a total order for } A \}.
-\end{eqnarray*}
-We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$.
-A quantifier-free $\L$-formula is defined by the following grammar:
-\begin{eqnarray*}
-\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
-&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
-&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
-\end{eqnarray*}
-where $x,y$ and $x_i$ are first order variables for every $i$. We now define an extended quantifier-free $\L$-formula as follows:
-\begin{eqnarray*}
-\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
-&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
-&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
-\end{eqnarray*}
-Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables.
-For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
-\begin{eqnarray*}
-f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $a_1,\ldots,a_q$, and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.
-
-Let $\X = (X_1,\dots,X_r)$, and let $\z$ be a tuple of variables.
-A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\X,\z)$ such that $f = f_{\exists \x \: \varphi(\x,\X,\z)}.$\\
-
-The {\em decision version} of a function $f$ is defined by the language $L_f = \{\A \mid f(\A) > 0\}$.
-\begin{theo}
-The decision version of a function in $\E1$ is in \textsc{P}.
-\end{theo}
-\begin{proof}
-Let $f$ be a function in $\E1$. Then there is a formula $\varphi(\x,\X,\z)$ such that
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $\z$ is an $m$-tuple of variables, $\e$ is an $m$-tuple of elements and $\x$ is a $k$-tuple of variables.
-Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible evaluations for $\z$.
-Let $\ell = \vert A \vert^k$ and $\d_1,\dots,d_\ell\in A^k$ be all possible evaluations for $\x$.
-Let $\varphi_\A(\X)$ be defined as follows:
-\begin{eqnarray*}
-\varphi_\A(\X) = \varphi(\d_1,\X,\e_1) \vee \cdots \vee \varphi(\d_1,\X,\e_n) \vee \varphi(\d_2,\X,\e_1) \vee \cdots \vee \varphi(\d_n,\X,\e_n).
-\end{eqnarray*}
-Note that $\varphi_\A(\X)$ will have at least one assignment for $\X$ iff $f(\A)>0$.
-Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction.
-Note that $\psi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$.
-Then, let $f'$ be defined as follows:
-\begin{eqnarray*}
-f'(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \psi_\A(\P) \} \mid.
-\end{eqnarray*}
-Note that $f' \in\#\Sigma_0$ \footnote[1]{(paper)}.
-Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
-However, it is showed in $^{[1]}$ that any counting problem in $\#\Sigma_0$ is computable in polynomial time, therefore, the decision version of $f$ is in P.
-\end{proof}
-
-\begin{theo}
-$\#\Sigma_1$ is closed under substraction $\Rightarrow$ P = NP.
-\end{theo}
-\begin{proof}
-\#3DNF $\in \#\Sigma_1$.
-Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment in a \#3DNF instance.
-Suppose that $F_{2^n}-F_{\#3DNF} \in \#\Sigma_1$.
-This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete.
-However, as we showed previously (Theorem 1), it is also in P.
-Then, co-NP $\subseteq$ P.
-\end{proof}
-
-\begin{theo}
-$\#\Sigma_1 \subsetneq \E1$
-\end{theo}
-\begin{proof}
-We will show that the $\E1$ function defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
-\end{proof}
-
-For a given function $f$, we define $f \dotminus 1$ as follows:
-\begin{eqnarray*}
-f \dotminus 1(\A) =
-\begin{cases}
-f(\A)-1, & \text{if }f(\A) > 0 \\
-0, & \text{if }f(\A) = 0.
-\end{cases}
-\end{eqnarray*}
-for every structure $\A$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
-
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1
-
-\begin{theo}
-$\E1$ is closed under substraction by one.
-\end{theo}
-\begin{proof}
-For every $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$, let $\z = (z_1,\dots,z_d)$ and let $\X = (X_1,\dots,X_r)$.
-There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$.
-This separates the proof in three cases:
-\begin{enumerate}
-
-% CASE 1
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
+Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
 \begin{eqnarray*}
@@
 Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E1.$
-
-%CASE 2
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
-\begin{eqnarray}
-\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
-\end{eqnarray}
-where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$.
-We also assume that $(\x_1,\dots,\x_n,\y) = \x$.
-As an example, the following formula is of this form:
-\begin{align*}
-\varphi(\x,\X) =  X_1(x_1,x_2) \wedge \neg X_1(x_3,x_4) \wedge (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 ) \wedge \forall z\big( S_1(x_1,z) \big),
-\end{align*}
-where $\x = (x_1,x_2,x_3,x_4)$ and $\X = (X_1)$.
-Here, $n = 1$, $\lambda(1) = \lambda(2) = 1$, $\x_1 = (x_1,x_2)$ and $\y = (x_3,x_4)$.
-Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_3,x_4)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4)$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z) \big)$.
-
-Similarly to the previous proof, we would like to eliminate the lexicographically smallest tuple of predicates that satisfies the formula \eqref{f1}.
-Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
-We now define
-\begin{multline*}
-\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to \varphi_{m,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
-\end{multline*}
-where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
-Note that $\alpha_{\min}$ is satisfied only by the lexicographically \linebreak smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
-Our new formula is
-\begin{multline}
-\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge\exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \\ \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg] \label{f2}.
-\end{multline}
-We now show a result by which the main proof will follow.
-\begin{lemma}  \label{lemmaone}
-$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
-\end{lemma}
-\begin{proof}
-Let $\A \in \Truc$.
-Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{\lambda(j)=i}\{\d_j\}$.
-We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
-\begin{enumerate}
-\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$.
-That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$).
-Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$.
-Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$.
-We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \z < \x_i$ or (3) $\theta(\x)\models \z > \x_i$.
-Considering that (2) and (3) are not possible since $\A\models\theta(\d)$, and both $\z$ and $\x_i$ are assigned the value $\d_i$, then $\theta(\x)\models \z = \x_i$.
-But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$ from which we conclude that $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which leads to a contradiction.
-\item[(b)] By the construction of $\P$, we see that 
-$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big).$$ However, the only possible assignment to $\alpha_{\min}(\x_1,\dots,\x_n)$ is $(\d_1,\dots,\d_n)$.
-Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
-\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$.
-Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $\P^\prime$, then $\A\models \neg P^\prime_i(\d_i)$.
-And so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $\P_i$ does not have.
-If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-On both cases, we have a contradiction.
-\end{enumerate}
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi(\x,\P)$, we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
-\end{proof}
-
-We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
-By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\hat{\varphi}_i(\x,\X)$ with $i\in\{1,\ldots,\hat{k}\}$ such that
-\begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \hat{\varphi}_1(\x,\X) \vee \hat{\varphi}_2(\x,\X) \vee \dots  \vee \hat{\varphi}_{\hat{k}}(\x,\X),
-\end{eqnarray*}
-where, for every $i\in\{1,\ldots,\hat{k}\}$, 
-\begin{eqnarray*}
-\hat{\varphi}_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \hat{\varphi}^{-}_i(\X,\y_i)  \wedge \beta_i(\x).
-\end{eqnarray*}
-The function $\lambda_i$ is defined analogously to the first part of the proof, the tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x.$ The formulas $\hat{\varphi}^{-}_i(\X,\y_i)$, $\beta_i(\x)$, are also defined analogously.
-Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables.
-Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$.
-Note that for every formula $\psi(\x)$,
-\begin{eqnarray*}
-\exists\x\:\psi(\x) &\equiv& \exists\x(\psi(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\psi(\x)\wedge\theta^{g(d)}(\x)).
-\end{eqnarray*}
-We define the following formulas $\check{\varphi}_i(\x,\X)$, for $i\in\{1,\ldots,\check{k}\},$ as follows: 
-\begin{eqnarray*}
-\check{\varphi}_1(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{g(d)}(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
-\check{\varphi}_{g(d)+1}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{2\cdot g(d)}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
-& \vdots & \\
-\check{\varphi}_{(\hat{k}-1) g(d)+1}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{\hat{k} \cdot g(d)}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^{g(d)}(\x).
-\end{eqnarray*}
-Note that $\check{k} = \hat{k} \cdot g(d).$ Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L-$structure $\A$, that is, every $i\in\{1,\ldots,\check{k}\}$ such that for every assignment $\s$ to $\x,$ $f_{\check{\varphi}_i(\s,\X)} = 0.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
-We order the remaining disjuncts and define the disjuncts $\varphi_i(\x,\X)$ such that for every $j\in\{1,\ldots,\check{k}\}$ where $\check{\varphi}_j(\x,\X)$ is satisfiable, there is an $i\in\{1,\ldots,k\}$ such that $\varphi_i(\x,\X) = \check{\varphi}_j(\x,\X).$
-
-As seen in lemma \ref{lemmaone}, for every $\L-$structure $\A$ and $i\in\{1,\ldots,k\}$, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$ but $f_{\varphi_i(\s,\X)}(\A) = 0,$ then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) = 0$, so it follows that for every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) > 0.$
-
-Now we can define the formula like we did on the first part of the proof. For every $i\in\{1,\ldots,k\}$,
-\begin{eqnarray*}
-\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
-\end{eqnarray*}
- Also, note that
-\begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X).
-\end{eqnarray*}
-
-Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\P$ such that $\A\models\varphi_2(\x,\P)$, and so on.
-As we already know how to exclude that assignment in the first disjunct, we now move to the next disjuncts.
-Similarly to the first part of the proof, let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
-\begin{multline*}
-\alpha^i_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_n,\w)\wedge\bigvee_{j=1}^n(\u_j\neq\v_j)\big)\to \varphi_{m_i,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
-\end{multline*}
-where $\u_j,\v_j,$ for $\{1,\ldots,n_i\}$ and $\w$ have the adequate number of variables. Also, let
-\begin{multline*}
-\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_n\bigg[\alpha^i_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
-\end{multline*}
-Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
-\begin{multline}
-\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge \\ (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
-\end{multline}
-Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$.
-Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone} for this sub-formula.
-As we did in Lemma \ref{lemmaone} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
-\begin{enumerate}
-\item[(a)] We know that $\A\models\exists\x\:\varphi_q(\x,\P),$ so it immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
-\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
-(1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
-(2) We can use the proof in Lemma \ref{lemmaone} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$.
-However, $\A\not\models\exists\x\:\hat{\psi}_q(\P)$ so $\A\not\models\exists\x\:\varphi^\prime_i(\P)$.
-And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
-\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
-As we deduce from the part (c) of Lemma \ref{lemmaone}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
-Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ Let $q^\prime$ be such that $\A\models\exists\x\:\varphi_{q^\prime}(\x,\P^\prime).$ By the choice of $q,$ then $q^\prime \geq q,$ but since $\P \neq \P^\prime,$ then $q^\prime > q.$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x),$ so $\A\models\psi_i(\P^\prime)$ for each $i$.
-But also, $\A\models\exists\x\:\alpha_q(\x),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which is a contradiction.
-\end{enumerate}
-
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
-
-% CASE 3
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_d)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^d$.
-We define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2:
-\begin{eqnarray*}
-\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\v_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
-\end{eqnarray*}
-where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_i + c_i = a_i$ for $i\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, $v_i$ has $q_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\v) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
-The formulas $\varphi^{-}_i(\X,\y_i,\v_i)$, $\beta_i(\x,\z)$, are also defined analogously.
-
-\end{enumerate}
-\end{proof}
-
-\end{document}

### 2015-04-08 tex/case2general.tex
8 abril

--- 2015-01-26 sharp-sigma-one-plus.tex
@@
- \documentclass[12pt]{article}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsthm} 
-\usepackage{fullpage}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{bm}
-
-\def\dotminus{\mathbin{\ooalign{\hss\raise1ex\hbox{.}\hss\cr
-  \mathsurround=0pt$-$}}}
-
-\def\E1{\#\Sigma_1^{+}}
-
-\def\Truc{\textsc{Struct}[\L]}
-
-\def\A{{\frak A}}
-\def\C{{\cal C}}
-\def\F{{\cal F}}
-\def\L{{\cal L}}
-\def\N{\mathbb{N}}
-\def\P{\vec{P}}
-\def\R{\vec{R}}
-\def\S{\vec{S}}
-\def\X{\vec{X}}
-\def\Y{\vec{Y}}
-\def\Z{\vec{Z}}
-%% a - arity of \X / arity of assignments \P to \X
-%% b - arity of predicates in \S
-%% c - arity of auxiliar predicates/variables
-\def\c{\vec{c}} %% super auxiliar elements
-\def\d{\vec{d}} %% counted elements
-\def\e{\vec{e}} %% counted elements
-%% f - counting function
-%% g - other functions
-%% h - other functions
-%% i - index
-%% j - index
-%% k - emergency index / size of tuple
-%% l - emergency index / size of tuple
-\def\l{\vec{\ell}}
-%% m - size of variable tuple
-%% n - size of predicate tuple
-%% o - not used
-%% p - 
-%% q - 
-%% r - size of \X / \P
-\def\s{\vec{s}}
-%% t - size of \S
-\def\t{\vec{t}}
-\def\u{\vec{u}} %% auxiliary variables
-\def\v{\vec{v}} %% auxiliary variables
-\def\w{\vec{w}} %% auxiliary variables
-\def\x{\vec{x}} %% quantified variables
-\def\y{\vec{y}} %% auxiliary variables
-\def\z{\vec{z}} %% open variables
-\def\ep{\vec{o}}
-\def\ga{\vec{p}}
-
-\newtheorem{theo}{Theorem}
-\newtheorem{lemma}[theo]{Lemma}
-\newtheorem{claim}[theo]{Claim}
-\newtheorem{coro}[theo]{Corollary}
-
-\begin{document}
-
-\begin{center}
-{ \LARGE \bf
-  Some properties of $\E1$
-}
-\end{center}
-
-We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
-\begin{eqnarray*}
-\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
-&& \leq \text{ is interpreted as a total order for } A \}.
-\end{eqnarray*}
-We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$.
-A quantifier-free $\L$-formula is defined by the following grammar:
-\begin{eqnarray*}
-\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
-&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
-&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
-\end{eqnarray*}
-where $x,y$ and $x_i$ are first order variables for every $i$. We now define an extended quantifier-free $\L$-formula as follows:
-\begin{eqnarray*}
-\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
-&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
-&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
-\end{eqnarray*}
-Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables.
-For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
-\begin{eqnarray*}
-f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, and $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.
-
-Let $\Y = (Y_1,\dots,Y_r)$, and let $\y$ be a tuple of variables.
-A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\exists \x \: \varphi(\x,\Y,\y)}.$\\
-
-The {\em decision version} of a function $f$ is defined by the language $L_f = \{\A \mid f(\A) > 0\}$.
-\begin{theo}
-The decision version of a function in $\E1$ is in \textsc{P}.
-\end{theo}
-\begin{proof}
-Let $f$ be a function in $\E1$. Then there is a formula $\varphi(\x,\X,\z)$ such that
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $\z$ is an $m$-tuple of variables, $\e$ is an $m$-tuple of elements and $\x$ is a $k$-tuple of variables.
-Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible evaluations for $\z$.
-Let $\ell = \vert A \vert^k$ and $\d_1,\dots,d_\ell\in A^k$ be all possible evaluations for $\x$.
-Let $\varphi_\A(\X)$ be defined as follows:
-\begin{eqnarray*}
-\varphi_\A(\X) = \varphi(\d_1,\X,\e_1) \vee \cdots \vee \varphi(\d_1,\X,\e_n) \vee \varphi(\d_2,\X,\e_1) \vee \cdots \vee \varphi(\d_n,\X,\e_n).
-\end{eqnarray*}
-Note that $\varphi_\A(\X)$ will have at least one assignment for $\X$ iff $f(\A)>0$.
-Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction.
-Note that $\psi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$.
-Then, let $f'$ be defined as follows:
-\begin{eqnarray*}
-f'(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \psi_\A(\P) \} \mid.
-\end{eqnarray*}
-Note that $f' \in\#\Sigma_0$ \footnote[1]{(paper)}.
-Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
-However, it is showed in $^{[1]}$ that any counting problem in $\#\Sigma_0$ is computable in polynomial time, therefore, the decision version of $f$ is in P.
-\end{proof}
-
-\begin{theo}
-$\#\Sigma_1$ is closed under substraction $\Rightarrow$ P = NP.
-\end{theo}
-\begin{proof}
-\#3DNF $\in \#\Sigma_1$.
-Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment in a \#3DNF instance.
-Suppose that $F_{2^n}-F_{\#3DNF} \in \#\Sigma_1$.
-This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete.
-However, as we showed previously (Theorem 1), it is also in P.
-Then, co-NP $\subseteq$ P.
-\end{proof}
-
-\begin{theo}
-$\#\Sigma_1 \subsetneq \E1$
-\end{theo}
-\begin{proof}
-We will show that the $\E1$ function defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
-\end{proof}
-
-For a given function $f$, we define $f \dotminus 1$ as follows:
-\begin{eqnarray*}
-f \dotminus 1(\A) =
-\begin{cases}
-f(\A)-1, & \text{if }f(\A) > 0 \\
-0, & \text{if }f(\A) = 0.
-\end{cases}
-\end{eqnarray*}
-for every $\L$-structure $\A \in \Truc$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
-
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1
-
-\begin{theo}
-$\E1$ is closed under substraction by one.
-\end{theo}
-\begin{proof}
-Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$.
-Without loss of generality, let $\z = (z_1,\dots,z_d)$ and let $\X = (X_1,\dots,X_r)$.
-There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$.
-This separates the proof in three cases:
-\begin{enumerate}
-
-% CASE 1
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\e \in A^d$.
-Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily.
-First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
-\begin{eqnarray*}
-\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
-\end{eqnarray*}
-This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
-\begin{eqnarray*}
-\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
-\end{eqnarray*}
-If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
-Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$.
-Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E1.$
-
-%CASE 2
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
-\begin{eqnarray}
-\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
-\end{eqnarray}
-where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
-Note that $\theta(\x)$ also mentions all variables in $\x$.
-We also assume that $(\x_1,\dots,\x_n,\y) = \x$.
-As an example, the following formula is of this form:
-\begin{multline*}
-\varphi(\x,\X) =  X_1(x_1,x_2) \wedge X_3(x_3) \wedge X_2(x_4,x_5) \wedge X_3(x_6) \wedge \neg X_1(x_7,x_8) \wedge \\ (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 ) \wedge \\ \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big),
-\end{multline*}
-where $\x = (x_1,x_2,x_3,x_4,x_5,x_6,x_7,x_8)$ and $\X = (X_1,X_2,X_3)$.
-Here, $n = 4$, $\lambda(1) = 1$, $\lambda(2) = \lambda(4) = 3$ and $\lambda(3) = 2$, $\x_1 = (x_1,x_2)$, $\x_2 = (x_3)$, $\x_3 = (x_4,x_5)$, $\x_4 = (x_6)$ and $\y = (x_7,x_8)$.
-Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_7,x_8)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 )$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2,x_5,x_8\},\{x_6,x_7\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big)$.
-
-Similarly to the previous proof, we would like to eliminate the lexicographically smallest tuple of predicates that satisfies the formula \eqref{f1}.
-Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
-We now define
-\begin{multline*}
-\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to \varphi_{m,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
-\end{multline*}
-where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
-Note that $\alpha_{\min}$ is satisfied only by the lexicographically \linebreak smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
-Our new formula is
-\begin{multline}
-\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge \\ \exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg] \label{f2}.
-\end{multline}
-We now show a result by which the main proof will follow.
-\begin{lemma}  \label{first}
-$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
-\end{lemma}
-\begin{proof}
-Let $\A \in \Truc$.
-Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{\lambda(j)=i}\{\d_j\}$.
-We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
-\begin{enumerate}
-\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$.
-That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$).
-Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$.
-Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$.
-We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \varphi_{<,a_i}(\z,\x_i)$ or (3) $\theta(\x)\models \varphi_{<,a_i}(\x_i,\z)$.
-Considering that (2) and (3) are not possible given that both $\z$ and $\x_i$ are assigned the value $\d_i$ and $\A\models\theta(\d)$, we have that $\theta(\x)\models \z = \x_i$.
-But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$, from which we conclude that $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which contradicts the initial assumption that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-\item[(b)] Note that if $\A \models \alpha_{\min}(\c_1,\ldots,\c_n)$, then necessarily $\c_i = \d_i$ for $i\in\{1,\ldots,n\}.$ However, by the construction of $\P$, we see that 
-$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \d_j\Big).$$ Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
-\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$.
-Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $P^\prime_i$.
-Then $\A\models \neg P^\prime_i(\d_i)$, and so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $P_i$ does not have.
-If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-On both cases, we have a contradiction.
-\end{enumerate}
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi(\x,\P)$, we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ it holds that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1,$ which was to be shown.
-\end{proof}
-
 We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
@@
 \end{eqnarray*}
-The function $\lambda_i$ is defined analogously to the first part of the proof, the tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x.$ The formulas $\gamma^{-}_i(\X,\y_i)$, $\delta_i(\x)$, are also defined analogously to the first part of the proof (refer to \eqref{f1}).
+The function $\lambda_i$ is defined analogously to $\lambda$ of the first part of the proof.
+The tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y_i) = \x.$ The formulas $\gamma^{-}_i(\X,\y_i)$, $\delta_i(\x)$ are defined analogously to $\varphi^{-}(\X,\y)$ and $\beta(\x)\,$\footnote[2]{Note that $\delta_i(\x)$ can include subformulas of the form $(\u = \u)$.} respectively (see \eqref{f1}).
 Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables.
@@
 \end{eqnarray*}
-Note that $m = \ell \cdot g(d).$ Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\Truc.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
-We use a function $\rho:\{1,\ldots,k\}\to\{1,\ldots,m\}$ such that $\varphi_i(\x,\X) = \xi_{\rho(i)}(\x,\X)$ is satisfiable, for every $i\in\{1,\ldots,k\}.$
+Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\Truc.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
+We use an injective\footnote[3]{We need this function to be injective because this way we can assure each one of the $k$ satisfiable disjuncts to be represented} function $\rho:\{1,\ldots,k\}\to\{1,\ldots,m\}$ such that $\varphi_i(\x,\X) = \xi_{\rho(i)}(\x,\X)$ is satisfiable, for every $i\in\{1,\ldots,k\}.$
 
-\begin{claim}
-For every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, then $f_{\varphi_i(\s,\X)}(\A) > 0.$
-\end{claim}
-\begin{proof}
-By contradiction, assume that there is an assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, and $f_{\varphi_i(\s,\X)}(\A) = 0.$ That is, there is no assignment $\R$ to $\X$ such that $\A\models\varphi_i(\s,\R).$ Since $\A\models\alpha_i(\s)$, it holds that $\A\models\theta_i(\s)$.
-Therefore, as we showed in part a) of Lemma \ref{first}, there is no assignment $\s^\prime$ to $\x$ such that $f_{\varphi_i(\s,\X)}(\A) > 0.$ This means $\varphi_i(\x,\X)$ is unsatisfiable, but, as we mentioned previously, all of such disjuncts have been eliminated, which is a contradiction.
-\end{proof}
-
-Now we can define the formula like we did on the first part of the proof. For every $i\in\{1,\ldots,k\}$,
+We can conclude that
+\begin{eqnarray*}
+\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X),
+\end{eqnarray*}
+and for every $i\in\{1,\ldots,k\}$,
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\y_i$ is a $p_i$-tuple of variables in $\x$, $\varphi_i^{-}(\X,\y_i)$ is a conjunction of negated predicates in $\X$, $\theta_i(\x)$ defines a total order on a partition of $\x$, and $\beta_i(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
-Also, note that
-\begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X).
-\end{eqnarray*}
+where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\y_i$ is a $p_i$-tuple of variables in $\x$, $\varphi_i^{-}(\X,\y_i)$ is a conjunction of negated predicates in $\X$, $\theta_i(\x)$ defines a total order on a partition of $\x$, and $\beta_i(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
+Let $\alpha_i(\x) = \theta_i(\x)\wedge\beta_i(\x).$
 
-Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\exists\x\:\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\P$ such that $\A\models\exists\x\:\varphi_2(\x,\P)$, and so on.
-As we already know how to exclude that assignment in the first disjunct, we now move to the next disjuncts.
-Similarly to the first part of the proof, let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
+%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
+%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
+
+\begin{claim} \label{alpha}
+Let $\A\in\Truc$ and $i\in\{1,\ldots,k\}$. For every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, it holds that $f_{\varphi_i(\s,\X)}(\A) > 0.$
+\end{claim}
+\begin{proof}
+Let $\s = (\s_1,\ldots,\s_{n_i},\t)$ such that $\A\models\alpha_i(\s)$, and let $\P = (\P_1,\ldots,\P_r)$ where $P_i = \bigcup_{j\in[1,n_i]:\:\lambda(j)=i}\{s_i\}$.
+We will show that $\A\models\varphi_i(\s,\P)$.
+By contradiction, suppose that $\A\not\models\exists\x\,\varphi_i(\x,\P)$.
+That is, there is no assignment $\e$ to $\x$ such that $\A\models\varphi(\e,\P).$ Following the same proof as in case (a) of Lemma \ref{first}, we conclude that $\varphi_i(\x,\P)$ is inconsistent.
+However, as we mentioned previously, all of such disjuncts have been eliminated, which leads to a contradiction.
+\end{proof}
+
+Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\exists\x\:\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\Q$ such that $\A\models\exists\x\:\varphi_2(\x,\Q)$, and so on.
+As we already know how to exclude that assignment in the first disjunct, we will now deal with the next disjuncts.
+Similarly to the first part of the proof, for each $i\in\{1,\ldots,k\}$ let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
 \begin{multline*}
@@
 \end{multline*}
-where $\u_j,\v_j,$ for $\{1,\ldots,n_i\}$ and $\w$ have the adequate number of variables. Also, let
+where $\u_j$ and $\v_j,$ have $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, and $\w$ has $p_i$ variables. Also, let
 \begin{multline*}
@@
 \end{multline*}
-Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
+Note that $\psi_i(\X)$ excludes only the lexicographically smallest tuple of predicates $\P$ such that $\A\models\exists\x\:\varphi_i(\P,\x),$ if there is at least one.
+In other words, every assignment $\P^\prime \neq \P$ is such that $\A\models\psi_i(\P^\prime).$ Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
 \begin{multline}
@@
 \end{multline}
-Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$.
-Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{first} for this sub-formula.
-As we did in Lemma \ref{first} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
-\begin{enumerate}
-\item[(a)] We know that $\A\models\exists\x\:\varphi_q(\x,\P),$ so it immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
-\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
-(1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
-(2) We can use the proof in Lemma \ref{first} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$.
-However, $\A\not\models\exists\x\:\hat{\psi}_q(\P)$ so $\A\not\models\exists\x\:\varphi^\prime_i(\P)$.
-And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
-\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
-As we deduce from the part (c) of Lemma \ref{first}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
-Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ Let $q^\prime$ be such that $\A\models\exists\x\:\varphi_{q^\prime}(\x,\P^\prime).$ By the choice of $q,$ then $q^\prime \geq q,$ but since $\P \neq \P^\prime,$ then $q^\prime > q.$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x),$ so $\A\models\psi_i(\P^\prime)$ for each $i$.
-But also, $\A\models\exists\x\:\alpha_q(\x),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which is a contradiction.
-\end{enumerate}
-
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
-
-% CASE 3
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$.
-We define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2:
-\begin{eqnarray*}
-\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\v_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
-\end{eqnarray*}
-where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_i + c_i = a_i$ for $i\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, $v_i$ has $q_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\v) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
-The formulas $\varphi^{-}_i(\X,\y_i,\v_i)$, $\beta_i(\x,\z)$, are also defined analogously.
-
-In this case, we mix both previous strategies, that is, we are going to {\em isolate} the lexicographically smallest predicate that satisfies the first satisfiable disjunct, and exclude the lexicographically smallest tuple that satisfies the evaluated disjunct.
-
-Let
-\begin{multline*}
-\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\u\exists\w(\alpha_i(\u,\w) \wedge \varphi_{<,p}(\w,\z)) \\.
-\end{multline*}
-
-\end{enumerate}
-\end{proof}
-
-\end{document}

### 2015-04-08 tex/case2generalproof.tex
8 abril

Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Assume that $\A\in\Truc$. Suppose first that there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least one assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\d = (\d_1,\dots,\d_{n_q},\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha_q(\d)$, where $\d_i$ is the corresponding assignment to $\x_i$, for every $i\in\{1,\dots,n_q\}$, and $\ep$ is an assignment for $\y$. Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j:\lambda_q(j)=i,j\in[1,n_q]}\{\d_j\}$. As we did in Lemma \ref{first} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies both (a) and (b)
\begin{enumerate}
\item[(a)] As we showed in part (a) of Lemma \ref{first}, if there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_q(\x,\R)$, then $\A\models\exists\x\:\varphi_q(\x,\P)$ for this particular $\P$. However, as we showed in Claim \ref{alpha}, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_q(\s)$, then there {\em is} such an assignment to $\X$. The assignment $\d$ to $\x$ satisfies that $\A\models\alpha_q(\d)$, so it holds that $\A\models\exists\x\:\varphi_q(\x,\P)$. It immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$

\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$. (1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$. (2) We can use the proof in Lemma \ref{first} to see that $\A\not\models\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula 
$$ \zeta_q(\X) =  (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{q-1}(\v)\vee\psi_q(\X)) $$ 
appears as a conjunct in $\varphi^\prime_i(\x,\X)$. However, by the choice of $q$, there is no $i\in\{1,\ldots,q-1\}$ such that $\A\models\exists\v\:\varphi_i(\v,\P)$, and also $\A\not\models\psi_q(\P)$. It follows that $\A\not\models\zeta_q(\P)$,  so $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$. And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$

\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b). As we deduce from the part (c) of Lemma \ref{first}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$. Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x)$, so $\A\models\psi_i(\P^\prime)$ for each $i$. But as we mentioned, also $\A\models\:\psi_q(\P^\prime),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which leads to a contradiction.
\end{enumerate}

With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0$, we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$

Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$. Let $\P$ be an arbitrary assignment to $\X$. Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\chi(\x,\P))$ for any formula $\chi(\x,\P)$. It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$. And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Hence, from the results in the previous paragraph, if $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}$, we have that $f^\prime = f\dotminus 1\in\E1$

### 2015-04-08 tex/case2part1.tex
8 abril

--- 2015-01-26 sharp-sigma-one-plus.tex
@@
- \documentclass[12pt]{article}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsthm} 
-\usepackage{fullpage}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{bm}
-
-\def\dotminus{\mathbin{\ooalign{\hss\raise1ex\hbox{.}\hss\cr
-  \mathsurround=0pt$-$}}}
-
-\def\E1{\#\Sigma_1^{+}}
-
-\def\Truc{\textsc{Struct}[\L]}
-
-\def\A{{\frak A}}
-\def\C{{\cal C}}
-\def\F{{\cal F}}
-\def\L{{\cal L}}
-\def\N{\mathbb{N}}
-\def\P{\vec{P}}
-\def\R{\vec{R}}
-\def\S{\vec{S}}
-\def\X{\vec{X}}
-\def\Y{\vec{Y}}
-\def\Z{\vec{Z}}
-%% a - arity of \X / arity of assignments \P to \X
-%% b - arity of predicates in \S
-%% c - arity of auxiliar predicates/variables
-\def\c{\vec{c}} %% super auxiliar elements
-\def\d{\vec{d}} %% counted elements
-\def\e{\vec{e}} %% counted elements
-%% f - counting function
-%% g - other functions
-%% h - other functions
-%% i - index
-%% j - index
-%% k - emergency index / size of tuple
-%% l - emergency index / size of tuple
-\def\l{\vec{\ell}}
-%% m - size of variable tuple
-%% n - size of predicate tuple
-%% o - not used
-%% p - 
-%% q - 
-%% r - size of \X / \P
-\def\s{\vec{s}}
-%% t - size of \S
-\def\t{\vec{t}}
-\def\u{\vec{u}} %% auxiliary variables
-\def\v{\vec{v}} %% auxiliary variables
-\def\w{\vec{w}} %% auxiliary variables
-\def\x{\vec{x}} %% quantified variables
-\def\y{\vec{y}} %% auxiliary variables
-\def\z{\vec{z}} %% open variables
-\def\ep{\vec{o}}
-\def\ga{\vec{p}}
-
-\newtheorem{theo}{Theorem}
-\newtheorem{lemma}[theo]{Lemma}
-\newtheorem{claim}[theo]{Claim}
-\newtheorem{coro}[theo]{Corollary}
-
-\begin{document}
-
-\begin{center}
-{ \LARGE \bf
-  Some properties of $\E1$
-}
-\end{center}
-
-We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
-\begin{eqnarray*}
-\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
-&& \leq \text{ is interpreted as a total order for } A \}.
-\end{eqnarray*}
-We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$.
-A quantifier-free $\L$-formula is defined by the following grammar:
-\begin{eqnarray*}
-\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
-&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
-&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
-\end{eqnarray*}
-where $x,y$ and $x_i$ are first order variables for every $i$. We now define an extended quantifier-free $\L$-formula as follows:
-\begin{eqnarray*}
-\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
-&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
-&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
-\end{eqnarray*}
-Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables.
-For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
-\begin{eqnarray*}
-f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, and $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.
-
-Let $\Y = (Y_1,\dots,Y_r)$, and let $\y$ be a tuple of variables.
-A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\exists \x \: \varphi(\x,\Y,\y)}.$\\
-
-The {\em decision version} of a function $f$ is defined by the language $L_f = \{\A \mid f(\A) > 0\}$.
-\begin{theo}
-The decision version of a function in $\E1$ is in \textsc{P}.
-\end{theo}
-\begin{proof}
-Let $f$ be a function in $\E1$. Then there is a formula $\varphi(\x,\X,\z)$ such that
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $\z$ is an $m$-tuple of variables, $\e$ is an $m$-tuple of elements and $\x$ is a $k$-tuple of variables.
-Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible evaluations for $\z$.
-Let $\ell = \vert A \vert^k$ and $\d_1,\dots,d_\ell\in A^k$ be all possible evaluations for $\x$.
-Let $\varphi_\A(\X)$ be defined as follows:
-\begin{eqnarray*}
-\varphi_\A(\X) = \varphi(\d_1,\X,\e_1) \vee \cdots \vee \varphi(\d_1,\X,\e_n) \vee \varphi(\d_2,\X,\e_1) \vee \cdots \vee \varphi(\d_n,\X,\e_n).
-\end{eqnarray*}
-Note that $\varphi_\A(\X)$ will have at least one assignment for $\X$ iff $f(\A)>0$.
-Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction.
-Note that $\psi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$.
-Then, let $f'$ be defined as follows:
-\begin{eqnarray*}
-f'(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \psi_\A(\P) \} \mid.
-\end{eqnarray*}
-Note that $f' \in\#\Sigma_0$ \footnote[1]{(paper)}.
-Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
-However, it is showed in $^{[1]}$ that any counting problem in $\#\Sigma_0$ is computable in polynomial time, therefore, the decision version of $f$ is in P.
-\end{proof}
-
-\begin{theo}
-$\#\Sigma_1$ is closed under substraction $\Rightarrow$ P = NP.
-\end{theo}
-\begin{proof}
-\#3DNF $\in \#\Sigma_1$.
-Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment in a \#3DNF instance.
-Suppose that $F_{2^n}-F_{\#3DNF} \in \#\Sigma_1$.
-This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete.
-However, as we showed previously (Theorem 1), it is also in P.
-Then, co-NP $\subseteq$ P.
-\end{proof}
-
-\begin{theo}
-$\#\Sigma_1 \subsetneq \E1$
-\end{theo}
-\begin{proof}
-We will show that the $\E1$ function defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
-\end{proof}
-
-For a given function $f$, we define $f \dotminus 1$ as follows:
-\begin{eqnarray*}
-f \dotminus 1(\A) =
-\begin{cases}
-f(\A)-1, & \text{if }f(\A) > 0 \\
-0, & \text{if }f(\A) = 0.
-\end{cases}
-\end{eqnarray*}
-for every $\L$-structure $\A \in \Truc$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
-
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1
-
-\begin{theo}
-$\E1$ is closed under substraction by one.
-\end{theo}
-\begin{proof}
-Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$.
-Without loss of generality, let $\z = (z_1,\dots,z_d)$ and let $\X = (X_1,\dots,X_r)$.
-There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$.
-This separates the proof in three cases:
-\begin{enumerate}
-
-% CASE 1
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\e \in A^d$.
-Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily.
-First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
-\begin{eqnarray*}
-\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
-\end{eqnarray*}
-This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
-\begin{eqnarray*}
-\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
-\end{eqnarray*}
-If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
-Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$.
-Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E1.$
-
-%CASE 2
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
+Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
 \begin{eqnarray*}
@@
 
-Similarly to the previous proof, we would like to eliminate the lexicographically smallest tuple of predicates that satisfies the formula \eqref{f1}.
+Similarly to the previous proof, we would like to eliminate the {\em lexicographically smallest}\footnote{We consider the lexicographically smallest tuple of predicates as the one in which its predicates contain the lexicographically smallest tuples and do not contain any more tuples than those} tuple of predicates that satisfies the formula \eqref{f1}.
 Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
@@
 We now show a result by which the main proof will follow.
-\begin{lemma}  \label{first}
-$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
-\end{lemma}
-\begin{proof}
-Let $\A \in \Truc$.
-Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{\lambda(j)=i}\{\d_j\}$.
-We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
-\begin{enumerate}
-\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$.
-That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$).
-Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$.
-Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$.
-We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \varphi_{<,a_i}(\z,\x_i)$ or (3) $\theta(\x)\models \varphi_{<,a_i}(\x_i,\z)$.
-Considering that (2) and (3) are not possible given that both $\z$ and $\x_i$ are assigned the value $\d_i$ and $\A\models\theta(\d)$, we have that $\theta(\x)\models \z = \x_i$.
-But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$, from which we conclude that $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which contradicts the initial assumption that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-\item[(b)] Note that if $\A \models \alpha_{\min}(\c_1,\ldots,\c_n)$, then necessarily $\c_i = \d_i$ for $i\in\{1,\ldots,n\}.$ However, by the construction of $\P$, we see that 
-$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \d_j\Big).$$ Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
-\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$.
-Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $P^\prime_i$.
-Then $\A\models \neg P^\prime_i(\d_i)$, and so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $P_i$ does not have.
-If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-On both cases, we have a contradiction.
-\end{enumerate}
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi(\x,\P)$, we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ it holds that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1,$ which was to be shown.
-\end{proof}
-
-We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
-By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\gamma_i(\x,\X)$ with $i\in\{1,\ldots,\ell\}$ such that
-\begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \gamma_1(\x,\X) \vee \gamma_2(\x,\X) \vee \dots  \vee \gamma_{\ell}(\x,\X),
-\end{eqnarray*}
-where, for every $i\in\{1,\ldots,\ell\}$, 
-\begin{eqnarray*}
-\gamma_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \gamma^{-}_i(\X,\y_i)  \wedge \delta_i(\x).
-\end{eqnarray*}
-The function $\lambda_i$ is defined analogously to the first part of the proof, the tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x.$ The formulas $\gamma^{-}_i(\X,\y_i)$, $\delta_i(\x)$, are also defined analogously to the first part of the proof (refer to \eqref{f1}).
-Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables.
-Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$.
-Note that for every formula $\eta(\x)$,
-\begin{eqnarray*}
-\exists\x\:\eta(\x) &\equiv& \exists\x(\eta(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\eta(\x)\wedge\theta^{g(d)}(\x)).
-\end{eqnarray*}
-We define the following formulas $\xi_i(\x,\X)$, for $i\in\{1,\ldots,m\},$ as follows: 
-\begin{eqnarray*}
-\xi_1(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\xi_{g(d)}(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
-\xi_{g(d)+1}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\xi_{2\cdot g(d)}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
-& \vdots & \\
-\xi_{(\ell-1) g(d)+1}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\xi_{\ell \cdot g(d)}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^{g(d)}(\x).
-\end{eqnarray*}
-Note that $m = \ell \cdot g(d).$ Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\Truc.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
-We use a function $\rho:\{1,\ldots,k\}\to\{1,\ldots,m\}$ such that $\varphi_i(\x,\X) = \xi_{\rho(i)}(\x,\X)$ is satisfiable, for every $i\in\{1,\ldots,k\}.$
-
-\begin{claim}
-For every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, then $f_{\varphi_i(\s,\X)}(\A) > 0.$
-\end{claim}
-\begin{proof}
-By contradiction, assume that there is an assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, and $f_{\varphi_i(\s,\X)}(\A) = 0.$ That is, there is no assignment $\R$ to $\X$ such that $\A\models\varphi_i(\s,\R).$ Since $\A\models\alpha_i(\s)$, it holds that $\A\models\theta_i(\s)$.
-Therefore, as we showed in part a) of Lemma \ref{first}, there is no assignment $\s^\prime$ to $\x$ such that $f_{\varphi_i(\s,\X)}(\A) > 0.$ This means $\varphi_i(\x,\X)$ is unsatisfiable, but, as we mentioned previously, all of such disjuncts have been eliminated, which is a contradiction.
-\end{proof}
-
-Now we can define the formula like we did on the first part of the proof. For every $i\in\{1,\ldots,k\}$,
-\begin{eqnarray*}
-\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
-\end{eqnarray*}
-where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\y_i$ is a $p_i$-tuple of variables in $\x$, $\varphi_i^{-}(\X,\y_i)$ is a conjunction of negated predicates in $\X$, $\theta_i(\x)$ defines a total order on a partition of $\x$, and $\beta_i(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
-Also, note that
-\begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X).
-\end{eqnarray*}
-
-Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\exists\x\:\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\P$ such that $\A\models\exists\x\:\varphi_2(\x,\P)$, and so on.
-As we already know how to exclude that assignment in the first disjunct, we now move to the next disjuncts.
-Similarly to the first part of the proof, let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
-\begin{multline*}
-\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_{n_i},\y)\wedge \\ \forall\v_1\cdots\forall\v_{n_i}\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_{n_i},\w)\wedge\bigvee_{j=1}^{n_i}(\u_j\neq\v_j)\big)\to \varphi_{m_i,<}((\u_1,\dots,\u_{n_i}),(\v_1,\dots,\v_{n_i}))\Big)\Big],
-\end{multline*}
-where $\u_j,\v_j,$ for $\{1,\ldots,n_i\}$ and $\w$ have the adequate number of variables. Also, let
-\begin{multline*}
-\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
-\end{multline*}
-Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
-\begin{multline}
-\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge \\ (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
-\end{multline}
-Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$.
-Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{first} for this sub-formula.
-As we did in Lemma \ref{first} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
-\begin{enumerate}
-\item[(a)] We know that $\A\models\exists\x\:\varphi_q(\x,\P),$ so it immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
-\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
-(1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
-(2) We can use the proof in Lemma \ref{first} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$.
-However, $\A\not\models\exists\x\:\hat{\psi}_q(\P)$ so $\A\not\models\exists\x\:\varphi^\prime_i(\P)$.
-And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
-\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
-As we deduce from the part (c) of Lemma \ref{first}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
-Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ Let $q^\prime$ be such that $\A\models\exists\x\:\varphi_{q^\prime}(\x,\P^\prime).$ By the choice of $q,$ then $q^\prime \geq q,$ but since $\P \neq \P^\prime,$ then $q^\prime > q.$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x),$ so $\A\models\psi_i(\P^\prime)$ for each $i$.
-But also, $\A\models\exists\x\:\alpha_q(\x),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which is a contradiction.
-\end{enumerate}
-
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
-
-% CASE 3
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$.
-We define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2:
-\begin{eqnarray*}
-\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\v_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
-\end{eqnarray*}
-where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_i + c_i = a_i$ for $i\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, $v_i$ has $q_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\v) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
-The formulas $\varphi^{-}_i(\X,\y_i,\v_i)$, $\beta_i(\x,\z)$, are also defined analogously.
-
-In this case, we mix both previous strategies, that is, we are going to {\em isolate} the lexicographically smallest predicate that satisfies the first satisfiable disjunct, and exclude the lexicographically smallest tuple that satisfies the evaluated disjunct.
-
-Let
-\begin{multline*}
-\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\u\exists\w(\alpha_i(\u,\w) \wedge \varphi_{<,p}(\w,\z)) \\.
-\end{multline*}
-
-\end{enumerate}
-\end{proof}
-
-\end{document}

### 2015-04-08 tex/case3.tex
8 abril

Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
\begin{eqnarray*}
f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
\end{eqnarray*}
for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$. In order to prove that $f\dotminus 1\in\E1$, we define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2, where
\begin{eqnarray*}
\varphi(\x,\X,\z) \equiv \varphi_1(\x,\X,\z) \vee \cdots \vee \varphi_k(\x,\X,\z),
\end{eqnarray*}
and
\begin{eqnarray*}
\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\w_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
\end{eqnarray*}
where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\ell} + c_{\ell} = a_{\ell}$ for $\ell\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, and $\w_i$ has $q_i$ variables. Furthermore, we have that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\w_i) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2. The formulas $\varphi^{-}_i(\X,\y_i,\w_i)$, $\beta_i(\x,\z)$, are also defined analogously.

In this case, we mix both the strategies in cases 1 and 2. That is, we are going to {\em isolate} the lexicographically smallest tuple of predicates that satisfies the first satisfiable disjunct, and then {\em exclude} the lexicographically smallest tuple that satisfies the isolated disjunct.

Let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and
\begin{multline*}
\alpha^i_{\min}(\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}) = \exists\y\exists\w\Big[ \alpha_i(\x_1,\dots,\x_{n_i},\y,\z_1,\dots,\z_{n_i},\w)\wedge \\ \forall\u_1\cdots\forall\u_{n_i}\forall\s\forall\v_1\cdots\forall\v_{n_i}\forall\t\Big(\big(\alpha_i(\u_1,\dots,\u_{n_i},\s,\v_1,\dots,\v_{n_i},\t)\wedge\bigvee_{j=1}^{n_i}(\x_j\neq\u_j \vee \z_j\neq\v_j)\big)\to \\ \varphi_{m_i,<}((\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}),(\u_1,\dots,\u_{n_i},\v_1,\dots,\v_{n_i}))\Big)\Big],
\end{multline*}
and let
\begin{multline*}
\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\w(\exists\u\:\varphi(\u,\X,\w) \wedge \varphi_{p,<}(\w,\z)).
\end{multline*}
Let $\P$ and $\e$ be assignments to $\X$ and $\z$. Note that $\psi_i(\X,\z)$ excludes $\P$ and $\e$ only if $\P$ is the {\em lexicographically smallest} tuple of predicates (Same as case 2) such that $\A\models\exists\x\:\varphi_i(\x,\P,\d)$, for some assignment $\d$ to $\z$, and $\e$ is the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_i(\x,\P,\e).$ We define $\varphi_i^\prime(\x,\X,\z)$ in the same way as case 2:
\begin{multline*}
\varphi_i^\prime(\x,\X,\z) = \varphi_i(\x,\X,\z) \wedge \psi_1(\X,\z) \wedge (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\psi_2(\X,\z)) \wedge \cdots \wedge \\ (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\cdots\vee\exists\u\exists\v\:\alpha_{i-1}(\u,\v)\vee\psi_i(\X,\z)).
\end{multline*}
Finally, let $\varphi^\prime(\x,\X,\z) = \bigvee_{i = 1}^k \varphi_i^\prime(\x,\X,\z)$.

Let $q$ be the least $i\in\{1,\ldots,k\}$ such that $\A\models\exists\x\:\varphi_i(\x,\R,\d)$ for some assignment $R$ to $\X$ and some assignment $\d$ to $\z$. Let $\P$ be the lexicographically smallest tuple of predicates such that $\A\models\exists\x\:\varphi_q(\x,\P,\d^\prime)$ for some assignment $\d^\prime$ to $\z$. Let $\e$ be the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_q(\x,\P,\e).$ This formula is such that (a) $\A\models\exists\x\:\varphi(\x,\P,\e)$, (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P,\e)$ and $\P$ and $\e$ are the only assignments that satisfy (a) and (b). The proof of this is analogous to case 2. Therefore, we conclude that $f_{\exists\x\,\varphi^\prime(\x,\X,\z)} = f_{\exists\x\,\varphi(\x,\X,\z)}\dotminus 1.$

### 2015-04-08 tex/lemmafirst.tex
8 abril

--- 2015-01-21 sharp-sigma-one-plus.tex
@@
-\documentclass[12pt]{article}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsthm} 
-\usepackage{fullpage}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{bm}
-
-\def\dotminus{\mathbin{\ooalign{\hss\raise1ex\hbox{.}\hss\cr
-  \mathsurround=0pt$-$}}}
-
-\def\E1{\#\Sigma_1^{+}}
-
-\def\Truc{\textsc{Struct}[\L]}
-
-\def\A{{\frak A}}
-\def\C{{\cal C}}
-\def\F{{\cal F}}
-\def\L{{\cal L}}
-\def\N{\mathbb{N}}
-\def\P{\vec{P}}
-\def\R{\vec{R}}
-\def\S{\vec{S}}
-\def\X{\vec{X}}
-\def\Y{\vec{Y}}
-\def\Z{\vec{Z}}
-%% a - arity of \X / arity of assignments \P to \X
-%% b - arity of predicates in \S
-%% c - arity of auxiliar predicates/variables
-\def\d{\vec{d}} %% counted elements
-\def\e{\vec{e}} %% counted elements
-%% f - counting function
-%% g - other functions
-%% h - other functions
-%% i - index
-%% j - index
-%% k - emergency index / size of tuple
-%% l - emergency index / size of tuple
-\def\l{\vec{\ell}}
-%% m - size of variable tuple
-%% n - size of predicate tuple
-%% o - not used
-%% p - 
-%% q - 
-%% r - size of \X / \P
-\def\s{\vec{s}}
-%% t - size of \S
-\def\u{\vec{u}} %% auxiliary variables
-\def\v{\vec{v}} %% auxiliary variables
-\def\w{\vec{w}} %% auxiliary variables
-\def\x{\vec{x}} %% quantified variables
-\def\y{\vec{y}} %% auxiliary variables
-\def\z{\vec{z}} %% open variables
-\def\ep{\vec{o}}
-\def\ga{\vec{p}}
-
-\newtheorem{theo}{Theorem}
-\newtheorem{lemma}[theo]{Lemma}
-\newtheorem{claim}[theo]{Claim}
-\newtheorem{coro}[theo]{Corollary}
-
-\begin{document}
-
-\begin{center}
-{ \LARGE \bf
-  Some properties of $\E1$
-}
-\end{center}
-
-We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
-\begin{eqnarray*}
-\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
-&& \leq \text{ is interpreted as a total order for } A \}.
-\end{eqnarray*}
-We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$.
-A quantifier-free $\L$-formula is defined by the following grammar:
-\begin{eqnarray*}
-\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
-&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
-&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
-\end{eqnarray*}
-where $x,y$ and $x_i$ are first order variables for every $i$. We now define an extended quantifier-free $\L$-formula as follows:
-\begin{eqnarray*}
-\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
-&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
-&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
-\end{eqnarray*}
-Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables.
-For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
-\begin{eqnarray*}
-f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $a_1,\ldots,a_q$, and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.
-
-Let $\X = (X_1,\dots,X_r)$, and let $\z$ be a tuple of variables.
-A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\X,\z)$ such that $f = f_{\exists \x \: \varphi(\x,\X,\z)}.$\\
-
-The {\em decision version} of a function $f$ is defined by the language $L_f = \{\A \mid f(\A) > 0\}$.
-\begin{theo}
-The decision version of a function in $\E1$ is in \textsc{P}.
-\end{theo}
-\begin{proof}
-Let $f$ be a function in $\E1$. Then there is a formula $\varphi(\x,\X,\z)$ such that
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $\z$ is an $m$-tuple of variables, $\e$ is an $m$-tuple of elements and $\x$ is a $k$-tuple of variables.
-Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible evaluations for $\z$.
-Let $\ell = \vert A \vert^k$ and $\d_1,\dots,d_\ell\in A^k$ be all possible evaluations for $\x$.
-Let $\varphi_\A(\X)$ be defined as follows:
-\begin{eqnarray*}
-\varphi_\A(\X) = \varphi(\d_1,\X,\e_1) \vee \cdots \vee \varphi(\d_1,\X,\e_n) \vee \varphi(\d_2,\X,\e_1) \vee \cdots \vee \varphi(\d_n,\X,\e_n).
-\end{eqnarray*}
-Note that $\varphi_\A(\X)$ will have at least one assignment for $\X$ iff $f(\A)>0$.
-Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction.
-Note that $\psi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$.
-Then, let $f'$ be defined as follows:
-\begin{eqnarray*}
-f'(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \psi_\A(\P) \} \mid.
-\end{eqnarray*}
-Note that $f' \in\#\Sigma_0$ \footnote[1]{(paper)}.
-Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
-However, it is showed in $^{[1]}$ that any counting problem in $\#\Sigma_0$ is computable in polynomial time, therefore, the decision version of $f$ is in P.
-\end{proof}
-
-\begin{theo}
-$\#\Sigma_1$ is closed under substraction $\Rightarrow$ P = NP.
-\end{theo}
-\begin{proof}
-\#3DNF $\in \#\Sigma_1$.
-Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment in a \#3DNF instance.
-Suppose that $F_{2^n}-F_{\#3DNF} \in \#\Sigma_1$.
-This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete.
-However, as we showed previously (Theorem 1), it is also in P.
-Then, co-NP $\subseteq$ P.
-\end{proof}
-
-\begin{theo}
-$\#\Sigma_1 \subsetneq \E1$
-\end{theo}
-\begin{proof}
-We will show that the $\E1$ function defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
-\end{proof}
-
-For a given function $f$, we define $f \dotminus 1$ as follows:
-\begin{eqnarray*}
-f \dotminus 1(\A) =
-\begin{cases}
-f(\A)-1, & \text{if }f(\A) > 0 \\
-0, & \text{if }f(\A) = 0.
-\end{cases}
-\end{eqnarray*}
-for every structure $\A$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
-
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1
-
-\begin{theo}
-$\E1$ is closed under substraction by one.
-\end{theo}
-\begin{proof}
-For every $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$, let $\z = (z_1,\dots,z_d)$ and let $\X = (X_1,\dots,X_r)$.
-There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$.
-This separates the proof in three cases:
-\begin{enumerate}
-
-% CASE 1
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\e \in A^d$.
-Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily.
-First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
-\begin{eqnarray*}
-\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
-\end{eqnarray*}
-This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
-\begin{eqnarray*}
-\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
-\end{eqnarray*}
-If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
-Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$.
-Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E1.$
-
-%CASE 2
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
-\begin{eqnarray}
-\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
-\end{eqnarray}
-where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$.
-We also assume that $(\x_1,\dots,\x_n,\y) = \x$.
-As an example, the following formula is of this form:
-\begin{align*}
-\varphi(\x,\X) =  X_1(x_1,x_2) \wedge \neg X_1(x_3,x_4) \wedge (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 ) \wedge \forall z\big( S_1(x_1,z) \big),
-\end{align*}
-where $\x = (x_1,x_2,x_3,x_4)$ and $\X = (X_1)$.
-Here, $n = 1$, $\lambda(1) = \lambda(2) = 1$, $\x_1 = (x_1,x_2)$ and $\y = (x_3,x_4)$.
-Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_3,x_4)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4)$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z) \big)$.
-
-Similarly to the previous proof, we would like to eliminate the lexicographically smallest tuple of predicates that satisfies the formula \eqref{f1}.
-Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
-We now define
-\begin{multline*}
-\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to \varphi_{m,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
-\end{multline*}
-where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
-Note that $\alpha_{\min}$ is satisfied only by the lexicographically \linebreak smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
-Our new formula is
-\begin{multline}
-\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge\exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \\ \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg] \label{f2}.
-\end{multline}
-We now show a result by which the main proof will follow.
-\begin{lemma}  \label{lemmaone}
+\begin{lemma} \label{first}
 $f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
@@
 Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{\lambda(j)=i}\{\d_j\}$.
+Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j\in[1,n]:\:\lambda(j)=i}\{\d_j\}$.
 We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
@@
 Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$.
-We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \z < \x_i$ or (3) $\theta(\x)\models \z > \x_i$.
-Considering that (2) and (3) are not possible since $\A\models\theta(\d)$, and both $\z$ and $\x_i$ are assigned the value $\d_i$, then $\theta(\x)\models \z = \x_i$.
-But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$ from which we conclude that $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which leads to a contradiction.
-\item[(b)] By the construction of $\P$, we see that 
-$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big).$$ However, the only possible assignment to $\alpha_{\min}(\x_1,\dots,\x_n)$ is $(\d_1,\dots,\d_n)$.
-Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
+We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \varphi_{<,a_i}(\z,\x_i)$ or (3) $\theta(\x)\models \varphi_{<,a_i}(\x_i,\z)$.
+Considering that (2) and (3) are not possible given that both $\z$ and $\x_i$ are assigned the value $\d_i$ and $\A\models\theta(\d)$, we have that $\theta(\x)\models \z = \x_i$.
+But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$, which means $\varphi(\x,\X)$ is inconsistent.
+That is, there's no structure $\A^\prime\in\Truc$ such that $\A^\prime\models\exists\x\:\varphi(\x,\R^\prime)$ for any assignment $\R^\prime$ to $\X$.
+In particular, $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which contradicts the initial assumption that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
+\item[(b)] Note that if $\A \models \alpha_{\min}(\c_1,\ldots,\c_n)$, then necessarily $\c_i = \d_i$ for $i\in\{1,\ldots,n\}.$ However, by the construction of $\P$, we see that 
+$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \d_j\Big).$$ Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
 \item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
 We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$.
-Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $\P^\prime$, then $\A\models \neg P^\prime_i(\d_i)$.
-And so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $\P_i$ does not have.
+Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $P^\prime_i$.
+Then $\A\models \neg P^\prime_i(\d_i)$, and so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
+Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $P_i$ does not have.
 If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
@@
 It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
+And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ it holds that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1,$ which was to be shown.
 \end{proof}
-
-We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
-By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\hat{\varphi}_i(\x,\X)$ with $i\in\{1,\ldots,\hat{k}\}$ such that
-\begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \hat{\varphi}_1(\x,\X) \vee \hat{\varphi}_2(\x,\X) \vee \dots  \vee \hat{\varphi}_{\hat{k}}(\x,\X),
-\end{eqnarray*}
-where, for every $i\in\{1,\ldots,\hat{k}\}$, 
-\begin{eqnarray*}
-\hat{\varphi}_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \hat{\varphi}^{-}_i(\X,\y_i)  \wedge \beta_i(\x).
-\end{eqnarray*}
-The function $\lambda_i$ is defined analogously to the first part of the proof, the tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x.$ The formulas $\hat{\varphi}^{-}_i(\X,\y_i)$, $\beta_i(\x)$, are also defined analogously.
-Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables.
-Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$.
-Note that for every formula $\psi(\x)$,
-\begin{eqnarray*}
-\exists\x\:\psi(\x) &\equiv& \exists\x(\psi(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\psi(\x)\wedge\theta^{g(d)}(\x)).
-\end{eqnarray*}
-We define the following formulas $\check{\varphi}_i(\x,\X)$, for $i\in\{1,\ldots,\check{k}\},$ as follows: 
-\begin{eqnarray*}
-\check{\varphi}_1(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{g(d)}(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
-\check{\varphi}_{g(d)+1}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{2\cdot g(d)}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
-& \vdots & \\
-\check{\varphi}_{(\hat{k}-1) g(d)+1}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{\hat{k} \cdot g(d)}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^{g(d)}(\x).
-\end{eqnarray*}
-Note that $\check{k} = \hat{k} \cdot g(d).$ Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L-$structure $\A$, that is, every $i\in\{1,\ldots,\check{k}\}$ such that for every assignment $\s$ to $\x,$ $f_{\check{\varphi}_i(\s,\X)} = 0.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
-We order the remaining disjuncts and define the disjuncts $\varphi_i(\x,\X)$ such that for every $j\in\{1,\ldots,\check{k}\}$ where $\check{\varphi}_j(\x,\X)$ is satisfiable, there is an $i\in\{1,\ldots,k\}$ such that $\varphi_i(\x,\X) = \check{\varphi}_j(\x,\X).$
-
-As seen in lemma \ref{lemmaone}, for every $\L-$structure $\A$ and $i\in\{1,\ldots,k\}$, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$ but $f_{\varphi_i(\s,\X)}(\A) = 0,$ then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) = 0$, so it follows that for every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) > 0.$
-
-Now we can define the formula like we did on the first part of the proof. For every $i\in\{1,\ldots,k\}$,
-\begin{eqnarray*}
-\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
-\end{eqnarray*}
- Also, note that
-\begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X).
-\end{eqnarray*}
-
-Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\P$ such that $\A\models\varphi_2(\x,\P)$, and so on.
-As we already know how to exclude that assignment in the first disjunct, we now move to the next disjuncts.
-Similarly to the first part of the proof, let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
-\begin{multline*}
-\alpha^i_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_n,\w)\wedge\bigvee_{j=1}^n(\u_j\neq\v_j)\big)\to \varphi_{m_i,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
-\end{multline*}
-where $\u_j,\v_j,$ for $\{1,\ldots,n_i\}$ and $\w$ have the adequate number of variables. Also, let
-\begin{multline*}
-\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_n\bigg[\alpha^i_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
-\end{multline*}
-Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
-\begin{multline}
-\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge \\ (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
-\end{multline}
-Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$.
-Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone} for this sub-formula.
-As we did in Lemma \ref{lemmaone} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
-\begin{enumerate}
-\item[(a)] We know that $\A\models\exists\x\:\varphi_q(\x,\P),$ so it immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
-\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
-(1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
-(2) We can use the proof in Lemma \ref{lemmaone} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$.
-However, $\A\not\models\exists\x\:\hat{\psi}_q(\P)$ so $\A\not\models\exists\x\:\varphi^\prime_i(\P)$.
-And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
-\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
-As we deduce from the part (c) of Lemma \ref{lemmaone}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
-Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ Let $q^\prime$ be such that $\A\models\exists\x\:\varphi_{q^\prime}(\x,\P^\prime).$ By the choice of $q,$ then $q^\prime \geq q,$ but since $\P \neq \P^\prime,$ then $q^\prime > q.$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x),$ so $\A\models\psi_i(\P^\prime)$ for each $i$.
-But also, $\A\models\exists\x\:\alpha_q(\x),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which is a contradiction.
-\end{enumerate}
-
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
-
-% CASE 3
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_d)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^d$.
-We define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2:
-\begin{eqnarray*}
-\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\v_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
-\end{eqnarray*}
-where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_i + c_i = a_i$ for $i\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, $v_i$ has $q_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\v) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
-The formulas $\varphi^{-}_i(\X,\y_i,\v_i)$, $\beta_i(\x,\z)$, are also defined analogously.
-
-\end{enumerate}
-\end{proof}
-
-\end{document}

### 2015-04-08 tex/sharp-sigma-one-plus.tex
8 abril

--- 2015-01-21 sharp-sigma-one-plus.tex
@@
 \def\A{{\frak A}}
+\def\B{{\frak B}}
 \def\C{{\cal C}}
@@
 \def\P{\vec{P}}
+\def\Q{\vec{Q}}
 \def\R{\vec{R}}
@@
 \newtheorem{theo}{Theorem}
-\newtheorem{lemma}[theo]{Lemma}
-\newtheorem{claim}[theo]{Claim}
-\newtheorem{coro}[theo]{Corollary}
+\newtheorem{lemma}{Lemma}[theo]
+\newtheorem{claim}{Claim}[theo]
+\newtheorem{coro}{Corollary}[theo]
 
@@
 \end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $a_1,\ldots,a_q$, and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.
+for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, and $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.
 
-Let $\X = (X_1,\dots,X_r)$, and let $\z$ be a tuple of variables.
-A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\X,\z)$ such that $f = f_{\exists \x \: \varphi(\x,\X,\z)}.$\\
+Let $\Y = (Y_1,\dots,Y_r)$, and let $\y$ be a tuple of variables.
+A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\exists \x \: \varphi(\x,\Y,\y)}.$\\
 
 The {\em decision version} of a function $f$ is defined by the language $L_f = \{\A \mid f(\A) > 0\}$.
-\begin{theo}
-The decision version of a function in $\E1$ is in \textsc{P}.
-\end{theo}
-\begin{proof}
-Let $f$ be a function in $\E1$. Then there is a formula $\varphi(\x,\X,\z)$ such that
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $\z$ is an $m$-tuple of variables, $\e$ is an $m$-tuple of elements and $\x$ is a $k$-tuple of variables.
-Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible evaluations for $\z$.
-Let $\ell = \vert A \vert^k$ and $\d_1,\dots,d_\ell\in A^k$ be all possible evaluations for $\x$.
-Let $\varphi_\A(\X)$ be defined as follows:
-\begin{eqnarray*}
-\varphi_\A(\X) = \varphi(\d_1,\X,\e_1) \vee \cdots \vee \varphi(\d_1,\X,\e_n) \vee \varphi(\d_2,\X,\e_1) \vee \cdots \vee \varphi(\d_n,\X,\e_n).
-\end{eqnarray*}
-Note that $\varphi_\A(\X)$ will have at least one assignment for $\X$ iff $f(\A)>0$.
-Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction.
-Note that $\psi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$.
-Then, let $f'$ be defined as follows:
-\begin{eqnarray*}
-f'(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \psi_\A(\P) \} \mid.
-\end{eqnarray*}
-Note that $f' \in\#\Sigma_0$ \footnote[1]{(paper)}.
-Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
-However, it is showed in $^{[1]}$ that any counting problem in $\#\Sigma_0$ is computable in polynomial time, therefore, the decision version of $f$ is in P.
-\end{proof}
 
-\begin{theo}
-$\#\Sigma_1$ is closed under substraction $\Rightarrow$ P = NP.
-\end{theo}
-\begin{proof}
-\#3DNF $\in \#\Sigma_1$.
-Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment in a \#3DNF instance.
-Suppose that $F_{2^n}-F_{\#3DNF} \in \#\Sigma_1$.
-This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete.
-However, as we showed previously (Theorem 1), it is also in P.
-Then, co-NP $\subseteq$ P.
-\end{proof}
+\input{th1}
 
-\begin{theo}
-$\#\Sigma_1 \subsetneq \E1$
-\end{theo}
-\begin{proof}
-We will show that the $\E1$ function defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
-\end{proof}
+\input{th2}
+
+\input{th3}
 
@@
 \end{eqnarray*}
-for every structure $\A$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
+for every $\L$-structure $\A \in \Truc$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
 
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1
-
-\begin{theo}
-$\E1$ is closed under substraction by one.
-\end{theo}
-\begin{proof}
-For every $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$, let $\z = (z_1,\dots,z_d)$ and let $\X = (X_1,\dots,X_r)$.
-There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$.
-This separates the proof in three cases:
-\begin{enumerate}
-
-% CASE 1
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\e \in A^d$.
-Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily.
-First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
-\begin{eqnarray*}
-\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
-\end{eqnarray*}
-This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
-\begin{eqnarray*}
-\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
-\end{eqnarray*}
-If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
-Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$.
-Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E1.$
-
-%CASE 2
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
-\begin{eqnarray}
-\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
-\end{eqnarray}
-where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$.
-We also assume that $(\x_1,\dots,\x_n,\y) = \x$.
-As an example, the following formula is of this form:
-\begin{align*}
-\varphi(\x,\X) =  X_1(x_1,x_2) \wedge \neg X_1(x_3,x_4) \wedge (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 ) \wedge \forall z\big( S_1(x_1,z) \big),
-\end{align*}
-where $\x = (x_1,x_2,x_3,x_4)$ and $\X = (X_1)$.
-Here, $n = 1$, $\lambda(1) = \lambda(2) = 1$, $\x_1 = (x_1,x_2)$ and $\y = (x_3,x_4)$.
-Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_3,x_4)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4)$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z) \big)$.
-
-Similarly to the previous proof, we would like to eliminate the lexicographically smallest tuple of predicates that satisfies the formula \eqref{f1}.
-Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
-We now define
-\begin{multline*}
-\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to \varphi_{m,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
-\end{multline*}
-where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
-Note that $\alpha_{\min}$ is satisfied only by the lexicographically \linebreak smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
-Our new formula is
-\begin{multline}
-\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge\exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \\ \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg] \label{f2}.
-\end{multline}
-We now show a result by which the main proof will follow.
-\begin{lemma}  \label{lemmaone}
-$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
-\end{lemma}
-\begin{proof}
-Let $\A \in \Truc$.
-Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{\lambda(j)=i}\{\d_j\}$.
-We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
-\begin{enumerate}
-\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$.
-That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$).
-Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$.
-Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$.
-We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \z < \x_i$ or (3) $\theta(\x)\models \z > \x_i$.
-Considering that (2) and (3) are not possible since $\A\models\theta(\d)$, and both $\z$ and $\x_i$ are assigned the value $\d_i$, then $\theta(\x)\models \z = \x_i$.
-But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$ from which we conclude that $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which leads to a contradiction.
-\item[(b)] By the construction of $\P$, we see that 
-$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big).$$ However, the only possible assignment to $\alpha_{\min}(\x_1,\dots,\x_n)$ is $(\d_1,\dots,\d_n)$.
-Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
-\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$.
-Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $\P^\prime$, then $\A\models \neg P^\prime_i(\d_i)$.
-And so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $\P_i$ does not have.
-If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-On both cases, we have a contradiction.
-\end{enumerate}
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi(\x,\P)$, we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
-\end{proof}
-
-We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
-By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\hat{\varphi}_i(\x,\X)$ with $i\in\{1,\ldots,\hat{k}\}$ such that
-\begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \hat{\varphi}_1(\x,\X) \vee \hat{\varphi}_2(\x,\X) \vee \dots  \vee \hat{\varphi}_{\hat{k}}(\x,\X),
-\end{eqnarray*}
-where, for every $i\in\{1,\ldots,\hat{k}\}$, 
-\begin{eqnarray*}
-\hat{\varphi}_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \hat{\varphi}^{-}_i(\X,\y_i)  \wedge \beta_i(\x).
-\end{eqnarray*}
-The function $\lambda_i$ is defined analogously to the first part of the proof, the tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x.$ The formulas $\hat{\varphi}^{-}_i(\X,\y_i)$, $\beta_i(\x)$, are also defined analogously.
-Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables.
-Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$.
-Note that for every formula $\psi(\x)$,
-\begin{eqnarray*}
-\exists\x\:\psi(\x) &\equiv& \exists\x(\psi(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\psi(\x)\wedge\theta^{g(d)}(\x)).
-\end{eqnarray*}
-We define the following formulas $\check{\varphi}_i(\x,\X)$, for $i\in\{1,\ldots,\check{k}\},$ as follows: 
-\begin{eqnarray*}
-\check{\varphi}_1(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{g(d)}(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
-\check{\varphi}_{g(d)+1}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{2\cdot g(d)}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
-& \vdots & \\
-\check{\varphi}_{(\hat{k}-1) g(d)+1}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{\hat{k} \cdot g(d)}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^{g(d)}(\x).
-\end{eqnarray*}
-Note that $\check{k} = \hat{k} \cdot g(d).$ Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L-$structure $\A$, that is, every $i\in\{1,\ldots,\check{k}\}$ such that for every assignment $\s$ to $\x,$ $f_{\check{\varphi}_i(\s,\X)} = 0.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
-We order the remaining disjuncts and define the disjuncts $\varphi_i(\x,\X)$ such that for every $j\in\{1,\ldots,\check{k}\}$ where $\check{\varphi}_j(\x,\X)$ is satisfiable, there is an $i\in\{1,\ldots,k\}$ such that $\varphi_i(\x,\X) = \check{\varphi}_j(\x,\X).$
-
-As seen in lemma \ref{lemmaone}, for every $\L-$structure $\A$ and $i\in\{1,\ldots,k\}$, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$ but $f_{\varphi_i(\s,\X)}(\A) = 0,$ then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) = 0$, so it follows that for every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) > 0.$
-
-Now we can define the formula like we did on the first part of the proof. For every $i\in\{1,\ldots,k\}$,
-\begin{eqnarray*}
-\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
-\end{eqnarray*}
- Also, note that
-\begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X).
-\end{eqnarray*}
-
-Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\P$ such that $\A\models\varphi_2(\x,\P)$, and so on.
-As we already know how to exclude that assignment in the first disjunct, we now move to the next disjuncts.
-Similarly to the first part of the proof, let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
-\begin{multline*}
-\alpha^i_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_n,\w)\wedge\bigvee_{j=1}^n(\u_j\neq\v_j)\big)\to \varphi_{m_i,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
-\end{multline*}
-where $\u_j,\v_j,$ for $\{1,\ldots,n_i\}$ and $\w$ have the adequate number of variables. Also, let
-\begin{multline*}
-\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_n\bigg[\alpha^i_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
-\end{multline*}
-Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
-\begin{multline}
-\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge \\ (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
-\end{multline}
-Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$.
-Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone} for this sub-formula.
-As we did in Lemma \ref{lemmaone} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
-\begin{enumerate}
-\item[(a)] We know that $\A\models\exists\x\:\varphi_q(\x,\P),$ so it immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
-\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
-(1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
-(2) We can use the proof in Lemma \ref{lemmaone} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$.
-However, $\A\not\models\exists\x\:\hat{\psi}_q(\P)$ so $\A\not\models\exists\x\:\varphi^\prime_i(\P)$.
-And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
-\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
-As we deduce from the part (c) of Lemma \ref{lemmaone}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
-Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ Let $q^\prime$ be such that $\A\models\exists\x\:\varphi_{q^\prime}(\x,\P^\prime).$ By the choice of $q,$ then $q^\prime \geq q,$ but since $\P \neq \P^\prime,$ then $q^\prime > q.$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x),$ so $\A\models\psi_i(\P^\prime)$ for each $i$.
-But also, $\A\models\exists\x\:\alpha_q(\x),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which is a contradiction.
-\end{enumerate}
-
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
-
-% CASE 3
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_d)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^d$.
-We define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2:
-\begin{eqnarray*}
-\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\v_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
-\end{eqnarray*}
-where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_i + c_i = a_i$ for $i\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, $v_i$ has $q_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\v) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
-The formulas $\varphi^{-}_i(\X,\y_i,\v_i)$, $\beta_i(\x,\z)$, are also defined analogously.
-
-\end{enumerate}
-\end{proof}
+\input{th4}
 

### 2015-04-08 tex/th1.tex
8 abril

--- 2015-01-21 sharp-sigma-one-plus.tex
@@
-\documentclass[12pt]{article}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsthm} 
-\usepackage{fullpage}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{bm}
-
-\def\dotminus{\mathbin{\ooalign{\hss\raise1ex\hbox{.}\hss\cr
-  \mathsurround=0pt$-$}}}
-
-\def\E1{\#\Sigma_1^{+}}
-
-\def\Truc{\textsc{Struct}[\L]}
-
-\def\A{{\frak A}}
-\def\C{{\cal C}}
-\def\F{{\cal F}}
-\def\L{{\cal L}}
-\def\N{\mathbb{N}}
-\def\P{\vec{P}}
-\def\R{\vec{R}}
-\def\S{\vec{S}}
-\def\X{\vec{X}}
-\def\Y{\vec{Y}}
-\def\Z{\vec{Z}}
-%% a - arity of \X / arity of assignments \P to \X
-%% b - arity of predicates in \S
-%% c - arity of auxiliar predicates/variables
-\def\d{\vec{d}} %% counted elements
-\def\e{\vec{e}} %% counted elements
-%% f - counting function
-%% g - other functions
-%% h - other functions
-%% i - index
-%% j - index
-%% k - emergency index / size of tuple
-%% l - emergency index / size of tuple
-\def\l{\vec{\ell}}
-%% m - size of variable tuple
-%% n - size of predicate tuple
-%% o - not used
-%% p - 
-%% q - 
-%% r - size of \X / \P
-\def\s{\vec{s}}
-%% t - size of \S
-\def\u{\vec{u}} %% auxiliary variables
-\def\v{\vec{v}} %% auxiliary variables
-\def\w{\vec{w}} %% auxiliary variables
-\def\x{\vec{x}} %% quantified variables
-\def\y{\vec{y}} %% auxiliary variables
-\def\z{\vec{z}} %% open variables
-\def\ep{\vec{o}}
-\def\ga{\vec{p}}
-
-\newtheorem{theo}{Theorem}
-\newtheorem{lemma}[theo]{Lemma}
-\newtheorem{claim}[theo]{Claim}
-\newtheorem{coro}[theo]{Corollary}
-
-\begin{document}
-
-\begin{center}
-{ \LARGE \bf
-  Some properties of $\E1$
-}
-\end{center}
-
-We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
-\begin{eqnarray*}
-\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
-&& \leq \text{ is interpreted as a total order for } A \}.
-\end{eqnarray*}
-We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$.
-A quantifier-free $\L$-formula is defined by the following grammar:
-\begin{eqnarray*}
-\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
-&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
-&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
-\end{eqnarray*}
-where $x,y$ and $x_i$ are first order variables for every $i$. We now define an extended quantifier-free $\L$-formula as follows:
-\begin{eqnarray*}
-\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
-&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
-&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
-\end{eqnarray*}
-Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables.
-For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
-\begin{eqnarray*}
-f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $a_1,\ldots,a_q$, and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.
-
-Let $\X = (X_1,\dots,X_r)$, and let $\z$ be a tuple of variables.
-A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\X,\z)$ such that $f = f_{\exists \x \: \varphi(\x,\X,\z)}.$\\
-
-The {\em decision version} of a function $f$ is defined by the language $L_f = \{\A \mid f(\A) > 0\}$.
-\begin{theo}
+\begin{theo} \label{decisionptime}
 The decision version of a function in $\E1$ is in \textsc{P}.
@@
 Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible evaluations for $\z$.
-Let $\ell = \vert A \vert^k$ and $\d_1,\dots,d_\ell\in A^k$ be all possible evaluations for $\x$.
+Let $\ell = \vert A \vert^k$ and $\d_1,\dots,\d_\ell\in A^k$ be all possible evaluations for $\x$.
 Let $\varphi_\A(\X)$ be defined as follows:
@@
 \end{eqnarray*}
-Note that $\varphi_\A(\X)$ will have at least one assignment for $\X$ iff $f(\A)>0$.
+Note that $\varphi_\A(\X)$ will have at least one assignment $\P$ for $\X$ such that $\A\models\varphi_\A(\P)$ iff $f(\A)>0$.
 Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction.
-Note that $\psi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$.
+Note as well that $\psi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$.
 Then, let $f'$ be defined as follows:
@@
 \end{eqnarray*}
-Note that $f' \in\#\Sigma_0$ \footnote[1]{(paper)}.
+Clearly $f' \in\#\Sigma_0$.
 Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
-However, it is showed in $^{[1]}$ that any counting problem in $\#\Sigma_0$ is computable in polynomial time, therefore, the decision version of $f$ is in P.
+However, it is showed in $^{[1]}$ that any counting problem in $\#\Sigma_0$ is computable in polynomial time.
+Therefore, the decision version of $f$ is in P.
 \end{proof}
-
-\begin{theo}
-$\#\Sigma_1$ is closed under substraction $\Rightarrow$ P = NP.
-\end{theo}
-\begin{proof}
-\#3DNF $\in \#\Sigma_1$.
-Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment in a \#3DNF instance.
-Suppose that $F_{2^n}-F_{\#3DNF} \in \#\Sigma_1$.
-This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete.
-However, as we showed previously (Theorem 1), it is also in P.
-Then, co-NP $\subseteq$ P.
-\end{proof}
-
-\begin{theo}
-$\#\Sigma_1 \subsetneq \E1$
-\end{theo}
-\begin{proof}
-We will show that the $\E1$ function defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
-\end{proof}
-
-For a given function $f$, we define $f \dotminus 1$ as follows:
-\begin{eqnarray*}
-f \dotminus 1(\A) =
-\begin{cases}
-f(\A)-1, & \text{if }f(\A) > 0 \\
-0, & \text{if }f(\A) = 0.
-\end{cases}
-\end{eqnarray*}
-for every structure $\A$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
-
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1
-
-\begin{theo}
-$\E1$ is closed under substraction by one.
-\end{theo}
-\begin{proof}
-For every $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$, let $\z = (z_1,\dots,z_d)$ and let $\X = (X_1,\dots,X_r)$.
-There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$.
-This separates the proof in three cases:
-\begin{enumerate}
-
-% CASE 1
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\e \in A^d$.
-Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily.
-First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
-\begin{eqnarray*}
-\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
-\end{eqnarray*}
-This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
-\begin{eqnarray*}
-\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
-\end{eqnarray*}
-If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
-Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$.
-Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E1.$
-
-%CASE 2
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
-\begin{eqnarray}
-\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
-\end{eqnarray}
-where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$.
-We also assume that $(\x_1,\dots,\x_n,\y) = \x$.
-As an example, the following formula is of this form:
-\begin{align*}
-\varphi(\x,\X) =  X_1(x_1,x_2) \wedge \neg X_1(x_3,x_4) \wedge (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 ) \wedge \forall z\big( S_1(x_1,z) \big),
-\end{align*}
-where $\x = (x_1,x_2,x_3,x_4)$ and $\X = (X_1)$.
-Here, $n = 1$, $\lambda(1) = \lambda(2) = 1$, $\x_1 = (x_1,x_2)$ and $\y = (x_3,x_4)$.
-Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_3,x_4)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4)$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z) \big)$.
-
-Similarly to the previous proof, we would like to eliminate the lexicographically smallest tuple of predicates that satisfies the formula \eqref{f1}.
-Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
-We now define
-\begin{multline*}
-\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to \varphi_{m,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
-\end{multline*}
-where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
-Note that $\alpha_{\min}$ is satisfied only by the lexicographically \linebreak smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
-Our new formula is
-\begin{multline}
-\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge\exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \\ \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg] \label{f2}.
-\end{multline}
-We now show a result by which the main proof will follow.
-\begin{lemma}  \label{lemmaone}
-$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
-\end{lemma}
-\begin{proof}
-Let $\A \in \Truc$.
-Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{\lambda(j)=i}\{\d_j\}$.
-We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
-\begin{enumerate}
-\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$.
-That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$).
-Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$.
-Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$.
-We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \z < \x_i$ or (3) $\theta(\x)\models \z > \x_i$.
-Considering that (2) and (3) are not possible since $\A\models\theta(\d)$, and both $\z$ and $\x_i$ are assigned the value $\d_i$, then $\theta(\x)\models \z = \x_i$.
-But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$ from which we conclude that $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which leads to a contradiction.
-\item[(b)] By the construction of $\P$, we see that 
-$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big).$$ However, the only possible assignment to $\alpha_{\min}(\x_1,\dots,\x_n)$ is $(\d_1,\dots,\d_n)$.
-Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
-\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$.
-Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $\P^\prime$, then $\A\models \neg P^\prime_i(\d_i)$.
-And so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $\P_i$ does not have.
-If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-On both cases, we have a contradiction.
-\end{enumerate}
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi(\x,\P)$, we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
-\end{proof}
-
-We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
-By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\hat{\varphi}_i(\x,\X)$ with $i\in\{1,\ldots,\hat{k}\}$ such that
-\begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \hat{\varphi}_1(\x,\X) \vee \hat{\varphi}_2(\x,\X) \vee \dots  \vee \hat{\varphi}_{\hat{k}}(\x,\X),
-\end{eqnarray*}
-where, for every $i\in\{1,\ldots,\hat{k}\}$, 
-\begin{eqnarray*}
-\hat{\varphi}_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \hat{\varphi}^{-}_i(\X,\y_i)  \wedge \beta_i(\x).
-\end{eqnarray*}
-The function $\lambda_i$ is defined analogously to the first part of the proof, the tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x.$ The formulas $\hat{\varphi}^{-}_i(\X,\y_i)$, $\beta_i(\x)$, are also defined analogously.
-Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables.
-Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$.
-Note that for every formula $\psi(\x)$,
-\begin{eqnarray*}
-\exists\x\:\psi(\x) &\equiv& \exists\x(\psi(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\psi(\x)\wedge\theta^{g(d)}(\x)).
-\end{eqnarray*}
-We define the following formulas $\check{\varphi}_i(\x,\X)$, for $i\in\{1,\ldots,\check{k}\},$ as follows: 
-\begin{eqnarray*}
-\check{\varphi}_1(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{g(d)}(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
-\check{\varphi}_{g(d)+1}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{2\cdot g(d)}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
-& \vdots & \\
-\check{\varphi}_{(\hat{k}-1) g(d)+1}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{\hat{k} \cdot g(d)}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^{g(d)}(\x).
-\end{eqnarray*}
-Note that $\check{k} = \hat{k} \cdot g(d).$ Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L-$structure $\A$, that is, every $i\in\{1,\ldots,\check{k}\}$ such that for every assignment $\s$ to $\x,$ $f_{\check{\varphi}_i(\s,\X)} = 0.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
-We order the remaining disjuncts and define the disjuncts $\varphi_i(\x,\X)$ such that for every $j\in\{1,\ldots,\check{k}\}$ where $\check{\varphi}_j(\x,\X)$ is satisfiable, there is an $i\in\{1,\ldots,k\}$ such that $\varphi_i(\x,\X) = \check{\varphi}_j(\x,\X).$
-
-As seen in lemma \ref{lemmaone}, for every $\L-$structure $\A$ and $i\in\{1,\ldots,k\}$, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$ but $f_{\varphi_i(\s,\X)}(\A) = 0,$ then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) = 0$, so it follows that for every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) > 0.$
-
-Now we can define the formula like we did on the first part of the proof. For every $i\in\{1,\ldots,k\}$,
-\begin{eqnarray*}
-\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
-\end{eqnarray*}
- Also, note that
-\begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X).
-\end{eqnarray*}
-
-Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\P$ such that $\A\models\varphi_2(\x,\P)$, and so on.
-As we already know how to exclude that assignment in the first disjunct, we now move to the next disjuncts.
-Similarly to the first part of the proof, let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
-\begin{multline*}
-\alpha^i_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_n,\w)\wedge\bigvee_{j=1}^n(\u_j\neq\v_j)\big)\to \varphi_{m_i,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
-\end{multline*}
-where $\u_j,\v_j,$ for $\{1,\ldots,n_i\}$ and $\w$ have the adequate number of variables. Also, let
-\begin{multline*}
-\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_n\bigg[\alpha^i_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
-\end{multline*}
-Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
-\begin{multline}
-\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge \\ (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
-\end{multline}
-Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$.
-Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone} for this sub-formula.
-As we did in Lemma \ref{lemmaone} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
-\begin{enumerate}
-\item[(a)] We know that $\A\models\exists\x\:\varphi_q(\x,\P),$ so it immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
-\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
-(1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
-(2) We can use the proof in Lemma \ref{lemmaone} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$.
-However, $\A\not\models\exists\x\:\hat{\psi}_q(\P)$ so $\A\not\models\exists\x\:\varphi^\prime_i(\P)$.
-And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
-\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
-As we deduce from the part (c) of Lemma \ref{lemmaone}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
-Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ Let $q^\prime$ be such that $\A\models\exists\x\:\varphi_{q^\prime}(\x,\P^\prime).$ By the choice of $q,$ then $q^\prime \geq q,$ but since $\P \neq \P^\prime,$ then $q^\prime > q.$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x),$ so $\A\models\psi_i(\P^\prime)$ for each $i$.
-But also, $\A\models\exists\x\:\alpha_q(\x),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which is a contradiction.
-\end{enumerate}
-
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
-
-% CASE 3
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_d)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^d$.
-We define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2:
-\begin{eqnarray*}
-\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\v_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
-\end{eqnarray*}
-where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_i + c_i = a_i$ for $i\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, $v_i$ has $q_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\v) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
-The formulas $\varphi^{-}_i(\X,\y_i,\v_i)$, $\beta_i(\x,\z)$, are also defined analogously.
-
-\end{enumerate}
-\end{proof}
-
-\end{document}

### 2015-04-08 tex/th2.tex
8 abril

--- 2015-01-21 sharp-sigma-one-plus.tex
@@
-\documentclass[12pt]{article}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsthm} 
-\usepackage{fullpage}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{bm}
-
-\def\dotminus{\mathbin{\ooalign{\hss\raise1ex\hbox{.}\hss\cr
-  \mathsurround=0pt$-$}}}
-
-\def\E1{\#\Sigma_1^{+}}
-
-\def\Truc{\textsc{Struct}[\L]}
-
-\def\A{{\frak A}}
-\def\C{{\cal C}}
-\def\F{{\cal F}}
-\def\L{{\cal L}}
-\def\N{\mathbb{N}}
-\def\P{\vec{P}}
-\def\R{\vec{R}}
-\def\S{\vec{S}}
-\def\X{\vec{X}}
-\def\Y{\vec{Y}}
-\def\Z{\vec{Z}}
-%% a - arity of \X / arity of assignments \P to \X
-%% b - arity of predicates in \S
-%% c - arity of auxiliar predicates/variables
-\def\d{\vec{d}} %% counted elements
-\def\e{\vec{e}} %% counted elements
-%% f - counting function
-%% g - other functions
-%% h - other functions
-%% i - index
-%% j - index
-%% k - emergency index / size of tuple
-%% l - emergency index / size of tuple
-\def\l{\vec{\ell}}
-%% m - size of variable tuple
-%% n - size of predicate tuple
-%% o - not used
-%% p - 
-%% q - 
-%% r - size of \X / \P
-\def\s{\vec{s}}
-%% t - size of \S
-\def\u{\vec{u}} %% auxiliary variables
-\def\v{\vec{v}} %% auxiliary variables
-\def\w{\vec{w}} %% auxiliary variables
-\def\x{\vec{x}} %% quantified variables
-\def\y{\vec{y}} %% auxiliary variables
-\def\z{\vec{z}} %% open variables
-\def\ep{\vec{o}}
-\def\ga{\vec{p}}
-
-\newtheorem{theo}{Theorem}
-\newtheorem{lemma}[theo]{Lemma}
-\newtheorem{claim}[theo]{Claim}
-\newtheorem{coro}[theo]{Corollary}
-
-\begin{document}
-
-\begin{center}
-{ \LARGE \bf
-  Some properties of $\E1$
-}
-\end{center}
-
-We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
-\begin{eqnarray*}
-\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
-&& \leq \text{ is interpreted as a total order for } A \}.
-\end{eqnarray*}
-We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$.
-A quantifier-free $\L$-formula is defined by the following grammar:
-\begin{eqnarray*}
-\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
-&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
-&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
-\end{eqnarray*}
-where $x,y$ and $x_i$ are first order variables for every $i$. We now define an extended quantifier-free $\L$-formula as follows:
-\begin{eqnarray*}
-\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
-&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
-&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
-\end{eqnarray*}
-Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables.
-For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
-\begin{eqnarray*}
-f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $a_1,\ldots,a_q$, and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.
-
-Let $\X = (X_1,\dots,X_r)$, and let $\z$ be a tuple of variables.
-A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\X,\z)$ such that $f = f_{\exists \x \: \varphi(\x,\X,\z)}.$\\
-
-The {\em decision version} of a function $f$ is defined by the language $L_f = \{\A \mid f(\A) > 0\}$.
-\begin{theo}
-The decision version of a function in $\E1$ is in \textsc{P}.
-\end{theo}
-\begin{proof}
-Let $f$ be a function in $\E1$. Then there is a formula $\varphi(\x,\X,\z)$ such that
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $\z$ is an $m$-tuple of variables, $\e$ is an $m$-tuple of elements and $\x$ is a $k$-tuple of variables.
-Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible evaluations for $\z$.
-Let $\ell = \vert A \vert^k$ and $\d_1,\dots,d_\ell\in A^k$ be all possible evaluations for $\x$.
-Let $\varphi_\A(\X)$ be defined as follows:
-\begin{eqnarray*}
-\varphi_\A(\X) = \varphi(\d_1,\X,\e_1) \vee \cdots \vee \varphi(\d_1,\X,\e_n) \vee \varphi(\d_2,\X,\e_1) \vee \cdots \vee \varphi(\d_n,\X,\e_n).
-\end{eqnarray*}
-Note that $\varphi_\A(\X)$ will have at least one assignment for $\X$ iff $f(\A)>0$.
-Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction.
-Note that $\psi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$.
-Then, let $f'$ be defined as follows:
-\begin{eqnarray*}
-f'(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \psi_\A(\P) \} \mid.
-\end{eqnarray*}
-Note that $f' \in\#\Sigma_0$ \footnote[1]{(paper)}.
-Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
-However, it is showed in $^{[1]}$ that any counting problem in $\#\Sigma_0$ is computable in polynomial time, therefore, the decision version of $f$ is in P.
-\end{proof}
-
 \begin{theo}
@@
 \begin{proof}
-\#3DNF $\in \#\Sigma_1$.
-Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment in a \#3DNF instance.
+Let $F_{\#3DNF}$ be a function that counts the satisfying assignments to a 3DNF formula.
+As shown in [1], $F_{\#3DNF} \in \#\Sigma_1$.
+Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment (satisfying or not) to a 3DNF formula.
 Suppose that $F_{2^n}-F_{\#3DNF} \in \#\Sigma_1$.
 This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete.
-However, as we showed previously (Theorem 1), it is also in P.
+However, as we showed previously in Theorem \ref{decisionptime}, it is also in P.
 Then, co-NP $\subseteq$ P.
 \end{proof}
-
-\begin{theo}
-$\#\Sigma_1 \subsetneq \E1$
-\end{theo}
-\begin{proof}
-We will show that the $\E1$ function defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
-\end{proof}
-
-For a given function $f$, we define $f \dotminus 1$ as follows:
-\begin{eqnarray*}
-f \dotminus 1(\A) =
-\begin{cases}
-f(\A)-1, & \text{if }f(\A) > 0 \\
-0, & \text{if }f(\A) = 0.
-\end{cases}
-\end{eqnarray*}
-for every structure $\A$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
-
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1
-
-\begin{theo}
-$\E1$ is closed under substraction by one.
-\end{theo}
-\begin{proof}
-For every $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$, let $\z = (z_1,\dots,z_d)$ and let $\X = (X_1,\dots,X_r)$.
-There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$.
-This separates the proof in three cases:
-\begin{enumerate}
-
-% CASE 1
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\e \in A^d$.
-Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily.
-First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
-\begin{eqnarray*}
-\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
-\end{eqnarray*}
-This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
-\begin{eqnarray*}
-\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
-\end{eqnarray*}
-If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
-Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$.
-Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E1.$
-
-%CASE 2
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
-\begin{eqnarray}
-\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
-\end{eqnarray}
-where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$.
-We also assume that $(\x_1,\dots,\x_n,\y) = \x$.
-As an example, the following formula is of this form:
-\begin{align*}
-\varphi(\x,\X) =  X_1(x_1,x_2) \wedge \neg X_1(x_3,x_4) \wedge (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 ) \wedge \forall z\big( S_1(x_1,z) \big),
-\end{align*}
-where $\x = (x_1,x_2,x_3,x_4)$ and $\X = (X_1)$.
-Here, $n = 1$, $\lambda(1) = \lambda(2) = 1$, $\x_1 = (x_1,x_2)$ and $\y = (x_3,x_4)$.
-Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_3,x_4)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4)$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z) \big)$.
-
-Similarly to the previous proof, we would like to eliminate the lexicographically smallest tuple of predicates that satisfies the formula \eqref{f1}.
-Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
-We now define
-\begin{multline*}
-\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to \varphi_{m,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
-\end{multline*}
-where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
-Note that $\alpha_{\min}$ is satisfied only by the lexicographically \linebreak smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
-Our new formula is
-\begin{multline}
-\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge\exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \\ \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg] \label{f2}.
-\end{multline}
-We now show a result by which the main proof will follow.
-\begin{lemma}  \label{lemmaone}
-$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
-\end{lemma}
-\begin{proof}
-Let $\A \in \Truc$.
-Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{\lambda(j)=i}\{\d_j\}$.
-We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
-\begin{enumerate}
-\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$.
-That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$).
-Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$.
-Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$.
-We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \z < \x_i$ or (3) $\theta(\x)\models \z > \x_i$.
-Considering that (2) and (3) are not possible since $\A\models\theta(\d)$, and both $\z$ and $\x_i$ are assigned the value $\d_i$, then $\theta(\x)\models \z = \x_i$.
-But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$ from which we conclude that $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which leads to a contradiction.
-\item[(b)] By the construction of $\P$, we see that 
-$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big).$$ However, the only possible assignment to $\alpha_{\min}(\x_1,\dots,\x_n)$ is $(\d_1,\dots,\d_n)$.
-Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
-\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$.
-Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $\P^\prime$, then $\A\models \neg P^\prime_i(\d_i)$.
-And so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $\P_i$ does not have.
-If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-On both cases, we have a contradiction.
-\end{enumerate}
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi(\x,\P)$, we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
-\end{proof}
-
-We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
-By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\hat{\varphi}_i(\x,\X)$ with $i\in\{1,\ldots,\hat{k}\}$ such that
-\begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \hat{\varphi}_1(\x,\X) \vee \hat{\varphi}_2(\x,\X) \vee \dots  \vee \hat{\varphi}_{\hat{k}}(\x,\X),
-\end{eqnarray*}
-where, for every $i\in\{1,\ldots,\hat{k}\}$, 
-\begin{eqnarray*}
-\hat{\varphi}_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \hat{\varphi}^{-}_i(\X,\y_i)  \wedge \beta_i(\x).
-\end{eqnarray*}
-The function $\lambda_i$ is defined analogously to the first part of the proof, the tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x.$ The formulas $\hat{\varphi}^{-}_i(\X,\y_i)$, $\beta_i(\x)$, are also defined analogously.
-Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables.
-Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$.
-Note that for every formula $\psi(\x)$,
-\begin{eqnarray*}
-\exists\x\:\psi(\x) &\equiv& \exists\x(\psi(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\psi(\x)\wedge\theta^{g(d)}(\x)).
-\end{eqnarray*}
-We define the following formulas $\check{\varphi}_i(\x,\X)$, for $i\in\{1,\ldots,\check{k}\},$ as follows: 
-\begin{eqnarray*}
-\check{\varphi}_1(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{g(d)}(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
-\check{\varphi}_{g(d)+1}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{2\cdot g(d)}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
-& \vdots & \\
-\check{\varphi}_{(\hat{k}-1) g(d)+1}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{\hat{k} \cdot g(d)}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^{g(d)}(\x).
-\end{eqnarray*}
-Note that $\check{k} = \hat{k} \cdot g(d).$ Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L-$structure $\A$, that is, every $i\in\{1,\ldots,\check{k}\}$ such that for every assignment $\s$ to $\x,$ $f_{\check{\varphi}_i(\s,\X)} = 0.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
-We order the remaining disjuncts and define the disjuncts $\varphi_i(\x,\X)$ such that for every $j\in\{1,\ldots,\check{k}\}$ where $\check{\varphi}_j(\x,\X)$ is satisfiable, there is an $i\in\{1,\ldots,k\}$ such that $\varphi_i(\x,\X) = \check{\varphi}_j(\x,\X).$
-
-As seen in lemma \ref{lemmaone}, for every $\L-$structure $\A$ and $i\in\{1,\ldots,k\}$, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$ but $f_{\varphi_i(\s,\X)}(\A) = 0,$ then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) = 0$, so it follows that for every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) > 0.$
-
-Now we can define the formula like we did on the first part of the proof. For every $i\in\{1,\ldots,k\}$,
-\begin{eqnarray*}
-\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
-\end{eqnarray*}
- Also, note that
-\begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X).
-\end{eqnarray*}
-
-Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\P$ such that $\A\models\varphi_2(\x,\P)$, and so on.
-As we already know how to exclude that assignment in the first disjunct, we now move to the next disjuncts.
-Similarly to the first part of the proof, let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
-\begin{multline*}
-\alpha^i_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_n,\w)\wedge\bigvee_{j=1}^n(\u_j\neq\v_j)\big)\to \varphi_{m_i,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
-\end{multline*}
-where $\u_j,\v_j,$ for $\{1,\ldots,n_i\}$ and $\w$ have the adequate number of variables. Also, let
-\begin{multline*}
-\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_n\bigg[\alpha^i_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
-\end{multline*}
-Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
-\begin{multline}
-\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge \\ (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
-\end{multline}
-Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$.
-Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone} for this sub-formula.
-As we did in Lemma \ref{lemmaone} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
-\begin{enumerate}
-\item[(a)] We know that $\A\models\exists\x\:\varphi_q(\x,\P),$ so it immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
-\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
-(1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
-(2) We can use the proof in Lemma \ref{lemmaone} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$.
-However, $\A\not\models\exists\x\:\hat{\psi}_q(\P)$ so $\A\not\models\exists\x\:\varphi^\prime_i(\P)$.
-And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
-\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
-As we deduce from the part (c) of Lemma \ref{lemmaone}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
-Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ Let $q^\prime$ be such that $\A\models\exists\x\:\varphi_{q^\prime}(\x,\P^\prime).$ By the choice of $q,$ then $q^\prime \geq q,$ but since $\P \neq \P^\prime,$ then $q^\prime > q.$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x),$ so $\A\models\psi_i(\P^\prime)$ for each $i$.
-But also, $\A\models\exists\x\:\alpha_q(\x),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which is a contradiction.
-\end{enumerate}
-
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
-
-% CASE 3
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_d)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^d$.
-We define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2:
-\begin{eqnarray*}
-\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\v_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
-\end{eqnarray*}
-where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_i + c_i = a_i$ for $i\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, $v_i$ has $q_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\v) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
-The formulas $\varphi^{-}_i(\X,\y_i,\v_i)$, $\beta_i(\x,\z)$, are also defined analogously.
-
-\end{enumerate}
-\end{proof}
-
-\end{document}

### 2015-04-08 tex/th3.tex
8 abril

\begin{theo}
$\#\Sigma_1 \subsetneq \E1$
\end{theo}
\begin{proof}
We will show that the $\E1$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$. By contradiction, suppose that it is. Let $\A = \langle A =\{1\},S^\A = \{1\},\leq^\A=\{(1,1)\} \rangle$. Then, $f(\A) = 1$. Now let $\B = \langle B =\{1,2\},S^\B = \{1\},\leq^\B=\{(1,1),(1,2),(2,2)\} \rangle$. Since $\#\Sigma_1$ is monotonic under addition of elements, $f(\B) \geq 1$. However, there is no assignment $s\in B$ to $x$ such that $\B\models\:\varphi(s)$, so $f(\B) = 0$, which leads to a contradiction.
\end{proof}

### 2015-04-08 tex/th4.tex
8 abril

--- 2015-01-21 sharp-sigma-one-plus.tex
@@
-\documentclass[12pt]{article}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsthm} 
-\usepackage{fullpage}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{bm}
-
-\def\dotminus{\mathbin{\ooalign{\hss\raise1ex\hbox{.}\hss\cr
-  \mathsurround=0pt$-$}}}
-
-\def\E1{\#\Sigma_1^{+}}
-
-\def\Truc{\textsc{Struct}[\L]}
-
-\def\A{{\frak A}}
-\def\C{{\cal C}}
-\def\F{{\cal F}}
-\def\L{{\cal L}}
-\def\N{\mathbb{N}}
-\def\P{\vec{P}}
-\def\R{\vec{R}}
-\def\S{\vec{S}}
-\def\X{\vec{X}}
-\def\Y{\vec{Y}}
-\def\Z{\vec{Z}}
-%% a - arity of \X / arity of assignments \P to \X
-%% b - arity of predicates in \S
-%% c - arity of auxiliar predicates/variables
-\def\d{\vec{d}} %% counted elements
-\def\e{\vec{e}} %% counted elements
-%% f - counting function
-%% g - other functions
-%% h - other functions
-%% i - index
-%% j - index
-%% k - emergency index / size of tuple
-%% l - emergency index / size of tuple
-\def\l{\vec{\ell}}
-%% m - size of variable tuple
-%% n - size of predicate tuple
-%% o - not used
-%% p - 
-%% q - 
-%% r - size of \X / \P
-\def\s{\vec{s}}
-%% t - size of \S
-\def\u{\vec{u}} %% auxiliary variables
-\def\v{\vec{v}} %% auxiliary variables
-\def\w{\vec{w}} %% auxiliary variables
-\def\x{\vec{x}} %% quantified variables
-\def\y{\vec{y}} %% auxiliary variables
-\def\z{\vec{z}} %% open variables
-\def\ep{\vec{o}}
-\def\ga{\vec{p}}
-
-\newtheorem{theo}{Theorem}
-\newtheorem{lemma}[theo]{Lemma}
-\newtheorem{claim}[theo]{Claim}
-\newtheorem{coro}[theo]{Corollary}
-
-\begin{document}
-
-\begin{center}
-{ \LARGE \bf
-  Some properties of $\E1$
-}
-\end{center}
-
-We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
-\begin{eqnarray*}
-\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
-&& \leq \text{ is interpreted as a total order for } A \}.
-\end{eqnarray*}
-We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$.
-A quantifier-free $\L$-formula is defined by the following grammar:
-\begin{eqnarray*}
-\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
-&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
-&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
-\end{eqnarray*}
-where $x,y$ and $x_i$ are first order variables for every $i$. We now define an extended quantifier-free $\L$-formula as follows:
-\begin{eqnarray*}
-\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
-&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
-&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
-\end{eqnarray*}
-Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables.
-For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
-\begin{eqnarray*}
-f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $a_1,\ldots,a_q$, and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.
-
-Let $\X = (X_1,\dots,X_r)$, and let $\z$ be a tuple of variables.
-A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\X,\z)$ such that $f = f_{\exists \x \: \varphi(\x,\X,\z)}.$\\
-
-The {\em decision version} of a function $f$ is defined by the language $L_f = \{\A \mid f(\A) > 0\}$.
-\begin{theo}
-The decision version of a function in $\E1$ is in \textsc{P}.
-\end{theo}
-\begin{proof}
-Let $f$ be a function in $\E1$. Then there is a formula $\varphi(\x,\X,\z)$ such that
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $\z$ is an $m$-tuple of variables, $\e$ is an $m$-tuple of elements and $\x$ is a $k$-tuple of variables.
-Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible evaluations for $\z$.
-Let $\ell = \vert A \vert^k$ and $\d_1,\dots,d_\ell\in A^k$ be all possible evaluations for $\x$.
-Let $\varphi_\A(\X)$ be defined as follows:
-\begin{eqnarray*}
-\varphi_\A(\X) = \varphi(\d_1,\X,\e_1) \vee \cdots \vee \varphi(\d_1,\X,\e_n) \vee \varphi(\d_2,\X,\e_1) \vee \cdots \vee \varphi(\d_n,\X,\e_n).
-\end{eqnarray*}
-Note that $\varphi_\A(\X)$ will have at least one assignment for $\X$ iff $f(\A)>0$.
-Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction.
-Note that $\psi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$.
-Then, let $f'$ be defined as follows:
-\begin{eqnarray*}
-f'(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \psi_\A(\P) \} \mid.
-\end{eqnarray*}
-Note that $f' \in\#\Sigma_0$ \footnote[1]{(paper)}.
-Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
-However, it is showed in $^{[1]}$ that any counting problem in $\#\Sigma_0$ is computable in polynomial time, therefore, the decision version of $f$ is in P.
-\end{proof}
-
-\begin{theo}
-$\#\Sigma_1$ is closed under substraction $\Rightarrow$ P = NP.
-\end{theo}
-\begin{proof}
-\#3DNF $\in \#\Sigma_1$.
-Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment in a \#3DNF instance.
-Suppose that $F_{2^n}-F_{\#3DNF} \in \#\Sigma_1$.
-This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete.
-However, as we showed previously (Theorem 1), it is also in P.
-Then, co-NP $\subseteq$ P.
-\end{proof}
-
-\begin{theo}
-$\#\Sigma_1 \subsetneq \E1$
-\end{theo}
-\begin{proof}
-We will show that the $\E1$ function defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
-\end{proof}
-
-For a given function $f$, we define $f \dotminus 1$ as follows:
-\begin{eqnarray*}
-f \dotminus 1(\A) =
-\begin{cases}
-f(\A)-1, & \text{if }f(\A) > 0 \\
-0, & \text{if }f(\A) = 0.
-\end{cases}
-\end{eqnarray*}
-for every structure $\A$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
-
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1
-
 \begin{theo}
@@
 \begin{enumerate}
+\item \input{case1.tex}
 
-% CASE 1
+\item \input{case2part1.tex}
+\input{lemmafirst.tex}
+\input{case2general.tex}
+\input{case2generalproof.tex}
 
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\e \in A^d$.
-Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily.
-First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
-\begin{eqnarray*}
-\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
-\end{eqnarray*}
-This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
-\begin{eqnarray*}
-\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
-\end{eqnarray*}
-If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
-Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$.
-Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E1.$
-
-%CASE 2
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
-\begin{eqnarray}
-\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
-\end{eqnarray}
-where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$.
-We also assume that $(\x_1,\dots,\x_n,\y) = \x$.
-As an example, the following formula is of this form:
-\begin{align*}
-\varphi(\x,\X) =  X_1(x_1,x_2) \wedge \neg X_1(x_3,x_4) \wedge (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 ) \wedge \forall z\big( S_1(x_1,z) \big),
-\end{align*}
-where $\x = (x_1,x_2,x_3,x_4)$ and $\X = (X_1)$.
-Here, $n = 1$, $\lambda(1) = \lambda(2) = 1$, $\x_1 = (x_1,x_2)$ and $\y = (x_3,x_4)$.
-Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_3,x_4)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4)$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z) \big)$.
-
-Similarly to the previous proof, we would like to eliminate the lexicographically smallest tuple of predicates that satisfies the formula \eqref{f1}.
-Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
-We now define
-\begin{multline*}
-\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to \varphi_{m,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
-\end{multline*}
-where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
-Note that $\alpha_{\min}$ is satisfied only by the lexicographically \linebreak smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
-Our new formula is
-\begin{multline}
-\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge\exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \\ \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg] \label{f2}.
-\end{multline}
-We now show a result by which the main proof will follow.
-\begin{lemma}  \label{lemmaone}
-$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
-\end{lemma}
-\begin{proof}
-Let $\A \in \Truc$.
-Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{\lambda(j)=i}\{\d_j\}$.
-We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
-\begin{enumerate}
-\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$.
-That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$).
-Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$.
-Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$.
-We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \z < \x_i$ or (3) $\theta(\x)\models \z > \x_i$.
-Considering that (2) and (3) are not possible since $\A\models\theta(\d)$, and both $\z$ and $\x_i$ are assigned the value $\d_i$, then $\theta(\x)\models \z = \x_i$.
-But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$ from which we conclude that $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which leads to a contradiction.
-\item[(b)] By the construction of $\P$, we see that 
-$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big).$$ However, the only possible assignment to $\alpha_{\min}(\x_1,\dots,\x_n)$ is $(\d_1,\dots,\d_n)$.
-Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
-\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$.
-Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $\P^\prime$, then $\A\models \neg P^\prime_i(\d_i)$.
-And so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $\P_i$ does not have.
-If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-On both cases, we have a contradiction.
-\end{enumerate}
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi(\x,\P)$, we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
-\end{proof}
-
-We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
-By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\hat{\varphi}_i(\x,\X)$ with $i\in\{1,\ldots,\hat{k}\}$ such that
-\begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \hat{\varphi}_1(\x,\X) \vee \hat{\varphi}_2(\x,\X) \vee \dots  \vee \hat{\varphi}_{\hat{k}}(\x,\X),
-\end{eqnarray*}
-where, for every $i\in\{1,\ldots,\hat{k}\}$, 
-\begin{eqnarray*}
-\hat{\varphi}_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \hat{\varphi}^{-}_i(\X,\y_i)  \wedge \beta_i(\x).
-\end{eqnarray*}
-The function $\lambda_i$ is defined analogously to the first part of the proof, the tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x.$ The formulas $\hat{\varphi}^{-}_i(\X,\y_i)$, $\beta_i(\x)$, are also defined analogously.
-Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables.
-Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$.
-Note that for every formula $\psi(\x)$,
-\begin{eqnarray*}
-\exists\x\:\psi(\x) &\equiv& \exists\x(\psi(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\psi(\x)\wedge\theta^{g(d)}(\x)).
-\end{eqnarray*}
-We define the following formulas $\check{\varphi}_i(\x,\X)$, for $i\in\{1,\ldots,\check{k}\},$ as follows: 
-\begin{eqnarray*}
-\check{\varphi}_1(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{g(d)}(\x,\X) &=& \hat{\varphi}_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
-\check{\varphi}_{g(d)+1}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{2\cdot g(d)}(\x,\X) &=& \hat{\varphi}_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
-& \vdots & \\
-\check{\varphi}_{(\hat{k}-1) g(d)+1}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^1(\x), \\
-& \vdots & \\
-\check{\varphi}_{\hat{k} \cdot g(d)}(\x,\X) &=& \hat{\varphi}_{\hat{k}}(\x,\X) \wedge \theta^{g(d)}(\x).
-\end{eqnarray*}
-Note that $\check{k} = \hat{k} \cdot g(d).$ Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L-$structure $\A$, that is, every $i\in\{1,\ldots,\check{k}\}$ such that for every assignment $\s$ to $\x,$ $f_{\check{\varphi}_i(\s,\X)} = 0.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
-We order the remaining disjuncts and define the disjuncts $\varphi_i(\x,\X)$ such that for every $j\in\{1,\ldots,\check{k}\}$ where $\check{\varphi}_j(\x,\X)$ is satisfiable, there is an $i\in\{1,\ldots,k\}$ such that $\varphi_i(\x,\X) = \check{\varphi}_j(\x,\X).$
-
-As seen in lemma \ref{lemmaone}, for every $\L-$structure $\A$ and $i\in\{1,\ldots,k\}$, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$ but $f_{\varphi_i(\s,\X)}(\A) = 0,$ then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) = 0$, so it follows that for every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, then $f_{\exists\x\:\varphi_i(\x,\X)}(\A) > 0.$
-
-Now we can define the formula like we did on the first part of the proof. For every $i\in\{1,\ldots,k\}$,
-\begin{eqnarray*}
-\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
-\end{eqnarray*}
- Also, note that
-\begin{eqnarray*}
-\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X).
-\end{eqnarray*}
-
-Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\P$ such that $\A\models\varphi_2(\x,\P)$, and so on.
-As we already know how to exclude that assignment in the first disjunct, we now move to the next disjuncts.
-Similarly to the first part of the proof, let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
-\begin{multline*}
-\alpha^i_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_n,\w)\wedge\bigvee_{j=1}^n(\u_j\neq\v_j)\big)\to \varphi_{m_i,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
-\end{multline*}
-where $\u_j,\v_j,$ for $\{1,\ldots,n_i\}$ and $\w$ have the adequate number of variables. Also, let
-\begin{multline*}
-\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_n\bigg[\alpha^i_{\min}(\u_1,\dots,\u_n) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
-\end{multline*}
-Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
-\begin{multline}
-\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge \\ (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
-\end{multline}
-Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$.
-Suppose first that there is at least an assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\P$ be an assignment to $\X$ defined as in Lemma \ref{lemmaone} for this sub-formula.
-As we did in Lemma \ref{lemmaone} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
-\begin{enumerate}
-\item[(a)] We know that $\A\models\exists\x\:\varphi_q(\x,\P),$ so it immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
-\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
-(1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
-(2) We can use the proof in Lemma \ref{lemmaone} to see that $\A\not\models\exists\x\:\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears in $\varphi^\prime_i(\x,\X)$.
-However, $\A\not\models\exists\x\:\hat{\psi}_q(\P)$ so $\A\not\models\exists\x\:\varphi^\prime_i(\P)$.
-And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
-\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
-As we deduce from the part (c) of Lemma \ref{lemmaone}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
-Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ Let $q^\prime$ be such that $\A\models\exists\x\:\varphi_{q^\prime}(\x,\P^\prime).$ By the choice of $q,$ then $q^\prime \geq q,$ but since $\P \neq \P^\prime,$ then $q^\prime > q.$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x),$ so $\A\models\psi_i(\P^\prime)$ for each $i$.
-But also, $\A\models\exists\x\:\alpha_q(\x),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which is a contradiction.
-\end{enumerate}
-
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1.$
-
-% CASE 3
-
-\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_d)$. That is,
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^d$.
-We define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2:
-\begin{eqnarray*}
-\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\v_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
-\end{eqnarray*}
-where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_i + c_i = a_i$ for $i\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, $v_i$ has $q_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\v) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
-The formulas $\varphi^{-}_i(\X,\y_i,\v_i)$, $\beta_i(\x,\z)$, are also defined analogously.
-
+\item \input{case3.tex}
 \end{enumerate}
 \end{proof}
-
-\end{document}

### 2015-04-22 tex/th1.tex
22 abr

@@
 \end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $\z$ is an $m$-tuple of variables, $\e$ is an $m$-tuple of elements and $\x$ is a $k$-tuple of variables.
-Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible evaluations for $\z$.
-Let $\ell = \vert A \vert^k$ and $\d_1,\dots,\d_\ell\in A^k$ be all possible evaluations for $\x$.
+where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables.
+Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible assignments for $\z$.
+Let $\ell = \vert A \vert^k$ and $\d_1,\dots,\d_\ell\in A^k$ be all possible assignations for $\x$.
+Note that $m$ and $k$ are fixed values, so both $\vert A \vert^k$ and $n=\vert A \vert^m$ are polynomial on the size of $\A$.
 Let $\varphi_\A(\X)$ be defined as follows:
@@
 Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction.
-Note as well that $\psi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$.
+We can do this since checking $\A\models\varphi(\d_i,\X,\e_i)$ can be done in polynomial time for each $i\in\{1,\ldots,n\}$.
+Note as well that $\varphi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$.
 Then, let $f'$ be defined as follows:
@@
 Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
-However, it is showed in $^{[1]}$ that any counting problem in $\#\Sigma_0$ is computable in polynomial time.
-Therefore, the decision version of $f$ is in P.
+However, it is showed in \ref{Hem} that any counting problem in $\#\Sigma_0$ is computable in polynomial time.
+Therefore, the decision version of $f$ is in {\sc P}.
 \end{proof}

### 2015-04-22 tex/th2.tex
22 abr

@@
 \begin{theo}
-$\#\Sigma_1$ is closed under substraction $\Rightarrow$ P = NP.
+If $\#\Sigma_1$ is closed under substraction, then {\sc P} = {\sc NP}.
 \end{theo}
 \begin{proof}
-Let $F_{\#3DNF}$ be a function that counts the satisfying assignments to a 3DNF formula.
-As shown in [1], $F_{\#3DNF} \in \#\Sigma_1$.
+Let $f_{\#3DNF}$ be a function that counts the satisfying assignments to a 3DNF formula.
+As shown in [1], $f_{\#3DNF} \in \#\Sigma_1$.
 Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Suppose that $F_{2^n}-F_{\#3DNF} \in \#\Sigma_1$.
+Suppose that $f_{2^n}-f_{\#3DNF} \in \#\Sigma_1$.
 This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete.

### 2015-04-29 tex/th1.tex
29 abr 1

@@
 Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible assignments for $\z$.
-Let $\ell = \vert A \vert^k$ and $\d_1,\dots,\d_\ell\in A^k$ be all possible assignations for $\x$.
-Note that $m$ and $k$ are fixed values, so both $\vert A \vert^k$ and $n=\vert A \vert^m$ are polynomial on the size of $\A$.
+Let $\ell = \vert A \vert^k$ and $\d_1,\dots,\d_\ell\in A^k$ be all possible assignments for $\x$.
+Note that $m$ and $k$ are fixed values, so both $\ell = \vert A \vert^k$ and $n=\vert A \vert^m$ are polynomial in the size of $\A$.
 Let $\varphi_\A(\X)$ be defined as follows:
@@
 Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
-However, it is showed in \ref{Hem} that any counting problem in $\#\Sigma_0$ is computable in polynomial time.
+However, it is showed in \cite{DBLP:journals/jcss/SalujaST95} that any counting problem in $\#\Sigma_0$ is computable in polynomial time.
 Therefore, the decision version of $f$ is in {\sc P}.

### 2015-05-01 tex/th1.tex
6 may

@@
 \begin{proof}
-Let $f$ be a function in $\E1$. Then there is a formula $\varphi(\x,\X,\z)$ such that
+Let $f$ be a function in $\E1$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables.
+where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables.
 Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible assignments for $\z$.
 Let $\ell = \vert A \vert^k$ and $\d_1,\dots,\d_\ell\in A^k$ be all possible assignments for $\x$.
-Note that $m$ and $k$ are fixed values, so both $\ell = \vert A \vert^k$ and $n=\vert A \vert^m$ are polynomial in the size of $\A$.
+Note that the values of $m$ and $k$ come from $\varphi(\x,\X,\z)$, so both $\ell = \vert A \vert^k$ and $n=\vert A \vert^m$ are polynomial in the size of $\A$.
 Let $\varphi_\A(\X)$ be defined as follows:
@@
 Note that $\varphi_\A(\X)$ will have at least one assignment $\P$ for $\X$ such that $\A\models\varphi_\A(\P)$ iff $f(\A)>0$.
-Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction.
-We can do this since checking $\A\models\varphi(\d_i,\X,\e_i)$ can be done in polynomial time for each $i\in\{1,\ldots,n\}$.
-Note as well that $\varphi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$.
-Then, let $f'$ be defined as follows:
+Let $\psi_\A(\X)$ be the following formula: For each freely quantified sub-formula $\beta(\x,\z)$ in $\varphi(\x,\X,\z)$ we check if $\A \models \beta(\d_i,\e_j)$, for every $i\in\{1,\ldots,\ell\}$ and $j\in\{1,\ldots,m\}$.
+If it is true, then we replace $\beta(\d_i,\e_j)$ with a tautology, and otherwise, with a contradiction.
+Let us give an example on how this would be done: \\
+
+Let $\varphi(\x,\X,\z) = \exists x_1,x_2( X_1(x_2) \wedge ( X_2(z_1,x_1) \wedge \forall y( S_1(y,z_1) )  ))$.
+In this case, $\x = (x_1,x_2)$, $\X = (X_1,X_2)$ and $\z = (z_1)$.
+Let $\A\in\Truc$ be $\A = \langle A,S^\A,\leq^\A \rangle$ where $A = \{1,2\}$, $S^\A = \{(1,1),(1,2),(2,2)\}$, and $1 \leq^\A 2$.
+We generate all the possible assignments for $\x$ and $\z$.
+And our new formula is:
+\begin{multline*}
+\varphi_\A(\X) = (X_1(1) \wedge ( X_2(1,1) \wedge \forall y( S_1(y,1) )  ) \vee (X_1(1) \wedge ( X_2(2,1) \wedge \forall y( S_1(y,2) )  ) \vee \\
+(X_1(2) \wedge ( X_2(1,1) \wedge \forall y( S_1(y,1) )  ) \vee (X_1(2) \wedge ( X_2(2,1) \wedge \forall y( S_1(y,2) )  ) \vee \\
+(X_1(1) \wedge ( X_2(1,2) \wedge \forall y( S_1(y,1) )  ) \vee (X_1(1) \wedge ( X_2(2,2) \wedge \forall y( S_1(y,2) )  ) \vee \\
+(X_1(2) \wedge ( X_2(1,2) \wedge \forall y( S_1(y,1) )  ) \vee (X_1(2) \wedge ( X_2(2,2) \wedge \forall y( S_1(y,2) )  ).
+\end{multline*}
+Now we replace the satisfied subformulas:
+\begin{multline*}
+\psi_\A(\X) = (X_1(1) \wedge X_2(1,1) \wedge \perp  ) \vee (X_1(1) \wedge  X_2(2,1) \wedge \top  ) \vee \\
+(X_1(2) \wedge  X_2(1,1) \wedge \perp  ) \vee (X_1(2) \wedge  X_2(2,1) \wedge \top  ) \vee \\
+(X_1(1) \wedge  X_2(1,2) \wedge \perp  ) \vee (X_1(1) \wedge  X_2(2,2) \wedge \top  ) \vee \\
+(X_1(2) \wedge  X_2(1,2) \wedge \perp  ) \vee (X_1(2) \wedge  X_2(2,2) \wedge \top  ),
+\end{multline*}
+which can be simplified even further:
+\begin{multline*}
+\psi_\A(\X) = (X_1(1) \wedge X_2(2,1)  ) \vee (X_1(2) \wedge  X_2(2,1) ) \vee \\ (X_1(1) \wedge  X_2(2,2)  ) \vee (X_1(2) \wedge  X_2(2,2)  )
+\end{multline*}
+Note that we can do this since checking $\A\models\varphi(\d_i,\X,\e_j)$ can be done in polynomial time for each $i,j$.
+Note as well that $\psi_\A(\X)$ is quantifier-free and its size is also polynomial to the size of $\A$.
+However, this still is not a valid $\L-$formula since there are no constants in $\L$.
+Let $P_a$ be a unary predicate that is interpreted as $P_a = \{a\}$ for every $a\in A$.
+Now let $\zeta_\A(\X,\u)$ be $\psi_\A(\X)$ with the conjunct $P_{a_1}(u_{a_1}) \wedge \cdots \wedge P_{a_{\vert A \vert}}(u_{a_{\vert A \vert}})$, where $u_a$ is a first order variable, and with each of its constants $a\in A$ replaced with $u_a$.
+In the example:
+\begin{multline*}
+\zeta_\A(\X,\u) = \big((X_1(u_1) \wedge X_2(u_2,u_1)  ) \vee (X_1(u_2) \wedge  X_2(u_2,u_1) ) \vee \\ (X_1(u_1) \wedge  X_2(u_2,u_2)  ) \vee (X_1(u_2) \wedge  X_2(u_2,u_2)  )\big) \wedge P_1(u_1) \wedge P_2(u_2),
+\end{multline*}
+where $\u = (u_1,u_2)$. Now, let $\A^\prime = \A \cup \{P_a\mid a\in A\}$, and let $f'$ be defined as follows:
 \begin{eqnarray*}
-f'(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \psi_\A(\P) \} \mid.
+f'(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \zeta_\A(\P,\e) \} \mid.
 \end{eqnarray*}
 Clearly $f' \in\#\Sigma_0$.
-Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
+Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A^\prime) > 0$, so computing $f'$ with the instance $\A^\prime$ is enough to solve the decision version of $f$.
 However, it is showed in \cite{DBLP:journals/jcss/SalujaST95} that any counting problem in $\#\Sigma_0$ is computable in polynomial time.

### 2015-05-01 tex/th2.tex
6 may

@@
 \begin{proof}
-Let $f_{\#3DNF}$ be a function that counts the satisfying assignments to a 3DNF formula.
-As shown in [1], $f_{\#3DNF} \in \#\Sigma_1$.
-Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Suppose that $f_{2^n}-f_{\#3DNF} \in \#\Sigma_1$.
-This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete.
+Suppose that $\#\Sigma_1$ is closed under substraction, that is, for each pair of functions $f,g\in\#\Sigma_1,$ there is an $h\in\#\Sigma_1$ such that $f(\A)-g(\A) = h(\A)$ for each $\A\in\Truc$.
+
+Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\psi$, where $S_i$ is a ternary relation such that $(a_1,a_2,a_3) \in S_i$ iff $(\neg a_1 \wedge \cdots \wedge\neg a_i \wedge a_{i+1} \wedge \cdots \wedge a_3)$ appears as a disjunct in $\psi$, for each $i\in\{1,\ldots,4\}$.
+
+Now let $f_{\#3DNF}$ be a function that counts the satisfying assignments to a 3DNF formula $\psi$. As shown in [1], $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
+$$
+\varphi(x,X) = (X(x) = X(x)).
+$$
+Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
+As we supposed initially, let $h\in\#\Sigma_1$ such that $f_{all}-f_{\#3DNF} = h$.
+For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ only if $\psi$ is a tautology, so the decision version of $f_{all}-f_{\#3DNF}$ is co-NP-complete.
 However, as we showed previously in Theorem \ref{decisionptime}, it is also in P.

### 2015-05-01 tex/th3.tex
22 abr
29 abr 1
6 may

@@
 Now let $\B = \langle B =\{1,2\},S^\B = \{1\},\leq^\B=\{(1,1),(1,2),(2,2)\} \rangle$.
-Since $\#\Sigma_1$ is monotonic under addition of elements, $f(\B) \geq 1$.
+Note that $\A$ is an induced substructure of $\B$.
+
+He have that for each function $g \in \#\Sigma_1$ and structures $\A_1,\A_2\in\Truc$, if $\A_1$ is an induced substructure of $\A_2$, then $f(\A_1) \leq f(\A_2)$ \cite{DBLP:journals/jcss/SalujaST95}.
+Therefore, $g(\B) \geq g(\A) = 1$.
 However, there is no assignment $s\in B$ to $x$ such that $\B\models\:\varphi(s)$, so $f(\B) = 0$, which leads to a contradiction.

### 2015-05-12 tex/sharp-sigma-one-plus.tex
22 abr
29 abr 1
12 may

@@
 \usepackage{amssymb}
+\usepackage{natbib}
+\usepackage{subfiles}
 \usepackage{bm}
@@
 
-\def\E1{\#\Sigma_1^{+}}
+\def\E1{\#\Sigma_1[\mbox{{\sc FO}}]}
 
@@
 
+\def\np{\mbox{{\sc NP}}}
+
 \newtheorem{theo}{Theorem}
@@
 \end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, and $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.
+for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements from $\A$.
 
-Let $\Y = (Y_1,\dots,Y_r)$, and let $\y$ be a tuple of variables.
 A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\exists \x \: \varphi(\x,\Y,\y)}.$\\
 
-The {\em decision version} of a function $f$ is defined by the language $L_f = \{\A \mid f(\A) > 0\}$.
+The {\em decision problem} associated to a function $f$ is defined by the language $L_f = \{\A \in \Truc \mid f(\A) > 0\}$.
 
 \input{th1}
+
+For a given pair of functions $f,g$, we define $f \dotminus g$ as follows:
+\begin{eqnarray*}
+(f \dotminus g)(\A) =
+\begin{cases}
+f(\A)-g(\A), & \text{if }f(\A)>g(\A) \\
+0, & \text{if }f(\A) \leq g(\A).
+\end{cases}
+\end{eqnarray*}
+for every $\L$-structure $\A \in \Truc$. A function class $\F$ is {\em closed under substraction} if for every pair of functions $f,g \in \F$, it holds that $f \dotminus g \in \F$.
 
@@
 
+\bibliographystyle{plain}
+\bibliography{SalujaST95}
+
 \end{document}

### 2015-05-12 tex/th1.tex
12 may

--- 2015-04-29 tex/th1.tex
@@
+%!TEX root = sharp-sigma-one-plus.tex
 \begin{theo} \label{decisionptime}
-The decision version of a function in $\E1$ is in \textsc{P}.
+The decision problem associated to a function in $\E1$ is in \textsc{P}.
 \end{theo}
 \begin{proof}
-Let $f$ be a function in $\E1$. Then there is a formula $\varphi(\x,\X,\z)$ such that
+Let $f$ be a function in $\E1$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables.
+where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables.
 Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible assignments for $\z$.
 Let $\ell = \vert A \vert^k$ and $\d_1,\dots,\d_\ell\in A^k$ be all possible assignments for $\x$.
-Note that $m$ and $k$ are fixed values, so both $\ell = \vert A \vert^k$ and $n=\vert A \vert^m$ are polynomial in the size of $\A$.
-Let $\varphi_\A(\X)$ be defined as follows:
+Note that the values of $m$ and $k$ come from $\varphi(\x,\X,\z)$, so both $\ell = \vert A \vert^k$ and $n=\vert A \vert^m$ are polynomial in the size of $\A$.
+Slightly abusing notation, let $\varphi_\A(\X)$ be defined as follows:
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-Note that $\varphi_\A(\X)$ will have at least one assignment $\P$ for $\X$ such that $\A\models\varphi_\A(\P)$ iff $f(\A)>0$.
-Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction.
-We can do this since checking $\A\models\varphi(\d_i,\X,\e_i)$ can be done in polynomial time for each $i\in\{1,\ldots,n\}$.
-Note as well that $\varphi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$.
-Then, let $f'$ be defined as follows:
+Note that $\varphi_\A(\X)$ will have at least one assignment $\P$ for $\X$ such that $\A\models\varphi_\A(\P)$ if and only if $f(\A)>0$.
+Let $\psi_\A(\X)$ be the following formula: For each FO-formula $\beta(\x,\z)$ in $\varphi(\x,\X,\z)$ we check if $\A \models \beta(\d_i,\e_j)$, for every $i\in\{1,\ldots,\ell\}$ and $j\in\{1,\ldots,m\}$.
+If it is true, then we replace $\beta(\d_i,\e_j)$ with a tautology, and otherwise, with a contradiction.
+Let us give an example on how this would be done: \\
+
+Let $\varphi(\x,\X,\z) = \exists x_1\exists x_2( X_1(x_2) \wedge ( X_2(z_1,x_1) \wedge \forall y( S_1(y,z_1) )  ))$.
+In this case, $\x = (x_1,x_2)$, $\X = (X_1,X_2)$ and $\z = (z_1)$.
+Let $\A = \langle A,S^\A,\leq^\A \rangle$, where $A = \{1,2\}$, $S^\A = \{(1,1),(1,2),(2,2)\}$, and $1 \leq^\A 2$.
+We generate all the possible assignments for $\x$ and $\z$.
+And our new formula is:
+\begin{multline*}
+\varphi_\A(\X) = (X_1(1) \wedge ( X_2(1,1) \wedge \forall y( S_1(y,1) )  ) \vee (X_1(1) \wedge ( X_2(2,1) \wedge \forall y( S_1(y,2) )  )\; \vee \\
+(X_1(2) \wedge ( X_2(1,1) \wedge \forall y( S_1(y,1) )  ) \vee (X_1(2) \wedge ( X_2(2,1) \wedge \forall y( S_1(y,2) )  )\; \vee \\
+(X_1(1) \wedge ( X_2(1,2) \wedge \forall y( S_1(y,1) )  ) \vee (X_1(1) \wedge ( X_2(2,2) \wedge \forall y( S_1(y,2) )  )\; \vee \\
+(X_1(2) \wedge ( X_2(1,2) \wedge \forall y( S_1(y,1) )  ) \vee (X_1(2) \wedge ( X_2(2,2) \wedge \forall y( S_1(y,2) )  ).
+\end{multline*}
+Now we replace the satisfied subformulas:
+\begin{multline*}
+\psi_\A(\X) = (X_1(1) \wedge X_2(1,1) \wedge \perp  ) \vee (X_1(1) \wedge  X_2(2,1) \wedge \top  ) \;\vee \\
+(X_1(2) \wedge  X_2(1,1) \wedge \perp  ) \vee (X_1(2) \wedge  X_2(2,1) \wedge \top  ) \;\vee \\
+(X_1(1) \wedge  X_2(1,2) \wedge \perp  ) \vee (X_1(1) \wedge  X_2(2,2) \wedge \top  ) \;\vee \\
+(X_1(2) \wedge  X_2(1,2) \wedge \perp  ) \vee (X_1(2) \wedge  X_2(2,2) \wedge \top  ),
+\end{multline*}
+which can be simplified even further:
+\begin{multline*}
+\psi_\A(\X) = (X_1(1) \wedge X_2(2,1)  ) \vee (X_1(2) \wedge  X_2(2,1) ) \;\vee \\ 
+(X_1(1) \wedge  X_2(2,2)  ) \vee (X_1(2) \wedge  X_2(2,2)  )
+\end{multline*}
+Note that we can do this since checking $\A\models\varphi(\d_i,\X,\e_j)$ can be done in polynomial time for each $i,j$, as $\varphi$ is assumed to be fixed.
+Note as well that $\psi_\A(\X)$ is quantifier-free and its size is also polynomial to the size of $\A$.
+However, this still is not a valid $\L-$formula since there are no constants in $\L$.
+Let $P_a$ be a unary predicate for every $a\in A$.
+Now let $\zeta_\A(\X,\u)$ be obtained from $\psi_\A(\X)$ by adding the conjunct $P_{a_1}(u_{a_1}) \wedge \cdots \wedge P_{a_{\vert A \vert}}(u_{a_{\vert A \vert}})$, where $u_a$ is a first order variable, and by replacing every constant $a\in A$ by $u_a$.
+In the example:
+\begin{multline*}
+\zeta_\A(\X,\u) = \big((X_1(u_1) \wedge X_2(u_2,u_1)  ) \vee (X_1(u_2) \wedge  X_2(u_2,u_1) )\; \vee \\ 
+(X_1(u_1) \wedge  X_2(u_2,u_2)  ) \vee (X_1(u_2) \wedge  X_2(u_2,u_2)  )\big) \wedge P_1(u_1) \wedge P_2(u_2),
+\end{multline*}
+where $\u = (u_1,u_2)$.
+Now, let $\L^\prime = \L \cup \{P_a\mid a\in A\}$ and $\A^\prime = \langle A,S^\A, \P^{\A^\prime}_{a_1},\ldots,\P^{\A^\prime}_{a_{\vert A \vert}}, \leq^\A \rangle$, where $\P^{\A^\prime}_a = \{a\}$ for each $a\in A$.
+Moreover, let $f^\prime$ be defined as follows:
 \begin{eqnarray*}
-f'(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \psi_\A(\P) \} \mid.
+f^\prime(\A^\prime) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \zeta_\A(\P,\e) \} \mid.
 \end{eqnarray*}
-Clearly $f' \in\#\Sigma_0$.
-Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
-However, it is showed in \cite{DBLP:journals/jcss/SalujaST95} that any counting problem in $\#\Sigma_0$ is computable in polynomial time.
-Therefore, the decision version of $f$ is in {\sc P}.
+Clearly $f^\prime \in\#\Sigma_0$.
+Also, note that for every $\A$, $f(\A) > 0$ if and only if $f^\prime(\A^\prime) > 0$, so computing $f^\prime$ with the instance $\A^\prime$ is enough to solve the decision version of $f$.
+Therefore, given that any counting problem in $\#\Sigma_0$ is computable in polynomial time\cite{DBLP:journals/jcss/SalujaST95}, we conclude that the decision version of $f$ is in {\sc P}.
 \end{proof}

### 2015-05-12 tex/th2.tex
12 may

--- 2015-04-22 tex/th2.tex
@@
 \begin{proof}
-Let $f_{\#3DNF}$ be a function that counts the satisfying assignments to a 3DNF formula.
-As shown in [1], $f_{\#3DNF} \in \#\Sigma_1$.
-Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Suppose that $f_{2^n}-f_{\#3DNF} \in \#\Sigma_1$.
-This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete.
-However, as we showed previously in Theorem \ref{decisionptime}, it is also in P.
-Then, co-NP $\subseteq$ P.
+Suppose that $\#\Sigma_1$ is closed under substraction, that is, for each pair of functions $f,g\in\#\Sigma_1,$ there is an $h\in\#\Sigma_1$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\Truc$.
+
+Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\psi$, where $A$ is the set of variables mentioned in $\psi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{1,2,3,4\}$:
+\begin{eqnarray*}
+S_1^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
+S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
+S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
+S_4^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\psi\}.
+\end{eqnarray*}
+Now let $f_{\#3DNF}$ be a function that counts the satisfying assignments to a 3DNF formula $\psi$. As shown in [1], $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
+$$
+\varphi(x,X) = (X(x) \vee \neg X(x)).
+$$
+Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
+As we supposed initially, let $h\in\#\Sigma_1$ such that $f_{all}-f_{\#3DNF} = h$.
+For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is {\sc co-NP}-complete.
+However, as we showed previously in Theorem \ref{decisionptime}, $L_h \in \textsc{P}$.
+Then, $\textsc{co-NP} \subseteq \textsc{P}$, from which we conclude that $\textsc{P} = \textsc{NP}$.
 \end{proof}

### 2015-05-12 tex/th3.tex
12 may

--- 2015-04-08 tex/th3.tex
@@
 Now let $\B = \langle B =\{1,2\},S^\B = \{1\},\leq^\B=\{(1,1),(1,2),(2,2)\} \rangle$.
-Since $\#\Sigma_1$ is monotonic under addition of elements, $f(\B) \geq 1$.
+Note that $\A$ is an induced substructure of $\B$.
+
+We have that for each function $g \in \#\Sigma_1$ and structures $\A_1,\A_2\in\Truc$, if $\A_1$ is an induced substructure of $\A_2$, then $g(\A_1) \leq g(\A_2)$ \cite{DBLP:journals/jcss/SalujaST95}.
+Therefore, $f(\B) \geq f(\A) = 1$.
 However, there is no assignment $s\in B$ to $x$ such that $\B\models\:\varphi(s)$, so $f(\B) = 0$, which leads to a contradiction.

### 2015-06-03 tex/th1.tex
2 jun

--- 2015-04-29 tex/th1.tex
@@
 \end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables.
-Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible assignments for $\z$.
-Let $\ell = \vert A \vert^k$ and $\d_1,\dots,\d_\ell\in A^k$ be all possible assignments for $\x$.
-Note that $m$ and $k$ are fixed values, so both $\ell = \vert A \vert^k$ and $n=\vert A \vert^m$ are polynomial in the size of $\A$.
-Let $\varphi_\A(\X)$ be defined as follows:
-\begin{eqnarray*}
-\varphi_\A(\X) = \varphi(\d_1,\X,\e_1) \vee \cdots \vee \varphi(\d_1,\X,\e_n) \vee \varphi(\d_2,\X,\e_1) \vee \cdots \vee \varphi(\d_n,\X,\e_n).
-\end{eqnarray*}
-Note that $\varphi_\A(\X)$ will have at least one assignment $\P$ for $\X$ such that $\A\models\varphi_\A(\P)$ iff $f(\A)>0$.
-Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction.
-We can do this since checking $\A\models\varphi(\d_i,\X,\e_i)$ can be done in polynomial time for each $i\in\{1,\ldots,n\}$.
-Note as well that $\varphi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$.
-Then, let $f'$ be defined as follows:
-\begin{eqnarray*}
-f'(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \psi_\A(\P) \} \mid.
-\end{eqnarray*}
-Clearly $f' \in\#\Sigma_0$.
-Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
-However, it is showed in \cite{DBLP:journals/jcss/SalujaST95} that any counting problem in $\#\Sigma_0$ is computable in polynomial time.
-Therefore, the decision version of $f$ is in {\sc P}.
+where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables.
+As said in \cite{DBLP:journals/jcss/SalujaST95}, a function $h:\Truc \to \N$ is in $\#\Sigma_0$ if there exists a quantifier-free $\L$-formula $\theta(\Y,\y)$ such that $h = f_{\theta(\Y,\y)}$.
+Let $\y = (\x,\z)$ and let $\psi(\X,\y) = \varphi(\x,\X,\z)$.
+Moreover, let $g = f_{\psi(\X,\y)}$.
+\begin{claim}
+For each $\A \in \Truc$, $f(\A) > 0$ iff $g(\A) > 0$.
+\end{claim}
+\begin{proof}
+($\Rightarrow$) Suppose $f(\A) > 0$.
+Let $\P$ and $\e$ be such that $\A \models \exists \x \, \varphi(\x,\P,\e)$.
+It follows that there is at least one $\d \in A^k$ such that $\A \models \varphi(\d,\P,\e) = \psi(\P,(\d,\e))$.
+Therefore, $g(\A) > 0$.
+($\Leftarrow$) Suppose $g(\A) > 0$.
+Let $\Q$ and $\c = (\c_1,\c_2)$, where $\c_1$ and $\c_2$ have $k$ and $m$ elements respectively, be such that $\A \models \psi(\Q,\c) = \varphi(\c_1,\Q,\c_2)$.
+Therefore, $\A \models \exists \x \, \varphi(\x,\Q,\c_2)$, from which we conclude that $f(\A) > 0$.
 \end{proof}
+Clearly $g \in\#\Sigma_0$.
+As we showed, computing $g$ is enough to solve the decision version of $f$.
+Therefore, given that any counting problem in $\#\Sigma_0$ is computable in polynomial time \cite{DBLP:journals/jcss/SalujaST95}, we conclude that the decision version of $f$ is in {\sc P}.
+\end{proof}

### 2015-06-03 tex/th2.tex
2 jun

@@
 \begin{theo}
-If $\#\Sigma_1$ is closed under substraction, then {\sc P} = {\sc NP}.
+If $\E1$ is closed under substraction, then {\sc P} = {\sc NP}.
 \end{theo}
 \begin{proof}
-Suppose that $\#\Sigma_1$ is closed under substraction, that is, for each pair of functions $f,g\in\#\Sigma_1,$ there is an $h\in\#\Sigma_1$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\Truc$.
+Suppose that $\E1$ is closed under substraction, that is, for each pair of functions $f,g\in \E1$, there is an $h\in\E1$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\Truc$.
 
@@
 Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-As we supposed initially, let $h\in\#\Sigma_1$ such that $f_{all}-f_{\#3DNF} = h$.
+As we supposed initially, let $h\in\E1$ such that $f_{all}-f_{\#3DNF} = h$.
 For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is {\sc co-NP}-complete.

### 2015-06-10 tex/sharp-sigma-one-plus.tex
2 jun
3 jun
10jun

@@
 
+\def\Eo{\#\Sigma_0[\mbox{{\sc FO}}]}
 \def\E1{\#\Sigma_1[\mbox{{\sc FO}}]}
@@
 \end{eqnarray*}
-where $x,y$ and $x_i$ are first order variables for every $i$. We now define an extended quantifier-free $\L$-formula as follows:
+where $x,y$ and $x_i$ are first order variables for every $i$. We now define an {\sc FO}-extended quantifier-free $\L$-formula as follows:
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements from $\A$.
+for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements from $\A$.\\
 
-A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\exists \x \: \varphi(\x,\Y,\y)}.$\\
+A function $f:\Truc \to \N$ is in $\Eo$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\Y,\y)$ such that $f = f_{\varphi(\Y,\y)}.$
+
+Similarly, a function $f:\Truc \to \N$ is in $\E1$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\exists \x \: \varphi(\x,\Y,\y)}.$\\
+
+\input{th0}
 

### 2015-06-10 tex/th0.tex
10jun

\begin{theo}
$\Eo \subseteq$ {\sc FP}.
\end{theo}
\begin{proof}
Let $f \in \Eo$, and let $\varphi(\X,\x)$ be and {\sc FO}-extended quantifier-free $\L$-formula such that:
\begin{eqnarray*}
f(\A) &=& \vert \{\langle \P,\e  \rangle \mid \A \models \varphi(\P,\e) \} \vert
\end{eqnarray*}
for each $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\e \in A^m$ and $\P = (P_1,\ldots,P_q)$ is a tuple of predicates. We will now show that counting $f(\A)$ can be done in polynomial time.

For each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$, let $R_{\beta}$ be a predicate of arity $m$. Let $\A^\prime = \langle A, \S^{\A}, R_{\beta}^{\A^\prime}, \leq^{\A} \rangle \in \Truc$, where $R_{\beta}^{\A^\prime} = \{\d \mid \A \models \beta(\d)\}$. Note that each $\beta(\x)$ is fixed in $\varphi(\X,\x)$, and for each $\d \in A^m$, checking whether $\A \models \beta(\d)$ can be done in polynomial time. Therefore, generating $R_{\beta}^{\A^\prime}$ can also be done in polynomial time.

Let $\psi(\X,\x)$ be obtained by replacing each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$ by $R_{\beta}(\x)$. Also, let $g = f_{\psi(\X,\x)}$. Note that for each tuple of predicates $\P$ and each $\e \in A^m$, $\A \models \varphi(\P,\e)$ if and only if $\A^\prime \models \psi(\P,\e)$, and so, $g(\A^\prime) = f(\A)$. However, $\psi(\X,\x)$ is a quantifier-free $\L$-formula, and therefore, $g \in \#\Sigma_0$. Since it is shown in \cite{DBLP:journals/jcss/SalujaST95} that $\#\Sigma_0 \subseteq$ {\sc FP}, we conclude that $f(\A)$ can be evaluated in polynomial time.

\end{proof}

### 2015-06-10 tex/th0not.tex
10jun

\begin{theo}
$\Eo \subseteq$ {\sc FP}.
\end{theo}
\begin{proof}
Let $f \in \Eo$, and let $\varphi(\X,\x)$ be and {\sc FO}-extended quantifier-free $\L$-formula such that:
\begin{eqnarray*}
f(\A) &=& \vert \{\langle \P,\e  \rangle \mid \A \models \varphi(\P,\e) \} \vert
\end{eqnarray*}
for each $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\z$ is an $m$-tuple of variables and $\P = (P_1,\ldots,P_q)$ is a tuple of predicates. Each $P_i$ has arity $c_i$, for $i \in \{1,\ldots,k\}$. We will now show that counting $f(\A)$ can be done in polynomial time.

Let $\psi_{\d}(\X) = \varphi(\X,\d)$ and $f_\A(\d) = \vert \{ \P \mid \A \models \psi_{\d}(\P) \} \vert$ for each $\d \in A^m$. Note that $f(\A) = \sum\limits_{\d\in A^m}f_\A(\d)$.

Now, let $\theta_{\d}(\X)$ be the following formula: For each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$ we check if $\A \models \beta(\d)$. If it is true, then we replace $\beta(\d)$ in $\psi_{\d}(\X)$ with a tautology, and otherwise, with a contradiction. Since checking if $\A \models \beta(\d)$ can be done in polynomial time, generating $\theta_{\d}(\X)$ can also be done in polynomial time. Note that for each tuple of predicates $\P$, $\A\models\theta_{\d}(\P)$ if and only if $\A\models\psi_{\d}(\P)$, so $f_\A(\d) = \vert \{ \P \mid \A\models \theta_{\d}(\P) \} \vert$. Also note that $\theta_{\d}(\X)$ has no {\sc FO}-formulas and therefore, counting $\vert \{ \P \mid \A \models \theta_{\d}(\P) \} \vert$ is as hard as counting some function in $\#\Sigma_0$. We will give a more rigorous proof of this statement.

Let $R_{\d}$ be a predicate of arity $m$, and let $\A^\prime = \langle A, \S^{\A}, R_{\d}^{\A^\prime}, \leq^{\A} \rangle \in \Truc$, where $R_{\d}^{\A^\prime} = \{\d\}$. Now let $\gamma_{\d}(\X,\x)$ be obtained by replacing each constant of $\d$ in $\theta_{\d}(\X)$ by its respective variable in $\x$. Let $\zeta(\X,\x) = R_{\d}(\x) \wedge \gamma_{\d}(\X,\x)$ and $g_{\d} = f_{\zeta(\X,\x)}$. Clearly, $g_{\d}(\A^\prime) = f_\A(\d)$. Note that $\zeta(\X,\x)$ is a quantifier-free formula and therefore, $g_{\d} \in \#\Sigma_0$.

In \cite{DBLP:journals/jcss/SalujaST95} it is shown that $\#\Sigma_0 \subseteq$ {\sc FP}, so counting $g_{\d}(\A^\prime)$ can be done in polynomial time. Therefore, counting $f(\A) = \sum\limits_{\d\in A^m}g_{\d}(\A^\prime)$ can also be done in polynomial time.
\end{proof}

### 2015-06-10 tex/th1.tex
10jun

@@
 \end{proof}
-Clearly $g \in\#\Sigma_0$.
-As we showed, computing $g$ is enough to solve the decision version of $f$.
-Therefore, given that any counting problem in $\#\Sigma_0$ is computable in polynomial time \cite{DBLP:journals/jcss/SalujaST95}, we conclude that the decision version of $f$ is in {\sc P}.
+Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula, so $g \in\Eo$.
+As we showed, computing $g$ is enough to solve the decision version of $f$, but as we showed in Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time, from which we conclude that the decision version of $f$ is in {\sc P}.
 \end{proof}

### 2015-06-10 tex/th1.tex
10jun

@@
 Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula, so $g \in\Eo$.
-As we showed, computing $g$ is enough to solve the decision version of $f$, but as we showed in Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time, from which we conclude that the decision version of $f$ is in {\sc P}.
+As we showed, computing $g$ is enough to solve the decision version of $f$, but as it is shown in Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time, from which we conclude that the decision version of $f$ is in {\sc P}.
 \end{proof}

### 2015-06-10 tex/th1.tex
10jun

--- 2015-06-10 tex/th1.tex
@@
 Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula, so $g \in\Eo$.
-As we showed, computing $g$ is enough to solve the decision version of $f$, but as we showed in Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time, from which we conclude that the decision version of $f$ is in {\sc P}.
+As we showed, for each $\A \in \Truc$, it holds that $\A \in L_f$ if and only if $g(\A) > 0$, but as it is shown in Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time.
+From this we conclude that the decision version of $f$ is in {\sc P}.
 \end{proof}

### 2015-06-24 examples.tex
examples added

 \documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm} 
\usepackage{fullpage}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bm}

\def\dotminus{\mathbin{\ooalign{\hss\raise1ex\hbox{.}\hss\cr
  \mathsurround=0pt$-$}}}

\def\Eone{\#\Sigma_1\textsc{[FO]}}
\def\Etwo{\#\Pi_1\textsc{[FO]-HORN}}
\def\Ethree{\#\Sigma_2\textsc{[FO]-HORN}}
\def\Efour{\#\Pi_2\textsc{[FO]-HORN}}

\def\Truc{\textsc{Struct}[\L]}

\def\A{{\frak A}}
\def\C{{\cal C}}
\def\F{{\cal F}}
\def\L{{\cal L}}
\def\N{\mathbb{N}}
\def\P{\vec{P}}
\def\R{\vec{R}}
\def\S{\vec{S}}
\def\X{\vec{X}}
\def\Y{\vec{Y}}
\def\Z{\vec{Z}}
%% a - arity of \X / arity of assignments \P to \X
%% b - arity of predicates in \S
%% c - arity of auxiliar predicates/variables
\def\c{\vec{c}} %% super auxiliar elements
\def\d{\vec{d}} %% counted elements
\def\e{\vec{e}} %% counted elements
%% f - counting function
%% g - other functions
%% h - other functions
%% i - index
%% j - index
%% k - emergency index / size of tuple
%% l - emergency index / size of tuple
\def\l{\vec{\ell}}
%% m - size of variable tuple
%% n - size of predicate tuple
%% o - not used
%% p - 
%% q - 
%% r - size of \X / \P
\def\s{\vec{s}}
%% t - size of \S
\def\t{\vec{t}}
\def\u{\vec{u}} %% auxiliary variables
\def\v{\vec{v}} %% auxiliary variables
\def\w{\vec{w}} %% auxiliary variables
\def\x{\vec{x}} %% quantified variables
\def\y{\vec{y}} %% auxiliary variables
\def\z{\vec{z}} %% open variables
\def\ep{\vec{o}}
\def\ga{\vec{p}}

\newtheorem{theo}{Theorem}
\newtheorem{lemma}[theo]{Lemma}
\newtheorem{claim}[theo]{Claim}
\newtheorem{coro}[theo]{Corollary}

\title{Some examples in extended logic counting classes}

\begin{document}
\maketitle

\section{{\sc \#3-DNF}}
A function $f:\Truc \to \N$ is in $\Eone$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\Y,\y)$ such that $f = f_{\varphi(\Y,\y)}$. We will now show that {\sc \#3-DNF} $\in\Eone$.\\

Let $\L = \{S_0,S_1,S_2,S_3,\leq\}$, and $\A = \langle A, S_0^\A, S_1^\A, S_2^\A, S_3^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc 3-DNF} formula $\psi$, where $A$ is the set of variables mentioned in $\psi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{0,1,2,3\}$:
\begin{eqnarray*}
S_0^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
S_1^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\psi\}.
\end{eqnarray*}

Now let $T$ be a unary second order variable, and
\begin{align*}
\varphi(T) =\ &\exists x \exists y \exists z \,(S_0(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z))\,\vee \\
&\exists x \exists y \exists z \,(S_1(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z))\,\vee \\
&\exists x \exists y \exists z \,(S_2(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z))\,\vee \\
&\exists x \exists y \exists z \,(S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z)).
\end{align*}
It is easily seen that each relation $T$ over $\A$ such that $\A\models\varphi(T)$ represents a valid assignment to the variables in $\psi$, and therefore, $f_{\varphi(T)}$ represents {\sc \#3-DNF}.

\section{{\sc \#3-HORN-SAT}}
A function $f:\Truc \to \N$ is in $\Etwo$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\forall \x\,\varphi(\x,\Y,\y)}$ and any second order variables only appear in Horn clauses. We will now show that {\sc \#3-HORN-SAT} $\in\Etwo$.\\

Let $\L = \{P_1,P_2,P_3,N_1,N_2,N_3,\leq\}$, and $\A = \langle A, P_1^\A,P_2^\A,P_3^\A,N_1^\A,N_2^\A,N_3^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc 3-HORN-SAT} formula $\psi$, where $A$ is the set of variables mentioned in $\psi$, and the relations in $\L$ are interpreted as follows:
\begin{eqnarray*}
P_1^\A &=& \{a\mid a \mbox{ appears as a conjunct in }\psi\},\\
P_2^\A &=& \{(a_1,a_2)\mid (a_1\vee \neg a_2) \mbox{ appears as a conjunct in }\psi\},\\
P_3^\A &=& \{(a_1,a_2,a_3)\mid (a_1\vee \neg a_2\vee \neg a_3) \mbox{ appears as a conjunct in }\psi\},\\
N_1^\A &=& \{a\mid (\neg a) \mbox{ appears as a conjunct in }\psi\},\\
N_2^\A &=& \{(a_1,a_2)\mid (\neg a_1\vee \neg a_2) \mbox{ appears as a conjunct in }\psi\},\\
N_3^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1\vee \neg a_2\vee \neg a_3) \mbox{ appears as a conjunct in }\psi\}.
\end{eqnarray*}

Now let $T$ be a unary second order variable, and
\begin{align*}
\varphi(T) =\ &\forall x\, (\neg P_1(x) \vee T(x))\,\wedge\\
&\forall x\forall y\, (\neg P_2(x,y) \vee T(x) \vee \neg T(y))\,\wedge\\
&\forall x\forall y\forall z\, (\neg P_3(x,y,z) \vee T(x) \vee \neg T(y) \vee \neg T(z))\,\wedge\\
&\forall x\, (\neg N_1(x) \vee \neg T(x))\,\wedge\\
&\forall x\forall y\, (\neg N_2(x,y) \vee \neg T(x) \vee \neg T(y))\,\wedge\\
&\forall x\forall y\forall z\, (\neg N_3(x,y,z) \vee \neg T(x) \vee \neg T(y) \vee \neg T(z))\,\wedge
\end{align*}

We can easily reorder $\varphi(T)$ as $\forall x \forall y \forall z\,\theta(x,y,z,T)$ where $\theta(x,y,z,T)$ is an {\sc FO}-extended quantifier-free $\L$-formula such that any second order variable only appears in a Horn clause. It is easily seen that each relation $T$ over $\A$ such that $\A\models\forall x \forall y \forall z\,\theta(x,y,z,T)$ represents a valid assignment to the variables in $\psi$, and therefore, $f_{\forall x \forall y \forall z\,\theta(x,y,z,T)}$ represents {\sc \#3-HORN-SAT}.

\section{{\sc \#DNF}}
A function $f:\Truc \to \N$ is in $\Ethree$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x^1,\x^2,\Y,\y)$ such that $f = f_{\exists\x^1\forall \x^2\,\varphi(\x^1,\x^2,\Y,\y)}$ and any second order variables only appear in Horn clauses. We will now show that {\sc \#DNF} $\in\Ethree$.

Let $\L = \{P,N,\leq\}$, and $\A = \langle A, P^\A, N^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc DNF} formula $\psi$, where $A$ is the set of variables mentioned in $\psi$, $P$ and $N$ is a binary relations described as follows:
\begin{eqnarray*}
P^\A &=& \{(c,\ell)\mid \ell \mbox{ appears in the $c$-th clause in }\psi\},\\
N^\A &=& \{(c,\ell)\mid \neg \ell \mbox{ appears in the $c$-th clause in }\psi\}.
\end{eqnarray*}

Now let $T$ be a unary second order variable, and
\begin{eqnarray*}
\varphi(T) &=& \exists c \,\Big(\forall \ell_1 (\neg P(c,\ell_1) \vee T(\ell_1)) \wedge \forall \ell_2 (\neg N(c,\ell_2) \vee \neg T(\ell_2))\Big). 
\end{eqnarray*}

We reorder $\varphi(T)$ as $\exists c \forall \ell_1 \forall \ell_2\, \theta(c,\ell_1,\ell_2,T)$ where $\theta(c,\ell_1,\ell_2,T)$ is an {\sc FO}-extended quantifier-free $\L$-formula such that any second order variable only appears in a Horn clause. It is easily seen that each relation $T$ over $\A$ such that $\A\models \exists c \forall \ell_1 \forall \ell_2\, \theta(c,\ell_1,\ell_2,T)$ represents a valid assignment to the variables in $\psi$, and therefore, $f_{\exists c \forall \ell_1 \forall \ell_2\, \theta(c,\ell_1,\ell_2,T)}$ represents {\sc \#DNF}.

\end{document}

### 2015-07-21 tex/case2generalproof.tex
Correcciones menores y comentarios.

@@
 Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Assume that $\A\in\Truc$.
-Suppose first that there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least one assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\d = (\d_1,\dots,\d_{n_q},\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha_q(\d)$, where $\d_i$ is the corresponding assignment to $\x_i$, for every $i\in\{1,\dots,n_q\}$, and $\ep$ is an assignment for $\y$.
+Suppose first that there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least index such that there exists an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_q(\x,\R^\prime).$ Let $\d = (\d_1,\dots,\d_{n_q},\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha_q(\d)$, where $\d_i$ is the corresponding assignment to $\x_i$, for every $i\in\{1,\dots,n_q\}$, and $\ep$ is an assignment for $\y$.
 Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j:\lambda_q(j)=i,j\in[1,n_q]}\{\d_j\}$.
@@
 But as we mentioned, also $\A\models\:\psi_q(\P^\prime),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which leads to a contradiction.
+\icristian{
+Al igual que un argumento anterior, esta demostración es mas clara si es directa y NO por contradicción.	
+}
 \end{enumerate}
@@
 
-Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$.
+Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for every $i\in\{1,\ldots,k\}$.
 Let $\P$ be an arbitrary assignment to $\X$.
@@
 It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Hence, from the results in the previous paragraph, if $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}$, we have that $f^\prime = f\dotminus 1\in\E1$
+And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Hence, from the results in the previous paragraph, if $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}$, we have that $f^\prime = f\dotminus 1\in\E1$.
+
+\icristian{
+Importante! Aqui falta argumentar porque la formula final esta en $\E1$.
+}

### 2015-07-21 tex/th1.tex
Correcciones menores y comentarios.

@@
 \begin{proof}
-Let $f$ be a function in $\E1$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
+Let $f$ be a function in $\E1$. Then there exists an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ where $\x$ is a $k$-tuple of variables and $\z$ is an $m$-tuple of variables such that
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables.
+for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$ and $A = \{a_1,\ldots,a_{\vert A \vert}\}$.
 As said in \cite{DBLP:journals/jcss/SalujaST95}, a function $h:\Truc \to \N$ is in $\#\Sigma_0$ if there exists a quantifier-free $\L$-formula $\theta(\Y,\y)$ such that $h = f_{\theta(\Y,\y)}$.
@@
 ($\Leftarrow$) Suppose $g(\A) > 0$.
-Let $\Q$ and $\c = (\c_1,\c_2)$, where $\c_1$ and $\c_2$ have $k$ and $m$ elements respectively, be such that $\A \models \psi(\Q,\c) = \varphi(\c_1,\Q,\c_2)$.
+Let $\Q$ and $\c = (\c_1,\c_2)$, where $\c_1$ and $\c_2$ have $k$ and $m$ elements respectively, such that $\A \models \psi(\Q,\c) = \varphi(\c_1,\Q,\c_2)$.
 Therefore, $\A \models \exists \x \, \varphi(\x,\Q,\c_2)$, from which we conclude that $f(\A) > 0$.
 \end{proof}
-Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula, so $g \in\Eo$.
+Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula and $g \in\Eo$.
 As we showed, for each $\A \in \Truc$, it holds that $\A \in L_f$ if and only if $g(\A) > 0$, but as it is shown in Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time.

### 2015-07-21 tex/th4.tex
Correcciones menores y comentarios.

@@
 \begin{proof}
-Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$.
-Without loss of generality, let $\z = (z_1,\dots,z_d)$ and let $\X = (X_1,\dots,X_r)$.
+Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$.
 There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$.

### 2015-08-04 tex/examples.tex
1 ago

--- 2015-06-24 examples.tex
@@
- \documentclass[12pt]{article}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsthm} 
-\usepackage{fullpage}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{bm}
+\section{Some examples in extended logic counting classes}
 
-\def\dotminus{\mathbin{\ooalign{\hss\raise1ex\hbox{.}\hss\cr
-  \mathsurround=0pt$-$}}}
-
-\def\Eone{\#\Sigma_1\textsc{[FO]}}
-\def\Etwo{\#\Pi_1\textsc{[FO]-HORN}}
-\def\Ethree{\#\Sigma_2\textsc{[FO]-HORN}}
-\def\Efour{\#\Pi_2\textsc{[FO]-HORN}}
-
-\def\Truc{\textsc{Struct}[\L]}
-
-\def\A{{\frak A}}
-\def\C{{\cal C}}
-\def\F{{\cal F}}
-\def\L{{\cal L}}
-\def\N{\mathbb{N}}
-\def\P{\vec{P}}
-\def\R{\vec{R}}
-\def\S{\vec{S}}
-\def\X{\vec{X}}
-\def\Y{\vec{Y}}
-\def\Z{\vec{Z}}
-%% a - arity of \X / arity of assignments \P to \X
-%% b - arity of predicates in \S
-%% c - arity of auxiliar predicates/variables
-\def\c{\vec{c}} %% super auxiliar elements
-\def\d{\vec{d}} %% counted elements
-\def\e{\vec{e}} %% counted elements
-%% f - counting function
-%% g - other functions
-%% h - other functions
-%% i - index
-%% j - index
-%% k - emergency index / size of tuple
-%% l - emergency index / size of tuple
-\def\l{\vec{\ell}}
-%% m - size of variable tuple
-%% n - size of predicate tuple
-%% o - not used
-%% p - 
-%% q - 
-%% r - size of \X / \P
-\def\s{\vec{s}}
-%% t - size of \S
-\def\t{\vec{t}}
-\def\u{\vec{u}} %% auxiliary variables
-\def\v{\vec{v}} %% auxiliary variables
-\def\w{\vec{w}} %% auxiliary variables
-\def\x{\vec{x}} %% quantified variables
-\def\y{\vec{y}} %% auxiliary variables
-\def\z{\vec{z}} %% open variables
-\def\ep{\vec{o}}
-\def\ga{\vec{p}}
-
-\newtheorem{theo}{Theorem}
-\newtheorem{lemma}[theo]{Lemma}
-\newtheorem{claim}[theo]{Claim}
-\newtheorem{coro}[theo]{Corollary}
-
-\title{Some examples in extended logic counting classes}
-
-\begin{document}
-\maketitle
-
-\section{{\sc \#3-DNF}}
+\subsection{{\sc \#3-DNF}}
 A function $f:\Truc \to \N$ is in $\Eone$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\Y,\y)$ such that $f = f_{\varphi(\Y,\y)}$.
@@
 
-\section{{\sc \#3-HORN-SAT}}
+\subsection{{\sc \#3-HORN-SAT}}
 A function $f:\Truc \to \N$ is in $\Etwo$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\forall \x\,\varphi(\x,\Y,\y)}$ and any second order variables only appear in Horn clauses.
@@
 
-\section{{\sc \#DNF}}
+\subsection{{\sc \#DNF}}
 A function $f:\Truc \to \N$ is in $\Ethree$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x^1,\x^2,\Y,\y)$ such that $f = f_{\exists\x^1\forall \x^2\,\varphi(\x^1,\x^2,\Y,\y)}$ and any second order variables only appear in Horn clauses.
@@
 It is easily seen that each relation $T$ over $\A$ such that $\A\models \exists c \forall \ell_1 \forall \ell_2\, \theta(c,\ell_1,\ell_2,T)$ represents a valid assignment to the variables in $\psi$, and therefore, $f_{\exists c \forall \ell_1 \forall \ell_2\, \theta(c,\ell_1,\ell_2,T)}$ represents {\sc \#DNF}.
-
-\end{document}

### 2015-08-04 tex/s-one-fo.tex
1 ago

\section{Some properties of $\E1$}

We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
\begin{eqnarray*}
	\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
	&& \leq \text{ is interpreted as a total order for } A \}.
\end{eqnarray*}
We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$. A quantifier-free $\L$-formula is defined by the following grammar:
\begin{eqnarray*}
	\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
	&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
	&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
\end{eqnarray*}
where $x,y$ and $x_i$ are first order variables for every $i$. We now define an {\sc FO}-extended quantifier-free $\L$-formula as follows:
\begin{eqnarray*}
	\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
	&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
	&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
\end{eqnarray*}
Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables. For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
\begin{eqnarray*}
	f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
\end{eqnarray*}
for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements from $\A$.\\

A function $f:\Truc \to \N$ is in $\Eo$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\Y,\y)$ such that $f = f_{\varphi(\Y,\y)}$.

Similarly, a function $f:\Truc \to \N$ is in $\E1$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\exists \x \: \varphi(\x,\Y,\y)}.$\\

\begin{theo} \label{fp1}
	$\Eo \subseteq$ {\sc FP}.
\end{theo}
\begin{proof}
	Let $f \in \Eo$, and let $\varphi(\X,\x)$ be and {\sc FO}-extended quantifier-free $\L$-formula such that:
	\begin{eqnarray*}
		f(\A) &=& \vert \{\langle \P,\e  \rangle \mid \A \models \varphi(\P,\e) \} \vert
	\end{eqnarray*}
	for each $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\e \in A^m$ and $\P = (P_1,\ldots,P_q)$ is a tuple of predicates. We will now show that counting $f(\A)$ can be done in polynomial time.

	For each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$, let $R_{\beta}$ be a predicate of arity $m$. Let $\A^\prime = \langle A, \S^{\A}, R_{\beta}^{\A^\prime}, \leq^{\A} \rangle \in \Truc$, where $R_{\beta}^{\A^\prime} = \{\d \mid \A \models \beta(\d)\}$. Note that each $\beta(\x)$ is fixed in $\varphi(\X,\x)$, and for each $\d \in A^m$, checking whether $\A \models \beta(\d)$ can be done in polynomial time. Therefore, generating $R_{\beta}^{\A^\prime}$ can also be done in polynomial time.

	Let $\psi(\X,\x)$ be obtained by replacing each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$ by $R_{\beta}(\x)$. Also, let $g = f_{\psi(\X,\x)}$. Note that for each tuple of predicates $\P$ and each $\e \in A^m$, $\A \models \varphi(\P,\e)$ if and only if $\A^\prime \models \psi(\P,\e)$, and so, $g(\A^\prime) = f(\A)$. However, $\psi(\X,\x)$ is a quantifier-free $\L$-formula, and therefore, $g \in \#\Sigma_0$. Since it is shown in \cite{DBLP:journals/jcss/SalujaST95} that $\#\Sigma_0 \subseteq$ {\sc FP}, we conclude that $f(\A)$ can be evaluated in polynomial time.

\end{proof}

The {\em decision problem} associated to a function $f$ is defined by the language $L_f = \{\A \in \Truc \mid f(\A) > 0\}$.

\begin{theo} \label{decisionptime}
	The decision problem associated to a function in $\E1$ is in \textsc{P}.
\end{theo}
\begin{proof}
	Let $f$ be a function in $\E1$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
	\begin{eqnarray*}
		f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
	\end{eqnarray*}
	where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables. As said in \cite{DBLP:journals/jcss/SalujaST95}, a function $h:\Truc \to \N$ is in $\#\Sigma_0$ if there exists a quantifier-free $\L$-formula $\theta(\Y,\y)$ such that $h = f_{\theta(\Y,\y)}$. Let $\y = (\x,\z)$ and let $\psi(\X,\y) = \varphi(\x,\X,\z)$. Moreover, let $g = f_{\psi(\X,\y)}$.
	\begin{claim}
		For each $\A \in \Truc$, $f(\A) > 0$ iff $g(\A) > 0$.
	\end{claim}
	\begin{proof}
		($\Rightarrow$) Suppose $f(\A) > 0$. Let $\P$ and $\e$ be such that $\A \models \exists \x \, \varphi(\x,\P,\e)$. It follows that there is at least one $\d \in A^k$ such that $\A \models \varphi(\d,\P,\e) = \psi(\P,(\d,\e))$. Therefore, $g(\A) > 0$. ($\Leftarrow$) Suppose $g(\A) > 0$. Let $\Q$ and $\c = (\c_1,\c_2)$, where $\c_1$ and $\c_2$ have $k$ and $m$ elements respectively, be such that $\A \models \psi(\Q,\c) = \varphi(\c_1,\Q,\c_2)$. Therefore, $\A \models \exists \x \, \varphi(\x,\Q,\c_2)$, from which we conclude that $f(\A) > 0$.
	\end{proof}
	Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula, so $g \in\Eo$. As we showed, for each $\A \in \Truc$, it holds that $\A \in L_f$ if and only if $g(\A) > 0$, but as it is shown in Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time. From this we conclude that the decision version of $f$ is in {\sc P}.
\end{proof}

For a given pair of functions $f,g$, we define $f \dotminus g$ as follows:
\begin{eqnarray*}
	(f \dotminus g)(\A) =
	\begin{cases}
		f(\A)-g(\A), & \text{if }f(\A)>g(\A) \\
		0, & \text{if }f(\A) \leq g(\A).
	\end{cases}
\end{eqnarray*}
for every $\L$-structure $\A \in \Truc$. A function class $\F$ is {\em closed under substraction} if for every pair of functions $f,g \in \F$, it holds that $f \dotminus g \in \F$.

\begin{theo}
	If $\E1$ is closed under substraction, then {\sc P} = {\sc NP}.
\end{theo}
\begin{proof}
	Suppose that $\E1$ is closed under substraction, that is, for each pair of functions $f,g\in \E1$, there is an $h\in\E1$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\Truc$.

	Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\psi$, where $A$ is the set of variables mentioned in $\psi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{1,2,3,4\}$:
	\begin{eqnarray*}
		S_1^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
		S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
		S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
		S_4^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\psi\}.
	\end{eqnarray*}
	Now let $f_{\#3DNF}$ be a function that counts the satisfying assignments to a 3DNF formula $\psi$. As shown in [1], $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
	$$
	\varphi(x,X) = (X(x) \vee \neg X(x)).
	$$
	Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula. As we supposed initially, let $h\in\E1$ such that $f_{all}-f_{\#3DNF} = h$. For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is {\sc co-NP}-complete. However, as we showed previously in Theorem \ref{decisionptime}, $L_h \in \textsc{P}$. Then, $\textsc{co-NP} \subseteq \textsc{P}$, from which we conclude that $\textsc{P} = \textsc{NP}$.
\end{proof}

\begin{theo}
	$\#\Sigma_1 \subsetneq \E1$
\end{theo}
\begin{proof}
	We will show that the $\E1$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$. By contradiction, suppose that it is. Let $\A = \langle A =\{1\},S^\A = \{1\},\leq^\A=\{(1,1)\} \rangle$. Then, $f(\A) = 1$. Now let $\B = \langle B =\{1,2\},S^\B = \{1\},\leq^\B=\{(1,1),(1,2),(2,2)\} \rangle$. Note that $\A$ is an induced substructure of $\B$.

	We have that for each function $g \in \#\Sigma_1$ and structures $\A_1,\A_2\in\Truc$, if $\A_1$ is an induced substructure of $\A_2$, then $g(\A_1) \leq g(\A_2)$ \cite{DBLP:journals/jcss/SalujaST95}. Therefore, $f(\B) \geq f(\A) = 1$. However, there is no assignment $s\in B$ to $x$ such that $\B\models\:\varphi(s)$, so $f(\B) = 0$, which leads to a contradiction.
\end{proof}

For a given function $f$, we define $f \dotminus 1$ as follows:
\begin{eqnarray*}
	f \dotminus 1(\A) =
	\begin{cases}
		f(\A)-1, & \text{if }f(\A) > 0 \\
		0, & \text{if }f(\A) = 0.
	\end{cases}
\end{eqnarray*}
for every $\L$-structure $\A \in \Truc$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.

\input{th4}

### 2015-08-04 tex/th4.tex
1 ago

--- 2015-01-30 sharp-sigma-one-plus.tex
@@
- \documentclass[12pt]{article}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amsthm} 
-\usepackage{fullpage}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{bm}
-
-\def\dotminus{\mathbin{\ooalign{\hss\raise1ex\hbox{.}\hss\cr
-  \mathsurround=0pt$-$}}}
-
-\def\E1{\#\Sigma_1^{+}}
-
-\def\Truc{\textsc{Struct}[\L]}
-
-\def\A{{\frak A}}
-\def\C{{\cal C}}
-\def\F{{\cal F}}
-\def\L{{\cal L}}
-\def\N{\mathbb{N}}
-\def\P{\vec{P}}
-\def\R{\vec{R}}
-\def\S{\vec{S}}
-\def\X{\vec{X}}
-\def\Y{\vec{Y}}
-\def\Z{\vec{Z}}
-%% a - arity of \X / arity of assignments \P to \X
-%% b - arity of predicates in \S
-%% c - arity of auxiliar predicates/variables
-\def\c{\vec{c}} %% super auxiliar elements
-\def\d{\vec{d}} %% counted elements
-\def\e{\vec{e}} %% counted elements
-%% f - counting function
-%% g - other functions
-%% h - other functions
-%% i - index
-%% j - index
-%% k - emergency index / size of tuple
-%% l - emergency index / size of tuple
-\def\l{\vec{\ell}}
-%% m - size of variable tuple
-%% n - size of predicate tuple
-%% o - not used
-%% p - 
-%% q - 
-%% r - size of \X / \P
-\def\s{\vec{s}}
-%% t - size of \S
-\def\t{\vec{t}}
-\def\u{\vec{u}} %% auxiliary variables
-\def\v{\vec{v}} %% auxiliary variables
-\def\w{\vec{w}} %% auxiliary variables
-\def\x{\vec{x}} %% quantified variables
-\def\y{\vec{y}} %% auxiliary variables
-\def\z{\vec{z}} %% open variables
-\def\ep{\vec{o}}
-\def\ga{\vec{p}}
-
-\newtheorem{theo}{Theorem}
-\newtheorem{lemma}[theo]{Lemma}
-\newtheorem{claim}[theo]{Claim}
-\newtheorem{coro}[theo]{Corollary}
-
-\begin{document}
-
-\begin{center}
-{ \LARGE \bf
-  Some properties of $\E1$
-}
-\end{center}
-
-We define the vocabulary $\L = \{ S_1,\dots,S_t, \leq \}$, where $S_1,\dots,S_t$ have arity $b_1,\dots,b_t$. Let
-\begin{eqnarray*}
-\Truc &=& \{\A \mid \A \text{ is an } \L \text{-structure with a finite domain } A \text{ such that} \\
-&& \leq \text{ is interpreted as a total order for } A \}.
-\end{eqnarray*}
-We also define a set of second order variables ${\cal X} = \{ X_i \mid i\in\N \}$ where $X_i$ has arity $a_i$, and for every $n \in \N$ there are infinite variables in ${\cal X}$ of arity $n$.
-A quantifier-free $\L$-formula is defined by the following grammar:
-\begin{eqnarray*}
-\varphi &::=& x = y \ \mid \ S_i(x_1,\dots,x_{b_i}), i \in \{1,\dots,t\} \ \mid \ x \leq y \ \mid \\
-&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \\ 
-&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi),
-\end{eqnarray*}
-where $x,y$ and $x_i$ are first order variables for every $i$. We now define an extended quantifier-free $\L$-formula as follows:
-\begin{eqnarray*}
-\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
-&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
-&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
-\end{eqnarray*}
-Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables.
-For every $\L$-formula $\psi(\Y,\y)$, we define the function $f_{\psi(\Y,\y)}:\Truc \to \N$ as follows:
-\begin{eqnarray*}
-f_{\psi(\Y,\y)}(\A) &=& \mid \{ \langle \P, \e \rangle \mid \A \models \psi(\P,\e) \} \mid,
-\end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_q)$ is a tuple of predicates of arity $c_1,\ldots,c_q$, and $P_i \subseteq A^{c_i}$ for every $i \in \{1,\ldots,q\}$, and $\e$ is a tuple of elements in $\A$.
-
-Let $\Y = (Y_1,\dots,Y_r)$, and let $\y$ be a tuple of variables.
-A function $f:\Truc \to \N$ is in $\E1$ if there exists an extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\exists \x \: \varphi(\x,\Y,\y)}.$\\
-
-The {\em decision version} of a function $f$ is defined by the language $L_f = \{\A \mid f(\A) > 0\}$.
-\begin{theo}
-The decision version of a function in $\E1$ is in \textsc{P}.
-\end{theo}
-\begin{proof}
-Let $f$ be a function in $\E1$. Then there is a formula $\varphi(\x,\X,\z)$ such that
-\begin{eqnarray*}
-f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $\z$ is an $m$-tuple of variables, $\e$ is an $m$-tuple of elements and $\x$ is a $k$-tuple of variables.
-Let $n = \vert A \vert^m$ and $\e_1,\dots,\e_n \in A^m$ be all possible evaluations for $\z$.
-Let $\ell = \vert A \vert^k$ and $\d_1,\dots,d_\ell\in A^k$ be all possible evaluations for $\x$.
-Let $\varphi_\A(\X)$ be defined as follows:
-\begin{eqnarray*}
-\varphi_\A(\X) = \varphi(\d_1,\X,\e_1) \vee \cdots \vee \varphi(\d_1,\X,\e_n) \vee \varphi(\d_2,\X,\e_1) \vee \cdots \vee \varphi(\d_n,\X,\e_n).
-\end{eqnarray*}
-Note that $\varphi_\A(\X)$ will have at least one assignment for $\X$ iff $f(\A)>0$.
-Let $\psi_\A(\X)$ be the formula that results of changing every satisfied sub-formula for a tautology and every non-satisfied sub-formula for a contradiction.
-Note that $\psi_\A(\X)$ is quantifier-free and its size is polynomial to the size of $\varphi(\x,\X,\z)$.
-Then, let $f'$ be defined as follows:
-\begin{eqnarray*}
-f'(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \psi_\A(\P) \} \mid.
-\end{eqnarray*}
-Note that $f' \in\#\Sigma_0$ \footnote[1]{(paper)}.
-Also, note that for every $\A$, $f(\A) > 0$ iff $f'(\A) > 0$, so computing $f'$ is enough to solve the decision version of $f$.
-However, it is showed in $^{[1]}$ that any counting problem in $\#\Sigma_0$ is computable in polynomial time, therefore, the decision version of $f$ is in P.
-\end{proof}
-
-\begin{theo}
-$\#\Sigma_1$ is closed under substraction $\Rightarrow$ P = NP.
-\end{theo}
-\begin{proof}
-\#3DNF $\in \#\Sigma_1$.
-Let $F_{2^n}$ be a $\#\Sigma_1$ function that counts every possible truth assignment in a \#3DNF instance.
-Suppose that $F_{2^n}-F_{\#3DNF} \in \#\Sigma_1$.
-This function equals 0 only if the instanced formula is a tautology, so the decision version of it is co-NP-complete.
-However, as we showed previously (Theorem 1), it is also in P.
-Then, co-NP $\subseteq$ P.
-\end{proof}
-
-\begin{theo}
-$\#\Sigma_1 \subsetneq \E1$
-\end{theo}
-\begin{proof}
-We will show that the $\E1$ function defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
-\end{proof}
-
-For a given function $f$, we define $f \dotminus 1$ as follows:
-\begin{eqnarray*}
-f \dotminus 1(\A) =
-\begin{cases}
-f(\A)-1, & \text{if }f(\A) > 0 \\
-0, & \text{if }f(\A) = 0.
-\end{cases}
-\end{eqnarray*}
-for every $\L$-structure $\A \in \Truc$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
-
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 
-%% F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1 F-1
-
 \begin{theo}
@@
 \begin{enumerate}
-
-% CASE 1
-
 \item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
@@
 Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E1.$
-
-%CASE 2
 
@@
 Considering that (2) and (3) are not possible given that both $\z$ and $\x_i$ are assigned the value $\d_i$ and $\A\models\theta(\d)$, we have that $\theta(\x)\models \z = \x_i$.
-But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$, from which we conclude that $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which contradicts the initial assumption that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
+But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$, which means $\varphi(\x,\X)$ is inconsistent.
+That is, there's no structure $\A^\prime\in\Truc$ such that $\A^\prime\models\exists\x\:\varphi(\x,\R^\prime)$ for any assignment $\R^\prime$ to $\X$.
+In particular, $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which contradicts the initial assumption that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
 		\item[(b)] Note that if $\A \models \alpha_{\min}(\c_1,\ldots,\c_n)$, then necessarily $\c_i = \d_i$ for $i\in\{1,\ldots,n\}.$ However, by the construction of $\P$, we see that 
@@
 \end{proof}
-
 We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
@@
 The function $\lambda_i$ is defined analogously to $\lambda$ of the first part of the proof.
-The tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y_i) = \x.$ The formulas $\gamma^{-}_i(\X,\y_i)$, $\delta_i(\x)$ are defined analogously to $\varphi^{-}(\X,\y)$ and $\beta(\x)$\footnote[2]{Note that $\delta_i(\x)$ can include subformulas of the form $(x_j = x_j)$.} respectively (see \eqref{f1}).
+The tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y_i) = \x.$ The formulas $\gamma^{-}_i(\X,\y_i)$, $\delta_i(\x)$ are defined analogously to $\varphi^{-}(\X,\y)$ and $\beta(\x)\,$\footnote[2]{Note that $\delta_i(\x)$ can include subformulas of the form $(\u = \u)$.} respectively (see \eqref{f1}).
 Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables.
@@
 Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\Truc.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
-We use an injective\footnote[3]{we need this function to be injective because that way we can assure each one of the $k$ satisfiable disjuncts to be represented} function $\rho:\{1,\ldots,k\}\to\{1,\ldots,m\}$ such that $\varphi_i(\x,\X) = \xi_{\rho(i)}(\x,\X)$ is satisfiable, for every $i\in\{1,\ldots,k\}.$
+We use an injective\footnote[3]{We need this function to be injective because this way we can assure each one of the $k$ satisfiable disjuncts to be represented} function $\rho:\{1,\ldots,k\}\to\{1,\ldots,m\}$ such that $\varphi_i(\x,\X) = \xi_{\rho(i)}(\x,\X)$ is satisfiable, for every $i\in\{1,\ldots,k\}.$
 
@@
 
-\begin{claim}
-Let $\A\in\Truc.$ For every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, it holds that $f_{\varphi_i(\s,\X)}(\A) > 0.$
+%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
+%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
+
+\begin{claim} \label{alpha}
+	Let $\A\in\Truc$ and $i\in\{1,\ldots,k\}$. For every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, it holds that $f_{\varphi_i(\s,\X)}(\A) > 0.$
 \end{claim}
 \begin{proof}
-By contradiction, assume that there is an assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, and $f_{\varphi_i(\s,\X)}(\A) = 0.$ That is, there is no assignment $\R$ to $\X$ such that $\A\models\varphi_i(\s,\R).$ Since $\A\models\alpha_i(\s)$, it holds that $\A\models\theta_i(\s)$.
-Therefore, as we showed in part a) of Lemma \ref{first}, there is no assignment $\s^\prime$ to $\x$ such that $f_{\varphi_i(\s,\X)}(\A) > 0.$ This means $\varphi_i(\x,\X)$ is unsatisfiable, but, as we mentioned previously, all of such disjuncts have been eliminated, which is a contradiction.
+	Let $\s = (\s_1,\ldots,\s_{n_i},\t)$ such that $\A\models\alpha_i(\s)$, and let $\P = (\P_1,\ldots,\P_r)$ where $P_i = \bigcup_{j\in[1,n_i]:\:\lambda(j)=i}\{s_i\}$.
+We will show that $\A\models\varphi_i(\s,\P)$.
+By contradiction, suppose that $\A\not\models\exists\x\,\varphi_i(\x,\P)$.
+That is, there is no assignment $\e$ to $\x$ such that $\A\models\varphi(\e,\P).$ Following the same proof as in case (a) of Lemma \ref{first}, we conclude that $\varphi_i(\x,\P)$ is inconsistent.
+However, as we mentioned previously, all of such disjuncts have been eliminated, which leads to a contradiction.
 \end{proof}
 
-Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\exists\x\:\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\P$ such that $\A\models\exists\x\:\varphi_2(\x,\P)$, and so on.
-As we already know how to exclude that assignment in the first disjunct, we now move to the next disjuncts.
-Similarly to the first part of the proof, let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
+Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\exists\x\:\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\Q$ such that $\A\models\exists\x\:\varphi_2(\x,\Q)$, and so on.
+As we already know how to exclude that assignment in the first disjunct, we will now deal with the next disjuncts.
+Similarly to the first part of the proof, for each $i\in\{1,\ldots,k\}$ let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
 \begin{multline*}
@@
 \end{multline*}
-where $\u_j,\v_j,$ for $\{1,\ldots,n_i\}$ and $\w$ have the corresponding number of variables. Also, let
+where $\u_j$ and $\v_j,$ have $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, and $\w$ has $p_i$ variables. Also, let
 \begin{multline*}
@@
 \end{multline*}
-Note that $\psi_i(\X)$ excludes only the lexicographically smallest predicate $\P$ such that $\A\models\exists\x\:\varphi_i(\X,\x),$ if there is at least one.
+Note that $\psi_i(\X)$ excludes only the lexicographically smallest tuple of predicates $\P$ such that $\A\models\exists\x\:\varphi_i(\P,\x),$ if there is at least one.
 In other words, every assignment $\P^\prime \neq \P$ is such that $\A\models\psi_i(\P^\prime).$ Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
@@
 \end{multline}
-Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Let $\A\in\Truc$.
-Suppose first that there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least one assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\d = (\d_1,\dots,\d_{n_q},\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha_q(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n_q\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{\lambda_q(j)=i}\{\d_j\}$.
-As we did in Lemma \ref{first} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies (a) and (b)
+Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Assume that $\A\in\Truc$.
+Suppose first that there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least one assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\d = (\d_1,\dots,\d_{n_q},\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha_q(\d)$, where $\d_i$ is the corresponding assignment to $\x_i$, for every $i\in\{1,\dots,n_q\}$, and $\ep$ is an assignment for $\y$.
+Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j:\lambda_q(j)=i,j\in[1,n_q]}\{\d_j\}$.
+As we did in Lemma \ref{first} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies both (a) and (b)
 \begin{enumerate}
-\item[(a)] As we showed in part (a) of Lemma \ref{first}, this assignment $\P$ to $\X$ is such that $\A\models\exists\x\:\varphi_q(\x,\P),$ if and only if there is at least one such an assignment.
-However, as we showed in Claim Six, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_q(\s)$, then there {\em is} such an assignment to $\X$.
-The assignment $\d$ to $\x$ satisfies that so we see that $\A\models\exists\x\:\varphi_q(\x,\P)$.
+	\item[(a)] As we showed in part (a) of Lemma \ref{first}, if there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_q(\x,\R)$, then $\A\models\exists\x\:\varphi_q(\x,\P)$ for this particular $\P$.
+However, as we showed in Claim \ref{alpha}, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_q(\s)$, then there {\em is} such an assignment to $\X$.
+The assignment $\d$ to $\x$ satisfies that $\A\models\alpha_q(\d)$, so it holds that $\A\models\exists\x\:\varphi_q(\x,\P)$.
 It immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
@@
 (1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
-(2) We can use the proof in Lemma \ref{first} to see that $\A\not\models\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula $$ \hat{\psi}_q(\X) =  (\exists\v\:\varphi_1(\v,\X)\vee\cdots\vee\exists\v\:\varphi_{q-1}(\v,\X)\vee\psi_q(\X)) $$ appears as a conjunct in $\varphi^\prime_i(\x,\X)$.
-However, since $\A\not\models\exists\v\:\varphi_i(\v,\P)$ for $i\in\{1,\ldots,q-1\}$ and $\A\not\models\psi_q(\P)$, it holds that $\A\not\models\hat{\psi}_q(\P)$, so $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$.
+(2) We can use the proof in Lemma \ref{first} to see that $\A\not\models\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula
+	$$ \zeta_q(\X) =  (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{q-1}(\v)\vee\psi_q(\X)) $$ 
+	appears as a conjunct in $\varphi^\prime_i(\x,\X)$.
+However, by the choice of $q$, there is no $i\in\{1,\ldots,q-1\}$ such that $\A\models\exists\v\:\varphi_i(\v,\P)$, and also $\A\not\models\psi_q(\P)$.
+It follows that $\A\not\models\zeta_q(\P)$,  so $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$.
 And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
@@
 As we deduce from the part (c) of Lemma \ref{first}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
-Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x),$ so $\A\models\psi_i(\P^\prime)$ for each $i$.
-But also, $\A\models\exists\x\:\psi_q(\P^\prime),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which is a contradiction.
+Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x)$, so $\A\models\psi_i(\P^\prime)$ for each $i$.
+But as we mentioned, also $\A\models\:\psi_q(\P^\prime),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which leads to a contradiction.
 \end{enumerate}
 
-With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ then $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
+With this, we conclude that for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0$, we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
 
@@
 It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
-And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Therefore, $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1 = f\dotminus 1.$
-
-Finally, let $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}.$ We conclude that $f^\prime = f\dotminus 1\in\E1$
-
-% CASE 3
+And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Hence, from the results in the previous paragraph, if $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}$, we have that $f^\prime = f\dotminus 1\in\E1$
 
@@
 for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$.
-We define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2, where
+In order to prove that $f\dotminus 1\in\E1$, we define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2, where
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\ell} + c_{\ell} = a_{\ell}$ for $\ell\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, $w_i$ has $q_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\w_i) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
+where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\ell} + c_{\ell} = a_{\ell}$ for $\ell\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, and $\w_i$ has $q_i$ variables.
+Furthermore, we have that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\w_i) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
 The formulas $\varphi^{-}_i(\X,\y_i,\w_i)$, $\beta_i(\x,\z)$, are also defined analogously.
 
-In this case, we mix both previous strategies, that is, we are going to {\em isolate} the lexicographically smallest predicate that satisfies the first satisfiable disjunct, and exclude the lexicographically smallest tuple that satisfies the evaluated disjunct.
+In this case, we mix both the strategies in cases 1 and 2.
+That is, we are going to {\em isolate} the lexicographically smallest tuple of predicates that satisfies the first satisfiable disjunct, and then {\em exclude} the lexicographically smallest tuple that satisfies the isolated disjunct.
 
-Let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$. Let
+Let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and
 \begin{multline*}
@@
 \begin{multline*}
-\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\w(\exists\u\:\varphi(\u,\X,\w) \wedge \varphi_{<,p}(\w,\z)).
+\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\w(\exists\u\:\varphi(\u,\X,\w) \wedge \varphi_{p,<}(\w,\z)).
 \end{multline*}
 Let $\P$ and $\e$ be assignments to $\X$ and $\z$.
-Note that $\psi_i(\X,\z)$ excludes $\P$ and $\e$ only if $\P$ is the lexicographically smallest predicate such that $\A\models\exists\x\:\varphi_i(\x,\P,\d)$, for some assignment $\d$ to $\z$, and $\e$ is the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_i(\x,\P,\e).$ We define $\varphi_i^\prime(\x,\X,\z)$ in the same way as case 2:
+Note that $\psi_i(\X,\z)$ excludes $\P$ and $\e$ only if $\P$ is the {\em lexicographically smallest} tuple of predicates (Same as case 2) such that $\A\models\exists\x\:\varphi_i(\x,\P,\d)$, for some assignment $\d$ to $\z$, and $\e$ is the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_i(\x,\P,\e).$ We define $\varphi_i^\prime(\x,\X,\z)$ in the same way as case 2:
 \begin{multline*}
@@
 Let $q$ be the least $i\in\{1,\ldots,k\}$ such that $\A\models\exists\x\:\varphi_i(\x,\R,\d)$ for some assignment $R$ to $\X$ and some assignment $\d$ to $\z$.
-Let $\P$ be the lexicographically smallest predicate such that $\A\models\exists\x\:\varphi_q(\x,\P,\d^\prime)$ for some assignment $\d^\prime$ to $\z$.
-Let $\e$ be the lexicographically smallest assignment to $z$ such that $\A\models\exists\x\:\varphi_q(\x,\P,\e).$ This formula is such that (a) $\A\models\exists\x\:\varphi(\x,\P,\e)$, (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P,\e)$ and $\P$ and $\e$ are the only assignments that satisfy (a) and (b).
-The proof of this is analogous to Case 2.
-We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X,\z)} = f_{\exists\x\,\varphi(\x,\X,\z)}\dotminus 1.$
-
+Let $\P$ be the lexicographically smallest tuple of predicates such that $\A\models\exists\x\:\varphi_q(\x,\P,\d^\prime)$ for some assignment $\d^\prime$ to $\z$.
+Let $\e$ be the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_q(\x,\P,\e).$ This formula is such that (a) $\A\models\exists\x\:\varphi(\x,\P,\e)$, (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P,\e)$ and $\P$ and $\e$ are the only assignments that satisfy (a) and (b).
+The proof of this is analogous to case 2.
+Therefore, we conclude that $f_{\exists\x\,\varphi^\prime(\x,\X,\z)} = f_{\exists\x\,\varphi(\x,\X,\z)}\dotminus 1.$
 \end{enumerate}
 \end{proof}
-
-\end{document}

### 2015-08-04 tex/case2general.tex

@@
 \end{eqnarray*}
-Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\Truc.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
+Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\Truc.$
+\icristian{
+Cuidado!
+Es indecidible verificar si una formula es no satisfacible.
+Lo que queremos decir acá es que verificar que $\left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x)$ (sin el $\beta$) es decidible.
+}
+Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
 We use an injective\footnote[3]{We need this function to be injective because this way we can assure each one of the $k$ satisfiable disjuncts to be represented} function $\rho:\{1,\ldots,k\}\to\{1,\ldots,m\}$ such that $\varphi_i(\x,\X) = \xi_{\rho(i)}(\x,\X)$ is satisfiable, for every $i\in\{1,\ldots,k\}.$
@@
 where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\y_i$ is a $p_i$-tuple of variables in $\x$, $\varphi_i^{-}(\X,\y_i)$ is a conjunction of negated predicates in $\X$, $\theta_i(\x)$ defines a total order on a partition of $\x$, and $\beta_i(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
+\icristian{De nuevo, necesitamos que $\beta_i(\x)$ mencione todas las variables?}
 Let $\alpha_i(\x) = \theta_i(\x)\wedge\beta_i(\x).$
@@
 \end{claim}
+\icristian{Esto lo colocaría como Lema ya que no estas afirmando nada.}
 \begin{proof}
@@
 As we already know how to exclude that assignment in the first disjunct, we will now deal with the next disjuncts.
+\icristian{Acá podriamos colocar la parte de la demostración que trata con el caso de una conjunción}
 Similarly to the first part of the proof, for each $i\in\{1,\ldots,k\}$ let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
@@
 \end{multline*}
+\icristian{
+Es necesario decir que $\forall\x\:\neg\alpha_i(\x)$?
+}
 Note that $\psi_i(\X)$ excludes only the lexicographically smallest tuple of predicates $\P$ such that $\A\models\exists\x\:\varphi_i(\P,\x),$ if there is at least one. 
+\icristian{
+De nuevo, yo sería mas preciso con este argumento del ``lexicographically smallest tuple of predicates''
+}
 In other words, every assignment $\P^\prime \neq \P$ is such that $\A\models\psi_i(\P^\prime).$ Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
@@
 \end{multline}
+\icristian{
+Yo representaría ``$(\exists\v\:\alpha_1(\v)\vee\psi_2(\X))$'' como ``$(\forall\v\:\neg \alpha_1(\v)\rightarrow \psi_2(\X))$''. Se entiende mejor que representa.
+}

### 2015-08-04 tex/case2part1.tex

@@
 \end{eqnarray*}
-for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
+for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. 
+\icristian{
+Yo cambiaria el orden de la demostracion.
+Primero mostraria como reescribir la formula en DNF agregando los ordenes totales y despues el caso de una conjuncion.
+Como esta ahora, parece un caso muy sintetico pero cuando uno ve la descomposicion le queda claro.
+}
+For the time being, suppose that
 \begin{eqnarray}
@@
 where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
-Note that $\theta(\x)$ also mentions all variables in $\x$.
-We also assume that $(\x_1,\dots,\x_n,\y) = \x$.
+\icristian{
+Porque necesitas que  $\beta(\x)$ mencione todas las variables? Me parece que esto no es necesario asumirlo.	
+}
+Note that $\theta(\x)$ also mentions all variables in $\x$. We also assume that $(\x_1,\dots,\x_n,\y) = \x$. 
+\icristian{
+Las suposiciones anteriores deberian ir directamente en la definicion y no como un ``comentario''.	
+}
 As an example, the following formula is of this form:
@@
 Similarly to the previous proof, we would like to eliminate the {\em lexicographically smallest}\footnote{We consider the lexicographically smallest tuple of predicates as the one in which its predicates contain the lexicographically smallest tuples and do not contain any more tuples than those} tuple of predicates that satisfies the formula \eqref{f1}.
-Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
-We now define
+\icristian{
+Esta explicacion habría que mejorarla ya que no es exactamente verdad que estas eliminando la tupla lexicograficamente menor de predicacos que satisface la formula. 	
+}
+Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$. We now define
 \begin{multline*}
@@
 \end{multline*}
+
 where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$. 
+\icristian{La definición de $\alpha$ es muy usada en la demostración por lo que yo la colocaría cuando se define $\theta$ y $\beta$.}
+\icristian{Yo recordaria lo que es $\varphi_{m,<}$.}
 Note that $\alpha_{\min}$ is satisfied only by the lexicographically \linebreak smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
@@
 \end{multline}
+\icristian{
+Esta formula se simplificaría si usamos que $\varphi(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)$.
+}
+\icristian{
+Yo explicaría aquí ambas disyunciones. Ayudaría mucho para entender la demostración.
+}
 We now show a result by which the main proof will follow.
+\icristian{
+``which the main proof will follow'' esto no es verdad, falta harto mas.
+}

### 2015-08-04 tex/case3.tex

@@
 \end{eqnarray*}
-where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\ell} + c_{\ell} = a_{\ell}$ for $\ell\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, and $\w_i$ has $q_i$ variables.
+where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\lambda_i(j)} + c_{\lambda_i(j)} = a_{\lambda_i(j)}$), $\y_i$ has $p_i$ variables, and $\w_i$ has $q_i$ variables.
 Furthermore, we have that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\w_i) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
+\icristian{
+Este orden total considera 	$\x,\z$ como un todo? Sería bueno recalcar esto.
+}
 The formulas $\varphi^{-}_i(\X,\y_i,\w_i)$, $\beta_i(\x,\z)$, are also defined analogously.
@@
 That is, we are going to {\em isolate} the lexicographically smallest tuple of predicates that satisfies the first satisfiable disjunct, and then {\em exclude} the lexicographically smallest tuple that satisfies the isolated disjunct.
+\icristian{
+Esta intuición no se entiende nada. Mejorar al igual que las intuiciones anteriores.
+}
 
@@
 Let $\P$ and $\e$ be assignments to $\X$ and $\z$.
-Note that $\psi_i(\X,\z)$ excludes $\P$ and $\e$ only if $\P$ is the {\em lexicographically smallest} tuple of predicates (Same as case 2) such that $\A\models\exists\x\:\varphi_i(\x,\P,\d)$, for some assignment $\d$ to $\z$, and $\e$ is the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_i(\x,\P,\e).$ We define $\varphi_i^\prime(\x,\X,\z)$ in the same way as case 2:
+Note that $\psi_i(\X,\z)$ excludes $\P$ and $\e$ only if $\P$ is the {\em lexicographically smallest} tuple of predicates (same as case 2) such that $\A\models\exists\x\:\varphi_i(\x,\P,\d)$, for some assignment $\d$ to $\z$, and $\e$ is the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_i(\x,\P,\e).$ We define $\varphi_i^\prime(\x,\X,\z)$ in the same way as case 2:
 \begin{multline*}
@@
 Let $q$ be the least $i\in\{1,\ldots,k\}$ such that $\A\models\exists\x\:\varphi_i(\x,\R,\d)$ for some assignment $R$ to $\X$ and some assignment $\d$ to $\z$.
-Let $\P$ be the lexicographically smallest tuple of predicates such that $\A\models\exists\x\:\varphi_q(\x,\P,\d^\prime)$ for some assignment $\d^\prime$ to $\z$.
+Let $\P$ be the lexicographically smallest tuple of predicates
+\icristian{
+Acá no esta bien definido que significa ``lexicographically smallest tuple of predicates''.
+También deberias hablar de ``$(\P, \z)$ be the lexicographically smallest tuple of predicates'' incluyendo a $\z$.
+}
+such that $\A\models\exists\x\:\varphi_q(\x,\P,\d^\prime)$ for some assignment $\d^\prime$ to $\z$.
 Let $\e$ be the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_q(\x,\P,\e).$ This formula is such that (a) $\A\models\exists\x\:\varphi(\x,\P,\e)$, (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P,\e)$ and $\P$ and $\e$ are the only assignments that satisfy (a) and (b).

### 2015-08-04 tex/lemmafirst.tex

@@
 \begin{proof}
-Let $\A \in \Truc$.
-Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
+Let $\A \in \Truc$. Consider two cases: 
+\icristian{
+Acá yo explicaría cuales son los dos casos para entender el roadmap de la demostración.	
+}
+assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
 Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
@@
 We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
+
 \begin{enumerate}
@@
 In particular, $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which contradicts the initial assumption that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
+\icristian{Este argumento es uno de los principales.
+Si cambiamos el orden de la demostración (colocando la descomposicion DNF primero) esta demo se podría colocar en el Claim 5.1 y después reusar acá}
+
 \item[(b)] Note that if $\A \models \alpha_{\min}(\c_1,\ldots,\c_n)$, then necessarily $\c_i = \d_i$ for $i\in\{1,\ldots,n\}.$ However, by the construction of $\P$, we see that 
 $$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \d_j\Big).$$ Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
+\icristian{
+Aqui yo recalcaría que 	$\A$ no modela la segunda parte de $\varphi^\prime(\x,\P)$.
+}
 \item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
 We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$.
-Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $P^\prime_i$.
-Then $\A\models \neg P^\prime_i(\d_i)$, and so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
+Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $P^\prime_{\lambda(i)}$.
+Then $\A\models \neg P^\prime_{\lambda(i)}(\d_i)$, and so,
+$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ 
+from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
 Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $P_i$ does not have.
-If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
+If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P'_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
 On both cases, we have a contradiction.
+\icristian{
+Esta demostración es mas clara si es directa, diciendo que para todo $\P^\prime \neq \P$ se cumple que $\A\models\exists\x\,\varphi(\x,\P^\prime)$ y $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
+}
 \end{enumerate}

### 2015-08-04 tex/th2.tex

@@
 As we supposed initially, let $h\in\E1$ such that $f_{all}-f_{\#3DNF} = h$.
-For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is {\sc co-NP}-complete.
+For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is {\sc coNP}-complete.
 However, as we showed previously in Theorem \ref{decisionptime}, $L_h \in \textsc{P}$.
-Then, $\textsc{co-NP} \subseteq \textsc{P}$, from which we conclude that $\textsc{P} = \textsc{NP}$.
+Then, $\textsc{coNP} \subseteq \textsc{P}$, from which we conclude that $\textsc{P} = \textsc{NP}$.
 \end{proof}

### 2015-08-04 tex/sharp-sigma-one-plus.tex
1ago

@@
 \usepackage{bm}
+
+\usepackage[textwidth=2cm,textsize=small]{todonotes}
+\newcommand{\cristian}[1]{\todo[color=blue!10]{{\bf Cristian:} #1}}
+\newcommand{\icristian}[1]{\todo[inline, color=blue!10]{{\bf Cristian:} #1}}
+\newcommand{\marcelo}[1]{\todo[color=red!10]{{\bf Marcelo:} #1}}
+\newcommand{\imarcelo}[1]{\todo[inline, color=red!10]{{\bf Marcelo:} #1}}
+\newcommand{\martin}[1]{\todo[color=green!10]{{\bf Martin:} #1}}
+\newcommand{\imartin}[1]{\todo[inline, color=green!10]{{\bf Martin:} #1}}
 
@@
 \end{eqnarray*}
+\icristian{Yo primero definiria la clase de ``$F$-extended quantifier-free $\L$-formula'' para una logica F sobre $\L$ y de ahi hablaria de quantifier free formula y FO-extended.
+Con esto tambien puedes definir altiro $\#\Sigma_1$, $\#\Sigma_0$ (que no estan definidas) y  $\#\Sigma_1[\mbox{{\sc FO}}]$.
+Esto tambien nos servira despues para otras logicas que queriamos extender.}
 Let $\Y = (Y_1,\dots,Y_q)$ be a tuple of second-order variables of arity $c_1,\ldots,c_q$, and let $\y$ be a tuple of first order variables.

### 2015-08-04 tex/examples.tex
3 ago

@@
 
-Let $\L = \{S_0,S_1,S_2,S_3,\leq\}$, and $\A = \langle A, S_0^\A, S_1^\A, S_2^\A, S_3^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc 3-DNF} formula $\psi$, where $A$ is the set of variables mentioned in $\psi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{0,1,2,3\}$:
+Let $\L = \{S_0,S_1,S_2,S_3,\leq\}$, and $\A = \langle A, S_0^\A, S_1^\A, S_2^\A, S_3^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc 3-DNF} formula $\Psi$, where $A$ is the set of variables mentioned in $\Psi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{0,1,2,3\}$:
 \begin{eqnarray*}
-S_0^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
-S_1^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
-S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
-S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\psi\}.
+S_0^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Psi\},\\
+S_1^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Psi\},\\
+S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Psi\},\\
+S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Psi\}.
 \end{eqnarray*}
@@
 \end{align*}
-It is easily seen that each relation $T$ over $\A$ such that $\A\models\varphi(T)$ represents a valid assignment to the variables in $\psi$, and therefore, $f_{\varphi(T)}$ represents {\sc \#3-DNF}.
+It is easily seen that each relation $T$ over $\A$ such that $\A\models\varphi(T)$ represents a valid assignment to the variables in $\Psi$, and therefore, $f_{\varphi(T)}$ represents {\sc \#3-DNF}.
 
@@
 
-Let $\L = \{P_1,P_2,P_3,N_1,N_2,N_3,\leq\}$, and $\A = \langle A, P_1^\A,P_2^\A,P_3^\A,N_1^\A,N_2^\A,N_3^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc 3-HORN-SAT} formula $\psi$, where $A$ is the set of variables mentioned in $\psi$, and the relations in $\L$ are interpreted as follows:
+Let $\L = \{P_1,P_2,P_3,N_1,N_2,N_3,\leq\}$, and $\A = \langle A, P_1^\A,P_2^\A,P_3^\A,N_1^\A,N_2^\A,N_3^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc 3-HORN-SAT} formula $\Psi$, where $A$ is the set of variables mentioned in $\Psi$, and the relations in $\L$ are interpreted as follows:
 \begin{eqnarray*}
-P_1^\A &=& \{a\mid a \mbox{ appears as a conjunct in }\psi\},\\
-P_2^\A &=& \{(a_1,a_2)\mid (a_1\vee \neg a_2) \mbox{ appears as a conjunct in }\psi\},\\
-P_3^\A &=& \{(a_1,a_2,a_3)\mid (a_1\vee \neg a_2\vee \neg a_3) \mbox{ appears as a conjunct in }\psi\},\\
-N_1^\A &=& \{a\mid (\neg a) \mbox{ appears as a conjunct in }\psi\},\\
-N_2^\A &=& \{(a_1,a_2)\mid (\neg a_1\vee \neg a_2) \mbox{ appears as a conjunct in }\psi\},\\
-N_3^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1\vee \neg a_2\vee \neg a_3) \mbox{ appears as a conjunct in }\psi\}.
+P_1^\A &=& \{a\mid a \mbox{ appears as a conjunct in }\Psi\},\\
+P_2^\A &=& \{(a_1,a_2)\mid (a_1\vee \neg a_2) \mbox{ appears as a conjunct in }\Psi\},\\
+P_3^\A &=& \{(a_1,a_2,a_3)\mid (a_1\vee \neg a_2\vee \neg a_3) \mbox{ appears as a conjunct in }\Psi\},\\
+N_1^\A &=& \{a\mid (\neg a) \mbox{ appears as a conjunct in }\Psi\},\\
+N_2^\A &=& \{(a_1,a_2)\mid (\neg a_1\vee \neg a_2) \mbox{ appears as a conjunct in }\Psi\},\\
+N_3^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1\vee \neg a_2\vee \neg a_3) \mbox{ appears as a conjunct in }\Psi\}.
 \end{eqnarray*}
@@
 We can easily reorder $\varphi(T)$ as $\forall x \forall y \forall z\,\theta(x,y,z,T)$ where $\theta(x,y,z,T)$ is an {\sc FO}-extended quantifier-free $\L$-formula such that any second order variable only appears in a Horn clause.
-It is easily seen that each relation $T$ over $\A$ such that $\A\models\forall x \forall y \forall z\,\theta(x,y,z,T)$ represents a valid assignment to the variables in $\psi$, and therefore, $f_{\forall x \forall y \forall z\,\theta(x,y,z,T)}$ represents {\sc \#3-HORN-SAT}.
+It is easily seen that each relation $T$ over $\A$ such that $\A\models\forall x \forall y \forall z\,\theta(x,y,z,T)$ represents a valid assignment to the variables in $\Psi$, and therefore, $f_{\forall x \forall y \forall z\,\theta(x,y,z,T)}$ represents {\sc \#3-HORN-SAT}.
 
@@
 
-Let $\L = \{P,N,\leq\}$, and $\A = \langle A, P^\A, N^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc DNF} formula $\psi$, where $A$ is the set of variables mentioned in $\psi$, $P$ and $N$ is a binary relations described as follows:
+Let $\L = \{P,N,\leq\}$, and $\A = \langle A, P^\A, N^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc DNF} formula $\Psi$, where $A$ is the set of variables mentioned in $\Psi$, $P$ and $N$ is a binary relations described as follows:
 \begin{eqnarray*}
-P^\A &=& \{(c,\ell)\mid \ell \mbox{ appears in the $c$-th clause in }\psi\},\\
-N^\A &=& \{(c,\ell)\mid \neg \ell \mbox{ appears in the $c$-th clause in }\psi\}.
+P^\A &=& \{(c,\ell)\mid \ell \mbox{ appears in the $c$-th clause in }\Psi\},\\
+N^\A &=& \{(c,\ell)\mid \neg \ell \mbox{ appears in the $c$-th clause in }\Psi\}.
 \end{eqnarray*}
@@
 We reorder $\varphi(T)$ as $\exists c \forall \ell_1 \forall \ell_2\, \theta(c,\ell_1,\ell_2,T)$ where $\theta(c,\ell_1,\ell_2,T)$ is an {\sc FO}-extended quantifier-free $\L$-formula such that any second order variable only appears in a Horn clause.
-It is easily seen that each relation $T$ over $\A$ such that $\A\models \exists c \forall \ell_1 \forall \ell_2\, \theta(c,\ell_1,\ell_2,T)$ represents a valid assignment to the variables in $\psi$, and therefore, $f_{\exists c \forall \ell_1 \forall \ell_2\, \theta(c,\ell_1,\ell_2,T)}$ represents {\sc \#DNF}.
+It is easily seen that each relation $T$ over $\A$ such that $\A\models \exists c \forall \ell_1 \forall \ell_2\, \theta(c,\ell_1,\ell_2,T)$ represents a valid assignment to the variables in $\Psi$, and therefore, $f_{\exists c \forall \ell_1 \forall \ell_2\, \theta(c,\ell_1,\ell_2,T)}$ represents {\sc \#DNF}.
+
+\subsection{{\sc \#HORN-SAT}}
+A function $f:\Truc \to \N$ is in $\Efour$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x^1,\x^2,\Y,\y)$ such that $f = f_{\forall\x^1\exists \x^2\,\varphi(\x^1,\x^2,\Y,\y)}$ and any second order variables only appear in Horn clauses.
+We will now show that {\sc \#DNF} $\in\Ethree$.
+
+Let $\L = \{P,N,\leq\}$, and $\A = \langle A, P^\A, N^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc HORN-SAT} formula $\Psi$, where $A$, $P$ and $N$ are defined as in the previous case.
+
+Now let $T$ be a unary second-order variable, and
+\begin{eqnarray*}
+	\psi(T) &=& \forall c \forall \ell\, [(P(c,\ell)\wedge\forall\ell'(N(c,\ell')\to T(\ell')))\to T(\ell)] \\
+	&\equiv& \forall c \forall \ell\, [\neg P(c,\ell)\vee \exists\ell'(N(c,\ell')\wedge \neg T(\ell'))\vee T(\ell)].
+\end{eqnarray*}
+Let $\neg X(c,\ell')$ be defined as $N(c,\ell')\wedge \neg T(\ell)$. That is,
+\begin{eqnarray*}
+	&& \forall c \forall \ell'\, [(N(c,\ell') \wedge \neg T(\ell')) \leftrightarrow \neg X(c,\ell')]\\
+	&\equiv& \forall c \forall \ell'\,[(\neg N(c,\ell')\vee T(\ell')\vee \neg X(c,\ell'))\wedge (X(c,\ell')\vee \neg T(\ell'))\wedge (X(c,\ell')\vee N(c,\ell'))].
+\end{eqnarray*}
+Finally, let
+\begin{multline*}
+	\varphi(T,X) = \forall c \forall \ell'\,[(\neg N(c,\ell')\vee T(\ell')\vee \neg X(c,\ell'))\wedge (X(c,\ell')\vee \neg T(\ell'))\wedge (X(c,\ell')\vee N(c,\ell'))]\,\wedge \\
+	\forall c \forall \ell\, [\neg P(c,\ell)\vee \exists\ell'\neg X(c,\ell')\vee T(\ell)]	.
+\end{multline*}
+This can be rewritten as
+\begin{multline*}
+	\varphi'(T,X) = \forall c \forall \ell \exists\ell'\,[(\neg N(c,\ell)\vee T(\ell)\vee \neg X(c,\ell))\wedge (X(c,\ell)\vee \neg T(\ell))\wedge (X(c,\ell)\vee N(c,\ell))\,\wedge \\
+	(\neg P(c,\ell)\vee \neg X(c,\ell')\vee T(\ell))].
+\end{multline*}
+Let $\theta(c,\ell,\ell',T,X)$ be such that $\varphi'(T,X) = \forall c \forall \ell \exists\ell'\,\theta(c,\ell,\ell',T,X)$.
+Note that $\theta(c,\ell,\ell',T,X)$ is an {\sc FO}-extended quantifier-free $\L$-formula such that any second order variable only appears in a Horn clause.
+First, note that each relation $T$ over $\A$ such that $\A\models \psi(T)$ represents a valid assignment to the variables in $\Psi$, and that in $\theta(c,\ell,\ell',T,X)$, $X$ is completely defined by the assignment to $T$.
+Therefore $f_{\psi(T)}(\A') = f_{\varphi'(T,X)}(\A')$ for each $\A'\in\Truc$, from which we conclude that $f_{\forall c \forall \ell \exists\ell'\,\theta(c,\ell,\ell',T,X)}$ represents {\sc \#HORN-SAT}.

### 2015-09-01 tex/classes.tex
1 sep

\section{Extended Horn counting hierarchies}

We define syntactically the classes $\EiE$ and $\EiU$ as following. First we define Horn clauses.
\begin{eqnarray*}
PL &::=& X_i(\x),\, i\in\N,\\
NL &::=& \neg X_i(\x),\, i\in\N,\\
NC &::=& NL \ \mid \ \alpha, \alpha \mbox{ is an {\sc FO}-formula over } \L \ \mid \ (NC \vee NC),\\
HC &::=& NC \ \mid \ (NC \vee PL),
\end{eqnarray*}
where $\x$ is a tuple with the corresponding number of variables. Now we define the classes' syntaxis inductively.
\begin{enumerate}
	\item $\EzeroE$:
	\begin{eqnarray*}
	E_0 &::=& HC \ \mid \ E_0 \wedge E_0.
	\end{eqnarray*}
	\item $\EzeroU$:
	\begin{eqnarray*}
	U_0 &::=& E_0.
	\end{eqnarray*}
	\item $\EipE$:
	\begin{eqnarray*}
	E_{i+1} &::=& U_i \ \mid \ \exists x \, E_{i+1}.
	\end{eqnarray*}
	\item $\EipU$:
	\begin{eqnarray*}
		U_{i+1} &::=& E_i \ \mid \ \forall x \, U_{i+1}.
	\end{eqnarray*}
\end{enumerate}
A function $f$ is in $\EiE$ (resp. $\EiU$) if there is an $\L$-formula $\varphi$ defined by the grammar $E_i$ (resp. $U_i$) such that $f = f_{\varphi}$.

We also define the classes $\EiEE$ and $\EiUE$ with the same grammar, but with the following changes:
\begin{eqnarray*}
	PL &::=& X_i(\x),\, i\in\N,\\
	NL &::=& \neg X_i(\x),\, i\in\N \ \mid \ \exists x\, NL\\
	NC &::=& NL \ \mid \ \alpha, \alpha \mbox{ is an {\sc FO}-formula over } \L \ \mid \ (NC \vee NC),\\
	HC &::=& NC \ \mid \ (NC \vee PL),
\end{eqnarray*}

### 2015-09-01 tex/examples.tex
1 sep

@@
 \subsection{{\sc \#3-DNF}}
-A function $f:\Truc \to \N$ is in $\Eone$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\Y,\y)$ such that $f = f_{\varphi(\Y,\y)}$.
-We will now show that {\sc \#3-DNF} $\in\Eone$.\\
+A function $f:\Truc \to \N$ is in $\EoneE$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\Y,\y)$ such that $f = f_{\varphi(\Y,\y)}$.
+We will now show that {\sc \#3-DNF} $\in\EoneE$.\\
 
@@
 \subsection{{\sc \#3-HORN-SAT}}
-A function $f:\Truc \to \N$ is in $\Etwo$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\forall \x\,\varphi(\x,\Y,\y)}$ and any second order variables only appear in Horn clauses.
-We will now show that {\sc \#3-HORN-SAT} $\in\Etwo$.\\
+A function $f:\Truc \to \N$ is in $\EoneU$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\forall \x\,\varphi(\x,\Y,\y)}$ and any second order variables only appear in Horn clauses.
+We will now show that {\sc \#3-HORN-SAT} $\in\EoneU$.\\
 
@@
 \subsection{{\sc \#DNF}}
-A function $f:\Truc \to \N$ is in $\Ethree$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x^1,\x^2,\Y,\y)$ such that $f = f_{\exists\x^1\forall \x^2\,\varphi(\x^1,\x^2,\Y,\y)}$ and any second order variables only appear in Horn clauses.
-We will now show that {\sc \#DNF} $\in\Ethree$.
+A function $f:\Truc \to \N$ is in $\EtwoE$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x^1,\x^2,\Y,\y)$ such that $f = f_{\exists\x^1\forall \x^2\,\varphi(\x^1,\x^2,\Y,\y)}$ and any second order variables only appear in Horn clauses.
+We will now show that {\sc \#DNF} $\in\EtwoE$.
 
@@
 \subsection{{\sc \#HORN-SAT}}
-A function $f:\Truc \to \N$ is in $\Efour$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x^1,\x^2,\Y,\y)$ such that $f = f_{\forall\x^1\exists \x^2\,\varphi(\x^1,\x^2,\Y,\y)}$ and any second order variables only appear in Horn clauses.
-We will now show that {\sc \#DNF} $\in\Ethree$.
+A function $f:\Truc \to \N$ is in $\EtwoU$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x^1,\x^2,\Y,\y)$ such that $f = f_{\forall\x^1\exists \x^2\,\varphi(\x^1,\x^2,\Y,\y)}$ and any second order variables only appear in Horn clauses.
+We will now show that {\sc \#DNF} $\in\EtwoE$.
 

### 2015-09-15 tex/classes.tex
15 sep

@@
 PL &::=& X_i(\x),\, i\in\N,\\
-NL &::=& \neg X_i(\x),\, i\in\N,\\
+NL &::=& \neg X_i(\x),\, i\in\N \ \mid \ \exists x\, NL,\\
 NC &::=& NL \ \mid \ \alpha, \alpha \mbox{ is an {\sc FO}-formula over } \L \ \mid \ (NC \vee NC),\\
@@
 
-We also define the classes $\EiEE$ and $\EiUE$ with the same grammar, but with the following changes:
-\begin{eqnarray*}
-	PL &::=& X_i(\x),\, i\in\N,\\
-	NL &::=& \neg X_i(\x),\, i\in\N \ \mid \ \exists x\, NL\\
-	NC &::=& NL \ \mid \ \alpha, \alpha \mbox{ is an {\sc FO}-formula over } \L \ \mid \ (NC \vee NC),\\
-	HC &::=& NC \ \mid \ (NC \vee PL),
-\end{eqnarray*}
+The class $\PE$ is defined in [?] as $\PE = \{f\mid f\in\#P \mbox{ and its decision version } L_{f}\in P\}$.
+
+\begin{theo}
+	$\EtwoE \subseteq \PE$
+\end{theo} 
+\begin{proof}
+	Let $f = f_{\varphi(\X,\z)}$ such that $\varphi(\X,\z)$ is defined by the grammar $E_2$.
+First we notice that, as stated in \cite{DBLP:journals/jcss/SalujaST95}, for every $\L$-formula $\varphi$, $f_{\varphi}\in\#P$.\\
+
+	We will now prove that there is a polynomial time algorithm that decides $L_{f}$.
+This is equivalent to decide, given $\A= \langle A, \S\A, \leq^\A \rangle\in\Truc$, whether there exist assignments $\P,\z$ such that $\A\models\varphi(\P,\z)$.
+This is the same as $\A\models\exists\X\exists\x\,\varphi(\X,\x)$.
+Since $\varphi(\X,\x)$ is defined by the grammar $E_2$, there exists $\psi(\X,\x,\y,\u)$ such that $\varphi(\X,\x) = \exists\x\exists\y\forall\u\exists\v\,\psi(\X,\x,\y,\u,\v)$, where every variable in $\v$ appears on a negated second-order literal $\neg X_i$, with $i\in\N$.
+
+	Given $\A\in\Truc$ we generate an equivalent $\L$-formula $\theta(\X,\x,\y)$ with a series of operations.
+First, we replace every instance of $\v$ by a disjunction of all $r$-tuples in $A^r$, where $r$ is the number of variables in $\v$.
+The remainder is still a {\sc Horn}-clause.
+Second, we replace every instance of $\u$ by a conjunction of all $s$-tulples in $A^s$, where $s$ is the number or variables in $\u$.
+The remainder is still a conjunction of {\sc Horn}-clauses.
+
+	Now we notice that $\zeta = \exists\X\exists\x\exists\y\,\theta(\X,\x,\y)$ is an existential second-order FO-formula, for which $\A\models\zeta$ can be decided in polynomial time.
+\end{proof}

### 2015-09-15 tex/examples.tex
15 sep

--- 2015-08-04 tex/examples.tex
@@
 \subsection{{\sc \#3-DNF}}
-A function $f:\Truc \to \N$ is in $\Eone$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\Y,\y)$ such that $f = f_{\varphi(\Y,\y)}$.
-We will now show that {\sc \#3-DNF} $\in\Eone$.\\
+We will now show that {\sc \#3-DNF} $\in\EoneE$.\\
 
@@
 \subsection{{\sc \#3-HORN-SAT}}
-A function $f:\Truc \to \N$ is in $\Etwo$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\forall \x\,\varphi(\x,\Y,\y)}$ and any second order variables only appear in Horn clauses.
-We will now show that {\sc \#3-HORN-SAT} $\in\Etwo$.\\
+We will now show that {\sc \#3-HORN-SAT} $\in\EoneU$.\\
 
@@
 \subsection{{\sc \#DNF}}
-A function $f:\Truc \to \N$ is in $\Ethree$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x^1,\x^2,\Y,\y)$ such that $f = f_{\exists\x^1\forall \x^2\,\varphi(\x^1,\x^2,\Y,\y)}$ and any second order variables only appear in Horn clauses.
-We will now show that {\sc \#DNF} $\in\Ethree$.
+We will now show that {\sc \#DNF} $\in\EtwoE$.\\
 
@@
 We reorder $\varphi(T)$ as $\exists c \forall \ell_1 \forall \ell_2\, \theta(c,\ell_1,\ell_2,T)$ where $\theta(c,\ell_1,\ell_2,T)$ is an {\sc FO}-extended quantifier-free $\L$-formula such that any second order variable only appears in a Horn clause.
-It is easily seen that each relation $T$ over $\A$ such that $\A\models \exists c \forall \ell_1 \forall \ell_2\, \theta(c,\ell_1,\ell_2,T)$ represents a valid assignment to the variables in $\Psi$, and therefore, $f_{\exists c \forall \ell_1 \forall \ell_2\, \theta(c,\ell_1,\ell_2,T)}$ represents {\sc \#DNF}.
+It is easily seen that each relation $T$ over $\A$ such that $\A\models \exists c \forall \ell_1 \forall \ell_2\, \theta(c,\ell_1,\ell_2,T)$ represents a valid assignment to the variables in $\Psi$, and therefore, $f_{\exists c \forall \ell_1 \forall \ell_2\, \theta(c,\ell_1,\ell_2,T)}$ represents {\sc \#DNF}.\\
+
+By extension, {\sc \#3-DNF} $\in\EtwoE$.
 
 \subsection{{\sc \#HORN-SAT}}
-A function $f:\Truc \to \N$ is in $\Efour$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x^1,\x^2,\Y,\y)$ such that $f = f_{\forall\x^1\exists \x^2\,\varphi(\x^1,\x^2,\Y,\y)}$ and any second order variables only appear in Horn clauses.
-We will now show that {\sc \#DNF} $\in\Ethree$.
+We will now show that {\sc \#DNF} $\in\EtwoE$.\\
 

### 2015-09-22 tex/examples.tex
22 sep

@@
 \begin{eqnarray*}
-P^\A &=& \{(c,\ell)\mid \ell \mbox{ appears in the $c$-th clause in }\Psi\},\\
-N^\A &=& \{(c,\ell)\mid \neg \ell \mbox{ appears in the $c$-th clause in }\Psi\}.
+P^\A &=& \{(c,\ell)\mid \ell \mbox{ appears in the $c$-th clause of }\Psi\},\\
+N^\A &=& \{(c,\ell)\mid \neg \ell \mbox{ appears in the $c$-th clause of }\Psi\}.
 \end{eqnarray*}
@@
 \subsection{{\sc \#HORN-SAT}}
-We will now show that {\sc \#DNF} $\in\EtwoE$.\\
+We will now show that {\sc \#HORN-SAT} $\in\EoneU$.\\
 

### 2015-09-22 tex/s-one-fo.tex
22 sep

@@
 \input{th4}
+
+\begin{theo}
+	{\sc \#DNF} is hard for $\E1$ under parsimonious reductions. 
+\end{theo}
+\begin{proof}
+	Let $f$ be an arbitrary function in $\E1$ and let $f_{\textsc{DNF}}$ be the function that characterizes {\sc \#DNF}.
+We will now show a function $h:\Truc\to L(P)$ such that for every $\A\in\Truc$, it holds that $f(\A) = f_{\textsc{DNF}}(h(\A))$, and which can be computed in polynomial time.
+Let $\psi$ be such that $f = f_{\psi}$.
+We separate the proof in three cases:
+	\begin{enumerate}
+		\item All free variables in $\psi$ are of first order.
+Let $\psi = \exists x\,\varphi(\x,\z)$ where $\x = (x_1,\ldots,x_c)$ and $\z = (z_1,\ldots,z_d)$.
+Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a propositional formula $\Phi = f(\A)$.
+First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\,\varphi(\x,\e)$.
+Let $n$ be the number of such assignments to $\z$.
+Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$.
+The DNF formula we return is
+		\[
+		\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
+		\]
+		which has exactly $n$ possible assignments.
+		\item There is only second-order free variables in $\psi$.
+Let $\psi = \exists x\,\varphi(\x,\X)$, where $\x = (x_1,\ldots,x_c)$ and $\X = (X_1,\ldots,X_r)$.
+Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a DNF-formula $\Phi = f(\A)$ as follows:
+
+		First, we convert $\varphi(\x,\X)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X)$. Let each $\varphi_{i}(\x,\X)$ be of the form
+		\[
+		\varphi_{i}(\x,\X) = X_{s_1}(\x_{s_1}) \wedge \cdots \wedge X_{s_m}(\x_{s_m}) \wedge \neg X_{t_1}(\x_{t_1}) \wedge \cdots \wedge \neg X_{t_n}(\x_{t_n}) \wedge \alpha_i(\x),
+		\]
+		where every $\x_{s_i}$ and $\x_{t_j}$ have the corresponding number of variables, and $\alpha(\x)$ is a FO-formula.
+
+		Then, let $\Gamma_i = \{\a\in A^c\mid \A\models\alpha(\a)\}$.
+There is exactly $\vert A \vert^c$ tuples to check, and each of those checks can be done in polynomial time, so this set can also be generated in polynomial time.
+We generate a propositional variable $p^{\ell}_{\a}$ for each $\a\in \bigcup_{j = 1}^c A^j$, and for each $\ell\in\{1,\ldots,r\}$.
+Let $\p$ be the tuple of all such variables.
+Also, let $\rho_{\ell}(\a)$ be the tuple that results of reordering $\a$ using the reordering of the variables of $\x$ in $\x_{\ell}$.
+
+		Finally, for each $\varphi_i(\x,\X)$ the DNF formula we return is
+		\[
+		\Phi_i = \bigvee_{\a\in\Gamma_i} p^{s_1}_{\rho_{s_1}(\a)} \wedge \cdots \wedge p^{s_m}_{\rho_{s_m}(\a)} \wedge \neg p^{t_1}_{\rho_{t_1}(\a)} \wedge \cdots \wedge \neg p^{t_n}_{\rho_{t_n}(\a)}.
+		\]
+		and $\Phi = \bigvee_{i = 1}^k \Phi_i$.
+
+		Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$.
+We can conclude that $f(\A) = f_{\textsc{DNF}}(\Phi)$.
+		\item Let $\psi = \exists x\,\varphi(\x,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\X = (X_1,\ldots,X_r)$ and $\z = (z_1,\ldots,z_d)$.
+Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$ we generate a DNF-formula $\Phi = f(\A)$ as follows:
+
+		As in the previous case, we convert $\varphi(\x,\X,\z)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X,\z)$. Let each $\varphi_{i}(\x,\X,\z)$ be of the form
+		\[
+		\varphi_{i}(\x,\X,\z) = X_{s_1}(\x_{s_1},\z_{s_1}) \wedge \cdots \wedge X_{s_m}(\x_{s_m},\z_{s_m}) \wedge \neg X_{t_1}(\x_{t_1},\z_{t_1}) \wedge \cdots \wedge \neg X_{t_n}(\x_{t_n},\z_{t_1}) \wedge \alpha_i(\x,\z),
+		\]
+		where every $\x_{s_i}$, $\z_{s_i}$, $\x_{t_j}$ and $\z_{t_j}$ have the corresponding number of variables, and $\alpha(\x)$ is a FO-formula.
+
+		We generate a propositional variable $p^{\e}_{\ell,\b}$ for each $\b\in \bigcup_{j = 1}^{c+d} A^j$, for each $\e\in A^d$, and for each $\ell\in\{1,\ldots,r\}$.
+Let $\p$ be the tuple of all such variables.
+Also, let $\rho_{\ell}(\a)$ be the tuple that results of reordering $\a$ using the reordering of the variables of $(\x,\z)$ in $(\x_{\ell},\z_{\ell})$.
+
+		For each $\e\in A^d$, we follow the following procedure.
+Eor each $\varphi_i(\x,\X)$, we generate the set $\Gamma^{\e}_i = \{\a\in A^c\mid \A\models\alpha(\a,\e)\}$, and so, the DNF formula we generate is
+		\[
+		\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{s_1,\rho_{s_1}(\a,\e)} \wedge \cdots \wedge p^{\e}_{s_m,\rho_{s_m}(\a,\e)} \wedge \neg p^{\e}_{t_1,\rho_{t_1}(\a,\e)} \wedge \cdots \wedge \neg p^{\e}_{t_n,\rho_{t_n}(\a,\e)} \wedge \beta_{\e},
+		\]
+		where $\beta_{\e}$ is defined as
+		\[
+		\beta_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e}} \bigwedge_\bigvee_{\b\in A^(c+d)} \bigwedge_{\ell = 1}^r p^{\e}_{\ell,\b}.	
+		\]
+	\end{enumerate}
+\end{proof}

### 2015-09-29 tex/s-one-fo.tex
29 sep

@@
 		\[
-		\varphi_{i}(\x,\X,\z) = X_{s_1}(\x_{s_1},\z_{s_1}) \wedge \cdots \wedge X_{s_m}(\x_{s_m},\z_{s_m}) \wedge \neg X_{t_1}(\x_{t_1},\z_{t_1}) \wedge \cdots \wedge \neg X_{t_n}(\x_{t_n},\z_{t_1}) \wedge \alpha_i(\x,\z),
+		\varphi_{i}(\x,\X,\z) = X_{s_1}(\y_{s_1}) \wedge \cdots \wedge X_{s_m}(\y_{s_m}) \wedge \neg X_{t_1}(\y_{t_1}) \wedge \cdots \wedge \neg X_{t_n}(\y_{t_1}) \wedge \alpha_i(\x,\z),
 		\]
-		where every $\x_{s_i}$, $\z_{s_i}$, $\x_{t_j}$ and $\z_{t_j}$ have the corresponding number of variables, and $\alpha(\x)$ is a FO-formula.
+		where every $\y_{s_i}$ and $\y_{t_j}$ have the corresponding number of variables in $(\x,\z)$, and $\alpha(\x)$ is a FO-formula.
 
 		We generate a propositional variable $p^{\e}_{\ell,\b}$ for each $\b\in \bigcup_{j = 1}^{c+d} A^j$, for each $\e\in A^d$, and for each $\ell\in\{1,\ldots,r\}$.
-Let $\p$ be the tuple of all such variables.
-Also, let $\rho_{\ell}(\a)$ be the tuple that results of reordering $\a$ using the reordering of the variables of $(\x,\z)$ in $(\x_{\ell},\z_{\ell})$.
+We also generate a $t_{\a}$ for each $\a$ Let $\p$ be the tuple of all such variables.
+Also, let $\rho_{\ell}$ be such that $\rho_{\ell}(\x,\z) = \y_{\ell}$
 
 		For each $\e\in A^d$, we follow the following procedure.
-Eor each $\varphi_i(\x,\X)$, we generate the set $\Gamma^{\e}_i = \{\a\in A^c\mid \A\models\alpha(\a,\e)\}$, and so, the DNF formula we generate is
+For each $\varphi_i(\x,\X)$, we generate the set $\Gamma^{\e}_i = \{\a\in A^c\mid \A\models\alpha(\a,\e)\}$, and so, the DNF formula we generate is
 		\[
-		\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{s_1,\rho_{s_1}(\a,\e)} \wedge \cdots \wedge p^{\e}_{s_m,\rho_{s_m}(\a,\e)} \wedge \neg p^{\e}_{t_1,\rho_{t_1}(\a,\e)} \wedge \cdots \wedge \neg p^{\e}_{t_n,\rho_{t_n}(\a,\e)} \wedge \beta_{\e},
+		\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{s_1,\rho_{s_1}(\a,\e)} \wedge \cdots \wedge p^{\e}_{s_m,\rho_{s_m}(\a,\e)} \wedge \neg p^{\e}_{t_1,\rho_{t_1}(\a,\e)} \wedge \cdots \wedge \neg p^{\e}_{t_n,\rho_{t_n}(\a,\e)} \wedge t_{\e},
 		\]
 		where $\beta_{\e}$ is defined as
-		\[
-		\beta_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e}} \bigwedge_\bigvee_{\b\in A^(c+d)} \bigwedge_{\ell = 1}^r p^{\e}_{\ell,\b}.	
-		\]
+
 	\end{enumerate}

### 2015-10-06 tex/examples.tex
6 sep

@@
 
-Let $\L = \{P,N,\leq\}$, and $\A = \langle A, P^\A, N^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc DNF} formula $\Psi$, where $A$ is the set of variables mentioned in $\Psi$, $P$ and $N$ is a binary relations described as follows:
+Let $\L = \{P,N,\leq\}$, and $\A = \langle A, P^\A, N^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc DNF} formula $\Psi$ with $n$ variables and $m$ clauses, where $A = \{1\ldots,n+m\}$, $P$ and $N$ is a binary relations described as follows:
 \begin{eqnarray*}
@@
 
-Let $\L = \{P,N,\leq\}$, and $\A = \langle A, P^\A, N^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc HORN-SAT} formula $\Psi$, where $A$, $P$ and $N$ are defined as in the previous case.
+Let $\L = \{P,N,F,\leq\}$, and $\A = \langle A, P^\A, N^\A, F^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc HORN-SAT} formula $\Psi$, where $A$, $P$ and $N$ are defined as in the previous case, and $F$ is defined as
+\[
+	F = \{ c \mid
+		\mbox{all literals in $c$ are negative}
+	\}
+\]
 
 Now let $T$ be a unary second-order variable, and
-\begin{eqnarray*}
-	\psi(T) &=& \forall c \forall \ell\, [(P(c,\ell)\wedge\forall\ell'(N(c,\ell')\to T(\ell')))\to T(\ell)] \\
-	&\equiv& \forall c \forall \ell\, [\neg P(c,\ell)\vee \exists\ell'(N(c,\ell')\wedge \neg T(\ell'))\vee T(\ell)].
-\end{eqnarray*}
+\begin{align*}
+	\psi(T) &=& \forall c\,[ F(c) \to \exists \ell'(N(c,\ell') \wedge \neg T(\ell'))]
+	\,\wedge
+	 \forall c \forall \ell\, [\neg F(c)\to(P(c,\ell)\wedge\forall\ell'(N(c,\ell')\to T(\ell')))\to T(\ell)] \\
+	 &\equiv& \forall c\,[ \neg F(c) \vee \exists \ell'(N(c,\ell') \wedge \neg T(\ell'))]
+	 \,\wedge
+	 \forall c \forall \ell\, [ F(c) \vee \neg P(c,\ell) \vee \exists\ell'(N(c,\ell') \wedge \neg T(\ell'))) \vee T(\ell)]
+\end{align*}
 Let $\neg X(c,\ell')$ be defined as $N(c,\ell')\wedge \neg T(\ell)$. That is,
@@
 	\varphi(T,X) = \forall c \forall \ell'\,[(\neg N(c,\ell')\vee T(\ell')\vee \neg X(c,\ell'))\wedge (X(c,\ell')\vee \neg T(\ell'))\wedge (X(c,\ell')\vee N(c,\ell'))]\,\wedge \\
-	\forall c \forall \ell\, [\neg P(c,\ell)\vee \exists\ell'\neg X(c,\ell')\vee T(\ell)]	.
+	\forall c\,[ \neg F(c) \vee \exists \ell'\neg X(c,\ell'))]
+	\wedge
+	\forall c \forall \ell\, [ F(c) \vee \neg P(c,\ell) \vee \exists\ell'\neg X(c,\ell') \vee T(\ell)].
 \end{multline*}
-This can be rewritten as
-\begin{multline*}
-	\varphi'(T,X) = \forall c \forall \ell \exists\ell'\,[(\neg N(c,\ell)\vee T(\ell)\vee \neg X(c,\ell))\wedge (X(c,\ell)\vee \neg T(\ell))\wedge (X(c,\ell)\vee N(c,\ell))\,\wedge \\
-	(\neg P(c,\ell)\vee \neg X(c,\ell')\vee T(\ell))].
-\end{multline*}
-Let $\theta(c,\ell,\ell',T,X)$ be such that $\varphi'(T,X) = \forall c \forall \ell \exists\ell'\,\theta(c,\ell,\ell',T,X)$.
-Note that $\theta(c,\ell,\ell',T,X)$ is an {\sc FO}-extended quantifier-free $\L$-formula such that any second order variable only appears in a Horn clause.
-First, note that each relation $T$ over $\A$ such that $\A\models \psi(T)$ represents a valid assignment to the variables in $\Psi$, and that in $\theta(c,\ell,\ell',T,X)$, $X$ is completely defined by the assignment to $T$.
-Therefore $f_{\psi(T)}(\A') = f_{\varphi'(T,X)}(\A')$ for each $\A'\in\Truc$, from which we conclude that $f_{\forall c \forall \ell \exists\ell'\,\theta(c,\ell,\ell',T,X)}$ represents {\sc \#HORN-SAT}.
+Let $\theta(c,\ell,T,X)$ be such that $\varphi'(T,X) = \forall c \forall \ell\,\theta(c,\ell,T,X)$.
+Note that $\theta(c,\ell,T,X)$ is an {\sc FO}-extended quantifier-free $\L$-formula such that any second order variable only appears in a Horn clause.
+First, note that each relation $T$ over $\A$ such that $\A\models \psi(T)$ represents a valid assignment to the variables in $\Psi$, and that in $\theta(c,\ell,T,X)$, $X$ is completely defined by the assignment to $T$.
+Therefore $f_{\psi(T)}(\A') = f_{\varphi'(T,X)}(\A')$ for each $\A'\in\Truc$, from which we conclude that $f_{\forall c \forall \ell\,\theta(c,\ell,T,X)}$ represents {\sc \#HORN-SAT}.
+
+\subsection{{\sc \#Disj-HORN-SAT}}
+We define the decision problem
+\[
+	\textsc{Disj-HORN-SAT} = \{\Phi \mid \Phi \mbox{ is a disjunction of Horn formulas and $\Phi$ is satisfiable}\},
+\]
+and the counting problem {\sc \#Disj-HORN-SAT} as a function that counts all satisfying assignments to a formula $\Phi$ which is a disjunction of Horn formulas.
+We will now show that {\sc \#Disj-HORN-SAT} $\in\EtwoE$.
+
+Let $\L = \{P,N,F,B,\leq\}$, and $\A = \langle A, P^\A, N^\A, F^\A, B^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc Disj-HORN-SAT} formula $\Psi$ with $n$ variables, $m$ clauses in total, and $k$ Horn formulas, where $A = \{1,\ldots,n+m+k\}$, $P$, $N$ and $F$ are defined as in the previous case, and $B$ is defined as:
+\[
+	B = \{ (c,f) \mid 
+		\mbox{the clause $c$ is in the Horn formula $f$}
+	\}.
+\]
+Now let $T$ be a unary second order variable, and
+\[
+	\psi(T) = \exists f\forall c\,(B(c,f)\to \forall\ell\,\theta(c,\ell,T,X)),
+\]
+where $\theta$ is defined as in the previous case.

### 2015-10-09 tex/classes.tex
22 sep
8 sep

@@
-\section{Extended Horn counting hierarchies}
+\section{The Extended Horn Counting Hierarchy}
 
-We define syntactically the classes $\EiE$ and $\EiU$ as following. First we define Horn clauses.
+We define syntactically the classes $\EiE$ and $\EiU$ as follows. First we define extended Horn clauses.
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $\x$ is a tuple with the corresponding number of variables. Now we define the classes' syntaxis inductively.
+where $\x$ is a tuple with the corresponding number of variables. Now we define the syntax of the classes inductively.
 \begin{enumerate}
@@
 	We will now prove that there is a polynomial time algorithm that decides $L_{f}$.
-This is equivalent to decide, given $\A= \langle A, \S\A, \leq^\A \rangle\in\Truc$, whether there exist assignments $\P,\z$ such that $\A\models\varphi(\P,\z)$.
+This is equivalent to decide, given $\A= \langle A, \S^\A, \leq^\A \rangle\in\Truc$, whether there exist assignments $\P,\z$ such that $\A\models\varphi(\P,\z)$.
 This is the same as $\A\models\exists\X\exists\x\,\varphi(\X,\x)$.
@@
 First, we replace every instance of $\v$ by a disjunction of all $r$-tuples in $A^r$, where $r$ is the number of variables in $\v$.
-The remainder is still a {\sc Horn}-clause.
+The result is still a Horn clause.
 Second, we replace every instance of $\u$ by a conjunction of all $s$-tulples in $A^s$, where $s$ is the number or variables in $\u$.
-The remainder is still a conjunction of {\sc Horn}-clauses.
+The result is still a conjunction of Horn clauses.
 

### 2015-10-09 tex/examples.tex
8 sep

@@
 
-\subsection{{\sc \#3-HORN-SAT}}
-We will now show that {\sc \#3-HORN-SAT} $\in\EoneU$.\\
+\subsection{{\sc \#3-Horn-Sat}}
+We will now show that {\sc \#3-Horn-Sat} $\in\EoneU$.\\
 
-Let $\L = \{P_1,P_2,P_3,N_1,N_2,N_3,\leq\}$, and $\A = \langle A, P_1^\A,P_2^\A,P_3^\A,N_1^\A,N_2^\A,N_3^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc 3-HORN-SAT} formula $\Psi$, where $A$ is the set of variables mentioned in $\Psi$, and the relations in $\L$ are interpreted as follows:
+Let $\L = \{P_1,P_2,P_3,N_1,N_2,N_3,\leq\}$, and $\A = \langle A, P_1^\A,P_2^\A,P_3^\A,N_1^\A,N_2^\A,N_3^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc 3-Horn-Sat} formula $\Psi$, where $A$ is the set of variables mentioned in $\Psi$, and the relations in $\L$ are interpreted as follows:
 \begin{eqnarray*}
@@
 We can easily reorder $\varphi(T)$ as $\forall x \forall y \forall z\,\theta(x,y,z,T)$ where $\theta(x,y,z,T)$ is an {\sc FO}-extended quantifier-free $\L$-formula such that any second order variable only appears in a Horn clause.
-It is easily seen that each relation $T$ over $\A$ such that $\A\models\forall x \forall y \forall z\,\theta(x,y,z,T)$ represents a valid assignment to the variables in $\Psi$, and therefore, $f_{\forall x \forall y \forall z\,\theta(x,y,z,T)}$ represents {\sc \#3-HORN-SAT}.
+It is easily seen that each relation $T$ over $\A$ such that $\A\models\forall x \forall y \forall z\,\theta(x,y,z,T)$ represents a valid assignment to the variables in $\Psi$, and therefore, $f_{\forall x \forall y \forall z\,\theta(x,y,z,T)}$ represents {\sc \#3-Horn-Sat}.
 
@@
 
-\subsection{{\sc \#HORN-SAT}}
-We will now show that {\sc \#HORN-SAT} $\in\EoneU$.\\
+\subsection{{\sc \#Horn-Sat}}
+We will now show that {\sc \#Horn-Sat} $\in\EoneU$.\\
 
-Let $\L = \{P,N,F,\leq\}$, and $\A = \langle A, P^\A, N^\A, F^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc HORN-SAT} formula $\Psi$, where $A$, $P$ and $N$ are defined as in the previous case, and $F$ is defined as
+Let $\L = \{P,N,F,\leq\}$, and $\A = \langle A, P^\A, N^\A, F^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc Horn-Sat} formula $\Psi$, where $A$, $P$ and $N$ are defined as in the previous case, and $F$ is defined as
 \[
@@
 First, note that each relation $T$ over $\A$ such that $\A\models \psi(T)$ represents a valid assignment to the variables in $\Psi$, and that in $\theta(c,\ell,T,X)$, $X$ is completely defined by the assignment to $T$.
-Therefore $f_{\psi(T)}(\A') = f_{\varphi'(T,X)}(\A')$ for each $\A'\in\Truc$, from which we conclude that $f_{\forall c \forall \ell\,\theta(c,\ell,T,X)}$ represents {\sc \#HORN-SAT}.
+Therefore $f_{\psi(T)}(\A') = f_{\varphi'(T,X)}(\A')$ for each $\A'\in\Truc$, from which we conclude that $f_{\forall c \forall \ell\,\theta(c,\ell,T,X)}$ represents {\sc \#Horn-Sat}.
 
-\subsection{{\sc \#Disj-HORN-SAT}}
+\subsection{{\sc \#Disj-Horn-Sat}}
 We define the decision problem
 \[
-	\textsc{Disj-HORN-SAT} = \{\Phi \mid \Phi \mbox{ is a disjunction of Horn formulas and $\Phi$ is satisfiable}\},
+	\textsc{Disj-Horn-Sat} = \{\Phi \mid \Phi \mbox{ is a disjunction of Horn formulas and $\Phi$ is satisfiable}\},
 \]
-and the counting problem {\sc \#Disj-HORN-SAT} as a function that counts all satisfying assignments to a formula $\Phi$ which is a disjunction of Horn formulas.
-We will now show that {\sc \#Disj-HORN-SAT} $\in\EtwoE$.
+and the counting problem {\sc \#Disj-Horn-Sat} as a function that counts all satisfying assignments to a formula $\Phi$ which is a disjunction of Horn formulas.
+We will now show that {\sc \#Disj-Horn-Sat} $\in\EtwoE$.
 
-Let $\L = \{P,N,F,B,\leq\}$, and $\A = \langle A, P^\A, N^\A, F^\A, B^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc Disj-HORN-SAT} formula $\Psi$ with $n$ variables, $m$ clauses in total, and $k$ Horn formulas, where $A = \{1,\ldots,n+m+k\}$, $P$, $N$ and $F$ are defined as in the previous case, and $B$ is defined as:
+Let $\L = \{P,N,F,B,\leq\}$, and $\A = \langle A, P^\A, N^\A, F^\A, B^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a {\sc Disj-Horn-Sat} formula $\Psi$ with $n$ variables, $m$ clauses in total, and $k$ Horn formulas, where $A = \{1,\ldots,n+m+k\}$, $P$, $N$ and $F$ are defined as in the previous case, and $B$ is defined as:
 \[

### 2015-10-09 tex/s-one-fo.tex
8 sep

@@
 		\item All free variables in $\psi$ are of first order.
-Let $\psi = \exists x\,\varphi(\x,\z)$ where $\x = (x_1,\ldots,x_c)$ and $\z = (z_1,\ldots,z_d)$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a propositional formula $\Phi = f(\A)$.
+Let $\psi = \exists \x\,\varphi(\x,\z)$ where $\x = (x_1,\ldots,x_c)$ and $\z = (z_1,\ldots,z_d)$.
+Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a propositional formula $\Phi = h(\A)$.
 First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\,\varphi(\x,\e)$.
@@
 		which has exactly $n$ possible assignments.
+
 		\item There is only second-order free variables in $\psi$.
-Let $\psi = \exists x\,\varphi(\x,\X)$, where $\x = (x_1,\ldots,x_c)$ and $\X = (X_1,\ldots,X_r)$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a DNF-formula $\Phi = f(\A)$ as follows:
+Let $\psi = \exists \x\,\varphi(\x,\X)$, where $\x = (x_1,\ldots,x_c)$ and $\X = (X_1,\ldots,X_r)$.
+Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a DNF-formula $\Phi = h(\A)$ as follows:
 
@@
 We can conclude that $f(\A) = f_{\textsc{DNF}}(\Phi)$.
-		\item Let $\psi = \exists x\,\varphi(\x,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\X = (X_1,\ldots,X_r)$ and $\z = (z_1,\ldots,z_d)$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$ we generate a DNF-formula $\Phi = f(\A)$ as follows:
+
+		\item Let $\psi = \exists \x\,\varphi(\x,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
+Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$ we generate a DNF-formula $\Phi = h(\A)$ as follows:
 
@@
 
-		We generate a propositional variable $p^{\e}_{\ell,\b}$ for each $\b\in \bigcup_{j = 1}^{c+d} A^j$, for each $\e\in A^d$, and for each $\ell\in\{1,\ldots,r\}$.
-We also generate a $t_{\a}$ for each $\a$ Let $\p$ be the tuple of all such variables.
-Also, let $\rho_{\ell}$ be such that $\rho_{\ell}(\x,\z) = \y_{\ell}$
+		We generate a propositional variable $p^{\e}_{\ell,\b}$ for each $\e\in A^d$, for each $\ell\in\{1,\ldots,r\}$, and for each $\b\in A^{c_{\ell}}$.
+Let $\p$ be the tuple of all such variables.
+We also generate a set $T = \{t_1,\ldots,t_{L}\}$ where $L = \lceil \log\vert A^d\vert \rceil$, and an arbitrary inyective function $\beta:A^d \to \{0,1\}^L$ where $\beta(\e) = (b^{\e}_1,\ldots,b^{\e}_L)$ for each $\e\in A^d$.
+Also, let $\rho_{\ell}$ be such that $\rho_{\ell}(\x,\z) = \y_{\ell}$.
 
 		For each $\e\in A^d$, we follow the following procedure.
-For each $\varphi_i(\x,\X)$, we generate the set $\Gamma^{\e}_i = \{\a\in A^c\mid \A\models\alpha(\a,\e)\}$, and so, the DNF formula we generate is
+For each $\varphi_i(\x,\X)$, we consider the set $\Gamma^{\e}_i = \{\a\in A^c\mid \A\models\alpha(\a,\e)\}$, and so, the DNF formula we generate is
 		\[
-		\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{s_1,\rho_{s_1}(\a,\e)} \wedge \cdots \wedge p^{\e}_{s_m,\rho_{s_m}(\a,\e)} \wedge \neg p^{\e}_{t_1,\rho_{t_1}(\a,\e)} \wedge \cdots \wedge \neg p^{\e}_{t_n,\rho_{t_n}(\a,\e)} \wedge t_{\e},
+		\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{s_1,\rho_{s_1}(\a,\e)} \wedge \cdots \wedge p^{\e}_{s_m,\rho_{s_m}(\a,\e)} \wedge \neg p^{\e}_{t_1,\rho_{t_1}(\a,\e)} \wedge \cdots \wedge \neg p^{\e}_{t_n,\rho_{t_n}(\a,\e)} \wedge \gamma_{\e} \wedge \theta_{\e}
 		\]
-		where $\beta_{\e}$ is defined as
+		where $\gamma_{\e}$ is defined as
+		\[
+			\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\b\in \bigcup_{j = 1}^{c+d} A^j} \bigwedge_{\ell = 1}^r p_{\ell,\b}^{\e'},
+		\]
+		and $\theta_{\e}$ is defined as
+		\[
+			\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
+		\]
+		Finally, we return $f(\A) = \Phi = \bigvee_{i = 1}^k \bigvee_{\e\in A^d} \Phi_i^{\e}$.
 
+		We will now show a function $g:\{\langle\P,\e\rangle \mid \A \models \exists\x\,\varphi(\x,\P,\e)\} \to \{\sigma \mid \sigma(\Phi) = 1\}$ and then prove that it is a bijection.
+Let $g$ be such that for each assignment $\langle \P,\e \rangle$, it holds that $g(\langle \P,\e \rangle) = \sigma$ where $\sigma$ is constructed as follows:
+
+		For each $t_i\in T$, let $\sigma(t_i) = b^{\e}_i$.
+For each $p_{\ell,\b}^{\e}$, where $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e}) = 1$ if and only if $\b \in P_{\ell}$.
+For each $p_{\ell,\b}^{\e'}$, where $\e'\in A^d\setminus\{\e\}$,  $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e'}) = 1$.
+Let $i$ be such that $\A\models \exists\x\,\varphi_i(\x,\P,\e)$.
+It is simple to see that from the way we constructed $\Phi_i^{\e}$, it holds that $\sigma(\Phi_i^{\e}) = 1$.
+Also, note that every different assignment $\langle \P',\e' \rangle$, it holds that $\sigma(\langle \P',\e' \rangle) \neq \sigma(\langle \P,\e \rangle)$.
+Therefore, $g$ is an injective function.
+Now, let $\sigma$ be an arbitrary assignment such that $\sigma(\Phi) = 1$.
+Let $i$ and $\e$ be such that $\sigma(\Phi^{\e}_i)$.
+Note that for every $\e'\in A^d\setminus\{\e\}$, it holds that $\sigma(\theta_{\e'}) = 0$, so there is exactly one possible $\e$ associated to $\sigma$.
+Consider the assignment $\P$ where $\b\in P_{\ell}$ if and only if $\sigma(p^{\e}_{\ell,\b}) = 1$, for each $\ell$ and $\b \in A^{c_{\ell}}$.
+Since there is at least one $\a\in\Gamma_i$ for which $\sigma$ satisfies its respective subformula, it holds that $\A\models\varphi_i(\a,\P,\e)$, which implies that $\A\models\exists\x\,\varphi(\x,\P,\e)$.
+From this we deduce that $g$ is surjective, and therefore a bijection.
+We conclude that $f(\A) = f_{\textsc{DNF}}(\Phi)$.
 	\end{enumerate}

### 2015-10-09 tex/examples.tex
9 oct

@@
 \subsection{{\sc \#Disj-Horn-Sat}}
-We define the decision problem
-\[
-	\textsc{Disj-Horn-Sat} = \{\Phi \mid \Phi \mbox{ is a disjunction of Horn formulas and $\Phi$ is satisfiable}\},
-\]
-and the counting problem {\sc \#Disj-Horn-Sat} as a function that counts all satisfying assignments to a formula $\Phi$ which is a disjunction of Horn formulas.
 We will now show that {\sc \#Disj-Horn-Sat} $\in\EtwoE$.

### 2015-10-09 tex/first_h.tex
9 oct

--- 2015-10-09 tex/s-one-fo.tex
@@
-\section{Some properties of $\E1$}
+\section{Extended Logic Counting Classes}
 
@@
 \end{eqnarray*}
-where $x,y$ and $x_i$ are first order variables for every $i$. We now define an {\sc FO}-extended quantifier-free $\L$-formula as follows:
+where $x,y$ and $x_i$ are first order variables for every $i$. For every logic $\mathcal{F}$, we define an $\mathcal{F}$-extended quantifier-free $\L$-formula as follows:
 \begin{eqnarray*}
-	\varphi &::=& \alpha, \alpha \text{ is an FO-formula over } \L  \ \mid \\
+	\varphi &::=& \alpha, \alpha \text{ is an $\mathcal{F}$-formula over } \L  \ \mid \\
 	&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
@@
 
-A function $f:\Truc \to \N$ is in $\Eo$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\Y,\y)$ such that $f = f_{\varphi(\Y,\y)}$.
+\subsection{The First {\sc FO}-Extended Counting Hierarchy}
 
-Similarly, a function $f:\Truc \to \N$ is in $\E1$ if there exists an {\sc FO}-extended quantifier-free $\L$-formula $\varphi(\x,\Y,\y)$ such that $f = f_{\exists \x \: \varphi(\x,\Y,\y)}.$\\
+We define syntactically the classes $\Eie$ and $\Eiu$ according to the following grammar:
+\begin{enumerate}
+	\item $\Eezero$:
+	\[
+		E_0 ::= \varphi \mid \varphi \mbox{ is an {\sc FO}-extended quantifier-free $\L$-formula }.
+	\]
+	\item $\Euzero$:
+		\[U_0 ::= E_0.\]
+	\item $\Eipe$:
+	\[	E_{i+1} ::= U_i \ \mid \ \exists x \, E_{i+1}.\]
+	\item $\Eipu$:
+		\[U_{i+1} ::= E_i \ \mid \ \forall x \, U_{i+1}.\]
+\end{enumerate}
+
+A function $f$ is in $\Eie$ (resp. $\Eiu$) if there is an $\L$-formula $\varphi$ defined by the grammar $E_i$ (resp. $U_i$) such that $f = f_{\varphi}$.
 
@@
 
-		\item There is only second-order free variables in $\psi$.
+		\item There are only second-order free variables in $\psi$.
 Let $\psi = \exists \x\,\varphi(\x,\X)$, where $\x = (x_1,\ldots,x_c)$ and $\X = (X_1,\ldots,X_r)$.

### 2015-10-09 tex/horn_h.tex
9 oct

--- 2015-10-09 tex/classes.tex
@@
 \end{proof}
+
+We define the decision problem
+\[
+\textsc{Disj-Horn-Sat} = \{\Phi \mid \Phi \mbox{ is a disjunction of Horn formulas and $\Phi$ is satisfiable}\},
+\]
+and the counting problem {\sc \#Disj-Horn-Sat} as a function that counts all satisfying assignments to a formula $\Phi$ which is a disjunction of Horn formulas.
+
+\begin{theo}
+	$\textsc{Disj-Horn-Sat}$ is hard for $\EtwoE$ under parsimonious reductions. 
+\end{theo} 
+\begin{proof}
+	Let $f$ be an arbitrary function in $\EtwoE$ and let $f_{\textsc{Disj-Horn-Sat}}$ be the function that characterizes {\sc \#Disj-Horn-Sat}.
+We will now show a function $h:\Truc\to L(P)$ such that for every $\A\in\Truc$, it holds that $f(\A) = f_{\textsc{Disj-Horn-Sat}}(h(\A))$, and which can be computed in polynomial time.
+Let $\psi$ be such that $f = f_{\psi}$.
+We separate the proof in three cases:\begin{enumerate}
+		\item All free variables in $\psi$ are of first order.
+Let $\psi = \exists \x \forall \y\,\varphi(\x,\y,\z)$ where $\z = (z_1,\ldots,z_d)$.
+Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a propositional formula $\Phi = h(\A)$.
+First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\forall \y\,\varphi(\x,\y,\e)$.
+Let $n$ be the number of such assignments to $\z$.
+Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$.
+The Disj-Horn-Sat formula we return is
+		\[
+		\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
+		\]
+		which has exactly $n$ possible assignments. Note that each $p_i$ and $\neg p_i$ is a Horn clause itself.
+
+		\item There are only second-order free variables in $\psi$.
+Let $\psi = \exists \x\forall \y\,\varphi(\x,\y,\X)$, where $\x = (x_1,\ldots,x_c)$, $\y = (y_1,\ldots,y_b)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
+Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a Disj-Horn-Sat-formula $\Phi = h(\A)$ as follows:
+
+		First, Let $\varphi(\x,\y,\X) = \bigwedge_{i = 1}^k \varphi_{i}(\x,\y,\X)$ where, without loss of generality, each $\varphi_{i}(\x,\y,\X)$ is of the form
+		\[
+		\varphi_{i}(\x,\y,\X) = X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n}) \vee \alpha_i(\x,\y),
+		\]
+		where $\u_{s}$ and every $\u_{t_j}$ have the corresponding number of variables in $(\x,\y)$, and $\alpha(\x,\y)$ is a FO-formula.
+For the rest of the proof, $X_s(\u_s)$ can be absent and it does not affect the proof.
+Note that this formula is equivalent to
+		\[
+		\neg\alpha_i(\x,\y) \to (X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n})),
+		\]
+
+		Then, for each $\a\in A^c$ let $\Gamma_i^{\a} = \{\b\in A^{b}\mid \A\models\neg\alpha(\a,\b)\}$.
+Note that this set can be generated in polynomial time.
+For each $\ell\in\{1,\ldots,r\}$, we generate a propositional variable $p^{\ell}_{\d}$ for each $\d\in A^{c_{\ell}}$.
+Let $\p$ be the tuple of all such variables.
+Also, let $\rho_{\ell}(\x,\y) = \u_{\ell}$, and $\rho_{\ell}(\d)$ be the tuple that results of reordering $\d$ using the reordering of the variables of $(\x,\y)$ in $\u_{\ell}$.
+
+		For each $\a\in A^c$ we generate a Horn formula
+		\[
+		\Phi_{\a}^i = \bigwedge_{\b\in\Gamma_i} p^{s}_{\rho_s(\a,\b)} \vee \neg p^{t_1}_{\rho_{t_1}(\a,\b)} \vee \cdots \vee \neg p^{t_n}_{\rho_{t_n}(\a,\b)}.
+		\]
+		and finally we return $\Phi = \bigvee_{\a\in A^c} \bigwedge_{i = 1}^k \Phi_{\a}^i$.
+
+		Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\forall\y\,\varphi(\x,\y,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$.
+We can conclude that $f(\A) = f_{\textsc{Disj-Horn-Sat}}(\Phi)$.
+
+		\item 
+	\end{enumerate}
+\end{proof}

### 2015-10-20 tex/first_h.tex
20 oct

@@
 Let $\p$ be the tuple of all such variables.
-We also generate a set $T = \{t_1,\ldots,t_{L}\}$ where $L = \lceil \log\vert A^d\vert \rceil$, and an arbitrary inyective function $\beta:A^d \to \{0,1\}^L$ where $\beta(\e) = (b^{\e}_1,\ldots,b^{\e}_L)$ for each $\e\in A^d$.
+We also generate a set $T = \{t_1,\ldots,t_{L}\}$ where $L = \lceil \log\vert A^d\vert \rceil$, and an arbitrary injective function $\beta:A^d \to \{0,1\}^L$ where $\beta(\e) = (b^{\e}_1,\ldots,b^{\e}_L)$ for each $\e\in A^d$.
 Also, let $\rho_{\ell}$ be such that $\rho_{\ell}(\x,\z) = \y_{\ell}$.
@@
 		\[
-			\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\b\in \bigcup_{j = 1}^{c+d} A^j} \bigwedge_{\ell = 1}^r p_{\ell,\b}^{\e'},
+			\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
 		\]
@@
 		\]
-		Finally, we return $f(\A) = \Phi = \bigvee_{i = 1}^k \bigvee_{\e\in A^d} \Phi_i^{\e}$.
+		Finally, we return $f(\A) = \Phi = \bigvee_{\e\in A^d} \bigvee_{i = 1}^k \Phi_i^{\e}$.
 

### 2015-10-20 tex/horn_h.tex
20 oct

@@
 Let $\psi = \exists \x\forall \y\,\varphi(\x,\y,\X)$, where $\x = (x_1,\ldots,x_c)$, $\y = (y_1,\ldots,y_b)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a Disj-Horn-Sat-formula $\Phi = h(\A)$ as follows:
+Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a Disj-Horn formula $\Phi = h(\A)$ as follows:
 
@@
 
-		\item 
+		\item Let $\psi = \exists \x\forall\y\,\varphi(\x,\y,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\y = (y_1,\ldots,y_b)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
+Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$ we generate a Disj-Horn formula $\Phi = h(\A)$ as follows:
+
+		As in the previous case, let $\varphi(\x,\y,\X,\z) = \bigwedge_{i = 1}^k$ where $\varphi(\x,\y,\X,\z)$ is of the form:
+		\[
+		\varphi_{i}(\x,\y,\X,\z) = X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n}) \vee \alpha_i(\x,\y,\z),
+		\]
+		where $\u_{s}$ and every $\u_{t_j}$ have the corresponding number of variables in $(\x,\y,\z)$, and $\alpha(\x,\y,\z)$ is a FO-formula.
+For the rest of the proof, $X_s(\u_s)$ can be absent and it does not affect the proof.
+Note that this formula is equivalent to
+		\[
+		\neg\alpha_i(\x,\y,\z) \to (X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n})).
+		\]
+
+		For each $\e\in A^d$, we do the following.
+For each $\a\in A^c$ let $\Gamma^{\e}_{i,\a} = \{\b \in A^b \mid \A\models \neg\alpha_i(\a,\b,\e) \}$.
+Note that this set can be generated in polynomial time.
+For each $\ell\in\{1,\ldots,r\}$ we generate a propositional variable $p^{\e}_{\ell,\d}$ for each $\d\in A^{c_{\ell}}$.
+Let $\p$ be the tuple of all such variables.
+We also generate a set $T = \{t_1,\ldots,t_{L}\}$ where $L = \lceil \log\vert A^d\vert \rceil$, and an arbitrary injective function $\beta:A^d \to \{0,1\}^L$ where $\beta(\e) = (b^{\e}_1,\ldots,b^{\e}_L)$ for each $\e\in A^d$.
+Also, let $\rho_{\ell}(\x,\y,\z) = \u_{\ell}$, and $\rho_{\ell}(\d)$ be the tuple that results of reordering $\d$ using the order of the variables in $(\x,\y,\z)$ in $\u_{\ell}$.
+And so, the Horn formula we generate is
+		\[
+			\Phi^{\e}_{i,\a} = \bigwedge_{\a\in\Gamma^{\e}_i} p^{\e}_{s,\rho_{s}(\a,\b,\e)} \vee p^{\e}_{t_1,\rho_{t_1}(\a,\b,\e)} \vee \cdots \vee p^{\e}_{t_n,\rho_{t_n}(\a,\b,\e)}.
+		\]
+		And finally we return $\Phi = \bigvee_{\e\in A^d}\bigvee_{\a\in A^c} (\bigwedge_{i=1}^k \Phi_{\e}^{i,\a} \wedge \gamma_{\e} \wedge \theta_{\e} )$, where $\gamma_{\e}$ is defined as
+		\[
+		\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
+		\]
+		and $\theta_{\e}$ is defined as
+		\[
+		\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
+		\]
+
 	\end{enumerate}

### 2015-11-02 tex/first_h.tex
2 nov

@@
 	\[
-		E_0 ::= \varphi \mid \varphi \mbox{ is an {\sc FO}-extended quantifier-free $\L$-formula }.
+		E_0 ::= \varphi , \varphi \mbox{ is an {\sc FO}-extended quantifier-free $\L$-formula }.
 	\]
@@
 	for each $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, where $\e \in A^m$ and $\P = (P_1,\ldots,P_q)$ is a tuple of predicates.
-We will now show that counting $f(\A)$ can be done in polynomial time.
+We will now show that computing $f(\A)$ can be done in polynomial time.
 
@@
 Note that for each tuple of predicates $\P$ and each $\e \in A^m$, $\A \models \varphi(\P,\e)$ if and only if $\A^\prime \models \psi(\P,\e)$, and so, $g(\A^\prime) = f(\A)$.
-However, $\psi(\X,\x)$ is a quantifier-free $\L$-formula, and therefore, $g \in \#\Sigma_0$.
+But, $\psi(\X,\x)$ is a quantifier-free $\L$-formula, and therefore, $g \in \#\Sigma_0$.
 Since it is shown in \cite{DBLP:journals/jcss/SalujaST95} that $\#\Sigma_0 \subseteq$ {\sc FP}, we conclude that $f(\A)$ can be evaluated in polynomial time.
@@
 	where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables.
-As said in \cite{DBLP:journals/jcss/SalujaST95}, a function $h:\Truc \to \N$ is in $\#\Sigma_0$ if there exists a quantifier-free $\L$-formula $\theta(\Y,\y)$ such that $h = f_{\theta(\Y,\y)}$.
 Let $\y = (\x,\z)$ and let $\psi(\X,\y) = \varphi(\x,\X,\z)$.
 Moreover, let $g = f_{\psi(\X,\y)}$.
-	\begin{claim}
+	\begin{claim} \label{mono}
 		For each $\A \in \Truc$, $f(\A) > 0$ iff $g(\A) > 0$.
@@
 Therefore, $g(\A) > 0$.
+
 		($\Leftarrow$) Suppose $g(\A) > 0$.
 Let $\Q$ and $\c = (\c_1,\c_2)$, where $\c_1$ and $\c_2$ have $k$ and $m$ elements respectively, be such that $\A \models \psi(\Q,\c) = \varphi(\c_1,\Q,\c_2)$.
-Therefore, $\A \models \exists \x \, \varphi(\x,\Q,\c_2)$, from which we conclude that $f(\A) > 0$.
+Then we have that $\A \models \exists \x \, \varphi(\x,\Q,\c_2)$, from which we conclude that $f(\A) > 0$.
 	\end{proof}
 	Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula, so $g \in\Eo$.
-As we showed, for each $\A \in \Truc$, it holds that $\A \in L_f$ if and only if $g(\A) > 0$, but as it is shown in Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time.
-From this we conclude that the decision version of $f$ is in {\sc P}.
+It follows from \ref{mono} that for each $\A \in \Truc$, it holds that $\A \in L_f$ if and only if $g(\A) > 0$.
+But by Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time.
+From this we conclude that the decision version of $f$ is in $\ptime$.
 \end{proof}
@@
 \begin{proof}
-	Suppose that $\E1$ is closed under substraction, that is, for each pair of functions $f,g\in \E1$, there is an $h\in\E1$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\Truc$.
+	Suppose that $\E1$ is closed under substraction, that is, for each pair of functions $f,g\in \E1$, there exists $h\in\E1$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\Truc$.
 
-	Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\psi$, where $A$ is the set of variables mentioned in $\psi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{1,2,3,4\}$:
+	Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\Phi$, where $A$ is the set of variables mentioned in $\Phi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{1,2,3,4\}$:
 	\begin{eqnarray*}
-		S_1^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
-		S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
-		S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\psi\},\\
-		S_4^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\psi\}.
+		S_1^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
+		S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
+		S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
+		S_4^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
 	\end{eqnarray*}
-	Now let $f_{\#3DNF}$ be a function that counts the satisfying assignments to a 3DNF formula $\psi$. As shown in [1], $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
+	Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
+	\begin{multline*}
+	\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
+	\end{multline*}
+	Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
 	$$
-	\varphi(x,X) = (X(x) \vee \neg X(x)).
+	\varphi(x,X) = (T(x) \vee \neg T(x)).
 	$$
 	Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-As we supposed initially, let $h\in\E1$ such that $f_{all}-f_{\#3DNF} = h$.
-For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is {\sc co-NP}-complete.
-However, as we showed previously in Theorem \ref{decisionptime}, $L_h \in \textsc{P}$.
-Then, $\textsc{co-NP} \subseteq \textsc{P}$, from which we conclude that $\textsc{P} = \textsc{NP}$.
+Given that $f_{\#3DNF}, f_{all} \in \E1$, we have by our initial assumption that $f_{all}-f_{\#3DNF} \in \E1$.
+Let $h\in\E1$ be such that $h = f_{all}-f_{\#3DNF}$.
+For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is $\conp$-complete.
+However, as we showed previously in Theorem \ref{decisionptime}, since $h\in\E1$, we have that $L_h \in \ptime$.
+Then, $\conp \subseteq \ptime$, from which we conclude that $\ptime = \np$.
 \end{proof}
@@
 \begin{proof}
-	Let $f$ be an arbitrary function in $\E1$ and let $f_{\textsc{DNF}}$ be the function that characterizes {\sc \#DNF}.
-We will now show a function $h:\Truc\to L(P)$ such that for every $\A\in\Truc$, it holds that $f(\A) = f_{\textsc{DNF}}(h(\A))$, and which can be computed in polynomial time.
-Let $\psi$ be such that $f = f_{\psi}$.
+	Let $f$ be an arbitrary function in $\E1$ and let $f_{\textsc{\#DNF}}$ be the function that defines {\sc \#DNF}.
+We will now show a function $h:\Truc\to \textsc{Struct}[\L_{\textsc{DNF}}]$, where $\L_{\textsc{DNF}} = \{N,P\}$, such that for every $\A\in\Truc$, it holds that $f(\A) = f_{\textsc{\#DNF}}(h(\A))$, and which can be computed in polynomial time.
+For every DNF formula $\Phi$ we define $\A_{\Phi} \in \textsc{Struct}[\L_{\textsc{DNF}}]$ as stated in \ref{dnfdef}.
+Let $\psi$ be such that $f = f_{\psi(\X,\z)}$.
 We separate the proof in three cases:
 	\begin{enumerate}
-		\item All free variables in $\psi$ are of first order.
-Let $\psi = \exists \x\,\varphi(\x,\z)$ where $\x = (x_1,\ldots,x_c)$ and $\z = (z_1,\ldots,z_d)$.
+		\item $\X = ()$.
+Let $\psi(\z) = \exists \x\,\varphi(\x,\z)$ where $\x = (x_1,\ldots,x_c)$ and $\z = (z_1,\ldots,z_d)$.
 Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a propositional formula $\Phi = h(\A)$.
@@
 Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$.
-The DNF formula we return is
+The DNF formula we generate is
 		\[
@@
 		\]
-		which has exactly $n$ possible assignments.
+		which has exactly $n$ possible assignments. Finally, we return $h(\A) = \A_{\Phi}$.
 
-		\item There are only second-order free variables in $\psi$.
-Let $\psi = \exists \x\,\varphi(\x,\X)$, where $\x = (x_1,\ldots,x_c)$ and $\X = (X_1,\ldots,X_r)$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a DNF-formula $\Phi = h(\A)$ as follows:
+		\item $\z = ()$.
+Let $\psi(\X) = \exists \x\,\varphi(\x,\X)$, where $\x = (x_1,\ldots,x_c)$ and $\X = (X_1,\ldots,X_r)$.
+Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$, we generate a DNF-formula $\Phi = h(\A)$ as follows.
+First, we convert $\varphi(\x,\X)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X)$.
+Let each $\varphi_{i}(\x,\X)$ be of the form
+		\[
+		\varphi_{i}(\x,\X) = X_{\pi_i(1)}(\x^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\x^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\x^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\x^{\nu}_{i,n_i}) \wedge \alpha_i(\x),
+		\]
+		where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$) is the rearrangement of $\x$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x)$ is a FO formula.
+We also define the tuple $\a^{\pi}_{i,j}$ ($\a^{\nu}_{i,j}$) as $\a$ rearranged in the same way as $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$), for each $\a\in A^c$.
 
-		First, we convert $\varphi(\x,\X)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X)$. Let each $\varphi_{i}(\x,\X)$ be of the form
+		We generate a propositional variable $p^{\ell}_{\a}$ for each $\a\in A^{c_{\ell}}$, and for each $\ell\in\{1,\ldots,r\}$, which will indicate if the tuple $\a$ is in the assignment to $X_{\ell}$.
+Let $\p$ be the tuple of all such variables.
+
+		Then, let $\Gamma_i = \{\a\in A^c\mid \A\models\alpha_i(\a)\}$.
+There is exactly $\vert A \vert^c$ tuples to check, and each of those checks can be done in polynomial time, so this set can also be generated in polynomial time.
+
+		Finally, for each $\varphi_i(\x,\X)$ the DNF formula we generate is
 		\[
-		\varphi_{i}(\x,\X) = X_{s_1}(\x_{s_1}) \wedge \cdots \wedge X_{s_m}(\x_{s_m}) \wedge \neg X_{t_1}(\x_{t_1}) \wedge \cdots \wedge \neg X_{t_n}(\x_{t_n}) \wedge \alpha_i(\x),
+		\Phi_i = \bigvee_{\a\in\Gamma_i} p^{\pi_i(1)}_{\a^{\pi}_{i,1}} \wedge \cdots \wedge p^{\pi_i(m_i)}_{\a^{\pi}_{i,m_i}} \wedge \neg p^{\nu_i(1)}_{\a^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\pi_i(n_i)}_{\a^{\nu}_{i,n_i}}.
 		\]
-		where every $\x_{s_i}$ and $\x_{t_j}$ have the corresponding number of variables, and $\alpha(\x)$ is a FO-formula.
+		and $\Phi = \bigvee_{i = 1}^k \Phi_i \vee \bigvee_{p\in\p} (p\wedge\neg p)$. Finally, we return $h(\A) = \A_{\Phi}$.
 
-		Then, let $\Gamma_i = \{\a\in A^c\mid \A\models\alpha(\a)\}$.
-There is exactly $\vert A \vert^c$ tuples to check, and each of those checks can be done in polynomial time, so this set can also be generated in polynomial time.
-We generate a propositional variable $p^{\ell}_{\a}$ for each $\a\in \bigcup_{j = 1}^c A^j$, and for each $\ell\in\{1,\ldots,r\}$.
-Let $\p$ be the tuple of all such variables.
-Also, let $\rho_{\ell}(\a)$ be the tuple that results of reordering $\a$ using the reordering of the variables of $\x$ in $\x_{\ell}$.
-
-		Finally, for each $\varphi_i(\x,\X)$ the DNF formula we return is
-		\[
-		\Phi_i = \bigvee_{\a\in\Gamma_i} p^{s_1}_{\rho_{s_1}(\a)} \wedge \cdots \wedge p^{s_m}_{\rho_{s_m}(\a)} \wedge \neg p^{t_1}_{\rho_{t_1}(\a)} \wedge \cdots \wedge \neg p^{t_n}_{\rho_{t_n}(\a)}.
-		\]
-		and $\Phi = \bigvee_{i = 1}^k \Phi_i$.
+		As an example, let $\psi^{*}(X_1,X_2) = \exists x \exists y\, (X_1(x) \wedge X_2(y,x) \wedge \neg X_1(y) \wedge x < y \wedge \neg\exists z\,(x<z\wedge z<y))$.
+We will compute $h(\B)$, where $\B = \langle \{1,2,3\}, < \rangle$, and we have that $1 < 2 < 3$.
+We generate the tuple of variables $\p = (p^1_1,p^1_2,p^1_3) \cup (p^2_{(i,j)})$ for $(i,j)\in \{1,2,3\}\times\{1,2,3\}$.
+Note that in this procedure we will only have one set $\Gamma_1 = \{(1,2),(2,3)\}$.
+And so, the DNF formula we generate is
+		\begin{multline*}
+		\Phi^{*} = (p^1_1\wedge p^2_{(2,1)} \wedge \neg p^1_2)\vee(p^1_2\wedge p^2_{(3,2)} \wedge \neg p^1_3)\, \vee \\ (p^1_1 \wedge \neg p^1_1) \vee(p^1_2 \wedge \neg p^1_2) \vee (p^1_3\wedge \neg p^1_3) \vee \bigvee_{i,j\in\{1,2,3\}}(p^2_{(i,j)}\wedge\neg p^2_{(i,j)})
+		\end{multline*}
 
@@
 		\item Let $\psi = \exists \x\,\varphi(\x,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$ we generate a DNF-formula $\Phi = h(\A)$ as follows:
-
-		As in the previous case, we convert $\varphi(\x,\X,\z)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X,\z)$. Let each $\varphi_{i}(\x,\X,\z)$ be of the form
+Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\Truc$ we generate a DNF-formula $\Phi = h(\A)$ as follows.
+As in the previous case, we convert $\varphi(\x,\X,\z)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X,\z)$.
+Let each $\varphi_{i}(\x,\X,\z)$ be of the form
 		\[
-		\varphi_{i}(\x,\X,\z) = X_{s_1}(\y_{s_1}) \wedge \cdots \wedge X_{s_m}(\y_{s_m}) \wedge \neg X_{t_1}(\y_{t_1}) \wedge \cdots \wedge \neg X_{t_n}(\y_{t_1}) \wedge \alpha_i(\x,\z),
+		\varphi_{i}(\x,\X,\z) = X_{\pi_i(1)}(\y^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\y^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\y^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\y^{\nu}_{i,n_i}) \wedge \alpha_i(\x,\z),
 		\]
-		where every $\y_{s_i}$ and $\y_{t_j}$ have the corresponding number of variables in $(\x,\z)$, and $\alpha(\x)$ is a FO-formula.
+		where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$) is the rearrangement of $(\x,\z)$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x,\z)$ is a FO formula.
+We also define the tuple $(\a,\e)^{\pi}_{i,j}$ ($(\a,\e)^{\nu}_{i,j}$) as $\a$ rearranged in the same way as $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$), for each $(\a,\e)\in A^{c+d}$.
 
 		We generate a propositional variable $p^{\e}_{\ell,\b}$ for each $\e\in A^d$, for each $\ell\in\{1,\ldots,r\}$, and for each $\b\in A^{c_{\ell}}$.
+This variable will be true if there is an assignation to $\P$ such that $\b\in P_{\ell}$ and $\A\models\psi(\P,\e)$.
 Let $\p$ be the tuple of all such variables.
@@
 		\[
-		\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{s_1,\rho_{s_1}(\a,\e)} \wedge \cdots \wedge p^{\e}_{s_m,\rho_{s_m}(\a,\e)} \wedge \neg p^{\e}_{t_1,\rho_{t_1}(\a,\e)} \wedge \cdots \wedge \neg p^{\e}_{t_n,\rho_{t_n}(\a,\e)} \wedge \gamma_{\e} \wedge \theta_{\e}
+		\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{\pi_i(1),(\a,\e)^{\pi}_{i,1}} \wedge \cdots \wedge p^{\e}_{\pi_i(m_i),(\a,\e)^{\pi}_{i,m_i}} \wedge \neg p^{\e}_{\nu_i(1),(\a,\e)^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\e}_{\pi_i(n_i),(\a,\e)^{\nu}_{i,n_i}} \wedge \gamma_{\e} \wedge \theta_{\e},
 		\]
@@
 		\]
-		Finally, we return $f(\A) = \Phi = \bigvee_{\e\in A^d} \bigvee_{i = 1}^k \Phi_i^{\e}$.
+		Finally,$\Phi = \bigvee_{\e\in A^d} \bigvee_{i = 1}^k \Phi_i^{\e} \vee \bigvee_{p\in\p} (p\wedge\neg p)$, and we return $f(\A) = \A_{\Phi}$.
 
-		We will now show a function $g:\{\langle\P,\e\rangle \mid \A \models \exists\x\,\varphi(\x,\P,\e)\} \to \{\sigma \mid \sigma(\Phi) = 1\}$ and then prove that it is a bijection.
-Let $g$ be such that for each assignment $\langle \P,\e \rangle$, it holds that $g(\langle \P,\e \rangle) = \sigma$ where $\sigma$ is constructed as follows:
-
+		As an example, let $\psi^{*}(X_1,X_2,z) = \exists x\, (X_1(x) \wedge X_2(z) \wedge \neg X_1(z) \wedge x < z)$.
+We will compute $h(\B)$, where $\B = \langle \{1,2,3\}, < \rangle$, and we have that $1 < 2 < 3$.
+We generate the tuple of variables $\p = (p^e_{1,1},p^e_{1,2},p^e_{1,3},p^e_{2,1},p^e_{2,2},p^e_{2,3})$, for $e\in \{1,2,3\}$.
+Here, we consider $L = 2$, $T = \{t_1,t_2\}$, and $\beta$ defined as $\beta(1) = (0,0)$, $\beta(2) = (0,1)$ and $\beta(3) = (1,0)$.
+In this procedure we generate the sets $\Gamma^1_1 = \emptyset$, $\Gamma^2_1 = \{1\}$, $\Gamma^3_1 = \{1,2\}$.
+Now we generate $\gamma_e$ for $e\in\{1,2,3\}$,
+		\[
+		\gamma_e = \bigwedge_{d\in\{1,2,3\}\setminus\{e\}} p^d_{1,1} \wedge p^d_{1,2} \wedge p^d_{1,3} \wedge p^d_{2,1} \wedge p^d_{2,2} \wedge p^d_{2,3},
+		\]
+		and $\theta_e$ for $e\in\{1,2,3\}$,
+		\begin{align*}
+		\theta_1 &= \neg t_1 \wedge \neg t_2,\\
+		\theta_2 &= \neg t_1 \wedge t_2,\\
+		\theta_3 &= t_1 \wedge \neg t_2.		
+		\end{align*}
+		Then, we generate the formulas $\Phi^e_1$, for $e\in\{1,2,3\}$,
+		\begin{align*}
+		\Phi^1_1 &= \square\\
+		\Phi^2_1 &= p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2,\\
+		\Phi^3_1 &= (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3),
+		\end{align*}
+		and so, the DNF formula we generate is
+		\begin{multline*}
+		\Phi = (p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2) \vee (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee \\ \bigvee_{e\in\{1,2,3\}}[(p^e_{1,1}\wedge\neg p^e_{1,1})\vee
+		(p^e_{1,2}\wedge\neg p^e_{1,2})\vee
+		(p^e_{1,3}\wedge\neg p^e_{1,3})\vee
+		(p^e_{2,1}\wedge\neg p^e_{2,1})\vee
+		(p^e_{2,2}\wedge\neg p^e_{2,2})\vee
+		(p^e_{2,3}\wedge\neg p^e_{2,3})].
+		\end{multline*}
+		To demonstrate that our reduction is correct, we will show a function $g:\{\langle\P,\e\rangle \mid \A \models \exists\x\,\varphi(\x,\P,\e)\} \to \{\sigma \mid \sigma(\Phi) = 1\}$ and then prove that it is a bijection.
+Let $g$ be such that for each assignment $\langle \P,\e \rangle$, it holds that $g(\langle \P,\e \rangle) = \sigma$ where $\sigma$ is constructed as follows.
 For each $t_i\in T$, let $\sigma(t_i) = b^{\e}_i$.
@@
 For each $p_{\ell,\b}^{\e'}$, where $\e'\in A^d\setminus\{\e\}$,  $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e'}) = 1$.
+
 		Let $i$ be such that $\A\models \exists\x\,\varphi_i(\x,\P,\e)$.
@@
 From this we deduce that $g$ is surjective, and therefore a bijection.
-We conclude that $f(\A) = f_{\textsc{DNF}}(\Phi)$.
+We conclude that $f(\A) = f_{\textsc{\#DNF}}(\A_{\Phi})$.
 	\end{enumerate}
 \end{proof}
+\begin{theo}
+	$\SUo$ with $n$ open first-order variables is properly contained in $\SUo$ with $n+1$ open first-order variables for $n\in\N$.  
+\end{theo}
+\begin{proof}
+	Suppose it is not properly contained for $n$.
+Let $\z = (z_1,\ldots,z_{n+1})$, let $\u = (u_1,\ldots,u_n)$ and let $\varphi(\z) = \bigwedge_{i = 1}^{n} (z_i < z_{i+1} \wedge \forall y(z_i\not<y \vee y\not<z_{i+1} )) \wedge \forall y(y \not< z_1)$.
+Note that $f_{\varphi(\z)}(\A) = 1$ if $\A$ has at least $n+1$ elements, and  $f_{\varphi(\z)}(\A) = 0$ otherwise, for each $\A\in\Truc$.
+Let $\forall\x\,\psi(\x,\X,\u)$ be such that $f_{\varphi(\z)} = f_{\forall\x\,\psi(\x,\X,\u)}$.
+Let $\A = \langle A, \S^{\A}, <^{\A} \rangle\in\Truc$ where $\vert A \vert = n+1$.
+Note that $f_{\forall\x\,\psi(\x,\X,\u)}(\A) = 1$.
+Let $(\P,\e)$ be such that $\A\models\forall\x\,\psi(\x,\P,\e)$.
+Let $a\in A$ be such that $a\not\in\e$.
+Consider $\B$ as the induced substructure of $\A$ with domain $A\setminus\{a\}$, and let $\Q$ be the corresponding induced sub-assignment of $\P$.
+As universal formulae are preserved under substructure, we have that $\A\models\forall\x\,\psi(\x,\Q,\e)$.
+This implies that $f_{\forall\x\,\psi(\x,\X,\u)}(\B) \geq 1$, which leads to a contradiction.
+\end{proof}

### 2015-12-15 tex/examples.tex
2 nov
15 dic

@@
 \subsection{{\sc \#3-DNF}}
-We will now show that {\sc \#3-DNF} $\in\EoneE$.\\
+We will now show that {\sc \#3-DNF} $\in\XE{1}$.\\
 
@@
 \subsection{{\sc \#3-Horn-Sat}}
-We will now show that {\sc \#3-Horn-Sat} $\in\EoneU$.\\
+We will now show that {\sc \#3-Horn-Sat} $\in\XU{1}$.\\
 
@@
 
-\subsection{{\sc \#DNF}}
-We will now show that {\sc \#DNF} $\in\EtwoE$.\\
+\subsection{{\sc \#DNF}} \label{dnfdef}
+We will now show that {\sc \#DNF} $\in\XE{2}$.\\
 
@@
 
-By extension, {\sc \#3-DNF} $\in\EtwoE$.
+By extension, {\sc \#3-DNF} $\in\XE{2}$.
 
 \subsection{{\sc \#Horn-Sat}}
-We will now show that {\sc \#Horn-Sat} $\in\EoneU$.\\
+We will now show that {\sc \#Horn-Sat} $\in\XU{1}$.\\
 
@@
 \subsection{{\sc \#Disj-Horn-Sat}}
-We will now show that {\sc \#Disj-Horn-Sat} $\in\EtwoE$.
+We will now show that {\sc \#Disj-Horn-Sat} $\in\XE{2}$.
 

### 2015-12-15 tex/first_h.tex
15 dic

@@
 
-We define syntactically the classes $\Eie$ and $\Eiu$ according to the following grammar:
+We define syntactically the classes $\XE{i}$ and $\XU{i}$ according to the following grammar:
 \begin{enumerate}
-	\item $\Eezero$:
+	\item $\XE{0}$:
 	\[
@@
 	\]
-	\item $\Euzero$:
+	\item $\XU{0}$:
 		\[U_0 ::= E_0.\]
-	\item $\Eipe$:
+	\item $\XE{i+1}$:
 	\[	E_{i+1} ::= U_i \ \mid \ \exists x \, E_{i+1}.\]
-	\item $\Eipu$:
+	\item $\XU{i+1}$:
 		\[U_{i+1} ::= E_i \ \mid \ \forall x \, U_{i+1}.\]
@@
 
-A function $f$ is in $\Eie$ (resp. $\Eiu$) if there is an $\L$-formula $\varphi$ defined by the grammar $E_i$ (resp. $U_i$) such that $f = f_{\varphi}$.
+A function $f$ is in $\XE{i}$ (resp. $\XU{i}$) if there is an $\L$-formula $\varphi$ defined by the grammar $E_i$ (resp. $U_i$) such that $f = f_{\varphi}$.
 
 \begin{theo} \label{fp1}
-	$\Eo \subseteq$ {\sc FP}.
+	$\XE{0} \subseteq$ {\sc FP}.
 \end{theo}
 \begin{proof}
-	Let $f \in \Eo$, and let $\varphi(\X,\x)$ be and {\sc FO}-extended quantifier-free $\L$-formula such that:
+	Let $f \in \XE{0}$, and let $\varphi(\X,\x)$ be and {\sc FO}-extended quantifier-free $\L$-formula such that:
 	\begin{eqnarray*}
@@
 	\end{proof}
-	Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula, so $g \in\Eo$.
+	Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula, so $g \in\XE{0}$.
 It follows from \ref{mono} that for each $\A \in \Truc$, it holds that $\A \in L_f$ if and only if $g(\A) > 0$.
@@
 \begin{theo}
-	$\SUo$ with $n$ open first-order variables is properly contained in $\SUo$ with $n+1$ open first-order variables for $n\in\N$.  
+	$\U{1}$ with $n$ open first-order variables is properly contained in $\U{1}$ with $n+1$ open first-order variables for $n\in\N$.  
 \end{theo}
@@
 Let $a\in A$ be such that $a\not\in\e$.
-Consider $\B$ as the induced substructure of $\A$ with domain $A\setminus\{a\}$, and let $\Q$ be the corresponding induced sub-assignment of $\P$.
-As universal formulae are preserved under substructure, we have that $\A\models\forall\x\,\psi(\x,\Q,\e)$.
-This implies that $f_{\forall\x\,\psi(\x,\X,\u)}(\B) \geq 1$, which leads to a contradiction.
+Consider the vocabulary $\L' = \L \cup \{\Q\}$, and let $\A' = \A \cup \langle \Q^{\A'} \rangle$ where $\Q^{\A'} = \P$.
+Note that $\A\models\varphi(\P,\e)$ if and only if $\A'\models\varphi(\e)$.
+Consider $\B'$ as the induced substructure of $\A'$ with domain $A\setminus\{a\}$, and let $\Q^{\B'}$ be the induced interpretation of $\Q$.
+Also let $\B$ be the induced substructure of $\A$ with domain $A\setminus\{a\}$.
+As universal formulae are preserved under substructure\footnote{Hodges, 1997.
+P.
+183}, we have that $\B'\models\forall\x\,\psi(\x,\e)$, which implies that $\B\models\forall\x\,\psi(\x,\Q^{\B'},\e)$.
+We conclude that $f_{\forall\x\,\psi(\x,\X,\u)}(\B) \geq 1$, which leads to a contradiction.
 \end{proof}

### 2015-12-15 tex/horn_h.tex
15 dic

@@
-\section{The Extended Horn Counting Hierarchy}
+\section{The Horn Counting Hierarchy}
 
-We define syntactically the classes $\EiE$ and $\EiU$ as follows. First we define extended Horn clauses.
+We define the syntactic classes $\XE{i}$ and $\XU{i}$ as follows. First we define extended Horn clauses.
 \begin{eqnarray*}
@@
 NC &::=& NL \ \mid \ \alpha, \alpha \mbox{ is an {\sc FO}-formula over } \L \ \mid \ (NC \vee NC),\\
-HC &::=& NC \ \mid \ (NC \vee PL),
+HC &::=& NC \ \mid \ (NC \vee PL) \ \mid \ PL,
 \end{eqnarray*}
-where $\x$ is a tuple with the corresponding number of variables. Now we define the syntax of the classes inductively.
+where $PL$ represents a positive literal, $NL$ is an \textit{extended} negative literal, $NC$ is an extended Horn clause, and $HC$ is an extended Horn formula.
+Now we define the syntax of the classes inductively.\textbf{}
 \begin{enumerate}
-	\item $\EzeroE$:
+	\item $\HE{0}$:
 	\begin{eqnarray*}
@@
 	\end{eqnarray*}
-	\item $\EzeroU$:
+	\item $\HU{0}$:
 	\begin{eqnarray*}
@@
 	\end{eqnarray*}
-	\item $\EipE$:
+	\item $\HE{i+1}$:
 	\begin{eqnarray*}
@@
 	\end{eqnarray*}
-	\item $\EipU$:
+	\item $\HU{i+1}$:
 	\begin{eqnarray*}
@@
 \end{enumerate}
-A function $f$ is in $\EiE$ (resp. $\EiU$) if there is an $\L$-formula $\varphi$ defined by the grammar $E_i$ (resp. $U_i$) such that $f = f_{\varphi}$.
+A function $f$ is in $\XE{i}$ (resp. $\XU{i}$) if there is an $\L$-formula $\varphi$ defined by the grammar $E_i$ (resp. $U_i$) such that $f = f_{\varphi}$.
 
-The class $\PE$ is defined in [?] as $\PE = \{f\mid f\in\#P \mbox{ and its decision version } L_{f}\in P\}$.
+The class $\PE$ is defined in \cite{DBLP:conf/mfcs/PagourtzisZ06} as $\PE = \{f\mid f\in\#P \mbox{ and its decision version } L_{f}\in P\}$.
 
 \begin{theo}
-	$\EtwoE \subseteq \PE$
+	$\HE{2} \subseteq \PE$
 \end{theo} 
@@
 
-	Now we notice that $\zeta = \exists\X\exists\x\exists\y\,\theta(\X,\x,\y)$ is an existential second-order FO-formula, for which $\A\models\zeta$ can be decided in polynomial time.
+	Now we notice that $\zeta = \exists\X\exists\x\exists\y\,\theta(\X,\x,\y)$ is an existential second-order FO-formula, for which $\A\models\zeta$ can be decided in polynomial time \cite{DBLP:conf/stacs/Gradel91}.
 \end{proof}
@@
 \begin{theo}
-	$\textsc{Disj-Horn-Sat}$ is hard for $\EtwoE$ under parsimonious reductions. 
+	$\textsc{Disj-Horn-Sat}$ is hard for $\XE{2}$ under parsimonious reductions. 
 \end{theo} 
 \begin{proof}
-	Let $f$ be an arbitrary function in $\EtwoE$ and let $f_{\textsc{Disj-Horn-Sat}}$ be the function that characterizes {\sc \#Disj-Horn-Sat}.
+	Let $f$ be an arbitrary function in $\XE{2}$ and let $f_{\textsc{Disj-Horn-Sat}}$ be the function that characterizes {\sc \#Disj-Horn-Sat}.
 We will now show a function $h:\Truc\to L(P)$ such that for every $\A\in\Truc$, it holds that $f(\A) = f_{\textsc{Disj-Horn-Sat}}(h(\A))$, and which can be computed in polynomial time.

### 2015-12-15 tex/pe.tex
15 dic

\section{Appendix - Problems in $\PE$}

\subsection{\#Clique}
Counting the number of cliques in a graph is in $\HU{1}$. Let $\L = \{E,\leq\}$, and $\A = \langle A, E^{\A}, \leq^{\A} \rangle$ be an $\L$-structure that represents a graph $G = (A,E^{\A})$.

Let $C$ be a unary second-order variable, and
\[
\varphi(C) = \forall x \forall y((C(x)\wedge C(y))\to E(x,y)).
\]
Therefore, $f_{\varphi(C)}(\A)$ counts the number of cliques in $G$.

\subsection{Counting the number of matchings in a graph}
This function (also known as the Hosoya index of the graph) is in $\HU{1}$. Let $\L = \{E,\leq\}$, and $\A = \langle A, E^{\A}, \leq^{\A} \rangle$ be an $\L$-structure that represents a graph $G = (A,E^{\A})$.

Let $M$ be a unary second-order variable, and
\begin{align*}
\varphi(M) &= \forall x \forall y (M(x,y)\to E(x,y))\,\wedge \\
& \forall x \forall y (M(x,y)\to x < y)\,\wedge \\
& \forall x \forall y \forall z(M(x,y)\wedge M(z,y) \to x = z)\,\wedge \\
& \forall x \forall y \forall z(M(x,y)\wedge M(x,z) \to y = z)\,\wedge \\
& \forall x \forall y \forall z(\neg M(x,y)\vee \neg M(y,z)).
\end{align*}
Therefore, $f_{\varphi(M)}(\A)$ counts the number of matchings in $G$.
\subsection{\#2SAT}

\subsection{Counting the number of paths between two given nodes in a graph}

\subsection{Counting the number of Eulerian paths in a graph}

\subsection{Counting the number of perfect matchings in a bipartite graph}

### 2016-04-06 tex/below.tex
'5abr'

\section{Counting classes below \#P}

We define $\totp$ using the following function. For each non-deterministic Turing machine $M$,
\[
	\tot_M(x) = (\text{\#paths of $M$ on input $x$}) - 1.
\]
A function $f:\{0,1\}^*\to\N$ is in $\totp$ if there exists a polynomial-time non-deterministic Turing machine $M$ such that $f = \tot_M$.

We define $\spanl$ using the following function. For each non-deterministic Turing machine $M$,
\[
	\sp_M(x) = \vert\{w \mid \text{$w$ is an output of $M$ on input $w$}\}\vert.
\]
A function $f:\{0,1\}^*\to\N$ is in $\spanl$ if there exists a log-space non-deterministic Turing machine $M$ such that $f = \sp_M$.

\begin{theo}
	$\spanl \subseteq \totp$.
\end{theo}
\begin{proof}
	Let $f:\{0,1\}^*\to\N$ be a function in $\spanl$ and let $M$ be a non-deterministic log-space Turing machine such that $f = \sp_M$. We will provide a polynomial-time non-deterministic Turing machine $M'$ such that for each $x\in\{0,1\}^*$ it holds that $\sp_M(x) = \tot_{M'}(x)$. First we note that given a configuration $s$ of $M$ on input $x$ checking whether there exists a final state of $M$ which is reachable from $s$ can be done in polynomial time. This implies that checking whether $f(x) > 0$ can also be done in polynomial time. We present this machine as a non-deterministic algorithm.

	\begin{algorithm}
		\caption{Calculate $\sp_M(x)$}
		\label{alg1}
		\begin{algorithmic}
			\REQUIRE $n \geq 0 \vee x \neq 0$
			\ENSURE $y = x^n$
			\STATE $y \leftarrow 1$
			\IF{$n < 0$}
			\STATE $X \leftarrow 1 / x$
			\STATE $N \leftarrow -n$
			\ELSE
			\STATE $X \leftarrow x$
			\STATE $N \leftarrow n$
			\ENDIF
			\WHILE{$N \neq 0$}
			\IF{$N$ is even}
			\STATE $X \leftarrow X \times X$
			\STATE $N \leftarrow N / 2$
			\ELSE[$N$ is odd]
			\STATE $y \leftarrow y \times X$
			\STATE $N \leftarrow N - 1$
			\ENDIF
			\ENDWHILE
		\end{algorithmic}
	\end{algorithm}

\end{proof}

### 2016-04-06 tex/below.tex
6 abr

@@
 	\begin{algorithm}
-		\caption{Calculate $\sp_M(x)$}
-		\label{alg1}
-		\begin{algorithmic}
-			\REQUIRE $n \geq 0 \vee x \neq 0$
-			\ENSURE $y = x^n$
-			\STATE $y \leftarrow 1$
-			\IF{$n < 0$}
-			\STATE $X \leftarrow 1 / x$
-			\STATE $N \leftarrow -n$
-			\ELSE
-			\STATE $X \leftarrow x$
-			\STATE $N \leftarrow n$
-			\ENDIF
-			\WHILE{$N \neq 0$}
-			\IF{$N$ is even}
-			\STATE $X \leftarrow X \times X$
-			\STATE $N \leftarrow N / 2$
-			\ELSE[$N$ is odd]
-			\STATE $y \leftarrow y \times X$
-			\STATE $N \leftarrow N - 1$
-			\ENDIF
-			\ENDWHILE
+
+	\begin{algorithmic}[]
+		\Begin
+		\If{$f(x) = 0$} stop
+		\Else \text{ choose from}
+		\State stop
+		\State call $\textsc{GenTree}_{f}(x,\{s_0\})$
+		\EndIf
+		\EndBegin
+		\Procedure{$\textsc{GenTree}_{f}(x,S)$}{}
+		\State $S_0 = \emptyset$, $S_1 = \emptyset$
+		\For{$s\in S$}
+		\State do a search in the configuration graph of $M$ with input $x$ from $s$:
+		\If{$s'$ is reached after writing 0 in the output tape} $S_0 \gets s'$
+		\EndIf
+		\If{$s'$ is reached after writing 1 in the output tape} $S_1 \gets s'$
+		\EndIf
+		\EndFor
+		\If{a final state from $M$ is reachable from some $s\in S_0$ and some $s'\in S_1$} choose between:
+		\State call $\textsc{GenTree}_{f}(x,S_0)$
+		\State call $\textsc{GenTree}_{f}(x,S_1)$
+		\ElsIf{a final state from $M$ is reachable from some $s\in S_i$ but not from any $s'\in S_{1-i}$}
+		\State call $\textsc{GenTree}_{f}(x,S_0)$
+		\ElsIf{no final state from $M$ is reachable from any $s\in S_0 \cup S_1$}
+		\State stop
+		\EndIf
+		\EndProcedure
 	\end{algorithmic}
+
 	\end{algorithm}
 
+	We use $s_0$ for the initial state of $M$ with input $x$.
+The algorithm can be executed in a nondeterministic polynomial-time machine, so we only need to show that the number of computation paths is equal to $\sp_M(x)$.
+
 \end{proof}

### 2016-04-13 tex/below.tex
13 abr

@@
 
-We define $\totp$ using the following function. For each non-deterministic Turing machine $M$,
+We define $\totp$ using the following function. For each non-deterministic Turing machine $M$ and $x\in\{0,1\}^*$,
 \[
@@
 
-We define $\spanl$ using the following function. For each non-deterministic Turing machine $M$,
+We define $\spanl$ using the following function. For each non-deterministic Turing machine $M$ and $x\in\{0,1\}^*$,
 \[
@@
 This implies that checking whether $f(x) > 0$ can also be done in polynomial time.
-We present this machine as a non-deterministic algorithm.
+We present this machine $M'$ as a non-deterministic algorithm.
 
@@
 		\State stop
-		\State call $\textsc{GenTree}_{f}(x,\{s_0\})$
+		\State call $\textsc{GenTree}_{f}(x,\{s_0\},\epsilon)$
 		\EndIf
 		\EndBegin
-		\Procedure{$\textsc{GenTree}_{f}(x,S)$}{}
+		\Procedure{$\textsc{GenTree}_{f}(x,S,y)$}{}
 		\State $S_0 = \emptyset$, $S_1 = \emptyset$
@@
 		\If{a final state from $M$ is reachable from some $s\in S_0$ and some $s'\in S_1$} choose between:
-		\State call $\textsc{GenTree}_{f}(x,S_0)$
-		\State call $\textsc{GenTree}_{f}(x,S_1)$
+		\State call $\textsc{GenTree}_{f}(x,S_0,y0)$
+		\State call $\textsc{GenTree}_{f}(x,S_1,y1)$
 		\ElsIf{a final state from $M$ is reachable from some $s\in S_i$ but not from any $s'\in S_{1-i}$}
-		\State call $\textsc{GenTree}_{f}(x,S_0)$
+		\State call $\textsc{GenTree}_{f}(x,S_i,yi)$
 		\ElsIf{no final state from $M$ is reachable from any $s\in S_0 \cup S_1$}
-		\State stop
+		\Return $y$
 		\EndIf
@@
 	We use $s_0$ for the initial state of $M$ with input $x$.
-The algorithm can be executed in a nondeterministic polynomial-time machine, so we only need to show that the number of computation paths is equal to $\sp_M(x)$.
+The algorithm can be executed in a nondeterministic polynomial-time machine, so we only need to show that the number of computation paths is equal to $\sp_M(x)+1$.
+
+	For each output $y$ consider some computation path $p$ of $M$ with input $x$ that outputs $y$.
+Note there is at least one computation path in $M'$ that follows $p$ and outputs $y$.
+Also, note that each distinct computation path of $M'$ outputs a different $y$, since in each branch the word differs in the last bit.
+Taking into account the dummy path in line 4, we conclude that the number of computation paths of $M'$ with input $x$ is equal to $\sp_M(x)+1$.
 
 \end{proof}
+
+Given a deterministic machine $M$ we define the function $f_M\{0,1\}^*\to\N$ as:
+\[
+	f_M(x) = \text{the output of $M$ on input $w$}
+\]
+for each $w\in\{0,1\}^*$. A function $f:\{0,1\}^*\to\N$ is in FP if there exists a deterministic polynomial-time machine $M$ such that $f = f_M$.
+\begin{theo} \label{diff}
+	For each function $f\in\shp$ there exists functions $g\in \totp$ and $h\in\text{FP}$ such that $f = g - h$.
+\end{theo}
+\begin{proof}
+	Let $f\in\shp$, and let $r:\{0,1\}^*\to\{0,1\}^*$ be such that for each $w\in\{0,1\}^*$ it holds that $f(w) = \shsat(r(w))$, and be computed in polynomial time. Let $\varphi_w = r(w)$.
+
+	Let $M$ be a non-deterministic machine that on input $w$ does the following: Computes $\varphi_w$.
+Let $n$ be the number of variables on $\varphi_w$.
+Then it chooses $x\in\{0,1\}$.
+If $x = 0$ then stop.
+Else, choose a valuation $\sigma$ of size $n$.
+If $\sigma\models\varphi_w$ then stop.
+Else, choose $y\in\{0,1\}$ then stop.
+Note that $\tot_M(w) = \shsat(\varphi_w)+2^n+1$.
+Let $g:\{0,1\}^*\to\N$ defined as $g(x) = \tot_M(x) - 1$.
+Clearly $g\in\totp$.
+
+	Let $M'$ be a deterministic machine that on input $w$ computes $\varphi_w$ and returns $2^n$. Let $h\in\{0,1\}^*\to\N$ defined as $f_M$. Since $M'$ runs in polynomial time, $h\in\text{FP}$.
+
+	Finally, for each $w\in\{0,1\}^*$ we have that $g(w) - h(w) = \shsat(\varphi_w) + 2^n - 2^n = f(w)$.
+\end{proof}

### 2016-04-13 tex/pe.tex
13 abr

@@
 \subsection{Counting the number of perfect matchings in a bipartite graph}
+
+\subsection{\shsat + $2^n$}
+
+This function takes a propositional formula $\varphi$ as an input and outputs the number of satisfying assignments of $\varphi$ plus $2^n$, where $n$ is the number of variables in $\varphi$.
+Since the decision version of this function is trivial, it is in $\shp$.
+We also show using the machine described in Theorem \ref{diff} that this function is in $\totp$.

### 2016-04-20 tex/below.tex
19 abr

@@
 \end{proof}
+
+\begin{theo}
+	$\XE{2} \subseteq \totp$
+\end{theo}
+\begin{proof}
+	Let $f:\Truc\to\N$ be a function in $\XE{2}$.
+Let $h$ be the function described in Theorem \ref{sigma2hard}.
+Consider a non-deterministic $M$ machine that on input $\A$ first computes $\varphi = h(\A)$ and then simulates the $\totp$ machine that computes $\shdhsat$.
+Since $h$ is a parsimonious reduction from $f$ to $\shdhsat$, it holds that $f(\A) = \tot_M(\A)+1$.
+\end{proof}

### 2016-04-20 tex/horn_h.tex
19 abr

@@
 \[
-\textsc{Disj-Horn-Sat} = \{\Phi \mid \Phi \mbox{ is a disjunction of Horn formulas and $\Phi$ is satisfiable}\},
+\shdhsat = \{\Phi \mid \Phi \mbox{ is a disjunction of Horn formulas and $\Phi$ is satisfiable}\},
 \]
-and the counting problem {\sc \#Disj-Horn-Sat} as a function that counts all satisfying assignments to a formula $\Phi$ which is a disjunction of Horn formulas.
+and the counting problem {\shdhsat } as a function that counts all satisfying assignments to a formula $\Phi$ which is a disjunction of Horn formulas.
 
-\begin{theo}
-	$\textsc{Disj-Horn-Sat}$ is hard for $\XE{2}$ under parsimonious reductions. 
+\begin{theo} \label{sigma2hard}
+	$\shdhsat$ is hard for $\XE{2}$ under parsimonious reductions. 
 \end{theo} 
 \begin{proof}
-	Let $f$ be an arbitrary function in $\XE{2}$ and let $f_{\textsc{Disj-Horn-Sat}}$ be the function that characterizes {\sc \#Disj-Horn-Sat}.
-We will now show a function $h:\Truc\to L(P)$ such that for every $\A\in\Truc$, it holds that $f(\A) = f_{\textsc{Disj-Horn-Sat}}(h(\A))$, and which can be computed in polynomial time.
+	Let $f$ be an arbitrary function in $\XE{2}$.
+We will now show a function $h:\Truc\to L(P)$ such that for every $\A\in\Truc$, it holds that $f(\A) = \shdhsat(h(\A))$, and which can be computed in polynomial time.
 Let $\psi$ be such that $f = f_{\psi}$.
@@
 Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$.
-The Disj-Horn-Sat formula we return is
+The disjunction of Horn formulas we return is
 		\[

### 2016-04-20 tex/pe.tex
19 abr

@@
 \section{Appendix - Problems in $\PE$}
+
+In this section we will show some problems that are known to be in $\PE$, a formula that captures the function, if known, and proof that it belongs to $\totp$.
 
@@
 
-\subsection{Counting the number of paths between two given nodes in a graph}
+\subsection{\#PATH}
+
+We define $\textsc{\#PATH}$ as a function such that for each non-directed graph $G$, a pair of its nodes $a,b$ and $k\in\N$, it holds that $\textsc{\#PATH}(G,a,b,k)$ is equal to the numbers of paths from $a$ to $b$ with size at most $k$.
+
+\begin{theo}
+	 $\textsc{\#PATH} \in \shl$
+\end{theo}
+\begin{proof}
+	Let $M$ be a non-deterministic machine that on input $(G,a,b,k)$ does the following: Start on $a$.
+On each cycle let $u$ be the previous vertex, and $M$ guesses the next vertex $v$, if $(u,v)\not\in E$, $M$ rejects.
+Else, $v = u$ and this ends the cycle.
+If it finds $b$, it accepts.
+If $k$ cycles have been made without finding $b$, it rejects.
+
+	Clearly, the number of accepting paths of $M$ with input $(G,a,b,k)$ is equal to $\textsc{\#PATH}(G,a,b,k)$.
+It can be seen that $M$ uses logarithmic space since at each point of the computation it only needs the value of $u$ and a cycle counter.
+\end{proof}
+
+\begin{coro}
+	 $\textsc{\#PATH} \in \totp$
+\end{coro}
+
+\subsection{\#SIMPLEPATH}
+
+We define $\textsc{\#SIMPLEPATH}$ as a function such that $\textsc{\#SIMPLEPATH}(G,a,b)$ counts the number of paths from $a$ to $b$ such that no vertex appears twice or more in the path.
+
+\begin{coro}
+	$\textsc{\#SIMPLEPATH} \in \totp$
+\end{coro}
+\begin{proof}
+	Let $M$ be a non-deterministic machine that on input $(G,a,b)$ does the following: If there is no path from $a$ to $b$, stop.
+Else, choose $i\in\{0,1\}$ if $i = 0$ stop.
+Else, start on $a$ and save the path $P=\emptyset$.
+On each cycle let $S = \emptyset$, let $u$ be the previous vertex and $P$ the current path.
+If $u = b$, stop.
+Else for each vertex $v$ such that $(u,v)\in E$, check whether there exists a path $Q$ from $v$ to $b$ such that no vertex of $Q$ appears in $P$.
+If it exists, add $v$ to $S$.
+Choose some vertex $v'\in S$, add it to $P$ and do $u = v'$.
+This ends the cycle.
+
+	Clearly the number of paths of $M$ on input $(G,a,b)$ is equal to $\textsc{\#SIMPLEPATH}(G,a,b)+1$. It can also be seen that $M$ runs on polynomial time.
+\end{proof}
 

### 2016-04-20 tex/below.tex
20 abr

@@
 	\begin{algorithm}	
-
+	\caption{Compute $f(x)+1$ paths}
 	\begin{algorithmic}[]
@@
 	\end{algorithmic}
-
 	\end{algorithm}
@@
 	We use $s_0$ for the initial state of $M$ with input $x$.
-The algorithm can be executed in a nondeterministic polynomial-time machine, so we only need to show that the number of computation paths is equal to $\sp_M(x)+1$.
+The algorithm can be executed in a nondeterministic polynomial-time machine, so we only need to show that the number of computation paths is equal to $\sp_M(x)-1$.
 

### 2016-04-20 tex/pe.tex
20 abr

@@
 Therefore, $f_{\varphi(C)}(\A)$ counts the number of cliques in $G$.
+
+We now show that \#Clique belongs to {\totp } with a non-deterministic machine $M$ which we present as a non-deterministic algorithm that takes $G = (V,E)$ as input.
+We further suppose that $V = \{1,...,n\}$.
+
+	\begin{algorithm}	
+		\caption{Compute $\text{\#Clique}(G)+1$ paths}
+		\begin{algorithmic}[]
+			\Begin
+			\If{$G = \emptyset$} stop
+			\Else \text{ choose} $v\in V\cup\{\perp\}$ 
+			\If{$v = \perp$} stop
+			\Else \text{ call} $\textsc{GenTree}(G,\emptyset)$
+			\EndIf
+			\EndIf
+			\EndBegin
+			\Procedure{$\textsc{GenTree}(G,S)$}{}
+			\State $u \gets$ the last vertex in $S$
+			\State $T = \emptyset$
+			\For{$v \in \{u+1,\dots,n\}$}
+			\If{$v$ is connected to each $u'\in S$}
+			\State $T \gets v$
+			\EndIf
+			\EndFor
+			\If{$T \neq \emptyset$} choose $v'\in T\cup\{\perp\}$
+			\If{$v' = \perp$} stop
+			\Else \text{ call} $\textsc{GenTree}(G,S\cup\{v'\})$
+			\EndIf
+			\Else \text{ stop}
+			\EndIf
+			\EndProcedure
+		\end{algorithmic}
+	\end{algorithm}
+
+At each point in the computation $S$ contains the vertices of a clique in $G$, and since the computation follows the order of the vertices in $G$, every clique is counted exactly once.
 
@@
 Therefore, $f_{\varphi(M)}(\A)$ counts the number of matchings in $G$.
+
 \subsection{\#2SAT}

### 2016-05-04 tex/below.tex
5 may

@@
 \end{proof}
+
+\begin{theo}
+	$\totp$ is closed under substraction of a polynomial on the size of the input.
+\end{theo}
+\begin{proof}
+	Let $f:\Truc\to\N$ be a function in $\totp$ and let $p$ be a polynomial.
+We will show a function $g\in\totp$ such that for each input $w$ it holds that $g(w) = \max(f(w) - p(\vert w \vert),0)$.
+Let $M$ be the non-deterministic Turing Machine such that $f(w) = \tot_M(w)+1$.
+We construct a non-deterministic Turing Machine that first computes $p(\vert w \vert)$.
+Then it simulates $M$ until it branches $p(\vert w \vert)$ times or finds the end of a path, following the first path of each branch.
+Each time it finds the end of a path, it subtracts one from $p(\vert w \vert)$ and backtracks until the previous branching path and continues repeating this process.
+After erasing $p(\vert w \vert)$ paths it continues simulating $M$.
+\end{proof}

### 2016-05-11 tex/below.tex
11 may

--- 2016-04-20 tex/below.tex
@@
 Else, choose $y\in\{0,1\}$ then stop.
-Note that $\tot_M(w) = \shsat(\varphi_w)+2^n+1$.
-Let $g:\{0,1\}^*\to\N$ defined as $g(x) = \tot_M(x) - 1$.
+Note that $\tot_M(w) = \shsat(\varphi_w)+2^n$.
+Let $g:\{0,1\}^*\to\N$ defined as $g(x) = \tot_M(x)$.
 Clearly $g\in\totp$.
@@
 Consider a non-deterministic $M$ machine that on input $\A$ first computes $\varphi = h(\A)$ and then simulates the $\totp$ machine that computes $\shdhsat$.
-Since $h$ is a parsimonious reduction from $f$ to $\shdhsat$, it holds that $f(\A) = \tot_M(\A)+1$.
+Since $h$ is a parsimonious reduction from $f$ to $\shdhsat$, it holds that $f(\A) = \tot_M(\A)$.
 \end{proof}
+
+\begin{theo}
+	$\totp$ is closed under subtraction of a polynomial on the size of the input.
+\end{theo}
+\begin{proof}
+	Let $f:\Truc\to\N$ be a function in $\totp$ and let $p$ be a polynomial.
+We will show a function $g\in\totp$ such that for each input $w$ it holds that $g(w) = \max(f(w) - p(\vert w \vert),0)$.
+Let $M$ be the non-deterministic Turing Machine such that $f(w) = \tot_M(w)$.
+
+	We construct a non-deterministic Turing Machine $M'$ that first computes $c = p(\vert w \vert)$.
+Then it deterministically simulates $M$ until it finds the end of a path, following the first path of each branch.
+Each time it finds the end of a path, it subtracts one from $c$ and saves the string of non-deterministic choices that were made.
+The machine then backtracks until the previous branching path and continues repeating the process.
+When all choices in a branch have been saved, it erases every path in that branch and saves only the choices until that point.
+If it happens that there are no more choices to make, and $c > 0$ the machine simply stops.
+If $c$ reaches 0, $M'$ simulates $M$ with input $w$.
+Each time the machine is about to do a non-deterministic choice it checks if it corresponds to any of the saved paths.
+If it is the case, it does not follow the branch.
+
+	We will now show that $g(w) = \tot_{M'}(w)$.
+Let $w\in\{0,1\}$.
+If $\tot_{M}(w) > p(\vert w \vert)$, then clearly $\tot_{M'}(w) = \tot_{M}(w) - p(\vert w \vert)$.
+If $\tot_{M}(w) \leq p(\vert w \vert)$, then it only computes the first simulation path and $\tot_{M'}(w) = 0$.
+\end{proof}

### 2016-05-25 tex/below.tex
25 may

@@
 A function $f:\{0,1\}^*\to\N$ is in $\spanl$ if there exists a log-space non-deterministic Turing machine $M$ such that $f = \sp_M$.
+
+Given a language $L$, $\totp^{L}$ is defined by all functions $f:\{0,1\}^*\to\N$ such that $f = tot_{M^L}$.
+\\
+
+By the following result, we note that $\shp$ can be considered an extension of $\totp$.
+
+\begin{theo}
+	$\shp \subseteq \totp^{\textsc{SAT}}$
+\end{theo}
+\begin{proof}
+	Let $f\in\shp$ and let $M$ be the non-deterministic polynomial-time machine such that $f = \acc_M$.
+We provide a machine $M'^{\textsc{SAT}}$ with an oracle of SAT using the following construction.
+On input $w$, $M'$ gives the oracle a formula $\varphi$ that can be satisfied if and only if $M$ has at least one accepting path on input $w$.
+If it answers {\it no}, the computation ends.
+If it answers {\it yes}, it chooses $x\in\{0,1\}$.
+If $x = 0$ stops, and if $x = 1$, then it does the following.
+$M'$ simulates $M$ until a non-deterministic choice has to be taken.
+Let $S$ be the set of configurations that are reached on each choice.
+Then, for each $s\in S$, $M'$ constructs a formula $\varphi_s$ that models the machine $M$ on input $w$, starting on configuration $s$, which is satisfiable if and only if $M$ has at least one accepting path.
+For each $s\in S$, the formula $\varphi_s$ is given to the oracle, and if it answers {\it yes}, the corresponding choice is taken non-deterministically.
+If it answers {\it no} the choice is ignored.
+This is repeated until no more choices can be made.
+It is clear that $\acc_M = \tot_{M'^{\textsc{SAT}}}$.
+We conclude that $f\in\totp^{\textsc{SAT}}$.
+\end{proof}
 
@@
 If $c$ reaches 0, $M'$ simulates $M$ with input $w$.
-Each time the machine is about to do a non-deterministic choice it checks if it corresponds to any of the saved paths.
+Each time the machine is about to do a non-deterministic choice $M'$ checks if the it corresponds to any of the saved paths.
 If it is the case, it does not follow the branch.
@@
 \end{proof}
+
+Given a polynomial $p$ and a function $f\{0,1\}^*\to\N$ we define the set
+\[
+	polycheck_{p,f} = \{w\mid f(w) \geq p(\vert w \vert) \}.
+\]
+We define $\shpc$ as the class of functions $f\in\shp$ such that for each polynomial $p$ it holds that $polycheck_{p,f}\in \text{PTIME}$.

### 2016-06-01 tex/below.tex
6 jun

@@
 By the following result, we note that $\shp$ can be considered an extension of $\totp$.
-
-\begin{theo}
-	$\shp \subseteq \totp^{\textsc{SAT}}$
-\end{theo}
-\begin{proof}
-	Let $f\in\shp$ and let $M$ be the non-deterministic polynomial-time machine such that $f = \acc_M$.
-We provide a machine $M'^{\textsc{SAT}}$ with an oracle of SAT using the following construction.
-On input $w$, $M'$ gives the oracle a formula $\varphi$ that can be satisfied if and only if $M$ has at least one accepting path on input $w$.
-If it answers {\it no}, the computation ends.
-If it answers {\it yes}, it chooses $x\in\{0,1\}$.
-If $x = 0$ stops, and if $x = 1$, then it does the following.
-$M'$ simulates $M$ until a non-deterministic choice has to be taken.
-Let $S$ be the set of configurations that are reached on each choice.
-Then, for each $s\in S$, $M'$ constructs a formula $\varphi_s$ that models the machine $M$ on input $w$, starting on configuration $s$, which is satisfiable if and only if $M$ has at least one accepting path.
-For each $s\in S$, the formula $\varphi_s$ is given to the oracle, and if it answers {\it yes}, the corresponding choice is taken non-deterministically.
-If it answers {\it no} the choice is ignored.
-This is repeated until no more choices can be made.
-It is clear that $\acc_M = \tot_{M'^{\textsc{SAT}}}$.
-We conclude that $f\in\totp^{\textsc{SAT}}$.
-\end{proof}
 
@@
 We define $\shpc$ as the class of functions $f\in\shp$ such that for each polynomial $p$ it holds that $polycheck_{p,f}\in \text{PTIME}$.
+
+We present a counting class diagram:
+
+\input{classdiag}
+
+Now we will prove the non-trivial inclusions.
+
+\begin{theo}
+	$\shp \subseteq \totp^{\textsc{SAT}}$
+\end{theo}
+\begin{proof}
+	Let $f\in\shp$ and let $M$ be the non-deterministic polynomial-time machine such that $f = \acc_M$.
+We provide a machine $M'^{\textsc{SAT}}$ with an oracle of SAT using the following construction.
+On input $w$, $M'$ gives the oracle a formula $\varphi$ that can be satisfied if and only if $M$ has at least one accepting path on input $w$.
+If it answers {\it no}, the computation ends.
+If it answers {\it yes}, it chooses $x\in\{0,1\}$.
+If $x = 0$ stops, and if $x = 1$, then it does the following.
+$M'$ simulates $M$ until a non-deterministic choice has to be taken.
+Let $S$ be the set of configurations that are reached on each choice.
+Then, for each $s\in S$, $M'$ constructs a formula $\varphi_s$ that models the machine $M$ on input $w$, starting on configuration $s$, which is satisfiable if and only if $M$ has at least one accepting path.
+For each $s\in S$, the formula $\varphi_s$ is given to the oracle, and if it answers {\it yes}, the corresponding choice is taken non-deterministically.
+If it answers {\it no} the choice is ignored.
+This is repeated until no more choices can be made.
+It is clear that $\acc_M = \tot_{M'^{\textsc{SAT}}}$.
+We conclude that $f\in\totp^{\textsc{SAT}}$.
+\end{proof}
+
+For a non-deterministic machine with an output tape $M$, we define $\sp!_M:\{0,1\}^*\to\N$ as follows:
+\[
+	\sp!_M(x) = \{w \mid \text{every path of $M$ with input $x$ that returns $w$ is an accepting path.} \}
+\]
+Similarly, we define $\stspanp$ (suggested pronunciation: ``strict span P'') as the class of functions $\sp!_M$ where $M$ is a non-deterministic polynomial time machine.
+
+To prove the inclusion $\spanp \subseteq \shp^{\np}$, we will use a more strong result.
+
+\begin{theo}
+	$\spanp \subseteq \stspanp$.
+\end{theo}
+\begin{proof}
+	Let $f \in \spanp$ and let $M$ be the non-deterministic machine for which $f = \sp_M$ that runs in $p(\vert x \vert)$ time.
+We construct a non-deterministic machine $M'$ that on input $w$ does the following.
+It simulates $M(x)$, and when the simulation finishes on a final state of $M$, stops.
+When it finishes on a non-final state, it appends a dummy sequence of characters, for example $0^{p(\vert x \vert)+1}$, to the output.
+Clearly $f = \sp!_{M'}$, and since $M'$ runs in polynomial time, we have that $f \in \stspanp$.
+\end{proof}
+
+\begin{theo}
+	$\stspanp = \shp^{\np}$
+\end{theo}
+\begin{proof}
+	To show that $\stspanp \subseteq \shp^{\np}$, consider some $f \in \stspanp$ and some non-deterministic machine $M$ for which $f = \sp!_M$ that runs in $p(\vert x\vert)$.
+Suppose every output of $M(x)$ has length $p(\vert x\vert)$.
+Consider the following language in NP:
+	\begin{multline*}
+		L = \{ (i,M,x,w) \mid M \text{ is a non-determinstic polynomial-time machine and,} \\ \text{either $i = 1$ and $w$ is given as output in some accepting path of $M(x)$}, \\ \text{or $i = 0$ and $w$ is given as output in some rejecting path of $M(x)$}\}.
+	\end{multline*}
+	 We construct a non-deterministic machine $M'$ with an oracle on $L$ that on input $x$ does the following.
+Guess some $w\in\{0,1\}^{p(\vert x\vert)}$ Give $(1,M,x,w)$ to the oracle.
+If it responds {\tt no}, then $M'$ answers {\tt no}.
+If the oracle responds {\tt yes}, then it proceeds giving $(0,\overline{M},x,w)$ to the oracle, where $\overline{M}$ is the complement of the machine $M$.
+If the oracle responds {\tt yes}, then $M'$ answers {\tt no}.
+If it responds {\tt no}, then $M'$ answers {\tt yes}.
+
+	 To show that $\shp^{\np} \subseteq \stspanp$, consider some $f \in \shp^{\np}$, some language $L\in\np$ and some non-deterministic machine $M^L$ for which $f = \acc_{M^L}$.
+We provide a non-deterministic machine $M'$ as the following non-deterministic procedure.
+Let $M_L$ the non-deterministic polynomial-time machine machine that computes $L$.
+	 \begin{algorithm}
+	 \caption{Compute $\acc_{M^L}(x)$ outputs that are accepted in each respective path}
+	 \begin{algorithmic}[]
+		 \Begin
+		 \State $w \gets \varepsilon$.
+		 \State Simulate $M^L(x)$:
+		 \While{running $M^L(x)$}
+		 \State $s\gets$ the current configuration of $M^L(x)$.
+		 \State $w \gets w \cdot s$.
+		 \If{$s$ is a call to the oracle with input $y$}
+		 \State Non-det. simulate $M_L(y)$:
+		 \InEach{accepting path} {\bf do} two branches:
+		 \State Continue running $M^L(x)$ on $s$ with the state $q_{\texttt{yes}}$.
+		 \State Continue running $M^L(x)$ on $s$ with the state $q_{\texttt{no}}$, but rejecting at the end
+		 \EndInEach
+		 \InEach{rejecting path}
+		 \State Continue running $M^L(x)$ on $s$ with the state $q_{\texttt{no}}$.
+		 \EndInEach
+		 \ElsIf{$\delta(s) \neq \emptyset$} choose from $\delta(s)$
+		 \State Continue simulating on that choice.
+		 \ElsIf{$s$ corresponds a final state}
+		 \State \Return {\tt yes} with output $w$, unless previously decided to be rejected.
+		 \Else
+		 \State \Return {\tt no} with output $w$.
+		 \EndIf
+		 \EndWhile
+		 \EndBegin
+	 \end{algorithmic}
+ 	 \end{algorithm}
+
+	 Let $x\in\{0,1\}^*$.
+To show that $\sp!_{M'}(x) = \acc_{M^L}(x)$, first note that for each strictly accepting output $w$ of $M'(x)$ there exists some accepting path $p$ of $M^L(x)$ such that $p = w$.
+Now note that for each accepting path $p$ of $M^L(x)$ there exists an output $w$ of $M'(x)$ such that $p = w$.
+This output is clearly accepted in some branch.
+To show that this output is not rejected in any branch, by contradiction suppose that it is.
+This branch cannot correspond to a rejecting path of $M^L(x)$ and then it must be rejected on line 11, from simulating a negative answer of the oracle in one of its accepting paths.
+This branch corresponds to a path that is never reachable since in $M^L(x)$ the oracle answered {\tt yes}, and we follow to a contradiction.
+\end{proof}

### 2016-06-03 tex/gi.tex
graph iso proof added

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsthm}
\usepackage{fullpage}

\def\shgi{\mbox{\#GI}}
\def\lgi{\mbox{GI}}
\def\fp{\mbox{FP}}
\def\aut{\mbox{aut}}
\def\iso{\mbox{iso}}

\title{Graph Isomorphism is Polynomially Equivalent to Counting Isomorphisms}

\newtheorem{theo}{Theorem}		
\newtheorem{lemma}[theo]{Lemma}
\newtheorem{claim}{Claim}[theo]
\newtheorem{coro}{Corollary}[theo]

\begin{document}

\maketitle

In this document we provide a comprehensive proof of Mathon's result that counting the isomorphisms from a pair of graphs can be done with a PTIME machine with an oracle to the graph isomorphism language.

Let us define the language $\lgi$ as:
\[
\lgi = \{(G_1,G_2)\mid \mbox{there exists an isomorphism $f$ from $G_1$ to $G_2$} \},
\]
and the function $\shgi$ as:
\[
\shgi(G_1,G_2) = \vert \{ f \mid \mbox{$f$ is an isomorphism from $G_1$ to $G_2$}  \} \vert.
\]
Before jumping to the main result, we provide some definitions.

For each graph $G$ and vertices $v_1,\ldots,v_k$ in $G$, we define $G_{v_1,\ldots,v_k}$ as a copy of $G$ with distinct labels on $v_1,\ldots,v_k$. This can be done by connecting each $v_i$ with some line of length $n+k$, where $\vert G \vert = n$.

For each graph $G$, we define $\aut(G) = \{g\mid \mbox{$g$ is an automorphism of $G$} \}$. Note that $(\aut(G),\circ)$ defines a group, where $(f_1\circ f_2)(x)$ is equal to $f_1(f_2(x))$. For each pair of graphs $G_1,G_2$ we also define $\iso(G_1,G_2) = \{f\vert \mbox{$f$ is an isomorphism from $G_1$ to $G_2$}\}$.

\begin{lemma}\label{aut}
	Let $G_1$ and $G_2$ be isomorphic graphs. We have that $\vert \iso(G_1,G_2) \vert = \vert \aut(G_1) \vert = \vert \aut(G_2) \vert$.
\end{lemma}
\begin{proof}
	Let $\phi:G_1\to G_2$ be some isomorphism from $G_1$ to $G_2$. First we will prove that $\vert \iso(G_1,G_2) \vert  = \vert \aut(G_1) \vert$ by providing a bijection from $\iso(G_1,G_2)$ to $\aut(G_1)$. Let $F:\iso(G_1,G_2)\to\aut(G_1)$ be defined as $F(f) = \phi^{-1}\circ f$. $F(f)$ is clearly an automorphism of $G_1$. Let $f_1,f_2\in\iso(G_1,G_2)$ be such that for some $v\in G_1$, $f_1(v)\neq f_2(v)$. Clearly $\phi^{-1}\circ f_1(v) \neq \phi^{-1}\circ f_2(v)$ so the function is injective. Now consider $G(g) = \phi \circ g$. For each automorphism $g\in\aut(G_1)$, there exists some $f\in\iso(G_1,G_2)$ such that $F(f) = g$, given by $f = G(g)$, so the function is surjection. This directly implies that $\vert \iso(G_2,G_1) \vert  = \vert \aut(G_2) \vert$. Clearly $\vert\iso(G_1,G_2)\vert = \vert\iso(G_2,G_1)\vert$, and this completes the proof.
\end{proof}

Given a graph $G$ and vertices $v_1,\ldots,v_k$ in $G$, we define the labelled graph $G_{v_1,\ldots,v_k}$ as a copy of $G$ with labels in $v_1,\ldots,v_k$. These labels provide the restriction that each automorphism $g$ in $G_{v_1,\ldots,v_k}$ is such that $g(v_i) = v_i$ for each $v_i$. This can be achieved by connecting each $v_i$ with a line of $n+i$ vertices, where $n = \vert G \vert$. We also have that this construction takes polynomial time. Clearly, $\aut(G_{v_1,\ldots,v_k})\subseteq\aut(G)$.

Given a graph $G$, we define 
\[
\pi_{v}^G = \{v'\mid \mbox{there is some automorphism $g$ of $G$ such that $g(v) = v'$}\}.
\]

\begin{lemma}
	For a given graph $G$ and some $v$ in $G$, $\pi_v^G$ is an equivalence class of $V$.
\end{lemma}
\begin{proof}
	We define the relation $\sigma\subseteq G\times G$ as $\sigma(u,v)$ if there is some automorphism $g$ in $G$ such that $g(u) = v$. This is clearly an equivalence relation. Let $v_1,\ldots,v_n$ be the vertices in $G$. We have that $\{\pi_{v_1}^G,\ldots,\pi_{v_1}^G\}$ are the equivalence classes of $\sigma$ over $G$.
\end{proof}

\begin{lemma}\label{induc}
	For a graph $G$ and vertices $v_1,\ldots,v_k$ in $G$, $\vert aut(G_{v_1,\ldots,v_{k-1}}) \vert = \vert \pi_{v_k}^{G_{v_1,\ldots,v_{k-1}}} \vert \cdot \vert aut(G_{v_1,\ldots,v_k}) \vert$.
\end{lemma}
\begin{proof}
	For each $v_i\in\pi^{G_{v_1,\ldots,v_{k-1}}}_{v_k}$, let $\phi_i$ be some automorphism of $G_{v_1,\ldots,v_{k-1}}$ such that $\phi_i(v_k) = v_i$. Let $\Phi$ be the set of all such $\phi_i$'s. By definition, $\vert \pi_{v_k}^{G_{v_1,\ldots,v_{k-1}}} \vert = \vert \Phi \vert$. To prove the initial statement, we will provide a bijection $F:aut(G_{v_1,\ldots,v_{k-1}}) \to \Phi\times aut(G_{v_1,\ldots,v_k})$. Consider $F(g) = (\phi_i, \phi_i^{-1}\circ g)$, where $g(v_k) = v_i$. By definition, $\phi_i\in\Phi$. We have that $\phi_i^{-1}\circ g$ is an automorphism of $G_{v_1,\ldots,v_{k-1}}$, and since $\phi_i^{-1}\circ g(v_k) = v_k$, we also have that $\phi_i^{-1}\circ g$ is an automorphism of $G_{v_1,\ldots,v_{k-1}}$. Since $\phi_i^{-1}\circ g_1 = \phi_i^{-1} \circ g_2$ implies that $g_1 = g_2$, this is an injection. To prove that this is a surjection, note that for each pair $(\phi_i,h)\in\Phi\times\aut(G_{v_1,\ldots,v_k})$, there exists $g = \phi_i\circ h$ such that $F(g) = h$. Since the function is a bijection, this completes the proof.
\end{proof}

Before we proceed to the main theorem, we prove a small lemma.

\begin{lemma}\label{polypi}
	For a graph $G$ and vertices $v_1,\ldots,v_k$ in $G$, $\delta(G,v_1,\ldots,v_k) = \vert \pi^{G_{v_1,\ldots,v_{k-1}}}_{v_k} \vert$ can be computed by a polynomial-time procedure with an oracle to $\lgi$.
\end{lemma}
\begin{proof}
	For a vertex $v\in G$, let $G^{v}$ be a copy of $G$ with a line of size $2n+1$ attached to $v$, where $\vert G \vert = n$. Let $S$ be the set of vertices in $G\setminus\{v_1,\ldots,v_k\}$. For each $v\in S$, we can check with the oracle in constant time if $G^{v}_{v_1,\ldots,v_{k-1}}$ and $G^{v_k}_{v_1,\ldots,v_{k-1}}$ are isomorphic. We have that this pair of graphs is isomorphic if and only if there is an automorphism $g$ of $G_{v_1,\ldots,v_{k-1}}$ such that $g(v_k) = v$, or equivalently, $v\in \pi^{G_{v_1,\ldots,v_{k-1}}}_{v_k}$. Repeating this call $\vert S \vert$ times gives us the exact value of $\vert \pi^{G_{v_1,\ldots,v_{k-1}}}_{v_k} \vert$.
\end{proof}

\begin{theo}
	$\shgi\in\fp^{\lgi}$.
\end{theo}
\begin{proof}
	We will construct a polynomial-time procedure with an oracle to $\lgi$ problem that given a pair of graphs $(G_1,G_2)$ computes $\shgi(G_1,G_2) = \vert \iso(G_1,G_2) \vert$. First, we verify if $G_1$ and $G_2$ are isomorphic with a single call to the oracle. If they are not, return 0. Let $v_1,\ldots,v_n$ be the vertices in $G_1$. From Lemma \ref{induc} We have that $\vert\aut(G_1)\vert = \vert \pi^{v_2}_{v_1} \vert \cdot \vert \pi^{v_3}_{v_1,v_2} \vert \cdots \vert \pi^{v_n}_{v_1,\ldots,v_n} \vert\cdot\vert\aut({G_1}_{v_1,\ldots,v_n}) \vert$. Note that $\vert\aut({G_1}_{v_1,\ldots,v_n}) \vert = 1$. Also note that from the result in Lemma \ref{polypi}, each $\vert\pi^{v_k}_{v_1,\ldots,v_{k-1}}\vert$ can be computed in polynomial time, and therefore, $\vert \aut(G_1) \vert$ can also be computed in polynomial time. From Lemma \ref{aut} we have that $\vert \aut(G_1) \vert = \vert \iso(G_1,G_2) \vert$ and this completes the proof.
\end{proof}

\end{document}

### 2016-06-04 tex/gi.tex
new ver gi.tex

@@
 \usepackage{fullpage}
+\usepackage{bm}
+
+\usepackage{amsfonts}
+\usepackage{amssymb}
+\usepackage{amsmath}
 
@@
 
-In this document we provide a comprehensive proof of Mathon's result that counting the isomorphisms from a pair of graphs can be done with a PTIME machine with an oracle to the graph isomorphism language.
+In this document we provide a comprehensive proof of Mathon's result that counting the isomorphisms from a pair of graphs can be done with a polynomial-time machine with an oracle to the graph isomorphism language.\cite{DBLP:journals/ipl/Mathon79}
 
@@
 \]
-Before jumping to the main result, we provide some definitions.
+Before going into the proof, we provide some definitions.
 
 For each graph $G$ and vertices $v_1,\ldots,v_k$ in $G$, we define $G_{v_1,\ldots,v_k}$ as a copy of $G$ with distinct labels on $v_1,\ldots,v_k$.
-This can be done by connecting each $v_i$ with some line of length $n+k$, where $\vert G \vert = n$.
+This can be done by connecting each $v_i$ with some line of length $n+i$, where $\vert G \vert = n$.
 
 For each graph $G$, we define $\aut(G) = \{g\mid \mbox{$g$ is an automorphism of $G$} \}$.
-Note that $(\aut(G),\circ)$ defines a group, where $(f_1\circ f_2)(x)$ is equal to $f_1(f_2(x))$.
+Note that $(\aut(G),\circ)$ defines a group, where $(f_1\circ f_2)$ represents the composite function of $f_1$ over $f_2$.
 For each pair of graphs $G_1,G_2$ we also define $\iso(G_1,G_2) = \{f\vert \mbox{$f$ is an isomorphism from $G_1$ to $G_2$}\}$.
+
+Before we proceed to the main theorem, we prove a series of lemma.
 
@@
 $F(f)$ is clearly an automorphism of $G_1$.
-Let $f_1,f_2\in\iso(G_1,G_2)$ be such that for some $v\in G_1$, $f_1(v)\neq f_2(v)$.
-Clearly $\phi^{-1}\circ f_1(v) \neq \phi^{-1}\circ f_2(v)$ so the function is injective.
-Now consider $G(g) = \phi \circ g$.
-For each automorphism $g\in\aut(G_1)$, there exists some $f\in\iso(G_1,G_2)$ such that $F(f) = g$, given by $f = G(g)$, so the function is surjection.
-This directly implies that $\vert \iso(G_2,G_1) \vert  = \vert \aut(G_2) \vert$.
+Since $\phi^{-1}\circ f_1 = \phi^{-1}\circ f_2$ implies that $f_1 = f_2$ the function is injective.
+For each $g\in\aut(G_1)$, there exists some $f\in\iso(G_1,G_2)$ such that $F(f) = g$, given by $f = \phi \circ g$, so the function is surjective.
+This directly gives us that $\vert \iso(G_2,G_1) \vert  = \vert \aut(G_2) \vert$.
 Clearly $\vert\iso(G_1,G_2)\vert = \vert\iso(G_2,G_1)\vert$, and this completes the proof.
@@
 Given a graph $G$ and vertices $v_1,\ldots,v_k$ in $G$, we define the labelled graph $G_{v_1,\ldots,v_k}$ as a copy of $G$ with labels in $v_1,\ldots,v_k$.
-These labels provide the restriction that each automorphism $g$ in $G_{v_1,\ldots,v_k}$ is such that $g(v_i) = v_i$ for each $v_i$.
+These labels provide the restriction that each automorphism $g$ in $G_{v_1,\ldots,v_k}$ is such that $g(v_i) = v_i$ for $1\leq i\leq k$.
 This can be achieved by connecting each $v_i$ with a line of $n+i$ vertices, where $n = \vert G \vert$.
 We also have that this construction takes polynomial time.
-Clearly, $\aut(G_{v_1,\ldots,v_k})\subseteq\aut(G)$.
+Furthermore, $\aut(G_{v_1,\ldots,v_k})\subseteq\aut(G)$.
 
-Given a graph $G$, we define 
+Given a graph $G$, we define the orbit of $v$ in $G$ as
 \[
@@
 \begin{lemma}
-	For a given graph $G$ and some $v$ in $G$, $\pi_v^G$ is an equivalence class of $V$.
+	For a given graph $G$ with a set of vertices $V$ and some $v\in V$, $\pi_v^G$ is an equivalence class of $V$.
 \end{lemma}
@@
 This is clearly an equivalence relation.
-Let $v_1,\ldots,v_n$ be the vertices in $G$.
+Let $V = \{v_1,\ldots,v_n\}$.
 We have that $\{\pi_{v_1}^G,\ldots,\pi_{v_1}^G\}$ are the equivalence classes of $\sigma$ over $G$.
@@
 	For each $v_i\in\pi^{G_{v_1,\ldots,v_{k-1}}}_{v_k}$, let $\phi_i$ be some automorphism of $G_{v_1,\ldots,v_{k-1}}$ such that $\phi_i(v_k) = v_i$.
-Let $\Phi$ be the set of all such $\phi_i$'s.
+Let $\Phi$ be the set of the chosen $\phi_i$'s.
 By definition, $\vert \pi_{v_k}^{G_{v_1,\ldots,v_{k-1}}} \vert = \vert \Phi \vert$.
@@
 \end{proof}
-
-Before we proceed to the main theorem, we prove a small lemma.
 
@@
 \end{proof}
+
+Now we proceed to the main result.
 
@@
 
+\bibliographystyle{plain}
+\bibliography{Mathon79}
+
 \end{document}

### 2016-06-22 tex/below.tex
22 jun

@@
 
-Now we will prove the non-trivial inclusions.
+Or new addition to the known diagram is the following inclusion.
 
 \begin{theo}
-	$\shp \subseteq \totp^{\textsc{SAT}}$
+	$\spanp \subseteq \totp^{\np}$
 \end{theo}
 \begin{proof}
-	Let $f\in\shp$ and let $M$ be the non-deterministic polynomial-time machine such that $f = \acc_M$.
-We provide a machine $M'^{\textsc{SAT}}$ with an oracle of SAT using the following construction.
-On input $w$, $M'$ gives the oracle a formula $\varphi$ that can be satisfied if and only if $M$ has at least one accepting path on input $w$.
-If it answers {\it no}, the computation ends.
-If it answers {\it yes}, it chooses $x\in\{0,1\}$.
-If $x = 0$ stops, and if $x = 1$, then it does the following.
-$M'$ simulates $M$ until a non-deterministic choice has to be taken.
-Let $S$ be the set of configurations that are reached on each choice.
-Then, for each $s\in S$, $M'$ constructs a formula $\varphi_s$ that models the machine $M$ on input $w$, starting on configuration $s$, which is satisfiable if and only if $M$ has at least one accepting path.
-For each $s\in S$, the formula $\varphi_s$ is given to the oracle, and if it answers {\it yes}, the corresponding choice is taken non-deterministically.
-If it answers {\it no} the choice is ignored.
-This is repeated until no more choices can be made.
-It is clear that $\acc_M = \tot_{M'^{\textsc{SAT}}}$.
-We conclude that $f\in\totp^{\textsc{SAT}}$.
+	Let $f\in\spanp$ and let $M$ be the non-deterministic polynomial-time machine for which $f = \sp_M$.
+We suppose without loss of generality that each output of $M(x)$ has length $p(\vert x \vert)$, where $p$ is a polynomial.
+We will provide a non-deterministic machine $M'^L$ where $L$ is the following language:
+	\begin{multline*}
+	L = \{(M,x,w_1,w_2) \mid \text{$M$ is a non-deterministic polynomial-time machine and} \\ \text{$M(x)$ gives some output in the lexicographic range $[w_1,w_2]$.} \}
+	\end{multline*}
+	The machine $M$ on input $x$ writes $w_1 = 0^{p(\vert x \vert)}, w_2 = 01^{p(\vert x \vert)-1}$, and asks the oracle $(M,x,w_1,w_2)$.
+If it answers yes, it generates a new branch and proceeds, if not, just proceeds.
+Then, write $w_1 = 10^{p(\vert x \vert)-1}, w_2 = 1^{p(\vert x \vert)}$ and asks the oracle $(M,x,w_1,w_2)$.
+If it answers yes, it generates a new branch and proceeds, if not, stop.
+TO BE CONTINUED.
 \end{proof}
@@
 Similarly, we define $\stspanp$ (suggested pronunciation: ``strict span P'') as the class of functions $\sp!_M$ where $M$ is a non-deterministic polynomial time machine.
-
-To prove the inclusion $\spanp \subseteq \shp^{\np}$, we will use a more strong result.
-
-\begin{theo}
-	$\spanp \subseteq \stspanp$.
-\end{theo}
-\begin{proof}
-	Let $f \in \spanp$ and let $M$ be the non-deterministic machine for which $f = \sp_M$ that runs in $p(\vert x \vert)$ time.
-We construct a non-deterministic machine $M'$ that on input $w$ does the following.
-It simulates $M(x)$, and when the simulation finishes on a final state of $M$, stops.
-When it finishes on a non-final state, it appends a dummy sequence of characters, for example $0^{p(\vert x \vert)+1}$, to the output.
-Clearly $f = \sp!_{M'}$, and since $M'$ runs in polynomial time, we have that $f \in \stspanp$.
-\end{proof}
 

### 2016-07-13 tex/below.tex
13 jul

@@
 We define $\totp$ using the following function. For each non-deterministic Turing machine $M$ and $x\in\{0,1\}^*$,
-\[
+$$
 	\tot_M(x) = (\text{\#paths of $M$ on input $x$}) - 1.
-\]
+$$
 A function $f:\{0,1\}^*\to\N$ is in $\totp$ if there exists a polynomial-time non-deterministic Turing machine $M$ such that $f = \tot_M$.
+
+For a non-deterministic Turing machine $M$ with an output tape, we define $\out_M(x)$ as the set of the outputs of each non-deterministic computation of $M$ on input $x$.
 
@@
 \[
-	\sp_M(x) = \vert\{w \mid \text{$w$ is an output of $M$ on input $w$}\}\vert.
+	\sp_M(x) = \vert\out_M(x)\vert.
 \]
@@
 
-Given a language $L$, $\totp^{L}$ is defined by all functions $f:\{0,1\}^*\to\N$ such that $f = tot_{M^L}$.
+Similarly, a function  $f:\{0,1\}^*\to\N$ is in $\spanp$ if there exists a polynomial-time non-deterministic Turing machine $M$ such that $f = \sp_M$.
+
+Given a language $L$, $\totp^{L}$ is defined by all functions $f:\{0,1\}^*\to\N$ such that $M$ is a non-deterministic polynomial machine and $f = \tot_{M^L}$.
 \\
 
-By the following result, we note that $\shp$ can be considered an extension of $\totp$.
-
-\begin{theo}
-	$\spanl \subseteq \totp$.
-\end{theo}
-\begin{proof}
-	Let $f:\{0,1\}^*\to\N$ be a function in $\spanl$ and let $M$ be a non-deterministic log-space Turing machine such that $f = \sp_M$.
-We will provide a polynomial-time non-deterministic Turing machine $M'$ such that for each $x\in\{0,1\}^*$ it holds that $\sp_M(x) = \tot_{M'}(x)$.
-First we note that given a configuration $s$ of $M$ on input $x$ checking whether there exists a final state of $M$ which is reachable from $s$ can be done in polynomial time.
-This implies that checking whether $f(x) > 0$ can also be done in polynomial time.
-We present this machine $M'$ as a non-deterministic algorithm.
-
-	\begin{algorithm}	
-	\caption{Compute $f(x)+1$ paths}
-	\begin{algorithmic}[]
-		\Begin
-		\If{$f(x) = 0$} stop
-		\Else \text{ choose from}
-		\State stop
-		\State call $\textsc{GenTree}_{f}(x,\{s_0\},\epsilon)$
-		\EndIf
-		\EndBegin
-		\Procedure{$\textsc{GenTree}_{f}(x,S,y)$}{}
-		\State $S_0 = \emptyset$, $S_1 = \emptyset$
-		\For{$s\in S$}
-		\State do a search in the configuration graph of $M$ with input $x$ from $s$:
-		\If{$s'$ is reached after writing 0 in the output tape} $S_0 \gets s'$
-		\EndIf
-		\If{$s'$ is reached after writing 1 in the output tape} $S_1 \gets s'$
-		\EndIf
-		\EndFor
-		\If{a final state from $M$ is reachable from some $s\in S_0$ and some $s'\in S_1$} choose between:
-		\State call $\textsc{GenTree}_{f}(x,S_0,y0)$
-		\State call $\textsc{GenTree}_{f}(x,S_1,y1)$
-		\ElsIf{a final state from $M$ is reachable from some $s\in S_i$ but not from any $s'\in S_{1-i}$}
-		\State call $\textsc{GenTree}_{f}(x,S_i,yi)$
-		\ElsIf{no final state from $M$ is reachable from any $s\in S_0 \cup S_1$}
-		\Return $y$
-		\EndIf
-		\EndProcedure
-	\end{algorithmic}
-	\end{algorithm}
-
-	We use $s_0$ for the initial state of $M$ with input $x$.
-The algorithm can be executed in a nondeterministic polynomial-time machine, so we only need to show that the number of computation paths is equal to $\sp_M(x)-1$.
-
-	For each output $y$ consider some computation path $p$ of $M$ with input $x$ that outputs $y$.
-Note there is at least one computation path in $M'$ that follows $p$ and outputs $y$.
-Also, note that each distinct computation path of $M'$ outputs a different $y$, since in each branch the word differs in the last bit.
-Taking into account the dummy path in line 4, we conclude that the number of computation paths of $M'$ with input $x$ is equal to $\sp_M(x)+1$.
-
-\end{proof}
+Now we will show some properties of $\totp$.
 
@@
 for each $w\in\{0,1\}^*$. A function $f:\{0,1\}^*\to\N$ is in FP if there exists a deterministic polynomial-time machine $M$ such that $f = f_M$.
+
+The following proof is added as a completion of a theorem of Pagourtzis and Zachos \cite{DBLP:conf/mfcs/PagourtzisZ06}.
 \begin{theo} \label{diff}
-	For each function $f\in\shp$ there exists functions $g\in \totp$ and $h\in\text{FP}$ such that $f = g - h$.
+	For each function $f\in\shp$ there exist functions $g\in \totp$ and $h\in\text{FP}$ such that $f = g - h$.
 \end{theo}
@@
 	We will now show that $g(w) = \tot_{M'}(w)$.
-Let $w\in\{0,1\}$.
+Let $w\in\{0,1\}^*$.
 If $\tot_{M}(w) > p(\vert w \vert)$, then clearly $\tot_{M'}(w) = \tot_{M}(w) - p(\vert w \vert)$.
@@
 
-We present a counting class diagram:
+We also define the class $\maxp$ (resp.
+$\minp$) as the class of functions $f:\{0,1\}^* \to \N$ where there exists some non-deterministic polynomial-time machine $M$ such that for each $x\in\{0,1\}^*$ it holds that $f(x) = \max(\out_M(x))$ (resp.
+$\min(\out_M(x))$).
+
+We present an updated class diagram of the classes surrounding $\totp$.
 
@@
 
-Or new addition to the known diagram is the following inclusion.
+Inclusions are indicated with arrows. Separation results under complexity assumptions are indicated with dashed arrows.
+
+The inclusion $\shl \subseteq \spanl$ and the fact that $\spanl \subseteq \fp$ implies $\ph = \text{P}^{\np}$ were showed by Àlvarez and Jenner $\cite{DBLP:journals/tcs/AlvarezJ93}$.
+The inclusions $\fp \subseteq \totp$ and $\totp \subseteq \shp$, and the fact that $\shp \subseteq \totp$ implies $\text{P} = \np$ were showed by Pagourtzis and Zachos in \cite{DBLP:conf/mfcs/PagourtzisZ06}.
+The inclusions $\fp^{\np} \subseteq \totp^{\np}$ and $\totp^{\np} \subseteq \shnp = \shp^{\np}$ are trivial relativizations.
+The inclusions $\minp \subseteq \fp^{\np}$, $\maxp \subseteq \fp^{\np}$, $\maxp \subseteq \spanp$ and $\shp \subseteq \spanp$, and the fact that $\spanp \subseteq \shp$ implies that $\np = \up$ were showed by Kobler et.
+al.
+in \cite{DBLP:journals/acta/KoblerST89}.
+The inclusion $\fp \subseteq fp^{\np}$ is trivial.
+The inclusion $\shl \subseteq \fp$ was showed in [COMPLETE THIS].
+
+\begin{theo}
+	$\spanl \subseteq \totp$.
+\end{theo}
+\begin{proof}
+	Let $f:\{0,1\}^*\to\N$ be a function in $\spanl$ and let $M$ be a non-deterministic log-space Turing machine such that $f = \sp_M$.
+We will provide a polynomial-time non-deterministic Turing machine $M'$ such that for each $x\in\{0,1\}^*$ it holds that $\sp_M(x) = \tot_{M'}(x)$.
+First we note that given a configuration $s$ of $M$ on input $x$ checking whether there exists a final state of $M$ which is reachable from $s$ can be done in polynomial time.
+This implies that checking whether $f(x) > 0$ can also be done in polynomial time.
+We present this machine $M'$ as a non-deterministic procedure in Algorithm \ref{spanltot}.
+
+	\begin{algorithm}	
+		\caption{Compute $f(x)+1$ paths} \label{spanltot}
+		\begin{algorithmic}[]
+			\Begin
+			\If{$f(x) = 0$} stop
+			\Else \text{ choose from}
+			\State stop
+			\State call $\textsc{GenTree}_{f}(x,\{s_0\},\epsilon)$
+			\EndIf
+			\EndBegin
+			\Procedure{$\textsc{GenTree}_{f}(x,S,y)$}{}
+			\State $S_0 = \emptyset$, $S_1 = \emptyset$
+			\For{$s\in S$}
+			\State do a search in the configuration graph of $M$ with input $x$ from $s$:
+			\If{$s'$ is reached after writing 0 in the output tape} $S_0 \gets s'$
+			\EndIf
+			\If{$s'$ is reached after writing 1 in the output tape} $S_1 \gets s'$
+			\EndIf
+			\EndFor
+			\If{a final state from $M$ is reachable from some $s\in S_0$ and some $s'\in S_1$} choose between:
+			\State call $\textsc{GenTree}_{f}(x,S_0,y0)$
+			\State call $\textsc{GenTree}_{f}(x,S_1,y1)$
+			\ElsIf{a final state from $M$ is reachable from some $s\in S_i$ but not from any $s'\in S_{1-i}$}
+			\State call $\textsc{GenTree}_{f}(x,S_i,yi)$
+			\ElsIf{no final state from $M$ is reachable from any $s\in S_0 \cup S_1$}
+			\Return $y$
+			\EndIf
+			\EndProcedure
+		\end{algorithmic}
+	\end{algorithm}
+
+	We use $s_0$ for the initial state of $M$ with input $x$.
+The algorithm can be executed in a nondeterministic polynomial-time machine, so we only need to show that the number of computation paths is equal to $\sp_M(x)-1$.
+
+	For each output $y \in \out_M(x)$ consider some computation path $p$ of $M$ with input $x$ that outputs $y$.
+Note there is at least one computation path in $M'$ that follows $p$ and outputs $y$.
+Also, note that each distinct computation path of $M'$ outputs a different $y$, since in each branch the word differs in the last bit.
+Taking into account the dummy path in line 4, we conclude that the number of computation paths of $M'$ with input $x$ is equal to $\sp_M(x)+1$.
+\end{proof}
 
@@
 	\end{multline*}
-	The machine $M$ on input $x$ writes $w_1 = 0^{p(\vert x \vert)}, w_2 = 01^{p(\vert x \vert)-1}$, and asks the oracle $(M,x,w_1,w_2)$.
-If it answers yes, it generates a new branch and proceeds, if not, just proceeds.
-Then, write $w_1 = 10^{p(\vert x \vert)-1}, w_2 = 1^{p(\vert x \vert)}$ and asks the oracle $(M,x,w_1,w_2)$.
-If it answers yes, it generates a new branch and proceeds, if not, stop.
-TO BE CONTINUED.
+	This is presented as a non-deterministic procedure in Algorithm \ref{spanptotnp}.
+
+		\begin{algorithm}	
+			\caption{Compute $f(x)+1$ paths} \label{spanptotnp}
+			\begin{algorithmic}[]
+				\Begin
+				\If{$f(x) = 0$} stop
+				\Else \text{ choose from}
+				\State stop
+				\State call $\textsc{GenTree}_{f}(x,\varepsilon,0)$
+				\EndIf
+				\EndBegin
+				\Procedure{$\textsc{GenTree}_{f}(x,w,m)$}{}
+				\If{$m = p(\vert x \vert)$}
+				\State stop
+				\Else
+				\State $w_1 = w0^{p(\vert x \vert)-m}$
+				\State $w_2 = w01^{p(\vert x \vert)-m-1}$
+				\If{$(M,x,w_1,w_2)\in L$}
+				\State call $\textsc{GenTree}_{f}(x,w0,m+1)$
+				\EndIf
+				\State $w_1 = w10^{p(\vert x \vert)-m-1}$
+				\State $w_2 = w1^{p(\vert x \vert)-m}$
+				\If{$(M,x,w_1,w_2)\in L$}
+				\State call $\textsc{GenTree}_{f}(x,w1,m+1)$
+				\EndIf
+				\EndIf
+				\EndProcedure
+			\end{algorithmic}
+		\end{algorithm}
+
+		Note that on input $x$, for each $w \in \out_M(x)$ there exists some branch in the algorithm that calls $\textsc{GenTree}_{f}(x,w,p(\vert x \vert))$ and then stops.
+Also note that for each branch in the algorithm that is not the dummy path in line 4, there is some $w$ in the last call to $\textsc{GenTree}_{f}$ that belongs to $\out_M(x)$.
+From this we conclude that $f(x) = \tot_{M'^L}(x)+1$.
 \end{proof}
@@
 Similarly, we define $\stspanp$ (suggested pronunciation: ``strict span P'') as the class of functions $\sp!_M$ where $M$ is a non-deterministic polynomial time machine.
+
+We provide a characterization of $\shnp$ that doesn't use oracles and lets each function in $\shnp$ be defined by a single non-deterministic machine.
+We contrast this result with the one showed by Kobler et.
+al.
+in \cite{DBLP:journals/acta/KoblerST89} that lets define each function in $\shnp$ without oracles using two machines.
 

### 2016-07-26 QSO/cap-comp.tex
Archivos de QSO

Assume that $\SR = (S, \adds, \mults, \zero, \one)$ is a semiring and $\theta$ is a sentence in $\qso(\SR)$. Then slightly abusing notation, we also use $\theta$ to denote a function from $\R$ to $\SR$ such that for every $\A \in \str[\R] $:
\begin{eqnarray*}
\theta(\enc(\A)) & = & \sem{\theta}(\A).
\end{eqnarray*}
\begin{definition}
Let $\SR$ be a semiring, $\FF$ a fragment of $\qso(\SR)$, $\CC$ a function complexity class over $\SR$ and $\KK$ a class of finite structures. Then $\FF$ {\em captures} $\CC$ over $\KK$ if:
\begin{itemize}
\item for every $\theta$ in $\FF$, there exists $f \in \CC$ such that $\res{\theta}{\KK} = \res{f}{\KK}$; and

\item for every $f \in \CC$, there exists $\theta$ in $\FF$ such that $\res{f}{\KK} = \res{\theta}{\KK}$.
\end{itemize}
\end{definition}
In the previous definition, if $\KK$ is the class of all finite structures, then we say that $\FF$ captures $\CC$ over the class of all structures, and if $\KK$is the class of all ordered finite structures, then we say that $\FF$ captures $\CC$ over the class of ordered structures. 
\begin{theorem}
$\eqso(\nat,\fo)$ captures $\shp$ over the class of ordered structures.
\end{theorem}

\begin{proposition}
$\eqso(\nat,\fo)$ does not capture $\shp$ over the class of all structures.
\end{proposition}

\begin{theorem}
$\eqso(\integ,\fo)$ captures $\gp$ over the class of ordered structures.
\end{theorem}

\begin{theorem}
$\eqso(\nat,\eso)$ captures $\spp$ over the class of ordered structures.
\end{theorem}

### 2016-07-26 QSO/logics.tex
Archivos de QSO

Fix a semiring $\SR = (S, \adds, \mults, \zero, \one)$. Then given a relational signature $\R$, the set of $\SR$-quantitative second-order logic formulas (or just $\qso(\SR)$-formulas) over $\R$ is given by the following grammar:
\begin{eqnarray}
\label{eq-def-qso}
  \theta & := & \varphi \ \mid \ s \ \mid \ (\theta \add \theta) \ \mid\ (\theta \mult \theta) \ \mid \ \sa{x} \theta \ \mid \ \pa{x} \theta \ \mid \ \sa{X} \theta \ \mid \ \pa{X} \theta 
\end{eqnarray}
where $\varphi$ is an $\so$-formula over $\R$, $s \in S$, $x \in \fv$ and $X \in \sv$.

Let $\R$ be a relational signature, $\A$ be a finite $\R$-structure with domain $A$, $v$ a first-order variable assignment for $\A$ and $V$ a second-order variable assignment for $\A$. Then the \emph{evaluation} of a $\qso(\SR)$-formula $\theta$ over $(\A, v, V)$ is defined as a function $\sem{\theta}$ that on input $(\A, v, V)$ returns a value in the semiring $\SR$. Formally, the function $\sem{\theta}$ is recursively defined as follows:
\begin{eqnarray*}
\sem{\varphi}(\A, v, V) & = & 
\begin{cases}
\one & \mbox{if } (\A, v, V) \models \varphi \\
\zero & \mbox{otherwise}
\end{cases}\\
\sem{s}(\A, v, V) & = & s \\
\sem{\theta_1 \add \theta_2}(\A, v, V) & = & \sem{\theta_1}(\A, v, V) \adds \sem{\theta_2}(\A, v, V)\\
\sem{\theta_1 \mult \theta_2}(\A, v, V) & = & \sem{\theta_1}(\A, v, V) \mults \sem{\theta_2}(\A, v, V)\\ 
\sem{\sa{x} \theta}(\A, v, V) & = & \bigadds_{a \in A} \sem{\theta}(\A,v[a/x],V)\\
\sem{\pa{x} \theta}(\A, v, V) & = & \bigmults_{a \in A} \sem{\theta}(\A,v[a/x],V)\\
\sem{\sa{X} \theta}(\A, v, V) & = & \bigadds_{B \subseteq A^{\arity(X)}} \sem{\theta}(\A, v, V[B/X])\\
\sem{\pa{X} \theta}(\A, v, V) & = & \bigmults_{B \subseteq A^{\arity(X)}} \sem{\theta}(\A, v, V[B/X])
\end{eqnarray*}
A formula in $\qso(\SR)$ can mention the usual quantifiers in $\so$ (that is, $\exists x$ and $\exists X$) and the quantifiers that make use of the two operations in the semiring (that is, $\Sigma x$, $\Pi x$, $\Sigma X$ and $\Pi X$), which are called {\em semiring quantifiers} . A $\qso(\SR)$-formula $\theta$ is said to be a \emph{sentence} if it does not have any free variable, that is, every variable in $\theta$ is under the scope of a usual quantifier or a semiring quantifier. It is important to notice that if $\theta$ is a $\qso(\SR)$-sentence over a relational signature $\R$, then for every finite $\R$-structure $\A$, first-order variable assignments $v_1$, $v_2$ for $\A$ and second-order variable assignments $V_1$, $V_2$ for $\A$, it holds that:
\begin{eqnarray*}
\sem{\theta}(\A, v_1, V_1) & = & \sem{\theta}(\A, v_2, V_2).
\end{eqnarray*}
Thus, in such a case we use the term $\sem{\theta}(\A)$ to denote $\sem{\theta}(\A, v, V)$, for some arbitrary first-order variable assignment $v$ for $\A$ and some arbitrary second-order variable assignment $V$ for $\A$.

In this paper, we consider several fragments of $\qso(\SR)$, which are obtained by restricting the syntax of the formula $\varphi$ in \eqref{eq-def-qso} or the use of the semiring quantifiers. Let $\eqso(\SR)$ be the fragment of $\qso(\SR)$ that is obtained by only allowing the semiring quantifiers $\Sigma x$, $\Pi x$ and $\Sigma X$ in \eqref{eq-def-qso}. Moreover, assuming that $\LL$ is a fragment of $\so$, let $\qso(\SR, \LL)$ be the fragment of $\qso(\SR)$ obtained by restricting $\varphi$ in \eqref{eq-def-qso} to be a formula in $\LL$, and let $\eqso(\SR, \LL)$ be the fragment of $\eqso(\SR)$ obtained by imposing the same restriction. In particular, in this paper we consider the following fragments $\LL$: $\fo$, which is obtained by disallowing the use of the second-order quantifier ($\exists X$) in the formula $\varphi$ in \eqref{eq-def-qso}, and $\eso$, which is obtained by allowing $\varphi$ in \eqref{eq-def-qso} to be a formula of the form $\exists X_1 \cdots \exists X_k \, \psi$, where $\psi$ is an $\fo$-formula.

### 2016-07-26 QSO/preliminaries.tex
Archivos de QSO

\subsection{Logics}

A relational signature $\R$ is a finite set $\{R_1, \ldots, R_k\}$, where each $R_i$ ($1 \leq i \leq k$) is a relation name with an associate arity greater than 0, which is denoted by $\arity(R_i)$. A finite structure over $\R$ (or just finite $\R$-structure) is a tuple $\A = \langle A, R_1^\A, \ldots, R_k^\A \rangle$ such that $A$ is a finite set and $R_i^\A \subseteq A^{\arity(R_i)}$ for every $i \in \{1, \ldots, k\}$. A finite $\R$-structure $\A$ is said to be ordered if $<$ is a binary predicate name in $\R$ and $<^\A$ is a linear order on $A$. Let $\str[\R]$ be the class of all finite $\R$-structures and $\ostr[\R]$ be the class of all ordered finite $\R$-structures.

From now on, assume given disjoint infinite sets $\fv$ and $\sv$ of first-order variables and second-order variables, respectively. Notice that every variable in $\sv$ has an associated arity, which is denoted by $\arity(X)$. Then given a relational signature $\R$, the set of second-order logic formulas ($\so$-formulas) over $\R$ is given by the following grammar:
\begin{eqnarray*}\ 
\varphi &:=& R(\bar u) \ \mid\  
X(\bar v)  \ \mid\ 
\neg \varphi \ \mid\ 
(\varphi \vee \varphi) \ \mid\ 
\exists x \, \varphi \ \mid\ 
\exists X \, \varphi
\end{eqnarray*}
where $R \in \R$, $\bar u$ is a tuple of (non-necessarily distinct) variables from $\fv$ whose length is $\arity(R)$, $X \in \sv$, $\bar v$ is a tuple of (non-necessarily distinct) variables from $\fv$ whose length is $\arity(X)$, and $x \in \fv$.

To define the semantics of $\so$, we need to introduce some terminology. Given a relational signature $\R$ and a finite $\R$-structure $\A$ with domain $A$, a first-order variable-assignment $v$ for $\A$ is a total function from $\fv$ to $A$, while a second-order variable-assignment $V$ for $\A$ is a total function with domain $\sv$ that maps each $X \in \sv$ to a subset of $A^{\arity(X)}$. Moreover, given a first-order variable-assignment $v$ for $\A$, $x \in \fv$ and $a \in A$, we denote by $v[a/x]$ a first-order variable-assignment such that $v[a/x](x) = a$ and $v[a/x](y) = v(y)$ for every $y \in \fv$ distinct from $x$. Similarly, given a second-order variable-assignment $V$ for $\A$, $X \in \sv$ and $B  \subseteq A^{\arity(X)}$, we denote by $V[B/X]$ a second-order variable-assignment such that $V[B/X](X) = B$ and $V[B/X](Y) = V(Y)$ for every $Y \in \sv$ distinct from $X$.

Assume that $\varphi$ is an $\so$-formula over a signature $\R$. Then given a finite $\R$-structure $\A$ with domain $A$, a first-order variable-assignment $v$ for $\A$ and a second-order variable-assignment $V$ for $\A$, we say that $(\A, v, V)$ satisfies $\varphi$, denoted by $(\A, v, V) \models \varphi$, if: (1) $\varphi$ is the formula $R(x_1, \ldots, x_\ell)$ and $(v(x_1), \ldots, v(x_\ell)) \in R^\A$; (2) $\varphi$ is the formula $X(x_1, \ldots, x_m)$ and $(v(x_1), \ldots, v(x_m)) \in V(X)$; (3) $\varphi$ is the formula $\neg \psi$ and it not the case that $(\A, v, V) \models \psi$; (4) $\varphi$ is the formula $(\varphi_1 \vee \varphi_2)$, and $(\A, v, V) \models \varphi_1$ or $(\A, v, V) \models \varphi_2$; (5) $\varphi$ is the formula $\exists x \, \psi$ and there exists $a \in A$ such that $(\A, v[a/x], V) \models \psi$; or (6) $\varphi$ is the formula $\exists X \, \psi$ and there exists $B \subseteq A^{\arity(X)}$ such that $(\A, v, V[B/X]) \models \psi$.  As usual, we consider the propositional operators $\wedge$, $\rightarrow$, and $\leftrightarrow$ that can be obtained from $\vee$ and $\neg$. 
%Moreover, we use the abbreviations $x \not \leq y$ and $x \notin X$ for the negation of the atoms $\leq$ and $\in$. 
%Finally, we consider standard abbreviations of formulas that can be defined in $\so$-logic (actually, $\fo$-logic) like $\first(x) := \fa{x} y \leq x$ and $\last(x) := \fa{x} x \leq y$ to denote the first and last element of the linear order $\leq$, respectively, and $\succesor(x,y) := x \leq y \wedge y \not\leq x \wedge \fa{z} ( z \leq x \vee y \leq z)$ to denote the successor relation.

\subsection{Semirings}

A semiring is a structure $\SR = (S, \adds, \mults, \zero, \one)$ where $(S, \adds, \zero)$ is a commutative monoid, $(S, \mults, \one)$ is a monoid, multiplication ($\mults$) distributes over addition ($\adds$), and $\zero \mults s = s \mults \zero = \zero$ for every $s \in S$. If the multiplication is commutative, we say that $\SR$ is commutative. In this paper, we always assume that $\SR$ is  commutative.
For the sake of simplification, we usually denote the set of elements $S$ with the name of the semiring $\SR$.
% and we say sentences like ``for all $s \in \SR$''.

In general, our definitions are given with respect to a generic semiring.
However, in some cases we consider some particular semiring like, for example, the semiring of \emph{natural numbers} $\nat = (\mathbb{N}, +, \cdot, 0, 1)$, the semiring of \emph{integers} $\integ = (\mathbb{Z}, +, \cdot, 0, 1)$, the \emph{boolean} semiring $\bln = (\{\false, \true\}, \vee, \wedge, \false, \true)$, or the \emph{tropical} semiring $\trop = (\nat \cup \{\infty\}, \min, +, \infty, 0)$.
For the last semiring, the $\min$-operator is defined by the linear order $0 < 1 < 2 < \cdots < \infty$ and the $+$-operator is just the addition over natural number with the exception that $v + \infty = \infty + v = \infty$ for all $v \in \nat \cup \{\infty\}$.

\subsection{Function complexity classes}
Assume that $\R = \{R_1, \ldots, R_k\}$ is a relational signature and $\A$ is a finite $\R$-structure with a domain $A$ containing $n$ elements, and assume that  $<$ is a linear order on $A$, say $a_1 < a_2 < \ldots < a_n$. For every $i \in \{1, \ldots, k\}$, define the encoding of $R_i^\A$, denoted by $\enc(R_i^\A)$, as the following binary string. Assume that $\ell = \arity(R_i)$ and consider an enumeration of the $\ell$-tuples over $A$ in the lexicographic order induced by $<$ (that is, $(a_1, \ldots, a_1, a_1)$, $(a_1, \ldots, a_1, a_2)$, $\ldots$, $(a_n, \ldots, a_n, a_{n-1})$, $(a_n, \ldots, a_n, a_n)$). Then let $\enc(R_i^\A)$ be a binary string of length $n^\ell$ such that the $i$-th bit of $\enc(R_i^\A)$ is 1 if the $i$-th tuple in the previous enumeration belongs to $R_i^\A$, and 0 otherwise. Moreover, define the encoding of $\A$, denoted by $\enc(\A)$, as the following binary string~\cite{L04}:
\begin{eqnarray*}
\enc(\A) & = & 0^n \, 1 \, \enc(R_1^\A) \, \cdots \, \enc(R_k^\A).
\end{eqnarray*}
Given a semiring $\SR = (S, \adds, \mults, \zero, \one)$ and a relational signature $\R$, a function $f$ is said to be from $\R$ to $\SR$ if $f$ is a total function from $\{\enc(\A) \mid \A \in \str[\R]\}$ to $S$. Moreover, a set $\CC$ of functions is said to be a {\em function complexity class over $\SR$} if every $f \in \CC$ is a function from some relational signature $\R$ to $\SR$.

If $\KK$ is a class of finite structures and $f$ is a function from a relational signature $\R$ to a semiring $\SR$, we denote by $\res{f}{\KK}$ the restriction of $f$ to $\KK$, that is, a function such that: the domain of $\res{f}{\KK}$ is $\{ \enc(\A) \mid \A \in \str[\R] \cap \KK\}$, and 
$\res{f}{\KK}(\enc(\A)) = f(\enc(\A))$ for every $\A \in \str[\R] \cap \KK$.

### 2016-08-03 QSO/cap-comp.tex
Nueva versión de QSO.

@@
-Assume that $\SR = (S, \adds, \mults, \zero, \one)$ is a semiring and $\theta$ is a sentence in $\qso(\SR)$.
-Then slightly abusing notation, we also use $\theta$ to denote a function from $\R$ to $\SR$ such that for every $\A \in \str[\R] $:
+Assume that $\alpha$ is a sentence in $\qso$. Then slightly abusing notation, we also use $\alpha$ to denote a function from $\R$ to $\bbN$ such that for every $\A \in \str[\R] $:
 \begin{eqnarray*}
-\theta(\enc(\A)) & = & \sem{\theta}(\A).
+\alpha(\enc(\A)) & = & \sem{\alpha}(\A).
 \end{eqnarray*}
 \begin{definition}
-Let $\SR$ be a semiring, $\FF$ a fragment of $\qso(\SR)$, $\CC$ a function complexity class over $\SR$ and $\KK$ a class of finite structures. Then $\FF$ {\em captures} $\CC$ over $\KK$ if:
+Let $\FF$ a fragment of $\qso$, $\CC$ a function complexity class over $\bbN$ and $\KK$ a class of finite structures. Then $\FF$ {\em captures} $\CC$ over $\KK$ if:
 \begin{itemize}
-\item for every $\theta$ in $\FF$, there exists $f \in \CC$ such that $\res{\theta}{\KK} = \res{f}{\KK}$; and
+\item for every $\alpha$ in $\FF$, there exists $f \in \CC$ such that $\res{\alpha}{\KK} = \res{f}{\KK}$; and
 
-\item for every $f \in \CC$, there exists $\theta$ in $\FF$ such that $\res{f}{\KK} = \res{\theta}{\KK}$.
+\item for every $f \in \CC$, there exists $\alpha$ in $\FF$ such that $\res{f}{\KK} = \res{\alpha}{\KK}$.
 \end{itemize}
@@
 In the previous definition, if $\KK$ is the class of all finite structures, then we say that $\FF$ captures $\CC$ over the class of all structures, and if $\KK$is the class of all ordered finite structures, then we say that $\FF$ captures $\CC$ over the class of ordered structures.
+
+\cristian{Aca hay que definir esto sobre estructuras finitas directamente. No veo el sentido de tener estructuras infinitas.}
+
 \begin{theorem}
-$\eqso(\nat,\fo)$ captures $\shp$ over the class of ordered structures.
+$\eqso(\fo)$ captures $\shp$ over the class of ordered structures.
 \end{theorem}
@@
 \begin{proposition}
-$\eqso(\nat,\fo)$ does not capture $\shp$ over the class of all structures.
+$\eqso(\fo)$ does not capture $\shp$ over the class of all structures.
 \end{proposition}
 
-\begin{theorem}
-$\eqso(\integ,\fo)$ captures $\gp$ over the class of ordered structures.
-\end{theorem}
+%\begin{theorem}
+%$\eqso(\integ,\fo)$ captures $\gp$ over the class of ordered structures.
+%\end{theorem}
 
 \begin{theorem}
-$\eqso(\nat,\eso)$ captures $\spp$ over the class of ordered structures.
+$\eqso(\eso)$ captures $\spp$ over the class of ordered structures.
 \end{theorem}

### 2016-08-03 QSO/logics.tex
Nueva versión de QSO.

@@
-Fix a semiring $\SR = (S, \adds, \mults, \zero, \one)$.
-Then given a relational signature $\R$, the set of $\SR$-quantitative second-order logic formulas (or just $\qso(\SR)$-formulas) over $\R$ is given by the following grammar:
+Consider $\bbN$ as the set of non-negative integers.
+Then given a relational signature $\R$, the set of quantitative second-order logic formulas (or just $\qso$-formulas) over $\R$ is given by the following grammar:
 \begin{eqnarray}
 \label{eq-def-qso}
-  \theta & := & \varphi \ \mid \ s \ \mid \ (\theta \add \theta) \ \mid\ (\theta \mult \theta) \ \mid \ \sa{x} \theta \ \mid \ \pa{x} \theta \ \mid \ \sa{X} \theta \ \mid \ \pa{X} \theta 
+  \alpha & := & \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha 
 \end{eqnarray}
-where $\varphi$ is an $\so$-formula over $\R$, $s \in S$, $x \in \fv$ and $X \in \sv$.
+where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
 
-Let $\R$ be a relational signature, $\A$ be a finite $\R$-structure with domain $A$, $v$ a first-order variable assignment for $\A$ and $V$ a second-order variable assignment for $\A$.
-Then the \emph{evaluation} of a $\qso(\SR)$-formula $\theta$ over $(\A, v, V)$ is defined as a function $\sem{\theta}$ that on input $(\A, v, V)$ returns a value in the semiring $\SR$.
-Formally, the function $\sem{\theta}$ is recursively defined as follows:
-\begin{eqnarray*}
+Let $\R$ be a relational signature, $\A$ be a finite $\R$-structure with domain $A$, $v$ a first-order assignment for $\A$ and $V$ a second-order assignment for $\A$.
+Then the \emph{evaluation} of a $\qso$-formula $\alpha$ over $(\A, v, V)$ is defined as a function $\sem{\alpha}$ that on input $(\A, v, V)$ returns a non-negative integer in $\bbN$.
+Formally, the function $\sem{\alpha}$ is recursively defined as follows:
+$$
+\renewcommand{\arraystretch}{1.7}
+\begin{array}{rcl} 
 \sem{\varphi}(\A, v, V) & = & 
 \begin{cases}
-\one & \mbox{if } (\A, v, V) \models \varphi \\
-\zero & \mbox{otherwise}
+1 & \mbox{if } (\A, v, V) \models \varphi \\
+0 & \mbox{otherwise}
 \end{cases}\\
 \sem{s}(\A, v, V) & = & s \\
-\sem{\theta_1 \add \theta_2}(\A, v, V) & = & \sem{\theta_1}(\A, v, V) \adds \sem{\theta_2}(\A, v, V)\\
-\sem{\theta_1 \mult \theta_2}(\A, v, V) & = & \sem{\theta_1}(\A, v, V) \mults \sem{\theta_2}(\A, v, V)\\ 
-\sem{\sa{x} \theta}(\A, v, V) & = & \bigadds_{a \in A} \sem{\theta}(\A,v[a/x],V)\\
-\sem{\pa{x} \theta}(\A, v, V) & = & \bigmults_{a \in A} \sem{\theta}(\A,v[a/x],V)\\
-\sem{\sa{X} \theta}(\A, v, V) & = & \bigadds_{B \subseteq A^{\arity(X)}} \sem{\theta}(\A, v, V[B/X])\\
-\sem{\pa{X} \theta}(\A, v, V) & = & \bigmults_{B \subseteq A^{\arity(X)}} \sem{\theta}(\A, v, V[B/X])
+\sem{\alpha_1 \add \alpha_2}(\A, v, V) & = & \sem{\alpha_1}(\A, v, V) + \sem{\alpha_2}(\A, v, V)\\
+\sem{\alpha_1 \mult \alpha_2}(\A, v, V) & = & \sem{\alpha_1}(\A, v, V) \cdot \sem{\alpha_2}(\A, v, V)\\ 
+\sem{\sa{x} \alpha}(\A, v, V) & = & \displaystyle \sum_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
+\sem{\pa{x} \alpha}(\A, v, V) & = & \displaystyle \prod_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
+\sem{\sa{X} \alpha}(\A, v, V) & = & \displaystyle \sum_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])\\
+\sem{\pa{X} \alpha}(\A, v, V) & = & \displaystyle \prod_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])
+\end{array}
+$$
+A formula in $\qso$ can mention the usual quantifiers in $\so$ (that is, $\exists x$ and $\exists X$) and the quantifiers that make use of addition and multiplication (that is, $\Sigma x$, $\Pi x$, $\Sigma X$ and $\Pi X$), which are called {\em quantitative quantifiers} .
+A $\qso$-formula $\alpha$ is said to be a \emph{sentence} if it does not have any free variable, that is, every variable in $\alpha$ is under the scope of a usual quantifier or a quantitative quantifier.
+It is important to notice that if $\alpha$ is a $\qso$-sentence over a relational signature $\R$, then for every finite $\R$-structure $\A$, first-order assignments $v_1$, $v_2$ for $\A$ and second-order assignments $V_1$, $V_2$ for $\A$, it holds that:
+\begin{eqnarray*}
+\sem{\alpha}(\A, v_1, V_1) & = & \sem{\alpha}(\A, v_2, V_2).
 \end{eqnarray*}
-A formula in $\qso(\SR)$ can mention the usual quantifiers in $\so$ (that is, $\exists x$ and $\exists X$) and the quantifiers that make use of the two operations in the semiring (that is, $\Sigma x$, $\Pi x$, $\Sigma X$ and $\Pi X$), which are called {\em semiring quantifiers} .
-A $\qso(\SR)$-formula $\theta$ is said to be a \emph{sentence} if it does not have any free variable, that is, every variable in $\theta$ is under the scope of a usual quantifier or a semiring quantifier.
-It is important to notice that if $\theta$ is a $\qso(\SR)$-sentence over a relational signature $\R$, then for every finite $\R$-structure $\A$, first-order variable assignments $v_1$, $v_2$ for $\A$ and second-order variable assignments $V_1$, $V_2$ for $\A$, it holds that:
-\begin{eqnarray*}
-\sem{\theta}(\A, v_1, V_1) & = & \sem{\theta}(\A, v_2, V_2).
-\end{eqnarray*}
-Thus, in such a case we use the term $\sem{\theta}(\A)$ to denote $\sem{\theta}(\A, v, V)$, for some arbitrary first-order variable assignment $v$ for $\A$ and some arbitrary second-order variable assignment $V$ for $\A$.
+Thus, in such a case we use the term $\sem{\alpha}(\A)$ to denote $\sem{\alpha}(\A, v, V)$, for some arbitrary first-order assignment $v$ for $\A$ and some arbitrary second-order assignment $V$ for $\A$.
 
-In this paper, we consider several fragments of $\qso(\SR)$, which are obtained by restricting the syntax of the formula $\varphi$ in \eqref{eq-def-qso} or the use of the semiring quantifiers.
-Let $\eqso(\SR)$ be the fragment of $\qso(\SR)$ that is obtained by only allowing the semiring quantifiers $\Sigma x$, $\Pi x$ and $\Sigma X$ in \eqref{eq-def-qso}.
-Moreover, assuming that $\LL$ is a fragment of $\so$, let $\qso(\SR, \LL)$ be the fragment of $\qso(\SR)$ obtained by restricting $\varphi$ in \eqref{eq-def-qso} to be a formula in $\LL$, and let $\eqso(\SR, \LL)$ be the fragment of $\eqso(\SR)$ obtained by imposing the same restriction.
+In this paper, we consider several fragments of $\qso$, which are obtained by restricting the syntax of the formula $\varphi$ in \eqref{eq-def-qso} or the use of the quantitative quantifiers.
+Let $\eqso$ be the fragment of $\qso$ that is obtained by only allowing the quantitative quantifiers $\Sigma x$, $\Pi x$ and $\Sigma X$ in \eqref{eq-def-qso}.
+Moreover, assuming that $\LL$ is a fragment of $\so$, let $\qso(\LL)$ be the fragment of $\qso$ obtained by restricting $\varphi$ in \eqref{eq-def-qso} to be a formula in $\LL$, and let $\eqso(\LL)$ be the fragment of $\eqso$ obtained by imposing the same restriction.
 In particular, in this paper we consider the following fragments $\LL$: $\fo$, which is obtained by disallowing the use of the second-order quantifier ($\exists X$) in the formula $\varphi$ in \eqref{eq-def-qso}, and $\eso$, which is obtained by allowing $\varphi$ in \eqref{eq-def-qso} to be a formula of the form $\exists X_1 \cdots \exists X_k \, \psi$, where $\psi$ is an $\fo$-formula.

### 2016-08-03 QSO/preliminaries.tex
Nueva versión de QSO.

@@
 To define the semantics of $\so$, we need to introduce some terminology.
-Given a relational signature $\R$ and a finite $\R$-structure $\A$ with domain $A$, a first-order variable-assignment $v$ for $\A$ is a total function from $\fv$ to $A$, while a second-order variable-assignment $V$ for $\A$ is a total function with domain $\sv$ that maps each $X \in \sv$ to a subset of $A^{\arity(X)}$.
-Moreover, given a first-order variable-assignment $v$ for $\A$, $x \in \fv$ and $a \in A$, we denote by $v[a/x]$ a first-order variable-assignment such that $v[a/x](x) = a$ and $v[a/x](y) = v(y)$ for every $y \in \fv$ distinct from $x$.
-Similarly, given a second-order variable-assignment $V$ for $\A$, $X \in \sv$ and $B  \subseteq A^{\arity(X)}$, we denote by $V[B/X]$ a second-order variable-assignment such that $V[B/X](X) = B$ and $V[B/X](Y) = V(Y)$ for every $Y \in \sv$ distinct from $X$.
+Given a relational signature $\R$ and a finite $\R$-structure $\A$ with domain $A$, a first-order assignment $v$ for $\A$ is a total function from $\fv$ to $A$, while a second-order assignment $V$ for $\A$ is a total function with domain $\sv$ that maps each $X \in \sv$ to a subset of $A^{\arity(X)}$.
+Moreover, given a first-order assignment $v$ for $\A$, $x \in \fv$ and $a \in A$, we denote by $v[a/x]$ a first-order assignment such that $v[a/x](x) = a$ and $v[a/x](y) = v(y)$ for every $y \in \fv$ distinct from $x$.
+Similarly, given a second-order assignment $V$ for $\A$, $X \in \sv$ and $B  \subseteq A^{\arity(X)}$, we denote by $V[B/X]$ a second-order assignment such that $V[B/X](X) = B$ and $V[B/X](Y) = V(Y)$ for every $Y \in \sv$ distinct from $X$.
 
 Assume that $\varphi$ is an $\so$-formula over a signature $\R$.
-Then given a finite $\R$-structure $\A$ with domain $A$, a first-order variable-assignment $v$ for $\A$ and a second-order variable-assignment $V$ for $\A$, we say that $(\A, v, V)$ satisfies $\varphi$, denoted by $(\A, v, V) \models \varphi$, if: (1) $\varphi$ is the formula $R(x_1, \ldots, x_\ell)$ and $(v(x_1), \ldots, v(x_\ell)) \in R^\A$; (2) $\varphi$ is the formula $X(x_1, \ldots, x_m)$ and $(v(x_1), \ldots, v(x_m)) \in V(X)$; (3) $\varphi$ is the formula $\neg \psi$ and it not the case that $(\A, v, V) \models \psi$; (4) $\varphi$ is the formula $(\varphi_1 \vee \varphi_2)$, and $(\A, v, V) \models \varphi_1$ or $(\A, v, V) \models \varphi_2$; (5) $\varphi$ is the formula $\exists x \, \psi$ and there exists $a \in A$ such that $(\A, v[a/x], V) \models \psi$; or (6) $\varphi$ is the formula $\exists X \, \psi$ and there exists $B \subseteq A^{\arity(X)}$ such that $(\A, v, V[B/X]) \models \psi$.
+Then given a finite $\R$-structure $\A$ with domain $A$, a first-order assignment $v$ for $\A$ and a second-order assignment $V$ for $\A$, we say that $(\A, v, V)$ satisfies $\varphi$, denoted by $(\A, v, V) \models \varphi$, if: (1) $\varphi$ is the formula $R(x_1, \ldots, x_\ell)$ and $(v(x_1), \ldots, v(x_\ell)) \in R^\A$; (2) $\varphi$ is the formula $X(x_1, \ldots, x_m)$ and $(v(x_1), \ldots, v(x_m)) \in V(X)$; (3) $\varphi$ is the formula $\neg \psi$ and it not the case that $(\A, v, V) \models \psi$; (4) $\varphi$ is the formula $(\varphi_1 \vee \varphi_2)$, and $(\A, v, V) \models \varphi_1$ or $(\A, v, V) \models \varphi_2$; (5) $\varphi$ is the formula $\exists x \, \psi$ and there exists $a \in A$ such that $(\A, v[a/x], V) \models \psi$; or (6) $\varphi$ is the formula $\exists X \, \psi$ and there exists $B \subseteq A^{\arity(X)}$ such that $(\A, v, V[B/X]) \models \psi$.
 As usual, we consider the propositional operators $\wedge$, $\rightarrow$, and $\leftrightarrow$ that can be obtained from $\vee$ and $\neg$.
@@
 %Finally, we consider standard abbreviations of formulas that can be defined in $\so$-logic (actually, $\fo$-logic) like $\first(x) := \fa{x} y \leq x$ and $\last(x) := \fa{x} x \leq y$ to denote the first and last element of the linear order $\leq$, respectively, and $\succesor(x,y) := x \leq y \wedge y \not\leq x \wedge \fa{z} ( z \leq x \vee y \leq z)$ to denote the successor relation.
-
-\subsection{Semirings}
-
-A semiring is a structure $\SR = (S, \adds, \mults, \zero, \one)$ where $(S, \adds, \zero)$ is a commutative monoid, $(S, \mults, \one)$ is a monoid, multiplication ($\mults$) distributes over addition ($\adds$), and $\zero \mults s = s \mults \zero = \zero$ for every $s \in S$.
-If the multiplication is commutative, we say that $\SR$ is commutative.
-In this paper, we always assume that $\SR$ is  commutative.
-For the sake of simplification, we usually denote the set of elements $S$ with the name of the semiring $\SR$.
-% and we say sentences like ``for all $s \in \SR$''.
-
-In general, our definitions are given with respect to a generic semiring.
-However, in some cases we consider some particular semiring like, for example, the semiring of \emph{natural numbers} $\nat = (\mathbb{N}, +, \cdot, 0, 1)$, the semiring of \emph{integers} $\integ = (\mathbb{Z}, +, \cdot, 0, 1)$, the \emph{boolean} semiring $\bln = (\{\false, \true\}, \vee, \wedge, \false, \true)$, or the \emph{tropical} semiring $\trop = (\nat \cup \{\infty\}, \min, +, \infty, 0)$.
-For the last semiring, the $\min$-operator is defined by the linear order $0 < 1 < 2 < \cdots < \infty$ and the $+$-operator is just the addition over natural number with the exception that $v + \infty = \infty + v = \infty$ for all $v \in \nat \cup \{\infty\}$.
 
@@
 \end{eqnarray*}
-Given a semiring $\SR = (S, \adds, \mults, \zero, \one)$ and a relational signature $\R$, a function $f$ is said to be from $\R$ to $\SR$ if $f$ is a total function from $\{\enc(\A) \mid \A \in \str[\R]\}$ to $S$.
-Moreover, a set $\CC$ of functions is said to be a {\em function complexity class over $\SR$} if every $f \in \CC$ is a function from some relational signature $\R$ to $\SR$.
+Given a set $\bbD$ and a relational signature $\R$, a function $f$ is said to be from $\R$ to $\bbD$ if $f$ is a total function from $\{\enc(\A) \mid \A \in \str[\R]\}$ to $\bbD$.
+Moreover, a set $\CC$ of functions over $\bbD$ is said to be a {\em function complexity class over $\bbD$} if every $f \in \CC$ is a function from some relational signature $\R$ to $\bbD$.
 

### 2016-08-03 QSO/preliminaries.tex
Cambios y presentación.

@@
 
-If $\KK$ is a class of finite structures and $f$ is a function from a relational signature $\R$ to a semiring $\SR$, we denote by $\res{f}{\KK}$ the restriction of $f$ to $\KK$, that is, a function such that: the domain of $\res{f}{\KK}$ is $\{ \enc(\A) \mid \A \in \str[\R] \cap \KK\}$, and
+If $\KK$ is a class of finite structures and $f$ is a function from a relational signature $\R$ to $\bbD$, we denote by $\res{f}{\KK}$ the restriction of $f$ to $\KK$, that is, a function such that: the domain of $\res{f}{\KK}$ is $\{ \enc(\A) \mid \A \in \str[\R] \cap \KK\}$, and
 $\res{f}{\KK}(\enc(\A)) = f(\enc(\A))$ for every $\A \in \str[\R] \cap \KK$.

### 2016-08-17 QSO/logics.tex
17 ago

@@
 In this paper, we consider several fragments of $\qso$, which are obtained by restricting the syntax of the formula $\varphi$ in \eqref{eq-def-qso} or the use of the quantitative quantifiers.
-Let $\eqso$ be the fragment of $\qso$ that is obtained by only allowing the quantitative quantifiers $\Sigma x$, $\Pi x$ and $\Sigma X$ in \eqref{eq-def-qso}.
+Let $\qfo$ be the fragment of $\qso$ obtained by only allowing the quantitative quantifiers $\Sigma x$, $\Pi x$ in \eqref{eq-def-qso}.
+Let $\eqso$ be a fragment of $\qso$ defined as $\qfo$ but also allowing the quantitative quantifier $\Sigma X$.
 Moreover, assuming that $\LL$ is a fragment of $\so$, let $\qso(\LL)$ be the fragment of $\qso$ obtained by restricting $\varphi$ in \eqref{eq-def-qso} to be a formula in $\LL$, and let $\eqso(\LL)$ be the fragment of $\eqso$ obtained by imposing the same restriction.
-In particular, in this paper we consider the following fragments $\LL$: $\fo$, which is obtained by disallowing the use of the second-order quantifier ($\exists X$) in the formula $\varphi$ in \eqref{eq-def-qso}, and $\eso$, which is obtained by allowing $\varphi$ in \eqref{eq-def-qso} to be a formula of the form $\exists X_1 \cdots \exists X_k \, \psi$, where $\psi$ is an $\fo$-formula.
+In particular, in this paper we consider the following fragments $\LL$: $\fo$, which is obtained by disallowing the use of the second-order quantifier ($\exists X$ or $\forall X$) in the formula $\varphi$ in \eqref{eq-def-qso}, and $\eso$, which is obtained by allowing $\varphi$ in \eqref{eq-def-qso} to be a formula of the form $\exists X_1 \cdots \exists X_k \, \psi$, where $\psi$ is an $\fo$-formula.
+
+\begin{theorem}
+	Let $\alpha$ be a $\qso$-formula in some fragment of $\qso$.
+There is a $\qso$ formula $\alpha'$ in the same fragment such that $\alpha'$ does not have any subformula of the form $(\beta_1 \cdot \beta_2)$ and $\sem{\alpha} = \sem{\alpha'}$.
+\end{theorem}
+\begin{proof}
+	Let $\{+, \cdot, \Sigma, \Pi\}$ be the set of {\it algebraic operators} in $\qso$.
+We will prove by induction over the number of algebraic operators in a $\qso$-formula $\alpha$ that there exists some $\qso$-formula $\alpha'$ that respects the stated condition.
+Let $\text{op}(\alpha)$ be the number of algebraic operators in $\alpha$.
+The case $\text{op}(\alpha) = 0$ is trivial.
+Let $n\in\nat$ be such that for each $\qso$-formula $\beta$ such that $\text{op}(\beta) < n$ there is some $\beta'$ that respects the condition.
+Let $\alpha$ be a $\qso$-formula such that $\text{op}(\alpha) = n$.
+Note that if $\alpha$ is of the form $\alpha = (\beta_1 + \beta_2)$, $\alpha = \Sigma x \beta$, $\alpha = \Pi x \beta$, $\alpha = \Sigma X \beta$ or $\alpha = \Pi X \beta$ the proof follows directly.
+Therefore we assume that $\alpha = (\beta_1 \cdot \beta_2)$.
+We separate the proof in two main cases:
+	\begin{enumerate}
+		\item At least one of $\beta_1, \beta_2$ is not of the form $\Pi x \gamma$, $\Pi X \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. Without loss of generality, let $\beta_1$ be such one.
+		\begin{enumerate}
+			\item $\beta_1 = s \in \nat$.
+Then we use $\alpha' = \beta_2 + \cdots + \beta_2$, where $\beta_2$ is repeated $s$ times, and if $s = 0$, take $\alpha' = \perp$.
+Since each of the formulas paired by $+$ has less than $n$ algebraic operators, the statement holds.
+			\item $\beta_1 = (\gamma_1 + \gamma_2)$.
+Then, we use $\alpha' = (\gamma_1\cdot\beta_2 + \gamma_2\cdot\beta_2)$ and each of the formulas paired by $+$ have less than $n$ algebraic operators so the statement holds.
+			\item $\beta_1 = \Sigma x \gamma$. Then, we use $\alpha' = \Sigma x (\gamma\cdot\beta_2)$ and as in the previous case the statement holds.
+			\item $\beta_1 = \Sigma X \gamma$, which we treat exactly as the previous one.
+		\end{enumerate}
+		\item Both of $\beta_1$ and $\beta_2$ are of the form $\Pi x \gamma$, $\Pi X \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. This case also separates in further cases:
+		\begin{enumerate}
+			\item $\beta_1 = \Sigma x \gamma_1(x)$ and $\beta_2 = \Sigma y \gamma_2(y)$. Then we use $\alpha' = \Sigma x(\gamma_1(x)\cdot\gamma_2(x))$, and the statement holds.
+			\item $\beta_1 = \Sigma X \gamma_1(X)$ and $\beta_2 = \Sigma Y \gamma_2(Y)$. [something about arities].
+			\item $\beta_1 = \Sigma X \gamma_1(X)$ and $\beta_2 = \Sigma x \gamma_2(x)$ (w.l.o.g.). [use singleton predicates].
+			\item $\beta_1 = \Sigma x \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). Then, we use $\alpha' = \Sigma x (\gamma \cdot \varphi)$, and the statement holds.
+			\item $\beta_1 = \Sigma X \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). We treat this case exactly as the previous one.
+			\item $\beta_1 = \varphi_1$ and $\beta_2 = \varphi_2$ both $\so$-formulas. Then, we use $\alpha' = (\varphi_1 \wedge \varphi_2)$ and the statement holds.
+		\end{enumerate}
+	\end{enumerate}
+\end{proof}

### 2016-08-17 QSO/preliminaries.tex
17 ago

@@
 
-A relational signature $\R$ is a finite set $\{R_1, \ldots, R_k\}$, where each $R_i$ ($1 \leq i \leq k$) is a relation name with an associate arity greater than 0, which is denoted by $\arity(R_i)$.
+A $\mathbb{N}$ relational signature $\R$ is a finite set $\{R_1, \ldots, R_k\}$, where each $R_i$ ($1 \leq i \leq k$) is a relation name with an associate arity greater than 0, which is denoted by $\arity(R_i)$.
 A finite structure over $\R$ (or just finite $\R$-structure) is a tuple $\A = \langle A, R_1^\A, \ldots, R_k^\A \rangle$ such that $A$ is a finite set and $R_i^\A \subseteq A^{\arity(R_i)}$ for every $i \in \{1, \ldots, k\}$.

### 2016-08-24 QSO/cap-comp.tex
23 ago

@@
 
+\begin{theorem} \label{captfp}
+	$\qfo(\lfp)$ captures $\fp$ over the class of ordered structures.
+\end{theorem}
+\begin{proof}
+	For the first condition, let $\alpha\in\qfo(\lfp)$, $\A\in\ostr$, $v$ and $V$ a first and second order assignment for $\A$, respectively.
+To evaluate $\sem{\alpha}(\A,v,V)$, we replace each first order sum and first order product by their corresponding expansion.
+This is, $\Sigma x \beta(x)$ is replaced by $(\beta(a_1)+\cdots+\beta(a_n))$, where $A = \{a_1,\ldots,a_n\}$, and $\Pi x \beta(x)$ is replaced by $(\beta(a_1)\cdot\,\cdots\,\cdot\beta(a_n))$.
+Then we replace each logic sub-formula $\varphi$ in $\alpha$ by their evaluated value, 0 or 1.
+Since each of this formulas is in $\lfp$, this can be done in polynomial time.
+The resulting formula is an arithmetic expression, which can be evaluated recursively in polynomial time.
+
+	For the second condition, let $f\in \fp$.
+Let $k\in\nat$ be such that $\vert f(\A) \vert = \vert A \vert^k$ for each $\A\in\ostr[\R]$.
+Let $\Phi(x_0,\ldots,x_{k-1})$ be a $\lfp$ formula such that $(\A,a_0,\ldots,a_{k-1})\models\Phi$ if and only if in the string $y = f(\A)$ there is 1 in the $(\vert A \vert^{k-1}m_{k-1} + \cdots + \vert A \vert^2 m_2 + \vert A \vert m_1 + m_0)$-th position from right to left, and where $a_i$ is the (0-indexed) $m_i$-th element of $A$.
+We use
+	$$
+	\alpha = \Sigma x_0 \cdots \Sigma x_{k-1} \Phi(x_0,\ldots,x_{k-1})\cdot\varphi_{k-1}(x_{k-1})\cdot\,\cdots\,\cdot\varphi_0(x_0).
+	$$
+	where $\varphi_i(x) = \Pi y[(y < x)\mapsto\Pi z_1\cdots\Pi z_i\,2]$ for $i > 0$ and $\varphi_0(x) = \Pi y[(y < x)\mapsto 2]$.
+Note that if $a$ is the  $m$-th element in $A$, then $\sem{\varphi_i(x)}(\A,a) = 2^{\vert A \vert^i m}$.
+Therefore, for each $(a_0,\ldots,a_{k-1})\in A^k$, we have that $\sem{\Phi(x_0,\ldots,x_{k-1})\cdot\varphi_{k-1}(x_{k-1})\,\cdots\,\varphi_0(x_0)}(\A,a_0,\ldots,a_{k-1}) = 2^{(\vert A \vert^{k-1}m_{k-1} + \cdots + \vert A \vert^2 m_2 + \vert A \vert m_1 + m_0)}$ if $(\A,a_0,\ldots,a_{k-1})\models\Phi$ and 0 otherwise, and adding these values gives $f(\A)$.
+\end{proof}
+
+\begin{theorem}
+	$\qfo(\pfp)$ captures $\nfpspace$ over the class of ordered structures.
+\end{theorem}
+\begin{proof}
+	For the first condition, we use the exact same procedure as in Theorem \ref{captfp}, noting that evaluating each of the $\pfp$ formulas can be done in polynomial space, and that the result is also polynomial on the size of the input.
+
+	For the second condition, note that for each function $f\in \nfpspace$, $\vert f(\A) \vert$ is polynomial on $\vert A \vert$.
+Thus, there is a $\pfp$ formula $\Phi$ that models the string $f(\A)$ by the bit.
+The rest of the proof is analogous to Theorem \ref{captfp}.
+\end{proof}
+
+\begin{theorem}
+	$\qfo(\tc)$ captures $\shl$ over the class of ordered structures.
+\end{theorem}
+\begin{proof}
+	For the first condition, let $\alpha\in\qfo(\tc)$, let $\A\in\ostr$ $v$ and $V$ first and second order assignments of $\A$ respectively.
+We evaluate $\sem{alpha}(\A,v,V)$ recursively.
+A logic formula $\varphi$ is evaluated...?
+
+	For the second condition, note that for each function $f\in \shl$, $\vert f(\A) \vert$ is polynomial on $\vert A \vert$.
+Thus, there is a $\tc$ formula $\Phi$ that models the string $f(\A)$ by the bit.
+The rest of the proof is analogous to Theorem \ref{captfp}.
+\end{proof}
+
 \begin{theorem}

### 2016-09-07 QSO/logics.tex
23 ago
7 sep

@@
 		\end{enumerate}
-		\item Both of $\beta_1$ and $\beta_2$ are of the form $\Pi x \gamma$, $\Pi X \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. This case also separates in further cases:
+		\item Both of $\beta_1$ and $\beta_2$ are of the form $\Pi x \gamma$, $\Pi X \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula.
+This case also separates in further cases.
+We define $(\varphi \mapsto \alpha) := (\neg\varphi + (\varphi\cdot\alpha))$ for which $\sem{(\varphi \mapsto \alpha)}(\A,v,V)$ equals $\sem{\alpha}(\A,v,V)$ when $(\A,v,V)\models\varphi$ and 1 otherwise.
 		\begin{enumerate}
@@
 \end{proof}
+
+We define an operator which extends the Least Fixed Point logic to counting.
+Recall that a fixed point operator is defined by a positive formula $\varphi(x_1,\ldots,x_k,R)$ where $R$ is a predicate or arity $k$.
+The operator $T_{\varphi}:2^{A^k} \to 2^{A^k}$ is defined as $T_{\varphi}(X) = \{(a_1,\ldots,a_k)\mid (\A,X)\models \varphi(a_1,\ldots,a_k,R) \}$, for each $X\subseteq A^k$ and $X^*\subseteq A^k$ is a fixed point of $T_{\varphi}$ if $T_{\varphi}(X) = X$.
+Then, the formula $[\textbf{lfp}_{(x_1,\ldots,x_k),R}\varphi(x_1,\ldots,x_k,R)](y_1,\ldots,y_k)$ is evaluated at the least fixed point of $T_{\varphi}$.
+
+We introduce the operator $[\textbf{alfp}_{(x_1,\ldots,x_k),R}\varphi(x_1,\ldots,x_k,R)\mid\alpha(u_1,\ldots,u_k,R,p)](y_1,\ldots,y_k)$ which uses a FO-formula $\varphi$ as well as a QSO-formula $\alpha$ and takes as input a placeholder variable $p$, which is interpreted as a constant.
+The function $\zeta_{\varphi,\alpha}:2^{A^k} \to \nat$ is defined as $\zeta_{\varphi,\alpha}(T_{\varphi}(X)) = \sem{\alpha'}(\A,R)$ where $\alpha'$ is obtained by replacing every instance of $p$ by $\zeta_{\varphi,\alpha}(X)$.

### 2016-09-07 QSO/logics.tex
9 sep

@@
 We introduce the operator $[\textbf{alfp}_{(x_1,\ldots,x_k),R}\varphi(x_1,\ldots,x_k,R)\mid\alpha(u_1,\ldots,u_k,R,p)](y_1,\ldots,y_k)$ which uses a FO-formula $\varphi$ as well as a QSO-formula $\alpha$ and takes as input a placeholder variable $p$, which is interpreted as a constant.
-The function $\zeta_{\varphi,\alpha}:2^{A^k} \to \nat$ is defined as $\zeta_{\varphi,\alpha}(T_{\varphi}(X)) = \sem{\alpha'}(\A,R)$ where $\alpha'$ is obtained by replacing every instance of $p$ by $\zeta_{\varphi,\alpha}(X)$.
+We define $T^{-1}_{\varphi}(X) = Y$ as the least set $Y$ for which $T_{\varphi}(Y) = X$.
+The function $\zeta_{\varphi,\alpha}:2^{A^k} \to \nat$ is defined as $\zeta_{\varphi,\alpha}(X) = \sem{\alpha'}(\A,R)$ where $\alpha'$ is obtained by replacing every instance of $p$ by $\zeta_{\varphi,\alpha}(T^{-1}_{\varphi}(X))$, and $\zeta_{\varphi,\alpha}(\emptyset) = 0$.
+Then the operator is evaluated as $$\sem{[\textbf{alfp}_{(x_1,\ldots,x_k),R}\varphi(x_1,\ldots,x_k,R)\mid\alpha(u_1,\ldots,u_k,R,p)](y_1,\ldots,y_k)}(y_1,\ldots,y_k) = \zeta_{\varphi,\alpha}(X^*)$$, where $X^*$ is the least fixed point of $T_{\varphi}$.

### 2016-09-14 QSO/logics.tex
14 sep

--- 2016-09-07 QSO/logics.tex
@@
 
-We introduce the operator $[\textbf{alfp}_{(x_1,\ldots,x_k),R}\varphi(x_1,\ldots,x_k,R)\mid\alpha(u_1,\ldots,u_k,R,p)](y_1,\ldots,y_k)$ which uses a FO-formula $\varphi$ as well as a QSO-formula $\alpha$ and takes as input a placeholder variable $p$, which is interpreted as a constant.
-The function $\zeta_{\varphi,\alpha}:2^{A^k} \to \nat$ is defined as $\zeta_{\varphi,\alpha}(T_{\varphi}(X)) = \sem{\alpha'}(\A,R)$ where $\alpha'$ is obtained by replacing every instance of $p$ by $\zeta_{\varphi,\alpha}(X)$.
+We introduce the operator $[\textbf{alfp}_{\varphi(x_1,\ldots,x_k,R)}\alpha(u_1,\ldots,u_k,R,\pi)](y_1,\ldots,y_k)$ which uses a FO-formula $\varphi$ as well as a QSO-formula $\alpha$ and takes as input a placeholder function $\pi$, which is interpreted as a constant.
+The partial function $\zeta_{\varphi,\alpha}:2^{A^k} \to \nat$ is such that  $\zeta_{\varphi,\alpha}(T_{\varphi}(X)) = \sem{\alpha'}(\A,R)$ where $\alpha'$ is obtained by replacing every instance of $\pi(\bar{z})$ by $\zeta_{\varphi,\alpha}(X)$, and $\zeta_{\varphi,\alpha}(\emptyset) = 0$.
+Then the operator is evaluated as $$\sem{[\textbf{alfp}_{\varphi(x_1,\ldots,x_k,R)}\alpha(u_1,\ldots,u_k,R,\pi)](y_1,\ldots,y_k)}(y_1,\ldots,y_k) = \zeta_{\varphi,\alpha}(X^*)$$, where $X^*$ is the least fixed point of $T_{\varphi}$.
+
+As an example, we will show a formula that counts the number of paths of size $n$ of a structure $\A$ with a binary relation $E$. First we define $\varphi(x,R)$:
+$$
+\varphi(x,R) = \forall y(x < y \vee x = y) \vee \exists z(R(z) \wedge \varphi_{succ}(z,x)),x
+$$
+where $\varphi_{succ}(x,y)$ is a formula that is satisfied by pairs $(x,y)$ that are consecutive in the order $<$.
+That is, $\varphi_{succ}(x,y) = x < y \wedge \forall z((x < z \wedge z < y) \to (x = z \vee z = y) )$.
+Now we define $\alpha(R)$ as:
+$$
+\alpha(x,y,R) = \neg \exists z(R(z)) + \Sigma z[E(x,z)\cdot \pi(z,y)].
+$$
+Then, the formula $[\alfp_{\varphi(x,R)} \alpha(x,y,R)](x,y)$ when evaluated on $(\A,a,b)$ will count the number of paths of size $n$ from $a$ to $b$.
+
+We also define the less powerful operator $\pth$ as follows.
+Given a positive formula $\varphi(x,R)$ and a formula $\psi(x,y)$ we define $[\pth_{\varphi(x,R)} \psi(x,y)](x,y)$ as a formula that when evaluated on $(\A, a, b)$ induces a graph over $\A$ with the binary relation determined by $\psi(x,y)$ and counts the number of paths from $a$ to $b$ of a size determined by the $\varphi(x,R)$.
+Formally, we define:
+$$
+\sem{[\pth_{\varphi(x,R)} \psi(x,y)](x,y)} = \sem{[\alfp_{\varphi(x,R)} \alpha_{\psi}(x,y,R)](x,y)},
+$$
+where $\alpha_{\psi}(x,y,R) = \neg \exists z(R(z)) + \Sigma z[\psi(x,z)\cdot \pi(z,y)]$.
+
+\begin{theorem}
+	Given a positive $\fo$ formula $\varphi(\bar{x},R)$ and a $\qfo$ formula $\alpha(\bar{x})$, there exists a $\qso$ formula $\beta(\bar{x})$ such that $\sem{[\alfp_{\varphi(\bar{x},R)} \alpha_{\psi}(\bar{x},R)](\bar{x})} = \sem{\beta(\bar{x})}$.
+\end{theorem}
+\begin{proof}
+
+\end{proof}

### 2016-09-16 QSO/logics.tex
15 sep

@@
 We define an operator which extends the Least Fixed Point logic to counting.
-Recall that a fixed point operator is defined by a positive formula $\varphi(x_1,\ldots,x_k,R)$ where $R$ is a predicate or arity $k$.
-The operator $T_{\varphi}:2^{A^k} \to 2^{A^k}$ is defined as $T_{\varphi}(X) = \{(a_1,\ldots,a_k)\mid (\A,X)\models \varphi(a_1,\ldots,a_k,R) \}$, for each $X\subseteq A^k$ and $X^*\subseteq A^k$ is a fixed point of $T_{\varphi}$ if $T_{\varphi}(X) = X$.
-Then, the formula $[\textbf{lfp}_{(x_1,\ldots,x_k),R}\varphi(x_1,\ldots,x_k,R)](y_1,\ldots,y_k)$ is evaluated at the least fixed point of $T_{\varphi}$.
+Recall that a fixed point operator is defined by a formula $\varphi(x_1,\ldots,x_k,R)$ which is positive on $R$ where $R$ is a predicate or arity $k$.
+The operator $T_{\varphi}:2^{A^k} \to 2^{A^k}$ is defined as $T_{\varphi}(X) = \{(a_1,\ldots,a_k)\mid (\A,X)\models \varphi(a_1,\ldots,a_k,R) \}$, for each $X\subseteq A^k$ and $X^*\subseteq A^k$ is a fixed point of $T_{\varphi}$ if $T_{\varphi}(X^*) = X^*$.
+Then, the formula $[\textbf{lfp}\,\varphi(x_1,\ldots,x_k,R)](y_1,\ldots,y_k)$ is evaluated at the least fixed point of $T_{\varphi}$.
 
-We introduce the operator $[\textbf{alfp}_{\varphi(x_1,\ldots,x_k,R)}\alpha(u_1,\ldots,u_k,R,\pi)](y_1,\ldots,y_k)$ which uses a FO-formula $\varphi$ as well as a QSO-formula $\alpha$ and takes as input a placeholder function $\pi$, which is interpreted as a constant.
-The partial function $\zeta_{\varphi,\alpha}:2^{A^k} \to \nat$ is such that  $\zeta_{\varphi,\alpha}(T_{\varphi}(X)) = \sem{\alpha'}(\A,R)$ where $\alpha'$ is obtained by replacing every instance of $\pi(\bar{z})$ by $\zeta_{\varphi,\alpha}(X)$, and $\zeta_{\varphi,\alpha}(\emptyset) = 0$.
-Then the operator is evaluated as $$\sem{[\textbf{alfp}_{\varphi(x_1,\ldots,x_k,R)}\alpha(u_1,\ldots,u_k,R,\pi)](y_1,\ldots,y_k)}(y_1,\ldots,y_k) = \zeta_{\varphi,\alpha}(X^*)$$, where $X^*$ is the least fixed point of $T_{\varphi}$.
+We extend QSO with the operator $[\textbf{alfp}_{\varphi(y_1,\ldots,y_k,R)}\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$.
+It is defined by a FO-formula $\varphi$ and a QFO-formula $\alpha$ which mentions a placeholder function $\pi$ with domain $\fv^{\ell}$.
+
+To formally characterize its semantics we define the partial function $\zeta_{\varphi,\alpha}:2^{A^k} \to (A^{\ell}\times\nat)$ for each $\varphi$ and $\alpha$ as follows.
+Let $\A$ be a structure.
+For $R = \emptyset$, it holds that $\zeta_{\varphi,\alpha}[\emptyset] = \sem{\alpha\mid_{\pi(\bar{u})\to 0}}$.
+For each $R \subseteq A^k$ such that $T_{\varphi}(R) \neq R$, let $\beta$ be a virtual $\qfo$ formula such that $\sem{\beta}(\A,v) = \zeta_{\varphi,\alpha}[R](v)$ for each first order assignment $v$, and it holds that $\zeta_{\varphi,\alpha}[T_{\varphi}(R)] = \sem{\alpha\mid_{\pi\to\beta}}$.
+
+For a given first order asignment $v$, the operator is evaluated as:
+$$
+[\textbf{alfp}_{\varphi(y_1,\ldots,y_k,R)}\alpha(x_1,\ldots,x_{\ell},R,\pi)](\A,v) = \zeta_{\varphi,\alpha}[X^*](v),
+$$
+where $X^*$ is the least fixed point of $\varphi$.
 

### 2016-09-16 QSO/logics.tex
15 sep

@@
 Let $\A$ be a structure.
-For $R = \emptyset$, it holds that $\zeta_{\varphi,\alpha}[\emptyset] = \sem{\alpha\mid_{\pi(\bar{u})\to 0}}$.
-For each $R \subseteq A^k$ such that $T_{\varphi}(R) \neq R$, let $\beta$ be a virtual $\qfo$ formula such that $\sem{\beta}(\A,v) = \zeta_{\varphi,\alpha}[R](v)$ for each first order assignment $v$, and it holds that $\zeta_{\varphi,\alpha}[T_{\varphi}(R)] = \sem{\alpha\mid_{\pi\to\beta}}$.
+For $R = \emptyset$, we define $\zeta_{\varphi,\alpha}[\emptyset] = \sem{\alpha\mid_{\pi(\bar{u})\to 0}}$.
+For each $R \subseteq A^k$ such that $T_{\varphi}(R) \neq R$, let $\beta$ be a virtual $\qfo$ formula with $\ell$ open variables such that $\sem{\beta}(\A,v) = \zeta_{\varphi,\alpha}[R](v)$ for each first order assignment $v$, and we define $\zeta_{\varphi,\alpha}[T_{\varphi}(R)] = \sem{\alpha\mid_{\pi(\bar{u})\to\beta(\bar{u})}}$.
 

### 2016-09-16 QSO/logics.tex
15 sep

@@
 
-We extend QSO with the operator $[\textbf{alfp}_{\varphi(y_1,\ldots,y_k,R)}\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$.
+We extend $\qfo$ with the operator $[\textbf{alfp}_{\varphi(y_1,\ldots,y_k,R)}\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$.
 It is defined by a FO-formula $\varphi$ and a QFO-formula $\alpha$ which mentions a placeholder function $\pi$ with domain $\fv^{\ell}$.
@@
 That is, $\varphi_{succ}(x,y) = x < y \wedge \forall z((x < z \wedge z < y) \to (x = z \vee z = y) )$.
-Now we define $\alpha(R)$ as:
+Now we define $\alpha(x,y,R,\pi)$ as:
 $$
-\alpha(x,y,R) = \neg \exists z(R(z)) + \Sigma z[E(x,z)\cdot \pi(z,y)].
+\alpha(x,y,R,\pi) = \neg \exists z(R(z)) + \Sigma z[E(x,z)\cdot \pi(z,y)].
 $$
-Then, the formula $[\alfp_{\varphi(x,R)} \alpha(x,y,R)](x,y)$ when evaluated on $(\A,a,b)$ will count the number of paths of size $n$ from $a$ to $b$.
+Then, the formula $[\alfp_{\varphi(x,R)} \alpha(x,y,R,\pi)](x,y)$ when evaluated on $(\A,a,b)$ will count the number of paths of size $n$ from $a$ to $b$.
 
@@
 $$
-\sem{[\pth_{\varphi(x,R)} \psi(x,y)](x,y)} = \sem{[\alfp_{\varphi(x,R)} \alpha_{\psi}(x,y,R)](x,y)},
+\sem{[\pth_{\varphi(x,R)} \psi(x,y)](x,y)} = \sem{[\alfp_{\varphi(x,R)} \alpha_{\psi}(x,y,R,\pi)](x,y)},
 $$
-where $\alpha_{\psi}(x,y,R) = \neg \exists z(R(z)) + \Sigma z[\psi(x,z)\cdot \pi(z,y)]$.
+where $\alpha_{\psi}(x,y,R,\pi) = \neg \exists z(R(z)) + \Sigma z[\psi(x,z)\cdot \pi(z,y)]$.
 

### 2016-09-20 QSO/logics.tex
21 sep

@@
 We define an operator which extends the Least Fixed Point logic to counting.
-Recall that a fixed point operator is defined by a formula $\varphi(x_1,\ldots,x_k,R)$ which is positive on $R$ where $R$ is a predicate or arity $k$.
-The operator $T_{\varphi}:2^{A^k} \to 2^{A^k}$ is defined as $T_{\varphi}(X) = \{(a_1,\ldots,a_k)\mid (\A,X)\models \varphi(a_1,\ldots,a_k,R) \}$, for each $X\subseteq A^k$ and $X^*\subseteq A^k$ is a fixed point of $T_{\varphi}$ if $T_{\varphi}(X^*) = X^*$.
-Then, the formula $[\textbf{lfp}\,\varphi(x_1,\ldots,x_k,R)](y_1,\ldots,y_k)$ is evaluated at the least fixed point of $T_{\varphi}$.
+Recall that a fixed point operator is defined by a formula $\varphi(x_1,\ldots,x_k,R)$ which is positive on $R$ where $R$ is a predicate of arity $k$.
+For a structure $\A$ with domain $A$, the operator $T_{\varphi}:2^{A^k} \to 2^{A^k}$ is defined as $T_{\varphi}(X) = \{(a_1,\ldots,a_k)\mid (\A,X)\models \varphi(a_1,\ldots,a_k,R) \}$, for each $X\subseteq A^k$.
+Let $T_0 = \emptyset$ and $T_{i+1} = T_{\varphi}(T_i)$ for each $i \in \nat$.
+Note that there exists $n\in \nat$ such that $T_{n+1} = T_n$.
+Semantically, $[\lfpop\,\varphi(x_1,\ldots,x_k,R)]$ is defined such that for each $(a_1,\ldots,a_k)\in A^k$, it holds that $\A\models[\lfpop\,\varphi(x_1,\ldots,x_k,R)](a_1,\ldots,a_k)$ if and only if $\A\models T_n(a_1,\ldots,a_k)$.
 
-We extend $\qfo$ with the operator $[\textbf{alfp}_{\varphi(y_1,\ldots,y_k,R)}\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$.
+We extend $\qfo$ with the operator $[\textbf{alfp}\,\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$.
 It is defined by a FO-formula $\varphi$ and a QFO-formula $\alpha$ which mentions a placeholder function $\pi$ with domain $\fv^{\ell}$.
+Let $\{T_i\}_{i\in\nat}$ and $n\in\nat$ be defined from $\varphi$ as it was for the $\lfp$ operator.
 
-To formally characterize its semantics we define the partial function $\zeta_{\varphi,\alpha}:2^{A^k} \to (A^{\ell}\times\nat)$ for each $\varphi$ and $\alpha$ as follows.
-Let $\A$ be a structure.
-For $R = \emptyset$, we define $\zeta_{\varphi,\alpha}[\emptyset] = \sem{\alpha\mid_{\pi(\bar{u})\to 0}}$.
-For each $R \subseteq A^k$ such that $T_{\varphi}(R) \neq R$, let $\beta$ be a virtual $\qfo$ formula with $\ell$ open variables such that $\sem{\beta}(\A,v) = \zeta_{\varphi,\alpha}[R](v)$ for each first order assignment $v$, and we define $\zeta_{\varphi,\alpha}[T_{\varphi}(R)] = \sem{\alpha\mid_{\pi(\bar{u})\to\beta(\bar{u})}}$.
+For each $\qfo$ formulas $\alpha, \beta, \gamma$, where $\beta$ and $\gamma$ have $\ell$ and $m$ first-order open variables respectively, $u_1,\ldots,u_{\ell} \in \fv$, and $v_1,\ldots,v_m \in \{u_1,\ldots,u_{\ell}\}$, we define $\alpha\mid_{\beta(u_1,\ldots,u_{\ell})\to\gamma(v_1,\ldots,v_{m})}$ as $\alpha$ where every instance of the subformula $\beta(u_1,\ldots,u_{\ell})$ is replaced by $\gamma(v_1,\ldots,v_{m})$.
+Moreover, for each $a\in A$, suppose that $a$ is the $p$-th element in the order $<^{\A}$.
+Then if $a$ is the first element, we define $\varphi_a(x) = \forall y(x < y \vee x = y)$, and if it is not we define:
+$$
+\varphi_a(x) = \exists x_1 \cdots \exists x_{p-1}[\bigwedge_{1\leq i,j < p}x_i\neq x_j \wedge \bigwedge_{i = 1}^p x_i < x \wedge \forall y(y < x \to \bigvee_{i = 1}^p y = x_i)].
+$$
 
-For a given first order asignment $v$, the operator is evaluated as:
+To formally characterize the semantics of $[\textbf{alfp}\,\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$. %. 
+We name $\T = \{T_0,\ldots,T_n\}$ and we define the operator $Z:\T \to \{f\mid f:A^{\ell}\to\nat\}$ as follows.
+For each $T_i \in \T$, we define a $\qfo$ formula $\beta_i$ as follows.
+If $i = 0$, then $\beta_i(u_1,\ldots,u_{\ell}) = 0$.
+If $i \geq 1$, then let $f = Z[T_{i-1}]$ and
 $$
-[\textbf{alfp}_{\varphi(y_1,\ldots,y_k,R)}\alpha(x_1,\ldots,x_{\ell},R,\pi)](\A,v) = \zeta_{\varphi,\alpha}[X^*](v),
+\beta_i(u_1,\ldots,u_{\ell}) = \bigplus_{(a_1,\ldots,a_{\ell})\in A^{\ell}} \varphi_{a_1}(u_1)\cdot\varphi_{a_2}(u_2)\,\cdots\,\varphi_{a_{\ell}}(u_{\ell})\cdot f(a_1,\ldots,a_{\ell}).
 $$
-where $X^*$ is the least fixed point of $\varphi$.
+Then, for each $T_i \in \T$, let $V$ be a second-order assignment for $\A$ that assigns $T_i$ to $R$, let $f: A^{\ell}\to\nat$ be such that for each $(a_1,\ldots,a_{\ell})\in A^{\ell}$ it holds $f(a_1,\ldots,a_{\ell}) = \sem{\alpha\mid_{\pi(u_1,\ldots,u_{\ell})\to \beta_i(u_1,\ldots,u_{\ell})}}(\A,v,V)$, where $v$ is a first-order assignment for $\A$ that satisfies $a_i = v(x_i)$ for each open $x_i$ in $\alpha$.
+We define $Z[T_i] = f$.
 
+For a given first order asignment $v$, let $a_i = v(x_i)$, let $f = Z[T_n]$, and the operator is evaluated as:
+$$
+[\textbf{alfp}\,\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](\A,v) = f(a_1,\ldots,a_{\ell}).
+$$
 As an example, we will show a formula that counts the number of paths of size $n$ of a structure $\A$ with a binary relation $E$. First we define $\varphi(x,R)$:

### 2016-09-30 QSO/logics.tex
30 sep

@@
 
-For each $\qfo$ formulas $\alpha, \beta, \gamma$, where $\beta$ and $\gamma$ have $\ell$ and $m$ first-order open variables respectively, $u_1,\ldots,u_{\ell} \in \fv$, and $v_1,\ldots,v_m \in \{u_1,\ldots,u_{\ell}\}$, we define $\alpha\mid_{\beta(u_1,\ldots,u_{\ell})\to\gamma(v_1,\ldots,v_{m})}$ as $\alpha$ where every instance of the subformula $\beta(u_1,\ldots,u_{\ell})$ is replaced by $\gamma(v_1,\ldots,v_{m})$.
+For each $\qfo$ formulas $\alpha, \beta, \gamma$, where $\beta$ and $\gamma$ have $\ell$ and $m$ first-order free variables respectively, $u_1,\ldots,u_{\ell} \in \fv$, and $v_1,\ldots,v_m \in \{u_1,\ldots,u_{\ell}\}$, we define $\alpha\mid_{\beta(u_1,\ldots,u_{\ell})\to\gamma(v_1,\ldots,v_{m})}$ as $\alpha$ where every instance of the subformula $\beta(u_1,\ldots,u_{\ell})$ is replaced by $\gamma(v_1,\ldots,v_{m})$.
 Moreover, for each $a\in A$, suppose that $a$ is the $p$-th element in the order $<^{\A}$.
@@
 $$
-\varphi_a(x) = \exists x_1 \cdots \exists x_{p-1}[\bigwedge_{1\leq i,j < p}x_i\neq x_j \wedge \bigwedge_{i = 1}^p x_i < x \wedge \forall y(y < x \to \bigvee_{i = 1}^p y = x_i)].
+\varphi_a(x) = \exists x_1 \cdots \exists x_{p-1}[\bigwedge_{1\leq i,j < p}x_i\neq x_j \wedge \bigwedge_{i = 1}^{p-1} x_i < x \wedge \forall y(y < x \to \bigvee_{i = 1}^{p-1} y = x_i)].
 $$
@@
 To formally characterize the semantics of $[\textbf{alfp}\,\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$. %. 
-We name $\T = \{T_0,\ldots,T_n\}$ and we define the operator $Z:\T \to \{f\mid f:A^{\ell}\to\nat\}$ as follows.
+We name $\T = \{T_0,\ldots,T_n\}$ and we define a sequence of functions $\zeta_0,\ldots,\zeta_n:A^{\ell}\to\nat$ as follows.
 For each $T_i \in \T$, we define a $\qfo$ formula $\beta_i$ as follows.
 If $i = 0$, then $\beta_i(u_1,\ldots,u_{\ell}) = 0$.
-If $i \geq 1$, then let $f = Z[T_{i-1}]$ and
+If $i \geq 1$, then
 $$
-\beta_i(u_1,\ldots,u_{\ell}) = \bigplus_{(a_1,\ldots,a_{\ell})\in A^{\ell}} \varphi_{a_1}(u_1)\cdot\varphi_{a_2}(u_2)\,\cdots\,\varphi_{a_{\ell}}(u_{\ell})\cdot f(a_1,\ldots,a_{\ell}).
+\beta_i(u_1,\ldots,u_{\ell}) = \bigplus_{(a_1,\ldots,a_{\ell})\in A^{\ell}} \varphi_{a_1}(u_1)\cdot\varphi_{a_2}(u_2)\,\cdots\,\varphi_{a_{\ell}}(u_{\ell})\cdot \zeta_{i-1}(a_1,\ldots,a_{\ell}).
 $$
-Then, for each $T_i \in \T$, let $V$ be a second-order assignment for $\A$ that assigns $T_i$ to $R$, let $f: A^{\ell}\to\nat$ be such that for each $(a_1,\ldots,a_{\ell})\in A^{\ell}$ it holds $f(a_1,\ldots,a_{\ell}) = \sem{\alpha\mid_{\pi(u_1,\ldots,u_{\ell})\to \beta_i(u_1,\ldots,u_{\ell})}}(\A,v,V)$, where $v$ is a first-order assignment for $\A$ that satisfies $a_i = v(x_i)$ for each open $x_i$ in $\alpha$.
-We define $Z[T_i] = f$.
+Then, for each $T_i \in \T$, let $V$ be a second-order assignment for $\A$ that assigns $T_i$ to $R$, let $\zeta_i: A^{\ell}\to\nat$ be such that for each $(a_1,\ldots,a_{\ell})\in A^{\ell}$ it holds $\zeta_i(a_1,\ldots,a_{\ell}) = \sem{\alpha\mid_{\pi(u_1,\ldots,u_{\ell})\to \beta_i(u_1,\ldots,u_{\ell})}}(\A,v,V)$, where $v$ is a first-order assignment for $\A$ that satisfies $a_i = v(x_i)$ for each free $x_i$ in $\alpha$.
 
-For a given first order asignment $v$, let $a_i = v(x_i)$, let $f = Z[T_n]$, and the operator is evaluated as:
+For a given first order asignment $v$ and a second order assignment $V$, let $a_i = v(x_i)$, the operator is then evaluated as:
 $$
-[\textbf{alfp}\,\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](\A,v) = f(a_1,\ldots,a_{\ell}).
+\sem{[\textbf{alfp}\,\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)]}(\A,v,V) = \zeta_n(a_1,\ldots,a_{\ell}).
 $$
@@
 $$
-\varphi(x,R) = \forall y(x < y \vee x = y) \vee \exists z(R(z) \wedge \varphi_{succ}(z,x)),x
+\varphi(x,R) = \forall y(x < y \vee x = y) \vee \exists z(R(z) \wedge \varphi_{succ}(z,x)),
 $$
@@
 $$
-\alpha(x,y,R,\pi) = \neg \exists z(R(z)) + \Sigma z[E(x,z)\cdot \pi(z,y)].
+\alpha(x,y,R,\pi) = (\neg \exists zR(z))\cdot(x = y) + \Sigma z[\pi(x,z)\cdot E(z,y)].
 $$
@@
 We also define the less powerful operator $\pth$ as follows.
-Given a positive formula $\varphi(x,R)$ and a formula $\psi(x,y)$ we define $[\pth_{\varphi(x,R)} \psi(x,y)](x,y)$ as a formula that when evaluated on $(\A, a, b)$ induces a graph over $\A$ with the binary relation determined by $\psi(x,y)$ and counts the number of paths from $a$ to $b$ of a size determined by the $\varphi(x,R)$.
-Formally, we define:
-$$
-\sem{[\pth_{\varphi(x,R)} \psi(x,y)](x,y)} = \sem{[\alfp_{\varphi(x,R)} \alpha_{\psi}(x,y,R,\pi)](x,y)},
-$$
-where $\alpha_{\psi}(x,y,R,\pi) = \neg \exists z(R(z)) + \Sigma z[\psi(x,z)\cdot \pi(z,y)]$.
+Let $\A$ be an ordered structure.
+Given a formula $\psi(\bar{x},\bar{y})$, where $\vert \bar{x} \vert = \vert \bar{y} \vert = k$ let ${\cal G} = ({\cal V},\cal{E})$ be induced graph over the set of vertices ${\cal V} = A^k$, and for every $\bar{a},\bar{b}\in A^k$ it holds that ${\cal E}(\bar{a},\bar{b})$ if and only if $\A \models \psi(\bar{a},\bar{b})$.
+To formalize the semantics for this operator, let $n = \vert A^k \vert$.
+For a given first order assignment $v$ and a second order asssignment $V$, let $\bar{a} = v(\bar{x})$ and $\bar{b} = v(\bar{y})$, and $\sem{[\pth\, \psi(\bar{x},\bar{y})]}(\A,v,V)$ will take the value of the number of paths of size $n$ from $\bar{a}$ to $\bar{b}$ in the graph ${\cal G}$.
 

### 2016-10-05 QSO/cap-comp.tex
15 sep
5 sep

@@
 	For the first condition, let $\alpha\in\qfo(\tc)$, let $\A\in\ostr$ $v$ and $V$ first and second order assignments of $\A$ respectively.
-We evaluate $\sem{alpha}(\A,v,V)$ recursively.
+We evaluate $\sem{\alpha}(\A,v,V)$ recursively.
 A logic formula $\varphi$ is evaluated...?
@@
 \end{proof}
+
+\begin{theorem}
+	$\rqfo(\fo)$ captures $\fp$ over the class of ordered structures.
+\end{theorem}
+
+\begin{theorem}
+	$\tqfo(\fo)$ captures $\shl$ over the class of ordered structures.
+\end{theorem}
+\begin{proof}
+	For the second condition, let $f \in \shl$.
+Let $M$ be a non-deterministic logspace machine such that $f(w) = \acc_M(w)$ Let ${\cal Q} = \{q_1,\ldots,q_{\ell}\}$ be the set of states of $M$, where $q_1$ is the initial state, and let $q_{\ell}$ be the only final state of $M$.
+Let $\Sigma = \{0,1\}$ wlog.
+and let $w\in\Sigma^n$.
+Let $w = w_1\cdots w_n$, with $w_i \in \{0,1\}$ and let $\A$ be the ordered structure that represents this word with $A = \{1,\ldots,n\}$ and an unary predicate $P(\cdot)$ where $i\in P$ if and only if $w_i = 1$.
+We notate $M(w)$ as the graph of configurations of $M$ running on input $w$.
+Suppose that $M$ has a reading tape and a working tape, and the latter uses space $c\cdot\log(n)$.
+
+	We represent configurations with a tuple of fixed size $k$.
+The formula $\varphi(\bar{x},\bar{y})$ describes consecutive configurations.
+The formula $\varphi_I(\bar{x})$ models that $\bar{x}$ is the initial configuration of $M(w)$.
+The formula $\varphi_F(\bar{x})$ models that $\bar{x}$ is a accepting (final) configuration of $M(w)$.
+The formula we construct is:
+	$$
+	\alpha = \sa{\bar{x}}\sa{\bar{y}}([\pth \varphi(\bar{u},\bar{v})](\bar{x},\bar{y})\cdot \varphi_I(\bar{x})\cdot\varphi_F(\bar{y})).
+	$$
+	Note that there is a total of $m = (n+2)\cdot (c\log(n)+2)\cdot 3^{c\log n} \cdot \ell$ distinct configurations in $M(w)$, so we will represent them in the following way.
+The first two elements of the tuple represent the position of the head in the reading tape.
+The next $e$ elements represent the position of the head in the working tape.
+We choose an appropriate $e$ such that $n^e \geq c\log(n)$ for each $n$.
+The next $2c$ elements represent the string in the working tape.
+The last $\ell$ elements represent the state in which the machine is in.
+To specify, let $\bar{a} = (a_1,\ldots,a_m)$ be a tuple that represents a configuration in $M(w)$.
+	\begin{itemize}
+	\item $(a_1,a_2)$ represent the position of the head in the reading tape.
+If $a_1 = 1$, then the head is in position $a_2$ (the $a_2$-th character in the input).
+If $a_1 = 2$, then $a_2 = 1$ represents that the head is at the beginning of the tape (position 0), and if $a_2 = 2$, then the head is at position $n+1$.
+	\item $(a_3,\ldots,a_{e+2})$ represent the position of the head in the working tape.
+Specifically, let $h$ be the value of $a_{e+2}a_{e+1}\cdots a_4 a_3$ in base $n$.
+Then the head is at position $h$ of the working tape.
+	\item $(a_{e+3},\ldots,a_{2+e+2c})$ represents the string that is in the working tape.
+Specifically, let $h$ be the value of $a_{2+e+2c}a_{1+e+2c}\cdots a_{e+4}a_{e+3}$ in base $n$.
+Then the string in the working tape is $h$ in base 3, where 2 is interpreted as $\texttt{B}$.
+	\item $(a_{3+e+2c},\ldots,a_{2+e+2c+\ell})$ represents the current state $q_i$. Every value in the tuple is $1$ except the element $a_{3+e+2c+i}$ which is 2.
+	\end{itemize}
+	Now we describe the formulas.
+	\begin{align*}
+	\varphi_I(\bar{x}) &= \varphi_v(\bar{x}) \wedge \exists u(\varphi_1(u) \wedge \varphi_{succ}(u,x_{3+e+2c})) \\
+	\varphi_F(\bar{x}) &= \varphi_v(\bar{x}) \wedge \exists u(\varphi_1(u) \wedge \varphi_{succ}(u,x_{2+e+2c+\ell})),
+	\end{align*}
+	where $\varphi_v(\bar{x})$ describes that $\bar{x}$ is a valid configuration of $M(w)$ and $\varphi_1(x)$ describes that $x$ is the first element of $A$.
+\end{proof}
+
+\begin{theorem}
+	$\tqso(\fo)$ captures $\shp$ over the class of ordered structures.
+\end{theorem}
+
+\begin{theorem}
+	($\qfo(\lfp)$ or $\qfo(\fo)$) with algebraic least fixed point operator captures $\totp$ over the class of ordered structures.
+\end{theorem}
 

### 2016-10-05 QSO/logics.tex
5 sep

@@
 Note that there exists $n\in \nat$ such that $T_{n+1} = T_n$.
-Semantically, $[\lfpop\,\varphi(x_1,\ldots,x_k,R)]$ is defined such that for each $(a_1,\ldots,a_k)\in A^k$, it holds that $\A\models[\lfpop\,\varphi(x_1,\ldots,x_k,R)](a_1,\ldots,a_k)$ if and only if $\A\models T_n(a_1,\ldots,a_k)$.
+Semantically, $[\lfpop\varphi(x_1,\ldots,x_k,R)]$ is defined such that for each $(a_1,\ldots,a_k)\in A^k$, it holds that $\A\models[\lfpop\varphi(x_1,\ldots,x_k,R)](a_1,\ldots,a_k)$ if and only if $\A\models T_n(a_1,\ldots,a_k)$.
 
-We extend $\qfo$ with the operator $[\textbf{alfp}\,\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$.
-It is defined by a FO-formula $\varphi$ and a QFO-formula $\alpha$ which mentions a placeholder function $\pi$ with domain $\fv^{\ell}$.
+We extend $\qfo$ with the operator $[\alfp\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$.
+It is defined by a FO-formula $\varphi$ and a formula $\alpha$ in $\qfo$ without the $\Pi$ operator, which mentions a placeholder function $\pi$ with domain $\fv^{\ell}$.
 Let $\{T_i\}_{i\in\nat}$ and $n\in\nat$ be defined from $\varphi$ as it was for the $\lfp$ operator.
+
+\martin{No estoy seguro si quitar el operador $\Pi$ de la fórmula o solo decir que la funcion $\pi$ no puede estar mencionada en ninguna instancia del operador $\Pi$.}
 
@@
 
-To formally characterize the semantics of $[\textbf{alfp}\,\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$. %. 
+To formally characterize the semantics of $[\alfp\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$. %. 
 We name $\T = \{T_0,\ldots,T_n\}$ and we define a sequence of functions $\zeta_0,\ldots,\zeta_n:A^{\ell}\to\nat$ as follows.
@@
 $$
-\sem{[\textbf{alfp}\,\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)]}(\A,v,V) = \zeta_n(a_1,\ldots,a_{\ell}).
+\sem{[\alfp\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)]}(\A,v,V) = \zeta_n(a_1,\ldots,a_{\ell}).
 $$
@@
 $$
-Then, the formula $[\alfp_{\varphi(x,R)} \alpha(x,y,R,\pi)](x,y)$ when evaluated on $(\A,a,b)$ will count the number of paths of size $n$ from $a$ to $b$.
+Then, the formula $[\alfp\varphi(x,R)\mid \alpha(x,y,R,\pi)](x,y)$ when evaluated on $(\A,a,b)$ will count the number of paths of size $n$ from $a$ to $b$.
+This operator lets us define the set of recursive $\qfo$ formulas over $\R$ ($\rqfo$-formulas) using the following grammar:
+\begin{eqnarray*}
+\label{eq-def-rqfo}
+\alpha & := & \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ [\alfp \varphi \mid \alpha]
+\end{eqnarray*}
+where $\varphi$ is an $\fo$-formula over $\R$, $s \in \bbN$ and $x \in \fv$.
 
-We also define the less powerful operator $\pth$ as follows.
+We also define the less powerful operator {\bf path} as follows.
 Let $\A$ be an ordered structure.
@@
 To formalize the semantics for this operator, let $n = \vert A^k \vert$.
-For a given first order assignment $v$ and a second order asssignment $V$, let $\bar{a} = v(\bar{x})$ and $\bar{b} = v(\bar{y})$, and $\sem{[\pth\, \psi(\bar{x},\bar{y})]}(\A,v,V)$ will take the value of the number of paths of size $n$ from $\bar{a}$ to $\bar{b}$ in the graph ${\cal G}$.
+For a given first order assignment $v$ and a second order asssignment $V$, let $\bar{a} = v(\bar{x})$ and $\bar{b} = v(\bar{y})$, and $\sem{[\pth\, \psi(\bar{x},\bar{y})]}(\A,v,V)$ will take the value of the number of paths of size less or equal to $n$ from $\bar{a}$ to $\bar{b}$ in the graph ${\cal G}$.
+This operator lets us define the set of transitive $\qfo$ formulas over $\R$ ($\tqfo$-formulas) using the following grammar:
+\begin{eqnarray*}
+	\label{eq-def-tqfo}	\alpha & := & \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ [\pth \varphi]
+\end{eqnarray*}
+where $\varphi$ is an $\fo$-formula over $\R$, $s \in \bbN$ and $x \in \fv$.
 
+We also define the set of transitive $\qso$ formulas ($\tqso$-formulas) using the following grammar:
+\begin{eqnarray*}
+	\label{eq-def-tqso}	\alpha & := & \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha \ \mid \ [\pth \varphi]
+\end{eqnarray*}
+where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
 \begin{theorem}
-	Given a positive $\fo$ formula $\varphi(\bar{x},R)$ and a $\qfo$ formula $\alpha(\bar{x})$, there exists a $\qso$ formula $\beta(\bar{x})$ such that $\sem{[\alfp_{\varphi(\bar{x},R)} \alpha_{\psi}(\bar{x},R)](\bar{x})} = \sem{\beta(\bar{x})}$.
+	Given a positive $\fo$ formula $\varphi(\bar{x},R)$ and a $\qfo$ formula $\alpha(\bar{x})$, there exists a $\qso$ formula $\beta(\bar{x})$ such that $\sem{[\alfp\varphi(\bar{x},R)\mid \alpha(\bar{x},R)](\bar{x})} = \sem{\beta(\bar{x})}$.
 \end{theorem}

### 2016-10-19 QSO/cap-comp.tex
19 oct

--- 2016-08-24 QSO/cap-comp.tex
@@
 \end{proof}
+
+\begin{theorem}
+	$\rqfo(\fo)$ captures $\fp$ over the class of ordered structures.
+\end{theorem}
+
+\begin{theorem}
+	$\tqfo(\fo)$ captures $\shl$ over the class of ordered structures.
+\end{theorem}
+\begin{proof}
+	For the second condition, let $f \in \shl$.
+We will address the case where $\R$ contains only one binary predicate $E$, and the rest of the cases can be deduced from this.
+Let $M$ be a non-deterministic logspace machine such that $f(\A) = \acc_M(\A)$ for each $\A \in \ostr[R]$.
+Suppose ${\cal Q} = \{q_1,\ldots,q_{\ell}\}$ is the set of states of $M$, where $q_1$ is the initial state, and $q_{\ell}$ is only final state of $M$.
+Let $n = \vert A \vert$ and let $w = \enc(\A) \in \{0,1\}^{n^2}$.
+We assume that $M$ with input $w$ uses space $s_M(w) < c\cdot\log(n)$ and furthermore, $s_M(w) < n-2$.
+We notate $M(w)$ as the graph of configurations of $M$ running on input $w$.
+
+	We represent configurations with a tuple of fixed size $k$.
+The formula $\varphi(\bar{x},\bar{y})$ describes a procedure that given a configuration generates a possible next configuration.
+The formula $\varphi_I(\bar{x})$ models that $\bar{x}$ is the initial configuration of $M(w)$.
+The formula $\varphi_F(\bar{x})$ models that $\bar{x}$ is an accepting (final) configuration of $M(w)$.
+The formula we construct is:
+	$$
+	\alpha = \sa{\bar{x}}\sa{\bar{y}}([\pth \varphi(\bar{u},\bar{v})](\bar{x},\bar{y})\cdot \varphi_I(\bar{x})\cdot\varphi_F(\bar{y})).
+	$$
+
+	To illustrate our idea, we will show a greatly simplified example.
+Consider a machine $M$ that works in exactly $\log_2(n)$ space and only allows 0 or 1 in the working tape.
+Consider an input $\A$ of size 16 (that is, $A = \{0,\ldots,9,A,\ldots,F\}$).
+Let some configuration $s$ have 0011 in the working tape, the head in the input tape is in position 26, and the head in the input tape is in position 2 (we consider 0-indexed positions).
+Also, $Q = \{q_1,\ldots,q_5\}$ and the current state is $q_3$.
+
+	As a first approach, we will use a 9-tuple $\bar{a} = (a_1,\ldots,a_9)$ to represent $s$.
+That is, $(a_1,a_2) = (1,A)$ represent the position of the head in the input tape, $a_3 = 2$ represents the position of the head in the working tape, $a_4 = C$ (1100b in base 16) represents the content of the working tape, and $(a_5,\ldots,a_9) = (0,0,1,0,0)$ represents the current state.
+Then $\bar{a} = (1,A,2,C,0,0,1,0,0)$ will represent $s$.
+It is not possible \martin{no se si ser\'a cierto esto} to find the $i$-th bit of an element $x$, or much less switch the $i$-th bit of an element, solely with $\fo$.
+So consider the following procedure.
+
+	\begin{algorithm}
+	\caption{If the $i$-th bit in $x$ is 1 replace it by 0}
+	\begin{algorithmic}
+		\State $b \gets bit(x,i)$
+		\State $\textbf{assert } b = 1$
+		\State $y = pow2(i)$
+		\State $z = substract(x,y)$
+		\State \Return $z$.
+	\end{algorithmic}
+	\end{algorithm}	
+	And the details for $bit$, $pow2$ and $substract$ are left for the reader as an exercise.
+
+	Note that there is an upper bound of $(n^2+2)\cdot (c\log(n)+2)\cdot 3^{c\log n} \cdot \ell$ distinct configurations in $M(w)$.
+Given a configuration $s\in M(w)$, we construct a tuple $\bar{a} = (a_0,\ldots,a_{k-1})$ that represents $s$.
+	\begin{itemize}
+	\item $(a_0,a_1,a_2)$ represent the position of the head in the reading tape.
+If $a_0 = 1$, then the head is in position $a_1a_2$ (the $a_1a_2$-th character in the input).
+If $a_1 = 2$, then $a_2 = 1$ represents that the head is at the beginning of the tape (position 0), and if $a_2 = 2$, then the head is at position $n+1$.
+	\item $a_3$ represents the position of the head in the working tape.
+	\item $(a_4,\ldots,a_{3+2c})$ represents the string that is in the working tape.
+Specifically, let $h$ be the value of $a_{3+2c}a_{2+2c}\cdots a_{5}a_{4}$ in base $n$.
+Then the string in the working tape is $h$ in base 3, where 2 is interpreted as $\texttt{B}$.
+	\item $(a_{4+2c},\ldots,a_{3+2c+\ell})$ represents the current state $q_i$. Every value in the tuple is $1$ except the element $a_{3+2c+i}$ which is 2.	
+	\end{itemize}
+
+	Now we describe the formulas.
+	\begin{align*}
+	\varphi_I(\bar{x}) &= \varphi_v(\bar{x}) \wedge \exists u(\varphi_1(u) \wedge \varphi_{succ}(u,x_{3+e+2c})), \\
+	\varphi_F(\bar{x}) &= \varphi_v(\bar{x}) \wedge \exists u(\varphi_1(u) \wedge \varphi_{succ}(u,x_{2+e+2c+\ell})),
+	\end{align*}
+	where $\varphi_v(\bar{x})$ describes that $\bar{x}$ is a valid configuration of $M(w)$ and $\varphi_1(x)$ describes that $x$ is the first element of $A$.
+	\begin{align*}
+	\varphi_v^1(\bar{x}) &= \exists u(\varphi_1(u) \wedge (x_1 = u \vee \varphi_{succ}(u,x_1)) \wedge \varphi_{succ}(u,x_1) \to (x_2 = u \vee \varphi_{succ}(u,x_2))), \\
+	\varphi_v^4(\bar{x}) &= \exists u(\varphi_1(u) \wedge \bigvee_{i = 1}^{\ell}(\varphi_{succ}(u,x_{2+e+2c+i})  \wedge \bigwedge\limits_{\substack{i = 1\\ j \neq i}}^{\ell} x_{2+e+2c+j} = u)).
+	\end{align*}
+\end{proof}
+
+\begin{theorem}
+	$\tqso(\fo)$ captures $\shp$ over the class of ordered structures.
+\end{theorem}
+
+\begin{theorem}
+	($\qfo(\lfp)$ or $\qfo(\fo)$) with algebraic least fixed point operator captures $\totp$ over the class of ordered structures.
+\end{theorem}
 

### 2016-10-26 QSO/cap-comp.tex
26 oct

@@
 Then $\bar{a} = (1,A,2,C,0,0,1,0,0)$ will represent $s$.
-It is not possible \martin{no se si ser\'a cierto esto} to find the $i$-th bit of an element $x$, or much less switch the $i$-th bit of an element, solely with $\fo$.
-So consider the following procedure.
 
-	\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 1 replace it by 0}
+	The problem that arises from this representation, is that to describe a transition in $M$ we need to read an arbitrary character in the working tape, and therefore we have to obtain the $a_3$-th bit in $a_4$.
+Furthermore, to get the following configuration, we need compute $a_4$ with the $a_3$-th bit flipped.
+This is generally not possible to describe with an $FO$ formula.
+Consider the following procedure that receives $x = a_4$ and $i = a_3$.
+
+	\begin{algorithm} \label{switch1to0}
+	\caption{If the $i$-th bit in $x$ is 1 replace it by 0 and return the result}
 	\begin{algorithmic}
-		\State $b \gets bit(x,i)$
-		\State $\textbf{assert } b = 1$
-		\State $y = pow2(i)$
-		\State $z = substract(x,y)$
-		\State \Return $z$.
+		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
+		\While{$j > 0$}
+		\State $v \gets 0$
+		\While{$u > 1$}
+		\State $u \gets u-2,\; v \gets v+1$
+		\EndWhile
+		\State $u\gets v,\; j \gets j-1$
+		\EndWhile
+		\While{$u > 1$}
+		\State $u \gets u-2$
+		\EndWhile
+		\State $\textbf{assert } u = 1$ \Comment{If $u \neq 1$ simply stop}	
+		\State $y \gets 1$ \Comment{Compute $2^i$ and store it in $y$}
+		\While{$i > 0$}
+		\State $z \gets 0$
+		\While{$y > 0$}
+		\State $z \gets z+2,\; y \gets y-1$
+		\EndWhile
+		\State $i \gets i-1,\; y \gets z$
+		\EndWhile
+		\While{$y > 0$} \Comment{Substract $y$ from $x$}
+		\State $x \gets x-1,\; y \gets y-1$
+		\EndWhile
+		\State \Return $x$.
 	\end{algorithmic}
 	\end{algorithm}	
-	And the details for $bit$, $pow2$ and $substract$ are left for the reader as an exercise.
+	Each of the instructions can be expressed with $\fo$, so our strategy is to use the $\pth$ operator to simulate the algorithm and then we can describe a transition using the processed value of $a_4$.
+Note that we need to describe three more procedures to simulate the transitions for $0\to 0$, $0 \to 1$ and $1\to 1$.
 
-	Note that there is an upper bound of $(n^2+2)\cdot (c\log(n)+2)\cdot 3^{c\log n} \cdot \ell$ distinct configurations in $M(w)$.
-Given a configuration $s\in M(w)$, we construct a tuple $\bar{a} = (a_0,\ldots,a_{k-1})$ that represents $s$.
-	\begin{itemize}
-	\item $(a_0,a_1,a_2)$ represent the position of the head in the reading tape.
-If $a_0 = 1$, then the head is in position $a_1a_2$ (the $a_1a_2$-th character in the input).
-If $a_1 = 2$, then $a_2 = 1$ represents that the head is at the beginning of the tape (position 0), and if $a_2 = 2$, then the head is at position $n+1$.
-	\item $a_3$ represents the position of the head in the working tape.
-	\item $(a_4,\ldots,a_{3+2c})$ represents the string that is in the working tape.
-Specifically, let $h$ be the value of $a_{3+2c}a_{2+2c}\cdots a_{5}a_{4}$ in base $n$.
-Then the string in the working tape is $h$ in base 3, where 2 is interpreted as $\texttt{B}$.
-	\item $(a_{4+2c},\ldots,a_{3+2c+\ell})$ represents the current state $q_i$. Every value in the tuple is $1$ except the element $a_{3+2c+i}$ which is 2.	
-	\end{itemize}
+	We will now describe how to simulate both the procedure and the transition.
+A procedure tuple $\bar{p} = (a_1,\ldots,a_{3+c+\ell},b_1,b_2,c_1,c_2,c_3,d_1,\ldots,d_{5c+2})$ represents the current configuration of $M(w)$ in $a_1,\ldots,a_{2+c+\ell}$, the values that will be read and written in the working tape in $b_1,b_2$, the instruction pointer in $c_1,c_2,c_3$ and the values stored in memory in $d_1,\ldots,d_{10c+2}$.
+In detail:
+	\begin{enumerate}
+		\item $a_1,a_2$ and $a_3$ represent the position of the head in the input tape and the working tape, respectively, $a_4,\ldots,a_{3+c}$ represent the content of the working tape and $a_{4+c},\ldots,a_{3+c+\ell}$ represent the current state in the current configuration that is being processed.
+		\item $b_1$ and $b_2$ are equal to the value that is being read in the working tape and the value that will be written in the working tape respectively.
+		\item $c_1,c_2,c_2$ represent the instruction pointer in the procedure. Only 8 different instructions are needed in the simulation.
+		\item Each value in memory of $x,y,z,u,v$ need $c$ elements to represent them and $i,j$ need only one. We map $(d_1\ldots,d_{c}) \to x$, $(d_{c+1}\ldots,d_{2c}) \to y$,
+		$(d_{2c+1}\ldots,d_{3c}) \to z$, $(d_{3c+1}\ldots,d_{4c}) \to u$,
+		$(d_{4c+1}\ldots,d_{5c}) \to v$, $d_{5c+1} \to i$ and $d_{5c+2}\to j$.
+	\end{enumerate}
+	For each transition $\delta \subseteq Q \times \{0,1\} \times \{0,1\} \times Q \times \{-1,=,+1\} \times \{0,1\} \times \{-1,=,+1\}$ we define a formula $\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$, where $\bar{x} = (x_1,\ldots,x_{3+c+\ell})$, $\bar{s} = (s_1,s_2)$, $\bar{w} = (w_1,w_2,w_3)$, $\bar{u} = (u_1,\ldots,u_{5c+2})$, $\bar{y} = (y_1,\ldots,y_{3+c+\ell})$, $\bar{t} = (t_1,t_2)$, $\bar{z} = (z_1,z_2,z_3)$ and $\bar{v} = (v_1,\ldots,v_{5c+2})$ that describes the procedure to compute the values in the configuration and the transition itself.
+We will describe the formula part by part.
+Suppose $\delta = (q_i,a,1,q_j,op_1,0,op_2)$, so we have to simulate Algorithm \ref{switch1to0}.
 
-	Now we describe the formulas.
-	\begin{align*}
-	\varphi_I(\bar{x}) &= \varphi_v(\bar{x}) \wedge \exists u(\varphi_1(u) \wedge \varphi_{succ}(u,x_{3+e+2c})), \\
-	\varphi_F(\bar{x}) &= \varphi_v(\bar{x}) \wedge \exists u(\varphi_1(u) \wedge \varphi_{succ}(u,x_{2+e+2c+\ell})),
-	\end{align*}
-	where $\varphi_v(\bar{x})$ describes that $\bar{x}$ is a valid configuration of $M(w)$ and $\varphi_1(x)$ describes that $x$ is the first element of $A$.
-	\begin{align*}
-	\varphi_v^1(\bar{x}) &= \exists u(\varphi_1(u) \wedge (x_1 = u \vee \varphi_{succ}(u,x_1)) \wedge \varphi_{succ}(u,x_1) \to (x_2 = u \vee \varphi_{succ}(u,x_2))), \\
-	\varphi_v^4(\bar{x}) &= \exists u(\varphi_1(u) \wedge \bigvee_{i = 1}^{\ell}(\varphi_{succ}(u,x_{2+e+2c+i})  \wedge \bigwedge\limits_{\substack{i = 1\\ j \neq i}}^{\ell} x_{2+e+2c+j} = u)).
-	\end{align*}
+	We start from instruction 0, which means that the procedure has not started yet and every value in the tuple is 0 except for the configuration values.
+It also initializes all the values in the tuple to 0 except for $x,u,i,j$.
+	\begin{multline*}
+	\varphi^{0,1}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{0,0}(s_1,s_2)\wedge\varphi^b_0(\bar{w}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_1(\bar{z})\, \wedge \\
+	\bigwedge_{i = 1}^c v_i = x_{3+i} \wedge \bigwedge_{i = c+1}^{2c} \varphi_0(v_i) \wedge \bigwedge_{i = 2c+1}^{3c} \varphi_0(v_i) \wedge \bigwedge_{i = 1}^c v_{3c+i} = x_{3+i} \wedge \bigwedge_{i = 4c+1}^{5c} \varphi_0(v_i) \wedge v_{5c+1} = x_3 \wedge v_{5c+2} = x_3.
+	\end{multline*}
+	Instruction 1 which checks whether the value of $j$ ($d_{5c+2}$ in the tuple) is more than 0 or not, and then proceeds to instruction 2 or 3 on each case.
+	\begin{multline*}
+	\varphi^{1,2}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
+	\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_1(\bar{w}) \wedge \neg \varphi_0(u_{5c+2}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_2(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i, \\
+	\varphi^{1,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
+	\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_1(\bar{w}) \wedge \varphi_0(u_{5c+2}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_3(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
+	\end{multline*}
+	Instruction 2 checks the value of $u$ ($d_{3c+1},\ldots,d_{4c}$ in the tuple).
+If it is $> 1$ then it substracts 2 from $u$ and adds 1 to $v$ ($d_{4c+1},\ldots,d_{5c}$ in the tuple), then repeats instruction 2.
+If it is equal to 0 or 1, then moves the value of $v$ to $u$, substracts 1 from $j$ and goes back to instruction 1.
+	\begin{multline*}
+	\varphi^{2,2}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
+	\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_2(\bar{w}) \wedge \neg \varphi^m_0(u_{3c+1},\ldots,u_{4c}) \wedge \neg \varphi^m_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_2(\bar{z})\, \wedge \\
+	\bigwedge_{i = 1}^c u_i = v_i \wedge
+	\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
+	\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
+	\varphi^m_{-2}(u_{3c+1},\ldots,u_{4c},v_{3c+1},\ldots,v_{4c})\, \wedge \\
+	\varphi^m_{+1}(u_{4c+1},\ldots,u_{5c},v_{4c+1},\ldots,v_{5c}) \wedge u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}. \\
+	\varphi^{2,1}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_2(\bar{w}) \wedge ( \varphi^m_0(u_{3c+1},\ldots,u_{4c}) \vee \varphi^m_1(u_{3c+1},\ldots,u_{4c})) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_1(\bar{z})\, \wedge \\
+	\bigwedge_{i = 1}^c u_i = v_i \wedge
+	\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
+	\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
+	\bigwedge_{i = 3c+1}^{4c} u_{c+i} = v_i \wedge
+	\varphi^m_0(v_{4c+1},\ldots,v_{5c}) \wedge
+	u_{5c+1} = v_{5c+1} \wedge \varphi_{-1}(u_{5c+2},v_{5c+2}).	
+	\end{multline*}
+	Instruction 3 calculates the value of $u \mod 2$, that is, it repeats instruction 3 until the value of $u$ is equal to 0 or 1.
+On each iteration, it substracts 2 from $u$.
+Moreover, if the value of $u$ at the end of the iterations is not 1 then there is no step defined.
+	\begin{multline*}
+	\varphi^{3,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_3(\bar{w}) \wedge \neg \varphi^m_0(u_{3c+1},\ldots,u_{4c}) \wedge \neg \varphi^m_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_3(\bar{z})\, \wedge \\
+	\bigwedge_{i = 1}^c u_i = v_i \wedge
+	\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
+	\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
+	\varphi^m_{-2}(u_{3c+1},\ldots,u_{4c},v_{3c+1},\ldots,v_{4c}) \wedge
+	\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
+	u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}, \\
+	\varphi^{3,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
+	\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_3(\bar{w}) \wedge \varphi^m_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_4(\bar{z})\, \wedge \\
+	\bigwedge_{i = 1}^c u_i = v_i \wedge
+	\varphi^m_1(v_{c+1},\ldots,v_{2c}) \wedge
+	\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
+	\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
+	\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
+	u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}
+	\end{multline*}
+	Instruction 4 checks the value of $i$ ($d_{5c+1}$ in the tuple.) If it is not 0 then goes to instruction 5 and if is 0 then goes to instruction 6.
+Moreover it initializes the value of $z$ ($d_{2c+1},\ldots,d_{3c}$ in the tuple) to 0 (which was 0 all along.)
+	\begin{multline*}
+	\varphi^{4,5}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_4(\bar{w}) \wedge \neg \varphi_0(u_{5c+1}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_5(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i, \\
+	\varphi^{4,6}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_4(\bar{w}) \wedge \varphi_0(u_{5c+1}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_6(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
+	\end{multline*}
+	Instruction 5 checks the value of $y$ ($d_{c+1},\ldots,d_{2c}$ in the tuple.) If it is more than 0 then it adds 2 to $z$ and substracts 1 from $y$, then repeats instruction 2.
+If it is not, then copies the value of $z$ to $y$ and subtracts 1 from $i$ and returns to instruction 4.
+	\begin{multline*}
+	\varphi^{5,5}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_5(\bar{w}) \wedge \neg \varphi^m_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_5(\bar{z}) \, \wedge \\
+	\bigwedge_{i = 1}^c u_i = v_i \wedge
+	\varphi^m_{-1}(u_{c+1},\ldots,u_{2c},v_{c+1},\ldots,v_{2c}) \wedge
+	\varphi^m_{+2}(u_{2c+1},\ldots,u_{3c},v_{2c+1},\ldots,v_{3c})\, \wedge \\
+	\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
+	\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
+	u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}, \\
+	\varphi^{5,4}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_5(\bar{w}) \wedge \varphi^m_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_4(\bar{z}) \, \wedge \\
+	\bigwedge_{i = 1}^c u_i = v_i \wedge
+	\bigwedge_{i = c+1}^{2c} u_{c+i} = v_i \wedge
+	\varphi^m_0(v_{2c+1},\ldots,v_{3c}) \wedge
+	\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
+	\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
+	\varphi_{-1}(u_{5c+1},v_{5c+1}) \wedge u_{5c+2} = v_{5c+2}
+	\end{multline*}
+	Instruction 6 checks the value of $y$. If it is more than 0, then subtracts 1 from $x$ and $y$ and repeats instruction 6. If it is not, then goes to instruction 7.
+	\begin{multline*}
+	\varphi^{6,6}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_6(\bar{w}) \wedge \neg \varphi^m_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_6(\bar{z}) \, \wedge \\
+	\varphi^m_{-1}(u_1,\ldots,u_c,v_i,\ldots,v_c) \wedge \varphi^m_{-1}(u_{c+1},\ldots,u_{2c},v_{c+1},\ldots,v_{2c}) \wedge \bigwedge_{i = 3c+1}^{5c+2} u_i = v_i]\,\vee \\
+	\varphi^{6,7}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_6(\bar{w}) \wedge \varphi^m_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_7(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
+	\end{multline*}
+	Instruction 7 stores the value of $x$ after the corresponding bit has been switched.
+Then we can define $\varphi_{\delta}$ which also simulates the actual transition.
+If $u$ equals 1, then copy what is stored in $x$ to $a_4,\ldots,a_{3+c}$, go from state $q_i$ to state $q_j$, and move the heads to their corresponding positions.
+	\begin{multline*}
+	\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 	[\bigwedge_{i = 1}^{3+c+\ell} x_i = y_i \wedge (\varphi^{0,1}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{1,2}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})\, \vee \\ \varphi^{1,3}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{2,2}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{2,1}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{3,3}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \\ \varphi^{3,4}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{4,5}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{4,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{5,5}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \\ \varphi^{5,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{6,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{6,7}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}))] \, \vee \\
+	[\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_7(\bar{w}) \wedge \varphi^m_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{0,0}(t_1,t_2) \wedge \varphi^b_0(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i \,\wedge \\
+	\varphi^2_{op_1}(x_1,x_2,y_1,y_2) \wedge \varphi_{op_2}(x_3,y_3) \wedge \bigwedge_{i = 1}^c u_i = x_{3+i} \wedge \varphi^q_i(x_{4+c},\ldots,x_{3+c+\ell}) \wedge \varphi^q_j(y_{4+c},\ldots,y_{3+c+\ell})].
+	\end{multline*}
+	Note that we also need to specify that the program we are following is Algorithm \ref{switch1to0} so we store $1,0$ in $b_1,b_2$ all along the procedure.
+We now describe the three other algorithms that compute the switches from $0\to 0$, $0\to 1$ and $1\to 1$.
 \end{proof}

### 2016-11-02 QSO/cap-comp.tex
2 nov

@@
 	\end{enumerate}
-	For each transition $\delta \subseteq Q \times \{0,1\} \times \{0,1\} \times Q \times \{-1,=,+1\} \times \{0,1\} \times \{-1,=,+1\}$ we define a formula $\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$, where $\bar{x} = (x_1,\ldots,x_{3+c+\ell})$, $\bar{s} = (s_1,s_2)$, $\bar{w} = (w_1,w_2,w_3)$, $\bar{u} = (u_1,\ldots,u_{5c+2})$, $\bar{y} = (y_1,\ldots,y_{3+c+\ell})$, $\bar{t} = (t_1,t_2)$, $\bar{z} = (z_1,z_2,z_3)$ and $\bar{v} = (v_1,\ldots,v_{5c+2})$ that describes the procedure to compute the values in the configuration and the transition itself.
+	For each transition $\delta \in \Delta \subseteq Q \times \{0,1\} \times \{0,1\} \times Q \times \{-1,=,+1\} \times \{0,1\} \times \{-1,=,+1\}$ we define a formula $\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$, where $\bar{x} = (x_1,\ldots,x_{3+c+\ell})$, $\bar{s} = (s_1,s_2)$, $\bar{w} = (w_1,w_2,w_3)$, $\bar{u} = (u_1,\ldots,u_{5c+2})$, $\bar{y} = (y_1,\ldots,y_{3+c+\ell})$, $\bar{t} = (t_1,t_2)$, $\bar{z} = (z_1,z_2,z_3)$ and $\bar{v} = (v_1,\ldots,v_{5c+2})$.
+The tuples $\bar{x}$ and $\bar{y}$ represent the current and next configuration of $M$ respectively, $\bar{s}$ and $\bar{y}$ indicate which algorithm is being followed, $\bar{w}$ and $\bar{z}$ represent the current and next instruction of the algorithm, $\bar{u}$ and $\bar{v}$ represent the current and next values in memory.
 We will describe the formula part by part.
 Suppose $\delta = (q_i,a,1,q_j,op_1,0,op_2)$, so we have to simulate Algorithm \ref{switch1to0}.
+
+	First we define some auxiliary formulas:
+	\begin{align*}
+	\varphi_0(x) &= \neg\exists y(y < x), \\
+	\varphi_{+1}(x,y) &= x < y \wedge \neg\exists z(x < z \wedge z < y), \\
+	\varphi_{-1}(x,y) &= \varphi_{+1}(y,x), \\
+	\varphi_1(x) &= \exists y(\varphi_0(y) \wedge \varphi_{+1}(y,x)), \\
+\text{for $i,j \in\{0,1\}$, }	\varphi_{i,j}(x,y) &= \varphi_i(x) \wedge \varphi_j(y), \\
+	\text{for each $k \leq 7$, }\varphi^b_k(x_1,x_2,x_3) &= \varphi_{a_1}(x_1) \wedge \varphi_{a_2}(x_2) \wedge \varphi_{a_3}(x_3)\text{, where $a_1a_2a_3$ is the value of $k$ in binary}, \\
+	\mu_{<}(x_1,\ldots,x_c,y_1,\ldots,y_c) &= \bigvee_{i = 1}^c ((\bigwedge_{j = 1}^{i-1} x_j = y_j) \wedge x_i < y_i), \\
+	\mu_{+1}(\bar{x},\bar{y}) &= \mu_{<}(\bar{x},\bar{y}) \wedge \neg\exists \bar{z}(\mu_{<}(\bar{x},\bar{z}) \wedge \mu_{<}(\bar{z},\bar{y})), \\
+	\mu_{+2}(\bar{x},\bar{y}) &= \exists\bar{z}(\mu_{+1}(\bar{x},\bar{z}) \wedge \mu_{+1}(\bar{z},\bar{y})),\\
+	\mu_{-1}(\bar{x},\bar{y}) &= \varphi_{+1}(\bar{y},\bar{x}),\\
+	\mu_{-2}(\bar{x},\bar{y}) &= \varphi_{+2}(\bar{y},\bar{x}),\\
+	\mu_{0}(\bar{x}) &= \neg\exists\bar{y}(\mu_{<}(\bar{y},\bar{x})),\\
+	\mu_{1}(\bar{x}) &= \exists\bar{y}(\mu_{0}(\bar{y})\wedge\mu_{+1}(\bar{y},\bar{x})),\\
+	\varphi^2_{<}(x_1,x_2,y_1,y_2) &= x_1 < y_1 \vee (x_1 = y_1 \wedge x_2 < y_2),\\
+	\varphi^2_{+1}(x_1,x_2,y_1,y_2), \varphi^2_{-1}(x_1,x_2,y_1,y_2) &\text{ defined analogously to $\mu_{+1}$ and $\mu_{-1}$},\\
+	\text{for each } q_i\in Q, \varphi^q_i(x_1,\ldots,x_{\ell}) &= \bigwedge_{\substack{j = 1 \\ j \neq i}}^{\ell} \varphi_0(x_j) \wedge \varphi_1(x_i), \\
+	\varphi^E_0(x_1,x_2) &= \neg E(x_1,x_2),\\		\varphi^E_1(x_1,x_2) &= E(x_1,x_2),\\
+	\end{align*}
 
@@
 	\varphi^{2,2}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-	\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_2(\bar{w}) \wedge \neg \varphi^m_0(u_{3c+1},\ldots,u_{4c}) \wedge \neg \varphi^m_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_2(\bar{z})\, \wedge \\
+	\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_2(\bar{w}) \wedge \neg \mu_0(u_{3c+1},\ldots,u_{4c}) \wedge \neg \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_2(\bar{z})\, \wedge \\
 	\bigwedge_{i = 1}^c u_i = v_i \wedge
@@
 	\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-	\varphi^m_{-2}(u_{3c+1},\ldots,u_{4c},v_{3c+1},\ldots,v_{4c})\, \wedge \\
-	\varphi^m_{+1}(u_{4c+1},\ldots,u_{5c},v_{4c+1},\ldots,v_{5c}) \wedge u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}. \\
-	\varphi^{2,1}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_2(\bar{w}) \wedge ( \varphi^m_0(u_{3c+1},\ldots,u_{4c}) \vee \varphi^m_1(u_{3c+1},\ldots,u_{4c})) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_1(\bar{z})\, \wedge \\
+	\mu_{-2}(u_{3c+1},\ldots,u_{4c},v_{3c+1},\ldots,v_{4c})\, \wedge \\
+	\mu_{+1}(u_{4c+1},\ldots,u_{5c},v_{4c+1},\ldots,v_{5c}) \wedge u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}. \\
+	\varphi^{2,1}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_2(\bar{w}) \wedge ( \mu_0(u_{3c+1},\ldots,u_{4c}) \vee \mu_1(u_{3c+1},\ldots,u_{4c})) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_1(\bar{z})\, \wedge \\
 	\bigwedge_{i = 1}^c u_i = v_i \wedge
@@
 	\bigwedge_{i = 3c+1}^{4c} u_{c+i} = v_i \wedge
-	\varphi^m_0(v_{4c+1},\ldots,v_{5c}) \wedge
+	\mu_0(v_{4c+1},\ldots,v_{5c}) \wedge
 	u_{5c+1} = v_{5c+1} \wedge \varphi_{-1}(u_{5c+2},v_{5c+2}).	
@@
 	\begin{multline*}
-	\varphi^{3,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_3(\bar{w}) \wedge \neg \varphi^m_0(u_{3c+1},\ldots,u_{4c}) \wedge \neg \varphi^m_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_3(\bar{z})\, \wedge \\
+	\varphi^{3,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_3(\bar{w}) \wedge \neg \mu_0(u_{3c+1},\ldots,u_{4c}) \wedge \neg \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_3(\bar{z})\, \wedge \\
 	\bigwedge_{i = 1}^c u_i = v_i \wedge
@@
 	\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-	\varphi^m_{-2}(u_{3c+1},\ldots,u_{4c},v_{3c+1},\ldots,v_{4c}) \wedge
+	\mu_{-2}(u_{3c+1},\ldots,u_{4c},v_{3c+1},\ldots,v_{4c}) \wedge
 	\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
@@
 	\varphi^{3,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-	\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_3(\bar{w}) \wedge \varphi^m_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_4(\bar{z})\, \wedge \\
+	\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_3(\bar{w}) \wedge \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_4(\bar{z})\, \wedge \\
 	\bigwedge_{i = 1}^c u_i = v_i \wedge
-	\varphi^m_1(v_{c+1},\ldots,v_{2c}) \wedge
+	\mu_1(v_{c+1},\ldots,v_{2c}) \wedge
 	\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
@@
 	\begin{multline*}
-	\varphi^{5,5}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_5(\bar{w}) \wedge \neg \varphi^m_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_5(\bar{z}) \, \wedge \\
+	\varphi^{5,5}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_5(\bar{w}) \wedge \neg \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_5(\bar{z}) \, \wedge \\
 	\bigwedge_{i = 1}^c u_i = v_i \wedge
-	\varphi^m_{-1}(u_{c+1},\ldots,u_{2c},v_{c+1},\ldots,v_{2c}) \wedge
-	\varphi^m_{+2}(u_{2c+1},\ldots,u_{3c},v_{2c+1},\ldots,v_{3c})\, \wedge \\
+	\mu_{-1}(u_{c+1},\ldots,u_{2c},v_{c+1},\ldots,v_{2c}) \wedge
+	\mu_{+2}(u_{2c+1},\ldots,u_{3c},v_{2c+1},\ldots,v_{3c})\, \wedge \\
 	\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
@@
 	u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}, \\
-	\varphi^{5,4}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_5(\bar{w}) \wedge \varphi^m_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_4(\bar{z}) \, \wedge \\
+	\varphi^{5,4}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_5(\bar{w}) \wedge \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_4(\bar{z}) \, \wedge \\
 	\bigwedge_{i = 1}^c u_i = v_i \wedge
 	\bigwedge_{i = c+1}^{2c} u_{c+i} = v_i \wedge
-	\varphi^m_0(v_{2c+1},\ldots,v_{3c}) \wedge
+	\mu_0(v_{2c+1},\ldots,v_{3c}) \wedge
 	\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
@@
 	\begin{multline*}
-	\varphi^{6,6}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_6(\bar{w}) \wedge \neg \varphi^m_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_6(\bar{z}) \, \wedge \\
-	\varphi^m_{-1}(u_1,\ldots,u_c,v_i,\ldots,v_c) \wedge \varphi^m_{-1}(u_{c+1},\ldots,u_{2c},v_{c+1},\ldots,v_{2c}) \wedge \bigwedge_{i = 3c+1}^{5c+2} u_i = v_i]\,\vee \\
-	\varphi^{6,7}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_6(\bar{w}) \wedge \varphi^m_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_7(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
+	\varphi^{6,6}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_6(\bar{w}) \wedge \neg \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_6(\bar{z}) \, \wedge \\
+	\mu_{-1}(u_1,\ldots,u_c,v_i,\ldots,v_c) \wedge \mu_{-1}(u_{c+1},\ldots,u_{2c},v_{c+1},\ldots,v_{2c}) \wedge \bigwedge_{i = 3c+1}^{5c+2} u_i = v_i]\,\vee \\
+	\varphi^{6,7}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_6(\bar{w}) \wedge \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_7(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
 	\end{multline*}
@@
 	\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 	[\bigwedge_{i = 1}^{3+c+\ell} x_i = y_i \wedge (\varphi^{0,1}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{1,2}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})\, \vee \\ \varphi^{1,3}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{2,2}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{2,1}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{3,3}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \\ \varphi^{3,4}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{4,5}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{4,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{5,5}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \\ \varphi^{5,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{6,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{6,7}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}))] \, \vee \\
-	[\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_7(\bar{w}) \wedge \varphi^m_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{0,0}(t_1,t_2) \wedge \varphi^b_0(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i \,\wedge \\
+	[\varphi^E_a(x_1,x_2) \wedge \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_7(\bar{w}) \wedge \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{0,0}(t_1,t_2) \wedge \varphi^b_0(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i \,\wedge \\
 	\varphi^2_{op_1}(x_1,x_2,y_1,y_2) \wedge \varphi_{op_2}(x_3,y_3) \wedge \bigwedge_{i = 1}^c u_i = x_{3+i} \wedge \varphi^q_i(x_{4+c},\ldots,x_{3+c+\ell}) \wedge \varphi^q_j(y_{4+c},\ldots,y_{3+c+\ell})].
@@
 We now describe the three other algorithms that compute the switches from $0\to 0$, $0\to 1$ and $1\to 1$.
+	\begin{algorithm} \label{switch0to0}
+		\caption{If the $i$-th bit in $x$ is 0 return $x$}
+		\begin{algorithmic}
+			\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
+			\While{$j > 0$}
+			\State $v \gets 0$
+			\While{$u > 1$}
+			\State $u \gets u-2,\; v \gets v+1$
+			\EndWhile
+			\State $u\gets v,\; j \gets j-1$
+			\EndWhile
+			\While{$u > 1$}
+			\State $u \gets u-2$
+			\EndWhile
+			\State $\textbf{assert } u = 0$ \Comment{If $u \neq 0$ simply stop}	
+			\State \Return $x$.
+		\end{algorithmic}
+	\end{algorithm}
+	\begin{algorithm} \label{switch0to1}
+		\caption{If the $i$-th bit in $x$ is 0 replace it by 1 and return the result}
+		\begin{algorithmic}
+			\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
+			\While{$j > 0$}
+			\State $v \gets 0$
+			\While{$u > 1$}
+			\State $u \gets u-2,\; v \gets v+1$
+			\EndWhile
+			\State $u\gets v,\; j \gets j-1$
+			\EndWhile
+			\While{$u > 1$}
+			\State $u \gets u-2$
+			\EndWhile
+			\State $\textbf{assert } u = 0$ \Comment{If $u \neq 0$ simply stop}	
+			\State $y \gets 1$ \Comment{Compute $2^i$ and store it in $y$}
+			\While{$i > 0$}
+			\State $z \gets 0$
+			\While{$y > 0$}
+			\State $z \gets z+2,\; y \gets y-1$
+			\EndWhile
+			\State $i \gets i-1,\; y \gets z$
+			\EndWhile
+			\While{$y > 0$} \Comment{Add $y$ to $x$}
+			\State $x \gets x+1,\; y \gets y-1$
+			\EndWhile
+			\State \Return $x$.
+		\end{algorithmic}
+	\end{algorithm}	
+	\begin{algorithm} \label{switch1to1}
+		\caption{If the $i$-th bit in $x$ is 1 return $x$}
+		\begin{algorithmic}
+			\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
+			\While{$j > 0$}
+			\State $v \gets 0$
+			\While{$u > 1$}
+			\State $u \gets u-2,\; v \gets v+1$
+			\EndWhile
+			\State $u\gets v,\; j \gets j-1$
+			\EndWhile
+			\While{$u > 1$}
+			\State $u \gets u-2$
+			\EndWhile
+			\State $\textbf{assert } u = 1$ \Comment{If $u \neq 0$ simply stop}	
+			\State \Return $x$.
+		\end{algorithmic}
+	\end{algorithm}
+	For the other three cases, where $\delta = (q_i,a,0,q_j,op_1,0,op_2)$, $\delta = (q_i,a,0,q_j,op_1,1,op_2)$ and $\delta = (q_i,a,1,q_j,op_1,1,op_2)$, $\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$ is defined analogously.
+Then, $\varphi(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$ is defined as:
+	$$
+	\varphi(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \bigvee_{\delta \in \Delta} \varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}).
+	$$
+	Finally we define $\varphi_I$ and $\varphi_F$:
+	\begin{align*}
+	\varphi_I(\bar{x},\bar{s},\bar{w},\bar{u}) &= \varphi_0(x_1) \wedge \varphi_0(x_2) \wedge \varphi_0(x_3) \wedge \bigwedge_{i = 4}^{3+c}\varphi_0(x_i) \wedge \varphi_1(x_{4+c}) \wedge \bigwedge_{i = 5+c}^{3+c+\ell} \varphi_0(x_i)\wedge \varphi_{0,0}(s_1,s_2)\ \wedge \varphi^b_0(\bar{w}) \wedge \bigwedge_{i = 1}^{5c+2}\varphi_0(u_i).
+\\
+	\varphi_F(\bar{x},\bar{s},\bar{w},\bar{u}) &= \varphi^q_{\ell}(x_{4+c},\ldots,x_{3+c+\ell}) \wedge \varphi_{0,0}(s_1,s_2)\ \wedge \varphi^b_0(\bar{w}) \wedge \bigwedge_{i = 1}^{5c+2}\varphi_0(u_i),
+	\end{align*}
+	and then $\sem{\alpha}(\A) = \sem{\sa{\bar{x}}\sa{\bar{y}}([\pth \varphi(\bar{u},\bar{v})](\bar{x},\bar{y})\cdot \varphi_I(\bar{x})\cdot\varphi_F(\bar{y}))}(\A) = \acc_M(\A)$.
+
 \end{proof}

### 2016-11-25 pe/pe.tex
pe

\documentclass[12pt]{article}
\usepackage{fullpage}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsmath}

\newtheorem{theorem}{Theorem}

\def\pe{\#\text{PE}}
\def\ptime{\text{P}}
\def\shp{\#\text{P}}
\def\acc{\text{acc}}
\def\Q{\mathcal{Q}}
\def\M{\mathcal{M}}
\def\N{\mathbb{N}}

\title{On $\pe$}

\begin{document}
	\maketitle

	In this document we describe the combined counting machine model. A combined-counting machine (CCM) is a tuple $(\Q,\Sigma,q_0,\delta,F)$, that follows the following rules. There exist pairwise disjoint sets $V,C,E$ such that $\Q = V\cup C \cup E$, where $q_0\in V$, $V \cap F = \emptyset$ and $C \subseteq F$. For each $q \in Q, a \in \Sigma$:
	\begin{itemize}
	\item If $q\in V$ and $\delta(q,a) = \{(q',b,X)\}$, then $q\in V$.
	\item If $q\in V$ and $\vert\delta(q,a)\vert > 1$, for each $(q',b,X) \in \delta(q,a)$ it holds that $q'\in C \cup E$, and there exists some $(q',c,Y) \in \delta(q,a)$ such that $q'\in C$.
	\item If $q\in C$, for each $(q',b,X) \in \delta(q,a)$ it holds that $1q'\in C \cup E$, and there exists some $(q',c,Y) \in \delta(q,a)$ such that $q'\in C$.
	\item If $q\in E$, for each $(q',b,X) \in \delta(q,a)$ it holds that $q'\in E$.
	\end{itemize}
	Our hypothesis characterizes $\pe$ using this machines.
	\begin{theorem}
		$f\in\pe$ iff there exists a polynomial-time CCM $\M$ such that $\acc_{\M} = f$.
	\end{theorem}

	We define the class $\pe^*$ as follows. A function $f:\Sigma^*\to\N$ is in $\pe^*$ if $f\in\shp$ and for each $k\in\N$, then the language $L_{f > k} = \{w\mid f(w) > k \}$ is in $\ptime$
\end{document}

### 2016-11-30 LICS/conclusions.tex
Archivos nuevos.

%!TEX root = main.tex

Here comes the conclusions.

### 2016-12-01 LICS/beyond.tex
paper sin apendice

--- 2016-10-05 QSO/logics.tex
@@
-Consider $\bbN$ as the set of non-negative integers.
-Then given a relational signature $\R$, the set of quantitative second-order logic formulas (or just $\qso$-formulas) over $\R$ is given by the following grammar:
-\begin{eqnarray}
-\label{eq-def-qso}
-  \alpha & := & \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha 
-\end{eqnarray}
+%!TEX root = main.tex
+
+\subsection{Transitive Logics}
+
+We define the operator {\bf path} as follows.
+Let $\A$ be an ordered structure.
+Given a formula $\psi(\bar{x},\bar{y})$, where $\vert \bar{x} \vert = \vert \bar{y} \vert = k$ let ${\cal G} = ({\cal V},\cal{E})$ be induced graph over the set of vertices ${\cal V} = A^k$, and for every $\bar{a},\bar{b}\in A^k$ it holds that ${\cal E}(\bar{a},\bar{b})$ if and only if $\A \models \psi(\bar{a},\bar{b})$.
+To formalize the semantics for this operator, let $n = \vert A^k \vert$.
+For a given first order assignment $v$ and a second order asssignment $V$, let $\bar{a} = v(\bar{x})$ and $\bar{b} = v(\bar{y})$, and $\sem{[\pth\, \psi(\bar{x},\bar{y})]}(\A,v,V)$ will take the value of the number of paths of size less or equal to $n$ from $\bar{a}$ to $\bar{b}$ in the graph ${\cal G}$.
+This operator lets us define the set of transitive $\qfo$ formulas over $\R$ ($\tqfo$-formulas) using the following grammar:
+\begin{multline*} 
+%	\label{eq-def-tqfo}
+	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \\ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ [\pth \varphi]
+\end{multline*}
+where $\varphi$ is an $\fo$-formula over $\R$, $s \in \bbN$ and $x \in \fv$.
+
+We also define the set of transitive $\qso$ formulas ($\tqso$-formulas) using the following grammar:
+\begin{multline*}
+%	\label{eq-def-tqso}
+	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \\ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha \ \mid \ [\pth \varphi]
+\end{multline*}
 where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
-
-Let $\R$ be a relational signature, $\A$ be a finite $\R$-structure with domain $A$, $v$ a first-order assignment for $\A$ and $V$ a second-order assignment for $\A$.
-Then the \emph{evaluation} of a $\qso$-formula $\alpha$ over $(\A, v, V)$ is defined as a function $\sem{\alpha}$ that on input $(\A, v, V)$ returns a non-negative integer in $\bbN$.
-Formally, the function $\sem{\alpha}$ is recursively defined as follows:
-$$
-\renewcommand{\arraystretch}{1.7}
-\begin{array}{rcl} 
-\sem{\varphi}(\A, v, V) & = & 
-\begin{cases}
-1 & \mbox{if } (\A, v, V) \models \varphi \\
-0 & \mbox{otherwise}
-\end{cases}\\
-\sem{s}(\A, v, V) & = & s \\
-\sem{\alpha_1 \add \alpha_2}(\A, v, V) & = & \sem{\alpha_1}(\A, v, V) + \sem{\alpha_2}(\A, v, V)\\
-\sem{\alpha_1 \mult \alpha_2}(\A, v, V) & = & \sem{\alpha_1}(\A, v, V) \cdot \sem{\alpha_2}(\A, v, V)\\ 
-\sem{\sa{x} \alpha}(\A, v, V) & = & \displaystyle \sum_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
-\sem{\pa{x} \alpha}(\A, v, V) & = & \displaystyle \prod_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
-\sem{\sa{X} \alpha}(\A, v, V) & = & \displaystyle \sum_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])\\
-\sem{\pa{X} \alpha}(\A, v, V) & = & \displaystyle \prod_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])
-\end{array}
-$$
-A formula in $\qso$ can mention the usual quantifiers in $\so$ (that is, $\exists x$ and $\exists X$) and the quantifiers that make use of addition and multiplication (that is, $\Sigma x$, $\Pi x$, $\Sigma X$ and $\Pi X$), which are called {\em quantitative quantifiers} .
-A $\qso$-formula $\alpha$ is said to be a \emph{sentence} if it does not have any free variable, that is, every variable in $\alpha$ is under the scope of a usual quantifier or a quantitative quantifier.
-It is important to notice that if $\alpha$ is a $\qso$-sentence over a relational signature $\R$, then for every finite $\R$-structure $\A$, first-order assignments $v_1$, $v_2$ for $\A$ and second-order assignments $V_1$, $V_2$ for $\A$, it holds that:
-\begin{eqnarray*}
-\sem{\alpha}(\A, v_1, V_1) & = & \sem{\alpha}(\A, v_2, V_2).
-\end{eqnarray*}
-Thus, in such a case we use the term $\sem{\alpha}(\A)$ to denote $\sem{\alpha}(\A, v, V)$, for some arbitrary first-order assignment $v$ for $\A$ and some arbitrary second-order assignment $V$ for $\A$.
-
-In this paper, we consider several fragments of $\qso$, which are obtained by restricting the syntax of the formula $\varphi$ in \eqref{eq-def-qso} or the use of the quantitative quantifiers.
-Let $\qfo$ be the fragment of $\qso$ obtained by only allowing the quantitative quantifiers $\Sigma x$, $\Pi x$ in \eqref{eq-def-qso}.
-Let $\eqso$ be a fragment of $\qso$ defined as $\qfo$ but also allowing the quantitative quantifier $\Sigma X$.
-Moreover, assuming that $\LL$ is a fragment of $\so$, let $\qso(\LL)$ be the fragment of $\qso$ obtained by restricting $\varphi$ in \eqref{eq-def-qso} to be a formula in $\LL$, and let $\eqso(\LL)$ be the fragment of $\eqso$ obtained by imposing the same restriction.
-In particular, in this paper we consider the following fragments $\LL$: $\fo$, which is obtained by disallowing the use of the second-order quantifier ($\exists X$ or $\forall X$) in the formula $\varphi$ in \eqref{eq-def-qso}, and $\eso$, which is obtained by allowing $\varphi$ in \eqref{eq-def-qso} to be a formula of the form $\exists X_1 \cdots \exists X_k \, \psi$, where $\psi$ is an $\fo$-formula.
+\begin{theorem}
+	Given a positive $\fo$ formula $\varphi(\bar{x},R)$ and a $\qfo$ formula $\alpha(\bar{x})$, there exists a $\qso$ formula $\beta(\bar{x})$ such that $\sem{[\alfp\varphi(\bar{x},R)\mid \alpha(\bar{x},R)](\bar{x})} = \sem{\beta(\bar{x})}$.
+\end{theorem}
 
 \begin{theorem}
-	Let $\alpha$ be a $\qso$-formula in some fragment of $\qso$.
-There is a $\qso$ formula $\alpha'$ in the same fragment such that $\alpha'$ does not have any subformula of the form $(\beta_1 \cdot \beta_2)$ and $\sem{\alpha} = \sem{\alpha'}$.
+	$\tqfo(\fo)$ captures $\shl$ over the class of ordered structures.
 \end{theorem}
-\begin{proof}
-	Let $\{+, \cdot, \Sigma, \Pi\}$ be the set of {\it algebraic operators} in $\qso$.
-We will prove by induction over the number of algebraic operators in a $\qso$-formula $\alpha$ that there exists some $\qso$-formula $\alpha'$ that respects the stated condition.
-Let $\text{op}(\alpha)$ be the number of algebraic operators in $\alpha$.
-The case $\text{op}(\alpha) = 0$ is trivial.
-Let $n\in\nat$ be such that for each $\qso$-formula $\beta$ such that $\text{op}(\beta) < n$ there is some $\beta'$ that respects the condition.
-Let $\alpha$ be a $\qso$-formula such that $\text{op}(\alpha) = n$.
-Note that if $\alpha$ is of the form $\alpha = (\beta_1 + \beta_2)$, $\alpha = \Sigma x \beta$, $\alpha = \Pi x \beta$, $\alpha = \Sigma X \beta$ or $\alpha = \Pi X \beta$ the proof follows directly.
-Therefore we assume that $\alpha = (\beta_1 \cdot \beta_2)$.
-We separate the proof in two main cases:
-	\begin{enumerate}
-		\item At least one of $\beta_1, \beta_2$ is not of the form $\Pi x \gamma$, $\Pi X \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. Without loss of generality, let $\beta_1$ be such one.
-		\begin{enumerate}
-			\item $\beta_1 = s \in \nat$.
-Then we use $\alpha' = \beta_2 + \cdots + \beta_2$, where $\beta_2$ is repeated $s$ times, and if $s = 0$, take $\alpha' = \perp$.
-Since each of the formulas paired by $+$ has less than $n$ algebraic operators, the statement holds.
-			\item $\beta_1 = (\gamma_1 + \gamma_2)$.
-Then, we use $\alpha' = (\gamma_1\cdot\beta_2 + \gamma_2\cdot\beta_2)$ and each of the formulas paired by $+$ have less than $n$ algebraic operators so the statement holds.
-			\item $\beta_1 = \Sigma x \gamma$. Then, we use $\alpha' = \Sigma x (\gamma\cdot\beta_2)$ and as in the previous case the statement holds.
-			\item $\beta_1 = \Sigma X \gamma$, which we treat exactly as the previous one.
-		\end{enumerate}
-		\item Both of $\beta_1$ and $\beta_2$ are of the form $\Pi x \gamma$, $\Pi X \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula.
-This case also separates in further cases.
-We define $(\varphi \mapsto \alpha) := (\neg\varphi + (\varphi\cdot\alpha))$ for which $\sem{(\varphi \mapsto \alpha)}(\A,v,V)$ equals $\sem{\alpha}(\A,v,V)$ when $(\A,v,V)\models\varphi$ and 1 otherwise.
-		\begin{enumerate}
-			\item $\beta_1 = \Sigma x \gamma_1(x)$ and $\beta_2 = \Sigma y \gamma_2(y)$. Then we use $\alpha' = \Sigma x(\gamma_1(x)\cdot\gamma_2(x))$, and the statement holds.
-			\item $\beta_1 = \Sigma X \gamma_1(X)$ and $\beta_2 = \Sigma Y \gamma_2(Y)$. [something about arities].
-			\item $\beta_1 = \Sigma X \gamma_1(X)$ and $\beta_2 = \Sigma x \gamma_2(x)$ (w.l.o.g.). [use singleton predicates].
-			\item $\beta_1 = \Sigma x \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). Then, we use $\alpha' = \Sigma x (\gamma \cdot \varphi)$, and the statement holds.
-			\item $\beta_1 = \Sigma X \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). We treat this case exactly as the previous one.
-			\item $\beta_1 = \varphi_1$ and $\beta_2 = \varphi_2$ both $\so$-formulas. Then, we use $\alpha' = (\varphi_1 \wedge \varphi_2)$ and the statement holds.
-		\end{enumerate}
-	\end{enumerate}
-\end{proof}
+
+\begin{theorem}
+	$\tqso(\fo)$ captures $\shp$ over the class of ordered structures.
+\end{theorem}
+
+\subsection{Recursive Logics}
 
@@
 
-\martin{No estoy seguro si quitar el operador $\Pi$ de la fórmula o solo decir que la funcion $\pi$ no puede estar mencionada en ninguna instancia del operador $\Pi$.}
-
 For each $\qfo$ formulas $\alpha, \beta, \gamma$, where $\beta$ and $\gamma$ have $\ell$ and $m$ first-order free variables respectively, $u_1,\ldots,u_{\ell} \in \fv$, and $v_1,\ldots,v_m \in \{u_1,\ldots,u_{\ell}\}$, we define $\alpha\mid_{\beta(u_1,\ldots,u_{\ell})\to\gamma(v_1,\ldots,v_{m})}$ as $\alpha$ where every instance of the subformula $\beta(u_1,\ldots,u_{\ell})$ is replaced by $\gamma(v_1,\ldots,v_{m})$.
@@
 Then if $a$ is the first element, we define $\varphi_a(x) = \forall y(x < y \vee x = y)$, and if it is not we define:
-$$
-\varphi_a(x) = \exists x_1 \cdots \exists x_{p-1}[\bigwedge_{1\leq i,j < p}x_i\neq x_j \wedge \bigwedge_{i = 1}^{p-1} x_i < x \wedge \forall y(y < x \to \bigvee_{i = 1}^{p-1} y = x_i)].
-$$
+\begin{multline*}
+\varphi_a(x) = \exists x_1 \cdots \exists x_{p-1}[\bigwedge_{1\leq i,j < p}x_i\neq x_j \wedge\,\\ \bigwedge_{i = 1}^{p-1} x_i < x  \wedge \forall y(y < x \to \bigvee_{i = 1}^{p-1} y = x_i)].
+\end{multline*}
 
@@
 If $i \geq 1$, then
-$$
-\beta_i(u_1,\ldots,u_{\ell}) = \bigplus_{(a_1,\ldots,a_{\ell})\in A^{\ell}} \varphi_{a_1}(u_1)\cdot\varphi_{a_2}(u_2)\,\cdots\,\varphi_{a_{\ell}}(u_{\ell})\cdot \zeta_{i-1}(a_1,\ldots,a_{\ell}).
-$$
+\begin{multline*}
+\beta_i(u_1,\ldots,u_{\ell}) = \mathop{+}_{(a_1,\ldots,a_{\ell})\in A^{\ell}} \varphi_{a_1}(u_1)\cdot\varphi_{a_2}(u_2)\,\cdots\,\\ \varphi_{a_{\ell}}(u_{\ell})\cdot \zeta_{i-1}(a_1,\ldots,a_{\ell}).
+\end{multline*}
 Then, for each $T_i \in \T$, let $V$ be a second-order assignment for $\A$ that assigns $T_i$ to $R$, let $\zeta_i: A^{\ell}\to\nat$ be such that for each $(a_1,\ldots,a_{\ell})\in A^{\ell}$ it holds $\zeta_i(a_1,\ldots,a_{\ell}) = \sem{\alpha\mid_{\pi(u_1,\ldots,u_{\ell})\to \beta_i(u_1,\ldots,u_{\ell})}}(\A,v,V)$, where $v$ is a first-order assignment for $\A$ that satisfies $a_i = v(x_i)$ for each free $x_i$ in $\alpha$.
@@
 For a given first order asignment $v$ and a second order assignment $V$, let $a_i = v(x_i)$, the operator is then evaluated as:
-$$
-\sem{[\alfp\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)]}(\A,v,V) = \zeta_n(a_1,\ldots,a_{\ell}).
-$$
+\begin{multline*}
+\sem{[\alfp\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)]}(\A,v,V) = \\ \zeta_n(a_1,\ldots,a_{\ell}).
+\end{multline*}
 As an example, we will show a formula that counts the number of paths of size $n$ of a structure $\A$ with a binary relation $E$. First we define $\varphi(x,R)$:
-$$
+\begin{multline*}
 \varphi(x,R) = \forall y(x < y \vee x = y) \vee \exists z(R(z) \wedge \varphi_{succ}(z,x)),
-$$
+\end{multline*}
 where $\varphi_{succ}(x,y)$ is a formula that is satisfied by pairs $(x,y)$ that are consecutive in the order $<$.
@@
 This operator lets us define the set of recursive $\qfo$ formulas over $\R$ ($\rqfo$-formulas) using the following grammar:
-\begin{eqnarray*}
+\begin{multline*}
 	\label{eq-def-rqfo}
-\alpha & := & \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ [\alfp \varphi \mid \alpha]
-\end{eqnarray*}
+	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid \\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ [\alfp \varphi \mid \alpha]
+\end{multline*}
 where $\varphi$ is an $\fo$-formula over $\R$, $s \in \bbN$ and $x \in \fv$.
 
-We also define the less powerful operator {\bf path} as follows.
-Let $\A$ be an ordered structure.
-Given a formula $\psi(\bar{x},\bar{y})$, where $\vert \bar{x} \vert = \vert \bar{y} \vert = k$ let ${\cal G} = ({\cal V},\cal{E})$ be induced graph over the set of vertices ${\cal V} = A^k$, and for every $\bar{a},\bar{b}\in A^k$ it holds that ${\cal E}(\bar{a},\bar{b})$ if and only if $\A \models \psi(\bar{a},\bar{b})$.
-To formalize the semantics for this operator, let $n = \vert A^k \vert$.
-For a given first order assignment $v$ and a second order asssignment $V$, let $\bar{a} = v(\bar{x})$ and $\bar{b} = v(\bar{y})$, and $\sem{[\pth\, \psi(\bar{x},\bar{y})]}(\A,v,V)$ will take the value of the number of paths of size less or equal to $n$ from $\bar{a}$ to $\bar{b}$ in the graph ${\cal G}$.
-This operator lets us define the set of transitive $\qfo$ formulas over $\R$ ($\tqfo$-formulas) using the following grammar:
-\begin{eqnarray*}
-	\label{eq-def-tqfo}	\alpha & := & \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ [\pth \varphi]
-\end{eqnarray*}
-where $\varphi$ is an $\fo$-formula over $\R$, $s \in \bbN$ and $x \in \fv$.
-
-We also define the set of transitive $\qso$ formulas ($\tqso$-formulas) using the following grammar:
-\begin{eqnarray*}
-	\label{eq-def-tqso}	\alpha & := & \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha \ \mid \ [\pth \varphi]
-\end{eqnarray*}
-where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
 \begin{theorem}
-	Given a positive $\fo$ formula $\varphi(\bar{x},R)$ and a $\qfo$ formula $\alpha(\bar{x})$, there exists a $\qso$ formula $\beta(\bar{x})$ such that $\sem{[\alfp\varphi(\bar{x},R)\mid \alpha(\bar{x},R)](\bar{x})} = \sem{\beta(\bar{x})}$.
+	$\rqfo(\fo)$ captures $\fp$ over the class of ordered structures.
 \end{theorem}
-\begin{proof}
-
-\end{proof}

### 2016-12-01 LICS/counting.tex
paper sin apendice

--- 2016-10-19 QSO/cap-comp.tex
@@
+%!TEX root = main.tex
+
 Assume that $\alpha$ is a sentence in $\qso$. Then slightly abusing notation, we also use $\alpha$ to denote a function from $\R$ to $\bbN$ such that for every $\A \in \str[\R] $:
@@
 
-\cristian{Aca hay que definir esto sobre estructuras finitas directamente. No veo el sentido de tener estructuras infinitas.}
-
 \begin{theorem} \label{captfp}
@@
 \end{theorem}
-\begin{proof}
-	For the first condition, let $\alpha\in\qfo(\lfp)$, $\A\in\ostr$, $v$ and $V$ a first and second order assignment for $\A$, respectively.
-To evaluate $\sem{\alpha}(\A,v,V)$, we replace each first order sum and first order product by their corresponding expansion.
-This is, $\Sigma x \beta(x)$ is replaced by $(\beta(a_1)+\cdots+\beta(a_n))$, where $A = \{a_1,\ldots,a_n\}$, and $\Pi x \beta(x)$ is replaced by $(\beta(a_1)\cdot\,\cdots\,\cdot\beta(a_n))$.
-Then we replace each logic sub-formula $\varphi$ in $\alpha$ by their evaluated value, 0 or 1.
-Since each of this formulas is in $\lfp$, this can be done in polynomial time.
-The resulting formula is an arithmetic expression, which can be evaluated recursively in polynomial time.
-
-	For the second condition, let $f\in \fp$.
-Let $k\in\nat$ be such that $\vert f(\A) \vert = \vert A \vert^k$ for each $\A\in\ostr[\R]$.
-Let $\Phi(x_0,\ldots,x_{k-1})$ be a $\lfp$ formula such that $(\A,a_0,\ldots,a_{k-1})\models\Phi$ if and only if in the string $y = f(\A)$ there is 1 in the $(\vert A \vert^{k-1}m_{k-1} + \cdots + \vert A \vert^2 m_2 + \vert A \vert m_1 + m_0)$-th position from right to left, and where $a_i$ is the (0-indexed) $m_i$-th element of $A$.
-We use
-	$$
-	\alpha = \Sigma x_0 \cdots \Sigma x_{k-1} \Phi(x_0,\ldots,x_{k-1})\cdot\varphi_{k-1}(x_{k-1})\cdot\,\cdots\,\cdot\varphi_0(x_0).
-	$$
-	where $\varphi_i(x) = \Pi y[(y < x)\mapsto\Pi z_1\cdots\Pi z_i\,2]$ for $i > 0$ and $\varphi_0(x) = \Pi y[(y < x)\mapsto 2]$.
-Note that if $a$ is the  $m$-th element in $A$, then $\sem{\varphi_i(x)}(\A,a) = 2^{\vert A \vert^i m}$.
-Therefore, for each $(a_0,\ldots,a_{k-1})\in A^k$, we have that $\sem{\Phi(x_0,\ldots,x_{k-1})\cdot\varphi_{k-1}(x_{k-1})\,\cdots\,\varphi_0(x_0)}(\A,a_0,\ldots,a_{k-1}) = 2^{(\vert A \vert^{k-1}m_{k-1} + \cdots + \vert A \vert^2 m_2 + \vert A \vert m_1 + m_0)}$ if $(\A,a_0,\ldots,a_{k-1})\models\Phi$ and 0 otherwise, and adding these values gives $f(\A)$.
-\end{proof}
 
@@
 \end{theorem}
-\begin{proof}
-	For the first condition, we use the exact same procedure as in Theorem \ref{captfp}, noting that evaluating each of the $\pfp$ formulas can be done in polynomial space, and that the result is also polynomial on the size of the input.
-
-	For the second condition, note that for each function $f\in \nfpspace$, $\vert f(\A) \vert$ is polynomial on $\vert A \vert$.
-Thus, there is a $\pfp$ formula $\Phi$ that models the string $f(\A)$ by the bit.
-The rest of the proof is analogous to Theorem \ref{captfp}.
-\end{proof}
-
-\begin{theorem}
-	$\qfo(\tc)$ captures $\shl$ over the class of ordered structures.
-\end{theorem}
-\begin{proof}
-	For the first condition, let $\alpha\in\qfo(\tc)$, let $\A\in\ostr$ $v$ and $V$ first and second order assignments of $\A$ respectively.
-We evaluate $\sem{\alpha}(\A,v,V)$ recursively.
-A logic formula $\varphi$ is evaluated...?
-
-	For the second condition, note that for each function $f\in \shl$, $\vert f(\A) \vert$ is polynomial on $\vert A \vert$.
-Thus, there is a $\tc$ formula $\Phi$ that models the string $f(\A)$ by the bit.
-The rest of the proof is analogous to Theorem \ref{captfp}.
-\end{proof}
-
-\begin{theorem}
-	$\rqfo(\fo)$ captures $\fp$ over the class of ordered structures.
-\end{theorem}
-
-\begin{theorem}
-	$\tqfo(\fo)$ captures $\shl$ over the class of ordered structures.
-\end{theorem}
-\begin{proof}
-	For the second condition, let $f \in \shl$.
-We will address the case where $\R$ contains only one binary predicate $E$, and the rest of the cases can be deduced from this.
-Let $M$ be a non-deterministic logspace machine such that $f(\A) = \acc_M(\A)$ for each $\A \in \ostr[R]$.
-Suppose ${\cal Q} = \{q_1,\ldots,q_{\ell}\}$ is the set of states of $M$, where $q_1$ is the initial state, and $q_{\ell}$ is only final state of $M$.
-Let $n = \vert A \vert$ and let $w = \enc(\A) \in \{0,1\}^{n^2}$.
-We assume that $M$ with input $w$ uses space $s_M(w) < c\cdot\log(n)$ and furthermore, $s_M(w) < n-2$.
-We notate $M(w)$ as the graph of configurations of $M$ running on input $w$.
-
-	We represent configurations with a tuple of fixed size $k$.
-The formula $\varphi(\bar{x},\bar{y})$ describes a procedure that given a configuration generates a possible next configuration.
-The formula $\varphi_I(\bar{x})$ models that $\bar{x}$ is the initial configuration of $M(w)$.
-The formula $\varphi_F(\bar{x})$ models that $\bar{x}$ is an accepting (final) configuration of $M(w)$.
-The formula we construct is:
-	$$
-	\alpha = \sa{\bar{x}}\sa{\bar{y}}([\pth \varphi(\bar{u},\bar{v})](\bar{x},\bar{y})\cdot \varphi_I(\bar{x})\cdot\varphi_F(\bar{y})).
-	$$
-
-	To illustrate our idea, we will show a greatly simplified example.
-Consider a machine $M$ that works in exactly $\log_2(n)$ space and only allows 0 or 1 in the working tape.
-Consider an input $\A$ of size 16 (that is, $A = \{0,\ldots,9,A,\ldots,F\}$).
-Let some configuration $s$ have 0011 in the working tape, the head in the input tape is in position 26, and the head in the input tape is in position 2 (we consider 0-indexed positions).
-Also, $Q = \{q_1,\ldots,q_5\}$ and the current state is $q_3$.
-
-	As a first approach, we will use a 9-tuple $\bar{a} = (a_1,\ldots,a_9)$ to represent $s$.
-That is, $(a_1,a_2) = (1,A)$ represent the position of the head in the input tape, $a_3 = 2$ represents the position of the head in the working tape, $a_4 = C$ (1100b in base 16) represents the content of the working tape, and $(a_5,\ldots,a_9) = (0,0,1,0,0)$ represents the current state.
-Then $\bar{a} = (1,A,2,C,0,0,1,0,0)$ will represent $s$.
-It is not possible \martin{no se si ser\'a cierto esto} to find the $i$-th bit of an element $x$, or much less switch the $i$-th bit of an element, solely with $\fo$.
-So consider the following procedure.
-
-	\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 1 replace it by 0}
-	\begin{algorithmic}
-		\State $b \gets bit(x,i)$
-		\State $\textbf{assert } b = 1$
-		\State $y = pow2(i)$
-		\State $z = substract(x,y)$
-		\State \Return $z$.
-	\end{algorithmic}
-	\end{algorithm}	
-	And the details for $bit$, $pow2$ and $substract$ are left for the reader as an exercise.
-
-	Note that there is an upper bound of $(n^2+2)\cdot (c\log(n)+2)\cdot 3^{c\log n} \cdot \ell$ distinct configurations in $M(w)$.
-Given a configuration $s\in M(w)$, we construct a tuple $\bar{a} = (a_0,\ldots,a_{k-1})$ that represents $s$.
-	\begin{itemize}
-	\item $(a_0,a_1,a_2)$ represent the position of the head in the reading tape.
-If $a_0 = 1$, then the head is in position $a_1a_2$ (the $a_1a_2$-th character in the input).
-If $a_1 = 2$, then $a_2 = 1$ represents that the head is at the beginning of the tape (position 0), and if $a_2 = 2$, then the head is at position $n+1$.
-	\item $a_3$ represents the position of the head in the working tape.
-	\item $(a_4,\ldots,a_{3+2c})$ represents the string that is in the working tape.
-Specifically, let $h$ be the value of $a_{3+2c}a_{2+2c}\cdots a_{5}a_{4}$ in base $n$.
-Then the string in the working tape is $h$ in base 3, where 2 is interpreted as $\texttt{B}$.
-	\item $(a_{4+2c},\ldots,a_{3+2c+\ell})$ represents the current state $q_i$. Every value in the tuple is $1$ except the element $a_{3+2c+i}$ which is 2.	
-	\end{itemize}
-
-	Now we describe the formulas.
-	\begin{align*}
-	\varphi_I(\bar{x}) &= \varphi_v(\bar{x}) \wedge \exists u(\varphi_1(u) \wedge \varphi_{succ}(u,x_{3+e+2c})), \\
-	\varphi_F(\bar{x}) &= \varphi_v(\bar{x}) \wedge \exists u(\varphi_1(u) \wedge \varphi_{succ}(u,x_{2+e+2c+\ell})),
-	\end{align*}
-	where $\varphi_v(\bar{x})$ describes that $\bar{x}$ is a valid configuration of $M(w)$ and $\varphi_1(x)$ describes that $x$ is the first element of $A$.
-	\begin{align*}
-	\varphi_v^1(\bar{x}) &= \exists u(\varphi_1(u) \wedge (x_1 = u \vee \varphi_{succ}(u,x_1)) \wedge \varphi_{succ}(u,x_1) \to (x_2 = u \vee \varphi_{succ}(u,x_2))), \\
-	\varphi_v^4(\bar{x}) &= \exists u(\varphi_1(u) \wedge \bigvee_{i = 1}^{\ell}(\varphi_{succ}(u,x_{2+e+2c+i})  \wedge \bigwedge\limits_{\substack{i = 1\\ j \neq i}}^{\ell} x_{2+e+2c+j} = u)).
-	\end{align*}
-\end{proof}
 
@@
 
-%\begin{theorem}
-%$\eqso(\integ,\fo)$ captures $\gp$ over the class of ordered structures.
-%\end{theorem}
+\begin{theorem}
+	$\eqso(\integ,\fo)$ captures $\gp$ over the class of ordered structures.
+\end{theorem}
 

### 2016-12-01 LICS/framework.tex
paper sin apendice

--- 2016-09-20 QSO/logics.tex
@@
+%!TEX root = main.tex
+
 Consider $\bbN$ as the set of non-negative integers.
 Then given a relational signature $\R$, the set of quantitative second-order logic formulas (or just $\qso$-formulas) over $\R$ is given by the following grammar:
-\begin{eqnarray}
+\begin{multline*}
 \label{eq-def-qso}
-  \alpha & := & \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha 
-\end{eqnarray}
+\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \\ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha 
+\end{multline*}
 where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
@@
 In particular, in this paper we consider the following fragments $\LL$: $\fo$, which is obtained by disallowing the use of the second-order quantifier ($\exists X$ or $\forall X$) in the formula $\varphi$ in \eqref{eq-def-qso}, and $\eso$, which is obtained by allowing $\varphi$ in \eqref{eq-def-qso} to be a formula of the form $\exists X_1 \cdots \exists X_k \, \psi$, where $\psi$ is an $\fo$-formula.
-
-\begin{theorem}
-	Let $\alpha$ be a $\qso$-formula in some fragment of $\qso$.
-There is a $\qso$ formula $\alpha'$ in the same fragment such that $\alpha'$ does not have any subformula of the form $(\beta_1 \cdot \beta_2)$ and $\sem{\alpha} = \sem{\alpha'}$.
-\end{theorem}
-\begin{proof}
-	Let $\{+, \cdot, \Sigma, \Pi\}$ be the set of {\it algebraic operators} in $\qso$.
-We will prove by induction over the number of algebraic operators in a $\qso$-formula $\alpha$ that there exists some $\qso$-formula $\alpha'$ that respects the stated condition.
-Let $\text{op}(\alpha)$ be the number of algebraic operators in $\alpha$.
-The case $\text{op}(\alpha) = 0$ is trivial.
-Let $n\in\nat$ be such that for each $\qso$-formula $\beta$ such that $\text{op}(\beta) < n$ there is some $\beta'$ that respects the condition.
-Let $\alpha$ be a $\qso$-formula such that $\text{op}(\alpha) = n$.
-Note that if $\alpha$ is of the form $\alpha = (\beta_1 + \beta_2)$, $\alpha = \Sigma x \beta$, $\alpha = \Pi x \beta$, $\alpha = \Sigma X \beta$ or $\alpha = \Pi X \beta$ the proof follows directly.
-Therefore we assume that $\alpha = (\beta_1 \cdot \beta_2)$.
-We separate the proof in two main cases:
-	\begin{enumerate}
-		\item At least one of $\beta_1, \beta_2$ is not of the form $\Pi x \gamma$, $\Pi X \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. Without loss of generality, let $\beta_1$ be such one.
-		\begin{enumerate}
-			\item $\beta_1 = s \in \nat$.
-Then we use $\alpha' = \beta_2 + \cdots + \beta_2$, where $\beta_2$ is repeated $s$ times, and if $s = 0$, take $\alpha' = \perp$.
-Since each of the formulas paired by $+$ has less than $n$ algebraic operators, the statement holds.
-			\item $\beta_1 = (\gamma_1 + \gamma_2)$.
-Then, we use $\alpha' = (\gamma_1\cdot\beta_2 + \gamma_2\cdot\beta_2)$ and each of the formulas paired by $+$ have less than $n$ algebraic operators so the statement holds.
-			\item $\beta_1 = \Sigma x \gamma$. Then, we use $\alpha' = \Sigma x (\gamma\cdot\beta_2)$ and as in the previous case the statement holds.
-			\item $\beta_1 = \Sigma X \gamma$, which we treat exactly as the previous one.
-		\end{enumerate}
-		\item Both of $\beta_1$ and $\beta_2$ are of the form $\Pi x \gamma$, $\Pi X \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula.
-This case also separates in further cases.
-We define $(\varphi \mapsto \alpha) := (\neg\varphi + (\varphi\cdot\alpha))$ for which $\sem{(\varphi \mapsto \alpha)}(\A,v,V)$ equals $\sem{\alpha}(\A,v,V)$ when $(\A,v,V)\models\varphi$ and 1 otherwise.
-		\begin{enumerate}
-			\item $\beta_1 = \Sigma x \gamma_1(x)$ and $\beta_2 = \Sigma y \gamma_2(y)$. Then we use $\alpha' = \Sigma x(\gamma_1(x)\cdot\gamma_2(x))$, and the statement holds.
-			\item $\beta_1 = \Sigma X \gamma_1(X)$ and $\beta_2 = \Sigma Y \gamma_2(Y)$. [something about arities].
-			\item $\beta_1 = \Sigma X \gamma_1(X)$ and $\beta_2 = \Sigma x \gamma_2(x)$ (w.l.o.g.). [use singleton predicates].
-			\item $\beta_1 = \Sigma x \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). Then, we use $\alpha' = \Sigma x (\gamma \cdot \varphi)$, and the statement holds.
-			\item $\beta_1 = \Sigma X \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). We treat this case exactly as the previous one.
-			\item $\beta_1 = \varphi_1$ and $\beta_2 = \varphi_2$ both $\so$-formulas. Then, we use $\alpha' = (\varphi_1 \wedge \varphi_2)$ and the statement holds.
-		\end{enumerate}
-	\end{enumerate}
-\end{proof}
-
-We define an operator which extends the Least Fixed Point logic to counting.
-Recall that a fixed point operator is defined by a formula $\varphi(x_1,\ldots,x_k,R)$ which is positive on $R$ where $R$ is a predicate of arity $k$.
-For a structure $\A$ with domain $A$, the operator $T_{\varphi}:2^{A^k} \to 2^{A^k}$ is defined as $T_{\varphi}(X) = \{(a_1,\ldots,a_k)\mid (\A,X)\models \varphi(a_1,\ldots,a_k,R) \}$, for each $X\subseteq A^k$.
-Let $T_0 = \emptyset$ and $T_{i+1} = T_{\varphi}(T_i)$ for each $i \in \nat$.
-Note that there exists $n\in \nat$ such that $T_{n+1} = T_n$.
-Semantically, $[\lfpop\,\varphi(x_1,\ldots,x_k,R)]$ is defined such that for each $(a_1,\ldots,a_k)\in A^k$, it holds that $\A\models[\lfpop\,\varphi(x_1,\ldots,x_k,R)](a_1,\ldots,a_k)$ if and only if $\A\models T_n(a_1,\ldots,a_k)$.
-
-We extend $\qfo$ with the operator $[\textbf{alfp}\,\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$.
-It is defined by a FO-formula $\varphi$ and a QFO-formula $\alpha$ which mentions a placeholder function $\pi$ with domain $\fv^{\ell}$.
-Let $\{T_i\}_{i\in\nat}$ and $n\in\nat$ be defined from $\varphi$ as it was for the $\lfp$ operator.
-
-For each $\qfo$ formulas $\alpha, \beta, \gamma$, where $\beta$ and $\gamma$ have $\ell$ and $m$ first-order open variables respectively, $u_1,\ldots,u_{\ell} \in \fv$, and $v_1,\ldots,v_m \in \{u_1,\ldots,u_{\ell}\}$, we define $\alpha\mid_{\beta(u_1,\ldots,u_{\ell})\to\gamma(v_1,\ldots,v_{m})}$ as $\alpha$ where every instance of the subformula $\beta(u_1,\ldots,u_{\ell})$ is replaced by $\gamma(v_1,\ldots,v_{m})$.
-Moreover, for each $a\in A$, suppose that $a$ is the $p$-th element in the order $<^{\A}$.
-Then if $a$ is the first element, we define $\varphi_a(x) = \forall y(x < y \vee x = y)$, and if it is not we define:
-$$
-\varphi_a(x) = \exists x_1 \cdots \exists x_{p-1}[\bigwedge_{1\leq i,j < p}x_i\neq x_j \wedge \bigwedge_{i = 1}^p x_i < x \wedge \forall y(y < x \to \bigvee_{i = 1}^p y = x_i)].
-$$
-
-To formally characterize the semantics of $[\textbf{alfp}\,\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$. %. 
-We name $\T = \{T_0,\ldots,T_n\}$ and we define the operator $Z:\T \to \{f\mid f:A^{\ell}\to\nat\}$ as follows.
-For each $T_i \in \T$, we define a $\qfo$ formula $\beta_i$ as follows.
-If $i = 0$, then $\beta_i(u_1,\ldots,u_{\ell}) = 0$.
-If $i \geq 1$, then let $f = Z[T_{i-1}]$ and
-$$
-\beta_i(u_1,\ldots,u_{\ell}) = \bigplus_{(a_1,\ldots,a_{\ell})\in A^{\ell}} \varphi_{a_1}(u_1)\cdot\varphi_{a_2}(u_2)\,\cdots\,\varphi_{a_{\ell}}(u_{\ell})\cdot f(a_1,\ldots,a_{\ell}).
-$$
-Then, for each $T_i \in \T$, let $V$ be a second-order assignment for $\A$ that assigns $T_i$ to $R$, let $f: A^{\ell}\to\nat$ be such that for each $(a_1,\ldots,a_{\ell})\in A^{\ell}$ it holds $f(a_1,\ldots,a_{\ell}) = \sem{\alpha\mid_{\pi(u_1,\ldots,u_{\ell})\to \beta_i(u_1,\ldots,u_{\ell})}}(\A,v,V)$, where $v$ is a first-order assignment for $\A$ that satisfies $a_i = v(x_i)$ for each open $x_i$ in $\alpha$.
-We define $Z[T_i] = f$.
-
-For a given first order asignment $v$, let $a_i = v(x_i)$, let $f = Z[T_n]$, and the operator is evaluated as:
-$$
-[\textbf{alfp}\,\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](\A,v) = f(a_1,\ldots,a_{\ell}).
-$$
-As an example, we will show a formula that counts the number of paths of size $n$ of a structure $\A$ with a binary relation $E$. First we define $\varphi(x,R)$:
-$$
-\varphi(x,R) = \forall y(x < y \vee x = y) \vee \exists z(R(z) \wedge \varphi_{succ}(z,x)),x
-$$
-where $\varphi_{succ}(x,y)$ is a formula that is satisfied by pairs $(x,y)$ that are consecutive in the order $<$.
-That is, $\varphi_{succ}(x,y) = x < y \wedge \forall z((x < z \wedge z < y) \to (x = z \vee z = y) )$.
-Now we define $\alpha(x,y,R,\pi)$ as:
-$$
-\alpha(x,y,R,\pi) = \neg \exists z(R(z)) + \Sigma z[E(x,z)\cdot \pi(z,y)].
-$$
-Then, the formula $[\alfp_{\varphi(x,R)} \alpha(x,y,R,\pi)](x,y)$ when evaluated on $(\A,a,b)$ will count the number of paths of size $n$ from $a$ to $b$.
-
-We also define the less powerful operator $\pth$ as follows.
-Given a positive formula $\varphi(x,R)$ and a formula $\psi(x,y)$ we define $[\pth_{\varphi(x,R)} \psi(x,y)](x,y)$ as a formula that when evaluated on $(\A, a, b)$ induces a graph over $\A$ with the binary relation determined by $\psi(x,y)$ and counts the number of paths from $a$ to $b$ of a size determined by the $\varphi(x,R)$.
-Formally, we define:
-$$
-\sem{[\pth_{\varphi(x,R)} \psi(x,y)](x,y)} = \sem{[\alfp_{\varphi(x,R)} \alpha_{\psi}(x,y,R,\pi)](x,y)},
-$$
-where $\alpha_{\psi}(x,y,R,\pi) = \neg \exists z(R(z)) + \Sigma z[\psi(x,z)\cdot \pi(z,y)]$.
-
-\begin{theorem}
-	Given a positive $\fo$ formula $\varphi(\bar{x},R)$ and a $\qfo$ formula $\alpha(\bar{x})$, there exists a $\qso$ formula $\beta(\bar{x})$ such that $\sem{[\alfp_{\varphi(\bar{x},R)} \alpha_{\psi}(\bar{x},R)](\bar{x})} = \sem{\beta(\bar{x})}$.
-\end{theorem}
-\begin{proof}
-
-\end{proof}

### 2016-12-01 LICS/preliminaries.tex
paper sin apendice

--- 2016-12-01 LICS/framework.tex
@@
 
-Consider $\bbN$ as the set of non-negative integers.
-Then given a relational signature $\R$, the set of quantitative second-order logic formulas (or just $\qso$-formulas) over $\R$ is given by the following grammar:
-\begin{multline*}
-\label{eq-def-qso}
-\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \\ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha 
-\end{multline*}
-where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
-
-Let $\R$ be a relational signature, $\A$ be a finite $\R$-structure with domain $A$, $v$ a first-order assignment for $\A$ and $V$ a second-order assignment for $\A$.
-Then the \emph{evaluation} of a $\qso$-formula $\alpha$ over $(\A, v, V)$ is defined as a function $\sem{\alpha}$ that on input $(\A, v, V)$ returns a non-negative integer in $\bbN$.
-Formally, the function $\sem{\alpha}$ is recursively defined as follows:
-$$
-\renewcommand{\arraystretch}{1.7}
-\begin{array}{rcl} 
-\sem{\varphi}(\A, v, V) & = & 
-\begin{cases}
-1 & \mbox{if } (\A, v, V) \models \varphi \\
-0 & \mbox{otherwise}
-\end{cases}\\
-\sem{s}(\A, v, V) & = & s \\
-\sem{\alpha_1 \add \alpha_2}(\A, v, V) & = & \sem{\alpha_1}(\A, v, V) + \sem{\alpha_2}(\A, v, V)\\
-\sem{\alpha_1 \mult \alpha_2}(\A, v, V) & = & \sem{\alpha_1}(\A, v, V) \cdot \sem{\alpha_2}(\A, v, V)\\ 
-\sem{\sa{x} \alpha}(\A, v, V) & = & \displaystyle \sum_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
-\sem{\pa{x} \alpha}(\A, v, V) & = & \displaystyle \prod_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
-\sem{\sa{X} \alpha}(\A, v, V) & = & \displaystyle \sum_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])\\
-\sem{\pa{X} \alpha}(\A, v, V) & = & \displaystyle \prod_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])
-\end{array}
-$$
-A formula in $\qso$ can mention the usual quantifiers in $\so$ (that is, $\exists x$ and $\exists X$) and the quantifiers that make use of addition and multiplication (that is, $\Sigma x$, $\Pi x$, $\Sigma X$ and $\Pi X$), which are called {\em quantitative quantifiers} .
-A $\qso$-formula $\alpha$ is said to be a \emph{sentence} if it does not have any free variable, that is, every variable in $\alpha$ is under the scope of a usual quantifier or a quantitative quantifier.
-It is important to notice that if $\alpha$ is a $\qso$-sentence over a relational signature $\R$, then for every finite $\R$-structure $\A$, first-order assignments $v_1$, $v_2$ for $\A$ and second-order assignments $V_1$, $V_2$ for $\A$, it holds that:
-\begin{eqnarray*}
-	\sem{\alpha}(\A, v_1, V_1) & = & \sem{\alpha}(\A, v_2, V_2).
-\end{eqnarray*}
-Thus, in such a case we use the term $\sem{\alpha}(\A)$ to denote $\sem{\alpha}(\A, v, V)$, for some arbitrary first-order assignment $v$ for $\A$ and some arbitrary second-order assignment $V$ for $\A$.
-
-In this paper, we consider several fragments of $\qso$, which are obtained by restricting the syntax of the formula $\varphi$ in \eqref{eq-def-qso} or the use of the quantitative quantifiers.
-Let $\qfo$ be the fragment of $\qso$ obtained by only allowing the quantitative quantifiers $\Sigma x$, $\Pi x$ in \eqref{eq-def-qso}.
-Let $\eqso$ be a fragment of $\qso$ defined as $\qfo$ but also allowing the quantitative quantifier $\Sigma X$.
-Moreover, assuming that $\LL$ is a fragment of $\so$, let $\qso(\LL)$ be the fragment of $\qso$ obtained by restricting $\varphi$ in \eqref{eq-def-qso} to be a formula in $\LL$, and let $\eqso(\LL)$ be the fragment of $\eqso$ obtained by imposing the same restriction.
-In particular, in this paper we consider the following fragments $\LL$: $\fo$, which is obtained by disallowing the use of the second-order quantifier ($\exists X$ or $\forall X$) in the formula $\varphi$ in \eqref{eq-def-qso}, and $\eso$, which is obtained by allowing $\varphi$ in \eqref{eq-def-qso} to be a formula of the form $\exists X_1 \cdots \exists X_k \, \psi$, where $\psi$ is an $\fo$-formula.
+Here come the preliminaries

### 2016-12-01 LICS/syntactic.tex
paper sin apendice

%!TEX root = main.tex

\subsection{Saluja Classes}

We define syntactically the classes $\XE{i}$ and $\XU{i}$ according to the following grammar:
\begin{enumerate}
	\item $\XE{0}$:
	\[
	E_0 ::= \varphi , \varphi \mbox{ is an {\sc FO}-extended quantifier-free $\L$-formula }.
	\]
	\item $\XU{0}$:
	\[U_0 ::= E_0.\]
	\item $\XE{i+1}$:
	\[	E_{i+1} ::= U_i \ \mid \ \exists x \, E_{i+1}.\]
	\item $\XU{i+1}$:
	\[U_{i+1} ::= E_i \ \mid \ \forall x \, U_{i+1}.\]
\end{enumerate}

A function $f$ is in $\XE{i}$ (resp. $\XU{i}$) if there is an $\L$-formula $\varphi$ defined by the grammar $E_i$ (resp. $U_i$) such that $f = f_{\varphi}$.

\begin{theorem} \label{fp1}
	$\XE{0} \subseteq$ {\sc FP}.
\end{theorem}

The {\em decision problem} associated to a function $f$ is defined by the language $L_f = \{\A \in \str \mid f(\A) > 0\}$.

\begin{theorem} \label{decisionptime}
	The decision problem associated to a function in $\E1$ is in \textsc{P}.
\end{theorem}

For a given pair of functions $f,g$, we define $f \dotminus g$ as follows:
\begin{eqnarray*}
	(f \dotminus g)(\A) =
	\begin{cases}
		f(\A)-g(\A), & \text{if }f(\A)>g(\A) \\
		0, & \text{if }f(\A) \leq g(\A).
	\end{cases}
\end{eqnarray*}
for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction} if for every pair of functions $f,g \in \F$, it holds that $f \dotminus g \in \F$.

\begin{theorem}
	If $\E1$ is closed under substraction, then {\sc P} = {\sc NP}.
\end{theorem}

\begin{theorem}
	$\#\Sigma_1 \subsetneq \E1$
\end{theorem}

For a given function $f$, we define $f \dotminus 1$ as follows:
\begin{eqnarray*}
	f \dotminus 1(\A) =
	\begin{cases}
		f(\A)-1, & \text{if }f(\A) > 0 \\
		0, & \text{if }f(\A) = 0.
	\end{cases}
\end{eqnarray*}
for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.

\begin{theorem}
	$\XE{1}$ is closed under substraction by one.
\end{theorem}

\begin{theorem}
	{\sc \#DNF} is hard for $\E1$ under parsimonious reductions. 
\end{theorem}

\begin{theorem}
	$\U{1}$ with $n$ open first-order variables is properly contained in $\U{1}$ with $n+1$ open first-order variables for $n\in\N$.  
\end{theorem}

\subsection{Horn Counting Classes}

We define the syntactic classes $\XE{i}$ and $\XU{i}$ as follows. First we define extended Horn clauses.
\begin{eqnarray*}
	PL &::=& X_i(\x),\, i\in\N,\\
	NL &::=& \neg X_i(\x),\, i\in\N \ \mid \ \exists x\, NL,\\
	NC &::=& NL \ \mid \ \alpha, \alpha \mbox{ is an {\sc FO}-formula over } \L \ \mid \ (NC \vee NC),\\
	HC &::=& NC \ \mid \ (NC \vee PL) \ \mid \ PL,
\end{eqnarray*}
where $PL$ represents a positive literal, $NL$ is an \textit{extended} negative literal, $NC$ is an extended Horn clause, and $HC$ is an extended Horn formula. Now we define the syntax of the classes inductively.\textbf{}
\begin{enumerate}
	\item $\HE{0}$:
	\begin{eqnarray*}
		E_0 &::=& HC \ \mid \ E_0 \wedge E_0.
	\end{eqnarray*}
	\item $\HU{0}$:
	\begin{eqnarray*}
		U_0 &::=& E_0.
	\end{eqnarray*}
	\item $\HE{i+1}$:
	\begin{eqnarray*}
		E_{i+1} &::=& U_i \ \mid \ \exists x \, E_{i+1}.
	\end{eqnarray*}
	\item $\HU{i+1}$:
	\begin{eqnarray*}
		U_{i+1} &::=& E_i \ \mid \ \forall x \, U_{i+1}.
	\end{eqnarray*}
\end{enumerate}
A function $f$ is in $\XE{i}$ (resp. $\XU{i}$) if there is an $\L$-formula $\varphi$ defined by the grammar $E_i$ (resp. $U_i$) such that $f = f_{\varphi}$.

The class $\pe$ is defined in \cite{DBLP:conf/mfcs/PagourtzisZ06} as $\pe = \{f\mid f\in\#P \mbox{ and its decision version } L_{f}\in P\}$.

\begin{theorem}
	$\HE{2} \subseteq \pe$
\end{theorem}

We define the decision problem
\begin{multline*}
\shdhsat = \{\Phi \mid \Phi \text{ is a disjunction of Horn} \\ \text{formulas and $\Phi$ is satisfiable}\},
\end{multline*}
and the counting problem {\shdhsat } as a function that counts all satisfying assignments to a formula $\Phi$ which is a disjunction of Horn formulas.

\begin{theorem} \label{sigma2hard}
	$\shdhsat$ is hard for $\XE{2}$ under parsimonious reductions. 
\end{theorem}

### 2016-12-02 LICS/appendix.tex
paper v.1

%!TEX root = main.tex

\textbf{Proof of Theorem \ref{captfp}}

For the first condition, let $\alpha\in\qfo(\lfp)$, $\A\in\ostr$, $v$ and $V$ a first and second order assignment for $\A$, respectively. To evaluate $\sem{\alpha}(\A,v,V)$, we replace each first order sum and first order product by their corresponding expansion. This is, $\Sigma x \beta(x)$ is replaced by $(\beta(a_1)+\cdots+\beta(a_n))$, where $A = \{a_1,\ldots,a_n\}$, and $\Pi x \beta(x)$ is replaced by $(\beta(a_1)\cdot\,\cdots\,\cdot\beta(a_n))$. Then we replace each logic sub-formula $\varphi$ in $\alpha$ by their evaluated value, 0 or 1. Since each of this formulas is in $\lfp$, this can be done in polynomial time. The resulting formula is an arithmetic expression, which can be evaluated recursively in polynomial time.

For the second condition, let $f\in \fp$. Let $k\in\nat$ be such that $\vert f(\A) \vert = \vert A \vert^k$ for each $\A\in\ostr[\R]$. Let $\Phi(x_0,\ldots,x_{k-1})$ be a $\lfp$ formula such that $(\A,a_0,\ldots,a_{k-1})\models\Phi$ if and only if in the string $y = f(\A)$ there is 1 in the $(\vert A \vert^{k-1}m_{k-1} + \cdots + \vert A \vert^2 m_2 + \vert A \vert m_1 + m_0)$-th position from right to left, and where $a_i$ is the (0-indexed) $m_i$-th element of $A$. We use
	$$
	\alpha = \Sigma x_0 \cdots \Sigma x_{k-1} \Phi(x_0,\ldots,x_{k-1})\cdot\varphi_{k-1}(x_{k-1})\cdot\,\cdots\,\cdot\varphi_0(x_0).
	$$
	where $\varphi_i(x) = \Pi y[(y < x)\mapsto\Pi z_1\cdots\Pi z_i\,2]$ for $i > 0$ and $\varphi_0(x) = \Pi y[(y < x)\mapsto 2]$. Note that if $a$ is the  $m$-th element in $A$, then $\sem{\varphi_i(x)}(\A,a) = 2^{\vert A \vert^i m}$. Therefore, for each $(a_0,\ldots,a_{k-1})\in A^k$, we have that $\sem{\Phi(x_0,\ldots,x_{k-1})\cdot\varphi_{k-1}(x_{k-1})\,\cdots\,\varphi_0(x_0)}(\A,a_0,\ldots,a_{k-1}) = 2^{(\vert A \vert^{k-1}m_{k-1} + \cdots + \vert A \vert^2 m_2 + \vert A \vert m_1 + m_0)}$ if $(\A,a_0,\ldots,a_{k-1})\models\Phi$ and 0 otherwise, and adding these values gives $f(\A)$.

\textbf{Proof of Theorem \ref{qfo-pfp-cap}}

For the first condition, we use the exact same procedure as in Theorem \ref{captfp}, noting that evaluating each of the $\pfp$ formulas can be done in polynomial space, and that the result is also polynomial on the size of the input.

For the second condition, note that for each function $f\in \nfpspace$, $\vert f(\A) \vert$ is polynomial on $\vert A \vert$. Thus, there is a $\pfp$ formula $\Phi$ that models the string $f(\A)$ by the bit. The rest of the proof is analogous to Theorem \ref{captfp}.

\textbf{Proof of Theorem \ref{eqso-fo-cap}}

\textbf{Proof of Theorem \ref{eqso-fo-cap-not}}

\textbf{Proof of Theorem \ref{eqso-eso-cap}}

\textbf{Proof of Theorem \ref{fp1}}

Let $f \in \XE{0}$, and let $\varphi(\X,\x)$ be and {\sc FO}-extended quantifier-free $\L$-formula such that:
\begin{eqnarray*}
	f(\A) &=& \vert \{\langle \P,\e  \rangle \mid \A \models \varphi(\P,\e) \} \vert
\end{eqnarray*}
for each $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^m$ and $\P = (P_1,\ldots,P_q)$ is a tuple of predicates. We will now show that computing $f(\A)$ can be done in polynomial time.

For each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$, let $R_{\beta}$ be a predicate of arity $m$. Let $\A^\prime = \langle A, \S^{\A}, R_{\beta}^{\A^\prime}, \leq^{\A} \rangle \in \str$, where $R_{\beta}^{\A^\prime} = \{\d \mid \A \models \beta(\d)\}$. Note that each $\beta(\x)$ is fixed in $\varphi(\X,\x)$, and for each $\d \in A^m$, checking whether $\A \models \beta(\d)$ can be done in polynomial time. Therefore, generating $R_{\beta}^{\A^\prime}$ can also be done in polynomial time.

Let $\psi(\X,\x)$ be obtained by replacing each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$ by $R_{\beta}(\x)$. Also, let $g = f_{\psi(\X,\x)}$. Note that for each tuple of predicates $\P$ and each $\e \in A^m$, $\A \models \varphi(\P,\e)$ if and only if $\A^\prime \models \psi(\P,\e)$, and so, $g(\A^\prime) = f(\A)$. But, $\psi(\X,\x)$ is a quantifier-free $\L$-formula, and therefore, $g \in \#\Sigma_0$. Since it is shown in \cite{DBLP:journals/jcss/SalujaST95} that $\#\Sigma_0 \subseteq$ {\sc FP}, we conclude that $f(\A)$ can be evaluated in polynomial time.

\textbf{Proof of Theorem \ref{decisionptime}}

Let $f$ be a function in $\E1$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
\begin{eqnarray*}
	f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
\end{eqnarray*}
where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables. Let $\y = (\x,\z)$ and let $\psi(\X,\y) = \varphi(\x,\X,\z)$. Moreover, let $g = f_{\psi(\X,\y)}$.
\begin{claim} \label{mono}
	For each $\A \in \str$, $f(\A) > 0$ iff $g(\A) > 0$.
\end{claim}
\begin{proof}
	($\Rightarrow$) Suppose $f(\A) > 0$. Let $\P$ and $\e$ be such that $\A \models \exists \x \, \varphi(\x,\P,\e)$. It follows that there is at least one $\d \in A^k$ such that $\A \models \varphi(\d,\P,\e) = \psi(\P,(\d,\e))$. Therefore, $g(\A) > 0$.

	($\Leftarrow$) Suppose $g(\A) > 0$. Let $\Q$ and $\c = (\c_1,\c_2)$, where $\c_1$ and $\c_2$ have $k$ and $m$ elements respectively, be such that $\A \models \psi(\Q,\c) = \varphi(\c_1,\Q,\c_2)$. Then we have that $\A \models \exists \x \, \varphi(\x,\Q,\c_2)$, from which we conclude that $f(\A) > 0$.
\end{proof}
Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula, so $g \in\XE{0}$. It follows from \ref{mono} that for each $\A \in \str$, it holds that $\A \in L_f$ if and only if $g(\A) > 0$. But by Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time. From this we conclude that the decision version of $f$ is in $\ptime$.

\textbf{Proof of Theorem \ref{sub-pnp}}

Suppose that $\E1$ is closed under substraction, that is, for each pair of functions $f,g\in \E1$, there exists $h\in\E1$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\str$.

Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\Phi$, where $A$ is the set of variables mentioned in $\Phi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{1,2,3,4\}$:
\begin{eqnarray*}
	S_1^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
	S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
	S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
	S_4^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
\end{eqnarray*}
Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
\begin{multline*}
\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
\end{multline*}
Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
$$
\varphi(x,X) = (T(x) \vee \neg T(x)).
$$
Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula. Given that $f_{\#3DNF}, f_{all} \in \E1$, we have by our initial assumption that $f_{all}-f_{\#3DNF} \in \E1$. Let $h\in\E1$ be such that $h = f_{all}-f_{\#3DNF}$. For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is $\conp$-complete. However, as we showed previously in Theorem \ref{decisionptime}, since $h\in\E1$, we have that $L_h \in \ptime$. Then, $\conp \subseteq \ptime$, from which we conclude that $\ptime = \np$.

\textbf{Proof of Theorem \ref{sigma1strict}}

We will show that the $\E1$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$. By contradiction, suppose that it is. Let $\A = \langle A =\{1\},S^\A = \{1\},\leq^\A=\{(1,1)\} \rangle$. Then, $f(\A) = 1$. Now let $\B = \langle B =\{1,2\},S^\B = \{1\},\leq^\B=\{(1,1),(1,2),(2,2)\} \rangle$. Note that $\A$ is an induced substructure of $\B$.

We have that for each function $g \in \#\Sigma_1$ and structures $\A_1,\A_2\in\str$, if $\A_1$ is an induced substructure of $\A_2$, then $g(\A_1) \leq g(\A_2)$ \cite{DBLP:journals/jcss/SalujaST95}. Therefore, $f(\B) \geq f(\A) = 1$. However, there is no assignment $s\in B$ to $x$ such that $\B\models\:\varphi(s)$, so $f(\B) = 0$, which leads to a contradiction.

\textbf{Proof of Theorem \ref{sigmafo-minusone}}

Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$. There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$. This separates the proof in three cases:
\begin{enumerate}
	\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
	\begin{eqnarray*}
		f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
	\end{eqnarray*}
	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^d$. Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily. First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
	\begin{eqnarray*}
		\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
	\end{eqnarray*}
	This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
	\begin{eqnarray*}
		\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
	\end{eqnarray*}
	If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$. Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$. Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E1.$

	\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
	\begin{eqnarray*}
		f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
	\end{eqnarray*}
	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
	\begin{eqnarray}
	\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
	\end{eqnarray}
	where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$. Note that $\theta(\x)$ also mentions all variables in $\x$. We also assume that $(\x_1,\dots,\x_n,\y) = \x$. As an example, the following formula is of this form:
	\begin{multline*}
	\varphi(\x,\X) =  X_1(x_1,x_2) \wedge X_3(x_3) \wedge X_2(x_4,x_5) \wedge X_3(x_6) \wedge \neg X_1(x_7,x_8) \wedge \\ (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 ) \wedge \\ \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big),
	\end{multline*}
	where $\x = (x_1,x_2,x_3,x_4,x_5,x_6,x_7,x_8)$ and $\X = (X_1,X_2,X_3)$. Here, $n = 4$, $\lambda(1) = 1$, $\lambda(2) = \lambda(4) = 3$ and $\lambda(3) = 2$, $\x_1 = (x_1,x_2)$, $\x_2 = (x_3)$, $\x_3 = (x_4,x_5)$, $\x_4 = (x_6)$ and $\y = (x_7,x_8)$. Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_7,x_8)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 )$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2,x_5,x_8\},\{x_6,x_7\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big)$.

	Similarly to the previous proof, we would like to eliminate the {\em lexicographically smallest}\footnote{We consider the lexicographically smallest tuple of predicates as the one in which its predicates contain the lexicographically smallest tuples and do not contain any more tuples than those} tuple of predicates that satisfies the formula \eqref{f1}. Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$. We now define
	\begin{multline*}
	\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to \varphi_{m,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
	\end{multline*}
	where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$. Note that $\alpha_{\min}$ is satisfied only by the lexicographically \linebreak smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$. Our new formula is
	\begin{multline}
	\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge \\ \exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg].
	\end{multline}
	We now show a result by which the main proof will follow.
	\begin{lemma} \label{first}
		$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
	\end{lemma}
	\begin{proof}
		Let $\A \in \str$. Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$. Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$. Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j\in[1,n]:\:\lambda(j)=i}\{\d_j\}$. We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
		\begin{enumerate}
			\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$. That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$). Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$. Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$. We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \varphi_{<,a_i}(\z,\x_i)$ or (3) $\theta(\x)\models \varphi_{<,a_i}(\x_i,\z)$. Considering that (2) and (3) are not possible given that both $\z$ and $\x_i$ are assigned the value $\d_i$ and $\A\models\theta(\d)$, we have that $\theta(\x)\models \z = \x_i$. But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$, which means $\varphi(\x,\X)$ is inconsistent. That is, there's no structure $\A^\prime\in\str$ such that $\A^\prime\models\exists\x\:\varphi(\x,\R^\prime)$ for any assignment $\R^\prime$ to $\X$. In particular, $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which contradicts the initial assumption that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
			\item[(b)] Note that if $\A \models \alpha_{\min}(\c_1,\ldots,\c_n)$, then necessarily $\c_i = \d_i$ for $i\in\{1,\ldots,n\}.$ However, by the construction of $\P$, we see that 
			$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \d_j\Big).$$ Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
			\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$. We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$. Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $P^\prime_i$. Then $\A\models \neg P^\prime_i(\d_i)$, and so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$. Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $P_i$ does not have. If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$. On both cases, we have a contradiction.
		\end{enumerate}
		With this, we conclude that for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$

		Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$. Let $\P$ be an arbitrary assignment to $\X$. Since $\A\not\models\exists\x\,\varphi(\x,\P)$, we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$. It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$. And so, for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ it holds that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1,$ which was to be shown.
	\end{proof}
	We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula. By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\gamma_i(\x,\X)$ with $i\in\{1,\ldots,\ell\}$ such that
	\begin{eqnarray*}
		\varphi(\x,\X) &\equiv& \gamma_1(\x,\X) \vee \gamma_2(\x,\X) \vee \dots  \vee \gamma_{\ell}(\x,\X),
	\end{eqnarray*}
	where, for every $i\in\{1,\ldots,\ell\}$, 
	\begin{eqnarray*}
		\gamma_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \gamma^{-}_i(\X,\y_i)  \wedge \delta_i(\x).
	\end{eqnarray*}
	The function $\lambda_i$ is defined analogously to $\lambda$ of the first part of the proof. The tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y_i) = \x.$ The formulas $\gamma^{-}_i(\X,\y_i)$, $\delta_i(\x)$ are defined analogously to $\varphi^{-}(\X,\y)$ and $\beta(\x)\,$\footnote[2]{Note that $\delta_i(\x)$ can include subformulas of the form $(\u = \u)$.} respectively (see \eqref{f1}). Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables. Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$. Note that for every formula $\eta(\x)$,
	\begin{eqnarray*}
		\exists\x\:\eta(\x) &\equiv& \exists\x(\eta(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\eta(\x)\wedge\theta^{g(d)}(\x)).
	\end{eqnarray*}
	We define the following formulas $\xi_i(\x,\X)$, for $i\in\{1,\ldots,m\}$ where $m = \ell \cdot g(d)$, as follows: 
	\begin{eqnarray*}
		\xi_1(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^1(\x), \\
		& \vdots & \\
		\xi_{g(d)}(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
		\xi_{g(d)+1}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^1(\x), \\
		& \vdots & \\
		\xi_{2\cdot g(d)}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
		& \vdots & \\
		\xi_{(\ell-1) g(d)+1}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^1(\x), \\
		& \vdots & \\
		\xi_{\ell \cdot g(d)}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^{g(d)}(\x).
	\end{eqnarray*}
	Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\str.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts. We use an injective\footnote[3]{We need this function to be injective because this way we can assure each one of the $k$ satisfiable disjuncts to be represented} function $\rho:\{1,\ldots,k\}\to\{1,\ldots,m\}$ such that $\varphi_i(\x,\X) = \xi_{\rho(i)}(\x,\X)$ is satisfiable, for every $i\in\{1,\ldots,k\}.$

	We can conclude that
	\begin{eqnarray*}
		\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X),
	\end{eqnarray*}
	and for every $i\in\{1,\ldots,k\}$,
	\begin{eqnarray*}
		\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
	\end{eqnarray*}
	where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\y_i$ is a $p_i$-tuple of variables in $\x$, $\varphi_i^{-}(\X,\y_i)$ is a conjunction of negated predicates in $\X$, $\theta_i(\x)$ defines a total order on a partition of $\x$, and $\beta_i(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$. Let $\alpha_i(\x) = \theta_i(\x)\wedge\beta_i(\x).$

	%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
	%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM

	\begin{claim} \label{alpha}
		Let $\A\in\str$ and $i\in\{1,\ldots,k\}$. For every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, it holds that $f_{\varphi_i(\s,\X)}(\A) > 0.$
	\end{claim}
	\begin{proof}
		Let $\s = (\s_1,\ldots,\s_{n_i},\t)$ such that $\A\models\alpha_i(\s)$, and let $\P = (\P_1,\ldots,\P_r)$ where $P_i = \bigcup_{j\in[1,n_i]:\:\lambda(j)=i}\{s_i\}$. We will show that $\A\models\varphi_i(\s,\P)$. By contradiction, suppose that $\A\not\models\exists\x\,\varphi_i(\x,\P)$. That is, there is no assignment $\e$ to $\x$ such that $\A\models\varphi(\e,\P).$ Following the same proof as in case (a) of Lemma \ref{first}, we conclude that $\varphi_i(\x,\P)$ is inconsistent. However, as we mentioned previously, all of such disjuncts have been eliminated, which leads to a contradiction.
	\end{proof}

	Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\exists\x\:\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\Q$ such that $\A\models\exists\x\:\varphi_2(\x,\Q)$, and so on. As we already know how to exclude that assignment in the first disjunct, we will now deal with the next disjuncts. Similarly to the first part of the proof, for each $i\in\{1,\ldots,k\}$ let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
	\begin{multline*}
	\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_{n_i},\y)\wedge \\ \forall\v_1\cdots\forall\v_{n_i}\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_{n_i},\w)\wedge\bigvee_{j=1}^{n_i}(\u_j\neq\v_j)\big)\to \varphi_{m_i,<}((\u_1,\dots,\u_{n_i}),(\v_1,\dots,\v_{n_i}))\Big)\Big],
	\end{multline*}
	where $\u_j$ and $\v_j,$ have $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, and $\w$ has $p_i$ variables. Also, let
	\begin{multline*}
	\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
	\end{multline*}
	Note that $\psi_i(\X)$ excludes only the lexicographically smallest tuple of predicates $\P$ such that $\A\models\exists\x\:\varphi_i(\P,\x),$ if there is at least one. In other words, every assignment $\P^\prime \neq \P$ is such that $\A\models\psi_i(\P^\prime).$ Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
	\begin{multline}
	\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge \\ (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
	\end{multline}
	Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Assume that $\A\in\str$. Suppose first that there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least one assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\d = (\d_1,\dots,\d_{n_q},\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha_q(\d)$, where $\d_i$ is the corresponding assignment to $\x_i$, for every $i\in\{1,\dots,n_q\}$, and $\ep$ is an assignment for $\y$. Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j:\lambda_q(j)=i,j\in[1,n_q]}\{\d_j\}$. As we did in Lemma \ref{first} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies both (a) and (b)
	\begin{enumerate}
		\item[(a)] As we showed in part (a) of Lemma \ref{first}, if there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_q(\x,\R)$, then $\A\models\exists\x\:\varphi_q(\x,\P)$ for this particular $\P$. However, as we showed in Claim \ref{alpha}, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_q(\s)$, then there {\em is} such an assignment to $\X$. The assignment $\d$ to $\x$ satisfies that $\A\models\alpha_q(\d)$, so it holds that $\A\models\exists\x\:\varphi_q(\x,\P)$. It immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$

		\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$. (1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$. (2) We can use the proof in Lemma \ref{first} to see that $\A\not\models\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula 
		$$ \zeta_q(\X) =  (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{q-1}(\v)\vee\psi_q(\X)) $$ 
		appears as a conjunct in $\varphi^\prime_i(\x,\X)$. However, by the choice of $q$, there is no $i\in\{1,\ldots,q-1\}$ such that $\A\models\exists\v\:\varphi_i(\v,\P)$, and also $\A\not\models\psi_q(\P)$. It follows that $\A\not\models\zeta_q(\P)$,  so $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$. And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$

		\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b). As we deduce from the part (c) of Lemma \ref{first}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$. Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x)$, so $\A\models\psi_i(\P^\prime)$ for each $i$. But as we mentioned, also $\A\models\:\psi_q(\P^\prime),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which leads to a contradiction.
	\end{enumerate}

	With this, we conclude that for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0$, we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$

	Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$. Let $\P$ be an arbitrary assignment to $\X$. Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\chi(\x,\P))$ for any formula $\chi(\x,\P)$. It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$. And so, for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Hence, from the results in the previous paragraph, if $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}$, we have that $f^\prime = f\dotminus 1\in\E1$

	\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
	\begin{eqnarray*}
		f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
	\end{eqnarray*}
	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$. In order to prove that $f\dotminus 1\in\E1$, we define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2, where
	\begin{eqnarray*}
		\varphi(\x,\X,\z) \equiv \varphi_1(\x,\X,\z) \vee \cdots \vee \varphi_k(\x,\X,\z),
	\end{eqnarray*}
	and
	\begin{eqnarray*}
		\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\w_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
	\end{eqnarray*}
	where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\ell} + c_{\ell} = a_{\ell}$ for $\ell\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, and $\w_i$ has $q_i$ variables. Furthermore, we have that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\w_i) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2. The formulas $\varphi^{-}_i(\X,\y_i,\w_i)$, $\beta_i(\x,\z)$, are also defined analogously.

	In this case, we mix both the strategies in cases 1 and 2. That is, we are going to {\em isolate} the lexicographically smallest tuple of predicates that satisfies the first satisfiable disjunct, and then {\em exclude} the lexicographically smallest tuple that satisfies the isolated disjunct.

	Let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and
	\begin{multline*}
	\alpha^i_{\min}(\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}) = \exists\y\exists\w\Big[ \alpha_i(\x_1,\dots,\x_{n_i},\y,\z_1,\dots,\z_{n_i},\w)\wedge \\ \forall\u_1\cdots\forall\u_{n_i}\forall\s\forall\v_1\cdots\forall\v_{n_i}\forall\t\Big(\big(\alpha_i(\u_1,\dots,\u_{n_i},\s,\v_1,\dots,\v_{n_i},\t)\wedge\bigvee_{j=1}^{n_i}(\x_j\neq\u_j \vee \z_j\neq\v_j)\big)\to \\ \varphi_{m_i,<}((\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}),(\u_1,\dots,\u_{n_i},\v_1,\dots,\v_{n_i}))\Big)\Big],
	\end{multline*}
	and let
	\begin{multline*}
	\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\w(\exists\u\:\varphi(\u,\X,\w) \wedge \varphi_{p,<}(\w,\z)).
	\end{multline*}
	Let $\P$ and $\e$ be assignments to $\X$ and $\z$. Note that $\psi_i(\X,\z)$ excludes $\P$ and $\e$ only if $\P$ is the {\em lexicographically smallest} tuple of predicates (Same as case 2) such that $\A\models\exists\x\:\varphi_i(\x,\P,\d)$, for some assignment $\d$ to $\z$, and $\e$ is the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_i(\x,\P,\e).$ We define $\varphi_i^\prime(\x,\X,\z)$ in the same way as case 2:
	\begin{multline*}
	\varphi_i^\prime(\x,\X,\z) = \varphi_i(\x,\X,\z) \wedge \psi_1(\X,\z) \wedge (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\psi_2(\X,\z)) \wedge \cdots \wedge \\ (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\cdots\vee\exists\u\exists\v\:\alpha_{i-1}(\u,\v)\vee\psi_i(\X,\z)).
	\end{multline*}
	Finally, let $\varphi^\prime(\x,\X,\z) = \bigvee_{i = 1}^k \varphi_i^\prime(\x,\X,\z)$.

	Let $q$ be the least $i\in\{1,\ldots,k\}$ such that $\A\models\exists\x\:\varphi_i(\x,\R,\d)$ for some assignment $R$ to $\X$ and some assignment $\d$ to $\z$. Let $\P$ be the lexicographically smallest tuple of predicates such that $\A\models\exists\x\:\varphi_q(\x,\P,\d^\prime)$ for some assignment $\d^\prime$ to $\z$. Let $\e$ be the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_q(\x,\P,\e).$ This formula is such that (a) $\A\models\exists\x\:\varphi(\x,\P,\e)$, (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P,\e)$ and $\P$ and $\e$ are the only assignments that satisfy (a) and (b). The proof of this is analogous to case 2. Therefore, we conclude that $f_{\exists\x\,\varphi^\prime(\x,\X,\z)} = f_{\exists\x\,\varphi(\x,\X,\z)}\dotminus 1.$
\end{enumerate}

\textbf{Proof of Theorem \ref{dnf-pars}}

Let $f$ be an arbitrary function in $\E1$ and let $f_{\textsc{\#DNF}}$ be the function that defines {\sc \#DNF}. We will now show a function $h:\str\to \textsc{Struct}[\L_{\textsc{DNF}}]$, where $\L_{\textsc{DNF}} = \{N,P\}$, such that for every $\A\in\str$, it holds that $f(\A) = f_{\textsc{\#DNF}}(h(\A))$, and which can be computed in polynomial time. For every DNF formula $\Phi$ we define $\A_{\Phi} \in \textsc{Struct}[\L_{\textsc{DNF}}]$ as stated in \ref{dnfdef}. Let $\psi$ be such that $f = f_{\psi(\X,\z)}$. We separate the proof in three cases:
\begin{enumerate}
	\item $\X = ()$. Let $\psi(\z) = \exists \x\,\varphi(\x,\z)$ where $\x = (x_1,\ldots,x_c)$ and $\z = (z_1,\ldots,z_d)$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a propositional formula $\Phi = h(\A)$. First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\,\varphi(\x,\e)$. Let $n$ be the number of such assignments to $\z$. Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$. The DNF formula we generate is
	\[
	\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
	\]
	which has exactly $n$ possible assignments. Finally, we return $h(\A) = \A_{\Phi}$.

	\item $\z = ()$. Let $\psi(\X) = \exists \x\,\varphi(\x,\X)$, where $\x = (x_1,\ldots,x_c)$ and $\X = (X_1,\ldots,X_r)$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a DNF-formula $\Phi = h(\A)$ as follows. First, we convert $\varphi(\x,\X)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X)$. Let each $\varphi_{i}(\x,\X)$ be of the form
	\[
	\varphi_{i}(\x,\X) = X_{\pi_i(1)}(\x^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\x^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\x^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\x^{\nu}_{i,n_i}) \wedge \alpha_i(\x),
	\]
	where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$) is the rearrangement of $\x$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x)$ is a FO formula. We also define the tuple $\a^{\pi}_{i,j}$ ($\a^{\nu}_{i,j}$) as $\a$ rearranged in the same way as $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$), for each $\a\in A^c$.

	We generate a propositional variable $p^{\ell}_{\a}$ for each $\a\in A^{c_{\ell}}$, and for each $\ell\in\{1,\ldots,r\}$, which will indicate if the tuple $\a$ is in the assignment to $X_{\ell}$. Let $\p$ be the tuple of all such variables.

	Then, let $\Gamma_i = \{\a\in A^c\mid \A\models\alpha_i(\a)\}$. There is exactly $\vert A \vert^c$ tuples to check, and each of those checks can be done in polynomial time, so this set can also be generated in polynomial time.

	Finally, for each $\varphi_i(\x,\X)$ the DNF formula we generate is
	\[
	\Phi_i = \bigvee_{\a\in\Gamma_i} p^{\pi_i(1)}_{\a^{\pi}_{i,1}} \wedge \cdots \wedge p^{\pi_i(m_i)}_{\a^{\pi}_{i,m_i}} \wedge \neg p^{\nu_i(1)}_{\a^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\pi_i(n_i)}_{\a^{\nu}_{i,n_i}}.
	\]
	and $\Phi = \bigvee_{i = 1}^k \Phi_i \vee \bigvee_{p\in\p} (p\wedge\neg p)$. Finally, we return $h(\A) = \A_{\Phi}$.

	As an example, let $\psi^{*}(X_1,X_2) = \exists x \exists y\, (X_1(x) \wedge X_2(y,x) \wedge \neg X_1(y) \wedge x < y \wedge \neg\exists z\,(x<z\wedge z<y))$. We will compute $h(\B)$, where $\B = \langle \{1,2,3\}, < \rangle$, and we have that $1 < 2 < 3$. We generate the tuple of variables $\p = (p^1_1,p^1_2,p^1_3) \cup (p^2_{(i,j)})$ for $(i,j)\in \{1,2,3\}\times\{1,2,3\}$. Note that in this procedure we will only have one set $\Gamma_1 = \{(1,2),(2,3)\}$. And so, the DNF formula we generate is 
	\begin{multline*}
	\Phi^{*} = (p^1_1\wedge p^2_{(2,1)} \wedge \neg p^1_2)\vee(p^1_2\wedge p^2_{(3,2)} \wedge \neg p^1_3)\, \vee \\ (p^1_1 \wedge \neg p^1_1) \vee(p^1_2 \wedge \neg p^1_2) \vee (p^1_3\wedge \neg p^1_3) \vee \bigvee_{i,j\in\{1,2,3\}}(p^2_{(i,j)}\wedge\neg p^2_{(i,j)})
	\end{multline*}

	Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$. We can conclude that $f(\A) = f_{\textsc{DNF}}(\Phi)$.

	\item Let $\psi = \exists \x\,\varphi(\x,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$ we generate a DNF-formula $\Phi = h(\A)$ as follows. As in the previous case, we convert $\varphi(\x,\X,\z)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X,\z)$. Let each $\varphi_{i}(\x,\X,\z)$ be of the form
	\[
	\varphi_{i}(\x,\X,\z) = X_{\pi_i(1)}(\y^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\y^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\y^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\y^{\nu}_{i,n_i}) \wedge \alpha_i(\x,\z),
	\]
	where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$) is the rearrangement of $(\x,\z)$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x,\z)$ is a FO formula. We also define the tuple $(\a,\e)^{\pi}_{i,j}$ ($(\a,\e)^{\nu}_{i,j}$) as $\a$ rearranged in the same way as $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$), for each $(\a,\e)\in A^{c+d}$.

	We generate a propositional variable $p^{\e}_{\ell,\b}$ for each $\e\in A^d$, for each $\ell\in\{1,\ldots,r\}$, and for each $\b\in A^{c_{\ell}}$. This variable will be true if there is an assignation to $\P$ such that $\b\in P_{\ell}$ and $\A\models\psi(\P,\e)$. Let $\p$ be the tuple of all such variables. We also generate a set $T = \{t_1,\ldots,t_{L}\}$ where $L = \lceil \log\vert A^d\vert \rceil$, and an arbitrary injective function $\beta:A^d \to \{0,1\}^L$ where $\beta(\e) = (b^{\e}_1,\ldots,b^{\e}_L)$ for each $\e\in A^d$. Also, let $\rho_{\ell}$ be such that $\rho_{\ell}(\x,\z) = \y_{\ell}$.

	For each $\e\in A^d$, we follow the following procedure. For each $\varphi_i(\x,\X)$, we consider the set $\Gamma^{\e}_i = \{\a\in A^c\mid \A\models\alpha(\a,\e)\}$, and so, the DNF formula we generate is
	\[
	\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{\pi_i(1),(\a,\e)^{\pi}_{i,1}} \wedge \cdots \wedge p^{\e}_{\pi_i(m_i),(\a,\e)^{\pi}_{i,m_i}} \wedge \neg p^{\e}_{\nu_i(1),(\a,\e)^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\e}_{\pi_i(n_i),(\a,\e)^{\nu}_{i,n_i}} \wedge \gamma_{\e} \wedge \theta_{\e},
	\]
	where $\gamma_{\e}$ is defined as
	\[
	\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
	\]
	and $\theta_{\e}$ is defined as
	\[
	\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
	\]
	Finally,$\Phi = \bigvee_{\e\in A^d} \bigvee_{i = 1}^k \Phi_i^{\e} \vee \bigvee_{p\in\p} (p\wedge\neg p)$, and we return $f(\A) = \A_{\Phi}$.

	As an example, let $\psi^{*}(X_1,X_2,z) = \exists x\, (X_1(x) \wedge X_2(z) \wedge \neg X_1(z) \wedge x < z)$. We will compute $h(\B)$, where $\B = \langle \{1,2,3\}, < \rangle$, and we have that $1 < 2 < 3$. We generate the tuple of variables $\p = (p^e_{1,1},p^e_{1,2},p^e_{1,3},p^e_{2,1},p^e_{2,2},p^e_{2,3})$, for $e\in \{1,2,3\}$. Here, we consider $L = 2$, $T = \{t_1,t_2\}$, and $\beta$ defined as $\beta(1) = (0,0)$, $\beta(2) = (0,1)$ and $\beta(3) = (1,0)$. In this procedure we generate the sets $\Gamma^1_1 = \emptyset$, $\Gamma^2_1 = \{1\}$, $\Gamma^3_1 = \{1,2\}$. Now we generate $\gamma_e$ for $e\in\{1,2,3\}$,
	\[
	\gamma_e = \bigwedge_{d\in\{1,2,3\}\setminus\{e\}} p^d_{1,1} \wedge p^d_{1,2} \wedge p^d_{1,3} \wedge p^d_{2,1} \wedge p^d_{2,2} \wedge p^d_{2,3},
	\]
	and $\theta_e$ for $e\in\{1,2,3\}$,
	\begin{align*}
	\theta_1 &= \neg t_1 \wedge \neg t_2,\\
	\theta_2 &= \neg t_1 \wedge t_2,\\
	\theta_3 &= t_1 \wedge \neg t_2.		
	\end{align*}
	Then, we generate the formulas $\Phi^e_1$, for $e\in\{1,2,3\}$,
	\begin{align*}
	\Phi^1_1 &= \square\\
	\Phi^2_1 &= p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2,\\
	\Phi^3_1 &= (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3),
	\end{align*}
	and so, the DNF formula we generate is
	\begin{multline*}
	\Phi = (p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2) \vee (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee \\ \bigvee_{e\in\{1,2,3\}}[(p^e_{1,1}\wedge\neg p^e_{1,1})\vee
	(p^e_{1,2}\wedge\neg p^e_{1,2})\vee
	(p^e_{1,3}\wedge\neg p^e_{1,3})\vee
	(p^e_{2,1}\wedge\neg p^e_{2,1})\vee
	(p^e_{2,2}\wedge\neg p^e_{2,2})\vee
	(p^e_{2,3}\wedge\neg p^e_{2,3})].
	\end{multline*}
	To demonstrate that our reduction is correct, we will show a function $g:\{\langle\P,\e\rangle \mid \A \models \exists\x\,\varphi(\x,\P,\e)\} \to \{\sigma \mid \sigma(\Phi) = 1\}$ and then prove that it is a bijection. Let $g$ be such that for each assignment $\langle \P,\e \rangle$, it holds that $g(\langle \P,\e \rangle) = \sigma$ where $\sigma$ is constructed as follows. For each $t_i\in T$, let $\sigma(t_i) = b^{\e}_i$. For each $p_{\ell,\b}^{\e}$, where $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e}) = 1$ if and only if $\b \in P_{\ell}$. For each $p_{\ell,\b}^{\e'}$, where $\e'\in A^d\setminus\{\e\}$,  $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e'}) = 1$.

	Let $i$ be such that $\A\models \exists\x\,\varphi_i(\x,\P,\e)$. It is simple to see that from the way we constructed $\Phi_i^{\e}$, it holds that $\sigma(\Phi_i^{\e}) = 1$. Also, note that every different assignment $\langle \P',\e' \rangle$, it holds that $\sigma(\langle \P',\e' \rangle) \neq \sigma(\langle \P,\e \rangle)$. Therefore, $g$ is an injective function. Now, let $\sigma$ be an arbitrary assignment such that $\sigma(\Phi) = 1$. Let $i$ and $\e$ be such that $\sigma(\Phi^{\e}_i)$. Note that for every $\e'\in A^d\setminus\{\e\}$, it holds that $\sigma(\theta_{\e'}) = 0$, so there is exactly one possible $\e$ associated to $\sigma$. Consider the assignment $\P$ where $\b\in P_{\ell}$ if and only if $\sigma(p^{\e}_{\ell,\b}) = 1$, for each $\ell$ and $\b \in A^{c_{\ell}}$. Since there is at least one $\a\in\Gamma_i$ for which $\sigma$ satisfies its respective subformula, it holds that $\A\models\varphi_i(\a,\P,\e)$, which implies that $\A\models\exists\x\,\varphi(\x,\P,\e)$. From this we deduce that $g$ is surjective, and therefore a bijection. We conclude that $f(\A) = f_{\textsc{\#DNF}}(\A_{\Phi})$.
\end{enumerate}

\textbf{Proof of Theorem \ref{nplusone-strict}}

Suppose it is not properly contained for $n$. Let $\z = (z_1,\ldots,z_{n+1})$, let $\u = (u_1,\ldots,u_n)$ and let $\varphi(\z) = \bigwedge_{i = 1}^{n} (z_i < z_{i+1} \wedge \forall y(z_i\not<y \vee y\not<z_{i+1} )) \wedge \forall y(y \not< z_1)$. Note that $f_{\varphi(\z)}(\A) = 1$ if $\A$ has at least $n+1$ elements, and  $f_{\varphi(\z)}(\A) = 0$ otherwise, for each $\A\in\str$. Let $\forall\x\,\psi(\x,\X,\u)$ be such that $f_{\varphi(\z)} = f_{\forall\x\,\psi(\x,\X,\u)}$. Let $\A = \langle A, \S^{\A}, <^{\A} \rangle\in\str$ where $\vert A \vert = n+1$. Note that $f_{\forall\x\,\psi(\x,\X,\u)}(\A) = 1$. Let $(\P,\e)$ be such that $\A\models\forall\x\,\psi(\x,\P,\e)$. Let $a\in A$ be such that $a\not\in\e$. Consider the vocabulary $\L' = \L \cup \{\Q\}$, and let $\A' = \A \cup \langle \Q^{\A'} \rangle$ where $\Q^{\A'} = \P$. Note that $\A\models\varphi(\P,\e)$ if and only if $\A'\models\varphi(\e)$. Consider $\B'$ as the induced substructure of $\A'$ with domain $A\setminus\{a\}$, and let $\Q^{\B'}$ be the induced interpretation of $\Q$. Also let $\B$ be the induced substructure of $\A$ with domain $A\setminus\{a\}$. As universal formulae are preserved under substructure\footnote{Hodges, 1997. P. 183}, we have that $\B'\models\forall\x\,\psi(\x,\e)$, which implies that $\B\models\forall\x\,\psi(\x,\Q^{\B'},\e)$. We conclude that $f_{\forall\x\,\psi(\x,\X,\u)}(\B) \geq 1$, which leads to a contradiction.

\textbf{Proof of Theorem \ref{sigma2-pe}}

Let $f = f_{\varphi(\X,\z)}$ such that $\varphi(\X,\z)$ is defined by the grammar $E_2$. First we notice that, as stated in \cite{DBLP:journals/jcss/SalujaST95}, for every $\L$-formula $\varphi$, $f_{\varphi}\in\shp$.\\

We will now prove that there is a polynomial time algorithm that decides $L_{f}$. This is equivalent to decide, given $\A= \langle A, \S^\A, \leq^\A \rangle\in\str$, whether there exist assignments $\P,\z$ such that $\A\models\varphi(\P,\z)$. This is the same as $\A\models\exists\X\exists\x\,\varphi(\X,\x)$. Since $\varphi(\X,\x)$ is defined by the grammar $E_2$, there exists $\psi(\X,\x,\y,\u)$ such that $\varphi(\X,\x) = \exists\x\exists\y\forall\u\exists\v\,\psi(\X,\x,\y,\u,\v)$, where every variable in $\v$ appears on a negated second-order literal $\neg X_i$, with $i\in\N$.

Given $\A\in\str$ we generate an equivalent $\L$-formula $\theta(\X,\x,\y)$ with a series of operations. First, we replace every instance of $\v$ by a disjunction of all $r$-tuples in $A^r$, where $r$ is the number of variables in $\v$. The result is still a Horn clause. Second, we replace every instance of $\u$ by a conjunction of all $s$-tulples in $A^s$, where $s$ is the number or variables in $\u$. The result is still a conjunction of Horn clauses.

Now we notice that $\zeta = \exists\X\exists\x\exists\y\,\theta(\X,\x,\y)$ is an existential second-order FO-formula, for which $\A\models\zeta$ can be decided in polynomial time \cite{DBLP:conf/stacs/Gradel91}.

\textbf{Proof of Theorem \ref{sigma2hard}}

Let $f$ be an arbitrary function in $\XE{2}$. We will now show a function $h:\str\to L(P)$ such that for every $\A\in\str$, it holds that $f(\A) = \shdhsat(h(\A))$, and which can be computed in polynomial time. Let $\psi$ be such that $f = f_{\psi}$. We separate the proof in three cases:\begin{enumerate}
	\item All free variables in $\psi$ are of first order. Let $\psi = \exists \x \forall \y\,\varphi(\x,\y,\z)$ where $\z = (z_1,\ldots,z_d)$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a propositional formula $\Phi = h(\A)$. First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\forall \y\,\varphi(\x,\y,\e)$. Let $n$ be the number of such assignments to $\z$. Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$. The disjunction of Horn formulas we return is
	\[
	\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
	\]
	which has exactly $n$ possible assignments. Note that each $p_i$ and $\neg p_i$ is a Horn clause itself.

	\item There are only second-order free variables in $\psi$. Let $\psi = \exists \x\forall \y\,\varphi(\x,\y,\X)$, where $\x = (x_1,\ldots,x_c)$, $\y = (y_1,\ldots,y_b)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a Disj-Horn formula $\Phi = h(\A)$ as follows:

	First, Let $\varphi(\x,\y,\X) = \bigwedge_{i = 1}^k \varphi_{i}(\x,\y,\X)$ where, without loss of generality, each $\varphi_{i}(\x,\y,\X)$ is of the form
	\[
	\varphi_{i}(\x,\y,\X) = X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n}) \vee \alpha_i(\x,\y),
	\]
	where $\u_{s}$ and every $\u_{t_j}$ have the corresponding number of variables in $(\x,\y)$, and $\alpha(\x,\y)$ is a FO-formula. For the rest of the proof, $X_s(\u_s)$ can be absent and it does not affect the proof. Note that this formula is equivalent to
	\[
	\neg\alpha_i(\x,\y) \to (X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n})),
	\]

	Then, for each $\a\in A^c$ let $\Gamma_i^{\a} = \{\b\in A^{b}\mid \A\models\neg\alpha(\a,\b)\}$. Note that this set can be generated in polynomial time. For each $\ell\in\{1,\ldots,r\}$, we generate a propositional variable $p^{\ell}_{\d}$ for each $\d\in A^{c_{\ell}}$. Let $\p$ be the tuple of all such variables. Also, let $\rho_{\ell}(\x,\y) = \u_{\ell}$, and $\rho_{\ell}(\d)$ be the tuple that results of reordering $\d$ using the reordering of the variables of $(\x,\y)$ in $\u_{\ell}$.

	For each $\a\in A^c$ we generate a Horn formula
	\[
	\Phi_{\a}^i = \bigwedge_{\b\in\Gamma_i} p^{s}_{\rho_s(\a,\b)} \vee \neg p^{t_1}_{\rho_{t_1}(\a,\b)} \vee \cdots \vee \neg p^{t_n}_{\rho_{t_n}(\a,\b)}.
	\]
	and finally we return $\Phi = \bigvee_{\a\in A^c} \bigwedge_{i = 1}^k \Phi_{\a}^i$.

	Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\forall\y\,\varphi(\x,\y,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$. We can conclude that $f(\A) = f_{\textsc{Disj-Horn-Sat}}(\Phi)$.

	\item Let $\psi = \exists \x\forall\y\,\varphi(\x,\y,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\y = (y_1,\ldots,y_b)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$. Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$ we generate a Disj-Horn formula $\Phi = h(\A)$ as follows:

	As in the previous case, let $\varphi(\x,\y,\X,\z) = \bigwedge_{i = 1}^k$ where $\varphi(\x,\y,\X,\z)$ is of the form:
	\[
	\varphi_{i}(\x,\y,\X,\z) = X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n}) \vee \alpha_i(\x,\y,\z),
	\]
	where $\u_{s}$ and every $\u_{t_j}$ have the corresponding number of variables in $(\x,\y,\z)$, and $\alpha(\x,\y,\z)$ is a FO-formula. For the rest of the proof, $X_s(\u_s)$ can be absent and it does not affect the proof. Note that this formula is equivalent to
	\[
	\neg\alpha_i(\x,\y,\z) \to (X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n})).
	\]

	For each $\e\in A^d$, we do the following. For each $\a\in A^c$ let $\Gamma^{\e}_{i,\a} = \{\b \in A^b \mid \A\models \neg\alpha_i(\a,\b,\e) \}$. Note that this set can be generated in polynomial time. For each $\ell\in\{1,\ldots,r\}$ we generate a propositional variable $p^{\e}_{\ell,\d}$ for each $\d\in A^{c_{\ell}}$. Let $\p$ be the tuple of all such variables. We also generate a set $T = \{t_1,\ldots,t_{L}\}$ where $L = \lceil \log\vert A^d\vert \rceil$, and an arbitrary injective function $\beta:A^d \to \{0,1\}^L$ where $\beta(\e) = (b^{\e}_1,\ldots,b^{\e}_L)$ for each $\e\in A^d$. Also, let $\rho_{\ell}(\x,\y,\z) = \u_{\ell}$, and $\rho_{\ell}(\d)$ be the tuple that results of reordering $\d$ using the order of the variables in $(\x,\y,\z)$ in $\u_{\ell}$. And so, the Horn formula we generate is
	\[
	\Phi^{\e}_{i,\a} = \bigwedge_{\a\in\Gamma^{\e}_i} p^{\e}_{s,\rho_{s}(\a,\b,\e)} \vee p^{\e}_{t_1,\rho_{t_1}(\a,\b,\e)} \vee \cdots \vee p^{\e}_{t_n,\rho_{t_n}(\a,\b,\e)}.
	\]
	And finally we return $\Phi = \bigvee_{\e\in A^d}\bigvee_{\a\in A^c} (\bigwedge_{i=1}^k \Phi_{\e}^{i,\a} \wedge \gamma_{\e} \wedge \theta_{\e} )$, where $\gamma_{\e}$ is defined as
	\[
	\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
	\]
	and $\theta_{\e}$ is defined as
	\[
	\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
	\]

\end{enumerate}

\textbf{Proof of Theorem \ref{so-rec}}

\textbf{Proof of Theorem \ref{tqfo-fo-cap}}

For the second condition, let $f \in \shl$. We will address the case where $\R$ contains only one binary predicate $E$, and the rest of the cases can be deduced from this. Let $M$ be a non-deterministic logspace machine such that $f(\A) = \acc_M(\A)$ for each $\A \in \ostr[R]$. Suppose ${\cal Q} = \{q_1,\ldots,q_{\ell}\}$ is the set of states of $M$, where $q_1$ is the initial state, and $q_{\ell}$ is only final state of $M$. Let $n = \vert A \vert$ and let $w = \enc(\A) \in \{0,1\}^{n^2}$. We assume that $M$ with input $w$ uses space $s_M(w) < c\cdot\log(n)$ and furthermore, $s_M(w) < n-2$. We notate $M(w)$ as the graph of configurations of $M$ running on input $w$.

We represent configurations with a tuple of fixed size. The formula $\varphi(\bar{x},\bar{y})$ describes a procedure that given a configuration generates a possible next configuration. The formula $\varphi_I(\bar{x})$ describes that $\bar{x}$ is the initial configuration of $M(w)$. The formula $\varphi_F(\bar{x})$ describes that $\bar{x}$ is an accepting (final) configuration of $M(w)$. The formula we construct is:
$$
\alpha = \sa{\bar{x}}\sa{\bar{y}}([\pth \varphi(\bar{u},\bar{v})](\bar{x},\bar{y})\cdot \varphi_I(\bar{x})\cdot\varphi_F(\bar{y})).
$$

To illustrate our idea, we will show a simplified example. Consider a machine $M$ that works in exactly $\log_2(n)$ space and only allows 0 or 1 in the working tape. Consider an input $\A$ of size 16 (that is, $A = \{0,\ldots,9,A,\ldots,F\}$). Let some configuration $s$ have 0011 in the working tape, the head in the input tape is in position 26, and the head in the input tape is in position 2 (we consider 0-indexed positions). Also, $Q = \{q_1,\ldots,q_5\}$ and the current state is $q_3$.

As a first approach, we will use a 9-tuple $\bar{a} = (a_1,\ldots,a_9)$ to represent $s$. That is, $(a_1,a_2) = (1,A)$ represent the position of the head in the input tape, $a_3 = 2$ represents the position of the head in the working tape, $a_4 = C$ (1100b in base 16) represents the content of the working tape, and $(a_5,\ldots,a_9) = (0,0,1,0,0)$ represents the current state. Then $\bar{a} = (1,A,2,C,0,0,1,0,0)$ will represent $s$.

The problem that arises from this representation, is that to describe a transition in $M$ we need to read an arbitrary character in the working tape, and therefore we have to obtain the $a_3$-th bit in $a_4$. Furthermore, to represent the following configuration, we need compute $a_4$ with the $a_3$-th bit flipped. This is generally not possible to describe with an $\fo$ formula. Consider the following procedure that receives $x = a_4$ and $i = a_3$.

\begin{algorithm}
	\caption{If the $i$-th bit in $x$ is 1 replace it by 0 and return the result}
	\label{switch1to0}
	\begin{algorithmic}
		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
		\While{$j > 0$}
		\State $v \gets 0$
		\While{$u > 1$}
		\State $u \gets u-2,\; v \gets v+1$
		\EndWhile
		\State $u\gets v,\; j \gets j-1$
		\EndWhile
		\While{$u > 1$}
		\State $u \gets u-2$
		\EndWhile
		\State $\textbf{assert } u = 1$ \Comment{If $u \neq 1$ simply stop}	
		\State $y \gets 1$ \Comment{Compute $2^i$ and store it in $y$}
		\While{$i > 0$}
		\State $z \gets 0$
		\While{$y > 0$}
		\State $z \gets z+2,\; y \gets y-1$
		\EndWhile
		\State $i \gets i-1,\; y \gets z$
		\EndWhile
		\While{$y > 0$} \Comment{Substract $y$ from $x$}
		\State $x \gets x-1,\; y \gets y-1$
		\EndWhile
		\State \Return $x$.
	\end{algorithmic}
\end{algorithm}	
Each of the instructions can be expressed with $\fo$, so our strategy is to use the $\pth$ operator to simulate the algorithm and then we can describe a transition using the processed value of $a_4$. Note that we need to describe three more procedures to simulate the transitions for $0\to 0$, $0 \to 1$ and $1\to 1$.

We will now describe how to simulate both the procedure and the transition. A procedure tuple $\bar{p} = (a_1,\ldots,a_{3+c+\ell},b_1,b_2,c_1,c_2,c_3,d_1,\ldots,d_{5c+2})$ represents the current configuration of $M(w)$ in $a_1,\ldots,a_{2+c+\ell}$, the values that will be read and written in the working tape in $b_1,b_2$, the instruction pointer in $c_1,c_2,c_3$ and the values stored in memory in $d_1,\ldots,d_{10c+2}$. In detail:
\begin{enumerate}
	\item $a_1,a_2$ and $a_3$ represent the position of the head in the input tape and the working tape, respectively, $a_4,\ldots,a_{3+c}$ represent the content of the working tape and $a_{4+c},\ldots,a_{3+c+\ell}$ represent the current state in the current configuration that is being processed.
	\item $b_1$ and $b_2$ are equal to the value that is being read in the working tape and the value that will be written in the working tape respectively.
	\item $c_1,c_2,c_2$ represent the instruction pointer in the procedure. Only 8 different instructions are needed in the simulation.
	\item Each value in memory of $x,y,z,u,v$ need $c$ elements to represent them and $i,j$ need only one. We map $(d_1\ldots,d_{c}) \to x$, $(d_{c+1}\ldots,d_{2c}) \to y$,
	$(d_{2c+1}\ldots,d_{3c}) \to z$, $(d_{3c+1}\ldots,d_{4c}) \to u$,
	$(d_{4c+1}\ldots,d_{5c}) \to v$, $d_{5c+1} \to i$ and $d_{5c+2}\to j$.
\end{enumerate}
For each transition $\delta \in \Delta \subseteq Q \times \{0,1\} \times \{0,1\} \times Q \times \{-1,=,+1\} \times \{0,1\} \times \{-1,=,+1\}$ we define a formula $\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$, where $\bar{x} = (x_1,\ldots,x_{3+c+\ell})$, $\bar{s} = (s_1,s_2)$, $\bar{w} = (w_1,w_2,w_3)$, $\bar{u} = (u_1,\ldots,u_{5c+2})$, $\bar{y} = (y_1,\ldots,y_{3+c+\ell})$, $\bar{t} = (t_1,t_2)$, $\bar{z} = (z_1,z_2,z_3)$ and $\bar{v} = (v_1,\ldots,v_{5c+2})$. The tuples $\bar{x}$ and $\bar{y}$ represent the current and next configuration of $M$ respectively, $\bar{s}$ and $\bar{y}$ indicate which algorithm is being followed, $\bar{w}$ and $\bar{z}$ represent the current and next instruction of the algorithm, $\bar{u}$ and $\bar{v}$ represent the current and next values in memory. We will describe the formula part by part. Suppose $\delta = (q_i,a,1,q_j,op_1,0,op_2)$, so we have to simulate Algorithm \ref{switch1to0}.

First we define some auxiliary formulas:
\begin{align*}
\varphi_0(x) &= \neg\exists y(y < x), \\
\varphi_{+1}(x,y) &= x < y \wedge \neg\exists z(x < z \wedge z < y), \\
\varphi_{-1}(x,y) &= \varphi_{+1}(y,x), \\
\varphi_1(x) &= \exists y(\varphi_0(y) \wedge \varphi_{+1}(y,x)), \\
\text{for $i,j \in\{0,1\}$, }	\varphi_{i,j}(x,y) &= \varphi_i(x) \wedge \varphi_j(y), \\
\text{for each $k \leq 7$, }\varphi^b_k(x_1,x_2,x_3) &= \varphi_{a_1}(x_1) \wedge \varphi_{a_2}(x_2) \wedge \varphi_{a_3}(x_3)\text{, where $a_1a_2a_3$ is the value of $k$ in binary}, \\
\mu_{<}(x_1,\ldots,x_c,y_1,\ldots,y_c) &= \bigvee_{i = 1}^c ((\bigwedge_{j = 1}^{i-1} x_j = y_j) \wedge x_i < y_i), \\
\mu_{+1}(\bar{x},\bar{y}) &= \mu_{<}(\bar{x},\bar{y}) \wedge \neg\exists \bar{z}(\mu_{<}(\bar{x},\bar{z}) \wedge \mu_{<}(\bar{z},\bar{y})), \\
\mu_{+2}(\bar{x},\bar{y}) &= \exists\bar{z}(\mu_{+1}(\bar{x},\bar{z}) \wedge \mu_{+1}(\bar{z},\bar{y})),\\
\mu_{-1}(\bar{x},\bar{y}) &= \varphi_{+1}(\bar{y},\bar{x}),\\
\mu_{-2}(\bar{x},\bar{y}) &= \varphi_{+2}(\bar{y},\bar{x}),\\
\mu_{0}(\bar{x}) &= \neg\exists\bar{y}(\mu_{<}(\bar{y},\bar{x})),\\
\mu_{1}(\bar{x}) &= \exists\bar{y}(\mu_{0}(\bar{y})\wedge\mu_{+1}(\bar{y},\bar{x})),\\
\varphi^2_{<}(x_1,x_2,y_1,y_2) &= x_1 < y_1 \vee (x_1 = y_1 \wedge x_2 < y_2),\\
\varphi^2_{+1}(x_1,x_2,y_1,y_2), \varphi^2_{-1}(x_1,x_2,y_1,y_2) &\text{ defined analogously to $\mu_{+1}$ and $\mu_{-1}$},\\
\text{for each } q_i\in Q, \varphi^q_i(x_1,\ldots,x_{\ell}) &= \bigwedge_{\substack{j = 1 \\ j \neq i}}^{\ell} \varphi_0(x_j) \wedge \varphi_1(x_i), \\
\varphi^E_0(x_1,x_2) &= \neg E(x_1,x_2),\\		\varphi^E_1(x_1,x_2) &= E(x_1,x_2),\\
\end{align*}

We start from instruction 0, which means that the procedure has not started yet and every value in the tuple is 0 except for the configuration values. It also initializes all the values in the tuple to 0 except for $x,u,i,j$.
\begin{multline*}
\varphi^{0,1}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{0,0}(s_1,s_2)\wedge\varphi^b_0(\bar{w}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_1(\bar{z})\, \wedge \\ 
\bigwedge_{i = 1}^c v_i = x_{3+i} \wedge \bigwedge_{i = c+1}^{2c} \varphi_0(v_i) \wedge \bigwedge_{i = 2c+1}^{3c} \varphi_0(v_i) \wedge \bigwedge_{i = 1}^c v_{3c+i} = x_{3+i} \wedge \bigwedge_{i = 4c+1}^{5c} \varphi_0(v_i) \wedge v_{5c+1} = x_3 \wedge v_{5c+2} = x_3.
\end{multline*}
Instruction 1 which checks whether the value of $j$ ($d_{5c+2}$ in the tuple) is more than 0 or not, and then proceeds to instruction 2 or 3 on each case.
\begin{multline*}
\varphi^{1,2}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_1(\bar{w}) \wedge \neg \varphi_0(u_{5c+2}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_2(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i, \\
\varphi^{1,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_1(\bar{w}) \wedge \varphi_0(u_{5c+2}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_3(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
\end{multline*}
Instruction 2 checks the value of $u$ ($d_{3c+1},\ldots,d_{4c}$ in the tuple). If it is $> 1$ then it substracts 2 from $u$ and adds 1 to $v$ ($d_{4c+1},\ldots,d_{5c}$ in the tuple), then repeats instruction 2. If it is equal to 0 or 1, then moves the value of $v$ to $u$, substracts 1 from $j$ and goes back to instruction 1.
\begin{multline*}
\varphi^{2,2}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_2(\bar{w}) \wedge \neg \mu_0(u_{3c+1},\ldots,u_{4c}) \wedge \neg \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_2(\bar{z})\, \wedge \\
\bigwedge_{i = 1}^c u_i = v_i \wedge
\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
\mu_{-2}(u_{3c+1},\ldots,u_{4c},v_{3c+1},\ldots,v_{4c})\, \wedge \\
\mu_{+1}(u_{4c+1},\ldots,u_{5c},v_{4c+1},\ldots,v_{5c}) \wedge u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}. \\
\varphi^{2,1}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_2(\bar{w}) \wedge ( \mu_0(u_{3c+1},\ldots,u_{4c}) \vee \mu_1(u_{3c+1},\ldots,u_{4c})) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_1(\bar{z})\, \wedge \\
\bigwedge_{i = 1}^c u_i = v_i \wedge
\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
\bigwedge_{i = 3c+1}^{4c} u_{c+i} = v_i \wedge
\mu_0(v_{4c+1},\ldots,v_{5c}) \wedge
u_{5c+1} = v_{5c+1} \wedge \varphi_{-1}(u_{5c+2},v_{5c+2}).	
\end{multline*}
Instruction 3 calculates the value of $u \mod 2$, that is, it repeats instruction 3 until the value of $u$ is equal to 0 or 1. On each iteration, it substracts 2 from $u$. Moreover, if the value of $u$ at the end of the iterations is not 1 then there is no step defined.
\begin{multline*}
\varphi^{3,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_3(\bar{w}) \wedge \neg \mu_0(u_{3c+1},\ldots,u_{4c}) \wedge \neg \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_3(\bar{z})\, \wedge \\
\bigwedge_{i = 1}^c u_i = v_i \wedge
\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
\mu_{-2}(u_{3c+1},\ldots,u_{4c},v_{3c+1},\ldots,v_{4c}) \wedge
\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}, \\
\varphi^{3,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_3(\bar{w}) \wedge \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_4(\bar{z})\, \wedge \\
\bigwedge_{i = 1}^c u_i = v_i \wedge
\mu_1(v_{c+1},\ldots,v_{2c}) \wedge
\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}
\end{multline*}
Instruction 4 checks the value of $i$ ($d_{5c+1}$ in the tuple.) If it is not 0 then goes to instruction 5 and if is 0 then goes to instruction 6. Moreover it initializes the value of $z$ ($d_{2c+1},\ldots,d_{3c}$ in the tuple) to 0 (which was 0 all along.)
\begin{multline*}
\varphi^{4,5}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_4(\bar{w}) \wedge \neg \varphi_0(u_{5c+1}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_5(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i, \\
\varphi^{4,6}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_4(\bar{w}) \wedge \varphi_0(u_{5c+1}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_6(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
\end{multline*}
Instruction 5 checks the value of $y$ ($d_{c+1},\ldots,d_{2c}$ in the tuple.) If it is more than 0 then it adds 2 to $z$ and substracts 1 from $y$, then repeats instruction 2. If it is not, then copies the value of $z$ to $y$ and subtracts 1 from $i$ and returns to instruction 4.
\begin{multline*}
\varphi^{5,5}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_5(\bar{w}) \wedge \neg \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_5(\bar{z}) \, \wedge \\
\bigwedge_{i = 1}^c u_i = v_i \wedge
\mu_{-1}(u_{c+1},\ldots,u_{2c},v_{c+1},\ldots,v_{2c}) \wedge
\mu_{+2}(u_{2c+1},\ldots,u_{3c},v_{2c+1},\ldots,v_{3c})\, \wedge \\
\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}, \\
\varphi^{5,4}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_5(\bar{w}) \wedge \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_4(\bar{z}) \, \wedge \\
\bigwedge_{i = 1}^c u_i = v_i \wedge
\bigwedge_{i = c+1}^{2c} u_{c+i} = v_i \wedge
\mu_0(v_{2c+1},\ldots,v_{3c}) \wedge
\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
\varphi_{-1}(u_{5c+1},v_{5c+1}) \wedge u_{5c+2} = v_{5c+2}
\end{multline*}
Instruction 6 checks the value of $y$. If it is more than 0, then subtracts 1 from $x$ and $y$ and repeats instruction 6. If it is not, then goes to instruction 7.
\begin{multline*}
\varphi^{6,6}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_6(\bar{w}) \wedge \neg \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_6(\bar{z}) \, \wedge \\
\mu_{-1}(u_1,\ldots,u_c,v_i,\ldots,v_c) \wedge \mu_{-1}(u_{c+1},\ldots,u_{2c},v_{c+1},\ldots,v_{2c}) \wedge \bigwedge_{i = 3c+1}^{5c+2} u_i = v_i]\,\vee \\
\varphi^{6,7}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_6(\bar{w}) \wedge \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_7(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
\end{multline*}
Instruction 7 stores the value of $x$ after the corresponding bit has been switched. Then we can define $\varphi_{\delta}$ which also simulates the actual transition. If $u$ equals 1, then copy what is stored in $x$ to $a_4,\ldots,a_{3+c}$, go from state $q_i$ to state $q_j$, and move the heads to their corresponding positions.
\begin{multline*}
\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 	[\bigwedge_{i = 1}^{3+c+\ell} x_i = y_i \wedge (\varphi^{0,1}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{1,2}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})\, \vee \\ \varphi^{1,3}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{2,2}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{2,1}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{3,3}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \\ \varphi^{3,4}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{4,5}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{4,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{5,5}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \\ \varphi^{5,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{6,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{6,7}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}))] \, \vee \\
[\varphi^E_a(x_1,x_2) \wedge \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_7(\bar{w}) \wedge \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{0,0}(t_1,t_2) \wedge \varphi^b_0(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i \,\wedge \\
\varphi^2_{op_1}(x_1,x_2,y_1,y_2) \wedge \varphi_{op_2}(x_3,y_3) \wedge \bigwedge_{i = 1}^c u_i = x_{3+i} \wedge \varphi^q_i(x_{4+c},\ldots,x_{3+c+\ell}) \wedge \varphi^q_j(y_{4+c},\ldots,y_{3+c+\ell})].
\end{multline*}

\begin{algorithm}
	\caption{If the $i$-th bit in $x$ is 0 return $x$} \label{switch0to0}
	\begin{algorithmic}
		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
		\While{$j > 0$}
		\State $v \gets 0$
		\While{$u > 1$}
		\State $u \gets u-2,\; v \gets v+1$
		\EndWhile
		\State $u\gets v,\; j \gets j-1$
		\EndWhile
		\While{$u > 1$}
		\State $u \gets u-2$
		\EndWhile
		\State $\textbf{assert } u = 0$ \Comment{If $u \neq 0$ simply stop}	
		\State \Return $x$.
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}
	\caption{If the $i$-th bit in $x$ is 0 replace it by 1 and return the result}
	\label{switch0to1}
	\begin{algorithmic}
		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
		\While{$j > 0$}
		\State $v \gets 0$
		\While{$u > 1$}
		\State $u \gets u-2,\; v \gets v+1$
		\EndWhile
		\State $u\gets v,\; j \gets j-1$
		\EndWhile
		\While{$u > 1$}
		\State $u \gets u-2$
		\EndWhile
		\State $\textbf{assert } u = 0$ \Comment{If $u \neq 0$ simply stop}	
		\State $y \gets 1$ \Comment{Compute $2^i$ and store it in $y$}
		\While{$i > 0$}
		\State $z \gets 0$
		\While{$y > 0$}
		\State $z \gets z+2,\; y \gets y-1$
		\EndWhile
		\State $i \gets i-1,\; y \gets z$
		\EndWhile
		\While{$y > 0$} \Comment{Add $y$ to $x$}
		\State $x \gets x+1,\; y \gets y-1$
		\EndWhile
		\State \Return $x$.
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}
	\caption{If the $i$-th bit in $x$ is 1 return $x$}
	\label{switch1to1}
	\begin{algorithmic}
		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
		\While{$j > 0$}
		\State $v \gets 0$
		\While{$u > 1$}
		\State $u \gets u-2,\; v \gets v+1$
		\EndWhile
		\State $u\gets v,\; j \gets j-1$
		\EndWhile
		\While{$u > 1$}
		\State $u \gets u-2$
		\EndWhile
		\State $\textbf{assert } u = 1$ \Comment{If $u \neq 0$ simply stop}	
		\State \Return $x$.
	\end{algorithmic}
\end{algorithm}
Note that we also need to specify that the program we are following is Algorithm \ref{switch1to0} so we store $1,0$ in $b_1,b_2$ all along the procedure. We describe the three other algorithms that compute the switches from $0\to 0$, $0\to 1$ and $1\to 1$ (Algorithms \ref{switch0to0}, \ref{switch0to1} and \ref{switch1to1}.)
For the other three cases, where $\delta = (q_i,a,0,q_j,op_1,0,op_2)$, $\delta = (q_i,a,0,q_j,op_1,1,op_2)$ and $\delta = (q_i,a,1,q_j,op_1,1,op_2)$, $\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$ is defined analogously. Then, $\varphi(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$ is defined as:
$$
\varphi(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \bigvee_{\delta \in \Delta} \varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}).
$$
Finally we define $\varphi_I$ and $\varphi_F$:
\begin{align*}
\varphi_I(\bar{x},\bar{s},\bar{w},\bar{u}) &= \varphi_0(x_1) \wedge \varphi_0(x_2) \wedge \varphi_0(x_3) \wedge \bigwedge_{i = 4}^{3+c}\varphi_0(x_i) \wedge \varphi_1(x_{4+c}) \wedge \bigwedge_{i = 5+c}^{3+c+\ell} \varphi_0(x_i)\wedge \varphi_{0,0}(s_1,s_2)\ \wedge \varphi^b_0(\bar{w}) \wedge \bigwedge_{i = 1}^{5c+2}\varphi_0(u_i). \\
\varphi_F(\bar{x},\bar{s},\bar{w},\bar{u}) &= \varphi^q_{\ell}(x_{4+c},\ldots,x_{3+c+\ell}) \wedge \varphi_{0,0}(s_1,s_2)\ \wedge \varphi^b_0(\bar{w}) \wedge \bigwedge_{i = 1}^{5c+2}\varphi_0(u_i),
\end{align*}
and then $\sem{\alpha}(\A) = \sem{\sa{\bar{x}}\sa{\bar{y}}([\pth \varphi(\bar{u},\bar{v})](\bar{x},\bar{y})\cdot \varphi_I(\bar{x})\cdot\varphi_F(\bar{y}))}(\A) = \acc_M(\A)$.

\textbf{Proof of Theorem \ref{tqso-fo-cap}}

\textbf{Proof of Theorem \ref{rqfo-fo-cap}}

### 2016-12-02 LICS/beyond.tex
paper v.1

@@
 where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
-\begin{theorem}
+\begin{theorem} \label{so-rec}
 	Given a positive $\fo$ formula $\varphi(\bar{x},R)$ and a $\qfo$ formula $\alpha(\bar{x})$, there exists a $\qso$ formula $\beta(\bar{x})$ such that $\sem{[\alfp\varphi(\bar{x},R)\mid \alpha(\bar{x},R)](\bar{x})} = \sem{\beta(\bar{x})}$.
@@
 
-\begin{theorem}
+\begin{theorem} \label{tqfo-fo-cap}
 	$\tqfo(\fo)$ captures $\shl$ over the class of ordered structures.
@@
 
-\begin{theorem}
+\begin{theorem} \label{tqso-fo-cap}
 	$\tqso(\fo)$ captures $\shp$ over the class of ordered structures.
@@
 
-\begin{theorem}
+\begin{theorem} \label{rqfo-fo-cap}
 	$\rqfo(\fo)$ captures $\fp$ over the class of ordered structures.

### 2016-12-02 LICS/counting.tex
paper v.1

@@
 
-\begin{theorem}
+\begin{theorem} \label{qfo-pfp-cap}
 	$\qfo(\pfp)$ captures $\nfpspace$ over the class of ordered structures.
@@
 
-\begin{theorem}
-	$\tqso(\fo)$ captures $\shp$ over the class of ordered structures.
-\end{theorem}
-
-\begin{theorem}
-	($\qfo(\lfp)$ or $\qfo(\fo)$) with algebraic least fixed point operator captures $\totp$ over the class of ordered structures.
-\end{theorem}
-
-\begin{theorem}
+\begin{theorem} \label{eqso-fo-cap}
 	$\eqso(\fo)$ captures $\shp$ over the class of ordered structures.
@@
 
-\begin{proposition}
+\begin{proposition} \label{eqso-fo-cap-not}
 	$\eqso(\fo)$ does not capture $\shp$ over the class of all structures.
@@
 
-\begin{theorem}
-	$\eqso(\integ,\fo)$ captures $\gp$ over the class of ordered structures.
-\end{theorem}
-
-\begin{theorem}
+\begin{theorem} \label{eqso-eso-cap}
 	$\eqso(\eso)$ captures $\spp$ over the class of ordered structures.

### 2016-12-02 LICS/syntactic.tex
paper v.1

@@
 
-\begin{theorem}
+\begin{theorem} \label{sub-pnp}
 	If $\E1$ is closed under substraction, then {\sc P} = {\sc NP}.
@@
 
-\begin{theorem}
+\begin{theorem} \label{sigma1strict}
 	$\#\Sigma_1 \subsetneq \E1$
@@
 
-\begin{theorem}
+\begin{theorem} \label{sigmafo-minusone}
 	$\XE{1}$ is closed under substraction by one.
@@
 
-\begin{theorem}
+\begin{theorem} \label{dnf-pars}
 	{\sc \#DNF} is hard for $\E1$ under parsimonious reductions. 
@@
 
-\begin{theorem}
+\begin{theorem} \label{nplusone-strict}
 	$\U{1}$ with $n$ open first-order variables is properly contained in $\U{1}$ with $n+1$ open first-order variables for $n\in\N$.  
@@
 
-\begin{theorem}
+\begin{theorem} \label{sigma2-pe}
 	$\HE{2} \subseteq \pe$

### 2016-12-05 LICS/appendix.tex
standarizing saluja

@@
 %!TEX root = main.tex
+
+\textbf{Proof of Theorem \ref{no-mult}}
+
+Let $\{+, \cdot, \Sigma, \Pi\}$ be the set of {\it algebraic operators} in $\qso$.
+We will prove by induction over the number of algebraic operators in a $\qso$-formula $\alpha$ that there exists some $\qso$-formula $\alpha'$ that respects the stated condition.
+Let $\text{op}(\alpha)$ be the number of algebraic operators in $\alpha$.
+The case $\text{op}(\alpha) = 0$ is trivial.
+Let $n\in\nat$ be such that for each $\qso$-formula $\beta$ such that $\text{op}(\beta) < n$ there is some $\beta'$ that respects the condition.
+Let $\alpha$ be a $\qso$-formula such that $\text{op}(\alpha) = n$.
+Note that if $\alpha$ is of the form $\alpha = (\beta_1 + \beta_2)$, $\alpha = \Sigma x \beta$, $\alpha = \Pi x \beta$, $\alpha = \Sigma X \beta$ or $\alpha = \Pi X \beta$ the proof follows directly.
+Therefore we assume that $\alpha = (\beta_1 \cdot \beta_2)$.
+We separate the proof in two main cases:
+\begin{enumerate}
+	\item At least one of $\beta_1, \beta_2$ is not of the form $\Pi x \gamma$, $\Pi X \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. Without loss of generality, let $\beta_1$ be such one.
+	\begin{enumerate}
+		\item $\beta_1 = s \in \nat$.
+Then we use $\alpha' = \beta_2 + \cdots + \beta_2$, where $\beta_2$ is repeated $s$ times, and if $s = 0$, take $\alpha' = \perp$.
+Since each of the formulas paired by $+$ has less than $n$ algebraic operators, the statement holds.
+		\item $\beta_1 = (\gamma_1 + \gamma_2)$.
+Then, we use $\alpha' = (\gamma_1\cdot\beta_2 + \gamma_2\cdot\beta_2)$ and each of the formulas paired by $+$ have less than $n$ algebraic operators so the statement holds.
+		\item $\beta_1 = \Sigma x \gamma$. Then, we use $\alpha' = \Sigma x (\gamma\cdot\beta_2)$ and as in the previous case the statement holds.
+		\item $\beta_1 = \Sigma X \gamma$, which we treat exactly as the previous one.
+	\end{enumerate}
+	\item Both of $\beta_1$ and $\beta_2$ are of the form $\Pi x \gamma$, $\Pi X \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula.
+This case also separates in further cases.
+We define $(\varphi \mapsto \alpha) := (\neg\varphi + (\varphi\cdot\alpha))$ for which $\sem{(\varphi \mapsto \alpha)}(\A,v,V)$ equals $\sem{\alpha}(\A,v,V)$ when $(\A,v,V)\models\varphi$ and 1 otherwise.
+	\begin{enumerate}
+		\item $\beta_1 = \Sigma x \gamma_1(x)$ and $\beta_2 = \Sigma y \gamma_2(y)$. Then we use $\alpha' = \Sigma x(\gamma_1(x)\cdot\gamma_2(x))$, and the statement holds.
+		\item $\beta_1 = \Sigma X \gamma_1(X)$ and $\beta_2 = \Sigma Y \gamma_2(Y)$. [something about arities].
+		\item $\beta_1 = \Sigma X \gamma_1(X)$ and $\beta_2 = \Sigma x \gamma_2(x)$ (w.l.o.g.). [use singleton predicates].
+		\item $\beta_1 = \Sigma x \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). Then, we use $\alpha' = \Sigma x (\gamma \cdot \varphi)$, and the statement holds.
+		\item $\beta_1 = \Sigma X \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). We treat this case exactly as the previous one.
+		\item $\beta_1 = \varphi_1$ and $\beta_2 = \varphi_2$ both $\so$-formulas. Then, we use $\alpha' = (\varphi_1 \wedge \varphi_2)$ and the statement holds.
+	\end{enumerate}
+\end{enumerate}
 
@@
 
+\textbf{Proof of Theorem \ref{saluja-eq}}
+
 \textbf{Proof of Theorem \ref{fp1}}
@@
 
-Let $f$ be a function in $\E1$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
+Let $f$ be a function in $\E{1}$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
 \begin{eqnarray*}
@@
 
-Suppose that $\E1$ is closed under substraction, that is, for each pair of functions $f,g\in \E1$, there exists $h\in\E1$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\str$.
+Suppose that $\E{1}$ is closed under substraction, that is, for each pair of functions $f,g\in \E{1}$, there exists $h\in\E{1}$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\str$.
 
@@
 Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Given that $f_{\#3DNF}, f_{all} \in \E1$, we have by our initial assumption that $f_{all}-f_{\#3DNF} \in \E1$.
-Let $h\in\E1$ be such that $h = f_{all}-f_{\#3DNF}$.
+Given that $f_{\#3DNF}, f_{all} \in \E{1}$, we have by our initial assumption that $f_{all}-f_{\#3DNF} \in \E{1}$.
+Let $h\in\E{1}$ be such that $h = f_{all}-f_{\#3DNF}$.
 For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is $\conp$-complete.
-However, as we showed previously in Theorem \ref{decisionptime}, since $h\in\E1$, we have that $L_h \in \ptime$.
+However, as we showed previously in Theorem \ref{decisionptime}, since $h\in\E{1}$, we have that $L_h \in \ptime$.
 Then, $\conp \subseteq \ptime$, from which we conclude that $\ptime = \np$.
@@
 
-We will show that the $\E1$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
+We will show that the $\E{1}$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
 By contradiction, suppose that it is.
@@
 \begin{enumerate}
-	\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
+	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
 	\begin{eqnarray*}
@@
 Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$.
-Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E1.$
+Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E{1}.$
 
-	\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
+	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
 	\begin{eqnarray*}
@@
 It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
-And so, for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Hence, from the results in the previous paragraph, if $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}$, we have that $f^\prime = f\dotminus 1\in\E1$
+And so, for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Hence, from the results in the previous paragraph, if $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}$, we have that $f^\prime = f\dotminus 1\in\E{1}$
 
-	\item Let $f \in \E1$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
+	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
 	\begin{eqnarray*}
@@
 	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$.
-In order to prove that $f\dotminus 1\in\E1$, we define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2, where
+In order to prove that $f\dotminus 1\in\E{1}$, we define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2, where
 	\begin{eqnarray*}
@@
 
-Let $f$ be an arbitrary function in $\E1$ and let $f_{\textsc{\#DNF}}$ be the function that defines {\sc \#DNF}.
+Let $f$ be an arbitrary function in $\E{1}$ and let $f_{\textsc{\#DNF}}$ be the function that defines {\sc \#DNF}.
 We will now show a function $h:\str\to \textsc{Struct}[\L_{\textsc{DNF}}]$, where $\L_{\textsc{DNF}} = \{N,P\}$, such that for every $\A\in\str$, it holds that $f(\A) = f_{\textsc{\#DNF}}(h(\A))$, and which can be computed in polynomial time.

### 2016-12-05 LICS/horn-counting.tex
standarizing saluja

--- 2016-12-01 LICS/syntactic.tex
@@
-%!TEX root = main.tex
-
-\subsection{Saluja Classes}
-
-We define syntactically the classes $\XE{i}$ and $\XU{i}$ according to the following grammar:
-\begin{enumerate}
-	\item $\XE{0}$:
-	\[
-	E_0 ::= \varphi , \varphi \mbox{ is an {\sc FO}-extended quantifier-free $\L$-formula }.
-	\]
-	\item $\XU{0}$:
-	\[U_0 ::= E_0.\]
-	\item $\XE{i+1}$:
-	\[	E_{i+1} ::= U_i \ \mid \ \exists x \, E_{i+1}.\]
-	\item $\XU{i+1}$:
-	\[U_{i+1} ::= E_i \ \mid \ \forall x \, U_{i+1}.\]
-\end{enumerate}
-
-A function $f$ is in $\XE{i}$ (resp. $\XU{i}$) if there is an $\L$-formula $\varphi$ defined by the grammar $E_i$ (resp. $U_i$) such that $f = f_{\varphi}$.
-
-\begin{theorem} \label{fp1}
-	$\XE{0} \subseteq$ {\sc FP}.
-\end{theorem}
-
-The {\em decision problem} associated to a function $f$ is defined by the language $L_f = \{\A \in \str \mid f(\A) > 0\}$.
-
-\begin{theorem} \label{decisionptime}
-	The decision problem associated to a function in $\E1$ is in \textsc{P}.
-\end{theorem}
-
-For a given pair of functions $f,g$, we define $f \dotminus g$ as follows:
-\begin{eqnarray*}
-	(f \dotminus g)(\A) =
-	\begin{cases}
-		f(\A)-g(\A), & \text{if }f(\A)>g(\A) \\
-		0, & \text{if }f(\A) \leq g(\A).
-	\end{cases}
-\end{eqnarray*}
-for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction} if for every pair of functions $f,g \in \F$, it holds that $f \dotminus g \in \F$.
-
-\begin{theorem}
-	If $\E1$ is closed under substraction, then {\sc P} = {\sc NP}.
-\end{theorem}
-
-\begin{theorem}
-	$\#\Sigma_1 \subsetneq \E1$
-\end{theorem}
-
-For a given function $f$, we define $f \dotminus 1$ as follows:
-\begin{eqnarray*}
-	f \dotminus 1(\A) =
-	\begin{cases}
-		f(\A)-1, & \text{if }f(\A) > 0 \\
-		0, & \text{if }f(\A) = 0.
-	\end{cases}
-\end{eqnarray*}
-for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
-
-\begin{theorem}
-	$\XE{1}$ is closed under substraction by one.
-\end{theorem}
-
-\begin{theorem}
-	{\sc \#DNF} is hard for $\E1$ under parsimonious reductions. 
-\end{theorem}
-
-\begin{theorem}
-	$\U{1}$ with $n$ open first-order variables is properly contained in $\U{1}$ with $n+1$ open first-order variables for $n\in\N$.  
-\end{theorem}
-
-\subsection{Horn Counting Classes}
-
-We define the syntactic classes $\XE{i}$ and $\XU{i}$ as follows. First we define extended Horn clauses.
+We define the fragments $\logeh{i}$ and $\loguh{i}$ as follows. First we define extended Horn clauses.
 \begin{eqnarray*}
@@
 	NL &::=& \neg X_i(\x),\, i\in\N \ \mid \ \exists x\, NL,\\
-	NC &::=& NL \ \mid \ \alpha, \alpha \mbox{ is an {\sc FO}-formula over } \L \ \mid \ (NC \vee NC),\\
+	NC &::=& NL \ \mid \ \alpha, \alpha \mbox{ is an {\sc FO}-formula} \ \mid \ (NC \vee NC),\\
 	HC &::=& NC \ \mid \ (NC \vee PL) \ \mid \ PL,
@@
 where $PL$ represents a positive literal, $NL$ is an \textit{extended} negative literal, $NC$ is an extended Horn clause, and $HC$ is an extended Horn formula.
-Now we define the syntax of the classes inductively.\textbf{}
-\begin{enumerate}
-	\item $\HE{0}$:
-	\begin{eqnarray*}
-		E_0 &::=& HC \ \mid \ E_0 \wedge E_0.
-	\end{eqnarray*}
-	\item $\HU{0}$:
-	\begin{eqnarray*}
-		U_0 &::=& E_0.
-	\end{eqnarray*}
-	\item $\HE{i+1}$:
-	\begin{eqnarray*}
-		E_{i+1} &::=& U_i \ \mid \ \exists x \, E_{i+1}.
-	\end{eqnarray*}
-	\item $\HU{i+1}$:
-	\begin{eqnarray*}
-		U_{i+1} &::=& E_i \ \mid \ \forall x \, U_{i+1}.
-	\end{eqnarray*}
-\end{enumerate}
-A function $f$ is in $\XE{i}$ (resp. $\XU{i}$) if there is an $\L$-formula $\varphi$ defined by the grammar $E_i$ (resp. $U_i$) such that $f = f_{\varphi}$.
-
-The class $\pe$ is defined in \cite{DBLP:conf/mfcs/PagourtzisZ06} as $\pe = \{f\mid f\in\#P \mbox{ and its decision version } L_{f}\in P\}$.
-
-\begin{theorem}
-	$\HE{2} \subseteq \pe$
+Now we define the syntax of the classes inductively.
+\begin{align*}
+\logeh{0} &::= HC \ \mid \ \logeh{0} \wedge \logeh{0}, \\
+\loguh{0} &::= \logeh{0}, \\ 
+\logeh{i+1} &::= \loguh{i} \ \mid \ \exists x\, \logeh{i+1}, \\
+\loguh{i+1} &::= \logeh{i} \ \mid \ \forall x\, \loguh{i+1}.
+\end{align*}
+\begin{theorem} \label{sigma2-pe}
+	$\eqso(\loguh{2}) \subseteq \pe$
 \end{theorem}
@@
 \begin{theorem} \label{sigma2hard}
-	$\shdhsat$ is hard for $\XE{2}$ under parsimonious reductions. 
+	$\shdhsat$ is hard for $\eqso(\loguh{2})$ under parsimonious reductions. 
 \end{theorem}

### 2016-12-05 LICS/preliminaries.tex
standarizing saluja

@@
 
-Here come the preliminaries
+A $\mathbb{N}$ relational signature $\R$ is a finite set $\{R_1, \ldots, R_k\}$, where each $R_i$ ($1 \leq i \leq k$) is a relation name with an associate arity greater than 0, which is denoted by $\arity(R_i)$.
+A finite structure over $\R$ (or just finite $\R$-structure) is a tuple $\A = \langle A, R_1^\A, \ldots, R_k^\A \rangle$ such that $A$ is a finite set and $R_i^\A \subseteq A^{\arity(R_i)}$ for every $i \in \{1, \ldots, k\}$.
+A finite $\R$-structure $\A$ is said to be ordered if $<$ is a binary predicate name in $\R$ and $<^\A$ is a linear order on $A$.
+Let $\str[\R]$ be the class of all finite $\R$-structures and $\ostr[\R]$ be the class of all ordered finite $\R$-structures.
+
+From now on, assume given disjoint infinite sets $\fv$ and $\sv$ of first-order variables and second-order variables, respectively.
+Notice that every variable in $\sv$ has an associated arity, which is denoted by $\arity(X)$.
+Then given a relational signature $\R$, the set of second-order logic formulas ($\so$-formulas) over $\R$ is given by the following grammar:
+\begin{eqnarray*}\ 
+	\varphi &:=& R(\bar u) \ \mid\  
+	X(\bar v)  \ \mid\ 
+	\neg \varphi \ \mid\ 
+	(\varphi \vee \varphi) \ \mid\ 
+	\exists x \, \varphi \ \mid\ 
+	\exists X \, \varphi
+\end{eqnarray*}
+where $R \in \R$, $\bar u$ is a tuple of (non-necessarily distinct) variables from $\fv$ whose length is $\arity(R)$, $X \in \sv$, $\bar v$ is a tuple of (non-necessarily distinct) variables from $\fv$ whose length is $\arity(X)$, and $x \in \fv$.
+
+To define the semantics of $\so$, we need to introduce some terminology.
+Given a relational signature $\R$ and a finite $\R$-structure $\A$ with domain $A$, a first-order assignment $v$ for $\A$ is a total function from $\fv$ to $A$, while a second-order assignment $V$ for $\A$ is a total function with domain $\sv$ that maps each $X \in \sv$ to a subset of $A^{\arity(X)}$.
+Moreover, given a first-order assignment $v$ for $\A$, $x \in \fv$ and $a \in A$, we denote by $v[a/x]$ a first-order assignment such that $v[a/x](x) = a$ and $v[a/x](y) = v(y)$ for every $y \in \fv$ distinct from $x$.
+Similarly, given a second-order assignment $V$ for $\A$, $X \in \sv$ and $B  \subseteq A^{\arity(X)}$, we denote by $V[B/X]$ a second-order assignment such that $V[B/X](X) = B$ and $V[B/X](Y) = V(Y)$ for every $Y \in \sv$ distinct from $X$.
+
+Assume that $\varphi$ is an $\so$-formula over a signature $\R$.
+Then given a finite $\R$-structure $\A$ with domain $A$, a first-order assignment $v$ for $\A$ and a second-order assignment $V$ for $\A$, we say that $(\A, v, V)$ satisfies $\varphi$, denoted by $(\A, v, V) \models \varphi$, if: (1) $\varphi$ is the formula $R(x_1, \ldots, x_\ell)$ and $(v(x_1), \ldots, v(x_\ell)) \in R^\A$; (2) $\varphi$ is the formula $X(x_1, \ldots, x_m)$ and $(v(x_1), \ldots, v(x_m)) \in V(X)$; (3) $\varphi$ is the formula $\neg \psi$ and it not the case that $(\A, v, V) \models \psi$; (4) $\varphi$ is the formula $(\varphi_1 \vee \varphi_2)$, and $(\A, v, V) \models \varphi_1$ or $(\A, v, V) \models \varphi_2$; (5) $\varphi$ is the formula $\exists x \, \psi$ and there exists $a \in A$ such that $(\A, v[a/x], V) \models \psi$; or (6) $\varphi$ is the formula $\exists X \, \psi$ and there exists $B \subseteq A^{\arity(X)}$ such that $(\A, v, V[B/X]) \models \psi$.
+As usual, we consider the propositional operators $\wedge$, $\rightarrow$, and $\leftrightarrow$ that can be obtained from $\vee$ and $\neg$.
+%Moreover, we use the abbreviations $x \not \leq y$ and $x \notin X$ for the negation of the atoms $\leq$ and $\in$. 
+%Finally, we consider standard abbreviations of formulas that can be defined in $\so$-logic (actually, $\fo$-logic) like $\first(x) := \fa{x} y \leq x$ and $\last(x) := \fa{x} x \leq y$ to denote the first and last element of the linear order $\leq$, respectively, and $\succesor(x,y) := x \leq y \wedge y \not\leq x \wedge \fa{z} ( z \leq x \vee y \leq z)$ to denote the successor relation.
+
+\subsection{Function complexity classes}
+Assume that $\R = \{R_1, \ldots, R_k\}$ is a relational signature and $\A$ is a finite $\R$-structure with a domain $A$ containing $n$ elements, and assume that  $<$ is a linear order on $A$, say $a_1 < a_2 < \ldots < a_n$.
+For every $i \in \{1, \ldots, k\}$, define the encoding of $R_i^\A$, denoted by $\enc(R_i^\A)$, as the following binary string.
+Assume that $\ell = \arity(R_i)$ and consider an enumeration of the $\ell$-tuples over $A$ in the lexicographic order induced by $<$ (that is, $(a_1, \ldots, a_1, a_1)$, $(a_1, \ldots, a_1, a_2)$, $\ldots$, $(a_n, \ldots, a_n, a_{n-1})$, $(a_n, \ldots, a_n, a_n)$).
+Then let $\enc(R_i^\A)$ be a binary string of length $n^\ell$ such that the $i$-th bit of $\enc(R_i^\A)$ is 1 if the $i$-th tuple in the previous enumeration belongs to $R_i^\A$, and 0 otherwise.
+Moreover, define the encoding of $\A$, denoted by $\enc(\A)$, as the following binary string~\cite{L04}:
+\begin{eqnarray*}
+	\enc(\A) & = & 0^n \, 1 \, \enc(R_1^\A) \, \cdots \, \enc(R_k^\A).
+\end{eqnarray*}
+Given a set $\bbD$ and a relational signature $\R$, a function $f$ is said to be from $\R$ to $\bbD$ if $f$ is a total function from $\{\enc(\A) \mid \A \in \str[\R]\}$ to $\bbD$.
+Moreover, a set $\CC$ of functions over $\bbD$ is said to be a {\em function complexity class over $\bbD$} if every $f \in \CC$ is a function from some relational signature $\R$ to $\bbD$.
+
+If $\KK$ is a class of finite structures and $f$ is a function from a relational signature $\R$ to $\bbD$, we denote by $\res{f}{\KK}$ the restriction of $f$ to $\KK$, that is, a function such that: the domain of $\res{f}{\KK}$ is $\{ \enc(\A) \mid \A \in \str[\R] \cap \KK\}$, and
+$\res{f}{\KK}(\enc(\A)) = f(\enc(\A))$ for every $\A \in \str[\R] \cap \KK$.

### 2016-12-05 LICS/saluja.tex
standarizing saluja

Saluja et. al \cite{DBLP:journals/jcss/SalujaST95} define a family of counting classes $\#\cL$ for each fragment $\cL$ of $\fo$. For a formula $\varphi(x,X)$, the function $f_{\varphi(x,X)}$ is defined as
\[
f_{\varphi(x,X)}(\A) = \vert \{\langle e,P\rangle\mid \A\models\varphi(e,P)\}\vert.
\]
for each $\A\in\str$. A function $f:\Sigma^*\to\nat$ is in $\#\cL$ if there exists an $\cL$ formula $\varphi(x,X)$ such that $f = f_{\varphi(x,X)}$.

\begin{theorem} \label{saluja-eq}
	$\eqso(\cL) = \#\cL$ for each fragment $\L$ of $\fo$.
\end{theorem}

For every logic $\cL$, we define an $\cL$-extended quantifier-free (QF) formula as follows:
\begin{eqnarray*}
	\varphi &::=& \alpha, \alpha \text{ is an $\cL$-formula} \ \mid \\
	&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
	&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
\end{eqnarray*}

We define syntactically the fragments $\logex{i}$ and $\logux{i}$ according to the following grammar:
\begin{align*}
\logex{0} = \logux{0} &::= \varphi , \varphi \mbox{ is an $\fo$-extended QF formula,} \\
\logex{i+1} &::= \logux{i} \ \mid \ \exists x\, \logex{i+1}, \\
\logux{i+1} &::= \logex{i} \ \mid \ \forall x\, \logux{i+1}.
\end{align*}

\begin{theorem} \label{fp1}
	$\qfo(\logex{0}) \subseteq$ {\sc FP}.
\end{theorem}

The {\em decision problem} associated to a function $f$ is defined by the language $L_f = \{\A \in \str \mid f(\A) > 0\}$.

\begin{theorem} \label{decisionptime}
	The decision problem associated to a function in $\eqso(\logex{1})$ is in \textsc{P}.
\end{theorem}

For a given pair of functions $f,g$, we define $f \dotminus g$ as follows:
\begin{eqnarray*}
	(f \dotminus g)(\A) =
	\begin{cases}
		f(\A)-g(\A), & \text{if }f(\A)>g(\A) \\
		0, & \text{if }f(\A) \leq g(\A).
	\end{cases}
\end{eqnarray*}
for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction} if for every pair of functions $f,g \in \F$, it holds that $f \dotminus g \in \F$.

\begin{theorem} \label{sub-pnp}
	If $\eqso(\loge{1})$ is closed under substraction, then {\sc P} = {\sc NP}.
\end{theorem}

\begin{theorem} \label{sigma1strict}
	$\eqso(\loge{1}) \subsetneq \eqso(\logex{1})$
\end{theorem}

For a given function $f$, we define $f \dotminus 1$ as follows:
\begin{eqnarray*}
	f \dotminus 1(\A) =
	\begin{cases}
		f(\A)-1, & \text{if }f(\A) > 0 \\
		0, & \text{if }f(\A) = 0.
	\end{cases}
\end{eqnarray*}
for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.

\begin{theorem} \label{sigmafo-minusone}
	$\eqso(\logex{1})$ is closed under substraction by one.
\end{theorem}

\begin{theorem} \label{dnf-pars}
	{\sc \#DNF} is hard for $\eqso(\loge{1})$ under parsimonious reductions. 
\end{theorem}

\begin{theorem} \label{nplusone-strict}
	$\U{1}$ with $n$ open first-order variables is properly contained in $\U{1}$ with $n+1$ open first-order variables for $n\in\N$.  
\end{theorem}

### 2016-12-05 LICS/syntactic.tex
standarizing saluja

--- 2016-12-01 LICS/syntactic.tex
@@
 
-We define syntactically the classes $\XE{i}$ and $\XU{i}$ according to the following grammar:
-\begin{enumerate}
-	\item $\XE{0}$:
-	\[
-	E_0 ::= \varphi , \varphi \mbox{ is an {\sc FO}-extended quantifier-free $\L$-formula }.
-	\]
-	\item $\XU{0}$:
-	\[U_0 ::= E_0.\]
-	\item $\XE{i+1}$:
-	\[	E_{i+1} ::= U_i \ \mid \ \exists x \, E_{i+1}.\]
-	\item $\XU{i+1}$:
-	\[U_{i+1} ::= E_i \ \mid \ \forall x \, U_{i+1}.\]
-\end{enumerate}
-
-A function $f$ is in $\XE{i}$ (resp. $\XU{i}$) if there is an $\L$-formula $\varphi$ defined by the grammar $E_i$ (resp. $U_i$) such that $f = f_{\varphi}$.
-
-\begin{theorem} \label{fp1}
-	$\XE{0} \subseteq$ {\sc FP}.
-\end{theorem}
-
-The {\em decision problem} associated to a function $f$ is defined by the language $L_f = \{\A \in \str \mid f(\A) > 0\}$.
-
-\begin{theorem} \label{decisionptime}
-	The decision problem associated to a function in $\E1$ is in \textsc{P}.
-\end{theorem}
-
-For a given pair of functions $f,g$, we define $f \dotminus g$ as follows:
-\begin{eqnarray*}
-	(f \dotminus g)(\A) =
-	\begin{cases}
-		f(\A)-g(\A), & \text{if }f(\A)>g(\A) \\
-		0, & \text{if }f(\A) \leq g(\A).
-	\end{cases}
-\end{eqnarray*}
-for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction} if for every pair of functions $f,g \in \F$, it holds that $f \dotminus g \in \F$.
-
-\begin{theorem}
-	If $\E1$ is closed under substraction, then {\sc P} = {\sc NP}.
-\end{theorem}
-
-\begin{theorem}
-	$\#\Sigma_1 \subsetneq \E1$
-\end{theorem}
-
-For a given function $f$, we define $f \dotminus 1$ as follows:
-\begin{eqnarray*}
-	f \dotminus 1(\A) =
-	\begin{cases}
-		f(\A)-1, & \text{if }f(\A) > 0 \\
-		0, & \text{if }f(\A) = 0.
-	\end{cases}
-\end{eqnarray*}
-for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
-
-\begin{theorem}
-	$\XE{1}$ is closed under substraction by one.
-\end{theorem}
-
-\begin{theorem}
-	{\sc \#DNF} is hard for $\E1$ under parsimonious reductions. 
-\end{theorem}
-
-\begin{theorem}
-	$\U{1}$ with $n$ open first-order variables is properly contained in $\U{1}$ with $n+1$ open first-order variables for $n\in\N$.  
-\end{theorem}
-
 \subsection{Horn Counting Classes}
-
-We define the syntactic classes $\XE{i}$ and $\XU{i}$ as follows. First we define extended Horn clauses.
-\begin{eqnarray*}
-	PL &::=& X_i(\x),\, i\in\N,\\
-	NL &::=& \neg X_i(\x),\, i\in\N \ \mid \ \exists x\, NL,\\
-	NC &::=& NL \ \mid \ \alpha, \alpha \mbox{ is an {\sc FO}-formula over } \L \ \mid \ (NC \vee NC),\\
-	HC &::=& NC \ \mid \ (NC \vee PL) \ \mid \ PL,
-\end{eqnarray*}
-where $PL$ represents a positive literal, $NL$ is an \textit{extended} negative literal, $NC$ is an extended Horn clause, and $HC$ is an extended Horn formula.
-Now we define the syntax of the classes inductively.\textbf{}
-\begin{enumerate}
-	\item $\HE{0}$:
-	\begin{eqnarray*}
-		E_0 &::=& HC \ \mid \ E_0 \wedge E_0.
-	\end{eqnarray*}
-	\item $\HU{0}$:
-	\begin{eqnarray*}
-		U_0 &::=& E_0.
-	\end{eqnarray*}
-	\item $\HE{i+1}$:
-	\begin{eqnarray*}
-		E_{i+1} &::=& U_i \ \mid \ \exists x \, E_{i+1}.
-	\end{eqnarray*}
-	\item $\HU{i+1}$:
-	\begin{eqnarray*}
-		U_{i+1} &::=& E_i \ \mid \ \forall x \, U_{i+1}.
-	\end{eqnarray*}
-\end{enumerate}
-A function $f$ is in $\XE{i}$ (resp. $\XU{i}$) if there is an $\L$-formula $\varphi$ defined by the grammar $E_i$ (resp. $U_i$) such that $f = f_{\varphi}$.
-
-The class $\pe$ is defined in \cite{DBLP:conf/mfcs/PagourtzisZ06} as $\pe = \{f\mid f\in\#P \mbox{ and its decision version } L_{f}\in P\}$.
-
-\begin{theorem}
-	$\HE{2} \subseteq \pe$
-\end{theorem}
-
-We define the decision problem
-\begin{multline*}
-\shdhsat = \{\Phi \mid \Phi \text{ is a disjunction of Horn} \\ \text{formulas and $\Phi$ is satisfiable}\},
-\end{multline*}
-and the counting problem {\shdhsat } as a function that counts all satisfying assignments to a formula $\Phi$ which is a disjunction of Horn formulas.
-
-\begin{theorem} \label{sigma2hard}
-	$\shdhsat$ is hard for $\XE{2}$ under parsimonious reductions. 
-\end{theorem}

### 2016-12-14 LICS/appendix.tex
avance en saluja

@@
 For the second condition, note that for each function $f\in \nfpspace$, $\vert f(\A) \vert$ is polynomial on $\vert A \vert$.
-Thus, there is a $\pfp$ formula $\Phi$ that models the string $f(\A)$ by the bit.
+Thus, there is a $\pfp$ formula $\Phi$ that models the string $f(\A)$ bitwise.
 The rest of the proof is analogous to Theorem \ref{captfp}.
@@
 
-\textbf{Proof of Theorem \ref{saluja-eq}}
+\textbf{Proof of Theorem \ref{one-sigma-zero}}
+
+Towards a contradiction, suppose that there exists a constant function $f(\A) = s > 0$ which is expressible in $\E{0}$.
+This is, there is a formula $\psi(x,X)$ such that $f_{\psi(x,X)}(\A) = \vert \{ \langle e,P\rangle \mid \A\models\psi(e,P) \} \vert = s$ for each $\A\in\ostr$.
+It can be seen that if $X$ contains more than zero variables, then the function $f_{\psi(x,X)}$ grows exponentially.
+We have now that $\psi(x)$ contains only first order variables.
+Consider some $\A \in \ostr$ of size $n$ and let $a_1,\ldots,a_s\in \A$ be the tuples that satisfy $\A \models \psi(a_i)$.
+Now consider a structure $\A'$ that is obtained by duplicating $\A$.
+This is, $\A'$ has size $2n$, its first half is isomorphic to its second half, and also to $\A$.
+We have that there is at least $2s$ tuples $a_i,b_i\in \A'$, $s$ on each half, such that $\A'\models\psi(a_i)$ and $\A'\models\psi(b_i)$, which leads to a contradiction.
+
+\textbf{Proof of Theorem \ref{mult-sigma-zero}}
+
+Let $\alpha \in \sqso(\loge{0})$ with binary product.
+This is, $\alpha$ is defined by the grammar $\alpha := \sa{x}\varphi \ \mid \ \sa{X}\varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \ (\alpha\mult\alpha)$, where $\varphi \in \loge{0}$.
+For convenience, we add the expressions $\sa{x}(\varphi\mult\varphi)$ and $\sa{X}(\varphi\mult\varphi)$ to the grammar.
+Let $k$ be the number of algebraic operators in $\alpha$.
+We will show by induction on $k$ that $\alpha$ is expressible without multiplication.
+If $k = 1$, then the possible cases are $\alpha = \sa{x}\varphi$, where $\varphi\in\loge{0}$, which is already product-free.
+If $k = 2$, then the only that is not two nested summations is $\alpha = \sa{x}(\varphi\mult\psi)$, which can be expressed as $\alpha = \sa{x}(\varphi\wedge\psi)$.
+Suppose that the hypothesis holds for $k = 1$ through $n-1$.
+Let $\alpha$ have $n$ algebraic operators.
+If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ where $\beta$ is an algebraic expression, then $\beta$ has $n-1$ algebraic operators and by hypothesis it can be expressed as product-free.
+If $\alpha = (\beta \mult \gamma)$ there are two cases.
+If at least one between $\beta$ and $\gamma$ has a product, then by hypothesis it can be expressed as product-free.
+Moreover, the new expression for $\alpha$ has strictly less than $n$ algebraic operators so by hypothesis it can be expressed as product-free.
+If there are no products in $\beta$ or $\gamma$, then let $\beta = \sa{x}\delta(x)$.
+We have that $\alpha$ can be expressed as $\sa{y}(\delta(y)\mult\gamma)$, where $y$ is a first-order variable that does not appear in $\beta$ or $\gamma$.
+Since $\delta(y)\mult\gamma$ has $n-1$ algebraic operators, it can be expressed as product-free, and then $\alpha$ can be expressed as product-free.
+
+\textbf{Proof of Theorem \ref{fo-prod-sigma-zero}}
+
+The resulting grammar is $\alpha := \sa{x}\varphi \ \mid \ \sa{X} \varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \pa{x}\alpha$.
+Consider a more restricted grammar that replaces $\pa{x}\alpha$ by $\pa{x}\sa{y}\varphi$.
+Let $\alpha = \pa{x}\sa{y}(x = y)$.
+This formula produces a constant function equal to 1, which as shown in Theorem \ref{one-sigma-zero} is not expressible in $\E{0}$.
 

### 2016-12-14 LICS/saluja.tex
avance en saluja

@@
 Saluja et.
-al \cite{DBLP:journals/jcss/SalujaST95} define a family of counting classes $\#\cL$ for each fragment $\cL$ of $\fo$.
+al \cite{DBLP:journals/jcss/SalujaST95} define a family of counting classes $\#\LL$ for each fragment $\LL$ of $\fo$.
 For a formula $\varphi(x,X)$, the function $f_{\varphi(x,X)}$ is defined as
@@
 \]
-for each $\A\in\str$. A function $f:\Sigma^*\to\nat$ is in $\#\cL$ if there exists an $\cL$ formula $\varphi(x,X)$ such that $f = f_{\varphi(x,X)}$.
+for each $\A\in\str$.
+A function $f:\Sigma^*\to\nat$ is in $\#\LL$ if there exists an $\LL$ formula $\varphi(x,X)$ such that $f = f_{\varphi(x,X)}$.
+Note that for each fragment $\LL$ of $\fo$, we have that $\#\LL \subseteq \eqso(\LL)$.
+In fact we need a more restricted fragment of $\eqso$.
+Let $\sqso(\LL)$ be a fragment of $\qso$ defined by the following grammar:
+\[
+\alpha := \sa{x}\varphi \ \mid \ \sa{X} \varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha,
+\]
+where $\varphi$ is an $\LL$-formula. It can be seen that $\sqso(\LL) = \#\LL$. We are interested in, for each $\fo$-fragment $\LL$, the biggest fragment of $\eqso$ that is contained in $\#\LL$.
 
-\begin{theorem} \label{saluja-eq}
-	$\eqso(\cL) = \#\cL$ for each fragment $\L$ of $\fo$.
+Let $\LL = \loge{0}$:
+\begin{theorem} \label{one-sigma-zero}
+	Positive constant functions are not expressible in $\E{0}$
+\end{theorem}
+\begin{corollary}
+	If a fragment of $\eqso(\loge{0})$ is contained in $\E{0}$, its grammar does not allow sole constants.
+\end{corollary}
+\begin{proposition}
+	Sum is not expressible in $\E{0}$
+\end{proposition}
+\begin{theorem} \label{mult-sigma-zero}
+	$\sqso(\loge{0})$ with binary product is contained in $\E{0}$.
+\end{theorem}
+\begin{theorem} \label{fo-prod-sigma-zero}
+	If an extension of $\sqso(\loge{0})$ is contained in $\E{0}$, its grammar does not allow first-order product.
 \end{theorem}
 
-For every logic $\cL$, we define an $\cL$-extended quantifier-free (QF) formula as follows:
+For every logic $\LL$, we define an $\LL$-extended quantifier-free (QF) formula as follows:
 \begin{eqnarray*}
-	\varphi &::=& \alpha, \alpha \text{ is an $\cL$-formula} \ \mid \\
+	\varphi &::=& \alpha, \alpha \text{ is an $\LL$-formula} \ \mid \\
 	&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\

### 2016-12-14 LICS/counting.tex
mas cambios ...

@@
 %!TEX root = main.tex
+
+\marcelo{Ahora solo consideramos la idea de capturar sobre estructuras ordenadas}
 
@@
 \begin{definition}
-	Let $\FF$ a fragment of $\qso$, $\CC$ a function complexity class over $\bbN$ and $\KK$ a class of finite structures. Then $\FF$ {\em captures} $\CC$ over $\KK$ if:
+	Let $\FF$ a fragment of $\qso$ and $\CC$ a function complexity class over $\bbN$. Then $\FF$ {\em captures} $\CC$ over the class of ordered structures if:
 	\begin{itemize}
-		\item for every $\alpha$ in $\FF$, there exists $f \in \CC$ such that $\res{\alpha}{\KK} = \res{f}{\KK}$; and
+		\item for every $\alpha$ in $\FF$, there exists $f \in \CC$ such that for every 
+		$\A \in \ostr[\R]$:  $\alpha(\enc(\A)) = f(\A)$; and
 
-		\item for every $f \in \CC$, there exists $\alpha$ in $\FF$ such that $\res{f}{\KK} = \res{\alpha}{\KK}$.
+		\item for every $f \in \CC$, there exists $\alpha$ in $\FF$ such that for every 
+		$\A \in \ostr[\R]$:  $f(\A) = \alpha(\enc(\A))$.
 	\end{itemize}
 \end{definition}
-In the previous definition, if $\KK$ is the class of all finite structures, then we say that $\FF$ captures $\CC$ over the class of all structures, and if $\KK$is the class of all ordered finite structures, then we say that $\FF$ captures $\CC$ over the class of ordered structures.
+%In the previous definition, if $\KK$ is the class of all finite structures, then we say that $\FF$ captures $\CC$ over the class of all structures, and if $\KK$is the class of all ordered finite structures, then we say that $\FF$ captures $\CC$ over the class of ordered structures.
 
@@
 
-\begin{proposition} \label{eqso-fo-cap-not}
-	$\eqso(\fo)$ does not capture $\shp$ over the class of all structures.
-\end{proposition}
-
 \begin{theorem} \label{eqso-eso-cap}

### 2016-12-14 LICS/preliminaries.tex
cambios ...
mas cambios ...

@@
 
+\marcelo{Agregar las definiciones de LFP y PFP}
 \subsection{Function complexity classes}
@@
 
-If $\KK$ is a class of finite structures and $f$ is a function from a relational signature $\R$ to $\bbD$, we denote by $\res{f}{\KK}$ the restriction of $f$ to $\KK$, that is, a function such that: the domain of $\res{f}{\KK}$ is $\{ \enc(\A) \mid \A \in \str[\R] \cap \KK\}$, and
-$\res{f}{\KK}(\enc(\A)) = f(\enc(\A))$ for every $\A \in \str[\R] \cap \KK$.
+\marcelo{Ya no consideramos clases de estructuras, hay que revisar que esta seccion tenga una notacion consistente}
+
+%If $\KK$ is a class of finite structures and $f$ is a function from a relational signature $\R$ to $\bbD$, we denote by $\res{f}{\KK}$ the restriction of $f$ to $\KK$, that is, a function such that: the domain of $\res{f}{\KK}$ is $\{ \enc(\A) \mid \A \in \str[\R] \cap \KK\}$, and
+%$\res{f}{\KK}(\enc(\A)) = f(\enc(\A))$ for every $\A \in \str[\R] \cap \KK$.

### 2016-12-14 LICS/preliminaries.tex
+

@@
 \end{eqnarray*}
-Given a set $\bbD$ and a relational signature $\R$, a function $f$ is said to be from $\R$ to $\bbD$ if $f$ is a total function from $\{\enc(\A) \mid \A \in \str[\R]\}$ to $\bbD$.
-Moreover, a set $\CC$ of functions over $\bbD$ is said to be a {\em function complexity class over $\bbD$} if every $f \in \CC$ is a function from some relational signature $\R$ to $\bbD$.
+Given a relational signature $\R$,  function $f$ is said to be defined over $\R$ if $f$ is a total function from $\{\enc(\A) \mid \A \in \ostr[\R]\}$ to $\nat$.
+Moreover, a set $\CC$ of functions is said to be a {\em function complexity class} if every $f \in \CC$ is a function defined over some relational signature $\R$.
 

### 2016-12-14 LICS/counting.tex
+

@@
 
-Assume that $\alpha$ is a sentence in $\qso$. Then slightly abusing notation, we also use $\alpha$ to denote a function from $\R$ to $\bbN$ such that for every $\A \in \str[\R] $:
+Assume that $\alpha$ is a sentence in $\qso$. Then slightly abusing notation, we also use $\alpha$ to denote a function defined over $\R$  such that for every $\A \in \ostr[\R] $:
 \begin{eqnarray*}
@@
 \begin{definition}
-	Let $\FF$ a fragment of $\qso$ and $\CC$ a function complexity class over $\bbN$. Then $\FF$ {\em captures} $\CC$ over the class of ordered structures if:
+	Let $\FF$ a fragment of $\qso$ and $\CC$ a function complexity class. Then $\FF$ {\em captures} $\CC$ if:
 	\begin{itemize}
 		\item for every $\alpha$ in $\FF$, there exists $f \in \CC$ such that for every 
-		$\A \in \ostr[\R]$:  $\alpha(\enc(\A)) = f(\A)$; and
+		$\A \in \ostr[\R]$:  $\alpha(\enc(\A)) = f(\enc(\A))$; and
 
 		\item for every $f \in \CC$, there exists $\alpha$ in $\FF$ such that for every 
-		$\A \in \ostr[\R]$:  $f(\A) = \alpha(\enc(\A))$.
+		$\A \in \ostr[\R]$:  $f(\enc(\A)) = \alpha(\enc(\A))$.
 	\end{itemize}

### 2016-12-14 LICS/counting.tex
++

@@
 
+\marcelo{Hacer una pasada a la demostracion del siguiente teorema considerando que solo usamos estructuras ordenadas, y teniendo en cuenta que las funciones estan definidas sobre relational signatures.}
+
 \begin{theorem} \label{captfp}
@@
 \end{theorem}
+
+\marcelo{Hacer una pasada a la demostracion del siguiente teorema basada en la demostracion para LFP.}
 

### 2016-12-15 LICS/appendix.tex
+
15 dic

@@
 
-\textbf{Proof of Theorem \ref{no-mult}}
+\subsection{Proof of Theorem \ref{no-mult}}
 
@@
 
-\textbf{Proof of Theorem \ref{captfp}}
+\subsection{Proof of Theorem \ref{captfp}}
 
-For the first condition, let $\alpha\in\qfo(\lfp)$, $\A\in\ostr$, $v$ and $V$ a first and second order assignment for $\A$, respectively.
-To evaluate $\sem{\alpha}(\A,v,V)$, we replace each first order sum and first order product by their corresponding expansion.
-This is, $\Sigma x \beta(x)$ is replaced by $(\beta(a_1)+\cdots+\beta(a_n))$, where $A = \{a_1,\ldots,a_n\}$, and $\Pi x \beta(x)$ is replaced by $(\beta(a_1)\cdot\,\cdots\,\cdot\beta(a_n))$.
-Then we replace each logic sub-formula $\varphi$ in $\alpha$ by their evaluated value, 0 or 1.
-Since each of this formulas is in $\lfp$, this can be done in polynomial time.
-The resulting formula is an arithmetic expression, which can be evaluated recursively in polynomial time.
+For the first condition, let $\alpha\in\qfo(\lfp)$ over some $\R$.
+Let $f$ be a function over $\R$ defined by the following procedure.
+Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{a_1,\ldots,a_n\}$.
+Replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
+This is, $\Sigma x \beta(x)$ is replaced by $(\beta(a_1)+\cdots+\beta(a_n))$ and $\Pi x \beta(x)$ is replaced by $(\beta(a_1)\cdot\,\cdots\,\cdot\beta(a_n))$.
+Then each sub-formula $\varphi\in\lfp$ in $\alpha$ is replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
+The resulting formula is an arithmetic expression of polynomial size which is evaluated and lastly given as output.
+Note that $f\in\fp$ and $f(\enc(\A)) = \alpha(\enc(\A))$.
 
-For the second condition, let $f\in \fp$.
-Let $k\in\nat$ be such that $\vert f(\A) \vert = \vert A \vert^k$ for each $\A\in\ostr[\R]$.
-Let $\Phi(x_0,\ldots,x_{k-1})$ be a $\lfp$ formula such that $(\A,a_0,\ldots,a_{k-1})\models\Phi$ if and only if in the string $y = f(\A)$ there is 1 in the $(\vert A \vert^{k-1}m_{k-1} + \cdots + \vert A \vert^2 m_2 + \vert A \vert m_1 + m_0)$-th position from right to left, and where $a_i$ is the (0-indexed) $m_i$-th element of $A$.
+For the second condition, let $f\in \fp$ defined over some $\R$.
+Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
+Let $\Phi(x_0,\ldots,x_{\ell-1})$ be a $\lfp$ formula that represents the following.
+For each $\A \in \ostr[\R]$, for convenience, suppose its domain is $A = \{0,\ldots,n-1\}$ with the usual order.
+Let $y$ be a string of size $n^{\ell}$ that represents the value of $f(\enc(\A))$ in binary.
+Then, for each $\ell$-tuple $(a_0,\ldots,a_{\ell-1}) \in A^{\ell}$, it holds that $(\A,a_0,\ldots,a_{\ell-1}) \models \Phi$ if and only if the $(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)$-th bit of $y$, from least to most significant, is 1.
 We use
 	$$
-	\alpha = \Sigma x_0 \cdots \Sigma x_{k-1} \Phi(x_0,\ldots,x_{k-1})\cdot\varphi_{k-1}(x_{k-1})\cdot\,\cdots\,\cdot\varphi_0(x_0).
+	\alpha = \Sigma x_0 \cdots \Sigma x_{\ell-1} \Phi(x_0,\ldots,x_{\ell-1})\cdot\varphi_{\ell-1}(x_{\ell-1})\cdot\,\cdots\,\cdot\varphi_0(x_0).
 	$$
 where $\varphi_i(x) = \Pi y[(y < x)\mapsto\Pi z_1\cdots\Pi z_i\,2]$ for $i > 0$ and $\varphi_0(x) = \Pi y[(y < x)\mapsto 2]$.
-Note that if $a$ is the  $m$-th element in $A$, then $\sem{\varphi_i(x)}(\A,a) = 2^{\vert A \vert^i m}$.
-Therefore, for each $(a_0,\ldots,a_{k-1})\in A^k$, we have that $\sem{\Phi(x_0,\ldots,x_{k-1})\cdot\varphi_{k-1}(x_{k-1})\,\cdots\,\varphi_0(x_0)}(\A,a_0,\ldots,a_{k-1}) = 2^{(\vert A \vert^{k-1}m_{k-1} + \cdots + \vert A \vert^2 m_2 + \vert A \vert m_1 + m_0)}$ if $(\A,a_0,\ldots,a_{k-1})\models\Phi$ and 0 otherwise, and adding these values gives $f(\A)$.
+Let $\A$ the previously defined structure.
+Note that for each $a\in A$, then $\sem{\varphi_i(x)}(\A,a) = 2^{n^i a}$.
+Therefore, for each $(a_0,\ldots,a_{\ell-1})\in A^\ell$, we have that
+$$
+\sem{\Phi(x_0,\ldots,x_{\ell-1})\cdot\varphi_{\ell-1}(x_{\ell-1})\,\cdots\,\varphi_0(x_0)}(\A,a_0,\ldots,a_{\ell-1}) = 
+\begin{cases}
+2^{(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)} &\text{if } (\A,a_0,\ldots,a_{\ell-1})\models\Phi \\
+0 &\text{otherwise.}
+\end{cases}
+$$
+and adding these values gives $f(\enc(\A))$. Then, $\alpha$ is in $\qso(\lfp)$ over $\R$ and $\alpha(\enc(\A)) = f(\enc(\A))$.
 
-\textbf{Proof of Theorem \ref{qfo-pfp-cap}}
+\subsection{Proof of Theorem \ref{qfo-pfp-cap}}
 
-For the first condition, we use the exact same procedure as in Theorem \ref{captfp}, noting that evaluating each of the $\pfp$ formulas can be done in polynomial space, and that the result is also polynomial on the size of the input.
+For the first condition, let $\alpha\in\qfo(\pfp)$ over some $\R$.
+Let $f$ be a function over $\R$ defined by the following procedure.
+Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{a_1,\ldots,a_n\}$.
+Replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
+This is, $\Sigma x \beta(x)$ is replaced by $(\beta(a_1)+\cdots+\beta(a_n))$ and $\Pi x \beta(x)$ is replaced by $(\beta(a_1)\cdot\,\cdots\,\cdot\beta(a_n))$.
+Then each sub-formula $\varphi\in\pfp$ in $\alpha$ is replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
+The resulting formula is an arithmetic expression of polynomial size which is evaluated and lastly given as output.
+Note that $f\in\nfpspace$ and $f(\enc(\A)) = \alpha(\enc(\A))$.
 
-For the second condition, note that for each function $f\in \nfpspace$, $\vert f(\A) \vert$ is polynomial on $\vert A \vert$.
-Thus, there is a $\pfp$ formula $\Phi$ that models the string $f(\A)$ bitwise.
-The rest of the proof is analogous to Theorem \ref{captfp}.
+For the second condition, let $f\in \nfpspace$ defined over some $\R$.
+Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
+Let $\Phi(x_0,\ldots,x_{\ell-1})$ be a $\pfp$ formula that represents the following.
+For each $\A \in \ostr[\R]$, for convenience, suppose its domain is $A = \{0,\ldots,n-1\}$ with the usual order.
+Let $y$ be a string of size $n^{\ell}$ that represents the value of $f(\enc(\A))$ in binary.
+Then, for each $\ell$-tuple $(a_0,\ldots,a_{\ell-1}) \in A^{\ell}$, it holds that $(\A,a_0,\ldots,a_{\ell-1}) \models \Phi$ if and only if the $(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)$-th bit of $y$, from least to most significant, is 1.
+We use
+$$
+\alpha = \Sigma x_0 \cdots \Sigma x_{\ell-1} \Phi(x_0,\ldots,x_{\ell-1})\cdot\varphi_{\ell-1}(x_{\ell-1})\cdot\,\cdots\,\cdot\varphi_0(x_0).
+$$
+where $\varphi_i(x) = \Pi y[(y < x)\mapsto\Pi z_1\cdots\Pi z_i\,2]$ for $i > 0$ and $\varphi_0(x) = \Pi y[(y < x)\mapsto 2]$.
+Let $\A$ the previously defined structure.
+Note that for each $a\in A$, then $\sem{\varphi_i(x)}(\A,a) = 2^{n^i a}$.
+Therefore, for each $(a_0,\ldots,a_{\ell-1})\in A^\ell$, we have that
+$$
+\sem{\Phi(x_0,\ldots,x_{\ell-1})\cdot\varphi_{\ell-1}(x_{\ell-1})\,\cdots\,\varphi_0(x_0)}(\A,a_0,\ldots,a_{\ell-1}) = 
+\begin{cases}
+2^{(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)} &\text{if } (\A,a_0,\ldots,a_{\ell-1})\models\Phi \\
+0 &\text{otherwise.}
+\end{cases}
+$$
+and adding these values gives $f(\enc(\A))$. Then, $\alpha$ is in $\qso(\pfp)$ over $\R$ and $\alpha(\enc(\A)) = f(\enc(\A))$.
 

### 2016-12-15 LICS/counting.tex
15 dic

@@
 	\begin{itemize}
-		\item for every $\alpha$ in $\FF$, there exists $f \in \CC$ such that for every 
+		\item for every $\alpha$ in $\FF$, let $\R$ be some signature in which $\alpha$ is in. There exists $f \in \CC$ such that for every 
 		$\A \in \ostr[\R]$:  $\alpha(\enc(\A)) = f(\enc(\A))$; and
 
-		\item for every $f \in \CC$, there exists $\alpha$ in $\FF$ such that for every 
+		\item for every $f \in \CC$, let $f$ be defined over $\R$. There exists $\alpha$ in $\FF$ such that for every 
 		$\A \in \ostr[\R]$:  $f(\enc(\A)) = \alpha(\enc(\A))$.
@@
 \marcelo{Hacer una pasada a la demostracion del siguiente teorema considerando que solo usamos estructuras ordenadas, y teniendo en cuenta que las funciones estan definidas sobre relational signatures.}
-
+\martin{done}
 \begin{theorem} \label{captfp}
@@
 \marcelo{Hacer una pasada a la demostracion del siguiente teorema basada en la demostracion para LFP.}
-
+\martin{done.}
 \begin{theorem} \label{qfo-pfp-cap}

### 2016-12-15 LICS/framework.tex
standarizing saluja
15 dic

@@
 Then given a relational signature $\R$, the set of quantitative second-order logic formulas (or just $\qso$-formulas) over $\R$ is given by the following grammar:
-\begin{multline*}
+\begin{multline}
 \label{eq-def-qso}
 \alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \\ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha 
-\end{multline*}
+\end{multline}
 where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
+If $\R$ is not mentioned, $\qso$ refers to the union of the sets of $\qso$ formulas over $\R$, for every relational signature $\R$.
 
-Let $\R$ be a relational signature, $\A$ be a finite $\R$-structure with domain $A$, $v$ a first-order assignment for $\A$ and $V$ a second-order assignment for $\A$.
+Let $\R$ be a relational signature, $\A$ be an ordered finite $\R$-structure with domain $A$, $v$ a first-order assignment for $\A$ and $V$ a second-order assignment for $\A$.
 Then the \emph{evaluation} of a $\qso$-formula $\alpha$ over $(\A, v, V)$ is defined as a function $\sem{\alpha}$ that on input $(\A, v, V)$ returns a non-negative integer in $\bbN$.
@@
 A $\qso$-formula $\alpha$ is said to be a \emph{sentence} if it does not have any free variable, that is, every variable in $\alpha$ is under the scope of a usual quantifier or a quantitative quantifier.
-It is important to notice that if $\alpha$ is a $\qso$-sentence over a relational signature $\R$, then for every finite $\R$-structure $\A$, first-order assignments $v_1$, $v_2$ for $\A$ and second-order assignments $V_1$, $V_2$ for $\A$, it holds that:
+It is important to notice that if $\alpha$ is a $\qso$-sentence over a relational signature $\R$, then for every ordered finite $\R$-structure $\A$, first-order assignments $v_1$, $v_2$ for $\A$ and second-order assignments $V_1$, $V_2$ for $\A$, it holds that:
 \begin{eqnarray*}
@@
 In particular, in this paper we consider the following fragments $\LL$: $\fo$, which is obtained by disallowing the use of the second-order quantifier ($\exists X$ or $\forall X$) in the formula $\varphi$ in \eqref{eq-def-qso}, and $\eso$, which is obtained by allowing $\varphi$ in \eqref{eq-def-qso} to be a formula of the form $\exists X_1 \cdots \exists X_k \, \psi$, where $\psi$ is an $\fo$-formula.
+
+\begin{theorem} \label{no-mult}
+	Let $\alpha$ be a $\qso$-formula in some fragment of $\qso$.
+There is a $\qso$ formula $\alpha'$ in the same fragment such that $\alpha'$ does not have any subformula of the form $(\beta_1 \cdot \beta_2)$ and $\sem{\alpha} = \sem{\alpha'}$.
+\end{theorem}

### 2016-12-15 LICS/preliminaries.tex
+
15 dic

@@
 
-A $\mathbb{N}$ relational signature $\R$ is a finite set $\{R_1, \ldots, R_k\}$, where each $R_i$ ($1 \leq i \leq k$) is a relation name with an associate arity greater than 0, which is denoted by $\arity(R_i)$.
+\marcelo{Dado que solo vamos a considerar estructuras ordenadas, hay que revisar y simplificar la notacion. Mantenemos las notaciones $\str[\R]$ y $\ostr[\R]$.}
+
+A $\mathbb{N}$ relational signature $\R$ is a finite set $\{R_1, \ldots, R_k\}$, where each $R_i$ ($1 \leq i \leq k$) is a relation name with an associated arity greater than 0, which is denoted by $\arity(R_i)$.
 A finite structure over $\R$ (or just finite $\R$-structure) is a tuple $\A = \langle A, R_1^\A, \ldots, R_k^\A \rangle$ such that $A$ is a finite set and $R_i^\A \subseteq A^{\arity(R_i)}$ for every $i \in \{1, \ldots, k\}$.
@@
 To define the semantics of $\so$, we need to introduce some terminology.
-Given a relational signature $\R$ and a finite $\R$-structure $\A$ with domain $A$, a first-order assignment $v$ for $\A$ is a total function from $\fv$ to $A$, while a second-order assignment $V$ for $\A$ is a total function with domain $\sv$ that maps each $X \in \sv$ to a subset of $A^{\arity(X)}$.
+Given a relational signature $\R$ and an ordered finite $\R$-structure $\A$ with domain $A$, a first-order assignment $v$ for $\A$ is a total function from $\fv$ to $A$, while a second-order assignment $V$ for $\A$ is a total function with domain $\sv$ that maps each $X \in \sv$ to a subset of $A^{\arity(X)}$.
 Moreover, given a first-order assignment $v$ for $\A$, $x \in \fv$ and $a \in A$, we denote by $v[a/x]$ a first-order assignment such that $v[a/x](x) = a$ and $v[a/x](y) = v(y)$ for every $y \in \fv$ distinct from $x$.
@@
 Assume that $\varphi$ is an $\so$-formula over a signature $\R$.
-Then given a finite $\R$-structure $\A$ with domain $A$, a first-order assignment $v$ for $\A$ and a second-order assignment $V$ for $\A$, we say that $(\A, v, V)$ satisfies $\varphi$, denoted by $(\A, v, V) \models \varphi$, if: (1) $\varphi$ is the formula $R(x_1, \ldots, x_\ell)$ and $(v(x_1), \ldots, v(x_\ell)) \in R^\A$; (2) $\varphi$ is the formula $X(x_1, \ldots, x_m)$ and $(v(x_1), \ldots, v(x_m)) \in V(X)$; (3) $\varphi$ is the formula $\neg \psi$ and it not the case that $(\A, v, V) \models \psi$; (4) $\varphi$ is the formula $(\varphi_1 \vee \varphi_2)$, and $(\A, v, V) \models \varphi_1$ or $(\A, v, V) \models \varphi_2$; (5) $\varphi$ is the formula $\exists x \, \psi$ and there exists $a \in A$ such that $(\A, v[a/x], V) \models \psi$; or (6) $\varphi$ is the formula $\exists X \, \psi$ and there exists $B \subseteq A^{\arity(X)}$ such that $(\A, v, V[B/X]) \models \psi$.
+Then given an ordered finite $\R$-structure $\A$ with domain $A$, a first-order assignment $v$ for $\A$ and a second-order assignment $V$ for $\A$, we say that $(\A, v, V)$ satisfies $\varphi$, denoted by $(\A, v, V) \models \varphi$, if: (1) $\varphi$ is the formula $R(x_1, \ldots, x_\ell)$ and $(v(x_1), \ldots, v(x_\ell)) \in R^\A$; (2) $\varphi$ is the formula $X(x_1, \ldots, x_m)$ and $(v(x_1), \ldots, v(x_m)) \in V(X)$; (3) $\varphi$ is the formula $\neg \psi$ and $(\A, v, V)$ does not satisfy $\psi$; (4) $\varphi$ is the formula $(\varphi_1 \vee \varphi_2)$, and $(\A, v, V) \models \varphi_1$ or $(\A, v, V) \models \varphi_2$; (5) $\varphi$ is the formula $\exists x \, \psi$ and there exists $a \in A$ such that $(\A, v[a/x], V) \models \psi$; or (6) $\varphi$ is the formula $\exists X \, \psi$ and there exists $B \subseteq A^{\arity(X)}$ such that $(\A, v, V[B/X]) \models \psi$.
 As usual, we consider the propositional operators $\wedge$, $\rightarrow$, and $\leftrightarrow$ that can be obtained from $\vee$ and $\neg$.
@@
 \subsection{Function complexity classes}
-Assume that $\R = \{R_1, \ldots, R_k\}$ is a relational signature and $\A$ is a finite $\R$-structure with a domain $A$ containing $n$ elements, and assume that  $<$ is a linear order on $A$, say $a_1 < a_2 < \ldots < a_n$.
+Assume that $\R = \{R_1, \ldots, R_k\}$ is a relational signature and $\A$ is an ordered finite $\R$-structure with a domain $A$ containing $n$ elements, and recall that  $<$ is a linear order on $A$, say $a_1 < a_2 < \ldots < a_n$.
 For every $i \in \{1, \ldots, k\}$, define the encoding of $R_i^\A$, denoted by $\enc(R_i^\A)$, as the following binary string.
@@
 \marcelo{Ya no consideramos clases de estructuras, hay que revisar que esta seccion tenga una notacion consistente}
+\martin{done}
+
+\marcelo{Agregar las definiciones de las clases usuales de funciones: FP, \#P, ...}
+\martin{done}
+
+We define the following function complexity classes: $\fp$ is defined as the class of functions computable in polynomial time.
+$\shp$ is defined as the class of functions that on an input are equal to the number of accepting paths of an $\np$ machine on that input.
+$\spp$ is defined as the class of functions that on an input are equal to the number of different outputs of an $\np$ transducer on that input.
+$\fpspace$ is the class of functions computable in polynomial space.
+$\nfpspace$ is the class of single-exponentially bounded functions computable in polynomial space.
 

### 2016-12-19 LICS/appendix.tex
19 dic

@@
 \textbf{Proof of Theorem \ref{eqso-fo-cap}}
+
+For condition (1) in \ref{cap-def}, let $\alpha \in \eqso(\fo)$ be defined over a relational signature $\R$.
+Let $\ell$ be the number of algebraic operators in $\alpha$.
+We will prove by induction over $\ell$ that there is a $\np$ machine $M$ such that $\alpha(\enc(\A))$ is equal to the number of accepting paths of $M$ on input $\enc(\A)$ for every $\A\in\ostr[\R]$.
+Let $\ell = 0$.
+Then $\alpha$ is either a formula in $\fo$ or a constant, and both functions can be easily computed in polynomial time.
+Suppose the hypothesis holds for $\ell$ from 1 through $n-1$.
+Let $\alpha$ have $n$ algebraic operators and let $\A$ be an ordered structure over $\R$ with domain $A$.
+If $\alpha = (\beta \add \gamma)$, then there exists a $\np$ machine for both $\beta$ and $\gamma$.
+Consider a machine that chooses between 0 or 1, if it is 0 (1), it calculates the function for $\beta$ ($\gamma$).
+The number of accepting paths in this $\np$ machine is equal to the function for $(\beta \add \gamma)$.
+If $\alpha = (\beta \mult \gamma)$, consider a machine that evaluates $\beta$, and in each accepting path, it continues evaluating $\gamma$.
+The number of accepting paths in this $\np$ machine is equal to the function for $(\beta \mult \gamma)$.
+If $\alpha = \sa{x}\beta$, consider a machine that chooses some $a\in A$ and then evaluates $\beta$ over $a$.
+If $\alpha = \pa{x}\beta$, consider a machine that evaluates $\beta$ consecutively over every element in $A$.
+If $\alpha = \sa{X}\beta$, consider a machine that guesses $P \subseteq A^{arity(X)}$ and then evaluates $\beta$ over $P$.
+This covers all possible cases for $\alpha$.
+Let $f$ be the function that this machine computes.
+We have that $f$ is a $\shp$ function over $\R$ and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
+
+For condition (2) in \ref{cap-def}, note that Saluja et al.
+\cite{DBLP:journals/jcss/SalujaST95} proved that $\shp = \#\fo$.
+We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso$ be its corresponding formula.
+Since this formula is also in $\eqso$, the condition holds.
 

### 2016-12-19 LICS/counting.tex
19 dic

@@
 \end{eqnarray*}
-\begin{definition}
-	Let $\FF$ a fragment of $\qso$ and $\CC$ a function complexity class. Then $\FF$ {\em captures} $\CC$ if:
+\begin{definition} \label{cap-def}
+	Let $\FF$ a fragment of $\qso$ and $\CC$ a function complexity class. Then $\FF$ {\em captures $\CC$ over ordered structures} if
+
 	\begin{itemize}
-		\item for every $\alpha$ in $\FF$, let $\R$ be some signature in which $\alpha$ is in. There exists $f \in \CC$ such that for every 
-		$\A \in \ostr[\R]$:  $\alpha(\enc(\A)) = f(\enc(\A))$; and
+	\item[(1)] for every $\alpha$ in $\FF$ over a signature $\R$. There exists $f \in \CC$ such that for every $\A \in \ostr[\R]$:  $\alpha(\enc(\A)) = f(\enc(\A))$; and
 
-		\item for every $f \in \CC$, let $f$ be defined over $\R$. There exists $\alpha$ in $\FF$ such that for every 
-		$\A \in \ostr[\R]$:  $f(\enc(\A)) = \alpha(\enc(\A))$.
+	\item[(2)] for every $f \in \CC$, let $f$ be defined over $\R$. There exists $\alpha$ in $\FF$ such that for every $\A \in \ostr[\R]$:  $f(\enc(\A)) = \alpha(\enc(\A))$.
 	\end{itemize} 
@@
 \begin{theorem} \label{captfp}
-	$\qfo(\lfp)$ captures $\fp$ over the class of ordered structures.
+	$\qfo(\lfp)$ captures $\fp$ over ordered structures.
 \end{theorem}
 
-\marcelo{Hacer una pasada a la demostracion del siguiente teorema basada en la demostracion para LFP.}
-\martin{done.}
 \begin{theorem} \label{qfo-pfp-cap}
-	$\qfo(\pfp)$ captures $\nfpspace$ over the class of ordered structures.
+	$\qfo(\pfp)$ captures $\nfpspace$ over ordered structures.
 \end{theorem}
 
+Saluja et.
+al \cite{DBLP:journals/jcss/SalujaST95} define a family of counting classes $\#\LL$ for each fragment $\LL$ of $\fo$.
+For a formula $\varphi(x,X)$ over $\R$, the function $f_{\varphi(x,X)}$ is defined as
+\[
+f_{\varphi(x,X)}(\A) = \vert \{\langle e,P\rangle\mid \A\models\varphi(e,P)\}\vert.
+\]
+for each $\A\in\ostr[\R]$.
+A function $f:\Sigma^*\to\nat$ is in $\#\LL$ if there exists an $\LL$ formula $\varphi(x,X)$ such that $f = f_{\varphi(x,X)}$.
+Let $\sqso(\LL)$ be a fragment of $\qso$ defined by the following grammar:
+\[
+\alpha := \sa{x}\varphi \ \mid \ \sa{X} \varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha,
+\]
+where $\varphi$ is an $\LL$-formula. It can be seen that $\sqso(\LL)$ captures $\#\LL$ over ordered structures and that $\sqso(\LL) \subseteq \eqso(\LL)$.
+
 \begin{theorem} \label{eqso-fo-cap}
-	$\eqso(\fo)$ captures $\shp$ over the class of ordered structures.
+	$\eqso(\fo)$ captures $\shp$ over ordered structures.
 \end{theorem}
@@
 \begin{theorem} \label{eqso-eso-cap}
-	$\eqso(\eso)$ captures $\spp$ over the class of ordered structures.
+	$\eqso(\eso)$ captures $\spp$ over ordered structures.
 \end{theorem}
+
+\begin{theorem} \label{qso-pfp-cap}
+	$\qso(\pfp)$ captures $\fpspace$ over ordered structures.
+\end{theorem}

### 2016-12-19 LICS/framework.tex
19 dic

@@
 
-Consider $\bbN$ as the set of non-negative integers.
-Then given a relational signature $\R$, the set of quantitative second-order logic formulas (or just $\qso$-formulas) over $\R$ is given by the following grammar:
+Given a relational signature $\R$, the set of quantitative second-order logic formulas (or just $\qso$-formulas) over $\R$ is given by the following grammar:
 \begin{multline}
@@
 Let $\R$ be a relational signature, $\A$ be an ordered finite $\R$-structure with domain $A$, $v$ a first-order assignment for $\A$ and $V$ a second-order assignment for $\A$.
-Then the \emph{evaluation} of a $\qso$-formula $\alpha$ over $(\A, v, V)$ is defined as a function $\sem{\alpha}$ that on input $(\A, v, V)$ returns a non-negative integer in $\bbN$.
+Then the \emph{evaluation} of a $\qso$-formula $\alpha$ over $(\A, v, V)$ is defined as a function $\sem{\alpha}$ that on input $(\A, v, V)$ returns a number in $\bbN$.
 Formally, the function $\sem{\alpha}$ is recursively defined as follows:

### 2016-12-19 LICS/preliminaries.tex
19 dic

@@
 
-\marcelo{Dado que solo vamos a considerar estructuras ordenadas, hay que revisar y simplificar la notacion. Mantenemos las notaciones $\str[\R]$ y $\ostr[\R]$.}
-
-A $\mathbb{N}$ relational signature $\R$ is a finite set $\{R_1, \ldots, R_k\}$, where each $R_i$ ($1 \leq i \leq k$) is a relation name with an associated arity greater than 0, which is denoted by $\arity(R_i)$.
+A relational signature $\R$ (or just signature) is a finite set $\{R_1, \ldots, R_k\}$, where each $R_i$ ($1 \leq i \leq k$) is a relation name with an associated arity greater than 0, which is denoted by $\arity(R_i)$.
 A finite structure over $\R$ (or just finite $\R$-structure) is a tuple $\A = \langle A, R_1^\A, \ldots, R_k^\A \rangle$ such that $A$ is a finite set and $R_i^\A \subseteq A^{\arity(R_i)}$ for every $i \in \{1, \ldots, k\}$.
@@
 \subsection{Function complexity classes}
-Assume that $\R = \{R_1, \ldots, R_k\}$ is a relational signature and $\A$ is an ordered finite $\R$-structure with a domain $A$ containing $n$ elements, and recall that  $<$ is a linear order on $A$, say $a_1 < a_2 < \ldots < a_n$.
+Assume that $\R = \{R_1, \ldots, R_k\}$ is a relational signature and $\A$ is an ordered finite $\R$-structure with a domain $A$ containing $n$ elements.
+Recall that  $<$ is a linear order on $A$, say $a_1 < a_2 < \ldots < a_n$.
 For every $i \in \{1, \ldots, k\}$, define the encoding of $R_i^\A$, denoted by $\enc(R_i^\A)$, as the following binary string.
@@
 \end{eqnarray*}
-Given a relational signature $\R$,  function $f$ is said to be defined over $\R$ if $f$ is a total function from $\{\enc(\A) \mid \A \in \ostr[\R]\}$ to $\nat$.
+Given a relational signature $\R$,  function $f$ is said to be {\em defined over $\R$} (or just {\em over $\R$}) if $f$ is a total function from $\{\enc(\A) \mid \A \in \ostr[\R]\}$ to $\nat$.
 Moreover, a set $\CC$ of functions is said to be a {\em function complexity class} if every $f \in \CC$ is a function defined over some relational signature $\R$.

### 2016-12-19 LICS/saluja.tex
19 dic

@@
-Saluja et.
-al \cite{DBLP:journals/jcss/SalujaST95} define a family of counting classes $\#\LL$ for each fragment $\LL$ of $\fo$.
-For a formula $\varphi(x,X)$, the function $f_{\varphi(x,X)}$ is defined as
-\[
-f_{\varphi(x,X)}(\A) = \vert \{\langle e,P\rangle\mid \A\models\varphi(e,P)\}\vert.
-\]
-for each $\A\in\str$.
-A function $f:\Sigma^*\to\nat$ is in $\#\LL$ if there exists an $\LL$ formula $\varphi(x,X)$ such that $f = f_{\varphi(x,X)}$.
-Note that for each fragment $\LL$ of $\fo$, we have that $\#\LL \subseteq \eqso(\LL)$.
-In fact we need a more restricted fragment of $\eqso$.
-Let $\sqso(\LL)$ be a fragment of $\qso$ defined by the following grammar:
-\[
-\alpha := \sa{x}\varphi \ \mid \ \sa{X} \varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha,
-\]
-where $\varphi$ is an $\LL$-formula. It can be seen that $\sqso(\LL) = \#\LL$. We are interested in, for each $\fo$-fragment $\LL$, the biggest fragment of $\eqso$ that is contained in $\#\LL$.
+We are interested in, for each $\fo$-fragment $\LL$, the biggest fragment of $\eqso$ that is contained in $\#\LL$.
 

### 2016-12-19 LICS/appendix.tex
19 dic

--- 2016-12-15 LICS/appendix.tex
@@
 
-\textbf{Proof of Theorem \ref{eqso-fo-cap}}
-
-\textbf{Proof of Theorem \ref{eqso-fo-cap-not}}
-
-\textbf{Proof of Theorem \ref{eqso-eso-cap}}
-
-\textbf{Proof of Theorem \ref{one-sigma-zero}}
+\subsection{Proof of Theorem \ref{eqso-fo-cap}}
+
+For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
+This machine, for each $\A\in\ostr[\R]$, each first-order assignment $v$ over $\A$ and each second-order assignment $V$ over $\A$, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths.
+Suppose $\A$ has domain $A$.
+If $\alpha$ is a $\fo$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial time, and accepts if and only if it holds true.
+If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
+If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
+If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
+If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
+If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
+If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
+This covers all possible cases for $\alpha$.
+Let $\alpha$ be a formula in $\eqso(\fo)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some first-order assignment $v$ over $\A$ and some second-order assignment $V$ over $\A$.
+We have that $f$ is a $\shp$ function over $\R$ and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
+
+For condition (2) in \ref{cap-def}, note that Saluja et al.
+\cite{DBLP:journals/jcss/SalujaST95} proved that $\shp = \#\fo$.
+We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso(\fo)$ be its corresponding formula.
+Since this formula is also in $\eqso(\fo)$, the condition holds.
+
+\subsection{Proof of Theorem \ref{eqso-eso-cap}}
+
+For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
+This machine, for each $\A\in\ostr[\R]$, each first-order assignment $v$ over $\A$ and each second-order assignment $V$ over $\A$ on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic outputs.
+Suppose $\A$ has domain $A$.
+On each of the steps the algorithm adds the current configuration to the output tape.
+If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time without modifying the output tape, and accepts if and only if the condition holds true.
+If $\alpha$ is a constant $s$, then the algorithm produces $s$ branches and accepts in all of them.
+If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
+If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
+If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
+If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
+If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
+This covers all possible cases for $\alpha$.
+Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some first-order assignment $v$ over $\A$ and some second-order assignment $V$ over $\A$.
+We have that $f$ is a $\spp$ function over $\R$ and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
+
+For condition (2) in \ref{cap-def}, note that Compton and Graedel \cite{DBLP:journals/jcss/ComptonG96} proved that $\spp = \#\eso$.
+We also have that $\sqso(\eso)$ captures $\#\eso$ over ordered structures so for each $f\in \spp$ let $\alpha \in \sqso(\eso)$ be its corresponding formula.
+Since this formula is also in $\eqso(\eso)$, the condition holds.
+
+\subsection{Proof of Theorem \ref{qso-pfp-cap}}
+
+For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
+This machine, for each $\A\in\ostr[\R]$, each first-order assignment $v$ over $\A$ and each second-order assignment $V$ over $\A$, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths.
+Suppose $\A$ has domain $A$.
+If $\alpha$ is a $\pfp$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space \cite{DBLP:conf/lics/AbiteboulV89}\martin{no estoy seguro de esta cita}, and accepts if and only if it holds true.
+If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
+If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
+If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
+If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
+If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
+If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
+If $\alpha = \pa{X}\beta$, it simulates $M_{\beta}$ on input $(\A,v,V[B/X])$ consecutively for each $B\in A^{arity(X)}$.
+This covers all possible cases for $\alpha$, and each of these steps can be computed in polynomial space.
+Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some first-order assignment $v$ over $\A$ and some second-order assignment $V$ over $\A$.
+We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$ \cite{DBLP:journals/siamcomp/Ladner89}, and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
+
+\subsection{Proof of Theorem \ref{one-sigma-zero}}
 

### 2016-12-19 LICS/counting.tex
19 dic

@@
 Saluja et.
-al \cite{DBLP:journals/jcss/SalujaST95} define a family of counting classes $\#\LL$ for each fragment $\LL$ of $\fo$.
+al \cite{DBLP:journals/jcss/SalujaST95} define a family of counting classes $\#\LL$ for each fragment $\LL$ of $\so$.
 For a formula $\varphi(x,X)$ over $\R$, the function $f_{\varphi(x,X)}$ is defined as
@@
 
+\begin{corollary} \label{eqso-lfp-cap}
+	$\eqso(\lfp)$ captures $\shp$ over ordered structures.
+\end{corollary}
+
 \begin{theorem} \label{eqso-eso-cap}

### 2016-12-19 LICS/appendix.tex
19 dic

@@
 %!TEX root = main.tex
+
+\subsection{Notation for the appendix}
+
+For a given signature $\R$, we define $\ostr[\R]^*$ as $$\ostr[\R]^* = \{(\A,v,V) \mid \A\in\ostr[\R]\text{, $v$ ($V$) is a first-order (second-order) assignment for $\A$}  \}$$.
+We denote $(\varphi \mapsto \alpha)$ for $(\neg\varphi + (\varphi\cdot\alpha))$.
+Note that $\sem{(\varphi \mapsto \alpha)}(\A,v,V)$ equals $\sem{\alpha}(\A,v,V)$ when $(\A,v,V)\models\varphi$ and 1 otherwise, for each $(\A,v,V) \in \ostr[\R]^*$.
 
@@
 	\end{enumerate}
-	\item Both of $\beta_1$ and $\beta_2$ are of the form $\Pi x \gamma$, $\Pi X \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula.
-This case also separates in further cases.
-We define $(\varphi \mapsto \alpha) := (\neg\varphi + (\varphi\cdot\alpha))$ for which $\sem{(\varphi \mapsto \alpha)}(\A,v,V)$ equals $\sem{\alpha}(\A,v,V)$ when $(\A,v,V)\models\varphi$ and 1 otherwise.
+	\item Both of $\beta_1$ and $\beta_2$ are of the form $\Pi x \gamma$, $\Pi X \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. This case also separates in further cases.
 	\begin{enumerate}
@@
 Let $y$ be a string of size $n^{\ell}$ that represents the value of $f(\enc(\A))$ in binary.
-Then, for each $\ell$-tuple $(a_0,\ldots,a_{\ell-1}) \in A^{\ell}$, it holds that $(\A,a_0,\ldots,a_{\ell-1}) \models \Phi$ if and only if the $(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)$-th bit of $y$, from least to most significant, is 1.
+Then, for each $\ell$-tuple $(a_0,\ldots,a_{\ell-1}) \in A^{\ell}$, it holds that $\A \models \Phi(a_0,\ldots,a_{\ell-1})$ if and only if the $(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)$-th bit of $y$, from least to most significant, is 1.
 We use
@@
 For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
-This machine, for each $\A\in\ostr[\R]$, each first-order assignment $v$ over $\A$ and each second-order assignment $V$ over $\A$, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths.
+This machine, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
 Suppose $\A$ has domain $A$.
@@
 This covers all possible cases for $\alpha$.
-Let $\alpha$ be a formula in $\eqso(\fo)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some first-order assignment $v$ over $\A$ and some second-order assignment $V$ over $\A$.
+Let $\alpha$ be a formula in $\eqso(\fo)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
 We have that $f$ is a $\shp$ function over $\R$ and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
@@
 For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
-This machine, for each $\A\in\ostr[\R]$, each first-order assignment $v$ over $\A$ and each second-order assignment $V$ over $\A$ on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic outputs.
+This machine, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic outputs for each $(\A,v,V) \in \ostr[\R]^*$.
 Suppose $\A$ has domain $A$.
@@
 This covers all possible cases for $\alpha$.
-Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some first-order assignment $v$ over $\A$ and some second-order assignment $V$ over $\A$.
+Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
 We have that $f$ is a $\spp$ function over $\R$ and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
@@
 For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
-This machine, for each $\A\in\ostr[\R]$, each first-order assignment $v$ over $\A$ and each second-order assignment $V$ over $\A$, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths.
+This machine, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
 Suppose $\A$ has domain $A$.
@@
 This covers all possible cases for $\alpha$, and each of these steps can be computed in polynomial space.
-Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some first-order assignment $v$ over $\A$ and some second-order assignment $V$ over $\A$.
+Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
 We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$ \cite{DBLP:journals/siamcomp/Ladner89}, and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
+
+For condition (2) in \ref{cap-def}...
 

### 2016-12-20 LICS/appendix.tex
20 dic

@@
 
-For a given signature $\R$, we define $\ostr[\R]^*$ as $$\ostr[\R]^* = \{(\A,v,V) \mid \A\in\ostr[\R]\text{, $v$ ($V$) is a first-order (second-order) assignment for $\A$}  \}$$.
+For a given signature $\R$, we define $\ostr[\R]^*$ as $$\ostr[\R]^* = \{(\A,v,V) \mid \A\in\ostr[\R]\text{, $v$ ($V$) is a first-order (second-order) assignment for $\A$}  \}.$$
 We denote $(\varphi \mapsto \alpha)$ for $(\neg\varphi + (\varphi\cdot\alpha))$.
-Note that $\sem{(\varphi \mapsto \alpha)}(\A,v,V)$ equals $\sem{\alpha}(\A,v,V)$ when $(\A,v,V)\models\varphi$ and 1 otherwise, for each $(\A,v,V) \in \ostr[\R]^*$.
+Note that $\sem{(\varphi \mapsto \alpha)}(\A,v,V)$ equals $\sem{\alpha}(\A,v,V)$ when $(\A,v,V)\models\varphi$ and equals 1 otherwise, for each $(\A,v,V) \in \ostr[\R]^*$.
+
+If $\bar{x} = (x_1,\ldots,x_m)$ and $\bar{y} = (y_1,\ldots,y_m)$ are tuples of first-order variables, we denote $\bar{x} < \bar{y}$ for the formula $\bigvee_{i = 1}^m[\bigwedge_{j = 1}^{i-1}x_j = y_j \wedge x_i < y_i]$.
 
@@
 Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
-Let $\Phi(x_0,\ldots,x_{\ell-1})$ be a $\lfp$ formula that represents the following.
+Let $\Phi(x_0,\ldots,x_{\ell-1})$ be a formula that represents the following.
 For each $\A \in \ostr[\R]$, for convenience, suppose its domain is $A = \{0,\ldots,n-1\}$ with the usual order.
@@
 Then, for each $\ell$-tuple $(a_0,\ldots,a_{\ell-1}) \in A^{\ell}$, it holds that $\A \models \Phi(a_0,\ldots,a_{\ell-1})$ if and only if the $(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)$-th bit of $y$, from least to most significant, is 1.
+Since this condition can be computed in polynomial time, and $\lfp$ captures $\ptime$, this formula can be expressed in $\lfp$.
 We use
 $$
-	\alpha = \Sigma x_0 \cdots \Sigma x_{\ell-1} \Phi(x_0,\ldots,x_{\ell-1})\cdot\varphi_{\ell-1}(x_{\ell-1})\cdot\,\cdots\,\cdot\varphi_0(x_0).
+\alpha = \sa{x_0} \cdots \sa{x_{\ell-1}} \Phi(x_0,\ldots,x_{\ell-1})\cdot\varphi_{\ell-1}(x_{\ell-1})\cdot\,\cdots\,\cdot\varphi_0(x_0).
 $$
@@
 \begin{cases}
-2^{(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)} &\text{if } (\A,a_0,\ldots,a_{\ell-1})\models\Phi \\
+2^{(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)} &\text{if } \A\models\Phi(a_0,\ldots,a_{\ell-1})\\
 0 &\text{otherwise.}
@@
 $$
-and adding these values gives $f(\enc(\A))$. Then, $\alpha$ is in $\qso(\lfp)$ over $\R$ and $\alpha(\enc(\A)) = f(\enc(\A))$.
+and adding these values gives $f(\enc(\A))$. Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\alpha(\enc(\A)) = f(\enc(\A))$. \qed
 
@@
 
-For the first condition, let $\alpha\in\qfo(\pfp)$ over some $\R$.
+For condition (2) in \ref{cap-def}, let $\alpha\in\qfo(\pfp)$ over some $\R$.
 Let $f$ be a function over $\R$ defined by the following procedure.
@@
 
-For the second condition, let $f\in \nfpspace$ defined over some $\R$.
+For condition (2) in \ref{cap-def}, let $f\in \nfpspace$ defined over some $\R$.
 Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
@@
 $$
-and adding these values gives $f(\enc(\A))$. Then, $\alpha$ is in $\qso(\pfp)$ over $\R$ and $\alpha(\enc(\A)) = f(\enc(\A))$.
+and adding these values gives $f(\enc(\A))$. Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\alpha(\enc(\A)) = f(\enc(\A))$. \qed
 
@@
 Since this formula is also in $\eqso(\fo)$, the condition holds.
+\qed
 
@@
 Since this formula is also in $\eqso(\eso)$, the condition holds.
+\qed
 
@@
 
-For condition (2) in \ref{cap-def}...
+For the second condition, let $f\in \fpspace$ defined over some $\R$.
+Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq 2^{n^\ell}$, where $\A$ has a domain of size $n$.
+Let $\Phi(X)$ be a $\pfp$ formula that represents the following.
+Consider an order over the subsets of $A^{\ell}$ given by the formula $$\varphi_{<}(X,Y) = \exists\bar{u}\big[\neg X(\bar{u})\wedge Y(\bar{u})\wedge \forall\bar{v}\big(
+\bar{u}<\bar{v}\to(X(\bar{u})\iff Y(\bar{v}))\big)\big].$$
+
+Let $B$ be the $p$-th subset with respect to this order and let $w$ be a string of size $2^{n^{\ell}}$ which represents the value of $f(\enc(\A))$ in binary.
+Then $\A\models\Phi(B)$ if and only if the $p$-th bit in $w$ is 1, from least to most significant.
+Since this condition can be verified in polynomial space, and $\pfp$ captures $\pspace$, then $\Phi$ can be expressed in $\pfp$.
+We define
+$\alpha := \sa{X}(\Phi(X)\mult\varphi(X))$, where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
+Note that for each $B\subseteq A^{\ell}$ such that $B$ is the $m$-th element in the given order, $\sem{\varphi(B)}(\A) = 2^m$.
+Therefore, $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
+\qed
 

### 2016-12-20 LICS/appendix.tex
20 dic

@@
 
-For the second condition, let $f\in \fp$ defined over some $\R$.
+For the second condition, let $f\in \fp$ defined over some signature $\R$.
 Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
 Let $\Phi(x_0,\ldots,x_{\ell-1})$ be a formula that represents the following.
-For each $\A \in \ostr[\R]$, for convenience, suppose its domain is $A = \{0,\ldots,n-1\}$ with the usual order.
-Let $y$ be a string of size $n^{\ell}$ that represents the value of $f(\enc(\A))$ in binary.
-Then, for each $\ell$-tuple $(a_0,\ldots,a_{\ell-1}) \in A^{\ell}$, it holds that $\A \models \Phi(a_0,\ldots,a_{\ell-1})$ if and only if the $(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)$-th bit of $y$, from least to most significant, is 1.
+Let $\A$ be an ordered structure over $\R$ with domain $A$.
+Consider a linear order over $A^{\ell}$ given by $\bar{a} < \bar{b}$, for any $\bar{a},\bar{b}\in A^{\ell}$.
+Let $w$ be a string of size $n^{\ell}$ the represents $f(\enc(\A))$ in binary.
+Let $(a_1,\ldots,a_{\ell})\in A^{\ell}$ be the $m$-th tuple in the given order (starting from 0).
+$\A\models\Phi(a_1,\ldots,a_n)$ if an only if the $m$-th bit in $w$ is 1, from least to most significant.
 Since this condition can be computed in polynomial time, and $\lfp$ captures $\ptime$, this formula can be expressed in $\lfp$.
@@
 $$
-\alpha = \sa{x_0} \cdots \sa{x_{\ell-1}} \Phi(x_0,\ldots,x_{\ell-1})\cdot\varphi_{\ell-1}(x_{\ell-1})\cdot\,\cdots\,\cdot\varphi_0(x_0).
-$$
-where $\varphi_i(x) = \Pi y[(y < x)\mapsto\Pi z_1\cdots\Pi z_i\,2]$ for $i > 0$ and $\varphi_0(x) = \Pi y[(y < x)\mapsto 2]$.
-Let $\A$ the previously defined structure.
-Note that for each $a\in A$, then $\sem{\varphi_i(x)}(\A,a) = 2^{n^i a}$.
-Therefore, for each $(a_0,\ldots,a_{\ell-1})\in A^\ell$, we have that
-$$
-\sem{\Phi(x_0,\ldots,x_{\ell-1})\cdot\varphi_{\ell-1}(x_{\ell-1})\,\cdots\,\varphi_0(x_0)}(\A,a_0,\ldots,a_{\ell-1}) = 
-\begin{cases}
-2^{(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)} &\text{if } \A\models\Phi(a_0,\ldots,a_{\ell-1})\\
-0 &\text{otherwise.}
-\end{cases}
-$$
-and adding these values gives $f(\enc(\A))$. Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\alpha(\enc(\A)) = f(\enc(\A))$. \qed
+\alpha = \sa{x_1} \cdots \sa{x_{\ell}} \Phi(x_1,\ldots,x_{\ell})\cdot\varphi(x_1,\ldots,x_{\ell}),
+$$
+where $\varphi(x_1,\ldots,x_{\ell}) := \pa{y_1}\cdots\pa{y_{\ell}}((y_1,\ldots,y_{\ell}) < (x_1,\ldots,x_{\ell}) \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
+Adding this values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
+Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\alpha(\enc(\A)) = f(\enc(\A))$.
+\qed
 
@@
 \subsection{Proof of Theorem \ref{one-sigma-zero}}
-
 Towards a contradiction, suppose that there exists a constant function $f(\A) = s > 0$ which is expressible in $\E{0}$.

### 2016-12-20 LICS/appendix.tex
20 dic

@@
 For a given signature $\R$, we define $\ostr[\R]^*$ as $$\ostr[\R]^* = \{(\A,v,V) \mid \A\in\ostr[\R]\text{, $v$ ($V$) is a first-order (second-order) assignment for $\A$}  \}.$$
-We denote $(\varphi \mapsto \alpha)$ for $(\neg\varphi + (\varphi\cdot\alpha))$.
-Note that $\sem{(\varphi \mapsto \alpha)}(\A,v,V)$ equals $\sem{\alpha}(\A,v,V)$ when $(\A,v,V)\models\varphi$ and equals 1 otherwise, for each $(\A,v,V) \in \ostr[\R]^*$.
-
+We denote $(\varphi \mapsto \alpha)$ for $(\neg\varphi + (\varphi\cdot\alpha))$. Note that for each $(\A,v,V) \in \ostr[\R]^*$, 
+$$
+\sem{(\varphi \mapsto \alpha)}(\A,v,V) = 
+\begin{cases}
+\sem{\alpha}(\A,v,V) &\text{if } (\A,v,V)\models\varphi,\\
+0 &\text{otherwise}.
+\end{cases}
+$$
 If $\bar{x} = (x_1,\ldots,x_m)$ and $\bar{y} = (y_1,\ldots,y_m)$ are tuples of first-order variables, we denote $\bar{x} < \bar{y}$ for the formula $\bigvee_{i = 1}^m[\bigwedge_{j = 1}^{i-1}x_j = y_j \wedge x_i < y_i]$.
@@
 
-For the first condition, let $\alpha\in\qfo(\lfp)$ over some $\R$.
+For the first condition, let $\alpha\in\qfo(\lfp)$ over some signature $\R$.
 Let $f$ be a function over $\R$ defined by the following procedure.
@@
 Let $(a_1,\ldots,a_{\ell})\in A^{\ell}$ be the $m$-th tuple in the given order (starting from 0).
-$\A\models\Phi(a_1,\ldots,a_n)$ if an only if the $m$-th bit in $w$ is 1, from least to most significant.
+$\A\models\Phi(a_1,\ldots,a_{\ell})$ if an only if the $m$-th bit in $w$ is 1, from least to most significant.
 Since this condition can be computed in polynomial time, and $\lfp$ captures $\ptime$, this formula can be expressed in $\lfp$.

### 2016-12-20 LICS/appendix.tex
20 dic

@@
 For a given signature $\R$, we define $\ostr[\R]^*$ as $$\ostr[\R]^* = \{(\A,v,V) \mid \A\in\ostr[\R]\text{, $v$ ($V$) is a first-order (second-order) assignment for $\A$}  \}.$$
-We denote $(\varphi \mapsto \alpha)$ for $(\neg\varphi + (\varphi\cdot\alpha))$. Note that for each $(\A,v,V) \in \ostr[\R]^*$, 
+The {\em conditional count} symbol $(\varphi \mapsto \alpha)$ is defined as $(\neg\varphi + (\varphi\cdot\alpha))$ for given $\so$ formula $\varphi$ and $\qso$ formula $\alpha$.
+Note that for each $(\A,v,V) \in \ostr[\R]^*$,
 $$
@@
 $$
+We will use the symbol $<$ also to denote the lexicographic order over same-sized tuples.
 If $\bar{x} = (x_1,\ldots,x_m)$ and $\bar{y} = (y_1,\ldots,y_m)$ are tuples of first-order variables, we denote $\bar{x} < \bar{y}$ for the formula $\bigvee_{i = 1}^m[\bigwedge_{j = 1}^{i-1}x_j = y_j \wedge x_i < y_i]$.
-
+Similarly, we use $=$ to denote equality between tuples, as $\bar{x} = \bar{y}$ denotes $\bigwedge_{i = 1}^m(x_i = y_i)$.
+
+If $\bar{x} = (x_1,\ldots,x_m)$ ($\bar{X} = (X_1,\ldots,X_m)$) is a tuple of first-order (second-order) variables, we denote $\sa{\bar{x}}\alpha$ for $\sa{x_1}\cdots\sa{x_m}\alpha$ and $\sa{\bar{X}}\alpha$ for $\sa{X_1}\cdots\sa{X_m}\alpha$ for each $\qso$ formula $\alpha$.
 \subsection{Proof of Theorem \ref{no-mult}}
 
-Let $\{+, \cdot, \Sigma, \Pi\}$ be the set of {\it algebraic operators} in $\qso$.
+Let $\{+, \cdot, \sa{}, \pa{}\}$ be the set of {\it algebraic operators} in $\qso$.
 We will prove by induction over the number of algebraic operators in a $\qso$-formula $\alpha$ that there exists some $\qso$-formula $\alpha'$ that respects the stated condition.
@@
 Let $\alpha$ be a $\qso$-formula such that $\text{op}(\alpha) = n$.
-Note that if $\alpha$ is of the form $\alpha = (\beta_1 + \beta_2)$, $\alpha = \Sigma x \beta$, $\alpha = \Pi x \beta$, $\alpha = \Sigma X \beta$ or $\alpha = \Pi X \beta$ the proof follows directly.
-Therefore we assume that $\alpha = (\beta_1 \cdot \beta_2)$.
+Note that if $\alpha$ is of the form $\alpha = (\beta_1 \add \beta_2)$, $\alpha = \sa{x} \beta$, $\alpha = \pa{x} \beta$, $\alpha = \sa{X} \beta$ or $\alpha = \pa{X} \beta$ the proof follows directly.
+Therefore we assume that $\alpha = (\beta_1 \mult \beta_2)$.
 We separate the proof in two main cases:
 \begin{enumerate}
-	\item At least one of $\beta_1, \beta_2$ is not of the form $\Pi x \gamma$, $\Pi X \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. Without loss of generality, let $\beta_1$ be such one.
+	\item At least one of $\beta_1, \beta_2$ is not of the form $\pa{x} \gamma$, $\pa{X} \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. Without loss of generality, let $\beta_1$ be such one.
 	\begin{enumerate}
 		\item $\beta_1 = s \in \nat$.
-Then we use $\alpha' = \beta_2 + \cdots + \beta_2$, where $\beta_2$ is repeated $s$ times, and if $s = 0$, take $\alpha' = \perp$.
-Since each of the formulas paired by $+$ has less than $n$ algebraic operators, the statement holds.
-		\item $\beta_1 = (\gamma_1 + \gamma_2)$.
-Then, we use $\alpha' = (\gamma_1\cdot\beta_2 + \gamma_2\cdot\beta_2)$ and each of the formulas paired by $+$ have less than $n$ algebraic operators so the statement holds.
-		\item $\beta_1 = \Sigma x \gamma$. Then, we use $\alpha' = \Sigma x (\gamma\cdot\beta_2)$ and as in the previous case the statement holds.
-		\item $\beta_1 = \Sigma X \gamma$, which we treat exactly as the previous one.
+Then we use $\alpha' = \beta_2 \add \cdots \add \beta_2$, where $\beta_2$ is repeated $s$ times, and if $s = 0$, take $\alpha' = \perp$.
+Since each of the formulas paired by $\add$ has less than $n$ algebraic operators, the statement holds.
+		\item $\beta_1 = (\gamma_1 \add \gamma_2)$.
+Then, we use $\alpha' = (\gamma_1\mult\beta_2 + \gamma_2\mult\beta_2)$ and each of the formulas paired by $\add$ have less than $n$ algebraic operators so the statement holds.
+		\item $\beta_1 = \sa{x} \gamma$. Then, we use $\alpha' = \sa{x} (\gamma\mult\beta_2)$ and as in the previous case the statement holds.
+		\item $\beta_1 = \sa{X} \gamma$, which we treat exactly as the previous one.
 	\end{enumerate}
-	\item Both of $\beta_1$ and $\beta_2$ are of the form $\Pi x \gamma$, $\Pi X \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. This case also separates in further cases.
+	\item Both of $\beta_1$ and $\beta_2$ are of the form $\pa{x} \gamma$, $\pa{X} \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. This case also separates in further cases.
 	\begin{enumerate}
-		\item $\beta_1 = \Sigma x \gamma_1(x)$ and $\beta_2 = \Sigma y \gamma_2(y)$. Then we use $\alpha' = \Sigma x(\gamma_1(x)\cdot\gamma_2(x))$, and the statement holds.
-		\item $\beta_1 = \Sigma X \gamma_1(X)$ and $\beta_2 = \Sigma Y \gamma_2(Y)$. [something about arities].
-		\item $\beta_1 = \Sigma X \gamma_1(X)$ and $\beta_2 = \Sigma x \gamma_2(x)$ (w.l.o.g.). [use singleton predicates].
-		\item $\beta_1 = \Sigma x \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). Then, we use $\alpha' = \Sigma x (\gamma \cdot \varphi)$, and the statement holds.
-		\item $\beta_1 = \Sigma X \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). We treat this case exactly as the previous one.
+		\item $\beta_1 = \pa{x} \gamma_1(x)$ and $\beta_2 = \pa{y} \gamma_2(y)$. Then we use $\alpha' = \pa{x}(\gamma_1(x)\mult\gamma_2(x))$, and the statement holds.
+		\item $\beta_1 = \pa{X} \gamma_1(X)$ and $\beta_2 = \pa{Y} \gamma_2(Y)$. [something about arities].
+		\item $\beta_1 = \pa{X} \gamma_1(X)$ and $\beta_2 = \pa{x} \gamma_2(x)$ (w.l.o.g.). [use singleton predicates].
+		\item $\beta_1 = \pa{x} \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). Then, we use $\alpha' = \pa{x} (\gamma \cdot \varphi)$, and the statement holds.
+		\item $\beta_1 = \pa{X} \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). We treat this case exactly as the previous one.
 		\item $\beta_1 = \varphi_1$ and $\beta_2 = \varphi_2$ both $\so$-formulas. Then, we use $\alpha' = (\varphi_1 \wedge \varphi_2)$ and the statement holds.
@@
 \subsection{Proof of Theorem \ref{captfp}}
+
+\martin{no revisar, es un copy-paste de arriba}
 
@@
 If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
+If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.%%If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
 If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
@@
 If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
+%% If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
 If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
@@
 
+\subsection{Proof of Theorem \ref{eqso-sigma-zero-in-fp}}
+
+Given a $\eqso(\loge{0})$-formula $\alpha$ over a signature $\R$ we will recursively construct a formula $\tau(\alpha)$ of the form $\beta_1\add\beta_2\add\cdots\add\beta_m$, where each $\beta_i$ is either $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{x}}\sa{\bar{X}}1$ or $\sa{\bar{x}}\sa{\bar{X}}\varphi$, where $\varphi$ is a formula in $\logex{0}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables, such that $\alpha(\enc(\A)) = \tau(\alpha)(\enc(\A))$ for each $\A\in\ostr[\R]$.
+First, each natural constant $s$ in $\alpha$ is replaced by $(1\add 1\add\cdots\add 1)$ ($s$ times).
+If $\alpha = 1$ or $\alpha = \varphi \in \logex{0}$, then $\tau(\alpha) = \alpha$.
+If $\alpha = (\beta \add \gamma)$ then $\tau(\alpha) = (\tau(\beta) \add \tau(\gamma))$.
+If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ this case separates in more cases: If $\alpha = \sa{x}\varphi$ or $\alpha = \sa{X}\varphi$, where $\varphi \in \logex{0}$, then $\tau(\alpha) = \alpha$.
+If $\alpha = \sa{x} 1$ or $\alpha = \sa{X} 1$ then $\tau(\alpha) = \alpha$.
+If $\alpha = \sa{x}(\beta\add\gamma)$ ($\alpha = \sa{X}(\beta\add\gamma)$) then $\tau(\alpha) = \tau(\sa{x}\beta) \add \tau(\sa{x}\gamma)$ ($\tau(\alpha) = \tau(\sa{X}\beta) \add \tau(\sa{X}\gamma)$).
+If $\alpha = \sa{x}\sa{y}\beta$ ($\alpha = \sa{x}\sa{X}\beta$) ($\alpha = \sa{X}\sa{Y}\beta$) then $\tau(\alpha) = \tau(\sa{x}\tau(\sa{y}\beta))$ ($\tau(\alpha) = \tau(\sa{x}\tau(\sa{X}\beta))$) ($\tau(\alpha) = \tau(\sa{X}\tau(\sa{Y}\beta))$).
+If $\alpha = \sa{X}\sa{x}\beta$ then $\tau(\alpha) = \tau(\sa{x}\tau(\sa{X}\beta))$.
+This covers all possible cases for $\alpha$.
+Let $\alpha$ be a $\eqso(\loge{0})$-formula over a signature $\R$.
+Consider the following procedure that receives $\enc(\A)$ as input, where $\A\in\ostr[\R]$ with domain $A$.
+Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_m$.
+For each $\beta_i$, if $\beta_i = 1$ or $\beta_i = \varphi$, then $\sem{\beta_i}(\A)$ can be computed in polynomial time.
+If $\beta_i = \sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{length(\bar{x})}$, if $\beta_i = \sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = 2^{\Sigma_{X\in\bar{X}}\vert A \vert^{arity(X)}}$, if $\beta_i = \sa{\bar{x}}\sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{length(\bar{x})}2^{\Sigma_{X\in\bar{X}}\vert A \vert^{arity(X)}}$, all of which can be computed in polynomial time.
+Otherwise, $\beta_i$ is $\pa{\bar{x}}\varphi$ or $\pa{\bar{X}}\pa{\bar{x}}\varphi$, and then $\sem{\beta_i}(\A) = f_{\varphi}(\enc(\A))$, and $f_{\varphi}\in \#\loge{0}$ which implies that $f_{\varphi}(\enc(\A))$ can be computed in polynomial time \cite{DBLP:journals/jcss/SalujaST95}.
+
+\subsection{Proof of Theorem \ref{eqso-sigma-one-in-eqso-pi-one}}
+
+Let $\alpha$ be a formula in $\eqso(\loge{1})$ over a signature $\R$.
+Let $\tau(\alpha) = \beta_1\add\beta_2\add\cdots\add\beta_m$ defined as in the proof of Theorem \ref{eqso-sigma-zero-in-fp}.
+This is, each $\beta_i$ is of the form $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{x}}\sa{\bar{X}}1$ or $\sa{\bar{x}}\sa{\bar{X}}\varphi$, where $\varphi$ is a formula in $\loge{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables.
+(Note that the form $\sa{\bar{X}}\varphi$ is not allowed in $\eqso(\loge{0})$.) For each $\beta_i$ we will construct a $\gamma_i$ such that $\gamma_i \in \eqso(\logu{1})$ and $\sem{\beta_i}(\A) = \sem{\gamma_i}(\A)$ for each $\A\in\ostr[\R]$.
+Since the forms $1$, $\sa{\bar{x}}1$, $\sa{\bar{X}}1$ and $\sa{\bar{x}}\sa{\bar{X}}1$ are already valid in $\eqso(\logu{1})$, we suppose $\beta_i$ is in one of the remaining forms.
+Without loss of generality, let $\beta_i = \sa{\bar{x}}\sa{\bar{X}}\exists\bar{y}\,\varphi(\bar{x},\bar{y},\bar{X})$ where $\varphi(\bar{x},\bar{y},\bar{X})$ is a quantifier-free formula.
+Similarly as in the proof by Saluja et al.
+\cite{DBLP:journals/jcss/SalujaST95}, we define $\gamma_i = \sa{\bar{x}}\sa{\bar{y}}\sa{\bar{X}}\varphi(\bar{x},\bar{y},\bar{X}) \wedge \forall\bar{z}(\varphi(\bar{x},\bar{z},\bar{X})\to(\bar{z} < \bar{y} \vee \bar{z} = \bar{y}))$ that fulfils the desired condition.
+The theorem holds true by using $\beta = \gamma_1\add\cdots\add\gamma_m$ which is in $\eqso(\logu{1})$
+
 \subsection{Proof of Theorem \ref{one-sigma-zero}}
+
 Towards a contradiction, suppose that there exists a constant function $f(\A) = s > 0$ which is expressible in $\E{0}$.

### 2016-12-20 LICS/framework.tex
20 dic

@@
 In this paper, we consider several fragments of $\qso$, which are obtained by restricting the syntax of the formula $\varphi$ in \eqref{eq-def-qso} or the use of the quantitative quantifiers.
-Let $\qfo$ be the fragment of $\qso$ obtained by only allowing the quantitative quantifiers $\Sigma x$, $\Pi x$ in \eqref{eq-def-qso}.
-Let $\eqso$ be a fragment of $\qso$ defined as $\qfo$ but also allowing the quantitative quantifier $\Sigma X$.
-Moreover, assuming that $\LL$ is a fragment of $\so$, let $\qso(\LL)$ be the fragment of $\qso$ obtained by restricting $\varphi$ in \eqref{eq-def-qso} to be a formula in $\LL$, and let $\eqso(\LL)$ be the fragment of $\eqso$ obtained by imposing the same restriction.
+Let $\qfo$ be a fragment of $\qso$ defined by the grammar in \eqref{eq-def-qso} but only the quantitative quantifiers $\Sigma x$ and $\Pi x$ are allowed.
+Let $\eqfo$ be the fragment of $\qso$ defined by the following grammar:
+$$
+\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid \ \sa{x} \alpha.
+$$ Let $\eqso$ be a fragment of $\qso$ defined as $\eqfo$ but also allowing the quantitative quantifier $\Sigma X$.
+Moreover, assuming that $\LL$ is a fragment of $\so$, let $\qso(\LL)$ be the fragment of $\qso$ obtained by restricting $\varphi$ in \eqref{eq-def-qso} to be a formula in $\LL$, and in general, if $\FF$ is a fragment of $\qso$, $\FF(\LL)$ be the fragment of $\FF$ obtained by imposing the same restriction.
 In particular, in this paper we consider the following fragments $\LL$: $\fo$, which is obtained by disallowing the use of the second-order quantifier ($\exists X$ or $\forall X$) in the formula $\varphi$ in \eqref{eq-def-qso}, and $\eso$, which is obtained by allowing $\varphi$ in \eqref{eq-def-qso} to be a formula of the form $\exists X_1 \cdots \exists X_k \, \psi$, where $\psi$ is an $\fo$-formula.

### 2016-12-20 LICS/saluja.tex
19 dic
20 dic

--- 2016-12-05 LICS/saluja.tex
@@
-Saluja et.
-al \cite{DBLP:journals/jcss/SalujaST95} define a family of counting classes $\#\cL$ for each fragment $\cL$ of $\fo$.
-For a formula $\varphi(x,X)$, the function $f_{\varphi(x,X)}$ is defined as
-\[
-f_{\varphi(x,X)}(\A) = \vert \{\langle e,P\rangle\mid \A\models\varphi(e,P)\}\vert.
-\]
-for each $\A\in\str$. A function $f:\Sigma^*\to\nat$ is in $\#\cL$ if there exists an $\cL$ formula $\varphi(x,X)$ such that $f = f_{\varphi(x,X)}$.
+% We are interested in, for each $\fo$-fragment $\LL$, the biggest fragment of $\eqso$ that is contained in $\#\LL$.
+%Let $\LL = \loge{0}$:
+%\begin{theorem} \label{one-sigma-zero}
+%	Positive constant functions are not expressible in $\E{0}$
+%\end{theorem}
+%\begin{corollary}
+%	If a fragment of $\eqso(\loge{0})$ is contained in $\E{0}$, its grammar does not allow sole constants.
+%\end{corollary}
+%\begin{conjecture}
+%	Sum is not expressible in $\E{0}$
+%\end{conjecture}
+%\begin{theorem} \label{mult-sigma-zero}
+%	$\sqso(\loge{0})$ with binary product is contained in $\E{0}$.
+%\end{theorem}
+%\begin{theorem} \label{fo-prod-sigma-zero}
+%	If an extension of $\sqso(\loge{0})$ is contained in $\E{0}$, its grammar does not allow first-order product.
+%\end{theorem}
 
-\begin{theorem} \label{saluja-eq}
-	$\eqso(\cL) = \#\cL$ for each fragment $\L$ of $\fo$.
+%For every logic $\LL$, we define an $\LL$-extended quantifier-free (QF) formula as follows:
+%\begin{eqnarray*}
+%	\varphi &::=& \alpha, \alpha \text{ is an $\LL$-formula} \ \mid \\
+%	&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
+%	&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
+%\end{eqnarray*}
+%
+%We define syntactically the fragments $\logex{i}$ and $\logux{i}$ according to the following grammar:
+%\begin{align*}
+%\logex{0} = \logux{0} &::= \varphi , \varphi \mbox{ is an $\fo$-extended QF formula,} \\
+%\logex{i+1} &::= \logux{i} \ \mid \ \exists x\, \logex{i+1}, \\
+%\logux{i+1} &::= \logex{i} \ \mid \ \forall x\, \logux{i+1}.
+%\end{align*}
+
+We see that many of the results in Saluja et. al. \cite{DBLP:journals/jcss/SalujaST95} for $\#\LL$ still apply in $\eqso(\LL)$ for a given fragment $\LL$:
+
+\begin{theorem} \label{eqso-sigma-zero-in-fp}
+	For every $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$, the function $f$ over $\R$ defined as $f(\enc(\A)) = \sem{\alpha}(\A)$ is in $\fp$.
 \end{theorem}
 
-For every logic $\cL$, we define an $\cL$-extended quantifier-free (QF) formula as follows:
-\begin{eqnarray*}
-	\varphi &::=& \alpha, \alpha \text{ is an $\cL$-formula} \ \mid \\
-	&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
-	&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
-\end{eqnarray*}
-
-We define syntactically the fragments $\logex{i}$ and $\logux{i}$ according to the following grammar:
-\begin{align*}
-\logex{0} = \logux{0} &::= \varphi , \varphi \mbox{ is an $\fo$-extended QF formula,} \\
-\logex{i+1} &::= \logux{i} \ \mid \ \exists x\, \logex{i+1}, \\
-\logux{i+1} &::= \logex{i} \ \mid \ \forall x\, \logux{i+1}.
-\end{align*}
-
-\begin{theorem} \label{fp1}
-	$\qfo(\logex{0}) \subseteq$ {\sc FP}.
+\begin{theorem} \label{eqso-sigma-one-in-eqso-pi-one}
+	For every $\eqso(\loge{1})$ formula $\alpha$ over a signature $\R$ there exists a $\eqso(\logu{1})$ formula $\beta$ over $\R$ such that $\sem{\alpha}(\A) = \sem{\beta}(\A)$ for every $\A\in\ostr[\R]$.
 \end{theorem}

### 2016-12-21 LICS/appendix.tex
21 dic

@@
 
-Given a $\eqso(\loge{0})$-formula $\alpha$ over a signature $\R$ we will recursively construct a formula $\tau(\alpha)$ of the form $\beta_1\add\beta_2\add\cdots\add\beta_m$, where each $\beta_i$ is either $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{x}}\sa{\bar{X}}1$ or $\sa{\bar{x}}\sa{\bar{X}}\varphi$, where $\varphi$ is a formula in $\logex{0}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables, such that $\alpha(\enc(\A)) = \tau(\alpha)(\enc(\A))$ for each $\A\in\ostr[\R]$.
+First, for a given signature $\R$ and logic fragment $\LL$ we define a {\em valid form in $\LL$} (or just {\em valid form}).
+A formula $\alpha$ is in valid form in $\LL$ if (1) it is equal to $\varphi$, $1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{x}}\sa{\bar{X}}\varphi$ or $\sa{\bar{x}}\sa{\bar{X}}1$ for some $\LL$ formula $\varphi$ and some tuple of first-order (second-order) variables $\bar{x}$ ($\bar{X}$), or (2) if it is equal to $(\beta\add\gamma)$ where $\beta$ and $\gamma$ are formulas in valid form.
+Given a $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$ we will recursively construct a formula $\tau(\alpha)$ in valid form such that $\alpha(\enc(\A)) = \tau(\alpha)(\enc(\A))$ for each $\A\in\ostr[\R]$.
 First, each natural constant $s$ in $\alpha$ is replaced by $(1\add 1\add\cdots\add 1)$ ($s$ times).
-If $\alpha = 1$ or $\alpha = \varphi \in \logex{0}$, then $\tau(\alpha) = \alpha$.
-If $\alpha = (\beta \add \gamma)$ then $\tau(\alpha) = (\tau(\beta) \add \tau(\gamma))$.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ this case separates in more cases: If $\alpha = \sa{x}\varphi$ or $\alpha = \sa{X}\varphi$, where $\varphi \in \logex{0}$, then $\tau(\alpha) = \alpha$.
-If $\alpha = \sa{x} 1$ or $\alpha = \sa{X} 1$ then $\tau(\alpha) = \alpha$.
-If $\alpha = \sa{x}(\beta\add\gamma)$ ($\alpha = \sa{X}(\beta\add\gamma)$) then $\tau(\alpha) = \tau(\sa{x}\beta) \add \tau(\sa{x}\gamma)$ ($\tau(\alpha) = \tau(\sa{X}\beta) \add \tau(\sa{X}\gamma)$).
-If $\alpha = \sa{x}\sa{y}\beta$ ($\alpha = \sa{x}\sa{X}\beta$) ($\alpha = \sa{X}\sa{Y}\beta$) then $\tau(\alpha) = \tau(\sa{x}\tau(\sa{y}\beta))$ ($\tau(\alpha) = \tau(\sa{x}\tau(\sa{X}\beta))$) ($\tau(\alpha) = \tau(\sa{X}\tau(\sa{Y}\beta))$).
-If $\alpha = \sa{X}\sa{x}\beta$ then $\tau(\alpha) = \tau(\sa{x}\tau(\sa{X}\beta))$.
+If $\alpha = 1$ or $\alpha = \varphi \in \loge{0}$, then $\tau(\alpha) = \alpha$.
+We assume that every sub-formula of $\alpha$ is in valid form.
+If $\alpha = (\beta \add \gamma)$ then $\tau(\alpha) = \alpha$.
+If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ this case separates in more cases: If $\alpha = \sa{x}\beta$ and $\beta$ is in one of the valid forms in (1), then $\tau(\alpha) = \alpha$.
+If $\alpha = \sa{X}\beta$ and $\beta$ is in the valid form $\sa{\bar{x}}\gamma$ then $\tau(\alpha) = \sa{\bar{x}}\sa{X}\gamma$; if $\beta$ it is in one of the remaining valid forms in (1), then $\tau(\alpha) = \alpha$.
+If $\alpha = \sa{x}\beta$ ($\alpha = \sa{X}\beta$) and $\beta = (\gamma_1\add\gamma_2)$, then $\tau(\alpha) = \tau(\sa{x}\gamma_1) \add \tau(\sa{x}\gamma_2)$ ($\tau(\alpha) = \tau(\sa{X}\gamma_1) \add \tau(\sa{X}\gamma_2)$).
 This covers all possible cases for $\alpha$.
-Let $\alpha$ be a $\eqso(\loge{0})$-formula over a signature $\R$.
+Let $\alpha$ be a $\eqso(\loge{0})$ formula over a signature $\R$.
 Consider the following procedure that receives $\enc(\A)$ as input, where $\A\in\ostr[\R]$ with domain $A$.
@@
 For each $\beta_i$, if $\beta_i = 1$ or $\beta_i = \varphi$, then $\sem{\beta_i}(\A)$ can be computed in polynomial time.
-If $\beta_i = \sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{length(\bar{x})}$, if $\beta_i = \sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = 2^{\Sigma_{X\in\bar{X}}\vert A \vert^{arity(X)}}$, if $\beta_i = \sa{\bar{x}}\sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{length(\bar{x})}2^{\Sigma_{X\in\bar{X}}\vert A \vert^{arity(X)}}$, all of which can be computed in polynomial time.
-Otherwise, $\beta_i$ is $\pa{\bar{x}}\varphi$ or $\pa{\bar{X}}\pa{\bar{x}}\varphi$, and then $\sem{\beta_i}(\A) = f_{\varphi}(\enc(\A))$, and $f_{\varphi}\in \#\loge{0}$ which implies that $f_{\varphi}(\enc(\A))$ can be computed in polynomial time \cite{DBLP:journals/jcss/SalujaST95}.
+If $\beta_i = \sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}$, if $\beta_i = \sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = 2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, if $\beta_i = \sa{\bar{x}}\sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, all of which can be computed in polynomial time.
+Otherwise, $\beta_i$ is $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, and then $\sem{\beta_i}(\A) = f_{\varphi}(\enc(\A))$, and $f_{\varphi}\in \#\loge{0}$ which implies that $f_{\varphi}(\enc(\A))$ can be computed in polynomial time \cite{DBLP:journals/jcss/SalujaST95}.
 
@@
 This is, each $\beta_i$ is of the form $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{x}}\sa{\bar{X}}1$ or $\sa{\bar{x}}\sa{\bar{X}}\varphi$, where $\varphi$ is a formula in $\loge{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables.
-(Note that the form $\sa{\bar{X}}\varphi$ is not allowed in $\eqso(\loge{0})$.) For each $\beta_i$ we will construct a $\gamma_i$ such that $\gamma_i \in \eqso(\logu{1})$ and $\sem{\beta_i}(\A) = \sem{\gamma_i}(\A)$ for each $\A\in\ostr[\R]$.
+For each $\beta_i$ we will construct a $\gamma_i$ such that $\gamma_i \in \eqso(\logu{1})$ and $\sem{\beta_i}(\A) = \sem{\gamma_i}(\A)$ for each $\A\in\ostr[\R]$.
 Since the forms $1$, $\sa{\bar{x}}1$, $\sa{\bar{X}}1$ and $\sa{\bar{x}}\sa{\bar{X}}1$ are already valid in $\eqso(\logu{1})$, we suppose $\beta_i$ is in one of the remaining forms.
@@
 	where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
-Note that $\alpha_{\min}$ is satisfied only by the lexicographically \linebreak smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
+Note that $\alpha_{\min}$ is satisfied only by the lexicographically smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
 Our new formula is

### 2016-12-23 LICS/appendix.tex

@@
 If $\bar{x} = (x_1,\ldots,x_m)$ and $\bar{y} = (y_1,\ldots,y_m)$ are tuples of first-order variables, we denote $\bar{x} < \bar{y}$ for the formula $\bigvee_{i = 1}^m[\bigwedge_{j = 1}^{i-1}x_j = y_j \wedge x_i < y_i]$.
-Similarly, we use $=$ to denote equality between tuples, as $\bar{x} = \bar{y}$ denotes $\bigwedge_{i = 1}^m(x_i = y_i)$.
+Similarly, we use $=$ to denote equality between tuples, as $\bar{x} = \bar{y}$ denotes $\bigwedge_{i = 1}^m(x_i = y_i)$, and also $\bar{x}\leq\bar{y}$ denotes $\bar{x} < \bar{y} \vee \bar{x} = \bar{y}$.
+We also denote $\min(\bar{x}) := \forall\bar{y}(\bar{y} < \bar{x}\vee \bar{y} = \bar{x})$.
 
 If $\bar{x} = (x_1,\ldots,x_m)$ ($\bar{X} = (X_1,\ldots,X_m)$) is a tuple of first-order (second-order) variables, we denote $\sa{\bar{x}}\alpha$ for $\sa{x_1}\cdots\sa{x_m}\alpha$ and $\sa{\bar{X}}\alpha$ for $\sa{X_1}\cdots\sa{X_m}\alpha$ for each $\qso$ formula $\alpha$.
+We also denote $\length(\bar{x}) = m$ ($\length(\bar{X}) = m$).
 \subsection{Proof of Theorem \ref{no-mult}}
@@
 Given a $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$ we will recursively construct a formula $\tau(\alpha)$ in valid form such that $\alpha(\enc(\A)) = \tau(\alpha)(\enc(\A))$ for each $\A\in\ostr[\R]$.
+
 First, each natural constant $s$ in $\alpha$ is replaced by $(1\add 1\add\cdots\add 1)$ ($s$ times).
@@
 This covers all possible cases for $\alpha$.
+Furthermore, suppose that $\LL$ is closed under conjunction and suppose $\alpha = (\beta\mult\gamma)$.
+If $\beta = (\delta_1\add\delta_2)$ (wlog), then $\tau(\alpha) = \tau(\delta_1\mult\gamma)\add\tau(\delta_2\mult\gamma)$.
+If $\beta = \sa{\bar{x}}\delta$, then $\tau(\alpha) = \tau(\sa{\bar{x}}\tau(\delta\mult\gamma))$.
+If $\beta = 1$, then $\tau(\alpha) = \gamma$.
+If $\beta = \varphi$ and $\gamma = \psi$, then $\tau(\alpha) = \varphi\wedge\psi$.
 Let $\alpha$ be a $\eqso(\loge{0})$ formula over a signature $\R$.
@@
 
-Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$.
+Let $\alpha$ be a $\eqso(\logex{1})$ formula over a signature $\R$.
+We will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
+Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_n$ be defined as in \ref{eqso-sigma-zero-in-fp}, that is, each $\beta_i$ is equal to $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{x}}\sa{\bar{X}}1$ or $\sa{\bar{x}}\sa{\bar{X}}\varphi$, where $\varphi$ is a formula in $\logex{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables.
+We further assume that each $\beta_i$ is either $\sa{\bar{x}}\varphi$ or $\sa{\bar{x}}\sa{\bar{X}}\varphi$, since each $1$, $\varphi$ and $\sa{\bar{X}}\varphi$ can be replaced by the formulas $\sa{x}\min(x)$, $\sa{x}(\varphi\wedge\min(x))$ and $\sa{x}\sa{\bar{X}}(\varphi\wedge\min(x))$ respectively, while maintaining the value of $\sem{\alpha}(\A)$.
+
+The intuition behind our reasoning is separated in three points: (1) For each $\beta_i$ of the form $\sa{\bar{x}}\varphi$, the formula $\kappa(\beta_i)$ will count every tuple $\bar{x}$ that satisfies $\varphi$ except for the lexicographically smallest one.
+(2) For each $\beta_i$ of the form $\sa{\bar{x}}\sa{\bar{X}}\varphi$, the formula $\kappa(\beta_i)$ will isolate the smallest $\bar{X}$ that satisfies $\varphi$, and exclude the lexicographically smallest tuple $\bar{x}$ that satisfies $\varphi(\bar{X})$.
+And (3) if $\alpha = (\beta\add\sa{\bar{x}}\varphi)$ or $\alpha = (\beta\add\sa{\bar{x}}\sa{\bar{X}}\varphi)$, the formula $\kappa(\alpha)$ will exclude the lexicographically smallest tuple that satisfies $\varphi$ if and only if $\beta$ is equal to 0.
+
+For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha) = \sa{\bar{x}}\exists\bar{y}[\varphi(\bar{x},\bar{y})\wedge\exists\bar{z}(\varphi(\bar{z},\bar{y})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfils the desired condition.
+
+For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{x}}\sa{\bar{X}}\exists\bar{y}\,\varphi(\bar{x},\bar{X},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha)$ procedurally as follows: Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
+We suppose that $\varphi$ is in a DNF form that leaves formulas that do not mention $\bar{X}$ as is.
+If it is not, we convert $\varphi$ to this form with a standard transformation algorithm.
+Let $\varphi(\bar{x},\bar{X},\bar{y}) = \bigvee_{i = 1}^m\varphi_i(\bar{x},\bar{X},\bar{y})$ where each $\varphi_i$ has the form:
+$$
+\varphi_i(\bar{x},\bar{X},\bar{y}) = (\text{conjunction of $X_i$'s}) \wedge (\text{conjunction of $\neg X_i$'s})  \wedge (\text{an $\fo$ formula that does not mention any $X_i$}).
+$$
+Define $\varphi_i^{+}$, $\varphi_i^{-}$ and $\varphi_i^{\fo}$ as the formulas mentioned above.
+
+In our procedure we assume that in $\varphi_i^{+}\wedge\varphi_i^{-}$ every first-order variable from $\bar{x},\bar{y}$ is mentioned at most once.
+If not, we add new first-order variables from $\fv$ to $\bar{y}$ so that no variable is repeated, replace them in $\varphi_i^{+}\wedge\varphi_i^{-}$ and add their respective equalities in $\varphi_i^{\fo}$ so that the new formula is equivalent.
+For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (x_1,x_2,x_3,x_4,y)$ and $\varphi_i := X(x_1,x_2) \wedge \neg X(x_3,x_4) \wedge x < y \wedge x_1 = x \wedge x_2 = y \wedge x_3 = x \wedge x_4 = x.$
+
+Furthermore, we assume that $\varphi_i^{\fo}$ defines an ordered partition of the variables in the tuple $(\bar{x},\bar{y})$.
+For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$.
+A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
+On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
+For a given $k$, we define $\cB_k$ as the number of possible ordered partitions for a set of size $k$. 
+If $\varphi_i^{\fo}$ does not define an ordered partition, then for $1 \leq j \leq \cB_{\length(\bar{x},\bar{y})}$ 
+let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
+The formula $\varphi_i(\bar{x},\bar{X},\bar{y})$ is converted into $\bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}}\varphi_i(\bar{x},\bar{X},\bar{y}) \wedge \theta^j(\bar{x},\bar{y})$.
+Consider each of these disjuncts as a new $\varphi_i$.
+
+Now let $\bar{u}$ be the tuple of every first-order variable mentioned in $\varphi_i^{+}$ and let $\bar{v}$ be such that $(\bar{x},\bar{y}) = (\bar{u},\bar{v})$.
+We define a formula $\mu_i$ that is satisfied only by the lexicographically smallest $\bar{u}$ that satisfies $\varphi_i^{+}$ for some $\bar{v}$:
+$$
+\mu_i(\bar{u}) = \exists\bar{v}\,\varphi_i^{\fo}(\bar{u},\bar{v})\wedge\forall\bar{u}'\forall\bar{v}'(\varphi_i^{\fo}(\bar{u}',\bar{v}')\to\bar{u}\leq\bar{u}').
+$$
+Lastly, we define $\kappa(\varphi_i)$:
+$$
+\kappa(\varphi_i) := 
+$$
+%Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$.
 There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$.
@@
 \begin{enumerate}
-	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
-	\begin{eqnarray*}
-		f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
-	\end{eqnarray*}
-	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^d$.
+%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
+%	\begin{eqnarray*}
+%		f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
+%	\end{eqnarray*}
+%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^d$.
 Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily.
 First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
-	\begin{eqnarray*}
-		\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
-	\end{eqnarray*}
-	This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
-	\begin{eqnarray*}
-		\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
-	\end{eqnarray*}
-	If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
+%	\begin{eqnarray*}
+%		\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
+%	\end{eqnarray*}
+%	This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
+%	\begin{eqnarray*}
+%		\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
+%	\end{eqnarray*}
+%	If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
 Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$.
 Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E{1}.$
-
+%	
 	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
@@
 	\begin{multline*}
-	\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to \varphi_{m,<}((\u_1,\dots,\u_n),(\v_1,\dots,\v_n))\Big)\Big],
+	\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to (\u_1,\dots,\u_n)<(\v_1,\dots,\v_n)\Big)\Big],
 	\end{multline*}
@@
 Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$.
-We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \varphi_{<,a_i}(\z,\x_i)$ or (3) $\theta(\x)\models \varphi_{<,a_i}(\x_i,\z)$.
+We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \z<\x_i$ or (3) $\theta(\x)\models \x_i<\z$.
 Considering that (2) and (3) are not possible given that both $\z$ and $\x_i$ are assigned the value $\d_i$ and $\A\models\theta(\d)$, we have that $\theta(\x)\models \z = \x_i$.
@@
 	\begin{multline*}
-	\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_{n_i},\y)\wedge \\ \forall\v_1\cdots\forall\v_{n_i}\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_{n_i},\w)\wedge\bigvee_{j=1}^{n_i}(\u_j\neq\v_j)\big)\to \varphi_{m_i,<}((\u_1,\dots,\u_{n_i}),(\v_1,\dots,\v_{n_i}))\Big)\Big],
+	\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_{n_i},\y)\wedge \\ \forall\v_1\cdots\forall\v_{n_i}\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_{n_i},\w)\wedge\bigvee_{j=1}^{n_i}(\u_j\neq\v_j)\big)\to (\u_1,\dots,\u_{n_i})<(\v_1,\dots,\v_{n_i})\Big)\Big],
 	\end{multline*}
@@
 	\begin{multline*}
-	\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \v \neq \u_\ell\Big) \bigg) \bigg].
+	\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge\limits_{\substack{\ell\in[1,n_i]:\\ \lambda_i(\ell) = j}} \v \neq \u_\ell\Big) \bigg) \bigg].
 	\end{multline*}
@@
 In other words, every assignment $\P^\prime \neq \P$ is such that $\A\models\psi_i(\P^\prime).$ Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
-	\begin{multline}
-	\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge \\ (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
-	\end{multline}
+	$$
+	\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
+	$$
 	Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Assume that $\A\in\str$.
@@
 	\begin{multline*}
-	\alpha^i_{\min}(\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}) = \exists\y\exists\w\Big[ \alpha_i(\x_1,\dots,\x_{n_i},\y,\z_1,\dots,\z_{n_i},\w)\wedge \\ \forall\u_1\cdots\forall\u_{n_i}\forall\s\forall\v_1\cdots\forall\v_{n_i}\forall\t\Big(\big(\alpha_i(\u_1,\dots,\u_{n_i},\s,\v_1,\dots,\v_{n_i},\t)\wedge\bigvee_{j=1}^{n_i}(\x_j\neq\u_j \vee \z_j\neq\v_j)\big)\to \\ \varphi_{m_i,<}((\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}),(\u_1,\dots,\u_{n_i},\v_1,\dots,\v_{n_i}))\Big)\Big],
+	\alpha^i_{\min}(\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}) = \exists\y\exists\w\Big[ \alpha_i(\x_1,\dots,\x_{n_i},\y,\z_1,\dots,\z_{n_i},\w)\wedge \\ \forall\u_1\cdots\forall\u_{n_i}\forall\s\forall\v_1\cdots\forall\v_{n_i}\forall\t\Big(\big(\alpha_i(\u_1,\dots,\u_{n_i},\s,\v_1,\dots,\v_{n_i},\t)\wedge\bigvee_{j=1}^{n_i}(\x_j\neq\u_j \vee \z_j\neq\v_j)\big)\to \\ (\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i})<(\u_1,\dots,\u_{n_i},\v_1,\dots,\v_{n_i})\Big)\Big],
 	\end{multline*}
@@
 	\begin{multline*}
-	\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\w(\exists\u\:\varphi(\u,\X,\w) \wedge \varphi_{p,<}(\w,\z)).
+	\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\w(\exists\u\:\varphi(\u,\X,\w) \wedge \w<\z).
 	\end{multline*}

### 2016-12-23 LICS/horn-counting.tex
pass on Horn section

@@
-We define the fragments $\logeh{i}$ and $\loguh{i}$ as follows. First we define extended Horn clauses.
+%!TEX root = main.tex
+\newcommand{\pP}{\textit{P}}
+\newcommand{\pN}{\textit{N}}
+\newcommand{\pV}{\textit{V}}
+\newcommand{\pT}{\textit{T}}
+\newcommand{\pA}{\textit{A}}
+\newcommand{\pNC}{\textit{NC}}
+
+A positive literal is a formula of the form $X(\x)$, where $X$ is a second-order variable and $\x$ is a tuple of first-order variables, and a negative literal is a formula of the form $\exists \v \, \neg X(\u,\v)$, where $\u$ and $\v$ are tuples of first-order variables.
+Given a relational signature $\R$, a clause over $\R$ is a formula of the form:
 \begin{eqnarray*}
-	PL &::=& X_i(\x),\, i\in\N,\\
-	NL &::=& \neg X_i(\x),\, i\in\N \ \mid \ \exists x\, NL,\\
-	NC &::=& NL \ \mid \ \alpha, \alpha \mbox{ is an {\sc FO}-formula} \ \mid \ (NC \vee NC),\\
-	HC &::=& NC \ \mid \ (NC \vee PL) \ \mid \ PL,
+\forall \x \, (\varphi_1 \vee \cdots \vee \varphi_n),
 \end{eqnarray*}
-where $PL$ represents a positive literal, $NL$ is an \textit{extended} negative literal, $NC$ is an extended Horn clause, and $HC$ is an extended Horn formula.
-Now we define the syntax of the classes inductively.
+where each $\varphi_i$ ($1 \leq i \leq n$) is either a positive literal or a negative literal or an \fo-formula over $\R$.
+A clause is said to be Horn if it contains at most one positive literal, and a formula is said to be Horn if it is a conjunction of Horn clauses over a relational signature $\R$.
+With this terminology, we define $\uhorn$ as the set of formulas $\psi$ such that $\psi$ is a conjunction of Horn clauses over a relational signature $\R$.
+
+\begin{proposition}\label{prop-uhorn-pe}
+$\eqso(\uhorn) \subseteq \pe$
+\end{proposition}
+
+\begin{example}
+Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot)\}$.
+This vocabulary is used as follows to encode a Horn formula.
+A fact $\pP(c,x)$ indicates that propositional variable $x$ is a disjunct in a clause $c$, while $\pN(c,x)$ indicates that $\neg x$ is a disjunct in $c$.
+Furthermore, $\pV(x)$ holds if  $x$ is a propositional variable, and $\pNC(c)$ holds if $c$ is a clause containing only negative literals, that is, $c$ is of the form $(\neg x_1 \vee \cdots \vee \neg x_n)$.
+
+To encode $\chsat$, we define an \so-formula $\varphi(\pT)$ over $\R$, where $\pT$ is a unary predicate, such that for every Horn formula $\theta$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \varphi(\pT)}(\A)$.
+In particular, $\pT(x)$ holds if and only if $x$ is a propositional variable that is assigned value 1.
+More specifically, $\varphi(\pT)$ is defined as follows:
 \begin{align*}
-\logeh{0} &::= HC \ \mid \ \logeh{0} \wedge \logeh{0}, \\
-\loguh{0} &::= \logeh{0}, \\ 
-\logeh{i+1} &::= \loguh{i} \ \mid \ \exists x\, \logeh{i+1}, \\
-\loguh{i+1} &::= \logeh{i} \ \mid \ \forall x\, \loguh{i+1}.
+&\forall x \, (\pT(x) \to \pV(x)) \ \wedge\\
+&\forall c \, (\pNC(c) \to \exists x \, (\pN(c,x) \wedge \neg \pT(x))) \ \wedge\\
+&\forall c \forall x \, ([\pP(c,x) \wedge \forall y \, (\pN(c,y) \to \pT(y))] \to \pT(x)).
 \end{align*}
-\begin{theorem} \label{sigma2-pe}
-	$\eqso(\loguh{2}) \subseteq \pe$
-\end{theorem}
+Given that $\uhorn$ is designed with the goal in mind of capturing $\chsat$, we expect $\varphi(\pT)$ to be a formula in $\uhorn$.
+However, if we rewrite it as a conjunction of clauses we obtain the following:
+\begin{align*}
+&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
+&\forall c \, (\neg \pNC(c) \vee \exists x \, (\pN(c,x) \wedge \neg \pT(x)))\ \wedge\\
+&\forall c \forall x \, (\neg \pP(c,x) \vee \exists y \, (\pN(c,y) \wedge \neg \pT(y)) \vee \pT(x)).
+\end{align*}
+The resulting formula $\varphi(\pT)$ is not in $\uhorn$, but it can be easily transformed into a formula in this class  by introducing an auxiliary binary predicate $\pA$ defined as follows:
+\begin{align*}
+\forall c \forall x \, (\neg \pA(c,x) \leftrightarrow [\pN(c,x) \wedge \neg \pT(x)]).
+\end{align*}
+In this way, we obtain the following formula $\psi(\pT,\pA)$ in $\uhorn$:
+\begin{align*}
+&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
+&\forall c \, (\neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
+&\forall c \forall x \, (\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
+&\forall c \forall x \, (\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
+&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
+&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{T}(x)).
+\end{align*}
+This formula effectively defines $\chsat$
+as for every Horn formula $\theta$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
+Therefore, we conclude that $\chsat \in \eqso(\uhorn)$.
+\end{example}
+We extend the definition of $\uhorn$ to allow existential quantification.
+More precisely, a formula $\varphi$ is in $\ehorn$ if $\varphi$ is of the form $\exists \bar x \, \psi$ with $\psi$ a Horn formula.
+Interestingly, it hold that $\cdnf \in \eqso(\ehorn)$ and
 
-We define the decision problem
+\begin{proposition}\label{prop-ehorn-pe}
+$\eqso(\ehorn) \subseteq \pe$.
+\end{proposition}
+A natural question at this point is whether in the definitions of $\uhorn$ and $\ehorn$, it is necessary to allow negative literals of the form $\exists \v \, \neg X(\u,\v)$.
+The following result shows that it is indeed the case:
+
+\begin{proposition}\label{prop-hsat-not-sigma2}
+$\chsat \not\in \eqso(\loge{2})$.
+\end{proposition}
+We conclude this section by showing that a natural extension of $\chsat$ is $\eqso(\ehorn)$-complete under parsimonious reductions. We define the decision problem:
 \begin{multline*}
-\shdhsat = \{\Phi \mid \Phi \text{ is a disjunction of Horn} \\ \text{formulas and $\Phi$ is satisfiable}\},
+\dhsat = \{\Phi \mid \Phi \text{ is a disjunction of}\\  \text{Horn formulas and $\Phi$ is satisfiable}\},
 \end{multline*}
-and the counting problem {\shdhsat } as a function that counts all satisfying assignments to a formula $\Phi$ which is a disjunction of Horn formulas.
+and the counting problem $\shdhsat$ as a function that counts all satisfying assignments of a formula $\Phi$ that is a disjunction of Horn formulas.
 
 \begin{theorem} \label{sigma2hard}
-	$\shdhsat$ is hard for $\eqso(\loguh{2})$ under parsimonious reductions. 
+	$\shdhsat$ is $\eqso(\ehorn)$-complete under parsimonious reductions. 
 \end{theorem}

### 2016-12-23 LICS/saluja.tex
Avance en sección de saluja.
Sección de Saluja primera versión.

@@
+%\begin{enumerate}
+%	\item Introduce the classes $\eqso(\LL)$.
+%	
+%	\item Study the properties of $\eqso(\LL)$.
+%	\begin{itemize}
+%		\item If $\LL$ is closed under conjunction, then binary product can be defined in $\eqso(\LL)$.
+%		
+%		\item If $\LL$ is closed under XXX, then the first order product can be defined in $\eqso(\LL)$.
+%		
+%		\item Every formula in $\eqso(\LL)$ can be rewrite in \emph{sum normal form}.
+%	\end{itemize}
+%	
+%	\item Introduce Saluja classes.
+%	
+%	\item Show the connection with Saluja classes.
+%	\begin{itemize}
+%		\item New classes are properly contained.
+%		\item Old classes are properly contained/equivalent in new classes. 
+%	\end{itemize}
+%	
+%	\item Study the complexity of new classes.
+%	\begin{itemize}
+%		\item $\eqso(\Sigma_0)$ is in FP.
+%		\item $\eqso(\Sigma_1)$ is $\#P$ easy.
+%		\item $\eqso(\Sigma_1)$ has free pass.
+%	\end{itemize}
+%	
+%	\item Closure properties.
+%	\begin{itemize}
+%		\item $\eqso(\Sigma_1)$ is not closed under substraction.
+%		\item $\eqso(\Sigma_1)$ is not closed under minus 1.
+%	\end{itemize}
+%	
+%\end{enumerate}
+
+Inspired by the connection between $\shp$ and $\#\fo$, the authors in~\cite{DBLP:journals/jcss/SalujaST95} studied subclasses of  $\#fo$ syntactical restricting the alternation quantifiers in boolean formulas defining the functions.
+Specifically, they define the classes $\E{i}$ and $\U{i}$ where functions are defined by $\fo$-formulas in $\loge{i}$ and $\logu{i}$, respectively, for every $i \geq 0$. 
+Interestingly, in~\cite{DBLP:journals/jcss/SalujaST95} they show that this function classes defined a finite hierarchy of the form:
+\[
+\E{0} \; \subsetneq \; \E{1} \; \subsetneq \; \U{1} \; \subsetneq \; \E{2} \; \subsetneq \; \U{2} \; = \; \#\fo 
+\]
+In this section, we study a natural extension of these classes in light of the $\eqso$ logic.
+Formally, we define the classes $\eqso(\loge{i})$ and $\eqso(\logu{i})$ where the boolean logic is restricted to $\loge{i}$ and $\logu{i}$, respectively.
+In the next result, we show that the prenex normal form hierarchy of $\eqso$ also defined a finite hierarchy similar than in~\cite{DBLP:journals/jcss/SalujaST95}.
+\begin{proposition}
+\begin{multline*}
+\eqso(\loge{0}) \; \subsetneq \; \eqso(\loge{1}) \; \subsetneq \; \eqso(\logu{1}) \; \subsetneq \; \\ \eqso(\loge{2}) \; \subsetneq \; \eqso(\logu{2}) \; = \; \eqso(\fo)
+\end{multline*}
+\end{proposition}
+A natural question until this point is what is the connection between $\E{i}$ and $\U{i}$ and their corresponding classes $\eqso(\loge{i})$ and $\eqso(\logu{i})$.
+Clearly, we have that $\E{i} \subseteq \eqso(\loge{i})$ and $\E{i} \subseteq \eqso(\logu{i})$.
+The following result shows that this inclusion is strict only for the first levels of the hierarchy.
+\begin{proposition}
+	For $i \leq 1$, it holds that $\E{i} \subsetneq \eqso(\loge{i})$ and $\E{i} \subsetneq \eqso(\logu{i})$. Moreover, for $i \geq 2$ we have that $\E{i} = \eqso(\loge{i})$ and $\E{i} = \eqso(\logu{i})$.
+\end{proposition}
+
+Now, we study the complexity classes describe by this hierarchies.
+As the following result shows, $\eqso(\loge{0})$ defines only tractable counting functions and $\eqso(\loge{1})$ intractable counting functions but with an tractable decision problems.
+\begin{proposition}
+All functions defined in $\eqso(\loge{0})$ and $\eqso(\loge{1})$ can be computed in $\fp$ and $\totp$, respectively. Furthermore, every function defined in $\eqso(\loge{1})$ has a FPTRAS.
+\end{proposition}
+Therefore, in terms of counting complexity, the $\eqso$-hierarchy behaves exactly the same as the $\#\fo$-hierarchy.
+
+The next step is to study the closure properties of $\eqso$-hierarchy.
+An advantage of the $\eqso$-hierarchy is that, by its language syntax, all the classes are closed under addition and first and second order sum. 
+Furthermore, by Theorem~\ref{theo:binary-prod} one can easily see that all the hierarchy is also closed under binary product. 
+Then a natural question is whether the hierarchy is closed under subtraction.
+Formally, for any pair of functions $f,g$, we define $f - g$ as the function such that $(f - g)(\A) = f(\A)-g(\A)$ whenever $f(\A)>g(\A)$ and $0$ otherwise.
+As the next result shows, all classes in the $\eqso$-hierarchy is not closed under subtraction unless ${\sc P} = {\sc NP}$
+\begin{theorem} \label{sub-pnp}
+If $\eqso(\loge{i})$ or $\eqso(\logu{i})$ is closed under subtraction for $i > 0$, then {\sc P} = {\sc NP}.
+\end{theorem}
+\cristian{Martin, el resultado que tienes en el apendice se generaliza trivialmente para todas las clases ya que todas contienen la clase $\eqso(\loge{0})$.}
+
+By the previous result, we know that functions in the $\eqso$ hierarchy are not closed under subtraction.
+Then, a natural restriction to this question is to ask whether these classes are closed under subtraction by one, namely, if $\CC$ is a class of functions and $f \in \CC$, is $f-1 \in \CC$ where $1$ is the constant function that outputs $1$ for every structure.
+We do not know $\eqso(\logex{1})$ is closed under subtraction by one. However, if we extend $\logex{1}$ with $\fo$ predicates.
+\begin{theorem} \label{sigmafo-minusone}
+	$\eqso(\logex{1})$ is closed under substraction by one.
+\end{theorem}
+
 % We are interested in, for each $\fo$-fragment $\LL$, the biggest fragment of $\eqso$ that is contained in $\#\LL$.
@@
 
-We see that many of the results in Saluja et. al. \cite{DBLP:journals/jcss/SalujaST95} for $\#\LL$ still apply in $\eqso(\LL)$ for a given fragment $\LL$:
+%We see that many of the results in Saluja et. al. \cite{DBLP:journals/jcss/SalujaST95} for $\#\LL$ still apply in $\eqso(\LL)$ for a given fragment $\LL$:
+%
+%\begin{theorem} \label{eqso-sigma-zero-in-fp}
+%	For every $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$, the function $f$ over $\R$ defined as $f(\enc(\A)) = \sem{\alpha}(\A)$ is in $\fp$.
+%\end{theorem}
+%
+%\begin{theorem} \label{eqso-sigma-one-in-eqso-pi-one}
+%	For every $\eqso(\loge{1})$ formula $\alpha$ over a signature $\R$ there exists a $\eqso(\logu{1})$ formula $\beta$ over $\R$ such that $\sem{\alpha}(\A) = \sem{\beta}(\A)$ for every $\A\in\ostr[\R]$.
+%\end{theorem}
+%
+%
+%The {\em decision problem} associated to a function $f$ is defined by the language $L_f = \{\A \in \str \mid f(\A) > 0\}$.
+%
+%\begin{theorem} \label{decisionptime}
+%	The decision problem associated to a function in $\eqso(\logex{1})$ is in \textsc{P}.
+%\end{theorem}
 
-\begin{theorem} \label{eqso-sigma-zero-in-fp}
-	For every $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$, the function $f$ over $\R$ defined as $f(\enc(\A)) = \sem{\alpha}(\A)$ is in $\fp$.
-\end{theorem}
-
-\begin{theorem} \label{eqso-sigma-one-in-eqso-pi-one}
-	For every $\eqso(\loge{1})$ formula $\alpha$ over a signature $\R$ there exists a $\eqso(\logu{1})$ formula $\beta$ over $\R$ such that $\sem{\alpha}(\A) = \sem{\beta}(\A)$ for every $\A\in\ostr[\R]$.
-\end{theorem}
-
-The {\em decision problem} associated to a function $f$ is defined by the language $L_f = \{\A \in \str \mid f(\A) > 0\}$.
-
-\begin{theorem} \label{decisionptime}
-	The decision problem associated to a function in $\eqso(\logex{1})$ is in \textsc{P}.
-\end{theorem}
-
-For a given pair of functions $f,g$, we define $f \dotminus g$ as follows:
-\begin{eqnarray*}
-	(f \dotminus g)(\A) =
-	\begin{cases}
-		f(\A)-g(\A), & \text{if }f(\A)>g(\A) \\
-		0, & \text{if }f(\A) \leq g(\A).
-	\end{cases}
-\end{eqnarray*}
-for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction} if for every pair of functions $f,g \in \F$, it holds that $f \dotminus g \in \F$.
-
-\begin{theorem} \label{sub-pnp}
-	If $\eqso(\loge{1})$ is closed under substraction, then {\sc P} = {\sc NP}.
-\end{theorem}
-
-\begin{theorem} \label{sigma1strict}
-	$\eqso(\loge{1}) \subsetneq \eqso(\logex{1})$
-\end{theorem}
-
-For a given function $f$, we define $f \dotminus 1$ as follows:
-\begin{eqnarray*}
-	f \dotminus 1(\A) =
-	\begin{cases}
-		f(\A)-1, & \text{if }f(\A) > 0 \\
-		0, & \text{if }f(\A) = 0.
-	\end{cases}
-\end{eqnarray*}
-for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
-
-\begin{theorem} \label{sigmafo-minusone}
-	$\eqso(\logex{1})$ is closed under substraction by one.
-\end{theorem}
-
-\begin{theorem} \label{dnf-pars}
-	{\sc \#DNF} is hard for $\eqso(\loge{1})$ under parsimonious reductions. 
-\end{theorem}
-
-\begin{theorem} \label{nplusone-strict}
-	$\U{1}$ with $n$ open first-order variables is properly contained in $\U{1}$ with $n+1$ open first-order variables for $n\in\N$.  
-\end{theorem}
+%For a given pair of functions $f,g$, we define $f \dotminus g$ as follows:
+%\begin{eqnarray*}
+%	(f \dotminus g)(\A) =
+%	\begin{cases}
+%		f(\A)-g(\A), & \text{if }f(\A)>g(\A) \\
+%		0, & \text{if }f(\A) \leq g(\A).
+%	\end{cases}
+%\end{eqnarray*}
+%for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction} if for every pair of functions $f,g \in \F$, it holds that $f \dotminus g \in \F$.
+%
+%\begin{theorem} \label{sub-pnp}
+%	If $\eqso(\loge{1})$ is closed under substraction, then {\sc P} = {\sc NP}.
+%\end{theorem}
+%
+%\begin{theorem} \label{sigma1strict}
+%	$\eqso(\loge{1}) \subsetneq \eqso(\logex{1})$
+%\end{theorem}
+%
+%For a given function $f$, we define $f \dotminus 1$ as follows:
+%\begin{eqnarray*}
+%	f \dotminus 1(\A) =
+%	\begin{cases}
+%		f(\A)-1, & \text{if }f(\A) > 0 \\
+%		0, & \text{if }f(\A) = 0.
+%	\end{cases}
+%\end{eqnarray*}
+%for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
+%
+%\begin{theorem} \label{sigmafo-minusone}
+%	$\eqso(\logex{1})$ is closed under substraction by one.
+%\end{theorem}
+%
+%\begin{theorem} \label{dnf-pars}
+%	{\sc \#DNF} is hard for $\eqso(\loge{1})$ under parsimonious reductions. 
+%\end{theorem}
+%
+%\begin{theorem} \label{nplusone-strict}
+%	$\U{1}$ with $n$ open first-order variables is properly contained in $\U{1}$ with $n+1$ open first-order variables for $n\in\N$.  
+%\end{theorem}

### 2016-12-23 LICS/syntactic.tex
Avance en sección de saluja.
Sección de Saluja primera versión.

@@
 
-\subsection{Saluja Classes}
+In this section we study the fragment of $\eqso(\LL)$ when $\LL$ is a boolean logic contained in $\fo$.
+
+We say that a formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\Sigma$-prenex normal form} if $\alpha$ is of the form:
+\[
+\alpha := \sa{\bar{X}} \sa{\bar{x}} \varphi(\bar{X}, \bar{x})
+\]
+where $\bar{X}$ and $\bar{x}$ is a sequence of $0$ or more second and first order variables, respectively.
+In particular, note that a boolean formula is always in $\Sigma$-prenex normal form.
+Then we say that a formula $\alpha$ is in \emph{sum normal form} if $\alpha$ is of the form $c + \Sigma_{i=1}^n \alpha_i$ where $c$ is a non-negative constant and each $\alpha_i$ is in $\Sigma$-prenex normal form.
+Next we show that each formula in $\eqso(\LL)$ can be converted in \emph{sum normal form} but not always in prenex normal form.
+
+\begin{theorem}
+	Every formula in $\eqso(\LL)$ can be rewritten in \emph{sum normal form}.
+Furthermore, there exists a language $\LL$ and a formula $\alpha$ such that there does not exist a formula in $\Sigma$-prenex normal form equivalent to $\alpha$.
+\end{theorem}
+
+An interesting question here is whether the multiplicative operators in $\qso$ can be defined in $\eqso(\LL)$.
+As the following result shows, if $\LL$ is closed under conjunction, then the binary product can be defined in  $\eqso(\LL)$.
+
+\begin{theorem}\label{theo:binary-prod}
+	If $\LL$ is closed under conjunction, then binary product can be defined in $\eqso(\LL)$.
+\end{theorem}
+
+\subsection{Prenex-normal form classes}
+\input{saluja}
 
 \subsection{Horn Counting Classes}
+\input{horn-counting}

### 2016-12-26 LICS/appendix.tex
26 dic

@@
 First, for a given signature $\R$ and logic fragment $\LL$ we define a {\em valid form in $\LL$} (or just {\em valid form}).
-A formula $\alpha$ is in valid form in $\LL$ if (1) it is equal to $\varphi$, $1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{x}}\sa{\bar{X}}\varphi$ or $\sa{\bar{x}}\sa{\bar{X}}1$ for some $\LL$ formula $\varphi$ and some tuple of first-order (second-order) variables $\bar{x}$ ($\bar{X}$), or (2) if it is equal to $(\beta\add\gamma)$ where $\beta$ and $\gamma$ are formulas in valid form.
+A formula $\alpha$ is in valid form in $\LL$ if (1) it is equal to $\varphi$, $1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}1$ for some $\LL$ formula $\varphi$ and some tuple of first-order (second-order) variables $\bar{x}$ ($\bar{X}$), or (2) if it is equal to $(\beta\add\gamma)$ where $\beta$ and $\gamma$ are formulas in valid form.
 Given a $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$ we will recursively construct a formula $\tau(\alpha)$ in valid form such that $\alpha(\enc(\A)) = \tau(\alpha)(\enc(\A))$ for each $\A\in\ostr[\R]$.
@@
 For each $\beta_i$, if $\beta_i = 1$ or $\beta_i = \varphi$, then $\sem{\beta_i}(\A)$ can be computed in polynomial time.
-If $\beta_i = \sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}$, if $\beta_i = \sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = 2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, if $\beta_i = \sa{\bar{x}}\sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, all of which can be computed in polynomial time.
+If $\beta_i = \sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}$, if $\beta_i = \sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = 2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, if $\beta_i = \sa{\bar{X}}\sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, all of which can be computed in polynomial time.
 Otherwise, $\beta_i$ is $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, and then $\sem{\beta_i}(\A) = f_{\varphi}(\enc(\A))$, and $f_{\varphi}\in \#\loge{0}$ which implies that $f_{\varphi}(\enc(\A))$ can be computed in polynomial time \cite{DBLP:journals/jcss/SalujaST95}.
@@
 Let $\tau(\alpha) = \beta_1\add\beta_2\add\cdots\add\beta_m$ defined as in the proof of Theorem \ref{eqso-sigma-zero-in-fp}.
-This is, each $\beta_i$ is of the form $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{x}}\sa{\bar{X}}1$ or $\sa{\bar{x}}\sa{\bar{X}}\varphi$, where $\varphi$ is a formula in $\loge{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables.
+This is, each $\beta_i$ is of the form $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}\sa{\bar{x}}1$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, where $\varphi$ is a formula in $\loge{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables.
 For each $\beta_i$ we will construct a $\gamma_i$ such that $\gamma_i \in \eqso(\logu{1})$ and $\sem{\beta_i}(\A) = \sem{\gamma_i}(\A)$ for each $\A\in\ostr[\R]$.
-Since the forms $1$, $\sa{\bar{x}}1$, $\sa{\bar{X}}1$ and $\sa{\bar{x}}\sa{\bar{X}}1$ are already valid in $\eqso(\logu{1})$, we suppose $\beta_i$ is in one of the remaining forms.
-Without loss of generality, let $\beta_i = \sa{\bar{x}}\sa{\bar{X}}\exists\bar{y}\,\varphi(\bar{x},\bar{y},\bar{X})$ where $\varphi(\bar{x},\bar{y},\bar{X})$ is a quantifier-free formula.
+Since the forms $1$, $\sa{\bar{x}}1$, $\sa{\bar{X}}1$ and $\sa{\bar{X}}\sa{\bar{x}}1$ are already valid in $\eqso(\logu{1})$, we suppose $\beta_i$ is in one of the remaining forms.
+Without loss of generality, let $\beta_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y},\bar{X})$ where $\varphi(\bar{x},\bar{y},\bar{X})$ is a quantifier-free formula.
 Similarly as in the proof by Saluja et al.
@@
 
-\subsection{Proof of Theorem \ref{one-sigma-zero}}
-
-Towards a contradiction, suppose that there exists a constant function $f(\A) = s > 0$ which is expressible in $\E{0}$.
+%\subsection{Proof of Theorem \ref{one-sigma-zero}}
+%
+%Towards a contradiction, suppose that there exists a constant function $f(\A) = s > 0$ which is expressible in $\E{0}$.
 This is, there is a formula $\psi(x,X)$ such that $f_{\psi(x,X)}(\A) = \vert \{ \langle e,P\rangle \mid \A\models\psi(e,P) \} \vert = s$ for each $\A\in\ostr$.
@@
 
-\textbf{Proof of Theorem \ref{mult-sigma-zero}}
-
-Let $\alpha \in \sqso(\loge{0})$ with binary product.
+%\textbf{Proof of Theorem \ref{mult-sigma-zero}}
+%
+%Let $\alpha \in \sqso(\loge{0})$ with binary product.
 This is, $\alpha$ is defined by the grammar $\alpha := \sa{x}\varphi \ \mid \ \sa{X}\varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \ (\alpha\mult\alpha)$, where $\varphi \in \loge{0}$.
@@
 Since $\delta(y)\mult\gamma$ has $n-1$ algebraic operators, it can be expressed as product-free, and then $\alpha$ can be expressed as product-free.
-
-\textbf{Proof of Theorem \ref{fo-prod-sigma-zero}}
-
-The resulting grammar is $\alpha := \sa{x}\varphi \ \mid \ \sa{X} \varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \pa{x}\alpha$.
+%
+%\textbf{Proof of Theorem \ref{fo-prod-sigma-zero}}
+%
+%The resulting grammar is $\alpha := \sa{x}\varphi \ \mid \ \sa{X} \varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \pa{x}\alpha$.
 Consider a more restricted grammar that replaces $\pa{x}\alpha$ by $\pa{x}\sa{y}\varphi$.
@@
 This formula produces a constant function equal to 1, which as shown in Theorem \ref{one-sigma-zero} is not expressible in $\E{0}$.
-
-\textbf{Proof of Theorem \ref{fp1}}
-
-Let $f \in \XE{0}$, and let $\varphi(\X,\x)$ be and {\sc FO}-extended quantifier-free $\L$-formula such that:
-\begin{eqnarray*}
-	f(\A) &=& \vert \{\langle \P,\e  \rangle \mid \A \models \varphi(\P,\e) \} \vert
-\end{eqnarray*}
-for each $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^m$ and $\P = (P_1,\ldots,P_q)$ is a tuple of predicates.
+%
+%\textbf{Proof of Theorem \ref{fp1}}
+%
+%Let $f \in \XE{0}$, and let $\varphi(\X,\x)$ be and {\sc FO}-extended quantifier-free $\L$-formula such that:
+%\begin{eqnarray*}
+%	f(\A) &=& \vert \{\langle \P,\e  \rangle \mid \A \models \varphi(\P,\e) \} \vert
+%\end{eqnarray*}
+%for each $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^m$ and $\P = (P_1,\ldots,P_q)$ is a tuple of predicates.
 We will now show that computing $f(\A)$ can be done in polynomial time.
-
-For each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$, let $R_{\beta}$ be a predicate of arity $m$.
+%
+%For each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$, let $R_{\beta}$ be a predicate of arity $m$.
 Let $\A^\prime = \langle A, \S^{\A}, R_{\beta}^{\A^\prime}, \leq^{\A} \rangle \in \str$, where $R_{\beta}^{\A^\prime} = \{\d \mid \A \models \beta(\d)\}$.
@@
 Therefore, generating $R_{\beta}^{\A^\prime}$ can also be done in polynomial time.
-
-Let $\psi(\X,\x)$ be obtained by replacing each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$ by $R_{\beta}(\x)$.
+%
+%Let $\psi(\X,\x)$ be obtained by replacing each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$ by $R_{\beta}(\x)$.
 Also, let $g = f_{\psi(\X,\x)}$.
@@
 
-\textbf{Proof of Theorem \ref{sigmafo-minusone}}
+\subsection{Proof of Theorem \ref{sigmafo-minusone}}
 
@@
 We will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_n$ be defined as in \ref{eqso-sigma-zero-in-fp}, that is, each $\beta_i$ is equal to $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{x}}\sa{\bar{X}}1$ or $\sa{\bar{x}}\sa{\bar{X}}\varphi$, where $\varphi$ is a formula in $\logex{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables.
-We further assume that each $\beta_i$ is either $\sa{\bar{x}}\varphi$ or $\sa{\bar{x}}\sa{\bar{X}}\varphi$, since each $1$, $\varphi$ and $\sa{\bar{X}}\varphi$ can be replaced by the formulas $\sa{x}\min(x)$, $\sa{x}(\varphi\wedge\min(x))$ and $\sa{x}\sa{\bar{X}}(\varphi\wedge\min(x))$ respectively, while maintaining the value of $\sem{\alpha}(\A)$.
+Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_n$ be defined as in \ref{eqso-sigma-zero-in-fp}, that is, each $\beta_i$ is equal to $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}\sa{\bar{x}}1$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, where $\varphi$ is a formula in $\logex{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables.
+We further assume that each $\beta_i$ is either $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, since each $1$, $\varphi$ and $\sa{\bar{X}}\varphi$ can be replaced by the formulas $\sa{x}\min(x)$, $\sa{x}(\varphi\wedge\min(x))$ and $\sa{x}\sa{\bar{X}}(\varphi\wedge\min(x))$ respectively, while maintaining the value of $\sem{\alpha}(\A)$.
 
 The intuition behind our reasoning is separated in three points: (1) For each $\beta_i$ of the form $\sa{\bar{x}}\varphi$, the formula $\kappa(\beta_i)$ will count every tuple $\bar{x}$ that satisfies $\varphi$ except for the lexicographically smallest one.
-(2) For each $\beta_i$ of the form $\sa{\bar{x}}\sa{\bar{X}}\varphi$, the formula $\kappa(\beta_i)$ will isolate the smallest $\bar{X}$ that satisfies $\varphi$, and exclude the lexicographically smallest tuple $\bar{x}$ that satisfies $\varphi(\bar{X})$.
-And (3) if $\alpha = (\beta\add\sa{\bar{x}}\varphi)$ or $\alpha = (\beta\add\sa{\bar{x}}\sa{\bar{X}}\varphi)$, the formula $\kappa(\alpha)$ will exclude the lexicographically smallest tuple that satisfies $\varphi$ if and only if $\beta$ is equal to 0.
+(2) For each $\beta_i$ of the form $\sa{\bar{X}}\sa{\bar{x}}\varphi$, the formula $\kappa(\beta_i)$ will isolate the smallest $\bar{X}$ that satisfies $\varphi$, and exclude the lexicographically smallest tuple $\bar{x}$ that satisfies $\varphi(\bar{X})$.
+And (3) if $\alpha = (\beta\add\sa{\bar{x}}\varphi)$ or $\alpha = (\beta\add\sa{\bar{X}}\sa{\bar{x}}\varphi)$, the formula $\kappa(\alpha)$ will exclude the lexicographically smallest tuple that satisfies $\varphi$ if and only if $\beta$ is equal to 0.
+
+\vspace{1em}
 
@@
 
-For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{x}}\sa{\bar{X}}\exists\bar{y}\,\varphi(\bar{x},\bar{X},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha)$ procedurally as follows: Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
+\vspace{1em}
+
+For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha)$ procedurally as follows: Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
 We suppose that $\varphi$ is in a DNF form that leaves formulas that do not mention $\bar{X}$ as is.
 If it is not, we convert $\varphi$ to this form with a standard transformation algorithm.
-Let $\varphi(\bar{x},\bar{X},\bar{y}) = \bigvee_{i = 1}^m\varphi_i(\bar{x},\bar{X},\bar{y})$ where each $\varphi_i$ has the form:
-$$
-\varphi_i(\bar{x},\bar{X},\bar{y}) = (\text{conjunction of $X_i$'s}) \wedge (\text{conjunction of $\neg X_i$'s})  \wedge (\text{an $\fo$ formula that does not mention any $X_i$}).
+Let $\varphi(\bar{X},\bar{x},\bar{y}) = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$ where each $\varphi_i$ has the form:
+$$
+\varphi_i(\bar{X},\bar{x},\bar{y}) = (\text{conjunction of $X_i$'s}) \wedge (\text{conjunction of $\neg X_i$'s})  \wedge (\text{an $\fo$ formula that does not mention any $X_i$}).
 $$
@@
 If not, we add new first-order variables from $\fv$ to $\bar{y}$ so that no variable is repeated, replace them in $\varphi_i^{+}\wedge\varphi_i^{-}$ and add their respective equalities in $\varphi_i^{\fo}$ so that the new formula is equivalent.
-For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (x_1,x_2,x_3,x_4,y)$ and $\varphi_i := X(x_1,x_2) \wedge \neg X(x_3,x_4) \wedge x < y \wedge x_1 = x \wedge x_2 = y \wedge x_3 = x \wedge x_4 = x.$
+For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
 
@@
 let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
-The formula $\varphi_i(\bar{x},\bar{X},\bar{y})$ is converted into $\bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}}\varphi_i(\bar{x},\bar{X},\bar{y}) \wedge \theta^j(\bar{x},\bar{y})$.
+The formula $\varphi_i(\bar{X},\bar{x},\bar{y})$ is converted into $\bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}}\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \theta^j(\bar{x},\bar{y})$.
 Consider each of these disjuncts as a new $\varphi_i$.
 
-Now let $\bar{u}$ be the tuple of every first-order variable mentioned in $\varphi_i^{+}$ and let $\bar{v}$ be such that $(\bar{x},\bar{y}) = (\bar{u},\bar{v})$.
-We define a formula $\mu_i$ that is satisfied only by the lexicographically smallest $\bar{u}$ that satisfies $\varphi_i^{+}$ for some $\bar{v}$:
-$$
-\mu_i(\bar{u}) = \exists\bar{v}\,\varphi_i^{\fo}(\bar{u},\bar{v})\wedge\forall\bar{u}'\forall\bar{v}'(\varphi_i^{\fo}(\bar{u}',\bar{v}')\to\bar{u}\leq\bar{u}').
-$$
-Lastly, we define $\kappa(\varphi_i)$:
-$$
-\kappa(\varphi_i) := 
-$$
-%Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$.
+After the previous assumptions, we do the following: for each $X_j\in \bar{X}$ we check every instance of $X_j(\bar{w})$ in $\varphi^{+}_i$ and every instance of $X_j(\bar{z})$ in $\varphi^{-}_i$, where $\bar{w}$ and $\bar{z}$ are subtuples of $(\bar{x},\bar{y})$.
+If the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{w} = \bar{z}$, the entire formula $\varphi_i$ is removed from $\varphi$.
+
+Now let $\bar{v}$ be the tuple of every first-order variable mentioned in $\varphi_i^{+}$ and let $\bar{u}$ be such that $(\bar{x},\bar{y}) = (\bar{u},\bar{v})$.
+We define a formula $\mu_i$ that is satisfied only by the lexicographically smallest $\bar{v}$ that satisfies $\varphi_i^{\fo}$ for some $\bar{u}$:
+$$
+\mu_i(\bar{v}) = \exists\bar{u}\,\varphi_i^{\fo}(\bar{u},\bar{v})\wedge\forall\bar{u}'\forall\bar{v}'(\varphi_i^{\fo}(\bar{u}',\bar{v}')\to\bar{v}\leq\bar{v}').
+$$
+The following is a pivotal formula for this proof.
+Consider some ordered finite structure $\A$ over $\R$.
+Given our conditions for $\varphi_i$, if $\A\models\varphi_i$ and $\A\models\varphi^{\fo}_i(\bar{u},\bar{v})$ then for that $\bar{v}$ we can define a $\bar{X}$ that has exactly the tuples mentioned in $\bar{v}$ and nothing else.
+This is {\em the smallest $\bar{X}$ that satisfies $\varphi_i$ over $\bar{v}$}.
+And if $\bar{v}$ is the lexicographically smallest such that $\A\models\varphi_i(\bar{v})$, then it is the {\em the smallest $\bar{X}$ that satisfies $\varphi_i$}.
+The following formula is satisfied by every pair $(\bar{X},\bar{x})$, except for the one formed by the smallest $\bar{X}$ that satisfies $\varphi_i$ and the lexicographically smallest $\bar{x}$ that satisfies $\varphi_i(\bar{X})$.
+And if no pair satisfies $\varphi_i$, the following formula is satisfied by every pair.
+$$
+\psi_i(\bar{X},\bar{x}) = \exists\bar{v}(\mu_i(\bar{v})\wedge(\neg\varphi^{+}_i(\bar{X},\bar{v})\vee\bigvee_{X \in \bar{X}} \exists\bar{z}(X(\bar{z}) \wedge \bigwedge\limits_{\substack{\text{instances of }X(\bar{w}) \\ \text{in }\varphi^{+}_i(\bar{X},\bar{v})}}\bar{w}\neq\bar{z}))) \vee \exists\bar{x}'(\exists\bar{y}'\varphi_i(\bar{X},\bar{x}',\bar{y}')\wedge \bar{x}'<\bar{x})\vee\neg\exists\bar{v}\mu_i(\bar{v}) .
+$$
+\begin{claim}
+	Let $\varphi_i(\bar{X},\bar{x},\bar{y}) = \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X},\bar{x},\bar{y})$ be a quantifier-free formula that respects all the mentioned assumptions.
+For a given ordered structure $\A$ with domain $A$, if $\A\models\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists some $\bar{B}$ over $\A$ and $\bar{b}\in A^{\length(\bar{x})}$ such that $\A\models\varphi_i(\bar{B},\bar{b},\bar{a})$ for some $\bar{a}\in A^{\length(\bar{y})}$.
+\end{claim}
+\begin{proof}
+	Let $\A$ be an ordered structure with domain $A$ and let $\bar{a}\in A^{\length(\bar{y})}$ and $\bar{b}\in A^{\length(\bar{x})}$ such that $\A\models\varphi^{\fo}_i(\bar{b},\bar{a})$.
+Define $\bar{B} = (B_1,\ldots,B_{\length(\bar{X})})$ as $B_j = \{\bar{c}\mid\bar{c}\text{ is a subtuple of $(\bar{b},\bar{a})$ and $X_j(\bar{c})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{b},\bar{a})$}\}$.
+Towards a contradiction, suppose that $\A\not\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
+By the choice of $\bar{a}$ and $\bar{b}$, and construction of $\bar{B}$ it is clear that $\A\models\varphi^{\fo}_i(\bar{b},\bar{a})\wedge\varphi^{+}_i(\bar{B},\bar{b},\bar{a})$, so we have that $\A\not\models\varphi^{-}_i(\bar{B},\bar{b},\bar{a})$.
+Let $B_t(\bar{c})$ be such that $X_t(\bar{c})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{b},\bar{a})$ and $\A\not\models\neg B_t(\bar{c})$.
+This is, $\bar{c}\in B_t$.
+By the construction of $\bar{B}$ we have that $X_t(\bar{c})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{b},\bar{a})$.
+From our assumptions, we have that (1) every first-order variable is mentioned at most once in $\varphi^{-}_i(\bar{X},\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X},\bar{x},\bar{y})$, (2) $\varphi^{\fo}_i$ defines an ordered partition over $(\bar{x},\bar{y})$, and (3) there are no $X_j(\bar{w})$ in $\varphi^{+}_i$ and $X_j(\bar{z})$ in $\varphi^{-}_i$ such that the ordered partition defined by $\varphi^{\fo}_i$ satisfies $\bar{x} = \bar{z}$.
+Since $\bar{c}$ is a subtuple of $(\bar{b},\bar{a})$ that satisfies $\varphi^{\fo}_i$, and $X_t(\bar{w})$ mentioned in $\varphi^{+}_i$ and $X_t(\bar{z})$ mentioned in $\varphi^{-}_i$ were both assigned the value $\bar{c}$, then the ordered partition satisfies $\bar{w} = \bar{z}$, which follows to a contradiction.
+\end{proof}
+We define $\chi_i = \exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$.
+Now recall that $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$, where each $\varphi_i$ satisfies all the previous assumptions.
+For each $\varphi_i$ we define:
+$$
+\varphi_i'(\bar{X},\bar{x},\bar{y}) = \varphi_i(\bar{X},\bar{x},\bar{y})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\vee\psi_2(\bar{X},\bar{x}))\wedge(\chi_1\vee\chi_2\vee\psi_2(\bar{X},\bar{x}))\wedge\cdots\wedge(
+\bigvee_{j = 1}^{j = i-1}\chi_j\vee\psi_i(\bar{X},\bar{x})),
+$$
+and lastly, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x},\bar{y})$.
+
+\vspace{1em}
+
+For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y}))$ for some algebraic formula $\beta$ and some quantifier-free formula $\varphi$, we define $\kappa(\alpha)$ as follows: First, perform the same transformations to $\varphi(\bar{X},\bar{x},\bar{y})$ as in the previous case.
+Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$ where each $\varphi_i$ satisfies the previous assumptions.
+We also use the previously defined formulas $\chi_i$ and $\psi_i$.
+
+We recursively define a function $\lambda$ as follows.
+If $\alpha = \sa{\bar{x}}\exists\bar{y}\varphi(\bar{x},\bar{y})$, then $\lambda(\alpha) = \exists\bar{x}'\exists\bar{y}'\varphi(\bar{x}',\bar{y}')$.
+If $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ and $\lambda(\alpha) = \chi_1\vee\cdots\vee\chi_m$ as each $\chi_i$ was previously defined.
+If $\alpha = (\beta + \gamma)$, then $\lambda(\alpha) = \lambda(\beta) \vee \lambda(\gamma)$.
+Note that for a given ordered structure $\A$, then $\A\models\lambda(\alpha)$ if and only if $\sem{\alpha}(\A) > 0$.
+
+Now, for each $\varphi_i$ we define:
+$$
+\varphi_i'(\bar{X},\bar{x},\bar{y}) = \varphi_i(\bar{X},\bar{x},\bar{y})\wedge[\lambda(\beta)\vee[\psi_1(\bar{X},\bar{x})\wedge(\chi_1\vee\psi_2(\bar{X},\bar{x}))\wedge(\chi_1\vee\chi_2\vee\psi_2(\bar{X},\bar{x}))\wedge\cdots\wedge(
+\bigvee_{j = 1}^{j = i-1}\chi_j\vee\psi_i(\bar{X},\bar{x}))]].
+$$
+And lastly $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x},\bar{y})$.
+%%Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$.
 There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$.
 This separates the proof in three cases:
-\begin{enumerate}
-%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
-%	\begin{eqnarray*}
-%		f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
-%	\end{eqnarray*}
-%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^d$.
+%\begin{enumerate}
+%%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
+%%	\begin{eqnarray*}
+%%		f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
+%%	\end{eqnarray*}
+%%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^d$.
 Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily.
 First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
-%	\begin{eqnarray*}
-%		\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
-%	\end{eqnarray*}
-%	This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
-%	\begin{eqnarray*}
-%		\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
-%	\end{eqnarray*}
-%	If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
+%%	\begin{eqnarray*}
+%%		\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
+%%	\end{eqnarray*}
+%%	This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
+%%	\begin{eqnarray*}
+%%		\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
+%%	\end{eqnarray*}
+%%	If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
 Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$.
 Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E{1}.$
-%	
-	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
-	\begin{eqnarray*}
-		f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
-	\end{eqnarray*}
-	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
-	\begin{eqnarray}
-	\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
-	\end{eqnarray}
-	where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
+%%	
+%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
+%	\begin{eqnarray*}
+%		f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
+%	\end{eqnarray*}
+%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
+%	\begin{eqnarray}
+%	\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
+%	\end{eqnarray}
+%	where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
 Note that $\theta(\x)$ also mentions all variables in $\x$.
@@
 As an example, the following formula is of this form:
-	\begin{multline*}
-	\varphi(\x,\X) =  X_1(x_1,x_2) \wedge X_3(x_3) \wedge X_2(x_4,x_5) \wedge X_3(x_6) \wedge \neg X_1(x_7,x_8) \wedge \\ (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 ) \wedge \\ \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big),
-	\end{multline*}
-	where $\x = (x_1,x_2,x_3,x_4,x_5,x_6,x_7,x_8)$ and $\X = (X_1,X_2,X_3)$.
+%	\begin{multline*}
+%	\varphi(\x,\X) =  X_1(x_1,x_2) \wedge X_3(x_3) \wedge X_2(x_4,x_5) \wedge X_3(x_6) \wedge \neg X_1(x_7,x_8) \wedge \\ (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 ) \wedge \\ \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big),
+%	\end{multline*}
+%	where $\x = (x_1,x_2,x_3,x_4,x_5,x_6,x_7,x_8)$ and $\X = (X_1,X_2,X_3)$.
 Here, $n = 4$, $\lambda(1) = 1$, $\lambda(2) = \lambda(4) = 3$ and $\lambda(3) = 2$, $\x_1 = (x_1,x_2)$, $\x_2 = (x_3)$, $\x_3 = (x_4,x_5)$, $\x_4 = (x_6)$ and $\y = (x_7,x_8)$.
 Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_7,x_8)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 )$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2,x_5,x_8\},\{x_6,x_7\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big)$.
-
-	Similarly to the previous proof, we would like to eliminate the {\em lexicographically smallest}\footnote{We consider the lexicographically smallest tuple of predicates as the one in which its predicates contain the lexicographically smallest tuples and do not contain any more tuples than those} tuple of predicates that satisfies the formula \eqref{f1}.
+%	
+%	Similarly to the previous proof, we would like to eliminate the {\em lexicographically smallest}\footnote{We consider the lexicographically smallest tuple of predicates as the one in which its predicates contain the lexicographically smallest tuples and do not contain any more tuples than those} tuple of predicates that satisfies the formula \eqref{f1}.
 Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
 We now define
-	\begin{multline*}
-	\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to (\u_1,\dots,\u_n)<(\v_1,\dots,\v_n)\Big)\Big],
-	\end{multline*}
-	where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
+%	\begin{multline*}
+%	\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to (\u_1,\dots,\u_n)<(\v_1,\dots,\v_n)\Big)\Big],
+%	\end{multline*}
+%	where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
 Note that $\alpha_{\min}$ is satisfied only by the lexicographically smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
 Our new formula is
-	\begin{multline}
-	\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge \\ \exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg].
-	\end{multline}
-	We now show a result by which the main proof will follow.
-	\begin{lemma} \label{first}
-		$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
-	\end{lemma}
-	\begin{proof}
-		Let $\A \in \str$.
+%	\begin{multline}
+%	\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge \\ \exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg].
+%	\end{multline}
+%	We now show a result by which the main proof will follow.
+%	\begin{lemma} \label{first}
+%		$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
+%	\end{lemma}
+%	\begin{proof}
+%		Let $\A \in \str$.
 Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
@@
 We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
-		\begin{enumerate}
-			\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$.
+%		\begin{enumerate}
+%			\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$.
 That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$).
@@
 In particular, $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which contradicts the initial assumption that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-			\item[(b)] Note that if $\A \models \alpha_{\min}(\c_1,\ldots,\c_n)$, then necessarily $\c_i = \d_i$ for $i\in\{1,\ldots,n\}.$ However, by the construction of $\P$, we see that 
-			$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \d_j\Big).$$ Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
-			\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
+%			\item[(b)] Note that if $\A \models \alpha_{\min}(\c_1,\ldots,\c_n)$, then necessarily $\c_i = \d_i$ for $i\in\{1,\ldots,n\}.$ However, by the construction of $\P$, we see that 
+%			$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \d_j\Big).$$ Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
+%			\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
 We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$.
@@
 On both cases, we have a contradiction.
-		\end{enumerate}
-		With this, we conclude that for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-		Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
+%		\end{enumerate}
+%		With this, we conclude that for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
+%		
+%		Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
 Let $\P$ be an arbitrary assignment to $\X$.
@@
 And so, for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ it holds that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1,$ which was to be shown.
-	\end{proof}
-	We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
+%	\end{proof}
+%	We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
 By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\gamma_i(\x,\X)$ with $i\in\{1,\ldots,\ell\}$ such that
-	\begin{eqnarray*}
-		\varphi(\x,\X) &\equiv& \gamma_1(\x,\X) \vee \gamma_2(\x,\X) \vee \dots  \vee \gamma_{\ell}(\x,\X),
-	\end{eqnarray*}
-	where, for every $i\in\{1,\ldots,\ell\}$, 
-	\begin{eqnarray*}
-		\gamma_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \gamma^{-}_i(\X,\y_i)  \wedge \delta_i(\x).
-	\end{eqnarray*}
-	The function $\lambda_i$ is defined analogously to $\lambda$ of the first part of the proof.
+%	\begin{eqnarray*}
+%		\varphi(\x,\X) &\equiv& \gamma_1(\x,\X) \vee \gamma_2(\x,\X) \vee \dots  \vee \gamma_{\ell}(\x,\X),
+%	\end{eqnarray*}
+%	where, for every $i\in\{1,\ldots,\ell\}$, 
+%	\begin{eqnarray*}
+%		\gamma_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \gamma^{-}_i(\X,\y_i)  \wedge \delta_i(\x).
+%	\end{eqnarray*}
+%	The function $\lambda_i$ is defined analogously to $\lambda$ of the first part of the proof.
 The tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y_i) = \x.$ The formulas $\gamma^{-}_i(\X,\y_i)$, $\delta_i(\x)$ are defined analogously to $\varphi^{-}(\X,\y)$ and $\beta(\x)\,$\footnote[2]{Note that $\delta_i(\x)$ can include subformulas of the form $(\u = \u)$.} respectively (see \eqref{f1}).
@@
 Note that for every formula $\eta(\x)$,
-	\begin{eqnarray*}
-		\exists\x\:\eta(\x) &\equiv& \exists\x(\eta(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\eta(\x)\wedge\theta^{g(d)}(\x)).
-	\end{eqnarray*}
-	We define the following formulas $\xi_i(\x,\X)$, for $i\in\{1,\ldots,m\}$ where $m = \ell \cdot g(d)$, as follows: 
-	\begin{eqnarray*}
-		\xi_1(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^1(\x), \\
-		& \vdots & \\
-		\xi_{g(d)}(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
-		\xi_{g(d)+1}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^1(\x), \\
-		& \vdots & \\
-		\xi_{2\cdot g(d)}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
-		& \vdots & \\
-		\xi_{(\ell-1) g(d)+1}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^1(\x), \\
-		& \vdots & \\
-		\xi_{\ell \cdot g(d)}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^{g(d)}(\x).
-	\end{eqnarray*}
-	Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\str.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
+%	\begin{eqnarray*}
+%		\exists\x\:\eta(\x) &\equiv& \exists\x(\eta(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\eta(\x)\wedge\theta^{g(d)}(\x)).
+%	\end{eqnarray*}
+%	We define the following formulas $\xi_i(\x,\X)$, for $i\in\{1,\ldots,m\}$ where $m = \ell \cdot g(d)$, as follows: 
+%	\begin{eqnarray*}
+%		\xi_1(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^1(\x), \\
+%		& \vdots & \\
+%		\xi_{g(d)}(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
+%		\xi_{g(d)+1}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^1(\x), \\
+%		& \vdots & \\
+%		\xi_{2\cdot g(d)}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
+%		& \vdots & \\
+%		\xi_{(\ell-1) g(d)+1}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^1(\x), \\
+%		& \vdots & \\
+%		\xi_{\ell \cdot g(d)}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^{g(d)}(\x).
+%	\end{eqnarray*}
+%	Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\str.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
 We use an injective\footnote[3]{We need this function to be injective because this way we can assure each one of the $k$ satisfiable disjuncts to be represented} function $\rho:\{1,\ldots,k\}\to\{1,\ldots,m\}$ such that $\varphi_i(\x,\X) = \xi_{\rho(i)}(\x,\X)$ is satisfiable, for every $i\in\{1,\ldots,k\}.$
-
-	We can conclude that
-	\begin{eqnarray*}
-		\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X),
-	\end{eqnarray*}
-	and for every $i\in\{1,\ldots,k\}$,
-	\begin{eqnarray*}
-		\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
-	\end{eqnarray*}
-	where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\y_i$ is a $p_i$-tuple of variables in $\x$, $\varphi_i^{-}(\X,\y_i)$ is a conjunction of negated predicates in $\X$, $\theta_i(\x)$ defines a total order on a partition of $\x$, and $\beta_i(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
+%	
+%	We can conclude that
+%	\begin{eqnarray*}
+%		\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X),
+%	\end{eqnarray*}
+%	and for every $i\in\{1,\ldots,k\}$,
+%	\begin{eqnarray*}
+%		\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
+%	\end{eqnarray*}
+%	where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\y_i$ is a $p_i$-tuple of variables in $\x$, $\varphi_i^{-}(\X,\y_i)$ is a conjunction of negated predicates in $\X$, $\theta_i(\x)$ defines a total order on a partition of $\x$, and $\beta_i(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
 Let $\alpha_i(\x) = \theta_i(\x)\wedge\beta_i(\x).$
-
-	%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
-	%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
-
-	\begin{claim} \label{alpha}
-		Let $\A\in\str$ and $i\in\{1,\ldots,k\}$. For every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, it holds that $f_{\varphi_i(\s,\X)}(\A) > 0.$
-	\end{claim}
-	\begin{proof}
-		Let $\s = (\s_1,\ldots,\s_{n_i},\t)$ such that $\A\models\alpha_i(\s)$, and let $\P = (\P_1,\ldots,\P_r)$ where $P_i = \bigcup_{j\in[1,n_i]:\:\lambda(j)=i}\{s_i\}$.
+%	
+%	%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
+%	%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
+%	
+%	\begin{claim} \label{alpha}
+%		Let $\A\in\str$ and $i\in\{1,\ldots,k\}$. For every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, it holds that $f_{\varphi_i(\s,\X)}(\A) > 0.$
+%	\end{claim}
+%	\begin{proof}
+%		Let $\s = (\s_1,\ldots,\s_{n_i},\t)$ such that $\A\models\alpha_i(\s)$, and let $\P = (\P_1,\ldots,\P_r)$ where $P_i = \bigcup_{j\in[1,n_i]:\:\lambda(j)=i}\{s_i\}$.
 We will show that $\A\models\varphi_i(\s,\P)$.
@@
 However, as we mentioned previously, all of such disjuncts have been eliminated, which leads to a contradiction.
-	\end{proof}
-
-	Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\exists\x\:\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\Q$ such that $\A\models\exists\x\:\varphi_2(\x,\Q)$, and so on.
+%	\end{proof}
+%	
+%	Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\exists\x\:\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\Q$ such that $\A\models\exists\x\:\varphi_2(\x,\Q)$, and so on.
 As we already know how to exclude that assignment in the first disjunct, we will now deal with the next disjuncts.
 Similarly to the first part of the proof, for each $i\in\{1,\ldots,k\}$ let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
-	\begin{multline*}
-	\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_{n_i},\y)\wedge \\ \forall\v_1\cdots\forall\v_{n_i}\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_{n_i},\w)\wedge\bigvee_{j=1}^{n_i}(\u_j\neq\v_j)\big)\to (\u_1,\dots,\u_{n_i})<(\v_1,\dots,\v_{n_i})\Big)\Big],
-	\end{multline*}
-	where $\u_j$ and $\v_j,$ have $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, and $\w$ has $p_i$ variables. Also, let
-	\begin{multline*}
-	\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge\limits_{\substack{\ell\in[1,n_i]:\\ \lambda_i(\ell) = j}} \v \neq \u_\ell\Big) \bigg) \bigg].
-	\end{multline*}
-	Note that $\psi_i(\X)$ excludes only the lexicographically smallest tuple of predicates $\P$ such that $\A\models\exists\x\:\varphi_i(\P,\x),$ if there is at least one.
+%	\begin{multline*}
+%	\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_{n_i},\y)\wedge \\ \forall\v_1\cdots\forall\v_{n_i}\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_{n_i},\w)\wedge\bigvee_{j=1}^{n_i}(\u_j\neq\v_j)\big)\to (\u_1,\dots,\u_{n_i})<(\v_1,\dots,\v_{n_i})\Big)\Big],
+%	\end{multline*}
+%	where $\u_j$ and $\v_j,$ have $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, and $\w$ has $p_i$ variables. Also, let
+%	\begin{multline*}
+%	\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge\limits_{\substack{\ell\in[1,n_i]:\\ \lambda_i(\ell) = j}} \v \neq \u_\ell\Big) \bigg) \bigg].
+%	\end{multline*}
+%	Note that $\psi_i(\X)$ excludes only the lexicographically smallest tuple of predicates $\P$ such that $\A\models\exists\x\:\varphi_i(\P,\x),$ if there is at least one.
 In other words, every assignment $\P^\prime \neq \P$ is such that $\A\models\psi_i(\P^\prime).$ Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
-	$$
-	\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
-	$$
-	Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Assume that $\A\in\str$.
+%	$$
+%	\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
+%	$$
+%	Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Assume that $\A\in\str$.
 Suppose first that there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least one assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\d = (\d_1,\dots,\d_{n_q},\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha_q(\d)$, where $\d_i$ is the corresponding assignment to $\x_i$, for every $i\in\{1,\dots,n_q\}$, and $\ep$ is an assignment for $\y$.
@@
 As we did in Lemma \ref{first} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies both (a) and (b)
-	\begin{enumerate}
-		\item[(a)] As we showed in part (a) of Lemma \ref{first}, if there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_q(\x,\R)$, then $\A\models\exists\x\:\varphi_q(\x,\P)$ for this particular $\P$.
+%	\begin{enumerate}
+%		\item[(a)] As we showed in part (a) of Lemma \ref{first}, if there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_q(\x,\R)$, then $\A\models\exists\x\:\varphi_q(\x,\P)$ for this particular $\P$.
 However, as we showed in Claim \ref{alpha}, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_q(\s)$, then there {\em is} such an assignment to $\X$.
@@
 It immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
-
-		\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
+%		
+%		\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
 (1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
 (2) We can use the proof in Lemma \ref{first} to see that $\A\not\models\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula
-		$$ \zeta_q(\X) =  (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{q-1}(\v)\vee\psi_q(\X)) $$ 
-		appears as a conjunct in $\varphi^\prime_i(\x,\X)$.
+%		$$ \zeta_q(\X) =  (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{q-1}(\v)\vee\psi_q(\X)) $$ 
+%		appears as a conjunct in $\varphi^\prime_i(\x,\X)$.
 However, by the choice of $q$, there is no $i\in\{1,\ldots,q-1\}$ such that $\A\models\exists\v\:\varphi_i(\v,\P)$, and also $\A\not\models\psi_q(\P)$.
@@
 And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
-
-		\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
+%		
+%		\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
 As we deduce from the part (c) of Lemma \ref{first}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
@@
 But as we mentioned, also $\A\models\:\psi_q(\P^\prime),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which leads to a contradiction.
-	\end{enumerate}
-
-	With this, we conclude that for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0$, we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-
-	Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$.
+%	\end{enumerate}
+%	
+%	With this, we conclude that for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0$, we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
+%	
+%	Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$.
 Let $\P$ be an arbitrary assignment to $\X$.
@@
 And so, for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Hence, from the results in the previous paragraph, if $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}$, we have that $f^\prime = f\dotminus 1\in\E{1}$
-
-	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
-	\begin{eqnarray*}
-		f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
-	\end{eqnarray*}
-	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$.
+%	
+%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
+%	\begin{eqnarray*}
+%		f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
+%	\end{eqnarray*}
+%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$.
 In order to prove that $f\dotminus 1\in\E{1}$, we define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2, where
-	\begin{eqnarray*}
-		\varphi(\x,\X,\z) \equiv \varphi_1(\x,\X,\z) \vee \cdots \vee \varphi_k(\x,\X,\z),
-	\end{eqnarray*}
-	and
-	\begin{eqnarray*}
-		\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\w_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
-	\end{eqnarray*}
-	where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\ell} + c_{\ell} = a_{\ell}$ for $\ell\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, and $\w_i$ has $q_i$ variables.
+%	\begin{eqnarray*}
+%		\varphi(\x,\X,\z) \equiv \varphi_1(\x,\X,\z) \vee \cdots \vee \varphi_k(\x,\X,\z),
+%	\end{eqnarray*}
+%	and
+%	\begin{eqnarray*}
+%		\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\w_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
+%	\end{eqnarray*}
+%	where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\ell} + c_{\ell} = a_{\ell}$ for $\ell\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, and $\w_i$ has $q_i$ variables.
 Furthermore, we have that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\w_i) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
 The formulas $\varphi^{-}_i(\X,\y_i,\w_i)$, $\beta_i(\x,\z)$, are also defined analogously.
-
-	In this case, we mix both the strategies in cases 1 and 2.
+%	
+%	In this case, we mix both the strategies in cases 1 and 2.
 That is, we are going to {\em isolate} the lexicographically smallest tuple of predicates that satisfies the first satisfiable disjunct, and then {\em exclude} the lexicographically smallest tuple that satisfies the isolated disjunct.
-
-	Let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and
-	\begin{multline*}
-	\alpha^i_{\min}(\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}) = \exists\y\exists\w\Big[ \alpha_i(\x_1,\dots,\x_{n_i},\y,\z_1,\dots,\z_{n_i},\w)\wedge \\ \forall\u_1\cdots\forall\u_{n_i}\forall\s\forall\v_1\cdots\forall\v_{n_i}\forall\t\Big(\big(\alpha_i(\u_1,\dots,\u_{n_i},\s,\v_1,\dots,\v_{n_i},\t)\wedge\bigvee_{j=1}^{n_i}(\x_j\neq\u_j \vee \z_j\neq\v_j)\big)\to \\ (\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i})<(\u_1,\dots,\u_{n_i},\v_1,\dots,\v_{n_i})\Big)\Big],
-	\end{multline*}
-	and let
-	\begin{multline*}
-	\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\w(\exists\u\:\varphi(\u,\X,\w) \wedge \w<\z).
-	\end{multline*}
-	Let $\P$ and $\e$ be assignments to $\X$ and $\z$.
+%	
+%	Let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and
+%	\begin{multline*}
+%	\alpha^i_{\min}(\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}) = \exists\y\exists\w\Big[ \alpha_i(\x_1,\dots,\x_{n_i},\y,\z_1,\dots,\z_{n_i},\w)\wedge \\ \forall\u_1\cdots\forall\u_{n_i}\forall\s\forall\v_1\cdots\forall\v_{n_i}\forall\t\Big(\big(\alpha_i(\u_1,\dots,\u_{n_i},\s,\v_1,\dots,\v_{n_i},\t)\wedge\bigvee_{j=1}^{n_i}(\x_j\neq\u_j \vee \z_j\neq\v_j)\big)\to \\ (\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i})<(\u_1,\dots,\u_{n_i},\v_1,\dots,\v_{n_i})\Big)\Big],
+%	\end{multline*}
+%	and let
+%	\begin{multline*}
+%	\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\w(\exists\u\:\varphi(\u,\X,\w) \wedge \w<\z).
+%	\end{multline*}
+%	Let $\P$ and $\e$ be assignments to $\X$ and $\z$.
 Note that $\psi_i(\X,\z)$ excludes $\P$ and $\e$ only if $\P$ is the {\em lexicographically smallest} tuple of predicates (Same as case 2) such that $\A\models\exists\x\:\varphi_i(\x,\P,\d)$, for some assignment $\d$ to $\z$, and $\e$ is the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_i(\x,\P,\e).$ We define $\varphi_i^\prime(\x,\X,\z)$ in the same way as case 2:
-	\begin{multline*}
-	\varphi_i^\prime(\x,\X,\z) = \varphi_i(\x,\X,\z) \wedge \psi_1(\X,\z) \wedge (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\psi_2(\X,\z)) \wedge \cdots \wedge \\ (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\cdots\vee\exists\u\exists\v\:\alpha_{i-1}(\u,\v)\vee\psi_i(\X,\z)).
-	\end{multline*}
-	Finally, let $\varphi^\prime(\x,\X,\z) = \bigvee_{i = 1}^k \varphi_i^\prime(\x,\X,\z)$.
-
-	Let $q$ be the least $i\in\{1,\ldots,k\}$ such that $\A\models\exists\x\:\varphi_i(\x,\R,\d)$ for some assignment $R$ to $\X$ and some assignment $\d$ to $\z$.
+%	\begin{multline*}
+%	\varphi_i^\prime(\x,\X,\z) = \varphi_i(\x,\X,\z) \wedge \psi_1(\X,\z) \wedge (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\psi_2(\X,\z)) \wedge \cdots \wedge \\ (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\cdots\vee\exists\u\exists\v\:\alpha_{i-1}(\u,\v)\vee\psi_i(\X,\z)).
+%	\end{multline*}
+%	Finally, let $\varphi^\prime(\x,\X,\z) = \bigvee_{i = 1}^k \varphi_i^\prime(\x,\X,\z)$.
+%	
+%	Let $q$ be the least $i\in\{1,\ldots,k\}$ such that $\A\models\exists\x\:\varphi_i(\x,\R,\d)$ for some assignment $R$ to $\X$ and some assignment $\d$ to $\z$.
 Let $\P$ be the lexicographically smallest tuple of predicates such that $\A\models\exists\x\:\varphi_q(\x,\P,\d^\prime)$ for some assignment $\d^\prime$ to $\z$.
@@
 Therefore, we conclude that $f_{\exists\x\,\varphi^\prime(\x,\X,\z)} = f_{\exists\x\,\varphi(\x,\X,\z)}\dotminus 1.$
-\end{enumerate}
+%\end{enumerate}
 

### 2016-12-27 LICS/saluja.tex
Cambio pequeño.
Nueva versión de Saluja.

@@
 
-Inspired by the connection between $\shp$ and $\#\fo$, the authors in~\cite{DBLP:journals/jcss/SalujaST95} studied subclasses of  $\#fo$ syntactical restricting the alternation quantifiers in boolean formulas defining the functions.
+Inspired by the connection between $\shp$ and $\sfo$, the authors in~\cite{DBLP:journals/jcss/SalujaST95} studied subclasses of  $\sfo$ syntactically restricting the alternation quantifiers in boolean formulas defining what we call the $\sfo$-hierarchy.
 Specifically, they define the classes $\E{i}$ and $\U{i}$ where functions are defined by $\fo$-formulas in $\loge{i}$ and $\logu{i}$, respectively, for every $i \geq 0$. 
@@
 Formally, we define the classes $\eqso(\loge{i})$ and $\eqso(\logu{i})$ where the boolean logic is restricted to $\loge{i}$ and $\logu{i}$, respectively.
-In the next result, we show that the prenex normal form hierarchy of $\eqso$ also defined a finite hierarchy similar than in~\cite{DBLP:journals/jcss/SalujaST95}.
+We denote each class by $\QE{i}$ and $\QU{i}$ for short.
+
+In the next result, we show that $\eqso$ also defined a finite hierarchy similar than in~\cite{DBLP:journals/jcss/SalujaST95} that we called the $\eqso$-hierarchy.
 \begin{proposition}
-\begin{multline*}
-\eqso(\loge{0}) \; \subsetneq \; \eqso(\loge{1}) \; \subsetneq \; \eqso(\logu{1}) \; \subsetneq \; \\ \eqso(\loge{2}) \; \subsetneq \; \eqso(\logu{2}) \; = \; \eqso(\fo)
-\end{multline*}
+\[
+\QE{0} \; \subsetneq \; \QE{1} \; \subsetneq \; \QU{1} \; \subsetneq \; \QE{2} \; \subsetneq \; \QU{2} \; = \; \eqso{\fo} 
+\]
 \end{proposition}
-A natural question until this point is what is the connection between $\E{i}$ and $\U{i}$ and their corresponding classes $\eqso(\loge{i})$ and $\eqso(\logu{i})$.
-Clearly, we have that $\E{i} \subseteq \eqso(\loge{i})$ and $\E{i} \subseteq \eqso(\logu{i})$.
-The following result shows that this inclusion is strict only for the first levels of the hierarchy.
+A natural question until this point is what is the connection between $\E{i}$ and $\U{i}$ and their corresponding classes $\QE{i}$ and $\QU{i}$.
+Clearly, we have that $\E{i} \subseteq \QE{i}$ and $\U{i} \subseteq \QU{i}$.
+Indeed, each formula $\varphi(\bar{X}, \bar{x})$ in $\E{i}$ or $\U{i}$ is equivalent to $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ which is in $\QE{i}$ or $\QU{i}$, respectively.
+Then it left to know whether the containment between the sharp and quantitative classes is strict or not. 
+For this, we study first whether each formula $\alpha$ in $\eqso(\LL)$ can be transformed in a \emph{prefix normal form} where all sums quantifiers are restricted to the beginning of the formula. 
+Formally, we say that a formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\Sigma$-prenex normal form} (or just prenex normal form) if $\alpha$ is of the form:
+\[
+\alpha := \sa{\bar{X}} \sa{\bar{x}} \varphi(\bar{X}, \bar{x})
+\]
+where $\bar{X}$ and $\bar{x}$ is a sequence of $0$ or more second and first order variables, respectively.
+In particular, note that a boolean formula is always in $\Sigma$-prenex normal form.
+Then we say that a formula $\alpha$ is in \emph{sum normal form} if $\alpha$ is of the form $c + \Sigma_{i=1}^n \alpha_i$ where $c$ is a non-negative constant and each $\alpha_i$ is in $\Sigma$-prenex normal form.
+Next we show that each formula in $\eqso(\LL)$ can be converted in \emph{sum normal form} but not always in prenex normal form.
+\begin{theorem}
+	Every formula in $\eqso(\LL)$ can be rewritten in \emph{sum normal form}.
+Furthermore, there exists a formula $\alpha \in \loge{1}$ such that there does not exist a formula in $\Sigma$-prenex normal form in $\QE{1}$ equivalent to $\alpha$.
+\end{theorem}
+
+By the previous result, we know that there exists logics $\LL$ where no $\Sigma$-prenex normal form exists. 
+Then an interesting question here is when can we convert a formula $\alpha$ in $\Sigma$-prenex normal form. 
+Of course, this would depend on the expressibility of $\LL$.
+Interestingly, if $\LL$ contains $\logu{1}$, then we can always convert any formula in $\Sigma$-prenex normal form. 
+\begin{theorem}
+	If $\logu{1} \subseteq \LL$, then for every formula $\alpha \in \eqso(\LL)$ there exists a formula $\beta \in \eqso(\LL)$ equivalent to $\alpha$ in $\Sigma$-prenex normal form.
+\end{theorem}
+
+The previous results gives the connection the connection between the hierarchy in \cite{DBLP:journals/jcss/SalujaST95} and the $\eqso$-hierarchy. 
+Indeed, if $\LL$ contains $\logu{1}$ we will have that $\sh{\LL}$ is equivalent to $\eqso(\LL)$ since each formula in $\eqso(\LL)$ can be converted in $\Sigma$-prefix normal form and, therefore, it is equivalent to a formula in $\sh{\LL}$.
+Unfortunately, this is not the case for $\loge{0}$ and $\loge{1}$ as the following result shows.
 \begin{proposition}
-	For $i \leq 1$, it holds that $\E{i} \subsetneq \eqso(\loge{i})$ and $\E{i} \subsetneq \eqso(\logu{i})$. Moreover, for $i \geq 2$ we have that $\E{i} = \eqso(\loge{i})$ and $\E{i} = \eqso(\logu{i})$.
+	The classes $\E{0}$ and $\E{1}$ are strictly contained in $\QE{0}$ and $\QE{1}$, respectively.
+Moreover, the classes $\U{1}$, $\E{2}$, and $\U{2}$ are equivalent with $\QU{1}$, $\QE{2}$, and $\QU{2}$, respectively.
 \end{proposition}
+The previous result shows that the classes $\QE{i}$ and $\QU{i}$ are more robust than the classes $\E{i}$ and $\U{i}$: they are closed under binary and sum quantifiers but the other not necessarily.
 
@@
 An advantage of the $\eqso$-hierarchy is that, by its language syntax, all the classes are closed under addition and first and second order sum.
-Furthermore, by Theorem~\ref{theo:binary-prod} one can easily see that all the hierarchy is also closed under binary product. 
-Then a natural question is whether the hierarchy is closed under subtraction.
+So, the first question is whether the multiplicative operators in $\qso$ can be defined in $\eqso(\LL)$.
+As the following result shows, if $\LL$ is closed under conjunction, then the binary product can be defined in  $\eqso(\LL)$.
+\begin{theorem}\label{theo:binary-prod}
+	If $\LL$ is closed under conjunction, then binary product can be defined in $\eqso(\LL)$.
+\end{theorem}
+The next question is whether the hierarchy is closed under subtraction.
 Formally, for any pair of functions $f,g$, we define $f - g$ as the function such that $(f - g)(\A) = f(\A)-g(\A)$ whenever $f(\A)>g(\A)$ and $0$ otherwise.
@@
 Then, a natural restriction to this question is to ask whether these classes are closed under subtraction by one, namely, if $\CC$ is a class of functions and $f \in \CC$, is $f-1 \in \CC$ where $1$ is the constant function that outputs $1$ for every structure.
-We do not know $\eqso(\logex{1})$ is closed under subtraction by one. However, if we extend $\logex{1}$ with $\fo$ predicates.
+We do not know $\E{1}$ is closed under subtraction by one. However, if we extend $\logex{1}$ with $\fo$ predicates we can show that this new fragment is closed under subtraction by one.
 \begin{theorem} \label{sigmafo-minusone}

### 2016-12-27 LICS/counting.tex
Cambios en las citas.

@@
 %!TEX root = main.tex
-
-\marcelo{Ahora solo consideramos la idea de capturar sobre estructuras ordenadas}
 
@@
 
-\marcelo{Hacer una pasada a la demostracion del siguiente teorema considerando que solo usamos estructuras ordenadas, y teniendo en cuenta que las funciones estan definidas sobre relational signatures.}
-\martin{done}
 \begin{theorem} \label{captfp}
@@
 
-Saluja et.
-al \cite{DBLP:journals/jcss/SalujaST95} define a family of counting classes $\#\LL$ for each fragment $\LL$ of $\so$.
-For a formula $\varphi(x,X)$ over $\R$, the function $f_{\varphi(x,X)}$ is defined as
+Saluja et. al \cite{SalujaST95} define a family of counting classes $\#\LL$ for each fragment $\LL$ of $\so$. For a formula $\varphi(x,X)$ over $\R$, the function $f_{\varphi(x,X)}$ is defined as
 \[

### 2016-12-27 LICS/saluja.tex

@@
 
-Inspired by the connection between $\shp$ and $\sfo$, the authors in~\cite{DBLP:journals/jcss/SalujaST95} studied subclasses of  $\sfo$ syntactically restricting the alternation quantifiers in boolean formulas defining what we call the $\sfo$-hierarchy.
+Inspired by the connection between $\shp$ and $\sfo$, the authors in~\cite{SalujaST95} studied subclasses of  $\sfo$ syntactically restricting the alternation quantifiers in boolean formulas defining what we call the $\sfo$-hierarchy.
 Specifically, they define the classes $\E{i}$ and $\U{i}$ where functions are defined by $\fo$-formulas in $\loge{i}$ and $\logu{i}$, respectively, for every $i \geq 0$. 
-Interestingly, in~\cite{DBLP:journals/jcss/SalujaST95} they show that this function classes defined a finite hierarchy of the form:
+Interestingly, in~\cite{SalujaST95} they show that this function classes defined a finite hierarchy of the form:
 \[
@@
 
-In the next result, we show that $\eqso$ also defined a finite hierarchy similar than in~\cite{DBLP:journals/jcss/SalujaST95} that we called the $\eqso$-hierarchy.
+In the next result, we show that $\eqso$ also defined a finite hierarchy similar than in~\cite{SalujaST95} that we called the $\eqso$-hierarchy.
 \begin{proposition}
@@
 
-The previous results gives the connection the connection between the hierarchy in \cite{DBLP:journals/jcss/SalujaST95} and the $\eqso$-hierarchy. 
+The previous results gives the connection the connection between the hierarchy in \cite{SalujaST95} and the $\eqso$-hierarchy. 
 Indeed, if $\LL$ contains $\logu{1}$ we will have that $\sh{\LL}$ is equivalent to $\eqso(\LL)$ since each formula in $\eqso(\LL)$ can be converted in $\Sigma$-prefix normal form and, therefore, it is equivalent to a formula in $\sh{\LL}$.
@@
 
-%We see that many of the results in Saluja et. al. \cite{DBLP:journals/jcss/SalujaST95} for $\#\LL$ still apply in $\eqso(\LL)$ for a given fragment $\LL$:
+%We see that many of the results in Saluja et. al. \cite{SalujaST95} for $\#\LL$ still apply in $\eqso(\LL)$ for a given fragment $\LL$:
 %

### 2016-12-27 LICS/framework.tex
Avance en sección 3.

@@
 
-Given a relational signature $\R$, the set of quantitative second-order logic formulas (or just $\qso$-formulas) over $\R$ is given by the following grammar:
-\begin{multline}
-\label{eq-def-qso}
-\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \\ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha 
-\end{multline}
+We introduce here the logic framework that we use for studying counting complexity classes. 
+This logic framework is based on the framework of Weighted Logics~\cite{DrosteG07} that has been used in the context of weighted automata for studying functions from words (or trees) to semirings. 
+We propose here to extend the framework of weighted logics to any relational structure and to restrict the semiring to natural numbers. 
+The extension to any relational structure will allow us to study general counting complexity classes and the restriction to the natural numbers will simplify the notation in this context (see Section~\ref{subsec:related} below for a more detailed discussion).
+
+Given a relational signature $\R$, the set of Quantitative Second-Order logic formulas (or just $\qso$-formulas) over $\R$ is given by the following grammar:
+%\[
+%\begin{array}{rcl}
+%\alpha & := & \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \ \\
+%& &  \sa{x} \alpha \ \mid \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha 
+%\end{array}
+%\]
+\begin{multline*}
+\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \\ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha 
+\end{multline*}
 where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
-If $\R$ is not mentioned, $\qso$ refers to the union of the sets of $\qso$ formulas over $\R$, for every relational signature $\R$.
+If $\R$ is not mentioned, $\qso$ refers to the union of the sets of $\qso$ formulas over $\R$, for every relational signature~$\R$.
+
+Note that the syntax of QSO formulas is divided in two levels. 
+The first level is composed by $\so$-formulas over $\R$ (called boolean formulas) and the second level is made by counting operators of addition and multiplication. 
+For this reason, the quantifiers in $\so$ (e.g.
+$\exists x$ or $\exists X$) are called boolean quantifiers and the quantifiers that make use of addition and multiplication (e.g.
+$\Sigma x$ or $\Pi X$) are called {\em quantitative quantifiers}.
+Furthermore, $\Sigma x$ and $\Sigma X$ are called the first- and second-order sum, and $\Pi x$ and $\Pi X$ the first- and second-order product.
+This division between boolean and quantitative levels is essential for understanding the difference between the logic and the counting part. 
+Furthermore, this will allow us later to parametrize both levels of the logic in order to capture different counting complexity classes.
 
@@
 Then the \emph{evaluation} of a $\qso$-formula $\alpha$ over $(\A, v, V)$ is defined as a function $\sem{\alpha}$ that on input $(\A, v, V)$ returns a number in $\bbN$.
-Formally, the function $\sem{\alpha}$ is recursively defined as follows:
-$$
-\renewcommand{\arraystretch}{1.7}
-\begin{array}{rcl} 
-\sem{\varphi}(\A, v, V) & = & 
+Formally, the function $\sem{\alpha}$ is recursively defined in Table~\ref{tab-semantics}.
+\begin{table}
+	\addtolength{\jot}{0.5em}
+	\begin{align*}
+	\sem{\varphi}(\A, v, V) & = 
 	\begin{cases}
@@
 	\end{cases}\\
-\sem{s}(\A, v, V) & = & s \\
-\sem{\alpha_1 \add \alpha_2}(\A, v, V) & = & \sem{\alpha_1}(\A, v, V) + \sem{\alpha_2}(\A, v, V)\\
-\sem{\alpha_1 \mult \alpha_2}(\A, v, V) & = & \sem{\alpha_1}(\A, v, V) \cdot \sem{\alpha_2}(\A, v, V)\\ 
-\sem{\sa{x} \alpha}(\A, v, V) & = & \displaystyle \sum_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
-\sem{\pa{x} \alpha}(\A, v, V) & = & \displaystyle \prod_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
-\sem{\sa{X} \alpha}(\A, v, V) & = & \displaystyle \sum_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])\\
-\sem{\pa{X} \alpha}(\A, v, V) & = & \displaystyle \prod_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])
-\end{array}
-$$
-A formula in $\qso$ can mention the usual quantifiers in $\so$ (that is, $\exists x$ and $\exists X$) and the quantifiers that make use of addition and multiplication (that is, $\Sigma x$, $\Pi x$, $\Sigma X$ and $\Pi X$), which are called {\em quantitative quantifiers} .
+	\sem{s}(\A, v, V) & = s \\
+	\sem{\alpha_1 \add \alpha_2}(\A, v, V) & = \sem{\alpha_1}(\A, v, V) + \sem{\alpha_2}(\A, v, V)\\
+	\sem{\alpha_1 \mult \alpha_2}(\A, v, V) & = \sem{\alpha_1}(\A, v, V) \cdot \sem{\alpha_2}(\A, v, V)\\ 
+	\sem{\sa{x} \alpha}(\A, v, V) & = \displaystyle \sum_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
+	\sem{\pa{x} \alpha}(\A, v, V) & = \displaystyle \prod_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
+	\sem{\sa{X} \alpha}(\A, v, V) & = \displaystyle \sum_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])\\
+	\sem{\pa{X} \alpha}(\A, v, V) & = \displaystyle \prod_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])
+	\end{align*}
+	\caption{The semantics of QSO formulas.}
+	\label{tab-semantics}
+	\vspace*{-20pt}
+\end{table}
+%$$
+%\renewcommand{\arraystretch}{1.7}
+%\begin{array}{rcl} 
+%\sem{\varphi}(\A, v, V) & = & 
+%\begin{cases}
+%1 & \mbox{if } (\A, v, V) \models \varphi \\
+%0 & \mbox{otherwise}
+%\end{cases}\\
+%\sem{s}(\A, v, V) & = & s \\
+%\sem{\alpha_1 \add \alpha_2}(\A, v, V) & = & \sem{\alpha_1}(\A, v, V) + \sem{\alpha_2}(\A, v, V)\\
+%\sem{\alpha_1 \mult \alpha_2}(\A, v, V) & = & \sem{\alpha_1}(\A, v, V) \cdot \sem{\alpha_2}(\A, v, V)\\ 
+%\sem{\sa{x} \alpha}(\A, v, V) & = & \displaystyle \sum_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
+%\sem{\pa{x} \alpha}(\A, v, V) & = & \displaystyle \prod_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
+%\sem{\sa{X} \alpha}(\A, v, V) & = & \displaystyle \sum_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])\\
+%\sem{\pa{X} \alpha}(\A, v, V) & = & \displaystyle \prod_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])
+%\end{array}
+%$$
 A $\qso$-formula $\alpha$ is said to be a \emph{sentence} if it does not have any free variable, that is, every variable in $\alpha$ is under the scope of a usual quantifier or a quantitative quantifier.
@@
 Thus, in such a case we use the term $\sem{\alpha}(\A)$ to denote $\sem{\alpha}(\A, v, V)$, for some arbitrary first-order assignment $v$ for $\A$ and some arbitrary second-order assignment $V$ for $\A$.
+\begin{example}
+Let $\bG = \{E(\cdot,\cdot)\}$ be the vocabulary for graphs and $\fG$ be an ordered finite $\bG$-structure encoding a non-directed graphs. 
+Suppose that we want to count the number of triangles in $\fG$. Then this can be defined with the following QSO-formula:
+\begin{multline*}
+\alpha_1 \ := \ \sa{x} \sa{y} \sa{z} ( E(x,y) \, \wedge \, E(y,z) \, \wedge \, E(z,x) \, \wedge \\
+x \leq y \, \wedge \, y \leq z )
+\end{multline*}
+In $\alpha_1$ we encode a triangle as an increasing sequence of nodes $\{x, y, z\}$ in order to count each triangle once. Then the boolean subformula  $E(x,y) \wedge E(y,z) \wedge E(z,x) \wedge
+x \leq y \wedge y \leq z$ is checking the triangle property by outputting $1$ if $\{x, y, z\}$ forms a triangle in $\fG$ and $0$ otherwise.
+Finally, the sum quantifiers in $\alpha_1$ aggregates all the values counting the number of triangles.
+
+Suppose now that we want to count the number of maximal cliques in $\fG$ (i.e.
+maximal set of nodes where every pair of nodes is connected).
+Given a second variable $X$ encoding a set, we can define that $X$ forms a clique in $\fG$ with the formula:
+$
+\op{clique}(X) := \fa{x} \fa{y} \left(X(x) \wedge Y(x)\right) \rightarrow E(x,y) 
+$.
+Then we can count the number of maximal cliques as follows:
+$$
+\alpha_2 := \sa{X} (\op{clique}(X) \wedge \fa{Y} (X \subset Y) \rightarrow \neg \op{clique}(Y) )
+$$ 
+Similar than for $\alpha_1$, in the boolean subformula of $\alpha_2$ we check whether $X$ is a maximal clique and with the sum quantifier we sum one for each set which is a maximal clique in $\fG$. 
+Note here that, in contrast to $\alpha_1$, in $\alpha_2$ we need second-order quantifiers both in the boolean and quantitative level.
+This is according to the inherent complexity of evaluating each formula: $\alpha_1$ defines an $\fp$ function where $\alpha_2$ defines a $\spp$-hard function~\cite{paper-that-shows-that-this-problem-is-SPANPhard}.
+\end{example}
+\begin{example}
+For a more involved example that includes multiplication, let $\bM = \{M(\cdot,\cdot)\}$ be the vocabulary for matrices where a structure $\fM$ over $\bM$ encodes a 0-1 matrix $A$ where $M(i,j)$ is true if $A_{i,j} = 1$ and $0$ otherwise.
+Suppose now that we want to compute the permanent of an $n$-by-$n$ 0-1 matrix $A$ which is defined by:
+$$
+\op{perm}(A) \; = \; \sum_{\sigma \in S_n} \prod_{i=1}^n A_{i, \sigma(i)}  
+$$
+where $S_n$ is the set of all permutations over $\{1, \ldots, n\}$.
+The permanent is an interesting function over matrices that has found many applications in different areas~\cite{permanent-applications} and one of the first function that was shown to be intractable for counting~\cite{Valiant79} (i.e.
+$\shp$-complete).
+\end{example}
 

### 2016-12-27 LICS/framework.tex
Mas avances en sección 3.

@@
 \end{multline*}
-In $\alpha_1$ we encode a triangle as an increasing sequence of nodes $\{x, y, z\}$ in order to count each triangle once. Then the boolean subformula  $E(x,y) \wedge E(y,z) \wedge E(z,x) \wedge
-x \leq y \wedge y \leq z$ is checking the triangle property by outputting $1$ if $\{x, y, z\}$ forms a triangle in $\fG$ and $0$ otherwise.
-Finally, the sum quantifiers in $\alpha_1$ aggregates all the values counting the number of triangles.
+For $\alpha_1$ we encode a triangle as an increasing sequence of nodes $\{x, y, z\}$ in order to count each triangle once. Then the boolean subformula  $E(x,y) \wedge E(y,z) \wedge E(z,x) \wedge
+x \leq y \wedge y \leq z$ is checking the triangle property, by outputting $1$ if $\{x, y, z\}$ forms a triangle in $\fG$ and $0$ otherwise.
+Finally, the sum quantifiers in $\alpha_1$ aggregates all the values, counting the number of triangles.
 
-Suppose now that we want to count the number of maximal cliques in $\fG$ (i.e.
-maximal set of nodes where every pair of nodes is connected).
-Given a second variable $X$ encoding a set, we can define that $X$ forms a clique in $\fG$ with the formula:
-$
-\op{clique}(X) := \fa{x} \fa{y} \left(X(x) \wedge Y(x)\right) \rightarrow E(x,y) 
-$.
-Then we can count the number of maximal cliques as follows:
+Suppose now that we want to count the number of cliques in $\fG$ (i.e. set of nodes where every pair of nodes is connected). We can easily compute this function with the following formula:
 $$
-\alpha_2 := \sa{X} (\op{clique}(X) \wedge \fa{Y} (X \subset Y) \rightarrow \neg \op{clique}(Y) )
+\alpha_2 := \sa{X} \big( \, \fa{x} \fa{y} \left(X(x) \wedge X(y)\right) \rightarrow E(x,y) \, \big)  
 $$ 
-Similar than for $\alpha_1$, in the boolean subformula of $\alpha_2$ we check whether $X$ is a maximal clique and with the sum quantifier we sum one for each set which is a maximal clique in $\fG$. 
-Note here that, in contrast to $\alpha_1$, in $\alpha_2$ we need second-order quantifiers both in the boolean and quantitative level.
-This is according to the inherent complexity of evaluating each formula: $\alpha_1$ defines an $\fp$ function where $\alpha_2$ defines a $\spp$-hard function~\cite{paper-that-shows-that-this-problem-is-SPANPhard}.
+Similar than for $\alpha_1$, in the boolean subformula of $\alpha_2$ we check whether $X$ is a clique and with the sum quantifier we add one for each clique in $\fG$. 
+Note here that, in contrast to $\alpha_1$, in $\alpha_2$ we need a second-order quantifier in the quantitative level.
+This is according to the inherent complexity of evaluating each formula: $\alpha_1$ defines an $\fp$ function where $\alpha_2$ defines a $\shp$-hard function~\cite{paper-that-shows-that-this-problem-is-sharpP-hard}.
 \end{example}
@@
 where $S_n$ is the set of all permutations over $\{1, \ldots, n\}$.
-The permanent is an interesting function over matrices that has found many applications in different areas~\cite{permanent-applications} and one of the first function that was shown to be intractable for counting~\cite{Valiant79} (i.e.
+The permanent is a relevant function over matrices that has found many applications in different areas~\cite{permanent-applications} and it was one of the first function that was shown to be difficult for counting~\cite{Valiant79} (i.e.
 $\shp$-complete).
+For a binary relation symbol $S$, let $\op{permut}(S)$ be the $\fo$-formula that is true iff $S$ is a permutation, i.e.
+bijective function, (the definition of $\op{permut}(S)$ is straightforward and we left this exercise to the reader).
+Then a $\qso$-formula for defining the permanent of a matrix is as follows:
+\[
+\alpha_3 := \sa{S} \op{permut}(S) \cdot \pa{x} (\ex{y} S(x,y) \wedge M(x,y))
+\]
+Intuitively, the subformula $\beta(s) := \pa{x} (\ex{y} S(x,y) \wedge M(x,y))$ calculates the value $\prod_{i=1}^n A_{i, \sigma(i)}$ whenever $S$ encodes a permutation $\sigma$.
+Then the subformula $\op{permut}(S) \cdot \beta(S)$ computes $\beta(S)$ when $S$ is a permutation and $0$ otherwise (i.e. $\op{permut}(S)$ behaves like a filter). 
+Finally, the second order sum aggregates these values iterating over all binary relations and calculating the permanent of the matrix.
+We would like to end this example by highlighting the similarity of $\alpha_3$ with the permanent formula. 
+Indeed, an advantage of $\qso$-formulas is that the first- and second-order quantifiers naturally reflect the operations used to define mathematical formulas in practice.
 \end{example}
 
-In this paper, we consider several fragments of $\qso$, which are obtained by restricting the syntax of the formula $\varphi$ in \eqref{eq-def-qso} or the use of the quantitative quantifiers.
+We consider several fragments of $\qso$, which are obtained by restricting the syntax of the formula $\varphi$ or the use of the quantitative quantifiers.
 Let $\qfo$ be a fragment of $\qso$ defined by the grammar in \eqref{eq-def-qso} but only the quantitative quantifiers $\Sigma x$ and $\Pi x$ are allowed.

### 2016-12-28 LICS/framework.tex
Version 1.0 de sección 3.

@@
 We introduce here the logic framework that we use for studying counting complexity classes. 
-This logic framework is based on the framework of Weighted Logics~\cite{DrosteG07} that has been used in the context of weighted automata for studying functions from words (or trees) to semirings. 
-We propose here to extend the framework of weighted logics to any relational structure and to restrict the semiring to natural numbers. 
-The extension to any relational structure will allow us to study general counting complexity classes and the restriction to the natural numbers will simplify the notation in this context (see Section~\ref{subsec:related} below for a more detailed discussion).
+This logic framework is based on the framework of Weighted Logics~\cite{DrosteG07} (WL) that has been used in the context of weighted automata for studying functions from words (or trees) to semirings.
+We propose here to use the framework of WL over any relational structure and to restrict the semiring to natural numbers. 
+The extension to any relational structure will allow us to study general counting complexity classes and the restriction to the natural numbers will simplify the notation in this context (see Section~\ref{sec:previous} below for a more detailed discussion).
 
@@
 $\Sigma x$ or $\Pi X$) are called {\em quantitative quantifiers}.
-Furthermore, $\Sigma x$ and $\Sigma X$ are called the first- and second-order sum, and $\Pi x$ and $\Pi X$ the first- and second-order product.
-This division between boolean and quantitative levels is essential for understanding the difference between the logic and the counting part. 
+Furthermore, $\Sigma x$ and $\Sigma X$ are called first- and second-order sum, and $\Pi x$ and $\Pi X$ the first- and second-order product, respectively.
+This division between boolean and quantitative levels is essential for understanding the difference between the logic and the quantitative part. 
 Furthermore, this will allow us later to parametrize both levels of the logic in order to capture different counting complexity classes.
@@
 A $\qso$-formula $\alpha$ is said to be a \emph{sentence} if it does not have any free variable, that is, every variable in $\alpha$ is under the scope of a usual quantifier or a quantitative quantifier.
-It is important to notice that if $\alpha$ is a $\qso$-sentence over a relational signature $\R$, then for every ordered finite $\R$-structure $\A$, first-order assignments $v_1$, $v_2$ for $\A$ and second-order assignments $V_1$, $V_2$ for $\A$, it holds that:
-\begin{eqnarray*}
-	\sem{\alpha}(\A, v_1, V_1) & = & \sem{\alpha}(\A, v_2, V_2).
-\end{eqnarray*}
+It is important to notice that if $\alpha$ is a $\qso$-sentence over a relational signature $\R$, then for every ordered finite $\R$-structure $\A$, first-order assignments $v_1$, $v_2$ for $\A$ and second-order assignments $V_1$, $V_2$ for $\A$, it holds that $\sem{\alpha}(\A, v_1, V_1) = \sem{\alpha}(\A, v_2, V_2)$.
 Thus, in such a case we use the term $\sem{\alpha}(\A)$ to denote $\sem{\alpha}(\A, v, V)$, for some arbitrary first-order assignment $v$ for $\A$ and some arbitrary second-order assignment $V$ for $\A$.
@@
 x \leq y \wedge y \leq z$ is checking the triangle property, by outputting $1$ if $\{x, y, z\}$ forms a triangle in $\fG$ and $0$ otherwise.
-Finally, the sum quantifiers in $\alpha_1$ aggregates all the values, counting the number of triangles.
+Finally, the sum quantifiers in $\alpha_1$ aggregates all the values, counting the number of triangles in $\fG$.
 
-Suppose now that we want to count the number of cliques in $\fG$ (i.e. set of nodes where every pair of nodes is connected). We can easily compute this function with the following formula:
+Suppose now that we want to count the number of cliques in~$\fG$ (i.e. set of nodes where every pair of nodes is connected). We can easily define this function with the following formula:
 $$
@@
 where $S_n$ is the set of all permutations over $\{1, \ldots, n\}$.
-The permanent is a relevant function over matrices that has found many applications in different areas~\cite{permanent-applications} and it was one of the first function that was shown to be difficult for counting~\cite{Valiant79} (i.e.
+The permanent is a relevant matrix function that has found many applications in different areas~\cite{permanent-applications} and it was one of the first function that was shown to be difficult for counting~\cite{Valiant79} (i.e.
 $\shp$-complete).
-For a binary relation symbol $S$, let $\op{permut}(S)$ be the $\fo$-formula that is true iff $S$ is a permutation, i.e.
-bijective function, (the definition of $\op{permut}(S)$ is straightforward and we left this exercise to the reader).
+For a binary relation symbol $S$, let $\op{permut}(S)$ be the $\fo$-formula that is true iff $S$ is a permutation, i.e. bijective function (the definition of $\op{permut}(S)$ is straightforward).
 Then a $\qso$-formula for defining the permanent of a matrix is as follows:
@@
 
-We consider several fragments of $\qso$, which are obtained by restricting the syntax of the formula $\varphi$ or the use of the quantitative quantifiers.
-Let $\qfo$ be a fragment of $\qso$ defined by the grammar in \eqref{eq-def-qso} but only the quantitative quantifiers $\Sigma x$ and $\Pi x$ are allowed.
-Let $\eqfo$ be the fragment of $\qso$ defined by the following grammar:
-$$
-\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid \ \sa{x} \alpha.
-$$ Let $\eqso$ be a fragment of $\qso$ defined as $\eqfo$ but also allowing the quantitative quantifier $\Sigma X$.
-Moreover, assuming that $\LL$ is a fragment of $\so$, let $\qso(\LL)$ be the fragment of $\qso$ obtained by restricting $\varphi$ in \eqref{eq-def-qso} to be a formula in $\LL$, and in general, if $\FF$ is a fragment of $\qso$, $\FF(\LL)$ be the fragment of $\FF$ obtained by imposing the same restriction.
-In particular, in this paper we consider the following fragments $\LL$: $\fo$, which is obtained by disallowing the use of the second-order quantifier ($\exists X$ or $\forall X$) in the formula $\varphi$ in \eqref{eq-def-qso}, and $\eso$, which is obtained by allowing $\varphi$ in \eqref{eq-def-qso} to be a formula of the form $\exists X_1 \cdots \exists X_k \, \psi$, where $\psi$ is an $\fo$-formula.
+We consider several fragments of $\qso$, which are obtained by restricting the syntax of the boolean formulas or the use of the quantitative quantifiers.
+In this direction, we denote by $\qfo$ the fragment of $\qso$ where second-order sum and product are not allowed. 
+For example, $\alpha_1$ is in $\qfo$ and $\alpha_2$ is not.
+Another interesting fragment are the formulas of $\qso$ where only sum operators and quantifiers are allowed. 
+Formally, we denote by $\eqso$ the fragment of $\qso$ where first- and second-order products (i.e. $\pa{x}$ and $\pa{X}$) are not allowed.
+For example, $\alpha_1$ and $\alpha_2$ are formulas of $\eqso$ but $\alpha_3$ is not. 
+We also consider fragments of $\qso$ by further restricting the boolean core of the logic.
+Let $\LL$ be a fragment of $\so$.
+We define the quantitative logic $\qso(\LL)$ to be the fragment of $\qso$ obtained by restricting $\varphi$ to be a formula in $\LL$. 
+In general, if $\FF$ is a fragment of $\qso$ we define $\FF(\LL)$ to be the fragment of $\FF$ obtained by restricting boolean formulas to $\LL$.
+For example, we can define the fragments $\qfo(\fo)$ or $\eqso(\fo)$ which are subfragments of $\qfo$ and $\eqso$ where $\varphi$ is restricted to $\fo$ formulas.
 
-\begin{theorem} \label{no-mult}
-	Let $\alpha$ be a $\qso$-formula in some fragment of $\qso$.
-There is a $\qso$ formula $\alpha'$ in the same fragment such that $\alpha'$ does not have any subformula of the form $(\beta_1 \cdot \beta_2)$ and $\sem{\alpha} = \sem{\alpha'}$.
-\end{theorem}
+In the next section, we variate $\FF$ and $\LL$ to capture several counting complexity classes.
+Before this, we devote the next subsection to show the connection of $\qso$ with previous frameworks for defining functions over relational structures.
+
+\subsection{Previous frameworks for quantitative functions} \label{sec:previous}
+
+In this subsection we discuss previous frameworks proposed in the literature and how they differ to our approach.
+We start by discussing the connection between $\qso$ and weighted logics (WL)~\cite{DrosteG07}. 
+As it was previously discussed, $\qso$ is a fragment of WL.
+The real difference is that we restrict the semiring used in WL to natural numbers in order to study counting complexity classes.
+Another difference of WL with our approach is that, as far as we know, this is the first paper to study weighted logics over general signature properly to do descriptive complexity of counting classes.
+Previous works on WL usually restrict the signature of the logic to strings, trees, and other specific structures (see \cite{droste2009handbook} for more examples), and did not study the logic over general structures.
+Furthermore, in this paper we propose further extensions for $\qso$ (see Section~\ref{sec:beyond}) which start to differ from previous approaches in WL.
+
+Another approach that resembles $\qso$ are counting logics~\cite{L04,counting-paper-here}. 
+Logics with counting operators usually extend $\fo$ with quantifiers that counts how many ways can be satisfied a formula over a second sort (i.e. natural numbers). 
+In contrast to our approach, counting operators are usually used for checking boolean properties over structures and not for outputting values (i.e. they do not define a function).
+Although there are some papers that uses this operator for counting~\cite{XXX}, they have not used the logic for capturing counting complexity classes.
+
+Finally, the work on~\cite{SalujaST95} and \cite{Gradel91} is of particular interest for our work. 
+In~\cite{SalujaST95}, it was proposed to define functions over structures by using open variables on SO-formulas and count how many extensions of the structure satisfy the formula (i.e.
+by instantiating the open variables).
+Formally, Saluja et.
+al \cite{SalujaST95} define a family of counting classes $\#\LL$ for each fragment $\LL$ of $\so$.
+For a formula $\varphi(\bar{x},\bar{X})$ over $\R$, the function $f_{\varphi(x,X)}$ is defined as
+$
+f_{\varphi}(\A) = \vert \{\langle \bar{a},\bar{A} \rangle\mid \A\models\varphi(\bar{a},\bar{A})\}\vert.
+$
+for each $\A\in\ostr[\R]$. Then a function $g:\ostr[\R]\to\nat$ is in $\#\LL$ if there exists an formula $\varphi(\bar{x},\bar{X})$ in $\LL$ such that $f = f_{\varphi}$. 
+In~\cite{SalujaST95}, it was shown several results for capturing counting complexity classes which are relevant for our work.
+We discuss and use these results in Sections~\ref{sec:complexity} and~\ref{sec:syntactic}.
+Note that for any formula $\varphi(\bar{x},\bar{X})$ it holds that $f_\varphi \equiv \sem{\sa{\bar{X}} \sa{\bar{x}} \varphi(\bar{x},\bar{X})}$, that is, the approach in \cite{SalujaST95} can be seen as a syntactical restriction of $\qso$.
+Thus, the advantage of our approach relies on the flexibility of defining  counting functions by alternating sum with product operators or, moreover, by introducing new quantitative operators (like in Section~\ref{sec:beyond}).
+Furthermore, in the next section we show how to capture some interesting classes that could never be captured by following the approach in~\cite{SalujaST95}.

### 2016-12-28 LICS/framework.tex
27 dic

--- 2016-12-27 LICS/framework.tex
@@
 
-We introduce here the logic framework that we use for studying counting complexity classes. 
-This logic framework is based on the framework of Weighted Logics~\cite{DrosteG07} that has been used in the context of weighted automata for studying functions from words (or trees) to semirings. 
-We propose here to extend the framework of weighted logics to any relational structure and to restrict the semiring to natural numbers. 
-The extension to any relational structure will allow us to study general counting complexity classes and the restriction to the natural numbers will simplify the notation in this context (see Section~\ref{subsec:related} below for a more detailed discussion).
+We define here the logic framework that we will use for studying counting complexity classes. 
+We propose here to use framework of weighted logics~\cite{}
 
-Given a relational signature $\R$, the set of Quantitative Second-Order logic formulas (or just $\qso$-formulas) over $\R$ is given by the following grammar:
-%\[
-%\begin{array}{rcl}
-%\alpha & := & \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \ \\
-%& &  \sa{x} \alpha \ \mid \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha 
-%\end{array}
-%\]
-\begin{multline*}
-\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \\ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha 
-\end{multline*}
+Given a relational signature $\R$, the set of quantitative second-order logic formulas (or just $\qso$-formulas) over $\R$ is given by the following grammar:
+\begin{multline}
+\label{eq-def-qso}
+\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \\ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha 
+\end{multline}
 where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
-If $\R$ is not mentioned, $\qso$ refers to the union of the sets of $\qso$ formulas over $\R$, for every relational signature~$\R$.
-
-Note that the syntax of QSO formulas is divided in two levels. 
-The first level is composed by $\so$-formulas over $\R$ (called boolean formulas) and the second level is made by counting operators of addition and multiplication. 
-For this reason, the quantifiers in $\so$ (e.g.
-$\exists x$ or $\exists X$) are called boolean quantifiers and the quantifiers that make use of addition and multiplication (e.g.
-$\Sigma x$ or $\Pi X$) are called {\em quantitative quantifiers}.
-Furthermore, $\Sigma x$ and $\Sigma X$ are called the first- and second-order sum, and $\Pi x$ and $\Pi X$ the first- and second-order product.
-This division between boolean and quantitative levels is essential for understanding the difference between the logic and the counting part. 
-Furthermore, this will allow us later to parametrize both levels of the logic in order to capture different counting complexity classes.
+If $\R$ is not mentioned, $\qso$ refers to the union of the sets of $\qso$ formulas over $\R$, for every relational signature $\R$.
 
@@
 Then the \emph{evaluation} of a $\qso$-formula $\alpha$ over $(\A, v, V)$ is defined as a function $\sem{\alpha}$ that on input $(\A, v, V)$ returns a number in $\bbN$.
-Formally, the function $\sem{\alpha}$ is recursively defined in Table~\ref{tab-semantics}.
-\begin{table}
-	\addtolength{\jot}{0.5em}
-	\begin{align*}
-	\sem{\varphi}(\A, v, V) & = 
+Formally, the function $\sem{\alpha}$ is recursively defined as follows:
+$$
+\renewcommand{\arraystretch}{1.7}
+\begin{array}{rcl} 
+\sem{\varphi}(\A, v, V) & = & 
 \begin{cases}
@@
 \end{cases}\\
-	\sem{s}(\A, v, V) & = s \\
-	\sem{\alpha_1 \add \alpha_2}(\A, v, V) & = \sem{\alpha_1}(\A, v, V) + \sem{\alpha_2}(\A, v, V)\\
-	\sem{\alpha_1 \mult \alpha_2}(\A, v, V) & = \sem{\alpha_1}(\A, v, V) \cdot \sem{\alpha_2}(\A, v, V)\\ 
-	\sem{\sa{x} \alpha}(\A, v, V) & = \displaystyle \sum_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
-	\sem{\pa{x} \alpha}(\A, v, V) & = \displaystyle \prod_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
-	\sem{\sa{X} \alpha}(\A, v, V) & = \displaystyle \sum_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])\\
-	\sem{\pa{X} \alpha}(\A, v, V) & = \displaystyle \prod_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])
-	\end{align*}
-	\caption{The semantics of QSO formulas.}
-	\label{tab-semantics}
-	\vspace*{-20pt}
-\end{table}
-%$$
-%\renewcommand{\arraystretch}{1.7}
-%\begin{array}{rcl} 
-%\sem{\varphi}(\A, v, V) & = & 
-%\begin{cases}
-%1 & \mbox{if } (\A, v, V) \models \varphi \\
-%0 & \mbox{otherwise}
-%\end{cases}\\
-%\sem{s}(\A, v, V) & = & s \\
-%\sem{\alpha_1 \add \alpha_2}(\A, v, V) & = & \sem{\alpha_1}(\A, v, V) + \sem{\alpha_2}(\A, v, V)\\
-%\sem{\alpha_1 \mult \alpha_2}(\A, v, V) & = & \sem{\alpha_1}(\A, v, V) \cdot \sem{\alpha_2}(\A, v, V)\\ 
-%\sem{\sa{x} \alpha}(\A, v, V) & = & \displaystyle \sum_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
-%\sem{\pa{x} \alpha}(\A, v, V) & = & \displaystyle \prod_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
-%\sem{\sa{X} \alpha}(\A, v, V) & = & \displaystyle \sum_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])\\
-%\sem{\pa{X} \alpha}(\A, v, V) & = & \displaystyle \prod_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])
-%\end{array}
-%$$
+\sem{s}(\A, v, V) & = & s \\
+\sem{\alpha_1 \add \alpha_2}(\A, v, V) & = & \sem{\alpha_1}(\A, v, V) + \sem{\alpha_2}(\A, v, V)\\
+\sem{\alpha_1 \mult \alpha_2}(\A, v, V) & = & \sem{\alpha_1}(\A, v, V) \cdot \sem{\alpha_2}(\A, v, V)\\ 
+\sem{\sa{x} \alpha}(\A, v, V) & = & \displaystyle \sum_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
+\sem{\pa{x} \alpha}(\A, v, V) & = & \displaystyle \prod_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
+\sem{\sa{X} \alpha}(\A, v, V) & = & \displaystyle \sum_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])\\
+\sem{\pa{X} \alpha}(\A, v, V) & = & \displaystyle \prod_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])
+\end{array}
+$$
+A formula in $\qso$ can mention the usual quantifiers in $\so$ (that is, $\exists x$ and $\exists X$) and the quantifiers that make use of addition and multiplication (that is, $\Sigma x$, $\Pi x$, $\Sigma X$ and $\Pi X$), which are called {\em quantitative quantifiers} .
 A $\qso$-formula $\alpha$ is said to be a \emph{sentence} if it does not have any free variable, that is, every variable in $\alpha$ is under the scope of a usual quantifier or a quantitative quantifier.
@@
 Thus, in such a case we use the term $\sem{\alpha}(\A)$ to denote $\sem{\alpha}(\A, v, V)$, for some arbitrary first-order assignment $v$ for $\A$ and some arbitrary second-order assignment $V$ for $\A$.
-\begin{example}
-Let $\bG = \{E(\cdot,\cdot)\}$ be the vocabulary for graphs and $\fG$ be an ordered finite $\bG$-structure encoding a non-directed graphs. 
-Suppose that we want to count the number of triangles in $\fG$. Then this can be defined with the following QSO-formula:
-\begin{multline*}
-\alpha_1 \ := \ \sa{x} \sa{y} \sa{z} ( E(x,y) \, \wedge \, E(y,z) \, \wedge \, E(z,x) \, \wedge \\
-x \leq y \, \wedge \, y \leq z )
-\end{multline*}
-In $\alpha_1$ we encode a triangle as an increasing sequence of nodes $\{x, y, z\}$ in order to count each triangle once. Then the boolean subformula  $E(x,y) \wedge E(y,z) \wedge E(z,x) \wedge
-x \leq y \wedge y \leq z$ is checking the triangle property by outputting $1$ if $\{x, y, z\}$ forms a triangle in $\fG$ and $0$ otherwise.
-Finally, the sum quantifiers in $\alpha_1$ aggregates all the values counting the number of triangles.
-
-Suppose now that we want to count the number of maximal cliques in $\fG$ (i.e.
-maximal set of nodes where every pair of nodes is connected).
-Given a second variable $X$ encoding a set, we can define that $X$ forms a clique in $\fG$ with the formula:
-$
-\op{clique}(X) := \fa{x} \fa{y} \left(X(x) \wedge Y(x)\right) \rightarrow E(x,y) 
-$.
-Then we can count the number of maximal cliques as follows:
-$$
-\alpha_2 := \sa{X} (\op{clique}(X) \wedge \fa{Y} (X \subset Y) \rightarrow \neg \op{clique}(Y) )
-$$ 
-Similar than for $\alpha_1$, in the boolean subformula of $\alpha_2$ we check whether $X$ is a maximal clique and with the sum quantifier we sum one for each set which is a maximal clique in $\fG$. 
-Note here that, in contrast to $\alpha_1$, in $\alpha_2$ we need second-order quantifiers both in the boolean and quantitative level.
-This is according to the inherent complexity of evaluating each formula: $\alpha_1$ defines an $\fp$ function where $\alpha_2$ defines a $\spp$-hard function~\cite{paper-that-shows-that-this-problem-is-SPANPhard}.
-\end{example}
-\begin{example}
-For a more involved example that includes multiplication, let $\bM = \{M(\cdot,\cdot)\}$ be the vocabulary for matrices where a structure $\fM$ over $\bM$ encodes a 0-1 matrix $A$ where $M(i,j)$ is true if $A_{i,j} = 1$ and $0$ otherwise.
-Suppose now that we want to compute the permanent of an $n$-by-$n$ 0-1 matrix $A$ which is defined by:
-$$
-\op{perm}(A) \; = \; \sum_{\sigma \in S_n} \prod_{i=1}^n A_{i, \sigma(i)}  
-$$
-where $S_n$ is the set of all permutations over $\{1, \ldots, n\}$.
-The permanent is an interesting function over matrices that has found many applications in different areas~\cite{permanent-applications} and one of the first function that was shown to be intractable for counting~\cite{Valiant79} (i.e.
-$\shp$-complete).
-\end{example}
 
@@
 
-\begin{theorem} \label{no-mult}
-	Let $\alpha$ be a $\qso$-formula in some fragment of $\qso$.
+%\begin{theorem} \label{no-mult}
+%	Let $\alpha$ be a $\qso$-formula in some fragment of $\qso$.
 There is a $\qso$ formula $\alpha'$ in the same fragment such that $\alpha'$ does not have any subformula of the form $(\beta_1 \cdot \beta_2)$ and $\sem{\alpha} = \sem{\alpha'}$.
-\end{theorem}
+%\end{theorem}

### 2016-12-28 LICS/horn-counting.tex
27 dic

@@
 \newcommand{\pNC}{\textit{NC}}
+\newcommand{\pD}{\textit{D}}
 
@@
 Given a relational signature $\R$, a clause over $\R$ is a formula of the form:
-\begin{eqnarray*}
+$$
 \forall \x \, (\varphi_1 \vee \cdots \vee \varphi_n),
-\end{eqnarray*}
-where each $\varphi_i$ ($1 \leq i \leq n$) is either a positive literal or a negative literal or an \fo-formula over $\R$.
+$$
+where each $\varphi_i$ ($1 \leq i \leq n$) is either a positive literal, a negative literal or an \fo-formula over $\R$.
 A clause is said to be Horn if it contains at most one positive literal, and a formula is said to be Horn if it is a conjunction of Horn clauses over a relational signature $\R$.
@@
 
-\begin{example}
+\begin{example} \label{ex-hornsat-esop1}
 Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot)\}$.
@@
 &\forall c \forall x \, (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{T}(x)).
+&\forall c \forall x \, (\textit{A}(c,x) \vee \neg\textit{T}(x)).
 \end{align*}

### 2016-12-28 LICS/syntactic.tex
27 dic

--- 2016-12-05 LICS/syntactic.tex
@@
 
-\subsection{Saluja Classes}
+In this section we study the fragment of $\eqso(\LL)$ when $\LL$ is a boolean logic contained in $\fo$.
+We show that by restricting $\LL$ we can find different subclasses below $\shp$ with interesting computational and closure properties.
+
+From this point on, for each fragment $\FF$ of $\qso$, we will also use $\FF$ to refer to the class of functions defined by the formulas in $\FF$.
+
+\subsection{Counting hierarchy below $\shp$}
+\input{saluja}
 
 \subsection{Horn Counting Classes}
+\input{horn-counting}

### 2016-12-28 LICS/appendix.tex
27 dic merge

@@
 
-\martin{no revisar, es un copy-paste de arriba}
-
 For the first condition, let $\alpha\in\qfo(\lfp)$ over some signature $\R$.
@@
 \subsection{Proof of Theorem \ref{qfo-pfp-cap}}
+
+\martin{no revisar, es un copy-paste de arriba}
 
@@
 For condition (2) in \ref{cap-def}, note that Saluja et al.
-\cite{DBLP:journals/jcss/SalujaST95} proved that $\shp = \#\fo$.
+\cite{SalujaST95} proved that $\shp = \#\fo$.
 We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso(\fo)$ be its corresponding formula.
@@
 
-For condition (2) in \ref{cap-def}, note that Compton and Graedel \cite{DBLP:journals/jcss/ComptonG96} proved that $\spp = \#\eso$.
+For condition (2) in \ref{cap-def}, note that Compton and Graedel \cite{ComptonG96} proved that $\spp = \#\eso$.
 We also have that $\sqso(\eso)$ captures $\#\eso$ over ordered structures so for each $f\in \spp$ let $\alpha \in \sqso(\eso)$ be its corresponding formula.
@@
 Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\pfp$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space \cite{DBLP:conf/lics/AbiteboulV89}\martin{no estoy seguro de esta cita}, and accepts if and only if it holds true.
+If $\alpha$ is a $\pfp$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space \cite{AbiteboulV89}\martin{no estoy seguro de esta cita}, and accepts if and only if it holds true.
 If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
@@
 Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$ \cite{DBLP:journals/siamcomp/Ladner89}, and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
+We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$ \cite{Ladner89}, and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
 
@@
 If $\beta_i = \sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}$, if $\beta_i = \sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = 2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, if $\beta_i = \sa{\bar{X}}\sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, all of which can be computed in polynomial time.
-Otherwise, $\beta_i$ is $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, and then $\sem{\beta_i}(\A) = f_{\varphi}(\enc(\A))$, and $f_{\varphi}\in \#\loge{0}$ which implies that $f_{\varphi}(\enc(\A))$ can be computed in polynomial time \cite{DBLP:journals/jcss/SalujaST95}.
+Otherwise, $\beta_i$ is $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, and then $\sem{\beta_i}(\A) = f_{\varphi}(\enc(\A))$, and $f_{\varphi}\in \#\loge{0}$ which implies that $f_{\varphi}(\enc(\A))$ can be computed in polynomial time \cite{SalujaST95}.
 
@@
 Similarly as in the proof by Saluja et al.
-\cite{DBLP:journals/jcss/SalujaST95}, we define $\gamma_i = \sa{\bar{x}}\sa{\bar{y}}\sa{\bar{X}}\varphi(\bar{x},\bar{y},\bar{X}) \wedge \forall\bar{z}(\varphi(\bar{x},\bar{z},\bar{X})\to(\bar{z} < \bar{y} \vee \bar{z} = \bar{y}))$ that fulfils the desired condition.
+\cite{SalujaST95}, we define $\gamma_i = \sa{\bar{x}}\sa{\bar{y}}\sa{\bar{X}}\varphi(\bar{x},\bar{y},\bar{X}) \wedge \forall\bar{z}(\varphi(\bar{x},\bar{z},\bar{X})\to(\bar{z} < \bar{y} \vee \bar{z} = \bar{y}))$ that fulfils the desired condition.
 The theorem holds true by using $\beta = \gamma_1\add\cdots\add\gamma_m$ which is in $\eqso(\logu{1})$
@@
 
-\textbf{Proof of Theorem \ref{sigma1strict}}
-
-We will show that the $\E{1}$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
+%\textbf{Proof of Theorem \ref{sigma1strict}}
+%
+%We will show that the $\E{1}$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
 By contradiction, suppose that it is.
@@
 Note that $\A$ is an induced substructure of $\B$.
-
-We have that for each function $g \in \#\Sigma_1$ and structures $\A_1,\A_2\in\str$, if $\A_1$ is an induced substructure of $\A_2$, then $g(\A_1) \leq g(\A_2)$ \cite{DBLP:journals/jcss/SalujaST95}.
+%
+%We have that for each function $g \in \#\Sigma_1$ and structures $\A_1,\A_2\in\str$, if $\A_1$ is an induced substructure of $\A_2$, then $g(\A_1) \leq g(\A_2)$ \cite{SalujaST95}.
 Therefore, $f(\B) \geq f(\A) = 1$.
@@
 $$
-\varphi_i(\bar{X},\bar{x},\bar{y}) = (\text{conjunction of $X_i$'s}) \wedge (\text{conjunction of $\neg X_i$'s})  \wedge (\text{an $\fo$ formula that does not mention any $X_i$}).
+\varphi_i(\bar{X},\bar{x},\bar{y}) = (\text{conjunction of $X_j$'s}) \wedge (\text{conjunction of $\neg X_j$'s})  \wedge (\text{an $\fo$ formula that does not mention any $X_j$}).
 $$
@@
 \begin{claim}
-	Let $\varphi_i(\bar{X},\bar{x},\bar{y}) = \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X},\bar{x},\bar{y})$ be a quantifier-free formula that respects all the mentioned assumptions.
-For a given ordered structure $\A$ with domain $A$, if $\A\models\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists some $\bar{B}$ over $\A$ and $\bar{b}\in A^{\length(\bar{x})}$ such that $\A\models\varphi_i(\bar{B},\bar{b},\bar{a})$ for some $\bar{a}\in A^{\length(\bar{y})}$.
+For a given ordered structure $\A$ such that $\A\models\exists\bar{X}\exists\bar{x}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, there is an assignment $(\bar{B},\bar{b})$ to $(\bar{X},\bar{x})$ that satisfies the following conditions (1) $\A\models\exists\bar{y}\,\varphi_i(\bar{B},\bar{b},\bar{y})$, (2) $\A\not\models\psi_i(\bar{B},\bar{b})$ and (3) this is the only assignment to $(\bar{X},\bar{x})$ that satisfies (1) and (2).
+\end{claim}
+\begin{proof}
+	proof here.
+\end{proof}
+The following claim identifies a crucial condition that is verifiable through a $\fo$. 
+\begin{claim}
+	Let $\varphi_i(\bar{X},\bar{x},\bar{y}) = \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X},\bar{x},\bar{y})$ be a $\logex{0}$ formula that satisfies all the mentioned assumptions.
+For a given ordered structure $\A$ it holds that $\A\models\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$ iff $\A\models\exists\bar{X}\exists\bar{x}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$.
 \end{claim}
@@
 $$
-And lastly $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x},\bar{y})$.
+And lastly $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x},\bar{y})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
+
 %%Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$.
@@
 %\end{enumerate}
-
-\textbf{Proof of Theorem \ref{dnf-pars}}
-
-Let $f$ be an arbitrary function in $\E{1}$ and let $f_{\textsc{\#DNF}}$ be the function that defines {\sc \#DNF}.
+%		
+%\textbf{Proof of Theorem \ref{dnf-pars}}
+%
+%Let $f$ be an arbitrary function in $\E{1}$ and let $f_{\textsc{\#DNF}}$ be the function that defines {\sc \#DNF}.
 We will now show a function $h:\str\to \textsc{Struct}[\L_{\textsc{DNF}}]$, where $\L_{\textsc{DNF}} = \{N,P\}$, such that for every $\A\in\str$, it holds that $f(\A) = f_{\textsc{\#DNF}}(h(\A))$, and which can be computed in polynomial time.
@@
 We separate the proof in three cases:
-\begin{enumerate}
-	\item $\X = ()$.
+%\begin{enumerate}
+%	\item $\X = ()$.
 Let $\psi(\z) = \exists \x\,\varphi(\x,\z)$ where $\x = (x_1,\ldots,x_c)$ and $\z = (z_1,\ldots,z_d)$.
@@
 The DNF formula we generate is
-	\[
-	\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
-	\]
-	which has exactly $n$ possible assignments. Finally, we return $h(\A) = \A_{\Phi}$.
-
-	\item $\z = ()$.
+%	\[
+%	\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
+%	\]
+%	which has exactly $n$ possible assignments. Finally, we return $h(\A) = \A_{\Phi}$.
+%	
+%	\item $\z = ()$.
 Let $\psi(\X) = \exists \x\,\varphi(\x,\X)$, where $\x = (x_1,\ldots,x_c)$ and $\X = (X_1,\ldots,X_r)$.
@@
 Let each $\varphi_{i}(\x,\X)$ be of the form
-	\[
-	\varphi_{i}(\x,\X) = X_{\pi_i(1)}(\x^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\x^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\x^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\x^{\nu}_{i,n_i}) \wedge \alpha_i(\x),
-	\]
-	where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$) is the rearrangement of $\x$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x)$ is a FO formula.
+%	\[
+%	\varphi_{i}(\x,\X) = X_{\pi_i(1)}(\x^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\x^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\x^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\x^{\nu}_{i,n_i}) \wedge \alpha_i(\x),
+%	\]
+%	where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$) is the rearrangement of $\x$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x)$ is a FO formula.
 We also define the tuple $\a^{\pi}_{i,j}$ ($\a^{\nu}_{i,j}$) as $\a$ rearranged in the same way as $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$), for each $\a\in A^c$.
-
-	We generate a propositional variable $p^{\ell}_{\a}$ for each $\a\in A^{c_{\ell}}$, and for each $\ell\in\{1,\ldots,r\}$, which will indicate if the tuple $\a$ is in the assignment to $X_{\ell}$.
+%	
+%	We generate a propositional variable $p^{\ell}_{\a}$ for each $\a\in A^{c_{\ell}}$, and for each $\ell\in\{1,\ldots,r\}$, which will indicate if the tuple $\a$ is in the assignment to $X_{\ell}$.
 Let $\p$ be the tuple of all such variables.
-
-	Then, let $\Gamma_i = \{\a\in A^c\mid \A\models\alpha_i(\a)\}$.
+%	
+%	Then, let $\Gamma_i = \{\a\in A^c\mid \A\models\alpha_i(\a)\}$.
 There is exactly $\vert A \vert^c$ tuples to check, and each of those checks can be done in polynomial time, so this set can also be generated in polynomial time.
-
-	Finally, for each $\varphi_i(\x,\X)$ the DNF formula we generate is
-	\[
-	\Phi_i = \bigvee_{\a\in\Gamma_i} p^{\pi_i(1)}_{\a^{\pi}_{i,1}} \wedge \cdots \wedge p^{\pi_i(m_i)}_{\a^{\pi}_{i,m_i}} \wedge \neg p^{\nu_i(1)}_{\a^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\pi_i(n_i)}_{\a^{\nu}_{i,n_i}}.
-	\]
-	and $\Phi = \bigvee_{i = 1}^k \Phi_i \vee \bigvee_{p\in\p} (p\wedge\neg p)$. Finally, we return $h(\A) = \A_{\Phi}$.
-
-	As an example, let $\psi^{*}(X_1,X_2) = \exists x \exists y\, (X_1(x) \wedge X_2(y,x) \wedge \neg X_1(y) \wedge x < y \wedge \neg\exists z\,(x<z\wedge z<y))$.
+%	
+%	Finally, for each $\varphi_i(\x,\X)$ the DNF formula we generate is
+%	\[
+%	\Phi_i = \bigvee_{\a\in\Gamma_i} p^{\pi_i(1)}_{\a^{\pi}_{i,1}} \wedge \cdots \wedge p^{\pi_i(m_i)}_{\a^{\pi}_{i,m_i}} \wedge \neg p^{\nu_i(1)}_{\a^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\pi_i(n_i)}_{\a^{\nu}_{i,n_i}}.
+%	\]
+%	and $\Phi = \bigvee_{i = 1}^k \Phi_i \vee \bigvee_{p\in\p} (p\wedge\neg p)$. Finally, we return $h(\A) = \A_{\Phi}$.
+%	
+%	As an example, let $\psi^{*}(X_1,X_2) = \exists x \exists y\, (X_1(x) \wedge X_2(y,x) \wedge \neg X_1(y) \wedge x < y \wedge \neg\exists z\,(x<z\wedge z<y))$.
 We will compute $h(\B)$, where $\B = \langle \{1,2,3\}, < \rangle$, and we have that $1 < 2 < 3$.
@@
 And so, the DNF formula we generate is
-	\begin{multline*}
-	\Phi^{*} = (p^1_1\wedge p^2_{(2,1)} \wedge \neg p^1_2)\vee(p^1_2\wedge p^2_{(3,2)} \wedge \neg p^1_3)\, \vee \\ (p^1_1 \wedge \neg p^1_1) \vee(p^1_2 \wedge \neg p^1_2) \vee (p^1_3\wedge \neg p^1_3) \vee \bigvee_{i,j\in\{1,2,3\}}(p^2_{(i,j)}\wedge\neg p^2_{(i,j)})
-	\end{multline*}
-
-	Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$.
+%	\begin{multline*}
+%	\Phi^{*} = (p^1_1\wedge p^2_{(2,1)} \wedge \neg p^1_2)\vee(p^1_2\wedge p^2_{(3,2)} \wedge \neg p^1_3)\, \vee \\ (p^1_1 \wedge \neg p^1_1) \vee(p^1_2 \wedge \neg p^1_2) \vee (p^1_3\wedge \neg p^1_3) \vee \bigvee_{i,j\in\{1,2,3\}}(p^2_{(i,j)}\wedge\neg p^2_{(i,j)})
+%	\end{multline*}
+%	
+%	Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$.
 We can conclude that $f(\A) = f_{\textsc{DNF}}(\Phi)$.
-
-	\item Let $\psi = \exists \x\,\varphi(\x,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
+%	
+%	\item Let $\psi = \exists \x\,\varphi(\x,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
 Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$ we generate a DNF-formula $\Phi = h(\A)$ as follows.
@@
 Let each $\varphi_{i}(\x,\X,\z)$ be of the form
-	\[
-	\varphi_{i}(\x,\X,\z) = X_{\pi_i(1)}(\y^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\y^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\y^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\y^{\nu}_{i,n_i}) \wedge \alpha_i(\x,\z),
-	\]
-	where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$) is the rearrangement of $(\x,\z)$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x,\z)$ is a FO formula.
+%	\[
+%	\varphi_{i}(\x,\X,\z) = X_{\pi_i(1)}(\y^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\y^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\y^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\y^{\nu}_{i,n_i}) \wedge \alpha_i(\x,\z),
+%	\]
+%	where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$) is the rearrangement of $(\x,\z)$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x,\z)$ is a FO formula.
 We also define the tuple $(\a,\e)^{\pi}_{i,j}$ ($(\a,\e)^{\nu}_{i,j}$) as $\a$ rearranged in the same way as $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$), for each $(\a,\e)\in A^{c+d}$.
-
-	We generate a propositional variable $p^{\e}_{\ell,\b}$ for each $\e\in A^d$, for each $\ell\in\{1,\ldots,r\}$, and for each $\b\in A^{c_{\ell}}$.
+%	
+%	We generate a propositional variable $p^{\e}_{\ell,\b}$ for each $\e\in A^d$, for each $\ell\in\{1,\ldots,r\}$, and for each $\b\in A^{c_{\ell}}$.
 This variable will be true if there is an assignation to $\P$ such that $\b\in P_{\ell}$ and $\A\models\psi(\P,\e)$.
@@
 Also, let $\rho_{\ell}$ be such that $\rho_{\ell}(\x,\z) = \y_{\ell}$.
-
-	For each $\e\in A^d$, we follow the following procedure.
+%	
+%	For each $\e\in A^d$, we follow the following procedure.
 For each $\varphi_i(\x,\X)$, we consider the set $\Gamma^{\e}_i = \{\a\in A^c\mid \A\models\alpha(\a,\e)\}$, and so, the DNF formula we generate is
-	\[
-	\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{\pi_i(1),(\a,\e)^{\pi}_{i,1}} \wedge \cdots \wedge p^{\e}_{\pi_i(m_i),(\a,\e)^{\pi}_{i,m_i}} \wedge \neg p^{\e}_{\nu_i(1),(\a,\e)^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\e}_{\pi_i(n_i),(\a,\e)^{\nu}_{i,n_i}} \wedge \gamma_{\e} \wedge \theta_{\e},
-	\]
-	where $\gamma_{\e}$ is defined as
-	\[
-	\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
-	\]
-	and $\theta_{\e}$ is defined as
-	\[
-	\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
-	\]
-	Finally,$\Phi = \bigvee_{\e\in A^d} \bigvee_{i = 1}^k \Phi_i^{\e} \vee \bigvee_{p\in\p} (p\wedge\neg p)$, and we return $f(\A) = \A_{\Phi}$.
-
-	As an example, let $\psi^{*}(X_1,X_2,z) = \exists x\, (X_1(x) \wedge X_2(z) \wedge \neg X_1(z) \wedge x < z)$.
+%	\[
+%	\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{\pi_i(1),(\a,\e)^{\pi}_{i,1}} \wedge \cdots \wedge p^{\e}_{\pi_i(m_i),(\a,\e)^{\pi}_{i,m_i}} \wedge \neg p^{\e}_{\nu_i(1),(\a,\e)^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\e}_{\pi_i(n_i),(\a,\e)^{\nu}_{i,n_i}} \wedge \gamma_{\e} \wedge \theta_{\e},
+%	\]
+%	where $\gamma_{\e}$ is defined as
+%	\[
+%	\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
+%	\]
+%	and $\theta_{\e}$ is defined as
+%	\[
+%	\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
+%	\]
+%	Finally,$\Phi = \bigvee_{\e\in A^d} \bigvee_{i = 1}^k \Phi_i^{\e} \vee \bigvee_{p\in\p} (p\wedge\neg p)$, and we return $f(\A) = \A_{\Phi}$.
+%	
+%	As an example, let $\psi^{*}(X_1,X_2,z) = \exists x\, (X_1(x) \wedge X_2(z) \wedge \neg X_1(z) \wedge x < z)$.
 We will compute $h(\B)$, where $\B = \langle \{1,2,3\}, < \rangle$, and we have that $1 < 2 < 3$.
@@
 Now we generate $\gamma_e$ for $e\in\{1,2,3\}$,
-	\[
-	\gamma_e = \bigwedge_{d\in\{1,2,3\}\setminus\{e\}} p^d_{1,1} \wedge p^d_{1,2} \wedge p^d_{1,3} \wedge p^d_{2,1} \wedge p^d_{2,2} \wedge p^d_{2,3},
-	\]
-	and $\theta_e$ for $e\in\{1,2,3\}$,
-	\begin{align*}
-	\theta_1 &= \neg t_1 \wedge \neg t_2,\\
-	\theta_2 &= \neg t_1 \wedge t_2,\\
-	\theta_3 &= t_1 \wedge \neg t_2.		
-	\end{align*}
-	Then, we generate the formulas $\Phi^e_1$, for $e\in\{1,2,3\}$,
-	\begin{align*}
-	\Phi^1_1 &= \square\\
-	\Phi^2_1 &= p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2,\\
-	\Phi^3_1 &= (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3),
-	\end{align*}
-	and so, the DNF formula we generate is
-	\begin{multline*}
-	\Phi = (p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2) \vee (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee \\ \bigvee_{e\in\{1,2,3\}}[(p^e_{1,1}\wedge\neg p^e_{1,1})\vee
-	(p^e_{1,2}\wedge\neg p^e_{1,2})\vee
-	(p^e_{1,3}\wedge\neg p^e_{1,3})\vee
-	(p^e_{2,1}\wedge\neg p^e_{2,1})\vee
-	(p^e_{2,2}\wedge\neg p^e_{2,2})\vee
-	(p^e_{2,3}\wedge\neg p^e_{2,3})].
-	\end{multline*}
-	To demonstrate that our reduction is correct, we will show a function $g:\{\langle\P,\e\rangle \mid \A \models \exists\x\,\varphi(\x,\P,\e)\} \to \{\sigma \mid \sigma(\Phi) = 1\}$ and then prove that it is a bijection.
+%	\[
+%	\gamma_e = \bigwedge_{d\in\{1,2,3\}\setminus\{e\}} p^d_{1,1} \wedge p^d_{1,2} \wedge p^d_{1,3} \wedge p^d_{2,1} \wedge p^d_{2,2} \wedge p^d_{2,3},
+%	\]
+%	and $\theta_e$ for $e\in\{1,2,3\}$,
+%	\begin{align*}
+%	\theta_1 &= \neg t_1 \wedge \neg t_2,\\
+%	\theta_2 &= \neg t_1 \wedge t_2,\\
+%	\theta_3 &= t_1 \wedge \neg t_2.		
+%	\end{align*}
+%	Then, we generate the formulas $\Phi^e_1$, for $e\in\{1,2,3\}$,
+%	\begin{align*}
+%	\Phi^1_1 &= \square\\
+%	\Phi^2_1 &= p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2,\\
+%	\Phi^3_1 &= (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3),
+%	\end{align*}
+%	and so, the DNF formula we generate is
+%	\begin{multline*}
+%	\Phi = (p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2) \vee (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee \\ \bigvee_{e\in\{1,2,3\}}[(p^e_{1,1}\wedge\neg p^e_{1,1})\vee
+%	(p^e_{1,2}\wedge\neg p^e_{1,2})\vee
+%	(p^e_{1,3}\wedge\neg p^e_{1,3})\vee
+%	(p^e_{2,1}\wedge\neg p^e_{2,1})\vee
+%	(p^e_{2,2}\wedge\neg p^e_{2,2})\vee
+%	(p^e_{2,3}\wedge\neg p^e_{2,3})].
+%	\end{multline*}
+%	To demonstrate that our reduction is correct, we will show a function $g:\{\langle\P,\e\rangle \mid \A \models \exists\x\,\varphi(\x,\P,\e)\} \to \{\sigma \mid \sigma(\Phi) = 1\}$ and then prove that it is a bijection.
 Let $g$ be such that for each assignment $\langle \P,\e \rangle$, it holds that $g(\langle \P,\e \rangle) = \sigma$ where $\sigma$ is constructed as follows.
@@
 For each $p_{\ell,\b}^{\e'}$, where $\e'\in A^d\setminus\{\e\}$,  $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e'}) = 1$.
-
-	Let $i$ be such that $\A\models \exists\x\,\varphi_i(\x,\P,\e)$.
+%	
+%	Let $i$ be such that $\A\models \exists\x\,\varphi_i(\x,\P,\e)$.
 It is simple to see that from the way we constructed $\Phi_i^{\e}$, it holds that $\sigma(\Phi_i^{\e}) = 1$.
@@
 We conclude that $f(\A) = f_{\textsc{\#DNF}}(\A_{\Phi})$.
-\end{enumerate}
-
-\textbf{Proof of Theorem \ref{nplusone-strict}}
-
-Suppose it is not properly contained for $n$.
+%\end{enumerate}
+%	
+%\textbf{Proof of Theorem \ref{nplusone-strict}}
+%
+%Suppose it is not properly contained for $n$.
 Let $\z = (z_1,\ldots,z_{n+1})$, let $\u = (u_1,\ldots,u_n)$ and let $\varphi(\z) = \bigwedge_{i = 1}^{n} (z_i < z_{i+1} \wedge \forall y(z_i\not<y \vee y\not<z_{i+1} )) \wedge \forall y(y \not< z_1)$.
@@
 We conclude that $f_{\forall\x\,\psi(\x,\X,\u)}(\B) \geq 1$, which leads to a contradiction.
-
-\textbf{Proof of Theorem \ref{sigma2-pe}}
-
-Let $f = f_{\varphi(\X,\z)}$ such that $\varphi(\X,\z)$ is defined by the grammar $E_2$.
-First we notice that, as stated in \cite{DBLP:journals/jcss/SalujaST95}, for every $\L$-formula $\varphi$, $f_{\varphi}\in\shp$.\\
-
-We will now prove that there is a polynomial time algorithm that decides $L_{f}$.
+%	
+%\textbf{Proof of Theorem \ref{sigma2-pe}}
+%
+%Let $f = f_{\varphi(\X,\z)}$ such that $\varphi(\X,\z)$ is defined by the grammar $E_2$.
+First we notice that, as stated in \cite{SalujaST95}, for every $\L$-formula $\varphi$, $f_{\varphi}\in\shp$.\\
+%
+%We will now prove that there is a polynomial time algorithm that decides $L_{f}$.
 This is equivalent to decide, given $\A= \langle A, \S^\A, \leq^\A \rangle\in\str$, whether there exist assignments $\P,\z$ such that $\A\models\varphi(\P,\z)$.
@@
 Since $\varphi(\X,\x)$ is defined by the grammar $E_2$, there exists $\psi(\X,\x,\y,\u)$ such that $\varphi(\X,\x) = \exists\x\exists\y\forall\u\exists\v\,\psi(\X,\x,\y,\u,\v)$, where every variable in $\v$ appears on a negated second-order literal $\neg X_i$, with $i\in\N$.
-
-Given $\A\in\str$ we generate an equivalent $\L$-formula $\theta(\X,\x,\y)$ with a series of operations.
+%
+%Given $\A\in\str$ we generate an equivalent $\L$-formula $\theta(\X,\x,\y)$ with a series of operations.
 First, we replace every instance of $\v$ by a disjunction of all $r$-tuples in $A^r$, where $r$ is the number of variables in $\v$.
@@
 The result is still a conjunction of Horn clauses.
-
-Now we notice that $\zeta = \exists\X\exists\x\exists\y\,\theta(\X,\x,\y)$ is an existential second-order FO-formula, for which $\A\models\zeta$ can be decided in polynomial time \cite{DBLP:conf/stacs/Gradel91}.
+%
+%Now we notice that $\zeta = \exists\X\exists\x\exists\y\,\theta(\X,\x,\y)$ is an existential second-order FO-formula, for which $\A\models\zeta$ can be decided in polynomial time \cite{Gradel91}.
 
@@
 
-Let $f$ be an arbitrary function in $\XE{2}$.
+First we prove that $\shdhsat$ is in $\eqso(\ehorn)$.
+Recall that each instance of $\shdhsat$ is a disjunction of Horn formulas.
+Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot), \pD(\cdot,\cdot)\}$.
+Each symbol in this vocabulary is used to indicate the same as in Example \ref{ex-hornsat-esop1}, with the addition of $\pD(d,c)$ which indicates that $c$ is a clause in the formula $d$.
+Recall that the formula
+\begin{align*}
+&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
+&\forall c \, (\neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
+&\forall c \forall x \, (\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
+&\forall c \forall x \, (\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
+&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
+&\forall c \forall x \, (\textit{A}(c,x) \vee \neg\textit{T}(x)).
+\end{align*}
+defines $\chsat$. We obtain the following formula $\psi(T,A)$ in $\ehorn$:
+\begin{align*}
+\exists d[&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
+&\forall c \, (\neg \pD(c,d)\vee \neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
+&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
+&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
+&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
+&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \neg\textit{T}(x))]
+\end{align*}
+effectively defines $\chsat$ as for every disjunction of Horn formulas $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
+Therefore, we conclude that $\shdhsat \in \eqso(\ehorn)$.
+
+\vspace{1em}
+We will now prove that $\shdhsat$ is hard for $\eqso$ over a signature $\R$ under parsimonious reductions.
+For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a procedure that computes a function $g_{\alpha}$.
+This function receives a finite $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
+We suppose that $\alpha$ is in sum normal form and:
+$$
+\alpha = c + \sum_{i = 1}^{\text{\#clauses}} \sa{\bar{X}}\sa{\bar{x}}\hat{\varphi}_i(\bar{X},\bar{x}),
+$$
+where each $\hat{\varphi}_i$ is a $\ehorn$ formula.
+
+Consider a finite $\R$-structure $\A$ with domain $A$.
+To simplify the proof, we extend our grammar to allow first-order constants.
+Consider each tuple $\bar{a}\in A^{\length(\bar{x})}$ as a tuple of first-order constants.
+The following formula defines the same function as $\alpha$:
+$$
+c + \sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length(\bar{x})}} \sa{\bar{X}}\hat{\varphi}_i(\bar{X},\bar{e}).
+$$
+We redefine this formula as $c + \sum_{i = 1}^{m} \sa{\bar{X}}\hat{\psi}_i(\bar{X})$, where each $\hat{\psi}_i$ is a formula in $\ehorn$ with the extended grammar.
+
+%\vspace{1em}
+%Let $\alpha$ be equal to some constant $c$. We define $g_{\alpha}$ as a function that, for each $\A\in\ostr[\R]$, outputs the following formula with exactly $c$ satisfying assignments:
+%$$
+%\bigvee_{i = 1}^{c}\neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_c.
+%$$
+
+Let the formula be equal to $c + \sum_{i = 1}^{m}\sa{\bar{X}}\exists\bar{y}\bigwedge_{j = 1}^{n}\forall\bar{z}\,\varphi^i_j(\bar{X},\bar{y},\bar{z})$ where each $\varphi_i$ is a Horn clause.
+The following formula defines the same function as $\alpha$:
+$$
+c + \sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length(\bar{y})}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\bar{z}}}\varphi^i_j(\bar{X},\bar{b},\bar{c}).
+$$
+Note that each $\fo$ formula in each $\varphi^i_j$ has no free variables.
+Therefore, we can evaluate each of these in polynomial time and replace them by $\perp$ and $\top$ where it corresponds.
+Each $\varphi^i_j$ will be of the form $\perp \vee\, \chi^i_j(\bar{X})$ or $\top \vee \chi^i_j(\bar{X})$ where $\chi^i_j$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$.
+The formulas of the form $\top \vee \chi^i_j(\bar{X})$ can be removed entirely, and the formulas of the form $\perp \vee\, \chi^i_j(\bar{X})$ can be replaced by $\chi^i_j(\bar{X})$.
+We obtain the formula
+$$
+c + \sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})
+$$
+where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$.
+
+We make the following observation.
+If $\alpha$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$, then the {\bf propositional formula} defined by $\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ has exactly $\sem{\alpha}(\A)$ assignments and is a disjunction of Horn formulas.
+
+%Let $f$ be an arbitrary function in $\XE{2}$.
 We will now show a function $h:\str\to L(P)$ such that for every $\A\in\str$, it holds that $f(\A) = \shdhsat(h(\A))$, and which can be computed in polynomial time.
@@
 We separate the proof in three cases:\begin{enumerate}
-	\item All free variables in $\psi$ are of first order.
+%	\item All free variables in $\psi$ are of first order.
 Let $\psi = \exists \x \forall \y\,\varphi(\x,\y,\z)$ where $\z = (z_1,\ldots,z_d)$.
@@
 The disjunction of Horn formulas we return is
-	\[
-	\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
-	\]
-	which has exactly $n$ possible assignments. Note that each $p_i$ and $\neg p_i$ is a Horn clause itself.
-
-	\item There are only second-order free variables in $\psi$.
+%	\[
+%	\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
+%	\]
+%	which has exactly $n$ possible assignments. Note that each $p_i$ and $\neg p_i$ is a Horn clause itself.
+%	
+%	\item There are only second-order free variables in $\psi$.
 Let $\psi = \exists \x\forall \y\,\varphi(\x,\y,\X)$, where $\x = (x_1,\ldots,x_c)$, $\y = (y_1,\ldots,y_b)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
 Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a Disj-Horn formula $\Phi = h(\A)$ as follows:
-
-	First, Let $\varphi(\x,\y,\X) = \bigwedge_{i = 1}^k \varphi_{i}(\x,\y,\X)$ where, without loss of generality, each $\varphi_{i}(\x,\y,\X)$ is of the form
-	\[
-	\varphi_{i}(\x,\y,\X) = X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n}) \vee \alpha_i(\x,\y),
-	\]
-	where $\u_{s}$ and every $\u_{t_j}$ have the corresponding number of variables in $(\x,\y)$, and $\alpha(\x,\y)$ is a FO-formula.
+%	
+%	First, Let $\varphi(\x,\y,\X) = \bigwedge_{i = 1}^k \varphi_{i}(\x,\y,\X)$ where, without loss of generality, each $\varphi_{i}(\x,\y,\X)$ is of the form
+%	\[
+%	\varphi_{i}(\x,\y,\X) = X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n}) \vee \alpha_i(\x,\y),
+%	\]
+%	where $\u_{s}$ and every $\u_{t_j}$ have the corresponding number of variables in $(\x,\y)$, and $\alpha(\x,\y)$ is a FO-formula.
 For the rest of the proof, $X_s(\u_s)$ can be absent and it does not affect the proof.
 Note that this formula is equivalent to
-	\[
-	\neg\alpha_i(\x,\y) \to (X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n})),
-	\]
-
-	Then, for each $\a\in A^c$ let $\Gamma_i^{\a} = \{\b\in A^{b}\mid \A\models\neg\alpha(\a,\b)\}$.
+%	\[
+%	\neg\alpha_i(\x,\y) \to (X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n})),
+%	\]
+%	
+%	Then, for each $\a\in A^c$ let $\Gamma_i^{\a} = \{\b\in A^{b}\mid \A\models\neg\alpha(\a,\b)\}$.
 Note that this set can be generated in polynomial time.
@@
 Also, let $\rho_{\ell}(\x,\y) = \u_{\ell}$, and $\rho_{\ell}(\d)$ be the tuple that results of reordering $\d$ using the reordering of the variables of $(\x,\y)$ in $\u_{\ell}$.
-
-	For each $\a\in A^c$ we generate a Horn formula
-	\[
-	\Phi_{\a}^i = \bigwedge_{\b\in\Gamma_i} p^{s}_{\rho_s(\a,\b)} \vee \neg p^{t_1}_{\rho_{t_1}(\a,\b)} \vee \cdots \vee \neg p^{t_n}_{\rho_{t_n}(\a,\b)}.
-	\]
-	and finally we return $\Phi = \bigvee_{\a\in A^c} \bigwedge_{i = 1}^k \Phi_{\a}^i$.
-
-	Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\forall\y\,\varphi(\x,\y,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$.
+%	
+%	For each $\a\in A^c$ we generate a Horn formula
+%	\[
+%	\Phi_{\a}^i = \bigwedge_{\b\in\Gamma_i} p^{s}_{\rho_s(\a,\b)} \vee \neg p^{t_1}_{\rho_{t_1}(\a,\b)} \vee \cdots \vee \neg p^{t_n}_{\rho_{t_n}(\a,\b)}.
+%	\]
+%	and finally we return $\Phi = \bigvee_{\a\in A^c} \bigwedge_{i = 1}^k \Phi_{\a}^i$.
+%	
+%	Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\forall\y\,\varphi(\x,\y,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$.
 We can conclude that $f(\A) = f_{\textsc{Disj-Horn-Sat}}(\Phi)$.
-
-	\item Let $\psi = \exists \x\forall\y\,\varphi(\x,\y,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\y = (y_1,\ldots,y_b)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
+%	
+%	\item Let $\psi = \exists \x\forall\y\,\varphi(\x,\y,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\y = (y_1,\ldots,y_b)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
 Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$ we generate a Disj-Horn formula $\Phi = h(\A)$ as follows:
-
-	As in the previous case, let $\varphi(\x,\y,\X,\z) = \bigwedge_{i = 1}^k$ where $\varphi(\x,\y,\X,\z)$ is of the form:
-	\[
-	\varphi_{i}(\x,\y,\X,\z) = X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n}) \vee \alpha_i(\x,\y,\z),
-	\]
-	where $\u_{s}$ and every $\u_{t_j}$ have the corresponding number of variables in $(\x,\y,\z)$, and $\alpha(\x,\y,\z)$ is a FO-formula.
+%	
+%	As in the previous case, let $\varphi(\x,\y,\X,\z) = \bigwedge_{i = 1}^k$ where $\varphi(\x,\y,\X,\z)$ is of the form:
+%	\[
+%	\varphi_{i}(\x,\y,\X,\z) = X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n}) \vee \alpha_i(\x,\y,\z),
+%	\]
+%	where $\u_{s}$ and every $\u_{t_j}$ have the corresponding number of variables in $(\x,\y,\z)$, and $\alpha(\x,\y,\z)$ is a FO-formula.
 For the rest of the proof, $X_s(\u_s)$ can be absent and it does not affect the proof.
 Note that this formula is equivalent to
-	\[
-	\neg\alpha_i(\x,\y,\z) \to (X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n})).
-	\]
-
-	For each $\e\in A^d$, we do the following.
+%	\[
+%	\neg\alpha_i(\x,\y,\z) \to (X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n})).
+%	\]
+%	
+%	For each $\e\in A^d$, we do the following.
 For each $\a\in A^c$ let $\Gamma^{\e}_{i,\a} = \{\b \in A^b \mid \A\models \neg\alpha_i(\a,\b,\e) \}$.
@@
 And so, the Horn formula we generate is
-	\[
-	\Phi^{\e}_{i,\a} = \bigwedge_{\a\in\Gamma^{\e}_i} p^{\e}_{s,\rho_{s}(\a,\b,\e)} \vee p^{\e}_{t_1,\rho_{t_1}(\a,\b,\e)} \vee \cdots \vee p^{\e}_{t_n,\rho_{t_n}(\a,\b,\e)}.
-	\]
-	And finally we return $\Phi = \bigvee_{\e\in A^d}\bigvee_{\a\in A^c} (\bigwedge_{i=1}^k \Phi_{\e}^{i,\a} \wedge \gamma_{\e} \wedge \theta_{\e} )$, where $\gamma_{\e}$ is defined as
-	\[
-	\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
-	\]
-	and $\theta_{\e}$ is defined as
-	\[
-	\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
-	\]
-
-\end{enumerate}
+%	\[
+%	\Phi^{\e}_{i,\a} = \bigwedge_{\a\in\Gamma^{\e}_i} p^{\e}_{s,\rho_{s}(\a,\b,\e)} \vee p^{\e}_{t_1,\rho_{t_1}(\a,\b,\e)} \vee \cdots \vee p^{\e}_{t_n,\rho_{t_n}(\a,\b,\e)}.
+%	\]
+%	And finally we return $\Phi = \bigvee_{\e\in A^d}\bigvee_{\a\in A^c} (\bigwedge_{i=1}^k \Phi_{\e}^{i,\a} \wedge \gamma_{\e} \wedge \theta_{\e} )$, where $\gamma_{\e}$ is defined as
+%	\[
+%	\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
+%	\]
+%	and $\theta_{\e}$ is defined as
+%	\[
+%	\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
+%	\]
+%	
+%\end{enumerate}
 

### 2016-12-28 LICS/preliminaries.tex
19 dic
Avance en sección 3.

@@
 
+\cristian{Agregar que la igualdad $=$ esta en nuestra lógica.}
+
+\cristian{Agregar acá que usamos las abreviaciones de $<$ o $\subset$.}
+
 \marcelo{Agregar las definiciones de LFP y PFP}
+
 \subsection{Function complexity classes}
@@
 
-\marcelo{Ya no consideramos clases de estructuras, hay que revisar que esta seccion tenga una notacion consistente}
-\martin{done}
-
-\marcelo{Agregar las definiciones de las clases usuales de funciones: FP, \#P, ...}
-\martin{done}
-
 We define the following function complexity classes: $\fp$ is defined as the class of functions computable in polynomial time.
@@
 $\fpspace$ is the class of functions computable in polynomial space.
+$\shpspace$ is defined analogously to $\shp$ but in polynomial space.
 $\nfpspace$ is the class of single-exponentially bounded functions computable in polynomial space.

### 2016-12-28 LICS/saluja.tex

@@
-%\begin{enumerate}
-%	\item Introduce the classes $\eqso(\LL)$.
-%	
-%	\item Study the properties of $\eqso(\LL)$.
-%	\begin{itemize}
-%		\item If $\LL$ is closed under conjunction, then binary product can be defined in $\eqso(\LL)$.
-%		
-%		\item If $\LL$ is closed under XXX, then the first order product can be defined in $\eqso(\LL)$.
-%		
-%		\item Every formula in $\eqso(\LL)$ can be rewrite in \emph{sum normal form}.
-%	\end{itemize}
-%	
-%	\item Introduce Saluja classes.
-%	
-%	\item Show the connection with Saluja classes.
-%	\begin{itemize}
-%		\item New classes are properly contained.
-%		\item Old classes are properly contained/equivalent in new classes. 
-%	\end{itemize}
-%	
-%	\item Study the complexity of new classes.
-%	\begin{itemize}
-%		\item $\eqso(\Sigma_0)$ is in FP.
-%		\item $\eqso(\Sigma_1)$ is $\#P$ easy.
-%		\item $\eqso(\Sigma_1)$ has free pass.
-%	\end{itemize}
-%	
-%	\item Closure properties.
-%	\begin{itemize}
-%		\item $\eqso(\Sigma_1)$ is not closed under substraction.
-%		\item $\eqso(\Sigma_1)$ is not closed under minus 1.
-%	\end{itemize}
-%	
-%\end{enumerate}
+%!TEX root = syntactic.tex
 
@@
 \[
-\QE{0} \; \subsetneq \; \QE{1} \; \subsetneq \; \QU{1} \; \subsetneq \; \QE{2} \; \subsetneq \; \QU{2} \; = \; \eqso{\fo} 
+\QE{0} \; \subsetneq \; \QE{1} \; \subsetneq \; \QU{1} \; \subsetneq \; \QE{2} \; \subsetneq \; \QU{2} \; = \; \eqso(\fo)
 \]
@@
 Indeed, each formula $\varphi(\bar{X}, \bar{x})$ in $\E{i}$ or $\U{i}$ is equivalent to $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ which is in $\QE{i}$ or $\QU{i}$, respectively.
-Then it left to know whether the containment between the sharp and quantitative classes is strict or not. 
+Then it is left to know whether the containment between the sharp and quantitative classes is strict or not. 
 For this, we study first whether each formula $\alpha$ in $\eqso(\LL)$ can be transformed in a \emph{prefix normal form} where all sums quantifiers are restricted to the beginning of the formula. 
@@
 
-The previous results gives the connection the connection between the hierarchy in \cite{SalujaST95} and the $\eqso$-hierarchy. 
+The previous results gives the connection between the hierarchy in \cite{SalujaST95} and the $\eqso$-hierarchy. 
 Indeed, if $\LL$ contains $\logu{1}$ we will have that $\sh{\LL}$ is equivalent to $\eqso(\LL)$ since each formula in $\eqso(\LL)$ can be converted in $\Sigma$-prefix normal form and, therefore, it is equivalent to a formula in $\sh{\LL}$.
@@
 
-By the previous result, we know that functions in the $\eqso$ hierarchy are not closed under subtraction.
+By the previous result, we know that functions in the $\eqso$ hierarchy are unlikely to be closed under subtraction.
 Then, a natural restriction to this question is to ask whether these classes are closed under subtraction by one, namely, if $\CC$ is a class of functions and $f \in \CC$, is $f-1 \in \CC$ where $1$ is the constant function that outputs $1$ for every structure.

### 2016-12-28 LICS/counting.tex
27 dic
Primera versión de Sección 4.

@@
 
-Assume that $\alpha$ is a sentence in $\qso$. Then slightly abusing notation, we also use $\alpha$ to denote a function defined over $\R$ such that for every $\A \in \ostr[\R] $:
+In this section, we show that by syntactically restricting $\qso$ one can capture different function complexity classes. 
+In other words, by using $\qso$ we can extend the theory of descriptive complexity~\cite{immerman2012descriptive} from decision problems to counting problems. 
+For this, we first formalize the notion of \emph{capturing} a function complexity class to then show how to capture classes like $\shp$, $\fp$, and $\fpspace$.
+
+Fix a relational signature $\R = \{R_1, \ldots, R_k\}$ and assume that $\A$ is an ordered finite $\R$-structure with a domain $A$ containing $n$ elements.
+Recall that  $<$ is a linear order on $A$, say $a_1 < a_2 < \ldots < a_n$.
+For every $i \in \{1, \ldots, k\}$, define the encoding of $R_i^\A$, denoted by $\enc(R_i^\A)$, as the following binary string.
+Assume that $\ell = \arity(R_i)$ and consider an enumeration of the $\ell$-tuples over $A$ in the lexicographic order induced by $<$.
+%(that is, $(a_1, \ldots, a_1, a_1)$, $(a_1, \ldots, a_1, a_2)$, $\ldots$, $(a_n, \ldots, a_n, a_{n-1})$, $(a_n, \ldots, a_n, a_n)$). 
+Then let $\enc(R_i^\A)$ be a binary string of length $n^\ell$ such that the $i$-th bit of $\enc(R_i^\A)$ is 1 if the $i$-th tuple in the previous enumeration belongs to $R_i^\A$, and 0 otherwise.
+Moreover, define the encoding of $\A$, denoted by $\enc(\A)$, as the following binary string~\cite{L04}:
 \begin{eqnarray*}
-	\alpha(\enc(\A)) & = & \sem{\alpha}(\A).
+	\enc(\A) & = & 0^n \, 1 \, \enc(R_1^\A) \, \cdots \, \enc(R_k^\A).
 \end{eqnarray*}
+%We define the class of all $\R$-functions, denoted by $\Func(\R)$, as the class of all functions $f: \ostr \rightarrow \bbN$.
+%Given a function complexity class $\CC$ (i.e.
+$f: \Sigma^* \rightarrow \bbN$ for every $f \in \CC$), we say that a function $f \in \Func(\R)$ can be computed in $\CC$ if there exists $g \in \CC$ such that $f(\A) = g(\enc(\A))$ for every $\A \in \ostr$.
+%Note that the function $g$ outputs $f$ for encodings of structures and can behave arbitrarily otherwise.
+We can now formalize the notion of capturing a function complexity class over $\R$-structures.
 \begin{definition} \label{cap-def}
-	Let $\FF$ a fragment of $\qso$ and $\CC$ a function complexity class. Then $\FF$ {\em captures $\CC$ over ordered structures} if
+	Let $\FF$ be a fragment of $\qso$ and $\CC$ a function complexity class. Then we say that {\em  $\FF$ captures $\CC$ over $\R$-structures} if the  following conditions are true:
+	\begin{enumerate}
+		\item for every $\alpha \in \FF$, there exists $f \in \CC$ such that $\sem{\alpha}(\A) = f(\enc(\A))$ for every $\A \in \ostr[\R]$.
 
-	\begin{itemize}
-	\item[(1)] for every $\alpha$ in $\FF$ over a signature $\R$. There exists $f \in \CC$ such that for every $\A \in \ostr[\R]$:  $\alpha(\enc(\A)) = f(\enc(\A))$; and
+		\item for every $f \in \CC$, there exists $\alpha \in \FF$ such that   $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A \in \ostr[\R]$.
+	\end{enumerate} 
+	Finally, we say that $\FF$ captures $\CC$ if $\FF$ captures $\CC$ over $\R$-structures for every relational signature $\R$.
+\end{definition}
+%For the sake of simplification, we denote the first condition by $\FF \subseteq \CC$ and the second condition by $\CC \subseteq \FF$.
+Note that in Definition~\ref{cap-def} the function $f \in \CC$ and the formula $\alpha \in \FF$ must coincide in all the strings that encode $\R$-structures.
+Since we want to capture $\CC$ over a fix set of structures (e.g.
+graphs, matrices), it is natural to just consider strings that encodes $\R$-structures.
+This definition is fairly standard in descriptive complexity \cite{L04,immerman2012descriptive} and all notions for capturing complexity classes restrict $f \in \CC$ similarly.
 
-	\item[(2)] for every $f \in \CC$, let $f$ be defined over $\R$. There exists $\alpha$ in $\FF$ such that for every $\A \in \ostr[\R]$:  $f(\enc(\A)) = \alpha(\enc(\A))$.
-	\end{itemize} 
-\end{definition}
-%In the previous definition, if $\KK$ is the class of all finite structures, then we say that $\FF$ captures $\CC$ over the class of all structures, and if $\KK$is the class of all ordered finite structures, then we say that $\FF$ captures $\CC$ over the class of ordered structures.
+Then, what functional complexity classes can be captured with fragments of $\qso$?
+For answering this question, it is reasonable to start with $\shp$ who is one of the most well-known functional complexity class~\cite{arora2009computational}. 
+Since $\shp$ has a strong similarity with $\np$, one could expect a ``Fagin-like'' Theorem~\cite{fagin1974generalized} for this class. 
+Actually, in~\cite{SalujaST95} it was shown that the class $\sfo$ captures $\shp$.
+In our setting, the class $\sfo$ is contained in $\eqso(\fo)$ which also captures $\shp$ as expected. 
+\begin{proposition} \label{eqso-fo-cap}
+	$\eqso(\fo)$ captures $\shp$ over ordered structures.
+\end{proposition}
+Recall that (see Section~\ref{sec:previous}) that every function class $\# \LL$ is contained in $\eqso(\LL)$.
+Thus, by~\cite{SalujaST95} it directly follows that every $\shp$-function can be defined in $\eqso(\LL)$.
+The other direction is also direct and follows by showing that $\shp$ is closed under first- and second-order sum.
+Moreover, these closure properties also hold if we change the boolean core $\fo$ to any polynomial-time logic like $\lfp$. This gives the following corollary.
+\begin{corollary} \label{eqso-lfp-cap}
+	$\eqso(\lfp)$ captures $\shp$ over ordered structures.
+\end{corollary}
+By following the same approach as~\cite{SalujaST95}, Compton et al.~\cite{ComptonG96} show that $\seso$ captures $\spp$.
+As one could expect, if we parametrize $\eqso$ with $\eso$, we can also capture~$\spp$.
+\begin{proposition} \label{eqso-eso-cap}
+	$\eqso(\eso)$ captures $\spp$ over ordered structures.
+\end{proposition}
+The previous result connecting fragments from $\# \LL$ with $\shp$ and $\spp$ is the further that we can go by using SO-formulas with free variables. 
+So, what happen with other counting complexity classes like $\fp$ and $\fpspace$?
+By trying to search for an $\# \LL$ class for capturing $\fp$ one can easily see that this is pointless.
+Indeed, by using free SO-variables one will easily capture $\shp$-hard functions (see Section~\ref{sec:syntactic}) and by only using free FO-variables one cannot capture exponential functions definable in $\fp$ (e.g.
+$2^{n}$).
+Therefore, it is impossible to capture classes like $\fp$ or $\fpspace$ if we follow previous approaches like~\cite{SalujaST95}.
 
+Instead, if we move from $\eqso$ we have other options for counting like first- and second-order products.
+Actually, if we combine $\qfo$ with $\lfp$ this is exactly what we need to capture $\fp$, the class of function that can be computed in~$\ptime$.
 \begin{theorem} \label{captfp}
@@
 \end{theorem}
+To capture $\fp$ one first shows that any formula in $\qfo(\lfp)$ can be evaluated in $\ptime$. 
+Indeed, $\lfp$ is a polynomial time logic~\cite{Immerman82,vardi1982complexity} and the sum and product quantifiers can be computed in $\fp$. 
+For the other direction, one has to use $\lfp$ to simulate the run of the $\fp$-machine and use the quantitative quantifiers to reconstruct the binary number built in the tape. 
+It is important to notice that the alternation between sum and products quantifiers is crucial for capturing $\fp$.
+That is, without these operators is difficult to capture $\fp$.
 
-\begin{theorem} \label{qfo-pfp-cap}
-	$\qfo(\pfp)$ captures $\nfpspace$ over ordered structures.
-\end{theorem}
-
-Saluja et. al \cite{SalujaST95} define a family of counting classes $\#\LL$ for each fragment $\LL$ of $\so$. For a formula $\varphi(x,X)$ over $\R$, the function $f_{\varphi(x,X)}$ is defined as
-\[
-f_{\varphi(x,X)}(\A) = \vert \{\langle e,P\rangle\mid \A\models\varphi(e,P)\}\vert.
-\]
-for each $\A\in\ostr[\R]$.
-A function $f:\Sigma^*\to\nat$ is in $\#\LL$ if there exists an $\LL$ formula $\varphi(x,X)$ such that $f = f_{\varphi(x,X)}$.
-Let $\sqso(\LL)$ be a fragment of $\qso$ defined by the following grammar:
-\[
-\alpha := \sa{x}\varphi \ \mid \ \sa{X} \varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha,
-\]
-where $\varphi$ is an $\LL$-formula. It can be seen that $\sqso(\LL)$ captures $\#\LL$ over ordered structures and that $\sqso(\LL) \subseteq \eqso(\LL)$.
-
-\begin{theorem} \label{eqso-fo-cap}
-	$\eqso(\fo)$ captures $\shp$ over ordered structures.
-\end{theorem}
-
-\begin{corollary} \label{eqso-lfp-cap}
-	$\eqso(\lfp)$ captures $\shp$ over ordered structures.
-\end{corollary}
-
-\begin{theorem} \label{eqso-eso-cap}
-	$\eqso(\eso)$ captures $\spp$ over ordered structures.
-\end{theorem}
-
+Another natural function complexity class is $\fpspace$~\cite{Ladner89}, the functional version of $\pspace$. 
+Of course, for capturing this class one needs a logical core powerful enough, like $\pfp$, for simulating the run of an $\pspace$-machine. 
+On the other hand, an $\fpspace$ machine can generate outputs of exponentially many bits. 
+For this reason, the second-order sum is not enough for the quantitative layer of a logic for $\fpspace$, that is, one needs to consider second-order products to produce big numbers. 
+This leads us to the fragment $\qso(\pfp)$ which naturally captures $\fpspace$. 
 \begin{theorem} \label{qso-pfp-cap}
@@
 \end{theorem}
+The proof of the previous theorem follows the same lines than the logical characterization of $\fp$: one shows that each function in $\qso(\pfp)$ can be computed in $\fpspace$ and, conversely, the output of an $\fpspace$ can be reconstructed by using $\pfp$ and second-order quantitative quantifiers.
+
+From the proof of the previous theorem a natural question follows: what happen if we use first-order quantitative quantifiers and $\pfp$?
+In~\cite{Ladner89}, Ladner also proposes the class $\nfpspace$ of all functions computed by $\pspace$-machines whose output is polynomially bounded. 
+Interestingly, if we restrict to FO-quantitative quantifiers we can also capture this class.
+\begin{corollary} \label{qfo-pfp-cap}
+	$\qfo(\pfp)$ captures $\nfpspace$ over ordered structures.
+\end{corollary}
+
+The results in this section validate $\qso$ as the logical framework for extending the theory of Descriptive Complexity to function computational complexity.
+In the next sections, we study subfragments of $\eqso$ and, moreover, show how to go beyond $\eqso$ to capture other computational classes.

### 2016-12-28 LICS/framework.tex
Primera versión de Sección 4.

--- 2016-12-28 LICS/framework.tex
@@
 
-We introduce here the logic framework that we use for studying counting complexity classes. 
+We introduce here the logic framework that we use for studying function complexity classes. 
 This logic framework is based on the framework of Weighted Logics~\cite{DrosteG07} (WL) that has been used in the context of weighted automata for studying functions from words (or trees) to semirings.
 We propose here to use the framework of WL over any relational structure and to restrict the semiring to natural numbers. 
-The extension to any relational structure will allow us to study general counting complexity classes and the restriction to the natural numbers will simplify the notation in this context (see Section~\ref{sec:previous} below for a more detailed discussion).
+The extension to any relational structure will allow us to study general function complexity classes and the restriction to the natural numbers will simplify the notation in this context (see Section~\ref{sec:previous} below for a more detailed discussion).
 
@@
 This division between boolean and quantitative levels is essential for understanding the difference between the logic and the quantitative part. 
-Furthermore, this will allow us later to parametrize both levels of the logic in order to capture different counting complexity classes.
+Furthermore, this will allow us later to parametrize both levels of the logic in order to capture different function complexity classes.
 
@@
 
-In the next section, we variate $\FF$ and $\LL$ to capture several counting complexity classes.
+In the next section, we variate $\FF$ and $\LL$ to capture several function complexity classes.
 Before this, we devote the next subsection to show the connection of $\qso$ with previous frameworks for defining functions over relational structures.
@@
 As it was previously discussed, $\qso$ is a fragment of WL.
-The real difference is that we restrict the semiring used in WL to natural numbers in order to study counting complexity classes.
+The real difference is that we restrict the semiring used in WL to natural numbers in order to study function complexity classes.
 Another difference of WL with our approach is that, as far as we know, this is the first paper to study weighted logics over general signature properly to do descriptive complexity of counting classes.
@@
 In contrast to our approach, counting operators are usually used for checking boolean properties over structures and not for outputting values (i.e. they do not define a function).
-Although there are some papers that uses this operator for counting~\cite{XXX}, they have not used the logic for capturing counting complexity classes.
+Although there are some papers that uses this operator for counting~\cite{XXX}, they have not used the logic for capturing function complexity classes.
 
-Finally, the work on~\cite{SalujaST95} and \cite{Gradel91} is of particular interest for our work. 
-In~\cite{SalujaST95}, it was proposed to define functions over structures by using open variables on SO-formulas and count how many extensions of the structure satisfy the formula (i.e.
-by instantiating the open variables).
+%\begin{theorem} \label{no-mult}
+%	Let $\alpha$ be a $\qso$-formula in some fragment of $\qso$.
+There is a $\qso$ formula $\alpha'$ in the same fragment such that $\alpha'$ does not have any subformula of the form $(\beta_1 \cdot \beta_2)$ and $\sem{\alpha} = \sem{\alpha'}$.
+%\end{theorem}
+
+Finally, the work on~\cite{SalujaST95} and \cite{compton1996logical} is of particular interest for our work. 
+In~\cite{SalujaST95}, it was proposed to define functions over structures by using free variables on SO-formulas and count how many extensions of the structure satisfy the formula (i.e.
+by instantiating the free variables).
 Formally, Saluja et.
@@
 for each $\A\in\ostr[\R]$. Then a function $g:\ostr[\R]\to\nat$ is in $\#\LL$ if there exists an formula $\varphi(\bar{x},\bar{X})$ in $\LL$ such that $f = f_{\varphi}$. 
-In~\cite{SalujaST95}, it was shown several results for capturing counting complexity classes which are relevant for our work.
+In~\cite{SalujaST95}, it was shown several results for capturing function complexity classes which are relevant for our work.
 We discuss and use these results in Sections~\ref{sec:complexity} and~\ref{sec:syntactic}.

### 2016-12-28 LICS/counting.tex
Nueva versión del apendice.

@@
 In our setting, the class $\sfo$ is contained in $\eqso(\fo)$ which also captures $\shp$ as expected. 
-\begin{proposition} \label{eqso-fo-cap}
+\begin{proposition} \label{prop:capture-shP}
 	$\eqso(\fo)$ captures $\shp$ over ordered structures.
@@
 Actually, if we combine $\qfo$ with $\lfp$ this is exactly what we need to capture $\fp$, the class of function that can be computed in~$\ptime$.
-\begin{theorem} \label{captfp}
+\begin{theorem} \label{theo:capture-fp}
 	$\qfo(\lfp)$ captures $\fp$ over ordered structures.
@@
 Interestingly, if we restrict to FO-quantitative quantifiers we can also capture this class.
-\begin{corollary} \label{qfo-pfp-cap}
+\begin{corollary} \label{cor:capture-fpspace-poly}
 	$\qfo(\pfp)$ captures $\nfpspace$ over ordered structures.

### 2016-12-28 LICS/appendix.tex
Nueva versión del apendice.
28 dic

@@
 If $\bar{x} = (x_1,\ldots,x_m)$ ($\bar{X} = (X_1,\ldots,X_m)$) is a tuple of first-order (second-order) variables, we denote $\sa{\bar{x}}\alpha$ for $\sa{x_1}\cdots\sa{x_m}\alpha$ and $\sa{\bar{X}}\alpha$ for $\sa{X_1}\cdots\sa{X_m}\alpha$ for each $\qso$ formula $\alpha$.
-We also denote $\length(\bar{x}) = m$ ($\length(\bar{X}) = m$).
-\subsection{Proof of Theorem \ref{no-mult}}
-
-Let $\{+, \cdot, \sa{}, \pa{}\}$ be the set of {\it algebraic operators} in $\qso$.
+We also use $\length(\bar{x})$ ($\length(\bar{X})$) to denote the number or variables in the tuple.
+%\subsection{Proof of Theorem \ref{no-mult}}
+%
+%Let $\{+, \cdot, \sa{}, \pa{}\}$ be the set of {\it algebraic operators} in $\qso$.
 We will prove by induction over the number of algebraic operators in a $\qso$-formula $\alpha$ that there exists some $\qso$-formula $\alpha'$ that respects the stated condition.
@@
 We separate the proof in two main cases:
-\begin{enumerate}
-	\item At least one of $\beta_1, \beta_2$ is not of the form $\pa{x} \gamma$, $\pa{X} \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. Without loss of generality, let $\beta_1$ be such one.
-	\begin{enumerate}
-		\item $\beta_1 = s \in \nat$.
+%\begin{enumerate}
+%	\item At least one of $\beta_1, \beta_2$ is not of the form $\pa{x} \gamma$, $\pa{X} \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula.
+Without loss of generality, let $\beta_1$ be such one.
+%	\begin{enumerate}
+%		\item $\beta_1 = s \in \nat$.
 Then we use $\alpha' = \beta_2 \add \cdots \add \beta_2$, where $\beta_2$ is repeated $s$ times, and if $s = 0$, take $\alpha' = \perp$.
 Since each of the formulas paired by $\add$ has less than $n$ algebraic operators, the statement holds.
-		\item $\beta_1 = (\gamma_1 \add \gamma_2)$.
+%		\item $\beta_1 = (\gamma_1 \add \gamma_2)$.
 Then, we use $\alpha' = (\gamma_1\mult\beta_2 + \gamma_2\mult\beta_2)$ and each of the formulas paired by $\add$ have less than $n$ algebraic operators so the statement holds.
-		\item $\beta_1 = \sa{x} \gamma$. Then, we use $\alpha' = \sa{x} (\gamma\mult\beta_2)$ and as in the previous case the statement holds.
-		\item $\beta_1 = \sa{X} \gamma$, which we treat exactly as the previous one.
-	\end{enumerate}
-	\item Both of $\beta_1$ and $\beta_2$ are of the form $\pa{x} \gamma$, $\pa{X} \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. This case also separates in further cases.
-	\begin{enumerate}
-		\item $\beta_1 = \pa{x} \gamma_1(x)$ and $\beta_2 = \pa{y} \gamma_2(y)$. Then we use $\alpha' = \pa{x}(\gamma_1(x)\mult\gamma_2(x))$, and the statement holds.
-		\item $\beta_1 = \pa{X} \gamma_1(X)$ and $\beta_2 = \pa{Y} \gamma_2(Y)$. [something about arities].
-		\item $\beta_1 = \pa{X} \gamma_1(X)$ and $\beta_2 = \pa{x} \gamma_2(x)$ (w.l.o.g.). [use singleton predicates].
-		\item $\beta_1 = \pa{x} \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). Then, we use $\alpha' = \pa{x} (\gamma \cdot \varphi)$, and the statement holds.
-		\item $\beta_1 = \pa{X} \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). We treat this case exactly as the previous one.
-		\item $\beta_1 = \varphi_1$ and $\beta_2 = \varphi_2$ both $\so$-formulas. Then, we use $\alpha' = (\varphi_1 \wedge \varphi_2)$ and the statement holds.
-	\end{enumerate}
-\end{enumerate}
+%		\item $\beta_1 = \sa{x} \gamma$. Then, we use $\alpha' = \sa{x} (\gamma\mult\beta_2)$ and as in the previous case the statement holds.
+%		\item $\beta_1 = \sa{X} \gamma$, which we treat exactly as the previous one.
+%	\end{enumerate}
+%	\item Both of $\beta_1$ and $\beta_2$ are of the form $\pa{x} \gamma$, $\pa{X} \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. This case also separates in further cases.
+%	\begin{enumerate}
+%		\item $\beta_1 = \pa{x} \gamma_1(x)$ and $\beta_2 = \pa{y} \gamma_2(y)$. Then we use $\alpha' = \pa{x}(\gamma_1(x)\mult\gamma_2(x))$, and the statement holds.
+%		\item $\beta_1 = \pa{X} \gamma_1(X)$ and $\beta_2 = \pa{Y} \gamma_2(Y)$. [something about arities].
+%		\item $\beta_1 = \pa{X} \gamma_1(X)$ and $\beta_2 = \pa{x} \gamma_2(x)$ (w.l.o.g.). [use singleton predicates].
+%		\item $\beta_1 = \pa{x} \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). Then, we use $\alpha' = \pa{x} (\gamma \cdot \varphi)$, and the statement holds.
+%		\item $\beta_1 = \pa{X} \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). We treat this case exactly as the previous one.
+%		\item $\beta_1 = \varphi_1$ and $\beta_2 = \varphi_2$ both $\so$-formulas. Then, we use $\alpha' = (\varphi_1 \wedge \varphi_2)$ and the statement holds.
+%	\end{enumerate}
+%\end{enumerate}
 
@@
 We will now prove that $\shdhsat$ is hard for $\eqso$ over a signature $\R$ under parsimonious reductions.
-For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a procedure that computes a function $g_{\alpha}$.
+For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a polynomial-time procedure that computes a function $g_{\alpha}$.
 This function receives a finite $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
@@
 $$
-\alpha = c + \sum_{i = 1}^{\text{\#clauses}} \sa{\bar{X}}\sa{\bar{x}}\hat{\varphi}_i(\bar{X},\bar{x}),
-$$
-where each $\hat{\varphi}_i$ is a $\ehorn$ formula.
+\alpha = c + \sum_{i = 1}^{\text{\#clauses}} \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigwedge_{j = 1}^{n}\forall\bar{z}\,\varphi^i_j(\bar{X},\bar{x},\bar{y},\bar{z}),
+$$
+where each $\varphi^i_j$ is a Horn clause.
 
@@
 To simplify the proof, we extend our grammar to allow first-order constants.
-Consider each tuple $\bar{a}\in A^{\length(\bar{x})}$ as a tuple of first-order constants.
+Consider each tuple $\bar{a}\in A^{\length(\bar{x})}$, each $\bar{b}\in A^{\length(\bar{y})}$ and each $\bar{c}\in A^{\length(\bar{z})}$ as a tuple of first-order constants.
 The following formula defines the same function as $\alpha$:
 $$
-c + \sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length(\bar{x})}} \sa{\bar{X}}\hat{\varphi}_i(\bar{X},\bar{e}).
-$$
-We redefine this formula as $c + \sum_{i = 1}^{m} \sa{\bar{X}}\hat{\psi}_i(\bar{X})$, where each $\hat{\psi}_i$ is a formula in $\ehorn$ with the extended grammar.
-
-%\vspace{1em}
-%Let $\alpha$ be equal to some constant $c$. We define $g_{\alpha}$ as a function that, for each $\A\in\ostr[\R]$, outputs the following formula with exactly $c$ satisfying assignments:
-%$$
-%\bigvee_{i = 1}^{c}\neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_c.
-%$$
-
-Let the formula be equal to $c + \sum_{i = 1}^{m}\sa{\bar{X}}\exists\bar{y}\bigwedge_{j = 1}^{n}\forall\bar{z}\,\varphi^i_j(\bar{X},\bar{y},\bar{z})$ where each $\varphi_i$ is a Horn clause.
-The following formula defines the same function as $\alpha$:
-$$
-c + \sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length(\bar{y})}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\bar{z}}}\varphi^i_j(\bar{X},\bar{b},\bar{c}).
-$$
-Note that each $\fo$ formula in each $\varphi^i_j$ has no free variables.
+c + \sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length(\bar{x})}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length(\bar{y})}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length(\bar{z})}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
+$$
+Note that each $\fo$ formula in each $\varphi^i_j$ over $(\bar{x},\bar{y},\bar{z})$ has no free variables.
 Therefore, we can evaluate each of these in polynomial time and replace them by $\perp$ and $\top$ where it corresponds.
-Each $\varphi^i_j$ will be of the form $\perp \vee\, \chi^i_j(\bar{X})$ or $\top \vee \chi^i_j(\bar{X})$ where $\chi^i_j$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$.
+Each $\varphi^i_j$ will be of the form $\perp \vee\, \chi^i_j(\bar{X})$ or $\top \vee \chi^i_j(\bar{X})$ where $\chi^i_j$ is a disjunction of $\neg X_{\ell}$'s and at most one $X_{\ell}$.
 The formulas of the form $\top \vee \chi^i_j(\bar{X})$ can be removed entirely, and the formulas of the form $\perp \vee\, \chi^i_j(\bar{X})$ can be replaced by $\chi^i_j(\bar{X})$.
@@
 
-We make the following observation.
-If $\alpha$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$, then the {\bf propositional formula} defined by $\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ has exactly $\sem{\alpha}(\A)$ assignments and is a disjunction of Horn formulas.
+Our idea for the rest of the proof is to define $g_{\alpha}$ for each $\alpha = \sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})$, for $\alpha = c$ and for $\alpha = \beta_1 + \cdots + \beta_m$ where each $\beta_i$ is in one of the two previous cases.
+
+If $\alpha$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ where $\psi_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$, then we obtain the {\bf propositional formula} $g_{\alpha}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulas.
+
+If $\alpha$ is equal to a constant $c$, then we define $g_{\alpha}(\A)$ as the following formula that has exactly $c$ satisfying assignments:
+$$
+g_{\alpha}(\A) = \bigvee_{i = 1}^{c}\neg t_1 \wedge \cdots \wedge \neg t_{i-1} \wedge t_i \wedge \neg t_{i+1} \wedge \cdots \wedge \neg p_c.
+$$ 
+If $\alpha = \beta_1 + \cdots + \beta_m$, let $g_{\beta_i}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\theta^i_{i,j}$ for each $\beta_i$ where each $\theta^i_{j,k}$ is a Horn clause.
+Let $\Theta_i = g_{\beta_i}(\A)$.
+We rename the variables in each $\Theta_i$ so none of them are mentioned more than once in total.
+We add $m$ new variables $t_1,\ldots,t_m$ and we define:
+\begin{align*}
+g_{\alpha}(\A) = &\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^1_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m}}}t) \wedge (t_1 \wedge \bigwedge_{\ell = 2}^{m} \neg t_{\ell})) \vee \\
+&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^2_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each $t$}\\ \text{ mentioned in}\\ \Theta_1,\Theta_3,\ldots,\Theta_{m}}}t) \wedge (t_2 \wedge \bigwedge\limits_{\substack{\ell = 1 \\ \ell \neq 2}}^{m} \neg t_{\ell})) \vee \cdots \vee\\
+&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^m_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m-1}}}t) \wedge (t_m \wedge \bigwedge_{\ell = 1}^{m-1} \neg t_{\ell})).
+\end{align*}
+The number of satisfying assignments for this formula is exactly the sum of satisfying assignments for every $g_{\beta_i}(\A)$.
 

### 2016-12-28 LICS/appendix-complexity.tex

--- 2016-12-26 LICS/appendix.tex
@@
-%!TEX root = main.tex
+\medskip
 
-\subsection{Notation for the appendix}
+%%% DEMOSTRACION DE SQSO(FO) y #P
+\subsection*{Proof of Proposition~\ref{prop:capture-shP}}
 
-For a given signature $\R$, we define $\ostr[\R]^*$ as $$\ostr[\R]^* = \{(\A,v,V) \mid \A\in\ostr[\R]\text{, $v$ ($V$) is a first-order (second-order) assignment for $\A$}  \}.$$
-The {\em conditional count} symbol $(\varphi \mapsto \alpha)$ is defined as $(\neg\varphi + (\varphi\cdot\alpha))$ for given $\so$ formula $\varphi$ and $\qso$ formula $\alpha$.
-Note that for each $(\A,v,V) \in \ostr[\R]^*$,
-$$
-\sem{(\varphi \mapsto \alpha)}(\A,v,V) = 
-\begin{cases}
-\sem{\alpha}(\A,v,V) &\text{if } (\A,v,V)\models\varphi,\\
-0 &\text{otherwise}.
-\end{cases}
-$$
-We will use the symbol $<$ also to denote the lexicographic order over same-sized tuples.
-If $\bar{x} = (x_1,\ldots,x_m)$ and $\bar{y} = (y_1,\ldots,y_m)$ are tuples of first-order variables, we denote $\bar{x} < \bar{y}$ for the formula $\bigvee_{i = 1}^m[\bigwedge_{j = 1}^{i-1}x_j = y_j \wedge x_i < y_i]$.
-Similarly, we use $=$ to denote equality between tuples, as $\bar{x} = \bar{y}$ denotes $\bigwedge_{i = 1}^m(x_i = y_i)$, and also $\bar{x}\leq\bar{y}$ denotes $\bar{x} < \bar{y} \vee \bar{x} = \bar{y}$.
-We also denote $\min(\bar{x}) := \forall\bar{y}(\bar{y} < \bar{x}\vee \bar{y} = \bar{x})$.
+We will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
+This machine, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
+Suppose $\A$ has domain $A$.
+If $\alpha$ is a $\fo$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial time, and accepts if and only if it holds true.
+If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
+If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
+%If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
+If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
+%If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$. 
+If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
+This covers all possible cases for $\alpha$.
+Let $\alpha$ be a formula in $\eqso(\fo)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
+We have that $f$ is a $\shp$-function over $\R$ and $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
 
-If $\bar{x} = (x_1,\ldots,x_m)$ ($\bar{X} = (X_1,\ldots,X_m)$) is a tuple of first-order (second-order) variables, we denote $\sa{\bar{x}}\alpha$ for $\sa{x_1}\cdots\sa{x_m}\alpha$ and $\sa{\bar{X}}\alpha$ for $\sa{X_1}\cdots\sa{X_m}\alpha$ for each $\qso$ formula $\alpha$.
-We also denote $\length(\bar{x}) = m$ ($\length(\bar{X}) = m$).
-\subsection{Proof of Theorem \ref{no-mult}}
+For the other direction, note that Saluja et al.~\cite{SalujaST95} proved that $\shp = \sfo$. 
+%We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso(\fo)$ be its corresponding formula. 
+Since a function in $\sfo$ can also be defined $\eqso(\fo)$ (see Section~\ref{sec:previous}), the condition holds. \qed
 
-Let $\{+, \cdot, \sa{}, \pa{}\}$ be the set of {\it algebraic operators} in $\qso$.
-We will prove by induction over the number of algebraic operators in a $\qso$-formula $\alpha$ that there exists some $\qso$-formula $\alpha'$ that respects the stated condition.
-Let $\text{op}(\alpha)$ be the number of algebraic operators in $\alpha$.
-The case $\text{op}(\alpha) = 0$ is trivial.
-Let $n\in\nat$ be such that for each $\qso$-formula $\beta$ such that $\text{op}(\beta) < n$ there is some $\beta'$ that respects the condition.
-Let $\alpha$ be a $\qso$-formula such that $\text{op}(\alpha) = n$.
-Note that if $\alpha$ is of the form $\alpha = (\beta_1 \add \beta_2)$, $\alpha = \sa{x} \beta$, $\alpha = \pa{x} \beta$, $\alpha = \sa{X} \beta$ or $\alpha = \pa{X} \beta$ the proof follows directly.
-Therefore we assume that $\alpha = (\beta_1 \mult \beta_2)$.
-We separate the proof in two main cases:
-\begin{enumerate}
-	\item At least one of $\beta_1, \beta_2$ is not of the form $\pa{x} \gamma$, $\pa{X} \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. Without loss of generality, let $\beta_1$ be such one.
-	\begin{enumerate}
-		\item $\beta_1 = s \in \nat$.
-Then we use $\alpha' = \beta_2 \add \cdots \add \beta_2$, where $\beta_2$ is repeated $s$ times, and if $s = 0$, take $\alpha' = \perp$.
-Since each of the formulas paired by $\add$ has less than $n$ algebraic operators, the statement holds.
-		\item $\beta_1 = (\gamma_1 \add \gamma_2)$.
-Then, we use $\alpha' = (\gamma_1\mult\beta_2 + \gamma_2\mult\beta_2)$ and each of the formulas paired by $\add$ have less than $n$ algebraic operators so the statement holds.
-		\item $\beta_1 = \sa{x} \gamma$. Then, we use $\alpha' = \sa{x} (\gamma\mult\beta_2)$ and as in the previous case the statement holds.
-		\item $\beta_1 = \sa{X} \gamma$, which we treat exactly as the previous one.
-	\end{enumerate}
-	\item Both of $\beta_1$ and $\beta_2$ are of the form $\pa{x} \gamma$, $\pa{X} \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. This case also separates in further cases.
-	\begin{enumerate}
-		\item $\beta_1 = \pa{x} \gamma_1(x)$ and $\beta_2 = \pa{y} \gamma_2(y)$. Then we use $\alpha' = \pa{x}(\gamma_1(x)\mult\gamma_2(x))$, and the statement holds.
-		\item $\beta_1 = \pa{X} \gamma_1(X)$ and $\beta_2 = \pa{Y} \gamma_2(Y)$. [something about arities].
-		\item $\beta_1 = \pa{X} \gamma_1(X)$ and $\beta_2 = \pa{x} \gamma_2(x)$ (w.l.o.g.). [use singleton predicates].
-		\item $\beta_1 = \pa{x} \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). Then, we use $\alpha' = \pa{x} (\gamma \cdot \varphi)$, and the statement holds.
-		\item $\beta_1 = \pa{X} \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). We treat this case exactly as the previous one.
-		\item $\beta_1 = \varphi_1$ and $\beta_2 = \varphi_2$ both $\so$-formulas. Then, we use $\alpha' = (\varphi_1 \wedge \varphi_2)$ and the statement holds.
-	\end{enumerate}
-\end{enumerate}
-
-\subsection{Proof of Theorem \ref{captfp}}
-
-\martin{no revisar, es un copy-paste de arriba}
+%%% DEMOSTRACION DE QFO(LFP) y FP
+\subsection*{Proof of Theorem~\ref{theo:capture-fp}}
 
@@
 The resulting formula is an arithmetic expression of polynomial size which is evaluated and lastly given as output.
-Note that $f\in\fp$ and $f(\enc(\A)) = \alpha(\enc(\A))$.
+Note that $f\in\fp$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
 
@@
 Adding this values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\alpha(\enc(\A)) = f(\enc(\A))$.
+Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
 \qed
+\medskip
 
-\subsection{Proof of Theorem \ref{qfo-pfp-cap}}
+%%% DEMOSTRACION DE QFO(PFP) y FPSPACE(poly)
+\subsection*{Proof of Corollary~\ref{cor:capture-fpspace-poly}}
 
-For condition (2) in \ref{cap-def}, let $\alpha\in\qfo(\pfp)$ over some $\R$.
+Let $\alpha\in\qfo(\pfp)$ over some $\R$.
 Let $f$ be a function over $\R$ defined by the following procedure.
@@
 The resulting formula is an arithmetic expression of polynomial size which is evaluated and lastly given as output.
-Note that $f\in\nfpspace$ and $f(\enc(\A)) = \alpha(\enc(\A))$.
+Note that $f\in\nfpspace$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
 
-For condition (2) in \ref{cap-def}, let $f\in \nfpspace$ defined over some $\R$.
+Now, let $f\in \nfpspace$ defined over some $\R$.
 Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
@@
 $$
-and adding these values gives $f(\enc(\A))$. Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\alpha(\enc(\A)) = f(\enc(\A))$. \qed
-
-\subsection{Proof of Theorem \ref{eqso-fo-cap}}
-
-For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\fo$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial time, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.%%If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Let $\alpha$ be a formula in $\eqso(\fo)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shp$ function over $\R$ and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
-
-For condition (2) in \ref{cap-def}, note that Saluja et al.
-\cite{DBLP:journals/jcss/SalujaST95} proved that $\shp = \#\fo$.
-We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso(\fo)$ be its corresponding formula.
-Since this formula is also in $\eqso(\fo)$, the condition holds.
-\qed
-
-\subsection{Proof of Theorem \ref{eqso-eso-cap}}
-
-For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic outputs for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-On each of the steps the algorithm adds the current configuration to the output tape.
-If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time without modifying the output tape, and accepts if and only if the condition holds true.
-If $\alpha$ is a constant $s$, then the algorithm produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-%% If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\spp$ function over $\R$ and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
-
-For condition (2) in \ref{cap-def}, note that Compton and Graedel \cite{DBLP:journals/jcss/ComptonG96} proved that $\spp = \#\eso$.
-We also have that $\sqso(\eso)$ captures $\#\eso$ over ordered structures so for each $f\in \spp$ let $\alpha \in \sqso(\eso)$ be its corresponding formula.
-Since this formula is also in $\eqso(\eso)$, the condition holds.
-\qed
-
-\subsection{Proof of Theorem \ref{qso-pfp-cap}}
-
-For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\pfp$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space \cite{DBLP:conf/lics/AbiteboulV89}\martin{no estoy seguro de esta cita}, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-If $\alpha = \pa{X}\beta$, it simulates $M_{\beta}$ on input $(\A,v,V[B/X])$ consecutively for each $B\in A^{arity(X)}$.
-This covers all possible cases for $\alpha$, and each of these steps can be computed in polynomial space.
-Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$ \cite{DBLP:journals/siamcomp/Ladner89}, and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
-
-For the second condition, let $f\in \fpspace$ defined over some $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq 2^{n^\ell}$, where $\A$ has a domain of size $n$.
-Let $\Phi(X)$ be a $\pfp$ formula that represents the following.
-Consider an order over the subsets of $A^{\ell}$ given by the formula $$\varphi_{<}(X,Y) = \exists\bar{u}\big[\neg X(\bar{u})\wedge Y(\bar{u})\wedge \forall\bar{v}\big(
-\bar{u}<\bar{v}\to(X(\bar{u})\iff Y(\bar{v}))\big)\big].$$
-
-Let $B$ be the $p$-th subset with respect to this order and let $w$ be a string of size $2^{n^{\ell}}$ which represents the value of $f(\enc(\A))$ in binary.
-Then $\A\models\Phi(B)$ if and only if the $p$-th bit in $w$ is 1, from least to most significant.
-Since this condition can be verified in polynomial space, and $\pfp$ captures $\pspace$, then $\Phi$ can be expressed in $\pfp$.
-We define
-$\alpha := \sa{X}(\Phi(X)\mult\varphi(X))$, where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
-Note that for each $B\subseteq A^{\ell}$ such that $B$ is the $m$-th element in the given order, $\sem{\varphi(B)}(\A) = 2^m$.
-Therefore, $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-\qed
-
-\subsection{Proof of Theorem \ref{eqso-sigma-zero-in-fp}}
-
-First, for a given signature $\R$ and logic fragment $\LL$ we define a {\em valid form in $\LL$} (or just {\em valid form}).
-A formula $\alpha$ is in valid form in $\LL$ if (1) it is equal to $\varphi$, $1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}1$ for some $\LL$ formula $\varphi$ and some tuple of first-order (second-order) variables $\bar{x}$ ($\bar{X}$), or (2) if it is equal to $(\beta\add\gamma)$ where $\beta$ and $\gamma$ are formulas in valid form.
-Given a $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$ we will recursively construct a formula $\tau(\alpha)$ in valid form such that $\alpha(\enc(\A)) = \tau(\alpha)(\enc(\A))$ for each $\A\in\ostr[\R]$.
-
-First, each natural constant $s$ in $\alpha$ is replaced by $(1\add 1\add\cdots\add 1)$ ($s$ times).
-If $\alpha = 1$ or $\alpha = \varphi \in \loge{0}$, then $\tau(\alpha) = \alpha$.
-We assume that every sub-formula of $\alpha$ is in valid form.
-If $\alpha = (\beta \add \gamma)$ then $\tau(\alpha) = \alpha$.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ this case separates in more cases: If $\alpha = \sa{x}\beta$ and $\beta$ is in one of the valid forms in (1), then $\tau(\alpha) = \alpha$.
-If $\alpha = \sa{X}\beta$ and $\beta$ is in the valid form $\sa{\bar{x}}\gamma$ then $\tau(\alpha) = \sa{\bar{x}}\sa{X}\gamma$; if $\beta$ it is in one of the remaining valid forms in (1), then $\tau(\alpha) = \alpha$.
-If $\alpha = \sa{x}\beta$ ($\alpha = \sa{X}\beta$) and $\beta = (\gamma_1\add\gamma_2)$, then $\tau(\alpha) = \tau(\sa{x}\gamma_1) \add \tau(\sa{x}\gamma_2)$ ($\tau(\alpha) = \tau(\sa{X}\gamma_1) \add \tau(\sa{X}\gamma_2)$).
-This covers all possible cases for $\alpha$.
-Furthermore, suppose that $\LL$ is closed under conjunction and suppose $\alpha = (\beta\mult\gamma)$.
-If $\beta = (\delta_1\add\delta_2)$ (wlog), then $\tau(\alpha) = \tau(\delta_1\mult\gamma)\add\tau(\delta_2\mult\gamma)$.
-If $\beta = \sa{\bar{x}}\delta$, then $\tau(\alpha) = \tau(\sa{\bar{x}}\tau(\delta\mult\gamma))$.
-If $\beta = 1$, then $\tau(\alpha) = \gamma$.
-If $\beta = \varphi$ and $\gamma = \psi$, then $\tau(\alpha) = \varphi\wedge\psi$.
-Let $\alpha$ be a $\eqso(\loge{0})$ formula over a signature $\R$.
-Consider the following procedure that receives $\enc(\A)$ as input, where $\A\in\ostr[\R]$ with domain $A$.
-Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_m$.
-For each $\beta_i$, if $\beta_i = 1$ or $\beta_i = \varphi$, then $\sem{\beta_i}(\A)$ can be computed in polynomial time.
-If $\beta_i = \sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}$, if $\beta_i = \sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = 2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, if $\beta_i = \sa{\bar{X}}\sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, all of which can be computed in polynomial time.
-Otherwise, $\beta_i$ is $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, and then $\sem{\beta_i}(\A) = f_{\varphi}(\enc(\A))$, and $f_{\varphi}\in \#\loge{0}$ which implies that $f_{\varphi}(\enc(\A))$ can be computed in polynomial time \cite{DBLP:journals/jcss/SalujaST95}.
-
-\subsection{Proof of Theorem \ref{eqso-sigma-one-in-eqso-pi-one}}
-
-Let $\alpha$ be a formula in $\eqso(\loge{1})$ over a signature $\R$.
-Let $\tau(\alpha) = \beta_1\add\beta_2\add\cdots\add\beta_m$ defined as in the proof of Theorem \ref{eqso-sigma-zero-in-fp}.
-This is, each $\beta_i$ is of the form $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}\sa{\bar{x}}1$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, where $\varphi$ is a formula in $\loge{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables.
-For each $\beta_i$ we will construct a $\gamma_i$ such that $\gamma_i \in \eqso(\logu{1})$ and $\sem{\beta_i}(\A) = \sem{\gamma_i}(\A)$ for each $\A\in\ostr[\R]$.
-Since the forms $1$, $\sa{\bar{x}}1$, $\sa{\bar{X}}1$ and $\sa{\bar{X}}\sa{\bar{x}}1$ are already valid in $\eqso(\logu{1})$, we suppose $\beta_i$ is in one of the remaining forms.
-Without loss of generality, let $\beta_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y},\bar{X})$ where $\varphi(\bar{x},\bar{y},\bar{X})$ is a quantifier-free formula.
-Similarly as in the proof by Saluja et al.
-\cite{DBLP:journals/jcss/SalujaST95}, we define $\gamma_i = \sa{\bar{x}}\sa{\bar{y}}\sa{\bar{X}}\varphi(\bar{x},\bar{y},\bar{X}) \wedge \forall\bar{z}(\varphi(\bar{x},\bar{z},\bar{X})\to(\bar{z} < \bar{y} \vee \bar{z} = \bar{y}))$ that fulfils the desired condition.
-The theorem holds true by using $\beta = \gamma_1\add\cdots\add\gamma_m$ which is in $\eqso(\logu{1})$
-
-%\subsection{Proof of Theorem \ref{one-sigma-zero}}
-%
-%Towards a contradiction, suppose that there exists a constant function $f(\A) = s > 0$ which is expressible in $\E{0}$.
-This is, there is a formula $\psi(x,X)$ such that $f_{\psi(x,X)}(\A) = \vert \{ \langle e,P\rangle \mid \A\models\psi(e,P) \} \vert = s$ for each $\A\in\ostr$.
-It can be seen that if $X$ contains more than zero variables, then the function $f_{\psi(x,X)}$ grows exponentially.
-We have now that $\psi(x)$ contains only first order variables.
-Consider some $\A \in \ostr$ of size $n$ and let $a_1,\ldots,a_s\in \A$ be the tuples that satisfy $\A \models \psi(a_i)$.
-Now consider a structure $\A'$ that is obtained by duplicating $\A$.
-This is, $\A'$ has size $2n$, its first half is isomorphic to its second half, and also to $\A$.
-We have that there is at least $2s$ tuples $a_i,b_i\in \A'$, $s$ on each half, such that $\A'\models\psi(a_i)$ and $\A'\models\psi(b_i)$, which leads to a contradiction.
-
-%\textbf{Proof of Theorem \ref{mult-sigma-zero}}
-%
-%Let $\alpha \in \sqso(\loge{0})$ with binary product.
-This is, $\alpha$ is defined by the grammar $\alpha := \sa{x}\varphi \ \mid \ \sa{X}\varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \ (\alpha\mult\alpha)$, where $\varphi \in \loge{0}$.
-For convenience, we add the expressions $\sa{x}(\varphi\mult\varphi)$ and $\sa{X}(\varphi\mult\varphi)$ to the grammar.
-Let $k$ be the number of algebraic operators in $\alpha$.
-We will show by induction on $k$ that $\alpha$ is expressible without multiplication.
-If $k = 1$, then the possible cases are $\alpha = \sa{x}\varphi$, where $\varphi\in\loge{0}$, which is already product-free.
-If $k = 2$, then the only that is not two nested summations is $\alpha = \sa{x}(\varphi\mult\psi)$, which can be expressed as $\alpha = \sa{x}(\varphi\wedge\psi)$.
-Suppose that the hypothesis holds for $k = 1$ through $n-1$.
-Let $\alpha$ have $n$ algebraic operators.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ where $\beta$ is an algebraic expression, then $\beta$ has $n-1$ algebraic operators and by hypothesis it can be expressed as product-free.
-If $\alpha = (\beta \mult \gamma)$ there are two cases.
-If at least one between $\beta$ and $\gamma$ has a product, then by hypothesis it can be expressed as product-free.
-Moreover, the new expression for $\alpha$ has strictly less than $n$ algebraic operators so by hypothesis it can be expressed as product-free.
-If there are no products in $\beta$ or $\gamma$, then let $\beta = \sa{x}\delta(x)$.
-We have that $\alpha$ can be expressed as $\sa{y}(\delta(y)\mult\gamma)$, where $y$ is a first-order variable that does not appear in $\beta$ or $\gamma$.
-Since $\delta(y)\mult\gamma$ has $n-1$ algebraic operators, it can be expressed as product-free, and then $\alpha$ can be expressed as product-free.
-%
-%\textbf{Proof of Theorem \ref{fo-prod-sigma-zero}}
-%
-%The resulting grammar is $\alpha := \sa{x}\varphi \ \mid \ \sa{X} \varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \pa{x}\alpha$.
-Consider a more restricted grammar that replaces $\pa{x}\alpha$ by $\pa{x}\sa{y}\varphi$.
-Let $\alpha = \pa{x}\sa{y}(x = y)$.
-This formula produces a constant function equal to 1, which as shown in Theorem \ref{one-sigma-zero} is not expressible in $\E{0}$.
-%
-%\textbf{Proof of Theorem \ref{fp1}}
-%
-%Let $f \in \XE{0}$, and let $\varphi(\X,\x)$ be and {\sc FO}-extended quantifier-free $\L$-formula such that:
-%\begin{eqnarray*}
-%	f(\A) &=& \vert \{\langle \P,\e  \rangle \mid \A \models \varphi(\P,\e) \} \vert
-%\end{eqnarray*}
-%for each $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^m$ and $\P = (P_1,\ldots,P_q)$ is a tuple of predicates.
-We will now show that computing $f(\A)$ can be done in polynomial time.
-%
-%For each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$, let $R_{\beta}$ be a predicate of arity $m$.
-Let $\A^\prime = \langle A, \S^{\A}, R_{\beta}^{\A^\prime}, \leq^{\A} \rangle \in \str$, where $R_{\beta}^{\A^\prime} = \{\d \mid \A \models \beta(\d)\}$.
-Note that each $\beta(\x)$ is fixed in $\varphi(\X,\x)$, and for each $\d \in A^m$, checking whether $\A \models \beta(\d)$ can be done in polynomial time.
-Therefore, generating $R_{\beta}^{\A^\prime}$ can also be done in polynomial time.
-%
-%Let $\psi(\X,\x)$ be obtained by replacing each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$ by $R_{\beta}(\x)$.
-Also, let $g = f_{\psi(\X,\x)}$.
-Note that for each tuple of predicates $\P$ and each $\e \in A^m$, $\A \models \varphi(\P,\e)$ if and only if $\A^\prime \models \psi(\P,\e)$, and so, $g(\A^\prime) = f(\A)$.
-But, $\psi(\X,\x)$ is a quantifier-free $\L$-formula, and therefore, $g \in \#\Sigma_0$.
-Since it is shown in \cite{DBLP:journals/jcss/SalujaST95} that $\#\Sigma_0 \subseteq$ {\sc FP}, we conclude that $f(\A)$ can be evaluated in polynomial time.
-
-\textbf{Proof of Theorem \ref{decisionptime}}
-
-Let $f$ be a function in $\E{1}$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
-\begin{eqnarray*}
-	f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables.
-Let $\y = (\x,\z)$ and let $\psi(\X,\y) = \varphi(\x,\X,\z)$.
-Moreover, let $g = f_{\psi(\X,\y)}$.
-\begin{claim} \label{mono}
-	For each $\A \in \str$, $f(\A) > 0$ iff $g(\A) > 0$.
-\end{claim}
-\begin{proof}
-	($\Rightarrow$) Suppose $f(\A) > 0$.
-Let $\P$ and $\e$ be such that $\A \models \exists \x \, \varphi(\x,\P,\e)$.
-It follows that there is at least one $\d \in A^k$ such that $\A \models \varphi(\d,\P,\e) = \psi(\P,(\d,\e))$.
-Therefore, $g(\A) > 0$.
-
-	($\Leftarrow$) Suppose $g(\A) > 0$.
-Let $\Q$ and $\c = (\c_1,\c_2)$, where $\c_1$ and $\c_2$ have $k$ and $m$ elements respectively, be such that $\A \models \psi(\Q,\c) = \varphi(\c_1,\Q,\c_2)$.
-Then we have that $\A \models \exists \x \, \varphi(\x,\Q,\c_2)$, from which we conclude that $f(\A) > 0$.
-\end{proof}
-Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula, so $g \in\XE{0}$.
-It follows from \ref{mono} that for each $\A \in \str$, it holds that $\A \in L_f$ if and only if $g(\A) > 0$.
-But by Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time.
-From this we conclude that the decision version of $f$ is in $\ptime$.
-
-\textbf{Proof of Theorem \ref{sub-pnp}}
-
-Suppose that $\E{1}$ is closed under substraction, that is, for each pair of functions $f,g\in \E{1}$, there exists $h\in\E{1}$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\str$.
-
-Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\Phi$, where $A$ is the set of variables mentioned in $\Phi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{1,2,3,4\}$:
-\begin{eqnarray*}
-	S_1^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-	S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-	S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-	S_4^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
-\end{eqnarray*}
-Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
-\begin{multline*}
-\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
-\end{multline*}
-Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
-$$
-\varphi(x,X) = (T(x) \vee \neg T(x)).
-$$
-Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Given that $f_{\#3DNF}, f_{all} \in \E{1}$, we have by our initial assumption that $f_{all}-f_{\#3DNF} \in \E{1}$.
-Let $h\in\E{1}$ be such that $h = f_{all}-f_{\#3DNF}$.
-For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is $\conp$-complete.
-However, as we showed previously in Theorem \ref{decisionptime}, since $h\in\E{1}$, we have that $L_h \in \ptime$.
-Then, $\conp \subseteq \ptime$, from which we conclude that $\ptime = \np$.
-
-\textbf{Proof of Theorem \ref{sigma1strict}}
-
-We will show that the $\E{1}$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
-By contradiction, suppose that it is.
-Let $\A = \langle A =\{1\},S^\A = \{1\},\leq^\A=\{(1,1)\} \rangle$.
-Then, $f(\A) = 1$.
-Now let $\B = \langle B =\{1,2\},S^\B = \{1\},\leq^\B=\{(1,1),(1,2),(2,2)\} \rangle$.
-Note that $\A$ is an induced substructure of $\B$.
-
-We have that for each function $g \in \#\Sigma_1$ and structures $\A_1,\A_2\in\str$, if $\A_1$ is an induced substructure of $\A_2$, then $g(\A_1) \leq g(\A_2)$ \cite{DBLP:journals/jcss/SalujaST95}.
-Therefore, $f(\B) \geq f(\A) = 1$.
-However, there is no assignment $s\in B$ to $x$ such that $\B\models\:\varphi(s)$, so $f(\B) = 0$, which leads to a contradiction.
-
-\subsection{Proof of Theorem \ref{sigmafo-minusone}}
-
-Let $\alpha$ be a $\eqso(\logex{1})$ formula over a signature $\R$.
-We will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_n$ be defined as in \ref{eqso-sigma-zero-in-fp}, that is, each $\beta_i$ is equal to $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}\sa{\bar{x}}1$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, where $\varphi$ is a formula in $\logex{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables.
-We further assume that each $\beta_i$ is either $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, since each $1$, $\varphi$ and $\sa{\bar{X}}\varphi$ can be replaced by the formulas $\sa{x}\min(x)$, $\sa{x}(\varphi\wedge\min(x))$ and $\sa{x}\sa{\bar{X}}(\varphi\wedge\min(x))$ respectively, while maintaining the value of $\sem{\alpha}(\A)$.
-
-The intuition behind our reasoning is separated in three points: (1) For each $\beta_i$ of the form $\sa{\bar{x}}\varphi$, the formula $\kappa(\beta_i)$ will count every tuple $\bar{x}$ that satisfies $\varphi$ except for the lexicographically smallest one.
-(2) For each $\beta_i$ of the form $\sa{\bar{X}}\sa{\bar{x}}\varphi$, the formula $\kappa(\beta_i)$ will isolate the smallest $\bar{X}$ that satisfies $\varphi$, and exclude the lexicographically smallest tuple $\bar{x}$ that satisfies $\varphi(\bar{X})$.
-And (3) if $\alpha = (\beta\add\sa{\bar{x}}\varphi)$ or $\alpha = (\beta\add\sa{\bar{X}}\sa{\bar{x}}\varphi)$, the formula $\kappa(\alpha)$ will exclude the lexicographically smallest tuple that satisfies $\varphi$ if and only if $\beta$ is equal to 0.
-
-\vspace{1em}
-
-For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha) = \sa{\bar{x}}\exists\bar{y}[\varphi(\bar{x},\bar{y})\wedge\exists\bar{z}(\varphi(\bar{z},\bar{y})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfils the desired condition.
-
-\vspace{1em}
-
-For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha)$ procedurally as follows: Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
-We suppose that $\varphi$ is in a DNF form that leaves formulas that do not mention $\bar{X}$ as is.
-If it is not, we convert $\varphi$ to this form with a standard transformation algorithm.
-Let $\varphi(\bar{X},\bar{x},\bar{y}) = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$ where each $\varphi_i$ has the form:
-$$
-\varphi_i(\bar{X},\bar{x},\bar{y}) = (\text{conjunction of $X_i$'s}) \wedge (\text{conjunction of $\neg X_i$'s})  \wedge (\text{an $\fo$ formula that does not mention any $X_i$}).
-$$
-Define $\varphi_i^{+}$, $\varphi_i^{-}$ and $\varphi_i^{\fo}$ as the formulas mentioned above.
-
-In our procedure we assume that in $\varphi_i^{+}\wedge\varphi_i^{-}$ every first-order variable from $\bar{x},\bar{y}$ is mentioned at most once.
-If not, we add new first-order variables from $\fv$ to $\bar{y}$ so that no variable is repeated, replace them in $\varphi_i^{+}\wedge\varphi_i^{-}$ and add their respective equalities in $\varphi_i^{\fo}$ so that the new formula is equivalent.
-For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
-
-Furthermore, we assume that $\varphi_i^{\fo}$ defines an ordered partition of the variables in the tuple $(\bar{x},\bar{y})$.
-For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$.
-A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
-On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
-For a given $k$, we define $\cB_k$ as the number of possible ordered partitions for a set of size $k$. 
-If $\varphi_i^{\fo}$ does not define an ordered partition, then for $1 \leq j \leq \cB_{\length(\bar{x},\bar{y})}$ 
-let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
-The formula $\varphi_i(\bar{X},\bar{x},\bar{y})$ is converted into $\bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}}\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \theta^j(\bar{x},\bar{y})$.
-Consider each of these disjuncts as a new $\varphi_i$.
-
-After the previous assumptions, we do the following: for each $X_j\in \bar{X}$ we check every instance of $X_j(\bar{w})$ in $\varphi^{+}_i$ and every instance of $X_j(\bar{z})$ in $\varphi^{-}_i$, where $\bar{w}$ and $\bar{z}$ are subtuples of $(\bar{x},\bar{y})$.
-If the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{w} = \bar{z}$, the entire formula $\varphi_i$ is removed from $\varphi$.
-
-Now let $\bar{v}$ be the tuple of every first-order variable mentioned in $\varphi_i^{+}$ and let $\bar{u}$ be such that $(\bar{x},\bar{y}) = (\bar{u},\bar{v})$.
-We define a formula $\mu_i$ that is satisfied only by the lexicographically smallest $\bar{v}$ that satisfies $\varphi_i^{\fo}$ for some $\bar{u}$:
-$$
-\mu_i(\bar{v}) = \exists\bar{u}\,\varphi_i^{\fo}(\bar{u},\bar{v})\wedge\forall\bar{u}'\forall\bar{v}'(\varphi_i^{\fo}(\bar{u}',\bar{v}')\to\bar{v}\leq\bar{v}').
-$$
-The following is a pivotal formula for this proof.
-Consider some ordered finite structure $\A$ over $\R$.
-Given our conditions for $\varphi_i$, if $\A\models\varphi_i$ and $\A\models\varphi^{\fo}_i(\bar{u},\bar{v})$ then for that $\bar{v}$ we can define a $\bar{X}$ that has exactly the tuples mentioned in $\bar{v}$ and nothing else.
-This is {\em the smallest $\bar{X}$ that satisfies $\varphi_i$ over $\bar{v}$}.
-And if $\bar{v}$ is the lexicographically smallest such that $\A\models\varphi_i(\bar{v})$, then it is the {\em the smallest $\bar{X}$ that satisfies $\varphi_i$}.
-The following formula is satisfied by every pair $(\bar{X},\bar{x})$, except for the one formed by the smallest $\bar{X}$ that satisfies $\varphi_i$ and the lexicographically smallest $\bar{x}$ that satisfies $\varphi_i(\bar{X})$.
-And if no pair satisfies $\varphi_i$, the following formula is satisfied by every pair.
-$$
-\psi_i(\bar{X},\bar{x}) = \exists\bar{v}(\mu_i(\bar{v})\wedge(\neg\varphi^{+}_i(\bar{X},\bar{v})\vee\bigvee_{X \in \bar{X}} \exists\bar{z}(X(\bar{z}) \wedge \bigwedge\limits_{\substack{\text{instances of }X(\bar{w}) \\ \text{in }\varphi^{+}_i(\bar{X},\bar{v})}}\bar{w}\neq\bar{z}))) \vee \exists\bar{x}'(\exists\bar{y}'\varphi_i(\bar{X},\bar{x}',\bar{y}')\wedge \bar{x}'<\bar{x})\vee\neg\exists\bar{v}\mu_i(\bar{v}) .
-$$
-\begin{claim}
-	Let $\varphi_i(\bar{X},\bar{x},\bar{y}) = \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X},\bar{x},\bar{y})$ be a quantifier-free formula that respects all the mentioned assumptions.
-For a given ordered structure $\A$ with domain $A$, if $\A\models\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists some $\bar{B}$ over $\A$ and $\bar{b}\in A^{\length(\bar{x})}$ such that $\A\models\varphi_i(\bar{B},\bar{b},\bar{a})$ for some $\bar{a}\in A^{\length(\bar{y})}$.
-\end{claim}
-\begin{proof}
-	Let $\A$ be an ordered structure with domain $A$ and let $\bar{a}\in A^{\length(\bar{y})}$ and $\bar{b}\in A^{\length(\bar{x})}$ such that $\A\models\varphi^{\fo}_i(\bar{b},\bar{a})$.
-Define $\bar{B} = (B_1,\ldots,B_{\length(\bar{X})})$ as $B_j = \{\bar{c}\mid\bar{c}\text{ is a subtuple of $(\bar{b},\bar{a})$ and $X_j(\bar{c})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{b},\bar{a})$}\}$.
-Towards a contradiction, suppose that $\A\not\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
-By the choice of $\bar{a}$ and $\bar{b}$, and construction of $\bar{B}$ it is clear that $\A\models\varphi^{\fo}_i(\bar{b},\bar{a})\wedge\varphi^{+}_i(\bar{B},\bar{b},\bar{a})$, so we have that $\A\not\models\varphi^{-}_i(\bar{B},\bar{b},\bar{a})$.
-Let $B_t(\bar{c})$ be such that $X_t(\bar{c})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{b},\bar{a})$ and $\A\not\models\neg B_t(\bar{c})$.
-This is, $\bar{c}\in B_t$.
-By the construction of $\bar{B}$ we have that $X_t(\bar{c})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{b},\bar{a})$.
-From our assumptions, we have that (1) every first-order variable is mentioned at most once in $\varphi^{-}_i(\bar{X},\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X},\bar{x},\bar{y})$, (2) $\varphi^{\fo}_i$ defines an ordered partition over $(\bar{x},\bar{y})$, and (3) there are no $X_j(\bar{w})$ in $\varphi^{+}_i$ and $X_j(\bar{z})$ in $\varphi^{-}_i$ such that the ordered partition defined by $\varphi^{\fo}_i$ satisfies $\bar{x} = \bar{z}$.
-Since $\bar{c}$ is a subtuple of $(\bar{b},\bar{a})$ that satisfies $\varphi^{\fo}_i$, and $X_t(\bar{w})$ mentioned in $\varphi^{+}_i$ and $X_t(\bar{z})$ mentioned in $\varphi^{-}_i$ were both assigned the value $\bar{c}$, then the ordered partition satisfies $\bar{w} = \bar{z}$, which follows to a contradiction.
-\end{proof}
-We define $\chi_i = \exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$.
-Now recall that $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$, where each $\varphi_i$ satisfies all the previous assumptions.
-For each $\varphi_i$ we define:
-$$
-\varphi_i'(\bar{X},\bar{x},\bar{y}) = \varphi_i(\bar{X},\bar{x},\bar{y})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\vee\psi_2(\bar{X},\bar{x}))\wedge(\chi_1\vee\chi_2\vee\psi_2(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigvee_{j = 1}^{j = i-1}\chi_j\vee\psi_i(\bar{X},\bar{x})),
-$$
-and lastly, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x},\bar{y})$.
-
-\vspace{1em}
-
-For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y}))$ for some algebraic formula $\beta$ and some quantifier-free formula $\varphi$, we define $\kappa(\alpha)$ as follows: First, perform the same transformations to $\varphi(\bar{X},\bar{x},\bar{y})$ as in the previous case.
-Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$ where each $\varphi_i$ satisfies the previous assumptions.
-We also use the previously defined formulas $\chi_i$ and $\psi_i$.
-
-We recursively define a function $\lambda$ as follows.
-If $\alpha = \sa{\bar{x}}\exists\bar{y}\varphi(\bar{x},\bar{y})$, then $\lambda(\alpha) = \exists\bar{x}'\exists\bar{y}'\varphi(\bar{x}',\bar{y}')$.
-If $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ and $\lambda(\alpha) = \chi_1\vee\cdots\vee\chi_m$ as each $\chi_i$ was previously defined.
-If $\alpha = (\beta + \gamma)$, then $\lambda(\alpha) = \lambda(\beta) \vee \lambda(\gamma)$.
-Note that for a given ordered structure $\A$, then $\A\models\lambda(\alpha)$ if and only if $\sem{\alpha}(\A) > 0$.
-
-Now, for each $\varphi_i$ we define:
-$$
-\varphi_i'(\bar{X},\bar{x},\bar{y}) = \varphi_i(\bar{X},\bar{x},\bar{y})\wedge[\lambda(\beta)\vee[\psi_1(\bar{X},\bar{x})\wedge(\chi_1\vee\psi_2(\bar{X},\bar{x}))\wedge(\chi_1\vee\chi_2\vee\psi_2(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigvee_{j = 1}^{j = i-1}\chi_j\vee\psi_i(\bar{X},\bar{x}))]].
-$$
-And lastly $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x},\bar{y})$.
-%%Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$.
-There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$.
-This separates the proof in three cases:
-%\begin{enumerate}
-%%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
-%%	\begin{eqnarray*}
-%%		f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
-%%	\end{eqnarray*}
-%%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^d$.
-Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily.
-First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
-%%	\begin{eqnarray*}
-%%		\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
-%%	\end{eqnarray*}
-%%	This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
-%%	\begin{eqnarray*}
-%%		\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
-%%	\end{eqnarray*}
-%%	If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
-Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$.
-Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E{1}.$
-%%	
-%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
-%	\begin{eqnarray*}
-%		f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
-%	\end{eqnarray*}
-%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
-%	\begin{eqnarray}
-%	\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
-%	\end{eqnarray}
-%	where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
-Note that $\theta(\x)$ also mentions all variables in $\x$.
-We also assume that $(\x_1,\dots,\x_n,\y) = \x$.
-As an example, the following formula is of this form:
-%	\begin{multline*}
-%	\varphi(\x,\X) =  X_1(x_1,x_2) \wedge X_3(x_3) \wedge X_2(x_4,x_5) \wedge X_3(x_6) \wedge \neg X_1(x_7,x_8) \wedge \\ (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 ) \wedge \\ \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big),
-%	\end{multline*}
-%	where $\x = (x_1,x_2,x_3,x_4,x_5,x_6,x_7,x_8)$ and $\X = (X_1,X_2,X_3)$.
-Here, $n = 4$, $\lambda(1) = 1$, $\lambda(2) = \lambda(4) = 3$ and $\lambda(3) = 2$, $\x_1 = (x_1,x_2)$, $\x_2 = (x_3)$, $\x_3 = (x_4,x_5)$, $\x_4 = (x_6)$ and $\y = (x_7,x_8)$.
-Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_7,x_8)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 )$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2,x_5,x_8\},\{x_6,x_7\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big)$.
-%	
-%	Similarly to the previous proof, we would like to eliminate the {\em lexicographically smallest}\footnote{We consider the lexicographically smallest tuple of predicates as the one in which its predicates contain the lexicographically smallest tuples and do not contain any more tuples than those} tuple of predicates that satisfies the formula \eqref{f1}.
-Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
-We now define
-%	\begin{multline*}
-%	\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to (\u_1,\dots,\u_n)<(\v_1,\dots,\v_n)\Big)\Big],
-%	\end{multline*}
-%	where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
-Note that $\alpha_{\min}$ is satisfied only by the lexicographically smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
-Our new formula is
-%	\begin{multline}
-%	\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge \\ \exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg].
-%	\end{multline}
-%	We now show a result by which the main proof will follow.
-%	\begin{lemma} \label{first}
-%		$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
-%	\end{lemma}
-%	\begin{proof}
-%		Let $\A \in \str$.
-Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j\in[1,n]:\:\lambda(j)=i}\{\d_j\}$.
-We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
-%		\begin{enumerate}
-%			\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$.
-That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$).
-Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$.
-Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$.
-We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \z<\x_i$ or (3) $\theta(\x)\models \x_i<\z$.
-Considering that (2) and (3) are not possible given that both $\z$ and $\x_i$ are assigned the value $\d_i$ and $\A\models\theta(\d)$, we have that $\theta(\x)\models \z = \x_i$.
-But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$, which means $\varphi(\x,\X)$ is inconsistent.
-That is, there's no structure $\A^\prime\in\str$ such that $\A^\prime\models\exists\x\:\varphi(\x,\R^\prime)$ for any assignment $\R^\prime$ to $\X$.
-In particular, $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which contradicts the initial assumption that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-%			\item[(b)] Note that if $\A \models \alpha_{\min}(\c_1,\ldots,\c_n)$, then necessarily $\c_i = \d_i$ for $i\in\{1,\ldots,n\}.$ However, by the construction of $\P$, we see that 
-%			$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \d_j\Big).$$ Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
-%			\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$.
-Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $P^\prime_i$.
-Then $\A\models \neg P^\prime_i(\d_i)$, and so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $P_i$ does not have.
-If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-On both cases, we have a contradiction.
-%		\end{enumerate}
-%		With this, we conclude that for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-%		
-%		Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi(\x,\P)$, we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
-And so, for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ it holds that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1,$ which was to be shown.
-%	\end{proof}
-%	We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
-By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\gamma_i(\x,\X)$ with $i\in\{1,\ldots,\ell\}$ such that
-%	\begin{eqnarray*}
-%		\varphi(\x,\X) &\equiv& \gamma_1(\x,\X) \vee \gamma_2(\x,\X) \vee \dots  \vee \gamma_{\ell}(\x,\X),
-%	\end{eqnarray*}
-%	where, for every $i\in\{1,\ldots,\ell\}$, 
-%	\begin{eqnarray*}
-%		\gamma_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \gamma^{-}_i(\X,\y_i)  \wedge \delta_i(\x).
-%	\end{eqnarray*}
-%	The function $\lambda_i$ is defined analogously to $\lambda$ of the first part of the proof.
-The tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y_i) = \x.$ The formulas $\gamma^{-}_i(\X,\y_i)$, $\delta_i(\x)$ are defined analogously to $\varphi^{-}(\X,\y)$ and $\beta(\x)\,$\footnote[2]{Note that $\delta_i(\x)$ can include subformulas of the form $(\u = \u)$.} respectively (see \eqref{f1}).
-Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables.
-Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$.
-Note that for every formula $\eta(\x)$,
-%	\begin{eqnarray*}
-%		\exists\x\:\eta(\x) &\equiv& \exists\x(\eta(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\eta(\x)\wedge\theta^{g(d)}(\x)).
-%	\end{eqnarray*}
-%	We define the following formulas $\xi_i(\x,\X)$, for $i\in\{1,\ldots,m\}$ where $m = \ell \cdot g(d)$, as follows: 
-%	\begin{eqnarray*}
-%		\xi_1(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^1(\x), \\
-%		& \vdots & \\
-%		\xi_{g(d)}(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
-%		\xi_{g(d)+1}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^1(\x), \\
-%		& \vdots & \\
-%		\xi_{2\cdot g(d)}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
-%		& \vdots & \\
-%		\xi_{(\ell-1) g(d)+1}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^1(\x), \\
-%		& \vdots & \\
-%		\xi_{\ell \cdot g(d)}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^{g(d)}(\x).
-%	\end{eqnarray*}
-%	Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\str.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
-We use an injective\footnote[3]{We need this function to be injective because this way we can assure each one of the $k$ satisfiable disjuncts to be represented} function $\rho:\{1,\ldots,k\}\to\{1,\ldots,m\}$ such that $\varphi_i(\x,\X) = \xi_{\rho(i)}(\x,\X)$ is satisfiable, for every $i\in\{1,\ldots,k\}.$
-%	
-%	We can conclude that
-%	\begin{eqnarray*}
-%		\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X),
-%	\end{eqnarray*}
-%	and for every $i\in\{1,\ldots,k\}$,
-%	\begin{eqnarray*}
-%		\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
-%	\end{eqnarray*}
-%	where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\y_i$ is a $p_i$-tuple of variables in $\x$, $\varphi_i^{-}(\X,\y_i)$ is a conjunction of negated predicates in $\X$, $\theta_i(\x)$ defines a total order on a partition of $\x$, and $\beta_i(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
-Let $\alpha_i(\x) = \theta_i(\x)\wedge\beta_i(\x).$
-%	
-%	%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
-%	%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
-%	
-%	\begin{claim} \label{alpha}
-%		Let $\A\in\str$ and $i\in\{1,\ldots,k\}$. For every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, it holds that $f_{\varphi_i(\s,\X)}(\A) > 0.$
-%	\end{claim}
-%	\begin{proof}
-%		Let $\s = (\s_1,\ldots,\s_{n_i},\t)$ such that $\A\models\alpha_i(\s)$, and let $\P = (\P_1,\ldots,\P_r)$ where $P_i = \bigcup_{j\in[1,n_i]:\:\lambda(j)=i}\{s_i\}$.
-We will show that $\A\models\varphi_i(\s,\P)$.
-By contradiction, suppose that $\A\not\models\exists\x\,\varphi_i(\x,\P)$.
-That is, there is no assignment $\e$ to $\x$ such that $\A\models\varphi(\e,\P).$ Following the same proof as in case (a) of Lemma \ref{first}, we conclude that $\varphi_i(\x,\P)$ is inconsistent.
-However, as we mentioned previously, all of such disjuncts have been eliminated, which leads to a contradiction.
-%	\end{proof}
-%	
-%	Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\exists\x\:\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\Q$ such that $\A\models\exists\x\:\varphi_2(\x,\Q)$, and so on.
-As we already know how to exclude that assignment in the first disjunct, we will now deal with the next disjuncts.
-Similarly to the first part of the proof, for each $i\in\{1,\ldots,k\}$ let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
-%	\begin{multline*}
-%	\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_{n_i},\y)\wedge \\ \forall\v_1\cdots\forall\v_{n_i}\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_{n_i},\w)\wedge\bigvee_{j=1}^{n_i}(\u_j\neq\v_j)\big)\to (\u_1,\dots,\u_{n_i})<(\v_1,\dots,\v_{n_i})\Big)\Big],
-%	\end{multline*}
-%	where $\u_j$ and $\v_j,$ have $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, and $\w$ has $p_i$ variables. Also, let
-%	\begin{multline*}
-%	\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge\limits_{\substack{\ell\in[1,n_i]:\\ \lambda_i(\ell) = j}} \v \neq \u_\ell\Big) \bigg) \bigg].
-%	\end{multline*}
-%	Note that $\psi_i(\X)$ excludes only the lexicographically smallest tuple of predicates $\P$ such that $\A\models\exists\x\:\varphi_i(\P,\x),$ if there is at least one.
-In other words, every assignment $\P^\prime \neq \P$ is such that $\A\models\psi_i(\P^\prime).$ Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
-%	$$
-%	\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
-%	$$
-%	Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Assume that $\A\in\str$.
-Suppose first that there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least one assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\d = (\d_1,\dots,\d_{n_q},\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha_q(\d)$, where $\d_i$ is the corresponding assignment to $\x_i$, for every $i\in\{1,\dots,n_q\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j:\lambda_q(j)=i,j\in[1,n_q]}\{\d_j\}$.
-As we did in Lemma \ref{first} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies both (a) and (b)
-%	\begin{enumerate}
-%		\item[(a)] As we showed in part (a) of Lemma \ref{first}, if there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_q(\x,\R)$, then $\A\models\exists\x\:\varphi_q(\x,\P)$ for this particular $\P$.
-However, as we showed in Claim \ref{alpha}, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_q(\s)$, then there {\em is} such an assignment to $\X$.
-The assignment $\d$ to $\x$ satisfies that $\A\models\alpha_q(\d)$, so it holds that $\A\models\exists\x\:\varphi_q(\x,\P)$.
-It immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
-%		
-%		\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
-(1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
-(2) We can use the proof in Lemma \ref{first} to see that $\A\not\models\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula
-%		$$ \zeta_q(\X) =  (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{q-1}(\v)\vee\psi_q(\X)) $$ 
-%		appears as a conjunct in $\varphi^\prime_i(\x,\X)$.
-However, by the choice of $q$, there is no $i\in\{1,\ldots,q-1\}$ such that $\A\models\exists\v\:\varphi_i(\v,\P)$, and also $\A\not\models\psi_q(\P)$.
-It follows that $\A\not\models\zeta_q(\P)$,  so $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$.
-And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
-%		
-%		\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
-As we deduce from the part (c) of Lemma \ref{first}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
-Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x)$, so $\A\models\psi_i(\P^\prime)$ for each $i$.
-But as we mentioned, also $\A\models\:\psi_q(\P^\prime),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which leads to a contradiction.
-%	\end{enumerate}
-%	
-%	With this, we conclude that for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0$, we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-%	
-%	Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\chi(\x,\P))$ for any formula $\chi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
-And so, for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Hence, from the results in the previous paragraph, if $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}$, we have that $f^\prime = f\dotminus 1\in\E{1}$
-%	
-%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
-%	\begin{eqnarray*}
-%		f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
-%	\end{eqnarray*}
-%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$.
-In order to prove that $f\dotminus 1\in\E{1}$, we define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2, where
-%	\begin{eqnarray*}
-%		\varphi(\x,\X,\z) \equiv \varphi_1(\x,\X,\z) \vee \cdots \vee \varphi_k(\x,\X,\z),
-%	\end{eqnarray*}
-%	and
-%	\begin{eqnarray*}
-%		\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\w_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
-%	\end{eqnarray*}
-%	where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\ell} + c_{\ell} = a_{\ell}$ for $\ell\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, and $\w_i$ has $q_i$ variables.
-Furthermore, we have that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\w_i) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
-The formulas $\varphi^{-}_i(\X,\y_i,\w_i)$, $\beta_i(\x,\z)$, are also defined analogously.
-%	
-%	In this case, we mix both the strategies in cases 1 and 2.
-That is, we are going to {\em isolate} the lexicographically smallest tuple of predicates that satisfies the first satisfiable disjunct, and then {\em exclude} the lexicographically smallest tuple that satisfies the isolated disjunct.
-%	
-%	Let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and
-%	\begin{multline*}
-%	\alpha^i_{\min}(\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}) = \exists\y\exists\w\Big[ \alpha_i(\x_1,\dots,\x_{n_i},\y,\z_1,\dots,\z_{n_i},\w)\wedge \\ \forall\u_1\cdots\forall\u_{n_i}\forall\s\forall\v_1\cdots\forall\v_{n_i}\forall\t\Big(\big(\alpha_i(\u_1,\dots,\u_{n_i},\s,\v_1,\dots,\v_{n_i},\t)\wedge\bigvee_{j=1}^{n_i}(\x_j\neq\u_j \vee \z_j\neq\v_j)\big)\to \\ (\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i})<(\u_1,\dots,\u_{n_i},\v_1,\dots,\v_{n_i})\Big)\Big],
-%	\end{multline*}
-%	and let
-%	\begin{multline*}
-%	\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\w(\exists\u\:\varphi(\u,\X,\w) \wedge \w<\z).
-%	\end{multline*}
-%	Let $\P$ and $\e$ be assignments to $\X$ and $\z$.
-Note that $\psi_i(\X,\z)$ excludes $\P$ and $\e$ only if $\P$ is the {\em lexicographically smallest} tuple of predicates (Same as case 2) such that $\A\models\exists\x\:\varphi_i(\x,\P,\d)$, for some assignment $\d$ to $\z$, and $\e$ is the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_i(\x,\P,\e).$ We define $\varphi_i^\prime(\x,\X,\z)$ in the same way as case 2:
-%	\begin{multline*}
-%	\varphi_i^\prime(\x,\X,\z) = \varphi_i(\x,\X,\z) \wedge \psi_1(\X,\z) \wedge (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\psi_2(\X,\z)) \wedge \cdots \wedge \\ (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\cdots\vee\exists\u\exists\v\:\alpha_{i-1}(\u,\v)\vee\psi_i(\X,\z)).
-%	\end{multline*}
-%	Finally, let $\varphi^\prime(\x,\X,\z) = \bigvee_{i = 1}^k \varphi_i^\prime(\x,\X,\z)$.
-%	
-%	Let $q$ be the least $i\in\{1,\ldots,k\}$ such that $\A\models\exists\x\:\varphi_i(\x,\R,\d)$ for some assignment $R$ to $\X$ and some assignment $\d$ to $\z$.
-Let $\P$ be the lexicographically smallest tuple of predicates such that $\A\models\exists\x\:\varphi_q(\x,\P,\d^\prime)$ for some assignment $\d^\prime$ to $\z$.
-Let $\e$ be the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_q(\x,\P,\e).$ This formula is such that (a) $\A\models\exists\x\:\varphi(\x,\P,\e)$, (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P,\e)$ and $\P$ and $\e$ are the only assignments that satisfy (a) and (b).
-The proof of this is analogous to case 2.
-Therefore, we conclude that $f_{\exists\x\,\varphi^\prime(\x,\X,\z)} = f_{\exists\x\,\varphi(\x,\X,\z)}\dotminus 1.$
-%\end{enumerate}
-
-\textbf{Proof of Theorem \ref{dnf-pars}}
-
-Let $f$ be an arbitrary function in $\E{1}$ and let $f_{\textsc{\#DNF}}$ be the function that defines {\sc \#DNF}.
-We will now show a function $h:\str\to \textsc{Struct}[\L_{\textsc{DNF}}]$, where $\L_{\textsc{DNF}} = \{N,P\}$, such that for every $\A\in\str$, it holds that $f(\A) = f_{\textsc{\#DNF}}(h(\A))$, and which can be computed in polynomial time.
-For every DNF formula $\Phi$ we define $\A_{\Phi} \in \textsc{Struct}[\L_{\textsc{DNF}}]$ as stated in \ref{dnfdef}.
-Let $\psi$ be such that $f = f_{\psi(\X,\z)}$.
-We separate the proof in three cases:
-\begin{enumerate}
-	\item $\X = ()$.
-Let $\psi(\z) = \exists \x\,\varphi(\x,\z)$ where $\x = (x_1,\ldots,x_c)$ and $\z = (z_1,\ldots,z_d)$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a propositional formula $\Phi = h(\A)$.
-First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\,\varphi(\x,\e)$.
-Let $n$ be the number of such assignments to $\z$.
-Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$.
-The DNF formula we generate is
-	\[
-	\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
-	\]
-	which has exactly $n$ possible assignments. Finally, we return $h(\A) = \A_{\Phi}$.
-
-	\item $\z = ()$.
-Let $\psi(\X) = \exists \x\,\varphi(\x,\X)$, where $\x = (x_1,\ldots,x_c)$ and $\X = (X_1,\ldots,X_r)$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a DNF-formula $\Phi = h(\A)$ as follows.
-First, we convert $\varphi(\x,\X)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X)$.
-Let each $\varphi_{i}(\x,\X)$ be of the form
-	\[
-	\varphi_{i}(\x,\X) = X_{\pi_i(1)}(\x^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\x^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\x^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\x^{\nu}_{i,n_i}) \wedge \alpha_i(\x),
-	\]
-	where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$) is the rearrangement of $\x$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x)$ is a FO formula.
-We also define the tuple $\a^{\pi}_{i,j}$ ($\a^{\nu}_{i,j}$) as $\a$ rearranged in the same way as $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$), for each $\a\in A^c$.
-
-	We generate a propositional variable $p^{\ell}_{\a}$ for each $\a\in A^{c_{\ell}}$, and for each $\ell\in\{1,\ldots,r\}$, which will indicate if the tuple $\a$ is in the assignment to $X_{\ell}$.
-Let $\p$ be the tuple of all such variables.
-
-	Then, let $\Gamma_i = \{\a\in A^c\mid \A\models\alpha_i(\a)\}$.
-There is exactly $\vert A \vert^c$ tuples to check, and each of those checks can be done in polynomial time, so this set can also be generated in polynomial time.
-
-	Finally, for each $\varphi_i(\x,\X)$ the DNF formula we generate is
-	\[
-	\Phi_i = \bigvee_{\a\in\Gamma_i} p^{\pi_i(1)}_{\a^{\pi}_{i,1}} \wedge \cdots \wedge p^{\pi_i(m_i)}_{\a^{\pi}_{i,m_i}} \wedge \neg p^{\nu_i(1)}_{\a^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\pi_i(n_i)}_{\a^{\nu}_{i,n_i}}.
-	\]
-	and $\Phi = \bigvee_{i = 1}^k \Phi_i \vee \bigvee_{p\in\p} (p\wedge\neg p)$. Finally, we return $h(\A) = \A_{\Phi}$.
-
-	As an example, let $\psi^{*}(X_1,X_2) = \exists x \exists y\, (X_1(x) \wedge X_2(y,x) \wedge \neg X_1(y) \wedge x < y \wedge \neg\exists z\,(x<z\wedge z<y))$.
-We will compute $h(\B)$, where $\B = \langle \{1,2,3\}, < \rangle$, and we have that $1 < 2 < 3$.
-We generate the tuple of variables $\p = (p^1_1,p^1_2,p^1_3) \cup (p^2_{(i,j)})$ for $(i,j)\in \{1,2,3\}\times\{1,2,3\}$.
-Note that in this procedure we will only have one set $\Gamma_1 = \{(1,2),(2,3)\}$.
-And so, the DNF formula we generate is
-	\begin{multline*}
-	\Phi^{*} = (p^1_1\wedge p^2_{(2,1)} \wedge \neg p^1_2)\vee(p^1_2\wedge p^2_{(3,2)} \wedge \neg p^1_3)\, \vee \\ (p^1_1 \wedge \neg p^1_1) \vee(p^1_2 \wedge \neg p^1_2) \vee (p^1_3\wedge \neg p^1_3) \vee \bigvee_{i,j\in\{1,2,3\}}(p^2_{(i,j)}\wedge\neg p^2_{(i,j)})
-	\end{multline*}
-
-	Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$.
-We can conclude that $f(\A) = f_{\textsc{DNF}}(\Phi)$.
-
-	\item Let $\psi = \exists \x\,\varphi(\x,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$ we generate a DNF-formula $\Phi = h(\A)$ as follows.
-As in the previous case, we convert $\varphi(\x,\X,\z)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X,\z)$.
-Let each $\varphi_{i}(\x,\X,\z)$ be of the form
-	\[
-	\varphi_{i}(\x,\X,\z) = X_{\pi_i(1)}(\y^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\y^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\y^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\y^{\nu}_{i,n_i}) \wedge \alpha_i(\x,\z),
-	\]
-	where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$) is the rearrangement of $(\x,\z)$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x,\z)$ is a FO formula.
-We also define the tuple $(\a,\e)^{\pi}_{i,j}$ ($(\a,\e)^{\nu}_{i,j}$) as $\a$ rearranged in the same way as $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$), for each $(\a,\e)\in A^{c+d}$.
-
-	We generate a propositional variable $p^{\e}_{\ell,\b}$ for each $\e\in A^d$, for each $\ell\in\{1,\ldots,r\}$, and for each $\b\in A^{c_{\ell}}$.
-This variable will be true if there is an assignation to $\P$ such that $\b\in P_{\ell}$ and $\A\models\psi(\P,\e)$.
-Let $\p$ be the tuple of all such variables.
-We also generate a set $T = \{t_1,\ldots,t_{L}\}$ where $L = \lceil \log\vert A^d\vert \rceil$, and an arbitrary injective function $\beta:A^d \to \{0,1\}^L$ where $\beta(\e) = (b^{\e}_1,\ldots,b^{\e}_L)$ for each $\e\in A^d$.
-Also, let $\rho_{\ell}$ be such that $\rho_{\ell}(\x,\z) = \y_{\ell}$.
-
-	For each $\e\in A^d$, we follow the following procedure.
-For each $\varphi_i(\x,\X)$, we consider the set $\Gamma^{\e}_i = \{\a\in A^c\mid \A\models\alpha(\a,\e)\}$, and so, the DNF formula we generate is
-	\[
-	\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{\pi_i(1),(\a,\e)^{\pi}_{i,1}} \wedge \cdots \wedge p^{\e}_{\pi_i(m_i),(\a,\e)^{\pi}_{i,m_i}} \wedge \neg p^{\e}_{\nu_i(1),(\a,\e)^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\e}_{\pi_i(n_i),(\a,\e)^{\nu}_{i,n_i}} \wedge \gamma_{\e} \wedge \theta_{\e},
-	\]
-	where $\gamma_{\e}$ is defined as
-	\[
-	\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
-	\]
-	and $\theta_{\e}$ is defined as
-	\[
-	\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
-	\]
-	Finally,$\Phi = \bigvee_{\e\in A^d} \bigvee_{i = 1}^k \Phi_i^{\e} \vee \bigvee_{p\in\p} (p\wedge\neg p)$, and we return $f(\A) = \A_{\Phi}$.
-
-	As an example, let $\psi^{*}(X_1,X_2,z) = \exists x\, (X_1(x) \wedge X_2(z) \wedge \neg X_1(z) \wedge x < z)$.
-We will compute $h(\B)$, where $\B = \langle \{1,2,3\}, < \rangle$, and we have that $1 < 2 < 3$.
-We generate the tuple of variables $\p = (p^e_{1,1},p^e_{1,2},p^e_{1,3},p^e_{2,1},p^e_{2,2},p^e_{2,3})$, for $e\in \{1,2,3\}$.
-Here, we consider $L = 2$, $T = \{t_1,t_2\}$, and $\beta$ defined as $\beta(1) = (0,0)$, $\beta(2) = (0,1)$ and $\beta(3) = (1,0)$.
-In this procedure we generate the sets $\Gamma^1_1 = \emptyset$, $\Gamma^2_1 = \{1\}$, $\Gamma^3_1 = \{1,2\}$.
-Now we generate $\gamma_e$ for $e\in\{1,2,3\}$,
-	\[
-	\gamma_e = \bigwedge_{d\in\{1,2,3\}\setminus\{e\}} p^d_{1,1} \wedge p^d_{1,2} \wedge p^d_{1,3} \wedge p^d_{2,1} \wedge p^d_{2,2} \wedge p^d_{2,3},
-	\]
-	and $\theta_e$ for $e\in\{1,2,3\}$,
-	\begin{align*}
-	\theta_1 &= \neg t_1 \wedge \neg t_2,\\
-	\theta_2 &= \neg t_1 \wedge t_2,\\
-	\theta_3 &= t_1 \wedge \neg t_2.		
-	\end{align*}
-	Then, we generate the formulas $\Phi^e_1$, for $e\in\{1,2,3\}$,
-	\begin{align*}
-	\Phi^1_1 &= \square\\
-	\Phi^2_1 &= p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2,\\
-	\Phi^3_1 &= (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3),
-	\end{align*}
-	and so, the DNF formula we generate is
-	\begin{multline*}
-	\Phi = (p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2) \vee (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee \\ \bigvee_{e\in\{1,2,3\}}[(p^e_{1,1}\wedge\neg p^e_{1,1})\vee
-	(p^e_{1,2}\wedge\neg p^e_{1,2})\vee
-	(p^e_{1,3}\wedge\neg p^e_{1,3})\vee
-	(p^e_{2,1}\wedge\neg p^e_{2,1})\vee
-	(p^e_{2,2}\wedge\neg p^e_{2,2})\vee
-	(p^e_{2,3}\wedge\neg p^e_{2,3})].
-	\end{multline*}
-	To demonstrate that our reduction is correct, we will show a function $g:\{\langle\P,\e\rangle \mid \A \models \exists\x\,\varphi(\x,\P,\e)\} \to \{\sigma \mid \sigma(\Phi) = 1\}$ and then prove that it is a bijection.
-Let $g$ be such that for each assignment $\langle \P,\e \rangle$, it holds that $g(\langle \P,\e \rangle) = \sigma$ where $\sigma$ is constructed as follows.
-For each $t_i\in T$, let $\sigma(t_i) = b^{\e}_i$.
-For each $p_{\ell,\b}^{\e}$, where $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e}) = 1$ if and only if $\b \in P_{\ell}$.
-For each $p_{\ell,\b}^{\e'}$, where $\e'\in A^d\setminus\{\e\}$,  $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e'}) = 1$.
-
-	Let $i$ be such that $\A\models \exists\x\,\varphi_i(\x,\P,\e)$.
-It is simple to see that from the way we constructed $\Phi_i^{\e}$, it holds that $\sigma(\Phi_i^{\e}) = 1$.
-Also, note that every different assignment $\langle \P',\e' \rangle$, it holds that $\sigma(\langle \P',\e' \rangle) \neq \sigma(\langle \P,\e \rangle)$.
-Therefore, $g$ is an injective function.
-Now, let $\sigma$ be an arbitrary assignment such that $\sigma(\Phi) = 1$.
-Let $i$ and $\e$ be such that $\sigma(\Phi^{\e}_i)$.
-Note that for every $\e'\in A^d\setminus\{\e\}$, it holds that $\sigma(\theta_{\e'}) = 0$, so there is exactly one possible $\e$ associated to $\sigma$.
-Consider the assignment $\P$ where $\b\in P_{\ell}$ if and only if $\sigma(p^{\e}_{\ell,\b}) = 1$, for each $\ell$ and $\b \in A^{c_{\ell}}$.
-Since there is at least one $\a\in\Gamma_i$ for which $\sigma$ satisfies its respective subformula, it holds that $\A\models\varphi_i(\a,\P,\e)$, which implies that $\A\models\exists\x\,\varphi(\x,\P,\e)$.
-From this we deduce that $g$ is surjective, and therefore a bijection.
-We conclude that $f(\A) = f_{\textsc{\#DNF}}(\A_{\Phi})$.
-\end{enumerate}
-
-\textbf{Proof of Theorem \ref{nplusone-strict}}
-
-Suppose it is not properly contained for $n$.
-Let $\z = (z_1,\ldots,z_{n+1})$, let $\u = (u_1,\ldots,u_n)$ and let $\varphi(\z) = \bigwedge_{i = 1}^{n} (z_i < z_{i+1} \wedge \forall y(z_i\not<y \vee y\not<z_{i+1} )) \wedge \forall y(y \not< z_1)$.
-Note that $f_{\varphi(\z)}(\A) = 1$ if $\A$ has at least $n+1$ elements, and  $f_{\varphi(\z)}(\A) = 0$ otherwise, for each $\A\in\str$.
-Let $\forall\x\,\psi(\x,\X,\u)$ be such that $f_{\varphi(\z)} = f_{\forall\x\,\psi(\x,\X,\u)}$.
-Let $\A = \langle A, \S^{\A}, <^{\A} \rangle\in\str$ where $\vert A \vert = n+1$.
-Note that $f_{\forall\x\,\psi(\x,\X,\u)}(\A) = 1$.
-Let $(\P,\e)$ be such that $\A\models\forall\x\,\psi(\x,\P,\e)$.
-Let $a\in A$ be such that $a\not\in\e$.
-Consider the vocabulary $\L' = \L \cup \{\Q\}$, and let $\A' = \A \cup \langle \Q^{\A'} \rangle$ where $\Q^{\A'} = \P$.
-Note that $\A\models\varphi(\P,\e)$ if and only if $\A'\models\varphi(\e)$.
-Consider $\B'$ as the induced substructure of $\A'$ with domain $A\setminus\{a\}$, and let $\Q^{\B'}$ be the induced interpretation of $\Q$.
-Also let $\B$ be the induced substructure of $\A$ with domain $A\setminus\{a\}$.
-As universal formulae are preserved under substructure\footnote{Hodges, 1997.
-P.
-183}, we have that $\B'\models\forall\x\,\psi(\x,\e)$, which implies that $\B\models\forall\x\,\psi(\x,\Q^{\B'},\e)$.
-We conclude that $f_{\forall\x\,\psi(\x,\X,\u)}(\B) \geq 1$, which leads to a contradiction.
-
-\textbf{Proof of Theorem \ref{sigma2-pe}}
-
-Let $f = f_{\varphi(\X,\z)}$ such that $\varphi(\X,\z)$ is defined by the grammar $E_2$.
-First we notice that, as stated in \cite{DBLP:journals/jcss/SalujaST95}, for every $\L$-formula $\varphi$, $f_{\varphi}\in\shp$.\\
-
-We will now prove that there is a polynomial time algorithm that decides $L_{f}$.
-This is equivalent to decide, given $\A= \langle A, \S^\A, \leq^\A \rangle\in\str$, whether there exist assignments $\P,\z$ such that $\A\models\varphi(\P,\z)$.
-This is the same as $\A\models\exists\X\exists\x\,\varphi(\X,\x)$.
-Since $\varphi(\X,\x)$ is defined by the grammar $E_2$, there exists $\psi(\X,\x,\y,\u)$ such that $\varphi(\X,\x) = \exists\x\exists\y\forall\u\exists\v\,\psi(\X,\x,\y,\u,\v)$, where every variable in $\v$ appears on a negated second-order literal $\neg X_i$, with $i\in\N$.
-
-Given $\A\in\str$ we generate an equivalent $\L$-formula $\theta(\X,\x,\y)$ with a series of operations.
-First, we replace every instance of $\v$ by a disjunction of all $r$-tuples in $A^r$, where $r$ is the number of variables in $\v$.
-The result is still a Horn clause.
-Second, we replace every instance of $\u$ by a conjunction of all $s$-tulples in $A^s$, where $s$ is the number or variables in $\u$.
-The result is still a conjunction of Horn clauses.
-
-Now we notice that $\zeta = \exists\X\exists\x\exists\y\,\theta(\X,\x,\y)$ is an existential second-order FO-formula, for which $\A\models\zeta$ can be decided in polynomial time \cite{DBLP:conf/stacs/Gradel91}.
-
-\textbf{Proof of Theorem \ref{sigma2hard}}
-
-Let $f$ be an arbitrary function in $\XE{2}$.
-We will now show a function $h:\str\to L(P)$ such that for every $\A\in\str$, it holds that $f(\A) = \shdhsat(h(\A))$, and which can be computed in polynomial time.
-Let $\psi$ be such that $f = f_{\psi}$.
-We separate the proof in three cases:\begin{enumerate}
-	\item All free variables in $\psi$ are of first order.
-Let $\psi = \exists \x \forall \y\,\varphi(\x,\y,\z)$ where $\z = (z_1,\ldots,z_d)$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a propositional formula $\Phi = h(\A)$.
-First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\forall \y\,\varphi(\x,\y,\e)$.
-Let $n$ be the number of such assignments to $\z$.
-Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$.
-The disjunction of Horn formulas we return is
-	\[
-	\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
-	\]
-	which has exactly $n$ possible assignments. Note that each $p_i$ and $\neg p_i$ is a Horn clause itself.
-
-	\item There are only second-order free variables in $\psi$.
-Let $\psi = \exists \x\forall \y\,\varphi(\x,\y,\X)$, where $\x = (x_1,\ldots,x_c)$, $\y = (y_1,\ldots,y_b)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a Disj-Horn formula $\Phi = h(\A)$ as follows:
-
-	First, Let $\varphi(\x,\y,\X) = \bigwedge_{i = 1}^k \varphi_{i}(\x,\y,\X)$ where, without loss of generality, each $\varphi_{i}(\x,\y,\X)$ is of the form
-	\[
-	\varphi_{i}(\x,\y,\X) = X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n}) \vee \alpha_i(\x,\y),
-	\]
-	where $\u_{s}$ and every $\u_{t_j}$ have the corresponding number of variables in $(\x,\y)$, and $\alpha(\x,\y)$ is a FO-formula.
-For the rest of the proof, $X_s(\u_s)$ can be absent and it does not affect the proof.
-Note that this formula is equivalent to
-	\[
-	\neg\alpha_i(\x,\y) \to (X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n})),
-	\]
-
-	Then, for each $\a\in A^c$ let $\Gamma_i^{\a} = \{\b\in A^{b}\mid \A\models\neg\alpha(\a,\b)\}$.
-Note that this set can be generated in polynomial time.
-For each $\ell\in\{1,\ldots,r\}$, we generate a propositional variable $p^{\ell}_{\d}$ for each $\d\in A^{c_{\ell}}$.
-Let $\p$ be the tuple of all such variables.
-Also, let $\rho_{\ell}(\x,\y) = \u_{\ell}$, and $\rho_{\ell}(\d)$ be the tuple that results of reordering $\d$ using the reordering of the variables of $(\x,\y)$ in $\u_{\ell}$.
-
-	For each $\a\in A^c$ we generate a Horn formula
-	\[
-	\Phi_{\a}^i = \bigwedge_{\b\in\Gamma_i} p^{s}_{\rho_s(\a,\b)} \vee \neg p^{t_1}_{\rho_{t_1}(\a,\b)} \vee \cdots \vee \neg p^{t_n}_{\rho_{t_n}(\a,\b)}.
-	\]
-	and finally we return $\Phi = \bigvee_{\a\in A^c} \bigwedge_{i = 1}^k \Phi_{\a}^i$.
-
-	Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\forall\y\,\varphi(\x,\y,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$.
-We can conclude that $f(\A) = f_{\textsc{Disj-Horn-Sat}}(\Phi)$.
-
-	\item Let $\psi = \exists \x\forall\y\,\varphi(\x,\y,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\y = (y_1,\ldots,y_b)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$ we generate a Disj-Horn formula $\Phi = h(\A)$ as follows:
-
-	As in the previous case, let $\varphi(\x,\y,\X,\z) = \bigwedge_{i = 1}^k$ where $\varphi(\x,\y,\X,\z)$ is of the form:
-	\[
-	\varphi_{i}(\x,\y,\X,\z) = X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n}) \vee \alpha_i(\x,\y,\z),
-	\]
-	where $\u_{s}$ and every $\u_{t_j}$ have the corresponding number of variables in $(\x,\y,\z)$, and $\alpha(\x,\y,\z)$ is a FO-formula.
-For the rest of the proof, $X_s(\u_s)$ can be absent and it does not affect the proof.
-Note that this formula is equivalent to
-	\[
-	\neg\alpha_i(\x,\y,\z) \to (X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n})).
-	\]
-
-	For each $\e\in A^d$, we do the following.
-For each $\a\in A^c$ let $\Gamma^{\e}_{i,\a} = \{\b \in A^b \mid \A\models \neg\alpha_i(\a,\b,\e) \}$.
-Note that this set can be generated in polynomial time.
-For each $\ell\in\{1,\ldots,r\}$ we generate a propositional variable $p^{\e}_{\ell,\d}$ for each $\d\in A^{c_{\ell}}$.
-Let $\p$ be the tuple of all such variables.
-We also generate a set $T = \{t_1,\ldots,t_{L}\}$ where $L = \lceil \log\vert A^d\vert \rceil$, and an arbitrary injective function $\beta:A^d \to \{0,1\}^L$ where $\beta(\e) = (b^{\e}_1,\ldots,b^{\e}_L)$ for each $\e\in A^d$.
-Also, let $\rho_{\ell}(\x,\y,\z) = \u_{\ell}$, and $\rho_{\ell}(\d)$ be the tuple that results of reordering $\d$ using the order of the variables in $(\x,\y,\z)$ in $\u_{\ell}$.
-And so, the Horn formula we generate is
-	\[
-	\Phi^{\e}_{i,\a} = \bigwedge_{\a\in\Gamma^{\e}_i} p^{\e}_{s,\rho_{s}(\a,\b,\e)} \vee p^{\e}_{t_1,\rho_{t_1}(\a,\b,\e)} \vee \cdots \vee p^{\e}_{t_n,\rho_{t_n}(\a,\b,\e)}.
-	\]
-	And finally we return $\Phi = \bigvee_{\e\in A^d}\bigvee_{\a\in A^c} (\bigwedge_{i=1}^k \Phi_{\e}^{i,\a} \wedge \gamma_{\e} \wedge \theta_{\e} )$, where $\gamma_{\e}$ is defined as
-	\[
-	\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
-	\]
-	and $\theta_{\e}$ is defined as
-	\[
-	\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
-	\]
-
-\end{enumerate}
-
-\textbf{Proof of Theorem \ref{so-rec}}
-
-\textbf{Proof of Theorem \ref{tqfo-fo-cap}}
-
-For the second condition, let $f \in \shl$.
-We will address the case where $\R$ contains only one binary predicate $E$, and the rest of the cases can be deduced from this.
-Let $M$ be a non-deterministic logspace machine such that $f(\A) = \acc_M(\A)$ for each $\A \in \ostr[R]$.
-Suppose ${\cal Q} = \{q_1,\ldots,q_{\ell}\}$ is the set of states of $M$, where $q_1$ is the initial state, and $q_{\ell}$ is only final state of $M$.
-Let $n = \vert A \vert$ and let $w = \enc(\A) \in \{0,1\}^{n^2}$.
-We assume that $M$ with input $w$ uses space $s_M(w) < c\cdot\log(n)$ and furthermore, $s_M(w) < n-2$.
-We notate $M(w)$ as the graph of configurations of $M$ running on input $w$.
-
-We represent configurations with a tuple of fixed size.
-The formula $\varphi(\bar{x},\bar{y})$ describes a procedure that given a configuration generates a possible next configuration.
-The formula $\varphi_I(\bar{x})$ describes that $\bar{x}$ is the initial configuration of $M(w)$.
-The formula $\varphi_F(\bar{x})$ describes that $\bar{x}$ is an accepting (final) configuration of $M(w)$.
-The formula we construct is:
-$$
-\alpha = \sa{\bar{x}}\sa{\bar{y}}([\pth \varphi(\bar{u},\bar{v})](\bar{x},\bar{y})\cdot \varphi_I(\bar{x})\cdot\varphi_F(\bar{y})).
-$$
-
-To illustrate our idea, we will show a simplified example.
-Consider a machine $M$ that works in exactly $\log_2(n)$ space and only allows 0 or 1 in the working tape.
-Consider an input $\A$ of size 16 (that is, $A = \{0,\ldots,9,A,\ldots,F\}$).
-Let some configuration $s$ have 0011 in the working tape, the head in the input tape is in position 26, and the head in the input tape is in position 2 (we consider 0-indexed positions).
-Also, $Q = \{q_1,\ldots,q_5\}$ and the current state is $q_3$.
-
-As a first approach, we will use a 9-tuple $\bar{a} = (a_1,\ldots,a_9)$ to represent $s$.
-That is, $(a_1,a_2) = (1,A)$ represent the position of the head in the input tape, $a_3 = 2$ represents the position of the head in the working tape, $a_4 = C$ (1100b in base 16) represents the content of the working tape, and $(a_5,\ldots,a_9) = (0,0,1,0,0)$ represents the current state.
-Then $\bar{a} = (1,A,2,C,0,0,1,0,0)$ will represent $s$.
-
-The problem that arises from this representation, is that to describe a transition in $M$ we need to read an arbitrary character in the working tape, and therefore we have to obtain the $a_3$-th bit in $a_4$.
-Furthermore, to represent the following configuration, we need compute $a_4$ with the $a_3$-th bit flipped.
-This is generally not possible to describe with an $\fo$ formula.
-Consider the following procedure that receives $x = a_4$ and $i = a_3$.
-
-\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 1 replace it by 0 and return the result}
-	\label{switch1to0}
-	\begin{algorithmic}
-		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
-		\While{$j > 0$}
-		\State $v \gets 0$
-		\While{$u > 1$}
-		\State $u \gets u-2,\; v \gets v+1$
-		\EndWhile
-		\State $u\gets v,\; j \gets j-1$
-		\EndWhile
-		\While{$u > 1$}
-		\State $u \gets u-2$
-		\EndWhile
-		\State $\textbf{assert } u = 1$ \Comment{If $u \neq 1$ simply stop}	
-		\State $y \gets 1$ \Comment{Compute $2^i$ and store it in $y$}
-		\While{$i > 0$}
-		\State $z \gets 0$
-		\While{$y > 0$}
-		\State $z \gets z+2,\; y \gets y-1$
-		\EndWhile
-		\State $i \gets i-1,\; y \gets z$
-		\EndWhile
-		\While{$y > 0$} \Comment{Substract $y$ from $x$}
-		\State $x \gets x-1,\; y \gets y-1$
-		\EndWhile
-		\State \Return $x$.
-	\end{algorithmic}
-\end{algorithm}	
-Each of the instructions can be expressed with $\fo$, so our strategy is to use the $\pth$ operator to simulate the algorithm and then we can describe a transition using the processed value of $a_4$.
-Note that we need to describe three more procedures to simulate the transitions for $0\to 0$, $0 \to 1$ and $1\to 1$.
-
-We will now describe how to simulate both the procedure and the transition.
-A procedure tuple $\bar{p} = (a_1,\ldots,a_{3+c+\ell},b_1,b_2,c_1,c_2,c_3,d_1,\ldots,d_{5c+2})$ represents the current configuration of $M(w)$ in $a_1,\ldots,a_{2+c+\ell}$, the values that will be read and written in the working tape in $b_1,b_2$, the instruction pointer in $c_1,c_2,c_3$ and the values stored in memory in $d_1,\ldots,d_{10c+2}$.
-In detail:
-\begin{enumerate}
-	\item $a_1,a_2$ and $a_3$ represent the position of the head in the input tape and the working tape, respectively, $a_4,\ldots,a_{3+c}$ represent the content of the working tape and $a_{4+c},\ldots,a_{3+c+\ell}$ represent the current state in the current configuration that is being processed.
-	\item $b_1$ and $b_2$ are equal to the value that is being read in the working tape and the value that will be written in the working tape respectively.
-	\item $c_1,c_2,c_2$ represent the instruction pointer in the procedure. Only 8 different instructions are needed in the simulation.
-	\item Each value in memory of $x,y,z,u,v$ need $c$ elements to represent them and $i,j$ need only one. We map $(d_1\ldots,d_{c}) \to x$, $(d_{c+1}\ldots,d_{2c}) \to y$,
-	$(d_{2c+1}\ldots,d_{3c}) \to z$, $(d_{3c+1}\ldots,d_{4c}) \to u$,
-	$(d_{4c+1}\ldots,d_{5c}) \to v$, $d_{5c+1} \to i$ and $d_{5c+2}\to j$.
-\end{enumerate}
-For each transition $\delta \in \Delta \subseteq Q \times \{0,1\} \times \{0,1\} \times Q \times \{-1,=,+1\} \times \{0,1\} \times \{-1,=,+1\}$ we define a formula $\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$, where $\bar{x} = (x_1,\ldots,x_{3+c+\ell})$, $\bar{s} = (s_1,s_2)$, $\bar{w} = (w_1,w_2,w_3)$, $\bar{u} = (u_1,\ldots,u_{5c+2})$, $\bar{y} = (y_1,\ldots,y_{3+c+\ell})$, $\bar{t} = (t_1,t_2)$, $\bar{z} = (z_1,z_2,z_3)$ and $\bar{v} = (v_1,\ldots,v_{5c+2})$.
-The tuples $\bar{x}$ and $\bar{y}$ represent the current and next configuration of $M$ respectively, $\bar{s}$ and $\bar{y}$ indicate which algorithm is being followed, $\bar{w}$ and $\bar{z}$ represent the current and next instruction of the algorithm, $\bar{u}$ and $\bar{v}$ represent the current and next values in memory.
-We will describe the formula part by part.
-Suppose $\delta = (q_i,a,1,q_j,op_1,0,op_2)$, so we have to simulate Algorithm \ref{switch1to0}.
-
-First we define some auxiliary formulas:
-\begin{align*}
-\varphi_0(x) &= \neg\exists y(y < x), \\
-\varphi_{+1}(x,y) &= x < y \wedge \neg\exists z(x < z \wedge z < y), \\
-\varphi_{-1}(x,y) &= \varphi_{+1}(y,x), \\
-\varphi_1(x) &= \exists y(\varphi_0(y) \wedge \varphi_{+1}(y,x)), \\
-\text{for $i,j \in\{0,1\}$, }	\varphi_{i,j}(x,y) &= \varphi_i(x) \wedge \varphi_j(y), \\
-\text{for each $k \leq 7$, }\varphi^b_k(x_1,x_2,x_3) &= \varphi_{a_1}(x_1) \wedge \varphi_{a_2}(x_2) \wedge \varphi_{a_3}(x_3)\text{, where $a_1a_2a_3$ is the value of $k$ in binary}, \\
-\mu_{<}(x_1,\ldots,x_c,y_1,\ldots,y_c) &= \bigvee_{i = 1}^c ((\bigwedge_{j = 1}^{i-1} x_j = y_j) \wedge x_i < y_i), \\
-\mu_{+1}(\bar{x},\bar{y}) &= \mu_{<}(\bar{x},\bar{y}) \wedge \neg\exists \bar{z}(\mu_{<}(\bar{x},\bar{z}) \wedge \mu_{<}(\bar{z},\bar{y})), \\
-\mu_{+2}(\bar{x},\bar{y}) &= \exists\bar{z}(\mu_{+1}(\bar{x},\bar{z}) \wedge \mu_{+1}(\bar{z},\bar{y})),\\
-\mu_{-1}(\bar{x},\bar{y}) &= \varphi_{+1}(\bar{y},\bar{x}),\\
-\mu_{-2}(\bar{x},\bar{y}) &= \varphi_{+2}(\bar{y},\bar{x}),\\
-\mu_{0}(\bar{x}) &= \neg\exists\bar{y}(\mu_{<}(\bar{y},\bar{x})),\\
-\mu_{1}(\bar{x}) &= \exists\bar{y}(\mu_{0}(\bar{y})\wedge\mu_{+1}(\bar{y},\bar{x})),\\
-\varphi^2_{<}(x_1,x_2,y_1,y_2) &= x_1 < y_1 \vee (x_1 = y_1 \wedge x_2 < y_2),\\
-\varphi^2_{+1}(x_1,x_2,y_1,y_2), \varphi^2_{-1}(x_1,x_2,y_1,y_2) &\text{ defined analogously to $\mu_{+1}$ and $\mu_{-1}$},\\
-\text{for each } q_i\in Q, \varphi^q_i(x_1,\ldots,x_{\ell}) &= \bigwedge_{\substack{j = 1 \\ j \neq i}}^{\ell} \varphi_0(x_j) \wedge \varphi_1(x_i), \\
-\varphi^E_0(x_1,x_2) &= \neg E(x_1,x_2),\\		\varphi^E_1(x_1,x_2) &= E(x_1,x_2),\\
-\end{align*}
-
-We start from instruction 0, which means that the procedure has not started yet and every value in the tuple is 0 except for the configuration values.
-It also initializes all the values in the tuple to 0 except for $x,u,i,j$.
-\begin{multline*}
-\varphi^{0,1}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{0,0}(s_1,s_2)\wedge\varphi^b_0(\bar{w}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_1(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c v_i = x_{3+i} \wedge \bigwedge_{i = c+1}^{2c} \varphi_0(v_i) \wedge \bigwedge_{i = 2c+1}^{3c} \varphi_0(v_i) \wedge \bigwedge_{i = 1}^c v_{3c+i} = x_{3+i} \wedge \bigwedge_{i = 4c+1}^{5c} \varphi_0(v_i) \wedge v_{5c+1} = x_3 \wedge v_{5c+2} = x_3.
-\end{multline*}
-Instruction 1 which checks whether the value of $j$ ($d_{5c+2}$ in the tuple) is more than 0 or not, and then proceeds to instruction 2 or 3 on each case.
-\begin{multline*}
-\varphi^{1,2}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_1(\bar{w}) \wedge \neg \varphi_0(u_{5c+2}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_2(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i, \\
-\varphi^{1,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_1(\bar{w}) \wedge \varphi_0(u_{5c+2}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_3(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
-\end{multline*}
-Instruction 2 checks the value of $u$ ($d_{3c+1},\ldots,d_{4c}$ in the tuple).
-If it is $> 1$ then it substracts 2 from $u$ and adds 1 to $v$ ($d_{4c+1},\ldots,d_{5c}$ in the tuple), then repeats instruction 2.
-If it is equal to 0 or 1, then moves the value of $v$ to $u$, substracts 1 from $j$ and goes back to instruction 1.
-\begin{multline*}
-\varphi^{2,2}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_2(\bar{w}) \wedge \neg \mu_0(u_{3c+1},\ldots,u_{4c}) \wedge \neg \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_2(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
-\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-\mu_{-2}(u_{3c+1},\ldots,u_{4c},v_{3c+1},\ldots,v_{4c})\, \wedge \\
-\mu_{+1}(u_{4c+1},\ldots,u_{5c},v_{4c+1},\ldots,v_{5c}) \wedge u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}. \\
-\varphi^{2,1}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_2(\bar{w}) \wedge ( \mu_0(u_{3c+1},\ldots,u_{4c}) \vee \mu_1(u_{3c+1},\ldots,u_{4c})) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_1(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
-\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-\bigwedge_{i = 3c+1}^{4c} u_{c+i} = v_i \wedge
-\mu_0(v_{4c+1},\ldots,v_{5c}) \wedge
-u_{5c+1} = v_{5c+1} \wedge \varphi_{-1}(u_{5c+2},v_{5c+2}).	
-\end{multline*}
-Instruction 3 calculates the value of $u \mod 2$, that is, it repeats instruction 3 until the value of $u$ is equal to 0 or 1.
-On each iteration, it substracts 2 from $u$.
-Moreover, if the value of $u$ at the end of the iterations is not 1 then there is no step defined.
-\begin{multline*}
-\varphi^{3,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_3(\bar{w}) \wedge \neg \mu_0(u_{3c+1},\ldots,u_{4c}) \wedge \neg \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_3(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
-\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-\mu_{-2}(u_{3c+1},\ldots,u_{4c},v_{3c+1},\ldots,v_{4c}) \wedge
-\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
-u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}, \\
-\varphi^{3,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_3(\bar{w}) \wedge \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_4(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\mu_1(v_{c+1},\ldots,v_{2c}) \wedge
-\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
-\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
-u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}
-\end{multline*}
-Instruction 4 checks the value of $i$ ($d_{5c+1}$ in the tuple.) If it is not 0 then goes to instruction 5 and if is 0 then goes to instruction 6.
-Moreover it initializes the value of $z$ ($d_{2c+1},\ldots,d_{3c}$ in the tuple) to 0 (which was 0 all along.)
-\begin{multline*}
-\varphi^{4,5}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_4(\bar{w}) \wedge \neg \varphi_0(u_{5c+1}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_5(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i, \\
-\varphi^{4,6}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_4(\bar{w}) \wedge \varphi_0(u_{5c+1}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_6(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
-\end{multline*}
-Instruction 5 checks the value of $y$ ($d_{c+1},\ldots,d_{2c}$ in the tuple.) If it is more than 0 then it adds 2 to $z$ and substracts 1 from $y$, then repeats instruction 2.
-If it is not, then copies the value of $z$ to $y$ and subtracts 1 from $i$ and returns to instruction 4.
-\begin{multline*}
-\varphi^{5,5}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_5(\bar{w}) \wedge \neg \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_5(\bar{z}) \, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\mu_{-1}(u_{c+1},\ldots,u_{2c},v_{c+1},\ldots,v_{2c}) \wedge
-\mu_{+2}(u_{2c+1},\ldots,u_{3c},v_{2c+1},\ldots,v_{3c})\, \wedge \\
-\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
-\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
-u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}, \\
-\varphi^{5,4}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_5(\bar{w}) \wedge \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_4(\bar{z}) \, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\bigwedge_{i = c+1}^{2c} u_{c+i} = v_i \wedge
-\mu_0(v_{2c+1},\ldots,v_{3c}) \wedge
-\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
-\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
-\varphi_{-1}(u_{5c+1},v_{5c+1}) \wedge u_{5c+2} = v_{5c+2}
-\end{multline*}
-Instruction 6 checks the value of $y$. If it is more than 0, then subtracts 1 from $x$ and $y$ and repeats instruction 6. If it is not, then goes to instruction 7.
-\begin{multline*}
-\varphi^{6,6}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_6(\bar{w}) \wedge \neg \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_6(\bar{z}) \, \wedge \\
-\mu_{-1}(u_1,\ldots,u_c,v_i,\ldots,v_c) \wedge \mu_{-1}(u_{c+1},\ldots,u_{2c},v_{c+1},\ldots,v_{2c}) \wedge \bigwedge_{i = 3c+1}^{5c+2} u_i = v_i]\,\vee \\
-\varphi^{6,7}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_6(\bar{w}) \wedge \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_7(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
-\end{multline*}
-Instruction 7 stores the value of $x$ after the corresponding bit has been switched.
-Then we can define $\varphi_{\delta}$ which also simulates the actual transition.
-If $u$ equals 1, then copy what is stored in $x$ to $a_4,\ldots,a_{3+c}$, go from state $q_i$ to state $q_j$, and move the heads to their corresponding positions.
-\begin{multline*}
-\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 	[\bigwedge_{i = 1}^{3+c+\ell} x_i = y_i \wedge (\varphi^{0,1}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{1,2}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})\, \vee \\ \varphi^{1,3}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{2,2}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{2,1}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{3,3}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \\ \varphi^{3,4}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{4,5}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{4,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{5,5}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \\ \varphi^{5,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{6,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{6,7}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}))] \, \vee \\
-[\varphi^E_a(x_1,x_2) \wedge \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_7(\bar{w}) \wedge \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{0,0}(t_1,t_2) \wedge \varphi^b_0(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i \,\wedge \\
-\varphi^2_{op_1}(x_1,x_2,y_1,y_2) \wedge \varphi_{op_2}(x_3,y_3) \wedge \bigwedge_{i = 1}^c u_i = x_{3+i} \wedge \varphi^q_i(x_{4+c},\ldots,x_{3+c+\ell}) \wedge \varphi^q_j(y_{4+c},\ldots,y_{3+c+\ell})].
-\end{multline*}
-
-\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 0 return $x$} \label{switch0to0}
-	\begin{algorithmic}
-		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
-		\While{$j > 0$}
-		\State $v \gets 0$
-		\While{$u > 1$}
-		\State $u \gets u-2,\; v \gets v+1$
-		\EndWhile
-		\State $u\gets v,\; j \gets j-1$
-		\EndWhile
-		\While{$u > 1$}
-		\State $u \gets u-2$
-		\EndWhile
-		\State $\textbf{assert } u = 0$ \Comment{If $u \neq 0$ simply stop}	
-		\State \Return $x$.
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 0 replace it by 1 and return the result}
-	\label{switch0to1}
-	\begin{algorithmic}
-		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
-		\While{$j > 0$}
-		\State $v \gets 0$
-		\While{$u > 1$}
-		\State $u \gets u-2,\; v \gets v+1$
-		\EndWhile
-		\State $u\gets v,\; j \gets j-1$
-		\EndWhile
-		\While{$u > 1$}
-		\State $u \gets u-2$
-		\EndWhile
-		\State $\textbf{assert } u = 0$ \Comment{If $u \neq 0$ simply stop}	
-		\State $y \gets 1$ \Comment{Compute $2^i$ and store it in $y$}
-		\While{$i > 0$}
-		\State $z \gets 0$
-		\While{$y > 0$}
-		\State $z \gets z+2,\; y \gets y-1$
-		\EndWhile
-		\State $i \gets i-1,\; y \gets z$
-		\EndWhile
-		\While{$y > 0$} \Comment{Add $y$ to $x$}
-		\State $x \gets x+1,\; y \gets y-1$
-		\EndWhile
-		\State \Return $x$.
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 1 return $x$}
-	\label{switch1to1}
-	\begin{algorithmic}
-		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
-		\While{$j > 0$}
-		\State $v \gets 0$
-		\While{$u > 1$}
-		\State $u \gets u-2,\; v \gets v+1$
-		\EndWhile
-		\State $u\gets v,\; j \gets j-1$
-		\EndWhile
-		\While{$u > 1$}
-		\State $u \gets u-2$
-		\EndWhile
-		\State $\textbf{assert } u = 1$ \Comment{If $u \neq 0$ simply stop}	
-		\State \Return $x$.
-	\end{algorithmic}
-\end{algorithm}
-Note that we also need to specify that the program we are following is Algorithm \ref{switch1to0} so we store $1,0$ in $b_1,b_2$ all along the procedure.
-We describe the three other algorithms that compute the switches from $0\to 0$, $0\to 1$ and $1\to 1$ (Algorithms \ref{switch0to0}, \ref{switch0to1} and \ref{switch1to1}.)
-For the other three cases, where $\delta = (q_i,a,0,q_j,op_1,0,op_2)$, $\delta = (q_i,a,0,q_j,op_1,1,op_2)$ and $\delta = (q_i,a,1,q_j,op_1,1,op_2)$, $\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$ is defined analogously.
-Then, $\varphi(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$ is defined as:
-$$
-\varphi(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \bigvee_{\delta \in \Delta} \varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}).
-$$
-Finally we define $\varphi_I$ and $\varphi_F$:
-\begin{align*}
-\varphi_I(\bar{x},\bar{s},\bar{w},\bar{u}) &= \varphi_0(x_1) \wedge \varphi_0(x_2) \wedge \varphi_0(x_3) \wedge \bigwedge_{i = 4}^{3+c}\varphi_0(x_i) \wedge \varphi_1(x_{4+c}) \wedge \bigwedge_{i = 5+c}^{3+c+\ell} \varphi_0(x_i)\wedge \varphi_{0,0}(s_1,s_2)\ \wedge \varphi^b_0(\bar{w}) \wedge \bigwedge_{i = 1}^{5c+2}\varphi_0(u_i).
-\\
-\varphi_F(\bar{x},\bar{s},\bar{w},\bar{u}) &= \varphi^q_{\ell}(x_{4+c},\ldots,x_{3+c+\ell}) \wedge \varphi_{0,0}(s_1,s_2)\ \wedge \varphi^b_0(\bar{w}) \wedge \bigwedge_{i = 1}^{5c+2}\varphi_0(u_i),
-\end{align*}
-and then $\sem{\alpha}(\A) = \sem{\sa{\bar{x}}\sa{\bar{y}}([\pth \varphi(\bar{u},\bar{v})](\bar{x},\bar{y})\cdot \varphi_I(\bar{x})\cdot\varphi_F(\bar{y}))}(\A) = \acc_M(\A)$.
-
-\textbf{Proof of Theorem \ref{tqso-fo-cap}}
-
-\textbf{Proof of Theorem \ref{rqfo-fo-cap}}
+and adding these values gives $f(\enc(\A))$. Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$. \qed

### 2016-12-28 LICS/appendix-notation.tex

--- 2016-12-26 LICS/appendix.tex
@@
-%!TEX root = main.tex
-
-\subsection{Notation for the appendix}
-
 For a given signature $\R$, we define $\ostr[\R]^*$ as $$\ostr[\R]^* = \{(\A,v,V) \mid \A\in\ostr[\R]\text{, $v$ ($V$) is a first-order (second-order) assignment for $\A$}  \}.$$
@@
 We also denote $\length(\bar{x}) = m$ ($\length(\bar{X}) = m$).
-\subsection{Proof of Theorem \ref{no-mult}}
-
-Let $\{+, \cdot, \sa{}, \pa{}\}$ be the set of {\it algebraic operators} in $\qso$.
-We will prove by induction over the number of algebraic operators in a $\qso$-formula $\alpha$ that there exists some $\qso$-formula $\alpha'$ that respects the stated condition.
-Let $\text{op}(\alpha)$ be the number of algebraic operators in $\alpha$.
-The case $\text{op}(\alpha) = 0$ is trivial.
-Let $n\in\nat$ be such that for each $\qso$-formula $\beta$ such that $\text{op}(\beta) < n$ there is some $\beta'$ that respects the condition.
-Let $\alpha$ be a $\qso$-formula such that $\text{op}(\alpha) = n$.
-Note that if $\alpha$ is of the form $\alpha = (\beta_1 \add \beta_2)$, $\alpha = \sa{x} \beta$, $\alpha = \pa{x} \beta$, $\alpha = \sa{X} \beta$ or $\alpha = \pa{X} \beta$ the proof follows directly.
-Therefore we assume that $\alpha = (\beta_1 \mult \beta_2)$.
-We separate the proof in two main cases:
-\begin{enumerate}
-	\item At least one of $\beta_1, \beta_2$ is not of the form $\pa{x} \gamma$, $\pa{X} \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. Without loss of generality, let $\beta_1$ be such one.
-	\begin{enumerate}
-		\item $\beta_1 = s \in \nat$.
-Then we use $\alpha' = \beta_2 \add \cdots \add \beta_2$, where $\beta_2$ is repeated $s$ times, and if $s = 0$, take $\alpha' = \perp$.
-Since each of the formulas paired by $\add$ has less than $n$ algebraic operators, the statement holds.
-		\item $\beta_1 = (\gamma_1 \add \gamma_2)$.
-Then, we use $\alpha' = (\gamma_1\mult\beta_2 + \gamma_2\mult\beta_2)$ and each of the formulas paired by $\add$ have less than $n$ algebraic operators so the statement holds.
-		\item $\beta_1 = \sa{x} \gamma$. Then, we use $\alpha' = \sa{x} (\gamma\mult\beta_2)$ and as in the previous case the statement holds.
-		\item $\beta_1 = \sa{X} \gamma$, which we treat exactly as the previous one.
-	\end{enumerate}
-	\item Both of $\beta_1$ and $\beta_2$ are of the form $\pa{x} \gamma$, $\pa{X} \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. This case also separates in further cases.
-	\begin{enumerate}
-		\item $\beta_1 = \pa{x} \gamma_1(x)$ and $\beta_2 = \pa{y} \gamma_2(y)$. Then we use $\alpha' = \pa{x}(\gamma_1(x)\mult\gamma_2(x))$, and the statement holds.
-		\item $\beta_1 = \pa{X} \gamma_1(X)$ and $\beta_2 = \pa{Y} \gamma_2(Y)$. [something about arities].
-		\item $\beta_1 = \pa{X} \gamma_1(X)$ and $\beta_2 = \pa{x} \gamma_2(x)$ (w.l.o.g.). [use singleton predicates].
-		\item $\beta_1 = \pa{x} \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). Then, we use $\alpha' = \pa{x} (\gamma \cdot \varphi)$, and the statement holds.
-		\item $\beta_1 = \pa{X} \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). We treat this case exactly as the previous one.
-		\item $\beta_1 = \varphi_1$ and $\beta_2 = \varphi_2$ both $\so$-formulas. Then, we use $\alpha' = (\varphi_1 \wedge \varphi_2)$ and the statement holds.
-	\end{enumerate}
-\end{enumerate}
-
-\subsection{Proof of Theorem \ref{captfp}}
-
-\martin{no revisar, es un copy-paste de arriba}
-
-For the first condition, let $\alpha\in\qfo(\lfp)$ over some signature $\R$.
-Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{a_1,\ldots,a_n\}$.
-Replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\Sigma x \beta(x)$ is replaced by $(\beta(a_1)+\cdots+\beta(a_n))$ and $\Pi x \beta(x)$ is replaced by $(\beta(a_1)\cdot\,\cdots\,\cdot\beta(a_n))$.
-Then each sub-formula $\varphi\in\lfp$ in $\alpha$ is replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size which is evaluated and lastly given as output.
-Note that $f\in\fp$ and $f(\enc(\A)) = \alpha(\enc(\A))$.
-
-For the second condition, let $f\in \fp$ defined over some signature $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
-Let $\Phi(x_0,\ldots,x_{\ell-1})$ be a formula that represents the following.
-Let $\A$ be an ordered structure over $\R$ with domain $A$.
-Consider a linear order over $A^{\ell}$ given by $\bar{a} < \bar{b}$, for any $\bar{a},\bar{b}\in A^{\ell}$.
-Let $w$ be a string of size $n^{\ell}$ the represents $f(\enc(\A))$ in binary.
-Let $(a_1,\ldots,a_{\ell})\in A^{\ell}$ be the $m$-th tuple in the given order (starting from 0).
-$\A\models\Phi(a_1,\ldots,a_{\ell})$ if an only if the $m$-th bit in $w$ is 1, from least to most significant.
-Since this condition can be computed in polynomial time, and $\lfp$ captures $\ptime$, this formula can be expressed in $\lfp$.
-We use
-$$
-\alpha = \sa{x_1} \cdots \sa{x_{\ell}} \Phi(x_1,\ldots,x_{\ell})\cdot\varphi(x_1,\ldots,x_{\ell}),
-$$
-where $\varphi(x_1,\ldots,x_{\ell}) := \pa{y_1}\cdots\pa{y_{\ell}}((y_1,\ldots,y_{\ell}) < (x_1,\ldots,x_{\ell}) \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
-Adding this values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\alpha(\enc(\A)) = f(\enc(\A))$.
-\qed
-
-\subsection{Proof of Theorem \ref{qfo-pfp-cap}}
-
-For condition (2) in \ref{cap-def}, let $\alpha\in\qfo(\pfp)$ over some $\R$.
-Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{a_1,\ldots,a_n\}$.
-Replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\Sigma x \beta(x)$ is replaced by $(\beta(a_1)+\cdots+\beta(a_n))$ and $\Pi x \beta(x)$ is replaced by $(\beta(a_1)\cdot\,\cdots\,\cdot\beta(a_n))$.
-Then each sub-formula $\varphi\in\pfp$ in $\alpha$ is replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size which is evaluated and lastly given as output.
-Note that $f\in\nfpspace$ and $f(\enc(\A)) = \alpha(\enc(\A))$.
-
-For condition (2) in \ref{cap-def}, let $f\in \nfpspace$ defined over some $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
-Let $\Phi(x_0,\ldots,x_{\ell-1})$ be a $\pfp$ formula that represents the following.
-For each $\A \in \ostr[\R]$, for convenience, suppose its domain is $A = \{0,\ldots,n-1\}$ with the usual order.
-Let $y$ be a string of size $n^{\ell}$ that represents the value of $f(\enc(\A))$ in binary.
-Then, for each $\ell$-tuple $(a_0,\ldots,a_{\ell-1}) \in A^{\ell}$, it holds that $\A \models \Phi(a_0,\ldots,a_{\ell-1})$ if and only if the $(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)$-th bit of $y$, from least to most significant, is 1.
-We use
-$$
-\alpha = \Sigma x_0 \cdots \Sigma x_{\ell-1} \Phi(x_0,\ldots,x_{\ell-1})\cdot\varphi_{\ell-1}(x_{\ell-1})\cdot\,\cdots\,\cdot\varphi_0(x_0).
-$$
-where $\varphi_i(x) = \Pi y[(y < x)\mapsto\Pi z_1\cdots\Pi z_i\,2]$ for $i > 0$ and $\varphi_0(x) = \Pi y[(y < x)\mapsto 2]$.
-Let $\A$ the previously defined structure.
-Note that for each $a\in A$, then $\sem{\varphi_i(x)}(\A,a) = 2^{n^i a}$.
-Therefore, for each $(a_0,\ldots,a_{\ell-1})\in A^\ell$, we have that
-$$
-\sem{\Phi(x_0,\ldots,x_{\ell-1})\cdot\varphi_{\ell-1}(x_{\ell-1})\,\cdots\,\varphi_0(x_0)}(\A,a_0,\ldots,a_{\ell-1}) = 
-\begin{cases}
-2^{(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)} &\text{if } (\A,a_0,\ldots,a_{\ell-1})\models\Phi \\
-0 &\text{otherwise.}
-\end{cases}
-$$
-and adding these values gives $f(\enc(\A))$. Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\alpha(\enc(\A)) = f(\enc(\A))$. \qed
-
-\subsection{Proof of Theorem \ref{eqso-fo-cap}}
-
-For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\fo$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial time, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.%%If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Let $\alpha$ be a formula in $\eqso(\fo)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shp$ function over $\R$ and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
-
-For condition (2) in \ref{cap-def}, note that Saluja et al.
-\cite{DBLP:journals/jcss/SalujaST95} proved that $\shp = \#\fo$.
-We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso(\fo)$ be its corresponding formula.
-Since this formula is also in $\eqso(\fo)$, the condition holds.
-\qed
-
-\subsection{Proof of Theorem \ref{eqso-eso-cap}}
-
-For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic outputs for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-On each of the steps the algorithm adds the current configuration to the output tape.
-If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time without modifying the output tape, and accepts if and only if the condition holds true.
-If $\alpha$ is a constant $s$, then the algorithm produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-%% If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\spp$ function over $\R$ and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
-
-For condition (2) in \ref{cap-def}, note that Compton and Graedel \cite{DBLP:journals/jcss/ComptonG96} proved that $\spp = \#\eso$.
-We also have that $\sqso(\eso)$ captures $\#\eso$ over ordered structures so for each $f\in \spp$ let $\alpha \in \sqso(\eso)$ be its corresponding formula.
-Since this formula is also in $\eqso(\eso)$, the condition holds.
-\qed
-
-\subsection{Proof of Theorem \ref{qso-pfp-cap}}
-
-For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\pfp$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space \cite{DBLP:conf/lics/AbiteboulV89}\martin{no estoy seguro de esta cita}, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-If $\alpha = \pa{X}\beta$, it simulates $M_{\beta}$ on input $(\A,v,V[B/X])$ consecutively for each $B\in A^{arity(X)}$.
-This covers all possible cases for $\alpha$, and each of these steps can be computed in polynomial space.
-Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$ \cite{DBLP:journals/siamcomp/Ladner89}, and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
-
-For the second condition, let $f\in \fpspace$ defined over some $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq 2^{n^\ell}$, where $\A$ has a domain of size $n$.
-Let $\Phi(X)$ be a $\pfp$ formula that represents the following.
-Consider an order over the subsets of $A^{\ell}$ given by the formula $$\varphi_{<}(X,Y) = \exists\bar{u}\big[\neg X(\bar{u})\wedge Y(\bar{u})\wedge \forall\bar{v}\big(
-\bar{u}<\bar{v}\to(X(\bar{u})\iff Y(\bar{v}))\big)\big].$$
-
-Let $B$ be the $p$-th subset with respect to this order and let $w$ be a string of size $2^{n^{\ell}}$ which represents the value of $f(\enc(\A))$ in binary.
-Then $\A\models\Phi(B)$ if and only if the $p$-th bit in $w$ is 1, from least to most significant.
-Since this condition can be verified in polynomial space, and $\pfp$ captures $\pspace$, then $\Phi$ can be expressed in $\pfp$.
-We define
-$\alpha := \sa{X}(\Phi(X)\mult\varphi(X))$, where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
-Note that for each $B\subseteq A^{\ell}$ such that $B$ is the $m$-th element in the given order, $\sem{\varphi(B)}(\A) = 2^m$.
-Therefore, $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-\qed
-
-\subsection{Proof of Theorem \ref{eqso-sigma-zero-in-fp}}
-
-First, for a given signature $\R$ and logic fragment $\LL$ we define a {\em valid form in $\LL$} (or just {\em valid form}).
-A formula $\alpha$ is in valid form in $\LL$ if (1) it is equal to $\varphi$, $1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}1$ for some $\LL$ formula $\varphi$ and some tuple of first-order (second-order) variables $\bar{x}$ ($\bar{X}$), or (2) if it is equal to $(\beta\add\gamma)$ where $\beta$ and $\gamma$ are formulas in valid form.
-Given a $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$ we will recursively construct a formula $\tau(\alpha)$ in valid form such that $\alpha(\enc(\A)) = \tau(\alpha)(\enc(\A))$ for each $\A\in\ostr[\R]$.
-
-First, each natural constant $s$ in $\alpha$ is replaced by $(1\add 1\add\cdots\add 1)$ ($s$ times).
-If $\alpha = 1$ or $\alpha = \varphi \in \loge{0}$, then $\tau(\alpha) = \alpha$.
-We assume that every sub-formula of $\alpha$ is in valid form.
-If $\alpha = (\beta \add \gamma)$ then $\tau(\alpha) = \alpha$.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ this case separates in more cases: If $\alpha = \sa{x}\beta$ and $\beta$ is in one of the valid forms in (1), then $\tau(\alpha) = \alpha$.
-If $\alpha = \sa{X}\beta$ and $\beta$ is in the valid form $\sa{\bar{x}}\gamma$ then $\tau(\alpha) = \sa{\bar{x}}\sa{X}\gamma$; if $\beta$ it is in one of the remaining valid forms in (1), then $\tau(\alpha) = \alpha$.
-If $\alpha = \sa{x}\beta$ ($\alpha = \sa{X}\beta$) and $\beta = (\gamma_1\add\gamma_2)$, then $\tau(\alpha) = \tau(\sa{x}\gamma_1) \add \tau(\sa{x}\gamma_2)$ ($\tau(\alpha) = \tau(\sa{X}\gamma_1) \add \tau(\sa{X}\gamma_2)$).
-This covers all possible cases for $\alpha$.
-Furthermore, suppose that $\LL$ is closed under conjunction and suppose $\alpha = (\beta\mult\gamma)$.
-If $\beta = (\delta_1\add\delta_2)$ (wlog), then $\tau(\alpha) = \tau(\delta_1\mult\gamma)\add\tau(\delta_2\mult\gamma)$.
-If $\beta = \sa{\bar{x}}\delta$, then $\tau(\alpha) = \tau(\sa{\bar{x}}\tau(\delta\mult\gamma))$.
-If $\beta = 1$, then $\tau(\alpha) = \gamma$.
-If $\beta = \varphi$ and $\gamma = \psi$, then $\tau(\alpha) = \varphi\wedge\psi$.
-Let $\alpha$ be a $\eqso(\loge{0})$ formula over a signature $\R$.
-Consider the following procedure that receives $\enc(\A)$ as input, where $\A\in\ostr[\R]$ with domain $A$.
-Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_m$.
-For each $\beta_i$, if $\beta_i = 1$ or $\beta_i = \varphi$, then $\sem{\beta_i}(\A)$ can be computed in polynomial time.
-If $\beta_i = \sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}$, if $\beta_i = \sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = 2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, if $\beta_i = \sa{\bar{X}}\sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, all of which can be computed in polynomial time.
-Otherwise, $\beta_i$ is $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, and then $\sem{\beta_i}(\A) = f_{\varphi}(\enc(\A))$, and $f_{\varphi}\in \#\loge{0}$ which implies that $f_{\varphi}(\enc(\A))$ can be computed in polynomial time \cite{DBLP:journals/jcss/SalujaST95}.
-
-\subsection{Proof of Theorem \ref{eqso-sigma-one-in-eqso-pi-one}}
-
-Let $\alpha$ be a formula in $\eqso(\loge{1})$ over a signature $\R$.
-Let $\tau(\alpha) = \beta_1\add\beta_2\add\cdots\add\beta_m$ defined as in the proof of Theorem \ref{eqso-sigma-zero-in-fp}.
-This is, each $\beta_i$ is of the form $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}\sa{\bar{x}}1$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, where $\varphi$ is a formula in $\loge{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables.
-For each $\beta_i$ we will construct a $\gamma_i$ such that $\gamma_i \in \eqso(\logu{1})$ and $\sem{\beta_i}(\A) = \sem{\gamma_i}(\A)$ for each $\A\in\ostr[\R]$.
-Since the forms $1$, $\sa{\bar{x}}1$, $\sa{\bar{X}}1$ and $\sa{\bar{X}}\sa{\bar{x}}1$ are already valid in $\eqso(\logu{1})$, we suppose $\beta_i$ is in one of the remaining forms.
-Without loss of generality, let $\beta_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y},\bar{X})$ where $\varphi(\bar{x},\bar{y},\bar{X})$ is a quantifier-free formula.
-Similarly as in the proof by Saluja et al.
-\cite{DBLP:journals/jcss/SalujaST95}, we define $\gamma_i = \sa{\bar{x}}\sa{\bar{y}}\sa{\bar{X}}\varphi(\bar{x},\bar{y},\bar{X}) \wedge \forall\bar{z}(\varphi(\bar{x},\bar{z},\bar{X})\to(\bar{z} < \bar{y} \vee \bar{z} = \bar{y}))$ that fulfils the desired condition.
-The theorem holds true by using $\beta = \gamma_1\add\cdots\add\gamma_m$ which is in $\eqso(\logu{1})$
-
-%\subsection{Proof of Theorem \ref{one-sigma-zero}}
-%
-%Towards a contradiction, suppose that there exists a constant function $f(\A) = s > 0$ which is expressible in $\E{0}$.
-This is, there is a formula $\psi(x,X)$ such that $f_{\psi(x,X)}(\A) = \vert \{ \langle e,P\rangle \mid \A\models\psi(e,P) \} \vert = s$ for each $\A\in\ostr$.
-It can be seen that if $X$ contains more than zero variables, then the function $f_{\psi(x,X)}$ grows exponentially.
-We have now that $\psi(x)$ contains only first order variables.
-Consider some $\A \in \ostr$ of size $n$ and let $a_1,\ldots,a_s\in \A$ be the tuples that satisfy $\A \models \psi(a_i)$.
-Now consider a structure $\A'$ that is obtained by duplicating $\A$.
-This is, $\A'$ has size $2n$, its first half is isomorphic to its second half, and also to $\A$.
-We have that there is at least $2s$ tuples $a_i,b_i\in \A'$, $s$ on each half, such that $\A'\models\psi(a_i)$ and $\A'\models\psi(b_i)$, which leads to a contradiction.
-
-%\textbf{Proof of Theorem \ref{mult-sigma-zero}}
-%
-%Let $\alpha \in \sqso(\loge{0})$ with binary product.
-This is, $\alpha$ is defined by the grammar $\alpha := \sa{x}\varphi \ \mid \ \sa{X}\varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \ (\alpha\mult\alpha)$, where $\varphi \in \loge{0}$.
-For convenience, we add the expressions $\sa{x}(\varphi\mult\varphi)$ and $\sa{X}(\varphi\mult\varphi)$ to the grammar.
-Let $k$ be the number of algebraic operators in $\alpha$.
-We will show by induction on $k$ that $\alpha$ is expressible without multiplication.
-If $k = 1$, then the possible cases are $\alpha = \sa{x}\varphi$, where $\varphi\in\loge{0}$, which is already product-free.
-If $k = 2$, then the only that is not two nested summations is $\alpha = \sa{x}(\varphi\mult\psi)$, which can be expressed as $\alpha = \sa{x}(\varphi\wedge\psi)$.
-Suppose that the hypothesis holds for $k = 1$ through $n-1$.
-Let $\alpha$ have $n$ algebraic operators.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ where $\beta$ is an algebraic expression, then $\beta$ has $n-1$ algebraic operators and by hypothesis it can be expressed as product-free.
-If $\alpha = (\beta \mult \gamma)$ there are two cases.
-If at least one between $\beta$ and $\gamma$ has a product, then by hypothesis it can be expressed as product-free.
-Moreover, the new expression for $\alpha$ has strictly less than $n$ algebraic operators so by hypothesis it can be expressed as product-free.
-If there are no products in $\beta$ or $\gamma$, then let $\beta = \sa{x}\delta(x)$.
-We have that $\alpha$ can be expressed as $\sa{y}(\delta(y)\mult\gamma)$, where $y$ is a first-order variable that does not appear in $\beta$ or $\gamma$.
-Since $\delta(y)\mult\gamma$ has $n-1$ algebraic operators, it can be expressed as product-free, and then $\alpha$ can be expressed as product-free.
-%
-%\textbf{Proof of Theorem \ref{fo-prod-sigma-zero}}
-%
-%The resulting grammar is $\alpha := \sa{x}\varphi \ \mid \ \sa{X} \varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \pa{x}\alpha$.
-Consider a more restricted grammar that replaces $\pa{x}\alpha$ by $\pa{x}\sa{y}\varphi$.
-Let $\alpha = \pa{x}\sa{y}(x = y)$.
-This formula produces a constant function equal to 1, which as shown in Theorem \ref{one-sigma-zero} is not expressible in $\E{0}$.
-%
-%\textbf{Proof of Theorem \ref{fp1}}
-%
-%Let $f \in \XE{0}$, and let $\varphi(\X,\x)$ be and {\sc FO}-extended quantifier-free $\L$-formula such that:
-%\begin{eqnarray*}
-%	f(\A) &=& \vert \{\langle \P,\e  \rangle \mid \A \models \varphi(\P,\e) \} \vert
-%\end{eqnarray*}
-%for each $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^m$ and $\P = (P_1,\ldots,P_q)$ is a tuple of predicates.
-We will now show that computing $f(\A)$ can be done in polynomial time.
-%
-%For each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$, let $R_{\beta}$ be a predicate of arity $m$.
-Let $\A^\prime = \langle A, \S^{\A}, R_{\beta}^{\A^\prime}, \leq^{\A} \rangle \in \str$, where $R_{\beta}^{\A^\prime} = \{\d \mid \A \models \beta(\d)\}$.
-Note that each $\beta(\x)$ is fixed in $\varphi(\X,\x)$, and for each $\d \in A^m$, checking whether $\A \models \beta(\d)$ can be done in polynomial time.
-Therefore, generating $R_{\beta}^{\A^\prime}$ can also be done in polynomial time.
-%
-%Let $\psi(\X,\x)$ be obtained by replacing each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$ by $R_{\beta}(\x)$.
-Also, let $g = f_{\psi(\X,\x)}$.
-Note that for each tuple of predicates $\P$ and each $\e \in A^m$, $\A \models \varphi(\P,\e)$ if and only if $\A^\prime \models \psi(\P,\e)$, and so, $g(\A^\prime) = f(\A)$.
-But, $\psi(\X,\x)$ is a quantifier-free $\L$-formula, and therefore, $g \in \#\Sigma_0$.
-Since it is shown in \cite{DBLP:journals/jcss/SalujaST95} that $\#\Sigma_0 \subseteq$ {\sc FP}, we conclude that $f(\A)$ can be evaluated in polynomial time.
-
-\textbf{Proof of Theorem \ref{decisionptime}}
-
-Let $f$ be a function in $\E{1}$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
-\begin{eqnarray*}
-	f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables.
-Let $\y = (\x,\z)$ and let $\psi(\X,\y) = \varphi(\x,\X,\z)$.
-Moreover, let $g = f_{\psi(\X,\y)}$.
-\begin{claim} \label{mono}
-	For each $\A \in \str$, $f(\A) > 0$ iff $g(\A) > 0$.
-\end{claim}
-\begin{proof}
-	($\Rightarrow$) Suppose $f(\A) > 0$.
-Let $\P$ and $\e$ be such that $\A \models \exists \x \, \varphi(\x,\P,\e)$.
-It follows that there is at least one $\d \in A^k$ such that $\A \models \varphi(\d,\P,\e) = \psi(\P,(\d,\e))$.
-Therefore, $g(\A) > 0$.
-
-	($\Leftarrow$) Suppose $g(\A) > 0$.
-Let $\Q$ and $\c = (\c_1,\c_2)$, where $\c_1$ and $\c_2$ have $k$ and $m$ elements respectively, be such that $\A \models \psi(\Q,\c) = \varphi(\c_1,\Q,\c_2)$.
-Then we have that $\A \models \exists \x \, \varphi(\x,\Q,\c_2)$, from which we conclude that $f(\A) > 0$.
-\end{proof}
-Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula, so $g \in\XE{0}$.
-It follows from \ref{mono} that for each $\A \in \str$, it holds that $\A \in L_f$ if and only if $g(\A) > 0$.
-But by Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time.
-From this we conclude that the decision version of $f$ is in $\ptime$.
-
-\textbf{Proof of Theorem \ref{sub-pnp}}
-
-Suppose that $\E{1}$ is closed under substraction, that is, for each pair of functions $f,g\in \E{1}$, there exists $h\in\E{1}$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\str$.
-
-Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\Phi$, where $A$ is the set of variables mentioned in $\Phi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{1,2,3,4\}$:
-\begin{eqnarray*}
-	S_1^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-	S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-	S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-	S_4^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
-\end{eqnarray*}
-Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
-\begin{multline*}
-\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
-\end{multline*}
-Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
-$$
-\varphi(x,X) = (T(x) \vee \neg T(x)).
-$$
-Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Given that $f_{\#3DNF}, f_{all} \in \E{1}$, we have by our initial assumption that $f_{all}-f_{\#3DNF} \in \E{1}$.
-Let $h\in\E{1}$ be such that $h = f_{all}-f_{\#3DNF}$.
-For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is $\conp$-complete.
-However, as we showed previously in Theorem \ref{decisionptime}, since $h\in\E{1}$, we have that $L_h \in \ptime$.
-Then, $\conp \subseteq \ptime$, from which we conclude that $\ptime = \np$.
-
-\textbf{Proof of Theorem \ref{sigma1strict}}
-
-We will show that the $\E{1}$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
-By contradiction, suppose that it is.
-Let $\A = \langle A =\{1\},S^\A = \{1\},\leq^\A=\{(1,1)\} \rangle$.
-Then, $f(\A) = 1$.
-Now let $\B = \langle B =\{1,2\},S^\B = \{1\},\leq^\B=\{(1,1),(1,2),(2,2)\} \rangle$.
-Note that $\A$ is an induced substructure of $\B$.
-
-We have that for each function $g \in \#\Sigma_1$ and structures $\A_1,\A_2\in\str$, if $\A_1$ is an induced substructure of $\A_2$, then $g(\A_1) \leq g(\A_2)$ \cite{DBLP:journals/jcss/SalujaST95}.
-Therefore, $f(\B) \geq f(\A) = 1$.
-However, there is no assignment $s\in B$ to $x$ such that $\B\models\:\varphi(s)$, so $f(\B) = 0$, which leads to a contradiction.
-
-\subsection{Proof of Theorem \ref{sigmafo-minusone}}
-
-Let $\alpha$ be a $\eqso(\logex{1})$ formula over a signature $\R$.
-We will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_n$ be defined as in \ref{eqso-sigma-zero-in-fp}, that is, each $\beta_i$ is equal to $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}\sa{\bar{x}}1$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, where $\varphi$ is a formula in $\logex{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables.
-We further assume that each $\beta_i$ is either $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, since each $1$, $\varphi$ and $\sa{\bar{X}}\varphi$ can be replaced by the formulas $\sa{x}\min(x)$, $\sa{x}(\varphi\wedge\min(x))$ and $\sa{x}\sa{\bar{X}}(\varphi\wedge\min(x))$ respectively, while maintaining the value of $\sem{\alpha}(\A)$.
-
-The intuition behind our reasoning is separated in three points: (1) For each $\beta_i$ of the form $\sa{\bar{x}}\varphi$, the formula $\kappa(\beta_i)$ will count every tuple $\bar{x}$ that satisfies $\varphi$ except for the lexicographically smallest one.
-(2) For each $\beta_i$ of the form $\sa{\bar{X}}\sa{\bar{x}}\varphi$, the formula $\kappa(\beta_i)$ will isolate the smallest $\bar{X}$ that satisfies $\varphi$, and exclude the lexicographically smallest tuple $\bar{x}$ that satisfies $\varphi(\bar{X})$.
-And (3) if $\alpha = (\beta\add\sa{\bar{x}}\varphi)$ or $\alpha = (\beta\add\sa{\bar{X}}\sa{\bar{x}}\varphi)$, the formula $\kappa(\alpha)$ will exclude the lexicographically smallest tuple that satisfies $\varphi$ if and only if $\beta$ is equal to 0.
-
-\vspace{1em}
-
-For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha) = \sa{\bar{x}}\exists\bar{y}[\varphi(\bar{x},\bar{y})\wedge\exists\bar{z}(\varphi(\bar{z},\bar{y})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfils the desired condition.
-
-\vspace{1em}
-
-For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha)$ procedurally as follows: Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
-We suppose that $\varphi$ is in a DNF form that leaves formulas that do not mention $\bar{X}$ as is.
-If it is not, we convert $\varphi$ to this form with a standard transformation algorithm.
-Let $\varphi(\bar{X},\bar{x},\bar{y}) = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$ where each $\varphi_i$ has the form:
-$$
-\varphi_i(\bar{X},\bar{x},\bar{y}) = (\text{conjunction of $X_i$'s}) \wedge (\text{conjunction of $\neg X_i$'s})  \wedge (\text{an $\fo$ formula that does not mention any $X_i$}).
-$$
-Define $\varphi_i^{+}$, $\varphi_i^{-}$ and $\varphi_i^{\fo}$ as the formulas mentioned above.
-
-In our procedure we assume that in $\varphi_i^{+}\wedge\varphi_i^{-}$ every first-order variable from $\bar{x},\bar{y}$ is mentioned at most once.
-If not, we add new first-order variables from $\fv$ to $\bar{y}$ so that no variable is repeated, replace them in $\varphi_i^{+}\wedge\varphi_i^{-}$ and add their respective equalities in $\varphi_i^{\fo}$ so that the new formula is equivalent.
-For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
-
-Furthermore, we assume that $\varphi_i^{\fo}$ defines an ordered partition of the variables in the tuple $(\bar{x},\bar{y})$.
-For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$.
-A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
-On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
-For a given $k$, we define $\cB_k$ as the number of possible ordered partitions for a set of size $k$. 
-If $\varphi_i^{\fo}$ does not define an ordered partition, then for $1 \leq j \leq \cB_{\length(\bar{x},\bar{y})}$ 
-let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
-The formula $\varphi_i(\bar{X},\bar{x},\bar{y})$ is converted into $\bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}}\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \theta^j(\bar{x},\bar{y})$.
-Consider each of these disjuncts as a new $\varphi_i$.
-
-After the previous assumptions, we do the following: for each $X_j\in \bar{X}$ we check every instance of $X_j(\bar{w})$ in $\varphi^{+}_i$ and every instance of $X_j(\bar{z})$ in $\varphi^{-}_i$, where $\bar{w}$ and $\bar{z}$ are subtuples of $(\bar{x},\bar{y})$.
-If the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{w} = \bar{z}$, the entire formula $\varphi_i$ is removed from $\varphi$.
-
-Now let $\bar{v}$ be the tuple of every first-order variable mentioned in $\varphi_i^{+}$ and let $\bar{u}$ be such that $(\bar{x},\bar{y}) = (\bar{u},\bar{v})$.
-We define a formula $\mu_i$ that is satisfied only by the lexicographically smallest $\bar{v}$ that satisfies $\varphi_i^{\fo}$ for some $\bar{u}$:
-$$
-\mu_i(\bar{v}) = \exists\bar{u}\,\varphi_i^{\fo}(\bar{u},\bar{v})\wedge\forall\bar{u}'\forall\bar{v}'(\varphi_i^{\fo}(\bar{u}',\bar{v}')\to\bar{v}\leq\bar{v}').
-$$
-The following is a pivotal formula for this proof.
-Consider some ordered finite structure $\A$ over $\R$.
-Given our conditions for $\varphi_i$, if $\A\models\varphi_i$ and $\A\models\varphi^{\fo}_i(\bar{u},\bar{v})$ then for that $\bar{v}$ we can define a $\bar{X}$ that has exactly the tuples mentioned in $\bar{v}$ and nothing else.
-This is {\em the smallest $\bar{X}$ that satisfies $\varphi_i$ over $\bar{v}$}.
-And if $\bar{v}$ is the lexicographically smallest such that $\A\models\varphi_i(\bar{v})$, then it is the {\em the smallest $\bar{X}$ that satisfies $\varphi_i$}.
-The following formula is satisfied by every pair $(\bar{X},\bar{x})$, except for the one formed by the smallest $\bar{X}$ that satisfies $\varphi_i$ and the lexicographically smallest $\bar{x}$ that satisfies $\varphi_i(\bar{X})$.
-And if no pair satisfies $\varphi_i$, the following formula is satisfied by every pair.
-$$
-\psi_i(\bar{X},\bar{x}) = \exists\bar{v}(\mu_i(\bar{v})\wedge(\neg\varphi^{+}_i(\bar{X},\bar{v})\vee\bigvee_{X \in \bar{X}} \exists\bar{z}(X(\bar{z}) \wedge \bigwedge\limits_{\substack{\text{instances of }X(\bar{w}) \\ \text{in }\varphi^{+}_i(\bar{X},\bar{v})}}\bar{w}\neq\bar{z}))) \vee \exists\bar{x}'(\exists\bar{y}'\varphi_i(\bar{X},\bar{x}',\bar{y}')\wedge \bar{x}'<\bar{x})\vee\neg\exists\bar{v}\mu_i(\bar{v}) .
-$$
-\begin{claim}
-	Let $\varphi_i(\bar{X},\bar{x},\bar{y}) = \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X},\bar{x},\bar{y})$ be a quantifier-free formula that respects all the mentioned assumptions.
-For a given ordered structure $\A$ with domain $A$, if $\A\models\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists some $\bar{B}$ over $\A$ and $\bar{b}\in A^{\length(\bar{x})}$ such that $\A\models\varphi_i(\bar{B},\bar{b},\bar{a})$ for some $\bar{a}\in A^{\length(\bar{y})}$.
-\end{claim}
-\begin{proof}
-	Let $\A$ be an ordered structure with domain $A$ and let $\bar{a}\in A^{\length(\bar{y})}$ and $\bar{b}\in A^{\length(\bar{x})}$ such that $\A\models\varphi^{\fo}_i(\bar{b},\bar{a})$.
-Define $\bar{B} = (B_1,\ldots,B_{\length(\bar{X})})$ as $B_j = \{\bar{c}\mid\bar{c}\text{ is a subtuple of $(\bar{b},\bar{a})$ and $X_j(\bar{c})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{b},\bar{a})$}\}$.
-Towards a contradiction, suppose that $\A\not\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
-By the choice of $\bar{a}$ and $\bar{b}$, and construction of $\bar{B}$ it is clear that $\A\models\varphi^{\fo}_i(\bar{b},\bar{a})\wedge\varphi^{+}_i(\bar{B},\bar{b},\bar{a})$, so we have that $\A\not\models\varphi^{-}_i(\bar{B},\bar{b},\bar{a})$.
-Let $B_t(\bar{c})$ be such that $X_t(\bar{c})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{b},\bar{a})$ and $\A\not\models\neg B_t(\bar{c})$.
-This is, $\bar{c}\in B_t$.
-By the construction of $\bar{B}$ we have that $X_t(\bar{c})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{b},\bar{a})$.
-From our assumptions, we have that (1) every first-order variable is mentioned at most once in $\varphi^{-}_i(\bar{X},\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X},\bar{x},\bar{y})$, (2) $\varphi^{\fo}_i$ defines an ordered partition over $(\bar{x},\bar{y})$, and (3) there are no $X_j(\bar{w})$ in $\varphi^{+}_i$ and $X_j(\bar{z})$ in $\varphi^{-}_i$ such that the ordered partition defined by $\varphi^{\fo}_i$ satisfies $\bar{x} = \bar{z}$.
-Since $\bar{c}$ is a subtuple of $(\bar{b},\bar{a})$ that satisfies $\varphi^{\fo}_i$, and $X_t(\bar{w})$ mentioned in $\varphi^{+}_i$ and $X_t(\bar{z})$ mentioned in $\varphi^{-}_i$ were both assigned the value $\bar{c}$, then the ordered partition satisfies $\bar{w} = \bar{z}$, which follows to a contradiction.
-\end{proof}
-We define $\chi_i = \exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$.
-Now recall that $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$, where each $\varphi_i$ satisfies all the previous assumptions.
-For each $\varphi_i$ we define:
-$$
-\varphi_i'(\bar{X},\bar{x},\bar{y}) = \varphi_i(\bar{X},\bar{x},\bar{y})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\vee\psi_2(\bar{X},\bar{x}))\wedge(\chi_1\vee\chi_2\vee\psi_2(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigvee_{j = 1}^{j = i-1}\chi_j\vee\psi_i(\bar{X},\bar{x})),
-$$
-and lastly, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x},\bar{y})$.
-
-\vspace{1em}
-
-For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y}))$ for some algebraic formula $\beta$ and some quantifier-free formula $\varphi$, we define $\kappa(\alpha)$ as follows: First, perform the same transformations to $\varphi(\bar{X},\bar{x},\bar{y})$ as in the previous case.
-Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$ where each $\varphi_i$ satisfies the previous assumptions.
-We also use the previously defined formulas $\chi_i$ and $\psi_i$.
-
-We recursively define a function $\lambda$ as follows.
-If $\alpha = \sa{\bar{x}}\exists\bar{y}\varphi(\bar{x},\bar{y})$, then $\lambda(\alpha) = \exists\bar{x}'\exists\bar{y}'\varphi(\bar{x}',\bar{y}')$.
-If $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ and $\lambda(\alpha) = \chi_1\vee\cdots\vee\chi_m$ as each $\chi_i$ was previously defined.
-If $\alpha = (\beta + \gamma)$, then $\lambda(\alpha) = \lambda(\beta) \vee \lambda(\gamma)$.
-Note that for a given ordered structure $\A$, then $\A\models\lambda(\alpha)$ if and only if $\sem{\alpha}(\A) > 0$.
-
-Now, for each $\varphi_i$ we define:
-$$
-\varphi_i'(\bar{X},\bar{x},\bar{y}) = \varphi_i(\bar{X},\bar{x},\bar{y})\wedge[\lambda(\beta)\vee[\psi_1(\bar{X},\bar{x})\wedge(\chi_1\vee\psi_2(\bar{X},\bar{x}))\wedge(\chi_1\vee\chi_2\vee\psi_2(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigvee_{j = 1}^{j = i-1}\chi_j\vee\psi_i(\bar{X},\bar{x}))]].
-$$
-And lastly $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x},\bar{y})$.
-%%Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$.
-There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$.
-This separates the proof in three cases:
-%\begin{enumerate}
-%%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
-%%	\begin{eqnarray*}
-%%		f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
-%%	\end{eqnarray*}
-%%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^d$.
-Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily.
-First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
-%%	\begin{eqnarray*}
-%%		\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
-%%	\end{eqnarray*}
-%%	This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
-%%	\begin{eqnarray*}
-%%		\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
-%%	\end{eqnarray*}
-%%	If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
-Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$.
-Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E{1}.$
-%%	
-%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
-%	\begin{eqnarray*}
-%		f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
-%	\end{eqnarray*}
-%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
-%	\begin{eqnarray}
-%	\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
-%	\end{eqnarray}
-%	where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
-Note that $\theta(\x)$ also mentions all variables in $\x$.
-We also assume that $(\x_1,\dots,\x_n,\y) = \x$.
-As an example, the following formula is of this form:
-%	\begin{multline*}
-%	\varphi(\x,\X) =  X_1(x_1,x_2) \wedge X_3(x_3) \wedge X_2(x_4,x_5) \wedge X_3(x_6) \wedge \neg X_1(x_7,x_8) \wedge \\ (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 ) \wedge \\ \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big),
-%	\end{multline*}
-%	where $\x = (x_1,x_2,x_3,x_4,x_5,x_6,x_7,x_8)$ and $\X = (X_1,X_2,X_3)$.
-Here, $n = 4$, $\lambda(1) = 1$, $\lambda(2) = \lambda(4) = 3$ and $\lambda(3) = 2$, $\x_1 = (x_1,x_2)$, $\x_2 = (x_3)$, $\x_3 = (x_4,x_5)$, $\x_4 = (x_6)$ and $\y = (x_7,x_8)$.
-Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_7,x_8)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 )$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2,x_5,x_8\},\{x_6,x_7\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big)$.
-%	
-%	Similarly to the previous proof, we would like to eliminate the {\em lexicographically smallest}\footnote{We consider the lexicographically smallest tuple of predicates as the one in which its predicates contain the lexicographically smallest tuples and do not contain any more tuples than those} tuple of predicates that satisfies the formula \eqref{f1}.
-Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
-We now define
-%	\begin{multline*}
-%	\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to (\u_1,\dots,\u_n)<(\v_1,\dots,\v_n)\Big)\Big],
-%	\end{multline*}
-%	where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
-Note that $\alpha_{\min}$ is satisfied only by the lexicographically smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
-Our new formula is
-%	\begin{multline}
-%	\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge \\ \exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg].
-%	\end{multline}
-%	We now show a result by which the main proof will follow.
-%	\begin{lemma} \label{first}
-%		$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
-%	\end{lemma}
-%	\begin{proof}
-%		Let $\A \in \str$.
-Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j\in[1,n]:\:\lambda(j)=i}\{\d_j\}$.
-We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
-%		\begin{enumerate}
-%			\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$.
-That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$).
-Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$.
-Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$.
-We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \z<\x_i$ or (3) $\theta(\x)\models \x_i<\z$.
-Considering that (2) and (3) are not possible given that both $\z$ and $\x_i$ are assigned the value $\d_i$ and $\A\models\theta(\d)$, we have that $\theta(\x)\models \z = \x_i$.
-But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$, which means $\varphi(\x,\X)$ is inconsistent.
-That is, there's no structure $\A^\prime\in\str$ such that $\A^\prime\models\exists\x\:\varphi(\x,\R^\prime)$ for any assignment $\R^\prime$ to $\X$.
-In particular, $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which contradicts the initial assumption that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-%			\item[(b)] Note that if $\A \models \alpha_{\min}(\c_1,\ldots,\c_n)$, then necessarily $\c_i = \d_i$ for $i\in\{1,\ldots,n\}.$ However, by the construction of $\P$, we see that 
-%			$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \d_j\Big).$$ Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
-%			\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$.
-Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $P^\prime_i$.
-Then $\A\models \neg P^\prime_i(\d_i)$, and so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $P_i$ does not have.
-If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-On both cases, we have a contradiction.
-%		\end{enumerate}
-%		With this, we conclude that for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-%		
-%		Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi(\x,\P)$, we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
-And so, for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ it holds that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1,$ which was to be shown.
-%	\end{proof}
-%	We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
-By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\gamma_i(\x,\X)$ with $i\in\{1,\ldots,\ell\}$ such that
-%	\begin{eqnarray*}
-%		\varphi(\x,\X) &\equiv& \gamma_1(\x,\X) \vee \gamma_2(\x,\X) \vee \dots  \vee \gamma_{\ell}(\x,\X),
-%	\end{eqnarray*}
-%	where, for every $i\in\{1,\ldots,\ell\}$, 
-%	\begin{eqnarray*}
-%		\gamma_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \gamma^{-}_i(\X,\y_i)  \wedge \delta_i(\x).
-%	\end{eqnarray*}
-%	The function $\lambda_i$ is defined analogously to $\lambda$ of the first part of the proof.
-The tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y_i) = \x.$ The formulas $\gamma^{-}_i(\X,\y_i)$, $\delta_i(\x)$ are defined analogously to $\varphi^{-}(\X,\y)$ and $\beta(\x)\,$\footnote[2]{Note that $\delta_i(\x)$ can include subformulas of the form $(\u = \u)$.} respectively (see \eqref{f1}).
-Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables.
-Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$.
-Note that for every formula $\eta(\x)$,
-%	\begin{eqnarray*}
-%		\exists\x\:\eta(\x) &\equiv& \exists\x(\eta(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\eta(\x)\wedge\theta^{g(d)}(\x)).
-%	\end{eqnarray*}
-%	We define the following formulas $\xi_i(\x,\X)$, for $i\in\{1,\ldots,m\}$ where $m = \ell \cdot g(d)$, as follows: 
-%	\begin{eqnarray*}
-%		\xi_1(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^1(\x), \\
-%		& \vdots & \\
-%		\xi_{g(d)}(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
-%		\xi_{g(d)+1}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^1(\x), \\
-%		& \vdots & \\
-%		\xi_{2\cdot g(d)}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
-%		& \vdots & \\
-%		\xi_{(\ell-1) g(d)+1}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^1(\x), \\
-%		& \vdots & \\
-%		\xi_{\ell \cdot g(d)}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^{g(d)}(\x).
-%	\end{eqnarray*}
-%	Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\str.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
-We use an injective\footnote[3]{We need this function to be injective because this way we can assure each one of the $k$ satisfiable disjuncts to be represented} function $\rho:\{1,\ldots,k\}\to\{1,\ldots,m\}$ such that $\varphi_i(\x,\X) = \xi_{\rho(i)}(\x,\X)$ is satisfiable, for every $i\in\{1,\ldots,k\}.$
-%	
-%	We can conclude that
-%	\begin{eqnarray*}
-%		\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X),
-%	\end{eqnarray*}
-%	and for every $i\in\{1,\ldots,k\}$,
-%	\begin{eqnarray*}
-%		\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
-%	\end{eqnarray*}
-%	where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\y_i$ is a $p_i$-tuple of variables in $\x$, $\varphi_i^{-}(\X,\y_i)$ is a conjunction of negated predicates in $\X$, $\theta_i(\x)$ defines a total order on a partition of $\x$, and $\beta_i(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
-Let $\alpha_i(\x) = \theta_i(\x)\wedge\beta_i(\x).$
-%	
-%	%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
-%	%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
-%	
-%	\begin{claim} \label{alpha}
-%		Let $\A\in\str$ and $i\in\{1,\ldots,k\}$. For every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, it holds that $f_{\varphi_i(\s,\X)}(\A) > 0.$
-%	\end{claim}
-%	\begin{proof}
-%		Let $\s = (\s_1,\ldots,\s_{n_i},\t)$ such that $\A\models\alpha_i(\s)$, and let $\P = (\P_1,\ldots,\P_r)$ where $P_i = \bigcup_{j\in[1,n_i]:\:\lambda(j)=i}\{s_i\}$.
-We will show that $\A\models\varphi_i(\s,\P)$.
-By contradiction, suppose that $\A\not\models\exists\x\,\varphi_i(\x,\P)$.
-That is, there is no assignment $\e$ to $\x$ such that $\A\models\varphi(\e,\P).$ Following the same proof as in case (a) of Lemma \ref{first}, we conclude that $\varphi_i(\x,\P)$ is inconsistent.
-However, as we mentioned previously, all of such disjuncts have been eliminated, which leads to a contradiction.
-%	\end{proof}
-%	
-%	Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\exists\x\:\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\Q$ such that $\A\models\exists\x\:\varphi_2(\x,\Q)$, and so on.
-As we already know how to exclude that assignment in the first disjunct, we will now deal with the next disjuncts.
-Similarly to the first part of the proof, for each $i\in\{1,\ldots,k\}$ let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
-%	\begin{multline*}
-%	\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_{n_i},\y)\wedge \\ \forall\v_1\cdots\forall\v_{n_i}\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_{n_i},\w)\wedge\bigvee_{j=1}^{n_i}(\u_j\neq\v_j)\big)\to (\u_1,\dots,\u_{n_i})<(\v_1,\dots,\v_{n_i})\Big)\Big],
-%	\end{multline*}
-%	where $\u_j$ and $\v_j,$ have $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, and $\w$ has $p_i$ variables. Also, let
-%	\begin{multline*}
-%	\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge\limits_{\substack{\ell\in[1,n_i]:\\ \lambda_i(\ell) = j}} \v \neq \u_\ell\Big) \bigg) \bigg].
-%	\end{multline*}
-%	Note that $\psi_i(\X)$ excludes only the lexicographically smallest tuple of predicates $\P$ such that $\A\models\exists\x\:\varphi_i(\P,\x),$ if there is at least one.
-In other words, every assignment $\P^\prime \neq \P$ is such that $\A\models\psi_i(\P^\prime).$ Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
-%	$$
-%	\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
-%	$$
-%	Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Assume that $\A\in\str$.
-Suppose first that there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least one assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\d = (\d_1,\dots,\d_{n_q},\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha_q(\d)$, where $\d_i$ is the corresponding assignment to $\x_i$, for every $i\in\{1,\dots,n_q\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j:\lambda_q(j)=i,j\in[1,n_q]}\{\d_j\}$.
-As we did in Lemma \ref{first} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies both (a) and (b)
-%	\begin{enumerate}
-%		\item[(a)] As we showed in part (a) of Lemma \ref{first}, if there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_q(\x,\R)$, then $\A\models\exists\x\:\varphi_q(\x,\P)$ for this particular $\P$.
-However, as we showed in Claim \ref{alpha}, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_q(\s)$, then there {\em is} such an assignment to $\X$.
-The assignment $\d$ to $\x$ satisfies that $\A\models\alpha_q(\d)$, so it holds that $\A\models\exists\x\:\varphi_q(\x,\P)$.
-It immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
-%		
-%		\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
-(1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
-(2) We can use the proof in Lemma \ref{first} to see that $\A\not\models\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula
-%		$$ \zeta_q(\X) =  (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{q-1}(\v)\vee\psi_q(\X)) $$ 
-%		appears as a conjunct in $\varphi^\prime_i(\x,\X)$.
-However, by the choice of $q$, there is no $i\in\{1,\ldots,q-1\}$ such that $\A\models\exists\v\:\varphi_i(\v,\P)$, and also $\A\not\models\psi_q(\P)$.
-It follows that $\A\not\models\zeta_q(\P)$,  so $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$.
-And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
-%		
-%		\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
-As we deduce from the part (c) of Lemma \ref{first}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
-Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x)$, so $\A\models\psi_i(\P^\prime)$ for each $i$.
-But as we mentioned, also $\A\models\:\psi_q(\P^\prime),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which leads to a contradiction.
-%	\end{enumerate}
-%	
-%	With this, we conclude that for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0$, we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-%	
-%	Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\chi(\x,\P))$ for any formula $\chi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
-And so, for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Hence, from the results in the previous paragraph, if $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}$, we have that $f^\prime = f\dotminus 1\in\E{1}$
-%	
-%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
-%	\begin{eqnarray*}
-%		f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
-%	\end{eqnarray*}
-%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$.
-In order to prove that $f\dotminus 1\in\E{1}$, we define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2, where
-%	\begin{eqnarray*}
-%		\varphi(\x,\X,\z) \equiv \varphi_1(\x,\X,\z) \vee \cdots \vee \varphi_k(\x,\X,\z),
-%	\end{eqnarray*}
-%	and
-%	\begin{eqnarray*}
-%		\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\w_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
-%	\end{eqnarray*}
-%	where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\ell} + c_{\ell} = a_{\ell}$ for $\ell\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, and $\w_i$ has $q_i$ variables.
-Furthermore, we have that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\w_i) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
-The formulas $\varphi^{-}_i(\X,\y_i,\w_i)$, $\beta_i(\x,\z)$, are also defined analogously.
-%	
-%	In this case, we mix both the strategies in cases 1 and 2.
-That is, we are going to {\em isolate} the lexicographically smallest tuple of predicates that satisfies the first satisfiable disjunct, and then {\em exclude} the lexicographically smallest tuple that satisfies the isolated disjunct.
-%	
-%	Let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and
-%	\begin{multline*}
-%	\alpha^i_{\min}(\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}) = \exists\y\exists\w\Big[ \alpha_i(\x_1,\dots,\x_{n_i},\y,\z_1,\dots,\z_{n_i},\w)\wedge \\ \forall\u_1\cdots\forall\u_{n_i}\forall\s\forall\v_1\cdots\forall\v_{n_i}\forall\t\Big(\big(\alpha_i(\u_1,\dots,\u_{n_i},\s,\v_1,\dots,\v_{n_i},\t)\wedge\bigvee_{j=1}^{n_i}(\x_j\neq\u_j \vee \z_j\neq\v_j)\big)\to \\ (\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i})<(\u_1,\dots,\u_{n_i},\v_1,\dots,\v_{n_i})\Big)\Big],
-%	\end{multline*}
-%	and let
-%	\begin{multline*}
-%	\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\w(\exists\u\:\varphi(\u,\X,\w) \wedge \w<\z).
-%	\end{multline*}
-%	Let $\P$ and $\e$ be assignments to $\X$ and $\z$.
-Note that $\psi_i(\X,\z)$ excludes $\P$ and $\e$ only if $\P$ is the {\em lexicographically smallest} tuple of predicates (Same as case 2) such that $\A\models\exists\x\:\varphi_i(\x,\P,\d)$, for some assignment $\d$ to $\z$, and $\e$ is the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_i(\x,\P,\e).$ We define $\varphi_i^\prime(\x,\X,\z)$ in the same way as case 2:
-%	\begin{multline*}
-%	\varphi_i^\prime(\x,\X,\z) = \varphi_i(\x,\X,\z) \wedge \psi_1(\X,\z) \wedge (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\psi_2(\X,\z)) \wedge \cdots \wedge \\ (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\cdots\vee\exists\u\exists\v\:\alpha_{i-1}(\u,\v)\vee\psi_i(\X,\z)).
-%	\end{multline*}
-%	Finally, let $\varphi^\prime(\x,\X,\z) = \bigvee_{i = 1}^k \varphi_i^\prime(\x,\X,\z)$.
-%	
-%	Let $q$ be the least $i\in\{1,\ldots,k\}$ such that $\A\models\exists\x\:\varphi_i(\x,\R,\d)$ for some assignment $R$ to $\X$ and some assignment $\d$ to $\z$.
-Let $\P$ be the lexicographically smallest tuple of predicates such that $\A\models\exists\x\:\varphi_q(\x,\P,\d^\prime)$ for some assignment $\d^\prime$ to $\z$.
-Let $\e$ be the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_q(\x,\P,\e).$ This formula is such that (a) $\A\models\exists\x\:\varphi(\x,\P,\e)$, (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P,\e)$ and $\P$ and $\e$ are the only assignments that satisfy (a) and (b).
-The proof of this is analogous to case 2.
-Therefore, we conclude that $f_{\exists\x\,\varphi^\prime(\x,\X,\z)} = f_{\exists\x\,\varphi(\x,\X,\z)}\dotminus 1.$
-%\end{enumerate}
-
-\textbf{Proof of Theorem \ref{dnf-pars}}
-
-Let $f$ be an arbitrary function in $\E{1}$ and let $f_{\textsc{\#DNF}}$ be the function that defines {\sc \#DNF}.
-We will now show a function $h:\str\to \textsc{Struct}[\L_{\textsc{DNF}}]$, where $\L_{\textsc{DNF}} = \{N,P\}$, such that for every $\A\in\str$, it holds that $f(\A) = f_{\textsc{\#DNF}}(h(\A))$, and which can be computed in polynomial time.
-For every DNF formula $\Phi$ we define $\A_{\Phi} \in \textsc{Struct}[\L_{\textsc{DNF}}]$ as stated in \ref{dnfdef}.
-Let $\psi$ be such that $f = f_{\psi(\X,\z)}$.
-We separate the proof in three cases:
-\begin{enumerate}
-	\item $\X = ()$.
-Let $\psi(\z) = \exists \x\,\varphi(\x,\z)$ where $\x = (x_1,\ldots,x_c)$ and $\z = (z_1,\ldots,z_d)$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a propositional formula $\Phi = h(\A)$.
-First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\,\varphi(\x,\e)$.
-Let $n$ be the number of such assignments to $\z$.
-Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$.
-The DNF formula we generate is
-	\[
-	\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
-	\]
-	which has exactly $n$ possible assignments. Finally, we return $h(\A) = \A_{\Phi}$.
-
-	\item $\z = ()$.
-Let $\psi(\X) = \exists \x\,\varphi(\x,\X)$, where $\x = (x_1,\ldots,x_c)$ and $\X = (X_1,\ldots,X_r)$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a DNF-formula $\Phi = h(\A)$ as follows.
-First, we convert $\varphi(\x,\X)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X)$.
-Let each $\varphi_{i}(\x,\X)$ be of the form
-	\[
-	\varphi_{i}(\x,\X) = X_{\pi_i(1)}(\x^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\x^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\x^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\x^{\nu}_{i,n_i}) \wedge \alpha_i(\x),
-	\]
-	where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$) is the rearrangement of $\x$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x)$ is a FO formula.
-We also define the tuple $\a^{\pi}_{i,j}$ ($\a^{\nu}_{i,j}$) as $\a$ rearranged in the same way as $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$), for each $\a\in A^c$.
-
-	We generate a propositional variable $p^{\ell}_{\a}$ for each $\a\in A^{c_{\ell}}$, and for each $\ell\in\{1,\ldots,r\}$, which will indicate if the tuple $\a$ is in the assignment to $X_{\ell}$.
-Let $\p$ be the tuple of all such variables.
-
-	Then, let $\Gamma_i = \{\a\in A^c\mid \A\models\alpha_i(\a)\}$.
-There is exactly $\vert A \vert^c$ tuples to check, and each of those checks can be done in polynomial time, so this set can also be generated in polynomial time.
-
-	Finally, for each $\varphi_i(\x,\X)$ the DNF formula we generate is
-	\[
-	\Phi_i = \bigvee_{\a\in\Gamma_i} p^{\pi_i(1)}_{\a^{\pi}_{i,1}} \wedge \cdots \wedge p^{\pi_i(m_i)}_{\a^{\pi}_{i,m_i}} \wedge \neg p^{\nu_i(1)}_{\a^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\pi_i(n_i)}_{\a^{\nu}_{i,n_i}}.
-	\]
-	and $\Phi = \bigvee_{i = 1}^k \Phi_i \vee \bigvee_{p\in\p} (p\wedge\neg p)$. Finally, we return $h(\A) = \A_{\Phi}$.
-
-	As an example, let $\psi^{*}(X_1,X_2) = \exists x \exists y\, (X_1(x) \wedge X_2(y,x) \wedge \neg X_1(y) \wedge x < y \wedge \neg\exists z\,(x<z\wedge z<y))$.
-We will compute $h(\B)$, where $\B = \langle \{1,2,3\}, < \rangle$, and we have that $1 < 2 < 3$.
-We generate the tuple of variables $\p = (p^1_1,p^1_2,p^1_3) \cup (p^2_{(i,j)})$ for $(i,j)\in \{1,2,3\}\times\{1,2,3\}$.
-Note that in this procedure we will only have one set $\Gamma_1 = \{(1,2),(2,3)\}$.
-And so, the DNF formula we generate is
-	\begin{multline*}
-	\Phi^{*} = (p^1_1\wedge p^2_{(2,1)} \wedge \neg p^1_2)\vee(p^1_2\wedge p^2_{(3,2)} \wedge \neg p^1_3)\, \vee \\ (p^1_1 \wedge \neg p^1_1) \vee(p^1_2 \wedge \neg p^1_2) \vee (p^1_3\wedge \neg p^1_3) \vee \bigvee_{i,j\in\{1,2,3\}}(p^2_{(i,j)}\wedge\neg p^2_{(i,j)})
-	\end{multline*}
-
-	Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$.
-We can conclude that $f(\A) = f_{\textsc{DNF}}(\Phi)$.
-
-	\item Let $\psi = \exists \x\,\varphi(\x,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$ we generate a DNF-formula $\Phi = h(\A)$ as follows.
-As in the previous case, we convert $\varphi(\x,\X,\z)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X,\z)$.
-Let each $\varphi_{i}(\x,\X,\z)$ be of the form
-	\[
-	\varphi_{i}(\x,\X,\z) = X_{\pi_i(1)}(\y^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\y^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\y^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\y^{\nu}_{i,n_i}) \wedge \alpha_i(\x,\z),
-	\]
-	where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$) is the rearrangement of $(\x,\z)$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x,\z)$ is a FO formula.
-We also define the tuple $(\a,\e)^{\pi}_{i,j}$ ($(\a,\e)^{\nu}_{i,j}$) as $\a$ rearranged in the same way as $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$), for each $(\a,\e)\in A^{c+d}$.
-
-	We generate a propositional variable $p^{\e}_{\ell,\b}$ for each $\e\in A^d$, for each $\ell\in\{1,\ldots,r\}$, and for each $\b\in A^{c_{\ell}}$.
-This variable will be true if there is an assignation to $\P$ such that $\b\in P_{\ell}$ and $\A\models\psi(\P,\e)$.
-Let $\p$ be the tuple of all such variables.
-We also generate a set $T = \{t_1,\ldots,t_{L}\}$ where $L = \lceil \log\vert A^d\vert \rceil$, and an arbitrary injective function $\beta:A^d \to \{0,1\}^L$ where $\beta(\e) = (b^{\e}_1,\ldots,b^{\e}_L)$ for each $\e\in A^d$.
-Also, let $\rho_{\ell}$ be such that $\rho_{\ell}(\x,\z) = \y_{\ell}$.
-
-	For each $\e\in A^d$, we follow the following procedure.
-For each $\varphi_i(\x,\X)$, we consider the set $\Gamma^{\e}_i = \{\a\in A^c\mid \A\models\alpha(\a,\e)\}$, and so, the DNF formula we generate is
-	\[
-	\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{\pi_i(1),(\a,\e)^{\pi}_{i,1}} \wedge \cdots \wedge p^{\e}_{\pi_i(m_i),(\a,\e)^{\pi}_{i,m_i}} \wedge \neg p^{\e}_{\nu_i(1),(\a,\e)^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\e}_{\pi_i(n_i),(\a,\e)^{\nu}_{i,n_i}} \wedge \gamma_{\e} \wedge \theta_{\e},
-	\]
-	where $\gamma_{\e}$ is defined as
-	\[
-	\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
-	\]
-	and $\theta_{\e}$ is defined as
-	\[
-	\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
-	\]
-	Finally,$\Phi = \bigvee_{\e\in A^d} \bigvee_{i = 1}^k \Phi_i^{\e} \vee \bigvee_{p\in\p} (p\wedge\neg p)$, and we return $f(\A) = \A_{\Phi}$.
-
-	As an example, let $\psi^{*}(X_1,X_2,z) = \exists x\, (X_1(x) \wedge X_2(z) \wedge \neg X_1(z) \wedge x < z)$.
-We will compute $h(\B)$, where $\B = \langle \{1,2,3\}, < \rangle$, and we have that $1 < 2 < 3$.
-We generate the tuple of variables $\p = (p^e_{1,1},p^e_{1,2},p^e_{1,3},p^e_{2,1},p^e_{2,2},p^e_{2,3})$, for $e\in \{1,2,3\}$.
-Here, we consider $L = 2$, $T = \{t_1,t_2\}$, and $\beta$ defined as $\beta(1) = (0,0)$, $\beta(2) = (0,1)$ and $\beta(3) = (1,0)$.
-In this procedure we generate the sets $\Gamma^1_1 = \emptyset$, $\Gamma^2_1 = \{1\}$, $\Gamma^3_1 = \{1,2\}$.
-Now we generate $\gamma_e$ for $e\in\{1,2,3\}$,
-	\[
-	\gamma_e = \bigwedge_{d\in\{1,2,3\}\setminus\{e\}} p^d_{1,1} \wedge p^d_{1,2} \wedge p^d_{1,3} \wedge p^d_{2,1} \wedge p^d_{2,2} \wedge p^d_{2,3},
-	\]
-	and $\theta_e$ for $e\in\{1,2,3\}$,
-	\begin{align*}
-	\theta_1 &= \neg t_1 \wedge \neg t_2,\\
-	\theta_2 &= \neg t_1 \wedge t_2,\\
-	\theta_3 &= t_1 \wedge \neg t_2.		
-	\end{align*}
-	Then, we generate the formulas $\Phi^e_1$, for $e\in\{1,2,3\}$,
-	\begin{align*}
-	\Phi^1_1 &= \square\\
-	\Phi^2_1 &= p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2,\\
-	\Phi^3_1 &= (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3),
-	\end{align*}
-	and so, the DNF formula we generate is
-	\begin{multline*}
-	\Phi = (p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2) \vee (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee \\ \bigvee_{e\in\{1,2,3\}}[(p^e_{1,1}\wedge\neg p^e_{1,1})\vee
-	(p^e_{1,2}\wedge\neg p^e_{1,2})\vee
-	(p^e_{1,3}\wedge\neg p^e_{1,3})\vee
-	(p^e_{2,1}\wedge\neg p^e_{2,1})\vee
-	(p^e_{2,2}\wedge\neg p^e_{2,2})\vee
-	(p^e_{2,3}\wedge\neg p^e_{2,3})].
-	\end{multline*}
-	To demonstrate that our reduction is correct, we will show a function $g:\{\langle\P,\e\rangle \mid \A \models \exists\x\,\varphi(\x,\P,\e)\} \to \{\sigma \mid \sigma(\Phi) = 1\}$ and then prove that it is a bijection.
-Let $g$ be such that for each assignment $\langle \P,\e \rangle$, it holds that $g(\langle \P,\e \rangle) = \sigma$ where $\sigma$ is constructed as follows.
-For each $t_i\in T$, let $\sigma(t_i) = b^{\e}_i$.
-For each $p_{\ell,\b}^{\e}$, where $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e}) = 1$ if and only if $\b \in P_{\ell}$.
-For each $p_{\ell,\b}^{\e'}$, where $\e'\in A^d\setminus\{\e\}$,  $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e'}) = 1$.
-
-	Let $i$ be such that $\A\models \exists\x\,\varphi_i(\x,\P,\e)$.
-It is simple to see that from the way we constructed $\Phi_i^{\e}$, it holds that $\sigma(\Phi_i^{\e}) = 1$.
-Also, note that every different assignment $\langle \P',\e' \rangle$, it holds that $\sigma(\langle \P',\e' \rangle) \neq \sigma(\langle \P,\e \rangle)$.
-Therefore, $g$ is an injective function.
-Now, let $\sigma$ be an arbitrary assignment such that $\sigma(\Phi) = 1$.
-Let $i$ and $\e$ be such that $\sigma(\Phi^{\e}_i)$.
-Note that for every $\e'\in A^d\setminus\{\e\}$, it holds that $\sigma(\theta_{\e'}) = 0$, so there is exactly one possible $\e$ associated to $\sigma$.
-Consider the assignment $\P$ where $\b\in P_{\ell}$ if and only if $\sigma(p^{\e}_{\ell,\b}) = 1$, for each $\ell$ and $\b \in A^{c_{\ell}}$.
-Since there is at least one $\a\in\Gamma_i$ for which $\sigma$ satisfies its respective subformula, it holds that $\A\models\varphi_i(\a,\P,\e)$, which implies that $\A\models\exists\x\,\varphi(\x,\P,\e)$.
-From this we deduce that $g$ is surjective, and therefore a bijection.
-We conclude that $f(\A) = f_{\textsc{\#DNF}}(\A_{\Phi})$.
-\end{enumerate}
-
-\textbf{Proof of Theorem \ref{nplusone-strict}}
-
-Suppose it is not properly contained for $n$.
-Let $\z = (z_1,\ldots,z_{n+1})$, let $\u = (u_1,\ldots,u_n)$ and let $\varphi(\z) = \bigwedge_{i = 1}^{n} (z_i < z_{i+1} \wedge \forall y(z_i\not<y \vee y\not<z_{i+1} )) \wedge \forall y(y \not< z_1)$.
-Note that $f_{\varphi(\z)}(\A) = 1$ if $\A$ has at least $n+1$ elements, and  $f_{\varphi(\z)}(\A) = 0$ otherwise, for each $\A\in\str$.
-Let $\forall\x\,\psi(\x,\X,\u)$ be such that $f_{\varphi(\z)} = f_{\forall\x\,\psi(\x,\X,\u)}$.
-Let $\A = \langle A, \S^{\A}, <^{\A} \rangle\in\str$ where $\vert A \vert = n+1$.
-Note that $f_{\forall\x\,\psi(\x,\X,\u)}(\A) = 1$.
-Let $(\P,\e)$ be such that $\A\models\forall\x\,\psi(\x,\P,\e)$.
-Let $a\in A$ be such that $a\not\in\e$.
-Consider the vocabulary $\L' = \L \cup \{\Q\}$, and let $\A' = \A \cup \langle \Q^{\A'} \rangle$ where $\Q^{\A'} = \P$.
-Note that $\A\models\varphi(\P,\e)$ if and only if $\A'\models\varphi(\e)$.
-Consider $\B'$ as the induced substructure of $\A'$ with domain $A\setminus\{a\}$, and let $\Q^{\B'}$ be the induced interpretation of $\Q$.
-Also let $\B$ be the induced substructure of $\A$ with domain $A\setminus\{a\}$.
-As universal formulae are preserved under substructure\footnote{Hodges, 1997.
-P.
-183}, we have that $\B'\models\forall\x\,\psi(\x,\e)$, which implies that $\B\models\forall\x\,\psi(\x,\Q^{\B'},\e)$.
-We conclude that $f_{\forall\x\,\psi(\x,\X,\u)}(\B) \geq 1$, which leads to a contradiction.
-
-\textbf{Proof of Theorem \ref{sigma2-pe}}
-
-Let $f = f_{\varphi(\X,\z)}$ such that $\varphi(\X,\z)$ is defined by the grammar $E_2$.
-First we notice that, as stated in \cite{DBLP:journals/jcss/SalujaST95}, for every $\L$-formula $\varphi$, $f_{\varphi}\in\shp$.\\
-
-We will now prove that there is a polynomial time algorithm that decides $L_{f}$.
-This is equivalent to decide, given $\A= \langle A, \S^\A, \leq^\A \rangle\in\str$, whether there exist assignments $\P,\z$ such that $\A\models\varphi(\P,\z)$.
-This is the same as $\A\models\exists\X\exists\x\,\varphi(\X,\x)$.
-Since $\varphi(\X,\x)$ is defined by the grammar $E_2$, there exists $\psi(\X,\x,\y,\u)$ such that $\varphi(\X,\x) = \exists\x\exists\y\forall\u\exists\v\,\psi(\X,\x,\y,\u,\v)$, where every variable in $\v$ appears on a negated second-order literal $\neg X_i$, with $i\in\N$.
-
-Given $\A\in\str$ we generate an equivalent $\L$-formula $\theta(\X,\x,\y)$ with a series of operations.
-First, we replace every instance of $\v$ by a disjunction of all $r$-tuples in $A^r$, where $r$ is the number of variables in $\v$.
-The result is still a Horn clause.
-Second, we replace every instance of $\u$ by a conjunction of all $s$-tulples in $A^s$, where $s$ is the number or variables in $\u$.
-The result is still a conjunction of Horn clauses.
-
-Now we notice that $\zeta = \exists\X\exists\x\exists\y\,\theta(\X,\x,\y)$ is an existential second-order FO-formula, for which $\A\models\zeta$ can be decided in polynomial time \cite{DBLP:conf/stacs/Gradel91}.
-
-\textbf{Proof of Theorem \ref{sigma2hard}}
-
-Let $f$ be an arbitrary function in $\XE{2}$.
-We will now show a function $h:\str\to L(P)$ such that for every $\A\in\str$, it holds that $f(\A) = \shdhsat(h(\A))$, and which can be computed in polynomial time.
-Let $\psi$ be such that $f = f_{\psi}$.
-We separate the proof in three cases:\begin{enumerate}
-	\item All free variables in $\psi$ are of first order.
-Let $\psi = \exists \x \forall \y\,\varphi(\x,\y,\z)$ where $\z = (z_1,\ldots,z_d)$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a propositional formula $\Phi = h(\A)$.
-First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\forall \y\,\varphi(\x,\y,\e)$.
-Let $n$ be the number of such assignments to $\z$.
-Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$.
-The disjunction of Horn formulas we return is
-	\[
-	\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
-	\]
-	which has exactly $n$ possible assignments. Note that each $p_i$ and $\neg p_i$ is a Horn clause itself.
-
-	\item There are only second-order free variables in $\psi$.
-Let $\psi = \exists \x\forall \y\,\varphi(\x,\y,\X)$, where $\x = (x_1,\ldots,x_c)$, $\y = (y_1,\ldots,y_b)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a Disj-Horn formula $\Phi = h(\A)$ as follows:
-
-	First, Let $\varphi(\x,\y,\X) = \bigwedge_{i = 1}^k \varphi_{i}(\x,\y,\X)$ where, without loss of generality, each $\varphi_{i}(\x,\y,\X)$ is of the form
-	\[
-	\varphi_{i}(\x,\y,\X) = X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n}) \vee \alpha_i(\x,\y),
-	\]
-	where $\u_{s}$ and every $\u_{t_j}$ have the corresponding number of variables in $(\x,\y)$, and $\alpha(\x,\y)$ is a FO-formula.
-For the rest of the proof, $X_s(\u_s)$ can be absent and it does not affect the proof.
-Note that this formula is equivalent to
-	\[
-	\neg\alpha_i(\x,\y) \to (X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n})),
-	\]
-
-	Then, for each $\a\in A^c$ let $\Gamma_i^{\a} = \{\b\in A^{b}\mid \A\models\neg\alpha(\a,\b)\}$.
-Note that this set can be generated in polynomial time.
-For each $\ell\in\{1,\ldots,r\}$, we generate a propositional variable $p^{\ell}_{\d}$ for each $\d\in A^{c_{\ell}}$.
-Let $\p$ be the tuple of all such variables.
-Also, let $\rho_{\ell}(\x,\y) = \u_{\ell}$, and $\rho_{\ell}(\d)$ be the tuple that results of reordering $\d$ using the reordering of the variables of $(\x,\y)$ in $\u_{\ell}$.
-
-	For each $\a\in A^c$ we generate a Horn formula
-	\[
-	\Phi_{\a}^i = \bigwedge_{\b\in\Gamma_i} p^{s}_{\rho_s(\a,\b)} \vee \neg p^{t_1}_{\rho_{t_1}(\a,\b)} \vee \cdots \vee \neg p^{t_n}_{\rho_{t_n}(\a,\b)}.
-	\]
-	and finally we return $\Phi = \bigvee_{\a\in A^c} \bigwedge_{i = 1}^k \Phi_{\a}^i$.
-
-	Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\forall\y\,\varphi(\x,\y,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$.
-We can conclude that $f(\A) = f_{\textsc{Disj-Horn-Sat}}(\Phi)$.
-
-	\item Let $\psi = \exists \x\forall\y\,\varphi(\x,\y,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\y = (y_1,\ldots,y_b)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$ we generate a Disj-Horn formula $\Phi = h(\A)$ as follows:
-
-	As in the previous case, let $\varphi(\x,\y,\X,\z) = \bigwedge_{i = 1}^k$ where $\varphi(\x,\y,\X,\z)$ is of the form:
-	\[
-	\varphi_{i}(\x,\y,\X,\z) = X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n}) \vee \alpha_i(\x,\y,\z),
-	\]
-	where $\u_{s}$ and every $\u_{t_j}$ have the corresponding number of variables in $(\x,\y,\z)$, and $\alpha(\x,\y,\z)$ is a FO-formula.
-For the rest of the proof, $X_s(\u_s)$ can be absent and it does not affect the proof.
-Note that this formula is equivalent to
-	\[
-	\neg\alpha_i(\x,\y,\z) \to (X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n})).
-	\]
-
-	For each $\e\in A^d$, we do the following.
-For each $\a\in A^c$ let $\Gamma^{\e}_{i,\a} = \{\b \in A^b \mid \A\models \neg\alpha_i(\a,\b,\e) \}$.
-Note that this set can be generated in polynomial time.
-For each $\ell\in\{1,\ldots,r\}$ we generate a propositional variable $p^{\e}_{\ell,\d}$ for each $\d\in A^{c_{\ell}}$.
-Let $\p$ be the tuple of all such variables.
-We also generate a set $T = \{t_1,\ldots,t_{L}\}$ where $L = \lceil \log\vert A^d\vert \rceil$, and an arbitrary injective function $\beta:A^d \to \{0,1\}^L$ where $\beta(\e) = (b^{\e}_1,\ldots,b^{\e}_L)$ for each $\e\in A^d$.
-Also, let $\rho_{\ell}(\x,\y,\z) = \u_{\ell}$, and $\rho_{\ell}(\d)$ be the tuple that results of reordering $\d$ using the order of the variables in $(\x,\y,\z)$ in $\u_{\ell}$.
-And so, the Horn formula we generate is
-	\[
-	\Phi^{\e}_{i,\a} = \bigwedge_{\a\in\Gamma^{\e}_i} p^{\e}_{s,\rho_{s}(\a,\b,\e)} \vee p^{\e}_{t_1,\rho_{t_1}(\a,\b,\e)} \vee \cdots \vee p^{\e}_{t_n,\rho_{t_n}(\a,\b,\e)}.
-	\]
-	And finally we return $\Phi = \bigvee_{\e\in A^d}\bigvee_{\a\in A^c} (\bigwedge_{i=1}^k \Phi_{\e}^{i,\a} \wedge \gamma_{\e} \wedge \theta_{\e} )$, where $\gamma_{\e}$ is defined as
-	\[
-	\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
-	\]
-	and $\theta_{\e}$ is defined as
-	\[
-	\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
-	\]
-
-\end{enumerate}
-
-\textbf{Proof of Theorem \ref{so-rec}}
-
-\textbf{Proof of Theorem \ref{tqfo-fo-cap}}
-
-For the second condition, let $f \in \shl$.
-We will address the case where $\R$ contains only one binary predicate $E$, and the rest of the cases can be deduced from this.
-Let $M$ be a non-deterministic logspace machine such that $f(\A) = \acc_M(\A)$ for each $\A \in \ostr[R]$.
-Suppose ${\cal Q} = \{q_1,\ldots,q_{\ell}\}$ is the set of states of $M$, where $q_1$ is the initial state, and $q_{\ell}$ is only final state of $M$.
-Let $n = \vert A \vert$ and let $w = \enc(\A) \in \{0,1\}^{n^2}$.
-We assume that $M$ with input $w$ uses space $s_M(w) < c\cdot\log(n)$ and furthermore, $s_M(w) < n-2$.
-We notate $M(w)$ as the graph of configurations of $M$ running on input $w$.
-
-We represent configurations with a tuple of fixed size.
-The formula $\varphi(\bar{x},\bar{y})$ describes a procedure that given a configuration generates a possible next configuration.
-The formula $\varphi_I(\bar{x})$ describes that $\bar{x}$ is the initial configuration of $M(w)$.
-The formula $\varphi_F(\bar{x})$ describes that $\bar{x}$ is an accepting (final) configuration of $M(w)$.
-The formula we construct is:
-$$
-\alpha = \sa{\bar{x}}\sa{\bar{y}}([\pth \varphi(\bar{u},\bar{v})](\bar{x},\bar{y})\cdot \varphi_I(\bar{x})\cdot\varphi_F(\bar{y})).
-$$
-
-To illustrate our idea, we will show a simplified example.
-Consider a machine $M$ that works in exactly $\log_2(n)$ space and only allows 0 or 1 in the working tape.
-Consider an input $\A$ of size 16 (that is, $A = \{0,\ldots,9,A,\ldots,F\}$).
-Let some configuration $s$ have 0011 in the working tape, the head in the input tape is in position 26, and the head in the input tape is in position 2 (we consider 0-indexed positions).
-Also, $Q = \{q_1,\ldots,q_5\}$ and the current state is $q_3$.
-
-As a first approach, we will use a 9-tuple $\bar{a} = (a_1,\ldots,a_9)$ to represent $s$.
-That is, $(a_1,a_2) = (1,A)$ represent the position of the head in the input tape, $a_3 = 2$ represents the position of the head in the working tape, $a_4 = C$ (1100b in base 16) represents the content of the working tape, and $(a_5,\ldots,a_9) = (0,0,1,0,0)$ represents the current state.
-Then $\bar{a} = (1,A,2,C,0,0,1,0,0)$ will represent $s$.
-
-The problem that arises from this representation, is that to describe a transition in $M$ we need to read an arbitrary character in the working tape, and therefore we have to obtain the $a_3$-th bit in $a_4$.
-Furthermore, to represent the following configuration, we need compute $a_4$ with the $a_3$-th bit flipped.
-This is generally not possible to describe with an $\fo$ formula.
-Consider the following procedure that receives $x = a_4$ and $i = a_3$.
-
-\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 1 replace it by 0 and return the result}
-	\label{switch1to0}
-	\begin{algorithmic}
-		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
-		\While{$j > 0$}
-		\State $v \gets 0$
-		\While{$u > 1$}
-		\State $u \gets u-2,\; v \gets v+1$
-		\EndWhile
-		\State $u\gets v,\; j \gets j-1$
-		\EndWhile
-		\While{$u > 1$}
-		\State $u \gets u-2$
-		\EndWhile
-		\State $\textbf{assert } u = 1$ \Comment{If $u \neq 1$ simply stop}	
-		\State $y \gets 1$ \Comment{Compute $2^i$ and store it in $y$}
-		\While{$i > 0$}
-		\State $z \gets 0$
-		\While{$y > 0$}
-		\State $z \gets z+2,\; y \gets y-1$
-		\EndWhile
-		\State $i \gets i-1,\; y \gets z$
-		\EndWhile
-		\While{$y > 0$} \Comment{Substract $y$ from $x$}
-		\State $x \gets x-1,\; y \gets y-1$
-		\EndWhile
-		\State \Return $x$.
-	\end{algorithmic}
-\end{algorithm}	
-Each of the instructions can be expressed with $\fo$, so our strategy is to use the $\pth$ operator to simulate the algorithm and then we can describe a transition using the processed value of $a_4$.
-Note that we need to describe three more procedures to simulate the transitions for $0\to 0$, $0 \to 1$ and $1\to 1$.
-
-We will now describe how to simulate both the procedure and the transition.
-A procedure tuple $\bar{p} = (a_1,\ldots,a_{3+c+\ell},b_1,b_2,c_1,c_2,c_3,d_1,\ldots,d_{5c+2})$ represents the current configuration of $M(w)$ in $a_1,\ldots,a_{2+c+\ell}$, the values that will be read and written in the working tape in $b_1,b_2$, the instruction pointer in $c_1,c_2,c_3$ and the values stored in memory in $d_1,\ldots,d_{10c+2}$.
-In detail:
-\begin{enumerate}
-	\item $a_1,a_2$ and $a_3$ represent the position of the head in the input tape and the working tape, respectively, $a_4,\ldots,a_{3+c}$ represent the content of the working tape and $a_{4+c},\ldots,a_{3+c+\ell}$ represent the current state in the current configuration that is being processed.
-	\item $b_1$ and $b_2$ are equal to the value that is being read in the working tape and the value that will be written in the working tape respectively.
-	\item $c_1,c_2,c_2$ represent the instruction pointer in the procedure. Only 8 different instructions are needed in the simulation.
-	\item Each value in memory of $x,y,z,u,v$ need $c$ elements to represent them and $i,j$ need only one. We map $(d_1\ldots,d_{c}) \to x$, $(d_{c+1}\ldots,d_{2c}) \to y$,
-	$(d_{2c+1}\ldots,d_{3c}) \to z$, $(d_{3c+1}\ldots,d_{4c}) \to u$,
-	$(d_{4c+1}\ldots,d_{5c}) \to v$, $d_{5c+1} \to i$ and $d_{5c+2}\to j$.
-\end{enumerate}
-For each transition $\delta \in \Delta \subseteq Q \times \{0,1\} \times \{0,1\} \times Q \times \{-1,=,+1\} \times \{0,1\} \times \{-1,=,+1\}$ we define a formula $\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$, where $\bar{x} = (x_1,\ldots,x_{3+c+\ell})$, $\bar{s} = (s_1,s_2)$, $\bar{w} = (w_1,w_2,w_3)$, $\bar{u} = (u_1,\ldots,u_{5c+2})$, $\bar{y} = (y_1,\ldots,y_{3+c+\ell})$, $\bar{t} = (t_1,t_2)$, $\bar{z} = (z_1,z_2,z_3)$ and $\bar{v} = (v_1,\ldots,v_{5c+2})$.
-The tuples $\bar{x}$ and $\bar{y}$ represent the current and next configuration of $M$ respectively, $\bar{s}$ and $\bar{y}$ indicate which algorithm is being followed, $\bar{w}$ and $\bar{z}$ represent the current and next instruction of the algorithm, $\bar{u}$ and $\bar{v}$ represent the current and next values in memory.
-We will describe the formula part by part.
-Suppose $\delta = (q_i,a,1,q_j,op_1,0,op_2)$, so we have to simulate Algorithm \ref{switch1to0}.
-
-First we define some auxiliary formulas:
-\begin{align*}
-\varphi_0(x) &= \neg\exists y(y < x), \\
-\varphi_{+1}(x,y) &= x < y \wedge \neg\exists z(x < z \wedge z < y), \\
-\varphi_{-1}(x,y) &= \varphi_{+1}(y,x), \\
-\varphi_1(x) &= \exists y(\varphi_0(y) \wedge \varphi_{+1}(y,x)), \\
-\text{for $i,j \in\{0,1\}$, }	\varphi_{i,j}(x,y) &= \varphi_i(x) \wedge \varphi_j(y), \\
-\text{for each $k \leq 7$, }\varphi^b_k(x_1,x_2,x_3) &= \varphi_{a_1}(x_1) \wedge \varphi_{a_2}(x_2) \wedge \varphi_{a_3}(x_3)\text{, where $a_1a_2a_3$ is the value of $k$ in binary}, \\
-\mu_{<}(x_1,\ldots,x_c,y_1,\ldots,y_c) &= \bigvee_{i = 1}^c ((\bigwedge_{j = 1}^{i-1} x_j = y_j) \wedge x_i < y_i), \\
-\mu_{+1}(\bar{x},\bar{y}) &= \mu_{<}(\bar{x},\bar{y}) \wedge \neg\exists \bar{z}(\mu_{<}(\bar{x},\bar{z}) \wedge \mu_{<}(\bar{z},\bar{y})), \\
-\mu_{+2}(\bar{x},\bar{y}) &= \exists\bar{z}(\mu_{+1}(\bar{x},\bar{z}) \wedge \mu_{+1}(\bar{z},\bar{y})),\\
-\mu_{-1}(\bar{x},\bar{y}) &= \varphi_{+1}(\bar{y},\bar{x}),\\
-\mu_{-2}(\bar{x},\bar{y}) &= \varphi_{+2}(\bar{y},\bar{x}),\\
-\mu_{0}(\bar{x}) &= \neg\exists\bar{y}(\mu_{<}(\bar{y},\bar{x})),\\
-\mu_{1}(\bar{x}) &= \exists\bar{y}(\mu_{0}(\bar{y})\wedge\mu_{+1}(\bar{y},\bar{x})),\\
-\varphi^2_{<}(x_1,x_2,y_1,y_2) &= x_1 < y_1 \vee (x_1 = y_1 \wedge x_2 < y_2),\\
-\varphi^2_{+1}(x_1,x_2,y_1,y_2), \varphi^2_{-1}(x_1,x_2,y_1,y_2) &\text{ defined analogously to $\mu_{+1}$ and $\mu_{-1}$},\\
-\text{for each } q_i\in Q, \varphi^q_i(x_1,\ldots,x_{\ell}) &= \bigwedge_{\substack{j = 1 \\ j \neq i}}^{\ell} \varphi_0(x_j) \wedge \varphi_1(x_i), \\
-\varphi^E_0(x_1,x_2) &= \neg E(x_1,x_2),\\		\varphi^E_1(x_1,x_2) &= E(x_1,x_2),\\
-\end{align*}
-
-We start from instruction 0, which means that the procedure has not started yet and every value in the tuple is 0 except for the configuration values.
-It also initializes all the values in the tuple to 0 except for $x,u,i,j$.
-\begin{multline*}
-\varphi^{0,1}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{0,0}(s_1,s_2)\wedge\varphi^b_0(\bar{w}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_1(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c v_i = x_{3+i} \wedge \bigwedge_{i = c+1}^{2c} \varphi_0(v_i) \wedge \bigwedge_{i = 2c+1}^{3c} \varphi_0(v_i) \wedge \bigwedge_{i = 1}^c v_{3c+i} = x_{3+i} \wedge \bigwedge_{i = 4c+1}^{5c} \varphi_0(v_i) \wedge v_{5c+1} = x_3 \wedge v_{5c+2} = x_3.
-\end{multline*}
-Instruction 1 which checks whether the value of $j$ ($d_{5c+2}$ in the tuple) is more than 0 or not, and then proceeds to instruction 2 or 3 on each case.
-\begin{multline*}
-\varphi^{1,2}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_1(\bar{w}) \wedge \neg \varphi_0(u_{5c+2}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_2(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i, \\
-\varphi^{1,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_1(\bar{w}) \wedge \varphi_0(u_{5c+2}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_3(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
-\end{multline*}
-Instruction 2 checks the value of $u$ ($d_{3c+1},\ldots,d_{4c}$ in the tuple).
-If it is $> 1$ then it substracts 2 from $u$ and adds 1 to $v$ ($d_{4c+1},\ldots,d_{5c}$ in the tuple), then repeats instruction 2.
-If it is equal to 0 or 1, then moves the value of $v$ to $u$, substracts 1 from $j$ and goes back to instruction 1.
-\begin{multline*}
-\varphi^{2,2}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_2(\bar{w}) \wedge \neg \mu_0(u_{3c+1},\ldots,u_{4c}) \wedge \neg \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_2(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
-\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-\mu_{-2}(u_{3c+1},\ldots,u_{4c},v_{3c+1},\ldots,v_{4c})\, \wedge \\
-\mu_{+1}(u_{4c+1},\ldots,u_{5c},v_{4c+1},\ldots,v_{5c}) \wedge u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}. \\
-\varphi^{2,1}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_2(\bar{w}) \wedge ( \mu_0(u_{3c+1},\ldots,u_{4c}) \vee \mu_1(u_{3c+1},\ldots,u_{4c})) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_1(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
-\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-\bigwedge_{i = 3c+1}^{4c} u_{c+i} = v_i \wedge
-\mu_0(v_{4c+1},\ldots,v_{5c}) \wedge
-u_{5c+1} = v_{5c+1} \wedge \varphi_{-1}(u_{5c+2},v_{5c+2}).	
-\end{multline*}
-Instruction 3 calculates the value of $u \mod 2$, that is, it repeats instruction 3 until the value of $u$ is equal to 0 or 1.
-On each iteration, it substracts 2 from $u$.
-Moreover, if the value of $u$ at the end of the iterations is not 1 then there is no step defined.
-\begin{multline*}
-\varphi^{3,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_3(\bar{w}) \wedge \neg \mu_0(u_{3c+1},\ldots,u_{4c}) \wedge \neg \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_3(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
-\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-\mu_{-2}(u_{3c+1},\ldots,u_{4c},v_{3c+1},\ldots,v_{4c}) \wedge
-\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
-u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}, \\
-\varphi^{3,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_3(\bar{w}) \wedge \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_4(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\mu_1(v_{c+1},\ldots,v_{2c}) \wedge
-\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
-\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
-u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}
-\end{multline*}
-Instruction 4 checks the value of $i$ ($d_{5c+1}$ in the tuple.) If it is not 0 then goes to instruction 5 and if is 0 then goes to instruction 6.
-Moreover it initializes the value of $z$ ($d_{2c+1},\ldots,d_{3c}$ in the tuple) to 0 (which was 0 all along.)
-\begin{multline*}
-\varphi^{4,5}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_4(\bar{w}) \wedge \neg \varphi_0(u_{5c+1}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_5(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i, \\
-\varphi^{4,6}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_4(\bar{w}) \wedge \varphi_0(u_{5c+1}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_6(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
-\end{multline*}
-Instruction 5 checks the value of $y$ ($d_{c+1},\ldots,d_{2c}$ in the tuple.) If it is more than 0 then it adds 2 to $z$ and substracts 1 from $y$, then repeats instruction 2.
-If it is not, then copies the value of $z$ to $y$ and subtracts 1 from $i$ and returns to instruction 4.
-\begin{multline*}
-\varphi^{5,5}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_5(\bar{w}) \wedge \neg \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_5(\bar{z}) \, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\mu_{-1}(u_{c+1},\ldots,u_{2c},v_{c+1},\ldots,v_{2c}) \wedge
-\mu_{+2}(u_{2c+1},\ldots,u_{3c},v_{2c+1},\ldots,v_{3c})\, \wedge \\
-\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
-\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
-u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}, \\
-\varphi^{5,4}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_5(\bar{w}) \wedge \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_4(\bar{z}) \, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\bigwedge_{i = c+1}^{2c} u_{c+i} = v_i \wedge
-\mu_0(v_{2c+1},\ldots,v_{3c}) \wedge
-\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
-\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
-\varphi_{-1}(u_{5c+1},v_{5c+1}) \wedge u_{5c+2} = v_{5c+2}
-\end{multline*}
-Instruction 6 checks the value of $y$. If it is more than 0, then subtracts 1 from $x$ and $y$ and repeats instruction 6. If it is not, then goes to instruction 7.
-\begin{multline*}
-\varphi^{6,6}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_6(\bar{w}) \wedge \neg \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_6(\bar{z}) \, \wedge \\
-\mu_{-1}(u_1,\ldots,u_c,v_i,\ldots,v_c) \wedge \mu_{-1}(u_{c+1},\ldots,u_{2c},v_{c+1},\ldots,v_{2c}) \wedge \bigwedge_{i = 3c+1}^{5c+2} u_i = v_i]\,\vee \\
-\varphi^{6,7}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_6(\bar{w}) \wedge \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_7(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
-\end{multline*}
-Instruction 7 stores the value of $x$ after the corresponding bit has been switched.
-Then we can define $\varphi_{\delta}$ which also simulates the actual transition.
-If $u$ equals 1, then copy what is stored in $x$ to $a_4,\ldots,a_{3+c}$, go from state $q_i$ to state $q_j$, and move the heads to their corresponding positions.
-\begin{multline*}
-\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 	[\bigwedge_{i = 1}^{3+c+\ell} x_i = y_i \wedge (\varphi^{0,1}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{1,2}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})\, \vee \\ \varphi^{1,3}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{2,2}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{2,1}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{3,3}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \\ \varphi^{3,4}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{4,5}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{4,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{5,5}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \\ \varphi^{5,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{6,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{6,7}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}))] \, \vee \\
-[\varphi^E_a(x_1,x_2) \wedge \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_7(\bar{w}) \wedge \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{0,0}(t_1,t_2) \wedge \varphi^b_0(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i \,\wedge \\
-\varphi^2_{op_1}(x_1,x_2,y_1,y_2) \wedge \varphi_{op_2}(x_3,y_3) \wedge \bigwedge_{i = 1}^c u_i = x_{3+i} \wedge \varphi^q_i(x_{4+c},\ldots,x_{3+c+\ell}) \wedge \varphi^q_j(y_{4+c},\ldots,y_{3+c+\ell})].
-\end{multline*}
-
-\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 0 return $x$} \label{switch0to0}
-	\begin{algorithmic}
-		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
-		\While{$j > 0$}
-		\State $v \gets 0$
-		\While{$u > 1$}
-		\State $u \gets u-2,\; v \gets v+1$
-		\EndWhile
-		\State $u\gets v,\; j \gets j-1$
-		\EndWhile
-		\While{$u > 1$}
-		\State $u \gets u-2$
-		\EndWhile
-		\State $\textbf{assert } u = 0$ \Comment{If $u \neq 0$ simply stop}	
-		\State \Return $x$.
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 0 replace it by 1 and return the result}
-	\label{switch0to1}
-	\begin{algorithmic}
-		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
-		\While{$j > 0$}
-		\State $v \gets 0$
-		\While{$u > 1$}
-		\State $u \gets u-2,\; v \gets v+1$
-		\EndWhile
-		\State $u\gets v,\; j \gets j-1$
-		\EndWhile
-		\While{$u > 1$}
-		\State $u \gets u-2$
-		\EndWhile
-		\State $\textbf{assert } u = 0$ \Comment{If $u \neq 0$ simply stop}	
-		\State $y \gets 1$ \Comment{Compute $2^i$ and store it in $y$}
-		\While{$i > 0$}
-		\State $z \gets 0$
-		\While{$y > 0$}
-		\State $z \gets z+2,\; y \gets y-1$
-		\EndWhile
-		\State $i \gets i-1,\; y \gets z$
-		\EndWhile
-		\While{$y > 0$} \Comment{Add $y$ to $x$}
-		\State $x \gets x+1,\; y \gets y-1$
-		\EndWhile
-		\State \Return $x$.
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 1 return $x$}
-	\label{switch1to1}
-	\begin{algorithmic}
-		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
-		\While{$j > 0$}
-		\State $v \gets 0$
-		\While{$u > 1$}
-		\State $u \gets u-2,\; v \gets v+1$
-		\EndWhile
-		\State $u\gets v,\; j \gets j-1$
-		\EndWhile
-		\While{$u > 1$}
-		\State $u \gets u-2$
-		\EndWhile
-		\State $\textbf{assert } u = 1$ \Comment{If $u \neq 0$ simply stop}	
-		\State \Return $x$.
-	\end{algorithmic}
-\end{algorithm}
-Note that we also need to specify that the program we are following is Algorithm \ref{switch1to0} so we store $1,0$ in $b_1,b_2$ all along the procedure.
-We describe the three other algorithms that compute the switches from $0\to 0$, $0\to 1$ and $1\to 1$ (Algorithms \ref{switch0to0}, \ref{switch0to1} and \ref{switch1to1}.)
-For the other three cases, where $\delta = (q_i,a,0,q_j,op_1,0,op_2)$, $\delta = (q_i,a,0,q_j,op_1,1,op_2)$ and $\delta = (q_i,a,1,q_j,op_1,1,op_2)$, $\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$ is defined analogously.
-Then, $\varphi(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$ is defined as:
-$$
-\varphi(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \bigvee_{\delta \in \Delta} \varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}).
-$$
-Finally we define $\varphi_I$ and $\varphi_F$:
-\begin{align*}
-\varphi_I(\bar{x},\bar{s},\bar{w},\bar{u}) &= \varphi_0(x_1) \wedge \varphi_0(x_2) \wedge \varphi_0(x_3) \wedge \bigwedge_{i = 4}^{3+c}\varphi_0(x_i) \wedge \varphi_1(x_{4+c}) \wedge \bigwedge_{i = 5+c}^{3+c+\ell} \varphi_0(x_i)\wedge \varphi_{0,0}(s_1,s_2)\ \wedge \varphi^b_0(\bar{w}) \wedge \bigwedge_{i = 1}^{5c+2}\varphi_0(u_i).
-\\
-\varphi_F(\bar{x},\bar{s},\bar{w},\bar{u}) &= \varphi^q_{\ell}(x_{4+c},\ldots,x_{3+c+\ell}) \wedge \varphi_{0,0}(s_1,s_2)\ \wedge \varphi^b_0(\bar{w}) \wedge \bigwedge_{i = 1}^{5c+2}\varphi_0(u_i),
-\end{align*}
-and then $\sem{\alpha}(\A) = \sem{\sa{\bar{x}}\sa{\bar{y}}([\pth \varphi(\bar{u},\bar{v})](\bar{x},\bar{y})\cdot \varphi_I(\bar{x})\cdot\varphi_F(\bar{y}))}(\A) = \acc_M(\A)$.
-
-\textbf{Proof of Theorem \ref{tqso-fo-cap}}
-
-\textbf{Proof of Theorem \ref{rqfo-fo-cap}}

### 2016-12-28 LICS/appendix.tex
28 dic

--- 2016-12-28 LICS/appendix.tex
@@
 We also denote $\length(\bar{x}) = m$ ($\length(\bar{X}) = m$).
+
 \subsection{Proof of Theorem \ref{no-mult}}
@@
 We separate the proof in two main cases:
+
 \begin{enumerate}
+
 	\item At least one of $\beta_1, \beta_2$ is not of the form $\pa{x} \gamma$, $\pa{X} \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. Without loss of generality, let $\beta_1$ be such one.
+
 	\begin{enumerate}
+
 		\item $\beta_1 = s \in \nat$.
@@
 Since each of the formulas paired by $\add$ has less than $n$ algebraic operators, the statement holds.
+
 		\item $\beta_1 = (\gamma_1 \add \gamma_2)$.
 Then, we use $\alpha' = (\gamma_1\mult\beta_2 + \gamma_2\mult\beta_2)$ and each of the formulas paired by $\add$ have less than $n$ algebraic operators so the statement holds.
+
 		\item $\beta_1 = \sa{x} \gamma$. Then, we use $\alpha' = \sa{x} (\gamma\mult\beta_2)$ and as in the previous case the statement holds.
+
 		\item $\beta_1 = \sa{X} \gamma$, which we treat exactly as the previous one.
+
 	\end{enumerate}
+
 	\item Both of $\beta_1$ and $\beta_2$ are of the form $\pa{x} \gamma$, $\pa{X} \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. This case also separates in further cases.
+
 	\begin{enumerate}
+
 		\item $\beta_1 = \pa{x} \gamma_1(x)$ and $\beta_2 = \pa{y} \gamma_2(y)$. Then we use $\alpha' = \pa{x}(\gamma_1(x)\mult\gamma_2(x))$, and the statement holds.
+
 		\item $\beta_1 = \pa{X} \gamma_1(X)$ and $\beta_2 = \pa{Y} \gamma_2(Y)$. [something about arities].
+
 		\item $\beta_1 = \pa{X} \gamma_1(X)$ and $\beta_2 = \pa{x} \gamma_2(x)$ (w.l.o.g.). [use singleton predicates].
+
 		\item $\beta_1 = \pa{x} \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). Then, we use $\alpha' = \pa{x} (\gamma \cdot \varphi)$, and the statement holds.
+
 		\item $\beta_1 = \pa{X} \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). We treat this case exactly as the previous one.
+
 		\item $\beta_1 = \varphi_1$ and $\beta_2 = \varphi_2$ both $\so$-formulas. Then, we use $\alpha' = (\varphi_1 \wedge \varphi_2)$ and the statement holds.
+
 	\end{enumerate}
+
 \end{enumerate}
+
+>>>>>>> 601d7da9913124a5a6c9adf12a847ca836fb40f5
+
+%\end{enumerate}
 
@@
 
-We make the following observation.
-If $\alpha$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$, then the {\bf propositional formula} defined by $\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ has exactly $\sem{\alpha}(\A)$ assignments and is a disjunction of Horn formulas.
+Our idea for the rest of the proof is to define $g_{\alpha}$ for each $\alpha = \sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})$, for $\alpha = c$ and for $\alpha = \beta_1 + \cdots + \beta_m$ where each $\beta_i$ is in one of the two previous cases.
+
+If $\alpha$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ where $\psi_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$, then we obtain the {\bf propositional formula} $g_{\alpha}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulas.
+
+If $\alpha$ is equal to a constant $c$, then we define $g_{\alpha}(\A)$ as the following formula that has exactly $c$ satisfying assignments:
+$$
+g_{\alpha}(\A) = \bigvee_{i = 1}^{c}\neg t_1 \wedge \cdots \wedge \neg t_{i-1} \wedge t_i \wedge \neg t_{i+1} \wedge \cdots \wedge \neg p_c.
+$$ 
+If $\alpha = \beta_1 + \cdots + \beta_m$, let $g_{\beta_i}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\theta^i_{i,j}$ for each $\beta_i$ where each $\theta^i_{j,k}$ is a Horn clause.
+Let $\Theta_i = g_{\beta_i}(\A)$.
+We rename the variables in each $\Theta_i$ so none of them are mentioned more than once in total.
+We add $m$ new variables $t_1,\ldots,t_m$ and we define:
+\begin{align*}
+g_{\alpha}(\A) = &\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^1_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m}}}t) \wedge (t_1 \wedge \bigwedge_{\ell = 2}^{m} \neg t_{\ell})) \vee \\
+&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^2_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each $t$}\\ \text{ mentioned in}\\ \Theta_1,\Theta_3,\ldots,\Theta_{m}}}t) \wedge (t_2 \wedge \bigwedge\limits_{\substack{\ell = 1 \\ \ell \neq 2}}^{m} \neg t_{\ell})) \vee \cdots \vee\\
+&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^m_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m-1}}}t) \wedge (t_m \wedge \bigwedge_{\ell = 1}^{m-1} \neg t_{\ell})).
+\end{align*}
+The number of satisfying assignments for this formula is exactly the sum of satisfying assignments for all $g_{\beta_i}(\A)$.
+This covers all possible cases for $\alpha$, and the complete procedure takes polynomial time.
 

### 2016-12-29 LICS/counting.tex
Arreglo apendice.

@@
 Moreover, these closure properties also hold if we change the boolean core $\fo$ to any polynomial-time logic like $\lfp$. This gives the following corollary.
-\begin{corollary} \label{eqso-lfp-cap}
+\begin{corollary} \label{prop:capture-shP-LFP}
 	$\eqso(\lfp)$ captures $\shp$ over ordered structures.
@@
 As one could expect, if we parametrize $\eqso$ with $\eso$, we can also capture~$\spp$.
-\begin{proposition} \label{eqso-eso-cap}
+\begin{proposition} \label{prop:capture-spanP}
 	$\eqso(\eso)$ captures $\spp$ over ordered structures.
@@
 This leads us to the fragment $\qso(\pfp)$ which naturally captures $\fpspace$. 
-\begin{theorem} \label{qso-pfp-cap}
+\begin{theorem} \label{theo:capture-fpspace}
 	$\qso(\pfp)$ captures $\fpspace$ over ordered structures.

### 2016-12-29 LICS/appendix-counting.tex

--- 2016-12-28 LICS/appendix.tex
@@
-%!TEX root = main.tex
+\medskip
 
-\subsection{Notation for the appendix}
+%%% DEMOSTRACION DE SQSO(FO) y #P
+\subsection*{Proof of Proposition~\ref{prop:capture-shP}}
 
-For a given signature $\R$, we define $\ostr[\R]^*$ as $$\ostr[\R]^* = \{(\A,v,V) \mid \A\in\ostr[\R]\text{, $v$ ($V$) is a first-order (second-order) assignment for $\A$}  \}.$$
-The {\em conditional count} symbol $(\varphi \mapsto \alpha)$ is defined as $(\neg\varphi + (\varphi\cdot\alpha))$ for given $\so$ formula $\varphi$ and $\qso$ formula $\alpha$.
-Note that for each $(\A,v,V) \in \ostr[\R]^*$,
-$$
-\sem{(\varphi \mapsto \alpha)}(\A,v,V) = 
-\begin{cases}
-\sem{\alpha}(\A,v,V) &\text{if } (\A,v,V)\models\varphi,\\
-0 &\text{otherwise}.
-\end{cases}
-$$
-We will use the symbol $<$ also to denote the lexicographic order over same-sized tuples.
-If $\bar{x} = (x_1,\ldots,x_m)$ and $\bar{y} = (y_1,\ldots,y_m)$ are tuples of first-order variables, we denote $\bar{x} < \bar{y}$ for the formula $\bigvee_{i = 1}^m[\bigwedge_{j = 1}^{i-1}x_j = y_j \wedge x_i < y_i]$.
-Similarly, we use $=$ to denote equality between tuples, as $\bar{x} = \bar{y}$ denotes $\bigwedge_{i = 1}^m(x_i = y_i)$, and also $\bar{x}\leq\bar{y}$ denotes $\bar{x} < \bar{y} \vee \bar{x} = \bar{y}$.
-We also denote $\min(\bar{x}) := \forall\bar{y}(\bar{y} < \bar{x}\vee \bar{y} = \bar{x})$.
+We will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
+This machine, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
+Suppose $\A$ has domain $A$.
+If $\alpha$ is a $\fo$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial time, and accepts if and only if it holds true.
+If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
+If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
+%If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
+If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
+%If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$. 
+If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
+This covers all possible cases for $\alpha$.
+Let $\alpha$ be a formula in $\eqso(\fo)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
+We have that $f$ is a $\shp$-function over $\R$ and $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
 
-If $\bar{x} = (x_1,\ldots,x_m)$ ($\bar{X} = (X_1,\ldots,X_m)$) is a tuple of first-order (second-order) variables, we denote $\sa{\bar{x}}\alpha$ for $\sa{x_1}\cdots\sa{x_m}\alpha$ and $\sa{\bar{X}}\alpha$ for $\sa{X_1}\cdots\sa{X_m}\alpha$ for each $\qso$ formula $\alpha$.
-We also denote $\length(\bar{x}) = m$ ($\length(\bar{X}) = m$).
-\subsection{Proof of Theorem \ref{no-mult}}
+For the other direction, note that Saluja et al.~\cite{SalujaST95} proved that $\shp = \sfo$. 
+%We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso(\fo)$ be its corresponding formula. 
+Since a function in $\sfo$ can also be defined $\eqso(\fo)$ (see Section~\ref{sec:previous}), the condition holds. \qed
 
-Let $\{+, \cdot, \sa{}, \pa{}\}$ be the set of {\it algebraic operators} in $\qso$.
-We will prove by induction over the number of algebraic operators in a $\qso$-formula $\alpha$ that there exists some $\qso$-formula $\alpha'$ that respects the stated condition.
-Let $\text{op}(\alpha)$ be the number of algebraic operators in $\alpha$.
-The case $\text{op}(\alpha) = 0$ is trivial.
-Let $n\in\nat$ be such that for each $\qso$-formula $\beta$ such that $\text{op}(\beta) < n$ there is some $\beta'$ that respects the condition.
-Let $\alpha$ be a $\qso$-formula such that $\text{op}(\alpha) = n$.
-Note that if $\alpha$ is of the form $\alpha = (\beta_1 \add \beta_2)$, $\alpha = \sa{x} \beta$, $\alpha = \pa{x} \beta$, $\alpha = \sa{X} \beta$ or $\alpha = \pa{X} \beta$ the proof follows directly.
-Therefore we assume that $\alpha = (\beta_1 \mult \beta_2)$.
-We separate the proof in two main cases:
-\begin{enumerate}
-	\item At least one of $\beta_1, \beta_2$ is not of the form $\pa{x} \gamma$, $\pa{X} \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. Without loss of generality, let $\beta_1$ be such one.
-	\begin{enumerate}
-		\item $\beta_1 = s \in \nat$.
-Then we use $\alpha' = \beta_2 \add \cdots \add \beta_2$, where $\beta_2$ is repeated $s$ times, and if $s = 0$, take $\alpha' = \perp$.
-Since each of the formulas paired by $\add$ has less than $n$ algebraic operators, the statement holds.
-		\item $\beta_1 = (\gamma_1 \add \gamma_2)$.
-Then, we use $\alpha' = (\gamma_1\mult\beta_2 + \gamma_2\mult\beta_2)$ and each of the formulas paired by $\add$ have less than $n$ algebraic operators so the statement holds.
-		\item $\beta_1 = \sa{x} \gamma$. Then, we use $\alpha' = \sa{x} (\gamma\mult\beta_2)$ and as in the previous case the statement holds.
-		\item $\beta_1 = \sa{X} \gamma$, which we treat exactly as the previous one.
-	\end{enumerate}
-	\item Both of $\beta_1$ and $\beta_2$ are of the form $\pa{x} \gamma$, $\pa{X} \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. This case also separates in further cases.
-	\begin{enumerate}
-		\item $\beta_1 = \pa{x} \gamma_1(x)$ and $\beta_2 = \pa{y} \gamma_2(y)$. Then we use $\alpha' = \pa{x}(\gamma_1(x)\mult\gamma_2(x))$, and the statement holds.
-		\item $\beta_1 = \pa{X} \gamma_1(X)$ and $\beta_2 = \pa{Y} \gamma_2(Y)$. [something about arities].
-		\item $\beta_1 = \pa{X} \gamma_1(X)$ and $\beta_2 = \pa{x} \gamma_2(x)$ (w.l.o.g.). [use singleton predicates].
-		\item $\beta_1 = \pa{x} \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). Then, we use $\alpha' = \pa{x} (\gamma \cdot \varphi)$, and the statement holds.
-		\item $\beta_1 = \pa{X} \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). We treat this case exactly as the previous one.
-		\item $\beta_1 = \varphi_1$ and $\beta_2 = \varphi_2$ both $\so$-formulas. Then, we use $\alpha' = (\varphi_1 \wedge \varphi_2)$ and the statement holds.
-	\end{enumerate}
-\end{enumerate}
+\medskip
 
-\subsection{Proof of Theorem \ref{captfp}}
+%%% DEMOSTRACION DE SQSO(ESO) y span-P
+\subsection{Proof of Proposition~\ref{prop:capture-spanP}}
+
+Similar than the previous proof, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
+This machine, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic outputs for each $(\A,v,V) \in \ostr[\R]^*$.
+Suppose $\A$ has domain $A$.
+On each of the steps the algorithm adds the current configuration to the output tape.
+\cristian{Para que necesitas esto?}
+If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time without modifying the output tape, and accepts if and only if the condition holds true.
+\cristian{Falta explicar mejor este caso y porque puedes ejecutar una $\eso$-formula.}
+If $\alpha$ is a constant $s$, then the algorithm produces $s$ branches and accepts in all of them.
+If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
+%If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$. 
+If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
+%If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$. 
+If $\alpha = \sa{X} \beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
+This covers all possible cases for $\alpha$.
+Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
+We have that $f$ is a $\spp$ function over $\R$ and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
+
+For the other direction, Compton et al.~\cite{ComptonG96} proved that $\spp = \#\eso$. Since a function in $\#\eso$ can also be defined in $\eqso(\eso)$, then $\eqso(\eso)$ captures $\spp$. \qed
+
+\medskip
+
+%%% DEMOSTRACION DE QFO(LFP) y FP
+\subsection*{Proof of Theorem~\ref{theo:capture-fp}}
 
@@
 Adding this values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\alpha(\enc(\A)) = f(\enc(\A))$.
+Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
+\qed
+\medskip
+
+%%% DEMOSTRACION DE QSO(PFP) y FPSPACE
+\subsection{Proof of Theorem~\ref{theo:capture-fpspace}}
+
+To show how to evaluate a $\qso(\pfp)$-formula, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
+This machine, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
+Suppose $\A$ has domain $A$.
+If $\alpha$ is a $\pfp$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space~\cite{L04}, and accepts if and only if it holds true.
+If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
+If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
+If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
+If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
+If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
+If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
+If $\alpha = \pa{X}\beta$, it simulates $M_{\beta}$ on input $(\A,v,V[B/X])$ consecutively for each $B\in A^{arity(X)}$.
+This covers all possible cases for $\alpha$, and each of these steps can be computed in polynomial space.
+Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
+We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$ \cite{Ladner89}, and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
+\cristian{Aca debes explicar mejor que hace la máquina.
+Es un copy paste de demostraciones anteriores y no queda claro porque puedes evaluar la formula en FPSPACE.
+Tampoco queda claro porque puedes usar no determinismo en FPSPACE.}
+
+For the second condition, let $f\in \fpspace$ defined over some $\R$.
+Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq 2^{n^\ell}$, where $\A$ has a domain of size $n$.
+Let $\Phi(X)$ be a $\pfp$ formula that represents the following.
+Consider an order over the subsets of $A^{\ell}$ given by the formula $$\varphi_{<}(X,Y) = \exists\bar{u}\big[\neg X(\bar{u})\wedge Y(\bar{u})\wedge \forall\bar{v}\big(
+\bar{u}<\bar{v}\to(X(\bar{u})\iff Y(\bar{v}))\big)\big].$$
+
+Let $B$ be the $p$-th subset with respect to this order and let $w$ be a string of size $2^{n^{\ell}}$ which represents the value of $f(\enc(\A))$ in binary.
+Then $\A\models\Phi(B)$ if and only if the $p$-th bit in $w$ is 1, from least to most significant.
+Since this condition can be verified in polynomial space, and $\pfp$ captures $\pspace$, then $\Phi$ can be expressed in $\pfp$.
+We define
+$\alpha := \sa{X}(\Phi(X)\mult\varphi(X))$, where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
+Note that for each $B\subseteq A^{\ell}$ such that $B$ is the $m$-th element in the given order, $\sem{\varphi(B)}(\A) = 2^m$.
+Therefore, $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
 \qed
 
-\subsection{Proof of Theorem \ref{qfo-pfp-cap}}
+\medskip
 
-\martin{no revisar, es un copy-paste de arriba}
+%%% DEMOSTRACION DE QFO(PFP) y FPSPACE(poly)
+\subsection*{Proof of Corollary~\ref{cor:capture-fpspace-poly}}
 
-For condition (2) in \ref{cap-def}, let $\alpha\in\qfo(\pfp)$ over some $\R$.
+Let $\alpha\in\qfo(\pfp)$ over some $\R$.
 Let $f$ be a function over $\R$ defined by the following procedure.
@@
 Then each sub-formula $\varphi\in\pfp$ in $\alpha$ is replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size which is evaluated and lastly given as output.
-Note that $f\in\nfpspace$ and $f(\enc(\A)) = \alpha(\enc(\A))$.
+\cristian{Explicar porque puedes hacer esto en FPSPACE(poly)}
+The resulting formula is an arithmetic expression of polynomial size which is evaluated and lastly given as output. Note that $f\in\nfpspace$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
 
-For condition (2) in \ref{cap-def}, let $f\in \nfpspace$ defined over some $\R$.
+Now, let $f\in \nfpspace$ defined over some $\R$.
 Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
@@
 $$
-and adding these values gives $f(\enc(\A))$. Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\alpha(\enc(\A)) = f(\enc(\A))$. \qed
-
-\subsection{Proof of Theorem \ref{eqso-fo-cap}}
-
-For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\fo$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial time, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.%%If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Let $\alpha$ be a formula in $\eqso(\fo)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shp$ function over $\R$ and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
-
-For condition (2) in \ref{cap-def}, note that Saluja et al.
-\cite{SalujaST95} proved that $\shp = \#\fo$.
-We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso(\fo)$ be its corresponding formula.
-Since this formula is also in $\eqso(\fo)$, the condition holds.
-\qed
-
-\subsection{Proof of Theorem \ref{eqso-eso-cap}}
-
-For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic outputs for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-On each of the steps the algorithm adds the current configuration to the output tape.
-If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time without modifying the output tape, and accepts if and only if the condition holds true.
-If $\alpha$ is a constant $s$, then the algorithm produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-%% If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\spp$ function over $\R$ and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
-
-For condition (2) in \ref{cap-def}, note that Compton and Graedel \cite{ComptonG96} proved that $\spp = \#\eso$.
-We also have that $\sqso(\eso)$ captures $\#\eso$ over ordered structures so for each $f\in \spp$ let $\alpha \in \sqso(\eso)$ be its corresponding formula.
-Since this formula is also in $\eqso(\eso)$, the condition holds.
-\qed
-
-\subsection{Proof of Theorem \ref{qso-pfp-cap}}
-
-For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\pfp$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space \cite{AbiteboulV89}\martin{no estoy seguro de esta cita}, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-If $\alpha = \pa{X}\beta$, it simulates $M_{\beta}$ on input $(\A,v,V[B/X])$ consecutively for each $B\in A^{arity(X)}$.
-This covers all possible cases for $\alpha$, and each of these steps can be computed in polynomial space.
-Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$ \cite{Ladner89}, and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
-
-For the second condition, let $f\in \fpspace$ defined over some $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq 2^{n^\ell}$, where $\A$ has a domain of size $n$.
-Let $\Phi(X)$ be a $\pfp$ formula that represents the following.
-Consider an order over the subsets of $A^{\ell}$ given by the formula $$\varphi_{<}(X,Y) = \exists\bar{u}\big[\neg X(\bar{u})\wedge Y(\bar{u})\wedge \forall\bar{v}\big(
-\bar{u}<\bar{v}\to(X(\bar{u})\iff Y(\bar{v}))\big)\big].$$
-
-Let $B$ be the $p$-th subset with respect to this order and let $w$ be a string of size $2^{n^{\ell}}$ which represents the value of $f(\enc(\A))$ in binary.
-Then $\A\models\Phi(B)$ if and only if the $p$-th bit in $w$ is 1, from least to most significant.
-Since this condition can be verified in polynomial space, and $\pfp$ captures $\pspace$, then $\Phi$ can be expressed in $\pfp$.
-We define
-$\alpha := \sa{X}(\Phi(X)\mult\varphi(X))$, where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
-Note that for each $B\subseteq A^{\ell}$ such that $B$ is the $m$-th element in the given order, $\sem{\varphi(B)}(\A) = 2^m$.
-Therefore, $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-\qed
-
-\subsection{Proof of Theorem \ref{eqso-sigma-zero-in-fp}}
-
-First, for a given signature $\R$ and logic fragment $\LL$ we define a {\em valid form in $\LL$} (or just {\em valid form}).
-A formula $\alpha$ is in valid form in $\LL$ if (1) it is equal to $\varphi$, $1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}1$ for some $\LL$ formula $\varphi$ and some tuple of first-order (second-order) variables $\bar{x}$ ($\bar{X}$), or (2) if it is equal to $(\beta\add\gamma)$ where $\beta$ and $\gamma$ are formulas in valid form.
-Given a $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$ we will recursively construct a formula $\tau(\alpha)$ in valid form such that $\alpha(\enc(\A)) = \tau(\alpha)(\enc(\A))$ for each $\A\in\ostr[\R]$.
-
-First, each natural constant $s$ in $\alpha$ is replaced by $(1\add 1\add\cdots\add 1)$ ($s$ times).
-If $\alpha = 1$ or $\alpha = \varphi \in \loge{0}$, then $\tau(\alpha) = \alpha$.
-We assume that every sub-formula of $\alpha$ is in valid form.
-If $\alpha = (\beta \add \gamma)$ then $\tau(\alpha) = \alpha$.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ this case separates in more cases: If $\alpha = \sa{x}\beta$ and $\beta$ is in one of the valid forms in (1), then $\tau(\alpha) = \alpha$.
-If $\alpha = \sa{X}\beta$ and $\beta$ is in the valid form $\sa{\bar{x}}\gamma$ then $\tau(\alpha) = \sa{\bar{x}}\sa{X}\gamma$; if $\beta$ it is in one of the remaining valid forms in (1), then $\tau(\alpha) = \alpha$.
-If $\alpha = \sa{x}\beta$ ($\alpha = \sa{X}\beta$) and $\beta = (\gamma_1\add\gamma_2)$, then $\tau(\alpha) = \tau(\sa{x}\gamma_1) \add \tau(\sa{x}\gamma_2)$ ($\tau(\alpha) = \tau(\sa{X}\gamma_1) \add \tau(\sa{X}\gamma_2)$).
-This covers all possible cases for $\alpha$.
-Furthermore, suppose that $\LL$ is closed under conjunction and suppose $\alpha = (\beta\mult\gamma)$.
-If $\beta = (\delta_1\add\delta_2)$ (wlog), then $\tau(\alpha) = \tau(\delta_1\mult\gamma)\add\tau(\delta_2\mult\gamma)$.
-If $\beta = \sa{\bar{x}}\delta$, then $\tau(\alpha) = \tau(\sa{\bar{x}}\tau(\delta\mult\gamma))$.
-If $\beta = 1$, then $\tau(\alpha) = \gamma$.
-If $\beta = \varphi$ and $\gamma = \psi$, then $\tau(\alpha) = \varphi\wedge\psi$.
-Let $\alpha$ be a $\eqso(\loge{0})$ formula over a signature $\R$.
-Consider the following procedure that receives $\enc(\A)$ as input, where $\A\in\ostr[\R]$ with domain $A$.
-Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_m$.
-For each $\beta_i$, if $\beta_i = 1$ or $\beta_i = \varphi$, then $\sem{\beta_i}(\A)$ can be computed in polynomial time.
-If $\beta_i = \sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}$, if $\beta_i = \sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = 2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, if $\beta_i = \sa{\bar{X}}\sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, all of which can be computed in polynomial time.
-Otherwise, $\beta_i$ is $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, and then $\sem{\beta_i}(\A) = f_{\varphi}(\enc(\A))$, and $f_{\varphi}\in \#\loge{0}$ which implies that $f_{\varphi}(\enc(\A))$ can be computed in polynomial time \cite{SalujaST95}.
-
-\subsection{Proof of Theorem \ref{eqso-sigma-one-in-eqso-pi-one}}
-
-Let $\alpha$ be a formula in $\eqso(\loge{1})$ over a signature $\R$.
-Let $\tau(\alpha) = \beta_1\add\beta_2\add\cdots\add\beta_m$ defined as in the proof of Theorem \ref{eqso-sigma-zero-in-fp}.
-This is, each $\beta_i$ is of the form $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}\sa{\bar{x}}1$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, where $\varphi$ is a formula in $\loge{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables.
-For each $\beta_i$ we will construct a $\gamma_i$ such that $\gamma_i \in \eqso(\logu{1})$ and $\sem{\beta_i}(\A) = \sem{\gamma_i}(\A)$ for each $\A\in\ostr[\R]$.
-Since the forms $1$, $\sa{\bar{x}}1$, $\sa{\bar{X}}1$ and $\sa{\bar{X}}\sa{\bar{x}}1$ are already valid in $\eqso(\logu{1})$, we suppose $\beta_i$ is in one of the remaining forms.
-Without loss of generality, let $\beta_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y},\bar{X})$ where $\varphi(\bar{x},\bar{y},\bar{X})$ is a quantifier-free formula.
-Similarly as in the proof by Saluja et al.
-\cite{SalujaST95}, we define $\gamma_i = \sa{\bar{x}}\sa{\bar{y}}\sa{\bar{X}}\varphi(\bar{x},\bar{y},\bar{X}) \wedge \forall\bar{z}(\varphi(\bar{x},\bar{z},\bar{X})\to(\bar{z} < \bar{y} \vee \bar{z} = \bar{y}))$ that fulfils the desired condition.
-The theorem holds true by using $\beta = \gamma_1\add\cdots\add\gamma_m$ which is in $\eqso(\logu{1})$
-
-%\subsection{Proof of Theorem \ref{one-sigma-zero}}
-%
-%Towards a contradiction, suppose that there exists a constant function $f(\A) = s > 0$ which is expressible in $\E{0}$.
-This is, there is a formula $\psi(x,X)$ such that $f_{\psi(x,X)}(\A) = \vert \{ \langle e,P\rangle \mid \A\models\psi(e,P) \} \vert = s$ for each $\A\in\ostr$.
-It can be seen that if $X$ contains more than zero variables, then the function $f_{\psi(x,X)}$ grows exponentially.
-We have now that $\psi(x)$ contains only first order variables.
-Consider some $\A \in \ostr$ of size $n$ and let $a_1,\ldots,a_s\in \A$ be the tuples that satisfy $\A \models \psi(a_i)$.
-Now consider a structure $\A'$ that is obtained by duplicating $\A$.
-This is, $\A'$ has size $2n$, its first half is isomorphic to its second half, and also to $\A$.
-We have that there is at least $2s$ tuples $a_i,b_i\in \A'$, $s$ on each half, such that $\A'\models\psi(a_i)$ and $\A'\models\psi(b_i)$, which leads to a contradiction.
-
-%\textbf{Proof of Theorem \ref{mult-sigma-zero}}
-%
-%Let $\alpha \in \sqso(\loge{0})$ with binary product.
-This is, $\alpha$ is defined by the grammar $\alpha := \sa{x}\varphi \ \mid \ \sa{X}\varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \ (\alpha\mult\alpha)$, where $\varphi \in \loge{0}$.
-For convenience, we add the expressions $\sa{x}(\varphi\mult\varphi)$ and $\sa{X}(\varphi\mult\varphi)$ to the grammar.
-Let $k$ be the number of algebraic operators in $\alpha$.
-We will show by induction on $k$ that $\alpha$ is expressible without multiplication.
-If $k = 1$, then the possible cases are $\alpha = \sa{x}\varphi$, where $\varphi\in\loge{0}$, which is already product-free.
-If $k = 2$, then the only that is not two nested summations is $\alpha = \sa{x}(\varphi\mult\psi)$, which can be expressed as $\alpha = \sa{x}(\varphi\wedge\psi)$.
-Suppose that the hypothesis holds for $k = 1$ through $n-1$.
-Let $\alpha$ have $n$ algebraic operators.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ where $\beta$ is an algebraic expression, then $\beta$ has $n-1$ algebraic operators and by hypothesis it can be expressed as product-free.
-If $\alpha = (\beta \mult \gamma)$ there are two cases.
-If at least one between $\beta$ and $\gamma$ has a product, then by hypothesis it can be expressed as product-free.
-Moreover, the new expression for $\alpha$ has strictly less than $n$ algebraic operators so by hypothesis it can be expressed as product-free.
-If there are no products in $\beta$ or $\gamma$, then let $\beta = \sa{x}\delta(x)$.
-We have that $\alpha$ can be expressed as $\sa{y}(\delta(y)\mult\gamma)$, where $y$ is a first-order variable that does not appear in $\beta$ or $\gamma$.
-Since $\delta(y)\mult\gamma$ has $n-1$ algebraic operators, it can be expressed as product-free, and then $\alpha$ can be expressed as product-free.
-%
-%\textbf{Proof of Theorem \ref{fo-prod-sigma-zero}}
-%
-%The resulting grammar is $\alpha := \sa{x}\varphi \ \mid \ \sa{X} \varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \pa{x}\alpha$.
-Consider a more restricted grammar that replaces $\pa{x}\alpha$ by $\pa{x}\sa{y}\varphi$.
-Let $\alpha = \pa{x}\sa{y}(x = y)$.
-This formula produces a constant function equal to 1, which as shown in Theorem \ref{one-sigma-zero} is not expressible in $\E{0}$.
-%
-%\textbf{Proof of Theorem \ref{fp1}}
-%
-%Let $f \in \XE{0}$, and let $\varphi(\X,\x)$ be and {\sc FO}-extended quantifier-free $\L$-formula such that:
-%\begin{eqnarray*}
-%	f(\A) &=& \vert \{\langle \P,\e  \rangle \mid \A \models \varphi(\P,\e) \} \vert
-%\end{eqnarray*}
-%for each $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^m$ and $\P = (P_1,\ldots,P_q)$ is a tuple of predicates.
-We will now show that computing $f(\A)$ can be done in polynomial time.
-%
-%For each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$, let $R_{\beta}$ be a predicate of arity $m$.
-Let $\A^\prime = \langle A, \S^{\A}, R_{\beta}^{\A^\prime}, \leq^{\A} \rangle \in \str$, where $R_{\beta}^{\A^\prime} = \{\d \mid \A \models \beta(\d)\}$.
-Note that each $\beta(\x)$ is fixed in $\varphi(\X,\x)$, and for each $\d \in A^m$, checking whether $\A \models \beta(\d)$ can be done in polynomial time.
-Therefore, generating $R_{\beta}^{\A^\prime}$ can also be done in polynomial time.
-%
-%Let $\psi(\X,\x)$ be obtained by replacing each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$ by $R_{\beta}(\x)$.
-Also, let $g = f_{\psi(\X,\x)}$.
-Note that for each tuple of predicates $\P$ and each $\e \in A^m$, $\A \models \varphi(\P,\e)$ if and only if $\A^\prime \models \psi(\P,\e)$, and so, $g(\A^\prime) = f(\A)$.
-But, $\psi(\X,\x)$ is a quantifier-free $\L$-formula, and therefore, $g \in \#\Sigma_0$.
-Since it is shown in \cite{DBLP:journals/jcss/SalujaST95} that $\#\Sigma_0 \subseteq$ {\sc FP}, we conclude that $f(\A)$ can be evaluated in polynomial time.
-
-\textbf{Proof of Theorem \ref{decisionptime}}
-
-Let $f$ be a function in $\E{1}$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
-\begin{eqnarray*}
-	f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables.
-Let $\y = (\x,\z)$ and let $\psi(\X,\y) = \varphi(\x,\X,\z)$.
-Moreover, let $g = f_{\psi(\X,\y)}$.
-\begin{claim} \label{mono}
-	For each $\A \in \str$, $f(\A) > 0$ iff $g(\A) > 0$.
-\end{claim}
-\begin{proof}
-	($\Rightarrow$) Suppose $f(\A) > 0$.
-Let $\P$ and $\e$ be such that $\A \models \exists \x \, \varphi(\x,\P,\e)$.
-It follows that there is at least one $\d \in A^k$ such that $\A \models \varphi(\d,\P,\e) = \psi(\P,(\d,\e))$.
-Therefore, $g(\A) > 0$.
-
-	($\Leftarrow$) Suppose $g(\A) > 0$.
-Let $\Q$ and $\c = (\c_1,\c_2)$, where $\c_1$ and $\c_2$ have $k$ and $m$ elements respectively, be such that $\A \models \psi(\Q,\c) = \varphi(\c_1,\Q,\c_2)$.
-Then we have that $\A \models \exists \x \, \varphi(\x,\Q,\c_2)$, from which we conclude that $f(\A) > 0$.
-\end{proof}
-Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula, so $g \in\XE{0}$.
-It follows from \ref{mono} that for each $\A \in \str$, it holds that $\A \in L_f$ if and only if $g(\A) > 0$.
-But by Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time.
-From this we conclude that the decision version of $f$ is in $\ptime$.
-
-\textbf{Proof of Theorem \ref{sub-pnp}}
-
-Suppose that $\E{1}$ is closed under substraction, that is, for each pair of functions $f,g\in \E{1}$, there exists $h\in\E{1}$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\str$.
-
-Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\Phi$, where $A$ is the set of variables mentioned in $\Phi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{1,2,3,4\}$:
-\begin{eqnarray*}
-	S_1^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-	S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-	S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-	S_4^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
-\end{eqnarray*}
-Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
-\begin{multline*}
-\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
-\end{multline*}
-Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
-$$
-\varphi(x,X) = (T(x) \vee \neg T(x)).
-$$
-Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Given that $f_{\#3DNF}, f_{all} \in \E{1}$, we have by our initial assumption that $f_{all}-f_{\#3DNF} \in \E{1}$.
-Let $h\in\E{1}$ be such that $h = f_{all}-f_{\#3DNF}$.
-For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is $\conp$-complete.
-However, as we showed previously in Theorem \ref{decisionptime}, since $h\in\E{1}$, we have that $L_h \in \ptime$.
-Then, $\conp \subseteq \ptime$, from which we conclude that $\ptime = \np$.
-
-%\textbf{Proof of Theorem \ref{sigma1strict}}
-%
-%We will show that the $\E{1}$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
-By contradiction, suppose that it is.
-Let $\A = \langle A =\{1\},S^\A = \{1\},\leq^\A=\{(1,1)\} \rangle$.
-Then, $f(\A) = 1$.
-Now let $\B = \langle B =\{1,2\},S^\B = \{1\},\leq^\B=\{(1,1),(1,2),(2,2)\} \rangle$.
-Note that $\A$ is an induced substructure of $\B$.
-%
-%We have that for each function $g \in \#\Sigma_1$ and structures $\A_1,\A_2\in\str$, if $\A_1$ is an induced substructure of $\A_2$, then $g(\A_1) \leq g(\A_2)$ \cite{SalujaST95}.
-Therefore, $f(\B) \geq f(\A) = 1$.
-However, there is no assignment $s\in B$ to $x$ such that $\B\models\:\varphi(s)$, so $f(\B) = 0$, which leads to a contradiction.
-
-\subsection{Proof of Theorem \ref{sigmafo-minusone}}
-
-Let $\alpha$ be a $\eqso(\logex{1})$ formula over a signature $\R$.
-We will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_n$ be defined as in \ref{eqso-sigma-zero-in-fp}, that is, each $\beta_i$ is equal to $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}\sa{\bar{x}}1$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, where $\varphi$ is a formula in $\logex{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables.
-We further assume that each $\beta_i$ is either $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, since each $1$, $\varphi$ and $\sa{\bar{X}}\varphi$ can be replaced by the formulas $\sa{x}\min(x)$, $\sa{x}(\varphi\wedge\min(x))$ and $\sa{x}\sa{\bar{X}}(\varphi\wedge\min(x))$ respectively, while maintaining the value of $\sem{\alpha}(\A)$.
-
-The intuition behind our reasoning is separated in three points: (1) For each $\beta_i$ of the form $\sa{\bar{x}}\varphi$, the formula $\kappa(\beta_i)$ will count every tuple $\bar{x}$ that satisfies $\varphi$ except for the lexicographically smallest one.
-(2) For each $\beta_i$ of the form $\sa{\bar{X}}\sa{\bar{x}}\varphi$, the formula $\kappa(\beta_i)$ will isolate the smallest $\bar{X}$ that satisfies $\varphi$, and exclude the lexicographically smallest tuple $\bar{x}$ that satisfies $\varphi(\bar{X})$.
-And (3) if $\alpha = (\beta\add\sa{\bar{x}}\varphi)$ or $\alpha = (\beta\add\sa{\bar{X}}\sa{\bar{x}}\varphi)$, the formula $\kappa(\alpha)$ will exclude the lexicographically smallest tuple that satisfies $\varphi$ if and only if $\beta$ is equal to 0.
-
-\vspace{1em}
-
-For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha) = \sa{\bar{x}}\exists\bar{y}[\varphi(\bar{x},\bar{y})\wedge\exists\bar{z}(\varphi(\bar{z},\bar{y})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfils the desired condition.
-
-\vspace{1em}
-
-For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha)$ procedurally as follows: Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
-We suppose that $\varphi$ is in a DNF form that leaves formulas that do not mention $\bar{X}$ as is.
-If it is not, we convert $\varphi$ to this form with a standard transformation algorithm.
-Let $\varphi(\bar{X},\bar{x},\bar{y}) = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$ where each $\varphi_i$ has the form:
-$$
-\varphi_i(\bar{X},\bar{x},\bar{y}) = (\text{conjunction of $X_j$'s}) \wedge (\text{conjunction of $\neg X_j$'s})  \wedge (\text{an $\fo$ formula that does not mention any $X_j$}).
-$$
-Define $\varphi_i^{+}$, $\varphi_i^{-}$ and $\varphi_i^{\fo}$ as the formulas mentioned above.
-
-In our procedure we assume that in $\varphi_i^{+}\wedge\varphi_i^{-}$ every first-order variable from $\bar{x},\bar{y}$ is mentioned at most once.
-If not, we add new first-order variables from $\fv$ to $\bar{y}$ so that no variable is repeated, replace them in $\varphi_i^{+}\wedge\varphi_i^{-}$ and add their respective equalities in $\varphi_i^{\fo}$ so that the new formula is equivalent.
-For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
-
-Furthermore, we assume that $\varphi_i^{\fo}$ defines an ordered partition of the variables in the tuple $(\bar{x},\bar{y})$.
-For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$.
-A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
-On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
-For a given $k$, we define $\cB_k$ as the number of possible ordered partitions for a set of size $k$. 
-If $\varphi_i^{\fo}$ does not define an ordered partition, then for $1 \leq j \leq \cB_{\length(\bar{x},\bar{y})}$ 
-let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
-The formula $\varphi_i(\bar{X},\bar{x},\bar{y})$ is converted into $\bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}}\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \theta^j(\bar{x},\bar{y})$.
-Consider each of these disjuncts as a new $\varphi_i$.
-
-After the previous assumptions, we do the following: for each $X_j\in \bar{X}$ we check every instance of $X_j(\bar{w})$ in $\varphi^{+}_i$ and every instance of $X_j(\bar{z})$ in $\varphi^{-}_i$, where $\bar{w}$ and $\bar{z}$ are subtuples of $(\bar{x},\bar{y})$.
-If the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{w} = \bar{z}$, the entire formula $\varphi_i$ is removed from $\varphi$.
-
-Now let $\bar{v}$ be the tuple of every first-order variable mentioned in $\varphi_i^{+}$ and let $\bar{u}$ be such that $(\bar{x},\bar{y}) = (\bar{u},\bar{v})$.
-We define a formula $\mu_i$ that is satisfied only by the lexicographically smallest $\bar{v}$ that satisfies $\varphi_i^{\fo}$ for some $\bar{u}$:
-$$
-\mu_i(\bar{v}) = \exists\bar{u}\,\varphi_i^{\fo}(\bar{u},\bar{v})\wedge\forall\bar{u}'\forall\bar{v}'(\varphi_i^{\fo}(\bar{u}',\bar{v}')\to\bar{v}\leq\bar{v}').
-$$
-The following is a pivotal formula for this proof.
-Consider some ordered finite structure $\A$ over $\R$.
-Given our conditions for $\varphi_i$, if $\A\models\varphi_i$ and $\A\models\varphi^{\fo}_i(\bar{u},\bar{v})$ then for that $\bar{v}$ we can define a $\bar{X}$ that has exactly the tuples mentioned in $\bar{v}$ and nothing else.
-This is {\em the smallest $\bar{X}$ that satisfies $\varphi_i$ over $\bar{v}$}.
-And if $\bar{v}$ is the lexicographically smallest such that $\A\models\varphi_i(\bar{v})$, then it is the {\em the smallest $\bar{X}$ that satisfies $\varphi_i$}.
-The following formula is satisfied by every pair $(\bar{X},\bar{x})$, except for the one formed by the smallest $\bar{X}$ that satisfies $\varphi_i$ and the lexicographically smallest $\bar{x}$ that satisfies $\varphi_i(\bar{X})$.
-And if no pair satisfies $\varphi_i$, the following formula is satisfied by every pair.
-$$
-\psi_i(\bar{X},\bar{x}) = \exists\bar{v}(\mu_i(\bar{v})\wedge(\neg\varphi^{+}_i(\bar{X},\bar{v})\vee\bigvee_{X \in \bar{X}} \exists\bar{z}(X(\bar{z}) \wedge \bigwedge\limits_{\substack{\text{instances of }X(\bar{w}) \\ \text{in }\varphi^{+}_i(\bar{X},\bar{v})}}\bar{w}\neq\bar{z}))) \vee \exists\bar{x}'(\exists\bar{y}'\varphi_i(\bar{X},\bar{x}',\bar{y}')\wedge \bar{x}'<\bar{x})\vee\neg\exists\bar{v}\mu_i(\bar{v}) .
-$$
-\begin{claim}
-For a given ordered structure $\A$ such that $\A\models\exists\bar{X}\exists\bar{x}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, there is an assignment $(\bar{B},\bar{b})$ to $(\bar{X},\bar{x})$ that satisfies the following conditions (1) $\A\models\exists\bar{y}\,\varphi_i(\bar{B},\bar{b},\bar{y})$, (2) $\A\not\models\psi_i(\bar{B},\bar{b})$ and (3) this is the only assignment to $(\bar{X},\bar{x})$ that satisfies (1) and (2).
-\end{claim}
-\begin{proof}
-	proof here.
-\end{proof}
-The following claim identifies a crucial condition that is verifiable through a $\fo$. 
-\begin{claim}
-	Let $\varphi_i(\bar{X},\bar{x},\bar{y}) = \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X},\bar{x},\bar{y})$ be a $\logex{0}$ formula that satisfies all the mentioned assumptions.
-For a given ordered structure $\A$ it holds that $\A\models\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$ iff $\A\models\exists\bar{X}\exists\bar{x}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$.
-\end{claim}
-\begin{proof}
-	Let $\A$ be an ordered structure with domain $A$ and let $\bar{a}\in A^{\length(\bar{y})}$ and $\bar{b}\in A^{\length(\bar{x})}$ such that $\A\models\varphi^{\fo}_i(\bar{b},\bar{a})$.
-Define $\bar{B} = (B_1,\ldots,B_{\length(\bar{X})})$ as $B_j = \{\bar{c}\mid\bar{c}\text{ is a subtuple of $(\bar{b},\bar{a})$ and $X_j(\bar{c})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{b},\bar{a})$}\}$.
-Towards a contradiction, suppose that $\A\not\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
-By the choice of $\bar{a}$ and $\bar{b}$, and construction of $\bar{B}$ it is clear that $\A\models\varphi^{\fo}_i(\bar{b},\bar{a})\wedge\varphi^{+}_i(\bar{B},\bar{b},\bar{a})$, so we have that $\A\not\models\varphi^{-}_i(\bar{B},\bar{b},\bar{a})$.
-Let $B_t(\bar{c})$ be such that $X_t(\bar{c})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{b},\bar{a})$ and $\A\not\models\neg B_t(\bar{c})$.
-This is, $\bar{c}\in B_t$.
-By the construction of $\bar{B}$ we have that $X_t(\bar{c})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{b},\bar{a})$.
-From our assumptions, we have that (1) every first-order variable is mentioned at most once in $\varphi^{-}_i(\bar{X},\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X},\bar{x},\bar{y})$, (2) $\varphi^{\fo}_i$ defines an ordered partition over $(\bar{x},\bar{y})$, and (3) there are no $X_j(\bar{w})$ in $\varphi^{+}_i$ and $X_j(\bar{z})$ in $\varphi^{-}_i$ such that the ordered partition defined by $\varphi^{\fo}_i$ satisfies $\bar{x} = \bar{z}$.
-Since $\bar{c}$ is a subtuple of $(\bar{b},\bar{a})$ that satisfies $\varphi^{\fo}_i$, and $X_t(\bar{w})$ mentioned in $\varphi^{+}_i$ and $X_t(\bar{z})$ mentioned in $\varphi^{-}_i$ were both assigned the value $\bar{c}$, then the ordered partition satisfies $\bar{w} = \bar{z}$, which follows to a contradiction.
-\end{proof}
-We define $\chi_i = \exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$.
-Now recall that $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$, where each $\varphi_i$ satisfies all the previous assumptions.
-For each $\varphi_i$ we define:
-$$
-\varphi_i'(\bar{X},\bar{x},\bar{y}) = \varphi_i(\bar{X},\bar{x},\bar{y})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\vee\psi_2(\bar{X},\bar{x}))\wedge(\chi_1\vee\chi_2\vee\psi_2(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigvee_{j = 1}^{j = i-1}\chi_j\vee\psi_i(\bar{X},\bar{x})),
-$$
-and lastly, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x},\bar{y})$.
-
-\vspace{1em}
-
-For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y}))$ for some algebraic formula $\beta$ and some quantifier-free formula $\varphi$, we define $\kappa(\alpha)$ as follows: First, perform the same transformations to $\varphi(\bar{X},\bar{x},\bar{y})$ as in the previous case.
-Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$ where each $\varphi_i$ satisfies the previous assumptions.
-We also use the previously defined formulas $\chi_i$ and $\psi_i$.
-
-We recursively define a function $\lambda$ as follows.
-If $\alpha = \sa{\bar{x}}\exists\bar{y}\varphi(\bar{x},\bar{y})$, then $\lambda(\alpha) = \exists\bar{x}'\exists\bar{y}'\varphi(\bar{x}',\bar{y}')$.
-If $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ and $\lambda(\alpha) = \chi_1\vee\cdots\vee\chi_m$ as each $\chi_i$ was previously defined.
-If $\alpha = (\beta + \gamma)$, then $\lambda(\alpha) = \lambda(\beta) \vee \lambda(\gamma)$.
-Note that for a given ordered structure $\A$, then $\A\models\lambda(\alpha)$ if and only if $\sem{\alpha}(\A) > 0$.
-
-Now, for each $\varphi_i$ we define:
-$$
-\varphi_i'(\bar{X},\bar{x},\bar{y}) = \varphi_i(\bar{X},\bar{x},\bar{y})\wedge[\lambda(\beta)\vee[\psi_1(\bar{X},\bar{x})\wedge(\chi_1\vee\psi_2(\bar{X},\bar{x}))\wedge(\chi_1\vee\chi_2\vee\psi_2(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigvee_{j = 1}^{j = i-1}\chi_j\vee\psi_i(\bar{X},\bar{x}))]].
-$$
-And lastly $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x},\bar{y})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
-
-%%Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$.
-There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$.
-This separates the proof in three cases:
-%\begin{enumerate}
-%%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
-%%	\begin{eqnarray*}
-%%		f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
-%%	\end{eqnarray*}
-%%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^d$.
-Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily.
-First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
-%%	\begin{eqnarray*}
-%%		\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
-%%	\end{eqnarray*}
-%%	This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
-%%	\begin{eqnarray*}
-%%		\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
-%%	\end{eqnarray*}
-%%	If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
-Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$.
-Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E{1}.$
-%%	
-%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
-%	\begin{eqnarray*}
-%		f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
-%	\end{eqnarray*}
-%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
-%	\begin{eqnarray}
-%	\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
-%	\end{eqnarray}
-%	where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
-Note that $\theta(\x)$ also mentions all variables in $\x$.
-We also assume that $(\x_1,\dots,\x_n,\y) = \x$.
-As an example, the following formula is of this form:
-%	\begin{multline*}
-%	\varphi(\x,\X) =  X_1(x_1,x_2) \wedge X_3(x_3) \wedge X_2(x_4,x_5) \wedge X_3(x_6) \wedge \neg X_1(x_7,x_8) \wedge \\ (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 ) \wedge \\ \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big),
-%	\end{multline*}
-%	where $\x = (x_1,x_2,x_3,x_4,x_5,x_6,x_7,x_8)$ and $\X = (X_1,X_2,X_3)$.
-Here, $n = 4$, $\lambda(1) = 1$, $\lambda(2) = \lambda(4) = 3$ and $\lambda(3) = 2$, $\x_1 = (x_1,x_2)$, $\x_2 = (x_3)$, $\x_3 = (x_4,x_5)$, $\x_4 = (x_6)$ and $\y = (x_7,x_8)$.
-Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_7,x_8)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 )$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2,x_5,x_8\},\{x_6,x_7\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big)$.
-%	
-%	Similarly to the previous proof, we would like to eliminate the {\em lexicographically smallest}\footnote{We consider the lexicographically smallest tuple of predicates as the one in which its predicates contain the lexicographically smallest tuples and do not contain any more tuples than those} tuple of predicates that satisfies the formula \eqref{f1}.
-Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
-We now define
-%	\begin{multline*}
-%	\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to (\u_1,\dots,\u_n)<(\v_1,\dots,\v_n)\Big)\Big],
-%	\end{multline*}
-%	where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
-Note that $\alpha_{\min}$ is satisfied only by the lexicographically smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
-Our new formula is
-%	\begin{multline}
-%	\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge \\ \exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg].
-%	\end{multline}
-%	We now show a result by which the main proof will follow.
-%	\begin{lemma} \label{first}
-%		$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
-%	\end{lemma}
-%	\begin{proof}
-%		Let $\A \in \str$.
-Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j\in[1,n]:\:\lambda(j)=i}\{\d_j\}$.
-We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
-%		\begin{enumerate}
-%			\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$.
-That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$).
-Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$.
-Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$.
-We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \z<\x_i$ or (3) $\theta(\x)\models \x_i<\z$.
-Considering that (2) and (3) are not possible given that both $\z$ and $\x_i$ are assigned the value $\d_i$ and $\A\models\theta(\d)$, we have that $\theta(\x)\models \z = \x_i$.
-But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$, which means $\varphi(\x,\X)$ is inconsistent.
-That is, there's no structure $\A^\prime\in\str$ such that $\A^\prime\models\exists\x\:\varphi(\x,\R^\prime)$ for any assignment $\R^\prime$ to $\X$.
-In particular, $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which contradicts the initial assumption that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-%			\item[(b)] Note that if $\A \models \alpha_{\min}(\c_1,\ldots,\c_n)$, then necessarily $\c_i = \d_i$ for $i\in\{1,\ldots,n\}.$ However, by the construction of $\P$, we see that 
-%			$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \d_j\Big).$$ Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
-%			\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$.
-Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $P^\prime_i$.
-Then $\A\models \neg P^\prime_i(\d_i)$, and so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $P_i$ does not have.
-If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-On both cases, we have a contradiction.
-%		\end{enumerate}
-%		With this, we conclude that for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-%		
-%		Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi(\x,\P)$, we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
-And so, for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ it holds that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1,$ which was to be shown.
-%	\end{proof}
-%	We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
-By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\gamma_i(\x,\X)$ with $i\in\{1,\ldots,\ell\}$ such that
-%	\begin{eqnarray*}
-%		\varphi(\x,\X) &\equiv& \gamma_1(\x,\X) \vee \gamma_2(\x,\X) \vee \dots  \vee \gamma_{\ell}(\x,\X),
-%	\end{eqnarray*}
-%	where, for every $i\in\{1,\ldots,\ell\}$, 
-%	\begin{eqnarray*}
-%		\gamma_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \gamma^{-}_i(\X,\y_i)  \wedge \delta_i(\x).
-%	\end{eqnarray*}
-%	The function $\lambda_i$ is defined analogously to $\lambda$ of the first part of the proof.
-The tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y_i) = \x.$ The formulas $\gamma^{-}_i(\X,\y_i)$, $\delta_i(\x)$ are defined analogously to $\varphi^{-}(\X,\y)$ and $\beta(\x)\,$\footnote[2]{Note that $\delta_i(\x)$ can include subformulas of the form $(\u = \u)$.} respectively (see \eqref{f1}).
-Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables.
-Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$.
-Note that for every formula $\eta(\x)$,
-%	\begin{eqnarray*}
-%		\exists\x\:\eta(\x) &\equiv& \exists\x(\eta(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\eta(\x)\wedge\theta^{g(d)}(\x)).
-%	\end{eqnarray*}
-%	We define the following formulas $\xi_i(\x,\X)$, for $i\in\{1,\ldots,m\}$ where $m = \ell \cdot g(d)$, as follows: 
-%	\begin{eqnarray*}
-%		\xi_1(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^1(\x), \\
-%		& \vdots & \\
-%		\xi_{g(d)}(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
-%		\xi_{g(d)+1}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^1(\x), \\
-%		& \vdots & \\
-%		\xi_{2\cdot g(d)}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
-%		& \vdots & \\
-%		\xi_{(\ell-1) g(d)+1}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^1(\x), \\
-%		& \vdots & \\
-%		\xi_{\ell \cdot g(d)}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^{g(d)}(\x).
-%	\end{eqnarray*}
-%	Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\str.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
-We use an injective\footnote[3]{We need this function to be injective because this way we can assure each one of the $k$ satisfiable disjuncts to be represented} function $\rho:\{1,\ldots,k\}\to\{1,\ldots,m\}$ such that $\varphi_i(\x,\X) = \xi_{\rho(i)}(\x,\X)$ is satisfiable, for every $i\in\{1,\ldots,k\}.$
-%	
-%	We can conclude that
-%	\begin{eqnarray*}
-%		\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X),
-%	\end{eqnarray*}
-%	and for every $i\in\{1,\ldots,k\}$,
-%	\begin{eqnarray*}
-%		\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
-%	\end{eqnarray*}
-%	where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\y_i$ is a $p_i$-tuple of variables in $\x$, $\varphi_i^{-}(\X,\y_i)$ is a conjunction of negated predicates in $\X$, $\theta_i(\x)$ defines a total order on a partition of $\x$, and $\beta_i(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
-Let $\alpha_i(\x) = \theta_i(\x)\wedge\beta_i(\x).$
-%	
-%	%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
-%	%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
-%	
-%	\begin{claim} \label{alpha}
-%		Let $\A\in\str$ and $i\in\{1,\ldots,k\}$. For every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, it holds that $f_{\varphi_i(\s,\X)}(\A) > 0.$
-%	\end{claim}
-%	\begin{proof}
-%		Let $\s = (\s_1,\ldots,\s_{n_i},\t)$ such that $\A\models\alpha_i(\s)$, and let $\P = (\P_1,\ldots,\P_r)$ where $P_i = \bigcup_{j\in[1,n_i]:\:\lambda(j)=i}\{s_i\}$.
-We will show that $\A\models\varphi_i(\s,\P)$.
-By contradiction, suppose that $\A\not\models\exists\x\,\varphi_i(\x,\P)$.
-That is, there is no assignment $\e$ to $\x$ such that $\A\models\varphi(\e,\P).$ Following the same proof as in case (a) of Lemma \ref{first}, we conclude that $\varphi_i(\x,\P)$ is inconsistent.
-However, as we mentioned previously, all of such disjuncts have been eliminated, which leads to a contradiction.
-%	\end{proof}
-%	
-%	Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\exists\x\:\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\Q$ such that $\A\models\exists\x\:\varphi_2(\x,\Q)$, and so on.
-As we already know how to exclude that assignment in the first disjunct, we will now deal with the next disjuncts.
-Similarly to the first part of the proof, for each $i\in\{1,\ldots,k\}$ let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
-%	\begin{multline*}
-%	\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_{n_i},\y)\wedge \\ \forall\v_1\cdots\forall\v_{n_i}\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_{n_i},\w)\wedge\bigvee_{j=1}^{n_i}(\u_j\neq\v_j)\big)\to (\u_1,\dots,\u_{n_i})<(\v_1,\dots,\v_{n_i})\Big)\Big],
-%	\end{multline*}
-%	where $\u_j$ and $\v_j,$ have $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, and $\w$ has $p_i$ variables. Also, let
-%	\begin{multline*}
-%	\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge\limits_{\substack{\ell\in[1,n_i]:\\ \lambda_i(\ell) = j}} \v \neq \u_\ell\Big) \bigg) \bigg].
-%	\end{multline*}
-%	Note that $\psi_i(\X)$ excludes only the lexicographically smallest tuple of predicates $\P$ such that $\A\models\exists\x\:\varphi_i(\P,\x),$ if there is at least one.
-In other words, every assignment $\P^\prime \neq \P$ is such that $\A\models\psi_i(\P^\prime).$ Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
-%	$$
-%	\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
-%	$$
-%	Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Assume that $\A\in\str$.
-Suppose first that there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least one assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\d = (\d_1,\dots,\d_{n_q},\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha_q(\d)$, where $\d_i$ is the corresponding assignment to $\x_i$, for every $i\in\{1,\dots,n_q\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j:\lambda_q(j)=i,j\in[1,n_q]}\{\d_j\}$.
-As we did in Lemma \ref{first} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies both (a) and (b)
-%	\begin{enumerate}
-%		\item[(a)] As we showed in part (a) of Lemma \ref{first}, if there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_q(\x,\R)$, then $\A\models\exists\x\:\varphi_q(\x,\P)$ for this particular $\P$.
-However, as we showed in Claim \ref{alpha}, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_q(\s)$, then there {\em is} such an assignment to $\X$.
-The assignment $\d$ to $\x$ satisfies that $\A\models\alpha_q(\d)$, so it holds that $\A\models\exists\x\:\varphi_q(\x,\P)$.
-It immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
-%		
-%		\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
-(1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
-(2) We can use the proof in Lemma \ref{first} to see that $\A\not\models\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula
-%		$$ \zeta_q(\X) =  (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{q-1}(\v)\vee\psi_q(\X)) $$ 
-%		appears as a conjunct in $\varphi^\prime_i(\x,\X)$.
-However, by the choice of $q$, there is no $i\in\{1,\ldots,q-1\}$ such that $\A\models\exists\v\:\varphi_i(\v,\P)$, and also $\A\not\models\psi_q(\P)$.
-It follows that $\A\not\models\zeta_q(\P)$,  so $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$.
-And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
-%		
-%		\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
-As we deduce from the part (c) of Lemma \ref{first}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
-Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x)$, so $\A\models\psi_i(\P^\prime)$ for each $i$.
-But as we mentioned, also $\A\models\:\psi_q(\P^\prime),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which leads to a contradiction.
-%	\end{enumerate}
-%	
-%	With this, we conclude that for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0$, we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-%	
-%	Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\chi(\x,\P))$ for any formula $\chi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
-And so, for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Hence, from the results in the previous paragraph, if $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}$, we have that $f^\prime = f\dotminus 1\in\E{1}$
-%	
-%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
-%	\begin{eqnarray*}
-%		f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
-%	\end{eqnarray*}
-%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$.
-In order to prove that $f\dotminus 1\in\E{1}$, we define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2, where
-%	\begin{eqnarray*}
-%		\varphi(\x,\X,\z) \equiv \varphi_1(\x,\X,\z) \vee \cdots \vee \varphi_k(\x,\X,\z),
-%	\end{eqnarray*}
-%	and
-%	\begin{eqnarray*}
-%		\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\w_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
-%	\end{eqnarray*}
-%	where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\ell} + c_{\ell} = a_{\ell}$ for $\ell\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, and $\w_i$ has $q_i$ variables.
-Furthermore, we have that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\w_i) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
-The formulas $\varphi^{-}_i(\X,\y_i,\w_i)$, $\beta_i(\x,\z)$, are also defined analogously.
-%	
-%	In this case, we mix both the strategies in cases 1 and 2.
-That is, we are going to {\em isolate} the lexicographically smallest tuple of predicates that satisfies the first satisfiable disjunct, and then {\em exclude} the lexicographically smallest tuple that satisfies the isolated disjunct.
-%	
-%	Let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and
-%	\begin{multline*}
-%	\alpha^i_{\min}(\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}) = \exists\y\exists\w\Big[ \alpha_i(\x_1,\dots,\x_{n_i},\y,\z_1,\dots,\z_{n_i},\w)\wedge \\ \forall\u_1\cdots\forall\u_{n_i}\forall\s\forall\v_1\cdots\forall\v_{n_i}\forall\t\Big(\big(\alpha_i(\u_1,\dots,\u_{n_i},\s,\v_1,\dots,\v_{n_i},\t)\wedge\bigvee_{j=1}^{n_i}(\x_j\neq\u_j \vee \z_j\neq\v_j)\big)\to \\ (\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i})<(\u_1,\dots,\u_{n_i},\v_1,\dots,\v_{n_i})\Big)\Big],
-%	\end{multline*}
-%	and let
-%	\begin{multline*}
-%	\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\w(\exists\u\:\varphi(\u,\X,\w) \wedge \w<\z).
-%	\end{multline*}
-%	Let $\P$ and $\e$ be assignments to $\X$ and $\z$.
-Note that $\psi_i(\X,\z)$ excludes $\P$ and $\e$ only if $\P$ is the {\em lexicographically smallest} tuple of predicates (Same as case 2) such that $\A\models\exists\x\:\varphi_i(\x,\P,\d)$, for some assignment $\d$ to $\z$, and $\e$ is the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_i(\x,\P,\e).$ We define $\varphi_i^\prime(\x,\X,\z)$ in the same way as case 2:
-%	\begin{multline*}
-%	\varphi_i^\prime(\x,\X,\z) = \varphi_i(\x,\X,\z) \wedge \psi_1(\X,\z) \wedge (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\psi_2(\X,\z)) \wedge \cdots \wedge \\ (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\cdots\vee\exists\u\exists\v\:\alpha_{i-1}(\u,\v)\vee\psi_i(\X,\z)).
-%	\end{multline*}
-%	Finally, let $\varphi^\prime(\x,\X,\z) = \bigvee_{i = 1}^k \varphi_i^\prime(\x,\X,\z)$.
-%	
-%	Let $q$ be the least $i\in\{1,\ldots,k\}$ such that $\A\models\exists\x\:\varphi_i(\x,\R,\d)$ for some assignment $R$ to $\X$ and some assignment $\d$ to $\z$.
-Let $\P$ be the lexicographically smallest tuple of predicates such that $\A\models\exists\x\:\varphi_q(\x,\P,\d^\prime)$ for some assignment $\d^\prime$ to $\z$.
-Let $\e$ be the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_q(\x,\P,\e).$ This formula is such that (a) $\A\models\exists\x\:\varphi(\x,\P,\e)$, (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P,\e)$ and $\P$ and $\e$ are the only assignments that satisfy (a) and (b).
-The proof of this is analogous to case 2.
-Therefore, we conclude that $f_{\exists\x\,\varphi^\prime(\x,\X,\z)} = f_{\exists\x\,\varphi(\x,\X,\z)}\dotminus 1.$
-%\end{enumerate}
-%		
-%\textbf{Proof of Theorem \ref{dnf-pars}}
-%
-%Let $f$ be an arbitrary function in $\E{1}$ and let $f_{\textsc{\#DNF}}$ be the function that defines {\sc \#DNF}.
-We will now show a function $h:\str\to \textsc{Struct}[\L_{\textsc{DNF}}]$, where $\L_{\textsc{DNF}} = \{N,P\}$, such that for every $\A\in\str$, it holds that $f(\A) = f_{\textsc{\#DNF}}(h(\A))$, and which can be computed in polynomial time.
-For every DNF formula $\Phi$ we define $\A_{\Phi} \in \textsc{Struct}[\L_{\textsc{DNF}}]$ as stated in \ref{dnfdef}.
-Let $\psi$ be such that $f = f_{\psi(\X,\z)}$.
-We separate the proof in three cases:
-%\begin{enumerate}
-%	\item $\X = ()$.
-Let $\psi(\z) = \exists \x\,\varphi(\x,\z)$ where $\x = (x_1,\ldots,x_c)$ and $\z = (z_1,\ldots,z_d)$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a propositional formula $\Phi = h(\A)$.
-First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\,\varphi(\x,\e)$.
-Let $n$ be the number of such assignments to $\z$.
-Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$.
-The DNF formula we generate is
-%	\[
-%	\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
-%	\]
-%	which has exactly $n$ possible assignments. Finally, we return $h(\A) = \A_{\Phi}$.
-%	
-%	\item $\z = ()$.
-Let $\psi(\X) = \exists \x\,\varphi(\x,\X)$, where $\x = (x_1,\ldots,x_c)$ and $\X = (X_1,\ldots,X_r)$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a DNF-formula $\Phi = h(\A)$ as follows.
-First, we convert $\varphi(\x,\X)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X)$.
-Let each $\varphi_{i}(\x,\X)$ be of the form
-%	\[
-%	\varphi_{i}(\x,\X) = X_{\pi_i(1)}(\x^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\x^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\x^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\x^{\nu}_{i,n_i}) \wedge \alpha_i(\x),
-%	\]
-%	where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$) is the rearrangement of $\x$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x)$ is a FO formula.
-We also define the tuple $\a^{\pi}_{i,j}$ ($\a^{\nu}_{i,j}$) as $\a$ rearranged in the same way as $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$), for each $\a\in A^c$.
-%	
-%	We generate a propositional variable $p^{\ell}_{\a}$ for each $\a\in A^{c_{\ell}}$, and for each $\ell\in\{1,\ldots,r\}$, which will indicate if the tuple $\a$ is in the assignment to $X_{\ell}$.
-Let $\p$ be the tuple of all such variables.
-%	
-%	Then, let $\Gamma_i = \{\a\in A^c\mid \A\models\alpha_i(\a)\}$.
-There is exactly $\vert A \vert^c$ tuples to check, and each of those checks can be done in polynomial time, so this set can also be generated in polynomial time.
-%	
-%	Finally, for each $\varphi_i(\x,\X)$ the DNF formula we generate is
-%	\[
-%	\Phi_i = \bigvee_{\a\in\Gamma_i} p^{\pi_i(1)}_{\a^{\pi}_{i,1}} \wedge \cdots \wedge p^{\pi_i(m_i)}_{\a^{\pi}_{i,m_i}} \wedge \neg p^{\nu_i(1)}_{\a^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\pi_i(n_i)}_{\a^{\nu}_{i,n_i}}.
-%	\]
-%	and $\Phi = \bigvee_{i = 1}^k \Phi_i \vee \bigvee_{p\in\p} (p\wedge\neg p)$. Finally, we return $h(\A) = \A_{\Phi}$.
-%	
-%	As an example, let $\psi^{*}(X_1,X_2) = \exists x \exists y\, (X_1(x) \wedge X_2(y,x) \wedge \neg X_1(y) \wedge x < y \wedge \neg\exists z\,(x<z\wedge z<y))$.
-We will compute $h(\B)$, where $\B = \langle \{1,2,3\}, < \rangle$, and we have that $1 < 2 < 3$.
-We generate the tuple of variables $\p = (p^1_1,p^1_2,p^1_3) \cup (p^2_{(i,j)})$ for $(i,j)\in \{1,2,3\}\times\{1,2,3\}$.
-Note that in this procedure we will only have one set $\Gamma_1 = \{(1,2),(2,3)\}$.
-And so, the DNF formula we generate is
-%	\begin{multline*}
-%	\Phi^{*} = (p^1_1\wedge p^2_{(2,1)} \wedge \neg p^1_2)\vee(p^1_2\wedge p^2_{(3,2)} \wedge \neg p^1_3)\, \vee \\ (p^1_1 \wedge \neg p^1_1) \vee(p^1_2 \wedge \neg p^1_2) \vee (p^1_3\wedge \neg p^1_3) \vee \bigvee_{i,j\in\{1,2,3\}}(p^2_{(i,j)}\wedge\neg p^2_{(i,j)})
-%	\end{multline*}
-%	
-%	Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$.
-We can conclude that $f(\A) = f_{\textsc{DNF}}(\Phi)$.
-%	
-%	\item Let $\psi = \exists \x\,\varphi(\x,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$ we generate a DNF-formula $\Phi = h(\A)$ as follows.
-As in the previous case, we convert $\varphi(\x,\X,\z)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X,\z)$.
-Let each $\varphi_{i}(\x,\X,\z)$ be of the form
-%	\[
-%	\varphi_{i}(\x,\X,\z) = X_{\pi_i(1)}(\y^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\y^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\y^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\y^{\nu}_{i,n_i}) \wedge \alpha_i(\x,\z),
-%	\]
-%	where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$) is the rearrangement of $(\x,\z)$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x,\z)$ is a FO formula.
-We also define the tuple $(\a,\e)^{\pi}_{i,j}$ ($(\a,\e)^{\nu}_{i,j}$) as $\a$ rearranged in the same way as $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$), for each $(\a,\e)\in A^{c+d}$.
-%	
-%	We generate a propositional variable $p^{\e}_{\ell,\b}$ for each $\e\in A^d$, for each $\ell\in\{1,\ldots,r\}$, and for each $\b\in A^{c_{\ell}}$.
-This variable will be true if there is an assignation to $\P$ such that $\b\in P_{\ell}$ and $\A\models\psi(\P,\e)$.
-Let $\p$ be the tuple of all such variables.
-We also generate a set $T = \{t_1,\ldots,t_{L}\}$ where $L = \lceil \log\vert A^d\vert \rceil$, and an arbitrary injective function $\beta:A^d \to \{0,1\}^L$ where $\beta(\e) = (b^{\e}_1,\ldots,b^{\e}_L)$ for each $\e\in A^d$.
-Also, let $\rho_{\ell}$ be such that $\rho_{\ell}(\x,\z) = \y_{\ell}$.
-%	
-%	For each $\e\in A^d$, we follow the following procedure.
-For each $\varphi_i(\x,\X)$, we consider the set $\Gamma^{\e}_i = \{\a\in A^c\mid \A\models\alpha(\a,\e)\}$, and so, the DNF formula we generate is
-%	\[
-%	\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{\pi_i(1),(\a,\e)^{\pi}_{i,1}} \wedge \cdots \wedge p^{\e}_{\pi_i(m_i),(\a,\e)^{\pi}_{i,m_i}} \wedge \neg p^{\e}_{\nu_i(1),(\a,\e)^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\e}_{\pi_i(n_i),(\a,\e)^{\nu}_{i,n_i}} \wedge \gamma_{\e} \wedge \theta_{\e},
-%	\]
-%	where $\gamma_{\e}$ is defined as
-%	\[
-%	\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
-%	\]
-%	and $\theta_{\e}$ is defined as
-%	\[
-%	\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
-%	\]
-%	Finally,$\Phi = \bigvee_{\e\in A^d} \bigvee_{i = 1}^k \Phi_i^{\e} \vee \bigvee_{p\in\p} (p\wedge\neg p)$, and we return $f(\A) = \A_{\Phi}$.
-%	
-%	As an example, let $\psi^{*}(X_1,X_2,z) = \exists x\, (X_1(x) \wedge X_2(z) \wedge \neg X_1(z) \wedge x < z)$.
-We will compute $h(\B)$, where $\B = \langle \{1,2,3\}, < \rangle$, and we have that $1 < 2 < 3$.
-We generate the tuple of variables $\p = (p^e_{1,1},p^e_{1,2},p^e_{1,3},p^e_{2,1},p^e_{2,2},p^e_{2,3})$, for $e\in \{1,2,3\}$.
-Here, we consider $L = 2$, $T = \{t_1,t_2\}$, and $\beta$ defined as $\beta(1) = (0,0)$, $\beta(2) = (0,1)$ and $\beta(3) = (1,0)$.
-In this procedure we generate the sets $\Gamma^1_1 = \emptyset$, $\Gamma^2_1 = \{1\}$, $\Gamma^3_1 = \{1,2\}$.
-Now we generate $\gamma_e$ for $e\in\{1,2,3\}$,
-%	\[
-%	\gamma_e = \bigwedge_{d\in\{1,2,3\}\setminus\{e\}} p^d_{1,1} \wedge p^d_{1,2} \wedge p^d_{1,3} \wedge p^d_{2,1} \wedge p^d_{2,2} \wedge p^d_{2,3},
-%	\]
-%	and $\theta_e$ for $e\in\{1,2,3\}$,
-%	\begin{align*}
-%	\theta_1 &= \neg t_1 \wedge \neg t_2,\\
-%	\theta_2 &= \neg t_1 \wedge t_2,\\
-%	\theta_3 &= t_1 \wedge \neg t_2.		
-%	\end{align*}
-%	Then, we generate the formulas $\Phi^e_1$, for $e\in\{1,2,3\}$,
-%	\begin{align*}
-%	\Phi^1_1 &= \square\\
-%	\Phi^2_1 &= p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2,\\
-%	\Phi^3_1 &= (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3),
-%	\end{align*}
-%	and so, the DNF formula we generate is
-%	\begin{multline*}
-%	\Phi = (p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2) \vee (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee \\ \bigvee_{e\in\{1,2,3\}}[(p^e_{1,1}\wedge\neg p^e_{1,1})\vee
-%	(p^e_{1,2}\wedge\neg p^e_{1,2})\vee
-%	(p^e_{1,3}\wedge\neg p^e_{1,3})\vee
-%	(p^e_{2,1}\wedge\neg p^e_{2,1})\vee
-%	(p^e_{2,2}\wedge\neg p^e_{2,2})\vee
-%	(p^e_{2,3}\wedge\neg p^e_{2,3})].
-%	\end{multline*}
-%	To demonstrate that our reduction is correct, we will show a function $g:\{\langle\P,\e\rangle \mid \A \models \exists\x\,\varphi(\x,\P,\e)\} \to \{\sigma \mid \sigma(\Phi) = 1\}$ and then prove that it is a bijection.
-Let $g$ be such that for each assignment $\langle \P,\e \rangle$, it holds that $g(\langle \P,\e \rangle) = \sigma$ where $\sigma$ is constructed as follows.
-For each $t_i\in T$, let $\sigma(t_i) = b^{\e}_i$.
-For each $p_{\ell,\b}^{\e}$, where $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e}) = 1$ if and only if $\b \in P_{\ell}$.
-For each $p_{\ell,\b}^{\e'}$, where $\e'\in A^d\setminus\{\e\}$,  $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e'}) = 1$.
-%	
-%	Let $i$ be such that $\A\models \exists\x\,\varphi_i(\x,\P,\e)$.
-It is simple to see that from the way we constructed $\Phi_i^{\e}$, it holds that $\sigma(\Phi_i^{\e}) = 1$.
-Also, note that every different assignment $\langle \P',\e' \rangle$, it holds that $\sigma(\langle \P',\e' \rangle) \neq \sigma(\langle \P,\e \rangle)$.
-Therefore, $g$ is an injective function.
-Now, let $\sigma$ be an arbitrary assignment such that $\sigma(\Phi) = 1$.
-Let $i$ and $\e$ be such that $\sigma(\Phi^{\e}_i)$.
-Note that for every $\e'\in A^d\setminus\{\e\}$, it holds that $\sigma(\theta_{\e'}) = 0$, so there is exactly one possible $\e$ associated to $\sigma$.
-Consider the assignment $\P$ where $\b\in P_{\ell}$ if and only if $\sigma(p^{\e}_{\ell,\b}) = 1$, for each $\ell$ and $\b \in A^{c_{\ell}}$.
-Since there is at least one $\a\in\Gamma_i$ for which $\sigma$ satisfies its respective subformula, it holds that $\A\models\varphi_i(\a,\P,\e)$, which implies that $\A\models\exists\x\,\varphi(\x,\P,\e)$.
-From this we deduce that $g$ is surjective, and therefore a bijection.
-We conclude that $f(\A) = f_{\textsc{\#DNF}}(\A_{\Phi})$.
-%\end{enumerate}
-%	
-%\textbf{Proof of Theorem \ref{nplusone-strict}}
-%
-%Suppose it is not properly contained for $n$.
-Let $\z = (z_1,\ldots,z_{n+1})$, let $\u = (u_1,\ldots,u_n)$ and let $\varphi(\z) = \bigwedge_{i = 1}^{n} (z_i < z_{i+1} \wedge \forall y(z_i\not<y \vee y\not<z_{i+1} )) \wedge \forall y(y \not< z_1)$.
-Note that $f_{\varphi(\z)}(\A) = 1$ if $\A$ has at least $n+1$ elements, and  $f_{\varphi(\z)}(\A) = 0$ otherwise, for each $\A\in\str$.
-Let $\forall\x\,\psi(\x,\X,\u)$ be such that $f_{\varphi(\z)} = f_{\forall\x\,\psi(\x,\X,\u)}$.
-Let $\A = \langle A, \S^{\A}, <^{\A} \rangle\in\str$ where $\vert A \vert = n+1$.
-Note that $f_{\forall\x\,\psi(\x,\X,\u)}(\A) = 1$.
-Let $(\P,\e)$ be such that $\A\models\forall\x\,\psi(\x,\P,\e)$.
-Let $a\in A$ be such that $a\not\in\e$.
-Consider the vocabulary $\L' = \L \cup \{\Q\}$, and let $\A' = \A \cup \langle \Q^{\A'} \rangle$ where $\Q^{\A'} = \P$.
-Note that $\A\models\varphi(\P,\e)$ if and only if $\A'\models\varphi(\e)$.
-Consider $\B'$ as the induced substructure of $\A'$ with domain $A\setminus\{a\}$, and let $\Q^{\B'}$ be the induced interpretation of $\Q$.
-Also let $\B$ be the induced substructure of $\A$ with domain $A\setminus\{a\}$.
-As universal formulae are preserved under substructure\footnote{Hodges, 1997.
-P.
-183}, we have that $\B'\models\forall\x\,\psi(\x,\e)$, which implies that $\B\models\forall\x\,\psi(\x,\Q^{\B'},\e)$.
-We conclude that $f_{\forall\x\,\psi(\x,\X,\u)}(\B) \geq 1$, which leads to a contradiction.
-%	
-%\textbf{Proof of Theorem \ref{sigma2-pe}}
-%
-%Let $f = f_{\varphi(\X,\z)}$ such that $\varphi(\X,\z)$ is defined by the grammar $E_2$.
-First we notice that, as stated in \cite{SalujaST95}, for every $\L$-formula $\varphi$, $f_{\varphi}\in\shp$.\\
-%
-%We will now prove that there is a polynomial time algorithm that decides $L_{f}$.
-This is equivalent to decide, given $\A= \langle A, \S^\A, \leq^\A \rangle\in\str$, whether there exist assignments $\P,\z$ such that $\A\models\varphi(\P,\z)$.
-This is the same as $\A\models\exists\X\exists\x\,\varphi(\X,\x)$.
-Since $\varphi(\X,\x)$ is defined by the grammar $E_2$, there exists $\psi(\X,\x,\y,\u)$ such that $\varphi(\X,\x) = \exists\x\exists\y\forall\u\exists\v\,\psi(\X,\x,\y,\u,\v)$, where every variable in $\v$ appears on a negated second-order literal $\neg X_i$, with $i\in\N$.
-%
-%Given $\A\in\str$ we generate an equivalent $\L$-formula $\theta(\X,\x,\y)$ with a series of operations.
-First, we replace every instance of $\v$ by a disjunction of all $r$-tuples in $A^r$, where $r$ is the number of variables in $\v$.
-The result is still a Horn clause.
-Second, we replace every instance of $\u$ by a conjunction of all $s$-tulples in $A^s$, where $s$ is the number or variables in $\u$.
-The result is still a conjunction of Horn clauses.
-%
-%Now we notice that $\zeta = \exists\X\exists\x\exists\y\,\theta(\X,\x,\y)$ is an existential second-order FO-formula, for which $\A\models\zeta$ can be decided in polynomial time \cite{Gradel91}.
-
-\textbf{Proof of Theorem \ref{sigma2hard}}
-
-First we prove that $\shdhsat$ is in $\eqso(\ehorn)$.
-Recall that each instance of $\shdhsat$ is a disjunction of Horn formulas.
-Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot), \pD(\cdot,\cdot)\}$.
-Each symbol in this vocabulary is used to indicate the same as in Example \ref{ex-hornsat-esop1}, with the addition of $\pD(d,c)$ which indicates that $c$ is a clause in the formula $d$.
-Recall that the formula
-\begin{align*}
-&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \neg\textit{T}(x)).
-\end{align*}
-defines $\chsat$. We obtain the following formula $\psi(T,A)$ in $\ehorn$:
-\begin{align*}
-\exists d[&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \pD(c,d)\vee \neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \neg\textit{T}(x))]
-\end{align*}
-effectively defines $\chsat$ as for every disjunction of Horn formulas $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
-Therefore, we conclude that $\shdhsat \in \eqso(\ehorn)$.
-
-\vspace{1em}
-We will now prove that $\shdhsat$ is hard for $\eqso$ over a signature $\R$ under parsimonious reductions.
-For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a procedure that computes a function $g_{\alpha}$.
-This function receives a finite $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
-We suppose that $\alpha$ is in sum normal form and:
-$$
-\alpha = c + \sum_{i = 1}^{\text{\#clauses}} \sa{\bar{X}}\sa{\bar{x}}\hat{\varphi}_i(\bar{X},\bar{x}),
-$$
-where each $\hat{\varphi}_i$ is a $\ehorn$ formula.
-
-Consider a finite $\R$-structure $\A$ with domain $A$.
-To simplify the proof, we extend our grammar to allow first-order constants.
-Consider each tuple $\bar{a}\in A^{\length(\bar{x})}$ as a tuple of first-order constants.
-The following formula defines the same function as $\alpha$:
-$$
-c + \sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length(\bar{x})}} \sa{\bar{X}}\hat{\varphi}_i(\bar{X},\bar{e}).
-$$
-We redefine this formula as $c + \sum_{i = 1}^{m} \sa{\bar{X}}\hat{\psi}_i(\bar{X})$, where each $\hat{\psi}_i$ is a formula in $\ehorn$ with the extended grammar.
-
-%\vspace{1em}
-%Let $\alpha$ be equal to some constant $c$. We define $g_{\alpha}$ as a function that, for each $\A\in\ostr[\R]$, outputs the following formula with exactly $c$ satisfying assignments:
-%$$
-%\bigvee_{i = 1}^{c}\neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_c.
-%$$
-
-Let the formula be equal to $c + \sum_{i = 1}^{m}\sa{\bar{X}}\exists\bar{y}\bigwedge_{j = 1}^{n}\forall\bar{z}\,\varphi^i_j(\bar{X},\bar{y},\bar{z})$ where each $\varphi_i$ is a Horn clause.
-The following formula defines the same function as $\alpha$:
-$$
-c + \sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length(\bar{y})}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\bar{z}}}\varphi^i_j(\bar{X},\bar{b},\bar{c}).
-$$
-Note that each $\fo$ formula in each $\varphi^i_j$ has no free variables.
-Therefore, we can evaluate each of these in polynomial time and replace them by $\perp$ and $\top$ where it corresponds.
-Each $\varphi^i_j$ will be of the form $\perp \vee\, \chi^i_j(\bar{X})$ or $\top \vee \chi^i_j(\bar{X})$ where $\chi^i_j$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$.
-The formulas of the form $\top \vee \chi^i_j(\bar{X})$ can be removed entirely, and the formulas of the form $\perp \vee\, \chi^i_j(\bar{X})$ can be replaced by $\chi^i_j(\bar{X})$.
-We obtain the formula
-$$
-c + \sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})
-$$
-where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$.
-
-We make the following observation.
-If $\alpha$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$, then the {\bf propositional formula} defined by $\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ has exactly $\sem{\alpha}(\A)$ assignments and is a disjunction of Horn formulas.
-
-%Let $f$ be an arbitrary function in $\XE{2}$.
-We will now show a function $h:\str\to L(P)$ such that for every $\A\in\str$, it holds that $f(\A) = \shdhsat(h(\A))$, and which can be computed in polynomial time.
-Let $\psi$ be such that $f = f_{\psi}$.
-We separate the proof in three cases:\begin{enumerate}
-%	\item All free variables in $\psi$ are of first order.
-Let $\psi = \exists \x \forall \y\,\varphi(\x,\y,\z)$ where $\z = (z_1,\ldots,z_d)$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a propositional formula $\Phi = h(\A)$.
-First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\forall \y\,\varphi(\x,\y,\e)$.
-Let $n$ be the number of such assignments to $\z$.
-Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$.
-The disjunction of Horn formulas we return is
-%	\[
-%	\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
-%	\]
-%	which has exactly $n$ possible assignments. Note that each $p_i$ and $\neg p_i$ is a Horn clause itself.
-%	
-%	\item There are only second-order free variables in $\psi$.
-Let $\psi = \exists \x\forall \y\,\varphi(\x,\y,\X)$, where $\x = (x_1,\ldots,x_c)$, $\y = (y_1,\ldots,y_b)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a Disj-Horn formula $\Phi = h(\A)$ as follows:
-%	
-%	First, Let $\varphi(\x,\y,\X) = \bigwedge_{i = 1}^k \varphi_{i}(\x,\y,\X)$ where, without loss of generality, each $\varphi_{i}(\x,\y,\X)$ is of the form
-%	\[
-%	\varphi_{i}(\x,\y,\X) = X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n}) \vee \alpha_i(\x,\y),
-%	\]
-%	where $\u_{s}$ and every $\u_{t_j}$ have the corresponding number of variables in $(\x,\y)$, and $\alpha(\x,\y)$ is a FO-formula.
-For the rest of the proof, $X_s(\u_s)$ can be absent and it does not affect the proof.
-Note that this formula is equivalent to
-%	\[
-%	\neg\alpha_i(\x,\y) \to (X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n})),
-%	\]
-%	
-%	Then, for each $\a\in A^c$ let $\Gamma_i^{\a} = \{\b\in A^{b}\mid \A\models\neg\alpha(\a,\b)\}$.
-Note that this set can be generated in polynomial time.
-For each $\ell\in\{1,\ldots,r\}$, we generate a propositional variable $p^{\ell}_{\d}$ for each $\d\in A^{c_{\ell}}$.
-Let $\p$ be the tuple of all such variables.
-Also, let $\rho_{\ell}(\x,\y) = \u_{\ell}$, and $\rho_{\ell}(\d)$ be the tuple that results of reordering $\d$ using the reordering of the variables of $(\x,\y)$ in $\u_{\ell}$.
-%	
-%	For each $\a\in A^c$ we generate a Horn formula
-%	\[
-%	\Phi_{\a}^i = \bigwedge_{\b\in\Gamma_i} p^{s}_{\rho_s(\a,\b)} \vee \neg p^{t_1}_{\rho_{t_1}(\a,\b)} \vee \cdots \vee \neg p^{t_n}_{\rho_{t_n}(\a,\b)}.
-%	\]
-%	and finally we return $\Phi = \bigvee_{\a\in A^c} \bigwedge_{i = 1}^k \Phi_{\a}^i$.
-%	
-%	Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\forall\y\,\varphi(\x,\y,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$.
-We can conclude that $f(\A) = f_{\textsc{Disj-Horn-Sat}}(\Phi)$.
-%	
-%	\item Let $\psi = \exists \x\forall\y\,\varphi(\x,\y,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\y = (y_1,\ldots,y_b)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$ we generate a Disj-Horn formula $\Phi = h(\A)$ as follows:
-%	
-%	As in the previous case, let $\varphi(\x,\y,\X,\z) = \bigwedge_{i = 1}^k$ where $\varphi(\x,\y,\X,\z)$ is of the form:
-%	\[
-%	\varphi_{i}(\x,\y,\X,\z) = X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n}) \vee \alpha_i(\x,\y,\z),
-%	\]
-%	where $\u_{s}$ and every $\u_{t_j}$ have the corresponding number of variables in $(\x,\y,\z)$, and $\alpha(\x,\y,\z)$ is a FO-formula.
-For the rest of the proof, $X_s(\u_s)$ can be absent and it does not affect the proof.
-Note that this formula is equivalent to
-%	\[
-%	\neg\alpha_i(\x,\y,\z) \to (X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n})).
-%	\]
-%	
-%	For each $\e\in A^d$, we do the following.
-For each $\a\in A^c$ let $\Gamma^{\e}_{i,\a} = \{\b \in A^b \mid \A\models \neg\alpha_i(\a,\b,\e) \}$.
-Note that this set can be generated in polynomial time.
-For each $\ell\in\{1,\ldots,r\}$ we generate a propositional variable $p^{\e}_{\ell,\d}$ for each $\d\in A^{c_{\ell}}$.
-Let $\p$ be the tuple of all such variables.
-We also generate a set $T = \{t_1,\ldots,t_{L}\}$ where $L = \lceil \log\vert A^d\vert \rceil$, and an arbitrary injective function $\beta:A^d \to \{0,1\}^L$ where $\beta(\e) = (b^{\e}_1,\ldots,b^{\e}_L)$ for each $\e\in A^d$.
-Also, let $\rho_{\ell}(\x,\y,\z) = \u_{\ell}$, and $\rho_{\ell}(\d)$ be the tuple that results of reordering $\d$ using the order of the variables in $(\x,\y,\z)$ in $\u_{\ell}$.
-And so, the Horn formula we generate is
-%	\[
-%	\Phi^{\e}_{i,\a} = \bigwedge_{\a\in\Gamma^{\e}_i} p^{\e}_{s,\rho_{s}(\a,\b,\e)} \vee p^{\e}_{t_1,\rho_{t_1}(\a,\b,\e)} \vee \cdots \vee p^{\e}_{t_n,\rho_{t_n}(\a,\b,\e)}.
-%	\]
-%	And finally we return $\Phi = \bigvee_{\e\in A^d}\bigvee_{\a\in A^c} (\bigwedge_{i=1}^k \Phi_{\e}^{i,\a} \wedge \gamma_{\e} \wedge \theta_{\e} )$, where $\gamma_{\e}$ is defined as
-%	\[
-%	\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
-%	\]
-%	and $\theta_{\e}$ is defined as
-%	\[
-%	\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
-%	\]
-%	
-%\end{enumerate}
-
-\textbf{Proof of Theorem \ref{so-rec}}
-
-\textbf{Proof of Theorem \ref{tqfo-fo-cap}}
-
-For the second condition, let $f \in \shl$.
-We will address the case where $\R$ contains only one binary predicate $E$, and the rest of the cases can be deduced from this.
-Let $M$ be a non-deterministic logspace machine such that $f(\A) = \acc_M(\A)$ for each $\A \in \ostr[R]$.
-Suppose ${\cal Q} = \{q_1,\ldots,q_{\ell}\}$ is the set of states of $M$, where $q_1$ is the initial state, and $q_{\ell}$ is only final state of $M$.
-Let $n = \vert A \vert$ and let $w = \enc(\A) \in \{0,1\}^{n^2}$.
-We assume that $M$ with input $w$ uses space $s_M(w) < c\cdot\log(n)$ and furthermore, $s_M(w) < n-2$.
-We notate $M(w)$ as the graph of configurations of $M$ running on input $w$.
-
-We represent configurations with a tuple of fixed size.
-The formula $\varphi(\bar{x},\bar{y})$ describes a procedure that given a configuration generates a possible next configuration.
-The formula $\varphi_I(\bar{x})$ describes that $\bar{x}$ is the initial configuration of $M(w)$.
-The formula $\varphi_F(\bar{x})$ describes that $\bar{x}$ is an accepting (final) configuration of $M(w)$.
-The formula we construct is:
-$$
-\alpha = \sa{\bar{x}}\sa{\bar{y}}([\pth \varphi(\bar{u},\bar{v})](\bar{x},\bar{y})\cdot \varphi_I(\bar{x})\cdot\varphi_F(\bar{y})).
-$$
-
-To illustrate our idea, we will show a simplified example.
-Consider a machine $M$ that works in exactly $\log_2(n)$ space and only allows 0 or 1 in the working tape.
-Consider an input $\A$ of size 16 (that is, $A = \{0,\ldots,9,A,\ldots,F\}$).
-Let some configuration $s$ have 0011 in the working tape, the head in the input tape is in position 26, and the head in the input tape is in position 2 (we consider 0-indexed positions).
-Also, $Q = \{q_1,\ldots,q_5\}$ and the current state is $q_3$.
-
-As a first approach, we will use a 9-tuple $\bar{a} = (a_1,\ldots,a_9)$ to represent $s$.
-That is, $(a_1,a_2) = (1,A)$ represent the position of the head in the input tape, $a_3 = 2$ represents the position of the head in the working tape, $a_4 = C$ (1100b in base 16) represents the content of the working tape, and $(a_5,\ldots,a_9) = (0,0,1,0,0)$ represents the current state.
-Then $\bar{a} = (1,A,2,C,0,0,1,0,0)$ will represent $s$.
-
-The problem that arises from this representation, is that to describe a transition in $M$ we need to read an arbitrary character in the working tape, and therefore we have to obtain the $a_3$-th bit in $a_4$.
-Furthermore, to represent the following configuration, we need compute $a_4$ with the $a_3$-th bit flipped.
-This is generally not possible to describe with an $\fo$ formula.
-Consider the following procedure that receives $x = a_4$ and $i = a_3$.
-
-\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 1 replace it by 0 and return the result}
-	\label{switch1to0}
-	\begin{algorithmic}
-		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
-		\While{$j > 0$}
-		\State $v \gets 0$
-		\While{$u > 1$}
-		\State $u \gets u-2,\; v \gets v+1$
-		\EndWhile
-		\State $u\gets v,\; j \gets j-1$
-		\EndWhile
-		\While{$u > 1$}
-		\State $u \gets u-2$
-		\EndWhile
-		\State $\textbf{assert } u = 1$ \Comment{If $u \neq 1$ simply stop}	
-		\State $y \gets 1$ \Comment{Compute $2^i$ and store it in $y$}
-		\While{$i > 0$}
-		\State $z \gets 0$
-		\While{$y > 0$}
-		\State $z \gets z+2,\; y \gets y-1$
-		\EndWhile
-		\State $i \gets i-1,\; y \gets z$
-		\EndWhile
-		\While{$y > 0$} \Comment{Substract $y$ from $x$}
-		\State $x \gets x-1,\; y \gets y-1$
-		\EndWhile
-		\State \Return $x$.
-	\end{algorithmic}
-\end{algorithm}	
-Each of the instructions can be expressed with $\fo$, so our strategy is to use the $\pth$ operator to simulate the algorithm and then we can describe a transition using the processed value of $a_4$.
-Note that we need to describe three more procedures to simulate the transitions for $0\to 0$, $0 \to 1$ and $1\to 1$.
-
-We will now describe how to simulate both the procedure and the transition.
-A procedure tuple $\bar{p} = (a_1,\ldots,a_{3+c+\ell},b_1,b_2,c_1,c_2,c_3,d_1,\ldots,d_{5c+2})$ represents the current configuration of $M(w)$ in $a_1,\ldots,a_{2+c+\ell}$, the values that will be read and written in the working tape in $b_1,b_2$, the instruction pointer in $c_1,c_2,c_3$ and the values stored in memory in $d_1,\ldots,d_{10c+2}$.
-In detail:
-\begin{enumerate}
-	\item $a_1,a_2$ and $a_3$ represent the position of the head in the input tape and the working tape, respectively, $a_4,\ldots,a_{3+c}$ represent the content of the working tape and $a_{4+c},\ldots,a_{3+c+\ell}$ represent the current state in the current configuration that is being processed.
-	\item $b_1$ and $b_2$ are equal to the value that is being read in the working tape and the value that will be written in the working tape respectively.
-	\item $c_1,c_2,c_2$ represent the instruction pointer in the procedure. Only 8 different instructions are needed in the simulation.
-	\item Each value in memory of $x,y,z,u,v$ need $c$ elements to represent them and $i,j$ need only one. We map $(d_1\ldots,d_{c}) \to x$, $(d_{c+1}\ldots,d_{2c}) \to y$,
-	$(d_{2c+1}\ldots,d_{3c}) \to z$, $(d_{3c+1}\ldots,d_{4c}) \to u$,
-	$(d_{4c+1}\ldots,d_{5c}) \to v$, $d_{5c+1} \to i$ and $d_{5c+2}\to j$.
-\end{enumerate}
-For each transition $\delta \in \Delta \subseteq Q \times \{0,1\} \times \{0,1\} \times Q \times \{-1,=,+1\} \times \{0,1\} \times \{-1,=,+1\}$ we define a formula $\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$, where $\bar{x} = (x_1,\ldots,x_{3+c+\ell})$, $\bar{s} = (s_1,s_2)$, $\bar{w} = (w_1,w_2,w_3)$, $\bar{u} = (u_1,\ldots,u_{5c+2})$, $\bar{y} = (y_1,\ldots,y_{3+c+\ell})$, $\bar{t} = (t_1,t_2)$, $\bar{z} = (z_1,z_2,z_3)$ and $\bar{v} = (v_1,\ldots,v_{5c+2})$.
-The tuples $\bar{x}$ and $\bar{y}$ represent the current and next configuration of $M$ respectively, $\bar{s}$ and $\bar{y}$ indicate which algorithm is being followed, $\bar{w}$ and $\bar{z}$ represent the current and next instruction of the algorithm, $\bar{u}$ and $\bar{v}$ represent the current and next values in memory.
-We will describe the formula part by part.
-Suppose $\delta = (q_i,a,1,q_j,op_1,0,op_2)$, so we have to simulate Algorithm \ref{switch1to0}.
-
-First we define some auxiliary formulas:
-\begin{align*}
-\varphi_0(x) &= \neg\exists y(y < x), \\
-\varphi_{+1}(x,y) &= x < y \wedge \neg\exists z(x < z \wedge z < y), \\
-\varphi_{-1}(x,y) &= \varphi_{+1}(y,x), \\
-\varphi_1(x) &= \exists y(\varphi_0(y) \wedge \varphi_{+1}(y,x)), \\
-\text{for $i,j \in\{0,1\}$, }	\varphi_{i,j}(x,y) &= \varphi_i(x) \wedge \varphi_j(y), \\
-\text{for each $k \leq 7$, }\varphi^b_k(x_1,x_2,x_3) &= \varphi_{a_1}(x_1) \wedge \varphi_{a_2}(x_2) \wedge \varphi_{a_3}(x_3)\text{, where $a_1a_2a_3$ is the value of $k$ in binary}, \\
-\mu_{<}(x_1,\ldots,x_c,y_1,\ldots,y_c) &= \bigvee_{i = 1}^c ((\bigwedge_{j = 1}^{i-1} x_j = y_j) \wedge x_i < y_i), \\
-\mu_{+1}(\bar{x},\bar{y}) &= \mu_{<}(\bar{x},\bar{y}) \wedge \neg\exists \bar{z}(\mu_{<}(\bar{x},\bar{z}) \wedge \mu_{<}(\bar{z},\bar{y})), \\
-\mu_{+2}(\bar{x},\bar{y}) &= \exists\bar{z}(\mu_{+1}(\bar{x},\bar{z}) \wedge \mu_{+1}(\bar{z},\bar{y})),\\
-\mu_{-1}(\bar{x},\bar{y}) &= \varphi_{+1}(\bar{y},\bar{x}),\\
-\mu_{-2}(\bar{x},\bar{y}) &= \varphi_{+2}(\bar{y},\bar{x}),\\
-\mu_{0}(\bar{x}) &= \neg\exists\bar{y}(\mu_{<}(\bar{y},\bar{x})),\\
-\mu_{1}(\bar{x}) &= \exists\bar{y}(\mu_{0}(\bar{y})\wedge\mu_{+1}(\bar{y},\bar{x})),\\
-\varphi^2_{<}(x_1,x_2,y_1,y_2) &= x_1 < y_1 \vee (x_1 = y_1 \wedge x_2 < y_2),\\
-\varphi^2_{+1}(x_1,x_2,y_1,y_2), \varphi^2_{-1}(x_1,x_2,y_1,y_2) &\text{ defined analogously to $\mu_{+1}$ and $\mu_{-1}$},\\
-\text{for each } q_i\in Q, \varphi^q_i(x_1,\ldots,x_{\ell}) &= \bigwedge_{\substack{j = 1 \\ j \neq i}}^{\ell} \varphi_0(x_j) \wedge \varphi_1(x_i), \\
-\varphi^E_0(x_1,x_2) &= \neg E(x_1,x_2),\\		\varphi^E_1(x_1,x_2) &= E(x_1,x_2),\\
-\end{align*}
-
-We start from instruction 0, which means that the procedure has not started yet and every value in the tuple is 0 except for the configuration values.
-It also initializes all the values in the tuple to 0 except for $x,u,i,j$.
-\begin{multline*}
-\varphi^{0,1}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{0,0}(s_1,s_2)\wedge\varphi^b_0(\bar{w}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_1(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c v_i = x_{3+i} \wedge \bigwedge_{i = c+1}^{2c} \varphi_0(v_i) \wedge \bigwedge_{i = 2c+1}^{3c} \varphi_0(v_i) \wedge \bigwedge_{i = 1}^c v_{3c+i} = x_{3+i} \wedge \bigwedge_{i = 4c+1}^{5c} \varphi_0(v_i) \wedge v_{5c+1} = x_3 \wedge v_{5c+2} = x_3.
-\end{multline*}
-Instruction 1 which checks whether the value of $j$ ($d_{5c+2}$ in the tuple) is more than 0 or not, and then proceeds to instruction 2 or 3 on each case.
-\begin{multline*}
-\varphi^{1,2}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_1(\bar{w}) \wedge \neg \varphi_0(u_{5c+2}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_2(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i, \\
-\varphi^{1,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_1(\bar{w}) \wedge \varphi_0(u_{5c+2}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_3(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
-\end{multline*}
-Instruction 2 checks the value of $u$ ($d_{3c+1},\ldots,d_{4c}$ in the tuple).
-If it is $> 1$ then it substracts 2 from $u$ and adds 1 to $v$ ($d_{4c+1},\ldots,d_{5c}$ in the tuple), then repeats instruction 2.
-If it is equal to 0 or 1, then moves the value of $v$ to $u$, substracts 1 from $j$ and goes back to instruction 1.
-\begin{multline*}
-\varphi^{2,2}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_2(\bar{w}) \wedge \neg \mu_0(u_{3c+1},\ldots,u_{4c}) \wedge \neg \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_2(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
-\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-\mu_{-2}(u_{3c+1},\ldots,u_{4c},v_{3c+1},\ldots,v_{4c})\, \wedge \\
-\mu_{+1}(u_{4c+1},\ldots,u_{5c},v_{4c+1},\ldots,v_{5c}) \wedge u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}. \\
-\varphi^{2,1}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_2(\bar{w}) \wedge ( \mu_0(u_{3c+1},\ldots,u_{4c}) \vee \mu_1(u_{3c+1},\ldots,u_{4c})) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_1(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
-\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-\bigwedge_{i = 3c+1}^{4c} u_{c+i} = v_i \wedge
-\mu_0(v_{4c+1},\ldots,v_{5c}) \wedge
-u_{5c+1} = v_{5c+1} \wedge \varphi_{-1}(u_{5c+2},v_{5c+2}).	
-\end{multline*}
-Instruction 3 calculates the value of $u \mod 2$, that is, it repeats instruction 3 until the value of $u$ is equal to 0 or 1.
-On each iteration, it substracts 2 from $u$.
-Moreover, if the value of $u$ at the end of the iterations is not 1 then there is no step defined.
-\begin{multline*}
-\varphi^{3,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_3(\bar{w}) \wedge \neg \mu_0(u_{3c+1},\ldots,u_{4c}) \wedge \neg \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_3(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
-\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-\mu_{-2}(u_{3c+1},\ldots,u_{4c},v_{3c+1},\ldots,v_{4c}) \wedge
-\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
-u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}, \\
-\varphi^{3,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_3(\bar{w}) \wedge \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_4(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\mu_1(v_{c+1},\ldots,v_{2c}) \wedge
-\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
-\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
-u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}
-\end{multline*}
-Instruction 4 checks the value of $i$ ($d_{5c+1}$ in the tuple.) If it is not 0 then goes to instruction 5 and if is 0 then goes to instruction 6.
-Moreover it initializes the value of $z$ ($d_{2c+1},\ldots,d_{3c}$ in the tuple) to 0 (which was 0 all along.)
-\begin{multline*}
-\varphi^{4,5}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_4(\bar{w}) \wedge \neg \varphi_0(u_{5c+1}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_5(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i, \\
-\varphi^{4,6}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_4(\bar{w}) \wedge \varphi_0(u_{5c+1}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_6(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
-\end{multline*}
-Instruction 5 checks the value of $y$ ($d_{c+1},\ldots,d_{2c}$ in the tuple.) If it is more than 0 then it adds 2 to $z$ and substracts 1 from $y$, then repeats instruction 2.
-If it is not, then copies the value of $z$ to $y$ and subtracts 1 from $i$ and returns to instruction 4.
-\begin{multline*}
-\varphi^{5,5}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_5(\bar{w}) \wedge \neg \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_5(\bar{z}) \, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\mu_{-1}(u_{c+1},\ldots,u_{2c},v_{c+1},\ldots,v_{2c}) \wedge
-\mu_{+2}(u_{2c+1},\ldots,u_{3c},v_{2c+1},\ldots,v_{3c})\, \wedge \\
-\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
-\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
-u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}, \\
-\varphi^{5,4}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_5(\bar{w}) \wedge \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_4(\bar{z}) \, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\bigwedge_{i = c+1}^{2c} u_{c+i} = v_i \wedge
-\mu_0(v_{2c+1},\ldots,v_{3c}) \wedge
-\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
-\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
-\varphi_{-1}(u_{5c+1},v_{5c+1}) \wedge u_{5c+2} = v_{5c+2}
-\end{multline*}
-Instruction 6 checks the value of $y$. If it is more than 0, then subtracts 1 from $x$ and $y$ and repeats instruction 6. If it is not, then goes to instruction 7.
-\begin{multline*}
-\varphi^{6,6}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_6(\bar{w}) \wedge \neg \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_6(\bar{z}) \, \wedge \\
-\mu_{-1}(u_1,\ldots,u_c,v_i,\ldots,v_c) \wedge \mu_{-1}(u_{c+1},\ldots,u_{2c},v_{c+1},\ldots,v_{2c}) \wedge \bigwedge_{i = 3c+1}^{5c+2} u_i = v_i]\,\vee \\
-\varphi^{6,7}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_6(\bar{w}) \wedge \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_7(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
-\end{multline*}
-Instruction 7 stores the value of $x$ after the corresponding bit has been switched.
-Then we can define $\varphi_{\delta}$ which also simulates the actual transition.
-If $u$ equals 1, then copy what is stored in $x$ to $a_4,\ldots,a_{3+c}$, go from state $q_i$ to state $q_j$, and move the heads to their corresponding positions.
-\begin{multline*}
-\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 	[\bigwedge_{i = 1}^{3+c+\ell} x_i = y_i \wedge (\varphi^{0,1}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{1,2}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})\, \vee \\ \varphi^{1,3}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{2,2}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{2,1}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{3,3}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \\ \varphi^{3,4}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{4,5}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{4,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{5,5}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \\ \varphi^{5,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{6,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{6,7}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}))] \, \vee \\
-[\varphi^E_a(x_1,x_2) \wedge \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_7(\bar{w}) \wedge \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{0,0}(t_1,t_2) \wedge \varphi^b_0(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i \,\wedge \\
-\varphi^2_{op_1}(x_1,x_2,y_1,y_2) \wedge \varphi_{op_2}(x_3,y_3) \wedge \bigwedge_{i = 1}^c u_i = x_{3+i} \wedge \varphi^q_i(x_{4+c},\ldots,x_{3+c+\ell}) \wedge \varphi^q_j(y_{4+c},\ldots,y_{3+c+\ell})].
-\end{multline*}
-
-\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 0 return $x$} \label{switch0to0}
-	\begin{algorithmic}
-		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
-		\While{$j > 0$}
-		\State $v \gets 0$
-		\While{$u > 1$}
-		\State $u \gets u-2,\; v \gets v+1$
-		\EndWhile
-		\State $u\gets v,\; j \gets j-1$
-		\EndWhile
-		\While{$u > 1$}
-		\State $u \gets u-2$
-		\EndWhile
-		\State $\textbf{assert } u = 0$ \Comment{If $u \neq 0$ simply stop}	
-		\State \Return $x$.
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 0 replace it by 1 and return the result}
-	\label{switch0to1}
-	\begin{algorithmic}
-		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
-		\While{$j > 0$}
-		\State $v \gets 0$
-		\While{$u > 1$}
-		\State $u \gets u-2,\; v \gets v+1$
-		\EndWhile
-		\State $u\gets v,\; j \gets j-1$
-		\EndWhile
-		\While{$u > 1$}
-		\State $u \gets u-2$
-		\EndWhile
-		\State $\textbf{assert } u = 0$ \Comment{If $u \neq 0$ simply stop}	
-		\State $y \gets 1$ \Comment{Compute $2^i$ and store it in $y$}
-		\While{$i > 0$}
-		\State $z \gets 0$
-		\While{$y > 0$}
-		\State $z \gets z+2,\; y \gets y-1$
-		\EndWhile
-		\State $i \gets i-1,\; y \gets z$
-		\EndWhile
-		\While{$y > 0$} \Comment{Add $y$ to $x$}
-		\State $x \gets x+1,\; y \gets y-1$
-		\EndWhile
-		\State \Return $x$.
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 1 return $x$}
-	\label{switch1to1}
-	\begin{algorithmic}
-		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
-		\While{$j > 0$}
-		\State $v \gets 0$
-		\While{$u > 1$}
-		\State $u \gets u-2,\; v \gets v+1$
-		\EndWhile
-		\State $u\gets v,\; j \gets j-1$
-		\EndWhile
-		\While{$u > 1$}
-		\State $u \gets u-2$
-		\EndWhile
-		\State $\textbf{assert } u = 1$ \Comment{If $u \neq 0$ simply stop}	
-		\State \Return $x$.
-	\end{algorithmic}
-\end{algorithm}
-Note that we also need to specify that the program we are following is Algorithm \ref{switch1to0} so we store $1,0$ in $b_1,b_2$ all along the procedure.
-We describe the three other algorithms that compute the switches from $0\to 0$, $0\to 1$ and $1\to 1$ (Algorithms \ref{switch0to0}, \ref{switch0to1} and \ref{switch1to1}.)
-For the other three cases, where $\delta = (q_i,a,0,q_j,op_1,0,op_2)$, $\delta = (q_i,a,0,q_j,op_1,1,op_2)$ and $\delta = (q_i,a,1,q_j,op_1,1,op_2)$, $\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$ is defined analogously.
-Then, $\varphi(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$ is defined as:
-$$
-\varphi(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \bigvee_{\delta \in \Delta} \varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}).
-$$
-Finally we define $\varphi_I$ and $\varphi_F$:
-\begin{align*}
-\varphi_I(\bar{x},\bar{s},\bar{w},\bar{u}) &= \varphi_0(x_1) \wedge \varphi_0(x_2) \wedge \varphi_0(x_3) \wedge \bigwedge_{i = 4}^{3+c}\varphi_0(x_i) \wedge \varphi_1(x_{4+c}) \wedge \bigwedge_{i = 5+c}^{3+c+\ell} \varphi_0(x_i)\wedge \varphi_{0,0}(s_1,s_2)\ \wedge \varphi^b_0(\bar{w}) \wedge \bigwedge_{i = 1}^{5c+2}\varphi_0(u_i).
-\\
-\varphi_F(\bar{x},\bar{s},\bar{w},\bar{u}) &= \varphi^q_{\ell}(x_{4+c},\ldots,x_{3+c+\ell}) \wedge \varphi_{0,0}(s_1,s_2)\ \wedge \varphi^b_0(\bar{w}) \wedge \bigwedge_{i = 1}^{5c+2}\varphi_0(u_i),
-\end{align*}
-and then $\sem{\alpha}(\A) = \sem{\sa{\bar{x}}\sa{\bar{y}}([\pth \varphi(\bar{u},\bar{v})](\bar{x},\bar{y})\cdot \varphi_I(\bar{x})\cdot\varphi_F(\bar{y}))}(\A) = \acc_M(\A)$.
-
-\textbf{Proof of Theorem \ref{tqso-fo-cap}}
-
-\textbf{Proof of Theorem \ref{rqfo-fo-cap}}
+and adding these values gives $f(\enc(\A))$. Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$. \qed

### 2016-12-29 LICS/preliminaries.tex

@@
 \subsection{Function complexity classes}
-Assume that $\R = \{R_1, \ldots, R_k\}$ is a relational signature and $\A$ is an ordered finite $\R$-structure with a domain $A$ containing $n$ elements.
-Recall that  $<$ is a linear order on $A$, say $a_1 < a_2 < \ldots < a_n$.
-For every $i \in \{1, \ldots, k\}$, define the encoding of $R_i^\A$, denoted by $\enc(R_i^\A)$, as the following binary string.
-Assume that $\ell = \arity(R_i)$ and consider an enumeration of the $\ell$-tuples over $A$ in the lexicographic order induced by $<$ (that is, $(a_1, \ldots, a_1, a_1)$, $(a_1, \ldots, a_1, a_2)$, $\ldots$, $(a_n, \ldots, a_n, a_{n-1})$, $(a_n, \ldots, a_n, a_n)$).
-Then let $\enc(R_i^\A)$ be a binary string of length $n^\ell$ such that the $i$-th bit of $\enc(R_i^\A)$ is 1 if the $i$-th tuple in the previous enumeration belongs to $R_i^\A$, and 0 otherwise.
-Moreover, define the encoding of $\A$, denoted by $\enc(\A)$, as the following binary string~\cite{L04}:
-\begin{eqnarray*}
-	\enc(\A) & = & 0^n \, 1 \, \enc(R_1^\A) \, \cdots \, \enc(R_k^\A).
-\end{eqnarray*}
-Given a relational signature $\R$,  function $f$ is said to be {\em defined over $\R$} (or just {\em over $\R$}) if $f$ is a total function from $\{\enc(\A) \mid \A \in \ostr[\R]\}$ to $\nat$.
-Moreover, a set $\CC$ of functions is said to be a {\em function complexity class} if every $f \in \CC$ is a function defined over some relational signature $\R$.
 

### 2016-12-29 LICS/saluja.tex

@@
 	Every formula in $\eqso(\LL)$ can be rewritten in \emph{sum normal form}.
-Furthermore, there exists a formula $\alpha \in \loge{1}$ such that there does not exist a formula in $\Sigma$-prenex normal form in $\QE{1}$ equivalent to $\alpha$.
+Furthermore, there exists a formula $\alpha \in \QE{1}$ such that there does not exist a formula in $\Sigma$-prenex normal form in $\QE{1}$ equivalent to $\alpha$.
 \end{theorem}

### 2016-12-29 LICS/appendix-counting.tex
29 dic

@@
 %%% DEMOSTRACION DE SQSO(ESO) y span-P
-\subsection{Proof of Proposition~\ref{prop:capture-spanP}}
+\subsection*{Proof of Proposition~\ref{prop:capture-spanP}}
 
 Similar than the previous proof, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic outputs for each $(\A,v,V) \in \ostr[\R]^*$.
+This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ distinct accepting outputs for each $(\A,v,V) \in \ostr[\R]^*$.
 Suppose $\A$ has domain $A$.
-On each of the steps the algorithm adds the current configuration to the output tape.
-\cristian{Para que necesitas esto?}
-If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time without modifying the output tape, and accepts if and only if the condition holds true.
-\cristian{Falta explicar mejor este caso y porque puedes ejecutar una $\eso$-formula.}
+If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time \cite{fagin1974generalized}, and accepts if and only if the condition holds true.
 If $\alpha$ is a constant $s$, then the algorithm produces $s$ branches and accepts in all of them. 
 If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.  
-%If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$. 
 If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$. 
-%If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$. 
-If $\alpha = \sa{X} \beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
+If $\alpha = \sa{X} \beta$, it chooses $B\in A^{\arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$. 
 This covers all possible cases for $\alpha$.
+Additionally, the algorithm produces a different output on each path. This can be done by printing the trace of all the non-deterministic choices.
+However, when the algorithm starts checking whether $(\A,v,V)\models\varphi$ for some $\eso$ formula $\varphi$, it stops printing in the output tape.
+This way the algorithm produces exactly one output from that point onwards.
 Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
@@
 
-For the other direction, Compton et al.~\cite{ComptonG96} proved that $\spp = \#\eso$. Since a function in $\#\eso$ can also be defined in $\eqso(\eso)$, then $\eqso(\eso)$ captures $\spp$. \qed
+For the other direction, Compton et al.~\cite{ComptonG96} proved that $\spp = \#\eso$.
+Since a function in $\#\eso$ can also be defined in $\eqso(\eso)$, then $\eqso(\eso)$ captures $\spp$ over ordered structures.
+\qed
 
@@
 where $\varphi(x_1,\ldots,x_{\ell}) := \pa{y_1}\cdots\pa{y_{\ell}}((y_1,\ldots,y_{\ell}) < (x_1,\ldots,x_{\ell}) \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
-Adding this values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
+Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
 Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
@@
 %%% DEMOSTRACION DE QSO(PFP) y FPSPACE
-\subsection{Proof of Theorem~\ref{theo:capture-fpspace}}
+\subsection*{Proof of Theorem~\ref{theo:capture-fpspace}}
 
 To show how to evaluate a $\qso(\pfp)$-formula, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
+This machine runs in non-deterministic polynomial space and, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
 Suppose $\A$ has domain $A$.
@@
 Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$ \cite{Ladner89}, and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
+We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$, by the fact that $\shpspace = \fpspace$ \cite{Ladner89}, and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
 \cristian{Aca debes explicar mejor que hace la máquina.
 Es un copy paste de demostraciones anteriores y no queda claro porque puedes evaluar la formula en FPSPACE.
-Tampoco queda claro porque puedes usar no determinismo en FPSPACE.}
+Tampoco queda claro porque puedes usar no determinismo en FPSPACE.} \martin{es un copy-paste porque hice la demostración para $\shpspace$, lo dejé un poco más claro ahora}
 
@@
 \subsection*{Proof of Corollary~\ref{cor:capture-fpspace-poly}}
+
+\martin{no revisar, voy a hacer un copy paste del teorema \ref{theo:capture-fp} cuando este listo}
 

### 2016-12-29 LICS/appendix.tex
29 dic

@@
 We also denote $\length(\bar{x}) = m$ ($\length(\bar{X}) = m$).
-
-\subsection{Proof of Theorem \ref{no-mult}}
+\subsection*{Proof of Theorem \ref{no-mult}}
 
@@
 
-\subsection{Proof of Theorem \ref{captfp}}
+\subsection*{Proof of Theorem \ref{captfp}}
 
@@
 
-\subsection{Proof of Theorem \ref{qfo-pfp-cap}}
+\subsection*{Proof of Theorem \ref{qfo-pfp-cap}}
 
@@
 
-\subsection{Proof of Theorem \ref{eqso-fo-cap}}
+\subsection*{Proof of Theorem \ref{eqso-fo-cap}}
 
@@
 
-\subsection{Proof of Theorem \ref{eqso-eso-cap}}
+\subsection*{Proof of Theorem \ref{eqso-eso-cap}}
 
@@
 
-\subsection{Proof of Theorem \ref{qso-pfp-cap}}
+\subsection*{Proof of Theorem \ref{qso-pfp-cap}}
 
@@
 
-\subsection{Proof of Theorem \ref{eqso-sigma-zero-in-fp}}
+\subsection*{Proof of Theorem \ref{eqso-sigma-zero-in-fp}}
 
@@
 
-\subsection{Proof of Theorem \ref{eqso-sigma-one-in-eqso-pi-one}}
+\subsection*{Proof of Theorem \ref{eqso-sigma-one-in-eqso-pi-one}}
 
@@
 
-%\subsection{Proof of Theorem \ref{one-sigma-zero}}
+%\subsection*{Proof of Theorem \ref{one-sigma-zero}}
 %
@@
 
-%\textbf{Proof of Theorem \ref{mult-sigma-zero}}
+%\subsection*{Proof of Theorem \ref{mult-sigma-zero}}
 %
@@
 %
-%\textbf{Proof of Theorem \ref{fo-prod-sigma-zero}}
+%\subsection*{Proof of Theorem \ref{fo-prod-sigma-zero}}
 %
@@
 %
-%\textbf{Proof of Theorem \ref{fp1}}
+%\subsection*{Proof of Theorem \ref{fp1}}
 %
@@
 
-\textbf{Proof of Theorem \ref{decisionptime}}
+\subsection*{Proof of Theorem \ref{decisionptime}}
 
@@
 
-\textbf{Proof of Theorem \ref{sub-pnp}}
+\subsection*{Proof of Theorem \ref{sub-pnp}}
 
@@
 
-%\textbf{Proof of Theorem \ref{sigma1strict}}
+%\subsection*{Proof of Theorem \ref{sigma1strict}}
 %
@@
 
-\subsection{Proof of Theorem \ref{sigmafo-minusone}}
+\subsection*{Proof of Theorem \ref{sigmafo-minusone}}
 
@@
 %		
-%\textbf{Proof of Theorem \ref{dnf-pars}}
+%\subsection*{Proof of Theorem \ref{dnf-pars}}
 %
@@
 %	
-%\textbf{Proof of Theorem \ref{nplusone-strict}}
+%\subsection*{Proof of Theorem \ref{nplusone-strict}}
 %
@@
 %	
-%\textbf{Proof of Theorem \ref{sigma2-pe}}
+%\subsection*{Proof of Theorem \ref{sigma2-pe}}
 %
@@
 
-\textbf{Proof of Theorem \ref{sigma2hard}}
+\subsection*{Proof of Theorem \ref{sigma2hard}}
 
@@
 $$
-Note that each $\fo$ formula in each $\varphi^i_j$ over $(\bar{x},\bar{y},\bar{z})$ has no free variables.
+Note that each $\fo$ formula over $(\bar{x},\bar{y},\bar{z})$ in each $\varphi^i_j$ has no free variables.
 Therefore, we can evaluate each of these in polynomial time and replace them by $\perp$ and $\top$ where it corresponds.
@@
 Let $\Theta_i = g_{\beta_i}(\A)$.
-We rename the variables in each $\Theta_i$ so none of them are mentioned more than once in total.
+We rename the variables in each $\Theta_i$ so none of them are mentioned in any other $\Theta_j$.
 We add $m$ new variables $t_1,\ldots,t_m$ and we define:
@@
 \end{align*}
-The number of satisfying assignments for this formula is exactly the sum of satisfying assignments for all $g_{\beta_i}(\A)$.
-This covers all possible cases for $\alpha$, and the complete procedure takes polynomial time.
+The formula is a disjunction of Horn formulas, and the number of satisfying assignments for this formula is exactly the sum of satisfying assignments for each $g_{\beta_i}(\A)$.
+This, at the same time, is equal to $\sem{\alpha}(\A)$.
+This covers all possible cases for $\alpha$, and the entire procedure takes polynomial time.
 
@@
 
-\textbf{Proof of Theorem \ref{so-rec}}
-
-\textbf{Proof of Theorem \ref{tqfo-fo-cap}}
+\subsection*{Proof of Theorem \ref{so-rec}}
+
+\subsection*{Proof of Theorem \ref{tqfo-fo-cap}}
 
@@
 
-\textbf{Proof of Theorem \ref{tqso-fo-cap}}
-
-\textbf{Proof of Theorem \ref{rqfo-fo-cap}}
+\subsection*{Proof of Theorem \ref{tqso-fo-cap}}
+
+\subsection*{Proof of Theorem \ref{rqfo-fo-cap}}

### 2016-12-29 LICS/saluja.tex
29 dic

@@
 In the next result, we show that $\eqso$ also defined a finite hierarchy similar than in~\cite{SalujaST95} that we called the $\eqso$-hierarchy.
-\begin{proposition}
+\begin{proposition} \label{prop:eqso-hierarchy-proper}
 \[
@@
 Next we show that each formula in $\eqso(\LL)$ can be converted in \emph{sum normal form} but not always in prenex normal form.
-\begin{theorem}
+\begin{theorem} \label{theo:eqso-sum-normal-form}
 	Every formula in $\eqso(\LL)$ can be rewritten in \emph{sum normal form}.
@@
 \end{theorem}
+\martin{Creo que el resultado de que existe una formula en $\QE{1}$ que no se puede pasar a prenex está mas relacionado con el teorema siguiente que con este.}
 
@@
 Interestingly, if $\LL$ contains $\logu{1}$, then we can always convert any formula in $\Sigma$-prenex normal form. 
-\begin{theorem}
+\begin{theorem} \label{theo:prenex}
 	If $\logu{1} \subseteq \LL$, then for every formula $\alpha \in \eqso(\LL)$ there exists a formula $\beta \in \eqso(\LL)$ equivalent to $\alpha$ in $\Sigma$-prenex normal form.
@@
 Unfortunately, this is not the case for $\loge{0}$ and $\loge{1}$ as the following result shows.
-\begin{proposition}
+\begin{proposition} \label{prop:sharp-Q-rel}
 	The classes $\E{0}$ and $\E{1}$ are strictly contained in $\QE{0}$ and $\QE{1}$, respectively.
@@
 As the following result shows, $\eqso(\loge{0})$ defines only tractable counting functions and $\eqso(\loge{1})$ intractable counting functions but with an tractable decision problems.
-\begin{proposition}
+\begin{proposition} \label{prop:qe0-fp-qe1-totp-fptras}
 All functions defined in $\eqso(\loge{0})$ and $\eqso(\loge{1})$ can be computed in $\fp$ and $\totp$, respectively. Furthermore, every function defined in $\eqso(\loge{1})$ has a FPTRAS.

### 2016-12-29 LICS/framework.tex
muchos cambios

@@
 \end{example}
-\begin{example}
+\begin{example}\label{exa-perm}
 For a more involved example that includes multiplication, let $\bM = \{M(\cdot,\cdot)\}$ be the vocabulary for matrices where a structure $\fM$ over $\bM$ encodes a 0-1 matrix $A$ where $M(i,j)$ is true if $A_{i,j} = 1$ and $0$ otherwise.
@@
 
-Finally, the work on~\cite{SalujaST95} and \cite{compton1996logical} is of particular interest for our work. 
+Finally, the work on~\cite{SalujaST95} and \cite{ComptonG96} is of particular interest for our work. 
 In~\cite{SalujaST95}, it was proposed to define functions over structures by using free variables on SO-formulas and count how many extensions of the structure satisfy the formula (i.e.

### 2016-12-29 LICS/saluja.tex
muchos cambios

--- 2016-12-29 LICS/saluja.tex
@@
 
-Inspired by the connection between $\shp$ and $\sfo$, the authors in~\cite{SalujaST95} studied subclasses of  $\sfo$ syntactically restricting the alternation quantifiers in boolean formulas defining what we call the $\sfo$-hierarchy.
-Specifically, they define the classes $\E{i}$ and $\U{i}$ where functions are defined by $\fo$-formulas in $\loge{i}$ and $\logu{i}$, respectively, for every $i \geq 0$. 
-Interestingly, in~\cite{SalujaST95} they show that this function classes defined a finite hierarchy of the form:
+Inspired by the connection between $\shp$ and $\sfo$, a hierarchy of subclases of $\sfo$ was introduced in~\cite{SalujaST95} 
+%studied subclasses of  $\sfo$ syntactically 
+by restricting the alternation of quantifiers in Boolean formulas.
+%, defining what we call 
+Specifically, the $\sfo$-hierarchy consists of the 
+%they define 
+the classes $\E{i}$ and $\U{i}$ for every $i \geq 0$, where $\E{i}$ (resp., $\U{i}$) is defined as $\sfo$ but restricting the formulas used to be in $\loge{i}$ (resp., $\logu{i}$).
+%is there exists a for are defined where functions are defined by $\fo$-formulas in $\loge{i}$ and $\logu{i}$, respectively, 
+%for every $i \geq 0$. 
+By definition we have that $\U{0} = \E{0}$. Moreover, it is shown in~\cite{SalujaST95} that:
+% this function classes defined a finite hierarchy of the form:
 \[
@@
 \]
-In this section, we study a natural extension of these classes in light of the $\eqso$ logic.
-Formally, we define the classes $\eqso(\loge{i})$ and $\eqso(\logu{i})$ where the boolean logic is restricted to $\loge{i}$ and $\logu{i}$, respectively.
-We denote each class by $\QE{i}$ and $\QU{i}$ for short.
-
-In the next result, we show that $\eqso$ also defined a finite hierarchy similar than in~\cite{SalujaST95} that we called the $\eqso$-hierarchy.
+In light of the framework introduced in this paper, natural extensions of these classes are obtained by considering 
+%in light of the $\eqso$ logic. Specifically, we consider 
+%the classes 
+$\eqso(\loge{i})$ and $\eqso(\logu{i})$ for every $i \geq 0$, which form the $\eqso$-hierarchy.
+%, where the boolean logic is restricted to $\loge{i}$ and $\logu{i}$, respectively. 
+%We denote each class by $\QE{i}$ and $\QU{i}$ for short. 
+As our first result, we show that in terms of containment the $\eqso$-hierarchy behaves as the $\sfo$-hierarchy:
+%$\eqso$ also defined a finite hierarchy similar than in~\cite{SalujaST95} that we called the $\eqso$-hierarchy.
 \begin{proposition}
-\[
-\QE{0} \; \subsetneq \; \QE{1} \; \subsetneq \; \QU{1} \; \subsetneq \; \QE{2} \; \subsetneq \; \QU{2} \; = \; \eqso(\fo)
-\]
+\begin{multline*}
+\; \QE{0} \; \subsetneq \; \QE{1} \; \subsetneq \; \QU{1} \; \subsetneq \\ \QE{2} \; \subsetneq \; \QU{2} \; = \; \eqso(\fo)
+\end{multline*}
 \end{proposition}
-A natural question until this point is what is the connection between $\E{i}$ and $\U{i}$ and their corresponding classes $\QE{i}$ and $\QU{i}$.
+As our second result, we establish precise connections between 
+%A natural question at this point is what is 
+%the connection between 
+$\E{i}$ and $\U{i}$ and their corresponding classes $\QE{i}$ and $\QU{i}$.
 Clearly, we have that $\E{i} \subseteq \QE{i}$ and $\U{i} \subseteq \QU{i}$.
-Indeed, each formula $\varphi(\bar{X}, \bar{x})$ in $\E{i}$ or $\U{i}$ is equivalent to $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ which is in $\QE{i}$ or $\QU{i}$, respectively.
-Then it is left to know whether the containment between the sharp and quantitative classes is strict or not. 
-For this, we study first whether each formula $\alpha$ in $\eqso(\LL)$ can be transformed in a \emph{prefix normal form} where all sums quantifiers are restricted to the beginning of the formula. 
-Formally, we say that a formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\Sigma$-prenex normal form} (or just prenex normal form) if $\alpha$ is of the form:
-\[
-\alpha := \sa{\bar{X}} \sa{\bar{x}} \varphi(\bar{X}, \bar{x})
-\]
-where $\bar{X}$ and $\bar{x}$ is a sequence of $0$ or more second and first order variables, respectively.
-In particular, note that a boolean formula is always in $\Sigma$-prenex normal form.
-Then we say that a formula $\alpha$ is in \emph{sum normal form} if $\alpha$ is of the form $c + \Sigma_{i=1}^n \alpha_i$ where $c$ is a non-negative constant and each $\alpha_i$ is in $\Sigma$-prenex normal form.
+Indeed, each formula $\varphi(\bar{X}, \bar{x})$ in $\E{i}$ is equivalent to the formula $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ in $\QE{i}$, and likewise for $\U{i}$ and $\QU{i}$.
+Then it is left to know whether these containments are strict.
+%between the sharp and quantitative classes is strict or not. 
+For this, we study first whether a formula $\alpha$ in $\eqso(\LL)$ can be transformed into a 
+%\emph{prenex 
+normal form where sum quantifiers are restricted to be at the beginning of the formula. 
+Formally, we say that a formula $\alpha$ in $\eqso(\LL)$ is in \emph{prenex normal form} 
+%(or just prenex normal form) 
+if $\alpha$ is of the form
+%\[
+%\alpha := 
+$\sa{\bar{X}} \sa{\bar{x}} \varphi(\bar{X}, \bar{x})$,
+%\]
+where $\bar{X}$ and $\bar{x}$ are sequences of zero or more second-order and first-order variables, respectively, and $\varphi(\bar{X}, \bar{x})$ is a formula in $\LL$.
+In particular, 
+%note that a 
+each Boolean formula is in 
+%$\Sigma$-
+prenex normal form.
+Moreover, we say that a formula $\alpha$ is in \emph{sum normal form} if $\alpha$ is of the form $c + \Sigma_{i=1}^n \alpha_i$, where $c$ is a non-negative constant and each $\alpha_i$ is in 
+%$\Sigma$-
+prenex normal form.
 Next we show that each formula in $\eqso(\LL)$ can be converted in \emph{sum normal form} but not always in prenex normal form.
-\begin{theorem}
+\begin{theorem}\label{theo-pnf-snf}
 Every formula in $\eqso(\LL)$ can be rewritten in \emph{sum normal form}.
-Furthermore, there exists a formula $\alpha \in \QE{1}$ such that there does not exist a formula in $\Sigma$-prenex normal form in $\QE{1}$ equivalent to $\alpha$.
+Furthermore, there exists a formula $\alpha \in \QE{1}$ that is not equivalent to any formula in $\QE{1}$ that is in prenex normal form.
+%such that there does not exist a formula in $\Sigma$-prenex normal form in $\QE{1}$ equivalent to $\alpha$.
 \end{theorem}
 
-By the previous result, we know that there exists logics $\LL$ where no $\Sigma$-prenex normal form exists. 
-Then an interesting question here is when can we convert a formula $\alpha$ in $\Sigma$-prenex normal form. 
-Of course, this would depend on the expressibility of $\LL$.
-Interestingly, if $\LL$ contains $\logu{1}$, then we can always convert any formula in $\Sigma$-prenex normal form. 
-\begin{theorem}
-	If $\logu{1} \subseteq \LL$, then for every formula $\alpha \in \eqso(\LL)$ there exists a formula $\beta \in \eqso(\LL)$ equivalent to $\alpha$ in $\Sigma$-prenex normal form.
+%By the previous result, we know that there exists logics $\LL$ where no $\Sigma$-prenex normal form exists. 
+%Then 
+An interesting question at this point is when a formula $\alpha \in \eqso(\LL)$ can be converted in 
+%$\Sigma$-
+prenex normal form. 
+%Of course, this would depend on the expressibility of $\LL$.
+Interestingly, if $\LL$ contains $\logu{1}$, then it can be shown that $\alpha$ can be converted in
+%we can always convert any formula in $\Sigma$-
+%prenex 
+this normal form. 
+\begin{theorem}\label{theo-pi1-pnf}
+	If $\logu{1} \subseteq \LL$, then 
+	%for 
+	every formula in
+	%$\alpha \in \eqso(\LL)$ 
+	$\eqso(\LL)$ can be rewritten in 
+	%there exists a formula $\beta \in \eqso(\LL)$ equivalent to $\alpha$ in $\Sigma$-
+	prenex normal form.
 \end{theorem}
 
-The previous results gives the connection between the hierarchy in \cite{SalujaST95} and the $\eqso$-hierarchy. 
-Indeed, if $\LL$ contains $\logu{1}$ we will have that $\sh{\LL}$ is equivalent to $\eqso(\LL)$ since each formula in $\eqso(\LL)$ can be converted in $\Sigma$-prefix normal form and, therefore, it is equivalent to a formula in $\sh{\LL}$.
-Unfortunately, this is not the case for $\loge{0}$ and $\loge{1}$ as the following result shows.
+Theorems \ref{theo-pnf-snf} and \ref{theo-pi1-pnf} are instrumental in answering our question of what is the relationship between the $\sfo$-hierarchy and the $\eqso$-hierarchy. 
+%The previous results gives the connection between the hierarchy in \cite{SalujaST95} and the $\eqso$-hierarchy. 
+Indeed, if $\LL$ contains $\logu{1}$, then we have that $\sh{\LL}$ is equal to $\eqso(\LL)$ since each formula in $\eqso(\LL)$ can be converted in prefix normal form and, therefore, it is equivalent to a formula in $\sh{\LL}$.
+The following proposition summarizes these results, also including the cases of $\E{0}$ and $\E{1}$.
+%Unfortunately, this is not the case for $\loge{0}$ and $\loge{1}$ as the following result shows.
 \begin{proposition}

### 2016-12-29 LICS/syntactic.tex
muchos cambios

@@
 %!TEX root = main.tex
+
+The class $\shp$ was introduced in \cite{Valiant79} to prove that computing the permanent of a matrix, as defined in Example \ref{exa-perm}, is a difficult problem.
+More specifically, it was shown in  \cite{Valiant79}  that this problem is $\shp$-complete.
+As a consequence of this result the problem of computing the number of perfect matchings in a bipartite graph was also shown to be $\shp$-complete.
+Since then  many counting problems have been proved to be $\shp$-complete \cite{V79b,PB83,P86,L86,BW91,HMRS98,BW05,DS12, PS13,PS14}.
+Among them, problems having easy decision counterparts play a fundamental role, as a counting problem with a hard decision version is expected to be hard.
+A first prominent example of such problems is counting the number of perfect matching in a bipartite graph, as it is well-known that the problem of verifying whether there exists a perfect matching in a bipartite graph can be solved in polynomial time.
+Other prominent examples of such problems include counting the number of: satisfying assignments of a 2-CNF propositional formula \cite{V79b}, satisfying assignments of a DNF propositional formula \cite{DHK05}, simple paths from a source node to a target node in a directed graph \cite{V79b}, extension of a partial order to a linear order \cite{BW91} and Eulerian cycles in an undirected graph \cite{BW05}.
+
+Counting problems with easy decision versions play a fundamental role in the search of efficient approximations algorithms for functions in $\shp$.
+A fully-polynomial randomized approximation scheme (FPRAS) for a function $f : \Sigma^* \to \bbN$ is a randomized algorithm ${\cal A} : \Sigma^* \times (0,1) \to \bbN$ such that: (1) for every string $x \in \Sigma^*$ and real value $\varepsilon \in (0,1)$, the probability that $|f(x) - {\cal A}(x,\varepsilon)| \leq \varepsilon \cdot f(x)$ is at least $\frac{3}{4}$, and (2) the running time of ${\cal A}$ is polynomial in the size of $x$ and $1/\varepsilon$ \cite{KL83}.
+Notably, there exist $\shp$-complete functions that can be efficiently approximated as they admit FPRAS; for instance, there exist FPRAS for the problems of counting the number of satisfying assignments of a DNF propositional formula \cite{KL83} and the number of perfect matchings of a bipartite graph \cite{JSV04}.
+A key observation here is that if a $\shp$-complete function admits an FPRAS, then its associated decision problem is in the complexity class $\bpp$ (Bounded-Error Probabilistic Polynomial-Time).
+Hence, under standard complexity-theoretical assumptions we can no hope for an FPRAS for a function in $\shp$ whose decision counterpart is $\np$-complete, and we have to concentrate on the class of counting problems with easy decision versions (in $\bpp$ or in a lower complexity class such as $\ptime$).
+
+The importance of the class counting problems with easy decision counterparts has motivated the search of robust definitions of classes of functions in $\shp$ with easy decision versions \cite{PagourtzisZ06}.
+In this section, we use the framework developed in this paper to address this problem.
+More specifically, we introduce in Section \ref{sec-hier-shp} a hierarchy of
+
+consider several fragments of $\eqso(\LL)$ where $\LL$ is a boolean logic contained in $\fo$, and we study
+
+% It should be noted that such class can be directly defined as the set functions f ?
+#P such that Lf ?
+P, which is denoted as #Pe in [50, 51].
+However, such a definition does not lead to a well-behaved and robust function complexity class.
+In particular, for every function f ?
+#P, we have that f + 1 is trivially in #Pe, which is an undesirable property.
+This has led to the introduction of the more robust class TotP, which is defined as the class of functions f for which there exists a non-deterministic Turing machine M running in polynomial time such that, f(x) is the result of subtracting 1 to the number of (non-necessarily accepting) runs of M with input x [40].
+In [51], it is proved that TotP ?
+#Pe and that TotP has a complete function problem under parsimonious reductions.
+However, no natural problem is known to be TotP-complete under this type of reductions [51].
 
@@
 
+\cite{OH93,FH08}
+
 From this point on, for each fragment $\FF$ of $\qso$, we will also use $\FF$ to refer to the class of functions defined by the formulas in $\FF$.
 
+\subsection{A counting hierarchy below $\shp$}
+\label{sec-hier-shp}
+\input{saluja}
+
 \subsection{Counting hierarchy below $\shp$}
-\input{saluja}
 

### 2016-12-29 LICS/appendix-syntactic.tex

--- 2016-12-29 LICS/appendix.tex
@@
-%!TEX root = main.tex
+\subsection*{Proof of Proposition \ref{prop:eqso-hierarchy-proper}} %V.1
+Classes QE0, QE1, QU1, QE2, QU2 = QFO form a properly contained hierarchy:\vspace{1em}
 
-\subsection{Notation for the appendix}
+\subsection*{Proof of Theorem \ref{theo:eqso-sum-normal-form}} %V.2
+Every formula in $\eqso(\LL)$ can be rewritten in \emph{sum normal form}.
+Furthermore, there exists a formula $\alpha \in \QE{1}$ such that there does not exist a formula in $\Sigma$-prenex normal form in $\QE{1}$ equivalent to $\alpha$:\vspace{1em}
 
-For a given signature $\R$, we define $\ostr[\R]^*$ as $$\ostr[\R]^* = \{(\A,v,V) \mid \A\in\ostr[\R]\text{, $v$ ($V$) is a first-order (second-order) assignment for $\A$}  \}.$$
-The {\em conditional count} symbol $(\varphi \mapsto \alpha)$ is defined as $(\neg\varphi + (\varphi\cdot\alpha))$ for given $\so$ formula $\varphi$ and $\qso$ formula $\alpha$.
-Note that for each $(\A,v,V) \in \ostr[\R]^*$,
-$$
-\sem{(\varphi \mapsto \alpha)}(\A,v,V) = 
-\begin{cases}
-\sem{\alpha}(\A,v,V) &\text{if } (\A,v,V)\models\varphi,\\
-0 &\text{otherwise}.
-\end{cases}
-$$
-We will use the symbol $<$ also to denote the lexicographic order over same-sized tuples.
-If $\bar{x} = (x_1,\ldots,x_m)$ and $\bar{y} = (y_1,\ldots,y_m)$ are tuples of first-order variables, we denote $\bar{x} < \bar{y}$ for the formula $\bigvee_{i = 1}^m[\bigwedge_{j = 1}^{i-1}x_j = y_j \wedge x_i < y_i]$.
-Similarly, we use $=$ to denote equality between tuples, as $\bar{x} = \bar{y}$ denotes $\bigwedge_{i = 1}^m(x_i = y_i)$, and also $\bar{x}\leq\bar{y}$ denotes $\bar{x} < \bar{y} \vee \bar{x} = \bar{y}$.
-We also denote $\min(\bar{x}) := \forall\bar{y}(\bar{y} < \bar{x}\vee \bar{y} = \bar{x})$.
+\subsection*{Proof of Theorem \ref{theo:prenex}}%V.3
+If $\logu{1} \subseteq \LL$, then for every formula $\alpha \in \eqso(\LL)$ there exists a formula $\beta \in \eqso(\LL)$ equivalent to $\alpha$ in $\Sigma$-prenex normal form:\vspace{1em}
 
-If $\bar{x} = (x_1,\ldots,x_m)$ ($\bar{X} = (X_1,\ldots,X_m)$) is a tuple of first-order (second-order) variables, we denote $\sa{\bar{x}}\alpha$ for $\sa{x_1}\cdots\sa{x_m}\alpha$ and $\sa{\bar{X}}\alpha$ for $\sa{X_1}\cdots\sa{X_m}\alpha$ for each $\qso$ formula $\alpha$.
-We also denote $\length(\bar{x}) = m$ ($\length(\bar{X}) = m$).
-\subsection*{Proof of Theorem \ref{no-mult}}
+\subsection*{Proof of Proposition \ref{prop:sharp-Q-rel}}%V.4
+The classes $\E{0}$ and $\E{1}$ are strictly contained in $\QE{0}$ and $\QE{1}$, respectively.
+Moreover, the classes $\U{1}$, $\E{2}$, and $\U{2}$ are equivalent with $\QU{1}$, $\QE{2}$, and $\QU{2}$, respectively:\vspace{1em}
 
-Let $\{+, \cdot, \sa{}, \pa{}\}$ be the set of {\it algebraic operators} in $\qso$.
-We will prove by induction over the number of algebraic operators in a $\qso$-formula $\alpha$ that there exists some $\qso$-formula $\alpha'$ that respects the stated condition.
-Let $\text{op}(\alpha)$ be the number of algebraic operators in $\alpha$.
-The case $\text{op}(\alpha) = 0$ is trivial.
-Let $n\in\nat$ be such that for each $\qso$-formula $\beta$ such that $\text{op}(\beta) < n$ there is some $\beta'$ that respects the condition.
-Let $\alpha$ be a $\qso$-formula such that $\text{op}(\alpha) = n$.
-Note that if $\alpha$ is of the form $\alpha = (\beta_1 \add \beta_2)$, $\alpha = \sa{x} \beta$, $\alpha = \pa{x} \beta$, $\alpha = \sa{X} \beta$ or $\alpha = \pa{X} \beta$ the proof follows directly.
-Therefore we assume that $\alpha = (\beta_1 \mult \beta_2)$.
-We separate the proof in two main cases:
+\subsection*{Proof of Proposition \ref{prop:qe0-fp-qe1-totp-fptras}} %V.5
+All functions defined in $\eqso(\loge{0})$ and $\eqso(\loge{1})$ can be computed in $\fp$ and $\totp$, respectively. Furthermore, every function defined in $\eqso(\loge{1})$ has a FPTRAS:\vspace{1em}
 
-\begin{enumerate}
+\subsection*{Proof of Theorem \ref{theo:binary-prod}}%V.6
+If $\LL$ is closed under conjunction, then binary product can be defined in $\eqso(\LL)$:\vspace{1em}
 
-	\item At least one of $\beta_1, \beta_2$ is not of the form $\pa{x} \gamma$, $\pa{X} \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. Without loss of generality, let $\beta_1$ be such one.
-
-	\begin{enumerate}
-
-		\item $\beta_1 = s \in \nat$.
-Then we use $\alpha' = \beta_2 \add \cdots \add \beta_2$, where $\beta_2$ is repeated $s$ times, and if $s = 0$, take $\alpha' = \perp$.
-Since each of the formulas paired by $\add$ has less than $n$ algebraic operators, the statement holds.
-
-		\item $\beta_1 = (\gamma_1 \add \gamma_2)$.
-Then, we use $\alpha' = (\gamma_1\mult\beta_2 + \gamma_2\mult\beta_2)$ and each of the formulas paired by $\add$ have less than $n$ algebraic operators so the statement holds.
-
-		\item $\beta_1 = \sa{x} \gamma$. Then, we use $\alpha' = \sa{x} (\gamma\mult\beta_2)$ and as in the previous case the statement holds.
-
-		\item $\beta_1 = \sa{X} \gamma$, which we treat exactly as the previous one.
-
-	\end{enumerate}
-
-	\item Both of $\beta_1$ and $\beta_2$ are of the form $\pa{x} \gamma$, $\pa{X} \gamma$, or $\varphi$, where $\varphi$ is an $\so$-formula. This case also separates in further cases.
-
-	\begin{enumerate}
-
-		\item $\beta_1 = \pa{x} \gamma_1(x)$ and $\beta_2 = \pa{y} \gamma_2(y)$. Then we use $\alpha' = \pa{x}(\gamma_1(x)\mult\gamma_2(x))$, and the statement holds.
-
-		\item $\beta_1 = \pa{X} \gamma_1(X)$ and $\beta_2 = \pa{Y} \gamma_2(Y)$. [something about arities].
-
-		\item $\beta_1 = \pa{X} \gamma_1(X)$ and $\beta_2 = \pa{x} \gamma_2(x)$ (w.l.o.g.). [use singleton predicates].
-
-		\item $\beta_1 = \pa{x} \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). Then, we use $\alpha' = \pa{x} (\gamma \cdot \varphi)$, and the statement holds.
-
-		\item $\beta_1 = \pa{X} \gamma$ and $\beta_2 = \varphi$, which is an $\so$-formula (w.l.o.g.). We treat this case exactly as the previous one.
-
-		\item $\beta_1 = \varphi_1$ and $\beta_2 = \varphi_2$ both $\so$-formulas. Then, we use $\alpha' = (\varphi_1 \wedge \varphi_2)$ and the statement holds.
-
-	\end{enumerate}
-
-\end{enumerate}
-
->>>>>>> 601d7da9913124a5a6c9adf12a847ca836fb40f5
-
-%\end{enumerate}
-
-\subsection*{Proof of Theorem \ref{captfp}}
-
-For the first condition, let $\alpha\in\qfo(\lfp)$ over some signature $\R$.
-Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{a_1,\ldots,a_n\}$.
-Replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\Sigma x \beta(x)$ is replaced by $(\beta(a_1)+\cdots+\beta(a_n))$ and $\Pi x \beta(x)$ is replaced by $(\beta(a_1)\cdot\,\cdots\,\cdot\beta(a_n))$.
-Then each sub-formula $\varphi\in\lfp$ in $\alpha$ is replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size which is evaluated and lastly given as output.
-Note that $f\in\fp$ and $f(\enc(\A)) = \alpha(\enc(\A))$.
-
-For the second condition, let $f\in \fp$ defined over some signature $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
-Let $\Phi(x_0,\ldots,x_{\ell-1})$ be a formula that represents the following.
-Let $\A$ be an ordered structure over $\R$ with domain $A$.
-Consider a linear order over $A^{\ell}$ given by $\bar{a} < \bar{b}$, for any $\bar{a},\bar{b}\in A^{\ell}$.
-Let $w$ be a string of size $n^{\ell}$ the represents $f(\enc(\A))$ in binary.
-Let $(a_1,\ldots,a_{\ell})\in A^{\ell}$ be the $m$-th tuple in the given order (starting from 0).
-$\A\models\Phi(a_1,\ldots,a_{\ell})$ if an only if the $m$-th bit in $w$ is 1, from least to most significant.
-Since this condition can be computed in polynomial time, and $\lfp$ captures $\ptime$, this formula can be expressed in $\lfp$.
-We use
-$$
-\alpha = \sa{x_1} \cdots \sa{x_{\ell}} \Phi(x_1,\ldots,x_{\ell})\cdot\varphi(x_1,\ldots,x_{\ell}),
-$$
-where $\varphi(x_1,\ldots,x_{\ell}) := \pa{y_1}\cdots\pa{y_{\ell}}((y_1,\ldots,y_{\ell}) < (x_1,\ldots,x_{\ell}) \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
-Adding this values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\alpha(\enc(\A)) = f(\enc(\A))$.
-\qed
-
-\subsection*{Proof of Theorem \ref{qfo-pfp-cap}}
-
-\martin{no revisar, es un copy-paste de arriba}
-
-For condition (2) in \ref{cap-def}, let $\alpha\in\qfo(\pfp)$ over some $\R$.
-Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{a_1,\ldots,a_n\}$.
-Replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\Sigma x \beta(x)$ is replaced by $(\beta(a_1)+\cdots+\beta(a_n))$ and $\Pi x \beta(x)$ is replaced by $(\beta(a_1)\cdot\,\cdots\,\cdot\beta(a_n))$.
-Then each sub-formula $\varphi\in\pfp$ in $\alpha$ is replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size which is evaluated and lastly given as output.
-Note that $f\in\nfpspace$ and $f(\enc(\A)) = \alpha(\enc(\A))$.
-
-For condition (2) in \ref{cap-def}, let $f\in \nfpspace$ defined over some $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
-Let $\Phi(x_0,\ldots,x_{\ell-1})$ be a $\pfp$ formula that represents the following.
-For each $\A \in \ostr[\R]$, for convenience, suppose its domain is $A = \{0,\ldots,n-1\}$ with the usual order.
-Let $y$ be a string of size $n^{\ell}$ that represents the value of $f(\enc(\A))$ in binary.
-Then, for each $\ell$-tuple $(a_0,\ldots,a_{\ell-1}) \in A^{\ell}$, it holds that $\A \models \Phi(a_0,\ldots,a_{\ell-1})$ if and only if the $(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)$-th bit of $y$, from least to most significant, is 1.
-We use
-$$
-\alpha = \Sigma x_0 \cdots \Sigma x_{\ell-1} \Phi(x_0,\ldots,x_{\ell-1})\cdot\varphi_{\ell-1}(x_{\ell-1})\cdot\,\cdots\,\cdot\varphi_0(x_0).
-$$
-where $\varphi_i(x) = \Pi y[(y < x)\mapsto\Pi z_1\cdots\Pi z_i\,2]$ for $i > 0$ and $\varphi_0(x) = \Pi y[(y < x)\mapsto 2]$.
-Let $\A$ the previously defined structure.
-Note that for each $a\in A$, then $\sem{\varphi_i(x)}(\A,a) = 2^{n^i a}$.
-Therefore, for each $(a_0,\ldots,a_{\ell-1})\in A^\ell$, we have that
-$$
-\sem{\Phi(x_0,\ldots,x_{\ell-1})\cdot\varphi_{\ell-1}(x_{\ell-1})\,\cdots\,\varphi_0(x_0)}(\A,a_0,\ldots,a_{\ell-1}) = 
-\begin{cases}
-2^{(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)} &\text{if } (\A,a_0,\ldots,a_{\ell-1})\models\Phi \\
-0 &\text{otherwise.}
-\end{cases}
-$$
-and adding these values gives $f(\enc(\A))$. Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\alpha(\enc(\A)) = f(\enc(\A))$. \qed
-
-\subsection*{Proof of Theorem \ref{eqso-fo-cap}}
-
-For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\fo$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial time, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.%%If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Let $\alpha$ be a formula in $\eqso(\fo)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shp$ function over $\R$ and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
-
-For condition (2) in \ref{cap-def}, note that Saluja et al.
-\cite{SalujaST95} proved that $\shp = \#\fo$.
-We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso(\fo)$ be its corresponding formula.
-Since this formula is also in $\eqso(\fo)$, the condition holds.
-\qed
-
-\subsection*{Proof of Theorem \ref{eqso-eso-cap}}
-
-For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic outputs for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-On each of the steps the algorithm adds the current configuration to the output tape.
-If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time without modifying the output tape, and accepts if and only if the condition holds true.
-If $\alpha$ is a constant $s$, then the algorithm produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-%% If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\spp$ function over $\R$ and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
-
-For condition (2) in \ref{cap-def}, note that Compton and Graedel \cite{ComptonG96} proved that $\spp = \#\eso$.
-We also have that $\sqso(\eso)$ captures $\#\eso$ over ordered structures so for each $f\in \spp$ let $\alpha \in \sqso(\eso)$ be its corresponding formula.
-Since this formula is also in $\eqso(\eso)$, the condition holds.
-\qed
-
-\subsection*{Proof of Theorem \ref{qso-pfp-cap}}
-
-For condition (1) in \ref{cap-def}, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\pfp$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space \cite{AbiteboulV89}\martin{no estoy seguro de esta cita}, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-If $\alpha = \pa{X}\beta$, it simulates $M_{\beta}$ on input $(\A,v,V[B/X])$ consecutively for each $B\in A^{arity(X)}$.
-This covers all possible cases for $\alpha$, and each of these steps can be computed in polynomial space.
-Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$ \cite{Ladner89}, and that for every $\A\in\ostr[\R]$: $f(\enc(\A)) = \alpha(\enc(\A))$.
-
-For the second condition, let $f\in \fpspace$ defined over some $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq 2^{n^\ell}$, where $\A$ has a domain of size $n$.
-Let $\Phi(X)$ be a $\pfp$ formula that represents the following.
-Consider an order over the subsets of $A^{\ell}$ given by the formula $$\varphi_{<}(X,Y) = \exists\bar{u}\big[\neg X(\bar{u})\wedge Y(\bar{u})\wedge \forall\bar{v}\big(
-\bar{u}<\bar{v}\to(X(\bar{u})\iff Y(\bar{v}))\big)\big].$$
-
-Let $B$ be the $p$-th subset with respect to this order and let $w$ be a string of size $2^{n^{\ell}}$ which represents the value of $f(\enc(\A))$ in binary.
-Then $\A\models\Phi(B)$ if and only if the $p$-th bit in $w$ is 1, from least to most significant.
-Since this condition can be verified in polynomial space, and $\pfp$ captures $\pspace$, then $\Phi$ can be expressed in $\pfp$.
-We define
-$\alpha := \sa{X}(\Phi(X)\mult\varphi(X))$, where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
-Note that for each $B\subseteq A^{\ell}$ such that $B$ is the $m$-th element in the given order, $\sem{\varphi(B)}(\A) = 2^m$.
-Therefore, $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-\qed
-
-\subsection*{Proof of Theorem \ref{eqso-sigma-zero-in-fp}}
-
-First, for a given signature $\R$ and logic fragment $\LL$ we define a {\em valid form in $\LL$} (or just {\em valid form}).
-A formula $\alpha$ is in valid form in $\LL$ if (1) it is equal to $\varphi$, $1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}1$ for some $\LL$ formula $\varphi$ and some tuple of first-order (second-order) variables $\bar{x}$ ($\bar{X}$), or (2) if it is equal to $(\beta\add\gamma)$ where $\beta$ and $\gamma$ are formulas in valid form.
-Given a $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$ we will recursively construct a formula $\tau(\alpha)$ in valid form such that $\alpha(\enc(\A)) = \tau(\alpha)(\enc(\A))$ for each $\A\in\ostr[\R]$.
-
-First, each natural constant $s$ in $\alpha$ is replaced by $(1\add 1\add\cdots\add 1)$ ($s$ times).
-If $\alpha = 1$ or $\alpha = \varphi \in \loge{0}$, then $\tau(\alpha) = \alpha$.
-We assume that every sub-formula of $\alpha$ is in valid form.
-If $\alpha = (\beta \add \gamma)$ then $\tau(\alpha) = \alpha$.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ this case separates in more cases: If $\alpha = \sa{x}\beta$ and $\beta$ is in one of the valid forms in (1), then $\tau(\alpha) = \alpha$.
-If $\alpha = \sa{X}\beta$ and $\beta$ is in the valid form $\sa{\bar{x}}\gamma$ then $\tau(\alpha) = \sa{\bar{x}}\sa{X}\gamma$; if $\beta$ it is in one of the remaining valid forms in (1), then $\tau(\alpha) = \alpha$.
-If $\alpha = \sa{x}\beta$ ($\alpha = \sa{X}\beta$) and $\beta = (\gamma_1\add\gamma_2)$, then $\tau(\alpha) = \tau(\sa{x}\gamma_1) \add \tau(\sa{x}\gamma_2)$ ($\tau(\alpha) = \tau(\sa{X}\gamma_1) \add \tau(\sa{X}\gamma_2)$).
-This covers all possible cases for $\alpha$.
-Furthermore, suppose that $\LL$ is closed under conjunction and suppose $\alpha = (\beta\mult\gamma)$.
-If $\beta = (\delta_1\add\delta_2)$ (wlog), then $\tau(\alpha) = \tau(\delta_1\mult\gamma)\add\tau(\delta_2\mult\gamma)$.
-If $\beta = \sa{\bar{x}}\delta$, then $\tau(\alpha) = \tau(\sa{\bar{x}}\tau(\delta\mult\gamma))$.
-If $\beta = 1$, then $\tau(\alpha) = \gamma$.
-If $\beta = \varphi$ and $\gamma = \psi$, then $\tau(\alpha) = \varphi\wedge\psi$.
-Let $\alpha$ be a $\eqso(\loge{0})$ formula over a signature $\R$.
-Consider the following procedure that receives $\enc(\A)$ as input, where $\A\in\ostr[\R]$ with domain $A$.
-Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_m$.
-For each $\beta_i$, if $\beta_i = 1$ or $\beta_i = \varphi$, then $\sem{\beta_i}(\A)$ can be computed in polynomial time.
-If $\beta_i = \sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}$, if $\beta_i = \sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = 2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, if $\beta_i = \sa{\bar{X}}\sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, all of which can be computed in polynomial time.
-Otherwise, $\beta_i$ is $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, and then $\sem{\beta_i}(\A) = f_{\varphi}(\enc(\A))$, and $f_{\varphi}\in \#\loge{0}$ which implies that $f_{\varphi}(\enc(\A))$ can be computed in polynomial time \cite{SalujaST95}.
-
-\subsection*{Proof of Theorem \ref{eqso-sigma-one-in-eqso-pi-one}}
-
-Let $\alpha$ be a formula in $\eqso(\loge{1})$ over a signature $\R$.
-Let $\tau(\alpha) = \beta_1\add\beta_2\add\cdots\add\beta_m$ defined as in the proof of Theorem \ref{eqso-sigma-zero-in-fp}.
-This is, each $\beta_i$ is of the form $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}\sa{\bar{x}}1$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, where $\varphi$ is a formula in $\loge{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables.
-For each $\beta_i$ we will construct a $\gamma_i$ such that $\gamma_i \in \eqso(\logu{1})$ and $\sem{\beta_i}(\A) = \sem{\gamma_i}(\A)$ for each $\A\in\ostr[\R]$.
-Since the forms $1$, $\sa{\bar{x}}1$, $\sa{\bar{X}}1$ and $\sa{\bar{X}}\sa{\bar{x}}1$ are already valid in $\eqso(\logu{1})$, we suppose $\beta_i$ is in one of the remaining forms.
-Without loss of generality, let $\beta_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y},\bar{X})$ where $\varphi(\bar{x},\bar{y},\bar{X})$ is a quantifier-free formula.
-Similarly as in the proof by Saluja et al.
-\cite{SalujaST95}, we define $\gamma_i = \sa{\bar{x}}\sa{\bar{y}}\sa{\bar{X}}\varphi(\bar{x},\bar{y},\bar{X}) \wedge \forall\bar{z}(\varphi(\bar{x},\bar{z},\bar{X})\to(\bar{z} < \bar{y} \vee \bar{z} = \bar{y}))$ that fulfils the desired condition.
-The theorem holds true by using $\beta = \gamma_1\add\cdots\add\gamma_m$ which is in $\eqso(\logu{1})$
-
-%\subsection*{Proof of Theorem \ref{one-sigma-zero}}
-%
-%Towards a contradiction, suppose that there exists a constant function $f(\A) = s > 0$ which is expressible in $\E{0}$.
-This is, there is a formula $\psi(x,X)$ such that $f_{\psi(x,X)}(\A) = \vert \{ \langle e,P\rangle \mid \A\models\psi(e,P) \} \vert = s$ for each $\A\in\ostr$.
-It can be seen that if $X$ contains more than zero variables, then the function $f_{\psi(x,X)}$ grows exponentially.
-We have now that $\psi(x)$ contains only first order variables.
-Consider some $\A \in \ostr$ of size $n$ and let $a_1,\ldots,a_s\in \A$ be the tuples that satisfy $\A \models \psi(a_i)$.
-Now consider a structure $\A'$ that is obtained by duplicating $\A$.
-This is, $\A'$ has size $2n$, its first half is isomorphic to its second half, and also to $\A$.
-We have that there is at least $2s$ tuples $a_i,b_i\in \A'$, $s$ on each half, such that $\A'\models\psi(a_i)$ and $\A'\models\psi(b_i)$, which leads to a contradiction.
-
-%\subsection*{Proof of Theorem \ref{mult-sigma-zero}}
-%
-%Let $\alpha \in \sqso(\loge{0})$ with binary product.
-This is, $\alpha$ is defined by the grammar $\alpha := \sa{x}\varphi \ \mid \ \sa{X}\varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \ (\alpha\mult\alpha)$, where $\varphi \in \loge{0}$.
-For convenience, we add the expressions $\sa{x}(\varphi\mult\varphi)$ and $\sa{X}(\varphi\mult\varphi)$ to the grammar.
-Let $k$ be the number of algebraic operators in $\alpha$.
-We will show by induction on $k$ that $\alpha$ is expressible without multiplication.
-If $k = 1$, then the possible cases are $\alpha = \sa{x}\varphi$, where $\varphi\in\loge{0}$, which is already product-free.
-If $k = 2$, then the only that is not two nested summations is $\alpha = \sa{x}(\varphi\mult\psi)$, which can be expressed as $\alpha = \sa{x}(\varphi\wedge\psi)$.
-Suppose that the hypothesis holds for $k = 1$ through $n-1$.
-Let $\alpha$ have $n$ algebraic operators.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ where $\beta$ is an algebraic expression, then $\beta$ has $n-1$ algebraic operators and by hypothesis it can be expressed as product-free.
-If $\alpha = (\beta \mult \gamma)$ there are two cases.
-If at least one between $\beta$ and $\gamma$ has a product, then by hypothesis it can be expressed as product-free.
-Moreover, the new expression for $\alpha$ has strictly less than $n$ algebraic operators so by hypothesis it can be expressed as product-free.
-If there are no products in $\beta$ or $\gamma$, then let $\beta = \sa{x}\delta(x)$.
-We have that $\alpha$ can be expressed as $\sa{y}(\delta(y)\mult\gamma)$, where $y$ is a first-order variable that does not appear in $\beta$ or $\gamma$.
-Since $\delta(y)\mult\gamma$ has $n-1$ algebraic operators, it can be expressed as product-free, and then $\alpha$ can be expressed as product-free.
-%
-%\subsection*{Proof of Theorem \ref{fo-prod-sigma-zero}}
-%
-%The resulting grammar is $\alpha := \sa{x}\varphi \ \mid \ \sa{X} \varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \pa{x}\alpha$.
-Consider a more restricted grammar that replaces $\pa{x}\alpha$ by $\pa{x}\sa{y}\varphi$.
-Let $\alpha = \pa{x}\sa{y}(x = y)$.
-This formula produces a constant function equal to 1, which as shown in Theorem \ref{one-sigma-zero} is not expressible in $\E{0}$.
-%
-%\subsection*{Proof of Theorem \ref{fp1}}
-%
-%Let $f \in \XE{0}$, and let $\varphi(\X,\x)$ be and {\sc FO}-extended quantifier-free $\L$-formula such that:
-%\begin{eqnarray*}
-%	f(\A) &=& \vert \{\langle \P,\e  \rangle \mid \A \models \varphi(\P,\e) \} \vert
-%\end{eqnarray*}
-%for each $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^m$ and $\P = (P_1,\ldots,P_q)$ is a tuple of predicates.
-We will now show that computing $f(\A)$ can be done in polynomial time.
-%
-%For each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$, let $R_{\beta}$ be a predicate of arity $m$.
-Let $\A^\prime = \langle A, \S^{\A}, R_{\beta}^{\A^\prime}, \leq^{\A} \rangle \in \str$, where $R_{\beta}^{\A^\prime} = \{\d \mid \A \models \beta(\d)\}$.
-Note that each $\beta(\x)$ is fixed in $\varphi(\X,\x)$, and for each $\d \in A^m$, checking whether $\A \models \beta(\d)$ can be done in polynomial time.
-Therefore, generating $R_{\beta}^{\A^\prime}$ can also be done in polynomial time.
-%
-%Let $\psi(\X,\x)$ be obtained by replacing each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$ by $R_{\beta}(\x)$.
-Also, let $g = f_{\psi(\X,\x)}$.
-Note that for each tuple of predicates $\P$ and each $\e \in A^m$, $\A \models \varphi(\P,\e)$ if and only if $\A^\prime \models \psi(\P,\e)$, and so, $g(\A^\prime) = f(\A)$.
-But, $\psi(\X,\x)$ is a quantifier-free $\L$-formula, and therefore, $g \in \#\Sigma_0$.
-Since it is shown in \cite{DBLP:journals/jcss/SalujaST95} that $\#\Sigma_0 \subseteq$ {\sc FP}, we conclude that $f(\A)$ can be evaluated in polynomial time.
-
-\subsection*{Proof of Theorem \ref{decisionptime}}
-
-Let $f$ be a function in $\E{1}$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
-\begin{eqnarray*}
-	f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
-\end{eqnarray*}
-where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables.
-Let $\y = (\x,\z)$ and let $\psi(\X,\y) = \varphi(\x,\X,\z)$.
-Moreover, let $g = f_{\psi(\X,\y)}$.
-\begin{claim} \label{mono}
-	For each $\A \in \str$, $f(\A) > 0$ iff $g(\A) > 0$.
-\end{claim}
-\begin{proof}
-	($\Rightarrow$) Suppose $f(\A) > 0$.
-Let $\P$ and $\e$ be such that $\A \models \exists \x \, \varphi(\x,\P,\e)$.
-It follows that there is at least one $\d \in A^k$ such that $\A \models \varphi(\d,\P,\e) = \psi(\P,(\d,\e))$.
-Therefore, $g(\A) > 0$.
-
-	($\Leftarrow$) Suppose $g(\A) > 0$.
-Let $\Q$ and $\c = (\c_1,\c_2)$, where $\c_1$ and $\c_2$ have $k$ and $m$ elements respectively, be such that $\A \models \psi(\Q,\c) = \varphi(\c_1,\Q,\c_2)$.
-Then we have that $\A \models \exists \x \, \varphi(\x,\Q,\c_2)$, from which we conclude that $f(\A) > 0$.
-\end{proof}
-Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula, so $g \in\XE{0}$.
-It follows from \ref{mono} that for each $\A \in \str$, it holds that $\A \in L_f$ if and only if $g(\A) > 0$.
-But by Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time.
-From this we conclude that the decision version of $f$ is in $\ptime$.
-
-\subsection*{Proof of Theorem \ref{sub-pnp}}
+\subsection*{Proof of Theorem \ref{sub-pnp}}%V.7
+If $\eqso(\loge{i})$ or $\eqso(\logu{i})$ is closed under subtraction for $i > 0$, then {\sc P} = {\sc NP}:\vspace{1em}
 
@@
 
-%\subsection*{Proof of Theorem \ref{sigma1strict}}
-%
-%We will show that the $\E{1}$ function $f$ defined by $\varphi(x_1) = (x_1 = x_1) \wedge \forall y \, S(y)$ is not in $\#\Sigma_1$.
-By contradiction, suppose that it is.
-Let $\A = \langle A =\{1\},S^\A = \{1\},\leq^\A=\{(1,1)\} \rangle$.
-Then, $f(\A) = 1$.
-Now let $\B = \langle B =\{1,2\},S^\B = \{1\},\leq^\B=\{(1,1),(1,2),(2,2)\} \rangle$.
-Note that $\A$ is an induced substructure of $\B$.
-%
-%We have that for each function $g \in \#\Sigma_1$ and structures $\A_1,\A_2\in\str$, if $\A_1$ is an induced substructure of $\A_2$, then $g(\A_1) \leq g(\A_2)$ \cite{SalujaST95}.
-Therefore, $f(\B) \geq f(\A) = 1$.
-However, there is no assignment $s\in B$ to $x$ such that $\B\models\:\varphi(s)$, so $f(\B) = 0$, which leads to a contradiction.
-
-\subsection*{Proof of Theorem \ref{sigmafo-minusone}}
+\subsection*{Proof of Theorem \ref{sigmafo-minusone}}%V.8
 
@@
 \end{proof}
-The following claim identifies a crucial condition that is verifiable through a $\fo$. 
+The following claim identifies a crucial condition that is verifiable with a $\fo$ formula. 
 \begin{claim}
@@
 
-%%Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$.
-There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$.
-This separates the proof in three cases:
-%\begin{enumerate}
-%%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\z)$, where $\z = (z_1,\dots,z_d)$. That is,
-%%	\begin{eqnarray*}
-%%		f(\A) &=& \mid \{ \langle\e\rangle \mid \A \models \exists \x \ \varphi(\x,\e) \} \mid,
-%%	\end{eqnarray*}
-%%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^d$.
-Our goal here is to eliminate the lexicographically smallest sequence of variables, which can be done easily.
-First, let $\y = (y_1,\dots,y_k)$, $\y\,^\prime = (y_1^\prime,\dots,y_k^\prime)$ and
-%%	\begin{eqnarray*}
-%%		\varphi_{k,<}(\y\,^\prime,\y) &=& \bigvee_{i = 1}^k \left( \bigwedge_{j=1}^{i-1} y_j^\prime = y_j \wedge y_i^\prime < y_i \right).
-%%	\end{eqnarray*}
-%%	This formula is true if $\y\,^\prime$ is lexicographically smaller than $\y$. Now, let $f'$ be defined by
-%%	\begin{eqnarray*}
-%%		\varphi^\prime(\x,\z) &=& \varphi(\x,\z) \wedge \exists \z\,^\prime (\varphi(\x,\z\,^\prime) \wedge \varphi_{d,<}(\z\,^\prime,\z ) ).
-%%	\end{eqnarray*}
-%%	If $f(\A)>0$, then $f'(\A)$ will count exactly one element less than $f(\A)$.
-Otherwise, if $f(\A)=0$, then $\A \not\models\exists \x\,\varphi(\x,\e)$ for every tuple $\e$ of elements in $A$, so $\A \not\models\exists \x\,\varphi^\prime(\x,\e)$ for every $\e$ and, therefore, $f'(\A)=0$.
-Hence, $f' = f\dotminus 1,$ from which we conclude that $f\dotminus 1\in\E{1}.$
-%%	
-%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X)$ where $\x = (x_1,\dots,x_d)$ and $\X = (X_1,\dots,X_r)$. That is,
-%	\begin{eqnarray*}
-%		f(\A) &=& \mid \{ \langle\P\rangle \mid \A \models \exists \x \ \varphi(\x,\P) \} \mid \label{f1},
-%	\end{eqnarray*}
-%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$ and $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$. For the time being, suppose that
-%	\begin{eqnarray}
-%	\varphi(\x,\X) &=& \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)
-%	\end{eqnarray}
-%	where $n$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda:\{1,\ldots,n\}\to\{1,\ldots,r\}$, $\y$ is a $p$-tuple of variables in $\x$, $\varphi^{-}(\X,\y)$ is a conjunction of negated predicates in $\X$, $\theta(\x)$ defines a total order on a partition of $\x$, and $\beta(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
-Note that $\theta(\x)$ also mentions all variables in $\x$.
-We also assume that $(\x_1,\dots,\x_n,\y) = \x$.
-As an example, the following formula is of this form:
-%	\begin{multline*}
-%	\varphi(\x,\X) =  X_1(x_1,x_2) \wedge X_3(x_3) \wedge X_2(x_4,x_5) \wedge X_3(x_6) \wedge \neg X_1(x_7,x_8) \wedge \\ (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 ) \wedge \\ \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big),
-%	\end{multline*}
-%	where $\x = (x_1,x_2,x_3,x_4,x_5,x_6,x_7,x_8)$ and $\X = (X_1,X_2,X_3)$.
-Here, $n = 4$, $\lambda(1) = 1$, $\lambda(2) = \lambda(4) = 3$ and $\lambda(3) = 2$, $\x_1 = (x_1,x_2)$, $\x_2 = (x_3)$, $\x_3 = (x_4,x_5)$, $\x_4 = (x_6)$ and $\y = (x_7,x_8)$.
-Moreover, $\varphi^{-}(\X,\y) = \neg X_1(x_7,x_8)$, $\theta(\x) = (x_1 < x_2 \wedge x_1 = x_3 \wedge x_1 = x_4 \wedge x_2 = x_8 \wedge x_2 = x_5 \wedge x_8 < x_6 \wedge x_6 = x_7 )$, which defines a total order on the partition of $\x$ $\{\{x_1,x_3,x_4\},\{x_2,x_5,x_8\},\{x_6,x_7\}\}$, and $\beta(\x) = \forall z\big( S_1(x_1,z,x_2) \wedge x_3 = x_3 \wedge x_4 = x_4 \wedge x_5 = x_5 \wedge x_6 = x_6 \wedge x_7 = x_7 \wedge x_8 = x_8 \big)$.
-%	
-%	Similarly to the previous proof, we would like to eliminate the {\em lexicographically smallest}\footnote{We consider the lexicographically smallest tuple of predicates as the one in which its predicates contain the lexicographically smallest tuples and do not contain any more tuples than those} tuple of predicates that satisfies the formula \eqref{f1}.
-Let $\u_i$ be a $a_{\lambda(i)}$-tuple of variables for every $i \in \{1,\dots,n\}$, and let $m = \sum_{i = 1}^n a_{\lambda(i)}$ be the number of variables of $(\x_1,\dots,\x_n)$.
-We now define
-%	\begin{multline*}
-%	\alpha_{\min}(\u_1,\dots,\u_n) = \exists\y\Big[ \alpha(\u_1,\dots,\u_n,\y)\wedge \\ \forall\v_1\cdots\forall\v_n\forall\w\Big(\big(\alpha(\v_1,\dots,\v_n,\w)\wedge\bigvee_{i=1}^n(\u_i\neq\v_i)\big)\to (\u_1,\dots,\u_n)<(\v_1,\dots,\v_n)\Big)\Big],
-%	\end{multline*}
-%	where $\alpha(\x) = \theta(\x) \wedge \beta(\x)$.
-Note that $\alpha_{\min}$ is satisfied only by the lexicographically smallest assignment $(\d_1,\dots,\d_n)$ to $(\x_1,\dots,\x_n)$ such that $\A\models\theta(\d_1,\dots,\d_n,\l)$ and $\A\models\beta(\d_1,\dots,\d_n,\l)$ for some $\l \in A^p$.
-Our new formula is
-%	\begin{multline}
-%	\varphi^\prime(\x,\X) = \left( \bigwedge_{i=1}^n X_{\lambda(i)}(\x_i) \right) \wedge \varphi^{-}(\X,\y) \wedge \theta(\x) \wedge \beta(\x)\wedge \\ \exists\u_1\cdots\exists\u_n\bigg[\alpha_{\min}(\u_1,\dots,\u_n) \wedge \bigg(\bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\u_i) \bigg) \vee \bigvee_{i=1}^r \exists \v\Big( X_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big) \bigg) \bigg].
-%	\end{multline}
-%	We now show a result by which the main proof will follow.
-%	\begin{lemma} \label{first}
-%		$f_{\exists \x \: \varphi^\prime(\x,\X)} = f_{\exists \x \: \varphi(\x,\X)} \dotminus 1$.
-%	\end{lemma}
-%	\begin{proof}
-%		Let $\A \in \str$.
-Consider two cases: assume first that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-Let $\d = (\d_1,\dots,\d_n,\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha(\d)$, where $\d_i$ is the respective assignment to $\x_i$, for every $i\in\{1,\dots,n\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j\in[1,n]:\:\lambda(j)=i}\{\d_j\}$.
-We will show that this assignment to $\X$ is such that (a) $\A\models\exists\x\,\varphi(\x,\P),$ (b) $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$ and (c) $\P$ is the only assignment that satisfies (a) and (b).
-%		\begin{enumerate}
-%			\item[(a)] By contradiction, suppose that $\A\not\models\exists\x\,\varphi(\x,\P)$.
-That is, there is no assignment $\s$ to $\x$ such that $\A\models\varphi(\s,\P).$ Since $\d$ is such that $\A\models \bigwedge_{i=1}^n P_{\lambda(i)}(\d_i)$ and $\A\models \alpha(\d)$, it follows that $\A\not\models\varphi^{-}(\P,\ep)$ (since $\alpha(\d)=\theta(\d)\wedge\beta(\d)$).
-Therefore, there is an $i\in\{1,\ldots,n\}$ such that $\neg P_{\lambda(i)}(\d_i)$ appears in $\varphi^{-}(\P,\ep)$.
-Then, there is an $a_{\lambda(i)}$-tuple $\z$ in $\y$ such that $\neg X_{\lambda(i)}(\z)$ appears in $\varphi^{-}(\X,\x)$.
-We know that either (1) $\theta(\x)\models \z = \x_i$, (2) $\theta(\x)\models \z<\x_i$ or (3) $\theta(\x)\models \x_i<\z$.
-Considering that (2) and (3) are not possible given that both $\z$ and $\x_i$ are assigned the value $\d_i$ and $\A\models\theta(\d)$, we have that $\theta(\x)\models \z = \x_i$.
-But if this is the case, then $X_{\lambda(i)}(\x_i), \neg X_{\lambda(i)}(\z)$ and $\z = \x_i$ are all logical consequences of $\varphi(\x,\X)$, which means $\varphi(\x,\X)$ is inconsistent.
-That is, there's no structure $\A^\prime\in\str$ such that $\A^\prime\models\exists\x\:\varphi(\x,\R^\prime)$ for any assignment $\R^\prime$ to $\X$.
-In particular, $\A\not\models\exists\x\,\varphi(\x,\R^\prime)$ for every possible assignment $\R^\prime$ to $\X$, which contradicts the initial assumption that $\A\models\exists\x\,\varphi(\x,\R)$ for some assignment $\R$ to $\X$.
-%			\item[(b)] Note that if $\A \models \alpha_{\min}(\c_1,\ldots,\c_n)$, then necessarily $\c_i = \d_i$ for $i\in\{1,\ldots,n\}.$ However, by the construction of $\P$, we see that 
-%			$$\A\not\models\bigvee_{i = 1}^{n}\neg P_{\lambda(i)}(\d_i) \text{ and that } \A\not\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \d_j\Big).$$ Then, $\A\not\models\exists\x\,\varphi^\prime(\x,\P)$.
-%			\item[(c)] By contradiction, let $\P^\prime \neq \P$ be such that $\A\models\exists\x\,\varphi(\x,\P^\prime)$ and $\A\not\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-We consider two cases: first, suppose that $\P^\prime$ is missing a tuple of $\P$.
-Let $i\in\{1,\ldots,n\}$ such that in $\d_i$ is not in $P^\prime_i$.
-Then $\A\models \neg P^\prime_i(\d_i)$, and so,$$\A\models \bigg(\bigvee_{i = 1}^{n}\neg X_{\lambda(i)}(\d_i) \bigg), $$ from which we conclude that $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-Second, suppose there is some predicate $\P^\prime_i$ in $\P^\prime$ which has a tuple that $P_i$ does not have.
-If this is the case, then $$\A\models\bigvee_{i=1}^r \exists \v\Big( P_i(\v) \wedge \bigwedge_{j\in[1,n]:\: \lambda(j) = i} \v \neq \u_j\Big),$$ so $\A\models\exists\x\,\varphi^\prime(\x,\P^\prime)$.
-On both cases, we have a contradiction.
-%		\end{enumerate}
-%		With this, we conclude that for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-%		
-%		Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\R)$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi(\x,\P)$, we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\psi(\x,\P))$ for any formula $\psi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime(\x,\R)$.
-And so, for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ it holds that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ We conclude that $f_{\exists\x\,\varphi^\prime(\x,\X)} = f_{\exists\x\,\varphi(\x,\X)}\dotminus 1,$ which was to be shown.
-%	\end{proof}
-%	We now continue with the general case, in which $\varphi(\x,\X)$ is an arbirtary extended quantifier-free $\L$-formula.
-By using a standard DNF transformation algorithm and considering FO-formulas over $\L$ as literals, we can find formulas $\gamma_i(\x,\X)$ with $i\in\{1,\ldots,\ell\}$ such that
-%	\begin{eqnarray*}
-%		\varphi(\x,\X) &\equiv& \gamma_1(\x,\X) \vee \gamma_2(\x,\X) \vee \dots  \vee \gamma_{\ell}(\x,\X),
-%	\end{eqnarray*}
-%	where, for every $i\in\{1,\ldots,\ell\}$, 
-%	\begin{eqnarray*}
-%		\gamma_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \gamma^{-}_i(\X,\y_i)  \wedge \delta_i(\x).
-%	\end{eqnarray*}
-%	The function $\lambda_i$ is defined analogously to $\lambda$ of the first part of the proof.
-The tuple $\x_{i,j}$ has $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, $\y_i$ has $p_i$ variables and are such that $(\x_{i,1},\ldots,\x_{i,n_i},\y_i) = \x.$ The formulas $\gamma^{-}_i(\X,\y_i)$, $\delta_i(\x)$ are defined analogously to $\varphi^{-}(\X,\y)$ and $\beta(\x)\,$\footnote[2]{Note that $\delta_i(\x)$ can include subformulas of the form $(\u = \u)$.} respectively (see \eqref{f1}).
-Let $g$ be a function that counts the number of possible orders over partitions on a $d$-tuple of variables.
-Let the formulas $\theta^i(\x)$ for $i\in\{1,\ldots,g(d)\}$ represent each of these orders over $\x$.
-Note that for every formula $\eta(\x)$,
-%	\begin{eqnarray*}
-%		\exists\x\:\eta(\x) &\equiv& \exists\x(\eta(\x)\wedge\theta^1(\x)) \vee \cdots \vee \exists\x(\eta(\x)\wedge\theta^{g(d)}(\x)).
-%	\end{eqnarray*}
-%	We define the following formulas $\xi_i(\x,\X)$, for $i\in\{1,\ldots,m\}$ where $m = \ell \cdot g(d)$, as follows: 
-%	\begin{eqnarray*}
-%		\xi_1(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^1(\x), \\
-%		& \vdots & \\
-%		\xi_{g(d)}(\x,\X) &=& \gamma_1(\x,\X) \wedge \theta^{g(d)}(\x), \\
-%		\xi_{g(d)+1}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^1(\x), \\
-%		& \vdots & \\
-%		\xi_{2\cdot g(d)}(\x,\X) &=& \gamma_2(\x,\X) \wedge \theta^{g(d)}(\x), \\
-%		& \vdots & \\
-%		\xi_{(\ell-1) g(d)+1}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^1(\x), \\
-%		& \vdots & \\
-%		\xi_{\ell \cdot g(d)}(\x,\X) &=& \gamma_{\ell}(\x,\X) \wedge \theta^{g(d)}(\x).
-%	\end{eqnarray*}
-%	Having every disjunct with a total order allows us to eliminate the ones that are unsatisfiable for every $\L$-structure $\A$, that is, each $i\in\{1,\ldots,m\}$ such that for every assignment $\s$ to $\x,$ $f_{\xi_i(\s,\X)}(\A) = 0,$ for every $\A\in\str.$ Let $k$ be the number of disjuncts that are left after eliminating the unsatisfiable disjuncts.
-We use an injective\footnote[3]{We need this function to be injective because this way we can assure each one of the $k$ satisfiable disjuncts to be represented} function $\rho:\{1,\ldots,k\}\to\{1,\ldots,m\}$ such that $\varphi_i(\x,\X) = \xi_{\rho(i)}(\x,\X)$ is satisfiable, for every $i\in\{1,\ldots,k\}.$
-%	
-%	We can conclude that
-%	\begin{eqnarray*}
-%		\varphi(\x,\X) &\equiv& \varphi_1(\x,\X) \vee \varphi_2(\x,\X) \vee \dots  \vee \varphi_k(\x,\X),
-%	\end{eqnarray*}
-%	and for every $i\in\{1,\ldots,k\}$,
-%	\begin{eqnarray*}
-%		\varphi_i(\x,\X) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i) \wedge \theta_i(\x) \wedge \beta_i(\x),
-%	\end{eqnarray*}
-%	where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\y_i$ is a $p_i$-tuple of variables in $\x$, $\varphi_i^{-}(\X,\y_i)$ is a conjunction of negated predicates in $\X$, $\theta_i(\x)$ defines a total order on a partition of $\x$, and $\beta_i(\x)$ is an FO-formula over $\L$ which mentions all variables in $\x$.
-Let $\alpha_i(\x) = \theta_i(\x)\wedge\beta_i(\x).$
-%	
-%	%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
-%	%% CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM CLAIM
-%	
-%	\begin{claim} \label{alpha}
-%		Let $\A\in\str$ and $i\in\{1,\ldots,k\}$. For every assignment $\s$ to $\x$ such that $\A\models\alpha_i(\s)$, it holds that $f_{\varphi_i(\s,\X)}(\A) > 0.$
-%	\end{claim}
-%	\begin{proof}
-%		Let $\s = (\s_1,\ldots,\s_{n_i},\t)$ such that $\A\models\alpha_i(\s)$, and let $\P = (\P_1,\ldots,\P_r)$ where $P_i = \bigcup_{j\in[1,n_i]:\:\lambda(j)=i}\{s_i\}$.
-We will show that $\A\models\varphi_i(\s,\P)$.
-By contradiction, suppose that $\A\not\models\exists\x\,\varphi_i(\x,\P)$.
-That is, there is no assignment $\e$ to $\x$ such that $\A\models\varphi(\e,\P).$ Following the same proof as in case (a) of Lemma \ref{first}, we conclude that $\varphi_i(\x,\P)$ is inconsistent.
-However, as we mentioned previously, all of such disjuncts have been eliminated, which leads to a contradiction.
-%	\end{proof}
-%	
-%	Our plan now is to exclude the lexicographically smallest assignment $\P$ such that $\A\models\exists\x\:\varphi_1(\x,\P)$, and if there is no such $\P$, exclude the lexicographically smallest $\Q$ such that $\A\models\exists\x\:\varphi_2(\x,\Q)$, and so on.
-As we already know how to exclude that assignment in the first disjunct, we will now deal with the next disjuncts.
-Similarly to the first part of the proof, for each $i\in\{1,\ldots,k\}$ let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and let
-%	\begin{multline*}
-%	\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) = \exists\y\Big[ \alpha_i(\u_1,\dots,\u_{n_i},\y)\wedge \\ \forall\v_1\cdots\forall\v_{n_i}\forall\w\Big(\big(\alpha_i(\v_1,\dots,\v_{n_i},\w)\wedge\bigvee_{j=1}^{n_i}(\u_j\neq\v_j)\big)\to (\u_1,\dots,\u_{n_i})<(\v_1,\dots,\v_{n_i})\Big)\Big],
-%	\end{multline*}
-%	where $\u_j$ and $\v_j,$ have $a_{\lambda_i(j)}$ variables for $j\in\{1,\ldots,n_i\}$, and $\w$ has $p_i$ variables. Also, let
-%	\begin{multline*}
-%	\psi_i(\X) = \forall\x\:\neg\alpha_i(\x) \vee \exists\u_1\cdots\exists\u_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i) \bigg) \vee \bigvee_{j=1}^r \exists \v\Big( X_j(\v) \wedge \bigwedge\limits_{\substack{\ell\in[1,n_i]:\\ \lambda_i(\ell) = j}} \v \neq \u_\ell\Big) \bigg) \bigg].
-%	\end{multline*}
-%	Note that $\psi_i(\X)$ excludes only the lexicographically smallest tuple of predicates $\P$ such that $\A\models\exists\x\:\varphi_i(\P,\x),$ if there is at least one.
-In other words, every assignment $\P^\prime \neq \P$ is such that $\A\models\psi_i(\P^\prime).$ Our new formula $\varphi_i^\prime(\x,\X)$ is defined as follows:
-%	$$
-%	\varphi_i^\prime(\x,\X) = \varphi_i(\x,\X) \wedge \psi_1(\X) \wedge (\exists\v\:\alpha_1(\v)\vee\psi_2(\X)) \wedge \cdots \wedge (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{i-1}(\v)\vee\psi_i(\X)).
-%	$$
-%	Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Assume that $\A\in\str$.
-Suppose first that there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least one assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\d = (\d_1,\dots,\d_{n_q},\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha_q(\d)$, where $\d_i$ is the corresponding assignment to $\x_i$, for every $i\in\{1,\dots,n_q\}$, and $\ep$ is an assignment for $\y$.
-Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j:\lambda_q(j)=i,j\in[1,n_q]}\{\d_j\}$.
-As we did in Lemma \ref{first} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies both (a) and (b)
-%	\begin{enumerate}
-%		\item[(a)] As we showed in part (a) of Lemma \ref{first}, if there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_q(\x,\R)$, then $\A\models\exists\x\:\varphi_q(\x,\P)$ for this particular $\P$.
-However, as we showed in Claim \ref{alpha}, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_q(\s)$, then there {\em is} such an assignment to $\X$.
-The assignment $\d$ to $\x$ satisfies that $\A\models\alpha_q(\d)$, so it holds that $\A\models\exists\x\:\varphi_q(\x,\P)$.
-It immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$
-%		
-%		\item[(b)] We will show that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for (1) $i\in\{1,\ldots,q-1\}$, and (2) $i\in\{q,\ldots,k\}$.
-(1) By the choice of $q$, it holds that $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$ for every $i\in\{1,\ldots,q-1\}$ since there is no possible assignment to $\X$ for any of their sub-formulas $\varphi_i(\x,\X)$.
-(2) We can use the proof in Lemma \ref{first} to see that $\A\not\models\psi_q(\P).$ For each $i\in\{q,\ldots,k\}$, the sub-formula
-%		$$ \zeta_q(\X) =  (\exists\v\:\alpha_1(\v)\vee\cdots\vee\exists\v\:\alpha_{q-1}(\v)\vee\psi_q(\X)) $$ 
-%		appears as a conjunct in $\varphi^\prime_i(\x,\X)$.
-However, by the choice of $q$, there is no $i\in\{1,\ldots,q-1\}$ such that $\A\models\exists\v\:\varphi_i(\v,\P)$, and also $\A\not\models\psi_q(\P)$.
-It follows that $\A\not\models\zeta_q(\P)$,  so $\A\not\models\exists\x\:\varphi^\prime_i(\x,\P)$.
-And so, we conclude that $\A\not\models\exists\x\:\varphi^\prime(\x,\P).$
-%		
-%		\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b).
-As we deduce from the part (c) of Lemma \ref{first}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$.
-Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x)$, so $\A\models\psi_i(\P^\prime)$ for each $i$.
-But as we mentioned, also $\A\models\:\psi_q(\P^\prime),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which leads to a contradiction.
-%	\end{enumerate}
-%	
-%	With this, we conclude that for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)>0$, we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = f_{\exists\x\,\varphi(\x,\X)}(\A)-1.$
-%	
-%	Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$.
-Let $\P$ be an arbitrary assignment to $\X$.
-Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\chi(\x,\P))$ for any formula $\chi(\x,\P)$.
-It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$.
-And so, for every $\A\in\str$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Hence, from the results in the previous paragraph, if $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}$, we have that $f^\prime = f\dotminus 1\in\E{1}$
-%	
-%	\item Let $f \in \E{1}$ be defined by an extended quantifier free $\L$-formula $\varphi(\x,\X,\z)$, where $\x = (x_1,\dots,x_d), \X = (X_1,\dots,X_r)$ and $\z = (z_1,\dots,z_p)$. That is,
-%	\begin{eqnarray*}
-%		f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \ \varphi(\x,\P,\e) \} \mid,
-%	\end{eqnarray*}
-%	for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\P = (P_1,\ldots,P_r)$, $P_i \subseteq A^{a_i}$ for every $i \in \{1,\ldots,r\}$ and $\e \in A^p$.
-In order to prove that $f\dotminus 1\in\E{1}$, we define the formulas $\varphi_i(\x,\X,\z)$ for $i\in\{1,\ldots,k\}$ in the same way as case 2, where
-%	\begin{eqnarray*}
-%		\varphi(\x,\X,\z) \equiv \varphi_1(\x,\X,\z) \vee \cdots \vee \varphi_k(\x,\X,\z),
-%	\end{eqnarray*}
-%	and
-%	\begin{eqnarray*}
-%		\varphi_i(\x,\X,\z) &=& \left( \bigwedge_{j=1}^{n_i} X_{\lambda_i(j)}(\x_{i,j},\z_{i,j}) \right) \wedge \varphi^{-}_i(\X,\y_i,\w_i) \wedge \theta_i(\x,\z) \wedge \beta_i(\x,\z),
-%	\end{eqnarray*}
-%	where $n_i$ is the number of times a non-negated variable in $\X$ is referred to, according to the function $\lambda_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, the tuple $\x_{i,j}$ has $b_{\lambda_i(j)}$ variables and the tuple $\z_{i,j}$ has $c_{\lambda_i(j)}$ for $j\in\{1,\ldots,n_i\}$ (note that $b_{\ell} + c_{\ell} = a_{\ell}$ for $\ell\in\{1,\ldots,r\}$), $\y_i$ has $p_i$ variables, and $\w_i$ has $q_i$ variables.
-Furthermore, we have that $(\x_{i,1},\ldots,\x_{i,n_i},\y) = \x$ and $(\z_{i,1},\ldots,\z_{i,n_i},\w_i) = \z.$ The formula $\theta(\x,\z)$ defines a total order, analogously to case 2.
-The formulas $\varphi^{-}_i(\X,\y_i,\w_i)$, $\beta_i(\x,\z)$, are also defined analogously.
-%	
-%	In this case, we mix both the strategies in cases 1 and 2.
-That is, we are going to {\em isolate} the lexicographically smallest tuple of predicates that satisfies the first satisfiable disjunct, and then {\em exclude} the lexicographically smallest tuple that satisfies the isolated disjunct.
-%	
-%	Let $m_i = \sum_{j = 1}^{n_i} a_{\lambda_i(j)}$ and
-%	\begin{multline*}
-%	\alpha^i_{\min}(\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i}) = \exists\y\exists\w\Big[ \alpha_i(\x_1,\dots,\x_{n_i},\y,\z_1,\dots,\z_{n_i},\w)\wedge \\ \forall\u_1\cdots\forall\u_{n_i}\forall\s\forall\v_1\cdots\forall\v_{n_i}\forall\t\Big(\big(\alpha_i(\u_1,\dots,\u_{n_i},\s,\v_1,\dots,\v_{n_i},\t)\wedge\bigvee_{j=1}^{n_i}(\x_j\neq\u_j \vee \z_j\neq\v_j)\big)\to \\ (\x_1,\dots,\x_{n_i},\z_1,\dots,\z_{n_i})<(\u_1,\dots,\u_{n_i},\v_1,\dots,\v_{n_i})\Big)\Big],
-%	\end{multline*}
-%	and let
-%	\begin{multline*}
-%	\psi_i(\X,\z) = \forall\x\forall\v\:\neg\alpha_i(\x,\v) \vee \exists\u_1\cdots\exists\u_{n_i}\exists\w_1\cdots\exists\w_{n_i}\bigg[\alpha^i_{\min}(\u_1,\dots,\u_{n_i},\w_1,\dots,\w_{n_i}) \\ \wedge \bigg(\bigg(\bigvee_{j = 1}^{n_i}\neg X_{\lambda_i(j)}(\u_i,\w_i) \bigg) \vee \bigvee_{j=1}^r \exists \s\exists \t\Big( X_j(\s,\t) \wedge \bigwedge_{\ell\in[1,n_i]: \lambda_i(\ell) = j} \s \neq \u_\ell \vee \t \neq \w_\ell\Big) \bigg) \bigg] \vee \\ \exists\w(\exists\u\:\varphi(\u,\X,\w) \wedge \w<\z).
-%	\end{multline*}
-%	Let $\P$ and $\e$ be assignments to $\X$ and $\z$.
-Note that $\psi_i(\X,\z)$ excludes $\P$ and $\e$ only if $\P$ is the {\em lexicographically smallest} tuple of predicates (Same as case 2) such that $\A\models\exists\x\:\varphi_i(\x,\P,\d)$, for some assignment $\d$ to $\z$, and $\e$ is the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_i(\x,\P,\e).$ We define $\varphi_i^\prime(\x,\X,\z)$ in the same way as case 2:
-%	\begin{multline*}
-%	\varphi_i^\prime(\x,\X,\z) = \varphi_i(\x,\X,\z) \wedge \psi_1(\X,\z) \wedge (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\psi_2(\X,\z)) \wedge \cdots \wedge \\ (\exists\u\exists\v\:\alpha_1(\u,\v)\vee\cdots\vee\exists\u\exists\v\:\alpha_{i-1}(\u,\v)\vee\psi_i(\X,\z)).
-%	\end{multline*}
-%	Finally, let $\varphi^\prime(\x,\X,\z) = \bigvee_{i = 1}^k \varphi_i^\prime(\x,\X,\z)$.
-%	
-%	Let $q$ be the least $i\in\{1,\ldots,k\}$ such that $\A\models\exists\x\:\varphi_i(\x,\R,\d)$ for some assignment $R$ to $\X$ and some assignment $\d$ to $\z$.
-Let $\P$ be the lexicographically smallest tuple of predicates such that $\A\models\exists\x\:\varphi_q(\x,\P,\d^\prime)$ for some assignment $\d^\prime$ to $\z$.
-Let $\e$ be the lexicographically smallest assignment to $\z$ such that $\A\models\exists\x\:\varphi_q(\x,\P,\e).$ This formula is such that (a) $\A\models\exists\x\:\varphi(\x,\P,\e)$, (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P,\e)$ and $\P$ and $\e$ are the only assignments that satisfy (a) and (b).
-The proof of this is analogous to case 2.
-Therefore, we conclude that $f_{\exists\x\,\varphi^\prime(\x,\X,\z)} = f_{\exists\x\,\varphi(\x,\X,\z)}\dotminus 1.$
-%\end{enumerate}
-%		
-%\subsection*{Proof of Theorem \ref{dnf-pars}}
-%
-%Let $f$ be an arbitrary function in $\E{1}$ and let $f_{\textsc{\#DNF}}$ be the function that defines {\sc \#DNF}.
-We will now show a function $h:\str\to \textsc{Struct}[\L_{\textsc{DNF}}]$, where $\L_{\textsc{DNF}} = \{N,P\}$, such that for every $\A\in\str$, it holds that $f(\A) = f_{\textsc{\#DNF}}(h(\A))$, and which can be computed in polynomial time.
-For every DNF formula $\Phi$ we define $\A_{\Phi} \in \textsc{Struct}[\L_{\textsc{DNF}}]$ as stated in \ref{dnfdef}.
-Let $\psi$ be such that $f = f_{\psi(\X,\z)}$.
-We separate the proof in three cases:
-%\begin{enumerate}
-%	\item $\X = ()$.
-Let $\psi(\z) = \exists \x\,\varphi(\x,\z)$ where $\x = (x_1,\ldots,x_c)$ and $\z = (z_1,\ldots,z_d)$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a propositional formula $\Phi = h(\A)$.
-First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\,\varphi(\x,\e)$.
-Let $n$ be the number of such assignments to $\z$.
-Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$.
-The DNF formula we generate is
-%	\[
-%	\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
-%	\]
-%	which has exactly $n$ possible assignments. Finally, we return $h(\A) = \A_{\Phi}$.
-%	
-%	\item $\z = ()$.
-Let $\psi(\X) = \exists \x\,\varphi(\x,\X)$, where $\x = (x_1,\ldots,x_c)$ and $\X = (X_1,\ldots,X_r)$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a DNF-formula $\Phi = h(\A)$ as follows.
-First, we convert $\varphi(\x,\X)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X)$.
-Let each $\varphi_{i}(\x,\X)$ be of the form
-%	\[
-%	\varphi_{i}(\x,\X) = X_{\pi_i(1)}(\x^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\x^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\x^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\x^{\nu}_{i,n_i}) \wedge \alpha_i(\x),
-%	\]
-%	where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$) is the rearrangement of $\x$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x)$ is a FO formula.
-We also define the tuple $\a^{\pi}_{i,j}$ ($\a^{\nu}_{i,j}$) as $\a$ rearranged in the same way as $\x^{\pi}_{i,j}$ ($\x^{\nu}_{i,j}$), for each $\a\in A^c$.
-%	
-%	We generate a propositional variable $p^{\ell}_{\a}$ for each $\a\in A^{c_{\ell}}$, and for each $\ell\in\{1,\ldots,r\}$, which will indicate if the tuple $\a$ is in the assignment to $X_{\ell}$.
-Let $\p$ be the tuple of all such variables.
-%	
-%	Then, let $\Gamma_i = \{\a\in A^c\mid \A\models\alpha_i(\a)\}$.
-There is exactly $\vert A \vert^c$ tuples to check, and each of those checks can be done in polynomial time, so this set can also be generated in polynomial time.
-%	
-%	Finally, for each $\varphi_i(\x,\X)$ the DNF formula we generate is
-%	\[
-%	\Phi_i = \bigvee_{\a\in\Gamma_i} p^{\pi_i(1)}_{\a^{\pi}_{i,1}} \wedge \cdots \wedge p^{\pi_i(m_i)}_{\a^{\pi}_{i,m_i}} \wedge \neg p^{\nu_i(1)}_{\a^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\pi_i(n_i)}_{\a^{\nu}_{i,n_i}}.
-%	\]
-%	and $\Phi = \bigvee_{i = 1}^k \Phi_i \vee \bigvee_{p\in\p} (p\wedge\neg p)$. Finally, we return $h(\A) = \A_{\Phi}$.
-%	
-%	As an example, let $\psi^{*}(X_1,X_2) = \exists x \exists y\, (X_1(x) \wedge X_2(y,x) \wedge \neg X_1(y) \wedge x < y \wedge \neg\exists z\,(x<z\wedge z<y))$.
-We will compute $h(\B)$, where $\B = \langle \{1,2,3\}, < \rangle$, and we have that $1 < 2 < 3$.
-We generate the tuple of variables $\p = (p^1_1,p^1_2,p^1_3) \cup (p^2_{(i,j)})$ for $(i,j)\in \{1,2,3\}\times\{1,2,3\}$.
-Note that in this procedure we will only have one set $\Gamma_1 = \{(1,2),(2,3)\}$.
-And so, the DNF formula we generate is
-%	\begin{multline*}
-%	\Phi^{*} = (p^1_1\wedge p^2_{(2,1)} \wedge \neg p^1_2)\vee(p^1_2\wedge p^2_{(3,2)} \wedge \neg p^1_3)\, \vee \\ (p^1_1 \wedge \neg p^1_1) \vee(p^1_2 \wedge \neg p^1_2) \vee (p^1_3\wedge \neg p^1_3) \vee \bigvee_{i,j\in\{1,2,3\}}(p^2_{(i,j)}\wedge\neg p^2_{(i,j)})
-%	\end{multline*}
-%	
-%	Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\,\varphi(\x,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$.
-We can conclude that $f(\A) = f_{\textsc{DNF}}(\Phi)$.
-%	
-%	\item Let $\psi = \exists \x\,\varphi(\x,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$ we generate a DNF-formula $\Phi = h(\A)$ as follows.
-As in the previous case, we convert $\varphi(\x,\X,\z)$ into an equivalent DNF formula $\bigvee_{i = 1}^k \varphi_{i}(\x,\X,\z)$.
-Let each $\varphi_{i}(\x,\X,\z)$ be of the form
-%	\[
-%	\varphi_{i}(\x,\X,\z) = X_{\pi_i(1)}(\y^{\pi}_{i,1}) \wedge \cdots \wedge X_{\pi_i(m_i)}(\y^{\pi}_{i,m_i}) \wedge \neg X_{\nu_i(1)}(\y^{\nu}_{i,1}) \wedge \cdots \wedge \neg X_{\nu_i(n_i)}(\y^{\nu}_{i,n_i}) \wedge \alpha_i(\x,\z),
-%	\]
-%	where $\pi_i:\{1,\ldots,n_i\}\to\{1,\ldots,r\}$, $\nu_i:\{1,\ldots,m_i\}\to\{1,\ldots,r\}$, $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$) is the rearrangement of $(\x,\z)$ in the $j$-th positive (negative) second order variable mentioned in $\varphi_i$, and $\alpha(\x,\z)$ is a FO formula.
-We also define the tuple $(\a,\e)^{\pi}_{i,j}$ ($(\a,\e)^{\nu}_{i,j}$) as $\a$ rearranged in the same way as $\y^{\pi}_{i,j}$ ($\y^{\nu}_{i,j}$), for each $(\a,\e)\in A^{c+d}$.
-%	
-%	We generate a propositional variable $p^{\e}_{\ell,\b}$ for each $\e\in A^d$, for each $\ell\in\{1,\ldots,r\}$, and for each $\b\in A^{c_{\ell}}$.
-This variable will be true if there is an assignation to $\P$ such that $\b\in P_{\ell}$ and $\A\models\psi(\P,\e)$.
-Let $\p$ be the tuple of all such variables.
-We also generate a set $T = \{t_1,\ldots,t_{L}\}$ where $L = \lceil \log\vert A^d\vert \rceil$, and an arbitrary injective function $\beta:A^d \to \{0,1\}^L$ where $\beta(\e) = (b^{\e}_1,\ldots,b^{\e}_L)$ for each $\e\in A^d$.
-Also, let $\rho_{\ell}$ be such that $\rho_{\ell}(\x,\z) = \y_{\ell}$.
-%	
-%	For each $\e\in A^d$, we follow the following procedure.
-For each $\varphi_i(\x,\X)$, we consider the set $\Gamma^{\e}_i = \{\a\in A^c\mid \A\models\alpha(\a,\e)\}$, and so, the DNF formula we generate is
-%	\[
-%	\Phi^{\e}_i = \bigvee_{\a\in\Gamma_i} p^{\e}_{\pi_i(1),(\a,\e)^{\pi}_{i,1}} \wedge \cdots \wedge p^{\e}_{\pi_i(m_i),(\a,\e)^{\pi}_{i,m_i}} \wedge \neg p^{\e}_{\nu_i(1),(\a,\e)^{\nu}_{i,1}} \wedge \cdots \wedge \neg p^{\e}_{\pi_i(n_i),(\a,\e)^{\nu}_{i,n_i}} \wedge \gamma_{\e} \wedge \theta_{\e},
-%	\]
-%	where $\gamma_{\e}$ is defined as
-%	\[
-%	\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
-%	\]
-%	and $\theta_{\e}$ is defined as
-%	\[
-%	\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
-%	\]
-%	Finally,$\Phi = \bigvee_{\e\in A^d} \bigvee_{i = 1}^k \Phi_i^{\e} \vee \bigvee_{p\in\p} (p\wedge\neg p)$, and we return $f(\A) = \A_{\Phi}$.
-%	
-%	As an example, let $\psi^{*}(X_1,X_2,z) = \exists x\, (X_1(x) \wedge X_2(z) \wedge \neg X_1(z) \wedge x < z)$.
-We will compute $h(\B)$, where $\B = \langle \{1,2,3\}, < \rangle$, and we have that $1 < 2 < 3$.
-We generate the tuple of variables $\p = (p^e_{1,1},p^e_{1,2},p^e_{1,3},p^e_{2,1},p^e_{2,2},p^e_{2,3})$, for $e\in \{1,2,3\}$.
-Here, we consider $L = 2$, $T = \{t_1,t_2\}$, and $\beta$ defined as $\beta(1) = (0,0)$, $\beta(2) = (0,1)$ and $\beta(3) = (1,0)$.
-In this procedure we generate the sets $\Gamma^1_1 = \emptyset$, $\Gamma^2_1 = \{1\}$, $\Gamma^3_1 = \{1,2\}$.
-Now we generate $\gamma_e$ for $e\in\{1,2,3\}$,
-%	\[
-%	\gamma_e = \bigwedge_{d\in\{1,2,3\}\setminus\{e\}} p^d_{1,1} \wedge p^d_{1,2} \wedge p^d_{1,3} \wedge p^d_{2,1} \wedge p^d_{2,2} \wedge p^d_{2,3},
-%	\]
-%	and $\theta_e$ for $e\in\{1,2,3\}$,
-%	\begin{align*}
-%	\theta_1 &= \neg t_1 \wedge \neg t_2,\\
-%	\theta_2 &= \neg t_1 \wedge t_2,\\
-%	\theta_3 &= t_1 \wedge \neg t_2.		
-%	\end{align*}
-%	Then, we generate the formulas $\Phi^e_1$, for $e\in\{1,2,3\}$,
-%	\begin{align*}
-%	\Phi^1_1 &= \square\\
-%	\Phi^2_1 &= p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2,\\
-%	\Phi^3_1 &= (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3),
-%	\end{align*}
-%	and so, the DNF formula we generate is
-%	\begin{multline*}
-%	\Phi = (p^2_{1,1} \wedge p^2_{2,2} \wedge \neg p^2_{1,2} \wedge \gamma_2 \wedge \theta_2) \vee (p^3_{1,1} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee (p^3_{1,2} \wedge p^3_{2,3} \wedge \neg p^3_{1,3} \wedge \gamma_3 \wedge \theta_3) \vee \\ \bigvee_{e\in\{1,2,3\}}[(p^e_{1,1}\wedge\neg p^e_{1,1})\vee
-%	(p^e_{1,2}\wedge\neg p^e_{1,2})\vee
-%	(p^e_{1,3}\wedge\neg p^e_{1,3})\vee
-%	(p^e_{2,1}\wedge\neg p^e_{2,1})\vee
-%	(p^e_{2,2}\wedge\neg p^e_{2,2})\vee
-%	(p^e_{2,3}\wedge\neg p^e_{2,3})].
-%	\end{multline*}
-%	To demonstrate that our reduction is correct, we will show a function $g:\{\langle\P,\e\rangle \mid \A \models \exists\x\,\varphi(\x,\P,\e)\} \to \{\sigma \mid \sigma(\Phi) = 1\}$ and then prove that it is a bijection.
-Let $g$ be such that for each assignment $\langle \P,\e \rangle$, it holds that $g(\langle \P,\e \rangle) = \sigma$ where $\sigma$ is constructed as follows.
-For each $t_i\in T$, let $\sigma(t_i) = b^{\e}_i$.
-For each $p_{\ell,\b}^{\e}$, where $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e}) = 1$ if and only if $\b \in P_{\ell}$.
-For each $p_{\ell,\b}^{\e'}$, where $\e'\in A^d\setminus\{\e\}$,  $\ell\in\{1,\ldots,r\}$ and $\b\in A^{c_{\ell}}$, let $\sigma(p_{\ell,\b}^{\e'}) = 1$.
-%	
-%	Let $i$ be such that $\A\models \exists\x\,\varphi_i(\x,\P,\e)$.
-It is simple to see that from the way we constructed $\Phi_i^{\e}$, it holds that $\sigma(\Phi_i^{\e}) = 1$.
-Also, note that every different assignment $\langle \P',\e' \rangle$, it holds that $\sigma(\langle \P',\e' \rangle) \neq \sigma(\langle \P,\e \rangle)$.
-Therefore, $g$ is an injective function.
-Now, let $\sigma$ be an arbitrary assignment such that $\sigma(\Phi) = 1$.
-Let $i$ and $\e$ be such that $\sigma(\Phi^{\e}_i)$.
-Note that for every $\e'\in A^d\setminus\{\e\}$, it holds that $\sigma(\theta_{\e'}) = 0$, so there is exactly one possible $\e$ associated to $\sigma$.
-Consider the assignment $\P$ where $\b\in P_{\ell}$ if and only if $\sigma(p^{\e}_{\ell,\b}) = 1$, for each $\ell$ and $\b \in A^{c_{\ell}}$.
-Since there is at least one $\a\in\Gamma_i$ for which $\sigma$ satisfies its respective subformula, it holds that $\A\models\varphi_i(\a,\P,\e)$, which implies that $\A\models\exists\x\,\varphi(\x,\P,\e)$.
-From this we deduce that $g$ is surjective, and therefore a bijection.
-We conclude that $f(\A) = f_{\textsc{\#DNF}}(\A_{\Phi})$.
-%\end{enumerate}
-%	
-%\subsection*{Proof of Theorem \ref{nplusone-strict}}
-%
-%Suppose it is not properly contained for $n$.
-Let $\z = (z_1,\ldots,z_{n+1})$, let $\u = (u_1,\ldots,u_n)$ and let $\varphi(\z) = \bigwedge_{i = 1}^{n} (z_i < z_{i+1} \wedge \forall y(z_i\not<y \vee y\not<z_{i+1} )) \wedge \forall y(y \not< z_1)$.
-Note that $f_{\varphi(\z)}(\A) = 1$ if $\A$ has at least $n+1$ elements, and  $f_{\varphi(\z)}(\A) = 0$ otherwise, for each $\A\in\str$.
-Let $\forall\x\,\psi(\x,\X,\u)$ be such that $f_{\varphi(\z)} = f_{\forall\x\,\psi(\x,\X,\u)}$.
-Let $\A = \langle A, \S^{\A}, <^{\A} \rangle\in\str$ where $\vert A \vert = n+1$.
-Note that $f_{\forall\x\,\psi(\x,\X,\u)}(\A) = 1$.
-Let $(\P,\e)$ be such that $\A\models\forall\x\,\psi(\x,\P,\e)$.
-Let $a\in A$ be such that $a\not\in\e$.
-Consider the vocabulary $\L' = \L \cup \{\Q\}$, and let $\A' = \A \cup \langle \Q^{\A'} \rangle$ where $\Q^{\A'} = \P$.
-Note that $\A\models\varphi(\P,\e)$ if and only if $\A'\models\varphi(\e)$.
-Consider $\B'$ as the induced substructure of $\A'$ with domain $A\setminus\{a\}$, and let $\Q^{\B'}$ be the induced interpretation of $\Q$.
-Also let $\B$ be the induced substructure of $\A$ with domain $A\setminus\{a\}$.
-As universal formulae are preserved under substructure\footnote{Hodges, 1997.
-P.
-183}, we have that $\B'\models\forall\x\,\psi(\x,\e)$, which implies that $\B\models\forall\x\,\psi(\x,\Q^{\B'},\e)$.
-We conclude that $f_{\forall\x\,\psi(\x,\X,\u)}(\B) \geq 1$, which leads to a contradiction.
-%	
-%\subsection*{Proof of Theorem \ref{sigma2-pe}}
-%
-%Let $f = f_{\varphi(\X,\z)}$ such that $\varphi(\X,\z)$ is defined by the grammar $E_2$.
-First we notice that, as stated in \cite{SalujaST95}, for every $\L$-formula $\varphi$, $f_{\varphi}\in\shp$.\\
-%
-%We will now prove that there is a polynomial time algorithm that decides $L_{f}$.
-This is equivalent to decide, given $\A= \langle A, \S^\A, \leq^\A \rangle\in\str$, whether there exist assignments $\P,\z$ such that $\A\models\varphi(\P,\z)$.
-This is the same as $\A\models\exists\X\exists\x\,\varphi(\X,\x)$.
-Since $\varphi(\X,\x)$ is defined by the grammar $E_2$, there exists $\psi(\X,\x,\y,\u)$ such that $\varphi(\X,\x) = \exists\x\exists\y\forall\u\exists\v\,\psi(\X,\x,\y,\u,\v)$, where every variable in $\v$ appears on a negated second-order literal $\neg X_i$, with $i\in\N$.
-%
-%Given $\A\in\str$ we generate an equivalent $\L$-formula $\theta(\X,\x,\y)$ with a series of operations.
-First, we replace every instance of $\v$ by a disjunction of all $r$-tuples in $A^r$, where $r$ is the number of variables in $\v$.
-The result is still a Horn clause.
-Second, we replace every instance of $\u$ by a conjunction of all $s$-tulples in $A^s$, where $s$ is the number or variables in $\u$.
-The result is still a conjunction of Horn clauses.
-%
-%Now we notice that $\zeta = \exists\X\exists\x\exists\y\,\theta(\X,\x,\y)$ is an existential second-order FO-formula, for which $\A\models\zeta$ can be decided in polynomial time \cite{Gradel91}.
+\subsection*{Proof of Proposition \ref{prop:uhorn-pe}} %V.9
 
-\subsection*{Proof of Theorem \ref{sigma2hard}}
+\subsection*{Proof of Proposition \ref{prop:ehorn-pe}} %V.11
+
+\subsection*{Proof of Proposition \ref{prop:hsat-not-sigma2}} %V.12
+$\eqso(\uhorn) \subseteq \totp$:\vspace{1em}
+
+\subsection*{Proof of Theorem \ref{sigma2hard}} %V.13
 
@@
 $$ 
-If $\alpha = \beta_1 + \cdots + \beta_m$, let $g_{\beta_i}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\theta^i_{i,j}$ for each $\beta_i$ where each $\theta^i_{j,k}$ is a Horn clause.
+If $\alpha = \beta_1 + \cdots + \beta_m$, let $g_{\beta_i}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\theta^i_{j,k}$ for each $\beta_i$ where each $\theta^i_{j,k}$ is a Horn clause.
 Let $\Theta_i = g_{\beta_i}(\A)$.
@@
 This covers all possible cases for $\alpha$, and the entire procedure takes polynomial time.
-
-%Let $f$ be an arbitrary function in $\XE{2}$.
-We will now show a function $h:\str\to L(P)$ such that for every $\A\in\str$, it holds that $f(\A) = \shdhsat(h(\A))$, and which can be computed in polynomial time.
-Let $\psi$ be such that $f = f_{\psi}$.
-We separate the proof in three cases:\begin{enumerate}
-%	\item All free variables in $\psi$ are of first order.
-Let $\psi = \exists \x \forall \y\,\varphi(\x,\y,\z)$ where $\z = (z_1,\ldots,z_d)$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a propositional formula $\Phi = h(\A)$.
-First we notice that this case is trivial since we can calculate this function in polynomial time, by checking for each $\e\in A^d$ if $\A\models\exists\x\forall \y\,\varphi(\x,\y,\e)$.
-Let $n$ be the number of such assignments to $\z$.
-Let $\p$ be an $n$-tuple of propositional variables where $\p = (p_1,\ldots,p_n)$.
-The disjunction of Horn formulas we return is
-%	\[
-%	\Phi = \bigvee_{i = 1}^n \neg p_1 \wedge \cdots \wedge \neg p_{i-1} \wedge p_i \wedge \neg p_{i+1} \wedge \cdots \wedge \neg p_n
-%	\]
-%	which has exactly $n$ possible assignments. Note that each $p_i$ and $\neg p_i$ is a Horn clause itself.
-%	
-%	\item There are only second-order free variables in $\psi$.
-Let $\psi = \exists \x\forall \y\,\varphi(\x,\y,\X)$, where $\x = (x_1,\ldots,x_c)$, $\y = (y_1,\ldots,y_b)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$, we generate a Disj-Horn formula $\Phi = h(\A)$ as follows:
-%	
-%	First, Let $\varphi(\x,\y,\X) = \bigwedge_{i = 1}^k \varphi_{i}(\x,\y,\X)$ where, without loss of generality, each $\varphi_{i}(\x,\y,\X)$ is of the form
-%	\[
-%	\varphi_{i}(\x,\y,\X) = X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n}) \vee \alpha_i(\x,\y),
-%	\]
-%	where $\u_{s}$ and every $\u_{t_j}$ have the corresponding number of variables in $(\x,\y)$, and $\alpha(\x,\y)$ is a FO-formula.
-For the rest of the proof, $X_s(\u_s)$ can be absent and it does not affect the proof.
-Note that this formula is equivalent to
-%	\[
-%	\neg\alpha_i(\x,\y) \to (X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n})),
-%	\]
-%	
-%	Then, for each $\a\in A^c$ let $\Gamma_i^{\a} = \{\b\in A^{b}\mid \A\models\neg\alpha(\a,\b)\}$.
-Note that this set can be generated in polynomial time.
-For each $\ell\in\{1,\ldots,r\}$, we generate a propositional variable $p^{\ell}_{\d}$ for each $\d\in A^{c_{\ell}}$.
-Let $\p$ be the tuple of all such variables.
-Also, let $\rho_{\ell}(\x,\y) = \u_{\ell}$, and $\rho_{\ell}(\d)$ be the tuple that results of reordering $\d$ using the reordering of the variables of $(\x,\y)$ in $\u_{\ell}$.
-%	
-%	For each $\a\in A^c$ we generate a Horn formula
-%	\[
-%	\Phi_{\a}^i = \bigwedge_{\b\in\Gamma_i} p^{s}_{\rho_s(\a,\b)} \vee \neg p^{t_1}_{\rho_{t_1}(\a,\b)} \vee \cdots \vee \neg p^{t_n}_{\rho_{t_n}(\a,\b)}.
-%	\]
-%	and finally we return $\Phi = \bigvee_{\a\in A^c} \bigwedge_{i = 1}^k \Phi_{\a}^i$.
-%	
-%	Note that for each assignment $\P$ to $\X$ such that $\A\models\exists\x\forall\y\,\varphi(\x,\y,\P)$ there is a corresponding assignment $\sigma$ to $\p$ where $\sigma(p^{\ell}_{\a}) = 1$ if and only if $\a\in P_{\ell}$.
-We can conclude that $f(\A) = f_{\textsc{Disj-Horn-Sat}}(\Phi)$.
-%	
-%	\item Let $\psi = \exists \x\forall\y\,\varphi(\x,\y,\X,\z)$, where $\x = (x_1,\ldots,x_c)$, $\y = (y_1,\ldots,y_b)$, $\z = (z_1,\ldots,z_d)$, $\X = (X_1,\ldots,X_r)$ and each $X_{\ell}$ has arity $c_{\ell}$.
-Given a structure $\A = \langle A, \S^{\A}, \leq^{\A} \rangle\in\str$ we generate a Disj-Horn formula $\Phi = h(\A)$ as follows:
-%	
-%	As in the previous case, let $\varphi(\x,\y,\X,\z) = \bigwedge_{i = 1}^k$ where $\varphi(\x,\y,\X,\z)$ is of the form:
-%	\[
-%	\varphi_{i}(\x,\y,\X,\z) = X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n}) \vee \alpha_i(\x,\y,\z),
-%	\]
-%	where $\u_{s}$ and every $\u_{t_j}$ have the corresponding number of variables in $(\x,\y,\z)$, and $\alpha(\x,\y,\z)$ is a FO-formula.
-For the rest of the proof, $X_s(\u_s)$ can be absent and it does not affect the proof.
-Note that this formula is equivalent to
-%	\[
-%	\neg\alpha_i(\x,\y,\z) \to (X_s(\u_{s}) \vee \neg X_{t_1}(\u_{t_1}) \vee \cdots \vee \neg X_{t_n}(\u_{t_n})).
-%	\]
-%	
-%	For each $\e\in A^d$, we do the following.
-For each $\a\in A^c$ let $\Gamma^{\e}_{i,\a} = \{\b \in A^b \mid \A\models \neg\alpha_i(\a,\b,\e) \}$.
-Note that this set can be generated in polynomial time.
-For each $\ell\in\{1,\ldots,r\}$ we generate a propositional variable $p^{\e}_{\ell,\d}$ for each $\d\in A^{c_{\ell}}$.
-Let $\p$ be the tuple of all such variables.
-We also generate a set $T = \{t_1,\ldots,t_{L}\}$ where $L = \lceil \log\vert A^d\vert \rceil$, and an arbitrary injective function $\beta:A^d \to \{0,1\}^L$ where $\beta(\e) = (b^{\e}_1,\ldots,b^{\e}_L)$ for each $\e\in A^d$.
-Also, let $\rho_{\ell}(\x,\y,\z) = \u_{\ell}$, and $\rho_{\ell}(\d)$ be the tuple that results of reordering $\d$ using the order of the variables in $(\x,\y,\z)$ in $\u_{\ell}$.
-And so, the Horn formula we generate is
-%	\[
-%	\Phi^{\e}_{i,\a} = \bigwedge_{\a\in\Gamma^{\e}_i} p^{\e}_{s,\rho_{s}(\a,\b,\e)} \vee p^{\e}_{t_1,\rho_{t_1}(\a,\b,\e)} \vee \cdots \vee p^{\e}_{t_n,\rho_{t_n}(\a,\b,\e)}.
-%	\]
-%	And finally we return $\Phi = \bigvee_{\e\in A^d}\bigvee_{\a\in A^c} (\bigwedge_{i=1}^k \Phi_{\e}^{i,\a} \wedge \gamma_{\e} \wedge \theta_{\e} )$, where $\gamma_{\e}$ is defined as
-%	\[
-%	\gamma_{\e} = \bigwedge_{\e'\in A^d\setminus\{\e\}} \bigwedge_{\ell = 1}^r \bigwedge_{\b\,\in A^{c_{\ell}}} p_{\ell,\b}^{\e'},
-%	\]
-%	and $\theta_{\e}$ is defined as
-%	\[
-%	\theta_{\e} = \bigwedge_{i\in[1,L]:\,b^{\e}_i = 0} \neg t_i\  \wedge \bigwedge_{i\in[1,L]:\,b^{\e}_i = 1} t_i.
-%	\]
-%	
-%\end{enumerate}
-
-\subsection*{Proof of Theorem \ref{so-rec}}
-
-\subsection*{Proof of Theorem \ref{tqfo-fo-cap}}
-
-For the second condition, let $f \in \shl$.
-We will address the case where $\R$ contains only one binary predicate $E$, and the rest of the cases can be deduced from this.
-Let $M$ be a non-deterministic logspace machine such that $f(\A) = \acc_M(\A)$ for each $\A \in \ostr[R]$.
-Suppose ${\cal Q} = \{q_1,\ldots,q_{\ell}\}$ is the set of states of $M$, where $q_1$ is the initial state, and $q_{\ell}$ is only final state of $M$.
-Let $n = \vert A \vert$ and let $w = \enc(\A) \in \{0,1\}^{n^2}$.
-We assume that $M$ with input $w$ uses space $s_M(w) < c\cdot\log(n)$ and furthermore, $s_M(w) < n-2$.
-We notate $M(w)$ as the graph of configurations of $M$ running on input $w$.
-
-We represent configurations with a tuple of fixed size.
-The formula $\varphi(\bar{x},\bar{y})$ describes a procedure that given a configuration generates a possible next configuration.
-The formula $\varphi_I(\bar{x})$ describes that $\bar{x}$ is the initial configuration of $M(w)$.
-The formula $\varphi_F(\bar{x})$ describes that $\bar{x}$ is an accepting (final) configuration of $M(w)$.
-The formula we construct is:
-$$
-\alpha = \sa{\bar{x}}\sa{\bar{y}}([\pth \varphi(\bar{u},\bar{v})](\bar{x},\bar{y})\cdot \varphi_I(\bar{x})\cdot\varphi_F(\bar{y})).
-$$
-
-To illustrate our idea, we will show a simplified example.
-Consider a machine $M$ that works in exactly $\log_2(n)$ space and only allows 0 or 1 in the working tape.
-Consider an input $\A$ of size 16 (that is, $A = \{0,\ldots,9,A,\ldots,F\}$).
-Let some configuration $s$ have 0011 in the working tape, the head in the input tape is in position 26, and the head in the input tape is in position 2 (we consider 0-indexed positions).
-Also, $Q = \{q_1,\ldots,q_5\}$ and the current state is $q_3$.
-
-As a first approach, we will use a 9-tuple $\bar{a} = (a_1,\ldots,a_9)$ to represent $s$.
-That is, $(a_1,a_2) = (1,A)$ represent the position of the head in the input tape, $a_3 = 2$ represents the position of the head in the working tape, $a_4 = C$ (1100b in base 16) represents the content of the working tape, and $(a_5,\ldots,a_9) = (0,0,1,0,0)$ represents the current state.
-Then $\bar{a} = (1,A,2,C,0,0,1,0,0)$ will represent $s$.
-
-The problem that arises from this representation, is that to describe a transition in $M$ we need to read an arbitrary character in the working tape, and therefore we have to obtain the $a_3$-th bit in $a_4$.
-Furthermore, to represent the following configuration, we need compute $a_4$ with the $a_3$-th bit flipped.
-This is generally not possible to describe with an $\fo$ formula.
-Consider the following procedure that receives $x = a_4$ and $i = a_3$.
-
-\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 1 replace it by 0 and return the result}
-	\label{switch1to0}
-	\begin{algorithmic}
-		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
-		\While{$j > 0$}
-		\State $v \gets 0$
-		\While{$u > 1$}
-		\State $u \gets u-2,\; v \gets v+1$
-		\EndWhile
-		\State $u\gets v,\; j \gets j-1$
-		\EndWhile
-		\While{$u > 1$}
-		\State $u \gets u-2$
-		\EndWhile
-		\State $\textbf{assert } u = 1$ \Comment{If $u \neq 1$ simply stop}	
-		\State $y \gets 1$ \Comment{Compute $2^i$ and store it in $y$}
-		\While{$i > 0$}
-		\State $z \gets 0$
-		\While{$y > 0$}
-		\State $z \gets z+2,\; y \gets y-1$
-		\EndWhile
-		\State $i \gets i-1,\; y \gets z$
-		\EndWhile
-		\While{$y > 0$} \Comment{Substract $y$ from $x$}
-		\State $x \gets x-1,\; y \gets y-1$
-		\EndWhile
-		\State \Return $x$.
-	\end{algorithmic}
-\end{algorithm}	
-Each of the instructions can be expressed with $\fo$, so our strategy is to use the $\pth$ operator to simulate the algorithm and then we can describe a transition using the processed value of $a_4$.
-Note that we need to describe three more procedures to simulate the transitions for $0\to 0$, $0 \to 1$ and $1\to 1$.
-
-We will now describe how to simulate both the procedure and the transition.
-A procedure tuple $\bar{p} = (a_1,\ldots,a_{3+c+\ell},b_1,b_2,c_1,c_2,c_3,d_1,\ldots,d_{5c+2})$ represents the current configuration of $M(w)$ in $a_1,\ldots,a_{2+c+\ell}$, the values that will be read and written in the working tape in $b_1,b_2$, the instruction pointer in $c_1,c_2,c_3$ and the values stored in memory in $d_1,\ldots,d_{10c+2}$.
-In detail:
-\begin{enumerate}
-	\item $a_1,a_2$ and $a_3$ represent the position of the head in the input tape and the working tape, respectively, $a_4,\ldots,a_{3+c}$ represent the content of the working tape and $a_{4+c},\ldots,a_{3+c+\ell}$ represent the current state in the current configuration that is being processed.
-	\item $b_1$ and $b_2$ are equal to the value that is being read in the working tape and the value that will be written in the working tape respectively.
-	\item $c_1,c_2,c_2$ represent the instruction pointer in the procedure. Only 8 different instructions are needed in the simulation.
-	\item Each value in memory of $x,y,z,u,v$ need $c$ elements to represent them and $i,j$ need only one. We map $(d_1\ldots,d_{c}) \to x$, $(d_{c+1}\ldots,d_{2c}) \to y$,
-	$(d_{2c+1}\ldots,d_{3c}) \to z$, $(d_{3c+1}\ldots,d_{4c}) \to u$,
-	$(d_{4c+1}\ldots,d_{5c}) \to v$, $d_{5c+1} \to i$ and $d_{5c+2}\to j$.
-\end{enumerate}
-For each transition $\delta \in \Delta \subseteq Q \times \{0,1\} \times \{0,1\} \times Q \times \{-1,=,+1\} \times \{0,1\} \times \{-1,=,+1\}$ we define a formula $\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$, where $\bar{x} = (x_1,\ldots,x_{3+c+\ell})$, $\bar{s} = (s_1,s_2)$, $\bar{w} = (w_1,w_2,w_3)$, $\bar{u} = (u_1,\ldots,u_{5c+2})$, $\bar{y} = (y_1,\ldots,y_{3+c+\ell})$, $\bar{t} = (t_1,t_2)$, $\bar{z} = (z_1,z_2,z_3)$ and $\bar{v} = (v_1,\ldots,v_{5c+2})$.
-The tuples $\bar{x}$ and $\bar{y}$ represent the current and next configuration of $M$ respectively, $\bar{s}$ and $\bar{y}$ indicate which algorithm is being followed, $\bar{w}$ and $\bar{z}$ represent the current and next instruction of the algorithm, $\bar{u}$ and $\bar{v}$ represent the current and next values in memory.
-We will describe the formula part by part.
-Suppose $\delta = (q_i,a,1,q_j,op_1,0,op_2)$, so we have to simulate Algorithm \ref{switch1to0}.
-
-First we define some auxiliary formulas:
-\begin{align*}
-\varphi_0(x) &= \neg\exists y(y < x), \\
-\varphi_{+1}(x,y) &= x < y \wedge \neg\exists z(x < z \wedge z < y), \\
-\varphi_{-1}(x,y) &= \varphi_{+1}(y,x), \\
-\varphi_1(x) &= \exists y(\varphi_0(y) \wedge \varphi_{+1}(y,x)), \\
-\text{for $i,j \in\{0,1\}$, }	\varphi_{i,j}(x,y) &= \varphi_i(x) \wedge \varphi_j(y), \\
-\text{for each $k \leq 7$, }\varphi^b_k(x_1,x_2,x_3) &= \varphi_{a_1}(x_1) \wedge \varphi_{a_2}(x_2) \wedge \varphi_{a_3}(x_3)\text{, where $a_1a_2a_3$ is the value of $k$ in binary}, \\
-\mu_{<}(x_1,\ldots,x_c,y_1,\ldots,y_c) &= \bigvee_{i = 1}^c ((\bigwedge_{j = 1}^{i-1} x_j = y_j) \wedge x_i < y_i), \\
-\mu_{+1}(\bar{x},\bar{y}) &= \mu_{<}(\bar{x},\bar{y}) \wedge \neg\exists \bar{z}(\mu_{<}(\bar{x},\bar{z}) \wedge \mu_{<}(\bar{z},\bar{y})), \\
-\mu_{+2}(\bar{x},\bar{y}) &= \exists\bar{z}(\mu_{+1}(\bar{x},\bar{z}) \wedge \mu_{+1}(\bar{z},\bar{y})),\\
-\mu_{-1}(\bar{x},\bar{y}) &= \varphi_{+1}(\bar{y},\bar{x}),\\
-\mu_{-2}(\bar{x},\bar{y}) &= \varphi_{+2}(\bar{y},\bar{x}),\\
-\mu_{0}(\bar{x}) &= \neg\exists\bar{y}(\mu_{<}(\bar{y},\bar{x})),\\
-\mu_{1}(\bar{x}) &= \exists\bar{y}(\mu_{0}(\bar{y})\wedge\mu_{+1}(\bar{y},\bar{x})),\\
-\varphi^2_{<}(x_1,x_2,y_1,y_2) &= x_1 < y_1 \vee (x_1 = y_1 \wedge x_2 < y_2),\\
-\varphi^2_{+1}(x_1,x_2,y_1,y_2), \varphi^2_{-1}(x_1,x_2,y_1,y_2) &\text{ defined analogously to $\mu_{+1}$ and $\mu_{-1}$},\\
-\text{for each } q_i\in Q, \varphi^q_i(x_1,\ldots,x_{\ell}) &= \bigwedge_{\substack{j = 1 \\ j \neq i}}^{\ell} \varphi_0(x_j) \wedge \varphi_1(x_i), \\
-\varphi^E_0(x_1,x_2) &= \neg E(x_1,x_2),\\		\varphi^E_1(x_1,x_2) &= E(x_1,x_2),\\
-\end{align*}
-
-We start from instruction 0, which means that the procedure has not started yet and every value in the tuple is 0 except for the configuration values.
-It also initializes all the values in the tuple to 0 except for $x,u,i,j$.
-\begin{multline*}
-\varphi^{0,1}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{0,0}(s_1,s_2)\wedge\varphi^b_0(\bar{w}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_1(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c v_i = x_{3+i} \wedge \bigwedge_{i = c+1}^{2c} \varphi_0(v_i) \wedge \bigwedge_{i = 2c+1}^{3c} \varphi_0(v_i) \wedge \bigwedge_{i = 1}^c v_{3c+i} = x_{3+i} \wedge \bigwedge_{i = 4c+1}^{5c} \varphi_0(v_i) \wedge v_{5c+1} = x_3 \wedge v_{5c+2} = x_3.
-\end{multline*}
-Instruction 1 which checks whether the value of $j$ ($d_{5c+2}$ in the tuple) is more than 0 or not, and then proceeds to instruction 2 or 3 on each case.
-\begin{multline*}
-\varphi^{1,2}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_1(\bar{w}) \wedge \neg \varphi_0(u_{5c+2}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_2(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i, \\
-\varphi^{1,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_1(\bar{w}) \wedge \varphi_0(u_{5c+2}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_3(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
-\end{multline*}
-Instruction 2 checks the value of $u$ ($d_{3c+1},\ldots,d_{4c}$ in the tuple).
-If it is $> 1$ then it substracts 2 from $u$ and adds 1 to $v$ ($d_{4c+1},\ldots,d_{5c}$ in the tuple), then repeats instruction 2.
-If it is equal to 0 or 1, then moves the value of $v$ to $u$, substracts 1 from $j$ and goes back to instruction 1.
-\begin{multline*}
-\varphi^{2,2}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_2(\bar{w}) \wedge \neg \mu_0(u_{3c+1},\ldots,u_{4c}) \wedge \neg \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_2(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
-\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-\mu_{-2}(u_{3c+1},\ldots,u_{4c},v_{3c+1},\ldots,v_{4c})\, \wedge \\
-\mu_{+1}(u_{4c+1},\ldots,u_{5c},v_{4c+1},\ldots,v_{5c}) \wedge u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}. \\
-\varphi^{2,1}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_2(\bar{w}) \wedge ( \mu_0(u_{3c+1},\ldots,u_{4c}) \vee \mu_1(u_{3c+1},\ldots,u_{4c})) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_1(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
-\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-\bigwedge_{i = 3c+1}^{4c} u_{c+i} = v_i \wedge
-\mu_0(v_{4c+1},\ldots,v_{5c}) \wedge
-u_{5c+1} = v_{5c+1} \wedge \varphi_{-1}(u_{5c+2},v_{5c+2}).	
-\end{multline*}
-Instruction 3 calculates the value of $u \mod 2$, that is, it repeats instruction 3 until the value of $u$ is equal to 0 or 1.
-On each iteration, it substracts 2 from $u$.
-Moreover, if the value of $u$ at the end of the iterations is not 1 then there is no step defined.
-\begin{multline*}
-\varphi^{3,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_3(\bar{w}) \wedge \neg \mu_0(u_{3c+1},\ldots,u_{4c}) \wedge \neg \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_3(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\bigwedge_{i = c+1}^{2c} u_i = v_i \wedge
-\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-\mu_{-2}(u_{3c+1},\ldots,u_{4c},v_{3c+1},\ldots,v_{4c}) \wedge
-\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
-u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}, \\
-\varphi^{3,3}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 
-\varphi_{1,0}(s_1,s_2) \wedge \varphi^b_3(\bar{w}) \wedge \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_4(\bar{z})\, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\mu_1(v_{c+1},\ldots,v_{2c}) \wedge
-\bigwedge_{i = 2c+1}^{3c} u_i = v_i \wedge
-\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
-\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
-u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}
-\end{multline*}
-Instruction 4 checks the value of $i$ ($d_{5c+1}$ in the tuple.) If it is not 0 then goes to instruction 5 and if is 0 then goes to instruction 6.
-Moreover it initializes the value of $z$ ($d_{2c+1},\ldots,d_{3c}$ in the tuple) to 0 (which was 0 all along.)
-\begin{multline*}
-\varphi^{4,5}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_4(\bar{w}) \wedge \neg \varphi_0(u_{5c+1}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_5(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i, \\
-\varphi^{4,6}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_4(\bar{w}) \wedge \varphi_0(u_{5c+1}) \wedge \varphi_{1,0}(t_1,t_2) \wedge 	\varphi^b_6(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
-\end{multline*}
-Instruction 5 checks the value of $y$ ($d_{c+1},\ldots,d_{2c}$ in the tuple.) If it is more than 0 then it adds 2 to $z$ and substracts 1 from $y$, then repeats instruction 2.
-If it is not, then copies the value of $z$ to $y$ and subtracts 1 from $i$ and returns to instruction 4.
-\begin{multline*}
-\varphi^{5,5}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_5(\bar{w}) \wedge \neg \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_5(\bar{z}) \, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\mu_{-1}(u_{c+1},\ldots,u_{2c},v_{c+1},\ldots,v_{2c}) \wedge
-\mu_{+2}(u_{2c+1},\ldots,u_{3c},v_{2c+1},\ldots,v_{3c})\, \wedge \\
-\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
-\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
-u_{5c+1} = v_{5c+1} \wedge u_{5c+2} = v_{5c+2}, \\
-\varphi^{5,4}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_5(\bar{w}) \wedge \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_4(\bar{z}) \, \wedge \\
-\bigwedge_{i = 1}^c u_i = v_i \wedge
-\bigwedge_{i = c+1}^{2c} u_{c+i} = v_i \wedge
-\mu_0(v_{2c+1},\ldots,v_{3c}) \wedge
-\bigwedge_{i = 3c+1}^{4c} u_i = v_i \wedge
-\bigwedge_{i = 4c+1}^{5c} u_i = v_i \wedge
-\varphi_{-1}(u_{5c+1},v_{5c+1}) \wedge u_{5c+2} = v_{5c+2}
-\end{multline*}
-Instruction 6 checks the value of $y$. If it is more than 0, then subtracts 1 from $x$ and $y$ and repeats instruction 6. If it is not, then goes to instruction 7.
-\begin{multline*}
-\varphi^{6,6}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_6(\bar{w}) \wedge \neg \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_6(\bar{z}) \, \wedge \\
-\mu_{-1}(u_1,\ldots,u_c,v_i,\ldots,v_c) \wedge \mu_{-1}(u_{c+1},\ldots,u_{2c},v_{c+1},\ldots,v_{2c}) \wedge \bigwedge_{i = 3c+1}^{5c+2} u_i = v_i]\,\vee \\
-\varphi^{6,7}_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_6(\bar{w}) \wedge \mu_0(u_{c+1},\ldots,u_{2c}) \wedge \varphi_{1,0}(t_1,t_2) \wedge \varphi^b_7(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i.
-\end{multline*}
-Instruction 7 stores the value of $x$ after the corresponding bit has been switched.
-Then we can define $\varphi_{\delta}$ which also simulates the actual transition.
-If $u$ equals 1, then copy what is stored in $x$ to $a_4,\ldots,a_{3+c}$, go from state $q_i$ to state $q_j$, and move the heads to their corresponding positions.
-\begin{multline*}
-\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = 	[\bigwedge_{i = 1}^{3+c+\ell} x_i = y_i \wedge (\varphi^{0,1}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{1,2}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})\, \vee \\ \varphi^{1,3}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{2,2}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{2,1}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{3,3}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \\ \varphi^{3,4}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{4,5}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{4,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{5,5}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \\ \varphi^{5,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{6,6}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) \vee \varphi^{6,7}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}))] \, \vee \\
-[\varphi^E_a(x_1,x_2) \wedge \varphi_{1,0}(s_1,s_2) \wedge \varphi^b_7(\bar{w}) \wedge \mu_1(u_{3c+1},\ldots,u_{4c}) \wedge \varphi_{0,0}(t_1,t_2) \wedge \varphi^b_0(\bar{z}) \wedge \bigwedge_{i = 1}^{5c+2} u_i = v_i \,\wedge \\
-\varphi^2_{op_1}(x_1,x_2,y_1,y_2) \wedge \varphi_{op_2}(x_3,y_3) \wedge \bigwedge_{i = 1}^c u_i = x_{3+i} \wedge \varphi^q_i(x_{4+c},\ldots,x_{3+c+\ell}) \wedge \varphi^q_j(y_{4+c},\ldots,y_{3+c+\ell})].
-\end{multline*}
-
-\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 0 return $x$} \label{switch0to0}
-	\begin{algorithmic}
-		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
-		\While{$j > 0$}
-		\State $v \gets 0$
-		\While{$u > 1$}
-		\State $u \gets u-2,\; v \gets v+1$
-		\EndWhile
-		\State $u\gets v,\; j \gets j-1$
-		\EndWhile
-		\While{$u > 1$}
-		\State $u \gets u-2$
-		\EndWhile
-		\State $\textbf{assert } u = 0$ \Comment{If $u \neq 0$ simply stop}	
-		\State \Return $x$.
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 0 replace it by 1 and return the result}
-	\label{switch0to1}
-	\begin{algorithmic}
-		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
-		\While{$j > 0$}
-		\State $v \gets 0$
-		\While{$u > 1$}
-		\State $u \gets u-2,\; v \gets v+1$
-		\EndWhile
-		\State $u\gets v,\; j \gets j-1$
-		\EndWhile
-		\While{$u > 1$}
-		\State $u \gets u-2$
-		\EndWhile
-		\State $\textbf{assert } u = 0$ \Comment{If $u \neq 0$ simply stop}	
-		\State $y \gets 1$ \Comment{Compute $2^i$ and store it in $y$}
-		\While{$i > 0$}
-		\State $z \gets 0$
-		\While{$y > 0$}
-		\State $z \gets z+2,\; y \gets y-1$
-		\EndWhile
-		\State $i \gets i-1,\; y \gets z$
-		\EndWhile
-		\While{$y > 0$} \Comment{Add $y$ to $x$}
-		\State $x \gets x+1,\; y \gets y-1$
-		\EndWhile
-		\State \Return $x$.
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
-	\caption{If the $i$-th bit in $x$ is 1 return $x$}
-	\label{switch1to1}
-	\begin{algorithmic}
-		\State $u \gets x,\; j \gets i$ \Comment{Get the $i$-th bit on $x$ and store it in $u$}
-		\While{$j > 0$}
-		\State $v \gets 0$
-		\While{$u > 1$}
-		\State $u \gets u-2,\; v \gets v+1$
-		\EndWhile
-		\State $u\gets v,\; j \gets j-1$
-		\EndWhile
-		\While{$u > 1$}
-		\State $u \gets u-2$
-		\EndWhile
-		\State $\textbf{assert } u = 1$ \Comment{If $u \neq 0$ simply stop}	
-		\State \Return $x$.
-	\end{algorithmic}
-\end{algorithm}
-Note that we also need to specify that the program we are following is Algorithm \ref{switch1to0} so we store $1,0$ in $b_1,b_2$ all along the procedure.
-We describe the three other algorithms that compute the switches from $0\to 0$, $0\to 1$ and $1\to 1$ (Algorithms \ref{switch0to0}, \ref{switch0to1} and \ref{switch1to1}.)
-For the other three cases, where $\delta = (q_i,a,0,q_j,op_1,0,op_2)$, $\delta = (q_i,a,0,q_j,op_1,1,op_2)$ and $\delta = (q_i,a,1,q_j,op_1,1,op_2)$, $\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$ is defined analogously.
-Then, $\varphi(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$ is defined as:
-$$
-\varphi(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}) = \bigvee_{\delta \in \Delta} \varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v}).
-$$
-Finally we define $\varphi_I$ and $\varphi_F$:
-\begin{align*}
-\varphi_I(\bar{x},\bar{s},\bar{w},\bar{u}) &= \varphi_0(x_1) \wedge \varphi_0(x_2) \wedge \varphi_0(x_3) \wedge \bigwedge_{i = 4}^{3+c}\varphi_0(x_i) \wedge \varphi_1(x_{4+c}) \wedge \bigwedge_{i = 5+c}^{3+c+\ell} \varphi_0(x_i)\wedge \varphi_{0,0}(s_1,s_2)\ \wedge \varphi^b_0(\bar{w}) \wedge \bigwedge_{i = 1}^{5c+2}\varphi_0(u_i).
-\\
-\varphi_F(\bar{x},\bar{s},\bar{w},\bar{u}) &= \varphi^q_{\ell}(x_{4+c},\ldots,x_{3+c+\ell}) \wedge \varphi_{0,0}(s_1,s_2)\ \wedge \varphi^b_0(\bar{w}) \wedge \bigwedge_{i = 1}^{5c+2}\varphi_0(u_i),
-\end{align*}
-and then $\sem{\alpha}(\A) = \sem{\sa{\bar{x}}\sa{\bar{y}}([\pth \varphi(\bar{u},\bar{v})](\bar{x},\bar{y})\cdot \varphi_I(\bar{x})\cdot\varphi_F(\bar{y}))}(\A) = \acc_M(\A)$.
-
-\subsection*{Proof of Theorem \ref{tqso-fo-cap}}
-
-\subsection*{Proof of Theorem \ref{rqfo-fo-cap}}

### 2016-12-30 LICS/saluja.tex
mas cambios

@@
 %, defining what we call 
-Specifically, the $\sfo$-hierarchy consists of the 
+Specifically, the \emph{$\sfo$-hierarchy} consists of the 
 %they define 
@@
 %for every $i \geq 0$. 
-By definition we have that $\U{0} = \E{0}$. Moreover, it is shown in~\cite{SalujaST95} that:
+By definition ,we have that $\U{0} = \E{0}$. Moreover, it is shown in~\cite{SalujaST95} that:
 % this function classes defined a finite hierarchy of the form:
@@
 %the classes 
-$\eqso(\loge{i})$ and $\eqso(\logu{i})$ for every $i \geq 0$, which form the $\eqso$-hierarchy.
+$\eqso(\loge{i})$ and $\eqso(\logu{i})$ for every $i \geq 0$, which form the \emph{$\eqso$-hierarchy}.
 %, where the boolean logic is restricted to $\loge{i}$ and $\logu{i}$, respectively. 
 %We denote each class by $\QE{i}$ and $\QU{i}$ for short. 
+Clearly, we have that $\E{i} \subseteq \QE{i}$ and $\U{i} \subseteq \QU{i}$.
+Indeed, each formula $\varphi(\bar{X}, \bar{x})$ in $\E{i}$ is equivalent to the formula $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ in $\QE{i}$, and likewise for $\U{i}$ and $\QU{i}$.
+But what is the exact relationship between these two hierarchies?
+%Then it is left to know whether these containments are strict.
+%between the sharp and quantitative classes is strict or not. 
+%For this, 
+To answer this question, we start by introducing two normal forms for $\eqso(\LL)$ that helps us to characterize the expressive power of this quantitative logic.
+%study first whether a formula $\alpha$ in $\eqso(\LL)$ can be transformed into a 
+%\emph{prenex 
+%normal form where sum quantifiers are restricted to be at the beginning of the formula. 
+%Formally, we say that 
+A formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-prenex normal form ($\LL$-PNF)} 
+%(or just prenex normal form) 
+if $\alpha$ is of the form
+%\[
+%\alpha := 
+$\sa{\bar{X}} \sa{\bar{x}} \varphi(\bar{X}, \bar{x})$,
+%\]
+where $\bar{X}$ and $\bar{x}$ are sequences of zero or more second-order and first-order variables, respectively, and $\varphi(\bar{X}, \bar{x})$ is a formula in $\LL$. Notice that 
+%a sentence in $\LL$ is in $\LL$-PNF, while 
+a formula $\varphi(\bar{X}, \bar{x})$ in $\#\LL$ is equivalent to the formula $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ in $\LL$-PNF. 
+%In particular, 
+%%note that a 
+%each Boolean formula is in 
+%%$\Sigma$-
+%prenex normal form.
+Moreover, a formula $\alpha$ is in \emph{$\LL$-sum normal form ($\LL$-SNF)} if $\alpha$ is of the form $c + \Sigma_{i=1}^n \alpha_i$, where $c$ is a non-negative constant and each $\alpha_i$ is in 
+%$\Sigma$-
+$\LL$-PNF.
+%prenex normal form.
+%The following results shows that 
+%%Next we show that 
+%each formula in $\eqso(\LL)$ can be converted in sum normal form.
+% but not always in prenex normal form.
+\begin{theorem}\label{theo-pnf-snf}
+Every formula in $\eqso(\LL)$ can be rewritten in $\LL$-SNF.
+%sum normal form.
+%such that there does not exist a formula in $\Sigma$-prenex normal form in $\QE{1}$ equivalent to $\alpha$.
+\end{theorem}
+%\martin{Creo que el resultado de que existe una formula en $\QE{1}$ que no se puede pasar a prenex está mas relacionado con el teorema siguiente que con este.}
+%By the previous result, we know that there exists logics $\LL$ where no $\Sigma$-prenex normal form exists. 
+%Then
+Therefore, to unveil the relationship between the $\sfo$-hierarchy and the $\eqso$-hierarchy, we need to understand the boundary between PNF and SNF. We do this in the following theorem. 
+%An interesting question at this point is when a formula $\alpha \in \eqso(\LL)$ can be converted in 
+%%$\Sigma$-
+%prenex normal form. 
+%%Of course, this would depend on the expressibility of $\LL$.
+%Interestingly, if $\LL$ contains $\logu{1}$, then it can be shown that $\alpha$ can be converted in
+%%we can always convert any formula in $\Sigma$-
+%%prenex 
+%this normal form. 
+\begin{theorem}\label{theo-pi1-pnf}
+There exists a formula $\alpha$ in $\QE{1}$ that is not equivalent to any formula in $\Sigma_1$-PNF. 
+%prenex normal form. 
+On the other hand, if $\logu{1} \subseteq \LL$, then 
+	%for 
+	every formula in
+	%$\alpha \in \eqso(\LL)$ 
+	$\eqso(\LL)$ can be rewritten in $\LL$-PNF. 
+	%there exists a formula $\beta \in \eqso(\LL)$ equivalent to $\alpha$ in $\Sigma$-
+%	prenex normal form.
+\end{theorem}
+
+\begin{figure*}
+
+\begin{center}
+\begin{tabular}{ccc}
+$\E{0}$ & $\subsetneq$ & 
+\end{tabular}
+\end{center}
+
+\caption{The relationship between the $\sfo$-hierarchy and the $\eqso$-hierarchy.\label{fir-sfo-eqso}}
+\end{figure*}
+
 As our first result, we show that in terms of containment the $\eqso$-hierarchy behaves as the $\sfo$-hierarchy:
@@
 $\E{i}$ and $\U{i}$ and their corresponding classes $\QE{i}$ and $\QU{i}$.
-Clearly, we have that $\E{i} \subseteq \QE{i}$ and $\U{i} \subseteq \QU{i}$.
-Indeed, each formula $\varphi(\bar{X}, \bar{x})$ in $\E{i}$ is equivalent to the formula $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ in $\QE{i}$, and likewise for $\U{i}$ and $\QU{i}$.
-Then it is left to know whether these containments are strict.
-%between the sharp and quantitative classes is strict or not. 
-For this, we study first whether a formula $\alpha$ in $\eqso(\LL)$ can be transformed into a 
-%\emph{prenex 
-normal form where sum quantifiers are restricted to be at the beginning of the formula. 
-Formally, we say that a formula $\alpha$ in $\eqso(\LL)$ is in \emph{prenex normal form} 
-%(or just prenex normal form) 
-if $\alpha$ is of the form
-%\[
-%\alpha := 
-$\sa{\bar{X}} \sa{\bar{x}} \varphi(\bar{X}, \bar{x})$,
-%\]
-where $\bar{X}$ and $\bar{x}$ are sequences of zero or more second-order and first-order variables, respectively, and $\varphi(\bar{X}, \bar{x})$ is a formula in $\LL$.
-In particular, 
-%note that a 
-each Boolean formula is in 
-%$\Sigma$-
-prenex normal form.
-Moreover, we say that a formula $\alpha$ is in \emph{sum normal form} if $\alpha$ is of the form $c + \Sigma_{i=1}^n \alpha_i$, where $c$ is a non-negative constant and each $\alpha_i$ is in 
-%$\Sigma$-
-prenex normal form.
-Next we show that each formula in $\eqso(\LL)$ can be converted in \emph{sum normal form} but not always in prenex normal form.
-\begin{theorem}\label{theo-pnf-snf}
-Every formula in $\eqso(\LL)$ can be rewritten in \emph{sum normal form}.
-Furthermore, there exists a formula $\alpha \in \QE{1}$ that is not equivalent to any formula in $\QE{1}$ that is in prenex normal form.
-%such that there does not exist a formula in $\Sigma$-prenex normal form in $\QE{1}$ equivalent to $\alpha$.
-\end{theorem}
-\martin{Creo que el resultado de que existe una formula en $\QE{1}$ que no se puede pasar a prenex está mas relacionado con el teorema siguiente que con este.}
-
-%By the previous result, we know that there exists logics $\LL$ where no $\Sigma$-prenex normal form exists. 
-%Then 
-An interesting question at this point is when a formula $\alpha \in \eqso(\LL)$ can be converted in 
-%$\Sigma$-
-prenex normal form. 
-%Of course, this would depend on the expressibility of $\LL$.
-Interestingly, if $\LL$ contains $\logu{1}$, then it can be shown that $\alpha$ can be converted in
-%we can always convert any formula in $\Sigma$-
-%prenex 
-this normal form. 
-\begin{theorem}\label{theo-pi1-pnf}
-	If $\logu{1} \subseteq \LL$, then 
-	%for 
-	every formula in
-	%$\alpha \in \eqso(\LL)$ 
-	$\eqso(\LL)$ can be rewritten in 
-	%there exists a formula $\beta \in \eqso(\LL)$ equivalent to $\alpha$ in $\Sigma$-
-	prenex normal form.
-\end{theorem}
 

### 2016-12-30 LICS/syntactic.tex

@@
 A first prominent example of such problems is counting the number of perfect matching in a bipartite graph, as it is well-known that the problem of verifying whether there exists a perfect matching in a bipartite graph can be solved in polynomial time.
-Other prominent examples of such problems include counting the number of: satisfying assignments of a 2-CNF propositional formula \cite{V79b}, satisfying assignments of a DNF propositional formula \cite{DHK05}, simple paths from a source node to a target node in a directed graph \cite{V79b}, extension of a partial order to a linear order \cite{BW91} and Eulerian cycles in an undirected graph \cite{BW05}.
+Other prominent examples of such problems include counting the number of: satisfying assignments of a 2-CNF propositional formula \cite{V79b}, satisfying assignments of a DNF propositional formula \cite{DHK05}, simple paths from a source node to a target node in a directed graph \cite{V79b}, extensions of a partial order to a linear order \cite{BW91} and Eulerian cycles in an undirected graph \cite{BW05}.
 
@@
 A key observation here is that if a $\shp$-complete function admits an FPRAS, then its associated decision problem is in the complexity class $\bpp$ (Bounded-Error Probabilistic Polynomial-Time).
-Hence, under standard complexity-theoretical assumptions we can no hope for an FPRAS for a function in $\shp$ whose decision counterpart is $\np$-complete, and we have to concentrate on the class of counting problems with easy decision versions (in $\bpp$ or in a lower complexity class such as $\ptime$).
+Hence, under standard complexity-theoretical assumptions we cannot hope for an FPRAS for a function in $\shp$ whose decision counterpart is $\np$-complete, and we have to concentrate on the class of counting problems with easy decision versions (in $\bpp$ or in a lower complexity class such as $\ptime$).
 

### 2016-12-30 LICS/saluja.tex
mas cambios ...

--- 2016-12-29 LICS/saluja.tex
@@
 Inspired by the connection between $\shp$ and $\sfo$, a hierarchy of subclases of $\sfo$ was introduced in~\cite{SalujaST95} 
-%studied subclasses of  $\sfo$ syntactically 
 by restricting the alternation of quantifiers in Boolean formulas.
-%, defining what we call 
-Specifically, the $\sfo$-hierarchy consists of the 
-%they define 
+Specifically, the \emph{$\sfo$-hierarchy} consists of the 
 the classes $\E{i}$ and $\U{i}$ for every $i \geq 0$, where $\E{i}$ (resp., $\U{i}$) is defined as $\sfo$ but restricting the formulas used to be in $\loge{i}$ (resp., $\logu{i}$).
-%is there exists a for are defined where functions are defined by $\fo$-formulas in $\loge{i}$ and $\logu{i}$, respectively, 
-%for every $i \geq 0$. 
-By definition we have that $\U{0} = \E{0}$. Moreover, it is shown in~\cite{SalujaST95} that:
-% this function classes defined a finite hierarchy of the form:
+By definition ,we have that $\U{0} = \E{0}$. Moreover, it is shown in~\cite{SalujaST95} that:
 \[
-\E{0} \; \subsetneq \; \E{1} \; \subsetneq \; \U{1} \; \subsetneq \; \E{2} \; \subsetneq \; \U{2} \; = \; \#\fo 
+\E{0} \; \subsetneq \; \E{1} \; \subsetneq \; \U{1} \; \subsetneq \; \E{2} \; \subsetneq \; \U{2} \; = \; \sfo 
 \]
 In light of the framework introduced in this paper, natural extensions of these classes are obtained by considering 
-%in light of the $\eqso$ logic. Specifically, we consider 
-%the classes 
-$\eqso(\loge{i})$ and $\eqso(\logu{i})$ for every $i \geq 0$, which form the $\eqso$-hierarchy.
-%, where the boolean logic is restricted to $\loge{i}$ and $\logu{i}$, respectively. 
-%We denote each class by $\QE{i}$ and $\QU{i}$ for short. 
-As our first result, we show that in terms of containment the $\eqso$-hierarchy behaves as the $\sfo$-hierarchy:
-%$\eqso$ also defined a finite hierarchy similar than in~\cite{SalujaST95} that we called the $\eqso$-hierarchy.
-\begin{proposition}
-\begin{multline*}
-\; \QE{0} \; \subsetneq \; \QE{1} \; \subsetneq \; \QU{1} \; \subsetneq \\ \QE{2} \; \subsetneq \; \QU{2} \; = \; \eqso(\fo)
-\end{multline*}
-\end{proposition}
-As our second result, we establish precise connections between 
-%A natural question at this point is what is 
-%the connection between 
-$\E{i}$ and $\U{i}$ and their corresponding classes $\QE{i}$ and $\QU{i}$.
+$\eqso(\loge{i})$ and $\eqso(\logu{i})$ for every $i \geq 0$, which form the \emph{$\eqso(\fo)$-hierarchy}.
 Clearly, we have that $\E{i} \subseteq \QE{i}$ and $\U{i} \subseteq \QU{i}$.
 Indeed, each formula $\varphi(\bar{X}, \bar{x})$ in $\E{i}$ is equivalent to the formula $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ in $\QE{i}$, and likewise for $\U{i}$ and $\QU{i}$.
-Then it is left to know whether these containments are strict.
-%between the sharp and quantitative classes is strict or not. 
-For this, we study first whether a formula $\alpha$ in $\eqso(\LL)$ can be transformed into a 
-%\emph{prenex 
-normal form where sum quantifiers are restricted to be at the beginning of the formula. 
-Formally, we say that a formula $\alpha$ in $\eqso(\LL)$ is in \emph{prenex normal form} 
-%(or just prenex normal form) 
+But what is the exact relationship between these two hierarchies?
+To answer this question, we start by introducing two normal forms for $\eqso(\LL)$ that helps us to characterize the expressive power of this quantitative logic.
+A formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-prenex normal form ($\LL$-PNF)} 
 if $\alpha$ is of the form
-%\[
-%\alpha := 
 $\sa{\bar{X}} \sa{\bar{x}} \varphi(\bar{X}, \bar{x})$,
-%\]
-where $\bar{X}$ and $\bar{x}$ are sequences of zero or more second-order and first-order variables, respectively, and $\varphi(\bar{X}, \bar{x})$ is a formula in $\LL$.
-In particular, 
-%note that a 
-each Boolean formula is in 
-%$\Sigma$-
-prenex normal form.
-Moreover, we say that a formula $\alpha$ is in \emph{sum normal form} if $\alpha$ is of the form $c + \Sigma_{i=1}^n \alpha_i$, where $c$ is a non-negative constant and each $\alpha_i$ is in 
-%$\Sigma$-
-prenex normal form.
-Next we show that each formula in $\eqso(\LL)$ can be converted in \emph{sum normal form} but not always in prenex normal form.
+where $\bar{X}$ and $\bar{x}$ are sequences of zero or more second-order and first-order variables, respectively, and $\varphi(\bar{X}, \bar{x})$ is a formula in $\LL$. Notice that 
+a formula $\varphi(\bar{X}, \bar{x})$ in $\sh{\LL}$ is equivalent to the formula $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ in $\LL$-PNF. 
+Moreover, a formula $\alpha$ is in \emph{$\LL$-sum normal form ($\LL$-SNF)} if $\alpha$ is of the form $c + \Sigma_{i=1}^n \alpha_i$, where $c$ is a non-negative constant and each $\alpha_i$ is in 
+$\LL$-PNF.
 \begin{theorem}\label{theo-pnf-snf}
-Every formula in $\eqso(\LL)$ can be rewritten in \emph{sum normal form}.
-Furthermore, there exists a formula $\alpha \in \QE{1}$ that is not equivalent to any formula in $\QE{1}$ that is in prenex normal form.
-%such that there does not exist a formula in $\Sigma$-prenex normal form in $\QE{1}$ equivalent to $\alpha$.
+Every formula in $\eqso(\LL)$ can be rewritten in $\LL$-SNF.
+\end{theorem}
+Therefore, to unveil the relationship between the $\sfo$-hierarchy and the $\eqso(\fo)$-hierarchy, we need to understand the boundary between PNF and SNF. We do this in the following theorem. 
+\begin{theorem}\label{theo-pi1-pnf}
+For $i = 0,1$, there exists a formula $\alpha_i$ in $\QE{i}$ that is not equivalent to any formula in $\Sigma_i$-PNF. 
+On the other hand, if $\logu{1} \subseteq \LL$, then 
+	every formula in
+	$\eqso(\LL)$ can be rewritten in $\LL$-PNF. 
 \end{theorem}
 
-%By the previous result, we know that there exists logics $\LL$ where no $\Sigma$-prenex normal form exists. 
-%Then 
-An interesting question at this point is when a formula $\alpha \in \eqso(\LL)$ can be converted in 
-%$\Sigma$-
-prenex normal form. 
-%Of course, this would depend on the expressibility of $\LL$.
-Interestingly, if $\LL$ contains $\logu{1}$, then it can be shown that $\alpha$ can be converted in
-%we can always convert any formula in $\Sigma$-
-%prenex 
-this normal form. 
-\begin{theorem}\label{theo-pi1-pnf}
-	If $\logu{1} \subseteq \LL$, then 
-	%for 
-	every formula in
-	%$\alpha \in \eqso(\LL)$ 
-	$\eqso(\LL)$ can be rewritten in 
-	%there exists a formula $\beta \in \eqso(\LL)$ equivalent to $\alpha$ in $\Sigma$-
-	prenex normal form.
-\end{theorem}
+\begin{figure*}
+\begin{center}
+\begin{tikzpicture}
+\node[rectw] (n1) {$\E{0}$};
+\node[rectw, right=0.5cm of n1] (n2) {};
+\node[rectw, above=0.5cm of n2] (n3) {$\E{1}$}
+	edge[draw=white] node {\rotatebox{45}{$\subsetneq$}} (n1);
+\node[rectw, below=0.5cm of n2] (n4) {$\QE{0}$}
+        edge[draw=white] node {\rotatebox{315}{$\subsetneq$}} (n1);
+\node[rectw, right=0.5cm of n2] (n5) {$\QE{1}$}
+        edge[draw=white] node {\rotatebox{315}{$\subsetneq$}} (n3)
+        edge[draw=white] node {\rotatebox{45}{$\subsetneq$}} (n4);
+\node[rectw, right=0.5cm of n5] (n6) {$\U{1}$}       
+        edge[draw=white] node {$\subsetneq$} (n5);
+\node[rectw, right=0.5cm of n6] (n7) {$\QU{1}$}       
+        edge[draw=white] node {$=$} (n6);
+\node[rectw, right=0.5cm of n7] (n8) {$\E{2}$}       
+        edge[draw=white] node {$\subsetneq$} (n7);
+\node[rectw, right=0.5cm of n8] (n9) {$\QE{2}$}       
+        edge[draw=white] node {$=$} (n8);        
+\node[rectw, right=0.5cm of n9] (n10) {$\U{2}$}       
+        edge[draw=white] node {$\subsetneq$} (n9);
+\node[rectw, right=0.5cm of n10] (n11) {$\QU{2}$}       
+        edge[draw=white] node {$=$} (n10); 
+\node[rectw, right=0.5cm of n11] (n12) {$\sfo$}       
+        edge[draw=white] node {$=$} (n11); 
+\end{tikzpicture}
+\end{center}
 
-Theorems \ref{theo-pnf-snf} and \ref{theo-pi1-pnf} are instrumental in answering our question of what is the relationship between the $\sfo$-hierarchy and the $\eqso$-hierarchy. 
-%The previous results gives the connection between the hierarchy in \cite{SalujaST95} and the $\eqso$-hierarchy. 
-Indeed, if $\LL$ contains $\logu{1}$, then we have that $\sh{\LL}$ is equal to $\eqso(\LL)$ since each formula in $\eqso(\LL)$ can be converted in prefix normal form and, therefore, it is equivalent to a formula in $\sh{\LL}$.
-The following proposition summarizes these results, also including the cases of $\E{0}$ and $\E{1}$.
-%Unfortunately, this is not the case for $\loge{0}$ and $\loge{1}$ as the following result shows.
+\caption{The relationship between the $\sfo$-hierarchy and the $\eqso(\fo)$-hierarchy, where $\E{1}$ and $\QE{0}$ are incomparable. \label{fig-sfo-eqso}}
+\end{figure*}
+
+As a consequence of Theorems \ref{theo-pnf-snf} and \ref{theo-pi1-pnf}, we obtain that $\E{i} \subsetneq \QE{i}$ for $i = 1,2$, and that $\sh{\LL} = \eqso(\LL)$ for every Boolean logic containing $\Pi_1$.
+The following proposition completes our picture of the relationship between the $\sfo$-hierarchy and the $\eqso(\fo)$-hierarchy, which is summarized in Figure \ref{fig-sfo-eqso}.
 \begin{proposition}
-	The classes $\E{0}$ and $\E{1}$ are strictly contained in $\QE{0}$ and $\QE{1}$, respectively.
-Moreover, the classes $\U{1}$, $\E{2}$, and $\U{2}$ are equivalent with $\QU{1}$, $\QE{2}$, and $\QU{2}$, respectively.
+The following properties hold:
+\begin{itemize}
+\item $\QE{0} \subsetneq \QE{1}$
+\item $\QE{1} \subsetneq \QU{1}$
+\item $\QE{0}$ and $\E{1}$ are incomparable, that is, $\E{1} \not\subseteq \QE{0}$ and $\QE{0} \not\subseteq \E{1}$
+\end{itemize}
 \end{proposition}
-The previous result shows that the classes $\QE{i}$ and $\QU{i}$ are more robust than the classes $\E{i}$ and $\U{i}$: they are closed under binary and sum quantifiers but the other not necessarily.
-
-Now, we study the complexity classes describe by this hierarchies.
-As the following result shows, $\eqso(\loge{0})$ defines only tractable counting functions and $\eqso(\loge{1})$ intractable counting functions but with an tractable decision problems.
-\begin{proposition}
-All functions defined in $\eqso(\loge{0})$ and $\eqso(\loge{1})$ can be computed in $\fp$ and $\totp$, respectively. Furthermore, every function defined in $\eqso(\loge{1})$ has a FPTRAS.
-\end{proposition}
-Therefore, in terms of counting complexity, the $\eqso$-hierarchy behaves exactly the same as the $\#\fo$-hierarchy.
-
-The next step is to study the closure properties of $\eqso$-hierarchy. 
-An advantage of the $\eqso$-hierarchy is that, by its language syntax, all the classes are closed under addition and first and second order sum.
-So, the first question is whether the multiplicative operators in $\qso$ can be defined in $\eqso(\LL)$.
-As the following result shows, if $\LL$ is closed under conjunction, then the binary product can be defined in  $\eqso(\LL)$.
-\begin{theorem}\label{theo:binary-prod}
-	If $\LL$ is closed under conjunction, then binary product can be defined in $\eqso(\LL)$.
-\end{theorem}
-The next question is whether the hierarchy is closed under subtraction.
-Formally, for any pair of functions $f,g$, we define $f - g$ as the function such that $(f - g)(\A) = f(\A)-g(\A)$ whenever $f(\A)>g(\A)$ and $0$ otherwise.
-As the next result shows, all classes in the $\eqso$-hierarchy is not closed under subtraction unless ${\sc P} = {\sc NP}$
-\begin{theorem} \label{sub-pnp}
-If $\eqso(\loge{i})$ or $\eqso(\logu{i})$ is closed under subtraction for $i > 0$, then {\sc P} = {\sc NP}.
-\end{theorem}
-\cristian{Martin, el resultado que tienes en el apendice se generaliza trivialmente para todas las clases ya que todas contienen la clase $\eqso(\loge{0})$.}
-
-By the previous result, we know that functions in the $\eqso$ hierarchy are unlikely to be closed under subtraction.
-Then, a natural restriction to this question is to ask whether these classes are closed under subtraction by one, namely, if $\CC$ is a class of functions and $f \in \CC$, is $f-1 \in \CC$ where $1$ is the constant function that outputs $1$ for every structure.
-We do not know $\E{1}$ is closed under subtraction by one. However, if we extend $\logex{1}$ with $\fo$ predicates we can show that this new fragment is closed under subtraction by one.
-\begin{theorem} \label{sigmafo-minusone}
-	$\eqso(\logex{1})$ is closed under substraction by one.
-\end{theorem}
-
-% We are interested in, for each $\fo$-fragment $\LL$, the biggest fragment of $\eqso$ that is contained in $\#\LL$.
-%Let $\LL = \loge{0}$:
-%\begin{theorem} \label{one-sigma-zero}
-%	Positive constant functions are not expressible in $\E{0}$
-%\end{theorem}
-%\begin{corollary}
-%	If a fragment of $\eqso(\loge{0})$ is contained in $\E{0}$, its grammar does not allow sole constants.
-%\end{corollary}
-%\begin{conjecture}
-%	Sum is not expressible in $\E{0}$
-%\end{conjecture}
-%\begin{theorem} \label{mult-sigma-zero}
-%	$\sqso(\loge{0})$ with binary product is contained in $\E{0}$.
-%\end{theorem}
-%\begin{theorem} \label{fo-prod-sigma-zero}
-%	If an extension of $\sqso(\loge{0})$ is contained in $\E{0}$, its grammar does not allow first-order product.
-%\end{theorem}
-
-%For every logic $\LL$, we define an $\LL$-extended quantifier-free (QF) formula as follows:
-%\begin{eqnarray*}
-%	\varphi &::=& \alpha, \alpha \text{ is an $\LL$-formula} \ \mid \\
-%	&& X_i(x_1,\dots,x_{a_i}), i\in\N \ \mid \ \\
-%	&& (\neg \varphi) \ \mid \ (\varphi \wedge \varphi) \ \mid \ (\varphi \vee \varphi).
-%\end{eqnarray*}
-%
-%We define syntactically the fragments $\logex{i}$ and $\logux{i}$ according to the following grammar:
-%\begin{align*}
-%\logex{0} = \logux{0} &::= \varphi , \varphi \mbox{ is an $\fo$-extended QF formula,} \\
-%\logex{i+1} &::= \logux{i} \ \mid \ \exists x\, \logex{i+1}, \\
-%\logux{i+1} &::= \logex{i} \ \mid \ \forall x\, \logux{i+1}.
-%\end{align*}
-
-%We see that many of the results in Saluja et. al. \cite{SalujaST95} for $\#\LL$ still apply in $\eqso(\LL)$ for a given fragment $\LL$:
-%
-%\begin{theorem} \label{eqso-sigma-zero-in-fp}
-%	For every $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$, the function $f$ over $\R$ defined as $f(\enc(\A)) = \sem{\alpha}(\A)$ is in $\fp$.
-%\end{theorem}
-%
-%\begin{theorem} \label{eqso-sigma-one-in-eqso-pi-one}
-%	For every $\eqso(\loge{1})$ formula $\alpha$ over a signature $\R$ there exists a $\eqso(\logu{1})$ formula $\beta$ over $\R$ such that $\sem{\alpha}(\A) = \sem{\beta}(\A)$ for every $\A\in\ostr[\R]$.
-%\end{theorem}
-%
-%
-%The {\em decision problem} associated to a function $f$ is defined by the language $L_f = \{\A \in \str \mid f(\A) > 0\}$.
-%
-%\begin{theorem} \label{decisionptime}
-%	The decision problem associated to a function in $\eqso(\logex{1})$ is in \textsc{P}.
-%\end{theorem}
-
-%For a given pair of functions $f,g$, we define $f \dotminus g$ as follows:
-%\begin{eqnarray*}
-%	(f \dotminus g)(\A) =
-%	\begin{cases}
-%		f(\A)-g(\A), & \text{if }f(\A)>g(\A) \\
-%		0, & \text{if }f(\A) \leq g(\A).
-%	\end{cases}
-%\end{eqnarray*}
-%for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction} if for every pair of functions $f,g \in \F$, it holds that $f \dotminus g \in \F$.
-%
-%\begin{theorem} \label{sub-pnp}
-%	If $\eqso(\loge{1})$ is closed under substraction, then {\sc P} = {\sc NP}.
-%\end{theorem}
-%
-%\begin{theorem} \label{sigma1strict}
-%	$\eqso(\loge{1}) \subsetneq \eqso(\logex{1})$
-%\end{theorem}
-%
-%For a given function $f$, we define $f \dotminus 1$ as follows:
-%\begin{eqnarray*}
-%	f \dotminus 1(\A) =
-%	\begin{cases}
-%		f(\A)-1, & \text{if }f(\A) > 0 \\
-%		0, & \text{if }f(\A) = 0.
-%	\end{cases}
-%\end{eqnarray*}
-%for every $\L$-structure $\A \in \str$. A function class $\F$ is {\em closed under substraction by one} if for every function $f \in \F$, it holds that $f \dotminus 1 \in \F$.
-%
-%\begin{theorem} \label{sigmafo-minusone}
-%	$\eqso(\logex{1})$ is closed under substraction by one.
-%\end{theorem}
-%
-%\begin{theorem} \label{dnf-pars}
-%	{\sc \#DNF} is hard for $\eqso(\loge{1})$ under parsimonious reductions. 
-%\end{theorem}
-%
-%\begin{theorem} \label{nplusone-strict}
-%	$\U{1}$ with $n$ open first-order variables is properly contained in $\U{1}$ with $n+1$ open first-order variables for $n\in\N$.  
-%\end{theorem}

### 2016-12-30 LICS/saluja.tex
mas

@@
 For $i = 0,1$, there exists a formula $\alpha_i$ in $\QE{i}$ that is not equivalent to any formula in $\Sigma_i$-PNF. 
-On the other hand, if $\logu{1} \subseteq \LL$, then 
+On the other hand, if $\logu{1} \subseteq \LL$ and $\LL$ is closed under conjunction and disjunction, then 
 	every formula in
@@
 
-As a consequence of Theorems \ref{theo-pnf-snf} and \ref{theo-pi1-pnf}, we obtain that $\E{i} \subsetneq \QE{i}$ for $i = 1,2$, and that $\sh{\LL} = \eqso(\LL)$ for every Boolean logic containing $\Pi_1$.
+As a consequence of Theorems \ref{theo-pnf-snf} and \ref{theo-pi1-pnf}, we obtain that $\E{i} \subsetneq \QE{i}$ for $i = 0,1$, and that $\sh{\LL} = \eqso(\LL)$ for $\LL = \Pi_1$, $\LL = \Sigma_2$ and $\LL = \Pi_2$.
 The following proposition completes our picture of the relationship between the $\sfo$-hierarchy and the $\eqso(\fo)$-hierarchy, which is summarized in Figure \ref{fig-sfo-eqso}.
-\begin{proposition}
+\begin{proposition}\label{prop-rest}
 The following properties hold:
 \begin{itemize}
-\item $\QE{0} \subsetneq \QE{1}$
+\item $\QE{0} \not\subseteq \E{0}$
+\item $\QE{0}$ and $\E{1}$ are incomparable, that is, $\E{1} \not\subseteq \QE{0}$ and $\QE{0} \not\subseteq \E{1}$
 \item $\QE{1} \subsetneq \QU{1}$
-\item $\QE{0}$ and $\E{1}$ are incomparable, that is, $\E{1} \not\subseteq \QE{0}$ and $\QE{0} \not\subseteq \E{1}$
 \end{itemize}

### 2016-12-30 LICS/syntactic-clo-horn.tex

--- 2016-12-28 LICS/horn-counting.tex
@@
 \newcommand{\pD}{\textit{D}}
+
+The previous result shows that the classes $\QE{i}$ and $\QU{i}$ are more robust than the classes $\E{i}$ and $\U{i}$: they are closed under binary and sum quantifiers but the other not necessarily.
+
+Now, we study the complexity classes describe by this hierarchies.
+As the following result shows, $\eqso(\loge{0})$ defines only tractable counting functions and $\eqso(\loge{1})$ intractable counting functions but with an tractable decision problems.
+\begin{proposition} \label{prop:qe0-fp-qe1-totp-fptras}
+All functions defined in $\eqso(\loge{0})$ and $\eqso(\loge{1})$ can be computed in $\fp$ and $\totp$, respectively. Furthermore, every function defined in $\eqso(\loge{1})$ has a FPTRAS.
+\end{proposition}
+Therefore, in terms of counting complexity, the $\eqso$-hierarchy behaves exactly the same as the $\#\fo$-hierarchy.
+
+The next step is to study the closure properties of $\eqso$-hierarchy. 
+An advantage of the $\eqso$-hierarchy is that, by its language syntax, all the classes are closed under addition and first and second order sum.
+So, the first question is whether the multiplicative operators in $\qso$ can be defined in $\eqso(\LL)$.
+As the following result shows, if $\LL$ is closed under conjunction, then the binary product can be defined in  $\eqso(\LL)$.
+\begin{theorem}\label{theo:binary-prod}
+	If $\LL$ is closed under conjunction, then binary product can be defined in $\eqso(\LL)$.
+\end{theorem}
+The next question is whether the hierarchy is closed under subtraction.
+Formally, for any pair of functions $f,g$, we define $f - g$ as the function such that $(f - g)(\A) = f(\A)-g(\A)$ whenever $f(\A)>g(\A)$ and $0$ otherwise.
+As the next result shows, all classes in the $\eqso$-hierarchy is not closed under subtraction unless ${\sc P} = {\sc NP}$
+\begin{theorem} \label{sub-pnp}
+If $\eqso(\loge{i})$ or $\eqso(\logu{i})$ is closed under subtraction for $i > 0$, then {\sc P} = {\sc NP}.
+\end{theorem}
+\cristian{Martin, el resultado que tienes en el apendice se generaliza trivialmente para todas las clases ya que todas contienen la clase $\eqso(\loge{0})$.}
+
+By the previous result, we know that functions in the $\eqso$ hierarchy are unlikely to be closed under subtraction.
+Then, a natural restriction to this question is to ask whether these classes are closed under subtraction by one, namely, if $\CC$ is a class of functions and $f \in \CC$, is $f-1 \in \CC$ where $1$ is the constant function that outputs $1$ for every structure.
+We do not know $\E{1}$ is closed under subtraction by one. However, if we extend $\logex{1}$ with $\fo$ predicates we can show that this new fragment is closed under subtraction by one.
+\begin{theorem} \label{sigmafo-minusone}
+	$\eqso(\logex{1})$ is closed under substraction by one.
+\end{theorem}
 
@@
 
-\begin{proposition}\label{prop-uhorn-pe}
-$\eqso(\uhorn) \subseteq \pe$
+\begin{proposition}\label{prop:uhorn-pe}
+$\eqso(\uhorn) \subseteq \totp$
 \end{proposition}
@@
 
-\begin{proposition}\label{prop-ehorn-pe}
-$\eqso(\ehorn) \subseteq \pe$.
+\begin{proposition}\label{prop:ehorn-pe}
+$\eqso(\ehorn) \subseteq \totp$.
 \end{proposition}
@@
 
-\begin{proposition}\label{prop-hsat-not-sigma2}
+\begin{proposition}\label{prop:hsat-not-sigma2}	
 $\chsat \not\in \eqso(\loge{2})$.

### 2016-12-31 LICS/saluja.tex
mas cammbios

@@
 the classes $\E{i}$ and $\U{i}$ for every $i \geq 0$, where $\E{i}$ (resp., $\U{i}$) is defined as $\sfo$ but restricting the formulas used to be in $\loge{i}$ (resp., $\logu{i}$).
-By definition ,we have that $\U{0} = \E{0}$. Moreover, it is shown in~\cite{SalujaST95} that:
+By definition, we have that $\U{0} = \E{0}$. Moreover, it is shown in~\cite{SalujaST95} that:
 \[
@@
 But what is the exact relationship between these two hierarchies?
-To answer this question, we start by introducing two normal forms for $\eqso(\LL)$ that helps us to characterize the expressive power of this quantitative logic.
+To answer this question, we first introduce two normal forms for $\eqso(\LL)$ that helps us to characterize the expressive power of this quantitative logic.
 A formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-prenex normal form ($\LL$-PNF)} 
@@
 a formula $\varphi(\bar{X}, \bar{x})$ in $\sh{\LL}$ is equivalent to the formula $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ in $\LL$-PNF. 
-Moreover, a formula $\alpha$ is in \emph{$\LL$-sum normal form ($\LL$-SNF)} if $\alpha$ is of the form $c + \Sigma_{i=1}^n \alpha_i$, where $c$ is a non-negative constant and each $\alpha_i$ is in 
+Moreover, a formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-sum normal form ($\LL$-SNF)} if $\alpha$ is of the form $c + \Sigma_{i=1}^n \alpha_i$, where $c$ is a non-negative constant and each $\alpha_i$ is in
 $\LL$-PNF.
@@
 
-As a consequence of Theorems \ref{theo-pnf-snf} and \ref{theo-pi1-pnf}, we obtain that $\E{i} \subsetneq \QE{i}$ for $i = 0,1$, and that $\sh{\LL} = \eqso(\LL)$ for $\LL = \Pi_1$, $\LL = \Sigma_2$ and $\LL = \Pi_2$.
-The following proposition completes our picture of the relationship between the $\sfo$-hierarchy and the $\eqso(\fo)$-hierarchy, which is summarized in Figure \ref{fig-sfo-eqso}.
+As a consequence of Theorems \ref{theo-pnf-snf} and \ref{theo-pi1-pnf}, we obtain that $\E{i} \subsetneq \QE{i}$ for $i = 0,1$, and that $\sh{\LL} = \eqso(\LL)$ for $\LL$ equal to  $\Pi_1$, $\Sigma_2$ or $\Pi_2$.
+The following proposition completes our picture of the relationship between the $\sfo$-hierarchy and the $\eqso(\fo)$-hierarchy.
 \begin{proposition}\label{prop-rest}
@@
 \begin{itemize}
-\item $\QE{0} \not\subseteq \E{0}$
 \item $\QE{0}$ and $\E{1}$ are incomparable, that is, $\E{1} \not\subseteq \QE{0}$ and $\QE{0} \not\subseteq \E{1}$
@@
 \end{proposition}
+The results of this section are summarized in Figure \ref{fig-sfo-eqso}.

### 2016-12-31 LICS/syntactic-clo.tex
mas cammbios

--- 2016-12-30 LICS/syntactic-clo-horn.tex
@@
 %!TEX root = main.tex
-\newcommand{\pP}{\textit{P}}
-\newcommand{\pN}{\textit{N}}
-\newcommand{\pV}{\textit{V}}
-\newcommand{\pT}{\textit{T}}
-\newcommand{\pA}{\textit{A}}
-\newcommand{\pNC}{\textit{NC}}
-\newcommand{\pD}{\textit{D}}
 
@@
 \end{theorem}
-
-A positive literal is a formula of the form $X(\x)$, where $X$ is a second-order variable and $\x$ is a tuple of first-order variables, and a negative literal is a formula of the form $\exists \v \, \neg X(\u,\v)$, where $\u$ and $\v$ are tuples of first-order variables.
-Given a relational signature $\R$, a clause over $\R$ is a formula of the form:
-$$
-\forall \x \, (\varphi_1 \vee \cdots \vee \varphi_n),
-$$
-where each $\varphi_i$ ($1 \leq i \leq n$) is either a positive literal, a negative literal or an \fo-formula over $\R$.
-A clause is said to be Horn if it contains at most one positive literal, and a formula is said to be Horn if it is a conjunction of Horn clauses over a relational signature $\R$.
-With this terminology, we define $\uhorn$ as the set of formulas $\psi$ such that $\psi$ is a conjunction of Horn clauses over a relational signature $\R$.
-
-\begin{proposition}\label{prop:uhorn-pe}
-$\eqso(\uhorn) \subseteq \totp$
-\end{proposition}
-
-\begin{example} \label{ex-hornsat-esop1}
-Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot)\}$.
-This vocabulary is used as follows to encode a Horn formula.
-A fact $\pP(c,x)$ indicates that propositional variable $x$ is a disjunct in a clause $c$, while $\pN(c,x)$ indicates that $\neg x$ is a disjunct in $c$.
-Furthermore, $\pV(x)$ holds if  $x$ is a propositional variable, and $\pNC(c)$ holds if $c$ is a clause containing only negative literals, that is, $c$ is of the form $(\neg x_1 \vee \cdots \vee \neg x_n)$.
-
-To encode $\chsat$, we define an \so-formula $\varphi(\pT)$ over $\R$, where $\pT$ is a unary predicate, such that for every Horn formula $\theta$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \varphi(\pT)}(\A)$.
-In particular, $\pT(x)$ holds if and only if $x$ is a propositional variable that is assigned value 1.
-More specifically, $\varphi(\pT)$ is defined as follows:
-\begin{align*}
-&\forall x \, (\pT(x) \to \pV(x)) \ \wedge\\
-&\forall c \, (\pNC(c) \to \exists x \, (\pN(c,x) \wedge \neg \pT(x))) \ \wedge\\
-&\forall c \forall x \, ([\pP(c,x) \wedge \forall y \, (\pN(c,y) \to \pT(y))] \to \pT(x)).
-\end{align*}
-Given that $\uhorn$ is designed with the goal in mind of capturing $\chsat$, we expect $\varphi(\pT)$ to be a formula in $\uhorn$.
-However, if we rewrite it as a conjunction of clauses we obtain the following:
-\begin{align*}
-&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \pNC(c) \vee \exists x \, (\pN(c,x) \wedge \neg \pT(x)))\ \wedge\\
-&\forall c \forall x \, (\neg \pP(c,x) \vee \exists y \, (\pN(c,y) \wedge \neg \pT(y)) \vee \pT(x)).
-\end{align*}
-The resulting formula $\varphi(\pT)$ is not in $\uhorn$, but it can be easily transformed into a formula in this class  by introducing an auxiliary binary predicate $\pA$ defined as follows:
-\begin{align*}
-\forall c \forall x \, (\neg \pA(c,x) \leftrightarrow [\pN(c,x) \wedge \neg \pT(x)]).
-\end{align*}
-In this way, we obtain the following formula $\psi(\pT,\pA)$ in $\uhorn$:
-\begin{align*}
-&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \neg\textit{T}(x)).
-\end{align*}
-This formula effectively defines $\chsat$
-as for every Horn formula $\theta$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
-Therefore, we conclude that $\chsat \in \eqso(\uhorn)$.
-\end{example}
-We extend the definition of $\uhorn$ to allow existential quantification.
-More precisely, a formula $\varphi$ is in $\ehorn$ if $\varphi$ is of the form $\exists \bar x \, \psi$ with $\psi$ a Horn formula.
-Interestingly, it hold that $\cdnf \in \eqso(\ehorn)$ and
-
-\begin{proposition}\label{prop:ehorn-pe}
-$\eqso(\ehorn) \subseteq \totp$.
-\end{proposition}
-A natural question at this point is whether in the definitions of $\uhorn$ and $\ehorn$, it is necessary to allow negative literals of the form $\exists \v \, \neg X(\u,\v)$.
-The following result shows that it is indeed the case:
-
-\begin{proposition}\label{prop:hsat-not-sigma2}	
-$\chsat \not\in \eqso(\loge{2})$.
-\end{proposition}
-We conclude this section by showing that a natural extension of $\chsat$ is $\eqso(\ehorn)$-complete under parsimonious reductions. We define the decision problem:
-\begin{multline*}
-\dhsat = \{\Phi \mid \Phi \text{ is a disjunction of}\\  \text{Horn formulas and $\Phi$ is satisfiable}\},
-\end{multline*}
-and the counting problem $\shdhsat$ as a function that counts all satisfying assignments of a formula $\Phi$ that is a disjunction of Horn formulas.
-
-\begin{theorem} \label{sigma2hard}
-	$\shdhsat$ is $\eqso(\ehorn)$-complete under parsimonious reductions. 
-\end{theorem}

### 2016-12-31 LICS/syntactic.tex
mas cambios ...
mas cammbios

@@
 More specifically, it was shown in  \cite{Valiant79}  that this problem is $\shp$-complete.
-As a consequence of this result the problem of computing the number of perfect matchings in a bipartite graph was also shown to be $\shp$-complete.
+As a consequence of this result the problem of computing the number of perfect matchings of a bipartite graph was also shown to be $\shp$-complete.
 Since then  many counting problems have been proved to be $\shp$-complete \cite{V79b,PB83,P86,L86,BW91,HMRS98,BW05,DS12, PS13,PS14}.
 Among them, problems having easy decision counterparts play a fundamental role, as a counting problem with a hard decision version is expected to be hard.
-A first prominent example of such problems is counting the number of perfect matching in a bipartite graph, as it is well-known that the problem of verifying whether there exists a perfect matching in a bipartite graph can be solved in polynomial time.
+A first prominent example of such problems is counting the number of perfect matching of a bipartite graph, as it is well-known that the problem of verifying whether there exists a perfect matching in a bipartite graph can be solved in polynomial time.
 Other prominent examples of such problems include counting the number of: satisfying assignments of a 2-CNF propositional formula \cite{V79b}, satisfying assignments of a DNF propositional formula \cite{DHK05}, simple paths from a source node to a target node in a directed graph \cite{V79b}, extensions of a partial order to a linear order \cite{BW91} and Eulerian cycles in an undirected graph \cite{BW05}.
@@
 Notably, there exist $\shp$-complete functions that can be efficiently approximated as they admit FPRAS; for instance, there exist FPRAS for the problems of counting the number of satisfying assignments of a DNF propositional formula \cite{KL83} and the number of perfect matchings of a bipartite graph \cite{JSV04}.
-A key observation here is that if a $\shp$-complete function admits an FPRAS, then its associated decision problem is in the complexity class $\bpp$ (Bounded-Error Probabilistic Polynomial-Time).
-Hence, under standard complexity-theoretical assumptions we cannot hope for an FPRAS for a function in $\shp$ whose decision counterpart is $\np$-complete, and we have to concentrate on the class of counting problems with easy decision versions (in $\bpp$ or in a lower complexity class such as $\ptime$).
+A key observation here is that if a $\shp$-complete function admits an FPRAS, then its associated decision problem is in the complexity class $\bpp$ (Bounded-Error Probabilistic Polynomial-Time \cite{G77}).
+Hence, under standard complexity-theoretical assumptions, we cannot hope for an FPRAS for a function in $\shp$ whose decision counterpart is $\np$-complete, and we have to concentrate on the class of counting problems with easy decision versions (in $\bpp$ or in a lower complexity class such as $\ptime$).
 
-The importance of the class counting problems with easy decision counterparts has motivated the search of robust definitions of classes of functions in $\shp$ with easy decision versions \cite{PagourtzisZ06}.
-In this section, we use the framework developed in this paper to address this problem.
-More specifically, we introduce in Section \ref{sec-hier-shp} a hierarchy of
+The importance of the class counting problems with easy decision counterparts has motivated the search of robust classes of functions in $\shp$ with this property \cite{PagourtzisZ06}.
+But the key question here is what should be considered a {\em robust} class.
+A first condition to be met is that the class should have a syntactic definition that allows for the existence of natural complete problems.
+Special attention has to be paid here to the notion of reduction used for completeness.
+Notice that under the notion of Cook reduction, originally used in \cite{Valiant79}, the problems $\cdnf$ and $\csat$ are $\shp$-complete.
+However, $\cdnf$ has an easy decision counterpart and admits an FPRAS, while $\csat$ does not satisfy these conditions unless $\ptime = \np$.
+A second condition to be met has to do with the closure properties satisfied by the class, which is a common theme when studying function complexity classes \cite{OH93,FH08}.
+As in the cases of $\ptime$ and $\np$ that are closed under intersection and union, we expect our class to be closed under multiplication and sum; in fact, if $f$ and $g$ are counting functions with easy decision versions, then we clearly expect $f + g$ and $f \cdot g$ to have easy decision counterparts.
+For a more elaborated closure property, assume that $\textit{sat\_one}$ is a function that returns one plus the number of satisfying assignments of a propositional formula.
+The decision counterpart of such function is the language of all proposicional formulas $\varphi$ such that $\textit{sat\_one}(\varphi) > 0$.
+Clearly $\textit{sat\_one}$ is a $\shp$-complete function whose decision counterpart is trivial.
+But should $\textit{sat\_one}$ be part of a robust class of counting function with easy decision versions?
+The key insight here is that if a function in $\shp$ has an easy decision counterpart $L$, then as $L \in \np$ we expect  to have a polynomial-time algorithm that verifies whether $x \in L$ by constructing witnesses for $x$.
+As an example of this considered all the counting functions mentioned in the previous paragraphs and, in particular, the decision counterpart of $\cdnf$ that can be solved by explicitly constructing in polynomial-time a satisfying assignment for an input propositional formula.
+Moreover, if such an algorithm for constructing witnesses exists, then we also expect to be able to manipulate such witnesses and in some cases to remove them.
+In other words, we expect a robust class $\CC$ of counting functions with easy decision versions to be closed under subtraction by one, that is, if $g \in \CC$, then the function $g - 1$ should also be in $\CC$.
+Notice that, unless $\ptime = \np$, no such class can contain the function $\textit{sat\_one}$ because $\textit{sat\_one} - 1$ counts the number of satisfying assignments of a propositional formula.
 
-consider several fragments of $\eqso(\LL)$ where $\LL$ is a boolean logic contained in $\fo$, and we study
+In this section, we use the framework developed in this paper to address the problem of defining a robust class of functions with easy decision versions.
+More specifically, we use the framework to introduce in Section \ref{sec-hier-shp} a syntactic hierarchy of function complexity classes contained in $\shp$.
+Then this hierarchy is used in Section \ref{sec-clo} to define a class of functions with easy decision versions and good closure properties, and in Section \ref{sec-horn} to define a class of functions with easy decision versions and
+natural complete problems.
 
@@
 
-In this section we study the fragment of $\eqso(\LL)$ when $\LL$ is a boolean logic contained in $\fo$.
+%
+%
+%In this section we study the fragment of $\eqso(\LL)$ when $\LL$ is a boolean logic contained in $\fo$.
 We show that by restricting $\LL$ we can find different subclasses below $\shp$ with interesting computational and closure properties.
+%
+%
+%
+%From this point on, for each fragment $\FF$ of $\qso$, we will also use $\FF$ to refer to the class of functions defined by the formulas in $\FF$.
 
-\cite{OH93,FH08}
-
-From this point on, for each fragment $\FF$ of $\qso$, we will also use $\FF$ to refer to the class of functions defined by the formulas in $\FF$.
-
-\subsection{A counting hierarchy below $\shp$}
+\subsection{A syntactic hierarchy of function complexity classes contained in $\shp$}
 \label{sec-hier-shp}
@@
 
-\subsection{Counting hierarchy below $\shp$}
+\subsection{Defining a class of functions with easy decision versions and good closure properties}
+\label{sec-clo}
+\input{syntactic-clo}
 
-\subsection{Horn Counting Classes}
-\input{horn-counting}
+\subsection{Defining a class of functions with easy decision versions and natural complete problems}
+\label{sec-horn}
+\input{syntactic-horn}

### 2016-12-31 LICS/syntactic.tex
sin comentarios

@@
 More specifically, it was shown in  \cite{Valiant79}  that this problem is $\shp$-complete.
-As a consequence of this result the problem of computing the number of perfect matchings of a bipartite graph was also shown to be $\shp$-complete.
+As a consequence of this result the $\shp$-completeness of the problem of computing the number of perfect matchings of a bipartite graph was established.
 Since then  many counting problems have been proved to be $\shp$-complete \cite{V79b,PB83,P86,L86,BW91,HMRS98,BW05,DS12, PS13,PS14}.
 Among them, problems having easy decision counterparts play a fundamental role, as a counting problem with a hard decision version is expected to be hard.
-A first prominent example of such problems is counting the number of perfect matching of a bipartite graph, as it is well-known that the problem of verifying whether there exists a perfect matching in a bipartite graph can be solved in polynomial time.
+Formally, the decision problem associated to a function $f : \Sigma^* \to \mathbb{N}$ is defined as $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$, and $f$ is said to have an easy decision version if $L_f \in \ptime$.
+A first prominent example of such functions is counting the number of perfect matching of a bipartite graph, as it is well-known that the problem of verifying whether there exists a perfect matching in a bipartite graph can be solved in polynomial time.
 Other prominent examples of such problems include counting the number of: satisfying assignments of a 2-CNF propositional formula \cite{V79b}, satisfying assignments of a DNF propositional formula \cite{DHK05}, simple paths from a source node to a target node in a directed graph \cite{V79b}, extensions of a partial order to a linear order \cite{BW91} and Eulerian cycles in an undirected graph \cite{BW05}.
@@
 But the key question here is what should be considered a {\em robust} class.
-A first condition to be met is that the class should have a syntactic definition that allows for the existence of natural complete problems.
+A first desirable condition is that the class has a syntactic definition that allows for the existence of natural complete problems \cite{P94}.
 Special attention has to be paid here to the notion of reduction used for completeness.
@@
 However, $\cdnf$ has an easy decision counterpart and admits an FPRAS, while $\csat$ does not satisfy these conditions unless $\ptime = \np$.
-A second condition to be met has to do with the closure properties satisfied by the class, which is a common theme when studying function complexity classes \cite{OH93,FH08}.
+Hence a more strict notion of reduction has to be considered; in particular, the notion of parsimonious reduction (to be defined later) satisfy that if a function $f$ is parsimoniously reducible to a function $g$, then $L_g \in \ptime$ implies that $L_f \in \ptime$ and the existence of an FPRAS for $g$ implies the existence of a FPRAS for $f$.
+A second desirable condition has to do with the closure properties satisfied by the class, which is a common theme when studying function complexity classes \cite{OH93,FH08}.
 As in the cases of $\ptime$ and $\np$ that are closed under intersection and union, we expect our class to be closed under multiplication and sum; in fact, if $f$ and $g$ are counting functions with easy decision versions, then we clearly expect $f + g$ and $f \cdot g$ to have easy decision counterparts.
 For a more elaborated closure property, assume that $\textit{sat\_one}$ is a function that returns one plus the number of satisfying assignments of a propositional formula.
-The decision counterpart of such function is the language of all proposicional formulas $\varphi$ such that $\textit{sat\_one}(\varphi) > 0$.
-Clearly $\textit{sat\_one}$ is a $\shp$-complete function whose decision counterpart is trivial.
+Clearly $\textit{sat\_one}$ is a $\shp$-complete function whose decision counterpart $L_{\textit{sat\_one}} = \{ \varphi \mid \textit{sat\_one}(\varphi) > 0 \}$ is trivial.
 But should $\textit{sat\_one}$ be part of a robust class of counting function with easy decision versions?

### 2016-12-31 LICS/preliminaries.tex
sin comentarios

@@
 where $R \in \R$, $\bar u$ is a tuple of (non-necessarily distinct) variables from $\fv$ whose length is $\arity(R)$, $X \in \sv$, $\bar v$ is a tuple of (non-necessarily distinct) variables from $\fv$ whose length is $\arity(X)$, and $x \in \fv$.
+
+\marcelo{Tenemos una definicion muy detallada de $\fo$ y $\so$, no es necesario para esta conferencia.
+Por otro lado seria bueno mencionar las definiciones de $\Sigma_i$ y $\Pi_i$, aunque son estandar es bueno decir que estamos considerando restricciones de $\fo$.}
 
@@
 
+\marcelo{La notacion $\tma_M(x)$ se usa en el paper, hay que definirla aqui.
+Donde vamos a usar la clase $\gp$?
+Para definir esta clase tambien necesitamos la notacion $\tmr_M(x)$, podriamos definirla aqui.
+}
+
 %If $\KK$ is a class of finite structures and $f$ is a function from a relational signature $\R$ to $\bbD$, we denote by $\res{f}{\KK}$ the restriction of $f$ to $\KK$, that is, a function such that: the domain of $\res{f}{\KK}$ is $\{ \enc(\A) \mid \A \in \str[\R] \cap \KK\}$, and

### 2016-12-31 LICS/syntactic-clo.tex
sin comentarios

@@
 
-The previous result shows that the classes $\QE{i}$ and $\QU{i}$ are more robust than the classes $\E{i}$ and $\U{i}$: they are closed under binary and sum quantifiers but the other not necessarily.
+Recall that a function $f : \Sigma^* \to \mathbb{N}$ has an easy decision counterpart if $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$ is a language in $\ptime$.
+As the goal of this section is to define a syntactic class of functions in $\shp$ with easy decision versions and good closure properties, we do not directly consider the semantic condition $L_f  \in \ptime$, but instead we consider a more restricted class of functions that has a simple syntactic definition.
+More precisely, a function $f : \Sigma^* \to \mathbb{N}$ is said to be in the complexity class $\totp$ \cite{PagourtzisZ06} if there exists a nondeterministic polynomial-time Turing Machine $M$ such that $f(x) = \tmt_M(x) - 1$ for every $x \in \Sigma^*$, where $\tmt_M(x)$ is the total number of runs of $M$ with input $x$.
+Notice that one is subtracted from $\tmt_M(x)$ to allow for $f(x) = 0$.
+Besides, notice that $\totp \subseteq \shp$ and that $f \in \totp$ implies that $L_f \in \ptime$.
 
-Now, we study the complexity classes describe by this hierarchies.
+The complexity class $\totp$ contains many important counting problems with easy decision counterparts, such as $\cpm$, $\ctwcnf$, $\cdnf$, $\chsat$ and $\csp$.
+Besides, $\totp$ has good closure properties as it is closed under sum, multiplication and subtraction by one.
+However, some functions in $\totp$ do not admit FPRAS under standard complexity-theoretical assumptions,\footnote{As an example consider the problem of counting the number of independent sets in a graph, and the widely believed assumption that $\np$ is not equal to the randomized complexity class $\rp$ (Randomized Polynomial-Time \cite{G77}).
+This counting problem is in $\totp$, and it is known that $\np = \rp$ if there exists an FPRAS for it \cite{DFJ02}.} and no natural complete problems are known for this class \cite{PagourtzisZ06}.
+Hence, we use the $\eqso(\fo)$-hierarchy to find restrictions of $\totp$ with good closure and approximation properties.
+
+%The previous result shows that the classes $\QE{i}$ and $\QU{i}$ are more robust than the classes $\E{i}$ and $\U{i}$: they are closed under binary and sum quantifiers but the other not necessarily.
+
+It was proved in \cite{SalujaST95} that every function in $\E{1}$ admits an FPRAS.
+Besides, it can be proved that $\E{1} \subseteq \totp$.
+However, this class is not closed under sum as shown in the following proposition.
+\begin{proposition}\label{prop-e1-nc}
+There exist functions $f, g \in \E{1}$ such that $(f + g) \not\in \E{1}$.
+\end{proposition}
+To overcome this limitation, one can consider the class $\QE{1}$, which is closed by under sum by definition.
+In fact, the following proposition shows that the same good properties as for $\E{1}$ hold for $\QE{1}$.
+\begin{proposition} \label{prop:qe0-fp-qe1-totp-fptras}
+$\QE{1} \subseteq \totp$ and every function in $\QE{1}$ has an FPRAS.
+\end{proposition}
+Moreover, from the following theorem one can conclude that $\QE{1}$ is closed under multiplication.
+\begin{theorem}\label{theo:binary-prod}
+	If $\LL$ is closed under conjunction, then $\eqso(\LL)$ is closed under multiplication.
+\end{theorem}
+Hence, it only remains to prove that $\QE{1}$ is closed under subtraction by one.
+Unfortunately, it is not clear whether this property holds; in fact, we conjecture that it is not the case.
+Thus, we need to find an extension of $\QE{1}$ that keeps all the previous properties and is closed under subtraction by one.
+It is important to notice that $\shp$ is believed not to be closed under subtraction by one by some complexity-theoretical assumption,\footnote{A decision problem $L$ is in the randomized complexity class $\cspp$ if there exists a nondeterministic polynomial-time Turing Machine $M$ such that for every $x \in L$ it holds that $\tma_M(x) - \tmr_M(x) = 2$, and for every $x \not\in L$ it holds that $\tma_M(x) = \tmr_M(x)$ \cite{OH93,FFK94}.
+It is believed that $\np \not\subseteq \cspp$, as there is no evidence that a Turing Machine with the property just described exists for the propositional satisfiability problem.
+However, if $\shp$ is closed under subtraction by one, then it holds that $\np \subseteq \cspp$ \cite{OH93}.} so the following proposition rules out any logic that extends $\Pi_1$.
+\begin{proposition}
+If $\Pi_1 \subseteq \LL$ and $\eqso(\LL)$ is closed under subtraction by one, then $\shp$ is closed under subtraction by one. 
+\end{proposition}
+Therefore, the desired extension has to be achieved by allowing some local extensions to $\Sigma_1$.
+More precisely, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulas over a relational signature $\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mentioned any second-order variable).
+With this extension we have that:
+\begin{theorem}\label{sigmafo-minusone}
+The class $\eqso(\logex{1})$ is closed under sum, product and subtraction by one. Moreover, $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
+\end{theorem}
+
+A natural question at this point is whether the property of being closed under subtraction by one could be generalized to proper subtraction.
+More precisely, a function complexity class $\CC$ is said to be closed under subtraction if for every $f,g \in \CC$, it holds that $(f \dotdiv g) \in \CC$, where $(f \dotdiv g)(x)$ is defined as $f(x) - g(x)$ if $f(x) \geq g(x)$, and as $0$ otherwise.
+The following theorem shows that such generalization does not work for any of the function complexity classes considered in this section.
+\begin{theorem} \label{sub-pnp}
+If any of the classes $\E{1}$, $\QE{1}$ or $\eqso(\logex{1})$ is closed under subtraction, then $\ptime = \np$.
+\end{theorem}
+
+\marcelo{La version anterior de este teorema decia lo siguiente: If $\eqso(\loge{i})$ or $\eqso(\logu{i})$ is closed under subtraction for $i > 0$, then $\ptime = \np$.
+La razon para considerar esta version mas general era un comentario de Cristian que decia que el resultado de Martin en el apendice se generaliza trivialmente para todas las clases ya que todas contienen la clase $\eqso(\loge{0})$.
+Me parece que esto es incorrecto ya que tambien se necesita que cada funcion en la clase tenga un problema de decision asociado en $\ptime$, lo que no es cierto para $\eqso(\logu{1})$ y las clases de mas arriba.
+Estas clases no deberian ser cerradas bajo resta pero por una razon distinta (por la misma razon por la que se cree que $\shp$ no es cerrado bajo resta), pero esto no me parece que valga la pena mencionarlo.}
+
+%
+%The next question is whether the hierarchy is closed under subtraction.
+Formally, for any pair of functions $f,g$, we define $f - g$ as the function such that $(f - g)(\A) = f(\A)-g(\A)$ whenever $f(\A)>g(\A)$ and $0$ otherwise.
+%As the next result shows, all classes in the $\eqso$-hierarchy is not closed under subtraction unless ${\sc P} = {\sc NP}$
+%\begin{theorem} \label{sub-pnp}
+%If $\eqso(\loge{i})$ or $\eqso(\logu{i})$ is closed under subtraction for $i > 0$, then {\sc P} = {\sc NP}.
+%\end{theorem}
+%\cristian{Martin, el resultado que tienes en el apendice se generaliza trivialmente para todas las clases ya que todas contienen la clase $\eqso(\loge{0})$.}
+
+%Now, we study the complexity classes describe by this hierarchies.
 As the following result shows, $\eqso(\loge{0})$ defines only tractable counting functions and $\eqso(\loge{1})$ intractable counting functions but with an tractable decision problems.
-\begin{proposition} \label{prop:qe0-fp-qe1-totp-fptras}
-All functions defined in $\eqso(\loge{0})$ and $\eqso(\loge{1})$ can be computed in $\fp$ and $\totp$, respectively. Furthermore, every function defined in $\eqso(\loge{1})$ has a FPTRAS.
-\end{proposition}
-Therefore, in terms of counting complexity, the $\eqso$-hierarchy behaves exactly the same as the $\#\fo$-hierarchy.
+%\begin{proposition} \label{prop:qe0-fp-qe1-totp-fptras}
+%All functions defined in $\eqso(\loge{0})$ and $\eqso(\loge{1})$ can be computed in $\fp$ and $\totp$, respectively. Furthermore, every function defined in $\eqso(\loge{1})$ has a FPTRAS.
+%\end{proposition}
+%Therefore, in terms of counting complexity, the $\eqso$-hierarchy behaves exactly the same as the $\#\fo$-hierarchy.
 
-The next step is to study the closure properties of $\eqso$-hierarchy. 
-An advantage of the $\eqso$-hierarchy is that, by its language syntax, all the classes are closed under addition and first and second order sum.
-So, the first question is whether the multiplicative operators in $\qso$ can be defined in $\eqso(\LL)$.
+%The next step is to study the closure properties of $\eqso$-hierarchy. 
+%An advantage of the $\eqso$-hierarchy is that, by its language syntax, all the classes are closed under addition and first and second order sum.
+%So, the first question is whether the multiplicative operators in $\qso$ can be defined in $\eqso(\LL)$.
 As the following result shows, if $\LL$ is closed under conjunction, then the binary product can be defined in  $\eqso(\LL)$.
-\begin{theorem}\label{theo:binary-prod}
-	If $\LL$ is closed under conjunction, then binary product can be defined in $\eqso(\LL)$.
-\end{theorem}
-The next question is whether the hierarchy is closed under subtraction.
-Formally, for any pair of functions $f,g$, we define $f - g$ as the function such that $(f - g)(\A) = f(\A)-g(\A)$ whenever $f(\A)>g(\A)$ and $0$ otherwise.
-As the next result shows, all classes in the $\eqso$-hierarchy is not closed under subtraction unless ${\sc P} = {\sc NP}$
-\begin{theorem} \label{sub-pnp}
-If $\eqso(\loge{i})$ or $\eqso(\logu{i})$ is closed under subtraction for $i > 0$, then {\sc P} = {\sc NP}.
-\end{theorem}
-\cristian{Martin, el resultado que tienes en el apendice se generaliza trivialmente para todas las clases ya que todas contienen la clase $\eqso(\loge{0})$.}
+%\begin{theorem}\label{theo:binary-prod}
+%	If $\LL$ is closed under conjunction, then binary product can be defined in $\eqso(\LL)$.
+%\end{theorem}
 
-By the previous result, we know that functions in the $\eqso$ hierarchy are unlikely to be closed under subtraction.
+%By the previous result, we know that functions in the $\eqso$ hierarchy are unlikely to be closed under subtraction.
 Then, a natural restriction to this question is to ask whether these classes are closed under subtraction by one, namely, if $\CC$ is a class of functions and $f \in \CC$, is $f-1 \in \CC$ where $1$ is the constant function that outputs $1$ for every structure.
-We do not know $\E{1}$ is closed under subtraction by one. However, if we extend $\logex{1}$ with $\fo$ predicates we can show that this new fragment is closed under subtraction by one.
-\begin{theorem} \label{sigmafo-minusone}
-	$\eqso(\logex{1})$ is closed under substraction by one.
-\end{theorem}
+%We do not know $\E{1}$ is closed under subtraction by one. However, if we extend $\logex{1}$ with $\fo$ predicates we can show that this new fragment is closed under subtraction by one.
+%\begin{theorem} \label{sigmafo-minusone}
+%	$\eqso(\logex{1})$ is closed under substraction by one.
+%\end{theorem}

### 2016-12-31 LICS/syntactic.tex
sin comentarios

@@
 More specifically, it was shown in  \cite{Valiant79}  that this problem is $\shp$-complete.
-As a consequence of this result the $\shp$-completeness of the problem of computing the number of perfect matchings of a bipartite graph was established.
-Since then  many counting problems have been proved to be $\shp$-complete \cite{V79b,PB83,P86,L86,BW91,HMRS98,BW05,DS12, PS13,PS14}.
+As a consequence of this result the $\shp$-completeness of the problem of computing the number of perfect matchings of a bipartite graph ($\cpm$) was established.
+Since then  many counting problems have been proved to be $\shp$-complete \cite{V79b,PB83,P86,L86,BW91,HMRS98,DG00,BW05,DS12, PS13,PS14}.
 Among them, problems having easy decision counterparts play a fundamental role, as a counting problem with a hard decision version is expected to be hard.
 Formally, the decision problem associated to a function $f : \Sigma^* \to \mathbb{N}$ is defined as $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$, and $f$ is said to have an easy decision version if $L_f \in \ptime$.
-A first prominent example of such functions is counting the number of perfect matching of a bipartite graph, as it is well-known that the problem of verifying whether there exists a perfect matching in a bipartite graph can be solved in polynomial time.
-Other prominent examples of such problems include counting the number of: satisfying assignments of a 2-CNF propositional formula \cite{V79b}, satisfying assignments of a DNF propositional formula \cite{DHK05}, simple paths from a source node to a target node in a directed graph \cite{V79b}, extensions of a partial order to a linear order \cite{BW91} and Eulerian cycles in an undirected graph \cite{BW05}.
+A first prominent example of such functions is $\cpm$, as it is well-known that the problem of verifying whether there exists a perfect matching in a bipartite graph can be solved in polynomial time.
+Other prominent examples of such problems include counting the number of: satisfying assignments of a 2-CNF propositional formula ($\ctwcnf$) \cite{V79b}, satisfying assignments of a DNF propositional formula ($\cdnf$) \cite{DHK05}, satisfying assignments of a Horn  propositional formula ($\chsat$) \cite{V79b}, simple paths from a source node to a target node in a directed graph ($\csp$) \cite{V79b}, extensions of a partial order to a linear order \cite{BW91} and Eulerian cycles in an undirected graph \cite{BW05}.
 
@@
 Notably, there exist $\shp$-complete functions that can be efficiently approximated as they admit FPRAS; for instance, there exist FPRAS for the problems of counting the number of satisfying assignments of a DNF propositional formula \cite{KL83} and the number of perfect matchings of a bipartite graph \cite{JSV04}.
-A key observation here is that if a $\shp$-complete function admits an FPRAS, then its associated decision problem is in the complexity class $\bpp$ (Bounded-Error Probabilistic Polynomial-Time \cite{G77}).
-Hence, under standard complexity-theoretical assumptions, we cannot hope for an FPRAS for a function in $\shp$ whose decision counterpart is $\np$-complete, and we have to concentrate on the class of counting problems with easy decision versions (in $\bpp$ or in a lower complexity class such as $\ptime$).
+A key observation here is that if a $\shp$-complete function admits an FPRAS, then its associated decision problem is in the randomized complexity class $\bpp$ (Bounded-Error Probabilistic Polynomial-Time \cite{G77}).
+Hence, under the widely believed assumption that  $\np \not\subseteq \bpp$, we cannot hope for an FPRAS for a function in $\shp$ whose decision counterpart is $\np$-complete, and we have to concentrate on the class of counting problems with easy decision versions (in $\bpp$ or in a lower complexity class such as $\ptime$).
 
@@
 Moreover, if such an algorithm for constructing witnesses exists, then we also expect to be able to manipulate such witnesses and in some cases to remove them.
-In other words, we expect a robust class $\CC$ of counting functions with easy decision versions to be closed under subtraction by one, that is, if $g \in \CC$, then the function $g - 1$ should also be in $\CC$.
-Notice that, unless $\ptime = \np$, no such class can contain the function $\textit{sat\_one}$ because $\textit{sat\_one} - 1$ counts the number of satisfying assignments of a propositional formula.
+In other words, we expect a robust class $\CC$ of counting functions with easy decision versions to be closed under subtraction by one, that is, if $g \in \CC$, then the function $g \dotdiv 1$ should also be in $\CC$, where $(g \dotdiv 1)(x)$ is defined as $g(x) - 1$ if $g(x) \geq 1$, and as $0$ otherwise.
+Notice that, unless $\ptime = \np$, no such class can contain the function $\textit{sat\_one}$ because $\textit{sat\_one} \dotdiv 1$ counts the number of satisfying assignments of a propositional formula.
 

### 2016-12-31 LICS/syntactic-horn.tex
mas cammbios
sin comentarios

--- 2016-12-28 LICS/horn-counting.tex
@@
 
-A positive literal is a formula of the form $X(\x)$, where $X$ is a second-order variable and $\x$ is a tuple of first-order variables, and a negative literal is a formula of the form $\exists \v \, \neg X(\u,\v)$, where $\u$ and $\v$ are tuples of first-order variables.
-Given a relational signature $\R$, a clause over $\R$ is a formula of the form:
-$$
-\forall \x \, (\varphi_1 \vee \cdots \vee \varphi_n),
-$$
-where each $\varphi_i$ ($1 \leq i \leq n$) is either a positive literal, a negative literal or an \fo-formula over $\R$.
-A clause is said to be Horn if it contains at most one positive literal, and a formula is said to be Horn if it is a conjunction of Horn clauses over a relational signature $\R$.
-With this terminology, we define $\uhorn$ as the set of formulas $\psi$ such that $\psi$ is a conjunction of Horn clauses over a relational signature $\R$.
+The goal of this section is to define a class of functions in $\shp$ with easy decision counterparts and natural complete problems.
+To this end, we consider the notion of parsimonious reduction.
+Formally, a function $f : \Sigma^* \to \N$ is parsimoniously reducible to a function $g : \Sigma^* \to \N$ if there exists a function $h: \Sigma^* \to \Sigma^*$ such that $h$ is computable in polynomial time and $f(x) = g(h(x))$ for every $x \in \Sigma^*$.
+As mentioned at the beginning of this section, if $f$ can be parsimoniously reduced to $g$, then $L_g \in \ptime$ implies that $L_f \in \ptime$ and the existence of an FPRAS for $g$ implies the existence of an FPRAS for $f$.
 
-\begin{proposition}\label{prop-uhorn-pe}
-$\eqso(\uhorn) \subseteq \pe$
-\end{proposition}
-
+In the previous section, we show that the class $\eqso(\logex{1})$ has good closure and approximation properties.
+Unfortunately, it is not clear whether it admits a {\em natural} complete problem under parsimonious reductions, where {\em natural} means any of the counting problems defined in this section or any other well-known counting problem (not one specifically designed to be complete for the class).
+Hence, in this section we follow a different approach to find a class of functions in $\shp$ with easy decision counterparts and natural complete problems, which is inspired by the approach followed in \cite{G92} that uses a restriction of second-order logic to Horn clauses for capturing $\ptime$ (over ordered structures).
+The following example shows how our approach works.
 \begin{example} \label{ex-hornsat-esop1}
@@
 
-To encode $\chsat$, we define an \so-formula $\varphi(\pT)$ over $\R$, where $\pT$ is a unary predicate, such that for every Horn formula $\theta$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \varphi(\pT)}(\A)$.
-In particular, $\pT(x)$ holds if and only if $x$ is a propositional variable that is assigned value 1.
+To define $\chsat$, we consider an \so-formula $\varphi(\pT)$ over $\R$, where $\pT$ is a unary predicate, such that for every Horn formula $\theta$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \varphi(\pT)}(\A)$.
+In particular, $\pT(x)$ holds if and only if $x$ is a propositional variable that is assigned value true.
 More specifically, $\varphi(\pT)$ is defined as follows:
@@
 \end{align*}
-Given that $\uhorn$ is designed with the goal in mind of capturing $\chsat$, we expect $\varphi(\pT)$ to be a formula in $\uhorn$.
+%Given that $\uhorn$ is designed with the goal in mind of capturing $\chsat$, we expect $\varphi(\pT)$ to be a formula in $\uhorn$.
 However, if we rewrite it as a conjunction of clauses we obtain the following:
+We can rewrite $\varphi(\pT)$ in the following way:
 \begin{align*}
@@
 \end{align*}
-The resulting formula $\varphi(\pT)$ is not in $\uhorn$, but it can be easily transformed into a formula in this class  by introducing an auxiliary binary predicate $\pA$ defined as follows:
+%The resulting formula $\varphi(\pT)$ is not in $\uhorn$, but it can be easily transformed into a formula in this class  by introducing an auxiliary binary predicate $\pA$ defined as follows:
+Moreover, by introducing an auxiliary predicate $\pA$ defined as:
 \begin{align*}
-\forall c \forall x \, (\neg \pA(c,x) \leftrightarrow [\pN(c,x) \wedge \neg \pT(x)]).
+\forall c \forall x \, (\neg \pA(c,x) \leftrightarrow [\pN(c,x) \wedge \neg \pT(x)]),
 \end{align*}
-In this way, we obtain the following formula $\psi(\pT,\pA)$ in $\uhorn$:
+we can translate $\varphi(\pT)$ into the following equivalent formula $\psi(\pT,\pA)$:
 \begin{align*}
@@
 \end{align*}
-This formula effectively defines $\chsat$
-as for every Horn formula $\theta$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
+More precisely, we have that:
+\begin{eqnarray*}
+\sem{\sa{\pT} \varphi(\pT)}(\A) &=& \sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A),
+\end{eqnarray*}
+ for every $\R$-structure $\A$ encoding a Horn formula.
+Therefore, the formula $\psi(\pT,\pA)$ also defines $\chsat$.
+More importantly, $\psi(\pT,\pA)$ resembles a conjunction of Horn clauses except for the use of negative literals of the form $\exists v \, \neg \textit{A}(u,v)$.
+%This formula effectively defines $\chsat$
+%as for every Horn formula $\theta$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
 Therefore, we conclude that $\chsat \in \eqso(\uhorn)$.
 \end{example}
-We extend the definition of $\uhorn$ to allow existential quantification.
+The previous example suggests that in order to define $\chsat$, we can use Horn formulas defined as follows. 
+A positive literal is a formula of the form $X(\x)$, where $X$ is a second-order variable and $\x$ is a tuple of first-order variables, and a negative literal is a formula of the form $\exists \v \, \neg X(\u,\v)$, where $\u$ and $\v$ are tuples of first-order variables.
+Given a relational signature $\R$, a clause over $\R$ is a formula of the form $\forall \x \, (\varphi_1 \vee \cdots \vee \varphi_n)$,
+where each $\varphi_i$ ($1 \leq i \leq n$) is either a positive literal, a negative literal or an \fo-formula over $\R$.
+A clause is said to be Horn if it contains at most one positive literal, and a formula is said to be Horn if it is a conjunction of Horn clauses over a relational signature $\R$.
+With this terminology, we define $\uhorn$ as the set of formulas $\psi$ such that $\psi$ is a conjunction of Horn clauses over a relational signature $\R$.
+
+We have that $\chsat \in \eqso(\uhorn)$. Moreover, the following proposition shows that $\eqso(\uhorn)$ forms a class of functions with easy decision counterparts.
+
+\begin{proposition}\label{prop:uhorn-pe}
+$\eqso(\uhorn) \subseteq \totp$
+\end{proposition}
+
+Thus, $\eqso(\uhorn)$ is a new alternative in our search for a class of functions in $\shp$ with easy decision counterparts and natural complete problems.
+Moreover, an even larger class for our search can be generated by extending the definition of $\uhorn$ with outermost existential quantification.
 More precisely, a formula $\varphi$ is in $\ehorn$ if $\varphi$ is of the form $\exists \bar x \, \psi$ with $\psi$ a Horn formula.
-Interestingly, it hold that $\cdnf \in \eqso(\ehorn)$ and
 
-\begin{proposition}\label{prop-ehorn-pe}
-$\eqso(\ehorn) \subseteq \pe$.
+\begin{proposition}\label{prop:ehorn-pe}
+$\eqso(\ehorn) \subseteq \totp$.
 \end{proposition}
-A natural question at this point is whether in the definitions of $\uhorn$ and $\ehorn$, it is necessary to allow negative literals of the form $\exists \v \, \neg X(\u,\v)$.
+Interestingly, we have that both $\chsat$ and $\cdnf$ belong to $\eqso(\ehorn)$. 
+An imperative question at this point is whether in the definitions of $\uhorn$ and $\ehorn$, it is necessary to allow negative literals of the form $\exists \v \, \neg X(\u,\v)$.
 The following result shows that it is indeed the case:
 
-\begin{proposition}\label{prop-hsat-not-sigma2}
+\begin{proposition}\label{prop:hsat-not-sigma2}	
 $\chsat \not\in \eqso(\loge{2})$.
 \end{proposition}
-We conclude this section by showing that a natural extension of $\chsat$ is $\eqso(\ehorn)$-complete under parsimonious reductions. We define the decision problem:
-\begin{multline*}
-\dhsat = \{\Phi \mid \Phi \text{ is a disjunction of}\\  \text{Horn formulas and $\Phi$ is satisfiable}\},
+
+We conclude this section by showing that $\eqso(\ehorn)$ is the class we were looking for, as not only every function in $\eqso(\ehorn)$ has an easy decision counterpart, but also $\eqso(\ehorn)$ admits a natural complete problem under parsimonious reductions.
+More precisely, define $\dhsat$ as the decision problem:\begin{multline*}
+\{\Phi \mid \Phi \text{ is a disjunction of}\\  \text{Horn formulas and $\Phi$ is satisfiable}\},
 \end{multline*}
-and the counting problem $\shdhsat$ as a function that counts all satisfying assignments of a formula $\Phi$ that is a disjunction of Horn formulas.
+and define $\shdhsat$ as the problem of counting the satisfying assignments of a formula $\Phi$ that is a disjunction of Horn formulas. Then we have that:
 

### 2016-12-31 LICS/appendix-syntactic.tex
30 dic
31 dic merge

@@
 Classes QE0, QE1, QU1, QE2, QU2 = QFO form a properly contained hierarchy:\vspace{1em}
+
+We give this proof in various parts.
+
+We show here that $\QE{0} \not\subseteq \E{1}$..
+By contradiction, suppose that there is a $\QE{0}$ formula $\alpha$ over some signature $\R$ such that defines the following function.
+For every finite $\R$-structure with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n - 1$.
+We use the following claim.
+\begin{claim}
+Let $\alpha = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. Then the function defined by $\alpha$ is either null, greater or equal to $n$, or is in $\Omega(n^2)$.
+\end{claim}
+\begin{proof}
+Suppose that the function defined by $\alpha$ is not $0$ and that $\varphi$ is in DNF.
+Furthermore, suppose $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$.
+Then $\alpha = \sa{\bar{x}} \varphi_1(\bar{x}) \vee \cdots \vee \varphi_n(\bar{x})$.
+Since $\alpha$ is not null, then some $\varphi_i$ must be satisfiable.
+This is, the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is not null.
+We will prove by induction on $\length(\bar{x})$ that the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is either greater or equal to $n$, or in $\Omega(n^2)$.
+We address the case $\length(\bar{x})= 1$, then $\alpha = \sa{x}\bigwedge\psi(x)$.
+If any $\psi(x) = (x = x)$ or $\neg(x < x)$, then we can eliminate it and we obtain the same function.
+If any $\psi = (x < x)$ or $\neg(x=x)$, then the function becomes null.
+If $\psi(x) = R(x,\ldots,x)$ for some $R\in\R$ the function becomes null for the structures we are considering.
+If $\psi(x) = \neg R(x,\ldots,x)$, we can eliminate it and for the structures we are considering we obtain the same function.
+The only possible $\alpha$ left is $\alpha = \sa{x}\top$ which is equal to the function $n$.
+This covers all possible cases for $\length(\bar{x}) = 1$.
+Now suppose that it holds for $\length(\bar{x}) = k$ and suppose $\alpha = \sa{\bar{x}}\bigwedge\psi(\bar{x})$ for $\length(\bar{x}) = k+1$.
+If any $\psi(\bar{x}) = (x_i = x_j)$ where $i \neq j$, then $\alpha$ describes the same function as $\alpha$ where $x_j$ has been replaced by $x_i$.
+In this formula the tuple of first-order variables has $k$ elements so the function it describes if one of the mentioned in the hypothesis.
+If $i = j$, then we can eliminate it and obtain the same function.
+If any $\psi(\bar{x}) = R(\bar{v})$ or $\neg R(\bar{v})$ where $\bar{v}$ is a sub-tuple of $\bar{x}$ then we can either eliminate it or the function becomes null, following the same argument as in the case $\length(\bar{x}) = 1$.
+If any $\psi(\bar{x}) = \neg(x_i = x_j)$ or $(x_i < x_j)$ where $i = j$, then the function becomes null.
+If any $\psi(\bar{x}) = \neg(x_i < x_j)$ where $i = j$, we can eliminate it.
+The remaining formulas in $\bigwedge\psi(\bar{x})$ are either $\neg(x_i = x_j)$, $(x_i<x_j)$ or $\neg(x_i<x_j)$.
+If the formula violates transitivity in $<$ (for example, $x < y \wedge y < z \wedge z < x$), then the function $\alpha$ describes is null.
+Therefore, there is some order over $\bar{x}$ that satisfies $\bigwedge\psi(\bar{x})$.
+Consider the formula that describes this order (like $x_1 < x_3 \wedge x_3 < x_4 \wedge x_4 < x_2$).
+The function $\alpha$ describes is greater or equal to the one this formula describes, which is exactly $\binom{n}{\length(\bar{x})}$ which is in $\Omega(n^{\length(\bar{x})}) \subseteq \Omega(n^2)$ if $\length(\bar{x}) > 1$.
+This concludes the proof of the claim.
+\end{proof}
+We suppose that $\alpha$ is in SNF, this is, $\alpha = c + \sum_{i = 1}^n\alpha_i$.
+Let $\mathbb{1}$ be a $\R$-structure with one element.
+Note that if $c > 0$, then $\alpha(\enc(\mathbb{1})) \geq c > 0$ which is not possible since $\alpha(\enc(\mathbb{1})) = 0$.
+Since $\alpha$ is not null, consider some $\alpha_i$ that describes a non-null function.
+Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, where $\varphi$ is quantifier-free.
+Note that if $\length(\bar{X}) > 0$, then the function $\alpha$ describes is in $\Omega(2^n)$, as it was proven by the authors in \cite{SalujaST95}.
+We have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$, as we proved in the claim, describes either some function greater or equal to $n$, or in $\Omega(n^2)$, which leads to a contradiction.
+Lastly, note that the formula $\sa{x}\exists y(x < y)$ is in $\E{0}$ and describes the function $n-1$, which concludes the proof.
+
+\vspace{1em}
+We show here that $\E{1}\not\subseteq\QE{0}$.
+By contradiction, suppose that the constant 1 is expressible in $\E{1}$.
+Let $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ describe this function for some quantifier-free formula $\varphi$.
 

### 2016-12-31 LICS/syntactic-horn.tex
31 dic merge

@@
 \end{proposition}
-
 We conclude this section by showing that $\eqso(\ehorn)$ is the class we were looking for, as not only every function in $\eqso(\ehorn)$ has an easy decision counterpart, but also $\eqso(\ehorn)$ admits a natural complete problem under parsimonious reductions.
-More precisely, define $\dhsat$ as the decision problem:\begin{multline*}
+More precisely, define $\dhsat$ as the decision problem:
+\begin{multline*}
 \{\Phi \mid \Phi \text{ is a disjunction of}\\  \text{Horn formulas and $\Phi$ is satisfiable}\},

### 2016-12-31 QSO/cap-comp.tex
31 dic merge

@@
 
-	We represent configurations with a tuple of fixed size $k$.
+	We represent configurations with a tuple of fixed size.
 The formula $\varphi(\bar{x},\bar{y})$ describes a procedure that given a configuration generates a possible next configuration.
-The formula $\varphi_I(\bar{x})$ models that $\bar{x}$ is the initial configuration of $M(w)$.
-The formula $\varphi_F(\bar{x})$ models that $\bar{x}$ is an accepting (final) configuration of $M(w)$.
+The formula $\varphi_I(\bar{x})$ describes that $\bar{x}$ is the initial configuration of $M(w)$.
+The formula $\varphi_F(\bar{x})$ describes that $\bar{x}$ is an accepting (final) configuration of $M(w)$.
 The formula we construct is:
@@
 
-	To illustrate our idea, we will show a greatly simplified example.
+	To illustrate our idea, we will show a simplified example.
 Consider a machine $M$ that works in exactly $\log_2(n)$ space and only allows 0 or 1 in the working tape.
@@
 	The problem that arises from this representation, is that to describe a transition in $M$ we need to read an arbitrary character in the working tape, and therefore we have to obtain the $a_3$-th bit in $a_4$.
-Furthermore, to get the following configuration, we need compute $a_4$ with the $a_3$-th bit flipped.
-This is generally not possible to describe with an $FO$ formula.
+Furthermore, to represent the following configuration, we need compute $a_4$ with the $a_3$-th bit flipped.
+This is generally not possible to describe with an $\fo$ formula.
 Consider the following procedure that receives $x = a_4$ and $i = a_3$.
 
-	\begin{algorithm} \label{switch1to0}
+	\begin{algorithm}
 	\caption{If the $i$-th bit in $x$ is 1 replace it by 0 and return the result}
+	\label{switch1to0}
 	\begin{algorithmic}
@@
 	\end{multline*}
-	Note that we also need to specify that the program we are following is Algorithm \ref{switch1to0} so we store $1,0$ in $b_1,b_2$ all along the procedure.
-We now describe the three other algorithms that compute the switches from $0\to 0$, $0\to 1$ and $1\to 1$.
-	\begin{algorithm} \label{switch0to0}
-		\caption{If the $i$-th bit in $x$ is 0 return $x$}
+
+	\begin{algorithm}
+		\caption{If the $i$-th bit in $x$ is 0 return $x$} \label{switch0to0}
 		\begin{algorithmic}
@@
 	\end{algorithm}
-	\begin{algorithm} \label{switch0to1}
+
+	\begin{algorithm}
 		\caption{If the $i$-th bit in $x$ is 0 replace it by 1 and return the result}
+		\label{switch0to1}
 		\begin{algorithmic}
@@
 	\end{algorithm}
-	\begin{algorithm} \label{switch1to1}
+
+	\begin{algorithm}
 		\caption{If the $i$-th bit in $x$ is 1 return $x$}
+		\label{switch1to1}
 		\begin{algorithmic}
@@
 	\end{algorithm}
+	Note that we also need to specify that the program we are following is Algorithm \ref{switch1to0} so we store $1,0$ in $b_1,b_2$ all along the procedure.
+We describe the three other algorithms that compute the switches from $0\to 0$, $0\to 1$ and $1\to 1$ (Algorithms \ref{switch0to0}, \ref{switch0to1} and \ref{switch1to1}.)
 	For the other three cases, where $\delta = (q_i,a,0,q_j,op_1,0,op_2)$, $\delta = (q_i,a,0,q_j,op_1,1,op_2)$ and $\delta = (q_i,a,1,q_j,op_1,1,op_2)$, $\varphi_{\delta}(\bar{x},\bar{s},\bar{w},\bar{u},\bar{y},\bar{t},\bar{z},\bar{v})$ is defined analogously.
@@
 
-%\begin{theorem}
-%$\eqso(\integ,\fo)$ captures $\gp$ over the class of ordered structures.
-%\end{theorem}
+\begin{theorem}
+$\eqso(\integ,\fo)$ captures $\gp$ over the class of ordered structures.
+\end{theorem}
 

### 2017-01-02 LICS/beyond.tex
sin comentarios

@@
 
-We define the operator {\bf path} as follows.
+It was shown in \cite{I86,I88} that first-order logic extended with a transitive closure operator captures $\nlog$.
+Inspired by this work, we extend the definition of $\qfo$ with an operator for counting the number of paths in a directed graph, and we show that it can be used to capture $\shl$.
+Besides, we show that the same idea can be used to extend $\qso$ allowing to capture harder complexity classes.
+
+Given a relation signature $\R$, the set of transitive $\qso$ formulas ($\tqso$-formulas) is defined by the following grammar:
+\begin{multline}
+	\label{eq-def-tqso}
+	\alpha := \varphi \, \mid \, s \, \mid \, (\alpha \add \alpha) \, \mid\, (\alpha \mult \alpha) \, \mid \, 
+	\sa{x} \alpha \, \mid\, \
+	\pa{x} \alpha \, \mid \\ 
+	\sa{X} \alpha \, \mid \, \pa{X} \alpha \, \mid \, [\pth \psi(\bar{x}, \bar{X},\bar{y}, \bar{Y})],
+\end{multline}
+where $\varphi$ is an $\so$-formula over $\R$, $\psi(\bar{x},\bar{X},\bar{y},\bar{Y})$ is an $\fo$-formula over $\R$, $\bar{x} = (x_1, \ldots, x_k)$, $\bar{y} = (y_1, \ldots, y_k)$ are tuples of pairwise distinct first-order variables, and $\bar{X} = (X_1, \ldots, X_\ell)$, $\bar{Y} = (Y_1, \ldots, Y_\ell)$ are tuples of pairwise distinct second-order variables such that $\arity(X_i) = \arity(Y_i) = m_i$ for every $i \in \{1, \ldots, \ell\}$.
+The semantics of $[\pth \psi(\bar{x},\bar{X},\bar{y}.\bar{Y})]$ is defined as follows.
+Given an $\R$-structure $\A$, define a (directed) graph $\cG_{\psi}(\A) = (N,E)$ such that $N = A^k \times \prod_{i=1}^\ell 2^{A^{m_i}}$, where $A$ is the domain of $\A$, and for every pair $(\bar b, \bar B)$, $(\bar c, \bar C)$ of elements of $N$, it holds that $((\bar b, \bar B), (\bar c, \bar C)) \in E$ if and only if $\A \models \psi(\bar b, \bar{B}, \bar c, \bar{C})$.
+Then given first-order and second-order assignments $v$, $V$ for $\A$, we have that $\sem{[\pth \psi(\bar{x},\bar{X}, \bar{y}, \bar{Y})]}(\A,v,V)$ is the number of paths in $\cG_\psi(\A)$ from $(v(\bar x), V(\bar X))$ to $(v(\bar y), V(\bar Y))$ whose length is at most $n = |N|$.
+
+As for the case of $\qso$, the logic $\tqso(\LL)$ is obtained by restricting $\varphi$ in \eqref{eq-def-tqso} to be a formula in $\LL$.
+Moreover, the logic $\tqfo$ is obtained by disallowing in \eqref{eq-def-tqso} formulas $\sa{X} \alpha$ and $\pa{X} \alpha$, and by only allowing  first-order free-variables in the formula $\psi$ used in $[\pth \psi]$ in \eqref{eq-def-tqso}.
+With this notation, we have the following results:
+
+\begin{theorem} \label{tqfo-shl}
+	$\tqfo(\fo)$ captures $\shl$ over the class of ordered structures.
+\end{theorem}
+
+\begin{theorem} \label{tqso-fo-fpsace}
+	$\tqso$ and $\tqso(\fo)$ captures $\fpspace$ over the class of ordered structures.
+\end{theorem}
+
+\marcelo{Puede que este equivocado, pero me parece que teniamos una forma de capturar $\shp$ usando el operator ${\bf path}$.
+Pero no logro recordar como se hacia esto, y me parece que lo que habiamos escrito antes en esta seccion estaba equivocado: ``$\tqso(\fo)$ captures $\shp$ over the class of ordered structures".
+Claro que yo puedo estar usando una definicion distinta de $\tqso(\fo)$.}
+
+%We also define the set of transitive $\qso$ formulas ($\tqso$-formulas) using the following grammar:
+%\begin{multline*}
+%%	\label{eq-def-tqso}
+%	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \\ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha \ \mid \ [\pth \varphi]
+%\end{multline*}
+%
+%
+% 
+%We define the operator {\bf path} as follows.
 Let $\A$ be an ordered structure.
@@
 To formalize the semantics for this operator, let $n = \vert A^k \vert$.
-For a given first order assignment $v$ and a second order asssignment $V$, let $\bar{a} = v(\bar{x})$ and $\bar{b} = v(\bar{y})$, and $\sem{[\pth\, \psi(\bar{x},\bar{y})]}(\A,v,V)$ will take the value of the number of paths of size less or equal to $n$ from $\bar{a}$ to $\bar{b}$ in the graph ${\cal G}$.
+%For a given first order assignment $v$ and a second order asssignment $V$, let $\bar{a} = v(\bar{x})$ and $\bar{b} = v(\bar{y})$, and $\sem{[\pth\, \psi(\bar{x},\bar{y})]}(\A,v,V)$ will take the value of the number of paths of size less or equal to $n$ from $\bar{a}$ to $\bar{b}$ in the graph ${\cal G}$.
 This operator lets us define the set of transitive $\qfo$ formulas over $\R$ ($\tqfo$-formulas) using the following grammar:
-\begin{multline*} 
-%	\label{eq-def-tqfo}
-	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \\ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ [\pth \varphi]
-\end{multline*}
-where $\varphi$ is an $\fo$-formula over $\R$, $s \in \bbN$ and $x \in \fv$.
-
-We also define the set of transitive $\qso$ formulas ($\tqso$-formulas) using the following grammar:
-\begin{multline*}
-%	\label{eq-def-tqso}
-	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \\ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha \ \mid \ [\pth \varphi]
-\end{multline*}
-where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
-\begin{theorem} \label{so-rec}
-	Given a positive $\fo$ formula $\varphi(\bar{x},R)$ and a $\qfo$ formula $\alpha(\bar{x})$, there exists a $\qso$ formula $\beta(\bar{x})$ such that $\sem{[\alfp\varphi(\bar{x},R)\mid \alpha(\bar{x},R)](\bar{x})} = \sem{\beta(\bar{x})}$.
-\end{theorem}
-
-\begin{theorem} \label{tqfo-fo-cap}
-	$\tqfo(\fo)$ captures $\shl$ over the class of ordered structures.
-\end{theorem}
-
-\begin{theorem} \label{tqso-fo-cap}
-	$\tqso(\fo)$ captures $\shp$ over the class of ordered structures.
-\end{theorem}
+%\begin{multline*} 
+%%	\label{eq-def-tqfo}
+%	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \\ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ [\pth \varphi]
+%\end{multline*}
+%where $\varphi$ is an $\fo$-formula over $\R$, $s \in \bbN$ and $x \in \fv$.
+%
+%We also define the set of transitive $\qso$ formulas ($\tqso$-formulas) using the following grammar:
+%\begin{multline*}
+%%	\label{eq-def-tqso}
+%	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \\ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha \ \mid \ [\pth \varphi]
+%\end{multline*}
+%where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
+%\begin{theorem} \label{so-rec}
+%	Given a positive $\fo$ formula $\varphi(\bar{x},R)$ and a $\qfo$ formula $\alpha(\bar{x})$, there exists a $\qso$ formula $\beta(\bar{x})$ such that $\sem{[\alfp\varphi(\bar{x},R)\mid \alpha(\bar{x},R)](\bar{x})} = \sem{\beta(\bar{x})}$.
+%\end{theorem}
+%
+%\begin{theorem} \label{tqfo-fo-cap}
+%	$\tqfo(\fo)$ captures $\shl$ over the class of ordered structures.
+%\end{theorem}
+%
+%\begin{theorem} \label{tqso-fo-cap}
+%	$\tqso(\fo)$ captures $\shp$ over the class of ordered structures.
+%\end{theorem}
 
@@
 
-We define an operator which extends the Least Fixed Point logic to counting.
-Recall that a fixed point operator is defined by a formula $\varphi(x_1,\ldots,x_k,R)$ which is positive on $R$ where $R$ is a predicate of arity $k$.
-For a structure $\A$ with domain $A$, the operator $T_{\varphi}:2^{A^k} \to 2^{A^k}$ is defined as $T_{\varphi}(X) = \{(a_1,\ldots,a_k)\mid (\A,X)\models \varphi(a_1,\ldots,a_k,R) \}$, for each $X\subseteq A^k$.
+We define an operator which extends least fixed point logic \cite{I86,vardi1982complexity} to allow counting.
+Recall that the least fixed point operator is defined by a formula $\psi(x_1,\ldots,x_k,R)$ that is positive on $R$, where $R$  is a predicate of arity $k$.
+For a structure $\A$ with domain $A$, the operator $T_{\varphi}:2^{A^k} \to 2^{A^k}$ is defined as $T_{\varphi}(X) = \{(a_1,\ldots,a_k)\mid (\A,X)\models \psi(a_1,\ldots,a_k,R) \}$, for each $X\subseteq A^k$.
 Let $T_0 = \emptyset$ and $T_{i+1} = T_{\varphi}(T_i)$ for each $i \in \nat$.
-Note that there exists $n\in \nat$ such that $T_{n+1} = T_n$.
-Semantically, $[\lfpop\varphi(x_1,\ldots,x_k,R)]$ is defined such that for each $(a_1,\ldots,a_k)\in A^k$, it holds that $\A\models[\lfpop\varphi(x_1,\ldots,x_k,R)](a_1,\ldots,a_k)$ if and only if $\A\models T_n(a_1,\ldots,a_k)$.
+Note that there exists $n\in \nat$ such that $T_{n+1} = T_n$ because $R$ is positive in $\psi(x_1,\ldots,x_k,R)$.
+Then the evaluation of $[\lfpop\psi(x_1,\ldots,x_k,R)]$ is defined by the fixed point $T_n$, that is, for every $(a_1,\ldots,a_k)\in A^k$, it holds that $\A\models[\lfpop\psi(x_1,\ldots,x_k,R)](a_1,\ldots,a_k)$ if and only if $(a_1,\ldots,a_k) \in T_n$.
 
-We extend $\qfo$ with the operator $[\alfp\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$.
-It is defined by a FO-formula $\varphi$ and a formula $\alpha$ in $\qfo$ without the $\Pi$ operator, which mentions a placeholder function $\pi$ with domain $\fv^{\ell}$.
-Let $\{T_i\}_{i\in\nat}$ and $n\in\nat$ be defined from $\varphi$ as it was for the $\lfp$ operator.
+We extend $\qfo$ with the operator $[\alfp\psi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](z_1,\ldots,z_{\ell})$, where $\psi(y_1,\ldots,y_k,R)$ 
+%It is defined by a 
+is an FO-formula 
+%$\varphi$ 
+and $\alpha$  is a $\qso$-formula 
+%in $\qfo$ without the $\Pi$ operator, which 
+that mentions a placeholder $\pi$ for a function of arity $\ell$.
+%with domain $\fv^{\ell}$. 
+To define the semantics of such operator 
+%$[\alfp\psi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$ 
+we need to introduce some terminology.
+Given a $\qso$-formula $\beta$ with $\ell$ first-order free variables and no second-order free variable, $\alpha|_{\pi\to\beta}$ is a $\qso$-formula obtained from $\alpha$ by replacing every occurrence $\pi(u_1, \ldots, u_\ell)$ of the placeholder by $\beta(u_1, \ldots, u_\ell)$.
+%For each $\qfo$ formulas $\alpha, \beta, \gamma$, where $\beta$ and $\gamma$ have $\ell$ and $m$ first-order free variables respectively, $u_1,\ldots,u_{\ell} \in \fv$, and $v_1,\ldots,v_m \in \{u_1,\ldots,u_{\ell}\}$, we define $\alpha\mid_{\beta(u_1,\ldots,u_{\ell})\to\gamma(v_1,\ldots,v_{m})}$ as $\alpha$ where every instance of the subformula $\beta(u_1,\ldots,u_{\ell})$ is replaced by $\gamma(v_1,\ldots,v_{m})$.
+Moreover, for every natural number $i \geq 1$, define a formula $\varphi_i(x)$ such that $\varphi_i(a)$ holds in a structure $\A$ if and only if $a$ is the $i$-th element in the linear order  $<^{\A}$.
+%$a\in A$, suppose that $a$ is the $p$-th element in the order $<^{\A}$. Then if $a$ is the first element, we define $\varphi_a(x) = \forall y(x < y \vee x = y)$, and if it is not we define:
+Thus, we have that $\varphi_1(x) = \forall y(x < y \vee x = y)$ and for every $p > 1$:
+\begin{multline*}
+\varphi_p(x) = \exists x_1 \cdots \exists x_{p-1}[\bigwedge_{1\leq i,j < p}x_i\neq x_j \wedge\,\\ \bigwedge_{i = 1}^{p-1} x_i < x  \wedge \forall y(y < x \to \bigvee_{i = 1}^{p-1} y = x_i)].
+\end{multline*}
+We are now ready to define the semantics of $[\alfp\psi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](z_1,\ldots,z_{\ell})$. 
+Assume that $\A$ is an $\R$-structure with domain $A$, and let $\{T_i\}_{i\in\nat}$ be as in the definition of $[\lfpop\psi(x_1,\ldots,x_k,R)]$.
+Moreover, suppose that $n\in\nat$ is the smallest value such that $T_{n+1} = T_n$. 
+%To formally characterize 
+%We name $\T = \{T_0,\ldots,T_n\}$ and 
+Then we define a sequence of functions $\zeta_0,\ldots,\zeta_n:A^{\ell}\to\nat$ as follows. 
+We first have that $\zeta_0(a_1, \ldots, a_\ell) = 0$ for every $(a_1, \ldots, a_\ell) \in A^\ell$.
+Now assume that function $\zeta_i$ has already been defined, where $0 \leq i < n$.
+By using this function define a formula $\beta_{i}(u_1, \ldots, u_\ell)$ as follows:
+%For each $T_i \in \T$, we define a $\qfo$ formula $\beta_i$ as follows. If $i = 0$, then $\beta_i(u_1,\ldots,u_{\ell}) = 0$. If $i \geq 1$, then
+\begin{multline*}
+%\beta_{i+1}(u_1,\ldots,u_{\ell}) = 
+ \mathop{+}_{(a_1,\ldots,a_{\ell})\in A^{\ell}} \varphi_{\sigma(a_1)}(u_1) \cdot \ldots \cdot \varphi_{\sigma(a_{\ell})}(u_{\ell})\cdot \zeta_{i}(a_1,\ldots,a_{\ell}), 
+\end{multline*}
+where $(u_1, \ldots, u_\ell)$ is a tuple of pairwise distinct first-order variables and $\sigma(\cdot)$ is a function that assigns to every $a \in A$ its position in the lineal order $<^\A$. 
+Then function $\zeta_{i+1}$ is defined as follows.
+Assuming that $(a_1, \ldots, a_\ell) \in A^\ell$, $v$ is a first-order assignment for $\A$ such that $v(z_i) = a_i$ for every $i \in \{1, \ldots, \ell\}$, and that $V$ is a second-order assignment for $\A$ such that $V(R) = T_{i+1}$, we have that:
+\begin{eqnarray*}
+\zeta_{i+1}(a_1, \ldots, a_\ell) & = & \sem{\alpha\mid_{\pi \to \beta}}(\A,v,V).
+\end{eqnarray*}
+%, for each $T_i \in \T$, let $V$ be a second-order assignment for $\A$ that assigns $T_i$ to $R$, let $\zeta_i: A^{\ell}\to\nat$ be such that for each $(a_1,\ldots,a_{\ell})\in A^{\ell}$ it holds $\zeta_i(a_1,\ldots,a_{\ell}) = \sem{\alpha\mid_{\pi(u_1,\ldots,u_{\ell})\to \beta_i(u_1,\ldots,u_{\ell})}}(\A,v,V)$, where $v$ is a first-order assignment for $\A$ that satisfies $a_i = v(x_i)$ for each free $x_i$ in $\alpha$.
+Finally, the semantics of $[\alfp\psi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](z_1,\ldots,z_{\ell})$ is defined by means of function $\zeta_n$:
+%For a given first order assignment $v$ and a second order assignment $V$, let $a_i = v(x_i)$, the operator is then evaluated as:
+\begin{align*}
+&\llbracket [\alfp\psi(y_1,\ldots,y_k,R)\mid\\
+&\hspace{5mm}\alpha(x_1,\ldots,x_{\ell},R,\pi)](z_1,\ldots,z_{\ell}) \rrbracket =
+%\sem{[\alfp\psi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)]}(\A,v,V) = \\ 
+\zeta_n(v(z_1),\ldots,v(z_{\ell})).
+\end{align*}
 
-For each $\qfo$ formulas $\alpha, \beta, \gamma$, where $\beta$ and $\gamma$ have $\ell$ and $m$ first-order free variables respectively, $u_1,\ldots,u_{\ell} \in \fv$, and $v_1,\ldots,v_m \in \{u_1,\ldots,u_{\ell}\}$, we define $\alpha\mid_{\beta(u_1,\ldots,u_{\ell})\to\gamma(v_1,\ldots,v_{m})}$ as $\alpha$ where every instance of the subformula $\beta(u_1,\ldots,u_{\ell})$ is replaced by $\gamma(v_1,\ldots,v_{m})$.
-Moreover, for each $a\in A$, suppose that $a$ is the $p$-th element in the order $<^{\A}$.
-Then if $a$ is the first element, we define $\varphi_a(x) = \forall y(x < y \vee x = y)$, and if it is not we define:
-\begin{multline*}
-\varphi_a(x) = \exists x_1 \cdots \exists x_{p-1}[\bigwedge_{1\leq i,j < p}x_i\neq x_j \wedge\,\\ \bigwedge_{i = 1}^{p-1} x_i < x  \wedge \forall y(y < x \to \bigvee_{i = 1}^{p-1} y = x_i)].
-\end{multline*}
-
-To formally characterize the semantics of $[\alfp\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$. %. 
-We name $\T = \{T_0,\ldots,T_n\}$ and we define a sequence of functions $\zeta_0,\ldots,\zeta_n:A^{\ell}\to\nat$ as follows.
-For each $T_i \in \T$, we define a $\qfo$ formula $\beta_i$ as follows.
-If $i = 0$, then $\beta_i(u_1,\ldots,u_{\ell}) = 0$.
-If $i \geq 1$, then
-\begin{multline*}
-\beta_i(u_1,\ldots,u_{\ell}) = \mathop{+}_{(a_1,\ldots,a_{\ell})\in A^{\ell}} \varphi_{a_1}(u_1)\cdot\varphi_{a_2}(u_2)\,\cdots\,\\ \varphi_{a_{\ell}}(u_{\ell})\cdot \zeta_{i-1}(a_1,\ldots,a_{\ell}).
-\end{multline*}
-Then, for each $T_i \in \T$, let $V$ be a second-order assignment for $\A$ that assigns $T_i$ to $R$, let $\zeta_i: A^{\ell}\to\nat$ be such that for each $(a_1,\ldots,a_{\ell})\in A^{\ell}$ it holds $\zeta_i(a_1,\ldots,a_{\ell}) = \sem{\alpha\mid_{\pi(u_1,\ldots,u_{\ell})\to \beta_i(u_1,\ldots,u_{\ell})}}(\A,v,V)$, where $v$ is a first-order assignment for $\A$ that satisfies $a_i = v(x_i)$ for each free $x_i$ in $\alpha$.
-
-For a given first order asignment $v$ and a second order assignment $V$, let $a_i = v(x_i)$, the operator is then evaluated as:
-\begin{multline*}
-\sem{[\alfp\varphi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)]}(\A,v,V) = \\ \zeta_n(a_1,\ldots,a_{\ell}).
-\end{multline*}
-As an example, we will show a formula that counts the number of paths of size $n$ of a structure $\A$ with a binary relation $E$. First we define $\varphi(x,R)$:
-\begin{multline*}
-\varphi(x,R) = \forall y(x < y \vee x = y) \vee \exists z(R(z) \wedge \varphi_{succ}(z,x)),
-\end{multline*}
+\begin{example}
+%As an example, 
+We would like to define a formula that, given a graph with $n$ nodes, counts the number of paths of length at most $n$ from a source node to a target node.
+To this end, assume that $\R = \{ E(\cdot,\cdot) \}$, and define $\psi(x,R)$ as follows:
+\begin{eqnarray*}
+%\psi(x,R) = 
+\forall y(x < y \vee x = y) \vee \exists z(R(z) \wedge \varphi_{succ}(z,x)),
+\end{eqnarray*}
 where $\varphi_{succ}(x,y)$ is a formula that is satisfied by pairs $(x,y)$ that are consecutive in the order $<$.
 That is, $\varphi_{succ}(x,y) = x < y \wedge \forall z((x < z \wedge z < y) \to (x = z \vee z = y) )$.
-Now we define $\alpha(x,y,R,\pi)$ as:
+Moreover, define $\alpha(x,y,R,\pi)$ as:
 $$
-\alpha(x,y,R,\pi) = (\neg \exists zR(z))\cdot(x = y) + \Sigma z[\pi(x,z)\cdot E(z,y)].
+%\alpha(x,y,R,\pi) = 
+(\neg \exists zR(z))\cdot(x = y) + \sa{z} [\pi(x,z)\cdot E(z,y)].
 $$
-Then, the formula $[\alfp\varphi(x,R)\mid \alpha(x,y,R,\pi)](x,y)$ when evaluated on $(\A,a,b)$ will count the number of paths of size $n$ from $a$ to $b$.
-This operator lets us define the set of recursive $\qfo$ formulas over $\R$ ($\rqfo$-formulas) using the following grammar:
+Then we have that $[\alfp\psi(x,R)\mid \alpha(x,y,R,\pi)](u,v)$ is our desired formula.
+In fact, assume that $\A$ is an $\R$-structure with $n$ elements in its domain, and assume $a,b$ are elements of $\A$.
+We have that the evaluation of $[\alfp\psi(x,R)\mid \alpha(x,y,R,\pi)](u,v)$ over $\A$ with $u$ replaced by $a$ and $v$ replaced by $b$ returns the number of paths in $\A$ from $a$ to $b$ of length at most $n$.
+%when evaluated on $(\A,a,b)$ will count the number of paths of size $n$ from $a$ to $b$.
+\end{example}
+
+Given a relation signature $\R$, the set of recursive $\qfo$ formulas ($\rqfo$-formulas) is defined by the following grammar:
+%This operator lets us define the set of recursive $\qfo$ formulas over $\R$ ($\rqfo$-formulas) using the following grammar:
 \begin{multline*}
-	\label{eq-def-rqfo}
+%	\label{eq-def-rqfo}
 	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid \\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ [\alfp \varphi \mid \alpha]
@@
 
+\marcelo{Vamos a permitir anidacion del operador $\alfp$? Esta gramatica lo permite.}
+
 \begin{theorem} \label{rqfo-fo-cap}

### 2017-01-02 LICS/beyond.tex
aun sin comentarios

@@
 We define an operator which extends least fixed point logic \cite{I86,vardi1982complexity} to allow counting. 
-Recall that the least fixed point operator is defined by a formula $\psi(x_1,\ldots,x_k,R)$ that is positive on $R$, where $R$  is a predicate of arity $k$.
+Fix a relational signature $\R$. Then the set of $\rqfo$ formulas over $\R$ is defined by the following grammar:
+\begin{multline*}
+	\alpha := \varphi \, \mid \, s \, \mid \, (\alpha \add \alpha) \, \mid\, (\alpha \mult \alpha) \, \mid \, 
+	\sa{x} \alpha \, \mid\\ 
+	\pa{x} \alpha \, \mid\,
+	\clfp{\psi(x_1,\ldots,x_k,R)}{\beta(y_1, \ldots, y_\ell, R, \pi)},
+\end{multline*}
+where $R$ is second-order variable of arity $k$, $\psi(x_1, \ldots, x_k, R)$ is an $\fo$-formula over $(\R \cup \{R\})$ that is positive on $R$ and $\beta(y_1, \ldots, y_\ell, R, \pi)$ is a $\qfo$-formula over $\R$ including a fresh symbol $\pi$ that represents a function of arity $\ell$, that is, every atomic formula in $\beta(y_1, \ldots, y_\ell, R, \pi)$ is either an $\fo$-formula over $\R$, a constant $s \in \N$ or a term of the form $\pi(u_1, \ldots, u_\ell)$ with $u_1, \ldots, u_\ell$ non-necessarily distinct first-order variables.
+The free variables of the formula $\clfp{\psi(x_1,\ldots,x_k,R)}{\beta(y_1, \ldots, y_\ell, R, \pi)}$
+are $y_1, \ldots, y_\ell$, in particular it does not have any second-order free variable.
+
+We need to introduce some terminology to define the semantics of the formula $\clfp{\psi(x_1,\ldots,x_k,R)}{\beta(y_1, \ldots, y_\ell, R, \pi)}$
+Given a $\qfo$-formula $\gamma$ with $\ell$ first-order free variables and no second-order free variable, define $\beta[\gamma/\pi]$ to be the $\qfo$-formula obtained from $\beta$ by replacing every occurrence of a term $\pi(u_1, \ldots, u_\ell)$ by $\gamma(u_1, \ldots, u_\ell)$.
+Let $\A$ be an $\R$-structure with domain $A$.
+Then for every $a \in A$, let $\varphi_a(x)$ be an $\fo$-formula such that $\A \models \varphi_a(b)$ if and only if $b = a$.
+Formally, assuming that $a$ is the $p$-th element of $A$ according to the linear order $<^\A$, we have that:
+%then $\varphi_a(x) = \forall y(x < y \vee x = y)$
+%for every natural number $i \geq 1$, let $\varphi_i(x)$ be an $\fo$-formula such that $\varphi_i(a)$ holds in an $\R$-structure $\A$ if and only if $a$ is the $i$-th element in the linear order  $<^{\A}$,
+%that is, $\varphi_1(x) = \forall y(x < y \vee x = y)$ and for every $p > 1$:
+\begin{multline*}
+\varphi_a(x) \ = \ \exists x_1 \cdots \exists x_{p-1}\bigg[\bigwedge_{i =1}^{p-2}x_i < x_{i+1} \wedge\,\\  
+x_{p-1} < x  \wedge \forall y(y < x \to \bigvee_{i = 1}^{p-1} y = x_i)\bigg].
+\end{multline*}
+Finally, given a function $f : A^\ell \to \N$, define $\qfo$-formula $\delta_{f,\A}(z_1, \ldots, z_\ell)$ as follows:
+\begin{multline*}
+ \mathop{+}_{(a_1,\ldots,a_{\ell})\in A^{\ell}} \ \varphi_{a_1}(z_1) \cdot \ldots \cdot \varphi_{a_{\ell}}(z_{\ell})\cdot f(a_1,\ldots,a_{\ell}), 
+\end{multline*}
+where $z_1, \ldots, z_\ell$ is a sequence of pairwise distinct first-order variables.
+We have that $\delta_{f,\A}(z_1, \ldots, z_\ell)$ encodes $f$ in the structure $\A$ since:
+\begin{eqnarray*}
+\sem{\delta_{f,\A}(z_1, \ldots, z_\ell)}(\A,v)  & = & f(v(z_1), \ldots, v(z_\ell)).
+\end{eqnarray*}
+%Recall that the least fixed point operator is defined by a formula $\psi(x_1,\ldots,x_k,R)$ that is positive on $R$, where $R$  is a predicate of arity $k$.
 For a structure $\A$ with domain $A$, the operator $T_{\varphi}:2^{A^k} \to 2^{A^k}$ is defined as $T_{\varphi}(X) = \{(a_1,\ldots,a_k)\mid (\A,X)\models \psi(a_1,\ldots,a_k,R) \}$, for each $X\subseteq A^k$.
@@
 Then the evaluation of $[\lfpop\psi(x_1,\ldots,x_k,R)]$ is defined by the fixed point $T_n$, that is, for every $(a_1,\ldots,a_k)\in A^k$, it holds that $\A\models[\lfpop\psi(x_1,\ldots,x_k,R)](a_1,\ldots,a_k)$ if and only if $(a_1,\ldots,a_k) \in T_n$.
-
-We extend $\qfo$ with the operator $[\alfp\psi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](z_1,\ldots,z_{\ell})$, where $\psi(y_1,\ldots,y_k,R)$ 
-%It is defined by a 
-is an FO-formula 
-%$\varphi$ 
-and $\alpha$  is a $\qso$-formula 
-%in $\qfo$ without the $\Pi$ operator, which 
-that mentions a placeholder $\pi$ for a function of arity $\ell$.
-%with domain $\fv^{\ell}$. 
-To define the semantics of such operator 
-%$[\alfp\psi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](x_1,\ldots,x_{\ell})$ 
-we need to introduce some terminology.
-Given a $\qso$-formula $\beta$ with $\ell$ first-order free variables and no second-order free variable, $\alpha|_{\pi\to\beta}$ is a $\qso$-formula obtained from $\alpha$ by replacing every occurrence $\pi(u_1, \ldots, u_\ell)$ of the placeholder by $\beta(u_1, \ldots, u_\ell)$.
-%For each $\qfo$ formulas $\alpha, \beta, \gamma$, where $\beta$ and $\gamma$ have $\ell$ and $m$ first-order free variables respectively, $u_1,\ldots,u_{\ell} \in \fv$, and $v_1,\ldots,v_m \in \{u_1,\ldots,u_{\ell}\}$, we define $\alpha\mid_{\beta(u_1,\ldots,u_{\ell})\to\gamma(v_1,\ldots,v_{m})}$ as $\alpha$ where every instance of the subformula $\beta(u_1,\ldots,u_{\ell})$ is replaced by $\gamma(v_1,\ldots,v_{m})$.
-Moreover, for every natural number $i \geq 1$, define a formula $\varphi_i(x)$ such that $\varphi_i(a)$ holds in a structure $\A$ if and only if $a$ is the $i$-th element in the linear order  $<^{\A}$.
-%$a\in A$, suppose that $a$ is the $p$-th element in the order $<^{\A}$. Then if $a$ is the first element, we define $\varphi_a(x) = \forall y(x < y \vee x = y)$, and if it is not we define:
-Thus, we have that $\varphi_1(x) = \forall y(x < y \vee x = y)$ and for every $p > 1$:
-\begin{multline*}
-\varphi_p(x) = \exists x_1 \cdots \exists x_{p-1}[\bigwedge_{1\leq i,j < p}x_i\neq x_j \wedge\,\\ \bigwedge_{i = 1}^{p-1} x_i < x  \wedge \forall y(y < x \to \bigvee_{i = 1}^{p-1} y = x_i)].
-\end{multline*}
-We are now ready to define the semantics of $[\alfp\psi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](z_1,\ldots,z_{\ell})$. 
-Assume that $\A$ is an $\R$-structure with domain $A$, and let $\{T_i\}_{i\in\nat}$ be as in the definition of $[\lfpop\psi(x_1,\ldots,x_k,R)]$.
-Moreover, suppose that $n\in\nat$ is the smallest value such that $T_{n+1} = T_n$. 
-%To formally characterize 
-%We name $\T = \{T_0,\ldots,T_n\}$ and 
-Then we define a sequence of functions $\zeta_0,\ldots,\zeta_n:A^{\ell}\to\nat$ as follows. 
-We first have that $\zeta_0(a_1, \ldots, a_\ell) = 0$ for every $(a_1, \ldots, a_\ell) \in A^\ell$.
-Now assume that function $\zeta_i$ has already been defined, where $0 \leq i < n$.
-By using this function define a formula $\beta_{i}(u_1, \ldots, u_\ell)$ as follows:
-%For each $T_i \in \T$, we define a $\qfo$ formula $\beta_i$ as follows. If $i = 0$, then $\beta_i(u_1,\ldots,u_{\ell}) = 0$. If $i \geq 1$, then
-\begin{multline*}
-%\beta_{i+1}(u_1,\ldots,u_{\ell}) = 
- \mathop{+}_{(a_1,\ldots,a_{\ell})\in A^{\ell}} \varphi_{\sigma(a_1)}(u_1) \cdot \ldots \cdot \varphi_{\sigma(a_{\ell})}(u_{\ell})\cdot \zeta_{i}(a_1,\ldots,a_{\ell}), 
-\end{multline*}
-where $(u_1, \ldots, u_\ell)$ is a tuple of pairwise distinct first-order variables and $\sigma(\cdot)$ is a function that assigns to every $a \in A$ its position in the lineal order $<^\A$. 
-Then function $\zeta_{i+1}$ is defined as follows.
-Assuming that $(a_1, \ldots, a_\ell) \in A^\ell$, $v$ is a first-order assignment for $\A$ such that $v(z_i) = a_i$ for every $i \in \{1, \ldots, \ell\}$, and that $V$ is a second-order assignment for $\A$ such that $V(R) = T_{i+1}$, we have that:
+We are now ready to define the semantics of the formula $\clfp{\psi(x_1,\ldots,x_k,R)}{\beta(y_1,\ldots,y_{\ell},R,\pi)}$,  whose free variables are $y_1$, $\ldots$, $y_\ell$.
+Assume that $\A$ is an $\R$-structure with domain $A$.
+Following the definition of the semantics of least fixed point logic \cite{I86,vardi1982complexity}, define an operator $T_{\varphi}:2^{A^k} \to 2^{A^k}$ such that:
+$$
+T_{\varphi}(X)  =  \{(a_1,\ldots,a_k) \in A^k \mid \A \models \psi(a_1,\ldots,a_k,X) \},
+$$
+for every $X \subseteq A^k$. This operator is used to defined 
+a sequence $T_0 \subsetneq T_1 \subsetneq \cdots \subsetneq T_n \subseteq A^k$  such that $n \geq 0$, $T_0 = \emptyset$, $T_{i+1} = T_{\varphi}(T_i)$ for every $i \in \{0, \ldots, n-1\}$, and $T_n = T_\varphi(T_n)$.
+We know that such a sequence exists as formula $\psi(x_1,\ldots,x_k,R)$ is positive on $R$.
+Now, the sequence $T_0 \subsetneq T_1 \subsetneq \cdots \subsetneq T_n$ is used to defined  a sequence of functions $f_0,f_1,\ldots,f_n:A^{\ell}\to\nat$.
+Formally, we have that $f_0(a_1, \ldots, a_\ell) = 0$ for every $(a_1, \ldots, a_\ell) \in A^\ell$.
+Moreover, assuming that $(a_1, \ldots, a_\ell) \in A^\ell$, $v$ is a first-order assignment for $\A$ such that $v(y_i) = a_i$ for every $i \in \{1, \ldots, \ell\}$, and that $V$ is a second-order assignment for $\A$ such that $V(R) = T_{i+1}$, we have that:
 \begin{eqnarray*}
-\zeta_{i+1}(a_1, \ldots, a_\ell) & = & \sem{\alpha\mid_{\pi \to \beta}}(\A,v,V).
+f_{i+1}(a_1, \ldots, a_\ell) & = & \sem{\beta[\delta_{f_i,\A}/\pi]}(\A,v,V).
 \end{eqnarray*}
 %, for each $T_i \in \T$, let $V$ be a second-order assignment for $\A$ that assigns $T_i$ to $R$, let $\zeta_i: A^{\ell}\to\nat$ be such that for each $(a_1,\ldots,a_{\ell})\in A^{\ell}$ it holds $\zeta_i(a_1,\ldots,a_{\ell}) = \sem{\alpha\mid_{\pi(u_1,\ldots,u_{\ell})\to \beta_i(u_1,\ldots,u_{\ell})}}(\A,v,V)$, where $v$ is a first-order assignment for $\A$ that satisfies $a_i = v(x_i)$ for each free $x_i$ in $\alpha$.
-Finally, the semantics of $[\alfp\psi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)](z_1,\ldots,z_{\ell})$ is defined by means of function $\zeta_n$:
+Finally, the semantics of the formula $\clfp{\psi(x_1,\ldots,x_k,R)}{\beta(y_1,\ldots,y_{\ell},R,\pi)}$ is defined by means of function $f_n$:
 %For a given first order assignment $v$ and a second order assignment $V$, let $a_i = v(x_i)$, the operator is then evaluated as:
 \begin{align*}
-&\llbracket [\alfp\psi(y_1,\ldots,y_k,R)\mid\\
-&\hspace{5mm}\alpha(x_1,\ldots,x_{\ell},R,\pi)](z_1,\ldots,z_{\ell}) \rrbracket =
-%\sem{[\alfp\psi(y_1,\ldots,y_k,R)\mid\alpha(x_1,\ldots,x_{\ell},R,\pi)]}(\A,v,V) = \\ 
-\zeta_n(v(z_1),\ldots,v(z_{\ell})).
+&\llbracket [{\bf lfp} \, \psi(x_1,\ldots,x_k,R) \,|\\
+&\hspace{10mm}\beta(y_1,\ldots,y_{\ell},R,\pi)]\rrbracket(\A,v) \ = \
+f_n(v(y_1),\ldots,v(y_{\ell})).
 \end{align*}
@@
 %As an example, 
-We would like to define a formula that, given a graph with $n$ nodes, counts the number of paths of length at most $n$ from a source node to a target node.
+We would like to define a formula that, given a graph $G$ with $n$ nodes and a pair of nodes $b$, $c$ in $G$, counts the number of paths of length at most $n$ from $b$ to $c$ in $G$.
 To this end, assume that $\R = \{ E(\cdot,\cdot) \}$, and define $\psi(x,R)$ as follows:
@@
 where $\varphi_{succ}(x,y)$ is a formula that is satisfied by pairs $(x,y)$ that are consecutive in the order $<$.
-That is, $\varphi_{succ}(x,y) = x < y \wedge \forall z((x < z \wedge z < y) \to (x = z \vee z = y) )$.
-Moreover, define $\alpha(x,y,R,\pi)$ as:
+That is, $\varphi_{succ}(x,y) = x < y \wedge \neg \exists z \, (x < z \wedge z < y)$.
+Moreover, define formula $\beta(x,y,R,\pi)$ as follows:
 $$
 %\alpha(x,y,R,\pi) = 
-(\neg \exists zR(z))\cdot(x = y) + \sa{z} [\pi(x,z)\cdot E(z,y)].
+%(\neg \exists zR(z))\cdot(x = y) 
+E(x,y) + \sa{z} [\pi(x,z)\cdot E(z,y)].
 $$
-Then we have that $[\alfp\psi(x,R)\mid \alpha(x,y,R,\pi)](u,v)$ is our desired formula.
-In fact, assume that $\A$ is an $\R$-structure with $n$ elements in its domain, and assume $a,b$ are elements of $\A$.
-We have that the evaluation of $[\alfp\psi(x,R)\mid \alpha(x,y,R,\pi)](u,v)$ over $\A$ with $u$ replaced by $a$ and $v$ replaced by $b$ returns the number of paths in $\A$ from $a$ to $b$ of length at most $n$.
-%when evaluated on $(\A,a,b)$ will count the number of paths of size $n$ from $a$ to $b$.
+Then we have that $\clfp{\psi(x,R)}{\beta(x,y,R,\pi)}$ defines our counting function.
+In fact, assume that $\A$ is an $\R$-structure with $n$ elements in its domain, $b,c$ are elements of $\A$ and $v$ is a first-order assignment over $\A$ such that $v(x) = b$ and $v(y) = c$.
+Then we have that $\sem{\clfp{\psi(x,R)}{\beta(x,y,R,\pi)}}(\A,v)$ is equal to the  number of paths in $\A$ from $b$ to $c$ of length at most $n$.
 \end{example}
 
-Given a relation signature $\R$, the set of recursive $\qfo$ formulas ($\rqfo$-formulas) is defined by the following grammar:
-%This operator lets us define the set of recursive $\qfo$ formulas over $\R$ ($\rqfo$-formulas) using the following grammar:
-\begin{multline*}
-%	\label{eq-def-rqfo}
-	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid \\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ [\alfp \varphi \mid \alpha]
-\end{multline*}
-where $\varphi$ is an $\fo$-formula over $\R$, $s \in \bbN$ and $x \in \fv$.
+\marcelo{Deberiamos mostrar como el ejemplo anterior se generaliza para definir $[\pth \psi(\bar x, \bar y)]$ en terminos del operador ${\bf lfp}$.}
 
-\marcelo{Vamos a permitir anidacion del operador $\alfp$? Esta gramatica lo permite.}
+It is well known that least fixed point logic is contained in second-order logic \cite{L04}. In the following theorem we show that the same holds in our case.
+\begin{theorem} \label{so-rec}
+$\rqfo \subseteq \qso$
+%	Given a positive $\fo$ formula $\varphi(\bar{x},R)$ and a $\qfo$ formula $\alpha(\bar{x})$, there exists a $\qso$ formula $\beta(\bar{x})$ such that $\sem{[\alfp\varphi(\bar{x},R)\mid \alpha(\bar{x},R)](\bar{x})} = \sem{\beta(\bar{x})}$.
+\end{theorem}
 
+Finally, the following is the main result of this section:
 \begin{theorem} \label{rqfo-fo-cap}
@@
 \end{theorem}
+
+\marcelo{Notese que estoy permitiendo en la formula $\beta(y_1, \ldots, y_\ell,R,\pi)$ tener subformulas $\pa{x} \alpha$. Esta bien el resultado con esto? O tenemos que eliminar estas formulas?}
+
+%Given a relation signature $\R$, the set of recursive $\qfo$ formulas ($\rqfo$-formulas) is defined by the following grammar:
+%%This operator lets us define the set of recursive $\qfo$ formulas over $\R$ ($\rqfo$-formulas) using the following grammar:
+%\begin{multline*}
+%%	\label{eq-def-rqfo}
+%	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid \\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ [\alfp \varphi \mid \alpha]
+%\end{multline*}
+%where $\varphi$ is an $\fo$-formula over $\R$, $s \in \bbN$ and $x \in \fv$.
+%
+%\marcelo{Vamos a permitir anidacion del operador $\alfp$? Esta gramatica lo permite.}
+%
+%\begin{theorem} \label{rqfo-fo-cap}
+%	$\rqfo(\fo)$ captures $\fp$ over the class of ordered structures.
+%\end{theorem}
+%

### 2017-01-02 LICS/appendix-syntactic.tex
added theo V.1 and half of V.2

@@
-\subsection*{Proof of Proposition \ref{prop:eqso-hierarchy-proper}} %V.1
+\subsection*{Proof of Theorem \ref{theo-pnf-snf}}
+Recall that a formula in $\eqso(\LL)$ is on the following grammar:
+$$
+\alpha = \varphi \ \mid \ s \ \mid \ (\alpha + \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \sa{X} \alpha,
+$$
+where $\varphi$ is a formula in $\LL$ and $s\in\nat$.
+We will show a recursive function $\tau$ such that for every $\eqso(\LL)$ formula $\alpha$, $\tau(\alpha)$ is a formula in $\LL$-SNF that describes the same function as $\alpha$.
+If $\alpha = \varphi$ or $\alpha = s$, then we define $\tau(\alpha) = \alpha$ which clearly describes the same function, and is already in $\LL$-SNF.
+If $\alpha = (\alpha_1 + \alpha_2)$, then $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$ which satisfies the condition.
+If $\alpha = \sa{x}\beta$, then let $\tau(\beta) = c + \sum_{i = 1}^{n}\beta_i$ where each $\beta_i$ is in $\LL$-PNF.
+First we replace $c$ with the formula $(\top \add \cdots \add \top)$ (c times) and we obtain a formula $\sum_{i = 1}^{m}\beta_i$ where each $\beta_i$ is in $\LL$-PNF.
+Since $\beta_i$ is in $\LL$-PNF, then $\sa{x}\beta_i$ also is, and then we define $\tau(\alpha) = \sum_{i = 1}^{m}\sa{x}\beta_i$ which satisfies the condition.
+If $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case.
+This covers all possible cases for $\alpha$ and we conclude the proof by taking $\tau(\alpha)$ as the desired rewrite of $\alpha$.
+
+\subsection*{Proof of theorem \ref{theo-pi1-pnf}}
+We divide the proof in three parts.
+
+\vspace{1em}
+First, we prove that the formula $\alpha_{0} = \sa{X(\cdot)}1 + 1$ is not equivalent to any formula in $\loge{0}$-PNF.
+Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ in $\loge{0}$-PNF that is equivalent to the $\eqso(\loge{0})$ formula $\sa{X(\cdot)}1 + 1$.
+Note that this defines the function $2^{n}+1$, where $n$ is the size of the input structure.
+Then, note that if $\varphi(\bar{X},\bar{x})$ is not satisfiable, then the function is null.
+Note also that if $\length(\bar{X}) = \length(\bar{x}) = 0$, then the function defined by $\alpha$ is never greater than 1.
+Now note that $\sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ is equivalent to a function in $\E{0}$.
+The authors in \cite{SalujaST95} proved that if $\length(\bar{X}) > 0$, then the function defined by the formula, for big enough structures, is always even.
+And since $\sa{\bar{x}}\varphi(\bar{x})$ defines a polynomially bounded function, we follow to a contradiction.
+
+\vspace{1em}
+Second, we prove that the formula $\alpha_{1} = 2$ is not equivalent to any formula in $\loge{1}$-PNF.
+Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to the $\eqso(\loge{1})$ formula $2$.
+Suppose that $\varphi$ is quantifier-free.
+Note first that if $\length(\bar{X}) = \length(\bar{x}) = 0$, then the function defined by $\alpha$ is never greater than 1.
+Consider an ordered structure $\A$.
+Since $\sem{\alpha}(\A) = 2$, there exist at least two assignations $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
+Now consider the ordered structure $\A'$ that is obtained by duplicating $\A$.
+This is, each half of $\A'$ is isomorphic to $\A$.
+Note that $\A'\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$ for $i = 1,2$ and there exists a third assignation $(\bar{B}_1',\bar{b}_1',\bar{a}_1')$ that is identical to $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$ but in the other half to the structure such that $\A'\models\varphi(\bar{B}_1',\bar{b}_1',\bar{a}_1')$.
+We have that $\sem{\alpha}(\A) \geq 3$, and we follow to a contradiction.
+
+\vspace{3em}
+
 Classes QE0, QE1, QU1, QE2, QU2 = QFO form a properly contained hierarchy:\vspace{1em}

### 2017-01-03 LICS/appendix-syntactic.tex
finished theo V.2 and V.3

@@
 If $\alpha = \sa{x}\beta$, then let $\tau(\beta) = c + \sum_{i = 1}^{n}\beta_i$ where each $\beta_i$ is in $\LL$-PNF.
-First we replace $c$ with the formula $(\top \add \cdots \add \top)$ (c times) and we obtain a formula $\sum_{i = 1}^{m}\beta_i$ where each $\beta_i$ is in $\LL$-PNF.
+First, if $c > 0$, we replace $c$ with the formula $(\top \add \cdots \add \top)$ (c times) and we obtain a formula $\sum_{i = 1}^{m}\beta_i$ where each $\beta_i$ is in $\LL$-PNF.
+If $c = 0$ the formula is already in that form.
 Since $\beta_i$ is in $\LL$-PNF, then $\sa{x}\beta_i$ also is, and then we define $\tau(\alpha) = \sum_{i = 1}^{m}\sa{x}\beta_i$ which satisfies the condition.
@@
 
-\vspace{3em}
+\vspace{1em}
+We now show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there is an equivalent formula $\beta$ in $\LL$-PNF.
+As in Theorem \ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such formula.
+As we showed, there exists an equivalent formula in $\LL$-SNF, so we assume that $\alpha$ is in that form.
+Let $\alpha = c + \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-SNF.
+First, as we did in Theorem \ref{theo-pnf-snf}, we replace the constant $c$ for $(\top \add \cdots \add \top)$ if $c > 0$ and we do nothing if $c = 0$, to obtain an equivalent formula $\sum_{i = 1}^m \alpha_i$.
+Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$.
+We replace each $\alpha_i$ for the equivalent formula $\sa{\bar{X}}\sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge\forall z\,Y(z) \wedge \forall z(y \leq z))$ to obtain that $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.
 
-Classes QE0, QE1, QU1, QE2, QU2 = QFO form a properly contained hierarchy:\vspace{1em}
+Now we begin describing the function $\tau$.
+If $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then the formula is already in $\LL$-PNF so we define $\tau(\alpha) = \alpha$.
+If $\alpha = \alpha_1 + \alpha_2$, then we assume that $\tau(\alpha_1) = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ and $\tau(\alpha_2) = \sa{\bar{X}}\sa{\bar{x}}\psi(\bar{X},\bar{x})$.
+The construction that we will provide for this function works by identifying a ``first'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ and a ``last'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$.
+These are identified by the following formulas:
+\begin{align*}
+\gamma_{\text{first}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z}), \\
+\gamma_{\text{last}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x}).
+\end{align*}
+In other words, the ``first'' assignment is the one where every second-order predicate is empty and the first-order assignment is the lexicographically smallest, and the ``last'' assignment is the one where every second-order predicate is full and the first-order assignment is the lexicographically greatest.
+We also need to identify assignments that are not first and are not last.
+We do this by negating the two formulas above, and we use a concatenation of the quantified first-order variables:
+\begin{align*}
+\gamma_{\text{not first}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{z}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{z}_i)), \\
+\gamma_{\text{not last}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{x} < \bar{z}_0 \vee \bigvee_{i = 1}^{\length(\bar{X})}\neg X(\bar{z}_i)),
+\end{align*}
+where $\bar{z} = (\bar{z}_0,\bar{z}_1,\ldots,\bar{z}_{\length(\bar{X})})$. The following formula is equivalent to $\alpha$:
+\begin{align}
+\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{not first}}(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{Y},\bar{y}))\vee\\
+&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))\vee\\
+&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{not last}}(\bar{Y},\bar{y}))\vee\\
+&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{last}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))].
+\end{align}
+To show that the formula is indeed equivalent to $\alpha$, note three facts.
+First, the formulas in lines (2) and (3) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$.
+Second, the formulas in lines (4) and (5) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
+And third, the four lines define pairwise disjoint assignments for $(\bar{X},\bar{x}),(\bar{Y},\bar{y})$.
 
-We give this proof in various parts.
+However, the formula is not yet in the correct form. By expanding it we obtain:
+\begin{align*}
+\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\exists\bar{v}(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i))\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
+&(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))\vee\\
+&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\exists\bar{w}(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i))\vee\\
+&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))].
+\end{align*}
+To construct an equivalent formula that is in the correct form, we define $\bar{u} = (\bar{v},\bar{w})$ and we replace the first-order quantifiers by a first-sum and count the first assignment to $\bar{v}$ and $\bar{w}$ that satisfies the formula.
+A similar construction was used in \cite{SalujaST95}.
+We obtain:
+\begin{align*}
+\sa{\bar{X}}\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i))\wedge\forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
+&(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))\vee\\
+&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee\\
+&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))].
+\end{align*}
+This covers all possible cases for $\alpha$.
 
-We show here that $\QE{0} \not\subseteq \E{1}$..
+Finally, consider a $\eqso(\LL)$ formula $\alpha$ in $\LL$-SNF.
+If $\alpha = \sum_{i = 1}^n\alpha_i$, then consider $\alpha = (\alpha_1 + (\sum_{i = 2}^n\alpha_i))$.
+Then we use $\tau(\alpha)$ as the rewrite of $\alpha$, which satisfies the condition in the hypothesis.
+
+\subsection{Proof of Theorem \ref{prop-rest}}
+We give this proof in three parts.
+
+\vspace{1em}
+First, we show that $\QE{0} \not\subseteq \E{1}$.
 By contradiction, suppose that there is a $\QE{0}$ formula $\alpha$ over some signature $\R$ such that defines the following function.
@@
 \vspace{1em}
-We show here that $\E{1}\not\subseteq\QE{0}$.
-By contradiction, suppose that the constant 1 is expressible in $\E{1}$.
-Let $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ describe this function for some quantifier-free formula $\varphi$.
+Now we show that $\E{1}\not\subseteq\QE{0}$.
+In Theorem \ref{theo-pi1-pnf} we proved that there is no formula in $\loge{1}$-PNF equivalent to the formula $\alpha = 2$.
+Every formula in $\E{1}$ can be expressed in $\loge{1}$-PNF, which implies that $2 \in \QE{0}$ and $2 \not\in \E{1}$.
+
+\vspace{1em}
+Lastly, we prove that $\eqso(\loge{1})\subsetneq\eqso(\logu{1})$.
+For inclusion, let $\alpha$ be a formula in $\eqso(\loge{1})$.
+Suppose that it is in $\loge{1}$-SNF.
+This is, $\alpha = c + \sum_{i = 1}^{n}\alpha_i$.
+Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, where $\varphi_i$ is quantifier-free, for each $\alpha_i$.
+We use the same construction used in \cite{SalujaST95}, and we obtain that the formula $\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \forall\bar{y}'(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')$ for every assignment to $(\bar{X},\bar{x})$.
+We do this replacement for each $\alpha_i$ and we obtain an equivalent formula in $\eqso(\logu{1})$.
+
+To prove that the inclusion is proper, consider the $\eqso(\logu{1})$ formula $\sa{x}\forall y(y = x)$. This formula defines the following function that takes an ordered structure $\A$ as input:
+$$
+\sem{\alpha}(\A) = 
+\begin{cases}
+1 &\A \text{ has one element}\\
+0 &\text{ otherwise}.
+\end{cases}
+$$
+Suppose that there exists an equivalent formula $\alpha$ in $\eqso(\loge{1})$.
+Also, suppose that it is in $\L$-PNF, so $\alpha = c + \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$.
+Since $\alpha$ takes the value 0 for some structures, $c$ must be 0.
+Consider a structure with one element $\mathfrak{1}$.
+We have that for some $i$, there exists an assignment $(\bar{B},\bar{b},\bar{a})$ for $(\bar{X},\bar{x},\bar{y})$ such that $\mathfrak{1}\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
+Consider now the structure $\mathfrak{2}$ that is obtained by duplicating $\mathfrak{1}$, as we did for Theorem \ref{theo-pi1-pnf}.
+Note that $\mathfrak{2}\models\varphi_i(\bar{B},\bar{b},\bar{a})$, which implies that $\sem{\alpha}(\mathfrak{2}) \geq 1$, which leads to a contradiction.
 

### 2017-01-03 LICS/appendix-syntactic.tex
added V.6, V.8, V.9

@@
 
-\subsection*{Proof of Theorem \ref{theo:eqso-sum-normal-form}} %V.2
-Every formula in $\eqso(\LL)$ can be rewritten in \emph{sum normal form}.
-Furthermore, there exists a formula $\alpha \in \QE{1}$ such that there does not exist a formula in $\Sigma$-prenex normal form in $\QE{1}$ equivalent to $\alpha$:\vspace{1em}
+\subsection*{Proof of Theorem \ref{theo:binary-prod}}
 
-\subsection*{Proof of Theorem \ref{theo:prenex}}%V.3
-If $\logu{1} \subseteq \LL$, then for every formula $\alpha \in \eqso(\LL)$ there exists a formula $\beta \in \eqso(\LL)$ equivalent to $\alpha$ in $\Sigma$-prenex normal form:\vspace{1em}
+Let $\LL$ be a fragment of $\so$ that is closed under conjunction.
+We will define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ with multiplication, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
+Furthermore, the formula $\tau(\alpha)$ is in $\LL$-SNF.
+First we replace each constant $s$ in $\alpha$ for $(\top \add \cdot \add \top)$ ($s$ times).
+If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$.
+We assume that for every $\beta$ that has less algebraic operators than $\alpha$, $\tau(\beta)$ is in $\LL$-SNF.
+If $\alpha = (\alpha_1 + \alpha_2)$ then we define $\tau(\alpha) = \tau(\alpha_1) + \tau(\alpha_2)$.
+If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$, then we define $\tau(\alpha)$ as the formula in $\LL$-SNF that is equivalent to $\sa{x}\tau(\beta)$ and to $\tau(\alpha) = \sa{X}\tau(\beta)$, respectively.
+If $\alpha = (\alpha_1 \cdot \alpha_2)$, we assume that each $\alpha_i$ is in $\LL$-SNF.
+We identify three cases.
+(1) Some $\alpha_i$ is equal to $\sum_{j = 1}^n\beta_j$ for $n > 1$.
+Suppose wlog.
+that it is $\alpha_1$.
+We then define $\tau(\alpha) = \sum_{j = 1}^n\tau(\beta_j\cdot\alpha_2)$.
+In the following cases, $\alpha_1$ and $\alpha_2$ are in $\LL$-SNF.
+(2) If some $\alpha_i$ is equal to $\sa{X}\beta$ or $\sa{x}\beta$, we define $\tau(\alpha)$ as the $\LL$-SNF formula that is equivalent to $\sa{x}\tau(\beta\cdot\alpha_2)$ and $\sa{X}\tau(\beta\cdot\alpha_2)$, respectively.
+The remaining case is (3) $\alpha_1 = \varphi_1$ and $\alpha_2 = \varphi_2$ where each $\varphi$ is an $\LL$ formula.
+Then we define $\tau(\alpha) = \varphi_1 \wedge \varphi_2$.
+This covers all possible cases for $\alpha$.
+For every formula $\alpha$ in $\eqso(\LL)$ extended by multiplication, we take $\tau(\alpha)$ which is equivalent and in unextended $\eqso(\LL)$.
+This concludes the proof.
 
-\subsection*{Proof of Proposition \ref{prop:sharp-Q-rel}}%V.4
-The classes $\E{0}$ and $\E{1}$ are strictly contained in $\QE{0}$ and $\QE{1}$, respectively.
-Moreover, the classes $\U{1}$, $\E{2}$, and $\U{2}$ are equivalent with $\QU{1}$, $\QE{2}$, and $\QU{2}$, respectively:\vspace{1em}
+\subsection*{Proof of Theorem \ref{sigmafo-minusone}}
 
-\subsection*{Proof of Proposition \ref{prop:qe0-fp-qe1-totp-fptras}} %V.5
-All functions defined in $\eqso(\loge{0})$ and $\eqso(\loge{1})$ can be computed in $\fp$ and $\totp$, respectively. Furthermore, every function defined in $\eqso(\loge{1})$ has a FPTRAS:\vspace{1em}
+We divide the proof in two parts.
 
-\subsection*{Proof of Theorem \ref{theo:binary-prod}}%V.6
-If $\LL$ is closed under conjunction, then binary product can be defined in $\eqso(\LL)$:\vspace{1em}
-
-\subsection*{Proof of Theorem \ref{sub-pnp}}%V.7
-If $\eqso(\loge{i})$ or $\eqso(\logu{i})$ is closed under subtraction for $i > 0$, then {\sc P} = {\sc NP}:\vspace{1em}
-
-Suppose that $\E{1}$ is closed under substraction, that is, for each pair of functions $f,g\in \E{1}$, there exists $h\in\E{1}$ such that $(f\dotminus g)(\A) = h(\A)$ for each $\A\in\str$.
-
-Let $\A = \langle A, S_1^\A, S_2^\A, S_3^\A, S_4^\A, \leq^\A \rangle$ be an $\L-$structure that represents an instance of a 3DNF formula $\Phi$, where $A$ is the set of variables mentioned in $\Phi$, $S_i^\A$ is a ternary relation described as follows, for each $i\in\{1,2,3,4\}$:
-\begin{eqnarray*}
-	S_1^\A &=& \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-	S_2^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-	S_3^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-	S_4^\A &=& \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
-\end{eqnarray*}
-Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
-\begin{multline*}
-\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
-\end{multline*}
-Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
-$$
-\varphi(x,X) = (T(x) \vee \neg T(x)).
-$$
-Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Given that $f_{\#3DNF}, f_{all} \in \E{1}$, we have by our initial assumption that $f_{all}-f_{\#3DNF} \in \E{1}$.
-Let $h\in\E{1}$ be such that $h = f_{all}-f_{\#3DNF}$.
-For each structure $\A$ that represents a 3DNF formula $\psi$, it holds that $h(\A) = f_{all}(\A)-f_{\#3DNF}(\A) = 0$ if and only if $\psi$ is a tautology, so the decision version $L_h$ of $f_{all}-f_{\#3DNF}$ is $\conp$-complete.
-However, as we showed previously in Theorem \ref{decisionptime}, since $h\in\E{1}$, we have that $L_h \in \ptime$.
-Then, $\conp \subseteq \ptime$, from which we conclude that $\ptime = \np$.
-
-\subsection*{Proof of Theorem \ref{sigmafo-minusone}}%V.8
+\vspace{1em}
+First, note that for every fragment $\LL$ of $\so$ the class $\eqso(\LL)$ is closed under sum by definition, and if $\LL$ is closed under conjunction, like $\logex{1}$ is, $\eqso(\LL)$ is closed under product.
+We are only left to prove that $\eqso(\logex{1})$ is closed under subtraction by one.
 
@@
 We will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_n$ be defined as in \ref{eqso-sigma-zero-in-fp}, that is, each $\beta_i$ is equal to $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}\sa{\bar{x}}1$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, where $\varphi$ is a formula in $\logex{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables.
-We further assume that each $\beta_i$ is either $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, since each $1$, $\varphi$ and $\sa{\bar{X}}\varphi$ can be replaced by the formulas $\sa{x}\min(x)$, $\sa{x}(\varphi\wedge\min(x))$ and $\sa{x}\sa{\bar{X}}(\varphi\wedge\min(x))$ respectively, while maintaining the value of $\sem{\alpha}(\A)$.
+We suppose that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
+Moreover, we assume that $\length(\bar{x}) > 0$ since we can replace each $\varphi_i$ for the equivalent formula $\sa{y}\varphi_i\wedge\min(y)$.
 
@@
 For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha)$ procedurally as follows: Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
-We suppose that $\varphi$ is in a DNF form that leaves formulas that do not mention $\bar{X}$ as is.
+We suppose that $\varphi$ is in a DNF form that leaves formulas that do not mention $\bar{X}$ intact.
 If it is not, we convert $\varphi$ to this form with a standard transformation algorithm.
@@
 $$
-\begin{claim}
+\begin{lemma}
 	For a given ordered structure $\A$ such that $\A\models\exists\bar{X}\exists\bar{x}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, there is an assignment $(\bar{B},\bar{b})$ to $(\bar{X},\bar{x})$ that satisfies the following conditions (1) $\A\models\exists\bar{y}\,\varphi_i(\bar{B},\bar{b},\bar{y})$, (2) $\A\not\models\psi_i(\bar{B},\bar{b})$ and (3) this is the only assignment to $(\bar{X},\bar{x})$ that satisfies (1) and (2).
-\end{claim}
+\end{lemma}
 \begin{proof}
-	proof here.
-\end{proof}
-The following claim identifies a crucial condition that is verifiable with a $\fo$ formula. 
+	Excusing the nested proof, we have a claim that identifies a crucial condition that is verifiable with a $\fo$ formula. 
 	\begin{claim}
@@
 	\end{proof}
+	We give the proof for the base theorem.
+Let $\A$ be an ordered structure with domain $A$.
+We clearly have that $\A\models\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$, so let $\bar{a}\in A^{\length(\bar{y})}$ and $\bar{b}\in A^{\length(\bar{x})}$ such that $\A\models\mu_i(\bar{b},\bar{a})$.
+We use the same construction in the claim and we obtain the assignment $(\bar{B},\bar{b},\bar{a})$ which satisfies the three conditions.
+\end{proof}
+
 We define $\chi_i = \exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$.
@@
 
-\subsection*{Proof of Proposition \ref{prop:uhorn-pe}} %V.9
+\vspace{1em}
+For the second part, for each $\alpha$ in $\eqso(\logex{0})$ over a signature $\R$ we will define a formula $\lambda(\alpha)$ in $\eqso(\loge{0})$ over a signature $\R_{\alpha}$, and a function $g_{\alpha}$ that receives a finite $\R$-structure $\A$ and outputs a finite $\R_{\alpha}$-structure $g_{\alpha}(\A)$.
+Let $\alpha$ be in $\eqso(\logex{0})$.
+The signature $\R_{\alpha}$ is defined by adding the symbol $R_{\psi}$ for every $\fo$ formula $\psi(\bar{z})$.
+Each symbol $R_{\psi}$ represents a predicate with arity $\length(\bar{z})$.
+Then, $\lambda(\alpha)$ is defined by replacing every $\fo$ formula $\psi(\bar{z})$ for $R_{\psi}(\bar{z})$.
+We now define the function $g_{\alpha}$.
+Let $\A$ be a $\R$-structure with domain $A$.
+Let $\A'$ be a $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A}$ empty.
+For each $\fo$-formula $\psi(\z)$ with $\length(\bar{z})$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length(\bar{z})}$.
+If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$.
+This concludes the construction of $\A'$.
+Note that the number of $\fo$ subformulas and each arity and tuple size is fixed to $\alpha$, so the entire procedure takes polynomial time over the size of the structure (and $g_{\alpha}(\A)$ has polynomial size over $\A$).
+We define $g_{\alpha}(\A) = \A'$ and we have that for each finite $\R$-structure $\A$: $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$.
+Therefore, we have a parsimonious reduction from $\alpha$ to $\lambda(\alpha)$, which is in $\eqso(\loge{1})$.
+
+To show that $\alpha$ is in $\totp$, consider a procedure that takes an input $\A$, converts it to $g_{\alpha}(\A)$ and simulates the $\totp$ procedure for $\lambda(\alpha)$ on input $g_{\alpha}(\A)$.
+This procedure generates exactly $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$ branches, and therefore $\alpha$ is in $\totp$.
+
+To show that $\alpha$ has an FPRAS, consider a procedure that takes an input $\A$, converts it to $g_{\alpha}(\A)$ and simulates the FPRAS for $\lambda(\alpha)$.
+This procedure also takes polynomial time and satisfies the condition.
+
+\subsection*{Proof of Theorem \ref{sub-pnp}}
+
+Towards a contradiction, suppose that any of the classes $\E{1}$, $\eqso(\loge{1})$, and $\eqso(\logex{1})$ is closed under subtraction.
+
+Let $\R$ include the symbols $S_1, S_2, S_3, S_4$, which describe the following properties.
+If a finite $\R$-structure $\A$ defines a 3DNF formula $\Phi$, then its domain is the set of variables mentioned in $\Phi$, and for each $i = 1,2,3,4$:
+\begin{align*}
+	S_1^\A &= \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
+	S_2^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
+	S_3^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
+	S_4^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
+\end{align*}
+Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
+\begin{multline*}
+\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
+\end{multline*}
+Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
+$$
+\varphi(x,X) = (T(x) \vee \neg T(x)).
+$$
+Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
+Given that $f_{\#3DNF}, f_{all} \in \E{1} \subseteq \eqso(\loge{1}) \subseteq \eqso(\logex{1}) \subseteq \totp$, and at least one of the classes is closed under subtraction, the function $f_{\#3DNF} - f_{all}$ is in $\totp$.
+However, note that for each $\R$-structure $\A$ that represents a formula $\Phi$, $(f_{\#3DNF} - f_{all})(\A) = 0$ if an only if $\Phi$ is a tautology.
+The decision version of this function is the $\np$-complete problem $\textsc{Tautology}$.
+And since the function is in $\totp$, its decision version is also in $\ptime$.
+We conclude that $\np \subseteq \ptime$.
 

### 2017-01-03 LICS/appendix-counting.tex
03 ene

@@
 We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$, by the fact that $\shpspace = \fpspace$ \cite{Ladner89}, and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-\cristian{Aca debes explicar mejor que hace la máquina.
-Es un copy paste de demostraciones anteriores y no queda claro porque puedes evaluar la formula en FPSPACE.
-Tampoco queda claro porque puedes usar no determinismo en FPSPACE.} \martin{es un copy-paste porque hice la demostración para $\shpspace$, lo dejé un poco más claro ahora}
 
@@
 Therefore, $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-\qed
 
@@
 
-\martin{no revisar, voy a hacer un copy paste del teorema \ref{theo:capture-fp} cuando este listo}
-
-Let $\alpha\in\qfo(\pfp)$ over some $\R$.
+For the first condition, let $\alpha\in\qfo(\pfp)$ over some signature $\R$.
 Let $f$ be a function over $\R$ defined by the following procedure.
@@
 Then each sub-formula $\varphi\in\pfp$ in $\alpha$ is replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-\cristian{Explicar porque puedes hacer esto en FPSPACE(poly)}
-The resulting formula is an arithmetic expression of polynomial size which is evaluated and lastly given as output. Note that $f\in\nfpspace$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
+The resulting formula is an arithmetic expression of polynomial size which is evaluated and lastly given as output.
+Note that $f\in\nfpspace$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
 
-Now, let $f\in \nfpspace$ defined over some $\R$.
+For the second condition, let $f\in \fpspace$ defined over some signature $\R$.
 Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
-Let $\Phi(x_0,\ldots,x_{\ell-1})$ be a $\pfp$ formula that represents the following.
-For each $\A \in \ostr[\R]$, for convenience, suppose its domain is $A = \{0,\ldots,n-1\}$ with the usual order.
-Let $y$ be a string of size $n^{\ell}$ that represents the value of $f(\enc(\A))$ in binary.
-Then, for each $\ell$-tuple $(a_0,\ldots,a_{\ell-1}) \in A^{\ell}$, it holds that $\A \models \Phi(a_0,\ldots,a_{\ell-1})$ if and only if the $(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)$-th bit of $y$, from least to most significant, is 1.
+Let $\Phi(x_0,\ldots,x_{\ell-1})$ be a formula that represents the following.
+Let $\A$ be an ordered structure over $\R$ with domain $A$.
+Consider a linear order over $A^{\ell}$ given by $\bar{a} < \bar{b}$, for any $\bar{a},\bar{b}\in A^{\ell}$.
+Let $w$ be a string of size $n^{\ell}$ the represents $f(\enc(\A))$ in binary.
+Let $(a_1,\ldots,a_{\ell})\in A^{\ell}$ be the $m$-th tuple in the given order (starting from 0).
+$\A\models\Phi(a_1,\ldots,a_{\ell})$ if an only if the $m$-th bit in $w$ is 1, from least to most significant.
+Since this condition can be computed in polynomial space, and $\pfp$ captures $\pspace$, this formula can be expressed in $\pfp$.
 We use
 $$
-\alpha = \Sigma x_0 \cdots \Sigma x_{\ell-1} \Phi(x_0,\ldots,x_{\ell-1})\cdot\varphi_{\ell-1}(x_{\ell-1})\cdot\,\cdots\,\cdot\varphi_0(x_0).
+\alpha = \sa{x_1} \cdots \sa{x_{\ell}} \Phi(x_1,\ldots,x_{\ell})\cdot\varphi(x_1,\ldots,x_{\ell}),
 $$
-where $\varphi_i(x) = \Pi y[(y < x)\mapsto\Pi z_1\cdots\Pi z_i\,2]$ for $i > 0$ and $\varphi_0(x) = \Pi y[(y < x)\mapsto 2]$.
-Let $\A$ the previously defined structure.
-Note that for each $a\in A$, then $\sem{\varphi_i(x)}(\A,a) = 2^{n^i a}$.
-Therefore, for each $(a_0,\ldots,a_{\ell-1})\in A^\ell$, we have that
-$$
-\sem{\Phi(x_0,\ldots,x_{\ell-1})\cdot\varphi_{\ell-1}(x_{\ell-1})\,\cdots\,\varphi_0(x_0)}(\A,a_0,\ldots,a_{\ell-1}) = 
-\begin{cases}
-2^{(n^{\ell-1}a_{\ell-1} + \cdots + n^2 a_2 + n a_1 + a_0)} &\text{if } (\A,a_0,\ldots,a_{\ell-1})\models\Phi \\
-0 &\text{otherwise.}
-\end{cases}
-$$
-and adding these values gives $f(\enc(\A))$. Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$. \qed
+where $\varphi(x_1,\ldots,x_{\ell}) := \pa{y_1}\cdots\pa{y_{\ell}}((y_1,\ldots,y_{\ell}) < (x_1,\ldots,x_{\ell}) \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
+Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
+Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.

### 2017-01-03 LICS/appendix-syntactic.tex
03 ene

@@
 Note that $\mathfrak{2}\models\varphi_i(\bar{B},\bar{b},\bar{a})$, which implies that $\sem{\alpha}(\mathfrak{2}) \geq 1$, which leads to a contradiction.
+
+\subsection*{Proof of Proposition \ref{prop-e1-nc}}
+
+Towards a contradiction, assume that the statement is false.
+This is, that $\E{1}$ is closed under binary sum.
+Consider the formula $\sa{x}(x = x)$ which is in $\E{1}$ over some signature $\R$.
+For every finite $\R$-structure $\A$ with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n$.
+From our assumption, there exists some formula in $\E{1}$ equivalent to the formula $\alpha \add \alpha$, which describes the function $2n$.
+Let $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ be this formula, where $\varphi$ is quantifier-free.
+Note that the function defined by this formula is equal or greater than the one defined by $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ divided by a polynomial factor.
+More specifically, for each ordered structure $\A$ with domain $A$, we have the following inequality:
+$$
+\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot \vert A \vert^{\length(\bar{y})} \geq \sem{\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A)
+$$
+Note that the formula $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ defines a function in $\E{0}$.
+It was shown by the authors in \cite{SalujaST95} that every function in $\E{0}$ grows exponentially over the size of the structure for large enough structures, when $\length(\bar{X}) > 0$.
+This function divided by a polynomial factor still grows exponentially.
+Therefore, for $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ we have that $\length(\bar{X}) = 0$.
+
+Now, for the formula $\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ note that if $\length(\bar{x}) = 0$, then the formula takes value at most 1, and if $\length(\bar{x}) = 1$, then the formula takes value at most $n$.
+Then we have that $\length(bar{x}) \geq 2$.
+Let $\psi = \bigvee_{i = 1}^{m}$ be an equivalent formula to $\varphi$ that is in DNF
 
@@
 
-\subsection*{Proof of Proposition \ref{prop:ehorn-pe}} %V.11
+\subsection*{Proof of Proposition \ref{prop:uhorn-pe}} %V.11
+This can be obtained directly by Proposition \ref{prop:ehorn-pe}, since $\eqso(\uhorn)\subseteq\eqso(\ehorn)\subseteq\totp$.
 

### 2017-01-04 LICS/appendix-syntactic.tex
04 ene

@@
 
-\subsection{Proof of Theorem \ref{prop-rest}}
+\subsection*{Proof of Theorem \ref{prop-rest}}
 We give this proof in three parts.
@@
 
-Now, for the formula $\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ note that if $\length(\bar{x}) = 0$, then the formula takes value at most 1, and if $\length(\bar{x}) = 1$, then the formula takes value at most $n$.
-Then we have that $\length(bar{x}) \geq 2$.
-Let $\psi = \bigvee_{i = 1}^{m}$ be an equivalent formula to $\varphi$ that is in DNF
+Now, for the formula $\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ consider a structure $\mathfrak{1}$ with only one element $a$.
+We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$, but the only possible assignment to $\bar{x}$ is the tuple $(a,\ldots,a)$ so $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which follows to a contradiction.
+
+\subsection*{Proof of Proposition \ref{prop:qe0-fp-qe1-totp-fptras}}
+
+We separate this proof in two parts
+
+\vspace{1em}
+The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$.
+This is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every finite $\R$-structure $\A$ with domain $A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$.
+We base our proof on this fact.
+
+For the first part, let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
+This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
+Consider the following nondeterministic procedure that on input $\enc(\A)$ generates $\sem{\alpha}(\A)$ branches.
+For each $\alpha_i = \varphi$, where $\varphi$ is a $\loge{1}$ formula, it checks if $\A\models\varphi$ in polynomial time and generates a new branch if that is the case.
+For each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi$, this formula is also in $\E{1}$.
+We use the reduction to $\cdnf$ provided in \cite{SalujaST95} and we obtain $g(\enc(\A))$, which is an instance to $\cdnf$.
+Since $\cdnf$ is also in $\totp$ \cite{PagourtzisZ06}, we simulate the corresponding nondeterministic procedure that generates exactly $\cdnf(\enc(g(\A)))$ branches.
+Since, $\fp\subseteq\totp$\cite{PagourtzisZ06}, each polynomially computable function is also in $\totp$, and then on each of these branches we simulate the corresponding nondeterministic procedure to generate $h(\vert A \vert)$ more.
+The number of branches for each $\alpha_i$ is $\sem{\alpha_i}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$, and the total number of branches in the procedure amounts to $\sem{\alpha}(\A)$.
+We conclude that $\alpha\in\totp$.
+
+\vspace{1em}
+For the second part, let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
+This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
+Note that each $\alpha_i$ that is equal to some $\loge{1}$ formula $\varphi$ has an FPRAS given by the procedure that simply checks if $\A\models\varphi$ and returns 1 if it does and 0 otherwise.
+Also, each remaining $\alpha_i$ has an FPRAS since $\alpha_i\in \E{1}$ \cite{SalujaST95}.
+If two functions have an FPRAS, then their sum also has one given by the procedure that simulates them both and sums the results.
+We conclude that $\alpha$ has an FPRAS.
 

### 2017-01-04 LICS/appendix-syntactic.tex
04 ene

@@
 
+\subsection*{Proof of Proposition \ref{pi-minusone}}
+
+Let $\LL$ be a fragment of $\fo$ that contains $\logu{1}$.
+Then we have that every function in $\U{1}$ is expressible in $\eqso(\LL)$.
+In particular, $\ctcnf \in \eqso(\LL)$.
+Suppose that $\eqso(\LL)$ is closed under subtraction by one.
+Then, the function $\ctcnf-1$, which counts the number of satisfying assignments of a 3-CNF formula minus one, is also in $\eqso(\LL)$.
+Note that $\eqso(\LL) \subseteq \eqso(\fo) = \shp$.
+We have that $\ctcnf$ is $\shp$-complete under parsimonious reductions\footnote{It can be easily verified that the standard reduction from SAT to 3-CNF (or 3-SAT) preserves the number of satisfying assignments}.
+Now, let $f$ be a function in $\shp$, and consider the nondeterministic polynomial-time procedure that on input $\enc(\A)$ computes the corresponding reduction to $\ctcnf$, name it $g(\enc(\A))$, and simulates the $\shp$ procedure for $\ctcnf-1$ on input $g(\enc(\A))$.
+We have that this is a $\shp$ procedure that computes $f-1$, from which we conclude that $\shp$ is closed under subtraction by one.
+
 \subsection*{Proof of Theorem \ref{sigmafo-minusone}}
@@
 
+\subsection*{Proof of Proposition \ref{prop:ehorn-pe}}
+Pagourtzis and Zachos mention a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula \cite{PagourtzisZ06}.
+This procedure can be easily extended to receive Horn formulas, and furthermore, a disjunction of Horn formulas.
+We can deduce that $\shdhsat$ is in $\totp$.
+
+As we show in Proposition \ref{sigma2hard}, $\shdhsat$ is complete for $\eqso(\ehorn)$ under parsimonious reductions.
+Let $\alpha$ be a formula in $\eqso(\ehorn)$ and let $g_{\alpha}$ be the reduction to $\shdhsat$.
+The $\totp$ procedure we construct, for each input $\enc(\A)$, is simply to compute $g_{\alpha}(\enc(\A))$, and then simulate the $\totp$ procedure for $\shdhsat$ on input $g_{\alpha}(\enc(\A))$.
+We conclude that $\alpha$ is in $\totp$.
+
 \subsection*{Proof of Proposition \ref{prop:hsat-not-sigma2}} %V.12
-$\eqso(\uhorn) \subseteq \totp$:\vspace{1em}
+We use a similar proof to the one provided by the authors in \cite{SalujaST95} to separate the classes $\E{2}$ and $\U{2}$.
+Suppose that the statement is false, this is, $\chsat \in \eqso(\loge{2})$.
+We consider the signature $\R$ that we used as the encoding for a Horn formula (Example \ref{ex-hornsat-esop1}) and that the formula $\alpha \in \eqso(\loge{2})$ follows the encoding in the same way.
+From what we proved in Theorem \ref{theo-pnf-snf}, we have that every formula in $\eqso(\loge{2})$ can be rewritten in $\loge{2}$-PNF, so we assume that $\alpha$ is in this form.
+Let $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\forall\bar{z}\,\varphi(\bar{X},\bar{x},\bar{y},\bar{z})$.
+Consider the following Horn formula $\Phi$:
+$$
+Phi = p \wedge \bigwedge_{i = 1}^n (t_i \vee p \to q) \wedge \neg q,
+$$
+where $n = \length(\bar{x}) + \length(\bar{y}) + 1$.
+Let $\A$ be the encoding of this formula.
+In the encoding, each variable appears as an element in the domain of $\A$.
+This formula is satisfiable, so $\sem{\alpha}(\A) \geq 1$.
+Let $(\bar{B},\bar{b},\bar{a})$ be an assignment to $(\bar{X},\bar{x},\bar{y})$ such that $\A\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
+Let $t_{\ell}$ be such that it does not appear in $\bar{b}$ or $\bar{a}$.
+Consider the induced substructure $\A'$ that is obtained by removing $t_{\ell}$ from $\A$ and $\bar{B}'$ as the subset of $\bar{B}$ obtained by deleting each appearance of $t_{\ell}$ in $\bar{B}$.
+We have that $\A'\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
+This is because each subformula of the form $\exists y \neg B_i$ is still true, and universal formulas are monotone over induced substructures.
+It follows that $\sem{\alpha}(\A') \geq 1$ which is not possible since $\A'$ encodes the formula
+$$
+Phi' = p \wedge \bigwedge_{i = 1}^{\ell-1} (t_i \vee p \to q) \wedge (p\to q) \wedge \bigwedge_{i = \ell+1}^{n} (t_i \vee p \to q) \wedge \neg q,
+$$
+which is unsatisfiable. We arrive to a contradiction and we conclude that $\chsat$ is not in $\eqso(\ehorn)$.
 

### 2017-01-04 LICS/syntactic-clo.tex
04 ene

@@
 However, if $\shp$ is closed under subtraction by one, then it holds that $\np \subseteq \cspp$ \cite{OH93}.} so the following proposition rules out any logic that extends $\Pi_1$.
-\begin{proposition}
+\begin{proposition} \label{pi-minusone}
 If $\Pi_1 \subseteq \LL$ and $\eqso(\LL)$ is closed under subtraction by one, then $\shp$ is closed under subtraction by one. 
 \end{proposition}
+\martin{para este teorema se necesita que $\LL$ sea un fragmento de $\fo$}
 Therefore, the desired extension has to be achieved by allowing some local extensions to $\Sigma_1$.

### 2017-01-05 LICS/counting.tex
sin comentarios
04 ene

@@
 We can now formalize the notion of capturing a function complexity class over $\R$-structures.
-\begin{definition} \label{cap-def}
+\begin{definition} \label{def:cap}
 	Let $\FF$ be a fragment of $\qso$ and $\CC$ a function complexity class. Then we say that {\em  $\FF$ captures $\CC$ over $\R$-structures} if the  following conditions are true:
@@
 %For the sake of simplification, we denote the first condition by $\FF \subseteq \CC$ and the second condition by $\CC \subseteq \FF$.
-Note that in Definition~\ref{cap-def} the function $f \in \CC$ and the formula $\alpha \in \FF$ must coincide in all the strings that encode $\R$-structures.
+Note that in Definition~\ref{def:cap} the function $f \in \CC$ and the formula $\alpha \in \FF$ must coincide in all the strings that encode $\R$-structures.
 Since we want to capture $\CC$ over a fix set of structures (e.g.
@@
 To capture $\fp$ one first shows that any formula in $\qfo(\lfp)$ can be evaluated in $\ptime$. 
-Indeed, $\lfp$ is a polynomial time logic~\cite{Immerman82,vardi1982complexity} and the sum and product quantifiers can be computed in $\fp$. 
+Indeed, $\lfp$ is a polynomial time logic~\cite{I86,vardi1982complexity} and the sum and product quantifiers can be computed in $\fp$. 
 For the other direction, one has to use $\lfp$ to simulate the run of the $\fp$-machine and use the quantitative quantifiers to reconstruct the binary number built in the tape. 

### 2017-01-05 LICS/syntactic-clo.tex
04 ene

@@
 \end{proposition}
-\martin{para este teorema se necesita que $\LL$ sea un fragmento de $\fo$}
+\martin{para este resultado se necesita que $\LL$ sea un fragmento de $\fo$}
 Therefore, the desired extension has to be achieved by allowing some local extensions to $\Sigma_1$.

### 2017-01-05 LICS/syntactic-clo.tex
Nueva versión de sección 5 y comienzo de la Intro.

--- 2016-12-31 LICS/syntactic-clo.tex
@@
 
+We use the \emph{$\eqso(\fo)$-hierarchy} to search for syntactic subclasses with good algorithmic and closure properties.  
+Before starting this search, we introduce a more strict notion of counting problems with easy decision versions.
 Recall that a function $f : \Sigma^* \to \mathbb{N}$ has an easy decision counterpart if $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$ is a language in $\ptime$.
 As the goal of this section is to define a syntactic class of functions in $\shp$ with easy decision versions and good closure properties, we do not directly consider the semantic condition $L_f  \in \ptime$, but instead we consider a more restricted class of functions that has a simple syntactic definition.
-More precisely, a function $f : \Sigma^* \to \mathbb{N}$ is said to be in the complexity class $\totp$ \cite{PagourtzisZ06} if there exists a nondeterministic polynomial-time Turing Machine $M$ such that $f(x) = \tmt_M(x) - 1$ for every $x \in \Sigma^*$, where $\tmt_M(x)$ is the total number of runs of $M$ with input $x$.
+More precisely, a function $f : \Sigma^* \to \mathbb{N}$ is said to be in the complexity class $\totp$~\cite{PagourtzisZ06} if there exists a nondeterministic polynomial-time Turing Machine $M$ such that $f(x) = \tmt_M(x) - 1$ for every $x \in \Sigma^*$, where $\tmt_M(x)$ is the total number of runs of $M$ with input $x$.
 Notice that one is subtracted from $\tmt_M(x)$ to allow for $f(x) = 0$.
@@
 
-The complexity class $\totp$ contains many important counting problems with easy decision counterparts, such as $\cpm$, $\ctwcnf$, $\cdnf$, $\chsat$ and $\csp$.
+The complexity class $\totp$ contains many important counting problems with easy decision counterparts, such as $\cpm$, $\cdnf$, and $\chsat$ among others~\cite{PagourtzisZ06}.
 Besides, $\totp$ has good closure properties as it is closed under sum, multiplication and subtraction by one.
@@
 This counting problem is in $\totp$, and it is known that $\np = \rp$ if there exists an FPRAS for it \cite{DFJ02}.} and no natural complete problems are known for this class \cite{PagourtzisZ06}.
-Hence, we use the $\eqso(\fo)$-hierarchy to find restrictions of $\totp$ with good closure and approximation properties.
+Hence, we use the $\eqso(\fo)$-hierarchy to find restrictions of $\totp$ with good approximation and closure properties.
 
@@
 
-It was proved in \cite{SalujaST95} that every function in $\E{1}$ admits an FPRAS.
-Besides, it can be proved that $\E{1} \subseteq \totp$.
-However, this class is not closed under sum as shown in the following proposition.
+It was proved in \cite{SalujaST95} that every function in $\E{1}$ admits an FPRAS. Besides, it can be proved that $\E{1} \subseteq \totp$. 
+However, we have to discard this class since it is not closed under sum and then is not robust under basic closure properties. 
 \begin{proposition}\label{prop-e1-nc}
@@
 To overcome this limitation, one can consider the class $\QE{1}$, which is closed by under sum by definition.
-In fact, the following proposition shows that the same good properties as for $\E{1}$ hold for $\QE{1}$.
+In fact, the following proposition shows that the same good properties as for $\E{1}$ hold for $\QE{1}$ plus being closed under sum and multiplication.
 \begin{proposition} \label{prop:qe0-fp-qe1-totp-fptras}
-$\QE{1} \subseteq \totp$ and every function in $\QE{1}$ has an FPRAS.
+$\QE{1} \subseteq \totp$ and every function in $\QE{1}$ has an FPRAS. Moreover, $\QE{1}$ is closed under sum and multiplication.
 \end{proposition}
-Moreover, from the following theorem one can conclude that $\QE{1}$ is closed under multiplication.
-\begin{theorem}\label{theo:binary-prod}
-	If $\LL$ is closed under conjunction, then $\eqso(\LL)$ is closed under multiplication.
-\end{theorem}
+%Moreover, from the following theorem one can conclude that $\QE{1}$ is closed under multiplication.
+%\begin{theorem}\label{theo:binary-prod}
+%	If $\LL$ is closed under conjunction, then $\eqso(\LL)$ is closed under multiplication.
+%\end{theorem}
 Hence, it only remains to prove that $\QE{1}$ is closed under subtraction by one.
@@
 Thus, we need to find an extension of $\QE{1}$ that keeps all the previous properties and is closed under subtraction by one.
-It is important to notice that $\shp$ is believed not to be closed under subtraction by one by some complexity-theoretical assumption,\footnote{A decision problem $L$ is in the randomized complexity class $\cspp$ if there exists a nondeterministic polynomial-time Turing Machine $M$ such that for every $x \in L$ it holds that $\tma_M(x) - \tmr_M(x) = 2$, and for every $x \not\in L$ it holds that $\tma_M(x) = \tmr_M(x)$ \cite{OH93,FFK94}.
+It is important to notice that $\shp$ is believed not to be closed under subtraction by one by some complexity-theoretical assumption\footnote{A decision problem $L$ is in the randomized complexity class $\cspp$ if there exists a nondeterministic polynomial-time Turing Machine $M$ such that for every $x \in L$ it holds that $\tma_M(x) - \tmr_M(x) = 2$, and for every $x \not\in L$ it holds that $\tma_M(x) = \tmr_M(x)$ \cite{OH93,FFK94}.
 It is believed that $\np \not\subseteq \cspp$, as there is no evidence that a Turing Machine with the property just described exists for the propositional satisfiability problem.
-However, if $\shp$ is closed under subtraction by one, then it holds that $\np \subseteq \cspp$ \cite{OH93}.} so the following proposition rules out any logic that extends $\Pi_1$.
+However, if $\shp$ is closed under subtraction by one, then it holds that $\np \subseteq \cspp$ \cite{OH93}.}.
+So, the following proposition rules out any logic that extends $\Pi_1$ for a possible extension of~$\QE{1}$.
 \begin{proposition}
-If $\Pi_1 \subseteq \LL$ and $\eqso(\LL)$ is closed under subtraction by one, then $\shp$ is closed under subtraction by one. 
+If $\Pi_1 \subseteq \LL$ and $\eqso(\LL)$ is closed under subtraction by one, then $\shp$ is closed under subtraction by~one. 
 \end{proposition}
@@
 More precisely, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulas over a relational signature $\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mentioned any second-order variable).
-With this extension we have that:
+With this extension we find the promised class.
 \begin{theorem}\label{sigmafo-minusone}
-The class $\eqso(\logex{1})$ is closed under sum, product and subtraction by one. Moreover, $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
+The class $\eqso(\logex{1})$ is closed under sum, multiplication and subtraction by one. Moreover, $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
 \end{theorem}
@@
 \end{theorem}
-
-\marcelo{La version anterior de este teorema decia lo siguiente: If $\eqso(\loge{i})$ or $\eqso(\logu{i})$ is closed under subtraction for $i > 0$, then $\ptime = \np$.
-La razon para considerar esta version mas general era un comentario de Cristian que decia que el resultado de Martin en el apendice se generaliza trivialmente para todas las clases ya que todas contienen la clase $\eqso(\loge{0})$.
-Me parece que esto es incorrecto ya que tambien se necesita que cada funcion en la clase tenga un problema de decision asociado en $\ptime$, lo que no es cierto para $\eqso(\logu{1})$ y las clases de mas arriba.
-Estas clases no deberian ser cerradas bajo resta pero por una razon distinta (por la misma razon por la que se cree que $\shp$ no es cerrado bajo resta), pero esto no me parece que valga la pena mencionarlo.}
 

### 2017-01-05 LICS/syntactic-horn.tex
Nueva versión de sección 5 y comienzo de la Intro.

@@
 \begin{align*}
-&\forall x \, (\pT(x) \to \pV(x)) \ \wedge\\
+\varphi(\pT) \; :=\;\;  &\forall x \, (\pT(x) \to \pV(x)) \ \wedge\\
 &\forall c \, (\pNC(c) \to \exists x \, (\pN(c,x) \wedge \neg \pT(x))) \ \wedge\\
@@
 \end{align*}
-we can translate $\varphi(\pT)$ into the following equivalent formula $\psi(\pT,\pA)$:
+we can translate $\varphi(\pT)$ into the following equivalent formula:
 \begin{align*}
-&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
+\psi(\pT,\pA) \; :=\;\;  &\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
 &\forall c \, (\neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
@@
 
-We have that $\chsat \in \eqso(\uhorn)$. Moreover, the following proposition shows that $\eqso(\uhorn)$ forms a class of functions with easy decision counterparts.
-
-\begin{proposition}\label{prop:uhorn-pe}
-$\eqso(\uhorn) \subseteq \totp$
-\end{proposition}
-
-Thus, $\eqso(\uhorn)$ is a new alternative in our search for a class of functions in $\shp$ with easy decision counterparts and natural complete problems.
+We have that $\chsat \in \eqso(\uhorn)$. Moreover, one can show that $\eqso(\uhorn)$ forms a class of functions with easy decision counterparts, namely, $\eqso(\uhorn) \subseteq \totp$.
+Thus, $\eqso(\uhorn)$ is a new alternative in our search for functions in $\shp$ with easy decision counterparts and natural complete problems.
 Moreover, an even larger class for our search can be generated by extending the definition of $\uhorn$ with outermost existential quantification.
@@
 An imperative question at this point is whether in the definitions of $\uhorn$ and $\ehorn$, it is necessary to allow negative literals of the form $\exists \v \, \neg X(\u,\v)$.
-The following result shows that it is indeed the case:
+Actually, this forces our Horn classes to be included in $\eqso(\logu{2})$ and not necessarily in $\eqso(\loge{2})$.
+The following result shows that this is indeed the case.
 
@@
 We conclude this section by showing that $\eqso(\ehorn)$ is the class we were looking for, as not only every function in $\eqso(\ehorn)$ has an easy decision counterpart, but also $\eqso(\ehorn)$ admits a natural complete problem under parsimonious reductions.
-More precisely, define $\dhsat$ as the decision problem:
+More precisely, define $\dhsat$ as:
 \begin{multline*}
-\{\Phi \mid \Phi \text{ is a disjunction of}\\  \text{Horn formulas and $\Phi$ is satisfiable}\},
+\{\Phi \mid \Phi \text{ is a disjunction of Horn formulas and satisfiable}\},
 \end{multline*}
+%\begin{multline*}
+%\{\Phi \mid \Phi \text{ is a disjunction of}\\  \text{Horn formulas and $\Phi$ is satisfiable}\},
+%\end{multline*}
 and define $\shdhsat$ as the problem of counting the satisfying assignments of a formula $\Phi$ that is a disjunction of Horn formulas. Then we have that:

### 2017-01-05 LICS/appendix-syntactic.tex

@@
 $$
-Phi = p \wedge \bigwedge_{i = 1}^n (t_i \vee p \to q) \wedge \neg q,
+\Phi = p \wedge \bigwedge_{i = 1}^n (t_i \wedge p \to q) \wedge \neg q,
 $$
@@
 $$
-Phi' = p \wedge \bigwedge_{i = 1}^{\ell-1} (t_i \vee p \to q) \wedge (p\to q) \wedge \bigwedge_{i = \ell+1}^{n} (t_i \vee p \to q) \wedge \neg q,
+\Phi' = p \wedge \bigwedge_{i = 1}^{\ell-1} (t_i \wedge p \to q) \wedge (p\to q) \wedge \bigwedge_{i = \ell+1}^{n} (t_i \wedge p \to q) \wedge \neg q,
 $$

### 2017-01-05 LICS/syntactic-clo.tex

@@
 \begin{proposition}
-If $\Pi_1 \subseteq \LL$ and $\eqso(\LL)$ is closed under subtraction by one, then $\shp$ is closed under subtraction by~one. 
+If $\Pi_1 \subseteq \LL \subseteq \fo$ and $\eqso(\LL)$ is closed under subtraction by one, then $\shp$ is closed under subtraction by~one. 
 \end{proposition}

### 2017-01-05 LICS/saluja.tex

@@
-%!TEX root = syntactic.tex
+%!TEX root = main.tex
 
@@
 a formula $\varphi(\bar{X}, \bar{x})$ in $\sh{\LL}$ is equivalent to the formula $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ in $\LL$-PNF. 
-Moreover, a formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-sum normal form ($\LL$-SNF)} if $\alpha$ is of the form $c + \Sigma_{i=1}^n \alpha_i$, where $c$ is a non-negative constant and each $\alpha_i$ is in
-$\LL$-PNF.
-\begin{theorem}\label{theo-pnf-snf}
+Moreover, a formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-sum normal form ($\LL$-SNF)} if $\alpha$ is of the form $\Sigma_{i=1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-PNF. 
+\begin{proposition}\label{theo-pnf-snf}
 Every formula in $\eqso(\LL)$ can be rewritten in $\LL$-SNF.
-\end{theorem}
+\end{proposition}
+If a formula is in $\LL$-PNF then clearly the formula is in $\LL$-SNF.
+Unfortunately, for some $\LL$ there exist formulas in $\eqso(\LL)$  that cannot be rewritten in $\LL$-PNF.
 Therefore, to unveil the relationship between the $\sfo$-hierarchy and the $\eqso(\fo)$-hierarchy, we need to understand the boundary between PNF and SNF. We do this in the following theorem. 
@@
 
-As a consequence of Theorems \ref{theo-pnf-snf} and \ref{theo-pi1-pnf}, we obtain that $\E{i} \subsetneq \QE{i}$ for $i = 0,1$, and that $\sh{\LL} = \eqso(\LL)$ for $\LL$ equal to  $\Pi_1$, $\Sigma_2$ or $\Pi_2$.
+As a consequence of Proposition~\ref{theo-pnf-snf} and Theorem~\ref{theo-pi1-pnf}, we obtain that $\E{i} \subsetneq \QE{i}$ for $i = 0,1$, and that $\sh{\LL} = \eqso(\LL)$ for $\LL$ equal to  $\Pi_1$, $\Sigma_2$ or $\Pi_2$.
 The following proposition completes our picture of the relationship between the $\sfo$-hierarchy and the $\eqso(\fo)$-hierarchy.
@@
 \end{proposition}
-The results of this section are summarized in Figure \ref{fig-sfo-eqso}.
+The relationship between both hierarchies is summarized in Figure \ref{fig-sfo-eqso}.
+One can see there that the difference between our hierarchy and the one proposed in~\cite{SalujaST95} only differ in~$\Sigma_0$ and~$\Sigma_1$. 
+Actually, we show next that this difference is very crucial for finding classes of functions with easy decision versions and good closure properties.

### 2017-01-05 LICS/syntactic.tex

@@
 
-The class $\shp$ was introduced in \cite{Valiant79} to prove that computing the permanent of a matrix, as defined in Example \ref{exa-perm}, is a difficult problem.
+The class $\shp$ was introduced in \cite{Valiant79} to prove that computing the permanent of a matrix, as defined in Example \ref{exa-perm}, is a $\shp$-complete problem.
+As a consequence of this result many counting problems have been proved to be $\shp$-complete \cite{V79b,PB83,P86,L86,BW91,HMRS98,DG00,BW05,DS12, PS13,PS14}.
+Among them, problems having easy decision counterparts play a fundamental role, as a counting problem with a hard decision version is expected to be hard.
+Formally, the decision problem associated to a function $f : \Sigma^* \to \mathbb{N}$ is defined as $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$, and $f$ is said to have an easy decision version if $L_f \in \ptime$.
+Many prominent examples satisfy this property like computing the number of: perfect matchings of a bipartite graph ($\cpm$)~\cite{Valiant79}, satisfying assignments of a DNF propositional formula ($\cdnf$)~\cite{DHK05}, satisfying assignments of a Horn  propositional formula ($\chsat$)~\cite{V79b}, among others.
+
+Counting problems with easy decision versions play a fundamental role in the search of efficient approximations algorithms for functions in $\shp$.
+A fully-polynomial randomized approximation scheme (FPRAS) for a function $f : \Sigma^* \to \bbN$ is a randomized algorithm ${\cal A} : \Sigma^* \times (0,1) \to \bbN$ such that: (1) for every string $x \in \Sigma^*$ and real value $\varepsilon \in (0,1)$, the probability that $|f(x) - {\cal A}(x,\varepsilon)| \leq \varepsilon \cdot f(x)$ is at least $\frac{3}{4}$, and (2) the running time of ${\cal A}$ is polynomial in the size of $x$ and $1/\varepsilon$ \cite{KL83}.
+Notably, there exist $\shp$-complete functions that can be efficiently approximated as they admit FPRAS; for instance, there exist FPRAS for $\cdnf$~\cite{KL83} and $\cpm$~\cite{JSV04}.
+A key observation here is that if a function $f$ admits an FPRAS, then $L_f$ is in the randomized complexity class $\bpp$~\cite{G77}.
+Hence, under the widely believed assumption that  $\np \not\subseteq \bpp$, we cannot hope for an FPRAS for a function in $\shp$ whose decision counterpart is $\np$-complete, and we have to concentrate on the class of counting problems with easy decision versions.
+
+The importance of the class of counting problems with easy decision counterparts has motivated the search of robust classes of functions in $\shp$ with this property \cite{PagourtzisZ06}.
+But the key question here is what should be considered a {\em robust} class.
+A first desirable condition has to do with the closure properties satisfied by the class, which is a common theme when studying function complexity classes \cite{OH93,FH08}.
+As in the cases of $\ptime$ and $\np$ that are closed under intersection and union, we expect our class to be closed under multiplication and sum.
+For a more elaborated closure property, assume that $\textit{sat\_one}$ is a function that returns one plus the number of satisfying assignments of a propositional formula.
+Clearly $\textit{sat\_one}$ is a $\shp$-complete function whose decision counterpart $L_{\textit{sat\_one}}$ is trivial.
+But should $\textit{sat\_one}$ be part of a robust class of counting function with easy decision versions?
+The key insight here is that if a function in $\shp$ has an easy decision counterpart $L$, then as $L \in \np$ we expect  to have a polynomial-time algorithm that verifies whether $x \in L$ by constructing witnesses for~$x$.
+%As an example of this considered all the counting functions mentioned in the previous paragraphs and, in particular, the decision counterpart of $\cdnf$ that can be solved by explicitly constructing in polynomial-time a satisfying assignment for an input propositional formula.
+Moreover, if such an algorithm for constructing witnesses exists, then we also expect to be able to manipulate such witnesses and in some cases to remove them.
+In other words, we expect a robust class $\CC$ of counting functions with easy decision versions to be closed under subtraction by one, that is, if $g \in \CC$, then the function $g \dotdiv 1$ should also be in $\CC$, where $(g \dotdiv 1)(x)$ is defined as $g(x) - 1$ if $g(x) \geq 1$, and as~$0$ otherwise.
+Notice that, unless $\ptime = \np$, no such class can contain the function $\textit{sat\_one}$ because $\textit{sat\_one} \dotdiv 1$ counts the number of satisfying assignments of a propositional formula.
+
+A second desirable condition of robustness is the existence of natural complete problems \cite{P94}.
+Special attention has to be paid here to the notion of reduction used for completeness.
+Notice that under the notion of Cook reduction, originally used in \cite{Valiant79}, the problems $\cdnf$ and $\csat$ are $\shp$-complete.
+However, $\cdnf$ has an easy decision counterpart and admits an FPRAS, while $\csat$ does not satisfy these conditions unless $\ptime = \np$.
+Hence a more strict notion of reduction has to be considered; in particular, the notion of parsimonious reduction (to be defined later) satisfy that if a function $f$ is parsimoniously reducible to a function $g$, then $L_g \in \ptime$ implies that $L_f \in \ptime$ and the existence of an FPRAS for $g$ implies the existence of a FPRAS for $f$.
+
+In this section, we use the framework developed in this paper to address the problem of defining a robust class of functions with easy decision versions.
+More specifically, we use the framework to introduce in Section \ref{sec-hier-shp} a syntactic hierarchy of function complexity classes contained in $\shp$.
+Then this hierarchy is used in Section \ref{sec-clo} to define a class of functions with easy decision versions and good closure properties, and in Section \ref{sec-horn} to define a class of functions with easy decision versions and
+natural complete problems.
+
+%The class $\shp$ was introduced in \cite{Valiant79} to prove that computing the permanent of a matrix, as defined in Example \ref{exa-perm}, is a difficult problem.
 More specifically, it was shown in  \cite{Valiant79}  that this problem is $\shp$-complete.
 As a consequence of this result the $\shp$-completeness of the problem of computing the number of perfect matchings of a bipartite graph ($\cpm$) was established.
-Since then  many counting problems have been proved to be $\shp$-complete \cite{V79b,PB83,P86,L86,BW91,HMRS98,DG00,BW05,DS12, PS13,PS14}.
+%Since then  many counting problems have been proved to be $\shp$-complete \cite{V79b,PB83,P86,L86,BW91,HMRS98,DG00,BW05,DS12, PS13,PS14}.
 Among them, problems having easy decision counterparts play a fundamental role, as a counting problem with a hard decision version is expected to be hard.
@@
 Other prominent examples of such problems include counting the number of: satisfying assignments of a 2-CNF propositional formula ($\ctwcnf$) \cite{V79b}, satisfying assignments of a DNF propositional formula ($\cdnf$) \cite{DHK05}, satisfying assignments of a Horn  propositional formula ($\chsat$) \cite{V79b}, simple paths from a source node to a target node in a directed graph ($\csp$) \cite{V79b}, extensions of a partial order to a linear order \cite{BW91} and Eulerian cycles in an undirected graph \cite{BW05}.
-
-Counting problems with easy decision versions play a fundamental role in the search of efficient approximations algorithms for functions in $\shp$.
+%
+%Counting problems with easy decision versions play a fundamental role in the search of efficient approximations algorithms for functions in $\shp$.
 A fully-polynomial randomized approximation scheme (FPRAS) for a function $f : \Sigma^* \to \bbN$ is a randomized algorithm ${\cal A} : \Sigma^* \times (0,1) \to \bbN$ such that: (1) for every string $x \in \Sigma^*$ and real value $\varepsilon \in (0,1)$, the probability that $|f(x) - {\cal A}(x,\varepsilon)| \leq \varepsilon \cdot f(x)$ is at least $\frac{3}{4}$, and (2) the running time of ${\cal A}$ is polynomial in the size of $x$ and $1/\varepsilon$ \cite{KL83}.
@@
 Hence, under the widely believed assumption that  $\np \not\subseteq \bpp$, we cannot hope for an FPRAS for a function in $\shp$ whose decision counterpart is $\np$-complete, and we have to concentrate on the class of counting problems with easy decision versions (in $\bpp$ or in a lower complexity class such as $\ptime$).
-
-The importance of the class counting problems with easy decision counterparts has motivated the search of robust classes of functions in $\shp$ with this property \cite{PagourtzisZ06}.
+%
+%The importance of the class counting problems with easy decision counterparts has motivated the search of robust classes of functions in $\shp$ with this property \cite{PagourtzisZ06}.
 But the key question here is what should be considered a {\em robust} class.
@@
 Hence a more strict notion of reduction has to be considered; in particular, the notion of parsimonious reduction (to be defined later) satisfy that if a function $f$ is parsimoniously reducible to a function $g$, then $L_g \in \ptime$ implies that $L_f \in \ptime$ and the existence of an FPRAS for $g$ implies the existence of a FPRAS for $f$.
-A second desirable condition has to do with the closure properties satisfied by the class, which is a common theme when studying function complexity classes \cite{OH93,FH08}.
+%A second desirable condition has to do with the closure properties satisfied by the class, which is a common theme when studying function complexity classes \cite{OH93,FH08}.
 As in the cases of $\ptime$ and $\np$ that are closed under intersection and union, we expect our class to be closed under multiplication and sum; in fact, if $f$ and $g$ are counting functions with easy decision versions, then we clearly expect $f + g$ and $f \cdot g$ to have easy decision counterparts.
@@
 Notice that, unless $\ptime = \np$, no such class can contain the function $\textit{sat\_one}$ because $\textit{sat\_one} \dotdiv 1$ counts the number of satisfying assignments of a propositional formula.
-
-In this section, we use the framework developed in this paper to address the problem of defining a robust class of functions with easy decision versions.
+%
+%In this section, we use the framework developed in this paper to address the problem of defining a robust class of functions with easy decision versions.
 More specifically, we use the framework to introduce in Section \ref{sec-hier-shp} a syntactic hierarchy of function complexity classes contained in $\shp$.
 Then this hierarchy is used in Section \ref{sec-clo} to define a class of functions with easy decision versions and good closure properties, and in Section \ref{sec-horn} to define a class of functions with easy decision versions and
-natural complete problems.
+%natural complete problems.
 
@@
 
-\subsection{A syntactic hierarchy of function complexity classes contained in $\shp$}
+%\subsection{A syntactic hierarchy of function complexity classes contained in $\shp$}
+\subsection{The $\eqso(\fo)$-hierarchy inside $\shp$}
 \label{sec-hier-shp}

### 2017-01-05 LICS/appendix-syntactic.tex
05 ene

@@
 where $\varphi$ is a formula in $\LL$ and $s\in\nat$.
-We will show a recursive function $\tau$ such that for every $\eqso(\LL)$ formula $\alpha$, $\tau(\alpha)$ is a formula in $\LL$-SNF that describes the same function as $\alpha$.
-If $\alpha = \varphi$ or $\alpha = s$, then we define $\tau(\alpha) = \alpha$ which clearly describes the same function, and is already in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$, then $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$ which satisfies the condition.
-If $\alpha = \sa{x}\beta$, then let $\tau(\beta) = c + \sum_{i = 1}^{n}\beta_i$ where each $\beta_i$ is in $\LL$-PNF.
-First, if $c > 0$, we replace $c$ with the formula $(\top \add \cdots \add \top)$ (c times) and we obtain a formula $\sum_{i = 1}^{m}\beta_i$ where each $\beta_i$ is in $\LL$-PNF.
-If $c = 0$ the formula is already in that form.
-Since $\beta_i$ is in $\LL$-PNF, then $\sa{x}\beta_i$ also is, and then we define $\tau(\alpha) = \sum_{i = 1}^{m}\sa{x}\beta_i$ which satisfies the condition.
+We will construct a recursive function $\tau$ such that for every $\eqso(\LL)$ formula $\alpha$, it outputs an equivalent formula $\tau(\alpha)$ which is in $\LL$-SNF.
+If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$ which is clearly equivalent and already in $\LL$-SNF.
+If $\alpha = s$, then we define $(\top \add \cdots \add \top)$ ($s$ times), which also satisfies the condition.
+We assume that for every sub-formula $\beta$ of $\alpha$, $\tau(\beta)$ is an equivalent formula in $\LL$-SNF.
+If $\alpha = (\alpha_1 + \alpha_2)$, then we define $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$.
+If $\alpha = \sa{x}\beta$, then let $\tau(\beta) = \sum_{i = 1}^{n}\beta_i$ where each $\beta_i$ is in $\LL$-PNF.
+We have that each $\beta_i$ also is in $\LL$-PNF, and then we define $\tau(\alpha) = \sum_{i = 1}^{m}\sa{x}\beta_i$ which satisfies the condition.
 If $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case.
@@
 As we showed, there exists an equivalent formula in $\LL$-SNF, so we assume that $\alpha$ is in that form.
-Let $\alpha = c + \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-SNF.
-First, as we did in Theorem \ref{theo-pnf-snf}, we replace the constant $c$ for $(\top \add \cdots \add \top)$ if $c > 0$ and we do nothing if $c = 0$, to obtain an equivalent formula $\sum_{i = 1}^m \alpha_i$.
+Let $\alpha = \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-SNF.
 Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$.
@@
 \end{proof}
-We suppose that $\alpha$ is in SNF, this is, $\alpha = c + \sum_{i = 1}^n\alpha_i$.
-Let $\mathbb{1}$ be a $\R$-structure with one element.
-Note that if $c > 0$, then $\alpha(\enc(\mathbb{1})) \geq c > 0$ which is not possible since $\alpha(\enc(\mathbb{1})) = 0$.
+We suppose that $\alpha$ is in SNF, this is, $\alpha = \sum_{i = 1}^n\alpha_i$.
 Since $\alpha$ is not null, consider some $\alpha_i$ that describes a non-null function.
@@
 
-We separate this proof in two parts
+We separate this proof in three parts
 
@@
 
-\subsection*{Proof of Theorem \ref{theo:binary-prod}}
+\vspace{2em}
+For the third part, we will prove that $\eqso(\loge{1})$ is closed under sum and multiplication.
+Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that the class is closed under multiplication.
+We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of $\so$.
 
-Let $\LL$ be a fragment of $\so$ that is closed under conjunction.
-We will define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ with multiplication, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
-Furthermore, the formula $\tau(\alpha)$ is in $\LL$-SNF.
+Let $\LL$ be a fragment {\bf closed under conjunction}.
+We will define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ extended by binary product, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
+In fact, the formula $\tau(\alpha)$ is in $\LL$-SNF.
 First we replace each constant $s$ in $\alpha$ for $(\top \add \cdot \add \top)$ ($s$ times).
@@
 This covers all possible cases for $\alpha$.
-For every formula $\alpha$ in $\eqso(\LL)$ extended by multiplication, we take $\tau(\alpha)$ which is equivalent and in unextended $\eqso(\LL)$.
+
+For every pair of formulas $\alpha,\beta$ in $\eqso(\LL)$, we have that their multiplication $(\alpha\cdot\beta)$ is a formula in the grammar $\eqso(\LL)$ extended by binary product, and so, there exists an equivalent formula $\tau(\alpha\cdot\beta)$ which is in unextended $\eqso(\LL)$.
+Since $\loge{1}$ is closed under conjunction, this also holds for $\eqso(\loge{1})$.
 This concludes the proof.
@@
 
-\subsection*{Proof of Proposition \ref{prop:uhorn-pe}} %V.11
-This can be obtained directly by Proposition \ref{prop:ehorn-pe}, since $\eqso(\uhorn)\subseteq\eqso(\ehorn)\subseteq\totp$.
-
 \subsection*{Proof of Proposition \ref{prop:ehorn-pe}}

### 2017-01-05 LICS/syntactic-clo.tex

--- 2017-01-05 LICS/syntactic-clo.tex
@@
 So, the following proposition rules out any logic that extends $\Pi_1$ for a possible extension of~$\QE{1}$.
-\begin{proposition}
-If $\Pi_1 \subseteq \LL$ and $\eqso(\LL)$ is closed under subtraction by one, then $\shp$ is closed under subtraction by~one. 
+\begin{proposition} \label{pi-minusone}
+If $\Pi_1 \subseteq \LL \subseteq \fo$ and $\eqso(\LL)$ is closed under subtraction by one, then $\shp$ is closed under subtraction by~one. 
 \end{proposition}

### 2017-01-06 LICS/beyond.tex

@@
 %!TEX root = main.tex
+
+We define the extension of $\qso$ with function symbols as follows.
+Assume that $\fs$ is an infinite set of function symbols, where each $h \in \fs$ has an associated arity, which is denoted by $\arity(h)$.
+Then the set of $\fqso$ formulas over a relational signature $\R$ is defined by the following grammar:
+\begin{multline}
+\label{eq-fqso}
+	\alpha := \varphi \ \mid \  s \  \mid \  h(x_1, \ldots, x_\ell) \  \mid \
+	(\alpha \add \alpha) \  \mid\  (\alpha \mult \alpha) \  \mid \\  
+	\sa{x} \alpha \  \mid \
+	\pa{x} \alpha \  \mid \
+	\sa{X} \alpha \  \mid \
+	\pa{X} \alpha,
+\end{multline}
+where $h \in \fs$, $\arity(h) = \ell$ and $x_1, \ldots, x_\ell$ is a sequence of (non-necessarily distinct) first-order variables.
+Given an $\R$-structure $\A$ with domain $A$, we say that $F$ is a function assignment for $\A$ if for every $h \in \fs$ with $\arity(h) = \ell$, we have that $F(h) :  A^\ell \to \N$.
+The notion of function assignment is used to extend the semantics of $\qso$ to the case of a formula of the form $h(x_1, \ldots, x_\ell)$.
+More precisely, given first-order and second-order assignments $v$ and $V$ for $\A$, respectively,
+we have that:
+\begin{eqnarray*}
+\sem{h(x_1, \ldots, x_\ell)}(\A,v,V,F) & = & F(h)(v(x_1),\ldots, v(x_\ell)).
+\end{eqnarray*}
+As for the case of $\qfo$, we define $\fqfo$ by restricting $\varphi$ in \eqref{eq-fqso} to be a first-order formula, and by disallowing quantifiers $\Sigma X$ and $\Pi X$ in \eqref{eq-fqso}.
+
+\subsection{Recursive Logics}
+
+We define an operator which extends Least Fixed Point Logic (LFP) \cite{I86,vardi1982complexity} to allow counting. 
+Fix a relational signature $\R$. Then the set of $\rqfo$ formulas over $\R$ is defined by the following grammar:
+\begin{multline*}
+	\alpha := \varphi \, \mid \, s \, \mid \, (\alpha \add \alpha) \, \mid\, (\alpha \mult \alpha) \, \mid \\
+	\sa{x} \alpha \, \mid\, 
+	\pa{x} \alpha \, \mid\,
+	\clfp{\beta(x_1, \ldots, x_\ell, h)},
+\end{multline*}
+where $x_1, \ldots, x_\ell$ is a sequence of pairwise distinct first-order variables and $\beta(x_1, \ldots, x_\ell, h)$ is an $\fqfo$-formula over $\R$ whose only function symbol is $h$ with $\arity(h) = \ell$.
+The free variables of the formula $\clfp{\beta(x_1,\ldots,x_\ell,h)}$ are $x_1, \ldots, x_\ell$; in particular, $h$ is not considered to be free in this formula.
+
+Fix and $\R$-structure with domain $A$ and a quantitative formula $\clfp{\beta(x_1,\ldots,x_\ell,h)}$, and assume that $\F$ is the set of functions $f :A^\ell \to \N$.
+To define the semantics of $\clfp{\beta(x_1,\ldots,x_\ell,h)}$, we first show how $\beta(x_1,\ldots,x_\ell,h)$ can be interpreted as an operator $T_{\beta}$ on the set $\F$.
+More precisely, for every $f \in \F$ and tuple $(a_1, \ldots, a_\ell) \in A^\ell$, the function $T_{\beta}(f)$ satisfies that:
+\begin{eqnarray*}
+T_{\beta}(f)(a_1, \ldots, a_\ell) & = & \sem{\beta(x_1, \ldots, x_\ell, h)}(\A,v,F),
+\end{eqnarray*}
+where $v$ is a first-order assignment  for $\A$ such that $v(x_i) = a_i$ for every $i \in \{1, \ldots, \ell\}$, and $F$ is a function assignment for $\A$ such that $F(h) = f$.
+
+As for the case of LFP, it would be natural to consider the partial order $\leq$ on $\F$ defined as $f \leq g$ if and only if $f(i) \leq g(i)$ for every $i \in \{1, \ldots, \ell\}$, and let the semantics of $\clfp{\beta(x_1,\ldots,x_\ell,h)}$ be the least fixed point of the operator $T_\beta$.
+However, $(\F, \leq)$ is not a complete lattice, so we do not have a Knaster-Tarski Theorem ensuring that such a fixed point exist.
+Instead, we generalize the semantics of LFP in the following way.
+Given a relation $R \subseteq A^\ell$, let $\chi_R$ be its characteristic function, that is $\chi_R(\bar a) = 1$ if $\bar a \in R$, and $\chi_R(\bar a) = 0$ otherwise.
+Moreover, given a function $f \in \F$, define the support of $f$, denoted by $\support(f)$, as $\{ \bar a \in A^\ell \mid f(\bar a) > 0 \}$.
+Notice that for a relation $R$ we have that $\support(\chi_R) = R$.
+In the definition of the semantics of LFP, an operator $T$ on relations is considered, and the semantics is defined in terms of the least fixed point for $T$, that is, a relation $R$ such that $T(R) = R$ and $R \subseteq S$ for every $S$ such that $T(S) = S$ \cite{I86,vardi1982complexity}.
+We can view the computation of this fixed point in terms of functions if we consider an operator $T^\star$ on characteristic functions defined as $T^\star(\chi_R) = \chi_{T(R)}$.
+In particular, the conditions defining the least fixed point of $T$ can be rewritten as follows in terms of $T^\star$: $\support(T^\star(\chi_R)) = \support(\chi_R)$ and $\support(\chi_R) \subseteq \support(\chi_S)$ for every $S$ such that
+$\support(T^\star(\chi_{S})) = \support(\chi_S)$.
+This condition can be generalized in the obvious way to the case of the operator $T_\beta$: a function $f \in \F$ is a {\em s-fixed point} of $T_{\beta}$ if $\support(T_\beta(f)) = \support(f)$, and $f$ is a {\em least s-fixed point} of $T_{\beta}$ if $f$ is a s-fixed point of $T_\beta$ and for every s-fixed point $g$ of $T_\beta$ it holds that $\support(f) \subseteq \support(g)$.
+The existence of such fixed point is ensured by the following lemma:
+\begin{lemma}\label{lem-support}
+If $f,g \in \F$ and $\support(f) \subseteq \support(g)$, then $\support(T_\beta(f)) \subseteq \support(T_\beta(g))$.
+\end{lemma}
+In fact, as for the case of LFP, this lemma gives us a simple way to compute the least s-fixed point of $T_\beta$.
+Let $f_0 \in \F$ be a function such that $f_0(\bar a) = 0$ for every $\bar a \in A^\ell$, and let function $f_{i+1}$ be defined as $T_\beta(f_i)$ for every $i \in \N$.
+Then there exists $j \geq 0$ such that $\support(f_j) = \support(T_\beta(f_j))$.
+Let $k$ be the smallest natural number such that $\support(f_{k}) = \support(T_\beta(f_k))$.
+We have that $f_k$ is a least s-fixed point of $T_\beta$, which is used to defined the semantics of $\clfp{\beta(x_1, \ldots, x_\ell, h)}$.
+More specifically, for an arbitrary first-order assignment $v$ for $\A$:
+\begin{eqnarray*}
+\sem{\clfp{\beta(x_1, \ldots, x_\ell, h)}}(\A,v) & = & f_{k}(v(x_1), \ldots, v(x_\ell))
+\end{eqnarray*}
+
+\begin{example}
+%As an example, 
+We would like to define a $\rqfo$-formula that, given a directed acyclic graph $G$ with $n$ nodes and a pair of nodes $b$, $c$ in $G$, counts the number of paths of length at most $n$ from $b$ to $c$ in $G$.
+To this end, assume that graphs are encoded using the relational signature $\R = \{ E(\cdot,\cdot) \}$, and then define formula $\alpha(x, y, f)$ as follows:
+$$
+%\alpha(x,y,R,\pi) = 
+%(\neg \exists zR(z))\cdot(x = y) 
+E(x,y) + \sa{z} f(x,z)\cdot E(z,y).
+$$
+We have that $\clfp{\alpha(x,y,f)}$ defines our counting function.
+In fact, assume that $\A$ is an $\R$-structure with $n$ elements in its domain encoding an acyclic directed graph.
+Moreover, assume that $b,c$ are elements of $\A$ and $v$ is a first-order assignment over $\A$ such that $v(x) = b$ and $v(y) = c$.
+Then we have that $\sem{\alpha(x,y,f)}(\A,v)$ is equal to the  number of paths in $\A$ from $b$ to $c$ of length at most $n$.
+
+Assume now that we need to extend our previous counting function to the case of arbitrary directed graphs. To this end, define the following auxiliary $\fo$-formulas:
+\begin{eqnarray*}
+\varphi_{\text{\rm first}}(x) & = & \neg \exists z (z < x)\\
+\varphi_{\text{succ}}(x,y) & = & x < y \wedge \neg \exists z ( x < z \wedge z < y).
+\end{eqnarray*}
+Moreover, define formula $\beta(x, y, u, g)$ as follows:
+\begin{multline*}
+(E(x,y) + \sa{z} g(x,z,u)\cdot E(z,y)) \cdot \varphi_{\text{\rm first}}(u) \ +\\
+\varphi_{\text{succ}}(v,u) \cdot g(x,y,v)
+\end{multline*}
+Then our extended counting function is defined by the following formula:
+$$
+\sa{u} (\varphi_{\text{\rm first}}(u) \wedge \clfp{\beta(x,y,u,g)}.
+$$ 
+In fact, the number of paths of length at most $n$ from a node $x$ to a node $y$ is recursively computed by using the formula $(E(x,y) + \sa{z} g(x,z,u)\cdot E(z,y)) \cdot \varphi_{\text{\rm first}}(u)$, which stores this value in $g(x,y,u)$ with $u$ the first element in the domain.
+The other formula $\varphi_{\text{succ}}(v,u) \cdot g(x,y,v)$ is used as a counter that allows to reach a fixed point in the support of function $g$ in $n$ steps.
+\end{example}
+
+%
+%
+%It is well known that least fixed point logic is contained in second-order logic \cite{L04}. In the following theorem we show that the same holds in our case.
+%\begin{theorem} \label{so-rec}
+%$\rqfo \subseteq \qso$
+%%	Given a positive $\fo$ formula $\varphi(\bar{x},R)$ and a $\qfo$ formula $\alpha(\bar{x})$, there exists a $\qso$ formula $\beta(\bar{x})$ such that $\sem{[\alfp\varphi(\bar{x},R)\mid \alpha(\bar{x},R)](\bar{x})} = \sem{\beta(\bar{x})}$.
+%\end{theorem}
+
+The following is the main result of this section:
+\begin{theorem} \label{rqfo-fo-cap}
+	$\rqfo(\fo)$ captures $\fp$ over the class of ordered structures.
+\end{theorem}
+
+\marcelo{Notese que estoy permitiendo en la formula $\beta(y_1, \ldots, y_\ell,R,\pi)$ tener subformulas $\pa{x} \alpha$. Esta bien el resultado con esto? O tenemos que eliminar estas formulas?}
+
+%Given a relation signature $\R$, the set of recursive $\qfo$ formulas ($\rqfo$-formulas) is defined by the following grammar:
+%%This operator lets us define the set of recursive $\qfo$ formulas over $\R$ ($\rqfo$-formulas) using the following grammar:
+%\begin{multline*}
+%%	\label{eq-def-rqfo}
+%	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid \\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ [\alfp \varphi \mid \alpha]
+%\end{multline*}
+%where $\varphi$ is an $\fo$-formula over $\R$, $s \in \bbN$ and $x \in \fv$.
+%
+%\marcelo{Vamos a permitir anidacion del operador $\alfp$? Esta gramatica lo permite.}
+%
+%\begin{theorem} \label{rqfo-fo-cap}
+%	$\rqfo(\fo)$ captures $\fp$ over the class of ordered structures.
+%\end{theorem}
+%
 
@@
 
-\marcelo{Puede que este equivocado, pero me parece que teniamos una forma de capturar $\shp$ usando el operator ${\bf path}$.
+\marcelo{Deberiamos mostrar como el ejemplo anterior se generaliza para definir $[\pth \psi(\bar x, \bar y)]$ en terminos del operador ${\bf lfp}$.}
+%
+%\marcelo{Puede que este equivocado, pero me parece que teniamos una forma de capturar $\shp$ usando el operator ${\bf path}$.
 Pero no logro recordar como se hacia esto, y me parece que lo que habiamos escrito antes en esta seccion estaba equivocado: ``$\tqso(\fo)$ captures $\shp$ over the class of ordered structures".
@@
 %\end{theorem}
-
-\subsection{Recursive Logics}
-
-We define an operator which extends least fixed point logic \cite{I86,vardi1982complexity} to allow counting. 
-Fix a relational signature $\R$. Then the set of $\rqfo$ formulas over $\R$ is defined by the following grammar:
-\begin{multline*}
-	\alpha := \varphi \, \mid \, s \, \mid \, (\alpha \add \alpha) \, \mid\, (\alpha \mult \alpha) \, \mid \, 
-	\sa{x} \alpha \, \mid\\ 
-	\pa{x} \alpha \, \mid\,
-	\clfp{\psi(x_1,\ldots,x_k,R)}{\beta(y_1, \ldots, y_\ell, R, \pi)},
-\end{multline*}
-where $R$ is second-order variable of arity $k$, $\psi(x_1, \ldots, x_k, R)$ is an $\fo$-formula over $(\R \cup \{R\})$ that is positive on $R$ and $\beta(y_1, \ldots, y_\ell, R, \pi)$ is a $\qfo$-formula over $\R$ including a fresh symbol $\pi$ that represents a function of arity $\ell$, that is, every atomic formula in $\beta(y_1, \ldots, y_\ell, R, \pi)$ is either an $\fo$-formula over $\R$, a constant $s \in \N$ or a term of the form $\pi(u_1, \ldots, u_\ell)$ with $u_1, \ldots, u_\ell$ non-necessarily distinct first-order variables.
-The free variables of the formula $\clfp{\psi(x_1,\ldots,x_k,R)}{\beta(y_1, \ldots, y_\ell, R, \pi)}$
-are $y_1, \ldots, y_\ell$, in particular it does not have any second-order free variable.
-
-We need to introduce some terminology to define the semantics of the formula $\clfp{\psi(x_1,\ldots,x_k,R)}{\beta(y_1, \ldots, y_\ell, R, \pi)}$
-Given a $\qfo$-formula $\gamma$ with $\ell$ first-order free variables and no second-order free variable, define $\beta[\gamma/\pi]$ to be the $\qfo$-formula obtained from $\beta$ by replacing every occurrence of a term $\pi(u_1, \ldots, u_\ell)$ by $\gamma(u_1, \ldots, u_\ell)$.
-Let $\A$ be an $\R$-structure with domain $A$.
-Then for every $a \in A$, let $\varphi_a(x)$ be an $\fo$-formula such that $\A \models \varphi_a(b)$ if and only if $b = a$.
-Formally, assuming that $a$ is the $p$-th element of $A$ according to the linear order $<^\A$, we have that:
-%then $\varphi_a(x) = \forall y(x < y \vee x = y)$
-%for every natural number $i \geq 1$, let $\varphi_i(x)$ be an $\fo$-formula such that $\varphi_i(a)$ holds in an $\R$-structure $\A$ if and only if $a$ is the $i$-th element in the linear order  $<^{\A}$,
-%that is, $\varphi_1(x) = \forall y(x < y \vee x = y)$ and for every $p > 1$:
-\begin{multline*}
-\varphi_a(x) \ = \ \exists x_1 \cdots \exists x_{p-1}\bigg[\bigwedge_{i =1}^{p-2}x_i < x_{i+1} \wedge\,\\  
-x_{p-1} < x  \wedge \forall y(y < x \to \bigvee_{i = 1}^{p-1} y = x_i)\bigg].
-\end{multline*}
-Finally, given a function $f : A^\ell \to \N$, define $\qfo$-formula $\delta_{f,\A}(z_1, \ldots, z_\ell)$ as follows:
-\begin{multline*}
- \mathop{+}_{(a_1,\ldots,a_{\ell})\in A^{\ell}} \ \varphi_{a_1}(z_1) \cdot \ldots \cdot \varphi_{a_{\ell}}(z_{\ell})\cdot f(a_1,\ldots,a_{\ell}), 
-\end{multline*}
-where $z_1, \ldots, z_\ell$ is a sequence of pairwise distinct first-order variables.
-We have that $\delta_{f,\A}(z_1, \ldots, z_\ell)$ encodes $f$ in the structure $\A$ since:
-\begin{eqnarray*}
-\sem{\delta_{f,\A}(z_1, \ldots, z_\ell)}(\A,v)  & = & f(v(z_1), \ldots, v(z_\ell)).
-\end{eqnarray*}
-%Recall that the least fixed point operator is defined by a formula $\psi(x_1,\ldots,x_k,R)$ that is positive on $R$, where $R$  is a predicate of arity $k$.
-For a structure $\A$ with domain $A$, the operator $T_{\varphi}:2^{A^k} \to 2^{A^k}$ is defined as $T_{\varphi}(X) = \{(a_1,\ldots,a_k)\mid (\A,X)\models \psi(a_1,\ldots,a_k,R) \}$, for each $X\subseteq A^k$.
-Let $T_0 = \emptyset$ and $T_{i+1} = T_{\varphi}(T_i)$ for each $i \in \nat$.
-Note that there exists $n\in \nat$ such that $T_{n+1} = T_n$ because $R$ is positive in $\psi(x_1,\ldots,x_k,R)$.
-Then the evaluation of $[\lfpop\psi(x_1,\ldots,x_k,R)]$ is defined by the fixed point $T_n$, that is, for every $(a_1,\ldots,a_k)\in A^k$, it holds that $\A\models[\lfpop\psi(x_1,\ldots,x_k,R)](a_1,\ldots,a_k)$ if and only if $(a_1,\ldots,a_k) \in T_n$.
-We are now ready to define the semantics of the formula $\clfp{\psi(x_1,\ldots,x_k,R)}{\beta(y_1,\ldots,y_{\ell},R,\pi)}$,  whose free variables are $y_1$, $\ldots$, $y_\ell$.
-Assume that $\A$ is an $\R$-structure with domain $A$.
-Following the definition of the semantics of least fixed point logic \cite{I86,vardi1982complexity}, define an operator $T_{\varphi}:2^{A^k} \to 2^{A^k}$ such that:
-$$
-T_{\varphi}(X)  =  \{(a_1,\ldots,a_k) \in A^k \mid \A \models \psi(a_1,\ldots,a_k,X) \},
-$$
-for every $X \subseteq A^k$. This operator is used to defined 
-a sequence $T_0 \subsetneq T_1 \subsetneq \cdots \subsetneq T_n \subseteq A^k$  such that $n \geq 0$, $T_0 = \emptyset$, $T_{i+1} = T_{\varphi}(T_i)$ for every $i \in \{0, \ldots, n-1\}$, and $T_n = T_\varphi(T_n)$.
-We know that such a sequence exists as formula $\psi(x_1,\ldots,x_k,R)$ is positive on $R$.
-Now, the sequence $T_0 \subsetneq T_1 \subsetneq \cdots \subsetneq T_n$ is used to defined  a sequence of functions $f_0,f_1,\ldots,f_n:A^{\ell}\to\nat$.
-Formally, we have that $f_0(a_1, \ldots, a_\ell) = 0$ for every $(a_1, \ldots, a_\ell) \in A^\ell$.
-Moreover, assuming that $(a_1, \ldots, a_\ell) \in A^\ell$, $v$ is a first-order assignment for $\A$ such that $v(y_i) = a_i$ for every $i \in \{1, \ldots, \ell\}$, and that $V$ is a second-order assignment for $\A$ such that $V(R) = T_{i+1}$, we have that:
-\begin{eqnarray*}
-f_{i+1}(a_1, \ldots, a_\ell) & = & \sem{\beta[\delta_{f_i,\A}/\pi]}(\A,v,V).
-\end{eqnarray*}
-%, for each $T_i \in \T$, let $V$ be a second-order assignment for $\A$ that assigns $T_i$ to $R$, let $\zeta_i: A^{\ell}\to\nat$ be such that for each $(a_1,\ldots,a_{\ell})\in A^{\ell}$ it holds $\zeta_i(a_1,\ldots,a_{\ell}) = \sem{\alpha\mid_{\pi(u_1,\ldots,u_{\ell})\to \beta_i(u_1,\ldots,u_{\ell})}}(\A,v,V)$, where $v$ is a first-order assignment for $\A$ that satisfies $a_i = v(x_i)$ for each free $x_i$ in $\alpha$.
-Finally, the semantics of the formula $\clfp{\psi(x_1,\ldots,x_k,R)}{\beta(y_1,\ldots,y_{\ell},R,\pi)}$ is defined by means of function $f_n$:
-%For a given first order assignment $v$ and a second order assignment $V$, let $a_i = v(x_i)$, the operator is then evaluated as:
-\begin{align*}
-&\llbracket [{\bf lfp} \, \psi(x_1,\ldots,x_k,R) \,|\\
-&\hspace{10mm}\beta(y_1,\ldots,y_{\ell},R,\pi)]\rrbracket(\A,v) \ = \
-f_n(v(y_1),\ldots,v(y_{\ell})).
-\end{align*}
-
-\begin{example}
-%As an example, 
-We would like to define a formula that, given a graph $G$ with $n$ nodes and a pair of nodes $b$, $c$ in $G$, counts the number of paths of length at most $n$ from $b$ to $c$ in $G$.
-To this end, assume that $\R = \{ E(\cdot,\cdot) \}$, and define $\psi(x,R)$ as follows:
-\begin{eqnarray*}
-%\psi(x,R) = 
-\forall y(x < y \vee x = y) \vee \exists z(R(z) \wedge \varphi_{succ}(z,x)),
-\end{eqnarray*}
-where $\varphi_{succ}(x,y)$ is a formula that is satisfied by pairs $(x,y)$ that are consecutive in the order $<$.
-That is, $\varphi_{succ}(x,y) = x < y \wedge \neg \exists z \, (x < z \wedge z < y)$.
-Moreover, define formula $\beta(x,y,R,\pi)$ as follows:
-$$
-%\alpha(x,y,R,\pi) = 
-%(\neg \exists zR(z))\cdot(x = y) 
-E(x,y) + \sa{z} [\pi(x,z)\cdot E(z,y)].
-$$
-Then we have that $\clfp{\psi(x,R)}{\beta(x,y,R,\pi)}$ defines our counting function.
-In fact, assume that $\A$ is an $\R$-structure with $n$ elements in its domain, $b,c$ are elements of $\A$ and $v$ is a first-order assignment over $\A$ such that $v(x) = b$ and $v(y) = c$.
-Then we have that $\sem{\clfp{\psi(x,R)}{\beta(x,y,R,\pi)}}(\A,v)$ is equal to the  number of paths in $\A$ from $b$ to $c$ of length at most $n$.
-\end{example}
-
-\marcelo{Deberiamos mostrar como el ejemplo anterior se generaliza para definir $[\pth \psi(\bar x, \bar y)]$ en terminos del operador ${\bf lfp}$.}
-
-It is well known that least fixed point logic is contained in second-order logic \cite{L04}. In the following theorem we show that the same holds in our case.
-\begin{theorem} \label{so-rec}
-$\rqfo \subseteq \qso$
-%	Given a positive $\fo$ formula $\varphi(\bar{x},R)$ and a $\qfo$ formula $\alpha(\bar{x})$, there exists a $\qso$ formula $\beta(\bar{x})$ such that $\sem{[\alfp\varphi(\bar{x},R)\mid \alpha(\bar{x},R)](\bar{x})} = \sem{\beta(\bar{x})}$.
-\end{theorem}
-
-Finally, the following is the main result of this section:
-\begin{theorem} \label{rqfo-fo-cap}
-	$\rqfo(\fo)$ captures $\fp$ over the class of ordered structures.
-\end{theorem}
-
-\marcelo{Notese que estoy permitiendo en la formula $\beta(y_1, \ldots, y_\ell,R,\pi)$ tener subformulas $\pa{x} \alpha$. Esta bien el resultado con esto? O tenemos que eliminar estas formulas?}
-
-%Given a relation signature $\R$, the set of recursive $\qfo$ formulas ($\rqfo$-formulas) is defined by the following grammar:
-%%This operator lets us define the set of recursive $\qfo$ formulas over $\R$ ($\rqfo$-formulas) using the following grammar:
-%\begin{multline*}
-%%	\label{eq-def-rqfo}
-%	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid \\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ [\alfp \varphi \mid \alpha]
-%\end{multline*}
-%where $\varphi$ is an $\fo$-formula over $\R$, $s \in \bbN$ and $x \in \fv$.
-%
-%\marcelo{Vamos a permitir anidacion del operador $\alfp$? Esta gramatica lo permite.}
-%
-%\begin{theorem} \label{rqfo-fo-cap}
-%	$\rqfo(\fo)$ captures $\fp$ over the class of ordered structures.
-%\end{theorem}
-%

### 2017-01-06 LICS/framework.tex
sin comentarios

@@
 If $\R$ is not mentioned, $\qso$ refers to the union of the sets of $\qso$ formulas over $\R$, for every relational signature~$\R$.
+\marcelo{En la gramatica de la logica deberiamos incluir la formula $\top$ que discutimos, la cual representa true y para la cual se tiene que $\sem{\top}(\A, v, V) = 1$. Les parece?}
 
@@
 
+\marcelo{IMPORTANTE: Tenemos que decir que $\qfo(\fo)$ permite variables libre de segundo orden en las formulas de primer orden. Hay que aclarar esto tambien para PFP y otras logicas.}
+
 In the next section, we variate $\FF$ and $\LL$ to capture several function complexity classes.

### 2017-01-06 LICS/beyond.tex
sc

@@
 Given an $\R$-structure $\A$ with domain $A$, we say that $F$ is a function assignment for $\A$ if for every $h \in \fs$ with $\arity(h) = \ell$, we have that $F(h) :  A^\ell \to \N$.
-The notion of function assignment is used to extend the semantics of $\qso$ to the case of a formula of the form $h(x_1, \ldots, x_\ell)$.
+The notion of function assignment is used to extend the semantics of $\qso$ to the case of a quantitative formula of the form $h(x_1, \ldots, x_\ell)$.
 More precisely, given first-order and second-order assignments $v$ and $V$ for $\A$, respectively,
@@
 \end{eqnarray*}
-As for the case of $\qfo$, we define $\fqfo$ by restricting $\varphi$ in \eqref{eq-fqso} to be a first-order formula, and by disallowing quantifiers $\Sigma X$ and $\Pi X$ in \eqref{eq-fqso}.
+As for the case of $\qfo$, we define $\fqfo$ disallowing quantifiers $\Sigma X$ and $\Pi X$ in \eqref{eq-fqso}, and we define $\fqfo(\fo)$ by also imposing the restriction that $\varphi$ in \eqref{eq-fqso} is a first-order formula.
 
-\subsection{Recursive Logics}
+\subsection{A logic with a fixed point operator}
 
 We define an operator which extends Least Fixed Point Logic (LFP) \cite{I86,vardi1982complexity} to allow counting. 
-Fix a relational signature $\R$. Then the set of $\rqfo$ formulas over $\R$ is defined by the following grammar:
+Fix a relational signature $\R$. Then the set of $\rqfo(\fo)$ formulas over $\R$ is defined by the following grammar:
 \begin{multline*}
@@
 \end{multline*}
-where $x_1, \ldots, x_\ell$ is a sequence of pairwise distinct first-order variables and $\beta(x_1, \ldots, x_\ell, h)$ is an $\fqfo$-formula over $\R$ whose only function symbol is $h$ with $\arity(h) = \ell$.
+where $\varphi$ is an $\fo$-formula, $x_1, \ldots, x_\ell$ is a sequence of pairwise distinct first-order variables and $\beta(x_1, \ldots, x_\ell, h)$ is an $\fqfo(\fo)$-formula over $\R$ whose only function symbol is $h$, with $\arity(h) = \ell$.
 The free variables of the formula $\clfp{\beta(x_1,\ldots,x_\ell,h)}$ are $x_1, \ldots, x_\ell$; in particular, $h$ is not considered to be free in this formula.
 
-Fix and $\R$-structure with domain $A$ and a quantitative formula $\clfp{\beta(x_1,\ldots,x_\ell,h)}$, and assume that $\F$ is the set of functions $f :A^\ell \to \N$.
-To define the semantics of $\clfp{\beta(x_1,\ldots,x_\ell,h)}$, we first show how $\beta(x_1,\ldots,x_\ell,h)$ can be interpreted as an operator $T_{\beta}$ on the set $\F$.
+Fix an $\R$-structure with domain $A$ and a quantitative formula $\clfp{\beta(x_1,\ldots,x_\ell,h)}$, and assume that $\F$ is the set of functions $f :A^\ell \to \N$.
+To define the semantics of $\clfp{\beta(x_1,\ldots,x_\ell,h)}$, we first show how $\beta(x_1,\ldots,x_\ell,h)$ can be interpreted as an operator $T_{\beta}$ on $\F$.
 More precisely, for every $f \in \F$ and tuple $(a_1, \ldots, a_\ell) \in A^\ell$, the function $T_{\beta}(f)$ satisfies that:
@@
 However, $(\F, \leq)$ is not a complete lattice, so we do not have a Knaster-Tarski Theorem ensuring that such a fixed point exist.
-Instead, we generalize the semantics of LFP in the following way.
+Instead, we generalize the semantics of LFP as follows.
+In the definition of the semantics of LFP, an operator $T$ on relations is considered, and the semantics is defined in terms of the least fixed point of $T$, that is, a relation $R$ such that: (a) $T(R) = R$, and (2) $R \subseteq S$ for every $S$ such that $T(S) = S$ \cite{I86,vardi1982complexity}.
+We can view $T$ as an operator on functions if we consider the characteristic function of a relation.
 Given a relation $R \subseteq A^\ell$, let $\chi_R$ be its characteristic function, that is $\chi_R(\bar a) = 1$ if $\bar a \in R$, and $\chi_R(\bar a) = 0$ otherwise.
+Then define an operator $T^\star$ on characteristic functions as $T^\star(\chi_R) = \chi_{T(R)}$.
+Moreover, we can rewrite the conditions defining a least fixed point of $T$ in terms of the operator $T^\star$ if we consider the notion of support of a function.
+Given a function $f \in \F$, define the support of $f$, denoted by $\support(f)$, as $\{ \bar a \in A^\ell \mid f(\bar a) > 0 \}$.
+Then given that $\support(\chi_R) = R$, we have that the conditions (a) and (b) are equivalent to the following conditions on $T^\star$:
+%Given a relation $R \subseteq A^\ell$, let $\chi_R$ be its characteristic function, that is $\chi_R(\bar a) = 1$ if $\bar a \in R$, and $\chi_R(\bar a) = 0$ otherwise.
 Moreover, given a function $f \in \F$, define the support of $f$, denoted by $\support(f)$, as $\{ \bar a \in A^\ell \mid f(\bar a) > 0 \}$.
 Notice that for a relation $R$ we have that $\support(\chi_R) = R$.
-In the definition of the semantics of LFP, an operator $T$ on relations is considered, and the semantics is defined in terms of the least fixed point for $T$, that is, a relation $R$ such that $T(R) = R$ and $R \subseteq S$ for every $S$ such that $T(S) = S$ \cite{I86,vardi1982complexity}.
-We can view the computation of this fixed point in terms of functions if we consider an operator $T^\star$ on characteristic functions defined as $T^\star(\chi_R) = \chi_{T(R)}$.
-In particular, the conditions defining the least fixed point of $T$ can be rewritten as follows in terms of $T^\star$: $\support(T^\star(\chi_R)) = \support(\chi_R)$ and $\support(\chi_R) \subseteq \support(\chi_S)$ for every $S$ such that
-$\support(T^\star(\chi_{S})) = \support(\chi_S)$.
-This condition can be generalized in the obvious way to the case of the operator $T_\beta$: a function $f \in \F$ is a {\em s-fixed point} of $T_{\beta}$ if $\support(T_\beta(f)) = \support(f)$, and $f$ is a {\em least s-fixed point} of $T_{\beta}$ if $f$ is a s-fixed point of $T_\beta$ and for every s-fixed point $g$ of $T_\beta$ it holds that $\support(f) \subseteq \support(g)$.
+In particular, the conditions defining the least fixed point of $T$ can be rewritten as follows in terms of $T^\star$:
+$\support(T^\star(\chi_R)) = \support(\chi_R)$, and $\support(\chi_R) \subseteq \support(\chi_S)$ for every $S$ such that  $\support(T^\star(\chi_{S})) = \support(\chi_S)$. 
+To define a notion of fixed point for $T_\beta$ we simply generalized these conditions. More precisely, %These conditions can be generalized in the obvious way to the case of the operator $T_\beta$: 
+a function $f \in \F$ is a {\em s-fixed point} of $T_{\beta}$ if $\support(T_\beta(f)) = \support(f)$, and $f$ is a {\em least s-fixed point} of $T_{\beta}$ if $f$ is a s-fixed point of $T_\beta$ and for every s-fixed point $g$ of $T_\beta$ it holds that $\support(f) \subseteq \support(g)$.
 The existence of such fixed point is ensured by the following lemma:
@@
 %As an example, 
-We would like to define a $\rqfo$-formula that, given a directed acyclic graph $G$ with $n$ nodes and a pair of nodes $b$, $c$ in $G$, counts the number of paths of length at most $n$ from $b$ to $c$ in $G$.
+We would like to define an $\rqfo(\fo)$-formula that, given a directed acyclic graph $G$ with $n$ nodes and a pair of nodes $b$, $c$ in $G$, counts the number of paths of length at most $n$ from $b$ to $c$ in $G$.
 To this end, assume that graphs are encoded using the relational signature $\R = \{ E(\cdot,\cdot) \}$, and then define formula $\alpha(x, y, f)$ as follows:
@@
 \end{example}
-
 %
@@
 %\end{theorem}
-
-The following is the main result of this section:
+In the following result we show how the fixed point operator introduced in this section can be used to capture (over the class of ordered structures) the fundamental class of functions that can be computed in polynomial time.
 \begin{theorem} \label{rqfo-fo-cap}

### 2017-01-06 LICS/appendix-counting.tex

@@
 Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq 2^{n^\ell}$, where $\A$ has a domain of size $n$.
-Let $\Phi(X)$ be a $\pfp$ formula that represents the following.
-Consider an order over the subsets of $A^{\ell}$ given by the formula $$\varphi_{<}(X,Y) = \exists\bar{u}\big[\neg X(\bar{u})\wedge Y(\bar{u})\wedge \forall\bar{v}\big(
-\bar{u}<\bar{v}\to(X(\bar{u})\iff Y(\bar{v}))\big)\big].$$
+Let $X$ be a second-order variable of arity $\ell$.
+Consider a linear order over predicates of arity $\ell$ given by the formula
+$$
+\varphi_{<}(X,Y) = \exists\bar{u}\big[\neg X(\bar{u})\wedge Y(\bar{u})\wedge \forall\bar{v}\big(
+\bar{u}<\bar{v}\to(X(\bar{u})\iff Y(\bar{v}))\big)\big].
+$$
+Consider a polynomial space algorithm over $\R\cup\{X\}$ that receives a $\R\cup\{X\}$-structure $\A'$.
+Let $X^{\A'}$ be the $p$-th subset in the given order (we consider 0-indexation).
+The algorithm simulates $f(\enc(\A))$ until $(2^{n^{\ell}}-p)$ characters have been written, but each character of the output is written over the last one.
+It accepts if and only if the last character written is $1$.
+Since this algorithm works in polynomial space, it can be described in $\pfp$ \cite{AbiteboulV89} by a formula $\Phi(X)$.
 
-Let $B$ be the $p$-th subset with respect to this order and let $w$ be a string of size $2^{n^{\ell}}$ which represents the value of $f(\enc(\A))$ in binary.
-Then $\A\models\Phi(B)$ if and only if the $p$-th bit in $w$ is 1, from least to most significant.
-Since this condition can be verified in polynomial space, and $\pfp$ captures $\pspace$, then $\Phi$ can be expressed in $\pfp$.
+Note that the formula also represents the following.
+For each $(\A,v,V)\in\ostr[\R]^*$, let $V(X)$ be the $p$-th subset with respect to this order and let $w$ be a string of size $2^{n^{\ell}}$ which represents the value of $f(\enc(\A))$ in binary.
+Then $(\A,v,V)\models\Phi(X)$ if and only if the $p$-th bit in $w$ is 1, from least to most significant.
 We define
-$\alpha := \sa{X}(\Phi(X)\mult\varphi(X))$, where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
+$$\alpha := \sa{X}(\Phi(X)\mult\varphi(X)),$$ where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
 Note that for each $B\subseteq A^{\ell}$ such that $B$ is the $m$-th element in the given order, $\sem{\varphi(B)}(\A) = 2^m$.

### 2017-01-06 LICS/extentions.tex

There are counting complexity classes that do not fit in our framework because either the output of a function in not a natural numbers (e.g. a negative number) or the class is not defined purely in terms of arithmetical operations (e.g. min and max).
To remedy this problem, we show here how to easily extend $\qso$ in order to capture these classes that go beyond our logical framework.

It is well-known that $\shp$ is not closed under subtraction, not either closed under subtraction by one, by some complexity-theoretical assumption~\cite{OH93}.
To overcome this issue, people introduced the class $\gp$~\cite{fenner1994gap} defined as the class of functions $f$ such that for some non-deterministic polynomial time Turing machine, $f(x)$ is the number of accepting runs minus the number of rejecting runs.
That is, $\gp$ is the closure of $\shp$ functions under subtraction and, by definition, it can compute negative values.

Given that our logical framework was built on top of non-negative values, we need to extend $\qso$ in order to describe functions in $\gp$. 
The most elegant way to do this is by allowing constants coming from $\bbZ$ instead of just $\bbN$. 
Formally, we define the logic $\qsoz$ whose syntax is the same as in (\ref{syntax}) and semantics is the same as in Table~\ref{tab-semantics} except that constant $s$ come from $\bbZ$.  
Similar than for $\qso$, we define the fragment $\eqsoz$ as the extension of $\eqso$ with negative constants.
\begin{example}
	Recall the setting of Example~\ref{ex:cliques} and suppose now that we want to compute the number of cliques in a graph that are not triangles. Then this can be easily done in $\qsoz$ with the following formula:
	\[
	\alpha_4 := \alpha_2 + (-1) \cdot \alpha_1 
	\]
\end{example}
Adding negative constants is a mild extension to allow subtraction in the logic. 
In fact, this is what we need to capture classes like $\gp$.
\begin{proposition} \label{prop:capture-gapp}
	$\eqsoz(\fo)$ captures $\gp$ over ordered structures.
\end{proposition}
The previous result can be proved straightforward from the characterization of $\shp$ and the definition of $\gp$. 
However, we believe that this is a relevant result for showing how robust is $\qso$ in order to capture diverse complexity classes.

### 2017-01-06 LICS/framework.tex

@@
 %\]
-\begin{multline*}
-\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \\ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha 
-\end{multline*}
+\begin{multline}
+\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \\ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha \label{syntax} 
+\end{multline}
 where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
@@
 Thus, in such a case we use the term $\sem{\alpha}(\A)$ to denote $\sem{\alpha}(\A, v, V)$, for some arbitrary first-order assignment $v$ for $\A$ and some arbitrary second-order assignment $V$ for $\A$.
-\begin{example}
+\begin{example}\label{ex:cliques}
 Let $\bG = \{E(\cdot,\cdot)\}$ be the vocabulary for graphs and $\fG$ be an ordered finite $\bG$-structure encoding a non-directed graphs. 

### 2017-01-06 LICS/extentions.tex
Avance en sección extended.

@@
 However, we believe that this is a relevant result for showing how robust is $\qso$ in order to capture diverse complexity classes.
+
+An interesting class of functions in complexity theory comes from considering the optimization version of a decision problem.
+For example, one can define MAX-SAT as the problem of determining the maximum number of clauses, of a given CNF formula, that can be made true by an assignment.
+Here, the problem is defined in terms of a maximization function and it is not known how to define this kind of function in $\shp$.
+To formalize this set of optimization problems, in~\cite{krentel1988complexity} the class $\optp$ was introduces as the class of functions computable by taking the maximum or minimum of the output values over all runs of a non-deterministic polynomial time machine with output (i.e.
+each run produces a binary string which is interpreted as a number).
+For instance, MAX-SAT is in $\optp$ with several other optimization version of $\np$-problems.
+Given that in~\cite{krentel1988complexity} the author do not make the distinction between $\max$ or $\min$, in 
+Given that
+
+\begin{multline}
+\alpha := \beta \ \mid \ s \ \mid \ (\, \max\{\alpha,\alpha\} \,) \ \mid\ (\, \min\{\alpha,\alpha\} \,) \ \mid \\ \maxa{x} \alpha \ \mid \ \mina{x} \alpha \ \mid \ \maxa{X} \alpha \ \mid \ \mina{X} \alpha
+\end{multline}

### 2017-01-06 LICS/beyond.tex

@@
 $$
-\sa{u} (\varphi_{\text{\rm first}}(u) \wedge \clfp{\beta(x,y,u,g)}.
+\sa{u} (\varphi_{\text{\rm first}}(u) \wedge \clfp{\beta(x,y,u,g)}).
 $$ 
@@
 
-\subsection{Transitive Logics}
+\subsection{A logic with an operator for counting paths}
 
@@
 
-\marcelo{Deberiamos mostrar como el ejemplo anterior se generaliza para definir $[\pth \psi(\bar x, \bar y)]$ en terminos del operador ${\bf lfp}$.}
+\begin{theorem} \label{tqfo-subseteq}
+	$\tqfo(\fo) \subseteq \rqfo(\fo)$.
+\end{theorem}
+Given tuples $\bar x = (x_1, \ldots, x_\ell)$, $\bar y = (y_1, \ldots, y_\ell)$ of pairwise distinct first-order variables, define formula $\varphi_{\text{\rm lex}}(\bar x, \bar y)$ as follows:
+$$
+\bigvee_{i = 1}^\ell \bigg(\bigwedge_{j = 1}^{i -1} x_i = y_i\bigg) \wedge x_i < y_i.
+$$
+That is, $\varphi_{\text{\rm lex}}(\bar x, \bar y)$ holds if $\bar x$ is smaller than $\bar y$ in the lexicographic order on tuples with $\ell$ elements induced by the built-in order of each structure.
+Moreover, define formula $\varphi_{\text{\rm succ}}(\bar x, \bar y)$ as follows:
+$$
+\varphi_{\text{\rm lex}}(\bar x, \bar y) \wedge \neg \exists \bar z \, (\varphi_{\text{\rm lex}}(\bar x, \bar z) \wedge \varphi_{\text{\rm lex}}(\bar z, \bar y)).
+$$
+That is, $\varphi_{\text{\rm succ}}(\bar x, \bar y)$ holds if $\bar y$ is the successor of $\bar x$ in the lexicographic order on tuples with $\ell$ elements.
+With this notation, define $\tqsos$ to be the restriction of $\tqso$ where each occurrence of the operator ${\bf path}$ is of the form:
+$$
+[\pth (\psi(\bar{x}, \bar{X},\bar{y}, \bar{Y}) \wedge \varphi_{\text{\rm succ}}(\bar x, \bar y))],
+$$
+where $\psi(\bar{x}, \bar{X},\bar{y}, \bar{Y})$ satisfies the same conditions as in grammar \eqref{eq-def-tqso}. Then we have that:
+\begin{theorem} \label{tqfo-shl}
+	$\tqsos(\fo)$ captures $\shp$ over the class of ordered structures.
+\end{theorem}
+
 %

### 2017-01-06 LICS/extentions.tex
Primera versión de extended classes.

--- 2017-01-06 LICS/extentions.tex
@@
-There are counting complexity classes that do not fit in our framework because either the output of a function in not a natural numbers (e.g.
+There exist counting complexity classes that do not fit in our framework because either the output of a function in not a natural numbers (e.g.
 a negative number) or the class is not defined purely in terms of arithmetical operations (e.g.
 min and max).
-To remedy this problem, we show here how to easily extend $\qso$ in order to capture these classes that go beyond our logical framework.
+To remedy this problem, we show here how to easily extend $\qso$ in order to capture these classes that go beyond sum and product over natural numbers.
 
-It is well-known that $\shp$ is not closed under subtraction, not either closed under subtraction by one, by some complexity-theoretical assumption~\cite{OH93}.
-To overcome this issue, people introduced the class $\gp$~\cite{fenner1994gap} defined as the class of functions $f$ such that for some non-deterministic polynomial time Turing machine, $f(x)$ is the number of accepting runs minus the number of rejecting runs.
-That is, $\gp$ is the closure of $\shp$ functions under subtraction and, by definition, it can compute negative values.
+It is well-known that $\shp$ is not closed under subtraction, not even closed under subtraction by one~\cite{OH93}.
+To overcome this issue, people introduced the class $\gp$~\cite{fenner1994gap} defined as the class of functions $f$ such that for some $\np$ machine, $f(x)$ is the number of accepting runs minus the number of rejecting runs.
+That is, $\gp$ is the closure of $\shp$ functions under subtraction and it can compute negative values by definition.
 
-Given that our logical framework was built on top of non-negative values, we need to extend $\qso$ in order to describe functions in $\gp$. 
+Given that our logical framework was built on top of natural numbers, we need to extend $\qso$ in order to describe functions in $\gp$. 
 The most elegant way to do this is by allowing constants coming from $\bbZ$ instead of just $\bbN$. 
-Formally, we define the logic $\qsoz$ whose syntax is the same as in (\ref{syntax}) and semantics is the same as in Table~\ref{tab-semantics} except that constant $s$ come from $\bbZ$.  
-Similar than for $\qso$, we define the fragment $\eqsoz$ as the extension of $\eqso$ with negative constants.
+Formally, we define the logic $\qsoz$ whose syntax is the same as in (\ref{syntax}) and whose semantics is the same as in Table~\ref{tab-semantics} except that the atomic formula $s$ (i.e.
+a constant) come from $\bbZ$.
+Similar than for $\qso$, we define the fragment $\eqsoz$ as the extension of $\eqso$ with constants in $\bbZ$.
 \begin{example}
@@
 	\[
-	\alpha_4 := \alpha_2 + (-1) \cdot \alpha_1 
+	\alpha_5 :=	\alpha_2 + (-1) \cdot \alpha_1 
 	\]
@@
 Adding negative constants is a mild extension to allow subtraction in the logic. 
-In fact, this is what we need to capture classes like $\gp$.
+In fact, this is exactly what we need to capture classes like $\gp$.
 \begin{proposition} \label{prop:capture-gapp}
@@
 \end{proposition}
-The previous result can be proved straightforward from the characterization of $\shp$ and the definition of $\gp$. 
-However, we believe that this is a relevant result for showing how robust is $\qso$ in order to capture diverse complexity classes.
+The previous result is straightforward to prove from the characterization of $\shp$ and the definition of $\gp$. 
+However, we believe that this is an interesting result for showing how robust is $\qso$ in order to capture diverse complexity classes.
+
+An interesting class of functions comes from considering the optimization version of a decision problem.
+For example, one can define MAX-SAT as the problem of determining the maximum number of clauses, of a given CNF formula, that can be made true by an assignment.
+Here, MAX-SAT is defined in terms of a maximization problem which in his essence differs from functions in $\shp$.
+To formalize this set of optimization problems, Krentel defined $\optp$~\cite{krentel1988complexity} as the class of functions computable by taking the maximum or minimum of the output values over all runs of an $\np$ machine with output (i.e.
+each run produces a binary string which is interpreted as a number).
+For instance, MAX-SAT is in $\optp$ as several others optimization version of $\np$-problems.
+Given that in~\cite{krentel1988complexity} the author do not make the distinction between $\max$ or $\min$, in~\cite{vollmer1995complexity} the authors define the classes $\maxp$ and $\minp$ as the max and min version of the problems in $\optp$ (i.e.
+$\optp = \maxp \cup \minp$).
+
+In order to capture classes of optimization functions, we can extend $\qso$ with $\max$ and $\min$ quantifiers (called $\optqso$) as follows. 
+Given a relational signature $\R$, the set of $\optqso$-formulas over $\R$ is given by extending the syntax in (\ref{syntax}) with the following operators:
+\begin{multline*}
+\max\{\alpha,\alpha\} \ \mid\ \min\{\alpha,\alpha\} \ \mid \\ \maxa{x} \alpha \ \mid \ \mina{x} \alpha \ \mid \ \maxa{X} \alpha \ \mid \ \mina{X} \alpha 
+\end{multline*}
+where $x \in \fv$ and $X \in \sv$.
+The semantics of the $\qso$-operators in $\optqso$ are defined as usual.
+Furthermore, the semantics of the max and min quantifiers are defined as the natural extension of the sum quantifiers in $\qso$ (see Table~\ref{tab-semantics}) by maximizing or minimizing (respectively) instead of summing.
+\begin{example}\label{ex:optqso}
+	Recall again the setting of Example~\ref{ex:cliques} and suppose now that we want to compute the size of the bigger clique in a graph. Then this can be define in $\optqso$ with the following formula:
+	\[
+\alpha_6 := \maxa{X} \left( \, \clique(X) \cdot \sa{z} X(z)  \, \right)
+	\]
+\end{example}
+Similar than for $\maxp$ and $\minp$, we have to distinguished between the $\max$ and $\min$ fragments of $\optqso$.
+For this, we define the fragment $\maxqso$ of all $\optqso$ formulas constructed from $\qfo$ operators and $\max$ operators (of first- and second-order).
+The class $\minqso$ is defined analogously changing $\max$ with $\min$.
+Note that in $\maxqso$ and $\minqso$ the second-order sum and product quantifiers are not allowed.
+For example, the formulas $\alpha_1$ and $\alpha_6$ are in $\maxqso$ but $\alpha_2$ and $\alpha_3$ are not.
+As one could expect, $\maxqso$ and $\minqso$ are the logics that we need to capture $\optp$.
+\begin{theorem} \label{theo:capture-optp}
+	$\maxqso(\fo)$ and $\minqso(\fo)$ captures $\maxp$ and $\minp$, respectively, over ordered structures.
+\end{theorem}
+It is important to mention that a similar result was proved in~\cite{kolaitis1994logical} for the class $\maxpb$ and $\minpb$ of problems in $\maxp$ and $\minp$ whose output value is polynomially bounded.
+Interestingly, the result is stronger here since our logic has the freedom to use sum and product quantifiers instead of using a max-and-count problem over boolean formulas. 
+Finally, it is straightforward to prove that our framework can also capture $\maxpb$ and $\minpb$ by disallowing the first-order quantifier in $\maxqso(\fo)$ and $\minqso(\fo)$, respectively.

### 2017-01-06 LICS/framework.tex
Primera versión de extended classes.

@@
 $$
-\alpha_2 := \sa{X} \big( \, \fa{x} \fa{y} \left(X(x) \wedge X(y)\right) \rightarrow E(x,y) \, \big)  
+\alpha_2 := \sa{X} \clique(X) 
 $$ 
+where $\clique(X) := \fa{x} \fa{y} \left(X(x) \wedge X(y)\right) \rightarrow E(x,y)$.
 Similar than for $\alpha_1$, in the boolean subformula of $\alpha_2$ we check whether $X$ is a clique and with the sum quantifier we add one for each clique in $\fG$. 

### 2017-01-06 LICS/appendix-counting.tex

@@
 Since a function in $\#\eso$ can also be defined in $\eqso(\eso)$, then $\eqso(\eso)$ captures $\spp$ over ordered structures.
-\qed
 
@@
 Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{a_1,\ldots,a_n\}$.
-Replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\Sigma x \beta(x)$ is replaced by $(\beta(a_1)+\cdots+\beta(a_n))$ and $\Pi x \beta(x)$ is replaced by $(\beta(a_1)\cdot\,\cdots\,\cdot\beta(a_n))$.
-Then each sub-formula $\varphi\in\lfp$ in $\alpha$ is replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
+Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{1,\ldots,n\}$.
+In the procedure we slightly extend the grammar of $\qfo(\lfp)$ to include constants.
+We replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
+This is, $\sa{x} \beta(x)$ is replaced by $(\beta(1)+\cdots+\beta(n))$ and $\pa{x}\beta(x)$ is replaced by $(\beta(1)\cdot\,\cdots\,\cdot\beta(n))$.
+Then each sub-formula $\varphi\in\lfp$ in $\alpha$ is evaluated in polynomial time and replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
 The resulting formula is an arithmetic expression of polynomial size which is evaluated and lastly given as output.
@@
 Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
-Let $\Phi(x_0,\ldots,x_{\ell-1})$ be a formula that represents the following.
-Let $\A$ be an ordered structure over $\R$ with domain $A$.
-Consider a linear order over $A^{\ell}$ given by $\bar{a} < \bar{b}$, for any $\bar{a},\bar{b}\in A^{\ell}$.
-Let $w$ be a string of size $n^{\ell}$ the represents $f(\enc(\A))$ in binary.
-Let $(a_1,\ldots,a_{\ell})\in A^{\ell}$ be the $m$-th tuple in the given order (starting from 0).
-$\A\models\Phi(a_1,\ldots,a_{\ell})$ if an only if the $m$-th bit in $w$ is 1, from least to most significant.
-Since this condition can be computed in polynomial time, and $\lfp$ captures $\ptime$, this formula can be expressed in $\lfp$.
-We use
+Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
+Consider a procedure that receives $\enc(\A)$ and an assignation $\bar{a}$ to $\bar{x}$.
+Let $m$ be the position of $\bar{a}$ in the lexicographic order of the tuples in $A^{\ell}$.
+The procedure then computes the $m$-th bit of $f(\enc(\A))$, from least to most significant.
+Since this procedure works in polynomial time, it can be described by an $\lfp$ formula $\Phi(\bar{x})$.
+Then we use
 $$
-\alpha = \sa{x_1} \cdots \sa{x_{\ell}} \Phi(x_1,\ldots,x_{\ell})\cdot\varphi(x_1,\ldots,x_{\ell}),
+\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot\varphi(\bar{x}),
 $$
-where $\varphi(x_1,\ldots,x_{\ell}) := \pa{y_1}\cdots\pa{y_{\ell}}((y_1,\ldots,y_{\ell}) < (x_1,\ldots,x_{\ell}) \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
+where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
 Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
 Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-\qed
+
 \medskip
@@
 Let $X^{\A'}$ be the $p$-th subset in the given order (we consider 0-indexation).
-The algorithm simulates $f(\enc(\A))$ until $(2^{n^{\ell}}-p)$ characters have been written, but each character of the output is written over the last one.
+The algorithm computes $f(\enc(\A))$ until $(2^{n^{\ell}}-p)$ characters have been written, but each character of the output is written over the last one.
 It accepts if and only if the last character written is $1$.
@@
 We define
-$$\alpha := \sa{X}(\Phi(X)\mult\varphi(X)),$$ where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
+$$
+\alpha := \sa{X}(\Phi(X)\mult\varphi(X)),
+$$ 
+where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
 Note that for each $B\subseteq A^{\ell}$ such that $B$ is the $m$-th element in the given order, $\sem{\varphi(B)}(\A) = 2^m$.
@@
 Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{a_1,\ldots,a_n\}$.
-Replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\Sigma x \beta(x)$ is replaced by $(\beta(a_1)+\cdots+\beta(a_n))$ and $\Pi x \beta(x)$ is replaced by $(\beta(a_1)\cdot\,\cdots\,\cdot\beta(a_n))$.
-Then each sub-formula $\varphi\in\pfp$ in $\alpha$ is replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
+Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{1,\ldots,n\}$.
+In the procedure we slightly extend the grammar of $\qfo(\pfp)$ to include constants.
+We replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
+This is, $\sa{x} \beta(x)$ is replaced by $(\beta(1)+\cdots+\beta(n))$ and $\pa{x}\beta(x)$ is replaced by $(\beta(1)\cdot\,\cdots\,\cdot\beta(n))$.
+Then each sub-formula $\varphi\in\pfp$ in $\alpha$ is evaluated in polynomial space and replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
 The resulting formula is an arithmetic expression of polynomial size which is evaluated and lastly given as output.
@@
 Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
-Let $\Phi(x_0,\ldots,x_{\ell-1})$ be a formula that represents the following.
-Let $\A$ be an ordered structure over $\R$ with domain $A$.
-Consider a linear order over $A^{\ell}$ given by $\bar{a} < \bar{b}$, for any $\bar{a},\bar{b}\in A^{\ell}$.
-Let $w$ be a string of size $n^{\ell}$ the represents $f(\enc(\A))$ in binary.
-Let $(a_1,\ldots,a_{\ell})\in A^{\ell}$ be the $m$-th tuple in the given order (starting from 0).
-$\A\models\Phi(a_1,\ldots,a_{\ell})$ if an only if the $m$-th bit in $w$ is 1, from least to most significant.
-Since this condition can be computed in polynomial space, and $\pfp$ captures $\pspace$, this formula can be expressed in $\pfp$.
-We use
+Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
+Consider a procedure that receives $\enc(\A)$ and an assignation $\bar{a}$ to $\bar{x}$.
+Let $m$ be the position of $\bar{a}$ in the lexicographic order of the tuples in $A^{\ell}$.
+The procedure then computes the $m$-th bit of $f(\enc(\A))$, from least to most significant.
+Since this procedure works in polynomial space, it can be described by an $\pfp$ formula $\Phi(\bar{x})$.
+Then we use
 $$
-\alpha = \sa{x_1} \cdots \sa{x_{\ell}} \Phi(x_1,\ldots,x_{\ell})\cdot\varphi(x_1,\ldots,x_{\ell}),
+\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot\varphi(\bar{x}),
 $$
-where $\varphi(x_1,\ldots,x_{\ell}) := \pa{y_1}\cdots\pa{y_{\ell}}((y_1,\ldots,y_{\ell}) < (x_1,\ldots,x_{\ell}) \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
+where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
 Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.

### 2017-01-06 LICS/appendix-syntactic.tex

@@
 $$
-\alpha = c + \sum_{i = 1}^{\text{\#clauses}} \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigwedge_{j = 1}^{n}\forall\bar{z}\,\varphi^i_j(\bar{X},\bar{x},\bar{y},\bar{z}),
+\alpha = \sum_{i = 1}^{\text{\#clauses}} \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigwedge_{j = 1}^{n}\forall\bar{z}\,\varphi^i_j(\bar{X},\bar{x},\bar{y},\bar{z}),
 $$
@@
 $$
-c + \sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length(\bar{x})}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length(\bar{y})}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length(\bar{z})}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
+\sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length(\bar{x})}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length(\bar{y})}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length(\bar{z})}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
 $$
@@
 $$
-c + \sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})
+\sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})
 $$

### 2017-01-06 LICS/preliminaries.tex
sc

--- 2016-12-29 LICS/preliminaries.tex
@@
 
+In this section, we introduce the main terminology used in this paper.
+
+\subsection{Second-order logic, LFP and PFP}
 A relational signature $\R$ (or just signature) is a finite set $\{R_1, \ldots, R_k\}$, where each $R_i$ ($1 \leq i \leq k$) is a relation name with an associated arity greater than 0, which is denoted by $\arity(R_i)$.
 A finite structure over $\R$ (or just finite $\R$-structure) is a tuple $\A = \langle A, R_1^\A, \ldots, R_k^\A \rangle$ such that $A$ is a finite set and $R_i^\A \subseteq A^{\arity(R_i)}$ for every $i \in \{1, \ldots, k\}$.
-A finite $\R$-structure $\A$ is said to be ordered if $<$ is a binary predicate name in $\R$ and $<^\A$ is a linear order on $A$.
-Let $\str[\R]$ be the class of all finite $\R$-structures and $\ostr[\R]$ be the class of all ordered finite $\R$-structures.
+In this paper we only consider finite structures, so we omit the word finite when referring to them.
+An $\R$-structure $\A$ is said to be ordered if $<$ is a binary predicate name in $\R$ and $<^\A$ is a linear order on $A$.
+Let $\str[\R]$ be the class of all $\R$-structures and $\ostr[\R]$ be the class of all ordered $\R$-structures.
 
@@
 Then given a relational signature $\R$, the set of second-order logic formulas ($\so$-formulas) over $\R$ is given by the following grammar:
-\begin{eqnarray*}\ 
-	\varphi &:=& R(\bar u) \ \mid\  
-	X(\bar v)  \ \mid\ 
+\begin{multline*}\ 
+	\varphi \ := \ x = y \ \mid \ R(\bar u) \ \mid \ \top  \ \mid\  
+	X(\bar v)  \ \mid\\ 
 	\neg \varphi \ \mid\ 
@@
 	\exists X \, \varphi
-\end{eqnarray*}
-where $R \in \R$, $\bar u$ is a tuple of (non-necessarily distinct) variables from $\fv$ whose length is $\arity(R)$, $X \in \sv$, $\bar v$ is a tuple of (non-necessarily distinct) variables from $\fv$ whose length is $\arity(X)$, and $x \in \fv$.
+\end{multline*}
+where $x,y \in \fv$, $R \in \R$, $\bar u$ is a tuple of (non-necessarily distinct) variables from $\fv$ whose length is $\arity(R)$, $\top$ is a reserved symbol to represent a tautology, $X \in \sv$, $\bar v$ is a tuple of (non-necessarily distinct) variables from $\fv$ whose length is $\arity(X)$, and $x \in \fv$.
 
-To define the semantics of $\so$, we need to introduce some terminology.
-Given a relational signature $\R$ and an ordered finite $\R$-structure $\A$ with domain $A$, a first-order assignment $v$ for $\A$ is a total function from $\fv$ to $A$, while a second-order assignment $V$ for $\A$ is a total function with domain $\sv$ that maps each $X \in \sv$ to a subset of $A^{\arity(X)}$.
+%\marcelo{Tenemos una definicion muy detallada de $\fo$ y $\so$, no es necesario para esta conferencia.
+Por otro lado seria bueno mencionar las definiciones de $\Sigma_i$ y $\Pi_i$, aunque son estandar es bueno decir que estamos considerando restricciones de $\fo$.}
+
+%\martin{Decir que $\LL$ incluye tautología y decir que Saluja no lo menciona}
+
+We assume that the reader is familiar with the semantics of $\so$, so we only introduce here some notation that will be used in this paper. 
+%To define the semantics of $\so$, we need to introduce some terminology. 
+Given a signature $\R$ and an $\R$-structure $\A$ with domain $A$, a first-order assignment $v$ for $\A$ is a total function from $\fv$ to $A$, while a second-order assignment $V$ for $\A$ is a total function with domain $\sv$ that maps each $X \in \sv$ to a subset of $A^{\arity(X)}$.
 Moreover, given a first-order assignment $v$ for $\A$, $x \in \fv$ and $a \in A$, we denote by $v[a/x]$ a first-order assignment such that $v[a/x](x) = a$ and $v[a/x](y) = v(y)$ for every $y \in \fv$ distinct from $x$.
 Similarly, given a second-order assignment $V$ for $\A$, $X \in \sv$ and $B  \subseteq A^{\arity(X)}$, we denote by $V[B/X]$ a second-order assignment such that $V[B/X](X) = B$ and $V[B/X](Y) = V(Y)$ for every $Y \in \sv$ distinct from $X$.
+We use notation $(\A, v, V) \models \varphi$ to indicate that structure $\A$ satisfies $\varphi$ under the variable assignments $v$ and $V$.
+In particular, we have that $(\A, v, V) \models \top$.
 
-Assume that $\varphi$ is an $\so$-formula over a signature $\R$.
+%Assume that $\varphi$ is an $\so$-formula over a signature $\R$.
 Then given an ordered finite $\R$-structure $\A$ with domain $A$, a first-order assignment $v$ for $\A$ and a second-order assignment $V$ for $\A$, we say that $(\A, v, V)$ satisfies $\varphi$, denoted by $(\A, v, V) \models \varphi$, if: (1) $\varphi$ is the formula $R(x_1, \ldots, x_\ell)$ and $(v(x_1), \ldots, v(x_\ell)) \in R^\A$; (2) $\varphi$ is the formula $X(x_1, \ldots, x_m)$ and $(v(x_1), \ldots, v(x_m)) \in V(X)$; (3) $\varphi$ is the formula $\neg \psi$ and $(\A, v, V)$ does not satisfy $\psi$; (4) $\varphi$ is the formula $(\varphi_1 \vee \varphi_2)$, and $(\A, v, V) \models \varphi_1$ or $(\A, v, V) \models \varphi_2$; (5) $\varphi$ is the formula $\exists x \, \psi$ and there exists $a \in A$ such that $(\A, v[a/x], V) \models \psi$; or (6) $\varphi$ is the formula $\exists X \, \psi$ and there exists $B \subseteq A^{\arity(X)}$ such that $(\A, v, V[B/X]) \models \psi$.
@@
 
-\cristian{Agregar que la igualdad $=$ esta en nuestra lógica.}
+%\cristian{Agregar que la igualdad $=$ esta en nuestra lógica.}
 
-\cristian{Agregar acá que usamos las abreviaciones de $<$ o $\subset$.}
+%\cristian{Agregar acá que usamos las abreviaciones de $<$ o $\subset$.}
 
@@
 
-\subsection{Function complexity classes}
+\cristian{Explicar acá que significa el containment de lógicas booleans, esto es, la notación $\U{1} \subseteq \LL$.}
+
+\subsection{Counting complexity classes}
 

### 2017-01-06 LICS/extentions.tex
sc

@@
 Interestingly, the result is stronger here since our logic has the freedom to use sum and product quantifiers instead of using a max-and-count problem over boolean formulas. 
-Finally, it is straightforward to prove that our framework can also capture $\maxpb$ and $\minpb$ by disallowing the first-order quantifier in $\maxqso(\fo)$ and $\minqso(\fo)$, respectively.
+Finally, it is straightforward to prove that our framework can also capture $\maxpb$ and $\minpb$ by disallowing the product $\Pi x$ in $\maxqso(\fo)$ and $\minqso(\fo)$, respectively.

### 2017-01-06 LICS/beyond.tex
arregle un label
sin comentarios

@@
 
-It was shown in \cite{I86,I88} that first-order logic extended with a transitive closure operator captures $\nlog$.
+It was shown in \cite{I86,I88} that $\fo$ extended with a transitive closure operator captures $\nlog$.
+Moreover, it was shown in \cite{I83} that $\so$ extended with a transitive closure operator captures $\pspace$.
 Inspired by this work, we extend the definition of $\qfo$ with an operator for counting the number of paths in a directed graph, and we show that it can be used to capture $\shl$.

### 2017-01-06 LICS/preliminaries.tex
sc

@@
 
+In this paper we consider several fragments of $\so$, in particular first-order logic ($\fo$), least fixed point logic (LFP) and partial fixed point logic (PFP) \cite{L04}.
+Moreover, for every $i \in \N$, we consider the fragment $\Sigma_i$ (resp., $\Pi_i$) of $\fo$, which is the set of $\fo$-formulas of the form
+$\exists \bar x_1 \forall \bar x_2 \cdots \exists \bar x_{i-1} \forall \bar x_{i} \, \psi$ (resp., 
+$\forall \bar x_1 \exists \bar x_2 \cdots \forall \bar x_{i-1} \exists \bar x_{i} \, \psi$) if $i$ is even and of the form
+$\exists \bar x_1 \forall \bar x_2 \cdots \forall \bar x_{i-1} \exists \bar x_{i} \, \psi$ (resp., 
+$\forall \bar x_1 \exists \bar x_2 \cdots \exists \bar x_{i-1} \forall \bar x_{i} \, \psi$) if $i$ is odd, where $\psi$ is a quantifier-free formula.
+Finally, we say that a fragment $\L_1$ is contained in a fragment $\L_2$, denoted by $\L_1 \subseteq \L_2$, if for every formula $\varphi$ in $\L_1$, there exists a formula $\psi$ in $\L_2$ such that $\varphi$ is logically equivalent to $\psi$.
+
 %Assume that $\varphi$ is an $\so$-formula over a signature $\R$.
@@
 
-\marcelo{Agregar las definiciones de LFP y PFP}
+%\marcelo{Agregar las definiciones de LFP y PFP}
 
-\cristian{Explicar acá que significa el containment de lógicas booleans, esto es, la notación $\U{1} \subseteq \LL$.}
+%\cristian{Explicar acá que significa el containment de lógicas booleans, esto es, la notación $\U{1} \subseteq \LL$.}
 
@@
 
-We define the following function complexity classes: $\fp$ is defined as the class of functions computable in polynomial time.
-$\shp$ is defined as the class of functions that on an input are equal to the number of accepting paths of an $\np$ machine on that input.
-$\spp$ is defined as the class of functions that on an input are equal to the number of different outputs of an $\np$ transducer on that input.
-$\fpspace$ is the class of functions computable in polynomial space.
+We consider the following counting complexity classes in this paper. 
+%We define the following function complexity classes: 
+$\fp$ is the class of functions computable in polynomial time, while $\fpspace$ is the class of functions computable in polynomial space.
+Moreover, $\nfpspace$ is the class of functions computable in polynomial space with output length bounded by a polynomial.
+Given a nondeterministic Turing Machine (NTM) $M$, let $\tma_M(x)$ be the number of accepting runs of $M$ with input $x$.
+Then $\shp$ is the class of functions $f$ for which there exists a polynomial-time NTM $M$ such that $f(x) = \tma_M(x)$ for every input $x$, while $\shl$ is the class of functions $f$ for which there exists a logarithmic-space NTM $M$ such that $f(x) = \tma_M(x)$ for every input $x$.
+Given an NTM $M$ with output tape, let $\tmo_M(x)$ be the number of distinct outputs of $M$ with input $x$ (notice that  $M$ produces an output if it halts in an accepting state).
+Then $\spp$ is the  class of functions $f$ for which there exists a polynomial-time NTM $M$ such that $f(x) = \tmo_M(x)$ for every input $x$.
+Notice that $\shp \subseteq \spp$, and this inclusion is believed to be strict.
+
+%as the class of functions that on an input are equal to the number of different outputs of an $\np$ transducer on that input.
+.
 $\shpspace$ is defined analogously to $\shp$ but in polynomial space.
@@
 
-\marcelo{La notacion $\tma_M(x)$ se usa en el paper, hay que definirla aqui.
+%\marcelo{La notacion $\tma_M(x)$ se usa en el paper, hay que definirla aqui.
 Donde vamos a usar la clase $\gp$?

### 2017-01-07 LICS/framework.tex
sc

--- 2017-01-06 LICS/framework.tex
@@
 
-We introduce here the logic framework that we use for studying function complexity classes. 
-This logic framework is based on the framework of Weighted Logics~\cite{DrosteG07} (WL) that has been used in the context of weighted automata for studying functions from words (or trees) to semirings.
+We introduce here the logical framework that we use for studying counting complexity classes. 
+This framework is based on the framework of Weighted Logics~\cite{DrosteG07} (WL) that has been used in the context of weighted automata for studying functions from words (or trees) to semirings. 
 We propose here to use the framework of WL over any relational structure and to restrict the semiring to natural numbers. 
-The extension to any relational structure will allow us to study general function complexity classes and the restriction to the natural numbers will simplify the notation in this context (see Section~\ref{sec:previous} below for a more detailed discussion).
+The extension to any relational structure will allow us to study general function complexity classes and the restriction to the natural numbers will simplify the notation in this context (see Section~\ref{sec:previous} for a more detailed discussion).
 
-Given a relational signature $\R$, the set of Quantitative Second-Order logic formulas (or just $\qso$-formulas) over $\R$ is given by the following grammar:
+Given a relational signature $\R$, the set of Quantitative Second-Order logic formulae (or just $\qso$-formulae) over $\R$ is given by the following grammar:
 %\[
@@
 where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
-If $\R$ is not mentioned, $\qso$ refers to the union of the sets of $\qso$ formulas over $\R$, for every relational signature~$\R$.
-\marcelo{En la gramatica de la logica deberiamos incluir la formula $\top$ que discutimos, la cual representa true y para la cual se tiene que $\sem{\top}(\A, v, V) = 1$. Les parece?}
+Moreover, if $\R$ is not mentioned, then $\qso$ refers to the union of the sets of $\qso$ formulae over all posible relational signatures.
+%\marcelo{En la gramatica de la logica deberiamos incluir la formula $\top$ que discutimos, la cual representa true y para la cual se tiene que $\sem{\top}(\A, v, V) = 1$. Les parece?}
 
-Note that the syntax of QSO formulas is divided in two levels. 
-The first level is composed by $\so$-formulas over $\R$ (called boolean formulas) and the second level is made by counting operators of addition and multiplication. 
+Note that the syntax of QSO formulae is divided in two levels. 
+The first level is composed by $\so$-formulae over $\R$ (called Boolean formulae) and the second level is made by counting operators of addition and multiplication. 
 For this reason, the quantifiers in $\so$ (e.g.
-$\exists x$ or $\exists X$) are called boolean quantifiers and the quantifiers that make use of addition and multiplication (e.g.
+$\exists x$ or $\exists X$) are called Boolean quantifiers and the quantifiers that make use of addition and multiplication (e.g.
 $\Sigma x$ or $\Pi X$) are called {\em quantitative quantifiers}.
-Furthermore, $\Sigma x$ and $\Sigma X$ are called first- and second-order sum, and $\Pi x$ and $\Pi X$ the first- and second-order product, respectively.
-This division between boolean and quantitative levels is essential for understanding the difference between the logic and the quantitative part. 
-Furthermore, this will allow us later to parametrize both levels of the logic in order to capture different function complexity classes.
+Furthermore, $\Sigma x$ and $\Sigma X$ are called first- and second-order sum, and $\Pi x$ and $\Pi X$ are called first- and second-order product, respectively.
+This division between Boolean and quantitative level is essential for understanding the difference between the logic and the quantitative part. 
+Furthermore, this will allow us later to parametrize both levels of the logic in order to capture different counting complexity classes.
 
-Let $\R$ be a relational signature, $\A$ be an ordered finite $\R$-structure with domain $A$, $v$ a first-order assignment for $\A$ and $V$ a second-order assignment for $\A$.
+Let $\R$ be a signature, $\A$ an $\R$-structure with domain $A$, $v$ a first-order assignment for $\A$ and $V$ a second-order assignment for $\A$.
 Then the \emph{evaluation} of a $\qso$-formula $\alpha$ over $(\A, v, V)$ is defined as a function $\sem{\alpha}$ that on input $(\A, v, V)$ returns a number in $\bbN$.
@@
 	\end{align*}
-	\caption{The semantics of QSO formulas.}
+	\caption{The semantics of QSO formulae.}
 	\label{tab-semantics}
@@
 A $\qso$-formula $\alpha$ is said to be a \emph{sentence} if it does not have any free variable, that is, every variable in $\alpha$ is under the scope of a usual quantifier or a quantitative quantifier.
-It is important to notice that if $\alpha$ is a $\qso$-sentence over a relational signature $\R$, then for every ordered finite $\R$-structure $\A$, first-order assignments $v_1$, $v_2$ for $\A$ and second-order assignments $V_1$, $V_2$ for $\A$, it holds that $\sem{\alpha}(\A, v_1, V_1) = \sem{\alpha}(\A, v_2, V_2)$.
+It is important to notice that if $\alpha$ is a $\qso$-sentence over a relational signature $\R$, then for every $\R$-structure $\A$, first-order assignments $v_1$, $v_2$ for $\A$ and second-order assignments $V_1$, $V_2$ for $\A$, it holds that $\sem{\alpha}(\A, v_1, V_1) = \sem{\alpha}(\A, v_2, V_2)$.
 Thus, in such a case we use the term $\sem{\alpha}(\A)$ to denote $\sem{\alpha}(\A, v, V)$, for some arbitrary first-order assignment $v$ for $\A$ and some arbitrary second-order assignment $V$ for $\A$.
 \begin{example}\label{ex:cliques}
-Let $\bG = \{E(\cdot,\cdot)\}$ be the vocabulary for graphs and $\fG$ be an ordered finite $\bG$-structure encoding a non-directed graphs. 
+Let $\bG = \{E(\cdot,\cdot),<\}$ be the vocabulary for graphs and $\fG$ be an ordered $\bG$-structure encoding a non-directed graph. 
 Suppose that we want to count the number of triangles in $\fG$. Then this can be defined with the following QSO-formula:
@@
 \alpha_1 \ := \ \sa{x} \sa{y} \sa{z} ( E(x,y) \, \wedge \, E(y,z) \, \wedge \, E(z,x) \, \wedge \\
-x \leq y \, \wedge \, y \leq z )
+x < y \, \wedge \, y < z )
 \end{multline*}
-For $\alpha_1$ we encode a triangle as an increasing sequence of nodes $\{x, y, z\}$ in order to count each triangle once. Then the boolean subformula  $E(x,y) \wedge E(y,z) \wedge E(z,x) \wedge
-x \leq y \wedge y \leq z$ is checking the triangle property, by outputting $1$ if $\{x, y, z\}$ forms a triangle in $\fG$ and $0$ otherwise.
+For $\alpha_1$ we encode a triangle as an increasing sequence of nodes $\{x, y, z\}$ in order to count each triangle once. Then the Boolean subformula  $E(x,y) \wedge E(y,z) \wedge E(z,x) \wedge
+x < y \wedge y < z$ is checking the triangle property, by outputting $1$ if $\{x, y, z\}$ forms a triangle in $\fG$ and $0$ otherwise.
 Finally, the sum quantifiers in $\alpha_1$ aggregates all the values, counting the number of triangles in $\fG$.
@@
 $$
-\alpha_2 := \sa{X} \big( \, \fa{x} \fa{y} \left(X(x) \wedge X(y)\right) \rightarrow E(x,y) \, \big)  
+\alpha_2 \ := \ \sa{X} \clique(X) ,
 $$ 
-Similar than for $\alpha_1$, in the boolean subformula of $\alpha_2$ we check whether $X$ is a clique and with the sum quantifier we add one for each clique in $\fG$. 
+where $\clique(X) := \forall x \forall y ((X(x) \wedge X(y) \wedge x \neq y)  \rightarrow E(x,y))$.
+Similar than for $\alpha_1$, in the Boolean subformula of $\alpha_2$ we check whether $X$ is a clique and with the sum quantifier we add one for each clique in $\fG$. 
 Note here that, in contrast to $\alpha_1$, in $\alpha_2$ we need a second-order quantifier in the quantitative level.
-This is according to the inherent complexity of evaluating each formula: $\alpha_1$ defines an $\fp$ function where $\alpha_2$ defines a $\shp$-hard function~\cite{paper-that-shows-that-this-problem-is-sharpP-hard}.
+This is according to the inherent complexity of evaluating each formula: $\alpha_1$ defines an $\fp$-function where $\alpha_2$ defines a $\shp$-complete function.
+%~\cite{paper-that-shows-that-this-problem-is-sharpP-hard}.
 \end{example}
 \begin{example}\label{exa-perm}
-For a more involved example that includes multiplication, let $\bM = \{M(\cdot,\cdot)\}$ be the vocabulary for matrices where a structure $\fM$ over $\bM$ encodes a 0-1 matrix $A$ where $M(i,j)$ is true if $A_{i,j} = 1$ and $0$ otherwise.
-Suppose now that we want to compute the permanent of an $n$-by-$n$ 0-1 matrix $A$ which is defined by:
-$$
-\op{perm}(A) \; = \; \sum_{\sigma \in S_n} \prod_{i=1}^n A_{i, \sigma(i)}  
-$$
+For a more involved example that includes multiplication, let $\bM = \{M(\cdot,\cdot)\}$ be a vocabulary for storing 0-1 matrices; in particular, a structure $\fM$ over $\bM$ encodes a 0-1 matrix $A$ as follows: if $A[i,j] = 1$, then $M(i,j)$ is true, otherwise $M(i.j)$ is false.
+Suppose now that we want to compute the permanent of an $n$-by-$n$ 0-1 matrix $A$, which is defined as:
+\begin{eqnarray*}
+\op{perm}(A) & = & \sum_{\sigma \in S_n} \prod_{i=1}^n A[i, \sigma(i)],  
+\end{eqnarray*}
 where $S_n$ is the set of all permutations over $\{1, \ldots, n\}$.
-The permanent is a relevant matrix function that has found many applications in different areas~\cite{permanent-applications} and it was one of the first function that was shown to be difficult for counting~\cite{Valiant79} (i.e.
-$\shp$-complete).
-For a binary relation symbol $S$, let $\op{permut}(S)$ be the $\fo$-formula that is true iff $S$ is a permutation, i.e. bijective function (the definition of $\op{permut}(S)$ is straightforward).
-Then a $\qso$-formula for defining the permanent of a matrix is as follows:
+The permanent is a fundamental function on matrices that has found many applications;
+% in different areas,
+%~\cite{permanent-applications},
+in fact, showing that this function is hard to compute was one of the main motivations behind the definition of the counting class $\shp$~\cite{Valiant79}.
+%and it was one of the first function that was shown to be difficult for counting~\cite{Valiant79} (i.e. $\shp$-complete).
+
+To define the permanent of a 0-1 matrix in $\qso$, assume that for a binary relation symbol $S$, $\op{permutation}(S)$ is an $\fo$-formula that is true if and only if $S$ is a permutation, that is, a total bijective function (the definition of $\op{permutation}(S)$ is straightforward).
+Then the following is a $\qso$-formula defining the permanent of a matrix:
 \[
-\alpha_3 := \sa{S} \op{permut}(S) \cdot \pa{x} (\ex{y} S(x,y) \wedge M(x,y))
+\alpha_3 := \sa{S} \op{permutation}(S) \cdot \pa{x} (\ex{y} S(x,y) \wedge M(x,y)).
 \]
-Intuitively, the subformula $\beta(s) := \pa{x} (\ex{y} S(x,y) \wedge M(x,y))$ calculates the value $\prod_{i=1}^n A_{i, \sigma(i)}$ whenever $S$ encodes a permutation $\sigma$.
-Then the subformula $\op{permut}(S) \cdot \beta(S)$ computes $\beta(S)$ when $S$ is a permutation and $0$ otherwise (i.e. $\op{permut}(S)$ behaves like a filter). 
+Intuitively, the subformula $\beta(s) := \pa{x} (\ex{y} S(x,y) \wedge M(x,y))$ calculates the value $\prod_{i=1}^n A[i, \sigma(i)]$ whenever $S$ encodes a permutation $\sigma$.
+Moreover, the subformula $\op{permutation}(S) \cdot \beta(S)$ returns $\beta(S)$ when $S$ is a permutation, and returns $0$ otherwise (i.e. $\op{permutation}(S)$ behaves like a filter). 
 Finally, the second order sum aggregates these values iterating over all binary relations and calculating the permanent of the matrix.
-We would like to end this example by highlighting the similarity of $\alpha_3$ with the permanent formula. 
-Indeed, an advantage of $\qso$-formulas is that the first- and second-order quantifiers naturally reflect the operations used to define mathematical formulas in practice.
+We would like to finish with this example by highlighting the similarity of $\alpha_3$ with the permanent formula. 
+Indeed, an advantage of $\qso$-formulae is that the first- and second-order quantifiers in the quantitative level naturally reflect the operations used to define mathematical formulae.
 \end{example}
 
-We consider several fragments of $\qso$, which are obtained by restricting the syntax of the boolean formulas or the use of the quantitative quantifiers.
+We consider several fragments of $\qso$, which are obtained by restricting the syntax of the Boolean formulae or the use of the quantitative quantifiers.
 In this direction, we denote by $\qfo$ the fragment of $\qso$ where second-order sum and product are not allowed. 
-For example, $\alpha_1$ is in $\qfo$ and $\alpha_2$ is not.
-Another interesting fragment are the formulas of $\qso$ where only sum operators and quantifiers are allowed. 
+For instance, for the $\qso$-formulae defined in Example \ref{ex:cliques}, we have that $\alpha_1$ is in $\qfo$ and $\alpha_2$ is not.
+Another interesting fragment of $\qso$ consists of the $\qso$-formulae where only sum operators and quantifiers are allowed. 
 Formally, we denote by $\eqso$ the fragment of $\qso$ where first- and second-order products (i.e. $\pa{x}$ and $\pa{X}$) are not allowed.
-For example, $\alpha_1$ and $\alpha_2$ are formulas of $\eqso$ but $\alpha_3$ is not. 
-We also consider fragments of $\qso$ by further restricting the boolean core of the logic.
-Let $\LL$ be a fragment of $\so$.
-We define the quantitative logic $\qso(\LL)$ to be the fragment of $\qso$ obtained by restricting $\varphi$ to be a formula in $\LL$. 
-In general, if $\FF$ is a fragment of $\qso$ we define $\FF(\LL)$ to be the fragment of $\FF$ obtained by restricting boolean formulas to $\LL$.
-For example, we can define the fragments $\qfo(\fo)$ or $\eqso(\fo)$ which are subfragments of $\qfo$ and $\eqso$ where $\varphi$ is restricted to $\fo$ formulas.
+For example, $\alpha_1$ and $\alpha_2$ in Example \ref{ex:cliques} are formulae of $\eqso$, while $\alpha_3$ in Example \ref{exa-perm} is not. 
+We also consider fragments of $\qso$ by further restricting the Boolean part of the logic.
+If $\LL$ is a fragment of $\so$, then we define the quantitative logic $\qso(\LL)$ to be the fragment of $\qso$ obtained by restricting $\varphi$ in \eqref{syntax} to be a formula in $\LL$.
+Moreover, we also restrict other fragments of $\qso$ by using the same idea.
+%In general, if $\FF$ is a fragment of $\qso$ we define $\FF(\LL)$ to be the fragment of $\FF$ obtained by restricting Boolean formulae to $\LL$.
+For example, we define $\qfo(\fo)$ to be the fragment of $\qfo$ obtained by restricting $\varphi$ in \eqref{syntax} to be an $\fo$-formula, and likewise for $\eqso(\fo)$.
+%, $\eqso(\fo)$ of $\qfo$ and $\eqso$ where $\varphi$ is restricted to $\fo$ formulae.
 
-\marcelo{IMPORTANTE: Tenemos que decir que $\qfo(\fo)$ permite variables libre de segundo orden en las formulas de primer orden. Hay que aclarar esto tambien para PFP y otras logicas.}
+%\marcelo{IMPORTANTE: Tenemos que decir que $\qfo(\fo)$ permite variables libre de segundo orden en las formulae de primer orden. Hay que aclarar esto tambien para PFP y otras logicas.}
 
-In the next section, we variate $\FF$ and $\LL$ to capture several function complexity classes.
-Before this, we devote the next subsection to show the connection of $\qso$ with previous frameworks for defining functions over relational structures.
+In the following section, we use different fragments of $\qso$ to capture counting complexity classes. 
+%variate $\FF$ and $\LL$ to capture several function complexity classes. 
+But before doing this, we 
+%devote the next subsection to 
+show the connection of $\qso$ with previous frameworks for defining functions over relational structures.
 
@@
 
-In this subsection we discuss previous frameworks proposed in the literature and how they differ to our approach.
+In this subsection, we discuss some previous frameworks proposed in the literature and how they differ from our approach.
 We start by discussing the connection between $\qso$ and weighted logics (WL)~\cite{DrosteG07}. 
 As it was previously discussed, $\qso$ is a fragment of WL.
-The real difference is that we restrict the semiring used in WL to natural numbers in order to study function complexity classes.
-Another difference of WL with our approach is that, as far as we know, this is the first paper to study weighted logics over general signature properly to do descriptive complexity of counting classes.
-Previous works on WL usually restrict the signature of the logic to strings, trees, and other specific structures (see \cite{droste2009handbook} for more examples), and did not study the logic over general structures.
-Furthermore, in this paper we propose further extensions for $\qso$ (see Section~\ref{sec:beyond}) which start to differ from previous approaches in WL.
+The main difference is that we restrict the semiring used in WL to natural numbers in order to study counting complexity classes.
+Another difference of WL with our approach is that, to the best of our knowledge, this is the first paper to study weighted logics over general relational signatures to do descriptive complexity for counting complexity classes.
+Previous works on WL usually restrict the signature of the logic to strings, trees, and other specific structures (see \cite{droste2009handbook} for more examples), and they did not study the logic over general structures.
+Furthermore, in this paper we propose further extensions for $\qso$ (see Section~\ref{sec:beyond}) which differ from previous approaches in WL.
 
-Another approach that resembles $\qso$ are counting logics~\cite{L04,counting-paper-here}. 
-Logics with counting operators usually extend $\fo$ with quantifiers that counts how many ways can be satisfied a formula over a second sort (i.e. natural numbers). 
-In contrast to our approach, counting operators are usually used for checking boolean properties over structures and not for outputting values (i.e. they do not define a function).
-Although there are some papers that uses this operator for counting~\cite{XXX}, they have not used the logic for capturing function complexity classes.
+Another approach that resembles $\qso$ are logics with counting~\cite{IL90,E97,GG98,L04}, which include operators that extend $\fo$ with quantifiers that allow to count in how many ways a formula  is satisfied (the result of this counting is a value of a second sort, in this case the  natural numbers).
+In contrast to our approach, counting operators are usually used for checking Boolean properties over structures and not for producing values (i.e. they do not define a function).
+In particular, we are not aware of any paper that uses this approach for capturing counting complexity classes.
+%Although there are some papers that uses this operator for counting~\cite{XXX}, they have not used the logic for capturing function complexity classes.
 
@@
 
-Finally, the work on~\cite{SalujaST95} and \cite{ComptonG96} is of particular interest for our work. 
-In~\cite{SalujaST95}, it was proposed to define functions over structures by using free variables on SO-formulas and count how many extensions of the structure satisfy the formula (i.e.
-by instantiating the free variables).
+Finally, the work in~\cite{SalujaST95} and \cite{ComptonG96} is of particular interest for our research. 
+In~\cite{SalujaST95}, it was proposed to define a function over a structure by using free variables in an SO-formula; in particular, the function is defined by the number of instantiations of the free variables that are satisfied by the structure.
+%and then counting how many extensions of an structure satisfy the formula (i.e. by instantiating the free variables). 
 Formally, Saluja et.
-al \cite{SalujaST95} define a family of counting classes $\#\LL$ for each fragment $\LL$ of $\so$.
-For a formula $\varphi(\bar{x},\bar{X})$ over $\R$, the function $f_{\varphi(x,X)}$ is defined as
+al \cite{SalujaST95} define a family of counting classes $\#\LL$ for a fragment $\LL$ of $\fo$.
+For a formula $\varphi(\bar{x},\bar{X})$ over $\R$, the function $f_{\varphi(\bar x, \bar X)}$ is defined as
 $
-f_{\varphi}(\A) = \vert \{\langle \bar{a},\bar{A} \rangle\mid \A\models\varphi(\bar{a},\bar{A})\}\vert.
+f_{\varphi(\bar x, \bar X)}(\A) = \vert \{(\bar{a},\bar{A}) \mid \A\models\varphi(\bar{a},\bar{A})\}\vert,
 $
-for each $\A\in\ostr[\R]$. Then a function $g:\ostr[\R]\to\nat$ is in $\#\LL$ if there exists an formula $\varphi(\bar{x},\bar{X})$ in $\LL$ such that $f = f_{\varphi}$. 
-In~\cite{SalujaST95}, it was shown several results for capturing function complexity classes which are relevant for our work.
+for every $\A\in\ostr[\R]$. Then a function $g:\ostr[\R]\to\nat$ is in $\#\LL$ if there exists a formula $\varphi(\bar{x},\bar{X})$ in $\LL$ such that $g = f_{\varphi(\bar x, \bar X)}$. 
+In~\cite{SalujaST95}, they were proved several results about capturing counting complexity classes which are relevant for our work.
 We discuss and use these results in Sections~\ref{sec:complexity} and~\ref{sec:syntactic}.
-Note that for any formula $\varphi(\bar{x},\bar{X})$ it holds that $f_\varphi \equiv \sem{\sa{\bar{X}} \sa{\bar{x}} \varphi(\bar{x},\bar{X})}$, that is, the approach in \cite{SalujaST95} can be seen as a syntactical restriction of $\qso$.
-Thus, the advantage of our approach relies on the flexibility of defining  counting functions by alternating sum with product operators or, moreover, by introducing new quantitative operators (like in Section~\ref{sec:beyond}).
-Furthermore, in the next section we show how to capture some interesting classes that could never be captured by following the approach in~\cite{SalujaST95}.
+Notice that for every formula $\varphi(\bar{x},\bar{X})$, it holds that $f_{\varphi(\bar{x},\bar{X})}$ is the same function as $\sem{\sa{\bar{X}} \sa{\bar{x}} \varphi(\bar{x},\bar{X})}$, that is, the approach in \cite{SalujaST95} can be seen as a syntactical restriction of our approach based on $\qso$.
+Thus, the advantage of our approach relies on the flexibility for defining  functions by alternating sum with product operators and, moreover, by introducing new quantitative operators (see Section~\ref{sec:beyond}).
+Furthermore, we show in the following section how to capture some fundamental classes that cannot be captured by following the approach in~\cite{SalujaST95}.

### 2017-01-07 LICS/preliminaries.tex
sc

@@
 Notice that every variable in $\sv$ has an associated arity, which is denoted by $\arity(X)$.
-Then given a relational signature $\R$, the set of second-order logic formulas ($\so$-formulas) over $\R$ is given by the following grammar:
+Then given a relational signature $\R$, the set of second-order logic formulae ($\so$-formulae) over $\R$ is given by the following grammar:
 \begin{multline*}\ 
@@
 In this paper we consider several fragments of $\so$, in particular first-order logic ($\fo$), least fixed point logic (LFP) and partial fixed point logic (PFP) \cite{L04}.
-Moreover, for every $i \in \N$, we consider the fragment $\Sigma_i$ (resp., $\Pi_i$) of $\fo$, which is the set of $\fo$-formulas of the form
+Moreover, for every $i \in \N$, we consider the fragment $\Sigma_i$ (resp., $\Pi_i$) of $\fo$, which is the set of $\fo$-formulae of the form
 $\exists \bar x_1 \forall \bar x_2 \cdots \exists \bar x_{i-1} \forall \bar x_{i} \, \psi$ (resp., 
@@
 %Moreover, we use the abbreviations $x \not \leq y$ and $x \notin X$ for the negation of the atoms $\leq$ and $\in$. 
-%Finally, we consider standard abbreviations of formulas that can be defined in $\so$-logic (actually, $\fo$-logic) like $\first(x) := \fa{x} y \leq x$ and $\last(x) := \fa{x} x \leq y$ to denote the first and last element of the linear order $\leq$, respectively, and $\succesor(x,y) := x \leq y \wedge y \not\leq x \wedge \fa{z} ( z \leq x \vee y \leq z)$ to denote the successor relation.
+%Finally, we consider standard abbreviations of formulae that can be defined in $\so$-logic (actually, $\fo$-logic) like $\first(x) := \fa{x} y \leq x$ and $\last(x) := \fa{x} x \leq y$ to denote the first and last element of the linear order $\leq$, respectively, and $\succesor(x,y) := x \leq y \wedge y \not\leq x \wedge \fa{z} ( z \leq x \vee y \leq z)$ to denote the successor relation.
 

### 2017-01-08 LICS/beyond.tex
Avances en sección 6.

@@
 
-We define the extension of $\qso$ with function symbols as follows.
+In the previous sections, we use weighted logics to give a framework for descriptive complexity of functions.
+Here, we go beyond weighted logics to add unexplored quantifiers at the quantitative level.
+Specifically, we give the first steps on defining quantitative recursion.
+This goal is not trivial for two reasons: (1) we want to add recursion over functions which, in general, it is a difficult task and (2) it is not clear what is the right notion of ``fixed point'' for recursive functions.
+Towards this goal, we show first how to extend $\qso$ with function symbols to later use them to define recursion at a quantitative level.
+Furthermore,  we define some sort of least fixed point that we called \emph{least s-fixed point} and, as a proof of concept, we show that it captures $\fp$.
+Further, we use our new tool to define an operator for counting paths in a graph, a natural generalization of the transitive closure operator~\cite{immerman1999descriptive}, and show that this captures the expressiveness of~$\shl$.
+
+For defining recursion, we first need to define an extension of $\qso$ with function symbols.
 Assume that $\fs$ is an infinite set of function symbols, where each $h \in \fs$ has an associated arity, which is denoted by $\arity(h)$.
@@
 where $h \in \fs$, $\arity(h) = \ell$ and $x_1, \ldots, x_\ell$ is a sequence of (non-necessarily distinct) first-order variables.
-Given an $\R$-structure $\A$ with domain $A$, we say that $F$ is a function assignment for $\A$ if for every $h \in \fs$ with $\arity(h) = \ell$, we have that $F(h) :  A^\ell \to \N$.
+Given an $\R$-structure $\A$ with domain $A$, we say that $F$ is a \emph{function assignment} for $\A$ if for every $h \in \fs$ with $\arity(h) = \ell$, we have that $F(h) :  A^\ell \to \N$.
 The notion of function assignment is used to extend the semantics of $\qso$ to the case of a quantitative formula of the form $h(x_1, \ldots, x_\ell)$.
@@
 \end{eqnarray*}
-As for the case of $\qfo$, we define $\fqfo$ disallowing quantifiers $\Sigma X$ and $\Pi X$ in \eqref{eq-fqso}, and we define $\fqfo(\fo)$ by also imposing the restriction that $\varphi$ in \eqref{eq-fqso} is a first-order formula.
+As for the case of $\qfo$, we define $\fqfo$ disallowing quantifiers $\Sigma X$ and $\Pi X$ in \eqref{eq-fqso}.
 
-\subsection{A logic with a fixed point operator}
+It worths noting that function symbols represent in $\fqso$ functions from tuples to numbers and cannot be compared to the classical notion of (boolean) function symbol in first-order logics~\cite{enderton2001mathematical}.
+Furthermore, one could see a function symbol as an ``oracle'' that is instantiated by the function assignment and is used to express another function.
+As far as we know, this is the first paper proposing this extension on weighted logics and it would be interesting to study this concept further (i.e.
+outside recursion).
 
 We define an operator which extends Least Fixed Point Logic (LFP) \cite{I86,vardi1982complexity} to allow counting. 
-Fix a relational signature $\R$. Then the set of $\rqfo(\fo)$ formulas over $\R$ is defined by the following grammar:
-\begin{multline*}
-	\alpha := \varphi \, \mid \, s \, \mid \, (\alpha \add \alpha) \, \mid\, (\alpha \mult \alpha) \, \mid \\
-	\sa{x} \alpha \, \mid\, 
-	\pa{x} \alpha \, \mid\,
-	\clfp{\beta(x_1, \ldots, x_\ell, h)},
-\end{multline*}
-where $\varphi$ is an $\fo$-formula, $x_1, \ldots, x_\ell$ is a sequence of pairwise distinct first-order variables and $\beta(x_1, \ldots, x_\ell, h)$ is an $\fqfo(\fo)$-formula over $\R$ whose only function symbol is $h$, with $\arity(h) = \ell$.
-The free variables of the formula $\clfp{\beta(x_1,\ldots,x_\ell,h)}$ are $x_1, \ldots, x_\ell$; in particular, $h$ is not considered to be free in this formula.
+We show how to define this operator in $\qfo(\fo)$ and we discuss how to extend these ideas to $\qso$ towards the end of the section. 
+Fix a relational signature $\R$.
+Then the set of $\rqfo(\fo)$ formulas (Recursive $\qfo$) over $\R$ is defined as an extension of $\qfo(\fo)$ by adding the operator $\clfp{\beta(\x, h)}$ where $\x = (x_1, \ldots, x_\ell)$ is a sequence of $\ell$ distinct first-order variables and $\beta(\x, h)$ is an $\fqfo(\fo)$-formula over $\R$ whose only function symbol is $h$, with $\arity(h) = \ell$.
+The free variables of the formula $\clfp{\beta(\x,h)}$ are $x_1, \ldots, x_\ell$; in particular, $h$ is not considered to be free in this formula.
 
-Fix an $\R$-structure with domain $A$ and a quantitative formula $\clfp{\beta(x_1,\ldots,x_\ell,h)}$, and assume that $\F$ is the set of functions $f :A^\ell \to \N$.
-To define the semantics of $\clfp{\beta(x_1,\ldots,x_\ell,h)}$, we first show how $\beta(x_1,\ldots,x_\ell,h)$ can be interpreted as an operator $T_{\beta}$ on $\F$.
-More precisely, for every $f \in \F$ and tuple $(a_1, \ldots, a_\ell) \in A^\ell$, the function $T_{\beta}(f)$ satisfies that:
+Fix an $\R$-structure with domain $A$ and a quantitative formula $\clfp{\beta(\x,h)}$, and assume that $\F$ is the set of functions $f :A^\ell \to \N$.
+To define the semantics of $\clfp{\beta(\x,h)}$, we first show how $\beta(\x,h)$ can be interpreted as an operator $T_{\beta}$ on $\F$.
+More precisely, for every $f \in \F$ and tuple $\a = (a_1, \ldots, a_{\ell}) \in A^\ell$, the function $T_{\beta}(f)$ satisfies that:
 \begin{eqnarray*}
-T_{\beta}(f)(a_1, \ldots, a_\ell) & = & \sem{\beta(x_1, \ldots, x_\ell, h)}(\A,v,F),
+T_{\beta}(f)(\a) & = & \sem{\beta(\x, h)}(\A,v,F),
 \end{eqnarray*}
@@
 
-As for the case of LFP, it would be natural to consider the partial order $\leq$ on $\F$ defined as $f \leq g$ if and only if $f(i) \leq g(i)$ for every $i \in \{1, \ldots, \ell\}$, and let the semantics of $\clfp{\beta(x_1,\ldots,x_\ell,h)}$ be the least fixed point of the operator $T_\beta$.
-However, $(\F, \leq)$ is not a complete lattice, so we do not have a Knaster-Tarski Theorem ensuring that such a fixed point exist.
+As for the case of LFP, it would be natural to consider the point-wise partial order $\leq$ on $\F$ defined as $f \leq g$ if, and only if, $f(i) \leq g(i)$ for every $i \in \{1, \ldots, \ell\}$, and let the semantics of $\clfp{\beta(\x,h)}$ be the least fixed point of the operator $T_\beta$.
+However, $(\F, \leq)$ is not a complete lattice, so we do not have a Knaster-Tarski Theorem ensuring that such a fixed point exists.
 Instead, we generalize the semantics of LFP as follows.
-In the definition of the semantics of LFP, an operator $T$ on relations is considered, and the semantics is defined in terms of the least fixed point of $T$, that is, a relation $R$ such that: (a) $T(R) = R$, and (2) $R \subseteq S$ for every $S$ such that $T(S) = S$ \cite{I86,vardi1982complexity}.
+In the definition of the semantics of LFP, an operator $T$ on relations is considered, and the semantics is defined in terms of the least fixed point of $T$, that is, a relation $R$ such that~\cite{I86,vardi1982complexity}:
+\begin{enumerate}
+ \item[(a)] $T(R) = R$, and 
+ \item[(b)]  $R \subseteq S$ for every $S$ such that $T(S) = S$.  
+\end{enumerate}
 We can view $T$ as an operator on functions if we consider the characteristic function of a relation.
@@
 In particular, the conditions defining the least fixed point of $T$ can be rewritten as follows in terms of $T^\star$:
-$\support(T^\star(\chi_R)) = \support(\chi_R)$, and $\support(\chi_R) \subseteq \support(\chi_S)$ for every $S$ such that  $\support(T^\star(\chi_{S})) = \support(\chi_S)$. 
+\begin{enumerate}
+	\item[(a)] $\support(T^\star(\chi_R)) = \support(\chi_R)$, and  
+	\item[(b)] $\support(\chi_R) \subseteq \support(\chi_S)$ for every $S$ such that  $\support(T^\star(\chi_{S})) = \support(\chi_S)$.  
+\end{enumerate}
 To define a notion of fixed point for $T_\beta$ we simply generalized these conditions. More precisely, %These conditions can be generalized in the obvious way to the case of the operator $T_\beta$: 
@@
 \end{lemma}
-In fact, as for the case of LFP, this lemma gives us a simple way to compute the least s-fixed point of $T_\beta$.
-Let $f_0 \in \F$ be a function such that $f_0(\bar a) = 0$ for every $\bar a \in A^\ell$, and let function $f_{i+1}$ be defined as $T_\beta(f_i)$ for every $i \in \N$.
+In fact, as for the case of LFP, this lemma gives us a simple way to compute a least s-fixed point of $T_\beta$.
+Let $f_0 \in \F$ be a function such that $f_0(\bar a) = 0$ for every $\bar a \in A^\ell$ (i.e.
+$f_0$ is the only function with empty support), and let function $f_{i+1}$ be defined as $T_\beta(f_i)$ for every $i \in \N$.
 Then there exists $j \geq 0$ such that $\support(f_j) = \support(T_\beta(f_j))$.
 Let $k$ be the smallest natural number such that $\support(f_{k}) = \support(T_\beta(f_k))$.
-We have that $f_k$ is a least s-fixed point of $T_\beta$, which is used to defined the semantics of $\clfp{\beta(x_1, \ldots, x_\ell, h)}$.
+We have that $f_k$ is a least s-fixed point of $T_\beta$, which is used to defined the semantics of $\clfp{\beta(\x, h)}$.
 More specifically, for an arbitrary first-order assignment $v$ for $\A$:
 \begin{eqnarray*}
-\sem{\clfp{\beta(x_1, \ldots, x_\ell, h)}}(\A,v) & = & f_{k}(v(x_1), \ldots, v(x_\ell))
+\sem{\clfp{\beta(\x, h)}}(\A,v) & = & f_{k}(v(\x))
 \end{eqnarray*}
 
-\begin{example}
+\begin{example} \label{ex:count-path}
 %As an example, 
@@
 
-Assume now that we need to extend our previous counting function to the case of arbitrary directed graphs. To this end, define the following auxiliary $\fo$-formulas:
-\begin{eqnarray*}
-\varphi_{\text{\rm first}}(x) & = & \neg \exists z (z < x)\\
-\varphi_{\text{succ}}(x,y) & = & x < y \wedge \neg \exists z ( x < z \wedge z < y).
-\end{eqnarray*}
-Moreover, define formula $\beta(x, y, u, g)$ as follows:
+Assume now that we need to extend our previous counting function to the case of arbitrary directed graphs.
+To this end, suppose that $\varphi_{\text{\rm first}}(x)$ and $\varphi_{\text{succ}}(x,y)$ are the $\fo$-formulas for defining the first and successor predicates, respectively, of the total-order $<$.
+Moreover, define formula $\beta(x, y, t, g)$ as follows:
 \begin{multline*}
-(E(x,y) + \sa{z} g(x,z,u)\cdot E(z,y)) \cdot \varphi_{\text{\rm first}}(u) \ +\\
-\varphi_{\text{succ}}(v,u) \cdot g(x,y,v)
+(E(x,y) + \sa{z} g(x,z,t)\cdot E(z,y)) \cdot \varphi_{\text{\rm first}}(t) \ +\\
+\sa{t'} \varphi_{\text{succ}}(t',t) \cdot \left(\sa{x'} \sa{y'} g(x',y',t') \right)
 \end{multline*}
@@
 $$
-\sa{u} (\varphi_{\text{\rm first}}(u) \wedge \clfp{\beta(x,y,u,g)}).
+\sa{t} (\varphi_{\text{\rm first}}(t) \wedge \clfp{\beta(x,y,t,g)}).
 $$ 
-In fact, the number of paths of length at most $n$ from a node $x$ to a node $y$ is recursively computed by using the formula $(E(x,y) + \sa{z} g(x,z,u)\cdot E(z,y)) \cdot \varphi_{\text{\rm first}}(u)$, which stores this value in $g(x,y,u)$ with $u$ the first element in the domain.
-The other formula $\varphi_{\text{succ}}(v,u) \cdot g(x,y,v)$ is used as a counter that allows to reach a fixed point in the support of function $g$ in $n$ steps.
+In fact, the number of paths of length at most $n$ from a node $x$ to a node $y$ is recursively computed by using the formula $(E(x,y) + \sa{z} g(x,z,t)\cdot E(z,y)) \cdot \varphi_{\text{\rm first}}(t)$, which stores this value in $g(x,y,t)$ with $t$ the first element in the domain.
+The other formula, the one with the filter $\varphi_{\text{succ}}(t',t)$, is used as a counter that allows to reach a fixed point in the support of function $g$ in $n$ steps.
 \end{example}
@@
 %\end{theorem} 
-In the following result we show how the fixed point operator introduced in this section can be used to capture (over the class of ordered structures) the fundamental class of functions that can be computed in polynomial time.
+Note that, in contrast to $\lfp$-logic, here we do not need to impose any positive restriction to the formula $\beta(\x,h)$.
+Indeed, since $\beta$ is constructed over the natural numbers based on the monotone operators of sum and product, it is not necessary to add any restriction to the base formula.
+In the following result we show how the fixed point operator introduced in this section can be used to capture the fundamental class of functions that can be computed in polynomial time.
 \begin{theorem} \label{rqfo-fo-cap}
@@
 \end{theorem}
-
-\marcelo{Notese que estoy permitiendo en la formula $\beta(y_1, \ldots, y_\ell,R,\pi)$ tener subformulas $\pa{x} \alpha$. Esta bien el resultado con esto? O tenemos que eliminar estas formulas?}
 
@@
 
-\subsection{A logic with an operator for counting paths}
-
 It was shown in \cite{I86,I88} that $\fo$ extended with a transitive closure operator captures $\nlog$.
 Moreover, it was shown in \cite{I83} that $\so$ extended with a transitive closure operator captures $\pspace$.
-Inspired by this work, we extend the definition of $\qfo$ with an operator for counting the number of paths in a directed graph, and we show that it can be used to capture $\shl$.
-Besides, we show that the same idea can be used to extend $\qso$ allowing to capture harder complexity classes.
+Inspired by this work, we extend the definition of $\qfo$ with an operator for counting the number of paths in a directed graph defined in terms of $\rqfo$.
+Then we show that it can be used to capture $\shl$. Besides, we show that the same idea can be used to extend $\qso$ allowing to capture harder complexity classes.
 
-Given a relation signature $\R$, the set of transitive $\qso$ formulas ($\tqso$-formulas) is defined by the following grammar:
-\begin{multline}
-	\label{eq-def-tqso}
-	\alpha := \varphi \, \mid \, s \, \mid \, (\alpha \add \alpha) \, \mid\, (\alpha \mult \alpha) \, \mid \, 
-	\sa{x} \alpha \, \mid\, \
-	\pa{x} \alpha \, \mid \\ 
-	\sa{X} \alpha \, \mid \, \pa{X} \alpha \, \mid \, [\pth \psi(\bar{x}, \bar{X},\bar{y}, \bar{Y})],
-\end{multline}
-where $\varphi$ is an $\so$-formula over $\R$, $\psi(\bar{x},\bar{X},\bar{y},\bar{Y})$ is an $\fo$-formula over $\R$, $\bar{x} = (x_1, \ldots, x_k)$, $\bar{y} = (y_1, \ldots, y_k)$ are tuples of pairwise distinct first-order variables, and $\bar{X} = (X_1, \ldots, X_\ell)$, $\bar{Y} = (Y_1, \ldots, Y_\ell)$ are tuples of pairwise distinct second-order variables such that $\arity(X_i) = \arity(Y_i) = m_i$ for every $i \in \{1, \ldots, \ell\}$.
-The semantics of $[\pth \psi(\bar{x},\bar{X},\bar{y}.\bar{Y})]$ is defined as follows.
-Given an $\R$-structure $\A$, define a (directed) graph $\cG_{\psi}(\A) = (N,E)$ such that $N = A^k \times \prod_{i=1}^\ell 2^{A^{m_i}}$, where $A$ is the domain of $\A$, and for every pair $(\bar b, \bar B)$, $(\bar c, \bar C)$ of elements of $N$, it holds that $((\bar b, \bar B), (\bar c, \bar C)) \in E$ if and only if $\A \models \psi(\bar b, \bar{B}, \bar c, \bar{C})$.
-Then given first-order and second-order assignments $v$, $V$ for $\A$, we have that $\sem{[\pth \psi(\bar{x},\bar{X}, \bar{y}, \bar{Y})]}(\A,v,V)$ is the number of paths in $\cG_\psi(\A)$ from $(v(\bar x), V(\bar X))$ to $(v(\bar y), V(\bar Y))$ whose length is at most $n = |N|$.
-
-As for the case of $\qso$, the logic $\tqso(\LL)$ is obtained by restricting $\varphi$ in \eqref{eq-def-tqso} to be a formula in $\LL$.
+Given a relation signature $\R$, the set of transitive $\qfo$ formulas ($\tqfo$-formulas) is defined as an extension of $\qfo$ with the operator:
+%\begin{multline}
+%	\label{eq-def-tqso}
+%	\alpha := \varphi \, \mid \, s \, \mid \, (\alpha \add \alpha) \, \mid\, (\alpha \mult \alpha) \, \mid \, 
+%	\sa{x} \alpha \, \mid\, \
+%	\pa{x} \alpha \, \mid \\ 
+%	\sa{X} \alpha \, \mid \, \pa{X} \alpha \, \mid \, [\pth \psi(\bar{x}, \bar{X},\bar{y}, \bar{Y})],
+%\end{multline}
+\[
+[\pth \psi(\bar{x},\bar{y})]
+\]
+where $\psi(\x, \y)$ is an $\fo$-formula over $\R$, $\bar{x} = (x_1, \ldots, x_k)$, $\bar{y} = (y_1, \ldots, y_k)$ are tuples of pairwise distinct first-order variables.
+The semantics of $[\pth \psi(\bar{x},\bar{y})]$ can easily be defined by using $\rqfo(\fo)$ as follows.
+Similar than Example~\ref{ex:count-path}, suppose that the formula $\psi(\bar{x},\bar{y})$ define a (directed) graph over tuples in $A^k$ ($A$ is the domain of $\A$) such that for every $\a, \b \in A^k$ there exists an edge $(\a, \b)$ if $\A \models \psi(\bar a, \bar b)$.
+Define formula $\beta_{\psi(\bar{x},\bar{y})}(\x, \y, \t, g)$ as follows:
+\begin{multline*}
+(\psi(\bar{x},\bar{y}) + \sa{\z} g(\x,\z,\t)\cdot \psi(\z,\y)) \cdot \varphi_{\text{\rm first}}(\t) \ +\\
+\sa{\t'} \varphi_{\text{succ}}(\t',\t) \cdot \left(\sa{\x'} \sa{\y'} g(\x',\y',\t') \right)
+\end{multline*}
+where $\varphi_{\text{\rm first}}$ and $\varphi_{\text{succ}}$ define now the first and successor predicates over the lexicographic order of tuples induced by~$<$.
+Then the semantics of the path operator is defined by the semantics of the following equivalent $\rqfo$ formula:
+\[
+[\pth \psi(\bar{x}, \bar{y})] := \sa{\t} (\varphi_{\text{\rm first}}(\t) \wedge \clfp{\beta_{\psi(\bar{x},\bar{y})}(\x,\y,\t,g)})
+\]
+In other words, $\sem{[\pth \psi(\bar{x}, \bar{y})]}(\A,v,V)$ is the number of paths in the graph induced by $\psi(\bar{x}, \bar{y})$ whose length is at most~$|A|$.
+%As for the case of $\qso$, the logic $\tqso(\LL)$ is obtained by restricting $\varphi$ in \eqref{eq-def-tqso} to be a formula in $\LL$.
 Moreover, the logic $\tqfo$ is obtained by disallowing in \eqref{eq-def-tqso} formulas $\sa{X} \alpha$ and $\pa{X} \alpha$, and by only allowing  first-order free-variables in the formula $\psi$ used in $[\pth \psi]$ in \eqref{eq-def-tqso}.
-With this notation, we have the following results:
+With this logic, we have the following results:
 
@@
 \end{theorem}
+
+\cristian{Aca voy}
 
@@
 where $\psi(\bar{x}, \bar{X},\bar{y}, \bar{Y})$ satisfies the same conditions as in grammar \eqref{eq-def-tqso}. Then we have that:
-\begin{theorem} \label{tqfo-shl}
+\begin{theorem} \label{tqsos-shp}
 	$\tqsos(\fo)$ captures $\shp$ over the class of ordered structures.

### 2017-01-08 LICS/counting.tex
sc

@@
 
-In this section, we show that by syntactically restricting $\qso$ one can capture different function complexity classes. 
-In other words, by using $\qso$ we can extend the theory of descriptive complexity~\cite{immerman2012descriptive} from decision problems to counting problems. 
-For this, we first formalize the notion of \emph{capturing} a function complexity class to then show how to capture classes like $\shp$, $\fp$, and $\fpspace$.
+In this section, we show that by syntactically restricting $\qso$ one can capture different counting complexity classes. 
+In other words, by using $\qso$ we can extend the theory of descriptive complexity~\cite{immerman1999descriptive} from decision problems to computation problems. 
+For this, we first formalize the notion of \emph{capturing} a complexity class of functions.
+%, and then show how to capture classes like $\shp$, $\fp$, and $\fpspace$.
 
-Fix a relational signature $\R = \{R_1, \ldots, R_k\}$ and assume that $\A$ is an ordered finite $\R$-structure with a domain $A$ containing $n$ elements.
+Fix a relational signature $\R = \{R_1, \ldots, R_k\}$ and assume that $\A$ is an ordered (finite) $\R$-structure with a domain $A = \{a_1, \ldots, a_n\}$.
+%$ containing $n$ elements. 
 Recall that  $<$ is a linear order on $A$, say $a_1 < a_2 < \ldots < a_n$.
@@
 %Note that the function $g$ outputs $f$ for encodings of structures and can behave arbitrarily otherwise.
-We can now formalize the notion of capturing a function complexity class over $\R$-structures.
+We can now formalize the notion of capturing a counting complexity class.
+%over $\R$-structures.
 \begin{definition} \label{def:cap}
-	Let $\FF$ be a fragment of $\qso$ and $\CC$ a function complexity class. Then we say that {\em  $\FF$ captures $\CC$ over $\R$-structures} if the  following conditions are true:
+	Let $\FF$ be a fragment of $\qso$ and $\CC$ a counting complexity class. Then {\em  $\FF$ captures $\CC$ over the class of ordered $\R$-structures} if the  following conditions hold:
 	\begin{enumerate}
@@
 	\end{enumerate} 
-	Finally, we say that $\FF$ captures $\CC$ if $\FF$ captures $\CC$ over $\R$-structures for every relational signature $\R$.
+	Moreover, {\em $\FF$ captures $\CC$ over ordered structures} if $\FF$ captures $\CC$ over the class of ordered $\R$-structures for every relational signature $\R$.
 \end{definition}
 %For the sake of simplification, we denote the first condition by $\FF \subseteq \CC$ and the second condition by $\CC \subseteq \FF$.
-Note that in Definition~\ref{def:cap} the function $f \in \CC$ and the formula $\alpha \in \FF$ must coincide in all the strings that encode $\R$-structures.
-Since we want to capture $\CC$ over a fix set of structures (e.g.
-graphs, matrices), it is natural to just consider strings that encodes $\R$-structures.
-This definition is fairly standard in descriptive complexity \cite{L04,immerman2012descriptive} and all notions for capturing complexity classes restrict $f \in \CC$ similarly.
+In Definition~\ref{def:cap}, function $f \in \CC$ and formula $\alpha \in \FF$ must coincide in all the strings that encode ordered $\R$-structures.
+Notice that this restriction is natural as we want to capture %Since we want to capture
+$\CC$ over a fix set of structures (e.g. graphs, matrices).
+%, it is natural to just consider strings that encodes $\R$-structures. 
+Moreover, this restriction is fairly standard in descriptive complexity \cite{immerman1999descriptive,L04}, and it has also been used in the previous work on capturing complexity classes of functions \cite{SalujaST95,ComptonG96}.
+%all notions for capturing complexity classes restrict $f \in \CC$ similarly.
 
-Then, what functional complexity classes can be captured with fragments of $\qso$?
-For answering this question, it is reasonable to start with $\shp$ who is one of the most well-known functional complexity class~\cite{arora2009computational}. 
+What counting complexity classes can be captured with fragments of $\qso$?
+For answering this question, it is reasonable to start with $\shp$, a well-known and widely-studied counting complexity class~\cite{arora2009computational}. 
 Since $\shp$ has a strong similarity with $\np$, one could expect a ``Fagin-like'' Theorem~\cite{fagin1974generalized} for this class. 
 Actually, in~\cite{SalujaST95} it was shown that the class $\sfo$ captures $\shp$.
-In our setting, the class $\sfo$ is contained in $\eqso(\fo)$ which also captures $\shp$ as expected. 
+In our setting, the class $\sfo$ is contained in $\eqso(\fo)$, which also captures $\shp$ as expected. 
 \begin{proposition} \label{prop:capture-shP}
@@
 \end{proposition}
-Recall that (see Section~\ref{sec:previous}) that every function class $\# \LL$ is contained in $\eqso(\LL)$.
-Thus, by~\cite{SalujaST95} it directly follows that every $\shp$-function can be defined in $\eqso(\LL)$.
-The other direction is also direct and follows by showing that $\shp$ is closed under first- and second-order sum.
+Recall  that every function class $\# \LL$ is contained in $\eqso(\LL)$ (see Section~\ref{sec:previous}).
+Thus, it directly follows from~\cite{SalujaST95}  that every $\shp$-function can be defined in $\eqso(\LL)$.
+The other direction of Proposition \ref{prop:capture-shP} is straightforward to prove, and it follows by the fact that $\shp$ is closed under first- and second-order sum.
 Moreover, these closure properties also hold if we change the boolean core $\fo$ to any polynomial-time logic like $\lfp$. This gives the following corollary.
@@
 \end{corollary}
-By following the same approach as~\cite{SalujaST95}, Compton et al.~\cite{ComptonG96} show that $\seso$ captures $\spp$.
+By following the same approach as~\cite{SalujaST95}, Compton and Gr\"adel~\cite{ComptonG96} show that $\seso$ captures $\spp$, where $\eso$ is the existential fragment of $\so$.
 As one could expect, if we parametrize $\eqso$ with $\eso$, we can also capture~$\spp$.
@@
 \end{proposition}
-The previous result connecting fragments from $\# \LL$ with $\shp$ and $\spp$ is the further that we can go by using SO-formulas with free variables. 
-So, what happen with other counting complexity classes like $\fp$ and $\fpspace$?
-By trying to search for an $\# \LL$ class for capturing $\fp$ one can easily see that this is pointless.
-Indeed, by using free SO-variables one will easily capture $\shp$-hard functions (see Section~\ref{sec:syntactic}) and by only using free FO-variables one cannot capture exponential functions definable in $\fp$ (e.g.
-$2^{n}$).
-Therefore, it is impossible to capture classes like $\fp$ or $\fpspace$ if we follow previous approaches like~\cite{SalujaST95}.
-
-Instead, if we move from $\eqso$ we have other options for counting like first- and second-order products.
-Actually, if we combine $\qfo$ with $\lfp$ this is exactly what we need to capture $\fp$, the class of function that can be computed in~$\ptime$.
+Can we capture $\fp$ by using $\# \LL$ for some fragment $\LL$ of $\so$?
+A first attempt could be based on the use of a fragment $\LL$ of $\so$ that capture either $\ptime$ or $\nlog$~\cite{G92}.
+Such an approach fails as $\# \LL$ can encode $\shp$-complete problems in both cases; in the first case, one can encode the problem of counting the number of satisfying assignments of a Horn  propositional formula, while in the second case one can encode the problem of counting the number of satisfying assignments of a 2-CNF propositional formula.
+A second attempt could be based then on considering a fragment $\LL$ of $\fo$.
+But even if we consider the existential fragment $\Sigma_1$ of $\fo$ the approach fails, as $\# \Sigma_1$ can encode $\shp$-complete problems like counting the number of satisfying assignments of a 3-DNF propositional formula\cite{SalujaST95}.
+One last attempt could be based on disallowing the use of second-order free variables in $\sfo$.
+But in this case one
+%The previous result connecting fragments from $\# \LL$ with $\shp$ and $\spp$ is the further that we can go by using SO-formulas with free variables. 
+%So, what happen with other counting complexity classes like $\fp$ and $\fpspace$?
+%By trying to search for an $\# \LL$ class for capturing $\fp$ one can easily see that this is pointless.
+%Indeed, 
+%By using free second-order variables one can easily capture $\shp$-hard functions (see Section~\ref{sec:syntactic}) and 
+%By only using free first-order variables one 
+cannot capture exponential functions definable in $\fp$ such as~$2^n$.
+%(e.g. $2^{n}$), so
+Hence, it is not  clear how to capture
+%impossible 
+%to capture 
+$\fp$ 
+%or $\fpspace$ 
+by following the approach proposed in~\cite{SalujaST95}. 
+%if we follow previous approaches like
+On the other hand, if we consider our framework and move out from $\eqso$, we have other options for counting like first- and second-order products.
+In fact, the combination of $\qfo$ with $\lfp$ is exactly what we need to capture $\fp$.
+% the class of function that can be computed in~$\ptime$.
 \begin{theorem} \label{theo:capture-fp}
@@
 \end{theorem}
-To capture $\fp$ one first shows that any formula in $\qfo(\lfp)$ can be evaluated in $\ptime$. 
-Indeed, $\lfp$ is a polynomial time logic~\cite{I86,vardi1982complexity} and the sum and product quantifiers can be computed in $\fp$. 
-For the other direction, one has to use $\lfp$ to simulate the run of the $\fp$-machine and use the quantitative quantifiers to reconstruct the binary number built in the tape. 
-It is important to notice that the alternation between sum and products quantifiers is crucial for capturing $\fp$.
-That is, without these operators is difficult to capture $\fp$.
+To prove this theorem, 
+%capture $\fp$ 
+one first shows that every formula in $\qfo(\lfp)$ can be evaluated in polynomial time. 
+Indeed, $\lfp$ is a polynomial-time logic~\cite{I86,vardi1982complexity}, and the sum and product quantifiers can also be computed in polynomial time. 
+For the other direction, one has to use $\qfo(\lfp)$ to simulate the run of a polynomial time TM $M$ computing a function, in particular using the quantitative quantifiers to reconstruct the natural number returned by $M$ in the output tape.
+%built in the ouptape. 
+It is important to notice that the alternation between sum and products quantifiers is crucial for this reconstructions and, thus, crucial for capturing $\fp$.
+%That is, without these operators is difficult to capture $\fp$.
 
-Another natural function complexity class is $\fpspace$~\cite{Ladner89}, the functional version of $\pspace$. 
-Of course, for capturing this class one needs a logical core powerful enough, like $\pfp$, for simulating the run of an $\pspace$-machine. 
-On the other hand, an $\fpspace$ machine can generate outputs of exponentially many bits. 
-For this reason, the second-order sum is not enough for the quantitative layer of a logic for $\fpspace$, that is, one needs to consider second-order products to produce big numbers. 
-This leads us to the fragment $\qso(\pfp)$ which naturally captures $\fpspace$. 
+At this point it is natural to ask whether one can extend the previous idea to capture $\fpspace$~\cite{Ladner89}, the class of functions computable in polynomial space. 
+%Another natural function complexity class is $\fpspace$~\cite{Ladner89}, the functional version of $\pspace$. 
+Of course, for capturing this class one needs a logical core powerful enough, like $\pfp$, for simulating the run of a polynomial-space TM.
+% $\pspace$-machine. 
+Moreover, 
+%On the other hand, 
+one also needs more powerful quantitative quantifiers as functions like $2^{2^n}$ can be computed in polynomial space,
+%such a TM can generatean $\fpspace$ machine can generate outputs of exponentially many bits. 
+so second-order sum is not enough for the quantitative layer of a logic for $\fpspace$.
+%For this reason, the second-order sum is not enough for the quantitative layer of a logic for $\fpspace$, that is, one needs to consider second-order products to produce big numbers. 
+In fact, by considering second-order product we obtain the fragment $\qso(\pfp)$ that 
+%naturally 
+captures $\fpspace$. 
 \begin{theorem} \label{theo:capture-fpspace}
@@
 \end{theorem}
-The proof of the previous theorem follows the same lines than the logical characterization of $\fp$: one shows that each function in $\qso(\pfp)$ can be computed in $\fpspace$ and, conversely, the output of an $\fpspace$ can be reconstructed by using $\pfp$ and second-order quantitative quantifiers.
+The proof of the previous theorem follows the same line as for the logical characterization of $\fp$: one shows that each function in $\qso(\pfp)$ can be computed in $\fpspace$ and, conversely, the output of a polynomial-space TM can be reconstructed by using $\pfp$ and second-order quantitative quantifiers.
 
-From the proof of the previous theorem a natural question follows: what happen if we use first-order quantitative quantifiers and $\pfp$?
-In~\cite{Ladner89}, Ladner also proposes the class $\nfpspace$ of all functions computed by $\pspace$-machines whose output is polynomially bounded. 
+From the proof of the previous theorem a natural question follows: what happens if we use first-order quantitative quantifiers and $\pfp$?
+In~\cite{Ladner89}, Ladner also introduced the class $\nfpspace$ of all functions computed by polynomial-space TMs 
+%$\pspace$-machines 
+with output length bounded by a polynomial.
+% output is polynomially bounded. 
 Interestingly, if we restrict to FO-quantitative quantifiers we can also capture this class.
@@
 
-The results in this section validate $\qso$ as the logical framework for extending the theory of Descriptive Complexity to function computational complexity.
-In the next sections, we study subfragments of $\eqso$ and, moreover, show how to go beyond $\eqso$ to capture other computational classes.
+The results of this section validate $\qso$ as an appropriate logical framework for extending the theory of Descriptive Complexity to counting complexity classes.
+In the following sections, we provide more arguments for this claim, by considering some fragments of $\eqso$ and, moreover, by showing how to go beyond $\eqso$ to capture other
+%counting complexity 
+classes.

### 2017-01-08 LICS/beyond.tex

@@
 Note that, in contrast to $\lfp$-logic, here we do not need to impose any positive restriction to the formula $\beta(\x,h)$.
-Indeed, since $\beta$ is constructed over the natural numbers based on the monotone operators of sum and product, it is not necessary to add any restriction to the base formula.
+Indeed, since $\beta$ is constructed from monotones operations over the natural numbers, it is not necessary to add any restriction to the base formula.
 In the following result we show how the fixed point operator introduced in this section can be used to capture the fundamental class of functions that can be computed in polynomial time.
@@
 It was shown in \cite{I86,I88} that $\fo$ extended with a transitive closure operator captures $\nlog$. 
-Moreover, it was shown in \cite{I83} that $\so$ extended with a transitive closure operator captures $\pspace$.
-Inspired by this work, we extend the definition of $\qfo$ with an operator for counting the number of paths in a directed graph defined in terms of $\rqfo$.
-Then we show that it can be used to capture $\shl$. Besides, we show that the same idea can be used to extend $\qso$ allowing to capture harder complexity classes.
+%Moreover, it was shown in \cite{I83} that $\so$ extended with a transitive closure operator captures $\pspace$. 
+Inspired by this work, we show that a restricted version of $\rqfo$ can be used to capture $\shl$, the counting version of $\nlog$. 
+Specifically, we use $\rqfo$ to define an operator for counting the number of paths in a directed graph to later show that it can be used to capture $\shl$. 
+%Besides, we show that the same idea can be used to extend $\qso$ allowing to capture harder complexity classes.
 
@@
 where $\psi(\x, \y)$ is an $\fo$-formula over $\R$, $\bar{x} = (x_1, \ldots, x_k)$, $\bar{y} = (y_1, \ldots, y_k)$ are tuples of pairwise distinct first-order variables.
-The semantics of $[\pth \psi(\bar{x},\bar{y})]$ can easily be defined by using $\rqfo(\fo)$ as follows.
-Similar than Example~\ref{ex:count-path}, suppose that the formula $\psi(\bar{x},\bar{y})$ define a (directed) graph over tuples in $A^k$ ($A$ is the domain of $\A$) such that for every $\a, \b \in A^k$ there exists an edge $(\a, \b)$ if $\A \models \psi(\bar a, \bar b)$.
+The semantics of $[\pth \psi(\bar{x},\bar{y})]$ can easily be defined in terms of $\rqfo(\fo)$ as follows.
+Similar than Example~\ref{ex:count-path}, the formula $\psi(\bar{x},\bar{y})$ indirectly defines a (directed) graph over tuples in $A^k$ such that for every $\a, \b \in A^k$ there exists an edge $(\a, \b)$ if $\A \models \psi(\bar a, \bar b)$.
 Define formula $\beta_{\psi(\bar{x},\bar{y})}(\x, \y, \t, g)$ as follows:
@@
 where $\varphi_{\text{\rm first}}$ and $\varphi_{\text{succ}}$ define now the first and successor predicates over the lexicographic order of tuples induced by~$<$.
-Then the semantics of the path operator is defined by the semantics of the following equivalent $\rqfo$ formula:
+Then the semantics of the path operator is defined by the following equivalent $\rqfo$ formula:
 \[

### 2017-01-08 LICS/beyond.tex
Nueva versión 6.

@@
 Specifically, we give the first steps on defining quantitative recursion.
-This goal is not trivial for two reasons: (1) we want to add recursion over functions which, in general, it is a difficult task and (2) it is not clear what is the right notion of ``fixed point'' for recursive functions.
+This goal is not trivial for two reasons: (1) we want to add recursion over functions and (2) it is not clear what could be the right notion of ``fixed point'' .
 Towards this goal, we show first how to extend $\qso$ with function symbols to later use them to define recursion at a quantitative level. 
-Furthermore,  we define some sort of least fixed point that we called \emph{least s-fixed point} and, as a proof of concept, we show that it captures $\fp$.
-Further, we use our new tool to define an operator for counting paths in a graph, a natural generalization of the transitive closure operator~\cite{immerman1999descriptive}, and show that this captures the expressiveness of~$\shl$.
+Furthermore, we define a natural generalization of (boolean) least fixed point for functions that we called \emph{least s-fixed point} and, as a proof of concept, we show that it captures $\fp$.
+Finally,we use our new tool to define an operator for counting paths in a graph, a natural generalization of the transitive closure operator~\cite{immerman1999descriptive}, and show that this captures the expressiveness of~$\shl$.
 
@@
 It worths noting that function symbols represent in $\fqso$ functions from tuples to numbers and cannot be compared to the classical notion of (boolean) function symbol in first-order logics~\cite{enderton2001mathematical}.
-Furthermore, one could see a function symbol as an ``oracle'' that is instantiated by the function assignment and is used to express another function.
-As far as we know, this is the first paper proposing this extension on weighted logics and it would be interesting to study this concept further (i.e.
-outside recursion).
+Furthermore, one could see a function symbol as an ``oracle'' that is instantiated by the function assignment. 
+As far as we know, this is the first paper proposing this extension on weighted logics and it would be interesting to study this concept further (i.e. outside recursion).
 
-We define an operator which extends Least Fixed Point Logic (LFP) \cite{I86,vardi1982complexity} to allow counting. 
-We show how to define this operator in $\qfo(\fo)$ and we discuss how to extend these ideas to $\qso$ towards the end of the section. 
+We define an operator which extends Least Fixed Point Logic (LFP) \cite{I86,vardi1982complexity} over $\fo$ to allow counting. 
+%We show how to define this operator in $\qfo(\fo)$ and we discuss how to extend these ideas to $\qso$ towards the end of the section. 
 Fix a relational signature $\R$.
@@
 Assume now that we need to extend our previous counting function to the case of arbitrary directed graphs.
-To this end, suppose that $\varphi_{\text{\rm first}}(x)$ and $\varphi_{\text{succ}}(x,y)$ are the $\fo$-formulas for defining the first and successor predicates, respectively, of the total-order $<$.
+To this end, suppose that $\varphi_{\text{\rm first}}(x)$ and $\varphi_{\text{succ}}(x,y)$ are the $\fo$-formulas for defining the first and successor predicates, respectively, of $<$.
 Moreover, define formula $\beta(x, y, t, g)$ as follows:
@@
 Note that, in contrast to $\lfp$-logic, here we do not need to impose any positive restriction to the formula $\beta(\x,h)$.
-Indeed, since $\beta$ is constructed from monotones operations over the natural numbers, it is not necessary to add any restriction to the base formula.
-In the following result we show how the fixed point operator introduced in this section can be used to capture the fundamental class of functions that can be computed in polynomial time.
+Indeed, since $\beta$ is constructed from monotones operations (i.e. sum and product) over the natural numbers, the resulting operator $T_{\beta}$ is monotone as well.
+
+Now that a least fixed point operator over functions is defined, the next step is to understand its expressive power.
+In the following result we show how lsfp-operator can be used to capture all functions that can be computed in polynomial time.
 \begin{theorem} \label{rqfo-fo-cap}
@@
 
+Our last goal in this section is to use the new characterization of $\fp$ to explore classes below it.
 It was shown in \cite{I86,I88} that $\fo$ extended with a transitive closure operator captures $\nlog$. 
@@
 Inspired by this work, we show that a restricted version of $\rqfo$ can be used to capture $\shl$, the counting version of $\nlog$. 
-Specifically, we use $\rqfo$ to define an operator for counting the number of paths in a directed graph to later show that it can be used to capture $\shl$. 
+Specifically, we use $\rqfo$ to define an operator for counting the number of paths in a directed graph which is exactly what we need for $\shl$. 
 %Besides, we show that the same idea can be used to extend $\qso$ allowing to capture harder complexity classes.
 
-Given a relation signature $\R$, the set of transitive $\qfo$ formulas ($\tqfo$-formulas) is defined as an extension of $\qfo$ with the operator:
+Given a relation signature $\R$, the set of transitive $\qfo$ formulas ($\tqfo$-formulas) is defined as an extension of $\qfo$ with the operator
 %\begin{multline}
@@
 %\end{multline}
-\[
+$
 [\pth \psi(\bar{x},\bar{y})]
-\]
-where $\psi(\x, \y)$ is an $\fo$-formula over $\R$, $\bar{x} = (x_1, \ldots, x_k)$, $\bar{y} = (y_1, \ldots, y_k)$ are tuples of pairwise distinct first-order variables.
+$ where $\psi(\x, \y)$ is an $\fo$-formula over $\R$, $\bar{x} = (x_1, \ldots, x_k)$, $\bar{y} = (y_1, \ldots, y_k)$ are tuples of pairwise distinct first-order variables.
 The semantics of $[\pth \psi(\bar{x},\bar{y})]$ can easily be defined in terms of $\rqfo(\fo)$ as follows.
-Similar than Example~\ref{ex:count-path}, the formula $\psi(\bar{x},\bar{y})$ indirectly defines a (directed) graph over tuples in $A^k$ such that for every $\a, \b \in A^k$ there exists an edge $(\a, \b)$ if $\A \models \psi(\bar a, \bar b)$.
-Define formula $\beta_{\psi(\bar{x},\bar{y})}(\x, \y, \t, g)$ as follows:
+Given an $\R$-structure $\A$ with domain $A$, define a (directed) graph $\cG_{\psi}(\A) = (N,E)$ such that $N = A^k$ and for every pair $\bar b, \bar c \in N$, it holds that $(\bar b, \bar c) \in E$ if, and only if, $\A \models \psi(\bar b, \bar c)$.
+Similar than for Example~\ref{ex:count-path}, we can count the paths of length at most $|A^k|$ with the formula $\beta_{\psi(\bar{x},\bar{y})}(\x, \y, \t, g)$:
 \begin{multline*}
-(\psi(\bar{x},\bar{y}) + \sa{\z} g(\x,\z,\t)\cdot \psi(\z,\y)) \cdot \varphi_{\text{\rm first}}(\t) \ +\\
-\sa{\t'} \varphi_{\text{succ}}(\t',\t) \cdot \left(\sa{\x'} \sa{\y'} g(\x',\y',\t') \right)
+(\psi(\bar{x},\bar{y}) + \sa{\z} g(\x,\z,\t)\cdot \psi(\z,\y)) \cdot \varphi_{\text{\rm first-lex}}(\t) \ +\\
+\sa{\t'} \varphi_{\text{succ-lex}}(\t',\t) \cdot \left(\sa{\x'} \sa{\y'} g(\x',\y',\t') \right)
 \end{multline*}
-where $\varphi_{\text{\rm first}}$ and $\varphi_{\text{succ}}$ define now the first and successor predicates over the lexicographic order of tuples induced by~$<$.
-Then the semantics of the path operator is defined by the following equivalent $\rqfo$ formula:
+where $\varphi_{\text{\rm first-lex}}$ and $\varphi_{\text{succ-lex}}$ are $\fo$-formulas for defining the first and successor predicates over tuples in $A^k$, following the lexicographic order induced by~$<$.
+Then the semantics of the path operator is defined $\rqfo$ formula:
 \[
@@
 \]
-In other words, $\sem{[\pth \psi(\bar{x}, \bar{y})]}(\A,v,V)$ is the number of paths in the graph induced by $\psi(\bar{x}, \bar{y})$ whose length is at most~$|A|$.
+In other words, $\sem{[\pth \psi(\bar{x}, \bar{y})]}(\A,v,V)$ counts the number of paths in the graph $\cG_{\psi}(\A)$ whose length is at most~$|A^k|$.
 %As for the case of $\qso$, the logic $\tqso(\LL)$ is obtained by restricting $\varphi$ in \eqref{eq-def-tqso} to be a formula in $\LL$.
 Moreover, the logic $\tqfo$ is obtained by disallowing in \eqref{eq-def-tqso} formulas $\sa{X} \alpha$ and $\pa{X} \alpha$, and by only allowing  first-order free-variables in the formula $\psi$ used in $[\pth \psi]$ in \eqref{eq-def-tqso}.
-With this logic, we have the following results:
-
+As it was previously said, the operator for counting paths is exactly what we need to capture $\shl$.
 \begin{theorem} \label{tqfo-shl}
@@
 \end{theorem}
+The last result perfectly illustrates the benefits of the logical framework for the descriptive complexity of counting functions. 
+The distinction of the language between boolean and quantitative allows us to define operators at the later that cannot be defined at the former. 
+Indeed, it is not clear how to capture $\shl$ with $\qfo(\tc)$\footnote{$\tc$ corresponds to $\fo$-logic extended with transitive closure}, or any other extension of $\qfo(\fo)$ at the boolean level.
+In contrast, with $\rqfo$ one can naturally define the path operator at the quantitative level and easily capture $\shl$.
+Therefore, Theorem~\ref{tqfo-shl} is good example of how to go beyond classical logic to have a better understanding of the expressivity of counting.
 
-\cristian{Aca voy}
+\begin{comment}
 
@@
 \end{theorem}
+
+\end{comment}
 

### 2017-01-08 LICS/extentions.tex
sc

@@
-There exist counting complexity classes that do not fit in our framework because either the output of a function in not a natural numbers (e.g.
+%!TEX root = main.tex
+
+There exist complexity classes that do not fit in our framework because either the output of a function in not a natural number (e.g.
 a negative number) or the class is not defined purely in terms of arithmetical operations (e.g.
-min and max).
-To remedy this problem, we show here how to easily extend $\qso$ in order to capture these classes that go beyond sum and product over natural numbers.
+min and max are used).
+To remedy this problem, we show here how $\qso$ can be easily extended  to capture such classes that go beyond sum and product over natural numbers.
 

### 2017-01-08 LICS/beyond.tex

@@
 The distinction of the language between boolean and quantitative allows us to define operators at the later that cannot be defined at the former. 
-Indeed, it is not clear how to capture $\shl$ with $\qfo(\tc)$\footnote{$\tc$ corresponds to $\fo$-logic extended with transitive closure}, or any other extension of $\qfo(\fo)$ at the boolean level.
+Indeed, it is not clear how to capture $\shl$ with $\qfo(\tc)$\footnote{$\tc$ corresponds to $\fo$-logic extended with transitive closure.}, or any other extension of $\qfo(\fo)$ at the boolean level.
 In contrast, with $\rqfo$ one can naturally define the path operator at the quantitative level and easily capture $\shl$.

### 2017-01-09 LICS/saluja.tex
sc

@@
 Inspired by the connection between $\shp$ and $\sfo$, a hierarchy of subclases of $\sfo$ was introduced in~\cite{SalujaST95} 
-by restricting the alternation of quantifiers in Boolean formulas.
+by restricting the alternation of quantifiers in Boolean formulae.
 Specifically, the \emph{$\sfo$-hierarchy} consists of the 
-the classes $\E{i}$ and $\U{i}$ for every $i \geq 0$, where $\E{i}$ (resp., $\U{i}$) is defined as $\sfo$ but restricting the formulas used to be in $\loge{i}$ (resp., $\logu{i}$).
+the classes $\E{i}$ and $\U{i}$ for every $i \geq 0$, where $\E{i}$ (resp., $\U{i}$) is defined as $\sfo$ but restricting the formulae used to be in $\loge{i}$ (resp., $\logu{i}$).
 By definition, we have that $\U{0} = \E{0}$. Moreover, it is shown in~\cite{SalujaST95} that:
@@
 If a formula is in $\LL$-PNF then clearly the formula is in $\LL$-SNF.
-Unfortunately, for some $\LL$ there exist formulas in $\eqso(\LL)$  that cannot be rewritten in $\LL$-PNF.
+Unfortunately, for some $\LL$ there exist formulae in $\eqso(\LL)$  that cannot be rewritten in $\LL$-PNF.
 Therefore, to unveil the relationship between the $\sfo$-hierarchy and the $\eqso(\fo)$-hierarchy, we need to understand the boundary between PNF and SNF. We do this in the following theorem. 
@@
 \end{proposition}
-The relationship between both hierarchies is summarized in Figure \ref{fig-sfo-eqso}.
-One can see there that the difference between our hierarchy and the one proposed in~\cite{SalujaST95} only differ in~$\Sigma_0$ and~$\Sigma_1$. 
-Actually, we show next that this difference is very crucial for finding classes of functions with easy decision versions and good closure properties.
+The relationship between the two hierarchies is summarized in Figure \ref{fig-sfo-eqso}.
+Our hierarchy and the one proposed in~\cite{SalujaST95} only differ in~$\Sigma_0$ and~$\Sigma_1$. 
+Interestingly, we show next that this difference is crucial for finding classes of functions with easy decision versions and good closure properties.

### 2017-01-09 LICS/syntactic-clo.tex
sc

@@
 
-We use the \emph{$\eqso(\fo)$-hierarchy} to search for syntactic subclasses with good algorithmic and closure properties.  
-Before starting this search, we introduce a more strict notion of counting problems with easy decision versions.
+We use the \emph{$\eqso(\fo)$-hierarchy} to define syntactic classes of functions with good algorithmic and closure properties.  
+But before 
+%starting our search
+doing this, we introduce a more strict notion of counting problem with easy decision version.
 Recall that a function $f : \Sigma^* \to \mathbb{N}$ has an easy decision counterpart if $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$ is a language in $\ptime$.
-As the goal of this section is to define a syntactic class of functions in $\shp$ with easy decision versions and good closure properties, we do not directly consider the semantic condition $L_f  \in \ptime$, but instead we consider a more restricted class of functions that has a simple syntactic definition.
-More precisely, a function $f : \Sigma^* \to \mathbb{N}$ is said to be in the complexity class $\totp$~\cite{PagourtzisZ06} if there exists a nondeterministic polynomial-time Turing Machine $M$ such that $f(x) = \tmt_M(x) - 1$ for every $x \in \Sigma^*$, where $\tmt_M(x)$ is the total number of runs of $M$ with input $x$.
+As the goal of this section is to define a syntactic class of functions in $\shp$ with easy decision versions and good closure properties, we do not directly consider the semantic condition $L_f  \in \ptime$, but instead we consider a more restricted
+%class of functions that has a simple 
+syntactic 
+%definition. 
+condition.
+More precisely, a function $f : \Sigma^* \to \mathbb{N}$ is said to be in the complexity class $\totp$~\cite{PagourtzisZ06} if there exists a  polynomial-time NTM $M$ such that $f(x) = \tmt_M(x) - 1$ for every $x \in \Sigma^*$, where $\tmt_M(x)$ is the total number of runs of $M$ with input $x$.
 Notice that one is subtracted from $\tmt_M(x)$ to allow for $f(x) = 0$.
@@
 It was proved in \cite{SalujaST95} that every function in $\E{1}$ admits an FPRAS. Besides, it can be proved that $\E{1} \subseteq \totp$. 
-However, we have to discard this class since it is not closed under sum and then is not robust under basic closure properties. 
+However, 
+%we have to discard 
+this class 
+%since it i
+is not closed under sum, and then it is not robust under basic closure properties. 
 \begin{proposition}\label{prop-e1-nc}
@@
 \end{proposition}
-To overcome this limitation, one can consider the class $\QE{1}$, which is closed by under sum by definition.
-In fact, the following proposition shows that the same good properties as for $\E{1}$ hold for $\QE{1}$ plus being closed under sum and multiplication.
+To overcome this limitation, one can consider the class $\QE{1}$, which is closed under sum by definition.
+In fact, the following proposition shows that the same good properties as for $\E{1}$ hold for $\QE{1}$, together with the fact that
+%plus being 
+it is closed under sum and multiplication.
 \begin{proposition} \label{prop:qe0-fp-qe1-totp-fptras}
@@
 Thus, we need to find an extension of $\QE{1}$ that keeps all the previous properties and is closed under subtraction by one.
-It is important to notice that $\shp$ is believed not to be closed under subtraction by one by some complexity-theoretical assumption\footnote{A decision problem $L$ is in the randomized complexity class $\cspp$ if there exists a nondeterministic polynomial-time Turing Machine $M$ such that for every $x \in L$ it holds that $\tma_M(x) - \tmr_M(x) = 2$, and for every $x \not\in L$ it holds that $\tma_M(x) = \tmr_M(x)$ \cite{OH93,FFK94}.
-It is believed that $\np \not\subseteq \cspp$, as there is no evidence that a Turing Machine with the property just described exists for the propositional satisfiability problem.
-However, if $\shp$ is closed under subtraction by one, then it holds that $\np \subseteq \cspp$ \cite{OH93}.}.
-So, the following proposition rules out any logic that extends $\Pi_1$ for a possible extension of~$\QE{1}$.
+It is important to notice that $\shp$ is believed not to be closed under subtraction by one by some complexity-theoretical assumption.\footnote{A decision problem $L$ is in the randomized complexity class $\cspp$ if there exists a polynomial-time NTM $M$ such that for every $x \in L$ it holds that $\tma_M(x) - \tmr_M(x) = 2$, and for every $x \not\in L$ it holds that $\tma_M(x) = \tmr_M(x)$ \cite{OH93,FFK94}.
+It is believed that $\np \not\subseteq \cspp$, as there is no evidence that a TM with the property just described exists for the propositional satisfiability problem.
+However, if $\shp$ is closed under subtraction by one, then it holds that $\np \subseteq \cspp$ \cite{OH93}.} So, the following proposition rules out any logic that extends $\Pi_1$ for a possible extension of~$\QE{1}$ with the desired closure property.
 \begin{proposition} \label{pi-minusone}
@@
 Therefore, the desired extension has to be achieved by allowing some local extensions to $\Sigma_1$.
-More precisely, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulas over a relational signature $\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mentioned any second-order variable).
-With this extension we find the promised class.
+More precisely, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulae over a relational signature $\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mentioned any second-order variable).
+With this extension we obtain a class with the desired properties.
+%find the promised class.
 \begin{theorem}\label{sigmafo-minusone}

### 2017-01-09 LICS/appendix-counting.tex
Verificación y arreglos en Appendice de Sección 4.

@@
 
-We will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
+We construct recursively a $\shp$-machine $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
 This machine, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
@@
 %If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
+If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
 %If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$. 
@@
 
-Similar than the previous proof, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
+Similar than the previous proof, we construct recursively a $\spp$ machine $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
 This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ distinct accepting outputs for each $(\A,v,V) \in \ostr[\R]^*$.
@@
 If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.  
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$. 
+If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$. 
 If $\alpha = \sa{X} \beta$, it chooses $B\in A^{\arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$. 
@@
 Then each sub-formula $\varphi\in\lfp$ in $\alpha$ is evaluated in polynomial time and replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size which is evaluated and lastly given as output.
+The resulting formula is an arithmetic expression of polynomial size (recall that $\alpha$ is fixed) which is evaluated and lastly given as output.
 Note that $f\in\fp$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
@@
 For the second condition, let $f\in \fp$ defined over some signature $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
+Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$ (i.e. $n^\ell$ is an upper bound for the output size), where $\A$ has a domain of size $n$.
 Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
-Consider a procedure that receives $\enc(\A)$ and an assignation $\bar{a}$ to $\bar{x}$.
+Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
 Let $m$ be the position of $\bar{a}$ in the lexicographic order of the tuples in $A^{\ell}$.
@@
 Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
+In other words, $\Phi(\bar{x})$ simulates the behavior of the $\fp$-machine and the formula $\alpha$ reconstruct the binary output.
 Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
@@
 
-To show how to evaluate a $\qso(\pfp)$-formula, we will construct a recursive non-deterministic algorithm $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
+To show how to evaluate a $\qso(\pfp)$-formula, we construct recursively a $\shpspace$-machine $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
 This machine runs in non-deterministic polynomial space and, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
@@
 If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$.
+If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
+If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$, continuing to the next $A$-value if the run of $M_{\beta}$ accepts.
 If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-If $\alpha = \pa{X}\beta$, it simulates $M_{\beta}$ on input $(\A,v,V[B/X])$ consecutively for each $B\in A^{arity(X)}$.
+If $\alpha = \pa{X}\beta$, it simulates $M_{\beta}$ on input $(\A,v,V[B/X])$ consecutively for each $B\in A^{arity(X)}$, again continuing to the next $A$-value if the run of $M_{\beta}$ accepts.
 This covers all possible cases for $\alpha$, and each of these steps can be computed in polynomial space.
@@
 For the second condition, let $f\in \fpspace$ defined over some $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq 2^{n^\ell}$, where $\A$ has a domain of size $n$.
+Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq 2^{n^\ell}$ (i.e.
+$2^{n^\ell}$ is an upper bound for the output size), where $\A$ has a domain of size $n$.
 Let $X$ be a second-order variable of arity $\ell$.
@@
 $$
-Consider a polynomial space algorithm over $\R\cup\{X\}$ that receives a $\R\cup\{X\}$-structure $\A'$.
-Let $X^{\A'}$ be the $p$-th subset in the given order (we consider 0-indexation).
-The algorithm computes $f(\enc(\A))$ until $(2^{n^{\ell}}-p)$ characters have been written, but each character of the output is written over the last one.
-It accepts if and only if the last character written is $1$.
-Since this algorithm works in polynomial space, it can be described in $\pfp$ \cite{AbiteboulV89} by a formula $\Phi(X)$.
-
-Note that the formula also represents the following.
-For each $(\A,v,V)\in\ostr[\R]^*$, let $V(X)$ be the $p$-th subset with respect to this order and let $w$ be a string of size $2^{n^{\ell}}$ which represents the value of $f(\enc(\A))$ in binary.
-Then $(\A,v,V)\models\Phi(X)$ if and only if the $p$-th bit in $w$ is 1, from least to most significant.
-We define
+Namely, we use relations to encode numbers with at most $2^{n^\ell}$-bits where the empty relation represents $0$ and the total-relation represents $2^{2^{n^\ell}}-1$.
+Furthermore, each relation $X$ indexes a position in the binary output of $f(\enc(\A))$ as follows.
+Consider a polynomial space algorithm over the $\R$ that receives as input an $\R$-structure $\A$ and a number $p$ encoded by a relation $X$.
+Then the algorithm accepts if, and only if, the $p$-th bit of $f(\enc(\A))$ is $1$.
+Since this procedure works in polynomial space, it can be described in $\pfp$ \cite{AbiteboulV89} by a formula $\Phi(X)$ where the free variable $X$ encodes the number $p$.
+Then, similar than the previous proof we define:
 $$
@@
 where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
-Note that for each $B\subseteq A^{\ell}$ such that $B$ is the $m$-th element in the given order, $\sem{\varphi(B)}(\A) = 2^m$.
+Note that for each $B\subseteq A^{\ell}$ such that $B$ is the $m$-th element in the order $\varphi_{<}(X,Y)$, it holds that $\sem{\varphi(B)}(\A) = 2^m$.
 Therefore, $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
@@
 Then each sub-formula $\varphi\in\pfp$ in $\alpha$ is evaluated in polynomial space and replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size which is evaluated and lastly given as output.
+The resulting formula is an arithmetic expression of polynomial size  (recall that $\alpha$ is fixed) which is evaluated and lastly given as output.
 Note that $f\in\nfpspace$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
@@
 Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
-Consider a procedure that receives $\enc(\A)$ and an assignation $\bar{a}$ to $\bar{x}$.
+Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
 Let $m$ be the position of $\bar{a}$ in the lexicographic order of the tuples in $A^{\ell}$.

### 2017-01-09 LICS/extentions.tex

@@
 
-It is well-known that $\shp$ is not closed under subtraction, not even closed under subtraction by one~\cite{OH93}.
-To overcome this issue, people introduced the class $\gp$~\cite{fenner1994gap} defined as the class of functions $f$ such that for some $\np$ machine, $f(x)$ is the number of accepting runs minus the number of rejecting runs.
-That is, $\gp$ is the closure of $\shp$ functions under subtraction and it can compute negative values by definition.
-
-Given that our logical framework was built on top of natural numbers, we need to extend $\qso$ in order to describe functions in $\gp$. 
+It is well-known that, under some reasonable complexity-theoretical assumptions, $\shp$ is not closed under subtraction, not even under subtraction by one~\cite{OH93}.
+To overcome this limitation, $\gp$ was introduced in~\cite{fenner1994gap} as the class of functions $f$ for which there exists a polynomial-time NTM $M$ such that $f(x) = \tma_M(x) - \tmr_M(x)$, where  $\tmr_M(x)$ is the number of rejecting runs of $M$ with input $x$.
+%such that for some $\np$ machine, $f(x)$ is the number of accepting runs minus the number of rejecting runs.
+That is, $\gp$ is the closure of $\shp$ functions under subtraction, and its functions can obviously take negative values.
+%it can compute negative values by definition. 
+Given that our logical framework was built on top of the natural numbers, we need to extend $\qso$ in order to %describe functions
+capture $\gp$. 
 The most elegant way to do this is by allowing constants coming from $\bbZ$ instead of just $\bbN$. 
-Formally, we define the logic $\qsoz$ whose syntax is the same as in (\ref{syntax}) and whose semantics is the same as in Table~\ref{tab-semantics} except that the atomic formula $s$ (i.e.
-a constant) come from $\bbZ$.
+Formally, we define the logic $\qsoz$ whose syntax is the same as in \eqref{syntax} and whose semantics is the same as in Table~\ref{tab-semantics} except that the atomic formula $s$ (i.e.
+a constant) comes from $\bbZ$.
 Similar than for $\qso$, we define the fragment $\eqsoz$ as the extension of $\eqso$ with constants in $\bbZ$.
@@
 Adding negative constants is a mild extension to allow subtraction in the logic. 
-In fact, this is exactly what we need to capture classes like $\gp$.
+Interestingly, this is exactly what we need to capture  $\gp$.
 \begin{proposition} \label{prop:capture-gapp}
@@
 \end{proposition}
-The previous result is straightforward to prove from the characterization of $\shp$ and the definition of $\gp$. 
-However, we believe that this is an interesting result for showing how robust is $\qso$ in order to capture diverse complexity classes.
+%The previous result is straightforward to prove from the characterization of $\shp$ and the definition of $\gp$. 
+%However, 
+%We 
+%believe 
+%think
+This is an interesting result that shows how robust is $\qso$ when capturing different complexity classes.
 
-An interesting class of functions comes from considering the optimization version of a decision problem.
-For example, one can define MAX-SAT as the problem of determining the maximum number of clauses, of a given CNF formula, that can be made true by an assignment.
-Here, MAX-SAT is defined in terms of a maximization problem which in his essence differs from functions in $\shp$.
-To formalize this set of optimization problems, Krentel defined $\optp$~\cite{krentel1988complexity} as the class of functions computable by taking the maximum or minimum of the output values over all runs of an $\np$ machine with output (i.e.
+A different class of functions comes from considering the optimization version of a decision problem.
+For example, one can define MAX-SAT as the problem of determining the maximum number of clauses, of a given CNF propositional formula, that can be made true by an assignment.
+Here, MAX-SAT is defined in terms of a maximization problem which in its essence differs from the functions in $\shp$.
+To formalize this set of optimization problems, Krentel defined $\optp$~\cite{krentel1988complexity} as the class of functions computable by taking the maximum or minimum of the output values over all runs of a polynomial-time NTM machine with output tape (i.e.
 each run produces a binary string which is interpreted as a number).
-For instance, MAX-SAT is in $\optp$ as several others optimization version of $\np$-problems.
-Given that in~\cite{krentel1988complexity} the author do not make the distinction between $\max$ or $\min$, in~\cite{vollmer1995complexity} the authors define the classes $\maxp$ and $\minp$ as the max and min version of the problems in $\optp$ (i.e.
+For instance, MAX-SAT is in $\optp$ as several others optimization versions of $\np$-problems.
+Given that in~\cite{krentel1988complexity} the author does not make the distinction between $\max$ or $\min$, in~\cite{vollmer1995complexity} the authors define the classes $\maxp$ and $\minp$ as the max and min version of the problems in $\optp$ (i.e.
 $\optp = \maxp \cup \minp$).
@@
 In order to capture classes of optimization functions, we can extend $\qso$ with $\max$ and $\min$ quantifiers (called $\optqso$) as follows. 
-Given a relational signature $\R$, the set of $\optqso$-formulas over $\R$ is given by extending the syntax in (\ref{syntax}) with the following operators:
+Given a relational signature $\R$, the set of $\optqso$-formulae over $\R$ is given by extending the syntax in (\ref{syntax}) with the following operators:
 \begin{multline*}
@@
 The semantics of the $\qso$-operators in $\optqso$ are defined as usual.
-Furthermore, the semantics of the max and min quantifiers are defined as the natural extension of the sum quantifiers in $\qso$ (see Table~\ref{tab-semantics}) by maximizing or minimizing (respectively) instead of summing.
+Furthermore, the semantics of the max and min quantifiers are defined as the natural extension of the sum quantifiers in $\qso$ (see Table~\ref{tab-semantics}) by maximizing or minimizing, respectively, instead of computing a sum.
 \begin{example}\label{ex:optqso}
-	Recall again the setting of Example~\ref{ex:cliques} and suppose now that we want to compute the size of the bigger clique in a graph. Then this can be define in $\optqso$ with the following formula:
+	Recall again the setting of Example~\ref{ex:cliques} and suppose now that we want to compute the size of the largest clique in a graph. This can be done in $\optqso$ with the following formula:
 	\[
@@
 	\]
+	Notice that formula $\sa{z} X(z)$ is used to compute the number of nodes in a clique $X$. 
 \end{example}
 Similar than for $\maxp$ and $\minp$, we have to distinguished between the $\max$ and $\min$ fragments of $\optqso$.
-For this, we define the fragment $\maxqso$ of all $\optqso$ formulas constructed from $\qfo$ operators and $\max$ operators (of first- and second-order).
+For this, we define the fragment $\maxqso$ of all $\optqso$ formulae constructed from $\qfo$ operators and $\max$-formulae $\max\{\alpha,\alpha\}$, $\maxa{x} \alpha$ and  $\maxa{X} \alpha$.
+%operators of first- and second-order. 
 The class $\minqso$ is defined analogously changing $\max$ with $\min$.
-Note that in $\maxqso$ and $\minqso$ the second-order sum and product quantifiers are not allowed.
-For example, the formulas $\alpha_1$ and $\alpha_6$ are in $\maxqso$ but $\alpha_2$ and $\alpha_3$ are not.
-As one could expect, $\maxqso$ and $\minqso$ are the logics that we need to capture $\optp$.
+Notice that in $\maxqso$ and $\minqso$, second-order sum and product are not allowed.
+For instance, formula $\alpha_6$ in Example \ref{ex:optqso} is in $\maxqso$.
+%the formulae $\alpha_1$ and $\alpha_6$ are in $\maxqso$ but $\alpha_2$ and $\alpha_3$ are not. 
+As one could expect, $\maxqso$ and $\minqso$ are the needed logics to capture $\maxp$ and $\minp$.
 \begin{theorem} \label{theo:capture-optp}
@@
 \end{theorem}
-It is important to mention that a similar result was proved in~\cite{kolaitis1994logical} for the class $\maxpb$ and $\minpb$ of problems in $\maxp$ and $\minp$ whose output value is polynomially bounded.
-Interestingly, the result is stronger here since our logic has the freedom to use sum and product quantifiers instead of using a max-and-count problem over boolean formulas. 
-Finally, it is straightforward to prove that our framework can also capture $\maxpb$ and $\minpb$ by disallowing the product $\Pi x$ in $\maxqso(\fo)$ and $\minqso(\fo)$, respectively.
+It is important to mention that a similar result was proved in~\cite{kolaitis1994logical} for the class $\maxpb$ (resp., $\minpb$) of problems in $\maxp$ (resp., $\minp$) whose output value is polynomially bounded.
+Interestingly, Theorem \ref{theo:capture-optp} is stronger since our logic has the freedom to use sum and product quantifiers, instead of using a max-and-count problem over Boolean formulae. 
+Finally, it is easy to prove that our framework can also capture $\maxpb$ and $\minpb$ by disallowing the product $\Pi x$ in $\maxqso(\fo)$ and $\minqso(\fo)$, respectively.

### 2017-01-09 LICS/syntactic-horn.tex

@@
 \end{example}
-The previous example suggests that in order to define $\chsat$, we can use Horn formulas defined as follows. 
+The previous example suggests that in order to define $\chsat$, we can use Horn formulae defined as follows. 
 A positive literal is a formula of the form $X(\x)$, where $X$ is a second-order variable and $\x$ is a tuple of first-order variables, and a negative literal is a formula of the form $\exists \v \, \neg X(\u,\v)$, where $\u$ and $\v$ are tuples of first-order variables.
@@
 A clause is said to be Horn if it contains at most one positive literal, and a formula is said to be Horn if it is a conjunction of Horn clauses over a relational signature $\R$.
-With this terminology, we define $\uhorn$ as the set of formulas $\psi$ such that $\psi$ is a conjunction of Horn clauses over a relational signature $\R$.
+With this terminology, we define $\uhorn$ as the set of formulae $\psi$ such that $\psi$ is a conjunction of Horn clauses over a relational signature $\R$.
 
@@
 We conclude this section by showing that $\eqso(\ehorn)$ is the class we were looking for, as not only every function in $\eqso(\ehorn)$ has an easy decision counterpart, but also $\eqso(\ehorn)$ admits a natural complete problem under parsimonious reductions.
-More precisely, define $\dhsat$ as:
-\begin{multline*}
-\{\Phi \mid \Phi \text{ is a disjunction of Horn formulas and satisfiable}\},
-\end{multline*}
+More precisely, define
+%$\dhsat$ as:
 %\begin{multline*}
-%\{\Phi \mid \Phi \text{ is a disjunction of}\\  \text{Horn formulas and $\Phi$ is satisfiable}\},
+%\{\Phi \mid \Phi \text{ is a disjunction of Horn formulae and satisfiable}\},
 %\end{multline*}
-and define $\shdhsat$ as the problem of counting the satisfying assignments of a formula $\Phi$ that is a disjunction of Horn formulas. Then we have that:
+%%\begin{multline*}
+%%\{\Phi \mid \Phi \text{ is a disjunction of}\\  \text{Horn formulae and $\Phi$ is satisfiable}\},
+%%\end{multline*}
+%and define 
+$\shdhsat$ as the problem of counting the satisfying assignments of a formula $\Phi$ that is a disjunction of Horn formulae. Then we have that:
 

### 2017-01-09 LICS/syntactic.tex

@@
 Formally, the decision problem associated to a function $f : \Sigma^* \to \mathbb{N}$ is defined as $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$, and $f$ is said to have an easy decision version if $L_f \in \ptime$.
-Many prominent examples satisfy this property like computing the number of: perfect matchings of a bipartite graph ($\cpm$)~\cite{Valiant79}, satisfying assignments of a DNF propositional formula ($\cdnf$)~\cite{DHK05}, satisfying assignments of a Horn  propositional formula ($\chsat$)~\cite{V79b}, among others.
+Many prominent examples satisfy this property, like computing the number of: perfect matchings of a bipartite graph ($\cpm$)~\cite{Valiant79}, satisfying assignments of a DNF propositional formula ($\cdnf$)~\cite{DHK05}, satisfying assignments of a Horn  propositional formula ($\chsat$)~\cite{V79b}, among others.
 
@@
 %
-%From this point on, for each fragment $\FF$ of $\qso$, we will also use $\FF$ to refer to the class of functions defined by the formulas in $\FF$.
+%From this point on, for each fragment $\FF$ of $\qso$, we will also use $\FF$ to refer to the class of functions defined by the formulae in $\FF$.
 

### 2017-01-09 LICS/appendix-syntactic.tex
Cambios en apendice.

@@
 If $\alpha = (\alpha_1 + \alpha_2)$, then we define $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$.
-If $\alpha = \sa{x}\beta$, then let $\tau(\beta) = \sum_{i = 1}^{n}\beta_i$ where each $\beta_i$ is in $\LL$-PNF.
-We have that each $\beta_i$ also is in $\LL$-PNF, and then we define $\tau(\alpha) = \sum_{i = 1}^{m}\sa{x}\beta_i$ which satisfies the condition.
+If $\alpha = \sa{x}\beta$, then let $\tau(\beta) = \sum_{i = 1}^{k}\beta_i$ for some $k$ where each $\beta_i$ is in $\LL$-PNF.
+By grouping together the terms in the sum, we define $\tau(\alpha) = \sum_{i = 1}^{m}\sa{x}\beta_i$ which satisfies the condition and is equivalent to $\alpha$.
 If $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case.
@@
 \vspace{1em}
-First, we prove that the formula $\alpha_{0} = \sa{X(\cdot)}1 + 1$ is not equivalent to any formula in $\loge{0}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ in $\loge{0}$-PNF that is equivalent to the $\eqso(\loge{0})$ formula $\sa{X(\cdot)}1 + 1$.
-Note that this defines the function $2^{n}+1$, where $n$ is the size of the input structure.
-Then, note that if $\varphi(\bar{X},\bar{x})$ is not satisfiable, then the function is null.
-Note also that if $\length(\bar{X}) = \length(\bar{x}) = 0$, then the function defined by $\alpha$ is never greater than 1.
-Now note that $\sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ is equivalent to a function in $\E{0}$.
-The authors in \cite{SalujaST95} proved that if $\length(\bar{X}) > 0$, then the function defined by the formula, for big enough structures, is always even.
-And since $\sa{\bar{x}}\varphi(\bar{x})$ defines a polynomially bounded function, we follow to a contradiction.
+First, we prove that the formula $\alpha_{0} = \left( \sa{X} 1 \right) + 1$ with $\arity(X) = 1$ (i.e.
+the function $2^{n}+1$, where $n$ is the size of the input structure) is not equivalent to any formula in $\loge{0}$-PNF.
+Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ in $\loge{0}$-PNF that is equivalent to the $\eqso(\loge{0})$ formula $\alpha_0$.
+In \cite{SalujaST95}, it was proved that if $\length(\bar{X}) > 0$, then the function defined by $\alpha$, for big enough structures, is always even which is not possible in our case.
+On the other hand, if $\alpha$ is of the form $\sa{\bar{x}}\varphi(\bar{x})$, then $\alpha$ defines a polynomially bounded function which also leads to a contradiction.
 
 \vspace{1em}
-Second, we prove that the formula $\alpha_{1} = 2$ is not equivalent to any formula in $\loge{1}$-PNF.
+Second, we prove that the formula $\alpha_{1} = 2$ (i.e.
+the constant $2$) is not equivalent to any formula in $\loge{1}$-PNF.
 Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to the $\eqso(\loge{1})$ formula $2$.
-Suppose that $\varphi$ is quantifier-free.
-Note first that if $\length(\bar{X}) = \length(\bar{x}) = 0$, then the function defined by $\alpha$ is never greater than 1.
-Consider an ordered structure $\A$.
-Since $\sem{\alpha}(\A) = 2$, there exist at least two assignations $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
+First, if $\length(\bar{X}) = \length(\bar{x}) = 0$, then the function defined by $\alpha$ is never greater than 1. 
+Therefore, suppose that $\length(\bar{X}) > 0$ or $\length(\bar{x}) > 0$, and consider any ordered structure $\A$. 
+Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
 Now consider the ordered structure $\A'$ that is obtained by duplicating $\A$.
 This is, each half of $\A'$ is isomorphic to $\A$.
-Note that $\A'\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$ for $i = 1,2$ and there exists a third assignation $(\bar{B}_1',\bar{b}_1',\bar{a}_1')$ that is identical to $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$ but in the other half to the structure such that $\A'\models\varphi(\bar{B}_1',\bar{b}_1',\bar{a}_1')$.
-We have that $\sem{\alpha}(\A) \geq 3$, and we follow to a contradiction.
+Note that $\A'\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$ for $i = 1,2$ and there exists a third assignments $(\bar{B}_1',\bar{b}_1',\bar{a}_1')$ that is isomorphic to $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$ but in the other half to the structure such that $\A'\models\varphi(\bar{B}_1',\bar{b}_1',\bar{a}_1')$.
+We have that $\sem{\alpha}(\A) \geq 3$ which is a contradiction.
 
@@
 Let $\alpha = \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-SNF.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$.
-We replace each $\alpha_i$ for the equivalent formula $\sa{\bar{X}}\sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge\forall z\,Y(z) \wedge \forall z(y \leq z))$ to obtain that $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.
+Without lost of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.
+If not, we replace each $\alpha_i$ for the equivalent formula $\sa{\bar{X}} \sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge\forall z\,Y(z) \wedge \forall z(y \leq z))$.
 
@@
 If $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then the formula is already in $\LL$-PNF so we define $\tau(\alpha) = \alpha$.
-If $\alpha = \alpha_1 + \alpha_2$, then we assume that $\tau(\alpha_1) = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ and $\tau(\alpha_2) = \sa{\bar{X}}\sa{\bar{x}}\psi(\bar{X},\bar{x})$.
+If $\alpha = \alpha_1 + \alpha_2$, then we assume that $\tau(\alpha_1) = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ and $\tau(\alpha_2) = \sa{\bar{Y}}\sa{\bar{y}}\psi(\bar{Y},\bar{y})$.
 The construction that we will provide for this function works by identifying a ``first'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ and a ``last'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$.
@@
 \end{align*}
+Similarly, we define the formulas $\gamma_{\text{first}}(\bar{Y},\bar{y})$ and $\gamma_{\text{last}}(\bar{Y},\bar{y})$ (for the sake of simplicity we reuse the names $\gamma_{\text{first}}$ and $\gamma_{\text{last}}$).
 In other words, the ``first'' assignment is the one where every second-order predicate is empty and the first-order assignment is the lexicographically smallest, and the ``last'' assignment is the one where every second-order predicate is full and the first-order assignment is the lexicographically greatest.
 We also need to identify assignments that are not first and are not last.
-We do this by negating the two formulas above, and we use a concatenation of the quantified first-order variables:
+We do this by negating the two formulas above and grouping together the first-order variables:
 \begin{align*}
@@
 \begin{align}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{not first}}(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{Y},\bar{y}))\vee\\
-&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{not last}}(\bar{Y},\bar{y}))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{last}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))].
+\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{not first}}(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{Y},\bar{y}))\vee \label{eq:partition1} \\
+&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))\vee \label{eq:partition2}\\
+&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{not last}}(\bar{Y},\bar{y}))\vee \label{eq:partition3}\\
+&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{last}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))]. \label{eq:partition4}
 \end{align}
-To show that the formula is indeed equivalent to $\alpha$, note three facts.
-First, the formulas in lines (2) and (3) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$.
-Second, the formulas in lines (4) and (5) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
-And third, the four lines define pairwise disjoint assignments for $(\bar{X},\bar{x}),(\bar{Y},\bar{y})$.
-
-However, the formula is not yet in the correct form. By expanding it we obtain:
+To show that the formula is indeed equivalent to $\alpha$, note that the formulas in lines (\ref{eq:partition1}) and (\ref{eq:partition2}) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$, and the formulas in lines (\ref{eq:partition3}) and (\ref{eq:partition4}) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
+Altogether, the four lines define pairwise disjoint assignments for $(\bar{X},\bar{x}),(\bar{Y},\bar{y})$.
+With this, it is straightforward to show that the above formula is equivalent to $\alpha$.
+However, the formula is not yet in the correct form since it has existential quantifiers in the subformulas $\gamma_{\text{not first}}$ and $\gamma_{\text{not last}}$.
+To solve this, first let take a close look to the complete formula:
 \begin{align*}
@@
 A similar construction was used in \cite{SalujaST95}.
-We obtain:
+Then the final formula equivalent to $\alpha$ is the following:
 \begin{align*}
-\sa{\bar{X}}\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i))\wedge\forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee\\
+\sa{\bar{X}}\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
+& \hspace{3cm}  \bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
+&
+(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))\vee
+\\
+&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
+& \hspace{3cm} (\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee \\
 &(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))].
 \end{align*}
-This covers all possible cases for $\alpha$.
-
 Finally, consider a $\eqso(\LL)$ formula $\alpha$ in $\LL$-SNF.
-If $\alpha = \sum_{i = 1}^n\alpha_i$, then consider $\alpha = (\alpha_1 + (\sum_{i = 2}^n\alpha_i))$.
-Then we use $\tau(\alpha)$ as the rewrite of $\alpha$, which satisfies the condition in the hypothesis.
+If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we can consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the condition in the hypothesis.
 

### 2017-01-09 LICS/beyond.tex

--- 2017-01-08 LICS/beyond.tex
@@
 
-In the previous sections, we use weighted logics to give a framework for descriptive complexity of functions.
-Here, we go beyond weighted logics to add unexplored quantifiers at the quantitative level.
-Specifically, we give the first steps on defining quantitative recursion.
-This goal is not trivial for two reasons: (1) we want to add recursion over functions and (2) it is not clear what could be the right notion of ``fixed point'' .
-Towards this goal, we show first how to extend $\qso$ with function symbols to later use them to define recursion at a quantitative level. 
-Furthermore, we define a natural generalization of (boolean) least fixed point for functions that we called \emph{least s-fixed point} and, as a proof of concept, we show that it captures $\fp$.
-Finally,we use our new tool to define an operator for counting paths in a graph, a natural generalization of the transitive closure operator~\cite{immerman1999descriptive}, and show that this captures the expressiveness of~$\shl$.
+In the previous sections, we use weighted logics to give a framework for descriptive complexity of counting complexity classes. Here, we go beyond weighted logics 
+%to add unexplored quantifiers at the quantitative level. Specifically, we 
+and give the first steps on defining recursion at the quantitative level.
+%quantitative recursion. 
+This goal is not trivial not only because we want to add recursion over functions, but also because 
+it is not clear what could be the right notion of ``fixed point''. 
+%for two reasons: (1) we want to add recursion over functions and (2) it is not clear what could be the right notion of ``fixed point'' . 
+To this end, we show first how to extend $\qso$ with function symbols that are later
+used 
+%to later use them to define recursion at a quantitative level. 
+%Furthermore, we 
+to define a natural generalization for functions of 
+%(boolean) 
+the notion of least fixed point of LFP, which is denoted as \emph{least s-fixed point}.
+% and, 
+ As a proof of concept, we show that how this notion can be used to captures $\fp$.
+Moreover, we use this concept
+%our new tool 
+to define an operator for counting paths in a graph, a natural generalization of the transitive closure operator~\cite{immerman1999descriptive}, and show that this gives rise to a logic that captures~$\shl$.
+%the expressiveness of
 
-For defining recursion, we first need to define an extension of $\qso$ with function symbols.
-Assume that $\fs$ is an infinite set of function symbols, where each $h \in \fs$ has an associated arity, which is denoted by $\arity(h)$.
-Then the set of $\fqso$ formulas over a relational signature $\R$ is defined by the following grammar:
+%For defining recursion, 
+We start
+%first need to define 
+by defining an extension of $\qso$ with function symbols. Assume that $\fs$ is an infinite set of function symbols, where each $h \in \fs$ has an associated arity
+%, which is 
+denoted by $\arity(h)$. Then the set of $\fqso$ formulas over a relational signature $\R$ is defined by the following grammar:
 \begin{multline}
@@
 
-It worths noting that function symbols represent in $\fqso$ functions from tuples to numbers and cannot be compared to the classical notion of (boolean) function symbol in first-order logics~\cite{enderton2001mathematical}.
-Furthermore, one could see a function symbol as an ``oracle'' that is instantiated by the function assignment. 
-As far as we know, this is the first paper proposing this extension on weighted logics and it would be interesting to study this concept further (i.e. outside recursion).
+It is worth noting that function symbols in $\fqso$ represent functions from tuples to natural numbers, so they are different from 
+%cannot be compared to 
+the classical notion of function symbol in $\fo$~\cite{enderton2001mathematical}. 
+%~$\shl$.first-order logics
+Furthermore, %one could see 
+a function symbol can be seen as an ``oracle'' that is instantiated by the function assignment. 
+%As far as we know, 
+To the best of our knowledge, this is the first paper to propose 
+%proposing 
+this extension on weighted logics, which should be further investigated. 
+% and it would be interesting to study this concept further (i.e. outside recursion).
 
-We define an operator which extends Least Fixed Point Logic (LFP) \cite{I86,vardi1982complexity} over $\fo$ to allow counting. 
+We define an 
+extension of 
+%operator which extends 
+LFP \cite{I86,vardi1982complexity} 
+%over $\fo$ to 
+to allow counting. 
 %We show how to define this operator in $\qfo(\fo)$ and we discuss how to extend these ideas to $\qso$ towards the end of the section. 
-Fix a relational signature $\R$.
-Then the set of $\rqfo(\fo)$ formulas (Recursive $\qfo$) over $\R$ is defined as an extension of $\qfo(\fo)$ by adding the operator $\clfp{\beta(\x, h)}$ where $\x = (x_1, \ldots, x_\ell)$ is a sequence of $\ell$ distinct first-order variables and $\beta(\x, h)$ is an $\fqfo(\fo)$-formula over $\R$ whose only function symbol is $h$, with $\arity(h) = \ell$.
+%Fix a relational signature $\R$. 
+More precisely, the set of $\rqfo(\fo)$ formulas over a relational signature $\R$, where $\rqfo$ stands for recursive $\qfo$, is defined as an extension of $\qfo(\fo)$ that includes the formula 
+%by adding the operator 
+$\clfp{\beta(\x, h)}$, where (1) $\x = (x_1, \ldots, x_\ell)$ is a sequence of $\ell$ distinct first-order variables, (2) $\beta(\x, h)$ is an $\fqfo(\fo)$-formula over $\R$ whose only function symbol is $h$, and (3) $\arity(h) = \ell$.
 The free variables of the formula $\clfp{\beta(\x,h)}$ are $x_1, \ldots, x_\ell$; in particular, $h$ is not considered to be free in this formula.
 
-Fix an $\R$-structure with domain $A$ and a quantitative formula $\clfp{\beta(\x,h)}$, and assume that $\F$ is the set of functions $f :A^\ell \to \N$.
+Fix an $\R$-structure with domain $A$ and a quantitative formula $\clfp{\beta(\x,h)}$ with $\arity(h) = \ell$, and assume that $\F$ is the set of functions $f :A^\ell \to \N$.
 To define the semantics of $\clfp{\beta(\x,h)}$, we first show how $\beta(\x,h)$ can be interpreted as an operator $T_{\beta}$ on $\F$.
@@
 Moreover, assume that $b,c$ are elements of $\A$ and $v$ is a first-order assignment over $\A$ such that $v(x) = b$ and $v(y) = c$.
-Then we have that $\sem{\alpha(x,y,f)}(\A,v)$ is equal to the  number of paths in $\A$ from $b$ to $c$ of length at most $n$.
+Then we have that $\sem{\clfp{\alpha(x,y,f)}}(\A,v)$ is equal to the  number of paths in $\A$ from $b$ to $c$ of length at most $n$.
 
@@
 In fact, the number of paths of length at most $n$ from a node $x$ to a node $y$ is recursively computed by using the formula $(E(x,y) + \sa{z} g(x,z,t)\cdot E(z,y)) \cdot \varphi_{\text{\rm first}}(t)$, which stores this value in $g(x,y,t)$ with $t$ the first element in the domain.
-The other formula, the one with the filter $\varphi_{\text{succ}}(t',t)$, is used as a counter that allows to reach a fixed point in the support of function $g$ in $n$ steps.
+The other formula $\sa{t'} \varphi_{\text{succ}}(t',t) \cdot \left(\sa{x'} \sa{y'} g(x',y',t') \right)$ is just an auxiliary artifact that is used
+%The other formula, the one with the filter $\varphi_{\text{succ}}(t',t)$, is used 
+as a counter to allow reaching a fixed point in the support of $g$ in $n$ steps.
+Notice that the use of the filter $\varphi_{\text{succ}}(t',t)$ prevents this formula for incrementing the value of $g(x,y,t)$ when $t$ is the first element in the domain.
 \end{example}
@@
 %\end{theorem} 
-Note that, in contrast to $\lfp$-logic, here we do not need to impose any positive restriction to the formula $\beta(\x,h)$.
+%Note that, 
+In contrast with $\lfp$, to reach a fixed point 
+%here 
+we do not need to impose any positive restriction on the formula $\beta(\x,h)$.
 Indeed, since $\beta$ is constructed from monotones operations (i.e. sum and product) over the natural numbers, the resulting operator $T_{\beta}$ is monotone as well.
@@
 Now that a least fixed point operator over functions is defined, the next step is to understand its expressive power.
-In the following result we show how lsfp-operator can be used to capture all functions that can be computed in polynomial time.
+In the following theorem, we show that this operator can be used to capture $\fp$.
+%result we show how lsfp-operator can be used to capture all functions that can be computed in polynomial time.
 \begin{theorem} \label{rqfo-fo-cap}
-	$\rqfo(\fo)$ captures $\fp$ over the class of ordered structures.
+	$\rqfo(\fo)$ captures $\fp$ over 
+	%the class of 
+	ordered structures.
 \end{theorem}
@@
 Inspired by this work, we show that a restricted version of $\rqfo$ can be used to capture $\shl$, the counting version of $\nlog$. 
-Specifically, we use $\rqfo$ to define an operator for counting the number of paths in a directed graph which is exactly what we need for $\shl$. 
+Specifically, we use $\rqfo$ to define an operator for counting the number of paths in a directed graph, which is what is needed to capture $\shl$. 
 %Besides, we show that the same idea can be used to extend $\qso$ allowing to capture harder complexity classes.
 
-Given a relation signature $\R$, the set of transitive $\qfo$ formulas ($\tqfo$-formulas) is defined as an extension of $\qfo$ with the operator
+Given a relation signature $\R$, the set of transitive $\qfo$ formulas ($\tqfo$-formulas) is defined as an extension of $\qfo$ with the formula
 %\begin{multline}
@@
 [\pth \psi(\bar{x},\bar{y})]
-$ where $\psi(\x, \y)$ is an $\fo$-formula over $\R$, $\bar{x} = (x_1, \ldots, x_k)$, $\bar{y} = (y_1, \ldots, y_k)$ are tuples of pairwise distinct first-order variables.
+$, where $\psi(\x, \y)$ is an $\fo$-formula over $\R$, and $\bar{x} = (x_1, \ldots, x_k)$, $\bar{y} = (y_1, \ldots, y_k)$ are tuples of pairwise distinct first-order variables.
 The semantics of $[\pth \psi(\bar{x},\bar{y})]$ can easily be defined in terms of $\rqfo(\fo)$ as follows.
 Given an $\R$-structure $\A$ with domain $A$, define a (directed) graph $\cG_{\psi}(\A) = (N,E)$ such that $N = A^k$ and for every pair $\bar b, \bar c \in N$, it holds that $(\bar b, \bar c) \in E$ if, and only if, $\A \models \psi(\bar b, \bar c)$.
-Similar than for Example~\ref{ex:count-path}, we can count the paths of length at most $|A^k|$ with the formula $\beta_{\psi(\bar{x},\bar{y})}(\x, \y, \t, g)$:
+Similar than for Example~\ref{ex:count-path}, we can count the paths of length at most $|A^k|$ in $ \cG_{\psi}(\A)$ with the formula $\beta_{\psi(\bar{x},\bar{y})}(\x, \y, \t, g)$:
 \begin{multline*}
 (\psi(\bar{x},\bar{y}) + \sa{\z} g(\x,\z,\t)\cdot \psi(\z,\y)) \cdot \varphi_{\text{\rm first-lex}}(\t) \ +\\
-\sa{\t'} \varphi_{\text{succ-lex}}(\t',\t) \cdot \left(\sa{\x'} \sa{\y'} g(\x',\y',\t') \right)
+\sa{\t'} \varphi_{\text{succ-lex}}(\t',\t) \cdot \left(\sa{\x'} \sa{\y'} g(\x',\y',\t') \right),
 \end{multline*}
-where $\varphi_{\text{\rm first-lex}}$ and $\varphi_{\text{succ-lex}}$ are $\fo$-formulas for defining the first and successor predicates over tuples in $A^k$, following the lexicographic order induced by~$<$.
-Then the semantics of the path operator is defined $\rqfo$ formula:
+where $\varphi_{\text{\rm first-lex}}$ and $\varphi_{\text{succ-lex}}$ are $\fo$-formulas defining the first and successor predicates over tuples in $A^k$, following the lexicographic order induced by~$<$.
+Then the semantics of the path operator can be defined by using the following definition of $[\pth \psi(\bar{x},\bar{y})]$ in $\rqfo$:
 \[
-[\pth \psi(\bar{x}, \bar{y})] := \sa{\t} (\varphi_{\text{\rm first}}(\t) \wedge \clfp{\beta_{\psi(\bar{x},\bar{y})}(\x,\y,\t,g)})
+[\pth \psi(\bar{x}, \bar{y})] \ :=\ \sa{\t} (\varphi_{\text{\rm first}}(\t) \wedge \clfp{\beta_{\psi(\bar{x},\bar{y})}(\x,\y,\t,g)}).
 \]
-In other words, $\sem{[\pth \psi(\bar{x}, \bar{y})]}(\A,v,V)$ counts the number of paths in the graph $\cG_{\psi}(\A)$ whose length is at most~$|A^k|$.
+In other words, $\sem{[\pth \psi(\bar{x}, \bar{y})]}(\A,v)$ counts the number of paths from $v(\bar x)$ to $v(\bar y)$ in the graph $\cG_{\psi}(\A)$ whose length is at most~$|A^k|$.
 %As for the case of $\qso$, the logic $\tqso(\LL)$ is obtained by restricting $\varphi$ in \eqref{eq-def-tqso} to be a formula in $\LL$.
@@
 \begin{theorem} \label{tqfo-shl}
-	$\tqfo(\fo)$ captures $\shl$ over the class of ordered structures.
+	$\tqfo(\fo)$ captures $\shl$ over ordered structures.
 \end{theorem}
-The last result perfectly illustrates the benefits of the logical framework for the descriptive complexity of counting functions. 
-The distinction of the language between boolean and quantitative allows us to define operators at the later that cannot be defined at the former. 
-Indeed, it is not clear how to capture $\shl$ with $\qfo(\tc)$\footnote{$\tc$ corresponds to $\fo$-logic extended with transitive closure}, or any other extension of $\qfo(\fo)$ at the boolean level.
+This last result 
+perfectly 
+illustrates the benefits of our logical framework for the development of descriptive complexity for counting %functions.
+complexity classes.  
+The distinction in the language between the Boolean and the quantitative level allows us to define operators at the later level that cannot be defined at the former. 
+As a example showing how fundamental this separation is, consider the issue of extending $\qfo(\fo)$ at the Boolean level in order to capture $\shl$.
+The natural alternative to do this is to use $\fo$ extending with a transitive closure operator, which is denoted by $\tc$.
+But then the problem is that for every language $L \in \nlog$, it holds that its characteristic function $\chi_L \in \qfo(\tc)$, where $\chi_L(w) = 1$ if $w \in L$, and $\chi_L(w) = 0$ otherwise.
+Thus, if we assume that
+
+$\qfo(\tc)$ can def
+In fact, it is not clear how $\shl$ could be capture by 
+to capture $\shl$ with $\qfo(\tc)$\footnote{$\tc$ corresponds to $\fo$-logic extended with transitive closure.}, or any other extension of $\qfo(\fo)$ at the boolean level.
 In contrast, with $\rqfo$ one can naturally define the path operator at the quantitative level and easily capture $\shl$.

### 2017-01-09 LICS/beyond.tex
sc

@@
 As a example showing how fundamental this separation is, consider the issue of extending $\qfo(\fo)$ at the Boolean level in order to capture $\shl$.
-The natural alternative to do this is to use $\fo$ extending with a transitive closure operator, which is denoted by $\tc$.
-But then the problem is that for every language $L \in \nlog$, it holds that its characteristic function $\chi_L \in \qfo(\tc)$, where $\chi_L(w) = 1$ if $w \in L$, and $\chi_L(w) = 0$ otherwise.
-Thus, if we assume that
-
-$\qfo(\tc)$ can def
-In fact, it is not clear how $\shl$ could be capture by 
-to capture $\shl$ with $\qfo(\tc)$\footnote{$\tc$ corresponds to $\fo$-logic extended with transitive closure.}, or any other extension of $\qfo(\fo)$ at the boolean level.
-In contrast, with $\rqfo$ one can naturally define the path operator at the quantitative level and easily capture $\shl$.
-Therefore, Theorem~\ref{tqfo-shl} is good example of how to go beyond classical logic to have a better understanding of the expressivity of counting.
+The natural alternative to do this is to use $\fo$ extended with a transitive closure operator, which is denoted by $\tc$.
+But then the problem is that for every language $L \in \nlog$, it holds that its characteristic function $\chi_L$ is in $\qfo(\tc)$, where $\chi_L(x) = 1$ if $x \in L$, and $\chi_L(x) = 0$ otherwise.
+Thus, if we assume that $\qfo(\tc)$ captures $\shl$ (over ordered structures), then we have that $\chi_L \in \nlog$ for every $L \in \nlog$.
+This implies that $\nlog = \ulog$,\footnote{A decision language $L$ is in $\ulog$ is there exists a logarithmic-space NTM $M$ accepting $L$ and satisfying that $\tma_M(x) = 1$ for every $x \in L$.} and thus contradicts the widely believed assumption that $\ulog \subsetneq \nlog$.
+%$\qfo(\tc)$ can def
+%In fact, it is not clear how $\shl$ could be capture by 
+%to capture $\shl$ with $\qfo(\tc)$\footnote{$\tc$ corresponds to $\fo$-logic extended with transitive closure.}, or any other extension of $\qfo(\fo)$ at the boolean level.
+%In contrast, with $\rqfo$ one can naturally define the path operator at the quantitative level and easily capture $\shl$.
+%Hence, Theorem~\ref{tqfo-shl} is good example of how to go beyond classical logic to have a better understanding of the expressivity of counting.
 

### 2017-01-09 LICS/appendix-counting.tex
Demostración de max y min.

@@
 Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\fo$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial time, and accepts if and only if it holds true.
+If $\alpha$ is a $\fo$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ deterministically in polynomial time, and accepts if and only if it holds true.
 If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
@@
 If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time \cite{fagin1974generalized}, and accepts if and only if the condition holds true.
-If $\alpha$ is a constant $s$, then the algorithm produces $s$ branches and accepts in all of them. 
+If $\alpha$ is a constant $s$, then the machine produces $s$ branches and accepts in all of them. 
 If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.  
@@
 This covers all possible cases for $\alpha$.
-Additionally, the algorithm produces a different output on each path. This can be done by printing the trace of all the non-deterministic choices.
-However, when the algorithm starts checking whether $(\A,v,V)\models\varphi$ for some $\eso$ formula $\varphi$, it stops printing in the output tape.
-This way the algorithm produces exactly one output from that point onwards.
+Additionally, the machine produces a different output on each path. This can be done by printing the trace of all the non-deterministic choices.
+However, when the machine starts checking whether $(\A,v,V)\models\varphi$ for some $\eso$ formula $\varphi$, it stops printing in the output tape.
+This way the machine produces exactly one output from that point onwards.
 Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
@@
 Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\pfp$-formula $\varphi$, then the algorithm checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space~\cite{L04}, and accepts if and only if it holds true.
+If $\alpha$ is a $\pfp$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space~\cite{L04}, and accepts if and only if it holds true.
 If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
@@
 Furthermore, each relation $X$ indexes a position in the binary output of $f(\enc(\A))$ as follows.
-Consider a polynomial space algorithm over the $\R$ that receives as input an $\R$-structure $\A$ and a number $p$ encoded by a relation $X$.
-Then the algorithm accepts if, and only if, the $p$-th bit of $f(\enc(\A))$ is $1$.
+Consider a polynomial space machine over the $\R$ that receives as input an $\R$-structure $\A$ and a number $p$ encoded by a relation $X$.
+Then the machine accepts if, and only if, the $p$-th bit of $f(\enc(\A))$ is $1$.
 Since this procedure works in polynomial space, it can be described in $\pfp$ \cite{AbiteboulV89} by a formula $\Phi(X)$ where the free variable $X$ encodes the number $p$.
@@
 Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
+
+%%% DEMOSTRACION DE GapP y SumQSO en Z  
+\subsection*{Proof of Proposition~\ref{prop:capture-gapp}}
+
+\cristian{FALTA LA DEMOSTRACION DE GapP y SumQSO en Z}
+
+%%% DEMOSTRACION DE optP y MaxQSO, MinQSO
+\subsection*{Proof of Theorem~\ref{theo:capture-optp}}
+
+Similar than the previous proofs, we construct recursively a non-deterministic polynomial time Turing machine $M_{\alpha}$ with output tape for each $\maxqso(\fo)$ formula $\alpha$ over a signature $\R$.
+This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ (in binary) over the output tape of some run, and this value is the maximum value over all runs.
+Suppose $\A$ has domain $A$.
+If $\alpha$ is a $\fo$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ in deterministic polynomial time, and output $1$ if, and only if, $(\A,v,V)$ satisfies $\varphi$.
+If $\alpha$ is a constant $s$, it outputs $s$ in binary over the output tape.
+If $\alpha = (\beta \add \gamma)$, then it simulates $M_{\beta}$ and $M_{\gamma}$ on input $(\A,v,V)$, adds the output of both machines and prints this value over the output tape.
+If $\alpha = (\beta \mult \gamma)$ or $\alpha = \max\{\beta,\gamma\}$, then it does the same than the previous case but it multiplies or maximizes, respectively, the outputs of both machines instead of adding.
+If $\alpha = \sa{x}\beta$, it iterates over all elements $a\in A$ simulating and adding the output of $M_{\beta}$ on input $(\A,v[a/x],V)$.
+$M_{\alpha}$ finally outputs the aggregated value in the output tape.
+If $\alpha = \pa{x}\beta$ or $\alpha = \maxa{x}\beta$, then $M_{\alpha}$ does the same idea than the previous case with the difference that the output of $M_{\beta}$ on input $(\A,v[a/x],V)$ is multiplied or maximized, respectively.
+If $\alpha = \maxa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
+This covers all possible cases for $\alpha$.
+Furthermore, it is straightforward to prove that each of these steps are correct and can be computed with a non-deterministic polynomial time Turing machine with output tape.
+Let $\alpha$ be a formula in $\maxqso$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the maximum run (with respect to the output value) of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
+We have that $f$ is a $\maxp$ function over $\R$ and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
+Finally, one can easily see that the same construction holds with $\minqso(\fo)$ by constructing a Turing machine that take the min over all runs instead of max.
+
+The proof for the other direction is similar than in \cite{kolaitis1994logical} extended with the ideas of previous proofs. Let $f\in \maxp$ be a function defined over some signature $\R$ and
+$\ell\in\nat$ such that $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$ for each $\A\in\ostr[\R]$ of size $n$.
+For $U \subseteq A^{\ell}$, we can interpret the encoding of $U$, i.e. $\enc(U)$, as the binary encoding of a number with $n^l$-bits. We denote this value by $\val(\enc(U))$.
+Then, given $\A\in\ostr[\R]$ with domain $A$ and $U \subseteq A^{\ell}$, consider the problem of checking whether $f(\enc(\A)) \geq \val(\enc(U))$. 
+Clearly, this is an $\np$-problem and, by Fagin's theorem, there exists a formula of the form $\ex{\bar{X}} \Phi(\bar{X}, Y)$ with $\Phi(\bar{X}, Y)$ in $\fo$-logic and $\arity(Y) = \ell$ such that $f(\enc(\A)) \geq \val(\enc(U))$ if, and only if, $(\A,v,V) \models \ex{\bar{X}} \Phi(\bar{X}, Y)$ with $V(Y) = U$.
+Then we can describe the function $f$ by the following $\maxqso$ formula:
+$$
+\alpha = \maxa{\bar{X}} \maxa{Y} \ \Phi(\bar{X}, Y) \cdot \big( \sa{\bar{x}} Y(\bar{x}) \cdot \pa{\bar{y}}(\bar{x} < \bar{y} \mapsto 2) \big).
+$$
+Note that, in contrast with previous proofs, we use $\bar{x} < \bar{y}$ instead of $\bar{y} < \bar{x}$ because the most significant bit in $\enc(U)$ correspond to the smallest tuple in $U$.  
+It is easy to check that $\Phi(\bar{X}, Y)$ simulates the NP-machine and, if $\Phi(\bar{X}, Y)$ holds, the formula to the right  reconstructs the binary output from the relation in $Y$.
+Then, $\alpha$ is in $\maxqso(\fo)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
+
+For the case of $\minqso(\fo)$ and a function $f \in \minp$, one has to follow the same approach but considering the problem of checking whether $f(\enc(\A)) \leq \val(\enc(U))$.
+Then, the formula $\alpha$ for $f$ will be:
+$$
+\alpha = \mina{\bar{X}} \mina{Y} \ \sa{\bar{x}} \big( (\Phi(\bar{X}, Y) \rightarrow Y(\bar{x})) \cdot \pa{\bar{y}}(\bar{x} < \bar{y} \mapsto 2)  \big).
+$$
+In this case, if the formula $\Phi(\bar{X}, Y)$ is false, then the output produced by the subformula inside the $\min$-quantifiers will be the biggest possible value (i.e. $2^{n^\ell}$).
+On the other hand, if $\Phi(\bar{X}, Y)$ holds, the subformula will produce $\val(\enc(U))$. 
+Similar than for $\max$, we conclude that $\alpha$ is in $\minqso(\fo)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.

### 2017-01-09 LICS/beyond.tex

@@
 %, which is 
-denoted by $\arity(h)$. Then the set of $\fqso$ formulas over a relational signature $\R$ is defined by the following grammar:
+denoted by $\arity(h)$. Then the set of $\fqso$ formulae over a relational signature $\R$ is defined by the following grammar:
 \begin{multline}
@@
 %Fix a relational signature $\R$. 
-More precisely, the set of $\rqfo(\fo)$ formulas over a relational signature $\R$, where $\rqfo$ stands for recursive $\qfo$, is defined as an extension of $\qfo(\fo)$ that includes the formula 
+More precisely, the set of $\rqfo(\fo)$ formulae over a relational signature $\R$, where $\rqfo$ stands for recursive $\qfo$, is defined as an extension of $\qfo(\fo)$ that includes the formula 
 %by adding the operator 
@@
 Assume now that we need to extend our previous counting function to the case of arbitrary directed graphs.
-To this end, suppose that $\varphi_{\text{\rm first}}(x)$ and $\varphi_{\text{succ}}(x,y)$ are the $\fo$-formulas for defining the first and successor predicates, respectively, of $<$.
+To this end, suppose that $\varphi_{\text{\rm first}}(x)$ and $\varphi_{\text{succ}}(x,y)$ are the $\fo$-formulae for defining the first and successor predicates, respectively, of $<$.
 Moreover, define formula $\beta(x, y, t, g)$ as follows:
@@
 $$
-\sa{t} (\varphi_{\text{\rm first}}(t) \wedge \clfp{\beta(x,y,t,g)}).
+\sa{t} (\varphi_{\text{\rm first}}(t) \cdot \clfp{\beta(x,y,t,g)}).
 $$ 
@@
 
-%Given a relation signature $\R$, the set of recursive $\qfo$ formulas ($\rqfo$-formulas) is defined by the following grammar:
-%%This operator lets us define the set of recursive $\qfo$ formulas over $\R$ ($\rqfo$-formulas) using the following grammar:
+%Given a relation signature $\R$, the set of recursive $\qfo$ formulae ($\rqfo$-formulae) is defined by the following grammar:
+%%This operator lets us define the set of recursive $\qfo$ formulae over $\R$ ($\rqfo$-formulae) using the following grammar:
 %\begin{multline*}
@@
 
-Given a relation signature $\R$, the set of transitive $\qfo$ formulas ($\tqfo$-formulas) is defined as an extension of $\qfo$ with the formula
+Given a relation signature $\R$, the set of transitive $\qfo$ formulae ($\tqfo$-formulae) is defined as an extension of $\qfo$ with the formula
 %\begin{multline}
@@
 \end{multline*}
-where $\varphi_{\text{\rm first-lex}}$ and $\varphi_{\text{succ-lex}}$ are $\fo$-formulas defining the first and successor predicates over tuples in $A^k$, following the lexicographic order induced by~$<$.
+where $\varphi_{\text{\rm first-lex}}$ and $\varphi_{\text{succ-lex}}$ are $\fo$-formulae defining the first and successor predicates over tuples in $A^k$, following the lexicographic order induced by~$<$.
 Then the semantics of the path operator can be defined by using the following definition of $[\pth \psi(\bar{x},\bar{y})]$ in $\rqfo$:
 \[
-[\pth \psi(\bar{x}, \bar{y})] \ :=\ \sa{\t} (\varphi_{\text{\rm first}}(\t) \wedge \clfp{\beta_{\psi(\bar{x},\bar{y})}(\x,\y,\t,g)}).
+[\pth \psi(\bar{x}, \bar{y})] \ :=\ \sa{\t} (\varphi_{\text{\rm first}}(\t) \cdot \clfp{\beta_{\psi(\bar{x},\bar{y})}(\x,\y,\t,g)}).
 \]
@@
 %As for the case of $\qso$, the logic $\tqso(\LL)$ is obtained by restricting $\varphi$ in \eqref{eq-def-tqso} to be a formula in $\LL$.
-Moreover, the logic $\tqfo$ is obtained by disallowing in \eqref{eq-def-tqso} formulas $\sa{X} \alpha$ and $\pa{X} \alpha$, and by only allowing  first-order free-variables in the formula $\psi$ used in $[\pth \psi]$ in \eqref{eq-def-tqso}.
+Moreover, the logic $\tqfo$ is obtained by disallowing in \eqref{eq-def-tqso} formulae $\sa{X} \alpha$ and $\pa{X} \alpha$, and by only allowing  first-order free-variables in the formula $\psi$ used in $[\pth \psi]$ in \eqref{eq-def-tqso}.
 As it was previously said, the operator for counting paths is exactly what we need to capture $\shl$.
@@
 
-%We also define the set of transitive $\qso$ formulas ($\tqso$-formulas) using the following grammar:
+%We also define the set of transitive $\qso$ formulae ($\tqso$-formulae) using the following grammar:
 %\begin{multline*}
@@
 %For a given first order assignment $v$ and a second order asssignment $V$, let $\bar{a} = v(\bar{x})$ and $\bar{b} = v(\bar{y})$, and $\sem{[\pth\, \psi(\bar{x},\bar{y})]}(\A,v,V)$ will take the value of the number of paths of size less or equal to $n$ from $\bar{a}$ to $\bar{b}$ in the graph ${\cal G}$.
-This operator lets us define the set of transitive $\qfo$ formulas over $\R$ ($\tqfo$-formulas) using the following grammar:
+This operator lets us define the set of transitive $\qfo$ formulae over $\R$ ($\tqfo$-formulae) using the following grammar:
 %\begin{multline*} 
@@
 %
-%We also define the set of transitive $\qso$ formulas ($\tqso$-formulas) using the following grammar:
+%We also define the set of transitive $\qso$ formulae ($\tqso$-formulae) using the following grammar:
 %\begin{multline*}

### 2017-01-09 LICS/counting.tex
sc

@@
 Then let $\enc(R_i^\A)$ be a binary string of length $n^\ell$ such that the $i$-th bit of $\enc(R_i^\A)$ is 1 if the $i$-th tuple in the previous enumeration belongs to $R_i^\A$, and 0 otherwise.
-Moreover, define the encoding of $\A$, denoted by $\enc(\A)$, as the following binary string~\cite{L04}:
-\begin{eqnarray*}
-	\enc(\A) & = & 0^n \, 1 \, \enc(R_1^\A) \, \cdots \, \enc(R_k^\A).
-\end{eqnarray*}
+Moreover, define the encoding of $\A$, denoted by $\enc(\A)$, as the
+%following binary string~\cite{L04}:
+%\begin{eqnarray*}
+	%\enc(\A) & = & 0^n \, 1 \, \enc(R_1^\A) \, \cdots \, \enc(R_k^\A).
+%\end{eqnarray*}
+$\enc(\A) =  0^n \, 1 \, \enc(R_1^\A) \, \cdots \, \enc(R_k^\A)$~\cite{L04}. 
 %We define the class of all $\R$-functions, denoted by $\Func(\R)$, as the class of all functions $f: \ostr \rightarrow \bbN$.
@@
 We can now formalize the notion of capturing a counting complexity class.
+From now on, we assume that every counting complexity class $\CC$ consists of functions $f : \{0,1\}^* \to \N$ that are invariant under isomorphism of structures.
+More precisely, given a relational signature $\R$, a function $f : \{0,1\}^* \to \N$ is invariant under $\R$-isomorphisms if for every pair $\A, \mathfrak{B} \in \ostr[\R]$ such that $\A$ and
+$\mathfrak{B}$ are isomorphic, it holds that $f(\enc(\A)) = f(\enc(\mathfrak{B}))$. It is worth noticing that this is a usual assumption in descriptive complexity~\cite{L04}.
 %over $\R$-structures.
@@
 
-		\item for every $f \in \CC$, there exists $\alpha \in \FF$ such that   $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A \in \ostr[\R]$.
+		\item for every $f \in \CC$ that is invariant under $\R$-isomorphisms, there exists $\alpha \in \FF$ such that   $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A \in \ostr[\R]$.
 	\end{enumerate} 

### 2017-01-09 LICS/appendix-counting.tex
Demostración lista.

@@
 
-%%% DEMOSTRACION DE GapP y SumQSO en Z  
-\subsection*{Proof of Proposition~\ref{prop:capture-gapp}}
-
-\cristian{FALTA LA DEMOSTRACION DE GapP y SumQSO en Z}
-
 %%% DEMOSTRACION DE optP y MaxQSO, MinQSO
@@
 
-The proof for the other direction is similar than in \cite{kolaitis1994logical} extended with the ideas of previous proofs. Let $f\in \maxp$ be a function defined over some signature $\R$ and
+The proof for the other direction is similar than in \cite{kolaitis1994logical} extended with the ideas of Theorem~\ref{theo:capture-fp}.
+Let $f\in \maxp$ be a function defined over some signature $\R$ and
 $\ell\in\nat$ such that $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$ for each $\A\in\ostr[\R]$ of size $n$.
-For $U \subseteq A^{\ell}$, we can interpret the encoding of $U$, i.e. $\enc(U)$, as the binary encoding of a number with $n^l$-bits. We denote this value by $\val(\enc(U))$.
-Then, given $\A\in\ostr[\R]$ with domain $A$ and $U \subseteq A^{\ell}$, consider the problem of checking whether $f(\enc(\A)) \geq \val(\enc(U))$. 
+For $U \subseteq A^{\ell}$, we can interpret the encoding of $U$ ($\enc(U)$) as the binary encoding of a number with $n^l$-bits. We denote this value by $\val(\enc(U))$.
+Then, given $\A\in\ostr[\R]$ and $U \subseteq A^{\ell}$, consider the problem of checking whether $f(\enc(\A)) \geq \val(\enc(U))$. 
 Clearly, this is an $\np$-problem and, by Fagin's theorem, there exists a formula of the form $\ex{\bar{X}} \Phi(\bar{X}, Y)$ with $\Phi(\bar{X}, Y)$ in $\fo$-logic and $\arity(Y) = \ell$ such that $f(\enc(\A)) \geq \val(\enc(U))$ if, and only if, $(\A,v,V) \models \ex{\bar{X}} \Phi(\bar{X}, Y)$ with $V(Y) = U$.
@@
 
-For the case of $\minqso(\fo)$ and a function $f \in \minp$, one has to follow the same approach but considering the problem of checking whether $f(\enc(\A)) \leq \val(\enc(U))$.
-Then, the formula $\alpha$ for $f$ will be:
+For the case of $\minqso(\fo)$ and a function $f \in \minp$, one has to follow the same approach but considering the $\np$-problem of checking whether $f(\enc(\A)) \leq \val(\enc(U))$.
+Then, the formula for describing $f$ is the following:
 $$

### 2017-01-09 LICS/syntactic-clo.tex
Acortando el paper.

@@
 
-A natural question at this point is whether the property of being closed under subtraction by one could be generalized to proper subtraction.
+%A natural question at this point is whether the property of being closed under subtraction by one could be generalized to proper subtraction.
 More precisely, a function complexity class $\CC$ is said to be closed under subtraction if for every $f,g \in \CC$, it holds that $(f \dotdiv g) \in \CC$, where $(f \dotdiv g)(x)$ is defined as $f(x) - g(x)$ if $f(x) \geq g(x)$, and as $0$ otherwise.
 The following theorem shows that such generalization does not work for any of the function complexity classes considered in this section.
-\begin{theorem} \label{sub-pnp}
-If any of the classes $\E{1}$, $\QE{1}$ or $\eqso(\logex{1})$ is closed under subtraction, then $\ptime = \np$.
-\end{theorem}
+%\begin{theorem} \label{sub-pnp}
+%If any of the classes $\E{1}$, $\QE{1}$ or $\eqso(\logex{1})$ is closed under subtraction, then $\ptime = \np$.
+%\end{theorem}
 

### 2017-01-09 LICS/beyond.tex

@@
 Thus, if we assume that $\qfo(\tc)$ captures $\shl$ (over ordered structures), then we have that $\chi_L \in \nlog$ for every $L \in \nlog$.
-This implies that $\nlog = \ulog$,\footnote{A decision language $L$ is in $\ulog$ is there exists a logarithmic-space NTM $M$ accepting $L$ and satisfying that $\tma_M(x) = 1$ for every $x \in L$.} and thus contradicts the widely believed assumption that $\ulog \subsetneq \nlog$.
+This implies that $\nlog = \ulog$,
+\footnote{A decision language $L$ is in $\ulog$ is there exists a logarithmic-space NTM $M$ accepting $L$ and satisfying that $\tma_M(x) = 1$ for every $x \in L$.} 
+and thus contradicts the widely believed assumption that $\ulog \subsetneq \nlog$.
 %$\qfo(\tc)$ can def

### 2017-01-09 LICS/conclusions.tex

@@
 
-Here comes the conclusions.
+We proposed a framework based on Weighted Logics to develop a descriptive complexity theory for complexity classes of functions.
+%In particular, we show how this framework can be used to capture fundamental counting complexity classes such as $\fp$, $\shp$ and $\fpspace$, among others.
+Moreover, we use it to define a hierarchy inside $\shp$, identifying counting complexity classes with good closure and approximation properties, and which admit natural complete problems.
+Finally, by adding recursion to the framework, we show how to capture lower counting complexity classes such as $\shl$.
+We consider the results of this paper as a first step in this direction.
+%towards a descriptive complexity theory for complexity classes of functions. ç
+In this sense, there are several directions for future research, some of which are mentioned here. 
+$\totp$ is an important counting complexity class for which we are missing a logical characterization.
+In the same way, we are missing logical characterization of harder fundamental complexity class such as $\spanl$.
+We would also like to define a larger syntactic subclass of $\shp$ where each function admits an FPRAS; notice that $\cpm$ is an important problem admitting an FPRAS\cite{JSV04} that is not included in the classes defined in Section \ref{sec-clo}.
+Moreover, by following the approach proposed in
+\cite{I83}, we would like to include second-order free variables in the operator for counting paths introduced in Section \ref{sec:beyond}, so to have alternative ways to capture $\fpspace$ and even $\shp$.
+Finally, the least fixed point operator introduced in Section \ref{sec:beyond} clearly deserves further investigation.

### 2017-01-09 LICS/counting.tex

@@
 The other direction of Proposition \ref{prop:capture-shP} is straightforward to prove, and it follows by the fact that $\shp$ is closed under first- and second-order sum.
-Moreover, these closure properties also hold if we change the boolean core $\fo$ to any polynomial-time logic like $\lfp$. This gives the following corollary.
-\begin{corollary} \label{prop:capture-shP-LFP}
-	$\eqso(\lfp)$ captures $\shp$ over ordered structures.
-\end{corollary}
+%Moreover, these closure properties also hold if we change the boolean core $\fo$ to any polynomial-time logic like $\lfp$. This gives the following corollary.
+%\begin{corollary} \label{prop:capture-shP-LFP}
+%	$\eqso(\lfp)$ captures $\shp$ over ordered structures.
+%\end{corollary}
+
 By following the same approach as~\cite{SalujaST95}, Compton and Gr\"adel~\cite{ComptonG96} show that $\seso$ captures $\spp$, where $\eso$ is the existential fragment of $\so$.

### 2017-01-09 LICS/extentions.tex

@@
 Adding negative constants is a mild extension to allow subtraction in the logic. 
-Interestingly, this is exactly what we need to capture  $\gp$.
-\begin{proposition} \label{prop:capture-gapp}
+It follows from our characterization of $\shp$ that this is exactly what we need to capture  $\gp$.
+\begin{corollary} \label{prop:capture-gapp}
 	$\eqsoz(\fo)$ captures $\gp$ over ordered structures.
-\end{proposition}
+\end{corollary}
 %The previous result is straightforward to prove from the characterization of $\shp$ and the definition of $\gp$. 

### 2017-01-09 LICS/saluja.tex

@@
 \node[rectw, right=0.5cm of n1] (n2) {};
-\node[rectw, above=0.5cm of n2] (n3) {$\E{1}$}
+\node[rectw, above=0.4cm of n2] (n3) {$\E{1}$}
 	edge[draw=white] node {\rotatebox{45}{$\subsetneq$}} (n1);
-\node[rectw, below=0.5cm of n2] (n4) {$\QE{0}$}
+\node[rectw, below=0.4cm of n2] (n4) {$\QE{0}$}
         edge[draw=white] node {\rotatebox{315}{$\subsetneq$}} (n1);
@@
 \end{center}
-
+\vspace{-0.4cm}
 \caption{The relationship between the $\sfo$-hierarchy and the $\eqso(\fo)$-hierarchy, where $\E{1}$ and $\QE{0}$ are incomparable. \label{fig-sfo-eqso}}

### 2017-01-09 LICS/beyond.tex
Acortando el paper.
Cortes varios y quedamos en 12 páginas.

@@
 %cannot be compared to 
-the classical notion of function symbol in $\fo$~\cite{enderton2001mathematical}. 
+the classical notion of function symbol in $\fo$~\cite{L04}. 
 %~$\shl$.first-order logics
@@
 $\clfp{\beta(\x, h)}$, where (1) $\x = (x_1, \ldots, x_\ell)$ is a sequence of $\ell$ distinct first-order variables, (2) $\beta(\x, h)$ is an $\fqfo(\fo)$-formula over $\R$ whose only function symbol is $h$, and (3) $\arity(h) = \ell$.
-The free variables of the formula $\clfp{\beta(\x,h)}$ are $x_1, \ldots, x_\ell$; in particular, $h$ is not considered to be free in this formula.
+The free variables of the formula $\clfp{\beta(\x,h)}$ are $x_1, \ldots, x_\ell$; in particular, $h$ is not considered to be free.
 
@@
 In the definition of the semantics of LFP, an operator $T$ on relations is considered, and the semantics is defined in terms of the least fixed point of $T$, that is, a relation $R$ such that~\cite{I86,vardi1982complexity}:
-\begin{enumerate}
- \item[(a)] $T(R) = R$, and 
- \item[(b)]  $R \subseteq S$ for every $S$ such that $T(S) = S$.  
-\end{enumerate}
+(a) $T(R) = R$, and (b) $R \subseteq S$ for every $S$ such that $T(S) = S$.  
+%\begin{enumerate}
+% \item[(a)] $T(R) = R$, and 
+% \item[(b)]  $R \subseteq S$ for every $S$ such that $T(S) = S$.  
+%\end{enumerate}
 We can view $T$ as an operator on functions if we consider the characteristic function of a relation.
@@
 In particular, the conditions defining the least fixed point of $T$ can be rewritten as follows in terms of $T^\star$:
-\begin{enumerate}
-	\item[(a)] $\support(T^\star(\chi_R)) = \support(\chi_R)$, and  
-	\item[(b)] $\support(\chi_R) \subseteq \support(\chi_S)$ for every $S$ such that  $\support(T^\star(\chi_{S})) = \support(\chi_S)$.  
-\end{enumerate}
+(a) $\support(T^\star(\chi_R)) = \support(\chi_R)$, and  (b) $\support(\chi_R) \subseteq \support(\chi_S)$ for every $S$ such that  $\support(T^\star(\chi_{S})) = \support(\chi_S)$.
+%\begin{enumerate}
+%	\item[(a)] $\support(T^\star(\chi_R)) = \support(\chi_R)$, and  
+%	\item[(b)] $\support(\chi_R) \subseteq \support(\chi_S)$ for every $S$ such that  $\support(T^\star(\chi_{S})) = \support(\chi_S)$.  
+%\end{enumerate}
 To define a notion of fixed point for $T_\beta$ we simply generalized these conditions. More precisely, %These conditions can be generalized in the obvious way to the case of the operator $T_\beta$: 
@@
 \end{multline*}
-Then our extended counting function is defined by the following formula:
+Then our extended counting function is defined by:
 $$

### 2017-01-09 LICS/counting.tex
Cortes varios y quedamos en 12 páginas.

@@
 Thus, it directly follows from~\cite{SalujaST95}  that every $\shp$-function can be defined in $\eqso(\LL)$.
-The other direction of Proposition \ref{prop:capture-shP} is straightforward to prove, and it follows by the fact that $\shp$ is closed under first- and second-order sum.
+The other direction of Proposition \ref{prop:capture-shP} follows by the fact that $\shp$ is closed under first- and second-order sum.
 %Moreover, these closure properties also hold if we change the boolean core $\fo$ to any polynomial-time logic like $\lfp$. This gives the following corollary.
@@
 \end{theorem}
-The proof of the previous theorem follows the same line as for the logical characterization of $\fp$: one shows that each function in $\qso(\pfp)$ can be computed in $\fpspace$ and, conversely, the output of a polynomial-space TM can be reconstructed by using $\pfp$ and second-order quantitative quantifiers.
+The proof of the previous theorem follows the same line as for the logical characterization of $\fp$: one shows that each function in $\qso(\pfp)$ can be computed in $\fpspace$ and, conversely, the output of a polynomial-space TM can be reconstructed by using $\pfp$ and quantitative quantifiers.
 

### 2017-01-09 LICS/framework.tex
Cortes varios y quedamos en 12 páginas.

@@
 where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
-Moreover, if $\R$ is not mentioned, then $\qso$ refers to the union of the sets of $\qso$ formulae over all posible relational signatures.
+Moreover, if $\R$ is not mentioned, then $\qso$ refers to the set of $\qso$ formulae over all posible relational signatures.
 %\marcelo{En la gramatica de la logica deberiamos incluir la formula $\top$ que discutimos, la cual representa true y para la cual se tiene que $\sem{\top}(\A, v, V) = 1$. Les parece?}
@@
 	\label{tab-semantics}
-	\vspace*{-20pt}
+	\vspace*{-21pt}
 \end{table}
@@
 Let $\bG = \{E(\cdot,\cdot),<\}$ be the vocabulary for graphs and $\fG$ be an ordered $\bG$-structure encoding a non-directed graph. 
-Suppose that we want to count the number of triangles in $\fG$. Then this can be defined with the following QSO-formula:
+Suppose that we want to count the number of triangles in $\fG$. Then this can be defined as follows:
 \begin{multline*}
@@
 For a more involved example that includes multiplication, let $\bM = \{M(\cdot,\cdot)\}$ be a vocabulary for storing 0-1 matrices; in particular, a structure $\fM$ over $\bM$ encodes a 0-1 matrix $A$ as follows: if $A[i,j] = 1$, then $M(i,j)$ is true, otherwise $M(i.j)$ is false.
-Suppose now that we want to compute the permanent of an $n$-by-$n$ 0-1 matrix $A$, which is defined as:
+Suppose now that we want to compute the permanent of an $n$-by-$n$ 0-1 matrix $A$, that is:
 \begin{eqnarray*}

### 2017-01-09 LICS/preliminaries.tex
Cortes varios y quedamos en 12 páginas.

@@
 
-In this section, we introduce the main terminology used in this paper.
+%In this section, 
+%We introduce here the main terminology used in the paper.
 
@@
 Similarly, given a second-order assignment $V$ for $\A$, $X \in \sv$ and $B  \subseteq A^{\arity(X)}$, we denote by $V[B/X]$ a second-order assignment such that $V[B/X](X) = B$ and $V[B/X](Y) = V(Y)$ for every $Y \in \sv$ distinct from $X$.
-We use notation $(\A, v, V) \models \varphi$ to indicate that structure $\A$ satisfies $\varphi$ under the variable assignments $v$ and $V$.
+We use notation $(\A, v, V) \models \varphi$ to indicate that structure $\A$ satisfies $\varphi$ under $v$ and $V$.
 In particular, we have that $(\A, v, V) \models \top$.

### 2017-01-09 LICS/syntactic-horn.tex
Cortes varios y quedamos en 12 páginas.

@@
 where each $\varphi_i$ ($1 \leq i \leq n$) is either a positive literal, a negative literal or an \fo-formula over $\R$.
-A clause is said to be Horn if it contains at most one positive literal, and a formula is said to be Horn if it is a conjunction of Horn clauses over a relational signature $\R$.
+A clause is said to be Horn if it contains at most one positive literal, and a formula is said to be Horn if it is a conjunction of Horn clauses.
 With this terminology, we define $\uhorn$ as the set of formulae $\psi$ such that $\psi$ is a conjunction of Horn clauses over a relational signature $\R$.

### 2017-01-09 LICS/beyond.tex
Cambios varios.

@@
 %, which is 
-denoted by $\arity(h)$. Then the set of $\fqso$ formulae over a relational signature $\R$ is defined by the following grammar:
+denoted by $\arity(h)$. Then the set of $\fqso$ formulae over a signature $\R$ is defined by the following grammar:
 \begin{multline}
@@
 %Fix a relational signature $\R$. 
-More precisely, the set of $\rqfo(\fo)$ formulae over a relational signature $\R$, where $\rqfo$ stands for recursive $\qfo$, is defined as an extension of $\qfo(\fo)$ that includes the formula 
+More precisely, the set of $\rqfo(\fo)$ formulae over a signature $\R$, where $\rqfo$ stands for recursive $\qfo$, is defined as an extension of $\qfo(\fo)$ that includes the formula 
 %by adding the operator 
@@
 We would like to define an $\rqfo(\fo)$-formula that, given a directed acyclic graph $G$ with $n$ nodes and a pair of nodes $b$, $c$ in $G$, counts the number of paths of length at most $n$ from $b$ to $c$ in $G$.
-To this end, assume that graphs are encoded using the relational signature $\R = \{ E(\cdot,\cdot) \}$, and then define formula $\alpha(x, y, f)$ as follows:
+To this end, assume that graphs are encoded using the signature $\R = \{ E(\cdot,\cdot) \}$, and then define formula $\alpha(x, y, f)$ as follows:
 $$

### 2017-01-09 LICS/counting.tex
Cambios varios.

@@
 
-Fix a relational signature $\R = \{R_1, \ldots, R_k\}$ and assume that $\A$ is an ordered (finite) $\R$-structure with a domain $A = \{a_1, \ldots, a_n\}$.
+Fix a signature $\R = \{R_1, \ldots, R_k\}$ and assume that $\A$ is an ordered (finite) $\R$-structure with a domain $A = \{a_1, \ldots, a_n\}$.
 %$ containing $n$ elements. 
@@
 From now on, we assume that every counting complexity class $\CC$ consists of functions $f : \{0,1\}^* \to \N$ that are invariant under isomorphism of structures.
-More precisely, given a relational signature $\R$, a function $f : \{0,1\}^* \to \N$ is invariant under $\R$-isomorphisms if for every pair $\A, \mathfrak{B} \in \ostr[\R]$ such that $\A$ and
+More precisely, given a signature $\R$, a function $f : \{0,1\}^* \to \N$ is invariant under $\R$-isomorphisms if for every pair $\A, \mathfrak{B} \in \ostr[\R]$ such that $\A$ and
 $\mathfrak{B}$ are isomorphic, it holds that $f(\enc(\A)) = f(\enc(\mathfrak{B}))$. It is worth noticing that this is a usual assumption in descriptive complexity~\cite{L04}.
@@
 \begin{definition} \label{def:cap}
-	Let $\FF$ be a fragment of $\qso$ and $\CC$ a counting complexity class. Then {\em  $\FF$ captures $\CC$ over the class of ordered $\R$-structures} if the  following conditions hold:
+	Let $\FF$ be a fragment of $\qso$ and $\CC$ a counting complexity class. Then {\em  $\FF$ captures $\CC$ over ordered $\R$-structures} if the  following conditions hold:
 	\begin{enumerate}
@@
 	\end{enumerate} 
-	Moreover, {\em $\FF$ captures $\CC$ over ordered structures} if $\FF$ captures $\CC$ over the class of ordered $\R$-structures for every relational signature $\R$.
+	Moreover, {\em $\FF$ captures $\CC$ over ordered structures} if $\FF$ captures~$\CC$ over ordered $\R$-structures for every signature~$\R$.
 \end{definition}

### 2017-01-09 LICS/extentions.tex
Cortes varios y quedamos en 12 páginas.
Cambios varios.

@@
 It is well-known that, under some reasonable complexity-theoretical assumptions, $\shp$ is not closed under subtraction, not even under subtraction by one~\cite{OH93}.
-To overcome this limitation, $\gp$ was introduced in~\cite{fenner1994gap} as the class of functions $f$ for which there exists a polynomial-time NTM $M$ such that $f(x) = \tma_M(x) - \tmr_M(x)$, where  $\tmr_M(x)$ is the number of rejecting runs of $M$ with input $x$.
+To overcome this limitation, $\gp$ was introduced in~\cite{FFK94} as the class of functions $f$ for which there exists a polynomial-time NTM $M$ such that $f(x) = \tma_M(x) - \tmr_M(x)$, where  $\tmr_M(x)$ is the number of rejecting runs of $M$ with input $x$.
 %such that for some $\np$ machine, $f(x)$ is the number of accepting runs minus the number of rejecting runs.
@@
 	Recall the setting of Example~\ref{ex:cliques} and suppose now that we want to compute the number of cliques in a graph that are not triangles.
-Then this can be easily done in $\qsoz$ with the following formula:
-	\[
+Then this can be easily done in $\qsoz$ with the formula:
+	$
 	\alpha_5 :=	\alpha_2 + (-1) \cdot \alpha_1 
-	\]
+	$.
 \end{example}
@@
 In order to capture classes of optimization functions, we can extend $\qso$ with $\max$ and $\min$ quantifiers (called $\optqso$) as follows. 
-Given a relational signature $\R$, the set of $\optqso$-formulae over $\R$ is given by extending the syntax in (\ref{syntax}) with the following operators:
+Given a signature $\R$, the set of $\optqso$-formulae over $\R$ is given by extending the syntax in (\ref{syntax}) with the following operators:
 \begin{multline*}
@@
 \begin{example}\label{ex:optqso}
-	Recall again the setting of Example~\ref{ex:cliques} and suppose now that we want to compute the size of the largest clique in a graph. This can be done in $\optqso$ with the following formula:
+	Recall again the setting of Example~\ref{ex:cliques} and suppose now that we want to compute the size of the largest clique in a graph. This can be done in $\optqso$ as follows:
 	\[

### 2017-01-09 LICS/syntactic-horn.tex
Cambios varios.

@@
 A positive literal is a formula of the form $X(\x)$, where $X$ is a second-order variable and $\x$ is a tuple of first-order variables, and a negative literal is a formula of the form $\exists \v \, \neg X(\u,\v)$, where $\u$ and $\v$ are tuples of first-order variables.
-Given a relational signature $\R$, a clause over $\R$ is a formula of the form $\forall \x \, (\varphi_1 \vee \cdots \vee \varphi_n)$,
-where each $\varphi_i$ ($1 \leq i \leq n$) is either a positive literal, a negative literal or an \fo-formula over $\R$.
+Given a signature $\R$, a clause over $\R$ is a formula of the form $\forall \x \, (\varphi_1 \vee \cdots \vee \varphi_n)$,
+where each $\varphi_i$ ($1 \leq i \leq n$) is either a positive literal, a negative literal or an \fo-formula.
 A clause is said to be Horn if it contains at most one positive literal, and a formula is said to be Horn if it is a conjunction of Horn clauses.
-With this terminology, we define $\uhorn$ as the set of formulae $\psi$ such that $\psi$ is a conjunction of Horn clauses over a relational signature $\R$.
+With this terminology, we define $\uhorn$ as the set of formulae $\psi$ such that $\psi$ is a conjunction of Horn clauses over a signature $\R$.
 

### 2017-01-09 LICS/framework.tex

@@
 where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
-Moreover, if $\R$ is not mentioned, then $\qso$ refers to the set of $\qso$ formulae over all posible relational signatures.
+Moreover, if $\R$ is not mentioned, then $\qso$ refers to the set of $\qso$ formulae over all possible relational signatures.
 %\marcelo{En la gramatica de la logica deberiamos incluir la formula $\top$ que discutimos, la cual representa true y para la cual se tiene que $\sem{\top}(\A, v, V) = 1$. Les parece?}
@@
 A $\qso$-formula $\alpha$ is said to be a \emph{sentence} if it does not have any free variable, that is, every variable in $\alpha$ is under the scope of a usual quantifier or a quantitative quantifier.
-It is important to notice that if $\alpha$ is a $\qso$-sentence over a relational signature $\R$, then for every $\R$-structure $\A$, first-order assignments $v_1$, $v_2$ for $\A$ and second-order assignments $V_1$, $V_2$ for $\A$, it holds that $\sem{\alpha}(\A, v_1, V_1) = \sem{\alpha}(\A, v_2, V_2)$.
+It is important to notice that if $\alpha$ is a $\qso$-sentence over a signature $\R$, then for every $\R$-structure $\A$, first-order assignments $v_1$, $v_2$ for $\A$ and second-order assignments $V_1$, $V_2$ for $\A$, it holds that $\sem{\alpha}(\A, v_1, V_1) = \sem{\alpha}(\A, v_2, V_2)$.
 Thus, in such a case we use the term $\sem{\alpha}(\A)$ to denote $\sem{\alpha}(\A, v, V)$, for some arbitrary first-order assignment $v$ for $\A$ and some arbitrary second-order assignment $V$ for $\A$.

### 2017-01-09 LICS/syntactic-clo.tex

@@
 Therefore, the desired extension has to be achieved by allowing some local extensions to $\Sigma_1$.
-More precisely, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulae over a relational signature $\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mentioned any second-order variable).
+More precisely, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulae over a signature $\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mentioned any second-order variable).
 With this extension we obtain a class with the desired properties.

### 2017-01-09 LICS/syntactic.tex
Acortando el paper.

@@
 The class $\shp$ was introduced in \cite{Valiant79} to prove that computing the permanent of a matrix, as defined in Example \ref{exa-perm}, is a $\shp$-complete problem.
-As a consequence of this result many counting problems have been proved to be $\shp$-complete \cite{V79b,PB83,P86,L86,BW91,HMRS98,DG00,BW05,DS12, PS13,PS14}.
+As a consequence of this result many counting problems have been proved to be $\shp$-complete~\cite{V79b,arora2009computational}.
+%\cite{V79b,PB83,P86,L86,BW91,HMRS98,DG00,BW05,DS12, PS13,PS14}.
 Among them, problems having easy decision counterparts play a fundamental role, as a counting problem with a hard decision version is expected to be hard.
 Formally, the decision problem associated to a function $f : \Sigma^* \to \mathbb{N}$ is defined as $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$, and $f$ is said to have an easy decision version if $L_f \in \ptime$.
-Many prominent examples satisfy this property, like computing the number of: perfect matchings of a bipartite graph ($\cpm$)~\cite{Valiant79}, satisfying assignments of a DNF propositional formula ($\cdnf$)~\cite{DHK05}, satisfying assignments of a Horn  propositional formula ($\chsat$)~\cite{V79b}, among others.
+Many prominent examples satisfy this property, like computing the number of: perfect matchings of a bipartite graph ($\cpm$)~\cite{Valiant79}, satisfying assignments of a DNF ($\cdnf$)~\cite{DHK05} or Horn ($\chsat$)~\cite{V79b} propositional formula, among others.
 

### 2017-01-10 LICS/framework.tex
comentarios y wording

@@
 To define the permanent of a 0-1 matrix in $\qso$, assume that for a binary relation symbol $S$, $\op{permutation}(S)$ is an $\fo$-formula that is true if and only if $S$ is a permutation, that is, a total bijective function (the definition of $\op{permutation}(S)$ is straightforward).
+\martin{abreviar $\op{permutation}$ a $\op{permut}$ ahorra un par de lineas}
 Then the following is a $\qso$-formula defining the permanent of a matrix:
@@
 \]
-Intuitively, the subformula $\beta(s) := \pa{x} (\ex{y} S(x,y) \wedge M(x,y))$ calculates the value $\prod_{i=1}^n A[i, \sigma(i)]$ whenever $S$ encodes a permutation $\sigma$.
+Intuitively, the subformula $\beta(S) := \pa{x} (\ex{y} S(x,y) \wedge M(x,y))$ calculates the value $\prod_{i=1}^n A[i, \sigma(i)]$ whenever $S$ encodes a permutation $\sigma$.
 Moreover, the subformula $\op{permutation}(S) \cdot \beta(S)$ returns $\beta(S)$ when $S$ is a permutation, and returns $0$ otherwise (i.e. $\op{permutation}(S)$ behaves like a filter). 
@@
 for every $\A\in\ostr[\R]$. Then a function $g:\ostr[\R]\to\nat$ is in $\#\LL$ if there exists a formula $\varphi(\bar{x},\bar{X})$ in $\LL$ such that $g = f_{\varphi(\bar x, \bar X)}$.
+\martin{La notacion $f_{\psi(x,X)}$ la inventamos}
 In~\cite{SalujaST95}, they were proved several results about capturing counting complexity classes which are relevant for our work.

### 2017-01-10 LICS/preliminaries.tex
Cambios varios.
comentarios y wording

@@
 Notice that every variable in $\sv$ has an associated arity, which is denoted by $\arity(X)$.
-Then given a relational signature $\R$, the set of second-order logic formulae ($\so$-formulae) over $\R$ is given by the following grammar:
+Then given a  signature $\R$, the set of second-order logic formulae ($\so$-formulae) over $\R$ is given by the following grammar:
 \begin{multline*}\ 
@@
 In particular, we have that $(\A, v, V) \models \top$.
+\martin{En vez de decir "for every $y\in \fv$ distinct from $x$" podría ser "for every $y\in\fv\setminus\{x\}$", y lo mismo con $Y\in\sv$.}
 
@@
 $\exists \bar x_1 \forall \bar x_2 \cdots \exists \bar x_{i-1} \forall \bar x_{i} \, \psi$ (resp., 
-$\forall \bar x_1 \exists \bar x_2 \cdots \forall \bar x_{i-1} \exists \bar x_{i} \, \psi$) if $i$ is even and of the form
+$\forall \bar x_1 \exists \bar x_2 \cdots \forall \bar x_{i-1} \exists \bar x_{i} \, \psi$) if $i$ is even, and of the form
 $\exists \bar x_1 \forall \bar x_2 \cdots \forall \bar x_{i-1} \exists \bar x_{i} \, \psi$ (resp., 
@@
 
-%If $\KK$ is a class of finite structures and $f$ is a function from a relational signature $\R$ to $\bbD$, we denote by $\res{f}{\KK}$ the restriction of $f$ to $\KK$, that is, a function such that: the domain of $\res{f}{\KK}$ is $\{ \enc(\A) \mid \A \in \str[\R] \cap \KK\}$, and
+%If $\KK$ is a class of finite structures and $f$ is a function from a signature $\R$ to $\bbD$, we denote by $\res{f}{\KK}$ the restriction of $f$ to $\KK$, that is, a function such that: the domain of $\res{f}{\KK}$ is $\{ \enc(\A) \mid \A \in \str[\R] \cap \KK\}$, and
 %$\res{f}{\KK}(\enc(\A)) = f(\enc(\A))$ for every $\A \in \str[\R] \cap \KK$.

### 2017-01-10 LICS/appendix-syntactic.tex

@@
 
-\subsection*{Proof of Theorem \ref{sub-pnp}}
-
-Towards a contradiction, suppose that any of the classes $\E{1}$, $\eqso(\loge{1})$, and $\eqso(\logex{1})$ is closed under subtraction.
-
-Let $\R$ include the symbols $S_1, S_2, S_3, S_4$, which describe the following properties.
+%\subsection*{Proof of Theorem \ref{sub-pnp}}
+%
+%Towards a contradiction, suppose that any of the classes $\E{1}$, $\eqso(\loge{1})$, and $\eqso(\logex{1})$ is closed under subtraction.
+%
+%Let $\R$ include the symbols $S_1, S_2, S_3, S_4$, which describe the following properties.
 If a finite $\R$-structure $\A$ defines a 3DNF formula $\Phi$, then its domain is the set of variables mentioned in $\Phi$, and for each $i = 1,2,3,4$:
-\begin{align*}
-	S_1^\A &= \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-	S_2^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-	S_3^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-	S_4^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
-\end{align*}
-Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
-\begin{multline*}
-\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
-\end{multline*}
-Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
-$$
-\varphi(x,X) = (T(x) \vee \neg T(x)).
-$$
-Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
+%\begin{align*}
+%	S_1^\A &= \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
+%	S_2^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
+%	S_3^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
+%	S_4^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
+%\end{align*}
+%Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
+%\begin{multline*}
+%\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
+%\end{multline*}
+%Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
+%$$
+%\varphi(x,X) = (T(x) \vee \neg T(x)).
+%$$
+%Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
 Given that $f_{\#3DNF}, f_{all} \in \E{1} \subseteq \eqso(\loge{1}) \subseteq \eqso(\logex{1}) \subseteq \totp$, and at least one of the classes is closed under subtraction, the function $f_{\#3DNF} - f_{all}$ is in $\totp$.

### 2017-01-10 LICS/preliminaries.tex
sc

@@
 In particular, we have that $(\A, v, V) \models \top$.
-\martin{En vez de decir "for every $y\in \fv$ distinct from $x$" podría ser "for every $y\in\fv\setminus\{x\}$", y lo mismo con $Y\in\sv$.}
+%\martin{En vez de decir "for every $y\in \fv$ distinct from $x$" podría ser "for every $y\in\fv\setminus\{x\}$", y lo mismo con $Y\in\sv$.}
 
-In this paper we consider several fragments of $\so$, in particular first-order logic ($\fo$), least fixed point logic (LFP) and partial fixed point logic (PFP) \cite{L04}.
+In this paper, we consider several fragments of $\so$, in particular first-order logic ($\fo$), least fixed point logic (LFP) and partial fixed point logic (PFP) \cite{L04}.
 Moreover, for every $i \in \N$, we consider the fragment $\Sigma_i$ (resp., $\Pi_i$) of $\fo$, which is the set of $\fo$-formulae of the form
@@
 Finally, we say that a fragment $\L_1$ is contained in a fragment $\L_2$, denoted by $\L_1 \subseteq \L_2$, if for every formula $\varphi$ in $\L_1$, there exists a formula $\psi$ in $\L_2$ such that $\varphi$ is logically equivalent to $\psi$.
+Besides, we say that $\L_1$ is properly contained in $\L_2$, denoted by $\L_1 \subsetneq \L_2$, if $\L_1 \subseteq \L_2$ and $\L_2 \not\subseteq \L_1$.
 
@@
 %We define the following function complexity classes: 
-$\fp$ is the class of functions computable in polynomial time, while $\fpspace$ is the class of functions computable in polynomial space.
-Moreover, $\nfpspace$ is the class of functions computable in polynomial space with output length bounded by a polynomial.
+$\fp$ is the class of functions $f : \Sigma^* \to \N$ computable in polynomial time, while $\fpspace$ is the class of functions $f : \Sigma^* \to \N$ computable in polynomial space. 
+%Moreover, $\nfpspace$ is the class of functions computable in polynomial space with output length bounded by a polynomial. 
 Given a nondeterministic Turing Machine (NTM) $M$, let $\tma_M(x)$ be the number of accepting runs of $M$ with input $x$.

### 2017-01-10 LICS/framework.tex
sc

--- 2017-01-09 LICS/framework.tex
@@
 We introduce here the logical framework that we use for studying counting complexity classes. 
-This framework is based on the framework of Weighted Logics~\cite{DrosteG07} (WL) that has been used in the context of weighted automata for studying functions from words (or trees) to semirings. 
+This framework is based on the framework of Weighted Logics (WL)~\cite{DrosteG07}  that has been used in the context of weighted automata for studying functions from words (or trees) to semirings. 
 We propose here to use the framework of WL over any relational structure and to restrict the semiring to natural numbers. 
-The extension to any relational structure will allow us to study general function complexity classes and the restriction to the natural numbers will simplify the notation in this context (see Section~\ref{sec:previous} for a more detailed discussion).
+The extension to any relational structure will allow us to study general counting complexity classes and the restriction to the natural numbers will simplify the notation in this context (see Section~\ref{sec:previous} for a more detailed discussion).
 
@@
 \begin{multline}
-\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \\ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha \label{syntax} 
+\alpha \ := \ \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \\ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha \label{syntax}
 \end{multline}
@@
 \end{multline*}
-For $\alpha_1$ we encode a triangle as an increasing sequence of nodes $\{x, y, z\}$ in order to count each triangle once. Then the Boolean subformula  $E(x,y) \wedge E(y,z) \wedge E(z,x) \wedge
-x < y \wedge y < z$ is checking the triangle property, by outputting $1$ if $\{x, y, z\}$ forms a triangle in $\fG$ and $0$ otherwise.
+We encode a triangle in $\alpha_1$ as an increasing sequence of nodes $\{x, y, z\}$, in order to count each triangle once. Then the Boolean subformula  $E(x,y) \wedge E(y,z) \wedge E(z,x) \wedge
+x < y \wedge y < z$ is checking the triangle property, by returning $1$ if $\{x, y, z\}$ forms a triangle in $\fG$ and $0$ otherwise.
 Finally, the sum quantifiers in $\alpha_1$ aggregates all the values, counting the number of triangles in $\fG$.
@@
 where $\clique(X) := \forall x \forall y ((X(x) \wedge X(y) \wedge x \neq y)  \rightarrow E(x,y))$.
-Similar than for $\alpha_1$, in the Boolean subformula of $\alpha_2$ we check whether $X$ is a clique and with the sum quantifier we add one for each clique in $\fG$. 
+Similar than for $\alpha_1$, in the Boolean subformula of $\alpha_2$ we check whether $X$ is a clique, and with the sum quantifier we add one for each clique in $\fG$. 
 Note here that, in contrast to $\alpha_1$, in $\alpha_2$ we need a second-order quantifier in the quantitative level.
@@
 \begin{example}\label{exa-perm}
-For a more involved example that includes multiplication, let $\bM = \{M(\cdot,\cdot)\}$ be a vocabulary for storing 0-1 matrices; in particular, a structure $\fM$ over $\bM$ encodes a 0-1 matrix $A$ as follows: if $A[i,j] = 1$, then $M(i,j)$ is true, otherwise $M(i.j)$ is false.
+For a more involved example that includes multiplication, let $\bM = \{M(\cdot,\cdot),<\}$ be a vocabulary for storing 0-1 matrices; in particular, a structure $\fM$ over $\bM$ encodes a 0-1 matrix $A$ as follows: if $A[i,j] = 1$, then $M(i,j)$ is true, otherwise $M(i.j)$ is false.
 Suppose now that we want to compute the permanent of an $n$-by-$n$ 0-1 matrix $A$, that is:
@@
 
-To define the permanent of a 0-1 matrix in $\qso$, assume that for a binary relation symbol $S$, $\op{permutation}(S)$ is an $\fo$-formula that is true if and only if $S$ is a permutation, that is, a total bijective function (the definition of $\op{permutation}(S)$ is straightforward).
+To define the permanent of a 0-1 matrix in $\qso$, assume that for a binary relation symbol $S$, $\op{permut}(S)$ is an $\fo$-formula that is true if and only if $S$ is a permutation, that is, a total bijective function (the definition of $\op{permut}(S)$ is straightforward).
+%\martin{abreviar $\op{permutation}$ a $\op{permut}$ ahorra un par de lineas}
 Then the following is a $\qso$-formula defining the permanent of a matrix:
 \[
-\alpha_3 := \sa{S} \op{permutation}(S) \cdot \pa{x} (\ex{y} S(x,y) \wedge M(x,y)).
+\alpha_3 := \sa{S} \op{permut}(S) \cdot \pa{x} (\ex{y} S(x,y) \wedge M(x,y)).
 \]
-Intuitively, the subformula $\beta(s) := \pa{x} (\ex{y} S(x,y) \wedge M(x,y))$ calculates the value $\prod_{i=1}^n A[i, \sigma(i)]$ whenever $S$ encodes a permutation $\sigma$.
-Moreover, the subformula $\op{permutation}(S) \cdot \beta(S)$ returns $\beta(S)$ when $S$ is a permutation, and returns $0$ otherwise (i.e. $\op{permutation}(S)$ behaves like a filter). 
+Intuitively, the subformula $\beta(S) := \pa{x} (\ex{y} S(x,y) \wedge M(x,y))$ calculates the value $\prod_{i=1}^n A[i, \sigma(i)]$ whenever $S$ encodes a permutation $\sigma$.
+Moreover, the subformula $\op{permut}(S) \cdot \beta(S)$ returns $\beta(S)$ when $S$ is a permutation, and returns $0$ otherwise (i.e. $\op{permut}(S)$ behaves like a filter). 
 Finally, the second order sum aggregates these values iterating over all binary relations and calculating the permanent of the matrix.
@@
 
-In this subsection, we discuss some previous frameworks proposed in the literature and how they differ from our approach.
+In this section, we discuss some previous frameworks proposed in the literature and how they differ from our approach.
 We start by discussing the connection between $\qso$ and weighted logics (WL)~\cite{DrosteG07}. 
@@
 The main difference is that we restrict the semiring used in WL to natural numbers in order to study counting complexity classes.
-Another difference of WL with our approach is that, to the best of our knowledge, this is the first paper to study weighted logics over general relational signatures to do descriptive complexity for counting complexity classes.
+Another difference of WL with our approach is that, to the best of our knowledge, this is the first paper to study weighted logics over general relational signatures, in order  to do descriptive complexity for counting complexity classes.
 Previous works on WL usually restrict the signature of the logic to strings, trees, and other specific structures (see \cite{droste2009handbook} for more examples), and they did not study the logic over general structures.
@@
 for every $\A\in\ostr[\R]$. Then a function $g:\ostr[\R]\to\nat$ is in $\#\LL$ if there exists a formula $\varphi(\bar{x},\bar{X})$ in $\LL$ such that $g = f_{\varphi(\bar x, \bar X)}$.
+%\martin{La notacion $f_{\psi(x,X)}$ la inventamos}
 In~\cite{SalujaST95}, they were proved several results about capturing counting complexity classes which are relevant for our work.
@@
 Notice that for every formula $\varphi(\bar{x},\bar{X})$, it holds that $f_{\varphi(\bar{x},\bar{X})}$ is the same function as $\sem{\sa{\bar{X}} \sa{\bar{x}} \varphi(\bar{x},\bar{X})}$, that is, the approach in \cite{SalujaST95} can be seen as a syntactical restriction of our approach based on $\qso$.
-Thus, the advantage of our approach relies on the flexibility for defining  functions by alternating sum with product operators and, moreover, by introducing new quantitative operators (see Section~\ref{sec:beyond}).
+Thus, the advantage of our approach relies on the flexibility to define functions by alternating sum with product operators and, moreover, by introducing new quantitative operators (see Section~\ref{sec:beyond}).
 Furthermore, we show in the following section how to capture some fundamental classes that cannot be captured by following the approach in~\cite{SalujaST95}.

### 2017-01-10 LICS/counting.tex
sc

@@
 Then let $\enc(R_i^\A)$ be a binary string of length $n^\ell$ such that the $i$-th bit of $\enc(R_i^\A)$ is 1 if the $i$-th tuple in the previous enumeration belongs to $R_i^\A$, and 0 otherwise.
-Moreover, define the encoding of $\A$, denoted by $\enc(\A)$, as the
+Moreover, define the encoding of $\A$, denoted by $\enc(\A)$, as the string
 %following binary string~\cite{L04}:
@@
 %\end{eqnarray*}
-$\enc(\A) =  0^n \, 1 \, \enc(R_1^\A) \, \cdots \, \enc(R_k^\A)$~\cite{L04}. 
+$0^n \, 1 \, \enc(R_1^\A) \, \cdots \, \enc(R_k^\A)$~\cite{L04}. 
 %We define the class of all $\R$-functions, denoted by $\Func(\R)$, as the class of all functions $f: \ostr \rightarrow \bbN$.
@@
 For answering this question, it is reasonable to start with $\shp$, a well-known and widely-studied counting complexity class~\cite{arora2009computational}. 
-Since $\shp$ has a strong similarity with $\np$, one could expect a ``Fagin-like'' Theorem~\cite{fagin1974generalized} for this class. 
+Since $\shp$ has a strong similarity with $\np$, one could expect a ``Fagin-like'' Theorem~\cite{F75} for this class. 
 Actually, in~\cite{SalujaST95} it was shown that the class $\sfo$ captures $\shp$.
@@
 %(e.g. $2^{n}$), so
-Hence, it is not  clear how to capture
+Thus, it is not  clear how to capture
 %impossible 
@@
 %built in the ouptape. 
-It is important to notice that the alternation between sum and products quantifiers is crucial for this reconstructions and, thus, crucial for capturing $\fp$.
+It is important to notice that the alternation between sum and product quantifiers is crucial for this reconstructions and, thus, crucial for capturing $\fp$.
 %That is, without these operators is difficult to capture $\fp$.
@@
 
-The results of this section validate $\qso$ as an appropriate logical framework for extending the theory of Descriptive Complexity to counting complexity classes.
+The results of this section validate $\qso$ as an appropriate logical framework for extending the theory of descriptive complexity to counting complexity classes.
 In the following sections, we provide more arguments for this claim, by considering some fragments of $\eqso$ and, moreover, by showing how to go beyond $\eqso$ to capture other

### 2017-01-10 LICS/extentions.tex
sc

@@
 \begin{example}
-	Recall the setting of Example~\ref{ex:cliques} and suppose now that we want to compute the number of cliques in a graph that are not triangles.
-Then this can be easily done in $\qsoz$ with the formula:
+	Recall the setting of Example~\ref{ex:cliques} and suppose now that we want to compute the number of cliques in a graph that are not triangles. This can be easily done in $\qsoz$ with the formula:
 	$
@@
 For instance, MAX-SAT is in $\optp$ as several others optimization versions of $\np$-problems.
-Given that in~\cite{krentel1988complexity} the author does not make the distinction between $\max$ or $\min$, in~\cite{vollmer1995complexity} the authors define the classes $\maxp$ and $\minp$ as the max and min version of the problems in $\optp$ (i.e.
+Given that in~\cite{krentel1988complexity} the author does not make the distinction between $\max$ and $\min$, in~\cite{vollmer1995complexity} the authors define the classes $\maxp$ and $\minp$ as the max and min version of the problems in $\optp$ (i.e.
 $\optp = \maxp \cup \minp$).
@@
 	\]
-	Notice that formula $\sa{z} X(z)$ is used to compute the number of nodes in a clique $X$. 
+	Notice that formula $\sa{z} X(z)$ is used to compute the number of nodes in a set $X$. 
 \end{example}
@@
 \begin{theorem} \label{theo:capture-optp}
-	$\maxqso(\fo)$ and $\minqso(\fo)$ captures $\maxp$ and $\minp$, respectively, over ordered structures.
+	$\maxqso(\fo)$ and $\minqso(\fo)$ capture $\maxp$ and $\minp$, respectively, over ordered structures.
 \end{theorem}

### 2017-01-10 LICS/extentions.tex
sc

@@
 each run produces a binary string which is interpreted as a number).
-For instance, MAX-SAT is in $\optp$ as several others optimization versions of $\np$-problems.
+For instance, MAX-SAT is in $\optp$ as many other optimization versions of $\np$-problems.
 Given that in~\cite{krentel1988complexity} the author does not make the distinction between $\max$ and $\min$, in~\cite{vollmer1995complexity} the authors define the classes $\maxp$ and $\minp$ as the max and min version of the problems in $\optp$ (i.e.
@@
 
-In order to capture classes of optimization functions, we can extend $\qso$ with $\max$ and $\min$ quantifiers (called $\optqso$) as follows. 
+In order to capture classes of optimization functions, we extend as follows $\qso$ with $\max$ and $\min$ quantifiers (called $\optqso$). 
 Given a signature $\R$, the set of $\optqso$-formulae over $\R$ is given by extending the syntax in (\ref{syntax}) with the following operators:

### 2017-01-10 LICS/framework.tex
sc

@@
 
-Suppose now that we want to count the number of cliques in~$\fG$ (i.e. set of nodes where every pair of nodes is connected). We can easily define this function with the following formula:
+Suppose now that we want to count the number of cliques in~$\fG$.
+%(i.e. set of nodes where every pair of nodes is connected). 
+We can 
+%easily 
+define this function with the following formula:
 $$
-\alpha_2 \ := \ \sa{X} \clique(X) ,
+\alpha_2 \ := \ \sa{X} \clique(X),
 $$ 
 where $\clique(X) := \forall x \forall y ((X(x) \wedge X(y) \wedge x \neq y)  \rightarrow E(x,y))$.
-Similar than for $\alpha_1$, in the Boolean subformula of $\alpha_2$ we check whether $X$ is a clique, and with the sum quantifier we add one for each clique in $\fG$. 
-Note here that, in contrast to $\alpha_1$, in $\alpha_2$ we need a second-order quantifier in the quantitative level.
-This is according to the inherent complexity of evaluating each formula: $\alpha_1$ defines an $\fp$-function where $\alpha_2$ defines a $\shp$-complete function.
+%Similar than for $\alpha_1$, 
+In the Boolean sub-formula of $\alpha_2$ we check whether $X$ is a clique, and with the sum quantifier we add one for each clique in $\fG$. 
+%Note here that, 
+But in contrast to $\alpha_1$, 
+in $\alpha_2$ we need a second-order quantifier in the quantitative level.
+This is according to the
+%inherent 
+complexity of evaluating each formula:
+%each formula: 
+$\alpha_1$ defines an $\fp$-function while $\alpha_2$ defines a $\shp$-complete function.
 %~\cite{paper-that-shows-that-this-problem-is-sharpP-hard}.

### 2017-01-10 LICS/syntactic-clo.tex
sc

@@
 It is important to notice that $\shp$ is believed not to be closed under subtraction by one by some complexity-theoretical assumption.\footnote{A decision problem $L$ is in the randomized complexity class $\cspp$ if there exists a polynomial-time NTM $M$ such that for every $x \in L$ it holds that $\tma_M(x) - \tmr_M(x) = 2$, and for every $x \not\in L$ it holds that $\tma_M(x) = \tmr_M(x)$ \cite{OH93,FFK94}.
-It is believed that $\np \not\subseteq \cspp$, as there is no evidence that a TM with the property just described exists for the propositional satisfiability problem.
+It is believed that $\np \not\subseteq \cspp$.
+%, as there is no evidence that a TM with the property just described exists for the propositional satisfiability problem. 
 However, if $\shp$ is closed under subtraction by one, then it holds that $\np \subseteq \cspp$ \cite{OH93}.} So, the following proposition rules out any logic that extends $\Pi_1$ for a possible extension of~$\QE{1}$ with the desired closure property.
@@
 \end{theorem}
+The proof that $\eqso(\logex{1})$ is closed under subtraction by one is the most involved of the paper.
+We think the main technique used in this proof, which is based on considering some witnesses of logarithmic size, is of independent interest.
 

### 2017-01-10 LICS/syntactic-horn.tex
sc

@@
 \begin{example} \label{ex-hornsat-esop1}
-Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot)\}$.
+Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot),<\}$.
 This vocabulary is used as follows to encode a Horn formula.
@@
 In particular, $\pT(x)$ holds if and only if $x$ is a propositional variable that is assigned value true.
-More specifically, $\varphi(\pT)$ is defined as follows:
+More specifically,
+%$\varphi(\pT)$ is defined as follows:
 \begin{align*}
@@
 \end{example}
-The previous example suggests that in order to define $\chsat$, we can use Horn formulae defined as follows. 
+The previous example suggests that to define $\chsat$, we can use Horn formulae defined as follows. 
 A positive literal is a formula of the form $X(\x)$, where $X$ is a second-order variable and $\x$ is a tuple of first-order variables, and a negative literal is a formula of the form $\exists \v \, \neg X(\u,\v)$, where $\u$ and $\v$ are tuples of first-order variables.
 Given a signature $\R$, a clause over $\R$ is a formula of the form $\forall \x \, (\varphi_1 \vee \cdots \vee \varphi_n)$,
-where each $\varphi_i$ ($1 \leq i \leq n$) is either a positive literal, a negative literal or an \fo-formula.
+where each $\varphi_i$ ($1 \leq i \leq n$) is either a positive literal, a negative literal or an \fo-formula over $\R$.
 A clause is said to be Horn if it contains at most one positive literal, and a formula is said to be Horn if it is a conjunction of Horn clauses.
-With this terminology, we define $\uhorn$ as the set of formulae $\psi$ such that $\psi$ is a conjunction of Horn clauses over a signature $\R$.
+With this terminology, we define $\uhorn$ as the set of formulae $\psi$ such that $\psi$ is a
+%conjunction of 
+Horn formula 
+%clauses 
+over a signature $\R$.
 
 We have that $\chsat \in \eqso(\uhorn)$. Moreover, one can show that $\eqso(\uhorn)$ forms a class of functions with easy decision counterparts, namely, $\eqso(\uhorn) \subseteq \totp$.
-Thus, $\eqso(\uhorn)$ is a new alternative in our search for functions in $\shp$ with easy decision counterparts and natural complete problems.
+Thus, $\eqso(\uhorn)$ is a new alternative in our search for a class of functions in $\shp$ with easy decision counterparts and natural complete problems.
 Moreover, an even larger class for our search can be generated by extending the definition of $\uhorn$ with outermost existential quantification.
-More precisely, a formula $\varphi$ is in $\ehorn$ if $\varphi$ is of the form $\exists \bar x \, \psi$ with $\psi$ a Horn formula.
+%More precisely, 
+Formally, a formula $\varphi$ is in $\ehorn$ if $\varphi$ is of the form $\exists \bar x \, \psi$ with $\psi$ a Horn formula.
 

### 2017-01-10 LICS/syntactic.tex
sc

@@
 In this section, we use the framework developed in this paper to address the problem of defining a robust class of functions with easy decision versions.
-More specifically, we use the framework to introduce in Section \ref{sec-hier-shp} a syntactic hierarchy of function complexity classes contained in $\shp$.
+More specifically, we use the framework to introduce in Section \ref{sec-hier-shp} a syntactic hierarchy of counting complexity classes contained in $\shp$.
 Then this hierarchy is used in Section \ref{sec-clo} to define a class of functions with easy decision versions and good closure properties, and in Section \ref{sec-horn} to define a class of functions with easy decision versions and

### 2017-01-10 LICS/beyond.tex
sc

@@
 
-In the previous sections, we use weighted logics to give a framework for descriptive complexity of counting complexity classes. Here, we go beyond weighted logics 
+%In the previous sections, 
+We have used weighted logics to give a framework for descriptive complexity of counting complexity classes. Here, we go beyond weighted logics 
 %to add unexplored quantifiers at the quantitative level. Specifically, we 
@@
 %(boolean) 
-the notion of least fixed point of LFP, which is denoted as \emph{least s-fixed point}.
+the notion of least fixed point of LFP.
+%, which is denoted as \emph{least s-fixed point}.
 % and, 
@@
 We would like to define an $\rqfo(\fo)$-formula that, given a directed acyclic graph $G$ with $n$ nodes and a pair of nodes $b$, $c$ in $G$, counts the number of paths of length at most $n$ from $b$ to $c$ in $G$.
-To this end, assume that graphs are encoded using the signature $\R = \{ E(\cdot,\cdot) \}$, and then define formula $\alpha(x, y, f)$ as follows:
+To this end, assume that graphs are encoded using the signature $\R = \{ E(\cdot,\cdot), < \}$, and then define formula $\alpha(x, y, f)$ as follows:
 $$
@@
 Thus, if we assume that $\qfo(\tc)$ captures $\shl$ (over ordered structures), then we have that $\chi_L \in \nlog$ for every $L \in \nlog$.
-This implies that $\nlog = \ulog$,
-\footnote{A decision language $L$ is in $\ulog$ is there exists a logarithmic-space NTM $M$ accepting $L$ and satisfying that $\tma_M(x) = 1$ for every $x \in L$.} 
+This implies that $\nlog = \ulog$,\footnote{A decision language $L$ is in $\ulog$ is there exists a logarithmic-space NTM $M$ accepting $L$ and satisfying that $\tma_M(x) = 1$ for every $x \in L$.}
 and thus contradicts the widely believed assumption that $\ulog \subsetneq \nlog$.

### 2017-01-11 LICS/appendix-syntactic.tex

@@
 
-Let $\LL$ be a fragment {\bf closed under conjunction}.
+\begin{lemma} \label{conj-mult}
+If $\LL$ is a fragment closed under conjunction, then $\eqso(\LL)$ is closed under binary multiplication.
+\end{lemma}
+\begin{proof}
 We will define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ extended by binary product, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
 In fact, the formula $\tau(\alpha)$ is in $\LL$-SNF.
-First we replace each constant $s$ in $\alpha$ for $(\top \add \cdot \add \top)$ ($s$ times).
+First we replace each constant $s$ in $\alpha$ for $(\top \add \cdots \add \top)$ ($s$ times).
 If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$.
@@
 This covers all possible cases for $\alpha$.
-
 For every pair of formulas $\alpha,\beta$ in $\eqso(\LL)$, we have that their multiplication $(\alpha\cdot\beta)$ is a formula in the grammar $\eqso(\LL)$ extended by binary product, and so, there exists an equivalent formula $\tau(\alpha\cdot\beta)$ which is in unextended $\eqso(\LL)$.
-Since $\loge{1}$ is closed under conjunction, this also holds for $\eqso(\loge{1})$.
-This concludes the proof.
+\end{proof}
+Since $\loge{1}$ is closed under conjunction, this also holds for $\eqso(\loge{1})$. This concludes the proof.
 
@@
 
-We divide the proof in two parts.
+{\bf Part 1.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
+This can be seen because $\eqso(\LL)$ is closed under sum for every fragment $\LL$ by definition, and since $\logex{1}$ is closed under conjunction, from Lemma \ref{conj-mult} it follows that $\eqso(\logex{1})$ is closed under multiplication.
 
 \vspace{1em}
-First, note that for every fragment $\LL$ of $\so$ the class $\eqso(\LL)$ is closed under sum by definition, and if $\LL$ is closed under conjunction, like $\logex{1}$ is, $\eqso(\LL)$ is closed under product.
-We are only left to prove that $\eqso(\logex{1})$ is closed under subtraction by one.
+{\bf Part 2.} We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
+We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Theorem \ref{prop:qe0-fp-qe1-totp-fptras}.
+First, we define a function that converts a formula $\alpha$ in $\eqso(\logex{1})$ over a signature $\R$ into a formula $\lambda(\alpha)$ in $\eqso(\loge{1})$ over a signature $\R_{\alpha}$.
+Afterwards, we define a function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an $\R_{\alpha}$-structure $g_{\alpha}(\A)$.
 
+Let $\alpha$ be in $\eqso(\logex{0})$.
+The signature $\R_{\alpha}$ is obtained by adding the symbol $R_{\psi}$, for every $\fo$ formula $\psi(\bar{z})$ in $\alpha$, to $\R$.
+Each symbol $R_{\psi}$ represents a predicate with arity $\length(\bar{z})$.
+Then, $\lambda(\alpha)$ is defined as $\alpha$ where each $\fo$ formula $\psi(\bar{z})$ has been replaced by $R_{\psi}(\bar{z})$.
+We now define the function $g_{\alpha}$ procedurally.
+Let $\A$ be a $\R$-structure with domain $A$.
+Let $\A'$ be an $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A}$ empty.
+For each $\fo$-formula $\psi(\z)$ with $\length(\bar{z})$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length(\bar{z})}$.
+If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$.
+This concludes the construction of $\A'$.
+Note that the number of $\fo$ subformulas and each arity and tuple size is fixed to $\alpha$, so computing this function takes polynomial time over the size of the structure.
+Moreover, the encoding of $\A'$ has polynomial size over the size of $\enc(\A)$.
+We define $g_{\alpha}(\A) = \A'$ and we have that for each $\R$-structure $\A$: $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$.
+Therefore, we have a parsimonious reduction from $\alpha$ to the $\eqso(\loge{1})$ formula $\lambda(\alpha)$.
+
+To show that $\alpha$ is in $\totp$, consider a procedure that takes an input $\enc(\A)$, converts it to $\enc(g_{\alpha}(\A))$ and simulates the $\totp$ procedure for $\lambda(\alpha)$ on input $\enc(g_{\alpha}(\A))$.
+This procedure generates exactly $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$ branches, and therefore $\alpha$ is in $\totp$.
+
+To show that $\alpha$ has an FPRAS, consider a procedure that takes an input $\enc(\A)$, converts it to $\enc(g_{\alpha}(\A))$ and simulates the FPRAS for $\lambda(\alpha)$.
+This procedure also takes polynomial time and satisfies the condition.
+
+\vspace{1em}
+{\bf Part 3.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one.
 Let $\alpha$ be a $\eqso(\logex{1})$ formula over a signature $\R$.
@@
 We suppose that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
-Moreover, we assume that $\length(\bar{x}) > 0$ since we can replace each $\varphi_i$ for the equivalent formula $\sa{y}\varphi_i\wedge\min(y)$.
+Moreover, we assume that $\length(\bar{x}) > 0$ and $\length(\bar{X}) > 0$ since we can replace each $\varphi_i$ for the equivalent formula $\sa{Y}\sa{y}\varphi_i\wedge\first(y)\wedge\forall z Y(z)$.
 
-The intuition behind our reasoning is separated in three points: (1) For each $\beta_i$ of the form $\sa{\bar{x}}\varphi$, the formula $\kappa(\beta_i)$ will count every tuple $\bar{x}$ that satisfies $\varphi$ except for the lexicographically smallest one.
-(2) For each $\beta_i$ of the form $\sa{\bar{X}}\sa{\bar{x}}\varphi$, the formula $\kappa(\beta_i)$ will isolate the smallest $\bar{X}$ that satisfies $\varphi$, and exclude the lexicographically smallest tuple $\bar{x}$ that satisfies $\varphi(\bar{X})$.
-And (3) if $\alpha = (\beta\add\sa{\bar{x}}\varphi)$ or $\alpha = (\beta\add\sa{\bar{X}}\sa{\bar{x}}\varphi)$, the formula $\kappa(\alpha)$ will exclude the lexicographically smallest tuple that satisfies $\varphi$ if and only if $\beta$ is equal to 0.
+The proof will be separated in two parts.
+(i) If $\alpha$ is in $\logex{1}$-PNF, this is, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$, then we define a formula $\varphi'$ that satisfies the following condition.
+For each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
+We define $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$.
+(ii) If $\alpha = \beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$, then we define a formula $\varphi'$ that satisfies the following condition.
+For each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$ {\bf and} $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
+We define $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$.
 
@@
 
-For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha) = \sa{\bar{x}}\exists\bar{y}[\varphi(\bar{x},\bar{y})\wedge\exists\bar{z}(\varphi(\bar{z},\bar{y})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfils the desired condition.
+{\em Section (i).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ where $\varphi$ is quantifier-free with the addition of atomic formulas in $\fo$.
+We call this syntax {\em extended quantifier-free} from now on.
+Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
+We convert $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m\varphi_i$ where each $\varphi_i$ is of the form:
+$$
+\varphi_i(\bar{X},\bar{x},\bar{y}) = \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge
+\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}} \wedge \underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}}.
+$$
+This can be done with a standard predicate-DNF transformation algorithm.
+We now define a series of transformations to $\varphi$ that will make each formula $\varphi$ satisfy the following three conditions.
+(a) Each first-order variable in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ is mentioned at most once.
+Moreover, in this formula no variable in $\bar{x}$ is mentioned.
+(b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
+We will define then what we mean by an ordered partition.
+(c) If $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
+\begin{itemize}
+	\item[(a)] Consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
+We add $\length(\bar{w})$ new variables $z_1,\ldots,z_{\length(\bar{w})}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length(\bar{w})})$.
+We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$.
+The formula $\varphi_i$ is now defined as
+	$$
+	\varphi_i(\bar{X},\bar{x},(\bar{y},\bar{z})) = (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{-}(\bar{X},\bar{x},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y}).
+	$$
+	We repeat this process for each instance of a $X_j(\bar{w})$ in $\varphi_i$, and we obtain a formula where none of the $X_j$'s acts over any variable in $\bar{x}$.
+We add the new first-order variables to $\bar{y}$ and we redefine $\varphi_i$ as
+	$$
+	\varphi_i(\bar{X},\bar{x},\bar{y}) = \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
+	$$
+	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
+	\item[(b)] An ordered partition on a set $S$ is defined by an equivalence relation $\sim$ over $S$, and a linear order over $S/\!\sim$.
+For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$.
+A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
+On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
+	For a given $k$, let $\cB_k$ be the number of possible ordered partitions for a set of size $k$. For $1 \leq j \leq \cB_{\length(\bar{x},\bar{y})}$ 
+	let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
+
+	The formula $\varphi(\bar{X},\bar{x},\bar{y})$ is then redefined as
+	$$
+	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}}[\theta^j(\bar{x},\bar{y})\wedge \varphi_i(\bar{X},\bar{x},\bar{y})],
+	$$
+	and each of these disjuncts is a $\varphi_i$. Therefore, each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
+	\item[(c)] For each $\varphi_i$, we do the following.
+If there exists an instance of $X_j(\bar{z})$ in $\varphi^{+}_i$, an instance of $\neg X_j(\bar{w})$ in $\varphi^{-}_i$ and the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$, then the entire formula $\varphi_i$ is removed from $\varphi$.
+\end{itemize}
+It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still an extended quantifier-free formula.
+From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y}) = \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c).
+Note the following.
+\begin{claim}
+For a given ordered structure $\A$, if there exists a first-order assignment $v$ for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists a second-order assignment $V$ for $\A$ such that  $(\A,V,v)\models\varphi_i(\bar{X},\bar{x},\bar{y})$.
+\end{claim}
+\begin{proof}
+Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
+Define $\bar{B} = (B_1,\ldots,B_{\length(\bar{X})})$ as $B_j = \{v(\bar{w})\mid\bar{w}\text{ is a subtuple of $\bar{y}$ and $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
+Towards a contradiction, suppose that $(\A,V,v)\not\models\varphi_i(\bar{X},\bar{x},\bar{y})$.
+By the choice of $v$, and construction of $V$ it is clear that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})\wedge\varphi^{+}_i(\bar{X},\bar{x},\bar{y})$, so we necessarily have that $(\A,V,v)\not\models\varphi^{-}_i(\bar{X},\bar{x},\bar{y})$.
+Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,v)\not\models\neg B_t(\bar{w})$.
+This is, $v(\bar{w})\in B_t$, but because of the way we constructed $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$.
+Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$.
+However, this violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
+\end{proof}
+We generalize the construction of $\bar{B}$ given in the claim.
+For a structure $\A$ and a first-order assignment $v$ for $\A$, then we define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length(\bar{X})})$ where each $B^v_j = \{v(\bar{w})\mid\bar{w}\text{ is a subtuple of $\bar{y}$ and $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
+Now, let $\bar{v}$ be the subtuple of $\bar{y}$ of every first-order variable mentioned in $\varphi_i^{+}$ and let $\bar{u}$ be a subtuple of $(\bar{x},\bar{y})$ such that concatenating $(\bar{v},\bar{u})$ is equal to $(\bar{x},\bar{y})$.
+We also define $v^*$ as a first-order assignment for $\A$ for which we have that $v^{*}(\bar{v})$ is the lexicographically smallest tuple that satisfies $(\A,v^{*})\models\exists\bar{u} \varphi^{\fo}_i(\bar{u},\bar{v})$.
+Lastly, we define $\bar{B}^* = \bar{B}^{v^{*}}$ and $V^*$ as some second-order assignment for $\A$ for which $V^*(\bar{X}) = \bar{B}^*$.
+
+We define a formula $\mu_i$ that is satisfied only by $v^*$:
+$$
+\mu_i(\bar{v}) = \exists\bar{u}\,\varphi_i^{\fo}(\bar{u},\bar{v})\wedge\forall\bar{u}'\forall\bar{v}'(\varphi_i^{\fo}(\bar{u}',\bar{v}')\to\bar{v}\leq\bar{v}').
+$$ 
+The following is a pivotal formula in our proof:
+\begin{align}
+\psi_i(\bar{X},\bar{x}) = &\exists\bar{v}\bigg(\overbrace{\mu_i(\bar{v})}^{(\blacktriangleleft)}\wedge\Big( \overbrace{\neg\varphi^{+}_i(\bar{X},\bar{v})}^{(\blacktriangle)} \vee \overbrace{\bigvee_{X \in \bar{X}} \exists\bar{z}(X(\bar{z}) \wedge \bigwedge\limits_{\substack{\text{subtuples }\bar{w} \text{ of }\bar{v} \text{ s.t.}\\ X(\bar{w})\text{ is mentioned} \\ \text{in }\varphi^{+}_i(\bar{X},\bar{v})}}\bar{w}\neq\bar{z}}^{(\blacktriangledown)})\Big)\bigg) \vee \label{minusone-1}\\ &\exists\bar{x}'(\exists\bar{y}'\varphi_i(\bar{X},\bar{x}',\bar{y}')\wedge \bar{x}'<\bar{x})\vee \label{minusone-2} \\
+&\neg\exists\bar{x}\exists\bar{y}\varphi^{\fo}_i(\bar{x},\bar{y}). \label{minusone-3}
+\end{align}
+Consider some $\A$.
+In detail, the formula \eqref{minusone-1} represents the following: In $(\blacktriangleleft)$, we make it so the only value for $\bar{v}$ we consider is $v^{*}(\bar{v})$.
+In $(\blacktriangle)$, we mention that the assignment for $\bar{X}$ satisfies the formula if does not contain at least one of the tuples in $v^{*}(\bar{v})$.
+In $(\blacktriangledown)$, we mention that in the assignment for $\bar{X}$ also satisfies the formula if it contains at least one tuple, in one of the $X$'s, that is not equal to the one in $v^{*}(\bar{v})$.
+Then, in the formula \eqref{minusone-2} we mention that the assignment for $\bar{x}$ satisfies the formula if it is not the lexicographically smallest assignment that satisfies $\varphi_i$ for that particular assignment for $\bar{X}$.
+Lastly, in \eqref{minusone-3}, we mention that the assignment for $(\bar{X},\bar{x})$ satisfies the formula if there are no assignments $V,v$ for $\A$ such that $(\A,V,v)\not\models \exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$.
+
+This way, the formula $\psi(\bar{X},\bar{x})$ functions as a sort of filter for valid assignments for $\bar{X}$ and $\bar{x}$.
+For each structure $\A$ we have that the only $V,v$ for which $(\A,V,v)\models\exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$ and also $(\A,V,v)\not\models\psi_i(\bar{X},\bar{x})$ is the one given by $v = v^*$ and $V = V^*$, should there exist one.
+%We prove this formally.
+%\begin{lemma}
+%	For a given ordered structure $\A$ such that $\A\models\exists\bar{X}\exists\bar{x}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, there is an assignment $(\bar{B},\bar{b})$ to $(\bar{X},\bar{x})$ that satisfies the following conditions (1) $\A\models\exists\bar{y}\,\varphi_i(\bar{B},\bar{b},\bar{y})$, (2) $\A\not\models\psi_i(\bar{B},\bar{b})$ and (3) this is the only assignment to $(\bar{X},\bar{x})$ that satisfies (1) and (2).
+%\end{lemma}
+%\begin{proof}
+%	Excusing the nested proof, we have a claim that identifies a crucial condition that is verifiable with a $\fo$ formula. 
+%	
+%	We give the proof for the base theorem.
+Let $\A$ be an ordered structure with domain $A$.
+We clearly have that $\A\models\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$, so let $\bar{a}\in A^{\length(\bar{y})}$ and $\bar{b}\in A^{\length(\bar{x})}$ such that $\A\models\mu_i(\bar{b},\bar{a})$.
+We use the same construction in the claim and we obtain the assignment $(\bar{B},\bar{b},\bar{a})$ which satisfies the three conditions.
+%\end{proof}
+
+We define $\chi_i = \neg\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$. Now recall that $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$. We define:
+$$
+\varphi'_1(\bar{X},\bar{x},\bar{y}) = \varphi_1(\bar{X},\bar{x},\bar{y})\wedge\psi_1(\bar{X},\bar{x}).
+$$
+This models that all the assignments that satisfy $\varphi_1$ also satisfy $\varphi'_1$, except for $(V^*,v^*)$. Then we define:
+$$
+\varphi'_2(\bar{X},\bar{x},\bar{y}) = \varphi_2(\bar{X},\bar{x},\bar{y})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
+$$
+This models that all the assignments that satisfy $\varphi_2$ also satisfy $\varphi'_2$, but if there was no assignment for $\varphi_1$, then we exclude $V^*,v^*$ for $\varphi_2$.
+Then we generalize this construction for each $\varphi_i$:
+$$
+\varphi_i'(\bar{X},\bar{x},\bar{y}) = \varphi_i(\bar{X},\bar{x},\bar{y})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_2(\bar{X},\bar{x}))\wedge\cdots\wedge(
+\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})),
+$$
+and lastly, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x},\bar{y})$, which satisfies our condition that $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.
 
@@
 
-For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha)$ procedurally as follows: Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
-We suppose that $\varphi$ is in a DNF form that leaves formulas that do not mention $\bar{X}$ intact.
-If it is not, we convert $\varphi$ to this form with a standard transformation algorithm.
-Let $\varphi(\bar{X},\bar{x},\bar{y}) = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$ where each $\varphi_i$ has the form:
-$$
-\varphi_i(\bar{X},\bar{x},\bar{y}) = (\text{conjunction of $X_j$'s}) \wedge (\text{conjunction of $\neg X_j$'s})  \wedge (\text{an $\fo$ formula that does not mention any $X_j$}).
-$$
-Define $\varphi_i^{+}$, $\varphi_i^{-}$ and $\varphi_i^{\fo}$ as the formulas mentioned above.
+{\em Section (ii).} Let $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y}))$ for some algebraic formula $\beta$ and some extended quantifier-free formula $\varphi$.
+We define $\kappa(\alpha)$ as follows: First, perform the same transformations to $\varphi(\bar{X},\bar{x},\bar{y})$ as in the previous section.
+Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c).
+We also use the previously defined formulas $\chi_i$ and $\psi_i$, for each $i$.
 
-In our procedure we assume that in $\varphi_i^{+}\wedge\varphi_i^{-}$ every first-order variable from $\bar{x},\bar{y}$ is mentioned at most once.
-If not, we add new first-order variables from $\fv$ to $\bar{y}$ so that no variable is repeated, replace them in $\varphi_i^{+}\wedge\varphi_i^{-}$ and add their respective equalities in $\varphi_i^{\fo}$ so that the new formula is equivalent.
-For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
-
-Furthermore, we assume that $\varphi_i^{\fo}$ defines an ordered partition of the variables in the tuple $(\bar{x},\bar{y})$.
-For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$.
-A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
-On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
-For a given $k$, we define $\cB_k$ as the number of possible ordered partitions for a set of size $k$. 
-If $\varphi_i^{\fo}$ does not define an ordered partition, then for $1 \leq j \leq \cB_{\length(\bar{x},\bar{y})}$ 
-let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
-The formula $\varphi_i(\bar{X},\bar{x},\bar{y})$ is converted into $\bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}}\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \theta^j(\bar{x},\bar{y})$.
-Consider each of these disjuncts as a new $\varphi_i$.
-
-After the previous assumptions, we do the following: for each $X_j\in \bar{X}$ we check every instance of $X_j(\bar{w})$ in $\varphi^{+}_i$ and every instance of $X_j(\bar{z})$ in $\varphi^{-}_i$, where $\bar{w}$ and $\bar{z}$ are subtuples of $(\bar{x},\bar{y})$.
-If the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{w} = \bar{z}$, the entire formula $\varphi_i$ is removed from $\varphi$.
-
-Now let $\bar{v}$ be the tuple of every first-order variable mentioned in $\varphi_i^{+}$ and let $\bar{u}$ be such that $(\bar{x},\bar{y}) = (\bar{u},\bar{v})$.
-We define a formula $\mu_i$ that is satisfied only by the lexicographically smallest $\bar{v}$ that satisfies $\varphi_i^{\fo}$ for some $\bar{u}$:
-$$
-\mu_i(\bar{v}) = \exists\bar{u}\,\varphi_i^{\fo}(\bar{u},\bar{v})\wedge\forall\bar{u}'\forall\bar{v}'(\varphi_i^{\fo}(\bar{u}',\bar{v}')\to\bar{v}\leq\bar{v}').
-$$
-The following is a pivotal formula for this proof.
-Consider some ordered finite structure $\A$ over $\R$.
-Given our conditions for $\varphi_i$, if $\A\models\varphi_i$ and $\A\models\varphi^{\fo}_i(\bar{u},\bar{v})$ then for that $\bar{v}$ we can define a $\bar{X}$ that has exactly the tuples mentioned in $\bar{v}$ and nothing else.
-This is {\em the smallest $\bar{X}$ that satisfies $\varphi_i$ over $\bar{v}$}.
-And if $\bar{v}$ is the lexicographically smallest such that $\A\models\varphi_i(\bar{v})$, then it is the {\em the smallest $\bar{X}$ that satisfies $\varphi_i$}.
-The following formula is satisfied by every pair $(\bar{X},\bar{x})$, except for the one formed by the smallest $\bar{X}$ that satisfies $\varphi_i$ and the lexicographically smallest $\bar{x}$ that satisfies $\varphi_i(\bar{X})$.
-And if no pair satisfies $\varphi_i$, the following formula is satisfied by every pair.
-$$
-\psi_i(\bar{X},\bar{x}) = \exists\bar{v}(\mu_i(\bar{v})\wedge(\neg\varphi^{+}_i(\bar{X},\bar{v})\vee\bigvee_{X \in \bar{X}} \exists\bar{z}(X(\bar{z}) \wedge \bigwedge\limits_{\substack{\text{instances of }X(\bar{w}) \\ \text{in }\varphi^{+}_i(\bar{X},\bar{v})}}\bar{w}\neq\bar{z}))) \vee \exists\bar{x}'(\exists\bar{y}'\varphi_i(\bar{X},\bar{x}',\bar{y}')\wedge \bar{x}'<\bar{x})\vee\neg\exists\bar{v}\mu_i(\bar{v}) .
-$$
-\begin{lemma}
-	For a given ordered structure $\A$ such that $\A\models\exists\bar{X}\exists\bar{x}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, there is an assignment $(\bar{B},\bar{b})$ to $(\bar{X},\bar{x})$ that satisfies the following conditions (1) $\A\models\exists\bar{y}\,\varphi_i(\bar{B},\bar{b},\bar{y})$, (2) $\A\not\models\psi_i(\bar{B},\bar{b})$ and (3) this is the only assignment to $(\bar{X},\bar{x})$ that satisfies (1) and (2).
-\end{lemma}
-\begin{proof}
-	Excusing the nested proof, we have a claim that identifies a crucial condition that is verifiable with a $\fo$ formula. 
-	\begin{claim}
-		Let $\varphi_i(\bar{X},\bar{x},\bar{y}) = \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X},\bar{x},\bar{y})$ be a $\logex{0}$ formula that satisfies all the mentioned assumptions.
-For a given ordered structure $\A$ it holds that $\A\models\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$ iff $\A\models\exists\bar{X}\exists\bar{x}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$.
-	\end{claim}
-	\begin{proof}
-		Let $\A$ be an ordered structure with domain $A$ and let $\bar{a}\in A^{\length(\bar{y})}$ and $\bar{b}\in A^{\length(\bar{x})}$ such that $\A\models\varphi^{\fo}_i(\bar{b},\bar{a})$.
-Define $\bar{B} = (B_1,\ldots,B_{\length(\bar{X})})$ as $B_j = \{\bar{c}\mid\bar{c}\text{ is a subtuple of $(\bar{b},\bar{a})$ and $X_j(\bar{c})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{b},\bar{a})$}\}$.
-Towards a contradiction, suppose that $\A\not\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
-By the choice of $\bar{a}$ and $\bar{b}$, and construction of $\bar{B}$ it is clear that $\A\models\varphi^{\fo}_i(\bar{b},\bar{a})\wedge\varphi^{+}_i(\bar{B},\bar{b},\bar{a})$, so we have that $\A\not\models\varphi^{-}_i(\bar{B},\bar{b},\bar{a})$.
-Let $B_t(\bar{c})$ be such that $X_t(\bar{c})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{b},\bar{a})$ and $\A\not\models\neg B_t(\bar{c})$.
-This is, $\bar{c}\in B_t$.
-By the construction of $\bar{B}$ we have that $X_t(\bar{c})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{b},\bar{a})$.
-From our assumptions, we have that (1) every first-order variable is mentioned at most once in $\varphi^{-}_i(\bar{X},\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X},\bar{x},\bar{y})$, (2) $\varphi^{\fo}_i$ defines an ordered partition over $(\bar{x},\bar{y})$, and (3) there are no $X_j(\bar{w})$ in $\varphi^{+}_i$ and $X_j(\bar{z})$ in $\varphi^{-}_i$ such that the ordered partition defined by $\varphi^{\fo}_i$ satisfies $\bar{x} = \bar{z}$.
-Since $\bar{c}$ is a subtuple of $(\bar{b},\bar{a})$ that satisfies $\varphi^{\fo}_i$, and $X_t(\bar{w})$ mentioned in $\varphi^{+}_i$ and $X_t(\bar{z})$ mentioned in $\varphi^{-}_i$ were both assigned the value $\bar{c}$, then the ordered partition satisfies $\bar{w} = \bar{z}$, which follows to a contradiction.
-	\end{proof}
-	We give the proof for the base theorem.
-Let $\A$ be an ordered structure with domain $A$.
-We clearly have that $\A\models\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$, so let $\bar{a}\in A^{\length(\bar{y})}$ and $\bar{b}\in A^{\length(\bar{x})}$ such that $\A\models\mu_i(\bar{b},\bar{a})$.
-We use the same construction in the claim and we obtain the assignment $(\bar{B},\bar{b},\bar{a})$ which satisfies the three conditions.
-\end{proof}
-
-We define $\chi_i = \exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$.
-Now recall that $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$, where each $\varphi_i$ satisfies all the previous assumptions.
-For each $\varphi_i$ we define:
-$$
-\varphi_i'(\bar{X},\bar{x},\bar{y}) = \varphi_i(\bar{X},\bar{x},\bar{y})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\vee\psi_2(\bar{X},\bar{x}))\wedge(\chi_1\vee\chi_2\vee\psi_2(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigvee_{j = 1}^{j = i-1}\chi_j\vee\psi_i(\bar{X},\bar{x})),
-$$
-and lastly, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x},\bar{y})$.
-
-\vspace{1em}
-
-For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y}))$ for some algebraic formula $\beta$ and some quantifier-free formula $\varphi$, we define $\kappa(\alpha)$ as follows: First, perform the same transformations to $\varphi(\bar{X},\bar{x},\bar{y})$ as in the previous case.
-Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$ where each $\varphi_i$ satisfies the previous assumptions.
-We also use the previously defined formulas $\chi_i$ and $\psi_i$.
-
-We recursively define a function $\lambda$ as follows.
-If $\alpha = \sa{\bar{x}}\exists\bar{y}\varphi(\bar{x},\bar{y})$, then $\lambda(\alpha) = \exists\bar{x}'\exists\bar{y}'\varphi(\bar{x}',\bar{y}')$.
-If $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ and $\lambda(\alpha) = \chi_1\vee\cdots\vee\chi_m$ as each $\chi_i$ was previously defined.
-If $\alpha = (\beta + \gamma)$, then $\lambda(\alpha) = \lambda(\beta) \vee \lambda(\gamma)$.
-Note that for a given ordered structure $\A$, then $\A\models\lambda(\alpha)$ if and only if $\sem{\alpha}(\A) > 0$.
+We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if and only if $\sem{\beta}(\A) = 0$.
+If $\beta = \sa{\bar{x}}\exists\bar{y}\varphi(\bar{x},\bar{y})$, then $\lambda(\beta) = \neg\exists\bar{x}'\exists\bar{y}'\varphi(\bar{x}',\bar{y}')$.
+If $\beta = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$ as each $\chi_i$ was previously defined.
+If $\beta = (\beta_1 + \beta_2)$, then $\lambda(\beta) = \lambda(\beta_1) \wedge \lambda(\beta_2)$.
 
@@
 $$
-\varphi_i'(\bar{X},\bar{x},\bar{y}) = \varphi_i(\bar{X},\bar{x},\bar{y})\wedge[\lambda(\beta)\vee[\psi_1(\bar{X},\bar{x})\wedge(\chi_1\vee\psi_2(\bar{X},\bar{x}))\wedge(\chi_1\vee\chi_2\vee\psi_2(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigvee_{j = 1}^{j = i-1}\chi_j\vee\psi_i(\bar{X},\bar{x}))]].
+\varphi_i'(\bar{X},\bar{x},\bar{y}) = \varphi_i(\bar{X},\bar{x},\bar{y})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_2(\bar{X},\bar{x}))\wedge\cdots\wedge(
+\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x}))\Big)\Big).
 $$
 And lastly $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x},\bar{y})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
-
-\vspace{1em}
-For the second part, for each $\alpha$ in $\eqso(\logex{0})$ over a signature $\R$ we will define a formula $\lambda(\alpha)$ in $\eqso(\loge{0})$ over a signature $\R_{\alpha}$, and a function $g_{\alpha}$ that receives a finite $\R$-structure $\A$ and outputs a finite $\R_{\alpha}$-structure $g_{\alpha}(\A)$.
-Let $\alpha$ be in $\eqso(\logex{0})$.
-The signature $\R_{\alpha}$ is defined by adding the symbol $R_{\psi}$ for every $\fo$ formula $\psi(\bar{z})$.
-Each symbol $R_{\psi}$ represents a predicate with arity $\length(\bar{z})$.
-Then, $\lambda(\alpha)$ is defined by replacing every $\fo$ formula $\psi(\bar{z})$ for $R_{\psi}(\bar{z})$.
-We now define the function $g_{\alpha}$.
-Let $\A$ be a $\R$-structure with domain $A$.
-Let $\A'$ be a $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A}$ empty.
-For each $\fo$-formula $\psi(\z)$ with $\length(\bar{z})$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length(\bar{z})}$.
-If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$.
-This concludes the construction of $\A'$.
-Note that the number of $\fo$ subformulas and each arity and tuple size is fixed to $\alpha$, so the entire procedure takes polynomial time over the size of the structure (and $g_{\alpha}(\A)$ has polynomial size over $\A$).
-We define $g_{\alpha}(\A) = \A'$ and we have that for each finite $\R$-structure $\A$: $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$.
-Therefore, we have a parsimonious reduction from $\alpha$ to $\lambda(\alpha)$, which is in $\eqso(\loge{1})$.
-
-To show that $\alpha$ is in $\totp$, consider a procedure that takes an input $\A$, converts it to $g_{\alpha}(\A)$ and simulates the $\totp$ procedure for $\lambda(\alpha)$ on input $g_{\alpha}(\A)$.
-This procedure generates exactly $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$ branches, and therefore $\alpha$ is in $\totp$.
-
-To show that $\alpha$ has an FPRAS, consider a procedure that takes an input $\A$, converts it to $g_{\alpha}(\A)$ and simulates the FPRAS for $\lambda(\alpha)$.
-This procedure also takes polynomial time and satisfies the condition.
 

### 2017-01-12 LICS/appendix-syntactic.tex
Avance temporal en minus one.

@@
 
-{\bf Part 1.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
+\noindent {\bf Closed under sum and multiplication.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
 This can be seen because $\eqso(\LL)$ is closed under sum for every fragment $\LL$ by definition, and since $\logex{1}$ is closed under conjunction, from Lemma \ref{conj-mult} it follows that $\eqso(\logex{1})$ is closed under multiplication.
@@
 \vspace{1em}
-{\bf Part 2.} We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
+\noindent {\bf Easy decision version and FRPRAS.}  We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
 We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Theorem \ref{prop:qe0-fp-qe1-totp-fptras}.
@@
 For each $\fo$-formula $\psi(\z)$ with $\length(\bar{z})$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length(\bar{z})}$.
-If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$.
+If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$ (this can be done in $\ptime$).
 This concludes the construction of $\A'$.
-Note that the number of $\fo$ subformulas and each arity and tuple size is fixed to $\alpha$, so computing this function takes polynomial time over the size of the structure.
+Note that the number of $\fo$ subformulas, arity and tuple size is fixed in $\alpha$, so computing this function takes polynomial time over the size of the structure.
 Moreover, the encoding of $\A'$ has polynomial size over the size of $\enc(\A)$.
@@
 
-To show that $\alpha$ is in $\totp$, consider a procedure that takes an input $\enc(\A)$, converts it to $\enc(g_{\alpha}(\A))$ and simulates the $\totp$ procedure for $\lambda(\alpha)$ on input $\enc(g_{\alpha}(\A))$.
-This procedure generates exactly $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$ branches, and therefore $\alpha$ is in $\totp$.
-
-To show that $\alpha$ has an FPRAS, consider a procedure that takes an input $\enc(\A)$, converts it to $\enc(g_{\alpha}(\A))$ and simulates the FPRAS for $\lambda(\alpha)$.
-This procedure also takes polynomial time and satisfies the condition.
+To show that $\alpha$ is in $\totp$, we can convert $\alpha$ and $\enc(\A)$ into $\lambda(\alpha)$ and $\enc(g_{\alpha}(\A))$, respectively, and run the procedure in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
+Similarly, to show that $\alpha$ has an FPRAS, we do the same as before and simulates the FPRAS for $\lambda(\alpha)$ in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
+These procedures also takes polynomial time and satisfies the condition.
 
 \vspace{1em}
-{\bf Part 3.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one.
-Let $\alpha$ be a $\eqso(\logex{1})$ formula over a signature $\R$.
-We will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-We suppose that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
-Moreover, we assume that $\length(\bar{x}) > 0$ and $\length(\bar{X}) > 0$ since we can replace each $\varphi_i$ for the equivalent formula $\sa{Y}\sa{y}\varphi_i\wedge\first(y)\wedge\forall z Y(z)$.
+\noindent {\bf Closed under subtraction by one.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. 
+For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
+Without lost of generality, we assume that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
+Moreover, we assume that $\length(\bar{x}) > 0$  since, if this is not the case, we can replace $\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
 
 The proof will be separated in two parts.
-(i) If $\alpha$ is in $\logex{1}$-PNF, this is, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$, then we define a formula $\varphi'$ that satisfies the following condition.
-For each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-We define $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$.
-(ii) If $\alpha = \beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$, then we define a formula $\varphi'$ that satisfies the following condition.
-For each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$ {\bf and} $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-We define $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$.
+In the first part, we suppose that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$.
+Then we will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
+From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ is the formula.
+In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ the sum of one or more formulas in $\logex{1}$-PNF.
+We define a formula $\varphi'$ that satisfies the following condition: if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ and $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
+From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be proven.
 
@@
 
-{\em Section (i).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ where $\varphi$ is quantifier-free with the addition of atomic formulas in $\fo$.
+\noindent {\em Part (1).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ where $\varphi$ is quantifier-free and with  atomic formulas in $\fo$.
 We call this syntax {\em extended quantifier-free} from now on.
+Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ (i.e.
+$\length(\bar{X}) = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}}\exists\bar{y}[\varphi(\bar{x},\bar{y})\wedge\exists\bar{z}(\varphi(\bar{z},\bar{y})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
+Therefore, for the rest of the proof we can assume that $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.
+
  Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
-We convert $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m\varphi_i$ where each $\varphi_i$ is of the form:
+To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in into a DNF formula. 
+More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i$ is a conjunction of atomic formulas or negation of atomic formulas.
+Furthermore, we suppose that each conjunct $\varphi_i$ has the~form:
 $$
 \varphi_i(\bar{X},\bar{x},\bar{y}) = \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
-\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}} \wedge \underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}}.
+\underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}} \wedge
+\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}}.
 $$
-This can be done with a standard predicate-DNF transformation algorithm.
-We now define a series of transformations to $\varphi$ that will make each formula $\varphi$ satisfy the following three conditions.
-(a) Each first-order variable in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ is mentioned at most once.
-Moreover, in this formula no variable in $\bar{x}$ is mentioned.
-(b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
-We will define then what we mean by an ordered partition.
-(c) If $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
+Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ can appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
+
+Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) each first-order variable in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ is mentioned at most once and, moreover, no variable in $\bar{x}$ is mentioned, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
+We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
 \begin{itemize}
-	\item[(a)] Consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
+	\item[(a)] In order to satisfy the first condition, consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
 We add $\length(\bar{w})$ new variables $z_1,\ldots,z_{\length(\bar{w})}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length(\bar{w})})$.
@@
 
-	The formula $\varphi(\bar{X},\bar{x},\bar{y})$ is then redefined as
+	Thus, the formula $\varphi(\bar{X},\bar{x},\bar{y})$ is then redefined as:
 	$$
@@
 	$$
-	and each of these disjuncts is a $\varphi_i$. Therefore, each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
-	\item[(c)] For each $\varphi_i$, we do the following.
+	Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
+	Then, by redefining the conjuncts $\varphi_i(\bar{X},\bar{x},\bar{y})$ of $\varphi$ to $\theta^j(\bar{x},\bar{y})\wedge \varphi_i(\bar{X},\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
+	\item[(c)] For showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
+$\varphi_i$ is inconsistent), we do the following.
 If there exists an instance of $X_j(\bar{z})$ in $\varphi^{+}_i$, an instance of $\neg X_j(\bar{w})$ in $\varphi^{-}_i$ and the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$, then the entire formula $\varphi_i$ is removed from $\varphi$.
@@
 From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y}) = \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c).
-Note the following.
-\begin{claim}
+\begin{claim}\label{claim:minusone}
 For a given ordered structure $\A$, if there exists a first-order assignment $v$ for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists a second-order assignment $V$ for $\A$ such that  $(\A,V,v)\models\varphi_i(\bar{X},\bar{x},\bar{y})$.
@@
 Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
-Define $\bar{B} = (B_1,\ldots,B_{\length(\bar{X})})$ as $B_j = \{v(\bar{w})\mid\bar{w}\text{ is a subtuple of $\bar{y}$ and $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
+Define $\bar{B} = (B_1,\ldots,B_{\length(\bar{X})})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
 Towards a contradiction, suppose that $(\A,V,v)\not\models\varphi_i(\bar{X},\bar{x},\bar{y})$.
 By the choice of $v$, and construction of $V$ it is clear that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})\wedge\varphi^{+}_i(\bar{X},\bar{x},\bar{y})$, so we necessarily have that $(\A,V,v)\not\models\varphi^{-}_i(\bar{X},\bar{x},\bar{y})$.
-Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,v)\not\models\neg B_t(\bar{w})$.
-This is, $v(\bar{w})\in B_t$, but because of the way we constructed $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$.
+Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,V, v)\not\models\neg X_t(\bar{w})$, namely, $v(\bar{w})\in B_t$. 
+However, by the construction of $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$.
 Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$.
-However, this violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
+This violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
 \end{proof}
-We generalize the construction of $\bar{B}$ given in the claim.
-For a structure $\A$ and a first-order assignment $v$ for $\A$, then we define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length(\bar{X})})$ where each $B^v_j = \{v(\bar{w})\mid\bar{w}\text{ is a subtuple of $\bar{y}$ and $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
-Now, let $\bar{v}$ be the subtuple of $\bar{y}$ of every first-order variable mentioned in $\varphi_i^{+}$ and let $\bar{u}$ be a subtuple of $(\bar{x},\bar{y})$ such that concatenating $(\bar{v},\bar{u})$ is equal to $(\bar{x},\bar{y})$.
-We also define $v^*$ as a first-order assignment for $\A$ for which we have that $v^{*}(\bar{v})$ is the lexicographically smallest tuple that satisfies $(\A,v^{*})\models\exists\bar{u} \varphi^{\fo}_i(\bar{u},\bar{v})$.
-Lastly, we define $\bar{B}^* = \bar{B}^{v^{*}}$ and $V^*$ as some second-order assignment for $\A$ for which $V^*(\bar{X}) = \bar{B}^*$.
 
-We define a formula $\mu_i$ that is satisfied only by $v^*$:
+The previous claim motivates the following definitions, in particular, a generalization of the sequence $\bar{B}$. 
+For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length(\bar{X})})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
+Moreover, let $\bar{v}$ be the sequence of all different first-order variable mentioned in $\varphi_i^{+}$ and let $\bar{u}$ be a sequence of variables disjoint from $\bar{v}$ such that $(\bar{v},\bar{u})$ and $(\bar{x},\bar{y})$ have the same set of variables.
+Then define $v^*$ to be the first-order assignment for $\A$ for which we have that $v^{*}(\bar{v})$ is the lexicographically smallest tuple that satisfies $(\A,v^{*})\models\exists\bar{u} \varphi^{\fo}_i(\bar{u},\bar{v})$.
+Indeed, one can define a formula $\mu_i$ that defines $v^*$ as follows:
 $$
@@
 $$ 
+In other words, we have that $(\A,v)\models \mu_i(\bar{v})$ if, and only if, $v(\bar{v}) = v^{*}(\bar{v})$ for every assignment $v$. 
+Finally, define $\bar{B}^* = \bar{B}^{v^{*}}$ and $V^*$ as some second-order assignment for $\A$ for which $V^*(\bar{X}) = \bar{B}^*$.
+
+Now, we have all the tools to introduce the main formula of the proof. 
+Intuitively, the pair $(V^*, v^*)$ is the minimum assignment that satisfies $\varphi_i(\bar{X}, \bar{x})$ and our candidate to be removed from the semantics of $\varphi(\bar{X}, \bar{x})$.
+Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ that has the following key property:
+\begin{center}
+if $\varphi_i(\bar{X}, \bar{x}, \bar{y})$ is satisfiable, then $(\A,V,v) \models\varphi_i(\bar{X},\bar{x},\bar{y})$ if, and only if, $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$ for every $(V,v)$.
+\end{center}
+That is, its exclude somehow the pair $(V^*, v^*)$. This property can be defined in logic as follows:
+\begin{align}
+\psi_i(\bar{X},\bar{x}) = & \;\; \big( \, \exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
+& \hspace{1cm} \Big( \; 
+\exists\bar{x}'(\exists\bar{y}'\varphi_i(\bar{X},\bar{x}',\bar{y}')\wedge \bar{x}'<\bar{x}) \;\;  \vee   \label{eq:part2} \\
+& \hspace{1.5cm} \exists\bar{v} (\mu_i(\bar{v})\wedge \big( \varphi^{+}_i(\bar{X},\bar{v}) \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big) \Big)  \label{eq:part3}
+\end{align}
+To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
+Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi_i(\bar{X},\bar{x},\bar{y})$.
+Then, the conclusion of the implication at (\ref{eq:part2}) and (\ref{eq:part3}), take care that the $\varphi_i$ is only satisfied by a pair $(V, v)$ different from $(V^*, v^*)$.
+Here, the first disjunct at (\ref{eq:part2}) check whether $(\A,V,v) \models\varphi_i(\bar{X},\bar{x},\bar{y})$ but $v(\bar{x})$ is not the lexicographically smallest tuple that have this property (i.e.
+$v(\bar{x}) \neq v^*(\bar{x})$).
+The second disjunct at (\ref{eq:part3}) is more involved and it can be explained as follows:
+
 The following is a pivotal formula in our proof:
+
 \begin{align}
@@
 \end{align}
+
 Consider some $\A$.

### 2017-01-12 LICS/appendix-syntactic.tex
Cambio menor.

@@
 It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still an extended quantifier-free formula.
-From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y}) = \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c).
+From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c).
+Furthermore, by condition (a) we have that:
+$$
+\varphi_i(\bar{X},\bar{x},\bar{y}) \; = \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
+$$
+where $\varphi^{+}$ and $\varphi^{-}_i$ do not depend on the variables $\bar{x}$.
 \begin{claim}\label{claim:minusone}
@@
 
-The previous claim motivates the following definitions, in particular, a generalization of the sequence $\bar{B}$. 
+The previous claim and proof motivates the following definitions, in particular, a generalization of the sequence $\bar{B}$. 
 For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length(\bar{X})})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
@@
 Then define $v^*$ to be the first-order assignment for $\A$ for which we have that $v^{*}(\bar{v})$ is the lexicographically smallest tuple that satisfies $(\A,v^{*})\models\exists\bar{u} \varphi^{\fo}_i(\bar{u},\bar{v})$.
-Indeed, one can define a formula $\mu_i$ that defines $v^*$ as follows:
+Indeed, one ca define a formula $\mu_i$ that defines $v^*$ as follows:
 $$

### 2017-01-12 LICS/appendix-syntactic.tex
Version nueva de minus one.

@@
 
-\noindent {\em Part (1).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ where $\varphi$ is quantifier-free and with  atomic formulas in $\fo$.
-We call this syntax {\em extended quantifier-free} from now on.
-Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ (i.e.
-$\length(\bar{X}) = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}}\exists\bar{y}[\varphi(\bar{x},\bar{y})\wedge\exists\bar{z}(\varphi(\bar{z},\bar{y})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
+\noindent {\em Part (1).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ where $\varphi$ is an  $\fo$-formula.
+Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}} \varphi(\bar{x})$ (i.e.
+$\length(\bar{X}) = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}} [\varphi(\bar{x})\wedge\exists\bar{z}(\varphi(\bar{z})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
 Therefore, for the rest of the proof we can assume that $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.
@@
 $$
-\varphi_i(\bar{X},\bar{x},\bar{y}) = \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
+\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
 \underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}} \wedge
@@
 
-Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) each first-order variable in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ is mentioned at most once and, moreover, no variable in $\bar{x}$ is mentioned, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
+Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ are mentioned in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
 We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
@@
 	$$
-	\varphi_i(\bar{X},\bar{x},(\bar{y},\bar{z})) = (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{-}(\bar{X},\bar{x},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y}).
+	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \exists \bar{z} \; (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{-}(\bar{X},\bar{x},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y})).
 	$$
@@
 	$$
-	\varphi_i(\bar{X},\bar{x},\bar{y}) = \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
+	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
 	$$
-	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
+	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = \exists \bar{y} \;  X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := \exists \bar{y} \;  X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
 	\item[(b)] An ordered partition on a set $S$ is defined by an equivalence relation $\sim$ over $S$, and a linear order over $S/\!\sim$.
@@
 
-	Thus, the formula $\varphi(\bar{X},\bar{x},\bar{y})$ is then redefined as:
+	Thus, the formula $\varphi_i(\bar{X},\bar{x})$ is then redefined as:
 	$$
-	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}}[\theta^j(\bar{x},\bar{y})\wedge \varphi_i(\bar{X},\bar{x},\bar{y})],
+	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}} \exists \bar{y} \; [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
 	$$
 	Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
-	Then, by redefining the conjuncts $\varphi_i(\bar{X},\bar{x},\bar{y})$ of $\varphi$ to $\theta^j(\bar{x},\bar{y})\wedge \varphi_i(\bar{X},\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
+	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  to $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
 	\item[(c)] For showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
@@
 \end{itemize}
-It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still an extended quantifier-free formula.
+It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$.
 From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c).
-Furthermore, by condition (a) we have that:
+Furthermore, it has the following form:
 $$
-\varphi_i(\bar{X},\bar{x},\bar{y}) \; = \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
+\varphi_i(\bar{X},\bar{x}) \; = \; \exists{\bar{y}} \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
 $$
-where $\varphi^{+}$ and $\varphi^{-}_i$ do not depend on the variables $\bar{x}$.
+where $\varphi^{+}$ and $\varphi^{-}_i$ do not depend on  $\bar{x}$.
 \begin{claim}\label{claim:minusone}
-For a given ordered structure $\A$, if there exists a first-order assignment $v$ for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists a second-order assignment $V$ for $\A$ such that  $(\A,V,v)\models\varphi_i(\bar{X},\bar{x},\bar{y})$.
+For a given ordered structure $\A$, if there exists a first-order assignment $v$ for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists a second-order assignment $V$ for $\A$ such that  $(\A,V,v)\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
 \end{claim}
@@
 Define $\bar{B} = (B_1,\ldots,B_{\length(\bar{X})})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
-Towards a contradiction, suppose that $(\A,V,v)\not\models\varphi_i(\bar{X},\bar{x},\bar{y})$.
-By the choice of $v$, and construction of $V$ it is clear that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})\wedge\varphi^{+}_i(\bar{X},\bar{x},\bar{y})$, so we necessarily have that $(\A,V,v)\not\models\varphi^{-}_i(\bar{X},\bar{x},\bar{y})$.
+Towards a contradiction, suppose that $(\A,V,v)\not\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
+By the choice of $v$, and construction of $V$ it is clear that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})\wedge\varphi^{+}_i(\bar{X},\bar{y})$, so we necessarily have that $(\A,V,v)\not\models\varphi^{-}_i(\bar{X},\bar{y})$.
 Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,V, v)\not\models\neg X_t(\bar{w})$, namely, $v(\bar{w})\in B_t$. 
@@
 
-The previous claim and proof motivates the following definitions, in particular, a generalization of the sequence $\bar{B}$. 
+The previous claim and proof motivates the following definitions.
 For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length(\bar{X})})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
-Moreover, let $\bar{v}$ be the sequence of all different first-order variable mentioned in $\varphi_i^{+}$ and let $\bar{u}$ be a sequence of variables disjoint from $\bar{v}$ such that $(\bar{v},\bar{u})$ and $(\bar{x},\bar{y})$ have the same set of variables.
-Then define $v^*$ to be the first-order assignment for $\A$ for which we have that $v^{*}(\bar{v})$ is the lexicographically smallest tuple that satisfies $(\A,v^{*})\models\exists\bar{u} \varphi^{\fo}_i(\bar{u},\bar{v})$.
-Indeed, one ca define a formula $\mu_i$ that defines $v^*$ as follows:
+One can easily check that for every assignments $V, v$ such that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$, then it holds that $(\A,\bar{B}^v,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all the assignment of $\bar{X}$ when $v$ is fixed.
+Therefore, by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
+For this, the idea is to choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
 $$
-\mu_i(\bar{v}) = \exists\bar{u}\,\varphi_i^{\fo}(\bar{u},\bar{v})\wedge\forall\bar{u}'\forall\bar{v}'(\varphi_i^{\fo}(\bar{u}',\bar{v}')\to\bar{v}\leq\bar{v}').
+\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge\forall\bar{x}'\forall\bar{y}'(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')).
 $$ 
-In other words, we have that $(\A,v)\models \mu_i(\bar{v})$ if, and only if, $v(\bar{v}) = v^{*}(\bar{v})$ for every assignment $v$. 
-Finally, define $\bar{B}^* = \bar{B}^{v^{*}}$ and $V^*$ as some second-order assignment for $\A$ for which $V^*(\bar{X}) = \bar{B}^*$.
+If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
+Furthermore, let $V^*$ the second order assignment that satisfies $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*$ the first order assignment such that $v^*(\bar{x}) = v(\bar{x})$.
+By the previous discussion, we have that $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.
 
 Now, we have all the tools to introduce the main formula of the proof. 
-Intuitively, the pair $(V^*, v^*)$ is the minimum assignment that satisfies $\varphi_i(\bar{X}, \bar{x})$ and our candidate to be removed from the semantics of $\varphi(\bar{X}, \bar{x})$.
-Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ that has the following key property:
-\begin{center}
-if $\varphi_i(\bar{X}, \bar{x}, \bar{y})$ is satisfiable, then $(\A,V,v) \models\varphi_i(\bar{X},\bar{x},\bar{y})$ if, and only if, $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$ for every $(V,v)$.
-\end{center}
-That is, its exclude somehow the pair $(V^*, v^*)$. This property can be defined in logic as follows:
+Intuitively, we want to exclude the assignment $(V^*, v^*)$ as a satisfying assignment for $\varphi_i(\bar{X},\bar{x})$.
+Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if, 
+if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, then $V \neq V^*$ or $v \neq v^*$. 
+That is, it excludes all pairs $(V, v)$ that coincide with $(V^*, v^*)$ in $(\bar{X}, \bar{x})$ whenever $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. This property can be defined as follows:
 \begin{align}
@@
 & \hspace{1cm} \Big( \; 
-\exists\bar{x}'(\exists\bar{y}'\varphi_i(\bar{X},\bar{x}',\bar{y}')\wedge \bar{x}'<\bar{x}) \;\;  \vee   \label{eq:part2} \\
-& \hspace{1.5cm} \exists\bar{v} (\mu_i(\bar{v})\wedge \big( \varphi^{+}_i(\bar{X},\bar{v}) \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big) \Big)  \label{eq:part3}
+\exists\bar{y} \;\; \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i^+(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\
+& \hspace{1.5cm} (\exists\bar{x}'\exists\bar{y}\, \varphi_i(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}) \;\; \Big)  \label{eq:part3}
 \end{align}
 To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
-Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi_i(\bar{X},\bar{x},\bar{y})$.
-Then, the conclusion of the implication at (\ref{eq:part2}) and (\ref{eq:part3}), take care that the $\varphi_i$ is only satisfied by a pair $(V, v)$ different from $(V^*, v^*)$.
-Here, the first disjunct at (\ref{eq:part2}) check whether $(\A,V,v) \models\varphi_i(\bar{X},\bar{x},\bar{y})$ but $v(\bar{x})$ is not the lexicographically smallest tuple that have this property (i.e.
-$v(\bar{x}) \neq v^*(\bar{x})$).
-The second disjunct at (\ref{eq:part3}) is more involved and it can be explained as follows:
+Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
+Then, the conclusion of the implication at (\ref{eq:part2}) and (\ref{eq:part3}), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
+Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by showing that if $\varphi_i(\bar{X},\bar{x})$ is satisfied then $V(\bar{X})$ must contain one element more that it is not in $V^*(\bar{X})$.
+The second disjunct (\ref{eq:part3}) is satisfied when $v(\bar{x})$ is not the lexicographically smallest tuple that satisfies $\varphi_i$ (i.e. $v(\bar{x}) \neq v^*(\bar{x})$).
 
-The following is a pivotal formula in our proof:
-
-\begin{align}
-\psi_i(\bar{X},\bar{x}) = &\exists\bar{v}\bigg(\overbrace{\mu_i(\bar{v})}^{(\blacktriangleleft)}\wedge\Big( \overbrace{\neg\varphi^{+}_i(\bar{X},\bar{v})}^{(\blacktriangle)} \vee \overbrace{\bigvee_{X \in \bar{X}} \exists\bar{z}(X(\bar{z}) \wedge \bigwedge\limits_{\substack{\text{subtuples }\bar{w} \text{ of }\bar{v} \text{ s.t.}\\ X(\bar{w})\text{ is mentioned} \\ \text{in }\varphi^{+}_i(\bar{X},\bar{v})}}\bar{w}\neq\bar{z}}^{(\blacktriangledown)})\Big)\bigg) \vee \label{minusone-1}\\ &\exists\bar{x}'(\exists\bar{y}'\varphi_i(\bar{X},\bar{x}',\bar{y}')\wedge \bar{x}'<\bar{x})\vee \label{minusone-2} \\
-&\neg\exists\bar{x}\exists\bar{y}\varphi^{\fo}_i(\bar{x},\bar{y}). \label{minusone-3}
-\end{align}
-
-Consider some $\A$.
-In detail, the formula \eqref{minusone-1} represents the following: In $(\blacktriangleleft)$, we make it so the only value for $\bar{v}$ we consider is $v^{*}(\bar{v})$.
-In $(\blacktriangle)$, we mention that the assignment for $\bar{X}$ satisfies the formula if does not contain at least one of the tuples in $v^{*}(\bar{v})$.
-In $(\blacktriangledown)$, we mention that in the assignment for $\bar{X}$ also satisfies the formula if it contains at least one tuple, in one of the $X$'s, that is not equal to the one in $v^{*}(\bar{v})$.
-Then, in the formula \eqref{minusone-2} we mention that the assignment for $\bar{x}$ satisfies the formula if it is not the lexicographically smallest assignment that satisfies $\varphi_i$ for that particular assignment for $\bar{X}$.
-Lastly, in \eqref{minusone-3}, we mention that the assignment for $(\bar{X},\bar{x})$ satisfies the formula if there are no assignments $V,v$ for $\A$ such that $(\A,V,v)\not\models \exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$.
-
-This way, the formula $\psi(\bar{X},\bar{x})$ functions as a sort of filter for valid assignments for $\bar{X}$ and $\bar{x}$.
-For each structure $\A$ we have that the only $V,v$ for which $(\A,V,v)\models\exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$ and also $(\A,V,v)\not\models\psi_i(\bar{X},\bar{x})$ is the one given by $v = v^*$ and $V = V^*$, should there exist one.
-%We prove this formally.
-%\begin{lemma}
-%	For a given ordered structure $\A$ such that $\A\models\exists\bar{X}\exists\bar{x}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, there is an assignment $(\bar{B},\bar{b})$ to $(\bar{X},\bar{x})$ that satisfies the following conditions (1) $\A\models\exists\bar{y}\,\varphi_i(\bar{B},\bar{b},\bar{y})$, (2) $\A\not\models\psi_i(\bar{B},\bar{b})$ and (3) this is the only assignment to $(\bar{X},\bar{x})$ that satisfies (1) and (2).
-%\end{lemma}
-%\begin{proof}
-%	Excusing the nested proof, we have a claim that identifies a crucial condition that is verifiable with a $\fo$ formula. 
-%	
-%	We give the proof for the base theorem.
-Let $\A$ be an ordered structure with domain $A$.
-We clearly have that $\A\models\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$, so let $\bar{a}\in A^{\length(\bar{y})}$ and $\bar{b}\in A^{\length(\bar{x})}$ such that $\A\models\mu_i(\bar{b},\bar{a})$.
-We use the same construction in the claim and we obtain the assignment $(\bar{B},\bar{b},\bar{a})$ which satisfies the three conditions.
-%\end{proof}
-
-We define $\chi_i = \neg\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$. Now recall that $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$. We define:
+We are ready to define the formula  $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$
+where each modified disjunct $\varphi_i'(\bar{X},\bar{x})$ is constructed as follows. 
+For the sake of simplification, define the auxiliary formula $\chi_i = \neg\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$. 
+This formula basically checks if $\varphi_i$ is satisfiable or not.
+Then, first define the formula:
 $$
-\varphi'_1(\bar{X},\bar{x},\bar{y}) = \varphi_1(\bar{X},\bar{x},\bar{y})\wedge\psi_1(\bar{X},\bar{x}).
+\varphi'_1(\bar{X},\bar{x}) = \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
 $$
@@
 $$
-\varphi'_2(\bar{X},\bar{x},\bar{y}) = \varphi_2(\bar{X},\bar{x},\bar{y})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
+\varphi'_2(\bar{X},\bar{x}) = \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
 $$
-This models that all the assignments that satisfy $\varphi_2$ also satisfy $\varphi'_2$, but if there was no assignment for $\varphi_1$, then we exclude $V^*,v^*$ for $\varphi_2$.
+This models that all the assignments that satisfy $\varphi_2$ also satisfy $\varphi'_2$, but if there was no assignment for $\varphi_1$, then we exclude $(V^*,v^*)$ for $\varphi_2$.
 Then we generalize this construction for each $\varphi_i$:
 $$
-\varphi_i'(\bar{X},\bar{x},\bar{y}) = \varphi_i(\bar{X},\bar{x},\bar{y})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_2(\bar{X},\bar{x}))\wedge\cdots\wedge(
+\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
 \bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})),
 $$
-and lastly, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x},\bar{y})$, which satisfies our condition that $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.
+By the construction of $\kappa(\alpha)$, one can easily check that  $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.
 
@@
 
-{\em Section (ii).} Let $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y}))$ for some algebraic formula $\beta$ and some extended quantifier-free formula $\varphi$.
-We define $\kappa(\alpha)$ as follows: First, perform the same transformations to $\varphi(\bar{X},\bar{x},\bar{y})$ as in the previous section.
-Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c).
+{\em Part (2).} Let $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ for some $\eqso(\logex{1})$ formula $\beta$. We define $\kappa(\alpha)$ as follows.
+First, perform the same transformations to $\varphi(\bar{X},\bar{x})$ as in the previous section.
+Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c).
 We also use the previously defined formulas $\chi_i$ and $\psi_i$, for each $i$.
@@
 We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if and only if $\sem{\beta}(\A) = 0$.
-If $\beta = \sa{\bar{x}}\exists\bar{y}\varphi(\bar{x},\bar{y})$, then $\lambda(\beta) = \neg\exists\bar{x}'\exists\bar{y}'\varphi(\bar{x}',\bar{y}')$.
-If $\beta = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$ as each $\chi_i$ was previously defined.
+If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg\exists\bar{x}'\varphi(\bar{x}')$.
+If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$ as each $\chi_i$ was previously defined.
 If $\beta = (\beta_1 + \beta_2)$, then $\lambda(\beta) = \lambda(\beta_1) \wedge \lambda(\beta_2)$.
-
 Now, for each $\varphi_i$ we define:
 $$
-\varphi_i'(\bar{X},\bar{x},\bar{y}) = \varphi_i(\bar{X},\bar{x},\bar{y})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_2(\bar{X},\bar{x}))\wedge\cdots\wedge(
+\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
 \bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x}))\Big)\Big).
 $$
-And lastly $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x},\bar{y})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
+Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
 

### 2017-01-12 LICS/appendix-syntactic.tex
12 en

--- 2017-01-12 LICS/appendix-syntactic.tex
@@
 Since $\alpha$ takes the value 0 for some structures, $c$ must be 0.
-Consider a structure with one element $\mathfrak{1}$.
+Consider a structure $\mathfrak{1}$ with one element.
 We have that for some $i$, there exists an assignment $(\bar{B},\bar{b},\bar{a})$ for $(\bar{X},\bar{x},\bar{y})$ such that $\mathfrak{1}\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
@@
 
-We separate this proof in three parts
-
-\vspace{1em}
 The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$.
@@
 
-For the first part, let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
+\vspace{1em}
+$\boldsymbol{\eqso(\loge{1})}$ {\bf is contained in \textsc{TotP}.} Let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
 This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
@@
 \vspace{1em}
-For the second part, let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
+{\bf Every function in} $\boldsymbol{\eqso(\loge{1})}$ {\bf has an FPRAS.}  let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
 This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
@@
 
-\vspace{2em}
-For the third part, we will prove that $\eqso(\loge{1})$ is closed under sum and multiplication.
-Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that the class is closed under multiplication.
+\vspace{1em}
+$\boldsymbol{\eqso(\loge{1})}$ {\bf is closed under sum and multiplication.} Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that the class is closed under multiplication.
 We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of $\so$.
@@
 
-\noindent {\bf Closed under sum and multiplication.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
+{\bf Part 1.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
 This can be seen because $\eqso(\LL)$ is closed under sum for every fragment $\LL$ by definition, and since $\logex{1}$ is closed under conjunction, from Lemma \ref{conj-mult} it follows that $\eqso(\logex{1})$ is closed under multiplication.
@@
 \vspace{1em}
-\noindent {\bf Easy decision version and FRPRAS.}  We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
+{\bf Part 2.} We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
 We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Theorem \ref{prop:qe0-fp-qe1-totp-fptras}.
@@
 For each $\fo$-formula $\psi(\z)$ with $\length(\bar{z})$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length(\bar{z})}$.
-If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$ (this can be done in $\ptime$).
+If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$.
 This concludes the construction of $\A'$.
-Note that the number of $\fo$ subformulas, arity and tuple size is fixed in $\alpha$, so computing this function takes polynomial time over the size of the structure.
+Note that the number of $\fo$ subformulas and each arity and tuple size is fixed to $\alpha$, so computing this function takes polynomial time over the size of the structure.
 Moreover, the encoding of $\A'$ has polynomial size over the size of $\enc(\A)$.
@@
 
-To show that $\alpha$ is in $\totp$, we can convert $\alpha$ and $\enc(\A)$ into $\lambda(\alpha)$ and $\enc(g_{\alpha}(\A))$, respectively, and run the procedure in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-Similarly, to show that $\alpha$ has an FPRAS, we do the same as before and simulates the FPRAS for $\lambda(\alpha)$ in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-These procedures also takes polynomial time and satisfies the condition.
+To show that $\alpha$ is in $\totp$, consider a procedure that takes an input $\enc(\A)$, converts it to $\enc(g_{\alpha}(\A))$ and simulates the $\totp$ procedure for $\lambda(\alpha)$ on input $\enc(g_{\alpha}(\A))$.
+This procedure generates exactly $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$ branches, and therefore $\alpha$ is in $\totp$.
+
+To show that $\alpha$ has an FPRAS, consider a procedure that takes an input $\enc(\A)$, converts it to $\enc(g_{\alpha}(\A))$ and simulates the FPRAS for $\lambda(\alpha)$.
+This procedure also takes polynomial time and satisfies the condition.
 
 \vspace{1em}
-\noindent {\bf Closed under subtraction by one.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. 
-For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Without lost of generality, we assume that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
-Moreover, we assume that $\length(\bar{x}) > 0$  since, if this is not the case, we can replace $\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
+{\bf Part 3.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one.
+Let $\alpha$ be a $\eqso(\logex{1})$ formula over a signature $\R$.
+We will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
+We suppose that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
+Moreover, we assume that $\length(\bar{x}) > 0$ and $\length(\bar{X}) > 0$ since we can replace each $\varphi_i$ for the equivalent formula $\sa{Y}\sa{y}\varphi_i\wedge\first(y)\wedge\forall z Y(z)$.
 
 The proof will be separated in two parts.
-In the first part, we suppose that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$.
-Then we will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ is the formula.
-In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ the sum of one or more formulas in $\logex{1}$-PNF.
-We define a formula $\varphi'$ that satisfies the following condition: if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ and $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be proven.
+(i) If $\alpha$ is in $\logex{1}$-PNF, this is, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$, then we define a formula $\varphi'$ that satisfies the following condition.
+For each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
+We define $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$.
+(ii) If $\alpha = \beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$, then we define a formula $\varphi'$ that satisfies the following condition.
+For each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$ {\bf and} $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
+We define $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$.
 
@@
 
-\noindent {\em Part (1).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ where $\varphi$ is quantifier-free and with  atomic formulas in $\fo$.
+{\em Section (i).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ where $\varphi$ is quantifier-free with the addition of atomic formulas in $\fo$.
 We call this syntax {\em extended quantifier-free} from now on.
-Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ (i.e.
-$\length(\bar{X}) = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}}\exists\bar{y}[\varphi(\bar{x},\bar{y})\wedge\exists\bar{z}(\varphi(\bar{z},\bar{y})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
-Therefore, for the rest of the proof we can assume that $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.
-
 Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
-To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in into a DNF formula. 
-More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i$ is a conjunction of atomic formulas or negation of atomic formulas.
-Furthermore, we suppose that each conjunct $\varphi_i$ has the~form:
+We convert $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m\varphi_i$ where each $\varphi_i$ is of the form:
 $$
 \varphi_i(\bar{X},\bar{x},\bar{y}) = \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge
-\underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}} \wedge
-\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}}.
+\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}} \wedge \underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}}.
 $$
-Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ can appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
-
-Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) each first-order variable in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ is mentioned at most once and, moreover, no variable in $\bar{x}$ is mentioned, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
-We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
+This can be done with a standard predicate-DNF transformation algorithm.
+We now define a series of transformations to $\varphi$ that will make each formula $\varphi$ satisfy the following three conditions.
+(a) Each first-order variable in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ is mentioned at most once.
+Moreover, in this formula no variable in $\bar{x}$ is mentioned.
+(b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
+We will define then what we mean by an ordered partition.
+(c) If $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
 \begin{itemize}
-	\item[(a)] In order to satisfy the first condition, consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
+	\item[(a)] Consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
 We add $\length(\bar{w})$ new variables $z_1,\ldots,z_{\length(\bar{w})}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length(\bar{w})})$.
@@
 
-	Thus, the formula $\varphi(\bar{X},\bar{x},\bar{y})$ is then redefined as:
+	The formula $\varphi(\bar{X},\bar{x},\bar{y})$ is then redefined as
 	$$
@@
 	$$
-	Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
-	Then, by redefining the conjuncts $\varphi_i(\bar{X},\bar{x},\bar{y})$ of $\varphi$ to $\theta^j(\bar{x},\bar{y})\wedge \varphi_i(\bar{X},\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
-	\item[(c)] For showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
-$\varphi_i$ is inconsistent), we do the following.
+	and each of these disjuncts is a $\varphi_i$. Therefore, each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
+	\item[(c)] For each $\varphi_i$, we do the following.
 If there exists an instance of $X_j(\bar{z})$ in $\varphi^{+}_i$, an instance of $\neg X_j(\bar{w})$ in $\varphi^{-}_i$ and the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$, then the entire formula $\varphi_i$ is removed from $\varphi$.
@@
 From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y}) = \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c).
-\begin{claim}\label{claim:minusone}
+Note the following.
+\begin{claim}
 For a given ordered structure $\A$, if there exists a first-order assignment $v$ for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists a second-order assignment $V$ for $\A$ such that  $(\A,V,v)\models\varphi_i(\bar{X},\bar{x},\bar{y})$.
@@
 Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
-Define $\bar{B} = (B_1,\ldots,B_{\length(\bar{X})})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
+Define $\bar{B} = (B_1,\ldots,B_{\length(\bar{X})})$ as $B_j = \{v(\bar{w})\mid\bar{w}\text{ is a subtuple of $\bar{y}$ and $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
 Towards a contradiction, suppose that $(\A,V,v)\not\models\varphi_i(\bar{X},\bar{x},\bar{y})$.
 By the choice of $v$, and construction of $V$ it is clear that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})\wedge\varphi^{+}_i(\bar{X},\bar{x},\bar{y})$, so we necessarily have that $(\A,V,v)\not\models\varphi^{-}_i(\bar{X},\bar{x},\bar{y})$.
-Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,V, v)\not\models\neg X_t(\bar{w})$, namely, $v(\bar{w})\in B_t$. 
-However, by the construction of $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$.
+Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,v)\not\models\neg B_t(\bar{w})$.
+This is, $v(\bar{w})\in B_t$, but because of the way we constructed $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$.
 Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$.
-This violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
+However, this violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
 \end{proof}
+We generalize the construction of $\bar{B}$ given in the claim.
+For a structure $\A$ and a first-order assignment $v$ for $\A$, then we define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length(\bar{X})})$ where each $B^v_j = \{v(\bar{w})\mid\bar{w}\text{ is a subtuple of $\bar{y}$ and $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
+Now, let $\bar{v}$ be the subtuple of $\bar{y}$ of every first-order variable mentioned in $\varphi_i^{+}$ and let $\bar{u}$ be a subtuple of $(\bar{x},\bar{y})$ such that concatenating $(\bar{v},\bar{u})$ is equal to $(\bar{x},\bar{y})$.
+We also define $v^*$ as a first-order assignment for $\A$ for which we have that $v^{*}(\bar{v})$ is the lexicographically smallest tuple that satisfies $(\A,v^{*})\models\exists\bar{u} \varphi^{\fo}_i(\bar{u},\bar{v})$.
+Lastly, we define $\bar{B}^* = \bar{B}^{v^{*}}$ and $V^*$ as some second-order assignment for $\A$ for which $V^*(\bar{X}) = \bar{B}^*$.
 
-The previous claim motivates the following definitions, in particular, a generalization of the sequence $\bar{B}$. 
-For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length(\bar{X})})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
-Moreover, let $\bar{v}$ be the sequence of all different first-order variable mentioned in $\varphi_i^{+}$ and let $\bar{u}$ be a sequence of variables disjoint from $\bar{v}$ such that $(\bar{v},\bar{u})$ and $(\bar{x},\bar{y})$ have the same set of variables.
-Then define $v^*$ to be the first-order assignment for $\A$ for which we have that $v^{*}(\bar{v})$ is the lexicographically smallest tuple that satisfies $(\A,v^{*})\models\exists\bar{u} \varphi^{\fo}_i(\bar{u},\bar{v})$.
-Indeed, one can define a formula $\mu_i$ that defines $v^*$ as follows:
+We define a formula $\mu_i$ that is satisfied only by $v^*$:
 $$
@@
 $$ 
-In other words, we have that $(\A,v)\models \mu_i(\bar{v})$ if, and only if, $v(\bar{v}) = v^{*}(\bar{v})$ for every assignment $v$. 
-Finally, define $\bar{B}^* = \bar{B}^{v^{*}}$ and $V^*$ as some second-order assignment for $\A$ for which $V^*(\bar{X}) = \bar{B}^*$.
-
-Now, we have all the tools to introduce the main formula of the proof. 
-Intuitively, the pair $(V^*, v^*)$ is the minimum assignment that satisfies $\varphi_i(\bar{X}, \bar{x})$ and our candidate to be removed from the semantics of $\varphi(\bar{X}, \bar{x})$.
-Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ that has the following key property:
-\begin{center}
-if $\varphi_i(\bar{X}, \bar{x}, \bar{y})$ is satisfiable, then $(\A,V,v) \models\varphi_i(\bar{X},\bar{x},\bar{y})$ if, and only if, $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$ for every $(V,v)$.
-\end{center}
-That is, its exclude somehow the pair $(V^*, v^*)$. This property can be defined in logic as follows:
-\begin{align}
-\psi_i(\bar{X},\bar{x}) = & \;\; \big( \, \exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
-& \hspace{1cm} \Big( \; 
-\exists\bar{x}'(\exists\bar{y}'\varphi_i(\bar{X},\bar{x}',\bar{y}')\wedge \bar{x}'<\bar{x}) \;\;  \vee   \label{eq:part2} \\
-& \hspace{1.5cm} \exists\bar{v} (\mu_i(\bar{v})\wedge \big( \varphi^{+}_i(\bar{X},\bar{v}) \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big) \Big)  \label{eq:part3}
-\end{align}
-To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
-Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi_i(\bar{X},\bar{x},\bar{y})$.
-Then, the conclusion of the implication at (\ref{eq:part2}) and (\ref{eq:part3}), take care that the $\varphi_i$ is only satisfied by a pair $(V, v)$ different from $(V^*, v^*)$.
-Here, the first disjunct at (\ref{eq:part2}) check whether $(\A,V,v) \models\varphi_i(\bar{X},\bar{x},\bar{y})$ but $v(\bar{x})$ is not the lexicographically smallest tuple that have this property (i.e.
-$v(\bar{x}) \neq v^*(\bar{x})$).
-The second disjunct at (\ref{eq:part3}) is more involved and it can be explained as follows:
-
 The following is a pivotal formula in our proof:
-
 \begin{align}
@@
 \end{align}
-
 Consider some $\A$.
@@
 This procedure can be easily extended to receive Horn formulas, and furthermore, a disjunction of Horn formulas.
-We can deduce that $\shdhsat$ is in $\totp$.
+Hence $\shdhsat$ is in $\totp$.
 

### 2017-01-12 LICS/appendix-syntactic.tex
length -> |x|

@@
 Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ in $\loge{0}$-PNF that is equivalent to the $\eqso(\loge{0})$ formula $\alpha_0$.
-In \cite{SalujaST95}, it was proved that if $\length(\bar{X}) > 0$, then the function defined by $\alpha$, for big enough structures, is always even which is not possible in our case.
+In \cite{SalujaST95}, it was proved that if $\length{\bar{X}} > 0$, then the function defined by $\alpha$, for big enough structures, is always even which is not possible in our case.
 On the other hand, if $\alpha$ is of the form $\sa{\bar{x}}\varphi(\bar{x})$, then $\alpha$ defines a polynomially bounded function which also leads to a contradiction.
@@
 Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to the $\eqso(\loge{1})$ formula $2$.
-First, if $\length(\bar{X}) = \length(\bar{x}) = 0$, then the function defined by $\alpha$ is never greater than 1. 
-Therefore, suppose that $\length(\bar{X}) > 0$ or $\length(\bar{x}) > 0$, and consider any ordered structure $\A$. 
+First, if $\length{\bar{X}} = \length{\bar{x}} = 0$, then the function defined by $\alpha$ is never greater than 1. 
+Therefore, suppose that $\length{\bar{X}} > 0$ or $\length{\bar{x}} > 0$, and consider any ordered structure $\A$. 
 Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
@@
 Let $\alpha = \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-SNF.
-Without lost of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.
+Without lost of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
 If not, we replace each $\alpha_i$ for the equivalent formula $\sa{\bar{X}} \sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge\forall z\,Y(z) \wedge \forall z(y \leq z))$.
@@
 \begin{align*}
-\gamma_{\text{first}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z}), \\
-\gamma_{\text{last}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x}).
+\gamma_{\text{first}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z}), \\
+\gamma_{\text{last}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x}).
 \end{align*}
@@
 \begin{align*}
-\gamma_{\text{not first}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{z}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{z}_i)), \\
-\gamma_{\text{not last}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{x} < \bar{z}_0 \vee \bigvee_{i = 1}^{\length(\bar{X})}\neg X(\bar{z}_i)),
+\gamma_{\text{not first}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{z}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{z}_i)), \\
+\gamma_{\text{not last}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{x} < \bar{z}_0 \vee \bigvee_{i = 1}^{\length{\bar{X}}}\neg X(\bar{z}_i)),
 \end{align*}
-where $\bar{z} = (\bar{z}_0,\bar{z}_1,\ldots,\bar{z}_{\length(\bar{X})})$. The following formula is equivalent to $\alpha$:
+where $\bar{z} = (\bar{z}_0,\bar{z}_1,\ldots,\bar{z}_{\length{\bar{X}}})$. The following formula is equivalent to $\alpha$:
 \begin{align}
@@
 \begin{align*}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\exists\bar{v}(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i))\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\exists\bar{w}(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))].
+\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\exists\bar{v}(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
+&(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))\vee\\
+&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\exists\bar{w}(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\vee\\
+&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))].
 \end{align*}
@@
 \begin{align*}
-\sa{\bar{X}}\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
-& \hspace{3cm}  \bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
+\sa{\bar{X}}\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
+& \hspace{3cm}  \bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
 &
-(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))\vee
+(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))\vee
 \\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
-& \hspace{3cm} (\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee \\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))].
+&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
+& \hspace{3cm} (\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee \\
+&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))].
 \end{align*}
@@
 \begin{proof}
-Suppose that the function defined by $\alpha$ is not $0$ and that $\varphi$ is in DNF.
-Furthermore, suppose $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$.
+Suppose that the function defined by $\alpha$ is not $0$ and that $\varphi$ is in DNF. Furthermore, suppose $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$.
 Then $\alpha = \sa{\bar{x}} \varphi_1(\bar{x}) \vee \cdots \vee \varphi_n(\bar{x})$.
@@
 This is, the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is not null.
-We will prove by induction on $\length(\bar{x})$ that the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is either greater or equal to $n$, or in $\Omega(n^2)$.
-We address the case $\length(\bar{x})= 1$, then $\alpha = \sa{x}\bigwedge\psi(x)$.
+We will prove by induction on $\length{\bar{x}}$ that the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is either greater or equal to $n$, or in $\Omega(n^2)$.
+We address the case $\length{\bar{x}}= 1$, then $\alpha = \sa{x}\bigwedge\psi(x)$.
 If any $\psi(x) = (x = x)$ or $\neg(x < x)$, then we can eliminate it and we obtain the same function.
@@
 The only possible $\alpha$ left is $\alpha = \sa{x}\top$ which is equal to the function $n$.
-This covers all possible cases for $\length(\bar{x}) = 1$.
-Now suppose that it holds for $\length(\bar{x}) = k$ and suppose $\alpha = \sa{\bar{x}}\bigwedge\psi(\bar{x})$ for $\length(\bar{x}) = k+1$.
+This covers all possible cases for $\length{\bar{x}} = 1$.
+Now suppose that it holds for $\length{\bar{x}} = k$ and suppose $\alpha = \sa{\bar{x}}\bigwedge\psi(\bar{x})$ for $\length{\bar{x}} = k+1$.
 If any $\psi(\bar{x}) = (x_i = x_j)$ where $i \neq j$, then $\alpha$ describes the same function as $\alpha$ where $x_j$ has been replaced by $x_i$.
@@
 If $i = j$, then we can eliminate it and obtain the same function.
-If any $\psi(\bar{x}) = R(\bar{v})$ or $\neg R(\bar{v})$ where $\bar{v}$ is a sub-tuple of $\bar{x}$ then we can either eliminate it or the function becomes null, following the same argument as in the case $\length(\bar{x}) = 1$.
+If any $\psi(\bar{x}) = R(\bar{v})$ or $\neg R(\bar{v})$ where $\bar{v}$ is a sub-tuple of $\bar{x}$ then we can either eliminate it or the function becomes null, following the same argument as in the case $\length{\bar{x}} = 1$.
 If any $\psi(\bar{x}) = \neg(x_i = x_j)$ or $(x_i < x_j)$ where $i = j$, then the function becomes null.
@@
 Consider the formula that describes this order (like $x_1 < x_3 \wedge x_3 < x_4 \wedge x_4 < x_2$).
-The function $\alpha$ describes is greater or equal to the one this formula describes, which is exactly $\binom{n}{\length(\bar{x})}$ which is in $\Omega(n^{\length(\bar{x})}) \subseteq \Omega(n^2)$ if $\length(\bar{x}) > 1$.
+The function $\alpha$ describes is greater or equal to the one this formula describes, which is exactly $\binom{n}{\length{\bar{x}}}$ which is in $\Omega(n^{\length{\bar{x}}}) \subseteq \Omega(n^2)$ if $\length{\bar{x}} > 1$.
 This concludes the proof of the claim.
@@
 Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, where $\varphi$ is quantifier-free.
-Note that if $\length(\bar{X}) > 0$, then the function $\alpha$ describes is in $\Omega(2^n)$, as it was proven by the authors in \cite{SalujaST95}.
+Note that if $\length{\bar{X}} > 0$, then the function $\alpha$ describes is in $\Omega(2^n)$, as it was proven by the authors in \cite{SalujaST95}.
 We have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$, as we proved in the claim, describes either some function greater or equal to $n$, or in $\Omega(n^2)$, which leads to a contradiction.
@@
 $$
-\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot \vert A \vert^{\length(\bar{y})} \geq \sem{\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A)
+\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot \vert A \vert^{\length{\bar{y}}} \geq \sem{\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A)
 $$
 Note that the formula $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ defines a function in $\E{0}$.
-It was shown by the authors in \cite{SalujaST95} that every function in $\E{0}$ grows exponentially over the size of the structure for large enough structures, when $\length(\bar{X}) > 0$.
+It was shown by the authors in \cite{SalujaST95} that every function in $\E{0}$ grows exponentially over the size of the structure for large enough structures, when $\length{\bar{X}} > 0$.
 This function divided by a polynomial factor still grows exponentially.
-Therefore, for $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ we have that $\length(\bar{X}) = 0$.
+Therefore, for $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ we have that $\length{\bar{X}} = 0$.
 
@@
 The signature $\R_{\alpha}$ is obtained by adding the symbol $R_{\psi}$, for every $\fo$ formula $\psi(\bar{z})$ in $\alpha$, to $\R$.
-Each symbol $R_{\psi}$ represents a predicate with arity $\length(\bar{z})$.
+Each symbol $R_{\psi}$ represents a predicate with arity $\length{\bar{z}}$.
 Then, $\lambda(\alpha)$ is defined as $\alpha$ where each $\fo$ formula $\psi(\bar{z})$ has been replaced by $R_{\psi}(\bar{z})$.
@@
 Let $\A'$ be an $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A}$ empty.
-For each $\fo$-formula $\psi(\z)$ with $\length(\bar{z})$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length(\bar{z})}$.
+For each $\fo$-formula $\psi(\z)$ with $\length{\bar{z}}$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length{\bar{z}}}$.
 If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$ (this can be done in $\ptime$).
@@
 Without lost of generality, we assume that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
-Moreover, we assume that $\length(\bar{x}) > 0$  since, if this is not the case, we can replace $\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
+Moreover, we assume that $\length{\bar{x}} > 0$  since, if this is not the case, we can replace $\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
 
@@
 Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}} \varphi(\bar{x})$ (i.e.
-$\length(\bar{X}) = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}} [\varphi(\bar{x})\wedge\exists\bar{z}(\varphi(\bar{z})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
-Therefore, for the rest of the proof we can assume that $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.
+$\length{\bar{X}} = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}} [\varphi(\bar{x})\wedge\exists\bar{z}(\varphi(\bar{z})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
+Therefore, for the rest of the proof we can assume that $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
 
- Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
+ Let $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$ and $\bar{X} = (X_1,\ldots,X_{\length{\bar{X}}})$.
 To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in into a DNF formula. 
@@
 	\item[(a)] In order to satisfy the first condition, consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
-We add $\length(\bar{w})$ new variables $z_1,\ldots,z_{\length(\bar{w})}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length(\bar{w})})$.
+We add $\length{\bar{w}}$ new variables $z_1,\ldots,z_{\length{\bar{w}}}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length{\bar{w}}})$.
 We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$.
@@
 On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
-	For a given $k$, let $\cB_k$ be the number of possible ordered partitions for a set of size $k$. For $1 \leq j \leq \cB_{\length(\bar{x},\bar{y})}$ 
+	For a given $k$, let $\cB_k$ be the number of possible ordered partitions for a set of size $k$. For $1 \leq j \leq \cB_{\length{(\bar{x},\bar{y})}}$ 
 	let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
@@
 	$$
-	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}} \exists \bar{y} \; [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
+	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length{(\bar{x},\bar{y})}}} \exists \bar{y} \; [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
 	$$
@@
 Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
-Define $\bar{B} = (B_1,\ldots,B_{\length(\bar{X})})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
+Define $\bar{B} = (B_1,\ldots,B_{\length{\bar{X}}})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
 Towards a contradiction, suppose that $(\A,V,v)\not\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
@@
 The previous claim and proof motivates the following definitions.
-For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length(\bar{X})})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
+For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length{\bar{X}}})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
 One can easily check that for every assignments $V, v$ such that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$, then it holds that $(\A,\bar{B}^v,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all the assignment of $\bar{X}$ when $v$ is fixed.
@@
 $$
-where $n = \length(\bar{x}) + \length(\bar{y}) + 1$.
+where $n = \length{\bar{x}} + \length{\bar{y}} + 1$.
 Let $\A$ be the encoding of this formula.
@@
 To simplify the proof, we extend our grammar to allow first-order constants.
-Consider each tuple $\bar{a}\in A^{\length(\bar{x})}$, each $\bar{b}\in A^{\length(\bar{y})}$ and each $\bar{c}\in A^{\length(\bar{z})}$ as a tuple of first-order constants.
+Consider each tuple $\bar{a}\in A^{\length{\bar{x}}}$, each $\bar{b}\in A^{\length{\bar{y}}}$ and each $\bar{c}\in A^{\length{\bar{z}}}$ as a tuple of first-order constants.
 The following formula defines the same function as $\alpha$:
 $$
-\sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length(\bar{x})}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length(\bar{y})}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length(\bar{z})}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
+\sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length{\bar{x}}}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length{\bar{y}}}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length{\bar{z}}}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
 $$

### 2017-01-12 LICS/appendix-syntactic.tex
Version nueva.

@@
 We define a formula $\varphi'$ that satisfies the following condition: if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ and $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be proven.
+From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be~proven.
 
@@
 
- Let $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$ and $\bar{X} = (X_1,\ldots,X_{\length{\bar{X}}})$.
-To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in into a DNF formula. 
-More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i$ is a conjunction of atomic formulas or negation of atomic formulas.
-Furthermore, we suppose that each conjunct $\varphi_i$ has the~form:
+To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in a DNF formula. 
+More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i(\bar{X}, \bar{x}) = \exists \bar{y} \, \varphi_i'(\bar{X}, \bar{x}, \bar{y})$ and $\varphi_i'(\bar{X}, \bar{x}, \bar{y})$ is a conjunction of atomic formulas or negation of atomic formulas.
+Furthermore, we suppose that each conjunct $\varphi_i'(\bar{X}, \bar{x}, \bar{y})$ has the~form:
 $$
-\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
+\varphi_i'(\bar{X}, \bar{x}, \bar{y}) =  \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
 \underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}} \wedge
@@
 $$
-Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ can appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
+Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ will appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
 
@@
 	Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
-	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  to $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
+	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  as $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
 	\item[(c)] For showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
@@
 It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$.
-From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c).
-Furthermore, it has the following form:
+From now on, we assume that each $\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \;  \varphi_i'(\bar{X},\bar{x}, \bar{y}) $ satisfies conditions (a), (b) and (c), and $\varphi_i'(\bar{X},\bar{x}, \bar{y})$ has the following~form:
 $$
-\varphi_i(\bar{X},\bar{x}) \; = \; \exists{\bar{y}} \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
+\varphi_i'(\bar{X},\bar{x}, \bar{y}) \; = \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
 $$
@@
 \begin{claim}\label{claim:minusone}
-For a given ordered structure $\A$, if there exists a first-order assignment $v$ for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists a second-order assignment $V$ for $\A$ such that  $(\A,V,v)\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
+For a given ordered structure $\A$ and a first-order assignment $v$ for $\A$, $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ if, and only if, there exists a second-order assignment $V$ for $\A$ such that $(\A,V,v)\models \varphi_i'(\bar{X},\bar{x}, \bar{y})$.
 \end{claim}
 \begin{proof}
-Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
+Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$, such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
 Define $\bar{B} = (B_1,\ldots,B_{\length{\bar{X}}})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
@@
 This violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
+
+For the other direction, if $(\A,V,v)\models \varphi_i'(\bar{X},\bar{x}, \bar{y})$ for a second order assignment $V$ for $\A$, then it is easy to check $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ since $\varphi^{\fo}_i(\bar{x},\bar{y})$ is a subformula of $\varphi_i'$.
 \end{proof}
@@
 For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length{\bar{X}}})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
-One can easily check that for every assignments $V, v$ such that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$, then it holds that $(\A,\bar{B}^v,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all the assignment of $\bar{X}$ when $v$ is fixed.
+One can easily check that for every assignments $(V, v)$ such that $(\A,V,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$, it holds that $(\A,\bar{B}^v,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all assignments of $\bar{X}$ when $v$ is fixed.
 Therefore, by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
@@
 If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
-Furthermore, let $V^*$ the second order assignment that satisfies $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*$ the first order assignment such that $v^*(\bar{x}) = v(\bar{x})$.
+Furthermore, let $V^*$ the second order assignment and $v^*$ the first order assignment that satisfy $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*(\bar{x}) = v(\bar{x})$.
 By the previous discussion, we have that $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.
 
-Now, we have all the tools to introduce the main formula of the proof. 
-Intuitively, we want to exclude the assignment $(V^*, v^*)$ as a satisfying assignment for $\varphi_i(\bar{X},\bar{x})$.
+Now, we have all the tools to introduce one of the main formulas in order to define $\kappa(\alpha)$. 
+Intuitively, we want to exclude the assignment $(V^*, v^*)$ from the satisfying assignments of $\varphi_i(\bar{X},\bar{x})$.
 Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if, 
-if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, then $V \neq V^*$ or $v \neq v^*$. 
-That is, it excludes all pairs $(V, v)$ that coincide with $(V^*, v^*)$ in $(\bar{X}, \bar{x})$ whenever $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. This property can be defined as follows:
+$V \neq V^*$ or $v \neq v^*$ whenever $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
+This property can be defined as follows:
 \begin{align}
@@
 & \hspace{1cm} \Big( \; 
-\exists\bar{y} \;\; \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i^+(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\
-& \hspace{1.5cm} (\exists\bar{x}'\exists\bar{y}\, \varphi_i(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}) \;\; \Big)  \label{eq:part3}
+\exists\bar{y}~\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i'(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\
+& \hspace{1.5cm} (\; \exists\bar{x}'\exists\bar{y}\; \varphi_i'(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x} \; ) \;\; \Big)  \label{eq:part3}
 \end{align}
 To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
-Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-Then, the conclusion of the implication at (\ref{eq:part2}) and (\ref{eq:part3}), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
-Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by showing that if $\varphi_i(\bar{X},\bar{x})$ is satisfied then $V(\bar{X})$ must contain one element more that it is not in $V^*(\bar{X})$.
+Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi_i'(\bar{X},\bar{x},\bar{y})$.
+Then, the conclusion of the implication (divided into (\ref{eq:part2}) and (\ref{eq:part3})), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
+Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by defining that if $\varphi_i'(\bar{X},\bar{x}, \bar{y})$ is satisfied then $V^*(\bar{X}) \subsetneq V(\bar{X})$. 
 The second disjunct (\ref{eq:part3}) is satisfied when $v(\bar{x})$ is not the lexicographically smallest tuple that satisfies $\varphi_i$ (i.e. $v(\bar{x}) \neq v^*(\bar{x})$).
+Finally, from the previous discussion one can easily check that $\psi_i(\bar{X},\bar{x})$ satisfies the desire property.
 
-We are ready to define the formula  $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$
-where each modified disjunct $\varphi_i'(\bar{X},\bar{x})$ is constructed as follows. 
+We are ready to define the formula  $\kappa(\alpha)$ as $\sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i^*(\bar{X},\bar{x})$
+where each modified disjunct $\varphi_i^*(\bar{X},\bar{x})$ is constructed as follows. 
 For the sake of simplification, define the auxiliary formula $\chi_i = \neg\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$. 
-This formula basically checks if $\varphi_i$ is satisfiable or not.
-Then, first define the formula:
+This formula basically checks if $\varphi_i$ is not satisfiable (recall Claim~\ref{claim:minusone}).
+Define the first formula $\varphi_1^*$ as:
 $$
-\varphi'_1(\bar{X},\bar{x}) = \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
+\varphi^*_1(\bar{X},\bar{x}) = \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
 $$
-This models that all the assignments that satisfy $\varphi_1$ also satisfy $\varphi'_1$, except for $(V^*,v^*)$. Then we define:
+This formula accepts all the assignments that satisfy $\varphi_1$, except for the assignment $(V^*,v^*)$ of $\varphi_1$. The second formula $\varphi_2^*$ is defined as:
 $$
-\varphi'_2(\bar{X},\bar{x}) = \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
+\varphi^*_2(\bar{X},\bar{x}) = \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
 $$
-This models that all the assignments that satisfy $\varphi_2$ also satisfy $\varphi'_2$, but if there was no assignment for $\varphi_1$, then we exclude $(V^*,v^*)$ for $\varphi_2$.
-Then we generalize this construction for each $\varphi_i$:
+This models all the assignments that satisfy $\varphi_2$, except for the assignment $(V^*,v^*)$ of $\varphi_1$.
+Moreover, if $\varphi_1$ is not satisfiable, then $\psi_1(\bar{X},\bar{x})$ and $\chi_1$ will hold, and the formula $\psi_2(\bar{X},\bar{x})$ will forbid the assignment $(V^*,v^*)$ of $\varphi_2$.
+One can easily generalize this construction for each $\varphi_i$ as follows:
 $$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
+\varphi_i^*(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
 \bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})),
 $$
-By the construction of $\kappa(\alpha)$, one can easily check that  $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.
+From the construction of $\kappa(\alpha)$, one can easily check that  $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.
 
@@
 {\em Part (2).} Let $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ for some $\eqso(\logex{1})$ formula $\beta$. We define $\kappa(\alpha)$ as follows.
-First, perform the same transformations to $\varphi(\bar{X},\bar{x})$ as in the previous section.
+First, rewrite $\varphi(\bar{X},\bar{x})$ as in the previous section.
 Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c).
-We also use the previously defined formulas $\chi_i$ and $\psi_i$, for each $i$.
+Also, consider the previously defined formulas $\chi_i$ and $\psi_i$, for each $i$.
 
-We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if and only if $\sem{\beta}(\A) = 0$.
+We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if, and only if, $\sem{\beta}(\A) = 0$.
 If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg\exists\bar{x}'\varphi(\bar{x}')$.
-If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$ as each $\chi_i$ was previously defined.
+If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and define $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$.
 If $\beta = (\beta_1 + \beta_2)$, then $\lambda(\beta) = \lambda(\beta_1) \wedge \lambda(\beta_2)$.
@@
 $$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
+\varphi_i^*(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
 \bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x}))\Big)\Big).
 $$
-Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
+Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i^*(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
 

### 2017-03-03 LMCS/appendix-counting.tex
journal primera version

--- 2017-01-09 LICS/appendix-counting.tex
@@
 Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time \cite{fagin1974generalized}, and accepts if and only if the condition holds true.
+If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time \cite{F75}, and accepts if and only if the condition holds true. 
 If $\alpha$ is a constant $s$, then the machine produces $s$ branches and accepts in all of them. 

### 2017-03-03 LMCS/appendix-syntactic.tex
journal primera version

--- 2017-01-12 LICS/appendix-syntactic.tex
@@
-\subsection*{Proof of Theorem \ref{theo-pnf-snf}}
 Recall that a formula in $\eqso(\LL)$ is on the following grammar:
@@
 This covers all possible cases for $\alpha$ and we conclude the proof by taking $\tau(\alpha)$ as the desired rewrite of $\alpha$.
-
-\subsection*{Proof of theorem \ref{theo-pi1-pnf}}
-We divide the proof in three parts.
-
-\vspace{1em}
-First, we prove that the formula $\alpha_{0} = \left( \sa{X} 1 \right) + 1$ with $\arity(X) = 1$ (i.e.
-the function $2^{n}+1$, where $n$ is the size of the input structure) is not equivalent to any formula in $\loge{0}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ in $\loge{0}$-PNF that is equivalent to the $\eqso(\loge{0})$ formula $\alpha_0$.
-In \cite{SalujaST95}, it was proved that if $\length(\bar{X}) > 0$, then the function defined by $\alpha$, for big enough structures, is always even which is not possible in our case.
-On the other hand, if $\alpha$ is of the form $\sa{\bar{x}}\varphi(\bar{x})$, then $\alpha$ defines a polynomially bounded function which also leads to a contradiction.
-
-\vspace{1em}
-Second, we prove that the formula $\alpha_{1} = 2$ (i.e.
-the constant $2$) is not equivalent to any formula in $\loge{1}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to the $\eqso(\loge{1})$ formula $2$.
-First, if $\length(\bar{X}) = \length(\bar{x}) = 0$, then the function defined by $\alpha$ is never greater than 1. 
-Therefore, suppose that $\length(\bar{X}) > 0$ or $\length(\bar{x}) > 0$, and consider any ordered structure $\A$. 
-Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
-Now consider the ordered structure $\A'$ that is obtained by duplicating $\A$.
-This is, each half of $\A'$ is isomorphic to $\A$.
-Note that $\A'\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$ for $i = 1,2$ and there exists a third assignments $(\bar{B}_1',\bar{b}_1',\bar{a}_1')$ that is isomorphic to $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$ but in the other half to the structure such that $\A'\models\varphi(\bar{B}_1',\bar{b}_1',\bar{a}_1')$.
-We have that $\sem{\alpha}(\A) \geq 3$ which is a contradiction.
-
-\vspace{1em}
-We now show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there is an equivalent formula $\beta$ in $\LL$-PNF.
-As in Theorem \ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such formula.
-As we showed, there exists an equivalent formula in $\LL$-SNF, so we assume that $\alpha$ is in that form.
-Let $\alpha = \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-SNF.
-Without lost of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.
-If not, we replace each $\alpha_i$ for the equivalent formula $\sa{\bar{X}} \sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge\forall z\,Y(z) \wedge \forall z(y \leq z))$.
-
-Now we begin describing the function $\tau$.
-If $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then the formula is already in $\LL$-PNF so we define $\tau(\alpha) = \alpha$.
-If $\alpha = \alpha_1 + \alpha_2$, then we assume that $\tau(\alpha_1) = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ and $\tau(\alpha_2) = \sa{\bar{Y}}\sa{\bar{y}}\psi(\bar{Y},\bar{y})$.
-The construction that we will provide for this function works by identifying a ``first'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ and a ``last'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$.
-These are identified by the following formulas:
-\begin{align*}
-\gamma_{\text{first}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z}), \\
-\gamma_{\text{last}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x}).
-\end{align*}
-Similarly, we define the formulas $\gamma_{\text{first}}(\bar{Y},\bar{y})$ and $\gamma_{\text{last}}(\bar{Y},\bar{y})$ (for the sake of simplicity we reuse the names $\gamma_{\text{first}}$ and $\gamma_{\text{last}}$).
-In other words, the ``first'' assignment is the one where every second-order predicate is empty and the first-order assignment is the lexicographically smallest, and the ``last'' assignment is the one where every second-order predicate is full and the first-order assignment is the lexicographically greatest.
-We also need to identify assignments that are not first and are not last.
-We do this by negating the two formulas above and grouping together the first-order variables:
-\begin{align*}
-\gamma_{\text{not first}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{z}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{z}_i)), \\
-\gamma_{\text{not last}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{x} < \bar{z}_0 \vee \bigvee_{i = 1}^{\length(\bar{X})}\neg X(\bar{z}_i)),
-\end{align*}
-where $\bar{z} = (\bar{z}_0,\bar{z}_1,\ldots,\bar{z}_{\length(\bar{X})})$. The following formula is equivalent to $\alpha$:
-\begin{align}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{not first}}(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{Y},\bar{y}))\vee \label{eq:partition1} \\
-&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))\vee \label{eq:partition2}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{not last}}(\bar{Y},\bar{y}))\vee \label{eq:partition3}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{last}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))]. \label{eq:partition4}
-\end{align}
-To show that the formula is indeed equivalent to $\alpha$, note that the formulas in lines (\ref{eq:partition1}) and (\ref{eq:partition2}) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$, and the formulas in lines (\ref{eq:partition3}) and (\ref{eq:partition4}) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
-Altogether, the four lines define pairwise disjoint assignments for $(\bar{X},\bar{x}),(\bar{Y},\bar{y})$.
-With this, it is straightforward to show that the above formula is equivalent to $\alpha$.
-However, the formula is not yet in the correct form since it has existential quantifiers in the subformulas $\gamma_{\text{not first}}$ and $\gamma_{\text{not last}}$.
-To solve this, first let take a close look to the complete formula:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\exists\bar{v}(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i))\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\exists\bar{w}(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))].
-\end{align*}
-To construct an equivalent formula that is in the correct form, we define $\bar{u} = (\bar{v},\bar{w})$ and we replace the first-order quantifiers by a first-sum and count the first assignment to $\bar{v}$ and $\bar{w}$ that satisfies the formula.
-A similar construction was used in \cite{SalujaST95}.
-Then the final formula equivalent to $\alpha$ is the following:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
-& \hspace{3cm}  \bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&
-(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))\vee
-\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
-& \hspace{3cm} (\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee \\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))].
-\end{align*}
-Finally, consider a $\eqso(\LL)$ formula $\alpha$ in $\LL$-SNF.
-If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we can consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the condition in the hypothesis.
-
-\subsection*{Proof of Theorem \ref{prop-rest}}
-We give this proof in three parts.
-
-\vspace{1em}
-First, we show that $\QE{0} \not\subseteq \E{1}$.
-By contradiction, suppose that there is a $\QE{0}$ formula $\alpha$ over some signature $\R$ such that defines the following function.
-For every finite $\R$-structure with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n - 1$.
-We use the following claim.
-\begin{claim}
-Let $\alpha = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. Then the function defined by $\alpha$ is either null, greater or equal to $n$, or is in $\Omega(n^2)$.
-\end{claim}
-\begin{proof}
-Suppose that the function defined by $\alpha$ is not $0$ and that $\varphi$ is in DNF.
-Furthermore, suppose $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$.
-Then $\alpha = \sa{\bar{x}} \varphi_1(\bar{x}) \vee \cdots \vee \varphi_n(\bar{x})$.
-Since $\alpha$ is not null, then some $\varphi_i$ must be satisfiable.
-This is, the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is not null.
-We will prove by induction on $\length(\bar{x})$ that the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is either greater or equal to $n$, or in $\Omega(n^2)$.
-We address the case $\length(\bar{x})= 1$, then $\alpha = \sa{x}\bigwedge\psi(x)$.
-If any $\psi(x) = (x = x)$ or $\neg(x < x)$, then we can eliminate it and we obtain the same function.
-If any $\psi = (x < x)$ or $\neg(x=x)$, then the function becomes null.
-If $\psi(x) = R(x,\ldots,x)$ for some $R\in\R$ the function becomes null for the structures we are considering.
-If $\psi(x) = \neg R(x,\ldots,x)$, we can eliminate it and for the structures we are considering we obtain the same function.
-The only possible $\alpha$ left is $\alpha = \sa{x}\top$ which is equal to the function $n$.
-This covers all possible cases for $\length(\bar{x}) = 1$.
-Now suppose that it holds for $\length(\bar{x}) = k$ and suppose $\alpha = \sa{\bar{x}}\bigwedge\psi(\bar{x})$ for $\length(\bar{x}) = k+1$.
-If any $\psi(\bar{x}) = (x_i = x_j)$ where $i \neq j$, then $\alpha$ describes the same function as $\alpha$ where $x_j$ has been replaced by $x_i$.
-In this formula the tuple of first-order variables has $k$ elements so the function it describes if one of the mentioned in the hypothesis.
-If $i = j$, then we can eliminate it and obtain the same function.
-If any $\psi(\bar{x}) = R(\bar{v})$ or $\neg R(\bar{v})$ where $\bar{v}$ is a sub-tuple of $\bar{x}$ then we can either eliminate it or the function becomes null, following the same argument as in the case $\length(\bar{x}) = 1$.
-If any $\psi(\bar{x}) = \neg(x_i = x_j)$ or $(x_i < x_j)$ where $i = j$, then the function becomes null.
-If any $\psi(\bar{x}) = \neg(x_i < x_j)$ where $i = j$, we can eliminate it.
-The remaining formulas in $\bigwedge\psi(\bar{x})$ are either $\neg(x_i = x_j)$, $(x_i<x_j)$ or $\neg(x_i<x_j)$.
-If the formula violates transitivity in $<$ (for example, $x < y \wedge y < z \wedge z < x$), then the function $\alpha$ describes is null.
-Therefore, there is some order over $\bar{x}$ that satisfies $\bigwedge\psi(\bar{x})$.
-Consider the formula that describes this order (like $x_1 < x_3 \wedge x_3 < x_4 \wedge x_4 < x_2$).
-The function $\alpha$ describes is greater or equal to the one this formula describes, which is exactly $\binom{n}{\length(\bar{x})}$ which is in $\Omega(n^{\length(\bar{x})}) \subseteq \Omega(n^2)$ if $\length(\bar{x}) > 1$.
-This concludes the proof of the claim.
-\end{proof}
-We suppose that $\alpha$ is in SNF, this is, $\alpha = \sum_{i = 1}^n\alpha_i$.
-Since $\alpha$ is not null, consider some $\alpha_i$ that describes a non-null function.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, where $\varphi$ is quantifier-free.
-Note that if $\length(\bar{X}) > 0$, then the function $\alpha$ describes is in $\Omega(2^n)$, as it was proven by the authors in \cite{SalujaST95}.
-We have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$, as we proved in the claim, describes either some function greater or equal to $n$, or in $\Omega(n^2)$, which leads to a contradiction.
-Lastly, note that the formula $\sa{x}\exists y(x < y)$ is in $\E{0}$ and describes the function $n-1$, which concludes the proof.
-
-\vspace{1em}
-Now we show that $\E{1}\not\subseteq\QE{0}$.
-In Theorem \ref{theo-pi1-pnf} we proved that there is no formula in $\loge{1}$-PNF equivalent to the formula $\alpha = 2$.
-Every formula in $\E{1}$ can be expressed in $\loge{1}$-PNF, which implies that $2 \in \QE{0}$ and $2 \not\in \E{1}$.
-
-\vspace{1em}
-Lastly, we prove that $\eqso(\loge{1})\subsetneq\eqso(\logu{1})$.
-For inclusion, let $\alpha$ be a formula in $\eqso(\loge{1})$.
-Suppose that it is in $\loge{1}$-SNF.
-This is, $\alpha = c + \sum_{i = 1}^{n}\alpha_i$.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, where $\varphi_i$ is quantifier-free, for each $\alpha_i$.
-We use the same construction used in \cite{SalujaST95}, and we obtain that the formula $\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \forall\bar{y}'(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')$ for every assignment to $(\bar{X},\bar{x})$.
-We do this replacement for each $\alpha_i$ and we obtain an equivalent formula in $\eqso(\logu{1})$.
-
-To prove that the inclusion is proper, consider the $\eqso(\logu{1})$ formula $\sa{x}\forall y(y = x)$. This formula defines the following function that takes an ordered structure $\A$ as input:
-$$
-\sem{\alpha}(\A) = 
-\begin{cases}
-1 &\A \text{ has one element}\\
-0 &\text{ otherwise}.
-\end{cases}
-$$
-Suppose that there exists an equivalent formula $\alpha$ in $\eqso(\loge{1})$.
-Also, suppose that it is in $\L$-PNF, so $\alpha = c + \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$.
-Since $\alpha$ takes the value 0 for some structures, $c$ must be 0.
-Consider a structure $\mathfrak{1}$ with one element.
-We have that for some $i$, there exists an assignment $(\bar{B},\bar{b},\bar{a})$ for $(\bar{X},\bar{x},\bar{y})$ such that $\mathfrak{1}\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
-Consider now the structure $\mathfrak{2}$ that is obtained by duplicating $\mathfrak{1}$, as we did for Theorem \ref{theo-pi1-pnf}.
-Note that $\mathfrak{2}\models\varphi_i(\bar{B},\bar{b},\bar{a})$, which implies that $\sem{\alpha}(\mathfrak{2}) \geq 1$, which leads to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop-e1-nc}}
-
-Towards a contradiction, assume that the statement is false.
-This is, that $\E{1}$ is closed under binary sum.
-Consider the formula $\sa{x}(x = x)$ which is in $\E{1}$ over some signature $\R$.
-For every finite $\R$-structure $\A$ with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n$.
-From our assumption, there exists some formula in $\E{1}$ equivalent to the formula $\alpha \add \alpha$, which describes the function $2n$.
-Let $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ be this formula, where $\varphi$ is quantifier-free.
-Note that the function defined by this formula is equal or greater than the one defined by $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ divided by a polynomial factor.
-More specifically, for each ordered structure $\A$ with domain $A$, we have the following inequality:
-$$
-\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot \vert A \vert^{\length(\bar{y})} \geq \sem{\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A)
-$$
-Note that the formula $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ defines a function in $\E{0}$.
-It was shown by the authors in \cite{SalujaST95} that every function in $\E{0}$ grows exponentially over the size of the structure for large enough structures, when $\length(\bar{X}) > 0$.
-This function divided by a polynomial factor still grows exponentially.
-Therefore, for $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ we have that $\length(\bar{X}) = 0$.
-
-Now, for the formula $\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ consider a structure $\mathfrak{1}$ with only one element $a$.
-We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$, but the only possible assignment to $\bar{x}$ is the tuple $(a,\ldots,a)$ so $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which follows to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop:qe0-fp-qe1-totp-fptras}}
-
-The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$.
-This is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every finite $\R$-structure $\A$ with domain $A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$.
-We base our proof on this fact.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is contained in \textsc{TotP}.} Let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Consider the following nondeterministic procedure that on input $\enc(\A)$ generates $\sem{\alpha}(\A)$ branches.
-For each $\alpha_i = \varphi$, where $\varphi$ is a $\loge{1}$ formula, it checks if $\A\models\varphi$ in polynomial time and generates a new branch if that is the case.
-For each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi$, this formula is also in $\E{1}$.
-We use the reduction to $\cdnf$ provided in \cite{SalujaST95} and we obtain $g(\enc(\A))$, which is an instance to $\cdnf$.
-Since $\cdnf$ is also in $\totp$ \cite{PagourtzisZ06}, we simulate the corresponding nondeterministic procedure that generates exactly $\cdnf(\enc(g(\A)))$ branches.
-Since, $\fp\subseteq\totp$\cite{PagourtzisZ06}, each polynomially computable function is also in $\totp$, and then on each of these branches we simulate the corresponding nondeterministic procedure to generate $h(\vert A \vert)$ more.
-The number of branches for each $\alpha_i$ is $\sem{\alpha_i}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$, and the total number of branches in the procedure amounts to $\sem{\alpha}(\A)$.
-We conclude that $\alpha\in\totp$.
-
-\vspace{1em}
-{\bf Every function in} $\boldsymbol{\eqso(\loge{1})}$ {\bf has an FPRAS.}  let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Note that each $\alpha_i$ that is equal to some $\loge{1}$ formula $\varphi$ has an FPRAS given by the procedure that simply checks if $\A\models\varphi$ and returns 1 if it does and 0 otherwise.
-Also, each remaining $\alpha_i$ has an FPRAS since $\alpha_i\in \E{1}$ \cite{SalujaST95}.
-If two functions have an FPRAS, then their sum also has one given by the procedure that simulates them both and sums the results.
-We conclude that $\alpha$ has an FPRAS.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is closed under sum and multiplication.} Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that the class is closed under multiplication.
-We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of $\so$.
-
-\begin{lemma} \label{conj-mult}
-If $\LL$ is a fragment closed under conjunction, then $\eqso(\LL)$ is closed under binary multiplication.
-\end{lemma}
-\begin{proof}
-We will define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ extended by binary product, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
-In fact, the formula $\tau(\alpha)$ is in $\LL$-SNF.
-First we replace each constant $s$ in $\alpha$ for $(\top \add \cdots \add \top)$ ($s$ times).
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$.
-We assume that for every $\beta$ that has less algebraic operators than $\alpha$, $\tau(\beta)$ is in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$ then we define $\tau(\alpha) = \tau(\alpha_1) + \tau(\alpha_2)$.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$, then we define $\tau(\alpha)$ as the formula in $\LL$-SNF that is equivalent to $\sa{x}\tau(\beta)$ and to $\tau(\alpha) = \sa{X}\tau(\beta)$, respectively.
-If $\alpha = (\alpha_1 \cdot \alpha_2)$, we assume that each $\alpha_i$ is in $\LL$-SNF.
-We identify three cases.
-(1) Some $\alpha_i$ is equal to $\sum_{j = 1}^n\beta_j$ for $n > 1$.
-Suppose wlog.
-that it is $\alpha_1$.
-We then define $\tau(\alpha) = \sum_{j = 1}^n\tau(\beta_j\cdot\alpha_2)$.
-In the following cases, $\alpha_1$ and $\alpha_2$ are in $\LL$-SNF.
-(2) If some $\alpha_i$ is equal to $\sa{X}\beta$ or $\sa{x}\beta$, we define $\tau(\alpha)$ as the $\LL$-SNF formula that is equivalent to $\sa{x}\tau(\beta\cdot\alpha_2)$ and $\sa{X}\tau(\beta\cdot\alpha_2)$, respectively.
-The remaining case is (3) $\alpha_1 = \varphi_1$ and $\alpha_2 = \varphi_2$ where each $\varphi$ is an $\LL$ formula.
-Then we define $\tau(\alpha) = \varphi_1 \wedge \varphi_2$.
-This covers all possible cases for $\alpha$.
-For every pair of formulas $\alpha,\beta$ in $\eqso(\LL)$, we have that their multiplication $(\alpha\cdot\beta)$ is a formula in the grammar $\eqso(\LL)$ extended by binary product, and so, there exists an equivalent formula $\tau(\alpha\cdot\beta)$ which is in unextended $\eqso(\LL)$.
-\end{proof}
-Since $\loge{1}$ is closed under conjunction, this also holds for $\eqso(\loge{1})$. This concludes the proof.
-
-\subsection*{Proof of Proposition \ref{pi-minusone}}
-
-Let $\LL$ be a fragment of $\fo$ that contains $\logu{1}$.
-Then we have that every function in $\U{1}$ is expressible in $\eqso(\LL)$.
-In particular, $\ctcnf \in \eqso(\LL)$.
-Suppose that $\eqso(\LL)$ is closed under subtraction by one.
-Then, the function $\ctcnf-1$, which counts the number of satisfying assignments of a 3-CNF formula minus one, is also in $\eqso(\LL)$.
-Note that $\eqso(\LL) \subseteq \eqso(\fo) = \shp$.
-We have that $\ctcnf$ is $\shp$-complete under parsimonious reductions\footnote{It can be easily verified that the standard reduction from SAT to 3-CNF (or 3-SAT) preserves the number of satisfying assignments}.
-Now, let $f$ be a function in $\shp$, and consider the nondeterministic polynomial-time procedure that on input $\enc(\A)$ computes the corresponding reduction to $\ctcnf$, name it $g(\enc(\A))$, and simulates the $\shp$ procedure for $\ctcnf-1$ on input $g(\enc(\A))$.
-We have that this is a $\shp$ procedure that computes $f-1$, from which we conclude that $\shp$ is closed under subtraction by one.
-
-\subsection*{Proof of Theorem \ref{sigmafo-minusone}}
-
-\noindent {\bf Closed under sum and multiplication.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
-This can be seen because $\eqso(\LL)$ is closed under sum for every fragment $\LL$ by definition, and since $\logex{1}$ is closed under conjunction, from Lemma \ref{conj-mult} it follows that $\eqso(\logex{1})$ is closed under multiplication.
-
-\vspace{1em}
-\noindent {\bf Easy decision version and FRPRAS.}  We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
-We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Theorem \ref{prop:qe0-fp-qe1-totp-fptras}.
-First, we define a function that converts a formula $\alpha$ in $\eqso(\logex{1})$ over a signature $\R$ into a formula $\lambda(\alpha)$ in $\eqso(\loge{1})$ over a signature $\R_{\alpha}$.
-Afterwards, we define a function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an $\R_{\alpha}$-structure $g_{\alpha}(\A)$.
-
-Let $\alpha$ be in $\eqso(\logex{0})$.
-The signature $\R_{\alpha}$ is obtained by adding the symbol $R_{\psi}$, for every $\fo$ formula $\psi(\bar{z})$ in $\alpha$, to $\R$.
-Each symbol $R_{\psi}$ represents a predicate with arity $\length(\bar{z})$.
-Then, $\lambda(\alpha)$ is defined as $\alpha$ where each $\fo$ formula $\psi(\bar{z})$ has been replaced by $R_{\psi}(\bar{z})$.
-We now define the function $g_{\alpha}$ procedurally.
-Let $\A$ be a $\R$-structure with domain $A$.
-Let $\A'$ be an $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A}$ empty.
-For each $\fo$-formula $\psi(\z)$ with $\length(\bar{z})$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length(\bar{z})}$.
-If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$ (this can be done in $\ptime$).
-This concludes the construction of $\A'$.
-Note that the number of $\fo$ subformulas, arity and tuple size is fixed in $\alpha$, so computing this function takes polynomial time over the size of the structure.
-Moreover, the encoding of $\A'$ has polynomial size over the size of $\enc(\A)$.
-We define $g_{\alpha}(\A) = \A'$ and we have that for each $\R$-structure $\A$: $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$.
-Therefore, we have a parsimonious reduction from $\alpha$ to the $\eqso(\loge{1})$ formula $\lambda(\alpha)$.
-
-To show that $\alpha$ is in $\totp$, we can convert $\alpha$ and $\enc(\A)$ into $\lambda(\alpha)$ and $\enc(g_{\alpha}(\A))$, respectively, and run the procedure in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-Similarly, to show that $\alpha$ has an FPRAS, we do the same as before and simulates the FPRAS for $\lambda(\alpha)$ in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-These procedures also takes polynomial time and satisfies the condition.
-
-\vspace{1em}
-\noindent {\bf Closed under subtraction by one.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. 
-For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Without lost of generality, we assume that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
-Moreover, we assume that $\length(\bar{x}) > 0$  since, if this is not the case, we can replace $\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
-
-The proof will be separated in two parts.
-In the first part, we suppose that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$.
-Then we will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ is the formula.
-In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ the sum of one or more formulas in $\logex{1}$-PNF.
-We define a formula $\varphi'$ that satisfies the following condition: if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ and $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be proven.
-
-\vspace{1em}
-
-\noindent {\em Part (1).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ where $\varphi$ is an  $\fo$-formula.
-Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}} \varphi(\bar{x})$ (i.e.
-$\length(\bar{X}) = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}} [\varphi(\bar{x})\wedge\exists\bar{z}(\varphi(\bar{z})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
-Therefore, for the rest of the proof we can assume that $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.
-
- Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
-To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in into a DNF formula. 
-More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i$ is a conjunction of atomic formulas or negation of atomic formulas.
-Furthermore, we suppose that each conjunct $\varphi_i$ has the~form:
-$$
-\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
-\underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}} \wedge
-\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}}.
-$$
-Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ can appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
-
-Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ are mentioned in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
-We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
-\begin{itemize}
-	\item[(a)] In order to satisfy the first condition, consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
-We add $\length(\bar{w})$ new variables $z_1,\ldots,z_{\length(\bar{w})}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length(\bar{w})})$.
-We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$.
-The formula $\varphi_i$ is now defined as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \exists \bar{z} \; (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{-}(\bar{X},\bar{x},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y})).
-	$$
-	We repeat this process for each instance of a $X_j(\bar{w})$ in $\varphi_i$, and we obtain a formula where none of the $X_j$'s acts over any variable in $\bar{x}$.
-We add the new first-order variables to $\bar{y}$ and we redefine $\varphi_i$ as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
-	$$
-	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = \exists \bar{y} \;  X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := \exists \bar{y} \;  X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
-	\item[(b)] An ordered partition on a set $S$ is defined by an equivalence relation $\sim$ over $S$, and a linear order over $S/\!\sim$.
-For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$.
-A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
-On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
-	For a given $k$, let $\cB_k$ be the number of possible ordered partitions for a set of size $k$. For $1 \leq j \leq \cB_{\length(\bar{x},\bar{y})}$ 
-	let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
-
-	Thus, the formula $\varphi_i(\bar{X},\bar{x})$ is then redefined as:
-	$$
-	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}} \exists \bar{y} \; [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
-	$$
-	Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
-	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  to $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
-	\item[(c)] For showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
-$\varphi_i$ is inconsistent), we do the following.
-If there exists an instance of $X_j(\bar{z})$ in $\varphi^{+}_i$, an instance of $\neg X_j(\bar{w})$ in $\varphi^{-}_i$ and the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$, then the entire formula $\varphi_i$ is removed from $\varphi$.
-\end{itemize}
-It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$.
-From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c).
-Furthermore, it has the following form:
-$$
-\varphi_i(\bar{X},\bar{x}) \; = \; \exists{\bar{y}} \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
-$$
-where $\varphi^{+}$ and $\varphi^{-}_i$ do not depend on  $\bar{x}$.
-\begin{claim}\label{claim:minusone}
-For a given ordered structure $\A$, if there exists a first-order assignment $v$ for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists a second-order assignment $V$ for $\A$ such that  $(\A,V,v)\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-\end{claim}
-\begin{proof}
-Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
-Define $\bar{B} = (B_1,\ldots,B_{\length(\bar{X})})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
-Towards a contradiction, suppose that $(\A,V,v)\not\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-By the choice of $v$, and construction of $V$ it is clear that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})\wedge\varphi^{+}_i(\bar{X},\bar{y})$, so we necessarily have that $(\A,V,v)\not\models\varphi^{-}_i(\bar{X},\bar{y})$.
-Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,V, v)\not\models\neg X_t(\bar{w})$, namely, $v(\bar{w})\in B_t$. 
-However, by the construction of $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$.
-Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$.
-This violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
-\end{proof}
-
-The previous claim and proof motivates the following definitions.
-For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length(\bar{X})})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
-One can easily check that for every assignments $V, v$ such that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$, then it holds that $(\A,\bar{B}^v,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all the assignment of $\bar{X}$ when $v$ is fixed.
-Therefore, by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
-For this, the idea is to choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
-$$
-\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge\forall\bar{x}'\forall\bar{y}'(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')).
-$$ 
-If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
-Furthermore, let $V^*$ the second order assignment that satisfies $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*$ the first order assignment such that $v^*(\bar{x}) = v(\bar{x})$.
-By the previous discussion, we have that $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.
-
-Now, we have all the tools to introduce the main formula of the proof. 
-Intuitively, we want to exclude the assignment $(V^*, v^*)$ as a satisfying assignment for $\varphi_i(\bar{X},\bar{x})$.
-Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if, 
-if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, then $V \neq V^*$ or $v \neq v^*$. 
-That is, it excludes all pairs $(V, v)$ that coincide with $(V^*, v^*)$ in $(\bar{X}, \bar{x})$ whenever $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. This property can be defined as follows:
-\begin{align}
-\psi_i(\bar{X},\bar{x}) = & \;\; \big( \, \exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
-& \hspace{1cm} \Big( \; 
-\exists\bar{y} \;\; \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i^+(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\
-& \hspace{1.5cm} (\exists\bar{x}'\exists\bar{y}\, \varphi_i(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}) \;\; \Big)  \label{eq:part3}
-\end{align}
-To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
-Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-Then, the conclusion of the implication at (\ref{eq:part2}) and (\ref{eq:part3}), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
-Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by showing that if $\varphi_i(\bar{X},\bar{x})$ is satisfied then $V(\bar{X})$ must contain one element more that it is not in $V^*(\bar{X})$.
-The second disjunct (\ref{eq:part3}) is satisfied when $v(\bar{x})$ is not the lexicographically smallest tuple that satisfies $\varphi_i$ (i.e. $v(\bar{x}) \neq v^*(\bar{x})$).
-
-We are ready to define the formula  $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$
-where each modified disjunct $\varphi_i'(\bar{X},\bar{x})$ is constructed as follows. 
-For the sake of simplification, define the auxiliary formula $\chi_i = \neg\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$. 
-This formula basically checks if $\varphi_i$ is satisfiable or not.
-Then, first define the formula:
-$$
-\varphi'_1(\bar{X},\bar{x}) = \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
-$$
-This models that all the assignments that satisfy $\varphi_1$ also satisfy $\varphi'_1$, except for $(V^*,v^*)$. Then we define:
-$$
-\varphi'_2(\bar{X},\bar{x}) = \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
-$$
-This models that all the assignments that satisfy $\varphi_2$ also satisfy $\varphi'_2$, but if there was no assignment for $\varphi_1$, then we exclude $(V^*,v^*)$ for $\varphi_2$.
-Then we generalize this construction for each $\varphi_i$:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})),
-$$
-By the construction of $\kappa(\alpha)$, one can easily check that  $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.
-
-\vspace{1em}
-
-{\em Part (2).} Let $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ for some $\eqso(\logex{1})$ formula $\beta$. We define $\kappa(\alpha)$ as follows.
-First, perform the same transformations to $\varphi(\bar{X},\bar{x})$ as in the previous section.
-Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c).
-We also use the previously defined formulas $\chi_i$ and $\psi_i$, for each $i$.
-
-We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if and only if $\sem{\beta}(\A) = 0$.
-If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg\exists\bar{x}'\varphi(\bar{x}')$.
-If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$ as each $\chi_i$ was previously defined.
-If $\beta = (\beta_1 + \beta_2)$, then $\lambda(\beta) = \lambda(\beta_1) \wedge \lambda(\beta_2)$.
-Now, for each $\varphi_i$ we define:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x}))\Big)\Big).
-$$
-Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
-
-%\subsection*{Proof of Theorem \ref{sub-pnp}}
-%
-%Towards a contradiction, suppose that any of the classes $\E{1}$, $\eqso(\loge{1})$, and $\eqso(\logex{1})$ is closed under subtraction.
-%
-%Let $\R$ include the symbols $S_1, S_2, S_3, S_4$, which describe the following properties.
-If a finite $\R$-structure $\A$ defines a 3DNF formula $\Phi$, then its domain is the set of variables mentioned in $\Phi$, and for each $i = 1,2,3,4$:
-%\begin{align*}
-%	S_1^\A &= \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_2^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_3^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_4^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
-%\end{align*}
-%Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
-%\begin{multline*}
-%\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
-%\end{multline*}
-%Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
-%$$
-%\varphi(x,X) = (T(x) \vee \neg T(x)).
-%$$
-%Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Given that $f_{\#3DNF}, f_{all} \in \E{1} \subseteq \eqso(\loge{1}) \subseteq \eqso(\logex{1}) \subseteq \totp$, and at least one of the classes is closed under subtraction, the function $f_{\#3DNF} - f_{all}$ is in $\totp$.
-However, note that for each $\R$-structure $\A$ that represents a formula $\Phi$, $(f_{\#3DNF} - f_{all})(\A) = 0$ if an only if $\Phi$ is a tautology.
-The decision version of this function is the $\np$-complete problem $\textsc{Tautology}$.
-And since the function is in $\totp$, its decision version is also in $\ptime$.
-We conclude that $\np \subseteq \ptime$.
-
-\subsection*{Proof of Proposition \ref{prop:ehorn-pe}}
-Pagourtzis and Zachos mention a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula \cite{PagourtzisZ06}.
-This procedure can be easily extended to receive Horn formulas, and furthermore, a disjunction of Horn formulas.
-Hence $\shdhsat$ is in $\totp$.
-
-As we show in Proposition \ref{sigma2hard}, $\shdhsat$ is complete for $\eqso(\ehorn)$ under parsimonious reductions.
-Let $\alpha$ be a formula in $\eqso(\ehorn)$ and let $g_{\alpha}$ be the reduction to $\shdhsat$.
-The $\totp$ procedure we construct, for each input $\enc(\A)$, is simply to compute $g_{\alpha}(\enc(\A))$, and then simulate the $\totp$ procedure for $\shdhsat$ on input $g_{\alpha}(\enc(\A))$.
-We conclude that $\alpha$ is in $\totp$.
-
-\subsection*{Proof of Proposition \ref{prop:hsat-not-sigma2}} %V.12
-We use a similar proof to the one provided by the authors in \cite{SalujaST95} to separate the classes $\E{2}$ and $\U{2}$.
-Suppose that the statement is false, this is, $\chsat \in \eqso(\loge{2})$.
-We consider the signature $\R$ that we used as the encoding for a Horn formula (Example \ref{ex-hornsat-esop1}) and that the formula $\alpha \in \eqso(\loge{2})$ follows the encoding in the same way.
-From what we proved in Theorem \ref{theo-pnf-snf}, we have that every formula in $\eqso(\loge{2})$ can be rewritten in $\loge{2}$-PNF, so we assume that $\alpha$ is in this form.
-Let $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\forall\bar{z}\,\varphi(\bar{X},\bar{x},\bar{y},\bar{z})$.
-Consider the following Horn formula $\Phi$:
-$$
-\Phi = p \wedge \bigwedge_{i = 1}^n (t_i \wedge p \to q) \wedge \neg q,
-$$
-where $n = \length(\bar{x}) + \length(\bar{y}) + 1$.
-Let $\A$ be the encoding of this formula.
-In the encoding, each variable appears as an element in the domain of $\A$.
-This formula is satisfiable, so $\sem{\alpha}(\A) \geq 1$.
-Let $(\bar{B},\bar{b},\bar{a})$ be an assignment to $(\bar{X},\bar{x},\bar{y})$ such that $\A\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-Let $t_{\ell}$ be such that it does not appear in $\bar{b}$ or $\bar{a}$.
-Consider the induced substructure $\A'$ that is obtained by removing $t_{\ell}$ from $\A$ and $\bar{B}'$ as the subset of $\bar{B}$ obtained by deleting each appearance of $t_{\ell}$ in $\bar{B}$.
-We have that $\A'\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-This is because each subformula of the form $\exists y \neg B_i$ is still true, and universal formulas are monotone over induced substructures.
-It follows that $\sem{\alpha}(\A') \geq 1$ which is not possible since $\A'$ encodes the formula
-$$
-\Phi' = p \wedge \bigwedge_{i = 1}^{\ell-1} (t_i \wedge p \to q) \wedge (p\to q) \wedge \bigwedge_{i = \ell+1}^{n} (t_i \wedge p \to q) \wedge \neg q,
-$$
-which is unsatisfiable. We arrive to a contradiction and we conclude that $\chsat$ is not in $\eqso(\ehorn)$.
-
-\subsection*{Proof of Theorem \ref{sigma2hard}} %V.13
-
-First we prove that $\shdhsat$ is in $\eqso(\ehorn)$.
-Recall that each instance of $\shdhsat$ is a disjunction of Horn formulas.
-Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot), \pD(\cdot,\cdot)\}$.
-Each symbol in this vocabulary is used to indicate the same as in Example \ref{ex-hornsat-esop1}, with the addition of $\pD(d,c)$ which indicates that $c$ is a clause in the formula $d$.
-Recall that the formula
-\begin{align*}
-&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \neg\textit{T}(x)).
-\end{align*}
-defines $\chsat$. We obtain the following formula $\psi(T,A)$ in $\ehorn$:
-\begin{align*}
-\exists d[&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \pD(c,d)\vee \neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \neg\textit{T}(x))]
-\end{align*}
-effectively defines $\chsat$ as for every disjunction of Horn formulas $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
-Therefore, we conclude that $\shdhsat \in \eqso(\ehorn)$.
-
-\vspace{1em}
-We will now prove that $\shdhsat$ is hard for $\eqso$ over a signature $\R$ under parsimonious reductions.
-For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a polynomial-time procedure that computes a function $g_{\alpha}$.
-This function receives a finite $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
-We suppose that $\alpha$ is in sum normal form and:
-$$
-\alpha = \sum_{i = 1}^{\text{\#clauses}} \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigwedge_{j = 1}^{n}\forall\bar{z}\,\varphi^i_j(\bar{X},\bar{x},\bar{y},\bar{z}),
-$$
-where each $\varphi^i_j$ is a Horn clause.
-
-Consider a finite $\R$-structure $\A$ with domain $A$.
-To simplify the proof, we extend our grammar to allow first-order constants.
-Consider each tuple $\bar{a}\in A^{\length(\bar{x})}$, each $\bar{b}\in A^{\length(\bar{y})}$ and each $\bar{c}\in A^{\length(\bar{z})}$ as a tuple of first-order constants.
-The following formula defines the same function as $\alpha$:
-$$
-\sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length(\bar{x})}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length(\bar{y})}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length(\bar{z})}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
-$$
-Note that each $\fo$ formula over $(\bar{x},\bar{y},\bar{z})$ in each $\varphi^i_j$ has no free variables.
-Therefore, we can evaluate each of these in polynomial time and replace them by $\perp$ and $\top$ where it corresponds.
-Each $\varphi^i_j$ will be of the form $\perp \vee\, \chi^i_j(\bar{X})$ or $\top \vee \chi^i_j(\bar{X})$ where $\chi^i_j$ is a disjunction of $\neg X_{\ell}$'s and at most one $X_{\ell}$.
-The formulas of the form $\top \vee \chi^i_j(\bar{X})$ can be removed entirely, and the formulas of the form $\perp \vee\, \chi^i_j(\bar{X})$ can be replaced by $\chi^i_j(\bar{X})$.
-We obtain the formula
-$$
-\sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})
-$$
-where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$.
-
-Our idea for the rest of the proof is to define $g_{\alpha}$ for each $\alpha = \sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})$, for $\alpha = c$ and for $\alpha = \beta_1 + \cdots + \beta_m$ where each $\beta_i$ is in one of the two previous cases.
-
-If $\alpha$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ where $\psi_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$, then we obtain the {\bf propositional formula} $g_{\alpha}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulas.
-
-If $\alpha$ is equal to a constant $c$, then we define $g_{\alpha}(\A)$ as the following formula that has exactly $c$ satisfying assignments:
-$$
-g_{\alpha}(\A) = \bigvee_{i = 1}^{c}\neg t_1 \wedge \cdots \wedge \neg t_{i-1} \wedge t_i \wedge \neg t_{i+1} \wedge \cdots \wedge \neg p_c.
-$$ 
-If $\alpha = \beta_1 + \cdots + \beta_m$, let $g_{\beta_i}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\theta^i_{j,k}$ for each $\beta_i$ where each $\theta^i_{j,k}$ is a Horn clause.
-Let $\Theta_i = g_{\beta_i}(\A)$.
-We rename the variables in each $\Theta_i$ so none of them are mentioned in any other $\Theta_j$.
-We add $m$ new variables $t_1,\ldots,t_m$ and we define:
-\begin{align*}
-g_{\alpha}(\A) = &\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^1_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m}}}t) \wedge (t_1 \wedge \bigwedge_{\ell = 2}^{m} \neg t_{\ell})) \vee \\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^2_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each $t$}\\ \text{ mentioned in}\\ \Theta_1,\Theta_3,\ldots,\Theta_{m}}}t) \wedge (t_2 \wedge \bigwedge\limits_{\substack{\ell = 1 \\ \ell \neq 2}}^{m} \neg t_{\ell})) \vee \cdots \vee\\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^m_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m-1}}}t) \wedge (t_m \wedge \bigwedge_{\ell = 1}^{m-1} \neg t_{\ell})).
-\end{align*}
-The formula is a disjunction of Horn formulas, and the number of satisfying assignments for this formula is exactly the sum of satisfying assignments for each $g_{\beta_i}(\A)$.
-This, at the same time, is equal to $\sem{\alpha}(\A)$.
-This covers all possible cases for $\alpha$, and the entire procedure takes polynomial time.

### 2017-03-03 LMCS/beyond.tex
journal primera version

--- 2017-01-10 LICS/beyond.tex
@@
 
-%In the previous sections, 
-We have used weighted logics to give a framework for descriptive complexity of counting complexity classes. Here, we go beyond weighted logics 
-%to add unexplored quantifiers at the quantitative level. Specifically, we 
-and give the first steps on defining recursion at the quantitative level.
-%quantitative recursion. 
+We have used weighted logics to give a framework for descriptive complexity of counting complexity classes.
+Here, we go beyond weighted logics and give the first steps on defining recursion at the quantitative level.
 This goal is not trivial not only because we want to add recursion over functions, but also because 
 it is not clear what could be the right notion of ``fixed point''. 
-%for two reasons: (1) we want to add recursion over functions and (2) it is not clear what could be the right notion of ``fixed point'' . 
 To this end, we show first how to extend $\qso$ with function symbols that are later
 used 
-%to later use them to define recursion at a quantitative level. 
-%Furthermore, we 
 to define a natural generalization for functions of 
-%(boolean) 
 the notion of least fixed point of LFP.
-%, which is denoted as \emph{least s-fixed point}.
-% and, 
  As a proof of concept, we show that how this notion can be used to captures $\fp$.
 Moreover, we use this concept
-%our new tool 
 to define an operator for counting paths in a graph, a natural generalization of the transitive closure operator~\cite{immerman1999descriptive}, and show that this gives rise to a logic that captures~$\shl$.
-%the expressiveness of
 
-%For defining recursion, 
-We start
-%first need to define 
-by defining an extension of $\qso$ with function symbols. Assume that $\fs$ is an infinite set of function symbols, where each $h \in \fs$ has an associated arity
-%, which is 
-denoted by $\arity(h)$. Then the set of $\fqso$ formulae over a signature $\R$ is defined by the following grammar:
-\begin{multline}
+We start by defining an extension of $\qso$ with function symbols.
+Assume that $\fs$ is an infinite set of function symbols, where each $h \in \fs$ has an associated arity denoted by $\arity(h)$.
+Then the set of $\fqso$ formulae over a signature $\R$ is defined by the following grammar:
+\begin{align}
 \label{eq-fqso}
-	\alpha := \varphi \ \mid \  s \  \mid \  h(x_1, \ldots, x_\ell) \  \mid \
-	(\alpha \add \alpha) \  \mid\  (\alpha \mult \alpha) \  \mid \\  
+	\alpha &:= \varphi \ \mid \  s \  \mid \  h(x_1, \ldots, x_\ell) \  \mid \
+	(\alpha \add \alpha) \  \mid\  (\alpha \mult \alpha) \  \mid 
 	\sa{x} \alpha \  \mid \
@@
 	\pa{X} \alpha,
-\end{multline}
+\end{align}
 where $h \in \fs$, $\arity(h) = \ell$ and $x_1, \ldots, x_\ell$ is a sequence of (non-necessarily distinct) first-order variables.
@@
 we have that:
-\begin{eqnarray*}
-\sem{h(x_1, \ldots, x_\ell)}(\A,v,V,F) & = & F(h)(v(x_1),\ldots, v(x_\ell)).
-\end{eqnarray*}
+\begin{align*}
+\sem{h(x_1, \ldots, x_\ell)}(\A,v,V,F) &= F(h)(v(x_1),\ldots, v(x_\ell)).
+\end{align*}
 As for the case of $\qfo$, we define $\fqfo$ disallowing quantifiers $\Sigma X$ and $\Pi X$ in \eqref{eq-fqso}.
 
-It is worth noting that function symbols in $\fqso$ represent functions from tuples to natural numbers, so they are different from 
-%cannot be compared to 
-the classical notion of function symbol in $\fo$~\cite{L04}. 
-%~$\shl$.first-order logics
-Furthermore, %one could see 
-a function symbol can be seen as an ``oracle'' that is instantiated by the function assignment. 
-%As far as we know, 
-To the best of our knowledge, this is the first paper to propose 
-%proposing 
-this extension on weighted logics, which should be further investigated. 
-% and it would be interesting to study this concept further (i.e. outside recursion).
+It is worth noting that function symbols in $\fqso$ represent functions from tuples to natural numbers, so they are different from the classical notion of function symbol in $\fo$~\cite{L04}. 
+Furthermore, a function symbol can be seen as an ``oracle'' that is instantiated by the function assignment. 
+To the best of our knowledge, this is the first paper to propose this extension on weighted logics, which should be further investigated.
 
-We define an 
-extension of 
-%operator which extends 
-LFP \cite{I86,vardi1982complexity} 
-%over $\fo$ to 
-to allow counting. 
-%We show how to define this operator in $\qfo(\fo)$ and we discuss how to extend these ideas to $\qso$ towards the end of the section. 
-%Fix a relational signature $\R$. 
-More precisely, the set of $\rqfo(\fo)$ formulae over a signature $\R$, where $\rqfo$ stands for recursive $\qfo$, is defined as an extension of $\qfo(\fo)$ that includes the formula 
-%by adding the operator 
-$\clfp{\beta(\x, h)}$, where (1) $\x = (x_1, \ldots, x_\ell)$ is a sequence of $\ell$ distinct first-order variables, (2) $\beta(\x, h)$ is an $\fqfo(\fo)$-formula over $\R$ whose only function symbol is $h$, and (3) $\arity(h) = \ell$.
+We define an extension of LFP \cite{I86,vardi1982complexity} to allow counting. 
+More precisely, the set of $\rqfo(\fo)$ formulae over a signature $\R$, where $\rqfo$ stands for recursive $\qfo$, is defined as an extension of $\qfo(\fo)$ that includes the formula $\clfp{\beta(\x, h)}$, where (1) $\x = (x_1, \ldots, x_\ell)$ is a sequence of $\ell$ distinct first-order variables, (2) $\beta(\x, h)$ is an $\fqfo(\fo)$-formula over $\R$ whose only function symbol is $h$, and (3) $\arity(h) = \ell$.
 The free variables of the formula $\clfp{\beta(\x,h)}$ are $x_1, \ldots, x_\ell$; in particular, $h$ is not considered to be free.
@@
 More precisely, for every $f \in \F$ and tuple $\a = (a_1, \ldots, a_{\ell}) \in A^\ell$, the function $T_{\beta}(f)$ satisfies that:
-\begin{eqnarray*}
-T_{\beta}(f)(\a) & = & \sem{\beta(\x, h)}(\A,v,F),
-\end{eqnarray*}
+\begin{align*}
+T_{\beta}(f)(\a) &= \sem{\beta(\x, h)}(\A,v,F),
+\end{align*}
 where $v$ is a first-order assignment  for $\A$ such that $v(x_i) = a_i$ for every $i \in \{1, \ldots, \ell\}$, and $F$ is a function assignment for $\A$ such that $F(h) = f$.
@@
 (a) $T(R) = R$, and (b) $R \subseteq S$ for every $S$ such that $T(S) = S$. 
-%\begin{enumerate}
-% \item[(a)] $T(R) = R$, and 
-% \item[(b)]  $R \subseteq S$ for every $S$ such that $T(S) = S$.  
-%\end{enumerate}
 We can view $T$ as an operator on functions if we consider the characteristic function of a relation.
@@
 Then given that $\support(\chi_R) = R$, we have that the conditions (a) and (b) are equivalent to the following conditions on $T^\star$:
-%Given a relation $R \subseteq A^\ell$, let $\chi_R$ be its characteristic function, that is $\chi_R(\bar a) = 1$ if $\bar a \in R$, and $\chi_R(\bar a) = 0$ otherwise.
-Moreover, given a function $f \in \F$, define the support of $f$, denoted by $\support(f)$, as $\{ \bar a \in A^\ell \mid f(\bar a) > 0 \}$.
-Notice that for a relation $R$ we have that $\support(\chi_R) = R$.
-In particular, the conditions defining the least fixed point of $T$ can be rewritten as follows in terms of $T^\star$:
 (a) $\support(T^\star(\chi_R)) = \support(\chi_R)$, and  (b) $\support(\chi_R) \subseteq \support(\chi_S)$ for every $S$ such that  $\support(T^\star(\chi_{S})) = \support(\chi_S)$.
-%\begin{enumerate}
-%	\item[(a)] $\support(T^\star(\chi_R)) = \support(\chi_R)$, and  
-%	\item[(b)] $\support(\chi_R) \subseteq \support(\chi_S)$ for every $S$ such that  $\support(T^\star(\chi_{S})) = \support(\chi_S)$.  
-%\end{enumerate}
-To define a notion of fixed point for $T_\beta$ we simply generalized these conditions. More precisely, %These conditions can be generalized in the obvious way to the case of the operator $T_\beta$: 
-a function $f \in \F$ is a {\em s-fixed point} of $T_{\beta}$ if $\support(T_\beta(f)) = \support(f)$, and $f$ is a {\em least s-fixed point} of $T_{\beta}$ if $f$ is a s-fixed point of $T_\beta$ and for every s-fixed point $g$ of $T_\beta$ it holds that $\support(f) \subseteq \support(g)$.
+To define a notion of fixed point for $T_\beta$ we simply generalized these conditions.
+More precisely, a function $f \in \F$ is a {\em s-fixed point} of $T_{\beta}$ if $\support(T_\beta(f)) = \support(f)$, and $f$ is a {\em least s-fixed point} of $T_{\beta}$ if $f$ is a s-fixed point of $T_\beta$ and for every s-fixed point $g$ of $T_\beta$ it holds that $\support(f) \subseteq \support(g)$.
 The existence of such fixed point is ensured by the following lemma:
-\begin{lemma}\label{lem-support}
+\begin{lem}\label{lem-support}
 If $f,g \in \F$ and $\support(f) \subseteq \support(g)$, then $\support(T_\beta(f)) \subseteq \support(T_\beta(g))$.
-\end{lemma}
+\end{lem}
+\proof
+\input{proof-lem-support}
+\qed
 In fact, as for the case of LFP, this lemma gives us a simple way to compute a least s-fixed point of $T_\beta$.
@@
 More specifically, for an arbitrary first-order assignment $v$ for $\A$:
-\begin{eqnarray*}
-\sem{\clfp{\beta(\x, h)}}(\A,v) & = & f_{k}(v(\x))
-\end{eqnarray*}
+\begin{align*}
+\sem{\clfp{\beta(\x, h)}}(\A,v) &= f_{k}(v(\x))
+\end{align*}
 
-\begin{example} \label{ex:count-path}
-%As an example, 
+\begin{exa} \label{ex:count-path}
 We would like to define an $\rqfo(\fo)$-formula that, given a directed acyclic graph $G$ with $n$ nodes and a pair of nodes $b$, $c$ in $G$, counts the number of paths of length at most $n$ from $b$ to $c$ in $G$.
@@
 $$ 
-%\alpha(x,y,R,\pi) = 
-%(\neg \exists zR(z))\cdot(x = y) 
 E(x,y) + \sa{z} f(x,z)\cdot E(z,y).
@@
 Moreover, define formula $\beta(x, y, t, g)$ as follows:
-\begin{multline*}
-(E(x,y) + \sa{z} g(x,z,t)\cdot E(z,y)) \cdot \varphi_{\text{\rm first}}(t) \ +\\
+\begin{align*}
+(E(x,y) + \sa{z} g(x,z,t)\cdot E(z,y)) \cdot \varphi_{\text{\rm first}}(t) \ +
 \sa{t'} \varphi_{\text{succ}}(t',t) \cdot \left(\sa{x'} \sa{y'} g(x',y',t') \right)
-\end{multline*}
+\end{align*}
 Then our extended counting function is defined by:
@@
 In fact, the number of paths of length at most $n$ from a node $x$ to a node $y$ is recursively computed by using the formula $(E(x,y) + \sa{z} g(x,z,t)\cdot E(z,y)) \cdot \varphi_{\text{\rm first}}(t)$, which stores this value in $g(x,y,t)$ with $t$ the first element in the domain.
-The other formula $\sa{t'} \varphi_{\text{succ}}(t',t) \cdot \left(\sa{x'} \sa{y'} g(x',y',t') \right)$ is just an auxiliary artifact that is used
-%The other formula, the one with the filter $\varphi_{\text{succ}}(t',t)$, is used 
-as a counter to allow reaching a fixed point in the support of $g$ in $n$ steps.
+The other formula $\sa{t'} \varphi_{\text{succ}}(t',t) \cdot \left(\sa{x'} \sa{y'} g(x',y',t') \right)$ is just an auxiliary artifact that is used as a counter to allow reaching a fixed point in the support of $g$ in $n$ steps.
 Notice that the use of the filter $\varphi_{\text{succ}}(t',t)$ prevents this formula for incrementing the value of $g(x,y,t)$ when $t$ is the first element in the domain.
-\end{example}
-%
-%
-%It is well known that least fixed point logic is contained in second-order logic \cite{L04}. In the following theorem we show that the same holds in our case.
-%\begin{theorem} \label{so-rec}
-%$\rqfo \subseteq \qso$
-%%	Given a positive $\fo$ formula $\varphi(\bar{x},R)$ and a $\qfo$ formula $\alpha(\bar{x})$, there exists a $\qso$ formula $\beta(\bar{x})$ such that $\sem{[\alfp\varphi(\bar{x},R)\mid \alpha(\bar{x},R)](\bar{x})} = \sem{\beta(\bar{x})}$.
-%\end{theorem} 
-%Note that, 
-In contrast with $\lfp$, to reach a fixed point 
-%here 
-we do not need to impose any positive restriction on the formula $\beta(\x,h)$.
+\end{exa}
+In contrast with $\lfp$, to reach a fixed point we do not need to impose any positive restriction on the formula $\beta(\x,h)$.
 Indeed, since $\beta$ is constructed from monotones operations (i.e. sum and product) over the natural numbers, the resulting operator $T_{\beta}$ is monotone as well.
@@
 In the following theorem, we show that this operator can be used to capture $\fp$.
-%result we show how lsfp-operator can be used to capture all functions that can be computed in polynomial time.
-\begin{theorem} \label{rqfo-fo-cap}
-	$\rqfo(\fo)$ captures $\fp$ over 
-	%the class of 
-	ordered structures.
-\end{theorem}
-
-%Given a relation signature $\R$, the set of recursive $\qfo$ formulae ($\rqfo$-formulae) is defined by the following grammar:
-%%This operator lets us define the set of recursive $\qfo$ formulae over $\R$ ($\rqfo$-formulae) using the following grammar:
-%\begin{multline*}
-%%	\label{eq-def-rqfo}
-%	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid \\ (\alpha \mult \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ [\alfp \varphi \mid \alpha]
-%\end{multline*}
-%where $\varphi$ is an $\fo$-formula over $\R$, $s \in \bbN$ and $x \in \fv$.
-%
-%\marcelo{Vamos a permitir anidacion del operador $\alfp$? Esta gramatica lo permite.}
-%
-%\begin{theorem} \label{rqfo-fo-cap}
-%	$\rqfo(\fo)$ captures $\fp$ over the class of ordered structures.
-%\end{theorem}
-%
-
+\begin{thm} \label{rqfo-fo-cap}
+	$\rqfo(\fo)$ captures $\fp$ over ordered structures.
+\end{thm}
+\proof
+\input{proof-rqfo-fo-cap}
+\qed
 Our last goal in this section is to use the new characterization of $\fp$ to explore classes below it.
 It was shown in \cite{I86,I88} that $\fo$ extended with a transitive closure operator captures $\nlog$. 
-%Moreover, it was shown in \cite{I83} that $\so$ extended with a transitive closure operator captures $\pspace$. 
 Inspired by this work, we show that a restricted version of $\rqfo$ can be used to capture $\shl$, the counting version of $\nlog$. 
 Specifically, we use $\rqfo$ to define an operator for counting the number of paths in a directed graph, which is what is needed to capture $\shl$.
-%Besides, we show that the same idea can be used to extend $\qso$ allowing to capture harder complexity classes.
 
-Given a relation signature $\R$, the set of transitive $\qfo$ formulae ($\tqfo$-formulae) is defined as an extension of $\qfo$ with the formula
-%\begin{multline}
-%	\label{eq-def-tqso}
-%	\alpha := \varphi \, \mid \, s \, \mid \, (\alpha \add \alpha) \, \mid\, (\alpha \mult \alpha) \, \mid \, 
-%	\sa{x} \alpha \, \mid\, \
-%	\pa{x} \alpha \, \mid \\ 
-%	\sa{X} \alpha \, \mid \, \pa{X} \alpha \, \mid \, [\pth \psi(\bar{x}, \bar{X},\bar{y}, \bar{Y})],
-%\end{multline}
-$
-[\pth \psi(\bar{x},\bar{y})]
-$, where $\psi(\x, \y)$ is an $\fo$-formula over $\R$, and $\bar{x} = (x_1, \ldots, x_k)$, $\bar{y} = (y_1, \ldots, y_k)$ are tuples of pairwise distinct first-order variables.
+Given a relation signature $\R$, the set of transitive $\qfo$ formulae ($\tqfo$-formulae) is defined as an extension of $\qfo$ with the formula $[\pth \psi(\bar{x},\bar{y})]$, where $\psi(\x, \y)$ is an $\fo$-formula over $\R$, and $\bar{x} = (x_1, \ldots, x_k)$, $\bar{y} = (y_1, \ldots, y_k)$ are tuples of pairwise distinct first-order variables.
 The semantics of $[\pth \psi(\bar{x},\bar{y})]$ can easily be defined in terms of $\rqfo(\fo)$ as follows.
@@
 Similar than for Example~\ref{ex:count-path}, we can count the paths of length at most $|A^k|$ in $ \cG_{\psi}(\A)$ with the formula $\beta_{\psi(\bar{x},\bar{y})}(\x, \y, \t, g)$:
-\begin{multline*}
-(\psi(\bar{x},\bar{y}) + \sa{\z} g(\x,\z,\t)\cdot \psi(\z,\y)) \cdot \varphi_{\text{\rm first-lex}}(\t) \ +\\
+\begin{align*}
+(\psi(\bar{x},\bar{y}) + \sa{\z} g(\x,\z,\t)\cdot \psi(\z,\y)) \cdot \varphi_{\text{\rm first-lex}}(\t) \ +
 \sa{\t'} \varphi_{\text{succ-lex}}(\t',\t) \cdot \left(\sa{\x'} \sa{\y'} g(\x',\y',\t') \right),
-\end{multline*}
+\end{align*}
 where $\varphi_{\text{\rm first-lex}}$ and $\varphi_{\text{succ-lex}}$ are $\fo$-formulae defining the first and successor predicates over tuples in $A^k$, following the lexicographic order induced by~$<$.
@@
 In other words, $\sem{[\pth \psi(\bar{x}, \bar{y})]}(\A,v)$ counts the number of paths from $v(\bar x)$ to $v(\bar y)$ in the graph $\cG_{\psi}(\A)$ whose length is at most~$|A^k|$.
-%As for the case of $\qso$, the logic $\tqso(\LL)$ is obtained by restricting $\varphi$ in \eqref{eq-def-tqso} to be a formula in $\LL$.
-Moreover, the logic $\tqfo$ is obtained by disallowing in \eqref{eq-def-tqso} formulae $\sa{X} \alpha$ and $\pa{X} \alpha$, and by only allowing  first-order free-variables in the formula $\psi$ used in $[\pth \psi]$ in \eqref{eq-def-tqso}.
 As it was previously said, the operator for counting paths is exactly what we need to capture $\shl$.
-\begin{theorem} \label{tqfo-shl}
+\begin{thm} \label{tqfo-shl}
 	$\tqfo(\fo)$ captures $\shl$ over ordered structures.
-\end{theorem}
-This last result 
-perfectly 
-illustrates the benefits of our logical framework for the development of descriptive complexity for counting %functions.
-complexity classes.  
+\end{thm}
+\proof
+\input{proof-tqfo-shl}
+\qed
+This last result perfectly illustrates the benefits of our logical framework for the development of descriptive complexity for counting complexity classes.  
 The distinction in the language between the Boolean and the quantitative level allows us to define operators at the later level that cannot be defined at the former. 
@@
 and thus contradicts the widely believed assumption that $\ulog \subsetneq \nlog$.
-%$\qfo(\tc)$ can def
-%In fact, it is not clear how $\shl$ could be capture by 
-%to capture $\shl$ with $\qfo(\tc)$\footnote{$\tc$ corresponds to $\fo$-logic extended with transitive closure.}, or any other extension of $\qfo(\fo)$ at the boolean level.
-%In contrast, with $\rqfo$ one can naturally define the path operator at the quantitative level and easily capture $\shl$.
-%Hence, Theorem~\ref{tqfo-shl} is good example of how to go beyond classical logic to have a better understanding of the expressivity of counting.
-
-\begin{comment}
-
-\begin{theorem} \label{tqso-fo-fpsace}
-	$\tqso$ and $\tqso(\fo)$ captures $\fpspace$ over the class of ordered structures.
-\end{theorem}
-
-\begin{theorem} \label{tqfo-subseteq}
-	$\tqfo(\fo) \subseteq \rqfo(\fo)$.
-\end{theorem}
-Given tuples $\bar x = (x_1, \ldots, x_\ell)$, $\bar y = (y_1, \ldots, y_\ell)$ of pairwise distinct first-order variables, define formula $\varphi_{\text{\rm lex}}(\bar x, \bar y)$ as follows:
-$$
-\bigvee_{i = 1}^\ell \bigg(\bigwedge_{j = 1}^{i -1} x_i = y_i\bigg) \wedge x_i < y_i.
-$$
-That is, $\varphi_{\text{\rm lex}}(\bar x, \bar y)$ holds if $\bar x$ is smaller than $\bar y$ in the lexicographic order on tuples with $\ell$ elements induced by the built-in order of each structure.
-Moreover, define formula $\varphi_{\text{\rm succ}}(\bar x, \bar y)$ as follows:
-$$
-\varphi_{\text{\rm lex}}(\bar x, \bar y) \wedge \neg \exists \bar z \, (\varphi_{\text{\rm lex}}(\bar x, \bar z) \wedge \varphi_{\text{\rm lex}}(\bar z, \bar y)).
-$$
-That is, $\varphi_{\text{\rm succ}}(\bar x, \bar y)$ holds if $\bar y$ is the successor of $\bar x$ in the lexicographic order on tuples with $\ell$ elements.
-With this notation, define $\tqsos$ to be the restriction of $\tqso$ where each occurrence of the operator ${\bf path}$ is of the form:
-$$
-[\pth (\psi(\bar{x}, \bar{X},\bar{y}, \bar{Y}) \wedge \varphi_{\text{\rm succ}}(\bar x, \bar y))],
-$$
-where $\psi(\bar{x}, \bar{X},\bar{y}, \bar{Y})$ satisfies the same conditions as in grammar \eqref{eq-def-tqso}. Then we have that:
-\begin{theorem} \label{tqsos-shp}
-	$\tqsos(\fo)$ captures $\shp$ over the class of ordered structures.
-\end{theorem}
-
-\end{comment}
-
-%
-%\marcelo{Puede que este equivocado, pero me parece que teniamos una forma de capturar $\shp$ usando el operator ${\bf path}$.
-Pero no logro recordar como se hacia esto, y me parece que lo que habiamos escrito antes en esta seccion estaba equivocado: ``$\tqso(\fo)$ captures $\shp$ over the class of ordered structures".
-Claro que yo puedo estar usando una definicion distinta de $\tqso(\fo)$.}
-
-%We also define the set of transitive $\qso$ formulae ($\tqso$-formulae) using the following grammar:
-%\begin{multline*}
-%%	\label{eq-def-tqso}
-%	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \\ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha \ \mid \ [\pth \varphi]
-%\end{multline*}
-%
-%
-% 
-%We define the operator {\bf path} as follows.
-Let $\A$ be an ordered structure.
-Given a formula $\psi(\bar{x},\bar{y})$, where $\vert \bar{x} \vert = \vert \bar{y} \vert = k$ let ${\cal G} = ({\cal V},\cal{E})$ be induced graph over the set of vertices ${\cal V} = A^k$, and for every $\bar{a},\bar{b}\in A^k$ it holds that ${\cal E}(\bar{a},\bar{b})$ if and only if $\A \models \psi(\bar{a},\bar{b})$.
-To formalize the semantics for this operator, let $n = \vert A^k \vert$.
-%For a given first order assignment $v$ and a second order asssignment $V$, let $\bar{a} = v(\bar{x})$ and $\bar{b} = v(\bar{y})$, and $\sem{[\pth\, \psi(\bar{x},\bar{y})]}(\A,v,V)$ will take the value of the number of paths of size less or equal to $n$ from $\bar{a}$ to $\bar{b}$ in the graph ${\cal G}$.
-This operator lets us define the set of transitive $\qfo$ formulae over $\R$ ($\tqfo$-formulae) using the following grammar:
-%\begin{multline*} 
-%%	\label{eq-def-tqfo}
-%	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \\ \mid \ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ [\pth \varphi]
-%\end{multline*}
-%where $\varphi$ is an $\fo$-formula over $\R$, $s \in \bbN$ and $x \in \fv$.
-%
-%We also define the set of transitive $\qso$ formulae ($\tqso$-formulae) using the following grammar:
-%\begin{multline*}
-%%	\label{eq-def-tqso}
-%	\alpha := \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \\ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha \ \mid \ [\pth \varphi]
-%\end{multline*}
-%where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
-%\begin{theorem} \label{so-rec}
-%	Given a positive $\fo$ formula $\varphi(\bar{x},R)$ and a $\qfo$ formula $\alpha(\bar{x})$, there exists a $\qso$ formula $\beta(\bar{x})$ such that $\sem{[\alfp\varphi(\bar{x},R)\mid \alpha(\bar{x},R)](\bar{x})} = \sem{\beta(\bar{x})}$.
-%\end{theorem}
-%
-%\begin{theorem} \label{tqfo-fo-cap}
-%	$\tqfo(\fo)$ captures $\shl$ over the class of ordered structures.
-%\end{theorem}
-%
-%\begin{theorem} \label{tqso-fo-cap}
-%	$\tqso(\fo)$ captures $\shp$ over the class of ordered structures.
-%\end{theorem}

### 2017-03-03 LMCS/conclusions.tex
journal primera version

--- 2017-01-09 LICS/conclusions.tex
@@
 In this sense, there are several directions for future research, some of which are mentioned here. 
-$\totp$ is an important counting complexity class for which we are missing a logical characterization.
-In the same way, we are missing logical characterization of harder fundamental complexity class such as $\spanl$.
+$\totp$ is an interesting counting complexity class as it naturally defines a class of functions in $\shp$ with easy decision counterparts.
+However, we do not have a logical characterization of this class.
+%for which we are missing a logical characterization. 
+In the same direction, we are missing logical characterizations of other fundamental complexity classes such as $\spanl$.
 We would also like to define a larger syntactic subclass of $\shp$ where each function admits an FPRAS; notice that $\cpm$ is an important problem admitting an FPRAS\cite{JSV04} that is not included in the classes defined in Section \ref{sec-clo}.

### 2017-03-03 LMCS/counting.tex
journal primera version

--- 2017-01-10 LICS/counting.tex
@@
 %over $\R$-structures.
-\begin{definition} \label{def:cap}
+\begin{defi} \label{def:cap}
 	Let $\FF$ be a fragment of $\qso$ and $\CC$ a counting complexity class. Then {\em  $\FF$ captures $\CC$ over ordered $\R$-structures} if the  following conditions hold:
@@
 	Moreover, {\em $\FF$ captures $\CC$ over ordered structures} if $\FF$ captures~$\CC$ over ordered $\R$-structures for every signature~$\R$.
-\end{definition}
+\end{defi}
 %For the sake of simplification, we denote the first condition by $\FF \subseteq \CC$ and the second condition by $\CC \subseteq \FF$.
@@
 In our setting, the class $\sfo$ is contained in $\eqso(\fo)$, which also captures $\shp$ as expected.
-\begin{proposition} \label{prop:capture-shP}
+
+For a given signature $\R$, we define $\ostr[\R]^*$ as
+\begin{multline*}
+\ostr[\R]^* = \{(\A,v,V) \mid \A\in\ostr[\R]\text{,} \\ \text{$v$ ($V$) is a first-order (second-order) assignment for $\A$}  \}.
+\end{multline*}
+
+\begin{prop} \label{prop:capture-shP}
 	$\eqso(\fo)$ captures $\shp$ over ordered structures.
-\end{proposition}
+\end{prop}
+\proof
+\input{proof-capture-shP} 
+\qed
 Recall  that every function class $\# \LL$ is contained in $\eqso(\LL)$ (see Section~\ref{sec:previous}).
@@
 The other direction of Proposition \ref{prop:capture-shP} follows by the fact that $\shp$ is closed under first- and second-order sum.
-%Moreover, these closure properties also hold if we change the boolean core $\fo$ to any polynomial-time logic like $\lfp$. This gives the following corollary.
-%\begin{corollary} \label{prop:capture-shP-LFP}
-%	$\eqso(\lfp)$ captures $\shp$ over ordered structures.
-%\end{corollary}
 
@@
 As one could expect, if we parametrize $\eqso$ with $\eso$, we can also capture~$\spp$.
-\begin{proposition} \label{prop:capture-spanP}
+\begin{prop} \label{prop:capture-spanP}
 	$\eqso(\eso)$ captures $\spp$ over ordered structures.
-\end{proposition}
+\end{prop}
+\proof
+\input{proof-capture-spanP}
+\qed
 Can we capture $\fp$ by using $\# \LL$ for some fragment $\LL$ of $\so$?
@@
 But in this case one
-%The previous result connecting fragments from $\# \LL$ with $\shp$ and $\spp$ is the further that we can go by using SO-formulas with free variables. 
-%So, what happen with other counting complexity classes like $\fp$ and $\fpspace$?
-%By trying to search for an $\# \LL$ class for capturing $\fp$ one can easily see that this is pointless.
-%Indeed, 
-%By using free second-order variables one can easily capture $\shp$-hard functions (see Section~\ref{sec:syntactic}) and 
-%By only using free first-order variables one 
 cannot capture exponential functions definable in $\fp$ such as~$2^n$.
-%(e.g. $2^{n}$), so
-Thus, it is not  clear how to capture
-%impossible 
-%to capture 
-$\fp$ 
-%or $\fpspace$ 
+Thus, it is not  clear how to capture $\fp$ 
 by following the approach proposed in~\cite{SalujaST95}. 
-%if we follow previous approaches like
 On the other hand, if we consider our framework and move out from $\eqso$, we have other options for counting like first- and second-order products.
 In fact, the combination of $\qfo$ with $\lfp$ is exactly what we need to capture $\fp$.
-% the class of function that can be computed in~$\ptime$.
-\begin{theorem} \label{theo:capture-fp}
+
+From now on, we will use the symbol $<$ also to denote the lexicographic order over same-sized tuples.
+If $\bar{x} = (x_1,\ldots,x_m)$ and $\bar{y} = (y_1,\ldots,y_m)$ are tuples of first-order variables, we denote $\bar{x} < \bar{y}$ for the formula $\bigvee_{i = 1}^m[\bigwedge_{j = 1}^{i-1}x_j = y_j \wedge x_i < y_i]$.
+Similarly, we use $=$ to denote equality between tuples, as $\bar{x} = \bar{y}$ denotes $\bigwedge_{i = 1}^m(x_i = y_i)$, and also $\bar{x}\leq\bar{y}$ denotes $\bar{x} < \bar{y} \vee \bar{x} = \bar{y}$.
+Furthermore, If $\bar{x} = (x_1,\ldots,x_m)$ ($\bar{X} = (X_1,\ldots,X_m)$) is a tuple of first-order (second-order) variables, we denote $\sa{\bar{x}}\alpha$ for $\sa{x_1}\cdots\sa{x_m}\alpha$ and $\sa{\bar{X}}\alpha$ for $\sa{X_1}\cdots\sa{X_m}\alpha$ for each $\qso$ formula $\alpha$.
+
+The {\em conditional count} symbol $(\varphi \mapsto \alpha)$ is defined as $(\neg\varphi + (\varphi\cdot\alpha))$ for given $\so$ formula $\varphi$ and $\qso$ formula $\alpha$.
+Note that for each $(\A,v,V) \in \ostr[\R]^*$,
+\[
+\sem{(\varphi \mapsto \alpha)}(\A,v,V) = 
+\begin{cases}
+\sem{\alpha}(\A,v,V) &\text{if } (\A,v,V)\models\varphi,\\
+0 &\text{otherwise}.
+\end{cases}
+\]
+\begin{thm} \label{theo:capture-fp}
 	$\qfo(\lfp)$ captures $\fp$ over ordered structures.
-\end{theorem}
+\end{thm}
+\proof
+\input{proof-capture-fp}
+\qed
 To prove this theorem, 
@@
 For the other direction, one has to use $\qfo(\lfp)$ to simulate the run of a polynomial time TM $M$ computing a function, in particular using the quantitative quantifiers to reconstruct the natural number returned by $M$ in the output tape.
-%built in the ouptape. 
 It is important to notice that the alternation between sum and product quantifiers is crucial for this reconstructions and, thus, crucial for capturing $\fp$.
-%That is, without these operators is difficult to capture $\fp$.
 
 At this point it is natural to ask whether one can extend the previous idea to capture $\fpspace$~\cite{Ladner89}, the class of functions computable in polynomial space. 
-%Another natural function complexity class is $\fpspace$~\cite{Ladner89}, the functional version of $\pspace$. 
 Of course, for capturing this class one needs a logical core powerful enough, like $\pfp$, for simulating the run of a polynomial-space TM.
-% $\pspace$-machine. 
 Moreover, 
-%On the other hand, 
 one also needs more powerful quantitative quantifiers as functions like $2^{2^n}$ can be computed in polynomial space,
-%such a TM can generatean $\fpspace$ machine can generate outputs of exponentially many bits. 
 so second-order sum is not enough for the quantitative layer of a logic for $\fpspace$.
-%For this reason, the second-order sum is not enough for the quantitative layer of a logic for $\fpspace$, that is, one needs to consider second-order products to produce big numbers. 
-In fact, by considering second-order product we obtain the fragment $\qso(\pfp)$ that 
-%naturally 
-captures $\fpspace$. 
-\begin{theorem} \label{theo:capture-fpspace}
+In fact, by considering second-order product we obtain the fragment $\qso(\pfp)$ that captures $\fpspace$. 
+\begin{thm} \label{theo:capture-fpspace}
 	$\qso(\pfp)$ captures $\fpspace$ over ordered structures.
-\end{theorem}
+\end{thm}
+\proof
+\input{proof-capture-fpspace}
+\qed
 The proof of the previous theorem follows the same line as for the logical characterization of $\fp$: one shows that each function in $\qso(\pfp)$ can be computed in $\fpspace$ and, conversely, the output of a polynomial-space TM can be reconstructed by using $\pfp$ and quantitative quantifiers.
@@
 In~\cite{Ladner89}, Ladner also introduced the class $\nfpspace$ of all functions computed by polynomial-space TMs 
-%$\pspace$-machines 
 with output length bounded by a polynomial.
-% output is polynomially bounded. 
 Interestingly, if we restrict to FO-quantitative quantifiers we can also capture this class.
-\begin{corollary} \label{cor:capture-fpspace-poly}
+\begin{cor} \label{cor:capture-fpspace-poly}
 	$\qfo(\pfp)$ captures $\nfpspace$ over ordered structures.
-\end{corollary}
+\end{cor}
+\proof
+\input{proof-capture-fpspace-poly}
+\qed
 
 The results of this section validate $\qso$ as an appropriate logical framework for extending the theory of descriptive complexity to counting complexity classes.
-In the following sections, we provide more arguments for this claim, by considering some fragments of $\eqso$ and, moreover, by showing how to go beyond $\eqso$ to capture other
-%counting complexity 
-classes.
+In the following sections, we provide more arguments for this claim, by considering some fragments of $\eqso$ and, moreover, by showing how to go beyond $\eqso$ to capture other classes.

### 2017-03-03 LMCS/extentions.tex
journal primera version

--- 2017-01-10 LICS/extentions.tex
@@
 To overcome this limitation, $\gp$ was introduced in~\cite{FFK94} as the class of functions $f$ for which there exists a polynomial-time NTM $M$ such that $f(x) = \tma_M(x) - \tmr_M(x)$, where  $\tmr_M(x)$ is the number of rejecting runs of $M$ with input $x$.
-%such that for some $\np$ machine, $f(x)$ is the number of accepting runs minus the number of rejecting runs.
 That is, $\gp$ is the closure of $\shp$ functions under subtraction, and its functions can obviously take negative values.
-%it can compute negative values by definition. 
-Given that our logical framework was built on top of the natural numbers, we need to extend $\qso$ in order to %describe functions
-capture $\gp$. 
+Given that our logical framework was built on top of the natural numbers, we need to extend $\qso$ in order to capture $\gp$. 
 The most elegant way to do this is by allowing constants coming from $\bbZ$ instead of just $\bbN$. 
@@
 Similar than for $\qso$, we define the fragment $\eqsoz$ as the extension of $\eqso$ with constants in $\bbZ$.
-\begin{example}
+\begin{exa}
 	Recall the setting of Example~\ref{ex:cliques} and suppose now that we want to compute the number of cliques in a graph that are not triangles. This can be easily done in $\qsoz$ with the formula:
@@
 	$.
-\end{example}
+\end{exa}
 Adding negative constants is a mild extension to allow subtraction in the logic. 
 It follows from our characterization of $\shp$ that this is exactly what we need to capture  $\gp$.
-\begin{corollary} \label{prop:capture-gapp}
+\begin{cor} \label{prop:capture-gapp}
 	$\eqsoz(\fo)$ captures $\gp$ over ordered structures.
-\end{corollary}
-%The previous result is straightforward to prove from the characterization of $\shp$ and the definition of $\gp$. 
-%However, 
-%We 
-%believe 
-%think
+\end{cor}
 This is an interesting result that shows how robust is $\qso$ when capturing different complexity classes.
@@
 Given a signature $\R$, the set of $\optqso$-formulae over $\R$ is given by extending the syntax in (\ref{syntax}) with the following operators:
-\begin{multline*}
-\max\{\alpha,\alpha\} \ \mid\ \min\{\alpha,\alpha\} \ \mid \\ \maxa{x} \alpha \ \mid \ \mina{x} \alpha \ \mid \ \maxa{X} \alpha \ \mid \ \mina{X} \alpha 
-\end{multline*}
+\begin{align*}
+\max\{\alpha,\alpha\} \ \mid\ \min\{\alpha,\alpha\} \ \mid \maxa{x} \alpha \ \mid \ \mina{x} \alpha \ \mid \ \maxa{X} \alpha \ \mid \ \mina{X} \alpha 
+\end{align*}
 where $x \in \fv$ and $X \in \sv$.
@@
 Furthermore, the semantics of the max and min quantifiers are defined as the natural extension of the sum quantifiers in $\qso$ (see Table~\ref{tab-semantics}) by maximizing or minimizing, respectively, instead of computing a sum.
-\begin{example}\label{ex:optqso}
+\begin{exa}\label{ex:optqso}
 	Recall again the setting of Example~\ref{ex:cliques} and suppose now that we want to compute the size of the largest clique in a graph. This can be done in $\optqso$ as follows:
@@
 	Notice that formula $\sa{z} X(z)$ is used to compute the number of nodes in a set $X$. 
-\end{example}
+\end{exa}
 Similar than for $\maxp$ and $\minp$, we have to distinguished between the $\max$ and $\min$ fragments of $\optqso$.
 For this, we define the fragment $\maxqso$ of all $\optqso$ formulae constructed from $\qfo$ operators and $\max$-formulae $\max\{\alpha,\alpha\}$, $\maxa{x} \alpha$ and  $\maxa{X} \alpha$.
-%operators of first- and second-order. 
 The class $\minqso$ is defined analogously changing $\max$ with $\min$.
@@
 For instance, formula $\alpha_6$ in Example \ref{ex:optqso} is in $\maxqso$.
-%the formulae $\alpha_1$ and $\alpha_6$ are in $\maxqso$ but $\alpha_2$ and $\alpha_3$ are not. 
 As one could expect, $\maxqso$ and $\minqso$ are the needed logics to capture $\maxp$ and $\minp$.
-\begin{theorem} \label{theo:capture-optp}
+\begin{thm} \label{theo:capture-optp}
 	$\maxqso(\fo)$ and $\minqso(\fo)$ capture $\maxp$ and $\minp$, respectively, over ordered structures.
-\end{theorem}
+\end{thm}
+\proof
+\input{proof-capture-optp}
+\qed
 It is important to mention that a similar result was proved in~\cite{kolaitis1994logical} for the class $\maxpb$ (resp., $\minpb$) of problems in $\maxp$ (resp., $\minp$) whose output value is polynomially bounded.

### 2017-03-03 LMCS/framework.tex
journal primera version

--- 2017-01-10 LICS/framework.tex
@@
 %\]
-\begin{multline}
-\alpha \ := \ \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \\ \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha \label{syntax}
-\end{multline}
+\begin{align}
+\alpha \ &:= \ \varphi \ \mid \ s \ \mid \ (\alpha \add \alpha) \ \mid\ (\alpha \mult \alpha) \ \mid \sa{x} \alpha \ \mid \ \pa{x} \alpha \ \mid \ \sa{X} \alpha \ \mid \ \pa{X} \alpha \label{syntax} 
+\end{align}
 where $\varphi$ is an $\so$-formula over $\R$, $s \in \bbN$, $x \in \fv$ and $X \in \sv$.
@@
 	\label{tab-semantics}
-	\vspace*{-21pt}
 \end{table}
-%$$
-%\renewcommand{\arraystretch}{1.7}
-%\begin{array}{rcl} 
-%\sem{\varphi}(\A, v, V) & = & 
-%\begin{cases}
-%1 & \mbox{if } (\A, v, V) \models \varphi \\
-%0 & \mbox{otherwise}
-%\end{cases}\\
-%\sem{s}(\A, v, V) & = & s \\
-%\sem{\alpha_1 \add \alpha_2}(\A, v, V) & = & \sem{\alpha_1}(\A, v, V) + \sem{\alpha_2}(\A, v, V)\\
-%\sem{\alpha_1 \mult \alpha_2}(\A, v, V) & = & \sem{\alpha_1}(\A, v, V) \cdot \sem{\alpha_2}(\A, v, V)\\ 
-%\sem{\sa{x} \alpha}(\A, v, V) & = & \displaystyle \sum_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
-%\sem{\pa{x} \alpha}(\A, v, V) & = & \displaystyle \prod_{a \in A} \sem{\alpha}(\A,v[a/x],V)\\
-%\sem{\sa{X} \alpha}(\A, v, V) & = & \displaystyle \sum_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])\\
-%\sem{\pa{X} \alpha}(\A, v, V) & = & \displaystyle \prod_{B \subseteq A^{\arity(X)}} \sem{\alpha}(\A, v, V[B/X])
-%\end{array}
-%$$
+
 A $\qso$-formula $\alpha$ is said to be a \emph{sentence} if it does not have any free variable, that is, every variable in $\alpha$ is under the scope of a usual quantifier or a quantitative quantifier.
@@
 Thus, in such a case we use the term $\sem{\alpha}(\A)$ to denote $\sem{\alpha}(\A, v, V)$, for some arbitrary first-order assignment $v$ for $\A$ and some arbitrary second-order assignment $V$ for $\A$.
-\begin{example}\label{ex:cliques}
+\begin{exa}\label{ex:cliques}
 Let $\bG = \{E(\cdot,\cdot),<\}$ be the vocabulary for graphs and $\fG$ be an ordered $\bG$-structure encoding a non-directed graph. 
 Suppose that we want to count the number of triangles in $\fG$. Then this can be defined as follows:
-\begin{multline*}
-\alpha_1 \ := \ \sa{x} \sa{y} \sa{z} ( E(x,y) \, \wedge \, E(y,z) \, \wedge \, E(z,x) \, \wedge \\
-x < y \, \wedge \, y < z )
-\end{multline*}
+\begin{align*}
+\alpha_1 \ &:= \ \sa{x} \sa{y} \sa{z} ( E(x,y) \, \wedge \, E(y,z) \, \wedge \, E(z,x) \, \wedge x < y \, \wedge \, y < z )
+\end{align*}
 We encode a triangle in $\alpha_1$ as an increasing sequence of nodes $\{x, y, z\}$, in order to count each triangle once. Then the Boolean subformula  $E(x,y) \wedge E(y,z) \wedge E(z,x) \wedge
@@
 
-Suppose now that we want to count the number of cliques in~$\fG$.
-%(i.e. set of nodes where every pair of nodes is connected). 
-We can 
-%easily 
-define this function with the following formula:
-$$
+Suppose now that we want to count the number of cliques in~$\fG$. We can define this function with the following formula:
+\[
 \alpha_2 \ := \ \sa{X} \clique(X),
-$$ 
+\] 
 where $\clique(X) := \forall x \forall y ((X(x) \wedge X(y) \wedge x \neq y)  \rightarrow E(x,y))$.
-%Similar than for $\alpha_1$, 
 In the Boolean sub-formula of $\alpha_2$ we check whether $X$ is a clique, and with the sum quantifier we add one for each clique in $\fG$. 
-%Note here that, 
 But in contrast to $\alpha_1$, 
@@
 This is according to the
-%inherent 
 complexity of evaluating each formula:
-%each formula: 
 $\alpha_1$ defines an $\fp$-function while $\alpha_2$ defines a $\shp$-complete function.
-%~\cite{paper-that-shows-that-this-problem-is-sharpP-hard}.
-\end{example}
-\begin{example}\label{exa-perm}
+\end{exa}
+\begin{exa}\label{exa-perm}
 For a more involved example that includes multiplication, let $\bM = \{M(\cdot,\cdot),<\}$ be a vocabulary for storing 0-1 matrices; in particular, a structure $\fM$ over $\bM$ encodes a 0-1 matrix $A$ as follows: if $A[i,j] = 1$, then $M(i,j)$ is true, otherwise $M(i.j)$ is false.
 Suppose now that we want to compute the permanent of an $n$-by-$n$ 0-1 matrix $A$, that is:
-\begin{eqnarray*}
-\op{perm}(A) & = & \sum_{\sigma \in S_n} \prod_{i=1}^n A[i, \sigma(i)],  
-\end{eqnarray*}
+\begin{align*}
+\op{perm}(A) &= \sum_{\sigma \in S_n} \prod_{i=1}^n A[i, \sigma(i)],  
+\end{align*}
 where $S_n$ is the set of all permutations over $\{1, \ldots, n\}$.
@@
 Indeed, an advantage of $\qso$-formulae is that the first- and second-order quantifiers in the quantitative level naturally reflect the operations used to define mathematical formulae.
-\end{example}
+\end{exa}
 
@@
 Moreover, we also restrict other fragments of $\qso$ by using the same idea.
-%In general, if $\FF$ is a fragment of $\qso$ we define $\FF(\LL)$ to be the fragment of $\FF$ obtained by restricting Boolean formulae to $\LL$.
 For example, we define $\qfo(\fo)$ to be the fragment of $\qfo$ obtained by restricting $\varphi$ in \eqref{syntax} to be an $\fo$-formula, and likewise for $\eqso(\fo)$.
-%, $\eqso(\fo)$ of $\qfo$ and $\eqso$ where $\varphi$ is restricted to $\fo$ formulae.
-
-%\marcelo{IMPORTANTE: Tenemos que decir que $\qfo(\fo)$ permite variables libre de segundo orden en las formulae de primer orden. Hay que aclarar esto tambien para PFP y otras logicas.}
 
 In the following section, we use different fragments of $\qso$ to capture counting complexity classes.
-%variate $\FF$ and $\LL$ to capture several function complexity classes. 
-But before doing this, we 
-%devote the next subsection to 
-show the connection of $\qso$ with previous frameworks for defining functions over relational structures.
+But before doing this, we show the connection of $\qso$ with previous frameworks for defining functions over relational structures.
 
@@
 In particular, we are not aware of any paper that uses this approach for capturing counting complexity classes.
-%Although there are some papers that uses this operator for counting~\cite{XXX}, they have not used the logic for capturing function complexity classes.
-
-%\begin{theorem} \label{no-mult}
-%	Let $\alpha$ be a $\qso$-formula in some fragment of $\qso$.
-There is a $\qso$ formula $\alpha'$ in the same fragment such that $\alpha'$ does not have any subformula of the form $(\beta_1 \cdot \beta_2)$ and $\sem{\alpha} = \sem{\alpha'}$.
-%\end{theorem}
 
@@
 In~\cite{SalujaST95}, it was proposed to define a function over a structure by using free variables in an SO-formula; in particular, the function is defined by the number of instantiations of the free variables that are satisfied by the structure.
-%and then counting how many extensions of an structure satisfy the formula (i.e. by instantiating the free variables). 
 Formally, Saluja et.
@@
 for every $\A\in\ostr[\R]$. Then a function $g:\ostr[\R]\to\nat$ is in $\#\LL$ if there exists a formula $\varphi(\bar{x},\bar{X})$ in $\LL$ such that $g = f_{\varphi(\bar x, \bar X)}$.
-%\martin{La notacion $f_{\psi(x,X)}$ la inventamos}
 In~\cite{SalujaST95}, they were proved several results about capturing counting complexity classes which are relevant for our work.

### 2017-03-03 LMCS/main.tex
journal primera version

\documentclass{lmcs} %%% last changed 2014-08-20

%% mandatory lists of keywords 
\keywords{MANDATORY list of keywords}

%% read in additional TeX-packages or personal macros here:
%% e.g. \usepackage{tikz}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{cite}
\usepackage{amsmath,amsfonts,amssymb,mathabx}
\usepackage{stmaryrd}
\input{macros}
%% define non-standard environments BEYOND the ones already supplied 
%% here, for example
\theoremstyle{plain}\newtheorem{satz}[thm]{Satz} %\crefname{satz}{Satz}{S\"atze}
%% Do NOT replace the proclamation environments lready provided by
%% your own.

\def\eg{{\em e.g.}}
\def\cf{{\em cf.}}

%% due to the dependence on amsart.cls, \begin{document} has to occur
%% BEFORE the title and author information:

\begin{document}

\title[Counting]{Descriptive Complexity for Counting Complexity Classes}
\titlecomment{{\lsuper*}OPTIONAL comment concerning the title, \eg, 
  if a variant or an extended abstract of the paper has appeared elsewhere.}

\author[M.~Arenas]{Marcelo Arenas}	%required
\address{PUC Chile}	%required
\email{marenas@ing.puc.cl}  %optional
%\thanks{thanks 1, optional.}	%optional

\author[M.~Munoz]{Martin Mu\~noz}	%optional
\address{PUC Chile}	%optional
\email{mmunos@uc.cl}  %optional
%\thanks{thanks 2, optional.}	%optional

\author[C.~Riveros]{Cristian Riveros}	%optional
\address{PUC Chile}	%optional
\email{cristian.riveros@uc.cl}  %optional
%\thanks{thanks 3, optional.}	%optional

%% etc.

%% required for running head on odd and even pages, use suitable
%% abbreviations in case of long titles and many authors:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% the abstract has to PRECEDE the command \maketitle:
%% be sure not to issue the \maketitle command twice!

\begin{abstract}
  \noindent Descriptive Complexity has been very successful in characterizing complexity classes of decision problems in terms of the properties definable in some logics. However, descriptive complexity for counting complexity classes, such as $\fp$ and $\shp$, has not been systematically studied, and it is not as developed as its decision counterpart. In this paper, we propose a framework based on Weighted Logics to address this issue. Specifically, by focusing on the natural numbers we obtain a logic called Quantitative Second Order Logics (QSO), and show how some of its fragments can be used to capture fundamental counting complexity classes such as $\fp$, $\shp$ and $\fpspace$, among others. We also use QSO to define a hierarchy inside $\shp$, identifying counting complexity classes with good closure and approximation properties, and which admit natural complete problems. Finally, we add recursion to QSO, and show how this extension naturally captures lower counting complexity classes such as $\shl$.
\end{abstract}

\maketitle

%% start the paper here:
\section{Introduction}
\input{introduction}

\section{Preliminaries} \label{sec:preliminaries}
\input{preliminaries}

\section{A logic for quantitative functions} \label{sec:logic}
\input{framework}

\section{Counting under $\qso$} \label{sec:complexity}
\input{counting}

\subsection{Extending $\qso$ to capture non-counting classes} \label{sec:extentions}
\input{extentions}

\section{Exploring the structure of $\shp$ through $\qso$} \label{sec:syntactic}
\input{syntactic}

\section{Adding recursion to QSO}\label{sec:beyond}
\input{beyond}

\section{Concluding remarks and future work}\label{sec:conclusions}
\input{conclusions}

\begin{thebibliography}{10}

\end{thebibliography}

\end{document}

### 2017-03-03 LMCS/preliminaries.tex
journal primera version

--- 2017-01-10 LICS/preliminaries.tex
@@
 Then given a  signature $\R$, the set of second-order logic formulae ($\so$-formulae) over $\R$ is given by the following grammar:
-\begin{multline*}\ 
-	\varphi \ := \ x = y \ \mid \ R(\bar u) \ \mid \ \top  \ \mid\  
-	X(\bar v)  \ \mid\\ 
+\begin{align*}\ 
+	\varphi \ &:= \ x = y \ \mid \ R(\bar u) \ \mid \ \top  \ \mid\  
+	X(\bar v)  \ \mid
 	\neg \varphi \ \mid\ 
@@
 	\exists X \, \varphi
-\end{multline*}
+ \end{align*}
 where $x,y \in \fv$, $R \in \R$, $\bar u$ is a tuple of (non-necessarily distinct) variables from $\fv$ whose length is $\arity(R)$, $\top$ is a reserved symbol to represent a tautology, $X \in \sv$, $\bar v$ is a tuple of (non-necessarily distinct) variables from $\fv$ whose length is $\arity(X)$, and $x \in \fv$.

### 2017-03-03 LMCS/proof-capture-fp.tex
journal primera version

--- 2017-03-03 LMCS/appendix-counting.tex
@@
-\medskip
-
-%%% DEMOSTRACION DE SQSO(FO) y #P
-\subsection*{Proof of Proposition~\ref{prop:capture-shP}}
-
-We construct recursively a $\shp$-machine $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\fo$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ deterministically in polynomial time, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-%If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-%If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$. 
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Let $\alpha$ be a formula in $\eqso(\fo)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shp$-function over $\R$ and $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the other direction, note that Saluja et al.~\cite{SalujaST95} proved that $\shp = \sfo$. 
-%We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso(\fo)$ be its corresponding formula. 
-Since a function in $\sfo$ can also be defined $\eqso(\fo)$ (see Section~\ref{sec:previous}), the condition holds. \qed
-
-\medskip
-
-%%% DEMOSTRACION DE SQSO(ESO) y span-P
-\subsection*{Proof of Proposition~\ref{prop:capture-spanP}}
-
-Similar than the previous proof, we construct recursively a $\spp$ machine $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ distinct accepting outputs for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time \cite{F75}, and accepts if and only if the condition holds true. 
-If $\alpha$ is a constant $s$, then the machine produces $s$ branches and accepts in all of them. 
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.  
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$. 
-If $\alpha = \sa{X} \beta$, it chooses $B\in A^{\arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$. 
-This covers all possible cases for $\alpha$.
-Additionally, the machine produces a different output on each path. This can be done by printing the trace of all the non-deterministic choices.
-However, when the machine starts checking whether $(\A,v,V)\models\varphi$ for some $\eso$ formula $\varphi$, it stops printing in the output tape.
-This way the machine produces exactly one output from that point onwards.
-Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\spp$ function over $\R$ and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the other direction, Compton et al.~\cite{ComptonG96} proved that $\spp = \#\eso$.
-Since a function in $\#\eso$ can also be defined in $\eqso(\eso)$, then $\eqso(\eso)$ captures $\spp$ over ordered structures.
-
-\medskip
-
-%%% DEMOSTRACION DE QFO(LFP) y FP
-\subsection*{Proof of Theorem~\ref{theo:capture-fp}}
-
 For the first condition, let $\alpha\in\qfo(\lfp)$ over some signature $\R$.
@@
 Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-\medskip
-
-%%% DEMOSTRACION DE QSO(PFP) y FPSPACE
-\subsection*{Proof of Theorem~\ref{theo:capture-fpspace}}
-
-To show how to evaluate a $\qso(\pfp)$-formula, we construct recursively a $\shpspace$-machine $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
-This machine runs in non-deterministic polynomial space and, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\pfp$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space~\cite{L04}, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$, continuing to the next $A$-value if the run of $M_{\beta}$ accepts.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-If $\alpha = \pa{X}\beta$, it simulates $M_{\beta}$ on input $(\A,v,V[B/X])$ consecutively for each $B\in A^{arity(X)}$, again continuing to the next $A$-value if the run of $M_{\beta}$ accepts.
-This covers all possible cases for $\alpha$, and each of these steps can be computed in polynomial space.
-Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$, by the fact that $\shpspace = \fpspace$ \cite{Ladner89}, and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the second condition, let $f\in \fpspace$ defined over some $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq 2^{n^\ell}$ (i.e.
-$2^{n^\ell}$ is an upper bound for the output size), where $\A$ has a domain of size $n$.
-Let $X$ be a second-order variable of arity $\ell$.
-Consider a linear order over predicates of arity $\ell$ given by the formula
-$$
-\varphi_{<}(X,Y) = \exists\bar{u}\big[\neg X(\bar{u})\wedge Y(\bar{u})\wedge \forall\bar{v}\big(
-\bar{u}<\bar{v}\to(X(\bar{u})\iff Y(\bar{v}))\big)\big].
-$$
-Namely, we use relations to encode numbers with at most $2^{n^\ell}$-bits where the empty relation represents $0$ and the total-relation represents $2^{2^{n^\ell}}-1$.
-Furthermore, each relation $X$ indexes a position in the binary output of $f(\enc(\A))$ as follows.
-Consider a polynomial space machine over the $\R$ that receives as input an $\R$-structure $\A$ and a number $p$ encoded by a relation $X$.
-Then the machine accepts if, and only if, the $p$-th bit of $f(\enc(\A))$ is $1$.
-Since this procedure works in polynomial space, it can be described in $\pfp$ \cite{AbiteboulV89} by a formula $\Phi(X)$ where the free variable $X$ encodes the number $p$.
-Then, similar than the previous proof we define:
-$$
-\alpha := \sa{X}(\Phi(X)\mult\varphi(X)),
-$$ 
-where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
-Note that for each $B\subseteq A^{\ell}$ such that $B$ is the $m$-th element in the order $\varphi_{<}(X,Y)$, it holds that $\sem{\varphi(B)}(\A) = 2^m$.
-Therefore, $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-\medskip
-
-%%% DEMOSTRACION DE QFO(PFP) y FPSPACE(poly)
-\subsection*{Proof of Corollary~\ref{cor:capture-fpspace-poly}}
-
-For the first condition, let $\alpha\in\qfo(\pfp)$ over some signature $\R$.
-Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{1,\ldots,n\}$.
-In the procedure we slightly extend the grammar of $\qfo(\pfp)$ to include constants.
-We replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\sa{x} \beta(x)$ is replaced by $(\beta(1)+\cdots+\beta(n))$ and $\pa{x}\beta(x)$ is replaced by $(\beta(1)\cdot\,\cdots\,\cdot\beta(n))$.
-Then each sub-formula $\varphi\in\pfp$ in $\alpha$ is evaluated in polynomial space and replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size  (recall that $\alpha$ is fixed) which is evaluated and lastly given as output.
-Note that $f\in\nfpspace$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
-
-For the second condition, let $f\in \fpspace$ defined over some signature $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
-Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
-Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
-Let $m$ be the position of $\bar{a}$ in the lexicographic order of the tuples in $A^{\ell}$.
-The procedure then computes the $m$-th bit of $f(\enc(\A))$, from least to most significant.
-Since this procedure works in polynomial space, it can be described by an $\pfp$ formula $\Phi(\bar{x})$.
-Then we use
-$$
-\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot\varphi(\bar{x}),
-$$
-where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
-Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-%%% DEMOSTRACION DE optP y MaxQSO, MinQSO
-\subsection*{Proof of Theorem~\ref{theo:capture-optp}}
-
-Similar than the previous proofs, we construct recursively a non-deterministic polynomial time Turing machine $M_{\alpha}$ with output tape for each $\maxqso(\fo)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ (in binary) over the output tape of some run, and this value is the maximum value over all runs.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\fo$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ in deterministic polynomial time, and output $1$ if, and only if, $(\A,v,V)$ satisfies $\varphi$.
-If $\alpha$ is a constant $s$, it outputs $s$ in binary over the output tape.
-If $\alpha = (\beta \add \gamma)$, then it simulates $M_{\beta}$ and $M_{\gamma}$ on input $(\A,v,V)$, adds the output of both machines and prints this value over the output tape.
-If $\alpha = (\beta \mult \gamma)$ or $\alpha = \max\{\beta,\gamma\}$, then it does the same than the previous case but it multiplies or maximizes, respectively, the outputs of both machines instead of adding.
-If $\alpha = \sa{x}\beta$, it iterates over all elements $a\in A$ simulating and adding the output of $M_{\beta}$ on input $(\A,v[a/x],V)$.
-$M_{\alpha}$ finally outputs the aggregated value in the output tape.
-If $\alpha = \pa{x}\beta$ or $\alpha = \maxa{x}\beta$, then $M_{\alpha}$ does the same idea than the previous case with the difference that the output of $M_{\beta}$ on input $(\A,v[a/x],V)$ is multiplied or maximized, respectively.
-If $\alpha = \maxa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Furthermore, it is straightforward to prove that each of these steps are correct and can be computed with a non-deterministic polynomial time Turing machine with output tape.
-Let $\alpha$ be a formula in $\maxqso$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the maximum run (with respect to the output value) of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\maxp$ function over $\R$ and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-Finally, one can easily see that the same construction holds with $\minqso(\fo)$ by constructing a Turing machine that take the min over all runs instead of max.
-
-The proof for the other direction is similar than in \cite{kolaitis1994logical} extended with the ideas of Theorem~\ref{theo:capture-fp}.
-Let $f\in \maxp$ be a function defined over some signature $\R$ and
-$\ell\in\nat$ such that $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$ for each $\A\in\ostr[\R]$ of size $n$.
-For $U \subseteq A^{\ell}$, we can interpret the encoding of $U$ ($\enc(U)$) as the binary encoding of a number with $n^l$-bits. We denote this value by $\val(\enc(U))$.
-Then, given $\A\in\ostr[\R]$ and $U \subseteq A^{\ell}$, consider the problem of checking whether $f(\enc(\A)) \geq \val(\enc(U))$. 
-Clearly, this is an $\np$-problem and, by Fagin's theorem, there exists a formula of the form $\ex{\bar{X}} \Phi(\bar{X}, Y)$ with $\Phi(\bar{X}, Y)$ in $\fo$-logic and $\arity(Y) = \ell$ such that $f(\enc(\A)) \geq \val(\enc(U))$ if, and only if, $(\A,v,V) \models \ex{\bar{X}} \Phi(\bar{X}, Y)$ with $V(Y) = U$.
-Then we can describe the function $f$ by the following $\maxqso$ formula:
-$$
-\alpha = \maxa{\bar{X}} \maxa{Y} \ \Phi(\bar{X}, Y) \cdot \big( \sa{\bar{x}} Y(\bar{x}) \cdot \pa{\bar{y}}(\bar{x} < \bar{y} \mapsto 2) \big).
-$$
-Note that, in contrast with previous proofs, we use $\bar{x} < \bar{y}$ instead of $\bar{y} < \bar{x}$ because the most significant bit in $\enc(U)$ correspond to the smallest tuple in $U$.  
-It is easy to check that $\Phi(\bar{X}, Y)$ simulates the NP-machine and, if $\Phi(\bar{X}, Y)$ holds, the formula to the right  reconstructs the binary output from the relation in $Y$.
-Then, $\alpha$ is in $\maxqso(\fo)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-For the case of $\minqso(\fo)$ and a function $f \in \minp$, one has to follow the same approach but considering the $\np$-problem of checking whether $f(\enc(\A)) \leq \val(\enc(U))$.
-Then, the formula for describing $f$ is the following:
-$$
-\alpha = \mina{\bar{X}} \mina{Y} \ \sa{\bar{x}} \big( (\Phi(\bar{X}, Y) \rightarrow Y(\bar{x})) \cdot \pa{\bar{y}}(\bar{x} < \bar{y} \mapsto 2)  \big).
-$$
-In this case, if the formula $\Phi(\bar{X}, Y)$ is false, then the output produced by the subformula inside the $\min$-quantifiers will be the biggest possible value (i.e. $2^{n^\ell}$).
-On the other hand, if $\Phi(\bar{X}, Y)$ holds, the subformula will produce $\val(\enc(U))$. 
-Similar than for $\max$, we conclude that $\alpha$ is in $\minqso(\fo)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.

### 2017-03-03 LMCS/proof-capture-fpspace-poly.tex
journal primera version

--- 2017-03-03 LMCS/appendix-counting.tex
@@
-\medskip
-
-%%% DEMOSTRACION DE SQSO(FO) y #P
-\subsection*{Proof of Proposition~\ref{prop:capture-shP}}
-
-We construct recursively a $\shp$-machine $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\fo$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ deterministically in polynomial time, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-%If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-%If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$. 
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Let $\alpha$ be a formula in $\eqso(\fo)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shp$-function over $\R$ and $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the other direction, note that Saluja et al.~\cite{SalujaST95} proved that $\shp = \sfo$. 
-%We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso(\fo)$ be its corresponding formula. 
-Since a function in $\sfo$ can also be defined $\eqso(\fo)$ (see Section~\ref{sec:previous}), the condition holds. \qed
-
-\medskip
-
-%%% DEMOSTRACION DE SQSO(ESO) y span-P
-\subsection*{Proof of Proposition~\ref{prop:capture-spanP}}
-
-Similar than the previous proof, we construct recursively a $\spp$ machine $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ distinct accepting outputs for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time \cite{F75}, and accepts if and only if the condition holds true. 
-If $\alpha$ is a constant $s$, then the machine produces $s$ branches and accepts in all of them. 
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.  
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$. 
-If $\alpha = \sa{X} \beta$, it chooses $B\in A^{\arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$. 
-This covers all possible cases for $\alpha$.
-Additionally, the machine produces a different output on each path. This can be done by printing the trace of all the non-deterministic choices.
-However, when the machine starts checking whether $(\A,v,V)\models\varphi$ for some $\eso$ formula $\varphi$, it stops printing in the output tape.
-This way the machine produces exactly one output from that point onwards.
-Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\spp$ function over $\R$ and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the other direction, Compton et al.~\cite{ComptonG96} proved that $\spp = \#\eso$.
-Since a function in $\#\eso$ can also be defined in $\eqso(\eso)$, then $\eqso(\eso)$ captures $\spp$ over ordered structures.
-
-\medskip
-
-%%% DEMOSTRACION DE QFO(LFP) y FP
-\subsection*{Proof of Theorem~\ref{theo:capture-fp}}
-
-For the first condition, let $\alpha\in\qfo(\lfp)$ over some signature $\R$.
-Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{1,\ldots,n\}$.
-In the procedure we slightly extend the grammar of $\qfo(\lfp)$ to include constants.
-We replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\sa{x} \beta(x)$ is replaced by $(\beta(1)+\cdots+\beta(n))$ and $\pa{x}\beta(x)$ is replaced by $(\beta(1)\cdot\,\cdots\,\cdot\beta(n))$.
-Then each sub-formula $\varphi\in\lfp$ in $\alpha$ is evaluated in polynomial time and replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size (recall that $\alpha$ is fixed) which is evaluated and lastly given as output.
-Note that $f\in\fp$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
-
-For the second condition, let $f\in \fp$ defined over some signature $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$ (i.e. $n^\ell$ is an upper bound for the output size), where $\A$ has a domain of size $n$.
-Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
-Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
-Let $m$ be the position of $\bar{a}$ in the lexicographic order of the tuples in $A^{\ell}$.
-The procedure then computes the $m$-th bit of $f(\enc(\A))$, from least to most significant.
-Since this procedure works in polynomial time, it can be described by an $\lfp$ formula $\Phi(\bar{x})$.
-Then we use
-$$
-\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot\varphi(\bar{x}),
-$$
-where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
-Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-In other words, $\Phi(\bar{x})$ simulates the behavior of the $\fp$-machine and the formula $\alpha$ reconstruct the binary output.
-Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-\medskip
-
-%%% DEMOSTRACION DE QSO(PFP) y FPSPACE
-\subsection*{Proof of Theorem~\ref{theo:capture-fpspace}}
-
-To show how to evaluate a $\qso(\pfp)$-formula, we construct recursively a $\shpspace$-machine $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
-This machine runs in non-deterministic polynomial space and, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\pfp$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space~\cite{L04}, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$, continuing to the next $A$-value if the run of $M_{\beta}$ accepts.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-If $\alpha = \pa{X}\beta$, it simulates $M_{\beta}$ on input $(\A,v,V[B/X])$ consecutively for each $B\in A^{arity(X)}$, again continuing to the next $A$-value if the run of $M_{\beta}$ accepts.
-This covers all possible cases for $\alpha$, and each of these steps can be computed in polynomial space.
-Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$, by the fact that $\shpspace = \fpspace$ \cite{Ladner89}, and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the second condition, let $f\in \fpspace$ defined over some $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq 2^{n^\ell}$ (i.e.
-$2^{n^\ell}$ is an upper bound for the output size), where $\A$ has a domain of size $n$.
-Let $X$ be a second-order variable of arity $\ell$.
-Consider a linear order over predicates of arity $\ell$ given by the formula
-$$
-\varphi_{<}(X,Y) = \exists\bar{u}\big[\neg X(\bar{u})\wedge Y(\bar{u})\wedge \forall\bar{v}\big(
-\bar{u}<\bar{v}\to(X(\bar{u})\iff Y(\bar{v}))\big)\big].
-$$
-Namely, we use relations to encode numbers with at most $2^{n^\ell}$-bits where the empty relation represents $0$ and the total-relation represents $2^{2^{n^\ell}}-1$.
-Furthermore, each relation $X$ indexes a position in the binary output of $f(\enc(\A))$ as follows.
-Consider a polynomial space machine over the $\R$ that receives as input an $\R$-structure $\A$ and a number $p$ encoded by a relation $X$.
-Then the machine accepts if, and only if, the $p$-th bit of $f(\enc(\A))$ is $1$.
-Since this procedure works in polynomial space, it can be described in $\pfp$ \cite{AbiteboulV89} by a formula $\Phi(X)$ where the free variable $X$ encodes the number $p$.
-Then, similar than the previous proof we define:
-$$
-\alpha := \sa{X}(\Phi(X)\mult\varphi(X)),
-$$ 
-where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
-Note that for each $B\subseteq A^{\ell}$ such that $B$ is the $m$-th element in the order $\varphi_{<}(X,Y)$, it holds that $\sem{\varphi(B)}(\A) = 2^m$.
-Therefore, $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-\medskip
-
-%%% DEMOSTRACION DE QFO(PFP) y FPSPACE(poly)
-\subsection*{Proof of Corollary~\ref{cor:capture-fpspace-poly}}
-
 For the first condition, let $\alpha\in\qfo(\pfp)$ over some signature $\R$.
@@
 Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-%%% DEMOSTRACION DE optP y MaxQSO, MinQSO
-\subsection*{Proof of Theorem~\ref{theo:capture-optp}}
-
-Similar than the previous proofs, we construct recursively a non-deterministic polynomial time Turing machine $M_{\alpha}$ with output tape for each $\maxqso(\fo)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ (in binary) over the output tape of some run, and this value is the maximum value over all runs.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\fo$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ in deterministic polynomial time, and output $1$ if, and only if, $(\A,v,V)$ satisfies $\varphi$.
-If $\alpha$ is a constant $s$, it outputs $s$ in binary over the output tape.
-If $\alpha = (\beta \add \gamma)$, then it simulates $M_{\beta}$ and $M_{\gamma}$ on input $(\A,v,V)$, adds the output of both machines and prints this value over the output tape.
-If $\alpha = (\beta \mult \gamma)$ or $\alpha = \max\{\beta,\gamma\}$, then it does the same than the previous case but it multiplies or maximizes, respectively, the outputs of both machines instead of adding.
-If $\alpha = \sa{x}\beta$, it iterates over all elements $a\in A$ simulating and adding the output of $M_{\beta}$ on input $(\A,v[a/x],V)$.
-$M_{\alpha}$ finally outputs the aggregated value in the output tape.
-If $\alpha = \pa{x}\beta$ or $\alpha = \maxa{x}\beta$, then $M_{\alpha}$ does the same idea than the previous case with the difference that the output of $M_{\beta}$ on input $(\A,v[a/x],V)$ is multiplied or maximized, respectively.
-If $\alpha = \maxa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Furthermore, it is straightforward to prove that each of these steps are correct and can be computed with a non-deterministic polynomial time Turing machine with output tape.
-Let $\alpha$ be a formula in $\maxqso$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the maximum run (with respect to the output value) of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\maxp$ function over $\R$ and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-Finally, one can easily see that the same construction holds with $\minqso(\fo)$ by constructing a Turing machine that take the min over all runs instead of max.
-
-The proof for the other direction is similar than in \cite{kolaitis1994logical} extended with the ideas of Theorem~\ref{theo:capture-fp}.
-Let $f\in \maxp$ be a function defined over some signature $\R$ and
-$\ell\in\nat$ such that $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$ for each $\A\in\ostr[\R]$ of size $n$.
-For $U \subseteq A^{\ell}$, we can interpret the encoding of $U$ ($\enc(U)$) as the binary encoding of a number with $n^l$-bits. We denote this value by $\val(\enc(U))$.
-Then, given $\A\in\ostr[\R]$ and $U \subseteq A^{\ell}$, consider the problem of checking whether $f(\enc(\A)) \geq \val(\enc(U))$. 
-Clearly, this is an $\np$-problem and, by Fagin's theorem, there exists a formula of the form $\ex{\bar{X}} \Phi(\bar{X}, Y)$ with $\Phi(\bar{X}, Y)$ in $\fo$-logic and $\arity(Y) = \ell$ such that $f(\enc(\A)) \geq \val(\enc(U))$ if, and only if, $(\A,v,V) \models \ex{\bar{X}} \Phi(\bar{X}, Y)$ with $V(Y) = U$.
-Then we can describe the function $f$ by the following $\maxqso$ formula:
-$$
-\alpha = \maxa{\bar{X}} \maxa{Y} \ \Phi(\bar{X}, Y) \cdot \big( \sa{\bar{x}} Y(\bar{x}) \cdot \pa{\bar{y}}(\bar{x} < \bar{y} \mapsto 2) \big).
-$$
-Note that, in contrast with previous proofs, we use $\bar{x} < \bar{y}$ instead of $\bar{y} < \bar{x}$ because the most significant bit in $\enc(U)$ correspond to the smallest tuple in $U$.  
-It is easy to check that $\Phi(\bar{X}, Y)$ simulates the NP-machine and, if $\Phi(\bar{X}, Y)$ holds, the formula to the right  reconstructs the binary output from the relation in $Y$.
-Then, $\alpha$ is in $\maxqso(\fo)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-For the case of $\minqso(\fo)$ and a function $f \in \minp$, one has to follow the same approach but considering the $\np$-problem of checking whether $f(\enc(\A)) \leq \val(\enc(U))$.
-Then, the formula for describing $f$ is the following:
-$$
-\alpha = \mina{\bar{X}} \mina{Y} \ \sa{\bar{x}} \big( (\Phi(\bar{X}, Y) \rightarrow Y(\bar{x})) \cdot \pa{\bar{y}}(\bar{x} < \bar{y} \mapsto 2)  \big).
-$$
-In this case, if the formula $\Phi(\bar{X}, Y)$ is false, then the output produced by the subformula inside the $\min$-quantifiers will be the biggest possible value (i.e. $2^{n^\ell}$).
-On the other hand, if $\Phi(\bar{X}, Y)$ holds, the subformula will produce $\val(\enc(U))$. 
-Similar than for $\max$, we conclude that $\alpha$ is in $\minqso(\fo)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.

### 2017-03-03 LMCS/proof-capture-fpspace.tex
journal primera version

--- 2017-03-03 LMCS/appendix-counting.tex
@@
-\medskip
-
-%%% DEMOSTRACION DE SQSO(FO) y #P
-\subsection*{Proof of Proposition~\ref{prop:capture-shP}}
-
-We construct recursively a $\shp$-machine $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\fo$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ deterministically in polynomial time, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-%If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-%If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$. 
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Let $\alpha$ be a formula in $\eqso(\fo)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shp$-function over $\R$ and $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the other direction, note that Saluja et al.~\cite{SalujaST95} proved that $\shp = \sfo$. 
-%We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso(\fo)$ be its corresponding formula. 
-Since a function in $\sfo$ can also be defined $\eqso(\fo)$ (see Section~\ref{sec:previous}), the condition holds. \qed
-
-\medskip
-
-%%% DEMOSTRACION DE SQSO(ESO) y span-P
-\subsection*{Proof of Proposition~\ref{prop:capture-spanP}}
-
-Similar than the previous proof, we construct recursively a $\spp$ machine $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ distinct accepting outputs for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time \cite{F75}, and accepts if and only if the condition holds true. 
-If $\alpha$ is a constant $s$, then the machine produces $s$ branches and accepts in all of them. 
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.  
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$. 
-If $\alpha = \sa{X} \beta$, it chooses $B\in A^{\arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$. 
-This covers all possible cases for $\alpha$.
-Additionally, the machine produces a different output on each path. This can be done by printing the trace of all the non-deterministic choices.
-However, when the machine starts checking whether $(\A,v,V)\models\varphi$ for some $\eso$ formula $\varphi$, it stops printing in the output tape.
-This way the machine produces exactly one output from that point onwards.
-Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\spp$ function over $\R$ and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the other direction, Compton et al.~\cite{ComptonG96} proved that $\spp = \#\eso$.
-Since a function in $\#\eso$ can also be defined in $\eqso(\eso)$, then $\eqso(\eso)$ captures $\spp$ over ordered structures.
-
-\medskip
-
-%%% DEMOSTRACION DE QFO(LFP) y FP
-\subsection*{Proof of Theorem~\ref{theo:capture-fp}}
-
-For the first condition, let $\alpha\in\qfo(\lfp)$ over some signature $\R$.
-Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{1,\ldots,n\}$.
-In the procedure we slightly extend the grammar of $\qfo(\lfp)$ to include constants.
-We replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\sa{x} \beta(x)$ is replaced by $(\beta(1)+\cdots+\beta(n))$ and $\pa{x}\beta(x)$ is replaced by $(\beta(1)\cdot\,\cdots\,\cdot\beta(n))$.
-Then each sub-formula $\varphi\in\lfp$ in $\alpha$ is evaluated in polynomial time and replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size (recall that $\alpha$ is fixed) which is evaluated and lastly given as output.
-Note that $f\in\fp$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
-
-For the second condition, let $f\in \fp$ defined over some signature $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$ (i.e. $n^\ell$ is an upper bound for the output size), where $\A$ has a domain of size $n$.
-Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
-Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
-Let $m$ be the position of $\bar{a}$ in the lexicographic order of the tuples in $A^{\ell}$.
-The procedure then computes the $m$-th bit of $f(\enc(\A))$, from least to most significant.
-Since this procedure works in polynomial time, it can be described by an $\lfp$ formula $\Phi(\bar{x})$.
-Then we use
-$$
-\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot\varphi(\bar{x}),
-$$
-where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
-Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-In other words, $\Phi(\bar{x})$ simulates the behavior of the $\fp$-machine and the formula $\alpha$ reconstruct the binary output.
-Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-\medskip
-
-%%% DEMOSTRACION DE QSO(PFP) y FPSPACE
-\subsection*{Proof of Theorem~\ref{theo:capture-fpspace}}
-
 To show how to evaluate a $\qso(\pfp)$-formula, we construct recursively a $\shpspace$-machine $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
@@
 Therefore, $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-\medskip
-
-%%% DEMOSTRACION DE QFO(PFP) y FPSPACE(poly)
-\subsection*{Proof of Corollary~\ref{cor:capture-fpspace-poly}}
-
-For the first condition, let $\alpha\in\qfo(\pfp)$ over some signature $\R$.
-Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{1,\ldots,n\}$.
-In the procedure we slightly extend the grammar of $\qfo(\pfp)$ to include constants.
-We replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\sa{x} \beta(x)$ is replaced by $(\beta(1)+\cdots+\beta(n))$ and $\pa{x}\beta(x)$ is replaced by $(\beta(1)\cdot\,\cdots\,\cdot\beta(n))$.
-Then each sub-formula $\varphi\in\pfp$ in $\alpha$ is evaluated in polynomial space and replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size  (recall that $\alpha$ is fixed) which is evaluated and lastly given as output.
-Note that $f\in\nfpspace$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
-
-For the second condition, let $f\in \fpspace$ defined over some signature $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
-Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
-Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
-Let $m$ be the position of $\bar{a}$ in the lexicographic order of the tuples in $A^{\ell}$.
-The procedure then computes the $m$-th bit of $f(\enc(\A))$, from least to most significant.
-Since this procedure works in polynomial space, it can be described by an $\pfp$ formula $\Phi(\bar{x})$.
-Then we use
-$$
-\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot\varphi(\bar{x}),
-$$
-where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
-Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-%%% DEMOSTRACION DE optP y MaxQSO, MinQSO
-\subsection*{Proof of Theorem~\ref{theo:capture-optp}}
-
-Similar than the previous proofs, we construct recursively a non-deterministic polynomial time Turing machine $M_{\alpha}$ with output tape for each $\maxqso(\fo)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ (in binary) over the output tape of some run, and this value is the maximum value over all runs.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\fo$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ in deterministic polynomial time, and output $1$ if, and only if, $(\A,v,V)$ satisfies $\varphi$.
-If $\alpha$ is a constant $s$, it outputs $s$ in binary over the output tape.
-If $\alpha = (\beta \add \gamma)$, then it simulates $M_{\beta}$ and $M_{\gamma}$ on input $(\A,v,V)$, adds the output of both machines and prints this value over the output tape.
-If $\alpha = (\beta \mult \gamma)$ or $\alpha = \max\{\beta,\gamma\}$, then it does the same than the previous case but it multiplies or maximizes, respectively, the outputs of both machines instead of adding.
-If $\alpha = \sa{x}\beta$, it iterates over all elements $a\in A$ simulating and adding the output of $M_{\beta}$ on input $(\A,v[a/x],V)$.
-$M_{\alpha}$ finally outputs the aggregated value in the output tape.
-If $\alpha = \pa{x}\beta$ or $\alpha = \maxa{x}\beta$, then $M_{\alpha}$ does the same idea than the previous case with the difference that the output of $M_{\beta}$ on input $(\A,v[a/x],V)$ is multiplied or maximized, respectively.
-If $\alpha = \maxa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Furthermore, it is straightforward to prove that each of these steps are correct and can be computed with a non-deterministic polynomial time Turing machine with output tape.
-Let $\alpha$ be a formula in $\maxqso$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the maximum run (with respect to the output value) of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\maxp$ function over $\R$ and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-Finally, one can easily see that the same construction holds with $\minqso(\fo)$ by constructing a Turing machine that take the min over all runs instead of max.
-
-The proof for the other direction is similar than in \cite{kolaitis1994logical} extended with the ideas of Theorem~\ref{theo:capture-fp}.
-Let $f\in \maxp$ be a function defined over some signature $\R$ and
-$\ell\in\nat$ such that $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$ for each $\A\in\ostr[\R]$ of size $n$.
-For $U \subseteq A^{\ell}$, we can interpret the encoding of $U$ ($\enc(U)$) as the binary encoding of a number with $n^l$-bits. We denote this value by $\val(\enc(U))$.
-Then, given $\A\in\ostr[\R]$ and $U \subseteq A^{\ell}$, consider the problem of checking whether $f(\enc(\A)) \geq \val(\enc(U))$. 
-Clearly, this is an $\np$-problem and, by Fagin's theorem, there exists a formula of the form $\ex{\bar{X}} \Phi(\bar{X}, Y)$ with $\Phi(\bar{X}, Y)$ in $\fo$-logic and $\arity(Y) = \ell$ such that $f(\enc(\A)) \geq \val(\enc(U))$ if, and only if, $(\A,v,V) \models \ex{\bar{X}} \Phi(\bar{X}, Y)$ with $V(Y) = U$.
-Then we can describe the function $f$ by the following $\maxqso$ formula:
-$$
-\alpha = \maxa{\bar{X}} \maxa{Y} \ \Phi(\bar{X}, Y) \cdot \big( \sa{\bar{x}} Y(\bar{x}) \cdot \pa{\bar{y}}(\bar{x} < \bar{y} \mapsto 2) \big).
-$$
-Note that, in contrast with previous proofs, we use $\bar{x} < \bar{y}$ instead of $\bar{y} < \bar{x}$ because the most significant bit in $\enc(U)$ correspond to the smallest tuple in $U$.  
-It is easy to check that $\Phi(\bar{X}, Y)$ simulates the NP-machine and, if $\Phi(\bar{X}, Y)$ holds, the formula to the right  reconstructs the binary output from the relation in $Y$.
-Then, $\alpha$ is in $\maxqso(\fo)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-For the case of $\minqso(\fo)$ and a function $f \in \minp$, one has to follow the same approach but considering the $\np$-problem of checking whether $f(\enc(\A)) \leq \val(\enc(U))$.
-Then, the formula for describing $f$ is the following:
-$$
-\alpha = \mina{\bar{X}} \mina{Y} \ \sa{\bar{x}} \big( (\Phi(\bar{X}, Y) \rightarrow Y(\bar{x})) \cdot \pa{\bar{y}}(\bar{x} < \bar{y} \mapsto 2)  \big).
-$$
-In this case, if the formula $\Phi(\bar{X}, Y)$ is false, then the output produced by the subformula inside the $\min$-quantifiers will be the biggest possible value (i.e. $2^{n^\ell}$).
-On the other hand, if $\Phi(\bar{X}, Y)$ holds, the subformula will produce $\val(\enc(U))$. 
-Similar than for $\max$, we conclude that $\alpha$ is in $\minqso(\fo)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.

### 2017-03-03 LMCS/proof-capture-optp.tex
journal primera version

--- 2017-03-03 LMCS/appendix-counting.tex
@@
-\medskip
-
-%%% DEMOSTRACION DE SQSO(FO) y #P
-\subsection*{Proof of Proposition~\ref{prop:capture-shP}}
-
-We construct recursively a $\shp$-machine $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\fo$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ deterministically in polynomial time, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-%If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-%If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$. 
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Let $\alpha$ be a formula in $\eqso(\fo)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shp$-function over $\R$ and $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the other direction, note that Saluja et al.~\cite{SalujaST95} proved that $\shp = \sfo$. 
-%We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso(\fo)$ be its corresponding formula. 
-Since a function in $\sfo$ can also be defined $\eqso(\fo)$ (see Section~\ref{sec:previous}), the condition holds. \qed
-
-\medskip
-
-%%% DEMOSTRACION DE SQSO(ESO) y span-P
-\subsection*{Proof of Proposition~\ref{prop:capture-spanP}}
-
-Similar than the previous proof, we construct recursively a $\spp$ machine $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ distinct accepting outputs for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time \cite{F75}, and accepts if and only if the condition holds true. 
-If $\alpha$ is a constant $s$, then the machine produces $s$ branches and accepts in all of them. 
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.  
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$. 
-If $\alpha = \sa{X} \beta$, it chooses $B\in A^{\arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$. 
-This covers all possible cases for $\alpha$.
-Additionally, the machine produces a different output on each path. This can be done by printing the trace of all the non-deterministic choices.
-However, when the machine starts checking whether $(\A,v,V)\models\varphi$ for some $\eso$ formula $\varphi$, it stops printing in the output tape.
-This way the machine produces exactly one output from that point onwards.
-Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\spp$ function over $\R$ and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the other direction, Compton et al.~\cite{ComptonG96} proved that $\spp = \#\eso$.
-Since a function in $\#\eso$ can also be defined in $\eqso(\eso)$, then $\eqso(\eso)$ captures $\spp$ over ordered structures.
-
-\medskip
-
-%%% DEMOSTRACION DE QFO(LFP) y FP
-\subsection*{Proof of Theorem~\ref{theo:capture-fp}}
-
-For the first condition, let $\alpha\in\qfo(\lfp)$ over some signature $\R$.
-Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{1,\ldots,n\}$.
-In the procedure we slightly extend the grammar of $\qfo(\lfp)$ to include constants.
-We replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\sa{x} \beta(x)$ is replaced by $(\beta(1)+\cdots+\beta(n))$ and $\pa{x}\beta(x)$ is replaced by $(\beta(1)\cdot\,\cdots\,\cdot\beta(n))$.
-Then each sub-formula $\varphi\in\lfp$ in $\alpha$ is evaluated in polynomial time and replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size (recall that $\alpha$ is fixed) which is evaluated and lastly given as output.
-Note that $f\in\fp$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
-
-For the second condition, let $f\in \fp$ defined over some signature $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$ (i.e. $n^\ell$ is an upper bound for the output size), where $\A$ has a domain of size $n$.
-Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
-Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
-Let $m$ be the position of $\bar{a}$ in the lexicographic order of the tuples in $A^{\ell}$.
-The procedure then computes the $m$-th bit of $f(\enc(\A))$, from least to most significant.
-Since this procedure works in polynomial time, it can be described by an $\lfp$ formula $\Phi(\bar{x})$.
-Then we use
-$$
-\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot\varphi(\bar{x}),
-$$
-where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
-Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-In other words, $\Phi(\bar{x})$ simulates the behavior of the $\fp$-machine and the formula $\alpha$ reconstruct the binary output.
-Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-\medskip
-
-%%% DEMOSTRACION DE QSO(PFP) y FPSPACE
-\subsection*{Proof of Theorem~\ref{theo:capture-fpspace}}
-
-To show how to evaluate a $\qso(\pfp)$-formula, we construct recursively a $\shpspace$-machine $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
-This machine runs in non-deterministic polynomial space and, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\pfp$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space~\cite{L04}, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$, continuing to the next $A$-value if the run of $M_{\beta}$ accepts.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-If $\alpha = \pa{X}\beta$, it simulates $M_{\beta}$ on input $(\A,v,V[B/X])$ consecutively for each $B\in A^{arity(X)}$, again continuing to the next $A$-value if the run of $M_{\beta}$ accepts.
-This covers all possible cases for $\alpha$, and each of these steps can be computed in polynomial space.
-Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$, by the fact that $\shpspace = \fpspace$ \cite{Ladner89}, and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the second condition, let $f\in \fpspace$ defined over some $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq 2^{n^\ell}$ (i.e.
-$2^{n^\ell}$ is an upper bound for the output size), where $\A$ has a domain of size $n$.
-Let $X$ be a second-order variable of arity $\ell$.
-Consider a linear order over predicates of arity $\ell$ given by the formula
-$$
-\varphi_{<}(X,Y) = \exists\bar{u}\big[\neg X(\bar{u})\wedge Y(\bar{u})\wedge \forall\bar{v}\big(
-\bar{u}<\bar{v}\to(X(\bar{u})\iff Y(\bar{v}))\big)\big].
-$$
-Namely, we use relations to encode numbers with at most $2^{n^\ell}$-bits where the empty relation represents $0$ and the total-relation represents $2^{2^{n^\ell}}-1$.
-Furthermore, each relation $X$ indexes a position in the binary output of $f(\enc(\A))$ as follows.
-Consider a polynomial space machine over the $\R$ that receives as input an $\R$-structure $\A$ and a number $p$ encoded by a relation $X$.
-Then the machine accepts if, and only if, the $p$-th bit of $f(\enc(\A))$ is $1$.
-Since this procedure works in polynomial space, it can be described in $\pfp$ \cite{AbiteboulV89} by a formula $\Phi(X)$ where the free variable $X$ encodes the number $p$.
-Then, similar than the previous proof we define:
-$$
-\alpha := \sa{X}(\Phi(X)\mult\varphi(X)),
-$$ 
-where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
-Note that for each $B\subseteq A^{\ell}$ such that $B$ is the $m$-th element in the order $\varphi_{<}(X,Y)$, it holds that $\sem{\varphi(B)}(\A) = 2^m$.
-Therefore, $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-\medskip
-
-%%% DEMOSTRACION DE QFO(PFP) y FPSPACE(poly)
-\subsection*{Proof of Corollary~\ref{cor:capture-fpspace-poly}}
-
-For the first condition, let $\alpha\in\qfo(\pfp)$ over some signature $\R$.
-Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{1,\ldots,n\}$.
-In the procedure we slightly extend the grammar of $\qfo(\pfp)$ to include constants.
-We replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\sa{x} \beta(x)$ is replaced by $(\beta(1)+\cdots+\beta(n))$ and $\pa{x}\beta(x)$ is replaced by $(\beta(1)\cdot\,\cdots\,\cdot\beta(n))$.
-Then each sub-formula $\varphi\in\pfp$ in $\alpha$ is evaluated in polynomial space and replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size  (recall that $\alpha$ is fixed) which is evaluated and lastly given as output.
-Note that $f\in\nfpspace$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
-
-For the second condition, let $f\in \fpspace$ defined over some signature $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
-Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
-Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
-Let $m$ be the position of $\bar{a}$ in the lexicographic order of the tuples in $A^{\ell}$.
-The procedure then computes the $m$-th bit of $f(\enc(\A))$, from least to most significant.
-Since this procedure works in polynomial space, it can be described by an $\pfp$ formula $\Phi(\bar{x})$.
-Then we use
-$$
-\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot\varphi(\bar{x}),
-$$
-where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
-Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-%%% DEMOSTRACION DE optP y MaxQSO, MinQSO
-\subsection*{Proof of Theorem~\ref{theo:capture-optp}}
-
 Similar than the previous proofs, we construct recursively a non-deterministic polynomial time Turing machine $M_{\alpha}$ with output tape for each $\maxqso(\fo)$ formula $\alpha$ over a signature $\R$.

### 2017-03-03 LMCS/proof-capture-shP.tex
journal primera version

--- 2017-03-03 LMCS/appendix-counting.tex
@@
-\medskip
-
-%%% DEMOSTRACION DE SQSO(FO) y #P
-\subsection*{Proof of Proposition~\ref{prop:capture-shP}}
-
 We construct recursively a $\shp$-machine $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
@@
 %We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso(\fo)$ be its corresponding formula. 
-Since a function in $\sfo$ can also be defined $\eqso(\fo)$ (see Section~\ref{sec:previous}), the condition holds. \qed
-
-\medskip
-
-%%% DEMOSTRACION DE SQSO(ESO) y span-P
-\subsection*{Proof of Proposition~\ref{prop:capture-spanP}}
-
-Similar than the previous proof, we construct recursively a $\spp$ machine $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ distinct accepting outputs for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time \cite{F75}, and accepts if and only if the condition holds true. 
-If $\alpha$ is a constant $s$, then the machine produces $s$ branches and accepts in all of them. 
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.  
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$. 
-If $\alpha = \sa{X} \beta$, it chooses $B\in A^{\arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$. 
-This covers all possible cases for $\alpha$.
-Additionally, the machine produces a different output on each path. This can be done by printing the trace of all the non-deterministic choices.
-However, when the machine starts checking whether $(\A,v,V)\models\varphi$ for some $\eso$ formula $\varphi$, it stops printing in the output tape.
-This way the machine produces exactly one output from that point onwards.
-Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\spp$ function over $\R$ and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the other direction, Compton et al.~\cite{ComptonG96} proved that $\spp = \#\eso$.
-Since a function in $\#\eso$ can also be defined in $\eqso(\eso)$, then $\eqso(\eso)$ captures $\spp$ over ordered structures.
-
-\medskip
-
-%%% DEMOSTRACION DE QFO(LFP) y FP
-\subsection*{Proof of Theorem~\ref{theo:capture-fp}}
-
-For the first condition, let $\alpha\in\qfo(\lfp)$ over some signature $\R$.
-Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{1,\ldots,n\}$.
-In the procedure we slightly extend the grammar of $\qfo(\lfp)$ to include constants.
-We replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\sa{x} \beta(x)$ is replaced by $(\beta(1)+\cdots+\beta(n))$ and $\pa{x}\beta(x)$ is replaced by $(\beta(1)\cdot\,\cdots\,\cdot\beta(n))$.
-Then each sub-formula $\varphi\in\lfp$ in $\alpha$ is evaluated in polynomial time and replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size (recall that $\alpha$ is fixed) which is evaluated and lastly given as output.
-Note that $f\in\fp$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
-
-For the second condition, let $f\in \fp$ defined over some signature $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$ (i.e. $n^\ell$ is an upper bound for the output size), where $\A$ has a domain of size $n$.
-Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
-Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
-Let $m$ be the position of $\bar{a}$ in the lexicographic order of the tuples in $A^{\ell}$.
-The procedure then computes the $m$-th bit of $f(\enc(\A))$, from least to most significant.
-Since this procedure works in polynomial time, it can be described by an $\lfp$ formula $\Phi(\bar{x})$.
-Then we use
-$$
-\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot\varphi(\bar{x}),
-$$
-where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
-Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-In other words, $\Phi(\bar{x})$ simulates the behavior of the $\fp$-machine and the formula $\alpha$ reconstruct the binary output.
-Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-\medskip
-
-%%% DEMOSTRACION DE QSO(PFP) y FPSPACE
-\subsection*{Proof of Theorem~\ref{theo:capture-fpspace}}
-
-To show how to evaluate a $\qso(\pfp)$-formula, we construct recursively a $\shpspace$-machine $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
-This machine runs in non-deterministic polynomial space and, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\pfp$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space~\cite{L04}, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$, continuing to the next $A$-value if the run of $M_{\beta}$ accepts.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-If $\alpha = \pa{X}\beta$, it simulates $M_{\beta}$ on input $(\A,v,V[B/X])$ consecutively for each $B\in A^{arity(X)}$, again continuing to the next $A$-value if the run of $M_{\beta}$ accepts.
-This covers all possible cases for $\alpha$, and each of these steps can be computed in polynomial space.
-Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$, by the fact that $\shpspace = \fpspace$ \cite{Ladner89}, and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the second condition, let $f\in \fpspace$ defined over some $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq 2^{n^\ell}$ (i.e.
-$2^{n^\ell}$ is an upper bound for the output size), where $\A$ has a domain of size $n$.
-Let $X$ be a second-order variable of arity $\ell$.
-Consider a linear order over predicates of arity $\ell$ given by the formula
-$$
-\varphi_{<}(X,Y) = \exists\bar{u}\big[\neg X(\bar{u})\wedge Y(\bar{u})\wedge \forall\bar{v}\big(
-\bar{u}<\bar{v}\to(X(\bar{u})\iff Y(\bar{v}))\big)\big].
-$$
-Namely, we use relations to encode numbers with at most $2^{n^\ell}$-bits where the empty relation represents $0$ and the total-relation represents $2^{2^{n^\ell}}-1$.
-Furthermore, each relation $X$ indexes a position in the binary output of $f(\enc(\A))$ as follows.
-Consider a polynomial space machine over the $\R$ that receives as input an $\R$-structure $\A$ and a number $p$ encoded by a relation $X$.
-Then the machine accepts if, and only if, the $p$-th bit of $f(\enc(\A))$ is $1$.
-Since this procedure works in polynomial space, it can be described in $\pfp$ \cite{AbiteboulV89} by a formula $\Phi(X)$ where the free variable $X$ encodes the number $p$.
-Then, similar than the previous proof we define:
-$$
-\alpha := \sa{X}(\Phi(X)\mult\varphi(X)),
-$$ 
-where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
-Note that for each $B\subseteq A^{\ell}$ such that $B$ is the $m$-th element in the order $\varphi_{<}(X,Y)$, it holds that $\sem{\varphi(B)}(\A) = 2^m$.
-Therefore, $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-\medskip
-
-%%% DEMOSTRACION DE QFO(PFP) y FPSPACE(poly)
-\subsection*{Proof of Corollary~\ref{cor:capture-fpspace-poly}}
-
-For the first condition, let $\alpha\in\qfo(\pfp)$ over some signature $\R$.
-Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{1,\ldots,n\}$.
-In the procedure we slightly extend the grammar of $\qfo(\pfp)$ to include constants.
-We replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\sa{x} \beta(x)$ is replaced by $(\beta(1)+\cdots+\beta(n))$ and $\pa{x}\beta(x)$ is replaced by $(\beta(1)\cdot\,\cdots\,\cdot\beta(n))$.
-Then each sub-formula $\varphi\in\pfp$ in $\alpha$ is evaluated in polynomial space and replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size  (recall that $\alpha$ is fixed) which is evaluated and lastly given as output.
-Note that $f\in\nfpspace$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
-
-For the second condition, let $f\in \fpspace$ defined over some signature $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
-Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
-Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
-Let $m$ be the position of $\bar{a}$ in the lexicographic order of the tuples in $A^{\ell}$.
-The procedure then computes the $m$-th bit of $f(\enc(\A))$, from least to most significant.
-Since this procedure works in polynomial space, it can be described by an $\pfp$ formula $\Phi(\bar{x})$.
-Then we use
-$$
-\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot\varphi(\bar{x}),
-$$
-where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
-Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-%%% DEMOSTRACION DE optP y MaxQSO, MinQSO
-\subsection*{Proof of Theorem~\ref{theo:capture-optp}}
-
-Similar than the previous proofs, we construct recursively a non-deterministic polynomial time Turing machine $M_{\alpha}$ with output tape for each $\maxqso(\fo)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ (in binary) over the output tape of some run, and this value is the maximum value over all runs.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\fo$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ in deterministic polynomial time, and output $1$ if, and only if, $(\A,v,V)$ satisfies $\varphi$.
-If $\alpha$ is a constant $s$, it outputs $s$ in binary over the output tape.
-If $\alpha = (\beta \add \gamma)$, then it simulates $M_{\beta}$ and $M_{\gamma}$ on input $(\A,v,V)$, adds the output of both machines and prints this value over the output tape.
-If $\alpha = (\beta \mult \gamma)$ or $\alpha = \max\{\beta,\gamma\}$, then it does the same than the previous case but it multiplies or maximizes, respectively, the outputs of both machines instead of adding.
-If $\alpha = \sa{x}\beta$, it iterates over all elements $a\in A$ simulating and adding the output of $M_{\beta}$ on input $(\A,v[a/x],V)$.
-$M_{\alpha}$ finally outputs the aggregated value in the output tape.
-If $\alpha = \pa{x}\beta$ or $\alpha = \maxa{x}\beta$, then $M_{\alpha}$ does the same idea than the previous case with the difference that the output of $M_{\beta}$ on input $(\A,v[a/x],V)$ is multiplied or maximized, respectively.
-If $\alpha = \maxa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Furthermore, it is straightforward to prove that each of these steps are correct and can be computed with a non-deterministic polynomial time Turing machine with output tape.
-Let $\alpha$ be a formula in $\maxqso$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the maximum run (with respect to the output value) of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\maxp$ function over $\R$ and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-Finally, one can easily see that the same construction holds with $\minqso(\fo)$ by constructing a Turing machine that take the min over all runs instead of max.
-
-The proof for the other direction is similar than in \cite{kolaitis1994logical} extended with the ideas of Theorem~\ref{theo:capture-fp}.
-Let $f\in \maxp$ be a function defined over some signature $\R$ and
-$\ell\in\nat$ such that $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$ for each $\A\in\ostr[\R]$ of size $n$.
-For $U \subseteq A^{\ell}$, we can interpret the encoding of $U$ ($\enc(U)$) as the binary encoding of a number with $n^l$-bits. We denote this value by $\val(\enc(U))$.
-Then, given $\A\in\ostr[\R]$ and $U \subseteq A^{\ell}$, consider the problem of checking whether $f(\enc(\A)) \geq \val(\enc(U))$. 
-Clearly, this is an $\np$-problem and, by Fagin's theorem, there exists a formula of the form $\ex{\bar{X}} \Phi(\bar{X}, Y)$ with $\Phi(\bar{X}, Y)$ in $\fo$-logic and $\arity(Y) = \ell$ such that $f(\enc(\A)) \geq \val(\enc(U))$ if, and only if, $(\A,v,V) \models \ex{\bar{X}} \Phi(\bar{X}, Y)$ with $V(Y) = U$.
-Then we can describe the function $f$ by the following $\maxqso$ formula:
-$$
-\alpha = \maxa{\bar{X}} \maxa{Y} \ \Phi(\bar{X}, Y) \cdot \big( \sa{\bar{x}} Y(\bar{x}) \cdot \pa{\bar{y}}(\bar{x} < \bar{y} \mapsto 2) \big).
-$$
-Note that, in contrast with previous proofs, we use $\bar{x} < \bar{y}$ instead of $\bar{y} < \bar{x}$ because the most significant bit in $\enc(U)$ correspond to the smallest tuple in $U$.  
-It is easy to check that $\Phi(\bar{X}, Y)$ simulates the NP-machine and, if $\Phi(\bar{X}, Y)$ holds, the formula to the right  reconstructs the binary output from the relation in $Y$.
-Then, $\alpha$ is in $\maxqso(\fo)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-For the case of $\minqso(\fo)$ and a function $f \in \minp$, one has to follow the same approach but considering the $\np$-problem of checking whether $f(\enc(\A)) \leq \val(\enc(U))$.
-Then, the formula for describing $f$ is the following:
-$$
-\alpha = \mina{\bar{X}} \mina{Y} \ \sa{\bar{x}} \big( (\Phi(\bar{X}, Y) \rightarrow Y(\bar{x})) \cdot \pa{\bar{y}}(\bar{x} < \bar{y} \mapsto 2)  \big).
-$$
-In this case, if the formula $\Phi(\bar{X}, Y)$ is false, then the output produced by the subformula inside the $\min$-quantifiers will be the biggest possible value (i.e. $2^{n^\ell}$).
-On the other hand, if $\Phi(\bar{X}, Y)$ holds, the subformula will produce $\val(\enc(U))$. 
-Similar than for $\max$, we conclude that $\alpha$ is in $\minqso(\fo)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
+Since a function in $\sfo$ can also be defined $\eqso(\fo)$ (see Section~\ref{sec:previous}), the condition holds.

### 2017-03-03 LMCS/proof-capture-spanP.tex
journal primera version

--- 2017-03-03 LMCS/appendix-counting.tex
@@
-\medskip
-
-%%% DEMOSTRACION DE SQSO(FO) y #P
-\subsection*{Proof of Proposition~\ref{prop:capture-shP}}
-
-We construct recursively a $\shp$-machine $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\fo$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ deterministically in polynomial time, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-%If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-%If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$. 
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Let $\alpha$ be a formula in $\eqso(\fo)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shp$-function over $\R$ and $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the other direction, note that Saluja et al.~\cite{SalujaST95} proved that $\shp = \sfo$. 
-%We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso(\fo)$ be its corresponding formula. 
-Since a function in $\sfo$ can also be defined $\eqso(\fo)$ (see Section~\ref{sec:previous}), the condition holds. \qed
-
-\medskip
-
-%%% DEMOSTRACION DE SQSO(ESO) y span-P
-\subsection*{Proof of Proposition~\ref{prop:capture-spanP}}
-
 Similar than the previous proof, we construct recursively a $\spp$ machine $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
@@
 Since a function in $\#\eso$ can also be defined in $\eqso(\eso)$, then $\eqso(\eso)$ captures $\spp$ over ordered structures.
-
-\medskip
-
-%%% DEMOSTRACION DE QFO(LFP) y FP
-\subsection*{Proof of Theorem~\ref{theo:capture-fp}}
-
-For the first condition, let $\alpha\in\qfo(\lfp)$ over some signature $\R$.
-Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{1,\ldots,n\}$.
-In the procedure we slightly extend the grammar of $\qfo(\lfp)$ to include constants.
-We replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\sa{x} \beta(x)$ is replaced by $(\beta(1)+\cdots+\beta(n))$ and $\pa{x}\beta(x)$ is replaced by $(\beta(1)\cdot\,\cdots\,\cdot\beta(n))$.
-Then each sub-formula $\varphi\in\lfp$ in $\alpha$ is evaluated in polynomial time and replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size (recall that $\alpha$ is fixed) which is evaluated and lastly given as output.
-Note that $f\in\fp$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
-
-For the second condition, let $f\in \fp$ defined over some signature $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$ (i.e. $n^\ell$ is an upper bound for the output size), where $\A$ has a domain of size $n$.
-Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
-Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
-Let $m$ be the position of $\bar{a}$ in the lexicographic order of the tuples in $A^{\ell}$.
-The procedure then computes the $m$-th bit of $f(\enc(\A))$, from least to most significant.
-Since this procedure works in polynomial time, it can be described by an $\lfp$ formula $\Phi(\bar{x})$.
-Then we use
-$$
-\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot\varphi(\bar{x}),
-$$
-where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
-Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-In other words, $\Phi(\bar{x})$ simulates the behavior of the $\fp$-machine and the formula $\alpha$ reconstruct the binary output.
-Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-\medskip
-
-%%% DEMOSTRACION DE QSO(PFP) y FPSPACE
-\subsection*{Proof of Theorem~\ref{theo:capture-fpspace}}
-
-To show how to evaluate a $\qso(\pfp)$-formula, we construct recursively a $\shpspace$-machine $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
-This machine runs in non-deterministic polynomial space and, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\pfp$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space~\cite{L04}, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$, continuing to the next $A$-value if the run of $M_{\beta}$ accepts.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-If $\alpha = \pa{X}\beta$, it simulates $M_{\beta}$ on input $(\A,v,V[B/X])$ consecutively for each $B\in A^{arity(X)}$, again continuing to the next $A$-value if the run of $M_{\beta}$ accepts.
-This covers all possible cases for $\alpha$, and each of these steps can be computed in polynomial space.
-Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$, by the fact that $\shpspace = \fpspace$ \cite{Ladner89}, and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the second condition, let $f\in \fpspace$ defined over some $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq 2^{n^\ell}$ (i.e.
-$2^{n^\ell}$ is an upper bound for the output size), where $\A$ has a domain of size $n$.
-Let $X$ be a second-order variable of arity $\ell$.
-Consider a linear order over predicates of arity $\ell$ given by the formula
-$$
-\varphi_{<}(X,Y) = \exists\bar{u}\big[\neg X(\bar{u})\wedge Y(\bar{u})\wedge \forall\bar{v}\big(
-\bar{u}<\bar{v}\to(X(\bar{u})\iff Y(\bar{v}))\big)\big].
-$$
-Namely, we use relations to encode numbers with at most $2^{n^\ell}$-bits where the empty relation represents $0$ and the total-relation represents $2^{2^{n^\ell}}-1$.
-Furthermore, each relation $X$ indexes a position in the binary output of $f(\enc(\A))$ as follows.
-Consider a polynomial space machine over the $\R$ that receives as input an $\R$-structure $\A$ and a number $p$ encoded by a relation $X$.
-Then the machine accepts if, and only if, the $p$-th bit of $f(\enc(\A))$ is $1$.
-Since this procedure works in polynomial space, it can be described in $\pfp$ \cite{AbiteboulV89} by a formula $\Phi(X)$ where the free variable $X$ encodes the number $p$.
-Then, similar than the previous proof we define:
-$$
-\alpha := \sa{X}(\Phi(X)\mult\varphi(X)),
-$$ 
-where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
-Note that for each $B\subseteq A^{\ell}$ such that $B$ is the $m$-th element in the order $\varphi_{<}(X,Y)$, it holds that $\sem{\varphi(B)}(\A) = 2^m$.
-Therefore, $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-\medskip
-
-%%% DEMOSTRACION DE QFO(PFP) y FPSPACE(poly)
-\subsection*{Proof of Corollary~\ref{cor:capture-fpspace-poly}}
-
-For the first condition, let $\alpha\in\qfo(\pfp)$ over some signature $\R$.
-Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{1,\ldots,n\}$.
-In the procedure we slightly extend the grammar of $\qfo(\pfp)$ to include constants.
-We replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\sa{x} \beta(x)$ is replaced by $(\beta(1)+\cdots+\beta(n))$ and $\pa{x}\beta(x)$ is replaced by $(\beta(1)\cdot\,\cdots\,\cdot\beta(n))$.
-Then each sub-formula $\varphi\in\pfp$ in $\alpha$ is evaluated in polynomial space and replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size  (recall that $\alpha$ is fixed) which is evaluated and lastly given as output.
-Note that $f\in\nfpspace$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
-
-For the second condition, let $f\in \fpspace$ defined over some signature $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
-Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
-Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
-Let $m$ be the position of $\bar{a}$ in the lexicographic order of the tuples in $A^{\ell}$.
-The procedure then computes the $m$-th bit of $f(\enc(\A))$, from least to most significant.
-Since this procedure works in polynomial space, it can be described by an $\pfp$ formula $\Phi(\bar{x})$.
-Then we use
-$$
-\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot\varphi(\bar{x}),
-$$
-where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
-Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-%%% DEMOSTRACION DE optP y MaxQSO, MinQSO
-\subsection*{Proof of Theorem~\ref{theo:capture-optp}}
-
-Similar than the previous proofs, we construct recursively a non-deterministic polynomial time Turing machine $M_{\alpha}$ with output tape for each $\maxqso(\fo)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ (in binary) over the output tape of some run, and this value is the maximum value over all runs.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\fo$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ in deterministic polynomial time, and output $1$ if, and only if, $(\A,v,V)$ satisfies $\varphi$.
-If $\alpha$ is a constant $s$, it outputs $s$ in binary over the output tape.
-If $\alpha = (\beta \add \gamma)$, then it simulates $M_{\beta}$ and $M_{\gamma}$ on input $(\A,v,V)$, adds the output of both machines and prints this value over the output tape.
-If $\alpha = (\beta \mult \gamma)$ or $\alpha = \max\{\beta,\gamma\}$, then it does the same than the previous case but it multiplies or maximizes, respectively, the outputs of both machines instead of adding.
-If $\alpha = \sa{x}\beta$, it iterates over all elements $a\in A$ simulating and adding the output of $M_{\beta}$ on input $(\A,v[a/x],V)$.
-$M_{\alpha}$ finally outputs the aggregated value in the output tape.
-If $\alpha = \pa{x}\beta$ or $\alpha = \maxa{x}\beta$, then $M_{\alpha}$ does the same idea than the previous case with the difference that the output of $M_{\beta}$ on input $(\A,v[a/x],V)$ is multiplied or maximized, respectively.
-If $\alpha = \maxa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Furthermore, it is straightforward to prove that each of these steps are correct and can be computed with a non-deterministic polynomial time Turing machine with output tape.
-Let $\alpha$ be a formula in $\maxqso$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the maximum run (with respect to the output value) of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\maxp$ function over $\R$ and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-Finally, one can easily see that the same construction holds with $\minqso(\fo)$ by constructing a Turing machine that take the min over all runs instead of max.
-
-The proof for the other direction is similar than in \cite{kolaitis1994logical} extended with the ideas of Theorem~\ref{theo:capture-fp}.
-Let $f\in \maxp$ be a function defined over some signature $\R$ and
-$\ell\in\nat$ such that $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$ for each $\A\in\ostr[\R]$ of size $n$.
-For $U \subseteq A^{\ell}$, we can interpret the encoding of $U$ ($\enc(U)$) as the binary encoding of a number with $n^l$-bits. We denote this value by $\val(\enc(U))$.
-Then, given $\A\in\ostr[\R]$ and $U \subseteq A^{\ell}$, consider the problem of checking whether $f(\enc(\A)) \geq \val(\enc(U))$. 
-Clearly, this is an $\np$-problem and, by Fagin's theorem, there exists a formula of the form $\ex{\bar{X}} \Phi(\bar{X}, Y)$ with $\Phi(\bar{X}, Y)$ in $\fo$-logic and $\arity(Y) = \ell$ such that $f(\enc(\A)) \geq \val(\enc(U))$ if, and only if, $(\A,v,V) \models \ex{\bar{X}} \Phi(\bar{X}, Y)$ with $V(Y) = U$.
-Then we can describe the function $f$ by the following $\maxqso$ formula:
-$$
-\alpha = \maxa{\bar{X}} \maxa{Y} \ \Phi(\bar{X}, Y) \cdot \big( \sa{\bar{x}} Y(\bar{x}) \cdot \pa{\bar{y}}(\bar{x} < \bar{y} \mapsto 2) \big).
-$$
-Note that, in contrast with previous proofs, we use $\bar{x} < \bar{y}$ instead of $\bar{y} < \bar{x}$ because the most significant bit in $\enc(U)$ correspond to the smallest tuple in $U$.  
-It is easy to check that $\Phi(\bar{X}, Y)$ simulates the NP-machine and, if $\Phi(\bar{X}, Y)$ holds, the formula to the right  reconstructs the binary output from the relation in $Y$.
-Then, $\alpha$ is in $\maxqso(\fo)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-For the case of $\minqso(\fo)$ and a function $f \in \minp$, one has to follow the same approach but considering the $\np$-problem of checking whether $f(\enc(\A)) \leq \val(\enc(U))$.
-Then, the formula for describing $f$ is the following:
-$$
-\alpha = \mina{\bar{X}} \mina{Y} \ \sa{\bar{x}} \big( (\Phi(\bar{X}, Y) \rightarrow Y(\bar{x})) \cdot \pa{\bar{y}}(\bar{x} < \bar{y} \mapsto 2)  \big).
-$$
-In this case, if the formula $\Phi(\bar{X}, Y)$ is false, then the output produced by the subformula inside the $\min$-quantifiers will be the biggest possible value (i.e. $2^{n^\ell}$).
-On the other hand, if $\Phi(\bar{X}, Y)$ holds, the subformula will produce $\val(\enc(U))$. 
-Similar than for $\max$, we conclude that $\alpha$ is in $\minqso(\fo)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.

### 2017-03-03 LMCS/proof-ehorn-pe.tex
journal primera version

--- 2017-01-12 LICS/appendix-syntactic.tex
@@
-\subsection*{Proof of Theorem \ref{theo-pnf-snf}}
-Recall that a formula in $\eqso(\LL)$ is on the following grammar:
-$$
-\alpha = \varphi \ \mid \ s \ \mid \ (\alpha + \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \sa{X} \alpha,
-$$
-where $\varphi$ is a formula in $\LL$ and $s\in\nat$.
-We will construct a recursive function $\tau$ such that for every $\eqso(\LL)$ formula $\alpha$, it outputs an equivalent formula $\tau(\alpha)$ which is in $\LL$-SNF.
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$ which is clearly equivalent and already in $\LL$-SNF.
-If $\alpha = s$, then we define $(\top \add \cdots \add \top)$ ($s$ times), which also satisfies the condition.
-We assume that for every sub-formula $\beta$ of $\alpha$, $\tau(\beta)$ is an equivalent formula in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$, then we define $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$.
-If $\alpha = \sa{x}\beta$, then let $\tau(\beta) = \sum_{i = 1}^{k}\beta_i$ for some $k$ where each $\beta_i$ is in $\LL$-PNF.
-By grouping together the terms in the sum, we define $\tau(\alpha) = \sum_{i = 1}^{m}\sa{x}\beta_i$ which satisfies the condition and is equivalent to $\alpha$.
-If $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case.
-This covers all possible cases for $\alpha$ and we conclude the proof by taking $\tau(\alpha)$ as the desired rewrite of $\alpha$.
-
-\subsection*{Proof of theorem \ref{theo-pi1-pnf}}
-We divide the proof in three parts.
-
-\vspace{1em}
-First, we prove that the formula $\alpha_{0} = \left( \sa{X} 1 \right) + 1$ with $\arity(X) = 1$ (i.e.
-the function $2^{n}+1$, where $n$ is the size of the input structure) is not equivalent to any formula in $\loge{0}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ in $\loge{0}$-PNF that is equivalent to the $\eqso(\loge{0})$ formula $\alpha_0$.
-In \cite{SalujaST95}, it was proved that if $\length(\bar{X}) > 0$, then the function defined by $\alpha$, for big enough structures, is always even which is not possible in our case.
-On the other hand, if $\alpha$ is of the form $\sa{\bar{x}}\varphi(\bar{x})$, then $\alpha$ defines a polynomially bounded function which also leads to a contradiction.
-
-\vspace{1em}
-Second, we prove that the formula $\alpha_{1} = 2$ (i.e.
-the constant $2$) is not equivalent to any formula in $\loge{1}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to the $\eqso(\loge{1})$ formula $2$.
-First, if $\length(\bar{X}) = \length(\bar{x}) = 0$, then the function defined by $\alpha$ is never greater than 1. 
-Therefore, suppose that $\length(\bar{X}) > 0$ or $\length(\bar{x}) > 0$, and consider any ordered structure $\A$. 
-Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
-Now consider the ordered structure $\A'$ that is obtained by duplicating $\A$.
-This is, each half of $\A'$ is isomorphic to $\A$.
-Note that $\A'\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$ for $i = 1,2$ and there exists a third assignments $(\bar{B}_1',\bar{b}_1',\bar{a}_1')$ that is isomorphic to $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$ but in the other half to the structure such that $\A'\models\varphi(\bar{B}_1',\bar{b}_1',\bar{a}_1')$.
-We have that $\sem{\alpha}(\A) \geq 3$ which is a contradiction.
-
-\vspace{1em}
-We now show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there is an equivalent formula $\beta$ in $\LL$-PNF.
-As in Theorem \ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such formula.
-As we showed, there exists an equivalent formula in $\LL$-SNF, so we assume that $\alpha$ is in that form.
-Let $\alpha = \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-SNF.
-Without lost of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.
-If not, we replace each $\alpha_i$ for the equivalent formula $\sa{\bar{X}} \sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge\forall z\,Y(z) \wedge \forall z(y \leq z))$.
-
-Now we begin describing the function $\tau$.
-If $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then the formula is already in $\LL$-PNF so we define $\tau(\alpha) = \alpha$.
-If $\alpha = \alpha_1 + \alpha_2$, then we assume that $\tau(\alpha_1) = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ and $\tau(\alpha_2) = \sa{\bar{Y}}\sa{\bar{y}}\psi(\bar{Y},\bar{y})$.
-The construction that we will provide for this function works by identifying a ``first'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ and a ``last'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$.
-These are identified by the following formulas:
-\begin{align*}
-\gamma_{\text{first}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z}), \\
-\gamma_{\text{last}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x}).
-\end{align*}
-Similarly, we define the formulas $\gamma_{\text{first}}(\bar{Y},\bar{y})$ and $\gamma_{\text{last}}(\bar{Y},\bar{y})$ (for the sake of simplicity we reuse the names $\gamma_{\text{first}}$ and $\gamma_{\text{last}}$).
-In other words, the ``first'' assignment is the one where every second-order predicate is empty and the first-order assignment is the lexicographically smallest, and the ``last'' assignment is the one where every second-order predicate is full and the first-order assignment is the lexicographically greatest.
-We also need to identify assignments that are not first and are not last.
-We do this by negating the two formulas above and grouping together the first-order variables:
-\begin{align*}
-\gamma_{\text{not first}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{z}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{z}_i)), \\
-\gamma_{\text{not last}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{x} < \bar{z}_0 \vee \bigvee_{i = 1}^{\length(\bar{X})}\neg X(\bar{z}_i)),
-\end{align*}
-where $\bar{z} = (\bar{z}_0,\bar{z}_1,\ldots,\bar{z}_{\length(\bar{X})})$. The following formula is equivalent to $\alpha$:
-\begin{align}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{not first}}(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{Y},\bar{y}))\vee \label{eq:partition1} \\
-&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))\vee \label{eq:partition2}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{not last}}(\bar{Y},\bar{y}))\vee \label{eq:partition3}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{last}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))]. \label{eq:partition4}
-\end{align}
-To show that the formula is indeed equivalent to $\alpha$, note that the formulas in lines (\ref{eq:partition1}) and (\ref{eq:partition2}) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$, and the formulas in lines (\ref{eq:partition3}) and (\ref{eq:partition4}) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
-Altogether, the four lines define pairwise disjoint assignments for $(\bar{X},\bar{x}),(\bar{Y},\bar{y})$.
-With this, it is straightforward to show that the above formula is equivalent to $\alpha$.
-However, the formula is not yet in the correct form since it has existential quantifiers in the subformulas $\gamma_{\text{not first}}$ and $\gamma_{\text{not last}}$.
-To solve this, first let take a close look to the complete formula:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\exists\bar{v}(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i))\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\exists\bar{w}(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))].
-\end{align*}
-To construct an equivalent formula that is in the correct form, we define $\bar{u} = (\bar{v},\bar{w})$ and we replace the first-order quantifiers by a first-sum and count the first assignment to $\bar{v}$ and $\bar{w}$ that satisfies the formula.
-A similar construction was used in \cite{SalujaST95}.
-Then the final formula equivalent to $\alpha$ is the following:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
-& \hspace{3cm}  \bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&
-(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))\vee
-\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
-& \hspace{3cm} (\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee \\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))].
-\end{align*}
-Finally, consider a $\eqso(\LL)$ formula $\alpha$ in $\LL$-SNF.
-If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we can consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the condition in the hypothesis.
-
-\subsection*{Proof of Theorem \ref{prop-rest}}
-We give this proof in three parts.
-
-\vspace{1em}
-First, we show that $\QE{0} \not\subseteq \E{1}$.
-By contradiction, suppose that there is a $\QE{0}$ formula $\alpha$ over some signature $\R$ such that defines the following function.
-For every finite $\R$-structure with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n - 1$.
-We use the following claim.
-\begin{claim}
-Let $\alpha = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. Then the function defined by $\alpha$ is either null, greater or equal to $n$, or is in $\Omega(n^2)$.
-\end{claim}
-\begin{proof}
-Suppose that the function defined by $\alpha$ is not $0$ and that $\varphi$ is in DNF.
-Furthermore, suppose $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$.
-Then $\alpha = \sa{\bar{x}} \varphi_1(\bar{x}) \vee \cdots \vee \varphi_n(\bar{x})$.
-Since $\alpha$ is not null, then some $\varphi_i$ must be satisfiable.
-This is, the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is not null.
-We will prove by induction on $\length(\bar{x})$ that the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is either greater or equal to $n$, or in $\Omega(n^2)$.
-We address the case $\length(\bar{x})= 1$, then $\alpha = \sa{x}\bigwedge\psi(x)$.
-If any $\psi(x) = (x = x)$ or $\neg(x < x)$, then we can eliminate it and we obtain the same function.
-If any $\psi = (x < x)$ or $\neg(x=x)$, then the function becomes null.
-If $\psi(x) = R(x,\ldots,x)$ for some $R\in\R$ the function becomes null for the structures we are considering.
-If $\psi(x) = \neg R(x,\ldots,x)$, we can eliminate it and for the structures we are considering we obtain the same function.
-The only possible $\alpha$ left is $\alpha = \sa{x}\top$ which is equal to the function $n$.
-This covers all possible cases for $\length(\bar{x}) = 1$.
-Now suppose that it holds for $\length(\bar{x}) = k$ and suppose $\alpha = \sa{\bar{x}}\bigwedge\psi(\bar{x})$ for $\length(\bar{x}) = k+1$.
-If any $\psi(\bar{x}) = (x_i = x_j)$ where $i \neq j$, then $\alpha$ describes the same function as $\alpha$ where $x_j$ has been replaced by $x_i$.
-In this formula the tuple of first-order variables has $k$ elements so the function it describes if one of the mentioned in the hypothesis.
-If $i = j$, then we can eliminate it and obtain the same function.
-If any $\psi(\bar{x}) = R(\bar{v})$ or $\neg R(\bar{v})$ where $\bar{v}$ is a sub-tuple of $\bar{x}$ then we can either eliminate it or the function becomes null, following the same argument as in the case $\length(\bar{x}) = 1$.
-If any $\psi(\bar{x}) = \neg(x_i = x_j)$ or $(x_i < x_j)$ where $i = j$, then the function becomes null.
-If any $\psi(\bar{x}) = \neg(x_i < x_j)$ where $i = j$, we can eliminate it.
-The remaining formulas in $\bigwedge\psi(\bar{x})$ are either $\neg(x_i = x_j)$, $(x_i<x_j)$ or $\neg(x_i<x_j)$.
-If the formula violates transitivity in $<$ (for example, $x < y \wedge y < z \wedge z < x$), then the function $\alpha$ describes is null.
-Therefore, there is some order over $\bar{x}$ that satisfies $\bigwedge\psi(\bar{x})$.
-Consider the formula that describes this order (like $x_1 < x_3 \wedge x_3 < x_4 \wedge x_4 < x_2$).
-The function $\alpha$ describes is greater or equal to the one this formula describes, which is exactly $\binom{n}{\length(\bar{x})}$ which is in $\Omega(n^{\length(\bar{x})}) \subseteq \Omega(n^2)$ if $\length(\bar{x}) > 1$.
-This concludes the proof of the claim.
-\end{proof}
-We suppose that $\alpha$ is in SNF, this is, $\alpha = \sum_{i = 1}^n\alpha_i$.
-Since $\alpha$ is not null, consider some $\alpha_i$ that describes a non-null function.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, where $\varphi$ is quantifier-free.
-Note that if $\length(\bar{X}) > 0$, then the function $\alpha$ describes is in $\Omega(2^n)$, as it was proven by the authors in \cite{SalujaST95}.
-We have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$, as we proved in the claim, describes either some function greater or equal to $n$, or in $\Omega(n^2)$, which leads to a contradiction.
-Lastly, note that the formula $\sa{x}\exists y(x < y)$ is in $\E{0}$ and describes the function $n-1$, which concludes the proof.
-
-\vspace{1em}
-Now we show that $\E{1}\not\subseteq\QE{0}$.
-In Theorem \ref{theo-pi1-pnf} we proved that there is no formula in $\loge{1}$-PNF equivalent to the formula $\alpha = 2$.
-Every formula in $\E{1}$ can be expressed in $\loge{1}$-PNF, which implies that $2 \in \QE{0}$ and $2 \not\in \E{1}$.
-
-\vspace{1em}
-Lastly, we prove that $\eqso(\loge{1})\subsetneq\eqso(\logu{1})$.
-For inclusion, let $\alpha$ be a formula in $\eqso(\loge{1})$.
-Suppose that it is in $\loge{1}$-SNF.
-This is, $\alpha = c + \sum_{i = 1}^{n}\alpha_i$.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, where $\varphi_i$ is quantifier-free, for each $\alpha_i$.
-We use the same construction used in \cite{SalujaST95}, and we obtain that the formula $\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \forall\bar{y}'(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')$ for every assignment to $(\bar{X},\bar{x})$.
-We do this replacement for each $\alpha_i$ and we obtain an equivalent formula in $\eqso(\logu{1})$.
-
-To prove that the inclusion is proper, consider the $\eqso(\logu{1})$ formula $\sa{x}\forall y(y = x)$. This formula defines the following function that takes an ordered structure $\A$ as input:
-$$
-\sem{\alpha}(\A) = 
-\begin{cases}
-1 &\A \text{ has one element}\\
-0 &\text{ otherwise}.
-\end{cases}
-$$
-Suppose that there exists an equivalent formula $\alpha$ in $\eqso(\loge{1})$.
-Also, suppose that it is in $\L$-PNF, so $\alpha = c + \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$.
-Since $\alpha$ takes the value 0 for some structures, $c$ must be 0.
-Consider a structure $\mathfrak{1}$ with one element.
-We have that for some $i$, there exists an assignment $(\bar{B},\bar{b},\bar{a})$ for $(\bar{X},\bar{x},\bar{y})$ such that $\mathfrak{1}\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
-Consider now the structure $\mathfrak{2}$ that is obtained by duplicating $\mathfrak{1}$, as we did for Theorem \ref{theo-pi1-pnf}.
-Note that $\mathfrak{2}\models\varphi_i(\bar{B},\bar{b},\bar{a})$, which implies that $\sem{\alpha}(\mathfrak{2}) \geq 1$, which leads to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop-e1-nc}}
-
-Towards a contradiction, assume that the statement is false.
-This is, that $\E{1}$ is closed under binary sum.
-Consider the formula $\sa{x}(x = x)$ which is in $\E{1}$ over some signature $\R$.
-For every finite $\R$-structure $\A$ with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n$.
-From our assumption, there exists some formula in $\E{1}$ equivalent to the formula $\alpha \add \alpha$, which describes the function $2n$.
-Let $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ be this formula, where $\varphi$ is quantifier-free.
-Note that the function defined by this formula is equal or greater than the one defined by $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ divided by a polynomial factor.
-More specifically, for each ordered structure $\A$ with domain $A$, we have the following inequality:
-$$
-\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot \vert A \vert^{\length(\bar{y})} \geq \sem{\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A)
-$$
-Note that the formula $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ defines a function in $\E{0}$.
-It was shown by the authors in \cite{SalujaST95} that every function in $\E{0}$ grows exponentially over the size of the structure for large enough structures, when $\length(\bar{X}) > 0$.
-This function divided by a polynomial factor still grows exponentially.
-Therefore, for $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ we have that $\length(\bar{X}) = 0$.
-
-Now, for the formula $\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ consider a structure $\mathfrak{1}$ with only one element $a$.
-We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$, but the only possible assignment to $\bar{x}$ is the tuple $(a,\ldots,a)$ so $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which follows to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop:qe0-fp-qe1-totp-fptras}}
-
-The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$.
-This is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every finite $\R$-structure $\A$ with domain $A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$.
-We base our proof on this fact.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is contained in \textsc{TotP}.} Let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Consider the following nondeterministic procedure that on input $\enc(\A)$ generates $\sem{\alpha}(\A)$ branches.
-For each $\alpha_i = \varphi$, where $\varphi$ is a $\loge{1}$ formula, it checks if $\A\models\varphi$ in polynomial time and generates a new branch if that is the case.
-For each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi$, this formula is also in $\E{1}$.
-We use the reduction to $\cdnf$ provided in \cite{SalujaST95} and we obtain $g(\enc(\A))$, which is an instance to $\cdnf$.
-Since $\cdnf$ is also in $\totp$ \cite{PagourtzisZ06}, we simulate the corresponding nondeterministic procedure that generates exactly $\cdnf(\enc(g(\A)))$ branches.
-Since, $\fp\subseteq\totp$\cite{PagourtzisZ06}, each polynomially computable function is also in $\totp$, and then on each of these branches we simulate the corresponding nondeterministic procedure to generate $h(\vert A \vert)$ more.
-The number of branches for each $\alpha_i$ is $\sem{\alpha_i}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$, and the total number of branches in the procedure amounts to $\sem{\alpha}(\A)$.
-We conclude that $\alpha\in\totp$.
-
-\vspace{1em}
-{\bf Every function in} $\boldsymbol{\eqso(\loge{1})}$ {\bf has an FPRAS.}  let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Note that each $\alpha_i$ that is equal to some $\loge{1}$ formula $\varphi$ has an FPRAS given by the procedure that simply checks if $\A\models\varphi$ and returns 1 if it does and 0 otherwise.
-Also, each remaining $\alpha_i$ has an FPRAS since $\alpha_i\in \E{1}$ \cite{SalujaST95}.
-If two functions have an FPRAS, then their sum also has one given by the procedure that simulates them both and sums the results.
-We conclude that $\alpha$ has an FPRAS.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is closed under sum and multiplication.} Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that the class is closed under multiplication.
-We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of $\so$.
-
-\begin{lemma} \label{conj-mult}
-If $\LL$ is a fragment closed under conjunction, then $\eqso(\LL)$ is closed under binary multiplication.
-\end{lemma}
-\begin{proof}
-We will define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ extended by binary product, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
-In fact, the formula $\tau(\alpha)$ is in $\LL$-SNF.
-First we replace each constant $s$ in $\alpha$ for $(\top \add \cdots \add \top)$ ($s$ times).
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$.
-We assume that for every $\beta$ that has less algebraic operators than $\alpha$, $\tau(\beta)$ is in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$ then we define $\tau(\alpha) = \tau(\alpha_1) + \tau(\alpha_2)$.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$, then we define $\tau(\alpha)$ as the formula in $\LL$-SNF that is equivalent to $\sa{x}\tau(\beta)$ and to $\tau(\alpha) = \sa{X}\tau(\beta)$, respectively.
-If $\alpha = (\alpha_1 \cdot \alpha_2)$, we assume that each $\alpha_i$ is in $\LL$-SNF.
-We identify three cases.
-(1) Some $\alpha_i$ is equal to $\sum_{j = 1}^n\beta_j$ for $n > 1$.
-Suppose wlog.
-that it is $\alpha_1$.
-We then define $\tau(\alpha) = \sum_{j = 1}^n\tau(\beta_j\cdot\alpha_2)$.
-In the following cases, $\alpha_1$ and $\alpha_2$ are in $\LL$-SNF.
-(2) If some $\alpha_i$ is equal to $\sa{X}\beta$ or $\sa{x}\beta$, we define $\tau(\alpha)$ as the $\LL$-SNF formula that is equivalent to $\sa{x}\tau(\beta\cdot\alpha_2)$ and $\sa{X}\tau(\beta\cdot\alpha_2)$, respectively.
-The remaining case is (3) $\alpha_1 = \varphi_1$ and $\alpha_2 = \varphi_2$ where each $\varphi$ is an $\LL$ formula.
-Then we define $\tau(\alpha) = \varphi_1 \wedge \varphi_2$.
-This covers all possible cases for $\alpha$.
-For every pair of formulas $\alpha,\beta$ in $\eqso(\LL)$, we have that their multiplication $(\alpha\cdot\beta)$ is a formula in the grammar $\eqso(\LL)$ extended by binary product, and so, there exists an equivalent formula $\tau(\alpha\cdot\beta)$ which is in unextended $\eqso(\LL)$.
-\end{proof}
-Since $\loge{1}$ is closed under conjunction, this also holds for $\eqso(\loge{1})$. This concludes the proof.
-
-\subsection*{Proof of Proposition \ref{pi-minusone}}
-
-Let $\LL$ be a fragment of $\fo$ that contains $\logu{1}$.
-Then we have that every function in $\U{1}$ is expressible in $\eqso(\LL)$.
-In particular, $\ctcnf \in \eqso(\LL)$.
-Suppose that $\eqso(\LL)$ is closed under subtraction by one.
-Then, the function $\ctcnf-1$, which counts the number of satisfying assignments of a 3-CNF formula minus one, is also in $\eqso(\LL)$.
-Note that $\eqso(\LL) \subseteq \eqso(\fo) = \shp$.
-We have that $\ctcnf$ is $\shp$-complete under parsimonious reductions\footnote{It can be easily verified that the standard reduction from SAT to 3-CNF (or 3-SAT) preserves the number of satisfying assignments}.
-Now, let $f$ be a function in $\shp$, and consider the nondeterministic polynomial-time procedure that on input $\enc(\A)$ computes the corresponding reduction to $\ctcnf$, name it $g(\enc(\A))$, and simulates the $\shp$ procedure for $\ctcnf-1$ on input $g(\enc(\A))$.
-We have that this is a $\shp$ procedure that computes $f-1$, from which we conclude that $\shp$ is closed under subtraction by one.
-
-\subsection*{Proof of Theorem \ref{sigmafo-minusone}}
-
-\noindent {\bf Closed under sum and multiplication.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
-This can be seen because $\eqso(\LL)$ is closed under sum for every fragment $\LL$ by definition, and since $\logex{1}$ is closed under conjunction, from Lemma \ref{conj-mult} it follows that $\eqso(\logex{1})$ is closed under multiplication.
-
-\vspace{1em}
-\noindent {\bf Easy decision version and FRPRAS.}  We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
-We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Theorem \ref{prop:qe0-fp-qe1-totp-fptras}.
-First, we define a function that converts a formula $\alpha$ in $\eqso(\logex{1})$ over a signature $\R$ into a formula $\lambda(\alpha)$ in $\eqso(\loge{1})$ over a signature $\R_{\alpha}$.
-Afterwards, we define a function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an $\R_{\alpha}$-structure $g_{\alpha}(\A)$.
-
-Let $\alpha$ be in $\eqso(\logex{0})$.
-The signature $\R_{\alpha}$ is obtained by adding the symbol $R_{\psi}$, for every $\fo$ formula $\psi(\bar{z})$ in $\alpha$, to $\R$.
-Each symbol $R_{\psi}$ represents a predicate with arity $\length(\bar{z})$.
-Then, $\lambda(\alpha)$ is defined as $\alpha$ where each $\fo$ formula $\psi(\bar{z})$ has been replaced by $R_{\psi}(\bar{z})$.
-We now define the function $g_{\alpha}$ procedurally.
-Let $\A$ be a $\R$-structure with domain $A$.
-Let $\A'$ be an $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A}$ empty.
-For each $\fo$-formula $\psi(\z)$ with $\length(\bar{z})$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length(\bar{z})}$.
-If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$ (this can be done in $\ptime$).
-This concludes the construction of $\A'$.
-Note that the number of $\fo$ subformulas, arity and tuple size is fixed in $\alpha$, so computing this function takes polynomial time over the size of the structure.
-Moreover, the encoding of $\A'$ has polynomial size over the size of $\enc(\A)$.
-We define $g_{\alpha}(\A) = \A'$ and we have that for each $\R$-structure $\A$: $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$.
-Therefore, we have a parsimonious reduction from $\alpha$ to the $\eqso(\loge{1})$ formula $\lambda(\alpha)$.
-
-To show that $\alpha$ is in $\totp$, we can convert $\alpha$ and $\enc(\A)$ into $\lambda(\alpha)$ and $\enc(g_{\alpha}(\A))$, respectively, and run the procedure in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-Similarly, to show that $\alpha$ has an FPRAS, we do the same as before and simulates the FPRAS for $\lambda(\alpha)$ in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-These procedures also takes polynomial time and satisfies the condition.
-
-\vspace{1em}
-\noindent {\bf Closed under subtraction by one.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. 
-For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Without lost of generality, we assume that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
-Moreover, we assume that $\length(\bar{x}) > 0$  since, if this is not the case, we can replace $\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
-
-The proof will be separated in two parts.
-In the first part, we suppose that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$.
-Then we will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ is the formula.
-In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ the sum of one or more formulas in $\logex{1}$-PNF.
-We define a formula $\varphi'$ that satisfies the following condition: if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ and $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be proven.
-
-\vspace{1em}
-
-\noindent {\em Part (1).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ where $\varphi$ is an  $\fo$-formula.
-Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}} \varphi(\bar{x})$ (i.e.
-$\length(\bar{X}) = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}} [\varphi(\bar{x})\wedge\exists\bar{z}(\varphi(\bar{z})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
-Therefore, for the rest of the proof we can assume that $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.
-
- Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
-To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in into a DNF formula. 
-More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i$ is a conjunction of atomic formulas or negation of atomic formulas.
-Furthermore, we suppose that each conjunct $\varphi_i$ has the~form:
-$$
-\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
-\underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}} \wedge
-\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}}.
-$$
-Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ can appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
-
-Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ are mentioned in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
-We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
-\begin{itemize}
-	\item[(a)] In order to satisfy the first condition, consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
-We add $\length(\bar{w})$ new variables $z_1,\ldots,z_{\length(\bar{w})}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length(\bar{w})})$.
-We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$.
-The formula $\varphi_i$ is now defined as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \exists \bar{z} \; (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{-}(\bar{X},\bar{x},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y})).
-	$$
-	We repeat this process for each instance of a $X_j(\bar{w})$ in $\varphi_i$, and we obtain a formula where none of the $X_j$'s acts over any variable in $\bar{x}$.
-We add the new first-order variables to $\bar{y}$ and we redefine $\varphi_i$ as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
-	$$
-	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = \exists \bar{y} \;  X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := \exists \bar{y} \;  X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
-	\item[(b)] An ordered partition on a set $S$ is defined by an equivalence relation $\sim$ over $S$, and a linear order over $S/\!\sim$.
-For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$.
-A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
-On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
-	For a given $k$, let $\cB_k$ be the number of possible ordered partitions for a set of size $k$. For $1 \leq j \leq \cB_{\length(\bar{x},\bar{y})}$ 
-	let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
-
-	Thus, the formula $\varphi_i(\bar{X},\bar{x})$ is then redefined as:
-	$$
-	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}} \exists \bar{y} \; [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
-	$$
-	Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
-	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  to $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
-	\item[(c)] For showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
-$\varphi_i$ is inconsistent), we do the following.
-If there exists an instance of $X_j(\bar{z})$ in $\varphi^{+}_i$, an instance of $\neg X_j(\bar{w})$ in $\varphi^{-}_i$ and the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$, then the entire formula $\varphi_i$ is removed from $\varphi$.
-\end{itemize}
-It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$.
-From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c).
-Furthermore, it has the following form:
-$$
-\varphi_i(\bar{X},\bar{x}) \; = \; \exists{\bar{y}} \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
-$$
-where $\varphi^{+}$ and $\varphi^{-}_i$ do not depend on  $\bar{x}$.
-\begin{claim}\label{claim:minusone}
-For a given ordered structure $\A$, if there exists a first-order assignment $v$ for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists a second-order assignment $V$ for $\A$ such that  $(\A,V,v)\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-\end{claim}
-\begin{proof}
-Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
-Define $\bar{B} = (B_1,\ldots,B_{\length(\bar{X})})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
-Towards a contradiction, suppose that $(\A,V,v)\not\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-By the choice of $v$, and construction of $V$ it is clear that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})\wedge\varphi^{+}_i(\bar{X},\bar{y})$, so we necessarily have that $(\A,V,v)\not\models\varphi^{-}_i(\bar{X},\bar{y})$.
-Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,V, v)\not\models\neg X_t(\bar{w})$, namely, $v(\bar{w})\in B_t$. 
-However, by the construction of $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$.
-Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$.
-This violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
-\end{proof}
-
-The previous claim and proof motivates the following definitions.
-For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length(\bar{X})})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
-One can easily check that for every assignments $V, v$ such that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$, then it holds that $(\A,\bar{B}^v,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all the assignment of $\bar{X}$ when $v$ is fixed.
-Therefore, by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
-For this, the idea is to choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
-$$
-\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge\forall\bar{x}'\forall\bar{y}'(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')).
-$$ 
-If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
-Furthermore, let $V^*$ the second order assignment that satisfies $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*$ the first order assignment such that $v^*(\bar{x}) = v(\bar{x})$.
-By the previous discussion, we have that $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.
-
-Now, we have all the tools to introduce the main formula of the proof. 
-Intuitively, we want to exclude the assignment $(V^*, v^*)$ as a satisfying assignment for $\varphi_i(\bar{X},\bar{x})$.
-Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if, 
-if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, then $V \neq V^*$ or $v \neq v^*$. 
-That is, it excludes all pairs $(V, v)$ that coincide with $(V^*, v^*)$ in $(\bar{X}, \bar{x})$ whenever $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. This property can be defined as follows:
-\begin{align}
-\psi_i(\bar{X},\bar{x}) = & \;\; \big( \, \exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
-& \hspace{1cm} \Big( \; 
-\exists\bar{y} \;\; \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i^+(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\
-& \hspace{1.5cm} (\exists\bar{x}'\exists\bar{y}\, \varphi_i(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}) \;\; \Big)  \label{eq:part3}
-\end{align}
-To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
-Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-Then, the conclusion of the implication at (\ref{eq:part2}) and (\ref{eq:part3}), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
-Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by showing that if $\varphi_i(\bar{X},\bar{x})$ is satisfied then $V(\bar{X})$ must contain one element more that it is not in $V^*(\bar{X})$.
-The second disjunct (\ref{eq:part3}) is satisfied when $v(\bar{x})$ is not the lexicographically smallest tuple that satisfies $\varphi_i$ (i.e. $v(\bar{x}) \neq v^*(\bar{x})$).
-
-We are ready to define the formula  $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$
-where each modified disjunct $\varphi_i'(\bar{X},\bar{x})$ is constructed as follows. 
-For the sake of simplification, define the auxiliary formula $\chi_i = \neg\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$. 
-This formula basically checks if $\varphi_i$ is satisfiable or not.
-Then, first define the formula:
-$$
-\varphi'_1(\bar{X},\bar{x}) = \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
-$$
-This models that all the assignments that satisfy $\varphi_1$ also satisfy $\varphi'_1$, except for $(V^*,v^*)$. Then we define:
-$$
-\varphi'_2(\bar{X},\bar{x}) = \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
-$$
-This models that all the assignments that satisfy $\varphi_2$ also satisfy $\varphi'_2$, but if there was no assignment for $\varphi_1$, then we exclude $(V^*,v^*)$ for $\varphi_2$.
-Then we generalize this construction for each $\varphi_i$:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})),
-$$
-By the construction of $\kappa(\alpha)$, one can easily check that  $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.
-
-\vspace{1em}
-
-{\em Part (2).} Let $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ for some $\eqso(\logex{1})$ formula $\beta$. We define $\kappa(\alpha)$ as follows.
-First, perform the same transformations to $\varphi(\bar{X},\bar{x})$ as in the previous section.
-Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c).
-We also use the previously defined formulas $\chi_i$ and $\psi_i$, for each $i$.
-
-We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if and only if $\sem{\beta}(\A) = 0$.
-If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg\exists\bar{x}'\varphi(\bar{x}')$.
-If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$ as each $\chi_i$ was previously defined.
-If $\beta = (\beta_1 + \beta_2)$, then $\lambda(\beta) = \lambda(\beta_1) \wedge \lambda(\beta_2)$.
-Now, for each $\varphi_i$ we define:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x}))\Big)\Big).
-$$
-Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
-
-%\subsection*{Proof of Theorem \ref{sub-pnp}}
-%
-%Towards a contradiction, suppose that any of the classes $\E{1}$, $\eqso(\loge{1})$, and $\eqso(\logex{1})$ is closed under subtraction.
-%
-%Let $\R$ include the symbols $S_1, S_2, S_3, S_4$, which describe the following properties.
-If a finite $\R$-structure $\A$ defines a 3DNF formula $\Phi$, then its domain is the set of variables mentioned in $\Phi$, and for each $i = 1,2,3,4$:
-%\begin{align*}
-%	S_1^\A &= \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_2^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_3^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_4^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
-%\end{align*}
-%Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
-%\begin{multline*}
-%\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
-%\end{multline*}
-%Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
-%$$
-%\varphi(x,X) = (T(x) \vee \neg T(x)).
-%$$
-%Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Given that $f_{\#3DNF}, f_{all} \in \E{1} \subseteq \eqso(\loge{1}) \subseteq \eqso(\logex{1}) \subseteq \totp$, and at least one of the classes is closed under subtraction, the function $f_{\#3DNF} - f_{all}$ is in $\totp$.
-However, note that for each $\R$-structure $\A$ that represents a formula $\Phi$, $(f_{\#3DNF} - f_{all})(\A) = 0$ if an only if $\Phi$ is a tautology.
-The decision version of this function is the $\np$-complete problem $\textsc{Tautology}$.
-And since the function is in $\totp$, its decision version is also in $\ptime$.
-We conclude that $\np \subseteq \ptime$.
-
-\subsection*{Proof of Proposition \ref{prop:ehorn-pe}}
 Pagourtzis and Zachos mention a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula \cite{PagourtzisZ06}.
@@
 We conclude that $\alpha$ is in $\totp$.
-
-\subsection*{Proof of Proposition \ref{prop:hsat-not-sigma2}} %V.12
-We use a similar proof to the one provided by the authors in \cite{SalujaST95} to separate the classes $\E{2}$ and $\U{2}$.
-Suppose that the statement is false, this is, $\chsat \in \eqso(\loge{2})$.
-We consider the signature $\R$ that we used as the encoding for a Horn formula (Example \ref{ex-hornsat-esop1}) and that the formula $\alpha \in \eqso(\loge{2})$ follows the encoding in the same way.
-From what we proved in Theorem \ref{theo-pnf-snf}, we have that every formula in $\eqso(\loge{2})$ can be rewritten in $\loge{2}$-PNF, so we assume that $\alpha$ is in this form.
-Let $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\forall\bar{z}\,\varphi(\bar{X},\bar{x},\bar{y},\bar{z})$.
-Consider the following Horn formula $\Phi$:
-$$
-\Phi = p \wedge \bigwedge_{i = 1}^n (t_i \wedge p \to q) \wedge \neg q,
-$$
-where $n = \length(\bar{x}) + \length(\bar{y}) + 1$.
-Let $\A$ be the encoding of this formula.
-In the encoding, each variable appears as an element in the domain of $\A$.
-This formula is satisfiable, so $\sem{\alpha}(\A) \geq 1$.
-Let $(\bar{B},\bar{b},\bar{a})$ be an assignment to $(\bar{X},\bar{x},\bar{y})$ such that $\A\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-Let $t_{\ell}$ be such that it does not appear in $\bar{b}$ or $\bar{a}$.
-Consider the induced substructure $\A'$ that is obtained by removing $t_{\ell}$ from $\A$ and $\bar{B}'$ as the subset of $\bar{B}$ obtained by deleting each appearance of $t_{\ell}$ in $\bar{B}$.
-We have that $\A'\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-This is because each subformula of the form $\exists y \neg B_i$ is still true, and universal formulas are monotone over induced substructures.
-It follows that $\sem{\alpha}(\A') \geq 1$ which is not possible since $\A'$ encodes the formula
-$$
-\Phi' = p \wedge \bigwedge_{i = 1}^{\ell-1} (t_i \wedge p \to q) \wedge (p\to q) \wedge \bigwedge_{i = \ell+1}^{n} (t_i \wedge p \to q) \wedge \neg q,
-$$
-which is unsatisfiable. We arrive to a contradiction and we conclude that $\chsat$ is not in $\eqso(\ehorn)$.
-
-\subsection*{Proof of Theorem \ref{sigma2hard}} %V.13
-
-First we prove that $\shdhsat$ is in $\eqso(\ehorn)$.
-Recall that each instance of $\shdhsat$ is a disjunction of Horn formulas.
-Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot), \pD(\cdot,\cdot)\}$.
-Each symbol in this vocabulary is used to indicate the same as in Example \ref{ex-hornsat-esop1}, with the addition of $\pD(d,c)$ which indicates that $c$ is a clause in the formula $d$.
-Recall that the formula
-\begin{align*}
-&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \neg\textit{T}(x)).
-\end{align*}
-defines $\chsat$. We obtain the following formula $\psi(T,A)$ in $\ehorn$:
-\begin{align*}
-\exists d[&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \pD(c,d)\vee \neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \neg\textit{T}(x))]
-\end{align*}
-effectively defines $\chsat$ as for every disjunction of Horn formulas $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
-Therefore, we conclude that $\shdhsat \in \eqso(\ehorn)$.
-
-\vspace{1em}
-We will now prove that $\shdhsat$ is hard for $\eqso$ over a signature $\R$ under parsimonious reductions.
-For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a polynomial-time procedure that computes a function $g_{\alpha}$.
-This function receives a finite $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
-We suppose that $\alpha$ is in sum normal form and:
-$$
-\alpha = \sum_{i = 1}^{\text{\#clauses}} \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigwedge_{j = 1}^{n}\forall\bar{z}\,\varphi^i_j(\bar{X},\bar{x},\bar{y},\bar{z}),
-$$
-where each $\varphi^i_j$ is a Horn clause.
-
-Consider a finite $\R$-structure $\A$ with domain $A$.
-To simplify the proof, we extend our grammar to allow first-order constants.
-Consider each tuple $\bar{a}\in A^{\length(\bar{x})}$, each $\bar{b}\in A^{\length(\bar{y})}$ and each $\bar{c}\in A^{\length(\bar{z})}$ as a tuple of first-order constants.
-The following formula defines the same function as $\alpha$:
-$$
-\sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length(\bar{x})}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length(\bar{y})}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length(\bar{z})}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
-$$
-Note that each $\fo$ formula over $(\bar{x},\bar{y},\bar{z})$ in each $\varphi^i_j$ has no free variables.
-Therefore, we can evaluate each of these in polynomial time and replace them by $\perp$ and $\top$ where it corresponds.
-Each $\varphi^i_j$ will be of the form $\perp \vee\, \chi^i_j(\bar{X})$ or $\top \vee \chi^i_j(\bar{X})$ where $\chi^i_j$ is a disjunction of $\neg X_{\ell}$'s and at most one $X_{\ell}$.
-The formulas of the form $\top \vee \chi^i_j(\bar{X})$ can be removed entirely, and the formulas of the form $\perp \vee\, \chi^i_j(\bar{X})$ can be replaced by $\chi^i_j(\bar{X})$.
-We obtain the formula
-$$
-\sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})
-$$
-where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$.
-
-Our idea for the rest of the proof is to define $g_{\alpha}$ for each $\alpha = \sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})$, for $\alpha = c$ and for $\alpha = \beta_1 + \cdots + \beta_m$ where each $\beta_i$ is in one of the two previous cases.
-
-If $\alpha$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ where $\psi_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$, then we obtain the {\bf propositional formula} $g_{\alpha}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulas.
-
-If $\alpha$ is equal to a constant $c$, then we define $g_{\alpha}(\A)$ as the following formula that has exactly $c$ satisfying assignments:
-$$
-g_{\alpha}(\A) = \bigvee_{i = 1}^{c}\neg t_1 \wedge \cdots \wedge \neg t_{i-1} \wedge t_i \wedge \neg t_{i+1} \wedge \cdots \wedge \neg p_c.
-$$ 
-If $\alpha = \beta_1 + \cdots + \beta_m$, let $g_{\beta_i}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\theta^i_{j,k}$ for each $\beta_i$ where each $\theta^i_{j,k}$ is a Horn clause.
-Let $\Theta_i = g_{\beta_i}(\A)$.
-We rename the variables in each $\Theta_i$ so none of them are mentioned in any other $\Theta_j$.
-We add $m$ new variables $t_1,\ldots,t_m$ and we define:
-\begin{align*}
-g_{\alpha}(\A) = &\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^1_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m}}}t) \wedge (t_1 \wedge \bigwedge_{\ell = 2}^{m} \neg t_{\ell})) \vee \\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^2_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each $t$}\\ \text{ mentioned in}\\ \Theta_1,\Theta_3,\ldots,\Theta_{m}}}t) \wedge (t_2 \wedge \bigwedge\limits_{\substack{\ell = 1 \\ \ell \neq 2}}^{m} \neg t_{\ell})) \vee \cdots \vee\\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^m_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m-1}}}t) \wedge (t_m \wedge \bigwedge_{\ell = 1}^{m-1} \neg t_{\ell})).
-\end{align*}
-The formula is a disjunction of Horn formulas, and the number of satisfying assignments for this formula is exactly the sum of satisfying assignments for each $g_{\beta_i}(\A)$.
-This, at the same time, is equal to $\sem{\alpha}(\A)$.
-This covers all possible cases for $\alpha$, and the entire procedure takes polynomial time.

### 2017-03-03 LMCS/proof-hsat-not-sigma2.tex
journal primera version

--- 2017-01-12 LICS/appendix-syntactic.tex
@@
-\subsection*{Proof of Theorem \ref{theo-pnf-snf}}
-Recall that a formula in $\eqso(\LL)$ is on the following grammar:
-$$
-\alpha = \varphi \ \mid \ s \ \mid \ (\alpha + \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \sa{X} \alpha,
-$$
-where $\varphi$ is a formula in $\LL$ and $s\in\nat$.
-We will construct a recursive function $\tau$ such that for every $\eqso(\LL)$ formula $\alpha$, it outputs an equivalent formula $\tau(\alpha)$ which is in $\LL$-SNF.
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$ which is clearly equivalent and already in $\LL$-SNF.
-If $\alpha = s$, then we define $(\top \add \cdots \add \top)$ ($s$ times), which also satisfies the condition.
-We assume that for every sub-formula $\beta$ of $\alpha$, $\tau(\beta)$ is an equivalent formula in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$, then we define $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$.
-If $\alpha = \sa{x}\beta$, then let $\tau(\beta) = \sum_{i = 1}^{k}\beta_i$ for some $k$ where each $\beta_i$ is in $\LL$-PNF.
-By grouping together the terms in the sum, we define $\tau(\alpha) = \sum_{i = 1}^{m}\sa{x}\beta_i$ which satisfies the condition and is equivalent to $\alpha$.
-If $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case.
-This covers all possible cases for $\alpha$ and we conclude the proof by taking $\tau(\alpha)$ as the desired rewrite of $\alpha$.
-
-\subsection*{Proof of theorem \ref{theo-pi1-pnf}}
-We divide the proof in three parts.
-
-\vspace{1em}
-First, we prove that the formula $\alpha_{0} = \left( \sa{X} 1 \right) + 1$ with $\arity(X) = 1$ (i.e.
-the function $2^{n}+1$, where $n$ is the size of the input structure) is not equivalent to any formula in $\loge{0}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ in $\loge{0}$-PNF that is equivalent to the $\eqso(\loge{0})$ formula $\alpha_0$.
-In \cite{SalujaST95}, it was proved that if $\length{\bar{X}} > 0$, then the function defined by $\alpha$, for big enough structures, is always even which is not possible in our case.
-On the other hand, if $\alpha$ is of the form $\sa{\bar{x}}\varphi(\bar{x})$, then $\alpha$ defines a polynomially bounded function which also leads to a contradiction.
-
-\vspace{1em}
-Second, we prove that the formula $\alpha_{1} = 2$ (i.e.
-the constant $2$) is not equivalent to any formula in $\loge{1}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to the $\eqso(\loge{1})$ formula $2$.
-First, if $\length{\bar{X}} = \length{\bar{x}} = 0$, then the function defined by $\alpha$ is never greater than 1. 
-Therefore, suppose that $\length{\bar{X}} > 0$ or $\length{\bar{x}} > 0$, and consider any ordered structure $\A$. 
-Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
-Now consider the ordered structure $\A'$ that is obtained by duplicating $\A$.
-This is, each half of $\A'$ is isomorphic to $\A$.
-Note that $\A'\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$ for $i = 1,2$ and there exists a third assignments $(\bar{B}_1',\bar{b}_1',\bar{a}_1')$ that is isomorphic to $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$ but in the other half to the structure such that $\A'\models\varphi(\bar{B}_1',\bar{b}_1',\bar{a}_1')$.
-We have that $\sem{\alpha}(\A) \geq 3$ which is a contradiction.
-
-\vspace{1em}
-We now show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there is an equivalent formula $\beta$ in $\LL$-PNF.
-As in Theorem \ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such formula.
-As we showed, there exists an equivalent formula in $\LL$-SNF, so we assume that $\alpha$ is in that form.
-Let $\alpha = \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-SNF.
-Without lost of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
-If not, we replace each $\alpha_i$ for the equivalent formula $\sa{\bar{X}} \sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge\forall z\,Y(z) \wedge \forall z(y \leq z))$.
-
-Now we begin describing the function $\tau$.
-If $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then the formula is already in $\LL$-PNF so we define $\tau(\alpha) = \alpha$.
-If $\alpha = \alpha_1 + \alpha_2$, then we assume that $\tau(\alpha_1) = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ and $\tau(\alpha_2) = \sa{\bar{Y}}\sa{\bar{y}}\psi(\bar{Y},\bar{y})$.
-The construction that we will provide for this function works by identifying a ``first'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ and a ``last'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$.
-These are identified by the following formulas:
-\begin{align*}
-\gamma_{\text{first}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z}), \\
-\gamma_{\text{last}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x}).
-\end{align*}
-Similarly, we define the formulas $\gamma_{\text{first}}(\bar{Y},\bar{y})$ and $\gamma_{\text{last}}(\bar{Y},\bar{y})$ (for the sake of simplicity we reuse the names $\gamma_{\text{first}}$ and $\gamma_{\text{last}}$).
-In other words, the ``first'' assignment is the one where every second-order predicate is empty and the first-order assignment is the lexicographically smallest, and the ``last'' assignment is the one where every second-order predicate is full and the first-order assignment is the lexicographically greatest.
-We also need to identify assignments that are not first and are not last.
-We do this by negating the two formulas above and grouping together the first-order variables:
-\begin{align*}
-\gamma_{\text{not first}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{z}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{z}_i)), \\
-\gamma_{\text{not last}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{x} < \bar{z}_0 \vee \bigvee_{i = 1}^{\length{\bar{X}}}\neg X(\bar{z}_i)),
-\end{align*}
-where $\bar{z} = (\bar{z}_0,\bar{z}_1,\ldots,\bar{z}_{\length{\bar{X}}})$. The following formula is equivalent to $\alpha$:
-\begin{align}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{not first}}(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{Y},\bar{y}))\vee \label{eq:partition1} \\
-&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))\vee \label{eq:partition2}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{not last}}(\bar{Y},\bar{y}))\vee \label{eq:partition3}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{last}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))]. \label{eq:partition4}
-\end{align}
-To show that the formula is indeed equivalent to $\alpha$, note that the formulas in lines (\ref{eq:partition1}) and (\ref{eq:partition2}) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$, and the formulas in lines (\ref{eq:partition3}) and (\ref{eq:partition4}) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
-Altogether, the four lines define pairwise disjoint assignments for $(\bar{X},\bar{x}),(\bar{Y},\bar{y})$.
-With this, it is straightforward to show that the above formula is equivalent to $\alpha$.
-However, the formula is not yet in the correct form since it has existential quantifiers in the subformulas $\gamma_{\text{not first}}$ and $\gamma_{\text{not last}}$.
-To solve this, first let take a close look to the complete formula:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\exists\bar{v}(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\exists\bar{w}(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))].
-\end{align*}
-To construct an equivalent formula that is in the correct form, we define $\bar{u} = (\bar{v},\bar{w})$ and we replace the first-order quantifiers by a first-sum and count the first assignment to $\bar{v}$ and $\bar{w}$ that satisfies the formula.
-A similar construction was used in \cite{SalujaST95}.
-Then the final formula equivalent to $\alpha$ is the following:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
-& \hspace{3cm}  \bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&
-(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))\vee
-\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
-& \hspace{3cm} (\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee \\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))].
-\end{align*}
-Finally, consider a $\eqso(\LL)$ formula $\alpha$ in $\LL$-SNF.
-If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we can consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the condition in the hypothesis.
-
-\subsection*{Proof of Theorem \ref{prop-rest}}
-We give this proof in three parts.
-
-\vspace{1em}
-First, we show that $\QE{0} \not\subseteq \E{1}$.
-By contradiction, suppose that there is a $\QE{0}$ formula $\alpha$ over some signature $\R$ such that defines the following function.
-For every finite $\R$-structure with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n - 1$.
-We use the following claim.
-\begin{claim}
-Let $\alpha = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. Then the function defined by $\alpha$ is either null, greater or equal to $n$, or is in $\Omega(n^2)$.
-\end{claim}
-\begin{proof}
-Suppose that the function defined by $\alpha$ is not $0$ and that $\varphi$ is in DNF. Furthermore, suppose $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$.
-Then $\alpha = \sa{\bar{x}} \varphi_1(\bar{x}) \vee \cdots \vee \varphi_n(\bar{x})$.
-Since $\alpha$ is not null, then some $\varphi_i$ must be satisfiable.
-This is, the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is not null.
-We will prove by induction on $\length{\bar{x}}$ that the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is either greater or equal to $n$, or in $\Omega(n^2)$.
-We address the case $\length{\bar{x}}= 1$, then $\alpha = \sa{x}\bigwedge\psi(x)$.
-If any $\psi(x) = (x = x)$ or $\neg(x < x)$, then we can eliminate it and we obtain the same function.
-If any $\psi = (x < x)$ or $\neg(x=x)$, then the function becomes null.
-If $\psi(x) = R(x,\ldots,x)$ for some $R\in\R$ the function becomes null for the structures we are considering.
-If $\psi(x) = \neg R(x,\ldots,x)$, we can eliminate it and for the structures we are considering we obtain the same function.
-The only possible $\alpha$ left is $\alpha = \sa{x}\top$ which is equal to the function $n$.
-This covers all possible cases for $\length{\bar{x}} = 1$.
-Now suppose that it holds for $\length{\bar{x}} = k$ and suppose $\alpha = \sa{\bar{x}}\bigwedge\psi(\bar{x})$ for $\length{\bar{x}} = k+1$.
-If any $\psi(\bar{x}) = (x_i = x_j)$ where $i \neq j$, then $\alpha$ describes the same function as $\alpha$ where $x_j$ has been replaced by $x_i$.
-In this formula the tuple of first-order variables has $k$ elements so the function it describes if one of the mentioned in the hypothesis.
-If $i = j$, then we can eliminate it and obtain the same function.
-If any $\psi(\bar{x}) = R(\bar{v})$ or $\neg R(\bar{v})$ where $\bar{v}$ is a sub-tuple of $\bar{x}$ then we can either eliminate it or the function becomes null, following the same argument as in the case $\length{\bar{x}} = 1$.
-If any $\psi(\bar{x}) = \neg(x_i = x_j)$ or $(x_i < x_j)$ where $i = j$, then the function becomes null.
-If any $\psi(\bar{x}) = \neg(x_i < x_j)$ where $i = j$, we can eliminate it.
-The remaining formulas in $\bigwedge\psi(\bar{x})$ are either $\neg(x_i = x_j)$, $(x_i<x_j)$ or $\neg(x_i<x_j)$.
-If the formula violates transitivity in $<$ (for example, $x < y \wedge y < z \wedge z < x$), then the function $\alpha$ describes is null.
-Therefore, there is some order over $\bar{x}$ that satisfies $\bigwedge\psi(\bar{x})$.
-Consider the formula that describes this order (like $x_1 < x_3 \wedge x_3 < x_4 \wedge x_4 < x_2$).
-The function $\alpha$ describes is greater or equal to the one this formula describes, which is exactly $\binom{n}{\length{\bar{x}}}$ which is in $\Omega(n^{\length{\bar{x}}}) \subseteq \Omega(n^2)$ if $\length{\bar{x}} > 1$.
-This concludes the proof of the claim.
-\end{proof}
-We suppose that $\alpha$ is in SNF, this is, $\alpha = \sum_{i = 1}^n\alpha_i$.
-Since $\alpha$ is not null, consider some $\alpha_i$ that describes a non-null function.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, where $\varphi$ is quantifier-free.
-Note that if $\length{\bar{X}} > 0$, then the function $\alpha$ describes is in $\Omega(2^n)$, as it was proven by the authors in \cite{SalujaST95}.
-We have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$, as we proved in the claim, describes either some function greater or equal to $n$, or in $\Omega(n^2)$, which leads to a contradiction.
-Lastly, note that the formula $\sa{x}\exists y(x < y)$ is in $\E{0}$ and describes the function $n-1$, which concludes the proof.
-
-\vspace{1em}
-Now we show that $\E{1}\not\subseteq\QE{0}$.
-In Theorem \ref{theo-pi1-pnf} we proved that there is no formula in $\loge{1}$-PNF equivalent to the formula $\alpha = 2$.
-Every formula in $\E{1}$ can be expressed in $\loge{1}$-PNF, which implies that $2 \in \QE{0}$ and $2 \not\in \E{1}$.
-
-\vspace{1em}
-Lastly, we prove that $\eqso(\loge{1})\subsetneq\eqso(\logu{1})$.
-For inclusion, let $\alpha$ be a formula in $\eqso(\loge{1})$.
-Suppose that it is in $\loge{1}$-SNF.
-This is, $\alpha = c + \sum_{i = 1}^{n}\alpha_i$.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, where $\varphi_i$ is quantifier-free, for each $\alpha_i$.
-We use the same construction used in \cite{SalujaST95}, and we obtain that the formula $\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \forall\bar{y}'(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')$ for every assignment to $(\bar{X},\bar{x})$.
-We do this replacement for each $\alpha_i$ and we obtain an equivalent formula in $\eqso(\logu{1})$.
-
-To prove that the inclusion is proper, consider the $\eqso(\logu{1})$ formula $\sa{x}\forall y(y = x)$. This formula defines the following function that takes an ordered structure $\A$ as input:
-$$
-\sem{\alpha}(\A) = 
-\begin{cases}
-1 &\A \text{ has one element}\\
-0 &\text{ otherwise}.
-\end{cases}
-$$
-Suppose that there exists an equivalent formula $\alpha$ in $\eqso(\loge{1})$.
-Also, suppose that it is in $\L$-PNF, so $\alpha = c + \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$.
-Since $\alpha$ takes the value 0 for some structures, $c$ must be 0.
-Consider a structure $\mathfrak{1}$ with one element.
-We have that for some $i$, there exists an assignment $(\bar{B},\bar{b},\bar{a})$ for $(\bar{X},\bar{x},\bar{y})$ such that $\mathfrak{1}\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
-Consider now the structure $\mathfrak{2}$ that is obtained by duplicating $\mathfrak{1}$, as we did for Theorem \ref{theo-pi1-pnf}.
-Note that $\mathfrak{2}\models\varphi_i(\bar{B},\bar{b},\bar{a})$, which implies that $\sem{\alpha}(\mathfrak{2}) \geq 1$, which leads to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop-e1-nc}}
-
-Towards a contradiction, assume that the statement is false.
-This is, that $\E{1}$ is closed under binary sum.
-Consider the formula $\sa{x}(x = x)$ which is in $\E{1}$ over some signature $\R$.
-For every finite $\R$-structure $\A$ with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n$.
-From our assumption, there exists some formula in $\E{1}$ equivalent to the formula $\alpha \add \alpha$, which describes the function $2n$.
-Let $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ be this formula, where $\varphi$ is quantifier-free.
-Note that the function defined by this formula is equal or greater than the one defined by $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ divided by a polynomial factor.
-More specifically, for each ordered structure $\A$ with domain $A$, we have the following inequality:
-$$
-\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot \vert A \vert^{\length{\bar{y}}} \geq \sem{\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A)
-$$
-Note that the formula $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ defines a function in $\E{0}$.
-It was shown by the authors in \cite{SalujaST95} that every function in $\E{0}$ grows exponentially over the size of the structure for large enough structures, when $\length{\bar{X}} > 0$.
-This function divided by a polynomial factor still grows exponentially.
-Therefore, for $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ we have that $\length{\bar{X}} = 0$.
-
-Now, for the formula $\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ consider a structure $\mathfrak{1}$ with only one element $a$.
-We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$, but the only possible assignment to $\bar{x}$ is the tuple $(a,\ldots,a)$ so $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which follows to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop:qe0-fp-qe1-totp-fptras}}
-
-The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$.
-This is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every finite $\R$-structure $\A$ with domain $A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$.
-We base our proof on this fact.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is contained in \textsc{TotP}.} Let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Consider the following nondeterministic procedure that on input $\enc(\A)$ generates $\sem{\alpha}(\A)$ branches.
-For each $\alpha_i = \varphi$, where $\varphi$ is a $\loge{1}$ formula, it checks if $\A\models\varphi$ in polynomial time and generates a new branch if that is the case.
-For each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi$, this formula is also in $\E{1}$.
-We use the reduction to $\cdnf$ provided in \cite{SalujaST95} and we obtain $g(\enc(\A))$, which is an instance to $\cdnf$.
-Since $\cdnf$ is also in $\totp$ \cite{PagourtzisZ06}, we simulate the corresponding nondeterministic procedure that generates exactly $\cdnf(\enc(g(\A)))$ branches.
-Since, $\fp\subseteq\totp$\cite{PagourtzisZ06}, each polynomially computable function is also in $\totp$, and then on each of these branches we simulate the corresponding nondeterministic procedure to generate $h(\vert A \vert)$ more.
-The number of branches for each $\alpha_i$ is $\sem{\alpha_i}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$, and the total number of branches in the procedure amounts to $\sem{\alpha}(\A)$.
-We conclude that $\alpha\in\totp$.
-
-\vspace{1em}
-{\bf Every function in} $\boldsymbol{\eqso(\loge{1})}$ {\bf has an FPRAS.}  let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Note that each $\alpha_i$ that is equal to some $\loge{1}$ formula $\varphi$ has an FPRAS given by the procedure that simply checks if $\A\models\varphi$ and returns 1 if it does and 0 otherwise.
-Also, each remaining $\alpha_i$ has an FPRAS since $\alpha_i\in \E{1}$ \cite{SalujaST95}.
-If two functions have an FPRAS, then their sum also has one given by the procedure that simulates them both and sums the results.
-We conclude that $\alpha$ has an FPRAS.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is closed under sum and multiplication.} Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that the class is closed under multiplication.
-We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of $\so$.
-
-\begin{lemma} \label{conj-mult}
-If $\LL$ is a fragment closed under conjunction, then $\eqso(\LL)$ is closed under binary multiplication.
-\end{lemma}
-\begin{proof}
-We will define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ extended by binary product, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
-In fact, the formula $\tau(\alpha)$ is in $\LL$-SNF.
-First we replace each constant $s$ in $\alpha$ for $(\top \add \cdots \add \top)$ ($s$ times).
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$.
-We assume that for every $\beta$ that has less algebraic operators than $\alpha$, $\tau(\beta)$ is in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$ then we define $\tau(\alpha) = \tau(\alpha_1) + \tau(\alpha_2)$.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$, then we define $\tau(\alpha)$ as the formula in $\LL$-SNF that is equivalent to $\sa{x}\tau(\beta)$ and to $\tau(\alpha) = \sa{X}\tau(\beta)$, respectively.
-If $\alpha = (\alpha_1 \cdot \alpha_2)$, we assume that each $\alpha_i$ is in $\LL$-SNF.
-We identify three cases.
-(1) Some $\alpha_i$ is equal to $\sum_{j = 1}^n\beta_j$ for $n > 1$.
-Suppose wlog.
-that it is $\alpha_1$.
-We then define $\tau(\alpha) = \sum_{j = 1}^n\tau(\beta_j\cdot\alpha_2)$.
-In the following cases, $\alpha_1$ and $\alpha_2$ are in $\LL$-SNF.
-(2) If some $\alpha_i$ is equal to $\sa{X}\beta$ or $\sa{x}\beta$, we define $\tau(\alpha)$ as the $\LL$-SNF formula that is equivalent to $\sa{x}\tau(\beta\cdot\alpha_2)$ and $\sa{X}\tau(\beta\cdot\alpha_2)$, respectively.
-The remaining case is (3) $\alpha_1 = \varphi_1$ and $\alpha_2 = \varphi_2$ where each $\varphi$ is an $\LL$ formula.
-Then we define $\tau(\alpha) = \varphi_1 \wedge \varphi_2$.
-This covers all possible cases for $\alpha$.
-For every pair of formulas $\alpha,\beta$ in $\eqso(\LL)$, we have that their multiplication $(\alpha\cdot\beta)$ is a formula in the grammar $\eqso(\LL)$ extended by binary product, and so, there exists an equivalent formula $\tau(\alpha\cdot\beta)$ which is in unextended $\eqso(\LL)$.
-\end{proof}
-Since $\loge{1}$ is closed under conjunction, this also holds for $\eqso(\loge{1})$. This concludes the proof.
-
-\subsection*{Proof of Proposition \ref{pi-minusone}}
-
-Let $\LL$ be a fragment of $\fo$ that contains $\logu{1}$.
-Then we have that every function in $\U{1}$ is expressible in $\eqso(\LL)$.
-In particular, $\ctcnf \in \eqso(\LL)$.
-Suppose that $\eqso(\LL)$ is closed under subtraction by one.
-Then, the function $\ctcnf-1$, which counts the number of satisfying assignments of a 3-CNF formula minus one, is also in $\eqso(\LL)$.
-Note that $\eqso(\LL) \subseteq \eqso(\fo) = \shp$.
-We have that $\ctcnf$ is $\shp$-complete under parsimonious reductions\footnote{It can be easily verified that the standard reduction from SAT to 3-CNF (or 3-SAT) preserves the number of satisfying assignments}.
-Now, let $f$ be a function in $\shp$, and consider the nondeterministic polynomial-time procedure that on input $\enc(\A)$ computes the corresponding reduction to $\ctcnf$, name it $g(\enc(\A))$, and simulates the $\shp$ procedure for $\ctcnf-1$ on input $g(\enc(\A))$.
-We have that this is a $\shp$ procedure that computes $f-1$, from which we conclude that $\shp$ is closed under subtraction by one.
-
-\subsection*{Proof of Theorem \ref{sigmafo-minusone}}
-
-\noindent {\bf Closed under sum and multiplication.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
-This can be seen because $\eqso(\LL)$ is closed under sum for every fragment $\LL$ by definition, and since $\logex{1}$ is closed under conjunction, from Lemma \ref{conj-mult} it follows that $\eqso(\logex{1})$ is closed under multiplication.
-
-\vspace{1em}
-\noindent {\bf Easy decision version and FRPRAS.}  We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
-We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Theorem \ref{prop:qe0-fp-qe1-totp-fptras}.
-First, we define a function that converts a formula $\alpha$ in $\eqso(\logex{1})$ over a signature $\R$ into a formula $\lambda(\alpha)$ in $\eqso(\loge{1})$ over a signature $\R_{\alpha}$.
-Afterwards, we define a function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an $\R_{\alpha}$-structure $g_{\alpha}(\A)$.
-
-Let $\alpha$ be in $\eqso(\logex{0})$.
-The signature $\R_{\alpha}$ is obtained by adding the symbol $R_{\psi}$, for every $\fo$ formula $\psi(\bar{z})$ in $\alpha$, to $\R$.
-Each symbol $R_{\psi}$ represents a predicate with arity $\length{\bar{z}}$.
-Then, $\lambda(\alpha)$ is defined as $\alpha$ where each $\fo$ formula $\psi(\bar{z})$ has been replaced by $R_{\psi}(\bar{z})$.
-We now define the function $g_{\alpha}$ procedurally.
-Let $\A$ be a $\R$-structure with domain $A$.
-Let $\A'$ be an $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A}$ empty.
-For each $\fo$-formula $\psi(\z)$ with $\length{\bar{z}}$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length{\bar{z}}}$.
-If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$ (this can be done in $\ptime$).
-This concludes the construction of $\A'$.
-Note that the number of $\fo$ subformulas, arity and tuple size is fixed in $\alpha$, so computing this function takes polynomial time over the size of the structure.
-Moreover, the encoding of $\A'$ has polynomial size over the size of $\enc(\A)$.
-We define $g_{\alpha}(\A) = \A'$ and we have that for each $\R$-structure $\A$: $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$.
-Therefore, we have a parsimonious reduction from $\alpha$ to the $\eqso(\loge{1})$ formula $\lambda(\alpha)$.
-
-To show that $\alpha$ is in $\totp$, we can convert $\alpha$ and $\enc(\A)$ into $\lambda(\alpha)$ and $\enc(g_{\alpha}(\A))$, respectively, and run the procedure in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-Similarly, to show that $\alpha$ has an FPRAS, we do the same as before and simulates the FPRAS for $\lambda(\alpha)$ in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-These procedures also takes polynomial time and satisfies the condition.
-
-\vspace{1em}
-\noindent {\bf Closed under subtraction by one.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. 
-For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Without lost of generality, we assume that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
-Moreover, we assume that $\length{\bar{x}} > 0$  since, if this is not the case, we can replace $\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
-
-The proof will be separated in two parts.
-In the first part, we suppose that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$.
-Then we will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ is the formula.
-In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ the sum of one or more formulas in $\logex{1}$-PNF.
-We define a formula $\varphi'$ that satisfies the following condition: if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ and $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be proven.
-
-\vspace{1em}
-
-\noindent {\em Part (1).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ where $\varphi$ is an  $\fo$-formula.
-Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}} \varphi(\bar{x})$ (i.e.
-$\length{\bar{X}} = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}} [\varphi(\bar{x})\wedge\exists\bar{z}(\varphi(\bar{z})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
-Therefore, for the rest of the proof we can assume that $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
-
- Let $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$ and $\bar{X} = (X_1,\ldots,X_{\length{\bar{X}}})$.
-To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in into a DNF formula. 
-More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i$ is a conjunction of atomic formulas or negation of atomic formulas.
-Furthermore, we suppose that each conjunct $\varphi_i$ has the~form:
-$$
-\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
-\underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}} \wedge
-\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}}.
-$$
-Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ can appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
-
-Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ are mentioned in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
-We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
-\begin{itemize}
-	\item[(a)] In order to satisfy the first condition, consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
-We add $\length{\bar{w}}$ new variables $z_1,\ldots,z_{\length{\bar{w}}}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length{\bar{w}}})$.
-We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$.
-The formula $\varphi_i$ is now defined as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \exists \bar{z} \; (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{-}(\bar{X},\bar{x},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y})).
-	$$
-	We repeat this process for each instance of a $X_j(\bar{w})$ in $\varphi_i$, and we obtain a formula where none of the $X_j$'s acts over any variable in $\bar{x}$.
-We add the new first-order variables to $\bar{y}$ and we redefine $\varphi_i$ as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
-	$$
-	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = \exists \bar{y} \;  X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := \exists \bar{y} \;  X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
-	\item[(b)] An ordered partition on a set $S$ is defined by an equivalence relation $\sim$ over $S$, and a linear order over $S/\!\sim$.
-For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$.
-A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
-On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
-	For a given $k$, let $\cB_k$ be the number of possible ordered partitions for a set of size $k$. For $1 \leq j \leq \cB_{\length{(\bar{x},\bar{y})}}$ 
-	let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
-
-	Thus, the formula $\varphi_i(\bar{X},\bar{x})$ is then redefined as:
-	$$
-	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length{(\bar{x},\bar{y})}}} \exists \bar{y} \; [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
-	$$
-	Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
-	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  to $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
-	\item[(c)] For showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
-$\varphi_i$ is inconsistent), we do the following.
-If there exists an instance of $X_j(\bar{z})$ in $\varphi^{+}_i$, an instance of $\neg X_j(\bar{w})$ in $\varphi^{-}_i$ and the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$, then the entire formula $\varphi_i$ is removed from $\varphi$.
-\end{itemize}
-It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$.
-From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c).
-Furthermore, it has the following form:
-$$
-\varphi_i(\bar{X},\bar{x}) \; = \; \exists{\bar{y}} \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
-$$
-where $\varphi^{+}$ and $\varphi^{-}_i$ do not depend on  $\bar{x}$.
-\begin{claim}\label{claim:minusone}
-For a given ordered structure $\A$, if there exists a first-order assignment $v$ for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists a second-order assignment $V$ for $\A$ such that  $(\A,V,v)\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-\end{claim}
-\begin{proof}
-Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
-Define $\bar{B} = (B_1,\ldots,B_{\length{\bar{X}}})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
-Towards a contradiction, suppose that $(\A,V,v)\not\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-By the choice of $v$, and construction of $V$ it is clear that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})\wedge\varphi^{+}_i(\bar{X},\bar{y})$, so we necessarily have that $(\A,V,v)\not\models\varphi^{-}_i(\bar{X},\bar{y})$.
-Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,V, v)\not\models\neg X_t(\bar{w})$, namely, $v(\bar{w})\in B_t$. 
-However, by the construction of $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$.
-Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$.
-This violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
-\end{proof}
-
-The previous claim and proof motivates the following definitions.
-For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length{\bar{X}}})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
-One can easily check that for every assignments $V, v$ such that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$, then it holds that $(\A,\bar{B}^v,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all the assignment of $\bar{X}$ when $v$ is fixed.
-Therefore, by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
-For this, the idea is to choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
-$$
-\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge\forall\bar{x}'\forall\bar{y}'(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')).
-$$ 
-If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
-Furthermore, let $V^*$ the second order assignment that satisfies $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*$ the first order assignment such that $v^*(\bar{x}) = v(\bar{x})$.
-By the previous discussion, we have that $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.
-
-Now, we have all the tools to introduce the main formula of the proof. 
-Intuitively, we want to exclude the assignment $(V^*, v^*)$ as a satisfying assignment for $\varphi_i(\bar{X},\bar{x})$.
-Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if, 
-if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, then $V \neq V^*$ or $v \neq v^*$. 
-That is, it excludes all pairs $(V, v)$ that coincide with $(V^*, v^*)$ in $(\bar{X}, \bar{x})$ whenever $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. This property can be defined as follows:
-\begin{align}
-\psi_i(\bar{X},\bar{x}) = & \;\; \big( \, \exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
-& \hspace{1cm} \Big( \; 
-\exists\bar{y} \;\; \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i^+(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\
-& \hspace{1.5cm} (\exists\bar{x}'\exists\bar{y}\, \varphi_i(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}) \;\; \Big)  \label{eq:part3}
-\end{align}
-To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
-Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-Then, the conclusion of the implication at (\ref{eq:part2}) and (\ref{eq:part3}), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
-Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by showing that if $\varphi_i(\bar{X},\bar{x})$ is satisfied then $V(\bar{X})$ must contain one element more that it is not in $V^*(\bar{X})$.
-The second disjunct (\ref{eq:part3}) is satisfied when $v(\bar{x})$ is not the lexicographically smallest tuple that satisfies $\varphi_i$ (i.e. $v(\bar{x}) \neq v^*(\bar{x})$).
-
-We are ready to define the formula  $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$
-where each modified disjunct $\varphi_i'(\bar{X},\bar{x})$ is constructed as follows. 
-For the sake of simplification, define the auxiliary formula $\chi_i = \neg\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$. 
-This formula basically checks if $\varphi_i$ is satisfiable or not.
-Then, first define the formula:
-$$
-\varphi'_1(\bar{X},\bar{x}) = \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
-$$
-This models that all the assignments that satisfy $\varphi_1$ also satisfy $\varphi'_1$, except for $(V^*,v^*)$. Then we define:
-$$
-\varphi'_2(\bar{X},\bar{x}) = \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
-$$
-This models that all the assignments that satisfy $\varphi_2$ also satisfy $\varphi'_2$, but if there was no assignment for $\varphi_1$, then we exclude $(V^*,v^*)$ for $\varphi_2$.
-Then we generalize this construction for each $\varphi_i$:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})),
-$$
-By the construction of $\kappa(\alpha)$, one can easily check that  $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.
-
-\vspace{1em}
-
-{\em Part (2).} Let $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ for some $\eqso(\logex{1})$ formula $\beta$. We define $\kappa(\alpha)$ as follows.
-First, perform the same transformations to $\varphi(\bar{X},\bar{x})$ as in the previous section.
-Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c).
-We also use the previously defined formulas $\chi_i$ and $\psi_i$, for each $i$.
-
-We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if and only if $\sem{\beta}(\A) = 0$.
-If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg\exists\bar{x}'\varphi(\bar{x}')$.
-If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$ as each $\chi_i$ was previously defined.
-If $\beta = (\beta_1 + \beta_2)$, then $\lambda(\beta) = \lambda(\beta_1) \wedge \lambda(\beta_2)$.
-Now, for each $\varphi_i$ we define:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x}))\Big)\Big).
-$$
-Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
-
-%\subsection*{Proof of Theorem \ref{sub-pnp}}
-%
-%Towards a contradiction, suppose that any of the classes $\E{1}$, $\eqso(\loge{1})$, and $\eqso(\logex{1})$ is closed under subtraction.
-%
-%Let $\R$ include the symbols $S_1, S_2, S_3, S_4$, which describe the following properties.
-If a finite $\R$-structure $\A$ defines a 3DNF formula $\Phi$, then its domain is the set of variables mentioned in $\Phi$, and for each $i = 1,2,3,4$:
-%\begin{align*}
-%	S_1^\A &= \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_2^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_3^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_4^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
-%\end{align*}
-%Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
-%\begin{multline*}
-%\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
-%\end{multline*}
-%Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
-%$$
-%\varphi(x,X) = (T(x) \vee \neg T(x)).
-%$$
-%Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Given that $f_{\#3DNF}, f_{all} \in \E{1} \subseteq \eqso(\loge{1}) \subseteq \eqso(\logex{1}) \subseteq \totp$, and at least one of the classes is closed under subtraction, the function $f_{\#3DNF} - f_{all}$ is in $\totp$.
-However, note that for each $\R$-structure $\A$ that represents a formula $\Phi$, $(f_{\#3DNF} - f_{all})(\A) = 0$ if an only if $\Phi$ is a tautology.
-The decision version of this function is the $\np$-complete problem $\textsc{Tautology}$.
-And since the function is in $\totp$, its decision version is also in $\ptime$.
-We conclude that $\np \subseteq \ptime$.
-
-\subsection*{Proof of Proposition \ref{prop:ehorn-pe}}
-Pagourtzis and Zachos mention a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula \cite{PagourtzisZ06}.
-This procedure can be easily extended to receive Horn formulas, and furthermore, a disjunction of Horn formulas.
-Hence $\shdhsat$ is in $\totp$.
-
-As we show in Proposition \ref{sigma2hard}, $\shdhsat$ is complete for $\eqso(\ehorn)$ under parsimonious reductions.
-Let $\alpha$ be a formula in $\eqso(\ehorn)$ and let $g_{\alpha}$ be the reduction to $\shdhsat$.
-The $\totp$ procedure we construct, for each input $\enc(\A)$, is simply to compute $g_{\alpha}(\enc(\A))$, and then simulate the $\totp$ procedure for $\shdhsat$ on input $g_{\alpha}(\enc(\A))$.
-We conclude that $\alpha$ is in $\totp$.
-
-\subsection*{Proof of Proposition \ref{prop:hsat-not-sigma2}} %V.12
 We use a similar proof to the one provided by the authors in \cite{SalujaST95} to separate the classes $\E{2}$ and $\U{2}$.
@@
 which is unsatisfiable. We arrive to a contradiction and we conclude that $\chsat$ is not in $\eqso(\ehorn)$.
-
-\subsection*{Proof of Theorem \ref{sigma2hard}} %V.13
-
-First we prove that $\shdhsat$ is in $\eqso(\ehorn)$.
-Recall that each instance of $\shdhsat$ is a disjunction of Horn formulas.
-Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot), \pD(\cdot,\cdot)\}$.
-Each symbol in this vocabulary is used to indicate the same as in Example \ref{ex-hornsat-esop1}, with the addition of $\pD(d,c)$ which indicates that $c$ is a clause in the formula $d$.
-Recall that the formula
-\begin{align*}
-&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \neg\textit{T}(x)).
-\end{align*}
-defines $\chsat$. We obtain the following formula $\psi(T,A)$ in $\ehorn$:
-\begin{align*}
-\exists d[&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \pD(c,d)\vee \neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \neg\textit{T}(x))]
-\end{align*}
-effectively defines $\chsat$ as for every disjunction of Horn formulas $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
-Therefore, we conclude that $\shdhsat \in \eqso(\ehorn)$.
-
-\vspace{1em}
-We will now prove that $\shdhsat$ is hard for $\eqso$ over a signature $\R$ under parsimonious reductions.
-For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a polynomial-time procedure that computes a function $g_{\alpha}$.
-This function receives a finite $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
-We suppose that $\alpha$ is in sum normal form and:
-$$
-\alpha = \sum_{i = 1}^{\text{\#clauses}} \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigwedge_{j = 1}^{n}\forall\bar{z}\,\varphi^i_j(\bar{X},\bar{x},\bar{y},\bar{z}),
-$$
-where each $\varphi^i_j$ is a Horn clause.
-
-Consider a finite $\R$-structure $\A$ with domain $A$.
-To simplify the proof, we extend our grammar to allow first-order constants.
-Consider each tuple $\bar{a}\in A^{\length{\bar{x}}}$, each $\bar{b}\in A^{\length{\bar{y}}}$ and each $\bar{c}\in A^{\length{\bar{z}}}$ as a tuple of first-order constants.
-The following formula defines the same function as $\alpha$:
-$$
-\sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length{\bar{x}}}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length{\bar{y}}}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length{\bar{z}}}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
-$$
-Note that each $\fo$ formula over $(\bar{x},\bar{y},\bar{z})$ in each $\varphi^i_j$ has no free variables.
-Therefore, we can evaluate each of these in polynomial time and replace them by $\perp$ and $\top$ where it corresponds.
-Each $\varphi^i_j$ will be of the form $\perp \vee\, \chi^i_j(\bar{X})$ or $\top \vee \chi^i_j(\bar{X})$ where $\chi^i_j$ is a disjunction of $\neg X_{\ell}$'s and at most one $X_{\ell}$.
-The formulas of the form $\top \vee \chi^i_j(\bar{X})$ can be removed entirely, and the formulas of the form $\perp \vee\, \chi^i_j(\bar{X})$ can be replaced by $\chi^i_j(\bar{X})$.
-We obtain the formula
-$$
-\sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})
-$$
-where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$.
-
-Our idea for the rest of the proof is to define $g_{\alpha}$ for each $\alpha = \sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})$, for $\alpha = c$ and for $\alpha = \beta_1 + \cdots + \beta_m$ where each $\beta_i$ is in one of the two previous cases.
-
-If $\alpha$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ where $\psi_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$, then we obtain the {\bf propositional formula} $g_{\alpha}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulas.
-
-If $\alpha$ is equal to a constant $c$, then we define $g_{\alpha}(\A)$ as the following formula that has exactly $c$ satisfying assignments:
-$$
-g_{\alpha}(\A) = \bigvee_{i = 1}^{c}\neg t_1 \wedge \cdots \wedge \neg t_{i-1} \wedge t_i \wedge \neg t_{i+1} \wedge \cdots \wedge \neg p_c.
-$$ 
-If $\alpha = \beta_1 + \cdots + \beta_m$, let $g_{\beta_i}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\theta^i_{j,k}$ for each $\beta_i$ where each $\theta^i_{j,k}$ is a Horn clause.
-Let $\Theta_i = g_{\beta_i}(\A)$.
-We rename the variables in each $\Theta_i$ so none of them are mentioned in any other $\Theta_j$.
-We add $m$ new variables $t_1,\ldots,t_m$ and we define:
-\begin{align*}
-g_{\alpha}(\A) = &\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^1_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m}}}t) \wedge (t_1 \wedge \bigwedge_{\ell = 2}^{m} \neg t_{\ell})) \vee \\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^2_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each $t$}\\ \text{ mentioned in}\\ \Theta_1,\Theta_3,\ldots,\Theta_{m}}}t) \wedge (t_2 \wedge \bigwedge\limits_{\substack{\ell = 1 \\ \ell \neq 2}}^{m} \neg t_{\ell})) \vee \cdots \vee\\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^m_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m-1}}}t) \wedge (t_m \wedge \bigwedge_{\ell = 1}^{m-1} \neg t_{\ell})).
-\end{align*}
-The formula is a disjunction of Horn formulas, and the number of satisfying assignments for this formula is exactly the sum of satisfying assignments for each $g_{\beta_i}(\A)$.
-This, at the same time, is equal to $\sem{\alpha}(\A)$.
-This covers all possible cases for $\alpha$, and the entire procedure takes polynomial time.

### 2017-03-03 LMCS/proof-lem-support.tex
journal primera version

{\bf For }$\boldsymbol{\fqfo(\fo).}$ Let $\R$ be a signature. We prove the statement for $\fqfo(\fo)$ over $\R$.

We prove inductively that for each formula $\beta(\bar{x},h)$ in $\fqfo(\fo)$ over $\R$, for a given pair of functions $f,g$ such that $\supp(f)\subseteq\supp(g)$, it holds that $\supp(T_{\beta}(f))\subseteq\supp(T_{\beta}(g))$. Let $\length{\bar{x}} = \ell$.

We separate the proof in each case determined by the $\fqfo$ grammar. For each of the following cases.
\begin{itemize}
\item[1.] $\beta$ is either equal to a constant $s$ or an $\fo$ formula $\varphi$. Then $h$ does not appear. Then, for each structure $\A$, each first-order assignment $v$ and functional assignments $F,G$ over $\A$, we have that $\sem{\beta(\bar{x},h)}(\A,v,F) = \sem{\beta(\bar{x},h)}(\A,v,G)$. As a result, $\supp(T_{\beta}(f)) = \supp(T_{\beta}(g))$ for every pair of functions $f,g$.
\item[2.] $\beta$ is equal to $h(\bar{y})$ for some subtuple $\bar{y}$ of $\bar{x}$. Then $T_{\beta}(f) = f$ and $T_{\beta}(g) = g$ and the condition holds trivially.
\end{itemize}
Suppose that the statement holds for each formula smaller than $\beta$.
\begin{itemize}
\item[3.] $\beta = (\beta_1 + \beta_2)$. It is easy to see that for each $\bar{a} \in A^{\ell}$ and function $f:A^{\ell}\to\nat$: $T_{\beta}(f)(\bar{a}) = T_{\beta_1}(f)(\bar{a}) + T_{\beta_2}(f)(\bar{a})$. Suppose $\supp(f)\subseteq\supp(g)$ and let $\bar{a} \in \supp(T_{\beta}(f))$, or in other words, $T_{\beta}(f)(\bar{a}) > 0$. Then, for some $\beta_i$ it holds that $T_{\beta_i}(f)(\bar{a}) > 0$. From the supposition we have that $T_{\beta_i}(g)(\bar{a}) > 0$ from which the statement follows.
\item[4.] $\beta = (\beta_1 \mult \beta_2)$. It is easy to see that for each $\bar{a}$ in $A^{\ell}$ and function $f:A^{\ell}\to\nat$: $T_{\beta}(f)(\bar{a}) = T_{\beta_1}(f)(\bar{a}) \mult T_{\beta_2}(f)(\bar{a})$. Suppose $\supp(f)\subseteq\supp(g)$ and let $\bar{a}$ be such that $T_{\beta}(f)(\bar{a}) > 0$. Then $T_{\beta_i}(f)(\bar{a}) > 0$ for both $\beta_i$. From the supposition we have that $T_{\beta_i}(g)(\bar{a}) > 0$ for both $\beta_i$ and the statement holds.
\item[5.] $\beta = \sa{y}\delta(y,\bar{x},h)$. Here we extend the grammar slightly to allow constants, and we use the notation $\delta[a/y]$ to denote the formula obtained by replacing each instance of $y$ by the constant $a$. It can be seen that $T_{\beta}(f)(\bar{a}) = \sum_{a \in A} T_{\delta[a/y]}(f)(\bar{a})$. Suppose $\supp(f)\subseteq\supp(g)$ and let $\bar{a}$ be such that $T_{\beta}(f)(\bar{a}) > 0$. Then for some $a\in A$ we have $T_{\delta[y/a]}(f)(\bar{a}) > 0$. The statement now follows as in the case 3.
\item[6.] $\beta = \pa{y}\delta(y,\bar{x},h)$. 
It can be seen that $T_{\beta}(f)(\bar{a}) = \prod_{a \in A} T_{\delta[a/y]}(f)(\bar{a})$. 
%If $T_{\beta}(f)(\bar{a}) > 0$, then $T_{\delta[y/a]}(f)(\bar{a}) > 0$ for each $a\in A$. As in the $\mult$ case, the statement follows directly.
The statement follows using the same argument from cases 4 and 5.
\end{itemize}
This covers all possible cases for $\beta$ and we finish the proof of the statement for $\fqfo(\fo)$.

\vspace{1em}
{\bf For} $\boldsymbol{\rqfo(\fo).}$ The only additional case is where $\beta = \clfp{\delta(\bar{y},h')}$ for some subtuple $\bar{y}$ of $\bar{x}$. We have that $\beta$ does not mention $h$, and so, the statement follows directly as we showed in the previous part of the proof.

### 2017-03-03 LMCS/proof-pi-minusone.tex
journal primera version

--- 2017-01-12 LICS/appendix-syntactic.tex
@@
-\subsection*{Proof of Theorem \ref{theo-pnf-snf}}
-Recall that a formula in $\eqso(\LL)$ is on the following grammar:
-$$
-\alpha = \varphi \ \mid \ s \ \mid \ (\alpha + \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \sa{X} \alpha,
-$$
-where $\varphi$ is a formula in $\LL$ and $s\in\nat$.
-We will construct a recursive function $\tau$ such that for every $\eqso(\LL)$ formula $\alpha$, it outputs an equivalent formula $\tau(\alpha)$ which is in $\LL$-SNF.
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$ which is clearly equivalent and already in $\LL$-SNF.
-If $\alpha = s$, then we define $(\top \add \cdots \add \top)$ ($s$ times), which also satisfies the condition.
-We assume that for every sub-formula $\beta$ of $\alpha$, $\tau(\beta)$ is an equivalent formula in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$, then we define $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$.
-If $\alpha = \sa{x}\beta$, then let $\tau(\beta) = \sum_{i = 1}^{k}\beta_i$ for some $k$ where each $\beta_i$ is in $\LL$-PNF.
-By grouping together the terms in the sum, we define $\tau(\alpha) = \sum_{i = 1}^{m}\sa{x}\beta_i$ which satisfies the condition and is equivalent to $\alpha$.
-If $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case.
-This covers all possible cases for $\alpha$ and we conclude the proof by taking $\tau(\alpha)$ as the desired rewrite of $\alpha$.
-
-\subsection*{Proof of theorem \ref{theo-pi1-pnf}}
-We divide the proof in three parts.
-
-\vspace{1em}
-First, we prove that the formula $\alpha_{0} = \left( \sa{X} 1 \right) + 1$ with $\arity(X) = 1$ (i.e.
-the function $2^{n}+1$, where $n$ is the size of the input structure) is not equivalent to any formula in $\loge{0}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ in $\loge{0}$-PNF that is equivalent to the $\eqso(\loge{0})$ formula $\alpha_0$.
-In \cite{SalujaST95}, it was proved that if $\length(\bar{X}) > 0$, then the function defined by $\alpha$, for big enough structures, is always even which is not possible in our case.
-On the other hand, if $\alpha$ is of the form $\sa{\bar{x}}\varphi(\bar{x})$, then $\alpha$ defines a polynomially bounded function which also leads to a contradiction.
-
-\vspace{1em}
-Second, we prove that the formula $\alpha_{1} = 2$ (i.e.
-the constant $2$) is not equivalent to any formula in $\loge{1}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to the $\eqso(\loge{1})$ formula $2$.
-First, if $\length(\bar{X}) = \length(\bar{x}) = 0$, then the function defined by $\alpha$ is never greater than 1. 
-Therefore, suppose that $\length(\bar{X}) > 0$ or $\length(\bar{x}) > 0$, and consider any ordered structure $\A$. 
-Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
-Now consider the ordered structure $\A'$ that is obtained by duplicating $\A$.
-This is, each half of $\A'$ is isomorphic to $\A$.
-Note that $\A'\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$ for $i = 1,2$ and there exists a third assignments $(\bar{B}_1',\bar{b}_1',\bar{a}_1')$ that is isomorphic to $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$ but in the other half to the structure such that $\A'\models\varphi(\bar{B}_1',\bar{b}_1',\bar{a}_1')$.
-We have that $\sem{\alpha}(\A) \geq 3$ which is a contradiction.
-
-\vspace{1em}
-We now show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there is an equivalent formula $\beta$ in $\LL$-PNF.
-As in Theorem \ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such formula.
-As we showed, there exists an equivalent formula in $\LL$-SNF, so we assume that $\alpha$ is in that form.
-Let $\alpha = \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-SNF.
-Without lost of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.
-If not, we replace each $\alpha_i$ for the equivalent formula $\sa{\bar{X}} \sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge\forall z\,Y(z) \wedge \forall z(y \leq z))$.
-
-Now we begin describing the function $\tau$.
-If $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then the formula is already in $\LL$-PNF so we define $\tau(\alpha) = \alpha$.
-If $\alpha = \alpha_1 + \alpha_2$, then we assume that $\tau(\alpha_1) = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ and $\tau(\alpha_2) = \sa{\bar{Y}}\sa{\bar{y}}\psi(\bar{Y},\bar{y})$.
-The construction that we will provide for this function works by identifying a ``first'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ and a ``last'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$.
-These are identified by the following formulas:
-\begin{align*}
-\gamma_{\text{first}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z}), \\
-\gamma_{\text{last}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x}).
-\end{align*}
-Similarly, we define the formulas $\gamma_{\text{first}}(\bar{Y},\bar{y})$ and $\gamma_{\text{last}}(\bar{Y},\bar{y})$ (for the sake of simplicity we reuse the names $\gamma_{\text{first}}$ and $\gamma_{\text{last}}$).
-In other words, the ``first'' assignment is the one where every second-order predicate is empty and the first-order assignment is the lexicographically smallest, and the ``last'' assignment is the one where every second-order predicate is full and the first-order assignment is the lexicographically greatest.
-We also need to identify assignments that are not first and are not last.
-We do this by negating the two formulas above and grouping together the first-order variables:
-\begin{align*}
-\gamma_{\text{not first}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{z}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{z}_i)), \\
-\gamma_{\text{not last}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{x} < \bar{z}_0 \vee \bigvee_{i = 1}^{\length(\bar{X})}\neg X(\bar{z}_i)),
-\end{align*}
-where $\bar{z} = (\bar{z}_0,\bar{z}_1,\ldots,\bar{z}_{\length(\bar{X})})$. The following formula is equivalent to $\alpha$:
-\begin{align}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{not first}}(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{Y},\bar{y}))\vee \label{eq:partition1} \\
-&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))\vee \label{eq:partition2}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{not last}}(\bar{Y},\bar{y}))\vee \label{eq:partition3}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{last}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))]. \label{eq:partition4}
-\end{align}
-To show that the formula is indeed equivalent to $\alpha$, note that the formulas in lines (\ref{eq:partition1}) and (\ref{eq:partition2}) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$, and the formulas in lines (\ref{eq:partition3}) and (\ref{eq:partition4}) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
-Altogether, the four lines define pairwise disjoint assignments for $(\bar{X},\bar{x}),(\bar{Y},\bar{y})$.
-With this, it is straightforward to show that the above formula is equivalent to $\alpha$.
-However, the formula is not yet in the correct form since it has existential quantifiers in the subformulas $\gamma_{\text{not first}}$ and $\gamma_{\text{not last}}$.
-To solve this, first let take a close look to the complete formula:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\exists\bar{v}(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i))\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\exists\bar{w}(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))].
-\end{align*}
-To construct an equivalent formula that is in the correct form, we define $\bar{u} = (\bar{v},\bar{w})$ and we replace the first-order quantifiers by a first-sum and count the first assignment to $\bar{v}$ and $\bar{w}$ that satisfies the formula.
-A similar construction was used in \cite{SalujaST95}.
-Then the final formula equivalent to $\alpha$ is the following:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
-& \hspace{3cm}  \bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&
-(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))\vee
-\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
-& \hspace{3cm} (\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee \\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))].
-\end{align*}
-Finally, consider a $\eqso(\LL)$ formula $\alpha$ in $\LL$-SNF.
-If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we can consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the condition in the hypothesis.
-
-\subsection*{Proof of Theorem \ref{prop-rest}}
-We give this proof in three parts.
-
-\vspace{1em}
-First, we show that $\QE{0} \not\subseteq \E{1}$.
-By contradiction, suppose that there is a $\QE{0}$ formula $\alpha$ over some signature $\R$ such that defines the following function.
-For every finite $\R$-structure with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n - 1$.
-We use the following claim.
-\begin{claim}
-Let $\alpha = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. Then the function defined by $\alpha$ is either null, greater or equal to $n$, or is in $\Omega(n^2)$.
-\end{claim}
-\begin{proof}
-Suppose that the function defined by $\alpha$ is not $0$ and that $\varphi$ is in DNF.
-Furthermore, suppose $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$.
-Then $\alpha = \sa{\bar{x}} \varphi_1(\bar{x}) \vee \cdots \vee \varphi_n(\bar{x})$.
-Since $\alpha$ is not null, then some $\varphi_i$ must be satisfiable.
-This is, the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is not null.
-We will prove by induction on $\length(\bar{x})$ that the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is either greater or equal to $n$, or in $\Omega(n^2)$.
-We address the case $\length(\bar{x})= 1$, then $\alpha = \sa{x}\bigwedge\psi(x)$.
-If any $\psi(x) = (x = x)$ or $\neg(x < x)$, then we can eliminate it and we obtain the same function.
-If any $\psi = (x < x)$ or $\neg(x=x)$, then the function becomes null.
-If $\psi(x) = R(x,\ldots,x)$ for some $R\in\R$ the function becomes null for the structures we are considering.
-If $\psi(x) = \neg R(x,\ldots,x)$, we can eliminate it and for the structures we are considering we obtain the same function.
-The only possible $\alpha$ left is $\alpha = \sa{x}\top$ which is equal to the function $n$.
-This covers all possible cases for $\length(\bar{x}) = 1$.
-Now suppose that it holds for $\length(\bar{x}) = k$ and suppose $\alpha = \sa{\bar{x}}\bigwedge\psi(\bar{x})$ for $\length(\bar{x}) = k+1$.
-If any $\psi(\bar{x}) = (x_i = x_j)$ where $i \neq j$, then $\alpha$ describes the same function as $\alpha$ where $x_j$ has been replaced by $x_i$.
-In this formula the tuple of first-order variables has $k$ elements so the function it describes if one of the mentioned in the hypothesis.
-If $i = j$, then we can eliminate it and obtain the same function.
-If any $\psi(\bar{x}) = R(\bar{v})$ or $\neg R(\bar{v})$ where $\bar{v}$ is a sub-tuple of $\bar{x}$ then we can either eliminate it or the function becomes null, following the same argument as in the case $\length(\bar{x}) = 1$.
-If any $\psi(\bar{x}) = \neg(x_i = x_j)$ or $(x_i < x_j)$ where $i = j$, then the function becomes null.
-If any $\psi(\bar{x}) = \neg(x_i < x_j)$ where $i = j$, we can eliminate it.
-The remaining formulas in $\bigwedge\psi(\bar{x})$ are either $\neg(x_i = x_j)$, $(x_i<x_j)$ or $\neg(x_i<x_j)$.
-If the formula violates transitivity in $<$ (for example, $x < y \wedge y < z \wedge z < x$), then the function $\alpha$ describes is null.
-Therefore, there is some order over $\bar{x}$ that satisfies $\bigwedge\psi(\bar{x})$.
-Consider the formula that describes this order (like $x_1 < x_3 \wedge x_3 < x_4 \wedge x_4 < x_2$).
-The function $\alpha$ describes is greater or equal to the one this formula describes, which is exactly $\binom{n}{\length(\bar{x})}$ which is in $\Omega(n^{\length(\bar{x})}) \subseteq \Omega(n^2)$ if $\length(\bar{x}) > 1$.
-This concludes the proof of the claim.
-\end{proof}
-We suppose that $\alpha$ is in SNF, this is, $\alpha = \sum_{i = 1}^n\alpha_i$.
-Since $\alpha$ is not null, consider some $\alpha_i$ that describes a non-null function.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, where $\varphi$ is quantifier-free.
-Note that if $\length(\bar{X}) > 0$, then the function $\alpha$ describes is in $\Omega(2^n)$, as it was proven by the authors in \cite{SalujaST95}.
-We have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$, as we proved in the claim, describes either some function greater or equal to $n$, or in $\Omega(n^2)$, which leads to a contradiction.
-Lastly, note that the formula $\sa{x}\exists y(x < y)$ is in $\E{0}$ and describes the function $n-1$, which concludes the proof.
-
-\vspace{1em}
-Now we show that $\E{1}\not\subseteq\QE{0}$.
-In Theorem \ref{theo-pi1-pnf} we proved that there is no formula in $\loge{1}$-PNF equivalent to the formula $\alpha = 2$.
-Every formula in $\E{1}$ can be expressed in $\loge{1}$-PNF, which implies that $2 \in \QE{0}$ and $2 \not\in \E{1}$.
-
-\vspace{1em}
-Lastly, we prove that $\eqso(\loge{1})\subsetneq\eqso(\logu{1})$.
-For inclusion, let $\alpha$ be a formula in $\eqso(\loge{1})$.
-Suppose that it is in $\loge{1}$-SNF.
-This is, $\alpha = c + \sum_{i = 1}^{n}\alpha_i$.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, where $\varphi_i$ is quantifier-free, for each $\alpha_i$.
-We use the same construction used in \cite{SalujaST95}, and we obtain that the formula $\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \forall\bar{y}'(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')$ for every assignment to $(\bar{X},\bar{x})$.
-We do this replacement for each $\alpha_i$ and we obtain an equivalent formula in $\eqso(\logu{1})$.
-
-To prove that the inclusion is proper, consider the $\eqso(\logu{1})$ formula $\sa{x}\forall y(y = x)$. This formula defines the following function that takes an ordered structure $\A$ as input:
-$$
-\sem{\alpha}(\A) = 
-\begin{cases}
-1 &\A \text{ has one element}\\
-0 &\text{ otherwise}.
-\end{cases}
-$$
-Suppose that there exists an equivalent formula $\alpha$ in $\eqso(\loge{1})$.
-Also, suppose that it is in $\L$-PNF, so $\alpha = c + \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$.
-Since $\alpha$ takes the value 0 for some structures, $c$ must be 0.
-Consider a structure $\mathfrak{1}$ with one element.
-We have that for some $i$, there exists an assignment $(\bar{B},\bar{b},\bar{a})$ for $(\bar{X},\bar{x},\bar{y})$ such that $\mathfrak{1}\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
-Consider now the structure $\mathfrak{2}$ that is obtained by duplicating $\mathfrak{1}$, as we did for Theorem \ref{theo-pi1-pnf}.
-Note that $\mathfrak{2}\models\varphi_i(\bar{B},\bar{b},\bar{a})$, which implies that $\sem{\alpha}(\mathfrak{2}) \geq 1$, which leads to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop-e1-nc}}
-
-Towards a contradiction, assume that the statement is false.
-This is, that $\E{1}$ is closed under binary sum.
-Consider the formula $\sa{x}(x = x)$ which is in $\E{1}$ over some signature $\R$.
-For every finite $\R$-structure $\A$ with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n$.
-From our assumption, there exists some formula in $\E{1}$ equivalent to the formula $\alpha \add \alpha$, which describes the function $2n$.
-Let $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ be this formula, where $\varphi$ is quantifier-free.
-Note that the function defined by this formula is equal or greater than the one defined by $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ divided by a polynomial factor.
-More specifically, for each ordered structure $\A$ with domain $A$, we have the following inequality:
-$$
-\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot \vert A \vert^{\length(\bar{y})} \geq \sem{\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A)
-$$
-Note that the formula $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ defines a function in $\E{0}$.
-It was shown by the authors in \cite{SalujaST95} that every function in $\E{0}$ grows exponentially over the size of the structure for large enough structures, when $\length(\bar{X}) > 0$.
-This function divided by a polynomial factor still grows exponentially.
-Therefore, for $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ we have that $\length(\bar{X}) = 0$.
-
-Now, for the formula $\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ consider a structure $\mathfrak{1}$ with only one element $a$.
-We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$, but the only possible assignment to $\bar{x}$ is the tuple $(a,\ldots,a)$ so $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which follows to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop:qe0-fp-qe1-totp-fptras}}
-
-The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$.
-This is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every finite $\R$-structure $\A$ with domain $A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$.
-We base our proof on this fact.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is contained in \textsc{TotP}.} Let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Consider the following nondeterministic procedure that on input $\enc(\A)$ generates $\sem{\alpha}(\A)$ branches.
-For each $\alpha_i = \varphi$, where $\varphi$ is a $\loge{1}$ formula, it checks if $\A\models\varphi$ in polynomial time and generates a new branch if that is the case.
-For each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi$, this formula is also in $\E{1}$.
-We use the reduction to $\cdnf$ provided in \cite{SalujaST95} and we obtain $g(\enc(\A))$, which is an instance to $\cdnf$.
-Since $\cdnf$ is also in $\totp$ \cite{PagourtzisZ06}, we simulate the corresponding nondeterministic procedure that generates exactly $\cdnf(\enc(g(\A)))$ branches.
-Since, $\fp\subseteq\totp$\cite{PagourtzisZ06}, each polynomially computable function is also in $\totp$, and then on each of these branches we simulate the corresponding nondeterministic procedure to generate $h(\vert A \vert)$ more.
-The number of branches for each $\alpha_i$ is $\sem{\alpha_i}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$, and the total number of branches in the procedure amounts to $\sem{\alpha}(\A)$.
-We conclude that $\alpha\in\totp$.
-
-\vspace{1em}
-{\bf Every function in} $\boldsymbol{\eqso(\loge{1})}$ {\bf has an FPRAS.}  let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Note that each $\alpha_i$ that is equal to some $\loge{1}$ formula $\varphi$ has an FPRAS given by the procedure that simply checks if $\A\models\varphi$ and returns 1 if it does and 0 otherwise.
-Also, each remaining $\alpha_i$ has an FPRAS since $\alpha_i\in \E{1}$ \cite{SalujaST95}.
-If two functions have an FPRAS, then their sum also has one given by the procedure that simulates them both and sums the results.
-We conclude that $\alpha$ has an FPRAS.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is closed under sum and multiplication.} Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that the class is closed under multiplication.
-We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of $\so$.
-
-\begin{lemma} \label{conj-mult}
-If $\LL$ is a fragment closed under conjunction, then $\eqso(\LL)$ is closed under binary multiplication.
-\end{lemma}
-\begin{proof}
-We will define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ extended by binary product, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
-In fact, the formula $\tau(\alpha)$ is in $\LL$-SNF.
-First we replace each constant $s$ in $\alpha$ for $(\top \add \cdots \add \top)$ ($s$ times).
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$.
-We assume that for every $\beta$ that has less algebraic operators than $\alpha$, $\tau(\beta)$ is in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$ then we define $\tau(\alpha) = \tau(\alpha_1) + \tau(\alpha_2)$.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$, then we define $\tau(\alpha)$ as the formula in $\LL$-SNF that is equivalent to $\sa{x}\tau(\beta)$ and to $\tau(\alpha) = \sa{X}\tau(\beta)$, respectively.
-If $\alpha = (\alpha_1 \cdot \alpha_2)$, we assume that each $\alpha_i$ is in $\LL$-SNF.
-We identify three cases.
-(1) Some $\alpha_i$ is equal to $\sum_{j = 1}^n\beta_j$ for $n > 1$.
-Suppose wlog.
-that it is $\alpha_1$.
-We then define $\tau(\alpha) = \sum_{j = 1}^n\tau(\beta_j\cdot\alpha_2)$.
-In the following cases, $\alpha_1$ and $\alpha_2$ are in $\LL$-SNF.
-(2) If some $\alpha_i$ is equal to $\sa{X}\beta$ or $\sa{x}\beta$, we define $\tau(\alpha)$ as the $\LL$-SNF formula that is equivalent to $\sa{x}\tau(\beta\cdot\alpha_2)$ and $\sa{X}\tau(\beta\cdot\alpha_2)$, respectively.
-The remaining case is (3) $\alpha_1 = \varphi_1$ and $\alpha_2 = \varphi_2$ where each $\varphi$ is an $\LL$ formula.
-Then we define $\tau(\alpha) = \varphi_1 \wedge \varphi_2$.
-This covers all possible cases for $\alpha$.
-For every pair of formulas $\alpha,\beta$ in $\eqso(\LL)$, we have that their multiplication $(\alpha\cdot\beta)$ is a formula in the grammar $\eqso(\LL)$ extended by binary product, and so, there exists an equivalent formula $\tau(\alpha\cdot\beta)$ which is in unextended $\eqso(\LL)$.
-\end{proof}
-Since $\loge{1}$ is closed under conjunction, this also holds for $\eqso(\loge{1})$. This concludes the proof.
-
-\subsection*{Proof of Proposition \ref{pi-minusone}}
-
 Let $\LL$ be a fragment of $\fo$ that contains $\logu{1}$.
@@
 We have that this is a $\shp$ procedure that computes $f-1$, from which we conclude that $\shp$ is closed under subtraction by one.
-
-\subsection*{Proof of Theorem \ref{sigmafo-minusone}}
-
-\noindent {\bf Closed under sum and multiplication.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
-This can be seen because $\eqso(\LL)$ is closed under sum for every fragment $\LL$ by definition, and since $\logex{1}$ is closed under conjunction, from Lemma \ref{conj-mult} it follows that $\eqso(\logex{1})$ is closed under multiplication.
-
-\vspace{1em}
-\noindent {\bf Easy decision version and FRPRAS.}  We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
-We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Theorem \ref{prop:qe0-fp-qe1-totp-fptras}.
-First, we define a function that converts a formula $\alpha$ in $\eqso(\logex{1})$ over a signature $\R$ into a formula $\lambda(\alpha)$ in $\eqso(\loge{1})$ over a signature $\R_{\alpha}$.
-Afterwards, we define a function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an $\R_{\alpha}$-structure $g_{\alpha}(\A)$.
-
-Let $\alpha$ be in $\eqso(\logex{0})$.
-The signature $\R_{\alpha}$ is obtained by adding the symbol $R_{\psi}$, for every $\fo$ formula $\psi(\bar{z})$ in $\alpha$, to $\R$.
-Each symbol $R_{\psi}$ represents a predicate with arity $\length(\bar{z})$.
-Then, $\lambda(\alpha)$ is defined as $\alpha$ where each $\fo$ formula $\psi(\bar{z})$ has been replaced by $R_{\psi}(\bar{z})$.
-We now define the function $g_{\alpha}$ procedurally.
-Let $\A$ be a $\R$-structure with domain $A$.
-Let $\A'$ be an $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A}$ empty.
-For each $\fo$-formula $\psi(\z)$ with $\length(\bar{z})$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length(\bar{z})}$.
-If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$ (this can be done in $\ptime$).
-This concludes the construction of $\A'$.
-Note that the number of $\fo$ subformulas, arity and tuple size is fixed in $\alpha$, so computing this function takes polynomial time over the size of the structure.
-Moreover, the encoding of $\A'$ has polynomial size over the size of $\enc(\A)$.
-We define $g_{\alpha}(\A) = \A'$ and we have that for each $\R$-structure $\A$: $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$.
-Therefore, we have a parsimonious reduction from $\alpha$ to the $\eqso(\loge{1})$ formula $\lambda(\alpha)$.
-
-To show that $\alpha$ is in $\totp$, we can convert $\alpha$ and $\enc(\A)$ into $\lambda(\alpha)$ and $\enc(g_{\alpha}(\A))$, respectively, and run the procedure in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-Similarly, to show that $\alpha$ has an FPRAS, we do the same as before and simulates the FPRAS for $\lambda(\alpha)$ in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-These procedures also takes polynomial time and satisfies the condition.
-
-\vspace{1em}
-\noindent {\bf Closed under subtraction by one.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. 
-For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Without lost of generality, we assume that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
-Moreover, we assume that $\length(\bar{x}) > 0$  since, if this is not the case, we can replace $\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
-
-The proof will be separated in two parts.
-In the first part, we suppose that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$.
-Then we will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ is the formula.
-In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ the sum of one or more formulas in $\logex{1}$-PNF.
-We define a formula $\varphi'$ that satisfies the following condition: if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ and $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be proven.
-
-\vspace{1em}
-
-\noindent {\em Part (1).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ where $\varphi$ is an  $\fo$-formula.
-Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}} \varphi(\bar{x})$ (i.e.
-$\length(\bar{X}) = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}} [\varphi(\bar{x})\wedge\exists\bar{z}(\varphi(\bar{z})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
-Therefore, for the rest of the proof we can assume that $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.
-
- Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
-To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in into a DNF formula. 
-More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i$ is a conjunction of atomic formulas or negation of atomic formulas.
-Furthermore, we suppose that each conjunct $\varphi_i$ has the~form:
-$$
-\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
-\underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}} \wedge
-\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}}.
-$$
-Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ can appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
-
-Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ are mentioned in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
-We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
-\begin{itemize}
-	\item[(a)] In order to satisfy the first condition, consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
-We add $\length(\bar{w})$ new variables $z_1,\ldots,z_{\length(\bar{w})}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length(\bar{w})})$.
-We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$.
-The formula $\varphi_i$ is now defined as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \exists \bar{z} \; (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{-}(\bar{X},\bar{x},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y})).
-	$$
-	We repeat this process for each instance of a $X_j(\bar{w})$ in $\varphi_i$, and we obtain a formula where none of the $X_j$'s acts over any variable in $\bar{x}$.
-We add the new first-order variables to $\bar{y}$ and we redefine $\varphi_i$ as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
-	$$
-	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = \exists \bar{y} \;  X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := \exists \bar{y} \;  X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
-	\item[(b)] An ordered partition on a set $S$ is defined by an equivalence relation $\sim$ over $S$, and a linear order over $S/\!\sim$.
-For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$.
-A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
-On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
-	For a given $k$, let $\cB_k$ be the number of possible ordered partitions for a set of size $k$. For $1 \leq j \leq \cB_{\length(\bar{x},\bar{y})}$ 
-	let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
-
-	Thus, the formula $\varphi_i(\bar{X},\bar{x})$ is then redefined as:
-	$$
-	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}} \exists \bar{y} \; [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
-	$$
-	Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
-	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  to $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
-	\item[(c)] For showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
-$\varphi_i$ is inconsistent), we do the following.
-If there exists an instance of $X_j(\bar{z})$ in $\varphi^{+}_i$, an instance of $\neg X_j(\bar{w})$ in $\varphi^{-}_i$ and the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$, then the entire formula $\varphi_i$ is removed from $\varphi$.
-\end{itemize}
-It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$.
-From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c).
-Furthermore, it has the following form:
-$$
-\varphi_i(\bar{X},\bar{x}) \; = \; \exists{\bar{y}} \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
-$$
-where $\varphi^{+}$ and $\varphi^{-}_i$ do not depend on  $\bar{x}$.
-\begin{claim}\label{claim:minusone}
-For a given ordered structure $\A$, if there exists a first-order assignment $v$ for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists a second-order assignment $V$ for $\A$ such that  $(\A,V,v)\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-\end{claim}
-\begin{proof}
-Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
-Define $\bar{B} = (B_1,\ldots,B_{\length(\bar{X})})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
-Towards a contradiction, suppose that $(\A,V,v)\not\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-By the choice of $v$, and construction of $V$ it is clear that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})\wedge\varphi^{+}_i(\bar{X},\bar{y})$, so we necessarily have that $(\A,V,v)\not\models\varphi^{-}_i(\bar{X},\bar{y})$.
-Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,V, v)\not\models\neg X_t(\bar{w})$, namely, $v(\bar{w})\in B_t$. 
-However, by the construction of $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$.
-Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$.
-This violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
-\end{proof}
-
-The previous claim and proof motivates the following definitions.
-For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length(\bar{X})})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
-One can easily check that for every assignments $V, v$ such that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$, then it holds that $(\A,\bar{B}^v,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all the assignment of $\bar{X}$ when $v$ is fixed.
-Therefore, by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
-For this, the idea is to choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
-$$
-\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge\forall\bar{x}'\forall\bar{y}'(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')).
-$$ 
-If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
-Furthermore, let $V^*$ the second order assignment that satisfies $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*$ the first order assignment such that $v^*(\bar{x}) = v(\bar{x})$.
-By the previous discussion, we have that $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.
-
-Now, we have all the tools to introduce the main formula of the proof. 
-Intuitively, we want to exclude the assignment $(V^*, v^*)$ as a satisfying assignment for $\varphi_i(\bar{X},\bar{x})$.
-Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if, 
-if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, then $V \neq V^*$ or $v \neq v^*$. 
-That is, it excludes all pairs $(V, v)$ that coincide with $(V^*, v^*)$ in $(\bar{X}, \bar{x})$ whenever $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. This property can be defined as follows:
-\begin{align}
-\psi_i(\bar{X},\bar{x}) = & \;\; \big( \, \exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
-& \hspace{1cm} \Big( \; 
-\exists\bar{y} \;\; \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i^+(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\
-& \hspace{1.5cm} (\exists\bar{x}'\exists\bar{y}\, \varphi_i(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}) \;\; \Big)  \label{eq:part3}
-\end{align}
-To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
-Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-Then, the conclusion of the implication at (\ref{eq:part2}) and (\ref{eq:part3}), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
-Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by showing that if $\varphi_i(\bar{X},\bar{x})$ is satisfied then $V(\bar{X})$ must contain one element more that it is not in $V^*(\bar{X})$.
-The second disjunct (\ref{eq:part3}) is satisfied when $v(\bar{x})$ is not the lexicographically smallest tuple that satisfies $\varphi_i$ (i.e. $v(\bar{x}) \neq v^*(\bar{x})$).
-
-We are ready to define the formula  $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$
-where each modified disjunct $\varphi_i'(\bar{X},\bar{x})$ is constructed as follows. 
-For the sake of simplification, define the auxiliary formula $\chi_i = \neg\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$. 
-This formula basically checks if $\varphi_i$ is satisfiable or not.
-Then, first define the formula:
-$$
-\varphi'_1(\bar{X},\bar{x}) = \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
-$$
-This models that all the assignments that satisfy $\varphi_1$ also satisfy $\varphi'_1$, except for $(V^*,v^*)$. Then we define:
-$$
-\varphi'_2(\bar{X},\bar{x}) = \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
-$$
-This models that all the assignments that satisfy $\varphi_2$ also satisfy $\varphi'_2$, but if there was no assignment for $\varphi_1$, then we exclude $(V^*,v^*)$ for $\varphi_2$.
-Then we generalize this construction for each $\varphi_i$:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})),
-$$
-By the construction of $\kappa(\alpha)$, one can easily check that  $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.
-
-\vspace{1em}
-
-{\em Part (2).} Let $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ for some $\eqso(\logex{1})$ formula $\beta$. We define $\kappa(\alpha)$ as follows.
-First, perform the same transformations to $\varphi(\bar{X},\bar{x})$ as in the previous section.
-Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c).
-We also use the previously defined formulas $\chi_i$ and $\psi_i$, for each $i$.
-
-We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if and only if $\sem{\beta}(\A) = 0$.
-If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg\exists\bar{x}'\varphi(\bar{x}')$.
-If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$ as each $\chi_i$ was previously defined.
-If $\beta = (\beta_1 + \beta_2)$, then $\lambda(\beta) = \lambda(\beta_1) \wedge \lambda(\beta_2)$.
-Now, for each $\varphi_i$ we define:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x}))\Big)\Big).
-$$
-Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
-
-%\subsection*{Proof of Theorem \ref{sub-pnp}}
-%
-%Towards a contradiction, suppose that any of the classes $\E{1}$, $\eqso(\loge{1})$, and $\eqso(\logex{1})$ is closed under subtraction.
-%
-%Let $\R$ include the symbols $S_1, S_2, S_3, S_4$, which describe the following properties.
-If a finite $\R$-structure $\A$ defines a 3DNF formula $\Phi$, then its domain is the set of variables mentioned in $\Phi$, and for each $i = 1,2,3,4$:
-%\begin{align*}
-%	S_1^\A &= \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_2^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_3^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_4^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
-%\end{align*}
-%Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
-%\begin{multline*}
-%\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
-%\end{multline*}
-%Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
-%$$
-%\varphi(x,X) = (T(x) \vee \neg T(x)).
-%$$
-%Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Given that $f_{\#3DNF}, f_{all} \in \E{1} \subseteq \eqso(\loge{1}) \subseteq \eqso(\logex{1}) \subseteq \totp$, and at least one of the classes is closed under subtraction, the function $f_{\#3DNF} - f_{all}$ is in $\totp$.
-However, note that for each $\R$-structure $\A$ that represents a formula $\Phi$, $(f_{\#3DNF} - f_{all})(\A) = 0$ if an only if $\Phi$ is a tautology.
-The decision version of this function is the $\np$-complete problem $\textsc{Tautology}$.
-And since the function is in $\totp$, its decision version is also in $\ptime$.
-We conclude that $\np \subseteq \ptime$.
-
-\subsection*{Proof of Proposition \ref{prop:ehorn-pe}}
-Pagourtzis and Zachos mention a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula \cite{PagourtzisZ06}.
-This procedure can be easily extended to receive Horn formulas, and furthermore, a disjunction of Horn formulas.
-Hence $\shdhsat$ is in $\totp$.
-
-As we show in Proposition \ref{sigma2hard}, $\shdhsat$ is complete for $\eqso(\ehorn)$ under parsimonious reductions.
-Let $\alpha$ be a formula in $\eqso(\ehorn)$ and let $g_{\alpha}$ be the reduction to $\shdhsat$.
-The $\totp$ procedure we construct, for each input $\enc(\A)$, is simply to compute $g_{\alpha}(\enc(\A))$, and then simulate the $\totp$ procedure for $\shdhsat$ on input $g_{\alpha}(\enc(\A))$.
-We conclude that $\alpha$ is in $\totp$.
-
-\subsection*{Proof of Proposition \ref{prop:hsat-not-sigma2}} %V.12
-We use a similar proof to the one provided by the authors in \cite{SalujaST95} to separate the classes $\E{2}$ and $\U{2}$.
-Suppose that the statement is false, this is, $\chsat \in \eqso(\loge{2})$.
-We consider the signature $\R$ that we used as the encoding for a Horn formula (Example \ref{ex-hornsat-esop1}) and that the formula $\alpha \in \eqso(\loge{2})$ follows the encoding in the same way.
-From what we proved in Theorem \ref{theo-pnf-snf}, we have that every formula in $\eqso(\loge{2})$ can be rewritten in $\loge{2}$-PNF, so we assume that $\alpha$ is in this form.
-Let $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\forall\bar{z}\,\varphi(\bar{X},\bar{x},\bar{y},\bar{z})$.
-Consider the following Horn formula $\Phi$:
-$$
-\Phi = p \wedge \bigwedge_{i = 1}^n (t_i \wedge p \to q) \wedge \neg q,
-$$
-where $n = \length(\bar{x}) + \length(\bar{y}) + 1$.
-Let $\A$ be the encoding of this formula.
-In the encoding, each variable appears as an element in the domain of $\A$.
-This formula is satisfiable, so $\sem{\alpha}(\A) \geq 1$.
-Let $(\bar{B},\bar{b},\bar{a})$ be an assignment to $(\bar{X},\bar{x},\bar{y})$ such that $\A\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-Let $t_{\ell}$ be such that it does not appear in $\bar{b}$ or $\bar{a}$.
-Consider the induced substructure $\A'$ that is obtained by removing $t_{\ell}$ from $\A$ and $\bar{B}'$ as the subset of $\bar{B}$ obtained by deleting each appearance of $t_{\ell}$ in $\bar{B}$.
-We have that $\A'\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-This is because each subformula of the form $\exists y \neg B_i$ is still true, and universal formulas are monotone over induced substructures.
-It follows that $\sem{\alpha}(\A') \geq 1$ which is not possible since $\A'$ encodes the formula
-$$
-\Phi' = p \wedge \bigwedge_{i = 1}^{\ell-1} (t_i \wedge p \to q) \wedge (p\to q) \wedge \bigwedge_{i = \ell+1}^{n} (t_i \wedge p \to q) \wedge \neg q,
-$$
-which is unsatisfiable. We arrive to a contradiction and we conclude that $\chsat$ is not in $\eqso(\ehorn)$.
-
-\subsection*{Proof of Theorem \ref{sigma2hard}} %V.13
-
-First we prove that $\shdhsat$ is in $\eqso(\ehorn)$.
-Recall that each instance of $\shdhsat$ is a disjunction of Horn formulas.
-Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot), \pD(\cdot,\cdot)\}$.
-Each symbol in this vocabulary is used to indicate the same as in Example \ref{ex-hornsat-esop1}, with the addition of $\pD(d,c)$ which indicates that $c$ is a clause in the formula $d$.
-Recall that the formula
-\begin{align*}
-&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \neg\textit{T}(x)).
-\end{align*}
-defines $\chsat$. We obtain the following formula $\psi(T,A)$ in $\ehorn$:
-\begin{align*}
-\exists d[&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \pD(c,d)\vee \neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \neg\textit{T}(x))]
-\end{align*}
-effectively defines $\chsat$ as for every disjunction of Horn formulas $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
-Therefore, we conclude that $\shdhsat \in \eqso(\ehorn)$.
-
-\vspace{1em}
-We will now prove that $\shdhsat$ is hard for $\eqso$ over a signature $\R$ under parsimonious reductions.
-For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a polynomial-time procedure that computes a function $g_{\alpha}$.
-This function receives a finite $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
-We suppose that $\alpha$ is in sum normal form and:
-$$
-\alpha = \sum_{i = 1}^{\text{\#clauses}} \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigwedge_{j = 1}^{n}\forall\bar{z}\,\varphi^i_j(\bar{X},\bar{x},\bar{y},\bar{z}),
-$$
-where each $\varphi^i_j$ is a Horn clause.
-
-Consider a finite $\R$-structure $\A$ with domain $A$.
-To simplify the proof, we extend our grammar to allow first-order constants.
-Consider each tuple $\bar{a}\in A^{\length(\bar{x})}$, each $\bar{b}\in A^{\length(\bar{y})}$ and each $\bar{c}\in A^{\length(\bar{z})}$ as a tuple of first-order constants.
-The following formula defines the same function as $\alpha$:
-$$
-\sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length(\bar{x})}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length(\bar{y})}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length(\bar{z})}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
-$$
-Note that each $\fo$ formula over $(\bar{x},\bar{y},\bar{z})$ in each $\varphi^i_j$ has no free variables.
-Therefore, we can evaluate each of these in polynomial time and replace them by $\perp$ and $\top$ where it corresponds.
-Each $\varphi^i_j$ will be of the form $\perp \vee\, \chi^i_j(\bar{X})$ or $\top \vee \chi^i_j(\bar{X})$ where $\chi^i_j$ is a disjunction of $\neg X_{\ell}$'s and at most one $X_{\ell}$.
-The formulas of the form $\top \vee \chi^i_j(\bar{X})$ can be removed entirely, and the formulas of the form $\perp \vee\, \chi^i_j(\bar{X})$ can be replaced by $\chi^i_j(\bar{X})$.
-We obtain the formula
-$$
-\sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})
-$$
-where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$.
-
-Our idea for the rest of the proof is to define $g_{\alpha}$ for each $\alpha = \sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})$, for $\alpha = c$ and for $\alpha = \beta_1 + \cdots + \beta_m$ where each $\beta_i$ is in one of the two previous cases.
-
-If $\alpha$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ where $\psi_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$, then we obtain the {\bf propositional formula} $g_{\alpha}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulas.
-
-If $\alpha$ is equal to a constant $c$, then we define $g_{\alpha}(\A)$ as the following formula that has exactly $c$ satisfying assignments:
-$$
-g_{\alpha}(\A) = \bigvee_{i = 1}^{c}\neg t_1 \wedge \cdots \wedge \neg t_{i-1} \wedge t_i \wedge \neg t_{i+1} \wedge \cdots \wedge \neg p_c.
-$$ 
-If $\alpha = \beta_1 + \cdots + \beta_m$, let $g_{\beta_i}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\theta^i_{j,k}$ for each $\beta_i$ where each $\theta^i_{j,k}$ is a Horn clause.
-Let $\Theta_i = g_{\beta_i}(\A)$.
-We rename the variables in each $\Theta_i$ so none of them are mentioned in any other $\Theta_j$.
-We add $m$ new variables $t_1,\ldots,t_m$ and we define:
-\begin{align*}
-g_{\alpha}(\A) = &\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^1_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m}}}t) \wedge (t_1 \wedge \bigwedge_{\ell = 2}^{m} \neg t_{\ell})) \vee \\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^2_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each $t$}\\ \text{ mentioned in}\\ \Theta_1,\Theta_3,\ldots,\Theta_{m}}}t) \wedge (t_2 \wedge \bigwedge\limits_{\substack{\ell = 1 \\ \ell \neq 2}}^{m} \neg t_{\ell})) \vee \cdots \vee\\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^m_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m-1}}}t) \wedge (t_m \wedge \bigwedge_{\ell = 1}^{m-1} \neg t_{\ell})).
-\end{align*}
-The formula is a disjunction of Horn formulas, and the number of satisfying assignments for this formula is exactly the sum of satisfying assignments for each $g_{\beta_i}(\A)$.
-This, at the same time, is equal to $\sem{\alpha}(\A)$.
-This covers all possible cases for $\alpha$, and the entire procedure takes polynomial time.

### 2017-03-03 LMCS/proof-prop-e1-nc.tex
journal primera version

--- 2017-01-12 LICS/appendix-syntactic.tex
@@
-\subsection*{Proof of Theorem \ref{theo-pnf-snf}}
-Recall that a formula in $\eqso(\LL)$ is on the following grammar:
-$$
-\alpha = \varphi \ \mid \ s \ \mid \ (\alpha + \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \sa{X} \alpha,
-$$
-where $\varphi$ is a formula in $\LL$ and $s\in\nat$.
-We will construct a recursive function $\tau$ such that for every $\eqso(\LL)$ formula $\alpha$, it outputs an equivalent formula $\tau(\alpha)$ which is in $\LL$-SNF.
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$ which is clearly equivalent and already in $\LL$-SNF.
-If $\alpha = s$, then we define $(\top \add \cdots \add \top)$ ($s$ times), which also satisfies the condition.
-We assume that for every sub-formula $\beta$ of $\alpha$, $\tau(\beta)$ is an equivalent formula in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$, then we define $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$.
-If $\alpha = \sa{x}\beta$, then let $\tau(\beta) = \sum_{i = 1}^{k}\beta_i$ for some $k$ where each $\beta_i$ is in $\LL$-PNF.
-By grouping together the terms in the sum, we define $\tau(\alpha) = \sum_{i = 1}^{m}\sa{x}\beta_i$ which satisfies the condition and is equivalent to $\alpha$.
-If $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case.
-This covers all possible cases for $\alpha$ and we conclude the proof by taking $\tau(\alpha)$ as the desired rewrite of $\alpha$.
-
-\subsection*{Proof of theorem \ref{theo-pi1-pnf}}
-We divide the proof in three parts.
-
-\vspace{1em}
-First, we prove that the formula $\alpha_{0} = \left( \sa{X} 1 \right) + 1$ with $\arity(X) = 1$ (i.e.
-the function $2^{n}+1$, where $n$ is the size of the input structure) is not equivalent to any formula in $\loge{0}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ in $\loge{0}$-PNF that is equivalent to the $\eqso(\loge{0})$ formula $\alpha_0$.
-In \cite{SalujaST95}, it was proved that if $\length{\bar{X}} > 0$, then the function defined by $\alpha$, for big enough structures, is always even which is not possible in our case.
-On the other hand, if $\alpha$ is of the form $\sa{\bar{x}}\varphi(\bar{x})$, then $\alpha$ defines a polynomially bounded function which also leads to a contradiction.
-
-\vspace{1em}
-Second, we prove that the formula $\alpha_{1} = 2$ (i.e.
-the constant $2$) is not equivalent to any formula in $\loge{1}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to the $\eqso(\loge{1})$ formula $2$.
-First, if $\length{\bar{X}} = \length{\bar{x}} = 0$, then the function defined by $\alpha$ is never greater than 1. 
-Therefore, suppose that $\length{\bar{X}} > 0$ or $\length{\bar{x}} > 0$, and consider any ordered structure $\A$. 
-Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
-Now consider the ordered structure $\A'$ that is obtained by duplicating $\A$.
-This is, each half of $\A'$ is isomorphic to $\A$.
-Note that $\A'\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$ for $i = 1,2$ and there exists a third assignments $(\bar{B}_1',\bar{b}_1',\bar{a}_1')$ that is isomorphic to $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$ but in the other half to the structure such that $\A'\models\varphi(\bar{B}_1',\bar{b}_1',\bar{a}_1')$.
-We have that $\sem{\alpha}(\A) \geq 3$ which is a contradiction.
-
-\vspace{1em}
-We now show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there is an equivalent formula $\beta$ in $\LL$-PNF.
-As in Theorem \ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such formula.
-As we showed, there exists an equivalent formula in $\LL$-SNF, so we assume that $\alpha$ is in that form.
-Let $\alpha = \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-SNF.
-Without lost of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
-If not, we replace each $\alpha_i$ for the equivalent formula $\sa{\bar{X}} \sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge\forall z\,Y(z) \wedge \forall z(y \leq z))$.
-
-Now we begin describing the function $\tau$.
-If $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then the formula is already in $\LL$-PNF so we define $\tau(\alpha) = \alpha$.
-If $\alpha = \alpha_1 + \alpha_2$, then we assume that $\tau(\alpha_1) = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ and $\tau(\alpha_2) = \sa{\bar{Y}}\sa{\bar{y}}\psi(\bar{Y},\bar{y})$.
-The construction that we will provide for this function works by identifying a ``first'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ and a ``last'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$.
-These are identified by the following formulas:
-\begin{align*}
-\gamma_{\text{first}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z}), \\
-\gamma_{\text{last}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x}).
-\end{align*}
-Similarly, we define the formulas $\gamma_{\text{first}}(\bar{Y},\bar{y})$ and $\gamma_{\text{last}}(\bar{Y},\bar{y})$ (for the sake of simplicity we reuse the names $\gamma_{\text{first}}$ and $\gamma_{\text{last}}$).
-In other words, the ``first'' assignment is the one where every second-order predicate is empty and the first-order assignment is the lexicographically smallest, and the ``last'' assignment is the one where every second-order predicate is full and the first-order assignment is the lexicographically greatest.
-We also need to identify assignments that are not first and are not last.
-We do this by negating the two formulas above and grouping together the first-order variables:
-\begin{align*}
-\gamma_{\text{not first}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{z}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{z}_i)), \\
-\gamma_{\text{not last}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{x} < \bar{z}_0 \vee \bigvee_{i = 1}^{\length{\bar{X}}}\neg X(\bar{z}_i)),
-\end{align*}
-where $\bar{z} = (\bar{z}_0,\bar{z}_1,\ldots,\bar{z}_{\length{\bar{X}}})$. The following formula is equivalent to $\alpha$:
-\begin{align}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{not first}}(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{Y},\bar{y}))\vee \label{eq:partition1} \\
-&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))\vee \label{eq:partition2}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{not last}}(\bar{Y},\bar{y}))\vee \label{eq:partition3}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{last}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))]. \label{eq:partition4}
-\end{align}
-To show that the formula is indeed equivalent to $\alpha$, note that the formulas in lines (\ref{eq:partition1}) and (\ref{eq:partition2}) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$, and the formulas in lines (\ref{eq:partition3}) and (\ref{eq:partition4}) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
-Altogether, the four lines define pairwise disjoint assignments for $(\bar{X},\bar{x}),(\bar{Y},\bar{y})$.
-With this, it is straightforward to show that the above formula is equivalent to $\alpha$.
-However, the formula is not yet in the correct form since it has existential quantifiers in the subformulas $\gamma_{\text{not first}}$ and $\gamma_{\text{not last}}$.
-To solve this, first let take a close look to the complete formula:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\exists\bar{v}(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\exists\bar{w}(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))].
-\end{align*}
-To construct an equivalent formula that is in the correct form, we define $\bar{u} = (\bar{v},\bar{w})$ and we replace the first-order quantifiers by a first-sum and count the first assignment to $\bar{v}$ and $\bar{w}$ that satisfies the formula.
-A similar construction was used in \cite{SalujaST95}.
-Then the final formula equivalent to $\alpha$ is the following:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
-& \hspace{3cm}  \bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&
-(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))\vee
-\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
-& \hspace{3cm} (\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee \\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))].
-\end{align*}
-Finally, consider a $\eqso(\LL)$ formula $\alpha$ in $\LL$-SNF.
-If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we can consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the condition in the hypothesis.
-
-\subsection*{Proof of Theorem \ref{prop-rest}}
-We give this proof in three parts.
-
-\vspace{1em}
-First, we show that $\QE{0} \not\subseteq \E{1}$.
-By contradiction, suppose that there is a $\QE{0}$ formula $\alpha$ over some signature $\R$ such that defines the following function.
-For every finite $\R$-structure with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n - 1$.
-We use the following claim.
-\begin{claim}
-Let $\alpha = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. Then the function defined by $\alpha$ is either null, greater or equal to $n$, or is in $\Omega(n^2)$.
-\end{claim}
-\begin{proof}
-Suppose that the function defined by $\alpha$ is not $0$ and that $\varphi$ is in DNF. Furthermore, suppose $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$.
-Then $\alpha = \sa{\bar{x}} \varphi_1(\bar{x}) \vee \cdots \vee \varphi_n(\bar{x})$.
-Since $\alpha$ is not null, then some $\varphi_i$ must be satisfiable.
-This is, the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is not null.
-We will prove by induction on $\length{\bar{x}}$ that the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is either greater or equal to $n$, or in $\Omega(n^2)$.
-We address the case $\length{\bar{x}}= 1$, then $\alpha = \sa{x}\bigwedge\psi(x)$.
-If any $\psi(x) = (x = x)$ or $\neg(x < x)$, then we can eliminate it and we obtain the same function.
-If any $\psi = (x < x)$ or $\neg(x=x)$, then the function becomes null.
-If $\psi(x) = R(x,\ldots,x)$ for some $R\in\R$ the function becomes null for the structures we are considering.
-If $\psi(x) = \neg R(x,\ldots,x)$, we can eliminate it and for the structures we are considering we obtain the same function.
-The only possible $\alpha$ left is $\alpha = \sa{x}\top$ which is equal to the function $n$.
-This covers all possible cases for $\length{\bar{x}} = 1$.
-Now suppose that it holds for $\length{\bar{x}} = k$ and suppose $\alpha = \sa{\bar{x}}\bigwedge\psi(\bar{x})$ for $\length{\bar{x}} = k+1$.
-If any $\psi(\bar{x}) = (x_i = x_j)$ where $i \neq j$, then $\alpha$ describes the same function as $\alpha$ where $x_j$ has been replaced by $x_i$.
-In this formula the tuple of first-order variables has $k$ elements so the function it describes if one of the mentioned in the hypothesis.
-If $i = j$, then we can eliminate it and obtain the same function.
-If any $\psi(\bar{x}) = R(\bar{v})$ or $\neg R(\bar{v})$ where $\bar{v}$ is a sub-tuple of $\bar{x}$ then we can either eliminate it or the function becomes null, following the same argument as in the case $\length{\bar{x}} = 1$.
-If any $\psi(\bar{x}) = \neg(x_i = x_j)$ or $(x_i < x_j)$ where $i = j$, then the function becomes null.
-If any $\psi(\bar{x}) = \neg(x_i < x_j)$ where $i = j$, we can eliminate it.
-The remaining formulas in $\bigwedge\psi(\bar{x})$ are either $\neg(x_i = x_j)$, $(x_i<x_j)$ or $\neg(x_i<x_j)$.
-If the formula violates transitivity in $<$ (for example, $x < y \wedge y < z \wedge z < x$), then the function $\alpha$ describes is null.
-Therefore, there is some order over $\bar{x}$ that satisfies $\bigwedge\psi(\bar{x})$.
-Consider the formula that describes this order (like $x_1 < x_3 \wedge x_3 < x_4 \wedge x_4 < x_2$).
-The function $\alpha$ describes is greater or equal to the one this formula describes, which is exactly $\binom{n}{\length{\bar{x}}}$ which is in $\Omega(n^{\length{\bar{x}}}) \subseteq \Omega(n^2)$ if $\length{\bar{x}} > 1$.
-This concludes the proof of the claim.
-\end{proof}
-We suppose that $\alpha$ is in SNF, this is, $\alpha = \sum_{i = 1}^n\alpha_i$.
-Since $\alpha$ is not null, consider some $\alpha_i$ that describes a non-null function.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, where $\varphi$ is quantifier-free.
-Note that if $\length{\bar{X}} > 0$, then the function $\alpha$ describes is in $\Omega(2^n)$, as it was proven by the authors in \cite{SalujaST95}.
-We have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$, as we proved in the claim, describes either some function greater or equal to $n$, or in $\Omega(n^2)$, which leads to a contradiction.
-Lastly, note that the formula $\sa{x}\exists y(x < y)$ is in $\E{0}$ and describes the function $n-1$, which concludes the proof.
-
-\vspace{1em}
-Now we show that $\E{1}\not\subseteq\QE{0}$.
-In Theorem \ref{theo-pi1-pnf} we proved that there is no formula in $\loge{1}$-PNF equivalent to the formula $\alpha = 2$.
-Every formula in $\E{1}$ can be expressed in $\loge{1}$-PNF, which implies that $2 \in \QE{0}$ and $2 \not\in \E{1}$.
-
-\vspace{1em}
-Lastly, we prove that $\eqso(\loge{1})\subsetneq\eqso(\logu{1})$.
-For inclusion, let $\alpha$ be a formula in $\eqso(\loge{1})$.
-Suppose that it is in $\loge{1}$-SNF.
-This is, $\alpha = c + \sum_{i = 1}^{n}\alpha_i$.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, where $\varphi_i$ is quantifier-free, for each $\alpha_i$.
-We use the same construction used in \cite{SalujaST95}, and we obtain that the formula $\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \forall\bar{y}'(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')$ for every assignment to $(\bar{X},\bar{x})$.
-We do this replacement for each $\alpha_i$ and we obtain an equivalent formula in $\eqso(\logu{1})$.
-
-To prove that the inclusion is proper, consider the $\eqso(\logu{1})$ formula $\sa{x}\forall y(y = x)$. This formula defines the following function that takes an ordered structure $\A$ as input:
-$$
-\sem{\alpha}(\A) = 
-\begin{cases}
-1 &\A \text{ has one element}\\
-0 &\text{ otherwise}.
-\end{cases}
-$$
-Suppose that there exists an equivalent formula $\alpha$ in $\eqso(\loge{1})$.
-Also, suppose that it is in $\L$-PNF, so $\alpha = c + \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$.
-Since $\alpha$ takes the value 0 for some structures, $c$ must be 0.
-Consider a structure $\mathfrak{1}$ with one element.
-We have that for some $i$, there exists an assignment $(\bar{B},\bar{b},\bar{a})$ for $(\bar{X},\bar{x},\bar{y})$ such that $\mathfrak{1}\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
-Consider now the structure $\mathfrak{2}$ that is obtained by duplicating $\mathfrak{1}$, as we did for Theorem \ref{theo-pi1-pnf}.
-Note that $\mathfrak{2}\models\varphi_i(\bar{B},\bar{b},\bar{a})$, which implies that $\sem{\alpha}(\mathfrak{2}) \geq 1$, which leads to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop-e1-nc}}
-
 Towards a contradiction, assume that the statement is false.
@@
 We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$, but the only possible assignment to $\bar{x}$ is the tuple $(a,\ldots,a)$ so $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which follows to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop:qe0-fp-qe1-totp-fptras}}
-
-The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$.
-This is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every finite $\R$-structure $\A$ with domain $A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$.
-We base our proof on this fact.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is contained in \textsc{TotP}.} Let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Consider the following nondeterministic procedure that on input $\enc(\A)$ generates $\sem{\alpha}(\A)$ branches.
-For each $\alpha_i = \varphi$, where $\varphi$ is a $\loge{1}$ formula, it checks if $\A\models\varphi$ in polynomial time and generates a new branch if that is the case.
-For each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi$, this formula is also in $\E{1}$.
-We use the reduction to $\cdnf$ provided in \cite{SalujaST95} and we obtain $g(\enc(\A))$, which is an instance to $\cdnf$.
-Since $\cdnf$ is also in $\totp$ \cite{PagourtzisZ06}, we simulate the corresponding nondeterministic procedure that generates exactly $\cdnf(\enc(g(\A)))$ branches.
-Since, $\fp\subseteq\totp$\cite{PagourtzisZ06}, each polynomially computable function is also in $\totp$, and then on each of these branches we simulate the corresponding nondeterministic procedure to generate $h(\vert A \vert)$ more.
-The number of branches for each $\alpha_i$ is $\sem{\alpha_i}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$, and the total number of branches in the procedure amounts to $\sem{\alpha}(\A)$.
-We conclude that $\alpha\in\totp$.
-
-\vspace{1em}
-{\bf Every function in} $\boldsymbol{\eqso(\loge{1})}$ {\bf has an FPRAS.}  let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Note that each $\alpha_i$ that is equal to some $\loge{1}$ formula $\varphi$ has an FPRAS given by the procedure that simply checks if $\A\models\varphi$ and returns 1 if it does and 0 otherwise.
-Also, each remaining $\alpha_i$ has an FPRAS since $\alpha_i\in \E{1}$ \cite{SalujaST95}.
-If two functions have an FPRAS, then their sum also has one given by the procedure that simulates them both and sums the results.
-We conclude that $\alpha$ has an FPRAS.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is closed under sum and multiplication.} Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that the class is closed under multiplication.
-We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of $\so$.
-
-\begin{lemma} \label{conj-mult}
-If $\LL$ is a fragment closed under conjunction, then $\eqso(\LL)$ is closed under binary multiplication.
-\end{lemma}
-\begin{proof}
-We will define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ extended by binary product, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
-In fact, the formula $\tau(\alpha)$ is in $\LL$-SNF.
-First we replace each constant $s$ in $\alpha$ for $(\top \add \cdots \add \top)$ ($s$ times).
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$.
-We assume that for every $\beta$ that has less algebraic operators than $\alpha$, $\tau(\beta)$ is in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$ then we define $\tau(\alpha) = \tau(\alpha_1) + \tau(\alpha_2)$.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$, then we define $\tau(\alpha)$ as the formula in $\LL$-SNF that is equivalent to $\sa{x}\tau(\beta)$ and to $\tau(\alpha) = \sa{X}\tau(\beta)$, respectively.
-If $\alpha = (\alpha_1 \cdot \alpha_2)$, we assume that each $\alpha_i$ is in $\LL$-SNF.
-We identify three cases.
-(1) Some $\alpha_i$ is equal to $\sum_{j = 1}^n\beta_j$ for $n > 1$.
-Suppose wlog.
-that it is $\alpha_1$.
-We then define $\tau(\alpha) = \sum_{j = 1}^n\tau(\beta_j\cdot\alpha_2)$.
-In the following cases, $\alpha_1$ and $\alpha_2$ are in $\LL$-SNF.
-(2) If some $\alpha_i$ is equal to $\sa{X}\beta$ or $\sa{x}\beta$, we define $\tau(\alpha)$ as the $\LL$-SNF formula that is equivalent to $\sa{x}\tau(\beta\cdot\alpha_2)$ and $\sa{X}\tau(\beta\cdot\alpha_2)$, respectively.
-The remaining case is (3) $\alpha_1 = \varphi_1$ and $\alpha_2 = \varphi_2$ where each $\varphi$ is an $\LL$ formula.
-Then we define $\tau(\alpha) = \varphi_1 \wedge \varphi_2$.
-This covers all possible cases for $\alpha$.
-For every pair of formulas $\alpha,\beta$ in $\eqso(\LL)$, we have that their multiplication $(\alpha\cdot\beta)$ is a formula in the grammar $\eqso(\LL)$ extended by binary product, and so, there exists an equivalent formula $\tau(\alpha\cdot\beta)$ which is in unextended $\eqso(\LL)$.
-\end{proof}
-Since $\loge{1}$ is closed under conjunction, this also holds for $\eqso(\loge{1})$. This concludes the proof.
-
-\subsection*{Proof of Proposition \ref{pi-minusone}}
-
-Let $\LL$ be a fragment of $\fo$ that contains $\logu{1}$.
-Then we have that every function in $\U{1}$ is expressible in $\eqso(\LL)$.
-In particular, $\ctcnf \in \eqso(\LL)$.
-Suppose that $\eqso(\LL)$ is closed under subtraction by one.
-Then, the function $\ctcnf-1$, which counts the number of satisfying assignments of a 3-CNF formula minus one, is also in $\eqso(\LL)$.
-Note that $\eqso(\LL) \subseteq \eqso(\fo) = \shp$.
-We have that $\ctcnf$ is $\shp$-complete under parsimonious reductions\footnote{It can be easily verified that the standard reduction from SAT to 3-CNF (or 3-SAT) preserves the number of satisfying assignments}.
-Now, let $f$ be a function in $\shp$, and consider the nondeterministic polynomial-time procedure that on input $\enc(\A)$ computes the corresponding reduction to $\ctcnf$, name it $g(\enc(\A))$, and simulates the $\shp$ procedure for $\ctcnf-1$ on input $g(\enc(\A))$.
-We have that this is a $\shp$ procedure that computes $f-1$, from which we conclude that $\shp$ is closed under subtraction by one.
-
-\subsection*{Proof of Theorem \ref{sigmafo-minusone}}
-
-\noindent {\bf Closed under sum and multiplication.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
-This can be seen because $\eqso(\LL)$ is closed under sum for every fragment $\LL$ by definition, and since $\logex{1}$ is closed under conjunction, from Lemma \ref{conj-mult} it follows that $\eqso(\logex{1})$ is closed under multiplication.
-
-\vspace{1em}
-\noindent {\bf Easy decision version and FRPRAS.}  We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
-We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Theorem \ref{prop:qe0-fp-qe1-totp-fptras}.
-First, we define a function that converts a formula $\alpha$ in $\eqso(\logex{1})$ over a signature $\R$ into a formula $\lambda(\alpha)$ in $\eqso(\loge{1})$ over a signature $\R_{\alpha}$.
-Afterwards, we define a function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an $\R_{\alpha}$-structure $g_{\alpha}(\A)$.
-
-Let $\alpha$ be in $\eqso(\logex{0})$.
-The signature $\R_{\alpha}$ is obtained by adding the symbol $R_{\psi}$, for every $\fo$ formula $\psi(\bar{z})$ in $\alpha$, to $\R$.
-Each symbol $R_{\psi}$ represents a predicate with arity $\length{\bar{z}}$.
-Then, $\lambda(\alpha)$ is defined as $\alpha$ where each $\fo$ formula $\psi(\bar{z})$ has been replaced by $R_{\psi}(\bar{z})$.
-We now define the function $g_{\alpha}$ procedurally.
-Let $\A$ be a $\R$-structure with domain $A$.
-Let $\A'$ be an $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A}$ empty.
-For each $\fo$-formula $\psi(\z)$ with $\length{\bar{z}}$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length{\bar{z}}}$.
-If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$ (this can be done in $\ptime$).
-This concludes the construction of $\A'$.
-Note that the number of $\fo$ subformulas, arity and tuple size is fixed in $\alpha$, so computing this function takes polynomial time over the size of the structure.
-Moreover, the encoding of $\A'$ has polynomial size over the size of $\enc(\A)$.
-We define $g_{\alpha}(\A) = \A'$ and we have that for each $\R$-structure $\A$: $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$.
-Therefore, we have a parsimonious reduction from $\alpha$ to the $\eqso(\loge{1})$ formula $\lambda(\alpha)$.
-
-To show that $\alpha$ is in $\totp$, we can convert $\alpha$ and $\enc(\A)$ into $\lambda(\alpha)$ and $\enc(g_{\alpha}(\A))$, respectively, and run the procedure in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-Similarly, to show that $\alpha$ has an FPRAS, we do the same as before and simulates the FPRAS for $\lambda(\alpha)$ in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-These procedures also takes polynomial time and satisfies the condition.
-
-\vspace{1em}
-\noindent {\bf Closed under subtraction by one.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. 
-For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Without lost of generality, we assume that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
-Moreover, we assume that $\length{\bar{x}} > 0$  since, if this is not the case, we can replace $\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
-
-The proof will be separated in two parts.
-In the first part, we suppose that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$.
-Then we will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ is the formula.
-In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ the sum of one or more formulas in $\logex{1}$-PNF.
-We define a formula $\varphi'$ that satisfies the following condition: if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ and $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be proven.
-
-\vspace{1em}
-
-\noindent {\em Part (1).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ where $\varphi$ is an  $\fo$-formula.
-Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}} \varphi(\bar{x})$ (i.e.
-$\length{\bar{X}} = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}} [\varphi(\bar{x})\wedge\exists\bar{z}(\varphi(\bar{z})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
-Therefore, for the rest of the proof we can assume that $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
-
- Let $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$ and $\bar{X} = (X_1,\ldots,X_{\length{\bar{X}}})$.
-To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in into a DNF formula. 
-More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i$ is a conjunction of atomic formulas or negation of atomic formulas.
-Furthermore, we suppose that each conjunct $\varphi_i$ has the~form:
-$$
-\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
-\underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}} \wedge
-\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}}.
-$$
-Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ can appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
-
-Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ are mentioned in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
-We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
-\begin{itemize}
-	\item[(a)] In order to satisfy the first condition, consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
-We add $\length{\bar{w}}$ new variables $z_1,\ldots,z_{\length{\bar{w}}}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length{\bar{w}}})$.
-We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$.
-The formula $\varphi_i$ is now defined as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \exists \bar{z} \; (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{-}(\bar{X},\bar{x},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y})).
-	$$
-	We repeat this process for each instance of a $X_j(\bar{w})$ in $\varphi_i$, and we obtain a formula where none of the $X_j$'s acts over any variable in $\bar{x}$.
-We add the new first-order variables to $\bar{y}$ and we redefine $\varphi_i$ as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
-	$$
-	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = \exists \bar{y} \;  X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := \exists \bar{y} \;  X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
-	\item[(b)] An ordered partition on a set $S$ is defined by an equivalence relation $\sim$ over $S$, and a linear order over $S/\!\sim$.
-For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$.
-A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
-On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
-	For a given $k$, let $\cB_k$ be the number of possible ordered partitions for a set of size $k$. For $1 \leq j \leq \cB_{\length{(\bar{x},\bar{y})}}$ 
-	let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
-
-	Thus, the formula $\varphi_i(\bar{X},\bar{x})$ is then redefined as:
-	$$
-	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length{(\bar{x},\bar{y})}}} \exists \bar{y} \; [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
-	$$
-	Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
-	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  to $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
-	\item[(c)] For showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
-$\varphi_i$ is inconsistent), we do the following.
-If there exists an instance of $X_j(\bar{z})$ in $\varphi^{+}_i$, an instance of $\neg X_j(\bar{w})$ in $\varphi^{-}_i$ and the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$, then the entire formula $\varphi_i$ is removed from $\varphi$.
-\end{itemize}
-It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$.
-From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c).
-Furthermore, it has the following form:
-$$
-\varphi_i(\bar{X},\bar{x}) \; = \; \exists{\bar{y}} \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
-$$
-where $\varphi^{+}$ and $\varphi^{-}_i$ do not depend on  $\bar{x}$.
-\begin{claim}\label{claim:minusone}
-For a given ordered structure $\A$, if there exists a first-order assignment $v$ for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists a second-order assignment $V$ for $\A$ such that  $(\A,V,v)\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-\end{claim}
-\begin{proof}
-Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
-Define $\bar{B} = (B_1,\ldots,B_{\length{\bar{X}}})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
-Towards a contradiction, suppose that $(\A,V,v)\not\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-By the choice of $v$, and construction of $V$ it is clear that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})\wedge\varphi^{+}_i(\bar{X},\bar{y})$, so we necessarily have that $(\A,V,v)\not\models\varphi^{-}_i(\bar{X},\bar{y})$.
-Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,V, v)\not\models\neg X_t(\bar{w})$, namely, $v(\bar{w})\in B_t$. 
-However, by the construction of $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$.
-Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$.
-This violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
-\end{proof}
-
-The previous claim and proof motivates the following definitions.
-For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length{\bar{X}}})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
-One can easily check that for every assignments $V, v$ such that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$, then it holds that $(\A,\bar{B}^v,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all the assignment of $\bar{X}$ when $v$ is fixed.
-Therefore, by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
-For this, the idea is to choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
-$$
-\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge\forall\bar{x}'\forall\bar{y}'(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')).
-$$ 
-If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
-Furthermore, let $V^*$ the second order assignment that satisfies $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*$ the first order assignment such that $v^*(\bar{x}) = v(\bar{x})$.
-By the previous discussion, we have that $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.
-
-Now, we have all the tools to introduce the main formula of the proof. 
-Intuitively, we want to exclude the assignment $(V^*, v^*)$ as a satisfying assignment for $\varphi_i(\bar{X},\bar{x})$.
-Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if, 
-if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, then $V \neq V^*$ or $v \neq v^*$. 
-That is, it excludes all pairs $(V, v)$ that coincide with $(V^*, v^*)$ in $(\bar{X}, \bar{x})$ whenever $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. This property can be defined as follows:
-\begin{align}
-\psi_i(\bar{X},\bar{x}) = & \;\; \big( \, \exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
-& \hspace{1cm} \Big( \; 
-\exists\bar{y} \;\; \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i^+(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\
-& \hspace{1.5cm} (\exists\bar{x}'\exists\bar{y}\, \varphi_i(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}) \;\; \Big)  \label{eq:part3}
-\end{align}
-To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
-Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-Then, the conclusion of the implication at (\ref{eq:part2}) and (\ref{eq:part3}), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
-Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by showing that if $\varphi_i(\bar{X},\bar{x})$ is satisfied then $V(\bar{X})$ must contain one element more that it is not in $V^*(\bar{X})$.
-The second disjunct (\ref{eq:part3}) is satisfied when $v(\bar{x})$ is not the lexicographically smallest tuple that satisfies $\varphi_i$ (i.e. $v(\bar{x}) \neq v^*(\bar{x})$).
-
-We are ready to define the formula  $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$
-where each modified disjunct $\varphi_i'(\bar{X},\bar{x})$ is constructed as follows. 
-For the sake of simplification, define the auxiliary formula $\chi_i = \neg\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$. 
-This formula basically checks if $\varphi_i$ is satisfiable or not.
-Then, first define the formula:
-$$
-\varphi'_1(\bar{X},\bar{x}) = \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
-$$
-This models that all the assignments that satisfy $\varphi_1$ also satisfy $\varphi'_1$, except for $(V^*,v^*)$. Then we define:
-$$
-\varphi'_2(\bar{X},\bar{x}) = \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
-$$
-This models that all the assignments that satisfy $\varphi_2$ also satisfy $\varphi'_2$, but if there was no assignment for $\varphi_1$, then we exclude $(V^*,v^*)$ for $\varphi_2$.
-Then we generalize this construction for each $\varphi_i$:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})),
-$$
-By the construction of $\kappa(\alpha)$, one can easily check that  $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.
-
-\vspace{1em}
-
-{\em Part (2).} Let $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ for some $\eqso(\logex{1})$ formula $\beta$. We define $\kappa(\alpha)$ as follows.
-First, perform the same transformations to $\varphi(\bar{X},\bar{x})$ as in the previous section.
-Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c).
-We also use the previously defined formulas $\chi_i$ and $\psi_i$, for each $i$.
-
-We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if and only if $\sem{\beta}(\A) = 0$.
-If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg\exists\bar{x}'\varphi(\bar{x}')$.
-If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$ as each $\chi_i$ was previously defined.
-If $\beta = (\beta_1 + \beta_2)$, then $\lambda(\beta) = \lambda(\beta_1) \wedge \lambda(\beta_2)$.
-Now, for each $\varphi_i$ we define:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x}))\Big)\Big).
-$$
-Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
-
-%\subsection*{Proof of Theorem \ref{sub-pnp}}
-%
-%Towards a contradiction, suppose that any of the classes $\E{1}$, $\eqso(\loge{1})$, and $\eqso(\logex{1})$ is closed under subtraction.
-%
-%Let $\R$ include the symbols $S_1, S_2, S_3, S_4$, which describe the following properties.
-If a finite $\R$-structure $\A$ defines a 3DNF formula $\Phi$, then its domain is the set of variables mentioned in $\Phi$, and for each $i = 1,2,3,4$:
-%\begin{align*}
-%	S_1^\A &= \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_2^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_3^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_4^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
-%\end{align*}
-%Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
-%\begin{multline*}
-%\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
-%\end{multline*}
-%Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
-%$$
-%\varphi(x,X) = (T(x) \vee \neg T(x)).
-%$$
-%Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Given that $f_{\#3DNF}, f_{all} \in \E{1} \subseteq \eqso(\loge{1}) \subseteq \eqso(\logex{1}) \subseteq \totp$, and at least one of the classes is closed under subtraction, the function $f_{\#3DNF} - f_{all}$ is in $\totp$.
-However, note that for each $\R$-structure $\A$ that represents a formula $\Phi$, $(f_{\#3DNF} - f_{all})(\A) = 0$ if an only if $\Phi$ is a tautology.
-The decision version of this function is the $\np$-complete problem $\textsc{Tautology}$.
-And since the function is in $\totp$, its decision version is also in $\ptime$.
-We conclude that $\np \subseteq \ptime$.
-
-\subsection*{Proof of Proposition \ref{prop:ehorn-pe}}
-Pagourtzis and Zachos mention a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula \cite{PagourtzisZ06}.
-This procedure can be easily extended to receive Horn formulas, and furthermore, a disjunction of Horn formulas.
-Hence $\shdhsat$ is in $\totp$.
-
-As we show in Proposition \ref{sigma2hard}, $\shdhsat$ is complete for $\eqso(\ehorn)$ under parsimonious reductions.
-Let $\alpha$ be a formula in $\eqso(\ehorn)$ and let $g_{\alpha}$ be the reduction to $\shdhsat$.
-The $\totp$ procedure we construct, for each input $\enc(\A)$, is simply to compute $g_{\alpha}(\enc(\A))$, and then simulate the $\totp$ procedure for $\shdhsat$ on input $g_{\alpha}(\enc(\A))$.
-We conclude that $\alpha$ is in $\totp$.
-
-\subsection*{Proof of Proposition \ref{prop:hsat-not-sigma2}} %V.12
-We use a similar proof to the one provided by the authors in \cite{SalujaST95} to separate the classes $\E{2}$ and $\U{2}$.
-Suppose that the statement is false, this is, $\chsat \in \eqso(\loge{2})$.
-We consider the signature $\R$ that we used as the encoding for a Horn formula (Example \ref{ex-hornsat-esop1}) and that the formula $\alpha \in \eqso(\loge{2})$ follows the encoding in the same way.
-From what we proved in Theorem \ref{theo-pnf-snf}, we have that every formula in $\eqso(\loge{2})$ can be rewritten in $\loge{2}$-PNF, so we assume that $\alpha$ is in this form.
-Let $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\forall\bar{z}\,\varphi(\bar{X},\bar{x},\bar{y},\bar{z})$.
-Consider the following Horn formula $\Phi$:
-$$
-\Phi = p \wedge \bigwedge_{i = 1}^n (t_i \wedge p \to q) \wedge \neg q,
-$$
-where $n = \length{\bar{x}} + \length{\bar{y}} + 1$.
-Let $\A$ be the encoding of this formula.
-In the encoding, each variable appears as an element in the domain of $\A$.
-This formula is satisfiable, so $\sem{\alpha}(\A) \geq 1$.
-Let $(\bar{B},\bar{b},\bar{a})$ be an assignment to $(\bar{X},\bar{x},\bar{y})$ such that $\A\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-Let $t_{\ell}$ be such that it does not appear in $\bar{b}$ or $\bar{a}$.
-Consider the induced substructure $\A'$ that is obtained by removing $t_{\ell}$ from $\A$ and $\bar{B}'$ as the subset of $\bar{B}$ obtained by deleting each appearance of $t_{\ell}$ in $\bar{B}$.
-We have that $\A'\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-This is because each subformula of the form $\exists y \neg B_i$ is still true, and universal formulas are monotone over induced substructures.
-It follows that $\sem{\alpha}(\A') \geq 1$ which is not possible since $\A'$ encodes the formula
-$$
-\Phi' = p \wedge \bigwedge_{i = 1}^{\ell-1} (t_i \wedge p \to q) \wedge (p\to q) \wedge \bigwedge_{i = \ell+1}^{n} (t_i \wedge p \to q) \wedge \neg q,
-$$
-which is unsatisfiable. We arrive to a contradiction and we conclude that $\chsat$ is not in $\eqso(\ehorn)$.
-
-\subsection*{Proof of Theorem \ref{sigma2hard}} %V.13
-
-First we prove that $\shdhsat$ is in $\eqso(\ehorn)$.
-Recall that each instance of $\shdhsat$ is a disjunction of Horn formulas.
-Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot), \pD(\cdot,\cdot)\}$.
-Each symbol in this vocabulary is used to indicate the same as in Example \ref{ex-hornsat-esop1}, with the addition of $\pD(d,c)$ which indicates that $c$ is a clause in the formula $d$.
-Recall that the formula
-\begin{align*}
-&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \neg\textit{T}(x)).
-\end{align*}
-defines $\chsat$. We obtain the following formula $\psi(T,A)$ in $\ehorn$:
-\begin{align*}
-\exists d[&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \pD(c,d)\vee \neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \neg\textit{T}(x))]
-\end{align*}
-effectively defines $\chsat$ as for every disjunction of Horn formulas $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
-Therefore, we conclude that $\shdhsat \in \eqso(\ehorn)$.
-
-\vspace{1em}
-We will now prove that $\shdhsat$ is hard for $\eqso$ over a signature $\R$ under parsimonious reductions.
-For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a polynomial-time procedure that computes a function $g_{\alpha}$.
-This function receives a finite $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
-We suppose that $\alpha$ is in sum normal form and:
-$$
-\alpha = \sum_{i = 1}^{\text{\#clauses}} \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigwedge_{j = 1}^{n}\forall\bar{z}\,\varphi^i_j(\bar{X},\bar{x},\bar{y},\bar{z}),
-$$
-where each $\varphi^i_j$ is a Horn clause.
-
-Consider a finite $\R$-structure $\A$ with domain $A$.
-To simplify the proof, we extend our grammar to allow first-order constants.
-Consider each tuple $\bar{a}\in A^{\length{\bar{x}}}$, each $\bar{b}\in A^{\length{\bar{y}}}$ and each $\bar{c}\in A^{\length{\bar{z}}}$ as a tuple of first-order constants.
-The following formula defines the same function as $\alpha$:
-$$
-\sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length{\bar{x}}}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length{\bar{y}}}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length{\bar{z}}}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
-$$
-Note that each $\fo$ formula over $(\bar{x},\bar{y},\bar{z})$ in each $\varphi^i_j$ has no free variables.
-Therefore, we can evaluate each of these in polynomial time and replace them by $\perp$ and $\top$ where it corresponds.
-Each $\varphi^i_j$ will be of the form $\perp \vee\, \chi^i_j(\bar{X})$ or $\top \vee \chi^i_j(\bar{X})$ where $\chi^i_j$ is a disjunction of $\neg X_{\ell}$'s and at most one $X_{\ell}$.
-The formulas of the form $\top \vee \chi^i_j(\bar{X})$ can be removed entirely, and the formulas of the form $\perp \vee\, \chi^i_j(\bar{X})$ can be replaced by $\chi^i_j(\bar{X})$.
-We obtain the formula
-$$
-\sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})
-$$
-where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$.
-
-Our idea for the rest of the proof is to define $g_{\alpha}$ for each $\alpha = \sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})$, for $\alpha = c$ and for $\alpha = \beta_1 + \cdots + \beta_m$ where each $\beta_i$ is in one of the two previous cases.
-
-If $\alpha$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ where $\psi_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$, then we obtain the {\bf propositional formula} $g_{\alpha}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulas.
-
-If $\alpha$ is equal to a constant $c$, then we define $g_{\alpha}(\A)$ as the following formula that has exactly $c$ satisfying assignments:
-$$
-g_{\alpha}(\A) = \bigvee_{i = 1}^{c}\neg t_1 \wedge \cdots \wedge \neg t_{i-1} \wedge t_i \wedge \neg t_{i+1} \wedge \cdots \wedge \neg p_c.
-$$ 
-If $\alpha = \beta_1 + \cdots + \beta_m$, let $g_{\beta_i}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\theta^i_{j,k}$ for each $\beta_i$ where each $\theta^i_{j,k}$ is a Horn clause.
-Let $\Theta_i = g_{\beta_i}(\A)$.
-We rename the variables in each $\Theta_i$ so none of them are mentioned in any other $\Theta_j$.
-We add $m$ new variables $t_1,\ldots,t_m$ and we define:
-\begin{align*}
-g_{\alpha}(\A) = &\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^1_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m}}}t) \wedge (t_1 \wedge \bigwedge_{\ell = 2}^{m} \neg t_{\ell})) \vee \\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^2_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each $t$}\\ \text{ mentioned in}\\ \Theta_1,\Theta_3,\ldots,\Theta_{m}}}t) \wedge (t_2 \wedge \bigwedge\limits_{\substack{\ell = 1 \\ \ell \neq 2}}^{m} \neg t_{\ell})) \vee \cdots \vee\\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^m_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m-1}}}t) \wedge (t_m \wedge \bigwedge_{\ell = 1}^{m-1} \neg t_{\ell})).
-\end{align*}
-The formula is a disjunction of Horn formulas, and the number of satisfying assignments for this formula is exactly the sum of satisfying assignments for each $g_{\beta_i}(\A)$.
-This, at the same time, is equal to $\sem{\alpha}(\A)$.
-This covers all possible cases for $\alpha$, and the entire procedure takes polynomial time.

### 2017-03-03 LMCS/proof-prop-rest.tex
journal primera version

--- 2017-01-12 LICS/appendix-syntactic.tex
@@
-\subsection*{Proof of Theorem \ref{theo-pnf-snf}}
-Recall that a formula in $\eqso(\LL)$ is on the following grammar:
-$$
-\alpha = \varphi \ \mid \ s \ \mid \ (\alpha + \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \sa{X} \alpha,
-$$
-where $\varphi$ is a formula in $\LL$ and $s\in\nat$.
-We will construct a recursive function $\tau$ such that for every $\eqso(\LL)$ formula $\alpha$, it outputs an equivalent formula $\tau(\alpha)$ which is in $\LL$-SNF.
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$ which is clearly equivalent and already in $\LL$-SNF.
-If $\alpha = s$, then we define $(\top \add \cdots \add \top)$ ($s$ times), which also satisfies the condition.
-We assume that for every sub-formula $\beta$ of $\alpha$, $\tau(\beta)$ is an equivalent formula in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$, then we define $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$.
-If $\alpha = \sa{x}\beta$, then let $\tau(\beta) = \sum_{i = 1}^{k}\beta_i$ for some $k$ where each $\beta_i$ is in $\LL$-PNF.
-By grouping together the terms in the sum, we define $\tau(\alpha) = \sum_{i = 1}^{m}\sa{x}\beta_i$ which satisfies the condition and is equivalent to $\alpha$.
-If $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case.
-This covers all possible cases for $\alpha$ and we conclude the proof by taking $\tau(\alpha)$ as the desired rewrite of $\alpha$.
-
-\subsection*{Proof of theorem \ref{theo-pi1-pnf}}
-We divide the proof in three parts.
-
-\vspace{1em}
-First, we prove that the formula $\alpha_{0} = \left( \sa{X} 1 \right) + 1$ with $\arity(X) = 1$ (i.e.
-the function $2^{n}+1$, where $n$ is the size of the input structure) is not equivalent to any formula in $\loge{0}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ in $\loge{0}$-PNF that is equivalent to the $\eqso(\loge{0})$ formula $\alpha_0$.
-In \cite{SalujaST95}, it was proved that if $\length{\bar{X}} > 0$, then the function defined by $\alpha$, for big enough structures, is always even which is not possible in our case.
-On the other hand, if $\alpha$ is of the form $\sa{\bar{x}}\varphi(\bar{x})$, then $\alpha$ defines a polynomially bounded function which also leads to a contradiction.
-
-\vspace{1em}
-Second, we prove that the formula $\alpha_{1} = 2$ (i.e.
-the constant $2$) is not equivalent to any formula in $\loge{1}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to the $\eqso(\loge{1})$ formula $2$.
-First, if $\length{\bar{X}} = \length{\bar{x}} = 0$, then the function defined by $\alpha$ is never greater than 1. 
-Therefore, suppose that $\length{\bar{X}} > 0$ or $\length{\bar{x}} > 0$, and consider any ordered structure $\A$. 
-Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
-Now consider the ordered structure $\A'$ that is obtained by duplicating $\A$.
-This is, each half of $\A'$ is isomorphic to $\A$.
-Note that $\A'\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$ for $i = 1,2$ and there exists a third assignments $(\bar{B}_1',\bar{b}_1',\bar{a}_1')$ that is isomorphic to $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$ but in the other half to the structure such that $\A'\models\varphi(\bar{B}_1',\bar{b}_1',\bar{a}_1')$.
-We have that $\sem{\alpha}(\A) \geq 3$ which is a contradiction.
-
-\vspace{1em}
-We now show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there is an equivalent formula $\beta$ in $\LL$-PNF.
-As in Theorem \ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such formula.
-As we showed, there exists an equivalent formula in $\LL$-SNF, so we assume that $\alpha$ is in that form.
-Let $\alpha = \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-SNF.
-Without lost of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
-If not, we replace each $\alpha_i$ for the equivalent formula $\sa{\bar{X}} \sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge\forall z\,Y(z) \wedge \forall z(y \leq z))$.
-
-Now we begin describing the function $\tau$.
-If $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then the formula is already in $\LL$-PNF so we define $\tau(\alpha) = \alpha$.
-If $\alpha = \alpha_1 + \alpha_2$, then we assume that $\tau(\alpha_1) = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ and $\tau(\alpha_2) = \sa{\bar{Y}}\sa{\bar{y}}\psi(\bar{Y},\bar{y})$.
-The construction that we will provide for this function works by identifying a ``first'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ and a ``last'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$.
-These are identified by the following formulas:
-\begin{align*}
-\gamma_{\text{first}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z}), \\
-\gamma_{\text{last}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x}).
-\end{align*}
-Similarly, we define the formulas $\gamma_{\text{first}}(\bar{Y},\bar{y})$ and $\gamma_{\text{last}}(\bar{Y},\bar{y})$ (for the sake of simplicity we reuse the names $\gamma_{\text{first}}$ and $\gamma_{\text{last}}$).
-In other words, the ``first'' assignment is the one where every second-order predicate is empty and the first-order assignment is the lexicographically smallest, and the ``last'' assignment is the one where every second-order predicate is full and the first-order assignment is the lexicographically greatest.
-We also need to identify assignments that are not first and are not last.
-We do this by negating the two formulas above and grouping together the first-order variables:
-\begin{align*}
-\gamma_{\text{not first}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{z}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{z}_i)), \\
-\gamma_{\text{not last}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{x} < \bar{z}_0 \vee \bigvee_{i = 1}^{\length{\bar{X}}}\neg X(\bar{z}_i)),
-\end{align*}
-where $\bar{z} = (\bar{z}_0,\bar{z}_1,\ldots,\bar{z}_{\length{\bar{X}}})$. The following formula is equivalent to $\alpha$:
-\begin{align}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{not first}}(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{Y},\bar{y}))\vee \label{eq:partition1} \\
-&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))\vee \label{eq:partition2}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{not last}}(\bar{Y},\bar{y}))\vee \label{eq:partition3}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{last}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))]. \label{eq:partition4}
-\end{align}
-To show that the formula is indeed equivalent to $\alpha$, note that the formulas in lines (\ref{eq:partition1}) and (\ref{eq:partition2}) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$, and the formulas in lines (\ref{eq:partition3}) and (\ref{eq:partition4}) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
-Altogether, the four lines define pairwise disjoint assignments for $(\bar{X},\bar{x}),(\bar{Y},\bar{y})$.
-With this, it is straightforward to show that the above formula is equivalent to $\alpha$.
-However, the formula is not yet in the correct form since it has existential quantifiers in the subformulas $\gamma_{\text{not first}}$ and $\gamma_{\text{not last}}$.
-To solve this, first let take a close look to the complete formula:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\exists\bar{v}(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\exists\bar{w}(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))].
-\end{align*}
-To construct an equivalent formula that is in the correct form, we define $\bar{u} = (\bar{v},\bar{w})$ and we replace the first-order quantifiers by a first-sum and count the first assignment to $\bar{v}$ and $\bar{w}$ that satisfies the formula.
-A similar construction was used in \cite{SalujaST95}.
-Then the final formula equivalent to $\alpha$ is the following:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
-& \hspace{3cm}  \bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&
-(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))\vee
-\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
-& \hspace{3cm} (\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee \\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))].
-\end{align*}
-Finally, consider a $\eqso(\LL)$ formula $\alpha$ in $\LL$-SNF.
-If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we can consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the condition in the hypothesis.
-
-\subsection*{Proof of Theorem \ref{prop-rest}}
 We give this proof in three parts.
@@
 Note that $\mathfrak{2}\models\varphi_i(\bar{B},\bar{b},\bar{a})$, which implies that $\sem{\alpha}(\mathfrak{2}) \geq 1$, which leads to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop-e1-nc}}
-
-Towards a contradiction, assume that the statement is false.
-This is, that $\E{1}$ is closed under binary sum.
-Consider the formula $\sa{x}(x = x)$ which is in $\E{1}$ over some signature $\R$.
-For every finite $\R$-structure $\A$ with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n$.
-From our assumption, there exists some formula in $\E{1}$ equivalent to the formula $\alpha \add \alpha$, which describes the function $2n$.
-Let $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ be this formula, where $\varphi$ is quantifier-free.
-Note that the function defined by this formula is equal or greater than the one defined by $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ divided by a polynomial factor.
-More specifically, for each ordered structure $\A$ with domain $A$, we have the following inequality:
-$$
-\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot \vert A \vert^{\length{\bar{y}}} \geq \sem{\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A)
-$$
-Note that the formula $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ defines a function in $\E{0}$.
-It was shown by the authors in \cite{SalujaST95} that every function in $\E{0}$ grows exponentially over the size of the structure for large enough structures, when $\length{\bar{X}} > 0$.
-This function divided by a polynomial factor still grows exponentially.
-Therefore, for $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ we have that $\length{\bar{X}} = 0$.
-
-Now, for the formula $\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ consider a structure $\mathfrak{1}$ with only one element $a$.
-We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$, but the only possible assignment to $\bar{x}$ is the tuple $(a,\ldots,a)$ so $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which follows to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop:qe0-fp-qe1-totp-fptras}}
-
-The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$.
-This is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every finite $\R$-structure $\A$ with domain $A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$.
-We base our proof on this fact.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is contained in \textsc{TotP}.} Let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Consider the following nondeterministic procedure that on input $\enc(\A)$ generates $\sem{\alpha}(\A)$ branches.
-For each $\alpha_i = \varphi$, where $\varphi$ is a $\loge{1}$ formula, it checks if $\A\models\varphi$ in polynomial time and generates a new branch if that is the case.
-For each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi$, this formula is also in $\E{1}$.
-We use the reduction to $\cdnf$ provided in \cite{SalujaST95} and we obtain $g(\enc(\A))$, which is an instance to $\cdnf$.
-Since $\cdnf$ is also in $\totp$ \cite{PagourtzisZ06}, we simulate the corresponding nondeterministic procedure that generates exactly $\cdnf(\enc(g(\A)))$ branches.
-Since, $\fp\subseteq\totp$\cite{PagourtzisZ06}, each polynomially computable function is also in $\totp$, and then on each of these branches we simulate the corresponding nondeterministic procedure to generate $h(\vert A \vert)$ more.
-The number of branches for each $\alpha_i$ is $\sem{\alpha_i}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$, and the total number of branches in the procedure amounts to $\sem{\alpha}(\A)$.
-We conclude that $\alpha\in\totp$.
-
-\vspace{1em}
-{\bf Every function in} $\boldsymbol{\eqso(\loge{1})}$ {\bf has an FPRAS.}  let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Note that each $\alpha_i$ that is equal to some $\loge{1}$ formula $\varphi$ has an FPRAS given by the procedure that simply checks if $\A\models\varphi$ and returns 1 if it does and 0 otherwise.
-Also, each remaining $\alpha_i$ has an FPRAS since $\alpha_i\in \E{1}$ \cite{SalujaST95}.
-If two functions have an FPRAS, then their sum also has one given by the procedure that simulates them both and sums the results.
-We conclude that $\alpha$ has an FPRAS.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is closed under sum and multiplication.} Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that the class is closed under multiplication.
-We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of $\so$.
-
-\begin{lemma} \label{conj-mult}
-If $\LL$ is a fragment closed under conjunction, then $\eqso(\LL)$ is closed under binary multiplication.
-\end{lemma}
-\begin{proof}
-We will define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ extended by binary product, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
-In fact, the formula $\tau(\alpha)$ is in $\LL$-SNF.
-First we replace each constant $s$ in $\alpha$ for $(\top \add \cdots \add \top)$ ($s$ times).
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$.
-We assume that for every $\beta$ that has less algebraic operators than $\alpha$, $\tau(\beta)$ is in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$ then we define $\tau(\alpha) = \tau(\alpha_1) + \tau(\alpha_2)$.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$, then we define $\tau(\alpha)$ as the formula in $\LL$-SNF that is equivalent to $\sa{x}\tau(\beta)$ and to $\tau(\alpha) = \sa{X}\tau(\beta)$, respectively.
-If $\alpha = (\alpha_1 \cdot \alpha_2)$, we assume that each $\alpha_i$ is in $\LL$-SNF.
-We identify three cases.
-(1) Some $\alpha_i$ is equal to $\sum_{j = 1}^n\beta_j$ for $n > 1$.
-Suppose wlog.
-that it is $\alpha_1$.
-We then define $\tau(\alpha) = \sum_{j = 1}^n\tau(\beta_j\cdot\alpha_2)$.
-In the following cases, $\alpha_1$ and $\alpha_2$ are in $\LL$-SNF.
-(2) If some $\alpha_i$ is equal to $\sa{X}\beta$ or $\sa{x}\beta$, we define $\tau(\alpha)$ as the $\LL$-SNF formula that is equivalent to $\sa{x}\tau(\beta\cdot\alpha_2)$ and $\sa{X}\tau(\beta\cdot\alpha_2)$, respectively.
-The remaining case is (3) $\alpha_1 = \varphi_1$ and $\alpha_2 = \varphi_2$ where each $\varphi$ is an $\LL$ formula.
-Then we define $\tau(\alpha) = \varphi_1 \wedge \varphi_2$.
-This covers all possible cases for $\alpha$.
-For every pair of formulas $\alpha,\beta$ in $\eqso(\LL)$, we have that their multiplication $(\alpha\cdot\beta)$ is a formula in the grammar $\eqso(\LL)$ extended by binary product, and so, there exists an equivalent formula $\tau(\alpha\cdot\beta)$ which is in unextended $\eqso(\LL)$.
-\end{proof}
-Since $\loge{1}$ is closed under conjunction, this also holds for $\eqso(\loge{1})$. This concludes the proof.
-
-\subsection*{Proof of Proposition \ref{pi-minusone}}
-
-Let $\LL$ be a fragment of $\fo$ that contains $\logu{1}$.
-Then we have that every function in $\U{1}$ is expressible in $\eqso(\LL)$.
-In particular, $\ctcnf \in \eqso(\LL)$.
-Suppose that $\eqso(\LL)$ is closed under subtraction by one.
-Then, the function $\ctcnf-1$, which counts the number of satisfying assignments of a 3-CNF formula minus one, is also in $\eqso(\LL)$.
-Note that $\eqso(\LL) \subseteq \eqso(\fo) = \shp$.
-We have that $\ctcnf$ is $\shp$-complete under parsimonious reductions\footnote{It can be easily verified that the standard reduction from SAT to 3-CNF (or 3-SAT) preserves the number of satisfying assignments}.
-Now, let $f$ be a function in $\shp$, and consider the nondeterministic polynomial-time procedure that on input $\enc(\A)$ computes the corresponding reduction to $\ctcnf$, name it $g(\enc(\A))$, and simulates the $\shp$ procedure for $\ctcnf-1$ on input $g(\enc(\A))$.
-We have that this is a $\shp$ procedure that computes $f-1$, from which we conclude that $\shp$ is closed under subtraction by one.
-
-\subsection*{Proof of Theorem \ref{sigmafo-minusone}}
-
-\noindent {\bf Closed under sum and multiplication.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
-This can be seen because $\eqso(\LL)$ is closed under sum for every fragment $\LL$ by definition, and since $\logex{1}$ is closed under conjunction, from Lemma \ref{conj-mult} it follows that $\eqso(\logex{1})$ is closed under multiplication.
-
-\vspace{1em}
-\noindent {\bf Easy decision version and FRPRAS.}  We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
-We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Theorem \ref{prop:qe0-fp-qe1-totp-fptras}.
-First, we define a function that converts a formula $\alpha$ in $\eqso(\logex{1})$ over a signature $\R$ into a formula $\lambda(\alpha)$ in $\eqso(\loge{1})$ over a signature $\R_{\alpha}$.
-Afterwards, we define a function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an $\R_{\alpha}$-structure $g_{\alpha}(\A)$.
-
-Let $\alpha$ be in $\eqso(\logex{0})$.
-The signature $\R_{\alpha}$ is obtained by adding the symbol $R_{\psi}$, for every $\fo$ formula $\psi(\bar{z})$ in $\alpha$, to $\R$.
-Each symbol $R_{\psi}$ represents a predicate with arity $\length{\bar{z}}$.
-Then, $\lambda(\alpha)$ is defined as $\alpha$ where each $\fo$ formula $\psi(\bar{z})$ has been replaced by $R_{\psi}(\bar{z})$.
-We now define the function $g_{\alpha}$ procedurally.
-Let $\A$ be a $\R$-structure with domain $A$.
-Let $\A'$ be an $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A}$ empty.
-For each $\fo$-formula $\psi(\z)$ with $\length{\bar{z}}$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length{\bar{z}}}$.
-If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$ (this can be done in $\ptime$).
-This concludes the construction of $\A'$.
-Note that the number of $\fo$ subformulas, arity and tuple size is fixed in $\alpha$, so computing this function takes polynomial time over the size of the structure.
-Moreover, the encoding of $\A'$ has polynomial size over the size of $\enc(\A)$.
-We define $g_{\alpha}(\A) = \A'$ and we have that for each $\R$-structure $\A$: $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$.
-Therefore, we have a parsimonious reduction from $\alpha$ to the $\eqso(\loge{1})$ formula $\lambda(\alpha)$.
-
-To show that $\alpha$ is in $\totp$, we can convert $\alpha$ and $\enc(\A)$ into $\lambda(\alpha)$ and $\enc(g_{\alpha}(\A))$, respectively, and run the procedure in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-Similarly, to show that $\alpha$ has an FPRAS, we do the same as before and simulates the FPRAS for $\lambda(\alpha)$ in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-These procedures also takes polynomial time and satisfies the condition.
-
-\vspace{1em}
-\noindent {\bf Closed under subtraction by one.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. 
-For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Without lost of generality, we assume that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
-Moreover, we assume that $\length{\bar{x}} > 0$  since, if this is not the case, we can replace $\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
-
-The proof will be separated in two parts.
-In the first part, we suppose that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$.
-Then we will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ is the formula.
-In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ the sum of one or more formulas in $\logex{1}$-PNF.
-We define a formula $\varphi'$ that satisfies the following condition: if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ and $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be proven.
-
-\vspace{1em}
-
-\noindent {\em Part (1).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ where $\varphi$ is an  $\fo$-formula.
-Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}} \varphi(\bar{x})$ (i.e.
-$\length{\bar{X}} = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}} [\varphi(\bar{x})\wedge\exists\bar{z}(\varphi(\bar{z})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
-Therefore, for the rest of the proof we can assume that $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
-
- Let $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$ and $\bar{X} = (X_1,\ldots,X_{\length{\bar{X}}})$.
-To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in into a DNF formula. 
-More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i$ is a conjunction of atomic formulas or negation of atomic formulas.
-Furthermore, we suppose that each conjunct $\varphi_i$ has the~form:
-$$
-\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
-\underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}} \wedge
-\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}}.
-$$
-Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ can appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
-
-Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ are mentioned in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
-We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
-\begin{itemize}
-	\item[(a)] In order to satisfy the first condition, consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
-We add $\length{\bar{w}}$ new variables $z_1,\ldots,z_{\length{\bar{w}}}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length{\bar{w}}})$.
-We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$.
-The formula $\varphi_i$ is now defined as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \exists \bar{z} \; (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{-}(\bar{X},\bar{x},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y})).
-	$$
-	We repeat this process for each instance of a $X_j(\bar{w})$ in $\varphi_i$, and we obtain a formula where none of the $X_j$'s acts over any variable in $\bar{x}$.
-We add the new first-order variables to $\bar{y}$ and we redefine $\varphi_i$ as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
-	$$
-	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = \exists \bar{y} \;  X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := \exists \bar{y} \;  X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
-	\item[(b)] An ordered partition on a set $S$ is defined by an equivalence relation $\sim$ over $S$, and a linear order over $S/\!\sim$.
-For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$.
-A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
-On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
-	For a given $k$, let $\cB_k$ be the number of possible ordered partitions for a set of size $k$. For $1 \leq j \leq \cB_{\length{(\bar{x},\bar{y})}}$ 
-	let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
-
-	Thus, the formula $\varphi_i(\bar{X},\bar{x})$ is then redefined as:
-	$$
-	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length{(\bar{x},\bar{y})}}} \exists \bar{y} \; [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
-	$$
-	Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
-	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  to $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
-	\item[(c)] For showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
-$\varphi_i$ is inconsistent), we do the following.
-If there exists an instance of $X_j(\bar{z})$ in $\varphi^{+}_i$, an instance of $\neg X_j(\bar{w})$ in $\varphi^{-}_i$ and the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$, then the entire formula $\varphi_i$ is removed from $\varphi$.
-\end{itemize}
-It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$.
-From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c).
-Furthermore, it has the following form:
-$$
-\varphi_i(\bar{X},\bar{x}) \; = \; \exists{\bar{y}} \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
-$$
-where $\varphi^{+}$ and $\varphi^{-}_i$ do not depend on  $\bar{x}$.
-\begin{claim}\label{claim:minusone}
-For a given ordered structure $\A$, if there exists a first-order assignment $v$ for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists a second-order assignment $V$ for $\A$ such that  $(\A,V,v)\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-\end{claim}
-\begin{proof}
-Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
-Define $\bar{B} = (B_1,\ldots,B_{\length{\bar{X}}})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
-Towards a contradiction, suppose that $(\A,V,v)\not\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-By the choice of $v$, and construction of $V$ it is clear that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})\wedge\varphi^{+}_i(\bar{X},\bar{y})$, so we necessarily have that $(\A,V,v)\not\models\varphi^{-}_i(\bar{X},\bar{y})$.
-Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,V, v)\not\models\neg X_t(\bar{w})$, namely, $v(\bar{w})\in B_t$. 
-However, by the construction of $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$.
-Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$.
-This violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
-\end{proof}
-
-The previous claim and proof motivates the following definitions.
-For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length{\bar{X}}})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
-One can easily check that for every assignments $V, v$ such that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$, then it holds that $(\A,\bar{B}^v,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all the assignment of $\bar{X}$ when $v$ is fixed.
-Therefore, by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
-For this, the idea is to choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
-$$
-\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge\forall\bar{x}'\forall\bar{y}'(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')).
-$$ 
-If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
-Furthermore, let $V^*$ the second order assignment that satisfies $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*$ the first order assignment such that $v^*(\bar{x}) = v(\bar{x})$.
-By the previous discussion, we have that $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.
-
-Now, we have all the tools to introduce the main formula of the proof. 
-Intuitively, we want to exclude the assignment $(V^*, v^*)$ as a satisfying assignment for $\varphi_i(\bar{X},\bar{x})$.
-Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if, 
-if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, then $V \neq V^*$ or $v \neq v^*$. 
-That is, it excludes all pairs $(V, v)$ that coincide with $(V^*, v^*)$ in $(\bar{X}, \bar{x})$ whenever $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. This property can be defined as follows:
-\begin{align}
-\psi_i(\bar{X},\bar{x}) = & \;\; \big( \, \exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
-& \hspace{1cm} \Big( \; 
-\exists\bar{y} \;\; \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i^+(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\
-& \hspace{1.5cm} (\exists\bar{x}'\exists\bar{y}\, \varphi_i(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}) \;\; \Big)  \label{eq:part3}
-\end{align}
-To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
-Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-Then, the conclusion of the implication at (\ref{eq:part2}) and (\ref{eq:part3}), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
-Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by showing that if $\varphi_i(\bar{X},\bar{x})$ is satisfied then $V(\bar{X})$ must contain one element more that it is not in $V^*(\bar{X})$.
-The second disjunct (\ref{eq:part3}) is satisfied when $v(\bar{x})$ is not the lexicographically smallest tuple that satisfies $\varphi_i$ (i.e. $v(\bar{x}) \neq v^*(\bar{x})$).
-
-We are ready to define the formula  $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$
-where each modified disjunct $\varphi_i'(\bar{X},\bar{x})$ is constructed as follows. 
-For the sake of simplification, define the auxiliary formula $\chi_i = \neg\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$. 
-This formula basically checks if $\varphi_i$ is satisfiable or not.
-Then, first define the formula:
-$$
-\varphi'_1(\bar{X},\bar{x}) = \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
-$$
-This models that all the assignments that satisfy $\varphi_1$ also satisfy $\varphi'_1$, except for $(V^*,v^*)$. Then we define:
-$$
-\varphi'_2(\bar{X},\bar{x}) = \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
-$$
-This models that all the assignments that satisfy $\varphi_2$ also satisfy $\varphi'_2$, but if there was no assignment for $\varphi_1$, then we exclude $(V^*,v^*)$ for $\varphi_2$.
-Then we generalize this construction for each $\varphi_i$:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})),
-$$
-By the construction of $\kappa(\alpha)$, one can easily check that  $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.
-
-\vspace{1em}
-
-{\em Part (2).} Let $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ for some $\eqso(\logex{1})$ formula $\beta$. We define $\kappa(\alpha)$ as follows.
-First, perform the same transformations to $\varphi(\bar{X},\bar{x})$ as in the previous section.
-Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c).
-We also use the previously defined formulas $\chi_i$ and $\psi_i$, for each $i$.
-
-We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if and only if $\sem{\beta}(\A) = 0$.
-If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg\exists\bar{x}'\varphi(\bar{x}')$.
-If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$ as each $\chi_i$ was previously defined.
-If $\beta = (\beta_1 + \beta_2)$, then $\lambda(\beta) = \lambda(\beta_1) \wedge \lambda(\beta_2)$.
-Now, for each $\varphi_i$ we define:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x}))\Big)\Big).
-$$
-Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
-
-%\subsection*{Proof of Theorem \ref{sub-pnp}}
-%
-%Towards a contradiction, suppose that any of the classes $\E{1}$, $\eqso(\loge{1})$, and $\eqso(\logex{1})$ is closed under subtraction.
-%
-%Let $\R$ include the symbols $S_1, S_2, S_3, S_4$, which describe the following properties.
-If a finite $\R$-structure $\A$ defines a 3DNF formula $\Phi$, then its domain is the set of variables mentioned in $\Phi$, and for each $i = 1,2,3,4$:
-%\begin{align*}
-%	S_1^\A &= \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_2^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_3^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_4^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
-%\end{align*}
-%Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
-%\begin{multline*}
-%\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
-%\end{multline*}
-%Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
-%$$
-%\varphi(x,X) = (T(x) \vee \neg T(x)).
-%$$
-%Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Given that $f_{\#3DNF}, f_{all} \in \E{1} \subseteq \eqso(\loge{1}) \subseteq \eqso(\logex{1}) \subseteq \totp$, and at least one of the classes is closed under subtraction, the function $f_{\#3DNF} - f_{all}$ is in $\totp$.
-However, note that for each $\R$-structure $\A$ that represents a formula $\Phi$, $(f_{\#3DNF} - f_{all})(\A) = 0$ if an only if $\Phi$ is a tautology.
-The decision version of this function is the $\np$-complete problem $\textsc{Tautology}$.
-And since the function is in $\totp$, its decision version is also in $\ptime$.
-We conclude that $\np \subseteq \ptime$.
-
-\subsection*{Proof of Proposition \ref{prop:ehorn-pe}}
-Pagourtzis and Zachos mention a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula \cite{PagourtzisZ06}.
-This procedure can be easily extended to receive Horn formulas, and furthermore, a disjunction of Horn formulas.
-Hence $\shdhsat$ is in $\totp$.
-
-As we show in Proposition \ref{sigma2hard}, $\shdhsat$ is complete for $\eqso(\ehorn)$ under parsimonious reductions.
-Let $\alpha$ be a formula in $\eqso(\ehorn)$ and let $g_{\alpha}$ be the reduction to $\shdhsat$.
-The $\totp$ procedure we construct, for each input $\enc(\A)$, is simply to compute $g_{\alpha}(\enc(\A))$, and then simulate the $\totp$ procedure for $\shdhsat$ on input $g_{\alpha}(\enc(\A))$.
-We conclude that $\alpha$ is in $\totp$.
-
-\subsection*{Proof of Proposition \ref{prop:hsat-not-sigma2}} %V.12
-We use a similar proof to the one provided by the authors in \cite{SalujaST95} to separate the classes $\E{2}$ and $\U{2}$.
-Suppose that the statement is false, this is, $\chsat \in \eqso(\loge{2})$.
-We consider the signature $\R$ that we used as the encoding for a Horn formula (Example \ref{ex-hornsat-esop1}) and that the formula $\alpha \in \eqso(\loge{2})$ follows the encoding in the same way.
-From what we proved in Theorem \ref{theo-pnf-snf}, we have that every formula in $\eqso(\loge{2})$ can be rewritten in $\loge{2}$-PNF, so we assume that $\alpha$ is in this form.
-Let $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\forall\bar{z}\,\varphi(\bar{X},\bar{x},\bar{y},\bar{z})$.
-Consider the following Horn formula $\Phi$:
-$$
-\Phi = p \wedge \bigwedge_{i = 1}^n (t_i \wedge p \to q) \wedge \neg q,
-$$
-where $n = \length{\bar{x}} + \length{\bar{y}} + 1$.
-Let $\A$ be the encoding of this formula.
-In the encoding, each variable appears as an element in the domain of $\A$.
-This formula is satisfiable, so $\sem{\alpha}(\A) \geq 1$.
-Let $(\bar{B},\bar{b},\bar{a})$ be an assignment to $(\bar{X},\bar{x},\bar{y})$ such that $\A\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-Let $t_{\ell}$ be such that it does not appear in $\bar{b}$ or $\bar{a}$.
-Consider the induced substructure $\A'$ that is obtained by removing $t_{\ell}$ from $\A$ and $\bar{B}'$ as the subset of $\bar{B}$ obtained by deleting each appearance of $t_{\ell}$ in $\bar{B}$.
-We have that $\A'\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-This is because each subformula of the form $\exists y \neg B_i$ is still true, and universal formulas are monotone over induced substructures.
-It follows that $\sem{\alpha}(\A') \geq 1$ which is not possible since $\A'$ encodes the formula
-$$
-\Phi' = p \wedge \bigwedge_{i = 1}^{\ell-1} (t_i \wedge p \to q) \wedge (p\to q) \wedge \bigwedge_{i = \ell+1}^{n} (t_i \wedge p \to q) \wedge \neg q,
-$$
-which is unsatisfiable. We arrive to a contradiction and we conclude that $\chsat$ is not in $\eqso(\ehorn)$.
-
-\subsection*{Proof of Theorem \ref{sigma2hard}} %V.13
-
-First we prove that $\shdhsat$ is in $\eqso(\ehorn)$.
-Recall that each instance of $\shdhsat$ is a disjunction of Horn formulas.
-Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot), \pD(\cdot,\cdot)\}$.
-Each symbol in this vocabulary is used to indicate the same as in Example \ref{ex-hornsat-esop1}, with the addition of $\pD(d,c)$ which indicates that $c$ is a clause in the formula $d$.
-Recall that the formula
-\begin{align*}
-&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \neg\textit{T}(x)).
-\end{align*}
-defines $\chsat$. We obtain the following formula $\psi(T,A)$ in $\ehorn$:
-\begin{align*}
-\exists d[&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \pD(c,d)\vee \neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \neg\textit{T}(x))]
-\end{align*}
-effectively defines $\chsat$ as for every disjunction of Horn formulas $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
-Therefore, we conclude that $\shdhsat \in \eqso(\ehorn)$.
-
-\vspace{1em}
-We will now prove that $\shdhsat$ is hard for $\eqso$ over a signature $\R$ under parsimonious reductions.
-For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a polynomial-time procedure that computes a function $g_{\alpha}$.
-This function receives a finite $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
-We suppose that $\alpha$ is in sum normal form and:
-$$
-\alpha = \sum_{i = 1}^{\text{\#clauses}} \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigwedge_{j = 1}^{n}\forall\bar{z}\,\varphi^i_j(\bar{X},\bar{x},\bar{y},\bar{z}),
-$$
-where each $\varphi^i_j$ is a Horn clause.
-
-Consider a finite $\R$-structure $\A$ with domain $A$.
-To simplify the proof, we extend our grammar to allow first-order constants.
-Consider each tuple $\bar{a}\in A^{\length{\bar{x}}}$, each $\bar{b}\in A^{\length{\bar{y}}}$ and each $\bar{c}\in A^{\length{\bar{z}}}$ as a tuple of first-order constants.
-The following formula defines the same function as $\alpha$:
-$$
-\sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length{\bar{x}}}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length{\bar{y}}}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length{\bar{z}}}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
-$$
-Note that each $\fo$ formula over $(\bar{x},\bar{y},\bar{z})$ in each $\varphi^i_j$ has no free variables.
-Therefore, we can evaluate each of these in polynomial time and replace them by $\perp$ and $\top$ where it corresponds.
-Each $\varphi^i_j$ will be of the form $\perp \vee\, \chi^i_j(\bar{X})$ or $\top \vee \chi^i_j(\bar{X})$ where $\chi^i_j$ is a disjunction of $\neg X_{\ell}$'s and at most one $X_{\ell}$.
-The formulas of the form $\top \vee \chi^i_j(\bar{X})$ can be removed entirely, and the formulas of the form $\perp \vee\, \chi^i_j(\bar{X})$ can be replaced by $\chi^i_j(\bar{X})$.
-We obtain the formula
-$$
-\sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})
-$$
-where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$.
-
-Our idea for the rest of the proof is to define $g_{\alpha}$ for each $\alpha = \sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})$, for $\alpha = c$ and for $\alpha = \beta_1 + \cdots + \beta_m$ where each $\beta_i$ is in one of the two previous cases.
-
-If $\alpha$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ where $\psi_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$, then we obtain the {\bf propositional formula} $g_{\alpha}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulas.
-
-If $\alpha$ is equal to a constant $c$, then we define $g_{\alpha}(\A)$ as the following formula that has exactly $c$ satisfying assignments:
-$$
-g_{\alpha}(\A) = \bigvee_{i = 1}^{c}\neg t_1 \wedge \cdots \wedge \neg t_{i-1} \wedge t_i \wedge \neg t_{i+1} \wedge \cdots \wedge \neg p_c.
-$$ 
-If $\alpha = \beta_1 + \cdots + \beta_m$, let $g_{\beta_i}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\theta^i_{j,k}$ for each $\beta_i$ where each $\theta^i_{j,k}$ is a Horn clause.
-Let $\Theta_i = g_{\beta_i}(\A)$.
-We rename the variables in each $\Theta_i$ so none of them are mentioned in any other $\Theta_j$.
-We add $m$ new variables $t_1,\ldots,t_m$ and we define:
-\begin{align*}
-g_{\alpha}(\A) = &\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^1_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m}}}t) \wedge (t_1 \wedge \bigwedge_{\ell = 2}^{m} \neg t_{\ell})) \vee \\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^2_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each $t$}\\ \text{ mentioned in}\\ \Theta_1,\Theta_3,\ldots,\Theta_{m}}}t) \wedge (t_2 \wedge \bigwedge\limits_{\substack{\ell = 1 \\ \ell \neq 2}}^{m} \neg t_{\ell})) \vee \cdots \vee\\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^m_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m-1}}}t) \wedge (t_m \wedge \bigwedge_{\ell = 1}^{m-1} \neg t_{\ell})).
-\end{align*}
-The formula is a disjunction of Horn formulas, and the number of satisfying assignments for this formula is exactly the sum of satisfying assignments for each $g_{\beta_i}(\A)$.
-This, at the same time, is equal to $\sem{\alpha}(\A)$.
-This covers all possible cases for $\alpha$, and the entire procedure takes polynomial time.

### 2017-03-03 LMCS/proof-qe0-fp-qe1-totp-fptras.tex
journal primera version

--- 2017-01-12 LICS/appendix-syntactic.tex
@@
-\subsection*{Proof of Theorem \ref{theo-pnf-snf}}
-Recall that a formula in $\eqso(\LL)$ is on the following grammar:
-$$
-\alpha = \varphi \ \mid \ s \ \mid \ (\alpha + \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \sa{X} \alpha,
-$$
-where $\varphi$ is a formula in $\LL$ and $s\in\nat$.
-We will construct a recursive function $\tau$ such that for every $\eqso(\LL)$ formula $\alpha$, it outputs an equivalent formula $\tau(\alpha)$ which is in $\LL$-SNF.
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$ which is clearly equivalent and already in $\LL$-SNF.
-If $\alpha = s$, then we define $(\top \add \cdots \add \top)$ ($s$ times), which also satisfies the condition.
-We assume that for every sub-formula $\beta$ of $\alpha$, $\tau(\beta)$ is an equivalent formula in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$, then we define $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$.
-If $\alpha = \sa{x}\beta$, then let $\tau(\beta) = \sum_{i = 1}^{k}\beta_i$ for some $k$ where each $\beta_i$ is in $\LL$-PNF.
-By grouping together the terms in the sum, we define $\tau(\alpha) = \sum_{i = 1}^{m}\sa{x}\beta_i$ which satisfies the condition and is equivalent to $\alpha$.
-If $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case.
-This covers all possible cases for $\alpha$ and we conclude the proof by taking $\tau(\alpha)$ as the desired rewrite of $\alpha$.
-
-\subsection*{Proof of theorem \ref{theo-pi1-pnf}}
-We divide the proof in three parts.
-
-\vspace{1em}
-First, we prove that the formula $\alpha_{0} = \left( \sa{X} 1 \right) + 1$ with $\arity(X) = 1$ (i.e.
-the function $2^{n}+1$, where $n$ is the size of the input structure) is not equivalent to any formula in $\loge{0}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ in $\loge{0}$-PNF that is equivalent to the $\eqso(\loge{0})$ formula $\alpha_0$.
-In \cite{SalujaST95}, it was proved that if $\length(\bar{X}) > 0$, then the function defined by $\alpha$, for big enough structures, is always even which is not possible in our case.
-On the other hand, if $\alpha$ is of the form $\sa{\bar{x}}\varphi(\bar{x})$, then $\alpha$ defines a polynomially bounded function which also leads to a contradiction.
-
-\vspace{1em}
-Second, we prove that the formula $\alpha_{1} = 2$ (i.e.
-the constant $2$) is not equivalent to any formula in $\loge{1}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to the $\eqso(\loge{1})$ formula $2$.
-First, if $\length(\bar{X}) = \length(\bar{x}) = 0$, then the function defined by $\alpha$ is never greater than 1. 
-Therefore, suppose that $\length(\bar{X}) > 0$ or $\length(\bar{x}) > 0$, and consider any ordered structure $\A$. 
-Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
-Now consider the ordered structure $\A'$ that is obtained by duplicating $\A$.
-This is, each half of $\A'$ is isomorphic to $\A$.
-Note that $\A'\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$ for $i = 1,2$ and there exists a third assignments $(\bar{B}_1',\bar{b}_1',\bar{a}_1')$ that is isomorphic to $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$ but in the other half to the structure such that $\A'\models\varphi(\bar{B}_1',\bar{b}_1',\bar{a}_1')$.
-We have that $\sem{\alpha}(\A) \geq 3$ which is a contradiction.
-
-\vspace{1em}
-We now show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there is an equivalent formula $\beta$ in $\LL$-PNF.
-As in Theorem \ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such formula.
-As we showed, there exists an equivalent formula in $\LL$-SNF, so we assume that $\alpha$ is in that form.
-Let $\alpha = \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-SNF.
-Without lost of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.
-If not, we replace each $\alpha_i$ for the equivalent formula $\sa{\bar{X}} \sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge\forall z\,Y(z) \wedge \forall z(y \leq z))$.
-
-Now we begin describing the function $\tau$.
-If $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then the formula is already in $\LL$-PNF so we define $\tau(\alpha) = \alpha$.
-If $\alpha = \alpha_1 + \alpha_2$, then we assume that $\tau(\alpha_1) = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ and $\tau(\alpha_2) = \sa{\bar{Y}}\sa{\bar{y}}\psi(\bar{Y},\bar{y})$.
-The construction that we will provide for this function works by identifying a ``first'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ and a ``last'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$.
-These are identified by the following formulas:
-\begin{align*}
-\gamma_{\text{first}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z}), \\
-\gamma_{\text{last}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x}).
-\end{align*}
-Similarly, we define the formulas $\gamma_{\text{first}}(\bar{Y},\bar{y})$ and $\gamma_{\text{last}}(\bar{Y},\bar{y})$ (for the sake of simplicity we reuse the names $\gamma_{\text{first}}$ and $\gamma_{\text{last}}$).
-In other words, the ``first'' assignment is the one where every second-order predicate is empty and the first-order assignment is the lexicographically smallest, and the ``last'' assignment is the one where every second-order predicate is full and the first-order assignment is the lexicographically greatest.
-We also need to identify assignments that are not first and are not last.
-We do this by negating the two formulas above and grouping together the first-order variables:
-\begin{align*}
-\gamma_{\text{not first}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{z}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{z}_i)), \\
-\gamma_{\text{not last}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{x} < \bar{z}_0 \vee \bigvee_{i = 1}^{\length(\bar{X})}\neg X(\bar{z}_i)),
-\end{align*}
-where $\bar{z} = (\bar{z}_0,\bar{z}_1,\ldots,\bar{z}_{\length(\bar{X})})$. The following formula is equivalent to $\alpha$:
-\begin{align}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{not first}}(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{Y},\bar{y}))\vee \label{eq:partition1} \\
-&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))\vee \label{eq:partition2}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{not last}}(\bar{Y},\bar{y}))\vee \label{eq:partition3}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{last}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))]. \label{eq:partition4}
-\end{align}
-To show that the formula is indeed equivalent to $\alpha$, note that the formulas in lines (\ref{eq:partition1}) and (\ref{eq:partition2}) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$, and the formulas in lines (\ref{eq:partition3}) and (\ref{eq:partition4}) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
-Altogether, the four lines define pairwise disjoint assignments for $(\bar{X},\bar{x}),(\bar{Y},\bar{y})$.
-With this, it is straightforward to show that the above formula is equivalent to $\alpha$.
-However, the formula is not yet in the correct form since it has existential quantifiers in the subformulas $\gamma_{\text{not first}}$ and $\gamma_{\text{not last}}$.
-To solve this, first let take a close look to the complete formula:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\exists\bar{v}(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i))\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\exists\bar{w}(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))].
-\end{align*}
-To construct an equivalent formula that is in the correct form, we define $\bar{u} = (\bar{v},\bar{w})$ and we replace the first-order quantifiers by a first-sum and count the first assignment to $\bar{v}$ and $\bar{w}$ that satisfies the formula.
-A similar construction was used in \cite{SalujaST95}.
-Then the final formula equivalent to $\alpha$ is the following:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length(\bar{X})}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
-& \hspace{3cm}  \bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&
-(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))\vee
-\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
-& \hspace{3cm} (\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length(\bar{Y})}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee \\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length(\bar{X})} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length(\bar{Y})} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))].
-\end{align*}
-Finally, consider a $\eqso(\LL)$ formula $\alpha$ in $\LL$-SNF.
-If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we can consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the condition in the hypothesis.
-
-\subsection*{Proof of Theorem \ref{prop-rest}}
-We give this proof in three parts.
-
-\vspace{1em}
-First, we show that $\QE{0} \not\subseteq \E{1}$.
-By contradiction, suppose that there is a $\QE{0}$ formula $\alpha$ over some signature $\R$ such that defines the following function.
-For every finite $\R$-structure with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n - 1$.
-We use the following claim.
-\begin{claim}
-Let $\alpha = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. Then the function defined by $\alpha$ is either null, greater or equal to $n$, or is in $\Omega(n^2)$.
-\end{claim}
-\begin{proof}
-Suppose that the function defined by $\alpha$ is not $0$ and that $\varphi$ is in DNF.
-Furthermore, suppose $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$.
-Then $\alpha = \sa{\bar{x}} \varphi_1(\bar{x}) \vee \cdots \vee \varphi_n(\bar{x})$.
-Since $\alpha$ is not null, then some $\varphi_i$ must be satisfiable.
-This is, the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is not null.
-We will prove by induction on $\length(\bar{x})$ that the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is either greater or equal to $n$, or in $\Omega(n^2)$.
-We address the case $\length(\bar{x})= 1$, then $\alpha = \sa{x}\bigwedge\psi(x)$.
-If any $\psi(x) = (x = x)$ or $\neg(x < x)$, then we can eliminate it and we obtain the same function.
-If any $\psi = (x < x)$ or $\neg(x=x)$, then the function becomes null.
-If $\psi(x) = R(x,\ldots,x)$ for some $R\in\R$ the function becomes null for the structures we are considering.
-If $\psi(x) = \neg R(x,\ldots,x)$, we can eliminate it and for the structures we are considering we obtain the same function.
-The only possible $\alpha$ left is $\alpha = \sa{x}\top$ which is equal to the function $n$.
-This covers all possible cases for $\length(\bar{x}) = 1$.
-Now suppose that it holds for $\length(\bar{x}) = k$ and suppose $\alpha = \sa{\bar{x}}\bigwedge\psi(\bar{x})$ for $\length(\bar{x}) = k+1$.
-If any $\psi(\bar{x}) = (x_i = x_j)$ where $i \neq j$, then $\alpha$ describes the same function as $\alpha$ where $x_j$ has been replaced by $x_i$.
-In this formula the tuple of first-order variables has $k$ elements so the function it describes if one of the mentioned in the hypothesis.
-If $i = j$, then we can eliminate it and obtain the same function.
-If any $\psi(\bar{x}) = R(\bar{v})$ or $\neg R(\bar{v})$ where $\bar{v}$ is a sub-tuple of $\bar{x}$ then we can either eliminate it or the function becomes null, following the same argument as in the case $\length(\bar{x}) = 1$.
-If any $\psi(\bar{x}) = \neg(x_i = x_j)$ or $(x_i < x_j)$ where $i = j$, then the function becomes null.
-If any $\psi(\bar{x}) = \neg(x_i < x_j)$ where $i = j$, we can eliminate it.
-The remaining formulas in $\bigwedge\psi(\bar{x})$ are either $\neg(x_i = x_j)$, $(x_i<x_j)$ or $\neg(x_i<x_j)$.
-If the formula violates transitivity in $<$ (for example, $x < y \wedge y < z \wedge z < x$), then the function $\alpha$ describes is null.
-Therefore, there is some order over $\bar{x}$ that satisfies $\bigwedge\psi(\bar{x})$.
-Consider the formula that describes this order (like $x_1 < x_3 \wedge x_3 < x_4 \wedge x_4 < x_2$).
-The function $\alpha$ describes is greater or equal to the one this formula describes, which is exactly $\binom{n}{\length(\bar{x})}$ which is in $\Omega(n^{\length(\bar{x})}) \subseteq \Omega(n^2)$ if $\length(\bar{x}) > 1$.
-This concludes the proof of the claim.
-\end{proof}
-We suppose that $\alpha$ is in SNF, this is, $\alpha = \sum_{i = 1}^n\alpha_i$.
-Since $\alpha$ is not null, consider some $\alpha_i$ that describes a non-null function.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, where $\varphi$ is quantifier-free.
-Note that if $\length(\bar{X}) > 0$, then the function $\alpha$ describes is in $\Omega(2^n)$, as it was proven by the authors in \cite{SalujaST95}.
-We have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$, as we proved in the claim, describes either some function greater or equal to $n$, or in $\Omega(n^2)$, which leads to a contradiction.
-Lastly, note that the formula $\sa{x}\exists y(x < y)$ is in $\E{0}$ and describes the function $n-1$, which concludes the proof.
-
-\vspace{1em}
-Now we show that $\E{1}\not\subseteq\QE{0}$.
-In Theorem \ref{theo-pi1-pnf} we proved that there is no formula in $\loge{1}$-PNF equivalent to the formula $\alpha = 2$.
-Every formula in $\E{1}$ can be expressed in $\loge{1}$-PNF, which implies that $2 \in \QE{0}$ and $2 \not\in \E{1}$.
-
-\vspace{1em}
-Lastly, we prove that $\eqso(\loge{1})\subsetneq\eqso(\logu{1})$.
-For inclusion, let $\alpha$ be a formula in $\eqso(\loge{1})$.
-Suppose that it is in $\loge{1}$-SNF.
-This is, $\alpha = c + \sum_{i = 1}^{n}\alpha_i$.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, where $\varphi_i$ is quantifier-free, for each $\alpha_i$.
-We use the same construction used in \cite{SalujaST95}, and we obtain that the formula $\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \forall\bar{y}'(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')$ for every assignment to $(\bar{X},\bar{x})$.
-We do this replacement for each $\alpha_i$ and we obtain an equivalent formula in $\eqso(\logu{1})$.
-
-To prove that the inclusion is proper, consider the $\eqso(\logu{1})$ formula $\sa{x}\forall y(y = x)$. This formula defines the following function that takes an ordered structure $\A$ as input:
-$$
-\sem{\alpha}(\A) = 
-\begin{cases}
-1 &\A \text{ has one element}\\
-0 &\text{ otherwise}.
-\end{cases}
-$$
-Suppose that there exists an equivalent formula $\alpha$ in $\eqso(\loge{1})$.
-Also, suppose that it is in $\L$-PNF, so $\alpha = c + \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$.
-Since $\alpha$ takes the value 0 for some structures, $c$ must be 0.
-Consider a structure $\mathfrak{1}$ with one element.
-We have that for some $i$, there exists an assignment $(\bar{B},\bar{b},\bar{a})$ for $(\bar{X},\bar{x},\bar{y})$ such that $\mathfrak{1}\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
-Consider now the structure $\mathfrak{2}$ that is obtained by duplicating $\mathfrak{1}$, as we did for Theorem \ref{theo-pi1-pnf}.
-Note that $\mathfrak{2}\models\varphi_i(\bar{B},\bar{b},\bar{a})$, which implies that $\sem{\alpha}(\mathfrak{2}) \geq 1$, which leads to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop-e1-nc}}
-
-Towards a contradiction, assume that the statement is false.
-This is, that $\E{1}$ is closed under binary sum.
-Consider the formula $\sa{x}(x = x)$ which is in $\E{1}$ over some signature $\R$.
-For every finite $\R$-structure $\A$ with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n$.
-From our assumption, there exists some formula in $\E{1}$ equivalent to the formula $\alpha \add \alpha$, which describes the function $2n$.
-Let $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ be this formula, where $\varphi$ is quantifier-free.
-Note that the function defined by this formula is equal or greater than the one defined by $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ divided by a polynomial factor.
-More specifically, for each ordered structure $\A$ with domain $A$, we have the following inequality:
-$$
-\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot \vert A \vert^{\length(\bar{y})} \geq \sem{\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A)
-$$
-Note that the formula $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ defines a function in $\E{0}$.
-It was shown by the authors in \cite{SalujaST95} that every function in $\E{0}$ grows exponentially over the size of the structure for large enough structures, when $\length(\bar{X}) > 0$.
-This function divided by a polynomial factor still grows exponentially.
-Therefore, for $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ we have that $\length(\bar{X}) = 0$.
-
-Now, for the formula $\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ consider a structure $\mathfrak{1}$ with only one element $a$.
-We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$, but the only possible assignment to $\bar{x}$ is the tuple $(a,\ldots,a)$ so $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which follows to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop:qe0-fp-qe1-totp-fptras}}
-
 The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$.
@@
 Since $\loge{1}$ is closed under conjunction, this also holds for $\eqso(\loge{1})$. This concludes the proof.
-
-\subsection*{Proof of Proposition \ref{pi-minusone}}
-
-Let $\LL$ be a fragment of $\fo$ that contains $\logu{1}$.
-Then we have that every function in $\U{1}$ is expressible in $\eqso(\LL)$.
-In particular, $\ctcnf \in \eqso(\LL)$.
-Suppose that $\eqso(\LL)$ is closed under subtraction by one.
-Then, the function $\ctcnf-1$, which counts the number of satisfying assignments of a 3-CNF formula minus one, is also in $\eqso(\LL)$.
-Note that $\eqso(\LL) \subseteq \eqso(\fo) = \shp$.
-We have that $\ctcnf$ is $\shp$-complete under parsimonious reductions\footnote{It can be easily verified that the standard reduction from SAT to 3-CNF (or 3-SAT) preserves the number of satisfying assignments}.
-Now, let $f$ be a function in $\shp$, and consider the nondeterministic polynomial-time procedure that on input $\enc(\A)$ computes the corresponding reduction to $\ctcnf$, name it $g(\enc(\A))$, and simulates the $\shp$ procedure for $\ctcnf-1$ on input $g(\enc(\A))$.
-We have that this is a $\shp$ procedure that computes $f-1$, from which we conclude that $\shp$ is closed under subtraction by one.
-
-\subsection*{Proof of Theorem \ref{sigmafo-minusone}}
-
-\noindent {\bf Closed under sum and multiplication.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
-This can be seen because $\eqso(\LL)$ is closed under sum for every fragment $\LL$ by definition, and since $\logex{1}$ is closed under conjunction, from Lemma \ref{conj-mult} it follows that $\eqso(\logex{1})$ is closed under multiplication.
-
-\vspace{1em}
-\noindent {\bf Easy decision version and FRPRAS.}  We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
-We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Theorem \ref{prop:qe0-fp-qe1-totp-fptras}.
-First, we define a function that converts a formula $\alpha$ in $\eqso(\logex{1})$ over a signature $\R$ into a formula $\lambda(\alpha)$ in $\eqso(\loge{1})$ over a signature $\R_{\alpha}$.
-Afterwards, we define a function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an $\R_{\alpha}$-structure $g_{\alpha}(\A)$.
-
-Let $\alpha$ be in $\eqso(\logex{0})$.
-The signature $\R_{\alpha}$ is obtained by adding the symbol $R_{\psi}$, for every $\fo$ formula $\psi(\bar{z})$ in $\alpha$, to $\R$.
-Each symbol $R_{\psi}$ represents a predicate with arity $\length(\bar{z})$.
-Then, $\lambda(\alpha)$ is defined as $\alpha$ where each $\fo$ formula $\psi(\bar{z})$ has been replaced by $R_{\psi}(\bar{z})$.
-We now define the function $g_{\alpha}$ procedurally.
-Let $\A$ be a $\R$-structure with domain $A$.
-Let $\A'$ be an $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A}$ empty.
-For each $\fo$-formula $\psi(\z)$ with $\length(\bar{z})$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length(\bar{z})}$.
-If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$ (this can be done in $\ptime$).
-This concludes the construction of $\A'$.
-Note that the number of $\fo$ subformulas, arity and tuple size is fixed in $\alpha$, so computing this function takes polynomial time over the size of the structure.
-Moreover, the encoding of $\A'$ has polynomial size over the size of $\enc(\A)$.
-We define $g_{\alpha}(\A) = \A'$ and we have that for each $\R$-structure $\A$: $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$.
-Therefore, we have a parsimonious reduction from $\alpha$ to the $\eqso(\loge{1})$ formula $\lambda(\alpha)$.
-
-To show that $\alpha$ is in $\totp$, we can convert $\alpha$ and $\enc(\A)$ into $\lambda(\alpha)$ and $\enc(g_{\alpha}(\A))$, respectively, and run the procedure in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-Similarly, to show that $\alpha$ has an FPRAS, we do the same as before and simulates the FPRAS for $\lambda(\alpha)$ in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-These procedures also takes polynomial time and satisfies the condition.
-
-\vspace{1em}
-\noindent {\bf Closed under subtraction by one.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. 
-For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Without lost of generality, we assume that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
-Moreover, we assume that $\length(\bar{x}) > 0$  since, if this is not the case, we can replace $\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
-
-The proof will be separated in two parts.
-In the first part, we suppose that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$.
-Then we will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ is the formula.
-In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ the sum of one or more formulas in $\logex{1}$-PNF.
-We define a formula $\varphi'$ that satisfies the following condition: if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ and $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be proven.
-
-\vspace{1em}
-
-\noindent {\em Part (1).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ where $\varphi$ is an  $\fo$-formula.
-Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}} \varphi(\bar{x})$ (i.e.
-$\length(\bar{X}) = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}} [\varphi(\bar{x})\wedge\exists\bar{z}(\varphi(\bar{z})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
-Therefore, for the rest of the proof we can assume that $\length(\bar{X}) > 0$ and $\length(\bar{x}) > 0$.
-
- Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$.
-To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in into a DNF formula. 
-More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i$ is a conjunction of atomic formulas or negation of atomic formulas.
-Furthermore, we suppose that each conjunct $\varphi_i$ has the~form:
-$$
-\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
-\underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}} \wedge
-\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}}.
-$$
-Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ can appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
-
-Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ are mentioned in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
-We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
-\begin{itemize}
-	\item[(a)] In order to satisfy the first condition, consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
-We add $\length(\bar{w})$ new variables $z_1,\ldots,z_{\length(\bar{w})}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length(\bar{w})})$.
-We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$.
-The formula $\varphi_i$ is now defined as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \exists \bar{z} \; (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{-}(\bar{X},\bar{x},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y})).
-	$$
-	We repeat this process for each instance of a $X_j(\bar{w})$ in $\varphi_i$, and we obtain a formula where none of the $X_j$'s acts over any variable in $\bar{x}$.
-We add the new first-order variables to $\bar{y}$ and we redefine $\varphi_i$ as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
-	$$
-	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = \exists \bar{y} \;  X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := \exists \bar{y} \;  X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
-	\item[(b)] An ordered partition on a set $S$ is defined by an equivalence relation $\sim$ over $S$, and a linear order over $S/\!\sim$.
-For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$.
-A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
-On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
-	For a given $k$, let $\cB_k$ be the number of possible ordered partitions for a set of size $k$. For $1 \leq j \leq \cB_{\length(\bar{x},\bar{y})}$ 
-	let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
-
-	Thus, the formula $\varphi_i(\bar{X},\bar{x})$ is then redefined as:
-	$$
-	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}} \exists \bar{y} \; [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
-	$$
-	Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
-	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  to $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
-	\item[(c)] For showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
-$\varphi_i$ is inconsistent), we do the following.
-If there exists an instance of $X_j(\bar{z})$ in $\varphi^{+}_i$, an instance of $\neg X_j(\bar{w})$ in $\varphi^{-}_i$ and the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$, then the entire formula $\varphi_i$ is removed from $\varphi$.
-\end{itemize}
-It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$.
-From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c).
-Furthermore, it has the following form:
-$$
-\varphi_i(\bar{X},\bar{x}) \; = \; \exists{\bar{y}} \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
-$$
-where $\varphi^{+}$ and $\varphi^{-}_i$ do not depend on  $\bar{x}$.
-\begin{claim}\label{claim:minusone}
-For a given ordered structure $\A$, if there exists a first-order assignment $v$ for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists a second-order assignment $V$ for $\A$ such that  $(\A,V,v)\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-\end{claim}
-\begin{proof}
-Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
-Define $\bar{B} = (B_1,\ldots,B_{\length(\bar{X})})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
-Towards a contradiction, suppose that $(\A,V,v)\not\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-By the choice of $v$, and construction of $V$ it is clear that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})\wedge\varphi^{+}_i(\bar{X},\bar{y})$, so we necessarily have that $(\A,V,v)\not\models\varphi^{-}_i(\bar{X},\bar{y})$.
-Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,V, v)\not\models\neg X_t(\bar{w})$, namely, $v(\bar{w})\in B_t$. 
-However, by the construction of $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$.
-Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$.
-This violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
-\end{proof}
-
-The previous claim and proof motivates the following definitions.
-For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length(\bar{X})})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
-One can easily check that for every assignments $V, v$ such that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$, then it holds that $(\A,\bar{B}^v,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all the assignment of $\bar{X}$ when $v$ is fixed.
-Therefore, by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
-For this, the idea is to choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
-$$
-\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge\forall\bar{x}'\forall\bar{y}'(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')).
-$$ 
-If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
-Furthermore, let $V^*$ the second order assignment that satisfies $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*$ the first order assignment such that $v^*(\bar{x}) = v(\bar{x})$.
-By the previous discussion, we have that $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.
-
-Now, we have all the tools to introduce the main formula of the proof. 
-Intuitively, we want to exclude the assignment $(V^*, v^*)$ as a satisfying assignment for $\varphi_i(\bar{X},\bar{x})$.
-Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if, 
-if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, then $V \neq V^*$ or $v \neq v^*$. 
-That is, it excludes all pairs $(V, v)$ that coincide with $(V^*, v^*)$ in $(\bar{X}, \bar{x})$ whenever $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. This property can be defined as follows:
-\begin{align}
-\psi_i(\bar{X},\bar{x}) = & \;\; \big( \, \exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
-& \hspace{1cm} \Big( \; 
-\exists\bar{y} \;\; \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i^+(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\
-& \hspace{1.5cm} (\exists\bar{x}'\exists\bar{y}\, \varphi_i(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}) \;\; \Big)  \label{eq:part3}
-\end{align}
-To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
-Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-Then, the conclusion of the implication at (\ref{eq:part2}) and (\ref{eq:part3}), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
-Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by showing that if $\varphi_i(\bar{X},\bar{x})$ is satisfied then $V(\bar{X})$ must contain one element more that it is not in $V^*(\bar{X})$.
-The second disjunct (\ref{eq:part3}) is satisfied when $v(\bar{x})$ is not the lexicographically smallest tuple that satisfies $\varphi_i$ (i.e. $v(\bar{x}) \neq v^*(\bar{x})$).
-
-We are ready to define the formula  $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$
-where each modified disjunct $\varphi_i'(\bar{X},\bar{x})$ is constructed as follows. 
-For the sake of simplification, define the auxiliary formula $\chi_i = \neg\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$. 
-This formula basically checks if $\varphi_i$ is satisfiable or not.
-Then, first define the formula:
-$$
-\varphi'_1(\bar{X},\bar{x}) = \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
-$$
-This models that all the assignments that satisfy $\varphi_1$ also satisfy $\varphi'_1$, except for $(V^*,v^*)$. Then we define:
-$$
-\varphi'_2(\bar{X},\bar{x}) = \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
-$$
-This models that all the assignments that satisfy $\varphi_2$ also satisfy $\varphi'_2$, but if there was no assignment for $\varphi_1$, then we exclude $(V^*,v^*)$ for $\varphi_2$.
-Then we generalize this construction for each $\varphi_i$:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})),
-$$
-By the construction of $\kappa(\alpha)$, one can easily check that  $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.
-
-\vspace{1em}
-
-{\em Part (2).} Let $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ for some $\eqso(\logex{1})$ formula $\beta$. We define $\kappa(\alpha)$ as follows.
-First, perform the same transformations to $\varphi(\bar{X},\bar{x})$ as in the previous section.
-Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c).
-We also use the previously defined formulas $\chi_i$ and $\psi_i$, for each $i$.
-
-We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if and only if $\sem{\beta}(\A) = 0$.
-If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg\exists\bar{x}'\varphi(\bar{x}')$.
-If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$ as each $\chi_i$ was previously defined.
-If $\beta = (\beta_1 + \beta_2)$, then $\lambda(\beta) = \lambda(\beta_1) \wedge \lambda(\beta_2)$.
-Now, for each $\varphi_i$ we define:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x}))\Big)\Big).
-$$
-Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
-
-%\subsection*{Proof of Theorem \ref{sub-pnp}}
-%
-%Towards a contradiction, suppose that any of the classes $\E{1}$, $\eqso(\loge{1})$, and $\eqso(\logex{1})$ is closed under subtraction.
-%
-%Let $\R$ include the symbols $S_1, S_2, S_3, S_4$, which describe the following properties.
-If a finite $\R$-structure $\A$ defines a 3DNF formula $\Phi$, then its domain is the set of variables mentioned in $\Phi$, and for each $i = 1,2,3,4$:
-%\begin{align*}
-%	S_1^\A &= \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_2^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_3^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_4^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
-%\end{align*}
-%Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
-%\begin{multline*}
-%\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
-%\end{multline*}
-%Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
-%$$
-%\varphi(x,X) = (T(x) \vee \neg T(x)).
-%$$
-%Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Given that $f_{\#3DNF}, f_{all} \in \E{1} \subseteq \eqso(\loge{1}) \subseteq \eqso(\logex{1}) \subseteq \totp$, and at least one of the classes is closed under subtraction, the function $f_{\#3DNF} - f_{all}$ is in $\totp$.
-However, note that for each $\R$-structure $\A$ that represents a formula $\Phi$, $(f_{\#3DNF} - f_{all})(\A) = 0$ if an only if $\Phi$ is a tautology.
-The decision version of this function is the $\np$-complete problem $\textsc{Tautology}$.
-And since the function is in $\totp$, its decision version is also in $\ptime$.
-We conclude that $\np \subseteq \ptime$.
-
-\subsection*{Proof of Proposition \ref{prop:ehorn-pe}}
-Pagourtzis and Zachos mention a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula \cite{PagourtzisZ06}.
-This procedure can be easily extended to receive Horn formulas, and furthermore, a disjunction of Horn formulas.
-Hence $\shdhsat$ is in $\totp$.
-
-As we show in Proposition \ref{sigma2hard}, $\shdhsat$ is complete for $\eqso(\ehorn)$ under parsimonious reductions.
-Let $\alpha$ be a formula in $\eqso(\ehorn)$ and let $g_{\alpha}$ be the reduction to $\shdhsat$.
-The $\totp$ procedure we construct, for each input $\enc(\A)$, is simply to compute $g_{\alpha}(\enc(\A))$, and then simulate the $\totp$ procedure for $\shdhsat$ on input $g_{\alpha}(\enc(\A))$.
-We conclude that $\alpha$ is in $\totp$.
-
-\subsection*{Proof of Proposition \ref{prop:hsat-not-sigma2}} %V.12
-We use a similar proof to the one provided by the authors in \cite{SalujaST95} to separate the classes $\E{2}$ and $\U{2}$.
-Suppose that the statement is false, this is, $\chsat \in \eqso(\loge{2})$.
-We consider the signature $\R$ that we used as the encoding for a Horn formula (Example \ref{ex-hornsat-esop1}) and that the formula $\alpha \in \eqso(\loge{2})$ follows the encoding in the same way.
-From what we proved in Theorem \ref{theo-pnf-snf}, we have that every formula in $\eqso(\loge{2})$ can be rewritten in $\loge{2}$-PNF, so we assume that $\alpha$ is in this form.
-Let $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\forall\bar{z}\,\varphi(\bar{X},\bar{x},\bar{y},\bar{z})$.
-Consider the following Horn formula $\Phi$:
-$$
-\Phi = p \wedge \bigwedge_{i = 1}^n (t_i \wedge p \to q) \wedge \neg q,
-$$
-where $n = \length(\bar{x}) + \length(\bar{y}) + 1$.
-Let $\A$ be the encoding of this formula.
-In the encoding, each variable appears as an element in the domain of $\A$.
-This formula is satisfiable, so $\sem{\alpha}(\A) \geq 1$.
-Let $(\bar{B},\bar{b},\bar{a})$ be an assignment to $(\bar{X},\bar{x},\bar{y})$ such that $\A\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-Let $t_{\ell}$ be such that it does not appear in $\bar{b}$ or $\bar{a}$.
-Consider the induced substructure $\A'$ that is obtained by removing $t_{\ell}$ from $\A$ and $\bar{B}'$ as the subset of $\bar{B}$ obtained by deleting each appearance of $t_{\ell}$ in $\bar{B}$.
-We have that $\A'\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-This is because each subformula of the form $\exists y \neg B_i$ is still true, and universal formulas are monotone over induced substructures.
-It follows that $\sem{\alpha}(\A') \geq 1$ which is not possible since $\A'$ encodes the formula
-$$
-\Phi' = p \wedge \bigwedge_{i = 1}^{\ell-1} (t_i \wedge p \to q) \wedge (p\to q) \wedge \bigwedge_{i = \ell+1}^{n} (t_i \wedge p \to q) \wedge \neg q,
-$$
-which is unsatisfiable. We arrive to a contradiction and we conclude that $\chsat$ is not in $\eqso(\ehorn)$.
-
-\subsection*{Proof of Theorem \ref{sigma2hard}} %V.13
-
-First we prove that $\shdhsat$ is in $\eqso(\ehorn)$.
-Recall that each instance of $\shdhsat$ is a disjunction of Horn formulas.
-Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot), \pD(\cdot,\cdot)\}$.
-Each symbol in this vocabulary is used to indicate the same as in Example \ref{ex-hornsat-esop1}, with the addition of $\pD(d,c)$ which indicates that $c$ is a clause in the formula $d$.
-Recall that the formula
-\begin{align*}
-&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \neg\textit{T}(x)).
-\end{align*}
-defines $\chsat$. We obtain the following formula $\psi(T,A)$ in $\ehorn$:
-\begin{align*}
-\exists d[&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \pD(c,d)\vee \neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \neg\textit{T}(x))]
-\end{align*}
-effectively defines $\chsat$ as for every disjunction of Horn formulas $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
-Therefore, we conclude that $\shdhsat \in \eqso(\ehorn)$.
-
-\vspace{1em}
-We will now prove that $\shdhsat$ is hard for $\eqso$ over a signature $\R$ under parsimonious reductions.
-For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a polynomial-time procedure that computes a function $g_{\alpha}$.
-This function receives a finite $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
-We suppose that $\alpha$ is in sum normal form and:
-$$
-\alpha = \sum_{i = 1}^{\text{\#clauses}} \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigwedge_{j = 1}^{n}\forall\bar{z}\,\varphi^i_j(\bar{X},\bar{x},\bar{y},\bar{z}),
-$$
-where each $\varphi^i_j$ is a Horn clause.
-
-Consider a finite $\R$-structure $\A$ with domain $A$.
-To simplify the proof, we extend our grammar to allow first-order constants.
-Consider each tuple $\bar{a}\in A^{\length(\bar{x})}$, each $\bar{b}\in A^{\length(\bar{y})}$ and each $\bar{c}\in A^{\length(\bar{z})}$ as a tuple of first-order constants.
-The following formula defines the same function as $\alpha$:
-$$
-\sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length(\bar{x})}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length(\bar{y})}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length(\bar{z})}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
-$$
-Note that each $\fo$ formula over $(\bar{x},\bar{y},\bar{z})$ in each $\varphi^i_j$ has no free variables.
-Therefore, we can evaluate each of these in polynomial time and replace them by $\perp$ and $\top$ where it corresponds.
-Each $\varphi^i_j$ will be of the form $\perp \vee\, \chi^i_j(\bar{X})$ or $\top \vee \chi^i_j(\bar{X})$ where $\chi^i_j$ is a disjunction of $\neg X_{\ell}$'s and at most one $X_{\ell}$.
-The formulas of the form $\top \vee \chi^i_j(\bar{X})$ can be removed entirely, and the formulas of the form $\perp \vee\, \chi^i_j(\bar{X})$ can be replaced by $\chi^i_j(\bar{X})$.
-We obtain the formula
-$$
-\sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})
-$$
-where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$.
-
-Our idea for the rest of the proof is to define $g_{\alpha}$ for each $\alpha = \sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})$, for $\alpha = c$ and for $\alpha = \beta_1 + \cdots + \beta_m$ where each $\beta_i$ is in one of the two previous cases.
-
-If $\alpha$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ where $\psi_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$, then we obtain the {\bf propositional formula} $g_{\alpha}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulas.
-
-If $\alpha$ is equal to a constant $c$, then we define $g_{\alpha}(\A)$ as the following formula that has exactly $c$ satisfying assignments:
-$$
-g_{\alpha}(\A) = \bigvee_{i = 1}^{c}\neg t_1 \wedge \cdots \wedge \neg t_{i-1} \wedge t_i \wedge \neg t_{i+1} \wedge \cdots \wedge \neg p_c.
-$$ 
-If $\alpha = \beta_1 + \cdots + \beta_m$, let $g_{\beta_i}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\theta^i_{j,k}$ for each $\beta_i$ where each $\theta^i_{j,k}$ is a Horn clause.
-Let $\Theta_i = g_{\beta_i}(\A)$.
-We rename the variables in each $\Theta_i$ so none of them are mentioned in any other $\Theta_j$.
-We add $m$ new variables $t_1,\ldots,t_m$ and we define:
-\begin{align*}
-g_{\alpha}(\A) = &\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^1_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m}}}t) \wedge (t_1 \wedge \bigwedge_{\ell = 2}^{m} \neg t_{\ell})) \vee \\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^2_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each $t$}\\ \text{ mentioned in}\\ \Theta_1,\Theta_3,\ldots,\Theta_{m}}}t) \wedge (t_2 \wedge \bigwedge\limits_{\substack{\ell = 1 \\ \ell \neq 2}}^{m} \neg t_{\ell})) \vee \cdots \vee\\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^m_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m-1}}}t) \wedge (t_m \wedge \bigwedge_{\ell = 1}^{m-1} \neg t_{\ell})).
-\end{align*}
-The formula is a disjunction of Horn formulas, and the number of satisfying assignments for this formula is exactly the sum of satisfying assignments for each $g_{\beta_i}(\A)$.
-This, at the same time, is equal to $\sem{\alpha}(\A)$.
-This covers all possible cases for $\alpha$, and the entire procedure takes polynomial time.

### 2017-03-03 LMCS/proof-rqfo-fo-cap.tex
journal primera version

Given the definition of the semantics of $\rqfo(\fo)$, it is clear that a fixed formula $\clfp{\beta(\x, h)}$ can be evaluated in polynomial time, from which it is possible to conclude that each fixed formula in $\rqfo(\fo)$ can be evaluated in polynomial time. Thus, to prove that $\rqfo(\fo)$ captures $\fp$, we only need to prove the second condition in Definition \ref{def:cap}.

\newcommand{\ttB}{\mathtt{B}}
\newcommand{\successor}{\text{succ}}

%For the second condition, 
Let $f$ be a function in $\fp$. We address the case when $f$ is defined for the encodings of the structures of a relational signature $\R = \{ E(\cdot, \cdot) \}$, as the proof for an arbitrary signature is analogous.
%$\R$ contains only one binary predicate $E$, and the remaining cases can be deduced from this. 
 Let $M$ be a deterministic polynomial-time TM with a working tape and an output tape, such that the output of $M$ on input $\enc(\A)$ is $f(\enc(\A))$ for each $\R$-structure $\A$. We assume that $M = (Q,\{0,1\},q_0,\delta)$, 
 %without final states, 
 where $Q = \{q_0,\ldots,q_{\ell}\}$, and $\delta:Q\times\{0,1,\ttB, \vdash\}\to Q\times\{0,1,\ttB, \vdash\}\times \{\leftarrow,\rightarrow\}\times\{0,1,\emptyset\}$ is a partial function. In particular, the tapes of $M$ are infinite to the right so the symbol $\vdash$ is used to indicate the first position in each tape, and $M$ does not have any final states, as it produces an output for each input. Moreover, the only allowed operations in the output tape are: 
 %The machine has an output tape and the only allowed operations in that tape on each step are 
 (1) writing 0 and moving the head one cell to the right, (2) writing 1 and moving the head one cell to the right, or (3) doing nothing. These operations are represented by the set $\{0,1,\emptyset\}$. Finally, assume that $M$, on input $\enc(\A)$ with domain $A = \{1,\dots,n\}$, executes exactly $n^k$ steps for some $k \geq 1$.

We construct a formula $\alpha$ in an extension of the grammar of $\rqfo(\fo)$ such that $\sem{\alpha}(\A) = f(\enc(\A))$ for each $\R$-structure $\A$. This extension allows defining the operator ${\bf lsfp}$ for multiple functions, analogously to the notion of simultaneous LFP~\cite{L04}.
%generalization of fixed-point operator with multiple predicates. 
Let $\bar{x} = (x_1,\ldots,x_k)$ and $\bar{t} = (t_1,\ldots,t_k)$. Then $\alpha$ is defined as:
\begin{align*}
\alpha = \sa{\bar{t}}\clfp{out(\bar{t}): \,&\alpha_{T_0}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out),\\
	&\alpha_{T_1}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out),\\
	&\alpha_{T_{\ttB}}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out),\\
	&\alpha_{T_{\vdash}}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out),\\
	&\alpha_{h}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out),\\
	&\alpha_{\hat{h}}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out),\\
	&\alpha_{s_{q_0}}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out),\\
	&\vdots \\
	&\alpha_{s_{q_{\ell}}}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out),\\
	&\alpha_{out}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out)}\mult \last(\bar{t}).
\end{align*}
Function $T_0$ is used to indicate whether the content of a cell of the working tape is 0 at some point of time, that is, $T_0(\bar{t},\bar{x}) > 0$ if the cell at position $\bar{x}$ of the working tape contains the symbol 0 at step $\bar{t}$, and $T_0(\bar{t},\bar{x}) = 0$ otherwise. Functions $T_1$, $T_{\ttB}$ and $T_{\vdash}$ are defined analogously. Function $h$ is used to indicate whether the head of the working tape is in some position at some point of time, that is, $h(\bar{t},\bar{x}) > 0$ if the head of the working tape is at position $\bar{x}$ at step $\bar{t}$, and $h(\bar{t},\bar{x}) = 0$ otherwise. 
Function $\hat{h}$ is used to indicate whether the head of the working tape is {\bf not} in some position at some point of time, that is, $\hat{h}(\bar{t},\bar{x}) > 0$ if the head of the working tape is {\bf not} at position $\bar{x}$ at step $\bar{t}$, and $h(\bar{t},\bar{x}) = 0$ otherwise. For each $i \in \{0, \ldots, \ell\}$, function 
$s_{q_i}$ is used to indicate whether the TM $M$ is in state $q_i$ at some point of time, that is, $s_{q_i}(\bar{t}) > 0$ if the TM $M$ is in state $q_i$ at step $\bar{t}$, and $s_{q_i}(\bar{t},\bar{x}) = 0$ otherwise. Finally, function $out$ stores the output of the TM $M$; in particular, $out(\bar t)$ is the value returned by $M$ when $\bar t$ is the last step (that is, when $\last(\bar t)$ holds).

Formulas $\alpha_{T_0}$, $\alpha_{T_1}$, $\alpha_{T_{\ttB}}$ and $\alpha_{T_{\vdash}}$ are defined as follows, assuming that $\bar y = (y_1, \ldots, y_k)$:
\begin{align*}
\alpha_{T_0}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\ 
&(\first(\bar{t}) \wedge \exists\bar{y}(\first(y_1,\ldots,y_{k-2})\wedge\neg E(y_{k-1},y_k) \wedge \successor(\bar{y},\bar{x}) ))+ \\
&\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult \hat{h}(\bar{t}',\bar{x}) \mult T_0(\bar{t}',\bar{x})) + \\
&\bigplus_{\delta(q,a) = (q',0,op,v)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')),\\
\alpha_{T_1}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\
&(\first(\bar{t}) \wedge \exists\bar{y}(\first(y_1,\ldots,y_{k-2})\wedge E(y_{k-1},y_k) \wedge \successor(\bar{y},\bar{x}) ))+ \\
&\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult \hat{h}(\bar{t}',\bar{x}) \mult T_1(\bar{t}',\bar{x})) + \\
&\bigplus_{\delta(q,a) = (q',1,op,v)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')),\\
\alpha_{T_{\ttB}}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\
&(\first(\bar{t})\wedge\exists\bar{y}\exists\bar{y}'(\first(y_1,\ldots,y_{k-2})\wedge\last(y_{k-1},y_k)\wedge\successor(\bar{y},\bar{y}')\wedge\bar{y}' < \bar{x})) + \\
&\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult \hat{h}(\bar{t}',\bar{x}) \mult T_{\ttB}(\bar{t}',\bar{x})) + \\
&\bigplus_{\delta(q,a) = (q',{\ttB},op,v)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')),\\
\alpha_{T_{\vdash}}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\
&(\first(\bar{t}) \wedge \first(\bar{x})) +
\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult \hat{h}(\bar{t}',\bar{x}) \mult T_{\vdash}(\bar{t}',\bar{x})) + \\
&\bigplus_{\delta(q,a) = (q',\vdash,op,v)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')).
\end{align*}
Formulas $\alpha_{h}$ and $\alpha_{\hat{h}}$ are defined as:
\begin{align*}
\alpha_{h}(&\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\
& (\first(\bar{t}) \wedge \successor(\bar{t},\bar{x})) + \\
&\bigplus_{\delta(q,a) = (q',b,\leftarrow,v)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult \successor(\bar{x},\bar{x}')\mult T_a(\bar{t}',\bar{x}') \mult h(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')) + \\
&\bigplus_{\delta(q,a) = (q',b,\rightarrow,v)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult \successor(\bar{x}',\bar{x})\mult T_a(\bar{t}',\bar{x}') \mult h(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')),\\
\alpha_{\hat{h}}(&\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\
& (\first(\bar{t}) \wedge \neg\successor(\bar{t},\bar{x})) + \\
&\bigplus_{\delta(q,a) = (q',b,\leftarrow,v)}\sa{\bar{t}'}\sa{\bar{x}'}\sa{\bar{x}''}(\successor(\bar{t}',\bar{t}) \mult T_a(\bar{t}',\bar{x}')  \mult h(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult \successor(\bar{x}'',\bar{x}') \mult (\bar{x} \neq \bar{x}'')) + \\
&\bigplus_{\delta(q,a) = (q',b,\rightarrow,v)}\sa{\bar{t}'}\sa{\bar{x}'}\sa{\bar{x}''}(\successor(\bar{t}',\bar{t}) \mult T_a(\bar{t}',\bar{x}')  \mult h(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult \successor(\bar{x}',\bar{x}'') \mult (\bar{x} \neq \bar{x}'')).
\end{align*}
Formula $\alpha_{q_0}$ is defined as:
\begin{align*}
\alpha_{q_0}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \first(\bar{t}) + \\
&\bigplus_{\delta(q,a) = (q_0,b,op,v)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult T_a(\bar{t}',\bar{x}') \mult h(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')).
\end{align*}
Moreover, for every $i \in \{1, \ldots, \ell\}$, formula $\alpha_{q_i}$ is defined as:
\begin{align*}
\alpha_{q_i}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\ &\bigplus_{\delta(q,a) = (q_i,b,op,v)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult T_a(\bar{t}',\bar{x}') \mult h(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')).
\end{align*}
Finally, formula $\alpha_{out}$ is defined as:
\begin{align*}
\alpha_{out}(\bar{t},&T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) =\\
&\bigplus_{\delta(q,a) = (q',b,op,0)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult 2\mult out(\bar{t}')) + \\
&\bigplus_{\delta(q,a) = (q',b,op,1)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult (2\mult out(\bar{t}')+1)) + \\ 
&\bigplus_{\delta(q,a) = (q',b,op,\emptyset)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult out(\bar{t}')).
\end{align*}
Clearly, at each iteration of the LSFP operator, the tuple $\bar{t}$ represents the step the machine is currently in. From the construction of the formula, and since the machine is deterministic, it can be seen that in each function $g\in\{T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h}\}$, at the $\bar{a}$-th iteration of the LSFP operator, it holds that $g(\bar{a},\bar{b}) \leq 1$ for each $\bar{b}\in A^k$, that $g(\bar{a}+1,\bar{b}) = 0$ for each $\bar{b}\in A^k$. Also, at the $\bar{a}$-th iteration, $g(\bar{a}) \leq 1$ and $g(\bar{a}+1) = 0$ for each $g\in\{s_{q_1},\ldots,s_{q_{\ell}}\}$. From this, we have that at each iteration $\bar{a}$ of the operator, $out(\bar{a})$ is equal to either $2\cdot out(\bar{a}-1)$, $2\cdot out(\bar{a}-1) + 1$, or $out(\bar{a}-1)$, which represents precisely the value in the output tape at each step of $M$ running on input $\enc(\A)$. From this argument, it can be seen that $\sem{\alpha}(\A) = f(\enc(\A))$.

\medskip

To conclude the proof, we show that for each formula $\alpha$ in the previously defined extension of $\rqfo(\fo)$, there exists an equivalent formula confirming to the grammar of $\rqfo(\fo)$ defined in Section \ref{sec:beyond}. It suffices to consider a formula $\alpha$ of the form 
$$
\alpha(\bar{x}_1) = \clfp{f_1(\bar{x}_1): \alpha_1(\bar{x}_1,f_1,\ldots,f_n),\alpha_2(\bar{x}_2,f_1,\ldots,f_n),\ldots,\alpha_n(\bar{x}_n,f_1,\ldots,f_n)},
$$
and show an equivalent formula defined by a LSFP operator which uses one formula less in its definition.

We construct the equivalent formula as follows. We use a new function symbol $f$ with arity $\length{\bar{x}_1} + \length{\bar{x}_2}$. For every $i \in \{1,\ldots,n\}$, let $\alpha_i'$ be the formula obtained by performing the following replacements in $\alpha_i$:
\begin{align*}
f_1(\bar{y}_1) &\text{ is replaced by } \sa{\bar{y}_2}f(\bar{y}_1,\bar{y}_2)\mult[\first(\bar{y}_1)\mult\last(\bar{y}_2)  \add (\neg\first(\bar{y}_1))\mult\first(\bar{y}_2)], \\
f_2(\bar{y}_2) &\text{ is replaced by } \sa{\bar{y}_1}f(\bar{y}_1,\bar{y}_2)\mult[\first(\bar{y}_1)\mult\first(\bar{y}_2)  \add \last(\bar{y}_1)\mult(\neg\first(\bar{y}_2))].
\end{align*}
Moreover, let $\beta$ be a formula defined as:
\begin{align*}
\beta(\bar{x}_1,\bar{x}_2) = \,&\alpha_1'(\bar{x}_1)\mult(\first(\bar{x}_1)\mult\last(\bar{x}_2)  \add (\neg\first(\bar{x}_1))\mult\first(\bar{x}_2)) \add \\ &\alpha_2'(\bar{x}_2)\mult(\first(\bar{x}_1)\mult\first(\bar{x}_2)  \add \last(\bar{x}_1)\mult(\neg\first(\bar{x}_2))).
\end{align*}
It can be seen that all values of $f_1$, besides the first one, are stored in the first assignment of $\bar{x}_2$, while the first value of $f_1$ is stored in the last assignment of $\bar{x}_2$. Moreover, all values of $f_2$, besides the first one, are stored in the last assignment of $\bar{x}_1$, while the first value of $f_2$ is stored in the first assignment of $\bar{x}_1$. 
%This schema is depicted in Table \ref{table-clfp}. 
We use formula $\beta$ to define the following formula:
\begin{align*}
\alpha'(\bar{x}_1) = \,&\sa{\bar{x}_2}\clfp{f(\bar{x}_1,\bar{x}_2): \beta(\bar{x}_1,\bar{x}_2,f,f_3,\ldots,f_n),\\
&\alpha'_3(\bar{x}_3,f,f_3,\ldots,f_n),\ldots,\alpha'_n(\bar{x}_n,f,f_3,\ldots,f_n)}\mult(\first(\bar{x}_1)\mult\last(\bar{x}_2)  \add (\neg\first(\bar{x}_1))\mult\first(\bar{x}_2))
\end{align*}

It is not difficult to see that $\alpha'(\bar{x}_1)$ is equivalent to $\alpha(\bar{x}_1)$, which concludes the proof.

### 2017-03-03 LMCS/proof-sigma2hard.tex
journal primera version

--- 2017-01-12 LICS/appendix-syntactic.tex
@@
-\subsection*{Proof of Theorem \ref{theo-pnf-snf}}
-Recall that a formula in $\eqso(\LL)$ is on the following grammar:
-$$
-\alpha = \varphi \ \mid \ s \ \mid \ (\alpha + \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \sa{X} \alpha,
-$$
-where $\varphi$ is a formula in $\LL$ and $s\in\nat$.
-We will construct a recursive function $\tau$ such that for every $\eqso(\LL)$ formula $\alpha$, it outputs an equivalent formula $\tau(\alpha)$ which is in $\LL$-SNF.
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$ which is clearly equivalent and already in $\LL$-SNF.
-If $\alpha = s$, then we define $(\top \add \cdots \add \top)$ ($s$ times), which also satisfies the condition.
-We assume that for every sub-formula $\beta$ of $\alpha$, $\tau(\beta)$ is an equivalent formula in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$, then we define $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$.
-If $\alpha = \sa{x}\beta$, then let $\tau(\beta) = \sum_{i = 1}^{k}\beta_i$ for some $k$ where each $\beta_i$ is in $\LL$-PNF.
-By grouping together the terms in the sum, we define $\tau(\alpha) = \sum_{i = 1}^{m}\sa{x}\beta_i$ which satisfies the condition and is equivalent to $\alpha$.
-If $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case.
-This covers all possible cases for $\alpha$ and we conclude the proof by taking $\tau(\alpha)$ as the desired rewrite of $\alpha$.
-
-\subsection*{Proof of theorem \ref{theo-pi1-pnf}}
-We divide the proof in three parts.
-
-\vspace{1em}
-First, we prove that the formula $\alpha_{0} = \left( \sa{X} 1 \right) + 1$ with $\arity(X) = 1$ (i.e.
-the function $2^{n}+1$, where $n$ is the size of the input structure) is not equivalent to any formula in $\loge{0}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ in $\loge{0}$-PNF that is equivalent to the $\eqso(\loge{0})$ formula $\alpha_0$.
-In \cite{SalujaST95}, it was proved that if $\length{\bar{X}} > 0$, then the function defined by $\alpha$, for big enough structures, is always even which is not possible in our case.
-On the other hand, if $\alpha$ is of the form $\sa{\bar{x}}\varphi(\bar{x})$, then $\alpha$ defines a polynomially bounded function which also leads to a contradiction.
-
-\vspace{1em}
-Second, we prove that the formula $\alpha_{1} = 2$ (i.e.
-the constant $2$) is not equivalent to any formula in $\loge{1}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to the $\eqso(\loge{1})$ formula $2$.
-First, if $\length{\bar{X}} = \length{\bar{x}} = 0$, then the function defined by $\alpha$ is never greater than 1. 
-Therefore, suppose that $\length{\bar{X}} > 0$ or $\length{\bar{x}} > 0$, and consider any ordered structure $\A$. 
-Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
-Now consider the ordered structure $\A'$ that is obtained by duplicating $\A$.
-This is, each half of $\A'$ is isomorphic to $\A$.
-Note that $\A'\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$ for $i = 1,2$ and there exists a third assignments $(\bar{B}_1',\bar{b}_1',\bar{a}_1')$ that is isomorphic to $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$ but in the other half to the structure such that $\A'\models\varphi(\bar{B}_1',\bar{b}_1',\bar{a}_1')$.
-We have that $\sem{\alpha}(\A) \geq 3$ which is a contradiction.
-
-\vspace{1em}
-We now show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there is an equivalent formula $\beta$ in $\LL$-PNF.
-As in Theorem \ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such formula.
-As we showed, there exists an equivalent formula in $\LL$-SNF, so we assume that $\alpha$ is in that form.
-Let $\alpha = \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-SNF.
-Without lost of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
-If not, we replace each $\alpha_i$ for the equivalent formula $\sa{\bar{X}} \sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge\forall z\,Y(z) \wedge \forall z(y \leq z))$.
-
-Now we begin describing the function $\tau$.
-If $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then the formula is already in $\LL$-PNF so we define $\tau(\alpha) = \alpha$.
-If $\alpha = \alpha_1 + \alpha_2$, then we assume that $\tau(\alpha_1) = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ and $\tau(\alpha_2) = \sa{\bar{Y}}\sa{\bar{y}}\psi(\bar{Y},\bar{y})$.
-The construction that we will provide for this function works by identifying a ``first'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ and a ``last'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$.
-These are identified by the following formulas:
-\begin{align*}
-\gamma_{\text{first}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z}), \\
-\gamma_{\text{last}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x}).
-\end{align*}
-Similarly, we define the formulas $\gamma_{\text{first}}(\bar{Y},\bar{y})$ and $\gamma_{\text{last}}(\bar{Y},\bar{y})$ (for the sake of simplicity we reuse the names $\gamma_{\text{first}}$ and $\gamma_{\text{last}}$).
-In other words, the ``first'' assignment is the one where every second-order predicate is empty and the first-order assignment is the lexicographically smallest, and the ``last'' assignment is the one where every second-order predicate is full and the first-order assignment is the lexicographically greatest.
-We also need to identify assignments that are not first and are not last.
-We do this by negating the two formulas above and grouping together the first-order variables:
-\begin{align*}
-\gamma_{\text{not first}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{z}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{z}_i)), \\
-\gamma_{\text{not last}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{x} < \bar{z}_0 \vee \bigvee_{i = 1}^{\length{\bar{X}}}\neg X(\bar{z}_i)),
-\end{align*}
-where $\bar{z} = (\bar{z}_0,\bar{z}_1,\ldots,\bar{z}_{\length{\bar{X}}})$. The following formula is equivalent to $\alpha$:
-\begin{align}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{not first}}(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{Y},\bar{y}))\vee \label{eq:partition1} \\
-&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))\vee \label{eq:partition2}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{not last}}(\bar{Y},\bar{y}))\vee \label{eq:partition3}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{last}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))]. \label{eq:partition4}
-\end{align}
-To show that the formula is indeed equivalent to $\alpha$, note that the formulas in lines (\ref{eq:partition1}) and (\ref{eq:partition2}) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$, and the formulas in lines (\ref{eq:partition3}) and (\ref{eq:partition4}) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
-Altogether, the four lines define pairwise disjoint assignments for $(\bar{X},\bar{x}),(\bar{Y},\bar{y})$.
-With this, it is straightforward to show that the above formula is equivalent to $\alpha$.
-However, the formula is not yet in the correct form since it has existential quantifiers in the subformulas $\gamma_{\text{not first}}$ and $\gamma_{\text{not last}}$.
-To solve this, first let take a close look to the complete formula:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\exists\bar{v}(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\exists\bar{w}(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))].
-\end{align*}
-To construct an equivalent formula that is in the correct form, we define $\bar{u} = (\bar{v},\bar{w})$ and we replace the first-order quantifiers by a first-sum and count the first assignment to $\bar{v}$ and $\bar{w}$ that satisfies the formula.
-A similar construction was used in \cite{SalujaST95}.
-Then the final formula equivalent to $\alpha$ is the following:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
-& \hspace{3cm}  \bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&
-(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))\vee
-\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
-& \hspace{3cm} (\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee \\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))].
-\end{align*}
-Finally, consider a $\eqso(\LL)$ formula $\alpha$ in $\LL$-SNF.
-If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we can consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the condition in the hypothesis.
-
-\subsection*{Proof of Theorem \ref{prop-rest}}
-We give this proof in three parts.
-
-\vspace{1em}
-First, we show that $\QE{0} \not\subseteq \E{1}$.
-By contradiction, suppose that there is a $\QE{0}$ formula $\alpha$ over some signature $\R$ such that defines the following function.
-For every finite $\R$-structure with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n - 1$.
-We use the following claim.
-\begin{claim}
-Let $\alpha = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. Then the function defined by $\alpha$ is either null, greater or equal to $n$, or is in $\Omega(n^2)$.
-\end{claim}
-\begin{proof}
-Suppose that the function defined by $\alpha$ is not $0$ and that $\varphi$ is in DNF. Furthermore, suppose $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$.
-Then $\alpha = \sa{\bar{x}} \varphi_1(\bar{x}) \vee \cdots \vee \varphi_n(\bar{x})$.
-Since $\alpha$ is not null, then some $\varphi_i$ must be satisfiable.
-This is, the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is not null.
-We will prove by induction on $\length{\bar{x}}$ that the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is either greater or equal to $n$, or in $\Omega(n^2)$.
-We address the case $\length{\bar{x}}= 1$, then $\alpha = \sa{x}\bigwedge\psi(x)$.
-If any $\psi(x) = (x = x)$ or $\neg(x < x)$, then we can eliminate it and we obtain the same function.
-If any $\psi = (x < x)$ or $\neg(x=x)$, then the function becomes null.
-If $\psi(x) = R(x,\ldots,x)$ for some $R\in\R$ the function becomes null for the structures we are considering.
-If $\psi(x) = \neg R(x,\ldots,x)$, we can eliminate it and for the structures we are considering we obtain the same function.
-The only possible $\alpha$ left is $\alpha = \sa{x}\top$ which is equal to the function $n$.
-This covers all possible cases for $\length{\bar{x}} = 1$.
-Now suppose that it holds for $\length{\bar{x}} = k$ and suppose $\alpha = \sa{\bar{x}}\bigwedge\psi(\bar{x})$ for $\length{\bar{x}} = k+1$.
-If any $\psi(\bar{x}) = (x_i = x_j)$ where $i \neq j$, then $\alpha$ describes the same function as $\alpha$ where $x_j$ has been replaced by $x_i$.
-In this formula the tuple of first-order variables has $k$ elements so the function it describes if one of the mentioned in the hypothesis.
-If $i = j$, then we can eliminate it and obtain the same function.
-If any $\psi(\bar{x}) = R(\bar{v})$ or $\neg R(\bar{v})$ where $\bar{v}$ is a sub-tuple of $\bar{x}$ then we can either eliminate it or the function becomes null, following the same argument as in the case $\length{\bar{x}} = 1$.
-If any $\psi(\bar{x}) = \neg(x_i = x_j)$ or $(x_i < x_j)$ where $i = j$, then the function becomes null.
-If any $\psi(\bar{x}) = \neg(x_i < x_j)$ where $i = j$, we can eliminate it.
-The remaining formulas in $\bigwedge\psi(\bar{x})$ are either $\neg(x_i = x_j)$, $(x_i<x_j)$ or $\neg(x_i<x_j)$.
-If the formula violates transitivity in $<$ (for example, $x < y \wedge y < z \wedge z < x$), then the function $\alpha$ describes is null.
-Therefore, there is some order over $\bar{x}$ that satisfies $\bigwedge\psi(\bar{x})$.
-Consider the formula that describes this order (like $x_1 < x_3 \wedge x_3 < x_4 \wedge x_4 < x_2$).
-The function $\alpha$ describes is greater or equal to the one this formula describes, which is exactly $\binom{n}{\length{\bar{x}}}$ which is in $\Omega(n^{\length{\bar{x}}}) \subseteq \Omega(n^2)$ if $\length{\bar{x}} > 1$.
-This concludes the proof of the claim.
-\end{proof}
-We suppose that $\alpha$ is in SNF, this is, $\alpha = \sum_{i = 1}^n\alpha_i$.
-Since $\alpha$ is not null, consider some $\alpha_i$ that describes a non-null function.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, where $\varphi$ is quantifier-free.
-Note that if $\length{\bar{X}} > 0$, then the function $\alpha$ describes is in $\Omega(2^n)$, as it was proven by the authors in \cite{SalujaST95}.
-We have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$, as we proved in the claim, describes either some function greater or equal to $n$, or in $\Omega(n^2)$, which leads to a contradiction.
-Lastly, note that the formula $\sa{x}\exists y(x < y)$ is in $\E{0}$ and describes the function $n-1$, which concludes the proof.
-
-\vspace{1em}
-Now we show that $\E{1}\not\subseteq\QE{0}$.
-In Theorem \ref{theo-pi1-pnf} we proved that there is no formula in $\loge{1}$-PNF equivalent to the formula $\alpha = 2$.
-Every formula in $\E{1}$ can be expressed in $\loge{1}$-PNF, which implies that $2 \in \QE{0}$ and $2 \not\in \E{1}$.
-
-\vspace{1em}
-Lastly, we prove that $\eqso(\loge{1})\subsetneq\eqso(\logu{1})$.
-For inclusion, let $\alpha$ be a formula in $\eqso(\loge{1})$.
-Suppose that it is in $\loge{1}$-SNF.
-This is, $\alpha = c + \sum_{i = 1}^{n}\alpha_i$.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, where $\varphi_i$ is quantifier-free, for each $\alpha_i$.
-We use the same construction used in \cite{SalujaST95}, and we obtain that the formula $\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \forall\bar{y}'(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')$ for every assignment to $(\bar{X},\bar{x})$.
-We do this replacement for each $\alpha_i$ and we obtain an equivalent formula in $\eqso(\logu{1})$.
-
-To prove that the inclusion is proper, consider the $\eqso(\logu{1})$ formula $\sa{x}\forall y(y = x)$. This formula defines the following function that takes an ordered structure $\A$ as input:
-$$
-\sem{\alpha}(\A) = 
-\begin{cases}
-1 &\A \text{ has one element}\\
-0 &\text{ otherwise}.
-\end{cases}
-$$
-Suppose that there exists an equivalent formula $\alpha$ in $\eqso(\loge{1})$.
-Also, suppose that it is in $\L$-PNF, so $\alpha = c + \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$.
-Since $\alpha$ takes the value 0 for some structures, $c$ must be 0.
-Consider a structure $\mathfrak{1}$ with one element.
-We have that for some $i$, there exists an assignment $(\bar{B},\bar{b},\bar{a})$ for $(\bar{X},\bar{x},\bar{y})$ such that $\mathfrak{1}\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
-Consider now the structure $\mathfrak{2}$ that is obtained by duplicating $\mathfrak{1}$, as we did for Theorem \ref{theo-pi1-pnf}.
-Note that $\mathfrak{2}\models\varphi_i(\bar{B},\bar{b},\bar{a})$, which implies that $\sem{\alpha}(\mathfrak{2}) \geq 1$, which leads to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop-e1-nc}}
-
-Towards a contradiction, assume that the statement is false.
-This is, that $\E{1}$ is closed under binary sum.
-Consider the formula $\sa{x}(x = x)$ which is in $\E{1}$ over some signature $\R$.
-For every finite $\R$-structure $\A$ with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n$.
-From our assumption, there exists some formula in $\E{1}$ equivalent to the formula $\alpha \add \alpha$, which describes the function $2n$.
-Let $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ be this formula, where $\varphi$ is quantifier-free.
-Note that the function defined by this formula is equal or greater than the one defined by $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ divided by a polynomial factor.
-More specifically, for each ordered structure $\A$ with domain $A$, we have the following inequality:
-$$
-\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot \vert A \vert^{\length{\bar{y}}} \geq \sem{\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A)
-$$
-Note that the formula $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ defines a function in $\E{0}$.
-It was shown by the authors in \cite{SalujaST95} that every function in $\E{0}$ grows exponentially over the size of the structure for large enough structures, when $\length{\bar{X}} > 0$.
-This function divided by a polynomial factor still grows exponentially.
-Therefore, for $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ we have that $\length{\bar{X}} = 0$.
-
-Now, for the formula $\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ consider a structure $\mathfrak{1}$ with only one element $a$.
-We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$, but the only possible assignment to $\bar{x}$ is the tuple $(a,\ldots,a)$ so $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which follows to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop:qe0-fp-qe1-totp-fptras}}
-
-The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$.
-This is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every finite $\R$-structure $\A$ with domain $A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$.
-We base our proof on this fact.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is contained in \textsc{TotP}.} Let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Consider the following nondeterministic procedure that on input $\enc(\A)$ generates $\sem{\alpha}(\A)$ branches.
-For each $\alpha_i = \varphi$, where $\varphi$ is a $\loge{1}$ formula, it checks if $\A\models\varphi$ in polynomial time and generates a new branch if that is the case.
-For each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi$, this formula is also in $\E{1}$.
-We use the reduction to $\cdnf$ provided in \cite{SalujaST95} and we obtain $g(\enc(\A))$, which is an instance to $\cdnf$.
-Since $\cdnf$ is also in $\totp$ \cite{PagourtzisZ06}, we simulate the corresponding nondeterministic procedure that generates exactly $\cdnf(\enc(g(\A)))$ branches.
-Since, $\fp\subseteq\totp$\cite{PagourtzisZ06}, each polynomially computable function is also in $\totp$, and then on each of these branches we simulate the corresponding nondeterministic procedure to generate $h(\vert A \vert)$ more.
-The number of branches for each $\alpha_i$ is $\sem{\alpha_i}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$, and the total number of branches in the procedure amounts to $\sem{\alpha}(\A)$.
-We conclude that $\alpha\in\totp$.
-
-\vspace{1em}
-{\bf Every function in} $\boldsymbol{\eqso(\loge{1})}$ {\bf has an FPRAS.}  let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Note that each $\alpha_i$ that is equal to some $\loge{1}$ formula $\varphi$ has an FPRAS given by the procedure that simply checks if $\A\models\varphi$ and returns 1 if it does and 0 otherwise.
-Also, each remaining $\alpha_i$ has an FPRAS since $\alpha_i\in \E{1}$ \cite{SalujaST95}.
-If two functions have an FPRAS, then their sum also has one given by the procedure that simulates them both and sums the results.
-We conclude that $\alpha$ has an FPRAS.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is closed under sum and multiplication.} Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that the class is closed under multiplication.
-We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of $\so$.
-
-\begin{lemma} \label{conj-mult}
-If $\LL$ is a fragment closed under conjunction, then $\eqso(\LL)$ is closed under binary multiplication.
-\end{lemma}
-\begin{proof}
-We will define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ extended by binary product, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
-In fact, the formula $\tau(\alpha)$ is in $\LL$-SNF.
-First we replace each constant $s$ in $\alpha$ for $(\top \add \cdots \add \top)$ ($s$ times).
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$.
-We assume that for every $\beta$ that has less algebraic operators than $\alpha$, $\tau(\beta)$ is in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$ then we define $\tau(\alpha) = \tau(\alpha_1) + \tau(\alpha_2)$.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$, then we define $\tau(\alpha)$ as the formula in $\LL$-SNF that is equivalent to $\sa{x}\tau(\beta)$ and to $\tau(\alpha) = \sa{X}\tau(\beta)$, respectively.
-If $\alpha = (\alpha_1 \cdot \alpha_2)$, we assume that each $\alpha_i$ is in $\LL$-SNF.
-We identify three cases.
-(1) Some $\alpha_i$ is equal to $\sum_{j = 1}^n\beta_j$ for $n > 1$.
-Suppose wlog.
-that it is $\alpha_1$.
-We then define $\tau(\alpha) = \sum_{j = 1}^n\tau(\beta_j\cdot\alpha_2)$.
-In the following cases, $\alpha_1$ and $\alpha_2$ are in $\LL$-SNF.
-(2) If some $\alpha_i$ is equal to $\sa{X}\beta$ or $\sa{x}\beta$, we define $\tau(\alpha)$ as the $\LL$-SNF formula that is equivalent to $\sa{x}\tau(\beta\cdot\alpha_2)$ and $\sa{X}\tau(\beta\cdot\alpha_2)$, respectively.
-The remaining case is (3) $\alpha_1 = \varphi_1$ and $\alpha_2 = \varphi_2$ where each $\varphi$ is an $\LL$ formula.
-Then we define $\tau(\alpha) = \varphi_1 \wedge \varphi_2$.
-This covers all possible cases for $\alpha$.
-For every pair of formulas $\alpha,\beta$ in $\eqso(\LL)$, we have that their multiplication $(\alpha\cdot\beta)$ is a formula in the grammar $\eqso(\LL)$ extended by binary product, and so, there exists an equivalent formula $\tau(\alpha\cdot\beta)$ which is in unextended $\eqso(\LL)$.
-\end{proof}
-Since $\loge{1}$ is closed under conjunction, this also holds for $\eqso(\loge{1})$. This concludes the proof.
-
-\subsection*{Proof of Proposition \ref{pi-minusone}}
-
-Let $\LL$ be a fragment of $\fo$ that contains $\logu{1}$.
-Then we have that every function in $\U{1}$ is expressible in $\eqso(\LL)$.
-In particular, $\ctcnf \in \eqso(\LL)$.
-Suppose that $\eqso(\LL)$ is closed under subtraction by one.
-Then, the function $\ctcnf-1$, which counts the number of satisfying assignments of a 3-CNF formula minus one, is also in $\eqso(\LL)$.
-Note that $\eqso(\LL) \subseteq \eqso(\fo) = \shp$.
-We have that $\ctcnf$ is $\shp$-complete under parsimonious reductions\footnote{It can be easily verified that the standard reduction from SAT to 3-CNF (or 3-SAT) preserves the number of satisfying assignments}.
-Now, let $f$ be a function in $\shp$, and consider the nondeterministic polynomial-time procedure that on input $\enc(\A)$ computes the corresponding reduction to $\ctcnf$, name it $g(\enc(\A))$, and simulates the $\shp$ procedure for $\ctcnf-1$ on input $g(\enc(\A))$.
-We have that this is a $\shp$ procedure that computes $f-1$, from which we conclude that $\shp$ is closed under subtraction by one.
-
-\subsection*{Proof of Theorem \ref{sigmafo-minusone}}
-
-\noindent {\bf Closed under sum and multiplication.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
-This can be seen because $\eqso(\LL)$ is closed under sum for every fragment $\LL$ by definition, and since $\logex{1}$ is closed under conjunction, from Lemma \ref{conj-mult} it follows that $\eqso(\logex{1})$ is closed under multiplication.
-
-\vspace{1em}
-\noindent {\bf Easy decision version and FRPRAS.}  We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
-We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Theorem \ref{prop:qe0-fp-qe1-totp-fptras}.
-First, we define a function that converts a formula $\alpha$ in $\eqso(\logex{1})$ over a signature $\R$ into a formula $\lambda(\alpha)$ in $\eqso(\loge{1})$ over a signature $\R_{\alpha}$.
-Afterwards, we define a function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an $\R_{\alpha}$-structure $g_{\alpha}(\A)$.
-
-Let $\alpha$ be in $\eqso(\logex{0})$.
-The signature $\R_{\alpha}$ is obtained by adding the symbol $R_{\psi}$, for every $\fo$ formula $\psi(\bar{z})$ in $\alpha$, to $\R$.
-Each symbol $R_{\psi}$ represents a predicate with arity $\length{\bar{z}}$.
-Then, $\lambda(\alpha)$ is defined as $\alpha$ where each $\fo$ formula $\psi(\bar{z})$ has been replaced by $R_{\psi}(\bar{z})$.
-We now define the function $g_{\alpha}$ procedurally.
-Let $\A$ be a $\R$-structure with domain $A$.
-Let $\A'$ be an $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A}$ empty.
-For each $\fo$-formula $\psi(\z)$ with $\length{\bar{z}}$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length{\bar{z}}}$.
-If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$ (this can be done in $\ptime$).
-This concludes the construction of $\A'$.
-Note that the number of $\fo$ subformulas, arity and tuple size is fixed in $\alpha$, so computing this function takes polynomial time over the size of the structure.
-Moreover, the encoding of $\A'$ has polynomial size over the size of $\enc(\A)$.
-We define $g_{\alpha}(\A) = \A'$ and we have that for each $\R$-structure $\A$: $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$.
-Therefore, we have a parsimonious reduction from $\alpha$ to the $\eqso(\loge{1})$ formula $\lambda(\alpha)$.
-
-To show that $\alpha$ is in $\totp$, we can convert $\alpha$ and $\enc(\A)$ into $\lambda(\alpha)$ and $\enc(g_{\alpha}(\A))$, respectively, and run the procedure in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-Similarly, to show that $\alpha$ has an FPRAS, we do the same as before and simulates the FPRAS for $\lambda(\alpha)$ in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-These procedures also takes polynomial time and satisfies the condition.
-
-\vspace{1em}
-\noindent {\bf Closed under subtraction by one.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. 
-For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Without lost of generality, we assume that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
-Moreover, we assume that $\length{\bar{x}} > 0$  since, if this is not the case, we can replace $\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
-
-The proof will be separated in two parts.
-In the first part, we suppose that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$.
-Then we will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ is the formula.
-In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ the sum of one or more formulas in $\logex{1}$-PNF.
-We define a formula $\varphi'$ that satisfies the following condition: if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ and $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be proven.
-
-\vspace{1em}
-
-\noindent {\em Part (1).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ where $\varphi$ is an  $\fo$-formula.
-Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}} \varphi(\bar{x})$ (i.e.
-$\length{\bar{X}} = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}} [\varphi(\bar{x})\wedge\exists\bar{z}(\varphi(\bar{z})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
-Therefore, for the rest of the proof we can assume that $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
-
- Let $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$ and $\bar{X} = (X_1,\ldots,X_{\length{\bar{X}}})$.
-To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in into a DNF formula. 
-More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i$ is a conjunction of atomic formulas or negation of atomic formulas.
-Furthermore, we suppose that each conjunct $\varphi_i$ has the~form:
-$$
-\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
-\underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}} \wedge
-\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}}.
-$$
-Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ can appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
-
-Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ are mentioned in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
-We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
-\begin{itemize}
-	\item[(a)] In order to satisfy the first condition, consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
-We add $\length{\bar{w}}$ new variables $z_1,\ldots,z_{\length{\bar{w}}}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length{\bar{w}}})$.
-We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$.
-The formula $\varphi_i$ is now defined as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \exists \bar{z} \; (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{-}(\bar{X},\bar{x},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y})).
-	$$
-	We repeat this process for each instance of a $X_j(\bar{w})$ in $\varphi_i$, and we obtain a formula where none of the $X_j$'s acts over any variable in $\bar{x}$.
-We add the new first-order variables to $\bar{y}$ and we redefine $\varphi_i$ as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
-	$$
-	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = \exists \bar{y} \;  X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := \exists \bar{y} \;  X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
-	\item[(b)] An ordered partition on a set $S$ is defined by an equivalence relation $\sim$ over $S$, and a linear order over $S/\!\sim$.
-For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$.
-A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
-On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
-	For a given $k$, let $\cB_k$ be the number of possible ordered partitions for a set of size $k$. For $1 \leq j \leq \cB_{\length{(\bar{x},\bar{y})}}$ 
-	let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
-
-	Thus, the formula $\varphi_i(\bar{X},\bar{x})$ is then redefined as:
-	$$
-	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length{(\bar{x},\bar{y})}}} \exists \bar{y} \; [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
-	$$
-	Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
-	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  to $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
-	\item[(c)] For showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
-$\varphi_i$ is inconsistent), we do the following.
-If there exists an instance of $X_j(\bar{z})$ in $\varphi^{+}_i$, an instance of $\neg X_j(\bar{w})$ in $\varphi^{-}_i$ and the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$, then the entire formula $\varphi_i$ is removed from $\varphi$.
-\end{itemize}
-It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$.
-From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c).
-Furthermore, it has the following form:
-$$
-\varphi_i(\bar{X},\bar{x}) \; = \; \exists{\bar{y}} \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
-$$
-where $\varphi^{+}$ and $\varphi^{-}_i$ do not depend on  $\bar{x}$.
-\begin{claim}\label{claim:minusone}
-For a given ordered structure $\A$, if there exists a first-order assignment $v$ for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists a second-order assignment $V$ for $\A$ such that  $(\A,V,v)\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-\end{claim}
-\begin{proof}
-Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
-Define $\bar{B} = (B_1,\ldots,B_{\length{\bar{X}}})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
-Towards a contradiction, suppose that $(\A,V,v)\not\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-By the choice of $v$, and construction of $V$ it is clear that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})\wedge\varphi^{+}_i(\bar{X},\bar{y})$, so we necessarily have that $(\A,V,v)\not\models\varphi^{-}_i(\bar{X},\bar{y})$.
-Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,V, v)\not\models\neg X_t(\bar{w})$, namely, $v(\bar{w})\in B_t$. 
-However, by the construction of $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$.
-Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$.
-This violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
-\end{proof}
-
-The previous claim and proof motivates the following definitions.
-For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length{\bar{X}}})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
-One can easily check that for every assignments $V, v$ such that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$, then it holds that $(\A,\bar{B}^v,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all the assignment of $\bar{X}$ when $v$ is fixed.
-Therefore, by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
-For this, the idea is to choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
-$$
-\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge\forall\bar{x}'\forall\bar{y}'(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')).
-$$ 
-If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
-Furthermore, let $V^*$ the second order assignment that satisfies $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*$ the first order assignment such that $v^*(\bar{x}) = v(\bar{x})$.
-By the previous discussion, we have that $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.
-
-Now, we have all the tools to introduce the main formula of the proof. 
-Intuitively, we want to exclude the assignment $(V^*, v^*)$ as a satisfying assignment for $\varphi_i(\bar{X},\bar{x})$.
-Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if, 
-if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, then $V \neq V^*$ or $v \neq v^*$. 
-That is, it excludes all pairs $(V, v)$ that coincide with $(V^*, v^*)$ in $(\bar{X}, \bar{x})$ whenever $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. This property can be defined as follows:
-\begin{align}
-\psi_i(\bar{X},\bar{x}) = & \;\; \big( \, \exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
-& \hspace{1cm} \Big( \; 
-\exists\bar{y} \;\; \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i^+(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\
-& \hspace{1.5cm} (\exists\bar{x}'\exists\bar{y}\, \varphi_i(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}) \;\; \Big)  \label{eq:part3}
-\end{align}
-To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
-Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-Then, the conclusion of the implication at (\ref{eq:part2}) and (\ref{eq:part3}), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
-Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by showing that if $\varphi_i(\bar{X},\bar{x})$ is satisfied then $V(\bar{X})$ must contain one element more that it is not in $V^*(\bar{X})$.
-The second disjunct (\ref{eq:part3}) is satisfied when $v(\bar{x})$ is not the lexicographically smallest tuple that satisfies $\varphi_i$ (i.e. $v(\bar{x}) \neq v^*(\bar{x})$).
-
-We are ready to define the formula  $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$
-where each modified disjunct $\varphi_i'(\bar{X},\bar{x})$ is constructed as follows. 
-For the sake of simplification, define the auxiliary formula $\chi_i = \neg\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$. 
-This formula basically checks if $\varphi_i$ is satisfiable or not.
-Then, first define the formula:
-$$
-\varphi'_1(\bar{X},\bar{x}) = \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
-$$
-This models that all the assignments that satisfy $\varphi_1$ also satisfy $\varphi'_1$, except for $(V^*,v^*)$. Then we define:
-$$
-\varphi'_2(\bar{X},\bar{x}) = \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
-$$
-This models that all the assignments that satisfy $\varphi_2$ also satisfy $\varphi'_2$, but if there was no assignment for $\varphi_1$, then we exclude $(V^*,v^*)$ for $\varphi_2$.
-Then we generalize this construction for each $\varphi_i$:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})),
-$$
-By the construction of $\kappa(\alpha)$, one can easily check that  $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.
-
-\vspace{1em}
-
-{\em Part (2).} Let $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ for some $\eqso(\logex{1})$ formula $\beta$. We define $\kappa(\alpha)$ as follows.
-First, perform the same transformations to $\varphi(\bar{X},\bar{x})$ as in the previous section.
-Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c).
-We also use the previously defined formulas $\chi_i$ and $\psi_i$, for each $i$.
-
-We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if and only if $\sem{\beta}(\A) = 0$.
-If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg\exists\bar{x}'\varphi(\bar{x}')$.
-If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$ as each $\chi_i$ was previously defined.
-If $\beta = (\beta_1 + \beta_2)$, then $\lambda(\beta) = \lambda(\beta_1) \wedge \lambda(\beta_2)$.
-Now, for each $\varphi_i$ we define:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x}))\Big)\Big).
-$$
-Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
-
-%\subsection*{Proof of Theorem \ref{sub-pnp}}
-%
-%Towards a contradiction, suppose that any of the classes $\E{1}$, $\eqso(\loge{1})$, and $\eqso(\logex{1})$ is closed under subtraction.
-%
-%Let $\R$ include the symbols $S_1, S_2, S_3, S_4$, which describe the following properties.
-If a finite $\R$-structure $\A$ defines a 3DNF formula $\Phi$, then its domain is the set of variables mentioned in $\Phi$, and for each $i = 1,2,3,4$:
-%\begin{align*}
-%	S_1^\A &= \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_2^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_3^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_4^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
-%\end{align*}
-%Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
-%\begin{multline*}
-%\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
-%\end{multline*}
-%Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
-%$$
-%\varphi(x,X) = (T(x) \vee \neg T(x)).
-%$$
-%Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Given that $f_{\#3DNF}, f_{all} \in \E{1} \subseteq \eqso(\loge{1}) \subseteq \eqso(\logex{1}) \subseteq \totp$, and at least one of the classes is closed under subtraction, the function $f_{\#3DNF} - f_{all}$ is in $\totp$.
-However, note that for each $\R$-structure $\A$ that represents a formula $\Phi$, $(f_{\#3DNF} - f_{all})(\A) = 0$ if an only if $\Phi$ is a tautology.
-The decision version of this function is the $\np$-complete problem $\textsc{Tautology}$.
-And since the function is in $\totp$, its decision version is also in $\ptime$.
-We conclude that $\np \subseteq \ptime$.
-
-\subsection*{Proof of Proposition \ref{prop:ehorn-pe}}
-Pagourtzis and Zachos mention a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula \cite{PagourtzisZ06}.
-This procedure can be easily extended to receive Horn formulas, and furthermore, a disjunction of Horn formulas.
-Hence $\shdhsat$ is in $\totp$.
-
-As we show in Proposition \ref{sigma2hard}, $\shdhsat$ is complete for $\eqso(\ehorn)$ under parsimonious reductions.
-Let $\alpha$ be a formula in $\eqso(\ehorn)$ and let $g_{\alpha}$ be the reduction to $\shdhsat$.
-The $\totp$ procedure we construct, for each input $\enc(\A)$, is simply to compute $g_{\alpha}(\enc(\A))$, and then simulate the $\totp$ procedure for $\shdhsat$ on input $g_{\alpha}(\enc(\A))$.
-We conclude that $\alpha$ is in $\totp$.
-
-\subsection*{Proof of Proposition \ref{prop:hsat-not-sigma2}} %V.12
-We use a similar proof to the one provided by the authors in \cite{SalujaST95} to separate the classes $\E{2}$ and $\U{2}$.
-Suppose that the statement is false, this is, $\chsat \in \eqso(\loge{2})$.
-We consider the signature $\R$ that we used as the encoding for a Horn formula (Example \ref{ex-hornsat-esop1}) and that the formula $\alpha \in \eqso(\loge{2})$ follows the encoding in the same way.
-From what we proved in Theorem \ref{theo-pnf-snf}, we have that every formula in $\eqso(\loge{2})$ can be rewritten in $\loge{2}$-PNF, so we assume that $\alpha$ is in this form.
-Let $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\forall\bar{z}\,\varphi(\bar{X},\bar{x},\bar{y},\bar{z})$.
-Consider the following Horn formula $\Phi$:
-$$
-\Phi = p \wedge \bigwedge_{i = 1}^n (t_i \wedge p \to q) \wedge \neg q,
-$$
-where $n = \length{\bar{x}} + \length{\bar{y}} + 1$.
-Let $\A$ be the encoding of this formula.
-In the encoding, each variable appears as an element in the domain of $\A$.
-This formula is satisfiable, so $\sem{\alpha}(\A) \geq 1$.
-Let $(\bar{B},\bar{b},\bar{a})$ be an assignment to $(\bar{X},\bar{x},\bar{y})$ such that $\A\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-Let $t_{\ell}$ be such that it does not appear in $\bar{b}$ or $\bar{a}$.
-Consider the induced substructure $\A'$ that is obtained by removing $t_{\ell}$ from $\A$ and $\bar{B}'$ as the subset of $\bar{B}$ obtained by deleting each appearance of $t_{\ell}$ in $\bar{B}$.
-We have that $\A'\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-This is because each subformula of the form $\exists y \neg B_i$ is still true, and universal formulas are monotone over induced substructures.
-It follows that $\sem{\alpha}(\A') \geq 1$ which is not possible since $\A'$ encodes the formula
-$$
-\Phi' = p \wedge \bigwedge_{i = 1}^{\ell-1} (t_i \wedge p \to q) \wedge (p\to q) \wedge \bigwedge_{i = \ell+1}^{n} (t_i \wedge p \to q) \wedge \neg q,
-$$
-which is unsatisfiable. We arrive to a contradiction and we conclude that $\chsat$ is not in $\eqso(\ehorn)$.
-
-\subsection*{Proof of Theorem \ref{sigma2hard}} %V.13
-
 First we prove that $\shdhsat$ is in $\eqso(\ehorn)$.

### 2017-03-03 LMCS/proof-sigmafo-minusone.tex
journal primera version

--- 2017-01-12 LICS/appendix-syntactic.tex
@@
-\subsection*{Proof of Theorem \ref{theo-pnf-snf}}
-Recall that a formula in $\eqso(\LL)$ is on the following grammar:
-$$
-\alpha = \varphi \ \mid \ s \ \mid \ (\alpha + \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \sa{X} \alpha,
-$$
-where $\varphi$ is a formula in $\LL$ and $s\in\nat$.
-We will construct a recursive function $\tau$ such that for every $\eqso(\LL)$ formula $\alpha$, it outputs an equivalent formula $\tau(\alpha)$ which is in $\LL$-SNF.
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$ which is clearly equivalent and already in $\LL$-SNF.
-If $\alpha = s$, then we define $(\top \add \cdots \add \top)$ ($s$ times), which also satisfies the condition.
-We assume that for every sub-formula $\beta$ of $\alpha$, $\tau(\beta)$ is an equivalent formula in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$, then we define $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$.
-If $\alpha = \sa{x}\beta$, then let $\tau(\beta) = \sum_{i = 1}^{k}\beta_i$ for some $k$ where each $\beta_i$ is in $\LL$-PNF.
-By grouping together the terms in the sum, we define $\tau(\alpha) = \sum_{i = 1}^{m}\sa{x}\beta_i$ which satisfies the condition and is equivalent to $\alpha$.
-If $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case.
-This covers all possible cases for $\alpha$ and we conclude the proof by taking $\tau(\alpha)$ as the desired rewrite of $\alpha$.
-
-\subsection*{Proof of theorem \ref{theo-pi1-pnf}}
-We divide the proof in three parts.
-
-\vspace{1em}
-First, we prove that the formula $\alpha_{0} = \left( \sa{X} 1 \right) + 1$ with $\arity(X) = 1$ (i.e.
-the function $2^{n}+1$, where $n$ is the size of the input structure) is not equivalent to any formula in $\loge{0}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ in $\loge{0}$-PNF that is equivalent to the $\eqso(\loge{0})$ formula $\alpha_0$.
-In \cite{SalujaST95}, it was proved that if $\length{\bar{X}} > 0$, then the function defined by $\alpha$, for big enough structures, is always even which is not possible in our case.
-On the other hand, if $\alpha$ is of the form $\sa{\bar{x}}\varphi(\bar{x})$, then $\alpha$ defines a polynomially bounded function which also leads to a contradiction.
-
-\vspace{1em}
-Second, we prove that the formula $\alpha_{1} = 2$ (i.e.
-the constant $2$) is not equivalent to any formula in $\loge{1}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to the $\eqso(\loge{1})$ formula $2$.
-First, if $\length{\bar{X}} = \length{\bar{x}} = 0$, then the function defined by $\alpha$ is never greater than 1. 
-Therefore, suppose that $\length{\bar{X}} > 0$ or $\length{\bar{x}} > 0$, and consider any ordered structure $\A$. 
-Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
-Now consider the ordered structure $\A'$ that is obtained by duplicating $\A$.
-This is, each half of $\A'$ is isomorphic to $\A$.
-Note that $\A'\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$ for $i = 1,2$ and there exists a third assignments $(\bar{B}_1',\bar{b}_1',\bar{a}_1')$ that is isomorphic to $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$ but in the other half to the structure such that $\A'\models\varphi(\bar{B}_1',\bar{b}_1',\bar{a}_1')$.
-We have that $\sem{\alpha}(\A) \geq 3$ which is a contradiction.
-
-\vspace{1em}
-We now show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there is an equivalent formula $\beta$ in $\LL$-PNF.
-As in Theorem \ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such formula.
-As we showed, there exists an equivalent formula in $\LL$-SNF, so we assume that $\alpha$ is in that form.
-Let $\alpha = \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-SNF.
-Without lost of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
-If not, we replace each $\alpha_i$ for the equivalent formula $\sa{\bar{X}} \sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge\forall z\,Y(z) \wedge \forall z(y \leq z))$.
-
-Now we begin describing the function $\tau$.
-If $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then the formula is already in $\LL$-PNF so we define $\tau(\alpha) = \alpha$.
-If $\alpha = \alpha_1 + \alpha_2$, then we assume that $\tau(\alpha_1) = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ and $\tau(\alpha_2) = \sa{\bar{Y}}\sa{\bar{y}}\psi(\bar{Y},\bar{y})$.
-The construction that we will provide for this function works by identifying a ``first'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ and a ``last'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$.
-These are identified by the following formulas:
-\begin{align*}
-\gamma_{\text{first}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z}), \\
-\gamma_{\text{last}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x}).
-\end{align*}
-Similarly, we define the formulas $\gamma_{\text{first}}(\bar{Y},\bar{y})$ and $\gamma_{\text{last}}(\bar{Y},\bar{y})$ (for the sake of simplicity we reuse the names $\gamma_{\text{first}}$ and $\gamma_{\text{last}}$).
-In other words, the ``first'' assignment is the one where every second-order predicate is empty and the first-order assignment is the lexicographically smallest, and the ``last'' assignment is the one where every second-order predicate is full and the first-order assignment is the lexicographically greatest.
-We also need to identify assignments that are not first and are not last.
-We do this by negating the two formulas above and grouping together the first-order variables:
-\begin{align*}
-\gamma_{\text{not first}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{z}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{z}_i)), \\
-\gamma_{\text{not last}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{x} < \bar{z}_0 \vee \bigvee_{i = 1}^{\length{\bar{X}}}\neg X(\bar{z}_i)),
-\end{align*}
-where $\bar{z} = (\bar{z}_0,\bar{z}_1,\ldots,\bar{z}_{\length{\bar{X}}})$. The following formula is equivalent to $\alpha$:
-\begin{align}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{not first}}(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{Y},\bar{y}))\vee \label{eq:partition1} \\
-&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))\vee \label{eq:partition2}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{not last}}(\bar{Y},\bar{y}))\vee \label{eq:partition3}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{last}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))]. \label{eq:partition4}
-\end{align}
-To show that the formula is indeed equivalent to $\alpha$, note that the formulas in lines (\ref{eq:partition1}) and (\ref{eq:partition2}) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$, and the formulas in lines (\ref{eq:partition3}) and (\ref{eq:partition4}) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
-Altogether, the four lines define pairwise disjoint assignments for $(\bar{X},\bar{x}),(\bar{Y},\bar{y})$.
-With this, it is straightforward to show that the above formula is equivalent to $\alpha$.
-However, the formula is not yet in the correct form since it has existential quantifiers in the subformulas $\gamma_{\text{not first}}$ and $\gamma_{\text{not last}}$.
-To solve this, first let take a close look to the complete formula:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\exists\bar{v}(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\exists\bar{w}(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))].
-\end{align*}
-To construct an equivalent formula that is in the correct form, we define $\bar{u} = (\bar{v},\bar{w})$ and we replace the first-order quantifiers by a first-sum and count the first assignment to $\bar{v}$ and $\bar{w}$ that satisfies the formula.
-A similar construction was used in \cite{SalujaST95}.
-Then the final formula equivalent to $\alpha$ is the following:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
-& \hspace{3cm}  \bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
-&
-(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))\vee
-\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
-& \hspace{3cm} (\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee \\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))].
-\end{align*}
-Finally, consider a $\eqso(\LL)$ formula $\alpha$ in $\LL$-SNF.
-If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we can consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the condition in the hypothesis.
-
-\subsection*{Proof of Theorem \ref{prop-rest}}
-We give this proof in three parts.
-
-\vspace{1em}
-First, we show that $\QE{0} \not\subseteq \E{1}$.
-By contradiction, suppose that there is a $\QE{0}$ formula $\alpha$ over some signature $\R$ such that defines the following function.
-For every finite $\R$-structure with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n - 1$.
-We use the following claim.
-\begin{claim}
-Let $\alpha = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. Then the function defined by $\alpha$ is either null, greater or equal to $n$, or is in $\Omega(n^2)$.
-\end{claim}
-\begin{proof}
-Suppose that the function defined by $\alpha$ is not $0$ and that $\varphi$ is in DNF. Furthermore, suppose $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$.
-Then $\alpha = \sa{\bar{x}} \varphi_1(\bar{x}) \vee \cdots \vee \varphi_n(\bar{x})$.
-Since $\alpha$ is not null, then some $\varphi_i$ must be satisfiable.
-This is, the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is not null.
-We will prove by induction on $\length{\bar{x}}$ that the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is either greater or equal to $n$, or in $\Omega(n^2)$.
-We address the case $\length{\bar{x}}= 1$, then $\alpha = \sa{x}\bigwedge\psi(x)$.
-If any $\psi(x) = (x = x)$ or $\neg(x < x)$, then we can eliminate it and we obtain the same function.
-If any $\psi = (x < x)$ or $\neg(x=x)$, then the function becomes null.
-If $\psi(x) = R(x,\ldots,x)$ for some $R\in\R$ the function becomes null for the structures we are considering.
-If $\psi(x) = \neg R(x,\ldots,x)$, we can eliminate it and for the structures we are considering we obtain the same function.
-The only possible $\alpha$ left is $\alpha = \sa{x}\top$ which is equal to the function $n$.
-This covers all possible cases for $\length{\bar{x}} = 1$.
-Now suppose that it holds for $\length{\bar{x}} = k$ and suppose $\alpha = \sa{\bar{x}}\bigwedge\psi(\bar{x})$ for $\length{\bar{x}} = k+1$.
-If any $\psi(\bar{x}) = (x_i = x_j)$ where $i \neq j$, then $\alpha$ describes the same function as $\alpha$ where $x_j$ has been replaced by $x_i$.
-In this formula the tuple of first-order variables has $k$ elements so the function it describes if one of the mentioned in the hypothesis.
-If $i = j$, then we can eliminate it and obtain the same function.
-If any $\psi(\bar{x}) = R(\bar{v})$ or $\neg R(\bar{v})$ where $\bar{v}$ is a sub-tuple of $\bar{x}$ then we can either eliminate it or the function becomes null, following the same argument as in the case $\length{\bar{x}} = 1$.
-If any $\psi(\bar{x}) = \neg(x_i = x_j)$ or $(x_i < x_j)$ where $i = j$, then the function becomes null.
-If any $\psi(\bar{x}) = \neg(x_i < x_j)$ where $i = j$, we can eliminate it.
-The remaining formulas in $\bigwedge\psi(\bar{x})$ are either $\neg(x_i = x_j)$, $(x_i<x_j)$ or $\neg(x_i<x_j)$.
-If the formula violates transitivity in $<$ (for example, $x < y \wedge y < z \wedge z < x$), then the function $\alpha$ describes is null.
-Therefore, there is some order over $\bar{x}$ that satisfies $\bigwedge\psi(\bar{x})$.
-Consider the formula that describes this order (like $x_1 < x_3 \wedge x_3 < x_4 \wedge x_4 < x_2$).
-The function $\alpha$ describes is greater or equal to the one this formula describes, which is exactly $\binom{n}{\length{\bar{x}}}$ which is in $\Omega(n^{\length{\bar{x}}}) \subseteq \Omega(n^2)$ if $\length{\bar{x}} > 1$.
-This concludes the proof of the claim.
-\end{proof}
-We suppose that $\alpha$ is in SNF, this is, $\alpha = \sum_{i = 1}^n\alpha_i$.
-Since $\alpha$ is not null, consider some $\alpha_i$ that describes a non-null function.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, where $\varphi$ is quantifier-free.
-Note that if $\length{\bar{X}} > 0$, then the function $\alpha$ describes is in $\Omega(2^n)$, as it was proven by the authors in \cite{SalujaST95}.
-We have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$, as we proved in the claim, describes either some function greater or equal to $n$, or in $\Omega(n^2)$, which leads to a contradiction.
-Lastly, note that the formula $\sa{x}\exists y(x < y)$ is in $\E{0}$ and describes the function $n-1$, which concludes the proof.
-
-\vspace{1em}
-Now we show that $\E{1}\not\subseteq\QE{0}$.
-In Theorem \ref{theo-pi1-pnf} we proved that there is no formula in $\loge{1}$-PNF equivalent to the formula $\alpha = 2$.
-Every formula in $\E{1}$ can be expressed in $\loge{1}$-PNF, which implies that $2 \in \QE{0}$ and $2 \not\in \E{1}$.
-
-\vspace{1em}
-Lastly, we prove that $\eqso(\loge{1})\subsetneq\eqso(\logu{1})$.
-For inclusion, let $\alpha$ be a formula in $\eqso(\loge{1})$.
-Suppose that it is in $\loge{1}$-SNF.
-This is, $\alpha = c + \sum_{i = 1}^{n}\alpha_i$.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, where $\varphi_i$ is quantifier-free, for each $\alpha_i$.
-We use the same construction used in \cite{SalujaST95}, and we obtain that the formula $\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \forall\bar{y}'(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')$ for every assignment to $(\bar{X},\bar{x})$.
-We do this replacement for each $\alpha_i$ and we obtain an equivalent formula in $\eqso(\logu{1})$.
-
-To prove that the inclusion is proper, consider the $\eqso(\logu{1})$ formula $\sa{x}\forall y(y = x)$. This formula defines the following function that takes an ordered structure $\A$ as input:
-$$
-\sem{\alpha}(\A) = 
-\begin{cases}
-1 &\A \text{ has one element}\\
-0 &\text{ otherwise}.
-\end{cases}
-$$
-Suppose that there exists an equivalent formula $\alpha$ in $\eqso(\loge{1})$.
-Also, suppose that it is in $\L$-PNF, so $\alpha = c + \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$.
-Since $\alpha$ takes the value 0 for some structures, $c$ must be 0.
-Consider a structure $\mathfrak{1}$ with one element.
-We have that for some $i$, there exists an assignment $(\bar{B},\bar{b},\bar{a})$ for $(\bar{X},\bar{x},\bar{y})$ such that $\mathfrak{1}\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
-Consider now the structure $\mathfrak{2}$ that is obtained by duplicating $\mathfrak{1}$, as we did for Theorem \ref{theo-pi1-pnf}.
-Note that $\mathfrak{2}\models\varphi_i(\bar{B},\bar{b},\bar{a})$, which implies that $\sem{\alpha}(\mathfrak{2}) \geq 1$, which leads to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop-e1-nc}}
-
-Towards a contradiction, assume that the statement is false.
-This is, that $\E{1}$ is closed under binary sum.
-Consider the formula $\sa{x}(x = x)$ which is in $\E{1}$ over some signature $\R$.
-For every finite $\R$-structure $\A$ with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n$.
-From our assumption, there exists some formula in $\E{1}$ equivalent to the formula $\alpha \add \alpha$, which describes the function $2n$.
-Let $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ be this formula, where $\varphi$ is quantifier-free.
-Note that the function defined by this formula is equal or greater than the one defined by $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ divided by a polynomial factor.
-More specifically, for each ordered structure $\A$ with domain $A$, we have the following inequality:
-$$
-\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot \vert A \vert^{\length{\bar{y}}} \geq \sem{\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A)
-$$
-Note that the formula $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ defines a function in $\E{0}$.
-It was shown by the authors in \cite{SalujaST95} that every function in $\E{0}$ grows exponentially over the size of the structure for large enough structures, when $\length{\bar{X}} > 0$.
-This function divided by a polynomial factor still grows exponentially.
-Therefore, for $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ we have that $\length{\bar{X}} = 0$.
-
-Now, for the formula $\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ consider a structure $\mathfrak{1}$ with only one element $a$.
-We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$, but the only possible assignment to $\bar{x}$ is the tuple $(a,\ldots,a)$ so $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which follows to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop:qe0-fp-qe1-totp-fptras}}
-
-The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$.
-This is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every finite $\R$-structure $\A$ with domain $A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$.
-We base our proof on this fact.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is contained in \textsc{TotP}.} Let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Consider the following nondeterministic procedure that on input $\enc(\A)$ generates $\sem{\alpha}(\A)$ branches.
-For each $\alpha_i = \varphi$, where $\varphi$ is a $\loge{1}$ formula, it checks if $\A\models\varphi$ in polynomial time and generates a new branch if that is the case.
-For each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi$, this formula is also in $\E{1}$.
-We use the reduction to $\cdnf$ provided in \cite{SalujaST95} and we obtain $g(\enc(\A))$, which is an instance to $\cdnf$.
-Since $\cdnf$ is also in $\totp$ \cite{PagourtzisZ06}, we simulate the corresponding nondeterministic procedure that generates exactly $\cdnf(\enc(g(\A)))$ branches.
-Since, $\fp\subseteq\totp$\cite{PagourtzisZ06}, each polynomially computable function is also in $\totp$, and then on each of these branches we simulate the corresponding nondeterministic procedure to generate $h(\vert A \vert)$ more.
-The number of branches for each $\alpha_i$ is $\sem{\alpha_i}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$, and the total number of branches in the procedure amounts to $\sem{\alpha}(\A)$.
-We conclude that $\alpha\in\totp$.
-
-\vspace{1em}
-{\bf Every function in} $\boldsymbol{\eqso(\loge{1})}$ {\bf has an FPRAS.}  let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Note that each $\alpha_i$ that is equal to some $\loge{1}$ formula $\varphi$ has an FPRAS given by the procedure that simply checks if $\A\models\varphi$ and returns 1 if it does and 0 otherwise.
-Also, each remaining $\alpha_i$ has an FPRAS since $\alpha_i\in \E{1}$ \cite{SalujaST95}.
-If two functions have an FPRAS, then their sum also has one given by the procedure that simulates them both and sums the results.
-We conclude that $\alpha$ has an FPRAS.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is closed under sum and multiplication.} Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that the class is closed under multiplication.
-We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of $\so$.
-
-\begin{lemma} \label{conj-mult}
-If $\LL$ is a fragment closed under conjunction, then $\eqso(\LL)$ is closed under binary multiplication.
-\end{lemma}
-\begin{proof}
-We will define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ extended by binary product, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
-In fact, the formula $\tau(\alpha)$ is in $\LL$-SNF.
-First we replace each constant $s$ in $\alpha$ for $(\top \add \cdots \add \top)$ ($s$ times).
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$.
-We assume that for every $\beta$ that has less algebraic operators than $\alpha$, $\tau(\beta)$ is in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$ then we define $\tau(\alpha) = \tau(\alpha_1) + \tau(\alpha_2)$.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$, then we define $\tau(\alpha)$ as the formula in $\LL$-SNF that is equivalent to $\sa{x}\tau(\beta)$ and to $\tau(\alpha) = \sa{X}\tau(\beta)$, respectively.
-If $\alpha = (\alpha_1 \cdot \alpha_2)$, we assume that each $\alpha_i$ is in $\LL$-SNF.
-We identify three cases.
-(1) Some $\alpha_i$ is equal to $\sum_{j = 1}^n\beta_j$ for $n > 1$.
-Suppose wlog.
-that it is $\alpha_1$.
-We then define $\tau(\alpha) = \sum_{j = 1}^n\tau(\beta_j\cdot\alpha_2)$.
-In the following cases, $\alpha_1$ and $\alpha_2$ are in $\LL$-SNF.
-(2) If some $\alpha_i$ is equal to $\sa{X}\beta$ or $\sa{x}\beta$, we define $\tau(\alpha)$ as the $\LL$-SNF formula that is equivalent to $\sa{x}\tau(\beta\cdot\alpha_2)$ and $\sa{X}\tau(\beta\cdot\alpha_2)$, respectively.
-The remaining case is (3) $\alpha_1 = \varphi_1$ and $\alpha_2 = \varphi_2$ where each $\varphi$ is an $\LL$ formula.
-Then we define $\tau(\alpha) = \varphi_1 \wedge \varphi_2$.
-This covers all possible cases for $\alpha$.
-For every pair of formulas $\alpha,\beta$ in $\eqso(\LL)$, we have that their multiplication $(\alpha\cdot\beta)$ is a formula in the grammar $\eqso(\LL)$ extended by binary product, and so, there exists an equivalent formula $\tau(\alpha\cdot\beta)$ which is in unextended $\eqso(\LL)$.
-\end{proof}
-Since $\loge{1}$ is closed under conjunction, this also holds for $\eqso(\loge{1})$. This concludes the proof.
-
-\subsection*{Proof of Proposition \ref{pi-minusone}}
-
-Let $\LL$ be a fragment of $\fo$ that contains $\logu{1}$.
-Then we have that every function in $\U{1}$ is expressible in $\eqso(\LL)$.
-In particular, $\ctcnf \in \eqso(\LL)$.
-Suppose that $\eqso(\LL)$ is closed under subtraction by one.
-Then, the function $\ctcnf-1$, which counts the number of satisfying assignments of a 3-CNF formula minus one, is also in $\eqso(\LL)$.
-Note that $\eqso(\LL) \subseteq \eqso(\fo) = \shp$.
-We have that $\ctcnf$ is $\shp$-complete under parsimonious reductions\footnote{It can be easily verified that the standard reduction from SAT to 3-CNF (or 3-SAT) preserves the number of satisfying assignments}.
-Now, let $f$ be a function in $\shp$, and consider the nondeterministic polynomial-time procedure that on input $\enc(\A)$ computes the corresponding reduction to $\ctcnf$, name it $g(\enc(\A))$, and simulates the $\shp$ procedure for $\ctcnf-1$ on input $g(\enc(\A))$.
-We have that this is a $\shp$ procedure that computes $f-1$, from which we conclude that $\shp$ is closed under subtraction by one.
-
-\subsection*{Proof of Theorem \ref{sigmafo-minusone}}
-
 \noindent {\bf Closed under sum and multiplication.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
@@
 Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i^*(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
-
-%\subsection*{Proof of Theorem \ref{sub-pnp}}
-%
-%Towards a contradiction, suppose that any of the classes $\E{1}$, $\eqso(\loge{1})$, and $\eqso(\logex{1})$ is closed under subtraction.
-%
-%Let $\R$ include the symbols $S_1, S_2, S_3, S_4$, which describe the following properties.
-If a finite $\R$-structure $\A$ defines a 3DNF formula $\Phi$, then its domain is the set of variables mentioned in $\Phi$, and for each $i = 1,2,3,4$:
-%\begin{align*}
-%	S_1^\A &= \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_2^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_3^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_4^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
-%\end{align*}
-%Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
-%\begin{multline*}
-%\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
-%\end{multline*}
-%Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
-%$$
-%\varphi(x,X) = (T(x) \vee \neg T(x)).
-%$$
-%Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Given that $f_{\#3DNF}, f_{all} \in \E{1} \subseteq \eqso(\loge{1}) \subseteq \eqso(\logex{1}) \subseteq \totp$, and at least one of the classes is closed under subtraction, the function $f_{\#3DNF} - f_{all}$ is in $\totp$.
-However, note that for each $\R$-structure $\A$ that represents a formula $\Phi$, $(f_{\#3DNF} - f_{all})(\A) = 0$ if an only if $\Phi$ is a tautology.
-The decision version of this function is the $\np$-complete problem $\textsc{Tautology}$.
-And since the function is in $\totp$, its decision version is also in $\ptime$.
-We conclude that $\np \subseteq \ptime$.
-
-\subsection*{Proof of Proposition \ref{prop:ehorn-pe}}
-Pagourtzis and Zachos mention a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula \cite{PagourtzisZ06}.
-This procedure can be easily extended to receive Horn formulas, and furthermore, a disjunction of Horn formulas.
-Hence $\shdhsat$ is in $\totp$.
-
-As we show in Proposition \ref{sigma2hard}, $\shdhsat$ is complete for $\eqso(\ehorn)$ under parsimonious reductions.
-Let $\alpha$ be a formula in $\eqso(\ehorn)$ and let $g_{\alpha}$ be the reduction to $\shdhsat$.
-The $\totp$ procedure we construct, for each input $\enc(\A)$, is simply to compute $g_{\alpha}(\enc(\A))$, and then simulate the $\totp$ procedure for $\shdhsat$ on input $g_{\alpha}(\enc(\A))$.
-We conclude that $\alpha$ is in $\totp$.
-
-\subsection*{Proof of Proposition \ref{prop:hsat-not-sigma2}} %V.12
-We use a similar proof to the one provided by the authors in \cite{SalujaST95} to separate the classes $\E{2}$ and $\U{2}$.
-Suppose that the statement is false, this is, $\chsat \in \eqso(\loge{2})$.
-We consider the signature $\R$ that we used as the encoding for a Horn formula (Example \ref{ex-hornsat-esop1}) and that the formula $\alpha \in \eqso(\loge{2})$ follows the encoding in the same way.
-From what we proved in Theorem \ref{theo-pnf-snf}, we have that every formula in $\eqso(\loge{2})$ can be rewritten in $\loge{2}$-PNF, so we assume that $\alpha$ is in this form.
-Let $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\forall\bar{z}\,\varphi(\bar{X},\bar{x},\bar{y},\bar{z})$.
-Consider the following Horn formula $\Phi$:
-$$
-\Phi = p \wedge \bigwedge_{i = 1}^n (t_i \wedge p \to q) \wedge \neg q,
-$$
-where $n = \length{\bar{x}} + \length{\bar{y}} + 1$.
-Let $\A$ be the encoding of this formula.
-In the encoding, each variable appears as an element in the domain of $\A$.
-This formula is satisfiable, so $\sem{\alpha}(\A) \geq 1$.
-Let $(\bar{B},\bar{b},\bar{a})$ be an assignment to $(\bar{X},\bar{x},\bar{y})$ such that $\A\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-Let $t_{\ell}$ be such that it does not appear in $\bar{b}$ or $\bar{a}$.
-Consider the induced substructure $\A'$ that is obtained by removing $t_{\ell}$ from $\A$ and $\bar{B}'$ as the subset of $\bar{B}$ obtained by deleting each appearance of $t_{\ell}$ in $\bar{B}$.
-We have that $\A'\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-This is because each subformula of the form $\exists y \neg B_i$ is still true, and universal formulas are monotone over induced substructures.
-It follows that $\sem{\alpha}(\A') \geq 1$ which is not possible since $\A'$ encodes the formula
-$$
-\Phi' = p \wedge \bigwedge_{i = 1}^{\ell-1} (t_i \wedge p \to q) \wedge (p\to q) \wedge \bigwedge_{i = \ell+1}^{n} (t_i \wedge p \to q) \wedge \neg q,
-$$
-which is unsatisfiable. We arrive to a contradiction and we conclude that $\chsat$ is not in $\eqso(\ehorn)$.
-
-\subsection*{Proof of Theorem \ref{sigma2hard}} %V.13
-
-First we prove that $\shdhsat$ is in $\eqso(\ehorn)$.
-Recall that each instance of $\shdhsat$ is a disjunction of Horn formulas.
-Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot), \pD(\cdot,\cdot)\}$.
-Each symbol in this vocabulary is used to indicate the same as in Example \ref{ex-hornsat-esop1}, with the addition of $\pD(d,c)$ which indicates that $c$ is a clause in the formula $d$.
-Recall that the formula
-\begin{align*}
-&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \neg\textit{T}(x)).
-\end{align*}
-defines $\chsat$. We obtain the following formula $\psi(T,A)$ in $\ehorn$:
-\begin{align*}
-\exists d[&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \pD(c,d)\vee \neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \neg\textit{T}(x))]
-\end{align*}
-effectively defines $\chsat$ as for every disjunction of Horn formulas $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
-Therefore, we conclude that $\shdhsat \in \eqso(\ehorn)$.
-
-\vspace{1em}
-We will now prove that $\shdhsat$ is hard for $\eqso$ over a signature $\R$ under parsimonious reductions.
-For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a polynomial-time procedure that computes a function $g_{\alpha}$.
-This function receives a finite $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
-We suppose that $\alpha$ is in sum normal form and:
-$$
-\alpha = \sum_{i = 1}^{\text{\#clauses}} \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigwedge_{j = 1}^{n}\forall\bar{z}\,\varphi^i_j(\bar{X},\bar{x},\bar{y},\bar{z}),
-$$
-where each $\varphi^i_j$ is a Horn clause.
-
-Consider a finite $\R$-structure $\A$ with domain $A$.
-To simplify the proof, we extend our grammar to allow first-order constants.
-Consider each tuple $\bar{a}\in A^{\length{\bar{x}}}$, each $\bar{b}\in A^{\length{\bar{y}}}$ and each $\bar{c}\in A^{\length{\bar{z}}}$ as a tuple of first-order constants.
-The following formula defines the same function as $\alpha$:
-$$
-\sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length{\bar{x}}}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length{\bar{y}}}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length{\bar{z}}}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
-$$
-Note that each $\fo$ formula over $(\bar{x},\bar{y},\bar{z})$ in each $\varphi^i_j$ has no free variables.
-Therefore, we can evaluate each of these in polynomial time and replace them by $\perp$ and $\top$ where it corresponds.
-Each $\varphi^i_j$ will be of the form $\perp \vee\, \chi^i_j(\bar{X})$ or $\top \vee \chi^i_j(\bar{X})$ where $\chi^i_j$ is a disjunction of $\neg X_{\ell}$'s and at most one $X_{\ell}$.
-The formulas of the form $\top \vee \chi^i_j(\bar{X})$ can be removed entirely, and the formulas of the form $\perp \vee\, \chi^i_j(\bar{X})$ can be replaced by $\chi^i_j(\bar{X})$.
-We obtain the formula
-$$
-\sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})
-$$
-where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$.
-
-Our idea for the rest of the proof is to define $g_{\alpha}$ for each $\alpha = \sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})$, for $\alpha = c$ and for $\alpha = \beta_1 + \cdots + \beta_m$ where each $\beta_i$ is in one of the two previous cases.
-
-If $\alpha$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ where $\psi_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$, then we obtain the {\bf propositional formula} $g_{\alpha}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulas.
-
-If $\alpha$ is equal to a constant $c$, then we define $g_{\alpha}(\A)$ as the following formula that has exactly $c$ satisfying assignments:
-$$
-g_{\alpha}(\A) = \bigvee_{i = 1}^{c}\neg t_1 \wedge \cdots \wedge \neg t_{i-1} \wedge t_i \wedge \neg t_{i+1} \wedge \cdots \wedge \neg p_c.
-$$ 
-If $\alpha = \beta_1 + \cdots + \beta_m$, let $g_{\beta_i}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\theta^i_{j,k}$ for each $\beta_i$ where each $\theta^i_{j,k}$ is a Horn clause.
-Let $\Theta_i = g_{\beta_i}(\A)$.
-We rename the variables in each $\Theta_i$ so none of them are mentioned in any other $\Theta_j$.
-We add $m$ new variables $t_1,\ldots,t_m$ and we define:
-\begin{align*}
-g_{\alpha}(\A) = &\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^1_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m}}}t) \wedge (t_1 \wedge \bigwedge_{\ell = 2}^{m} \neg t_{\ell})) \vee \\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^2_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each $t$}\\ \text{ mentioned in}\\ \Theta_1,\Theta_3,\ldots,\Theta_{m}}}t) \wedge (t_2 \wedge \bigwedge\limits_{\substack{\ell = 1 \\ \ell \neq 2}}^{m} \neg t_{\ell})) \vee \cdots \vee\\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^m_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m-1}}}t) \wedge (t_m \wedge \bigwedge_{\ell = 1}^{m-1} \neg t_{\ell})).
-\end{align*}
-The formula is a disjunction of Horn formulas, and the number of satisfying assignments for this formula is exactly the sum of satisfying assignments for each $g_{\beta_i}(\A)$.
-This, at the same time, is equal to $\sem{\alpha}(\A)$.
-This covers all possible cases for $\alpha$, and the entire procedure takes polynomial time.

### 2017-03-03 LMCS/proof-theo-pi1-pnf.tex
journal primera version

--- 2017-01-12 LICS/appendix-syntactic.tex
@@
-\subsection*{Proof of Theorem \ref{theo-pnf-snf}}
-Recall that a formula in $\eqso(\LL)$ is on the following grammar:
-$$
-\alpha = \varphi \ \mid \ s \ \mid \ (\alpha + \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \sa{X} \alpha,
-$$
-where $\varphi$ is a formula in $\LL$ and $s\in\nat$.
-We will construct a recursive function $\tau$ such that for every $\eqso(\LL)$ formula $\alpha$, it outputs an equivalent formula $\tau(\alpha)$ which is in $\LL$-SNF.
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$ which is clearly equivalent and already in $\LL$-SNF.
-If $\alpha = s$, then we define $(\top \add \cdots \add \top)$ ($s$ times), which also satisfies the condition.
-We assume that for every sub-formula $\beta$ of $\alpha$, $\tau(\beta)$ is an equivalent formula in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$, then we define $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$.
-If $\alpha = \sa{x}\beta$, then let $\tau(\beta) = \sum_{i = 1}^{k}\beta_i$ for some $k$ where each $\beta_i$ is in $\LL$-PNF.
-By grouping together the terms in the sum, we define $\tau(\alpha) = \sum_{i = 1}^{m}\sa{x}\beta_i$ which satisfies the condition and is equivalent to $\alpha$.
-If $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case.
-This covers all possible cases for $\alpha$ and we conclude the proof by taking $\tau(\alpha)$ as the desired rewrite of $\alpha$.
-
-\subsection*{Proof of theorem \ref{theo-pi1-pnf}}
 We divide the proof in three parts.
@@
 If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we can consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the condition in the hypothesis.
-
-\subsection*{Proof of Theorem \ref{prop-rest}}
-We give this proof in three parts.
-
-\vspace{1em}
-First, we show that $\QE{0} \not\subseteq \E{1}$.
-By contradiction, suppose that there is a $\QE{0}$ formula $\alpha$ over some signature $\R$ such that defines the following function.
-For every finite $\R$-structure with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n - 1$.
-We use the following claim.
-\begin{claim}
-Let $\alpha = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. Then the function defined by $\alpha$ is either null, greater or equal to $n$, or is in $\Omega(n^2)$.
-\end{claim}
-\begin{proof}
-Suppose that the function defined by $\alpha$ is not $0$ and that $\varphi$ is in DNF. Furthermore, suppose $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$.
-Then $\alpha = \sa{\bar{x}} \varphi_1(\bar{x}) \vee \cdots \vee \varphi_n(\bar{x})$.
-Since $\alpha$ is not null, then some $\varphi_i$ must be satisfiable.
-This is, the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is not null.
-We will prove by induction on $\length{\bar{x}}$ that the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is either greater or equal to $n$, or in $\Omega(n^2)$.
-We address the case $\length{\bar{x}}= 1$, then $\alpha = \sa{x}\bigwedge\psi(x)$.
-If any $\psi(x) = (x = x)$ or $\neg(x < x)$, then we can eliminate it and we obtain the same function.
-If any $\psi = (x < x)$ or $\neg(x=x)$, then the function becomes null.
-If $\psi(x) = R(x,\ldots,x)$ for some $R\in\R$ the function becomes null for the structures we are considering.
-If $\psi(x) = \neg R(x,\ldots,x)$, we can eliminate it and for the structures we are considering we obtain the same function.
-The only possible $\alpha$ left is $\alpha = \sa{x}\top$ which is equal to the function $n$.
-This covers all possible cases for $\length{\bar{x}} = 1$.
-Now suppose that it holds for $\length{\bar{x}} = k$ and suppose $\alpha = \sa{\bar{x}}\bigwedge\psi(\bar{x})$ for $\length{\bar{x}} = k+1$.
-If any $\psi(\bar{x}) = (x_i = x_j)$ where $i \neq j$, then $\alpha$ describes the same function as $\alpha$ where $x_j$ has been replaced by $x_i$.
-In this formula the tuple of first-order variables has $k$ elements so the function it describes if one of the mentioned in the hypothesis.
-If $i = j$, then we can eliminate it and obtain the same function.
-If any $\psi(\bar{x}) = R(\bar{v})$ or $\neg R(\bar{v})$ where $\bar{v}$ is a sub-tuple of $\bar{x}$ then we can either eliminate it or the function becomes null, following the same argument as in the case $\length{\bar{x}} = 1$.
-If any $\psi(\bar{x}) = \neg(x_i = x_j)$ or $(x_i < x_j)$ where $i = j$, then the function becomes null.
-If any $\psi(\bar{x}) = \neg(x_i < x_j)$ where $i = j$, we can eliminate it.
-The remaining formulas in $\bigwedge\psi(\bar{x})$ are either $\neg(x_i = x_j)$, $(x_i<x_j)$ or $\neg(x_i<x_j)$.
-If the formula violates transitivity in $<$ (for example, $x < y \wedge y < z \wedge z < x$), then the function $\alpha$ describes is null.
-Therefore, there is some order over $\bar{x}$ that satisfies $\bigwedge\psi(\bar{x})$.
-Consider the formula that describes this order (like $x_1 < x_3 \wedge x_3 < x_4 \wedge x_4 < x_2$).
-The function $\alpha$ describes is greater or equal to the one this formula describes, which is exactly $\binom{n}{\length{\bar{x}}}$ which is in $\Omega(n^{\length{\bar{x}}}) \subseteq \Omega(n^2)$ if $\length{\bar{x}} > 1$.
-This concludes the proof of the claim.
-\end{proof}
-We suppose that $\alpha$ is in SNF, this is, $\alpha = \sum_{i = 1}^n\alpha_i$.
-Since $\alpha$ is not null, consider some $\alpha_i$ that describes a non-null function.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, where $\varphi$ is quantifier-free.
-Note that if $\length{\bar{X}} > 0$, then the function $\alpha$ describes is in $\Omega(2^n)$, as it was proven by the authors in \cite{SalujaST95}.
-We have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$, as we proved in the claim, describes either some function greater or equal to $n$, or in $\Omega(n^2)$, which leads to a contradiction.
-Lastly, note that the formula $\sa{x}\exists y(x < y)$ is in $\E{0}$ and describes the function $n-1$, which concludes the proof.
-
-\vspace{1em}
-Now we show that $\E{1}\not\subseteq\QE{0}$.
-In Theorem \ref{theo-pi1-pnf} we proved that there is no formula in $\loge{1}$-PNF equivalent to the formula $\alpha = 2$.
-Every formula in $\E{1}$ can be expressed in $\loge{1}$-PNF, which implies that $2 \in \QE{0}$ and $2 \not\in \E{1}$.
-
-\vspace{1em}
-Lastly, we prove that $\eqso(\loge{1})\subsetneq\eqso(\logu{1})$.
-For inclusion, let $\alpha$ be a formula in $\eqso(\loge{1})$.
-Suppose that it is in $\loge{1}$-SNF.
-This is, $\alpha = c + \sum_{i = 1}^{n}\alpha_i$.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, where $\varphi_i$ is quantifier-free, for each $\alpha_i$.
-We use the same construction used in \cite{SalujaST95}, and we obtain that the formula $\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \forall\bar{y}'(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')$ for every assignment to $(\bar{X},\bar{x})$.
-We do this replacement for each $\alpha_i$ and we obtain an equivalent formula in $\eqso(\logu{1})$.
-
-To prove that the inclusion is proper, consider the $\eqso(\logu{1})$ formula $\sa{x}\forall y(y = x)$. This formula defines the following function that takes an ordered structure $\A$ as input:
-$$
-\sem{\alpha}(\A) = 
-\begin{cases}
-1 &\A \text{ has one element}\\
-0 &\text{ otherwise}.
-\end{cases}
-$$
-Suppose that there exists an equivalent formula $\alpha$ in $\eqso(\loge{1})$.
-Also, suppose that it is in $\L$-PNF, so $\alpha = c + \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$.
-Since $\alpha$ takes the value 0 for some structures, $c$ must be 0.
-Consider a structure $\mathfrak{1}$ with one element.
-We have that for some $i$, there exists an assignment $(\bar{B},\bar{b},\bar{a})$ for $(\bar{X},\bar{x},\bar{y})$ such that $\mathfrak{1}\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
-Consider now the structure $\mathfrak{2}$ that is obtained by duplicating $\mathfrak{1}$, as we did for Theorem \ref{theo-pi1-pnf}.
-Note that $\mathfrak{2}\models\varphi_i(\bar{B},\bar{b},\bar{a})$, which implies that $\sem{\alpha}(\mathfrak{2}) \geq 1$, which leads to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop-e1-nc}}
-
-Towards a contradiction, assume that the statement is false.
-This is, that $\E{1}$ is closed under binary sum.
-Consider the formula $\sa{x}(x = x)$ which is in $\E{1}$ over some signature $\R$.
-For every finite $\R$-structure $\A$ with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n$.
-From our assumption, there exists some formula in $\E{1}$ equivalent to the formula $\alpha \add \alpha$, which describes the function $2n$.
-Let $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ be this formula, where $\varphi$ is quantifier-free.
-Note that the function defined by this formula is equal or greater than the one defined by $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ divided by a polynomial factor.
-More specifically, for each ordered structure $\A$ with domain $A$, we have the following inequality:
-$$
-\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot \vert A \vert^{\length{\bar{y}}} \geq \sem{\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A)
-$$
-Note that the formula $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ defines a function in $\E{0}$.
-It was shown by the authors in \cite{SalujaST95} that every function in $\E{0}$ grows exponentially over the size of the structure for large enough structures, when $\length{\bar{X}} > 0$.
-This function divided by a polynomial factor still grows exponentially.
-Therefore, for $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ we have that $\length{\bar{X}} = 0$.
-
-Now, for the formula $\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ consider a structure $\mathfrak{1}$ with only one element $a$.
-We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$, but the only possible assignment to $\bar{x}$ is the tuple $(a,\ldots,a)$ so $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which follows to a contradiction.
-
-\subsection*{Proof of Proposition \ref{prop:qe0-fp-qe1-totp-fptras}}
-
-The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$.
-This is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every finite $\R$-structure $\A$ with domain $A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$.
-We base our proof on this fact.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is contained in \textsc{TotP}.} Let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Consider the following nondeterministic procedure that on input $\enc(\A)$ generates $\sem{\alpha}(\A)$ branches.
-For each $\alpha_i = \varphi$, where $\varphi$ is a $\loge{1}$ formula, it checks if $\A\models\varphi$ in polynomial time and generates a new branch if that is the case.
-For each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi$, this formula is also in $\E{1}$.
-We use the reduction to $\cdnf$ provided in \cite{SalujaST95} and we obtain $g(\enc(\A))$, which is an instance to $\cdnf$.
-Since $\cdnf$ is also in $\totp$ \cite{PagourtzisZ06}, we simulate the corresponding nondeterministic procedure that generates exactly $\cdnf(\enc(g(\A)))$ branches.
-Since, $\fp\subseteq\totp$\cite{PagourtzisZ06}, each polynomially computable function is also in $\totp$, and then on each of these branches we simulate the corresponding nondeterministic procedure to generate $h(\vert A \vert)$ more.
-The number of branches for each $\alpha_i$ is $\sem{\alpha_i}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$, and the total number of branches in the procedure amounts to $\sem{\alpha}(\A)$.
-We conclude that $\alpha\in\totp$.
-
-\vspace{1em}
-{\bf Every function in} $\boldsymbol{\eqso(\loge{1})}$ {\bf has an FPRAS.}  let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
-Note that each $\alpha_i$ that is equal to some $\loge{1}$ formula $\varphi$ has an FPRAS given by the procedure that simply checks if $\A\models\varphi$ and returns 1 if it does and 0 otherwise.
-Also, each remaining $\alpha_i$ has an FPRAS since $\alpha_i\in \E{1}$ \cite{SalujaST95}.
-If two functions have an FPRAS, then their sum also has one given by the procedure that simulates them both and sums the results.
-We conclude that $\alpha$ has an FPRAS.
-
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is closed under sum and multiplication.} Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that the class is closed under multiplication.
-We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of $\so$.
-
-\begin{lemma} \label{conj-mult}
-If $\LL$ is a fragment closed under conjunction, then $\eqso(\LL)$ is closed under binary multiplication.
-\end{lemma}
-\begin{proof}
-We will define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ extended by binary product, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
-In fact, the formula $\tau(\alpha)$ is in $\LL$-SNF.
-First we replace each constant $s$ in $\alpha$ for $(\top \add \cdots \add \top)$ ($s$ times).
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$.
-We assume that for every $\beta$ that has less algebraic operators than $\alpha$, $\tau(\beta)$ is in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$ then we define $\tau(\alpha) = \tau(\alpha_1) + \tau(\alpha_2)$.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$, then we define $\tau(\alpha)$ as the formula in $\LL$-SNF that is equivalent to $\sa{x}\tau(\beta)$ and to $\tau(\alpha) = \sa{X}\tau(\beta)$, respectively.
-If $\alpha = (\alpha_1 \cdot \alpha_2)$, we assume that each $\alpha_i$ is in $\LL$-SNF.
-We identify three cases.
-(1) Some $\alpha_i$ is equal to $\sum_{j = 1}^n\beta_j$ for $n > 1$.
-Suppose wlog.
-that it is $\alpha_1$.
-We then define $\tau(\alpha) = \sum_{j = 1}^n\tau(\beta_j\cdot\alpha_2)$.
-In the following cases, $\alpha_1$ and $\alpha_2$ are in $\LL$-SNF.
-(2) If some $\alpha_i$ is equal to $\sa{X}\beta$ or $\sa{x}\beta$, we define $\tau(\alpha)$ as the $\LL$-SNF formula that is equivalent to $\sa{x}\tau(\beta\cdot\alpha_2)$ and $\sa{X}\tau(\beta\cdot\alpha_2)$, respectively.
-The remaining case is (3) $\alpha_1 = \varphi_1$ and $\alpha_2 = \varphi_2$ where each $\varphi$ is an $\LL$ formula.
-Then we define $\tau(\alpha) = \varphi_1 \wedge \varphi_2$.
-This covers all possible cases for $\alpha$.
-For every pair of formulas $\alpha,\beta$ in $\eqso(\LL)$, we have that their multiplication $(\alpha\cdot\beta)$ is a formula in the grammar $\eqso(\LL)$ extended by binary product, and so, there exists an equivalent formula $\tau(\alpha\cdot\beta)$ which is in unextended $\eqso(\LL)$.
-\end{proof}
-Since $\loge{1}$ is closed under conjunction, this also holds for $\eqso(\loge{1})$. This concludes the proof.
-
-\subsection*{Proof of Proposition \ref{pi-minusone}}
-
-Let $\LL$ be a fragment of $\fo$ that contains $\logu{1}$.
-Then we have that every function in $\U{1}$ is expressible in $\eqso(\LL)$.
-In particular, $\ctcnf \in \eqso(\LL)$.
-Suppose that $\eqso(\LL)$ is closed under subtraction by one.
-Then, the function $\ctcnf-1$, which counts the number of satisfying assignments of a 3-CNF formula minus one, is also in $\eqso(\LL)$.
-Note that $\eqso(\LL) \subseteq \eqso(\fo) = \shp$.
-We have that $\ctcnf$ is $\shp$-complete under parsimonious reductions\footnote{It can be easily verified that the standard reduction from SAT to 3-CNF (or 3-SAT) preserves the number of satisfying assignments}.
-Now, let $f$ be a function in $\shp$, and consider the nondeterministic polynomial-time procedure that on input $\enc(\A)$ computes the corresponding reduction to $\ctcnf$, name it $g(\enc(\A))$, and simulates the $\shp$ procedure for $\ctcnf-1$ on input $g(\enc(\A))$.
-We have that this is a $\shp$ procedure that computes $f-1$, from which we conclude that $\shp$ is closed under subtraction by one.
-
-\subsection*{Proof of Theorem \ref{sigmafo-minusone}}
-
-\noindent {\bf Closed under sum and multiplication.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
-This can be seen because $\eqso(\LL)$ is closed under sum for every fragment $\LL$ by definition, and since $\logex{1}$ is closed under conjunction, from Lemma \ref{conj-mult} it follows that $\eqso(\logex{1})$ is closed under multiplication.
-
-\vspace{1em}
-\noindent {\bf Easy decision version and FRPRAS.}  We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
-We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Theorem \ref{prop:qe0-fp-qe1-totp-fptras}.
-First, we define a function that converts a formula $\alpha$ in $\eqso(\logex{1})$ over a signature $\R$ into a formula $\lambda(\alpha)$ in $\eqso(\loge{1})$ over a signature $\R_{\alpha}$.
-Afterwards, we define a function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an $\R_{\alpha}$-structure $g_{\alpha}(\A)$.
-
-Let $\alpha$ be in $\eqso(\logex{0})$.
-The signature $\R_{\alpha}$ is obtained by adding the symbol $R_{\psi}$, for every $\fo$ formula $\psi(\bar{z})$ in $\alpha$, to $\R$.
-Each symbol $R_{\psi}$ represents a predicate with arity $\length{\bar{z}}$.
-Then, $\lambda(\alpha)$ is defined as $\alpha$ where each $\fo$ formula $\psi(\bar{z})$ has been replaced by $R_{\psi}(\bar{z})$.
-We now define the function $g_{\alpha}$ procedurally.
-Let $\A$ be a $\R$-structure with domain $A$.
-Let $\A'$ be an $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A}$ empty.
-For each $\fo$-formula $\psi(\z)$ with $\length{\bar{z}}$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length{\bar{z}}}$.
-If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$ (this can be done in $\ptime$).
-This concludes the construction of $\A'$.
-Note that the number of $\fo$ subformulas, arity and tuple size is fixed in $\alpha$, so computing this function takes polynomial time over the size of the structure.
-Moreover, the encoding of $\A'$ has polynomial size over the size of $\enc(\A)$.
-We define $g_{\alpha}(\A) = \A'$ and we have that for each $\R$-structure $\A$: $\sem{\alpha}(\A) = \sem{\lambda(\alpha)}(g_{\alpha}(\A))$.
-Therefore, we have a parsimonious reduction from $\alpha$ to the $\eqso(\loge{1})$ formula $\lambda(\alpha)$.
-
-To show that $\alpha$ is in $\totp$, we can convert $\alpha$ and $\enc(\A)$ into $\lambda(\alpha)$ and $\enc(g_{\alpha}(\A))$, respectively, and run the procedure in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-Similarly, to show that $\alpha$ has an FPRAS, we do the same as before and simulates the FPRAS for $\lambda(\alpha)$ in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-These procedures also takes polynomial time and satisfies the condition.
-
-\vspace{1em}
-\noindent {\bf Closed under subtraction by one.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. 
-For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Without lost of generality, we assume that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
-Moreover, we assume that $\length{\bar{x}} > 0$  since, if this is not the case, we can replace $\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
-
-The proof will be separated in two parts.
-In the first part, we suppose that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$.
-Then we will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ is the formula.
-In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ the sum of one or more formulas in $\logex{1}$-PNF.
-We define a formula $\varphi'$ that satisfies the following condition: if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ and $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be proven.
-
-\vspace{1em}
-
-\noindent {\em Part (1).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ where $\varphi$ is an  $\fo$-formula.
-Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}} \varphi(\bar{x})$ (i.e.
-$\length{\bar{X}} = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}} [\varphi(\bar{x})\wedge\exists\bar{z}(\varphi(\bar{z})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
-Therefore, for the rest of the proof we can assume that $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
-
- Let $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$ and $\bar{X} = (X_1,\ldots,X_{\length{\bar{X}}})$.
-To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in into a DNF formula. 
-More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i$ is a conjunction of atomic formulas or negation of atomic formulas.
-Furthermore, we suppose that each conjunct $\varphi_i$ has the~form:
-$$
-\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \underbrace{\varphi_i^{\fo}(\bar{x},\bar{y})}_{\text{an $\fo$ formula}} \wedge 
-\underbrace{\varphi_i^{+}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $X_j$'s}} \wedge
-\underbrace{\varphi_i^{-}(\bar{X},\bar{x},\bar{y})}_{\text{conjunction of $\neg X_j$'s}}.
-$$
-Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ can appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
-
-Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ are mentioned in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
-We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
-\begin{itemize}
-	\item[(a)] In order to satisfy the first condition, consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
-We add $\length{\bar{w}}$ new variables $z_1,\ldots,z_{\length{\bar{w}}}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length{\bar{w}}})$.
-We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$.
-The formula $\varphi_i$ is now defined as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \exists \bar{z} \; (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{-}(\bar{X},\bar{x},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y})).
-	$$
-	We repeat this process for each instance of a $X_j(\bar{w})$ in $\varphi_i$, and we obtain a formula where none of the $X_j$'s acts over any variable in $\bar{x}$.
-We add the new first-order variables to $\bar{y}$ and we redefine $\varphi_i$ as
-	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
-	$$
-	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = \exists \bar{y} \;  X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := \exists \bar{y} \;  X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
-	\item[(b)] An ordered partition on a set $S$ is defined by an equivalence relation $\sim$ over $S$, and a linear order over $S/\!\sim$.
-For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$.
-A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
-On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
-	For a given $k$, let $\cB_k$ be the number of possible ordered partitions for a set of size $k$. For $1 \leq j \leq \cB_{\length{(\bar{x},\bar{y})}}$ 
-	let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
-
-	Thus, the formula $\varphi_i(\bar{X},\bar{x})$ is then redefined as:
-	$$
-	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length{(\bar{x},\bar{y})}}} \exists \bar{y} \; [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
-	$$
-	Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
-	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  to $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
-	\item[(c)] For showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
-$\varphi_i$ is inconsistent), we do the following.
-If there exists an instance of $X_j(\bar{z})$ in $\varphi^{+}_i$, an instance of $\neg X_j(\bar{w})$ in $\varphi^{-}_i$ and the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$, then the entire formula $\varphi_i$ is removed from $\varphi$.
-\end{itemize}
-It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$.
-From now on, we assume that each $\varphi_i(\bar{X},\bar{x},\bar{y})$ satisfies conditions (a), (b) and (c).
-Furthermore, it has the following form:
-$$
-\varphi_i(\bar{X},\bar{x}) \; = \; \exists{\bar{y}} \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
-$$
-where $\varphi^{+}$ and $\varphi^{-}_i$ do not depend on  $\bar{x}$.
-\begin{claim}\label{claim:minusone}
-For a given ordered structure $\A$, if there exists a first-order assignment $v$ for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$, then there exists a second-order assignment $V$ for $\A$ such that  $(\A,V,v)\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-\end{claim}
-\begin{proof}
-Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$ such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
-Define $\bar{B} = (B_1,\ldots,B_{\length{\bar{X}}})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
-Towards a contradiction, suppose that $(\A,V,v)\not\models \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-By the choice of $v$, and construction of $V$ it is clear that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})\wedge\varphi^{+}_i(\bar{X},\bar{y})$, so we necessarily have that $(\A,V,v)\not\models\varphi^{-}_i(\bar{X},\bar{y})$.
-Let $X_t$ be such that $\neg X_t(\bar{w})$ is mentioned in $\varphi^{-}_i(\bar{X},\bar{y})$ and $(\A,V, v)\not\models\neg X_t(\bar{w})$, namely, $v(\bar{w})\in B_t$. 
-However, by the construction of $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$.
-Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$.
-This violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
-\end{proof}
-
-The previous claim and proof motivates the following definitions.
-For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length{\bar{X}}})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
-One can easily check that for every assignments $V, v$ such that $(\A,V,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$, then it holds that $(\A,\bar{B}^v,v)\models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all the assignment of $\bar{X}$ when $v$ is fixed.
-Therefore, by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
-For this, the idea is to choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
-$$
-\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge\forall\bar{x}'\forall\bar{y}'(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')).
-$$ 
-If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
-Furthermore, let $V^*$ the second order assignment that satisfies $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*$ the first order assignment such that $v^*(\bar{x}) = v(\bar{x})$.
-By the previous discussion, we have that $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.
-
-Now, we have all the tools to introduce the main formula of the proof. 
-Intuitively, we want to exclude the assignment $(V^*, v^*)$ as a satisfying assignment for $\varphi_i(\bar{X},\bar{x})$.
-Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if, 
-if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, then $V \neq V^*$ or $v \neq v^*$. 
-That is, it excludes all pairs $(V, v)$ that coincide with $(V^*, v^*)$ in $(\bar{X}, \bar{x})$ whenever $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. This property can be defined as follows:
-\begin{align}
-\psi_i(\bar{X},\bar{x}) = & \;\; \big( \, \exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
-& \hspace{1cm} \Big( \; 
-\exists\bar{y} \;\; \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i^+(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\
-& \hspace{1.5cm} (\exists\bar{x}'\exists\bar{y}\, \varphi_i(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}) \;\; \Big)  \label{eq:part3}
-\end{align}
-To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
-Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y})$.
-Then, the conclusion of the implication at (\ref{eq:part2}) and (\ref{eq:part3}), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
-Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by showing that if $\varphi_i(\bar{X},\bar{x})$ is satisfied then $V(\bar{X})$ must contain one element more that it is not in $V^*(\bar{X})$.
-The second disjunct (\ref{eq:part3}) is satisfied when $v(\bar{x})$ is not the lexicographically smallest tuple that satisfies $\varphi_i$ (i.e. $v(\bar{x}) \neq v^*(\bar{x})$).
-
-We are ready to define the formula  $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$
-where each modified disjunct $\varphi_i'(\bar{X},\bar{x})$ is constructed as follows. 
-For the sake of simplification, define the auxiliary formula $\chi_i = \neg\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$. 
-This formula basically checks if $\varphi_i$ is satisfiable or not.
-Then, first define the formula:
-$$
-\varphi'_1(\bar{X},\bar{x}) = \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
-$$
-This models that all the assignments that satisfy $\varphi_1$ also satisfy $\varphi'_1$, except for $(V^*,v^*)$. Then we define:
-$$
-\varphi'_2(\bar{X},\bar{x}) = \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
-$$
-This models that all the assignments that satisfy $\varphi_2$ also satisfy $\varphi'_2$, but if there was no assignment for $\varphi_1$, then we exclude $(V^*,v^*)$ for $\varphi_2$.
-Then we generalize this construction for each $\varphi_i$:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})),
-$$
-By the construction of $\kappa(\alpha)$, one can easily check that  $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.
-
-\vspace{1em}
-
-{\em Part (2).} Let $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ for some $\eqso(\logex{1})$ formula $\beta$. We define $\kappa(\alpha)$ as follows.
-First, perform the same transformations to $\varphi(\bar{X},\bar{x})$ as in the previous section.
-Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c).
-We also use the previously defined formulas $\chi_i$ and $\psi_i$, for each $i$.
-
-We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if and only if $\sem{\beta}(\A) = 0$.
-If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg\exists\bar{x}'\varphi(\bar{x}')$.
-If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$ as each $\chi_i$ was previously defined.
-If $\beta = (\beta_1 + \beta_2)$, then $\lambda(\beta) = \lambda(\beta_1) \wedge \lambda(\beta_2)$.
-Now, for each $\varphi_i$ we define:
-$$
-\varphi_i'(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x}))\Big)\Big).
-$$
-Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i'(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
-
-%\subsection*{Proof of Theorem \ref{sub-pnp}}
-%
-%Towards a contradiction, suppose that any of the classes $\E{1}$, $\eqso(\loge{1})$, and $\eqso(\logex{1})$ is closed under subtraction.
-%
-%Let $\R$ include the symbols $S_1, S_2, S_3, S_4$, which describe the following properties.
-If a finite $\R$-structure $\A$ defines a 3DNF formula $\Phi$, then its domain is the set of variables mentioned in $\Phi$, and for each $i = 1,2,3,4$:
-%\begin{align*}
-%	S_1^\A &= \{(a_1,a_2,a_3)\mid (\neg a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_2^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge \neg a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_3^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge \neg a_3) \mbox{ appears as a disjunct in }\Phi\},\\
-%	S_4^\A &= \{(a_1,a_2,a_3)\mid ( a_1 \wedge  a_2 \wedge  a_3) \mbox{ appears as a disjunct in }\Phi\}.
-%\end{align*}
-%Now we define $f_{\#3DNF} = f_{\psi(T)}$ where
-%\begin{multline*}
-%\psi(T) = \exists x \exists y \exists z\, [(S_1(x,y,z) \wedge \neg T(x) \wedge \neg T(y) \wedge \neg T(z)) \vee (S_2(x,y,z) \wedge T(x) \wedge \neg T(y) \wedge \neg T(z)) \, \vee \\ (S_3(x,y,z) \wedge T(x) \wedge T(y) \wedge \neg T(z)) \vee (S_4(x,y,z) \wedge T(x) \wedge T(y) \wedge T(z))].
-%\end{multline*}
-%Note that $f_{\#3DNF} \in \#\Sigma_1$. Let $f_{all} = f_{\exists x\:\varphi(x,X)}$, where
-%$$
-%\varphi(x,X) = (T(x) \vee \neg T(x)).
-%$$
-%Note that $f_{all}$ counts every possible truth assignment (satisfying or not) to a 3DNF formula.
-Given that $f_{\#3DNF}, f_{all} \in \E{1} \subseteq \eqso(\loge{1}) \subseteq \eqso(\logex{1}) \subseteq \totp$, and at least one of the classes is closed under subtraction, the function $f_{\#3DNF} - f_{all}$ is in $\totp$.
-However, note that for each $\R$-structure $\A$ that represents a formula $\Phi$, $(f_{\#3DNF} - f_{all})(\A) = 0$ if an only if $\Phi$ is a tautology.
-The decision version of this function is the $\np$-complete problem $\textsc{Tautology}$.
-And since the function is in $\totp$, its decision version is also in $\ptime$.
-We conclude that $\np \subseteq \ptime$.
-
-\subsection*{Proof of Proposition \ref{prop:ehorn-pe}}
-Pagourtzis and Zachos mention a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula \cite{PagourtzisZ06}.
-This procedure can be easily extended to receive Horn formulas, and furthermore, a disjunction of Horn formulas.
-Hence $\shdhsat$ is in $\totp$.
-
-As we show in Proposition \ref{sigma2hard}, $\shdhsat$ is complete for $\eqso(\ehorn)$ under parsimonious reductions.
-Let $\alpha$ be a formula in $\eqso(\ehorn)$ and let $g_{\alpha}$ be the reduction to $\shdhsat$.
-The $\totp$ procedure we construct, for each input $\enc(\A)$, is simply to compute $g_{\alpha}(\enc(\A))$, and then simulate the $\totp$ procedure for $\shdhsat$ on input $g_{\alpha}(\enc(\A))$.
-We conclude that $\alpha$ is in $\totp$.
-
-\subsection*{Proof of Proposition \ref{prop:hsat-not-sigma2}} %V.12
-We use a similar proof to the one provided by the authors in \cite{SalujaST95} to separate the classes $\E{2}$ and $\U{2}$.
-Suppose that the statement is false, this is, $\chsat \in \eqso(\loge{2})$.
-We consider the signature $\R$ that we used as the encoding for a Horn formula (Example \ref{ex-hornsat-esop1}) and that the formula $\alpha \in \eqso(\loge{2})$ follows the encoding in the same way.
-From what we proved in Theorem \ref{theo-pnf-snf}, we have that every formula in $\eqso(\loge{2})$ can be rewritten in $\loge{2}$-PNF, so we assume that $\alpha$ is in this form.
-Let $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\forall\bar{z}\,\varphi(\bar{X},\bar{x},\bar{y},\bar{z})$.
-Consider the following Horn formula $\Phi$:
-$$
-\Phi = p \wedge \bigwedge_{i = 1}^n (t_i \wedge p \to q) \wedge \neg q,
-$$
-where $n = \length{\bar{x}} + \length{\bar{y}} + 1$.
-Let $\A$ be the encoding of this formula.
-In the encoding, each variable appears as an element in the domain of $\A$.
-This formula is satisfiable, so $\sem{\alpha}(\A) \geq 1$.
-Let $(\bar{B},\bar{b},\bar{a})$ be an assignment to $(\bar{X},\bar{x},\bar{y})$ such that $\A\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-Let $t_{\ell}$ be such that it does not appear in $\bar{b}$ or $\bar{a}$.
-Consider the induced substructure $\A'$ that is obtained by removing $t_{\ell}$ from $\A$ and $\bar{B}'$ as the subset of $\bar{B}$ obtained by deleting each appearance of $t_{\ell}$ in $\bar{B}$.
-We have that $\A'\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-This is because each subformula of the form $\exists y \neg B_i$ is still true, and universal formulas are monotone over induced substructures.
-It follows that $\sem{\alpha}(\A') \geq 1$ which is not possible since $\A'$ encodes the formula
-$$
-\Phi' = p \wedge \bigwedge_{i = 1}^{\ell-1} (t_i \wedge p \to q) \wedge (p\to q) \wedge \bigwedge_{i = \ell+1}^{n} (t_i \wedge p \to q) \wedge \neg q,
-$$
-which is unsatisfiable. We arrive to a contradiction and we conclude that $\chsat$ is not in $\eqso(\ehorn)$.
-
-\subsection*{Proof of Theorem \ref{sigma2hard}} %V.13
-
-First we prove that $\shdhsat$ is in $\eqso(\ehorn)$.
-Recall that each instance of $\shdhsat$ is a disjunction of Horn formulas.
-Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot), \pD(\cdot,\cdot)\}$.
-Each symbol in this vocabulary is used to indicate the same as in Example \ref{ex-hornsat-esop1}, with the addition of $\pD(d,c)$ which indicates that $c$ is a clause in the formula $d$.
-Recall that the formula
-\begin{align*}
-&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \neg\textit{T}(x)).
-\end{align*}
-defines $\chsat$. We obtain the following formula $\psi(T,A)$ in $\ehorn$:
-\begin{align*}
-\exists d[&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \pD(c,d)\vee \neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \neg\textit{T}(x))]
-\end{align*}
-effectively defines $\chsat$ as for every disjunction of Horn formulas $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
-Therefore, we conclude that $\shdhsat \in \eqso(\ehorn)$.
-
-\vspace{1em}
-We will now prove that $\shdhsat$ is hard for $\eqso$ over a signature $\R$ under parsimonious reductions.
-For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a polynomial-time procedure that computes a function $g_{\alpha}$.
-This function receives a finite $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
-We suppose that $\alpha$ is in sum normal form and:
-$$
-\alpha = \sum_{i = 1}^{\text{\#clauses}} \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigwedge_{j = 1}^{n}\forall\bar{z}\,\varphi^i_j(\bar{X},\bar{x},\bar{y},\bar{z}),
-$$
-where each $\varphi^i_j$ is a Horn clause.
-
-Consider a finite $\R$-structure $\A$ with domain $A$.
-To simplify the proof, we extend our grammar to allow first-order constants.
-Consider each tuple $\bar{a}\in A^{\length{\bar{x}}}$, each $\bar{b}\in A^{\length{\bar{y}}}$ and each $\bar{c}\in A^{\length{\bar{z}}}$ as a tuple of first-order constants.
-The following formula defines the same function as $\alpha$:
-$$
-\sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length{\bar{x}}}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length{\bar{y}}}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length{\bar{z}}}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
-$$
-Note that each $\fo$ formula over $(\bar{x},\bar{y},\bar{z})$ in each $\varphi^i_j$ has no free variables.
-Therefore, we can evaluate each of these in polynomial time and replace them by $\perp$ and $\top$ where it corresponds.
-Each $\varphi^i_j$ will be of the form $\perp \vee\, \chi^i_j(\bar{X})$ or $\top \vee \chi^i_j(\bar{X})$ where $\chi^i_j$ is a disjunction of $\neg X_{\ell}$'s and at most one $X_{\ell}$.
-The formulas of the form $\top \vee \chi^i_j(\bar{X})$ can be removed entirely, and the formulas of the form $\perp \vee\, \chi^i_j(\bar{X})$ can be replaced by $\chi^i_j(\bar{X})$.
-We obtain the formula
-$$
-\sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})
-$$
-where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$.
-
-Our idea for the rest of the proof is to define $g_{\alpha}$ for each $\alpha = \sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})$, for $\alpha = c$ and for $\alpha = \beta_1 + \cdots + \beta_m$ where each $\beta_i$ is in one of the two previous cases.
-
-If $\alpha$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ where $\psi_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$, then we obtain the {\bf propositional formula} $g_{\alpha}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulas.
-
-If $\alpha$ is equal to a constant $c$, then we define $g_{\alpha}(\A)$ as the following formula that has exactly $c$ satisfying assignments:
-$$
-g_{\alpha}(\A) = \bigvee_{i = 1}^{c}\neg t_1 \wedge \cdots \wedge \neg t_{i-1} \wedge t_i \wedge \neg t_{i+1} \wedge \cdots \wedge \neg p_c.
-$$ 
-If $\alpha = \beta_1 + \cdots + \beta_m$, let $g_{\beta_i}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\theta^i_{j,k}$ for each $\beta_i$ where each $\theta^i_{j,k}$ is a Horn clause.
-Let $\Theta_i = g_{\beta_i}(\A)$.
-We rename the variables in each $\Theta_i$ so none of them are mentioned in any other $\Theta_j$.
-We add $m$ new variables $t_1,\ldots,t_m$ and we define:
-\begin{align*}
-g_{\alpha}(\A) = &\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^1_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m}}}t) \wedge (t_1 \wedge \bigwedge_{\ell = 2}^{m} \neg t_{\ell})) \vee \\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^2_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each $t$}\\ \text{ mentioned in}\\ \Theta_1,\Theta_3,\ldots,\Theta_{m}}}t) \wedge (t_2 \wedge \bigwedge\limits_{\substack{\ell = 1 \\ \ell \neq 2}}^{m} \neg t_{\ell})) \vee \cdots \vee\\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^m_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m-1}}}t) \wedge (t_m \wedge \bigwedge_{\ell = 1}^{m-1} \neg t_{\ell})).
-\end{align*}
-The formula is a disjunction of Horn formulas, and the number of satisfying assignments for this formula is exactly the sum of satisfying assignments for each $g_{\beta_i}(\A)$.
-This, at the same time, is equal to $\sem{\alpha}(\A)$.
-This covers all possible cases for $\alpha$, and the entire procedure takes polynomial time.

### 2017-03-03 LMCS/proof-tqfo-shl.tex
journal primera version

$\boldsymbol{\tqfo(\fo)}$ {\bf can be computed in} $\boldsymbol{\shl.}$
Let $\R$ be some relational signature. Let $\alpha$ be a formula in $\tqfo(\fo)$. We will construct a nondeterministic logspace algorithm $M_{\alpha}$ that on input $\enc(\A)$, where a first-order assignment $v$ is being stored in memory, accepts in $\sem{\alpha}(\A,v)$ paths. Suppose the domain of $\A$ is $A = \{1,\ldots,n\}$. The algorithm needs $c\mult\log_2(n)$ bits of memory to store $v$, where $c$ is the total number of first-order variables in $\alpha$. If $\alpha = \varphi$, we check if $(\A,v)\models\varphi$ in deterministic logarithmic space, and accept if and only if it does. If $\alpha = s$, we generate $s$ branches and accept in all of them. If $\alpha = (\alpha_1 + \alpha_2)$, we simulate $M_{\alpha_1}$ and $M_{\alpha_2}$ on separate branches. If $\alpha = (\alpha_1\mult\alpha_2)$, we simulate $\alpha_1$ and if it accepts, instead of doing so, we simulate $\alpha_2$. If $\alpha = \sa{x}\beta$, for each $a\in A$ we generate a different branch where we simulate $M_{\beta}$ while storing $v[a/x]$. If $\alpha = \pa{x}\beta$, we simulate $M_{\beta}$ while storing $v[1/n]$, and on each accepting branch, instead of accepting we replace the assignment on $x$ to 2, to simulate $M_{\beta}$ while storing $v[2/x]$, and so on. If $\alpha = [\pth \varphi(\bar{x},\bar{y})]$ where $\varphi$ is an $\fo$ formula, we simulate the $\shl$ procedure that counts the number of paths for a graph of a given size. This procedure starts by setting $\bar{a} = v(\bar{x})$. On each iteration, nondeterministically chooses an assignment $\bar{a}$ for $\bar{x}$, continues if $(\A,v)\models\varphi(\bar{a}',\bar{a})$ where $\bar{a}'$ is the previously chosen value for $\bar{a}$, and rejects otherwise. If at any point we obtain that $\bar{a} = v(\bar{y})$, we generate an accepting branch, and continue simulating the procedure in the current branch. We simulate $n^{\length{\bar{x}}}$ iterations of the procedure, and this generates exactly $\sem{[\pth \varphi(\bar{x},\bar{y})]}(\A,v)$ accepting branches. This ends the construction of the algorithm. Consider $f$ as the $\shl$ function associated to this procedure and we have that for each finite $\R$-structure $\A$: $f(\enc(\A)) = \sem{\alpha}(\A)$.

\vspace{1em}
$\boldsymbol{\shl}$ {\bf can be modelled in }$\boldsymbol{\tqfo(\fo).}$ Let $f$ be a function in $\shl$ and let $M$ be a nondeterministic logspace machine such that $\acc_{M}(\enc(\A)) = f(\enc(\A))$. We assume that $M$ has only one accepting state and upon accepting it immediately stops. Moreover, we assume that there exists only one accepting configuration altogether. We make use of transitive closure logic ($\tc$) to simplify our proof~\cite{G07}. We have that $\tc$ captures $\nlog$\cite{I83}, so there exists a formula such that $\A\models\varphi$ if and only if $M$ accepts $\enc(\A)$. This formula can be expressed as:
$$
\varphi = \exists\bar{u}\exists\bar{z}(\first(\bar{u})\wedge \psi_{\bf acc}(\bar{z})\wedge[{\bf tc}_{\bar{x},\bar{y}}\,\psi_{\bf next}(\bar{x},\bar{y})](\bar{u},\bar{z})),
$$
where $\psi_{\bf acc}(\bar{z})$ is an $\fo$ formula that expresses that $\bar{z}$ is an accepting configuration, and $\psi_{\bf next}(\bar{x},\bar{y})$ is an $\fo$ formula that expresses that $\bar{y}$ is the next configuration from $\bar{x}$~\cite{G07}. Here, there is a 1-1 correspondence between configurations of $M$ and assignments to $\bar{z}$. As a consequence, given a structure $\A$, and a first-order assignment $v$ to $\A$ where $v(\bar{x})$ is the starting configuration and $v(\bar{y})$ is the sole accepting configuration, the value of $\sem{[\pth\psi_{\bf next}(\bar{x},\bar{y})]}(\A,v)$ is equal to $\acc_M(\enc(\A))$.
Finally, we define the $\tqfo(\fo)$ formula
$$
\alpha = \sa{\bar{u}}\sa{\bar{z}}(\first(\bar{u})\mult\psi_{\bf acc}(\bar{z})\mult[\pth \psi_{\bf next}(\bar{u},\bar{z})]),
$$
which satisfies $\sem{\alpha}(\A) = f(\enc(\A))$ for each structure $\A$. This concludes the proof.

### 2017-03-03 LMCS/saluja.tex
journal primera version

--- 2017-01-09 LICS/saluja.tex
@@
 Moreover, a formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-sum normal form ($\LL$-SNF)} if $\alpha$ is of the form $\Sigma_{i=1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-PNF. 
-\begin{proposition}\label{theo-pnf-snf}
+\begin{prop}\label{theo-pnf-snf}
 Every formula in $\eqso(\LL)$ can be rewritten in $\LL$-SNF.
-\end{proposition}
+\end{prop}
+\proof
+\input{proof-theo-pnf-snf}
+\qed
 If a formula is in $\LL$-PNF then clearly the formula is in $\LL$-SNF.
@@
 Therefore, to unveil the relationship between the $\sfo$-hierarchy and the $\eqso(\fo)$-hierarchy, we need to understand the boundary between PNF and SNF. We do this in the following theorem. 
-\begin{theorem}\label{theo-pi1-pnf}
+\begin{thm}\label{theo-pi1-pnf}
 For $i = 0,1$, there exists a formula $\alpha_i$ in $\QE{i}$ that is not equivalent to any formula in $\Sigma_i$-PNF. 
-On the other hand, if $\logu{1} \subseteq \LL$ and $\LL$ is closed under conjunction and disjunction, then 
-	every formula in
-	$\eqso(\LL)$ can be rewritten in $\LL$-PNF. 
-\end{theorem}
+On the other hand, if $\logu{1} \subseteq \LL$ and $\LL$ is closed under conjunction and disjunction, then every formula in $\eqso(\LL)$ can be rewritten in $\LL$-PNF.
+
+From now on, for every first-order tuple $\bar{x}$ (second-order tuple $\bar{X}$) $\length{\bar{x}}$ as the size of $\bar{x}$ ($\length{\bar{X}}$ as the size of $\bar{X}$).
+\end{thm}
+\proof
+\input{proof-theo-pi1-pnf}
+\qed
 
@@
 \caption{The relationship between the $\sfo$-hierarchy and the $\eqso(\fo)$-hierarchy, where $\E{1}$ and $\QE{0}$ are incomparable. \label{fig-sfo-eqso}}
+\vspace{-0.1cm}
 \end{figure*}
@@
 The following proposition completes our picture of the relationship between the $\sfo$-hierarchy and the $\eqso(\fo)$-hierarchy.
-\begin{proposition}\label{prop-rest}
+\begin{prop}\label{prop-rest}
 The following properties hold:
@@
 \end{itemize}
-\end{proposition}
+\end{prop}
+\proof
+\input{proof-prop-rest}
+\qed
 The relationship between the two hierarchies is summarized in Figure \ref{fig-sfo-eqso}.

### 2017-03-03 LMCS/syntactic-clo.tex
journal primera version

--- 2017-01-10 LICS/syntactic-clo.tex
@@
 We use the \emph{$\eqso(\fo)$-hierarchy} to define syntactic classes of functions with good algorithmic and closure properties.  
-But before 
-%starting our search
-doing this, we introduce a more strict notion of counting problem with easy decision version.
+But before doing this, we introduce a more strict notion of counting problem with easy decision version.
 Recall that a function $f : \Sigma^* \to \mathbb{N}$ has an easy decision counterpart if $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$ is a language in $\ptime$.
-As the goal of this section is to define a syntactic class of functions in $\shp$ with easy decision versions and good closure properties, we do not directly consider the semantic condition $L_f  \in \ptime$, but instead we consider a more restricted
-%class of functions that has a simple 
-syntactic 
-%definition. 
-condition.
+As the goal of this section is to define a syntactic class of functions in $\shp$ with easy decision versions and good closure properties, we do not directly consider the semantic condition $L_f  \in \ptime$, but instead we consider a more restricted syntactic condition.
 More precisely, a function $f : \Sigma^* \to \mathbb{N}$ is said to be in the complexity class $\totp$~\cite{PagourtzisZ06} if there exists a  polynomial-time NTM $M$ such that $f(x) = \tmt_M(x) - 1$ for every $x \in \Sigma^*$, where $\tmt_M(x)$ is the total number of runs of $M$ with input $x$.
@@
 
-%The previous result shows that the classes $\QE{i}$ and $\QU{i}$ are more robust than the classes $\E{i}$ and $\U{i}$: they are closed under binary and sum quantifiers but the other not necessarily.
-
 It was proved in \cite{SalujaST95} that every function in $\E{1}$ admits an FPRAS. Besides, it can be proved that $\E{1} \subseteq \totp$. 
-However, 
-%we have to discard 
-this class 
-%since it i
-is not closed under sum, and then it is not robust under basic closure properties. 
-\begin{proposition}\label{prop-e1-nc}
+However, this class is not closed under sum, and then it is not robust under basic closure properties. 
+\begin{prop}\label{prop-e1-nc}
 There exist functions $f, g \in \E{1}$ such that $(f + g) \not\in \E{1}$.
-\end{proposition}
+\end{prop}
+\proof
+\input{proof-prop-e1-nc}
+\qed
 To overcome this limitation, one can consider the class $\QE{1}$, which is closed under sum by definition.
-In fact, the following proposition shows that the same good properties as for $\E{1}$ hold for $\QE{1}$, together with the fact that
-%plus being 
-it is closed under sum and multiplication.
-\begin{proposition} \label{prop:qe0-fp-qe1-totp-fptras}
+In fact, the following proposition shows that the same good properties as for $\E{1}$ hold for $\QE{1}$, together with the fact that it is closed under sum and multiplication.
+\begin{prop} \label{prop:qe0-fp-qe1-totp-fptras}
 $\QE{1} \subseteq \totp$ and every function in $\QE{1}$ has an FPRAS. Moreover, $\QE{1}$ is closed under sum and multiplication.
-\end{proposition}
-%Moreover, from the following theorem one can conclude that $\QE{1}$ is closed under multiplication.
-%\begin{theorem}\label{theo:binary-prod}
-%	If $\LL$ is closed under conjunction, then $\eqso(\LL)$ is closed under multiplication.
-%\end{theorem}
+\end{prop}
+\proof
+\input{proof-qe0-fp-qe1-totp-fptras}
+\qed
 Hence, it only remains to prove that $\QE{1}$ is closed under subtraction by one.
@@
 It is believed that $\np \not\subseteq \cspp$.
-%, as there is no evidence that a TM with the property just described exists for the propositional satisfiability problem. 
 However, if $\shp$ is closed under subtraction by one, then it holds that $\np \subseteq \cspp$ \cite{OH93}.} So, the following proposition rules out any logic that extends $\Pi_1$ for a possible extension of~$\QE{1}$ with the desired closure property.
-\begin{proposition} \label{pi-minusone}
+\begin{prop} \label{pi-minusone}
 If $\Pi_1 \subseteq \LL \subseteq \fo$ and $\eqso(\LL)$ is closed under subtraction by one, then $\shp$ is closed under subtraction by~one. 
-\end{proposition}
+\end{prop}
+\proof
+\input{proof-pi-minusone}
+\qed
 Therefore, the desired extension has to be achieved by allowing some local extensions to $\Sigma_1$.
@@
 With this extension we obtain a class with the desired properties.
-%find the promised class.
-\begin{theorem}\label{sigmafo-minusone}
+\begin{thm}\label{sigmafo-minusone}
 The class $\eqso(\logex{1})$ is closed under sum, multiplication and subtraction by one. Moreover, $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
-\end{theorem}
+\end{thm}
+\proof
+\input{proof-sigmafo-minusone}
+\qed
 The proof that $\eqso(\logex{1})$ is closed under subtraction by one is the most involved of the paper.
 We think the main technique used in this proof, which is based on considering some witnesses of logarithmic size, is of independent interest.
-
-%A natural question at this point is whether the property of being closed under subtraction by one could be generalized to proper subtraction.
-More precisely, a function complexity class $\CC$ is said to be closed under subtraction if for every $f,g \in \CC$, it holds that $(f \dotdiv g) \in \CC$, where $(f \dotdiv g)(x)$ is defined as $f(x) - g(x)$ if $f(x) \geq g(x)$, and as $0$ otherwise.
-The following theorem shows that such generalization does not work for any of the function complexity classes considered in this section.
-%\begin{theorem} \label{sub-pnp}
-%If any of the classes $\E{1}$, $\QE{1}$ or $\eqso(\logex{1})$ is closed under subtraction, then $\ptime = \np$.
-%\end{theorem}
-
-%
-%The next question is whether the hierarchy is closed under subtraction.
-Formally, for any pair of functions $f,g$, we define $f - g$ as the function such that $(f - g)(\A) = f(\A)-g(\A)$ whenever $f(\A)>g(\A)$ and $0$ otherwise.
-%As the next result shows, all classes in the $\eqso$-hierarchy is not closed under subtraction unless ${\sc P} = {\sc NP}$
-%\begin{theorem} \label{sub-pnp}
-%If $\eqso(\loge{i})$ or $\eqso(\logu{i})$ is closed under subtraction for $i > 0$, then {\sc P} = {\sc NP}.
-%\end{theorem}
-%\cristian{Martin, el resultado que tienes en el apendice se generaliza trivialmente para todas las clases ya que todas contienen la clase $\eqso(\loge{0})$.}
-
-%Now, we study the complexity classes describe by this hierarchies.
-As the following result shows, $\eqso(\loge{0})$ defines only tractable counting functions and $\eqso(\loge{1})$ intractable counting functions but with an tractable decision problems.
-%\begin{proposition} \label{prop:qe0-fp-qe1-totp-fptras}
-%All functions defined in $\eqso(\loge{0})$ and $\eqso(\loge{1})$ can be computed in $\fp$ and $\totp$, respectively. Furthermore, every function defined in $\eqso(\loge{1})$ has a FPTRAS.
-%\end{proposition}
-%Therefore, in terms of counting complexity, the $\eqso$-hierarchy behaves exactly the same as the $\#\fo$-hierarchy.
-
-%The next step is to study the closure properties of $\eqso$-hierarchy. 
-%An advantage of the $\eqso$-hierarchy is that, by its language syntax, all the classes are closed under addition and first and second order sum.
-%So, the first question is whether the multiplicative operators in $\qso$ can be defined in $\eqso(\LL)$.
-As the following result shows, if $\LL$ is closed under conjunction, then the binary product can be defined in  $\eqso(\LL)$.
-%\begin{theorem}\label{theo:binary-prod}
-%	If $\LL$ is closed under conjunction, then binary product can be defined in $\eqso(\LL)$.
-%\end{theorem}
-
-%By the previous result, we know that functions in the $\eqso$ hierarchy are unlikely to be closed under subtraction.
-Then, a natural restriction to this question is to ask whether these classes are closed under subtraction by one, namely, if $\CC$ is a class of functions and $f \in \CC$, is $f-1 \in \CC$ where $1$ is the constant function that outputs $1$ for every structure.
-%We do not know $\E{1}$ is closed under subtraction by one. However, if we extend $\logex{1}$ with $\fo$ predicates we can show that this new fragment is closed under subtraction by one.
-%\begin{theorem} \label{sigmafo-minusone}
-%	$\eqso(\logex{1})$ is closed under substraction by one.
-%\end{theorem}

### 2017-03-03 LMCS/syntactic-horn.tex
journal primera version

--- 2017-01-10 LICS/syntactic-horn.tex
@@
 The following example shows how our approach works.
-\begin{example} \label{ex-hornsat-esop1}
+\begin{exa} \label{ex-hornsat-esop1}
 Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot),<\}$.
@@
 More precisely, we have that:
-\begin{eqnarray*}
-\sem{\sa{\pT} \varphi(\pT)}(\A) &=& \sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A),
-\end{eqnarray*}
+\begin{align*}
+\sem{\sa{\pT} \varphi(\pT)}(\A) &= \sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A),
+\end{align*}
  for every $\R$-structure $\A$ encoding a Horn formula.
@@
 More importantly, $\psi(\pT,\pA)$ resembles a conjunction of Horn clauses except for the use of negative literals of the form $\exists v \, \neg \textit{A}(u,v)$.
-%This formula effectively defines $\chsat$
-%as for every Horn formula $\theta$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
-Therefore, we conclude that $\chsat \in \eqso(\uhorn)$.
-\end{example}
+\end{exa}
 The previous example suggests that to define $\chsat$, we can use Horn formulae defined as follows. 
@@
 A clause is said to be Horn if it contains at most one positive literal, and a formula is said to be Horn if it is a conjunction of Horn clauses.
-With this terminology, we define $\uhorn$ as the set of formulae $\psi$ such that $\psi$ is a
-%conjunction of 
-Horn formula 
-%clauses 
-over a signature $\R$.
+With this terminology, we define $\uhorn$ as the set of formulae $\psi$ such that $\psi$ is a Horn formula over a signature $\R$.
 
@@
 Moreover, an even larger class for our search can be generated by extending the definition of $\uhorn$ with outermost existential quantification.
-%More precisely, 
 Formally, a formula $\varphi$ is in $\ehorn$ if $\varphi$ is of the form $\exists \bar x \, \psi$ with $\psi$ a Horn formula.
 
-\begin{proposition}\label{prop:ehorn-pe}
+\begin{prop}\label{prop:ehorn-pe}
 $\eqso(\ehorn) \subseteq \totp$.
-\end{proposition}
+\end{prop}
+\proof
+\input{proof-ehorn-pe}
+\qed
 Interestingly, we have that both $\chsat$ and $\cdnf$ belong to $\eqso(\ehorn)$. 
@@
 
-\begin{proposition}\label{prop:hsat-not-sigma2}	
+\begin{prop}\label{prop:hsat-not-sigma2}	
 $\chsat \not\in \eqso(\loge{2})$.
-\end{proposition}
+\end{prop}
+\proof
+\input{proof-hsat-not-sigma2}
+\qed
 We conclude this section by showing that $\eqso(\ehorn)$ is the class we were looking for, as not only every function in $\eqso(\ehorn)$ has an easy decision counterpart, but also $\eqso(\ehorn)$ admits a natural complete problem under parsimonious reductions.
 More precisely, define
-%$\dhsat$ as:
-%\begin{multline*}
-%\{\Phi \mid \Phi \text{ is a disjunction of Horn formulae and satisfiable}\},
-%\end{multline*}
-%%\begin{multline*}
-%%\{\Phi \mid \Phi \text{ is a disjunction of}\\  \text{Horn formulae and $\Phi$ is satisfiable}\},
-%%\end{multline*}
-%and define 
 $\shdhsat$ as the problem of counting the satisfying assignments of a formula $\Phi$ that is a disjunction of Horn formulae. Then we have that:
 
-\begin{theorem} \label{sigma2hard}
+\begin{thm} \label{sigma2hard}
 	$\shdhsat$ is $\eqso(\ehorn)$-complete under parsimonious reductions. 
-\end{theorem}
+\end{thm}
+\proof
+\input{proof-sigma2hard}
+\qed

### 2017-03-03 LMCS/syntactic.tex
journal primera version

--- 2017-01-10 LICS/syntactic.tex
@@
 
-The class $\shp$ was introduced in \cite{Valiant79} to prove that computing the permanent of a matrix, as defined in Example \ref{exa-perm}, is a $\shp$-complete problem.
-As a consequence of this result many counting problems have been proved to be $\shp$-complete~\cite{V79b,arora2009computational}.
-%\cite{V79b,PB83,P86,L86,BW91,HMRS98,DG00,BW05,DS12, PS13,PS14}.
-Among them, problems having easy decision counterparts play a fundamental role, as a counting problem with a hard decision version is expected to be hard.
-Formally, the decision problem associated to a function $f : \Sigma^* \to \mathbb{N}$ is defined as $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$, and $f$ is said to have an easy decision version if $L_f \in \ptime$.
-Many prominent examples satisfy this property, like computing the number of: perfect matchings of a bipartite graph ($\cpm$)~\cite{Valiant79}, satisfying assignments of a DNF ($\cdnf$)~\cite{DHK05} or Horn ($\chsat$)~\cite{V79b} propositional formula, among others.
-
-Counting problems with easy decision versions play a fundamental role in the search of efficient approximations algorithms for functions in $\shp$.
-A fully-polynomial randomized approximation scheme (FPRAS) for a function $f : \Sigma^* \to \bbN$ is a randomized algorithm ${\cal A} : \Sigma^* \times (0,1) \to \bbN$ such that: (1) for every string $x \in \Sigma^*$ and real value $\varepsilon \in (0,1)$, the probability that $|f(x) - {\cal A}(x,\varepsilon)| \leq \varepsilon \cdot f(x)$ is at least $\frac{3}{4}$, and (2) the running time of ${\cal A}$ is polynomial in the size of $x$ and $1/\varepsilon$ \cite{KL83}.
-Notably, there exist $\shp$-complete functions that can be efficiently approximated as they admit FPRAS; for instance, there exist FPRAS for $\cdnf$~\cite{KL83} and $\cpm$~\cite{JSV04}.
-A key observation here is that if a function $f$ admits an FPRAS, then $L_f$ is in the randomized complexity class $\bpp$~\cite{G77}.
-Hence, under the widely believed assumption that  $\np \not\subseteq \bpp$, we cannot hope for an FPRAS for a function in $\shp$ whose decision counterpart is $\np$-complete, and we have to concentrate on the class of counting problems with easy decision versions.
-
-The importance of the class of counting problems with easy decision counterparts has motivated the search of robust classes of functions in $\shp$ with this property \cite{PagourtzisZ06}.
-But the key question here is what should be considered a {\em robust} class.
-A first desirable condition has to do with the closure properties satisfied by the class, which is a common theme when studying function complexity classes \cite{OH93,FH08}.
-As in the cases of $\ptime$ and $\np$ that are closed under intersection and union, we expect our class to be closed under multiplication and sum.
-For a more elaborated closure property, assume that $\textit{sat\_one}$ is a function that returns one plus the number of satisfying assignments of a propositional formula.
-Clearly $\textit{sat\_one}$ is a $\shp$-complete function whose decision counterpart $L_{\textit{sat\_one}}$ is trivial.
-But should $\textit{sat\_one}$ be part of a robust class of counting function with easy decision versions?
-The key insight here is that if a function in $\shp$ has an easy decision counterpart $L$, then as $L \in \np$ we expect  to have a polynomial-time algorithm that verifies whether $x \in L$ by constructing witnesses for~$x$.
-%As an example of this considered all the counting functions mentioned in the previous paragraphs and, in particular, the decision counterpart of $\cdnf$ that can be solved by explicitly constructing in polynomial-time a satisfying assignment for an input propositional formula.
-Moreover, if such an algorithm for constructing witnesses exists, then we also expect to be able to manipulate such witnesses and in some cases to remove them.
-In other words, we expect a robust class $\CC$ of counting functions with easy decision versions to be closed under subtraction by one, that is, if $g \in \CC$, then the function $g \dotdiv 1$ should also be in $\CC$, where $(g \dotdiv 1)(x)$ is defined as $g(x) - 1$ if $g(x) \geq 1$, and as~$0$ otherwise.
-Notice that, unless $\ptime = \np$, no such class can contain the function $\textit{sat\_one}$ because $\textit{sat\_one} \dotdiv 1$ counts the number of satisfying assignments of a propositional formula.
-
-A second desirable condition of robustness is the existence of natural complete problems \cite{P94}.
-Special attention has to be paid here to the notion of reduction used for completeness.
-Notice that under the notion of Cook reduction, originally used in \cite{Valiant79}, the problems $\cdnf$ and $\csat$ are $\shp$-complete.
-However, $\cdnf$ has an easy decision counterpart and admits an FPRAS, while $\csat$ does not satisfy these conditions unless $\ptime = \np$.
-Hence a more strict notion of reduction has to be considered; in particular, the notion of parsimonious reduction (to be defined later) satisfy that if a function $f$ is parsimoniously reducible to a function $g$, then $L_g \in \ptime$ implies that $L_f \in \ptime$ and the existence of an FPRAS for $g$ implies the existence of a FPRAS for $f$.
-
-In this section, we use the framework developed in this paper to address the problem of defining a robust class of functions with easy decision versions.
-More specifically, we use the framework to introduce in Section \ref{sec-hier-shp} a syntactic hierarchy of counting complexity classes contained in $\shp$.
-Then this hierarchy is used in Section \ref{sec-clo} to define a class of functions with easy decision versions and good closure properties, and in Section \ref{sec-horn} to define a class of functions with easy decision versions and
-natural complete problems.
-
-%The class $\shp$ was introduced in \cite{Valiant79} to prove that computing the permanent of a matrix, as defined in Example \ref{exa-perm}, is a difficult problem.
-More specifically, it was shown in  \cite{Valiant79}  that this problem is $\shp$-complete.
-As a consequence of this result the $\shp$-completeness of the problem of computing the number of perfect matchings of a bipartite graph ($\cpm$) was established.
-%Since then  many counting problems have been proved to be $\shp$-complete \cite{V79b,PB83,P86,L86,BW91,HMRS98,DG00,BW05,DS12, PS13,PS14}.
-Among them, problems having easy decision counterparts play a fundamental role, as a counting problem with a hard decision version is expected to be hard.
-Formally, the decision problem associated to a function $f : \Sigma^* \to \mathbb{N}$ is defined as $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$, and $f$ is said to have an easy decision version if $L_f \in \ptime$.
-A first prominent example of such functions is $\cpm$, as it is well-known that the problem of verifying whether there exists a perfect matching in a bipartite graph can be solved in polynomial time.
-Other prominent examples of such problems include counting the number of: satisfying assignments of a 2-CNF propositional formula ($\ctwcnf$) \cite{V79b}, satisfying assignments of a DNF propositional formula ($\cdnf$) \cite{DHK05}, satisfying assignments of a Horn  propositional formula ($\chsat$) \cite{V79b}, simple paths from a source node to a target node in a directed graph ($\csp$) \cite{V79b}, extensions of a partial order to a linear order \cite{BW91} and Eulerian cycles in an undirected graph \cite{BW05}.
-%
-%Counting problems with easy decision versions play a fundamental role in the search of efficient approximations algorithms for functions in $\shp$.
-A fully-polynomial randomized approximation scheme (FPRAS) for a function $f : \Sigma^* \to \bbN$ is a randomized algorithm ${\cal A} : \Sigma^* \times (0,1) \to \bbN$ such that: (1) for every string $x \in \Sigma^*$ and real value $\varepsilon \in (0,1)$, the probability that $|f(x) - {\cal A}(x,\varepsilon)| \leq \varepsilon \cdot f(x)$ is at least $\frac{3}{4}$, and (2) the running time of ${\cal A}$ is polynomial in the size of $x$ and $1/\varepsilon$ \cite{KL83}.
-Notably, there exist $\shp$-complete functions that can be efficiently approximated as they admit FPRAS; for instance, there exist FPRAS for the problems of counting the number of satisfying assignments of a DNF propositional formula \cite{KL83} and the number of perfect matchings of a bipartite graph \cite{JSV04}.
-A key observation here is that if a $\shp$-complete function admits an FPRAS, then its associated decision problem is in the randomized complexity class $\bpp$ (Bounded-Error Probabilistic Polynomial-Time \cite{G77}).
-Hence, under the widely believed assumption that  $\np \not\subseteq \bpp$, we cannot hope for an FPRAS for a function in $\shp$ whose decision counterpart is $\np$-complete, and we have to concentrate on the class of counting problems with easy decision versions (in $\bpp$ or in a lower complexity class such as $\ptime$).
-%
-%The importance of the class counting problems with easy decision counterparts has motivated the search of robust classes of functions in $\shp$ with this property \cite{PagourtzisZ06}.
-But the key question here is what should be considered a {\em robust} class.
-A first desirable condition is that the class has a syntactic definition that allows for the existence of natural complete problems \cite{P94}.
-Special attention has to be paid here to the notion of reduction used for completeness.
-Notice that under the notion of Cook reduction, originally used in \cite{Valiant79}, the problems $\cdnf$ and $\csat$ are $\shp$-complete.
-However, $\cdnf$ has an easy decision counterpart and admits an FPRAS, while $\csat$ does not satisfy these conditions unless $\ptime = \np$.
-Hence a more strict notion of reduction has to be considered; in particular, the notion of parsimonious reduction (to be defined later) satisfy that if a function $f$ is parsimoniously reducible to a function $g$, then $L_g \in \ptime$ implies that $L_f \in \ptime$ and the existence of an FPRAS for $g$ implies the existence of a FPRAS for $f$.
-%A second desirable condition has to do with the closure properties satisfied by the class, which is a common theme when studying function complexity classes \cite{OH93,FH08}.
-As in the cases of $\ptime$ and $\np$ that are closed under intersection and union, we expect our class to be closed under multiplication and sum; in fact, if $f$ and $g$ are counting functions with easy decision versions, then we clearly expect $f + g$ and $f \cdot g$ to have easy decision counterparts.
-For a more elaborated closure property, assume that $\textit{sat\_one}$ is a function that returns one plus the number of satisfying assignments of a propositional formula.
-Clearly $\textit{sat\_one}$ is a $\shp$-complete function whose decision counterpart $L_{\textit{sat\_one}} = \{ \varphi \mid \textit{sat\_one}(\varphi) > 0 \}$ is trivial.
-But should $\textit{sat\_one}$ be part of a robust class of counting function with easy decision versions?
-The key insight here is that if a function in $\shp$ has an easy decision counterpart $L$, then as $L \in \np$ we expect  to have a polynomial-time algorithm that verifies whether $x \in L$ by constructing witnesses for $x$.
-As an example of this considered all the counting functions mentioned in the previous paragraphs and, in particular, the decision counterpart of $\cdnf$ that can be solved by explicitly constructing in polynomial-time a satisfying assignment for an input propositional formula.
-Moreover, if such an algorithm for constructing witnesses exists, then we also expect to be able to manipulate such witnesses and in some cases to remove them.
-In other words, we expect a robust class $\CC$ of counting functions with easy decision versions to be closed under subtraction by one, that is, if $g \in \CC$, then the function $g \dotdiv 1$ should also be in $\CC$, where $(g \dotdiv 1)(x)$ is defined as $g(x) - 1$ if $g(x) \geq 1$, and as $0$ otherwise.
-Notice that, unless $\ptime = \np$, no such class can contain the function $\textit{sat\_one}$ because $\textit{sat\_one} \dotdiv 1$ counts the number of satisfying assignments of a propositional formula.
-%
-%In this section, we use the framework developed in this paper to address the problem of defining a robust class of functions with easy decision versions.
-More specifically, we use the framework to introduce in Section \ref{sec-hier-shp} a syntactic hierarchy of function complexity classes contained in $\shp$.
-Then this hierarchy is used in Section \ref{sec-clo} to define a class of functions with easy decision versions and good closure properties, and in Section \ref{sec-horn} to define a class of functions with easy decision versions and
-%natural complete problems.
-
-% It should be noted that such class can be directly defined as the set functions f ?
-#P such that Lf ?
-P, which is denoted as #Pe in [50, 51].
-However, such a definition does not lead to a well-behaved and robust function complexity class.
-In particular, for every function f ?
-#P, we have that f + 1 is trivially in #Pe, which is an undesirable property.
-This has led to the introduction of the more robust class TotP, which is defined as the class of functions f for which there exists a non-deterministic Turing machine M running in polynomial time such that, f(x) is the result of subtracting 1 to the number of (non-necessarily accepting) runs of M with input x [40].
-In [51], it is proved that TotP ?
-#Pe and that TotP has a complete function problem under parsimonious reductions.
-However, no natural problem is known to be TotP-complete under this type of reductions [51].
-
-%
-%
-%In this section we study the fragment of $\eqso(\LL)$ when $\LL$ is a boolean logic contained in $\fo$.
-We show that by restricting $\LL$ we can find different subclasses below $\shp$ with interesting computational and closure properties.
-%
-%
-%
-%From this point on, for each fragment $\FF$ of $\qso$, we will also use $\FF$ to refer to the class of functions defined by the formulae in $\FF$.
-
-%\subsection{A syntactic hierarchy of function complexity classes contained in $\shp$}
 \subsection{The $\eqso(\fo)$-hierarchy inside $\shp$}

### 2017-03-07 LMCS/framework.tex
journam 7 mar

@@
 \]
-Intuitively, the subformula $\beta(S) := \pa{x} (\ex{y} S(x,y) \wedge M(x,y))$ calculates the value $\prod_{i=1}^n A[i, \sigma(i)]$ whenever $S$ encodes a permutation $\sigma$.
+Intuitively, the subformula $\beta(S) := \pa{x} (\ex{y} S(x,y) \wedge M(x,y))$ calculates the value \linebreak $\prod_{i=1}^n A[i, \sigma(i)]$ whenever $S$ encodes a permutation $\sigma$.
 Moreover, the subformula $\op{permut}(S) \cdot \beta(S)$ returns $\beta(S)$ when $S$ is a permutation, and returns $0$ otherwise (i.e. $\op{permut}(S)$ behaves like a filter). 

### 2017-03-07 LMCS/proof-qe0-fp-qe1-totp-fptras.tex
journam 7 mar

@@
 
-\begin{lemma} \label{conj-mult}
+\begin{lem} \label{conj-mult}
 	If $\LL$ is a fragment closed under conjunction, then $\eqso(\LL)$ is closed under binary multiplication.
-\end{lemma}
-\begin{proof}
+\end{lem}
+\proof
 	We will define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ extended by binary product, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
@@
 For every pair of formulas $\alpha,\beta$ in $\eqso(\LL)$, we have that their multiplication $(\alpha\cdot\beta)$ is a formula in the grammar $\eqso(\LL)$ extended by binary product, and so, there exists an equivalent formula $\tau(\alpha\cdot\beta)$ which is in unextended $\eqso(\LL)$.
-\end{proof}
+\qed
 Since $\loge{1}$ is closed under conjunction, this also holds for $\eqso(\loge{1})$. This concludes the proof.

### 2017-03-07 LMCS/proof-sigmafo-minusone.tex
journam 7 mar

@@
+%!TEX root = main.tex
+
 \noindent {\bf Closed under sum and multiplication.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
@@
 where $\varphi^{+}$ and $\varphi^{-}_i$ do not depend on  $\bar{x}$.
-\begin{claim}\label{claim:minusone}
+\begin{clm}\label{claim:minusone}
 	For a given ordered structure $\A$ and a first-order assignment $v$ for $\A$, $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ if, and only if, there exists a second-order assignment $V$ for $\A$ such that $(\A,V,v)\models \varphi_i'(\bar{X},\bar{x}, \bar{y})$.
-\end{claim}
-\begin{proof}
+\end{clm}
+\proof
 	Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$, such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
@@
 	For the other direction, if $(\A,V,v)\models \varphi_i'(\bar{X},\bar{x}, \bar{y})$ for a second order assignment $V$ for $\A$, then it is easy to check $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ since $\varphi^{\fo}_i(\bar{x},\bar{y})$ is a subformula of $\varphi_i'$.
-\end{proof}
+\qed
 
@@
 For this, the idea is to choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
-$$
+\[
 \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge\forall\bar{x}'\forall\bar{y}'(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')).
-$$ 
+\] 
 If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
@@
 \begin{align}
-\psi_i(\bar{X},\bar{x}) = & \;\; \big( \, \exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
-& \hspace{1cm} \Big( \; 
+\psi_i(\bar{X},\bar{x}) = &\big( \, \exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
+&\Big( \; 
 \exists\bar{y}~\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i'(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\
-& \hspace{1.5cm} (\; \exists\bar{x}'\exists\bar{y}\; \varphi_i'(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x} \; ) \;\; \Big)  \label{eq:part3}
+&(\; \exists\bar{x}'\exists\bar{y}\; \varphi_i'(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x} \; ) \;\; \Big)  \label{eq:part3}
 \end{align}
@@
 Define the first formula $\varphi_1^*$ as:
-$$
+\[
 \varphi^*_1(\bar{X},\bar{x}) = \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
-$$
+\]
 This formula accepts all the assignments that satisfy $\varphi_1$, except for the assignment $(V^*,v^*)$ of $\varphi_1$. The second formula $\varphi_2^*$ is defined as:
-$$
+\[
 \varphi^*_2(\bar{X},\bar{x}) = \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
-$$
+\]
 This models all the assignments that satisfy $\varphi_2$, except for the assignment $(V^*,v^*)$ of $\varphi_1$.
@@
 One can easily generalize this construction for each $\varphi_i$ as follows:
-$$
-\varphi_i^*(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
+\begin{multline*}
+\varphi_i^*(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\,\wedge \\ (\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
 \bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})),
-$$
+\end{multline*}
 From the construction of $\kappa(\alpha)$, one can easily check that  $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A)-1$ for each $\A$.
@@
 Now, for each $\varphi_i$ we define:
-$$
-\varphi_i^*(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
+\begin{multline*}
+\varphi_i^*(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\,\wedge \\
+(\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
 \bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x}))\Big)\Big).
-$$
+\end{multline*}
 Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i^*(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.

### 2017-03-07 LMCS/saluja.tex
journam 7 mar

@@
 \begin{figure*}
+%\begin{center}
+%\begin{tikzpicture}
+%\node[rectw] (n1) {$\E{0}$};
+%\node[rectw, right=0.5cm of n1] (n2) {};
+%\node[rectw, above=0.3cm of n2] (n3) {$\E{1}$}
+%	edge[draw=white] node {\rotatebox{45}{$\subsetneq$}} (n1);
+%\node[rectw, below=0.3cm of n2] (n4) {$\QE{0}$}
+%        edge[draw=white] node {\rotatebox{315}{$\subsetneq$}} (n1);
+%\node[rectw, right=0.3cm of n2] (n5) {$\QE{1}$}
+%        edge[draw=white] node {\rotatebox{315}{$\subsetneq$}} (n3)
+%         edge[draw=white] node {\rotatebox{45}{$\subsetneq$}} (n4);
+%\node[rectw, right=0.3cm of n5] (n6) {$\U{1}$}       
+%        edge[draw=white] node {$\subsetneq$} (n5);
+%\node[rectw, right=0.3cm of n6] (n7) {$\QU{1}$}       
+%        edge[draw=white] node {$=$} (n6);
+%\node[rectw, right=0.3cm of n7] (n8) {$\E{2}$}       
+%        edge[draw=white] node {$\subsetneq$} (n7);
+%\node[rectw, right=0.3cm of n8] (n9) {$\QE{2}$}       
+%        edge[draw=white] node {$=$} (n8);        
+%\node[rectw, right=0.3cm of n9] (n10) {$\U{2}$}       
+%        edge[draw=white] node {$\subsetneq$} (n9);
+%\node[rectw, right=0.3cm of n10] (n11) {$\QU{2}$}       
+%        edge[draw=white] node {$=$} (n10); 
+%\node[rectw, right=0.3cm of n11] (n12) {$\sfo$}       
+%        edge[draw=white] node {$=$} (n11); 
+%\end{tikzpicture}
+%\end{center}
 \begin{center}
@@
 	\node[rectw, right=0.5cm of n1] (n2) {};
-\node[rectw, above=0.4cm of n2] (n3) {$\E{1}$}
+	\node[rectw, above=0.3cm of n2] (n3) {$\E{1}$}
 		edge[draw=white] node {\rotatebox{45}{$\subsetneq$}} (n1);
-\node[rectw, below=0.4cm of n2] (n4) {$\QE{0}$}
+	\node[rectw, below=0.5cm of n2] (n4) {$\QE{0}$}
 		edge[draw=white] node {\rotatebox{315}{$\subsetneq$}} (n1);
@@
 		edge[draw=white] node {\rotatebox{45}{$\subsetneq$}} (n4);
-\node[rectw, right=0.5cm of n5] (n6) {$\U{1}$}       
+	\node[rectw, right=0.8cm of n5] (n6) {$\U{1}$}       
 		edge[draw=white] node {$\subsetneq$} (n5);
-\node[rectw, right=0.5cm of n6] (n7) {$\QU{1}$}       
-        edge[draw=white] node {$=$} (n6);
-\node[rectw, right=0.5cm of n7] (n8) {$\E{2}$}       
-        edge[draw=white] node {$\subsetneq$} (n7);
-\node[rectw, right=0.5cm of n8] (n9) {$\QE{2}$}       
-        edge[draw=white] node {$=$} (n8);        
-\node[rectw, right=0.5cm of n9] (n10) {$\U{2}$}       
-        edge[draw=white] node {$\subsetneq$} (n9);
-\node[rectw, right=0.5cm of n10] (n11) {$\QU{2}$}       
+	\node[rectw, below=0.3cm of n6] (n7) {$\QU{1}$}       
+		edge[draw=white] node {\rotatebox{90}{$=$}} (n6);
+	\node[rectw, right=1.0cm of n6] (n8) {$\E{2}$}       
+		edge[draw=white] node {$\subsetneq$} (n6);
+	\node[rectw, below=0.3cm of n8] (n9) {$\QE{2}$}       
+		edge[draw=white] node {\rotatebox{90}{$=$}} (n8);        
+	\node[rectw, right=1.0cm of n8] (n10) {$\U{2}$}       
+		edge[draw=white] node {$\subsetneq$} (n8);
+	\node[rectw, below=0.3cm of n10] (n11) {$\QU{2}$}       
+		edge[draw=white] node {\rotatebox{90}{$=$}} (n10); 
+	\node[rectw, right=0.5cm of n10] (n12) {$\sfo$}       
 		edge[draw=white] node {$=$} (n10); 
-\node[rectw, right=0.5cm of n11] (n12) {$\sfo$}       
-        edge[draw=white] node {$=$} (n11); 
 	\end{tikzpicture}

### 2017-03-08 LMCS/saluja.tex
journal 8 mar

@@
 \end{center}
+\vspace{1cm}
+\begin{center}
+	\begin{tikzpicture}
+	\node[rectw] (n1) {$\E{0}$};
+	\node[rectw, right=1.2cm of n1] (n2) {};
+	\node[rectw, above=0.2cm of n2] (n3) {$\E{1}$}
+	edge[draw=white] node {\rotatebox{45}{$\subsetneq$}} (n1);
+	\node[rectw, below=0.1cm of n2] (n4) {$\QE{0}$}
+	edge[draw=white] node {\rotatebox{315}{$\subsetneq$}} (n1);
+	\node[rectw, right=1.2cm of n2] (n5) {$\QE{1}$}
+	edge[draw=white] node {\rotatebox{315}{$\subsetneq$}} (n3)
+	edge[draw=white] node {\rotatebox{45}{$\subsetneq$}} (n4);
+	\node[rectw, right=1.3cm of n5] (n6) {\rotatebox{90}{$=$}}       
+	edge[draw=white] node {$\subsetneq$} (n5);
+	\node[rectw, above=0cm of n6] (n13) {$\U{1}$};
+	\node[rectw, below=0cm of n6] (n7) {$\QU{1}$};      
+	\node[rectw, right=1.8cm of n6] (n8) {\rotatebox{90}{$=$}}       
+	edge[draw=white] node {$\subsetneq$} (n6);
+	\node[rectw, above=0cm of n8] (n14) {$\E{2}$};
+	\node[rectw, below=0cm of n8] (n9) {$\QE{2}$};       
+	\node[rectw, right=1.8cm of n8] (n10) {\rotatebox{90}{$=$}}       
+	edge[draw=white] node {$\subsetneq$} (n8);
+	\node[rectw, above=0cm of n10] (n15) {$\U{2}$};
+	\node[rectw, below=0cm of n10] (n11) {$\QU{2}$};      
+	edge[draw=white] node {\rotatebox{90}{$=$}} (n10); 
+	\node[rectw, right=0.5cm of n10] (n12) {$\sfo$}       
+	edge[draw=white] node {$=$} (n10); 
+	\end{tikzpicture}
+\end{center}
 \vspace{-0.4cm}

### 2017-03-23 LMCS/counting.tex
journal 23 mar

@@
 We can now formalize the notion of capturing a counting complexity class.
-From now on, we assume that every counting complexity class $\CC$ consists of functions $f : \{0,1\}^* \to \N$ that are invariant under isomorphism of structures.
-More precisely, given a signature $\R$, a function $f : \{0,1\}^* \to \N$ is invariant under $\R$-isomorphisms if for every pair $\A, \mathfrak{B} \in \ostr[\R]$ such that $\A$ and
-$\mathfrak{B}$ are isomorphic, it holds that $f(\enc(\A)) = f(\enc(\mathfrak{B}))$. It is worth noticing that this is a usual assumption in descriptive complexity~\cite{L04}.
-%over $\R$-structures.
 \begin{defi} \label{def:cap}
@@
 
-		\item for every $f \in \CC$ that is invariant under $\R$-isomorphisms, there exists $\alpha \in \FF$ such that   $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A \in \ostr[\R]$.
+		\item for every $f \in \CC$, there exists $\alpha \in \FF$ such that   $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A \in \ostr[\R]$.
 	\end{enumerate} 
@@
 
-For a given signature $\R$, we define $\ostr[\R]^*$ as
-\begin{multline*}
-\ostr[\R]^* = \{(\A,v,V) \mid \A\in\ostr[\R]\text{,} \\ \text{$v$ ($V$) is a first-order (second-order) assignment for $\A$}  \}.
-\end{multline*}
-
 \begin{prop} \label{prop:capture-shP}
@@
 \qed
-Recall  that every function class $\# \LL$ is contained in $\eqso(\LL)$ (see Section~\ref{sec:previous}).
-Thus, it directly follows from~\cite{SalujaST95}  that every $\shp$-function can be defined in $\eqso(\LL)$.
-The other direction of Proposition \ref{prop:capture-shP} follows by the fact that $\shp$ is closed under first- and second-order sum.
 
@@
 Similarly, we use $=$ to denote equality between tuples, as $\bar{x} = \bar{y}$ denotes $\bigwedge_{i = 1}^m(x_i = y_i)$, and also $\bar{x}\leq\bar{y}$ denotes $\bar{x} < \bar{y} \vee \bar{x} = \bar{y}$.
-Furthermore, If $\bar{x} = (x_1,\ldots,x_m)$ ($\bar{X} = (X_1,\ldots,X_m)$) is a tuple of first-order (second-order) variables, we denote $\sa{\bar{x}}\alpha$ for $\sa{x_1}\cdots\sa{x_m}\alpha$ and $\sa{\bar{X}}\alpha$ for $\sa{X_1}\cdots\sa{X_m}\alpha$ for each $\qso$ formula $\alpha$.
+%Furthermore, If $\bar{x} = (x_1,\ldots,x_m)$ ($\bar{X} = (X_1,\ldots,X_m)$) is a tuple of first-order (second-order) variables, we denote $\sa{\bar{x}}\alpha$ for $\sa{x_1}\cdots\sa{x_m}\alpha$ and $\sa{\bar{X}}\alpha$ for $\sa{X_1}\cdots\sa{X_m}\alpha$ for each $\qso$ formula $\alpha$.
 
 The {\em conditional count} symbol $(\varphi \mapsto \alpha)$ is defined as $(\neg\varphi + (\varphi\cdot\alpha))$ for given $\so$ formula $\varphi$ and $\qso$ formula $\alpha$.
-Note that for each $(\A,v,V) \in \ostr[\R]^*$,
+Note that for each $\A \in \ostr[\R]$, and each first-order (second-order) assignment $v$ ($V$) over $\A$,
 \[
@@
 \sem{\alpha}(\A,v,V) &\text{if } (\A,v,V)\models\varphi,\\
-0 &\text{otherwise}.
+1 &\text{otherwise}.
 \end{cases}
@@
 \qed
-To prove this theorem, 
+%To prove this theorem, 
 %capture $\fp$ 
-one first shows that every formula in $\qfo(\lfp)$ can be evaluated in polynomial time. 
-Indeed, $\lfp$ is a polynomial-time logic~\cite{I86,vardi1982complexity}, and the sum and product quantifiers can also be computed in polynomial time. 
-For the other direction, one has to use $\qfo(\lfp)$ to simulate the run of a polynomial time TM $M$ computing a function, in particular using the quantitative quantifiers to reconstruct the natural number returned by $M$ in the output tape.
-It is important to notice that the alternation between sum and product quantifiers is crucial for this reconstructions and, thus, crucial for capturing $\fp$.
+%one first shows that every formula in $\qfo(\lfp)$ can be evaluated in polynomial time. 
+%Indeed, $\lfp$ is a polynomial-time logic~\cite{I86,vardi1982complexity}, and the sum and product quantifiers can also be computed in polynomial time. 
+%For the other direction, one has to use $\qfo(\lfp)$ to simulate the run of a polynomial time TM $M$ computing a function, in particular using the quantitative quantifiers to reconstruct the natural number returned by $M$ in the output tape.
+%It is important to notice that the alternation between sum and product quantifiers is crucial for this reconstructions and, thus, crucial for capturing $\fp$.
 
@@
 \qed
-The proof of the previous theorem follows the same line as for the logical characterization of $\fp$: one shows that each function in $\qso(\pfp)$ can be computed in $\fpspace$ and, conversely, the output of a polynomial-space TM can be reconstructed by using $\pfp$ and quantitative quantifiers.
+%The proof of the previous theorem follows the same line as for the logical characterization of $\fp$: one shows that each function in $\qso(\pfp)$ can be computed in $\fpspace$ and, conversely, the output of a polynomial-space TM can be reconstructed by using $\pfp$ and quantitative quantifiers.
 

### 2017-03-23 LMCS/introduction.tex
journal 23 mar

%!TEX root = main.tex

%\marcelo{Enfatizar el rol fundamental de la logica para obtener los resultados, por ejemplo para obtener cerrado bajo menos uno}
%
%\marcelo{Poner un comentario sobre funciones totales versus parciales, dado que estamos considerando clases con funciones totales}
%
%\marcelo{Enfatizar el rol fundamental de la logica para obtener los resultados, por ejemplo para obtener cerrado bajo menos uno}
%
%Strategy:
%\begin{enumerate}
%	\item Descriptive complexity and application.
%	
%	\item Counting complexity classes. 
%	
%	\item Our contribution in terms of logic.
%	
%	\item sharP and its structure. 
%	
%	\item Syntactic classes with good properties. 
%\end{enumerate}
%
%\cristian{Aqui empieza la intro.}

The goal of descriptive complexity is to measure the complexity of a problem in terms of the logical constructors needed to express it~\cite{immerman1999descriptive}. 
The starting point of this branch of complexity theory is Fagin's theorem~\cite{F75}, which states that $\np$ is equal to existential second-order logic. Since then, many more complexity classes have been characterized in terms of logics (see \cite{G07} for a survey) and descriptive complexity has found a variety of applications in different areas~\cite{immerman1999descriptive, L04}.
For instance, Fagin's theorem was the key ingredient to define the class {\sc MaxSNP}~\cite{PY91}, which was later shown to be a fundamental class in the study of hardness of approximation \cite{ALMSS98}. 
It is important to mention here that the definition of {\sc MaxSNP} would not have been possible without the machine-independent point of view of descriptive complexity, as pointed out in~\cite{PY91}.

Counting problems differ from decision problems in that the value of a function has to be computed.
More generally, a counting problem corresponds to computing a function $f$ from a set of instances (e.g. graphs, formulae, etc) to natural numbers.\footnote{This value is usually associated to counting the number of solutions
	%for a given instance 
	in a search problem, but here we consider a more general definition.} 
The study of counting problems has given rise to a rich theory of counting complexity classes \cite{HV95,F97,arora2009computational}. Some of these classes are natural counterparts of some classes of decision problems; for example, $\fp$ 
%(resp., $\fpspace$) 
is the class of all functions that can be computed in polynomial time, 
%(resp., polynomial space), 
the natural counterpart of $\ptime$.
% ($\pspace$ resp.). 
However, other function complexity classes have emerged from the need to understand the complexity of some computation problems for which little can be said if their decision counterparts are considered. This is the case of the class $\shp$, a counting complexity class introduced in \cite{Valiant79} to prove that natural problems like counting the number of satisfying assignments of a propositional formula or the number of perfect matchings of a bipartite graph~\cite{Valiant79} are difficult, namely, $\shp$-complete.
Starting from $\shp$,
many more natural 
%the zoo of 
counting complexity classes have been defined, such as 
%was open with other natural counting classes like 
$\shl$, $\spp$ and $\gp$~\cite{HV95,F97}.
%among others~\cite{HV95,F97}.

Although counting problems play a prominent role in computational complexity, descriptive complexity for this type of problems has not been systematically studied and it is not as developed as for the case of decision problems. Insightful characterizations of $\shp$ and some of its extensions have been provided \cite{SalujaST95,ComptonG96}. However, these characterizations do not define function problems in terms of a logic, but instead in terms of some counting problems associated to a logic like $\fo$. Thus, it is not clear how these characterizations can be used to provide a general descriptive complexity framework for counting complexity classes like $\fp$ and $\fpspace$ (the class of functions computable in polynomial space). 
%It should be mentioned that logical definability has also been studied for the case of optimization problems \cite{KT94} and computation over the real numbers \cite{GM95}. As for the previous cases, it is not clear how these approaches can be extended to provide logical characterizations of a variety of function complexity classes.

In this paper, we propose to study the descriptive complexity of counting complexity classes in terms of Weighted Logics (WL)~\cite{DrosteG07}, a general logical framework that combines Boolean formulae (e.g. in $\fo$ or $\so$) with operations over a fixed semiring (e.g. $\bbN$). 
Specifically, we propose a restriction of WL over natural numbers, called Quantitative Second Order Logics (QSO), and study its expressive power for defining counting complexity classes over ordered structures. 
As a proof of concept, we show that natural syntactical fragments of $\qso$ captures counting complexity classes like $\shp$, $\spp$, $\fp$ and $\fpspace$.
Furthermore, by slightly extending the framework we can prove that $\qso$ can also capture classes like $\gp$ and $\optp$, showing the robustness of our approach.

The next step is to use the machine-independent point of view of $\qso$ to search for subclasses of $\shp$ with some fundamental properties.
%inside $\shp$. 
The question here is, what properties are desirable for a subclass of $\shp$?
First, it is desirable to have a class of counting problems whose associated decision versions are tractable, in the sense that one can decide in 
%$\ptime$ 
polynomial time whether the value of the function is greater than $0$. 
In fact, this requirement is crucial in order to find
%have any chance of finding 
efficient approximation algorithms for a given function (see Section~\ref{sec:syntactic}).
Second, we expect that the class is closed under basic arithmetical operations like sum, multiplication and subtraction by one. 
This is a common topic for counting complexity classes; for example, it is known that $\shp$ is not closed under subtraction by one (under some complexity-theoretical assumption). 
Finally, we want a class with natural complete problems, which characterize all problems in it.
%from the complexity point of view.

In this paper, we give the first results towards defining subclasses of $\shp$ that are robust in terms of existence of efficient approximations, having good closure properties, and existence of natural complete problems. 
Specifically, we introduce a syntactic hierarchy inside $\shp$, called $\eqso(\fo)$-hierarchy, and we show that it is closely related to the $\fo$-hierarchy introduced in~\cite{SalujaST95}. 
Looking inside the $\eqso(\fo)$-hierarchy, we propose the class $\eqso(\logex{1})$ and show that every function in it has a tractable associated decision version, and it is closed under sum, multiplication, and subtraction by one.
%minus one. 
Unfortunately, it is not clear whether this class admits 
%we cannot show a
a natural complete problem.
% for $\eqso(\logex{1})$.
Thus, 
%Despite of this 
we also introduce a Horn-style syntactic class, inspired by~\cite{G92},
%the approach in~\cite{G92}, 
that has tractable associated decision versions and a natural complete problem.

After studying the 
%internal 
structure of $\shp$, we move beyond $\qso$ by introducing new quantifiers. 
By adding variables for functions on top of $\qso$, we introduce a quantitative least fixed point operator to the logic. 
Adding finite recursion to a numerical setting is subtle since functions over natural numbers can easily diverge without finding any fixed point. 
By using the support of the functions, we give a natural halting condition that generalizes the least fixed point operator of Boolean logics.
%\martin{creo que la noción de {\em support} no es muy conocida. Yo cambiaría "By using the support of the functions" por algo que no lo d\'e por sabido como "By using the {\em support} operator over functions"}
Then, with a quantitative recursion at hand we show how to capture $\fp$ from a different perspective and, moreover, how to restrict recursion to capture lower complexity classes 
%below $\fp$ 
such as~$\shl$, the counting version of $\nlog$.
%(Nondeterministic Logarithmic-space).

\smallskip

\noindent{\bf Organisation.} The main terminology used in the paper is given in Section~\ref{sec:preliminaries}. Then the logical framework is introduced in Section~\ref{sec:logic}, and it is used to capture standard counting complexity classes in Section~\ref{sec:complexity}. The structure of $\shp$ is studied in Section~\ref{sec:syntactic}. Section~\ref{sec:beyond} is devoted to define recursion in $\qso$, and to show how to capture classes below $\fp$. 
Finally, we give some concluding remarks in Section~\ref{sec:conclusions}. 
Due to the lack of space, the %complete 
proofs of the results of the paper can be found at 
%a separate appendix that can be downloaded from 
\verb+http://marenas.sitios.ing.uc.cl/ls17.pdf+

### 2017-03-23 LMCS/preliminaries.tex
journal 23 mar

@@
  \end{align*}
-where $x,y \in \fv$, $R \in \R$, $\bar u$ is a tuple of (non-necessarily distinct) variables from $\fv$ whose length is $\arity(R)$, $\top$ is a reserved symbol to represent a tautology, $X \in \sv$, $\bar v$ is a tuple of (non-necessarily distinct) variables from $\fv$ whose length is $\arity(X)$, and $x \in \fv$.
+where $x,y \in \fv$, $R \in \R$, $\bar u$ is a tuple of (not necessarily distinct) variables from $\fv$ whose length is $\arity(R)$, $\top$ is a reserved symbol to represent a tautology, $X \in \sv$, $\bar v$ is a tuple of (not necessarily distinct) variables from $\fv$ whose length is $\arity(X)$, and $x \in \fv$.
 

### 2017-03-23 LMCS/proof-capture-fp.tex
journal 23 mar

@@
-For the first condition, let $\alpha\in\qfo(\lfp)$ over some signature $\R$.
-Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{1,\ldots,n\}$.
-In the procedure we slightly extend the grammar of $\qfo(\lfp)$ to include constants.
-We replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\sa{x} \beta(x)$ is replaced by $(\beta(1)+\cdots+\beta(n))$ and $\pa{x}\beta(x)$ is replaced by $(\beta(1)\cdot\,\cdots\,\cdot\beta(n))$.
-Then each sub-formula $\varphi\in\lfp$ in $\alpha$ is evaluated in polynomial time and replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size (recall that $\alpha$ is fixed) which is evaluated and lastly given as output.
-Note that $f\in\fp$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
+For the first condition, let $\alpha\in\qfo(\lfp)$ over some signature $\R$, defined by the grammar in \ref{syntax}.
+Notice that for each $\lfp$ formula $\varphi$,  checking whether $\A\models\varphi$ can be done in deterministic polynomial time on the size of $\A$ [cite here].
+Also, the constant function $s$ can be trivially simulated in $\fp$.
+These facts, together with closure properties of $\fp$ of polynomial sum and product [cite here?] suffice to show that the function represented by $\alpha$ is in $\fp$.
 
-For the second condition, let $f\in \fp$ defined over some signature $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$ (i.e. $n^\ell$ is an upper bound for the output size), where $\A$ has a domain of size $n$.
+For the second condition, let $f\in \fp$ and consider some signature $\R$.
+Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq \size{\A}^\ell$ (i.e. $\size{\A}^\ell$ is an upper bound for the output size).
 Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
-Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
+Define a language
+\[
+L = \{(\A,a_1,\ldots,a_{\ell})\mid a_1,\ldots,a_{\ell}\in A \text{ and the } (a_1,\ldots,a_{\ell})\text{-th bit of }f(\enc(\A))\text{ is 1}\}.
+\]
+
+%Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
 Let $m$ be the position of $\bar{a}$ in the lexicographic order of the tuples in $A^{\ell}$.
 The procedure then computes the $m$-th bit of $f(\enc(\A))$, from least to most significant.
-Since this procedure works in polynomial time, it can be described by an $\lfp$ formula $\Phi(\bar{x})$.
+Since this language is in $\ptime$, by \cite{I86} there exists a formula $\Phi(\bar{x})$ in $\lfp$ such that $\A\models\Phi(\bar{a})$ if and only if $(\A,\bar{a})\in L$. 
 Then we use
@@
 $$
-where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
+where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2)$.
+This formula takes the value $2^m$ if there exists $m$ tuples in $A^{\ell}$ that are smaller than $\bar{x}$.
 Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.

### 2017-03-23 LMCS/proof-capture-fpspace-poly.tex
journal 23 mar

@@
-For the first condition, let $\alpha\in\qfo(\pfp)$ over some signature $\R$.
-Let $f$ be a function over $\R$ defined by the following procedure.
-Let $\enc(\A)$ be an input, where $\A$ is an ordered structure over $\R$ with domain $A = \{1,\ldots,n\}$.
-In the procedure we slightly extend the grammar of $\qfo(\pfp)$ to include constants.
-We replace each first order sum and first order product in $\alpha$ by an expansion using the elements in $A$.
-This is, $\sa{x} \beta(x)$ is replaced by $(\beta(1)+\cdots+\beta(n))$ and $\pa{x}\beta(x)$ is replaced by $(\beta(1)\cdot\,\cdots\,\cdot\beta(n))$.
-Then each sub-formula $\varphi\in\pfp$ in $\alpha$ is evaluated in polynomial space and replaced by 1 if $\A\models\varphi$ and by 0 otherwise.
-The resulting formula is an arithmetic expression of polynomial size  (recall that $\alpha$ is fixed) which is evaluated and lastly given as output.
-Note that $f\in\nfpspace$ and $f(\enc(\A)) = \sem{\alpha}(\A)$.
+For the first condition, [analogous to 4.4]
 
-For the second condition, let $f\in \fpspace$ defined over some signature $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$, where $\A$ has a domain of size $n$.
-Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
-Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
-Let $m$ be the position of $\bar{a}$ in the lexicographic order of the tuples in $A^{\ell}$.
-The procedure then computes the $m$-th bit of $f(\enc(\A))$, from least to most significant.
-Since this procedure works in polynomial space, it can be described by an $\pfp$ formula $\Phi(\bar{x})$.
-Then we use
-$$
-\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot\varphi(\bar{x}),
-$$
-where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2).$ Note that if $\bar{a} \in A^{\ell}$ is the $m$-th tuple in the given order (starting from 0), then $\sem{\varphi(\bar{a})}(\A) = 2^{m}$.
-Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-Then, $\alpha$ is in $\qfo(\pfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
+For the second condition, [analogous to 4.4]

### 2017-03-23 LMCS/proof-capture-shP.tex
journal 23 mar

@@
-We construct recursively a $\shp$-machine $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
+Saluja et al.~\cite{SalujaST95} proved that $\shp = \sfo$.
+Hence, given that every function in $\sfo$ can also be defined $\eqso(\fo)$ (see Section~\ref{sec:previous}), we have that every function in $\shp$ can be defined in $\eqso(\fo)$.
+
+For the other direction, let $\alpha\in\eqso(\fo)$ over some signature $\R$, defined by the grammar in \ref{syntax}.
+Notice that given a $\fo$ formula $\varphi$, checking whether $\A\models\varphi$ can be done in deterministic polynomial time on the size of $\A$.
+Also, the constant function $s$ for some $s\in\nat$ can be trivially simulated in $\shp$.
+These facts, together with the closure under exponential sum and polynomial product of $\shp$\cite{F97}, suffice to show that the function represented $\alpha$ is also in $\shp$.
+
+%We construct recursively a $\shp$-machine $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.
 This machine, on input $(\A,v,V)$ accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
@@
 If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-%If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-%If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$. 
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
+%If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
+%If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
 This covers all possible cases for $\alpha$.
@@
 We have that $f$ is a $\shp$-function over $\R$ and $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the other direction, note that Saluja et al.~\cite{SalujaST95} proved that $\shp = \sfo$. 
-%We also have that $\sqso(\fo)$ captures $\#\fo$ over ordered structures so for each $f\in \shp$ let $\alpha \in \sqso(\fo)$ be its corresponding formula. 
-Since a function in $\sfo$ can also be defined $\eqso(\fo)$ (see Section~\ref{sec:previous}), the condition holds.

### 2017-03-23 LMCS/proof-capture-spanP.tex
journal 23 mar

@@
-Similar than the previous proof, we construct recursively a $\spp$ machine $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ distinct accepting outputs for each $(\A,v,V) \in \ostr[\R]^*$.
+Compton et al.~\cite{ComptonG96} proved that $\spp = \#\eso$. Since a function in $\#\eso$ can also be defined in $\eqso(\eso)$, then $\eqso(\eso)$ captures $\spp$ over ordered structures.
+
+For the other direction, notice that given an $\eso$ formula $\varphi$, checking whether $\A\models\varphi$ can be done in non-deterministic polynomial time on the size of $\A$ [Fagin's].
+To ensure that this $\spp$ machine adds at most one different output, the same string is written in each non-deterministic path.
+Furthermore, the constant function $f = s$ for some $s\in\nat$ can be trivially simulated in $\spp$.
+These facts, together with the closure properties of $\spp$ [where?] suffice to show that the function represented by $\alpha$ is also in $\spp$.
+
+%Similar than the previous proof, we construct recursively a $\spp$ machine $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.
+This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ distinct accepting outputs for each $\A \in \ostr[\R]$.
 Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time \cite{F75}, and accepts if and only if the condition holds true. 
-If $\alpha$ is a constant $s$, then the machine produces $s$ branches and accepts in all of them. 
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.  
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$. 
-If $\alpha = \sa{X} \beta$, it chooses $B\in A^{\arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$. 
-This covers all possible cases for $\alpha$.
-Additionally, the machine produces a different output on each path. This can be done by printing the trace of all the non-deterministic choices.
-However, when the machine starts checking whether $(\A,v,V)\models\varphi$ for some $\eso$ formula $\varphi$, it stops printing in the output tape.
+%If $\alpha$ is a $\eso$-formula $\varphi$ it checks if $(\A,v,V)\models\varphi$ non-deterministically in polynomial time \cite{F75}, and accepts if and only if the condition holds true. 
+%If $\alpha$ is a constant $s$, then the machine produces $s$ branches and accepts in all of them. 
+%If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.  
+%If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$. 
+%If $\alpha = \sa{X} \beta$, it chooses $B\in A^{\arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$. 
+%This covers all possible cases for $\alpha$.
+%Additionally, the machine produces a different output on each path. This can be done by printing the trace of all the non-deterministic choices.
+%However, when the machine starts checking whether $(\A,v,V)\models\varphi$ for some $\eso$ formula $\varphi$, it stops printing in the output tape.
 This way the machine produces exactly one output from that point onwards.
-Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\spp$ function over $\R$ and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the other direction, Compton et al.~\cite{ComptonG96} proved that $\spp = \#\eso$.
-Since a function in $\#\eso$ can also be defined in $\eqso(\eso)$, then $\eqso(\eso)$ captures $\spp$ over ordered structures.
+%Let $\alpha$ be a formula in $\eqso(\eso)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the number of accepting outputs of $M_{\alpha}$ on input $(\A,v,V)$ for some $\A \in \ostr[\R]$.
+%We have that $f$ is a $\spp$ function over $\R$ and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.

### 2017-03-27 LMCS/proof-capture-fp.tex
journal 27 mar

@@
-For the first condition, let $\alpha\in\qfo(\lfp)$ over some signature $\R$, defined by the grammar in \ref{syntax}.
-Notice that for each $\lfp$ formula $\varphi$,  checking whether $\A\models\varphi$ can be done in deterministic polynomial time on the size of $\A$ [cite here].
+For condition (1), notice that for each $\lfp$ formula $\varphi$, checking whether $\A\models\varphi$ can be done in deterministic polynomial time on the size of $\A$\cite{I83}.
 Also, the constant function $s$ can be trivially simulated in $\fp$.
-These facts, together with closure properties of $\fp$ of polynomial sum and product [cite here?] suffice to show that the function represented by $\alpha$ is in $\fp$.
+Furthermore, $\fp$ is closed under polynomial sum and multiplication.
+This suffices to show that the condition holds.
 
-For the second condition, let $f\in \fp$ and consider some signature $\R$.
+For condition (2), let $f\in \fp$ and consider some signature $\R$.
 Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq \size{\A}^\ell$ (i.e. $\size{\A}^\ell$ is an upper bound for the output size).
-Let $\bar{x} = (x_1,\ldots,x_{\ell})$.
+Let $\bar{x} = (x_1,\ldots,x_{\ell})$ and $\bar{a} = (a_1,\ldots,a_{\ell})$.
 Define a language
@@
 \]
-
 %Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
@@
 The procedure then computes the $m$-th bit of $f(\enc(\A))$, from least to most significant.
-Since this language is in $\ptime$, by \cite{I86} there exists a formula $\Phi(\bar{x})$ in $\lfp$ such that $\A\models\Phi(\bar{a})$ if and only if $(\A,\bar{a})\in L$. 
+Since this language is in $\ptime$, by \cite{I83} there exists a formula $\Phi(\bar{x})$ in $\lfp$ such that $\A\models\Phi(\bar{a})$ if and only if $(\A,\bar{a})\in L$. 
 Then we use
@@
 where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2)$.
-This formula takes the value $2^m$ if there exists $m$ tuples in $A^{\ell}$ that are smaller than $\bar{x}$.
+This formula takes the value $2^m$ if there exist $m$ tuples in $A^{\ell}$ that are smaller than $\bar{x}$.
 Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-In other words, $\Phi(\bar{x})$ simulates the behavior of the $\fp$-machine and the formula $\alpha$ reconstruct the binary output.
+In other words, $\Phi(\bar{x})$ simulates the behavior of the $\fp$-machine and the formula $\alpha$ reconstructs the binary output.
 Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.

### 2017-03-27 LMCS/proof-capture-fpspace-poly.tex
journal 27 mar

@@
-For the first condition, [analogous to 4.4]
-
-For the second condition, [analogous to 4.4]
+In this proof, both conditions are analogous to theorem \ref{theo:capture-fp}.
+For condition (1), each $\pfp$ formula $\varphi$ can be evaluated in $\pspace$.
+The mentioned closure properties are also possessed by $\nfpspace$.
+For condition (2) we define the same variables and the same language $L$, which in this case is in $\pspace$.
+The same construction of $\alpha$, which in turn is in $\qfo(\pfp)$, can be used to show that the condition holds.

### 2017-03-27 LMCS/proof-capture-fpspace.tex
journal 27 mar

@@
-To show how to evaluate a $\qso(\pfp)$-formula, we construct recursively a $\shpspace$-machine $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
+For condition (1), notice that each $\pfp$ formula can be evaluated in deterministic polynomial space.
+Also, the constant function $f = s$ can be trivially simulated in $\fpspace$ for each $s\in\nat$.
+Furthermore, $\fpspace$ is closed under exponential sum and multiplication.
+This suffices to prove that the condition holds.
+
+%To show how to evaluate a $\qso(\pfp)$-formula, we construct recursively a $\shpspace$-machine $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
 This machine runs in non-deterministic polynomial space and, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
@@
 If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
+%If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
 If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$, continuing to the next $A$-value if the run of $M_{\beta}$ accepts.
@@
 
-For the second condition, let $f\in \fpspace$ defined over some $\R$.
+For the condition (2), we provide a similar proof to theorem \ref{theo:capture-fp}.
+Let $f\in \fpspace$ defined over some $\R$.
 Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq 2^{n^\ell}$ (i.e.
@@
 Furthermore, each relation $X$ indexes a position in the binary output of $f(\enc(\A))$ as follows.
-Consider a polynomial space machine over the $\R$ that receives as input an $\R$-structure $\A$ and a number $p$ encoded by a relation $X$.
+%Consider a polynomial space machine over the $\R$ that receives as input an $\R$-structure $\A$ and a number $p$ encoded by a relation $X$.
 Then the machine accepts if, and only if, the $p$-th bit of $f(\enc(\A))$ is $1$.
-Since this procedure works in polynomial space, it can be described in $\pfp$ \cite{AbiteboulV89} by a formula $\Phi(X)$ where the free variable $X$ encodes the number $p$.
+Define a language
+\[
+L = \{(\A,B)\mid B \subseteq A^{\ell}\text{ and the $B$-th bit of $f(\enc(\A))$ is 1}\}.
+\]
+Since this language is in $\pspace$, it can be described in $\pfp$ \cite{AbiteboulV89} by a formula $\Phi(X)$ where the free variable $X$ encodes the number $p$.
 Then, similar than the previous proof we define:
@@
 where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
-Note that for each $B\subseteq A^{\ell}$ such that $B$ is the $m$-th element in the order $\varphi_{<}(X,Y)$, it holds that $\sem{\varphi(B)}(\A) = 2^m$.
-Therefore, $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
+As in theorem \ref{theo:capture-fp}, this formula takes the value $2^m$ if there exist $m$ predicates that are smaller than $X$.
+Using an analogous argument, we conclude that $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.

### 2017-03-27 LMCS/proof-capture-shP.tex
journal 27 mar

@@
-Saluja et al.~\cite{SalujaST95} proved that $\shp = \sfo$.
-Hence, given that every function in $\sfo$ can also be defined $\eqso(\fo)$ (see Section~\ref{sec:previous}), we have that every function in $\shp$ can be defined in $\eqso(\fo)$.
+In this proof, as we will in the following capture theorems, we refer to definition \ref{def:cap}.
+First, we prove condition (2).
+Saluja et al.
+proved that $\shp = \sfo$\cite{SalujaST95}.
+Hence, given that every function in $\sfo$ can be trivially defined as a formula in $\eqso(\fo)$ (see Section~\ref{sec:previous}), the condition holds.
 
-For the other direction, let $\alpha\in\eqso(\fo)$ over some signature $\R$, defined by the grammar in \ref{syntax}.
+For condition (1), let $\alpha\in\eqso(\fo)$ over some signature $\R$, defined by the grammar in \ref{syntax}.
 Notice that given a $\fo$ formula $\varphi$, checking whether $\A\models\varphi$ can be done in deterministic polynomial time on the size of $\A$.
 Also, the constant function $s$ for some $s\in\nat$ can be trivially simulated in $\shp$.
-These facts, together with the closure under exponential sum and polynomial product of $\shp$\cite{F97}, suffice to show that the function represented $\alpha$ is also in $\shp$.
+These facts, together with the closures under exponential sum and polynomial product of $\shp$\cite{F97}, suffice to show that the function represented by $\alpha$ is in $\shp$.
 

### 2017-03-27 LMCS/proof-capture-spanP.tex
journal 27 mar

@@
-Compton et al.~\cite{ComptonG96} proved that $\spp = \#\eso$. Since a function in $\#\eso$ can also be defined in $\eqso(\eso)$, then $\eqso(\eso)$ captures $\spp$ over ordered structures.
+To prove the condition (2), we use the fact that $\spp = \#\eso$. The condition holds using the same argument as in proposition \ref{prop:capture-shP}.
 
-For the other direction, notice that given an $\eso$ formula $\varphi$, checking whether $\A\models\varphi$ can be done in non-deterministic polynomial time on the size of $\A$ [Fagin's].
-To ensure that this $\spp$ machine adds at most one different output, the same string is written in each non-deterministic path.
+For condition (1), notice that given an $\eso$ formula $\varphi$, checking whether $\A\models\varphi$ can be done in non-deterministic polynomial time on the size of $\A$\cite{fagin1974generalized}.
+To ensure that this $\spp$ machine adds at most one different output, the same string is written as output in each non-deterministic path.
 Furthermore, the constant function $f = s$ for some $s\in\nat$ can be trivially simulated in $\spp$.
-These facts, together with the closure properties of $\spp$ [where?] suffice to show that the function represented by $\alpha$ is also in $\spp$.
+The condition holds analogously to proposition \ref{prop:capture-shP} since $\spp$ possesses the mentioned closure properties as well\cite{OH93}.
 

### 2017-03-29 LMCS/proof-capture-fpspace.tex
journal 29 mar

@@
-For condition (1), notice that each $\pfp$ formula can be evaluated in deterministic polynomial space.
-Also, the constant function $f = s$ can be trivially simulated in $\fpspace$ for each $s\in\nat$.
-Furthermore, $\fpspace$ is closed under exponential sum and multiplication.
-This suffices to prove that the condition holds.
+For condition (1), notice that each $\pfp$ formula can be evaluated in deterministic polynomial space, that the constant function $s$ can be trivially simulated in $\fpspace$ for each $s\in\nat$, and that $\fpspace$ is closed under exponential sum and multiplication.
+This suffices to show that the condition holds.
 

### 2017-03-29 LMCS/proof-capture-optp.tex
journal 29 mar

@@
-Similar than the previous proofs, we construct recursively a non-deterministic polynomial time Turing machine $M_{\alpha}$ with output tape for each $\maxqso(\fo)$ formula $\alpha$ over a signature $\R$.
+In this proof we construct recursively a non-deterministic polynomial time Turing machine $M_{\alpha}$ with output tape for each $\maxqso(\fo)$ formula $\alpha$ over a signature $\R$.
 This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ (in binary) over the output tape of some run, and this value is the maximum value over all runs.

### 2017-03-29 LMCS/proof-capture-spanP.tex
journal 29 mar

@@
-To prove the condition (2), we use the fact that $\spp = \#\eso$. The condition holds using the same argument as in proposition \ref{prop:capture-shP}.
+To prove the condition (2), we use the fact that $\spp = \#(\eso)$. The condition holds using the same argument as in proposition \ref{prop:capture-shP}.
 
 For condition (1), notice that given an $\eso$ formula $\varphi$, checking whether $\A\models\varphi$ can be done in non-deterministic polynomial time on the size of $\A$\cite{fagin1974generalized}.
-To ensure that this $\spp$ machine adds at most one different output, the same string is written as output in each non-deterministic path.
-Furthermore, the constant function $f = s$ for some $s\in\nat$ can be trivially simulated in $\spp$.
+The $\spp$ machine we construct to evaluate the formula produces at most one different output, and does this by writing the same string as output in each non-deterministic path.
+Furthermore, the constant function $s$ for some $s\in\nat$ can be trivially simulated in $\spp$.
 The condition holds analogously to proposition \ref{prop:capture-shP} since $\spp$ possesses the mentioned closure properties as well\cite{OH93}.

### 2017-03-29 LMCS/proof-theo-pi1-pnf.tex
journam 7 mar
journal 29 mar

@@
+%!TEX root = main.tex
+
 We divide the proof in three parts.
@@
 First, we prove that the formula $\alpha_{0} = \left( \sa{X} 1 \right) + 1$ with $\arity(X) = 1$ (i.e.
-the function $2^{n}+1$, where $n$ is the size of the input structure) is not equivalent to any formula in $\loge{0}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ in $\loge{0}$-PNF that is equivalent to the $\eqso(\loge{0})$ formula $\alpha_0$.
-In \cite{SalujaST95}, it was proved that if $\length{\bar{X}} > 0$, then the function defined by $\alpha$, for big enough structures, is always even which is not possible in our case.
+the function $2^{\vert\A\vert}+1$) is not equivalent to any formula in $\loge{0}$-PNF.
+Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ in $\loge{0}$-PNF that is equivalent to $\alpha_0$.
+In \cite{SalujaST95}, it was proved that if $\length{\bar{X}} > 0$, the function defined by $\alpha$ is always even for big enough structures, which is not possible in our case.
 On the other hand, if $\alpha$ is of the form $\sa{\bar{x}}\varphi(\bar{x})$, then $\alpha$ defines a polynomially bounded function which also leads to a contradiction.
@@
 Second, we prove that the formula $\alpha_{1} = 2$ (i.e.
-the constant $2$) is not equivalent to any formula in $\loge{1}$-PNF.
+the constant function $2$) is not equivalent to any formula in $\loge{1}$-PNF.
 Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to the $\eqso(\loge{1})$ formula $2$.
 First, if $\length{\bar{X}} = \length{\bar{x}} = 0$, then the function defined by $\alpha$ is never greater than 1. 
-Therefore, suppose that $\length{\bar{X}} > 0$ or $\length{\bar{x}} > 0$, and consider any ordered structure $\A$. 
+Therefore, suppose that $\length{\bar{X}} > 0$ or $\length{\bar{x}} > 0$, and consider some ordered structure $\A$. 
 Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
-Now consider the ordered structure $\A'$ that is obtained by duplicating $\A$.
-This is, each half of $\A'$ is isomorphic to $\A$.
-Note that $\A'\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$ for $i = 1,2$ and there exists a third assignments $(\bar{B}_1',\bar{b}_1',\bar{a}_1')$ that is isomorphic to $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$ but in the other half to the structure such that $\A'\models\varphi(\bar{B}_1',\bar{b}_1',\bar{a}_1')$.
-We have that $\sem{\alpha}(\A) \geq 3$ which is a contradiction.
+Now consider the ordered structure $\A'$ that is obtained by concatenating $\A$ twice.
+Indeed, each half of $\A'$ is isomorphic to $\A$.
+Note that $\A'\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$ for $i = 1,2$ and there exists a third assignment $(\bar{B}_1',\bar{b}_1',\bar{a}_1')$ that is isomorphic to $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, in the other half of the structure, such that $\A'\models\varphi(\bar{B}_1',\bar{b}_1',\bar{a}_1')$.
+We have that $\sem{\alpha}(\A) \geq 3$ which leads to a contradiction.
 
@@
 \begin{align*}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\exists\bar{v}(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
+\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\exists\bar{v}(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z}))\vee\\
 &(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))\vee\\
@@
 \begin{align*}
-\sa{\bar{X}}\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
-& \hspace{3cm}  \bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})\vee\\
+\sa{\bar{X}}&\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[ \\
+&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
+&\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z}))\vee\\
 &
@@
 &(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
-& \hspace{3cm} (\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee \\
+&(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee \\
 &(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))].

### 2017-03-29 LMCS/proof-theo-pnf-snf.tex
journal 29 mar

Recall that a formula in $\eqso(\LL)$ is written on the following grammar:
\[
\alpha = \varphi \ \mid \ s \ \mid \ (\alpha + \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \sa{X} \alpha,
\]
where $\varphi$ is a formula in $\LL$ and $s\in\nat$. We will construct a recursive function $\tau$ such that for every $\eqso(\LL)$ formula $\alpha$, it outputs an equivalent formula $\tau(\alpha)$ which is in $\LL$-SNF. If $\alpha = \varphi$, let $\tau(\alpha) = \alpha$. If $\alpha = s$, let $\tau(\alpha) = (\top \add \cdots \add \top)$ ($s$ times). We assume that for every sub-formula $\beta$ in $\alpha$, $\tau(\beta)$ is an equivalent formula in $\LL$-SNF. If $\alpha = (\alpha_1 + \alpha_2)$, let $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$. If $\alpha = \sa{x}\beta$, then $\tau(\beta) = \sum_{i = 1}^{k}\beta_i$ for some $k$ where each $\beta_i$ is in $\LL$-PNF. We define $\tau(\alpha) = \sum_{i = 1}^{k}\sa{x}\beta_i$. If $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case. This covers all possible cases for $\alpha$ and we conclude the proof by taking $\tau(\alpha)$ as the desired rewrite of $\alpha$.

### 2017-04-07 LMCS/proof-theo-pi1-pnf.tex
journal 7 abr

@@
 As in Theorem \ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such formula.
-As we showed, there exists an equivalent formula in $\LL$-SNF, so we assume that $\alpha$ is in that form.
-Let $\alpha = \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-SNF.
-Without lost of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
-If not, we replace each $\alpha_i$ for the equivalent formula $\sa{\bar{X}} \sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge\forall z\,Y(z) \wedge \forall z(y \leq z))$.
+Assume that $\alpha$ is in $\LL$-SNF.
+Let $\alpha = \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-PNF.
+Without loss of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
+If not, we replace each $\alpha_i$ by the equivalent formula $\sa{\bar{X}} \sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge\forall z\,Y(z) \wedge \forall z(y \leq z))$.
 
@@
 If $\alpha = \alpha_1 + \alpha_2$, then we assume that $\tau(\alpha_1) = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ and $\tau(\alpha_2) = \sa{\bar{Y}}\sa{\bar{y}}\psi(\bar{Y},\bar{y})$.
-The construction that we will provide for this function works by identifying a ``first'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ and a ``last'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$.
-These are identified by the following formulas:
+Our construction works by identifying a ``first'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ and a ``last'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ using the following formulas:
 \begin{align*}
@@
 In other words, the ``first'' assignment is the one where every second-order predicate is empty and the first-order assignment is the lexicographically smallest, and the ``last'' assignment is the one where every second-order predicate is full and the first-order assignment is the lexicographically greatest.
-We also need to identify assignments that are not first and are not last.
+We also need to identify the assignments that are not first and the ones that are not last.
 We do this by negating the two formulas above and grouping together the first-order variables:
@@
 With this, it is straightforward to show that the above formula is equivalent to $\alpha$.
-However, the formula is not yet in the correct form since it has existential quantifiers in the subformulas $\gamma_{\text{not first}}$ and $\gamma_{\text{not last}}$.
-To solve this, first let take a close look to the complete formula:
+However, the formula is not yet in the correct form since it has existential quantifiers in the sub-formulas $\gamma_{\text{not first}}$ and $\gamma_{\text{not last}}$.
+To solve this, first let us take a close look to the complete formula:
 \begin{align*}
@@
 \begin{align*}
-\sa{\bar{X}}&\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}[ \\
-&(\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
-&\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z}))\vee\\
+\sa{\bar{X}}&\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}\big[ \\
+&[\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
+&\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})]\vee\\
 &
-(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))\vee
+[\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}')]\vee
 \\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
-&(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}')\vee \\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}'))].
+&[\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
+&(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}']\vee \\
+&[\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}')]\big].
 \end{align*}
+
 Finally, consider a $\eqso(\LL)$ formula $\alpha$ in $\LL$-SNF.
-If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we can consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the condition in the hypothesis.
+If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we can consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the hypothesis.

### 2017-04-07 LICS/beyond.tex
minor corrections

@@
 % and, 
- As a proof of concept, we show that how this notion can be used to captures $\fp$.
+ As a proof of concept, we show how this notion can be used to capture $\fp$.
 Moreover, we use this concept
@@
 \end{multline}
-where $h \in \fs$, $\arity(h) = \ell$ and $x_1, \ldots, x_\ell$ is a sequence of (non-necessarily distinct) first-order variables.
+where $h \in \fs$, $\arity(h) = \ell$ and $x_1, \ldots, x_\ell$ is a sequence of (not necessarily distinct) first-order variables.
 Given an $\R$-structure $\A$ with domain $A$, we say that $F$ is a \emph{function assignment} for $\A$ if for every $h \in \fs$ with $\arity(h) = \ell$, we have that $F(h) :  A^\ell \to \N$.
@@
 %proposing 
-this extension on weighted logics, which should be further investigated. 
+this extension of weighted logics, which should be further investigated. 
 % and it would be interesting to study this concept further (i.e. outside recursion).
@@
 we do not need to impose any positive restriction on the formula $\beta(\x,h)$.
-Indeed, since $\beta$ is constructed from monotones operations (i.e. sum and product) over the natural numbers, the resulting operator $T_{\beta}$ is monotone as well.
+Indeed, since $\beta$ is constructed from monotone operations (i.e. sum and product) over the natural numbers, the resulting operator $T_{\beta}$ is monotone as well.
 
@@
 
-Given a relation signature $\R$, the set of transitive $\qfo$ formulae ($\tqfo$-formulae) is defined as an extension of $\qfo$ with the formula
+Given a relational signature $\R$, the set of transitive $\qfo$ formulae ($\tqfo$-formulae) is defined as an extension of $\qfo$ with the formula
 %\begin{multline}
@@
 Given an $\R$-structure $\A$ with domain $A$, define a (directed) graph $\cG_{\psi}(\A) = (N,E)$ such that $N = A^k$ and for every pair $\bar b, \bar c \in N$, it holds that $(\bar b, \bar c) \in E$ if, and only if, $\A \models \psi(\bar b, \bar c)$.
-Similar than for Example~\ref{ex:count-path}, we can count the paths of length at most $|A^k|$ in $ \cG_{\psi}(\A)$ with the formula $\beta_{\psi(\bar{x},\bar{y})}(\x, \y, \t, g)$:
+Just as we did for Example~\ref{ex:count-path}, we can count the paths of length at most $|A^k|$ in $ \cG_{\psi}(\A)$ with the formula $\beta_{\psi(\bar{x},\bar{y})}(\x, \y, \t, g)$:
 \begin{multline*}
@@
 But then the problem is that for every language $L \in \nlog$, it holds that its characteristic function $\chi_L$ is in $\qfo(\tc)$, where $\chi_L(x) = 1$ if $x \in L$, and $\chi_L(x) = 0$ otherwise.
-Thus, if we assume that $\qfo(\tc)$ captures $\shl$ (over ordered structures), then we have that $\chi_L \in \nlog$ for every $L \in \nlog$.
-This implies that $\nlog = \ulog$,\footnote{A decision language $L$ is in $\ulog$ is there exists a logarithmic-space NTM $M$ accepting $L$ and satisfying that $\tma_M(x) = 1$ for every $x \in L$.}
-and thus contradicts the widely believed assumption that $\ulog \subsetneq \nlog$.
+Thus, if we assume that $\qfo(\tc)$ captures $\shl$ (over ordered structures), then we have that $\chi_L \in \shl$ for every $L \in \nlog$.
+This would imply that $\nlog = \ulog$,\footnote{A decision language $L$ is in $\ulog$ is there exists a logarithmic-space NTM $M$ accepting $L$ and satisfying that $\tma_M(x) = 1$ for every $x \in L$.} which is believed to be so\cite{Reinhardt97}.
+%and thus contradicts the widely believed assumption that $\ulog \subsetneq \nlog$.
 %$\qfo(\tc)$ can def

### 2017-04-07 LICS/counting.tex
minor corrections

@@
 %(that is, $(a_1, \ldots, a_1, a_1)$, $(a_1, \ldots, a_1, a_2)$, $\ldots$, $(a_n, \ldots, a_n, a_{n-1})$, $(a_n, \ldots, a_n, a_n)$). 
-Then let $\enc(R_i^\A)$ be a binary string of length $n^\ell$ such that the $i$-th bit of $\enc(R_i^\A)$ is 1 if the $i$-th tuple in the previous enumeration belongs to $R_i^\A$, and 0 otherwise.
+Then let $\enc(R_i^\A)$ be a binary string of length $n^\ell$ such that the $i$-th bit of $\enc(R_i^\A)$ is 1 if the $i$-th tuple in the enumeration belongs to $R_i^\A$, and 0 otherwise.
 Moreover, define the encoding of $\A$, denoted by $\enc(\A)$, as the string
@@
 Notice that this restriction is natural as we want to capture %Since we want to capture
-$\CC$ over a fix set of structures (e.g. graphs, matrices).
+$\CC$ over a fixed set of structures (e.g. graphs, matrices).
 %, it is natural to just consider strings that encodes $\R$-structures. 
@@
 \end{proposition}
-Recall  that every function class $\# \LL$ is contained in $\eqso(\LL)$ (see Section~\ref{sec:previous}).
-Thus, it directly follows from~\cite{SalujaST95}  that every $\shp$-function can be defined in $\eqso(\LL)$.
+Recall  that every function class $\# \LL$ is contained in $\eqso(\LL)$, for some fragment $\LL$ of $\fo$.
+(see Section~\ref{sec:previous}).
+Thus, it directly follows from~\cite{SalujaST95}  that every $\shp$-function can be defined in $\eqso(\fo)$.
 The other direction of Proposition \ref{prop:capture-shP} follows by the fact that $\shp$ is closed under first- and second-order sum.
@@
 Can we capture $\fp$ by using $\# \LL$ for some fragment $\LL$ of $\so$?
-A first attempt could be based on the use of a fragment $\LL$ of $\so$ that capture either $\ptime$ or $\nlog$~\cite{G92}.
+A first attempt could be based on the use of a fragment $\LL$ of $\so$ that captures either $\ptime$ or $\nlog$~\cite{G92}.
 Such an approach fails as $\# \LL$ can encode $\shp$-complete problems in both cases; in the first case, one can encode the problem of counting the number of satisfying assignments of a Horn  propositional formula, while in the second case one can encode the problem of counting the number of satisfying assignments of a 2-CNF propositional formula.
@@
 %built in the ouptape. 
-It is important to notice that the alternation between sum and product quantifiers is crucial for this reconstructions and, thus, crucial for capturing $\fp$.
+It is important to notice that the alternation between sum and product quantifiers is crucial for this reconstruction and, thus, crucial for capturing $\fp$.
 %That is, without these operators is difficult to capture $\fp$.

### 2017-04-07 LICS/extentions.tex
minor corrections

@@
 
-There exist complexity classes that do not fit in our framework because either the output of a function in not a natural number (e.g.
+There exist complexity classes that do not fit in our framework because either the output of a function is not a natural number (e.g.
 a negative number) or the class is not defined purely in terms of arithmetical operations (e.g.
@@
 a constant) comes from $\bbZ$.
-Similar than for $\qso$, we define the fragment $\eqsoz$ as the extension of $\eqso$ with constants in $\bbZ$.
+Just as we did for $\qso$, we define the fragment $\eqsoz$ as the extension of $\eqso$ with constants in $\bbZ$.
 \begin{example}
@@
 %think
-This is an interesting result that shows how robust is $\qso$ when capturing different complexity classes.
+This is an interesting result that shows how robust $\qso$ is when capturing different complexity classes.
 
@@
 each run produces a binary string which is interpreted as a number).
-For instance, MAX-SAT is in $\optp$ as many other optimization versions of $\np$-problems.
+For instance, MAX-SAT is in $\optp$ as many other optimization versions of $\np$-problems are.
 Given that in~\cite{krentel1988complexity} the author does not make the distinction between $\max$ and $\min$, in~\cite{vollmer1995complexity} the authors define the classes $\maxp$ and $\minp$ as the max and min version of the problems in $\optp$ (i.e.
@@
 \end{example}
-Similar than for $\maxp$ and $\minp$, we have to distinguished between the $\max$ and $\min$ fragments of $\optqso$.
+Similar than for $\maxp$ and $\minp$, we have to distinguish between the $\max$ and $\min$ fragments of $\optqso$.
 For this, we define the fragment $\maxqso$ of all $\optqso$ formulae constructed from $\qfo$ operators and $\max$-formulae $\max\{\alpha,\alpha\}$, $\maxa{x} \alpha$ and  $\maxa{X} \alpha$.
@@
 %the formulae $\alpha_1$ and $\alpha_6$ are in $\maxqso$ but $\alpha_2$ and $\alpha_3$ are not. 
-As one could expect, $\maxqso$ and $\minqso$ are the needed logics to capture $\maxp$ and $\minp$.
+As one could expect, $\maxqso$ and $\minqso$ are the logics needed to capture $\maxp$ and $\minp$.
 \begin{theorem} \label{theo:capture-optp}

### 2017-04-07 LICS/framework.tex
minor corrections

@@
 \begin{example}\label{ex:cliques}
-Let $\bG = \{E(\cdot,\cdot),<\}$ be the vocabulary for graphs and $\fG$ be an ordered $\bG$-structure encoding a non-directed graph. 
+Let $\bG = \{E(\cdot,\cdot),<\}$ be the vocabulary for graphs and $\fG$ be an ordered $\bG$-structure encoding an undirected graph. 
 Suppose that we want to count the number of triangles in $\fG$. Then this can be defined as follows:
@@
 %\martin{La notacion $f_{\psi(x,X)}$ la inventamos}
-In~\cite{SalujaST95}, they were proved several results about capturing counting complexity classes which are relevant for our work.
+In~\cite{SalujaST95}, several results were proved about capturing counting complexity classes which are relevant for our work.
 We discuss and use these results in Sections~\ref{sec:complexity} and~\ref{sec:syntactic}.

### 2017-04-07 LICS/preliminaries.tex
minor corrections

@@
 \end{multline*}
-where $x,y \in \fv$, $R \in \R$, $\bar u$ is a tuple of (non-necessarily distinct) variables from $\fv$ whose length is $\arity(R)$, $\top$ is a reserved symbol to represent a tautology, $X \in \sv$, $\bar v$ is a tuple of (non-necessarily distinct) variables from $\fv$ whose length is $\arity(X)$, and $x \in \fv$.
+where $x,y \in \fv$, $R \in \R$, $\bar u$ is a tuple of (not necessarily distinct) variables from $\fv$ whose length is $\arity(R)$, $\top$ is a reserved symbol to represent a tautology, $X \in \sv$, $\bar v$ is a tuple of (not necessarily distinct) variables from $\fv$ whose length is $\arity(X)$, and $x \in \fv$.
 

### 2017-04-07 LICS/syntactic-clo.tex
minor corrections

@@
 Therefore, the desired extension has to be achieved by allowing some local extensions to $\Sigma_1$.
-More precisely, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulae over a signature $\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mentioned any second-order variable).
+More precisely, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulae over a signature $\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mention any second-order variable).
 With this extension we obtain a class with the desired properties.

### 2017-04-07 LICS/syntactic.tex
minor corrections

@@
 
-Counting problems with easy decision versions play a fundamental role in the search of efficient approximations algorithms for functions in $\shp$.
+Counting problems with easy decision versions play a fundamental role in the search of efficient approximation algorithms for functions in $\shp$.
 A fully-polynomial randomized approximation scheme (FPRAS) for a function $f : \Sigma^* \to \bbN$ is a randomized algorithm ${\cal A} : \Sigma^* \times (0,1) \to \bbN$ such that: (1) for every string $x \in \Sigma^*$ and real value $\varepsilon \in (0,1)$, the probability that $|f(x) - {\cal A}(x,\varepsilon)| \leq \varepsilon \cdot f(x)$ is at least $\frac{3}{4}$, and (2) the running time of ${\cal A}$ is polynomial in the size of $x$ and $1/\varepsilon$ \cite{KL83}.
@@
 Clearly $\textit{sat\_one}$ is a $\shp$-complete function whose decision counterpart $L_{\textit{sat\_one}}$ is trivial.
-But should $\textit{sat\_one}$ be part of a robust class of counting function with easy decision versions?
+But should $\textit{sat\_one}$ be part of a robust class of counting functions with easy decision versions?
 The key insight here is that if a function in $\shp$ has an easy decision counterpart $L$, then as $L \in \np$ we expect  to have a polynomial-time algorithm that verifies whether $x \in L$ by constructing witnesses for~$x$.
@@
 However, $\cdnf$ has an easy decision counterpart and admits an FPRAS, while $\csat$ does not satisfy these conditions unless $\ptime = \np$.
-Hence a more strict notion of reduction has to be considered; in particular, the notion of parsimonious reduction (to be defined later) satisfy that if a function $f$ is parsimoniously reducible to a function $g$, then $L_g \in \ptime$ implies that $L_f \in \ptime$ and the existence of an FPRAS for $g$ implies the existence of a FPRAS for $f$.
+Hence a more strict notion of reduction has to be considered; in particular, the notion of parsimonious reduction (to be defined later) satisfies that if a function $f$ is parsimoniously reducible to a function $g$, then $L_g \in \ptime$ implies that $L_f \in \ptime$ and the existence of an FPRAS for $g$ implies the existence of a FPRAS for $f$.
 

### 2017-04-11 LICS/beyond.tex
camera-ready 11 abr

@@
 complexity classes.  
-The distinction in the language between the Boolean and the quantitative level allows us to define operators at the later level that cannot be defined at the former. 
+The distinction in the language between the Boolean and the quantitative level allows us to define operators at the latter level that cannot be defined at the former. 
 As a example showing how fundamental this separation is, consider the issue of extending $\qfo(\fo)$ at the Boolean level in order to capture $\shl$.

### 2017-04-11 LICS/counting.tex
camera-ready 11 abr

@@
 
-In this section, we show that by syntactically restricting $\qso$ one can capture different counting complexity classes. 
+In this section, we show that by syntactically restricting [or extending] $\qso$ one can capture different counting complexity classes. 
 In other words, by using $\qso$ we can extend the theory of descriptive complexity~\cite{immerman1999descriptive} from decision problems to computation problems. 
@@
 We can now formalize the notion of capturing a counting complexity class. 
-From now on, we assume that every counting complexity class $\CC$ consists of functions $f : \{0,1\}^* \to \N$ that are invariant under isomorphism of structures.
+%From now on, we assume that every counting complexity class $\CC$ consists of functions $f : \{0,1\}^* \to \N$ that are invariant under isomorphism of structures.
 More precisely, given a signature $\R$, a function $f : \{0,1\}^* \to \N$ is invariant under $\R$-isomorphisms if for every pair $\A, \mathfrak{B} \in \ostr[\R]$ such that $\A$ and
-$\mathfrak{B}$ are isomorphic, it holds that $f(\enc(\A)) = f(\enc(\mathfrak{B}))$. It is worth noticing that this is a usual assumption in descriptive complexity~\cite{L04}.
+%$\mathfrak{B}$ are isomorphic, it holds that $f(\enc(\A)) = f(\enc(\mathfrak{B}))$. It is worth noticing that this is a usual assumption in descriptive complexity~\cite{L04}.
 %over $\R$-structures.
@@
 %all notions for capturing complexity classes restrict $f \in \CC$ similarly.
+
+[comment about unordered structures]
 

### 2017-04-11 LICS/framework.tex
camera-ready 11 abr

@@
 
-We consider several fragments of $\qso$, which are obtained by restricting the syntax of the Boolean formulae or the use of the quantitative quantifiers.
+We consider several fragments [and extensions] of $\qso$, which are obtained by restricting the syntax of the Boolean formulae or the use of the quantitative quantifiers.
 In this direction, we denote by $\qfo$ the fragment of $\qso$ where second-order sum and product are not allowed. 
@@
 For example, $\alpha_1$ and $\alpha_2$ in Example \ref{ex:cliques} are formulae of $\eqso$, while $\alpha_3$ in Example \ref{exa-perm} is not. 
-We also consider fragments of $\qso$ by further restricting the Boolean part of the logic.
-If $\LL$ is a fragment of $\so$, then we define the quantitative logic $\qso(\LL)$ to be the fragment of $\qso$ obtained by restricting $\varphi$ in \eqref{syntax} to be a formula in $\LL$.
-Moreover, we also restrict other fragments of $\qso$ by using the same idea.
+We also consider fragments [and extensions] of $\qso$ by further restricting the Boolean part of the logic.
+If $\LL$ is a fragment [or extension] of $\so$, then we define the quantitative logic $\qso(\LL)$ to be the fragment of $\qso$ obtained by restricting $\varphi$ in \eqref{syntax} to be a formula in $\LL$.
+Moreover, we also [consider] other fragments [or extension] of $\qso$ by using the same idea.
 %In general, if $\FF$ is a fragment of $\qso$ we define $\FF(\LL)$ to be the fragment of $\FF$ obtained by restricting Boolean formulae to $\LL$.
@@
 
-In the following section, we use different fragments of $\qso$ to capture counting complexity classes. 
+In the following section, we use different fragments [and extensions] of $\qso$ to capture counting complexity classes. 
 %variate $\FF$ and $\LL$ to capture several function complexity classes. 
@@
 al \cite{SalujaST95} define a family of counting classes $\#\LL$ for a fragment $\LL$ of $\fo$.
-For a formula $\varphi(\bar{x},\bar{X})$ over $\R$, the function $f_{\varphi(\bar x, \bar X)}$ is defined as
+[This definition can also be expanded to include extensions of $\fo$ as it was done in \cite{ComptonG96}.] For a formula $\varphi(\bar{x},\bar{X})$ over $\R$, the function $f_{\varphi(\bar x, \bar X)}$ is defined as
 $

### 2017-04-11 LICS/preliminaries.tex
camera-ready 11 abr

@@
 
-In this paper, we consider several fragments of $\so$, in particular first-order logic ($\fo$), least fixed point logic (LFP) and partial fixed point logic (PFP) \cite{L04}.
+In this paper, we consider several fragments [and extensions] of $\so$, in particular first-order logic ($\fo$), least fixed point logic (LFP) and partial fixed point logic (PFP) \cite{L04}.
 Moreover, for every $i \in \N$, we consider the fragment $\Sigma_i$ (resp., $\Pi_i$) of $\fo$, which is the set of $\fo$-formulae of the form

### 2017-04-11 LICS/syntactic-clo.tex
camera-ready 11 abr

@@
 Therefore, the desired extension has to be achieved by allowing some local extensions to $\Sigma_1$.
-More precisely, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulae over a signature $\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mention any second-order variable).
+More precisely, we define $\logex{1}$ as follows.
+An $\fo$-extended quantifier free formula is defined by the grammar
+\[
+\varphi \,:=\, x = y \,\mid\, R(\bar{u}) \,\mid\, X(\bar{v}) \,\mid\, \top \,\mid\, \neg\varphi \,\mid\, (\varphi\vee\varphi) \,\mid\, \psi(\bar{w})
+\]
+where the characterization of the symbols is the same as the one mentioned in the definition of second-order formulae, and $\psi$ is an $\fo$-formula.
+A formula in $\logex{1}$ has the form $\exists\bar{x}\,\varphi(\bar{x})$, where $\varphi$ is an $\fo$-extended quantifier free formula.
 With this extension we obtain a class with the desired properties.

### 2017-04-11 LICS/syntactic.tex
camera-ready 11 abr

@@
 Formally, the decision problem associated to a function $f : \Sigma^* \to \mathbb{N}$ is defined as $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$, and $f$ is said to have an easy decision version if $L_f \in \ptime$.
-Many prominent examples satisfy this property, like computing the number of: perfect matchings of a bipartite graph ($\cpm$)~\cite{Valiant79}, satisfying assignments of a DNF ($\cdnf$)~\cite{DHK05} or Horn ($\chsat$)~\cite{V79b} propositional formula, among others.
+Many prominent examples satisfy this property, like computing the number of: perfect matchings of a bipartite graph ($\cpm$)~\cite{Valiant79}, satisfying assignments of a DNF ($\cdnf$)~\cite{DHK05,KL83} or Horn ($\chsat$)~\cite{V79b} propositional formula, among others.
 

### 2017-04-12 LICS/beyond.tex
Cambios después de conversar con Martin.

@@
 Thus, if we assume that $\qfo(\tc)$ captures $\shl$ (over ordered structures), then we have that $\chi_L \in \shl$ for every $L \in \nlog$.
-This would imply that $\nlog = \ulog$,\footnote{A decision language $L$ is in $\ulog$ is there exists a logarithmic-space NTM $M$ accepting $L$ and satisfying that $\tma_M(x) = 1$ for every $x \in L$.} which is believed to be so\cite{Reinhardt97}.
+This would imply that $\nlog = \ulog$,\footnote{A decision language $L$ is in $\ulog$ is there exists a logarithmic-space NTM $M$ accepting $L$ and satisfying that $\tma_M(x) = 1$ for every $x \in L$.} solving an open problem \cite{Reinhardt97}.
 %and thus contradicts the widely believed assumption that $\ulog \subsetneq \nlog$.

### 2017-04-12 LICS/counting.tex
camera-ready 11 abr
Cambios después de conversar con Martin.

@@
 
-In this section, we show that by syntactically restricting [or extending] $\qso$ one can capture different counting complexity classes. 
+In this section, we show that by syntactically restricting $\qso$ one can capture different counting complexity classes. 
 In other words, by using $\qso$ we can extend the theory of descriptive complexity~\cite{immerman1999descriptive} from decision problems to computation problems. 
@@
 
-[comment about unordered structures]
+\martin{comment about unordered structures}
 
@@
 A first attempt could be based on the use of a fragment $\LL$ of $\so$ that captures either $\ptime$ or $\nlog$~\cite{G92}.
-Such an approach fails as $\# \LL$ can encode $\shp$-complete problems in both cases; in the first case, one can encode the problem of counting the number of satisfying assignments of a Horn  propositional formula, while in the second case one can encode the problem of counting the number of satisfying assignments of a 2-CNF propositional formula.
+Such an approach fails as $\# \LL$ can encode $\shp$-complete problems in both cases since they contain $\sfo$.
+%; in the first case, one can encode the problem of counting the number of satisfying assignments of a Horn  propositional formula, while in the second case one can encode the problem of counting the number of satisfying assignments of a 2-CNF propositional formula.
 A second attempt could be based then on considering a fragment $\LL$ of $\fo$. 
@@
 \end{theorem}
+\martin{rewrite two following paragraphs}
+
 To prove this theorem, 
@@
 %Another natural function complexity class is $\fpspace$~\cite{Ladner89}, the functional version of $\pspace$. 
-Of course, for capturing this class one needs a logical core powerful enough, like $\pfp$, for simulating the run of a polynomial-space TM.
+For capturing this class one can use a logical core powerful enough, like $\pfp$, for simulating the run of a polynomial-space TM.
 % $\pspace$-machine. 

### 2017-04-12 LICS/preliminaries.tex
Cambios después de conversar con Martin.

--- 2017-04-07 LICS/preliminaries.tex
@@
 
-In this paper, we consider several fragments of $\so$, in particular first-order logic ($\fo$), least fixed point logic (LFP) and partial fixed point logic (PFP) \cite{L04}.
+In this paper, we consider several fragments and extensions of $\so$, in particular first-order logic ($\fo$), least fixed point logic (LFP) and partial fixed point logic (PFP) \cite{L04}.
 Moreover, for every $i \in \N$, we consider the fragment $\Sigma_i$ (resp., $\Pi_i$) of $\fo$, which is the set of $\fo$-formulae of the form

### 2017-04-13 LICS/syntactic-clo.tex
Cambios después de conversar con Martin.
camera-ready

--- 2017-04-07 LICS/syntactic-clo.tex
@@
 \end{proposition}
+\martin{Sugerencia:
+
+	we define $\logex{1}$ as $\Sigma_1$ but allowing arbitrary quantifiers in sub-formulas that have only first-order variables.
+This is, atomic formulae include arbitrary first-order formulae.
+[Tal vez incluir un ejemplo] }
 Therefore, the desired extension has to be achieved by allowing some local extensions to $\Sigma_1$.

### 2017-04-13 LICS/todo-list.tex
camera-ready

%!TEX root = main.tex

\section*{\LARGE \textbf{TODO LIST}}

\bigskip

Reviewer 2

\begin{itemize}

\item[$\checkmark$] p1, c2, l12: "type of problem" not "problems"

\item[$\checkmark$] p2, c1, l23: "on top of" not "in"

\item[$\checkmark$] p2, c2, l9 and l12 and again at p10, c1, l2: "not necessarily" rather than "non-"

\item[$\checkmark$] Example III.1: "undirected graph" rather than "non-directed".

\item[$\checkmark$] p4, c2, l-23: "several results were proved" rather than "they were
proved..."

\item[$\checkmark$] p5, c1, l6: delete the word "previous".

\item I am puzzled by the definition of invariance under R-isomorphisms
that appears just before Definition IV.1. You have assumed that
structures are ordered, and the definition of the encoding enc is
based on the resulting lexicographical order. Under these
conditions, if A and B are isomorphic structures, then enc(A) =
enc(B), so any function would be invariant.

You say that the assumption is standard in descriptive complexity.
But there the issue of isomorphism invariance only arises when we
want to define properties of *unordered* structures. For such a
structure A, enc(A) may depend on the particular choice of order
and we want to consider properties that are invariant under this
choice.

In short, because you only consider ordered structures, I don't
think you need to define isomorphism-invariance and this assumption
can also be removed from Definition IV.1(2).
\martin{retire donde mencionaba la idea de invariante a isomorfismo y esta marcado donde hay que agregar el comentario de estructuras ordenadas o no ordenadas}

\item[$\checkmark$] p5, c1, l-23: "fixed" rather than "fix".

\item[$\checkmark$] p5, c1, l-9 to l-7: you haven't said what \cal{L} is? At least for
the third occurrence of this symbol, I think you actually mean FO.

\item[$\checkmark$] p5, c2, l5: "captures" rather than "capture".

\item p5, c2, l5 to l10: this explantion that \#L (when L is LFP or TC)
can express some \#P-complete problems seems out of place when you
have already mentioned (just before Prop. IV.2) that \#FO captures
all of \#P.
\martin{esta marcado donde hay que reescribir esto}

\item[$\checkmark$] p5, c2, l-18: "reconstruction" rather than "reconstructions".

\item Also, is the claim that sum-product alternation "is crucial" for
this something that you can prove? That is, is there a proof that
you cannot capture FP without the alternation?
\martin{esta marcado donde hay que reescribir esto}

\item p5, c2, l-14 to l-12: the sentence beginning with "Of course". Why
is this obvious? I can imagine that the powerful quantitative
mechanisms introduced could compensate for the lack of a powerful
logical core. So, again, this looks to me like a statement that
asks for proof.
\martin{esta marcado donde hay que reescribir esto}

\item p2, c2, l-22: this seems to say that PFP is a fragment of
second-order logic. But, surele we do not know this is the case.

\item Theorem IV.5. You defined QSO(L) for fragments L of second-order
logic, and PFP is not such a fragment, so at this point it is
undefined. Of course, I can see how to define it, but this should
be stated more carefully.
\martin{suger\'i cambiar 'fragments' por 'fragments and extensions' pero no fue muy acogida por el autor R.}

\item[$\checkmark$] p6, c1, l19: "is" not "in".

\item[$\checkmark$] p6, c1, l-13; and p6, c2, l-15; and p11, c2, l-15: "Similar than" sounds wrong. I would use "Just
as..."

\item[$\checkmark$] p6, c1, l-2: "QSO is" rather than "is QSO".

\item[$\checkmark$] p6, c2, l13: "as are many" instead of "as many"

\item[$\checkmark$] p6, c2, l-15: "distinguish" not "distinguished"

\item[$\checkmark$] p6, c2, l-6: "logics needed" rather than "needed logics"

\item[$\checkmark$] p7, c1, l27: Is \cal{A} the algorithm or the function?

\item[$\checkmark$] p7, c1, l-2: "functions" not "function"

\item[$\checkmark$] p7, c2, l22: "satisfies" rather than "satisfy"

\item[$\checkmark$] p9, c1, l13: "mention" not "mentioned"

\item[$\checkmark$] p9, c2, l-2 to l-1: perhaps just "the set of Horn formulae over a
signature R" rather than "the set of formulae psi such that psi is
a Horn formula over a signature R".

\item[$\checkmark$] p10, c1, l-12: delete "that" before "how"

\item[$\checkmark$] p10, c1, l-11: "capture" not "captures"

\item[$\checkmark$] p10, c2, l19: "of weighted logics" not "on"

\item[$\checkmark$] p11, c2, l12: "monotone" not "monotones"

\item[$\checkmark$] p11, c2, l29: "relational" not "relation"

\item[$\checkmark$] p12, c1, l18: "the widely believed assumption" that UL is properly
contained in NL? I don't think this assumption is widely believed
among complexity theorists. I think UL = NL is more widely
believed. See the references:
1. Klaus Reinhardt, Eric Allender: Making Nondeterminism
Unambiguous. SIAM J. Comput. 29(4): 1118-1131 (2000)
2. Eric Allender, Klaus Reinhardt, Shiyu Zhou: Isolation, Matching,
and Counting Uniform and Nonuniform Upper Bounds. 
J. Comput. Syst. Sci. 59(2): 164-181 (1999).
\martin{corregido en el ultimo parrafo antes de las conclusiones}
\end{itemize}

Reviewer 3

\begin{itemize}

\item[$\checkmark$] What is the strength of full QSO without any fixed-point operator? This is an apparent gap in the story you are telling.
\martin{esta mencionado en las conclusiones}

\item[$\checkmark$] After Corollary IV.8: "This is an interesting result..." Generally, I think that you should let the reader decide which results are interesting and which are not. I recommend to cut the "interesting" here.

\item[$\checkmark$] First paragraph of Section V: I find it weird that you cite [6] for the problem \#DNF. This problem has been studied long before; it was at least already known to be \#P-complete in [24], but I would guess Valiant already remarked in his early papers that the problem is \#P-complete (although I have not checked this).
\martin{cada vez que se menciona [6] por dnf lo cambie por [6,24]}

\item after Proposition V.6: Please make the definition of $\Sigma_1[FO]$ more readable. I had a hard time parsing it.
\martin{agregué una nota de cómo se puede hacer un poco más legible y además más corto}

\item[$\checkmark$] Typo: "this notion can be used to captures" -> capture

\item[$\checkmark$] After Theorem VI.4: "at the later level" -> latter

\item[$\checkmark$] I did not get the definition of min(x) in the appendix. Should the $<$ be $>$? Is this just supposed to define the lexicographically smallest of all tuples, so $a^m$ where a is the minimal element of the domain?

\end{itemize}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

### 2017-04-13 LICS/syntactic-clo.tex
camera-ready 13 abr

@@
 
-	we define $\logex{1}$ as $\Sigma_1$ but allowing arbitrary quantifiers in sub-formulas that have only first-order variables.
-This is, atomic formulae include arbitrary first-order formulae.
-[Tal vez incluir un ejemplo] }
+	we define $\logex{1}$ as $\Sigma_1$ but allowing arbitrary quantifiers in sub-formulas that have only first-order variables. [Tal vez agregar un ejemplo para que se entienda de donde viene el [FO] ]}
 Therefore, the desired extension has to be achieved by allowing some local extensions to $\Sigma_1$.

### 2017-04-17 LICS/beyond.tex
camera-ready
Cambios de lo faltante (revisado con Marcelo).

@@
 Thus, if we assume that $\qfo(\tc)$ captures $\shl$ (over ordered structures), then we have that $\chi_L \in \shl$ for every $L \in \nlog$.
-This would imply that $\nlog = \ulog$,\footnote{A decision language $L$ is in $\ulog$ is there exists a logarithmic-space NTM $M$ accepting $L$ and satisfying that $\tma_M(x) = 1$ for every $x \in L$.} solving an open problem \cite{Reinhardt97}.
+This would imply that $\nlog = \ulog$,\footnote{A decision language $L$ is in $\ulog$ is there exists a logarithmic-space NTM $M$ accepting $L$ and satisfying that $\tma_M(x) = 1$ for every $x \in L$.} solving an outstanding open problem \cite{Reinhardt97}.
+
 %and thus contradicts the widely believed assumption that $\ulog \subsetneq \nlog$.

### 2017-04-17 LICS/counting.tex
camera-ready 13 abr
Cambios de lo faltante (revisado con Marcelo).

@@
 
-\martin{comment about unordered structures}
-
 What counting complexity classes can be captured with fragments of $\qso$?
@@
 Can we capture $\fp$ by using $\# \LL$ for some fragment $\LL$ of $\so$? 
-A first attempt could be based on the use of a fragment $\LL$ of $\so$ that captures either $\ptime$ or $\nlog$~\cite{G92}.
+%A first attempt could be based on the use of a fragment $\LL$ of $\so$ that captures either $\ptime$ or $\nlog$~\cite{G92}.
 Such an approach fails as $\# \LL$ can encode $\shp$-complete problems in both cases since they contain $\sfo$.
 %; in the first case, one can encode the problem of counting the number of satisfying assignments of a Horn  propositional formula, while in the second case one can encode the problem of counting the number of satisfying assignments of a 2-CNF propositional formula.
-A second attempt could be based then on considering a fragment $\LL$ of $\fo$. 
+A first attempt could be based on considering a fragment $\LL$ of $\fo$. 
 But even if we consider the existential fragment $\Sigma_1$ of $\fo$ the approach fails, as $\# \Sigma_1$ can encode $\shp$-complete problems like counting the number of satisfying assignments of a 3-DNF propositional formula\cite{SalujaST95}.
-One last attempt could be based on disallowing the use of second-order free variables in $\sfo$.
+A second attempt could be based on disallowing the use of second-order free variables in $\sfo$.
 But in this case one
@@
 \end{theorem}
-\martin{rewrite two following paragraphs}
 
@@
 %built in the ouptape. 
-It is important to notice that the alternation between sum and product quantifiers is crucial for this reconstruction and, thus, crucial for capturing $\fp$.
+It is important to notice that the alternation between sum and product quantifiers is used for this reconstruction.
+%and, thus, crucial for capturing $\fp$.
 %That is, without these operators is difficult to capture $\fp$.

### 2017-04-17 LICS/syntactic-clo.tex
Cambios de lo faltante (revisado con Marcelo).

--- 2017-04-11 LICS/syntactic-clo.tex
@@
 Therefore, the desired extension has to be achieved by allowing some local extensions to $\Sigma_1$.
-More precisely, we define $\logex{1}$ as follows.
-An $\fo$-extended quantifier free formula is defined by the grammar
-\[
-\varphi \,:=\, x = y \,\mid\, R(\bar{u}) \,\mid\, X(\bar{v}) \,\mid\, \top \,\mid\, \neg\varphi \,\mid\, (\varphi\vee\varphi) \,\mid\, \psi(\bar{w})
-\]
-where the characterization of the symbols is the same as the one mentioned in the definition of second-order formulae, and $\psi$ is an $\fo$-formula.
-A formula in $\logex{1}$ has the form $\exists\bar{x}\,\varphi(\bar{x})$, where $\varphi$ is an $\fo$-extended quantifier free formula.
+More precisely, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulae over a signature $\R$ to be of the form either (1) $u = v$, or (2) $X(\bar u)$, where $X$ is a second-order variable, or (3) $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, (3) does not mention any second-order variable).
 With this extension we obtain a class with the desired properties.

### 2017-07-21 LMCS/framework.tex
Revisión de sección 2 y 3.

@@
 
-Note that the syntax of QSO formulae is divided in two levels. 
+The syntax of QSO formulae is divided in two levels. 
 The first level is composed by $\so$-formulae over $\R$ (called Boolean formulae) and the second level is made by counting operators of addition and multiplication. 
@@
 
-Let $\R$ be a signature, $\A$ an $\R$-structure with domain $A$, $v$ a first-order assignment for $\A$ and $V$ a second-order assignment for $\A$.
-Then the \emph{evaluation} of a $\qso$-formula $\alpha$ over $(\A, v, V)$ is defined as a function $\sem{\alpha}$ that on input $(\A, v, V)$ returns a number in $\bbN$.
-Formally, the function $\sem{\alpha}$ is recursively defined in Table~\ref{tab-semantics}.
 \begin{table}
@@
 \end{table}
-
+Let $\R$ be a signature, $\A$ an $\R$-structure with domain $A$, $v$ a first-order assignment for $\A$ and $V$ a second-order assignment for $\A$.
+Then the \emph{evaluation} of a $\qso$-formula $\alpha$ over $(\A, v, V)$ is defined as a function $\sem{\alpha}$ that on input $(\A, v, V)$ returns a number in $\bbN$.
+Formally, the function $\sem{\alpha}$ is recursively defined in Table~\ref{tab-semantics}.
 A $\qso$-formula $\alpha$ is said to be a \emph{sentence} if it does not have any free variable, that is, every variable in $\alpha$ is under the scope of a usual quantifier or a quantitative quantifier.
@@
 \] 
-where $\clique(X) := \forall x \forall y ((X(x) \wedge X(y) \wedge x \neq y)  \rightarrow E(x,y))$.
+where $\clique(X) := \fa{x} \fa{y} ((X(x) \wedge X(y) \wedge x \neq y)  \rightarrow E(x,y))$.
 In the Boolean sub-formula of $\alpha_2$ we check whether $X$ is a clique, and with the sum quantifier we add one for each clique in $\fG$. 
@@
 \begin{exa}\label{exa-perm}
-For a more involved example that includes multiplication, let $\bM = \{M(\cdot,\cdot),<\}$ be a vocabulary for storing 0-1 matrices; in particular, a structure $\fM$ over $\bM$ encodes a 0-1 matrix $A$ as follows: if $A[i,j] = 1$, then $M(i,j)$ is true, otherwise $M(i.j)$ is false.
+For an example that includes multiplication, let $\bM = \{M(\cdot,\cdot),<\}$ be a vocabulary for storing 0-1 matrices; in particular, a structure $\fM$ over $\bM$ encodes a 0-1 matrix $A$ as follows: if $A[i,j] = 1$, then $M(i,j)$ is true, otherwise $M(i.j)$ is false.
 Suppose now that we want to compute the permanent of an $n$-by-$n$ 0-1 matrix $A$, that is:
@@
 %~\cite{permanent-applications},
-in fact, showing that this function is hard to compute was one of the main motivations behind the definition of the counting class $\shp$~\cite{Valiant79}.
+in fact, showing that this function is hard to compute was one of the main motivations behind the definition of the class $\shp$~\cite{Valiant79}.
 %and it was one of the first function that was shown to be difficult for counting~\cite{Valiant79} (i.e. $\shp$-complete).
 
-To define the permanent of a 0-1 matrix in $\qso$, assume that for a binary relation symbol $S$, $\op{permut}(S)$ is an $\fo$-formula that is true if and only if $S$ is a permutation, that is, a total bijective function (the definition of $\op{permut}(S)$ is straightforward).
+To define the permanent of a 0-1 matrix in $\qso$, assume that for a binary relation symbol $S$, $\op{permut}(S)$ is an $\fo$-formula that is true if, and only if, $S$ is a permutation, namely, a total bijective function (the definition of $\op{permut}(S)$ is straightforward).
 %\martin{abreviar $\op{permutation}$ a $\op{permut}$ ahorra un par de lineas}
@@
 
-We consider several fragments of $\qso$, which are obtained by restricting the syntax of the Boolean formulae or the use of the quantitative quantifiers.
+We consider several fragments or extensions of $\qso$, which are obtained by restricting the syntax of the Boolean formulae or the use of the quantitative quantifiers.
 In this direction, we denote by $\qfo$ the fragment of $\qso$ where second-order sum and product are not allowed. 
 For instance, for the $\qso$-formulae defined in Example \ref{ex:cliques}, we have that $\alpha_1$ is in $\qfo$ and $\alpha_2$ is not.
-Another interesting fragment of $\qso$ consists of the $\qso$-formulae where only sum operators and quantifiers are allowed. 
+Another interesting fragment of $\qso$ consists of the $\qso$-formulae where only sum operators and sum quantifiers are allowed. 
 Formally, we denote by $\eqso$ the fragment of $\qso$ where first- and second-order products (i.e. $\pa{x}$ and $\pa{X}$) are not allowed.
@@
 
-In the following section, we use different fragments of $\qso$ to capture counting complexity classes.
+In the following section, we use different fragments or extensions of $\qso$ to capture counting complexity classes.
 But before doing this, we show the connection of $\qso$ with previous frameworks for defining functions over relational structures.
@@
 $
-f_{\varphi(\bar x, \bar X)}(\A) = \vert \{(\bar{a},\bar{A}) \mid \A\models\varphi(\bar{a},\bar{A})\}\vert,
+f_{\varphi(\bar x, \bar X)}(\A) = \vert \{(\bar{a},\bar{A}) \mid \A\models\varphi(\bar{a},\bar{A})\}\vert
 $
 for every $\A\in\ostr[\R]$. Then a function $g:\ostr[\R]\to\nat$ is in $\#\LL$ if there exists a formula $\varphi(\bar{x},\bar{X})$ in $\LL$ such that $g = f_{\varphi(\bar x, \bar X)}$.
-In~\cite{SalujaST95}, they were proved several results about capturing counting complexity classes which are relevant for our work.
+In~\cite{SalujaST95}, they proved several results about capturing counting complexity classes which are relevant for our work.
 We discuss and use these results in Sections~\ref{sec:complexity} and~\ref{sec:syntactic}.
@@
 Thus, the advantage of our approach relies on the flexibility to define functions by alternating sum with product operators and, moreover, by introducing new quantitative operators (see Section~\ref{sec:beyond}).
-Furthermore, we show in the following section how to capture some fundamental classes that cannot be captured by following the approach in~\cite{SalujaST95}.
+Furthermore, we show in the next section how to capture some classes that cannot be captured by following the approach in~\cite{SalujaST95}.

### 2017-07-21 LMCS/preliminaries.tex
Revisión de sección 2 y 3.

@@
 A finite structure over $\R$ (or just finite $\R$-structure) is a tuple $\A = \langle A, R_1^\A, \ldots, R_k^\A \rangle$ such that $A$ is a finite set and $R_i^\A \subseteq A^{\arity(R_i)}$ for every $i \in \{1, \ldots, k\}$.
-In this paper we only consider finite structures, so we omit the word finite when referring to them.
-An $\R$-structure $\A$ is said to be ordered if $<$ is a binary predicate name in $\R$ and $<^\A$ is a linear order on $A$.
-Let $\str[\R]$ be the class of all $\R$-structures and $\ostr[\R]$ be the class of all ordered $\R$-structures.
+Further, an $\R$-structure $\A$ is said to be ordered if $<$ is a binary predicate name in $\R$ and $<^\A$ is a linear order on $A$.
+We denote by $\ostr[\R]$ the class of all finite ordered $\R$-structures. 
+In this paper we only consider finite ordered structures, so we will usually omit the word finite or ordered when referring to them.
 
@@
 	(\varphi \vee \varphi) \ \mid\ 
-	\exists x \, \varphi \ \mid\ 
-	\exists X \, \varphi
+	\ex{x} \varphi \ \mid\ 
+	\ex{X} \varphi
  \end{align*}
@@
 We use notation $(\A, v, V) \models \varphi$ to indicate that structure $\A$ satisfies $\varphi$ under $v$ and $V$.
-In particular, we have that $(\A, v, V) \models \top$.
+% In particular, we have that $(\A, v, V) \models \top$.
 %\martin{En vez de decir "for every $y\in \fv$ distinct from $x$" podría ser "for every $y\in\fv\setminus\{x\}$", y lo mismo con $Y\in\sv$.}
 
-In this paper, we consider several fragments of $\so$, in particular first-order logic ($\fo$), least fixed point logic (LFP) and partial fixed point logic (PFP) \cite{L04}.
+In this paper, we consider several fragments or extensions of $\so$ like first-order logic~($\fo$), least fixed point logic (LFP) and partial fixed point logic (PFP) \cite{L04}.
 Moreover, for every $i \in \N$, we consider the fragment $\Sigma_i$ (resp., $\Pi_i$) of $\fo$, which is the set of $\fo$-formulae of the form
@@
 
-We consider the following counting complexity classes in this paper. 
+We consider several counting complexity classes in this paper, some of the are recalled here~(see \cite{F97,hemaspaandra2013complexity}).
 %We define the following function complexity classes: 

### 2017-07-21 LMCS/counting.tex
journal 27 mar
Avances en Sección 4.

@@
 In this section, we show that by syntactically restricting $\qso$ one can capture different counting complexity classes. 
-In other words, by using $\qso$ we can extend the theory of descriptive complexity~\cite{immerman1999descriptive} from decision problems to computation problems. 
+In other words, by using $\qso$ we can extend the theory of descriptive complexity~\cite{immerman1999descriptive} from decision problems to counting problems. 
 For this, we first formalize the notion of \emph{capturing} a complexity class of functions.
@@
 
-Fix a signature $\R = \{R_1, \ldots, R_k\}$ and assume that $\A$ is an ordered (finite) $\R$-structure with a domain $A = \{a_1, \ldots, a_n\}$.
-%$ containing $n$ elements. 
-Recall that  $<$ is a linear order on $A$, say $a_1 < a_2 < \ldots < a_n$.
+Fix a signature $\R = \{R_1, \ldots, R_k\}$ and assume that $\A$ is an ordered $\R$-structure with a domain $A = \{a_1, \ldots, a_n\}$ and $a_1 < a_2 < \ldots < a_n$.
 For every $i \in \{1, \ldots, k\}$, define the encoding of $R_i^\A$, denoted by $\enc(R_i^\A)$, as the following binary string.
@@
 Then let $\enc(R_i^\A)$ be a binary string of length $n^\ell$ such that the $i$-th bit of $\enc(R_i^\A)$ is 1 if the $i$-th tuple in the previous enumeration belongs to $R_i^\A$, and 0 otherwise.
-Moreover, define the encoding of $\A$, denoted by $\enc(\A)$, as the string
+Moreover, define the encoding of $\A$, denoted by $\enc(\A)$, as the string~\cite{L04}:
 %following binary string~\cite{L04}:
@@
 %\end{eqnarray*}
-$0^n \, 1 \, \enc(R_1^\A) \, \cdots \, \enc(R_k^\A)$~\cite{L04}. 
+$$
+0^n \, 1 \, \enc(R_1^\A) \, \cdots \, \enc(R_k^\A).
+$$
 %We define the class of all $\R$-functions, denoted by $\Func(\R)$, as the class of all functions $f: \ostr \rightarrow \bbN$.
@@
 by following the approach proposed in~\cite{SalujaST95}. 
-On the other hand, if we consider our framework and move out from $\eqso$, we have other options for counting like first- and second-order products.
+On the other hand, if we consider our framework and move out from $\eqso$, we have other alternatives for counting like first- and second-order products.
 In fact, the combination of $\qfo$ with $\lfp$ is exactly what we need to capture $\fp$.
-
-From now on, we will use the symbol $<$ also to denote the lexicographic order over same-sized tuples.
-If $\bar{x} = (x_1,\ldots,x_m)$ and $\bar{y} = (y_1,\ldots,y_m)$ are tuples of first-order variables, we denote $\bar{x} < \bar{y}$ for the formula $\bigvee_{i = 1}^m[\bigwedge_{j = 1}^{i-1}x_j = y_j \wedge x_i < y_i]$.
-Similarly, we use $=$ to denote equality between tuples, as $\bar{x} = \bar{y}$ denotes $\bigwedge_{i = 1}^m(x_i = y_i)$, and also $\bar{x}\leq\bar{y}$ denotes $\bar{x} < \bar{y} \vee \bar{x} = \bar{y}$.
-%Furthermore, If $\bar{x} = (x_1,\ldots,x_m)$ ($\bar{X} = (X_1,\ldots,X_m)$) is a tuple of first-order (second-order) variables, we denote $\sa{\bar{x}}\alpha$ for $\sa{x_1}\cdots\sa{x_m}\alpha$ and $\sa{\bar{X}}\alpha$ for $\sa{X_1}\cdots\sa{X_m}\alpha$ for each $\qso$ formula $\alpha$.
-
-The {\em conditional count} symbol $(\varphi \mapsto \alpha)$ is defined as $(\neg\varphi + (\varphi\cdot\alpha))$ for given $\so$ formula $\varphi$ and $\qso$ formula $\alpha$.
-Note that for each $\A \in \ostr[\R]$, and each first-order (second-order) assignment $v$ ($V$) over $\A$,
-\[
-\sem{(\varphi \mapsto \alpha)}(\A,v,V) = 
-\begin{cases}
-\sem{\alpha}(\A,v,V) &\text{if } (\A,v,V)\models\varphi,\\
-1 &\text{otherwise}.
-\end{cases}
-\]
 \begin{thm} \label{theo:capture-fp}

### 2017-07-21 LMCS/proof-capture-fp.tex
Avances en Sección 4.

@@
-For condition (1), notice that for each $\lfp$ formula $\varphi$, checking whether $\A\models\varphi$ can be done in deterministic polynomial time on the size of $\A$\cite{I83}.
-Also, the constant function $s$ can be trivially simulated in $\fp$.
-Furthermore, $\fp$ is closed under polynomial sum and multiplication.
-This suffices to show that the condition holds.
+In the next proof and the followings, we will use the symbol $<$ also to denote the lexicographic order over same-sized tuples.
+Formally, for $\bar{x} = (x_1,\ldots,x_m)$ and $\bar{y} = (y_1,\ldots,y_m)$ we denote by $\bar{x} < \bar{y}$ the formula:
+$$
+\bigvee_{i = 1}^m \, \bigwedge_{j = 1}^{i-1} \, (x_j = y_j \wedge x_i < y_i).
+$$
+Similarly, we use $=$ to denote equality between tuples (e.g. $\bar{x} = \bar{y}$) and $\leq$ to denote $\bar{x} < \bar{y} \vee \bar{x} = \bar{y}$.
+%Furthermore, If $\bar{x} = (x_1,\ldots,x_m)$ ($\bar{X} = (X_1,\ldots,X_m)$) is a tuple of first-order (second-order) variables, we denote $\sa{\bar{x}}\alpha$ for $\sa{x_1}\cdots\sa{x_m}\alpha$ and $\sa{\bar{X}}\alpha$ for $\sa{X_1}\cdots\sa{X_m}\alpha$ for each $\qso$ formula $\alpha$.
+We also use some syntactic sugar in $\qso$ to simplify formulas.
+For example, we use {\em conditional count} symbol $(\varphi \mapsto \alpha)$ defined as $ (\varphi\cdot\alpha) + \neg\varphi$ for any Boolean formula $\varphi$ and any quantitative formula $\alpha$.
+Note that for each $\A \in \ostr[\R]$, and each first-order (second-order) assignment $v$ ($V$) over $\A$:
+\[
+\sem{(\varphi \mapsto \alpha)}(\A,v,V) = 
+\begin{cases}
+\sem{\alpha}(\A,v,V) &\text{if } (\A,v,V)\models\varphi,\\
+1 &\text{otherwise}.
+\end{cases}
+\]
 
-For condition (2), let $f\in \fp$ and consider some signature $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq \size{\A}^\ell$ (i.e. $\size{\A}^\ell$ is an upper bound for the output size).
-Let $\bar{x} = (x_1,\ldots,x_{\ell})$ and $\bar{a} = (a_1,\ldots,a_{\ell})$.
-Define a language
+Now we prove Theorem~\ref{theo:capture-fp}.
+For condition (1), recall that checking whether $\A\models\varphi$ for any $\lfp$ formula $\varphi$ can be done in deterministic polynomial time on the size of $\A$\cite{I83}.
+%Also, the constant function $s$ can be trivially simulated in $\fp$. 
+Furthermore, it is easy to check that $\fp$ is closed under polynomial sum and multiplication. 
+We conclude that any formula in $\qfo(\lfp)$ can be computed in $\fp$.	
+For condition (2), let $\R$ be a signature, $f\in \fp$
+and $\ell\in\nat$ such that $\log_2\left(f(\enc(\A)) \right) \leq \size{\A}^\ell$ for every $\A\in\ostr[\R]$ (i.e. $\size{\A}^\ell$ is an upper bound for the output size of $f$ over $\A$).
+Consider the language:
 \[
-L = \{(\A,a_1,\ldots,a_{\ell})\mid a_1,\ldots,a_{\ell}\in A \text{ and the } (a_1,\ldots,a_{\ell})\text{-th bit of }f(\enc(\A))\text{ is 1}\}.
+L = \{(\A,\bar{a})\mid \bar{a} \in A^l \text{ and the } \bar{a}\text{-th bit of }f(\enc(\A))\text{ is 1}\}.
 \]
+where $\bar{a}$ encodes a number by following the lexicographic order over $A^l$.
 %Consider a procedure that receives $\enc(\A)$ and an assignment $\bar{a}$ to $\bar{x}$.
@@
 Since this language is in $\ptime$, by \cite{I83} there exists a formula $\Phi(\bar{x})$ in $\lfp$ such that $\A\models\Phi(\bar{a})$ if and only if $(\A,\bar{a})\in L$. 
-Then we use
+Then we use the following formula to encode $f$:
 $$
-\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot\varphi(\bar{x}),
+\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2)
 $$
-where $\varphi(\bar{x}) := \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2)$.
-This formula takes the value $2^m$ if there exist $m$ tuples in $A^{\ell}$ that are smaller than $\bar{x}$.
-Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
-In other words, $\Phi(\bar{x})$ simulates the behavior of the $\fp$-machine and the formula $\alpha$ reconstructs the binary output.
-Then, $\alpha$ is in $\qfo(\lfp)$ over $\R$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
+Here, the subformula $\pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2)$ takes the value $2^m$ if there exist $m$ tuples in $A^{\ell}$ that are smaller than $\bar{x}$.
+Summing these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
+In other words, $\Phi(\bar{x})$ simulates the behavior of the $\fp$-machine and the formula $\alpha$ reconstructs the binary output bit by bit.
+Then, $\alpha$ is in $\qfo(\lfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.

### 2017-07-21 LMCS/proof-capture-shP.tex
Avances en Sección 4.

@@
-In this proof, as we will in the following capture theorems, we refer to definition \ref{def:cap}.
-First, we prove condition (2).
-Saluja et al.
+We briefly explain that the two conditions of Definition~\ref{def:cap} are satisfied.
+First, for condition (2) Saluja et al.
 proved that $\shp = \sfo$\cite{SalujaST95}.
-Hence, given that every function in $\sfo$ can be trivially defined as a formula in $\eqso(\fo)$ (see Section~\ref{sec:previous}), the condition holds.
-
-For condition (1), let $\alpha\in\eqso(\fo)$ over some signature $\R$, defined by the grammar in \ref{syntax}.
-Notice that given a $\fo$ formula $\varphi$, checking whether $\A\models\varphi$ can be done in deterministic polynomial time on the size of $\A$.
-Also, the constant function $s$ for some $s\in\nat$ can be trivially simulated in $\shp$.
+Hence, given that every function in $\sfo$ can be trivially defined as a formula in $\eqso(\fo)$ (see Section~\ref{sec:previous}) then condition~(2) holds.
+For condition (1), let $\alpha\in\eqso(\fo)$ over some signature $\R$.
+Given a $\fo$ formula $\varphi$, checking whether $\A\models\varphi$ can be done in deterministic polynomial time on the size of $\A$ and the constant function $s$ can be trivially simulated in $\shp$.
 These facts, together with the closures under exponential sum and polynomial product of $\shp$\cite{F97}, suffice to show that the function represented by $\alpha$ is in $\shp$.
-
 %We construct recursively a $\shp$-machine $M_{\alpha}$ for each $\eqso(\fo)$ formula $\alpha$ over a signature $\R$.

### 2017-07-21 LMCS/proof-capture-spanP.tex
Avances en Sección 4.

@@
-To prove the condition (2), we use the fact that $\spp = \#(\eso)$. The condition holds using the same argument as in proposition \ref{prop:capture-shP}.
-
+To prove the condition (2), we use the fact that $\spp = \#(\eso)$. The condition holds using the same argument as in Proposition~\ref{prop:capture-shP}.
 For condition (1), notice that given an $\eso$ formula $\varphi$, checking whether $\A\models\varphi$ can be done in non-deterministic polynomial time on the size of $\A$\cite{fagin1974generalized}. 
-The $\spp$ machine we construct to evaluate the formula produces at most one different output, and does this by writing the same string as output in each non-deterministic path.
-Furthermore, the constant function $s$ for some $s\in\nat$ can be trivially simulated in $\spp$.
-The condition holds analogously to proposition \ref{prop:capture-shP} since $\spp$ possesses the mentioned closure properties as well\cite{OH93}.
-
+Therefore, a $\spp$ machine for $\varphi$ will simulate the non-deterministic polynomial time machine and produce the same string as output in each accepting non-deterministic run.
+Furthermore, the constant function $s$ for some $s\in\nat$ can be trivially simulated in $\spp$ and, thus, condition (1) holds analogously to Proposition \ref{prop:capture-shP} since $\spp$ is also closed under exponential sum and polynomial product~\cite{OH93}.
 %Similar than the previous proof, we construct recursively a $\spp$ machine $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.

### 2017-07-21 LMCS/proof-capture-fp.tex
Avance sección 4.

@@
 $$
-Similarly, we use $=$ to denote equality between tuples (e.g. $\bar{x} = \bar{y}$) and $\leq$ to denote $\bar{x} < \bar{y} \vee \bar{x} = \bar{y}$.
+Similarly, we use $\bar{x} = \bar{y}$ to denote equality between tuples and $\bar{x} \leq \bar{y}$  to denote $\bar{x} < \bar{y} \vee \bar{x} = \bar{y}$.
 %Furthermore, If $\bar{x} = (x_1,\ldots,x_m)$ ($\bar{X} = (X_1,\ldots,X_m)$) is a tuple of first-order (second-order) variables, we denote $\sa{\bar{x}}\alpha$ for $\sa{x_1}\cdots\sa{x_m}\alpha$ and $\sa{\bar{X}}\alpha$ for $\sa{X_1}\cdots\sa{X_m}\alpha$ for each $\qso$ formula $\alpha$.
-We also use some syntactic sugar in $\qso$ to simplify formulas.
-For example, we use {\em conditional count} symbol $(\varphi \mapsto \alpha)$ defined as $ (\varphi\cdot\alpha) + \neg\varphi$ for any Boolean formula $\varphi$ and any quantitative formula $\alpha$.
+We also use some syntactic sugar in $\qso$ to simplify formulas like the {\em conditional count} symbol $(\varphi \mapsto \alpha)$ defined as $ (\varphi\cdot\alpha) + \neg\varphi$ for any Boolean formula $\varphi$ and any quantitative formula $\alpha$.
 Note that for each $\A \in \ostr[\R]$, and each first-order (second-order) assignment $v$ ($V$) over $\A$:

### 2017-07-24 LMCS/counting.tex
Paso completo a sección 4.

@@
 one also needs more powerful quantitative quantifiers as functions like $2^{2^n}$ can be computed in polynomial space,
-so second-order sum is not enough for the quantitative layer of a logic for $\fpspace$.
+so $\eqso$ is not enough for the quantitative layer of a logic for $\fpspace$.
 In fact, by considering second-order product we obtain the fragment $\qso(\pfp)$ that captures $\fpspace$. 

### 2017-07-24 LMCS/extentions.tex
Paso completo a sección 4.

@@
 a negative number) or the class is not defined purely in terms of arithmetical operations (e.g.
-min and max are used).
+min and max).
 To remedy this problem, we show here how $\qso$ can be easily extended  to capture such classes that go beyond sum and product over natural numbers.
@@
 \end{cor}
-This is an interesting result that shows how robust is $\qso$ when capturing different complexity classes.
+This is an interesting result that shows how robust and versatile is $\qso$ for capturing different counting complexity classes even beyond $\bbN$.
 
@@
 For instance, MAX-SAT is in $\optp$ as many other optimization versions of $\np$-problems.
-Given that in~\cite{krentel1988complexity} the author does not make the distinction between $\max$ and $\min$, in~\cite{vollmer1995complexity} the authors define the classes $\maxp$ and $\minp$ as the max and min version of the problems in $\optp$ (i.e.
+Given that in~\cite{krentel1988complexity} Krentel did not make the distinction between $\max$ and $\min$, in~\cite{vollmer1995complexity} they defined the classes $\maxp$ and $\minp$ as the max and min version of the problems in $\optp$ (i.e.
 $\optp = \maxp \cup \minp$).
@@
 The semantics of the $\qso$-operators in $\optqso$ are defined as usual.
-Furthermore, the semantics of the max and min quantifiers are defined as the natural extension of the sum quantifiers in $\qso$ (see Table~\ref{tab-semantics}) by maximizing or minimizing, respectively, instead of computing a sum.
+Furthermore, the semantics of the max and min quantifiers are defined as the natural extension of the sum quantifiers in $\qso$ (see Table~\ref{tab-semantics}) by maximizing or minimizing, respectively, instead of computing a sum or a product.
 \begin{exa}\label{ex:optqso}
@@
 \qed
-It is important to mention that a similar result was proved in~\cite{kolaitis1994logical} for the class $\maxpb$ (resp., $\minpb$) of problems in $\maxp$ (resp., $\minp$) whose output value is polynomially bounded.
+It is important to mention that a similar result, following the framework of~\cite{SalujaST95}, was proved in~\cite{kolaitis1994logical} for the class $\maxpb$ (resp., $\minpb$) of problems in $\maxp$ (resp., $\minp$) whose output value is polynomially bounded.
 Interestingly, Theorem \ref{theo:capture-optp} is stronger since our logic has the freedom to use sum and product quantifiers, instead of using a max-and-count problem over Boolean formulae. 

### 2017-07-24 LMCS/proof-capture-fp.tex
Paso completo a sección 4.

@@
-In the next proof and the followings, we will use the symbol $<$ also to denote the lexicographic order over same-sized tuples.
+In this proof and the followings, we will reuse the symbol $<$ to denote the lexicographic order over same-sized tuples.
 Formally, for $\bar{x} = (x_1,\ldots,x_m)$ and $\bar{y} = (y_1,\ldots,y_m)$ we denote by $\bar{x} < \bar{y}$ the formula:
@@
 %Furthermore, If $\bar{x} = (x_1,\ldots,x_m)$ ($\bar{X} = (X_1,\ldots,X_m)$) is a tuple of first-order (second-order) variables, we denote $\sa{\bar{x}}\alpha$ for $\sa{x_1}\cdots\sa{x_m}\alpha$ and $\sa{\bar{X}}\alpha$ for $\sa{X_1}\cdots\sa{X_m}\alpha$ for each $\qso$ formula $\alpha$.
-We also use some syntactic sugar in $\qso$ to simplify formulas like the {\em conditional count} symbol $(\varphi \mapsto \alpha)$ defined as $ (\varphi\cdot\alpha) + \neg\varphi$ for any Boolean formula $\varphi$ and any quantitative formula $\alpha$.
+We will also use some syntactic sugar in $\qso$ to simplify formulas.
+Specifically, we will use the {\em conditional count} symbol $(\varphi \mapsto \alpha)$ defined as $ (\varphi\cdot\alpha) + \neg\varphi$ for any Boolean formula $\varphi$ and any quantitative formula $\alpha$.
 Note that for each $\A \in \ostr[\R]$, and each first-order (second-order) assignment $v$ ($V$) over $\A$:
@@
 Furthermore, it is easy to check that $\fp$ is closed under polynomial sum and multiplication. 
-We conclude that any formula in $\qfo(\lfp)$ can be computed in $\fp$.	
-For condition (2), let $\R$ be a signature, $f\in \fp$
+We conclude then that any formula in $\qfo(\lfp)$ can be computed in $\fp$.	
+For condition~(2), let $\R$ be a signature, $f\in \fp$
 and $\ell\in\nat$ such that $\log_2\left(f(\enc(\A)) \right) \leq \size{\A}^\ell$ for every $\A\in\ostr[\R]$ (i.e. $\size{\A}^\ell$ is an upper bound for the output size of $f$ over $\A$).
@@
 The procedure then computes the $m$-th bit of $f(\enc(\A))$, from least to most significant.
-Since this language is in $\ptime$, by \cite{I83} there exists a formula $\Phi(\bar{x})$ in $\lfp$ such that $\A\models\Phi(\bar{a})$ if and only if $(\A,\bar{a})\in L$. 
-Then we use the following formula to encode $f$:
+Clearly, the language $L$ is in $\ptime$ and by \cite{I83} there exists a formula $\Phi(\bar{x})$ in $\lfp$ such that $\A\models\Phi(\bar{a})$ if, and only if, $(\A,\bar{a})\in L$. 
+We use then the following formula to encode $f$:
 $$
-\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot \pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2)
-$$
-Here, the subformula $\pa{\bar{y}}(\bar{y} < \bar{x} \mapsto 2)$ takes the value $2^m$ if there exist $m$ tuples in $A^{\ell}$ that are smaller than $\bar{x}$.
+\alpha = \sa{\bar{x}} \Phi(\bar{x})\cdot \pa{\bar{y}}(\bar{y} < \bar{x}) \mapsto 2)$$
+Note that the subformula $\pa{\bar{y}}(\bar{y} < \bar{x}) \mapsto 2$ takes the value $2^m$ if there exist $m$ tuples in $A^{\ell}$ that are smaller than $\bar{x}$.
 Summing these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
 In other words, $\Phi(\bar{x})$ simulates the behavior of the $\fp$-machine and the formula $\alpha$ reconstructs the binary output bit by bit.
-Then, $\alpha$ is in $\qfo(\lfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
+Then $\alpha$ is in $\qfo(\lfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.

### 2017-07-24 LMCS/proof-capture-optp.tex
Paso completo a sección 4.

@@
-In this proof we construct recursively a non-deterministic polynomial time Turing machine $M_{\alpha}$ with output tape for each $\maxqso(\fo)$ formula $\alpha$ over a signature $\R$.
-This machine, on input $(\A,v,V)$, non-deterministically produces $\sem{\alpha}(\A,v,V)$ (in binary) over the output tape of some run, and this value is the maximum value over all runs.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\fo$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ in deterministic polynomial time, and output $1$ if, and only if, $(\A,v,V)$ satisfies $\varphi$.
-If $\alpha$ is a constant $s$, it outputs $s$ in binary over the output tape.
-If $\alpha = (\beta \add \gamma)$, then it simulates $M_{\beta}$ and $M_{\gamma}$ on input $(\A,v,V)$, adds the output of both machines and prints this value over the output tape.
-If $\alpha = (\beta \mult \gamma)$ or $\alpha = \max\{\beta,\gamma\}$, then it does the same than the previous case but it multiplies or maximizes, respectively, the outputs of both machines instead of adding.
-If $\alpha = \sa{x}\beta$, it iterates over all elements $a\in A$ simulating and adding the output of $M_{\beta}$ on input $(\A,v[a/x],V)$.
-$M_{\alpha}$ finally outputs the aggregated value in the output tape.
-If $\alpha = \pa{x}\beta$ or $\alpha = \maxa{x}\beta$, then $M_{\alpha}$ does the same idea than the previous case with the difference that the output of $M_{\beta}$ on input $(\A,v[a/x],V)$ is multiplied or maximized, respectively.
-If $\alpha = \maxa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-This covers all possible cases for $\alpha$.
-Furthermore, it is straightforward to prove that each of these steps are correct and can be computed with a non-deterministic polynomial time Turing machine with output tape.
-Let $\alpha$ be a formula in $\maxqso$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the maximum run (with respect to the output value) of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\maxp$ function over $\R$ and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-Finally, one can easily see that the same construction holds with $\minqso(\fo)$ by constructing a Turing machine that take the min over all runs instead of max.
-
+It is straightforward to prove that $\maxp$ can compute any $\fo$-formula, is closed under first-order sum and product, and second-order maximization. 
+Therefore, condition (1) in Definition~\ref{def:cap} follows similar than in the previous characterizations. Furthermore, one can easily see that the same holds with $\minqso(\fo)$.
 The proof for the other direction is similar than in \cite{kolaitis1994logical} extended with the ideas of Theorem~\ref{theo:capture-fp}.
 Let $f\in \maxp$ be a function defined over some signature $\R$ and
-$\ell\in\nat$ such that $\lceil\log_2 f(\enc(\A)) \rceil \leq n^\ell$ for each $\A\in\ostr[\R]$ of size $n$.
-For $U \subseteq A^{\ell}$, we can interpret the encoding of $U$ ($\enc(U)$) as the binary encoding of a number with $n^l$-bits. We denote this value by $\val(\enc(U))$.
+$\ell\in\nat$ such that $\lceil\log_2 f(\enc(\A)) \rceil \leq |\A|^\ell$ for each $\A\in\ostr[\R]$.
+For $U \subseteq A^{\ell}$, we can interpret the encoding of $U$ ($\enc(U)$) as the binary encoding of a number with $|\A|^\ell$-bits. We denote this value by $\val(\enc(U))$.
 Then, given $\A\in\ostr[\R]$ and $U \subseteq A^{\ell}$, consider the problem of checking whether $f(\enc(\A)) \geq \val(\enc(U))$. 
-Clearly, this is an $\np$-problem and, by Fagin's theorem, there exists a formula of the form $\ex{\bar{X}} \Phi(\bar{X}, Y)$ with $\Phi(\bar{X}, Y)$ in $\fo$-logic and $\arity(Y) = \ell$ such that $f(\enc(\A)) \geq \val(\enc(U))$ if, and only if, $(\A,v,V) \models \ex{\bar{X}} \Phi(\bar{X}, Y)$ with $V(Y) = U$.
-Then we can describe the function $f$ by the following $\maxqso$ formula:
+Clearly, this is an $\np$-problem and, by Fagin's theorem, there exists a formula of the form $\ex{\bar{X}} \Phi(\bar{X}, Y)$ with $\Phi(\bar{X}, Y)$ in $\fo$ and $\arity(Y) = \ell$ such that $f(\enc(\A)) \geq \val(\enc(U))$ if, and only if, $(\A,v,V) \models \ex{\bar{X}} \Phi(\bar{X}, Y)$ with $V(Y) = U$.
+Then we can describe $f$ by the following $\maxqso$ formula:
 $$
@@
 $$
-In this case, if the formula $\Phi(\bar{X}, Y)$ is false, then the output produced by the subformula inside the $\min$-quantifiers will be the biggest possible value (i.e. $2^{n^\ell}$).
+In this case, if the formula $\Phi(\bar{X}, Y)$ is false, then the output produced by the subformula inside the $\min$-quantifiers will be the biggest possible value (i.e. $2^{{|\A|}^\ell}$).
 On the other hand, if $\Phi(\bar{X}, Y)$ holds, the subformula will produce $\val(\enc(U))$. 

### 2017-07-27 LMCS/proof-prop-rest.tex
journam 7 mar
Pasada por sección 5.1.

@@
 We give this proof in three parts.
-
-\vspace{1em}
-First, we show that $\QE{0} \not\subseteq \E{1}$.
-By contradiction, suppose that there is a $\QE{0}$ formula $\alpha$ over some signature $\R$ such that defines the following function.
-For every finite $\R$-structure with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n - 1$.
-We use the following claim.
-\begin{claim}
+First, we show that $\E{1} \not\subseteq \QE{0}$.
+By contradiction, let $\R$ be the empty signature and suppose that there is a $\QE{0}$ formula $\alpha$ over $\R$ such that defines the following function: for every finite $\R$-structure with $n$ elements, $\alpha(\enc(\A)) = n - 1$.
+We need the following claim.
+\begin{clm}
 	Let $\alpha = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. Then the function defined by $\alpha$ is either null, greater or equal to $n$, or is in $\Omega(n^2)$.
-\end{claim}
-\begin{proof}
-	Suppose that the function defined by $\alpha$ is not $0$ and that $\varphi$ is in DNF. Furthermore, suppose $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$.
-	Then $\alpha = \sa{\bar{x}} \varphi_1(\bar{x}) \vee \cdots \vee \varphi_n(\bar{x})$.
-Since $\alpha$ is not null, then some $\varphi_i$ must be satisfiable.
-This is, the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is not null.
-	We will prove by induction on $\length{\bar{x}}$ that the function defined by $\sa{\bar{x}}\varphi(\bar{x})$ is either greater or equal to $n$, or in $\Omega(n^2)$.
-	We address the case $\length{\bar{x}}= 1$, then $\alpha = \sa{x}\bigwedge\psi(x)$.
-If any $\psi(x) = (x = x)$ or $\neg(x < x)$, then we can eliminate it and we obtain the same function.
+\end{clm}
+\proof
+	Suppose that the function defined by $\alpha$ is not $0$ and that $\varphi$ is in DNF, namely, $\alpha = \sa{\bar{x}} \varphi_1(\bar{x}) \vee \cdots \vee \varphi_n(\bar{x})$.
+Furthermore, suppose $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$.
+Since $\alpha$ is not null, then some $\varphi_i$ must be satisfiable and suppose that $\varphi_i(\bar{x}) = \bigwedge\psi(\bar{x})$ where each $\psi(\bar{x})$ is a literal (i.e.
+an atomic formula or its negation).
+	We will prove by induction on $\length{\bar{x}}$ that the function defined by $\sa{\bar{x}} \bigwedge\psi(\bar{x})$ is either greater or equal to $n$, or in $\Omega(n^2)$.
+	For $\length{\bar{x}}= 1$ and $\alpha = \sa{x}\bigwedge\psi(x)$, if any $\psi(x) = (x = x)$ or $\neg(x < x)$, then we can eliminate it and we obtain the same function.
 If any $\psi = (x < x)$ or $\neg(x=x)$, then the function becomes null.
-If $\psi(x) = R(x,\ldots,x)$ for some $R\in\R$ the function becomes null for the structures we are considering.
-If $\psi(x) = \neg R(x,\ldots,x)$, we can eliminate it and for the structures we are considering we obtain the same function.
 The only possible $\alpha$ left is $\alpha = \sa{x}\top$ which is equal to the function $n$.
@@
 If any $\psi(\bar{x}) = (x_i = x_j)$ where $i \neq j$, then $\alpha$ describes the same function as $\alpha$ where $x_j$ has been replaced by $x_i$.
-In this formula the tuple of first-order variables has $k$ elements so the function it describes if one of the mentioned in the hypothesis.
-If $i = j$, then we can eliminate it and obtain the same function.
-If any $\psi(\bar{x}) = R(\bar{v})$ or $\neg R(\bar{v})$ where $\bar{v}$ is a sub-tuple of $\bar{x}$ then we can either eliminate it or the function becomes null, following the same argument as in the case $\length{\bar{x}} = 1$.
-If any $\psi(\bar{x}) = \neg(x_i = x_j)$ or $(x_i < x_j)$ where $i = j$, then the function becomes null.
-If any $\psi(\bar{x}) = \neg(x_i < x_j)$ where $i = j$, we can eliminate it.
-The remaining formulas in $\bigwedge\psi(\bar{x})$ are either $\neg(x_i = x_j)$, $(x_i<x_j)$ or $\neg(x_i<x_j)$.
-If the formula violates transitivity in $<$ (for example, $x < y \wedge y < z \wedge z < x$), then the function $\alpha$ describes is null.
-Therefore, there is some order over $\bar{x}$ that satisfies $\bigwedge\psi(\bar{x})$.
-Consider the formula that describes this order (like $x_1 < x_3 \wedge x_3 < x_4 \wedge x_4 < x_2$).
-The function $\alpha$ describes is greater or equal to the one this formula describes, which is exactly $\binom{n}{\length{\bar{x}}}$ which is in $\Omega(n^{\length{\bar{x}}}) \subseteq \Omega(n^2)$ if $\length{\bar{x}} > 1$.
-This concludes the proof of the claim.
-\end{proof}
-We suppose that $\alpha$ is in SNF, this is, $\alpha = \sum_{i = 1}^n\alpha_i$.
+In this formula the tuple of first-order variables has now $k$ elements so, by induction, the function satisfies our conclusions.
+Furthermore, if any $\psi(\bar{x}) = (x_i = x_j)$ or $\neg (x_i < x_j)$ where $i = j$, then we can eliminate $\psi(\bar{x})$ and obtain the same function, and if any $\psi(\bar{x}) = \neg(x_i = x_j)$ or $(x_i < x_j)$ where $i = j$, then the function becomes null.
+	The remaining formulas in $\bigwedge\psi(\bar{x})$ are either $\neg(x_i = x_j)$, $(x_i<x_j)$ or $\neg(x_i<x_j)$ with $i \neq j$.
+If the formula violates transitivity in $<$ (e.g.
+$x < y \wedge y < z \wedge z < x$), then $\alpha$ defines the null function.
+Therefore, there must be some order over $\bar{x}$ that satisfies $\bigwedge\psi(\bar{x})$ and this order can be chosen in  $\binom{n}{\length{\bar{x}}}$ different ways over the domain of $\A$.
+This implies that $\sem{\alpha} \in \Omega(n^{\length{\bar{x}}}) \subseteq \Omega(n^2)$ if $\length{\bar{x}} > 1$ and concludes the proof of the claim.
+\qed
+Coming back on proving that $\alpha$ does not define the $(n-1)$-function, suppose that $\alpha$ is in SNF, namely, $\alpha = \sum_{i = 1}^k \alpha_i$ for some fix $k$.
 Since $\alpha$ is not null, consider some $\alpha_i$ that describes a non-null function.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, where $\varphi$ is quantifier-free.
-Note that if $\length{\bar{X}} > 0$, then the function $\alpha$ describes is in $\Omega(2^n)$, as it was proven by the authors in \cite{SalujaST95}.
-We have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$, as we proved in the claim, describes either some function greater or equal to $n$, or in $\Omega(n^2)$, which leads to a contradiction.
-Lastly, note that the formula $\sa{x}\exists y(x < y)$ is in $\E{0}$ and describes the function $n-1$, which concludes the proof.
+Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ where $\varphi$ is quantifier-free.
+Note that if $\length{\bar{X}} > 0$, then the function $\sem{\alpha}$ is in $\Omega(2^n)$, as it was proven in~\cite{SalujaST95}.
+Therefore, we have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$ and, as we proved in the above claim, it describes either some function greater or equal to $n$, or in $\Omega(n^2)$, which leads to a contradiction.
+To conclude, note that the formula $\sa{x} \ex{y} (x < y)$ is in $\E{1}$ and describes the function $n-1$.
 
-\vspace{1em}
-Now we show that $\E{1}\not\subseteq\QE{0}$.
+Now we show that $\QE{0} \not\subseteq \E{1}$.
 In Theorem \ref{theo-pi1-pnf} we proved that there is no formula in $\loge{1}$-PNF equivalent to the formula $\alpha = 2$.
@@
 
-\vspace{1em}
-Lastly, we prove that $\eqso(\loge{1})\subsetneq\eqso(\logu{1})$.
+Finally, we prove that $\eqso(\loge{1})\subsetneq\eqso(\logu{1})$.
 For inclusion, let $\alpha$ be a formula in $\eqso(\loge{1})$.
-Suppose that it is in $\loge{1}$-SNF.
-This is, $\alpha = c + \sum_{i = 1}^{n}\alpha_i$.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$, where $\varphi_i$ is quantifier-free, for each $\alpha_i$.
-We use the same construction used in \cite{SalujaST95}, and we obtain that the formula $\exists\bar{y}\,\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \forall\bar{y}'(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')$ for every assignment to $(\bar{X},\bar{x})$.
+Suppose that it is in $\loge{1}$-SNF, namely, $\alpha = c + \sum_{i = 1}^{n}\alpha_i$.
+Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\ex{\bar{y}}\varphi_i(\bar{X},\bar{x},\bar{y})$, where $\varphi_i$ is quantifier-free for each $\alpha_i$.
+We use the same construction used in \cite{SalujaST95}, and we obtain that the formula $\ex{\bar{y}}\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \fa{\bar{y}'}(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')$ for every assignment to $(\bar{X},\bar{x})$.
 We do this replacement for each $\alpha_i$ and we obtain an equivalent formula in $\eqso(\logu{1})$.
 
-To prove that the inclusion is proper, consider the $\eqso(\logu{1})$ formula $\sa{x}\forall y(y = x)$. This formula defines the following function that takes an ordered structure $\A$ as input:
+To prove that the inclusion is proper, consider the $\eqso(\logu{1})$ formula $\sa{x} \fa{y}(y = x)$. This formula defines the following function over each ordered structure $\A$:
 $$
@@
 Suppose that there exists an equivalent formula $\alpha$ in $\eqso(\loge{1})$.
-Also, suppose that it is in $\L$-PNF, so $\alpha = c + \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\varphi_i(\bar{X},\bar{x},\bar{y})$.
+Also, suppose that it is in $\L$-PNF, so $\alpha = c + \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\ex{\bar{y}}\varphi_i(\bar{X},\bar{x},\bar{y})$.
 Since $\alpha$ takes the value 0 for some structures, $c$ must be 0.
-Consider a structure $\mathfrak{1}$ with one element.
-We have that for some $i$, there exists an assignment $(\bar{B},\bar{b},\bar{a})$ for $(\bar{X},\bar{x},\bar{y})$ such that $\mathfrak{1}\models\varphi_i(\bar{B},\bar{b},\bar{a})$.
-Consider now the structure $\mathfrak{2}$ that is obtained by duplicating $\mathfrak{1}$, as we did for Theorem \ref{theo-pi1-pnf}.
-Note that $\mathfrak{2}\models\varphi_i(\bar{B},\bar{b},\bar{a})$, which implies that $\sem{\alpha}(\mathfrak{2}) \geq 1$, which leads to a contradiction.
+Consider a structure $\A'$ with one element.
+We have that for some $i$, there exists an assignment $(\bar{B},\bar{b},\bar{a})$ for $(\bar{X},\bar{x},\bar{y})$ such that $\A' \models\varphi_i(\bar{B},\bar{b},\bar{a})$.
+Consider now the structure $\A''$ that is obtained by duplicating $\A'$, as we did for Theorem \ref{theo-pi1-pnf}.
+Note that $\A''\models\varphi_i(\bar{B},\bar{b},\bar{a})$, which implies that $\sem{\alpha}(\A' \uplus \A'') > 1$, which leads to a contradiction.

### 2017-07-27 LMCS/proof-theo-pi1-pnf.tex
Pasada por sección 5.1.

@@
 
+From now on, for every first-order tuple $\bar{x}$ or second-order tuple $\bar{X}$ we write $\length{\bar{x}}$ or $\length{\bar{X}}$ as the size of $\bar{x}$ or $\bar{X}$ respectively.
 We divide the proof in three parts.
 
-\vspace{1em}
 First, we prove that the formula $\alpha_{0} = \left( \sa{X} 1 \right) + 1$ with $\arity(X) = 1$ (i.e.
@@
 
-\vspace{1em}
 Second, we prove that the formula $\alpha_{1} = 2$ (i.e.
 the constant function $2$) is not equivalent to any formula in $\loge{1}$-PNF.
-Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to the $\eqso(\loge{1})$ formula $2$.
+Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to $\alpha_1$.
 First, if $\length{\bar{X}} = \length{\bar{x}} = 0$, then the function defined by $\alpha$ is never greater than 1. 
 Therefore, suppose that $\length{\bar{X}} > 0$ or $\length{\bar{x}} > 0$, and consider some ordered structure $\A$. 
-Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
-Now consider the ordered structure $\A'$ that is obtained by concatenating $\A$ twice.
+Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
+Now consider the ordered structure $\A'$ that is obtained by taking the disjoint union of $\A$ twice.
 Indeed, each half of $\A'$ is isomorphic to $\A$.
 Note that $\A'\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$ for $i = 1,2$ and there exists a third assignment $(\bar{B}_1',\bar{b}_1',\bar{a}_1')$ that is isomorphic to $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, in the other half of the structure, such that $\A'\models\varphi(\bar{B}_1',\bar{b}_1',\bar{a}_1')$.
-We have that $\sem{\alpha}(\A) \geq 3$ which leads to a contradiction.
+We have that $\sem{\alpha}(\A') \geq 3$ which leads to a contradiction.
 
-\vspace{1em}
-We now show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there is an equivalent formula $\beta$ in $\LL$-PNF.
-As in Theorem \ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such formula.
-Assume that $\alpha$ is in $\LL$-SNF.
-Let $\alpha = \sum_{i = 1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-PNF.
+For the last part of the proof, we show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there exists an equivalent formula in $\LL$-PNF.
+Similar than in Theorem~\ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such a formula.
+Assume that $\alpha = \sum_{i = 1}^n \alpha_i$ is in $\LL$-SNF where each $\alpha_i$ is in $\LL$-PNF.
 Without loss of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
-If not, we replace each $\alpha_i$ by the equivalent formula $\sa{\bar{X}} \sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge\forall z\,Y(z) \wedge \forall z(y \leq z))$.
-
+If not, we can replace each $\alpha_i$ by the equivalent formula:
+$$
+\sa{\bar{X}} \sa{Y}\sa{\bar{x}}\sa{y}(\varphi_i(\bar{X},\bar{x})\wedge \fa{z} Y(z) \wedge \fa{z} z \leq y).
+$$
 Now we begin describing the function $\tau$.
@@
 \begin{align*}
-\gamma_{\text{first}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z}), \\
-\gamma_{\text{last}}(\bar{X},\bar{x}) &= \bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x}).
+\gamma_{\text{first}}(\bar{X},\bar{x}) & \; = \;  \bigwedge_{i = 1}^{\length{\bar{X}}} \fa{\bar{z}}\neg X_i(\bar{z}) \wedge \fa{\bar{z}}(\bar{x}\leq\bar{z}), \\
+\gamma_{\text{last}}(\bar{X},\bar{x}) & \; = \;  \bigwedge_{i = 1}^{\length{\bar{X}}} \fa{\bar{z}} X_i(\bar{z}) \wedge \fa{\bar{z}}(\bar{z}\leq\bar{x}).
 \end{align*}
-Similarly, we define the formulas $\gamma_{\text{first}}(\bar{Y},\bar{y})$ and $\gamma_{\text{last}}(\bar{Y},\bar{y})$ (for the sake of simplicity we reuse the names $\gamma_{\text{first}}$ and $\gamma_{\text{last}}$).
+Similarly, we can define the formulas $\gamma_{\text{first}}(\bar{Y},\bar{y})$ and $\gamma_{\text{last}}(\bar{Y},\bar{y})$.
 In other words, the ``first'' assignment is the one where every second-order predicate is empty and the first-order assignment is the lexicographically smallest, and the ``last'' assignment is the one where every second-order predicate is full and the first-order assignment is the lexicographically greatest.
@@
 \begin{align*}
-\gamma_{\text{not first}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{z}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{z}_i)), \\
-\gamma_{\text{not last}}(\bar{X},\bar{x}) &= \exists\bar{z}(\bar{x} < \bar{z}_0 \vee \bigvee_{i = 1}^{\length{\bar{X}}}\neg X(\bar{z}_i)),
+\gamma_{\text{not-first}}(\bar{X},\bar{x}) & \; = \; \ex{\bar{z}}(\bar{z}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{z}_i)), \\
+\gamma_{\text{not-last}}(\bar{X},\bar{x}) & \; = \; \ex{\bar{z}}(\bar{x} < \bar{z}_0 \vee \bigvee_{i = 1}^{\length{\bar{X}}}\neg X(\bar{z}_i)),
 \end{align*}
-where $\bar{z} = (\bar{z}_0,\bar{z}_1,\ldots,\bar{z}_{\length{\bar{X}}})$. The following formula is equivalent to $\alpha$:
+where $\bar{z} = (\bar{z}_0,\bar{z}_1,\ldots,\bar{z}_{\length{\bar{X}}})$. Then the following formula is equivalent to $\alpha$:
 \begin{align}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{not first}}(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{Y},\bar{y}))\vee \label{eq:partition1} \\
+\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{not-first}}(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{Y},\bar{y}))\vee \label{eq:partition1} \\
 &(\varphi(\bar{X},\bar{x})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))\vee \label{eq:partition2}\\
-&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{not last}}(\bar{Y},\bar{y}))\vee \label{eq:partition3}\\
+&(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{first}}(\bar{X},\bar{x})\wedge\gamma_{\text{not-last}}(\bar{Y},\bar{y}))\vee \label{eq:partition3}\\
 &(\psi(\bar{Y},\bar{y})\wedge\gamma_{\text{last}}(\bar{X},\bar{x})\wedge\gamma_{\text{last}}(\bar{Y},\bar{y}))]. \label{eq:partition4}
 \end{align}
+
 To show that the formula is indeed equivalent to $\alpha$, note that the formulas in lines (\ref{eq:partition1}) and (\ref{eq:partition2}) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$, and the formulas in lines (\ref{eq:partition3}) and (\ref{eq:partition4}) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
@@
 With this, it is straightforward to show that the above formula is equivalent to $\alpha$.
-However, the formula is not yet in the correct form since it has existential quantifiers in the sub-formulas $\gamma_{\text{not first}}$ and $\gamma_{\text{not last}}$.
-To solve this, first let us take a close look to the complete formula:
-\begin{align*}
-\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\exists\bar{v}(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z}))\vee\\
-&(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\exists\bar{w}(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\vee\\
-&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))].
-\end{align*}
-To construct an equivalent formula that is in the correct form, we define $\bar{u} = (\bar{v},\bar{w})$ and we replace the first-order quantifiers by a first-sum and count the first assignment to $\bar{v}$ and $\bar{w}$ that satisfies the formula.
+However, the formula is not yet in the correct form since it has existential quantifiers in the sub-formulas $\gamma_{\text{not-first}}$ and $\gamma_{\text{not-last}}$.
+To solve this, we can replace each existential quantifier by a first order sum that counts just the first assignment that satisfies the inner formula and this can be defined in $\logu{1}$.
 A similar construction was used in \cite{SalujaST95}.
-Then the final formula equivalent to $\alpha$ is the following:
-\begin{align*}
-\sa{\bar{X}}&\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}\big[ \\
-&[\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
-&\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})]\vee\\
-&
-[\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}')]\vee
-\\
-&[\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
-&(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}']\vee \\
-&[\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}')]\big].
-\end{align*}
 
@@
 If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we can consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the hypothesis.
+%To show that the formula is indeed equivalent to $\alpha$, note that the formulas in lines (\ref{eq:partition1}) and (\ref{eq:partition2}) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$, and the formulas in lines (\ref{eq:partition3}) and (\ref{eq:partition4}) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
+Altogether, the four lines define pairwise disjoint assignments for $(\bar{X},\bar{x}),(\bar{Y},\bar{y})$.
+With this, it is straightforward to show that the above formula is equivalent to $\alpha$.
+However, the formula is not yet in the correct form since it has existential quantifiers in the sub-formulas $\gamma_{\text{not-first}}$ and $\gamma_{\text{not-last}}$.
+To solve this, first let us take a close look to the complete formula:
+%\begin{align*}
+%\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{Y}}\sa{\bar{y}}[&(\varphi(\bar{X},\bar{x})\wedge\exists\bar{v}(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z}))\vee\\
+%&(\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))\vee\\
+%&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\exists\bar{w}(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\vee\\
+%&(\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y}))].
+%\end{align*}
+%To construct an equivalent formula that is in the correct form, we define $\bar{u} = (\bar{v},\bar{w})$ and we replace the first-order quantifiers by a first-sum and count the first assignment to $\bar{v}$ and $\bar{w}$ that satisfies the formula.
+A similar construction was used in \cite{SalujaST95}.
+Then the final formula equivalent to $\alpha$ is the following:
+%\begin{align*}
+%\sa{\bar{X}}&\sa{\bar{Y}}\sa{\bar{x}}\sa{\bar{y}}\sa{\bar{u}}\big[ \\
+%&[\varphi(\bar{X},\bar{x})\wedge(\bar{v}_0 < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i))\wedge \forall\bar{u}'((\bar{v}_0' < \bar{x} \vee \bigvee_{i = 1}^{\length{\bar{X}}}X(\bar{v}_i'))\to\bar{u}\leq\bar{u}') \wedge \\
+%&\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z}\neg Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{y}\leq\bar{z})]\vee\\
+%&
+%[\varphi(\bar{X},\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}')]\vee
+%\\
+%&[\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z}\neg X_i(\bar{z}) \wedge \forall\bar{z}(\bar{x}\leq\bar{z})\wedge \\
+%&(\bar{y} < \bar{w}_0 \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i))\wedge\forall\bar{u}'(\bar{y} < \bar{w}_0' \vee \bigvee_{i = 1}^{\length{\bar{Y}}}\neg Y(\bar{w}_i'))\to\bar{u}\leq\bar{u}']\vee \\
+%&[\psi(\bar{Y},\bar{y})\wedge\bigwedge_{i = 1}^{\length{\bar{X}}} \forall\bar{z} X_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{x})\wedge\bigwedge_{i = 1}^{\length{\bar{Y}}} \forall\bar{z} Y_i(\bar{z}) \wedge \forall\bar{z}(\bar{z}\leq\bar{y})\wedge\forall\bar{u}'(\bar{u}\leq\bar{u}')]\big].
+%\end{align*}
+%
+%Finally, consider a $\eqso(\LL)$ formula $\alpha$ in $\LL$-SNF.
+If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we can consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the hypothesis.

### 2017-07-27 LMCS/proof-theo-pnf-snf.tex
Pasada por sección 5.1.

@@
-Recall that a formula in $\eqso(\LL)$ is written on the following grammar:
+Recall that a formula in $\eqso(\LL)$ is defined by the following grammar:
 \[
-\alpha = \varphi \ \mid \ s \ \mid \ (\alpha + \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \sa{X} \alpha,
+\alpha = \varphi \ \mid \ s \ \mid \ (\alpha + \alpha) \ \mid \ \sa{x} \alpha \ \mid \ \sa{X} \alpha
 \]
 where $\varphi$ is a formula in $\LL$ and $s\in\nat$. 
-We will construct a recursive function $\tau$ such that for every $\eqso(\LL)$ formula $\alpha$, it outputs an equivalent formula $\tau(\alpha)$ which is in $\LL$-SNF.
-If $\alpha = \varphi$, let $\tau(\alpha) = \alpha$.
-If $\alpha = s$, let $\tau(\alpha) = (\top \add \cdots \add \top)$ ($s$ times).
-We assume that for every sub-formula $\beta$ in $\alpha$, $\tau(\beta)$ is an equivalent formula in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$, let $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$.
-If $\alpha = \sa{x}\beta$, then $\tau(\beta) = \sum_{i = 1}^{k}\beta_i$ for some $k$ where each $\beta_i$ is in $\LL$-PNF.
-We define $\tau(\alpha) = \sum_{i = 1}^{k}\sa{x}\beta_i$.
-If $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case.
+To find an equivalent formula in $\LL$-SNF for every $\alpha \in \eqso(\LL)$, we give a recursive function $\tau$ such that $\tau(\alpha)$ is in $\LL$-SNF and $\tau(\alpha) \equiv \alpha$. 
+Specifically, if $\alpha = \varphi$, define $\tau(\alpha) = \alpha$ and if $\alpha = s$, define $\tau(\alpha) = (\top \add \overset{\text{$s$-times}}{\ldots} \add \top)$.
+If $\alpha = (\alpha_1 + \alpha_2)$, define $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$ and if $\alpha = \sa{x}\beta$, define $\tau(\alpha) = \sum_{i = 1}^{k}\sa{x}\beta_i$ where $\tau(\beta) = \sum_{i = 1}^{k}\beta_i$ for some $k$ and each $\beta_i$ is in $\LL$-PNF.
+Finally, if $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case.
 This covers all possible cases for $\alpha$ and we conclude the proof by taking $\tau(\alpha)$ as the desired rewrite of $\alpha$.

### 2017-07-27 LMCS/saluja.tex
Pasada por sección 5.1.

@@
 On the other hand, if $\logu{1} \subseteq \LL$ and $\LL$ is closed under conjunction and disjunction, then every formula in $\eqso(\LL)$ can be rewritten in $\LL$-PNF. 
-
-From now on, for every first-order tuple $\bar{x}$ (second-order tuple $\bar{X}$) $\length{\bar{x}}$ as the size of $\bar{x}$ ($\length{\bar{X}}$ as the size of $\bar{X}$).
 \end{thm}
@@
 		edge[draw=white] node {\rotatebox{45}{$\subsetneq$}} (n4);
-	\node[rectw, right=0.8cm of n5] (n6) {$\U{1}$}       
+	\node[rectw, right=0.8cm of n5] (n6) {$\QU{1}$}       
 		edge[draw=white] node {$\subsetneq$} (n5);
-	\node[rectw, below=0.3cm of n6] (n7) {$\QU{1}$}       
+	\node[rectw, below=0.3cm of n6] (n7) {$\U{1}$}       
 		edge[draw=white] node {\rotatebox{90}{$=$}} (n6);
-	\node[rectw, right=1.0cm of n6] (n8) {$\E{2}$}       
+	\node[rectw, right=1.0cm of n6] (n8) {$\QE{2}$}       
 		edge[draw=white] node {$\subsetneq$} (n6);
-	\node[rectw, below=0.3cm of n8] (n9) {$\QE{2}$}       
+	\node[rectw, below=0.3cm of n8] (n9) {$\E{2}$}       
 		edge[draw=white] node {\rotatebox{90}{$=$}} (n8);        
-	\node[rectw, right=1.0cm of n8] (n10) {$\U{2}$}       
+	\node[rectw, right=1.0cm of n8] (n10) {$\QU{2}$}       
 		edge[draw=white] node {$\subsetneq$} (n8);
-	\node[rectw, below=0.3cm of n10] (n11) {$\QU{2}$}       
+	\node[rectw, below=0.3cm of n10] (n11) {$\U{2}$}       
 		edge[draw=white] node {\rotatebox{90}{$=$}} (n10); 
@@
 \end{center}
-\vspace{1cm}
-\begin{center}
-	\begin{tikzpicture}
-	\node[rectw] (n1) {$\E{0}$};
-	\node[rectw, right=1.2cm of n1] (n2) {};
-	\node[rectw, above=0.2cm of n2] (n3) {$\E{1}$}
-	edge[draw=white] node {\rotatebox{45}{$\subsetneq$}} (n1);
-	\node[rectw, below=0.1cm of n2] (n4) {$\QE{0}$}
-	edge[draw=white] node {\rotatebox{315}{$\subsetneq$}} (n1);
-	\node[rectw, right=1.2cm of n2] (n5) {$\QE{1}$}
-	edge[draw=white] node {\rotatebox{315}{$\subsetneq$}} (n3)
-	edge[draw=white] node {\rotatebox{45}{$\subsetneq$}} (n4);
-	\node[rectw, right=1.3cm of n5] (n6) {\rotatebox{90}{$=$}}       
-	edge[draw=white] node {$\subsetneq$} (n5);
-	\node[rectw, above=0cm of n6] (n13) {$\U{1}$};
-	\node[rectw, below=0cm of n6] (n7) {$\QU{1}$};      
-	\node[rectw, right=1.8cm of n6] (n8) {\rotatebox{90}{$=$}}       
-	edge[draw=white] node {$\subsetneq$} (n6);
-	\node[rectw, above=0cm of n8] (n14) {$\E{2}$};
-	\node[rectw, below=0cm of n8] (n9) {$\QE{2}$};       
-	\node[rectw, right=1.8cm of n8] (n10) {\rotatebox{90}{$=$}}       
-	edge[draw=white] node {$\subsetneq$} (n8);
-	\node[rectw, above=0cm of n10] (n15) {$\U{2}$};
-	\node[rectw, below=0cm of n10] (n11) {$\QU{2}$};      
-	edge[draw=white] node {\rotatebox{90}{$=$}} (n10); 
-	\node[rectw, right=0.5cm of n10] (n12) {$\sfo$}       
-	edge[draw=white] node {$=$} (n10); 
-	\end{tikzpicture}
-\end{center}
-\vspace{-0.4cm}
+%\vspace{1cm}
+%\begin{center}
+%	\begin{tikzpicture}
+%	\node[rectw] (n1) {$\E{0}$};
+%	\node[rectw, right=1.2cm of n1] (n2) {};
+%	\node[rectw, above=0.2cm of n2] (n3) {$\E{1}$}
+%	edge[draw=white] node {\rotatebox{45}{$\subsetneq$}} (n1);
+%	\node[rectw, below=0.1cm of n2] (n4) {$\QE{0}$}
+%	edge[draw=white] node {\rotatebox{315}{$\subsetneq$}} (n1);
+%	\node[rectw, right=1.2cm of n2] (n5) {$\QE{1}$}
+%	edge[draw=white] node {\rotatebox{315}{$\subsetneq$}} (n3)
+%	edge[draw=white] node {\rotatebox{45}{$\subsetneq$}} (n4);
+%	\node[rectw, right=1.3cm of n5] (n6) {\rotatebox{90}{$=$}}       
+%	edge[draw=white] node {$\subsetneq$} (n5);
+%	\node[rectw, above=0cm of n6] (n13) {$\U{1}$};
+%	\node[rectw, below=0cm of n6] (n7) {$\QU{1}$};      
+%	\node[rectw, right=1.8cm of n6] (n8) {\rotatebox{90}{$=$}}       
+%	edge[draw=white] node {$\subsetneq$} (n6);
+%	\node[rectw, above=0cm of n8] (n14) {$\E{2}$};
+%	\node[rectw, below=0cm of n8] (n9) {$\QE{2}$};       
+%	\node[rectw, right=1.8cm of n8] (n10) {\rotatebox{90}{$=$}}       
+%	edge[draw=white] node {$\subsetneq$} (n8);
+%	\node[rectw, above=0cm of n10] (n15) {$\U{2}$};
+%	\node[rectw, below=0cm of n10] (n11) {$\QU{2}$};      
+%	edge[draw=white] node {\rotatebox{90}{$=$}} (n10); 
+%	\node[rectw, right=0.5cm of n10] (n12) {$\sfo$}       
+%	edge[draw=white] node {$=$} (n10); 
+%	\end{tikzpicture}
+%\end{center}
 \caption{The relationship between the $\sfo$-hierarchy and the $\eqso(\fo)$-hierarchy, where $\E{1}$ and $\QE{0}$ are incomparable. \label{fig-sfo-eqso}}
@@
 \begin{itemize}
-\item $\QE{0}$ and $\E{1}$ are incomparable, that is, $\E{1} \not\subseteq \QE{0}$ and $\QE{0} \not\subseteq \E{1}$
-\item $\QE{1} \subsetneq \QU{1}$
+\item $\QE{0}$ and $\E{1}$ are incomparable, that is, $\E{1} \not\subseteq \QE{0}$ and $\QE{0} \not\subseteq \E{1}$,
+\item $\QE{1} \subsetneq \QU{1}$.
 \end{itemize}

### 2017-08-02 LMCS/proof-pi-minusone.tex
Avance en sección 5.2.

@@
 Then, the function $\ctcnf-1$, which counts the number of satisfying assignments of a 3-CNF formula minus one, is also in $\eqso(\LL)$.
-Note that $\eqso(\LL) \subseteq \eqso(\fo) = \shp$.
-We have that $\ctcnf$ is $\shp$-complete under parsimonious reductions\footnote{It can be easily verified that the standard reduction from SAT to 3-CNF (or 3-SAT) preserves the number of satisfying assignments}.
-Now, let $f$ be a function in $\shp$, and consider the nondeterministic polynomial-time procedure that on input $\enc(\A)$ computes the corresponding reduction to $\ctcnf$, name it $g(\enc(\A))$, and simulates the $\shp$ procedure for $\ctcnf-1$ on input $g(\enc(\A))$.
-We have that this is a $\shp$ procedure that computes $f-1$, from which we conclude that $\shp$ is closed under subtraction by one.
+Recall also that $\eqso(\LL) \subseteq \eqso(\fo) = \shp$ and that $\ctcnf$ is $\shp$-complete under parsimonious reductions\footnote{It can be easily verified that the standard reduction from SAT to 3-CNF (or 3-SAT) preserves the number of satisfying assignments}.
+Let $f$ be a function in $\shp$, and consider the nondeterministic polynomial-time procedure that on input $\enc(\A)$ computes the corresponding reduction $g(\enc(\A))$ into $\ctcnf$ and simulates the $\shp$ procedure for $\ctcnf-1$ on input $g(\enc(\A))$.
+This is a $\shp$ procedure that computes $f-1$, from which we conclude that $\shp$ is closed under subtraction by one.

### 2017-08-02 LMCS/proof-prop-e1-nc.tex
Avance en sección 5.2.

@@
-Towards a contradiction, assume that the statement is false.
-This is, that $\E{1}$ is closed under binary sum.
-Consider the formula $\sa{x}(x = x)$ which is in $\E{1}$ over some signature $\R$.
-For every finite $\R$-structure $\A$ with $n$ elements, and where every predicate in $\R$ is empty, $\alpha(\enc(\A)) = n$.
+Towards a contradiction, assume that the statement is false, namely, $\E{1}$ is closed under binary sum.
+Consider the formula $\sa{x}(x = x) \in \E{1}$ over some signature $\R$.
+Clearly, for every finite $\R$-structure $\A$ with $n$ elements it holds that $\alpha(\enc(\A)) = n$.
 From our assumption, there exists some formula in $\E{1}$ equivalent to the formula $\alpha \add \alpha$, which describes the function $2n$.
-Let $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ be this formula, where $\varphi$ is quantifier-free.
+Let $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ be this formula, where $\varphi$ is a quantifier-free first-order formula.
 Note that the function defined by this formula is equal or greater than the one defined by $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ divided by a polynomial factor.
-More specifically, for each ordered structure $\A$ with domain $A$, we have the following inequality:
+More specifically, for each ordered structure $\A$ with $n$ elements, we have the following inequality:
 $$
-\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot \vert A \vert^{\length{\bar{y}}} \geq \sem{\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A)
+\sem{\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A)
+\leq 
+\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot  n^{\length{\bar{y}}} \leq 2n^{\length{\bar{y}}+1} 
 $$
-Note that the formula $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ defines a function in $\E{0}$.
-It was shown by the authors in \cite{SalujaST95} that every function in $\E{0}$ grows exponentially over the size of the structure for large enough structures, when $\length{\bar{X}} > 0$.
-This function divided by a polynomial factor still grows exponentially.
-Therefore, for $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ we have that $\length{\bar{X}} = 0$.
+Note that the formula $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ defines a function in $\E{0}$ and it was shown by the authors in \cite{SalujaST95} that every function in $\E{0}$ grows exponentially over the size of the structure for large enough structures, when $\length{\bar{X}} > 0$.
+This contradict the above inequality and, therefore, for $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ we have that $\length{\bar{X}} = 0$.
 
-Now, for the formula $\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ consider a structure $\mathfrak{1}$ with only one element $a$.
-We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$, but the only possible assignment to $\bar{x}$ is the tuple $(a,\ldots,a)$ so $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which follows to a contradiction.
+Now, if the formula for defining $\alpha + \alpha$ is of the form $\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ consider a structure $\mathfrak{1}$ with only one element $a$.
+We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$, but the only possible assignment to $\bar{x}$ is the tuple $(a,\ldots,a)$ so $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which leads to a contradiction.

### 2017-08-02 LMCS/proof-qe0-fp-qe1-totp-fptras.tex
Avance en sección 5.2.

@@
 This is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every finite $\R$-structure $\A$ with domain $A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$.
-We base our proof on this fact.
+We use this fact in the next arguments.
 
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is contained in \textsc{TotP}.} Let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
+To show that $\eqso(\loge{1})$ is contained  in \textsc{TotP}, let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
 This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
@@
 Since, $\fp\subseteq\totp$\cite{PagourtzisZ06}, each polynomially computable function is also in $\totp$, and then on each of these branches we simulate the corresponding nondeterministic procedure to generate $h(\vert A \vert)$ more.
-The number of branches for each $\alpha_i$ is $\sem{\alpha_i}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$, and the total number of branches in the procedure amounts to $\sem{\alpha}(\A)$.
+The number of branches for each $\alpha_i$ is $\sem{\alpha_i}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$, and the total number of branches is equal to $\sem{\alpha}(\A)$.
 We conclude that $\alpha\in\totp$.
 
-\vspace{1em}
-{\bf Every function in} $\boldsymbol{\eqso(\loge{1})}$ {\bf has an FPRAS.}  let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
+To show that every function in $\eqso(\loge{1})$ has an FPRAS,  let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF.
 This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF.
@@
 
-\vspace{1em}
-$\boldsymbol{\eqso(\loge{1})}$ {\bf is closed under sum and multiplication.} Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that the class is closed under multiplication.
+Finally, we show that $\eqso(\loge{1})$ is closed under sum and multiplication.
+Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that is closed under multiplication.
 We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of $\so$.
@@
 \proof
-	We will define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ extended by binary product, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
+	We define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ extended by binary product, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
 In fact, the formula $\tau(\alpha)$ is in $\LL$-SNF.
-First we replace each constant $s$ in $\alpha$ for $(\top \add \cdots \add \top)$ ($s$ times).
+First we replace each constant $s$ with $(\top \add \cdots \add \top)$ ($s$ times).
 If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$.
@@
 This covers all possible cases for $\alpha$.
-For every pair of formulas $\alpha,\beta$ in $\eqso(\LL)$, we have that their multiplication $(\alpha\cdot\beta)$ is a formula in the grammar $\eqso(\LL)$ extended by binary product, and so, there exists an equivalent formula $\tau(\alpha\cdot\beta)$ which is in unextended $\eqso(\LL)$.
+For every pair of formulas $\alpha,\beta$ in $\eqso(\LL)$, we have that their multiplication $(\alpha\cdot\beta)$ is a formula in the grammar $\eqso(\LL)$ extended by binary product, and so, there exists an equivalent formula $\tau(\alpha\cdot\beta)$ which is in $\eqso(\LL)$.
 \qed
-Since $\loge{1}$ is closed under conjunction, this also holds for $\eqso(\loge{1})$. This concludes the proof.
+Since $\loge{1}$ is closed under conjunction, then Lemma~\ref{conj-mult} also holds for $\eqso(\loge{1})$. This concludes the proof.

### 2017-08-02 LMCS/proof-sigmafo-minusone.tex
Avance en sección 5.2.

@@
 
-\noindent {\bf Closed under sum and multiplication.} We show here that $\eqso(\logex{1})$ is closed under sum and multiplication.
-This can be seen because $\eqso(\LL)$ is closed under sum for every fragment $\LL$ by definition, and since $\logex{1}$ is closed under conjunction, from Lemma \ref{conj-mult} it follows that $\eqso(\logex{1})$ is closed under multiplication.
+For the sake of readability, we divide the proof in three parts. The last part, i.e. subtraction by one, is probably the most technical proof of the paper.
 
-\vspace{1em}
-\noindent {\bf Easy decision version and FRPRAS.}  We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
-We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Theorem \ref{prop:qe0-fp-qe1-totp-fptras}.
+\medskip
+
+\noindent {\bf Closed under sum and multiplication.} By the previous results, it is straightforward to prove that $\eqso(\logex{1})$ is closed under sum and multiplication.
+Indeed, $\eqso(\LL)$ is closed under sum by definition for every fragment $\LL$, and since $\logex{1}$ is closed under conjunction, from Lemma \ref{conj-mult} it follows that $\eqso(\logex{1})$ is closed under multiplication.
+\medskip
+
+\noindent {\bf Easy decision version and FPRAS.}  We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
+We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
 First, we define a function that converts a formula $\alpha$ in $\eqso(\logex{1})$ over a signature $\R$ into a formula $\lambda(\alpha)$ in $\eqso(\loge{1})$ over a signature $\R_{\alpha}$.
@@
 
-Let $\alpha$ be in $\eqso(\logex{0})$.
-The signature $\R_{\alpha}$ is obtained by adding the symbol $R_{\psi}$, for every $\fo$ formula $\psi(\bar{z})$ in $\alpha$, to $\R$.
+Let $\alpha$ be in $\eqso(\logex{1})$.
+The signature $\R_{\alpha}$ is obtained by adding the symbol $R_{\psi}$ to $\R$ for every $\fo$ formula $\psi(\bar{z})$ in $\alpha$.
 Each symbol $R_{\psi}$ represents a predicate with arity $\length{\bar{z}}$.
 Then, $\lambda(\alpha)$ is defined as $\alpha$ where each $\fo$ formula $\psi(\bar{z})$ has been replaced by $R_{\psi}(\bar{z})$.
-We now define the function $g_{\alpha}$ procedurally.
+We now define the function $g_{\alpha}$ with a polynomial time procedure.
 Let $\A$ be a $\R$-structure with domain $A$.
@@
 For each $\fo$-formula $\psi(\z)$ with $\length{\bar{z}}$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length{\bar{z}}}$.
-If $\A\models\psi(\bar{a})$, then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$ (this can be done in $\ptime$).
+If $\A\models\psi(\bar{a})$ (this can be done in $\ptime$), then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$ .
 This concludes the construction of $\A'$.
@@
 Similarly, to show that $\alpha$ has an FPRAS, we do the same as before and simulates the FPRAS for $\lambda(\alpha)$ in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-These procedures also takes polynomial time and satisfies the condition.
+These procedures also takes polynomial time and satisfies the required conditions.
 
-\vspace{1em}
+\medskip
+
 \noindent {\bf Closed under subtraction by one.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. 
-For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Without lost of generality, we assume that $\alpha$ is in $\logex{1}$-SNF and $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
-Moreover, we assume that $\length{\bar{x}} > 0$  since, if this is not the case, we can replace $\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
+For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$-formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
+Without lost of generality, we assume that $\alpha$ is in $\logex{1}$-SNF, that is, $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
+Moreover, we assume that $\length{\bar{x}} > 0$  since, if this is not the case, we can replace $\sa{\bar{X}}\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
 
 The proof will be separated in two parts.
-In the first part, we suppose that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$.
+In the first part, we will assume that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$.
 Then we will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ is the formula.
+From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ will be the desired formula.
 In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ the sum of one or more formulas in $\logex{1}$-PNF.
@@
 
-\vspace{1em}
+\smallskip
 
@@
 Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}} \varphi(\bar{x})$ (i.e.
-$\length{\bar{X}} = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}} [\varphi(\bar{x})\wedge\exists\bar{z}(\varphi(\bar{z})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
+$\length{\bar{X}} = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}} [\varphi(\bar{x})\wedge \ex{\bar{z}}(\varphi(\bar{z})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
 Therefore, for the rest of the proof we can assume that $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
@@
 To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in a DNF formula. 
-More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i(\bar{X}, \bar{x}) = \exists \bar{y} \, \varphi_i'(\bar{X}, \bar{x}, \bar{y})$ and $\varphi_i'(\bar{X}, \bar{x}, \bar{y})$ is a conjunction of atomic formulas or negation of atomic formulas.
-Furthermore, we suppose that each conjunct $\varphi_i'(\bar{X}, \bar{x}, \bar{y})$ has the~form:
+More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i(\bar{X}, \bar{x}) = \ex{\bar{y}} \varphi_i'(\bar{X}, \bar{x}, \bar{y})$ and $\varphi_i'(\bar{X}, \bar{x}, \bar{y})$ is a conjunction of atomic formulas or negation of atomic formulas.
+Furthermore, we assume that each $\varphi_i'(\bar{X}, \bar{x}, \bar{y})$ has the~form:
 $$
@@
 $$
-Note that atomic formulas of the form $R(\bar{z})$ where $R \in \R$ will appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
+Note that atomic formulas, like $R(\bar{z})$ for $R \in \R$, will appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
 
-Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ are mentioned in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
+Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ are mentioned in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the formal definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
 We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
-\begin{itemize}
+\begin{itemize} \itemsep1mm
 	\item[(a)] In order to satisfy the first condition, consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
 We add $\length{\bar{w}}$ new variables $z_1,\ldots,z_{\length{\bar{w}}}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length{\bar{w}}})$.
-We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$.
-The formula $\varphi_i$ is now defined as
+We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$ (denoted by $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})[X_j(\bar{w}) \leftarrow X_j(\bar{z})]$ and then the formula $\varphi_i$ is equivalently defined as:
 	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \exists \bar{z} \; (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{-}(\bar{X},\bar{x},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y})).
+	\varphi_i(\bar{X},\bar{x}) \ := \ \ex{\bar{y}} \ex{\bar{z}} (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y})[X_j(\bar{w}) \leftarrow X_j(\bar{z})] \wedge
+	\varphi_i^{-}(\bar{X},\bar{x},\bar{y}).
 	$$
 	We repeat this process for each instance of a $X_j(\bar{w})$ in $\varphi_i$, and we obtain a formula where none of the $X_j$'s acts over any variable in $\bar{x}$.
-We add the new first-order variables to $\bar{y}$ and we redefine $\varphi_i$ as
+We add the new first-order variables to $\bar{y}$ and we redefine $\varphi_i$ as:
 	$$
-	\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \; \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
+	\varphi_i(\bar{X},\bar{x}) \ := \  \ex{\bar{y}} \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
 	$$
-	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = \exists \bar{y} \;  X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := \exists \bar{y} \;  X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$
+	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = \ex{\bar{y}}  x < y \wedge  X(x,y)\wedge \neg X(x,x)$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and:
+	$$
+	\varphi_i \ := \ \ex{\bar{y}} v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x \wedge x < y  \wedge  X(v_1,v_2) \wedge \neg X(v_3,v_4).
+	$$ 
 	\item[(b)] An ordered partition on a set $S$ is defined by an equivalence relation $\sim$ over $S$, and a linear order over $S/\!\sim$.
@@
 	For a given $k$, let $\cB_k$ be the number of possible ordered partitions for a set of size $k$. For $1 \leq j \leq \cB_{\length{(\bar{x},\bar{y})}}$ 
-	let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$.
-
-	Thus, the formula $\varphi_i(\bar{X},\bar{x})$ is then redefined as:
+	let $\theta^j(\bar{x},\bar{y})$ be the formula that defines the $j$-th ordered partition over $(\bar{x},\bar{y})$. 
+	Thus, the formula $\varphi(\bar{X},\bar{x})$ is then redefined as:
 	$$
-	\bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length{(\bar{x},\bar{y})}}} \exists \bar{y} \; [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
+	\varphi(\bar{X},\bar{x}) \ := \ \bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length{(\bar{x},\bar{y})}}} \ex{\bar{y}} [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
 	$$
 	Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
-	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  as $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$.
-	\item[(c)] For showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
+	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  as $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ forces an ordered partition over the variables in $(\bar{x},\bar{y})$.
+	\item[(c)] Finally, for showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
 $\varphi_i$ is inconsistent), we do the following.
@@
 It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$.
-From now on, we assume that each $\varphi_i(\bar{X},\bar{x}) = \exists \bar{y} \;  \varphi_i'(\bar{X},\bar{x}, \bar{y}) $ satisfies conditions (a), (b) and (c), and $\varphi_i'(\bar{X},\bar{x}, \bar{y})$ has the following~form:
+From now on, we assume that each $\varphi_i(\bar{X},\bar{x}) = \ex{\bar{y}}  \varphi_i'(\bar{X},\bar{x}, \bar{y}) $ satisfies conditions (a), (b) and (c), and $\varphi_i'(\bar{X},\bar{x}, \bar{y})$ has the following~form:
 $$
@@
 \begin{clm}\label{claim:minusone}
-	For a given ordered structure $\A$ and a first-order assignment $v$ for $\A$, $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ if, and only if, there exists a second-order assignment $V$ for $\A$ such that $(\A,V,v)\models \varphi_i'(\bar{X},\bar{x}, \bar{y})$.
+	For an ordered structure $\A$ and a first-order assignment $v$ for $\A$, $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ if, and only if, there exists an assignment $V$ for $\A$ such that $(\A,V,v)\models \varphi_i'(\bar{X},\bar{x}, \bar{y})$.
 \end{clm}
@@
 \[
-\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge\forall\bar{x}'\forall\bar{y}'(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')).
+\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge \fa{\bar{x}'} \fa{\bar{y}'}(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')).
 \] 
@@
 \begin{align}
-\psi_i(\bar{X},\bar{x}) = &\big( \, \exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
+\psi_i(\bar{X},\bar{x}) = &\big( \, \ex{\bar{x}} \ex{\bar{y}} \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
 &\Big( \; 
-\exists\bar{y}~\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i'(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \exists\bar{z}\, ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\
-&(\; \exists\bar{x}'\exists\bar{y}\; \varphi_i'(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x} \; ) \;\; \Big)  \label{eq:part3}
+\ex{\bar{y}} \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i'(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \ex{\bar{z}} ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\
+&(\; \ex{\bar{x}'} \ex{\bar{y}} \varphi_i'(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x} \; ) \;\; \Big)  \label{eq:part3}
 \end{align}
 To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
-Indeed, by Claim~\ref{claim:minusone} we know that if $\exists\bar{x}\exists\bar{y}\, \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi_i'(\bar{X},\bar{x},\bar{y})$.
+Indeed, by Claim~\ref{claim:minusone} we know that if $\ex{\bar{x}} \ex{\bar{y}} \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi_i'(\bar{X},\bar{x},\bar{y})$.
 Then, the conclusion of the implication (divided into (\ref{eq:part2}) and (\ref{eq:part3})), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
@@
 where each modified disjunct $\varphi_i^*(\bar{X},\bar{x})$ is constructed as follows. 
-For the sake of simplification, define the auxiliary formula $\chi_i = \neg\exists\bar{x}\exists\bar{y}\,\varphi^{\fo}_i(\bar{x},\bar{y})$. 
+For the sake of simplification, define the auxiliary formula $\chi_i = \neg \ex{\bar{x}} \ex{\bar{y}}\varphi^{\fo}_i(\bar{x},\bar{y})$. 
 This formula basically checks if $\varphi_i$ is not satisfiable (recall Claim~\ref{claim:minusone}).
@@
 We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if, and only if, $\sem{\beta}(\A) = 0$.
-If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg\exists\bar{x}'\varphi(\bar{x}')$.
+If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg \ex{\bar{x}'} \varphi(\bar{x}')$.
 If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and define $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$.

### 2017-08-02 LMCS/syntactic-clo.tex
Avance en sección 5.2.

@@
 Besides, $\totp$ has good closure properties as it is closed under sum, multiplication and subtraction by one.
-However, some functions in $\totp$ do not admit FPRAS under standard complexity-theoretical assumptions,\footnote{As an example consider the problem of counting the number of independent sets in a graph, and the widely believed assumption that $\np$ is not equal to the randomized complexity class $\rp$ (Randomized Polynomial-Time \cite{G77}).
-This counting problem is in $\totp$, and it is known that $\np = \rp$ if there exists an FPRAS for it \cite{DFJ02}.} and no natural complete problems are known for this class \cite{PagourtzisZ06}.
+However, some functions in $\totp$ do not admit FPRAS under standard complexity-theoretical assumptions\footnote{As an example consider the problem of counting the number of independent sets in a graph, and the widely believed assumption that $\np$ is not equal to the randomized complexity class $\rp$ (Randomized Polynomial-Time \cite{G77}).
+This counting problem is in $\totp$, and it is known that $\np = \rp$ if there exists an FPRAS for it \cite{DFJ02}.}, and no natural complete problems are known for this class \cite{PagourtzisZ06}.
 Hence, we use the $\eqso(\fo)$-hierarchy to find restrictions of $\totp$ with good approximation and closure properties.
@@
 Thus, we need to find an extension of $\QE{1}$ that keeps all the previous properties and is closed under subtraction by one.
-It is important to notice that $\shp$ is believed not to be closed under subtraction by one by some complexity-theoretical assumption.\footnote{A decision problem $L$ is in the randomized complexity class $\cspp$ if there exists a polynomial-time NTM $M$ such that for every $x \in L$ it holds that $\tma_M(x) - \tmr_M(x) = 2$, and for every $x \not\in L$ it holds that $\tma_M(x) = \tmr_M(x)$ \cite{OH93,FFK94}.
+It is important to notice that $\shp$ is believed not to be closed under subtraction by one by some complexity-theoretical assumption\footnote{A decision problem $L$ is in the randomized complexity class $\cspp$ if there exists a polynomial-time NTM $M$ such that for every $x \in L$ it holds that $\tma_M(x) - \tmr_M(x) = 2$, and for every $x \not\in L$ it holds that $\tma_M(x) = \tmr_M(x)$ \cite{OH93,FFK94}.
 It is believed that $\np \not\subseteq \cspp$.
-However, if $\shp$ is closed under subtraction by one, then it holds that $\np \subseteq \cspp$ \cite{OH93}.} So, the following proposition rules out any logic that extends $\Pi_1$ for a possible extension of~$\QE{1}$ with the desired closure property.
+However, if $\shp$ is closed under subtraction by one, then it holds that $\np \subseteq \cspp$ \cite{OH93}.}.
+So, the following proposition rules out any logic that extends $\Pi_1$ for a possible extension of~$\QE{1}$ with the desired closure property.
 \begin{prop} \label{pi-minusone}
@@
 \qed
-Therefore, the desired extension has to be achieved by allowing some local extensions to $\Sigma_1$.
-More precisely, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulae over a signature $\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mentioned any second-order variable).
+Therefore, the desired extension has to be achieved by allowing some local extensions to~$\Sigma_1$.
+More precisely, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulae over a signature~$\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mentioned any second-order variable).
 With this extension we obtain a class with the desired properties.

### 2017-08-04 LMCS/proof-sigmafo-minusone.tex
Sección 5.2 lista.

@@
 
-\smallskip
+\medskip
 
@@
 	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  as $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ forces an ordered partition over the variables in $(\bar{x},\bar{y})$.
-	\item[(c)] Finally, for showing that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
+	\item[(c)] Finally, to rewrite the formula such that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
 $\varphi_i$ is inconsistent), we do the following.
@@
 It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$.
-From now on, we assume that each $\varphi_i(\bar{X},\bar{x}) = \ex{\bar{y}}  \varphi_i'(\bar{X},\bar{x}, \bar{y}) $ satisfies conditions (a), (b) and (c), and $\varphi_i'(\bar{X},\bar{x}, \bar{y})$ has the following~form:
+From now on, we assume that each $\varphi_i(\bar{X},\bar{x}) = \ex{\bar{y}}  \varphi_i'(\bar{X},\bar{x}, \bar{y}) $ satisfies conditions (a), (b) and (c), and $\varphi_i'(\bar{X},\bar{x}, \bar{y})$ has the~form:
 $$
@@
 \begin{clm}\label{claim:minusone}
-	For an ordered structure $\A$ and a first-order assignment $v$ for $\A$, $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ if, and only if, there exists an assignment $V$ for $\A$ such that $(\A,V,v)\models \varphi_i'(\bar{X},\bar{x}, \bar{y})$.
+	For an ordered structure $\A$ and a FO assignment $v$ for $\A$, $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ if, and only if, there exists a SO assignment $V$ for $\A$ such that $(\A,V,v)\models \varphi_i'(\bar{X},\bar{x}, \bar{y})$.
 \end{clm}
@@
 Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$.
-This violates condition (c) since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
+This violates condition (c) above since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
 
@@
 One can easily check that for every assignments $(V, v)$ such that $(\A,V,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$, it holds that $(\A,\bar{B}^v,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all assignments of $\bar{X}$ when $v$ is fixed.
-Therefore, by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
-For this, the idea is to choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
+This motivates the main idea of Part (1): by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
+For this, we choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
 \[
@@
 If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
-Furthermore, let $V^*$ the second order assignment and $v^*$ the first order assignment that satisfy $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*(\bar{x}) = v(\bar{x})$.
-By the previous discussion, we have that $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.
+Furthermore, let $V^*$ be the second order assignment and $v^*$ the first order assignment that satisfy $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*(\bar{x}) = v(\bar{x})$.
+By the previous discussion, $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.
 
-Now, we have all the tools to introduce one of the main formulas in order to define $\kappa(\alpha)$. 
+Now, we have all the ingredients in order to define $\kappa(\alpha)$. 
 Intuitively, we want to exclude the assignment $(V^*, v^*)$ from the satisfying assignments of $\varphi_i(\bar{X},\bar{x})$.
-Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if, 
-$V \neq V^*$ or $v \neq v^*$ whenever $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
+Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if, if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, then
+$V \neq V^*$ or $v \neq v^*$. 
 This property can be defined as follows:
 \begin{align}
-\psi_i(\bar{X},\bar{x}) = &\big( \, \ex{\bar{x}} \ex{\bar{y}} \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
+\psi_i(\bar{X},\bar{x}) \; := \;  &\big( \, \ex{\bar{x}} \ex{\bar{y}} \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
 &\Big( \; 
-\ex{\bar{y}} \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \varphi_i'(\bar{X},\bar{x}, \bar{y})  \rightarrow \bigvee_{X \in \bar{X}} \ex{\bar{z}} ( \, X(\bar{z}) \wedge \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})}\bar{w}\neq\bar{z} \, ) \big)  \;\;\;\;  \vee   \label{eq:part2} \\
+\ex{\bar{y}} \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \; \varphi_i'(\bar{X},\bar{x}, \bar{y})  \rightarrow \!\!
+\bigvee_{X \in \bar{X}} \ex{\bar{z}} [ \, X(\bar{z}) \wedge   \!\!\!\!\!\!\!\!\!\!
+\bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})} \!\!\!\!\!\!\!\!\!\!\!
+\bar{w}\neq\bar{z} \, ] \; \big) \;  \vee   \label{eq:part2} \\
 &(\; \ex{\bar{x}'} \ex{\bar{y}} \varphi_i'(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x} \; ) \;\; \Big)  \label{eq:part3}
@@
 \[
-\varphi^*_1(\bar{X},\bar{x}) = \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
+\varphi^*_1(\bar{X},\bar{x}) \; := \; \varphi_1(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x}).
 \]
@@
 \[
-\varphi^*_2(\bar{X},\bar{x}) = \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
+\varphi^*_2(\bar{X},\bar{x}) \; := \; \varphi_2(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\wedge(\chi_1\to\psi_2(\bar{X},\bar{x})).
 \]
@@
 \begin{multline*}
-\varphi_i^*(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\,\wedge \\ (\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
-\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})),
+\varphi_i^*(\bar{X},\bar{x}) \; := \; \varphi_i(\bar{X},\bar{x})\wedge\psi_1(\bar{X},\bar{x})\,\wedge \\ (\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
+\bigwedge_{j = 1}^{j = i-1}\chi_j\to\psi_i(\bar{X},\bar{x})).
 \end{multline*}
@@
 
-\smallskip
+\medskip
 
-{\em Part (2).} Let $\alpha = (\beta + \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ for some $\eqso(\logex{1})$ formula $\beta$. We define $\kappa(\alpha)$ as follows.
-First, rewrite $\varphi(\bar{X},\bar{x})$ as in the previous section.
-Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c).
-Also, consider the previously defined formulas $\chi_i$ and $\psi_i$, for each $i$.
-
+\noindent {\em Part (2).} Let $\alpha = \beta + \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ for some $\eqso(\logex{1})$ formula $\beta$. We define $\kappa(\alpha)$ as follows.
+First, rewrite $\varphi(\bar{X},\bar{x})$ as in Part (1).
+Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) defined above.
+Also, consider the previously defined formulas $\chi_i$ and $\psi_i$, for each $i \leq m$.
 We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if, and only if, $\sem{\beta}(\A) = 0$.
 If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg \ex{\bar{x}'} \varphi(\bar{x}')$.
-If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and define $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$.
-If $\beta = (\beta_1 + \beta_2)$, then $\lambda(\beta) = \lambda(\beta_1) \wedge \lambda(\beta_2)$.
-Now, for each $\varphi_i$ we define:
+If $\beta = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$, then
+%let $\varphi = \bigvee_{i = 1}^{m}\varphi_i$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) of the previous section, and 
+define $\lambda(\beta) = \chi_1\wedge \cdots\wedge\chi_m$. If $\beta = (\beta_1 + \beta_2)$, then $\lambda(\beta) = \lambda(\beta_1) \wedge \lambda(\beta_2)$.
+Now, following the same ideas as in Part (1) we define a formula $\varphi_i^*(\bar{X},\bar{x})$ that removes the minimal $(V^*, v^*)$ of $\varphi_i$ whenever $\beta$ cannot be satisfied (i.e.
+$\lambda(\beta)$ is true).
+Formally, we define $\varphi_i^*$ as follows:
 \begin{multline*}
-\varphi_i^*(\bar{X},\bar{x}) = \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\,\wedge \\
+\varphi_i^*(\bar{X},\bar{x}) \ := \ \varphi_i(\bar{X},\bar{x})\wedge\Big(\lambda(\beta)\to\Big(\psi_1(\bar{X},\bar{x})\,\wedge \\
 (\chi_1\to\psi_2(\bar{X},\bar{x}))\wedge((\chi_1\wedge\chi_2)\to\psi_3(\bar{X},\bar{x}))\wedge\cdots\wedge(
@@
 \end{multline*}
-Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i^*(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired condition.
+Finally, $\kappa(\alpha)$ is defined as $\kappa(\alpha) = \kappa(\beta) + \sa{\bar{X}}\sa{\bar{x}} \bigvee_{i = 1}^m\varphi_i^*(\bar{X},\bar{x})$, which is in $\eqso(\logex{1})$ and satisfies the desired conditions.
+This concludes the proof.

### 2017-08-04 LMCS/syntactic-horn.tex
Avance en sección 5.3.

@@
 To define $\chsat$, we consider an \so-formula $\varphi(\pT)$ over $\R$, where $\pT$ is a unary predicate, such that for every Horn formula $\theta$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \varphi(\pT)}(\A)$.
-In particular, $\pT(x)$ holds if and only if $x$ is a propositional variable that is assigned value true.
+In particular, $\pT(x)$ holds if, and only if, $x$ is a propositional variable that is assigned value true.
 More specifically,
@@
 \begin{align*}
-\varphi(\pT) \; :=\;\;  &\forall x \, (\pT(x) \to \pV(x)) \ \wedge\\
-&\forall c \, (\pNC(c) \to \exists x \, (\pN(c,x) \wedge \neg \pT(x))) \ \wedge\\
-&\forall c \forall x \, ([\pP(c,x) \wedge \forall y \, (\pN(c,y) \to \pT(y))] \to \pT(x)).
+\varphi(\pT) \; :=\;\;  & \fa{x} (\pT(x) \to \pV(x)) \ \wedge\\
+& \fa{c}  (\pNC(c) \to \ex{x} (\pN(c,x) \wedge \neg \pT(x))) \ \wedge\\
+& \fa{c} \fa{x} ([\pP(c,x) \wedge \fa{y} (\pN(c,y) \to \pT(y))] \to \pT(x)).
 \end{align*}
@@
 \begin{align*}
-&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \pNC(c) \vee \exists x \, (\pN(c,x) \wedge \neg \pT(x)))\ \wedge\\
-&\forall c \forall x \, (\neg \pP(c,x) \vee \exists y \, (\pN(c,y) \wedge \neg \pT(y)) \vee \pT(x)).
+& \fa{x}  (\neg \pT(x) \vee \pV(x)) \ \wedge\\
+& \fa{c}  (\neg \pNC(c) \vee \ex{x} (\pN(c,x) \wedge \neg \pT(x)))\ \wedge\\
+& \fa{c} \fa{x}  (\neg \pP(c,x) \vee \ex{y} (\pN(c,y) \wedge \neg \pT(y)) \vee \pT(x)).
 \end{align*}
@@
 \begin{align*}
-\forall c \forall x \, (\neg \pA(c,x) \leftrightarrow [\pN(c,x) \wedge \neg \pT(x)]),
+\fa{c} \fa{x}  (\neg \pA(c,x) \leftrightarrow [\pN(c,x) \wedge \neg \pT(x)]),
 \end{align*}
@@
 \begin{align*}
-\psi(\pT,\pA) \; :=\;\;  &\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \neg\textit{T}(x)).
+\psi(\pT,\pA) \; := \;\;  & \fa{x} (\neg \pT(x) \vee \pV(x)) \ \wedge\\
+& \fa{c} (\neg \textit{NC}(c) \vee \ex{x} \neg \textit{A}(c,x)) \ \wedge\\
+& \fa{c} \fa{x}  (\neg \textit{P}(c,x) \vee \ex{y} \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
+& \fa{c} \fa{x} (\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge \\
+& \fa{c} \fa{x} (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
+& \fa{c} \fa{x} (\textit{A}(c,x) \vee \neg\textit{T}(x)).
 \end{align*}
@@
 Therefore, the formula $\psi(\pT,\pA)$ also defines $\chsat$.
-More importantly, $\psi(\pT,\pA)$ resembles a conjunction of Horn clauses except for the use of negative literals of the form $\exists v \, \neg \textit{A}(u,v)$.
+More importantly, $\psi(\pT,\pA)$ resembles a conjunction of Horn clauses except for the use of negative literals of the form $\ex{v} \neg \textit{A}(u,v)$.
 \end{exa}
 The previous example suggests that to define $\chsat$, we can use Horn formulae defined as follows. 
-A positive literal is a formula of the form $X(\x)$, where $X$ is a second-order variable and $\x$ is a tuple of first-order variables, and a negative literal is a formula of the form $\exists \v \, \neg X(\u,\v)$, where $\u$ and $\v$ are tuples of first-order variables.
-Given a signature $\R$, a clause over $\R$ is a formula of the form $\forall \x \, (\varphi_1 \vee \cdots \vee \varphi_n)$,
+A positive literal is a formula of the form $X(\x)$, where $X$ is a second-order variable and $\x$ is a tuple of first-order variables, and a negative literal is a formula of the form $\ex{\v} \neg X(\u,\v)$, where $\u$ and $\v$ are tuples of first-order variables.
+Given a signature $\R$, a clause over $\R$ is a formula of the form $\fa{\x} (\varphi_1 \vee \cdots \vee \varphi_n)$,
 where each $\varphi_i$ ($1 \leq i \leq n$) is either a positive literal, a negative literal or an \fo-formula over $\R$.
@@
 
-We have that $\chsat \in \eqso(\uhorn)$. Moreover, one can show that $\eqso(\uhorn)$ forms a class of functions with easy decision counterparts, namely, $\eqso(\uhorn) \subseteq \totp$.
+As we have seen, we have that $\chsat \in \eqso(\uhorn)$.
+Moreover, one can show that $\eqso(\uhorn)$ forms a class of functions with easy decision counterparts, namely, $\eqso(\uhorn) \subseteq \totp$.
 Thus, $\eqso(\uhorn)$ is a new alternative in our search for a class of functions in $\shp$ with easy decision counterparts and natural complete problems.
 Moreover, an even larger class for our search can be generated by extending the definition of $\uhorn$ with outermost existential quantification.
-Formally, a formula $\varphi$ is in $\ehorn$ if $\varphi$ is of the form $\exists \bar x \, \psi$ with $\psi$ a Horn formula.
+Formally, a formula $\varphi$ is in $\ehorn$ if $\varphi$ is of the form $\ex{\bar x} \psi$ with $\psi$ a Horn formula.
 
@@
 Interestingly, we have that both $\chsat$ and $\cdnf$ belong to $\eqso(\ehorn)$. 
-An imperative question at this point is whether in the definitions of $\uhorn$ and $\ehorn$, it is necessary to allow negative literals of the form $\exists \v \, \neg X(\u,\v)$.
+An imperative question at this point is whether in the definitions of $\uhorn$ and $\ehorn$, it is necessary to allow negative literals of the form $\ex{\v} \neg X(\u,\v)$.
 Actually, this forces our Horn classes to be included in $\eqso(\logu{2})$ and not necessarily in $\eqso(\loge{2})$.

### 2017-08-07 LMCS/proof-ehorn-pe.tex
Lista sección 5.3.

@@
-Pagourtzis and Zachos mention a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula \cite{PagourtzisZ06}.
+%Let $\shdhsat$ be the problem of counting the number of satisfying assignments of a disjunction of Horn formulas. 
+In~\cite{PagourtzisZ06}, Pagourtzis and Zachos gave a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula.
 This procedure can be easily extended to receive Horn formulas, and furthermore, a disjunction of Horn formulas.
 Hence $\shdhsat$ is in $\totp$.
-
-As we show in Proposition \ref{sigma2hard}, $\shdhsat$ is complete for $\eqso(\ehorn)$ under parsimonious reductions.
+As we saw in Theorem~\ref{sigma2hard}, $\shdhsat$ is complete for $\eqso(\ehorn)$ under parsimonious reductions.
 Let $\alpha$ be a formula in $\eqso(\ehorn)$ and let $g_{\alpha}$ be the reduction to $\shdhsat$.
-The $\totp$ procedure we construct, for each input $\enc(\A)$, is simply to compute $g_{\alpha}(\enc(\A))$, and then simulate the $\totp$ procedure for $\shdhsat$ on input $g_{\alpha}(\enc(\A))$.
-We conclude that $\alpha$ is in $\totp$.
+Then the $\totp$ procedure consist in computing $g_{\alpha}(\enc(\A))$ for each input $\enc(\A)$ and then simulate the $\totp$ procedure for $\shdhsat$ on input $g_{\alpha}(\enc(\A))$.
+Therefore, we conclude that $\alpha$ defines a function in~$\totp$.

### 2017-08-07 LMCS/proof-hsat-not-sigma2.tex
Lista sección 5.3.

@@
-We use a similar proof to the one provided by the authors in \cite{SalujaST95} to separate the classes $\E{2}$ and $\U{2}$.
+%We use a similar proof to the one provided by the authors in \cite{SalujaST95} to separate the classes $\E{2}$ and $\U{2}$. 
 Suppose that the statement is false, this is, $\chsat \in \eqso(\loge{2})$.
-We consider the signature $\R$ that we used as the encoding for a Horn formula (Example \ref{ex-hornsat-esop1}) and that the formula $\alpha \in \eqso(\loge{2})$ follows the encoding in the same way.
-From what we proved in Theorem \ref{theo-pnf-snf}, we have that every formula in $\eqso(\loge{2})$ can be rewritten in $\loge{2}$-PNF, so we assume that $\alpha$ is in this form.
-Let $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\forall\bar{z}\,\varphi(\bar{X},\bar{x},\bar{y},\bar{z})$.
-Consider the following Horn formula $\Phi$:
+Consider the signature $\R$ from Example~\ref{ex-hornsat-esop1} and let $\alpha \in \eqso(\loge{2})$ be a formula over $\R$ that defines $\chsat$.
+By Proposition~\ref{theo-pnf-snf} we know that every formula in $\eqso(\loge{2})$ can be rewritten in $\loge{2}$-PNF, so we can assume that $\alpha$ is of the form $\sa{\bar{X}}\sa{\bar{x}} \ex{\bar{y}} \fa{\bar{z}}\varphi(\bar{X},\bar{x},\bar{y},\bar{z})$.
+Now, consider the following Horn formula:
 $$
-\Phi = p \wedge \bigwedge_{i = 1}^n (t_i \wedge p \to q) \wedge \neg q,
+\Phi \ = \ p \wedge \bigwedge_{i = 1}^n (t_i \wedge p \to q) \wedge \neg q,
 $$
-where $n = \length{\bar{x}} + \length{\bar{y}} + 1$.
-Let $\A$ be the encoding of this formula.
-In the encoding, each variable appears as an element in the domain of $\A$.
-This formula is satisfiable, so $\sem{\alpha}(\A) \geq 1$.
-Let $(\bar{B},\bar{b},\bar{a})$ be an assignment to $(\bar{X},\bar{x},\bar{y})$ such that $\A\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-Let $t_{\ell}$ be such that it does not appear in $\bar{b}$ or $\bar{a}$.
-Consider the induced substructure $\A'$ that is obtained by removing $t_{\ell}$ from $\A$ and $\bar{B}'$ as the subset of $\bar{B}$ obtained by deleting each appearance of $t_{\ell}$ in $\bar{B}$.
-We have that $\A'\models\forall\bar{z}\,\varphi(\bar{B},\bar{b},\bar{a},\bar{z})$.
-This is because each subformula of the form $\exists y \neg B_i$ is still true, and universal formulas are monotone over induced substructures.
-It follows that $\sem{\alpha}(\A') \geq 1$ which is not possible since $\A'$ encodes the formula
+such that $n = \length{\bar{x}} + \length{\bar{y}} + 1$ and let $\A_{\Phi}$ be the encoding of this formula over $\R$. 
+%In the encoding, each variable appears as an element in the domain of $\A$. 
+One can easily check that $\Phi$ is satisfiable, so $\sem{\alpha}(\A_{\Phi}) \geq 1$.
+Let $(\bar{B},\bar{b},\bar{a})$ be an assignment to $(\bar{X},\bar{x},\bar{y})$ such that $\A_{\Phi} \models \fa{\bar{z}} \varphi(\bar{B},\bar{b},\bar{a},\bar{z})$ and let $t_{\ell}$ be such that it does not appear in $\bar{b}$ or $\bar{a}$ (recall that $n > \length{\bar{x}} + \length{\bar{y}}$).
+Consider the induced substructure $\A_{\Phi}'$ that is obtained by removing $t_{\ell}$ from $\A_{\Phi}$ and $\bar{B}'$ as the subset of $\bar{B}$ obtained by deleting each appearance of $t_{\ell}$ in $\bar{B}$.
+We have that $\A_{\Phi}'\models \fa{\bar{z}}\varphi(\bar{B}',\bar{b},\bar{a},\bar{z})$ since universal formulas are monotone over induced substructures.
+Then it follows that $\sem{\alpha}(\A_{\Phi}') \geq 1$ which is not possible since $\A_{\Phi}'$ encodes the formula
 $$
@@
 $$
-which is unsatisfiable. We arrive to a contradiction and we conclude that $\chsat$ is not in $\eqso(\ehorn)$.
+which is unsatisfiable. This leads to a contradiction and we conclude that $\chsat$ is not in $\eqso(\ehorn)$.

### 2017-08-07 LMCS/proof-sigma2hard.tex
Lista sección 5.3.

@@
 Recall that each instance of $\shdhsat$ is a disjunction of Horn formulas.
-Let $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot), \pD(\cdot,\cdot)\}$.
+Let $\R$ be a relational signature such that $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot), \pD(\cdot,\cdot)\}$.
 Each symbol in this vocabulary is used to indicate the same as in Example \ref{ex-hornsat-esop1}, with the addition of $\pD(d,c)$ which indicates that $c$ is a clause in the formula $d$.
-Recall that the formula
+Recall that the formula $\sa{\pT} \sa{\pA} \psi(\pT,\pA)$
+defines $\chsat$. In order to encode $\shdhsat$, we extend $\psi(\pT,\pA)$ by adding the information of $\pD(d,c)$ as follows:
 \begin{align*}
-&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\textit{A}(c,x) \vee \neg\textit{T}(x)).
+\psi'(T,A) \ := \ \ex{d} \big[ \ & \fa{x} (\neg \pT(x) \vee \pV(x)) \ \wedge\\
+& \fa{c} (\neg \pD(c,d)\vee \neg \textit{NC}(c) \vee  \ex{x} \neg \textit{A}(c,x)) \ \wedge\\
+& \fa{c} \fa{x} (\neg \pD(c,d)\vee\neg \textit{P}(c,x) \vee \ex{y} \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
+& \fa{c} \fa{x} (\neg \pD(c,d)\vee\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
+& \fa{c} \fa{x}  (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
+& \fa{c} \fa{x} (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \neg\textit{T}(x)) \ \big]
 \end{align*}
-defines $\chsat$. We obtain the following formula $\psi(T,A)$ in $\ehorn$:
-\begin{align*}
-\exists d[&\forall x \, (\neg \pT(x) \vee \pV(x)) \ \wedge\\
-&\forall c \, (\neg \pD(c,d)\vee \neg \textit{NC}(c) \vee \exists x \, \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{P}(c,x) \vee \exists y \, \neg \textit{A}(c,y) \vee \textit{T}(x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\neg \textit{N}(c,x) \vee \textit{T}(x) \vee \neg \textit{A}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-&\forall c \forall x \, (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \neg\textit{T}(x))]
-\end{align*}
-effectively defines $\chsat$ as for every disjunction of Horn formulas $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi(\pT,\pA)}(\A)$.
+One can check that $\psi'(T,A)'$ effectively defines $\shdhsat$ as for every disjunction of Horn formulas $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi'(\pT,\pA)}(\A)$.
 Therefore, we conclude that $\shdhsat \in \eqso(\ehorn)$.
 
-\vspace{1em}
-We will now prove that $\shdhsat$ is hard for $\eqso$ over a signature $\R$ under parsimonious reductions.
-For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a polynomial-time procedure that computes a function $g_{\alpha}$.
-This function receives a finite $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
-We suppose that $\alpha$ is in sum normal form and:
+Next, we will prove that $\shdhsat$ is hard for $\eqso(\ehorn)$ over a signature~$\R$ under parsimonious reductions.
+For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a polynomial-time function $g_{\alpha}$ that  receives an $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
+By Proposition~\ref{theo-pnf-snf}, we can assume that $\alpha$ is of the form:
 $$
-\alpha = \sum_{i = 1}^{\text{\#clauses}} \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\bigwedge_{j = 1}^{n}\forall\bar{z}\,\varphi^i_j(\bar{X},\bar{x},\bar{y},\bar{z}),
+\alpha \ = \ \sum_{i = 1}^{m} \sa{\bar{X}}\sa{\bar{x}} \ex{\bar{y}} \bigwedge_{j = 1}^{n} \fa{\bar{z}} \varphi^i_j(\bar{X},\bar{x},\bar{y},\bar{z}),
 $$
-where each $\varphi^i_j$ is a Horn clause.
+where each $\varphi^i_j$ is a Horn-clause.
+Now, consider a finite $\R$-structure $\A$ with domain $A$. 
+The first transformation of $\alpha$ and $\A$ towards a disjunction of horn-formulas is to replace each first-order quantifier (i.e.
+$\Sigma{\bar{x}}$,  $\exists\bar{y}$, and $\forall\bar{z}$) with first-order constants.
+Specifically, the following formula defines the same function as $\alpha$ and is of polynomial size with respect to $\A$:
+$$
+\alpha_{\A} \ = \ \sum_{i = 1}^{m} \sum_{\bar{a}\in A^{\length{\bar{x}}}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length{\bar{y}}}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length{\bar{z}}}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
+$$
+Note that each first-order subformula in $\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c})$ has no free variable and, therefore, we can evaluate each of them in polynomial time and replace by $\perp$ and $\top$ wherever it corresponds.
+In other words, in polynomial time we can replace $\varphi^i_j$ by a conjunction of $\neg X_{h}$'s and at most one $X_{h}$.
+After simplifying, grouping and reordering the previous formula, we can obtain an equivalent formula $\alpha_{\A}'$ of the form:
+$$
+\alpha_{\A}' \ := \ \sum_{i = 1}^{m'}\sa{\bar{X}}\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi^{i}_{j,k}(\bar{X})
+$$
+where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{h}$'s and at most one $X_{h}$.
+The reader can easily check that $\sem{\alpha}(\A) = \sem{\alpha_{\A}'}(\A)$.
 
-Consider a finite $\R$-structure $\A$ with domain $A$.
-To simplify the proof, we extend our grammar to allow first-order constants.
-Consider each tuple $\bar{a}\in A^{\length{\bar{x}}}$, each $\bar{b}\in A^{\length{\bar{y}}}$ and each $\bar{c}\in A^{\length{\bar{z}}}$ as a tuple of first-order constants.
-The following formula defines the same function as $\alpha$:
+The idea for the rest of the proof is to show how to obtain $g_{\alpha}(\A)$, i.e. an instance of $\shdhsat$, from $\alpha_{\A}'$.
+First, if $\alpha_{\A}'$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi_{j,k}(\bar{X})$, then we can define $g_{\alpha}(\A)$ equal to the propositional formula $\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulas.
+Otherwise, if $m' > 1$ we can use $m'$ new fresh variables $t_1,\ldots,t_{m'}$ and define:
 $$
-\sum_{i = 1}^{\#clauses} \sum_{\bar{a}\in A^{\length{\bar{x}}}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length{\bar{y}}}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length{\bar{z}}}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
+g_{\alpha}(\A) \ := \ \bigvee_{i=1}^{m'} \, \bigvee_{j = 1}^{n_1'} \, \bigwedge_{k = 1}^{n_2'} \, \psi^i_{j,k}(\bar{X}) \wedge t_i \wedge \bigwedge_{l \neq i} \neg t_l
 $$ 
-Note that each $\fo$ formula over $(\bar{x},\bar{y},\bar{z})$ in each $\varphi^i_j$ has no free variables.
-Therefore, we can evaluate each of these in polynomial time and replace them by $\perp$ and $\top$ where it corresponds.
-Each $\varphi^i_j$ will be of the form $\perp \vee\, \chi^i_j(\bar{X})$ or $\top \vee \chi^i_j(\bar{X})$ where $\chi^i_j$ is a disjunction of $\neg X_{\ell}$'s and at most one $X_{\ell}$.
-The formulas of the form $\top \vee \chi^i_j(\bar{X})$ can be removed entirely, and the formulas of the form $\perp \vee\, \chi^i_j(\bar{X})$ can be replaced by $\chi^i_j(\bar{X})$.
-We obtain the formula
-$$
-\sum_{i = 1}^{m}\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})
-$$
-where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$.
-
-Our idea for the rest of the proof is to define $g_{\alpha}$ for each $\alpha = \sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi^{i}_{j,k}(\bar{X})$, for $\alpha = c$ and for $\alpha = \beta_1 + \cdots + \beta_m$ where each $\beta_i$ is in one of the two previous cases.
-
-If $\alpha$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ where $\psi_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$'s and zero or one $X_{\ell}$, then we obtain the {\bf propositional formula} $g_{\alpha}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulas.
-
-If $\alpha$ is equal to a constant $c$, then we define $g_{\alpha}(\A)$ as the following formula that has exactly $c$ satisfying assignments:
-$$
-g_{\alpha}(\A) = \bigvee_{i = 1}^{c}\neg t_1 \wedge \cdots \wedge \neg t_{i-1} \wedge t_i \wedge \neg t_{i+1} \wedge \cdots \wedge \neg p_c.
-$$ 
-If $\alpha = \beta_1 + \cdots + \beta_m$, let $g_{\beta_i}(\A) = \bigvee_{j = 1}^{\#d}\bigwedge_{k = 1}^{\#c}\theta^i_{j,k}$ for each $\beta_i$ where each $\theta^i_{j,k}$ is a Horn clause.
-Let $\Theta_i = g_{\beta_i}(\A)$.
-We rename the variables in each $\Theta_i$ so none of them are mentioned in any other $\Theta_j$.
-We add $m$ new variables $t_1,\ldots,t_m$ and we define:
-\begin{align*}
-g_{\alpha}(\A) = &\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^1_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m}}}t) \wedge (t_1 \wedge \bigwedge_{\ell = 2}^{m} \neg t_{\ell})) \vee \\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^2_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each $t$}\\ \text{ mentioned in}\\ \Theta_1,\Theta_3,\ldots,\Theta_{m}}}t) \wedge (t_2 \wedge \bigwedge\limits_{\substack{\ell = 1 \\ \ell \neq 2}}^{m} \neg t_{\ell})) \vee \cdots \vee\\
-&\bigvee_{i = 1}^{\#d}(\bigwedge_{j = 1}^{\#c}\theta^m_{i,j} \wedge (\bigwedge\limits_{\substack{\text{each } t\\ \text{ mentioned in}\\ \Theta_2,\ldots,\Theta_{m-1}}}t) \wedge (t_m \wedge \bigwedge_{\ell = 1}^{m-1} \neg t_{\ell})).
-\end{align*}
-The formula is a disjunction of Horn formulas, and the number of satisfying assignments for this formula is exactly the sum of satisfying assignments for each $g_{\beta_i}(\A)$.
-This, at the same time, is equal to $\sem{\alpha}(\A)$.
+over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \} \cup \{t_1,\ldots,t_{m'}\}$.
+Intuitively, variables $t_1,\ldots,t_{m'}$ are used to count from $1$ to $m'$ each subformula $\sa{\bar{X}}\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi^{i}_{j,k}(\bar{X})$ in $\alpha_{\A}'$.
+One can easily check that $g_{\alpha}(\A)$ is a disjunction of Horn formulas, and the number of satisfying assignments is exactly $\sem{\alpha}(\A)$.
 This covers all possible cases for $\alpha$, and the entire procedure takes polynomial time.

### 2017-08-07 LMCS/syntactic-horn.tex
Lista sección 5.3.

@@
 \end{prop}
-\proof
-\input{proof-ehorn-pe}
-\qed
+For the sake of simplification, we postpone the proof the previous proposition after the proof of Theorem~\ref{sigma2hard}.
+
 Interestingly, we have that both $\chsat$ and $\cdnf$ belong to $\eqso(\ehorn)$. 
@@
 \qed
+
+\noindent{\emph{Proof of Proposition~\ref{prop:ehorn-pe}.}}
+\input{proof-ehorn-pe}
+\qed

### 2017-08-07 LMCS/proof-sigma2hard.tex
Cambios menores.

@@
 \end{align*}
-One can check that $\psi'(T,A)'$ effectively defines $\shdhsat$ as for every disjunction of Horn formulas $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi'(\pT,\pA)}(\A)$.
+One can check that $\psi'(T,A)$ effectively defines $\shdhsat$ as for every disjunction of Horn formulas $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi'(\pT,\pA)}(\A)$.
 Therefore, we conclude that $\shdhsat \in \eqso(\ehorn)$.

### 2017-08-18 LMCS/proof-sigma2hard.tex
[journal] Cambios menores y arreglado error en la demo de sigma2hard

@@
+%!TEX root = main.tex
+
 First we prove that $\shdhsat$ is in $\eqso(\ehorn)$.
@@
 Each symbol in this vocabulary is used to indicate the same as in Example \ref{ex-hornsat-esop1}, with the addition of $\pD(d,c)$ which indicates that $c$ is a clause in the formula $d$.
-Recall that the formula $\sa{\pT} \sa{\pA} \psi(\pT,\pA)$
+Define $\psi$ as in Example \ref{ex-hornsat-esop1} such that $\sa{\pT} \sa{\pA} \psi(\pT,\pA)$
 defines $\chsat$. In order to encode $\shdhsat$, we extend $\psi(\pT,\pA)$ by adding the information of $\pD(d,c)$ as follows:
@@
 & \fa{c} \fa{x}  (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \textit{N}(c,x)) \ \wedge\\
-& \fa{c} \fa{x} (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \neg\textit{T}(x)) \ \big]
+& \fa{c} \fa{x} (\neg \pD(c,d)\vee\textit{A}(c,x) \vee \neg\textit{T}(x)) \ \big].
 \end{align*}
@@
 $$
-\alpha \ = \ \sum_{i = 1}^{m} \sa{\bar{X}}\sa{\bar{x}} \ex{\bar{y}} \bigwedge_{j = 1}^{n} \fa{\bar{z}} \varphi^i_j(\bar{X},\bar{x},\bar{y},\bar{z}),
+\alpha \ = \ \sum_{i = 1}^{m} \sa{\bar{X}_{i}}\sa{\bar{x}} \ex{\bar{y}} \bigwedge_{j = 1}^{n} \fa{\bar{z}} \varphi^i_j(\bar{X}_{i},\bar{x},\bar{y},\bar{z}),
 $$
-where each $\varphi^i_j$ is a Horn-clause.
+where each $\varphi^i_j$ is a Horn-clause, and each $\bar{X}_{i}$ is a sequence of second-order variables.
+Consider $\bar{X}$ as the union of all $\bar{X}_{i}$. We replace each of the $m$ addends in $\alpha$ for
+$$
+\sa{\bar{X}}\sa{\bar{x}} \ex{\bar{y}} \bigwedge_{j = 1}^{n} \fa{\bar{z}} \varphi^i_j(\bar{X}_{i},\bar{x},\bar{y},\bar{z})\wedge\bigwedge_{X\not\in\bar{X}_{i}}\fa{\bar{u}}X(\bar{u}),
+$$
+whose sum is equivalent to $\alpha$.
 Now, consider a finite $\R$-structure $\A$ with domain $A$. 
-The first transformation of $\alpha$ and $\A$ towards a disjunction of horn-formulas is to replace each first-order quantifier (i.e.
-$\Sigma{\bar{x}}$,  $\exists\bar{y}$, and $\forall\bar{z}$) with first-order constants.
-Specifically, the following formula defines the same function as $\alpha$ and is of polynomial size with respect to $\A$:
+The next transformation of $\alpha$ and $\A$ towards a disjunction of horn-formulas is to expand each first-order quantifier (i.e.
+$\Sigma{\bar{x}}$,  $\exists\bar{y}$, and $\forall\bar{z}$) as we replace their variables with first-order constants.
+Specifically, we obtain the following formula which defines the same function as $\alpha$ and is of polynomial size with respect to $\A$:
 $$
-\alpha_{\A} \ = \ \sum_{i = 1}^{m} \sum_{\bar{a}\in A^{\length{\bar{x}}}} \sa{\bar{X}}\bigvee_{\bar{b}\in A^{\length{\bar{y}}}}\bigwedge_{j = 1}^{n}\bigwedge_{\bar{c}\in A^{\length{\bar{z}}}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
+\alpha_{\A} \ = \ \sum_{i = 1}^{m} \sum_{\bar{a}\,\in A^{\length{\bar{x}}}} \sa{\bar{X}}\bigvee_{\bar{b}\,\in A^{\length{\bar{y}}}}\bigwedge_{j = 1}^{n'}\bigwedge_{\bar{c}\,\in A^{\length{\bar{z}}}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
 $$
-Note that each first-order subformula in $\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c})$ has no free variable and, therefore, we can evaluate each of them in polynomial time and replace by $\perp$ and $\top$ wherever it corresponds.
-In other words, in polynomial time we can replace $\varphi^i_j$ by a conjunction of $\neg X_{h}$'s and at most one $X_{h}$.
+Note that each first-order sub-formula in $\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c})$ has no free variables and, therefore, we can evaluate each of them in polynomial time and replace by $\perp$ and $\top$ wherever it corresponds.
+In other words, in polynomial time we can replace $\varphi^i_j$ by a disjunction of $\neg X_{h}$ and at most one $X_{h}$, evaluated on constants.
 After simplifying, grouping and reordering the previous formula, we can obtain an equivalent formula $\alpha_{\A}'$ of the form:
@@
 $$
-where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{h}$'s and at most one $X_{h}$.
+where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{h}$ and at most one $X_{h}$.
 The reader can easily check that $\sem{\alpha}(\A) = \sem{\alpha_{\A}'}(\A)$.
@@
 $$
-g_{\alpha}(\A) \ := \ \bigvee_{i=1}^{m'} \, \bigvee_{j = 1}^{n_1'} \, \bigwedge_{k = 1}^{n_2'} \, \psi^i_{j,k}(\bar{X}) \wedge t_i \wedge \bigwedge_{l \neq i} \neg t_l
+g_{\alpha}(\A) \ := \ \bigvee_{i=1}^{m'} \, \bigvee_{j = 1}^{n_1'} \, \bigwedge_{k = 1}^{n_2'} \, \psi^i_{j,k}(\bar{X}) \wedge t_i \wedge \bigwedge_{\ell \neq i} \neg t_{\ell}
 $$ 

### 2017-08-18 LMCS/proof-theo-pi1-pnf.tex
[journal] Cambios menores y arreglado error en la demo de sigma2hard

@@
 Second, we prove that the formula $\alpha_{1} = 2$ (i.e.
-the constant function $2$) is not equivalent to any formula in $\loge{1}$-PNF.
+$\sem{\alpha_{1}}$ is the constant function $2$) is not equivalent to any formula in $\loge{1}$-PNF.
 Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to $\alpha_1$.
@@
 Therefore, suppose that $\length{\bar{X}} > 0$ or $\length{\bar{x}} > 0$, and consider some ordered structure $\A$. 
-Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
+Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
 Now consider the ordered structure $\A'$ that is obtained by taking the disjoint union of $\A$ twice.
@@
 
-For the last part of the proof, we show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there exists an equivalent formula in $\LL$-PNF.
-Similar than in Theorem~\ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such a formula.
+For the latter part of the proof, we show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there exists an equivalent formula in $\LL$-PNF.
+Similarly to Theorem~\ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such a formula.
 Assume that $\alpha = \sum_{i = 1}^n \alpha_i$ is in $\LL$-SNF where each $\alpha_i$ is in $\LL$-PNF.
 Without loss of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
-If not, we can replace each $\alpha_i$ by the equivalent formula:
+If not, we can replace each $\alpha_i$ by the equivalent formula
 $$
@@
 Finally, consider a $\eqso(\LL)$ formula $\alpha$ in $\LL$-SNF.
-If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we can consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the hypothesis.
+If $\alpha = \sum_{i = 1}^n\alpha_i$, then by induction we consider $\alpha = \alpha_1 + (\sum_{i = 2}^n\alpha_i)$ and use $\tau(\alpha_1 + \tau(\sum_{i = 2}^n\alpha_i))$ as the rewrite of $\alpha$, which satisfies the hypothesis.
 %To show that the formula is indeed equivalent to $\alpha$, note that the formulas in lines (\ref{eq:partition1}) and (\ref{eq:partition2}) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$, and the formulas in lines (\ref{eq:partition3}) and (\ref{eq:partition4}) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.

### 2017-08-18 LMCS/saluja.tex
[journal] Cambios menores y arreglado error en la demo de sigma2hard

@@
 To answer this question, we first introduce two normal forms for $\eqso(\LL)$ that helps us to characterize the expressive power of this quantitative logic.
-A formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-prenex normal form ($\LL$-PNF)} 
+A formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-prenex normal form} ($\LL$-PNF) 
 if $\alpha$ is of the form
 $\sa{\bar{X}} \sa{\bar{x}} \varphi(\bar{X}, \bar{x})$,
-where $\bar{X}$ and $\bar{x}$ are sequences of zero or more second-order and first-order variables, respectively, and $\varphi(\bar{X}, \bar{x})$ is a formula in $\LL$. Notice that 
+where $\bar{X}$ and $\bar{x}$ are sequences of zero or more second-order and first-order variables, respectively, (as expected, $\sa{\bar{X}}\!$ is the respective nesting of $\sa{X}$'s) and $\varphi(\bar{X}, \bar{x})$ is a formula in $\LL$.
+Notice that
 a formula $\varphi(\bar{X}, \bar{x})$ in $\sh{\LL}$ is equivalent to the formula $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ in $\LL$-PNF. 
-Moreover, a formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-sum normal form ($\LL$-SNF)} if $\alpha$ is of the form $\Sigma_{i=1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-PNF. 
+Moreover, a formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-sum normal form} ($\LL$-SNF) if $\alpha$ is of the form $\sum_{i=1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-PNF. 
 \begin{prop}\label{theo-pnf-snf}

### 2017-09-27 LMCS/proof-theo-pi1-pnf.tex
[journal] Cambios menores en Teorema 5.2 (de reescribibilidad en PNF)

@@
 In \cite{SalujaST95}, it was proved that if $\length{\bar{X}} > 0$, the function defined by $\alpha$ is always even for big enough structures, which is not possible in our case.
-On the other hand, if $\alpha$ is of the form $\sa{\bar{x}}\varphi(\bar{x})$, then $\alpha$ defines a polynomially bounded function which also leads to a contradiction.
+On the other hand, if $\alpha$ is of the form $\sa{\bar{x}}\varphi(\bar{x})$, then $\alpha$ defines a polynomially bounded function which leads to a contradiction.
 

### 2017-09-27 LMCS/proof-sigma2hard.tex
[journal] Cambios menores en demo Sigma2Hard

@@
 
-Next, we will prove that $\shdhsat$ is hard for $\eqso(\ehorn)$ over a signature~$\R$ under parsimonious reductions.
+Next, we prove that $\shdhsat$ is hard for $\eqso(\ehorn)$ over a signature~$\R$ under parsimonious reductions.
 For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a polynomial-time function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
@@
 Now, consider a finite $\R$-structure $\A$ with domain $A$. 
-The next transformation of $\alpha$ and $\A$ towards a disjunction of horn-formulas is to expand each first-order quantifier (i.e.
+The next transformation of $\alpha$ and $\A$ towards a disjunction of Horn-formulas is to expand each first-order quantifier (i.e.
 $\Sigma{\bar{x}}$,  $\exists\bar{y}$, and $\forall\bar{z}$) as we replace their variables with first-order constants.

### 2017-09-27 LMCS/proof-prop-rest.tex
[journal] Simplificacion en la demo de Proposicion 5.3 (SigmaQSOSigma0 y #Sigma1 son imcomparables)

@@
 First, we show that $\E{1} \not\subseteq \QE{0}$.
-By contradiction, let $\R$ be the empty signature and suppose that there is a $\QE{0}$ formula $\alpha$ over $\R$ such that defines the following function: for every finite $\R$-structure with $n$ elements, $\alpha(\enc(\A)) = n - 1$.
-We need the following claim.
+Towards a contradiction, let $\R = \{ < \}$ and suppose that there is a $\QE{0}$ formula $\alpha$ over $\R$ which is equivalent to the $\E{1}$ formula $\sa{x} \ex{y} (x < y)$.
+This is, for every finite $\R$-structure $\A$ with $n$ elements, $\sem{\alpha}(\A) = n - 1$.
+We will use the following claim.
 \begin{clm}
-	Let $\alpha = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. Then the function defined by $\alpha$ is either null, greater or equal to $n$, or is in $\Omega(n^2)$.
+	Let $\alpha = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. Then $\alpha$ defines a function that is either null, greater or equal to $n$, or is in $\Omega(n^2)$.
 \end{clm}
 \proof
-	Suppose that the function defined by $\alpha$ is not $0$ and that $\varphi$ is in DNF, namely, $\alpha = \sa{\bar{x}} \varphi_1(\bar{x}) \vee \cdots \vee \varphi_n(\bar{x})$.
-Furthermore, suppose $\bar{x} = (x_1,\ldots,x_{\length{\bar{x}}})$.
-Since $\alpha$ is not null, then some $\varphi_i$ must be satisfiable and suppose that $\varphi_i(\bar{x}) = \bigwedge\psi(\bar{x})$ where each $\psi(\bar{x})$ is a literal (i.e.
-an atomic formula or its negation).
-	We will prove by induction on $\length{\bar{x}}$ that the function defined by $\sa{\bar{x}} \bigwedge\psi(\bar{x})$ is either greater or equal to $n$, or in $\Omega(n^2)$.
-	For $\length{\bar{x}}= 1$ and $\alpha = \sa{x}\bigwedge\psi(x)$, if any $\psi(x) = (x = x)$ or $\neg(x < x)$, then we can eliminate it and we obtain the same function.
-If any $\psi = (x < x)$ or $\neg(x=x)$, then the function becomes null.
-The only possible $\alpha$ left is $\alpha = \sa{x}\top$ which is equal to the function $n$.
-This covers all possible cases for $\length{\bar{x}} = 1$.
-	Now suppose that it holds for $\length{\bar{x}} = k$ and suppose $\alpha = \sa{\bar{x}}\bigwedge\psi(\bar{x})$ for $\length{\bar{x}} = k+1$.
-If any $\psi(\bar{x}) = (x_i = x_j)$ where $i \neq j$, then $\alpha$ describes the same function as $\alpha$ where $x_j$ has been replaced by $x_i$.
-In this formula the tuple of first-order variables has now $k$ elements so, by induction, the function satisfies our conclusions.
-Furthermore, if any $\psi(\bar{x}) = (x_i = x_j)$ or $\neg (x_i < x_j)$ where $i = j$, then we can eliminate $\psi(\bar{x})$ and obtain the same function, and if any $\psi(\bar{x}) = \neg(x_i = x_j)$ or $(x_i < x_j)$ where $i = j$, then the function becomes null.
-	The remaining formulas in $\bigwedge\psi(\bar{x})$ are either $\neg(x_i = x_j)$, $(x_i<x_j)$ or $\neg(x_i<x_j)$ with $i \neq j$.
-If the formula violates transitivity in $<$ (e.g.
-$x < y \wedge y < z \wedge z < x$), then $\alpha$ defines the null function.
-Therefore, there must be some order over $\bar{x}$ that satisfies $\bigwedge\psi(\bar{x})$ and this order can be chosen in  $\binom{n}{\length{\bar{x}}}$ different ways over the domain of $\A$.
-This implies that $\sem{\alpha} \in \Omega(n^{\length{\bar{x}}}) \subseteq \Omega(n^2)$ if $\length{\bar{x}} > 1$ and concludes the proof of the claim.
+	Note that each atomic sub-formula in $\varphi(\bar{x})$ has either the form $(x = y)$, $(x < y)$, $\top$ or a negation thereof, where $x,y\in\bar{x}$.
+Suppose $\sem{\alpha}$ is not null and consider some $\R$-structure $\A$ such that $\sem{\alpha}(\A) > 0$.
+Let $\bar{a}$ be an assignment to $\bar{x}$ such that $\A\models\varphi(\bar{a})$.
+It can be seen that each assignment $\bar{a}'$ that has the same order over its variables\footnote{For example, if $\bar{a} = (a_1,a_2,a_3,a_4)$, such order may be $a_1 > a_3 = a_4 > a_2$.} as $\bar{a}$ also satisfies $\A\models\varphi(\bar{a}')$.
+If this order has $k$ partitions then $\binom{n}{k}$ assignments for $\bar{x}$ satisfy this order, and therefore, $\sem{\alpha} \geq \binom{n}{k}$.
+If $k = 1$, then $\binom{n}{k} = n$, and if $k \geq 2$, then $\binom{n}{k} \in \Omega(n^2)$, which proves the claim.
 \qed
-Coming back on proving that $\alpha$ does not define the $(n-1)$-function, suppose that $\alpha$ is in SNF, namely, $\alpha = \sum_{i = 1}^k \alpha_i$ for some fix $k$.
+To prove that $\alpha$ does not define the $(n-1)$-function, suppose that $\alpha$ is in SNF, namely, $\alpha = \sum_{i = 1}^k \alpha_i$ for some fix $k$.
 Since $\alpha$ is not null, consider some $\alpha_i$ that describes a non-null function.
@@
 Note that if $\length{\bar{X}} > 0$, then the function $\sem{\alpha}$ is in $\Omega(2^n)$, as it was proven in~\cite{SalujaST95}.
-Therefore, we have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$ and, as we proved in the above claim, it describes either some function greater or equal to $n$, or in $\Omega(n^2)$, which leads to a contradiction.
-To conclude, note that the formula $\sa{x} \ex{y} (x < y)$ is in $\E{1}$ and describes the function $n-1$.
+Therefore, we have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$ and, as we proved in the claim above, it describes either some function greater or equal to $n$, or in $\Omega(n^2)$, which leads to a contradiction.
 

### 2017-10-03 LMCS/proof-qe0-fp-qe1-totp-fptras.tex
[journal] Simplficacion en demo de Lemma 5.7 (cerrado bajo multiplicacion)

@@
 The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$.
-This is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every finite $\R$-structure $\A$ with domain $A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$.
-We use this fact in the next arguments.
+This is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every $\R$-structure $\A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\size{\A})$.
+We use this fact in the following arguments.
 
@@
 Since $\cdnf$ is also in $\totp$ \cite{PagourtzisZ06}, we simulate the corresponding nondeterministic procedure that generates exactly $\cdnf(\enc(g(\A)))$ branches.
-Since, $\fp\subseteq\totp$\cite{PagourtzisZ06}, each polynomially computable function is also in $\totp$, and then on each of these branches we simulate the corresponding nondeterministic procedure to generate $h(\vert A \vert)$ more.
-The number of branches for each $\alpha_i$ is $\sem{\alpha_i}(\A) = \cdnf(\enc(g(\A)))\cdot h(\vert A \vert)$, and the total number of branches is equal to $\sem{\alpha}(\A)$.
+Since, $\fp\subseteq\totp$\cite{PagourtzisZ06}, each polynomially computable function is also in $\totp$, and then on each of these branches we simulate the corresponding nondeterministic procedure to generate $h(\size{\A})$ more.
+The number of branches for each $\alpha_i$ is $\sem{\alpha_i}(\A) = \cdnf(\enc(g(\A)))\cdot h(\size{\A})$, and the total number of branches is equal to $\sem{\alpha}(\A)$.
 We conclude that $\alpha\in\totp$.
@@
 Finally, we show that $\eqso(\loge{1})$ is closed under sum and multiplication.
-Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that is closed under multiplication.
+Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that it is closed under multiplication.
 We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of $\so$.
@@
 \proof
-	We define a recursive function $\tau$ that receives a formula $\alpha$ over the grammar of $\eqso(\LL)$ extended by binary product, and outputs an equivalent formula $\tau(\alpha)$ over the unextended grammar of $\eqso(\LL)$.
-In fact, the formula $\tau(\alpha)$ is in $\LL$-SNF.
-First we replace each constant $s$ with $(\top \add \cdots \add \top)$ ($s$ times).
-If $\alpha = \varphi$, then we define $\tau(\alpha) = \alpha$.
-We assume that for every $\beta$ that has less algebraic operators than $\alpha$, $\tau(\beta)$ is in $\LL$-SNF.
-If $\alpha = (\alpha_1 + \alpha_2)$ then we define $\tau(\alpha) = \tau(\alpha_1) + \tau(\alpha_2)$.
-If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$, then we define $\tau(\alpha)$ as the formula in $\LL$-SNF that is equivalent to $\sa{x}\tau(\beta)$ and to $\tau(\alpha) = \sa{X}\tau(\beta)$, respectively.
-If $\alpha = (\alpha_1 \cdot \alpha_2)$, we assume that each $\alpha_i$ is in $\LL$-SNF.
-We identify three cases.
-(1) Some $\alpha_i$ is equal to $\sum_{j = 1}^n\beta_j$ for $n > 1$.
-Suppose wlog.
-that it is $\alpha_1$.
-We then define $\tau(\alpha) = \sum_{j = 1}^n\tau(\beta_j\cdot\alpha_2)$.
-In the following cases, $\alpha_1$ and $\alpha_2$ are in $\LL$-SNF.
-(2) If some $\alpha_i$ is equal to $\sa{X}\beta$ or $\sa{x}\beta$, we define $\tau(\alpha)$ as the $\LL$-SNF formula that is equivalent to $\sa{x}\tau(\beta\cdot\alpha_2)$ and $\sa{X}\tau(\beta\cdot\alpha_2)$, respectively.
-The remaining case is (3) $\alpha_1 = \varphi_1$ and $\alpha_2 = \varphi_2$ where each $\varphi$ is an $\LL$ formula.
-Then we define $\tau(\alpha) = \varphi_1 \wedge \varphi_2$.
-This covers all possible cases for $\alpha$.
-For every pair of formulas $\alpha,\beta$ in $\eqso(\LL)$, we have that their multiplication $(\alpha\cdot\beta)$ is a formula in the grammar $\eqso(\LL)$ extended by binary product, and so, there exists an equivalent formula $\tau(\alpha\cdot\beta)$ which is in $\eqso(\LL)$.
+	Given two formulas $\alpha, \beta$ in $\eqso(\LL)$ we will show a formula in the grammar which is equivalent to $(\alpha\mult \beta)$.
+From what was proven in Proposition \ref{theo-pnf-snf}, we may assume that $\alpha$ and $\beta$ are in $\LL$-SNF.
+Let $\alpha = \sum_{i = 1}^n\sa{\bar{X}_i}\sa{\bar{x}_i}\varphi_i(\bar{X}_i,\bar{x}_i)$, and $\beta = \sum_{i = 1}^m\sa{\bar{Y}_i}\sa{\bar{y}_i}\psi_i(\bar{Y}_i,\bar{y}_i)$.
+Expanding the product in $(\alpha\mult \beta)$ and reorganizing results in the equivalent formula
+	$$
+	\sum_{i = 1}^n\sum_{j = 1}^m\sa{\bar{X}_i}\sa{\bar{Y}_j}\sa{\bar{x}_i}\sa{\bar{y}_j}(\varphi_i(\bar{X}_i,\bar{x}_i)\wedge\psi_i(\bar{Y}_j,\bar{y}_j)),
+	$$
+	which is in $\LL$-SNF, and therefore, in $\eqso(\LL)$.
 \qed
-Since $\loge{1}$ is closed under conjunction, then Lemma~\ref{conj-mult} also holds for $\eqso(\loge{1})$. This concludes the proof.
+Since $\loge{1}$ is closed under conjunction, then Lemma~\ref{conj-mult} holds for $\eqso(\loge{1})$. This concludes the proof.

### 2017-10-04 LMCS/proof-capture-fp.tex
[journal] cambios menores en demos

@@
-In this proof and the followings, we will reuse the symbol $<$ to denote the lexicographic order over same-sized tuples.
+In this and the following proofs, we will reuse the symbol $<$ to denote the lexicographic order over same-sized tuples. 
 Formally, for $\bar{x} = (x_1,\ldots,x_m)$ and $\bar{y} = (y_1,\ldots,y_m)$ we denote by $\bar{x} < \bar{y}$ the formula:
@@
 \]
-
+Furthermore, we use $\size{\A}$ to denote the size of an $\R$-structure $\A$.
 Now we prove Theorem~\ref{theo:capture-fp}. 
@@
 Note that the subformula $\pa{\bar{y}}(\bar{y} < \bar{x}) \mapsto 2$ takes the value $2^m$ if there exist $m$ tuples in $A^{\ell}$ that are smaller than $\bar{x}$.
-Summing these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
+Adding these values for each $\bar{a}\in A^{\ell}$ gives exactly $f(\enc(\A))$.
 In other words, $\Phi(\bar{x})$ simulates the behavior of the $\fp$-machine and the formula $\alpha$ reconstructs the binary output bit by bit.

### 2017-10-04 LMCS/proof-prop-rest.tex
[journal] cambios menores en demos

@@
 Towards a contradiction, let $\R = \{ < \}$ and suppose that there is a $\QE{0}$ formula $\alpha$ over $\R$ which is equivalent to the $\E{1}$ formula $\sa{x} \ex{y} (x < y)$. 
-This is, for every finite $\R$-structure $\A$ with $n$ elements, $\sem{\alpha}(\A) = n - 1$.
-We will use the following claim.
+This is, for every finite $\R$-structure $\A$, $\sem{\alpha}(\A) = \size{\A} - 1$.
+
+Suppose that $\alpha$ is in SNF, namely, $\alpha = \sum_{i = 1}^k \alpha_i$ for some fixed $k$. 
+Since $\alpha$ is not null, consider some $\alpha_i$ that describes a non-null function. 
+Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ where $\varphi$ is quantifier-free. 
+Note that if $\length{\bar{X}} > 0$, then the function $\sem{\alpha}$ is in $\Omega(2^{\size{\A}})$, as it was proven in~\cite{SalujaST95}. 
+Therefore, we have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$. 
+We conclude our proof with the following claim.
 \begin{clm}
-	Let $\alpha = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. Then $\alpha$ defines a function that is either null, greater or equal to $n$, or is in $\Omega(n^2)$.
+	Let $\alpha = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. 
+	Then the function $\sem{\alpha}$ is either null, greater or equal to $n$, or is in $\Omega(n^2)$, where $n$ is the size of the input structure.
 \end{clm}
 \proof
-	Note that each atomic sub-formula in $\varphi(\bar{x})$ has either the form $(x = y)$, $(x < y)$, $\top$ or a negation thereof, where $x,y\in\bar{x}$.
+	Note that each atomic sub-formula in $\varphi(\bar{x})$ is either $(x = y)$, $(x < y)$, $\top$ or a negation thereof, where $x,y\in\bar{x}$. 
 	Suppose $\sem{\alpha}$ is not null and consider some $\R$-structure $\A$ such that $\sem{\alpha}(\A) > 0$. 
 	Let $\bar{a}$ be an assignment to $\bar{x}$ such that $\A\models\varphi(\bar{a})$. 
-It can be seen that each assignment $\bar{a}'$ that has the same order over its variables\footnote{For example, if $\bar{a} = (a_1,a_2,a_3,a_4)$, such order may be $a_1 > a_3 = a_4 > a_2$.} as $\bar{a}$ also satisfies $\A\models\varphi(\bar{a}')$.
-If this order has $k$ partitions then $\binom{n}{k}$ assignments for $\bar{x}$ satisfy this order, and therefore, $\sem{\alpha} \geq \binom{n}{k}$.
-If $k = 1$, then $\binom{n}{k} = n$, and if $k \geq 2$, then $\binom{n}{k} \in \Omega(n^2)$, which proves the claim.
+	It can be seen that each assignment $\bar{a}'$ that has the same order over its variables\footnote{For example, if $\bar{a} = (a_1,a_2,a_3,a_4)$, such order may be $a_1 > a_3 = a_4 > a_2$, which has tree partitions.}	as $\bar{a}$ also satisfies $\A\models\varphi(\bar{a}')$.
+	If this order has $k$ partitions then $\binom{\size{\A}}{k}$ assignments for $\bar{x}$ satisfy this order, and therefore, $\sem{\alpha} \geq \binom{\size{\A}}{k}$. 
+	If $k = 1$, then $\binom{\size{\A}}{k} = \size{\A}$, and if $k \geq 2$, then $\binom{\size{\A}}{k} \in \Omega(\size{\A}^2)$, which proves the claim.
 \qed
-To prove that $\alpha$ does not define the $(n-1)$-function, suppose that $\alpha$ is in SNF, namely, $\alpha = \sum_{i = 1}^k \alpha_i$ for some fix $k$.
-Since $\alpha$ is not null, consider some $\alpha_i$ that describes a non-null function.
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ where $\varphi$ is quantifier-free.
-Note that if $\length{\bar{X}} > 0$, then the function $\sem{\alpha}$ is in $\Omega(2^n)$, as it was proven in~\cite{SalujaST95}.
-Therefore, we have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$ and, as we proved in the claim above, it describes either some function greater or equal to $n$, or in $\Omega(n^2)$, which leads to a contradiction.
 
@@
 Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\ex{\bar{y}}\varphi_i(\bar{X},\bar{x},\bar{y})$, where $\varphi_i$ is quantifier-free for each $\alpha_i$. 
-We use the same construction used in \cite{SalujaST95}, and we obtain that the formula $\ex{\bar{y}}\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \fa{\bar{y}'}(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')$ for every assignment to $(\bar{X},\bar{x})$.
-We do this replacement for each $\alpha_i$ and we obtain an equivalent formula in $\eqso(\logu{1})$.
+We use the same construction used in \cite{SalujaST95}, and we obtain that the formula $\ex{\bar{y}}\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,[\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \fa{\bar{y}'}(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')]$ for every assignment to $(\bar{X},\bar{x})$.
+Doing this replacement for each $\alpha_i$ renders an equivalent formula in $\eqso(\logu{1})$.
 
-To prove that the inclusion is proper, consider the $\eqso(\logu{1})$ formula $\sa{x} \fa{y}(y = x)$. This formula defines the following function over each ordered structure $\A$:
+To prove that the inclusion is proper, consider the $\eqso(\logu{1})$ formula $\sa{x} \fa{y}(y = x)$. 
+This formula defines the following function over each ordered structure $\A$:
 $$
@@
 Suppose that there exists an equivalent formula $\alpha$ in $\eqso(\loge{1})$. 
-Also, suppose that it is in $\L$-PNF, so $\alpha = c + \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\ex{\bar{y}}\varphi_i(\bar{X},\bar{x},\bar{y})$.
-Since $\alpha$ takes the value 0 for some structures, $c$ must be 0.
+Also, suppose that it is in $\L$-PNF, so $\alpha = \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\ex{\bar{y}}\varphi_i(\bar{X},\bar{x},\bar{y})$. 
 Consider a structure $\A'$ with one element. 

### 2018-05-06 LMCS/beyond.tex
journal 23 mar
Pasada sobre la seccion 6 y el resto del paper

@@
 the notion of least fixed point of LFP.
- As a proof of concept, we show that how this notion can be used to captures $\fp$.
+ As a proof of concept, we show that this notion can be used to captures $\fp$.
 Moreover, we use this concept
@@
 \end{align}
-where $h \in \fs$, $\arity(h) = \ell$ and $x_1, \ldots, x_\ell$ is a sequence of (non-necessarily distinct) first-order variables.
+where $h \in \fs$, $\arity(h) = \ell$ and $x_1, \ldots, x_\ell$ is a sequence of (not necessarily distinct) first-order variables.
 Given an $\R$-structure $\A$ with domain $A$, we say that $F$ is a \emph{function assignment} for $\A$ if for every $h \in \fs$ with $\arity(h) = \ell$, we have that $F(h) :  A^\ell \to \N$.
@@
 Furthermore, a function symbol can be seen as an ``oracle'' that is instantiated by the function assignment. 
-To the best of our knowledge, this is the first paper to propose this extension on weighted logics, which should be further investigated.
+To the best of our knowledge, this is the first article to propose this extension on weighted logics, which we think should be further investigated.
 
@@
 \begin{lem}\label{lem-support}
-If $f,g \in \F$ and $\support(f) \subseteq \support(g)$, then $\support(T_\beta(f)) \subseteq \support(T_\beta(g))$.
+Let $h \in \fs$ such that $\arity(h) = \ell$, and $\beta$ be an $\fqfo(\fo)$-formula over a signature $\R$ such that only $h$ can occur as a function symbol in $\beta$.
+Moreover, let $\A$ be an $\R$-structure with domain $A$, $f,g : A^\ell \to \mathbb{N}$ and $F,G$ be function assignments such that $F(h) = f$ and $F(h) = g$.
+If $\support(f) \subseteq \support(g)$, then for every first-order and second-order assignments $v$ and $V$, respectively, it holds that:
+\begin{center}
+if $\sem{\beta}(\A,v,V,F) > 0$, then $\sem{\beta}(\A,v,V,G) > 0$.
+\end{center}
 \end{lem}
@@
 \qed
-In fact, as for the case of LFP, this lemma gives us a simple way to compute a least s-fixed point of $T_\beta$.
+
+In the particular case of an $\rqfo(\fo)$-formula $\clfp{\beta(\x, h)}$, Lemma  \ref{lem-support} tell us that if $f,g \in \F$ and $\support(f) \subseteq \support(g)$, then $\support(T_\beta(f)) \subseteq \support(T_\beta(g))$.
+Hence, as for the case of LFP, this lemma gives us a simple way to compute a least s-fixed point of $T_\beta$.
 Let $f_0 \in \F$ be a function such that $f_0(\bar a) = 0$ for every $\bar a \in A^\ell$ (i.e.
@@
 We would like to define an $\rqfo(\fo)$-formula that, given a directed acyclic graph $G$ with $n$ nodes and a pair of nodes $b$, $c$ in $G$, counts the number of paths of length at most $n$ from $b$ to $c$ in $G$.
-To this end, assume that graphs are encoded using the signature $\R = \{ E(\cdot,\cdot), < \}$, and then define formula $\alpha(x, y, f)$ as follows:
+To this end, assume that graphs are encoded using the signature $\R = \{ E(\cdot,\cdot)\}$, and then define formula $\alpha(x, y, f)$ as follows:
 $$ 
@@
 Assume now that we need to extend our previous counting function to the case of arbitrary directed graphs.
-To this end, suppose that $\varphi_{\text{\rm first}}(x)$ and $\varphi_{\text{succ}}(x,y)$ are the $\fo$-formulae for defining the first and successor predicates, respectively, of $<$.
-Moreover, define formula $\beta(x, y, t, g)$ as follows:
+To this end, suppose that $\varphi_{\text{\rm first}}(x)$ and $\varphi_{\text{succ}}(x,y)$ are $\fo$-formulae defining the first element of $<$ and the successor relation associated to $<$, respectively.\footnote{Recall that in this paper we consider ordered $\R$-structures.} Moreover, define formula $\beta(x, y, t, g)$ as follows:
 \begin{align*}
@@
 Notice that the use of the filter $\varphi_{\text{succ}}(t',t)$ prevents this formula for incrementing the value of $g(x,y,t)$ when $t$ is the first element in the domain.
+\qed
 \end{exa}
 In contrast with $\lfp$, to reach a fixed point we do not need to impose any positive restriction on the formula $\beta(\x,h)$.
-Indeed, since $\beta$ is constructed from monotones operations (i.e. sum and product) over the natural numbers, the resulting operator $T_{\beta}$ is monotone as well.
+In fact, since $\beta$ is constructed from monotones operations (sum and product) over the natural numbers, the resulting operator $T_{\beta}$ is monotone as well.
 
@@
 Inspired by this work, we show that a restricted version of $\rqfo$ can be used to capture $\shl$, the counting version of $\nlog$. 
-Specifically, we use $\rqfo$ to define an operator for counting the number of paths in a directed graph, which is what is needed to capture $\shl$.
+Specifically, we use $\rqfo$ to define an operator for counting the number of paths in a directed graph, which is what is needed to capture~$\shl$.
 
@@
 Then the semantics of the path operator can be defined by using the following definition of $[\pth \psi(\bar{x},\bar{y})]$ in $\rqfo$:
-\[
-[\pth \psi(\bar{x}, \bar{y})] \ :=\ \sa{\t} (\varphi_{\text{\rm first}}(\t) \cdot \clfp{\beta_{\psi(\bar{x},\bar{y})}(\x,\y,\t,g)}).
-\]
+\begin{eqnarray*}
+[\pth \psi(\bar{x}, \bar{y})] & := & \sa{\t} (\varphi_{\text{\rm first}}(\t) \cdot \clfp{\beta_{\psi(\bar{x},\bar{y})}(\x,\y,\t,g)}).
+\end{eqnarray*}
 In other words, $\sem{[\pth \psi(\bar{x}, \bar{y})]}(\A,v)$ counts the number of paths from $v(\bar x)$ to $v(\bar y)$ in the graph $\cG_{\psi}(\A)$ whose length is at most~$|A^k|$.
-As it was previously said, the operator for counting paths is exactly what we need to capture $\shl$.
+As it was mentioned, the operator for counting paths is exactly what we need to capture $\shl$.
 \begin{thm} \label{tqfo-shl}
@@
 \qed
-This last result perfectly illustrates the benefits of our logical framework for the development of descriptive complexity for counting complexity classes.  
-The distinction in the language between the Boolean and the quantitative level allows us to define operators at the later level that cannot be defined at the former. 
-As a example showing how fundamental this separation is, consider the issue of extending $\qfo(\fo)$ at the Boolean level in order to capture $\shl$.
+%This last result perfectly illustrates the benefits of our logical framework for the development of descriptive complexity for counting complexity classes.  
+%The distinction in the language between the Boolean and the quantitative level allows us to define operators at the later level that cannot be defined at the former. 
+%As a example showing how fundamental this separation is, consider the issue of extending $\qfo(\fo)$ at the Boolean level in order to capture $\shl$.
 The natural alternative to do this is to use $\fo$ extended with a transitive closure operator, which is denoted by $\tc$.
@@
 This implies that $\nlog = \ulog$,\footnote{A decision language $L$ is in $\ulog$ is there exists a logarithmic-space NTM $M$ accepting $L$ and satisfying that $\tma_M(x) = 1$ for every $x \in L$.}
-and thus contradicts the widely believed assumption that $\ulog \subsetneq \nlog$.
+%and thus contradicts the widely believed assumption that $\ulog \subsetneq \nlog$.
+This last result 
+perfectly 
+illustrates the benefits of our logical framework for the development of descriptive complexity for counting %functions.
+complexity classes.  
+The distinction in the language between the Boolean and the quantitative level allows us to define operators at the latter level that cannot be defined at the former. 
+As a example showing how fundamental this separation is, consider the issue of extending $\qfo(\fo)$ at the Boolean level in order to capture $\shl$.
+The natural alternative to do this is to use $\fo$ extended with a transitive closure operator, which is denoted by $\tc$.
+But then the problem is that for every language $L \in \nlog$, it holds that its characteristic function $\chi_L$ is in $\qfo(\tc)$, where $\chi_L(x) = 1$ if $x \in L$, and $\chi_L(x) = 0$ otherwise.
+Thus, if we assume that $\qfo(\tc)$ captures $\shl$ (over ordered structures), then we have that $\chi_L \in \shl$ for every $L \in \nlog$.
+This would imply that $\nlog = \ulog$,\footnote{A decision language $L$ is in $\ulog$ is there exists a logarithmic-space NTM $M$ accepting $L$ and satisfying that $\tma_M(x) = 1$ for every $x \in L$.} solving an outstanding open problem \cite{Reinhardt97}.

### 2018-05-06 LMCS/proof-lem-support.tex
Pasada sobre la seccion 6 y el resto del paper

@@
-{\bf For }$\boldsymbol{\fqfo(\fo).}$ Let $\R$ be a signature. We prove the statement for $\fqfo(\fo)$ over $\R$.
+%{\bf For }$\boldsymbol{\fqfo(\fo).}$ Let $\R$ be a signature. We prove the statement for $\fqfo(\fo)$ over $\R$.
 
-We prove inductively that for each formula $\beta(\bar{x},h)$ in $\fqfo(\fo)$ over $\R$, for a given pair of functions $f,g$ such that $\supp(f)\subseteq\supp(g)$, it holds that $\supp(T_{\beta}(f))\subseteq\supp(T_{\beta}(g))$.
+We prove the lemma by induction on the structure of $\beta$. First we need to consider the base cases.
+% inductively that for each formula $\beta(\bar{x},h)$ in $\fqfo(\fo)$ over $\R$, for a given pair of functions $f,g$ such that $\supp(f)\subseteq\supp(g)$, it holds that $\supp(T_{\beta}(f))\subseteq\supp(T_{\beta}(g))$.
 Let $\length{\bar{x}} = \ell$.
-
-We separate the proof in each case determined by the $\fqfo$ grammar. For each of the following cases.
-\begin{itemize}
-\item[1.] $\beta$ is either equal to a constant $s$ or an $\fo$ formula $\varphi$.
-Then $h$ does not appear.
+%We separate the proof in each case determined by the $\fqfo$ grammar. For each of the following cases.
+\begin{enumerate}
+\item Assume that $\beta$ is either a constant $s \in \mathbb{N}$ or an $\fo$-formula $\varphi$.
+In both cases function symbol $h$ is not mentioned, so $\sem{\beta}(\A,v,V,F) = \sem{\beta}(\A,v,V,G)$ and it trivially holds that if $\sem{\beta}(\A,v,V,F) > 0$, then $\sem{\beta}(\A,v,V,G) > 0$.
+%Then $h$ does not appear.
 Then, for each structure $\A$, each first-order assignment $v$ and functional assignments $F,G$ over $\A$, we have that $\sem{\beta(\bar{x},h)}(\A,v,F) = \sem{\beta(\bar{x},h)}(\A,v,G)$.
 As a result, $\supp(T_{\beta}(f)) = \supp(T_{\beta}(g))$ for every pair of functions $f,g$.
-\item[2.] $\beta$ is equal to $h(\bar{y})$ for some subtuple $\bar{y}$ of $\bar{x}$. Then $T_{\beta}(f) = f$ and $T_{\beta}(g) = g$ and the condition holds trivially.
-\end{itemize}
-Suppose that the statement holds for each formula smaller than $\beta$.
-\begin{itemize}
-\item[3.] $\beta = (\beta_1 + \beta_2)$.
-It is easy to see that for each $\bar{a} \in A^{\ell}$ and function $f:A^{\ell}\to\nat$: $T_{\beta}(f)(\bar{a}) = T_{\beta_1}(f)(\bar{a}) + T_{\beta_2}(f)(\bar{a})$.
+
+\item Assume that $\beta$ is equal to $h(y_1, \ldots, y_\ell)$, where $y_1$, $\ldots$, $y_\ell$ is a sequence of (non-necessarily pairwise distinct) variables.
+Let $\bar a = (v(y_1), \ldots, v(y_\ell))$.
+Then we have that $\sem{\beta}(\A,v,V,F) = F(h)(v(y_1), \ldots, v(y_\ell)) = f(\bar a)$ and $\sem{\beta}(\A,v,V,G) = g(\bar a)$.
+Given that $\supp(f) \subseteq \supp(g)$, if $f(\bar a) > 0$, then $g(\bar a) > 0$.
+Hence, we conclude that if $\sem{\beta}(\A,v,V,F) > 0$, then $\sem{\beta}(\A,v,V,G) > 0$.
+\end{enumerate}
+We now consider the inductive steps. Assume that the property holds for $\fqfo(\fo)$-formulae $\beta_1$, $\beta_2$ and $\delta$
+%Suppose that the statement holds for each formula smaller than $\beta$.
+\begin{enumerate}
+\setcounter{enumi}{2}
+\item Assume that $\beta = (\beta_1 + \beta_2)$. If $\sem{\beta}(\A,v,V,F) > 0$, then 
+$\sem{\beta_1}(\A,v,V,F) > 0$ or $\sem{\beta_2}(\A,v,V,F) > 0$.
+Thus, by induction hypothesis we conclude that $\sem{\beta_1}(\A,v,V,G) > 0$ or $\sem{\beta_2}(\A,v,V,G) > 0$.
+Hence, we have that $\sem{\beta}(\A,v,V,G) > 0$.
+%It is easy to see that for each $\bar{a} \in A^{\ell}$ and function $f:A^{\ell}\to\nat$: $T_{\beta}(f)(\bar{a}) = T_{\beta_1}(f)(\bar{a}) + T_{\beta_2}(f)(\bar{a})$.
 Suppose $\supp(f)\subseteq\supp(g)$ and let $\bar{a} \in \supp(T_{\beta}(f))$, or in other words, $T_{\beta}(f)(\bar{a}) > 0$.
@@
 From the supposition we have that $T_{\beta_i}(g)(\bar{a}) > 0$ from which the statement follows.
-\item[4.] $\beta = (\beta_1 \mult \beta_2)$.
+
+\item Assume that $\beta = (\beta_1 \mult \beta_2)$. If $\sem{\beta}(\A,v,V,F) > 0$, then 
+$\sem{\beta_1}(\A,v,V,F) > 0$ and $\sem{\beta_2}(\A,v,V,F) > 0$.
+Thus, by induction hypothesis we conclude that $\sem{\beta_1}(\A,v,V,G) > 0$ and $\sem{\beta_2}(\A,v,V,G) > 0$.
+Hence, we have that $\sem{\beta}(\A,v,V,G) > 0$.
+
+%\item[4.] $\beta = (\beta_1 \mult \beta_2)$.
 It is easy to see that for each $\bar{a}$ in $A^{\ell}$ and function $f:A^{\ell}\to\nat$: $T_{\beta}(f)(\bar{a}) = T_{\beta_1}(f)(\bar{a}) \mult T_{\beta_2}(f)(\bar{a})$.
@@
 From the supposition we have that $T_{\beta_i}(g)(\bar{a}) > 0$ for both $\beta_i$ and the statement holds.
-\item[5.] $\beta = \sa{y}\delta(y,\bar{x},h)$.
-Here we extend the grammar slightly to allow constants, and we use the notation $\delta[a/y]$ to denote the formula obtained by replacing each instance of $y$ by the constant $a$.
+
+\item Suppose that $\beta = \sa{x} \delta$.
+Then we have that $\sem{\beta}(\A,v,V,F) = \sum_{a \in A} \sem{\delta}(\A,v[a/x],V,F)$ and $\sem{\beta}(\A,v,V,G) = \sum_{a \in A} \sem{\delta}(\A,v[a/x],V,G)$.
+Thus, if we assume that $\sem{\beta}(\A,v,V,F) > 0$, then there exists $a \in A$ such that $\sem{\delta}(\A,v[a/x],V,F) > 0$.
+Hence, by induction hypothesis we have that $\sem{\delta}(\A,v[a/x],V,G) > 0$ and, therefore, we conclude that $\sem{\beta}(\A,v,V,G) >0$.
+
+%Here we extend the grammar slightly to allow constants, and we use the notation $\delta[a/y]$ to denote the formula obtained by replacing each instance of $y$ by the constant $a$.
 It can be seen that $T_{\beta}(f)(\bar{a}) = \sum_{a \in A} T_{\delta[a/y]}(f)(\bar{a})$.
@@
 The statement now follows as in the case 3.
-\item[6.] $\beta = \pa{y}\delta(y,\bar{x},h)$. 
-It can be seen that $T_{\beta}(f)(\bar{a}) = \prod_{a \in A} T_{\delta[a/y]}(f)(\bar{a})$. 
+
+\item Suppose that $\beta = \pa{x} \delta$.
+Then we have that $\sem{\beta}(\A,v,V,F) = \prod_{a \in A} \sem{\delta}(\A,v[a/x],V,F)$ and $\sem{\beta}(\A,v,V,G) = \prod_{a \in A} \sem{\delta}(\A,v[a/x],V,G)$.
+Thus, if we assume that $\sem{\beta}(\A,v,V,F) > 0$, then $\sem{\delta}(\A,v[a/x],V,F) > 0$ for every $a \in A$.
+Hence, by induction hypothesis we have that $\sem{\delta}(\A,v[a/x],V,G) > 0$ for every $a \in A$, and, therefore, we conclude that $\sem{\beta}(\A,v,V,G) >0$.
+
+%\item[6.] $\beta = \pa{y}\delta(y,\bar{x},h)$. 
+%It can be seen that $T_{\beta}(f)(\bar{a}) = \prod_{a \in A} T_{\delta[a/y]}(f)(\bar{a})$. 
 %If $T_{\beta}(f)(\bar{a}) > 0$, then $T_{\delta[y/a]}(f)(\bar{a}) > 0$ for each $a\in A$. As in the $\mult$ case, the statement follows directly.
-The statement follows using the same argument from cases 4 and 5.
-\end{itemize}
-This covers all possible cases for $\beta$ and we finish the proof of the statement for $\fqfo(\fo)$.
-
-\vspace{1em}
-{\bf For} $\boldsymbol{\rqfo(\fo).}$ The only additional case is where $\beta = \clfp{\delta(\bar{y},h')}$ for some subtuple $\bar{y}$ of $\bar{x}$.
+%The statement follows using the same argument from cases 4 and 5.
+\end{enumerate}
+%This covers all possible cases for $\beta$ and we finish the proof of the statement for $\fqfo(\fo)$.
+%
+%\vspace{1em}
+%{\bf For} $\boldsymbol{\rqfo(\fo).}$ The only additional case is where $\beta = \clfp{\delta(\bar{y},h')}$ for some subtuple $\bar{y}$ of $\bar{x}$.
 We have that $\beta$ does not mention $h$, and so, the statement follows directly as we showed in the previous part of the proof.

### 2018-05-06 LMCS/proof-rqfo-fo-cap.tex
Pasada sobre la seccion 6 y el resto del paper

@@
-Given the definition of the semantics of $\rqfo(\fo)$, it is clear that a fixed formula $\clfp{\beta(\x, h)}$ can be evaluated in polynomial time, from which it is possible to conclude that each fixed formula in $\rqfo(\fo)$ can be evaluated in polynomial time.
-Thus, to prove that $\rqfo(\fo)$ captures $\fp$, we only need to prove the second condition in Definition \ref{def:cap}.
+%!TEX root = main.tex
 
@@
 \newcommand{\successor}{\text{succ}}
+\newcommand{\out}{\text{\it out}}
+
+Given the definition of the semantics of $\rqfo(\fo)$, it is clear that a fixed formula $\clfp{\beta(\x, h)}$ can be evaluated in polynomial time, from which it is possible to conclude that each fixed formula in $\rqfo(\fo)$ can be evaluated in polynomial time.
+Thus, to prove that $\rqfo(\fo)$ captures $\fp$, we only need to prove the second condition in Definition \ref{def:cap}.
 
@@
 These operations are represented by the set $\{0,1,\emptyset\}$.
-Finally, assume that $M$, on input $\enc(\A)$ with domain $A = \{1,\dots,n\}$, executes exactly $n^k$ steps for some $k \geq 1$.
+Finally, assume that $M$, on input $\enc(\A)$ with domain $A = \{1,\dots,n\}$, executes exactly $n^k$ steps for a fixed $k \geq 1$.
 
@@
 \begin{align*}
-\alpha = \sa{\bar{t}}\clfp{out(\bar{t}): \,&\alpha_{T_0}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out),\\
-	&\alpha_{T_1}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out),\\
-	&\alpha_{T_{\ttB}}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out),\\
-	&\alpha_{T_{\vdash}}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out),\\
-	&\alpha_{h}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out),\\
-	&\alpha_{\hat{h}}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out),\\
-	&\alpha_{s_{q_0}}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out),\\
+\alpha = \sa{\bar{t}}\clfp{\out(\bar{t}): \,&\alpha_{T_0}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out),\\
+	&\alpha_{T_1}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out),\\
+	&\alpha_{T_{\ttB}}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out),\\
+	&\alpha_{T_{\vdash}}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out),\\
+	&\alpha_{h}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out),\\
+	&\alpha_{\hat{h}}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out),\\
+	&\alpha_{s_{q_0}}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out),\\
 	&\vdots \\
-	&\alpha_{s_{q_{\ell}}}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out),\\
-	&\alpha_{out}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out)}\mult \last(\bar{t}).
+	&\alpha_{s_{q_{\ell}}}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out),\\
+	&\alpha_{\out}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out)}\mult \last(\bar{t}).
 \end{align*}
-Function $T_0$ is used to indicate whether the content of a cell of the working tape is 0 at some point of time, that is, $T_0(\bar{t},\bar{x}) > 0$ if the cell at position $\bar{x}$ of the working tape contains the symbol 0 at step $\bar{t}$, and $T_0(\bar{t},\bar{x}) = 0$ otherwise.
+In this formula, $T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h}$ are functions symbols of arity $2 \cdot k$, while $s_{q_0},\ldots,s_{q_{\ell}},\out$ are function symbols of arity $k$.
+For each one of these function symbols $f$, there is a formula $\alpha_f$ defining how the values of $f$ are updated when computing the fixed point.
+For example, $\alpha_{T_0}$ is used to define the values of function $T_0$.
+Notice that the values of each function $f$ in $\alpha$ depend on the values of the other functions, that is why we talk about a simultaneous computation.
+Besides, notice that notation $\clfp{\out(\bar{t}) : \ldots}$ is used to indicate that (1) the free variables of the formula are $\bar t = (t_1, \ldots, t_k)$, and (2) once the least fixed point has been computed, the value of $\clfp{\out(\bar{t}) : \ldots}$ for an assignment $\bar a$ to $\bar t$ is given by the value of function $\out$ on $\bar a$ in the fixed point.
+Finally, it is important to notice that $\alpha$ can be transformed into a proper $\rqfo(\fo)$ formula by using the same techniques to prove that a simultaneous LFP can be transformed into a single LFP~\cite{L04}; in particular, functions symbols $T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out$ are replaced by a single function $f$ of arity $6 \cdot 2 \cdot k + (\ell + 1) \cdot k + k = (\ell+ 14) \cdot k$.
+
+In the formula $\alpha$, function $T_0$ is used to indicate whether the content of a cell of the working tape is 0 at some point of time, that is, $T_0(\bar{t},\bar{x}) > 0$ if the cell at position $\bar{x}$ of the working tape contains the symbol 0 at step $\bar{t}$, and $T_0(\bar{t},\bar{x}) = 0$ otherwise.
 Functions $T_1$, $T_{\ttB}$ and $T_{\vdash}$ are defined analogously.
 Function $h$ is used to indicate whether the head of the working tape is in some position at some point of time, that is, $h(\bar{t},\bar{x}) > 0$ if the head of the working tape is at position $\bar{x}$ at step $\bar{t}$, and $h(\bar{t},\bar{x}) = 0$ otherwise.
-Function $\hat{h}$ is used to indicate whether the head of the working tape is {\bf not} in some position at some point of time, that is, $\hat{h}(\bar{t},\bar{x}) > 0$ if the head of the working tape is {\bf not} at position $\bar{x}$ at step $\bar{t}$, and $h(\bar{t},\bar{x}) = 0$ otherwise.
+Function $\hat{h}$ is used to indicate whether the head of the working tape is {\bf not} in some position at some point of time, that is, $\hat{h}(\bar{t},\bar{x}) > 0$ if the head of the working tape is {\bf not} at position $\bar{x}$ at step $\bar{t}$, and $\hat{h}(\bar{t},\bar{x}) = 0$ otherwise.
 For each $i \in \{0, \ldots, \ell\}$, function
-$s_{q_i}$ is used to indicate whether the TM $M$ is in state $q_i$ at some point of time, that is, $s_{q_i}(\bar{t}) > 0$ if the TM $M$ is in state $q_i$ at step $\bar{t}$, and $s_{q_i}(\bar{t},\bar{x}) = 0$ otherwise.
-Finally, function $out$ stores the output of the TM $M$; in particular, $out(\bar t)$ is the value returned by $M$ when $\bar t$ is the last step (that is, when $\last(\bar t)$ holds).
+$s_{q_i}$ is used to indicate whether the TM $M$ is in state $q_i$ at some point of time, that is, $s_{q_i}(\bar{t}) > 0$ if the TM $M$ is in state $q_i$ at step $\bar{t}$, and $s_{q_i}(\bar{t}) = 0$ otherwise.
+Function $\out$ stores the output of the TM $M$; in particular, $\out(\bar t)$ is the value returned by $M$ when $\bar t$ is the last step.
+Finally, assuming that $\varphi_{\text{\rm last}}(x)$ is an $\fo$-formulae defining the last element of $<$, we have that $\last(\bar t) = \bigwedge_{i=1}^n \varphi_{\text{\rm last}}(t_i)$, that is, $\last(\bar t)$ holds if $\bar t$ is the last step.
+Therefore, the use of $\last(\bar t)$ in $\alpha$ allows us to return the output of the TM $M$.
 
-Formulas $\alpha_{T_0}$, $\alpha_{T_1}$, $\alpha_{T_{\ttB}}$ and $\alpha_{T_{\vdash}}$ are defined as follows, assuming that $\bar y = (y_1, \ldots, y_k)$:
+As in Example \ref{ex:count-path}, assume that $\varphi_{\text{\rm first}}(x)$ and $\varphi_{\text{succ}}(x,y)$ are $\fo$-formulae defining the first element of $<$ and the successor relation associated to $<$, respectively.
+Moreover, assume that formulae $\varphi_{\text{succ}}(x,y)$ and $\varphi_{\text{\rm last}}(x)$ are used to defined
+an $\fo$-formula $\succesor(\bar{t},\bar{t'})$ such that $\succesor(\bar{t},\bar{t'})$ holds if $\bar t'$ is the successor step of $\bar t$ (that is, $\succesor(\cdot,\cdot)$ is the successor relation associated to the lexicographical order induced by $<$ on the tuples with $k$ elements).
+Then formula $\alpha_{T_{\vdash}}$ is defined as follows:
+\begin{multline*}
+\alpha_{T_{\vdash}}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out) = \\
+(\first(\bar{t}) \wedge \first(\bar{x})) +
+\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult \hat{h}(\bar{t}',\bar{x}) \mult T_{\vdash}(\bar{t}',\bar{x})) \ + \\
+\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,\vdash,\_,\_)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')).
+\end{multline*}
+Notation $\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,\vdash,\_,\_)}$ in this formula is used to indicate that we are considering a sum over all pairs $(q,a) \in Q \times \{0,1,\ttB,\vdash\}$ such that $\delta(q,a) = (q',\vdash,u,v)$ for some $q' \in Q$, $u \in \{\leftarrow,\rightarrow\}$ and $v \in \{0,1,\emptyset\}$.
+Formulas $\alpha_{T_0}$, $\alpha_{T_1}$ and $\alpha_{T_{\ttB}}$ are defined analogously; for the sake of completeness, we only show here how $\alpha_{T_0}$ is defined, assuming that $\bar y = (y_1, \ldots, y_k)$:
+\begin{multline*}
+\alpha_{T_0}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out) = \\ 
+(\first(\bar{t}) \wedge \exists\bar{y}(\varphi_{\text{\rm first}}(y_1) \wedge \cdots \wedge \varphi_{\text{\rm first}}(y_{k-2}) \wedge\neg E(y_{k-1},y_k) \wedge \successor(\bar{y},\bar{x}) ))\ + \\
+\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult \hat{h}(\bar{t}',\bar{x}) \mult T_0(\bar{t}',\bar{x})) + \\
+\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,0,\_,\_)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')).
+%\alpha_{T_1}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\
+%&(\first(\bar{t}) \wedge \exists\bar{y}(\first(y_1,\ldots,y_{k-2})\wedge E(y_{k-1},y_k) \wedge \successor(\bar{y},\bar{x}) ))+ \\
+%&\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult \hat{h}(\bar{t}',\bar{x}) \mult T_1(\bar{t}',\bar{x})) + \\
+%&\bigplus_{\delta(q,a) = (q',1,op,v)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')),\\
+%\alpha_{T_{\ttB}}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\
+%&(\first(\bar{t})\wedge\exists\bar{y}\exists\bar{y}'(\first(y_1,\ldots,y_{k-2})\wedge\last(y_{k-1},y_k)\wedge\successor(\bar{y},\bar{y}')\wedge\bar{y}' < \bar{x})) + \\
+%&\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult \hat{h}(\bar{t}',\bar{x}) \mult T_{\ttB}(\bar{t}',\bar{x})) + \\
+%&\bigplus_{\delta(q,a) = (q',{\ttB},op,v)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')),\\
+\end{multline*}
+Notice that in the initial step, the encoding of the structure $\A$ has to be placed on the working tape of $M$ from the second position since the symbol $\vdash$ is placed in the first position.
+This is the reason why we use tuple $\bar y$ and define $\bar x$ as the successor of $\bar y$.
+Formulas $\alpha_{h}$ is defined as follows:
 \begin{align*}
-\alpha_{T_0}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\ 
-&(\first(\bar{t}) \wedge \exists\bar{y}(\first(y_1,\ldots,y_{k-2})\wedge\neg E(y_{k-1},y_k) \wedge \successor(\bar{y},\bar{x}) ))+ \\
-&\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult \hat{h}(\bar{t}',\bar{x}) \mult T_0(\bar{t}',\bar{x})) + \\
-&\bigplus_{\delta(q,a) = (q',0,op,v)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')),\\
-\alpha_{T_1}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\
-&(\first(\bar{t}) \wedge \exists\bar{y}(\first(y_1,\ldots,y_{k-2})\wedge E(y_{k-1},y_k) \wedge \successor(\bar{y},\bar{x}) ))+ \\
-&\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult \hat{h}(\bar{t}',\bar{x}) \mult T_1(\bar{t}',\bar{x})) + \\
-&\bigplus_{\delta(q,a) = (q',1,op,v)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')),\\
-\alpha_{T_{\ttB}}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\
-&(\first(\bar{t})\wedge\exists\bar{y}\exists\bar{y}'(\first(y_1,\ldots,y_{k-2})\wedge\last(y_{k-1},y_k)\wedge\successor(\bar{y},\bar{y}')\wedge\bar{y}' < \bar{x})) + \\
-&\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult \hat{h}(\bar{t}',\bar{x}) \mult T_{\ttB}(\bar{t}',\bar{x})) + \\
-&\bigplus_{\delta(q,a) = (q',{\ttB},op,v)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')),\\
-\alpha_{T_{\vdash}}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\
-&(\first(\bar{t}) \wedge \first(\bar{x})) +
-\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult \hat{h}(\bar{t}',\bar{x}) \mult T_{\vdash}(\bar{t}',\bar{x})) + \\
-&\bigplus_{\delta(q,a) = (q',\vdash,op,v)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')).
+\alpha_{h}(\bar{t},\bar{x},&T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out) = \\
+&(\first(\bar{t}) \wedge \successor(\bar{t},\bar{x})) + \\
+&\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,\_,\leftarrow,\_)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult \successor(\bar{x},\bar{x}')\mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')) \ + \\
+&\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,\_,\rightarrow,\_)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult \successor(\bar{x}',\bar{x})\mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')).
 \end{align*}
-Formulas $\alpha_{h}$ and $\alpha_{\hat{h}}$ are defined as:
+Similarly, formula $\alpha_{\hat{h}}$ is defined as follows:
 \begin{align*}
-\alpha_{h}(&\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\
-& (\first(\bar{t}) \wedge \successor(\bar{t},\bar{x})) + \\
-&\bigplus_{\delta(q,a) = (q',b,\leftarrow,v)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult \successor(\bar{x},\bar{x}')\mult T_a(\bar{t}',\bar{x}') \mult h(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')) + \\
-&\bigplus_{\delta(q,a) = (q',b,\rightarrow,v)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult \successor(\bar{x}',\bar{x})\mult T_a(\bar{t}',\bar{x}') \mult h(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')),\\
-\alpha_{\hat{h}}(&\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\
-& (\first(\bar{t}) \wedge \neg\successor(\bar{t},\bar{x})) + \\
-&\bigplus_{\delta(q,a) = (q',b,\leftarrow,v)}\sa{\bar{t}'}\sa{\bar{x}'}\sa{\bar{x}''}(\successor(\bar{t}',\bar{t}) \mult T_a(\bar{t}',\bar{x}')  \mult h(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult \successor(\bar{x}'',\bar{x}') \mult (\bar{x} \neq \bar{x}'')) + \\
-&\bigplus_{\delta(q,a) = (q',b,\rightarrow,v)}\sa{\bar{t}'}\sa{\bar{x}'}\sa{\bar{x}''}(\successor(\bar{t}',\bar{t}) \mult T_a(\bar{t}',\bar{x}')  \mult h(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult \successor(\bar{x}',\bar{x}'') \mult (\bar{x} \neq \bar{x}'')).
+\alpha_{\hat{h}}(\bar{t},\bar{x},&T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out) = \\
+&(\first(\bar{t}) \wedge \neg\successor(\bar{t},\bar{x})) + \\
+&\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,\_,\leftarrow,\_)}\sa{\bar{t}'}\sa{\bar{x}'}\sa{\bar{x}''}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \, \mult \\
+&\hspace{250pt} \successor(\bar{x}'',\bar{x}') \mult (\bar{x} \neq \bar{x}'')) \ + \\
+&\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,\_,\rightarrow,\_)}\sa{\bar{t}'}\sa{\bar{x}'}\sa{\bar{x}''}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \, \mult \\
+&\hspace{250pt} \successor(\bar{x}',\bar{x}'') \mult (\bar{x} \neq \bar{x}'')).
 \end{align*}
+Notice that in the definition of $\alpha_{\hat{h}}$, $\fo$-formula $\bar{x} \neq \bar{x}''$ is used to indicate that tuples $\bar{x}$ and $\bar{x}''$ are distinct (that is, there exists $i \in \{1, \ldots, k\}$ such that the $i$-th components of $\bar{x}$ and $\bar{x}''$ are distinct).
 Formula $\alpha_{q_0}$ is defined as:
+\begin{multline*}
+\alpha_{q_0}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out) \ = \ \first(\bar{t}) \ + \\
+\bigplus_{(q,a) \,:\, \delta(q,a) = (q_0,\_,\_,\_)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult  s_{q}(\bar{t}')).
+\end{multline*}
+Moreover, for every $i \in \{1, \ldots, \ell\}$, formula $\alpha_{q_i}$ is defined analogously.
+%\begin{align*}
+%\alpha_{q_i}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\ &\bigplus_{\delta(q,a) = (q_i,b,op,v)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult T_a(\bar{t}',\bar{x}') \mult h(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')).
+%\end{align*}
+Finally, formula $\alpha_{\out}$ is defined as:
 \begin{align*}
-\alpha_{q_0}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \first(\bar{t}) + \\
-&\bigplus_{\delta(q,a) = (q_0,b,op,v)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult T_a(\bar{t}',\bar{x}') \mult h(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')).
+\alpha_{\out}(&\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out) =\\
+&\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,\_,\_,0)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult 2\mult \out(\bar{t}')) \ + \\
+&\bigplus_{(q,a) \,:\,\delta(q,a) = (\_,\_,\_,1)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult (2\mult \out(\bar{t}')+1)) \ + \\
+&\bigplus_{(q,a) \,:\,\delta(q,a) = (\_,\_,\_,\emptyset)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult \out(\bar{t}')).
 \end{align*}
-Moreover, for every $i \in \{1, \ldots, \ell\}$, formula $\alpha_{q_i}$ is defined as:
-\begin{align*}
-\alpha_{q_i}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\ &\bigplus_{\delta(q,a) = (q_i,b,op,v)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult T_a(\bar{t}',\bar{x}') \mult h(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')).
-\end{align*}
-Finally, formula $\alpha_{out}$ is defined as:
-\begin{align*}
-\alpha_{out}(\bar{t},&T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) =\\
-&\bigplus_{\delta(q,a) = (q',b,op,0)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult 2\mult out(\bar{t}')) + \\
-&\bigplus_{\delta(q,a) = (q',b,op,1)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult (2\mult out(\bar{t}')+1)) + \\
-&\bigplus_{\delta(q,a) = (q',b,op,\emptyset)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult out(\bar{t}')).
-\end{align*}
-Clearly, at each iteration of the LSFP operator, the tuple $\bar{t}$ represents the step the machine is currently in.
-From the construction of the formula, and since the machine is deterministic, it can be seen that in each function $g\in\{T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h}\}$, at the $\bar{a}$-th iteration of the LSFP operator, it holds that $g(\bar{a},\bar{b}) \leq 1$ for each $\bar{b}\in A^k$, that $g(\bar{a}+1,\bar{b}) = 0$ for each $\bar{b}\in A^k$.
-Also, at the $\bar{a}$-th iteration, $g(\bar{a}) \leq 1$ and $g(\bar{a}+1) = 0$ for each $g\in\{s_{q_1},\ldots,s_{q_{\ell}}\}$.
-From this, we have that at each iteration $\bar{a}$ of the operator, $out(\bar{a})$ is equal to either $2\cdot out(\bar{a}-1)$, $2\cdot out(\bar{a}-1) + 1$, or $out(\bar{a}-1)$, which represents precisely the value in the output tape at each step of $M$ running on input $\enc(\A)$.
-From this argument, it can be seen that $\sem{\alpha}(\A) = f(\enc(\A))$.
-
-\medskip
-
-To conclude the proof, we show that for each formula $\alpha$ in the previously defined extension of $\rqfo(\fo)$, there exists an equivalent formula confirming to the grammar of $\rqfo(\fo)$ defined in Section \ref{sec:beyond}.
+Clearly, at each iteration of the fixed point operator, the tuple $\bar{t}$ represents the step the machine is currently in.
+Assume that $\bar{a}$, $\bar{a}'$ are tuples with $k$ elements such that $\bar{a}'$ is the successor of $\bar{a}$.
+From the construction of the formula, and since the machine is deterministic, it can be seen that for each function $f \in\{T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h}\}$, at the $\bar{a}$-th iteration of the fixed point operator, it holds that $f(\bar{a},\bar{b}) \leq 1$ and  $f(\bar{a}',\bar{b}) = 0$ for every $\bar{b}\in A^k$.
+In the same way, it can be seen that for each function $g \in\{s_{q_1},\ldots,s_{q_{\ell}}\}$, at the $\bar{a}$-th iteration of the fixed point operator, it holds that $g(\bar{a}) \leq 1$ and $g(\bar{a}') = 0$.
+From this, we have that at the iteration $\bar{a}'$ of the fixed point operator, $\out(\bar{a}')$ is equal to either $2\cdot \out(\bar{a})$, $2\cdot \out(\bar{a}') + 1$, or $\out(\bar{a}')$, which represents precisely the value in the output tape at each step of $M$ running on the input $\enc(\A)$.
+From this argument, it can be seen that $\sem{\alpha}(\A) = f(\enc(\A))$, which concludes the proof of the theorem.
+%
+%\medskip
+%
+%To conclude the proof, we show that for each formula $\alpha$ in the previously defined extension of $\rqfo(\fo)$, there exists an equivalent formula confirming to the grammar of $\rqfo(\fo)$ defined in Section \ref{sec:beyond}.
 It suffices to consider a formula $\alpha$ of the form
-$$
-\alpha(\bar{x}_1) = \clfp{f_1(\bar{x}_1): \alpha_1(\bar{x}_1,f_1,\ldots,f_n),\alpha_2(\bar{x}_2,f_1,\ldots,f_n),\ldots,\alpha_n(\bar{x}_n,f_1,\ldots,f_n)},
-$$
-and show an equivalent formula defined by a LSFP operator which uses one formula less in its definition.
-
-We construct the equivalent formula as follows.
+%$$
+%\alpha(\bar{x}_1) = \clfp{f_1(\bar{x}_1): \alpha_1(\bar{x}_1,f_1,\ldots,f_n),\alpha_2(\bar{x}_2,f_1,\ldots,f_n),\ldots,\alpha_n(\bar{x}_n,f_1,\ldots,f_n)},
+%$$
+%and show an equivalent formula defined by a LSFP operator which uses one formula less in its definition.
+%
+%We construct the equivalent formula as follows.
 We use a new function symbol $f$ with arity $\length{\bar{x}_1} + \length{\bar{x}_2}$.
 For every $i \in \{1,\ldots,n\}$, let $\alpha_i'$ be the formula obtained by performing the following replacements in $\alpha_i$:
-\begin{align*}
-f_1(\bar{y}_1) &\text{ is replaced by } \sa{\bar{y}_2}f(\bar{y}_1,\bar{y}_2)\mult[\first(\bar{y}_1)\mult\last(\bar{y}_2)  \add (\neg\first(\bar{y}_1))\mult\first(\bar{y}_2)], \\
-f_2(\bar{y}_2) &\text{ is replaced by } \sa{\bar{y}_1}f(\bar{y}_1,\bar{y}_2)\mult[\first(\bar{y}_1)\mult\first(\bar{y}_2)  \add \last(\bar{y}_1)\mult(\neg\first(\bar{y}_2))].
-\end{align*}
-Moreover, let $\beta$ be a formula defined as:
-\begin{align*}
-\beta(\bar{x}_1,\bar{x}_2) = \,&\alpha_1'(\bar{x}_1)\mult(\first(\bar{x}_1)\mult\last(\bar{x}_2)  \add (\neg\first(\bar{x}_1))\mult\first(\bar{x}_2)) \add \\ &\alpha_2'(\bar{x}_2)\mult(\first(\bar{x}_1)\mult\first(\bar{x}_2)  \add \last(\bar{x}_1)\mult(\neg\first(\bar{x}_2))).
-\end{align*}
-It can be seen that all values of $f_1$, besides the first one, are stored in the first assignment of $\bar{x}_2$, while the first value of $f_1$ is stored in the last assignment of $\bar{x}_2$.
+%\begin{align*}
+%f_1(\bar{y}_1) &\text{ is replaced by } \sa{\bar{y}_2}f(\bar{y}_1,\bar{y}_2)\mult[\first(\bar{y}_1)\mult\last(\bar{y}_2)  \add (\neg\first(\bar{y}_1))\mult\first(\bar{y}_2)], \\
+%f_2(\bar{y}_2) &\text{ is replaced by } \sa{\bar{y}_1}f(\bar{y}_1,\bar{y}_2)\mult[\first(\bar{y}_1)\mult\first(\bar{y}_2)  \add \last(\bar{y}_1)\mult(\neg\first(\bar{y}_2))].
+%\end{align*}
+%Moreover, let $\beta$ be a formula defined as:
+%\begin{align*}
+%\beta(\bar{x}_1,\bar{x}_2) = \,&\alpha_1'(\bar{x}_1)\mult(\first(\bar{x}_1)\mult\last(\bar{x}_2)  \add (\neg\first(\bar{x}_1))\mult\first(\bar{x}_2)) \add \\ &\alpha_2'(\bar{x}_2)\mult(\first(\bar{x}_1)\mult\first(\bar{x}_2)  \add \last(\bar{x}_1)\mult(\neg\first(\bar{x}_2))).
+%\end{align*}
+%It can be seen that all values of $f_1$, besides the first one, are stored in the first assignment of $\bar{x}_2$, while the first value of $f_1$ is stored in the last assignment of $\bar{x}_2$.
 Moreover, all values of $f_2$, besides the first one, are stored in the last assignment of $\bar{x}_1$, while the first value of $f_2$ is stored in the first assignment of $\bar{x}_1$.
-%This schema is depicted in Table \ref{table-clfp}. 
-We use formula $\beta$ to define the following formula:
-\begin{align*}
-\alpha'(\bar{x}_1) = \,&\sa{\bar{x}_2}\clfp{f(\bar{x}_1,\bar{x}_2): \beta(\bar{x}_1,\bar{x}_2,f,f_3,\ldots,f_n),\\
-&\alpha'_3(\bar{x}_3,f,f_3,\ldots,f_n),\ldots,\alpha'_n(\bar{x}_n,f,f_3,\ldots,f_n)}\mult(\first(\bar{x}_1)\mult\last(\bar{x}_2)  \add (\neg\first(\bar{x}_1))\mult\first(\bar{x}_2))
-\end{align*}
-
-It is not difficult to see that $\alpha'(\bar{x}_1)$ is equivalent to $\alpha(\bar{x}_1)$, which concludes the proof.
+%%This schema is depicted in Table \ref{table-clfp}. 
+%We use formula $\beta$ to define the following formula:
+%\begin{align*}
+%\alpha'(\bar{x}_1) = \,&\sa{\bar{x}_2}\clfp{f(\bar{x}_1,\bar{x}_2): \beta(\bar{x}_1,\bar{x}_2,f,f_3,\ldots,f_n),\\
+%&\alpha'_3(\bar{x}_3,f,f_3,\ldots,f_n),\ldots,\alpha'_n(\bar{x}_n,f,f_3,\ldots,f_n)}\mult(\first(\bar{x}_1)\mult\last(\bar{x}_2)  \add (\neg\first(\bar{x}_1))\mult\first(\bar{x}_2))
+%\end{align*}
+%
+%It is not difficult to see that $\alpha'(\bar{x}_1)$ is equivalent to $\alpha(\bar{x}_1)$, which concludes the proof.

### 2018-05-06 LMCS/proof-tqfo-shl.tex
Pasada sobre la seccion 6 y el resto del paper

@@
-$\boldsymbol{\tqfo(\fo)}$ {\bf can be computed in} $\boldsymbol{\shl.}$
-Let $\R$ be some relational signature.
-Let $\alpha$ be a formula in $\tqfo(\fo)$.
-We will construct a nondeterministic logspace algorithm $M_{\alpha}$ that on input $\enc(\A)$, where a first-order assignment $v$ is being stored in memory, accepts in $\sem{\alpha}(\A,v)$ paths.
-Suppose the domain of $\A$ is $A = \{1,\ldots,n\}$.
-The algorithm needs $c\mult\log_2(n)$ bits of memory to store $v$, where $c$ is the total number of first-order variables in $\alpha$.
-If $\alpha = \varphi$, we check if $(\A,v)\models\varphi$ in deterministic logarithmic space, and accept if and only if it does.
-If $\alpha = s$, we generate $s$ branches and accept in all of them.
+%!TEX root = main.tex
+
+%&$\boldsymbol{\tqfo(\fo)}$ {\bf can be computed in} $\boldsymbol{\shl.}$
+First, we show that every formula in $\tqfo(\fo)$ defines a function that is in $\shl$.
+Let $\R$ be a relational signature and $\alpha$ a formula over $\R$ in $\tqfo(\fo)$.
+Next we construct a logarithmic-space nondeterministic Turing Machine $M_{\alpha}$ that on input $(\enc(\A),v)$, where $\A$ is an $\R$-structure and $v$ is a first-order assignment for $\A$, has $\sem{\alpha}(\A,v)$ accepting runs (so that we can conclude that the function defined by $\alpha$ is in $\shl$).
+Suppose that the domain of $\A$ is $A = \{1,\ldots,n\}$.
+The TM $M_{\alpha}$ needs $\ell \mult\log_2(n)$ bits of memory to store the first-order variables occurring in $\alpha$, where $\ell$ is the number of variables occurring in this formula (which is the same as the number of variables in the domain of $v$).
+If $\alpha = \varphi$, where $\varphi$ is an $\fo$-formula, then we check if $(\A,v)\models\varphi$ in deterministic logarithmic space, and accept if and only if this condition holds.
+If $\alpha = s$, where $s$ is a fixed natural number, then we generate $s$ possible runs and accept in all of them.
 If $\alpha = (\alpha_1 + \alpha_2)$, we simulate $M_{\alpha_1}$ and $M_{\alpha_2}$ on separate branches.
-If $\alpha = (\alpha_1\mult\alpha_2)$, we simulate $\alpha_1$ and if it accepts, instead of doing so, we simulate $\alpha_2$.
-If $\alpha = \sa{x}\beta$, for each $a\in A$ we generate a different branch where we simulate $M_{\beta}$ while storing $v[a/x]$.
-If $\alpha = \pa{x}\beta$, we simulate $M_{\beta}$ while storing $v[1/n]$, and on each accepting branch, instead of accepting we replace the assignment on $x$ to 2, to simulate $M_{\beta}$ while storing $v[2/x]$, and so on.
-If $\alpha = [\pth \varphi(\bar{x},\bar{y})]$ where $\varphi$ is an $\fo$ formula, we simulate the $\shl$ procedure that counts the number of paths for a graph of a given size.
-This procedure starts by setting $\bar{a} = v(\bar{x})$.
-On each iteration, nondeterministically chooses an assignment $\bar{a}$ for $\bar{x}$, continues if $(\A,v)\models\varphi(\bar{a}',\bar{a})$ where $\bar{a}'$ is the previously chosen value for $\bar{a}$, and rejects otherwise.
-If at any point we obtain that $\bar{a} = v(\bar{y})$, we generate an accepting branch, and continue simulating the procedure in the current branch.
+If $\alpha = (\alpha_1\mult\alpha_2)$, we simulate $M_{\alpha_1}$ and if it accepts, then instead of accepting we simulate $M_{\alpha_2}$.
+If $\alpha = \sa{x}\beta$, for each $a\in A$ we generate a different run where we simulate $M_{\beta}$ with input $v[a/x]$.
+If $\alpha = \pa{x}\beta$, we simulate $M_{\beta}$ with input $v[1/n]$, and on each accepting run, instead of accepting we replace the assignment of $x$ to 2, to simulate $M_{\beta}$ with input $v[2/x]$, and so on.
+If $\alpha = [\pth \varphi(\bar{x},\bar{y})]$, where $\varphi$ is an $\fo$-formula, then we simulate the $\shl$ procedure that counts the number of paths of a given length from a source to a target node in an input graph (where the length is at most the number of nodes in the graph).
+%This procedure starts by setting $\bar{a} = v(\bar{x})$.
+On each iteration, it nondeterministically chooses an assignment $\bar{b}$ for $\bar{x}$, and continues if $(\A,v)\models\varphi(\bar{a},\bar{b})$ where $\bar{a}$ is the previously chosen value for $\bar{x}$, and it rejects otherwise.
+If at any point we obtain that the current value for $\bar x$ is  $\bar{a}$ and $\bar a = v(\bar{y})$, we generate an accepting branch, and continue simulating the procedure in the current branch.
 We simulate $n^{\length{\bar{x}}}$ iterations of the procedure, and this generates exactly $\sem{[\pth \varphi(\bar{x},\bar{y})]}(\A,v)$ accepting branches.
@@
 
-\vspace{1em}
-$\boldsymbol{\shl}$ {\bf can be modelled in }$\boldsymbol{\tqfo(\fo).}$ Let $f$ be a function in $\shl$ and let $M$ be a nondeterministic logspace machine such that $\acc_{M}(\enc(\A)) = f(\enc(\A))$.
-We assume that $M$ has only one accepting state and upon accepting it immediately stops.
-Moreover, we assume that there exists only one accepting configuration altogether.
+%\vspace{1em}
+%$\boldsymbol{\shl}$ {\bf can be modelled in }$\boldsymbol{\tqfo(\fo).}$ 
+Second, we show that every function in $\shl$ can be encoded by a formula in $\tqfo(\fo)$.
+Let $f$ be a function in $\shl$ and $M$ a logarithmic-space nondeterministic  Turing Machine such that $\tma_M(\enc(\A)) = f(\enc(\A))$.
+We assume that $M$ has only one accepting state, and that no transition is defined for this state.
+Moreover, we assume that there exists only one accepting configuration.
 We make use of transitive closure logic ($\tc$) to simplify our proof~\cite{G07}.
-We have that $\tc$ captures $\nlog$\cite{I83}, so there exists a formula such that $\A\models\varphi$ if and only if $M$ accepts $\enc(\A)$.
+We have that $\tc$ captures $\nlog$\cite{I83}, so that there exists a formula $\varphi$ in $\tc$ such that $\A\models\varphi$ if and only if $M$ accepts $\enc(\A)$.
 This formula can be expressed as:
@@
 $$
-where $\psi_{\bf acc}(\bar{z})$ is an $\fo$ formula that expresses that $\bar{z}$ is an accepting configuration, and $\psi_{\bf next}(\bar{x},\bar{y})$ is an $\fo$ formula that expresses that $\bar{y}$ is the next configuration from $\bar{x}$~\cite{G07}.
-Here, there is a 1-1 correspondence between configurations of $M$ and assignments to $\bar{z}$.
-As a consequence, given a structure $\A$, and a first-order assignment $v$ to $\A$ where $v(\bar{x})$ is the starting configuration and $v(\bar{y})$ is the sole accepting configuration, the value of $\sem{[\pth\psi_{\bf next}(\bar{x},\bar{y})]}(\A,v)$ is equal to $\acc_M(\enc(\A))$.
-Finally, we define the $\tqfo(\fo)$ formula
-$$
-\alpha = \sa{\bar{u}}\sa{\bar{z}}(\first(\bar{u})\mult\psi_{\bf acc}(\bar{z})\mult[\pth \psi_{\bf next}(\bar{u},\bar{z})]),
-$$
-which satisfies $\sem{\alpha}(\A) = f(\enc(\A))$ for each structure $\A$. This concludes the proof.
+where $\psi_{\bf acc}(\bar{z})$ is an $\fo$-formula that indicates that $\bar{z}$ is an accepting configuration, and $\psi_{\bf next}(\bar{x},\bar{y})$ is an $\fo$-formula that indicates that $\bar{y}$ is a successor configuration of $\bar{x}$~\cite{G07}.
+Here, there is a one-to-one correspondence between configurations of $M$ and assignments to $\bar{z}$.
+As a consequence, given a structure $\A$ and a first-order assignment $v$ for $\A$, where $v(\bar{x})$ is the starting configuration and $v(\bar{y})$ is the sole accepting configuration, the value of $\sem{[\pth\psi_{\bf next}(\bar{x},\bar{y})]}(\A,v)$ is equal to $\tma_M(\enc(\A))$.
+Therefore, the $\tqfo(\fo)$-formula
+$
+\alpha = \sa{\bar{u}}\sa{\bar{z}}(\first(\bar{u})\mult\psi_{\bf acc}(\bar{z})\mult[\pth \psi_{\bf next}(\bar{u},\bar{z})])
+$
+satisfies that $\sem{\alpha}(\A) = f(\enc(\A))$. This concludes the proof of the theorem.

### 2018-05-07 LMCS/beyond.tex
proofreading of Section 6

@@
 Then the set of $\fqso$ formulae over a signature $\R$ is defined by the following grammar:
-\begin{align}
+\begin{multline}
 \label{eq-fqso}
-	\alpha &:= \varphi \ \mid \  s \  \mid \  h(x_1, \ldots, x_\ell) \  \mid \
-	(\alpha \add \alpha) \  \mid\  (\alpha \mult \alpha) \  \mid 
+	\alpha \ :=  \ \varphi \ \mid \  s \  \mid \  h(x_1, \ldots, x_\ell) \  \mid \
+	(\alpha \add \alpha) \  \mid\  (\alpha \mult \alpha) \  \mid \\
 	\sa{x} \alpha \  \mid \
@@
 	\pa{X} \alpha,
-\end{align}
+\end{multline}
 where $h \in \fs$, $\arity(h) = \ell$ and $x_1, \ldots, x_\ell$ is a sequence of (not necessarily distinct) first-order variables.
@@
 \begin{lem}\label{lem-support}
-Let $h \in \fs$ such that $\arity(h) = \ell$, and $\beta$ be an $\fqfo(\fo)$-formula over a signature $\R$ such that only $h$ can occur as a function symbol in $\beta$.
+Let $h \in \fs$ such that $\arity(h) = \ell$, and $\beta$ be an $\fqfo(\fo)$-formula over a signature $\R$ such that if a function symbol occurs in $\beta$, then this function symbol is $h$.
 Moreover, let $\A$ be an $\R$-structure with domain $A$, $f,g : A^\ell \to \mathbb{N}$ and $F,G$ be function assignments such that $F(h) = f$ and $F(h) = g$.
@@
 \begin{align*}
-\sem{\clfp{\beta(\x, h)}}(\A,v) &= f_{k}(v(\x))
+\sem{\clfp{\beta(\x, h)}}(\A,v) &= f_{k}(v(\x)).
 \end{align*}
@@
 In other words, $\sem{[\pth \psi(\bar{x}, \bar{y})]}(\A,v)$ counts the number of paths from $v(\bar x)$ to $v(\bar y)$ in the graph $\cG_{\psi}(\A)$ whose length is at most~$|A^k|$.
-As it was mentioned, the operator for counting paths is exactly what we need to capture $\shl$.
+As it was mentioned before, the operator for counting paths is exactly what we need to capture $\shl$.
 \begin{thm} \label{tqfo-shl}

### 2018-05-07 LMCS/introduction.tex
[journal] Cambios menores y arreglado error en la demo de sigma2hard
Listo para enviar ...

@@
 
+It is important to mention that this paper is an extension of the conference article \cite{AMR17}.
+In this version, we have included the complete proofs of all the results in the paper, paying special attention in showing the main techniques used to establish them.
+Besides, we have simplified some of the terminology used in \cite{AMR17}, with the goal of presenting the main notions studied in the paper in a simple way.
+
+\smallskip
 \smallskip
@@
 Finally, we give some concluding remarks in Section~\ref{sec:conclusions}.
-Due to the lack of space, the %complete 
-proofs of the results of the paper can be found at 
-%a separate appendix that can be downloaded from 
-\verb+http://marenas.sitios.ing.uc.cl/ls17.pdf+
+%Due to the lack of space, the proofs of the results of the paper can be found at a separate appendix that can be downloaded from \verb+http://marenas.sitios.ing.uc.cl/ls17.pdf

### 2018-05-07 LMCS/main.tex
journal 8 mar
journal 23 mar
Paso completo a sección 4.
Pasada sobre la seccion 6 y el resto del paper
Listo para enviar ...

@@
 %% mandatory lists of keywords 
-\keywords{MANDATORY list of keywords}
+%\keywords{MANDATORY list of keywords}
 
@@
 
+%\allowdisplaybreaks
+
 %% due to the dependence on amsart.cls, \begin{document} has to occur
@@
 \title[Counting]{Descriptive Complexity for Counting Complexity Classes}
-\titlecomment{{\lsuper*}OPTIONAL comment concerning the title, \eg, 
-  if a variant or an extended abstract of the paper has appeared elsewhere.}
+%\titlecomment{{\lsuper*}OPTIONAL comment concerning the title, \eg, 
+%  if a variant or an extended abstract of the paper has appeared elsewhere.}
 
 \author[M.~Arenas]{Marcelo Arenas}	%required
-\address{PUC Chile}	%required
+\address{Pontificia Universidad Cat\'olica de Chile}	%required
 \email{marenas@ing.puc.cl}  %optional
@@
 \author[M.~Munoz]{Martin Mu\~noz}	%optional
-\address{PUC Chile}	%optional
+\address{Pontificia Universidad Cat\'olica de Chile}	%optional
 \email{mmunos@uc.cl}  %optional
@@
 \author[C.~Riveros]{Cristian Riveros}	%optional
-\address{PUC Chile}	%optional
+\address{Pontificia Universidad Cat\'olica de Chile}	%optional
 \email{cristian.riveros@uc.cl}  %optional
@@
 
-\subsection{Extending $\qso$ to capture non-counting classes} \label{sec:extentions}
+\subsection{Extending $\qso$ to capture classes beyond counting} \label{sec:extentions}
 \input{extentions}
@@
 
-\begin{thebibliography}{10}
-
-\end{thebibliography}
+\bibliographystyle{plain}
+\bibliography{biblio}
 

### 2018-05-07 LMCS/send/proof-capture-fpspace.tex
En el directorio send esta la version enviada a arXiv

For the first condition of Definition~\ref{def:cap}, notice that each $\pfp$ formula can be evaluated in deterministic polynomial space, the constant function $s$ can be trivially simulated in $\fpspace$, and $\fpspace$ is closed under exponential sum and multiplication. This suffices to show that the condition holds.
For the second condition, the proof is similar than in Theorem~\ref{theo:capture-fp}. Let $f\in \fpspace$ defined over some $\R$ and $\ell\in\nat$ such that $\log_2\left( f(\enc(\A)) \right) \leq 2^{{|\A|}^\ell}$ for every $\A\in\ostr[\R]$  (i.e. $2^{{|\A|}^\ell}$ is an upper bound for the output size). Let $X$ be a second-order variable of arity $\ell$. Consider the linear order induced by $<$ over predicates of arity $\ell$ which can be defined by the following formula:
$$
\varphi_{<}(X,Y) = \ex{\bar{u}}\big[\neg X(\bar{u})\wedge Y(\bar{u})\wedge \fa{\bar{v}}\big(
\bar{u}<\bar{v}\to(X(\bar{u})\iff Y(\bar{v}))\big)\big].
$$
Namely, we use relations to encode numbers with at most $2^{{|\A|}^\ell}$-bits where the empty relation represents $0$ and the total-relation represents $2^{2^{{|\A|}^\ell}}-1$.
Furthermore, we can use a relation~$X$ to index a position in the binary output of $f(\enc(\A))$ as follows.
%Consider a polynomial space machine over the $\R$ that receives as input an $\R$-structure $\A$ and a number $p$ encoded by a relation $X$. Then the machine accepts if, and only if, the $p$-th bit of $f(\enc(\A))$ is $1$. 
Define the language:
\[
L = \{(\A,B)\mid B \subseteq A^{\ell}\text{ and the $B$-th bit of $f(\enc(\A))$ is 1}\}.
\]
Since $L$ is in $\pspace$, it can be specified in $\pfp$ \cite{AbiteboulV89} by a formula $\Phi(X)$ where the free variable $X$ encodes relation $B$ in $L$. Then, similar than the previous proof we define:
$$
\alpha := \sa{X} \Phi(X)\mult  \pa{Y}(\varphi_{<}(Y,X)\mapsto 2).
$$ 
where $\pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$ takes the value $2^m$ if there exist $m$ predicates that are smaller than~$X$ and $\alpha$ reconstruct the output of $f(\enc(\A))$ by simulating $f$ with $\Phi(X)$. Using an analogous argument, we conclude that $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.

### 2018-05-07 LMCS/send/proof-prop-e1-nc.tex
En el directorio send esta la version enviada a arXiv

Towards a contradiction, assume that $\E{1}$ is closed under binary sum. 
Consider the formula $\alpha = \sa{x}(x = x) \in \E{1}$ over some signature $\R$. 
This defines the function $\sem{\alpha}(\A) = \size{\A}$. 
From our assumption, there exists some formula in $\E{1}$ equivalent to the formula $\alpha \add \alpha$, which describes the function $2\size{\A}$. 
Let $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ be this formula, where $\varphi$ is in first-order and quantifier-free. 
For each $\R$-structure $\A$, we have the following inequality:
$$
\sem{\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A)
\leq 
\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot  \size{\A}^{\length{\bar{y}}} \leq 2\size{\A}^{\length{\bar{y}}+1} 
$$
Note that the formula $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ defines a function in $\E{0}$. 
Therefore, as it was proven in \cite{SalujaST95}, if $\length{\bar{X}} > 0$ then the function is in $\Omega(2^{\size{\A}})$, which violates the inequality.

We now have that $\length{\bar{X}} = 0$. 
Consider a structure $\mathfrak{1}$ with only one element. 
We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$, but since the structure has only one element, there is only one possible assignment to $\bar{x}$. 
And so, $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which leads to a contradiction.

### 2018-05-07 LMCS/send/syntactic.tex
En el directorio send esta la version enviada a arXiv

--- 2017-04-11 LICS/syntactic.tex
@@
 Counting problems with easy decision versions play a fundamental role in the search of efficient approximation algorithms for functions in $\shp$.
-A fully-polynomial randomized approximation scheme (FPRAS) for a function $f : \Sigma^* \to \bbN$ is a randomized algorithm ${\cal A} : \Sigma^* \times (0,1) \to \bbN$ such that: (1) for every string $x \in \Sigma^*$ and real value $\varepsilon \in (0,1)$, the probability that $|f(x) - {\cal A}(x,\varepsilon)| \leq \varepsilon \cdot f(x)$ is at least $\frac{3}{4}$, and (2) the running time of ${\cal A}$ is polynomial in the size of $x$ and $1/\varepsilon$ \cite{KL83}.
+A fully-polynomial randomized approximation scheme (FPRAS) for a function $f : \Sigma^* \to \bbN$ is a randomized algorithm $\mathcal{A} : \Sigma^* \times (0,1) \to \bbN$ such that: (1) for every string $x \in \Sigma^*$ and real value $\varepsilon \in (0,1)$, the probability that $|f(x) - \mathcal{A}(x,\varepsilon)| \leq \varepsilon \cdot f(x)$ is at least $\frac{3}{4}$, and (2) the running time of $\mathcal{A}$ is polynomial in the size of $x$ and $1/\varepsilon$ \cite{KL83}.
 Notably, there exist $\shp$-complete functions that can be efficiently approximated as they admit FPRAS; for instance, there exist FPRAS for $\cdnf$~\cite{KL83} and $\cpm$~\cite{JSV04}.
@@
 
-A second desirable condition of robustness is the existence of natural complete problems \cite{P94}.
+A second desirable condition of robustness is the existence of natural complete problems~\cite{P94}.
 Special attention has to be paid here to the notion of reduction used for completeness.
@@
 
-%The class $\shp$ was introduced in \cite{Valiant79} to prove that computing the permanent of a matrix, as defined in Example \ref{exa-perm}, is a difficult problem.
-More specifically, it was shown in  \cite{Valiant79}  that this problem is $\shp$-complete.
-As a consequence of this result the $\shp$-completeness of the problem of computing the number of perfect matchings of a bipartite graph ($\cpm$) was established.
-%Since then  many counting problems have been proved to be $\shp$-complete \cite{V79b,PB83,P86,L86,BW91,HMRS98,DG00,BW05,DS12, PS13,PS14}.
-Among them, problems having easy decision counterparts play a fundamental role, as a counting problem with a hard decision version is expected to be hard.
-Formally, the decision problem associated to a function $f : \Sigma^* \to \mathbb{N}$ is defined as $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$, and $f$ is said to have an easy decision version if $L_f \in \ptime$.
-A first prominent example of such functions is $\cpm$, as it is well-known that the problem of verifying whether there exists a perfect matching in a bipartite graph can be solved in polynomial time.
-Other prominent examples of such problems include counting the number of: satisfying assignments of a 2-CNF propositional formula ($\ctwcnf$) \cite{V79b}, satisfying assignments of a DNF propositional formula ($\cdnf$) \cite{DHK05}, satisfying assignments of a Horn  propositional formula ($\chsat$) \cite{V79b}, simple paths from a source node to a target node in a directed graph ($\csp$) \cite{V79b}, extensions of a partial order to a linear order \cite{BW91} and Eulerian cycles in an undirected graph \cite{BW05}.
-%
-%Counting problems with easy decision versions play a fundamental role in the search of efficient approximations algorithms for functions in $\shp$.
-A fully-polynomial randomized approximation scheme (FPRAS) for a function $f : \Sigma^* \to \bbN$ is a randomized algorithm ${\cal A} : \Sigma^* \times (0,1) \to \bbN$ such that: (1) for every string $x \in \Sigma^*$ and real value $\varepsilon \in (0,1)$, the probability that $|f(x) - {\cal A}(x,\varepsilon)| \leq \varepsilon \cdot f(x)$ is at least $\frac{3}{4}$, and (2) the running time of ${\cal A}$ is polynomial in the size of $x$ and $1/\varepsilon$ \cite{KL83}.
-Notably, there exist $\shp$-complete functions that can be efficiently approximated as they admit FPRAS; for instance, there exist FPRAS for the problems of counting the number of satisfying assignments of a DNF propositional formula \cite{KL83} and the number of perfect matchings of a bipartite graph \cite{JSV04}.
-A key observation here is that if a $\shp$-complete function admits an FPRAS, then its associated decision problem is in the randomized complexity class $\bpp$ (Bounded-Error Probabilistic Polynomial-Time \cite{G77}).
-Hence, under the widely believed assumption that  $\np \not\subseteq \bpp$, we cannot hope for an FPRAS for a function in $\shp$ whose decision counterpart is $\np$-complete, and we have to concentrate on the class of counting problems with easy decision versions (in $\bpp$ or in a lower complexity class such as $\ptime$).
-%
-%The importance of the class counting problems with easy decision counterparts has motivated the search of robust classes of functions in $\shp$ with this property \cite{PagourtzisZ06}.
-But the key question here is what should be considered a {\em robust} class.
-A first desirable condition is that the class has a syntactic definition that allows for the existence of natural complete problems \cite{P94}.
-Special attention has to be paid here to the notion of reduction used for completeness.
-Notice that under the notion of Cook reduction, originally used in \cite{Valiant79}, the problems $\cdnf$ and $\csat$ are $\shp$-complete.
-However, $\cdnf$ has an easy decision counterpart and admits an FPRAS, while $\csat$ does not satisfy these conditions unless $\ptime = \np$.
-Hence a more strict notion of reduction has to be considered; in particular, the notion of parsimonious reduction (to be defined later) satisfy that if a function $f$ is parsimoniously reducible to a function $g$, then $L_g \in \ptime$ implies that $L_f \in \ptime$ and the existence of an FPRAS for $g$ implies the existence of a FPRAS for $f$.
-%A second desirable condition has to do with the closure properties satisfied by the class, which is a common theme when studying function complexity classes \cite{OH93,FH08}.
-As in the cases of $\ptime$ and $\np$ that are closed under intersection and union, we expect our class to be closed under multiplication and sum; in fact, if $f$ and $g$ are counting functions with easy decision versions, then we clearly expect $f + g$ and $f \cdot g$ to have easy decision counterparts.
-For a more elaborated closure property, assume that $\textit{sat\_one}$ is a function that returns one plus the number of satisfying assignments of a propositional formula.
-Clearly $\textit{sat\_one}$ is a $\shp$-complete function whose decision counterpart $L_{\textit{sat\_one}} = \{ \varphi \mid \textit{sat\_one}(\varphi) > 0 \}$ is trivial.
-But should $\textit{sat\_one}$ be part of a robust class of counting function with easy decision versions?
-The key insight here is that if a function in $\shp$ has an easy decision counterpart $L$, then as $L \in \np$ we expect  to have a polynomial-time algorithm that verifies whether $x \in L$ by constructing witnesses for $x$.
-As an example of this considered all the counting functions mentioned in the previous paragraphs and, in particular, the decision counterpart of $\cdnf$ that can be solved by explicitly constructing in polynomial-time a satisfying assignment for an input propositional formula.
-Moreover, if such an algorithm for constructing witnesses exists, then we also expect to be able to manipulate such witnesses and in some cases to remove them.
-In other words, we expect a robust class $\CC$ of counting functions with easy decision versions to be closed under subtraction by one, that is, if $g \in \CC$, then the function $g \dotdiv 1$ should also be in $\CC$, where $(g \dotdiv 1)(x)$ is defined as $g(x) - 1$ if $g(x) \geq 1$, and as $0$ otherwise.
-Notice that, unless $\ptime = \np$, no such class can contain the function $\textit{sat\_one}$ because $\textit{sat\_one} \dotdiv 1$ counts the number of satisfying assignments of a propositional formula.
-%
-%In this section, we use the framework developed in this paper to address the problem of defining a robust class of functions with easy decision versions.
-More specifically, we use the framework to introduce in Section \ref{sec-hier-shp} a syntactic hierarchy of function complexity classes contained in $\shp$.
-Then this hierarchy is used in Section \ref{sec-clo} to define a class of functions with easy decision versions and good closure properties, and in Section \ref{sec-horn} to define a class of functions with easy decision versions and
-%natural complete problems.
-
-% It should be noted that such class can be directly defined as the set functions f ?
-#P such that Lf ?
-P, which is denoted as #Pe in [50, 51].
-However, such a definition does not lead to a well-behaved and robust function complexity class.
-In particular, for every function f ?
-#P, we have that f + 1 is trivially in #Pe, which is an undesirable property.
-This has led to the introduction of the more robust class TotP, which is defined as the class of functions f for which there exists a non-deterministic Turing machine M running in polynomial time such that, f(x) is the result of subtracting 1 to the number of (non-necessarily accepting) runs of M with input x [40].
-In [51], it is proved that TotP ?
-#Pe and that TotP has a complete function problem under parsimonious reductions.
-However, no natural problem is known to be TotP-complete under this type of reductions [51].
-
-%
-%
-%In this section we study the fragment of $\eqso(\LL)$ when $\LL$ is a boolean logic contained in $\fo$.
-We show that by restricting $\LL$ we can find different subclasses below $\shp$ with interesting computational and closure properties.
-%
-%
-%
-%From this point on, for each fragment $\FF$ of $\qso$, we will also use $\FF$ to refer to the class of functions defined by the formulae in $\FF$.
-
-%\subsection{A syntactic hierarchy of function complexity classes contained in $\shp$}
 \subsection{The $\eqso(\fo)$-hierarchy inside $\shp$}

### 2018-12-08 LMCS/todo-list.tex
added TODO list

% !TeX root = main.tex

\section*{\LARGE \textbf{TODO LIST}}

\bigskip

{\bf Technical}

\begin{itemize}
	\setlength\itemsep{0.5em}
\item p.2, first paragraph: ``However, other function classes have emerged from the need to understand
the complexity of some computation problems for which little can be said if their
decision counterparts are considered.'' Why can only little be said about the decision counterparts
of the examples mentionend afterwards?
\item p.7, first sentence: ``fragments or extensions [...] obtained by restricting'': extensions are not
obtained by restricting something
\item p.10, before Thm. 4.5: ``by considering second-order product'': We need both second-order
product and sum (at least both are used in Thm. 4.5)
\item p.10, last line: ``$B$-th bit'': mention in what way $B$ is interpreted as a number
\item p.12: ``To formalize this set of optimization problems'': class instead of set
\item p.14: ``we have to concentrate on the class of counting problems with easy decision versions'':
here, easy decision version means ``in \bpp''. Before, it was defined as ``in P''. Should be
consistent.
\item p.14: ``as expected, $\Sigma \bar{X}$ is the respective nesting of $\Sigma X.$’s'': why is this explained for secondorder
quantifiers but not for first-order quantifiers?
\item p.14: $\sum_{i=1}^n\alpha_i$: explain that this is notation for $\alpha_1+\cdots+\alpha_n$
\item p.17: ``Since $\alpha$ is not null'' $\to$ ``Since $\sem{\alpha}$ is not the identically zero function'' (or similar)
\item p.17/18: In the proof of Prop 5.3, the vocabulary only contains the numerical predicate $<$. When encoding structures as binary words, numerical predicates are implicit (and not
encoded), though. What is the intended meaning here?
\item p.18, after proof of Claim 5.4: ``which implies that 2 $\in \eqso(\Sigma_0)$ and $2 \not\in \E{1}$'': $2 \in \eqso(\Sigma_0)$
does not follow from the argument before as stated in this sentence, but is trivial
\item p.18: ``Also, suppose that it is in $\L$-PNF'': $\L$ should be $\LL$ , PNF should be SNF
\item p.19, end of proof of Prop. 5.5: ``which violates the inequality'' $\to$ ``which violates the inequality
for large structures''
\item p.20: ``we will show a formula in the grammar'' $\to$ ``... in the logic''
\item p.20: ``Since $\Sigma_1$ is closed under conjunction, then Lemma 5.7 holds for ...'' $\to$ ``Since $\Sigma_1$ is
closed under cnojunction, Lemma 5.7 can be applied to ...'' (or similar)
\item p.22: ``and leaving each $R_{\psi}^{\A}$ empty'' $\to$ $R_{\psi}^{\A'}$
\item p.22: ``To show that $\alpha$ is in \totp'' $\to$ ``To show that the function defined by $\alpha$ ...'' (or similar)
\item p.22: ``Let $\alpha$ = ... where $\varphi$ is an FO-formula'' $\to$ ``... where $\varphi$ is an $\Sigma_1$[FO]-formula''
\item p.23, formulae: Maybe put (large) parantheses around the relativization of quantifiers, for
example the part of formula $\varphi_i$ after the quantifiers
\item p.24, formula 5.6: subscript of large conjunction: $\varphi_i^{+}(\bar{X},\bar{v})$ $\to$ $\varphi_i^{+}(\bar{X},\bar{y})$
\item p.25: ``We construct a function $\lambda$ that receives a quantitative formula $\beta$ ...'' $\to$ ``... that receives
a formula $\beta \in \eqso(\logex{1})$ ...''
\item p.25, last sentence: ``which is based on considering some witnesses of logarithmic size'' could
be explained a little more / made more clear
\item p.26: ``... auxiliary predicate $A$ defined as'' $\to$ ``... auxiliary predicate $A$ which will be defined
as ... in the formula'' (or similar)
\item p.28: ``since universal formulas are monotone over induced substructures'': ``monotone over
induced substructures'' should be reformulated
\item p.28: ``Next we prove that ... is hard for ... over a signature $\R$ ...'' sounds like it only works
for one specific vocabulary. Maybe simply ``over any signature $\R$''
\item p.28, formula $\alpha_{\A}$: second big conjunction should have $\bar{z}'$ in the subscript instead of $\bar{z}$ (because
the quantifier $\forall u$ is handled by this as well if I understand correctly)
\item p.29: ``In other words, in poly time we can replace $\varphi_j^i$ by a disjunction of ....'': Some of these
formulae might evaluate to $\perp$ or $\top$ altogether and can be removed (potentially together with
the surrounding conjunction)
\item p.29: ``Intuitively $t_1,\ldots,t_{m'}$ are used to count from 1 to $m'$ each subformula ...'': Should be
reformulated, maybe rather write that they are used to have disjoint sets of propositional
assignments for the different disjuncts of the outermost disjunction (corresponding to the
summands in the original formula)
\item p.30, proof of Prop. 5.12: For $\dhsat \in \totp$, [35] is cited. It’s stated that a \totp-procedure
for DNF-SAT is given there and it can be adapted to \dhsat. While the
cited paper contains the result for DNF-SAT, the procedure is not explicitly given as far as
I can see, so I can't check whether it is adaptable. On the other hand, membership of any
function counting the number of satisfying assignments for a class of propositional formulae
with poly-time decision version in \totp should hold by a similar argument as membership in
DelayP. Maybe change the citation or give a short proof sketch.
\martin{Punto posiblemente largo}
\item p.31, definition of $\leq$ on $\F$: ``$f(i) \leq g(i)$ for every $i \in \{1,\ldots,\ell\}$'': Inputs for $f$ and $g$ should be
elements of $A^{\ell}$, not $\{1,\ldots,\ell\}$
\item p.32/33, example 6.2: I don't see why the formula given for dag's does not work for arbitrary
graphs: In arbitrary graphs, the fixpoint is still reached after at most $n-1$ steps, since the
maximal length of any shortest path in any graph is $n-1$. On the other hand, assuming that
I am mistaken and the support does not become stable within $n-1$ steps: Using the formula
given for arbitrary graphs, the support of the constructed function restricted to those tuples
that have first in their last component is the same as the support of the constructed function
using the formula for dags in each step of the recursion. Hence, the formula should not be
able to fix such a problem.
\item p.33: ``executes exactly $n^k$ steps for a fixed $k \geq 1$'' → ``executes exactly $n^k$
steps on large
inputs for a fixed $k \geq 1$'' (at least it does not work for $n = 1$) (small inputs can be handled
seperately)
\item p.35 and p.36: While $\sum$ was used for finite sums in \qso-formulae throughout the paper, now
a large +-sign is used. This might be confusing
\item p.36: $\alpha_{q_i}$ used instead of $\alpha_{s_{q_{i}}}$
\item p.36: ``Clearly, at each iteration of the fixed point operator, the tuple $\bar{t}$ represents the step the
machine is currently in.'': This is quite vague. Maybe something like ``Clearly, tuple $\bar{t}$ encodes
the number of steps the machine has done and in each iteration of the fixed point operator,
one timestep of the machine is executed''
\item p.36: ``at the $\bar{a}$-th iteration of the fixed point operator, it holds that $f(\bar{a}, \bar{b}) \leq 1$ and $f(\bar{a}', \bar{b}) = 0$ for every $\bar{b} \in A^k$. 
In the same way, it can be seen that for each function $g \in \ldots$ it holds that $g(\bar{a}) \leq 1$ and $g(\bar{a}')= 0$. 
I think that more properties of these functions should be mentioned
in order to argue that the defined function is the function computed by the machine:
\begin{itemize}
\item[-] For each $\bar{b}$, only one of the functions $T_0$,$T_1$,$T_B$,$T_{\vdash}$ outputs 1 on input ($\bar{a}, \bar{b})$
\item[-] For each $\bar{b}$, only one of the functions $h,\hat{h}$ outputs 1 on input $(\bar{a}, \bar{b})$
\item[-] There is exactly one $\bar{b}$ with $h(\bar{a}, \bar{b}) = 1$
\item[–] There is exatcly one $i$ such that $s_{q_i}(\bar{a}) = 1$
\end{itemize}
(It could also generally be argued that the functions are 0/1-functions encoding the intended
predicates and are 0 when the timestep is the successor of the current timestep)
\item p.36: ``is defined as an extension of QFO with the formula ...'' $\to$	 ``is defined as the extension
of QFO by the additional operator ...'' (or similar)
\end{itemize}
\vspace{1em}
{\bf Grammatical/Language}
\begin{itemize}
	\setlength\itemsep{0.5em}
	\item generally: both formulae and formulas are used as plural of formula---should be consistent
	\item p.2, in the middle: we propose a restriction: Should be something like ``we propose to restrict
	WL to the natural numbers as fixed semiring, calling this restriction ...''
	\item p.3, first paragraph of 2.1: ``omit the word finite or ordered'' $\to$ ``omit the words finite and
	ordered''
	\item p5, description of syntax of QSO: ``This division between Boolean and quantitative level
	is essential for understanding the difference between the logic and the quantitative part.''
	reformulate this?
	\item p.5, next sentence: ``this will allow us later'' $\to$ ``this will later allow us''
	\item p.6, second to last sentence: ``similarity of $\alpha_3$ with'': ``similarity to'' is much more common
	than ``similarity with''
	\item p.7, next sentence: ''in this direction'' $\to$ ``in this regard''
	\item p.7, last sentence before 3.1: ''connection of QSO with'' $\to$ ``connection of QSO to''
	\item p.7, middle: ``Another difference of WL with our approach'' $\to$ ``Another difference between
	WL and our approach''
	\item p.9, first sentence: ``then condition (2) holds'': omit the word then or reformulate the sentence
	\item p.9, first paragraph: ``the constant function $s$ can be trivially simulated'' $\to$ ``any constant
	function $s$ can be trivially simulated''
	\item p.9, proof of Prop. 4.3: ``To prove the condition (2)'' $\to$ ``To prove condition (2)''
	\item throughout the paper: ``similar than'' $\to$ ``similar to'' / ``similar than for X, we do Y'' $\to$
	``similarly as for''. Examples:
	\begin{itemize}
	\item[-] p.10, proof of Thm. 4.5: ``the proof is similar than in Theorem 4.4'' $\to$ ``the proof is
	similar to the proof of Theorem 4.4''	
	\item[-] p.10, second sentence: ``similar than the previous proof'' $\to$ ``similar as in the previous
	proof''
	\item[-] p.11, last sentence before example 4.7: ``Similar than for QSO'' $\to$ ``Similarly as for QSO''
	\end{itemize}
	\item p.11, end of proof of Thm. 4.5: ``Using an analogous argument'': state what the argument is
	analogous to (the argument in Thm. 4.4)
	\item p.12, after Corollary 4.8: ``shows how robust and versatile is QSO'' $\to$ ``shows how robust and
	versatile QSO is''
	\item p.12, after Corollary 4.8: ``counting complexity classes [...] even beyond $\nat$'' - make more precise
	what ``beyond $\nat$'' refers to
	\item p.12, ``given CNF propositional formula, that can be made true'': no comma
	\item p.12: ``in [43] they defined'' $\to$ something like ``$<$authors$>$ [43] defined''
	\item p.12: ``we extend as follows QSO'' $\to$ ``we extend QSO by max and min quantifiers as follows''
	\item p.12: ``MinQSO is defined analogously changing max with min'' $\to$ ``replacing max by min''
	\item p.12: ``the same holds with MinQSO(FO)'' $\to$ ``the same holds for ...''
	\item p.13: ``in contrast with'' $\to$ ``in contrast to''
	\item p.13: ``follow the same approach but considering'' $\to$ ``but consider''
	\item p.13: ``in the search of efficient approximation algorithms'' $\to$ ``in the search for''
	\item p.14: ``the search of robust classes'' $\to$ ''the search for robust classes''
	\item p.14: ''the first desirable condition has to do with the closure properties satisfied by the class'':
	Reformulate this sentence
	\item p.14: ``As in the cases of P and NP that are closed'' $\to$ ``Analogously to the cases of P and
	NP, which are closed under intersection and union, ...''
	\item p.16: ``such that for both, $A\mod\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$'' $\to$ ''such that for $i \in \{1,2\}$: ... ''
	\item p.16: ``Similarly to Theorem 5.1, we show ..'' $\to$ ``Similarly as in the proof of Theorem 5.1, we
	show''
	\item p.19: ``However, this class is not closed under sum, and then it is not robust under basic
	closure properties'': Should be reformulated
	\item p.19: ``Let ... be this formula, where $\varphi$ is in first-order and quantifier-free'' $\to$ ``... where $\varphi$ is
	first-order and quantifier-free''
	\item several occurrences: ``This is, ...'' $\to$ ``That is, ...'', for example two times on p.20
	\item p.20: ``we focus only in proving'' $\to$ ``we focus only on proving''
	\item p.20: ``We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of FO.''
	$\to$ ``We prove this for the more general case of $\eqso(\LL)$ with $\LL$ being a fragment of SO.''
	(or similar)

	\item p.20: ``we will show a formula'' $\to$ ``give'' or ``construct'' instead of ``show''
	\item p.20/21: ``$\shp$ is believed not to be closed under subtraction by one by some complexitytheoretic
	assumption'' $\to$ ``... due to some complexity-theoretic assumption'' (or similar)
	\item p.21, next sentence: ``for a possible extension'' $\to$ ``as a possible extension''
	\item p.21: ``the desired extension has to be achieved by ... . More precisely, we define...'': The second
	sentence doesn’t make the statement of the first one more precise, but rather explains the
	actual definiton (which is chosen in accordance with the sentence before)
	\item p.21: ``a parsimonious reduction from a function'' $\to$ ``a parsimonious reduction from any
	function''
	\item p.21: ``we define a function that converts a formula $\alpha$ in...'' $\to$ ``that converts a given formula
	$\alpha$...''
	\item p.22: ``we iterate for every tuple'' $\to$ ``we iterate over all tuples''
	\item p.22: ``Then we will show how to define a formula $\varphi'$
	...'' $\to$ ``We will show how to define ...''
	\item p.22: ''no variable from $\bar{x}$ are mentioned'' $\to$ ``... is mentioned''
	\item p.24: ``if, and only if, if X, then Y'': This construction makes it unneccesarily hard to parse
	the sentence
	\item p.24: ``Then, the conclusion of the implication...'' $\to$ ``The conclusion of the implication...''
	\item p.25: ``For the sake of simplification...'' $\to$ ``For the sake of simplicity...''
	\item p.26: ``In the previous section, we show ...'' $\to$ ``In the previous section we showed ...''
	\item p.27: ``we postpone the proof the previous proposition after ...'' $\to$ ``we postpone the proof of
	the previous proposition to after ...''
	\item p.29: ``replace by $\perp$ and $\top$ \underline{wherever it corresponds}'': Should be reformulated
	\item p.30: ``consist in computing $g_{\alpha}(\enc(\A))$ for each input ...'' $\to$ ``... on input ...''
	\item p.30, same sentence: ``and then simulate'' $\to$ ``and then simulating''
	\item p.33: ``In contrast with LFP'' $\to$ ``In contrast to LFP''
	\item p.33: ``These operations are represented by the set $\{0,1,\emptyset\}$, respectively.'' $\to$ ``... by 0, 1 and
	$\emptyset$, respectively''
	\item p.34: ``Notice that the values of each function $f$ in $\alpha$'' $\to$ ``Notice that the values of each
	function in $\alpha$''
	\item p.34: ``that is why we talk about ...'' $\to$ ``which is why we talk about''
	\item p.34: ``Besides, notice that notation ...'' $\to$ ``Besides, notice that the notation ...''
	\item p.34: ``Function $h$ is used to indicate whether the head of the working tape is in some position
	at some point of time'' $\to$ ``... whether the head of the working tape is in a specific position
	at a certain point of time'' (or similar)
	\item p.36: ``Given a relation signature $\R$'' $\to$ ``... relational signature ...''
	\item p.37: ``As it was mentioned before'' $\to$ ``As mentioned before''
	\item p.37: ``Next we construct a logarithmic-space ....'' $\to$ ``We construct a logarithmic-space ...''
	6
	\item p.38: ``In this sense, there are several directions for future research'' $\to$ ``Consequently, there
	are ...'' (or similar)
	\item p.38: ``In the same direction, ...'' $\to$ ``Similarly, ...'' (or similar)
	\item p.38: ``so to have'' $\to$ ``in order to have'' (or similar)
\end{itemize}
\vspace{1em}
{\bf Grammatical/Language}
\begin{itemize}
	\setlength\itemsep{0.5em}
	\item p.1, footnote: ``clases'' should be ``classes''
	\item throughout the paper: plural used with verb ending on s. Examples:
	\begin{itemize}
	\item[-] p.2, middle: ``fragments of QSO captures''
	\item[-] p.6, example 3.1: ``the sum quantifiers in $\alpha_1$ aggregates''
	\end{itemize}
	\item p4, beginning of 2.2: ``some of \underline{the} are recalled here''
	\item p6, example 3.2, first paragraph: ``otherwise $M(i.j)$'', dot used instead of comma
	\item throughout the paper: a instead of an (espacially: a FO formula/assignment, a SO formula/assignment).
	Example:
	\begin{itemize}
		\item[-] p.9, third sentence: ``Given a FO formula'' $\to$ ``Given an FO formula''
	\item[-] p.38: ``As a example'' $\to$ ``As an example''
\end{itemize}
	\item throughout the paper: singular used with verb not ending on s. Examples:
	\begin{itemize}
	\item[-] p.9, after proof of Prop. 4.3: ``use of a fragment [...] that capture''
	\item[-] p.11, end of proof of Thm. 4.5: ``and $\alpha$ reconstruct''
	\item p.9, middle: ``A second attempt could be based then'' $\to$ ``A second attempt could then be
	based''
\end{itemize}
	\item p.11, 4.1: ``output of a function \underline{in} not a natural number''
	\item p.12: ``we have to dinstinguished''
	\item p.14, beginning of 5.1: ``subclases''
	\item p.21: ``it does not mentioned'' $\to$ ``it does not mention''
	\item p.22: ``Without lost of generality'' $\to$ ``Without loss of generality''
	\item p.24: ``The previous claim and proof motivates the following definitions.'' $\to$ ``The previous
	claim and proof motivate the following definition.''
	\item p.30: ``to define a natural generalization for functions of the notion of least fixed point of
	LFP'' $\to$ ``to define a natural generalization of LFP to functions'' (or similar)
	\item p.30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
	\item p.31: ``we simply generalized these conditions'' $\to$ ``we simply generalize these conditions''
	\item p.32: ``which is used to defined'' $\to$ ``which is used to define''
	\item p.33: ``since $\beta$ is constructed from monotones operations'' $\to$ ``... monotone ...''
	\item p.35: ``Formulas $\alpha_h$ is defined ...'' $\to$ ``Formula $\alpha_h$ is defined ...''
	7
	\item p.37: ``we simulate $M_\beta$ with input $v[1/n]$'' $\to$ ``... with input $v[1/x]$''
\end{itemize}
\vspace{1em}
{\bf Other}
\begin{itemize}
	\setlength\itemsep{0.5em}
	\item running title is ``COUNTING''
	\item on definition of functions, \texttt{\symbol{92}colon} should be used instead of :
	\item In [1], a similar hierarchy to the one by Saluja et al. was introduced and studied. This work
	should be cited.
\end{itemize}

### 2018-12-08 LMCS/send/proof-capture-fpspace-poly.tex

In this proof, both conditions are analogous to Theorem~\ref{theo:capture-fp} and~\ref{theo:capture-fpspace}. For the first condition, each $\pfp$ formula $\varphi$ can be evaluated in $\pspace$ and the class is closed under first sum and product. For the second condition, we use the same language $L$ defined in the proof of Theorem~\ref{theo:capture-fp}, which in this case is in $\pspace$. The same construction of $\alpha$, which in turn is in $\qfo(\pfp)$, is used to show that the condition holds.

### 2018-12-08 LMCS/send/proof-lem-support.tex

--- 2018-05-06 LMCS/proof-lem-support.tex
@@
 \item Assume that $\beta$ is either a constant $s \in \mathbb{N}$ or an $\fo$-formula $\varphi$.
-In both cases function symbol $h$ is not mentioned, so $\sem{\beta}(\A,v,V,F) = \sem{\beta}(\A,v,V,G)$ and it trivially holds that if $\sem{\beta}(\A,v,V,F) > 0$, then $\sem{\beta}(\A,v,V,G) > 0$.
+In both cases, function symbol $h$ is not mentioned, so $\sem{\beta}(\A,v,V,F) = \sem{\beta}(\A,v,V,G)$ and it trivially holds that if $\sem{\beta}(\A,v,V,F) > 0$, then $\sem{\beta}(\A,v,V,G) > 0$.
 %Then $h$ does not appear.

### 2018-12-08 LMCS/send/proof-rqfo-fo-cap.tex

--- 2018-05-06 LMCS/proof-rqfo-fo-cap.tex
@@
  (1) writing 0 and moving the head one cell to the right, (2) writing 1 and moving the head one cell to the right, or (3) doing nothing.
-These operations are represented by the set $\{0,1,\emptyset\}$.
+These operations are represented by the set $\{0,1,\emptyset\}$, respectively.
 Finally, assume that $M$, on input $\enc(\A)$ with domain $A = \{1,\dots,n\}$, executes exactly $n^k$ steps for a fixed $k \geq 1$.
@@
 \begin{align*}
-\alpha = \sa{\bar{t}}\clfp{\out(\bar{t}): \,&\alpha_{T_0}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out),\\
+\alpha \ = \ \sa{\bar{t}}\clfp{\out(\bar{t}): \,&\alpha_{T_0}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out),\\
 	&\alpha_{T_1}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out),\\
@@
 Notice that the values of each function $f$ in $\alpha$ depend on the values of the other functions, that is why we talk about a simultaneous computation.
-Besides, notice that notation $\clfp{\out(\bar{t}) : \ldots}$ is used to indicate that (1) the free variables of the formula are $\bar t = (t_1, \ldots, t_k)$, and (2) once the least fixed point has been computed, the value of $\clfp{\out(\bar{t}) : \ldots}$ for an assignment $\bar a$ to $\bar t$ is given by the value of function $\out$ on $\bar a$ in the fixed point.
-Finally, it is important to notice that $\alpha$ can be transformed into a proper $\rqfo(\fo)$ formula by using the same techniques to prove that a simultaneous LFP can be transformed into a single LFP~\cite{L04}; in particular, functions symbols $T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out$ are replaced by a single function $f$ of arity $6 \cdot 2 \cdot k + (\ell + 1) \cdot k + k = (\ell+ 14) \cdot k$.
+Besides, notice that notation $\clfp{\out(\bar{t}) : \ldots}$ is used to indicate that (1) the free variables of the formula are $\bar t = (t_1, \ldots, t_k)$, and (2) once the least fixed point has been computed, the value of $\clfp{\out(\bar{t}) : \ldots}$ for an assignment $\bar a$ to $\bar t$ is given by the value (in the fixed point) of function $\out$  on $\bar a$.
+Finally, it is important to notice that $\alpha$ can be transformed into a proper $\rqfo(\fo)$ formula by using the same techniques used to prove that simultaneous LFP has the same expressiveness as LFP~\cite{L04}; in particular, functions symbols $T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out$ are replaced by a single function $f$ of arity $6 \cdot 2 \cdot k + (\ell + 1) \cdot k + k = (\ell+ 14) \cdot k$.
 
@@
 Function $\out$ stores the output of the TM $M$; in particular, $\out(\bar t)$ is the value returned by $M$ when $\bar t$ is the last step.
-Finally, assuming that $\varphi_{\text{\rm last}}(x)$ is an $\fo$-formulae defining the last element of $<$, we have that $\last(\bar t) = \bigwedge_{i=1}^n \varphi_{\text{\rm last}}(t_i)$, that is, $\last(\bar t)$ holds if $\bar t$ is the last step.
+Finally, assuming that $\varphi_{\text{\rm last}}(x)$ is an $\fo$-formula defining the last element of $<$, we have that $\last(\bar t) = \bigwedge_{i=1}^k \varphi_{\text{\rm last}}(t_i)$, that is, $\last(\bar t)$ holds if $\bar t$ is the last step.
 Therefore, the use of $\last(\bar t)$ in $\alpha$ allows us to return the output of the TM $M$.
 
-As in Example \ref{ex:count-path}, assume that $\varphi_{\text{\rm first}}(x)$ and $\varphi_{\text{succ}}(x,y)$ are $\fo$-formulae defining the first element of $<$ and the successor relation associated to $<$, respectively.
-Moreover, assume that formulae $\varphi_{\text{succ}}(x,y)$ and $\varphi_{\text{\rm last}}(x)$ are used to defined
-an $\fo$-formula $\succesor(\bar{t},\bar{t'})$ such that $\succesor(\bar{t},\bar{t'})$ holds if $\bar t'$ is the successor step of $\bar t$ (that is, $\succesor(\cdot,\cdot)$ is the successor relation associated to the lexicographical order induced by $<$ on the tuples with $k$ elements).
+As in Example \ref{ex:count-path}, assume that $\varphi_{\text{\rm first}}(x)$ is a $\fo$-formula defining the first element of $<$.
+Moreover, assume that $\first(\bar t)$ and $\succesor(\bar{t},\bar{t'})$ are $\fo$-formulae such that
+$\first(\bar t)$ holds if $\bar t$ is the first step and $\succesor(\bar{t},\bar{t'})$ holds if $\bar t'$ is the successor step of $\bar t$ (that is, $\succesor(\cdot,\cdot)$ is the successor relation associated to the lexicographical order induced by $<$ on the tuples with $k$ elements).
 Then formula $\alpha_{T_{\vdash}}$ is defined as follows:
@@
 Notation $\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,\vdash,\_,\_)}$ in this formula is used to indicate that we are considering a sum over all pairs $(q,a) \in Q \times \{0,1,\ttB,\vdash\}$ such that $\delta(q,a) = (q',\vdash,u,v)$ for some $q' \in Q$, $u \in \{\leftarrow,\rightarrow\}$ and $v \in \{0,1,\emptyset\}$.
-Formulas $\alpha_{T_0}$, $\alpha_{T_1}$ and $\alpha_{T_{\ttB}}$ are defined analogously; for the sake of completeness, we only show here how $\alpha_{T_0}$ is defined, assuming that $\bar y = (y_1, \ldots, y_k)$:
+Formulas $\alpha_{T_0}$, $\alpha_{T_1}$ and $\alpha_{T_{\ttB}}$ are defined analogously; for the sake of presentation, we only show here how $\alpha_{T_0}$ is defined, assuming that $\bar y = (y_1, \ldots, y_k)$:
 \begin{multline*}
@@
 \begin{multline*}
-\alpha_{q_0}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out) \ = \ \first(\bar{t}) \ + \\
+\alpha_{q_0}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out) \ = \\ 
+\first(\bar{t}) \ + \
 \bigplus_{(q,a) \,:\, \delta(q,a) = (q_0,\_,\_,\_)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult  s_{q}(\bar{t}')).
@@
 In the same way, it can be seen that for each function $g \in\{s_{q_1},\ldots,s_{q_{\ell}}\}$, at the $\bar{a}$-th iteration of the fixed point operator, it holds that $g(\bar{a}) \leq 1$ and $g(\bar{a}') = 0$.
-From this, we have that at the iteration $\bar{a}'$ of the fixed point operator, $\out(\bar{a}')$ is equal to either $2\cdot \out(\bar{a})$, $2\cdot \out(\bar{a}') + 1$, or $\out(\bar{a}')$, which represents precisely the value in the output tape at each step of $M$ running on the input $\enc(\A)$.
+From this, we have that at the iteration $\bar{a}'$ of the fixed point operator, $\out(\bar{a}')$ is equal to either $2\cdot \out(\bar{a})$, $2\cdot \out(\bar{a}) + 1$, or $\out(\bar{a})$, which represents precisely the value in the output tape at each step of $M$ running on the input $\enc(\A)$.
 From this argument, it can be seen that $\sem{\alpha}(\A) = f(\enc(\A))$, which concludes the proof of the theorem.

### 2018-12-09 LMCS/beyond.tex
all direct replacements by reviewer 1

@@
-%!TEX root = main.tex
+% !TeX root = main.tex
 
@@
 the notion of least fixed point of LFP.
- As a proof of concept, we show that this notion can be used to captures $\fp$.
+ As a proof of concept, we show that this notion can be used to capture $\fp$.
 Moreover, we use this concept
@@
 
-As for the case of LFP, it would be natural to consider the point-wise partial order $\leq$ on $\F$ defined as $f \leq g$ if, and only if, $f(i) \leq g(i)$ for every $i \in \{1, \ldots, \ell\}$, and let the semantics of $\clfp{\beta(\x,h)}$ be the least fixed point of the operator $T_\beta$.
+As for the case of LFP, it would be natural to consider the point-wise partial order $\leq$ on $\F$ defined as $f \leq g$ if, and only if, $f(\bar{a}) \leq g(\bar{a})$ for every $\bar{a} \in A^{\ell}$, and let the semantics of $\clfp{\beta(\x,h)}$ be the least fixed point of the operator $T_\beta$.
 However, $(\F, \leq)$ is not a complete lattice, so we do not have a Knaster-Tarski Theorem ensuring that such a fixed point exists.
@@
 (a) $\support(T^\star(\chi_R)) = \support(\chi_R)$, and  (b) $\support(\chi_R) \subseteq \support(\chi_S)$ for every $S$ such that  $\support(T^\star(\chi_{S})) = \support(\chi_S)$.
-To define a notion of fixed point for $T_\beta$ we simply generalized these conditions.
+To define a notion of fixed point for $T_\beta$ we simply generalize these conditions.
 More precisely, a function $f \in \F$ is a {\em s-fixed point} of $T_{\beta}$ if $\support(T_\beta(f)) = \support(f)$, and $f$ is a {\em least s-fixed point} of $T_{\beta}$ if $f$ is a s-fixed point of $T_\beta$ and for every s-fixed point $g$ of $T_\beta$ it holds that $\support(f) \subseteq \support(g)$.
@@
 Let $k$ be the smallest natural number such that $\support(f_{k}) = \support(T_\beta(f_k))$.
-We have that $f_k$ is a least s-fixed point of $T_\beta$, which is used to defined the semantics of $\clfp{\beta(\x, h)}$.
+We have that $f_k$ is a least s-fixed point of $T_\beta$, which is used to define the semantics of $\clfp{\beta(\x, h)}$.
 More specifically, for an arbitrary first-order assignment $v$ for $\A$:
@@
 \end{exa}
-In contrast with $\lfp$, to reach a fixed point we do not need to impose any positive restriction on the formula $\beta(\x,h)$.
-In fact, since $\beta$ is constructed from monotones operations (sum and product) over the natural numbers, the resulting operator $T_{\beta}$ is monotone as well.
+In contrast to $\lfp$, to reach a fixed point we do not need to impose any positive restriction on the formula $\beta(\x,h)$.
+In fact, since $\beta$ is constructed from monotone operations (sum and product) over the natural numbers, the resulting operator $T_{\beta}$ is monotone as well.
 
@@
 
-Given a relation signature $\R$, the set of transitive $\qfo$ formulae ($\tqfo$-formulae) is defined as an extension of $\qfo$ with the formula $[\pth \psi(\bar{x},\bar{y})]$, where $\psi(\x, \y)$ is an $\fo$-formula over $\R$, and $\bar{x} = (x_1, \ldots, x_k)$, $\bar{y} = (y_1, \ldots, y_k)$ are tuples of pairwise distinct first-order variables.
+Given a relational signature $\R$, the set of transitive $\qfo$ formulae ($\tqfo$-formulae) is defined as an extension of $\qfo$ with the formula $[\pth \psi(\bar{x},\bar{y})]$, where $\psi(\x, \y)$ is an $\fo$-formula over $\R$, and $\bar{x} = (x_1, \ldots, x_k)$, $\bar{y} = (y_1, \ldots, y_k)$ are tuples of pairwise distinct first-order variables.
 The semantics of $[\pth \psi(\bar{x},\bar{y})]$ can easily be defined in terms of $\rqfo(\fo)$ as follows.
@@
 In other words, $\sem{[\pth \psi(\bar{x}, \bar{y})]}(\A,v)$ counts the number of paths from $v(\bar x)$ to $v(\bar y)$ in the graph $\cG_{\psi}(\A)$ whose length is at most~$|A^k|$.
-As it was mentioned before, the operator for counting paths is exactly what we need to capture $\shl$.
+As mentioned before, the operator for counting paths is exactly what we need to capture $\shl$.
 \begin{thm} \label{tqfo-shl}
@@
 The distinction in the language between the Boolean and the quantitative level allows us to define operators at the latter level that cannot be defined at the former. 
-As a example showing how fundamental this separation is, consider the issue of extending $\qfo(\fo)$ at the Boolean level in order to capture $\shl$.
+As an example showing how fundamental this separation is, consider the issue of extending $\qfo(\fo)$ at the Boolean level in order to capture $\shl$.
 The natural alternative to do this is to use $\fo$ extended with a transitive closure operator, which is denoted by $\tc$.

### 2018-12-09 LMCS/extentions.tex
all direct replacements by reviewer 1

@@
-%!TEX root = main.tex
+% !TeX root = main.tex
 
-There exist complexity classes that do not fit in our framework because either the output of a function in not a natural number (e.g.
+There exist complexity classes that do not fit in our framework because either the output of a function is not a natural number (e.g.
 a negative number) or the class is not defined purely in terms of arithmetical operations (e.g.
@@
 	$
-	\alpha_5 :=	\alpha_2 + (-1) \cdot \alpha_1 
-	$.
+	\alpha_5 :=	\alpha_2 + (-1) \cdot \alpha_1. 
+	$ \qed
 \end{exa}
@@
 \end{cor}
-This is an interesting result that shows how robust and versatile is $\qso$ for capturing different counting complexity classes even beyond $\bbN$.
+This is an interesting result that shows how robust and versatile $\qso$ is for capturing different counting complexity classes even beyond $\bbN$.
 
 A different class of functions comes from considering the optimization version of a decision problem.
-For example, one can define MAX-SAT as the problem of determining the maximum number of clauses, of a given CNF propositional formula, that can be made true by an assignment.
+For example, one can define MAX-SAT as the problem of determining the maximum number of clauses, of a given CNF propositional formula that can be made true by an assignment.
 Here, MAX-SAT is defined in terms of a maximization problem which in its essence differs from the functions in $\shp$.
-To formalize this set of optimization problems, Krentel defined $\optp$~\cite{krentel1988complexity} as the class of functions computable by taking the maximum or minimum of the output values over all runs of a polynomial-time NTM machine with output tape (i.e.
+To formalize this class of optimization problems, Krentel defined $\optp$~\cite{krentel1988complexity} as the class of functions computable by taking the maximum or minimum of the output values over all runs of a polynomial-time NTM machine with output tape (i.e.
 each run produces a binary string which is interpreted as a number).
@@
 
-In order to capture classes of optimization functions, we extend as follows $\qso$ with $\max$ and $\min$ quantifiers (called $\optqso$). 
+In order to capture classes of optimization functions, we extend $\qso$ with $\max$ and $\min$ quantifiers as follows (called $\optqso$). 
 Given a signature $\R$, the set of $\optqso$-formulae over $\R$ is given by extending the syntax in (\ref{syntax}) with the following operators:
@@
 	\]
-	Notice that formula $\sa{z} X(z)$ is used to compute the number of nodes in a set $X$. 
+	Notice that formula $\sa{z} X(z)$ is used to compute the number of nodes in a set $X$.  \qed
 \end{exa}
-Similar than for $\maxp$ and $\minp$, we have to distinguished between the $\max$ and $\min$ fragments of $\optqso$.
+Similar than for $\maxp$ and $\minp$, we have to distinguish between the $\max$ and $\min$ fragments of $\optqso$.
 For this, we define the fragment $\maxqso$ of all $\optqso$ formulae constructed from $\qfo$ operators and $\max$-formulae $\max\{\alpha,\alpha\}$, $\maxa{x} \alpha$ and  $\maxa{X} \alpha$.
-The class $\minqso$ is defined analogously changing $\max$ with $\min$.
+The class $\minqso$ is defined analogously replacing $\max$ by $\min$.
 Notice that in $\maxqso$ and $\minqso$, second-order sum and product are not allowed.

### 2018-12-09 LMCS/framework.tex
all direct replacements by reviewer 1

@@
 This division between Boolean and quantitative level is essential for understanding the difference between the logic and the quantitative part. 
-Furthermore, this will allow us later to parametrize both levels of the logic in order to capture different counting complexity classes.
+Furthermore, this will later allow us to parametrize both levels of the logic in order to capture different counting complexity classes.
 
@@
 complexity of evaluating each formula:
-$\alpha_1$ defines an $\fp$-function while $\alpha_2$ defines a $\shp$-complete function.
+$\alpha_1$ defines an $\fp$-function while $\alpha_2$ defines a $\shp$-complete function. \qed
 \end{exa}
 \begin{exa}\label{exa-perm}
-For an example that includes multiplication, let $\bM = \{M(\cdot,\cdot),<\}$ be a vocabulary for storing 0-1 matrices; in particular, a structure $\fM$ over $\bM$ encodes a 0-1 matrix $A$ as follows: if $A[i,j] = 1$, then $M(i,j)$ is true, otherwise $M(i.j)$ is false.
+For an example that includes multiplication, let $\bM = \{M(\cdot,\cdot),<\}$ be a vocabulary for storing 0-1 matrices; in particular, a structure $\fM$ over $\bM$ encodes a 0-1 matrix $A$ as follows: if $A[i,j] = 1$, then $M(i,j)$ is true, otherwise $M(i,j)$ is false.
 Suppose now that we want to compute the permanent of an $n$-by-$n$ 0-1 matrix $A$, that is:
@@
 Finally, the second order sum aggregates these values iterating over all binary relations and calculating the permanent of the matrix.
-We would like to finish with this example by highlighting the similarity of $\alpha_3$ with the permanent formula. 
-Indeed, an advantage of $\qso$-formulae is that the first- and second-order quantifiers in the quantitative level naturally reflect the operations used to define mathematical formulae.
+We would like to finish with this example by highlighting the similarity of $\alpha_3$ to the permanent formula. 
+Indeed, an advantage of $\qso$-formulae is that the first- and second-order quantifiers in the quantitative level naturally reflect the operations used to define mathematical formulae. \qed
 \end{exa}
@@
 We consider several fragments or extensions of $\qso$, which are obtained by restricting the syntax of the Boolean formulae or the use of the quantitative quantifiers.
-In this direction, we denote by $\qfo$ the fragment of $\qso$ where second-order sum and product are not allowed. 
+In this regard, we denote by $\qfo$ the fragment of $\qso$ where second-order sum and product are not allowed. 
 For instance, for the $\qso$-formulae defined in Example \ref{ex:cliques}, we have that $\alpha_1$ is in $\qfo$ and $\alpha_2$ is not.
@@
 In the following section, we use different fragments or extensions of $\qso$ to capture counting complexity classes.
-But before doing this, we show the connection of $\qso$ with previous frameworks for defining functions over relational structures.
+But before doing this, we show the connection of $\qso$ to previous frameworks for defining functions over relational structures.
 
@@
 The main difference is that we restrict the semiring used in WL to natural numbers in order to study counting complexity classes.
-Another difference of WL with our approach is that, to the best of our knowledge, this is the first paper to study weighted logics over general relational signatures, in order  to do descriptive complexity for counting complexity classes.
+Another difference between WL and our approach is that, to the best of our knowledge, this is the first paper to study weighted logics over general relational signatures, in order  to do descriptive complexity for counting complexity classes.
 Previous works on WL usually restrict the signature of the logic to strings, trees, and other specific structures (see \cite{droste2009handbook} for more examples), and they did not study the logic over general structures.

### 2018-12-09 LMCS/preliminaries.tex
all direct replacements by reviewer 1

@@
 We denote by $\ostr[\R]$ the class of all finite ordered $\R$-structures. 
-In this paper we only consider finite ordered structures, so we will usually omit the word finite or ordered when referring to them.
+In this paper we only consider finite ordered structures, so we will usually omit the words finite and ordered when referring to them.
 
@@
 
-We consider several counting complexity classes in this paper, some of the are recalled here~(see \cite{F97,hemaspaandra2013complexity}).
+We consider several counting complexity classes in this paper, some of them are recalled here~(see \cite{F97,hemaspaandra2013complexity}).
 %We define the following function complexity classes: 

### 2018-12-09 LMCS/proof-capture-optp.tex
all direct replacements by reviewer 1

@@
 It is straightforward to prove that $\maxp$ can compute any $\fo$-formula, is closed under first-order sum and product, and second-order maximization. 
-Therefore, condition (1) in Definition~\ref{def:cap} follows similar than in the previous characterizations. Furthermore, one can easily see that the same holds with $\minqso(\fo)$.
+Therefore, condition (1) in Definition~\ref{def:cap} follows similar than in the previous characterizations. Furthermore, one can easily see that the same holds for $\minqso(\fo)$.
 The proof for the other direction is similar than in \cite{kolaitis1994logical} extended with the ideas of Theorem~\ref{theo:capture-fp}.
@@
 $$
-Note that, in contrast with previous proofs, we use $\bar{x} < \bar{y}$ instead of $\bar{y} < \bar{x}$ because the most significant bit in $\enc(U)$ correspond to the smallest tuple in $U$.  
+Note that, in contrast to previous proofs, we use $\bar{x} < \bar{y}$ instead of $\bar{y} < \bar{x}$ because the most significant bit in $\enc(U)$ correspond to the smallest tuple in $U$.  
 It is easy to check that $\Phi(\bar{X}, Y)$ simulates the NP-machine and, if $\Phi(\bar{X}, Y)$ holds, the formula to the right  reconstructs the binary output from the relation in $Y$.
@@
 
-For the case of $\minqso(\fo)$ and a function $f \in \minp$, one has to follow the same approach but considering the $\np$-problem of checking whether $f(\enc(\A)) \leq \val(\enc(U))$.
+For the case of $\minqso(\fo)$ and a function $f \in \minp$, one has to follow the same approach but consider the $\np$-problem of checking whether $f(\enc(\A)) \leq \val(\enc(U))$.
 Then, the formula for describing $f$ is the following:

### 2018-12-09 LMCS/proof-capture-shP.tex
all direct replacements by reviewer 1

@@
 For condition (1), let $\alpha\in\eqso(\fo)$ over some signature $\R$.
-Given a $\fo$ formula $\varphi$, checking whether $\A\models\varphi$ can be done in deterministic polynomial time on the size of $\A$ and the constant function $s$ can be trivially simulated in $\shp$.
+Given an $\fo$ formula $\varphi$, checking whether $\A\models\varphi$ can be done in deterministic polynomial time on the size of $\A$ and any constant function $s$ can be trivially simulated in $\shp$.
 These facts, together with the closures under exponential sum and polynomial product of $\shp$\cite{F97}, suffice to show that the function represented by $\alpha$ is in $\shp$.

### 2018-12-09 LMCS/proof-capture-spanP.tex
all direct replacements by reviewer 1

@@
-To prove the condition (2), we use the fact that $\spp = \#(\eso)$. The condition holds using the same argument as in Proposition~\ref{prop:capture-shP}.
+To prove condition (2), we use the fact that $\spp = \#(\eso)$. The condition holds using the same argument as in Proposition~\ref{prop:capture-shP}.
 For condition (1), notice that given an $\eso$ formula $\varphi$, checking whether $\A\models\varphi$ can be done in non-deterministic polynomial time on the size of $\A$\cite{fagin1974generalized}. 
 Therefore, a $\spp$ machine for $\varphi$ will simulate the non-deterministic polynomial time machine and produce the same string as output in each accepting non-deterministic run.
-Furthermore, the constant function $s$ for some $s\in\nat$ can be trivially simulated in $\spp$ and, thus, condition (1) holds analogously to Proposition \ref{prop:capture-shP} since $\spp$ is also closed under exponential sum and polynomial product~\cite{OH93}.
+Furthermore, any constant function $s$ can be trivially simulated in $\spp$ and, thus, condition (1) holds analogously to Proposition \ref{prop:capture-shP} since $\spp$ is also closed under exponential sum and polynomial product~\cite{OH93}.
 %Similar than the previous proof, we construct recursively a $\spp$ machine $M_{\alpha}$ for each $\eqso(\eso)$ formula $\alpha$ over a signature $\R$.

### 2018-12-09 LMCS/proof-ehorn-pe.tex
all direct replacements by reviewer 1

@@
 Let $\alpha$ be a formula in $\eqso(\ehorn)$ and let $g_{\alpha}$ be the reduction to $\shdhsat$.
-Then the $\totp$ procedure consist in computing $g_{\alpha}(\enc(\A))$ for each input $\enc(\A)$ and then simulate the $\totp$ procedure for $\shdhsat$ on input $g_{\alpha}(\enc(\A))$.
+Then the $\totp$ procedure consist in computing $g_{\alpha}(\enc(\A))$ on input $\enc(\A)$ and then simulating the $\totp$ procedure for $\shdhsat$ on input $g_{\alpha}(\enc(\A))$.
 Therefore, we conclude that $\alpha$ defines a function in~$\totp$.

### 2018-12-09 LMCS/proof-prop-e1-nc.tex
all direct replacements by reviewer 1

@@
-Towards a contradiction, assume that the statement is false, namely, $\E{1}$ is closed under binary sum.
-Consider the formula $\sa{x}(x = x) \in \E{1}$ over some signature $\R$.
-Clearly, for every finite $\R$-structure $\A$ with $n$ elements it holds that $\alpha(\enc(\A)) = n$.
-From our assumption, there exists some formula in $\E{1}$ equivalent to the formula $\alpha \add \alpha$, which describes the function $2n$.
-Let $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ be this formula, where $\varphi$ is a quantifier-free first-order formula.
-Note that the function defined by this formula is equal or greater than the one defined by $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ divided by a polynomial factor.
-More specifically, for each ordered structure $\A$ with $n$ elements, we have the following inequality:
+Towards a contradiction, assume that $\E{1}$ is closed under binary sum. 
+Consider the formula $\alpha = \sa{x}(x = x) \in \E{1}$ over some signature $\R$. 
+This defines the function $\sem{\alpha}(\A) = \size{\A}$. 
+From our assumption, there exists some formula in $\E{1}$ equivalent to the formula $\alpha \add \alpha$, which describes the function $2\size{\A}$. 
+Let $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ be this formula, where $\varphi$ is first-order and quantifier-free. 
+For each $\R$-structure $\A$, we have the following inequality:
 $$
@@
 \leq 
-\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot  n^{\length{\bar{y}}} \leq 2n^{\length{\bar{y}}+1} 
+\sem{\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})}(\A) \cdot  \size{\A}^{\length{\bar{y}}} \leq 2\size{\A}^{\length{\bar{y}}+1} 
 $$
-Note that the formula $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ defines a function in $\E{0}$ and it was shown by the authors in \cite{SalujaST95} that every function in $\E{0}$ grows exponentially over the size of the structure for large enough structures, when $\length{\bar{X}} > 0$.
-This contradict the above inequality and, therefore, for $\sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ we have that $\length{\bar{X}} = 0$.
+Note that the formula $\sa{\bar{X}}\sa{\bar{x}}\sa{\bar{y}}\,\varphi(\bar{X},\bar{x},\bar{y})$ defines a function in $\E{0}$. 
+Therefore, as it was proven in \cite{SalujaST95}, if $\length{\bar{X}} > 0$ then the function is in $\Omega(2^{\size{\A}})$, which violates the inequality for large structures.
 
-Now, if the formula for defining $\alpha + \alpha$ is of the form $\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})$ consider a structure $\mathfrak{1}$ with only one element $a$.
-We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$, but the only possible assignment to $\bar{x}$ is the tuple $(a,\ldots,a)$ so $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which leads to a contradiction.
+We now have that $\length{\bar{X}} = 0$.
+Consider a structure $\mathfrak{1}$ with only one element.
+We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$ $\sem{\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which implies that $\length{\bar{x}} > 0$.
+But since the structure has only one element, there is only one possible assignment to $\bar{x}$.
+And so, $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which leads to a contradiction.

### 2018-12-09 LMCS/proof-prop-rest.tex
all direct replacements by reviewer 1

@@
 Now we show that $\QE{0} \not\subseteq \E{1}$.
-In Theorem \ref{theo-pi1-pnf} we proved that there is no formula in $\loge{1}$-PNF equivalent to the formula $\alpha = 2$. 
-Every formula in $\E{1}$ can be expressed in $\loge{1}$-PNF, which implies that $2 \in \QE{0}$ and $2 \not\in \E{1}$.
+Note that every formula in $\E{1}$ can be expressed in $\loge{1}$-PNF.
+However, in Theorem \ref{theo-pi1-pnf} we proved that there is no formula in $\loge{1}$-PNF equivalent to the formula $\alpha = 2$.
+We obtain that $2 \in \QE{0}$ and $2 \not\in \E{1}$.
 
@@
 For inclusion, let $\alpha$ be a formula in $\eqso(\loge{1})$. 
-Suppose that it is in $\loge{1}$-SNF, namely, $\alpha = c + \sum_{i = 1}^{n}\alpha_i$. 
-Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\ex{\bar{y}}\varphi_i(\bar{X},\bar{x},\bar{y})$, where $\varphi_i$ is quantifier-free for each $\alpha_i$. 
-We use the same construction used in \cite{SalujaST95}, and we obtain that the formula $\ex{\bar{y}}\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,[\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \fa{\bar{y}'}(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')]$ for every assignment to $(\bar{X},\bar{x})$.
-Doing this replacement for each $\alpha_i$ renders an equivalent formula in $\eqso(\logu{1})$.
+Suppose that it is in $\loge{1}$-SNF, namely, $\alpha = \sum_{i = 1}^{n}\alpha_i$. 
+Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\ex{\bar{y}}\varphi_i(\bar{X},\bar{x},\bar{y})$, where each $\varphi_i$ is quantifier-free. 
+We use the same construction used in \cite{SalujaST95} for each $\alpha_i$, and we obtain that the formula $\ex{\bar{y}}\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,[\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \fa{\bar{y}'}(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')]$ for every assignment to $(\bar{X},\bar{x})$.
+We do this replacement for each $\alpha_i$, and we obtain an equivalent formula to $\alpha$ in $\eqso(\logu{1})$.
 
@@
 Suppose that there exists an equivalent formula $\alpha$ in $\eqso(\loge{1})$. 
-Also, suppose that it is in $\L$-PNF, so $\alpha = \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\ex{\bar{y}}\varphi_i(\bar{X},\bar{x},\bar{y})$. 
+Also, suppose that it is in $\LL$-SNF, so $\alpha = \sum_{i = 1}^n\sa{\bar{X}}\sa{\bar{x}}\ex{\bar{y}}\varphi_i(\bar{X},\bar{x},\bar{y})$. 
 Consider a structure $\A'$ with one element. 

### 2018-12-09 LMCS/proof-qe0-fp-qe1-totp-fptras.tex
all direct replacements by reviewer 1

@@
 Finally, we show that $\eqso(\loge{1})$ is closed under sum and multiplication. 
-Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only in proving that it is closed under multiplication.
-We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of $\so$.
+Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only on proving that it is closed under multiplication. 
+We prove this for the more general case of $\eqso(\LL)$ with $\LL$ being a fragment of SO.
 
@@
 \proof
-	Given two formulas $\alpha, \beta$ in $\eqso(\LL)$ we will show a formula in the grammar which is equivalent to $(\alpha\mult \beta)$.
+	Given two formulas $\alpha, \beta$ in $\eqso(\LL)$ we will construct a formula in the logic which is equivalent to $(\alpha\mult \beta)$. 
 	From what was proven in Proposition \ref{theo-pnf-snf}, we may assume that $\alpha$ and $\beta$ are in $\LL$-SNF. 
@@
 \qed
-Since $\loge{1}$ is closed under conjunction, then Lemma~\ref{conj-mult} holds for $\eqso(\loge{1})$. This concludes the proof.
+Since $\loge{1}$ is closed under conjunction, then Lemma~\ref{conj-mult} holds for $\eqso(\loge{1})$. 
+This concludes the proof.

### 2018-12-09 LMCS/proof-rqfo-fo-cap.tex
all direct replacements by reviewer 1

@@
  (1) writing 0 and moving the head one cell to the right, (2) writing 1 and moving the head one cell to the right, or (3) doing nothing.
-These operations are represented by the set $\{0,1,\emptyset\}$.
+These operations are represented by 0, 1, and $\emptyset$, respectively.
 Finally, assume that $M$, on input $\enc(\A)$ with domain $A = \{1,\dots,n\}$, executes exactly $n^k$ steps for a fixed $k \geq 1$.
@@
 For example, $\alpha_{T_0}$ is used to define the values of function $T_0$.
-Notice that the values of each function $f$ in $\alpha$ depend on the values of the other functions, that is why we talk about a simultaneous computation.
-Besides, notice that notation $\clfp{\out(\bar{t}) : \ldots}$ is used to indicate that (1) the free variables of the formula are $\bar t = (t_1, \ldots, t_k)$, and (2) once the least fixed point has been computed, the value of $\clfp{\out(\bar{t}) : \ldots}$ for an assignment $\bar a$ to $\bar t$ is given by the value of function $\out$ on $\bar a$ in the fixed point.
-Finally, it is important to notice that $\alpha$ can be transformed into a proper $\rqfo(\fo)$ formula by using the same techniques to prove that a simultaneous LFP can be transformed into a single LFP~\cite{L04}; in particular, functions symbols $T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out$ are replaced by a single function $f$ of arity $6 \cdot 2 \cdot k + (\ell + 1) \cdot k + k = (\ell+ 14) \cdot k$.
+Notice that the values of each function in $\alpha$ depend on the values of the other functions, which is why we talk about a simultaneous computation.
+Besides, notice that the notation $\clfp{\out(\bar{t}) : \ldots}$ is used to indicate that (1) the free variables of the formula are $\bar t = (t_1, \ldots, t_k)$, and (2) once the least fixed point has been computed, the value of $\clfp{\out(\bar{t}) : \ldots}$ for an assignment $\bar a$ to $\bar t$ is given by the value (in the fixed point) of function $\out$  on $\bar a$.
+Finally, it is important to notice that $\alpha$ can be transformed into a proper $\rqfo(\fo)$ formula by using the same techniques used to prove that simultaneous LFP has the same expressiveness as LFP~\cite{L04}; in particular, functions symbols $T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out$ are replaced by a single function $f$ of arity $6 \cdot 2 \cdot k + (\ell + 1) \cdot k + k = (\ell+ 14) \cdot k$.
 
@@
 Function $\out$ stores the output of the TM $M$; in particular, $\out(\bar t)$ is the value returned by $M$ when $\bar t$ is the last step.
-Finally, assuming that $\varphi_{\text{\rm last}}(x)$ is an $\fo$-formulae defining the last element of $<$, we have that $\last(\bar t) = \bigwedge_{i=1}^n \varphi_{\text{\rm last}}(t_i)$, that is, $\last(\bar t)$ holds if $\bar t$ is the last step.
+Finally, assuming that $\varphi_{\text{\rm last}}(x)$ is an $\fo$-formula defining the last element of $<$, we have that $\last(\bar t) = \bigwedge_{i=1}^k \varphi_{\text{\rm last}}(t_i)$, that is, $\last(\bar t)$ holds if $\bar t$ is the last step.
 Therefore, the use of $\last(\bar t)$ in $\alpha$ allows us to return the output of the TM $M$.
 
-As in Example \ref{ex:count-path}, assume that $\varphi_{\text{\rm first}}(x)$ and $\varphi_{\text{succ}}(x,y)$ are $\fo$-formulae defining the first element of $<$ and the successor relation associated to $<$, respectively.
-Moreover, assume that formulae $\varphi_{\text{succ}}(x,y)$ and $\varphi_{\text{\rm last}}(x)$ are used to defined
-an $\fo$-formula $\succesor(\bar{t},\bar{t'})$ such that $\succesor(\bar{t},\bar{t'})$ holds if $\bar t'$ is the successor step of $\bar t$ (that is, $\succesor(\cdot,\cdot)$ is the successor relation associated to the lexicographical order induced by $<$ on the tuples with $k$ elements).
+As in Example \ref{ex:count-path}, assume that $\varphi_{\text{\rm first}}(x)$ is an $\fo$-formula defining the first element of $<$.
+Moreover, assume that $\first(\bar t)$ and $\succesor(\bar{t},\bar{t'})$ are $\fo$-formulae such that
+$\first(\bar t)$ holds if $\bar t$ is the first step and $\succesor(\bar{t},\bar{t'})$ holds if $\bar t'$ is the successor step of $\bar t$ (that is, $\succesor(\cdot,\cdot)$ is the successor relation associated to the lexicographical order induced by $<$ on the tuples with $k$ elements).
 Then formula $\alpha_{T_{\vdash}}$ is defined as follows:
@@
 This is the reason why we use tuple $\bar y$ and define $\bar x$ as the successor of $\bar y$.
-Formulas $\alpha_{h}$ is defined as follows:
+Formula $\alpha_{h}$ is defined as follows:
 \begin{align*}

### 2018-12-09 LMCS/proof-sigmafo-minusone.tex
all direct replacements by reviewer 1

@@
 \noindent {\bf Easy decision version and FPRAS.}  We show here that $\eqso(\logex{1}) \subseteq \totp$ and every function in $\eqso(\logex{1})$ has an FPRAS.
-We do this by showing a parsimonious reduction from a function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
-First, we define a function that converts a formula $\alpha$ in $\eqso(\logex{1})$ over a signature $\R$ into a formula $\lambda(\alpha)$ in $\eqso(\loge{1})$ over a signature $\R_{\alpha}$.
+We do this by showing a parsimonious reduction from any function in $\eqso(\logex{1})$ to some function in $\eqso(\loge{1})$, and using the result of Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
+First, we define a function that converts any formula $\alpha$ in $\eqso(\logex{1})$ over a signature $\R$ into a formula $\lambda(\alpha)$ in $\eqso(\loge{1})$ over a signature $\R_{\alpha}$.
 Afterwards, we define a function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an $\R_{\alpha}$-structure $g_{\alpha}(\A)$.
@@
 Let $\A$ be a $\R$-structure with domain $A$.
-Let $\A'$ be an $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A}$ empty.
-For each $\fo$-formula $\psi(\z)$ with $\length{\bar{z}}$ open first-order variables, we iterate for every tuple $\bar{a} \in A^{\length{\bar{z}}}$.
+Let $\A'$ be an $\R_{\alpha}$-structure obtained by copying $\A$ and leaving each $R_{\psi}^{\A'}$ empty.
+For each $\fo$-formula $\psi(\z)$ with $\length{\bar{z}}$ open first-order variables, we iterate over all tuples $\bar{a} \in A^{\length{\bar{z}}}$.
 If $\A\models\psi(\bar{a})$ (this can be done in $\ptime$), then the tuple $\bar{a}$ is added to $R_{\psi}^{\A'}$ .
@@
 For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$-formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
-Without lost of generality, we assume that $\alpha$ is in $\logex{1}$-SNF, that is, $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
+Without loss of generality, we assume that $\alpha$ is in $\logex{1}$-SNF, that is, $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
 Moreover, we assume that $\length{\bar{x}} > 0$  since, if this is not the case, we can replace $\sa{\bar{X}}\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
@@
 In the first part, we will assume that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$.
-Then we will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
+We will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
 From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ will be the desired formula.
@@
 
-\noindent {\em Part (1).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ where $\varphi$ is an  $\fo$-formula.
+\noindent {\em Part (1).}  Let $\alpha =  \sa{\bar{X}}\sa{\bar{x}} \varphi(\bar{X},\bar{x})$ where $\varphi$ is a $\logex{1}$-formula.
 Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}} \varphi(\bar{x})$ (i.e.
@@
 
-Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ are mentioned in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the formal definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
+Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ is mentioned in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the formal definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
 We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
@@
 \begin{clm}\label{claim:minusone}
-	For an ordered structure $\A$ and a FO assignment $v$ for $\A$, $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ if, and only if, there exists a SO assignment $V$ for $\A$ such that $(\A,V,v)\models \varphi_i'(\bar{X},\bar{x}, \bar{y})$.
+	For an ordered structure $\A$ and an FO assignment $v$ for $\A$, $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ if, and only if, there exists an SO assignment $V$ for $\A$ such that $(\A,V,v)\models \varphi_i'(\bar{X},\bar{x}, \bar{y})$.
 \end{clm}
@@
 
-The previous claim and proof motivates the following definitions.
+The previous claim and proof motivate the following definition.
 For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length{\bar{X}}})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
@@
 \bigvee_{X \in \bar{X}} \ex{\bar{z}} [ \, X(\bar{z}) \wedge   \!\!\!\!\!\!\!\!\!\!
-\bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{v})} \!\!\!\!\!\!\!\!\!\!\!
+\bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{y})} \!\!\!\!\!\!\!\!\!\!\!
 \bar{w}\neq\bar{z} \, ] \; \big) \;  \vee   \label{eq:part2} \\
@@
 Indeed, by Claim~\ref{claim:minusone} we know that if $\ex{\bar{x}} \ex{\bar{y}} \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi_i'(\bar{X},\bar{x},\bar{y})$.
-Then, the conclusion of the implication (divided into (\ref{eq:part2}) and (\ref{eq:part3})), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
+The conclusion of the implication (divided into (\ref{eq:part2}) and (\ref{eq:part3})), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
 Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by defining that if $\varphi_i'(\bar{X},\bar{x}, \bar{y})$ is satisfied then $V^*(\bar{X}) \subsetneq V(\bar{X})$. 
@@
 where each modified disjunct $\varphi_i^*(\bar{X},\bar{x})$ is constructed as follows. 
-For the sake of simplification, define the auxiliary formula $\chi_i = \neg \ex{\bar{x}} \ex{\bar{y}}\varphi^{\fo}_i(\bar{x},\bar{y})$. 
+For the sake of simplicity, define the auxiliary formula $\chi_i = \neg \ex{\bar{x}} \ex{\bar{y}}\varphi^{\fo}_i(\bar{x},\bar{y})$. 
 This formula basically checks if $\varphi_i$ is not satisfiable (recall Claim~\ref{claim:minusone}).
@@
 Also, consider the previously defined formulas $\chi_i$ and $\psi_i$, for each $i \leq m$.
-We construct a function $\lambda$ that receives a quantitative formula $\beta$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if, and only if, $\sem{\beta}(\A) = 0$.
+We construct a function $\lambda$ that receives a formula $\beta \in \eqso(\logex{1})$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if, and only if, $\sem{\beta}(\A) = 0$.
 If $\beta = \sa{\bar{x}} \varphi(\bar{x})$, then $\lambda(\beta) = \neg \ex{\bar{x}'} \varphi(\bar{x}')$.

### 2018-12-09 LMCS/proof-theo-pi1-pnf.tex
all direct replacements by reviewer 1

@@
-%!TEX root = main.tex
+% !TeX root = main.tex
 
-From now on, for every first-order tuple $\bar{x}$ or second-order tuple $\bar{X}$ we write $\length{\bar{x}}$ or $\length{\bar{X}}$ as the size of $\bar{x}$ or $\bar{X}$ respectively.
+From now on, for every first-order tuple $\bar{x}$ or second-order tuple $\bar{X}$ we write $\length{\bar{x}}$ or $\length{\bar{X}}$ as the number of variables in $\bar{x}$ or $\bar{X}$ respectively. 
 We divide the proof in three parts.
@@
 
-Second, we prove that the formula $\alpha_{1} = 2$ (i.e.
-$\sem{\alpha_{1}}$ is the constant function $2$) is not equivalent to any formula in $\loge{1}$-PNF.
+Second, we prove that the formula $\alpha_{1} = 2$ (i.e. $\sem{\alpha_{1}}$ is the constant function $2$) is not equivalent to any formula in $\loge{1}$-PNF. 
 Suppose that there exists some formula $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\, \varphi(\bar{X},\bar{x},\bar{y})$ in $\loge{1}$-PNF that is equivalent to $\alpha_1$. 
@@
 Therefore, suppose that $\length{\bar{X}} > 0$ or $\length{\bar{x}} > 0$, and consider some ordered structure $\A$. 
-Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
+Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for $i\in\{1,2\}$: $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
 Now consider the ordered structure $\A'$ that is obtained by taking the disjoint union of $\A$ twice. 
@@
 Note that $\A'\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$ for $i = 1,2$ and there exists a third assignment $(\bar{B}_1',\bar{b}_1',\bar{a}_1')$ that is isomorphic to $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, in the other half of the structure, such that $\A'\models\varphi(\bar{B}_1',\bar{b}_1',\bar{a}_1')$.
-We have that $\sem{\alpha}(\A') \geq 3$ which leads to a contradiction.
+As a result, we have that $\sem{\alpha}(\A') \geq 3$ which leads to a contradiction.
 
-For the latter part of the proof, we show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there exists an equivalent formula in $\LL$-PNF.
-Similarly to Theorem~\ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such a formula.
+For the last part of the proof, we show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there exists an equivalent formula in $\LL$-PNF.
+Similarly as in the proof of Theorem~\ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such a formula. 
 Assume that $\alpha = \sum_{i = 1}^n \alpha_i$ is in $\LL$-SNF where each $\alpha_i$ is in $\LL$-PNF. 
 Without loss of generality, we assume that each $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi_i(\bar{X},\bar{x})$ with $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$. 
-If not, we can replace each $\alpha_i$ by the equivalent formula
+If that is not the case, we can replace each $\alpha_i$ by the equivalent formula
 $$
@@
 To show that the formula is indeed equivalent to $\alpha$, note that the formulas in lines (\ref{eq:partition1}) and (\ref{eq:partition2}) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$, and the formulas in lines (\ref{eq:partition3}) and (\ref{eq:partition4}) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
-Altogether, the four lines define pairwise disjoint assignments for $(\bar{X},\bar{x}),(\bar{Y},\bar{y})$.
+Altogether the four lines define pairwise disjoint assignments for $(\bar{X},\bar{x}),(\bar{Y},\bar{y})$. 
 With this, it is straightforward to show that the above formula is equivalent to $\alpha$. 

### 2018-12-09 LMCS/proof-tqfo-shl.tex
all direct replacements by reviewer 1

@@
 Let $\R$ be a relational signature and $\alpha$ a formula over $\R$ in $\tqfo(\fo)$.
-Next we construct a logarithmic-space nondeterministic Turing Machine $M_{\alpha}$ that on input $(\enc(\A),v)$, where $\A$ is an $\R$-structure and $v$ is a first-order assignment for $\A$, has $\sem{\alpha}(\A,v)$ accepting runs (so that we can conclude that the function defined by $\alpha$ is in $\shl$).
+We construct a logarithmic-space nondeterministic Turing Machine $M_{\alpha}$ that on input $(\enc(\A),v)$, where $\A$ is an $\R$-structure and $v$ is a first-order assignment for $\A$, has $\sem{\alpha}(\A,v)$ accepting runs (so that we can conclude that the function defined by $\alpha$ is in $\shl$).
 Suppose that the domain of $\A$ is $A = \{1,\ldots,n\}$.
@@
 If $\alpha = \sa{x}\beta$, for each $a\in A$ we generate a different run where we simulate $M_{\beta}$ with input $v[a/x]$.
-If $\alpha = \pa{x}\beta$, we simulate $M_{\beta}$ with input $v[1/n]$, and on each accepting run, instead of accepting we replace the assignment of $x$ to 2, to simulate $M_{\beta}$ with input $v[2/x]$, and so on.
+If $\alpha = \pa{x}\beta$, we simulate $M_{\beta}$ with input $v[1/x]$, and on each accepting run, instead of accepting we replace the assignment of $x$ to 2, to simulate $M_{\beta}$ with input $v[2/x]$, and so on.
 If $\alpha = [\pth \varphi(\bar{x},\bar{y})]$, where $\varphi$ is an $\fo$-formula, then we simulate the $\shl$ procedure that counts the number of paths of a given length from a source to a target node in an input graph (where the length is at most the number of nodes in the graph).
@@
 $$
-\varphi = \exists\bar{u}\exists\bar{z}(\first(\bar{u})\wedge \psi_{\bf acc}(\bar{z})\wedge[{\bf tc}_{\bar{x},\bar{y}}\,\psi_{\bf next}(\bar{x},\bar{y})](\bar{u},\bar{z})),
+\varphi = \exists\bar{u}\exists\bar{z}(\psi_{\text{initial}}(\bar{u})\wedge \psi_{\text{acc}}(\bar{z})\wedge[{\bf tc}_{\bar{x},\bar{y}}\,\psi_{\text{next}}(\bar{x},\bar{y})](\bar{u},\bar{z})),
 $$
-where $\psi_{\bf acc}(\bar{z})$ is an $\fo$-formula that indicates that $\bar{z}$ is an accepting configuration, and $\psi_{\bf next}(\bar{x},\bar{y})$ is an $\fo$-formula that indicates that $\bar{y}$ is a successor configuration of $\bar{x}$~\cite{G07}.
+where $\psi_{\text{initial}}(\bar{u})$ is an $\fo$-formula that indicates that $\bar{u}$ is the initial configuration, $\psi_{\text{acc}}(\bar{z})$ is an $\fo$-formula that indicates that $\bar{z}$ is an accepting configuration, and $\psi_{\text{next}}(\bar{x},\bar{y})$ is an $\fo$-formula that indicates that $\bar{y}$ is a successor configuration of $\bar{x}$~\cite{G07}.
 Here, there is a one-to-one correspondence between configurations of $M$ and assignments to $\bar{z}$.
-As a consequence, given a structure $\A$ and a first-order assignment $v$ for $\A$, where $v(\bar{x})$ is the starting configuration and $v(\bar{y})$ is the sole accepting configuration, the value of $\sem{[\pth\psi_{\bf next}(\bar{x},\bar{y})]}(\A,v)$ is equal to $\tma_M(\enc(\A))$.
+As a consequence, given a structure $\A$ and a first-order assignment $v$ for $\A$, where $v(\bar{x})$ is the starting configuration and $v(\bar{y})$ is the sole accepting configuration, the value of $\sem{[\pth\psi_{\text{next}}(\bar{x},\bar{y})]}(\A,v)$ is equal to $\tma_M(\enc(\A))$.
 Therefore, the $\tqfo(\fo)$-formula
 $
-\alpha = \sa{\bar{u}}\sa{\bar{z}}(\first(\bar{u})\mult\psi_{\bf acc}(\bar{z})\mult[\pth \psi_{\bf next}(\bar{u},\bar{z})])
+\alpha = \sa{\bar{u}}\sa{\bar{z}}(\psi_{\text{initial}}(\bar{u})\mult\psi_{\text{acc}}(\bar{z})\mult[\pth \psi_{\text{next}}(\bar{u},\bar{z})])
 $

### 2018-12-09 LMCS/saluja.tex
all direct replacements by reviewer 1

@@
 
-Inspired by the connection between $\shp$ and $\sfo$, a hierarchy of subclases of $\sfo$ was introduced in~\cite{SalujaST95} 
+Inspired by the connection between $\shp$ and $\sfo$, a hierarchy of subclasses of $\sfo$ was introduced in~\cite{SalujaST95} 
 by restricting the alternation of quantifiers in Boolean formulae.

### 2018-12-09 LMCS/syntactic-clo.tex
all direct replacements by reviewer 1

@@
 However, if $\shp$ is closed under subtraction by one, then it holds that $\np \subseteq \cspp$ \cite{OH93}.}.
-So, the following proposition rules out any logic that extends $\Pi_1$ for a possible extension of~$\QE{1}$ with the desired closure property.
+So, the following proposition rules out any logic that extends $\Pi_1$ as a possible extension of~$\QE{1}$ with the desired closure property.
 \begin{prop} \label{pi-minusone}
@@
 Therefore, the desired extension has to be achieved by allowing some local extensions to~$\Sigma_1$.
-More precisely, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulae over a signature~$\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mentioned any second-order variable).
+More precisely, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulae over a signature~$\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mention any second-order variable).
 With this extension we obtain a class with the desired properties.

### 2018-12-09 LMCS/syntactic-horn.tex
all direct replacements by reviewer 1

@@
 
-In the previous section, we show that the class $\eqso(\logex{1})$ has good closure and approximation properties.
+In the previous section, we showed that the class $\eqso(\logex{1})$ has good closure and approximation properties.
 Unfortunately, it is not clear whether it admits a {\em natural} complete problem under parsimonious reductions, where {\em natural} means any of the counting problems defined in this section or any other well-known counting problem (not one specifically designed to be complete for the class).
@@
 More importantly, $\psi(\pT,\pA)$ resembles a conjunction of Horn clauses except for the use of negative literals of the form $\ex{v} \neg \textit{A}(u,v)$.
+\qed
 \end{exa}
@@
 \end{prop}
-For the sake of simplification, we postpone the proof the previous proposition after the proof of Theorem~\ref{sigma2hard}.
+For the sake of simplicity, we postpone the proof of the previous proposition to after the proof of Theorem~\ref{sigma2hard}.
 

### 2018-12-09 LMCS/syntactic.tex
all direct replacements by reviewer 1

@@
 %!TEX root = main.tex
+
+The class $\shp$ was introduced in \cite{Valiant79} to prove that computing the permanent of a matrix, as defined in Example \ref{exa-perm}, is a $\shp$-complete problem.
+As a consequence of this result many counting problems have been proved to be $\shp$-complete~\cite{V79b,arora2009computational}.
+%\cite{V79b,PB83,P86,L86,BW91,HMRS98,DG00,BW05,DS12, PS13,PS14}.
+Among them, problems having easy decision counterparts play a fundamental role, as a counting problem with a hard decision version is expected to be hard.
+Formally, the decision problem associated to a function $f : \Sigma^* \to \mathbb{N}$ is defined as $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$, and $f$ is said to have an easy decision version if $L_f \in \ptime$.
+Many prominent examples satisfy this property, like computing the number of: perfect matchings of a bipartite graph ($\cpm$)~\cite{Valiant79}, satisfying assignments of a DNF ($\cdnf$)~\cite{DHK05,KL83} or Horn ($\chsat$)~\cite{V79b} propositional formula, among others.
+
+Counting problems with easy decision versions play a fundamental role in the search for efficient approximation algorithms for functions in $\shp$.
+A fully-polynomial randomized approximation scheme (FPRAS) for a function $f : \Sigma^* \to \bbN$ is a randomized algorithm $\mathcal{A} : \Sigma^* \times (0,1) \to \bbN$ such that: (1) for every string $x \in \Sigma^*$ and real value $\varepsilon \in (0,1)$, the probability that $|f(x) - \mathcal{A}(x,\varepsilon)| \leq \varepsilon \cdot f(x)$ is at least $\frac{3}{4}$, and (2) the running time of $\mathcal{A}$ is polynomial in the size of $x$ and $1/\varepsilon$ \cite{KL83}.
+Notably, there exist $\shp$-complete functions that can be efficiently approximated as they admit FPRAS; for instance, there exist FPRAS for $\cdnf$~\cite{KL83} and $\cpm$~\cite{JSV04}.
+A key observation here is that if a function $f$ admits an FPRAS, then $L_f$ is in the randomized complexity class $\bpp$~\cite{G77}.
+Hence, under the widely believed assumption that  $\np \not\subseteq \bpp$, we cannot hope for an FPRAS for a function in $\shp$ whose decision counterpart is $\np$-complete, and we have to concentrate on the class of counting problems with easy decision versions.
+
+The importance of the class of counting problems with easy decision counterparts has motivated the search for robust classes of functions in $\shp$ with this property \cite{PagourtzisZ06}.
+But the key question here is what should be considered a {\em robust} class.
+A first desirable condition has to do with the closure properties satisfied by the class, which is a common theme when studying function complexity classes \cite{OH93,FH08}.
+Analogously to the cases of $\ptime$ and $\np$, which are closed under intersection and union, we expect our class to be closed under multiplication and sum.
+For a more elaborated closure property, assume that $\textit{sat\_one}$ is a function that returns one plus the number of satisfying assignments of a propositional formula.
+Clearly $\textit{sat\_one}$ is a $\shp$-complete function whose decision counterpart $L_{\textit{sat\_one}}$ is trivial.
+But should $\textit{sat\_one}$ be part of a robust class of counting functions with easy decision versions?
+The key insight here is that if a function in $\shp$ has an easy decision counterpart $L$, then as $L \in \np$ we expect  to have a polynomial-time algorithm that verifies whether $x \in L$ by constructing witnesses for~$x$.
+%As an example of this considered all the counting functions mentioned in the previous paragraphs and, in particular, the decision counterpart of $\cdnf$ that can be solved by explicitly constructing in polynomial-time a satisfying assignment for an input propositional formula.
+Moreover, if such an algorithm for constructing witnesses exists, then we also expect to be able to manipulate such witnesses and in some cases to remove them.
+In other words, we expect a robust class $\CC$ of counting functions with easy decision versions to be closed under subtraction by one, that is, if $g \in \CC$, then the function $g \dotdiv 1$ should also be in $\CC$, where $(g \dotdiv 1)(x)$ is defined as $g(x) - 1$ if $g(x) \geq 1$, and as~$0$ otherwise.
+Notice that, unless $\ptime = \np$, no such class can contain the function $\textit{sat\_one}$ because $\textit{sat\_one} \dotdiv 1$ counts the number of satisfying assignments of a propositional formula.
+
+A second desirable condition of robustness is the existence of natural complete problems~\cite{P94}.
+Special attention has to be paid here to the notion of reduction used for completeness.
+Notice that under the notion of Cook reduction, originally used in \cite{Valiant79}, the problems $\cdnf$ and $\csat$ are $\shp$-complete.
+However, $\cdnf$ has an easy decision counterpart and admits an FPRAS, while $\csat$ does not satisfy these conditions unless $\ptime = \np$.
+Hence a more strict notion of reduction has to be considered; in particular, the notion of parsimonious reduction (to be defined later) satisfies that if a function $f$ is parsimoniously reducible to a function $g$, then $L_g \in \ptime$ implies that $L_f \in \ptime$ and the existence of an FPRAS for $g$ implies the existence of a FPRAS for $f$.
+
+In this section, we use the framework developed in this paper to address the problem of defining a robust class of functions with easy decision versions.
+More specifically, we use the framework to introduce in Section \ref{sec-hier-shp} a syntactic hierarchy of counting complexity classes contained in $\shp$.
+Then this hierarchy is used in Section \ref{sec-clo} to define a class of functions with easy decision versions and good closure properties, and in Section \ref{sec-horn} to define a class of functions with easy decision versions and
+natural complete problems.
 

### 2018-12-09 LMCS/todo-list.tex
all direct replacements by reviewer 1

@@
 \item p.10, last line: ``$B$-th bit'': mention in what way $B$ is interpreted as a number
-\item p.12: ``To formalize this set of optimization problems'': class instead of set
+\item[$\checkmark$] p.12: ``To formalize this set of optimization problems'': class instead of set
 \item p.14: ``we have to concentrate on the class of counting problems with easy decision versions'':
@@
 does not follow from the argument before as stated in this sentence, but is trivial
-\item p.18: ``Also, suppose that it is in $\L$-PNF'': $\L$ should be $\LL$ , PNF should be SNF
-\item p.19, end of proof of Prop. 5.5: ``which violates the inequality'' $\to$ ``which violates the inequality
+\item[$\checkmark$] p.18: ``Also, suppose that it is in $\L$-PNF'': $\L$ should be $\LL$ , PNF should be SNF
+\item[$\checkmark$] p.19, end of proof of Prop. 5.5: ``which violates the inequality'' $\to$ ``which violates the inequality
 for large structures''
-\item p.20: ``we will show a formula in the grammar'' $\to$ ``... in the logic''
+\item[$\checkmark$] p.20: ``we will show a formula in the grammar'' $\to$ ``... in the logic''
 \item p.20: ``Since $\Sigma_1$ is closed under conjunction, then Lemma 5.7 holds for ...'' $\to$ ``Since $\Sigma_1$ is
 closed under cnojunction, Lemma 5.7 can be applied to ...'' (or similar)
-\item p.22: ``and leaving each $R_{\psi}^{\A}$ empty'' $\to$ $R_{\psi}^{\A'}$
+\item[$\checkmark$] p.22: ``and leaving each $R_{\psi}^{\A}$ empty'' $\to$ $R_{\psi}^{\A'}$
 \item p.22: ``To show that $\alpha$ is in \totp'' $\to$ ``To show that the function defined by $\alpha$ ...'' (or similar)
-\item p.22: ``Let $\alpha$ = ... where $\varphi$ is an FO-formula'' $\to$ ``... where $\varphi$ is an $\Sigma_1$[FO]-formula''
+\item[$\checkmark$] p.22: ``Let $\alpha$ = ... where $\varphi$ is an FO-formula'' $\to$ ``... where $\varphi$ is an $\Sigma_1$[FO]-formula''
 \item p.23, formulae: Maybe put (large) parantheses around the relativization of quantifiers, for
 example the part of formula $\varphi_i$ after the quantifiers
-\item p.24, formula 5.6: subscript of large conjunction: $\varphi_i^{+}(\bar{X},\bar{v})$ $\to$ $\varphi_i^{+}(\bar{X},\bar{y})$
-\item p.25: ``We construct a function $\lambda$ that receives a quantitative formula $\beta$ ...'' $\to$ ``... that receives
+\item[$\checkmark$] p.24, formula 5.6: subscript of large conjunction: $\varphi_i^{+}(\bar{X},\bar{v})$ $\to$ $\varphi_i^{+}(\bar{X},\bar{y})$
+\item[$\checkmark$] p.25: ``We construct a function $\lambda$ that receives a quantitative formula $\beta$ ...'' $\to$ ``... that receives
 a formula $\beta \in \eqso(\logex{1})$ ...''
@@
 DelayP. Maybe change the citation or give a short proof sketch.
-\martin{Punto posiblemente largo}
-\item p.31, definition of $\leq$ on $\F$: ``$f(i) \leq g(i)$ for every $i \in \{1,\ldots,\ell\}$'': Inputs for $f$ and $g$ should be
+\item[$\checkmark$] p.31, definition of $\leq$ on $\F$: ``$f(i) \leq g(i)$ for every $i \in \{1,\ldots,\ell\}$'': Inputs for $f$ and $g$ should be
 elements of $A^{\ell}$, not $\{1,\ldots,\ell\}$
@@
 	WL to the natural numbers as fixed semiring, calling this restriction ...''
-	\item p.3, first paragraph of 2.1: ``omit the word finite or ordered'' $\to$ ``omit the words finite and
+	\item[$\checkmark$] p.3, first paragraph of 2.1: ``omit the word finite or ordered'' $\to$ ``omit the words finite and
 	ordered''
@@
 	reformulate this?
-	\item p.5, next sentence: ``this will allow us later'' $\to$ ``this will later allow us''
-	\item p.6, second to last sentence: ``similarity of $\alpha_3$ with'': ``similarity to'' is much more common
+	\item[$\checkmark$] p.5, next sentence: ``this will allow us later'' $\to$ ``this will later allow us''
+	\item[$\checkmark$] p.6, second to last sentence: ``similarity of $\alpha_3$ with'': ``similarity to'' is much more common
 	than ``similarity with''
-	\item p.7, next sentence: ''in this direction'' $\to$ ``in this regard''
-	\item p.7, last sentence before 3.1: ''connection of QSO with'' $\to$ ``connection of QSO to''
-	\item p.7, middle: ``Another difference of WL with our approach'' $\to$ ``Another difference between
+	\item[$\checkmark$] p.7, next sentence: ''in this direction'' $\to$ ``in this regard''
+	\item[$\checkmark$] p.7, last sentence before 3.1: ''connection of QSO with'' $\to$ ``connection of QSO to''
+	\item[$\checkmark$] p.7, middle: ``Another difference of WL with our approach'' $\to$ ``Another difference between
 	WL and our approach''
 	\item p.9, first sentence: ``then condition (2) holds'': omit the word then or reformulate the sentence
-	\item p.9, first paragraph: ``the constant function $s$ can be trivially simulated'' $\to$ ``any constant
+	\item[$\checkmark$] p.9, first paragraph: ``the constant function $s$ can be trivially simulated'' $\to$ ``any constant
 	function $s$ can be trivially simulated''
-	\item p.9, proof of Prop. 4.3: ``To prove the condition (2)'' $\to$ ``To prove condition (2)''
+	\item[$\checkmark$] p.9, proof of Prop. 4.3: ``To prove the condition (2)'' $\to$ ``To prove condition (2)''
 	\item throughout the paper: ``similar than'' $\to$ ``similar to'' / ``similar than for X, we do Y'' $\to$
@@
 	analogous to (the argument in Thm. 4.4)
-	\item p.12, after Corollary 4.8: ``shows how robust and versatile is QSO'' $\to$ ``shows how robust and
+	\item[$\checkmark$] p.12, after Corollary 4.8: ``shows how robust and versatile is QSO'' $\to$ ``shows how robust and
 	versatile QSO is''
@@
 	what ``beyond $\nat$'' refers to
-	\item p.12, ``given CNF propositional formula, that can be made true'': no comma
+	\item[$\checkmark$] p.12, ``given CNF propositional formula, that can be made true'': no comma
 	\item p.12: ``in [43] they defined'' $\to$ something like ``$<$authors$>$ [43] defined''
-	\item p.12: ``we extend as follows QSO'' $\to$ ``we extend QSO by max and min quantifiers as follows''
-	\item p.12: ``MinQSO is defined analogously changing max with min'' $\to$ ``replacing max by min''
-	\item p.12: ``the same holds with MinQSO(FO)'' $\to$ ``the same holds for ...''
-	\item p.13: ``in contrast with'' $\to$ ``in contrast to''
-	\item p.13: ``follow the same approach but considering'' $\to$ ``but consider''
-	\item p.13: ``in the search of efficient approximation algorithms'' $\to$ ``in the search for''
-	\item p.14: ``the search of robust classes'' $\to$ ''the search for robust classes''
+	\item[$\checkmark$] p.12: ``we extend as follows QSO'' $\to$ ``we extend QSO by max and min quantifiers as follows''
+	\item[$\checkmark$] p.12: ``MinQSO is defined analogously changing max with min'' $\to$ ``replacing max by min''
+	\item[$\checkmark$] p.12: ``the same holds with MinQSO(FO)'' $\to$ ``the same holds for ...''
+	\item[$\checkmark$] p.13: ``in contrast with'' $\to$ ``in contrast to''
+	\item[$\checkmark$] p.13: ``follow the same approach but considering'' $\to$ ``but consider''
+	\item[$\checkmark$] p.13: ``in the search of efficient approximation algorithms'' $\to$ ``in the search for''
+	\item[$\checkmark$] p.14: ``the search of robust classes'' $\to$ ''the search for robust classes''
 	\item p.14: ''the first desirable condition has to do with the closure properties satisfied by the class'':
 	Reformulate this sentence
-	\item p.14: ``As in the cases of P and NP that are closed'' $\to$ ``Analogously to the cases of P and
+	\item[$\checkmark$] p.14: ``As in the cases of P and NP that are closed'' $\to$ ``Analogously to the cases of P and
 	NP, which are closed under intersection and union, ...''
-	\item p.16: ``such that for both, $A\mod\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$'' $\to$ ''such that for $i \in \{1,2\}$: ... ''
-	\item p.16: ``Similarly to Theorem 5.1, we show ..'' $\to$ ``Similarly as in the proof of Theorem 5.1, we
+	\item[$\checkmark$] p.16: ``such that for both, $A\mod\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$'' $\to$ ''such that for $i \in \{1,2\}$: ... ''
+	\item[$\checkmark$] p.16: ``Similarly to Theorem 5.1, we show ..'' $\to$ ``Similarly as in the proof of Theorem 5.1, we
 	show''
@@
 	closure properties'': Should be reformulated
-	\item p.19: ``Let ... be this formula, where $\varphi$ is in first-order and quantifier-free'' $\to$ ``... where $\varphi$ is
+	\item[$\checkmark$] p.19: ``Let ... be this formula, where $\varphi$ is in first-order and quantifier-free'' $\to$ ``... where $\varphi$ is
 	first-order and quantifier-free''
 	\item several occurrences: ``This is, ...'' $\to$ ``That is, ...'', for example two times on p.20
-	\item p.20: ``we focus only in proving'' $\to$ ``we focus only on proving''
-	\item p.20: ``We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of FO.''
+	\item[$\checkmark$] p.20: ``we focus only in proving'' $\to$ ``we focus only on proving''
+	\item[$\checkmark$] p.20: ``We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of FO.''
 	$\to$ ``We prove this for the more general case of $\eqso(\LL)$ with $\LL$ being a fragment of SO.''
@@
 
-	\item p.20: ``we will show a formula'' $\to$ ``give'' or ``construct'' instead of ``show''
+	\item[$\checkmark$] p.20: ``we will show a formula'' $\to$ ``give'' or ``construct'' instead of ``show''
 	\item p.20/21: ``$\shp$ is believed not to be closed under subtraction by one by some complexitytheoretic
 	assumption'' $\to$ ``... due to some complexity-theoretic assumption'' (or similar)
-	\item p.21, next sentence: ``for a possible extension'' $\to$ ``as a possible extension''
+	\item[$\checkmark$] p.21, next sentence: ``for a possible extension'' $\to$ ``as a possible extension''
 	\item p.21: ``the desired extension has to be achieved by ... . More precisely, we define...'': The second
@@
 	actual definiton (which is chosen in accordance with the sentence before)
-	\item p.21: ``a parsimonious reduction from a function'' $\to$ ``a parsimonious reduction from any
+	\item[$\checkmark$] p.21: ``a parsimonious reduction from a function'' $\to$ ``a parsimonious reduction from any
 	function''
-	\item p.21: ``we define a function that converts a formula $\alpha$ in...'' $\to$ ``that converts a given formula
+	\item[$\checkmark$] p.21: ``we define a function that converts a formula $\alpha$ in...'' $\to$ ``that converts a given formula
 	$\alpha$...''
-	\item p.22: ``we iterate for every tuple'' $\to$ ``we iterate over all tuples''
-	\item p.22: ``Then we will show how to define a formula $\varphi'$
+	\item[$\checkmark$] p.22: ``we iterate for every tuple'' $\to$ ``we iterate over all tuples''
+	\item[$\checkmark$] p.22: ``Then we will show how to define a formula $\varphi'$
 	...'' $\to$ ``We will show how to define ...''
-	\item p.22: ''no variable from $\bar{x}$ are mentioned'' $\to$ ``... is mentioned''
+	\item[$\checkmark$] p.22: ''no variable from $\bar{x}$ are mentioned'' $\to$ ``... is mentioned''
 	\item p.24: ``if, and only if, if X, then Y'': This construction makes it unneccesarily hard to parse
 	the sentence
-	\item p.24: ``Then, the conclusion of the implication...'' $\to$ ``The conclusion of the implication...''
-	\item p.25: ``For the sake of simplification...'' $\to$ ``For the sake of simplicity...''
-	\item p.26: ``In the previous section, we show ...'' $\to$ ``In the previous section we showed ...''
-	\item p.27: ``we postpone the proof the previous proposition after ...'' $\to$ ``we postpone the proof of
+	\item[$\checkmark$] p.24: ``Then, the conclusion of the implication...'' $\to$ ``The conclusion of the implication...''
+	\item[$\checkmark$] p.25: ``For the sake of simplification...'' $\to$ ``For the sake of simplicity...''
+	\item[$\checkmark$] p.26: ``In the previous section, we show ...'' $\to$ ``In the previous section we showed ...''
+	\item[$\checkmark$] p.27: ``we postpone the proof the previous proposition after ...'' $\to$ ``we postpone the proof of
 	the previous proposition to after ...''
 	\item p.29: ``replace by $\perp$ and $\top$ \underline{wherever it corresponds}'': Should be reformulated
-	\item p.30: ``consist in computing $g_{\alpha}(\enc(\A))$ for each input ...'' $\to$ ``... on input ...''
-	\item p.30, same sentence: ``and then simulate'' $\to$ ``and then simulating''
-	\item p.33: ``In contrast with LFP'' $\to$ ``In contrast to LFP''
-	\item p.33: ``These operations are represented by the set $\{0,1,\emptyset\}$, respectively.'' $\to$ ``... by 0, 1 and
+	\item[$\checkmark$] p.30: ``consist in computing $g_{\alpha}(\enc(\A))$ for each input ...'' $\to$ ``... on input ...''
+	\item[$\checkmark$] p.30, same sentence: ``and then simulate'' $\to$ ``and then simulating''
+	\item[$\checkmark$] p.33: ``In contrast with LFP'' $\to$ ``In contrast to LFP''
+	\item[$\checkmark$] p.33: ``These operations are represented by the set $\{0,1,\emptyset\}$, respectively.'' $\to$ ``... by 0, 1 and
 	$\emptyset$, respectively''
-	\item p.34: ``Notice that the values of each function $f$ in $\alpha$'' $\to$ ``Notice that the values of each
+	\item[$\checkmark$] p.34: ``Notice that the values of each function $f$ in $\alpha$'' $\to$ ``Notice that the values of each
 	function in $\alpha$''
-	\item p.34: ``that is why we talk about ...'' $\to$ ``which is why we talk about''
-	\item p.34: ``Besides, notice that notation ...'' $\to$ ``Besides, notice that the notation ...''
+	\item[$\checkmark$] p.34: ``that is why we talk about ...'' $\to$ ``which is why we talk about''
+	\item[$\checkmark$] p.34: ``Besides, notice that notation ...'' $\to$ ``Besides, notice that the notation ...''
 	\item p.34: ``Function $h$ is used to indicate whether the head of the working tape is in some position
@@
 	at a certain point of time'' (or similar)
-	\item p.36: ``Given a relation signature $\R$'' $\to$ ``... relational signature ...''
-	\item p.37: ``As it was mentioned before'' $\to$ ``As mentioned before''
-	\item p.37: ``Next we construct a logarithmic-space ....'' $\to$ ``We construct a logarithmic-space ...''
+	\item[$\checkmark$] p.36: ``Given a relation signature $\R$'' $\to$ ``... relational signature ...''
+	\item[$\checkmark$] p.37: ``As it was mentioned before'' $\to$ ``As mentioned before''
+	\item[$\checkmark$] p.37: ``Next we construct a logarithmic-space ....'' $\to$ ``We construct a logarithmic-space ...''
 	6
@@
 	\end{itemize}
-	\item p4, beginning of 2.2: ``some of \underline{the} are recalled here''
-	\item p6, example 3.2, first paragraph: ``otherwise $M(i.j)$'', dot used instead of comma
-	\item throughout the paper: a instead of an (espacially: a FO formula/assignment, a SO formula/assignment).
+	\item[$\checkmark$] p4, beginning of 2.2: ``some of \underline{the} are recalled here''
+	\item[$\checkmark$] p6, example 3.2, first paragraph: ``otherwise $M(i.j)$'', dot used instead of comma
+	\item[$\checkmark$] throughout the paper: a instead of an (espacially: a FO formula/assignment, a SO formula/assignment).
 	Example:
@@
 	\item[-] p.11, end of proof of Thm. 4.5: ``and $\alpha$ reconstruct''
-	\item p.9, middle: ``A second attempt could be based then'' $\to$ ``A second attempt could then be
+\end{itemize}
+	\item[$\checkmark$] p.9, middle: ``A second attempt could be based then'' $\to$ ``A second attempt could then be
 	based''
-\end{itemize}
-	\item p.11, 4.1: ``output of a function \underline{in} not a natural number''
-	\item p.12: ``we have to dinstinguished''
-	\item p.14, beginning of 5.1: ``subclases''
-	\item p.21: ``it does not mentioned'' $\to$ ``it does not mention''
-	\item p.22: ``Without lost of generality'' $\to$ ``Without loss of generality''
-	\item p.24: ``The previous claim and proof motivates the following definitions.'' $\to$ ``The previous
+	\item[$\checkmark$] p.11, 4.1: ``output of a function \underline{in} not a natural number''
+	\item[$\checkmark$] p.12: ``we have to dinstinguished''
+	\item[$\checkmark$] p.14, beginning of 5.1: ``subclases''
+	\item[$\checkmark$] p.21: ``it does not mentioned'' $\to$ ``it does not mention''
+	\item[$\checkmark$] p.22: ``Without lost of generality'' $\to$ ``Without loss of generality''
+	\item[$\checkmark$] p.24: ``The previous claim and proof motivates the following definitions.'' $\to$ ``The previous
 	claim and proof motivate the following definition.''
@@
 	LFP'' $\to$ ``to define a natural generalization of LFP to functions'' (or similar)
-	\item p.30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
-	\item p.31: ``we simply generalized these conditions'' $\to$ ``we simply generalize these conditions''
-	\item p.32: ``which is used to defined'' $\to$ ``which is used to define''
-	\item p.33: ``since $\beta$ is constructed from monotones operations'' $\to$ ``... monotone ...''
-	\item p.35: ``Formulas $\alpha_h$ is defined ...'' $\to$ ``Formula $\alpha_h$ is defined ...''
+	\item[$\checkmark$] p.30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
+	\item[$\checkmark$] p.31: ``we simply generalized these conditions'' $\to$ ``we simply generalize these conditions''
+	\item[$\checkmark$] p.32: ``which is used to defined'' $\to$ ``which is used to define''
+	\item[$\checkmark$] p.33: ``since $\beta$ is constructed from monotones operations'' $\to$ ``... monotone ...''
+	\item[$\checkmark$] p.35: ``Formulas $\alpha_h$ is defined ...'' $\to$ ``Formula $\alpha_h$ is defined ...''
 	7
-	\item p.37: ``we simulate $M_\beta$ with input $v[1/n]$'' $\to$ ``... with input $v[1/x]$''
+	\item[$\checkmark$] p.37: ``we simulate $M_\beta$ with input $v[1/n]$'' $\to$ ``... with input $v[1/x]$''
 \end{itemize}

### 2018-12-10 LMCS/beyond.tex
added reviewer 2 to todo list

@@
 \begin{exa} \label{ex:count-path}
-We would like to define an $\rqfo(\fo)$-formula that, given a directed acyclic graph $G$ with $n$ nodes and a pair of nodes $b$, $c$ in $G$, counts the number of paths of length at most $n$ from $b$ to $c$ in $G$.
+We would like to define an $\rqfo(\fo)$-formula that, given a directed acyclic graph $G$ with $n$ nodes and a pair of nodes $b$, $c$ in $G$, counts the number of paths of length less than $n$ from $b$ to $c$ in $G$.
 To this end, assume that graphs are encoded using the signature $\R = \{ E(\cdot,\cdot)\}$, and then define formula $\alpha(x, y, f)$ as follows:

### 2018-12-10 LMCS/proof-capture-fpspace.tex
added reviewer 2 to todo list

@@
-For condition (1), notice that each $\pfp$ formula can be evaluated in deterministic polynomial space, that the constant function $s$ can be trivially simulated in $\fpspace$ for each $s\in\nat$, and that $\fpspace$ is closed under exponential sum and multiplication.
+For the first condition of Definition~\ref{def:cap}, notice that each $\pfp$ formula can be evaluated in deterministic polynomial space, the constant function $s$ can be trivially simulated in $\fpspace$, and $\fpspace$ is closed under exponential sum and multiplication.
 This suffices to show that the condition holds.
-
-%To show how to evaluate a $\qso(\pfp)$-formula, we construct recursively a $\shpspace$-machine $M_{\alpha}$ for each $\qso(\pfp)$ formula $\alpha$ over a signature $\R$.
-This machine runs in non-deterministic polynomial space and, on input $(\A,v,V)$, accepts in $\sem{\alpha}(\A,v,V)$ of its non-deterministic paths for each $(\A,v,V) \in \ostr[\R]^*$.
-Suppose $\A$ has domain $A$.
-If $\alpha$ is a $\pfp$-formula $\varphi$, then the machine checks if $(\A,v,V)\models\varphi$ deterministically in polynomial space~\cite{L04}, and accepts if and only if it holds true.
-If $\alpha$ is a constant $s$, it produces $s$ branches and accepts in all of them.
-If $\alpha = (\beta \add \gamma)$, then it chooses between 0 or 1, if it is 0 (1), it simulates $M_{\beta}$ ($M_{\gamma}$) on input $(\A,v,V)$.
-If $\alpha = (\beta \mult \gamma)$, it simulates $M_{\beta}$ on input $(\A,v,V)$ and on each accepting path, it continues simulating $M_{\gamma}$ on input $(\A,v,V)$.
-%If $\alpha = \sa{x}\beta$, it chooses $a\in A$ non-deterministically and simulates $M_{\beta}$ on input $(\A,v[a/x],V)$.
-If $\alpha = \pa{x}\beta$, it simulates $M_{\beta}$ on input $(\A,v[a/x],V)$ consecutively for each $a\in A$, continuing to the next $A$-value if the run of $M_{\beta}$ accepts.
-If $\alpha = \sa{X}\beta$, it chooses $B\in A^{arity(X)}$ and simulates $M_{\beta}$ on input $(\A,v,V[B/X])$.
-If $\alpha = \pa{X}\beta$, it simulates $M_{\beta}$ on input $(\A,v,V[B/X])$ consecutively for each $B\in A^{arity(X)}$, again continuing to the next $A$-value if the run of $M_{\beta}$ accepts.
-This covers all possible cases for $\alpha$, and each of these steps can be computed in polynomial space.
-Let $\alpha$ be a formula in $\qso(\pfp)$ over a signature $\R$ and let $f$ be a function over $\R$ such that $f(\enc(\A))$ is equal to the accepting paths of $M_{\alpha}$ on input $(\A,v,V)$ for some $(\A,v,V) \in \ostr[\R]^*$.
-We have that $f$ is a $\shpspace$ function over $\R$, which implies that $f$ is also a $\fpspace$ function over $\R$, by the fact that $\shpspace = \fpspace$ \cite{Ladner89}, and that $f(\enc(\A)) = \sem{\alpha}(\A)$ for every $\A\in\ostr[\R]$.
-
-For the condition (2), we provide a similar proof to theorem \ref{theo:capture-fp}.
-Let $f\in \fpspace$ defined over some $\R$.
-Let $\ell\in\nat$ be such that for each $\A\in\ostr[\R]$, $\lceil\log_2 f(\enc(\A)) \rceil \leq 2^{n^\ell}$ (i.e.
-$2^{n^\ell}$ is an upper bound for the output size), where $\A$ has a domain of size $n$.
+For the second condition, the proof is similar than in Theorem~\ref{theo:capture-fp}.
+Let $f\in \fpspace$ defined over some $\R$ and $\ell\in\nat$ such that $\log_2\left( f(\enc(\A)) \right) \leq 2^{{|\A|}^\ell}$ for every $\A\in\ostr[\R]$  (i.e.
+$2^{{|\A|}^\ell}$ is an upper bound for the output size).
 Let $X$ be a second-order variable of arity $\ell$.
-Consider a linear order over predicates of arity $\ell$ given by the formula
+Consider the linear order induced by $<$ over predicates of arity $\ell$ which can be defined by the following formula:
 $$
-\varphi_{<}(X,Y) = \exists\bar{u}\big[\neg X(\bar{u})\wedge Y(\bar{u})\wedge \forall\bar{v}\big(
+\varphi_{<}(X,Y) = \ex{\bar{u}}\big[\neg X(\bar{u})\wedge Y(\bar{u})\wedge \fa{\bar{v}}\big(
 \bar{u}<\bar{v}\to(X(\bar{u})\iff Y(\bar{v}))\big)\big].
 $$
-Namely, we use relations to encode numbers with at most $2^{n^\ell}$-bits where the empty relation represents $0$ and the total-relation represents $2^{2^{n^\ell}}-1$.
-Furthermore, each relation $X$ indexes a position in the binary output of $f(\enc(\A))$ as follows.
+Namely, we use relations to encode numbers with at most $2^{{|\A|}^\ell}$ bits where the empty relation represents $0$ and the total-relation represents $2^{2^{{|\A|}^\ell}}-1$.
+Furthermore, we can use a relation~$X$ to index a position in the binary output of $f(\enc(\A))$ as follows.
 %Consider a polynomial space machine over the $\R$ that receives as input an $\R$-structure $\A$ and a number $p$ encoded by a relation $X$.
 Then the machine accepts if, and only if, the $p$-th bit of $f(\enc(\A))$ is $1$.
-Define a language
+Define the language:
 \[
@@
 \]
-Since this language is in $\pspace$, it can be described in $\pfp$ \cite{AbiteboulV89} by a formula $\Phi(X)$ where the free variable $X$ encodes the number $p$.
+Since $L$ is in $\pspace$, it can be specified in $\pfp$ \cite{AbiteboulV89} by a formula $\Phi(X)$ where the free variable $X$ encodes relation $B$ in $L$.
 Then, similar than the previous proof we define:
 $$
-\alpha := \sa{X}(\Phi(X)\mult\varphi(X)),
+\alpha := \sa{X} \Phi(X)\mult  \pa{Y}(\varphi_{<}(Y,X)\mapsto 2).
 $$ 
-where $\varphi(X) = \pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$.
-As in theorem \ref{theo:capture-fp}, this formula takes the value $2^m$ if there exist $m$ predicates that are smaller than $X$.
+where $\pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$ takes the value $2^m$ if there exist $m$ predicates that are smaller than~$X$ and $\alpha$ reconstruct the output of $f(\enc(\A))$ by simulating $f$ with $\Phi(X)$.
 Using an analogous argument, we conclude that $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.

### 2018-12-10 LMCS/proof-sigma2hard.tex
added reviewer 2 to todo list

@@
 where each $\varphi^i_j$ is a Horn-clause, and each $\bar{X}_{i}$ is a sequence of second-order variables.
-Consider $\bar{X}$ as the union of all $\bar{X}_{i}$. We replace each of the $m$ addends in $\alpha$ for
+Consider $\bar{X}$ as the union of all $\bar{X}_{i}$. We replace each of the $m$ addends in $\alpha$ with
 $$
@@
 $$
-Note that each first-order sub-formula in $\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c})$ has no free variables and, therefore, we can evaluate each of them in polynomial time and replace by $\perp$ and $\top$ wherever it corresponds.
-In other words, in polynomial time we can replace $\varphi^i_j$ by a disjunction of $\neg X_{h}$ and at most one $X_{h}$, evaluated on constants.
+Note that each first-order sub-formula in $\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c})$ has no free variables and, therefore, we can evaluate each of them in polynomial time and replace with $\perp$ and $\top$ wherever it corresponds.
+In other words, in polynomial time we can replace $\varphi^i_j$ with a disjunction of $\neg X_{h}$ and at most one $X_{h}$, evaluated on constants.
 After simplifying, grouping and reordering the previous formula, we can obtain an equivalent formula $\alpha_{\A}'$ of the form:

### 2018-12-10 LMCS/proof-sigmafo-minusone.tex
added reviewer 2 to todo list

@@
 Without loss of generality, we assume that $\alpha$ is in $\logex{1}$-SNF, that is, $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
-Moreover, we assume that $\length{\bar{x}} > 0$  since, if this is not the case, we can replace $\sa{\bar{X}}\varphi_i$ for the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
+Moreover, we assume that $\length{\bar{x}} > 0$  since, if this is not the case, we can replace $\sa{\bar{X}}\varphi_i$ with the equivalent formula $\sa{\bar{X}} \sa{y} \varphi_i \wedge\first(y)$.
 
@@
 Intuitively, we want to exclude the assignment $(V^*, v^*)$ from the satisfying assignments of $\varphi_i(\bar{X},\bar{x})$.
-Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if, if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, then
-$V \neq V^*$ or $v \neq v^*$. 
+Towards this goal, we can define a formula $\psi_i(\bar{X},\bar{x})$ such that, if $\varphi_i(\bar{X}, \bar{x})$ is satisfiable, $(\A, V, v) \models \psi_i(\bar{X},\bar{x})$ if, and only if either $V \neq V^*$ or $v \neq v^*$.
 This property can be defined as follows:

### 2018-12-10 LMCS/proof-tqfo-shl.tex
added reviewer 2 to todo list

@@
 Suppose that the domain of $\A$ is $A = \{1,\ldots,n\}$.
-The TM $M_{\alpha}$ needs $\ell \mult\log_2(n)$ bits of memory to store the first-order variables occurring in $\alpha$, where $\ell$ is the number of variables occurring in this formula (which is the same as the number of variables in the domain of $v$).
+The TM $M_{\alpha}$ needs $\ell \mult\log_2(n)$ bits of memory to store the first-order variables that appear in $\alpha$, where $\ell$ is the number of different variables in this formula (and also the size of the domain of $v$).
 If $\alpha = \varphi$, where $\varphi$ is an $\fo$-formula, then we check if $(\A,v)\models\varphi$ in deterministic logarithmic space, and accept if and only if this condition holds.
-If $\alpha = s$, where $s$ is a fixed natural number, then we generate $s$ possible runs and accept in all of them.
+If $\alpha = s$, where $s$ is a fixed natural number, then we generate $s$ branches and accept in all of them.
 If $\alpha = (\alpha_1 + \alpha_2)$, we simulate $M_{\alpha_1}$ and $M_{\alpha_2}$ on separate branches.
 If $\alpha = (\alpha_1\mult\alpha_2)$, we simulate $M_{\alpha_1}$ and if it accepts, then instead of accepting we simulate $M_{\alpha_2}$.
-If $\alpha = \sa{x}\beta$, for each $a\in A$ we generate a different run where we simulate $M_{\beta}$ with input $v[a/x]$.
-If $\alpha = \pa{x}\beta$, we simulate $M_{\beta}$ with input $v[1/x]$, and on each accepting run, instead of accepting we replace the assignment of $x$ to 2, to simulate $M_{\beta}$ with input $v[2/x]$, and so on.
-If $\alpha = [\pth \varphi(\bar{x},\bar{y})]$, where $\varphi$ is an $\fo$-formula, then we simulate the $\shl$ procedure that counts the number of paths of a given length from a source to a target node in an input graph (where the length is at most the number of nodes in the graph).
+If $\alpha = \sa{x}\beta$, for each $a\in A$ we generate a different branch where we simulate $M_{\beta}$ with input $v[a/x]$.
+If $\alpha = \pa{x}\beta$, we simulate $M_{\beta}$ with input $v[1/x]$, and on each accepting run, instead of accepting we simulate $M_{\beta}$ with input $v[2/x]$, and so on.
+If $\alpha = [\pth \varphi(\bar{x},\bar{y})]$, where $\varphi$ is an $\fo$-formula, we simulate the $\shl$ procedure that counts the number of paths of a given length from a source to a target node in an input graph (where the length is at most the number of nodes in the graph).
 %This procedure starts by setting $\bar{a} = v(\bar{x})$.

### 2018-12-10 LMCS/todo-list.tex
added reviewer 2 to todo list

@@
 \section*{\LARGE \textbf{TODO LIST}}
+
+\bigskip
+
+{\bf Reviewer 1}
 
@@
 	\item[$\checkmark$] p.22: ''no variable from $\bar{x}$ are mentioned'' $\to$ ``... is mentioned''
-	\item p.24: ``if, and only if, if X, then Y'': This construction makes it unneccesarily hard to parse
+	\item[$\checkmark$] p.24: ``if, and only if, if X, then Y'': This construction makes it unneccesarily hard to parse
 	the sentence
@@
 	7
-	\item[$\checkmark$] p.37: ``we simulate $M_\beta$ with input $v[1/n]$'' $\to$ ``... with input $v[1/x]$''
+	\item[$\checkmark$] p.37: ``we simulate $M_\beta$ on input $v[1/n]$'' $\to$ ``... with input $v[1/x]$''
 \end{itemize}
@@
 	\item running title is ``COUNTING''
-	\item on definition of functions, \texttt{\symbol{92}colon} should be used instead of :
+	\item on definition of functions, \verb|\colon| should be used instead of :
 	\item In [1], a similar hierarchy to the one by Saluja et al. was introduced and studied. This work
@@
 \end{itemize}
+
+\bigskip
+
+{\bf Reviewer 2}
+
+\begin{itemize}
+		\setlength\itemsep{0.5em}
+\item In several proofs, you rely on disjoint union of solutions and cartesian product of solutions.
+Indeed, its gives a polynomial time counting reduction.
+Can you make connections with (+,*) circuits ?
+It would be similar to what is done in A Circuit-Based Approach to Efficient Enumeration (Antoine Amarilli, Pierre Bourhis, Louis Jachiet, Stefan Mengel) where simple circuits (dDNNF, which are basically disjoint unions and cartesian products) are used to capture the solutions of MSO queries.
+
+\item In general I would be interested to know the relationship of your couting classes with their enumeration counterpart. 
+For instance there is a paper adapting [39] to enumeration: Enumeration Complexity of Logical Query Problems with Second-order Variables (Arnaud Durand and Yann Strozecki).
+In particular they consider class with some Horn structure to capture easy enumeration problems, it may have some connection with the class $\sigma$ QSO($\Pi_1$ Horn) you introduce.
+
+\item For the most involved proofs, you could describe informally your proof strategy in a few lines to help the reader.
+You may also add some example to illustrate the different fragment of QSO (you already do it sometimes and it is very helpful).
+
+\item[$\checkmark$] page 4:
+-some of the $\to$ some of them 
+\item[$\checkmark$] page 10: 
+-$2^{|A|^l}$-bits $\to$ $2^{|A|^l}$ bits
+\item[$\checkmark$] page 21:
+-it does not mentioned $\to$ it does not mention
+\item page 22: 
+-you make a long proof about some closure under subtraction by one.
+You should put the result in a theorem envirpnment to make what you are
+proving more evident.
+\item[$\checkmark$] -replace .... for $\to$ replace ... by;  there are several instances of
+that
+\item[$\checkmark$] -$\phi$ is an FO-formula $\to$ a $\Sigma_1[FO]$ formula
+\item -you explain that you rewrite your formula to obtain conditions a,b and c, 
+you should give an idea how these conditions will be used
+\item[$\checkmark$] page 23:
+a SO, a FO  an SO, an FO
+\item page 24: 
+if, if $\phi$$\to$ $\phi$
+\item page 25: 
+-you explain that the proof relies on witnesses of logarithmic size.
+It was not clear to me while reading the proof. Can you explain more ?
+\item page 28:
+-the complete problem for $\Sigma$ QSO($\Sigma_2$ Horn) would be complete also for $\#\Sigma-2$ Horn. You should mention here that you do not need sum here
+and that the framework of [39] is enough for this result.
+\item[?] page 30:
+-not, trivial $\to$ not trivial
+\item[$\checkmark$] page 32: 
+- paths of length at most n $\to$ in an acylic graph, all paths are of size less than n
+\item page 33:
+-there is a proof symbol \verb|\qed| but it does not end a well defined proof environment
+
+\item page 38: 
+the last sentence on fixed point operator is to vague. Either expand it to 
+say what king of question on fixed point operator you want to investigate or remove it	
+\end{itemize}

### 2018-12-27 LMCS/sendrev/proof-theo-pi1-pnf.tex
carpeta con version send corregida

--- 2018-12-09 LMCS/proof-theo-pi1-pnf.tex
@@
-% !TeX root = main.tex
+%!TEX root = main.tex
 
@@
 Therefore, suppose that $\length{\bar{X}} > 0$ or $\length{\bar{x}} > 0$, and consider some ordered structure $\A$. 
-Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for $i\in\{1,2\}$: $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
+Since $\sem{\alpha}(\A) = 2$, there exist at least two assignments $(\bar{B}_1,\bar{b}_1,\bar{a}_1)$, $(\bar{B}_2,\bar{b}_2,\bar{a}_2)$ to $(\bar{X},\bar{x},\bar{y})$ such that for both, $\A\models\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$.
 Now consider the ordered structure $\A'$ that is obtained by taking the disjoint union of $\A$ twice. 
@@
 For the last part of the proof, we show that if $\LL$ contains $\logu{1}$ and is closed under conjunction and disjunction, then for every formula $\alpha$ in $\eqso(\LL)$ there exists an equivalent formula in $\LL$-PNF.
-Similarly as in the proof of Theorem~\ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such a formula. 
+Similarly to Theorem~\ref{theo-pnf-snf}, we show a recursive function $\tau$ that produces such a formula. 
 Assume that $\alpha = \sum_{i = 1}^n \alpha_i$ is in $\LL$-SNF where each $\alpha_i$ is in $\LL$-PNF. 

### 2018-12-27 LMCS/sendrev/proof-theo-pnf-snf.tex
carpeta con version send corregida

--- 2017-07-27 LMCS/proof-theo-pnf-snf.tex
@@
 To find an equivalent formula in $\LL$-SNF for every $\alpha \in \eqso(\LL)$, we give a recursive function $\tau$ such that $\tau(\alpha)$ is in $\LL$-SNF and $\tau(\alpha) \equiv \alpha$. 
-Specifically, if $\alpha = \varphi$, define $\tau(\alpha) = \alpha$ and if $\alpha = s$, define $\tau(\alpha) = (\top \add \overset{\text{$s$-times}}{\ldots} \add \top)$.
-If $\alpha = (\alpha_1 + \alpha_2)$, define $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$ and if $\alpha = \sa{x}\beta$, define $\tau(\alpha) = \sum_{i = 1}^{k}\sa{x}\beta_i$ where $\tau(\beta) = \sum_{i = 1}^{k}\beta_i$ for some $k$ and each $\beta_i$ is in $\LL$-PNF.
-Finally, if $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case.
+Specifically, if $\alpha = \varphi$, define $\tau(\alpha) = \alpha$; 
+if $\alpha = s$, define $\tau(\alpha) = (\top \add \overset{\text{$s $ times}}{\ldots} \add \top)$;
+if $\alpha = (\alpha_1 + \alpha_2)$, define $\tau(\alpha) = (\tau(\alpha_1) + \tau(\alpha_2))$;
+if $\alpha = \sa{x}\beta$, assume $\tau(\beta) = \sum_{i = 1}^{k}\beta_i$ such that each $\beta_i$ is in $\LL$-PNF, and define $\tau(\alpha) = \sum_{i = 1}^{k}\sa{x}\beta_i$;
+and if $\alpha = \sa{X}\beta$, then we proceed analogously as in the previous case.
 This covers all possible cases for $\alpha$ and we conclude the proof by taking $\tau(\alpha)$ as the desired rewrite of $\alpha$.

### 2018-12-27 LMCS/sendrev/proof-tqfo-shl.tex
carpeta con version send corregida

--- 2018-12-09 LMCS/proof-tqfo-shl.tex
@@
 Let $\R$ be a relational signature and $\alpha$ a formula over $\R$ in $\tqfo(\fo)$.
-We construct a logarithmic-space nondeterministic Turing Machine $M_{\alpha}$ that on input $(\enc(\A),v)$, where $\A$ is an $\R$-structure and $v$ is a first-order assignment for $\A$, has $\sem{\alpha}(\A,v)$ accepting runs (so that we can conclude that the function defined by $\alpha$ is in $\shl$).
+Next we construct a logarithmic-space nondeterministic Turing Machine $M_{\alpha}$ that on input $(\enc(\A),v)$, where $\A$ is an $\R$-structure and $v$ is a first-order assignment for $\A$, has $\sem{\alpha}(\A,v)$ accepting runs (so that we can conclude that the function defined by $\alpha$ is in $\shl$).
 Suppose that the domain of $\A$ is $A = \{1,\ldots,n\}$.
@@
 If $\alpha = \sa{x}\beta$, for each $a\in A$ we generate a different run where we simulate $M_{\beta}$ with input $v[a/x]$.
-If $\alpha = \pa{x}\beta$, we simulate $M_{\beta}$ with input $v[1/x]$, and on each accepting run, instead of accepting we replace the assignment of $x$ to 2, to simulate $M_{\beta}$ with input $v[2/x]$, and so on.
+If $\alpha = \pa{x}\beta$, we simulate $M_{\beta}$ with input $v[1/n]$, and on each accepting run, instead of accepting we replace the assignment of $x$ to 2, to simulate $M_{\beta}$ with input $v[2/x]$, and so on.
 If $\alpha = [\pth \varphi(\bar{x},\bar{y})]$, where $\varphi$ is an $\fo$-formula, then we simulate the $\shl$ procedure that counts the number of paths of a given length from a source to a target node in an input graph (where the length is at most the number of nodes in the graph).

### 2018-12-27 LMCS/sendrev/proof-prop-e1-nc.tex
correcciones en la carpeta sendrev

--- 2018-12-09 LMCS/proof-prop-e1-nc.tex
@@
 Consider a structure $\mathfrak{1}$ with only one element. 
-We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$ $\sem{\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which implies that $\length{\bar{x}} > 0$.
-But since the structure has only one element, there is only one possible assignment to $\bar{x}$.
+We have that $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) = 2$, but since the structure has only one element, there is only one possible assignment to $\bar{x}$. 
 And so, $\sem{\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y})}(\mathfrak{1}) \leq 1$, which leads to a contradiction.

### 2018-12-27 LMCS/sendrev/proof-prop-rest.tex
correcciones en la carpeta sendrev

--- 2017-10-04 LMCS/proof-prop-rest.tex
@@
 We use the same construction used in \cite{SalujaST95}, and we obtain that the formula $\ex{\bar{y}}\varphi_i(\bar{X},\bar{x},\bar{y})$ is equivalent to $\sa{\bar{y}}\,[\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \fa{\bar{y}'}(\varphi_i(\bar{X},\bar{x},\bar{y}')\to\bar{y}\leq\bar{y}')]$ for every assignment to $(\bar{X},\bar{x})$.
-Doing this replacement for each $\alpha_i$ renders an equivalent formula in $\eqso(\logu{1})$.
+We do this replacement for each $\alpha_i$, and we obtain an equivalent formula to $\alpha$ in $\eqso(\logu{1})$.
 

### 2019-01-06 LMCS/sendrev/main.tex
carpeta con version send corregida
correcciones en la carpeta sendrev
mas revisiones

--- 2018-05-07 LMCS/main.tex
@@
 %% mandatory lists of keywords 
-%\keywords{MANDATORY list of keywords}
+\keywords{Function complexity clases, descriptive complexity}
+\ACMCCS{
+[{\bf Theory of computation}]:~Logic---Finite Model Theory
+[{\bf Theory of computation}]:~Computational complexity and cryptography---Complexity classes}
 
@@
 \usepackage{stmaryrd}
+\usepackage[textwidth=2cm,textsize=small]{todonotes}
+\usepackage{enumitem}
 \input{macros}
@@
 
-\title[Counting]{Descriptive Complexity for Counting Complexity Classes}
+\title[Descriptive Complexity for Counting Complexity Classes]{Descriptive Complexity for Counting Complexity Classes}
 %\titlecomment{{\lsuper*}OPTIONAL comment concerning the title, \eg, 
@@
 
+\newpage
+\input{todo-list}
+
 \end{document}

### 2019-01-06 LMCS/sendrev/proof-qe0-fp-qe1-totp-fptras.tex
correcciones en la carpeta sendrev
mas revisiones

--- 2018-12-09 LMCS/proof-qe0-fp-qe1-totp-fptras.tex
@@
+% !TeX root = main.tex
+
 The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$. 
-This is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every $\R$-structure $\A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\size{\A})$.
+That is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every $\R$-structure $\A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\size{\A})$.
 We use this fact in the following arguments.
@@
 To show that $\eqso(\loge{1})$ is contained  in \textsc{TotP}, let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF. 
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF. 
+That is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF. 
 Consider the following nondeterministic procedure that on input $\enc(\A)$ generates $\sem{\alpha}(\A)$ branches. 
@@
 To show that every function in $\eqso(\loge{1})$ has an FPRAS,  let $\alpha$ be a $\eqso(\loge{1})$ formula and assume that it is in $\loge{1}$-SNF. 
-This is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF. 
+That is, $\alpha = \sum_{i = 1}^n\alpha_i$ where each $\alpha_i$ is in $\loge{1}$-PNF. 
 Note that each $\alpha_i$ that is equal to some $\loge{1}$ formula $\varphi$ has an FPRAS given by the procedure that simply checks if $\A\models\varphi$ and returns 1 if it does and 0 otherwise. 
@@
 Since $\eqso(\loge{1})$ is closed under sum by definition, we focus only on proving that it is closed under multiplication. 
-We prove this for the more general case of $\eqso(\LL)$ with $\LL$ being a fragment of SO.
+We prove this for the more general case of $\eqso(\LL)$ with $\LL$ being a fragment of $\so$.
 

### 2019-01-06 LMCS/sendrev/proof-rqfo-fo-cap.tex
correcciones en la carpeta sendrev
mas revisiones

--- 2018-12-09 LMCS/proof-rqfo-fo-cap.tex
@@
 
-In the formula $\alpha$, function $T_0$ is used to indicate whether the content of a cell of the working tape is 0 at some point of time, that is, $T_0(\bar{t},\bar{x}) > 0$ if the cell at position $\bar{x}$ of the working tape contains the symbol 0 at step $\bar{t}$, and $T_0(\bar{t},\bar{x}) = 0$ otherwise.
+In the formula $\alpha$, function $T_0$ is used to indicate whether the content of a cell of the working tape is 0 at a certain point of time, that is, $T_0(\bar{t},\bar{x}) > 0$ if the cell at position $\bar{x}$ of the working tape contains the symbol 0 at step $\bar{t}$, and $T_0(\bar{t},\bar{x}) = 0$ otherwise.
 Functions $T_1$, $T_{\ttB}$ and $T_{\vdash}$ are defined analogously.
-Function $h$ is used to indicate whether the head of the working tape is in some position at some point of time, that is, $h(\bar{t},\bar{x}) > 0$ if the head of the working tape is at position $\bar{x}$ at step $\bar{t}$, and $h(\bar{t},\bar{x}) = 0$ otherwise.
-Function $\hat{h}$ is used to indicate whether the head of the working tape is {\bf not} in some position at some point of time, that is, $\hat{h}(\bar{t},\bar{x}) > 0$ if the head of the working tape is {\bf not} at position $\bar{x}$ at step $\bar{t}$, and $\hat{h}(\bar{t},\bar{x}) = 0$ otherwise.
-For each $i \in \{0, \ldots, \ell\}$, function
-$s_{q_i}$ is used to indicate whether the TM $M$ is in state $q_i$ at some point of time, that is, $s_{q_i}(\bar{t}) > 0$ if the TM $M$ is in state $q_i$ at step $\bar{t}$, and $s_{q_i}(\bar{t}) = 0$ otherwise.
+Function $h$ is used to indicate whether the head of the working tape is in a specific position at a certain point of time, that is, $h(\bar{t},\bar{x}) > 0$ if the head of the working tape is at position $\bar{x}$ at step $\bar{t}$, and $h(\bar{t},\bar{x}) = 0$ otherwise.
+Function $\hat{h}$ is used to indicate whether the head of the working tape is {\bf not} in a position at a particular point of time, that is, $\hat{h}(\bar{t},\bar{x}) > 0$ if the head of the working tape is {\bf not} at position $\bar{x}$ at step $\bar{t}$, and $\hat{h}(\bar{t},\bar{x}) = 0$ otherwise.
+For each $i \in \{0, \ldots, \ell\}$, function $s_{q_i}$ is used to indicate whether the TM $M$ is in state $q_i$ at a certain point of time, that is, $s_{q_i}(\bar{t}) > 0$ if the TM $M$ is in state $q_i$ at step $\bar{t}$, and $s_{q_i}(\bar{t}) = 0$ otherwise.
 Function $\out$ stores the output of the TM $M$; in particular, $\out(\bar t)$ is the value returned by $M$ when $\bar t$ is the last step.

### 2019-01-06 LMCS/sendrev/syntactic-clo.tex
correcciones en la carpeta sendrev
mas revisiones

--- 2018-12-09 LMCS/syntactic-clo.tex
@@
 Thus, we need to find an extension of $\QE{1}$ that keeps all the previous properties and is closed under subtraction by one.
-It is important to notice that $\shp$ is believed not to be closed under subtraction by one by some complexity-theoretical assumption\footnote{A decision problem $L$ is in the randomized complexity class $\cspp$ if there exists a polynomial-time NTM $M$ such that for every $x \in L$ it holds that $\tma_M(x) - \tmr_M(x) = 2$, and for every $x \not\in L$ it holds that $\tma_M(x) = \tmr_M(x)$ \cite{OH93,FFK94}.
+It is important to notice that $\shp$ is believed not to be closed under subtraction by one due to some complexity-theoretical assumption\footnote{A decision problem $L$ is in the randomized complexity class $\cspp$ if there exists a polynomial-time NTM $M$ such that for every $x \in L$ it holds that $\tma_M(x) - \tmr_M(x) = 2$, and for every $x \not\in L$ it holds that $\tma_M(x) = \tmr_M(x)$ \cite{OH93,FFK94}.
 It is believed that $\np \not\subseteq \cspp$.

### 2019-01-06 LMCS/sendrev/todo-list.tex
mas revisiones

--- 2018-12-10 LMCS/todo-list.tex
@@
 	first-order and quantifier-free''
-	\item several occurrences: ``This is, ...'' $\to$ ``That is, ...'', for example two times on p.20
+	\item[$\checkmark$] several occurrences: ``This is, ...'' $\to$ ``That is, ...'', for example two times on p.20
 	\item[$\checkmark$] p.20: ``we focus only in proving'' $\to$ ``we focus only on proving''
@@
 	\item[$\checkmark$] p.20: ``we will show a formula'' $\to$ ``give'' or ``construct'' instead of ``show''
-	\item p.20/21: ``$\shp$ is believed not to be closed under subtraction by one by some complexitytheoretic
+	\item[$\checkmark$] p.20/21: ``$\shp$ is believed not to be closed under subtraction by one by some complexitytheoretic
 	assumption'' $\to$ ``... due to some complexity-theoretic assumption'' (or similar)
@@
 	\item[$\checkmark$] p.34: ``Besides, notice that notation ...'' $\to$ ``Besides, notice that the notation ...''
-	\item p.34: ``Function $h$ is used to indicate whether the head of the working tape is in some position
+	\item[$\checkmark$] p.34: ``Function $h$ is used to indicate whether the head of the working tape is in some position
 	at some point of time'' $\to$ ``... whether the head of the working tape is in a specific position
@@
 	6
-	\item p.38: ``In this sense, there are several directions for future research'' $\to$ ``Consequently, there
+	\item[?] p.38: ``In this sense, there are several directions for future research'' $\to$ ``Consequently, there
 	are ...'' (or similar)
-	\item p.38: ``In the same direction, ...'' $\to$ ``Similarly, ...'' (or similar)
-	\item p.38: ``so to have'' $\to$ ``in order to have'' (or similar)
+	\item[$\checkmark$] p.38: ``In the same direction, ...'' $\to$ ``Similarly, ...'' (or similar)
+	\item[$\checkmark$] p.38: ``so to have'' $\to$ ``in order to have'' (or similar)
 \end{itemize}
@@
 	\setlength\itemsep{0.5em}
-	\item running title is ``COUNTING''
+	\item[$\checkmark$] running title is ``COUNTING''
 	\item on definition of functions, \verb|\colon| should be used instead of :

### 2019-01-19 LMCS/sendrev/response-letter.tex
carta de respuesta + correcciones varias

\documentclass[a4paper]{article}

\usepackage{fullpage}
\usepackage{cite}
\usepackage{amsmath,amsfonts,amssymb,mathabx}
\usepackage{stmaryrd}
\usepackage[textwidth=2cm,textsize=small]{todonotes}
\usepackage{enumitem}
\input{macrosplain}

\def\eg{{\em e.g.}}
\def\cf{{\em cf.}}

\newcommand{\red}[1]{{\bf\textcolor{red}{#1}}}

\begin{document}

\title{Response letter to the reviews of the paper: Descriptive Complexity for Counting Complexity Classes}

\maketitle

We would like to thank the anonymous reviewers for their useful comments.
Below are our answers to comments or suggestions made about our paper.
We highlight the reviewer comment in red and then write our answer.

\section*{Reviewer 1}

\subsection*{Technical}

\begin{itemize}
	\setlength\itemsep{0.5em}
	\item \red{p.2, first paragraph: ``However, other function classes have emerged from the need to understand	the complexity of some computation problems for which little can be said if their decision counterparts are considered.'' Why can only little be said about the decision counterparts of the examples mentionend[sic] afterwards?}

	This was meant to say that the complexity of the decision counterparts of the problems mentioned afterwards did not provide a full understanding of the complexity of the problem in its computation version. This sentence has been reformulated to explain that idea.

%	\item p.7, first sentence: ``fragments or extensions [...] obtained by restricting'': extensions are not obtained by restricting something
%	\item p.10, before Thm. 4.5: ``by considering second-order product'': We need both second-order product and sum (at least both are used in Thm. 4.5)
	\item \red{p.10, last line: ``$B$-th bit'': mention in what way $B$ is interpreted as a number}

	This proof was rewitten to explain this notion better.

%	\item[$\checkmark$] p.12: ``To formalize this set of optimization problems'': class instead of set
%	\item p.14: ``we have to concentrate on the class of counting problems with easy decision versions'': here, easy decision version means ``in \bpp''. Before, it was defined as ``in P''. Should be consistent.
%	\item p.14: ``as expected, $\Sigma \bar{X}$ is the respective nesting of $\Sigma X.$’s'': why is this explained for second order 	quantifiers but not for first-order quantifiers?
%	\item p.14: $\sum_{i=1}^n\alpha_i$: explain that this is notation for $\alpha_1+\cdots+\alpha_n$
%	\item p.17: ``Since $\alpha$ is not null'' $\to$ ``Since $\sem{\alpha}$ is not the identically zero function'' (or similar)
	\item \red{p.17/18: In the proof of Prop 5.3, the vocabulary only contains the numerical predicate $<$. When encoding structures as binary words, numerical predicates are implicit (and not encoded), though. What is the intended meaning here?}

	For ordered structures the vocabulary should always include $<$. The encoding function in the paper was revised to make this more clear.

%	\item p.18, after proof of Claim 5.4: ``which implies that 2 $\in \eqso(\Sigma_0)$ and $2 \not\in \E{1}$'': $2 \in \eqso(\Sigma_0)$ does not follow from the argument before as stated in this sentence, but is trivial
%	\item[$\checkmark$] p.18: ``Also, suppose that it is in $\L$-PNF'': $\L$ should be $\LL$ , PNF should be SNF
%	\item[$\checkmark$] p.19, end of proof of Prop. 5.5: ``which violates the inequality'' $\to$ ``which violates the inequality 	for large structures''
%	\item[$\checkmark$] p.20: ``we will show a formula in the grammar'' $\to$ ``... in the logic''
%	\item p.20: ``Since $\Sigma_1$ is closed under conjunction, then Lemma 5.7 holds for ...'' $\to$ ``Since $\Sigma_1$ is closed under cnojunction, Lemma 5.7 can be applied to ...'' (or similar)
%	\item[$\checkmark$] p.22: ``and leaving each $R_{\psi}^{\A}$ empty'' $\to$ $R_{\psi}^{\A'}$
%	\item p.22: ``To show that $\alpha$ is in \totp'' $\to$ ``To show that the function defined by $\alpha$ ...'' (or similar)
%	\item[$\checkmark$] p.22: ``Let $\alpha$ = ... where $\varphi$ is an FO-formula'' $\to$ ``... where $\varphi$ is an $\Sigma_1$[FO]-formula''
	\item p.23, formulae: Maybe put (large) parantheses around the relativization of quantifiers, for	example the part of formula $\varphi_i$ after the quantifiers
	\item[$\checkmark$] p.24, formula 5.6: subscript of large conjunction: $\varphi_i^{+}(\bar{X},\bar{v})$ $\to$ $\varphi_i^{+}(\bar{X},\bar{y})$
	\item[$\checkmark$] p.25: ``We construct a function $\lambda$ that receives a quantitative formula $\beta$ ...'' $\to$ ``... that receives
	a formula $\beta \in \eqso(\logex{1})$ ...''
	\item p.25, last sentence: ``which is based on considering some witnesses of logarithmic size'' could
	be explained a little more / made more clear
	\item p.26: ``... auxiliary predicate $A$ defined as'' $\to$ ``... auxiliary predicate $A$ which will be defined
	as ... in the formula'' (or similar)
	\item p.28: ``since universal formulas are monotone over induced substructures'': ``monotone over
	induced substructures'' should be reformulated
	\item p.28: ``Next we prove that ... is hard for ... over a signature $\R$ ...'' sounds like it only works
	for one specific vocabulary. Maybe simply ``over any signature $\R$''
	\item p.28, formula $\alpha_{\A}$: second big conjunction should have $\bar{z}'$ in the subscript instead of $\bar{z}$ (because
	the quantifier $\forall u$ is handled by this as well if I understand correctly)
	\item p.29: ``In other words, in poly time we can replace $\varphi_j^i$ by a disjunction of ....'': Some of these
	formulae might evaluate to $\perp$ or $\top$ altogether and can be removed (potentially together with
	the surrounding conjunction)
	\item p.29: ``Intuitively $t_1,\ldots,t_{m'}$ are used to count from 1 to $m'$ each subformula ...'': Should be
	reformulated, maybe rather write that they are used to have disjoint sets of propositional
	assignments for the different disjuncts of the outermost disjunction (corresponding to the
	summands in the original formula)
	\item p.30, proof of Prop. 5.12: For $\dhsat \in \totp$, [35] is cited. It’s stated that a \totp-procedure
	for DNF-SAT is given there and it can be adapted to \dhsat. While the
	cited paper contains the result for DNF-SAT, the procedure is not explicitly given as far as
	I can see, so I can't check whether it is adaptable. On the other hand, membership of any
	function counting the number of satisfying assignments for a class of propositional formulae
	with poly-time decision version in \totp should hold by a similar argument as membership in
	DelayP. Maybe change the citation or give a short proof sketch.
	\item[$\checkmark$] p.31, definition of $\leq$ on $\F$: ``$f(i) \leq g(i)$ for every $i \in \{1,\ldots,\ell\}$'': Inputs for $f$ and $g$ should be
	elements of $A^{\ell}$, not $\{1,\ldots,\ell\}$
	\item p.32/33, example 6.2: I don't see why the formula given for dag's does not work for arbitrary
	graphs: In arbitrary graphs, the fixpoint is still reached after at most $n-1$ steps, since the
	maximal length of any shortest path in any graph is $n-1$. On the other hand, assuming that
	I am mistaken and the support does not become stable within $n-1$ steps: Using the formula
	given for arbitrary graphs, the support of the constructed function restricted to those tuples
	that have first in their last component is the same as the support of the constructed function
	using the formula for dags in each step of the recursion. Hence, the formula should not be
	able to fix such a problem.
	\item p.33: ``executes exactly $n^k$ steps for a fixed $k \geq 1$'' → ``executes exactly $n^k$
	steps on large
	inputs for a fixed $k \geq 1$'' (at least it does not work for $n = 1$) (small inputs can be handled
	seperately)
	\item p.35 and p.36: While $\sum$ was used for finite sums in \qso-formulae throughout the paper, now
	a large +-sign is used. This might be confusing
	\item p.36: $\alpha_{q_i}$ used instead of $\alpha_{s_{q_{i}}}$
	\item p.36: ``Clearly, at each iteration of the fixed point operator, the tuple $\bar{t}$ represents the step the
	machine is currently in.'': This is quite vague. Maybe something like ``Clearly, tuple $\bar{t}$ encodes
	the number of steps the machine has done and in each iteration of the fixed point operator,
	one timestep of the machine is executed''
	\item p.36: ``at the $\bar{a}$-th iteration of the fixed point operator, it holds that $f(\bar{a}, \bar{b}) \leq 1$ and $f(\bar{a}', \bar{b}) = 0$ for every $\bar{b} \in A^k$. 
	In the same way, it can be seen that for each function $g \in \ldots$ it holds that $g(\bar{a}) \leq 1$ and $g(\bar{a}')= 0$. 
	I think that more properties of these functions should be mentioned
	in order to argue that the defined function is the function computed by the machine:
	\begin{itemize}
		\item[-] For each $\bar{b}$, only one of the functions $T_0$,$T_1$,$T_B$,$T_{\vdash}$ outputs 1 on input ($\bar{a}, \bar{b})$
		\item[-] For each $\bar{b}$, only one of the functions $h,\hat{h}$ outputs 1 on input $(\bar{a}, \bar{b})$
		\item[-] There is exactly one $\bar{b}$ with $h(\bar{a}, \bar{b}) = 1$
		\item[–] There is exatcly one $i$ such that $s_{q_i}(\bar{a}) = 1$
	\end{itemize}
	(It could also generally be argued that the functions are 0/1-functions encoding the intended
	predicates and are 0 when the timestep is the successor of the current timestep)
	\item p.36: ``is defined as an extension of QFO with the formula ...'' $\to$	 ``is defined as the extension
	of QFO by the additional operator ...'' (or similar)
\end{itemize}
\vspace{1em}
\subsection*{Grammatical/Language}
\begin{itemize}
	\setlength\itemsep{0.5em}
	\item generally: both formulae and formulas are used as plural of formula---should be consistent
	\item p.2, in the middle: we propose a restriction: Should be something like ``we propose to restrict
	WL to the natural numbers as fixed semiring, calling this restriction ...''
	\item[$\checkmark$] p.3, first paragraph of 2.1: ``omit the word finite or ordered'' $\to$ ``omit the words finite and
	ordered''
	\item p5, description of syntax of QSO: ``This division between Boolean and quantitative level
	is essential for understanding the difference between the logic and the quantitative part.''
	reformulate this?
	\item[$\checkmark$] p.5, next sentence: ``this will allow us later'' $\to$ ``this will later allow us''
	\item[$\checkmark$] p.6, second to last sentence: ``similarity of $\alpha_3$ with'': ``similarity to'' is much more common
	than ``similarity with''
	\item[$\checkmark$] p.7, next sentence: ''in this direction'' $\to$ ``in this regard''
	\item[$\checkmark$] p.7, last sentence before 3.1: ''connection of QSO with'' $\to$ ``connection of QSO to''
	\item[$\checkmark$] p.7, middle: ``Another difference of WL with our approach'' $\to$ ``Another difference between
	WL and our approach''
	\item p.9, first sentence: ``then condition (2) holds'': omit the word then or reformulate the sentence
	\item[$\checkmark$] p.9, first paragraph: ``the constant function $s$ can be trivially simulated'' $\to$ ``any constant
	function $s$ can be trivially simulated''
	\item[$\checkmark$] p.9, proof of Prop. 4.3: ``To prove the condition (2)'' $\to$ ``To prove condition (2)''
	\item throughout the paper: ``similar than'' $\to$ ``similar to'' / ``similar than for X, we do Y'' $\to$
	``similarly as for''. Examples:
	\begin{itemize}
		\item[-] p.10, proof of Thm. 4.5: ``the proof is similar than in Theorem 4.4'' $\to$ ``the proof is
		similar to the proof of Theorem 4.4''	
		\item[-] p.10, second sentence: ``similar than the previous proof'' $\to$ ``similar as in the previous
		proof''
		\item[-] p.11, last sentence before example 4.7: ``Similar than for QSO'' $\to$ ``Similarly as for QSO''
	\end{itemize}
	\item p.11, end of proof of Thm. 4.5: ``Using an analogous argument'': state what the argument is
	analogous to (the argument in Thm. 4.4)
	\item[$\checkmark$] p.12, after Corollary 4.8: ``shows how robust and versatile is QSO'' $\to$ ``shows how robust and
	versatile QSO is''
	\item p.12, after Corollary 4.8: ``counting complexity classes [...] even beyond $\nat$'' - make more precise
	what ``beyond $\nat$'' refers to
	\item[$\checkmark$] p.12, ``given CNF propositional formula, that can be made true'': no comma
	\item p.12: ``in [43] they defined'' $\to$ something like ``$<$authors$>$ [43] defined''
	\item[$\checkmark$] p.12: ``we extend as follows QSO'' $\to$ ``we extend QSO by max and min quantifiers as follows''
	\item[$\checkmark$] p.12: ``MinQSO is defined analogously changing max with min'' $\to$ ``replacing max by min''
	\item[$\checkmark$] p.12: ``the same holds with MinQSO(FO)'' $\to$ ``the same holds for ...''
	\item[$\checkmark$] p.13: ``in contrast with'' $\to$ ``in contrast to''
	\item[$\checkmark$] p.13: ``follow the same approach but considering'' $\to$ ``but consider''
	\item[$\checkmark$] p.13: ``in the search of efficient approximation algorithms'' $\to$ ``in the search for''
	\item[$\checkmark$] p.14: ``the search of robust classes'' $\to$ ''the search for robust classes''
	\item p.14: ''the first desirable condition has to do with the closure properties satisfied by the class'':
	Reformulate this sentence
	\item[$\checkmark$] p.14: ``As in the cases of P and NP that are closed'' $\to$ ``Analogously to the cases of P and
	NP, which are closed under intersection and union, ...''
	\item[$\checkmark$] p.16: ``such that for both, $A\mod\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$'' $\to$ ''such that for $i \in \{1,2\}$: ... ''
	\item[$\checkmark$] p.16: ``Similarly to Theorem 5.1, we show ..'' $\to$ ``Similarly as in the proof of Theorem 5.1, we
	show''
	\item p.19: ``However, this class is not closed under sum, and then it is not robust under basic
	closure properties'': Should be reformulated
	\item[$\checkmark$] p.19: ``Let ... be this formula, where $\varphi$ is in first-order and quantifier-free'' $\to$ ``... where $\varphi$ is
	first-order and quantifier-free''
	\item[$\checkmark$] several occurrences: ``This is, ...'' $\to$ ``That is, ...'', for example two times on p.20
	\item[$\checkmark$] p.20: ``we focus only in proving'' $\to$ ``we focus only on proving''
	\item[$\checkmark$] p.20: ``We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of FO.''
	$\to$ ``We prove this for the more general case of $\eqso(\LL)$ with $\LL$ being a fragment of SO.''
	(or similar)

	\item[$\checkmark$] p.20: ``we will show a formula'' $\to$ ``give'' or ``construct'' instead of ``show''
	\item[$\checkmark$] p.20/21: ``$\shp$ is believed not to be closed under subtraction by one by some complexitytheoretic
	assumption'' $\to$ ``... due to some complexity-theoretic assumption'' (or similar)
	\item[$\checkmark$] p.21, next sentence: ``for a possible extension'' $\to$ ``as a possible extension''
	\item p.21: ``the desired extension has to be achieved by ... . More precisely, we define...'': The second
	sentence doesn’t make the statement of the first one more precise, but rather explains the
	actual definiton (which is chosen in accordance with the sentence before)
	\item[$\checkmark$] p.21: ``a parsimonious reduction from a function'' $\to$ ``a parsimonious reduction from any
	function''
	\item[$\checkmark$] p.21: ``we define a function that converts a formula $\alpha$ in...'' $\to$ ``that converts a given formula
	$\alpha$...''
	\item[$\checkmark$] p.22: ``we iterate for every tuple'' $\to$ ``we iterate over all tuples''
	\item[$\checkmark$] p.22: ``Then we will show how to define a formula $\varphi'$
	...'' $\to$ ``We will show how to define ...''
	\item[$\checkmark$] p.22: ''no variable from $\bar{x}$ are mentioned'' $\to$ ``... is mentioned''
	\item[$\checkmark$] p.24: ``if, and only if, if X, then Y'': This construction makes it unneccesarily hard to parse
	the sentence
	\item[$\checkmark$] p.24: ``Then, the conclusion of the implication...'' $\to$ ``The conclusion of the implication...''
	\item[$\checkmark$] p.25: ``For the sake of simplification...'' $\to$ ``For the sake of simplicity...''
	\item[$\checkmark$] p.26: ``In the previous section, we show ...'' $\to$ ``In the previous section we showed ...''
	\item[$\checkmark$] p.27: ``we postpone the proof the previous proposition after ...'' $\to$ ``we postpone the proof of
	the previous proposition to after ...''
	\item p.29: ``replace by $\perp$ and $\top$ \underline{wherever it corresponds}'': Should be reformulated
	\item[$\checkmark$] p.30: ``consist in computing $g_{\alpha}(\enc(\A))$ for each input ...'' $\to$ ``... on input ...''
	\item[$\checkmark$] p.30, same sentence: ``and then simulate'' $\to$ ``and then simulating''
	\item[$\checkmark$] p.33: ``In contrast with LFP'' $\to$ ``In contrast to LFP''
	\item[$\checkmark$] p.33: ``These operations are represented by the set $\{0,1,\emptyset\}$, respectively.'' $\to$ ``... by 0, 1 and
	$\emptyset$, respectively''
	\item[$\checkmark$] p.34: ``Notice that the values of each function $f$ in $\alpha$'' $\to$ ``Notice that the values of each
	function in $\alpha$''
	\item[$\checkmark$] p.34: ``that is why we talk about ...'' $\to$ ``which is why we talk about''
	\item[$\checkmark$] p.34: ``Besides, notice that notation ...'' $\to$ ``Besides, notice that the notation ...''
	\item[$\checkmark$] p.34: ``Function $h$ is used to indicate whether the head of the working tape is in some position
	at some point of time'' $\to$ ``... whether the head of the working tape is in a specific position
	at a certain point of time'' (or similar)
	\item[$\checkmark$] p.36: ``Given a relation signature $\R$'' $\to$ ``... relational signature ...''
	\item[$\checkmark$] p.37: ``As it was mentioned before'' $\to$ ``As mentioned before''
	\item[$\checkmark$] p.37: ``Next we construct a logarithmic-space ....'' $\to$ ``We construct a logarithmic-space ...''
	6
	\item[?] p.38: ``In this sense, there are several directions for future research'' $\to$ ``Consequently, there
	are ...'' (or similar)
	\item[$\checkmark$] p.38: ``In the same direction, ...'' $\to$ ``Similarly, ...'' (or similar)
	\item[$\checkmark$] p.38: ``so to have'' $\to$ ``in order to have'' (or similar)
\end{itemize}
\vspace{1em}
\subsection*{Typos}
\begin{itemize}
	\setlength\itemsep{0.5em}
	\item p.1, footnote: ``clases'' should be ``classes''
	\item throughout the paper: plural used with verb ending on s. Examples:
	\begin{itemize}
		\item[-] p.2, middle: ``fragments of QSO captures''
		\item[-] p.6, example 3.1: ``the sum quantifiers in $\alpha_1$ aggregates''
	\end{itemize}
	\item[$\checkmark$] p4, beginning of 2.2: ``some of \underline{the} are recalled here''
	\item[$\checkmark$] p6, example 3.2, first paragraph: ``otherwise $M(i.j)$'', dot used instead of comma
	\item[$\checkmark$] throughout the paper: a instead of an (espacially: a FO formula/assignment, a SO formula/assignment).
	Example:
	\begin{itemize}
		\item[-] p.9, third sentence: ``Given a FO formula'' $\to$ ``Given an FO formula''
		\item[-] p.38: ``As a example'' $\to$ ``As an example''
	\end{itemize}
	\item throughout the paper: singular used with verb not ending on s. Examples:
	\begin{itemize}
		\item[-] p.9, after proof of Prop. 4.3: ``use of a fragment [...] that capture''
		\item[-] p.11, end of proof of Thm. 4.5: ``and $\alpha$ reconstruct''
	\end{itemize}
	\item[$\checkmark$] p.9, middle: ``A second attempt could be based then'' $\to$ ``A second attempt could then be
	based''
	\item[$\checkmark$] p.11, 4.1: ``output of a function \underline{in} not a natural number''
	\item[$\checkmark$] p.12: ``we have to dinstinguished''
	\item[$\checkmark$] p.14, beginning of 5.1: ``subclases''
	\item[$\checkmark$] p.21: ``it does not mentioned'' $\to$ ``it does not mention''
	\item[$\checkmark$] p.22: ``Without lost of generality'' $\to$ ``Without loss of generality''
	\item[$\checkmark$] p.24: ``The previous claim and proof motivates the following definitions.'' $\to$ ``The previous
	claim and proof motivate the following definition.''
	\item p.30: ``to define a natural generalization for functions of the notion of least fixed point of
	LFP'' $\to$ ``to define a natural generalization of LFP to functions'' (or similar)
	\item[$\checkmark$] p.30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
	\item[$\checkmark$] p.31: ``we simply generalized these conditions'' $\to$ ``we simply generalize these conditions''
	\item[$\checkmark$] p.32: ``which is used to defined'' $\to$ ``which is used to define''
	\item[$\checkmark$] p.33: ``since $\beta$ is constructed from monotones operations'' $\to$ ``... monotone ...''
	\item[$\checkmark$] p.35: ``Formulas $\alpha_h$ is defined ...'' $\to$ ``Formula $\alpha_h$ is defined ...''
	7
	\item[$\checkmark$] p.37: ``we simulate $M_\beta$ on input $v[1/n]$'' $\to$ ``... with input $v[1/x]$''
\end{itemize}
\vspace{1em}
{\bf Other}
\begin{itemize}
	\setlength\itemsep{0.5em}
	\item[$\checkmark$] running title is ``COUNTING''
	\item on definition of functions, \verb|\colon| should be used instead of :
	\item In [1], a similar hierarchy to the one by Saluja et al. was introduced and studied. This work
	should be cited.
\end{itemize}

\bigskip

\section*{Reviewer 2}

\begin{itemize}
	\setlength\itemsep{0.5em}
	\item In several proofs, you rely on disjoint union of solutions and cartesian product of solutions. Indeed, its gives a polynomial time counting reduction. Can you make connections with (+,*) circuits ? It would be similar to what is done in A Circuit-Based Approach to Efficient Enumeration (Antoine Amarilli, Pierre Bourhis, Louis Jachiet, Stefan Mengel) where simple circuits (dDNNF, which are basically disjoint unions and cartesian products) are used to capture the solutions of MSO queries.

	\item In general I would be interested to know the relationship of your couting classes with their enumeration counterpart. 
	For instance there is a paper adapting [39] to enumeration: Enumeration Complexity of Logical Query Problems with Second-order Variables (Arnaud Durand and Yann Strozecki). In particular they consider class with some Horn structure to capture easy enumeration problems, it may have some connection with the class $\sigma$ QSO($\Pi_1$ Horn) you introduce.

	\item For the most involved proofs, you could describe informally your proof strategy in a few lines to help the reader. You may also add some example to illustrate the different fragment of QSO (you already do it sometimes and it is very helpful).

	\item[$\checkmark$] page 4:
	-some of the $\to$ some of them 
	\item[$\checkmark$] page 10: 
	-$2^{|A|^l}$-bits $\to$ $2^{|A|^l}$ bits
	\item[$\checkmark$] page 21:
	-it does not mentioned $\to$ it does not mention
	\item page 22: 
	-you make a long proof about some closure under subtraction by one.
	You should put the result in a theorem envirpnment to make what you are
	proving more evident.
	\item[$\checkmark$] -replace .... for $\to$ replace ... by;  there are several instances of
	that
	\item[$\checkmark$] -$\phi$ is an FO-formula $\to$ a $\Sigma_1[FO]$ formula
	\item -you explain that you rewrite your formula to obtain conditions a,b and c, 
	you should give an idea how these conditions will be used
	\item[$\checkmark$] page 23:
	a SO, a FO  an SO, an FO
	\item page 24: 
	if, if $\phi$$\to$ $\phi$
	\item page 25: 
	-you explain that the proof relies on witnesses of logarithmic size.
	It was not clear to me while reading the proof. Can you explain more ?
	\item page 28:
	-the complete problem for $\Sigma$ QSO($\Sigma_2$ Horn) would be complete also for $\#\Sigma-2$ Horn. You should mention here that you do not need sum here
	and that the framework of [39] is enough for this result.
	\item[?] page 30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''

	\item[$\checkmark$] page 32: 
	- paths of length at most n $\to$ in an acylic graph, all paths are of size less than n
	\item page 33:
	-there is a proof symbol \verb|\qed| but it does not end a well defined proof environment

	\item page 38: 
	the last sentence on fixed point operator is to vague. Either expand it to 
	say what king of question on fixed point operator you want to investigate or remove it	
\end{itemize}

\end{document}

### 2019-01-19 LMCS/sendrev/counting.tex
correcciones en la carpeta sendrev
carta de respuesta + correcciones varias

--- 2017-07-24 LMCS/counting.tex
@@
 
-Fix a signature $\R = \{R_1, \ldots, R_k\}$ and assume that $\A$ is an ordered $\R$-structure with a domain $A = \{a_1, \ldots, a_n\}$ and $a_1 < a_2 < \ldots < a_n$.
-For every $i \in \{1, \ldots, k\}$, define the encoding of $R_i^\A$, denoted by $\enc(R_i^\A)$, as the following binary string.
+Fix a signature $\R = \{R_1, \ldots, R_k\}$ and assume that $\A$ is an ordered $\R$-structure with a domain $A = \{a_1, \ldots, a_n\}$, $R_k =\, <$, and $a_1 <^{\A} a_2 <^{\A} \ldots <^{\A} a_n$.
+For every $i \in \{1, \ldots, k-1\}$, define the encoding of $R_i^\A$, denoted by $\enc(R_i^\A)$, as the following binary string.
 Assume that $\ell = \arity(R_i)$ and consider an enumeration of the $\ell$-tuples over $A$ in the lexicographic order induced by $<$.
@@
 $$
-0^n \, 1 \, \enc(R_1^\A) \, \cdots \, \enc(R_k^\A).
+0^n \, 1 \, \enc(R_1^\A) \, \cdots \, \enc(R_{k-1}^\A).
 $$
+\martin{modifique la definicion para que no sea necesario codificar el $<$.}
 %We define the class of all $\R$-functions, denoted by $\Func(\R)$, as the class of all functions $f: \ostr \rightarrow \bbN$.
@@
 	\end{enumerate} 
-	Moreover, {\em $\FF$ captures $\CC$ over ordered structures} if $\FF$ captures~$\CC$ over ordered $\R$-structures for every signature~$\R$.
+	Moreover, {\em $\FF$ captures $\CC$ over ordered structures} if $\FF$ captures~$\CC$ over ordered $\R$-structures for every signature~$\R$. \qed
 \end{defi}
@@
 Such an approach fails as $\# \LL$ can encode $\shp$-complete problems in both cases; in the first case, one can encode the problem of counting the number of satisfying assignments of a Horn  propositional formula, while in the second case one can encode the problem of counting the number of satisfying assignments of a 2-CNF propositional formula.
-A second attempt could be based then on considering a fragment $\LL$ of $\fo$.
+A second attempt could then be based on considering a fragment $\LL$ of $\fo$.
 But even if we consider the existential fragment $\Sigma_1$ of $\fo$ the approach fails, as $\# \Sigma_1$ can encode $\shp$-complete problems like counting the number of satisfying assignments of a 3-DNF propositional formula\cite{SalujaST95}.

### 2019-01-19 LMCS/sendrev/framework.tex
carpeta con version send corregida
correcciones en la carpeta sendrev
carta de respuesta + correcciones varias

--- 2018-12-09 LMCS/framework.tex
@@
 
-We consider several fragments or extensions of $\qso$, which are obtained by restricting the syntax of the Boolean formulae or the use of the quantitative quantifiers.
+We consider several fragments of $\qso$, which are obtained by restricting the syntax of the Boolean formulae or the use of the quantitative quantifiers, and we consider some extensions that are obtained by adding recursive operators to $\qso$.
 In this regard, we denote by $\qfo$ the fragment of $\qso$ where second-order sum and product are not allowed. 

### 2019-01-19 LMCS/sendrev/introduction.tex
mas revisiones
carta de respuesta + correcciones varias

--- 2018-05-07 LMCS/introduction.tex
@@
 % ($\pspace$ resp.). 
-However, other function complexity classes have emerged from the need to understand the complexity of some computation problems for which little can be said if their decision counterparts are considered.
-This is the case of the class $\shp$, a counting complexity class introduced in \cite{Valiant79} to prove that natural problems like counting the number of satisfying assignments of a propositional formula or the number of perfect matchings of a bipartite graph~\cite{Valiant79} are difficult, namely, $\shp$-complete.
+However, other function complexity classes have emerged from the need to prove that some functions are difficult to compute, even though their decision counterparts can be solved efficiently.
+This is the case of the class $\shp$, a counting complexity class introduced in \cite{Valiant79} to prove that natural problems like counting the number of satisfying assignments of a propositional formula in DNF or the number of perfect matchings of a bipartite graph~\cite{Valiant79} are difficult, namely, $\shp$-complete, even though their decision counterparts can be solved in polynomial time.
 Starting from $\shp$,

### 2019-01-19 LMCS/sendrev/proof-capture-fpspace.tex
correcciones en la carpeta sendrev
carta de respuesta + correcciones varias

--- 2018-12-10 LMCS/proof-capture-fpspace.tex
@@
+% !TeX root = main.tex
+
 For the first condition of Definition~\ref{def:cap}, notice that each $\pfp$ formula can be evaluated in deterministic polynomial space, the constant function $s$ can be trivially simulated in $\fpspace$, and $\fpspace$ is closed under exponential sum and multiplication.
 This suffices to show that the condition holds.
-For the second condition, the proof is similar than in Theorem~\ref{theo:capture-fp}.
+For the second condition, the proof is similar to the proof of Theorem~\ref{theo:capture-fp}.
 Let $f\in \fpspace$ defined over some $\R$ and $\ell\in\nat$ such that $\log_2\left( f(\enc(\A)) \right) \leq 2^{{|\A|}^\ell}$ for every $\A\in\ostr[\R]$  (i.e.
-$2^{{|\A|}^\ell}$ is an upper bound for the output size).
+$2^{{|\A|}^\ell}$ is an upper bound for the size of the output).
 Let $X$ be a second-order variable of arity $\ell$.
@@
 $$
-Namely, we use relations to encode numbers with at most $2^{{|\A|}^\ell}$ bits where the empty relation represents $0$ and the total-relation represents $2^{2^{{|\A|}^\ell}}-1$.
+Namely, we use predicates to encode a number that will have most $2^{{|\A|}^\ell}$ bits.
+We define this encoding through the function $\tau\colon 2^{A^\ell}\to\nat$, such that $\tau(B)$ is equal to the number of predicates in $2^{A^\ell}$ that are smaller than $B$ with respect to the induced order.
+For example, we have that $\tau(\emptyset) = 0$ and $\tau(A^{\ell}) = 2^{{|\A|}^\ell}-1$.
 Furthermore, we can use a relation~$X$ to index a position in the binary output of $f(\enc(\A))$ as follows.
@@
 \[
-L = \{(\A,B)\mid B \subseteq A^{\ell}\text{ and the $B$-th bit of $f(\enc(\A))$ is 1}\}.
+L = \{(\A,B)\mid B \subseteq A^{\ell}\text{ and the $\tau(B)$-th bit of $f(\enc(\A))$ is 1}\}.
 \]
-Since $L$ is in $\pspace$, it can be specified in $\pfp$ \cite{AbiteboulV89} by a formula $\Phi(X)$ where the free variable $X$ encodes relation $B$ in $L$.
-Then, similar than the previous proof we define:
+Since $L$ is in $\pspace$, it can be specified in $\pfp$ \cite{AbiteboulV89} by a formula $\Phi(X)$ such that $\A\models\Phi(B)$ if and only if $(\A,B)\in L$.
+Then, similarly as for the previous proof we define:
 $$
@@
 $$ 
-where $\pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$ takes the value $2^m$ if there exist $m$ predicates that are smaller than~$X$ and $\alpha$ reconstruct the output of $f(\enc(\A))$ by simulating $f$ with $\Phi(X)$.
-Using an analogous argument, we conclude that $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
+where $\pa{Y}(\varphi_{<}(Y,X)\mapsto 2)$ takes the value $2^{\tau(X)}$ and $\alpha$ reconstructs the output of $f(\enc(\A))$.
+Using an argument analogous to the previous proof, we conclude that $\alpha\in\qso(\pfp)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
+\martin{Reescrib\'i varias l\'ineas de esta demostraci\'on}

### 2019-01-19 LMCS/sendrev/proof-prop-rest.tex
mas revisiones
carta de respuesta + correcciones varias

@@
+%!TEX root = main.tex
+
 We give this proof in three parts.
 First, we show that $\E{1} \not\subseteq \QE{0}$. 
-Towards a contradiction, let $\R = \{ < \}$ and suppose that there is a $\QE{0}$ formula $\alpha$ over $\R$ which is equivalent to the $\E{1}$ formula $\sa{x} \ex{y} (x < y)$. 
-This is, for every finite $\R$-structure $\A$, $\sem{\alpha}(\A) = \size{\A} - 1$.
-
+For this inclusion to be true, it is required to hold for an arbitrary ordered relational signature $\R$, so it suffices to show it is not for at least one.
+Let $\R$ be the minimal ordered signature that contains only the relation name $<$.
+Suppose that there is a $\QE{0}$ formula $\alpha$ over $\R$ which is equivalent to the $\E{1}$ formula $\sa{x} \ex{y} (x < y)$. 
+That is, for every finite $\R$-structure $\A$, $\sem{\alpha}(\A) = \size{\A} - 1$.
+\martin{reformul\'e por qu\'e elijo ese R}
 Suppose that $\alpha$ is in SNF, namely, $\alpha = \sum_{i = 1}^k \alpha_i$ for some fixed $k$. 
-Since $\alpha$ is not null, consider some $\alpha_i$ that describes a non-null function. 
+Since $\sem{\alpha}$ is not the identically zero function, consider some $\alpha_i$ that describes a non-null function. 
 Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ where $\varphi$ is quantifier-free. 
@@
 	Let $\bar{a}$ be an assignment to $\bar{x}$ such that $\A\models\varphi(\bar{a})$. 
-	It can be seen that each assignment $\bar{a}'$ that has the same order over its variables\footnote{For example, if $\bar{a} = (a_1,a_2,a_3,a_4)$, such order may be $a_1 > a_3 = a_4 > a_2$, which has tree partitions.}	as $\bar{a}$ also satisfies $\A\models\varphi(\bar{a}')$.
-	If this order has $k$ partitions then $\binom{\size{\A}}{k}$ assignments for $\bar{x}$ satisfy this order, and therefore, $\sem{\alpha} \geq \binom{\size{\A}}{k}$. 
+	It can be seen that each assignment $\bar{a}'$ that has the same ordering over its variables\footnote{For example, if $\bar{a} = (a_1,a_2,a_3,a_4)$, such ordering may be $a_1 > a_3 = a_4 > a_2$, which has tree partitions.}	as $\bar{a}$ also satisfies $\A\models\varphi(\bar{a}')$.
+	If this ordering has $k$ partitions then $\binom{\size{\A}}{k}$ assignments for $\bar{x}$ will satisfy it, and therefore, $\sem{\alpha} \geq \binom{\size{\A}}{k}$. 
 	If $k = 1$, then $\binom{\size{\A}}{k} = \size{\A}$, and if $k \geq 2$, then $\binom{\size{\A}}{k} \in \Omega(\size{\A}^2)$, which proves the claim.
 \qed
-
+\martin{cambie order por ordering, creo que se entiende mejor}
 Now we show that $\QE{0} \not\subseteq \E{1}$. 
 In Theorem \ref{theo-pi1-pnf} we proved that there is no formula in $\loge{1}$-PNF equivalent to the formula $\alpha = 2$. 
-Every formula in $\E{1}$ can be expressed in $\loge{1}$-PNF, which implies that $2 \in \QE{0}$ and $2 \not\in \E{1}$.
+Every formula in $\E{1}$ can be expressed in $\loge{1}$-PNF, which implies that $2 \not\in \E{1}$.
+Therefore, given that $2 \in \QE{0}$ by the definition of this logic, we conclude that $\QE{0} \not\subseteq \E{1}$.
 

### 2019-01-19 LMCS/sendrev/proof-sigmafo-minusone.tex
correcciones en la carpeta sendrev
carta de respuesta + correcciones varias

--- 2018-12-10 LMCS/proof-sigmafo-minusone.tex
@@
 
-To show that $\alpha$ is in $\totp$, we can convert $\alpha$ and $\enc(\A)$ into $\lambda(\alpha)$ and $\enc(g_{\alpha}(\A))$, respectively, and run the procedure in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
+To show that the function defined by $\alpha$ is in $\totp$, we can convert $\alpha$ and $\enc(\A)$ into $\lambda(\alpha)$ and $\enc(g_{\alpha}(\A))$, respectively, and run the procedure in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
 Similarly, to show that $\alpha$ has an FPRAS, we do the same as before and simulates the FPRAS for $\lambda(\alpha)$ in Proposition~\ref{prop:qe0-fp-qe1-totp-fptras}.
@@
 	$$
-	\varphi_i(\bar{X},\bar{x}) \ := \ \ex{\bar{y}} \ex{\bar{z}} (\bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y})) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y})[X_j(\bar{w}) \leftarrow X_j(\bar{z})] \wedge
-	\varphi_i^{-}(\bar{X},\bar{x},\bar{y}).
+	\varphi_i(\bar{X},\bar{x}) \ := \ \ex{\bar{y}} \ex{\bar{z}} \big( \bar{z} = \bar{w} \wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{x},\bar{y})[X_j(\bar{w}) \leftarrow X_j(\bar{z})] \wedge
+	\varphi_i^{-}(\bar{X},\bar{x},\bar{y}) \big).
 	$$
@@
 	$$
-	\varphi_i(\bar{X},\bar{x}) \ := \  \ex{\bar{y}} \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y}).
+	\varphi_i(\bar{X},\bar{x}) \ := \  \ex{\bar{y}} [ \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})].
 	$$
-	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = \ex{\bar{y}}  x < y \wedge  X(x,y)\wedge \neg X(x,x)$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and:
+	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = \ex{\bar{y}} [ x < y \wedge  X(x,y)\wedge \neg X(x,x)]$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and:
 	$$
-	\varphi_i \ := \ \ex{\bar{y}} v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x \wedge x < y  \wedge  X(v_1,v_2) \wedge \neg X(v_3,v_4).
+	\varphi_i \ := \ \ex{\bar{y}}[ v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x \wedge x < y  \wedge  X(v_1,v_2) \wedge \neg X(v_3,v_4)].
 	$$ 
@@
 \[
-\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge \fa{\bar{x}'} \fa{\bar{y}'}(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')).
+\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge \fa{\bar{x}'} \fa{\bar{y}'}[\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')].
 \] 
@@
 &\Big( \; 
-\ex{\bar{y}} \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \; \varphi_i'(\bar{X},\bar{x}, \bar{y})  \rightarrow \!\!
+\ex{\bar{y}}[ \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \; \varphi_i'(\bar{X},\bar{x}, \bar{y})  \rightarrow \!\!
 \bigvee_{X \in \bar{X}} \ex{\bar{z}} [ \, X(\bar{z}) \wedge   \!\!\!\!\!\!\!\!\!\!
 \bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{y})} \!\!\!\!\!\!\!\!\!\!\!
-\bar{w}\neq\bar{z} \, ] \; \big) \;  \vee   \label{eq:part2} \\
-&(\; \ex{\bar{x}'} \ex{\bar{y}} \varphi_i'(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x} \; ) \;\; \Big)  \label{eq:part3}
+\bar{w}\neq\bar{z} \, ] \; \big)] \;  \vee   \label{eq:part2} \\
+&(\; \ex{\bar{x}'} \ex{\bar{y}}[ \varphi_i'(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}] \; ) \;\; \Big)  \label{eq:part3}
 \end{align}
+\martin{agregu\'e parentesis de aqui hacia arriba}
 To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 

### 2019-01-19 LMCS/sendrev/saluja.tex
correcciones en la carpeta sendrev
carta de respuesta + correcciones varias

--- 2018-12-09 LMCS/saluja.tex
@@
 $\sa{\bar{X}} \sa{\bar{x}} \varphi(\bar{X}, \bar{x})$,
-where $\bar{X}$ and $\bar{x}$ are sequences of zero or more second-order and first-order variables, respectively, (as expected, $\sa{\bar{X}}\!$ is the respective nesting of $\sa{X}$'s) and $\varphi(\bar{X}, \bar{x})$ is a formula in $\LL$.
+where $\bar{X}$ and $\bar{x}$ are sequences of zero or more second-order and first-order variables, respectively, (as expected, $\sa{\bar{X}}\!$ and $\sa{\bar{x}}\!$ are the respective nestings of $\sa{X}$'s and $\sa{x}$'s) and $\varphi(\bar{X}, \bar{x})$ is a formula in $\LL$.
 Notice that
 a formula $\varphi(\bar{X}, \bar{x})$ in $\sh{\LL}$ is equivalent to the formula $\sa{\bar X} \sa{\bar x} \varphi(\bar{X}, \bar{x})$ in $\LL$-PNF. 
-Moreover, a formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-sum normal form} ($\LL$-SNF) if $\alpha$ is of the form $\sum_{i=1}^n \alpha_i$ where each $\alpha_i$ is in $\LL$-PNF. 
+Moreover, a formula $\alpha$ in $\eqso(\LL)$ is in \emph{$\LL$-sum normal form} ($\LL$-SNF) if $\alpha$ is of the form $\sum_{i=1}^n \alpha_i$, where this is a shorthand notation for $\alpha_1+\cdots+\alpha_n$, and each $\alpha_i$ is in $\LL$-PNF.
 \begin{prop}\label{theo-pnf-snf}

### 2019-01-19 LMCS/sendrev/todo-list.tex
carta de respuesta + correcciones varias

@@
-% !TeX root = main.tex
+\documentclass[a4paper]{article}
 
-\section*{\LARGE \textbf{TODO LIST}}
+\usepackage{fullpage}
+\usepackage{cite}
+\usepackage{amsmath,amsfonts,amssymb,mathabx}
+\usepackage{stmaryrd}
+\usepackage[textwidth=2cm,textsize=small]{todonotes}
+\usepackage{enumitem}
+\input{macros}
 
-\bigskip
+\def\eg{{\em e.g.}}
+\def\cf{{\em cf.}}
 
-{\bf Reviewer 1}
+\begin{document}
 
-\bigskip
+\title{TODO LIST}
+\date{}
 
-{\bf Technical}
+\maketitle
+
+\section*{Reviewer 1}
+
+\subsection*{Technical}
 
@@
 	\setlength\itemsep{0.5em}
-\item p.2, first paragraph: ``However, other function classes have emerged from the need to understand
+\item[$\checkmark$] p.2, first paragraph: ``However, other function classes have emerged from the need to understand
 the complexity of some computation problems for which little can be said if their
@@
 of the examples mentionend afterwards?
-\item p.7, first sentence: ``fragments or extensions [...] obtained by restricting'': extensions are not
+\item[$\checkmark$] p.7, first sentence: ``fragments or extensions [...] obtained by restricting'': extensions are not
 obtained by restricting something
-\item p.10, before Thm. 4.5: ``by considering second-order product'': We need both second-order
+\item[$\checkmark$] p.10, before Thm. 4.5: ``by considering second-order product'': We need both second-order
 product and sum (at least both are used in Thm. 4.5)
-\item p.10, last line: ``$B$-th bit'': mention in what way $B$ is interpreted as a number
+\item[$\checkmark$] p.10, last line: ``$B$-th bit'': mention in what way $B$ is interpreted as a number
 \item[$\checkmark$] p.12: ``To formalize this set of optimization problems'': class instead of set
-\item p.14: ``we have to concentrate on the class of counting problems with easy decision versions'':
+\item[$\checkmark$] p.14: ``we have to concentrate on the class of counting problems with easy decision versions'':
 here, easy decision version means ``in \bpp''. Before, it was defined as ``in P''. Should be
 consistent.
-\item p.14: ``as expected, $\Sigma \bar{X}$ is the respective nesting of $\Sigma X.$’s'': why is this explained for secondorder
+\item[$\checkmark$] p.14: ``as expected, $\Sigma \bar{X}$ is the respective nesting of $\Sigma X.$’s'': why is this explained for secondorder
 quantifiers but not for first-order quantifiers?
-\item p.14: $\sum_{i=1}^n\alpha_i$: explain that this is notation for $\alpha_1+\cdots+\alpha_n$
-\item p.17: ``Since $\alpha$ is not null'' $\to$ ``Since $\sem{\alpha}$ is not the identically zero function'' (or similar)
+\item[$\checkmark$] p.14: $\sum_{i=1}^n\alpha_i$: explain that this is notation for $\alpha_1+\cdots+\alpha_n$
+\item[$\checkmark$] p.17: ``Since $\alpha$ is not null'' $\to$ ``Since $\sem{\alpha}$ is not the identically zero function'' (or similar)
 \item p.17/18: In the proof of Prop 5.3, the vocabulary only contains the numerical predicate $<$. When encoding structures as binary words, numerical predicates are implicit (and not
 encoded), though. What is the intended meaning here?
-\item p.18, after proof of Claim 5.4: ``which implies that 2 $\in \eqso(\Sigma_0)$ and $2 \not\in \E{1}$'': $2 \in \eqso(\Sigma_0)$
+\item[$\checkmark$] p.18, after proof of Claim 5.4: ``which implies that 2 $\in \eqso(\Sigma_0)$ and $2 \not\in \E{1}$'': $2 \in \eqso(\Sigma_0)$
 does not follow from the argument before as stated in this sentence, but is trivial
@@
 \item[$\checkmark$] p.20: ``we will show a formula in the grammar'' $\to$ ``... in the logic''
-\item p.20: ``Since $\Sigma_1$ is closed under conjunction, then Lemma 5.7 holds for ...'' $\to$ ``Since $\Sigma_1$ is
+\item[$\checkmark$] p.20: ``Since $\Sigma_1$ is closed under conjunction, then Lemma 5.7 holds for ...'' $\to$ ``Since $\Sigma_1$ is
 closed under cnojunction, Lemma 5.7 can be applied to ...'' (or similar)
 \item[$\checkmark$] p.22: ``and leaving each $R_{\psi}^{\A}$ empty'' $\to$ $R_{\psi}^{\A'}$
-\item p.22: ``To show that $\alpha$ is in \totp'' $\to$ ``To show that the function defined by $\alpha$ ...'' (or similar)
+\item[$\checkmark$] p.22: ``To show that $\alpha$ is in \totp'' $\to$ ``To show that the function defined by $\alpha$ ...'' (or similar)
 \item[$\checkmark$] p.22: ``Let $\alpha$ = ... where $\varphi$ is an FO-formula'' $\to$ ``... where $\varphi$ is an $\Sigma_1$[FO]-formula''
@@
 \vspace{1em}
-{\bf Grammatical/Language}
+\subsection*{Grammatical/Language}
 \begin{itemize}
@@
 \vspace{1em}
-{\bf Grammatical/Language}
+\subsection*{Typos}
 \begin{itemize}
@@
 
-{\bf Reviewer 2}
+\section*{Reviewer 2}
 
@@
 and that the framework of [39] is enough for this result.
-\item[?] page 30:
--not, trivial $\to$ not trivial
+\item[?] page 30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
+
 \item[$\checkmark$] page 32: 
@@
 \end{itemize}
+
+\end{document}

### 2019-01-21 LMCS/sendrev/response-letter.tex
reviewer 1 - technical casi listo

@@
 \section*{Reviewer 1}
-
-\subsection*{Technical}
 
@@
 %	\item[$\checkmark$] p.22: ``Let $\alpha$ = ... where $\varphi$ is an FO-formula'' $\to$ ``... where $\varphi$ is an $\Sigma_1$[FO]-formula''
-	\item p.23, formulae: Maybe put (large) parantheses around the relativization of quantifiers, for	example the part of formula $\varphi_i$ after the quantifiers
-	\item[$\checkmark$] p.24, formula 5.6: subscript of large conjunction: $\varphi_i^{+}(\bar{X},\bar{v})$ $\to$ $\varphi_i^{+}(\bar{X},\bar{y})$
-	\item[$\checkmark$] p.25: ``We construct a function $\lambda$ that receives a quantitative formula $\beta$ ...'' $\to$ ``... that receives
-	a formula $\beta \in \eqso(\logex{1})$ ...''
-	\item p.25, last sentence: ``which is based on considering some witnesses of logarithmic size'' could
-	be explained a little more / made more clear
-	\item p.26: ``... auxiliary predicate $A$ defined as'' $\to$ ``... auxiliary predicate $A$ which will be defined
-	as ... in the formula'' (or similar)
-	\item p.28: ``since universal formulas are monotone over induced substructures'': ``monotone over
-	induced substructures'' should be reformulated
-	\item p.28: ``Next we prove that ... is hard for ... over a signature $\R$ ...'' sounds like it only works
-	for one specific vocabulary. Maybe simply ``over any signature $\R$''
-	\item p.28, formula $\alpha_{\A}$: second big conjunction should have $\bar{z}'$ in the subscript instead of $\bar{z}$ (because
-	the quantifier $\forall u$ is handled by this as well if I understand correctly)
-	\item p.29: ``In other words, in poly time we can replace $\varphi_j^i$ by a disjunction of ....'': Some of these
-	formulae might evaluate to $\perp$ or $\top$ altogether and can be removed (potentially together with
-	the surrounding conjunction)
-	\item p.29: ``Intuitively $t_1,\ldots,t_{m'}$ are used to count from 1 to $m'$ each subformula ...'': Should be
-	reformulated, maybe rather write that they are used to have disjoint sets of propositional
-	assignments for the different disjuncts of the outermost disjunction (corresponding to the
-	summands in the original formula)
-	\item p.30, proof of Prop. 5.12: For $\dhsat \in \totp$, [35] is cited. It’s stated that a \totp-procedure
-	for DNF-SAT is given there and it can be adapted to \dhsat. While the
-	cited paper contains the result for DNF-SAT, the procedure is not explicitly given as far as
-	I can see, so I can't check whether it is adaptable. On the other hand, membership of any
-	function counting the number of satisfying assignments for a class of propositional formulae
-	with poly-time decision version in \totp should hold by a similar argument as membership in
-	DelayP. Maybe change the citation or give a short proof sketch.
-	\item[$\checkmark$] p.31, definition of $\leq$ on $\F$: ``$f(i) \leq g(i)$ for every $i \in \{1,\ldots,\ell\}$'': Inputs for $f$ and $g$ should be
-	elements of $A^{\ell}$, not $\{1,\ldots,\ell\}$
-	\item p.32/33, example 6.2: I don't see why the formula given for dag's does not work for arbitrary
-	graphs: In arbitrary graphs, the fixpoint is still reached after at most $n-1$ steps, since the
-	maximal length of any shortest path in any graph is $n-1$. On the other hand, assuming that
-	I am mistaken and the support does not become stable within $n-1$ steps: Using the formula
-	given for arbitrary graphs, the support of the constructed function restricted to those tuples
-	that have first in their last component is the same as the support of the constructed function
-	using the formula for dags in each step of the recursion. Hence, the formula should not be
-	able to fix such a problem.
-	\item p.33: ``executes exactly $n^k$ steps for a fixed $k \geq 1$'' → ``executes exactly $n^k$
-	steps on large
-	inputs for a fixed $k \geq 1$'' (at least it does not work for $n = 1$) (small inputs can be handled
-	seperately)
-	\item p.35 and p.36: While $\sum$ was used for finite sums in \qso-formulae throughout the paper, now
-	a large +-sign is used. This might be confusing
-	\item p.36: $\alpha_{q_i}$ used instead of $\alpha_{s_{q_{i}}}$
-	\item p.36: ``Clearly, at each iteration of the fixed point operator, the tuple $\bar{t}$ represents the step the
-	machine is currently in.'': This is quite vague. Maybe something like ``Clearly, tuple $\bar{t}$ encodes
-	the number of steps the machine has done and in each iteration of the fixed point operator,
-	one timestep of the machine is executed''
-	\item p.36: ``at the $\bar{a}$-th iteration of the fixed point operator, it holds that $f(\bar{a}, \bar{b}) \leq 1$ and $f(\bar{a}', \bar{b}) = 0$ for every $\bar{b} \in A^k$. 
+%	\item p.23, formulae: Maybe put (large) parantheses around the relativization of quantifiers, for	example the part of formula $\varphi_i$ after the quantifiers
+%	\item[$\checkmark$] p.24, formula 5.6: subscript of large conjunction: $\varphi_i^{+}(\bar{X},\bar{v})$ $\to$ $\varphi_i^{+}(\bar{X},\bar{y})$
+%	\item[$\checkmark$] p.25: ``We construct a function $\lambda$ that receives a quantitative formula $\beta$ ...'' $\to$ ``... that receives a formula $\beta \in \eqso(\logex{1})$ ...''
+	\item \red{p.25, last sentence: ``which is based on considering some witnesses of logarithmic size'' could be explained a little more / made more clear}
+
+	WIP
+
+%	\item p.26: ``... auxiliary predicate $A$ defined as'' $\to$ ``... auxiliary predicate $A$ which will be defined as ... in the formula'' (or similar)
+%	\item p.28: ``since universal formulas are monotone over induced substructures'': ``monotone over induced substructures'' should be reformulated
+%	\item p.28: ``Next we prove that ... is hard for ... over a signature $\R$ ...'' sounds like it only works for one specific vocabulary. Maybe simply ``over any signature $\R$''
+%	\item p.28, formula $\alpha_{\A}$: second big conjunction should have $\bar{z}'$ in the subscript instead of $\bar{z}$ (because the quantifier $\forall u$ is handled by this as well if I understand correctly)
+%	\item p.29: ``In other words, in poly time we can replace $\varphi_j^i$ by a disjunction of ....'': Some of these formulae might evaluate to $\perp$ or $\top$ altogether and can be removed (potentially together with the surrounding conjunction)
+%	\item p.29: ``Intuitively $t_1,\ldots,t_{m'}$ are used to count from 1 to $m'$ each subformula ...'': Should be reformulated, maybe rather write that they are used to have disjoint sets of propositional assignments for the different disjuncts of the outermost disjunction (corresponding to the summands in the original formula)
+	\item \red{p.30, proof of Prop.
+5.12: For $\dhsat \in \totp$, [35] is cited.
+It’s stated that a \totp-procedure for DNF-SAT is given there and it can be adapted to \dhsat.
+While the cited paper contains the result for DNF-SAT, the procedure is not explicitly given as far as I can see, so I can't check whether it is adaptable.
+On the other hand, membership of any function counting the number of satisfying assignments for a class of propositional formulae with poly-time decision version in \totp should hold by a similar argument as membership in DelayP.
+Maybe change the citation or give a short proof sketch.}
+
+	A proof was added.
+
+%	\item[$\checkmark$] p.31, definition of $\leq$ on $\F$: ``$f(i) \leq g(i)$ for every $i \in \{1,\ldots,\ell\}$'': Inputs for $f$ and $g$ should be elements of $A^{\ell}$, not $\{1,\ldots,\ell\}$
+	\item \red{p.32/33, example 6.2: I don't see why the formula given for dag's does not work for arbitrary graphs: In arbitrary graphs, the fixpoint is still reached after at most $n-1$ steps, since the maximal length of any shortest path in any graph is $n-1$.
+On the other hand, assuming that I am mistaken and the support does not become stable within $n-1$ steps: Using the formula	given for arbitrary graphs, the support of the constructed function restricted to those tuples that have first in their last component is the same as the support of the constructed function using the formula for dags in each step of the recursion.
+Hence, the formula should not be able to fix such a problem.}
+
+	In a graph that is only one cycle of size $n$, namely $C_n$, the fixed point is reached after step $n$.
+
+%	\item p.33: ``executes exactly $n^k$ steps for a fixed $k \geq 1$'' → ``executes exactly $n^k$ steps on large inputs for a fixed $k \geq 1$'' (at least it does not work for $n = 1$) (small inputs can be handled seperately)
+%	\item p.35 and p.36: While $\sum$ was used for finite sums in \qso-formulae throughout the paper, now a large +-sign is used. This might be confusing
+%	\item p.36: $\alpha_{q_i}$ used instead of $\alpha_{s_{q_{i}}}$
+%	\item p.36: ``Clearly, at each iteration of the fixed point operator, the tuple $\bar{t}$ represents the step the machine is currently in.'': This is quite vague.
+Maybe something like ``Clearly, tuple $\bar{t}$ encodes the number of steps the machine has done and in each iteration of the fixed point operator, one timestep of the machine is executed''
+%	\item p.36: ``at the $\bar{a}$-th iteration of the fixed point operator, it holds that $f(\bar{a}, \bar{b}) \leq 1$ and $f(\bar{a}', \bar{b}) = 0$ for every $\bar{b} \in A^k$.
 In the same way, it can be seen that for each function $g \in \ldots$ it holds that $g(\bar{a}) \leq 1$ and $g(\bar{a}')= 0$.
-	I think that more properties of these functions should be mentioned
-	in order to argue that the defined function is the function computed by the machine:
-	\begin{itemize}
-		\item[-] For each $\bar{b}$, only one of the functions $T_0$,$T_1$,$T_B$,$T_{\vdash}$ outputs 1 on input ($\bar{a}, \bar{b})$
-		\item[-] For each $\bar{b}$, only one of the functions $h,\hat{h}$ outputs 1 on input $(\bar{a}, \bar{b})$
-		\item[-] There is exactly one $\bar{b}$ with $h(\bar{a}, \bar{b}) = 1$
-		\item[–] There is exatcly one $i$ such that $s_{q_i}(\bar{a}) = 1$
-	\end{itemize}
-	(It could also generally be argued that the functions are 0/1-functions encoding the intended
-	predicates and are 0 when the timestep is the successor of the current timestep)
-	\item p.36: ``is defined as an extension of QFO with the formula ...'' $\to$	 ``is defined as the extension
-	of QFO by the additional operator ...'' (or similar)
+I think that more properties of these functions should be mentioned in order to argue that the defined function is the function computed by the machine:
+%	\begin{itemize}
+%		\item[-] For each $\bar{b}$, only one of the functions $T_0$,$T_1$,$T_B$,$T_{\vdash}$ outputs 1 on input ($\bar{a}, \bar{b})$
+%		\item[-] For each $\bar{b}$, only one of the functions $h,\hat{h}$ outputs 1 on input $(\bar{a}, \bar{b})$
+%		\item[-] There is exactly one $\bar{b}$ with $h(\bar{a}, \bar{b}) = 1$
+%		\item[–] There is exatcly one $i$ such that $s_{q_i}(\bar{a}) = 1$
+%	\end{itemize}
+%	(It could also generally be argued that the functions are 0/1-functions encoding the intended predicates and are 0 when the timestep is the successor of the current timestep)
+%	\item p.36: ``is defined as an extension of QFO with the formula ...'' $\to$ ``is defined as the extension of QFO by the additional operator ...'' (or similar)
 \end{itemize}
@@
 	WL to the natural numbers as fixed semiring, calling this restriction ...''
-	\item[$\checkmark$] p.3, first paragraph of 2.1: ``omit the word finite or ordered'' $\to$ ``omit the words finite and
-	ordered''
+%	\item[$\checkmark$] p.3, first paragraph of 2.1: ``omit the word finite or ordered'' $\to$ ``omit the words finite and ordered''
 	\item p5, description of syntax of QSO: ``This division between Boolean and quantitative level
@@
 	reformulate this?
-	\item[$\checkmark$] p.5, next sentence: ``this will allow us later'' $\to$ ``this will later allow us''
-	\item[$\checkmark$] p.6, second to last sentence: ``similarity of $\alpha_3$ with'': ``similarity to'' is much more common
+%	\item[$\checkmark$] p.5, next sentence: ``this will allow us later'' $\to$ ``this will later allow us''
+%	\item[$\checkmark$] p.6, second to last sentence: ``similarity of $\alpha_3$ with'': ``similarity to'' is much more common
 	than ``similarity with''
-	\item[$\checkmark$] p.7, next sentence: ''in this direction'' $\to$ ``in this regard''
-	\item[$\checkmark$] p.7, last sentence before 3.1: ''connection of QSO with'' $\to$ ``connection of QSO to''
-	\item[$\checkmark$] p.7, middle: ``Another difference of WL with our approach'' $\to$ ``Another difference between
+%	\item[$\checkmark$] p.7, next sentence: ''in this direction'' $\to$ ``in this regard''
+%	\item[$\checkmark$] p.7, last sentence before 3.1: ''connection of QSO with'' $\to$ ``connection of QSO to''
+%	\item[$\checkmark$] p.7, middle: ``Another difference of WL with our approach'' $\to$ ``Another difference between
 	WL and our approach''
 	\item p.9, first sentence: ``then condition (2) holds'': omit the word then or reformulate the sentence
-	\item[$\checkmark$] p.9, first paragraph: ``the constant function $s$ can be trivially simulated'' $\to$ ``any constant
-	function $s$ can be trivially simulated''
-	\item[$\checkmark$] p.9, proof of Prop. 4.3: ``To prove the condition (2)'' $\to$ ``To prove condition (2)''
-	\item throughout the paper: ``similar than'' $\to$ ``similar to'' / ``similar than for X, we do Y'' $\to$
-	``similarly as for''. Examples:
-	\begin{itemize}
-		\item[-] p.10, proof of Thm. 4.5: ``the proof is similar than in Theorem 4.4'' $\to$ ``the proof is
-		similar to the proof of Theorem 4.4''	
-		\item[-] p.10, second sentence: ``similar than the previous proof'' $\to$ ``similar as in the previous
-		proof''
-		\item[-] p.11, last sentence before example 4.7: ``Similar than for QSO'' $\to$ ``Similarly as for QSO''
-	\end{itemize}
-	\item p.11, end of proof of Thm. 4.5: ``Using an analogous argument'': state what the argument is
-	analogous to (the argument in Thm. 4.4)
-	\item[$\checkmark$] p.12, after Corollary 4.8: ``shows how robust and versatile is QSO'' $\to$ ``shows how robust and
-	versatile QSO is''
-	\item p.12, after Corollary 4.8: ``counting complexity classes [...] even beyond $\nat$'' - make more precise
-	what ``beyond $\nat$'' refers to
-	\item[$\checkmark$] p.12, ``given CNF propositional formula, that can be made true'': no comma
+%	\item[$\checkmark$] p.9, first paragraph: ``the constant function $s$ can be trivially simulated'' $\to$ ``any constant function $s$ can be trivially simulated''
+%	\item[$\checkmark$] p.9, proof of Prop. 4.3: ``To prove the condition (2)'' $\to$ ``To prove condition (2)''
+%	\item throughout the paper: ``similar than'' $\to$ ``similar to'' / ``similar than for X, we do Y'' $\to$ ``similarly as for''. Examples:
+%	\begin{itemize}
+%		\item[-] p.10, proof of Thm. 4.5: ``the proof is similar than in Theorem 4.4'' $\to$ ``the proof is
+%		similar to the proof of Theorem 4.4''	
+%		\item[-] p.10, second sentence: ``similar than the previous proof'' $\to$ ``similar as in the previous
+%		proof''
+%		\item[-] p.11, last sentence before example 4.7: ``Similar than for QSO'' $\to$ ``Similarly as for QSO''
+%	\end{itemize}
+	\item p.11, end of proof of Thm. 4.5: ``Using an analogous argument'': state what the argument is analogous to (the argument in Thm. 4.4)
+%	\item[$\checkmark$] p.12, after Corollary 4.8: ``shows how robust and versatile is QSO'' $\to$ ``shows how robust and versatile QSO is''
+	\item p.12, after Corollary 4.8: ``counting complexity classes [...] even beyond $\nat$'' - make more precise what ``beyond $\nat$'' refers to
+%	\item[$\checkmark$] p.12, ``given CNF propositional formula, that can be made true'': no comma
 	\item p.12: ``in [43] they defined'' $\to$ something like ``$<$authors$>$ [43] defined''
-	\item[$\checkmark$] p.12: ``we extend as follows QSO'' $\to$ ``we extend QSO by max and min quantifiers as follows''
-	\item[$\checkmark$] p.12: ``MinQSO is defined analogously changing max with min'' $\to$ ``replacing max by min''
-	\item[$\checkmark$] p.12: ``the same holds with MinQSO(FO)'' $\to$ ``the same holds for ...''
-	\item[$\checkmark$] p.13: ``in contrast with'' $\to$ ``in contrast to''
-	\item[$\checkmark$] p.13: ``follow the same approach but considering'' $\to$ ``but consider''
-	\item[$\checkmark$] p.13: ``in the search of efficient approximation algorithms'' $\to$ ``in the search for''
-	\item[$\checkmark$] p.14: ``the search of robust classes'' $\to$ ''the search for robust classes''
+%	\item[$\checkmark$] p.12: ``we extend as follows QSO'' $\to$ ``we extend QSO by max and min quantifiers as follows''
+%	\item[$\checkmark$] p.12: ``MinQSO is defined analogously changing max with min'' $\to$ ``replacing max by min''
+%	\item[$\checkmark$] p.12: ``the same holds with MinQSO(FO)'' $\to$ ``the same holds for ...''
+%	\item[$\checkmark$] p.13: ``in contrast with'' $\to$ ``in contrast to''
+%	\item[$\checkmark$] p.13: ``follow the same approach but considering'' $\to$ ``but consider''
+%	\item[$\checkmark$] p.13: ``in the search of efficient approximation algorithms'' $\to$ ``in the search for''
+%	\item[$\checkmark$] p.14: ``the search of robust classes'' $\to$ ''the search for robust classes''
 	\item p.14: ''the first desirable condition has to do with the closure properties satisfied by the class'':
 	Reformulate this sentence
-	\item[$\checkmark$] p.14: ``As in the cases of P and NP that are closed'' $\to$ ``Analogously to the cases of P and
-	NP, which are closed under intersection and union, ...''
-	\item[$\checkmark$] p.16: ``such that for both, $A\mod\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$'' $\to$ ''such that for $i \in \{1,2\}$: ... ''
-	\item[$\checkmark$] p.16: ``Similarly to Theorem 5.1, we show ..'' $\to$ ``Similarly as in the proof of Theorem 5.1, we
+%	\item[$\checkmark$] p.14: ``As in the cases of P and NP that are closed'' $\to$ ``Analogously to the cases of P and NP, which are closed under intersection and union, ...''
+%	\item[$\checkmark$] p.16: ``such that for both, $A\mod\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$'' $\to$ ''such that for $i \in \{1,2\}$: ... ''
+%	\item[$\checkmark$] p.16: ``Similarly to Theorem 5.1, we show ..'' $\to$ ``Similarly as in the proof of Theorem 5.1, we
 	show''
@@
 	closure properties'': Should be reformulated
-	\item[$\checkmark$] p.19: ``Let ... be this formula, where $\varphi$ is in first-order and quantifier-free'' $\to$ ``... where $\varphi$ is
-	first-order and quantifier-free''
-	\item[$\checkmark$] several occurrences: ``This is, ...'' $\to$ ``That is, ...'', for example two times on p.20
-	\item[$\checkmark$] p.20: ``we focus only in proving'' $\to$ ``we focus only on proving''
-	\item[$\checkmark$] p.20: ``We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of FO.''
-	$\to$ ``We prove this for the more general case of $\eqso(\LL)$ with $\LL$ being a fragment of SO.''
-	(or similar)
-
-	\item[$\checkmark$] p.20: ``we will show a formula'' $\to$ ``give'' or ``construct'' instead of ``show''
-	\item[$\checkmark$] p.20/21: ``$\shp$ is believed not to be closed under subtraction by one by some complexitytheoretic
-	assumption'' $\to$ ``... due to some complexity-theoretic assumption'' (or similar)
-	\item[$\checkmark$] p.21, next sentence: ``for a possible extension'' $\to$ ``as a possible extension''
-	\item p.21: ``the desired extension has to be achieved by ... . More precisely, we define...'': The second
-	sentence doesn’t make the statement of the first one more precise, but rather explains the
-	actual definiton (which is chosen in accordance with the sentence before)
-	\item[$\checkmark$] p.21: ``a parsimonious reduction from a function'' $\to$ ``a parsimonious reduction from any
-	function''
-	\item[$\checkmark$] p.21: ``we define a function that converts a formula $\alpha$ in...'' $\to$ ``that converts a given formula
-	$\alpha$...''
-	\item[$\checkmark$] p.22: ``we iterate for every tuple'' $\to$ ``we iterate over all tuples''
-	\item[$\checkmark$] p.22: ``Then we will show how to define a formula $\varphi'$
-	...'' $\to$ ``We will show how to define ...''
-	\item[$\checkmark$] p.22: ''no variable from $\bar{x}$ are mentioned'' $\to$ ``... is mentioned''
-	\item[$\checkmark$] p.24: ``if, and only if, if X, then Y'': This construction makes it unneccesarily hard to parse
-	the sentence
-	\item[$\checkmark$] p.24: ``Then, the conclusion of the implication...'' $\to$ ``The conclusion of the implication...''
-	\item[$\checkmark$] p.25: ``For the sake of simplification...'' $\to$ ``For the sake of simplicity...''
-	\item[$\checkmark$] p.26: ``In the previous section, we show ...'' $\to$ ``In the previous section we showed ...''
-	\item[$\checkmark$] p.27: ``we postpone the proof the previous proposition after ...'' $\to$ ``we postpone the proof of
-	the previous proposition to after ...''
-	\item p.29: ``replace by $\perp$ and $\top$ \underline{wherever it corresponds}'': Should be reformulated
-	\item[$\checkmark$] p.30: ``consist in computing $g_{\alpha}(\enc(\A))$ for each input ...'' $\to$ ``... on input ...''
-	\item[$\checkmark$] p.30, same sentence: ``and then simulate'' $\to$ ``and then simulating''
-	\item[$\checkmark$] p.33: ``In contrast with LFP'' $\to$ ``In contrast to LFP''
-	\item[$\checkmark$] p.33: ``These operations are represented by the set $\{0,1,\emptyset\}$, respectively.'' $\to$ ``... by 0, 1 and
-	$\emptyset$, respectively''
-	\item[$\checkmark$] p.34: ``Notice that the values of each function $f$ in $\alpha$'' $\to$ ``Notice that the values of each
-	function in $\alpha$''
-	\item[$\checkmark$] p.34: ``that is why we talk about ...'' $\to$ ``which is why we talk about''
-	\item[$\checkmark$] p.34: ``Besides, notice that notation ...'' $\to$ ``Besides, notice that the notation ...''
-	\item[$\checkmark$] p.34: ``Function $h$ is used to indicate whether the head of the working tape is in some position
-	at some point of time'' $\to$ ``... whether the head of the working tape is in a specific position
-	at a certain point of time'' (or similar)
-	\item[$\checkmark$] p.36: ``Given a relation signature $\R$'' $\to$ ``... relational signature ...''
-	\item[$\checkmark$] p.37: ``As it was mentioned before'' $\to$ ``As mentioned before''
-	\item[$\checkmark$] p.37: ``Next we construct a logarithmic-space ....'' $\to$ ``We construct a logarithmic-space ...''
-	6
-	\item[?] p.38: ``In this sense, there are several directions for future research'' $\to$ ``Consequently, there
-	are ...'' (or similar)
-	\item[$\checkmark$] p.38: ``In the same direction, ...'' $\to$ ``Similarly, ...'' (or similar)
-	\item[$\checkmark$] p.38: ``so to have'' $\to$ ``in order to have'' (or similar)
+%	\item[$\checkmark$] p.19: ``Let ... be this formula, where $\varphi$ is in first-order and quantifier-free'' $\to$ ``... where $\varphi$ is first-order and quantifier-free''
+%	\item[$\checkmark$] several occurrences: ``This is, ...'' $\to$ ``That is, ...'', for example two times on p.20
+%	\item[$\checkmark$] p.20: ``we focus only in proving'' $\to$ ``we focus only on proving''
+%	\item[$\checkmark$] p.20: ``We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of FO.''
+	$\to$ ``We prove this for the more general case of $\eqso(\LL)$ with $\LL$ being a fragment of SO.'' (or similar)
+%	\item[$\checkmark$] p.20: ``we will show a formula'' $\to$ ``give'' or ``construct'' instead of ``show''
+%	\item[$\checkmark$] p.20/21: ``$\shp$ is believed not to be closed under subtraction by one by some complexitytheoretic assumption'' $\to$ ``...
+due to some complexity-theoretic assumption'' (or similar)
+%	\item[$\checkmark$] p.21, next sentence: ``for a possible extension'' $\to$ ``as a possible extension''
+	\item p.21: ``the desired extension has to be achieved by ...
+.
+More precisely, we define...'': The second sentence doesn’t make the statement of the first one more precise, but rather explains the actual definiton (which is chosen in accordance with the sentence before)
+%	\item[$\checkmark$] p.21: ``a parsimonious reduction from a function'' $\to$ ``a parsimonious reduction from any function''
+%	\item[$\checkmark$] p.21: ``we define a function that converts a formula $\alpha$ in...'' $\to$ ``that converts a given formula $\alpha$...''
+%	\item[$\checkmark$] p.22: ``we iterate for every tuple'' $\to$ ``we iterate over all tuples''
+%	\item[$\checkmark$] p.22: ``Then we will show how to define a formula $\varphi'$ ...'' $\to$ ``We will show how to define ...''
+%	\item[$\checkmark$] p.22: ''no variable from $\bar{x}$ are mentioned'' $\to$ ``... is mentioned''
+%	\item[$\checkmark$] p.24: ``if, and only if, if X, then Y'': This construction makes it unneccesarily hard to parse the sentence
+%	\item[$\checkmark$] p.24: ``Then, the conclusion of the implication...'' $\to$ ``The conclusion of the implication...''
+%	\item[$\checkmark$] p.25: ``For the sake of simplification...'' $\to$ ``For the sake of simplicity...''
+%	\item[$\checkmark$] p.26: ``In the previous section, we show ...'' $\to$ ``In the previous section we showed ...''
+%	\item[$\checkmark$] p.27: ``we postpone the proof the previous proposition after ...'' $\to$ ``we postpone the proof of the previous proposition to after ...''
+%	\item p.29: ``replace by $\perp$ and $\top$ \underline{wherever it corresponds}'': Should be reformulated
+%	\item[$\checkmark$] p.30: ``consist in computing $g_{\alpha}(\enc(\A))$ for each input ...'' $\to$ ``... on input ...''
+%	\item[$\checkmark$] p.30, same sentence: ``and then simulate'' $\to$ ``and then simulating''
+%	\item[$\checkmark$] p.33: ``In contrast with LFP'' $\to$ ``In contrast to LFP''
+%	\item[$\checkmark$] p.33: ``These operations are represented by the set $\{0,1,\emptyset\}$, respectively.'' $\to$ ``... by 0, 1 and $\emptyset$, respectively''
+%	\item[$\checkmark$] p.34: ``Notice that the values of each function $f$ in $\alpha$'' $\to$ ``Notice that the values of each function in $\alpha$''
+%	\item[$\checkmark$] p.34: ``that is why we talk about ...'' $\to$ ``which is why we talk about''
+%	\item[$\checkmark$] p.34: ``Besides, notice that notation ...'' $\to$ ``Besides, notice that the notation ...''
+%	\item[$\checkmark$] p.34: ``Function $h$ is used to indicate whether the head of the working tape is in some position at some point of time'' $\to$ ``...
+whether the head of the working tape is in a specific position at a certain point of time'' (or similar)
+%	\item[$\checkmark$] p.36: ``Given a relation signature $\R$'' $\to$ ``... relational signature ...''
+%	\item[$\checkmark$] p.37: ``As it was mentioned before'' $\to$ ``As mentioned before''
+%	\item[$\checkmark$] p.37: ``Next we construct a logarithmic-space ....'' $\to$ ``We construct a logarithmic-space ...''
+	\item[?] p.38: ``In this sense, there are several directions for future research'' $\to$ ``Consequently, there are ...'' (or similar)
+%	\item[$\checkmark$] p.38: ``In the same direction, ...'' $\to$ ``Similarly, ...'' (or similar)
+%	\item[$\checkmark$] p.38: ``so to have'' $\to$ ``in order to have'' (or similar)
 \end{itemize}
 \vspace{1em}
-\subsection*{Typos}
+%\subsection*{Typos}
+%\begin{itemize}
+%	\setlength\itemsep{0.5em}
+%	\item p.1, footnote: ``clases'' should be ``classes''
+%	\item throughout the paper: plural used with verb ending on s. Examples:
+%	\begin{itemize}
+%		\item[-] p.2, middle: ``fragments of QSO captures''
+%		\item[-] p.6, example 3.1: ``the sum quantifiers in $\alpha_1$ aggregates''
+%	\end{itemize}
+%	\item[$\checkmark$] p4, beginning of 2.2: ``some of \underline{the} are recalled here''
+%	\item[$\checkmark$] p6, example 3.2, first paragraph: ``otherwise $M(i.j)$'', dot used instead of comma
+%	\item[$\checkmark$] throughout the paper: a instead of an (espacially: a FO formula/assignment, a SO formula/assignment).
+%	Example:
+%	\begin{itemize}
+%		\item[-] p.9, third sentence: ``Given a FO formula'' $\to$ ``Given an FO formula''
+%		\item[-] p.38: ``As a example'' $\to$ ``As an example''
+%	\end{itemize}
+%	\item throughout the paper: singular used with verb not ending on s. Examples:
+%	\begin{itemize}
+%		\item[-] p.9, after proof of Prop. 4.3: ``use of a fragment [...] that capture''
+%		\item[-] p.11, end of proof of Thm. 4.5: ``and $\alpha$ reconstruct''
+%	\end{itemize}
+%	\item[$\checkmark$] p.9, middle: ``A second attempt could be based then'' $\to$ ``A second attempt could then be
+%	based''
+%	\item[$\checkmark$] p.11, 4.1: ``output of a function \underline{in} not a natural number''
+%	\item[$\checkmark$] p.12: ``we have to dinstinguished''
+%	\item[$\checkmark$] p.14, beginning of 5.1: ``subclases''
+%	\item[$\checkmark$] p.21: ``it does not mentioned'' $\to$ ``it does not mention''
+%	\item[$\checkmark$] p.22: ``Without lost of generality'' $\to$ ``Without loss of generality''
+%	\item[$\checkmark$] p.24: ``The previous claim and proof motivates the following definitions.'' $\to$ ``The previous
+%	claim and proof motivate the following definition.''
+%	\item p.30: ``to define a natural generalization for functions of the notion of least fixed point of
+%	LFP'' $\to$ ``to define a natural generalization of LFP to functions'' (or similar)
+%	\item[$\checkmark$] p.30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
+%	\item[$\checkmark$] p.31: ``we simply generalized these conditions'' $\to$ ``we simply generalize these conditions''
+%	\item[$\checkmark$] p.32: ``which is used to defined'' $\to$ ``which is used to define''
+%	\item[$\checkmark$] p.33: ``since $\beta$ is constructed from monotones operations'' $\to$ ``... monotone ...''
+%	\item[$\checkmark$] p.35: ``Formulas $\alpha_h$ is defined ...'' $\to$ ``Formula $\alpha_h$ is defined ...''
+%	7
+%	\item[$\checkmark$] p.37: ``we simulate $M_\beta$ on input $v[1/n]$'' $\to$ ``... with input $v[1/x]$''
+%\end{itemize}
+\vspace{1em}
 \begin{itemize}
 	\setlength\itemsep{0.5em}
-	\item p.1, footnote: ``clases'' should be ``classes''
-	\item throughout the paper: plural used with verb ending on s. Examples:
-	\begin{itemize}
-		\item[-] p.2, middle: ``fragments of QSO captures''
-		\item[-] p.6, example 3.1: ``the sum quantifiers in $\alpha_1$ aggregates''
-	\end{itemize}
-	\item[$\checkmark$] p4, beginning of 2.2: ``some of \underline{the} are recalled here''
-	\item[$\checkmark$] p6, example 3.2, first paragraph: ``otherwise $M(i.j)$'', dot used instead of comma
-	\item[$\checkmark$] throughout the paper: a instead of an (espacially: a FO formula/assignment, a SO formula/assignment).
-	Example:
-	\begin{itemize}
-		\item[-] p.9, third sentence: ``Given a FO formula'' $\to$ ``Given an FO formula''
-		\item[-] p.38: ``As a example'' $\to$ ``As an example''
-	\end{itemize}
-	\item throughout the paper: singular used with verb not ending on s. Examples:
-	\begin{itemize}
-		\item[-] p.9, after proof of Prop. 4.3: ``use of a fragment [...] that capture''
-		\item[-] p.11, end of proof of Thm. 4.5: ``and $\alpha$ reconstruct''
-	\end{itemize}
-	\item[$\checkmark$] p.9, middle: ``A second attempt could be based then'' $\to$ ``A second attempt could then be
-	based''
-	\item[$\checkmark$] p.11, 4.1: ``output of a function \underline{in} not a natural number''
-	\item[$\checkmark$] p.12: ``we have to dinstinguished''
-	\item[$\checkmark$] p.14, beginning of 5.1: ``subclases''
-	\item[$\checkmark$] p.21: ``it does not mentioned'' $\to$ ``it does not mention''
-	\item[$\checkmark$] p.22: ``Without lost of generality'' $\to$ ``Without loss of generality''
-	\item[$\checkmark$] p.24: ``The previous claim and proof motivates the following definitions.'' $\to$ ``The previous
-	claim and proof motivate the following definition.''
-	\item p.30: ``to define a natural generalization for functions of the notion of least fixed point of
-	LFP'' $\to$ ``to define a natural generalization of LFP to functions'' (or similar)
-	\item[$\checkmark$] p.30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
-	\item[$\checkmark$] p.31: ``we simply generalized these conditions'' $\to$ ``we simply generalize these conditions''
-	\item[$\checkmark$] p.32: ``which is used to defined'' $\to$ ``which is used to define''
-	\item[$\checkmark$] p.33: ``since $\beta$ is constructed from monotones operations'' $\to$ ``... monotone ...''
-	\item[$\checkmark$] p.35: ``Formulas $\alpha_h$ is defined ...'' $\to$ ``Formula $\alpha_h$ is defined ...''
-	7
-	\item[$\checkmark$] p.37: ``we simulate $M_\beta$ on input $v[1/n]$'' $\to$ ``... with input $v[1/x]$''
-\end{itemize}
-\vspace{1em}
-{\bf Other}
-\begin{itemize}
-	\setlength\itemsep{0.5em}
-	\item[$\checkmark$] running title is ``COUNTING''
-	\item on definition of functions, \verb|\colon| should be used instead of :
-	\item In [1], a similar hierarchy to the one by Saluja et al. was introduced and studied. This work
-	should be cited.
+%	\item[$\checkmark$] running title is ``COUNTING''
+%	\item on definition of functions, \verb|\colon| should be used instead of :
+	\item In [1], a similar hierarchy to the one by Saluja et al. was introduced and studied. This work should be cited.
 \end{itemize}
@@
 
-	\item[$\checkmark$] page 4:
-	-some of the $\to$ some of them 
-	\item[$\checkmark$] page 10: 
-	-$2^{|A|^l}$-bits $\to$ $2^{|A|^l}$ bits
-	\item[$\checkmark$] page 21:
-	-it does not mentioned $\to$ it does not mention
-	\item page 22: 
-	-you make a long proof about some closure under subtraction by one.
-	You should put the result in a theorem envirpnment to make what you are
-	proving more evident.
-	\item[$\checkmark$] -replace .... for $\to$ replace ... by;  there are several instances of
-	that
+%	\item[$\checkmark$] page 4: -some of the $\to$ some of them 
+%	\item[$\checkmark$] page 10:  -$2^{|A|^l}$-bits $\to$ $2^{|A|^l}$ bits
+%	\item[$\checkmark$] page 21: -it does not mentioned $\to$ it does not mention
+%	\item page 22: -you make a long proof about some closure under subtraction by one. You should put the result in a theorem envirpnment to make what you are proving more evident.
+%	\item[$\checkmark$] -replace .... for $\to$ replace ... by;  there are several instances of that
 	\item[$\checkmark$] -$\phi$ is an FO-formula $\to$ a $\Sigma_1[FO]$ formula
@@
 	\item[?] page 30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
-
-	\item[$\checkmark$] page 32: 
-	- paths of length at most n $\to$ in an acylic graph, all paths are of size less than n
+%	\item[$\checkmark$] page 32: - paths of length at most n $\to$ in an acylic graph, all paths are of size less than n
 	\item page 33:

### 2019-01-21 LMCS/sendrev/beyond.tex
correcciones en la carpeta sendrev
reviewer 1 - technical casi listo

--- 2018-12-10 LMCS/beyond.tex
@@
 
-Given a relational signature $\R$, the set of transitive $\qfo$ formulae ($\tqfo$-formulae) is defined as an extension of $\qfo$ with the formula $[\pth \psi(\bar{x},\bar{y})]$, where $\psi(\x, \y)$ is an $\fo$-formula over $\R$, and $\bar{x} = (x_1, \ldots, x_k)$, $\bar{y} = (y_1, \ldots, y_k)$ are tuples of pairwise distinct first-order variables.
+Given a relational signature $\R$, the set of transitive $\qfo$ formulae ($\tqfo$-formulae) is defined as the extension of $\qfo$ by the additional operator $[\pth \psi(\bar{x},\bar{y})]$, where $\psi(\x, \y)$ is an $\fo$-formula over $\R$, and $\bar{x} = (x_1, \ldots, x_k)$, $\bar{y} = (y_1, \ldots, y_k)$ are tuples of pairwise distinct first-order variables.
 The semantics of $[\pth \psi(\bar{x},\bar{y})]$ can easily be defined in terms of $\rqfo(\fo)$ as follows.

### 2019-01-21 LMCS/sendrev/framework.tex
reviewer 1 - technical casi listo

@@
 $\Sigma x$ or $\Pi X$) are called {\em quantitative quantifiers}.
-Furthermore, $\Sigma x$ and $\Sigma X$ are called first- and second-order sum, and $\Pi x$ and $\Pi X$ are called first- and second-order product, respectively.
+Furthermore, $\Sigma x$ and $\Sigma X$ are called first- and second-order sum, whereas $\Pi x$ and $\Pi X$ are called first- and second-order product, respectively.
 This division between Boolean and quantitative level is essential for understanding the difference between the logic and the quantitative part. 

### 2019-01-21 LMCS/sendrev/proof-ehorn-pe.tex
correcciones en la carpeta sendrev
reviewer 1 - technical casi listo

--- 2018-12-09 LMCS/proof-ehorn-pe.tex
@@
 %Let $\shdhsat$ be the problem of counting the number of satisfying assignments of a disjunction of Horn formulas. 
-In~\cite{PagourtzisZ06}, Pagourtzis and Zachos gave a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula.
+%In~\cite{PagourtzisZ06}, Pagourtzis and Zachos gave a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula.
 This procedure can be easily extended to receive Horn formulas, and furthermore, a disjunction of Horn formulas.
 Hence $\shdhsat$ is in $\totp$.
-As we saw in Theorem~\ref{sigma2hard}, $\shdhsat$ is complete for $\eqso(\ehorn)$ under parsimonious reductions.
+First we will prove that $\shdhsat$ is in $\totp$.
+For this consider a non-deterministic procedure that receives a $\dhsat$ formula $\Phi$ as input and does the following.
+First check if $\Phi$ is satisfiable.
+If it is not, stop, and if it is, create a dummy branch that simply stops.
+Back in the main branch pick a variable $x$ in $\Phi$ and create two formulas $\Phi_0$ and $\Phi_1$ where $x$ has been replaced by $\perp$ or $\top$ respectively.
+If only one of these is satisfiable, continue on this branch with the respective $\Phi_i$, and if both are, create a new branch for $\Phi_1$ and continue on this branch with $\Phi_0$.
+On each branch repeat the same instructions until no variables are left to replace.
+Since $\dhsat$ is in $\ptime$ each formula check can be done in $\ptime$, so the procedure takes at most $h(\vert\Phi\vert)\vert\Phi\vert$ steps, for some polynomial function $h$, and it produces exactly $\shdhsat(\Phi)+1$ branches.
+Hence $\shdhsat$ is in $\totp$.
+For the next part of the proof, notice that in Theorem~\ref{sigma2hard} we proved that $\shdhsat$ is complete for $\eqso(\ehorn)$ under parsimonious reductions.
 Let $\alpha$ be a formula in $\eqso(\ehorn)$ and let $g_{\alpha}$ be the reduction to $\shdhsat$.
@@
 Therefore, we conclude that $\alpha$ defines a function in~$\totp$.
+\martin{agregu\'e una demostraci\'on de que $\shdhsat$ esta en $\totp$}

### 2019-01-21 LMCS/sendrev/proof-hsat-not-sigma2.tex
mas revisiones
reviewer 1 - technical casi listo

--- 2017-08-07 LMCS/proof-hsat-not-sigma2.tex
@@
 %We use a similar proof to the one provided by the authors in \cite{SalujaST95} to separate the classes $\E{2}$ and $\U{2}$. 
-Suppose that the statement is false, this is, $\chsat \in \eqso(\loge{2})$.
+Suppose that the statement is false, that is, $\chsat \in \eqso(\loge{2})$.
 Consider the signature $\R$ from Example~\ref{ex-hornsat-esop1} and let $\alpha \in \eqso(\loge{2})$ be a formula over $\R$ that defines $\chsat$.
@@
 Consider the induced substructure $\A_{\Phi}'$ that is obtained by removing $t_{\ell}$ from $\A_{\Phi}$ and $\bar{B}'$ as the subset of $\bar{B}$ obtained by deleting each appearance of $t_{\ell}$ in $\bar{B}$.
-We have that $\A_{\Phi}'\models \fa{\bar{z}}\varphi(\bar{B}',\bar{b},\bar{a},\bar{z})$ since universal formulas are monotone over induced substructures.
-Then it follows that $\sem{\alpha}(\A_{\Phi}') \geq 1$ which is not possible since $\A_{\Phi}'$ encodes the formula
+Given that if a universal formula holds in a structure $\A$, then it holds in every induced substructure of $\A$, we have that $\A_{\Phi}'\models \fa{\bar{z}}\varphi(\bar{B}',\bar{b},\bar{a},\bar{z})$.
+And so, it follows that $\sem{\alpha}(\A_{\Phi}') \geq 1$ which is not possible since $\A_{\Phi}'$ encodes the formula
 $$

### 2019-01-21 LMCS/sendrev/proof-rqfo-fo-cap.tex
reviewer 1 - technical casi listo

@@
 These operations are represented by 0, 1, and $\emptyset$, respectively.
-Finally, assume that $M$, on input $\enc(\A)$ with domain $A = \{1,\dots,n\}$, executes exactly $n^k$ steps for a fixed $k \geq 1$.
+Finally, assume that $M$, on input $\enc(\A)$ with domain $A = \{1,\dots,n\}$, executes exactly $n^k$ steps on large inputs for a fixed $k \geq 1$.
+We ignore small inputs since they can be handled separately.
+ \martin{tom\'e esta correcci\'on}
 
@@
 \sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult \hat{h}(\bar{t}',\bar{x}) \mult T_{\vdash}(\bar{t}',\bar{x})) \ + \\
-\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,\vdash,\_,\_)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')).
+\sum_{(q,a) \,:\, \delta(q,a) = (\_,\vdash,\_,\_)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')).
 \end{multline*}
-Notation $\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,\vdash,\_,\_)}$ in this formula is used to indicate that we are considering a sum over all pairs $(q,a) \in Q \times \{0,1,\ttB,\vdash\}$ such that $\delta(q,a) = (q',\vdash,u,v)$ for some $q' \in Q$, $u \in \{\leftarrow,\rightarrow\}$ and $v \in \{0,1,\emptyset\}$.
+Notation $\sum_{(q,a) \,:\, \delta(q,a) = (\_,\vdash,\_,\_)}$ in this formula is used to indicate that we are considering a sum over all pairs $(q,a) \in Q \times \{0,1,\ttB,\vdash\}$ such that $\delta(q,a) = (q',\vdash,u,v)$ for some $q' \in Q$, $u \in \{\leftarrow,\rightarrow\}$ and $v \in \{0,1,\emptyset\}$.
 Formulas $\alpha_{T_0}$, $\alpha_{T_1}$ and $\alpha_{T_{\ttB}}$ are defined analogously; for the sake of presentation, we only show here how $\alpha_{T_0}$ is defined, assuming that $\bar y = (y_1, \ldots, y_k)$:
@@
 \sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult \hat{h}(\bar{t}',\bar{x}) \mult T_0(\bar{t}',\bar{x})) + \\
-\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,0,\_,\_)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')).
+\sum_{(q,a) \,:\, \delta(q,a) = (\_,0,\_,\_)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')).
 %\alpha_{T_1}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\
@@
 %&\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult \hat{h}(\bar{t}',\bar{x}) \mult T_1(\bar{t}',\bar{x})) + \\
-%&\bigplus_{\delta(q,a) = (q',1,op,v)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')),\\
+%&\sum_{\delta(q,a) = (q',1,op,v)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')),\\
 %\alpha_{T_{\ttB}}(\bar{t},\bar{x},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\
@@
 %&\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult \hat{h}(\bar{t}',\bar{x}) \mult T_{\ttB}(\bar{t}',\bar{x})) + \\
-%&\bigplus_{\delta(q,a) = (q',{\ttB},op,v)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')),\\
+%&\sum_{\delta(q,a) = (q',{\ttB},op,v)}\sa{\bar{t}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}) \mult T_a(\bar{t}',\bar{x}) \mult s_{q}(\bar{t}')),\\
 \end{multline*}
@@
 &(\first(\bar{t}) \wedge \successor(\bar{t},\bar{x})) + \\
-&\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,\_,\leftarrow,\_)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult \successor(\bar{x},\bar{x}')\mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')) \ + \\
-&\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,\_,\rightarrow,\_)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult \successor(\bar{x}',\bar{x})\mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')).
+&\sum_{(q,a) \,:\, \delta(q,a) = (\_,\_,\leftarrow,\_)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult \successor(\bar{x},\bar{x}')\mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')) \ + \\
+&\sum_{(q,a) \,:\, \delta(q,a) = (\_,\_,\rightarrow,\_)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult \successor(\bar{x}',\bar{x})\mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')).
 \end{align*}
@@
 &(\first(\bar{t}) \wedge \neg\successor(\bar{t},\bar{x})) + \\
-&\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,\_,\leftarrow,\_)}\sa{\bar{t}'}\sa{\bar{x}'}\sa{\bar{x}''}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \, \mult \\
+&\sum_{(q,a) \,:\, \delta(q,a) = (\_,\_,\leftarrow,\_)}\sa{\bar{t}'}\sa{\bar{x}'}\sa{\bar{x}''}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \, \mult \\
 &\hspace{250pt} \successor(\bar{x}'',\bar{x}') \mult (\bar{x} \neq \bar{x}'')) \ + \\
-&\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,\_,\rightarrow,\_)}\sa{\bar{t}'}\sa{\bar{x}'}\sa{\bar{x}''}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \, \mult \\
+&\sum_{(q,a) \,:\, \delta(q,a) = (\_,\_,\rightarrow,\_)}\sa{\bar{t}'}\sa{\bar{x}'}\sa{\bar{x}''}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \, \mult \\
 &\hspace{250pt} \successor(\bar{x}',\bar{x}'') \mult (\bar{x} \neq \bar{x}'')).
@@
 Notice that in the definition of $\alpha_{\hat{h}}$, $\fo$-formula $\bar{x} \neq \bar{x}''$ is used to indicate that tuples $\bar{x}$ and $\bar{x}''$ are distinct (that is, there exists $i \in \{1, \ldots, k\}$ such that the $i$-th components of $\bar{x}$ and $\bar{x}''$ are distinct).
-Formula $\alpha_{q_0}$ is defined as:
+Formula $\alpha_{s_{q_0}}$ is defined as:
 \begin{multline*}
-\alpha_{q_0}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out) \ = \\ 
+\alpha_{s_{q_0}}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out) \ = \\ 
 \first(\bar{t}) \ + \
-\bigplus_{(q,a) \,:\, \delta(q,a) = (q_0,\_,\_,\_)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult  s_{q}(\bar{t}')).
+\sum_{(q,a) \,:\, \delta(q,a) = (q_0,\_,\_,\_)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult  s_{q}(\bar{t}')).
 \end{multline*}
-Moreover, for every $i \in \{1, \ldots, \ell\}$, formula $\alpha_{q_i}$ is defined analogously.
+Moreover, for every $i \in \{1, \ldots, \ell\}$, formula $\alpha_{s_{q_i}}$ is defined analogously.
 %\begin{align*}
-%\alpha_{q_i}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\ &\bigplus_{\delta(q,a) = (q_i,b,op,v)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult T_a(\bar{t}',\bar{x}') \mult h(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')).
+%\alpha_{q_i}(\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},&h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},out) = \\ &\sum_{\delta(q,a) = (q_i,b,op,v)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult T_a(\bar{t}',\bar{x}') \mult h(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}')).
 %\end{align*}
@@
 \alpha_{\out}(&\bar{t},T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h},s_{q_0},\ldots,s_{q_{\ell}},\out) =\\
-&\bigplus_{(q,a) \,:\, \delta(q,a) = (\_,\_,\_,0)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult 2\mult \out(\bar{t}')) \ + \\
-&\bigplus_{(q,a) \,:\,\delta(q,a) = (\_,\_,\_,1)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult (2\mult \out(\bar{t}')+1)) \ + \\
-&\bigplus_{(q,a) \,:\,\delta(q,a) = (\_,\_,\_,\emptyset)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult \out(\bar{t}')).
+&\sum_{(q,a) \,:\, \delta(q,a) = (\_,\_,\_,0)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult 2\mult \out(\bar{t}')) \ + \\
+&\sum_{(q,a) \,:\,\delta(q,a) = (\_,\_,\_,1)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult (2\mult \out(\bar{t}')+1)) \ + \\
+&\sum_{(q,a) \,:\,\delta(q,a) = (\_,\_,\_,\emptyset)}\sa{\bar{t}'}\sa{\bar{x}'}(\successor(\bar{t}',\bar{t}) \mult h(\bar{t}',\bar{x}') \mult T_a(\bar{t}',\bar{x}') \mult s_{q}(\bar{t}') \mult \out(\bar{t}')).
 \end{align*}
-Clearly, at each iteration of the fixed point operator, the tuple $\bar{t}$ represents the step the machine is currently in.
+%Clearly, at each iteration of the fixed point operator, the tuple $\bar{t}$ represents the step the machine is currently in. 
+Clearly, tuple $\bar{t}$ encodes the number of steps the machine has done and in each iteration of the fixed point operator, one timestep of the machine is executed.
 Assume that $\bar{a}$, $\bar{a}'$ are tuples with $k$ elements such that $\bar{a}'$ is the successor of $\bar{a}$. 
 From the construction of the formula, and since the machine is deterministic, it can be seen that for each function $f \in\{T_0,T_1,T_{\ttB},T_{\vdash},h,\hat{h}\}$, at the $\bar{a}$-th iteration of the fixed point operator, it holds that $f(\bar{a},\bar{b}) \leq 1$ and  $f(\bar{a}',\bar{b}) = 0$ for every $\bar{b}\in A^k$.
-In the same way, it can be seen that for each function $g \in\{s_{q_1},\ldots,s_{q_{\ell}}\}$, at the $\bar{a}$-th iteration of the fixed point operator, it holds that $g(\bar{a}) \leq 1$ and $g(\bar{a}') = 0$.
+In the same way, it can be seen that (1) for each function $g \in\{s_{q_1},\ldots,s_{q_{\ell}}\}$, at the $\bar{a}$-th iteration of the fixed point operator, it holds that $g(\bar{a}) \leq 1$ and $g(\bar{a}') = 0$, that (2) for each $\bar{b}$, only one of the functions $T_0$,$T_1$,$T_B$,$T_{\vdash}$ outputs 1 on input ($\bar{a}, \bar{b})$, (3) for each $\bar{b}$, only one of the functions $h,\hat{h}$ outputs 1 on input $(\bar{a}, \bar{b})$, (4) there is exactly one $\bar{b}$ with $h(\bar{a}, \bar{b}) = 1$, and (5) there is exactly one $i$ such that $s_{q_i}(\bar{a}) = 1$.
 From this, we have that at the iteration $\bar{a}'$ of the fixed point operator, $\out(\bar{a}')$ is equal to either $2\cdot \out(\bar{a})$, $2\cdot \out(\bar{a}) + 1$, or $\out(\bar{a})$, which represents precisely the value in the output tape at each step of $M$ running on the input $\enc(\A)$.
 From this argument, it can be seen that $\sem{\alpha}(\A) = f(\enc(\A))$, which concludes the proof of the theorem.
+\martin{agregue la correccion verbatim}
 %

### 2019-01-21 LMCS/sendrev/proof-sigma2hard.tex
correcciones en la carpeta sendrev
reviewer 1 - technical casi listo

--- 2017-09-27 LMCS/proof-sigma2hard.tex
@@
 First we prove that $\shdhsat$ is in $\eqso(\ehorn)$.
-Recall that each instance of $\shdhsat$ is a disjunction of Horn formulas.
+Recall that each instance of $\shdhsat$ is a disjunction of Horn formulae.
 Let $\R$ be a relational signature such that $\R = \{\pP(\cdot,\cdot), \pN(\cdot,\cdot), \pV(\cdot), \pNC(\cdot), \pD(\cdot,\cdot)\}$.
@@
 
-Next, we prove that $\shdhsat$ is hard for $\eqso(\ehorn)$ over a signature~$\R$ under parsimonious reductions.
+Next, we prove that $\shdhsat$ is hard for $\eqso(\ehorn)$ over any signature~$\R$ under parsimonious reductions.
 For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a polynomial-time function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
@@
 $$
-where each $\varphi^i_j$ is a Horn-clause, and each $\bar{X}_{i}$ is a sequence of second-order variables.
-Consider $\bar{X}$ as the union of all $\bar{X}_{i}$. We replace each of the $m$ addends in $\alpha$ for
+where each $\varphi^i_j$ is a Horn clause, and each $\bar{X}_{i}$ is a sequence of second-order variables.
+Consider $\bar{X}$ as the union of all $\bar{X}_{i}$. We replace each of the $m$ summands in $\alpha$ with
 $$
-\sa{\bar{X}}\sa{\bar{x}} \ex{\bar{y}} \bigwedge_{j = 1}^{n} \fa{\bar{z}} \varphi^i_j(\bar{X}_{i},\bar{x},\bar{y},\bar{z})\wedge\bigwedge_{X\not\in\bar{X}_{i}}\fa{\bar{u}}X(\bar{u}),
+\sa{\bar{X}}\sa{\bar{x}} \ex{\bar{y}} \bigwedge_{j = 1}^{n} \fa{\bar{z}} \varphi^i_j(\bar{X}_{i},\bar{x},\bar{y},\bar{z})\wedge\bigwedge_{X\not\in\bar{X}_{i}}\!\!\fa{\bar{u}}X(\bar{u}),
 $$
@@
 $$
-\alpha_{\A} \ = \ \sum_{i = 1}^{m} \sum_{\bar{a}\,\in A^{\length{\bar{x}}}} \sa{\bar{X}}\bigvee_{\bar{b}\,\in A^{\length{\bar{y}}}}\bigwedge_{j = 1}^{n'}\bigwedge_{\bar{c}\,\in A^{\length{\bar{z}}}}\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c}).
+\alpha_{\A} \ = \ \sum_{i = 1}^{m} \sum_{\bar{a}\,\in A^{\length{\bar{x}}}} \sa{\bar{X}}\bigvee_{\bar{b}\,\in A^{\length{\bar{y}}}}\bigwedge_{j = 1}^n\bigwedge_{\bar{c}\,\in A^{\length{\bar{z}}}}\varphi^i_j(\bar{X}_i,\bar{a},\bar{b},\bar{c})\wedge \bigwedge_{X \not\in \bar{X}_i}\bigwedge_{\bar{e}\,\in A^{\arity(X)}}\!\!\!\!\!\!\!
+X(\bar{e}).
 $$
-Note that each first-order sub-formula in $\varphi^i_j(\bar{X},\bar{a},\bar{b},\bar{c})$ has no free variables and, therefore, we can evaluate each of them in polynomial time and replace by $\perp$ and $\top$ wherever it corresponds.
-In other words, in polynomial time we can replace $\varphi^i_j$ by a disjunction of $\neg X_{h}$ and at most one $X_{h}$, evaluated on constants.
+\martin{reescribi esta formula}
+Note that each first-order subformula in $\varphi^i_j(\bar{X}_i,\bar{a},\bar{b},\bar{c})$ has no free variables and, therefore, we can evaluate each of them in polynomial time and easily rewrite $\alpha_{\A}$ to an equivalent formula that does not have any first-order subformula.
+In other words, in polynomial time we can replace $\varphi^i_j$ with a disjunction of $\neg X_{\ell}$ and at most one $X_{\ell}$, evaluated on constants.
 After simplifying, grouping and reordering the previous formula, we can obtain an equivalent formula $\alpha_{\A}'$ of the form:
@@
 $$
-where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{h}$ and at most one $X_{h}$.
+where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$ and at most one $X_{\ell}$.
 The reader can easily check that $\sem{\alpha}(\A) = \sem{\alpha_{\A}'}(\A)$.
@@
 First, if $\alpha_{\A}'$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi_{j,k}(\bar{X})$, then we can define $g_{\alpha}(\A)$ equal to the propositional formula $\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulas.
-Otherwise, if $m' > 1$ we can use $m'$ new fresh variables $t_1,\ldots,t_{m'}$ and define:
+Otherwise, if $m' > 1$ we can use $m'$ fresh new variables $t_1,\ldots,t_{m'}$ and define:
 $$
@@
 over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \} \cup \{t_1,\ldots,t_{m'}\}$.
-Intuitively, variables $t_1,\ldots,t_{m'}$ are used to count from $1$ to $m'$ each subformula $\sa{\bar{X}}\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi^{i}_{j,k}(\bar{X})$ in $\alpha_{\A}'$.
+Variables $t_1,\ldots,t_{m'}$ are used to have disjoint sets of propositional assignments for the different disjuncts of the outermost disjunction, which correspond to the summands in the original formula.
 One can easily check that $g_{\alpha}(\A)$ is a disjunction of Horn formulas, and the number of satisfying assignments is exactly $\sem{\alpha}(\A)$.

### 2019-01-21 LMCS/sendrev/proof-sigmafo-minusone.tex
reviewer 1 - technical casi listo

@@
-% !TeX root = main.tex
+%!TEX root = main.tex
 
-For the sake of readability, we divide the proof in three parts. The last part, i.e. subtraction by one, is probably the most technical proof of the paper.
+For the sake of readability, we divide the proof in three parts.
+The last part, i.e.
+subtraction by one, is probably the most technical proof of the paper.
+Furthermore, since this proof is several pages long, a line was drawn on the left margin to visually differentiate the proof from the rest of the paper.
+\martin{Ac\'a menciono la l\'inea del margen izquierdo}
 
@@
 This concludes the construction of $\A'$.
-Note that the number of $\fo$ subformulas, arity and tuple size is fixed in $\alpha$, so computing this function takes polynomial time over the size of the structure.
+Note that the number of $\fo$ subformulae, arity and tuple size is fixed in $\alpha$, so computing this function takes polynomial time over the size of the structure.
 Moreover, the encoding of $\A'$ has polynomial size over the size of $\enc(\A)$.
@@
 We will show how to define a formula $\varphi'$ that satisfies the following condition: for each $\A$, if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ for some $V$ and $v$ over $\A$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From this, we clearly have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ will be the desired formula.
-In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ the sum of one or more formulas in $\logex{1}$-PNF.
+From this, we will have that $\kappa(\alpha) = \sa{\bar{X}}\sa{\bar{x}}\varphi'$ is the desired formula.
+In the second part, we suppose that $\alpha$ is of the form $\beta + \sa{\bar{X}}\sa{\bar{x}}\varphi$ with $\beta$ being the sum of one or more formulae in $\logex{1}$-PNF.
 We define a formula $\varphi'$ that satisfies the following condition: if $(\A,V,v)\models \varphi(\bar{X},\bar{x})$ and $\sem{\beta}(\A) = 0$, then there exists exactly one assignment to $(\bar{X},\bar{x})$ that satisfies $\varphi$ and not $\varphi'$.
-From here, we can define $\kappa(\alpha)$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be~proven.
+From here, we can define $\kappa$ recursively as  $\kappa(\alpha) = \kappa(\beta) +  \sa{\bar{X}}\sa{\bar{x}}\varphi'$ and the property of subtraction by one will be~proven.
 
@@
 Note that, if $\alpha$ is of the form $\alpha = \sa{\bar{x}} \varphi(\bar{x})$ (i.e.
-$\length{\bar{X}} = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}} [\varphi(\bar{x})\wedge \ex{\bar{z}}(\varphi(\bar{z})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfills the desired condition.
+$\length{\bar{X}} = 0$), we can define $\kappa(\alpha) = \sa{\bar{x}} [\varphi(\bar{x})\wedge \ex{\bar{z}}(\varphi(\bar{z})\wedge\bar{z} < \bar{x})]$, which is in $\eqso(\logex{1})$ and fulfils the desired property.
 Therefore, for the rest of the proof we can assume that $\length{\bar{X}} > 0$ and $\length{\bar{x}} > 0$.
@@
 To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in a DNF formula. 
-More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i(\bar{X}, \bar{x}) = \ex{\bar{y}} \varphi_i'(\bar{X}, \bar{x}, \bar{y})$ and $\varphi_i'(\bar{X}, \bar{x}, \bar{y})$ is a conjunction of atomic formulas or negation of atomic formulas.
+More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i(\bar{X}, \bar{x}) = \ex{\bar{y}} \varphi_i'(\bar{X}, \bar{x}, \bar{y})$ and $\varphi_i'(\bar{X}, \bar{x}, \bar{y})$ is a conjunction of atomic formulae or negation of atomic formulae.
 Furthermore, we assume that each $\varphi_i'(\bar{X}, \bar{x}, \bar{y})$ has the~form:
@@
 $$
-Note that atomic formulas, like $R(\bar{z})$ for $R \in \R$, will appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
+Note that atomic formulae, like $R(\bar{z})$ for $R \in \R$, will appear in the subformula $\varphi_i^{\fo}(\bar{x},\bar{y})$.
 
-Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ is mentioned in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines an ordered partition over the variables in $(\bar{x},\bar{y})$ (see below for the formal definition of ordered partition) and (c) if $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned, then the ordered partition should not satisfy $\bar{z} = \bar{w}$.
+Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ appears in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines a weak ordering over the variables in $(\bar{x},\bar{y})$ (the precise definition of a weak ordering is detailed below) and (c) if both $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ appear in the formula, the weak ordering should not satisfy $\bar{z} = \bar{w}$.
 We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
-\begin{itemize} \itemsep1mm
-	\item[(a)] In order to satisfy the first condition, consider some instance of a $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
-We add $\length{\bar{w}}$ new variables $z_1,\ldots,z_{\length{\bar{w}}}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length{\bar{w}}})$.
+
+\medskip
+
+\noindent {\em (a) No variable from $\bar{x}$ appears in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$.} In order to satisfy this condition, consider some instance of $X_j(\bar{w})$ in $\varphi_i$, where $\bar{w}$ is a subtuple of $(\bar{x},\bar{y})$.
+Add $\length{\bar{w}}$ new variables $z_1,\ldots,z_{\length{\bar{w}}}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length{\bar{w}}})$.
 We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$ (denoted by $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})[X_j(\bar{w}) \leftarrow X_j(\bar{z})]$ and then the formula $\varphi_i$ is equivalently defined as:
@@
 $$
-	\varphi_i(\bar{X},\bar{x}) \ := \  \ex{\bar{y}} [ \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})].
+\varphi_i(\bar{X},\bar{x}) \ := \  \ex{\bar{y}} \big( \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})\big).
 $$
-	For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = \ex{\bar{y}} [ x < y \wedge  X(x,y)\wedge \neg X(x,x)]$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and:
+For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = \ex{\bar{y}} \big( x < y \wedge  X(x,y)\wedge \neg X(x,x)\big)$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and:
 $$
-	\varphi_i \ := \ \ex{\bar{y}}[ v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x \wedge x < y  \wedge  X(v_1,v_2) \wedge \neg X(v_3,v_4)].
+\varphi_i \ := \ \ex{\bar{y}}\big( v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x \wedge x < y  \wedge  X(v_1,v_2) \wedge \neg X(v_3,v_4)\big).
 $$
-	\item[(b)] An ordered partition on a set $S$ is defined by an equivalence relation $\sim$ over $S$, and a linear order over $S/\!\sim$.
+\noindent {\em (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines a weak ordering over the variables in $(\bar{x},\bar{y})$.} A weak ordering on a set $S$ is defined by an equivalence relation $\sim$ over $S$, and a linear order over $S/\!\sim$.
 For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$.
-A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
-On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
-	For a given $k$, let $\cB_k$ be the number of possible ordered partitions for a set of size $k$. For $1 \leq j \leq \cB_{\length{(\bar{x},\bar{y})}}$ 
-	let $\theta^j(\bar{x},\bar{y})$ be the formula that defines the $j$-th ordered partition over $(\bar{x},\bar{y})$. 
-	Thus, the formula $\varphi(\bar{X},\bar{x})$ is then redefined as:
+A possible weak ordering would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$.
+On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define a weak ordering since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
+For a given $k$, let $\cB_k$ be the number of possible weak orderings for a set of size $k$. For $1 \leq j \leq \cB_{\length{(\bar{x},\bar{y})}}$ 
+let $\theta^j(\bar{x},\bar{y})$ be the formula that defines the $j$-th weak ordering over $(\bar{x},\bar{y})$. 
+The formula $\varphi(\bar{X},\bar{x})$ is thus redefined as:
 $$
-	\varphi(\bar{X},\bar{x}) \ := \ \bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length{(\bar{x},\bar{y})}}} \ex{\bar{y}} [\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})],
+\varphi(\bar{X},\bar{x}) \ := \ \bigvee_{i = 1}^m \bigvee_{j = 1}^{\cB_{\length{(\bar{x},\bar{y})}}} \ex{\bar{y}} \big(\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y}) \wedge \varphi_i^{-}(\bar{X},\bar{y}) \wedge \varphi_i^{+}(\bar{X},\bar{y})\big),
 $$
 Note that each $\theta^j(\bar{x},\bar{y})$ is an $\fo$-formula.
-	Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  as $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ forces an ordered partition over the variables in $(\bar{x},\bar{y})$.
-	\item[(c)] Finally, to rewrite the formula such that no $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ are mentioned in $\varphi_i$ with $\bar{z}$ and $\bar{w}$ equivalent in the ordered partition (i.e.
-$\varphi_i$ is inconsistent), we do the following.
-If there exists an instance of $X_j(\bar{z})$ in $\varphi^{+}_i$, an instance of $\neg X_j(\bar{w})$ in $\varphi^{-}_i$ and the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$, then the entire formula $\varphi_i$ is removed from $\varphi$.
-\end{itemize}
+Then, by redefining $\varphi_i^{\fo}(\bar{x},\bar{y})$  as $\theta^j(\bar{x},\bar{y})\wedge \varphi_i^{\fo}(\bar{x},\bar{y})$, we can suppose that each $\varphi_i^{\fo}(\bar{x},\bar{y})$ forces a weak ordering over the variables in $(\bar{x},\bar{y})$.
+
+\noindent {\em (c) If both $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ appear in the formula, the weak ordering should not satisfy $\bar{z} = \bar{w}$.} 
+%Finally, to rewrite the formula such that no $X_j(\bar{z})$ or $\neg X_j(\bar{w})$ appear in $\varphi_i$ while $\bar{z}$ and $\bar{w}$ are made equal in the weak ordering (which would make $\varphi_i$ inconsistent), we do the following.
+If there exists an instance of $X_j(\bar{z})$ in $\varphi^{+}_i$, an instance of $\neg X_j(\bar{w})$ in $\varphi^{-}_i$ and the weak ordering in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$, then the entire formula $\varphi_i$ is removed from $\varphi$.
 It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$.
@@
 \proof
-	Let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$, such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
+	To prove the {\em only if} direction, let $\A$ be an ordered structure with domain $A$ and let $v$ be a first-order assignment for $\A$, such that $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$.
 	Define $\bar{B} = (B_1,\ldots,B_{\length{\bar{X}}})$ as $B_j = \{v(\bar{w})\mid \text{ $X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$, and let $V$ be a second-order assignment for which $V(\bar{X}) = \bar{B}$.
@@
 	However, by the construction of $B_t$, there exists a subtuple $\bar{z}$ of $\bar{y}$ such that $X_t(\bar{z})$ appears in $\varphi^{+}_i(\bar{X},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$.
-Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$.
+Since $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ and $v(\bar{z}) = v(\bar{w})$, then the weak ordering in $\varphi^{\fo}_i$ satisfies $\bar{z} = \bar{w}$.
 This violates condition (c) above since $\neg X_t(\bar{w})$ appears in $\varphi^{-}_i$ and $X_t(\bar{z})$ appears in $\varphi^{+}_i$, which leads to a contradiction.
 
-	For the other direction, if $(\A,V,v)\models \varphi_i'(\bar{X},\bar{x}, \bar{y})$ for a second order assignment $V$ for $\A$, then it is easy to check $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ since $\varphi^{\fo}_i(\bar{x},\bar{y})$ is a subformula of $\varphi_i'$.
+%	For the other direction, if $(\A,V,v)\models \varphi_i'(\bar{X},\bar{x}, \bar{y})$ for a second order assignment $V$ for $\A$, then it is easy to check $(\A,v)\models\varphi^{\fo}_i(\bar{x},\bar{y})$ since $\varphi^{\fo}_i(\bar{x},\bar{y})$ is a subformula of $\varphi_i'$.
+The other direction is trivial since $\varphi^{\fo}_i(\bar{x},\bar{y})$ is a subformula of $\varphi_i'$.
 \qed
@@
 For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length{\bar{X}}})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
-One can easily check that for every assignments $(V, v)$ such that $(\A,V,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$, it holds that $(\A,\bar{B}^v,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all assignments of $\bar{X}$ when $v$ is fixed.
-This motivates the main idea of Part (1): by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
+One can easily check that for every assignment $(V, v)$ such that $(\A,V,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$, it holds that $(\A,\bar{B}^v,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all assignments of $\bar{X}$ when $v$ is fixed.
+This offers an insight into the main idea of Part (1): by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 
 For this, we choose the minimal $v$ that satisfies $\varphi^{\fo}_i(\bar{x},\bar{y})$ which can be defined with the following formula:
 \[
-\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge \fa{\bar{x}'} \fa{\bar{y}'}[\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')].
+\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge \fa{\bar{x}'} \fa{\bar{y}'}\big(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')\big).
 \] 
 If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
-Furthermore, let $V^*$ be the second order assignment and $v^*$ the first order assignment that satisfy $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*(\bar{x}) = v(\bar{x})$.
+Furthermore, let $V^*$ be the second-order assignment and $v^*$ the first-order assignment that satisfy $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*(\bar{x}) = v(\bar{x})$.
 By the previous discussion, $(\A,V^*,v^*)\models\varphi_i(\bar{X},\bar{x})$.
@@
 \psi_i(\bar{X},\bar{x}) \; := \;  &\big( \, \ex{\bar{x}} \ex{\bar{y}} \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
-&\Big( \; 
-\ex{\bar{y}}[ \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \; \varphi_i'(\bar{X},\bar{x}, \bar{y})  \rightarrow \!\!
-\bigvee_{X \in \bar{X}} \ex{\bar{z}} [ \, X(\bar{z}) \wedge   \!\!\!\!\!\!\!\!\!\!
-\bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{y})} \!\!\!\!\!\!\!\!\!\!\!
-\bar{w}\neq\bar{z} \, ] \; \big)] \;  \vee   \label{eq:part2} \\
-&(\; \ex{\bar{x}'} \ex{\bar{y}}[ \varphi_i'(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}] \; ) \;\; \Big)  \label{eq:part3}
+&\Big( 
+\ex{\bar{y}}\!\big[ \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \, \varphi_i'(\bar{X},\bar{x}, \bar{y})  \rightarrow \!\!
+\bigvee_{X \in \bar{X}}\!
+\ex{\bar{z}}\!( \, X(\bar{z}) \wedge   \!\!\!\!\!\!\!\!\!\!\!\!
+\bigwedge\limits_{X(\bar{w}) \, \in \, \varphi^{+}_i(\bar{X},\bar{y})} \!\!\!\!\!\!\!\!\!\!\!\!
+\bar{w}\neq\bar{z} \, ) \, \big)\big] \;  \vee \!
+\label{eq:part2} \\
+&\; \ex{\bar{x}'} \ex{\bar{y}}\big( \varphi_i'(\bar{X},\bar{x}',\bar{y})\wedge \bar{x}' < \bar{x}\big) \; \;\; \Big)  \label{eq:part3}
 \end{align}
-\martin{agregu\'e parentesis de aqui hacia arriba}
+\martin{agregu\'e par\'entesis en varias f\'ormulas en esta p\'agina y las anteriores}
 To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
-Indeed, by Claim~\ref{claim:minusone} we know that if $\ex{\bar{x}} \ex{\bar{y}} \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exists assignments $V$ and $v$ such that $(\A,V,v) \models\varphi_i'(\bar{X},\bar{x},\bar{y})$.
-The conclusion of the implication (divided into (\ref{eq:part2}) and (\ref{eq:part3})), take care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
+Indeed, by Claim~\ref{claim:minusone} we know that if $\ex{\bar{x}} \ex{\bar{y}} \varphi^{\fo}_i(\bar{x},\bar{y})$ is true, then there exist assignments $V$ and $v$ such that $(\A,V,v) \models\varphi_i'(\bar{X},\bar{x},\bar{y})$.
+The conclusion of the implication (divided into (\ref{eq:part2}) and (\ref{eq:part3})), takes care that $V(\bar{X}) \neq V^*(\bar{X})$ or $v(\bar{x}) \neq v^*(\bar{x})$.
 Here, the first disjunct (\ref{eq:part2}) checks that $V(\bar{X}) \neq V^*(\bar{X})$ by defining that if $\varphi_i'(\bar{X},\bar{x}, \bar{y})$ is satisfied then $V^*(\bar{X}) \subsetneq V(\bar{X})$. 
 The second disjunct (\ref{eq:part3}) is satisfied when $v(\bar{x})$ is not the lexicographically smallest tuple that satisfies $\varphi_i$ (i.e. $v(\bar{x}) \neq v^*(\bar{x})$).
-Finally, from the previous discussion one can easily check that $\psi_i(\bar{X},\bar{x})$ satisfies the desire property.
+Finally, from the previous discussion one can easily check that $\psi_i(\bar{X},\bar{x})$ satisfies the desired property.
 
@@
 This models all the assignments that satisfy $\varphi_2$, except for the assignment $(V^*,v^*)$ of $\varphi_1$.
-Moreover, if $\varphi_1$ is not satisfiable, then $\psi_1(\bar{X},\bar{x})$ and $\chi_1$ will hold, and the formula $\psi_2(\bar{X},\bar{x})$ will forbid the assignment $(V^*,v^*)$ of $\varphi_2$.
-One can easily generalize this construction for each $\varphi_i$ as follows:
+Moreover, if $\varphi_1$ is not satisfiable, then $\psi_1(\bar{X},\bar{x})$ and $\chi_1$ will hold, and the formula $\psi_2(\bar{X},\bar{x})$ will exclude the assignment $(V^*,v^*)$ of $\varphi_2$.
+This construction can be generalized for each $\varphi_i$ as follows:
 \begin{multline*}
@@
 Let $\varphi = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x})$ where each $\varphi_i$ satisfies conditions (a), (b) and (c) defined above.
-Also, consider the previously defined formulas $\chi_i$ and $\psi_i$, for each $i \leq m$.
+Also, consider the previously defined formulae $\chi_i$ and $\psi_i$, for each $i \leq m$.
 We construct a function $\lambda$ that receives a formula $\beta \in \eqso(\logex{1})$ and produces a logic formula $\lambda(\beta)$ that satisfies $\A\models\lambda(\beta)$ if, and only if, $\sem{\beta}(\A) = 0$.

### 2019-01-21 LMCS/sendrev/todo-list.tex
reviewer 1 - technical casi listo

@@
 \usepackage{enumitem}
-\input{macros}
+\input{macrosplain}
 
@@
 \item[$\checkmark$] p.17: ``Since $\alpha$ is not null'' $\to$ ``Since $\sem{\alpha}$ is not the identically zero function'' (or similar)
-\item p.17/18: In the proof of Prop 5.3, the vocabulary only contains the numerical predicate $<$. When encoding structures as binary words, numerical predicates are implicit (and not
+\item[$\checkmark$] p.17/18: In the proof of Prop 5.3, the vocabulary only contains the numerical predicate $<$. When encoding structures as binary words, numerical predicates are implicit (and not
 encoded), though. What is the intended meaning here?
@@
 \item[$\checkmark$] p.22: ``Let $\alpha$ = ... where $\varphi$ is an FO-formula'' $\to$ ``... where $\varphi$ is an $\Sigma_1$[FO]-formula''
-\item p.23, formulae: Maybe put (large) parantheses around the relativization of quantifiers, for
+\item[$\checkmark$] p.23, formulae: Maybe put (large) parantheses around the relativization of quantifiers, for
 example the part of formula $\varphi_i$ after the quantifiers
@@
 be explained a little more / made more clear
-\item p.26: ``... auxiliary predicate $A$ defined as'' $\to$ ``... auxiliary predicate $A$ which will be defined
+\item[$\checkmark$] p.26: ``... auxiliary predicate $A$ defined as'' $\to$ ``... auxiliary predicate $A$ which will be defined
 as ... in the formula'' (or similar)
-\item p.28: ``since universal formulas are monotone over induced substructures'': ``monotone over
+\item[$\checkmark$] p.28: ``since universal formulas are monotone over induced substructures'': ``monotone over
 induced substructures'' should be reformulated
-\item p.28: ``Next we prove that ... is hard for ... over a signature $\R$ ...'' sounds like it only works
+\item[$\checkmark$] p.28: ``Next we prove that ... is hard for ... over a signature $\R$ ...'' sounds like it only works
 for one specific vocabulary. Maybe simply ``over any signature $\R$''
-\item p.28, formula $\alpha_{\A}$: second big conjunction should have $\bar{z}'$ in the subscript instead of $\bar{z}$ (because
+\item[$\checkmark$] p.28, formula $\alpha_{\A}$: second big conjunction should have $\bar{z}'$ in the subscript instead of $\bar{z}$ (because
 the quantifier $\forall u$ is handled by this as well if I understand correctly)
-\item p.29: ``In other words, in poly time we can replace $\varphi_j^i$ by a disjunction of ....'': Some of these
+\item[$\checkmark$] p.29: ``In other words, in poly time we can replace $\varphi_j^i$ by a disjunction of ....'': Some of these
 formulae might evaluate to $\perp$ or $\top$ altogether and can be removed (potentially together with
 the surrounding conjunction)
-\item p.29: ``Intuitively $t_1,\ldots,t_{m'}$ are used to count from 1 to $m'$ each subformula ...'': Should be
+\item[$\checkmark$] p.29: ``Intuitively $t_1,\ldots,t_{m'}$ are used to count from 1 to $m'$ each subformula ...'': Should be
 reformulated, maybe rather write that they are used to have disjoint sets of propositional
@@
 summands in the original formula)
-\item p.30, proof of Prop. 5.12: For $\dhsat \in \totp$, [35] is cited. It’s stated that a \totp-procedure
+\item[$\checkmark$] p.30, proof of Prop. 5.12: For $\dhsat \in \totp$, [35] is cited. It’s stated that a \totp-procedure
 for DNF-SAT is given there and it can be adapted to \dhsat. While the
@@
 able to fix such a problem.
-\item p.33: ``executes exactly $n^k$ steps for a fixed $k \geq 1$'' → ``executes exactly $n^k$
+\item[$\checkmark$] p.33: ``executes exactly $n^k$ steps for a fixed $k \geq 1$'' → ``executes exactly $n^k$
 steps on large
@@
 seperately)
-\item p.35 and p.36: While $\sum$ was used for finite sums in \qso-formulae throughout the paper, now
+\item[$\checkmark$] p.35 and p.36: While $\sum$ was used for finite sums in \qso-formulae throughout the paper, now
 a large +-sign is used. This might be confusing
-\item p.36: $\alpha_{q_i}$ used instead of $\alpha_{s_{q_{i}}}$
-\item p.36: ``Clearly, at each iteration of the fixed point operator, the tuple $\bar{t}$ represents the step the
+\item[$\checkmark$] p.36: $\alpha_{q_i}$ used instead of $\alpha_{s_{q_{i}}}$
+\item[$\checkmark$] p.36: ``Clearly, at each iteration of the fixed point operator, the tuple $\bar{t}$ represents the step the
 machine is currently in.'': This is quite vague. Maybe something like ``Clearly, tuple $\bar{t}$ encodes
@@
 one timestep of the machine is executed''
-\item p.36: ``at the $\bar{a}$-th iteration of the fixed point operator, it holds that $f(\bar{a}, \bar{b}) \leq 1$ and $f(\bar{a}', \bar{b}) = 0$ for every $\bar{b} \in A^k$. 
+\item[$\checkmark$] p.36: ``at the $\bar{a}$-th iteration of the fixed point operator, it holds that $f(\bar{a}, \bar{b}) \leq 1$ and $f(\bar{a}', \bar{b}) = 0$ for every $\bar{b} \in A^k$. 
 In the same way, it can be seen that for each function $g \in \ldots$ it holds that $g(\bar{a}) \leq 1$ and $g(\bar{a}')= 0$. 
@@
 predicates and are 0 when the timestep is the successor of the current timestep)
-\item p.36: ``is defined as an extension of QFO with the formula ...'' $\to$	 ``is defined as the extension
+\item[$\checkmark$] p.36: ``is defined as an extension of QFO with the formula ...'' $\to$	 ``is defined as the extension
 of QFO by the additional operator ...'' (or similar)

### 2019-01-21 LMCS/sendrev/proof-ehorn-pe.tex
cambiar formulas por formulae

@@
 If it is not, stop, and if it is, create a dummy branch that simply stops.
-Back in the main branch pick a variable $x$ in $\Phi$ and create two formulas $\Phi_0$ and $\Phi_1$ where $x$ has been replaced by $\perp$ or $\top$ respectively.
+Back in the main branch pick a variable $x$ in $\Phi$ and create two formulae $\Phi_0$ and $\Phi_1$ where $x$ has been replaced by $\perp$ or $\top$ respectively.
 If only one of these is satisfiable, continue on this branch with the respective $\Phi_i$, and if both are, create a new branch for $\Phi_1$ and continue on this branch with $\Phi_0$.

### 2019-01-21 LMCS/sendrev/proof-qe0-fp-qe1-totp-fptras.tex
carta de respuesta + correcciones varias
cambiar formulas por formulae

@@
 \proof
-	Given two formulas $\alpha, \beta$ in $\eqso(\LL)$ we will construct a formula in the logic which is equivalent to $(\alpha\mult \beta)$. 
+	Given two formulae $\alpha, \beta$ in $\eqso(\LL)$ we will construct a formula in the logic which is equivalent to $(\alpha\mult \beta)$. 
 	From what was proven in Proposition \ref{theo-pnf-snf}, we may assume that $\alpha$ and $\beta$ are in $\LL$-SNF. 
@@
 \qed
-Since $\loge{1}$ is closed under conjunction, then Lemma~\ref{conj-mult} holds for $\eqso(\loge{1})$. 
+Since $\loge{1}$ is closed under conjunction, then Lemma~\ref{conj-mult} can be applied to $\eqso(\loge{1})$. 
 This concludes the proof.

### 2019-01-21 LMCS/sendrev/proof-rqfo-fo-cap.tex
cambiar formulas por formulae

@@
 Notation $\sum_{(q,a) \,:\, \delta(q,a) = (\_,\vdash,\_,\_)}$ in this formula is used to indicate that we are considering a sum over all pairs $(q,a) \in Q \times \{0,1,\ttB,\vdash\}$ such that $\delta(q,a) = (q',\vdash,u,v)$ for some $q' \in Q$, $u \in \{\leftarrow,\rightarrow\}$ and $v \in \{0,1,\emptyset\}$.
-Formulas $\alpha_{T_0}$, $\alpha_{T_1}$ and $\alpha_{T_{\ttB}}$ are defined analogously; for the sake of presentation, we only show here how $\alpha_{T_0}$ is defined, assuming that $\bar y = (y_1, \ldots, y_k)$:
+Formulae $\alpha_{T_0}$, $\alpha_{T_1}$ and $\alpha_{T_{\ttB}}$ are defined analogously; for the sake of presentation, we only show here how $\alpha_{T_0}$ is defined, assuming that $\bar y = (y_1, \ldots, y_k)$:
 \begin{multline*}

### 2019-01-21 LMCS/sendrev/proof-sigma2hard.tex
cambiar formulas por formulae

@@
 \end{align*}
-One can check that $\psi'(T,A)$ effectively defines $\shdhsat$ as for every disjunction of Horn formulas $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi'(\pT,\pA)}(\A)$.
+One can check that $\psi'(T,A)$ effectively defines $\shdhsat$ as for every disjunction of Horn formulae $\theta = \theta_1\vee\cdots\vee\theta_m$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \sa{\pA} \psi'(\pT,\pA)}(\A)$.
 Therefore, we conclude that $\shdhsat \in \eqso(\ehorn)$.
@@
 Now, consider a finite $\R$-structure $\A$ with domain $A$. 
-The next transformation of $\alpha$ and $\A$ towards a disjunction of Horn-formulas is to expand each first-order quantifier (i.e.
+The next transformation of $\alpha$ and $\A$ towards a disjunction of Horn-formulae is to expand each first-order quantifier (i.e.
 $\Sigma{\bar{x}}$,  $\exists\bar{y}$, and $\forall\bar{z}$) as we replace their variables with first-order constants.
@@
 The idea for the rest of the proof is to show how to obtain $g_{\alpha}(\A)$, i.e. an instance of $\shdhsat$, from $\alpha_{\A}'$.
-First, if $\alpha_{\A}'$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi_{j,k}(\bar{X})$, then we can define $g_{\alpha}(\A)$ equal to the propositional formula $\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulas.
+First, if $\alpha_{\A}'$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi_{j,k}(\bar{X})$, then we can define $g_{\alpha}(\A)$ equal to the propositional formula $\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulae.
 Otherwise, if $m' > 1$ we can use $m'$ fresh new variables $t_1,\ldots,t_{m'}$ and define:
@@
 Variables $t_1,\ldots,t_{m'}$ are used to have disjoint sets of propositional assignments for the different disjuncts of the outermost disjunction, which correspond to the summands in the original formula.
-One can easily check that $g_{\alpha}(\A)$ is a disjunction of Horn formulas, and the number of satisfying assignments is exactly $\sem{\alpha}(\A)$.
+One can easily check that $g_{\alpha}(\A)$ is a disjunction of Horn formulae, and the number of satisfying assignments is exactly $\sem{\alpha}(\A)$.
 This covers all possible cases for $\alpha$, and the entire procedure takes polynomial time.

### 2019-01-21 LMCS/sendrev/proof-theo-pi1-pnf.tex
correcciones en la carpeta sendrev
cambiar formulas por formulae

--- 2018-12-09 LMCS/proof-theo-pi1-pnf.tex
@@
 If $\alpha = \alpha_1 + \alpha_2$, then we assume that $\tau(\alpha_1) = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ and $\tau(\alpha_2) = \sa{\bar{Y}}\sa{\bar{y}}\psi(\bar{Y},\bar{y})$. 
-Our construction works by identifying a ``first'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ and a ``last'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ using the following formulas:
+Our construction works by identifying a ``first'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ and a ``last'' assignment for both $(\bar{X},\bar{x})$ and $(\bar{Y},\bar{y})$ using the following formulae:
 \begin{align*}
@@
 \end{align*}
-Similarly, we can define the formulas $\gamma_{\text{first}}(\bar{Y},\bar{y})$ and $\gamma_{\text{last}}(\bar{Y},\bar{y})$.
+Similarly, we can define the formulae $\gamma_{\text{first}}(\bar{Y},\bar{y})$ and $\gamma_{\text{last}}(\bar{Y},\bar{y})$.
 In other words, the ``first'' assignment is the one where every second-order predicate is empty and the first-order assignment is the lexicographically smallest, and the ``last'' assignment is the one where every second-order predicate is full and the first-order assignment is the lexicographically greatest.
 We also need to identify the assignments that are not first and the ones that are not last. 
-We do this by negating the two formulas above and grouping together the first-order variables:
+We do this by negating the two formulae above and grouping together the first-order variables:
 \begin{align*}
@@
 
-To show that the formula is indeed equivalent to $\alpha$, note that the formulas in lines (\ref{eq:partition1}) and (\ref{eq:partition2}) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$, and the formulas in lines (\ref{eq:partition3}) and (\ref{eq:partition4}) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
+To show that the formula is indeed equivalent to $\alpha$, note that the formulae in lines (\ref{eq:partition1}) and (\ref{eq:partition2}) form a partition over the assignments of $(\bar{X},\bar{x})$, while fixing an assignment for $(\bar{Y},\bar{y})$, and the formulae in lines (\ref{eq:partition3}) and (\ref{eq:partition4}) form a partition over the assignments of $(\bar{Y},\bar{y})$, while fixing an assignment for $(\bar{X},\bar{x})$.
 Altogether the four lines define pairwise disjoint assignments for $(\bar{X},\bar{x}),(\bar{Y},\bar{y})$. 
 With this, it is straightforward to show that the above formula is equivalent to $\alpha$. 
-However, the formula is not yet in the correct form since it has existential quantifiers in the sub-formulas $\gamma_{\text{not-first}}$ and $\gamma_{\text{not-last}}$. 
+However, the formula is not yet in the correct form since it has existential quantifiers in the sub-formulae $\gamma_{\text{not-first}}$ and $\gamma_{\text{not-last}}$. 
 To solve this, we can replace each existential quantifier by a first order sum that counts just the first assignment that satisfies the inner formula and this can be defined in $\logu{1}$. 

### 2019-01-21 LMCS/sendrev/beyond.tex
cambiar : por \colon

@@
 where $h \in \fs$, $\arity(h) = \ell$ and $x_1, \ldots, x_\ell$ is a sequence of (not necessarily distinct) first-order variables.
-Given an $\R$-structure $\A$ with domain $A$, we say that $F$ is a \emph{function assignment} for $\A$ if for every $h \in \fs$ with $\arity(h) = \ell$, we have that $F(h) :  A^\ell \to \N$.
+Given an $\R$-structure $\A$ with domain $A$, we say that $F$ is a \emph{function assignment} for $\A$ if for every $h \in \fs$ with $\arity(h) = \ell$, we have that $F(h)\colon A^\ell \to \N$.
 The notion of function assignment is used to extend the semantics of $\qso$ to the case of a quantitative formula of the form $h(x_1, \ldots, x_\ell)$.
@@
 
-Fix an $\R$-structure with domain $A$ and a quantitative formula $\clfp{\beta(\x,h)}$ with $\arity(h) = \ell$, and assume that $\F$ is the set of functions $f :A^\ell \to \N$.
+Fix an $\R$-structure with domain $A$ and a quantitative formula $\clfp{\beta(\x,h)}$ with $\arity(h) = \ell$, and assume that $\F$ is the set of functions $f\colon A^\ell \to \N$.
 To define the semantics of $\clfp{\beta(\x,h)}$, we first show how $\beta(\x,h)$ can be interpreted as an operator $T_{\beta}$ on $\F$.
@@
 Let $h \in \fs$ such that $\arity(h) = \ell$, and $\beta$ be an $\fqfo(\fo)$-formula over a signature $\R$ such that if a function symbol occurs in $\beta$, then this function symbol is $h$.
-Moreover, let $\A$ be an $\R$-structure with domain $A$, $f,g : A^\ell \to \mathbb{N}$ and $F,G$ be function assignments such that $F(h) = f$ and $F(h) = g$.
+Moreover, let $\A$ be an $\R$-structure with domain $A$, $f,g\colon A^\ell \to \mathbb{N}$ and $F,G$ be function assignments such that $F(h) = f$ and $F(h) = g$.
 If $\support(f) \subseteq \support(g)$, then for every first-order and second-order assignments $v$ and $V$, respectively, it holds that:

### 2019-01-21 LMCS/sendrev/framework.tex
cambiar : por \colon

@@
 $
-for every $\A\in\ostr[\R]$. Then a function $g:\ostr[\R]\to\nat$ is in $\#\LL$ if there exists a formula $\varphi(\bar{x},\bar{X})$ in $\LL$ such that $g = f_{\varphi(\bar x, \bar X)}$.
+for every $\A\in\ostr[\R]$. Then a function $g\colon \ostr[\R]\to\nat$ is in $\#\LL$ if there exists a formula $\varphi(\bar{x},\bar{X})$ in $\LL$ such that $g = f_{\varphi(\bar x, \bar X)}$.
 In~\cite{SalujaST95}, they proved several results about capturing counting complexity classes which are relevant for our work.

### 2019-01-21 LMCS/sendrev/preliminaries.tex
correcciones en la carpeta sendrev
cambiar : por \colon

--- 2018-12-09 LMCS/preliminaries.tex
@@
 %We define the following function complexity classes: 
-$\fp$ is the class of functions $f : \Sigma^* \to \N$ computable in polynomial time, while $\fpspace$ is the class of functions $f : \Sigma^* \to \N$ computable in polynomial space. 
+$\fp$ is the class of functions $f \colon \Sigma^* \to \N$ computable in polynomial time, while $\fpspace$ is the class of functions $f \colon \Sigma^* \to \N$ computable in polynomial space. 
 %Moreover, $\nfpspace$ is the class of functions computable in polynomial space with output length bounded by a polynomial. 

### 2019-01-21 LMCS/sendrev/proof-qe0-fp-qe1-totp-fptras.tex
cambiar : por \colon

@@
 The authors in \cite{SalujaST95} proved that there exists a {\em product reduction} from every function in $\E{1}$ to a restricted version of $\cdnf$. 
-That is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g:\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h:\nat\to\nat$ such that for every $\R$-structure $\A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\size{\A})$.
+That is, if $\alpha\in\E{1}$ over some signature $\R$, there exist polynomially computable functions $g\colon\ostr[\R]\to\ostr[\R_{\text{DNF}}]$ and $h\colon\nat\to\nat$ such that for every $\R$-structure $\A$, it holds that $\sem{\alpha}(\A) = \cdnf(\enc(g(\A)))\cdot h(\size{\A})$.
 We use this fact in the following arguments.

### 2019-01-21 LMCS/sendrev/syntactic-horn.tex
correcciones en la carpeta sendrev
reviewer 1 - technical casi listo
cambiar : por \colon

--- 2018-12-09 LMCS/syntactic-horn.tex
@@
 To this end, we consider the notion of parsimonious reduction.
-Formally, a function $f : \Sigma^* \to \N$ is parsimoniously reducible to a function $g : \Sigma^* \to \N$ if there exists a function $h: \Sigma^* \to \Sigma^*$ such that $h$ is computable in polynomial time and $f(x) = g(h(x))$ for every $x \in \Sigma^*$.
+Formally, a function $f\colon\Sigma^* \to \N$ is parsimoniously reducible to a function $g\colon\Sigma^* \to \N$ if there exists a function $h\colon\Sigma^* \to \Sigma^*$ such that $h$ is computable in polynomial time and $f(x) = g(h(x))$ for every $x \in \Sigma^*$.
 As mentioned at the beginning of this section, if $f$ can be parsimoniously reduced to $g$, then $L_g \in \ptime$ implies that $L_f \in \ptime$ and the existence of an FPRAS for $g$ implies the existence of an FPRAS for $f$.
@@
 %The resulting formula $\varphi(\pT)$ is not in $\uhorn$, but it can be easily transformed into a formula in this class  by introducing an auxiliary binary predicate $\pA$ defined as follows:
-Moreover, by introducing an auxiliary predicate $\pA$ defined as:
+Moreover, by introducing an auxiliary predicate $\pA$ which will be defined as it appears in the formula
 \begin{align*}

### 2019-01-21 LMCS/sendrev/syntactic.tex
carta de respuesta + correcciones varias
cambiar : por \colon

--- 2018-12-09 LMCS/syntactic.tex
@@
 Among them, problems having easy decision counterparts play a fundamental role, as a counting problem with a hard decision version is expected to be hard.
-Formally, the decision problem associated to a function $f : \Sigma^* \to \mathbb{N}$ is defined as $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$, and $f$ is said to have an easy decision version if $L_f \in \ptime$.
+Formally, the decision problem associated to a function $f\colon\Sigma^* \to \mathbb{N}$ is defined as $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$, and $f$ is said to have an easy decision version if $L_f \in \ptime$.
 Many prominent examples satisfy this property, like computing the number of: perfect matchings of a bipartite graph ($\cpm$)~\cite{Valiant79}, satisfying assignments of a DNF ($\cdnf$)~\cite{DHK05,KL83} or Horn ($\chsat$)~\cite{V79b} propositional formula, among others.
@@
 Counting problems with easy decision versions play a fundamental role in the search for efficient approximation algorithms for functions in $\shp$. 
-A fully-polynomial randomized approximation scheme (FPRAS) for a function $f : \Sigma^* \to \bbN$ is a randomized algorithm $\mathcal{A} : \Sigma^* \times (0,1) \to \bbN$ such that: (1) for every string $x \in \Sigma^*$ and real value $\varepsilon \in (0,1)$, the probability that $|f(x) - \mathcal{A}(x,\varepsilon)| \leq \varepsilon \cdot f(x)$ is at least $\frac{3}{4}$, and (2) the running time of $\mathcal{A}$ is polynomial in the size of $x$ and $1/\varepsilon$ \cite{KL83}.
+A fully-polynomial randomized approximation scheme (FPRAS) for a function $f\colon \Sigma^* \to \bbN$ is a randomized algorithm $\mathcal{A} \colon \Sigma^* \times (0,1) \to \bbN$ such that: (1) for every string $x \in \Sigma^*$ and real value $\varepsilon \in (0,1)$, the probability that $|f(x) - \mathcal{A}(x,\varepsilon)| \leq \varepsilon \cdot f(x)$ is at least $\frac{3}{4}$, and (2) the running time of $\mathcal{A}$ is polynomial in the size of $x$ and $1/\varepsilon$ \cite{KL83}.
 Notably, there exist $\shp$-complete functions that can be efficiently approximated as they admit FPRAS; for instance, there exist FPRAS for $\cdnf$~\cite{KL83} and $\cpm$~\cite{JSV04}. 
@@
 Hence, under the widely believed assumption that  $\np \not\subseteq \bpp$, we cannot hope for an FPRAS for a function in $\shp$ whose decision counterpart is $\np$-complete, and we have to concentrate on the class of counting problems with easy decision versions.
+That is, with decision versions in P.
 

### 2019-01-21 LMCS/sendrev/response-letter.tex
varias reescrituras gramaticales

@@
 \end{itemize}
-\vspace{1em}
-\subsection*{Grammatical/Language}
-\begin{itemize}
-	\setlength\itemsep{0.5em}
-	\item generally: both formulae and formulas are used as plural of formula---should be consistent
-	\item p.2, in the middle: we propose a restriction: Should be something like ``we propose to restrict
-	WL to the natural numbers as fixed semiring, calling this restriction ...''
+%\vspace{1em}
+%\subsection*{Grammatical/Language}
+%\begin{itemize}
+%	\setlength\itemsep{0.5em}
+%	\item generally: both formulae and formulas are used as plural of formula---should be consistent
+%	\item p.2, in the middle: we propose a restriction: Should be something like ``we propose to restrict WL to the natural numbers as fixed semiring, calling this restriction ...''
 %	\item[$\checkmark$] p.3, first paragraph of 2.1: ``omit the word finite or ordered'' $\to$ ``omit the words finite and ordered''
-	\item p5, description of syntax of QSO: ``This division between Boolean and quantitative level
-	is essential for understanding the difference between the logic and the quantitative part.''
-	reformulate this?
+%	\item p5, description of syntax of QSO: ``This division between Boolean and quantitative level is essential for understanding the difference between the logic and the quantitative part.'' reformulate this?
 %	\item[$\checkmark$] p.5, next sentence: ``this will allow us later'' $\to$ ``this will later allow us''
-%	\item[$\checkmark$] p.6, second to last sentence: ``similarity of $\alpha_3$ with'': ``similarity to'' is much more common
-	than ``similarity with''
+%	\item[$\checkmark$] p.6, second to last sentence: ``similarity of $\alpha_3$ with'': ``similarity to'' is much more common than ``similarity with''
 %	\item[$\checkmark$] p.7, next sentence: ''in this direction'' $\to$ ``in this regard''
 %	\item[$\checkmark$] p.7, last sentence before 3.1: ''connection of QSO with'' $\to$ ``connection of QSO to''
-%	\item[$\checkmark$] p.7, middle: ``Another difference of WL with our approach'' $\to$ ``Another difference between
-	WL and our approach''
-	\item p.9, first sentence: ``then condition (2) holds'': omit the word then or reformulate the sentence
+%	\item[$\checkmark$] p.7, middle: ``Another difference of WL with our approach'' $\to$ ``Another difference between WL and our approach''
+%	\item p.9, first sentence: ``then condition (2) holds'': omit the word then or reformulate the sentence
 %	\item[$\checkmark$] p.9, first paragraph: ``the constant function $s$ can be trivially simulated'' $\to$ ``any constant function $s$ can be trivially simulated''
@@
 %	\end{itemize}
-	\item p.11, end of proof of Thm. 4.5: ``Using an analogous argument'': state what the argument is analogous to (the argument in Thm. 4.4)
+%	\item p.11, end of proof of Thm. 4.5: ``Using an analogous argument'': state what the argument is analogous to (the argument in Thm. 4.4)
 %	\item[$\checkmark$] p.12, after Corollary 4.8: ``shows how robust and versatile is QSO'' $\to$ ``shows how robust and versatile QSO is''
-	\item p.12, after Corollary 4.8: ``counting complexity classes [...] even beyond $\nat$'' - make more precise what ``beyond $\nat$'' refers to
+%	\item p.12, after Corollary 4.8: ``counting complexity classes [...] even beyond $\nat$'' - make more precise what ``beyond $\nat$'' refers to
 %	\item[$\checkmark$] p.12, ``given CNF propositional formula, that can be made true'': no comma
-	\item p.12: ``in [43] they defined'' $\to$ something like ``$<$authors$>$ [43] defined''
+%	\item p.12: ``in [43] they defined'' $\to$ something like ``$<$authors$>$ [43] defined''
 %	\item[$\checkmark$] p.12: ``we extend as follows QSO'' $\to$ ``we extend QSO by max and min quantifiers as follows''
@@
 %	\item[$\checkmark$] p.14: ``the search of robust classes'' $\to$ ''the search for robust classes''
-	\item p.14: ''the first desirable condition has to do with the closure properties satisfied by the class'':
-	Reformulate this sentence
+%	\item p.14: ''the first desirable condition has to do with the closure properties satisfied by the class'': Reformulate this sentence
 %	\item[$\checkmark$] p.14: ``As in the cases of P and NP that are closed'' $\to$ ``Analogously to the cases of P and NP, which are closed under intersection and union, ...''
 %	\item[$\checkmark$] p.16: ``such that for both, $A\mod\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$'' $\to$ ''such that for $i \in \{1,2\}$: ... ''
-%	\item[$\checkmark$] p.16: ``Similarly to Theorem 5.1, we show ..'' $\to$ ``Similarly as in the proof of Theorem 5.1, we
-	show''
-	\item p.19: ``However, this class is not closed under sum, and then it is not robust under basic
-	closure properties'': Should be reformulated
+%	\item[$\checkmark$] p.16: ``Similarly to Theorem 5.1, we show ..'' $\to$ ``Similarly as in the proof of Theorem 5.1, we 	show''
+%	\item p.19: ``However, this class is not closed under sum, and then it is not robust under basic	closure properties'': Should be reformulated
 %	\item[$\checkmark$] p.19: ``Let ... be this formula, where $\varphi$ is in first-order and quantifier-free'' $\to$ ``... where $\varphi$ is first-order and quantifier-free''
@@
 %	\item[$\checkmark$] p.20: ``we focus only in proving'' $\to$ ``we focus only on proving''
-%	\item[$\checkmark$] p.20: ``We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of FO.''
-	$\to$ ``We prove this for the more general case of $\eqso(\LL)$ with $\LL$ being a fragment of SO.'' (or similar)
+%	\item[$\checkmark$] p.20: ``We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of FO.'' $\to$ ``We prove this for the more general case of $\eqso(\LL)$ with $\LL$ being a fragment of SO.'' (or similar)
 %	\item[$\checkmark$] p.20: ``we will show a formula'' $\to$ ``give'' or ``construct'' instead of ``show''
@@
 %	\item[$\checkmark$] p.21, next sentence: ``for a possible extension'' $\to$ ``as a possible extension''
-	\item p.21: ``the desired extension has to be achieved by ...
+%	\item p.21: ``the desired extension has to be achieved by ...
 .
@@
 %	\item[$\checkmark$] p.37: ``Next we construct a logarithmic-space ....'' $\to$ ``We construct a logarithmic-space ...''
-	\item[?] p.38: ``In this sense, there are several directions for future research'' $\to$ ``Consequently, there are ...'' (or similar)
+%	\item[?] p.38: ``In this sense, there are several directions for future research'' $\to$ ``Consequently, there are ...'' (or similar)
 %	\item[$\checkmark$] p.38: ``In the same direction, ...'' $\to$ ``Similarly, ...'' (or similar)
 %	\item[$\checkmark$] p.38: ``so to have'' $\to$ ``in order to have'' (or similar)
-\end{itemize}
+%\end{itemize}
 \vspace{1em}

### 2019-01-21 LMCS/sendrev/beyond.tex
varias reescrituras gramaticales

@@
 Given an $\R$-structure $\A$ with domain $A$, define a (directed) graph $\cG_{\psi}(\A) = (N,E)$ such that $N = A^k$ and for every pair $\bar b, \bar c \in N$, it holds that $(\bar b, \bar c) \in E$ if, and only if, $\A \models \psi(\bar b, \bar c)$.
-Similar than for Example~\ref{ex:count-path}, we can count the paths of length at most $|A^k|$ in $ \cG_{\psi}(\A)$ with the formula $\beta_{\psi(\bar{x},\bar{y})}(\x, \y, \t, g)$:
+As was the case for Example~\ref{ex:count-path}, we can count the paths of length at most $|A^k|$ in $ \cG_{\psi}(\A)$ with the formula $\beta_{\psi(\bar{x},\bar{y})}(\x, \y, \t, g)$:
 \begin{align*}

### 2019-01-21 LMCS/sendrev/extentions.tex
correcciones en la carpeta sendrev
varias reescrituras gramaticales

--- 2018-12-09 LMCS/extentions.tex
@@
 a constant) comes from $\bbZ$.
-Similar than for $\qso$, we define the fragment $\eqsoz$ as the extension of $\eqso$ with constants in $\bbZ$.
+Similarly as for $\qso$, we define the fragment $\eqsoz$ as the extension of $\eqso$ with constants in $\bbZ$.
 \begin{exa}
@@
 \end{cor}
-This is an interesting result that shows how robust and versatile $\qso$ is for capturing different counting complexity classes even beyond $\bbN$.
+This is an interesting result that shows how robust and versatile $\qso$ is for capturing different counting complexity classes whose functions are not restricted to $\bbN$.
 
@@
 For instance, MAX-SAT is in $\optp$ as many other optimization versions of $\np$-problems.
-Given that in~\cite{krentel1988complexity} Krentel did not make the distinction between $\max$ and $\min$, in~\cite{vollmer1995complexity} they defined the classes $\maxp$ and $\minp$ as the max and min version of the problems in $\optp$ (i.e.
+Given that in~\cite{krentel1988complexity} Krentel did not make the distinction between $\max$ and $\min$, Vollmer and Wagner~\cite{vollmer1995complexity} defined the classes $\maxp$ and $\minp$ as the max and min version of the problems in $\optp$ (i.e.
 $\optp = \maxp \cup \minp$).
@@
 \end{exa}
-Similar than for $\maxp$ and $\minp$, we have to distinguish between the $\max$ and $\min$ fragments of $\optqso$.
+Similarly as for $\maxp$ and $\minp$, we have to distinguish between the $\max$ and $\min$ fragments of $\optqso$.
 For this, we define the fragment $\maxqso$ of all $\optqso$ formulae constructed from $\qfo$ operators and $\max$-formulae $\max\{\alpha,\alpha\}$, $\maxa{x} \alpha$ and  $\maxa{X} \alpha$.

### 2019-01-21 LMCS/sendrev/framework.tex
varias reescrituras gramaticales

@@
 Furthermore, $\Sigma x$ and $\Sigma X$ are called first- and second-order sum, whereas $\Pi x$ and $\Pi X$ are called first- and second-order product, respectively.
-This division between Boolean and quantitative level is essential for understanding the difference between the logic and the quantitative part. 
+%This division between Boolean and quantitative level is essential for understanding the difference between the logic and the quantitative part. 
+This separation between the Boolean and quantitative levels is essential for understanding the difference between the logic and the quantitative halves of the framework.
+\martin{creo que eso quer\'ia decir esta oraci\'on}
 Furthermore, this will later allow us to parametrize both levels of the logic in order to capture different counting complexity classes.

### 2019-01-21 LMCS/sendrev/introduction.tex
varias reescrituras gramaticales

@@
 $\bbN$).
-Specifically, we propose a restriction of WL over natural numbers, called Quantitative Second Order Logics (QSO), and study its expressive power for defining counting complexity classes over ordered structures.
+Specifically, we propose to restrict WL to the natural numbers as a fixed semiring, calling this restriction Quantitative Second Order Logics (QSO), and study its expressive power for defining counting complexity classes over ordered structures.
 As a proof of concept, we show that natural syntactical fragments of $\qso$ captures counting complexity classes like $\shp$, $\spp$, $\fp$ and $\fpspace$.

### 2019-01-21 LMCS/sendrev/proof-capture-optp.tex
correcciones en la carpeta sendrev
varias reescrituras gramaticales

--- 2018-12-09 LMCS/proof-capture-optp.tex
@@
 It is straightforward to prove that $\maxp$ can compute any $\fo$-formula, is closed under first-order sum and product, and second-order maximization. 
-Therefore, condition (1) in Definition~\ref{def:cap} follows similar than in the previous characterizations. Furthermore, one can easily see that the same holds for $\minqso(\fo)$.
-The proof for the other direction is similar than in \cite{kolaitis1994logical} extended with the ideas of Theorem~\ref{theo:capture-fp}.
+Therefore, condition (1) in Definition~\ref{def:cap} follows similarly as in the previous characterizations. Furthermore, one can easily see that the same holds for $\minqso(\fo)$.
+The proof for the other direction is similar to the one described in \cite{kolaitis1994logical} extended with the ideas of Theorem~\ref{theo:capture-fp}.
 Let $f\in \maxp$ be a function defined over some signature $\R$ and
@@
 On the other hand, if $\Phi(\bar{X}, Y)$ holds, the subformula will produce $\val(\enc(U))$. 
-Similar than for $\max$, we conclude that $\alpha$ is in $\minqso(\fo)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.
+In a similar way as in $\max$, we conclude that $\alpha$ is in $\minqso(\fo)$ and $\sem{\alpha}(\A) = f(\enc(\A))$.

### 2019-01-21 LMCS/sendrev/proof-capture-shP.tex
correcciones en la carpeta sendrev
varias reescrituras gramaticales

--- 2018-12-09 LMCS/proof-capture-shP.tex
@@
-We briefly explain that the two conditions of Definition~\ref{def:cap} are satisfied.
+We briefly explain how the two conditions of Definition~\ref{def:cap} are satisfied.
 First, for condition (2) Saluja et al.
 proved that $\shp = \sfo$\cite{SalujaST95}.
-Hence, given that every function in $\sfo$ can be trivially defined as a formula in $\eqso(\fo)$ (see Section~\ref{sec:previous}) then condition~(2) holds.
+Hence, given that every function in $\sfo$ can be trivially defined as a formula in $\eqso(\fo)$ (see Section~\ref{sec:previous}), condition~(2) holds.
 For condition (1), let $\alpha\in\eqso(\fo)$ over some signature $\R$.

### 2019-01-21 LMCS/sendrev/syntactic-clo.tex
reviewer 1 - technical casi listo
varias reescrituras gramaticales

@@
 But before doing this, we introduce a more strict notion of counting problem with easy decision version.
-Recall that a function $f : \Sigma^* \to \mathbb{N}$ has an easy decision counterpart if $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$ is a language in $\ptime$.
+Recall that a function $f\colon\Sigma^* \to \mathbb{N}$ has an easy decision counterpart if $L_f = \{ x \in \Sigma^* \mid f(x) > 0 \}$ is a language in $\ptime$.
 As the goal of this section is to define a syntactic class of functions in $\shp$ with easy decision versions and good closure properties, we do not directly consider the semantic condition $L_f  \in \ptime$, but instead we consider a more restricted syntactic condition.
-More precisely, a function $f : \Sigma^* \to \mathbb{N}$ is said to be in the complexity class $\totp$~\cite{PagourtzisZ06} if there exists a  polynomial-time NTM $M$ such that $f(x) = \tmt_M(x) - 1$ for every $x \in \Sigma^*$, where $\tmt_M(x)$ is the total number of runs of $M$ with input $x$.
+More precisely, a function $f\colon\Sigma^* \to \mathbb{N}$ is said to be in the complexity class $\totp$~\cite{PagourtzisZ06} if there exists a  polynomial-time NTM $M$ such that $f(x) = \tmt_M(x) - 1$ for every $x \in \Sigma^*$, where $\tmt_M(x)$ is the total number of runs of $M$ with input $x$.
 Notice that one is subtracted from $\tmt_M(x)$ to allow for $f(x) = 0$.
@@
 It was proved in \cite{SalujaST95} that every function in $\E{1}$ admits an FPRAS. Besides, it can be proved that $\E{1} \subseteq \totp$. 
-However, this class is not closed under sum, and then it is not robust under basic closure properties. 
+However, this class is not closed under sum, so it is not robust under the basic closure properties we are looking for. 
 \begin{prop}\label{prop-e1-nc}
@@
 Therefore, the desired extension has to be achieved by allowing some local extensions to~$\Sigma_1$.
-More precisely, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulae over a signature~$\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mention any second-order variable).
+To this end, we define $\logex{1}$ as $\Sigma_1$ but allowing atomic formulae over a signature~$\R$ to be of the form either $u = v$ or $X(\bar u)$, where $X$ is a second-order variable, or $\varphi(\bar u)$, where $\varphi(\bar u)$ is a first-order formula over $\R$ (in particular, it does not mention any second-order variable).
 With this extension we obtain a class with the desired properties.
@@
 \end{thm}
+\begin{leftbar}
 \proof
@@
 \qed
+\end{leftbar}
 The proof that $\eqso(\logex{1})$ is closed under subtraction by one is the most involved of the paper.
 We think the main technique used in this proof, which is based on considering some witnesses of logarithmic size, is of independent interest.
+\martin{Esta correcci\'on se me fue en collera}

### 2019-01-21 LMCS/sendrev/syntactic.tex
varias reescrituras gramaticales

@@
 But the key question here is what should be considered a {\em robust} class.
-A first desirable condition has to do with the closure properties satisfied by the class, which is a common theme when studying function complexity classes \cite{OH93,FH08}.
+A first desirable condition is related to its closure properties, which is a common theme when studying function complexity classes \cite{OH93,FH08}.
 Analogously to the cases of $\ptime$ and $\np$, which are closed under intersection and union, we expect our class to be closed under multiplication and sum.

### 2019-01-21 LMCS/sendrev/todo-list.tex
varias reescrituras gramaticales

@@
 	\setlength\itemsep{0.5em}
-	\item generally: both formulae and formulas are used as plural of formula---should be consistent
-	\item p.2, in the middle: we propose a restriction: Should be something like ``we propose to restrict
-	WL to the natural numbers as fixed semiring, calling this restriction ...''
+	\item[$\checkmark$] generally: both formulae and formulas are used as plural of formula---should be consistent
+	\item[$\checkmark$] p.2, in the middle: we propose a restriction: Should be something like ``we propose to restrict WL to the natural numbers as fixed semiring, calling this restriction ...''
 	\item[$\checkmark$] p.3, first paragraph of 2.1: ``omit the word finite or ordered'' $\to$ ``omit the words finite and
 	ordered''
-	\item p5, description of syntax of QSO: ``This division between Boolean and quantitative level
-	is essential for understanding the difference between the logic and the quantitative part.''
-	reformulate this?
+	\item[$\checkmark$] p5, description of syntax of QSO: ``This division between Boolean and quantitative level is essential for understanding the difference between the logic and the quantitative part.'' reformulate this?
 	\item[$\checkmark$] p.5, next sentence: ``this will allow us later'' $\to$ ``this will later allow us''
@@
 	WL and our approach''
-	\item p.9, first sentence: ``then condition (2) holds'': omit the word then or reformulate the sentence
+	\item[$\checkmark$] p.9, first sentence: ``then condition (2) holds'': omit the word then or reformulate the sentence
 	\item[$\checkmark$] p.9, first paragraph: ``the constant function $s$ can be trivially simulated'' $\to$ ``any constant
@@
 	\item[$\checkmark$] p.9, proof of Prop. 4.3: ``To prove the condition (2)'' $\to$ ``To prove condition (2)''
-	\item throughout the paper: ``similar than'' $\to$ ``similar to'' / ``similar than for X, we do Y'' $\to$
+	\item[$\checkmark$] throughout the paper: ``similar than'' $\to$ ``similar to'' / ``similar than for X, we do Y'' $\to$
 	``similarly as for''. Examples:
@@
 	\end{itemize}
-	\item p.11, end of proof of Thm. 4.5: ``Using an analogous argument'': state what the argument is
+	\item[$\checkmark$] p.11, end of proof of Thm. 4.5: ``Using an analogous argument'': state what the argument is
 	analogous to (the argument in Thm. 4.4)
@@
 	versatile QSO is''
-	\item p.12, after Corollary 4.8: ``counting complexity classes [...] even beyond $\nat$'' - make more precise
+	\item[$\checkmark$] p.12, after Corollary 4.8: ``counting complexity classes [...] even beyond $\nat$'' - make more precise
 	what ``beyond $\nat$'' refers to
 	\item[$\checkmark$] p.12, ``given CNF propositional formula, that can be made true'': no comma
-	\item p.12: ``in [43] they defined'' $\to$ something like ``$<$authors$>$ [43] defined''
+	\item[$\checkmark$] p.12: ``in [43] they defined'' $\to$ something like ``$<$authors$>$ [43] defined''
 	\item[$\checkmark$] p.12: ``we extend as follows QSO'' $\to$ ``we extend QSO by max and min quantifiers as follows''
@@
 	\item[$\checkmark$] p.14: ``the search of robust classes'' $\to$ ''the search for robust classes''
-	\item p.14: ''the first desirable condition has to do with the closure properties satisfied by the class'':
+	\item[$\checkmark$] p.14: ''the first desirable condition has to do with the closure properties satisfied by the class'':
 	Reformulate this sentence
@@
 	show''
-	\item p.19: ``However, this class is not closed under sum, and then it is not robust under basic
+	\item[$\checkmark$] p.19: ``However, this class is not closed under sum, and then it is not robust under basic
 	closure properties'': Should be reformulated
@@
 	\item[$\checkmark$] p.21, next sentence: ``for a possible extension'' $\to$ ``as a possible extension''
-	\item p.21: ``the desired extension has to be achieved by ... . More precisely, we define...'': The second
+	\item[$\checkmark$] p.21: ``the desired extension has to be achieved by ... . More precisely, we define...'': The second
 	sentence doesn’t make the statement of the first one more precise, but rather explains the
@@
 	the previous proposition to after ...''
-	\item p.29: ``replace by $\perp$ and $\top$ \underline{wherever it corresponds}'': Should be reformulated
+	\item[$\checkmark$] p.29: ``replace by $\perp$ and $\top$ \underline{wherever it corresponds}'': Should be reformulated
 	\item[$\checkmark$] p.30: ``consist in computing $g_{\alpha}(\enc(\A))$ for each input ...'' $\to$ ``... on input ...''
@@
 	6
-	\item[?] p.38: ``In this sense, there are several directions for future research'' $\to$ ``Consequently, there
+	\item[$\checkmark$] p.38: ``In this sense, there are several directions for future research'' $\to$ ``Consequently, there
 	are ...'' (or similar)
@@
 	\item[$\checkmark$] running title is ``COUNTING''
-	\item on definition of functions, \verb|\colon| should be used instead of :
+	\item[$\checkmark$] on definition of functions, \verb|\colon| should be used instead of :
 	\item In [1], a similar hierarchy to the one by Saluja et al. was introduced and studied. This work
@@
 -it does not mentioned $\to$ it does not mention
-\item page 22: 
+\item[$\checkmark$] page 22: 
 -you make a long proof about some closure under subtraction by one.

### 2019-01-21 LMCS/sendrev/beyond.tex
un par de typos

@@
 
-We have used weighted logics to give a framework for descriptive complexity of counting complexity classes.
+We have used weighted logics to give a framework for descriptive complexity of counting complexity~classes.
 Here, we go beyond weighted logics and give the first steps on defining recursion at the quantitative level.
@@
 used 
-to define a natural generalization for functions of 
-the notion of least fixed point of LFP.
+to define a natural generalization of LFP to functions.
  As a proof of concept, we show that this notion can be used to capture $\fp$.

### 2019-01-22 LMCS/sendrev/response-letter.tex
correcciones Reviewer 2

@@
 
-	In a graph that is only one cycle of size $n$, namely $C_n$, the fixed point is reached after step $n$.
+	The problem with the first formula for directed graphs is precisely the opposite.
+The fixed point could be reached {\em before} we want it to.
+The fixed point is always reached after $k$ steps where $k$ is the size of the maximum shortest path between two vertices in the graph, and if the graph has a cycle in it, there might be paths of size $k' \in [k+1,n]$ that wouldn't be counted because the fixed point was reached earlier.
+
+%	In a graph that is only one cycle of size $n$, namely $C_n$, the fixed point is reached after step $n$.
 
@@
 %	\item[$\checkmark$] -replace .... for $\to$ replace ... by;  there are several instances of that
-	\item[$\checkmark$] -$\phi$ is an FO-formula $\to$ a $\Sigma_1[FO]$ formula
-	\item -you explain that you rewrite your formula to obtain conditions a,b and c, 
-	you should give an idea how these conditions will be used
-	\item[$\checkmark$] page 23:
-	a SO, a FO  an SO, an FO
-	\item page 24: 
-	if, if $\phi$$\to$ $\phi$
-	\item page 25: 
-	-you explain that the proof relies on witnesses of logarithmic size.
-	It was not clear to me while reading the proof. Can you explain more ?
-	\item page 28:
-	-the complete problem for $\Sigma$ QSO($\Sigma_2$ Horn) would be complete also for $\#\Sigma-2$ Horn. You should mention here that you do not need sum here
-	and that the framework of [39] is enough for this result.
-	\item[?] page 30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
+%	\item[$\checkmark$] -$\phi$ is an FO-formula $\to$ a $\Sigma_1[FO]$ formula
+%	\item -you explain that you rewrite your formula to obtain conditions a,b and c, you should give an idea how these conditions will be used
+%	\item[$\checkmark$] page 23: a SO, a FO  an SO, an FO
+%	\item page 24: 	if, if $\phi$$\to$ $\phi$
+%	\item page 25: 	-you explain that the proof relies on witnesses of logarithmic size.	It was not clear to me while reading the proof. Can you explain more ?
+	\item page 28:	-the complete problem for $\Sigma$ QSO($\Sigma_2$ Horn) would be complete also for $\#\Sigma-2$ Horn.
+You should mention here that you do not need sum here	and that the framework of [39] is enough for this result.
+%	\item[?] page 30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
 %	\item[$\checkmark$] page 32: - paths of length at most n $\to$ in an acylic graph, all paths are of size less than n
-	\item page 33:
-	-there is a proof symbol \verb|\qed| but it does not end a well defined proof environment
-
-	\item page 38: 
-	the last sentence on fixed point operator is to vague. Either expand it to 
-	say what king of question on fixed point operator you want to investigate or remove it	
+%	\item page 33: 	-there is a proof symbol \verb|\qed| but it does not end a well defined proof environment
+	\item page 38: the last sentence on fixed point operator is to vague. Either expand it to say what king of question on fixed point operator you want to investigate or remove it	
 \end{itemize}

### 2019-01-22 LMCS/sendrev/proof-sigmafo-minusone.tex
correcciones Reviewer 2

@@
 
-The proof will be separated in two parts.
+{\em Proof outline:} The proof will be separated in two parts.
 In the first part, we will assume that  $\alpha$ is in $\logex{1}$-PNF, namely, $\alpha = \sa{\bar{X}}\sa{\bar{x}}\varphi$ for some $\varphi$ in $\logex{1}$.
@@
 
-To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ in a DNF formula. 
+To simplify the analysis of $\varphi$, the first step is to rewrite $\varphi$ as a DNF formula. 
 More precisely, we rewrite $\varphi$ into an equivalent formula of the form $\bigvee_{i = 1}^m \varphi_i$ for some $m\in \bbN$ where each $\varphi_i(\bar{X}, \bar{x}) = \ex{\bar{y}} \varphi_i'(\bar{X}, \bar{x}, \bar{y})$ and $\varphi_i'(\bar{X}, \bar{x}, \bar{y})$ is a conjunction of atomic formulae or negation of atomic formulae.
@@
 
-Now, we define a series of rewritings to $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ appears in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines a weak ordering over the variables in $(\bar{x},\bar{y})$ (the precise definition of a weak ordering is detailed below) and (c) if both $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ appear in the formula, the weak ordering should not satisfy $\bar{z} = \bar{w}$.
+Now, we define a series of rewrites of $\varphi$ that will make each formula $\varphi_i$ satisfy the following three conditions: (a) no variable from $\bar{x}$ appears in $\varphi_i^{-}(\bar{X},\bar{x},\bar{y})\wedge\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$, (b) $\varphi_i^{\fo}(\bar{x},\bar{y})$ defines a weak ordering over the variables in $(\bar{x},\bar{y})$ (the precise definition of a weak ordering is detailed below) and (c) if both $X_j(\bar{z})$ and $\neg X_j(\bar{w})$ appear in the formula, the weak ordering should not satisfy $\bar{z} = \bar{w}$.
+When these conditions are met, $\varphi_i'(\bar{X}, \bar{x}, \bar{y})$ will be satisfiable if and only if $\varphi_i^{\fo}(\bar{x},\bar{y})$ is satisfiable.
+This is proven in Claim~\ref{claim:minusone}.
 We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
+\martin{Peque\~na explicaci\'on de para qu\'e sirven las condiciones a b y c}
 
@@
 Add $\length{\bar{w}}$ new variables $z_1,\ldots,z_{\length{\bar{w}}}$ to the formula and let $\bar{z} = (z_1,\ldots,z_{\length{\bar{w}}})$.
-We redefine $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$ (denoted by $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})[X_j(\bar{w}) \leftarrow X_j(\bar{z})]$ and then the formula $\varphi_i$ is equivalently defined as:
+We rewrite $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})$ by replacing $X_j(\bar{w})$ with $X_j(\bar{z})$ (denoted by $\varphi_i^{+}(\bar{X},\bar{x},\bar{y})[X_j(\bar{w}) \leftarrow X_j(\bar{z})]$) and then the formula $\varphi_i$ is equivalently defined as:
 $$
@@
 It is important to notice that the resulting $\varphi$ is equivalent to the initial one, and it is still a formula in $\eqso(\logex{1})$.
-From now on, we assume that each $\varphi_i(\bar{X},\bar{x}) = \ex{\bar{y}}  \varphi_i'(\bar{X},\bar{x}, \bar{y}) $ satisfies conditions (a), (b) and (c), and $\varphi_i'(\bar{X},\bar{x}, \bar{y})$ has the~form:
+From now on, we assume that each $\varphi_i(\bar{X},\bar{x}) = \ex{\bar{y}}  \varphi_i'(\bar{X},\bar{x}, \bar{y}) $ satisfies conditions (a), (b) and (c), and moreover, $\varphi_i'(\bar{X},\bar{x}, \bar{y})$ has the~form:
 $$
-\varphi_i'(\bar{X},\bar{x}, \bar{y}) \; = \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}) 
+\varphi_i'(\bar{X},\bar{x}, \bar{y}) \; = \; \varphi^{\fo}_i(\bar{x},\bar{y}) \wedge \varphi^{+}_i(\bar{X}, \bar{y}) \wedge \varphi^{-}_i(\bar{X},\bar{y}).
 $$
-where $\varphi^{+}$ and $\varphi^{-}_i$ do not depend on $\bar{x}$.
+Note that neither one of $\varphi^{+}$ or $\varphi^{-}_i$ depends on $\bar{x}$.
 \begin{clm}\label{claim:minusone}
@@
 
-The previous claim and proof motivate the following definition.
-For a structure $\A$ and a first-order assignment $v$ for $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length{\bar{X}}})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
-One can easily check that for every assignment $(V, v)$ such that $(\A,V,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$, it holds that $(\A,\bar{B}^v,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$, namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all assignments of $\bar{X}$ when $v$ is fixed.
+The previous claim its subsequent proof motivate the following definition.
+For a first-order assignment $v$ over $\A$, define $\bar{B}^v = (B^v_1,\ldots,B^v_{\length{\bar{X}}})$ where each $B^v_j = \{v(\bar{w}) \mid \text{$X_j(\bar{w})$ is mentioned in $\varphi^{+}_i(\bar{X},\bar{y})$}\}$.
+One can easily check that for every assignment $(V, v)$ such that $(\A,V,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$, it holds that $(\A,\bar{B}^v,v)\models \varphi_i'(\bar{X}, \bar{x}, \bar{y})$ and $\bar{B}^v \subseteq V(\bar{X})$.
+Namely, $\bar{B}^v$ is a valid candidate for $\bar{X}$ and, furthermore, it is contained in all satisfying assignments of $\bar{X}$ when $v$ is fixed.
 This offers an insight into the main idea of Part (1): by choosing one particular $v$ the plan is to remove $\bar{B}^v$ as an assignment over $\bar{X}$ in $\varphi_i$. 

### 2019-01-22 LMCS/sendrev/todo-list.tex
un par de typos
correcciones Reviewer 2

@@
 be explained a little more / made more clear
+\martin{no logr\'e completar esta correcci\'on}
 \item[$\checkmark$] p.26: ``... auxiliary predicate $A$ defined as'' $\to$ ``... auxiliary predicate $A$ which will be defined
@@
 elements of $A^{\ell}$, not $\{1,\ldots,\ell\}$
-\item p.32/33, example 6.2: I don't see why the formula given for dag's does not work for arbitrary
+\item[?] p.32/33, example 6.2: I don't see why the formula given for dag's does not work for arbitrary
 graphs: In arbitrary graphs, the fixpoint is still reached after at most $n-1$ steps, since the
@@
 able to fix such a problem.
+\martin{me gustar\'ia tener feedback para saber si mi correcci\'on est\'a bien}
 \item[$\checkmark$] p.33: ``executes exactly $n^k$ steps for a fixed $k \geq 1$'' → ``executes exactly $n^k$
@@
 	\setlength\itemsep{0.5em}
-	\item p.1, footnote: ``clases'' should be ``classes''
+	\item[$\checkmark$] p.1, footnote: ``clases'' should be ``classes''
 	\item throughout the paper: plural used with verb ending on s. Examples:
@@
 	\end{itemize}
+	\martin{no alcanc\'e a revisar esta ni la de cuatro puntos m\'as adelante}
 	\item[$\checkmark$] p4, beginning of 2.2: ``some of \underline{the} are recalled here''
@@
 \end{itemize}
+	\martin{lo que menciono arriba}
 	\item[$\checkmark$] p.9, middle: ``A second attempt could be based then'' $\to$ ``A second attempt could then be
@@
 	claim and proof motivate the following definition.''
-	\item p.30: ``to define a natural generalization for functions of the notion of least fixed point of
+	\item[$\checkmark$] p.30: ``to define a natural generalization for functions of the notion of least fixed point of
 	LFP'' $\to$ ``to define a natural generalization of LFP to functions'' (or similar)
@@
 	\item[$\checkmark$] on definition of functions, \verb|\colon| should be used instead of :
-	\item In [1], a similar hierarchy to the one by Saluja et al. was introduced and studied. This work
-	should be cited.
+	\item In [1], a similar hierarchy to the one by Saluja et al. was introduced and studied. This work should be cited.
 \end{itemize}
@@
 You may also add some example to illustrate the different fragment of QSO (you already do it sometimes and it is very helpful).
+\martin{me gustar\'ia que conversemos qu\'e demostraciones necesitan esto, y si el outline de la demostraci\'on de -1 basta. Tambi\'en qu\'e fragmentos de QSO necesitan un ejemplo.}
 
@@
 \item[$\checkmark$] -$\phi$ is an FO-formula $\to$ a $\Sigma_1[FO]$ formula
-\item -you explain that you rewrite your formula to obtain conditions a,b and c, 
+\item[$\checkmark$] -you explain that you rewrite your formula to obtain conditions a,b and c, 
 you should give an idea how these conditions will be used
@@
 a SO, a FO  an SO, an FO
-\item page 24: 
+\item[$\checkmark$] page 24: 
 if, if $\phi$$\to$ $\phi$
@@
 It was not clear to me while reading the proof. Can you explain more ?
+\martin{no logr\'e corregir esto, es la misma acotaci\'on que el Reviewer 1}
 \item page 28:
@@
 and that the framework of [39] is enough for this result.
-\item[?] page 30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
+\martin{conversemos esto}
+\item[$\checkmark$] page 30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
 
@@
 - paths of length at most n $\to$ in an acylic graph, all paths are of size less than n
-\item page 33:
+\item[$\checkmark$] page 33:
 -there is a proof symbol \verb|\qed| but it does not end a well defined proof environment
@@
 say what king of question on fixed point operator you want to investigate or remove it	
+\martin{conversemos esto}
 \end{itemize}

### 2019-01-22 LMCS/sendrev/beyond.tex
correcciones Reviewer 2
un par de correcciones y algunos problemas de compatibilidad con el la instruccion !TEX = main

@@
 Moreover, assume that $b,c$ are elements of $\A$ and $v$ is a first-order assignment over $\A$ such that $v(x) = b$ and $v(y) = c$.
-Then we have that $\sem{\clfp{\alpha(x,y,f)}}(\A,v)$ is equal to the  number of paths in $\A$ from $b$ to $c$ of length at most $n$.
+Then we have that $\sem{\clfp{\alpha(x,y,f)}}(\A,v)$ is equal to the number of paths in $\A$ from $b$ to $c$ of length at most $n-1$.
 
@@
 Notice that the use of the filter $\varphi_{\text{succ}}(t',t)$ prevents this formula for incrementing the value of $g(x,y,t)$ when $t$ is the first element in the domain.
-\qed
 \end{exa}
+\martin{El Reviewer 1 hizo un comentario de este ejemplo, lo respond\'i en la carta}
 In contrast to $\lfp$, to reach a fixed point we do not need to impose any positive restriction on the formula $\beta(\x,h)$.

### 2019-01-25 LMCS/sendrev/example-3dnf-minusone.tex
cambios menores en -1 y agregue un ejemplo del testigo tamano llog

%!TEX root = main.tex

Let $\R = \{P_0(\cdot,\cdot,\cdot), P_1(\cdot,\cdot,\cdot), P_2(\cdot,\cdot,\cdot), P_3(\cdot,\cdot,\cdot)\}$. This vocabulary is used as follows to encode a 3-DNF formula. Relations $P_0,P_1,P_2,P_3$ evaluated on a tuple $(x,y,z)$ hold if the disjuncts $(x\wedge y \wedge z),(x\wedge y \wedge \neg z),(x\wedge \neg y \wedge \neg z),(\neg x\wedge\neg y \wedge\neg z)$ are present in the formula, respectively.
To define $\ctdnf$, we consider an $\so$-formula $\varphi(T)$ over $\R$, where $T$ is an unary predicate, such that for every 3-DNF formula $\theta$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sa{T}\varphi(T)$. Particularly, $T(x)$ holds if, and only if, $x$ is assigned value true. In detail,
\begin{align*}
\varphi(T) = \ex{x}\ex{y}\ex{z}\big(\,
&(P_0(x,y,z)\wedge \ \; T(x)\wedge \ \; T(y)\wedge \ \; T(z))\,\vee\\
&(P_1(x,y,z)\wedge \ \; T(x)\wedge \ \; T(y)\wedge \!\neg T(z))\,\vee\\
&(P_2(x,y,z)\wedge \ \; T(x)\wedge\!\neg T(y)\wedge \!\neg T(z))\,\vee\\
&(P_3(x,y,z)\wedge \!\neg T(x)\wedge \!\neg T(y)\wedge \!\neg T(z))\,\big).
\end{align*}
This formula is in $\loge{1}$, so $\ctdnf$ is in $\eqso(\loge{1})$, and indeed, also in $\eqso(\logex{1})$.

We will apply a simplified version of our construction to $\varphi$ to illustrate how it restricts a single assignment to $T$. First, note that we assumed the formula had the form $\sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ where $\bar{x}$ was not empty, so we start by doing the corresponding replacement. Consider the formula is $\varphi(T,u)$ and in each disjunct the expression $\wedge \first(u)$ is added. The next step of our construction was rewriting each disjunct so a sequence of conditions (that we called {\em a, b} and {\em c}) held. After the rewritings have been done, one of the disjuncts will be
\[
\varphi'_{\ell}(T,u,x,y,z) = \underbrace{P_0(x,y,z)\wedge \theta(u,x,y,z)\wedge \first(u)}_{\varphi^{\fo}_{\ell}(u,x,y,z)}\wedge \underbrace{T(x)\wedge T(y)\wedge T(z)}_{\varphi^{+}_{\ell}(T,x,y,z)},
\]
Where $\theta(u,x,y,z)$ is a weak ordering for $(u,x,y,z)$. Furthermore,
\begin{multline*}
\text{$min$-}\varphi^{\fo}_{\ell}(u,x,y,z) = P_0(x,y,z)\wedge \theta(u,x,y,z)\wedge \first(u)\, \wedge \\ \fa{u'}\fa{x'}\fa{y'}\fa{z'}\big(P_0(x',y',z')\wedge \theta(u',x',y',z')\wedge \first(u')\to(u\leq u' \wedge \langle x,y,z\rangle\leq \langle x',y',z'\rangle)\big),
\end{multline*}
and lastly,
\begin{align*}
\psi_{\ell}(T,u) = \; &\ex{u'}\ex{x}\ex{y}\ex{z}(P_0(x,y,z)\wedge \theta(u',x,y,z)\wedge \first(u'))\,\to \\
&\bigg(\ex{x}\ex{y}\ex{z}\Big[\text{$min$-}\varphi^{\fo}_{\ell}(u,x,y,z)\,\wedge \\ 
&\hspace{5em}\big(\,T(x)\wedge T(y)\wedge T(z)\to\ex{w}(T(w)\wedge w \neq x \wedge w\neq y \wedge w\neq z)\big)\Big]\, \vee\\
&\ex{u'}\ex{x}\ex{y}\ex{z}(\varphi'_{\ell}(T,u',x,y,z)\wedge u' < u)\bigg).
\end{align*}
Note that we omitted the $\varphi^{\fo}_{\ell}(u,x,y,z)$ part around the beginning of the third line of the formula because it is redundant. The crux of our construction are the second and third lines in the formula. Given a first-order assignment $v$ that satisfies $\text{$min$-}\varphi^{\fo}_{\ell}(u,x,y,z)$, consider the tuple $\langle v(x),v(y),v(z)\rangle$. We end our example by bringing attention to the fact that this tuple is a logarithmic-sized witness which effectively identifies, and lets us restrict our desired assignment to $T$.
\martin{este ejemplo sirve para ayudar a entender la idea del testigo de tama\~no logar\'itmico}

### 2019-01-25 LMCS/sendrev/syntactic-clo.tex
un par de correcciones y algunos problemas de compatibilidad con el la instruccion !TEX = main
cambios menores en -1 y agregue un ejemplo del testigo tamano llog

@@
 The proof that $\eqso(\logex{1})$ is closed under subtraction by one is the most involved of the paper.
-We think the main technique used in this proof, which is based on considering some witnesses of logarithmic size, is of independent interest.
-\martin{Esta correcci\'on se me fue en collera}
+%We think the main technique used in this proof, which is based on considering some witnesses of logarithmic size, is of independent interest.
+The main technique in this proof involves using witnesses of logarithmic size, namely $v^{*}(\bar{y})$, which we believe could prove to be useful by itself. 
+We will now borrow an example from Saluja et. al~\cite{SalujaST95} to illustrate how the construction described in the proof effectively restricts the desired assignments on a formula.
+\begin{exa}
+\input{example-3dnf-minusone}
+\end{exa}

### 2019-02-27 LMCS/sendrev/framework.tex
sc

@@
 %This division between Boolean and quantitative level is essential for understanding the difference between the logic and the quantitative part. 
-This separation between the Boolean and quantitative levels is essential for understanding the difference between the logic and the quantitative halves of the framework.
+This separation between the Boolean and quantitative levels is essential for understanding the difference between the logic and the quantitative parts of the framework.
 \martin{creo que eso quer\'ia decir esta oraci\'on}
+\marcelo{OK}
 Furthermore, this will later allow us to parametrize both levels of the logic in order to capture different counting complexity classes.

### 2019-02-27 LMCS/sendrev/proof-prop-rest.tex
sc

@@
 
-We give this proof in three parts.
+We divide this proof into three parts.
 First, we show that $\E{1} \not\subseteq \QE{0}$. 
-For this inclusion to be true, it is required to hold for an arbitrary ordered relational signature $\R$, so it suffices to show it is not for at least one.
-Let $\R$ be the minimal ordered signature that contains only the relation name $<$.
-Suppose that there is a $\QE{0}$ formula $\alpha$ over $\R$ which is equivalent to the $\E{1}$ formula $\sa{x} \ex{y} (x < y)$. 
+For this inclusion to be true, it is required to hold for an arbitrary ordered relational signature $\R$, so it suffices to show that it  is not true for at least one such a signature.
+Let $\R$ be the ordered signature that contains only the relation name $<$.
+Suppose that there is a $\QE{0}$ formula $\alpha$ over $\R$ which is equivalent to the $\E{1}$-formula $\sa{x} \ex{y} (x < y)$. 
 That is, for every finite $\R$-structure $\A$, $\sem{\alpha}(\A) = \size{\A} - 1$.
 \martin{reformul\'e por qu\'e elijo ese R}
+\marcelo{OK}
 Suppose that $\alpha$ is in SNF, namely, $\alpha = \sum_{i = 1}^k \alpha_i$ for some fixed $k$. 
@@
 Let $\alpha_i = \sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ where $\varphi$ is quantifier-free. 
-Note that if $\length{\bar{X}} > 0$, then the function $\sem{\alpha}$ is in $\Omega(2^{\size{\A}})$, as it was proven in~\cite{SalujaST95}. 
+Notice that if $\length{\bar{X}} > 0$, then the function $\sem{\alpha}$ is in $\Omega(2^{\size{\A}})$, as it was proven in~\cite{SalujaST95}. 
 Therefore, we have that $\alpha_i = \sa{\bar{x}}\varphi(\bar{x})$. 
-We conclude our proof with the following claim.
+We conclude our proof with the following claim, from which we conclude that $\alpha = \sum_{i = 1}^k \alpha_i$ cannot be equivalent to the $\E{1}$-formula $\sa{x} \ex{y} (x < y)$. 
 \begin{clm}
-	Let $\alpha = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. 
-	Then the function $\sem{\alpha}$ is either null, greater or equal to $n$, or is in $\Omega(n^2)$, where $n$ is the size of the input structure.
+	Let $\beta = \sa{\bar{x}}\varphi(\bar{x})$	where $\varphi$ is quantifier free. 
+	Then the function $\sem{\beta}$ is either null, greater or equal to $n$, or is in $\Omega(n^2)$, where $n$ is the size of the input structure.
 \end{clm}
 \proof
-	Note that each atomic sub-formula in $\varphi(\bar{x})$ is either $(x = y)$, $(x < y)$, $\top$ or a negation thereof, where $x,y\in\bar{x}$. 
-	Suppose $\sem{\alpha}$ is not null and consider some $\R$-structure $\A$ such that $\sem{\alpha}(\A) > 0$. 
-	Let $\bar{a}$ be an assignment to $\bar{x}$ such that $\A\models\varphi(\bar{a})$. 
-	It can be seen that each assignment $\bar{a}'$ that has the same ordering over its variables\footnote{For example, if $\bar{a} = (a_1,a_2,a_3,a_4)$, such ordering may be $a_1 > a_3 = a_4 > a_2$, which has tree partitions.}	as $\bar{a}$ also satisfies $\A\models\varphi(\bar{a}')$.
-	If this ordering has $k$ partitions then $\binom{\size{\A}}{k}$ assignments for $\bar{x}$ will satisfy it, and therefore, $\sem{\alpha} \geq \binom{\size{\A}}{k}$. 
-	If $k = 1$, then $\binom{\size{\A}}{k} = \size{\A}$, and if $k \geq 2$, then $\binom{\size{\A}}{k} \in \Omega(\size{\A}^2)$, which proves the claim.
+Assume that $\bar x = (x_1, \ldots, x_m)$, and notice that each atomic sub-formula in $\varphi(\bar{x})$ is either $(x_i = x_j)$, $(x_i < x_j)$, $\top$ or a negation thereof, where $i,j \in \{1, \ldots, m\}$.
+	Suppose $\sem{\beta}$ is not null and consider some $\R$-structure $\A$ such that $\sem{\beta}(\A) > 0$. Hence, there exists an assignment $\bar a = (a_1, \ldots, a_m)$
+%	Let $\bar{a}$ be an assignment 
+	for $\bar{x}$ such that $\A\models\varphi(\bar{a})$.
+	Given this assignment, define an equivalence relation $\sim$ on $\{x_1, \ldots, x_m\}$ as follows: $x_i \sim x_j$ if and only if $a_i = a_j$, and assume that $\sim$ partitions $\{x_1, \ldots, x_m\}$ into $\ell$ equivalence classes, where $\ell \geq 1$.
+%	It can be seen that each assignment $\bar{a}'$ that has the same ordering over its variables\footnote{For example, if $\bar{a} = (a_1,a_2,a_3,a_4)$, such ordering may be $a_1 > a_3 = a_4 > a_2$, which has tree partitions.}	as $\bar{a}$ also satisfies $\A\models\varphi(\bar{a}')$.
+%	If this ordering has $k$ partitions then $\binom{\size{\A}}{k}$ assignments for $\bar{x}$ will satisfy it, and therefore, $\sem{\beta} \geq \binom{\size{\A}}{k}$. 
+Then we have that there exist at least $\binom{\size{\A}}{\ell}$ assignments $\bar b$ for $\bar{x}$ such that $\A\models\varphi(\bar{b})$.
+Thus, given that if $\ell = 1$, then $\binom{\size{\A}}{\ell} = \size{\A}$, and if $\ell \geq 2$, then $\binom{\size{\A}}{\ell} \in \Omega(\size{\A}^2)$, we conclude that the claim holds.
 \qed
 \martin{cambie order por ordering, creo que se entiende mejor}
+\marcelo{me parece que las nociones de ordering y particion no quedaban claras, por esto defini de manera precisa la nocion de particion y cambie la demostracion}
 Now we show that $\QE{0} \not\subseteq \E{1}$. 

### 2019-02-28 LMCS/sendrev/example-3dnf-minusone.tex
sc

@@
 Let $\R = \{P_0(\cdot,\cdot,\cdot), P_1(\cdot,\cdot,\cdot), P_2(\cdot,\cdot,\cdot), P_3(\cdot,\cdot,\cdot)\}$.
-This vocabulary is used as follows to encode a 3-DNF formula.
-Relations $P_0,P_1,P_2,P_3$ evaluated on a tuple $(x,y,z)$ hold if the disjuncts $(x\wedge y \wedge z),(x\wedge y \wedge \neg z),(x\wedge \neg y \wedge \neg z),(\neg x\wedge\neg y \wedge\neg z)$ are present in the formula, respectively.
+This vocabulary is used as follows to encode a 3-DNF formula $\theta$.
+Relations $P_0,P_1,P_2,P_3$ evaluated on a tuple $(x,y,z)$ hold if the formulae $(x\wedge y \wedge z),(x\wedge y \wedge \neg z),(x\wedge \neg y \wedge \neg z),(\neg x\wedge\neg y \wedge\neg z)$ occur in $\theta$, respectively.
 To define $\ctdnf$, we consider an $\so$-formula $\varphi(T)$ over $\R$, where $T$ is an unary predicate, such that for every 3-DNF formula $\theta$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sa{T}\varphi(T)$.
-Particularly, $T(x)$ holds if, and only if, $x$ is assigned value true.
-In detail,
+More specifically, $T(x)$ holds if and only if  $x$ is assigned value true, so that $\varphi(T)$ is defined as follows:
 \begin{align*}
@@
 
-We will apply a simplified version of our construction to $\varphi$ to illustrate how it restricts a single assignment to $T$.
-First, note that we assumed the formula had the form $\sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ where $\bar{x}$ was not empty, so we start by doing the corresponding replacement.
+We apply a simplified version of our construction to $\varphi$ to illustrate how it restricts a single assignment to $T$.
+First, notice that we assumed the formula had the form $\sa{\bar{X}}\sa{\bar{x}}\varphi(\bar{X},\bar{x})$ where $\bar{x}$ was not empty, so we start by doing the corresponding replacement.
 Consider the formula is $\varphi(T,u)$ and in each disjunct the expression $\wedge \first(u)$ is added.

### 2019-02-28 LMCS/sendrev/proof-qe0-fp-qe1-totp-fptras.tex
un par de correcciones y algunos problemas de compatibilidad con el la instruccion !TEX = main
sc

@@
 \qed
-Since $\loge{1}$ is closed under conjunction, then Lemma~\ref{conj-mult} can be applied to $\eqso(\loge{1})$. 
-This concludes the proof.
+Since $\loge{1}$ is closed under conjunction, we have that Lemma~\ref{conj-mult} can be applied to $\eqso(\loge{1})$, and we can deduce that $\eqso(\loge{1})$ is closed under multiplication. 
+This concludes the proof of the proposition.

### 2019-02-28 LMCS/sendrev/proof-sigmafo-minusone.tex
cambios menores en -1 y agregue un ejemplo del testigo tamano llog
sc

@@
 
-For the sake of readability, we divide the proof in three parts.
+For the sake of readability, we divide the proof into three parts.
 The last part, i.e.
-subtraction by one, is probably the most technical proof of the paper.
-Furthermore, since this proof is several pages long, a line was drawn on the left margin to visually differentiate the proof from the rest of the paper.
+subtraction by one, is the most technical proof in the paper.
+Since this proof is several pages long, a line was drawn on the left margin to visually differentiate it from the rest of the paper.
 \martin{Ac\'a menciono la l\'inea del margen izquierdo}
+\marcelo{OK}
 
@@
 \noindent {\bf Closed under subtraction by one.} We prove here that $\eqso(\logex{1})$ is closed under subtraction by one. 
-For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$-formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
+For this, given $\alpha \in \eqso(\logex{1})$ over a signature $\R$, we will define a $\eqso(\logex{1})$-formula $\kappa(\alpha)$ such that for each finite structure $\A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$.
 Without loss of generality, we assume that $\alpha$ is in $\logex{1}$-SNF, that is, $\alpha = \sum_{i = 1}^{n}\sa{\bar{X}}\sa{\bar{x}}\varphi_i$ where each $\varphi_i$ is in $\logex{1}$.
@@
 \martin{Peque\~na explicaci\'on de para qu\'e sirven las condiciones a b y c}
+\marcelo{OK}
 
@@
 \[
-\text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge \fa{\bar{x}'} \fa{\bar{y}'}\big(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')\big).
+\text{{\it min}-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \; = \;  \,\varphi_i^{\fo}(\bar{x},\bar{y})\wedge \fa{\bar{x}'} \fa{\bar{y}'}\big(\varphi_i^{\fo}(\bar{x}',\bar{y}')\to ( \bar{x}\leq\bar{x}' \wedge  \bar{y}\leq\bar{y}')\big).
 \] 
-If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
+If $\varphi^{\fo}_i$ is satisfiable, let $v$ be the only assignment such that  $(\A,v)\models \text{{\it min}-}\varphi^{\fo}_i(\bar{x}, \bar{y})$. 
 Furthermore, let $V^*$ be the second-order assignment and $v^*$ the first-order assignment that satisfy $V^*(\bar{X}) = \bar{B}^{v}$ and $v^*(\bar{x}) = v(\bar{x})$.
@@
 \begin{align}
-\psi_i(\bar{X},\bar{x}) \; := \;  &\big( \, \ex{\bar{x}} \ex{\bar{y}} \varphi^{\fo}_i(\bar{x},\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
+\psi_i(\bar{X},\bar{x}) \; := \;  &\big( \, \ex{\bar{x}'} \ex{\bar{y}} \varphi^{\fo}_i(\bar{x}',\bar{y}) \,\big) \rightarrow  \label{eq:part1} \\
 &\Big( 
-\ex{\bar{y}}\!\big[ \text{$min$-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \, \varphi_i'(\bar{X},\bar{x}, \bar{y})  \rightarrow \!\!
+\ex{\bar{y}}\!\big[ \text{{\it min}-}\varphi^{\fo}_i(\bar{x}, \bar{y}) \wedge \big( \, \varphi_i'(\bar{X},\bar{x}, \bar{y})  \rightarrow \!\!
 \bigvee_{X \in \bar{X}}\!
@@
 \martin{agregu\'e par\'entesis en varias f\'ormulas en esta p\'agina y las anteriores}
+\marcelo{OK}
 To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 

### 2019-02-28 LMCS/sendrev/syntactic-clo.tex
sc

--- 2019-01-21 LMCS/sendrev/syntactic-clo.tex
@@
 \end{leftbar}
-The proof that $\eqso(\logex{1})$ is closed under subtraction by one is the most involved of the paper.
-We think the main technique used in this proof, which is based on considering some witnesses of logarithmic size, is of independent interest.
-\martin{Esta correcci\'on se me fue en collera}
+%The proof that $\eqso(\logex{1})$ is closed under subtraction by one is the most involved of the paper.
+%%We think the main technique used in this proof, which is based on considering some witnesses of logarithmic size, is of independent interest.
+%The main technique in this proof involves using witnesses of logarithmic size, namely $v^{*}(\bar{y})$, which we believe could prove to be useful by itself. 
+%We will now borrow an example from Saluja et. al~\cite{SalujaST95} to illustrate how the construction described in the proof effectively restricts the desired assignments on a formula.
+%\begin{exa}
+%\input{example-3dnf-minusone}
+%\end{exa}
+%
+%
+The proof that $\eqso(\logex{1})$ is closed under subtraction by one is the most involved in the paper.
+%We think the main technique used in this proof, which is based on considering some witnesses of logarithmic size, is of independent interest.
+A key insight in this proof is the fact that if a formula in $\eqso(\logex{1})$ has a satisfying assignment over a structure, then it also has a satisfying assignment over this structure that is of logarithmic size, and which can be characterized and removed by using some fixed formulae.
+To give more intuition about this idea, let us consider the case of $\ctdnf$, and the way Saluja et\text{.} al \cite{SalujaST95}  propose to encode this problem in $\E{1}$ and, thus, also in $\eqso(\logex{1})$.
+Let $\R = \{P_0(\cdot,\cdot,\cdot), P_1(\cdot,\cdot,\cdot), P_2(\cdot,\cdot,\cdot), P_3(\cdot,\cdot,\cdot),<\}$ and $\theta$ be a propositional formula in 3-DNF.
+Then $\theta$ is encoded as an $\R$-structure $\A_\theta$ as follows.
+The domain of $\A_\theta$ is the set of propositional variables occurring in $\theta$, and for every tuple $(a,b,c)$ of propositional variables occurring in $\theta$, we have that: (i) $P_0(a,b,c)$ holds if $(a\wedge b \wedge c)$ is a disjunct in $\theta$;
+(ii) $P_1(a,b,c)$ holds if $(a\wedge b \wedge \neg c)$ is a disjunct in $\theta$;
+(iii) $P_2(a,b,c)$ holds if $(a\wedge \neg b \wedge \neg c)$ is a disjunct in $\theta$; and 
+(iv) $P_3(a,b,c)$ holds if $(\neg a \wedge \neg b \wedge \neg c)$ is a disjunct in $\theta$.
+Moreover, to define $\ctdnf$, we consider a fixed $\so$-formula $\varphi(T)$ over $\R$, where $T$ is a unary predicate, such that the number of satisfying assignments of $\theta$ is equal to the evaluation of $\sa{T}\varphi(T)$ over $\A_\theta$.
+More specifically, $T(a)$ holds if and only if  $a$ is assigned value true, so that $\varphi(T)$ is defined as follows:
+\begin{align*}
+\varphi(T) = \ex{x}\ex{y}\ex{z}\big(\,
+&(P_0(x,y,z)\wedge T(x)\wedge  T(y)\wedge T(z))\ \vee\\
+&(P_1(x,y,z)\wedge T(x)\wedge  T(y)\wedge \neg T(z))\ \vee\\
+&(P_2(x,y,z)\wedge T(x)\wedge \neg T(y)\wedge \neg T(z))\ \vee\\
+&(P_3(x,y,z)\wedge \neg T(x)\wedge \neg T(y)\wedge \neg T(z))\big).
+\end{align*}
+Let us now focus on the first disjunct $(P_0(x,y,z)\wedge T(x)\wedge  T(y)\wedge T(z))$ of $\varphi(T)$.
+Assuming that a propositional formula $(a\wedge b \wedge c)$ is a disjunct in $\theta$, we have that $(\A_\theta,V) \models \varphi(T)$ for every assignment $V$ for $T$ such that $\{a,b,c\} \subseteq V(T)$.
+Notice that some of these assignments, such as the one that assigns to $T$ all the variables occurring in $\theta$, are of linear size in the size of $\A_\theta$.
+However, $\varphi(T)$ also admits satisfying assignments of logarithmic size in the size of $\A_\theta$, such as $V(T) = \{a,b,c\}$.
+A key idea in the proof is the fact that the minimum of such small witnesses (under the lexicographic order induced by~$<$) can be identified by using a fixed formula:
+\begin{multline*}
+\alpha(x,y,z) \ = \ P_0(x,y,z) \wedge  \fa{x'}\fa{y'}\fa{z'} \big(P_0(x',y',z')\ \to\\
+\big((x < x') \vee (x = x' \wedge y < y') \vee (x = x' \wedge y = y' \wedge z < z')\big)\big),
+\end{multline*}
+and then it can be removed also by using a fixed formula:
+\begin{align*}
+\beta(T) \ = \ \, & \ex{x} \ex{y} \ex{z} \big(P_0(x,y,z)\wedge T(x)\wedge  T(y)\wedge T(z)\big)\ \wedge \\
+& \ex{x'}\ex{y'}\ex{z'} \big(\alpha(x',y',z')\ \wedge \\
+& \hspace{23mm} (T(x')\wedge T(y')\wedge T(z')\to \ex{w}(T(w) \wedge w \neq x' \wedge w\neq y' \wedge w'\neq z))\big)
+\end{align*}
+In particular, the combination of these two formulae forces $T$ to represent a satisfying assignment for $\theta$ that is different from the set $\{a_0,b_0,c_0\}$, where $(a_0,b_0,c_0)$ is the minimum tuple under the lexicographic order induced by~$<$ on the set of tuples $(a,b,c)$ such that $(a \wedge b \wedge c)$ is a disjunct of $\theta$.
+In the proof we generalize and properly formalize this idea, thus using the existence of logarithmic size witnesses for the formulae in $\eqso(\logex{1})$ to prove that this class is closed under subtraction by one.
+We think that the existence of such witnesses is a fundamental property of this class that deserves to be further investigated.
+
+\martin{este ejemplo sirve para ayudar a entender la idea del testigo de tama\~no logar\'itmico}
+\marcelo{reescribi el ejemplo para hacerlo mas simple y concentrarnos en la propiedad de los testigos de tama\~no logaritmico, por favor verificar si se entiende}

### 2019-02-28 LMCS/sendrev/syntactic-clo.tex
pass on the 3-DNF example

@@
 However, $\varphi(T)$ also admits satisfying assignments of logarithmic size in the size of $\A_\theta$, such as $V(T) = \{a,b,c\}$.
-A key idea in the proof is the fact that the minimum of such small witnesses (under the lexicographic order induced by~$<$) can be identified by using a fixed formula:
+A key idea in the proof is the fact that the minimum of such small witnesses (under the lexicographic order induced by~$<$) can be identified by using a fixed formula in $\logex{1}$:
 \begin{multline*}
@@
 \end{multline*}
-and then it can be removed also by using a fixed formula:
+and then it can be removed also by using a fixed formula in $\logex{1}$:
 \begin{align*}
@@
 & \ex{x'}\ex{y'}\ex{z'} \big(\alpha(x',y',z')\ \wedge \\
-& \hspace{23mm} (T(x')\wedge T(y')\wedge T(z')\to \ex{w}(T(w) \wedge w \neq x' \wedge w\neq y' \wedge w'\neq z))\big)
+& \hspace{23mm} (T(x')\wedge T(y')\wedge T(z')\to \ex{w}(T(w) \wedge w \neq x' \wedge w\neq y' \wedge w'\neq z))\big).
 \end{align*}
 In particular, the combination of these two formulae forces $T$ to represent a satisfying assignment for $\theta$ that is different from the set $\{a_0,b_0,c_0\}$, where $(a_0,b_0,c_0)$ is the minimum tuple under the lexicographic order induced by~$<$ on the set of tuples $(a,b,c)$ such that $(a \wedge b \wedge c)$ is a disjunct of $\theta$.
-In the proof we generalize and properly formalize this idea, thus using the existence of logarithmic size witnesses for the formulae in $\eqso(\logex{1})$ to prove that this class is closed under subtraction by one.
+In the proof, we generalize and properly formalize this idea, thus using the existence of logarithmic size witnesses for the formulae in $\eqso(\logex{1})$ to prove that this class is closed under subtraction by one.
 We think that the existence of such witnesses is a fundamental property of this class that deserves to be further investigated.

### 2019-02-28 LMCS/sendrev/proof-ehorn-pe.tex
pass on section 5

--- 2018-12-09 LMCS/proof-ehorn-pe.tex
@@
+%!TEX root = main.tex
+%
 %Let $\shdhsat$ be the problem of counting the number of satisfying assignments of a disjunction of Horn formulas. 
-In~\cite{PagourtzisZ06}, Pagourtzis and Zachos gave a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula.
+%In~\cite{PagourtzisZ06}, Pagourtzis and Zachos gave a $\totp$ procedure that computes the number of satisfying assignments of a DNF formula.
 This procedure can be easily extended to receive Horn formulas, and furthermore, a disjunction of Horn formulas.
 Hence $\shdhsat$ is in $\totp$.
-As we saw in Theorem~\ref{sigma2hard}, $\shdhsat$ is complete for $\eqso(\ehorn)$ under parsimonious reductions.
+%
+As we mentioned before, $\totp$ is closed under parsimonious reductions, so we only need to show that 
+%First, we prove that 
+$\shdhsat$ is in $\totp$.
+For this consider a non-deterministic procedure that receives a $\dhsat$ formula $\Phi$ as input and does the following.
+First it checks whether $\Phi$ is satisfiable.
+If it is not, then it stops; otherwise, it creates a dummy branch that simply stops, and continues in the main branch.
+More precisely, it picks in the main branch a propositional variable $x$ in $\Phi$ and creates two formulae $\Phi_0$ and $\Phi_1$, where $x$ has been replaced by $\perp$ and $\top$, respectively.
+If only one of these is satisfiable, it continues on this branch with the respective $\Phi_i$, and if both are satisfiable, then it creates a new branch for $\Phi_1$ and continues on this branch with $\Phi_0$.
+On each branch, it repeats the same instructions until no variables are left to replace.
+Since $\dhsat$ is in $\ptime$, all the aforementioned checks can be done in polynomial time, so that the procedure takes at most $h(n)$ steps in each branch, where $n$ is the size of $\Phi$ and $h$ is some fixed polynomial.
+Moreover, the algorithm produces exactly $\shdhsat(\Phi)+1$ branches, from which we conclude that $\shdhsat$ is in $\totp$.
+%For the next part of the proof, notice that in Theorem~\ref{sigma2hard} we establish that $\shdhsat$ is complete for $\eqso(\ehorn)$ under parsimonious reductions.
 Let $\alpha$ be a formula in $\eqso(\ehorn)$ and let $g_{\alpha}$ be the reduction to $\shdhsat$.
@@
 Therefore, we conclude that $\alpha$ defines a function in~$\totp$.
+\martin{agregu\'e una demostraci\'on de que $\shdhsat$ esta en $\totp$}
+\marcelo{simplifique la demostracion, por favor revisar}

### 2019-02-28 LMCS/sendrev/proof-sigma2hard.tex
pass on section 5

@@
 
-Next, we prove that $\shdhsat$ is hard for $\eqso(\ehorn)$ over any signature~$\R$ under parsimonious reductions.
-For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we will define a polynomial-time function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
+Next, we prove that $\shdhsat$ is hard for $\eqso(\ehorn)$ over each signature~$\R$ under parsimonious reductions.
+For each $\eqso(\ehorn)$ formula $\alpha$ over $\R$, we define a polynomial-time function $g_{\alpha}$ that receives an $\R$-structure $\A$ and outputs an instance of $\shdhsat$ such that $\sem{\alpha}(\A) = \shdhsat(g_{\alpha}(\A))$.
 By Proposition~\ref{theo-pnf-snf}, we can assume that $\alpha$ is of the form:
@@
 $$
-\sa{\bar{X}}\sa{\bar{x}} \ex{\bar{y}} \bigwedge_{j = 1}^{n} \fa{\bar{z}} \varphi^i_j(\bar{X}_{i},\bar{x},\bar{y},\bar{z})\wedge\bigwedge_{X\not\in\bar{X}_{i}}\!\!\fa{\bar{u}}X(\bar{u}),
+\sa{\bar{X}}\sa{\bar{x}} \ex{\bar{y}} \bigg( \bigwedge_{j = 1}^{n} \fa{\bar{z}} \varphi^i_j(\bar{X}_{i},\bar{x},\bar{y},\bar{z})\wedge\bigwedge_{X\not\in\bar{X}_{i}}\!\!\fa{\bar{u}}X(\bar{u})\bigg),
 $$
@@
 The next transformation of $\alpha$ and $\A$ towards a disjunction of Horn-formulae is to expand each first-order quantifier (i.e.
-$\Sigma{\bar{x}}$,  $\exists\bar{y}$, and $\forall\bar{z}$) as we replace their variables with first-order constants.
-Specifically, we obtain the following formula which defines the same function as $\alpha$ and is of polynomial size with respect to $\A$:
+$\Sigma{\bar{x}}$,  $\exists\bar{y}$, and $\forall\bar{z}$) by replacing variables with constants.
+%their variables with first-order constants. 
+More specifically, we obtain the following formula that defines the same function as $\alpha$, and it is of polynomial size in the size of $\A$ (recall that $\alpha$ is fixed):
 $$
-\alpha_{\A} \ = \ \sum_{i = 1}^{m} \sum_{\bar{a}\,\in A^{\length{\bar{x}}}} \sa{\bar{X}}\bigvee_{\bar{b}\,\in A^{\length{\bar{y}}}}\bigwedge_{j = 1}^n\bigwedge_{\bar{c}\,\in A^{\length{\bar{z}}}}\varphi^i_j(\bar{X}_i,\bar{a},\bar{b},\bar{c})\wedge \bigwedge_{X \not\in \bar{X}_i}\bigwedge_{\bar{e}\,\in A^{\arity(X)}}\!\!\!\!\!\!\!
-X(\bar{e}).
+\alpha_{\A} \ = \ \sum_{i = 1}^{m} \sum_{\bar{a}\,\in A^{\length{\bar{x}}}} \sa{\bar{X}}\bigvee_{\bar{b}\,\in A^{\length{\bar{y}}}}\bigg(\bigwedge_{j = 1}^n\bigwedge_{\bar{c}\,\in A^{\length{\bar{z}}}}\varphi^i_j(\bar{X}_i,\bar{a},\bar{b},\bar{c})\wedge \bigwedge_{X \not\in \bar{X}_i}\bigwedge_{\bar{e}\,\in A^{\arity(X)}}\!\!\!\!\!\!\!
+X(\bar{e})\bigg).
 $$
 \martin{reescribi esta formula}
-Note that each first-order subformula in $\varphi^i_j(\bar{X}_i,\bar{a},\bar{b},\bar{c})$ has no free variables and, therefore, we can evaluate each of them in polynomial time and easily rewrite $\alpha_{\A}$ to an equivalent formula that does not have any first-order subformula.
-In other words, in polynomial time we can replace $\varphi^i_j$ with a disjunction of $\neg X_{\ell}$ and at most one $X_{\ell}$, evaluated on constants.
+\marcelo{OK. Le agregue los parentesis grandes a esta formula y la anterior para que quedara mas claro el scope del cuantificador existencial y la disyuncion}
+Notice that each first-order subformula in $\varphi^i_j(\bar{X}_i,\bar{a},\bar{b},\bar{c})$ has no free variables and, therefore, we can evaluate each of them in polynomial time and easily rewrite $\alpha_{\A}$ to an equivalent formula that does not have any first-order subformula.
+In other words, in polynomial time we can replace $\varphi^i_j$ with a disjunction of negative literals of the form $\neg X_{\ell}(\bar d)$ and at most one positive literal of the form $X_{\ell}(\bar d)$, where $\bar d$ is a tuple of constants.
 After simplifying, grouping and reordering the previous formula, we can obtain an equivalent formula $\alpha_{\A}'$ of the form:
 $$
-\alpha_{\A}' \ := \ \sum_{i = 1}^{m'}\sa{\bar{X}}\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi^{i}_{j,k}(\bar{X})
+\alpha_{\A}' \ = \ \sum_{i = 1}^{m'}\sa{\bar{X}}\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi^{i}_{j,k}(\bar{X}),
 $$
-where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of $\neg X_{\ell}$ and at most one $X_{\ell}$.
-The reader can easily check that $\sem{\alpha}(\A) = \sem{\alpha_{\A}'}(\A)$.
+where every $\psi^{i}_{j,k}(\bar{X})$ is a disjunction of 
+%$\neg X_{\ell}$ and at most one $X_{\ell}$.
+negative literals of the form $\neg X_{\ell}(\bar d)$ and at most one positive literal of the form $X_{\ell}(\bar d)$, where $\bar d$ is a tuple of constants. 
+%It is straightforward to check that~$\sem{\alpha}(\A) = \sem{\alpha_{\A}'}(\A)$.
 
 The idea for the rest of the proof is to show how to obtain $g_{\alpha}(\A)$, i.e. an instance of $\shdhsat$, from $\alpha_{\A}'$.
-First, if $\alpha_{\A}'$ is equal to $\sa{\bar{X}}\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi_{j,k}(\bar{X})$, then we can define $g_{\alpha}(\A)$ equal to the propositional formula $\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$ which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulae.
-Otherwise, if $m' > 1$ we can use $m'$ fresh new variables $t_1,\ldots,t_{m'}$ and define:
+First, if $m' = 1$ and $\alpha_{\A}' = \sa{\bar{X}}\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi_{j,k}(\bar{X})$, then we can define $g_{\alpha}(\A)$ as the propositional formula $\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi_{j,k}(\bar{X})$ over the propositional alphabet $\{X(\bar{e}) \mid X \in \bar{X} \text{ and } \bar{e}\in A^{\arity(X)} \}$.
+It is straightforward to see that $\bigvee_{j = 1}^{n_1'}\bigwedge_{k = 1}^{n_2'}\psi_{j,k}(\bar{X})$ is a disjunction of Horn formulae, and its number of satisfying assignments is exactly $\sem{\alpha}(\A)$.
+%which has exactly $\sem{\alpha}(\A)$ satisfying assignments and is precisely a disjunction of Horn formulae.
+Otherwise, if $m' > 1$, then we can use $m'$ fresh new variables $t_1,\ldots,t_{m'}$ and define:
 $$
@@
 Variables $t_1,\ldots,t_{m'}$ are used to have disjoint sets of propositional assignments for the different disjuncts of the outermost disjunction, which correspond to the summands in the original formula.
-One can easily check that $g_{\alpha}(\A)$ is a disjunction of Horn formulae, and the number of satisfying assignments is exactly $\sem{\alpha}(\A)$.
+One can easily check that $g_{\alpha}(\A)$ is a disjunction of Horn formulae, and that the number of satisfying assignments of $g_{\alpha}(\A)$ is exactly $\sem{\alpha}(\A)$.
 This covers all possible cases for $\alpha$, and the entire procedure takes polynomial time.
+
+\marcelo{Hice algunos cambios en esta demostracion (teorema \ref{sigma2hard}), por favor revisen con cuidado si todo esta bien}

### 2019-02-28 LMCS/sendrev/syntactic-clo.tex
pass on section 5

@@
 (iv) $P_3(a,b,c)$ holds if $(\neg a \wedge \neg b \wedge \neg c)$ is a disjunct in $\theta$.
-Moreover, to define $\ctdnf$, we consider a fixed $\so$-formula $\varphi(T)$ over $\R$, where $T$ is a unary predicate, such that the number of satisfying assignments of $\theta$ is equal to the evaluation of $\sa{T}\varphi(T)$ over $\A_\theta$.
-More specifically, $T(a)$ holds if and only if  $a$ is assigned value true, so that $\varphi(T)$ is defined as follows:
+Moreover, to define $\ctdnf$, we consider a fixed $\so$-formula $\varphi(T)$ over $\R$, where $T$ is a unary predicate, such that the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{T}\varphi(T)}(\A_\theta)$.
+More specifically, $T(a)$ holds if and only if  $a$ is assigned value true, so that $\varphi(T)$ is defined as:
 \begin{align*}
-\varphi(T) = \ex{x}\ex{y}\ex{z}\big(\,
+\varphi(T) \ := \ \ex{x}\ex{y}\ex{z}\big(\,
 &(P_0(x,y,z)\wedge T(x)\wedge  T(y)\wedge T(z))\ \vee\\
@@
 \begin{multline*}
-\alpha(x,y,z) \ = \ P_0(x,y,z) \wedge  \fa{x'}\fa{y'}\fa{z'} \big(P_0(x',y',z')\ \to\\
+\alpha(x,y,z) \ := \ P_0(x,y,z) \wedge  \fa{x'}\fa{y'}\fa{z'} \big(P_0(x',y',z')\ \to\\
 \big((x < x') \vee (x = x' \wedge y < y') \vee (x = x' \wedge y = y' \wedge z < z')\big)\big),
@@
 \begin{align*}
-\beta(T) \ = \ \, & \ex{x} \ex{y} \ex{z} \big(P_0(x,y,z)\wedge T(x)\wedge  T(y)\wedge T(z)\big)\ \wedge \\
+\beta(T) \ := \ \, & \ex{x} \ex{y} \ex{z} \big(P_0(x,y,z)\wedge T(x)\wedge  T(y)\wedge T(z)\big)\ \wedge \\
 & \ex{x'}\ex{y'}\ex{z'} \big(\alpha(x',y',z')\ \wedge \\

### 2019-02-28 LMCS/sendrev/syntactic-horn.tex
pass on section 5

@@
 To define $\chsat$, we consider an \so-formula $\varphi(\pT)$ over $\R$, where $\pT$ is a unary predicate, such that for every Horn formula $\theta$ encoded by an $\R$-structure $\A$, the number of satisfying assignments of $\theta$ is equal to $\sem{\sa{\pT} \varphi(\pT)}(\A)$.
-In particular, $\pT(x)$ holds if, and only if, $x$ is a propositional variable that is assigned value true.
+In particular, $\pT(x)$ holds if and only if $x$ is a propositional variable that is assigned value true.
 More specifically,
@@
 %The resulting formula $\varphi(\pT)$ is not in $\uhorn$, but it can be easily transformed into a formula in this class  by introducing an auxiliary binary predicate $\pA$ defined as follows:
-Moreover, by introducing an auxiliary predicate $\pA$ which will be defined as it appears in the formula
+Moreover, by introducing an auxiliary predicate $\pA$ defined as 
 \begin{align*}
@@
 \end{prop}
-For the sake of simplicity, we postpone the proof of the previous proposition to after the proof of Theorem~\ref{sigma2hard}.
+In this section, we identify a complete problem for $\eqso(\ehorn)$ under parsimonious reductions.
+Hence, to prove that $\eqso(\ehorn) \subseteq \totp$, it is enough to prove that such a problem is in $\totp$, as $\totp$ is closed under parsimonious reductions.
+We give this proof at the end of this section, after the complete problem has been identified.
+%For the sake of simplicity, we postpone the proof of the previous proposition to after the proof of Theorem~\ref{sigma2hard}.
 
@@
 \qed
-We conclude this section by showing that $\eqso(\ehorn)$ is the class we were looking for, as not only every function in $\eqso(\ehorn)$ has an easy decision counterpart, but also $\eqso(\ehorn)$ admits a natural complete problem under parsimonious reductions.
+%We conclude this section by showing 
+Next we show that $\eqso(\ehorn)$ is the class we were looking for, as not only every function in $\eqso(\ehorn)$ has an easy decision counterpart, but also $\eqso(\ehorn)$ admits a natural complete problem under parsimonious reductions.
 More precisely, define
@@
 \qed
+Now that we have a complete problem for $\eqso(\ehorn)$, we can provide a simple proof of Proposition~\ref{prop:ehorn-pe}.
+
+\medskip
 

### 2019-02-28 LMCS/sendrev/beyond.tex
Solving some pending issues of reviewer 1

@@
 We would like to define an $\rqfo(\fo)$-formula that, given a directed acyclic graph $G$ with $n$ nodes and a pair of nodes $b$, $c$ in $G$, counts the number of paths of length less than $n$ from $b$ to $c$ in $G$.
-To this end, assume that graphs are encoded using the signature $\R = \{ E(\cdot,\cdot)\}$, and then define formula $\alpha(x, y, f)$ as follows:
+To this end, assume that graphs are encoded using the signature $\R = \{ E(\cdot,\cdot),<\}$, and then define formula $\alpha(x, y, f)$ as follows:
 $$ 

### 2019-02-28 LMCS/sendrev/introduction.tex
un par de correcciones y algunos problemas de compatibilidad con el la instruccion !TEX = main
Solving some pending issues of reviewer 1

@@
 Specifically, we propose to restrict WL to the natural numbers as a fixed semiring, calling this restriction Quantitative Second Order Logics (QSO), and study its expressive power for defining counting complexity classes over ordered structures.
-As a proof of concept, we show that natural syntactical fragments of $\qso$ captures counting complexity classes like $\shp$, $\spp$, $\fp$ and $\fpspace$.
+As a proof of concept, we show that natural syntactical fragments of $\qso$ capture counting complexity classes like $\shp$, $\spp$, $\fp$ and $\fpspace$.
 Furthermore, by slightly extending the framework we can prove that $\qso$ can also capture classes like $\gp$ and $\optp$, showing the robustness of our approach.

### 2019-02-28 LMCS/sendrev/todo-list.tex
un par de correcciones y algunos problemas de compatibilidad con el la instruccion !TEX = main
Solving some pending issues of reviewer 1

@@
 a formula $\beta \in \eqso(\logex{1})$ ...''
-\item p.25, last sentence: ``which is based on considering some witnesses of logarithmic size'' could
+\item[$\checkmark$] p.25, last sentence: ``which is based on considering some witnesses of logarithmic size'' could
 be explained a little more / made more clear
-\martin{no logr\'e completar esta correcci\'on}
+\marcelo{Esto quedo listo con el ejemplo que puso Martin y la pasada que hice sobre el}
+
 \item[$\checkmark$] p.26: ``... auxiliary predicate $A$ defined as'' $\to$ ``... auxiliary predicate $A$ which will be defined
@@
 	\item[$\checkmark$] p.1, footnote: ``clases'' should be ``classes''
-	\item throughout the paper: plural used with verb ending on s. Examples:
+	\item[$\checkmark$] throughout the paper: plural used with verb ending on s. Examples:
 	\begin{itemize}
-	\item[-] p.2, middle: ``fragments of QSO captures''
-	\item[-] p.6, example 3.1: ``the sum quantifiers in $\alpha_1$ aggregates''
+	\item[$\checkmark$] p.2, middle: ``fragments of QSO captures''
+	\item[$\checkmark$] p.6, example 3.1: ``the sum quantifiers in $\alpha_1$ aggregates''
 	\end{itemize}
 	\martin{no alcanc\'e a revisar esta ni la de cuatro puntos m\'as adelante}
+	\marcelo{corregi estos errores en el paper}
 	\item[$\checkmark$] p4, beginning of 2.2: ``some of \underline{the} are recalled here''
@@
 \end{itemize}
-	\item throughout the paper: singular used with verb not ending on s. Examples:
+	\item[$\checkmark$] throughout the paper: singular used with verb not ending on s. Examples:
 	\begin{itemize}
-	\item[-] p.9, after proof of Prop. 4.3: ``use of a fragment [...] that capture''
-	\item[-] p.11, end of proof of Thm. 4.5: ``and $\alpha$ reconstruct''
+	\item[$\checkmark$] p.9, after proof of Prop. 4.3: ``use of a fragment [...] that capture''
+	\item[$\checkmark$] p.11, end of proof of Thm. 4.5: ``and $\alpha$ reconstruct''
 \end{itemize}
 	\martin{lo que menciono arriba}
+	\marcelo{tambien corregi esto}
 	\item[$\checkmark$] p.9, middle: ``A second attempt could be based then'' $\to$ ``A second attempt could then be
@@
 	\item[$\checkmark$] on definition of functions, \verb|\colon| should be used instead of :
-	\item In [1], a similar hierarchy to the one by Saluja et al. was introduced and studied. This work should be cited.
+	\item In \cite{DBLP:conf/csl/0001HKV16}, a similar hierarchy to the one by Saluja et al. was introduced and studied. This work should be cited.
 \end{itemize}
@@
 
+\bibliographystyle{plain}
+\bibliography{0001HKV16}
+
 \end{document}

### 2019-03-01 LMCS/sendrev/response-letter.tex
un par de correcciones y algunos problemas de compatibilidad con el la instruccion !TEX = main
Revisión de Reviewer 2.

@@
 %	\item on definition of functions, \verb|\colon| should be used instead of :
-	\item In [1], a similar hierarchy to the one by Saluja et al. was introduced and studied. This work should be cited.
+	\item In \cite{DBLP:conf/csl/0001HKV16}, a similar hierarchy to the one by Saluja et al. was introduced and studied. This work should be cited.
 \end{itemize}
@@
 	\setlength\itemsep{0.5em}
-	\item In several proofs, you rely on disjoint union of solutions and cartesian product of solutions.
+	\item \red{In several proofs, you rely on disjoint union of solutions and cartesian product of solutions.
 Indeed, its gives a polynomial time counting reduction.
@@
 It would be similar to what is done in A Circuit-Based Approach to Efficient Enumeration (Antoine Amarilli, Pierre Bourhis, Louis Jachiet, Stefan Mengel) where simple circuits (dDNNF, which are basically disjoint unions and cartesian products) are used to capture the solutions of MSO queries.
+In general I would be interested to know the relationship of your couting classes with their enumeration counterpart.
+	For instance there is a paper adapting [39] to enumeration: Enumeration Complexity of Logical Query Problems with Second-order Variables (Arnaud Durand and Yann Strozecki).
+In particular they consider class with some Horn structure to capture easy enumeration problems, it may have some connection with the class $\sigma$ QSO($\Pi_1$ Horn) you introduce.}
 
-	\item In general I would be interested to know the relationship of your couting classes with their enumeration counterpart. 
-	For instance there is a paper adapting [39] to enumeration: Enumeration Complexity of Logical Query Problems with Second-order Variables (Arnaud Durand and Yann Strozecki).
-In particular they consider class with some Horn structure to capture easy enumeration problems, it may have some connection with the class $\sigma$ QSO($\Pi_1$ Horn) you introduce.
+	We thank the reviewer for showing us the literature regarding enumeration problems, in particular, for the paper from Durand and Strozecki.
+Unfortunately, the relation between our framework with enumeration complexity classes goes beyond the scope of the paper.
+Nevertheless, in the last section we added a final paragraph pointing out this connection as an interesting open question.
 
-	\item For the most involved proofs, you could describe informally your proof strategy in a few lines to help the reader.
-You may also add some example to illustrate the different fragment of QSO (you already do it sometimes and it is very helpful).
+	\item \red{For the most involved proofs, you could describe informally your proof strategy in a few lines to help the reader.
+You may also add some example to illustrate the different fragment of QSO (you already do it sometimes and it is very helpful).}
+
+	Although we agree that adding more proof sketches and examples would help, the paper is already too long for adding more additional content.
+Furthermore, as the reviewer mentioned we already have this for the main definitions and proofs (e.g.
+proof of substraction by 1 at Theorem 5.9).
+For this reason, we decided to keep the paper as it is in terms of more examples and proof sketches.
 
@@
 %	\item page 24: 	if, if $\phi$$\to$ $\phi$
-%	\item page 25: 	-you explain that the proof relies on witnesses of logarithmic size.	It was not clear to me while reading the proof. Can you explain more ?
-	\item page 28:	-the complete problem for $\Sigma$ QSO($\Sigma_2$ Horn) would be complete also for $\#\Sigma-2$ Horn.
-You should mention here that you do not need sum here	and that the framework of [39] is enough for this result.
+	\item \red{page 25: 	-you explain that the proof relies on witnesses of logarithmic size.	It was not clear to me while reading the proof. Can you explain more ?}
+
+	This is discussed in the response to Reviewer 1.
+
+	\item \red{page 28:	The complete problem for $\Sigma\text{QSO}(\Sigma_2\text{-Horn})$ would be complete also for $\#(\Sigma_2\text{-Horn})$.
+You should mention here that you do not need sum here and that the framework of [39] is enough for this result.}
+
+	Indeed, from the proof of Theorem 5.14 one can also derive that $\#(\Sigma_2\text{-Horn})$ is equivalent to $\Sigma\text{QSO}(\Sigma_2\text{-Horn})$ and, thus, the framework of [39] is enough for defining this class of problems.
+At the end of Section 5, we added a paragraph mentioning this.
+
 %	\item[?] page 30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
@@
 %	\item page 33: 	-there is a proof symbol \verb|\qed| but it does not end a well defined proof environment
-	\item page 38: the last sentence on fixed point operator is to vague. Either expand it to say what king of question on fixed point operator you want to investigate or remove it	
+	\item \red{page 38: the last sentence on fixed point operator is to vague. Either expand it to say what king of question on fixed point operator you want to investigate or remove it.}
+
+	We agree with the reviewer and we decided to remove this sentence. 
 \end{itemize}
 
+\bibliographystyle{plain}
+\bibliography{0001HKV16}
+
 \end{document}

### 2019-03-01 LMCS/sendrev/todo-list.tex
Revisión de Reviewer 2.

@@
 
+\usepackage[utf8]{inputenc}
+\usepackage[english]{babel}
 \usepackage{fullpage}
@@
 \martin{me gustar\'ia tener feedback para saber si mi correcci\'on est\'a bien}
-\item[$\checkmark$] p.33: ``executes exactly $n^k$ steps for a fixed $k \geq 1$'' → ``executes exactly $n^k$
+\item[$\checkmark$] p.33: ``executes exactly $n^k$ steps for a fixed $k \geq 1$'' $\rightarrow$ ``executes exactly $n^k$
 steps on large
@@
 		\setlength\itemsep{0.5em}
-\item In several proofs, you rely on disjoint union of solutions and cartesian product of solutions.
+\item[$\checkmark$] In several proofs, you rely on disjoint union of solutions and cartesian product of solutions.
 Indeed, its gives a polynomial time counting reduction.
@@
 
-\item In general I would be interested to know the relationship of your couting classes with their enumeration counterpart. 
+\item[$\checkmark$] In general I would be interested to know the relationship of your couting classes with their enumeration counterpart. 
 For instance there is a paper adapting [39] to enumeration: Enumeration Complexity of Logical Query Problems with Second-order Variables (Arnaud Durand and Yann Strozecki).
@@
 
-\item For the most involved proofs, you could describe informally your proof strategy in a few lines to help the reader.
+\cristian{No tiene mucho sentido estos comentarios del reviewer ya que va más alla del paper.
+Para dejarlo tranquilo, agregue un último comentario en las conclusiones diciendo que sería interesante estudiar estos problemas en el contexto de enumeración.
+También agregue la respuesta a la carta.}
+
+\item[$\checkmark$] For the most involved proofs, you could describe informally your proof strategy in a few lines to help the reader.
 You may also add some example to illustrate the different fragment of QSO (you already do it sometimes and it is very helpful).
 \martin{me gustar\'ia que conversemos qu\'e demostraciones necesitan esto, y si el outline de la demostraci\'on de -1 basta. Tambi\'en qu\'e fragmentos de QSO necesitan un ejemplo.}
+
+\cristian{El paper ya tiene demasiados ejemplos y ya esta demasiado largo. En la carta le respondí que lamentablemente, debido a que el paper esta demasiado largo, decidimos no agregar mas ejemplos.}
 
@@
 if, if $\phi$$\to$ $\phi$
-\item page 25: 
+\item[$\checkmark$] page 25: 
 -you explain that the proof relies on witnesses of logarithmic size.
@@
 \martin{no logr\'e corregir esto, es la misma acotaci\'on que el Reviewer 1}
-\item page 28:
+\cristian{Le respondí diciendo que es mencionado en la respuesta al otro reviewer.}
+\item[$\checkmark$] page 28:
 -the complete problem for $\Sigma$ QSO($\Sigma_2$ Horn) would be complete also for $\#\Sigma-2$ Horn. You should mention here that you do not need sum here
@@
 \martin{conversemos esto}
+\cristian{Agregue un parrafo al final de la sección 5 mencionando esto.
+De hecho, desde la demostración uno puede demostrar que las dos classes son iguales (la suma de afuera no te agrega expresividad).
+Le respondí esto mismo en la carta.}
 \item[$\checkmark$] page 30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
@@
 
-\item page 38: 
+\item[$\checkmark$] page 38: 
 the last sentence on fixed point operator is to vague. Either expand it to 
@@
 \martin{conversemos esto}
+\cristian{Tiene razón y eliminé esta oración al final del paper.}
 \end{itemize}

### 2019-03-01 LMCS/sendrev/response-letter.tex
sc

@@
 	The problem with the first formula for directed graphs is precisely the opposite.
-The fixed point could be reached {\em before} we want it to.
-The fixed point is always reached after $k$ steps where $k$ is the size of the maximum shortest path between two vertices in the graph, and if the graph has a cycle in it, there might be paths of size $k' \in [k+1,n]$ that wouldn't be counted because the fixed point was reached earlier.
+The fixed point could be reached too soon without counting all paths of length at most $n-1$.
+The fixed point is always reached after $k$ steps, where $k$ is the size of the maximum shortest path between two vertices in the graph.
+If the graph has a cycle in it, there might be paths of size $k' \in \{k+1,\ldots,n-1\}$ that would not be counted because the fixed point was reached earlier.
+That is why we need a different formula for the general case.
+We added a comment about this in the paper.
 

### 2019-03-01 LMCS/sendrev/beyond.tex
sc

--- 2019-01-22 LMCS/sendrev/beyond.tex
@@
 We would like to define an $\rqfo(\fo)$-formula that, given a directed acyclic graph $G$ with $n$ nodes and a pair of nodes $b$, $c$ in $G$, counts the number of paths of length less than $n$ from $b$ to $c$ in $G$.
-To this end, assume that graphs are encoded using the signature $\R = \{ E(\cdot,\cdot)\}$, and then define formula $\alpha(x, y, f)$ as follows:
-$$ 
+To this end, assume that graphs are encoded using the signature $\R = \{ E(\cdot,\cdot),<\}$, and then define formula $\alpha(x, y, f)$ as follows:
+\begin{eqnarray}\label{eq-count-paths-acyclic}
 E(x,y) + \sa{z} f(x,z)\cdot E(z,y).
-$$
+\end{eqnarray}
 We have that $\clfp{\alpha(x,y,f)}$ defines our counting function.
@@
 
-Assume now that we need to extend our previous counting function to the case of arbitrary directed graphs.
-To this end, suppose that $\varphi_{\text{\rm first}}(x)$ and $\varphi_{\text{succ}}(x,y)$ are $\fo$-formulae defining the first element of $<$ and the successor relation associated to $<$, respectively.\footnote{Recall that in this paper we consider ordered $\R$-structures.} Moreover, define formula $\beta(x, y, t, g)$ as follows:
+Assume now that we need to count the number of paths of length less than $n$ from a node $b$ to a node $c$ in a directed graph that is not necessarily acyclic. 
+%extend our previous counting function to the case of arbitrary directed graphs. 
+Then we cannot use formula \eqref{eq-count-paths-acyclic}, as the fixed point could be reached too soon without counting all paths of length at most $n-1$.
+The fixed point of \eqref{eq-count-paths-acyclic} is always reached after $k$ steps, where $k$ is the size of the maximum shortest path between two vertices in the graph.
+If the graph has a cycle in it, there might be paths of size $k' \in \{k+1,\ldots,n-1\}$ that would not be counted because the fixed point was reached earlier.
+Thus, we need a more involved formula in the general case, which is given below.
+
+%To this end, 
+Suppose that $\varphi_{\text{\rm first}}(x)$ and $\varphi_{\text{succ}}(x,y)$ are $\fo$-formulae defining the first element of $<$ and the successor relation associated to $<$, respectively.
+%\footnote{Recall that in this paper we consider ordered $\R$-structures.} 
+Moreover, define formula $\beta(x, y, t, g)$ as follows:
 \begin{align*}
@@
 Notice that the use of the filter $\varphi_{\text{succ}}(t',t)$ prevents this formula for incrementing the value of $g(x,y,t)$ when $t$ is the first element in the domain.
+\qed
 \end{exa}
 \martin{El Reviewer 1 hizo un comentario de este ejemplo, lo respond\'i en la carta}
+\marcelo{Me parece bien la respuesta en la carta. Agregue esta respuesta en el paper, ayuda a entender mejor el ejemplo.}
 In contrast to $\lfp$, to reach a fixed point we do not need to impose any positive restriction on the formula $\beta(\x,h)$.

### 2019-03-01 LMCS/sendrev/todo-list.tex
sc

@@
 elements of $A^{\ell}$, not $\{1,\ldots,\ell\}$
-\item[?] p.32/33, example 6.2: I don't see why the formula given for dag's does not work for arbitrary
+\item[$\checkmark$] p.32/33, example 6.2: I don't see why the formula given for dag's does not work for arbitrary
 graphs: In arbitrary graphs, the fixpoint is still reached after at most $n-1$ steps, since the
@@
 \martin{me gustar\'ia tener feedback para saber si mi correcci\'on est\'a bien}
+\marcelo{Esta bien la respuesta en la carta. La agregue al ejemplo para que quedara mas claro}
 \item[$\checkmark$] p.33: ``executes exactly $n^k$ steps for a fixed $k \geq 1$'' $\rightarrow$ ``executes exactly $n^k$

### 2019-03-02 LMCS/sendrev/syntactic-clo.tex
adding a comment about complete problems for totp

@@
 Besides, $\totp$ has good closure properties as it is closed under sum, multiplication and subtraction by one.
-However, some functions in $\totp$ do not admit FPRAS under standard complexity-theoretical assumptions\footnote{As an example consider the problem of counting the number of independent sets in a graph, and the widely believed assumption that $\np$ is not equal to the randomized complexity class $\rp$ (Randomized Polynomial-Time \cite{G77}).
-This counting problem is in $\totp$, and it is known that $\np = \rp$ if there exists an FPRAS for it \cite{DFJ02}.}, and no natural complete problems are known for this class \cite{PagourtzisZ06}.
+However, some functions in $\totp$ do not admit FPRAS under standard complexity-theoretical assumptions.\footnote{As an example consider the problem of counting the number of independent sets in a graph, and the widely believed assumption that $\np$ is not equal to the randomized complexity class $\rp$ (Randomized Polynomial-Time \cite{G77}).
+This counting problem is in $\totp$, and it is known that $\np = \rp$ if there exists an FPRAS for it \cite{DFJ02}.}
+%, and no natural complete problems are known for this class \cite{PagourtzisZ06}. 
 Hence, we use the $\eqso(\fo)$-hierarchy to find restrictions of $\totp$ with good approximation and closure properties.

### 2019-03-02 LMCS/sendrev/syntactic-horn.tex
Revisión de Reviewer 2.
adding a comment about complete problems for totp

@@
 Unfortunately, it is not clear whether it admits a {\em natural} complete problem under parsimonious reductions, where {\em natural} means any of the counting problems defined in this section or any other well-known counting problem (not one specifically designed to be complete for the class).
-Hence, in this section we follow a different approach to find a class of functions in $\shp$ with easy decision counterparts and natural complete problems, which is inspired by the approach followed in \cite{G92} that uses a restriction of second-order logic to Horn clauses for capturing $\ptime$ (over ordered structures).
+On the other hand, $\totp$ admits a natural complete problem under parsimonious reductions, which is the problem of counting the number of inputs accepted by a monotone circuit~\cite{BCPPZ17}.
+However, the notion of monotone circuit used in \cite{BCPPZ17} does not correspond with the usual notion of monotone circuit~\cite{GS90}, that is, circuits with AND and OR gates but without negation.
+In this sense, we still lack a class of functions in $\shp$ with easy decision counterparts and a complete problem that is well known and has been widely studied.
+In this section, we follow a different approach to find such a class,
+%Hence, in this section we follow a different approach to find a class of functions in $\shp$ with easy decision counterparts and natural complete problems, 
+which is inspired by the approach followed in \cite{G92} that uses a restriction of second-order logic to Horn clauses for capturing $\ptime$ (over ordered structures).
 The following example shows how our approach works.
+\marcelo{Cambie este parrafo para mencionar un resultado nuevo de Pagourtzis et al. donde muestran un problema completo para Totp, que es un poco mas natural}
 \begin{exa} \label{ex-hornsat-esop1}
@@
 \qed
+
+Finally, it is important to mention that from the previous proof one can easily derive that $\eqso(\ehorn) \equiv \#(\ehorn)$.
+Therefore, the framework in~\cite{SalujaST95} is enough for defining the class of problems in $\eqso(\ehorn)$.

### 2019-03-04 LMCS/sendrev/framework.tex
Solving some pending issues of reviewer 1
Cambio en biblio.

@@
 x < y \wedge y < z$ is checking the triangle property, by returning $1$ if $\{x, y, z\}$ forms a triangle in $\fG$ and $0$ otherwise.
-Finally, the sum quantifiers in $\alpha_1$ aggregates all the values, counting the number of triangles in $\fG$.
+Finally, the sum quantifiers in $\alpha_1$ aggregate all the values, counting the number of triangles in $\fG$.
 
@@
 
-Finally, the work in~\cite{SalujaST95} and \cite{ComptonG96} is of particular interest for our research. 
+Finally, the work in~\cite{SalujaST95,ComptonG96,0001HKV16} is of particular interest for our research. 
 In~\cite{SalujaST95}, it was proposed to define a function over a structure by using free variables in an SO-formula; in particular, the function is defined by the number of instantiations of the free variables that are satisfied by the structure.

### 2019-03-05 LMCS/sendrev/response-letter.tex
pass on the response letter

@@
 \def\cf{{\em cf.}}
+\newcommand{\ans}{{\bf Answer:\ }}
+\newcommand{\cm}[1]{{\bf Comment:\ } #1}
 
@@
 
-\title{Response letter to the reviews of the paper: Descriptive Complexity for Counting Complexity Classes}
+\title{Response letter to the reviews of the paper ``Descriptive Complexity for Counting Complexity Classes''}
 
@@
 
-We would like to thank the anonymous reviewers for their useful comments.
-Below are our answers to comments or suggestions made about our paper.
-We highlight the reviewer comment in red and then write our answer.
+We thank very much the anonymous reviewers for their useful comments.
+We carefully took into consideration all their comments, and implemented their suggestions (in particular, all the typos mentioned in the reviews were fixed).
+Below please find our answers to their main comments.
 
@@
 	\setlength\itemsep{0.5em}
-	\item \red{p.2, first paragraph: ``However, other function classes have emerged from the need to understand	the complexity of some computation problems for which little can be said if their decision counterparts are considered.'' Why can only little be said about the decision counterparts of the examples mentionend[sic] afterwards?}
+	\item \cm{p.2, first paragraph: ``However, other function classes have emerged from the need to understand	the complexity of some computation problems for which little can be said if their decision counterparts are considered.'' Why can only little be said about the decision counterparts of the examples mentionend[sic] afterwards?}
 
-	This was meant to say that the complexity of the decision counterparts of the problems mentioned afterwards did not provide a full understanding of the complexity of the problem in its computation version.
-This sentence has been reformulated to explain that idea.
+	\ans What we intended to say was that there is a significant difference in terms of complexity between the computation and decision versions of the problems mentioned afterwards, so no much can be inferred about the complexity of these computation problems from their decision counterparts.
+This sentence was reformulated to explain that idea.
 
@@
 %	\item p.10, before Thm. 4.5: ``by considering second-order product'': We need both second-order product and sum (at least both are used in Thm. 4.5)
-	\item \red{p.10, last line: ``$B$-th bit'': mention in what way $B$ is interpreted as a number}
+	\item \cm{p.10, last line: ``$B$-th bit'': mention in what way $B$ is interpreted as a number}
 
-	This proof was rewitten to explain this notion better.
+	\ans This proof was rewritten to explain this notion better.
 
@@
 %	\item p.17: ``Since $\alpha$ is not null'' $\to$ ``Since $\sem{\alpha}$ is not the identically zero function'' (or similar)
-	\item \red{p.17/18: In the proof of Prop 5.3, the vocabulary only contains the numerical predicate $<$.
+	\item \cm{p.17/18: In the proof of Prop 5.3, the vocabulary only contains the numerical predicate $<$.
 When encoding structures as binary words, numerical predicates are implicit (and not encoded), though.
@@
 
-	For ordered structures the vocabulary should always include $<$. The encoding function in the paper was revised to make this more clear.
+	\ans For ordered structures the vocabulary should always include $<$. The encoding function in the paper was revised to make this more clear.
 
@@
 %	\item[$\checkmark$] p.25: ``We construct a function $\lambda$ that receives a quantitative formula $\beta$ ...'' $\to$ ``... that receives a formula $\beta \in \eqso(\logex{1})$ ...''
-	\item \red{p.25, last sentence: ``which is based on considering some witnesses of logarithmic size'' could be explained a little more / made more clear}
+	\item \cm{p.25, last sentence: ``which is based on considering some witnesses of logarithmic size'' could be explained a little more / made more clear}
 
-	WIP
+	\ans After the proof of Theorem 5.9, we added an example based on \#3-DNF with a detailed explanation of the notion of witness of logarithmic size.
 
@@
 %	\item p.29: ``Intuitively $t_1,\ldots,t_{m'}$ are used to count from 1 to $m'$ each subformula ...'': Should be reformulated, maybe rather write that they are used to have disjoint sets of propositional assignments for the different disjuncts of the outermost disjunction (corresponding to the summands in the original formula)
-	\item \red{p.30, proof of Prop.
+	\item \cm{p.30, proof of Prop.
 5.12: For $\dhsat \in \totp$, [35] is cited.
@@
 
-	A proof was added.
+	\ans A proof that $\dhsat \in \totp$ was included in  the paper.
 
 %	\item[$\checkmark$] p.31, definition of $\leq$ on $\F$: ``$f(i) \leq g(i)$ for every $i \in \{1,\ldots,\ell\}$'': Inputs for $f$ and $g$ should be elements of $A^{\ell}$, not $\{1,\ldots,\ell\}$
-	\item \red{p.32/33, example 6.2: I don't see why the formula given for dag's does not work for arbitrary graphs: In arbitrary graphs, the fixpoint is still reached after at most $n-1$ steps, since the maximal length of any shortest path in any graph is $n-1$.
+	\item \cm{p.32/33, example 6.2: I don't see why the formula given for dag's does not work for arbitrary graphs: In arbitrary graphs, the fixpoint is still reached after at most $n-1$ steps, since the maximal length of any shortest path in any graph is $n-1$.
 On the other hand, assuming that I am mistaken and the support does not become stable within $n-1$ steps: Using the formula	given for arbitrary graphs, the support of the constructed function restricted to those tuples that have first in their last component is the same as the support of the constructed function using the formula for dags in each step of the recursion.
@@
 
-	The problem with the first formula for directed graphs is precisely the opposite.
-The fixed point could be reached too soon without counting all paths of length at most $n-1$.
+	\ans The problem with the first formula for directed graphs that are not acyclic is precisely the opposite.
+The fixed point could be reached too soon without counting all paths of length at most $n-1$ (for example, consider a clique with $n$ nodes).
 The fixed point is always reached after $k$ steps, where $k$ is the size of the maximum shortest path between two vertices in the graph.
@@
 We added a comment about this in the paper.
+
+\item In \cite{DBLP:conf/csl/0001HKV16}, a similar hierarchy to the one by Saluja et al. was introduced and studied. This work should be cited.
+
+\ans This article is now cited in the paper.
 
@@
 \end{itemize}
+%%\vspace{1em}
+%%\subsection*{Grammatical/Language}
+%%\begin{itemize}
+%%	\setlength\itemsep{0.5em}
+%%	\item generally: both formulae and formulas are used as plural of formula---should be consistent
+%%	\item p.2, in the middle: we propose a restriction: Should be something like ``we propose to restrict WL to the natural numbers as fixed semiring, calling this restriction ...''
+%%	\item[$\checkmark$] p.3, first paragraph of 2.1: ``omit the word finite or ordered'' $\to$ ``omit the words finite and ordered''
+%%	\item p5, description of syntax of QSO: ``This division between Boolean and quantitative level is essential for understanding the difference between the logic and the quantitative part.'' reformulate this?
+%%	\item[$\checkmark$] p.5, next sentence: ``this will allow us later'' $\to$ ``this will later allow us''
+%%	\item[$\checkmark$] p.6, second to last sentence: ``similarity of $\alpha_3$ with'': ``similarity to'' is much more common than ``similarity with''
+%%	\item[$\checkmark$] p.7, next sentence: ''in this direction'' $\to$ ``in this regard''
+%%	\item[$\checkmark$] p.7, last sentence before 3.1: ''connection of QSO with'' $\to$ ``connection of QSO to''
+%%	\item[$\checkmark$] p.7, middle: ``Another difference of WL with our approach'' $\to$ ``Another difference between WL and our approach''
+%%	\item p.9, first sentence: ``then condition (2) holds'': omit the word then or reformulate the sentence
+%%	\item[$\checkmark$] p.9, first paragraph: ``the constant function $s$ can be trivially simulated'' $\to$ ``any constant function $s$ can be trivially simulated''
+%%	\item[$\checkmark$] p.9, proof of Prop. 4.3: ``To prove the condition (2)'' $\to$ ``To prove condition (2)''
+%%	\item throughout the paper: ``similar than'' $\to$ ``similar to'' / ``similar than for X, we do Y'' $\to$ ``similarly as for''. Examples:
+%%	\begin{itemize}
+%%		\item[-] p.10, proof of Thm. 4.5: ``the proof is similar than in Theorem 4.4'' $\to$ ``the proof is
+%%		similar to the proof of Theorem 4.4''	
+%%		\item[-] p.10, second sentence: ``similar than the previous proof'' $\to$ ``similar as in the previous
+%%		proof''
+%%		\item[-] p.11, last sentence before example 4.7: ``Similar than for QSO'' $\to$ ``Similarly as for QSO''
+%%	\end{itemize}
+%%	\item p.11, end of proof of Thm. 4.5: ``Using an analogous argument'': state what the argument is analogous to (the argument in Thm. 4.4)
+%%	\item[$\checkmark$] p.12, after Corollary 4.8: ``shows how robust and versatile is QSO'' $\to$ ``shows how robust and versatile QSO is''
+%%	\item p.12, after Corollary 4.8: ``counting complexity classes [...] even beyond $\nat$'' - make more precise what ``beyond $\nat$'' refers to
+%%	\item[$\checkmark$] p.12, ``given CNF propositional formula, that can be made true'': no comma
+%%	\item p.12: ``in [43] they defined'' $\to$ something like ``$<$authors$>$ [43] defined''
+%%	\item[$\checkmark$] p.12: ``we extend as follows QSO'' $\to$ ``we extend QSO by max and min quantifiers as follows''
+%%	\item[$\checkmark$] p.12: ``MinQSO is defined analogously changing max with min'' $\to$ ``replacing max by min''
+%%	\item[$\checkmark$] p.12: ``the same holds with MinQSO(FO)'' $\to$ ``the same holds for ...''
+%%	\item[$\checkmark$] p.13: ``in contrast with'' $\to$ ``in contrast to''
+%%	\item[$\checkmark$] p.13: ``follow the same approach but considering'' $\to$ ``but consider''
+%%	\item[$\checkmark$] p.13: ``in the search of efficient approximation algorithms'' $\to$ ``in the search for''
+%%	\item[$\checkmark$] p.14: ``the search of robust classes'' $\to$ ''the search for robust classes''
+%%	\item p.14: ''the first desirable condition has to do with the closure properties satisfied by the class'': Reformulate this sentence
+%%	\item[$\checkmark$] p.14: ``As in the cases of P and NP that are closed'' $\to$ ``Analogously to the cases of P and NP, which are closed under intersection and union, ...''
+%%	\item[$\checkmark$] p.16: ``such that for both, $A\mod\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$'' $\to$ ''such that for $i \in \{1,2\}$: ... ''
+%%	\item[$\checkmark$] p.16: ``Similarly to Theorem 5.1, we show ..'' $\to$ ``Similarly as in the proof of Theorem 5.1, we 	show''
+%%	\item p.19: ``However, this class is not closed under sum, and then it is not robust under basic	closure properties'': Should be reformulated
+%%	\item[$\checkmark$] p.19: ``Let ... be this formula, where $\varphi$ is in first-order and quantifier-free'' $\to$ ``... where $\varphi$ is first-order and quantifier-free''
+%%	\item[$\checkmark$] several occurrences: ``This is, ...'' $\to$ ``That is, ...'', for example two times on p.20
+%%	\item[$\checkmark$] p.20: ``we focus only in proving'' $\to$ ``we focus only on proving''
+%%	\item[$\checkmark$] p.20: ``We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of FO.'' $\to$ ``We prove this for the more general case of $\eqso(\LL)$ with $\LL$ being a fragment of SO.'' (or similar)
+%%	\item[$\checkmark$] p.20: ``we will show a formula'' $\to$ ``give'' or ``construct'' instead of ``show''
+%%	\item[$\checkmark$] p.20/21: ``$\shp$ is believed not to be closed under subtraction by one by some complexitytheoretic assumption'' $\to$ ``...
+due to some complexity-theoretic assumption'' (or similar)
+%%	\item[$\checkmark$] p.21, next sentence: ``for a possible extension'' $\to$ ``as a possible extension''
+%%	\item p.21: ``the desired extension has to be achieved by ...
+.
+More precisely, we define...'': The second sentence doesn’t make the statement of the first one more precise, but rather explains the actual definiton (which is chosen in accordance with the sentence before)
+%%	\item[$\checkmark$] p.21: ``a parsimonious reduction from a function'' $\to$ ``a parsimonious reduction from any function''
+%%	\item[$\checkmark$] p.21: ``we define a function that converts a formula $\alpha$ in...'' $\to$ ``that converts a given formula $\alpha$...''
+%%	\item[$\checkmark$] p.22: ``we iterate for every tuple'' $\to$ ``we iterate over all tuples''
+%%	\item[$\checkmark$] p.22: ``Then we will show how to define a formula $\varphi'$ ...'' $\to$ ``We will show how to define ...''
+%%	\item[$\checkmark$] p.22: ''no variable from $\bar{x}$ are mentioned'' $\to$ ``... is mentioned''
+%%	\item[$\checkmark$] p.24: ``if, and only if, if X, then Y'': This construction makes it unneccesarily hard to parse the sentence
+%%	\item[$\checkmark$] p.24: ``Then, the conclusion of the implication...'' $\to$ ``The conclusion of the implication...''
+%%	\item[$\checkmark$] p.25: ``For the sake of simplification...'' $\to$ ``For the sake of simplicity...''
+%%	\item[$\checkmark$] p.26: ``In the previous section, we show ...'' $\to$ ``In the previous section we showed ...''
+%%	\item[$\checkmark$] p.27: ``we postpone the proof the previous proposition after ...'' $\to$ ``we postpone the proof of the previous proposition to after ...''
+%%	\item p.29: ``replace by $\perp$ and $\top$ \underline{wherever it corresponds}'': Should be reformulated
+%%	\item[$\checkmark$] p.30: ``consist in computing $g_{\alpha}(\enc(\A))$ for each input ...'' $\to$ ``... on input ...''
+%%	\item[$\checkmark$] p.30, same sentence: ``and then simulate'' $\to$ ``and then simulating''
+%%	\item[$\checkmark$] p.33: ``In contrast with LFP'' $\to$ ``In contrast to LFP''
+%%	\item[$\checkmark$] p.33: ``These operations are represented by the set $\{0,1,\emptyset\}$, respectively.'' $\to$ ``... by 0, 1 and $\emptyset$, respectively''
+%%	\item[$\checkmark$] p.34: ``Notice that the values of each function $f$ in $\alpha$'' $\to$ ``Notice that the values of each function in $\alpha$''
+%%	\item[$\checkmark$] p.34: ``that is why we talk about ...'' $\to$ ``which is why we talk about''
+%%	\item[$\checkmark$] p.34: ``Besides, notice that notation ...'' $\to$ ``Besides, notice that the notation ...''
+%%	\item[$\checkmark$] p.34: ``Function $h$ is used to indicate whether the head of the working tape is in some position at some point of time'' $\to$ ``...
+whether the head of the working tape is in a specific position at a certain point of time'' (or similar)
+%%	\item[$\checkmark$] p.36: ``Given a relation signature $\R$'' $\to$ ``... relational signature ...''
+%%	\item[$\checkmark$] p.37: ``As it was mentioned before'' $\to$ ``As mentioned before''
+%%	\item[$\checkmark$] p.37: ``Next we construct a logarithmic-space ....'' $\to$ ``We construct a logarithmic-space ...''
+%%	\item[?] p.38: ``In this sense, there are several directions for future research'' $\to$ ``Consequently, there are ...'' (or similar)
+%%	\item[$\checkmark$] p.38: ``In the same direction, ...'' $\to$ ``Similarly, ...'' (or similar)
+%%	\item[$\checkmark$] p.38: ``so to have'' $\to$ ``in order to have'' (or similar)
+%%\end{itemize}
 %\vspace{1em}
-%\subsection*{Grammatical/Language}
+%%\subsection*{Typos}
+%%\begin{itemize}
+%%	\setlength\itemsep{0.5em}
+%%	\item p.1, footnote: ``clases'' should be ``classes''
+%%	\item throughout the paper: plural used with verb ending on s. Examples:
+%%	\begin{itemize}
+%%		\item[-] p.2, middle: ``fragments of QSO captures''
+%%		\item[-] p.6, example 3.1: ``the sum quantifiers in $\alpha_1$ aggregates''
+%%	\end{itemize}
+%%	\item[$\checkmark$] p4, beginning of 2.2: ``some of \underline{the} are recalled here''
+%%	\item[$\checkmark$] p6, example 3.2, first paragraph: ``otherwise $M(i.j)$'', dot used instead of comma
+%%	\item[$\checkmark$] throughout the paper: a instead of an (espacially: a FO formula/assignment, a SO formula/assignment).
+%%	Example:
+%%	\begin{itemize}
+%%		\item[-] p.9, third sentence: ``Given a FO formula'' $\to$ ``Given an FO formula''
+%%		\item[-] p.38: ``As a example'' $\to$ ``As an example''
+%%	\end{itemize}
+%%	\item throughout the paper: singular used with verb not ending on s. Examples:
+%%	\begin{itemize}
+%%		\item[-] p.9, after proof of Prop. 4.3: ``use of a fragment [...] that capture''
+%%		\item[-] p.11, end of proof of Thm. 4.5: ``and $\alpha$ reconstruct''
+%%	\end{itemize}
+%%	\item[$\checkmark$] p.9, middle: ``A second attempt could be based then'' $\to$ ``A second attempt could then be
+%%	based''
+%%	\item[$\checkmark$] p.11, 4.1: ``output of a function \underline{in} not a natural number''
+%%	\item[$\checkmark$] p.12: ``we have to dinstinguished''
+%%	\item[$\checkmark$] p.14, beginning of 5.1: ``subclases''
+%%	\item[$\checkmark$] p.21: ``it does not mentioned'' $\to$ ``it does not mention''
+%%	\item[$\checkmark$] p.22: ``Without lost of generality'' $\to$ ``Without loss of generality''
+%%	\item[$\checkmark$] p.24: ``The previous claim and proof motivates the following definitions.'' $\to$ ``The previous
+%%	claim and proof motivate the following definition.''
+%%	\item p.30: ``to define a natural generalization for functions of the notion of least fixed point of
+%%	LFP'' $\to$ ``to define a natural generalization of LFP to functions'' (or similar)
+%%	\item[$\checkmark$] p.30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
+%%	\item[$\checkmark$] p.31: ``we simply generalized these conditions'' $\to$ ``we simply generalize these conditions''
+%%	\item[$\checkmark$] p.32: ``which is used to defined'' $\to$ ``which is used to define''
+%%	\item[$\checkmark$] p.33: ``since $\beta$ is constructed from monotones operations'' $\to$ ``... monotone ...''
+%%	\item[$\checkmark$] p.35: ``Formulas $\alpha_h$ is defined ...'' $\to$ ``Formula $\alpha_h$ is defined ...''
+%%	7
+%%	\item[$\checkmark$] p.37: ``we simulate $M_\beta$ on input $v[1/n]$'' $\to$ ``... with input $v[1/x]$''
+%%\end{itemize}
+%\vspace{1em}
 %\begin{itemize}
 %	\setlength\itemsep{0.5em}
-%	\item generally: both formulae and formulas are used as plural of formula---should be consistent
-%	\item p.2, in the middle: we propose a restriction: Should be something like ``we propose to restrict WL to the natural numbers as fixed semiring, calling this restriction ...''
-%	\item[$\checkmark$] p.3, first paragraph of 2.1: ``omit the word finite or ordered'' $\to$ ``omit the words finite and ordered''
-%	\item p5, description of syntax of QSO: ``This division between Boolean and quantitative level is essential for understanding the difference between the logic and the quantitative part.'' reformulate this?
-%	\item[$\checkmark$] p.5, next sentence: ``this will allow us later'' $\to$ ``this will later allow us''
-%	\item[$\checkmark$] p.6, second to last sentence: ``similarity of $\alpha_3$ with'': ``similarity to'' is much more common than ``similarity with''
-%	\item[$\checkmark$] p.7, next sentence: ''in this direction'' $\to$ ``in this regard''
-%	\item[$\checkmark$] p.7, last sentence before 3.1: ''connection of QSO with'' $\to$ ``connection of QSO to''
-%	\item[$\checkmark$] p.7, middle: ``Another difference of WL with our approach'' $\to$ ``Another difference between WL and our approach''
-%	\item p.9, first sentence: ``then condition (2) holds'': omit the word then or reformulate the sentence
-%	\item[$\checkmark$] p.9, first paragraph: ``the constant function $s$ can be trivially simulated'' $\to$ ``any constant function $s$ can be trivially simulated''
-%	\item[$\checkmark$] p.9, proof of Prop. 4.3: ``To prove the condition (2)'' $\to$ ``To prove condition (2)''
-%	\item throughout the paper: ``similar than'' $\to$ ``similar to'' / ``similar than for X, we do Y'' $\to$ ``similarly as for''. Examples:
-%	\begin{itemize}
-%		\item[-] p.10, proof of Thm. 4.5: ``the proof is similar than in Theorem 4.4'' $\to$ ``the proof is
-%		similar to the proof of Theorem 4.4''	
-%		\item[-] p.10, second sentence: ``similar than the previous proof'' $\to$ ``similar as in the previous
-%		proof''
-%		\item[-] p.11, last sentence before example 4.7: ``Similar than for QSO'' $\to$ ``Similarly as for QSO''
-%	\end{itemize}
-%	\item p.11, end of proof of Thm. 4.5: ``Using an analogous argument'': state what the argument is analogous to (the argument in Thm. 4.4)
-%	\item[$\checkmark$] p.12, after Corollary 4.8: ``shows how robust and versatile is QSO'' $\to$ ``shows how robust and versatile QSO is''
-%	\item p.12, after Corollary 4.8: ``counting complexity classes [...] even beyond $\nat$'' - make more precise what ``beyond $\nat$'' refers to
-%	\item[$\checkmark$] p.12, ``given CNF propositional formula, that can be made true'': no comma
-%	\item p.12: ``in [43] they defined'' $\to$ something like ``$<$authors$>$ [43] defined''
-%	\item[$\checkmark$] p.12: ``we extend as follows QSO'' $\to$ ``we extend QSO by max and min quantifiers as follows''
-%	\item[$\checkmark$] p.12: ``MinQSO is defined analogously changing max with min'' $\to$ ``replacing max by min''
-%	\item[$\checkmark$] p.12: ``the same holds with MinQSO(FO)'' $\to$ ``the same holds for ...''
-%	\item[$\checkmark$] p.13: ``in contrast with'' $\to$ ``in contrast to''
-%	\item[$\checkmark$] p.13: ``follow the same approach but considering'' $\to$ ``but consider''
-%	\item[$\checkmark$] p.13: ``in the search of efficient approximation algorithms'' $\to$ ``in the search for''
-%	\item[$\checkmark$] p.14: ``the search of robust classes'' $\to$ ''the search for robust classes''
-%	\item p.14: ''the first desirable condition has to do with the closure properties satisfied by the class'': Reformulate this sentence
-%	\item[$\checkmark$] p.14: ``As in the cases of P and NP that are closed'' $\to$ ``Analogously to the cases of P and NP, which are closed under intersection and union, ...''
-%	\item[$\checkmark$] p.16: ``such that for both, $A\mod\varphi(\bar{B}_i,\bar{b}_i,\bar{a}_i)$'' $\to$ ''such that for $i \in \{1,2\}$: ... ''
-%	\item[$\checkmark$] p.16: ``Similarly to Theorem 5.1, we show ..'' $\to$ ``Similarly as in the proof of Theorem 5.1, we 	show''
-%	\item p.19: ``However, this class is not closed under sum, and then it is not robust under basic	closure properties'': Should be reformulated
-%	\item[$\checkmark$] p.19: ``Let ... be this formula, where $\varphi$ is in first-order and quantifier-free'' $\to$ ``... where $\varphi$ is first-order and quantifier-free''
-%	\item[$\checkmark$] several occurrences: ``This is, ...'' $\to$ ``That is, ...'', for example two times on p.20
-%	\item[$\checkmark$] p.20: ``we focus only in proving'' $\to$ ``we focus only on proving''
-%	\item[$\checkmark$] p.20: ``We prove this for a more general case for $\eqso(\LL)$ where $\LL$ is a fragment of FO.'' $\to$ ``We prove this for the more general case of $\eqso(\LL)$ with $\LL$ being a fragment of SO.'' (or similar)
-%	\item[$\checkmark$] p.20: ``we will show a formula'' $\to$ ``give'' or ``construct'' instead of ``show''
-%	\item[$\checkmark$] p.20/21: ``$\shp$ is believed not to be closed under subtraction by one by some complexitytheoretic assumption'' $\to$ ``...
-due to some complexity-theoretic assumption'' (or similar)
-%	\item[$\checkmark$] p.21, next sentence: ``for a possible extension'' $\to$ ``as a possible extension''
-%	\item p.21: ``the desired extension has to be achieved by ...
-.
-More precisely, we define...'': The second sentence doesn’t make the statement of the first one more precise, but rather explains the actual definiton (which is chosen in accordance with the sentence before)
-%	\item[$\checkmark$] p.21: ``a parsimonious reduction from a function'' $\to$ ``a parsimonious reduction from any function''
-%	\item[$\checkmark$] p.21: ``we define a function that converts a formula $\alpha$ in...'' $\to$ ``that converts a given formula $\alpha$...''
-%	\item[$\checkmark$] p.22: ``we iterate for every tuple'' $\to$ ``we iterate over all tuples''
-%	\item[$\checkmark$] p.22: ``Then we will show how to define a formula $\varphi'$ ...'' $\to$ ``We will show how to define ...''
-%	\item[$\checkmark$] p.22: ''no variable from $\bar{x}$ are mentioned'' $\to$ ``... is mentioned''
-%	\item[$\checkmark$] p.24: ``if, and only if, if X, then Y'': This construction makes it unneccesarily hard to parse the sentence
-%	\item[$\checkmark$] p.24: ``Then, the conclusion of the implication...'' $\to$ ``The conclusion of the implication...''
-%	\item[$\checkmark$] p.25: ``For the sake of simplification...'' $\to$ ``For the sake of simplicity...''
-%	\item[$\checkmark$] p.26: ``In the previous section, we show ...'' $\to$ ``In the previous section we showed ...''
-%	\item[$\checkmark$] p.27: ``we postpone the proof the previous proposition after ...'' $\to$ ``we postpone the proof of the previous proposition to after ...''
-%	\item p.29: ``replace by $\perp$ and $\top$ \underline{wherever it corresponds}'': Should be reformulated
-%	\item[$\checkmark$] p.30: ``consist in computing $g_{\alpha}(\enc(\A))$ for each input ...'' $\to$ ``... on input ...''
-%	\item[$\checkmark$] p.30, same sentence: ``and then simulate'' $\to$ ``and then simulating''
-%	\item[$\checkmark$] p.33: ``In contrast with LFP'' $\to$ ``In contrast to LFP''
-%	\item[$\checkmark$] p.33: ``These operations are represented by the set $\{0,1,\emptyset\}$, respectively.'' $\to$ ``... by 0, 1 and $\emptyset$, respectively''
-%	\item[$\checkmark$] p.34: ``Notice that the values of each function $f$ in $\alpha$'' $\to$ ``Notice that the values of each function in $\alpha$''
-%	\item[$\checkmark$] p.34: ``that is why we talk about ...'' $\to$ ``which is why we talk about''
-%	\item[$\checkmark$] p.34: ``Besides, notice that notation ...'' $\to$ ``Besides, notice that the notation ...''
-%	\item[$\checkmark$] p.34: ``Function $h$ is used to indicate whether the head of the working tape is in some position at some point of time'' $\to$ ``...
-whether the head of the working tape is in a specific position at a certain point of time'' (or similar)
-%	\item[$\checkmark$] p.36: ``Given a relation signature $\R$'' $\to$ ``... relational signature ...''
-%	\item[$\checkmark$] p.37: ``As it was mentioned before'' $\to$ ``As mentioned before''
-%	\item[$\checkmark$] p.37: ``Next we construct a logarithmic-space ....'' $\to$ ``We construct a logarithmic-space ...''
-%	\item[?] p.38: ``In this sense, there are several directions for future research'' $\to$ ``Consequently, there are ...'' (or similar)
-%	\item[$\checkmark$] p.38: ``In the same direction, ...'' $\to$ ``Similarly, ...'' (or similar)
-%	\item[$\checkmark$] p.38: ``so to have'' $\to$ ``in order to have'' (or similar)
+%%	\item[$\checkmark$] running title is ``COUNTING''
+%%	\item on definition of functions, \verb|\colon| should be used instead of :
+%	\item In \cite{DBLP:conf/csl/0001HKV16}, a similar hierarchy to the one by Saluja et al. was introduced and studied. This work should be cited.
 %\end{itemize}
-\vspace{1em}
-%\subsection*{Typos}
-%\begin{itemize}
-%	\setlength\itemsep{0.5em}
-%	\item p.1, footnote: ``clases'' should be ``classes''
-%	\item throughout the paper: plural used with verb ending on s. Examples:
-%	\begin{itemize}
-%		\item[-] p.2, middle: ``fragments of QSO captures''
-%		\item[-] p.6, example 3.1: ``the sum quantifiers in $\alpha_1$ aggregates''
-%	\end{itemize}
-%	\item[$\checkmark$] p4, beginning of 2.2: ``some of \underline{the} are recalled here''
-%	\item[$\checkmark$] p6, example 3.2, first paragraph: ``otherwise $M(i.j)$'', dot used instead of comma
-%	\item[$\checkmark$] throughout the paper: a instead of an (espacially: a FO formula/assignment, a SO formula/assignment).
-%	Example:
-%	\begin{itemize}
-%		\item[-] p.9, third sentence: ``Given a FO formula'' $\to$ ``Given an FO formula''
-%		\item[-] p.38: ``As a example'' $\to$ ``As an example''
-%	\end{itemize}
-%	\item throughout the paper: singular used with verb not ending on s. Examples:
-%	\begin{itemize}
-%		\item[-] p.9, after proof of Prop. 4.3: ``use of a fragment [...] that capture''
-%		\item[-] p.11, end of proof of Thm. 4.5: ``and $\alpha$ reconstruct''
-%	\end{itemize}
-%	\item[$\checkmark$] p.9, middle: ``A second attempt could be based then'' $\to$ ``A second attempt could then be
-%	based''
-%	\item[$\checkmark$] p.11, 4.1: ``output of a function \underline{in} not a natural number''
-%	\item[$\checkmark$] p.12: ``we have to dinstinguished''
-%	\item[$\checkmark$] p.14, beginning of 5.1: ``subclases''
-%	\item[$\checkmark$] p.21: ``it does not mentioned'' $\to$ ``it does not mention''
-%	\item[$\checkmark$] p.22: ``Without lost of generality'' $\to$ ``Without loss of generality''
-%	\item[$\checkmark$] p.24: ``The previous claim and proof motivates the following definitions.'' $\to$ ``The previous
-%	claim and proof motivate the following definition.''
-%	\item p.30: ``to define a natural generalization for functions of the notion of least fixed point of
-%	LFP'' $\to$ ``to define a natural generalization of LFP to functions'' (or similar)
-%	\item[$\checkmark$] p.30: ``this notion can be used to captures FP'' $\to$ ``... to capture FP''
-%	\item[$\checkmark$] p.31: ``we simply generalized these conditions'' $\to$ ``we simply generalize these conditions''
-%	\item[$\checkmark$] p.32: ``which is used to defined'' $\to$ ``which is used to define''
-%	\item[$\checkmark$] p.33: ``since $\beta$ is constructed from monotones operations'' $\to$ ``... monotone ...''
-%	\item[$\checkmark$] p.35: ``Formulas $\alpha_h$ is defined ...'' $\to$ ``Formula $\alpha_h$ is defined ...''
-%	7
-%	\item[$\checkmark$] p.37: ``we simulate $M_\beta$ on input $v[1/n]$'' $\to$ ``... with input $v[1/x]$''
-%\end{itemize}
-\vspace{1em}
-\begin{itemize}
-	\setlength\itemsep{0.5em}
-%	\item[$\checkmark$] running title is ``COUNTING''
-%	\item on definition of functions, \verb|\colon| should be used instead of :
-	\item In \cite{DBLP:conf/csl/0001HKV16}, a similar hierarchy to the one by Saluja et al. was introduced and studied. This work should be cited.
-\end{itemize}
 
@@
 	\setlength\itemsep{0.5em}
-	\item \red{In several proofs, you rely on disjoint union of solutions and cartesian product of solutions.
+	\item \cm{In several proofs, you rely on disjoint union of solutions and cartesian product of solutions.
 Indeed, its gives a polynomial time counting reduction.
-Can you make connections with (+,*) circuits ?
+Can you make connections with (+,*) circuits?
 It would be similar to what is done in A Circuit-Based Approach to Efficient Enumeration (Antoine Amarilli, Pierre Bourhis, Louis Jachiet, Stefan Mengel) where simple circuits (dDNNF, which are basically disjoint unions and cartesian products) are used to capture the solutions of MSO queries.
@@
 
-	We thank the reviewer for showing us the literature regarding enumeration problems, in particular, for the paper from Durand and Strozecki.
-Unfortunately, the relation between our framework with enumeration complexity classes goes beyond the scope of the paper.
-Nevertheless, in the last section we added a final paragraph pointing out this connection as an interesting open question.
+	\ans We thank the reviewer for pointing us the literature regarding enumeration problems, in particular, for the paper by Durand and Strozecki.
+The relationship between our framework with enumeration complexity classes is a very interesting question to answer, but it goes beyond the scope of the paper.
+In the last section, we added a final paragraph pointing out this connection as an interesting open problem that should be further investigated.
 
-	\item \red{For the most involved proofs, you could describe informally your proof strategy in a few lines to help the reader.
+	\item \cm{For the most involved proofs, you could describe informally your proof strategy in a few lines to help the reader.
 You may also add some example to illustrate the different fragment of QSO (you already do it sometimes and it is very helpful).}
 
-	Although we agree that adding more proof sketches and examples would help, the paper is already too long for adding more additional content.
+	\ans We added an example based on \#-3DNF after the proof of Theorem 5.9, to help the reader to understand the main idea behind the most involved proof in the paper.
+As mentioned by the reviewer, we tried to have examples illustrating the main notions used in the paper.
+However, we would prefer not to add even more examples, as the paper is already quite long.
+
+	% Although we agree that adding more proof sketches and examples would help, the paper is already too long for adding more additional content.
 Furthermore, as the reviewer mentioned we already have this for the main definitions and proofs (e.g.
@@
 %	\item page 24: 	if, if $\phi$$\to$ $\phi$
-	\item \red{page 25: 	-you explain that the proof relies on witnesses of logarithmic size.	It was not clear to me while reading the proof. Can you explain more ?}
+	\item \cm{page 25: 	-you explain that the proof relies on witnesses of logarithmic size.	It was not clear to me while reading the proof. Can you explain more ?}
 
-	This is discussed in the response to Reviewer 1.
+	\ans After the proof of Theorem 5.9, we added an example based on \#3-DNF with a detailed explanation of the notion of witness of logarithmic size.
+	%This is discussed in the response to Reviewer 1.
 
-	\item \red{page 28:	The complete problem for $\Sigma\text{QSO}(\Sigma_2\text{-Horn})$ would be complete also for $\#(\Sigma_2\text{-Horn})$.
+	\item \cm{page 28: The complete problem for $\Sigma\text{QSO}(\Sigma_2\text{-Horn})$ would be complete also for $\#(\Sigma_2\text{-Horn})$.
 You should mention here that you do not need sum here and that the framework of [39] is enough for this result.}
 
-	Indeed, from the proof of Theorem 5.14 one can also derive that $\#(\Sigma_2\text{-Horn})$ is equivalent to $\Sigma\text{QSO}(\Sigma_2\text{-Horn})$ and, thus, the framework of [39] is enough for defining this class of problems.
+	\ans Indeed, from the proof of Theorem 5.14 one can also conclude that $\#(\Sigma_2\text{-Horn})$ is equivalent to $\Sigma\text{QSO}(\Sigma_2\text{-Horn})$ and, thus, the framework proposed in [39] is enough for defining this class of problems.
 At the end of Section 5, we added a paragraph mentioning this.
@@
 %	\item page 33: 	-there is a proof symbol \verb|\qed| but it does not end a well defined proof environment
-	\item \red{page 38: the last sentence on fixed point operator is to vague. Either expand it to say what king of question on fixed point operator you want to investigate or remove it.}
+	\item \cm{page 38: the last sentence on fixed point operator is to vague. Either expand it to say what kind of question on fixed point operator you want to investigate or remove it.}
 
-	We agree with the reviewer and we decided to remove this sentence. 
+	\ans We agree with the reviewer, and we decided to remove this sentence. 
 \end{itemize}

### 2019-03-05 LMCS/sendrev/response-letter.tex
Pequena modificación.

@@
 
-	\ans What we intended to say was that there is a significant difference in terms of complexity between the computation and decision versions of the problems mentioned afterwards, so no much can be inferred about the complexity of these computation problems from their decision counterparts.
+	\ans The intention of this sentence was that there is a significant difference in terms of complexity between the computation and decision versions of the problems mentioned afterwards (e.g.
+$\#\textsc{P}$-complete versus $\textsc{PTIME}$), so no much can be inferred about the complexity of these computation problems from their decision counterparts.
 This sentence was reformulated to explain that idea.

### 2019-03-05 LMCS/sendrev/beyond.tex
removing comments

@@
 \end{exa}
-\martin{El Reviewer 1 hizo un comentario de este ejemplo, lo respond\'i en la carta}
-\marcelo{Me parece bien la respuesta en la carta. Agregue esta respuesta en el paper, ayuda a entender mejor el ejemplo.}
+%\martin{El Reviewer 1 hizo un comentario de este ejemplo, lo respond\'i en la carta}
+%\marcelo{Me parece bien la respuesta en la carta. Agregue esta respuesta en el paper, ayuda a entender mejor el ejemplo.}
+
 In contrast to $\lfp$, to reach a fixed point we do not need to impose any positive restriction on the formula $\beta(\x,h)$.

### 2019-03-05 LMCS/sendrev/counting.tex
un par de correcciones y algunos problemas de compatibilidad con el la instruccion !TEX = main
sc
Solving some pending issues of reviewer 1
removing comments

@@
 $$
-\martin{modifique la definicion para que no sea necesario codificar el $<$.}
+
+%\martin{modifique la definicion para que no sea necesario codificar el $<$.}
+%\marcelo{OK}
 %We define the class of all $\R$-functions, denoted by $\Func(\R)$, as the class of all functions $f: \ostr \rightarrow \bbN$.

### 2019-03-05 LMCS/sendrev/response-letter.tex
removing comments

@@
 We thank very much the anonymous reviewers for their useful comments.
-We carefully took into consideration all their comments, and implemented their suggestions (in particular, all the typos mentioned in the reviews were fixed).
+We carefully took into consideration all of them, and implemented their suggestions (in particular, all the typos mentioned in the reviews were fixed).
 Below please find our answers to their main comments.

### 2019-03-05 LMCS/sendrev/framework.tex
removing comments

@@
 This separation between the Boolean and quantitative levels is essential for understanding the difference between the logic and the quantitative parts of the framework.
-\martin{creo que eso quer\'ia decir esta oraci\'on}
-\marcelo{OK}
+%\martin{creo que eso quer\'ia decir esta oraci\'on}
+%\marcelo{OK}
 Furthermore, this will later allow us to parametrize both levels of the logic in order to capture different counting complexity classes.

### 2019-03-05 LMCS/sendrev/proof-prop-rest.tex
removing comments

@@
 That is, for every finite $\R$-structure $\A$, $\sem{\alpha}(\A) = \size{\A} - 1$.
-\martin{reformul\'e por qu\'e elijo ese R}
-\marcelo{OK}
+%\martin{reformul\'e por qu\'e elijo ese R}
+%\marcelo{OK}
+
 Suppose that $\alpha$ is in SNF, namely, $\alpha = \sum_{i = 1}^k \alpha_i$ for some fixed $k$. 
@@
 \qed
-\martin{cambie order por ordering, creo que se entiende mejor}
-\marcelo{me parece que las nociones de ordering y particion no quedaban claras, por esto defini de manera precisa la nocion de particion y cambie la demostracion}
+%\martin{cambie order por ordering, creo que se entiende mejor}
+%\marcelo{me parece que las nociones de ordering y particion no quedaban claras, por esto defini de manera precisa la nocion de particion y cambie la demostracion}
+
 Now we show that $\QE{0} \not\subseteq \E{1}$. 

### 2019-03-05 LMCS/sendrev/proof-sigma2hard.tex
removing comments

@@
 $$
-\martin{reescribi esta formula}
-\marcelo{OK. Le agregue los parentesis grandes a esta formula y la anterior para que quedara mas claro el scope del cuantificador existencial y la disyuncion}
+%\martin{reescribi esta formula}
+%\marcelo{OK. Le agregue los parentesis grandes a esta formula y la anterior para que quedara mas claro el scope del cuantificador existencial y la disyuncion}
+
 Notice that each first-order subformula in $\varphi^i_j(\bar{X}_i,\bar{a},\bar{b},\bar{c})$ has no free variables and, therefore, we can evaluate each of them in polynomial time and easily rewrite $\alpha_{\A}$ to an equivalent formula that does not have any first-order subformula.
@@
 
-\marcelo{Hice algunos cambios en esta demostracion (teorema \ref{sigma2hard}), por favor revisen con cuidado si todo esta bien}
+%\marcelo{Hice algunos cambios en esta demostracion (teorema \ref{sigma2hard}), por favor revisen con cuidado si todo esta bien}

### 2019-03-05 LMCS/sendrev/proof-sigmafo-minusone.tex
removing comments

@@
 Since this proof is several pages long, a line was drawn on the left margin to visually differentiate it from the rest of the paper.
-\martin{Ac\'a menciono la l\'inea del margen izquierdo}
-\marcelo{OK}
-
 \medskip
@@
 We explain below how to rewrite $\varphi_i$ in order to satisfy each condition.
-\martin{Peque\~na explicaci\'on de para qu\'e sirven las condiciones a b y c}
-\marcelo{OK}
+%\martin{Peque\~na explicaci\'on de para qu\'e sirven las condiciones a b y c}
+%\marcelo{OK}
 
@@
 \end{align}
-\martin{agregu\'e par\'entesis en varias f\'ormulas en esta p\'agina y las anteriores}
-\marcelo{OK}
+%\martin{agregu\'e par\'entesis en varias f\'ormulas en esta p\'agina y las anteriores}
+%\marcelo{OK}
+
 To understand the formula, first notice that the premise of the implication at (\ref{eq:part1}) is true if, and only if, $\varphi_i(\bar{X}, \bar{x})$ is satisfiable. 
